### 2016-10-16 docs/dg_penalty/main.tex
adding dg penalty docs

\pdfoutput=1

%\documentclass[preprint,10pt]{elsarticle}
\documentclass[10pt]{article}
%\documentclass[review]{siamart0216}

\usepackage{fullpage}
\usepackage{hyperref}
%\let\proof\relax
%\let\endproof\relax
\usepackage{amsmath,amssymb,amsfonts,mathrsfs}%,amsthm}
\usepackage[titletoc,toc,title]{appendix}

%\usepackage{lineno}
\usepackage{array} 
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{mathtools}
\usepackage{pdfpages}
\usepackage[textsize=footnotesize,color=green]{todonotes}
\usepackage{bm}
%\usepackage{tikz}
\usepackage[normalem]{ulem}
\usepackage{hhline}

%% ====================================== alg package
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{algorithmicx}
\algblock{ParFor}{EndParFor}
% customising the new block
\algnewcommand\algorithmicparfor{\textbf{parfor}}
\algnewcommand\algorithmicpardo{\textbf{do}}
\algnewcommand\algorithmicendparfor{\textbf{end\ parfor}}
\algrenewtext{ParFor}[1]{\algorithmicparfor\ #1\ \algorithmicpardo}
\algrenewtext{EndParFor}{\algorithmicendparfor}
%% ====================================== end alg package

\usepackage{graphicx}
\usepackage{subfig}
\usepackage{color}

%% ====================================== graphics

\usepackage{pgfplots}
\usepackage{pgfplotstable}
\definecolor{markercolor}{RGB}{124.9, 255, 160.65}
\pgfplotsset{width=10cm,compat=1.9}
\pgfplotsset{
tick label style={font=\small},
label style={font=\small},
legend style={font=\small}
}

\usetikzlibrary{calc}

%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
\newcommand{\logLogSlopeTriangle}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        \pgfmathsetmacro{\xArel}{#1}
        \pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{\yArel}
        \pgfmathsetmacro{\xCrel}{\xArel}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=north] {1}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=west] {#4}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

\newcommand{\logLogSlopeTriangleNeg}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        \pgfmathsetmacro{\xArel}{#1}
        \pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{\yArel}
        \pgfmathsetmacro{\xCrel}{\xArel}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=.5,anchor=south] {1}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=west] {#4}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
\newcommand{\logLogSlopeTriangleFlipNeg}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        %\pgfmathsetmacro{\xArel}{#1}
        %\pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{#3}
        \pgfmathsetmacro{\xCrel}{#1}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

	\pgfmathsetmacro{\xArel}{\xBrel}
        \pgfmathsetmacro{\yArel}{\yCrel}

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=east] {#4}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=north] {1}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
\newcommand{\logLogSlopeTriangleFlip}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        %\pgfmathsetmacro{\xArel}{#1}
        %\pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{#3}
        \pgfmathsetmacro{\xCrel}{#1}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

	\pgfmathsetmacro{\xArel}{\xBrel}
        \pgfmathsetmacro{\yArel}{\yCrel}

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=east] {#4}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=south] {1}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

%% ====================================== macros

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

\newcommand{\vect}[1]{\ensuremath\boldsymbol{#1}}
\newcommand{\tensor}[1]{\underline{\bm{#1}}}
\newcommand{\del}{\triangle}
\newcommand{\curl}{\grad \times}
\renewcommand{\div}{\grad \cdot}
\newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}{\rm #2}}}
\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
\newcommand{\pdd}[2]{\frac{\partial^2#1}{\partial#2^2}}

\newcommand{\bs}[1]{\boldsymbol{#1}}

\newcommand{\equaldef}{\stackrel{\mathrm{def}}{=}}

\newcommand{\tablab}[1]{\label{tab:#1}}
\newcommand{\tabref}[1]{Table~\ref{tab:#1}}

\newcommand{\theolab}[1]{\label{theo:#1}}
\newcommand{\theoref}[1]{\ref{theo:#1}}
\newcommand{\eqnlab}[1]{\label{eq:#1}}
\newcommand{\eqnref}[1]{\eqref{eq:#1}}
\newcommand{\seclab}[1]{\label{sec:#1}}
\newcommand{\secref}[1]{\ref{sec:#1}}
\newcommand{\lemlab}[1]{\label{lem:#1}}
\newcommand{\lemref}[1]{\ref{lem:#1}}

\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\snor}[1]{\left| #1 \right|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRb}[1]{\left| #1 \right|}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}
\newcommand{\LRceil}[1]{\left\lceil #1 \right\rceil}

\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\Div} {\ensuremath{\nabla\cdot}}
\newcommand{\Nel} {\ensuremath{{N^\text{el}}}}
\newcommand{\jump}[1] {\ensuremath{\LRs{\![#1]\!}}}
\newcommand{\avg}[1] {\ensuremath{\LRc{\!\{#1\}\!}}}
\newcommand{\uh}{\widehat{u}}
\newcommand{\Bh}{\widehat{B}}
\newcommand{\fnh}{\widehat{f}_n}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\Lk}{L^2\LRp{D^k}}
\newcommand{\Dhat}{\widehat{D}}
\newcommand{\Lhat}{L^2\LRp{\Dhat}}
\newcommand{\pO}{\partial\Omega}
\newcommand{\Gh}{\Gamma_h}
\newcommand{\Gm}{\Gamma_{-}}
\newcommand{\Gp}{\Gamma_{+}}
\newcommand{\Go}{\Gamma_0}
\newcommand{\Oh}{\Omega_h}

%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{lemma}[theorem]{Lemma}
%\newtheorem{proposition}[theorem]{Proposition}
%\newtheorem{corollary}[theorem]{Corollary}

%\newenvironment{definition}[1][Definition]{\begin{trivlist}
%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
%\newenvironment{example}[1][Example]{\begin{trivlist}
%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\newcommand{\eval}[2][\right]{\relax
  \ifx#1\right\relax \left.\fi#2#1\rvert}

\def\etal{{\it et al.~}}

\def\arr#1#2#3#4{\left[
\begin{array}{cc}
#1 & #2\\
#3 & #4\\
\end{array}
\right]}
\def\vectwo#1#2{\left[
\begin{array}{c}
#1\\
#2\\
\end{array}
\right]}
\def\vecthree#1#2#3{\left[
\begin{array}{c}
#1\\
#2\\
#3\\
\end{array}
\right]}
\def\vectfour#1#2#3#4{\left[
\begin{array}{c}
#1\\
#2\\
#3\\
#4\\
\end{array}
\right]}

\newcommand{\G} {\Gamma}
\newcommand{\Gin} {\Gamma_{in}}
\newcommand{\Gout} {\Gamma_{out}}

\newcommand{\note}[1]{{\color{blue}#1}}
\newcommand{\remark}[1]{\textbf{\color{red}#1}}

\newcommand{\tri}{{\rm tri}}
\newcommand{\sqr}{{\rm quad}}

\newcommand{\refhex}{\widehat{\mathcal{H}}}
\newcommand{\reftet}{\widehat{\mathcal{T}}}
\newcommand{\refwedg}{\widehat{\mathcal{W}}}

\newcommand{\refpyr}{\widehat{\mathcal{P}}}
\newcommand{\refpyrf}{\widehat{\mathcal{P}}^f}
\newcommand{\refpyrfq}{\widehat{\mathcal{P}}^{\sqr}}
\newcommand{\refpyrft}{\widehat{\mathcal{P}}^{\tri}}

\newcommand{\hex}{{\mathcal{H}}}
\newcommand{\tet}{{\mathcal{T}}}
\newcommand{\wedg}{{\mathcal{W}}}

\newcommand{\pyr}{{\mathcal{P}}}
\newcommand{\pyrf}{{\mathcal{P}}^f}
\newcommand{\pyrfq}{{\mathcal{P}}^{\sqr}}
\newcommand{\pyrft}{{\mathcal{P}}^{\tri}}

\newcommand{\LinfDk}{L^{\infty}\LRp{D^k}}

\newcommand{\diag}[1]{{\rm diag}\LRp{#1}}

\newcommand{\half}{1/2}

\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

%% d in integrand
\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}}

\makeatletter
\renewcommand\d[1]{\mspace{6mu}\mathrm{d}#1\@ifnextchar\d{\mspace{-3mu}}{}}
\makeatother

\newcommand{\squareface}{[-1,1]^2}

\date{}
\author{Jesse Chan}
\title{On the penalty parameter in first order DG formulations}

\begin{document}

\maketitle
\tableofcontents

%\begin{abstract}
%
%\end{abstract}

\section{Introduction}

\section{DG numerical  fluxes}

We consider a first order system of hyperbolic equations
\[
\bm{A}_0\pd{\bm{U}}{t} + \sum_{i=1}^d \pd{\bm{F}_i(\bm{U})}{\bm{x}_i} = 0,
\]
which may alternatively be written as
\[
\bm{A}_0\pd{\bm{U}}{t} + \sum_{i=1}^d \pd{\LRp{\bm{A}_{i}U}}{\bm{x}_i}, \qquad \bm{A}_{i} = \pd{\bm{F}_i(\bm{U})}{\bm{U}}
\]
where $\bm{A}_{i}$ are symmetric matrices.  The semi-discrete DG formulation for such systems may be written as 
\[
\sum_{D^k \in \Oh} \LRp{\bm{A}_0\pd{\bm{U}}{t},\bm{V} }_{\Lk} = \sum_{D^k \in \Oh}\LRp{ \LRp{\bm{F}_i(\bm{U}), \bm{V}_{,i}}_{\Lk} - \LRa{\bm{F}_i^*(\bm{U})\bm{n}_i,\bm{V}}_{\partial D^k}}
\]
where $\bm{n}$ is the outward normal on a face $f$ of $D^k$, and $\bm{F^*}$ is a numerical flux depending defined on shared faces between two elements.

For convergence, $\bm{F^*} = \bm{F^*}(\bm{U})$ must be consistent such that, for exact solutions $\bm{U}$, 
\[
\bm{F^*} = \bm{F}(\bm{U}).
\]
Let $f$ be a shared face between two elements $D^{k,+}$ and $D^{k,-}$, and let $\bm{F}^+, \bm{F}^-$ be evaluations of $\bm{F}(\bm{U})$ restricted to $D^{k,+}$ and $D^{k,-}$, respectively.  Typical DG fluxes are defined as the sum of a consistent averaging of $\bm{F}^+$ and $\bm{F}^-$ and a penalization term
\begin{equation}
\bm{F^*} = \avg{\bm{F}_n(\bm{U})} - \bm{W}\jump{\bm{U}},
\label{eq:flux}
\end{equation}
where $\bm{W}$ is some positive-definite matrix, which is required for energy stability.

The upwind numerical flux is a well-known flux of the form (\ref{eq:flux}).  For some normal vector $\bm{n}$, let $\jump{\bm{A}_n} = \sum_{i=1}^d \jump{\bm{A}_i\bm{n}_i}$.  By \note{add citation}, $\jump{\bm{A}_n}$ contains real eigenvalues, and admits an eigenvalue decomposition
\[
\jump{\bm{A}_n} = \bm{V}{\bm{\Lambda}}\bm{V}^{-1}, \qquad \bm{\Lambda} = 
\left(\begin{array}{ccc}
\lambda_1 & & \\
& \ddots & \\
& & \lambda_d
\end{array}\right).
\]
For problems with continuous coefficients, the upwind numerical flux can be defined as
\[
\bm{F^*} = \bm{A}^+\bm{U}^- + \bm{A}^- \bm{U}^+,
\]
where the matrices $\bm{A}^+,\bm{A}^-$ are constructed from the positive and negative eigenvalues 
\begin{align*}
\bm{A}^+ &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} + \LRb{\bm{\Lambda}}} \bm{V}^{-1}\\
\bm{A}^- &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} - \LRb{\bm{\Lambda}}} \bm{V}^{-1},
\end{align*}
and $\LRb{\Lambda}$ is the diagonal matrix whose entries consist of the absolute values of the eigenvalues $\LRb{\lambda_i}$.  This can be rewritten as
\[
\bm{F^*} = \avg{\bm{A}\bm{U}} - \bm{V}\LRb{\bm{\Lambda}}\bm{V}^{-1} \jump{\bm{U}}.  
\]

An alternative to upwind fluxes are penalty fluxes, which penalize appropriately defined jumps of the solution.  One such penalty flux is given as
\[
\bm{F^*} = \avg{\bm{A}\bm{U}} - \bm{A}^T \bm{W} \jump{\bm{A}\bm{U}}.  
\]
where $\bm{W}$ is some positive-definite weighting matrix.  In contrast to Lax-Friedrichs fluxes, the penalty flux enforce a weaker continuity.

\subsection{Example: acoustic wave equation}

Consider the isotropic acoustic wave equation in pressure-velocity form
\begin{align*}
\pd{p}{t} &= \Div \bm{u}\\
\pd{\bm{u}}{t} &= \Grad p.
\end{align*}
Let $\bm{U}$ denote the group variable $\bm{U} = (p,u,v)$, where $u$ and $v$ are the $x$ and $y$ components of velocity.  Then, in two dimensions, the isotropic wave equation is given as
\[
\pd{\bm{U}}{t} + \pd{\bm{A}_x\bm{U}}{x} + \pd{\bm{A}_y\bm{U}}{y} = 0, \qquad \bm{A}_x = 
\left(\begin{array}{ccc}
0 & 1 & 0\\
1 & 0 & 0\\
0 & 0 & 0
\end{array}
\right), \qquad 
\bm{A}_y = 
\left(\begin{array}{ccc}
0 & 0 & 1\\
0 & 0 & 0\\
1 & 0 & 0
\end{array}
\right).
\]
The normal flux matrix $\bm{A}_n$ is then
\[
\bm{A}_n = 
\left(\begin{array}{ccc}
0 & \bm{n}_x & \bm{n}_y\\
\bm{n}_x & 0 & 0\\
\bm{n}_y & 0 & 0
\end{array}
\right)
\]
implying that the penalty fluxes (with $\bm{W} = \tau\bm{I}$) are 
\[
\avg{\bm{A}_n \bm{U}} - \bm{A}_n^T \jump{\bm{A}_n\bm{U}} = \left(\begin{array}{c}
\avg{\bm{u}_n}\\
\avg{p \bm{n}_x}\\
\avg{p \bm{n}_y}
\end{array}
\right) - 
\tau\left(\begin{array}{c}
\jump{p}\\
\jump{\bm{u_n}}\bm{n}_x\\
\jump{\bm{u_n}}\bm{n}_y
\end{array}
\right)
\]
For $\tau = 1$, these fluxes coincide with the upwind fluxes.

\section{Dependence of spectra on the penalty parameter}

\subsection{Eigenmodes}

\note{Reproduce Gershgorin proof here.}

\subsection{Eigenvalues}

Let $\bm{A} = \bm{B} + \bm{C}$ be the DG discretization matrix, which is assumed to be the sum of a skew-symmetric matrix $\bm{B}$ and symmetric, negative-definite matrix $\bm{C}$.

Let $\lambda, \bm{u}$ be an eigenpair of $\bm{A}$. Let $\lambda = \alpha + i\beta$ and $\bm{u} = \bm{v} + i\bm{w}$, where $\alpha,\beta$ and $\bm{v},\bm{w}$ are the real and imaginary parts of $\lambda$ and $\bm{u}$, respectively.  Then, expanding and grouping terms in
\[
\bm{A}(\bm{v} + i\bm{w}) = (\alpha + i\beta) (\bm{v} + i\bm{w})
\]
we have that
\begin{align*}
\bm{A}\bm{v} &= \alpha\bm{v}-\beta\bm{w}\\
\bm{A}\bm{w} &= \beta\bm{v}+\alpha\bm{w}.
\end{align*}
Assuming that $\bm{u}^*\bm{u} = \nor{\bm{v}}^2 + \nor{\bm{w}}^2 = 1$, multiplying both sides by $\bm{v}^T,\bm{w}^T$ and straightforward manipulations %yield
%\begin{align*}
%\bm{v}^T\bm{A}\bm{v} &= \alpha\nor{\bm{v}}^2-\beta\bm{v}^T\bm{w}\\
%\bm{w}^T\bm{A}\bm{w} &= \beta\bm{w}^T{\bm{v}}+\alpha\nor{\bm{w}}^2.  
%\end{align*}
%\begin{align*}
%\bm{v}^T\bm{A}\bm{v} + \bm{w}^T\bm{A}\bm{w} &= \alpha\\
%\bm{v}^T\bm{A}\bm{w} - \bm{w}^T\bm{A}\bm{w} &= \beta.
%\end{align*}
using the skew-symmetry of $\bm{B}$ and symmetry of $\bm{C}$ yields
\begin{align*}
\tau\LRp{\bm{v}^T\bm{C}\bm{v} + \bm{w}^T\bm{C}\bm{w}} &= \alpha\\
\bm{v}^T\bm{B}\bm{w} - \bm{w}^T\bm{B}\bm{w} &= \beta.
\end{align*}
Since $\bm{C}$ is symmetric, negative-definite, and independent of $\tau$, the quantity
\[
\LRb{\bm{v}^T\bm{C}\bm{v} + \bm{w}^T\bm{C}\bm{w}} \leq \nor{\bm{C}}.
\]
Since this is bounded independently of $\tau$, we conclude that $\LRb{\alpha} = O(\tau)$.

\note{Use Gershgorin result to show that $\alpha$ diverges for eigenmodes the non-conforming subspace.}

\section{Numerical experiments}

\bibliographystyle{unsrt}
\bibliography{dgpenalty}

\end{document}

### 2016-10-16 docs/dg_penalty/main.tex
updating main with more sketch of proof

@@
 \avg{\bm{u}_n}\\
-\avg{p \bm{n}_x}\\
-\avg{p \bm{n}_y}
+\avg{p }\bm{n}_x\\
+\avg{p }\bm{n}_y
 \end{array}
@@
 
-\subsection{Eigenmodes}
+\note{Reproduce Gershgorin proof here.}
+We split the DG approximation space into conforming and non-conforming parts, which induces a block matrix
+\[
+\bm{K} = \left(\begin{array}{cc}
+\bm{A} & \bm{B}\\
+-\bm{B}^T & \bm{C} + \tau \bm{S}
+\end{array}\right).
+\]
+We will adapt the proof presented in \cite{Warburton20063205} to show that the spectra of $\bm{K}$ splits into two sets as $\tau\rightarrow \infty$, the first of which approaches the eigenvalues of $\bm{A}$, and the second of which moves left of the imaginary axis at a rate of $O(\tau)$.
 
-\note{Reproduce Gershgorin proof here.}
-
-\subsection{Eigenvalues}
-
-Let $\bm{A} = \bm{B} + \bm{C}$ be the DG discretization matrix, which is assumed to be the sum of a skew-symmetric matrix $\bm{B}$ and symmetric, negative-definite matrix $\bm{C}$.
-
-Let $\lambda, \bm{u}$ be an eigenpair of $\bm{A}$.
+We first consider the spectra of $\bm{C} + \tau\bm{S}$.
+By the construction of the DG formulation $\bm{C}$ is skew-symmetric, while $\bm{S}$ is symmetric and negative-definite.
+Let $\lambda, \bm{u}$ be an eigenpair of $\bm{C} + \tau{\bm{S}}$ corresponding to the largest $N^D$ eigenvalues.
 Let $\lambda = \alpha + i\beta$ and $\bm{u} = \bm{v} + i\bm{w}$, where $\alpha,\beta$ and $\bm{v},\bm{w}$ are the real and imaginary parts of $\lambda$ and $\bm{u}$, respectively.
@@
 \[
-\bm{A}(\bm{v} + i\bm{w}) = (\alpha + i\beta) (\bm{v} + i\bm{w})
+(\bm{C}+\tau\bm{S})(\bm{v} + i\bm{w}) = (\alpha + i\beta) (\bm{v} + i\bm{w})
 \]
@@
 \begin{align*}
-\bm{A}\bm{v} &= \alpha\bm{v}-\beta\bm{w}\\
-\bm{A}\bm{w} &= \beta\bm{v}+\alpha\bm{w}.
+(\bm{C}+\tau\bm{S})\bm{v} &= \alpha\bm{v}-\beta\bm{w}\\
+(\bm{C}+\tau\bm{S})\bm{w} &= \beta\bm{v}+\alpha\bm{w}.
 \end{align*}
@@
 %\end{align*}
-using the skew-symmetry of $\bm{B}$ and symmetry of $\bm{C}$ yields
+using the skew-symmetry of $\bm{C}$ and symmetry of $\bm{S}$ yields
 \begin{align*}
-\tau\LRp{\bm{v}^T\bm{C}\bm{v} + \bm{w}^T\bm{C}\bm{w}} &= \alpha\\
-\bm{v}^T\bm{B}\bm{w} - \bm{w}^T\bm{B}\bm{w} &= \beta.
+\bm{v}^T\bm{C}\bm{w} - \bm{w}^T\bm{C}\bm{w} &= \beta\\
+\tau\LRp{\bm{v}^T\bm{S}\bm{v} + \bm{w}^T\bm{S}\bm{w}} &= \alpha.
 \end{align*}
-Since $\bm{C}$ is symmetric, negative-definite, and independent of $\tau$, the quantity
+Since $\bm{v}^T\bm{C}\bm{w} - \bm{w}^T\bm{C}\bm{w}= 2\bm{w}^T\bm{C}\bm{v} = \beta$ is independent of $\tau$ and $\bm{v},\bm{w}$ are normalized, $\beta$ remains bounded as $\tau\rightarrow \infty$.
+Similarly, since $\bm{S}$ is symmetric, negative-definite, and independent of $\tau$, the quantity
 \[
-\LRb{\bm{v}^T\bm{C}\bm{v} + \bm{w}^T\bm{C}\bm{w}} \leq \nor{\bm{C}}.
+\LRb{\bm{v}^T\bm{S}\bm{v} + \bm{w}^T\bm{S}\bm{w}} = \LRb{\frac{\alpha}{\tau}}
 \]
-Since this is bounded independently of $\tau$, we conclude that $\LRb{\alpha} = O(\tau)$.
+is bounded independently of $\tau$, and we conclude that $\LRb{\alpha} = O(\tau)$.
+Thus, as $\tau$ increases, the eigenvalues of $\bm{C} + \tau\bm{S}$ are shifted further and further left of the imaginary axis.
 
-\note{Use Gershgorin result to show that $\alpha$ diverges for eigenmodes the non-conforming subspace.}
+\subsection{Bifurcation of eigenvalue sets}
+
+Following \cite{Warburton20063205}, let $\bm{U}$ and $\bm{Q}$ be unitary matrices whose columns contain the eigenvectors of $\bm{A}$ and $\bm{C} + \tau \bm{S}$, respectively.
+Then, applying a block diagonal similarity transform yields
+\[
+\left(\begin{array}{cc}
+\bm{U}^* & \\
+& \bm{Q}^*
+\end{array}\right)
+\left(\begin{array}{cc}
+\bm{A} & \bm{B}\\
+-\bm{B}^T & \bm{C} + \tau \bm{S}
+\end{array}\right)
+\left(\begin{array}{cc}
+\bm{U} & \\
+& \bm{Q}
+\end{array}\right)
+ = \left(\begin{array}{cc}
+\bm{\Lambda}_C & \bm{U}^T\bm{B}\bm{Q}\\
+-\bm{Q}^T\bm{B}^T\bm{U} & \bm{\Lambda}_D
+\end{array}\right),
+\]
+where $\bm{\Lambda}_C,\bm{\Lambda}_D$ are diagonal matrices whose entries consist of eigenvalues of $\bm{A},\bm{C} + \tau\bm{S}$, respectively.
+Since $\bm{B}$ is independent of $\tau$, $\nor{\bm{U}^T\bm{B}\bm{Q}}$ can be bounded independently of $\tau$, assuming that $\bm{U}$ and $\bm{Q}$ are normalized.
+Gerschgorin's theorem applied to $\bm{\Lambda}_C,\bm{\Lambda}_D$ then implies that the eigenvalues of $\bm{K}$ are contained in two sets of discs with radii independent of $\tau$.
+The first set of discs are centered around $\lambda^C_i$ for $i = 1,\ldots,N^C$, while the second set of discs are centered around $\lambda^D_i$ for $i = 1,\ldots,N^D$, where ${\rm Re}(\lambda^D_i) = O(\tau)$.
+
+Consider now the $N^C$ eigenvalues with the smallest magnitude.
+By the Gerschgorin argument, these must remain bounded as $\tau \rightarrow \infty$.
+Let $\bm{W} = (\bm{W}_C,\bm{W}_D)$ and $\Lambda_C$ be the matrix of eigenvectors and eigenvalues corresponding to these $N^C$ smallest eigenvalues, such that
+\[
+\left(\begin{array}{cc}
+\bm{A} & \bm{B}\\
+-\bm{B}^T & \bm{C} + \tau \bm{S}
+\end{array}\right)
+\left(\begin{array}{c}
+\bm{W}_C
+\\
+\bm{W}_D
+\end{array}\right) = 
+\left(\begin{array}{c}
+\bm{A}\bm{W}_C + \bm{B}\bm{W}_D\\
+-\bm{B}^T\bm{W}_C + \bm{C}\bm{W}_D + \tau \bm{S}\bm{W}_D
+\end{array}\right)
+= 
+\bm{\Lambda}_C
+\left(\begin{array}{c}
+\bm{W}_C
+\\
+\bm{W}_D
+\end{array}\right) 
+\]
+This implies 
+\[
+\tau \nor{\bm{S}\bm{W}_D} = \nor{\bm{\Lambda}_C \bm{W}_D + \bm{B}^T\bm{W}_C  - \bm{C}\bm{W}_D}.
+\]
+Since $\Lambda_C$ remains bounded as $\tau\rightarrow \infty$, the right hand side is bounded independently of $\tau$ under normalization of $\bm{W}_C,\bm{W}_D$, and we conclude that the non-conforming
+\[
+\nor{\bm{S}\bm{W}_D} = O(1/\tau).
+\]
+As a consequence, as $\tau\rightarrow \infty$, the smallest $N^C$ eigenvalues of $\bm{K}$ converge to the purely imaginary eigenvalues of $\bm{A}$.
+These correspond to a discretization using a conforming approximation space.
 

### 2016-10-20 docs/dg_penalty/main.tex
updating note

@@
 \usetikzlibrary{calc}
-
-%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
-\newcommand{\logLogSlopeTriangle}[5]
-{
-    % #1. Relative offset in x direction.
-    % #2. Width in x direction, so xA-xB.
-    % #3. Relative offset in y direction.
-    % #4. Slope d(y)/d(log10(x)).
-    % #5. Plot options.
-
-    \pgfplotsextra
-    {
-        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
-        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
-        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
-        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}
-
-        % Calculate auxilliary quantities, in relative sense.
-        \pgfmathsetmacro{\xArel}{#1}
-        \pgfmathsetmacro{\yArel}{#3}
-        \pgfmathsetmacro{\xBrel}{#1-#2}
-        \pgfmathsetmacro{\yBrel}{\yArel}
-        \pgfmathsetmacro{\xCrel}{\xArel}
-
-        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
-        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
-        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.
-
-        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
-        \coordinate (A) at (rel axis cs:\xArel,\yArel);
-        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
-        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);
-
-        % Draw slope triangle.
-        \draw[#5]   (A)-- node[pos=0.5,anchor=north] {1}
-                    (B)-- 
-                    (C)-- node[pos=0.5,anchor=west] {#4}
-                    cycle;
-    }
-}
-%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
-
-\newcommand{\logLogSlopeTriangleNeg}[5]
-{
-    % #1. Relative offset in x direction.
-    % #2. Width in x direction, so xA-xB.
-    % #3. Relative offset in y direction.
-    % #4. Slope d(y)/d(log10(x)).
-    % #5. Plot options.
-
-    \pgfplotsextra
-    {
-        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
-        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
-        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
-        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}
-
-        % Calculate auxilliary quantities, in relative sense.
-        \pgfmathsetmacro{\xArel}{#1}
-        \pgfmathsetmacro{\yArel}{#3}
-        \pgfmathsetmacro{\xBrel}{#1-#2}
-        \pgfmathsetmacro{\yBrel}{\yArel}
-        \pgfmathsetmacro{\xCrel}{\xArel}
-
-        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
-        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
-        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.
-
-        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
-        \coordinate (A) at (rel axis cs:\xArel,\yArel);
-        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
-        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);
-
-        % Draw slope triangle.
-        \draw[#5]   (A)-- node[pos=.5,anchor=south] {1}
-                    (B)-- 
-                    (C)-- node[pos=0.5,anchor=west] {#4}
-                    cycle;
-    }
-}
-%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
-
-%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
-\newcommand{\logLogSlopeTriangleFlipNeg}[5]
-{
-    % #1. Relative offset in x direction.
-    % #2. Width in x direction, so xA-xB.
-    % #3. Relative offset in y direction.
-    % #4. Slope d(y)/d(log10(x)).
-    % #5. Plot options.
-
-    \pgfplotsextra
-    {
-        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
-        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
-        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
-        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}
-
-        % Calculate auxilliary quantities, in relative sense.
-        %\pgfmathsetmacro{\xArel}{#1}
-        %\pgfmathsetmacro{\yArel}{#3}
-        \pgfmathsetmacro{\xBrel}{#1-#2}
-        \pgfmathsetmacro{\yBrel}{#3}
-        \pgfmathsetmacro{\xCrel}{#1}
-
-        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
-        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
-        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.
-
-	\pgfmathsetmacro{\xArel}{\xBrel}
-        \pgfmathsetmacro{\yArel}{\yCrel}
-
-        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
-        \coordinate (A) at (rel axis cs:\xArel,\yArel);
-        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
-        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);
-
-        % Draw slope triangle.
-        \draw[#5]   (A)-- node[pos=0.5,anchor=east] {#4}
-                    (B)-- 
-                    (C)-- node[pos=0.5,anchor=north] {1}
-                    cycle;
-    }
-}
-%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
-
-%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
-\newcommand{\logLogSlopeTriangleFlip}[5]
-{
-    % #1. Relative offset in x direction.
-    % #2. Width in x direction, so xA-xB.
-    % #3. Relative offset in y direction.
-    % #4. Slope d(y)/d(log10(x)).
-    % #5. Plot options.
-
-    \pgfplotsextra
-    {
-        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
-        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
-        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
-        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}
-
-        % Calculate auxilliary quantities, in relative sense.
-        %\pgfmathsetmacro{\xArel}{#1}
-        %\pgfmathsetmacro{\yArel}{#3}
-        \pgfmathsetmacro{\xBrel}{#1-#2}
-        \pgfmathsetmacro{\yBrel}{#3}
-        \pgfmathsetmacro{\xCrel}{#1}
-
-        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
-        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
-        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.
-
-	\pgfmathsetmacro{\xArel}{\xBrel}
-        \pgfmathsetmacro{\yArel}{\yCrel}
-
-        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
-        \coordinate (A) at (rel axis cs:\xArel,\yArel);
-        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
-        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);
-
-        % Draw slope triangle.
-        \draw[#5]   (A)-- node[pos=0.5,anchor=east] {#4}
-                    (B)-- 
-                    (C)-- node[pos=0.5,anchor=south] {1}
-                    cycle;
-    }
-}
-%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
-
-%% ====================================== macros
+\usepackage{stmaryrd}
 
@@
 \newcommand{\Nel} {\ensuremath{{N^\text{el}}}}
-\newcommand{\jump}[1] {\ensuremath{\LRs{\![#1]\!}}}
+\newcommand{\jump}[1] {\ensuremath{\llbracket#1\rrbracket}}
 \newcommand{\avg}[1] {\ensuremath{\LRc{\!\{#1\}\!}}}
@@
 \date{}
-\author{Jesse Chan}
-\title{On the penalty parameter in first order DG formulations}
+\author{Jesse Chan, T. Warburton}
+\title{On penalty flux parameters in first order DG formulations}
 
@@
 \[
-\sum_{D^k \in \Oh} \LRp{\bm{A}_0\pd{\bm{U}}{t},\bm{V} }_{\Lk} = \sum_{D^k \in \Oh}\LRp{ \LRp{\bm{F}_i(\bm{U}), \bm{V}_{,i}}_{\Lk} - \LRa{\bm{F}_i^*(\bm{U})\bm{n}_i,\bm{V}}_{\partial D^k}}
+\sum_{D^k \in \Oh} \LRp{\bm{A}_0\pd{\bm{U}}{t},\bm{V} }_{\Lk} = \sum_{D^k \in \Oh}\LRp{ \LRp{\bm{F}_i(\bm{U}), \bm{V}_{,i}}_{\Lk} - \LRa{\bm{F}_n^*(\bm{U})\bm{n}_i,\bm{V}}_{\partial D^k}}
 \]
-where $\bm{n}$ is the outward normal on a face $f$ of $D^k$, and $\bm{F^*}$ is a numerical flux depending defined on shared faces between two elements.
+where $\bm{n}$ is the outward normal on a face $f$ of $D^k$, and $\bm{F^*}_n$ is a numerical flux depending defined on shared faces between two elements.
 
-For convergence, $\bm{F^*} = \bm{F^*}(\bm{U})$ must be consistent such that, for exact solutions $\bm{U}$, 
+%For convergence, $\bm{F^*}_n = \bm{F^*}_n(\bm{U})$ must be consistent such that, for exact solutions $\bm{U}$, 
+%\[
+%\bm{F^*}_n = \bm{F}(\bm{U})\cdot \bm{n}.
+%\]
+
+Let $f$ be a shared face between two elements $D^{k,+}$ and $D^{k,-}$, and let $\bm{F}^+, \bm{F}^-$ be evaluations of $\bm{F}(\bm{U})$ restricted to $D^{k,+}$ and $D^{k,-}$, respectively.
+The average and jump of a vector valued function over $f$ are then defined as
 \[
-\bm{F^*} = \bm{F}(\bm{U}).
+\avg{\bm{F}} = \frac{\bm{F}^+ + \bm{F}^-}{2}, \qquad \jump{\bm{F}} = \bm{F}^+ - \bm{F}^-.
 \]
-Let $f$ be a shared face between two elements $D^{k,+}$ and $D^{k,-}$, and let $\bm{F}^+, \bm{F}^-$ be evaluations of $\bm{F}(\bm{U})$ restricted to $D^{k,+}$ and $D^{k,-}$, respectively.
-Typical DG fluxes are defined as the sum of a consistent averaging of $\bm{F}^+$ and $\bm{F}^-$ and a penalization term
+Typical numerical fluxes are defined as the sum of the average of $\bm{F}^+$ and $\bm{F}^-$ and a penalization term
 \begin{equation}
-\bm{F^*} = \avg{\bm{F}_n(\bm{U})} - \bm{W}\jump{\bm{U}},
+\bm{F^*}_n = \avg{\bm{F}(\bm{U})}\cdot \bm{n} - \bm{G}\LRp{\jump{\bm{U}}},
 \label{eq:flux}
 \end{equation}
-where $\bm{W}$ is some positive-definite matrix, which is required for energy stability.
+where $\bm{G}$ is some convex function.
 
 The upwind numerical flux is a well-known flux of the form (\ref{eq:flux}).
-For some normal vector $\bm{n}$, let $\jump{\bm{A}_n} = \sum_{i=1}^d \jump{\bm{A}_i\bm{n}_i}$.
-By \note{add citation}, $\jump{\bm{A}_n}$ contains real eigenvalues, and admits an eigenvalue decomposition
+For some normal vector $\bm{n}$, let ${\bm{A}}_n = \sum_{i=1}^d {\bm{A}_i\bm{n}_i}$.
+By \note{add citation}, ${\bm{A}}_n$ contains real eigenvalues, and admits an eigenvalue decomposition
 \[
-\jump{\bm{A}_n} = \bm{V}{\bm{\Lambda}}\bm{V}^{-1}, \qquad \bm{\Lambda} = 
+\bm{A}_n = \bm{V}{\bm{\Lambda}}\bm{V}^{-1}, \qquad \bm{\Lambda} = 
 \left(\begin{array}{ccc}
@@
 \[
-\bm{F^*} = \bm{A}^+\bm{U}^- + \bm{A}^- \bm{U}^+,
+\bm{F^*}_n = \bm{A}_n^+\bm{U}^- + \bm{A_n}^- \bm{U}^+,
 \]
-where the matrices $\bm{A}^+,\bm{A}^-$ are constructed from the positive and negative eigenvalues 
+where the matrices $\bm{A_n}^+,\bm{A_n}^-$ are constructed from the positive and negative eigenvalues 
 \begin{align*}
-\bm{A}^+ &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} + \LRb{\bm{\Lambda}}} \bm{V}^{-1}\\
-\bm{A}^- &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} - \LRb{\bm{\Lambda}}} \bm{V}^{-1},
+\bm{A}_n^+ &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} + \LRb{\bm{\Lambda}}} \bm{V}^{-1}\\
+\bm{A}_n^- &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} - \LRb{\bm{\Lambda}}} \bm{V}^{-1},
 \end{align*}
-and $\LRb{\Lambda}$ is the diagonal matrix whose entries consist of the absolute values of the eigenvalues $\LRb{\lambda_i}$.  This can be rewritten as
+and $\LRb{\bm{\Lambda}}$ is the diagonal matrix whose entries consist of the absolute values of the eigenvalues $\LRb{\lambda_i}$.
+This can be rewritten as the sum of the central flux and a penalization
 \[
-\bm{F^*} = \avg{\bm{A}\bm{U}} - \bm{V}\LRb{\bm{\Lambda}}\bm{V}^{-1} \jump{\bm{U}}.  
+\bm{F^*}_n = \avg{\bm{A}_n\bm{U}} - \tau\bm{V}\LRb{\bm{\Lambda}}\bm{V}^{-1} \jump{\bm{U}}.  
+\]
+For problems with discontinuous coefficients, this may be generalized by solving a Riemann problem to yield
+\[
+\bm{F^*}_n = \avg{\bm{A}_n\bm{U}} - \tau \jump{{\bm{V}\LRb{\bm{\Lambda}}\bm{V}^{-1}\bm{U}}}.  
 \]
 
-An alternative to upwind fluxes are penalty fluxes, which penalize appropriately defined jumps of the solution.  One such penalty flux is given as
+An alternative to upwind fluxes are penalty fluxes, which penalize appropriately defined jumps of the solution.  One such penalty flux is 
 \[
-\bm{F^*} = \avg{\bm{A}\bm{U}} - \bm{A}^T \bm{W} \jump{\bm{A}\bm{U}}.  
+\bm{F^*}_n = \avg{\bm{A}_n\bm{U}} - \bm{A}_n^T \bm{W} \jump{\bm{A}_n\bm{U}}.  
 \]
-where $\bm{W}$ is some positive-definite weighting matrix.  In contrast to Lax-Friedrichs fluxes, the penalty flux enforce a weaker continuity.
+where $\bm{W}$ is a diagonal weighting matrix with positive entries.
+We take $\bm{W} = \tau \bm{I}$ in the following sections; however, the entries can be made to vary.
+For example, taking $\bm{W} = \LRb{\bm{\Lambda}}^{-1}$ recovers the upwind fluxes \note{prove this}.
+%\[
+%\bm{F^*}_n = \avg{\bm{A}_n\bm{U}} - \bm{A}_n^T \LRb{\bm{Lambda}}^{-1} \jump{\bm{A}_n\bm{U}}.  
+%\]
 
-\subsection{Example: acoustic wave equation}
+\subsection{Examples: acoustic wave equation}
 
@@
 \[
-\avg{\bm{A}_n \bm{U}} - \bm{A}_n^T \jump{\bm{A}_n\bm{U}} = \left(\begin{array}{c}
+\avg{\bm{A}_n \bm{U}} - \tau\bm{A}_n^T \jump{\bm{A}_n\bm{U}} = \left(\begin{array}{c}
 \avg{\bm{u}_n}\\
@@
 \]
-For $\tau = 1$, these fluxes coincide with the upwind fluxes.
+For $\tau = 1$, these fluxes coincide with the upwind fluxes for isotropic media.
+We note that the penalty flux affects only the normal component of the velocity; this is in contrast to Lax-Friedrichs-type fluxes, which penalize component-wise discontinuities in the velocity.
+
+\note{Skew-symmetric discretization for constant coefficient problems \cite{kopriva2014energy}.
+For wave propagation, can wrap variable coefficients into $A_0$ \cite{chan2016weight1,chan2016weight2}}.
 
@@
 
-\note{Reproduce Gershgorin proof here.}
-We split the DG approximation space into conforming and non-conforming parts, which induces a block matrix
+\note{Define spaces and show that $\bm{S}$ is symmetric definite based on DG discretization.}
+
+We adapt the proof presented in \cite{Warburton20063205} to show that the spectra of $\bm{K}$ splits into two sets as $\tau\rightarrow \infty$, the first of which approaches the eigenvalues of $\bm{A}$, and the second of which moves left of the imaginary axis at a rate of $O(\tau)$.
+
+We begin by splitting the DG approximation space into a conforming part $V^C$ and a non-conforming part $V^{NC}$.  Let $V$ be the DG approximation space
+\[
+V = \LRc{ u \in L^2(\Oh) : \left.u\right|_{D^k} \in P^N(D^k)}.  
+\]
+The conforming space $V^{C}$ is then defined as
+\begin{equation}
+V^{C} = \LRc{ u(\bm{x}) \in V: }
+\label{eq:conf}
+\end{equation}
+We can similarly define the non-conforming approximation space $V^{NC}$ as the orthogonal complement of $V^{C}$, such that the full approximation space $V$ is 
+\[
+V = V^C \oplus V^{NC}.  
+\]
+
+This induces a block decomposition
 \[
@@
 \]
-We will adapt the proof presented in \cite{Warburton20063205} to show that the spectra of $\bm{K}$ splits into two sets as $\tau\rightarrow \infty$, the first of which approaches the eigenvalues of $\bm{A}$, and the second of which moves left of the imaginary axis at a rate of $O(\tau)$.
-
 We first consider the spectra of $\bm{C} + \tau\bm{S}$.
@@
 
-\subsection{Bifurcation of eigenvalue sets}
-
+We may now show how the spectra of $\bm{K}$ behaves as $\tau \rightarrow \infty$.
 Following \cite{Warburton20063205}, let $\bm{U}$ and $\bm{Q}$ be unitary matrices whose columns contain the eigenvectors of $\bm{A}$ and $\bm{C} + \tau \bm{S}$, respectively.
@@
 By the Gerschgorin argument, these must remain bounded as $\tau \rightarrow \infty$.
-Let $\bm{W} = (\bm{W}_C,\bm{W}_D)$ and $\Lambda_C$ be the matrix of eigenvectors and eigenvalues corresponding to these $N^C$ smallest eigenvalues, such that
+Let $\bm{W} = (\bm{W}_C,\bm{W}_D)$ and $\bm{\Lambda}_C$ be the matrix of eigenvectors and eigenvalues corresponding to these $N^C$ smallest eigenvalues, such that
 \[
@@
 \]
-Since $\Lambda_C$ remains bounded as $\tau\rightarrow \infty$, the right hand side is bounded independently of $\tau$ under normalization of $\bm{W}_C,\bm{W}_D$, and we conclude that the non-conforming
+Since $\Lambda_C$ remains bounded as $\tau\rightarrow \infty$, the right hand side is bounded independently of $\tau$ under normalization of $\bm{W}_C,\bm{W}_D$, implying that the non-conforming component $\bm{W}^D$ satisfies
 \[
-\nor{\bm{S}\bm{W}_D} = O(1/\tau).
+C\nor{\bm{W}_D} \leq \nor{\bm{S}\bm{W}_D} = O(1/\tau).
 \]
-As a consequence, as $\tau\rightarrow \infty$, the smallest $N^C$ eigenvalues of $\bm{K}$ converge to the purely imaginary eigenvalues of $\bm{A}$.
-These correspond to a discretization using a conforming approximation space.
+As a consequence, $\bm{W}_D \rightarrow 0$ as $\tau\rightarrow \infty$, and the smallest $N^C$ eigenvalues of $\bm{K}$ converge to the purely imaginary eigenvalues of $\bm{A}$ at a rate of $O(1/\tau)$.
+These correspond to a discretization using the conforming approximation space (\ref{eq:conf}).
 
 \section{Numerical experiments}
+
+The above analysis illustrates the behavior of the spectra of the DG discretization for the asymptotic cases when $\tau = 0$ or $\tau \rightarrow \infty$.
+However, it is less clear how the spectra of $\bm{K}$ behaves for $\tau = O(1)$.
+We rely instead on numerical experiments in one and two space dimensions to illustrate behaviors for the advection and acoustic wave equations.
+
+\subsection{1D experiments}
+
+We consider first the scalar advection equation in 1D with periodic boundary conditions.
+For $\tau = 0$, the eigenvalues of the DG discretization matrix are purely imaginary.
+Figure~\ref{fig:track1D} shows the paths taken by these eigenvalues as $\tau$ increases from zero to $\tau = 4$.
+As predicted, a subset of eigenvalues diverges towards the left half plane, which we refer to as \emph{divergent}.
+The corresponding eigenmodes are shown in Figure~\ref{fig:trackmodes1D}, with inter-element jumps of these eigenfunctions increasing as $\tau$ increases.
+
+\begin{figure}
+\centering
+\subfloat[Paths of eigenvalues as $\tau$ increases]{\includegraphics[width=.46\textwidth]{figs/trackedEigs.png}}
+\hspace{1em}
+\subfloat[Zoom of eigenvalues near imaginary axis]{\includegraphics[width=.44\textwidth]{figs/trackedEigsZoom.png}}
+\caption{Eigenvalue paths for DG advection with $\tau \in [0,4]$ on a mesh of 8 elements of degree $N=3$.
+Eigenvalues are overlaid on these paths for $\tau = 0$, $\tau = 1$, and $\tau = 4$.
+The zoomed in view near the imaginary axis shows the return of spurious modes to the imaginary axis for $\tau$ sufficiently large.
+}
+\label{fig:track1D}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigsDiverge1.png}}
+\hspace{.5em}
+\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigsDiverge2.png}}
+\hspace{.5em}
+\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigsDiverge3.png}}\\
+\subfloat[$\tau = 0$]{\includegraphics[width=.32\textwidth]{figs/tauModesDiverge1.png}}
+\hspace{.5em}
+\subfloat[$\tau = 1$]{\includegraphics[width=.32\textwidth]{figs/tauModesDiverge2.png}}
+\hspace{.5em}
+\subfloat[$\tau = 10$]{\includegraphics[width=.32\textwidth]{figs/tauModesDiverge3.png}}
+\caption{Behavior of divergent eigenvalues for advection as $\tau$ increases.
+The order of approximation is $N=3$ on a mesh of 4 elements.} %The displayed eigenfunctions correspond to the highlighted eigenvalue.
+}
+\label{fig:trackmodes1D}
+\end{figure}
+
+However, for sufficiently large $\tau$, a subset of eigenvalues return to the imaginary axis.
+Figure~\ref{fig:spurious} illustrates that, as $\tau$ increases, the magnitude of the inter-element jumps present in these modes decreases, and the mode approaches a continuous function.
+These eigenvalues which return to the imaginary axis correspond to a second set of \emph{spurious} modes, consisting of higher frequency components with sharp peaks, which have previously been observed in high order $C^0$ finite element methods \cite{ainsworth2014dispersive,hughes2014finite}.
+
+\begin{figure}
+\centering
+\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigs1.png}}
+\hspace{.5em}
+\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigs2.png}}
+\hspace{.5em}
+\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigs3.png}}\\
+\subfloat[$\tau = 0$]{\includegraphics[width=.32\textwidth]{figs/tauModes1.png}}
+\hspace{.5em}
+\subfloat[$\tau = 1$]{\includegraphics[width=.32\textwidth]{figs/tauModes2.png}}
+\hspace{.5em}
+\subfloat[$\tau = 10$]{\includegraphics[width=.32\textwidth]{figs/tauModes3.png}}
+\caption{Return behavior of DG advection eigenvalues for sufficiently large $\tau$.
+For $\tau = 1$ certain eigenvalues have negative real part, and the corresponding eigenmodes are damped.
+For $\tau \gg 1$, certain eigenvalues return to the imaginary axis as spurious modes in conforming discretizations.
+}
+\label{fig:spurious}
+\end{figure}
+
+We repeat the above experiment for the acoustic wave equation in 1D with reflection boundary conditions on $[-1,1]$, and compare against the exact eigenvalues $\lambda_i = i \frac{\pi}{2}$ for $i = 1,2,\ldots$.
+The behavior of the eigenvalues as $\tau$ increases (Figure~\ref{fig:trackModesWave}) is similar to the behavior observed for the scalar advection equation.
+An examination of the eigenmodes corresponding to divergent and spurious eigenvalues as $\tau \rightarrow \infty$ shows that these modes behave similarly to the spurious modes observed for the scalar advection equation, with individual pressure and velocity components converging to continuous functions.
+
+%\begin{figure}
+%\centering
+%\subfloat{\includegraphics[width=.475\textwidth]{figs/tauEigs1.png}}
+%\hspace{.5em}
+%\subfloat{\includegraphics[width=.475\textwidth]{figs/tauEigs2.png}}
+%\caption{ }
+%\label{fig:trackModesWave}
+%\end{figure}
+
+\subsection{2D experiments}
+
+In two space dimensions, taking $\tau\rightarrow \infty$ again results in spurious modes which return to the imaginary axis.
+The corresponding eigenvalues converge to conforming functions.
+For the acoustic wave equation, $p$ is continuous across faces, edges, and vertices, implying that $p\in H^1(\Oh)$, while only the normal component of $\bm{u}$ is continuous across faces, implying that $\bm{u} \in H({\rm div}; \Oh)$.
+
+Uniform triangular mesh resulting from bisecting each element of a $2\times 2$ uniform quadrilateral mesh.  
+%For the advection equation, 
+%\begin{figure}
+%\centering
+%\subfloat{\includegraphics[width=.475\textwidth]{figs/trackedAdvecEigs.png}}
+%\hspace{.5em}
+%\subfloat{\includegraphics[width=.465\textwidth]{figs/trackedAdvecEigsZoom1.png}}
+%\caption{Behavior of eigenvalues for advection }
+%\label{fig:trackModesAdvec}
+%\end{figure}
+
+\begin{figure}
+\centering
+\subfloat[$\tau = 1$]{\includegraphics[width=.3\textwidth]{figs/waveEigs1.png}}
+\hspace{.5em}
+\subfloat[$\tau = 10$]{\includegraphics[width=.3\textwidth]{figs/waveEigs2.png}}
+\hspace{.5em}
+\subfloat[$\tau = 100$]{\includegraphics[width=.3\textwidth]{figs/waveEigs3.png}}
+\caption{Behavior of eigenvalues for the acoustic wave equation in two dimensions. Note the changing scale of the real axis.}
+\label{fig:waveeigs}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfloat[$\tau = 0,1, 50$, zoomed in view]{\includegraphics[height=.3\textheight]{figs/waveEigs2Dbox.png}}
+\hspace{2em}
+\subfloat[$\tau= 0,1,50$, returning eigenvalue paths]{\includegraphics[height=.3\textheight]{figs/trackedWaveEigs2D.png}}
+\caption{Behavior of eigenvalues for the acoustic wave equation in two dimensions.  Divergent eigenvalues in the far left half plane are not shown.  }
+\label{fig:trackEigsWave}
+\end{figure}
+
+\begin{figure}
+\centering
+\subfloat[$\tau = .1$, $\lambda = -0.2379 + 8.7528i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_p1.png}}
+\hspace{.5em}
+\subfloat[$\tau = 1$, $\lambda = -1.0145 + 8.9270i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_p2.png}}
+\hspace{.5em}
+\subfloat[$\tau = 50$, $\lambda = -0.0437 + 7.6167i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_p3.png}}
+\caption{Behavior of spurious returning modes for the acoustic wave equation in two dimensions.  }
+\label{fig:trackModesWave}
+\end{figure}
+
+\section{Conclusions}
 

### 2016-10-21 docs/dg_penalty/main.tex
updating note with more

@@
 \maketitle
-\tableofcontents
+%\tableofcontents
 
-%\begin{abstract}
-%
-%\end{abstract}
+\begin{abstract}
+Abstract goes here.  Mention assumptions: continuously varying material data, skew-symmetric formulations.  
+\end{abstract}
 
 \section{Introduction}
+
+\note{Intro}
 
@@
 
+\note{Define $\Oh, \Gh$, derive global formulation.}
+
 %For convergence, $\bm{F^*}_n = \bm{F^*}_n(\bm{U})$ must be consistent such that, for exact solutions $\bm{U}$, 
@@
 \]
-Typical numerical fluxes are defined as the sum of the average of $\bm{F}^+$ and $\bm{F}^-$ and a penalization term
+Typical numerical fluxes are defined as the sum of the average of $\bm{F}^+$ and $\bm{F}^-$ and a penalization term $\bm{G(\bm{U}^+,\bm{U}^-)}$
 \begin{equation}
-\bm{F^*}_n = \avg{\bm{F}(\bm{U})}\cdot \bm{n} - \bm{G}\LRp{\jump{\bm{U}}},
+\bm{F^*}_n = \avg{\bm{F}(\bm{U})}\cdot \bm{n} - \bm{G(\bm{U}^+,\bm{U}^-)}.
 \label{eq:flux}
 \end{equation}
-where $\bm{G}$ is some convex function.
 
 The upwind numerical flux is a well-known flux of the form (\ref{eq:flux}).
-For some normal vector $\bm{n}$, let ${\bm{A}}_n = \sum_{i=1}^d {\bm{A}_i\bm{n}_i}$.
+For an outward normal vector $\bm{n}$, let ${\bm{A}}_n = \sum_{i=1}^d {\bm{A}_i\bm{n}_i}$.
 By \note{add citation}, ${\bm{A}}_n$ contains real eigenvalues, and admits an eigenvalue decomposition
 \[
-\bm{A}_n = \bm{V}{\bm{\Lambda}}\bm{V}^{-1}, \qquad \bm{\Lambda} = 
+\bm{A}_n = \bm{V}{\bm{\Lambda}}\bm{V}^T, \qquad \bm{\Lambda} = 
 \left(\begin{array}{ccc}
@@
 \]
-For problems with continuous coefficients, the upwind numerical flux can be defined as
+For problems with continuous coefficients, the upwind numerical flux over a face $f \in \Gh$ can be defined  as
 \[
@@
 \begin{align*}
-\bm{A}_n^+ &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} + \LRb{\bm{\Lambda}}} \bm{V}^{-1}\\
-\bm{A}_n^- &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} - \LRb{\bm{\Lambda}}} \bm{V}^{-1},
+\bm{A}_n^+ &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} + \LRb{\bm{\Lambda}}} \bm{V}^T\\
+\bm{A}_n^- &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} - \LRb{\bm{\Lambda}}} \bm{V}^T,
 \end{align*}
@@
 \[
-\bm{F^*}_n = \avg{\bm{A}_n\bm{U}} - \tau\bm{V}\LRb{\bm{\Lambda}}\bm{V}^{-1} \jump{\bm{U}}.  
+\bm{F^*}_n = \avg{\bm{A}_n\bm{U}} - \bm{V}\LRb{\bm{\Lambda}}\bm{V}^T \jump{\bm{U}}.  
 \]
@@
 \[
-\bm{F^*}_n = \avg{\bm{A}_n\bm{U}} - \tau \jump{{\bm{V}\LRb{\bm{\Lambda}}\bm{V}^{-1}\bm{U}}}.  
+\bm{F^*}_n = \avg{\bm{A}_n\bm{U}} - \jump{{\bm{V}\LRb{\bm{\Lambda}}\bm{V}^T\bm{U}}}.  
 \]
-
 An alternative to upwind fluxes are penalty fluxes, which penalize appropriately defined jumps of the solution.  One such penalty flux is 
 \[
-\bm{F^*}_n = \avg{\bm{A}_n\bm{U}} - \bm{A}_n^T \bm{W} \jump{\bm{A}_n\bm{U}}.  
+\bm{F^*}_n = \avg{\bm{A}_n\bm{U}} - \tau \bm{A}_n^T \jump{\bm{A}_n\bm{U}}.  
 \]
-where $\bm{W}$ is a diagonal weighting matrix with positive entries.
-We take $\bm{W} = \tau \bm{I}$ in the following sections; however, the entries can be made to vary.
-For example, taking $\bm{W} = \LRb{\bm{\Lambda}}^{-1}$ recovers the upwind fluxes \note{prove this}.
-%\[
-%\bm{F^*}_n = \avg{\bm{A}_n\bm{U}} - \bm{A}_n^T \LRb{\bm{Lambda}}^{-1} \jump{\bm{A}_n\bm{U}}.  
-%\]
+where $\tau$ is proportional to $\lambda_{\max}^{-1}$.
+This choice of $\tau$ ensures that the penalty term is of the same order of magnitude as the upwind penalization, and that the CFL restriction for both methods is identical.
 
-\subsection{Examples: acoustic wave equation}
+\note{Skew-symmetric discretization for constant coefficient problems \cite{kopriva2014energy}.
+Extendable to Maxwells \cite{warburton2013low} and elasticity \cite{wilcox2010high,ye2016discontinuous}.}
 
-Consider the isotropic acoustic wave equation in pressure-velocity form
+\subsection{Conforming and non-conforming spaces}
+
+Let $V$ denote the piecewise polynomial approximation space 
+\[
+V = \LRc{ \bm{U} \in\LRp{ L^2(\Oh)}^d : \left.\bm{U}\right|_{D^k} \in \LRp{P^N(D^k)}^d}.  
+\]
+We define a decomposition of $V$ into a conforming part $V^C$ and a non-conforming part $V^{NC}$ based on the penalization term $\bm{G}(\bm{U}^+,\bm{U}^-)$.
+The conforming part is defined as the null space of $\bm{G}(\bm{U}^+,\bm{U}^-)$
+\begin{equation}
+V^{C} = \LRc{ \bm{U}(\bm{x}) \in V: \quad  \bm{G}(\bm{U}^+,\bm{U}^-) = 0, \quad \forall f \in \Gh},
+\label{eq:conf}
+\end{equation}
+while the non-conforming approximation space $V^{NC}$ is defined as $L^2$ orthogonal complement of $V^{C}$ in $V$
+\[
+V^{NC} = \LRc{ \bm{U}(\bm{x}) \in V: \LRp{\bm{u},\bm{v}}_{L^2(\Oh)} = 0, \quad \forall \bm{v}\in V^C}.
+\]
+\note{For problems where $\bm{A}_i$ are spatially continuous, $\bm{V}\LRb{\bm{\Lambda}}\bm{V}^T \jump{\bm{U}} = 0$ if and only if $\bm{A}_n^T \bm{A}_n\jump{\bm{U}} = 0$.  Both conditions enforce that 
+\[
+\bm{v}_i^T \jump{\bm{U}} = 0, \qquad \lambda_i \neq 0.
+\]
+}
+As a result, the conforming and non-conforming spaces induced by upwind and penalty fluxes are identical.
+
+For problems with discontinuous material data, it is less straightforward to show equivalence between the conforming spaces $V^C$ resulting from upwind and penalty fluxes.
+However, for many practical wave propagation settings, spatially varying coefficients are incorporated into the matrix $\bm{A_0}$, resulting in constant $\bm{A}_i$.
+For both smoothly varying and discontinuous $\bm{A}_0$, it is possible to formulate energy stable DG methods with penalty fluxes \cite{mercerat2015nodal, chan2016weight1}.
+Additionally, DG methods using penalty fluxes are observed to perform similarly to DG methods with upwind fluxes for  wave propagation problems with discontinuous coefficients \cite{warburton2013low, chan2016weight1, ye2016discontinuous}.
+
+\subsection{Scalar advection and acoustic wave propagation}
+\label{sec:confexamples}
+We illustrate the conforming spaces induced by penalty fluxes using the scalar advection equation and acoustic wave equation as concrete examples.  The scalar advection equation is given as
+\[
+\pd{u}{t} + \div\LRp{\bm{\beta}u} = 0.
+\]
+where $\bm{\beta}$ is the direction of advection.
+For continuous advection vectors such that $\bm{\beta}(\bm{x}) \neq 0$, the penalty flux is equivalent to a parametrized upwind flux \cite{hesthaven2007nodal}
+\[
+\bm{F}^*_n = \avg{\bm{\beta}_n u} - \tau\LRb{\bm{\beta}_n}\jump{u}.
+\]
+The conforming space $V^C$ induced by this flux is then
+\[
+V^C = \LRc{ u \in L^2\LRp{\Oh} : \quad \left.u\right|_{D^k} \in P^N(D^k), \quad \LRb{\bm{\beta}_n}\jump{u} = 0}.
+\]
+In one dimension, this simply implies that $u$ is $C^0$ continuous across element boundaries.
+In higher dimensions, this implies that $u$ is continuous along streamlines or directions $\bm{d}(\bm{x})$ where $\bm{\beta}\cdot \bm{d} \neq 0$.
+
+Next, we consider the acoustic wave equation in pressure-velocity form
 \begin{align*}
-\pd{p}{t} &= \Div \bm{u}\\
-\pd{\bm{u}}{t} &= \Grad p.
+\frac{1}{c^2}\pd{p}{t} &= \Div \bm{u}\\
+\pd{\bm{u}}{t} &= \Grad p,
 \end{align*}
-Let $\bm{U}$ denote the group variable $\bm{U} = (p,u,v)$, where $u$ and $v$ are the $x$ and $y$ components of velocity.  Then, in two dimensions, the isotropic wave equation is given as
+where $c^2(\bm{x})$ is the wavespeed.
+Let $\bm{U}$ denote the group variable $\bm{U} = (p,u,v)$, where $u$ and $v$ are the $x$ and $y$ components of velocity.
+Then, in two dimensions, the isotropic wave equation is given as
 \[
@@
 \]
-For $\tau = 1$, these fluxes coincide with the upwind fluxes for isotropic media.
-We note that the penalty flux affects only the normal component of the velocity; this is in contrast to Lax-Friedrichs-type fluxes, which penalize component-wise discontinuities in the velocity.
+For $\tau = 1$, these fluxes coincide with the upwind fluxes for continuously varying media.
+In this case, the conforming subspace $V^{C}$ induced by the penalty flux consists of a scalar (pressure) component $V^C_p$ and a vector (velocity) component $V^C_{\bm{u}}$.
+The pressure component $p$ satisfies $\jump{p} = 0$ over all faces $f\in \Gh$.
+For polynomial approximation spaces, this implies that $p$ is continuous across faces, edges, and vertices, and $V^C_p$ is the standard $C^0$ continuous piecewise polynomial finite element space.
+For the velocity component of $V^C$, the penalty fluxes enforce normal continuity, such that $\jump{\bm{u}}\cdot \bm{n} = 0$ over all faces $f \in \Gh$.
+Since each component of $\bm{u}$ is approximated from $P^N(D^k)$, this implies that $V^C_{\bm{u}}$ is the $H({\rm div})$-conforming Brezzi-Douglas-Marini finite element space \cite{brezzi1985two,boffi2013mixed}.\footnote{It is possible to recover $V^C_{\bm{u}}$ as the Raviart-Thomas finite element space by approximating each component of $\bm{u}$ from a smaller polynomial space \cite{kirby2004algorithm}.}
 
-\note{Skew-symmetric discretization for constant coefficient problems \cite{kopriva2014energy}.
-For wave propagation, can wrap variable coefficients into $A_0$ \cite{chan2016weight1,chan2016weight2}}.
+Penalty fluxes can also be interpreted as Lax-Friedrichs fluxes applied to scalar and vector variables separately.  In contrast, for a component-wise Lax-Friedrichs flux 
+\[
+\avg{\bm{A}_n \bm{U}} - \tau\bm{A}_n^T \jump{\bm{A}_n\bm{U}} = \left(\begin{array}{c}
+\avg{\bm{u}_n}\\
+\avg{p }\bm{n}_x\\
+\avg{p }\bm{n}_y
+\end{array}
+\right) - 
+\tau\left(\begin{array}{c}
+\jump{p}\\
+\jump{\bm{u}_x}\\
+\jump{\bm{u}_y}
+\end{array}
+\right),
+\]
+the conforming space $V^C_{\bm{u}}$ becomes the space of vector piecewise polynomial functions which are $C^0$ continuous in both components, resulting in overly-constrained continuity conditions for $\bm{u}$.
 
@@
 
-We begin by splitting the DG approximation space into a conforming part $V^C$ and a non-conforming part $V^{NC}$.  Let $V$ be the DG approximation space
-\[
-V = \LRc{ u \in L^2(\Oh) : \left.u\right|_{D^k} \in P^N(D^k)}.  
-\]
-The conforming space $V^{C}$ is then defined as
-\begin{equation}
-V^{C} = \LRc{ u(\bm{x}) \in V: }
-\label{eq:conf}
-\end{equation}
-We can similarly define the non-conforming approximation space $V^{NC}$ as the orthogonal complement of $V^{C}$, such that the full approximation space $V$ is 
-\[
-V = V^C \oplus V^{NC}.  
-\]
-
+\note{finish}
 This induces a block decomposition
@@
 \end{align*}
-Assuming that $\bm{u}^*\bm{u} = \nor{\bm{v}}^2 + \nor{\bm{w}}^2 = 1$, multiplying both sides by $\bm{v}^T,\bm{w}^T$ and straightforward manipulations %yield
+Assuming that $\nor{\bm{u}}^2 = \nor{\bm{v}}^2 + \nor{\bm{w}}^2 = 1$, multiplying both sides by $\bm{v}^T,\bm{w}^T$ and straightforward manipulations %yield
 %\begin{align*}
@@
 For $\tau = 0$, the eigenvalues of the DG discretization matrix are purely imaginary.
+Paths taken by eigenvalues as $\tau$ increases are determined by sampling spectra over a sufficiently fine set of $\tau$ and using a particle tracking method \cite{simpletracker}.
 Figure~\ref{fig:track1D} shows the paths taken by these eigenvalues as $\tau$ increases from zero to $\tau = 4$.
-As predicted, a subset of eigenvalues diverges towards the left half plane, which we refer to as \emph{divergent}.
+As predicted, a subset of \emph{divergent} eigenvalues move further left of the imaginary axis as $\tau$ increases.
 The corresponding eigenmodes are shown in Figure~\ref{fig:trackmodes1D}, with inter-element jumps of these eigenfunctions increasing as $\tau$ increases.
@@
 However, for sufficiently large $\tau$, a subset of eigenvalues return to the imaginary axis.
-Figure~\ref{fig:spurious} illustrates that, as $\tau$ increases, the magnitude of the inter-element jumps present in these modes decreases, and the mode approaches a continuous function.
-These eigenvalues which return to the imaginary axis correspond to a second set of \emph{spurious} modes, consisting of higher frequency components with sharp peaks, which have previously been observed in high order $C^0$ finite element methods \cite{ainsworth2014dispersive,hughes2014finite}.
+Figure~\ref{fig:spurious} illustrates that, as $\tau$ increases, the magnitude of the inter-element jumps present in these modes decreases, and the mode approaches a $C^0$ continuous function.
+These eigenvalues which move right to return to the imaginary axis correspond to a second set of \emph{spurious} modes.
+These spurious modes, corresponding to under-resolved modes of the exact underlying operator, consist of higher frequency components with sharp peaks and have previously been observed in high order $C^0$ finite element methods \cite{ainsworth2014dispersive,hughes2014finite}.
 
@@
 
-In two space dimensions, taking $\tau\rightarrow \infty$ again results in spurious modes which return to the imaginary axis.
-The corresponding eigenvalues converge to conforming functions.
-For the acoustic wave equation, $p$ is continuous across faces, edges, and vertices, implying that $p\in H^1(\Oh)$, while only the normal component of $\bm{u}$ is continuous across faces, implying that $\bm{u} \in H({\rm div}; \Oh)$.
+In two space dimensions, taking $\tau\rightarrow \infty$ again results in spurious modes which return to the imaginary axis, with the corresponding eigenmodes approaching conforming functions.  
+We illustrate this through numerical experiments for the acoustic wave equation and advection in two space dimensions.
+In all numerical experiments, we use $N=3$ and a uniform triangular mesh resulting from bisecting each element of a $2\times 2$ uniform quadrilateral mesh.
 
-Uniform triangular mesh resulting from bisecting each element of a $2\times 2$ uniform quadrilateral mesh.  
-%For the advection equation, 
-%\begin{figure}
-%\centering
-%\subfloat{\includegraphics[width=.475\textwidth]{figs/trackedAdvecEigs.png}}
-%\hspace{.5em}
-%\subfloat{\includegraphics[width=.465\textwidth]{figs/trackedAdvecEigsZoom1.png}}
-%\caption{Behavior of eigenvalues for advection }
-%\label{fig:trackModesAdvec}
-%\end{figure}
+Figure~\ref{fig:waveeigs} shows the spectra of the DG discretization matrix for the acoustic wave equation for $\tau = 1, 10, 50$.
+Like the one-dimensional case, a subset of eigenvalues diverge towards the left half plane as $\tau$ increases.
+However, unlike the one-dimensional case, these eigenvalues collapse towards the real line as $\tau$ increases.
 
@@
 
+Similarly, as predicted, a subset of eigenvalues returns towards the imaginary axis as $\tau \rightarrow \infty$.
+These are shown in Figure~\ref{fig:trackEigsWave}, along with traced paths taken by each eigenvalue as $\tau$ increases.
+
 \begin{figure}
 \centering
-\subfloat[$\tau = 0,1, 50$, zoomed in view]{\includegraphics[height=.3\textheight]{figs/waveEigs2Dbox.png}}
+\subfloat[$\tau = 0,1, 50$, zoomed in view]{\includegraphics[height=.3\textheight]{figs/waveEigs2Dbox.png}\label{subfig:big}}
 \hspace{2em}
-\subfloat[$\tau= 0,1,50$, returning eigenvalue paths]{\includegraphics[height=.3\textheight]{figs/trackedWaveEigs2D.png}}
-\caption{Behavior of eigenvalues for the acoustic wave equation in two dimensions.  Divergent eigenvalues in the far left half plane are not shown.  }
+\subfloat[$\tau= 0,1,50$, returning eigenvalue paths]{\includegraphics[height=.3\textheight]{figs/trackedWaveEigs2D.png}\label{subfig:box}}
+\caption{Behavior of eigenvalues for the acoustic wave equation in two dimensions.
+Figure~\ref{subfig:box} shows a zoom of the boxed region in Figure~\ref{subfig:big}, with overlaid eigenvalue paths as $\tau$ increases.
+Divergent eigenvalues in the far left half plane are not shown.
+}
 \label{fig:trackEigsWave}
 \end{figure}
+
+The pressure and velocity components of the corresponding acoustic eigenmodes converge to conforming approximations.
+As noted in Section~\ref{sec:confexamples}, this implies that pressure components lie in $H^1(\Oh)$ and are continuous across faces, edges, and vertices.
+Figure~\ref{fig:trackModesWave} shows the behavior of the pressure component of an eigenmode corresponding to a spurious returning eigenvalue for $\tau = .1, 1, 100$.
+As $\tau$ increases and the eigenvalue approaches the imaginary axis, the eigenmode approaches a $C^0$ continuous function with a high frequencies and sharp peak, similar to the spurious eigenmodes observed for the one-dimensional case in Figure~\ref{fig:spurious}.
 
@@
 \hspace{.5em}
-\subfloat[$\tau = 50$, $\lambda = -0.0437 + 7.6167i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_p3.png}}
-\caption{Behavior of spurious returning modes for the acoustic wave equation in two dimensions.  }
+\subfloat[$\tau = 100$, $\lambda = -0.0437 + 7.6167i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_p3.png}}
+\caption{Behavior of the real part of the pressure component of spurious eigenmodes for the acoustic wave equation in two dimensions.  }
 \label{fig:trackModesWave}
@@
 
+The non-zero velocity components of spurious eigenmodes converge to approximations in $H({\rm div}; \Oh)$, such that only the normal component of velocity is continuous across faces.
+Figure~\ref{fig:trackModesWaveU} shows visualizations of these eigenmodes.
+For large $\tau$, normal jumps of the velocity component vanish.
+% approach an $H({\rm div};\Oh)$-conforming approximation.
+
+\begin{figure}
+\centering
+\subfloat[$\tau = .1$, $\lambda = -0.6009 +11.3209i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_u1.png}}
+\hspace{.5em}
+\subfloat[$\tau = 1$, $\lambda =  -1.9601 + 9.5629i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_u2.png}}
+\hspace{.5em}
+\subfloat[$\tau = 100$, $\lambda = -0.0634 + 7.7670i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_u3.png}}
+\caption{Behavior of the real part of the velocity component of spurious eigenmodes for the acoustic wave equation in two dimensions.  }
+\label{fig:trackModesWaveU}
+\end{figure}
+
+For the advection equation with periodic boundary conditions, the spectra of the DG discretization behaves similarly.
+In all following experiments, we use advection vector $\beta = (1,0)$.
+For $\tau=0$, all eigenvalues lie on the imaginary axis, and as $\tau$ increases, the spectra splits into eigenvalues which diverge towards the left half plane and eigenvalues which return to the imaginary axis.
+Figure~\ref{fig:trackModesAdvec} shows the spectra of the DG matrix for $\tau = 0,1,50$, with eigenvalue paths for increasing values of $\tau$ overlaid.
+
+\begin{figure}
+\centering
+\subfloat{\includegraphics[width=.475\textwidth]{figs/trackedAdvecEigs.png}}
+\hspace{.5em}
+\subfloat{\includegraphics[width=.465\textwidth]{figs/trackedAdvecEigsZoom1.png}}
+\caption{Eigenvalue paths for the DG discretization of advection as $\tau$ increases.  Divergent eigenvalues in the far left half plane are not shown.  }
+\label{fig:trackModesAdvec}
+\end{figure}
+
+Unlike spurious conforming modes present for the acoustic wave equation, spurious conforming modes for the advection equation satisfy $\jump{\beta_n u} = 0$, and allow discontinuities along faces which are tangential to flow directions.
+\note{Find cite on regularity of advection solutions.}
+
+\begin{figure}
+\centering
+\subfloat[$\tau = 0$, $\lambda = 0 + 7.9246i$]{\includegraphics[width=.3\textwidth]{figs/advecEig2D_u1.png}}
+\hspace{.5em}
+\subfloat[$\tau = 1$, $\lambda =  -0.8278 + 6.7831i$]{\includegraphics[width=.3\textwidth]{figs/advecEig2D_u2.png}}
+\hspace{.5em}
+\subfloat[$\tau = 100$, $\lambda = -0.0239 + 5.9282i$]{\includegraphics[width=.3\textwidth]{figs/advecEig2D_u3.png}}
+\caption{Behavior of the real part of spurious eigenmodes for the advection equation with $\beta = (1,0)$. }
+\label{fig:trackModesAdvecU}
+\end{figure}
+
 \section{Conclusions}
+
+\note{Penalized DG methods as sweet spot treatments of spurious modes, somewhere between central DG and conforming discretizations.}
 

### 2016-10-23 docs/dg_penalty/main.tex
updating note

@@
 
+\newcommand{\mi}{\mathrm{i}} %% roman "i"
 \newcommand{\mb}[1]{\mathbf{#1}}
@@
 
-%\newtheorem{theorem}{Theorem}[section]
-%\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}[theorem]{Lemma}
 %\newtheorem{proposition}[theorem]{Proposition}
@@
 
-\section{DG numerical  fluxes}
+\section{Semi-discrete discontinuous Galerkin formulation}
 
-We consider a first order system of hyperbolic equations
+We consider linear first order hyperbolic systems of PDEs in $\mathbb{R}^d$
+\begin{equation}
+\bm{A}_0(\bm{x})\pd{\bm{U}}{t} + \sum_{i=1}^d \pd{\LRp{\bm{A}_{i}\bm{U}}}{\bm{x}_i} = 0
+\label{eq:cons}
+\end{equation}
+with solution components $\bm{U} = (u_1, \ldots, u_M)^T$.
+In this work, we assume that $\bm{A}_0$ is symmetric and positive-definite, and that the matrices $\bm{A}_i$ are symmetric or symmetrizable.
+In this work, we assume that $\bm{A}_i$ are constant, though $\bm{A}_0(\bm{x})$ may vary spatially.
+These assumptions are true for many PDEs governing wave propagation in heterogeneous media, including such as the acoustic wave equation \cite{chan2016weight1}, linear elasticity \cite{de2008interior,wilcox2010high,ye2016discontinuous}, and Maxwell's equations \cite{hesthaven2002nodal,grote2007interior,warburton2013low}.
+%\note{Restrict to constant coeff $\bm{A}_i$?}
+
+We assume the domain $\Omega$ is triangulated exactly using a mesh $\Oh$ consisting of elements $D^k$.  We define the $L^2$ inner product over elements $D^k$ and their boundaries $\partial D^k$ as
 \[
-\bm{A}_0\pd{\bm{U}}{t} + \sum_{i=1}^d \pd{\bm{F}_i(\bm{U})}{\bm{x}_i} = 0,
+\LRp{u,v}_{D^k} = \int_{D^k} u(\bm{x})v(\bm{x}) \diff x, \qquad \LRa{u,v}_{\partial D^k} = \int_{\partial D^k} u(\bm{x})v(\bm{x}) \diff x.
 \]
-which may alternatively be written as
+A semi-discrete DG formulation for (\ref{eq:cons}) may be written as 
+\begin{equation}
+\sum_{D^k \in \Oh} \LRp{\bm{A}_0\pd{\bm{U}}{t},\bm{V} }_{\Lk} = \sum_{D^k \in \Oh}\LRp{ \LRp{\bm{A}_i(\bm{U}), \bm{V}_{,i}}_{\Lk} - \LRa{(\bm{A}_n\bm{U})^*,\bm{V}}_{\partial D^k}}
+\label{eq:formulation}
+\end{equation}
+where $(\bm{A}_n\bm{U})^*$ is a numerical flux defined on boundary faces and shared faces between elements.
+
+%We assume that $\bm{A}_{i}$ are symmetric matrices.
+
+\subsection{Penalization terms in numerical fluxes}
+
+Let $f$ be a shared face between two elements $D^{k,+}$ and $D^{k,-}$ with outward normal $\bm{n}^-$ on $D^{k,-}$, and let $\bm{U}^+, \bm{U}^-$ be evaluations of $\bm{U}$ restricted to $D^{k,+}$ and $D^{k,-}$, respectively.
+The average of both scalar and vector-valued functions is defined as
 \[
-\bm{A}_0\pd{\bm{U}}{t} + \sum_{i=1}^d \pd{\LRp{\bm{A}_{i}U}}{\bm{x}_i}, \qquad \bm{A}_{i} = \pd{\bm{F}_i(\bm{U})}{\bm{U}}
+\avg{u} = \frac{{u}^+ + u^-}{2},
 \]
-where $\bm{A}_{i}$ are symmetric matrices.  The semi-discrete DG formulation for such systems may be written as 
+and the jump of scalar and vector-valued functions over $f$ are defined as
 \[
-\sum_{D^k \in \Oh} \LRp{\bm{A}_0\pd{\bm{U}}{t},\bm{V} }_{\Lk} = \sum_{D^k \in \Oh}\LRp{ \LRp{\bm{F}_i(\bm{U}), \bm{V}_{,i}}_{\Lk} - \LRa{\bm{F}_n^*(\bm{U})\bm{n}_i,\bm{V}}_{\partial D^k}}
+\jump{u} = u^+ \bm{n}^+ + u^-\bm{n}^-, \qquad \jump{\bm{u}} = \bm{u}^+\cdot \bm{n}^+ + \bm{u}^-\cdot\bm{n}^-.  
 \]
-where $\bm{n}$ is the outward normal on a face $f$ of $D^k$, and $\bm{F^*}_n$ is a numerical flux depending defined on shared faces between two elements.
-
-\note{Define $\Oh, \Gh$, derive global formulation.}
-
-%For convergence, $\bm{F^*}_n = \bm{F^*}_n(\bm{U})$ must be consistent such that, for exact solutions $\bm{U}$, 
-%\[
-%\bm{F^*}_n = \bm{F}(\bm{U})\cdot \bm{n}.
-%\]
-
-Let $f$ be a shared face between two elements $D^{k,+}$ and $D^{k,-}$, and let $\bm{F}^+, \bm{F}^-$ be evaluations of $\bm{F}(\bm{U})$ restricted to $D^{k,+}$ and $D^{k,-}$, respectively.
-The average and jump of a vector valued function over $f$ are then defined as
-\[
-\avg{\bm{F}} = \frac{\bm{F}^+ + \bm{F}^-}{2}, \qquad \jump{\bm{F}} = \bm{F}^+ - \bm{F}^-.
-\]
-Typical numerical fluxes are defined as the sum of the average of $\bm{F}^+$ and $\bm{F}^-$ and a penalization term $\bm{G(\bm{U}^+,\bm{U}^-)}$
+Typical numerical fluxes are defined as the sum of the average of $\bm{U}^+$ and $\bm{U}^-$ and a penalization term $\bm{G(\bm{U}^+,\bm{U}^-)}$
 \begin{equation}
-\bm{F^*}_n = \avg{\bm{F}(\bm{U})}\cdot \bm{n} - \bm{G(\bm{U}^+,\bm{U}^-)}.
+(\bm{A}_n\bm{U})^* = \avg{\bm{A}_n\bm{U}}\cdot \bm{n}^- - \bm{G(\bm{U}^+,\bm{U}^-)}.
 \label{eq:flux}
 \end{equation}
-
+\note{Clean up jump notation here.}
 The upwind numerical flux is a well-known flux of the form (\ref{eq:flux}).
-For an outward normal vector $\bm{n}$, let ${\bm{A}}_n = \sum_{i=1}^d {\bm{A}_i\bm{n}_i}$.
-By \note{add citation}, ${\bm{A}}_n$ contains real eigenvalues, and admits an eigenvalue decomposition
+For an outward normal vector $\bm{n}^-$, let ${\bm{A}}_n = \sum_{i=1}^d {\bm{A}_i\bm{n}^-_i}$.
+By symmetry of $\bm{A}_i$, ${\bm{A}}_n$ contains real eigenvalues, and admits an eigenvalue decomposition
 \[
@@
 \[
-\bm{F^*}_n = \bm{A}_n^+\bm{U}^- + \bm{A_n}^- \bm{U}^+,
+(\bm{A}_n\bm{U})^* = \bm{A}_n^+\bm{U}^- + \bm{A_n}^- \bm{U}^+,
 \]
@@
 \[
-\bm{F^*}_n = \avg{\bm{A}_n\bm{U}} - \bm{V}\LRb{\bm{\Lambda}}\bm{V}^T \jump{\bm{U}}.  
+(\bm{A}_n\bm{U})^* = \avg{\bm{A}_n\bm{U}} - \bm{V}\LRb{\bm{\Lambda}}\bm{V}^T \jump{\bm{U}}.  
 \]
-For problems with discontinuous coefficients, this may be generalized by solving a Riemann problem to yield
+%For problems with discontinuous coefficients, this may be generalized by solving a Riemann problem to yield
+%\[
+%\bm{F^*}_n = \avg{\bm{A}_n\bm{U}} - \jump{{\bm{V}\LRb{\bm{\Lambda}}\bm{V}^T\bm{U}}}.  
+%\]
+We consider penalty fluxes as an alternative to upwind fluxes, which penalize appropriately defined jumps of the solution.  One such penalty flux is 
 \[
-\bm{F^*}_n = \avg{\bm{A}_n\bm{U}} - \jump{{\bm{V}\LRb{\bm{\Lambda}}\bm{V}^T\bm{U}}}.  
-\]
-An alternative to upwind fluxes are penalty fluxes, which penalize appropriately defined jumps of the solution.  One such penalty flux is 
-\[
-\bm{F^*}_n = \avg{\bm{A}_n\bm{U}} - \tau \bm{A}_n^T \jump{\bm{A}_n\bm{U}}.  
+(\bm{A}_n\bm{U})^* = \avg{\bm{A}_n\bm{U}} - \frac{\tau}{2} \bm{A}_n^T \jump{\bm{A}_n\bm{U}}.  
 \]
 where $\tau$ is proportional to $\lambda_{\max}^{-1}$.
-This choice of $\tau$ ensures that the penalty term is of the same order of magnitude as the upwind penalization, and that the CFL restriction for both methods is identical.
+This choice of $\tau$ ensures that the penalty term is of the same order of magnitude as the upwind penalization, and that the CFL restriction for both upwind and penalty fluxes are identical.
+Unlike upwind fluxes, penalty fluxes can be used when a diagonalization of $\bm{A}_n$ is not explicitly known, such as linear elasticity in anisotropic media \cite{ye2016discontinuous}.
 
-\note{Skew-symmetric discretization for constant coefficient problems \cite{kopriva2014energy}.
-Extendable to Maxwells \cite{warburton2013low} and elasticity \cite{wilcox2010high,ye2016discontinuous}.}
+\section{Dependence of spectra on the penalty parameter}
+
+The remainder of this work focuses on the influence of $\tau$ on the spectra of DG discretization matrices.
 
@@
 \[
-V = \LRc{ \bm{U} \in\LRp{ L^2(\Oh)}^d : \left.\bm{U}\right|_{D^k} \in \LRp{P^N(D^k)}^d}.  
+V = \LRc{ \bm{U} \in\LRp{ L^2(\Omega)}^d : \left.\bm{U}\right|_{D^k} \in \LRp{P^N(D^k)}^d}.  
 \]
@@
 \[
-V^{NC} = \LRc{ \bm{U}(\bm{x}) \in V: \LRp{\bm{u},\bm{v}}_{L^2(\Oh)} = 0, \quad \forall \bm{v}\in V^C}.
+V^{NC} = \LRc{ \bm{U}(\bm{x}) \in V: \LRp{\bm{u},\bm{v}}_{L^2(\Omega)} = 0, \quad \forall \bm{v}\in V^C}.
 \]
-\note{For problems where $\bm{A}_i$ are spatially continuous, $\bm{V}\LRb{\bm{\Lambda}}\bm{V}^T \jump{\bm{U}} = 0$ if and only if $\bm{A}_n^T \bm{A}_n\jump{\bm{U}} = 0$.  Both conditions enforce that 
+When the coefficient matrices $\bm{A}_i$ are spatially continuous, the upwind flux condition $\bm{V}\LRb{\bm{\Lambda}}\bm{V}^T \jump{\bm{U}} = 0$ and the penalty flux condition $\bm{A}_n^T \jump{\bm{A}_n\bm{U}} = 0$ are both equivalent to
 \[
-\bm{v}_i^T \jump{\bm{U}} = 0, \qquad \lambda_i \neq 0.
+\bm{v}_i^T \jump{\bm{U}} = 0, \qquad \lambda_i \neq 0,
 \]
-}
-As a result, the conforming and non-conforming spaces induced by upwind and penalty fluxes are identical.
+where $\lambda_i, \bm{v}_i$ are an eigenvalue and eigenvector of $\bm{A}_n$.  As a result, the conforming and non-conforming spaces induced by upwind and penalty fluxes are identical.
 
@@
 \[
-\bm{F}^*_n = \avg{\bm{\beta}_n u} - \tau\LRb{\bm{\beta}_n}\jump{u}.
+(\bm{A}_n\bm{U})^* = \avg{\bm{\beta}_n u} - \tau\LRb{\bm{\beta}_n}\jump{u}.
 \]
@@
 \[
-V^C = \LRc{ u \in L^2\LRp{\Oh} : \quad \left.u\right|_{D^k} \in P^N(D^k), \quad \LRb{\bm{\beta}_n}\jump{u} = 0}.
+V^C = \LRc{ u \in L^2\LRp{\Omega} : \quad \left.u\right|_{D^k} \in P^N(D^k), \quad \LRb{\bm{\beta}_n}\jump{u} = 0}.
 \]
-In one dimension, this simply implies that $u$ is $C^0$ continuous across element boundaries.
+In one dimension, this simply implies that $u$ is $C^0$-continuous across element boundaries.
 In higher dimensions, this implies that $u$ is continuous along streamlines or directions $\bm{d}(\bm{x})$ where $\bm{\beta}\cdot \bm{d} \neq 0$.
@@
 \left(\begin{array}{ccc}
-0 & \bm{n}_x & \bm{n}_y\\
-\bm{n}_x & 0 & 0\\
-\bm{n}_y & 0 & 0
+0 & \bm{n}^-_x & \bm{n}^-_y\\
+\bm{n}^-_x & 0 & 0\\
+\bm{n}^-_y & 0 & 0
 \end{array}
@@
 \]
-implying that the penalty fluxes (with $\bm{W} = \tau\bm{I}$) are 
+implying that the penalty fluxes are 
 \[
-\avg{\bm{A}_n \bm{U}} - \tau\bm{A}_n^T \jump{\bm{A}_n\bm{U}} = \left(\begin{array}{c}
+\avg{\bm{A}_n \bm{U}} - \frac{\tau}{2}\bm{A}_n^T \jump{\bm{A}_n\bm{U}} = \left(\begin{array}{c}
 \avg{\bm{u}_n}\\
-\avg{p }\bm{n}_x\\
-\avg{p }\bm{n}_y
+\avg{p }\bm{n}^-_x\\
+\avg{p }\bm{n}^-_y
 \end{array}
 \right) - 
-\tau\left(\begin{array}{c}
+\frac{\tau}{2}\left(\begin{array}{c}
 \jump{p}\\
-\jump{\bm{u_n}}\bm{n}_x\\
-\jump{\bm{u_n}}\bm{n}_y
+\jump{\bm{u_n}}\bm{n}^-_x\\
+\jump{\bm{u_n}}\bm{n}^-_y
 \end{array}
@@
 For polynomial approximation spaces, this implies that $p$ is continuous across faces, edges, and vertices, and $V^C_p$ is the standard $C^0$ continuous piecewise polynomial finite element space.
-For the velocity component of $V^C$, the penalty fluxes enforce normal continuity, such that $\jump{\bm{u}}\cdot \bm{n} = 0$ over all faces $f \in \Gh$.
+
+For the velocity component $V_{\bm{u}}^C$, the penalty fluxes enforce normal continuity, such that $\jump{\bm{u}}\cdot \bm{n}^- = 0$ over all faces $f \in \Gh$.
 Since each component of $\bm{u}$ is approximated from $P^N(D^k)$, this implies that $V^C_{\bm{u}}$ is the $H({\rm div})$-conforming Brezzi-Douglas-Marini finite element space \cite{brezzi1985two,boffi2013mixed}.\footnote{It is possible to recover $V^C_{\bm{u}}$ as the Raviart-Thomas finite element space by approximating each component of $\bm{u}$ from a smaller polynomial space \cite{kirby2004algorithm}.}
@@
 \[
-\avg{\bm{A}_n \bm{U}} - \tau\bm{A}_n^T \jump{\bm{A}_n\bm{U}} = \left(\begin{array}{c}
+(\bm{A}_n\bm{U})^* = \left(\begin{array}{c}
 \avg{\bm{u}_n}\\
-\avg{p }\bm{n}_x\\
-\avg{p }\bm{n}_y
+\avg{p }\bm{n}^-_x\\
+\avg{p }\bm{n}^-_y
 \end{array}
 \right) - 
-\tau\left(\begin{array}{c}
+\frac{\tau}{2}\left(\begin{array}{c}
 \jump{p}\\
@@
 \]
-the conforming space $V^C_{\bm{u}}$ becomes the space of vector piecewise polynomial functions which are $C^0$ continuous in both components, resulting in overly-constrained continuity conditions for $\bm{u}$.
+the conforming velocity space $V^C_{\bm{u}}$ becomes the space of vector piecewise polynomial functions which are $C^0$ continuous in both $x$ and $y$ components, resulting in over-constrained continuity conditions for $\bm{u}$.
 
-\section{Dependence of spectra on the penalty parameter}
+\subsection{Behavior as $\tau\rightarrow\infty$}
 
-\note{Define spaces and show that $\bm{S}$ is symmetric definite based on DG discretization.}
+Let $\bm{K}$ be the matrix resulting from the discretization of the DG formulation (\ref{eq:formulation}).  We are interested in eigenvalues and eigenvectors of 
+\[
+\bm{K}\bm{u} = \lambda_i\bm{M}\bm{u}
+\]
+where $\bm{M}$ is the $L^2$ mass matrix.
+We adapt the approach of Warburton and Embree in \cite{Warburton20063205} to (\ref{eq:block}) to show that the spectra of $\bm{K}$ splits into two sets of eigenvalues as $\tau\rightarrow \infty$, the first of which approaches the eigenvalues of $\bm{A}$, and the second of which diverges with real part approaching $-\infty$ as $\tau\rightarrow \infty$.
 
-We adapt the proof presented in \cite{Warburton20063205} to show that the spectra of $\bm{K}$ splits into two sets as $\tau\rightarrow \infty$, the first of which approaches the eigenvalues of $\bm{A}$, and the second of which moves left of the imaginary axis at a rate of $O(\tau)$.
-
-\note{finish}
-This induces a block decomposition
-\[
+Let $\LRc{\bm{\Phi}_i }_{i=1}^{N^{C}}$ and $\LRc{\bm{\Psi}_i }_{i=1}^{N^{NC}}$ be $L^2$-orthogonal bases for $\bm{V}^C$ and $\bm{V}^{NC}$, respectively.
+These spaces then induce a block decomposition of the DG discretization matrix $\bm{K}$
+\begin{equation}
 \bm{K} = \left(\begin{array}{cc}
 \bm{A} & \bm{B}\\
--\bm{B}^T & \bm{C} + \tau \bm{S}
-\end{array}\right).
+\tilde{\bm{B}}^T & \bm{C} + \tau \bm{S}
+\end{array}\right),
+\label{eq:block}
+\end{equation}
+where the blocks $\bm{A},\bm{B},\bm{C}$ are defined as
+\begin{align*}
+\bm{A}_{mn} &= \sum_{D^k\in \Oh} \LRp{\bm{A}_i\bm{\Phi}_n, \bm{\Phi}_{m,i}}_{D^k}, \qquad 1\leq m,n \leq N^C\\
+\bm{B}_{mn} &= \sum_{D^k\in \Oh} \LRp{\bm{A}_i\bm{\Psi}_n, \bm{\Phi}_{m,i}}_{D^k} - \LRa{\avg{\bm{A}_n\bm{\Psi}_i}, {\bm{\Phi}_i}}_{\partial D^k} \qquad 1\leq m \leq N^C, \quad 1\leq n \leq N^{NC} \\
+\bm{C}_{mn}  &= -\sum_{D^k\in \Oh} \LRa{\avg{\bm{A}_n\bm{\Psi}_i}, {\bm{\Psi}_i}}_{\partial D^k}, \qquad 1 \leq m,n \leq N^{NC}\\
+\bm{S}_{mn}  &= \frac{1}{2}\sum_{D^k\in \Oh} \LRa{\bm{A}_n^T\jump{\bm{A}_n \bm{\Psi_n}}, {\bm{\Psi}_m}}_{\partial D^k}, \qquad 1 \leq m,n \leq N^{NC}.  
+\end{align*}
+For exact quadrature, constant $\bm{A}_i$, and $\tau = 0$, the DG formulation is skew-symmetric \cite{warburton2013low, chan2015gpu, kopriva2014energy}, implying that $\bm{A},\bm{C}$ are skew-symmetric and that $\tilde{\bm{B}} = -\bm{B}$.
+Additionally, the entries of the penalization matrix $\bm{S}$ can be rewritten as
+\[
+\bm{S}_{mn}  &= \frac{1}{2}\sum_{D^k\in \Oh} \LRa{\jump{\bm{A}_n \bm{\Psi}_n}, \bm{A}_n{\bm{\Psi}_m}}_{\partial D^k} = -\frac{1}{2}\sum_{f \in \Gh} \LRa{\jump{\bm{A}_n \bm{\Psi}_n}, \jump{\bm{A}_n{\bm{\Psi}_m}}}_{f},
 \]
-We first consider the spectra of $\bm{C} + \tau\bm{S}$.
-By the construction of the DG formulation $\bm{C}$ is skew-symmetric, while $\bm{S}$ is symmetric and negative-definite.
-Let $\lambda, \bm{u}$ be an eigenpair of $\bm{C} + \tau{\bm{S}}$ corresponding to the largest $N^D$ eigenvalues.
-Let $\lambda = \alpha + i\beta$ and $\bm{u} = \bm{v} + i\bm{w}$, where $\alpha,\beta$ and $\bm{v},\bm{w}$ are the real and imaginary parts of $\lambda$ and $\bm{u}$, respectively.
+implying that $\bm{S}$ is symmetric and negative-definite.
+
+%The generalized eigenvalue problem $\bm{K}\bm{u}=\lambda_i\bm{M}\bm{u}$ is then equivalent to the eigenvalue problem $\bm{K}\bm{u} = \lambda_i\bm{u}$.  
+%We begin by considering the spectra of $\bm{C} + \tau\bm{S}$.  
+%Let $\lambda, \bm{u}$ be an eigenpair of $\bm{C} + \tau{\bm{S}}$, such that $\lambda = \alpha + i\beta$ and $\bm{u} = \bm{v} + i\bm{w}$, where $\alpha,\beta$ and $\bm{v},\bm{w}$ are the real and imaginary parts of $\lambda$ and $\bm{u}$, respectively.
 Then, expanding and grouping terms in
-\[
-(\bm{C}+\tau\bm{S})(\bm{v} + i\bm{w}) = (\alpha + i\beta) (\bm{v} + i\bm{w})
-\]
-we have that
-\begin{align*}
-(\bm{C}+\tau\bm{S})\bm{v} &= \alpha\bm{v}-\beta\bm{w}\\
-(\bm{C}+\tau\bm{S})\bm{w} &= \beta\bm{v}+\alpha\bm{w}.
-\end{align*}
-Assuming that $\nor{\bm{u}}^2 = \nor{\bm{v}}^2 + \nor{\bm{w}}^2 = 1$, multiplying both sides by $\bm{v}^T,\bm{w}^T$ and straightforward manipulations %yield
+%\[
+%(\bm{C}+\tau\bm{S})(\bm{v} + i\bm{w}) = (\alpha + i\beta) (\bm{v} + i\bm{w})
+%\]
+%and using the fact that $\bm{C} + \tau\bm{S}$ is real-valued, we have that
 %\begin{align*}
-%\bm{v}^T\bm{A}\bm{v} &= \alpha\nor{\bm{v}}^2-\beta\bm{v}^T\bm{w}\\
-%\bm{w}^T\bm{A}\bm{w} &= \beta\bm{w}^T{\bm{v}}+\alpha\nor{\bm{w}}^2.  
+%(\bm{C}+\tau\bm{S})\bm{v} &= \alpha\bm{v}-\beta\bm{w}\\
+%(\bm{C}+\tau\bm{S})\bm{w} &= \beta\bm{v}+\alpha\bm{w}.
 %\end{align*}
+%%Assuming that $\nor{\bm{u}}^2 = \nor{\bm{v}}^2 + \nor{\bm{w}}^2 = 1$, multiplying both sides by $\bm{v}^T,\bm{w}^T$ and straightforward manipulations %yield
+%%\begin{align*}
+%%\bm{v}^T\bm{A}\bm{v} &= \alpha\nor{\bm{v}}^2-\beta\bm{v}^T\bm{w}\\
+%%\bm{w}^T\bm{A}\bm{w} &= \beta\bm{w}^T{\bm{v}}+\alpha\nor{\bm{w}}^2.  
+%%\end{align*}
+%%\begin{align*}
+%%\bm{v}^T\bm{A}\bm{v} + \bm{w}^T\bm{A}\bm{w} &= \alpha\\
+%%\bm{v}^T\bm{A}\bm{w} - \bm{w}^T\bm{A}\bm{w} &= \beta.
+%%\end{align*}
+%The symmetry of $\bm{S}$ then yields
 %\begin{align*}
-%\bm{v}^T\bm{A}\bm{v} + \bm{w}^T\bm{A}\bm{w} &= \alpha\\
-%\bm{v}^T\bm{A}\bm{w} - \bm{w}^T\bm{A}\bm{w} &= \beta.
+%\bm{v}^T\bm{C}\bm{w} - \bm{w}^T\bm{C}\bm{w} &= \beta\\
+%%\LRp{\bm{v}^T\bm{C}\bm{v} + \bm{w}^T\bm{C}\bm{w}} + 
+%\tau\LRp{\bm{v}^T\bm{S}\bm{v} + \bm{w}^T\bm{S}\bm{w}} &= \alpha.
 %\end{align*}
-using the skew-symmetry of $\bm{C}$ and symmetry of $\bm{S}$ yields
-\begin{align*}
-\bm{v}^T\bm{C}\bm{w} - \bm{w}^T\bm{C}\bm{w} &= \beta\\
-\tau\LRp{\bm{v}^T\bm{S}\bm{v} + \bm{w}^T\bm{S}\bm{w}} &= \alpha.
-\end{align*}
-Since $\bm{v}^T\bm{C}\bm{w} - \bm{w}^T\bm{C}\bm{w}= 2\bm{w}^T\bm{C}\bm{v} = \beta$ is independent of $\tau$ and $\bm{v},\bm{w}$ are normalized, $\beta$ remains bounded as $\tau\rightarrow \infty$.
+%Since $\bm{v}^T\bm{C}\bm{w} - \bm{w}^T\bm{C}\bm{w}= \beta$ is independent of $\tau$ and $\bm{v},\bm{w}$ are normalized, $\beta$ must remain bounded as $\tau\rightarrow \infty$.
 Similarly, since $\bm{S}$ is symmetric, negative-definite, and independent of $\tau$, the quantity
-\[
-\LRb{\bm{v}^T\bm{S}\bm{v} + \bm{w}^T\bm{S}\bm{w}} = \LRb{\frac{\alpha}{\tau}}
-\]
-is bounded independently of $\tau$, and we conclude that $\LRb{\alpha} = O(\tau)$.
+%\[
+%\LRb{\bm{v}^T\bm{S}\bm{v} + \bm{w}^T\bm{S}\bm{w}} = \LRb{\frac{\alpha}{\tau}}
+%\]
+%is bounded independently of $\tau$, and we conclude that $\LRb{\alpha} = O(\tau)$.
 Thus, as $\tau$ increases, the eigenvalues of $\bm{C} + \tau\bm{S}$ are shifted further and further left of the imaginary axis.
@@
 We may now show how the spectra of $\bm{K}$ behaves as $\tau \rightarrow \infty$.
-Following \cite{Warburton20063205}, let $\bm{U}$ and $\bm{Q}$ be unitary matrices whose columns contain the eigenvectors of $\bm{A}$ and $\bm{C} + \tau \bm{S}$, respectively.
-Then, applying a block diagonal similarity transform yields
+Since $\bm{A}$ is skew-symmetric and $\bm{S}$ is symmetric, they are diagonalizable under unitary matrices $\bm{U}$ and $\bm{Q}$, whose columns contain the eigenvectors of $\bm{A}$ and $\bm{S}$, respectively.
+%\note{FIX $C + \tau S$ Gerschgorin part}
+Following \cite{Warburton20063205}, we apply a block diagonal similarity transform to yield
 \[
-\left(\begin{array}{cc}
+\tilde{\bm{K}} = \left(\begin{array}{cc}
 \bm{U}^* & \\
@@
  = \left(\begin{array}{cc}
-\bm{\Lambda}_C & \bm{U}^T\bm{B}\bm{Q}\\
--\bm{Q}^T\bm{B}^T\bm{U} & \bm{\Lambda}_D
+\bm{\Lambda}^C & \bm{U}^*\bm{B}\bm{Q}\\
+-\bm{Q}^*\bm{B}^T\bm{U} & \bm{Q}^*\bm{C}\bm{Q} + \tau \bm{\Lambda}^{S}
 \end{array}\right),
 \]
-where $\bm{\Lambda}_C,\bm{\Lambda}_D$ are diagonal matrices whose entries consist of eigenvalues of $\bm{A},\bm{C} + \tau\bm{S}$, respectively.
-Since $\bm{B}$ is independent of $\tau$, $\nor{\bm{U}^T\bm{B}\bm{Q}}$ can be bounded independently of $\tau$, assuming that $\bm{U}$ and $\bm{Q}$ are normalized.
-Gerschgorin's theorem applied to $\bm{\Lambda}_C,\bm{\Lambda}_D$ then implies that the eigenvalues of $\bm{K}$ are contained in two sets of discs with radii independent of $\tau$.
-The first set of discs are centered around $\lambda^C_i$ for $i = 1,\ldots,N^C$, while the second set of discs are centered around $\lambda^D_i$ for $i = 1,\ldots,N^D$, where ${\rm Re}(\lambda^D_i) = O(\tau)$.
+where $\bm{\Lambda}^C,\bm{\Lambda}^{S}$ are diagonal matrices whose entries consist of eigenvalues of $\bm{A}$ and $\bm{S}$.
+
+Since $\bm{Q}^*\bm{C}\bm{Q}$ is skew-symmetric and its diagonal has zero real part, $\bm{Q}^*\bm{C}\bm{Q} + \tau \bm{\Lambda}^{S}$ has diagonal entries $\tau \lambda^S_j + \mi\gamma_j$, where $\lambda^S_j$ is real and negative and $\gamma_j$ is independent of $\tau$.
+Thus, the diagonal of $\tilde{\bm{K}}$ consists of $N^C$ purely imaginary values and $N^{NC}$ values of the form $\tau \lambda^S_j + \mi\gamma_j$.
+Since $\bm{B}$ is independent of $\tau$, the entries of $\bm{U}^*\bm{B}\bm{Q}$ are independent of $\tau$, assuming that $\bm{U}$ and $\bm{Q}$ are normalized.
+Gerschgorin's theorem applied to $\tilde{\bm{K}}$ then implies that the eigenvalues of $\bm{K}$ are contained in two sets of discs with radii independent of $\tau$.
+The first set of discs are centered around $\lambda^C_j$ for $j = 1,\ldots,N^C$, while the second set of discs are centered around $\tau\lambda^S_j + \mi\gamma_j$ for $j = 1,\ldots,N^{NC}$.
+For sufficiently large $\tau$, these disks must become disjoint, implying that the union of disks centered around  $\lambda^C_i$ contains exactly $N^C$ eigenvalues, while the union of disks centered around  $\lambda^{NC}_i$ contain exactly $N^{NC}$ eigenvalues \cite{horn2012matrix}.
+In other words, the spectra of $\bm{K}$ diverges into two distinct sets of $N^C$ and $N^{NC}$ eigenvalues each as $\tau \rightarrow \infty$, with the set of $N^{NC}$ eigenvalues containing negative real parts with magnitude $O(\tau)$.
 
@@
 By the Gerschgorin argument, these must remain bounded as $\tau \rightarrow \infty$.
-Let $\bm{W} = (\bm{W}_C,\bm{W}_D)$ and $\bm{\Lambda}_C$ be the matrix of eigenvectors and eigenvalues corresponding to these $N^C$ smallest eigenvalues, such that
+Let $\bm{W} = (\bm{W}^C,\bm{W}^{NC})$ and $\bm{\Lambda}^C$ be the matrix of eigenvectors and eigenvalues corresponding to these $N^C$ smallest eigenvalues, such that
 \[
@@
 \left(\begin{array}{c}
-\bm{W}_C
+\bm{W}^C
 \\
-\bm{W}_D
+\bm{W}^{NC}
 \end{array}\right) = 
 \left(\begin{array}{c}
-\bm{A}\bm{W}_C + \bm{B}\bm{W}_D\\
--\bm{B}^T\bm{W}_C + \bm{C}\bm{W}_D + \tau \bm{S}\bm{W}_D
+\bm{A}\bm{W}^C + \bm{B}\bm{W}^{NC}\\
+-\bm{B}^T\bm{W}^C + \bm{C}\bm{W}^{NC} + \tau \bm{S}\bm{W}^{NC}
 \end{array}\right)
@@
 \left(\begin{array}{c}
-\bm{W}_C
+\bm{W}^C
 \\
-\bm{W}_D
+\bm{W}^{NC}
 \end{array}\right) 
@@
 \[
-\tau \nor{\bm{S}\bm{W}_D} = \nor{\bm{\Lambda}_C \bm{W}_D + \bm{B}^T\bm{W}_C  - \bm{C}\bm{W}_D}.
+\tau \nor{\bm{S}\bm{W}^{NC}} = \nor{\bm{\Lambda}^C \bm{W}^{NC} + \bm{B}^T\bm{W}^C  - \bm{C}\bm{W}^{NC}}.
 \]
-Since $\Lambda_C$ remains bounded as $\tau\rightarrow \infty$, the right hand side is bounded independently of $\tau$ under normalization of $\bm{W}_C,\bm{W}_D$, implying that the non-conforming component $\bm{W}^D$ satisfies
+Since $\bm{\Lambda}^C$ remains bounded as $\tau\rightarrow \infty$, the right hand side is bounded independently of $\tau$ under normalization of $\bm{W}^C,\bm{W}^{NC}$, implying that the non-conforming component $\bm{W}^{NC}$ satisfies
 \[
-C\nor{\bm{W}_D} \leq \nor{\bm{S}\bm{W}_D} = O(1/\tau).
+\sqrt{\lambda^{NC}_{\min}}\nor{\bm{W}^{NC}} \leq \nor{\bm{S}\bm{W}^{NC}} = O(1/\tau)
 \]
-As a consequence, $\bm{W}_D \rightarrow 0$ as $\tau\rightarrow \infty$, and the smallest $N^C$ eigenvalues of $\bm{K}$ converge to the purely imaginary eigenvalues of $\bm{A}$ at a rate of $O(1/\tau)$.
+where $\lambda^{NC}_{\min} > 0$ is the smallest eigenvalue of $\bm{S}$.
+As a consequence, $\bm{W}^{NC} \rightarrow 0$ as $\tau\rightarrow \infty$, and the smallest $N^C$ eigenvalues of $\bm{K}$ converge to the eigenvalues of $\bm{A}$ at a rate of $O(1/\tau)$.
 These correspond to a discretization using the conforming approximation space (\ref{eq:conf}).
+
+These results are summarized in the following lemma:
+\begin{lemma}
+As $\tau \rightarrow \infty$, the spectrum of the DG discretization matrix $\bm{K}$ decouples into two sets of eigenvalues $\LRc{\lambda^C_1,\ldots,\lambda^C_{N^C}}$ and $\LRc{\lambda^{NC}_1,\ldots,\lambda^{NC}_{N^{NC}}}$.
+The eigenvalues $\LRc{\lambda^{NC}_1,\ldots,\lambda^{NC}_{N^{NC}}}$ diverge towards the left half plane, with $\LRb{{\rm Re}(\lambda_i)} = O(\tau)$, while  the eigenvalues $\LRc{\lambda^C_1,\ldots,\lambda^C_{N^C}}$ converge to the eigenvalues of $\bm{A}$ at the rate $O(1/\tau)$.
+\label{lemma:eig}
+\end{lemma}
+We note that, while we consider skew-symmetric DG formulations in this work, Lemma~\ref{lemma:eig} is straightforward to generalize to any DG formulations where $\bm{S}$ is symmetric negative definite and $\bm{A}$ is diagonalizable.
 
@@
 The above analysis illustrates the behavior of the spectra of the DG discretization for the asymptotic cases when $\tau = 0$ or $\tau \rightarrow \infty$.
-However, it is less clear how the spectra of $\bm{K}$ behaves for $\tau = O(1)$.
+However, it is less clear how the spectra of $\bm{K}$ behaves for $\tau \approx 1$.
 We rely instead on numerical experiments in one and two space dimensions to illustrate behaviors for the advection and acoustic wave equations.
+In particular, these experiments indicate that certain modes with negative real part for $\tau = O(1)$ return to the imaginary axis as $\tau \rightarrow \infty$.
+Since eigenvalues with a negative real part correspond to dissipation in time-domain simulations, this implies that these under-resolved modes become undamped as $\tau$ grows sufficiently large.
 
@@
 These eigenvalues which move right to return to the imaginary axis correspond to a second set of \emph{spurious} modes.
-These spurious modes, corresponding to under-resolved modes of the exact underlying operator, consist of higher frequency components with sharp peaks and have previously been observed in high order $C^0$ finite element methods \cite{ainsworth2014dispersive,hughes2014finite}.
+These modes are related to spurious modes observed for high order $C^0$ finite element methods \cite{ainsworth2014dispersive}, which consist of higher frequency components with sharp peaks.
+Since the real part of these eigenvalues approaches zero as $\tau\rightarrow \infty$, these modes become undamped and persist as spurious solution components in time-domain simulations \cite{hughes2014finite}
 
@@
 Figure~\ref{fig:waveeigs} shows the spectra of the DG discretization matrix for the acoustic wave equation for $\tau = 1, 10, 50$.
-Like the one-dimensional case, a subset of eigenvalues diverge towards the left half plane as $\tau$ increases.
-However, unlike the one-dimensional case, these eigenvalues collapse towards the real line as $\tau$ increases.
+Like the one-dimensional case, a subset of eigenvalues diverge towards the left half plane as $\tau$ increases, though unlike the one-dimensional case, these eigenvalues collapse towards the real line as $\tau$ increases.
+Similarly, as predicted, a subset of eigenvalues returns towards the imaginary axis as $\tau \rightarrow \infty$.
+These are shown in Figure~\ref{fig:trackEigsWave}, along with traced paths taken by each eigenvalue as $\tau$ increases.
 
@@
 \end{figure}
-
-Similarly, as predicted, a subset of eigenvalues returns towards the imaginary axis as $\tau \rightarrow \infty$.
-These are shown in Figure~\ref{fig:trackEigsWave}, along with traced paths taken by each eigenvalue as $\tau$ increases.
 
@@
 As $\tau$ increases and the eigenvalue approaches the imaginary axis, the eigenmode approaches a $C^0$ continuous function with a high frequencies and sharp peak, similar to the spurious eigenmodes observed for the one-dimensional case in Figure~\ref{fig:spurious}.
+Likewise, the non-zero velocity components of spurious eigenmodes converge to approximations in $H({\rm div}; \Oh)$, such that only the normal component of velocity is continuous across faces.
+Figure~\ref{fig:trackModesWaveU} shows visualizations of these eigenmodes.
+For large $\tau$, normal jumps of the velocity component vanish.
+% approach an $H({\rm div};\Oh)$-conforming approximation.
 
@@
 \end{figure}
-
-The non-zero velocity components of spurious eigenmodes converge to approximations in $H({\rm div}; \Oh)$, such that only the normal component of velocity is continuous across faces.
-Figure~\ref{fig:trackModesWaveU} shows visualizations of these eigenmodes.
-For large $\tau$, normal jumps of the velocity component vanish.
-% approach an $H({\rm div};\Oh)$-conforming approximation.
 
@@
 
-Unlike spurious conforming modes present for the acoustic wave equation, spurious conforming modes for the advection equation satisfy $\jump{\beta_n u} = 0$, and allow discontinuities along faces which are tangential to flow directions.
-\note{Find cite on regularity of advection solutions.}
+Unlike spurious conforming modes present for the acoustic wave equation, spurious conforming modes for the advection equation satisfy $\jump{\beta_n u} = 0$, and allow discontinuities along faces which are tangential to flow directions, as shown in Figure~\ref{fig:trackModesAdvecU}.
+%\note{Find cite on regularity of advection solutions.}
 
@@
 
-\note{Penalized DG methods as sweet spot treatments of spurious modes, somewhere between central DG and conforming discretizations.}
+For symmetric linear hyperbolic systems of PDEs, penalty fluxes are an alternative to upwind fluxes for DG methods.
+Both penalty and upwind fluxes can be interpreted as weakly enforcing continuity conditions implied by the underlying PDE.
+As the penalty parameter $\tau$ increases, we show that the eigenvalues of the DG discretization matrix split into two sets, with one set of eigenvalues converging to eigenvalues of a conforming discretization and the other set diverging with real part approaching $-\infty$.
+
+Numerical experiments show that, for $\tau = O(1)$, there exist eigenvalues with negative real part which return to the imaginary axis as $\tau \rightarrow 0$ or $\tau\rightarrow \infty$.
+The corresponding eigenmodes of such eigenvalues can be interpreted as spurious under-resolved modes which persist in time-domain simulations.
+These results suggest that the use of upwind and penalty fluxes damps spurious modes present not only for $\tau = 0$ (central fluxes), but also for $\tau \rightarrow\infty$ (a conforming discretization).
 

### 2016-11-30 docs/elasticity_notes/main.tex
adding elasticity paper

\pdfoutput=1

%\documentclass[preprint,10pt]{elsarticle}
%\documentclass[10pt]{article}
%\documentclass[review]{siamart0216}
\documentclass{siamart0216}

\usepackage{fullpage}
\usepackage{hyperref}
%\let\proof\relax
%\let\endproof\relax
\usepackage{amsmath,amssymb,amsfonts,mathrsfs}%,amsthm}
\usepackage[titletoc,toc,title]{appendix}

%\usepackage{lineno}
\usepackage{array} 
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{mathtools}
\usepackage{pdfpages}
\usepackage[textsize=footnotesize,color=green]{todonotes}
\usepackage{bm}
%\usepackage{tikz}
\usepackage[normalem]{ulem}
\usepackage{hhline}

%% ====================================== alg package
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{algorithmicx}
\algblock{ParFor}{EndParFor}
% customising the new block
\algnewcommand\algorithmicparfor{\textbf{parfor}}
\algnewcommand\algorithmicpardo{\textbf{do}}
\algnewcommand\algorithmicendparfor{\textbf{end\ parfor}}
\algrenewtext{ParFor}[1]{\algorithmicparfor\ #1\ \algorithmicpardo}
\algrenewtext{EndParFor}{\algorithmicendparfor}
%% ====================================== end alg package

\usepackage{graphicx}
\usepackage{subfig}
\usepackage{color}

%% ====================================== graphics

\usepackage{pgfplots}
\usepackage{pgfplotstable}
\definecolor{markercolor}{RGB}{124.9, 255, 160.65}
\pgfplotsset{width=10cm,compat=1.9}
\pgfplotsset{
tick label style={font=\small},
label style={font=\small},
legend style={font=\small}
}

\usetikzlibrary{calc}

\usetikzlibrary{calc}

%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
\newcommand{\logLogSlopeTriangle}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        \pgfmathsetmacro{\xArel}{#1}
        \pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{\yArel}
        \pgfmathsetmacro{\xCrel}{\xArel}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=north] {1}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=west] {#4}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

\newcommand{\logLogSlopeTriangleNeg}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        \pgfmathsetmacro{\xArel}{#1}
        \pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{\yArel}
        \pgfmathsetmacro{\xCrel}{\xArel}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=.5,anchor=south] {1}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=west] {#4}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
\newcommand{\logLogSlopeTriangleFlipNeg}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        %\pgfmathsetmacro{\xArel}{#1}
        %\pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{#3}
        \pgfmathsetmacro{\xCrel}{#1}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

	\pgfmathsetmacro{\xArel}{\xBrel}
        \pgfmathsetmacro{\yArel}{\yCrel}

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=east] {#4}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=north] {1}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
\newcommand{\logLogSlopeTriangleFlip}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        %\pgfmathsetmacro{\xArel}{#1}
        %\pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{#3}
        \pgfmathsetmacro{\xCrel}{#1}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

	\pgfmathsetmacro{\xArel}{\xBrel}
        \pgfmathsetmacro{\yArel}{\yCrel}

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=east] {#4}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=south] {1}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

\usepackage{stmaryrd}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

\newcommand{\vect}[1]{\ensuremath\boldsymbol{#1}}
\newcommand{\tensor}[1]{\underline{\bm{#1}}}
\newcommand{\del}{\triangle}
\newcommand{\curl}{\grad \times}
\renewcommand{\div}{\grad \cdot}
\newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}{\rm #2}}}
\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
\newcommand{\pdd}[2]{\frac{\partial^2#1}{\partial#2^2}}

\newcommand{\bs}[1]{\boldsymbol{#1}}

\newcommand{\equaldef}{\stackrel{\mathrm{def}}{=}}

\newcommand{\mi}{\mathrm{i}} %% roman "i"
\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRb}[1]{\left| #1 \right|}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}
\newcommand{\LRceil}[1]{\left\lceil #1 \right\rceil}
\newcommand{\LRl}[1]{\left. #1 \right|}

%\newcommand{\cond}[1]{\kappa\LRp{#1}}
\newcommand{\cond}[2]{\nor{#1}_{#2}\nor{{#1}^{-1}}_{#2}}

\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\Div} {\ensuremath{\nabla\cdot}}
\newcommand{\jump}[1] {\ensuremath{\llbracket#1\rrbracket}}
\newcommand{\avg}[1] {\ensuremath{\LRc{\!\{#1\}\!}}}

\renewcommand{\Oh}{{\Omega_h}}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\Lk}{L^2\LRp{D^k}}
\newcommand{\Ldk}{L^2\LRp{\partial D^k}}
\newcommand{\Dhat}{\widehat{D}}
\newcommand{\Lhat}{L^2\LRp{\Dhat}}

%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{lemma}[theorem]{Lemma}
%\newtheorem{proposition}[theorem]{Proposition}
%\newtheorem{corollary}[theorem]{Corollary}

%\newenvironment{definition}[1][Definition]{\begin{trivlist}
%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
%\newenvironment{example}[1][Example]{\begin{trivlist}
%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\newcommand{\eval}[2][\right]{\relax
  \ifx#1\right\relax \left.\fi#2#1\rvert}

\def\etal{{\it et al.~}}

\newcommand{\note}[1]{{\color{blue}#1}}
\newcommand{\remark}[1]{\textbf{\color{red}#1}}

\newcommand{\LinfDk}{L^{\infty}\LRp{D^k}}

\newcommand{\diag}[1]{{\rm diag}\LRp{#1}}

\newcommand{\half}{1/2}

\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

%% d in integrand
\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}}

\makeatletter
\renewcommand\d[1]{\mspace{6mu}\mathrm{d}#1\@ifnextchar\d{\mspace{-3mu}}{}}
\makeatother

\date{}
\author{Jesse Chan}
\title{Weight-adjusted discontinuous Galerkin methods: matrix-valued weights with application to elastic wave propagation}

\begin{document}

\maketitle
%\tableofcontents

\begin{abstract}
Weight-adjusted inner products \cite{chan2016weight1,chan2016weight2} replace weighted $L^2$ inner products with easily invertible weight-adjusted approximations.  When paired with discontinuous Galerkin (DG) methods, the result is a low storage, energy stable, and (for sufficiently regular weights) high order accurate method for wave propagation in the presence of general heterogeneous media and curvilinear meshes.  In this work, we extend weight-adjusted DG (WADG) methods to weighted inner products with matrix-valued weights, focusing on the linear elastic wave equation as an application.  We present a new DG formulation based on the discretization of the symmetric form of the linear elasticity wave equation, with upwind-like dissipation incorporated through simple penalty terms in the numerical fluxes.  This results in provably stable formulations for arbitrary anisotropic stiffness tensors and curvilinear meshes.  A convergence analysis is given, and numerical results confirm the stability and high order accuracy of WADG for several problems in elastic wave propagation.  
\end{abstract}

\section{Introduction}

There exist many high order energy stable discontinuous Galerkin formulations for the elastic wave equations in first order form.  Wilcox et al.\ introduced a discontinuous Galerkin spectral element method on hexahedral meshes

SEM \cite{komatitsch2010high,wilcox2010high}

High order DG \cite{kaser2006arbitrary, dumbser2006arbitrary, de2007arbitrary,ye2016discontinuous}

Interior penalty formulations.  Analysis: Riviere et al.\ \cite{riviere2003discontinuous,riviere2007discontinuous} and Delcourte et al.\ \cite{delcourte2009high,delcourte2015analysis}, dispersion analysis \cite{de2008interior,antonietti2016high}.

GPU-accelerated DG methods \cite{klockner2009nodal,godel2010gpu,godel2010scalability}.  Also elastic wave propagation and seismic wave applications \cite{modave2015nodal,modave2016gpu}

High order approximations of material data \cite{castro2010seismic,mercerat2015nodal}, but not low storage or not energy stable.

Extending mass lumping without dispersive effects \cite{christon1999influence,guermond2013correction}.  Reduces to exact mass matrix inversion for constant weights.

A simpler extension of WADG \cite{chan2016weight1,chan2016weight2} to elasticity and weight-adjusted inner products for vector-valued functions.

Outline: present elasticity and an energy stable DG method, motivate need for alternative to inverting large dense coupled mass matrix, introduce weight-adjusted inner products.

\section{Symmetric form of the elastic wave equation }

We begin with the linear elastic wave equation under small deformations in a domain $\Omega$.  These equations can be written a first order velocity-stress  system 
\begin{align*}
\rho \pd{\bm{v}}{t} &= \Div{\bm{S}}\\
\pd{\bm{S}}{t} &= \frac{1}{2} \bar{\bm{C}}\LRp{\Grad\bm{v} + \Grad\bm{v}^T},
\end{align*}
where $\bm{v}$ is the velocity, $\bm{S}$ is the stress tensor, $\rho$ is density, and $\bar{\bm{C}}$ is the symmetric constitutive stiffness tensor relating stress and strain.  These equations can also be rewritten as a symmetric hyperbolic system of PDEs \cite{hughes1978classical}  
\begin{align}
\rho \pd{\bm{v}}{t} &= \sum_{i=1}^d \bm{A}_i^T \pd{\bm{\sigma}}{\bm{x}_i}\nonumber \\
\bm{C}^{-1} \pd{\bm{\sigma}}{t} &= \sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i}.
\label{eq:symelas}
\end{align}
where $\bm{C}$ is the symmetric matrix form of the constitutive tensor $\bar{\bm{C}}$ and $\sigma$ is a vector of length $N_d = \frac{d(d+1)}{2}$, the number of unique entries of the stress tensor $\bm{S}$.  We note that the matrices $\bm{A}_i$ are constant, while $\rho$ and $\bm{C}$ can vary spatially.  Furthermore, we will assume that $\rho$ and $\bm{C}$ are positive-definite and bounded such that 
\begin{align*}
\rho_{\min} &\leq \rho(\bm{x}) \leq \rho_{\max} \\
0 < c_{\min} \leq \bm{x}^T\bm{C}\bm{x} &\leq c_{\max} < \infty, \qquad \bm{x} \in \mathbb{R}^{N_d},
\end{align*}
which also implies that $\bm{C}^{-1}$ is positive-definite and bounded.

In two dimensions, $\bm{v} = (\bm{v}_1, \bm{v}_2)^T$ and $\bm{\sigma}= (\sigma_{xx},\sigma_{yy},\sigma_{xy})^T$ 
\[
\bm{S} = \LRp{\begin{array}{cc}
\sigma_{xx} & \sigma_{xy}\\
\sigma_{xy} & \sigma_{yy}
\end{array}}.
\]
The matrices $\bm{A}_i$ are given as
\[
\bm{A}_1 = \left(\begin{array}{cc}
1 & 0\\
0 & 0\\
0 & 1
\end{array}\right), \qquad 
\bm{A}_2 = \left(\begin{array}{cc}
0 & 0\\
0 & 1\\
1 & 0
\end{array}\right).
\]

In three dimensions, the velocity is $\bm{v} = (\bm{v}_1,\bm{v}_2,\bm{v}_3)^T$, while $\bm{\sigma}= (\sigma_{xx},\sigma_{yy},\sigma_{zz},\sigma_{yz},\sigma_{xy},\sigma_{xz})^T$ denotes the unique entries of the stress tensor $\bm{S}$
\[
\bm{S} = \LRp{\begin{array}{ccc}
\sigma_{xx} & \sigma_{xy} & \sigma_{xz}\\
\sigma_{xy} & \sigma_{yy}& \sigma_{yz}\\
\sigma_{xz} & \sigma_{yz}& \sigma_{zz}
\end{array}}.
\]
The matrices $\bm{A}_i$ are then
\[
\bm{A}_1 = \left(\begin{array}{ccc}
1 & 0 & 0\\
0 & 0& 0\\
0 & 0& 0\\
0 & 0& 0\\
0 & 0 & 1\\
0 & 1 & 0
\end{array}\right), \qquad 
\bm{A}_2 = \left(\begin{array}{ccc}
0 & 0 & 0\\
0 & 1 & 0\\
0 & 0 & 0\\
0 & 0 & 1\\
0 & 0 & 0\\
1 & 0 & 0
\end{array}\right), \qquad
\bm{A}_3 = \left(\begin{array}{ccc}
0 & 0 & 0\\
0 & 0 & 0\\
0 & 0 & 1\\
0 & 1 & 0\\
1 & 0 & 0\\
0 & 0 & 0
\end{array}\right)
\]

In general anisotropic media, $\bm{C}$ is symmetric and positive-definite.  For two-dimensional isotropic media, $\bm{C}$ is given as
\[
\bm{C}= \LRp{
\begin{array}{ccc}
2\mu+\lambda  &  \lambda  &     0\\
\lambda &       2\mu+\lambda &      0\\
          0 &      0           &     {\mu}
\end{array}}, 
\]
where $\lambda,\mu$ are Lame parameters.  For three-dimensional isotropic media, $\bm{C}$ is given instead by
\[
\bm{C}= \LRp{
\begin{array}{cccccc}
2\mu+\lambda  &  \lambda  &     \lambda  & & & \\
\lambda &   2\mu+\lambda & \lambda   & & & \\
\lambda & \lambda &   2\mu+\lambda &    & &\\
 &   &    & {\mu} & &\\
&  &   &    & {\mu} &\\
& &  &   &    & {\mu} \\
\end{array}}.  
\]
We will consider both spatially varying isotropic and anisotropic media in this work.

\section{An energy stable discontinuous Galerkin variational formulation}

The advantage of the symmetric first order formulation of the elastic wave equations (\ref{eq:symelas}) is that it is straightforward to derive an energy stable discontinuous Galerkin formulation.  We assume that the domain $\Omega$ is Lipschitz and exactly triangulated by a mesh $\Omega_h$, which consists of elements $D^k$.  We further assume that each element $D^k$ is the image of a reference element $\widehat{D}$ under the local elemental mapping 
\[
\bm{x}^k = \bm{\Phi}^k \widehat{\bm{x}},
\]
where $\bm{x}^k = \LRc{x^k,y^k,z^k}$ denote physical coordinates on $D^k$ and $\widehat{\bm{x}} = \LRc{\widehat{x},\widehat{y},\widehat{z}}$ denote coordinates on the reference element.  We denote the determinant of the Jacobian of $\bm{\Phi}^k$ as $J^k$.

We will approximate components over each element $D^k$ from an approximation space $V_h\LRp{D^k}$, which we defined as
\[
V_h\LRp{D^k} = \bm{\Phi}^k \circ V_h\LRp{\widehat{D}}.
\]
where $V_h\LRp{\widehat{D}}$ is an approximation space defined on the reference element.  For the remainder of this work, we will take $V_h\LRp{\widehat{D}} = P^N\LRp{\widehat{D}}$, where $P^N(\LRp{\widehat{D}})$ is the polynomial space of total degree $N$ on the reference simplex.  In two dimensions, 
\[
P^N\LRp{\widehat{D}} = \LRc{ \widehat{x}^i \widehat{y}^j, \quad 0 \leq i + j \leq N},
\] 
while in three dimensions, 
\[
P^N\LRp{\widehat{D}} = \LRc{ \widehat{x}^i \widehat{y}^j \widehat{z}^k, \quad 0 \leq i + j +k \leq N}.  
\]

We denote the $L^2$ inner product and norm over $D^k$ by $\LRp{\cdot,\cdot}_{L^2\LRp{D^k}}$, such that
\[
\LRp{\bm{f},\bm{g}}_{\Lk} = \int_{D^k} \bm{f}\cdot\bm{g} \diff x = \int_{\widehat{D}} \bm{f}\cdot\bm{g} J^k, \qquad \nor{\bm{f}}_{\Lk}^2 = \LRp{\bm{f},\bm{f}}_{\Lk},
\]
where $\bm{f},\bm{g}$ are vector-valued functions.  Global $L^2$ inner products and norms are using local $L^2$ inner products and norms
\[
\LRp{\bm{f},\bm{g}}_{\Oh} = \sum_{D^k\in \Oh} \int_{D^k} \LRp{\bm{f},\bm{g}}_{\Lk}, \qquad \nor{\bm{f}}_{\Oh}^2 = \sum_{D^k\in \Oh} \nor{\bm{f}}^2_{\Lk}.
\]

Let $f$ be a face of an element $D^{k}$ with neighboring element $D^{k,+}$ and unit outward normal $\bm{n}$.  We define $u^-$ and $u^+$ denote interior and exterior values of a discontinuous function $u$ such that
\[
{u}^- = \LRl{u}_{f \cap D^k}, \qquad {u}^+ = \LRl{u}_{f \cap  \partial D^{k,+}}.  
\]
The jump and average of a scalar function $u\in V_h\LRp{\Omega_h}$ over $f$ are then defined as
\[
\jump{u} = u^+ - u^-, \qquad \avg{u} = \frac{u^+ + u^-}{2}.
\]
Jumps and averages of a vector-valued functions $\bm{u}\in \mathbb{R}^m$ and $\bm{S}\in \mathbb{R}^{m\times n}$ are then defined component-wise 
\[
\LRp{\jump{\bm{u}}}_i = \jump{\bm{u}_i}, \qquad 1\leq i \leq m, \qquad \LRp{\jump{\bm{S}}}_{ij} = \jump{\bm{S}_i}, \qquad 1\leq i \leq m, \quad 1\leq j \leq n.
\]

We can now specify a DG formulation for the linear elastic wave equation (\ref{eq:symelas}).  Symmetric hyperbolic systems readily admit a DG formulation based on penalty fluxes, as detailed in \cite{chan2016short}.  For the linear elastic wave equation in symmetric first order form, this formulation is given as
\begin{align}
\sum_{D^k\in \Oh}\LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= \sum_{D^k\in \Oh}\LRp{\LRp{ \sum_{i=1}^d \bm{A}_i^T\pd{\bm{\sigma}}{\bm{x}_i},\bm{w}}_{\Lk}  + \LRa{\frac{1}{2}\bm{A}_n^T\jump{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk}} \nonumber\\
\sum_{D^k\in \Oh}\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \sum_{D^k\in \Oh}\LRp{\LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk}},
\label{eq:dgform}
\end{align}
where $\tau_v,\tau_\sigma \geq 0$ are penalty parameters and $\bm{A}_n$ is the normal matrix defined on a face $f$ as $\bm{A}_n = \sum_{i=1}^d \bm{n}_i \bm{A}_i$.  In two dimensions, $\bm{A}_n$ is 
\[
\bm{A}_n =  \LRp{\begin{array}{cc}
{n}_x & 0\\
 0 & {n}_y\\
{n}_y & {n}_x
\end{array}}.
\]
while in three dimensions, $\bm{A}_n$ is
\[
\bm{A}_n =  \LRp{\begin{array}{ccc}
{n}_x & 0 & 0\\
 0 & {n}_y & 0\\
0&  0 & {n}_z \\
0 &{n}_z & {n}_y\\
{n}_z & 0 & {n}_x\\
{n}_y & {n}_z & 0\\
\end{array}}.  
\]

\subsection{Boundary conditions}

In this work, we assume boundary conditions on velocity and traction of the form
\[
\bm{v} = \bm{v}_{\rm bc}, \qquad \bm{S}\bm{n} = \bm{t}_{\rm bc}
\]
where $\bm{v}_{\rm bc}$ and $\bm{t}_{\rm bc}$ are given values.  Traction boundary conditions where $\bm{t}_{\rm bc} = 0$ are referred to as free-surface boundary conditions.  We follow \cite{leveque2002finite,wilcox2010high} and impose boundary conditions on the DG formulation through exterior values and jumps of the solution.  Boundary conditions on the normal component of the stress can be imposed by noting that the numerical flux contains the term $\jump{\bm{A}_n^T\bm{\sigma}} = \bm{S}\bm{n}$.

For a face which lies on a boundary, velocity boundary conditions are imposed by setting 
\begin{align*}
\jump{\bm{v}} &= 2\LRp{\bm{v}_{\rm bc} - \bm{v}^- }, \qquad \jump{\bm{A}_n^T\bm{\sigma}}=\jump{\bm{S}\bm{n}} = 0, 
\end{align*}
while traction boundary conditions are imposed by
\begin{align*}
\jump{\bm{A}_n^T\bm{\sigma}} = \jump{\bm{S}\bm{n}} = 2\LRp{\bm{t}_{\rm bc} - \bm{S}^-\bm{n} }= 2\LRp{\bm{t}_{\rm bc} - \bm{A}_n^T\bm{\sigma}^- }, \qquad \jump{\bm{v}} = 0.
\end{align*}
The numerical fluxes are then computed using these modified jumps.

For problems which involve the truncation of infinite or large domains, absorbing boundary conditions are required.  For such cases, we impose simple non-reflective boundary conditions \cite{leveque2002finite} through jumps 
\begin{align*}
\jump{\bm{A}_n^T\bm{\sigma}} = \jump{\bm{S}\bm{n}} = - \bm{S}^-\bm{n} = -\bm{A}_n^T\bm{\sigma}^- , \qquad \jump{\bm{v}} = - \bm{v}^-. 
\end{align*}
We note that more accurate absorbing conditions can be imposed using, for example, perfectly matched layers \cite{berenger1994perfectly} or high order absorbing boundary conditions \cite{hagstrom2004new,modave2016high}.

%Easiest way to impose BCs: set $U^+$ values, compute central fluxes and then compute penalty fluxes via $\bm{A}_n^T\bm{A}_n \jump{U} = \bm{A}_n \jump{F}$.

\subsection{Energy stability}

It is straightforward to show that the formulation (\ref{eq:dgform}) is energy stable for zero velocity and traction boundary conditions, as well as non-reflective boundary conditions.  Integrating by parts the velocity equations yields
\begin{align}
\LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= -\LRp{ \sum_{i=1}^d \bm{\sigma},\bm{A}_i \pd{\bm{w}}{\bm{x}_i}}_{\Lk}  + \LRa{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}} \nonumber\\
\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}} \label{eq:dgform}
\end{align}
Taking $(\bm{w},\bm{q}) = (\bm{v},\bm{\sigma})$ and adding both equations together yields
\begin{align*}
\sum_{D^k\in \Oh}\frac{1}{2}\pd{}{t}&\LRp{\LRp{\rho \bm{v},\bm{v}}_{\Lk} + \LRp{\bm{C}^{-1} \bm{\sigma},\bm{\sigma}}_{\Lk}} =  \\
&\sum_{D^k\in \Omega_h}\LRa{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{v}}_{\partial D^k}+ \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{\sigma}}_{\partial D^k}  \\
&\sum_{D^k\in \Omega_h}\sum_{f\in \partial D^k}\int_{f}{\bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v}} + {\frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma}}.
\end{align*}
Let $\Gamma_h$ denote the set of unique faces in $\Oh$, and let $\Gamma_v,\Gamma_\sigma,\Gamma_{\rm abc}$ denote the parts of the boundary where velocity, traction, and non-reflective boundary conditions are imposed, respectively.  We separate surface terms into contributions from interior shared faces and from boundary faces.  On interior shared faces, 
\begin{align*}
&\sum_{f\in \Gamma_h \setminus \partial \Omega} \int_{f}{\bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v}} + {\frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma}}\\
&\leq - \sum_{f \in\Gamma_h \setminus \partial \Omega} \int_{f} \frac{\tau_v}{2}\LRb{\bm{A}_n\jump{\bm{v}}}^2 + \frac{\tau_\sigma}{2}\LRb{\bm{A}_n^T\jump{\bm{\sigma}}}^2.  
%\label{eq:stability}
\end{align*}
For faces which lie on the boundary $\Gamma_v$ where velocity boundary conditions are imposed, $\jump{\bm{v}} = -2\bm{v}^-$, $\jump{\bm{A}_n^T\bm{\sigma}} = 0$, and $\bm{A}_n^T\avg{\bm{\sigma}} = \bm{Sn}^-$, implying that 
\begin{align*}
&\sum_{f\in \Gamma_v} \int_f \bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v} + \frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma}\\
&= \sum_{f\in \Gamma_v} \int_f \bm{A}_n^T \bm{\sigma}^-\bm{v}^- - \bm{A}_n \bm{v}^- \bm{\sigma}^- - \tau_v \bm{A}_n^T\bm{A}_n \LRb{\bm{v}^-}^2 = - \sum_{f\in \Gamma_v} \int_f \tau_v \bm{A}_n^T\bm{A}_n \LRb{\bm{v}^-}^2 .
\end{align*}
For faces which lie on $\Gamma_\sigma$, $\bm{A}_n^T\jump{\bm{\sigma}} = -2\bm{A}_n^T\bm{\sigma}^-$, $\bm{A}_n^T\avg{\bm{\sigma}} = 0$, and $\jump{\bm{v}} = 0$, yielding a similar contribution
\[
\sum_{f\in \Gamma_\sigma} \int_f \bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v} + \frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma} = - \sum_{f\in \Gamma_\sigma} \int_f \tau_\sigma \bm{A}_n\bm{A}_n^T \LRb{\bm{\sigma}^-}^2.
\]
Finally, for faces in $\Gamma_{\rm abc}$ we have $\bm{A}_n^T\avg{\bm{\sigma}} =\frac{1}{2} \bm{A}_n^T\bm{\sigma}^- $, $\bm{A}_n^T\jump{\bm{\sigma}} = -\bm{A}_n^T\bm{\sigma}^-$, and $\jump{\bm{v}} = -\bm{v}^-$, yielding
\begin{align*}
\sum_{f\in \Gamma_{\rm abc}} &\int_f \bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v} + \frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma} =\\
& - \sum_{f\in \Gamma_{\rm abc}} \int_f \frac{\tau_v}{2} \bm{A}_n^T\bm{A}_n \LRb{\bm{v}^-}^2 + \frac{\tau_\sigma}{2} \bm{A}_n\bm{A}_n^T \LRb{\bm{\sigma}^-}^2.  
\end{align*}
Combining these face contributions yields
\begin{align}
&\sum_{D^k\in \Oh} \frac{1}{2}\pd{}{t}\LRp{\LRp{\rho \bm{v},\bm{v}}_{\Lk} + \LRp{\bm{C}^{-1} \bm{\sigma},\bm{\sigma}}_{\Lk}} \leq - \sum_{f \in\Gamma_h \setminus \partial \Omega} \int_{f} \frac{\tau_v}{2}\LRb{\bm{A}_n\jump{\bm{v}}}^2 + \frac{\tau_\sigma}{2}\LRb{\bm{A}_n^T\jump{\bm{\sigma}}}^2\nonumber\\
& - \sum_{f\in \Gamma_v} \int_f \tau_v  \LRb{\bm{A}_n\bm{v}^-}^2  - \sum_{f\in \Gamma_\sigma} \int_f \tau_\sigma  \LRb{\bm{A}_n^T\bm{\sigma}^-}^2- \sum_{f\in \Gamma_{\rm abc}} \int_f \frac{\tau_v}{2} \LRb{\bm{A}_n\bm{v}^-}^2 + \frac{\tau_\sigma}{2} \LRb{\bm{A}_n^T\bm{\sigma}^-}^2 \leq 0.  
\label{eq:stability}
\end{align}

Since $\rho$ and $\bm{C}^{-1}$ are positive definite, the left hand side of (\ref{eq:stability}) is an $L^2$-equivalent norm on $(\bm{u},\bm{\sigma})$ 
\[
C_1 \nor{(\bm{u},\bm{\sigma})}_{\L}^2 \leq \sum_{D^k\in \Oh}\frac{1}{2}\pd{}{t}&\LRp{\LRp{\rho \bm{v},\bm{v}}_{\Lk} + \LRp{\bm{C}^{-1} \bm{\sigma},\bm{\sigma}}_{\Lk}} \leq C_2 \nor{(\bm{u},\bm{\sigma})}_{\L}^2
\]
for $\nor{(\bm{u},\bm{\sigma})}_{\L}^2 = \nor{\bm{u}}^2_{\L} + \nor{\bm{\sigma}}^2_{\L}$.  
% $\nor{(\bm{u},\bm{\sigma})}_{\L}$ is 
%\[
%\nor{(\bm{u},\bm{\sigma})}_{\L}^2 =\nor{(\bm{u},\bm{\sigma})}_{\L}^2 = \nor{\bm{u}}^2_{\L} + \nor{\bm{\sigma}}^2_{\L}.
%\]
This implies that the DG formulation is energy stable and that the magnitude of the DG solution $(\bm{u},\bm{\sigma})$ is non-increasing in time.  This also shows that dissipation present for positive penalization constants $\tau_v,\tau_{\sigma}$ acts on non-conforming components with non-zero jumps $\bm{A}_n^T\jump{\bm{v}}$ and $\bm{A}_n\jump{\bm{\sigma}}$.  In fact, it was shown in \cite{chan2016short} that, in the limit as $\tau_v,\tau_\sigma \rightarrow\infty$, the eigenspaces of DG discretizations split into a conforming part consisting of $\bm{u},\bm{\sigma}$ which satisfy
\[
\bm{A}_n\jump{\bm{u}} = 0, \qquad \bm{A}_n^T\jump{\bm{\sigma}} = 0
\]
and an non-conforming part (defined through the $L^2$ orthogonal complement) corresponding to eigenvalues contain real parts which approach $-\infty$.  For the linear elastic wave equations, these conditions are equivalent to the requirement of $C^0$ continuity for $\bm{u}$ and normal continuity of the stress tensor $\jump{\bm{S}}\bm{n} = 0$.
%\begin{align*}
%\jump{\bm{S} \bm{n}} = 0.
%\jump{\bm{\sigma}_{xx} \bm{n}_x + \bm{\sigma}_{xy} \bm{n}_y + \bm{\sigma}_{xz} \bm{n}_z} &= 0\\
%\jump{\bm{\sigma}_{xy} \bm{n}_x + \bm{\sigma}_{yy} \bm{n}_y + \bm{\sigma}_{yz} \bm{n}_z} &= 0\\
%\jump{\bm{\sigma}_{xz} \bm{n}_x + \bm{\sigma}_{yz} \bm{n}_y + \bm{\sigma}_{zz} \bm{n}_z} &= 0.
%\end{align*}

While this DG formulation is energy stable, 
%Symmetric hyperbolic system
%\[
%\bm{A}_0\pd{\bm{v}}{t} = \sum_{i=1}^d \bm{A}_i\pd{\bm{v}}{\bm{x}_i}.
%\]
%The matrix $\bm{A}_0$ is given as
%\[
%\bm{A}_0 = \LRp{
%\begin{array}{cc}
%\rho\bm{I} &\\
%& \bm{C}^{-1}
%\end{array}
%}.
%\]

%The matrices $\bm{A}_1,\bm{A}_2$ are defined as
%\[
%\bm{A}_1 = \LRp{\begin{array}{cc}
%& \bm{A}_x^T\\
%\bm{A}_x
%\end{array}}, \qquad
%\bm{A}_2 = \LRp{\begin{array}{cc}
%& \bm{A}_y^T\\
%\bm{A}_y
%\end{array}},
%\qquad
%\bm{A}_x = \left(\begin{array}{cc}
%1 & 0\\
%0 & 0\\
%0 & 1
%\end{array}\right), \qquad 
%\bm{A}_y = \left(\begin{array}{cc}
%0 & 0\\
%0 & 1\\
%1 & 0
%\end{array}\right)
%\]
%This implies that the RHS terms are 
%\[
%\bm{A}_1 \pd{\bm{v}}{x} + \bm{A}_2 \pd{\bm{v}}{y}  = \LRp{\begin{array}{c}
%\Div \bm{\sigma}_x \\
%\Div \bm{\sigma}_y \\
%\pd{u_1}{x} \\
%\pd{u_2}{y} \\
%\pd{u_2}{x} + \pd{u_1}{y}
%\end{array}}
%\]
%
%The normal matrix $\bm{A}_n$ is
%\[
%\bm{A}_x {n}_x + \bm{A}_y{n}_y = 
%\LRp{\begin{array}{cc}
%& \bm{A}_n^T\\
%\bm{A}_n
%\end{array}}, \qquad
%\bm{A}_n = \LRp{\begin{array}{cc}
%{n}_x & 0\\
% 0 & {n}_y\\
%{n}_y & {n}_x
%\end{array}}
%\]
%We use penalty fluxes defined by 
%\[
%%\LRp{\bm{A}_n\bm{v}}^* = \bm{A}_0^{-1}\bm{A}_n\avg{\bm{v}} - \tau \bm{A}_0^{-1}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}
%\LRp{\bm{A}_n\bm{v}}^* =\bm{A}_n\avg{\bm{v}} - \bm{\tau} \bm{A}_n^T\bm{A}_n\jump{\bm{v}}.
%\]
%The central flux terms for elasticity are then given as
%\[
%\bm{A}_n\avg{\bm{v}} = \LRp{\begin{array}{c}
%\avg{\bm{\sigma}_x} \cdot \bm{n}\\
%\avg{\bm{\sigma}_y} \cdot \bm{n}\\
%\avg{u_1}{n}_x\\
%\avg{u_2}{n}_y\\
%\avg{u_2}{n}_x + \avg{u_1}{n}_y
%\end{array}}
%\]
%while the penalty flux term is
%\[
%\bm{A}_n^T\bm{A}_n\jump{\bm{v}} = 
%\LRp{\begin{array}{ccccc}
%1 & n_xn_y & & &\\
%n_xn_y & 1 & & &\\
% & & n_x^2 & 0 & n_xn_y \\
%  & & 0 & n_y^2 & n_xn_y \\
%   & & n_xn_y & n_xn_y & 1 
%\end{array}}
%\LRp{\begin{array}{c}
%\jump{u_1}\\
%\jump{u_2}\\
%\jump{\sigma_{xx}}\\
%\jump{\sigma_{yy}}\\
%\jump{\sigma_{xy}}
%\end{array}}
%=
%\LRp{\begin{array}{c}
%\jump{u_1} + \jump{u_2}n_xn_y\\
%n_xn_y\jump{u_1} + \jump{u_2}\\
%\LRp{\jump{\bm{\sigma}_x}\cdot\bm{n}}n_x\\
%\LRp{\jump{\bm{\sigma}_y}\cdot\bm{n}}n_y\\
%\jump{\sigma_{xy}} + n_x n_y \LRp{\jump{\sigma_{xx}} + \jump{\sigma_{yy}}}
%\end{array}}.
%\]
%We note that $\bm{A}_n^T\bm{A}_n$ has eigenvalues $1 \pm n_x n_y$ and thus has norm less than $3/2$.  
%We choose the penalty parameter $\bm{\tau}$ as a diagonal matrix such that $\bm{A}_0^{-1}\bm{\tau}\bm{A}_n^T\bm{A}_n = O(1)$.  In this case, $\tau_v = O\LRp{\frac{1}{\rho}}$, while $\tau_{\sigma} = O\LRp{\frac{1}{2\mu + \lambda}}$.

%\subsection{3D formulation}
%\[
%\bm{A}_0\pd{\bm{v}}{t} = \bm{A}_1\pd{\bm{v}}{x} + \bm{A}_2\pd{\bm{v}}{x} + \bm{A}_3\pd{\bm{v}}{x}
%\]
%where $\bm{v} = (u_1,u_2,u_3,{\sigma}_{xx},{\sigma}_{yy},{\sigma}_{zz},\sigma_{yz},\sigma_{xz},\sigma_{xy})$.  We will also denote
%\[
%\bm{\sigma} = \LRp{\bm{\sigma}_x \quad \bm{\sigma}_y} \quad \bm{\sigma}_z
%= 
%\LRp{\begin{array}{ccc}
%\sigma_{xx} & \sigma_{xy} & \sigma_{xz}\\
%\sigma_{xy} & \sigma_{yy} & \sigma_{yz}\\
%\sigma_{xz} & \sigma_{yz} & \sigma_{zz}
%\end{array}}.
%\]
%
%In general anisotropic media, $\bm{C}$ is symmetric and positive-definite.  For three-dimensional isotropic media, $\bm{C}$ is given as
%\[
%\bm{C}= \LRp{
%\begin{array}{cccccc}
%2\mu+\lambda  &  \lambda  &     \lambda  & & & &\\
%\lambda &   2\mu+\lambda & \lambda   & & & &\\
%\lambda & \lambda &   2\mu+\lambda &    & & &\\
% &   &    & \frac{\mu}{2} & &\\
%&  &   &    & \frac{\mu}{2} &\\
%& &  &   &    & \frac{\mu}{2} \\
%\end{array}}, 
%\]
%where $\lambda,\mu$ are Lame parameters.  
%
%The matrices $\bm{A}_1,\bm{A}_2,\bm{A}_3$ are defined as
%\[
%\bm{A}_1 = \LRp{\begin{array}{cc}
%& \bm{A}_x^T\\
%\bm{A}_x
%\end{array}}, \qquad
%\bm{A}_2 = \LRp{\begin{array}{cc}
%& \bm{A}_y^T\\
%\bm{A}_y
%\end{array}},
%\bm{A}_3 = \LRp{\begin{array}{cc}
%& \bm{A}_z^T\\
%\bm{A}_z
%\end{array}},
%\]
%with $\bm{A}_x,\bm{A}_y,\bm{A}_z$
%\[
%\bm{A}_x = \left(\begin{array}{ccc}
%1 & 0 & 0\\
%0 & 0& 0\\
%0 & 0& 0\\
%0 & 0& 0\\
%0 & 0 & 1\\
%0 & 1 & 0
%\end{array}\right), \qquad 
%\bm{A}_y = \left(\begin{array}{ccc}
%0 & 0 & 0\\
%0 & 1 & 0\\
%0 & 0 & 0\\
%0 & 0 & 1\\
%0 & 0 & 0\\
%1 & 0 & 0
%\end{array}\right), \qquad
%\bm{A}_z = \left(\begin{array}{ccc}
%0 & 0 & 0\\
%0 & 0 & 0\\
%0 & 0 & 1\\
%0 & 1 & 0\\
%1 & 0 & 0\\
%0 & 0 & 0
%\end{array}\right)
%\]
%This implies that the RHS terms are 
%\[
%\bm{A}_1 \pd{\bm{v}}{x} + \bm{A}_2 \pd{\bm{v}}{y} +\bm{A}_3 \pd{\bm{v}}{z}  = \LRp{\begin{array}{c}
%\Div \bm{\sigma}_x \\
%\Div \bm{\sigma}_y \\
%\Div \bm{\sigma}_z \\
%\pd{u_1}{x} \\
%\pd{u_2}{y} \\
%\pd{u_3}{z} \\
%\pd{u_3}{y} + \pd{u_2}{z}\\
%\pd{u_3}{x} + \pd{u_1}{z}\\
%\pd{u_2}{x} + \pd{u_1}{y}
%\end{array}}.
%\]
%
%The normal matrix $\bm{A}_n$ is
%\[
%\bm{A}_x {n}_x + \bm{A}_y{n}_y = 
%\LRp{\begin{array}{cc}
%& \bm{A}_n^T\\
%\bm{A}_n
%\end{array}}, \qquad
%\bm{A}_n = \LRp{\begin{array}{ccc}
%{n}_x & 0 & 0\\
% 0 & {n}_y & 0\\
%0&  0 & {n}_z \\
%0 &{n}_z & {n}_y\\
%{n}_z & 0 & {n}_x\\
%{n}_y & {n}_z & 0\\
%\end{array}}
%\]
%We use penalty fluxes defined by 
%\[
%%\LRp{\bm{A}_n\bm{v}}^* = \bm{A}_0^{-1}\bm{A}_n\avg{\bm{v}} - \tau \bm{A}_0^{-1}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}
%\LRp{\bm{A}_n\bm{v}}^* =\bm{A}_n\avg{\bm{v}} - \bm{\tau} \bm{A}_n^T\bm{A}_n\jump{\bm{v}}.
%\]
%The central flux terms for elasticity are then given as
%\[
%\bm{A}_n\avg{\bm{v}} = \LRp{\begin{array}{c}
%\avg{\bm{\sigma}_x} \cdot \bm{n}\\
%\avg{\bm{\sigma}_y} \cdot \bm{n}\\
%\avg{\bm{\sigma}_z} \cdot \bm{n}\\
%\avg{u_1}{n}_x\\
%\avg{u_2}{n}_y\\
%\avg{u_3}{n}_z\\
%\avg{u_3}{n}_y + \avg{u_2}{n}_z\\
%\avg{u_3}{n}_x + \avg{u_1}{n}_z\\
%\avg{u_2}{n}_x + \avg{u_1}{n}_y
%\end{array}}
%\]
%The penalty flux term is evaluated simply by re-applying $\bm{A}_n = \bm{A}_n^T$.  Expanded out, it yields
%\begin{align*}
%\bm{A}_n^T\bm{A}_n\jump{\bm{v}} &= 
%\LRp{\begin{array}{ccc|ccc|ccc}
%1 & n_xn_y & n_xn_z   &&&&&&\\
%n_xn_y & 1 & n_yn_z   &&&&&&\\
%n_xn_z & n_yn_z & 1   &&&&&&\\
%\hline\\
%&&&  n_x^2&&&&n_xn_z&n_xn_y\\
%&&&  &n_y^2&&n_xn_z&&n_xn_y\\
%&&&  &&n_z^2&n_xn_z&n_xn_y&\\
%\hline\\
%&&&  &n_xn_z& n_xn_y& n_y^2+n_z^2&n_xn_y&n_xn_z\\
%&&&  n_xn_z && n_xn_y& n_xn_y& n_x^2+n_z^2&n_yn_z\\
%&&&  n_xn_z &n_xn_y && n_xn_z& n_yn_z& n_x^2+n_y^2\\
%\end{array}}
%\LRp{\begin{array}{c}
%\jump{u_1}\\
%\jump{u_2}\\
%\jump{u_3}\\
%\jump{\sigma_{xx}}\\
%\jump{\sigma_{yy}}\\
%\jump{\sigma_{zz}}\\
%\jump{\sigma_{yz}}\\
%\jump{\sigma_{xz}}\\
%\jump{\sigma_{xy}}
%\end{array}}\\
%&=
%\LRp{\begin{array}{c}
%\jump{u_1} + \jump{u_2}n_xn_y + \jump{u_3}n_xn_z\\
%n_xn_y\jump{u_1} + \jump{u_2} + n_yn_z\jump{u_3}\\
%n_xn_z\jump{u_1} + n_yn_z\jump{u_2} + \jump{u_3}\\
%\LRp{\jump{\bm{\sigma}_x}\cdot\bm{n}}n_x\\
%\LRp{\jump{\bm{\sigma}_y}\cdot\bm{n}}n_y\\
%\LRp{\jump{\bm{\sigma}_z}\cdot\bm{n}}n_z\\
%\jump{\sigma_{yz}}(n_y^2+n_z^2) + \LRp{\jump{\sigma_{xz}}n_y + \jump{\sigma_{xy}}n_z}n_x + \LRp{\jump{\sigma_{yy}}n_y + \jump{\sigma_{zz}}n_y}n_z\\
%\jump{\sigma_{xz}}(n_x^2+n_z^2) + \LRp{\jump{\sigma_{xx}}n_z + \jump{\sigma_{yz}}n_y}n_x + \LRp{\jump{\sigma_{xy}}n_y + \jump{\sigma_{zz}}n_x}n_z\\
%\jump{\sigma_{xy}}(n_x^2+n_y^2) + \LRp{\jump{\sigma_{xx}}n_y + \jump{\sigma_{yy}}n_y}n_x + \LRp{\jump{\sigma_{xz}}n_y + \jump{\sigma_{yz}}n_x}n_z\\
%\end{array}}.
%\end{align*}
%
%We choose the penalty parameter $\bm{\tau}$ as a diagonal matrix such that $\bm{A}_0^{-1}\bm{\tau}\bm{A}_n^T\bm{A}_n = O(1)$.  In this case, $\tau_v = O\LRp{\frac{1}{\rho}}$, while $\tau_{\sigma} = O\LRp{\frac{1}{2\mu + \lambda}}$.

\section{Weight-adjusted approximation of vector-valued inner products}

We also introduce the standard Lebesgue $L^p$ norms over a general domain $\Omega$
\begin{align*}
\nor{u}_{L^p\LRp{\Omega}} &= \LRp{\int_{\Omega} u^p}^{1/p} \qquad 1 \leq p < \infty \\
\nor{u}_{L^{\infty}\LRp{\Omega}} &= \inf\LRc{C \geq 0: \LRb{u\LRp{\bm{x}}} \leq C \quad \forall \bm{x}\in \Omega},
\end{align*}
and the associated $L^p$ spaces
\begin{align*}
L^p\LRp{\Omega} &= \LRc{u: \Omega\rightarrow \mathbb{R}, \quad \nor{u}_{L^p\LRp{\Omega}} < \infty} \qquad 1\leq p < \infty \\
L^{\infty}\LRp{\Omega} &= \LRc{u: \Omega\rightarrow \mathbb{R}, \quad \nor{u}_{L^{\infty}\LRp{\Omega}} < \infty}.
\end{align*}
The $L^p$ Sobolev seminorms and norms of degree $s$ are then defined 
\begin{align*}
\LRb{u}_{W^{s,p}\LRp{\Omega}} &= \LRp{\sum_{\LRb{\alpha}= s} \nor{ D^{\alpha} u}_{L^p\LRp{\Omega}}^p}^{1/p}, \qquad \LRb{u}_{W^{s,\infty}\LRp{\Omega}} = \max_{\LRb{\alpha}= s} \nor{D^{\alpha}u}_{L^{\infty}\LRp{\Omega}}\\
\nor{u}_{W^{s,p}\LRp{\Omega}} &= \LRp{\sum_{\LRb{\alpha}\leq s} \nor{ D^{\alpha} u}_{L^p\LRp{\Omega}}^p}^{1/p}, \qquad \nor{u}_{W^{s,\infty}\LRp{\Omega}} = \max_{\LRb{\alpha}\leq s} \nor{D^{\alpha}u}_{L^{\infty}\LRp{\Omega}}.
\end{align*}
where $\alpha = \LRc{\alpha_1,\alpha_2,\alpha_3}$ is a multi-index such that
\[
D^{\alpha}u = \pd{^{\alpha_1}}{x^{\alpha_1}}\pd{^{\alpha_2}}{y^{\alpha_2}}\pd{^{\alpha_3}}{z^{\alpha_3}} u,
\]

We define $D^{\bm{\alpha}}\bm{v}$  as the component-wise derivative of $\bm{v}$ with respect to a $d$-dimensional multi-index $\bm{\alpha}$
\[
\LRp{D^{\bm{\alpha}}\bm{v}}_i = \frac{\partial^{\alpha_1}}{\partial x^{\alpha_1}}\frac{\partial^{\alpha_2}}{\partial x^{\alpha_2}}\frac{\partial^{\alpha_3}}{\partial x^{\alpha_3}}\bm{v}_i.
\]

We generalize the weight-adjusted scalar inner products introduced in \cite{chan2016weight1,chan2016weight2} to vector-valued inner products with matrix weights.

Vector Sobolev norms for $\bm{v}(\bm{x}) \in \mathbb{R}^m$ can be defined as 
\begin{align*}
\nor{\bm{v}}_{W^{k,p}}^p &= {\sum_{i=1}^m \nor{\bm{v}_i}_{W^{k,2}}^p} = \sum_{\LRb{\alpha} \leq k} \nor{D^{\alpha}\bm{v}}^p_{L^p}, \qquad 1 \leq p < \infty,\\
\nor{\bm{v}}_{W^{k,\infty}} &= \max_i \nor{\bm{v}_i}_{W^{k,\infty}}.
\end{align*}

Let $\bm{W}$ be a symmetric positive-definite matrix such that
\[
0 < w_{\min} \leq  \nor{\bm{W}(\bm{x})}_2 \leq w_{\max} < \infty, \qquad 0 < \tilde{w}_{\min} \leq  \nor{\bm{W}^{-1}(\bm{x})}_2 \leq \tilde{w}_{\max} < \infty, \qquad \forall \bm{x} \in \Omega.
\]
We define a $k$th order Sobolev norm for matrix-valued weight function $\bm{W}(\bm{x})$ in terms of the induced $p$-norm
\begin{align*}
\nor{\bm{W}(\bm{x})}_{k,p}^p &= \sum_{\LRb{\bm{\alpha}} \leq k} \sup_{\bm{x}} \nor{D^{\bm{\alpha}}\bm{W}(\bm{x})}^p_p
\end{align*}

While this norm is not sub-multiplicative, the following bound holds
\begin{align*}
\nor{\bm{W}{\bm{v}}}_{W^{k,p}}^p &= \sum_{\LRb{\alpha}\leq k} \nor{D^{\alpha}\LRp{\bm{W}\bm{v}}}_{L^p}^p 
%\leq C_N \sum_{\LRb{\alpha}\leq k} \nor{\sum_{\LRb{\beta}\leq \alpha} \LRp{D^{\beta}\bm{W}}\LRp{D^{\alpha-\beta}\bm{v}} }_{L^p}^p\\
\leq C_N \int  \sum_{\LRb{\alpha}\leq k} \sum_{\LRb{\beta}\leq \LRb{\alpha}} \nor{\LRp{D^{\beta}\bm{W}}\LRp{D^{\alpha-\beta}\bm{v}} }_{p}^p \\
&\leq C_N\int \LRp{\sum_{\LRb{\alpha}\leq k} \nor{\LRp{D^{\alpha}\bm{W}}}}^p \LRp{\sum_{\LRb{\alpha}\leq k} \nor{D^{\alpha}\bm{v}}_{p} }^p\\
&\leq C_N \nor{\bm{W}}_{k,p}^p \nor{\bm{v}}_{k,p}^p,
\end{align*}
where we have used Leibniz's rule, Cauchy-Schwarz, and the arithmetic-geometric mean inequality.

To show that weight-adjusted DG methods are high order accurate, we make use of a weighted interpolation estimate.  The scalar version of this estimate is derived in \cite{warburton2013low,chan2016weight1}.

\begin{theorem}[Theorem 3.1 in \cite{chan2016weight1}]
Let $D^k$ be a quasi-regular element with representative size $h = {\rm diam}\LRp{D^k}$. For $N \geq 0$, $w\in W^{N+1,\infty}$, and $u\in W^{N+1,2}$, 
\[
\nor{u - \frac{1}{w}\Pi_N(wu)}_{L^2\LRp{D^k}} \leq C_N h^{N+1} \nor{{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{w}}_{L^\infty}\nor{w}_{W^{N+1,\infty}}\nor{u}_{W^{N+1,2}}.
\]
\label{thm:wproj}
\end{theorem}

For $w = 1$, this reduces to the standard interpolation estimate
\[
\nor{u-\Pi_N u}_{L^2\LRp{D^k}} \leq Ch^{N+1}\nor{u}_{W^{N+1,2}}.
\]  
Theorem~\ref{thm:Wproj} extends Theorem~\ref{thm:wproj} to matrix weights by computing weighted interpolation estimates for the quantity $\bm{W}^{-1}\Pi_N\LRp{\bm{W}\bm{v}}$.

\begin{theorem}
Let $D^k$ be a quasi-regular element with representative size $h = {\rm diam}\LRp{D^k}$. For $N \geq 0$, $\bm{W} \in \LRp{W^{N+1,\infty}}^{d\times d}$, and $\bm{v}\in \LRp{W^{N+1,2}}^d$, 
\[
\nor{\bm{v} - \bm{W}^{-1}\Pi_N\LRp{\bm{W}\bm{v}}}_{L^2} \leq Ch^{N+1} \nor{{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk}\tilde{w}_{\max}\nor{\bm{W}}_{N+1} \nor{\bm{v}}_{W^{N+1,2}}
\]
\label{thm:Wproj}
\end{theorem}
\begin{proof}
The proof is the similar to the scalar case.  Using vector-valued versions of Bramble-Hilbert and a scaling argument for quasi-regular elements yields
\begin{align*}
\nor{\bm{v} - \bm{W}^{-1}\Pi_N\LRp{\bm{W}\bm{v}}}_{L^2} &\leq C_1 \nor{\sqrt{J}}_{\LinfDk} \sup_{\bm{x}}\nor{\bm{W}^{-1}}_{2} \nor{\bm{W}\bm{v} - \Pi_N\LRp{\bm{W}\bm{v}}}_{L^2}\\
&\leq C_1 \nor{{\sqrt{J}}}_{\LinfDk} \sup_{\bm{x}}\nor{\bm{W}^{-1}}_{2} \LRb{\bm{W}\bm{v}}_{W^{N+1,2}\LRp{\Dhat}}\\
&\leq C_2 h^{N+1}\nor{{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk} \sup_{\bm{x}}\nor{\bm{W}^{-1}}_{2} \nor{\bm{W}\bm{v}}_{W^{N+1,2}\LRp{D^k}}\\
&\leq C_3 h^{N+1}\nor{{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk} \tilde{w}_{\max} \nor{\bm{W}}_{W^{N+1,2}}\nor{\bm{v}}_{W^{N+1,2}\LRp{D^k}}.
\end{align*}
\end{proof}

\subsection{Properties of weight-adjusted vector-valued inner products}

Let $\Pi_N$ be the $L^2$ projection applied to each component of a vector.  Define the operator $T_{\bm{W}}$
\[
T_{\bm{W}}\bm{v} = \Pi_N\LRp{\bm{W}\bm{v}} 
\]
The inverse operator $T^{-1}_{\bm{W}}$ is defined implicitly via
\[
\LRp{\bm{W}T^{-1}_{\bm{W}}\bm{v},\bm{w}} = \LRp{\bm{v},\bm{w}} .
\]
These definitions generalize $T_{\bm{W}}$ to matrix-valued $\bm{W}$.  Conveniently, all properties of $T_{\bm{W}}$ for scalar $\bm{W}$ carry over from \cite{chan2016weight1}.
\begin{lemma}
Let $\Pi_N$ denote the component-wise $L^2$ projection, and let $\bm{W} \in \LRp{L^{\infty}}^m$.  Then, $T_{\bm{W}}$ satisfies the following properties:
\begin{enumerate}
\vspace{.5em}
\item $T_{\bm{W}}^{-1}T_{\bm{W}} = \Pi_N$ \label{p1}
\vspace{.5em}
\item $\Pi_N T_{\bm{W}}^{-1} = T_{\bm{W}}^{-1}\Pi_N  =  T_{\bm{W}}^{-1}$ \label{p2}
\vspace{.5em}
\item $\nor{T^{-1}_{\bm{W}}}_{L^2\LRp{D^k}} \leq \tilde{w}_{\max}$. \label{p4}
\vspace{.5em}
\item $\LRp{T^{-1}_{\bm{W}} \bm{v},\bm{w}}$ forms an inner product on $\LRp{P^N}^m \times \LRp{P^N}^m$, which is equivalent to the $L^2$ inner product with equivalence constants $C_1 = \tilde{w}_{\min}, C_2 = \tilde{w}_{\max}$. \label{p3}
\vspace{.5em}
\end{enumerate}
\label{lemma:props}
\end{lemma}
\begin{proof}
The proofs of properties~\ref{p1} and \ref{p2} are consequences of the definition of $T_{\bm{W}}$, and are identical to proofs for the scalar case.  Property~\ref{p4} is a straightforward extension from the scalar case.  Let $\bm{v}\in \LRp{P^N}^m$.  Then, 
\[
\LRp{T^{-1}_{\bm{W}} \bm{v},\bm{v}} = \LRp{\bm{W}^{-1}\bm{W}T^{-1}_{\bm{W}} \bm{v},\bm{v}} \leq \sup_{\bm{x}}\nor{\bm{W}^{-1}(\bm{x})}\LRp{\bm{W}T^{-1}_{\bm{W}} \bm{v},\bm{v}} = \tilde{w}_{\max} \nor{\bm{v}}_{L^2}^2.
\]
Property~\ref{p3} then simply requires the lower bound
\[
\LRp{T^{-1}_{\bm{W}} \bm{v},\bm{v}} = \LRp{\bm{W}^{-1}\bm{W}T^{-1}_{\bm{W}} \bm{v},\bm{v}} \geq \inf_{\bm{x}}\nor{\bm{W}^{-1}(\bm{x})}\nor{\bm{v}}_{L^2}^2 = \tilde{w}_{\min}\nor{\bm{v}}_{L^2}^2 .  
\]
\end{proof}

Using Theorem~\ref{thm:Wproj}, we may also show that the matrix-valued weight-adjusted inner product is also high order accurate for sufficiently regular $\bm{W}$.
\begin{theorem}
Let $D^k$ be a quasi-regular element with representative size $h = {\rm diam}\LRp{D^k}$. For $N > 0$
\[
\nor{ \bm{W}\bm{v} - {T}^{-1}_{\bm{W}^{-1}}\bm{v} }_{L^2\LRp{D^k} } \leq C_{\bm{W}} h^{N+1} \nor{\bm{v}}_{W^{N+1,2}\LRp{D^k}}
\]
with constant $C_{\bm{W}}$ depending on $\bm{W}$ and $N$
\[
%C_{\bm{W}} = C_N \cond{\bm{W}}{2,\infty} \nor{\bm{W}}_{W^{N+1,2}\LRp{D^k}}.  
C_{\bm{W}} = C_N \nor{\sqrt{J}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk} \tilde{w}_{\max}{w}_{\max} \nor{\bm{W}}_{W^{N+1,2}\LRp{D^k}}.  
\]
\end{theorem}
\begin{proof}
The proof is again similar to the scalar case.  The triangle inequality gives
\[
\nor{ \bm{W}\bm{v} - {T}^{-1}_{\bm{W}^{-1}}\bm{v} }_{L^2\LRp{D^k} } \leq \nor{ \bm{W}\bm{v} - \Pi_N(\bm{W}\bm{v})}_{L^2\LRp{D^k}} + \nor{\Pi_N(\bm{W}\bm{v}) - T^{-1}_{\bm{W}^{-1}}\bm{v} }_{L^2\LRp{D^k} }.  
\]
The former term is bounded by interpolation estimates and by arguments used in the proof of Theorem~\ref{thm:Wproj}.  The latter term is bounded 
\begin{align*}
\nor{\Pi_N(\bm{W}\bm{v})- T^{-1}_{\bm{W}^{-1}}\bm{v} }_{L^2\LRp{D^k}} &= \nor{T^{-1}_{\bm{W}^{-1}} T_{\bm{W}^{-1}}\Pi_N(\bm{W}\bm{v})- T^{-1}_{\bm{W}^{-1}}\Pi_N\bm{v} }_{L^2\LRp{D^k} } \\
&= \nor{T^{-1}_{\bm{W}^{-1}} \Pi_N \LRp{T_{\bm{W}^{-1}}\Pi_N(\bm{W}\bm{v})}- T^{-1}_{\bm{W}^{-1}}\Pi_N\bm{v} }_{L^2\LRp{D^k} } \\
&\leq C_N \nor{T^{-1}_{\bm{W}^{-1}}} \nor{ T_{\bm{W}^{-1}}\Pi_N(\bm{W}\bm{v})-\Pi_N\bm{v} }_{L^2\LRp{D^k} }\\
&\leq C_N \nor{T^{-1}_{\bm{W}^{-1}}} \nor{ \Pi_N\LRp{\bm{W}^{-1}\Pi_N(\bm{W}\bm{v})} - \Pi_N\bm{v} }_{L^2\LRp{D^k} }\\
&\leq C_N {w}_{\max} \nor{ \bm{W}^{-1}\Pi_N(\bm{W}\bm{v}) - \bm{v} }_{L^2\LRp{D^k} }.
\end{align*}
where we have used $\nor{T_{\bm{W}^{-1}}^{-1}} \leq {w}_{\max}$ (Lemma~\ref{lemma:props}) and $\nor{\Pi_N}_{L^2\LRp{D^k}} = 1$.  
An application of Theorem~\ref{thm:Wproj} to bound $\nor{ \bm{W}^{-1}\Pi_N(\bm{W}\bm{v}) - \bm{v} }_{L^2\LRp{D^k} }$ completes the proof.
\end{proof}

%We note that the local conservation property of DG is lost when using weight-adjusted inner products.  However,

\section{A weight-adjusted discontinuous Galerkin method for elastic wave propagation}

Recall the energy stable formulation for DG 
\begin{align*}
\LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= \LRp{ \sum_{i=1}^d \bm{A}_i^T\pd{\bm{\sigma}}{\bm{x}_i},\bm{w}}_{\Lk}  + \LRa{\frac{1}{2}\bm{A}_n^T\jump{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{v}},\bm{w}}\\
\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{\sigma}},\bm{q}}.
\end{align*}
The weight-adjusted DG method simply replaces the weighted $L^2$ inner products appearing in the left hand side of this formulation with weight-adjusted approximations
\begin{align*}
\LRp{T^{-1}_{1/\rho} \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= \LRp{ \sum_{i=1}^d \bm{A}_i^T\pd{\bm{\sigma}}{\bm{x}_i},\bm{w}}_{\Lk}  + \LRa{\frac{1}{2}\bm{A}_n^T\jump{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{v}},\bm{w}}\\
\LRp{T^{-1}_{\bm{C}} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{\sigma}},\bm{q}}.
\end{align*}
We note that, since the right hand side of the WADG formulation is identical to the right hand side of the DG formulation (\ref{eq:dgform}), WADG preserves a variant of the energy stability (\ref{eq:stability}).  
\begin{align*}
\frac{1}{2}\pd{}{t}\LRp{\LRp{T^{-1}_{1/\rho} \bm{v},\bm{v}}_{\Omega_h} + \LRp{T_{\bm{C}}^{-1} \bm{\sigma},\bm{\sigma}}_{\Omega_h}} \leq - \sum_{f \in \Gamma_h} \int_{f} \frac{\tau_v}{2}\LRb{\bm{B_n}\jump{\bm{v}}}^2 + \frac{\tau_\sigma}{2}\LRb{\bm{B_n}^T\jump{\bm{\sigma}}}^2.  
\end{align*}

We note that, for isotropic media, another way to address sub-element variations in material parameters is to diagonalize the matrix $\bm{C}$ through a change of variables \cite{etienne2010hp}.  This results in uncoupled weighted mass matrices, which can be treated using weight-adjusted approximations for scalar weights.  However, this approach is more difficult to apply when $\bm{C}$ is anisotropic.

The matrix form of DG 
\begin{align*}
\bm{M}_{\rho}\bm{v}_i &= \bm{A}_{{v}_i} \bm{v}_i, \qquad i = 1,\ldots,d\\
\bm{M}_{\bm{C}}\bm{\sigma} &= \bm{A}_{{\sigma}} \bm{\sigma}.
\end{align*}
where the matrices $\bm{A}_{\bm{v}_i}$ and $\bm{A}_{\sigma}$ are induced by the discretization of the right hand side.

In $d$ dimensions, the \emph{coupled} mass matrix is given as 
\[
\LRp{\bm{M}_{\bm{C}^{-1}}}_{i_1,\ldots,i_{N_d},j_1,\ldots,j_{N_d}} = \int_{D^k} \LRp{\begin{array}{c}\phi_{i_1} \\ \vdots \\ \phi_{i_{N_d}}\end{array}}^T\bm{C}^{-1} \LRp{\begin{array}{c}\phi_{j_1}\\ \vdots \\ \phi_{j_{N_d}}\end{array}}.
\]
The storage and inversion of such matrices is expensive,

Weight-adjusted approximation 
\begin{equation}
\bm{M}_{\bm{C}^{-1}} \approx \LRp{\bm{I} \otimes \bm{M}} \bm{M}_{\bm{C}}^{-1} \LRp{\bm{I} \otimes \bm{M}}.
\label{eq:wadgmc}
\end{equation}

We note that, when $\lambda$ and $\mu$ are piecewise constant, the weight-adjusted inner product coincides with exact inversion of the coupled mass matrix.  In this case, $\bm{M}_{\bm{C}}$ reduces to the Kronecker product of the inverse stiffness tensor and the local mass matrix 
\[
\bm{M}_{\bm{C}^{-1}} = \bm{C}^{-1}\otimes \bm{M},
\]
which can be explicitly inverted as
\[
\bm{M}^{-1}_{\bm{C}^{-1}} = \bm{C}\otimes \bm{M}^{-1}.  
\]
Substituting $\bm{M}_{\bm{C}^{-1}}^{-1} = \bm{C}\otimes \bm{M}^{-1}$ into the weight adjusted approximation (\ref{eq:wadgmc}) recovers this expression.% for $\bm{M}_C^{-1}$.

\subsection{Energy stability on curvilinear meshes}

On affine elements, the skew-symmetric formulation is equivalent to the ``strong'' DG formulation 
This formulation yields an efficient quadrature-free evaluation in terms of derivative and lift matrices, which can be further accelerated by exploiting structure under certain bases \cite{chan2015gpu}.

Need to directly discretize skew-symmetric formulation for energy stability on curvilinear meshes.

We note that this formulation is \emph{a-priori} stable, in the sense that a discrete version of energy stability still holds if the $L^2$ inner products are replaced with discrete inner products computed through inexact quadrature rules.  \note{useful for curvilinear meshes}.

\subsection{Convergence analysis}

%Nodal collocation does not work for arbitrary wavespeeds.  Example: wave equation, $c^2 = 1 + .9\sin(8\pi x)\sin(8\pi y)$.

\section{Numerical experiments}

\subsection{Accuracy of weight-adjusted inner products}

Show matrix WADG approximation of $T^{-1}_{\bm{W}}\bm{u}$ is close to $\bm{W}^{-1}\bm{u}$.

\subsection{Verification of stability}

Eigenvalue plots with $\tau = 0$ and $\tau = O(1)$ for both square and curvilinear grids with random coefficients close to zero and 1.

\subsection{Analytic solutions}

We begin by studying the accuracy and convergence of weight-adjusted DG method for several analytical solutions in linear elasticity.  In all cases, the solution in expressed terms of the displacement vector $\bm{v}(\bm{x},t)$.  Initial conditions for velocity and stress are computed through 
\[
\bm{w}(\bm{x},t) = \pd{\bm{v}}{t}, \qquad \bm{\sigma} = \bm{C}\bm{E} = \bm{C}\frac{1}{2}\LRp{\Grad\bm{v} + \Grad\bm{v}^T}.  
\]
Unless otherwise stated, we report relative $L^2$ errors for all components of the solution 
\[
\frac{\nor{\bm{v}-\bm{v}_h}_{L^2\LRp{\Omega}}}{\nor{\bm{v}}_{L^2\LRp{\Omega}}} = \frac{\LRp{\sum_{i=1}^m \nor{\bm{v}_i-\bm{v}^2_{i,h}}_{L^2\LRp{\Omega}}}^{1/2}}{\LRp{\sum_{i=1}^m\nor{\bm{v}_i}^2_{L^2\LRp{\Omega}}}^{1/2}}.
\]

\subsubsection{Harmonic oscillation of a square}

We first examine convergence  on a unit square domain with  $\lambda = \mu = \rho = 1$.   The displacement is given by
\begin{align*}
u_1(x,y,t) &= \cos(\omega\pi t)\cos(\pi x)\sin(\pi y)\\
u_2(x,y,t) &= -\cos(\omega\pi t)\sin(\pi x)\cos(\pi y),
\end{align*}
where $\omega = \sqrt{2\mu}$.  Figure~\ref{fig:harmonic_osc} shows $L^2$ errors computed at time $T = 1$, using uniform triangular meshes constructed by bisecting a uniform mesh of quadrilaterals along the diagonal.  For $N = 1,\ldots, 5$, optimal $O(h^{N+1})$ rates of convergence are observed.

\begin{figure}
\centering
\begin{tikzpicture}
\begin{loglogaxis}[
	legend cell align=left,
	width=.49\textwidth,
    xlabel={Mesh size $h$},
       ylabel={$L^2$ errors}, 
    xmin=.04, xmax=1,
    ymin=1e-12, ymax=1,
    legend pos=south east,
    xmajorgrids=true,
    ymajorgrids=true,
    grid style=dashed,
]

\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.1732)(0.25,0.0439)(0.125,0.0110)(0.0625,0.0027)};
%\node at (axis cs:.04,.0027) {$N = 1$};

\logLogSlopeTriangleFlip{0.32}{0.125}{0.815}{2}{blue}

\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.0150)(0.25,0.0020)(0.125,2.4911e-04)(0.0625,3.1277e-05)};
%\node at (axis cs:.04,3e-5) {$N = 2$};

\logLogSlopeTriangleFlip{0.32}{0.125}{0.66}{3}{red}

\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
coordinates{(0.5, 0.0013)(0.25,8.3910e-05)(0.125,5.3840e-06)(0.0625,3.2015e-07)};
%\node at (axis cs:.04,3e-7) {$N = 3$};

\logLogSlopeTriangleFlip{0.32}{0.125}{0.505}{4}{black}

\addplot[color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
coordinates{(0.5,1.0687e-04)(0.25,3.3041e-06)(0.125,9.7715e-08)(0.0625, 2.8750e-09)};
%\node at (axis cs:.03,1.5e-13) {$a = 10^{-4}$};
\logLogSlopeTriangleFlip{0.32}{0.125}{0.34}{5}{magenta}

\addplot[color=violet,mark=pentagon*,semithick, mark options={fill=markercolor}]
coordinates{(0.5,5.9754e-06)(0.25,9.3644e-08)(0.125,1.3980e-09)(0.0625,2.0686e-11)};

\logLogSlopeTriangleFlip{0.32}{0.125}{0.17}{6}{violet}

%\node at (axis cs:.03,6.8e-10) {$u$ discontinuous};
\legend{$N=1$,$N=2$,$N=3$,$N=4$,$N=5$}
\end{loglogaxis}
\end{tikzpicture}
\caption{Convergence of $L^2$ errors for harmonic oscillation.}
\label{fig:harmonic_osc}
\end{figure}

\subsubsection{Rayleigh and Lamb waves}

Next, we examine the convergence of WADG for Rayleigh and Lamb waves, both of which test the imposition of traction-free boundary conditions.

Rayleigh waves are elastic surface waves which decay exponentially away from the surface.  These waves are given by the displacement vector
\begin{align*}
\bm{v}\LRp{x,y,t} &= e^{-\omega x \sqrt{1-\xi^2}}
\LRp{
\begin{array}{c}
\cos(\omega(y+c_r t)) \\
\sqrt{1-\xi^2}\sin(\omega(y+c_r t))
\end{array}
}\\
&+ \LRp{\frac{\xi^2}{2}-1}e^{-\omega x\sqrt{1-\frac{\xi^2\mu}{2\mu + \lambda}}}
\LRp{
\begin{array}{c}
\cos(\omega(y + c_r t))\\
{\sin(\omega(y+c_r t))}/{\sqrt{1-\frac{\xi^2\mu}{2\mu + \lambda}}}
\end{array}
},
\end{align*}
where $\omega$ is the wavespeed, $c_r$ is the Rayleigh phase velocity $c_r = \xi \sqrt{\mu}$, and $\xi$ satisfies
\[
\sqrt{1-\xi^2}\sqrt{1 - \frac{\xi^2\mu}{2\mu+\lambda}} - \LRp{\frac{\xi^2}{2}-1}^2 = 0.
\]
In our computations, we use $\rho = \mu = \lambda = 1$, $\xi = 0.949554083888034$, and $\omega = 2\pi$.  We solve on the domain $[0,2]\times[0,1]$ using a sequence of uniform triangular meshes, and enforce traction-free boundary conditions at $x=0$ and exact Dirichlet boundary conditions at $x=2$.  Periodic boundary conditions are applied at $y = 0$ and $y = 1$.

Lamb waves are supported by elastic waveguides with traction-free (free surface) boundary conditions at the top and bottom of the domain.  The displacement of these waves 
\begin{align*}
\bm{v}_1\LRp{x,y,t} &= \LRp{-kB_1 \cos(p y)- qB_2 \cos(q y)} \sin(k x - \omega t)\\
\bm{v}_2\LRp{x,y,t} &= \LRp{-pB_1 \sin(py) + kB_2 \sin(qy)} \cos(kx - \omega t)
\end{align*}
where $k$ is the wavenumber and $\omega$ is the frequency, and the constants $p$ and $q$ are defined as
\[
p^2 = \frac{\omega^2}{2\mu + \lambda} - k^2, \qquad q^2 = \frac{\omega^2}{\mu} - k^2.
\]
The wavenumber $k$ and frequency $\omega$ are related through a dispersion relation.  The ratio of the amplitudes $B_1/ B_2$ can be determined using other parameters, implying that $B_1, B_2$ are unique up to a scaling constant.  In our experiments, we use $\rho = \mu = 1$, $\lambda = 2$, $k = 2\pi$.  For these values, $\omega = 13.137063197233$, $B_1 = 126.1992721468$ and $B_2 = 53.88807700007$.  We solve on the domain $[-1,1]\times[-1/2,1/2]$, with traction-free boundary conditions at $y = \pm 1/2$ and periodic boundary conditions at $x = \pm 1$.

Figure~\ref{fig:rayleigh} shows $L^2$ errors for both waves at time $T = 1$.  Observed convergence rates fall between the optimal rate of $O(h^{N+1})$ and theoretical rate of $O(h^{N+1/2})$ \cite{johnson1986analysis,richter1988optimal}.

\begin{figure}
\centering
\subfloat[Rayleigh wave]{
\begin{tikzpicture}
\begin{loglogaxis}[
    legend cell align=left,
    width=.475\textwidth,
    xlabel={Mesh size $h$},
    ylabel={$L^2$ errors}, 
    xmin=.04, xmax=1,
    ymin=1e-8, ymax=1.5,
    legend pos=south east,
    xmajorgrids=true,
    ymajorgrids=true,
    grid style=dashed,
]

\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.9569)(0.25,0.4509)(0.125,0.1234)(0.0625,0.0292)};%\node at (axis cs:.04,.0027) {$N = 1$};
\logLogSlopeTriangleFlip{0.32}{0.125}{0.83}{2}{blue}

\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.3483)(0.25,0.0521)(0.125,0.0072)(0.0625,9.3104e-04)};
\logLogSlopeTriangleFlip{0.32}{0.125}{0.66}{3}{red}

\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.0759)(0.25,  0.0072)(0.125,5.3457e-04)(0.0625,3.6842e-05)};
\logLogSlopeTriangleFlip{0.32}{0.125}{0.49}{4}{black}

\addplot[color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.0169)(0.25,  7.7478e-04)(0.125,2.6817e-05)(0.0625,8.6998e-07)};
\logLogSlopeTriangleFlip{0.32}{0.125}{0.31}{5}{magenta}

\addplot[color=violet,mark=pentagon*,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.0031)(0.25,6.7730e-05)(0.125,1.235333472480438e-06)(0.0625,2.023726469227763e-08)};

\logLogSlopeTriangleFlip{0.32}{0.125}{0.12}{6}{violet}

%\node at (axis cs:.03,6.8e-10) {$u$ discontinuous};
\legend{$N=1$,$N=2$,$N=3$,$N=4$,$N=5$}
\end{loglogaxis}
\end{tikzpicture}
}
\subfloat[Lamb wave]{
\begin{tikzpicture}
\begin{loglogaxis}[
	legend cell align=left,
	width=.475\textwidth,
    xlabel={Mesh size $h$},
%       ylabel={$L^2$ errors}, 
    xmin=.04, xmax=1,
    ymin=2e-7, ymax=1.5,
    legend pos=south east,
    xmajorgrids=true,
    ymajorgrids=true,
    grid style=dashed,
]

\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.9924)(0.25,0.6613)(0.125,0.2713)(0.0625,0.0737)};%\node at (axis cs:.04,.0027) {$N = 1$};
\logLogSlopeTriangleFlip{0.32}{0.125}{0.85}{2}{blue}

\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.6614)(0.25,0.2338)(0.125,0.0205)(0.0625,0.0021)};
\logLogSlopeTriangleFlip{0.32}{0.125}{0.65}{3}{red}

\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.4237)(0.25, 0.0280)(0.125, 0.0018)(0.0625,1.3754e-04)};
\logLogSlopeTriangleFlip{0.32}{0.125}{0.475}{4}{black}

\addplot[color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
coordinates{(0.5, 0.1700 )(0.25,0.0054 )(0.125,1.9613e-04)(0.0625,6.7755e-06)};
\logLogSlopeTriangleFlip{0.32}{0.125}{0.295}{5}{magenta}

\addplot[color=violet,mark=pentagon*,semithick, mark options={fill=markercolor}]
coordinates{(0.5, 0.0393)(0.25,9.0649e-04)(0.125, 1.7466e-05)(0.0625,3.1730e-07)};
\logLogSlopeTriangleFlip{0.32}{0.125}{0.1075}{6}{violet}

%\node at (axis cs:.03,6.8e-10) {$u$ discontinuous};
\legend{$N=1$,$N=2$,$N=3$,$N=4$,$N=5$}
\end{loglogaxis}
\end{tikzpicture}
\label{fig:rayleigh}
}
\caption{Convergence of $L^2$ errors for Rayleigh and Lamb wave solutions.}
\end{figure}

%We also examine the convergence of WADG for Rayleigh waves for near-incompressible materials with $\mu / \lambda \ll 1$.  We compute $L^2$ errors at final time $T = 1/(10\sqrt{\mu})$.

\subsubsection{Stoneley waves}

A Stoneley wave is supported along the interface between two solids \cite{stoneley1924elastic}.  Like Rayleigh waves, Stoneley waves decay exponentially away from the interface, and test the effectiveness of numerical fluxes across interfaces.  We follow \cite{wilcox2010high,appelo2015energy} and use discontinuous media defined by
\[
(\rho,\lambda,\mu) = \begin{cases}
(10 ,3,3) & y > 0\\
(1,1,1) & y < 0.
\end{cases}.
\]
The displacement vector for a Stoneley wave is then given by
\begin{align*}
\bm{v}_1(x,y,t) &= \begin{cases}
{\rm Re}\LRp{\LRp{ikB_1e^{-kb_{1p}y} + kb_{1s}B_2 e^{-kb_{1s}y}}e^{i(ky-\omega t)}}, & y > 0\\
{\rm Re}\LRp{\LRp{-k b_{1p}B_1e^{-k b_{1p} y} + i kB_2e^{-kb_{1s}y}}e^{i(kx-\omega t)}}, & y < 0
\end{cases}\\
\bm{v}_2(x,y,t) &= \begin{cases}
{\rm Re}\LRp{\LRp{ikB_3 e^{k b_{2p} y} - kb_{2s}B_4 e^{k b_{2s}y }} e^{i(kx-\omega t)}}, & y > 0\\
{\rm Re}\LRp{\LRp{k b_{2p} B_3 e^{k b_{2p}y} + ikB_4 e^{k b_{2s}y}}e^{i(kx-\omega t)}}, & y < 0
\end{cases},
\end{align*}
where $c_{st}$ is the Stoneley wave speed, and 
\[
k = \omega/ c_{st}, \qquad b_{jp} = \sqrt{1 - \frac{c_{st}^2}{(2\mu_j + \lambda_j)/\rho_j}}, \qquad  b_{js} = \sqrt{1 - \frac{c_{st}^2}{(\mu_j)/\rho_j}}, \qquad j = 1,2.
\]
The Stoneley wave speed $c_{st}$ can be determined based material parameters and interface conditions, and the amplitudes $B_1,B_2,B_3,B_4$ are determined from $c_{st}$ up to scaling by a constant.  For the parameters used in this study, we take $c_{st} = 0.546981324213884$, $B_1 = i0.2952173626624, B_2 = -0.6798795208473, B_3 = i0.5220044931212$, and $B_4 = -0.9339639688697$.  We assume $k =1$, which gives $\omega=c_{st}$.

We solve on the domain $[-1,1]\times [-5,5]$, and enforce Dirichlet boundary conditions at all boundaries using the exact solution.  Figure~\ref{fig:stoneley} shows $L^2$ errors for two uniform meshes of triangles constructed by bisecting a quadrilateral mesh of $K_{\rm 1D} \times 5K_{\rm 1D}$ elements. Figure~\ref{subfig:stoneley1} shows errors when $K_{\rm 1D}$ is even and the mesh is fitted to the interface at $y = 0$, while Figure~\ref{subfig:stoneley2} shows errors when $K_{\rm 1D}$ is odd and the interface cuts through element interiors.  When the mesh is fitted to the interface, we observe optimal $O(h^{N+1})$ rates of convergence for $N > 1$.

When the mesh is not fitted to the interface exactly, we compute the application of the weight-adjusted mass matrix inverse using a quadrature rule from Xiao and Gimbutas \cite{xiao2010quadrature} which is exact for degree $2N+1$ polynomials.  Since the values of $\rho,\mu$, and $\lambda$ are positive at all quadrature points, the method is energy stable.  However, since the exact solution is discontinuous, we expect an $O(1)$ error in elements cut by the interface, resulting in $L^2$ errors which converge at rate $O(h^{1/2})$.  We note that, when using piecewise constant approximations of $\mu$ and $\lambda$, we observe the same $O(h^{1/2})$ convergence rate, though errors are roughly twice as large in magnitude.  \note{Discuss need to extend WADG to low regularity weights.}

\begin{figure}
\centering
\subfloat[Fitted interface]{
\begin{tikzpicture}
\begin{loglogaxis}[
    legend cell align=left,
    legend style={legend pos=south east, font=\tiny},
    width=.475\textwidth,
    xlabel={Mesh size $h$},
    ylabel={$L^2$ errors}, 
    xmin=.04, xmax=1,
    ymin=2.5e-10, ymax=.15,
    xmajorgrids=true,
    ymajorgrids=true,
    grid style=dashed,
]

\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.0961)(0.25,0.0321)(0.125,0.0108)(0.0625,0.0034)}; 
\logLogSlopeTriangleFlip{0.32}{0.125}{0.84}{2}{blue}

\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.0088)(0.25,0.0013)(0.125,1.9177e-04)(0.0625, 2.8444e-05)};
\logLogSlopeTriangleFlip{0.32}{0.125}{0.62}{3}{red}

\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
coordinates{(0.5, 8.9050e-04)(0.25,7.5838e-05)(0.125,4.9595e-06)(0.0625,3.4440e-07)};
\logLogSlopeTriangleFlip{0.32}{0.125}{0.41}{4}{black}

\addplot[color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
coordinates{(0.5,4.2348e-05)(0.25, 1.4232e-06)(0.125, 4.8606e-08)(0.0625,1.7358e-09)};
\logLogSlopeTriangleFlip{0.32}{0.125}{0.16}{5}{magenta}

\addplot[color=violet,mark=pentagon*,semithick, mark options={fill=markercolor}]
coordinates{(0.5,2.7994e-06)(0.25,4.6958e-08)(0.125,7.9344e-10)};
\logLogSlopeTriangleFlip{0.5}{0.125}{0.1}{6}{violet}

%\node at (axis cs:.03,6.8e-10) {$u$ discontinuous};
\legend{$N=1$,$N=2$,$N=3$,$N=4$,$N=5$}
\end{loglogaxis}
\end{tikzpicture}
\label{subfig:stoneley1}
}
\subfloat[Non-fitted interface]{
\begin{tikzpicture}
\begin{loglogaxis}[
    legend cell align=left,
    legend style={legend pos=south east, font=\tiny},
    width=.475\textwidth,
    xlabel={Mesh size $h$},
%    ylabel={$L^2$ errors}, 
    xmin=.04, xmax=1,
    ymin=1e-2, ymax=.2,
    xmajorgrids=true,
    ymajorgrids=true,
    grid style=dashed,
]

\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.1507)(0.25,0.0796)(0.125,0.0495)(0.0625,0.0331)}; 
%\logLogSlopeTriangleFlip{0.32}{0.125}{0.84}{1}{blue}

\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.0820)(0.25,0.0594)(0.125,0.0427)(0.0625, 0.0307)};
%\logLogSlopeTriangleFlip{0.32}{0.125}{0.62}{3}{red}

\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
coordinates{(0.5, 0.0640)(0.25,0.0435)(0.125,0.0306)(0.0625,0.0216)};
%\logLogSlopeTriangleFlip{0.32}{0.125}{0.41}{4}{black}

\addplot[color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
coordinates{(0.5, 0.0593)(0.25, 0.0436)(0.125, 0.0318)(0.0625,0.0229)};
\logLogSlopeTriangle{0.32}{0.125}{0.225}{1/2}{black}

%\addplot[dashed,color=blue,mark=*,semithick, mark options={fill=markercolor}]
%coordinates{(0.5,0.1648)(0.25,0.0927)(0.125,0.0607)(0.0625,0.0417)}; 
%
%\addplot[dashed,color=red,mark=square*,semithick, mark options={fill=markercolor}]
%coordinates{(0.5,0.1200)(0.25, 0.0855)(0.125,0.0597)(0.0625,0.0421)};
%
%\addplot[dashed,color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
%coordinates{(0.5, 0.1399)(0.25,0.0942)(0.125,0.0660)(0.0625,0.0463)};
%
%\addplot[dashed,color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
%coordinates{(0.5,0.1410)(0.25, 0.0984)(0.125, 0.0697)(0.0625, 0.0490)};

%\addplot[color=violet,mark=pentagon*,semithick, mark options={fill=markercolor}]
%coordinates{(0.5,2.7994e-06)(0.25,4.6958e-08)(0.125,7.9344e-10)};
%\logLogSlopeTriangleFlip{0.5}{0.125}{0.1}{6}{violet}

%\node at (axis cs:.03,6.8e-10) {$u$ discontinuous};
\legend{$N=1$,$N=2$,$N=3$,$N=4$}
\end{loglogaxis}
\end{tikzpicture}
\label{subfig:stoneley2}
}
\caption{Convergence of WADG for a Stoneley wave using a fitted mesh aligned with the interface and a non-fitted mesh where the interface does not lie exactly on an element boundary. }
\label{fig:stoneley}
\end{figure}

\subsubsection{Harmonic oscillation of an annulus}

This example tests the convergence of WADG for curvilinear mappings.

\subsection{Application examples}

\subsubsection{Stiff inclusion}

Leveque stiff inclusion example.

\subsubsection{Heterogeneous anisotropic material}

From \cite{de2007arbitrary}: discontinuous media with anisotropy on one side, isotropy on the other.

Also do a similar case where random noise is added to both sides and the anisotropy is more strongly discontinuous.

\subsection{Three-dimensional example}

Add GPU computations here.

\section{Conclusions}

Generally applicable to problems with spatially varying matrices $\bm{A}_{0}(\bm{x})$, such as in Maxwell's equations for cloaking problems \cite{li2012time}, as well as in vector-valued $H({\rm div})$ or $H({\rm curl})$ functions under matrix-valued mappings such as Piola transforms.

Extension to hybrid meshes is straightforward in 2D \cite{smith2010elastic}. For 3D, care must be taken with pyramids \cite{chan2015orthogonal,chan2015gpu}.

Formulation allows for the reduction of constant-coefficient RHS costs at high orders using BBDG \cite{chan2015bbdg}.

\section{Acknowledgments}

The author gratefully thanks Thomas Hagstrom, Tim Warburton, Axel Modave, and Ruichao Ye for informative discussions.

\bibliographystyle{unsrt}
\bibliography{dgpenalty}

\end{document}

### 2016-12-02 docs/elasticity_notes/main.tex
adding more elasticity paper updates

@@
 \renewcommand{\L}{L^2\LRp{\Omega}}
-\newcommand{\Lk}{L^2\LRp{D^k}}
-\newcommand{\Ldk}{L^2\LRp{\partial D^k}}
+%\newcommand{\Lk}{\Lk}
+%\newcommand{\Ldk}{L^2\LRp{\partial D^k}}
+\newcommand{\Lk}{D^k}
+\newcommand{\Ldk}{\partial D^k}
 \newcommand{\Dhat}{\widehat{D}}
@@
 
-\newcommand{\note}[1]{{\color{blue}#1}}
+\newcommand{\note}[1]{{\color{blue}{#1}}}
 \newcommand{\remark}[1]{\textbf{\color{red}#1}}
@@
 \author{Jesse Chan}
-\title{Weight-adjusted discontinuous Galerkin methods: matrix-valued weights with application to elastic wave propagation}
+\title{Weight-adjusted discontinuous Galerkin methods: matrix-valued weights and linear elastic wave propagation}
 
@@
 \maketitle
-%\tableofcontents
 
@@
 
+%\tableofcontents
+
 \section{Introduction}
@@
 
-SEM \cite{komatitsch2010high,wilcox2010high}
-
-High order DG \cite{kaser2006arbitrary, dumbser2006arbitrary, de2007arbitrary,ye2016discontinuous}
-
-Interior penalty formulations.
-Analysis: Riviere et al.\ \cite{riviere2003discontinuous,riviere2007discontinuous} and Delcourte et al.\ \cite{delcourte2009high,delcourte2015analysis}, dispersion analysis \cite{de2008interior,antonietti2016high}.
-
-GPU-accelerated DG methods \cite{klockner2009nodal,godel2010gpu,godel2010scalability}.  Also elastic wave propagation and seismic wave applications \cite{modave2015nodal,modave2016gpu}
-
-High order approximations of material data \cite{castro2010seismic,mercerat2015nodal}, but not low storage or not energy stable.
+SEM \cite{komatitsch2010high}
+
+DG-SEM \cite{wilcox2010high}
+
+High order DG \cite{kaser2006arbitrary, dumbser2006arbitrary, de2007arbitrary, ye2016discontinuous}
+
+Interior penalty formulations.  Analysis: Riviere et al.\ \cite{riviere2003discontinuous,riviere2007discontinuous} and Delcourte et al.\ \cite{delcourte2009high,delcourte2015analysis}
+
+GPU-accelerated DG methods \cite{klockner2009nodal,godel2010gpu,godel2010scalability}.  Also elastic wave propagation and seismic wave applications \cite{modave2015nodal, modave2016gpu}
+
+High order approximations of material data \cite{castro2010seismic, mercerat2015nodal}, but not low storage or not energy stable.
 
@@
 
-We begin with the linear elastic wave equation under small deformations in a domain $\Omega$.  These equations can be written a first order velocity-stress  system 
-\begin{align*}
-\rho \pd{\bm{v}}{t} &= \Div{\bm{S}}\\
+We begin with the linear elastic wave equation under small deformations in a domain $\Omega$.
+These equations can be written a first order velocity-stress system for velocity $\bm{v}$ and symmetric stress tensor $\bm{S}$
+\begin{align*}
+\rho \pd{\bm{v}}{t} &= \Div{\bm{S}} + \bm{f}\\
 \pd{\bm{S}}{t} &= \frac{1}{2} \bar{\bm{C}}\LRp{\Grad\bm{v} + \Grad\bm{v}^T},
 \end{align*}
-where $\bm{v}$ is the velocity, $\bm{S}$ is the stress tensor, $\rho$ is density, and $\bar{\bm{C}}$ is the symmetric constitutive stiffness tensor relating stress and strain.
+where $\bm{f}$ is the body force per unit volume, $\rho$ is density, and $\bar{\bm{C}}$ is the symmetric constitutive stiffness tensor relating stress and strain.
 These equations can also be rewritten as a symmetric hyperbolic system of PDEs \cite{hughes1978classical}
 \begin{align}
-\rho \pd{\bm{v}}{t} &= \sum_{i=1}^d \bm{A}_i^T \pd{\bm{\sigma}}{\bm{x}_i}\nonumber \\
+\rho \pd{\bm{v}}{t} &= \sum_{i=1}^d \bm{A}_i^T \pd{\bm{\sigma}}{\bm{x}_i}\nonumber + \bm{f}\\
 \bm{C}^{-1} \pd{\bm{\sigma}}{t} &= \sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i}.
@@
 
-\section{An energy stable discontinuous Galerkin variational formulation}
+\section{An energy stable discontinuous Galerkin formulation for elastic wave propagation}
 
@@
 where $\bm{x}^k = \LRc{x^k,y^k,z^k}$ denote physical coordinates on $D^k$ and $\widehat{\bm{x}} = \LRc{\widehat{x},\widehat{y},\widehat{z}}$ denote coordinates on the reference element.
-We denote the determinant of the Jacobian of $\bm{\Phi}^k$ as $J^k$.
-
-We will approximate components over each element $D^k$ from an approximation space $V_h\LRp{D^k}$, which we defined as
+We denote the determinant of the Jacobian of $\bm{\Phi}^k$ as $J$, and refer to it as the Jacobian for the remainder of this work.
+
+We will approximate components over each element $D^k$ from an approximation space $V_h\LRp{D^k}$, which we define as the composition of the mapping $\bm{\Phi}^k$ and a reference approximation space $V_h\LRp{\widehat{D}}$
 \[
@@
 \]
-where $V_h\LRp{\widehat{D}}$ is an approximation space defined on the reference element.
-For the remainder of this work, we will take $V_h\LRp{\widehat{D}} = P^N\LRp{\widehat{D}}$, where $P^N(\LRp{\widehat{D}})$ is the polynomial space of total degree $N$ on the reference simplex.
+For the remainder of this work, we will take $V_h\LRp{\widehat{D}} = P^N\LRp{\widehat{D}}$, where $P^N\LRp{\widehat{D}}$ is the polynomial space of total degree $N$ on the reference simplex.
 In two dimensions,
@@
 \] 
-while in three dimensions, 
+and in three dimensions, 
 \[
@@
 
-We denote the $L^2$ inner product and norm over $D^k$ by $\LRp{\cdot,\cdot}_{L^2\LRp{D^k}}$, such that
-\[
-\LRp{\bm{f},\bm{g}}_{\Lk} = \int_{D^k} \bm{f}\cdot\bm{g} \diff x = \int_{\widehat{D}} \bm{f}\cdot\bm{g} J^k, \qquad \nor{\bm{f}}_{\Lk}^2 = \LRp{\bm{f},\bm{f}}_{\Lk},
+We denote the $L^2$ inner product and norm over $D^k$ by $\LRp{\cdot,\cdot}_{\Lk}$, such that
+\[
+\LRp{\bm{f},\bm{g}}_{\Lk} = \int_{D^k} \bm{f}\cdot\bm{g} \diff x = \int_{\widehat{D}} \bm{f}\cdot\bm{g} J, \qquad \nor{\bm{f}}_{\Lk}^2 = \LRp{\bm{f},\bm{f}}_{\Lk},
 \]
@@
 \begin{align}
-\sum_{D^k\in \Oh}\LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= \sum_{D^k\in \Oh}\LRp{\LRp{ \sum_{i=1}^d \bm{A}_i^T\pd{\bm{\sigma}}{\bm{x}_i},\bm{w}}_{\Lk}  + \LRa{\frac{1}{2}\bm{A}_n^T\jump{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk}} \nonumber\\
-\sum_{D^k\in \Oh}\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \sum_{D^k\in \Oh}\LRp{\LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk}},
+\sum_{D^k\in \Oh} \LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= 
+\sum_{D^k\in \Oh} \LRp{\LRp{ \sum_{i=1}^d \bm{A}_i^T\pd{\bm{\sigma}}{\bm{x}_i} + \bm{f},\bm{w}}_{\Lk}  + \LRa{\frac{1}{2}\bm{A}_n^T\jump{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk}} \nonumber\\
+\sum_{D^k\in \Oh}\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &=
+\sum_{D^k\in \Oh}\LRp{\LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk}},
 \label{eq:dgform}
 \end{align}
-where $\tau_v,\tau_\sigma \geq 0$ are penalty parameters and $\bm{A}_n$ is the normal matrix defined on a face $f$ as $\bm{A}_n = \sum_{i=1}^d \bm{n}_i \bm{A}_i$.  In two dimensions, $\bm{A}_n$ is 
+for all $\bm{w},\bm{q}\in V_h\LRp{\Omega_h}$.
+Here, $\tau_v,\tau_\sigma \geq 0$ are penalty parameters and $\bm{A}_n$ is the normal matrix defined on a face $f$ as $\bm{A}_n = \sum_{i=1}^d \bm{n}_i \bm{A}_i$.
+In two dimensions, $\bm{A}_n$ is
 \[
@@
 \end{align*}
-The numerical fluxes are then computed using these modified jumps.
-
 For problems which involve the truncation of infinite or large domains, absorbing boundary conditions are required.
@@
 
+In all cases, boundary conditions are imposed by computing numerical fluxes using these modified jumps.  This imposition guarantees energy stability for all three sets of boundary conditions.
+
 %Easiest way to impose BCs: set $U^+$ values, compute central fluxes and then compute penalty fluxes via $\bm{A}_n^T\bm{A}_n \jump{U} = \bm{A}_n \jump{F}$.
@@
 It is straightforward to show that the formulation (\ref{eq:dgform}) is energy stable for zero velocity and traction boundary conditions, as well as non-reflective boundary conditions.
-Integrating by parts the velocity equations yields
-\begin{align}
-\LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= -\LRp{ \sum_{i=1}^d \bm{\sigma},\bm{A}_i \pd{\bm{w}}{\bm{x}_i}}_{\Lk}  + \LRa{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}} \nonumber\\
-\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}} \label{eq:dgform}
-\end{align}
+Integrating by parts the velocity equations gives
+\begin{align*}
+\LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= -\LRp{ \sum_{i=1}^d \bm{\sigma},\bm{A}_i \pd{\bm{w}}{\bm{x}_i}}_{\Lk}  + \LRa{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk} \nonumber\\
+\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk}
+\end{align*}
 Taking $(\bm{w},\bm{q}) = (\bm{v},\bm{\sigma})$ and adding both equations together yields
@@
 We separate surface terms into contributions from interior shared faces and from boundary faces.
-On interior shared faces,
+On an interior shared face, we sum contributions from the two adjacent elements to yield
 \begin{align*}
@@
 &\leq - \sum_{f \in\Gamma_h \setminus \partial \Omega} \int_{f} \frac{\tau_v}{2}\LRb{\bm{A}_n\jump{\bm{v}}}^2 + \frac{\tau_\sigma}{2}\LRb{\bm{A}_n^T\jump{\bm{\sigma}}}^2.  
-%\label{eq:stability}
+%\label{thm:stability}
 \end{align*}
@@
 \end{align*}
-Combining these face contributions yields
+Combining all face contributions together gives the following result:
+\begin{theorem}
+The DG formulation (\ref{eq:dgform}) is energy stable for $\tau_v,\tau_\sigma \geq 0$, in the sense that
 \begin{align}
@@
 & - \sum_{f\in \Gamma_v} \int_f \tau_v  \LRb{\bm{A}_n\bm{v}^-}^2  - \sum_{f\in \Gamma_\sigma} \int_f \tau_\sigma  \LRb{\bm{A}_n^T\bm{\sigma}^-}^2- \sum_{f\in \Gamma_{\rm abc}} \int_f \frac{\tau_v}{2} \LRb{\bm{A}_n\bm{v}^-}^2 + \frac{\tau_\sigma}{2} \LRb{\bm{A}_n^T\bm{\sigma}^-}^2 \leq 0.
-\label{eq:stability}
+\label{thm:stability}
 \end{align}
-
-Since $\rho$ and $\bm{C}^{-1}$ are positive definite, the left hand side of (\ref{eq:stability}) is an $L^2$-equivalent norm on $(\bm{u},\bm{\sigma})$ 
-\[
-C_1 \nor{(\bm{u},\bm{\sigma})}_{\L}^2 \leq \sum_{D^k\in \Oh}\frac{1}{2}\pd{}{t}&\LRp{\LRp{\rho \bm{v},\bm{v}}_{\Lk} + \LRp{\bm{C}^{-1} \bm{\sigma},\bm{\sigma}}_{\Lk}} \leq C_2 \nor{(\bm{u},\bm{\sigma})}_{\L}^2
-\]
-for $\nor{(\bm{u},\bm{\sigma})}_{\L}^2 = \nor{\bm{u}}^2_{\L} + \nor{\bm{\sigma}}^2_{\L}$.  
-% $\nor{(\bm{u},\bm{\sigma})}_{\L}$ is 
-%\[
-%\nor{(\bm{u},\bm{\sigma})}_{\L}^2 =\nor{(\bm{u},\bm{\sigma})}_{\L}^2 = \nor{\bm{u}}^2_{\L} + \nor{\bm{\sigma}}^2_{\L}.
-%\]
-This implies that the DG formulation is energy stable and that the magnitude of the DG solution $(\bm{u},\bm{\sigma})$ is non-increasing in time.
+\end{theorem}
+
+Since $\rho$ and $\bm{C}^{-1}$ are positive definite, the left hand side of (\ref{thm:stability}) is an $L^2$-equivalent norm on $(\bm{u},\bm{\sigma})$, and Theorem~\ref{thm:stability} implies that the magnitude of the DG solution $(\bm{u},\bm{\sigma})$ is non-increasing in time.
 This also shows that dissipation present for positive penalization constants $\tau_v,\tau_{\sigma}$ acts on non-conforming components with non-zero jumps $\bm{A}_n^T\jump{\bm{v}}$ and $\bm{A}_n\jump{\bm{\sigma}}$.
@@
 
-While this DG formulation is energy stable, 
-%Symmetric hyperbolic system
-%\[
-%\bm{A}_0\pd{\bm{v}}{t} = \sum_{i=1}^d \bm{A}_i\pd{\bm{v}}{\bm{x}_i}.
-%\]
-%The matrix $\bm{A}_0$ is given as
-%\[
-%\bm{A}_0 = \LRp{
-%\begin{array}{cc}
-%\rho\bm{I} &\\
-%& \bm{C}^{-1}
-%\end{array}
-%}.
-%\]
-
-%The matrices $\bm{A}_1,\bm{A}_2$ are defined as
-%\[
-%\bm{A}_1 = \LRp{\begin{array}{cc}
-%& \bm{A}_x^T\\
-%\bm{A}_x
-%\end{array}}, \qquad
-%\bm{A}_2 = \LRp{\begin{array}{cc}
-%& \bm{A}_y^T\\
-%\bm{A}_y
-%\end{array}},
-%\qquad
-%\bm{A}_x = \left(\begin{array}{cc}
-%1 & 0\\
-%0 & 0\\
-%0 & 1
-%\end{array}\right), \qquad 
-%\bm{A}_y = \left(\begin{array}{cc}
-%0 & 0\\
-%0 & 1\\
-%1 & 0
-%\end{array}\right)
-%\]
+\subsection{The semi-discrete matrix system for DG}
+
+The solution to (\ref{eq:dgform}) can be approximated by discretizing in space and using an explicit time integrator, which requires only evaluations of local contributions over $D^k$ to the DG formulation
+\begin{align}
+\LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= {\LRp{ \sum_{i=1}^d \bm{A}_i^T\pd{\bm{\sigma}}{\bm{x}_i},\bm{w}}_{\Lk}  + \LRa{\frac{1}{2}\bm{A}_n^T\jump{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk}} \nonumber\\
+\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= {\LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk}}.
+\label{eq:dgloc} 
+\end{align}
+
+Let $\LRc{\phi_i}_{i=1}^{N_p}$ be a basis for $P^N\LRp{\widehat{D}}$.  We define the reference mass matrix $\widehat{\bm{M}}$ and the physical mass matrix $\bm{M}$ for an element $D^k$ as
+\[
+\LRp{\widehat{\bm{M}}}_{ij} = \int_{\widehat{D}}\phi_j\phi_i \diff{\widehat{\bm{x}}}, \qquad \LRp{\bm{M}}_{ij} = \int_{D^k} \phi_j\phi_i \diff{\bm{x}} = \int_{\widehat{D}}\phi_j\phi_i J\diff{ \widehat{\bm{x}}}.
+\]
+For affine mappings, $J$ is constant and $\bm{M} = J\widehat{\bm{M}}$.  We also define weak differentiation matrices $\bm{S}_{k}$ and face mass matrix $\bm{M}_f$ such that
+\[
+\LRp{\bm{S}_{k}}_{ij} =  \int_{D^k} \pd{\phi_j}{\bm{x}_k}\phi_i \diff{\bm{x}}, \qquad \LRp{\bm{M}_{f}}_{ij} =  \int_{f} \phi_j \phi_i \diff{\bm{x}} = \int_{\widehat{f}} \phi_j \phi_i J^f\diff{\widehat{\bm{x}}},
+\]
+where $J^f$ is the Jacobian of the mapping from a reference face $\widehat{f}$ to $f$.
+For affinely mapped simplices, $J^f$ is also constant and $\bm{M}_f = J^f \widehat{\bm{M}}_f$, where the definition of the reference face mass matrix $\widehat{\bm{M}}_f$ is analogous to the definition of the reference mass matrix $\widehat{\bm{M}}$.
+
+Finally, we define weighted mass matrices.
+Let $w(\bm{x})\in \mathbb{R}$ and $\bm{W}(\bm{x})\in \mathbb{R}^{m\times n}$.
+Then, scalar and matrix-weighted mass matrices $\bm{M}_w$ and $\bm{M}_{\bm{W}}$ are defined through
+\[
+\LRp{\bm{M}_w}_{ij} = \int_{D^k} w(\bm{x}) \phi_j(\bm{x})  \phi_i(\bm{x}) \diff x, \qquad \bm{M}_{\bm{W}} = \LRp{\begin{array}{ccc}
+\bm{M}_{\bm{W}_{1,1}}& \ldots & \bm{M}_{\bm{W}_{1,n}}\\
+\vdots & \ddots & \vdots\\
+\bm{M}_{\bm{W}_{m,1}} & \ldots & \bm{M}_{\bm{W}_{m,n}}\\
+\end{array}},
+\]
+where $\bm{M}_{\bm{W}_{i,j}}$ is the scalar weighted mass matrix weighted by the $(i,j)$ entry of $\bm{W}$.
+
+Local contributions to the DG variational form may then be evaluated in a quadrature-free manner using these matrices.
+Let $\bm{\Sigma}_i,\bm{V}_i$ denote vectors containing degrees of freedom for solution components $\bm{\sigma}_i, \bm{v}_i$ such that
+\begin{align*}
+\bm{\sigma}_{i}(\bm{x},t) &= \sum_{j=1}^{N_p} \bm{\Sigma}_i(t) \phi_j(\bm{x}), \qquad 1 \leq i \leq N_d\\ 
+\bm{v}_{i}(\bm{x},t) &= \sum_{j=1}^{N_p} \bm{V}_i(t) \phi_j(\bm{x}), \qquad 1 \leq i \leq d. 
+\end{align*}
+Then, the local DG formulation can be written as a block system of ordinary differential equations (ODEs)
+%\begin{align}
+%\bm{M}_{\rho} \pd{\bm{V}_n}{t} &= \sum_{i=1}^d \sum_{j=1}^{N_d}\LRp{\bm{A}_i}_{mn} \bm{S}_{i} \bm{\Sigma}_m + \sum_{f\in \partial D^k}\bm{M}_f\bm{F}_{v,n} , \qquad n = 1,\ldots, d\nonumber\\
+%\sum_{n=1}^{N_d} \bm{M}_{\bm{C}^{-1}_{mn}} \pd{\bm{\Sigma}_n}{t} &= \sum_{i=1}^d \sum_{n=1}^{d} \LRp{\bm{A}_i}_{mn}\bm{S}_{i}\bm{V}_n  +  \sum_{f\in \partial D^k}\bm{M}_f\bm{F}_{\sigma,m}, \qquad m = 1,\ldots,N_d,
+%\label{eq:dgmat}
+%\end{align}
+%where $\bm{F}_{v,n}$ is a vector representing the $n$th component of the velocity numerical flux, $\bm{F}_{\sigma,m}$ is a vector representing the $m$th component of the stress numerical flux, and $\bm{M}_{\rho},\bm{M}_{\bm{C}^{-1}_{mn}}$ are mass matrices weighted with respect to density and the $(m,n)$ entry of $\bm{C}$, respectively.
+We can rewrite this in
+by concatenating $\bm{\Sigma}_i,\bm{V}_i$ into single vectors $\bm{\Sigma},\bm{V}$ and using the Kronecker product $\otimes$
+\begin{align}
+\bm{M}_{\rho \bm{I}} \pd{\bm{V}}{t} &= \sum_{i=1}^d  \LRp{\bm{A}_i^T\otimes \bm{S}_{i}} \bm{\Sigma} + \sum_{f\in \partial D^k}\LRp{\bm{I}\otimes \bm{M}_f}\bm{F}_{v}\nonumber\\
+\bm{M}_{\bm{C}^{-1}} \pd{\bm{\Sigma}}{t} &= \sum_{i=1}^d \LRp{\bm{A}_i\otimes \bm{S}_{i}}\bm{V}  +  \sum_{f\in \partial D^k}\LRp{\bm{I}\otimes\bm{M}_f}\bm{F}_{\sigma}. 
+\label{eq:dgmat}
+\end{align}
+where $\bm{M}_f\bm{F}_{v}$ and $\bm{M}_f\bm{F}_{\sigma}$ represent contributions to the velocity and stress equations from surface integrals.
+
+In order to apply standard time integration methods, we must invert $\bm{M}_{\rho\bm{I}}$ and $\bm{M}_{\bm{C}^{-1}}$ to isolate $\pd{\bm{v}}{t}$ and $\pd{\bm{\sigma}}{t}$ on the left hand side.
+While the inversion of $\bm{M}_{\rho\bm{I}}$ and $\bm{M}_{\bm{C}^{-1}}$ can be parallelized from element to element, doing so typically requires either the precomputation and storage of large dense matrix inverses or the on-the-fly construction and solution of a large dense matrix at every timestep.
+The former option requires a large amount of storage, while the latter option is computationally expensive and difficult to parallelize.
+This cost can be overcome for $\rho, \bm{C}$ which are constant over an element $D^k$, in which case $\bm{M}_{\rho\bm{I}}$ is block diagonal with identical blocks $\bm{M}_\rho = \rho\bm{M}$, while $\bm{M}_{\bm{C}^{-1}}$ reduces to
+\[
+\bm{M}_{\bm{C}^{-1}} = \LRp{\begin{array}{ccc}
+\bm{C}_{1,1}^{-1}\bm{M}& \ldots & \bm{C}^{-1}_{1,N_d}\bm{M}\\
+\vdots & \ddots & \vdots\\
+\bm{C}_{N_d,1}^{-1}\bm{M} & \ldots & \bm{C}^{-1}_{N_d,N_d}\bm{M}\\
+\end{array}} 
+= \LRp{\bm{C}^{-1}\otimes \bm{M}}.
+\]
+Then, $\bm{M}^{-1}_\rho = \frac{1}{\rho}\bm{M}^{-1} = \frac{1}{J\rho}\widehat{\bm{M}}^{-1}$, and $\bm{M}^{-1}_{\bm{C}^{-1}} = \bm{C}\otimes \bm{M}^{-1} = \bm{C}\otimes \LRp{\frac{1}{J}\widehat{\bm{M}}^{-1}}$, and each matrix inverse can be applied using the inverse of the reference mass matrix $\widehat{\bm{M}}^{-1}$ and the values of $\rho, \bm{C}$, and $J$ on each element.
+Applying this observation to \ref{eq:dgmat} then yields the following system of ODEs
+\begin{align*}
+\pd{\bm{V}}{t} &= \sum_{i=1}^d  \LRp{\frac{1}{\rho}\bm{A}_i^T\otimes \bm{D}_{i}} \bm{\Sigma} + \sum_{f\in \partial D^k}\LRp{\frac{1}{\rho}\bm{I}\otimes \bm{L}_f}\bm{F}_{v}\\
+\pd{\bm{\Sigma}}{t} &= \sum_{i=1}^d \LRp{\bm{C}\bm{A}_i\otimes \bm{D}_{i}}\bm{V}  +  \sum_{f\in \partial D^k}\LRp{\bm{C}\otimes\bm{L}_f}\bm{F}_{\sigma},
+\end{align*}
+where we have introduced the differentiation matrix $\bm{D}_i = \bm{M}^{-1}\bm{S}_i$ and lift matrix $\bm{L}_f = \bm{M}^{-1}\bm{M}_f$.
+For affine elements, both derivative and lift matrices may be applied using only geometric factors and reference derivative and lift matrices.
+
+Unfortunately, if $\rho$ and $\bm{C}$ vary spatially within an element, the above approach can no longer be used to invert $\bm{M}_\rho$ and $\bm{M}_{\bm{C}^{-1}}$ in an efficient and low-storage manner.
+%One can avoid these issues by using mass lumping techniques, where nodal points for Lagrange basis functions are co-located at quadrature points on simplices \cite{chin1999higher, cohen2001higher}; however, these points are only available up to order $N\leq 4$ in three dimensions.
+Additionally, these techniques can require the use of enriched polynomial spaces due to a mismatch between the number of points and dimension of $P^N$.
+Mass-lumping can also be done using interior quadratures with $N_p$ points \cite{williams2014symmetric}, though these are defined only up to $N=6$ in three dimensions, and proving high order accuracy.
+We address these issues by approximating the weighted mass matrix and corresponding weighted $L^2$ inner product with a weight-adjusted approximation which is low storage and simple to invert, energy stable, and provably high order accurate for spatially varying weights $\rho, \bm{C}$ with sufficiently regularity.
+
 %This implies that the RHS terms are 
@@
 %\]
-%
-%The normal matrix $\bm{A}_n$ is
-%\[
-%\bm{A}_x {n}_x + \bm{A}_y{n}_y = 
-%\LRp{\begin{array}{cc}
-%& \bm{A}_n^T\\
-%\bm{A}_n
-%\end{array}}, \qquad
-%\bm{A}_n = \LRp{\begin{array}{cc}
-%{n}_x & 0\\
-% 0 & {n}_y\\
-%{n}_y & {n}_x
-%\end{array}}
-%\]
-%We use penalty fluxes defined by 
-%\[
-%%\LRp{\bm{A}_n\bm{v}}^* = \bm{A}_0^{-1}\bm{A}_n\avg{\bm{v}} - \tau \bm{A}_0^{-1}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}
-%\LRp{\bm{A}_n\bm{v}}^* =\bm{A}_n\avg{\bm{v}} - \bm{\tau} \bm{A}_n^T\bm{A}_n\jump{\bm{v}}.
-%\]
+
 %The central flux terms for elasticity are then given as
@@
 
-%\subsection{3D formulation}
-%\[
-%\bm{A}_0\pd{\bm{v}}{t} = \bm{A}_1\pd{\bm{v}}{x} + \bm{A}_2\pd{\bm{v}}{x} + \bm{A}_3\pd{\bm{v}}{x}
-%\]
-%where $\bm{v} = (u_1,u_2,u_3,{\sigma}_{xx},{\sigma}_{yy},{\sigma}_{zz},\sigma_{yz},\sigma_{xz},\sigma_{xy})$.  We will also denote
-%\[
-%\bm{\sigma} = \LRp{\bm{\sigma}_x \quad \bm{\sigma}_y} \quad \bm{\sigma}_z
-%= 
-%\LRp{\begin{array}{ccc}
-%\sigma_{xx} & \sigma_{xy} & \sigma_{xz}\\
-%\sigma_{xy} & \sigma_{yy} & \sigma_{yz}\\
-%\sigma_{xz} & \sigma_{yz} & \sigma_{zz}
-%\end{array}}.
-%\]
-%
-%In general anisotropic media, $\bm{C}$ is symmetric and positive-definite.  For three-dimensional isotropic media, $\bm{C}$ is given as
-%\[
-%\bm{C}= \LRp{
-%\begin{array}{cccccc}
-%2\mu+\lambda  &  \lambda  &     \lambda  & & & &\\
-%\lambda &   2\mu+\lambda & \lambda   & & & &\\
-%\lambda & \lambda &   2\mu+\lambda &    & & &\\
-% &   &    & \frac{\mu}{2} & &\\
-%&  &   &    & \frac{\mu}{2} &\\
-%& &  &   &    & \frac{\mu}{2} \\
-%\end{array}}, 
-%\]
-%where $\lambda,\mu$ are Lame parameters.  
-%
-%The matrices $\bm{A}_1,\bm{A}_2,\bm{A}_3$ are defined as
-%\[
-%\bm{A}_1 = \LRp{\begin{array}{cc}
-%& \bm{A}_x^T\\
-%\bm{A}_x
-%\end{array}}, \qquad
-%\bm{A}_2 = \LRp{\begin{array}{cc}
-%& \bm{A}_y^T\\
-%\bm{A}_y
-%\end{array}},
-%\bm{A}_3 = \LRp{\begin{array}{cc}
-%& \bm{A}_z^T\\
-%\bm{A}_z
-%\end{array}},
-%\]
-%with $\bm{A}_x,\bm{A}_y,\bm{A}_z$
-%\[
-%\bm{A}_x = \left(\begin{array}{ccc}
-%1 & 0 & 0\\
-%0 & 0& 0\\
-%0 & 0& 0\\
-%0 & 0& 0\\
-%0 & 0 & 1\\
-%0 & 1 & 0
-%\end{array}\right), \qquad 
-%\bm{A}_y = \left(\begin{array}{ccc}
-%0 & 0 & 0\\
-%0 & 1 & 0\\
-%0 & 0 & 0\\
-%0 & 0 & 1\\
-%0 & 0 & 0\\
-%1 & 0 & 0
-%\end{array}\right), \qquad
-%\bm{A}_z = \left(\begin{array}{ccc}
-%0 & 0 & 0\\
-%0 & 0 & 0\\
-%0 & 0 & 1\\
-%0 & 1 & 0\\
-%1 & 0 & 0\\
-%0 & 0 & 0
-%\end{array}\right)
-%\]
 %This implies that the RHS terms are 
@@
 %\end{array}}.
-%\]
-%
-%The normal matrix $\bm{A}_n$ is
-%\[
-%\bm{A}_x {n}_x + \bm{A}_y{n}_y = 
-%\LRp{\begin{array}{cc}
-%& \bm{A}_n^T\\
-%\bm{A}_n
-%\end{array}}, \qquad
-%\bm{A}_n = \LRp{\begin{array}{ccc}
-%{n}_x & 0 & 0\\
-% 0 & {n}_y & 0\\
-%0&  0 & {n}_z \\
-%0 &{n}_z & {n}_y\\
-%{n}_z & 0 & {n}_x\\
-%{n}_y & {n}_z & 0\\
-%\end{array}}
-%\]
-%We use penalty fluxes defined by 
-%\[
-%%\LRp{\bm{A}_n\bm{v}}^* = \bm{A}_0^{-1}\bm{A}_n\avg{\bm{v}} - \tau \bm{A}_0^{-1}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}
-%\LRp{\bm{A}_n\bm{v}}^* =\bm{A}_n\avg{\bm{v}} - \bm{\tau} \bm{A}_n^T\bm{A}_n\jump{\bm{v}}.
 %\]
@@
 
-\section{Weight-adjusted approximation of vector-valued inner products}
-
-We also introduce the standard Lebesgue $L^p$ norms over a general domain $\Omega$
-\begin{align*}
-\nor{u}_{L^p\LRp{\Omega}} &= \LRp{\int_{\Omega} u^p}^{1/p} \qquad 1 \leq p < \infty \\
-\nor{u}_{L^{\infty}\LRp{\Omega}} &= \inf\LRc{C \geq 0: \LRb{u\LRp{\bm{x}}} \leq C \quad \forall \bm{x}\in \Omega},
-\end{align*}
-and the associated $L^p$ spaces
-\begin{align*}
-L^p\LRp{\Omega} &= \LRc{u: \Omega\rightarrow \mathbb{R}, \quad \nor{u}_{L^p\LRp{\Omega}} < \infty} \qquad 1\leq p < \infty \\
-L^{\infty}\LRp{\Omega} &= \LRc{u: \Omega\rightarrow \mathbb{R}, \quad \nor{u}_{L^{\infty}\LRp{\Omega}} < \infty}.
-\end{align*}
-The $L^p$ Sobolev seminorms and norms of degree $s$ are then defined 
+\section{Weight-adjusted inner products for matrix-valued weights}
+
+Weight-adjusted inner products are high order accurate approximations of weighted $L^2$ inner products.
+These can be interpreted as generalizations of mass lumping techniques, reducing to mass lumping when integrals are evaluated with appropriate quadrature rules.
+These weight-adjusted inner products result in weight-adjusted mass matrices, whose inverses approximate the inverse of a weighted $L^2$ mass matrix.
+
+We wish to apply weight-adjusted approximations to avoid the inversion of $\bm{M}_\rho$ and $\bm{M}_{\bm{C}^{-1}}$.
+Approximating the inverse of $\bm{M}_\rho$ can be done using weight-adjusted approximations for scalar weights \cite{chan2016weight1,chan2016weight2}, which we review in Section~\ref{sec:swadg}.
+We then extend scalar weight-adjusted approximations to matrix-valued weights in Section~\ref{sec:mwadg} to approximate the inverse of $\bm{M}_{\bm{C}^{-1}}$.
+
+\subsection{Scalar weight adjusted inner products}
+\label{sec:swadg}
+
+We introduce standard Lebesgue $L^p$ norms and their associated $L^p$ spaces over a general domain $\Omega$
+\begin{align*}
+\nor{u}_{L^p\LRp{\Omega}} = \LRp{\int_{\Omega} u^p}^{1/p}, \qquad  
+L^p\LRp{\Omega} = \LRc{u: \Omega\rightarrow \mathbb{R}, \quad \nor{u}_{L^p\LRp{\Omega}} < \infty} 
+\end{align*}
+for $1 \leq p < \infty$.  For $p = \infty$, these are defined as
+\begin{align*}
+\nor{u}_{L^{\infty}\LRp{\Omega}} = \inf\LRc{C \geq 0: \LRb{u\LRp{\bm{x}}} \leq C \quad \forall \bm{x}\in \Omega}, \qquad L^{\infty}\LRp{\Omega} &= \LRc{u: \Omega\rightarrow \mathbb{R}, \quad \nor{u}_{L^{\infty}\LRp{\Omega}} < \infty}.
+\end{align*}
+These induce $L^p$ Sobolev seminorms and norms of degree $s$ 
 \begin{align*}
@@
 
-We define $D^{\bm{\alpha}}\bm{v}$  as the component-wise derivative of $\bm{v}$ with respect to a $d$-dimensional multi-index $\bm{\alpha}$
-\[
-\LRp{D^{\bm{\alpha}}\bm{v}}_i = \frac{\partial^{\alpha_1}}{\partial x^{\alpha_1}}\frac{\partial^{\alpha_2}}{\partial x^{\alpha_2}}\frac{\partial^{\alpha_3}}{\partial x^{\alpha_3}}\bm{v}_i.
-\]
-
-We generalize the weight-adjusted scalar inner products introduced in \cite{chan2016weight1,chan2016weight2} to vector-valued inner products with matrix weights.
-
-Vector Sobolev norms for $\bm{v}(\bm{x}) \in \mathbb{R}^m$ can be defined as 
+We define two operators $T_w: L^2\LRp{D^k} \rightarrow P^N\LRp{D^k}$ and $T^{-1}_w: L^2\LRp{D^k} \rightarrow P^N\LRp{D^k}$ such that
+\[
+T_wu = \Pi_N(wu), \qquad \LRp{wT^{-1}_wu,v}_{\Lk} = \LRp{u,v}_{\Lk}.  
+\]
+A weighted $L^2$ inner product $\LRp{wu,v}_{\Lk}$ can be approximated by a weight-adjusted inner product
+\[
+\LRp{wu,v}_{\Lk} = \LRp{T_wu,v}_{\Lk}\approx \LRp{T^{-1}_{1/w}u,v}_{\Lk}.  
+\]
+based on the observation that $T^{-1}_{1/w}u \approx uw$.  This approximation is made precise by the following estimates for errors in approximations of the product $uw$ and weighted moments:
+\begin{theorem}[Theorem 5 in \cite{chan2016weight1}]
+Let $D^k$ be quasi-regular with representative size $h = {\rm diam}\LRp{D^k}$.  For $N \geq 0$, $w\in W^{N+1,\infty}\LRp{D^k}$, and $u \in W^{N+1,2}\LRp{D^k}$, 
+\begin{align}
+\nor{{u}{w}-T_{w}u}_{L^2\LRp{D^k}} &\leq C_wh^{N+1} \nor{u}_{W^{N+1,2}\LRp{D^k}},\\
+\nor{{u}{w}-T^{-1}_{1/w}u}_{L^2\LRp{D^k}} &\leq C_wh^{N+1}   \nor{u}_{W^{N+1,2}\LRp{D^k}}.
+\end{align}
+where $C_w$ depends on $w$ as follows:
+\[
+C_w = C\nor{w}_{L^{\infty}\LRp{D^k}}\nor{\frac{1}{w}}_{L^{\infty}\LRp{D^k}} \nor{w}_{W^{N+1,\infty}\LRp{D^k}}.
+\]
+\end{theorem}
+\begin{theorem}[Theorem 6 in \cite{chan2016weight1}]
+Let $u\in W^{N+1,2}\LRp{D^k}$, $w\in W^{N+1,\infty}\LRp{D^k}$, and $v \in P^M\LRp{D^k}$ for $0 \leq M\leq N$; then
+\begin{align*}
+&\LRb{\LRp{{w}u,v}_{L^2\LRp{D^k}} - \waip{u,v}{w}} \\
+&\leq Ch^{2N+2 - M} \nor{{w}}_{L^{\infty}\LRp{D^k}}\nor{\frac{1}{w}}_{L^{\infty}\LRp{D^k}}^2 \nor{w}^2_{W^{N+1,\infty}\LRp{D^k}}  \nor{u}_{W^{N+1,2}\LRp{D^k}} \nor{v}_{L^{\infty}\LRp{D^k}}.
+\end{align*}
+\end{theorem}
+
+\subsection{Extension to matrix weights}
+\label{sec:mwadg}
+
+We now aim to generalize weight-adjusted inner products to the case of matrix-valued weights.
+We first define appropriate generalizations of norms used in Section~\ref{sec:swadg} to vector-valued functions.
+Let $D^{\bm{\alpha}}\bm{v}$ denote component-wise differentiation of $\bm{v}$ with respect to a $d$-dimensional multi-index $\bm{\alpha}$.
+Then, vector $L^p$ Sobolev norms for $\bm{v}(\bm{x}) \in \mathbb{R}^m$ can be defined as
 \begin{align*}
@@
 \end{align*}
-
-Let $\bm{W}$ be a symmetric positive-definite matrix such that
+The corresponding Sobolev spaces $W^{k,p}$ and $W^{k,\infty}$ are defined similar to the scalar case.
+
+Let $\bm{W}(\bm{x})$ be a matrix-valued weight function which is pointwise symmetric positive-definite matrix
 \[
@@
 \]
-We define a $k$th order Sobolev norm for matrix-valued weight function $\bm{W}(\bm{x})$ in terms of the induced $p$-norm
+We define a $k$th order Sobolev norm for $\bm{W}(\bm{x})$ in terms of the induced $p$-norm
 \begin{align*}
@@
 \end{align*}
-
 While this norm is not sub-multiplicative, the following bound holds
@@
 
-To show that weight-adjusted DG methods are high order accurate, we make use of a weighted interpolation estimate.
-The scalar version of this estimate is derived in \cite{warburton2013low,chan2016weight1}.
-
+To show that weight-adjusted DG methods are high order accurate, we make use of a scalar weighted interpolation estimate derived in \cite{warburton2013low,chan2016weight1}.  
 \begin{theorem}[Theorem 3.1 in \cite{chan2016weight1}]
@@
 \[
-\nor{u - \frac{1}{w}\Pi_N(wu)}_{L^2\LRp{D^k}} \leq C_N h^{N+1} \nor{{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{w}}_{L^\infty}\nor{w}_{W^{N+1,\infty}}\nor{u}_{W^{N+1,2}}.
+\nor{u - \frac{1}{w}\Pi_N(wu)}_{\Lk} \leq C_N h^{N+1} \nor{{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{w}}_{L^\infty}\nor{w}_{W^{N+1,\infty}}\nor{u}_{W^{N+1,2}}.
 \]
@@
 
-For $w = 1$, this reduces to the standard interpolation estimate
-\[
-\nor{u-\Pi_N u}_{L^2\LRp{D^k}} \leq Ch^{N+1}\nor{u}_{W^{N+1,2}}.
-\]  
-Theorem~\ref{thm:Wproj} extends Theorem~\ref{thm:wproj} to matrix weights by computing weighted interpolation estimates for the quantity $\bm{W}^{-1}\Pi_N\LRp{\bm{W}\bm{v}}$.
-
+The following theorem extends Theorem~\ref{thm:wproj} to matrix weights by computing weighted interpolation estimates for the quantity $\bm{W}^{-1}\Pi_N\LRp{\bm{W}\bm{v}}$.  
 \begin{theorem}
@@
 
-\subsection{Properties of weight-adjusted vector-valued inner products}
-
-Let $\Pi_N$ be the $L^2$ projection applied to each component of a vector.  Define the operator $T_{\bm{W}}$
-\[
-T_{\bm{W}}\bm{v} = \Pi_N\LRp{\bm{W}\bm{v}} 
+\subsubsection{Properties of weight-adjusted vector-valued inner products}
+\label{sec:mwadgprop}
+
+Let $\Pi_N\bm{u}$ be defined as the $L^2$ projection applied to each component of the vector-valued function $\bm{u}$.  We then define the operator $T_{\bm{W}}$ analogously to the scalar case
+\[
+T_{\bm{W}}\bm{v} = \Pi_N\LRp{\bm{W}\bm{v}}.  
 \]
@@
 \]
-These definitions generalize $T_{\bm{W}}$ to matrix-valued $\bm{W}$.  Conveniently, all properties of $T_{\bm{W}}$ for scalar $\bm{W}$ carry over from \cite{chan2016weight1}.
+This definition is a straightforward generalization of $T^{-1}_w}$ to matrix-valued weights $\bm{W}$.
+Conveniently, all properties of $T_{w},T^{-1}_{w}$ for scalar $w(\bm{x})$ carry over to matrix weights $\bm{W}(\bm{x})$ as well \cite{chan2016weight1}.
 \begin{lemma}
@@
 \vspace{.5em}
-\item $\nor{T^{-1}_{\bm{W}}}_{L^2\LRp{D^k}} \leq \tilde{w}_{\max}$. \label{p4}
+\item $\nor{T^{-1}_{\bm{W}}}_{\Lk} \leq \tilde{w}_{\max}$. \label{p4}
 \vspace{.5em}
@@
 \[
-\nor{ \bm{W}\bm{v} - {T}^{-1}_{\bm{W}^{-1}}\bm{v} }_{L^2\LRp{D^k} } \leq C_{\bm{W}} h^{N+1} \nor{\bm{v}}_{W^{N+1,2}\LRp{D^k}}
+\nor{ \bm{W}\bm{v} - {T}^{-1}_{\bm{W}^{-1}}\bm{v} }_{\Lk } \leq C_{\bm{W}} h^{N+1} \nor{\bm{v}}_{W^{N+1,2}\LRp{D^k}}
 \]
@@
 \[
-\nor{ \bm{W}\bm{v} - {T}^{-1}_{\bm{W}^{-1}}\bm{v} }_{L^2\LRp{D^k} } \leq \nor{ \bm{W}\bm{v} - \Pi_N(\bm{W}\bm{v})}_{L^2\LRp{D^k}} + \nor{\Pi_N(\bm{W}\bm{v}) - T^{-1}_{\bm{W}^{-1}}\bm{v} }_{L^2\LRp{D^k} }.
+\nor{ \bm{W}\bm{v} - {T}^{-1}_{\bm{W}^{-1}}\bm{v} }_{\Lk } \leq \nor{ \bm{W}\bm{v} - \Pi_N(\bm{W}\bm{v})}_{\Lk} + \nor{\Pi_N(\bm{W}\bm{v}) - T^{-1}_{\bm{W}^{-1}}\bm{v} }_{\Lk }.  
 \]
@@
 \begin{align*}
-\nor{\Pi_N(\bm{W}\bm{v})- T^{-1}_{\bm{W}^{-1}}\bm{v} }_{L^2\LRp{D^k}} &= \nor{T^{-1}_{\bm{W}^{-1}} T_{\bm{W}^{-1}}\Pi_N(\bm{W}\bm{v})- T^{-1}_{\bm{W}^{-1}}\Pi_N\bm{v} }_{L^2\LRp{D^k} } \\
-&= \nor{T^{-1}_{\bm{W}^{-1}} \Pi_N \LRp{T_{\bm{W}^{-1}}\Pi_N(\bm{W}\bm{v})}- T^{-1}_{\bm{W}^{-1}}\Pi_N\bm{v} }_{L^2\LRp{D^k} } \\
-&\leq C_N \nor{T^{-1}_{\bm{W}^{-1}}} \nor{ T_{\bm{W}^{-1}}\Pi_N(\bm{W}\bm{v})-\Pi_N\bm{v} }_{L^2\LRp{D^k} }\\
-&\leq C_N \nor{T^{-1}_{\bm{W}^{-1}}} \nor{ \Pi_N\LRp{\bm{W}^{-1}\Pi_N(\bm{W}\bm{v})} - \Pi_N\bm{v} }_{L^2\LRp{D^k} }\\
-&\leq C_N {w}_{\max} \nor{ \bm{W}^{-1}\Pi_N(\bm{W}\bm{v}) - \bm{v} }_{L^2\LRp{D^k} }.
-\end{align*}
-where we have used $\nor{T_{\bm{W}^{-1}}^{-1}} \leq {w}_{\max}$ (Lemma~\ref{lemma:props}) and $\nor{\Pi_N}_{L^2\LRp{D^k}} = 1$.  
-An application of Theorem~\ref{thm:Wproj} to bound $\nor{ \bm{W}^{-1}\Pi_N(\bm{W}\bm{v}) - \bm{v} }_{L^2\LRp{D^k} }$ completes the proof.
+\nor{\Pi_N(\bm{W}\bm{v})- T^{-1}_{\bm{W}^{-1}}\bm{v} }_{\Lk} &= \nor{T^{-1}_{\bm{W}^{-1}} T_{\bm{W}^{-1}}\Pi_N(\bm{W}\bm{v})- T^{-1}_{\bm{W}^{-1}}\Pi_N\bm{v} }_{\Lk } \\
+&= \nor{T^{-1}_{\bm{W}^{-1}} \Pi_N \LRp{T_{\bm{W}^{-1}}\Pi_N(\bm{W}\bm{v})}- T^{-1}_{\bm{W}^{-1}}\Pi_N\bm{v} }_{\Lk } \\
+&\leq C_N \nor{T^{-1}_{\bm{W}^{-1}}} \nor{ T_{\bm{W}^{-1}}\Pi_N(\bm{W}\bm{v})-\Pi_N\bm{v} }_{\Lk }\\
+&\leq C_N \nor{T^{-1}_{\bm{W}^{-1}}} \nor{ \Pi_N\LRp{\bm{W}^{-1}\Pi_N(\bm{W}\bm{v})} - \Pi_N\bm{v} }_{\Lk }\\
+&\leq C_N {w}_{\max} \nor{ \bm{W}^{-1}\Pi_N(\bm{W}\bm{v}) - \bm{v} }_{\Lk }.
+\end{align*}
+where we have used $\nor{T_{\bm{W}^{-1}}^{-1}} \leq {w}_{\max}$ (Lemma~\ref{lemma:props}) and $\nor{\Pi_N}_{\Lk} = 1$.  
+An application of Theorem~\ref{thm:Wproj} to bound $\nor{ \bm{W}^{-1}\Pi_N(\bm{W}\bm{v}) - \bm{v} }_{\Lk }$ completes the proof.
 \end{proof}
 
+\note{This is a good generalization because defining $T^{-1}_w \bm{v} = T^{-1}_{w\bm{I}}\bm{v}$ recovers estimates for scalar weight-adjusted inner products.}
+
 %We note that the local conservation property of DG is lost when using weight-adjusted inner products.  However,
 
-\section{A weight-adjusted discontinuous Galerkin method for elastic wave propagation}
-
-Recall the energy stable formulation for DG 
-\begin{align*}
-\LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= \LRp{ \sum_{i=1}^d \bm{A}_i^T\pd{\bm{\sigma}}{\bm{x}_i},\bm{w}}_{\Lk}  + \LRa{\frac{1}{2}\bm{A}_n^T\jump{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{v}},\bm{w}}\\
-\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{\sigma}},\bm{q}}.
-\end{align*}
-The weight-adjusted DG method simply replaces the weighted $L^2$ inner products appearing in the left hand side of this formulation with weight-adjusted approximations
-\begin{align*}
-\LRp{T^{-1}_{1/\rho} \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= \LRp{ \sum_{i=1}^d \bm{A}_i^T\pd{\bm{\sigma}}{\bm{x}_i},\bm{w}}_{\Lk}  + \LRa{\frac{1}{2}\bm{A}_n^T\jump{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{v}},\bm{w}}\\
-\LRp{T^{-1}_{\bm{C}} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{\sigma}},\bm{q}}.
-\end{align*}
-We note that, since the right hand side of the WADG formulation is identical to the right hand side of the DG formulation (\ref{eq:dgform}), WADG preserves a variant of the energy stability (\ref{eq:stability}).
-\begin{align*}
-\frac{1}{2}\pd{}{t}\LRp{\LRp{T^{-1}_{1/\rho} \bm{v},\bm{v}}_{\Omega_h} + \LRp{T_{\bm{C}}^{-1} \bm{\sigma},\bm{\sigma}}_{\Omega_h}} \leq - \sum_{f \in \Gamma_h} \int_{f} \frac{\tau_v}{2}\LRb{\bm{B_n}\jump{\bm{v}}}^2 + \frac{\tau_\sigma}{2}\LRb{\bm{B_n}^T\jump{\bm{\sigma}}}^2.
-\end{align*}
+\subsubsection{Approximation of weighted mass matrix inverses}
+
+The advantage of using weight-adjusted inner products is that the corresponding weight-adjusted mass matrices are straightforward to invert.
+For scalar weights, the weight-adjusted mass matrix approximates the weighted $L^2$ mass matrix and its inverse by
+\[
+\bm{M}_w \approx \bm{M}\bm{M}^{-1}_{1/w}\bm{M}, \qquad \bm{M}^{-1}_w \approx \bm{M}^{-1}\bm{M}_{1/w}\bm{M}^{-1}.
+\]
+By evaluating $\bm{M}_{1/w}$ in a matrix-free fashion (for example, using a sufficiently accurate quadrature rule), the inverse of the weight-adjusted mass matrix $\bm{M}^{-1}\bm{M}_{1/w}\bm{M}^{-1}$ can be applied in a low storage fashion.
+
+For weight-adjusted inner products with matrix-valued weights, the corresponding weight-adjusted mass matrices approximate weighted $L^2$ mass matrices and inverses in a similar fashion
+\begin{align}
+\bm{M}_{\bm{W}} \approx \LRp{\bm{I} \otimes \bm{M}} \bm{M}^{-1}_{\bm{W}^{-1}} \LRp{\bm{I} \otimes \bm{M}} \nonumber\\
+\bm{M}_{\bm{W}}^{-1} \approx \LRp{\bm{I} \otimes \bm{M}^{-1}} \bm{M}_{\bm{W}^{-1}} \LRp{\bm{I} \otimes \bm{M}^{-1}}.
+\label{eq:wadgmc}
+\end{align}
+
+We note that, when $\bm{W}$ is constant over $D^k$, $\bm{M}_{\bm{W}}$ reduces to the Kronecker product of the inverse stiffness tensor and the local mass matrix 
+\[
+\bm{M}_{\bm{W}} = \bm{W}\otimes \bm{M}, \qquad \bm{M}_{\bm{W}}^{-1} = \bm{W}^{-1}\otimes \bm{M}^{-1}.
+\] 
+In this case, we also have $\bm{M}_{\bm{W}^{-1}} = \bm{W}^{-1}\otimes \bm{M}$.
+Substituting this explicit inverse into the weight adjusted mass matrix inverse $\LRp{\bm{I} \otimes \bm{M}^{-1}} \bm{M}_{\bm{W}^{-1}} \LRp{\bm{I} \otimes \bm{M}^{-1}}$ in (\ref{eq:wadgmc}) recovers the exact inversion of $\bm{M}_{\bm{W}}$.
+
+\section{An energy stable weight-adjusted discontinuous Galerkin formulation for elastic wave propagation}
+\label{sec:wadgelas}
+%Recall the energy stable formulation for DG 
+%\begin{align*}
+%\LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= \LRp{ \sum_{i=1}^d \bm{A}_i^T\pd{\bm{\sigma}}{\bm{x}_i},\bm{w}}_{\Lk}  + \LRa{\frac{1}{2}\bm{A}_n^T\jump{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{v}},\bm{w}}_{\Ldk}\\
+%\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{\sigma}},\bm{q}}_{\Ldk}.
+%\end{align*}
+We construct a weight-adjusted DG method by simply replacing the weighted $L^2$ inner products appearing in the left hand side of the local DG formulation (\ref{eq:dgloc}) with weight-adjusted approximations
+\begin{align*}
+\LRp{T^{-1}_{1/\rho} \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= \LRp{ \sum_{i=1}^d \bm{A}_i^T\pd{\bm{\sigma}}{\bm{x}_i},\bm{w}}_{\Lk}  + \LRa{\frac{1}{2}\bm{A}_n^T\jump{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{v}},\bm{w}}_{\Ldk}\\
+\LRp{T^{-1}_{\bm{C}} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{\sigma}},\bm{q}}_{\Ldk}.
+\end{align*}
+Since the right hand side of the WADG formulation is identical to the right hand side of the DG formulation (\ref{eq:dgform}), WADG preserves a variant of the energy stability in Theorem~\ref{thm:stability}
+\begin{align*}
+&\sum_{D^k\in \Oh} \frac{1}{2}\pd{}{t}\LRp{\LRp{T^{-1}_{1/\rho} \bm{v},\bm{v}}_{\Lk} + \LRp{T^{-1}_{\bm{C}} \bm{\sigma},\bm{\sigma}}_{\Lk}} \leq - \sum_{f \in\Gamma_h \setminus \partial \Omega} \int_{f} \frac{\tau_v}{2}\LRb{\bm{A}_n\jump{\bm{v}}}^2 + \frac{\tau_\sigma}{2}\LRb{\bm{A}_n^T\jump{\bm{\sigma}}}^2\nonumber\\
+& - \sum_{f\in \Gamma_v} \int_f \tau_v  \LRb{\bm{A}_n\bm{v}^-}^2  - \sum_{f\in \Gamma_\sigma} \int_f \tau_\sigma  \LRb{\bm{A}_n^T\bm{\sigma}^-}^2- \sum_{f\in \Gamma_{\rm abc}} \int_f \frac{\tau_v}{2} \LRb{\bm{A}_n\bm{v}^-}^2 + \frac{\tau_\sigma}{2} \LRb{\bm{A}_n^T\bm{\sigma}^-}^2 \leq 0.
+\end{align*}
+
+The substitution of weight-adjusted inner products for weighted $L^2$ inner products replaces the weighted $L^2$ mass matrices in (\ref{eq:dgmat}) by their weight-adjusted approximations.  
+%\begin{align*}
+%\bm{M}_{\rho \bm{I}} \pd{\bm{V}}{t} &= \sum_{i=1}^d  \LRp{\bm{A}_i^T\otimes \bm{S}_{i}} \bm{\Sigma} + \sum_{f\in \partial D^k}\LRp{\bm{I}\otimes \bm{M}_f}\bm{F}_{v}\nonumber\\
+%\bm{M}_{\bm{C}^{-1}} \pd{\bm{\Sigma}}{t} &= \sum_{i=1}^d \LRp{\bm{A}_i\otimes \bm{S}_{i}}\bm{V}  +  \sum_{f\in \partial D^k}\LRp{\bm{I}\otimes\bm{M}_f}\bm{F}_{\sigma}. 
+%\end{align*}
+Inverting these weight-adjusted mass matrices yields the following system of ODEs for $\bm{V},\bm{\Sigma}$
+\begin{align*}
+\pd{\bm{V}}{t} &= \LRp{\bm{I}\otimes\bm{M}^{-1}} \bm{M}_{\LRp{\rho^{-1}\bm{I}}} \LRp{\sum_{i=1}^d  \LRp{\bm{A}_i^T\otimes \bm{D}_{i}} \bm{\Sigma} + \sum_{f\in \partial D^k}\LRp{\bm{I}\otimes \bm{L}_f}\bm{F}_{v}} \nonumber\\
+\pd{\bm{\Sigma}}{t} &= \LRp{\bm{I}\otimes\bm{M}^{-1}} \bm{M}_{\bm{C}}  \LRp{\sum_{i=1}^d \LRp{\bm{A}_i\otimes \bm{D}_{i}}\bm{V}  +  \sum_{f\in \partial D^k}\LRp{\bm{I}\otimes\bm{L}_f}\bm{F}_{\sigma}}.
+\end{align*}
+In practice, the matrices $\LRp{\bm{I}\otimes\bm{M}^{-1}} \bm{M}_{\LRp{\rho^{-1}\bm{I}}}$ and $\LRp{\bm{I}\otimes\bm{M}^{-1}} \bm{M}_{\bm{C}}$ are applied to a vector using quadrature.
+After fusing operations together, this procedure can be boiled down to multiplication by two rectangular matrices.
+
+%The matrix form of DG 
+%\begin{align*}
+%\bm{M}_{\rho \bm{I}} \pd{\bm{V}}{t} &= \sum_{i=1}^d  \LRp{\bm{A}_i^T\otimes \bm{S}_{i}} \bm{\Sigma} + \sum_{f\in \partial D^k}\LRp{\bm{I}\otimes \bm{M}_f}\bm{F}_{v}\nonumber\\
+%\bm{M}_{\bm{C}^{-1}} \pd{\bm{\Sigma}}{t} &= \sum_{i=1}^d \LRp{\bm{A}_i\otimes \bm{S}_{i}}\bm{V}  +  \sum_{f\in \partial D^k}\LRp{\bm{I}\otimes\bm{M}_f}\bm{F}_{\sigma}. 
+%\end{align*}
+%where the matrices $\bm{A}_{\bm{v}_i}$ and $\bm{A}_{\sigma}$ are induced by the discretization of the right hand side.
 
 We note that, for isotropic media, another way to address sub-element variations in material parameters is to diagonalize the matrix $\bm{C}$ through a change of variables \cite{etienne2010hp}.
-This results in uncoupled weighted mass matrices, which can be treated using weight-adjusted approximations for scalar weights.
+This results in a local system of ODEs with only scalar weighted mass matrices \cite{mercerat2015nodal}, which can be treated using scalar weight-adjusted approximations.
 However, this approach is more difficult to apply when $\bm{C}$ is anisotropic.
 
-The matrix form of DG 
-\begin{align*}
-\bm{M}_{\rho}\bm{v}_i &= \bm{A}_{{v}_i} \bm{v}_i, \qquad i = 1,\ldots,d\\
-\bm{M}_{\bm{C}}\bm{\sigma} &= \bm{A}_{{\sigma}} \bm{\sigma}.
-\end{align*}
-where the matrices $\bm{A}_{\bm{v}_i}$ and $\bm{A}_{\sigma}$ are induced by the discretization of the right hand side.
-
-In $d$ dimensions, the \emph{coupled} mass matrix is given as 
-\[
-\LRp{\bm{M}_{\bm{C}^{-1}}}_{i_1,\ldots,i_{N_d},j_1,\ldots,j_{N_d}} = \int_{D^k} \LRp{\begin{array}{c}\phi_{i_1} \\ \vdots \\ \phi_{i_{N_d}}\end{array}}^T\bm{C}^{-1} \LRp{\begin{array}{c}\phi_{j_1}\\ \vdots \\ \phi_{j_{N_d}}\end{array}}.
-\]
-The storage and inversion of such matrices is expensive,
-
-Weight-adjusted approximation 
-\begin{equation}
-\bm{M}_{\bm{C}^{-1}} \approx \LRp{\bm{I} \otimes \bm{M}} \bm{M}_{\bm{C}}^{-1} \LRp{\bm{I} \otimes \bm{M}}.
-\label{eq:wadgmc}
-\end{equation}
-
-We note that, when $\lambda$ and $\mu$ are piecewise constant, the weight-adjusted inner product coincides with exact inversion of the coupled mass matrix.
-In this case, $\bm{M}_{\bm{C}}$ reduces to the Kronecker product of the inverse stiffness tensor and the local mass matrix
-\[
-\bm{M}_{\bm{C}^{-1}} = \bm{C}^{-1}\otimes \bm{M},
-\]
-which can be explicitly inverted as
-\[
-\bm{M}^{-1}_{\bm{C}^{-1}} = \bm{C}\otimes \bm{M}^{-1}.  
-\]
-Substituting $\bm{M}_{\bm{C}^{-1}}^{-1} = \bm{C}\otimes \bm{M}^{-1}$ into the weight adjusted approximation (\ref{eq:wadgmc}) recovers this expression.% for $\bm{M}_C^{-1}$.
-
-\subsection{Energy stability on curvilinear meshes}
-
-On affine elements, the skew-symmetric formulation is equivalent to the ``strong'' DG formulation 
-This formulation yields an efficient quadrature-free evaluation in terms of derivative and lift matrices, which can be further accelerated by exploiting structure under certain bases \cite{chan2015gpu}.
-
-Need to directly discretize skew-symmetric formulation for energy stability on curvilinear meshes.
-
-We note that this formulation is \emph{a-priori} stable, in the sense that a discrete version of energy stability still holds if the $L^2$ inner products are replaced with discrete inner products computed through inexact quadrature rules.
-\note{useful for curvilinear meshes}.
+One drawback of the analysis presented in this work is that accuracy is not guaranteed in the isotropic incompressible limit $\mu / \lambda \rightarrow 0$.
+In this case, the stiffness matrix $\bm{C}$ becomes singular, and the constant $\tilde{w}_{\max}$ in the upper bound on $\nor{\bm{C}}$ blows up.
+However, numerical experiments in Section~\ref{sec:incomp} suggest that the accuracy of WADG  for linear elastic wave propagation does not degrade significantly in the incompressible limit.
+
+\subsection{An energy stable formulation for curvilinear meshes}
+
+We have shown the energy stability of the DG formulation (\ref{eq:dgform}) on meshes of affine elements by assuming exact integration of all terms.
+However, energy stability can still be guaranteed if integrals are evaluated inexactly using quadrature.
+Instead of discretizing the ``strong'' DG formulation (\ref{eq:dgform}), we discretize the ``skew-symmetric'' formulation, where the right hand side of the velocity equations is integrated by parts,\footnote{The choice of which equation to integrate by parts is arbitrary, since integrating the stress equations by parts also results in an energy stable formulation.
+} resulting in the formulation
+\begin{align*}
+\int_{\Dhat} T^{-1}_{(\rho J)^{-1}} \pd{\bm{v}}{t}\bm{w} &= -\int_{\Dhat}\LRp{\sum_{i=1}^d \bm{\sigma},\bm{A}_i \pd{\bm{w}}{\bm{x}_i}}J  + \sum_{f \in \partial D^k} \int_{\widehat{f}}\LRp{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}}\bm{w}J^f \nonumber\\
+\int_{\Dhat}  T^{-1}_{\LRp{J^{-1}\bm{C}}} \pd{\bm{\sigma}}{t}\bm{q} &= \int_{\Dhat} \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}J + \sum_{f\in \partial D^k}\int_{\widehat{f}}\LRp{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}}\bm{q}J^f
+%\LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= -\LRp{ \sum_{i=1}^d \bm{\sigma},\bm{A}_i \pd{\bm{w}}{\bm{x}_i}}_{\Lk}  + \LRa{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk} \nonumber\\
+%\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk}.
+\end{align*}
+where we have incorporated spatial variations of $J$ into the definition of the weights on the left hand side.
+
+Since the proof of stability in Theorem~\ref{thm:stability} requires only algebraic manipulations of this formulation, the same energy stability result holds for this formulation.
+More importantly, a discrete version of energy stability is still guaranteed in the presence of inexact quadrature, where integrals in Theorem~\ref{thm:stability} are replaced with quadrature approximations.
+This is especially important for discretizations on curvilinear meshes, where the exact integration of spatially varying geometric factors and Jacobians can be either inhibitively expensive for high order curvilinear mappings or impossible for rational mappings \cite{engvall2016isogeometric, michoski2015foundations}.
+
+We note that, to ensure energy stability on curved and non-affine elements, the ``skew-symmetric'' formulation must be evaluated explicitly using quadrature, which is typically more expensive than quadrature-free evaluations used to evaluate the ``strong'' DG formulation (\ref{eq:dgform}).
+These costs can be slightly reduced for curvilinear meshes by evaluating the DG formulation using the ``skew-symmetric'' formulation on curvilinear elements and the more efficient ``strong'' DG formulation on affine elements \cite{chan2016weight2}.
 
@@
 
-\subsection{Accuracy of weight-adjusted inner products}
-
-Show matrix WADG approximation of $T^{-1}_{\bm{W}}\bm{u}$ is close to $\bm{W}^{-1}\bm{u}$.
+The following sections present several numerical experiments validating the stability and accuracy of the proposed method in two and three dimensions.
 
@@
 
-Eigenvalue plots with $\tau = 0$ and $\tau = O(1)$ for both square and curvilinear grids with random coefficients close to zero and 1.
+blah
+
+\note{Eigenvalue plots with $\tau = 0$ and $\tau = O(1)$ for both square and curvilinear grids with random coefficients close to zero and 1.  Change in solution norm as time goes on.  }
 
@@
 
-We begin by studying the accuracy and convergence of weight-adjusted DG method for several analytical solutions in linear elasticity.
-In all cases, the solution in expressed terms of the displacement vector $\bm{v}(\bm{x},t)$.
+Next, we study the accuracy and convergence of weight-adjusted DG method for several analytical solutions in linear elasticity.
+In all cases, the solution in expressed terms of the displacement vector $\bm{u}(\bm{x},t)$.
 Initial conditions for velocity and stress are computed through
 \[
-\bm{w}(\bm{x},t) = \pd{\bm{v}}{t}, \qquad \bm{\sigma} = \bm{C}\bm{E} = \bm{C}\frac{1}{2}\LRp{\Grad\bm{v} + \Grad\bm{v}^T}.  
-\]
-Unless otherwise stated, we report relative $L^2$ errors for all components of the solution 
-\[
-\frac{\nor{\bm{v}-\bm{v}_h}_{L^2\LRp{\Omega}}}{\nor{\bm{v}}_{L^2\LRp{\Omega}}} = \frac{\LRp{\sum_{i=1}^m \nor{\bm{v}_i-\bm{v}^2_{i,h}}_{L^2\LRp{\Omega}}}^{1/2}}{\LRp{\sum_{i=1}^m\nor{\bm{v}_i}^2_{L^2\LRp{\Omega}}}^{1/2}}.
+\bm{v}(\bm{x},t) = \pd{\bm{u}}{t}, \qquad \bm{\sigma} = \bm{C}\bm{E} = \bm{C}\frac{1}{2}\LRp{\Grad\bm{u} + \Grad\bm{u}^T}.  
+\]
+Unless otherwise stated, we report relative $L^2$ errors for all components of the solution $\bm{U} = (\bm{v},\bm{\sigma})$ 
+\[
+\frac{\nor{\bm{U}-\bm{U}_h}_{L^2\LRp{\Omega}}}{\nor{\bm{U}}_{L^2\LRp{\Omega}}} = \frac{\LRp{\sum_{i=1}^m \nor{\bm{U}_i-\bm{U}_{i,h}}^2_{L^2\LRp{\Omega}}}^{1/2}}{\LRp{\sum_{i=1}^m\nor{\bm{U}_i}^2_{L^2\LRp{\Omega}}}^{1/2}}.
 \]
@@
 Figure~\ref{fig:rayleigh} shows $L^2$ errors for both waves at time $T = 1$.
-Observed convergence rates fall between the optimal rate of $O(h^{N+1})$ and theoretical rate of $O(h^{N+1/2})$ \cite{johnson1986analysis,richter1988optimal}.
+Observed convergence rates fall between the optimal rate of $O(h^{N+1})$ and theoretical rate of $O(h^{N+1/2})$ \cite{johnson1986analysis}.
 
@@
 
-%We also examine the convergence of WADG for Rayleigh waves for near-incompressible materials with $\mu / \lambda \ll 1$.  We compute $L^2$ errors at final time $T = 1/(10\sqrt{\mu})$.
+\subsubsection{Rayleigh waves in near-incompressible materials}
+\label{sec:incomp}
+
+As noted in Section~\ref{sec:wadgelas}, error estimates for isotropic elasticity no longer hold in the incompressible limit with $\mu / \lambda \rightarrow 0$ limit due to the fact that $\bm{C}$ becomes singular.
+We use the propagation of Rayleigh waves to examine the behavior of WADG for near-incompressible materials.
+We follow \cite{sjogreen2012fourth,appelo2015energy} and fix $\lambda = 1$ and set $\mu = 1, .1, .01, .001, .0001$.
+Since the Rayleigh wave propagates with speed proportional to $\sqrt{\mu}$, we compute $L^2$ errors at the final time $1 / (4\sqrt{\mu})$ to ensure a fair comparison between solutions at different values of $\mu$.
+
+%Table~\ref{tab:incomp} shows total relative errors and relative errors for $\bm{v}$, $\bm{\sigma}$ individually at different orders and mesh sizes.
+While the relative errors for $\bm{\sigma}$ grow as $\mu/\lambda \rightarrow 0$, total errors and errors for $\bm{v}$ remain roughly the same magnitude for all values of $\mu$.
+Table~\ref{tab:incomp} shows relative errors for $\bm{v}$, $\bm{\sigma}$ at different orders and mesh sizes.
+The relative errors for $\bm{\sigma}$ grow as $\mu/\lambda \rightarrow 0$.
+This is not surprising, as the constant in the error estimates of Section~\ref{sec:mwadgprop} blows up in the incompressible limit.
+However, because the magnitude of $\bm{\sigma}$ also decreases as $\mu/\lambda\rightarrow 0$, relative errors for $\bm{v}$ remain roughly the same magnitude as the material approaches incompressibility.
+
+\begin{table}
+\centering
+\subfloat[Relative error in $\bm{v}$]{
+\begin{tabular}{|c||c|c|c|c|c|}
+\hline
+ & $\mu = 1$ & $\mu = .1$ & $\mu = .01$ & $\mu = .001$ & \mu = .0001$  \\
+\hhline{|=|=|=|=|=|=|}
+$N = 2$, $h = 1/4$ &  0.0288 &  0.0270 &  0.0309 & 0.0401 & 0.0568\\
+\hline
+$N = 3$, $h = 1/4$ & 0.0030 & 0.0027 &0.0030 & 0.0037 & 0.0049\\
+\hline
+$N = 4$, $h = 1/4$ & 2.7705e-04 & 2.5060e-04 & 2.9185e-04 &3.9779e-04 &5.2470e-04\\
+\hhline{|=|=|=|=|=|=|}
+$N = 2$, $h = 1/8$ & 0.0028  & 0.0025  & 0.0029 & 0.0036 & 0.0047 \\
+\hline
+$N = 3$, $h = 1/8$ & 1.8315e-04 & 1.6960e-04 &1.8524e-04 &2.2522e-04 & 2.8301e-04\\
+\hline
+$N = 4$, $h = 1/8$ & 8.2285e-06 & 7.5813e-06 &8.1174e-06 & 1.0762e-05&1.4886e-05\\
+\hline
+\end{tabular}
+}\\
+\subfloat[Relative error in $\bm{\sigma}$]{
+\begin{tabular}{|c||c|c|c|c|c|}
+\hline
+ & $\mu = 1$ & $\mu = .1$ & $\mu = .01$ & $\mu = .001$& \mu = .0001$ \\
+\hhline{|=|=|=|=|=|=|}
+$N = 2$, $h = 1/4$ & 0.0656 &  0.0754 &  0.1214 & 0.2134 & 0.4145\\
+\hline
+$N = 3$, $h = 1/4$ & 0.0091 & 0.0108 & 0.0182 & 0.0321& 0.0576\\
+\hline
+$N = 4$, $h = 1/4$ & 9.7408e-04 & 0.0012 &  0.0022 & 0.0041 &0.0068\\
+\hhline{|=|=|=|=|=|=|}
+$N = 2$, $h = 1/8$ & 0.0093 & 0.0116 &0.0204  & 0.0379 &0.0765  \\
+\hline
+$N = 3$, $h = 1/8$ & 6.8288e-04 & 8.0145e-04 & 0.0015 & 0.0031 &0.0056\\
+\hline
+$N = 4$, $h = 1/8$ &3.4231e-05 &  4.3510e-05 & 8.8747e-05 & 1.9771e-04& 3.6053e-04 \\
+\hline
+\end{tabular}
+}
+%\\
+%\subfloat[Total relative error ]{
+%\begin{tabular}{|c||c|c|c|c|c|}
+%\hline
+% & $\mu = 1$ & $\mu = .1$ & $\mu = .01$ & $\mu = .001$ & \mu = .0001$ \\
+%\hhline{|=|=|=|=|=|=|}
+%$N = 2$, $h = 1/4$ & 0.0539 & 0.0369  &  0.0339 & 0.0409 & 0.0571\\
+%\hline
+%$N = 3$, $h = 1/4$ & 0.0073 & 0.0046 &0.0037 &0.0039 & 0.0050\\
+%\hline
+%$N = 4$, $h = 1/4$ & 7.7419e-04 & 4.8735e-04 & 3.9193e-04 &4.2768e-04 &5.3111e-04\\
+%\hhline{|=|=|=|=|=|=|}
+%$N = 2$, $h = 1/8$ & 0.0074 & 0.0048  & 0.0038 &  0.0038 & 0.0048  \\
+%\hline
+%$N = 3$, $h = 1/8$ & 5.4120e-04 & 3.2760e-04 &2.6070e-04 & 2.5465e-04 & 2.9104e-04\\
+%\hline
+%$N = 4$, $h = 1/8$ & 2.7006e-05 & 1.7103e-05 &1.3344e-05 &1.3133e-05&1.5507e-05\\
+%\hline
+%\end{tabular}
+%}
+\caption{Behavior of WADG for linear elastic wave propagation in the incompressible limit as $\mu / \lambda \rightarrow 0$.  Errors are shown for various orders and mesh resolutions.  }
+\label{tab:incomp}
+\end{table}
+
+%\note{Examine energy stability as $\mu/\lambda\rightarrow 0$.  Matrix $\bm{C}$ becomes singular.}
 
@@
 We note that, when using piecewise constant approximations of $\mu$ and $\lambda$, we observe the same $O(h^{1/2})$ convergence rate, though errors are roughly twice as large in magnitude.
-\note{Discuss need to extend WADG to low regularity weights.}
+\note{Discuss possibly extending theory for WADG to low regularity weights.}
 
@@
 
-Extension to hybrid meshes is straightforward in 2D \cite{smith2010elastic}. For 3D, care must be taken with pyramids \cite{chan2015orthogonal,chan2015gpu}.
-
+Extension to hybrid meshes in 3D, care must be taken with pyramids \cite{chan2015orthogonal,chan2015gpu}.
+
+which can be further accelerated by exploiting structure under Bernstein-Bezier bases \cite{chan2015gpu}. 
 Formulation allows for the reduction of constant-coefficient RHS costs at high orders using BBDG \cite{chan2015bbdg}.
@@
 
-The author gratefully thanks Thomas Hagstrom, Tim Warburton, Axel Modave, and Ruichao Ye for informative discussions.
+The author gratefully thanks Thomas Hagstrom, Tim Warburton, Axel Modave, and Ruichao Ye for helpful and informative discussions.
 

### 2016-12-21 docs/dg_penalty/main.tex
local commits (not sure wtf this other stuff is)

@@
 
-%\documentclass[preprint,10pt]{elsarticle}
-\documentclass[10pt]{article}
-%\documentclass[review]{siamart0216}
+%\documentclass{amsart}
+\documentclass[preprint,10pt]{elsarticle}
+%\documentclass[10pt]{article}
+%\documentclass{siamart0216}
 
@@
 \author{Jesse Chan, T. Warburton}
-\title{On penalty flux parameters in first order DG formulations}
+\title{A short note on the penalty flux parameter for first order discontinuous Galerkin formulations}
 
@@
 
-\maketitle
-%\tableofcontents
+%\maketitle
+\begin{frontmatter}
+\title{A short note on the penalty flux parameter for first order discontinuous Galerkin formulations}
+
+\author[rice]{Jesse Chan\corref{cor1}}
+\ead{Jesse.Chan@caam.rice.edu}
+\cortext[cor1]{Principal Corresponding author}
+\author[vt]{T. Warburton}
+\ead{tcew@vt.edu}
+\address[rice]{Department of Computational and Applied Mathematics, Rice University, 6100 Main St, Houston, TX, 77005}
+\address[vt]{Department of Mathematics, Virginia Tech, 225 Stanger Street, Blacksburg, VA 24061-0123}
+
+\address{}
 
 \begin{abstract}
-Abstract goes here.  Mention assumptions: continuously varying material data, skew-symmetric formulations.  
+Penalty fluxes are dissipative numerical fluxes for high order discontinuous Galerkin (DG) methods which depend on a penalization parameter \cite{warburton2013low, ye2016discontinuous}.
+We investigate the dependence of the spectra of high order DG discretizations on this parameter, and show that as its value increases, the spectra of the DG discretization splits into two disjoint sets of eigenvalues.
+One set converges to the eigenvalues of a conforming discretization, while the other set corresponds to spurious eigenvalues which are damped proportionally to $\tau$.
+Numerical experiments also demonstrate that undamped spurious modes present in both in the limit of zero and large penalization parameters are damped for moderate values of the upwind parameter.
 \end{abstract}
+\end{frontmatter}
+
+%\begin{abstract}
+%Penalty fluxes are dissipative numerical fluxes for high order discontinuous Galerkin (DG) methods which depend on a penalization parameter \cite{warburton2013low, ye2016discontinuous}.
+We investigate the dependence of the spectra of high order DG discretizations on this parameter, and show that as its value increases, the spectra of the DG discretization splits into two disjoint sets of eigenvalues.
+One set converges to the eigenvalues of a conforming discretization, while the other set corresponds to spurious eigenvalues which are damped proportionally to $\tau$.
+Numerical experiments also demonstrate that undamped spurious modes present in both in the limit of zero and large penalization parameters are damped for moderate values of the upwind parameter.
+%\end{abstract}
 
@@
 
-\note{Intro}
+High order discontinuous Galerkin (DG) methods are a popular choice of discretization for simulations of time-domain wave propagation, due to their low numerical dispersion and their ability to accomodate unstructured meshes \cite{grote2006discontinuous,hesthaven2007nodal}.
+For PDEs in second order form, a common choice of DG method is the symmetric interior penalty (SIPG) discontinuous Galerkin method \cite{riviere2008discontinuous}, which involves a penalization parameter $\tau$.
+The role of this penalty parameter was investigated in \cite{Warburton20063205} for discretizations of the Maxwells equations.
+There, it was shown that in the limit of $\tau\rightarrow\infty$, the spectra of the DG discretization matrix splits into two disjoint sets of eigenvalues, with one set converging to eigenvalues of a conforming discretization and one set associated with divergent spurious eigenvalues.
+
+We consider penalty fluxes for discretizations of first order systems of hyperbolic PDEs, which depend on some parameter $\tau \geq 0$.
+For $\tau = 0$, these fluxes reduce to the energy-conserving and non-dissipative central fluxes \cite{fezoui2005convergence}.
+For $\tau > 0$, penalty fluxes are related to upwind fluxes for problems with continuous coefficients.
+We extend the results of \cite{Warburton20063205} to DG formulations for first order systems with penalty fluxes and show that the spectra of the DG discretization matrix again splits into two disjoint sets of eigenvalues as $\tau\rightarrow\infty$.
+The first set of eigenvalues correspond to the eigenvalues of a conforming discretization, while the second set of eigenvalues have real parts approaching $-\infty$ as $\tau\rightarrow \infty$ and correspond to damped spurious modes.
+Numerical experiments verify these results for the advection and acoustic wave equations in one and two dimensions.
+Numerical experiments also show that, while increasing $\tau$ pushes most eigenvalues further left of the imaginary axis, the real parts of certain eigenvalues converge to zero as $\tau\rightarrow\infty$.
+This implies that taking a moderate penalty parameter $\tau = O(1)$ damps spurious modes which remain otherwise undamped in time-dependent simulations with either central fluxes ($\tau = 0$) and conforming discretizations ($\tau \rightarrow \infty$).
 
@@
 \begin{equation}
-\bm{A}_0(\bm{x})\pd{\bm{U}}{t} + \sum_{i=1}^d \pd{\LRp{\bm{A}_{i}\bm{U}}}{\bm{x}_i} = 0
+\pd{\bm{U}}{t} + \sum_{i=1}^d \bm{A}_i\pd{{\bm{U}}}{\bm{x}_i} = 0
 \label{eq:cons}
 \end{equation}
-with solution components $\bm{U} = (u_1, \ldots, u_M)^T$.
-In this work, we assume that $\bm{A}_0$ is symmetric and positive-definite, and that the matrices $\bm{A}_i$ are symmetric or symmetrizable.
-In this work, we assume that $\bm{A}_i$ are constant, though $\bm{A}_0(\bm{x})$ may vary spatially.
-These assumptions are true for many PDEs governing wave propagation in heterogeneous media, including such as the acoustic wave equation \cite{chan2016weight1}, linear elasticity \cite{de2008interior,wilcox2010high,ye2016discontinuous}, and Maxwell's equations \cite{hesthaven2002nodal,grote2007interior,warburton2013low}.
-%\note{Restrict to constant coeff $\bm{A}_i$?}
-
-We assume the domain $\Omega$ is triangulated exactly using a mesh $\Oh$ consisting of elements $D^k$.  We define the $L^2$ inner product over elements $D^k$ and their boundaries $\partial D^k$ as
+with solution components $\bm{U} = (u_1, \ldots, u_M)^T$ and symmetric and spatially constant matrices $\bm{A}_i$.
+We assume the domain $\Omega$ is triangulated exactly using a mesh $\Oh$ consisting of elements $D^k$.
+We define the $L^2$ inner product over elements $D^k$ and their boundaries $\partial D^k$ as
 \[
@@
 \begin{equation}
-\sum_{D^k \in \Oh} \LRp{\bm{A}_0\pd{\bm{U}}{t},\bm{V} }_{\Lk} = \sum_{D^k \in \Oh}\LRp{ \LRp{\bm{A}_i(\bm{U}), \bm{V}_{,i}}_{\Lk} - \LRa{(\bm{A}_n\bm{U})^*,\bm{V}}_{\partial D^k}}
+\sum_{D^k \in \Oh} \LRp{\pd{\bm{U}}{t},\bm{V} }_{\Lk} = \sum_{D^k \in \Oh}\LRp{ \LRp{\bm{A}_i(\bm{U}), \bm{V}_{,i}}_{\Lk} - \LRa{(\bm{A}_n\bm{U})^*,\bm{V}}_{\partial D^k}}
 \label{eq:formulation}
@@
 where $(\bm{A}_n\bm{U})^*$ is a numerical flux defined on boundary faces and shared faces between elements.
+
+We note that, while the assumptions on constant $\bm{A}_i$ are restrictive, the results in this work are straightforward to generalize to a slight modification of (\ref{eq:cons})
+\begin{equation}
+\bm{A}_0\pd{\bm{U}}{t} + \sum_{i=1}^d \bm{A}_i\pd{{\bm{U}}}{\bm{x}_i} = \bm{f}.
+\label{eq:consA}
+\end{equation}
+We assume $\bm{A}_0 = \bm{I}$ in this work for brevity; however, all results are straightforward to extend to systems of the form for any $\bm{A}_0 = \bm{A}_0(\bm{x})$ which may vary spatially and is pointwise symmetric and positive-definite.
+Many models of wave propagation in general heterogeneous media can be represented in the form of (\ref{eq:consA}), including the acoustic wave equation \cite{chan2016weight1}, linear elasticity \cite{de2008interior,wilcox2010high,ye2016discontinuous}, and Maxwell's equations \cite{hesthaven2002nodal,hesthaven2004high,grote2007interior,warburton2013low}.
+%Results also follow for matrices $\bm{A}_i$ may be symmetrized through an appropriate change of variables \cite{kopriva2014energy}.
 
@@
 
-Let $f$ be a shared face between two elements $D^{k,+}$ and $D^{k,-}$ with outward normal $\bm{n}^-$ on $D^{k,-}$, and let $\bm{U}^+, \bm{U}^-$ be evaluations of $\bm{U}$ restricted to $D^{k,+}$ and $D^{k,-}$, respectively.
-The average of both scalar and vector-valued functions is defined as
+Let $f$ be a shared face between a given element $D^{k,-}$ and its neighbor $D^{k,+}$.
+We denote the outward normal on the face $f$ of $D^{k,-}$ as $\bm{n}$, and let $\bm{U}^+, \bm{U}^-$ be evaluations of $\bm{U}$ restricted to $D^{k,+}$ and $D^{k,-}$, respectively.
+The jump and average scalar-valued functions is defined as
 \[
-\avg{u} = \frac{{u}^+ + u^-}{2},
+\avg{u} = \frac{{u}^+ + u^-}{2}, \qquad \jump{u} = u^+ - u^-.
 \]
-and the jump of scalar and vector-valued functions over $f$ are defined as
+Define the normal matrix 
 \[
-\jump{u} = u^+ \bm{n}^+ + u^-\bm{n}^-, \qquad \jump{\bm{u}} = \bm{u}^+\cdot \bm{n}^+ + \bm{u}^-\cdot\bm{n}^-.  
+{\bm{A}}_n = \sum_{i=1}^d {\bm{A}_i\bm{n}_i}.
 \]
-Typical numerical fluxes are defined as the sum of the average of $\bm{U}^+$ and $\bm{U}^-$ and a penalization term $\bm{G(\bm{U}^+,\bm{U}^-)}$
-\begin{equation}
-(\bm{A}_n\bm{U})^* = \avg{\bm{A}_n\bm{U}}\cdot \bm{n}^- - \bm{G(\bm{U}^+,\bm{U}^-)}.
-\label{eq:flux}
-\end{equation}
-\note{Clean up jump notation here.}
-The upwind numerical flux is a well-known flux of the form (\ref{eq:flux}).
-For an outward normal vector $\bm{n}^-$, let ${\bm{A}}_n = \sum_{i=1}^d {\bm{A}_i\bm{n}^-_i}$.
+The penalty numerical flux is then defined as
+\[
+(\bm{A}_n\bm{U})^* = \bm{A}_n\avg{\bm{U}} - \frac{\tau}{2} \bm{A}_n^T \bm{A}_n\jump{\bm{U}}.
+\]
+where $\tau \geq 0$.
+We compare the penalty flux to the well-known upwind numerical flux for the case of continuous coefficients.
 By symmetry of $\bm{A}_i$, ${\bm{A}}_n$ contains real eigenvalues, and admits an eigenvalue decomposition
@@
 \]
-For problems with continuous coefficients, the upwind numerical flux over a face $f \in \Gh$ can be defined  as
+For problems with continuous coefficients, the upwind numerical flux over a face $f \in \Gh$ can then be defined  as
 \[
@@
 \[
-(\bm{A}_n\bm{U})^* = \avg{\bm{A}_n\bm{U}} - \bm{V}\LRb{\bm{\Lambda}}\bm{V}^T \jump{\bm{U}}.  
+(\bm{A}_n\bm{U})^* = \bm{A}_n\avg{\bm{U}} - \frac{1}{2}\bm{V}\LRb{\bm{\Lambda}}\bm{V}^T \jump{\bm{U}}.  
 \]
@@
 %\]
-We consider penalty fluxes as an alternative to upwind fluxes, which penalize appropriately defined jumps of the solution.  One such penalty flux is 
+We note that including a positive-definite weight matrix $\bm{W}$ into the penalty term
 \[
-(\bm{A}_n\bm{U})^* = \avg{\bm{A}_n\bm{U}} - \frac{\tau}{2} \bm{A}_n^T \jump{\bm{A}_n\bm{U}}.  
+\frac{\tau}{2} \bm{A}_n^T \bm{W}\bm{A}_n\jump{\bm{U}}
 \]
-where $\tau$ is proportional to $\lambda_{\max}^{-1}$.
-This choice of $\tau$ ensures that the penalty term is of the same order of magnitude as the upwind penalization, and that the CFL restriction for both upwind and penalty fluxes are identical.
-Unlike upwind fluxes, penalty fluxes can be used when a diagonalization of $\bm{A}_n$ is not explicitly known, such as linear elasticity in anisotropic media \cite{ye2016discontinuous}.
+recovers the upwind fluxes for $\tau = 1$ and $\bm{W} = \bm{V}\LRb{\bm{\Lambda}}^{\dagger}\bm{V}^{-1}$, where $\bm{A}^{\dagger}$ is the pseudo-inverse of $\bm{A}$.
+However, the advantage of penalty fluxes is that, unlike upwind fluxes, they can be used when explicit knowledge of a diagonalization of $\bm{A}_n$ is not known, such as for linear elasticity in anisotropic media \cite{ye2016discontinuous}.
+Thus, we assume $\bm{W} = \bm{I}$ and require that $\bm{V}\LRb{\bm{\Lambda}}\bm{V}^T$ and $\tau\bm{A}_n^T\bm{A}_n$ are of the same magnitude, which implies that $\tau$ should be proportional to $1/\LRp{\max_i\LRb{\lambda_i}}$.
+This ensures that the penalty term is of the same magnitude as the upwind penalization term, and that CFL restrictions in explicit time-stepping methods will be identical for both upwind and penalty fluxes.
 
@@
 The remainder of this work focuses on the influence of $\tau$ on the spectra of DG discretization matrices.
+We will show that, as $\tau$ increases, the spectra splits into two sets of eigenvalues corresponding respectively to conforming and spurious modes, where conformity is defined based on $\bm{A}_n$.
 
-\subsection{Conforming and non-conforming spaces}
+\subsection{Conforming and non-conforming approximation spaces}
 
@@
 \]
-We define a decomposition of $V$ into a conforming part $V^C$ and a non-conforming part $V^{NC}$ based on the penalization term $\bm{G}(\bm{U}^+,\bm{U}^-)$.
-The conforming part is defined as the null space of $\bm{G}(\bm{U}^+,\bm{U}^-)$
+We decompose $V$ into a conforming space $V^C$ and a non-conforming space $V^{NC}$ based on the penalization term.
+The conforming space is defined to be elements of $V$ for which the penalization term vanishes over all faces of all elements.
+For penalty fluxes, this implies $V^C$ is
 \begin{equation}
-V^{C} = \LRc{ \bm{U}(\bm{x}) \in V: \quad  \bm{G}(\bm{U}^+,\bm{U}^-) = 0, \quad \forall f \in \Gh},
+V^{C} = \LRc{ \bm{U}(\bm{x}) \in V: \quad  \bm{A}_n\jump{\bm{U}} = 0, \quad \forall f \in \partial D^k, \quad \forall D^k\in \Oh}.
 \label{eq:conf}
 \end{equation}
-while the non-conforming approximation space $V^{NC}$ is defined as $L^2$ orthogonal complement of $V^{C}$ in $V$
+The non-conforming approximation space $V^{NC}$ is defined as $L^2$ orthogonal complement of $V^{C}$ in $V$
 \[
@@
 \]
-When the coefficient matrices $\bm{A}_i$ are spatially continuous, the upwind flux condition $\bm{V}\LRb{\bm{\Lambda}}\bm{V}^T \jump{\bm{U}} = 0$ and the penalty flux condition $\bm{A}_n^T \jump{\bm{A}_n\bm{U}} = 0$ are both equivalent to
+A conforming space could also be defined using the upwind flux by seeking $\bm{U}$ such that $\bm{V}\LRb{\bm{\Lambda}}\bm{V}^T \jump{\bm{U}} = 0$.
+When the coefficient matrices $\bm{A}_i$ are spatially continuous, the condition $\bm{V}\LRb{\bm{\Lambda}}\bm{V}^T \jump{\bm{U}} = 0$ and the penalty flux condition $\bm{A}_n^T \bm{A}_n\jump{\bm{U}} = 0$ are both equivalent to
 \[
@@
 
-For problems with discontinuous material data, it is less straightforward to show equivalence between the conforming spaces $V^C$ resulting from upwind and penalty fluxes.
-However, for many practical wave propagation settings, spatially varying coefficients are incorporated into the matrix $\bm{A_0}$, resulting in constant $\bm{A}_i$.
-For both smoothly varying and discontinuous $\bm{A}_0$, it is possible to formulate energy stable DG methods with penalty fluxes \cite{mercerat2015nodal, chan2016weight1}.
-Additionally, DG methods using penalty fluxes are observed to perform similarly to DG methods with upwind fluxes for  wave propagation problems with discontinuous coefficients \cite{warburton2013low, chan2016weight1, ye2016discontinuous}.
+We note that, for problems with discontinuous material data, it is less straightforward to show equivalence between the conforming spaces $V^C$ resulting from upwind and penalty fluxes.
+However, penalty fluxes can still be applied in the presence of discontinuous coefficients by incorporating these spatial variations into the matrix $\bm{A_0}$.
+When paired with penalty fluxes, this approach results in energy stable DG methods \cite{chan2016weight1}.
+Additionally, DG methods with penalty fluxes are observed to perform similarly to DG methods with upwind fluxes for several realistic wave propagation problems with discontinuous coefficients \cite{ye2016discontinuous}.
 
@@
 \label{sec:confexamples}
-We illustrate the conforming spaces induced by penalty fluxes using the scalar advection equation and acoustic wave equation as concrete examples.  The scalar advection equation is given as
+We illustrate the conforming spaces induced by penalty fluxes using the scalar advection equation and acoustic wave equation as examples.  The scalar advection equation is given as
 \[
-\pd{u}{t} + \div\LRp{\bm{\beta}u} = 0.
+\pd{u}{t} + \Div\LRp{\bm{\beta}u} = 0.
 \]
-where $\bm{\beta}$ is the direction of advection.
-For continuous advection vectors such that $\bm{\beta}(\bm{x}) \neq 0$, the penalty flux is equivalent to a parametrized upwind flux \cite{hesthaven2007nodal}
+where $\bm{\beta}$ is the direction of advection.  For continuous advection vectors such that $\bm{\beta}(\bm{x}) \neq 0$, the penalty flux is 
 \[
-(\bm{A}_n\bm{U})^* = \avg{\bm{\beta}_n u} - \tau\LRb{\bm{\beta}_n}\jump{u}.
+(\bm{A}_n\bm{U})^* = \bm{\beta}_n\avg{ u} - \tau\LRb{\bm{\beta}_n}\jump{u}.
 \]
@@
 \left(\begin{array}{ccc}
-0 & \bm{n}^-_x & \bm{n}^-_y\\
-\bm{n}^-_x & 0 & 0\\
-\bm{n}^-_y & 0 & 0
+0 & \bm{n}_x & \bm{n}_y\\
+\bm{n}_x & 0 & 0\\
+\bm{n}_y & 0 & 0
 \end{array}
@@
 \[
-\avg{\bm{A}_n \bm{U}} - \frac{\tau}{2}\bm{A}_n^T \jump{\bm{A}_n\bm{U}} = \left(\begin{array}{c}
+\bm{A}_n \avg{\bm{U}} - \frac{\tau}{2}\bm{A}_n^T\bm{A}_n \jump{\bm{U}} = \left(\begin{array}{c}
 \avg{\bm{u}_n}\\
-\avg{p }\bm{n}^-_x\\
-\avg{p }\bm{n}^-_y
+\avg{p }\bm{n}_x\\
+\avg{p }\bm{n}_y
 \end{array}
@@
 \jump{p}\\
-\jump{\bm{u_n}}\bm{n}^-_x\\
-\jump{\bm{u_n}}\bm{n}^-_y
+\jump{\bm{u_n}}\bm{n}_x\\
+\jump{\bm{u_n}}\bm{n}_y
 \end{array}
@@
 \]
-For $\tau = 1$, these fluxes coincide with the upwind fluxes for continuously varying media.
+Since the eigenvalues of $\bm{A}_n$ are $-1,0,1$, we take the penalty parameter to be $\tau = 1 / (\max_i |\lambda_i|) = 1$.
+Additionally, for $\tau = 1$, the penalty fluxes coincide with the upwind fluxes for continuously varying media.
 In this case, the conforming subspace $V^{C}$ induced by the penalty flux consists of a scalar (pressure) component $V^C_p$ and a vector (velocity) component $V^C_{\bm{u}}$.
@@
 For polynomial approximation spaces, this implies that $p$ is continuous across faces, edges, and vertices, and $V^C_p$ is the standard $C^0$ continuous piecewise polynomial finite element space.
-
-For the velocity component $V_{\bm{u}}^C$, the penalty fluxes enforce normal continuity, such that $\jump{\bm{u}}\cdot \bm{n}^- = 0$ over all faces $f \in \Gh$.
-Since each component of $\bm{u}$ is approximated from $P^N(D^k)$, this implies that $V^C_{\bm{u}}$ is the $H({\rm div})$-conforming Brezzi-Douglas-Marini finite element space \cite{brezzi1985two,boffi2013mixed}.\footnote{It is possible to recover $V^C_{\bm{u}}$ as the Raviart-Thomas finite element space by approximating each component of $\bm{u}$ from a smaller polynomial space \cite{kirby2004algorithm}.}
+For the velocity component $V_{\bm{u}}^C$, the penalty fluxes enforce normal continuity, such that $\jump{\bm{u}}\cdot \bm{n} = 0$ over all faces $f \in \Gh$.
+Since each component of $\bm{u}$ is approximated from $P^N(D^k)$, this implies that $V^C_{\bm{u}}$ is the $H({\rm div})$-conforming Brezzi-Douglas-Marini finite element space \cite{brezzi1985two,boffi2013mixed}.\footnote{It is possible to recover $V^C_{\bm{u}}$ as the Raviart-Thomas finite element space by approximating each component of $\bm{u}$ from a different polynomial space \cite{kirby2004algorithm}.}   These reflect minimal continuity conditions which result from requiring $\Div \bm{u} \in \LRp{L^2\LRp{\Oh}}^d$.
 
@@
 \avg{\bm{u}_n}\\
-\avg{p }\bm{n}^-_x\\
-\avg{p }\bm{n}^-_y
+\avg{p }\bm{n}_x\\
+\avg{p }\bm{n}_y
 \end{array}
@@
 \end{align*}
-For exact quadrature, constant $\bm{A}_i$, and $\tau = 0$, the DG formulation is skew-symmetric \cite{warburton2013low, chan2015gpu, kopriva2014energy}, implying that $\bm{A},\bm{C}$ are skew-symmetric and that $\tilde{\bm{B}} = -\bm{B}$.
-Additionally, the entries of the penalization matrix $\bm{S}$ can be rewritten as
+For exact quadrature, constant $\bm{A}_i$, and $\tau = 0$, the DG formulation is skew-symmetric \cite{warburton2013low, chan2015gpu, chan2016weight2, kopriva2014energy}, implying that $\bm{A},\bm{C}$ are skew-symmetric and that $\tilde{\bm{B}} = -\bm{B}$.
+Additionally, the entries of the penalization matrix $\bm{S}$ can be rewritten as a sum over all unique faces of elements in the mesh
 \[
@@
 \]
-implying that $\bm{S}$ is symmetric and negative-definite.
+for consistently defined jumps and normal vectors $\bm{n}$  over each unique face.  This implies $\bm{S}$ is symmetric and negative-definite.
 
@@
 As a consequence, $\bm{W}^{NC} \rightarrow 0$ as $\tau\rightarrow \infty$, and the smallest $N^C$ eigenvalues of $\bm{K}$ converge to the eigenvalues of $\bm{A}$ at a rate of $O(1/\tau)$.
-These correspond to a discretization using the conforming approximation space (\ref{eq:conf}).
+These eigenvalues of $\bm{A}$ correspond to a discretization using the conforming approximation space (\ref{eq:conf}).
 
@@
 As $\tau \rightarrow \infty$, the spectrum of the DG discretization matrix $\bm{K}$ decouples into two sets of eigenvalues $\LRc{\lambda^C_1,\ldots,\lambda^C_{N^C}}$ and $\LRc{\lambda^{NC}_1,\ldots,\lambda^{NC}_{N^{NC}}}$.
-The eigenvalues $\LRc{\lambda^{NC}_1,\ldots,\lambda^{NC}_{N^{NC}}}$ diverge towards the left half plane, with $\LRb{{\rm Re}(\lambda_i)} = O(\tau)$, while  the eigenvalues $\LRc{\lambda^C_1,\ldots,\lambda^C_{N^C}}$ converge to the eigenvalues of $\bm{A}$ at the rate $O(1/\tau)$.
+The eigenvalues $\LRc{\lambda^{NC}_1,\ldots,\lambda^{NC}_{N^{NC}}}$ diverge towards the left half plane, with $\LRb{{\rm Re}(\lambda_i)} = O(\tau)$, while  the eigenvalues $\LRc{\lambda^C_1,\ldots,\lambda^C_{N^C}}$ converge to the eigenvalues of $\bm{A}$ with rate $O(1/\tau)$.
 \label{lemma:eig}
@@
 
-The above analysis illustrates the behavior of the spectra of the DG discretization for the asymptotic cases when $\tau = 0$ or $\tau \rightarrow \infty$.
+Lemma~\ref{lemma:eig} illustrates the behavior of the spectra of the DG discretization for the asymptotic cases when $\tau = 0$ or $\tau \rightarrow \infty$.
 However, it is less clear how the spectra of $\bm{K}$ behaves for $\tau \approx 1$.
@@
 \centering
-\subfloat[Paths of eigenvalues as $\tau$ increases]{\includegraphics[width=.46\textwidth]{figs/trackedEigs.png}}
+\subfloat[Paths of eigenvalues as $\tau$ increases]{\includegraphics[width=.48\textwidth]{figs/trackedEigs.png}}
 \hspace{1em}
-\subfloat[Zoom of eigenvalues near imaginary axis]{\includegraphics[width=.44\textwidth]{figs/trackedEigsZoom.png}}
+\subfloat[Zoom of eigenvalues near imaginary axis]{\includegraphics[width=.435\textwidth]{figs/trackedEigsZoom.png}}
 \caption{Eigenvalue paths for DG advection with $\tau \in [0,4]$ on a mesh of 8 elements of degree $N=3$.
@@
 \subfloat[$\tau = 10$]{\includegraphics[width=.32\textwidth]{figs/tauModesDiverge3.png}}
-\caption{Behavior of divergent eigenvalues for advection as $\tau$ increases.
+\caption{Behavior of a mode for the advection equation corresponding to a divergent eigenvalue (indicated by the red star) as $\tau$ increases.
 The order of approximation is $N=3$ on a mesh of 4 elements.} %The displayed eigenfunctions correspond to the highlighted eigenvalue.
@@
 These eigenvalues which move right to return to the imaginary axis correspond to a second set of \emph{spurious} modes.
-These modes are related to spurious modes observed for high order $C^0$ finite element methods \cite{ainsworth2014dispersive}, which consist of higher frequency components with sharp peaks.
+These modes closely resemble back-propagating spurious modes observed for high order $C^0$ finite element methods \cite{ainsworth2014dispersive}, which consist of higher frequency components with sharp peaks.
 Since the real part of these eigenvalues approaches zero as $\tau\rightarrow \infty$, these modes become undamped and persist as spurious solution components in time-domain simulations \cite{hughes2014finite}
@@
 
-We repeat the above experiment for the acoustic wave equation in 1D with reflection boundary conditions on $[-1,1]$, and compare against the exact eigenvalues $\lambda_i = i \frac{\pi}{2}$ for $i = 1,2,\ldots$.
-The behavior of the eigenvalues as $\tau$ increases (Figure~\ref{fig:trackModesWave}) is similar to the behavior observed for the scalar advection equation.
-An examination of the eigenmodes corresponding to divergent and spurious eigenvalues as $\tau \rightarrow \infty$ shows that these modes behave similarly to the spurious modes observed for the scalar advection equation, with individual pressure and velocity components converging to continuous functions.
+Figure~\ref{fig:spurious} also indicates that taking $\tau = O(1)$ suppresses spurious non-dissipative modes associated with both $\tau = 0$ (central fluxes) and $\tau\rightarrow \infty$ (conforming discretizations).
+Let $\lambda^{\tau}_i, \bm{v}^{\tau}_i$ be eigenvalues and eigenvectors of the DG discretization matrix $\bm{K}$ with penalty $\tau$.
+The exact solution to the semi-discrete formulation is
+\[
+\bm{u}(t) =\bm{V}^{\tau}\LRp{ \sum_j c_j e^{-\lambda_i^{\tau} t} \bm{v}^{\tau}_j}, \qquad \bm{u}(0) = \sum_j c_j\bm{v}^{\tau}_j,
+\]
+which illustrates how modes $\bm{v}^{\tau}_j$ are damped for $t > 0$ if the corresponding eigenvalue $\lambda^{\tau}_j$ has a negative real part .
+We observe that, when spurious modes of the solution for $\tau = 0$ or $\tau \gg 1$ are expanded in the modes for $\tau = 1$
+\[
+\bm{v}^{\tau}_i = \sum_j c^\tau_j \bm{v}^{1}_j, 
+\]
+the largest components correspond to eigenvalues with larger negative real parts and heavily damped modes.
+As an example, we consider the spurious mode illustrated in Figure~\ref{fig:spurious}.
+The absolute values of the coefficients $c^\tau_j$ of this mode are plotted in Figure~\ref{fig:cmodes} for both $\tau = 0$ and $\tau = 100$.
+Only the four coefficients larger than $10^{-13}$ are shown.
+For comparison, the negative real values of the corresponding eigenvalues $\lambda^1_j$ are also overlaid in Figure~\ref{fig:cmodes}.
+The two largest coefficients correspond to eigenvalues with large negative real parts and more strongly damped modes.
+The two remaining coefficients correspond to less highly damped eigenvalues with smaller negative real parts, but are an order of magnitude smaller than the two large coefficients.
+\begin{figure}
+\centering
+\subfloat[$c_j^0$ (spurious modes for  $\tau = 0$)]{\includegraphics[width=.45\textwidth]{figs/ceigs0.png}}
+\hspace{1em}
+\subfloat[$c_j^{100}$ (spurious modes for $\tau = 100$)]{\includegraphics[width=.45\textwidth]{figs/ceigs100.png}}
+\caption{Coefficients $c^0_j$ and $c^{100}_j$ of the spurious mode in Figure~\ref{fig:spurious} when expanded in the modes of $\bm{K}$ for $\tau=1$.
+Coefficients with magnitude less than $10^{-13}$ are not shown.
+The negative real parts of the corresponding eigenvalues ${\rm Re}(\lambda^1_j)$, which determine the magnitude of damping applied to the $j$th component, are overlaid.
+}
+\label{fig:cmodes}
+\end{figure}
+%\note{Show that both $\tau = 0, \infty$ spurious modes have large damped contributions.}
+
+The above experiments were repeated for the acoustic wave equation in 1D with reflection boundary conditions on $[-1,1]$, with results similar to those observed for the scalar advection equation.
+%and an examination of the eigenmodes corresponding to divergent and spurious eigenvalues as $\tau \rightarrow \infty$ shows that these modes behave similarly to the spurious modes observed for the scalar advection equation, with individual pressure and velocity components converging to continuous functions.
 
@@
 In two space dimensions, taking $\tau\rightarrow \infty$ again results in spurious modes which return to the imaginary axis, with the corresponding eigenmodes approaching conforming functions.
-We illustrate this through numerical experiments for the acoustic wave equation and advection in two space dimensions.
-In all numerical experiments, we use $N=3$ and a uniform triangular mesh resulting from bisecting each element of a $2\times 2$ uniform quadrilateral mesh.
+We illustrate this through numerical experiments with the acoustic wave equation and advection equation in two space dimensions.
+In all numerical experiments, we use $N=3$ and a uniform triangular mesh resulting from bisecting each element of a uniform quadrilateral mesh.
 
@@
 Likewise, the non-zero velocity components of spurious eigenmodes converge to approximations in $H({\rm div}; \Oh)$, such that only the normal component of velocity is continuous across faces.
-Figure~\ref{fig:trackModesWaveU} shows visualizations of these eigenmodes.
+
+%Figure~\ref{fig:trackModesWaveU} shows visualizations of these eigenmodes.
 For large $\tau$, normal jumps of the velocity component vanish.
@@
 
-\begin{figure}
-\centering
-\subfloat[$\tau = .1$, $\lambda = -0.6009 +11.3209i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_u1.png}}
-\hspace{.5em}
-\subfloat[$\tau = 1$, $\lambda =  -1.9601 + 9.5629i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_u2.png}}
-\hspace{.5em}
-\subfloat[$\tau = 100$, $\lambda = -0.0634 + 7.7670i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_u3.png}}
-\caption{Behavior of the real part of the velocity component of spurious eigenmodes for the acoustic wave equation in two dimensions.  }
-\label{fig:trackModesWaveU}
-\end{figure}
+%\begin{figure}
+%\centering
+%\subfloat[$\tau = .1$, $\lambda = -0.6009 +11.3209i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_u1.png}}
+%\hspace{.5em}
+%\subfloat[$\tau = 1$, $\lambda =  -1.9601 + 9.5629i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_u2.png}}
+%\hspace{.5em}
+%\subfloat[$\tau = 100$, $\lambda = -0.0634 + 7.7670i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_u3.png}}
+%\caption{Behavior of the real part of the velocity component of spurious eigenmodes for the acoustic wave equation in two dimensions.  }
+%\label{fig:trackModesWaveU}
+%\end{figure}
 
@@
 Figure~\ref{fig:trackModesAdvec} shows the spectra of the DG matrix for $\tau = 0,1,50$, with eigenvalue paths for increasing values of $\tau$ overlaid.
+Unlike spurious conforming modes present for the acoustic wave equation, spurious conforming modes for the advection equation satisfy $\LRb{\bm{\beta}_n}\jump{ u} = 0$, and allow discontinuities along faces which lie tangential to flow directions (Figure~\ref{fig:trackModesAdvecU}).
 
@@
 \end{figure}
-
-Unlike spurious conforming modes present for the acoustic wave equation, spurious conforming modes for the advection equation satisfy $\jump{\beta_n u} = 0$, and allow discontinuities along faces which are tangential to flow directions, as shown in Figure~\ref{fig:trackModesAdvecU}.
-%\note{Find cite on regularity of advection solutions.}
 
@@
 
-Numerical experiments show that, for $\tau = O(1)$, there exist eigenvalues with negative real part which return to the imaginary axis as $\tau \rightarrow 0$ or $\tau\rightarrow \infty$.
+Numerical experiments also demonstrate that, for $\tau = O(1)$, there exist eigenvalues with negative real part which return to the imaginary axis as $\tau \rightarrow 0$ or $\tau\rightarrow \infty$.
 The corresponding eigenmodes of such eigenvalues can be interpreted as spurious under-resolved modes which persist in time-domain simulations.
-These results suggest that the use of upwind and penalty fluxes damps spurious modes present not only for $\tau = 0$ (central fluxes), but also for $\tau \rightarrow\infty$ (a conforming discretization).
+These results qualify how the use of upwind and penalty fluxes dampens spurious modes which would otherwise persist for both $\tau = 0$ (central fluxes) and $\tau \rightarrow\infty$ (a conforming discretization).
 

### 2016-12-21 docs/elasticity_notes/main.tex
local commits (not sure wtf this other stuff is)

@@
 \renewcommand{\div}{\grad \cdot}
+
+\newcommand{\bs}[1]{\boldsymbol{#1}}
+\newcommand{\equaldef}{\stackrel{\mathrm{def}}{=}}
+
 \newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}{\rm #2}}}
@@
 \newcommand{\pdd}[2]{\frac{\partial^2#1}{\partial#2^2}}
-
-\newcommand{\bs}[1]{\boldsymbol{#1}}
-
-\newcommand{\equaldef}{\stackrel{\mathrm{def}}{=}}
-
-\newcommand{\mi}{\mathrm{i}} %% roman "i"
 \newcommand{\mb}[1]{\mathbf{#1}}
@@
 
-\renewcommand{\Oh}{{\Omega_h}}
+\newcommand{\Oh}{{\Omega_h}}
 \renewcommand{\L}{L^2\LRp{\Omega}}
-%\newcommand{\Lk}{\Lk}
+%\newcommand{\Lk}{L^2\LRp{D^k}}
 %\newcommand{\Ldk}{L^2\LRp{\partial D^k}}
@@
 \author{Jesse Chan}
-\title{Weight-adjusted discontinuous Galerkin methods: matrix-valued weights and linear elastic wave propagation}
+\title{Weight-adjusted discontinuous Galerkin methods: matrix-valued weights and elastic wave propagation in heterogeneous media}
 
@@
 In this work, we extend weight-adjusted DG (WADG) methods to weighted inner products with matrix-valued weights, focusing on the linear elastic wave equation as an application.
-We present a new DG formulation based on the discretization of the symmetric form of the linear elasticity wave equation, with upwind-like dissipation incorporated through simple penalty terms in the numerical fluxes.
-This results in provably stable formulations for arbitrary anisotropic stiffness tensors and curvilinear meshes.
+We present a unified DG formulation for isotropic and anisotropic media based on the discretization of the symmetric form of the linear elasticity wave equation, with upwind-like dissipation incorporated through simple penalty terms in the numerical fluxes.
+This formulation is provably stable for arbitrary stiffness tensors and curvilinear meshes.
 A convergence analysis is given, and numerical results confirm the stability and high order accuracy of WADG for several problems in elastic wave propagation.
@@
 
-There exist many high order energy stable discontinuous Galerkin formulations for the elastic wave equations in first order form.
-Wilcox et al.\ introduced a discontinuous Galerkin spectral element method on hexahedral meshes
-
-SEM \cite{komatitsch2010high}
-
-DG-SEM \cite{wilcox2010high}
-
-High order DG \cite{kaser2006arbitrary, dumbser2006arbitrary, de2007arbitrary, ye2016discontinuous}
-
-Interior penalty formulations.  Analysis: Riviere et al.\ \cite{riviere2003discontinuous,riviere2007discontinuous} and Delcourte et al.\ \cite{delcourte2009high,delcourte2015analysis}
-
-GPU-accelerated DG methods \cite{klockner2009nodal,godel2010gpu,godel2010scalability}.  Also elastic wave propagation and seismic wave applications \cite{modave2015nodal, modave2016gpu}
-
-High order approximations of material data \cite{castro2010seismic, mercerat2015nodal}, but not low storage or not energy stable.
-
-Extending mass lumping without dispersive effects \cite{christon1999influence,guermond2013correction}.  Reduces to exact mass matrix inversion for constant weights.
-
-A simpler extension of WADG \cite{chan2016weight1,chan2016weight2} to elasticity and weight-adjusted inner products for vector-valued functions.
-
-Outline: present elasticity and an energy stable DG method, motivate need for alternative to inverting large dense coupled mass matrix, introduce weight-adjusted inner products.
+Efficient and accurate simulations of elastic wave propagation form the foundation for various seismic applications, including imaging and inversion techniques, as well as rupture and earthquake simulations.
+The most common methods for the simulation of waves are finite difference methods \cite{virieux1986p}; however, finite element methods have garnered interest due to their low numerical dispersion and ability to accomodate geometrically flexible unstructured meshes.
+
+Typical methods for time-domain wave propagation utilize explicit time stepping methods, since the hyperbolic partial differential equations (PDEs) governing wave propagation typically admit less restrictive stable timestep restrictions than parabolic equations.
+However, finite element methods require the inversion of a global mass matrix when paired with explicit time integrators.
+On hexahedral meshes, this can be avoided by using spectral element methods (SEM), which utilize mass lumping to produce a diagonal mass matrix \cite{komatitsch1998spectral}.
+This global inversion can also be avoided through the use of discontinuous Galerkin (DG) methods, which result in a locally invertible block diagonal mass matrix.
+Due to difficulties in extending mass-lumping techniques to tetrahedra, high order DG methods are often employed for seismic simulations which use simplicial meshes \cite{kaser2006arbitrary, dumbser2006arbitrary, de2007arbitrary,delcourte2009high,delcourte2015analysis,ye2016discontinuous}.
+Such methods also lend themselves well to efficient implementations using Graphics Processing Units (GPUs) \cite{klockner2009nodal, modave2015nodal, modave2016gpu, chan2015gpu}.
+
+Most high order DG methods on simplicial meshes assume that material coefficients are constant over each element, which can result in spurious reflections and the loss of high order accuracy \cite{castro2010seismic,mercerat2015nodal,bencomo2015discontinuous}.
+This limitation can be overcome for DG methods by incorporating sub-element variation in media through weighted mass matrices, which results in an energy stable method \cite{mercerat2015nodal, bencomo2015discontinuous}.
+For DG methods on hexahedral meshes, mass lumping reduces the construction of weighted mass matrices to entry-wise scalings of a diagonal mass matrix.
+However, on tetrahedral meshes, this approach necessitates the pre-computation and storage of factorizations for each local mass matrix, which greatly increases both storage costs and data transferred at high orders of approximation.
+These costs are especially problematic for accelerator architectures such as GPUs, which possess limited memory.
+%Additionally, excessive storage costs can require reduced problem sizes and under-utilization of accelerator cards.
+
+Storage costs associated with weighted mass matrices can be avoided by approximating weighted $L^2$ inner products using weight-adjusted inner products, which result in easily invertible approximations to weighted mass matrices \cite{chan2016weight1,chan2016weight2}.
+For sufficiently regular weights, high order accuracy is also retained.
+When paired with an energy stable DG formulation, weight-adjusted inner products result in weight-adjusted DG methods (WADG), which preserve energy stability and high order accuracy while retaining a low asymptotic storage cost.
+Additionally, unlike mass-lumping techniques, WADG does not rely on the use of inexact quadrature rules, and reduces to the exact inversion of mass matrices for constant weights.
+
+Weight-adjusted DG methods have been applied to both acoustic wave propagation in heterogeneous media and on curvilinear meshes \cite{chan2016weight1,chan2016weight2}.
+Both cases involve scalar weighting functions.
+In this work, we extend weight-adjusted inner products to matrix-valued weights, and demonstrate the usefulness of this generalization to elastic wave propagation in arbitrary media.
+A new energy stable DG formulation is introduced based on the symmetric form of the elastic wave equations, with upwind-like numerical dissipation introduced through simple penalty fluxes \cite{warburton2013low,ye2016discontinuous}.
+
+This work proceeds as follows.
+Sections~\ref{sec:symelas} and \ref{sec:dgform} begin by presenting an energy stable DG formulation with simple penalty fluxes for the elastic wave equation in symmetric hyperbolic form, and discusses issues related to storage and inversion of local mass matrices for material coefficients with sub-element variations.
+Section~\ref{sec:mwadg} extends weight-adjusted approximations to weighted $L^2$ inner products and mass matrices to the case of matrix-valued weights, and proves interpolation estimates which account for the regularity of the matrix weight.
+These results are incorporated into a weight-adjusted DG method for the linear elastic wave equations in Section~\ref{sec:wadgelas}.
+Finally, numerical results in Section~\ref{sec:numerical} demonstrate the accuracy of this method for several analytic solutions in linear elasticity.
 
 \section{Symmetric form of the elastic wave equation }
-
-We begin with the linear elastic wave equation under small deformations in a domain $\Omega$.
+\label{sec:symelas}
+We begin with the linear elastic wave equation under small deformations in a domain $\Omega \in \mathbb{R}^d$.
 These equations can be written a first order velocity-stress system for velocity $\bm{v}$ and symmetric stress tensor $\bm{S}$
@@
 \rho \pd{\bm{v}}{t} &= \Div{\bm{S}} + \bm{f}\\
-\pd{\bm{S}}{t} &= \frac{1}{2} \bar{\bm{C}}\LRp{\Grad\bm{v} + \Grad\bm{v}^T},
-\end{align*}
-where $\bm{f}$ is the body force per unit volume, $\rho$ is density, and $\bar{\bm{C}}$ is the symmetric constitutive stiffness tensor relating stress and strain.
-These equations can also be rewritten as a symmetric hyperbolic system of PDEs \cite{hughes1978classical}
+\pd{\bm{S}}{t} &= \frac{1}{2} \tensor{\bm{C}}\LRp{\Grad\bm{v} + \Grad\bm{v}^T},
+\end{align*}
+where $\bm{f}$ is the body force per unit volume, $\rho$ is density, and $\tensor{\bm{C}}$ is the symmetric constitutive stiffness tensor relating stress and strain.
+We rewrite these equations as a symmetric hyperbolic system of PDEs \cite{hughes1978classical}
 \begin{align}
@@
 \end{align}
-where $\bm{C}$ is the symmetric matrix form of the constitutive tensor $\bar{\bm{C}}$ and $\sigma$ is a vector of length $N_d = \frac{d(d+1)}{2}$, the number of unique entries of the stress tensor $\bm{S}$.
+where $\bm{C}$ is the symmetric matrix form of the constitutive tensor $\tensor{\bm{C}}$ and $\sigma$ is a vector of length $N_d = \frac{d(d+1)}{2}$, the number of unique entries of the stress tensor $\bm{S}$ in $d$ dimensions.
 We note that the matrices $\bm{A}_i$ are constant, while $\rho$ and $\bm{C}$ can vary spatially.
-Furthermore, we will assume that $\rho$ and $\bm{C}$ are positive-definite and bounded such that
-\begin{align*}
-\rho_{\min} &\leq \rho(\bm{x}) \leq \rho_{\max} \\
-0 < c_{\min} \leq \bm{x}^T\bm{C}\bm{x} &\leq c_{\max} < \infty, \qquad \bm{x} \in \mathbb{R}^{N_d},
-\end{align*}
-which also implies that $\bm{C}^{-1}$ is positive-definite and bounded.
+Furthermore, we will assume that $\rho$ and $\bm{C}$ are positive-definite and bounded pointwise such that
+\begin{align*}
+0 < \rho_{\min} &\leq \rho(\bm{x}) \leq \rho_{\max} < \infty, \qquad \forall \bm{x} \in \mathbb{R}^d,\\
+0 < c_{\min} \leq \bm{u}^T\bm{C}(\bm{x})\bm{u} &\leq c_{\max} < \infty, \qquad \forall \bm{u} \in \mathbb{R}^{N_d}, \quad \forall \bm{x} \in \mathbb{R}^d,
+\end{align*}
+which also implies that $\bm{C}^{-1}$ is pointwise positive-definite and bounded.
 
@@
 \sigma_{xy} & \sigma_{yy}
-\end{array}}.
-\]
-The matrices $\bm{A}_i$ are given as
+\end{array}},
+\]
+while the matrices $\bm{A}_i$ are 
 \[
@@
 \]
-
 In three dimensions, the velocity is $\bm{v} = (\bm{v}_1,\bm{v}_2,\bm{v}_3)^T$, while $\bm{\sigma}= (\sigma_{xx},\sigma_{yy},\sigma_{zz},\sigma_{yz},\sigma_{xy},\sigma_{xz})^T$ denotes the unique entries of the stress tensor $\bm{S}$
@@
 \[
-\bm{C}= \LRp{
-\begin{array}{cccccc}
-2\mu+\lambda  &  \lambda  &     \lambda  & & & \\
-\lambda &   2\mu+\lambda & \lambda   & & & \\
-\lambda & \lambda &   2\mu+\lambda &    & &\\
- &   &    & {\mu} & &\\
-&  &   &    & {\mu} &\\
-& &  &   &    & {\mu} \\
+%\bm{C}= 
+%\LRp{
+%\begin{array}{cccccc}
+%2\mu+\lambda  &  \lambda  &     \lambda  & & & \\
+%\lambda &   2\mu+\lambda & \lambda   & & & \\
+%\lambda & \lambda &   2\mu+\lambda &    & &\\
+% &   &    & {\mu} & &\\
+%&  &   &    & {\mu} &\\
+%& &  &   &    & {\mu} \\
+%\end{array}}.  
+\LRp{\begin{array}{cccc}
+2\mu+\lambda  &  \lambda  &     \lambda   & \\
+\lambda &   2\mu+\lambda & \lambda   &  \\
+\lambda & \lambda &   2\mu+\lambda &    \\
+ &   &    & {\mu}\bm{I}^{3\times3} \\
 \end{array}}.  
@@
 \section{An energy stable discontinuous Galerkin formulation for elastic wave propagation}
-
+\label{sec:dgform}
 The advantage of the symmetric first order formulation of the elastic wave equations (\ref{eq:symelas}) is that it is straightforward to derive an energy stable discontinuous Galerkin formulation.
@@
 \]
-
 We denote the $L^2$ inner product and norm over $D^k$ by $\LRp{\cdot,\cdot}_{\Lk}$, such that
@@
 \]
+We define also the $L^2$ inner product and norm over the boundary $\partial D^k$ of an element
+\[
+\LRa{u,v}_{\Ldk} = \int_{\partial D^k}uv\diff{x} = \sum_{f\in \partial D^k}\int_{\widehat{f}}uvJ^f\diff{\widehat{x}}, \qquad \nor{u}_{\Ldk}^2 = \LRa{u,u},
+\]
+where $J^f$ is the Jacobian of the mapping from a reference face (triangle) $\widehat{f}$ to a physical face $f$ of an element.
 
@@
 
-We can now specify a DG formulation for the linear elastic wave equation (\ref{eq:symelas}).
-Symmetric hyperbolic systems readily admit a DG formulation based on penalty fluxes, as detailed in \cite{chan2016short}.
+We can now specify a unified DG formulation for the linear elastic wave equation (\ref{eq:symelas}) in either isotropic or anisotropic media.
+Symmetric hyperbolic systems readily admit a DG formulation based on penalty fluxes \cite{chan2016short}.
 For the linear elastic wave equation in symmetric first order form, this formulation is given as
@@
 &\sum_{f\in \Gamma_h \setminus \partial \Omega} \int_{f}{\bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v}} + {\frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma}}\\
-&\leq - \sum_{f \in\Gamma_h \setminus \partial \Omega} \int_{f} \frac{\tau_v}{2}\LRb{\bm{A}_n\jump{\bm{v}}}^2 + \frac{\tau_\sigma}{2}\LRb{\bm{A}_n^T\jump{\bm{\sigma}}}^2.  
+&= - \sum_{f \in\Gamma_h \setminus \partial \Omega} \int_{f} \frac{\tau_v}{2}\LRb{\bm{A}_n\jump{\bm{v}}}^2 + \frac{\tau_\sigma}{2}\LRb{\bm{A}_n^T\jump{\bm{\sigma}}}^2.  
 %\label{thm:stability}
@@
 \begin{align}
-&\sum_{D^k\in \Oh} \frac{1}{2}\pd{}{t}\LRp{\LRp{\rho \bm{v},\bm{v}}_{\Lk} + \LRp{\bm{C}^{-1} \bm{\sigma},\bm{\sigma}}_{\Lk}} \leq - \sum_{f \in\Gamma_h \setminus \partial \Omega} \int_{f} \frac{\tau_v}{2}\LRb{\bm{A}_n\jump{\bm{v}}}^2 + \frac{\tau_\sigma}{2}\LRb{\bm{A}_n^T\jump{\bm{\sigma}}}^2\nonumber\\
+&\sum_{D^k\in \Oh} \frac{1}{2}\pd{}{t}\LRp{\LRp{\rho \bm{v},\bm{v}}_{\Lk} + \LRp{\bm{C}^{-1} \bm{\sigma},\bm{\sigma}}_{\Lk}} = - \sum_{f \in\Gamma_h \setminus \partial \Omega} \int_{f} \frac{\tau_v}{2}\LRb{\bm{A}_n\jump{\bm{v}}}^2 + \frac{\tau_\sigma}{2}\LRb{\bm{A}_n^T\jump{\bm{\sigma}}}^2\nonumber\\
 & - \sum_{f\in \Gamma_v} \int_f \tau_v  \LRb{\bm{A}_n\bm{v}^-}^2  - \sum_{f\in \Gamma_\sigma} \int_f \tau_\sigma  \LRb{\bm{A}_n^T\bm{\sigma}^-}^2- \sum_{f\in \Gamma_{\rm abc}} \int_f \frac{\tau_v}{2} \LRb{\bm{A}_n\bm{v}^-}^2 + \frac{\tau_\sigma}{2} \LRb{\bm{A}_n^T\bm{\sigma}^-}^2 \leq 0.
@@
 Then, $\bm{M}^{-1}_\rho = \frac{1}{\rho}\bm{M}^{-1} = \frac{1}{J\rho}\widehat{\bm{M}}^{-1}$, and $\bm{M}^{-1}_{\bm{C}^{-1}} = \bm{C}\otimes \bm{M}^{-1} = \bm{C}\otimes \LRp{\frac{1}{J}\widehat{\bm{M}}^{-1}}$, and each matrix inverse can be applied using the inverse of the reference mass matrix $\widehat{\bm{M}}^{-1}$ and the values of $\rho, \bm{C}$, and $J$ on each element.
-Applying this observation to \ref{eq:dgmat} then yields the following system of ODEs
+Applying this observation to \ref{eq:dgmat} then yields the following local system of ODEs
 \begin{align*}
@@
 Mass-lumping can also be done using interior quadratures with $N_p$ points \cite{williams2014symmetric}, though these are defined only up to $N=6$ in three dimensions, and proving high order accuracy.
-We address these issues by approximating the weighted mass matrix and corresponding weighted $L^2$ inner product with a weight-adjusted approximation which is low storage and simple to invert, energy stable, and provably high order accurate for spatially varying weights $\rho, \bm{C}$ with sufficiently regularity.
+For isotropic media, one way to address sub-element variations in material parameters is to diagonalize the matrix $\bm{C}$ through a change of variables \cite{etienne2010hp}.
+This results in a local system of ODEs with only scalar weighted mass matrices \cite{mercerat2015nodal}, which can be treated using scalar weight-adjusted approximations.
+We take a different approach in addressing these issues by approximating the matrix-weighted mass matrix $\bm{M}_{\bm{C}^{-1}}$ and corresponding matrix-weighted $L^2$ inner product with a weight-adjusted approximation which is low storage, simple to invert, energy stable, and provably high order accurate for spatially varying weights $\rho, \bm{C}$ with sufficiently regularity.
 
@@
 \section{Weight-adjusted inner products for matrix-valued weights}
-
+\label{sec:mwadg}
 Weight-adjusted inner products are high order accurate approximations of weighted $L^2$ inner products.
@@
 \]
-based on the observation that $T^{-1}_{1/w}u \approx uw$.  This approximation is made precise by the following estimates for errors in approximations of the product $uw$ and weighted moments:
+based on the observation that $T^{-1}_{1/w}u \approx uw$.
+This approximation is made precise by the following estimates for approximations of the product $uw$ and weighted moments on affinely mapped elements:
 \begin{theorem}[Theorem 5 in \cite{chan2016weight1}]
@@
 \begin{align*}
-&\LRb{\LRp{{w}u,v}_{L^2\LRp{D^k}} - \waip{u,v}{w}} \\
+&\LRb{\LRp{{w}u,v}_{L^2\LRp{D^k}} - \LRp{T^{-1}_{1/w}u,v}} \\
 &\leq Ch^{2N+2 - M} \nor{{w}}_{L^{\infty}\LRp{D^k}}\nor{\frac{1}{w}}_{L^{\infty}\LRp{D^k}}^2 \nor{w}^2_{W^{N+1,\infty}\LRp{D^k}}  \nor{u}_{W^{N+1,2}\LRp{D^k}} \nor{v}_{L^{\infty}\LRp{D^k}}.
@@
 \end{theorem}
+Both theorems rely on a scalar weighted interpolation estimate derived in \cite{warburton2013low,chan2016weight1} for a general non-affine element $D^k$.  
+\begin{theorem}[Theorem 3.1 in \cite{chan2016weight1}]
+Let $D^k$ be a quasi-regular element with representative size $h = {\rm diam}\LRp{D^k}$. For $N \geq 0$, $w\in W^{N+1,\infty}$, and $u\in W^{N+1,2}$, 
+\[
+\nor{u - \frac{1}{w}\Pi_N(wu)}_{\Lk} \leq C_N h^{N+1} \nor{{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{w}}_{L^\infty}\nor{w}_{W^{N+1,\infty}}\nor{u}_{W^{N+1,2}}.
+\]
+\label{thm:wproj}
+\end{theorem}
 
 \subsection{Extension to matrix weights}
-\label{sec:mwadg}
 
@@
 \begin{align*}
-\nor{\bm{W}(\bm{x})}_{k,p}^p &= \sum_{\LRb{\bm{\alpha}} \leq k} \sup_{\bm{x}} \nor{D^{\bm{\alpha}}\bm{W}(\bm{x})}^p_p
+\nor{\bm{W}(\bm{x})}_{k,p,\infty}^p &= \sum_{\LRb{\bm{\alpha}} \leq k} \sup_{\bm{x}} \nor{D^{\bm{\alpha}}\bm{W}(\bm{x})}^p_p
 \end{align*}
@@
 &\leq C_N\int \LRp{\sum_{\LRb{\alpha}\leq k} \nor{\LRp{D^{\alpha}\bm{W}}}}^p \LRp{\sum_{\LRb{\alpha}\leq k} \nor{D^{\alpha}\bm{v}}_{p} }^p\\
-&\leq C_N \nor{\bm{W}}_{k,p}^p \nor{\bm{v}}_{k,p}^p,
+&\leq C_N \nor{\bm{W}}_{k,p,\infty}^p \nor{\bm{v}}_{k,p}^p,
 \end{align*}
 where we have used Leibniz's rule, Cauchy-Schwarz, and the arithmetic-geometric mean inequality.
-
-To show that weight-adjusted DG methods are high order accurate, we make use of a scalar weighted interpolation estimate derived in \cite{warburton2013low,chan2016weight1}.  
-\begin{theorem}[Theorem 3.1 in \cite{chan2016weight1}]
-Let $D^k$ be a quasi-regular element with representative size $h = {\rm diam}\LRp{D^k}$. For $N \geq 0$, $w\in W^{N+1,\infty}$, and $u\in W^{N+1,2}$, 
-\[
-\nor{u - \frac{1}{w}\Pi_N(wu)}_{\Lk} \leq C_N h^{N+1} \nor{{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{w}}_{L^\infty}\nor{w}_{W^{N+1,\infty}}\nor{u}_{W^{N+1,2}}.
-\]
-\label{thm:wproj}
-\end{theorem}
 
@@
 \[
-\nor{\bm{v} - \bm{W}^{-1}\Pi_N\LRp{\bm{W}\bm{v}}}_{L^2} \leq Ch^{N+1} \nor{{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk}\tilde{w}_{\max}\nor{\bm{W}}_{N+1} \nor{\bm{v}}_{W^{N+1,2}}
-\]
-\label{thm:Wproj}
+\nor{\bm{v} - \bm{W}^{-1}\Pi_N\LRp{\bm{W}\bm{v}}}_{L^2} \leq Ch^{N+1} \nor{{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk}\tilde{w}_{\max}\nor{\bm{W}}_{N+1,2,\infty} \nor{\bm{v}}_{W^{N+1,2}}
+\]
+\label{thm:wwproj}
 \end{theorem}
@@
 &\leq C_2 h^{N+1}\nor{{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk} \sup_{\bm{x}}\nor{\bm{W}^{-1}}_{2} \nor{\bm{W}\bm{v}}_{W^{N+1,2}\LRp{D^k}}\\
-&\leq C_3 h^{N+1}\nor{{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk} \tilde{w}_{\max} \nor{\bm{W}}_{W^{N+1,2}}\nor{\bm{v}}_{W^{N+1,2}\LRp{D^k}}.
+&\leq C_3 h^{N+1}\nor{{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk} \tilde{w}_{\max} \nor{\bm{W}}_{{N+1,2,\infty}}\nor{\bm{v}}_{W^{N+1,2}\LRp{D^k}}.
 \end{align*}
@@
 
-\subsubsection{Properties of weight-adjusted vector-valued inner products}
+\subsubsection{Weight-adjusted approximations with matrix weights}
 \label{sec:mwadgprop}
@@
 \]
-This definition is a straightforward generalization of $T^{-1}_w}$ to matrix-valued weights $\bm{W}$.
+This definition is a straightforward generalization of $T^{-1}_{w}$ to matrix-valued weights $\bm{W}$.
 Conveniently, all properties of $T_{w},T^{-1}_{w}$ for scalar $w(\bm{x})$ carry over to matrix weights $\bm{W}(\bm{x})$ as well \cite{chan2016weight1}.
@@
 
-Using Theorem~\ref{thm:Wproj}, we may also show that the matrix-valued weight-adjusted inner product is also high order accurate for sufficiently regular $\bm{W}$.
+Using Theorem~\ref{thm:wwproj}, we may also show that the matrix-valued weight-adjusted inner product is also high order accurate for sufficiently regular $\bm{W}$.
 \begin{theorem}
@@
 %C_{\bm{W}} = C_N \cond{\bm{W}}{2,\infty} \nor{\bm{W}}_{W^{N+1,2}\LRp{D^k}}.  
-C_{\bm{W}} = C_N \nor{\sqrt{J}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk} \tilde{w}_{\max}{w}_{\max} \nor{\bm{W}}_{W^{N+1,2}\LRp{D^k}}.  
-\]
+C_{\bm{W}} = C_N \nor{\sqrt{J}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk} \tilde{w}_{\max}{w}_{\max} \nor{\bm{W}}_{{N+1,2,\infty}}.  
+\]
+\label{thm:west}
 \end{theorem}
@@
 \]
-The former term is bounded by interpolation estimates and by arguments used in the proof of Theorem~\ref{thm:Wproj}.  The latter term is bounded 
+The former term is bounded by interpolation estimates and by arguments used in the proof of Theorem~\ref{thm:wwproj}.  The latter term is bounded 
 \begin{align*}
@@
 where we have used $\nor{T_{\bm{W}^{-1}}^{-1}} \leq {w}_{\max}$ (Lemma~\ref{lemma:props}) and $\nor{\Pi_N}_{\Lk} = 1$.  
-An application of Theorem~\ref{thm:Wproj} to bound $\nor{ \bm{W}^{-1}\Pi_N(\bm{W}\bm{v}) - \bm{v} }_{\Lk }$ completes the proof.
+An application of Theorem~\ref{thm:wwproj} to bound $\nor{ \bm{W}^{-1}\Pi_N(\bm{W}\bm{v}) - \bm{v} }_{\Lk }$ completes the proof.
 \end{proof}
 
-\note{This is a good generalization because defining $T^{-1}_w \bm{v} = T^{-1}_{w\bm{I}}\bm{v}$ recovers estimates for scalar weight-adjusted inner products.}
+We note that these estimates are tight, in the sense that we recover estimates for scalar weight-adjusted inner products when the matrix weight is taken to be $\bm{W} = w\bm{I}$.  
+%\note{This is a good generalization because defining $T^{-1}_w \bm{v} = T^{-1}_{w\bm{I}}\bm{v}$ recovers estimates for scalar weight-adjusted inner products.}
 
@@
 \] 
-In this case, we also have $\bm{M}_{\bm{W}^{-1}} = \bm{W}^{-1}\otimes \bm{M}$.
-Substituting this explicit inverse into the weight adjusted mass matrix inverse $\LRp{\bm{I} \otimes \bm{M}^{-1}} \bm{M}_{\bm{W}^{-1}} \LRp{\bm{I} \otimes \bm{M}^{-1}}$ in (\ref{eq:wadgmc}) recovers the exact inversion of $\bm{M}_{\bm{W}}$.
+In this case, we also have $\bm{M}_{\bm{W}^{-1}} = \bm{W}^{-1}\otimes \bm{M}$, and substituting this explicit inverse into the weight adjusted mass matrix inverse $\LRp{\bm{I} \otimes \bm{M}^{-1}} \bm{M}_{\bm{W}^{-1}} \LRp{\bm{I} \otimes \bm{M}^{-1}}$ in (\ref{eq:wadgmc}) recovers the exact inversion of $\bm{M}_{\bm{W}}$.
 
@@
 %\end{align*}
-Inverting these weight-adjusted mass matrices yields the following system of ODEs for $\bm{V},\bm{\Sigma}$
+Inverting these weight-adjusted mass matrices yields the following local system of ODEs for $\bm{V},\bm{\Sigma}$
 \begin{align*}
@@
 \pd{\bm{\Sigma}}{t} &= \LRp{\bm{I}\otimes\bm{M}^{-1}} \bm{M}_{\bm{C}}  \LRp{\sum_{i=1}^d \LRp{\bm{A}_i\otimes \bm{D}_{i}}\bm{V}  +  \sum_{f\in \partial D^k}\LRp{\bm{I}\otimes\bm{L}_f}\bm{F}_{\sigma}}.
-\end{align*}
-In practice, the matrices $\LRp{\bm{I}\otimes\bm{M}^{-1}} \bm{M}_{\LRp{\rho^{-1}\bm{I}}}$ and $\LRp{\bm{I}\otimes\bm{M}^{-1}} \bm{M}_{\bm{C}}$ are applied to a vector using quadrature.
-After fusing operations together, this procedure can be boiled down to multiplication by two rectangular matrices.
-
-%The matrix form of DG 
-%\begin{align*}
-%\bm{M}_{\rho \bm{I}} \pd{\bm{V}}{t} &= \sum_{i=1}^d  \LRp{\bm{A}_i^T\otimes \bm{S}_{i}} \bm{\Sigma} + \sum_{f\in \partial D^k}\LRp{\bm{I}\otimes \bm{M}_f}\bm{F}_{v}\nonumber\\
-%\bm{M}_{\bm{C}^{-1}} \pd{\bm{\Sigma}}{t} &= \sum_{i=1}^d \LRp{\bm{A}_i\otimes \bm{S}_{i}}\bm{V}  +  \sum_{f\in \partial D^k}\LRp{\bm{I}\otimes\bm{M}_f}\bm{F}_{\sigma}. 
-%\end{align*}
-%where the matrices $\bm{A}_{\bm{v}_i}$ and $\bm{A}_{\sigma}$ are induced by the discretization of the right hand side.
-
-We note that, for isotropic media, another way to address sub-element variations in material parameters is to diagonalize the matrix $\bm{C}$ through a change of variables \cite{etienne2010hp}.
-This results in a local system of ODEs with only scalar weighted mass matrices \cite{mercerat2015nodal}, which can be treated using scalar weight-adjusted approximations.
-However, this approach is more difficult to apply when $\bm{C}$ is anisotropic.
+\label{eq:wadgmat}
+\end{align*}
+In practice, the matrices $\LRp{\bm{I}\otimes\bm{M}^{-1}} \bm{M}_{\LRp{\rho^{-1}\bm{I}}}$ and $\LRp{\bm{I}\otimes\bm{M}^{-1}} \bm{M}_{\bm{C}}$ are applied in a matrix-free fashion using reference element matrices and values of $\rho,\bm{C}$ at quadrature points.
+After fusing operations together, this procedure can be boiled down to multiplication by two rectangular matrices \cite{chan2016weight2}.
 
 One drawback of the analysis presented in this work is that accuracy is not guaranteed in the isotropic incompressible limit $\mu / \lambda \rightarrow 0$.
-In this case, the stiffness matrix $\bm{C}$ becomes singular, and the constant $\tilde{w}_{\max}$ in the upper bound on $\nor{\bm{C}}$ blows up.
-However, numerical experiments in Section~\ref{sec:incomp} suggest that the accuracy of WADG  for linear elastic wave propagation does not degrade significantly in the incompressible limit.
-
-\subsection{An energy stable formulation for curvilinear meshes}
+In this case, the stiffness matrix $\bm{C}$ becomes singular, and the constant $\tilde{w}_{\max}$ in the upper bound on $\nor{\bm{C}^{-1}}$ blows up.
+However, numerical experiments in Section~\ref{sec:incomp} suggest that, while taking $\mu/\lambda \approx 0$ (or $\bm{C}$ nearly singular) results in larger relative errors for $\bm{\sigma}$, the accuracy of the WADG solution for $\bm{v}$ does not degrade significantly for near-incompressible materials.
+
+\subsection{Energy stability on curvilinear meshes}
 
@@
 
-%Nodal collocation does not work for arbitrary wavespeeds.  Example: wave equation, $c^2 = 1 + .9\sin(8\pi x)\sin(8\pi y)$.
+Using estimates from Section~\ref{sec:mwadgprop}, we can extend the semi-discrete convergence analysis in \cite{houston2002discontinuous,warburton2013low,chan2016weight1} to linear elastic wave propagation on meshes of affine elements.
+Techniques in \cite{warburton2013low} can be used to extend this analysis to curvilinear elements.
+
+In terms of group variables $\bm{U} = \LRp{\bm{v},\bm{\sigma}}$ and $\bm{V} = \LRp{\bm{w},\bm{q}} \in \LRp{V_h\LRp{\Oh}}^d\times\LRp{V_h\LRp{\Oh}}^{N_d}$, the variational formulation can be written as
+\begin{align*}
+&\LRp{T^{-1}_{\bm{A}_0^{-1}}\pd{\bm{U}}{t},\bm{V}}_{\Omega} + a(\bm{U},\bm{V}) + b(\bm{U},\bm{V}) = (\bm{f},\bm{V})\\
+a(\bm{U},\bm{V}) &= \sum_{D^k\in \Oh} \LRp{-\LRp{ \sum_{i=1}^d \bm{\sigma},\bm{A}_i \pd{\bm{w}}{\bm{x}_i}}_{\Lk}  + \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk}} \\
+b(\bm{U},\bm{V}) &= \sum_{D^k\in \Oh}\LRp{ \LRa{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk} },
+%a(\bm{U},\bm{V}) &= \sum_{D^k\in \Oh} \LRp{\LRp{ \sum_{i=1}^d \bm{A}_i^T\pd{\bm{\sigma}}{\bm{x}_i},\bm{w}}_{\Lk}  + \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk}} \\
+%b(\bm{U},\bm{V}) &= \sum_{D^k\in \Oh}\LRp{ \LRa{\bm{A}_n^T\jump{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk} },
+\end{align*}
+where $\bm{A}_0 = {\rm diag}\LRp{\rho\bm{I},\bm{C}^{-1}}$.  Energy stability implies that
+\begin{align*}
+&b(\bm{U},\bm{U}) = \frac{1}{2}\sum_{D^k\in\Oh} \LRp{\tau_v\nor{\jump{\bm{A}_n\bm{v}}}_{\Ldk}^2 + \tau_\sigma\nor{\jump{\bm{A}^T_n\bm{\sigma}}}_{\Ldk}^2}\\
+&\LRp{T^{-1}_{\bm{A}_0^{-1}}\pd{\bm{U}}{t},\bm{U}}_{\Omega} + b(\bm{U},\bm{U}) =  (\bm{f},\bm{U}).
+\end{align*}
+Let $\bm{U},\bm{U}_h$ denote the exact and discrete WADG solutions, respectively.  Since the DG formulation (\ref{eq:dgform}) is consistent, these solutions satisfy
+\begin{align}
+\LRp{\bm{A}_0\pd{\bm{U}}{t},\bm{V}} + a(\bm{U},\bm{V}) + b(\bm{U},\bm{V}) &= (\bm{f},\bm{V}) \nonumber\\
+\LRp{T_{\bm{A}_0^{-1}}^{-1}\pd{\bm{U}_h}{t},\bm{V}} + a(\bm{U}_h,\bm{V}) + b(\bm{U}_h,\bm{V}) &= (\bm{f},\bm{V})
+\label{eq:dgvswadg}
+\end{align}
+for all $\bm{V} \in \LRp{V_h\LRp{\Oh}}^d\times\LRp{V_h\LRp{\Oh}}^{N_d}$.  We decompose the error $\bm{U}-\bm{U}_h$ into a projection error $\bm{\epsilon}$ and discretization error $\bm{\eta}$.
+\[
+\bm{U}-\bm{U}_h  = \LRp{\Pi_N\bm{U}-\bm{U}_h}-\LRp{\Pi_N\bm{U}-\bm{U}} = \bm{\eta} - \bm{\epsilon}.
+\]
+We also introduce a consistency error $\bm{\delta} = \bm{A}_0\bm{U} - T_{\bm{A}_0^{-1}}^{-1}\bm{U}$ resulting from the approximation of $\bm{A}_0\bm{U}$ by a weight-adjusted inner product
+\[
+\bm{A}_0\pd{\bm{U}}{t} - T_{\bm{A}_0^{-1}}^{-1}\pd{\bm{U}_h}{t} = \pd{}{t}\LRp{\bm{A}_0\bm{U} - T_{\bm{A}_0^{-1}}^{-1}\bm{U}} + \pd{}{t}T_{\bm{A}_0^{-1}}^{-1}\LRp{\Pi_N\bm{U}-\bm{U}_h} = \pd{\bm{\delta}}{t} + \pd{}{t}\LRp{T_{\bm{A}_0^{-1}}^{-1}\bm{\eta}}
+\]  
+where we have used that $\bm{T}_{\bm{A}_0^{-1}}^{-1} = \bm{T}_{\bm{A}_0^{-1}}^{-1}\Pi_N$.  Subtracting the DG and WADG formulations in (\ref{eq:dgvswadg}) and setting $\bm{V} = \bm{\eta}$ then yields 
+\begin{align}
+\frac{1}{2}\pd{}{t}\LRp{T^{-1}_{\bm{A}_0^{-1}}\bm{\eta},\bm{\eta}}_{\Omega} + b(\bm{\eta},\bm{\eta}) &= \LRp{\pd{\bm{\delta}}{t},\bm{\eta}} + a(\bm{\epsilon},\bm{\eta}) + b(\bm{\epsilon},\bm{\eta}),
+\label{eq:erreqn}
+\end{align}
+where we have used that $a(\bm{\eta},\bm{\eta}) = 0$ by skew symmetry.
+
+We bound $a(\bm{\epsilon},\bm{\eta}) + b(\bm{\epsilon},\bm{\eta})$ on the right hand side by integrating by parts the stress equation and using the component-wise $L^2$ orthogonality of $\bm{\epsilon}$ to derivatives of $\bm{\eta}$.
+This reduces the term $a(\bm{\epsilon},\bm{\eta}) + b(\bm{v},\bm{\eta})$ to surface contributions over each element, which can be combined with contributions from neighboring elements
+\begin{align*}
+ \sum_{D^k\in \Oh}&\LRp{ \LRa{\bm{A}_n^T\avg{\bm{\epsilon}_\sigma} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{\epsilon}_v},\bm{\eta}_v}_{\Ldk} + \LRa{\bm{A}_n\avg{\bm{\epsilon}_v} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\epsilon}_\sigma},\bm{\eta}_\sigma}_{\Ldk}}\\
+&= \sum_{D^k\in \Oh}\LRp{ 
+\LRa{\avg{\bm{\epsilon}_\sigma} - \frac{\tau_v}{2}\bm{A}_n\jump{\bm{\epsilon}_v},\bm{A}_n\jump{\bm{\eta}_v}}_{\Ldk} + \LRa{\avg{\bm{\epsilon}_v} - \frac{\tau_\sigma}{2}\bm{A}_n^T\jump{\bm{\epsilon}_\sigma},\bm{A}_n^T\jump{\bm{\eta}_{\sigma}}}_{\Ldk}}\\ %- \LRa{\bm{A}_n^T\jump{\bm{\epsilon}_\sigma},\frac{\tau_{\sigma}}{2}\bm{A}_n^T\jump{\bm{\eta}}_\sigma}_{\Ldk} }\\
+&\leq \sum_{D^k\in \Oh}\nor{\avg{\bm{\epsilon}_\sigma} - \frac{\tau_v}{2}\bm{A}_n\jump{\bm{\epsilon}_v}}_{\Ldk} \nor{\bm{A}_n\jump{\bm{\eta}_v}}_{\Ldk} + \nor{\avg{\bm{\epsilon}_v} - \frac{\tau_\sigma}{2}\bm{A}_n^T\jump{\bm{\epsilon}_\sigma}}_{\Ldk}\nor{\bm{A}_n^T\jump{\bm{\eta}_{\sigma}}}_{\Ldk}\\
+%&\leq \LRp{\nor{\avg{\bm{\epsilon}_\sigma} - \frac{\tau_v}{2}\bm{A}_n\jump{\bm{\epsilon}_v}}_{\Ldk}  + \nor{\avg{\bm{\epsilon}_v} - \frac{\tau_\sigma}{2}\bm{A}_n^T\jump{\bm{\epsilon}_\sigma}}_{\LDk}} +  \LRp{\nor{\bm{A}_n\jump{\bm{\eta}_v}}_{\Ldk}^2 + \nor{\bm{A}_n^T\jump{\bm{\eta}_{\sigma}}}_{\Ldk}^2}\\
+&\leq C_\tau \sum_{D^k\in \Oh}\nor{\bm{\epsilon}}_{\Ldk}\frac{1}{2}\LRp{\tau_v\nor{\bm{A}_n\jump{\bm{\eta}_v}}_{\Ldk}^2 + \tau_\sigma\nor{\bm{A}_n^T\jump{\bm{\eta}_{\sigma}}}_{\Ldk}^2}^{1/2}
+\end{align*}
+where $C_\tau$ is proportional to $\max\LRp{\tau_v, \tau_\sigma}$.  Using Young's inequality with $\alpha \geq 1/\max\LRp{\tau_v,\tau_\sigma,1}$ sufficiently large yields the following bound 
+\[
+\LRb{a(\bm{\epsilon},\bm{\eta}) + b(\bm{\epsilon},\bm{\eta})} \leq \frac{C}{4\alpha}{b(\bm{\eta},\bm{\eta})} + \frac{C\alpha}{4} \sum_{D^k\in \Oh}\nor{\bm{\epsilon}}_{\Ldk}^2.
+\]
+%where $\bm{E}_v,\bm{E}_\sigma,\bm{\eta}_v,\bm{\eta}_\sigma$ denote respective components of $\bm{E},\bm{\eta}$.  
+This can be used to bound (\ref{eq:erreqn}).  Using Young's inequality on the remaining terms and absorbing terms involving $\bm{\eta}$ into the left hand side yields
+\[
+\frac{1}{2}\pd{}{t}\LRp{T^{-1}_{\bm{A}_0^{-1}}\bm{\eta},\bm{\eta}}_{\Omega} + \LRp{1-\frac{C}{4\alpha}} b(\bm{\eta},\bm{\eta}) \leq \nor{\pd{\bm{\delta}}{t}}_{\Lk} + \sum_{D^k \in \Oh}\frac{ C\alpha}{4} \nor{\bm{\epsilon}}_{\Ldk}^2.
+\]
+Integrating from $[0,T]$, and using the arithmetic-geometric mean inequality yields
+\[
+\frac{1}{2}\LRp{T^{-1}_{\bm{A}_0^{-1}}\bm{\eta},\bm{\eta}}_{\Omega} + \LRp{1-\frac{C}{4\alpha}}\int_0^T  b(\bm{\eta},\bm{\eta}) \leq \int_0^T\nor{\pd{\bm{\delta}}{t}}_{\Lk} + \int_0^T\sum_{D^k \in \Oh}C\nor{\bm{\epsilon}}_{\Ldk}.
+\]
+Using Theorem~\ref{thm:west} and a standard $hp$ trace inequality $\nor{u}_{\Ldk} \leq Ch^{-1/2} \nor{u}_{\L}$ \cite{chan2015gpu} then yields
+\[
+\nor{\bm{\eta}} \leq C_{\bm{W}} T \sup_{t\in [0,T]}\LRp{\nor{\pd{\bm{\delta}}{t}} + \nor{\bm{\epsilon}}h^{-1/2}}.
+\]
+Bounding each term on the right using interpolation estimates for sufficiently regular $\bm{U}$ yields the final result
+\[
+\nor{\bm{U}-\bm{U}_h} \leq (C_1 + C_2 T) h^{N + 1/2}\LRp{\nor{\bm{U}}_{W^{N+1,2}} + C_3 h^{1/2}\nor{\pd{\bm{U}}{t}}_{W^{N+1,2}}}.
+\]
+where $C_2$ depends on $N$ and $C_3$ depends on $\nor{\bm{W}}_{{N+1,2,\infty}}$.
+\note{Todo - finish}
 
 \section{Numerical experiments}
-
+\label{sec:numerical}
 The following sections present several numerical experiments validating the stability and accuracy of the proposed method in two and three dimensions.
-
-\subsection{Verification of stability}
-
-blah
-
-\note{Eigenvalue plots with $\tau = 0$ and $\tau = O(1)$ for both square and curvilinear grids with random coefficients close to zero and 1.  Change in solution norm as time goes on.  }
+The energy stability of the method is also confirmed for examples with sub-cell variations in heterogeneous media and curvilinear meshes.
+The convergence of the new DG formulation in piecewise constant isotropic media is confirmed using analytic solutions, while the convergence of the method for high order approximations of heterogeneous media is confirmed using a fine grid reference solution.
+Finally, the method is applied to problems with anisotropy and stiffness matrices $\bm{C}$ with sub-element variations.
+
+In all experiments, we follow \cite{chan2016weight1} and compute application of weight-adjusted mass matrices using a quadrature exact for polynomials of degree $(2N+1)$.
+Time integration is performed using the low-storage 4th order five-stage Runge-Kutta scheme of Carpenter and Kennedy \cite{carpenter1994fourth}, and the timestep is chosen based on the global restriction
+\[
+dt = \min_{f} \frac{C_{\rm CFL}}{  \nor{\bm{C}(\bm{x})}_{2,\infty,\Omega} C_N \nor{J^f}_{L^{\infty}(f)}\nor{J^{-1}}_{L^{\infty}(f)}}
+\]
+where $C_N = O(N^2)$ is the order-dependent constant in the surface polynomial trace inequality \cite{chan2015gpu}, $C_{\rm CFL}$ is a tunable global CFL constant, and $\nor{\bm{C}}_{2,\infty,\Omega}$ denotes the spatial supremum of the $2$-norm of the stiffness matrix $\bm{C}(\bm{x})$.
+\[
+\nor{\bm{C}}_{2,\infty,\Omega} = \sup_{\bm{x}\in\Omega}\nor{\bm{C}(\bm{x})}_{2}.  
+\]
+
+\subsection{Spectra and choice of penalty parameter}
+
+We first verify the energy stability of the WADG method for arbitrary heterogeneous media.
+We use a stiffness matrix constructed using random numbers and similarity transforms, such that at every quadrature point, $\bm{C}(\bm{x}) = \bm{U}\bm{D}\bm{U}^{T}$, where $\bm{D}$ is a diagonal matrix with positive entries $d_{\min} \leq \bm{D}_{ii} \leq d_{\max}$ and $\bm{U}$ is a random unitary matrix.
+Let $\bm{A}_h$ denote the matrix induced by the global semi-discrete DG formulation, such that the time evolution of the solution $\bm{v},\bm{\sigma}$ is governed by
+\[
+\pd{\bm{Q}}{t} = \bm{A}_h\bm{Q},
+\]
+with $\bm{Q}$ denotes a vector of degrees of freedom for $(\bm{v},\bm{\sigma})$.
+Figure~\ref{fig:spectra} shows computed eigenvalues of $\bm{A}_h$ for $[d_{\min},d_{\max}] = [1/10,1]$ and $[10^{-4},1]$, $\tau_v=\tau_\sigma = 0$ and $\tau_v=\tau_\sigma = 1$ under discretization parameters $N=4$ and $h = 1/4$.
+
+\begin{figure}
+\centering
+\subfloat[$\tau_v = \tau_\sigma = 0$]{\includegraphics[width=.475\textwidth]{../figs/spectra_tau0.png}}
+%\hspace{1em}
+\subfloat[$\tau_v = \tau_\sigma = 1$]{\includegraphics[width=.475\textwidth]{../figs/spectra_tau1.png}}
+\caption{Spectra for $N=4$ and $h = 1/4$, using randomly chosen $\bm{C}(\bm{x})$ with eigenvalues between $[d_{\min}, 1]$ at each quadrature point.
+For both $\tau_v = \tau_\sigma = 0$ and $\tau_v = \tau_\sigma = 1$, the largest real part of the spectra is $O(10^{-14})$.}
+\label{fig:spectra}
+\end{figure}
+
+We note that, for $\tau_v = \tau_\sigma = 1$, $\nor{\bm{A}_h}$ is approximately twice as large compared to when $\tau_v = \tau_\sigma = 0$.
+This can result in a more restrictive stability condition on the timestep.
+We observe that the growth in $\nor{\bm{A}_h}$ for $\tau_v,\tau_\sigma = 1$ is due to the presence of eigenvalues with large negative real parts, which implies that larger timesteps may be used under time integrators with extended stability regions along the real axis \cite{abdulle2002fourth}.
+
+Alternatively, we can avoid growth in $\nor{\bm{A}_h}$ by scaling $\tau$.
+To avoid artificially increasing the stiffness of (\ref{eq:wadgmat}), it was argued in \cite{chan2016short} that the magnitude of the penalty parameter should depend on the largest singular value of $\bm{A}_n$, which is $O(1)$ in two and three dimensions.
+However, this does not factor in the effect of $\rho$ and $\bm{C}^{-1}$ appearing in the left hand side.
+These coefficients can be taken into account by scaling only the stress penalization parameter with the average of local max norms over a face $f$
+\[
+\tau_{\sigma,f} \approx \frac{2}{\nor{\bm{C}^+(\bm{x})}_{2,\infty,f}+\nor{\bm{C}^-(\bm{x})}_{2,\infty,f}},
+\]
+where $\bm{C}^+,\bm{C}^-$ denote values of the stiffness matrix on elements which share the face $f$.
+It may also be advantageous to rescale $\tau_\sigma$ for each individual component of stress based on the rows of $\bm{C}$.
+Under this rescaling of the stress penalty parameter, we observe numerically that the norm of $\bm{A}_h$ remains roughly the same magnitude as for $\tau_p = \tau_\sigma = 0$.
+%Since we do not observe significant changes in computed $L^2$, we adopt this rescaling of $\tau_\sigma$ in all reported results.
+%We note that this rescaling of the penalty parameters also avoids other issues associated with over-penalization, such as the reintroduction of spurious modes which would otherwise be damped under less aggressive penalties \cite{chan2016short}.
+
+In the following experiments, we use $\tau_v = \tau_\sigma = 1$.  However, we do not observe significant changes in computed $L^2$ errors under the aforementioned rescaling of $\tau_\sigma$.
 
@@
 Next, we study the accuracy and convergence of weight-adjusted DG method for several analytical solutions in linear elasticity.
-In all cases, the solution in expressed terms of the displacement vector $\bm{u}(\bm{x},t)$.
+In all cases, the solution in expressed terms of the displacement vector $\bm{u}(\bm{x},t) = (u_1,\ldots,u_d)$.
 Initial conditions for velocity and stress are computed through
@@
 
-We first examine convergence  on a unit square domain with  $\lambda = \mu = \rho = 1$.   The displacement is given by
+We first examine convergence  on a unit square domain with  $\lambda = \mu = \rho = 1$.   The components of the displacement vector is given by
 \begin{align*}
@@
        ylabel={$L^2$ errors}, 
-    xmin=.04, xmax=1,
+    xmin=.025, xmax=1,
     ymin=1e-12, ymax=1,
@@
 \addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.1732)(0.25,0.0439)(0.125,0.0110)(0.0625,0.0027)};
-%\node at (axis cs:.04,.0027) {$N = 1$};
-
-\logLogSlopeTriangleFlip{0.32}{0.125}{0.815}{2}{blue}
+coordinates{(0.5,0.162198)(0.25,0.0389974)(0.125,0.0095562)(0.0625,0.00227802)};
+\logLogSlopeTriangleFlip{0.4}{0.125}{0.81}{2.05}{blue}
 
 \addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.0150)(0.25,0.0020)(0.125,2.4911e-04)(0.0625,3.1277e-05)};
-%\node at (axis cs:.04,3e-5) {$N = 2$};
-
-\logLogSlopeTriangleFlip{0.32}{0.125}{0.66}{3}{red}
+coordinates{(0.5,0.0128844)(0.25,0.00156487)(0.125,0.000197047)(0.0625,2.42153e-05)};
+\logLogSlopeTriangleFlip{0.4}{0.125}{0.65}{3.01}{red}
 
 \addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5, 0.0013)(0.25,8.3910e-05)(0.125,5.3840e-06)(0.0625,3.2015e-07)};
-%\node at (axis cs:.04,3e-7) {$N = 3$};
-
-\logLogSlopeTriangleFlip{0.32}{0.125}{0.505}{4}{black}
+coordinates{(0.5,0.00110186)(0.25,6.9321e-05)(0.125,4.31386e-06)(0.0625,2.49229e-07)};
+\logLogSlopeTriangleFlip{0.4}{0.125}{0.49}{4.06}{black}
 
 \addplot[color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,1.0687e-04)(0.25,3.3041e-06)(0.125,9.7715e-08)(0.0625, 2.8750e-09)};
-%\node at (axis cs:.03,1.5e-13) {$a = 10^{-4}$};
-\logLogSlopeTriangleFlip{0.32}{0.125}{0.34}{5}{magenta}
+coordinates{(0.5,8.42986e-05)(0.25,2.47716e-06)(0.125,7.21337e-08)(0.0625,2.17472e-09)};
+\logLogSlopeTriangleFlip{0.4}{0.125}{0.325}{5.08}{magenta}
 
 \addplot[color=violet,mark=pentagon*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,5.9754e-06)(0.25,9.3644e-08)(0.125,1.3980e-09)(0.0625,2.0686e-11)};
-
-\logLogSlopeTriangleFlip{0.32}{0.125}{0.17}{6}{violet}
-
-%\node at (axis cs:.03,6.8e-10) {$u$ discontinuous};
+coordinates{(0.5,4.91233e-06)(0.25,7.57947e-08)(0.125,1.10789e-09)(0.0625,1.63538e-11)};
+\logLogSlopeTriangleFlip{0.4}{0.125}{0.15}{6.09}{violet}
+
 \legend{$N=1$,$N=2$,$N=3$,$N=4$,$N=5$}
@@
 \begin{align*}
-\bm{v}\LRp{x,y,t} &= e^{-\omega x \sqrt{1-\xi^2}}
+\bm{u}\LRp{x,y,t} &= e^{-\omega x \sqrt{1-\xi^2}}
 \LRp{
@@
 
-Lamb waves are supported by elastic waveguides with traction-free (free surface) boundary conditions at the top and bottom of the domain.  The displacement of these waves 
-\begin{align*}
-\bm{v}_1\LRp{x,y,t} &= \LRp{-kB_1 \cos(p y)- qB_2 \cos(q y)} \sin(k x - \omega t)\\
-\bm{v}_2\LRp{x,y,t} &= \LRp{-pB_1 \sin(py) + kB_2 \sin(qy)} \cos(kx - \omega t)
+Lamb waves are supported by elastic waveguides with traction-free (free surface) boundary conditions at the top and bottom of the domain.  The displacement of these waves is given by
+\begin{align*}
+{u}_1\LRp{x,y,t} &= \LRp{-kB_1 \cos(p y)- qB_2 \cos(q y)} \sin(k x - \omega t)\\
+{u}_2\LRp{x,y,t} &= \LRp{-pB_1 \sin(py) + kB_2 \sin(qy)} \cos(kx - \omega t)
 \end{align*}
@@
 Figure~\ref{fig:rayleigh} shows $L^2$ errors for both waves at time $T = 1$.
-Observed convergence rates fall between the optimal rate of $O(h^{N+1})$ and theoretical rate of $O(h^{N+1/2})$ \cite{johnson1986analysis}.
+Computed convergence rates fall between the optimal rate of $O(h^{N+1})$ and theoretical rate of $O(h^{N+1/2})$ \cite{johnson1986analysis}.
 
@@
     ylabel={$L^2$ errors}, 
-    xmin=.04, xmax=1,
+    xmin=.025, xmax=1,
     ymin=1e-8, ymax=1.5,
@@
 \addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.9569)(0.25,0.4509)(0.125,0.1234)(0.0625,0.0292)};%\node at (axis cs:.04,.0027) {$N = 1$};
-\logLogSlopeTriangleFlip{0.32}{0.125}{0.83}{2}{blue}
+coordinates{(0.5,1.04483)(0.25,0.342112)(0.125,0.102275)(0.0625,0.0277191)};    
+\logLogSlopeTriangleFlip{.39}{0.11}{0.815}{1.81}{blue}
 
 \addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.3483)(0.25,0.0521)(0.125,0.0072)(0.0625,9.3104e-04)};
-\logLogSlopeTriangleFlip{0.32}{0.125}{0.66}{3}{red}
+coordinates{(0.5,0.287061)(0.25,0.0563744)(0.125,0.00785962)(0.0625,0.000992348)};    
+\logLogSlopeTriangleFlip{.38}{0.1}{0.64}{2.91}{red}
 
 \addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.0759)(0.25,  0.0072)(0.125,5.3457e-04)(0.0625,3.6842e-05)};
-\logLogSlopeTriangleFlip{0.32}{0.125}{0.49}{4}{black}
+coordinates{(0.5,0.0736249)(0.25,0.00747192)(0.125,0.000529189)(0.0625,3.53252e-05)};   
+\logLogSlopeTriangleFlip{.38}{0.1}{0.475}{3.86}{black}
 
 \addplot[color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.0169)(0.25,  7.7478e-04)(0.125,2.6817e-05)(0.0625,8.6998e-07)};
-\logLogSlopeTriangleFlip{0.32}{0.125}{0.31}{5}{magenta}
+coordinates{(0.5,0.0199958)(0.25,0.000774494)(0.125,2.70714e-05)(0.0625,8.78659e-07)};   
+\logLogSlopeTriangleFlip{.38}{0.1}{0.285}{4.89}{magenta}
 
 \addplot[color=violet,mark=pentagon*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.0031)(0.25,6.7730e-05)(0.125,1.235333472480438e-06)(0.0625,2.023726469227763e-08)};
-
-\logLogSlopeTriangleFlip{0.32}{0.125}{0.12}{6}{violet}
+coordinates{(0.5,0.00304235)(0.25,6.78379e-05)(0.125,1.18747e-06)(0.0625,1.92431e-08)};    
+\logLogSlopeTriangleFlip{.38}{0.1}{0.09}{5.89}{violet}
 
@@
 %       ylabel={$L^2$ errors}, 
-    xmin=.04, xmax=1,
+    xmin=.025, xmax=1,
     ymin=2e-7, ymax=1.5,
@@
 \addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.9924)(0.25,0.6613)(0.125,0.2713)(0.0625,0.0737)};%\node at (axis cs:.04,.0027) {$N = 1$};
-\logLogSlopeTriangleFlip{0.32}{0.125}{0.85}{2}{blue}
+coordinates{(0.5,0.97515)(0.25,0.667267)(0.125,0.26441)(0.0625,0.0643182)};   
+\logLogSlopeTriangleFlip{.39}{0.11}{0.835}{2.04}{blue}
 
 \addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.6614)(0.25,0.2338)(0.125,0.0205)(0.0625,0.0021)};
-\logLogSlopeTriangleFlip{0.32}{0.125}{0.65}{3}{red}
+coordinates{(0.5,0.659983)(0.25,0.191328)(0.125,0.0172492)(0.0625,0.00168533)};   
+\logLogSlopeTriangleFlip{.38}{0.1}{0.62}{3.41}{red}
 
 \addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.4237)(0.25, 0.0280)(0.125, 0.0018)(0.0625,1.3754e-04)};
-\logLogSlopeTriangleFlip{0.32}{0.125}{0.475}{4}{black}
-
+coordinates{(0.5,0.366154)(0.25,0.024884)(0.125,0.0014833)(0.0625,0.000103456)};    
+\logLogSlopeTriangleFlip{.38}{0.1}{0.45}{3.96}{black}
 \addplot[color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5, 0.1700 )(0.25,0.0054 )(0.125,1.9613e-04)(0.0625,6.7755e-06)};
-\logLogSlopeTriangleFlip{0.32}{0.125}{0.295}{5}{magenta}
+coordinates{(0.5,0.141902)(0.25,0.00437419)(0.125,0.000154364)(0.0625,5.20702e-06)};   
+\logLogSlopeTriangleFlip{.38}{0.1}{0.26}{4.86}{magenta}
 
 \addplot[color=violet,mark=pentagon*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5, 0.0393)(0.25,9.0649e-04)(0.125, 1.7466e-05)(0.0625,3.1730e-07)};
-\logLogSlopeTriangleFlip{0.32}{0.125}{0.1075}{6}{violet}
-
-%\node at (axis cs:.03,6.8e-10) {$u$ discontinuous};
+coordinates{(0.5,0.0327909)(0.25,0.000752122)(0.125,1.34499e-05)(0.0625,2.18614e-07)};   
+\logLogSlopeTriangleFlip{0.38}{0.1}{0.075}{5.87}{violet}
+
 \legend{$N=1$,$N=2$,$N=3$,$N=4$,$N=5$}
@@
 However, because the magnitude of $\bm{\sigma}$ also decreases as $\mu/\lambda\rightarrow 0$, relative errors for $\bm{v}$ remain roughly the same magnitude as the material approaches incompressibility.
+Decreasing $\mu$ by four orders of magnitude results in up to a ten-fold increase in relative error for $\bm{\sigma}$, but less than a two-fold increase in error for $\bm{v}$.
 
@@
 \hline
- & $\mu = 1$ & $\mu = .1$ & $\mu = .01$ & $\mu = .001$ & \mu = .0001$  \\
+ & $\mu = 1$ & $\mu = .1$ & $\mu = .01$ & $\mu = .001$ & $\mu = .0001$  \\
 \hhline{|=|=|=|=|=|=|}
-$N = 2$, $h = 1/4$ &  0.0288 &  0.0270 &  0.0309 & 0.0401 & 0.0568\\
+$N = 2$, $h = 1/4$ &  3.1063e-02  &  2.6972e-02  & 3.0858e-02 &  4.0103e-02  & 5.6835e-02\\
 \hline
-$N = 3$, $h = 1/4$ & 0.0030 & 0.0027 &0.0030 & 0.0037 & 0.0049\\
+$N = 3$, $h = 1/4$ & 3.1677e-03 &  2.6848e-03 &  2.9854e-03 &  3.6960e-03 &  4.9362e-03\\
 \hline
-$N = 4$, $h = 1/4$ & 2.7705e-04 & 2.5060e-04 & 2.9185e-04 &3.9779e-04 &5.2470e-04\\
+$N = 4$, $h = 1/4$ & 2.8726e-04 &  2.4990e-04 &  2.9142e-04 &  3.9774e-04 &  5.2469e-04\\
+
 \hhline{|=|=|=|=|=|=|}
-$N = 2$, $h = 1/8$ & 0.0028  & 0.0025  & 0.0029 & 0.0036 & 0.0047 \\
+$N = 2$, $h = 1/8$ & 3.1819e-03  & 2.5476e-03 &  2.8877e-03 &  3.5608e-03  & 4.7010e-03 \\
 \hline
-$N = 3$, $h = 1/8$ & 1.8315e-04 & 1.6960e-04 &1.8524e-04 &2.2522e-04 & 2.8301e-04\\
+$N = 3$, $h = 1/8$ & 1.8867e-04  & 1.6925e-04  & 1.8509e-04  & 2.2520e-04  & 2.8301e-04\\
 \hline
-$N = 4$, $h = 1/8$ & 8.2285e-06 & 7.5813e-06 &8.1174e-06 & 1.0762e-05&1.4886e-05\\
+$N = 4$, $h = 1/8$ & 8.4999e-06  & 7.5750e-06 &  8.1094e-06  & 1.0760e-05  & 1.4886e-05\\
 \hline         
@@
 \hline
- & $\mu = 1$ & $\mu = .1$ & $\mu = .01$ & $\mu = .001$& \mu = .0001$ \\
+ & $\mu = 1$ & $\mu = .1$ & $\mu = .01$ & $\mu = .001$& $\mu = .0001$ \\
 \hhline{|=|=|=|=|=|=|}
-$N = 2$, $h = 1/4$ & 0.0656 &  0.0754 &  0.1214 & 0.2134 & 0.4145\\
+$N = 2$, $h = 1/4$ & 6.8250e-02  & 7.4130e-02 &  1.2104e-01 &  2.1333e-01 &  4.1451e-01\\
 \hline
-$N = 3$, $h = 1/4$ & 0.0091 & 0.0108 & 0.0182 & 0.0321& 0.0576\\
+$N = 3$, $h = 1/4$ & 9.2980e-03  & 1.0685e-02 &  1.8109e-02 &  3.2046e-02 &  5.7612e-02\\
 \hline
-$N = 4$, $h = 1/4$ & 9.7408e-04 & 0.0012 &  0.0022 & 0.0041 &0.0068\\
+$N = 4$, $h = 1/4$ & 9.7251e-04  & 1.1687e-03  & 2.1941e-03 &  4.1378e-03 &  6.8404e-03\\
 \hhline{|=|=|=|=|=|=|}
-$N = 2$, $h = 1/8$ & 0.0093 & 0.0116 &0.0204  & 0.0379 &0.0765  \\
+$N = 2$, $h = 1/8$ & 9.8138e-03  & 1.1429e-02  & 2.0331e-02 &  3.7889e-02 & 7.6516e-02\\
 \hline
-$N = 3$, $h = 1/8$ & 6.8288e-04 & 8.0145e-04 & 0.0015 & 0.0031 &0.0056\\
+$N = 3$, $h = 1/8$ & 6.6596e-04  & 7.8157e-04  & 1.5353e-03  & 3.1239e-03 &  5.6341e-03\\
 \hline
-$N = 4$, $h = 1/8$ &3.4231e-05 &  4.3510e-05 & 8.8747e-05 & 1.9771e-04& 3.6053e-04 \\
+$N = 4$, $h = 1/8$ & 3.4272e-05  & 4.2639e-05 &  8.8393e-05  & 1.9764e-04 &  3.6052e-04 \\
 \hline
@@
 \begin{align*}
-\bm{v}_1(x,y,t) &= \begin{cases}
+u_1(x,y,t) &= \begin{cases}
 {\rm Re}\LRp{\LRp{ikB_1e^{-kb_{1p}y} + kb_{1s}B_2 e^{-kb_{1s}y}}e^{i(ky-\omega t)}}, & y > 0\\
@@
 \end{cases}\\
-\bm{v}_2(x,y,t) &= \begin{cases}
+u_2(x,y,t) &= \begin{cases}
 {\rm Re}\LRp{\LRp{ikB_3 e^{k b_{2p} y} - kb_{2s}B_4 e^{k b_{2s}y }} e^{i(kx-\omega t)}}, & y > 0\\
@@
 Figure~\ref{subfig:stoneley1} shows errors when $K_{\rm 1D}$ is even and the mesh is fitted to the interface at $y = 0$, while Figure~\ref{subfig:stoneley2} shows errors when $K_{\rm 1D}$ is odd and the interface cuts through element interiors.
-When the mesh is fitted to the interface, we observe optimal $O(h^{N+1})$ rates of convergence for $N > 1$.
+When the mesh is fitted to the interface, computed convergence rates are between $O(h^{N+1/2})$ and $O(h^{N+1})$.
 
@@
 Since the values of $\rho,\mu$, and $\lambda$ are positive at all quadrature points, the method is energy stable.
-However, since the exact solution is discontinuous, we expect an $O(1)$ error in elements cut by the interface, resulting in $L^2$ errors which converge at rate $O(h^{1/2})$.
+However, since the exact solution is discontinuous, the error in elements cut by the interface is $O(1)$, resulting in $L^2$ errors which converge at rate $O(h^{1/2})$.
 We note that, when using piecewise constant approximations of $\mu$ and $\lambda$, we observe the same $O(h^{1/2})$ convergence rate, though errors are roughly twice as large in magnitude.
-\note{Discuss possibly extending theory for WADG to low regularity weights.}
 
@@
     ylabel={$L^2$ errors}, 
-    xmin=.04, xmax=1,
-    ymin=2.5e-10, ymax=.15,
+    xmin=.025, xmax=1,
+    ymin=.75e-11, ymax=.25,
     xmajorgrids=true,
@@
 \addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.0961)(0.25,0.0321)(0.125,0.0108)(0.0625,0.0034)}; 
-\logLogSlopeTriangleFlip{0.32}{0.125}{0.84}{2}{blue}
+coordinates{(0.5,0.108456)(0.25,0.0385149)(0.125,0.0138288)(0.0625,0.0043988)};    
+\logLogSlopeTriangleFlip{.4}{0.125}{0.86}{1.65}{blue}
 
 \addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.0088)(0.25,0.0013)(0.125,1.9177e-04)(0.0625, 2.8444e-05)};
-\logLogSlopeTriangleFlip{0.32}{0.125}{0.62}{3}{red}
+coordinates{(0.5,0.00887301)(0.25,0.00128102)(0.125,0.000186756)(0.0625,2.73376e-05)};    
+\logLogSlopeTriangleFlip{.4}{0.125}{0.65}{2.77}{red}
 
 \addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5, 8.9050e-04)(0.25,7.5838e-05)(0.125,4.9595e-06)(0.0625,3.4440e-07)};
-\logLogSlopeTriangleFlip{0.32}{0.125}{0.41}{4}{black}
+coordinates{(0.5,0.000930299)(0.25,7.92165e-05)(0.125,5.2534e-06)(0.0625,3.67676e-07)};    
+\logLogSlopeTriangleFlip{.4}{0.125}{0.475}{3.84}{black}
 
 \addplot[color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,4.2348e-05)(0.25, 1.4232e-06)(0.125, 4.8606e-08)(0.0625,1.7358e-09)};
-\logLogSlopeTriangleFlip{0.32}{0.125}{0.16}{5}{magenta}
+coordinates{(0.5,4.34823e-05)(0.25,1.45784e-06)(0.125,4.92754e-08)(0.0625,1.74105e-09)};   
+\logLogSlopeTriangleFlip{.4}{0.125}{0.26}{4.82}{magenta}
 
 \addplot[color=violet,mark=pentagon*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,2.7994e-06)(0.25,4.6958e-08)(0.125,7.9344e-10)};
-\logLogSlopeTriangleFlip{0.5}{0.125}{0.1}{6}{violet}
+coordinates{(0.5,2.97005e-06)(0.25,4.94512e-08)(0.125,8.48451e-10)(0.0625,1.47109e-11)};  
+\logLogSlopeTriangleFlip{0.4}{0.125}{0.075}{ 5.85}{violet}
 
@@
 %    ylabel={$L^2$ errors}, 
-    xmin=.04, xmax=1,
-    ymin=1e-2, ymax=.2,
+    xmin=.025, xmax=1,
+    ymin=1e-2, ymax=.25,
     xmajorgrids=true,
@@
 coordinates{(0.5, 0.0593)(0.25, 0.0436)(0.125, 0.0318)(0.0625,0.0229)};
-\logLogSlopeTriangle{0.32}{0.125}{0.225}{1/2}{black}
-
-%\addplot[dashed,color=blue,mark=*,semithick, mark options={fill=markercolor}]
-%coordinates{(0.5,0.1648)(0.25,0.0927)(0.125,0.0607)(0.0625,0.0417)}; 
-%
-%\addplot[dashed,color=red,mark=square*,semithick, mark options={fill=markercolor}]
-%coordinates{(0.5,0.1200)(0.25, 0.0855)(0.125,0.0597)(0.0625,0.0421)};
-%
-%\addplot[dashed,color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
-%coordinates{(0.5, 0.1399)(0.25,0.0942)(0.125,0.0660)(0.0625,0.0463)};
-%
-%\addplot[dashed,color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
-%coordinates{(0.5,0.1410)(0.25, 0.0984)(0.125, 0.0697)(0.0625, 0.0490)};
-
-%\addplot[color=violet,mark=pentagon*,semithick, mark options={fill=markercolor}]
-%coordinates{(0.5,2.7994e-06)(0.25,4.6958e-08)(0.125,7.9344e-10)};
-%\logLogSlopeTriangleFlip{0.5}{0.125}{0.1}{6}{violet}
+\logLogSlopeTriangle{0.425}{0.125}{0.225}{1/2}{black}
 
@@
 
-\subsubsection{Harmonic oscillation of an annulus}
-
-This example tests the convergence of WADG for curvilinear mappings.
+\subsubsection{Convergence to a reference solution}
+
+Finally, we examine the accuracy of the WADG method for smoothly varying heterogeneous media by comparing against a reference spectral method solution with order $N = 50$ on a unit square $[-1,1]^2$.
+Boundary conditions are enforced weakly through DG-like fluxes \cite{wilcox2010high}.
+We use a heterogeneous isotropic medium with $\rho,\lambda, \mu$ set to
+\[
+\rho(\bm{x}) = 1, \qquad \lambda(\bm{x}) = 1 + .25\sin(\pi x)\sin(\pi y), \qquad \mu(\bm{x}) = 1 + .25\cos(\pi x) \cos(\pi y).  
+\]
+Initial stresses are set to zero, while the initial velocity is set to 
+\[
+v_1(\bm{x}) = \cos(\pi x)\sin(\pi y), \qquad v_2(\bm{x}) = -\sin(\pi x)\cos(\pi y).
+\]
+Zero traction boundary conditions are enforced on all domain boundaries.
+Figure~\ref{fig:refsol} shows $L^2$ errors with respect to the reference solution at time $T=1/2$ for different mesh sizes and orders of approximation.
+Computed convergence rates fall between the optimal $O(h^{N+1})$ and predicted $O(h^{N+1/2})$.
+\begin{figure}
+\centering
+\begin{tikzpicture}
+\begin{loglogaxis}[
+    legend cell align=left,
+    legend style={legend pos=south east, font=\tiny},
+    width=.475\textwidth,
+    xlabel={Mesh size $h$},
+%    ylabel={$L^2$ errors}, 
+    xmin=.025, xmax=1,
+    ymin=1e-7, ymax=1.25,
+    xmajorgrids=true,
+    ymajorgrids=true,
+    grid style=dashed,
+]
+
+\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
+coordinates{(0.5,1)(0.25,0.473741)(0.125,0.141342)(0.0625,0.0355409)};    
+\logLogSlopeTriangleFlip{0.375}{0.11}{0.81}{1.99}{blue}
+
+\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
+coordinates{(0.5,0.484875)(0.25,0.0686547)(0.125,0.00767686)(0.0625,0.000933392)};   
+\logLogSlopeTriangleFlip{0.375}{0.11}{0.59}{3.04}{red}
+
+\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
+coordinates{(0.5,0.141415)(0.25,0.0126263)(0.125,0.0008361)(0.0625,5.57276e-05)};    
+\logLogSlopeTriangleFlip{0.35}{0.09}{0.42}{3.91}{black}
+
+\addplot[color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
+coordinates{(0.5,0.0452398)(0.25,0.00182005)(0.125,9.0346e-05)(0.0625,3.08518e-06)};   
+\logLogSlopeTriangleFlip{0.35}{0.08}{0.255}{4.87}{magenta}
+
+\addplot[color=violet,mark=diamond*,semithick, mark options={fill=markercolor}]
+coordinates{(0.5,0.0186937)(0.25,0.000684602)(0.125,1.12013e-05)(0.0625,2.13796e-07)};    
+\logLogSlopeTriangleFlip{0.35}{0.075}{0.1}{5.71}{violet}
+
+\legend{$N=1$,$N=2$,$N=3$,$N=4$,$N=5$}
+\end{loglogaxis}
+\end{tikzpicture}
+\caption{Convergence of WADG under mesh refinement to a reference $N=50$ spectral method solution with smoothly varying heterogeneous media for $N=1,\ldots,5$.}
+\label{fig:refsol}
+\end{figure}
+
+%\subsubsection{Harmonic oscillation of an annulus}
+%
+%This example tests the convergence of WADG for curvilinear mappings.
 
@@
 
+Next, we investigate the behavior of WADG for linear elasticity for several application examples.
+
 \subsubsection{Stiff inclusion}
 
-Leveque stiff inclusion example.
+A common test of methods for linear elastic wave propagation is the stiff inclusion problem \cite{leveque2002finite,kaser2006arbitrary,appelo2015energy}.
+We solve on a domain $[-1,1]\times [-.5, .5]$ with a rectangular inclusion located at $[-.5,.5]\times[-.1,.1]$.
+Outside of the inclusion, material parameters are taken to be
+\[
+\rho = 1, \qquad \mu = 1,\qquad \lambda = 2.
+\]
+Within the inclusion, material parameters are taken to be
+\[
+\rho = 1, \qquad \mu = 100,\qquad \lambda = 200,
+\]
+such that the wave speed in the rectangular inclusion is ten times that of the wave speed outside.  A pulse is generated through velocity boundary conditions at $x = -1$
+\[
+v_1(x,y,t) = \begin{cases}
+\sin(\pi t / t_0), & t < t_0\\
+0, & t \geq t_0
+\end{cases}, \qquad
+v_2(x,y,t) = 0.
+\]
+In our experiments, we take $t_0 = .025$.
+Traction free boundary conditions are enforced at all other domain boundaries.
+A wave is generated when the pulse hits the inclusion, bouncing back and forth and producing reflected waves which propagate outward from the stiff region.
+
+\begin{figure}
+\centering
+\subfloat[$\LRb{\bm{\sigma}_{xx}+\bm{\sigma}_{yy}}$]{\includegraphics[width=.475\textwidth]{../figs/inclusion_aligned.png}}
+\hspace{2em}
+\subfloat[$\LRb{\bm{\sigma}_{xy}}$]{\includegraphics[width=.475\textwidth]{../figs/inclusion_aligned_shear.png}}
+%\subfloat[Aligned mesh (zoom)]{\includegraphics[height=.24\textwidth]{../figs/inclusion_aligned_zoom.png}}
+%\\
+%\subfloat[Un-aligned mesh]{\includegraphics[height=.24\textwidth]{../figs/inclusion_unaligned.png}}
+%\hspace{2em}
+%\subfloat[Un-aligned mesh (zoom)]{\includegraphics[height=.24\textwidth]{../figs/inclusion_unaligned_zoom.png}}
+\caption{Values of different solution fields at $T=.4$ for the stiff inclusion problem of Leveque \cite{leveque2002finite} with order of approximation $N=5$.}
+\label{fig:incl}
+\end{figure}
+
+%We consider both aligned meshes where the inclusion boundary coincides with interior faces and an un-aligned mesh where the inclusion cuts through element interiors.
+Both are constructed by dividing each element of a quadrilateral meshes along the diagonal to produce triangular meshes.
+The aligned mesh is constructed using $100 \times 50$ quadrilateral elements in the $x$ and $y$ coordinates, respectively, while the un-aligned mesh was constructed using $101 \times 51$ elements.
+Figure~\ref{fig:incl} shows values of $\LRb{\bm{\sigma}_{xx} + \bm{\sigma}_{yy}}$ at final time $T=.4$, along with a zoomed-in view along the interface.
+The results for the interface-aligned mesh show good agreement with results in the literature \cite{leveque2002finite,kaser2006arbitrary,appelo2015energy}.
+%The sub-element resolution of $\mu,\lambda$, an error is still incurred when the mesh is un-aligned with the interface
+%For the un-aligned mesh, the sub-element resolution of material data is clearly visible, though an interface error is still incurred by the under-resolution of the interface \cite{symes2009interface}.
+
+We construct a mesh by dividing each element of a quadrilateral meshes along the diagonal to produce triangular meshes, using $100 \times 50$ quadrilateral elements in the $x$ and $y$ coordinates, respectively.
+Figure~\ref{fig:incl} shows values of $\LRb{\bm{\sigma}_{xx} + \bm{\sigma}_{yy}}$ at final time $T=.4$.
+Following \cite{leveque2002finite}, a nonlinear color scale is used in order to distinguish small-amplitude waves.
+The results for the interface-aligned mesh show qualitatively good agreement with results in the literature \cite{leveque2002finite,kaser2006arbitrary,appelo2015energy}.
+
+%\note{Finish}
 
@@
 
-From \cite{de2007arbitrary}: discontinuous media with anisotropy on one side, isotropy on the other.
-
-Also do a similar case where random noise is added to both sides and the anisotropy is more strongly discontinuous.
-
-\subsection{Three-dimensional example}
-
-Add GPU computations here.
+We next examine a model wave propagation problem in heterogeneous anisotropy media \cite{carcione1988wave,komatitsch2000simulation,de2007arbitrary}.
+The density $\rho = 7100$ is constant over the domain, while the entries of the stiffness matrix $\bm{C}$ are taken to be
+\begin{align*}
+\bm{C}_{11} &= .165, \quad \bm{C}_{12} = .05, \quad \bm{C}_{22} = .062, \quad \bm{C}_{33} = .0396, \qquad x < 0\\
+\bm{C}_{11} &= .165, \quad \bm{C}_{12} = .0858, \quad \bm{C}_{22} = .165, \quad \bm{C}_{33} = .0396, \qquad x > 0,
+\end{align*}
+with the remaining entries determined by symmetry or set to zero if unspecified.
+For $ x < 0$, this corresponds to an anisotropic material, while for $x > 0$, this corresponds to an isotropic material with $\mu = .0396, \lambda = .0858$.
+
+The computational domain is taken to be $[-.16, .16]^2$, and following \cite{komatitsch2000simulation,de2007arbitrary}, we use $N=5$ and a triangular mesh constructed by subdividing a grid of $32\times 32$ uniform quadrilaterals.
+This mesh corresponds to an element size of $.005$, which was used in \cite{komatitsch2000simulation,de2007arbitrary}.
+Forcing is applied to the $y$-component of the velocity by a Ricker pulse point source
+\[
+f(\bm{x},t) = \LRp{1 - 2(\pi f_0 (t-t_0))^2} e^{-(\pi f_0 (t-t_0))^2}\delta(x-x_0),
+\]
+where $x_0 = -.02$, $f_0 = .17$, and $t_0 = 1/f_0$.\footnote{All values and units are adapted from \cite{komatitsch2000simulation,de2007arbitrary}, and correspond to units of meters, kg, and microseconds.}
+
+\begin{figure}
+\centering
+%\subfloat[$\bm{v}_1$]{\includegraphics[width=.45\textwidth]{}}
+%\subfloat[$\bm{v}_2$]{\includegraphics[width=.45\textwidth]{}}
+\subfloat[$T=30 \mu s$]{\includegraphics[width=.4\textwidth]{../figs/aniso.png}}
+\hspace{2em}
+\subfloat[$T=60 \mu s$]{\includegraphics[width=.4\textwidth]{../figs/aniso.png}}
+\caption{Example of wave propagation in heterogeneous anisotropic media.  The vertical component of the velocity $\bm{v}_2$ is shown at $T = 30$ and $T=60$ microseconds. }
+\end{figure}
+
+\subsection{A three-dimensional example}
+
+Finally, we present a three-dimensional example of elastic wave propagation in heterogeneous media with sub-element variations, anisotropy, and a discontinuity across an interface.
+We consider elastic wave propagation on the bi-unit cube $[-1,1]^3$, with isotropic media for $z < 0$ and transversely isotropic media for $z > 0$.
+For isotropic media, we take $\mu(\bm{x}) = \lambda(\bm{x}) = w(\bm{x})$, where
+\[
+w(\bm{x}) = 1 + \cos(k\pi x)\cos(k\pi y)\cos(k\pi z).  
+\]
+For the transversely isotropic media, we take the entries of the stiffness matrix $\bm{C}$ to be 
+\begin{align*}
+\bm{C}_{11} &= \frac{2\mu + \lambda}{4}\\
+\bm{C}_{12} &= \frac{2\mu + \lambda}{4}\\
+\bm{C}_{13} &= \frac{2\mu + \lambda}{4}\\
+\bm{C}_{22} &= \frac{2\mu + \lambda}{4}\\
+\bm{C}_{23} &= \frac{2\mu + \lambda}{4}\\
+\bm{C}_{33} &= \frac{2\mu + \lambda}{4}
+\end{align*}
+
+We follow the implementation of GPU-accelerated DG methods outlined in \cite{klockner2009nodal}, which breaks the computational work in each timestep \emph{volume} and \emph{surface} kernels for the evaluation of the DG formulation and an \emph{update} step to apply a time integration method.
+For this work, apply the weight-adjusted mass matrix inverse within the update kernel as well.
+Strategies for volume and surface kernels follow \cite{klockner2009nodal}, while computational approaches for WADG are outlined in \cite{chan2016weight2}.
+More detailed computational performance analyses will be performed in future work.
+All three-dimensional computations are performed on an AMD Tahiti GPU.
 
 \section{Conclusions}
-
-Generally applicable to problems with spatially varying matrices $\bm{A}_{0}(\bm{x})$, such as in Maxwell's equations for cloaking problems \cite{li2012time}, as well as in vector-valued $H({\rm div})$ or $H({\rm curl})$ functions under matrix-valued mappings such as Piola transforms.
-
-Extension to hybrid meshes in 3D, care must be taken with pyramids \cite{chan2015orthogonal,chan2015gpu}.
-
-which can be further accelerated by exploiting structure under Bernstein-Bezier bases \cite{chan2015gpu}. 
-Formulation allows for the reduction of constant-coefficient RHS costs at high orders using BBDG \cite{chan2015bbdg}.
+\label{sec:conclusions}
+
+This work presents a weight-adjusted discontinuous Galerkin (WADG) method for the linear elastic wave equations with arbitrary heterogeneous media.
+The method is energy stable and high order accurate for arbitrary stiffness matrices, and a slight modification results in an energy stable method for curvilinear meshes as well.
+The penalty numerical fluxes for this formulation are simple to derive and implement, and their lack of dependence on the stiffness matrix allows for a unified treatment of isotropic and anisotropic media.
+Numerical examples confirm the accuracy of this method for analytic solutions in of the elastic wave equations, as well as its high order accuracy with respect to a reference solution for smoothly varying heterogeneous media.
+The behavior of this method for several application examples in also shows good agreement with existing results in the literature for both isotropic and anisotropic heterogeneous media.
+
+We note that the implementation of this method reduces to the application of the weight-adjusted mass matrix inverse and the evaluation of constant-coefficient terms in the DG formulation.
+Fast methods based on Bernstein-Bezier bases for the application of derivative and lift matrices for constant-coefficient terms can reduce this latter cost, especially at high orders of approximation \cite{chan2015bbdg}.
+Finally, we note that the approach presented here is more generally applicable to problems involving a weighted mass matrix with a matrix-valued weight, which we will explore in the future.
+These include hyperbolic systems of PDEs with a right hand side of the form $\bm{A}_{0}(\bm{x})\pd{\bm{U}}{t}$, such as Maxwell's equations for cloaking problems \cite{li2012time}.
+
+%Extension to hybrid meshes in 3D, care must be taken with pyramids \cite{chan2015orthogonal, chan2015gpu}.
 
@@
 
-The author gratefully thanks Thomas Hagstrom, Tim Warburton, Axel Modave, and Ruichao Ye for helpful and informative discussions.
+The author gratefully thanks Thomas Hagstrom, Tim Warburton, Axel Modave, Ruichao Ye, Zheng Wang, and Mario Bencomo for helpful and informative discussions.
 

### 2016-12-21 cpp/src/WaveOKL3d.cpp

@@ -105,6 +105,7 @@ occa::memory c_quad_ids;
 occa::kernel rk_update_BBWADGq;
 
 // ================ curvilinear specific kernels ====================
+
 occa::kernel rk_volume_WADG;
 occa::kernel rk_surface_WADG;
 occa::kernel rk_volume_planar;

### 2016-12-21 matlab/Codes2D/MakeCylinder2D.m

@@ -72,7 +72,7 @@ for n=1:NCurveFaces  % deform specified faces
   ids = find(abs(1-vr)>1e-7); % warp and blend
   if(f==1) blend = -(r(ids)+s(ids))./(1-vr(ids)); end;
   if(f==2) blend =      +(r(ids)+1)./(1-vr(ids)); end;
-  if(f==3) blend = -(r(ids)+s(ids))./(1-vr(ids)); end;
+  if(f==3) blend = -(r(ids)+s(ids))./(1-vr(ids)); end;    
 
   x(ids,k) = x(ids,k)+blend.*vdx(ids);
   y(ids,k) = y(ids,k)+blend.*vdy(ids);

### 2016-12-21 matlab/Elasticity2D_Stonely.m

@@ -1,12 +1,14 @@
-function Elasticity2D_Stonely
+function L2err = Elasticity2D_Stonely(N,K1D)
 
 % clear all, clear
 clear -global *
 
 Globals2D
 
-K1D = 4;
-N = 3;
+if nargin==0
+    K1D = 4;
+    N = 3;
+end
 c_flag = 0;
 FinalTime = 1;
 
@@ -128,6 +130,9 @@ tau{2} = tau0./rhoF;
 tau{3} = tau0./(2*muF + lambdaF);
 tau{4} = tau0./(2*muF + lambdaF);
 tau{5} = tau0./(2*muF + lambdaF);
+% for fld = 1:Nfld
+%     tau{fld} = tau0*ones(size(muF));
+% end
     
 
 rho = Vq*rho;
@@ -288,13 +293,13 @@ end
 
 % compute time step size
 CN = (N+1)^2/2; % guessing...
-dt = 2/(sqrt(max((2*mu(:)+lambda(:))./rho(:)))*CN*max(Fscale(:)));
+dt = .5/(sqrt(max((2*mu(:)+lambda(:))./rho(:)))*CN*max(Fscale(:)));
 
 % outer time step loop
 tstep = 0;
 
 
-figure
+% figure
 % colormap(gray)
 % colormap(hot)
 while (time<FinalTime)
@@ -314,7 +319,7 @@ while (time<FinalTime)
         
     end;
     
-    if 1 && mod(tstep,10)==0
+    if nargin==0 && mod(tstep,10)==0
         clf        
         p = U{3}+U{4}; % trace(S)
 %         p = U{1};
@@ -353,7 +358,7 @@ errU = sqrt(sum(wqJ(:).*err(:)))/sqrt(sum(wqJ(:).*uq(:)));
 serr = (Vq*U{3} - sxx(xq,yq,FinalTime)).^2 + (Vq*U{4} - syy(xq,yq,FinalTime)).^2 + (Vq*U{5} - sxy(xq,yq,FinalTime)).^2;
 sq = sxx(xq,yq,FinalTime).^2 + syy(xq,yq,FinalTime).^2 + sxy(xq,yq,FinalTime).^2;
 errS = sqrt(sum(wqJ(:).*serr(:)))/sqrt(sum(wqJ(:).*sq(:)));
-
+if nargin==0
 figure
[…]

### 2016-12-21 matlab/TV1D.m

@@ -7,10 +7,10 @@ for i = 1:N
     TV = TV + abs(u(i,:) - u(i+1,:));
 end
 
-tol = 1e-3; %max(abs(u - repmat(mean(u,1),N+1,1)))
+tol = 1e-2; %max(abs(u - repmat(mean(u,1),N+1,1)))
 for i = 2:N
     flag1 = abs(u(i,:)-u(i-1,:)) > tol;
     flag2 = abs(u(i+1,:)-u(i,:)) > tol;
-    TV = TV + abs(sign(u(i,:) - u(i-1,:)).*flag1 - sign(u(i+1,:)- u(i,:)).*flag2); % sign variations at each node    
+%     TV = TV + abs(sign(u(i,:) - u(i-1,:)).*flag1 - sign(u(i+1,:)- u(i,:)).*flag2); % sign variations at each node    
 %     TV = TV + abs(u(i+1,:) - 2*u(i,:) + u(i-1,:)); % sign variations at each node    
 end

### 2016-12-21 matlab/Vandermonde2DQuad.m

@@ -4,7 +4,7 @@ Vp = zeros(length(xp(:)), (N+1)*(N+1));
 sk = 1;
 for i=0:N    
     for j=0:N        
-        Vp(:,sk) = JacobiP(xp(:), 0, 0, i).*JacobiP(yp(:), 0, 0, j);        
+        Vp(:,sk) = JacobiP(xp(:), 0, 0, j).*JacobiP(yp(:), 0, 0, i);
         sk = sk+1;        
     end    
 end

### 2016-12-21 matlab/Wave1D.m

@@ -2,81 +2,30 @@
 
 function [l2err]= Wave1D(smax,delta,p0sigma,p0approx)
 
-if nargin < 2
-    smax = 100;
-    delta = .125;
-    p0sigma = 0;
-    p0approx = 0;
-end
-
 % Driver script for solving the 1D advection equations
 Globals1D;
 
 % Order of polymomials used for approximation
-N = 4;
+N = 5;
 
 % Generate simple mesh
-K1D = 32;
-[Nv, VX, K, EToV] = MeshGen1D(-1.0,1.0,K1D);
+a = .1;
+K1D = 2*a/.0025;
+[Nv, VX, K, EToV] = MeshGen1D(-a,a,K1D);
 
 % Initialize solver and construct grid and metric
 StartUp1D;
 
 
-% vmapP(1) = vmapM(end); % hack for periodic
-% vmapP(end) = vmapM(1); % hack for periodic
+rp = linspace(-1,1,100);
+Vp = Vandermonde1D(N,rp)/V;
+xp = Vp*x;
 
 % Set initial conditions
 % p = cos(pi/2*x);
-p = exp(-100*x.^2);
+% p = exp(-100*x.^2);
+p = zeros(size(x));
 u = zeros(size(x));
-
-global sigmax
-
-M = 3;
-L = 1-delta;
-sigmax = (x > L).*(x-L).^M + (x < -L).*(-x-L).^M;
-
-P1 = eye(N+1);
-
-% sigmax = (x > L).*(smax./(delta - (x-L)) - smax/delta) + (x < -L).*(smax./(delta - (-x-L)) - smax/delta);
-% sigmax(sigmax > 1e7) = 0;
-
-if p0sigma
-    for e = 1:K
-        sigmax(:,e) = mean(sigmax(:,e));
-    end
-end
-
-sigmax = smax*sigmax/max(sigmax(:));
-
-if p0approx
-    P1 = V*diag([1;zeros(N,1)])*inv(V); % reduce to const mode
-end
-
-sk = 1;
-for e = 1:K
-    if max(sigmax(:,e)) > 1e-8
-        pmlK(sk) = e;
-        sk = sk + 1;
-    end
-end
-nonPML = setdiff(1:K,pmlK);
-
-% sigmax(:) = smax;
-% sigmax(:,nonPML) = 0;
-
-% reverse profile
-if 0
-    sigmax = smax-sigmax; 
[…]

### 2016-12-21 matlab/Wave1D_SEM.m

@@ -6,13 +6,15 @@ function Wave1D_SEM
 Globals1D;
 
 % Order of polymomials used for approximation
-N = 1;
+N = 8;
 
-CG = 1;
+CG = 0;
+% Solve Problem
+FinalTime = 8;
 
 % Generate simple mesh
-% K1D = 16;
-K1D = 64;
+K1D = 8;
+% K1D = 64;
 [Nv, VX, K, EToV] = MeshGen1D(-1.0,1.0,K1D);
 
 % Initialize solver and construct grid and metric
@@ -26,7 +28,6 @@ Vp = Vandermonde1D(N,rp)/V;
 xp = Vp*x;
 
 % CG
-
 pairs = sort([vmapM vmapP],2);
 [~,id] = unique(pairs(:,1));
 pairs = pairs(id,:);
@@ -48,10 +49,6 @@ pex = @(x) 1-(x > -1/3) + exp(-100*(x-1/3).^2);
 p = pex(x);
 % p = exp(-100*x.^2);
 u = zeros(size(x));
-
-
-% Solve Problem
-FinalTime = 8;
 
 time = 0;
 
@@ -99,6 +96,28 @@ for tstep=1:Nsteps
     end
 end;
 
+% s = [ones(N-1,1); .5; 0];
+% a = .5; kc = 1;
+% s = ones(N+1,1); s(kc:N+1) = 1-a*(((kc:N+1) - kc)./(N-kc)).^2;
+% plot(0:N,s);return
+% F = V*diag(s)/V;
+
+re = linspace(-1,1,N+1)';
+Ve = Vandermonde1D(N,re)/V;
+xe = Ve*x;
+VB = bern_basis_1D(N,r);
+VBe = bern_basis_1D(N,re);
+F = VB*Ve;
+% F = get_BB_P1smoother(N);
+
+TV = TV1D(VB\p) + TV1D(VB\u);
+[~,id] = max(TV);
+ids = id-2:id+1; % refine in neighborhood
+
+plim = p;
+plim(:,ids) = F*plim(:,ids);
+% [plim ids] = p1limit(plim);
+
 figure
 % plot(x(:),p(:),'o','linewidth',2,'markersize',8,'MarkerFaceColor','g');
 hold on
@@ -108,9 +127,24 @@ plot(xp(:),pex(xp(:)),'-','linewidth',2);
 axis([-1 1 -.5 1.5])
 hold off
 grid on
-legend('FEM approximation','Exact solution')
[…]

### 2016-12-21 matlab/Wave2D_spectra.m

@@ -1,11 +1,11 @@
-function Advec2D_cfl
+function Wave2D_spectra
 
 Globals2D
 global bx by Vq Pq Vfq Pfq Vrq Vsq Prq Psq
 
 
 % Polynomial order used for approximation
-N = 2;
+N = 3;
 
 % Read in Mesh
 % [Nv, VX, VY, K, EToV] = MeshReaderGambit2D('Grid/Other/squarereg.neu');
@@ -66,7 +66,7 @@ dtscale = dtscale2D; dt = min(dtscale)*rmin*2/3
 if 1
     e = zeros(3*Np*K,1);
     A = zeros(3*Np*K);
-    for tau = [1e6]
+    for tau = [10]
         for i = 1:3*Np*K
             e(i) = 1;
             ids = 1:Np*K;
@@ -85,7 +85,7 @@ if 1
         end
         [W D] = eig(A);
         d = diag(D);
-        plot(d,'o')
+        plot(real(d),imag(d),'o')
         axis equal
 %         xlim([-100 1])
         %axis([-100 1 -50 50])
@@ -93,7 +93,7 @@ if 1
         title(sprintf('tau = %d\n',tau))
         drawnow
     end
-    
+    keyboard
     ids = find(abs(real(d)) < 1);
     d = d(ids);
     W = W(:,ids);
@@ -119,14 +119,12 @@ if 1
         color_line3(xp,yp,vv,vv,'.')
         view(3)
         pause
-    end
+    end       
     
-    
-    
-    [W2 D2] = eig(P*A);
-    d2 = diag(D2);
-    plot(d2,'x')
-    return
+%     [W2 D2] = eig(P*A);
+%     d2 = diag(D2);
+%     plot(d2,'x')
+%     return
 end
 
 %%

### 2016-12-21 matlab/bb_filters.m

@@ -34,13 +34,13 @@ up = Vp*u;
 % plot(xp(:),up(:),'r-.','linewidth',2)
 sum(u,1)
 W = Ve;
-W = get_BB_P1smoother(N);
-W = get_BB_smoother(N);
+% W = get_BB_P1smoother(N);
+% W = get_BB_smoother(N);
 
-a = 5; s = 2; nc = 2; n = linspace(0,1,N+1);
-d = exp(-a*((n-n(nc))./(1-n(nc))).^s);  d(1:nc) = 1;
-% clf;plot(n,d);return
-W = T*diag(d)/T;
+% a = 5; s = 2; nc = 2; n = linspace(0,1,N+1);
+% d = exp(-a*((n-n(nc))./(1-n(nc))).^s);  d(1:nc) = 1;
+% % clf;plot(n,d);return
+% W = T*diag(d)/T;
 
 [TVK] = TV1D(u);
 TV = sum(TVK(:));
@@ -55,11 +55,11 @@ u(:,ids) = W*u(:,ids);
 % u(:,ids) = W*(u(:,ids)-ulin(:,ids));
 % umodal = T\u(:,ids); u(:,ids) = u(:,ids) - T(:,1:2)*umodal(1:2,:) + ulin(:,ids);
 sum(u,1)
-uBp = Vp*u;
+uBp = Vp*f(xe);
 plot(xp(:),uBp(:),'r-.','linewidth',2)
 % plot(xe(:),u(:),'o--','linewidth',2,'markersize',10,'MarkerFaceColor','g')
 
-
+return
 % plot([x(1,:); x(end,:)],repmat(TVK,2,1),'ko-','linewidth',2,'MarkerFaceColor','g')
 %legend('Exact function','Polynomial interpolant','Bernstein coefficients','Location','NorthWest')
 % legend('Exact function','Polynomial interpolant','Shock detector')%,'Location','NorthWest')

### 2016-12-21 matlab/scratch/check_TV_2D.m

@@ -2,20 +2,23 @@ function check_TV_2D
 
 Globals2D
 
-N = 4;
+N = 5;
 
 [Nv, VX, VY, K, EToV] = MeshReaderGambit2D('Grid/Other/block2.neu');
+% [Nv, VY, VX, K, EToV] = MeshReaderGambit2D('Grid/CFD/kovA02.neu');
+% VX = VX - .5; VY = VY - .25;
 
-% [Nv, VX, VY, K, EToV] = MeshReaderGambit2D('Grid/Other/lshape.neu');
 % [Nv VX VY K EToV] = unif_tri_mesh(4);
 StartUp2D
 
-Hrefine2D(1:K);
-StartUp2D
-
+nref = 1;
+for ref = 1:nref
+    Hrefine2D(1:K);
+    StartUp2D
+end
 [r s] = Nodes2D(N); [r s] = xytors(r,s);
 [re se] = EquiNodes2D(N); [re se] = xytors(re,se);
-[rp sp] = EquiNodes2D(N+4); [rp sp] = xytors(rp,sp);
+[rp sp] = EquiNodes2D(25); [rp sp] = xytors(rp,sp);
 
 Vp = Vandermonde2D(N,rp,sp)/V;
 xp = Vp*x; yp = Vp*y;
@@ -26,26 +29,31 @@ xe = Ve*x; ye = Ve*y;
 Dr = V\Vr; Ds = V\Vs;
 Ve = bern_basis_tri(N,re,se);
 
-f = @(x,y) (y > .5*sin(pi*x)) + exp(sin(pi*x+y)) + sin(2*pi*(x+y));
+rad = @(x,y) sqrt(x.^2+y.^2);
+% f = @(x,y) (y > .5*sin(pi*x)) + exp(sin(pi*(.5*x+y))).*sin(4*pi*x).*sin(4*pi*y);
 % f = @(x,y) y > .1;
 % f = @(x,y) sin(25*pi*x+exp(y));
-% r = @(x,y) sqrt(x.^2+y.^2);
-% f = @(x,y) 1./sqrt(r(x,y));
+f = @(x,y) (rad(x+2/3,y) < 1/4) + 4*abs(rad(x,y)-1/4).*(rad(x,y) < 1/4) + exp(-36*rad(x-2/3,y).^2);
 
 Vp = bern_basis_tri(N,rp,sp);
 
 u = V\f(x,y);
 
-% up = Vp*u;
-% color_line3(xp,yp,up,up,'.'); 
+up = Vp*u;
[…]

### 2016-12-31 docs/elasticity_notes/main.tex
updates to elas paper

@@
 \newcommand{\mc}[1]{\mathcal{#1}}
+\newcommand{\snor}[1]{\left| #1 \right|}
 \newcommand{\nor}[1]{\left\| #1 \right\|}
@@
 \renewcommand{\L}{L^2\LRp{\Omega}}
-%\newcommand{\Lk}{L^2\LRp{D^k}}
-%\newcommand{\Ldk}{L^2\LRp{\partial D^k}}
-\newcommand{\Lk}{D^k}
-\newcommand{\Ldk}{\partial D^k}
+\newcommand{\Lk}{L^2\LRp{D^k}}
+\newcommand{\Ldk}{L^2\LRp{\partial D^k}}
+%\newcommand{\Lk}{D^k}
+%\newcommand{\Ldk}{\partial D^k}
 \newcommand{\Dhat}{\widehat{D}}
@@
 \begin{abstract}
-Weight-adjusted inner products \cite{chan2016weight1,chan2016weight2} replace weighted $L^2$ inner products with easily invertible weight-adjusted approximations.
-When paired with discontinuous Galerkin (DG) methods, the result is a low storage, energy stable, and (for sufficiently regular weights) high order accurate method for wave propagation in the presence of general heterogeneous media and curvilinear meshes.
-In this work, we extend weight-adjusted DG (WADG) methods to weighted inner products with matrix-valued weights, focusing on the linear elastic wave equation as an application.
-We present a unified DG formulation for isotropic and anisotropic media based on the discretization of the symmetric form of the linear elasticity wave equation, with upwind-like dissipation incorporated through simple penalty terms in the numerical fluxes.
-This formulation is provably stable for arbitrary stiffness tensors and curvilinear meshes.
-A convergence analysis is given, and numerical results confirm the stability and high order accuracy of WADG for several problems in elastic wave propagation.
+Weight-adjusted inner products \cite{chan2016weight1,chan2016weight2} are easily invertible approximations to weighted $L^2$ inner products.
+These approximations can be paired with a discontinuous Galerkin (DG) discretization to produce a time-domain method for wave propagation which is low storage, energy stable, and high order accurate for arbitrary heterogeneous media and curvilinear meshes.
+In this work, we extend weight-adjusted DG (WADG) methods to the case of matrix-valued weights, with the linear elastic wave equation as an application.
+We present a DG formulation of the symmetric form of the linear elastic wave equation, with upwind-like dissipation incorporated through simple penalty fluxes.
+A semi-discrete convergence analysis is given, and numerical results confirm the stability and high order accuracy of WADG for several problems in elastic wave propagation.
 \end{abstract}
@@
 
-Efficient and accurate simulations of elastic wave propagation form the foundation for various seismic applications, including imaging and inversion techniques, as well as rupture and earthquake simulations.
-The most common methods for the simulation of waves are finite difference methods \cite{virieux1986p}; however, finite element methods have garnered interest due to their low numerical dispersion and ability to accomodate geometrically flexible unstructured meshes.
-
-Typical methods for time-domain wave propagation utilize explicit time stepping methods, since the hyperbolic partial differential equations (PDEs) governing wave propagation typically admit less restrictive stable timestep restrictions than parabolic equations.
+Efficient and accurate methods for elastic wave propagation form a foundation for a broad range of applications, from seismic and medical imaging to rupture and earthquake simulation.
+Finite differences are the most common choice of method \cite{virieux1986p}; however, finite element methods have garnered interest due to their low numerical dispersion and ability to accommodate geometrically flexible unstructured meshes.
+
+Typical methods for time-domain wave propagation utilize explicit time stepping, since the hyperbolic partial differential equations (PDEs) which govern wave propagation admit a reasonable stable time-step restriction.
 However, finite element methods require the inversion of a global mass matrix when paired with explicit time integrators.
-On hexahedral meshes, this can be avoided by using spectral element methods (SEM), which utilize mass lumping to produce a diagonal mass matrix \cite{komatitsch1998spectral}.
-This global inversion can also be avoided through the use of discontinuous Galerkin (DG) methods, which result in a locally invertible block diagonal mass matrix.
-Due to difficulties in extending mass-lumping techniques to tetrahedra, high order DG methods are often employed for seismic simulations which use simplicial meshes \cite{kaser2006arbitrary, dumbser2006arbitrary, de2007arbitrary,delcourte2009high,delcourte2015analysis,ye2016discontinuous}.
-Such methods also lend themselves well to efficient implementations using Graphics Processing Units (GPUs) \cite{klockner2009nodal, modave2015nodal, modave2016gpu, chan2015gpu}.
+Spectral element methods (SEM) sidestep this issue on hexahedral meshes by utilizing a mass lumped diagonal mass matrix \cite{komatitsch1998spectral}.
+The inversion of a globally coupled matrix can also be avoided through the use of discontinuous Galerkin (DG) methods, which result in a locally invertible block diagonal mass matrices.
+Due to difficulties in extending mass-lumping techniques from hexahedra to tetrahedra, high order DG methods are often employed for seismic simulations which require the use of simplicial meshes \cite{kaser2006arbitrary, dumbser2006arbitrary, de2007arbitrary,delcourte2009high,delcourte2015analysis,ye2016discontinuous}.
+High order DG methods also lend themselves well to efficient implementations using Graphics Processing Units (GPUs) \cite{klockner2009nodal, modave2015nodal, modave2016gpu, chan2015gpu}.
 
 Most high order DG methods on simplicial meshes assume that material coefficients are constant over each element, which can result in spurious reflections and the loss of high order accuracy \cite{castro2010seismic,mercerat2015nodal,bencomo2015discontinuous}.
-This limitation can be overcome for DG methods by incorporating sub-element variation in media through weighted mass matrices, which results in an energy stable method \cite{mercerat2015nodal, bencomo2015discontinuous}.
-For DG methods on hexahedral meshes, mass lumping reduces the construction of weighted mass matrices to entry-wise scalings of a diagonal mass matrix.
+This limitation can be overcome by incorporating sub-element heterogeneities into weighted mass matrices, resulting in a DG method which is both high order accurate and energy stable \cite{mercerat2015nodal, bencomo2015discontinuous}.
+On tetrahedral meshes, this approach necessitates the pre-computation and storage of factorizations or inverses for each local mass matrix, which greatly increases both storage costs and data transferred at high orders of approximation.
+These costs are especially problematic for accelerator architectures such as GPUs, which possess limited memory.
+%For DG methods on hexahedral meshes, mass lumping reduces the construction of weighted mass matrices to entry-wise scalings of a diagonal mass matrix.
 However, on tetrahedral meshes, this approach necessitates the pre-computation and storage of factorizations for each local mass matrix, which greatly increases both storage costs and data transferred at high orders of approximation.
 These costs are especially problematic for accelerator architectures such as GPUs, which possess limited memory.
-%Additionally, excessive storage costs can require reduced problem sizes and under-utilization of accelerator cards.
 
@@
 For sufficiently regular weights, high order accuracy is also retained.
-When paired with an energy stable DG formulation, weight-adjusted inner products result in weight-adjusted DG methods (WADG), which preserve energy stability and high order accuracy while retaining a low asymptotic storage cost.
-Additionally, unlike mass-lumping techniques, WADG does not rely on the use of inexact quadrature rules, and reduces to the exact inversion of mass matrices for constant weights.
-
-Weight-adjusted DG methods have been applied to both acoustic wave propagation in heterogeneous media and on curvilinear meshes \cite{chan2016weight1,chan2016weight2}.
-Both cases involve scalar weighting functions.
+When paired with an energy stable DG formulation, these approximations result in weight-adjusted DG methods (WADG), which preserve energy stability and high order accuracy while retaining a low asymptotic storage cost.
+Additionally, unlike mass-lumping techniques, WADG methods do not rely on the use of inexact quadrature rules, and reduce to the exact inversion of mass matrices for constant weights.
+
+Weight-adjusted DG methods have been applied to acoustic wave propagation in heterogeneous media and on curvilinear meshes \cite{chan2016weight1,chan2016weight2}.
+Both of these previous applications have involved scalar weighting functions.
 In this work, we extend weight-adjusted inner products to matrix-valued weights, and demonstrate the usefulness of this generalization to elastic wave propagation in arbitrary media.
-A new energy stable DG formulation is introduced based on the symmetric form of the elastic wave equations, with upwind-like numerical dissipation introduced through simple penalty fluxes \cite{warburton2013low,ye2016discontinuous}.
-
-This work proceeds as follows.
-Sections~\ref{sec:symelas} and \ref{sec:dgform} begin by presenting an energy stable DG formulation with simple penalty fluxes for the elastic wave equation in symmetric hyperbolic form, and discusses issues related to storage and inversion of local mass matrices for material coefficients with sub-element variations.
-Section~\ref{sec:mwadg} extends weight-adjusted approximations to weighted $L^2$ inner products and mass matrices to the case of matrix-valued weights, and proves interpolation estimates which account for the regularity of the matrix weight.
+A new energy stable DG formulation is introduced based on the symmetric form of the elastic wave equations, with upwind-like numerical dissipation introduced through simple penalty fluxes \cite{warburton2013low}.
+In contrast to the fluxes proposed in \cite{ye2016discontinuous}, the penalty fluxes used here are independent of material coefficients.
+
+This work proceeds as follows: Sections~\ref{sec:symelas} and \ref{sec:dgform} present an energy stable DG formulation with simple penalty fluxes for the symmetric hyperbolic form of the elastic wave equation, and discuss issues related to storage and inversion of local mass matrices for material coefficients with sub-element variations.
+Section~\ref{sec:mwadg} extends weight-adjusted approximations to weighted $L^2$ inner products and mass matrices to the case of matrix-valued weights, and provides interpolation estimates which account for the regularity of the matrix weight.
 These results are incorporated into a weight-adjusted DG method for the linear elastic wave equations in Section~\ref{sec:wadgelas}.
-Finally, numerical results in Section~\ref{sec:numerical} demonstrate the accuracy of this method for several analytic solutions in linear elasticity.
-
-\section{Symmetric form of the elastic wave equation }
+Finally, numerical results in Section~\ref{sec:numerical} demonstrate the accuracy of this method for several problems in linear elasticity.
+
+\section{Symmetric form of the elastic wave equation}
 \label{sec:symelas}
-We begin with the linear elastic wave equation under small deformations in a domain $\Omega \in \mathbb{R}^d$.
+We begin with the linear elastic wave equation in a domain $\Omega \in \mathbb{R}^d$.
 These equations can be written a first order velocity-stress system for velocity $\bm{v}$ and symmetric stress tensor $\bm{S}$
@@
 where $\bm{f}$ is the body force per unit volume, $\rho$ is density, and $\tensor{\bm{C}}$ is the symmetric constitutive stiffness tensor relating stress and strain.
-We rewrite these equations as a symmetric hyperbolic system of PDEs \cite{hughes1978classical}
+We rewrite these equations as a symmetric hyperbolic system of PDEs \cite{hughes1978classical} using Voigt notation
 \begin{align}
@@
 where $\bm{C}$ is the symmetric matrix form of the constitutive tensor $\tensor{\bm{C}}$ and $\sigma$ is a vector of length $N_d = \frac{d(d+1)}{2}$, the number of unique entries of the stress tensor $\bm{S}$ in $d$ dimensions.
-We note that the matrices $\bm{A}_i$ are constant, while $\rho$ and $\bm{C}$ can vary spatially.
+We note that the matrices $\bm{A}_i$ are spatially constant, while $\rho,\bm{C}$, and $\bm{C}^{-1}$ can vary spatially.
 Furthermore, we will assume that $\rho$ and $\bm{C}$ are positive-definite and bounded pointwise such that
 \begin{align*}
-0 < \rho_{\min} &\leq \rho(\bm{x}) \leq \rho_{\max} < \infty, \qquad \forall \bm{x} \in \mathbb{R}^d,\\
-0 < c_{\min} \leq \bm{u}^T\bm{C}(\bm{x})\bm{u} &\leq c_{\max} < \infty, \qquad \forall \bm{u} \in \mathbb{R}^{N_d}, \quad \forall \bm{x} \in \mathbb{R}^d,
-\end{align*}
-which also implies that $\bm{C}^{-1}$ is pointwise positive-definite and bounded.
+0 &< \rho_{\min} \leq \rho(\bm{x}) \leq \rho_{\max} < \infty, \\
+0 &< c_{\min} \leq \bm{u}^T\bm{C}(\bm{x})\bm{u} \leq c_{\max} < \infty\\
+0 &< \tilde{c}_{\min} \leq \bm{u}^T\bm{C}^{-1}(\bm{x})\bm{u} \leq \tilde{c}_{\max} < \infty
+%, \qquad \forall \bm{u} \in \mathbb{R}^{N_d}, \quad \forall \bm{x} \in \mathbb{R}^d.
+\end{align*}
+for all $\bm{x}\in \mathbb{R}^d$ and all $\bm{u} \in \mathbb{R}^{N_d}$.
 
@@
 \]
-In three dimensions, the velocity is $\bm{v} = (\bm{v}_1,\bm{v}_2,\bm{v}_3)^T$, while $\bm{\sigma}= (\sigma_{xx},\sigma_{yy},\sigma_{zz},\sigma_{yz},\sigma_{xy},\sigma_{xz})^T$ denotes the unique entries of the stress tensor $\bm{S}$
+In three dimensions, the velocity is $\bm{v} = (\bm{v}_1,\bm{v}_2,\bm{v}_3)^T$, while $\bm{\sigma}= (\sigma_{xx},\sigma_{yy},\sigma_{zz},\sigma_{yz},\sigma_{xz},\sigma_{xy})^T$ denotes the unique entries of the stress tensor $\bm{S}$
 \[
@@
 
-We will approximate components over each element $D^k$ from an approximation space $V_h\LRp{D^k}$, which we define as the composition of the mapping $\bm{\Phi}^k$ and a reference approximation space $V_h\LRp{\widehat{D}}$
+We will approximate solution components over each element $D^k$ from an approximation space $V_h\LRp{D^k}$, which we define as the composition of the mapping $\bm{\Phi}^k$ and a reference approximation space $V_h\LRp{\widehat{D}}$
 \[
@@
 \]
+The global approximation space $V_h\LRp{\Oh}$ is then defined as the direct sum of elemental approximation spaces
+\[
+V_h\LRp{\Oh} = \bigoplus_{D^k}V_h\LRp{D^k}.  
+\]
 For the remainder of this work, we will take $V_h\LRp{\widehat{D}} = P^N\LRp{\widehat{D}}$, where $P^N\LRp{\widehat{D}}$ is the polynomial space of total degree $N$ on the reference simplex.
-In two dimensions,
+In two dimensions, $P^N$ on a triangle is
 \[
@@
 \] 
-and in three dimensions, 
+and in three dimensions, $P^N$ on a tetrahedron is
 \[
@@
 \[
-\LRp{\bm{f},\bm{g}}_{\Oh} = \sum_{D^k\in \Oh} \int_{D^k} \LRp{\bm{f},\bm{g}}_{\Lk}, \qquad \nor{\bm{f}}_{\Oh}^2 = \sum_{D^k\in \Oh} \nor{\bm{f}}^2_{\Lk}.
+\LRp{\bm{f},\bm{g}}_{\L} = \sum_{D^k\in \Oh} \int_{D^k} \LRp{\bm{f},\bm{g}}_{\Lk}, \qquad \nor{\bm{f}}_{\L}^2 = \sum_{D^k\in \Oh} \nor{\bm{f}}^2_{\Lk}.
 \]
@@
 \[
-\LRa{u,v}_{\Ldk} = \int_{\partial D^k}uv\diff{x} = \sum_{f\in \partial D^k}\int_{\widehat{f}}uvJ^f\diff{\widehat{x}}, \qquad \nor{u}_{\Ldk}^2 = \LRa{u,u},
-\]
-where $J^f$ is the Jacobian of the mapping from a reference face (triangle) $\widehat{f}$ to a physical face $f$ of an element.
+\LRa{u,v}_{\Ldk} = \int_{\partial D^k}uv\diff{x} = \sum_{f\in \partial D^k}\int_{\widehat{f}}uvJ^f\diff{\widehat{x}}, \qquad \nor{u}_{\Ldk}^2 = \LRa{u,u}_{\Ldk},
+\]
+where $J^f$ is the Jacobian of the mapping from a reference face $\widehat{f}$ to a physical face $f$ of an element.
 
@@
 \[
-{u}^- = \LRl{u}_{f \cap D^k}, \qquad {u}^+ = \LRl{u}_{f \cap  \partial D^{k,+}}.  
+{u}^- = \LRl{u}_{f \cap \partial D^k}, \qquad {u}^+ = \LRl{u}_{f \cap  \partial D^{k,+}}.  
 \]
@@
 
-We can now specify a unified DG formulation for the linear elastic wave equation (\ref{eq:symelas}) in either isotropic or anisotropic media.
+We can now specify a DG formulation for the linear elastic wave equation (\ref{eq:symelas}).
 Symmetric hyperbolic systems readily admit a DG formulation based on penalty fluxes \cite{chan2016short}.
@@
 {n}_z & 0 & {n}_x\\
-{n}_y & {n}_z & 0\\
+{n}_y & {n}_x & 0\\
 \end{array}}.  
 \]
+
+We note that, unlike the penalty DG formulation given in \cite{ye2016discontinuous}, material coefficients $\rho,\bm{C}$ appear only on the left hand side of (\ref{eq:dgform}).
+Thus, efficient techniques for constant coefficient formulations \cite{chan2015bbdg} can be used to evaluate the right hand side of the formulation, even in the presence of sub-element variations in $\rho,\bm{C}$.
+% and numerical fluxes are independent of these material coefficients if $\tau_v,\tau_\sigma$ are also independent of $\rho,\bm{C}$.
+This implies that, even in the presence of sub-element variations in $\rho,\bm{C}$, fast constant coefficient techniques \cite{chan2015bbdg} can be used in the evaluation of the right hand side of the formulation.
 
@@
 We follow \cite{leveque2002finite,wilcox2010high} and impose boundary conditions on the DG formulation through exterior values and jumps of the solution.
-Boundary conditions on the normal component of the stress can be imposed by noting that the numerical flux contains the term $\jump{\bm{A}_n^T\bm{\sigma}} = \bm{S}\bm{n}$.
+Boundary conditions on the normal component of the stress can be imposed by noting that the numerical flux contains the term $\jump{\bm{A}_n^T\bm{\sigma}} = \jump{\bm{S}\bm{n}}$.
 
@@
 \end{align*}
-while traction boundary conditions are imposed by
+while traction boundary conditions are enforced through
 \begin{align*}
@@
 
-In all cases, boundary conditions are imposed by computing numerical fluxes using these modified jumps.  This imposition guarantees energy stability for all three sets of boundary conditions.
+In all cases, boundary conditions are imposed by computing numerical fluxes using these modified jumps.
+This imposition guarantees energy stability for free surface, non-reflective, and homogeneous velocity boundary conditions.
 
@@
 
-It is straightforward to show that the formulation (\ref{eq:dgform}) is energy stable for zero velocity and traction boundary conditions, as well as non-reflective boundary conditions.
-Integrating by parts the velocity equations gives
+%It is straightforward to show that the formulation  is energy stable for zero velocity and traction boundary conditions, as well as non-reflective boundary conditions.  
+Integrating the velocity equations of (\ref{eq:dgform}) by parts gives
 \begin{align*}
@@
 \begin{align*}
-\sum_{D^k\in \Oh}\frac{1}{2}\pd{}{t}&\LRp{\LRp{\rho \bm{v},\bm{v}}_{\Lk} + \LRp{\bm{C}^{-1} \bm{\sigma},\bm{\sigma}}_{\Lk}} =  \\
-&\sum_{D^k\in \Omega_h}\LRa{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{v}}_{\partial D^k}+ \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{\sigma}}_{\partial D^k}  \\
-&\sum_{D^k\in \Omega_h}\sum_{f\in \partial D^k}\int_{f}{\bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v}} + {\frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma}}.
+\sum_{D^k\in \Oh}&\frac{1}{2}\pd{}{t}\LRp{\LRp{\rho \bm{v},\bm{v}}_{\Lk} + \LRp{\bm{C}^{-1} \bm{\sigma},\bm{\sigma}}_{\Lk}} \\
+&=\sum_{D^k\in \Omega_h}\LRa{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{v}}_{\partial D^k}+ \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{\sigma}}_{\partial D^k}  \\
+&=\sum_{D^k\in \Omega_h}\sum_{f\in \partial D^k}\int_{f}{\bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v}} + {\frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma}}.
 \end{align*}
@@
 & - \sum_{f\in \Gamma_v} \int_f \tau_v  \LRb{\bm{A}_n\bm{v}^-}^2  - \sum_{f\in \Gamma_\sigma} \int_f \tau_\sigma  \LRb{\bm{A}_n^T\bm{\sigma}^-}^2- \sum_{f\in \Gamma_{\rm abc}} \int_f \frac{\tau_v}{2} \LRb{\bm{A}_n\bm{v}^-}^2 + \frac{\tau_\sigma}{2} \LRb{\bm{A}_n^T\bm{\sigma}^-}^2 \leq 0.
+\label{eq:thm:stability}
+\end{align}
 \label{thm:stability}
-\end{align}
 \end{theorem}
 
-Since $\rho$ and $\bm{C}^{-1}$ are positive definite, the left hand side of (\ref{thm:stability}) is an $L^2$-equivalent norm on $(\bm{u},\bm{\sigma})$, and Theorem~\ref{thm:stability} implies that the magnitude of the DG solution $(\bm{u},\bm{\sigma})$ is non-increasing in time.
+Since $\rho$ and $\bm{C}^{-1}$ are positive definite, the left hand side of (\ref{eq:thm:stability}) is an $L^2$-equivalent norm on $(\bm{v},\bm{\sigma})$, and Theorem~\ref{thm:stability} implies that the magnitude of the DG solution $(\bm{v},\bm{\sigma})$ is non-increasing in time.
 This also shows that dissipation present for positive penalization constants $\tau_v,\tau_{\sigma}$ acts on non-conforming components with non-zero jumps $\bm{A}_n^T\jump{\bm{v}}$ and $\bm{A}_n\jump{\bm{\sigma}}$.
@@
 and an non-conforming part (defined through the $L^2$ orthogonal complement) corresponding to eigenvalues contain real parts which approach $-\infty$.
-For the linear elastic wave equations, these conditions are equivalent to the requirement of $C^0$ continuity for $\bm{u}$ and normal continuity of the stress tensor $\jump{\bm{S}}\bm{n} = 0$.
+For the linear elastic wave equations, these conditions are equivalent to requirements of $C^0$ continuity for $\bm{u}$ and normal continuity of the stress tensor $\jump{\bm{S}}\bm{n} = 0$.
 %\begin{align*}
@@
 \end{align}
-where $\bm{M}_f\bm{F}_{v}$ and $\bm{M}_f\bm{F}_{\sigma}$ represent contributions to the velocity and stress equations from surface integrals.
+where $\bm{F}_{v}$ and $\bm{F}_{\sigma}$ denote degrees of freedom for velocity and stress numerical fluxes.
 
 In order to apply standard time integration methods, we must invert $\bm{M}_{\rho\bm{I}}$ and $\bm{M}_{\bm{C}^{-1}}$ to isolate $\pd{\bm{v}}{t}$ and $\pd{\bm{\sigma}}{t}$ on the left hand side.
-While the inversion of $\bm{M}_{\rho\bm{I}}$ and $\bm{M}_{\bm{C}^{-1}}$ can be parallelized from element to element, doing so typically requires either the precomputation and storage of large dense matrix inverses or the on-the-fly construction and solution of a large dense matrix at every timestep.
+While the inversion of $\bm{M}_{\rho\bm{I}}$ and $\bm{M}_{\bm{C}^{-1}}$ can be parallelized from element to element, doing so typically requires either the precomputation and storage of large dense matrix inverses or the on-the-fly construction and solution of a large dense matrix system at every time-step.
 The former option requires a large amount of storage, while the latter option is computationally expensive and difficult to parallelize.
@@
 \]
-Then, $\bm{M}^{-1}_\rho = \frac{1}{\rho}\bm{M}^{-1} = \frac{1}{J\rho}\widehat{\bm{M}}^{-1}$, and $\bm{M}^{-1}_{\bm{C}^{-1}} = \bm{C}\otimes \bm{M}^{-1} = \bm{C}\otimes \LRp{\frac{1}{J}\widehat{\bm{M}}^{-1}}$, and each matrix inverse can be applied using the inverse of the reference mass matrix $\widehat{\bm{M}}^{-1}$ and the values of $\rho, \bm{C}$, and $J$ on each element.
-Applying this observation to \ref{eq:dgmat} then yields the following local system of ODEs
+Then, $\bm{M}^{-1}_\rho = \frac{1}{\rho}\bm{M}^{-1} = \frac{1}{J\rho}\widehat{\bm{M}}^{-1}$, and $\bm{M}^{-1}_{\bm{C}^{-1}} = \bm{C}\otimes \bm{M}^{-1} = \bm{C}\otimes \LRp{\frac{1}{J}\widehat{\bm{M}}^{-1}}$, and each matrix inverse can be applied using the inverse of the reference mass matrix $\widehat{\bm{M}}^{-1}$ and the values of $\rho, \bm{C}$, and $J$ over each element.
+Applying this observation to (\ref{eq:dgmat}) then yields the following local system of ODEs
 \begin{align*}
@@
 This results in a local system of ODEs with only scalar weighted mass matrices \cite{mercerat2015nodal}, which can be treated using scalar weight-adjusted approximations.
-We take a different approach in addressing these issues by approximating the matrix-weighted mass matrix $\bm{M}_{\bm{C}^{-1}}$ and corresponding matrix-weighted $L^2$ inner product with a weight-adjusted approximation which is low storage, simple to invert, energy stable, and provably high order accurate for spatially varying weights $\rho, \bm{C}$ with sufficiently regularity.
+We take a different approach in addressing these issues and approximate the matrix-weighted $L^2$ inner product (and corresponding matrix-weighted mass matrix $\bm{M}_{\bm{C}^{-1}}$) using a weight-adjusted approximation which is low storage, simple to invert, energy stable, and provably high order accurate for spatially varying weights $\rho, \bm{C}$ with sufficiently regularity.
 
@@
 \end{align}
-where $C_w$ depends on $w$ as follows:
-\[
-C_w = C\nor{w}_{L^{\infty}\LRp{D^k}}\nor{\frac{1}{w}}_{L^{\infty}\LRp{D^k}} \nor{w}_{W^{N+1,\infty}\LRp{D^k}}.
-\]
+where $C_w = C\nor{w}_{L^{\infty}\LRp{D^k}}\nor{\frac{1}{w}}_{L^{\infty}\LRp{D^k}} \nor{w}_{W^{N+1,\infty}\LRp{D^k}}$.
 \end{theorem}
-\begin{theorem}[Theorem 6 in \cite{chan2016weight1}]
-Let $u\in W^{N+1,2}\LRp{D^k}$, $w\in W^{N+1,\infty}\LRp{D^k}$, and $v \in P^M\LRp{D^k}$ for $0 \leq M\leq N$; then
-\begin{align*}
-&\LRb{\LRp{{w}u,v}_{L^2\LRp{D^k}} - \LRp{T^{-1}_{1/w}u,v}} \\
-&\leq Ch^{2N+2 - M} \nor{{w}}_{L^{\infty}\LRp{D^k}}\nor{\frac{1}{w}}_{L^{\infty}\LRp{D^k}}^2 \nor{w}^2_{W^{N+1,\infty}\LRp{D^k}}  \nor{u}_{W^{N+1,2}\LRp{D^k}} \nor{v}_{L^{\infty}\LRp{D^k}}.
-\end{align*}
-\end{theorem}
-Both theorems rely on a scalar weighted interpolation estimate derived in \cite{warburton2013low,chan2016weight1} for a general non-affine element $D^k$.  
+
+%\begin{theorem}[Theorem 6 in \cite{chan2016weight1}]
+%Let $u\in W^{N+1,2}\LRp{D^k}$, $w\in W^{N+1,\infty}\LRp{D^k}$, and $v \in P^M\LRp{D^k}$ for $0 \leq M\leq N$; then
+%\begin{align*}
+%&\LRb{\LRp{{w}u,v}_{L^2\LRp{D^k}} - \LRp{T^{-1}_{1/w}u,v}} \\
+%&\leq Ch^{2N+2 - M} \nor{{w}}_{L^{\infty}\LRp{D^k}}\nor{\frac{1}{w}}_{L^{\infty}\LRp{D^k}}^2 \nor{w}^2_{W^{N+1,\infty}\LRp{D^k}}  \nor{u}_{W^{N+1,2}\LRp{D^k}} \nor{v}_{L^{\infty}\LRp{D^k}}.
+%\end{align*}
+%\end{theorem}
+%Both theorems rely on a scalar weighted interpolation estimate derived in \cite{warburton2013low,chan2016weight1} for a general non-affine element $D^k$.  
+This theorem relies on a scalar weighted interpolation estimate derived in \cite{warburton2013low,chan2016weight1} for a general non-affine element $D^k$.  
 \begin{theorem}[Theorem 3.1 in \cite{chan2016weight1}]
-Let $D^k$ be a quasi-regular element with representative size $h = {\rm diam}\LRp{D^k}$. For $N \geq 0$, $w\in W^{N+1,\infty}$, and $u\in W^{N+1,2}$, 
-\[
-\nor{u - \frac{1}{w}\Pi_N(wu)}_{\Lk} \leq C_N h^{N+1} \nor{{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{w}}_{L^\infty}\nor{w}_{W^{N+1,\infty}}\nor{u}_{W^{N+1,2}}.
-\]
+Let $D^k$ be a quasi-regular element with representative size $h = {\rm diam}\LRp{D^k}$. For $N \geq 0$, $w\in W^{N+1,\infty}\LRp{D^k}$, and $u\in W^{N+1,2}\LRp{D^k}$, 
+\begin{align*}
+&\nor{u - \frac{1}{w}\Pi_N(wu)}_{\Lk} \leq \\
+&C h^{N+1} \nor{{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{w}}_{\LinfDk}\nor{w}_{W^{N+1,\infty}\LRp{D^k}}\nor{u}_{W^{N+1,2}\LRp{D^k}}.
+\end{align*}
 \label{thm:wproj}
@@
 
-We now aim to generalize weight-adjusted inner products to the case of matrix-valued weights.
+We now generalize weight-adjusted inner products to the case of matrix-valued weights.
 We first define appropriate generalizations of norms used in Section~\ref{sec:swadg} to vector-valued functions.
@@
 \begin{align*}
-\nor{\bm{v}}_{W^{k,p}}^p &= {\sum_{i=1}^m \nor{\bm{v}_i}_{W^{k,2}}^p} = \sum_{\LRb{\alpha} \leq k} \nor{D^{\alpha}\bm{v}}^p_{L^p}, \qquad 1 \leq p < \infty,\\
-\nor{\bm{v}}_{W^{k,\infty}} &= \max_i \nor{\bm{v}_i}_{W^{k,\infty}}.
-\end{align*}
-The corresponding Sobolev spaces $W^{k,p}$ and $W^{k,\infty}$ are defined similar to the scalar case.
-
-Let $\bm{W}(\bm{x})$ be a matrix-valued weight function which is pointwise symmetric positive-definite matrix
+\LRb{\bm{v}}_{W^{k,p}}^p &= {\sum_{i=1}^m \LRb{\bm{v}_i}_{W^{k,p}}^p}, \qquad \nor{\bm{v}}_{W^{k,p}}^p = {\sum_{i=1}^m \nor{\bm{v}_i}_{W^{k,p}}^p} \qquad 1 \leq p < \infty,\\
+\LRb{\bm{v}}_{W^{k,\infty}} &= \max_i \LRb{\bm{v}_i}_{W^{k,\infty}}, \qquad \nor{\bm{v}}_{W^{k,\infty}} = \max_i \nor{\bm{v}_i}_{W^{k,\infty}}.
+\end{align*}
+The corresponding Sobolev spaces $W^{k,p}$ and $W^{k,\infty}$ are defined similarly to the scalar case.
+
+Let $\bm{W}(\bm{x})$ be a matrix-valued weight function which is pointwise symmetric positive-definite 
 \[
@@
 \end{align*}
-While this norm is not sub-multiplicative, the following bound holds
+where $D^{\bm{\alpha}}\bm{W}(\bm{x})$ again denotes component-wise differentiation.  While this norm is not sub-multiplicative, the following bound holds
 \begin{align*}
@@
 \leq C_N \int  \sum_{\LRb{\alpha}\leq k} \sum_{\LRb{\beta}\leq \LRb{\alpha}} \nor{\LRp{D^{\beta}\bm{W}}\LRp{D^{\alpha-\beta}\bm{v}} }_{p}^p \\
-&\leq C_N\int \LRp{\sum_{\LRb{\alpha}\leq k} \nor{\LRp{D^{\alpha}\bm{W}}}}^p \LRp{\sum_{\LRb{\alpha}\leq k} \nor{D^{\alpha}\bm{v}}_{p} }^p\\
+&\leq C_N\int \LRp{\sum_{\LRb{\alpha}\leq k} \nor{\LRp{D^{\alpha}\bm{W}}}_p}^p \LRp{\sum_{\LRb{\alpha}\leq k} \nor{D^{\alpha}\bm{v}}_{p} }^p\\
 &\leq C_N \nor{\bm{W}}_{k,p,\infty}^p \nor{\bm{v}}_{k,p}^p,
@@
 \begin{theorem}
-Let $D^k$ be a quasi-regular element with representative size $h = {\rm diam}\LRp{D^k}$. For $N \geq 0$, $\bm{W} \in \LRp{W^{N+1,\infty}}^{d\times d}$, and $\bm{v}\in \LRp{W^{N+1,2}}^d$, 
-\[
-\nor{\bm{v} - \bm{W}^{-1}\Pi_N\LRp{\bm{W}\bm{v}}}_{L^2} \leq Ch^{N+1} \nor{{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk}\tilde{w}_{\max}\nor{\bm{W}}_{N+1,2,\infty} \nor{\bm{v}}_{W^{N+1,2}}
+Let $D^k$ be a quasi-regular element with representative size $h = {\rm diam}\LRp{D^k}$.
+For $N \geq 0$, $\bm{W} \in \LRp{W^{N+1,\infty}\LRp{D^k}}^{d\times d}$, and $\bm{v}\in \LRp{W^{N+1,2}\LRp{D^k}}^d$,
+\[
+\nor{\bm{v} - \bm{W}^{-1}\Pi_N\LRp{\bm{W}\bm{v}}}_{\Lk} \leq Ch^{N+1} \nor{{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk}\tilde{w}_{\max}\nor{\bm{W}}_{N+1,2,\infty} \nor{\bm{v}}_{W^{N+1,2}\LRp{D^k}}
 \]
@@
 \begin{align*}
-\nor{\bm{v} - \bm{W}^{-1}\Pi_N\LRp{\bm{W}\bm{v}}}_{L^2} &\leq C_1 \nor{\sqrt{J}}_{\LinfDk} \sup_{\bm{x}}\nor{\bm{W}^{-1}}_{2} \nor{\bm{W}\bm{v} - \Pi_N\LRp{\bm{W}\bm{v}}}_{L^2}\\
+\nor{\bm{v} - \bm{W}^{-1}\Pi_N\LRp{\bm{W}\bm{v}}}_{\Lk} &\leq C_1 \nor{\sqrt{J}}_{\LinfDk} \sup_{\bm{x}}\nor{\bm{W}^{-1}}_{2} \nor{\bm{W}\bm{v} - \Pi_N\LRp{\bm{W}\bm{v}}}_{\Lk}\\
 &\leq C_1 \nor{{\sqrt{J}}}_{\LinfDk} \sup_{\bm{x}}\nor{\bm{W}^{-1}}_{2} \LRb{\bm{W}\bm{v}}_{W^{N+1,2}\LRp{\Dhat}}\\
@@
 \[
-\LRp{\bm{W}T^{-1}_{\bm{W}}\bm{v},\bm{w}} = \LRp{\bm{v},\bm{w}} .
+\LRp{\bm{W}T^{-1}_{\bm{W}}\bm{v},\bm{w}}_{\Lk} = \LRp{\bm{v},\bm{w}}_{\Lk}.
 \]
 This definition is a straightforward generalization of $T^{-1}_{w}$ to matrix-valued weights $\bm{W}$.
-Conveniently, all properties of $T_{w},T^{-1}_{w}$ for scalar $w(\bm{x})$ carry over to matrix weights $\bm{W}(\bm{x})$ as well \cite{chan2016weight1}.
+Conveniently, all properties of $T_{w},T^{-1}_{w}$ for scalar $w(\bm{x})$ \cite{chan2016weight1} carry over to matrix weights $\bm{W}(\bm{x})$ as well.
 \begin{lemma}
@@
 \vspace{.5em}
-\item $\LRp{T^{-1}_{\bm{W}} \bm{v},\bm{w}}$ forms an inner product on $\LRp{P^N}^m \times \LRp{P^N}^m$, which is equivalent to the $L^2$ inner product with equivalence constants $C_1 = \tilde{w}_{\min}, C_2 = \tilde{w}_{\max}$.
+\item $\LRp{T^{-1}_{\bm{W}} \bm{v},\bm{w}}_{\Lk}$ forms an inner product on $\LRp{P^N}^m \times \LRp{P^N}^m$, which is equivalent to the $L^2$ inner product with equivalence constants $C_1 = \tilde{w}_{\min}, C_2 = \tilde{w}_{\max}$.
 \label{p3}
@@
 \[
-\LRp{T^{-1}_{\bm{W}} \bm{v},\bm{v}} = \LRp{\bm{W}^{-1}\bm{W}T^{-1}_{\bm{W}} \bm{v},\bm{v}} \leq \sup_{\bm{x}}\nor{\bm{W}^{-1}(\bm{x})}\LRp{\bm{W}T^{-1}_{\bm{W}} \bm{v},\bm{v}} = \tilde{w}_{\max} \nor{\bm{v}}_{L^2}^2.
+\LRp{T^{-1}_{\bm{W}} \bm{v},\bm{v}}_{\Lk} = \LRp{\bm{W}^{-1}\bm{W}T^{-1}_{\bm{W}} \bm{v},\bm{v}}_{\Lk} \leq \sup_{\bm{x}}\nor{\bm{W}^{-1}(\bm{x})}\LRp{\bm{W}T^{-1}_{\bm{W}} \bm{v},\bm{v}}_{\Lk} = \tilde{w}_{\max} \nor{\bm{v}}_{\Lk}^2.
 \]
@@
 \[
-\LRp{T^{-1}_{\bm{W}} \bm{v},\bm{v}} = \LRp{\bm{W}^{-1}\bm{W}T^{-1}_{\bm{W}} \bm{v},\bm{v}} \geq \inf_{\bm{x}}\nor{\bm{W}^{-1}(\bm{x})}\nor{\bm{v}}_{L^2}^2 = \tilde{w}_{\min}\nor{\bm{v}}_{L^2}^2 .  
+\LRp{T^{-1}_{\bm{W}} \bm{v},\bm{v}}_{\Lk} = \LRp{\bm{W}^{-1}\bm{W}T^{-1}_{\bm{W}} \bm{v},\bm{v}}_{\Lk} \geq \inf_{\bm{x}}\nor{\bm{W}^{-1}(\bm{x})}\nor{\bm{v}}_{\Lk}^2 = \tilde{w}_{\min}\nor{\bm{v}}_{\Lk}^2 .
 \]
@@
 \begin{proof}
-The proof is again similar to the scalar case.  The triangle inequality gives
+The proof follows the scalar case.  The triangle inequality gives
 \[
@@
 
-We note that these estimates are tight, in the sense that we recover estimates for scalar weight-adjusted inner products when the matrix weight is taken to be $\bm{W} = w\bm{I}$.  
-%\note{This is a good generalization because defining $T^{-1}_w \bm{v} = T^{-1}_{w\bm{I}}\bm{v}$ recovers estimates for scalar weight-adjusted inner products.}
+We note that these estimates are tight, in the sense that estimates for scalar weight-adjusted inner products are recovered when the matrix weight is taken to be $\bm{W} = w\bm{I}$.
 
@@
 
-The substitution of weight-adjusted inner products for weighted $L^2$ inner products replaces the weighted $L^2$ mass matrices in (\ref{eq:dgmat}) by their weight-adjusted approximations.  
+The use of weight-adjusted inner products replaces the weighted $L^2$ mass matrices in (\ref{eq:dgmat}) by their weight-adjusted approximations.  
 %\begin{align*}
@@
 
-One drawback of the analysis presented in this work is that accuracy is not guaranteed in the isotropic incompressible limit $\mu / \lambda \rightarrow 0$.
+One drawback of the analysis presented in this work is that accuracy of the weight-adjusted approximation is not guaranteed in the incompressible limit $\mu / \lambda \rightarrow 0$.
 In this case, the stiffness matrix $\bm{C}$ becomes singular, and the constant $\tilde{w}_{\max}$ in the upper bound on $\nor{\bm{C}^{-1}}$ blows up.
@@
 
-Since the proof of stability in Theorem~\ref{thm:stability} requires only algebraic manipulations of this formulation, the same energy stability result holds for this formulation.
-More importantly, a discrete version of energy stability is still guaranteed in the presence of inexact quadrature, where integrals in Theorem~\ref{thm:stability} are replaced with quadrature approximations.
+The proof of energy stability in Theorem~\ref{thm:stability} requires only algebraic manipulations of this formulation, and does not require that integration-by-parts holds discretely.
+This implies that a discrete version of energy stability is still guaranteed in the presence of inexact quadrature, where integrals in Theorem~\ref{thm:stability} are replaced with quadrature approximations.
 This is especially important for discretizations on curvilinear meshes, where the exact integration of spatially varying geometric factors and Jacobians can be either inhibitively expensive for high order curvilinear mappings or impossible for rational mappings \cite{engvall2016isogeometric, michoski2015foundations}.
@@
 We note that, to ensure energy stability on curved and non-affine elements, the ``skew-symmetric'' formulation must be evaluated explicitly using quadrature, which is typically more expensive than quadrature-free evaluations used to evaluate the ``strong'' DG formulation (\ref{eq:dgform}).
-These costs can be slightly reduced for curvilinear meshes by evaluating the DG formulation using the ``skew-symmetric'' formulation on curvilinear elements and the more efficient ``strong'' DG formulation on affine elements \cite{chan2016weight2}.
+These costs can be slightly reduced for most curvilinear meshes by evaluating the DG formulation using the ``skew-symmetric'' formulation on curvilinear elements and the more efficient ``strong'' DG formulation on affine elements \cite{chan2016weight2}.
 
@@
 
-In terms of group variables $\bm{U} = \LRp{\bm{v},\bm{\sigma}}$ and $\bm{V} = \LRp{\bm{w},\bm{q}} \in \LRp{V_h\LRp{\Oh}}^d\times\LRp{V_h\LRp{\Oh}}^{N_d}$, the variational formulation can be written as
-\begin{align*}
-&\LRp{T^{-1}_{\bm{A}_0^{-1}}\pd{\bm{U}}{t},\bm{V}}_{\Omega} + a(\bm{U},\bm{V}) + b(\bm{U},\bm{V}) = (\bm{f},\bm{V})\\
+Let $\bm{U},\bm{U}_h$ denote the exact and discrete WADG solutions, respectively.  We will assume that $\bm{U},\pd{\bm{U}}{t}$ are sufficiently regular such that
+\[
+\nor{\bm{U}}_{W^{N+1,2}\LRp{\Oh}}} < \infty, \qquad \nor{\pd{\bm{U}}{t}}_{W^{N+1,2}\LRp{\Oh}} < \infty,
+\]
+where we define $\nor{\bm{U}}_{W^{N+1,2}\LRp{\Oh}}}^2 = \sum_k \nor{\bm{U}}_{W^{N+1,2}\LRp{D^k}}}^2$.
+
+%We will also use Gronwall's inequality:
+%\begin{lemma}[Lemma 4.4 in \cite{dolejvsi2002finite}]
+%Let $y,q,r,z \in C[0,T]$.  Then, 
+%\[
+%y(T) + q(T) \leq z(T) + \int_0^T r(s) y(s) \diff{s}
+%\]
+%implies
+%\[
+%y(T) + q(T) + \int_0^T  r(s)q(s) e^{\int_s^T r}\diff{s} \leq z(T) + \int_0^T  r(s)z(s)e^{\int_s^T r}\diff{s}.
+%\]
+%\label{lemma:gron}
+%\end{lemma}
+
+In terms of group variables $\bm{U} = \LRp{\bm{v},\bm{\sigma}}$ and $\bm{V} = \LRp{\bm{w},\bm{q}} \in \LRp{V_h\LRp{\Oh}}^d\times\LRp{V_h\LRp{\Oh}}^{N_d}$, the WADG formulation can be written as
+\begin{align*}
+&\LRp{T^{-1}_{\bm{A}_0^{-1}}\pd{\bm{U}}{t},\bm{V}}_{\L} + a(\bm{U},\bm{V}) + b(\bm{U},\bm{V}) = (\bm{f},\bm{V})\\
 a(\bm{U},\bm{V}) &= \sum_{D^k\in \Oh} \LRp{-\LRp{ \sum_{i=1}^d \bm{\sigma},\bm{A}_i \pd{\bm{w}}{\bm{x}_i}}_{\Lk}  + \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk}} \\
@@
 \end{align*}
-where $\bm{A}_0 = {\rm diag}\LRp{\rho\bm{I},\bm{C}^{-1}}$.  Energy stability implies that
-\begin{align*}
-&b(\bm{U},\bm{U}) = \frac{1}{2}\sum_{D^k\in\Oh} \LRp{\tau_v\nor{\jump{\bm{A}_n\bm{v}}}_{\Ldk}^2 + \tau_\sigma\nor{\jump{\bm{A}^T_n\bm{\sigma}}}_{\Ldk}^2}\\
-&\LRp{T^{-1}_{\bm{A}_0^{-1}}\pd{\bm{U}}{t},\bm{U}}_{\Omega} + b(\bm{U},\bm{U}) =  (\bm{f},\bm{U}).
-\end{align*}
-Let $\bm{U},\bm{U}_h$ denote the exact and discrete WADG solutions, respectively.  Since the DG formulation (\ref{eq:dgform}) is consistent, these solutions satisfy
+where $\bm{A}_0(\bm{x}) = {\rm diag}\LRp{\rho(\bm{x})\bm{I}^{d\times d},\bm{C}^{-1}(\bm{x})}$.  Energy stability implies that
+\begin{align*}
+&b(\bm{U},\bm{U}) = \sum_{D^k\in\Oh} \LRp{\frac{\tau_v}{2}\nor{\jump{\bm{A}_n\bm{v}}}_{\Ldk}^2 + \frac{\tau_\sigma}{2}\nor{\jump{\bm{A}^T_n\bm{\sigma}}}_{\Ldk}^2}\\
+&\LRp{T^{-1}_{\bm{A}_0^{-1}}\pd{\bm{U}}{t},\bm{U}}_{\L} + b(\bm{U},\bm{U}) =  (\bm{f},\bm{U}).
+\end{align*}
+Since the DG formulation (\ref{eq:dgform}) is consistent, these solutions satisfy
 \begin{align}
-\LRp{\bm{A}_0\pd{\bm{U}}{t},\bm{V}} + a(\bm{U},\bm{V}) + b(\bm{U},\bm{V}) &= (\bm{f},\bm{V}) \nonumber\\
-\LRp{T_{\bm{A}_0^{-1}}^{-1}\pd{\bm{U}_h}{t},\bm{V}} + a(\bm{U}_h,\bm{V}) + b(\bm{U}_h,\bm{V}) &= (\bm{f},\bm{V})
+\LRp{\bm{A}_0\pd{\bm{U}}{t},\bm{V}}_{\L} + a(\bm{U},\bm{V}) + b(\bm{U},\bm{V}) &= (\bm{f},\bm{V}) \nonumber\\
+\LRp{T_{\bm{A}_0^{-1}}^{-1}\pd{\bm{U}_h}{t},\bm{V}}_{\L} + a(\bm{U}_h,\bm{V}) + b(\bm{U}_h,\bm{V}) &= (\bm{f},\bm{V})
 \label{eq:dgvswadg}
@@
 \]
+We assume that $\bm{U}_h(\bm{x},0)$ is the $L^2$ projection of the exact initial condition, such that $\left.\bm{\eta}\right|_{t=0} = 0$.
 We also introduce a consistency error $\bm{\delta} = \bm{A}_0\bm{U} - T_{\bm{A}_0^{-1}}^{-1}\bm{U}$ resulting from the approximation of $\bm{A}_0\bm{U}$ by a weight-adjusted inner product
@@
 \begin{align}
-\frac{1}{2}\pd{}{t}\LRp{T^{-1}_{\bm{A}_0^{-1}}\bm{\eta},\bm{\eta}}_{\Omega} + b(\bm{\eta},\bm{\eta}) &= \LRp{\pd{\bm{\delta}}{t},\bm{\eta}} + a(\bm{\epsilon},\bm{\eta}) + b(\bm{\epsilon},\bm{\eta}),
+\frac{1}{2}\pd{}{t}\LRp{T^{-1}_{\bm{A}_0^{-1}}\bm{\eta},\bm{\eta}}_{\L} + b(\bm{\eta},\bm{\eta}) &= \LRp{\pd{\bm{\delta}}{t},\bm{\eta}}_{\L} + a(\bm{\epsilon},\bm{\eta}) + b(\bm{\epsilon},\bm{\eta}),
 \label{eq:erreqn}
@@
 We bound $a(\bm{\epsilon},\bm{\eta}) + b(\bm{\epsilon},\bm{\eta})$ on the right hand side by integrating by parts the stress equation and using the component-wise $L^2$ orthogonality of $\bm{\epsilon}$ to derivatives of $\bm{\eta}$.
-This reduces the term $a(\bm{\epsilon},\bm{\eta}) + b(\bm{v},\bm{\eta})$ to surface contributions over each element, which can be combined with contributions from neighboring elements
+This reduces the term $a(\bm{\epsilon},\bm{\eta}) + b(\bm{v},\bm{\eta})$ to surface contributions over each element, which can be combined with contributions from neighboring elements to yield
 \begin{align*}
  \sum_{D^k\in \Oh}&\LRp{ \LRa{\bm{A}_n^T\avg{\bm{\epsilon}_\sigma} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{\epsilon}_v},\bm{\eta}_v}_{\Ldk} + \LRa{\bm{A}_n\avg{\bm{\epsilon}_v} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\epsilon}_\sigma},\bm{\eta}_\sigma}_{\Ldk}}\\
-&= \sum_{D^k\in \Oh}\LRp{ 
+&= \frac{1}{2}\sum_{D^k\in \Oh}\LRp{ 
 \LRa{\avg{\bm{\epsilon}_\sigma} - \frac{\tau_v}{2}\bm{A}_n\jump{\bm{\epsilon}_v},\bm{A}_n\jump{\bm{\eta}_v}}_{\Ldk} + \LRa{\avg{\bm{\epsilon}_v} - \frac{\tau_\sigma}{2}\bm{A}_n^T\jump{\bm{\epsilon}_\sigma},\bm{A}_n^T\jump{\bm{\eta}_{\sigma}}}_{\Ldk}}\\ %- \LRa{\bm{A}_n^T\jump{\bm{\epsilon}_\sigma},\frac{\tau_{\sigma}}{2}\bm{A}_n^T\jump{\bm{\eta}}_\sigma}_{\Ldk} }\\
-&\leq \sum_{D^k\in \Oh}\nor{\avg{\bm{\epsilon}_\sigma} - \frac{\tau_v}{2}\bm{A}_n\jump{\bm{\epsilon}_v}}_{\Ldk} \nor{\bm{A}_n\jump{\bm{\eta}_v}}_{\Ldk} + \nor{\avg{\bm{\epsilon}_v} - \frac{\tau_\sigma}{2}\bm{A}_n^T\jump{\bm{\epsilon}_\sigma}}_{\Ldk}\nor{\bm{A}_n^T\jump{\bm{\eta}_{\sigma}}}_{\Ldk}\\
+&\leq \frac{1}{2}\sum_{D^k\in \Oh}\nor{\avg{\bm{\epsilon}_\sigma} - \frac{\tau_v}{2}\bm{A}_n\jump{\bm{\epsilon}_v}}_{\Ldk} \nor{\bm{A}_n\jump{\bm{\eta}_v}}_{\Ldk} + \nor{\avg{\bm{\epsilon}_v} - \frac{\tau_\sigma}{2}\bm{A}_n^T\jump{\bm{\epsilon}_\sigma}}_{\Ldk}\nor{\bm{A}_n^T\jump{\bm{\eta}_{\sigma}}}_{\Ldk}\\
 %&\leq \LRp{\nor{\avg{\bm{\epsilon}_\sigma} - \frac{\tau_v}{2}\bm{A}_n\jump{\bm{\epsilon}_v}}_{\Ldk}  + \nor{\avg{\bm{\epsilon}_v} - \frac{\tau_\sigma}{2}\bm{A}_n^T\jump{\bm{\epsilon}_\sigma}}_{\LDk}} +  \LRp{\nor{\bm{A}_n\jump{\bm{\eta}_v}}_{\Ldk}^2 + \nor{\bm{A}_n^T\jump{\bm{\eta}_{\sigma}}}_{\Ldk}^2}\\
-&\leq C_\tau \sum_{D^k\in \Oh}\nor{\bm{\epsilon}}_{\Ldk}\frac{1}{2}\LRp{\tau_v\nor{\bm{A}_n\jump{\bm{\eta}_v}}_{\Ldk}^2 + \tau_\sigma\nor{\bm{A}_n^T\jump{\bm{\eta}_{\sigma}}}_{\Ldk}^2}^{1/2}
-\end{align*}
-where $C_\tau$ is proportional to $\max\LRp{\tau_v, \tau_\sigma}$.  Using Young's inequality with $\alpha \geq 1/\max\LRp{\tau_v,\tau_\sigma,1}$ sufficiently large yields the following bound 
-\[
-\LRb{a(\bm{\epsilon},\bm{\eta}) + b(\bm{\epsilon},\bm{\eta})} \leq \frac{C}{4\alpha}{b(\bm{\eta},\bm{\eta})} + \frac{C\alpha}{4} \sum_{D^k\in \Oh}\nor{\bm{\epsilon}}_{\Ldk}^2.
-\]
-%where $\bm{E}_v,\bm{E}_\sigma,\bm{\eta}_v,\bm{\eta}_\sigma$ denote respective components of $\bm{E},\bm{\eta}$.  
-This can be used to bound (\ref{eq:erreqn}).  Using Young's inequality on the remaining terms and absorbing terms involving $\bm{\eta}$ into the left hand side yields
-\[
-\frac{1}{2}\pd{}{t}\LRp{T^{-1}_{\bm{A}_0^{-1}}\bm{\eta},\bm{\eta}}_{\Omega} + \LRp{1-\frac{C}{4\alpha}} b(\bm{\eta},\bm{\eta}) \leq \nor{\pd{\bm{\delta}}{t}}_{\Lk} + \sum_{D^k \in \Oh}\frac{ C\alpha}{4} \nor{\bm{\epsilon}}_{\Ldk}^2.
-\]
-Integrating from $[0,T]$, and using the arithmetic-geometric mean inequality yields
-\[
-\frac{1}{2}\LRp{T^{-1}_{\bm{A}_0^{-1}}\bm{\eta},\bm{\eta}}_{\Omega} + \LRp{1-\frac{C}{4\alpha}}\int_0^T  b(\bm{\eta},\bm{\eta}) \leq \int_0^T\nor{\pd{\bm{\delta}}{t}}_{\Lk} + \int_0^T\sum_{D^k \in \Oh}C\nor{\bm{\epsilon}}_{\Ldk}.
-\]
-Using Theorem~\ref{thm:west} and a standard $hp$ trace inequality $\nor{u}_{\Ldk} \leq Ch^{-1/2} \nor{u}_{\L}$ \cite{chan2015gpu} then yields
-\[
-\nor{\bm{\eta}} \leq C_{\bm{W}} T \sup_{t\in [0,T]}\LRp{\nor{\pd{\bm{\delta}}{t}} + \nor{\bm{\epsilon}}h^{-1/2}}.
-\]
-Bounding each term on the right using interpolation estimates for sufficiently regular $\bm{U}$ yields the final result
-\[
-\nor{\bm{U}-\bm{U}_h} \leq (C_1 + C_2 T) h^{N + 1/2}\LRp{\nor{\bm{U}}_{W^{N+1,2}} + C_3 h^{1/2}\nor{\pd{\bm{U}}{t}}_{W^{N+1,2}}}.
-\]
-where $C_2$ depends on $N$ and $C_3$ depends on $\nor{\bm{W}}_{{N+1,2,\infty}}$.
-\note{Todo - finish}
+&\leq C_\tau \sum_{D^k\in \Oh}\nor{\bm{\epsilon}}_{\Ldk} \LRp{\frac{\tau_v}{2}\nor{\bm{A}_n\jump{\bm{\eta}_v}}_{\Ldk}^2 + \frac{\tau_\sigma}{2}\nor{\bm{A}_n^T\jump{\bm{\eta}_{\sigma}}}_{\Ldk}^2}^{1/2}
+\end{align*}
+where $C_\tau$ is proportional to $\max\LRp{\tau_v, \tau_\sigma}$.  Using Young's inequality with $\alpha = C_\tau / 2$ yields the following bound 
+\[
+\LRb{a(\bm{\epsilon},\bm{\eta}) + b(\bm{\epsilon},\bm{\eta})} \leq {b(\bm{\eta},\bm{\eta})} + \frac{C_\tau^2 }{4} \sum_{D^k\in \Oh}\nor{\bm{\epsilon}}_{\Ldk}^2.
+\]
+Applying this to (\ref{eq:erreqn}) and using Cauchy-Schwarz on $\LRp{\pd{\bm{\delta}}{t},\bm{\eta}}_{\L}$ then yields
+\[
+\frac{1}{2}\pd{}{t}\LRp{T^{-1}_{\bm{A}_0^{-1}}\bm{\eta},\bm{\eta}}_{\L} + b(\bm{\eta},\bm{\eta}) \leq \nor{\pd{\bm{\delta}}{t}}_{\L} \nor{\bm{\eta}}_{\L} +  b(\bm{\eta},\bm{\eta}) + \sum_{D^k \in \Oh}\frac{ C_\tau^2}{4} \nor{\bm{\epsilon}}_{\Ldk}^2.
+\]
+The right hand side term involving $\bm{\epsilon}$ can be bounded using interpolation estimates and a standard $hp$ trace inequality $\nor{u}_{\Ldk} \leq Ch^{-1/2} \nor{u}_{\Lk}$ \cite{warburton2003constants,chan2015gpu}
+\[
+\sum_{D^k \in \Oh} \nor{\bm{\epsilon}}_{\Ldk}^2 \leq \sum_{D^k \in \Oh} C h^{-1} \nor{\bm{\epsilon}}_{\Lk}^2 = Ch^{-1} \nor{\bm{\epsilon}}^2_{\L} \leq Ch^{2N+1} \nor{\bm{U}}^2_{W^{N+1,2}}.  
+\]
+We apply this estimate and eliminate factors of $b(\bm{\eta},\bm{\eta})$ on both sides.  Then, integrating from $[0,T]$ and using Lemma~\ref{lemma:props} yields
+\begin{equation}
+A_{\min}^2 \nor{\bm{\eta}}^2_{\L} \leq \int_0^T \nor{\pd{\bm{\delta}}{t}}_{\L} \nor{\bm{\eta}}_{\L} + C h^{2N+1} \nor{\bm{U}}^2_{W^{N+1,2}(\Omega)}.
+\label{eq:intermed}
+\end{equation}
+%where $C$ depends on $\tau_v,\tau_\sigma, N$, and 
+where $A_{\min} =\min\LRp{ \rho_{\min}, \tilde{c}_{\min}}$.  Since $\bm{A}_0$ is independent of $t$, Theorem~\ref{thm:west} gives
+\[
+\nor{\pd{\bm{\delta}}{t}}_{\L} = \nor{\bm{A}_0 \pd{\bm{U}}{t} - T_{\bm{A}_0^{-1}}^{-1}\pd{\bm{U}}{t}}_{\L} \leq C h^{N+1} {A}_{\max} \nor{\bm{A}_0}_{W^{N+1,\infty}(\Oh)}\nor{\pd{\bm{U}}{t}}_{\L}.
+\]
+where $A_{\max} = \max\LRp{\frac{\rho_{\max}}{\rho_{\min}}, c_{\max} \tilde{c}_{\max}}$.  
+Using this estimate and Young's inequality on $\nor{\pd{\bm{\delta}}{t}}_{\L} \nor{\bm{\eta}}_{\L}$ gives
+\[
+\nor{\bm{\eta}}_{\L}^2 \leq \int_0^T \nor{\bm{\eta}}^2_{\L} + C\LRp{h^{2N+2} \LRp{\frac{A_{\max}}{A_{\min}}}^2\nor{\bm{A}_0}_{W^{N+1,\infty}(\Oh)}^2\nor{\pd{\bm{U}}{t}}_{\L}^2  + h^{2N+1} \nor{\bm{U}}^2_{W^{N+1,2}(\Oh)}}.
+\]
+An application of Gronwall's inequality \cite{riviere2008discontinuous} and the arithmetic-geometric mean inequality then gives 
+\[
+\nor{\bm{\eta}}_{\L} \leq C e^{T} h^{N+1/2} \sup_{t \in [0,T]}\LRp{h^{1/2} \LRp{\frac{A_{\max}}{A_{\min}}}\nor{\bm{A}_0}_{W^{N+1,\infty}(\Oh)}\nor{\pd{\bm{U}}{t}}_{\L}  +  \nor{\bm{U}}_{W^{N+1,2}(\Oh)}}.
+\]
+The triangle inequality gives the final estimate
+\[
+\nor{\bm{U}-\bm{U}_h} \leq C\LRp{1 + e^{T}} h^{N + 1/2}\sup_{t \in [0,T]}\LRp{ h^{1/2}C_A\nor{\pd{\bm{U}}{t}}_{W^{N+1,2}(\Oh)}+\nor{\bm{U}}_{W^{N+1,2}(\Oh)}}.
+\]
+where $C_A = \LRp{\frac{A_{\max}}{A_{\min}}}\nor{\bm{A}_0}_{W^{N+1,\infty}(\Oh)}$.
+
+%We note that the exponential dependence of the constant $C$ on $T$ is due to the appearance of the non-conformity error $\pd{\bm{\delta}}{t}$ on the right hand side.
+If the weighted $L^2$ inner products in the DG formulation (\ref{eq:dgform}) are used instead of weight-adjusted approximations, then $\bm{\delta} = 0$.
+Then, taking the supremum over $t$ of (\ref{eq:intermed}) gives
+%\[
+%\nor{\bm{\eta}}_{\L} \leq C T \sup_{t\in [0,T]}  h^{N+1/2} \nor{\bm{U}}_{W^{N+1,2}(\Omega)},
+%\]
+%and the triangle inequality yields the estimate
+%\[
+%\nor{\bm{U}-\bm{U}_h} \leq C(1 + T) h^{N + 1/2} \LRp{h^{1/2}\nor{\bm{U}(\bm{x},0)}_{W^{N+1,2}} + \sup_{t\in[0,T]} \nor{\bm{U}}_{W^{N+1,2}} }.
+%\]
+%Similar estimates for DG and WADG were derived for the acoustic wave equation in \cite{chan2016weight1}.
+These estimates imply that the discrepancy between DG and WADG solutions should increase for long-time simulations; however, this is not observed in numerical experiments, suggesting that the bound may be sub-optimal in $T$.
 
@@
 
-In all experiments, we follow \cite{chan2016weight1} and compute application of weight-adjusted mass matrices using a quadrature exact for polynomials of degree $(2N+1)$.
-Time integration is performed using the low-storage 4th order five-stage Runge-Kutta scheme of Carpenter and Kennedy \cite{carpenter1994fourth}, and the timestep is chosen based on the global restriction
-\[
-dt = \min_{f} \frac{C_{\rm CFL}}{  \nor{\bm{C}(\bm{x})}_{2,\infty,\Omega} C_N \nor{J^f}_{L^{\infty}(f)}\nor{J^{-1}}_{L^{\infty}(f)}}
-\]
-where $C_N = O(N^2)$ is the order-dependent constant in the surface polynomial trace inequality \cite{chan2015gpu}, $C_{\rm CFL}$ is a tunable global CFL constant, and $\nor{\bm{C}}_{2,\infty,\Omega}$ denotes the spatial supremum of the $2$-norm of the stiffness matrix $\bm{C}(\bm{x})$.
-\[
-\nor{\bm{C}}_{2,\infty,\Omega} = \sup_{\bm{x}\in\Omega}\nor{\bm{C}(\bm{x})}_{2}.  
-\]
+In all experiments, we follow \cite{chan2016weight1} and compute application of weight-adjusted mass matrices using a quadrature exact for polynomials of degree $(2N+1)$ \cite{xiao2010quadrature}.
+Time integration is performed using the low-storage 4th order five-stage Runge-Kutta scheme of Carpenter and Kennedy \cite{carpenter1994fourth}, and the time-step is chosen based on the global restriction
+\[
+dt = \min_{k} \frac{C_{\rm CFL}}{  \sup_{\bm{x}\in\Omega}\nor{\bm{C}(\bm{x})}_{2} C_N \nor{J^f}_{L^{\infty}\LRp{\partial D^k}}\nor{J^{-1}}_{L^{\infty}\LRp{D^k}}}
+\]
+where $C_N = O(N^2)$ is the order-dependent constant in the surface polynomial trace inequality \cite{chan2015gpu} and $C_{\rm CFL}$ is a tunable global CFL constant.
+Finally, in all following experiments, we use $\tau_v = \tau_\sigma = 1$ unless specified otherwise.
 
@@
 We first verify the energy stability of the WADG method for arbitrary heterogeneous media.
-We use a stiffness matrix constructed using random numbers and similarity transforms, such that at every quadrature point, $\bm{C}(\bm{x}) = \bm{U}\bm{D}\bm{U}^{T}$, where $\bm{D}$ is a diagonal matrix with positive entries $d_{\min} \leq \bm{D}_{ii} \leq d_{\max}$ and $\bm{U}$ is a random unitary matrix.
+We use a stiffness matrix constructed using similarity transforms, such that at every quadrature point, $\bm{C}(\bm{x}) = \bm{U}\bm{D}\bm{U}^{T}$, where $\bm{D}$ is a diagonal matrix with random positive entries $d_{\min} \leq \bm{D}_{ii} \leq d_{\max}$ and $\bm{U}$ is a random unitary matrix.
 Let $\bm{A}_h$ denote the matrix induced by the global semi-discrete DG formulation, such that the time evolution of the solution $\bm{v},\bm{\sigma}$ is governed by
@@
 Figure~\ref{fig:spectra} shows computed eigenvalues of $\bm{A}_h$ for $[d_{\min},d_{\max}] = [1/10,1]$ and $[10^{-4},1]$, $\tau_v=\tau_\sigma = 0$ and $\tau_v=\tau_\sigma = 1$ under discretization parameters $N=4$ and $h = 1/4$.
+In both cases, the largest real part of any eigenvalue is $O(10^{-14})$, verifying the energy stability of the WADG discretization for arbitrary media.
 
@@
 
-We note that, for $\tau_v = \tau_\sigma = 1$, $\nor{\bm{A}_h}$ is approximately twice as large compared to when $\tau_v = \tau_\sigma = 0$.
-This can result in a more restrictive stability condition on the timestep.
-We observe that the growth in $\nor{\bm{A}_h}$ for $\tau_v,\tau_\sigma = 1$ is due to the presence of eigenvalues with large negative real parts, which implies that larger timesteps may be used under time integrators with extended stability regions along the real axis \cite{abdulle2002fourth}.
-
-Alternatively, we can avoid growth in $\nor{\bm{A}_h}$ by scaling $\tau$.
+For practical simulations, the choice of $\tau_v, \tau_\sigma$ remains to be specified.
+Increasing $\tau_v,\tau_\sigma$ results in greater damping of under-resolved spurious components of the solution \cite{chan2016short}; however, a naive selection of these penalty parameters can result in an overly restrictive time-step restriction for stability.
+We wish to choose $\tau_v,\tau_\sigma$ as large as possible without increasing $\nor{\bm{A}_h}$.
+
 To avoid artificially increasing the stiffness of (\ref{eq:wadgmat}), it was argued in \cite{chan2016short} that the magnitude of the penalty parameter should depend on the largest singular value of $\bm{A}_n$, which is $O(1)$ in two and three dimensions.
-However, this does not factor in the effect of $\rho$ and $\bm{C}^{-1}$ appearing in the left hand side.
-These coefficients can be taken into account by scaling only the stress penalization parameter with the average of local max norms over a face $f$
-\[
-\tau_{\sigma,f} \approx \frac{2}{\nor{\bm{C}^+(\bm{x})}_{2,\infty,f}+\nor{\bm{C}^-(\bm{x})}_{2,\infty,f}},
-\]
-where $\bm{C}^+,\bm{C}^-$ denote values of the stiffness matrix on elements which share the face $f$.
+However, this suggestion does not factor in the effect of $\rho$ and $\bm{C}^{-1}$ appearing in the left hand side.
+These material coefficients can be taken into account by scaling the penalty parameters such that flux terms are dimensionally consistent.
+One example of such a scaling is
+\[
+\tau_v = \gamma_v \sup_{\bm{x} \in f} \sqrt{\nor{\avg{\bm{C}(\bm{x})}}\avg{\rho(\bm{x})}}, \qquad \tau_\sigma =  \gamma_\sigma  \sup_{\bm{x} \in f} \frac{1}{\sqrt{\nor{\avg{\bm{C}(\bm{x})}}\avg{\rho(\bm{x})}}},
+\]  
+where the supremum is taken locally over each face and $\gamma_v, \gamma_\sigma$ are dimensionless constants.
+%We adopt this scaling for all remaining experiments, with $\gamma_v = 1/2$, $\gamma_\sigma = 1$
+
+%
+
+%We note that, for $\tau_v = \tau_\sigma = 1$, $\nor{\bm{A}_h}$ is over twice as large compared to when $\tau_v = \tau_\sigma = 0$.
+This can result in a more restrictive stability condition on the time-step.
+We observe that the growth in $\nor{\bm{A}_h}$ for $\tau_v,\tau_\sigma = 1$ is due to the presence of eigenvalues with large negative real parts, which suggests that larger time-steps may be used under time integrators with extended stability regions along the real axis \cite{abdulle2002fourth}.
+
+%Under this scaling, taking $\gamma_v = 1, \gamma_\sigma / d$, where $d$ is the spatial dimension
+
+%These coefficients can be taken into account by scaling only the stress penalization parameter with the average of local max norms over a face $f$
+%\[
+%\tau_{\sigma,f} \approx \frac{2}{\nor{\bm{C}^+(\bm{x})}_{2,\infty,f}+\nor{\bm{C}^-(\bm{x})}_{2,\infty,f}},
+%\]
+%where $\bm{C}^+,\bm{C}^-$ denote values of the stiffness matrix on elements which share the face $f$.
 It may also be advantageous to rescale $\tau_\sigma$ for each individual component of stress based on the rows of $\bm{C}$.
 Under this rescaling of the stress penalty parameter, we observe numerically that the norm of $\bm{A}_h$ remains roughly the same magnitude as for $\tau_p = \tau_\sigma = 0$.
-%Since we do not observe significant changes in computed $L^2$, we adopt this rescaling of $\tau_\sigma$ in all reported results.
-%We note that this rescaling of the penalty parameters also avoids other issues associated with over-penalization, such as the reintroduction of spurious modes which would otherwise be damped under less aggressive penalties \cite{chan2016short}.
-
-In the following experiments, we use $\tau_v = \tau_\sigma = 1$.  However, we do not observe significant changes in computed $L^2$ errors under the aforementioned rescaling of $\tau_\sigma$.
+We do not observe significant changes in computed $L^2$ errors under the aforementioned rescaling of $\tau_\sigma$.
 
@@
 Next, we study the accuracy and convergence of weight-adjusted DG method for several analytical solutions in linear elasticity.
-In all cases, the solution in expressed terms of the displacement vector $\bm{u}(\bm{x},t) = (u_1,\ldots,u_d)$.
+In all cases, the solution is expressed terms of the displacement vector $\bm{u}(\bm{x},t) = (u_1,\ldots,u_d)$.
 Initial conditions for velocity and stress are computed through
 \[
-\bm{v}(\bm{x},t) = \pd{\bm{u}}{t}, \qquad \bm{\sigma} = \bm{C}\bm{E} = \bm{C}\frac{1}{2}\LRp{\Grad\bm{u} + \Grad\bm{u}^T}.  
+\bm{v}(\bm{x},t) = \pd{\bm{u}}{t}, \qquad \bm{\sigma} = \bm{C}\frac{1}{2}\LRp{\Grad\bm{u} + \Grad\bm{u}^T}.  
 \]
@@
 
-We first examine convergence  on a unit square domain with  $\lambda = \mu = \rho = 1$.   The components of the displacement vector is given by
+We first examine convergence  on a unit square domain with  $\lambda = \mu = \rho = 1$.   The components of the displacement vector are given by
 \begin{align*}
@@
 \]
-In our computations, we use $\rho = \mu = \lambda = 1$, $\xi = 0.949554083888034$, and $\omega = 2\pi$.
+In our computations, we use $\rho = \mu = \lambda = 1$, $\xi = 0.949554083888034$, and $\omega = 2\pi$ \cite{wilcox2010high}.
 We solve on the domain $[0,2]\times[0,1]$ using a sequence of uniform triangular meshes, and enforce traction-free boundary conditions at $x=0$ and exact Dirichlet boundary conditions at $x=2$.
@@
 In our experiments, we use $\rho = \mu = 1$, $\lambda = 2$, $k = 2\pi$.
-For these values, $\omega = 13.137063197233$, $B_1 = 126.1992721468$ and $B_2 = 53.88807700007$.
+For these values, $\omega = 13.137063197233$, $B_1 = 126.1992721468$ and $B_2 = 53.88807700007$ \cite{wilcox2010high}.
 We solve on the domain $[-1,1]\times[-1/2,1/2]$, with traction-free boundary conditions at $y = \pm 1/2$ and periodic boundary conditions at $x = \pm 1$.
@@
 \end{tikzpicture}
+}
 \label{fig:rayleigh}
-}
 \caption{Convergence of $L^2$ errors for Rayleigh and Lamb wave solutions.}
@@
 
-As noted in Section~\ref{sec:wadgelas}, error estimates for isotropic elasticity no longer hold in the incompressible limit with $\mu / \lambda \rightarrow 0$ limit due to the fact that $\bm{C}$ becomes singular.
+As noted in Section~\ref{sec:wadgelas}, error estimates for isotropic elasticity no longer hold in the incompressible limit $\mu / \lambda \rightarrow 0$ due to the fact that $\bm{C}$ becomes singular.
 We use the propagation of Rayleigh waves to examine the behavior of WADG for near-incompressible materials.
@@
 This is not surprising, as the constant in the error estimates of Section~\ref{sec:mwadgprop} blows up in the incompressible limit.
-However, because the magnitude of $\bm{\sigma}$ also decreases as $\mu/\lambda\rightarrow 0$, relative errors for $\bm{v}$ remain roughly the same magnitude as the material approaches incompressibility.
+However, because the magnitude of $\bm{\sigma}$ also decreases as $\mu/\lambda\rightarrow 0$, relative errors for $\bm{v}$ remain roughly the same magnitude for near-incompressible materials.
 Decreasing $\mu$ by four orders of magnitude results in up to a ten-fold increase in relative error for $\bm{\sigma}$, but less than a two-fold increase in error for $\bm{v}$.
@@
 
-%\note{Examine energy stability as $\mu/\lambda\rightarrow 0$.  Matrix $\bm{C}$ becomes singular.}
-
 \subsubsection{Stoneley waves}
@@
 The Stoneley wave speed $c_{st}$ can be determined based material parameters and interface conditions, and the amplitudes $B_1,B_2,B_3,B_4$ are determined from $c_{st}$ up to scaling by a constant.
-For the parameters used in this study, we take $c_{st} = 0.546981324213884$, $B_1 = i0.2952173626624, B_2 = -0.6798795208473, B_3 = i0.5220044931212$, and $B_4 = -0.9339639688697$.
+For the parameters used in this study, we take $c_{st} = 0.546981324213884$, $B_1 = i0.2952173626624, B_2 = -0.6798795208473, B_3 = i0.5220044931212$, and $B_4 = -0.9339639688697$ \cite{wilcox2010high}.
 We assume $k =1$, which gives $\omega=c_{st}$.
@@
 Figure~\ref{subfig:stoneley1} shows errors when $K_{\rm 1D}$ is even and the mesh is fitted to the interface at $y = 0$, while Figure~\ref{subfig:stoneley2} shows errors when $K_{\rm 1D}$ is odd and the interface cuts through element interiors.
+
 When the mesh is fitted to the interface, computed convergence rates are between $O(h^{N+1/2})$ and $O(h^{N+1})$.
-
 When the mesh is not fitted to the interface exactly, we compute the application of the weight-adjusted mass matrix inverse using a quadrature rule from Xiao and Gimbutas \cite{xiao2010quadrature} which is exact for degree $2N+1$ polynomials.
@@
 }
-\caption{Convergence of WADG for a Stoneley wave using a fitted mesh aligned with the interface and a non-fitted mesh where the interface does not lie exactly on an element boundary. }
+\caption{Convergence of WADG for a Stoneley wave using a fitted mesh aligned with the interface (Figure~\ref{subfig:stoneley1}) and a non-fitted mesh where the interface does not lie exactly on an element boundary (Figure~\ref{subfig:stoneley2}).
+}
 \label{fig:stoneley}
@@
 Finally, we examine the accuracy of the WADG method for smoothly varying heterogeneous media by comparing against a reference spectral method solution with order $N = 50$ on a unit square $[-1,1]^2$.
-Boundary conditions are enforced weakly through DG-like fluxes \cite{wilcox2010high}.
+Boundary conditions are enforced weakly through numerical fluxes \cite{wilcox2010high}.
 We use a heterogeneous isotropic medium with $\rho,\lambda, \mu$ set to
@@
 
-Next, we investigate the behavior of WADG for linear elasticity for several application examples.
+We next demonstrate the accuracy and flexibility of WADG for several application-based problems in linear elasticity with heterogeneity and anisotropy.
 
@@
 
-A common test of methods for linear elastic wave propagation is the stiff inclusion problem \cite{leveque2002finite,kaser2006arbitrary,appelo2015energy}.
-We solve on a domain $[-1,1]\times [-.5, .5]$ with a rectangular inclusion located at $[-.5,.5]\times[-.1,.1]$.
-Outside of the inclusion, material parameters are taken to be
+The stiff inclusion problem  is a common test of methods for linear elastic wave propagation \cite{leveque2002finite,kaser2006arbitrary,appelo2015energy}, where an inclusion with higher wavespeed is embedded within a non-stiff region.
+Waves which reach this region of high wavespeed are transmitted through the inclusion, bouncing back and forth within the region.
+This vibration then produces waves which propagate outward from the inclusion.
+
+We solve on a domain $[-1,1]\times [-.5, .5]$ with a rectangular inclusion located at $[-.5,.5]\times[-.1,.1]$.  Outside of the inclusion, material parameters are taken to be 
 \[
@@
 \]
-In our experiments, we take $t_0 = .025$.
-Traction free boundary conditions are enforced at all other domain boundaries.
-A wave is generated when the pulse hits the inclusion, bouncing back and forth and producing reflected waves which propagate outward from the stiff region.
+In our experiments, we take $t_0 = .025$.  Traction free boundary conditions are enforced at all other domain boundaries.
 
@@
 
-We construct a mesh by dividing each element of a quadrilateral meshes along the diagonal to produce triangular meshes, using $100 \times 50$ quadrilateral elements in the $x$ and $y$ coordinates, respectively.
+We construct a uniform triangular mesh by dividing each element of a quadrilateral meshes along the diagonal to produce triangular meshes, using $100 \times 50$ quadrilateral elements in the $x$ and $y$ coordinates, respectively.
+Penalty parameters of $\tau_v = \tau_\sigma = 1/2$ are chosen.
 Figure~\ref{fig:incl} shows values of $\LRb{\bm{\sigma}_{xx} + \bm{\sigma}_{yy}}$ at final time $T=.4$.
 Following \cite{leveque2002finite}, a nonlinear color scale is used in order to distinguish small-amplitude waves.
-The results for the interface-aligned mesh show qualitatively good agreement with results in the literature \cite{leveque2002finite,kaser2006arbitrary,appelo2015energy}.
+The results show qualitatively good agreement with results in the literature \cite{leveque2002finite,kaser2006arbitrary,appelo2015energy}.
 
@@
 
-The computational domain is taken to be $[-.16, .16]^2$, and following \cite{komatitsch2000simulation,de2007arbitrary}, we use $N=5$ and a triangular mesh constructed by subdividing a grid of $32\times 32$ uniform quadrilaterals.
-This mesh corresponds to an element size of $.005$, which was used in \cite{komatitsch2000simulation,de2007arbitrary}.
-Forcing is applied to the $y$-component of the velocity by a Ricker pulse point source
+The computational domain is taken to be $[-.32, .32]^2$, and following \cite{komatitsch2000simulation,de2007arbitrary}, we use $N=5$ and a triangular mesh of $32768$ elements constructed by subdividing a grid of $128\times 128$ uniform quadrilaterals.
+This mesh corresponds to an element size of $.005$, as used in \cite{komatitsch2000simulation,de2007arbitrary}.
+Forcing is applied to the $y$-component of the velocity by a Ricker wavelet point source
 \[
@@
 \]
-where $x_0 = -.02$, $f_0 = .17$, and $t_0 = 1/f_0$.\footnote{All values and units are adapted from \cite{komatitsch2000simulation,de2007arbitrary}, and correspond to units of meters, kg, and microseconds.}
-
+where $x_0 = -.02$, $f_0 = .17$, and $t_0 = 1/f_0$.\footnote{All values and units are adapted from \cite{komatitsch2000simulation,de2007arbitrary}, and correspond to units of meters, kg, and microseconds.}  The penalty parameters are taken to be $\tau_\rho = \tau_\sigma = 1/2$.
+Figure~\ref{fig:aniso} shows the $y$-component of velocity $\bm{v}_2$ at times $T = 30 \mu s$ (zoomed in) and $T= 60 \mu s$.
+Both results show qualitative agreement with reference results from \cite{carcione1988wave,komatitsch2000simulation,de2007arbitrary}.
 \begin{figure}
@@
 %\subfloat[$\bm{v}_2$]{\includegraphics[width=.45\textwidth]{}}
-\subfloat[$T=30 \mu s$]{\includegraphics[width=.4\textwidth]{../figs/aniso.png}}
-\hspace{2em}
-\subfloat[$T=60 \mu s$]{\includegraphics[width=.4\textwidth]{../figs/aniso.png}}
-\caption{Example of wave propagation in heterogeneous anisotropic media.  The vertical component of the velocity $\bm{v}_2$ is shown at $T = 30$ and $T=60$ microseconds. }
+\subfloat[$T=30 \mu s$ (zoomed in)]{\includegraphics[height=18em]{../figs/aniso1.png}}
+\hspace{3em}
+\subfloat[$T=60 \mu s$]{\includegraphics[height=18.25em]{../figs/aniso2.png}}
+\caption{An example of wave propagation in heterogeneous anisotropic media.  The vertical component of the velocity $\bm{v}_2$ is shown at $T = 30$ and $T=60$ microseconds. }
+\label{fig:aniso}
 \end{figure}
@@
 
-Finally, we present a three-dimensional example of elastic wave propagation in heterogeneous media with sub-element variations, anisotropy, and a discontinuity across an interface.
-We consider elastic wave propagation on the bi-unit cube $[-1,1]^3$, with isotropic media for $z < 0$ and transversely isotropic media for $z > 0$.
-For isotropic media, we take $\mu(\bm{x}) = \lambda(\bm{x}) = w(\bm{x})$, where
-\[
-w(\bm{x}) = 1 + \cos(k\pi x)\cos(k\pi y)\cos(k\pi z).  
-\]
-For the transversely isotropic media, we take the entries of the stiffness matrix $\bm{C}$ to be 
-\begin{align*}
-\bm{C}_{11} &= \frac{2\mu + \lambda}{4}\\
-\bm{C}_{12} &= \frac{2\mu + \lambda}{4}\\
-\bm{C}_{13} &= \frac{2\mu + \lambda}{4}\\
-\bm{C}_{22} &= \frac{2\mu + \lambda}{4}\\
-\bm{C}_{23} &= \frac{2\mu + \lambda}{4}\\
-\bm{C}_{33} &= \frac{2\mu + \lambda}{4}
-\end{align*}
-
-We follow the implementation of GPU-accelerated DG methods outlined in \cite{klockner2009nodal}, which breaks the computational work in each timestep \emph{volume} and \emph{surface} kernels for the evaluation of the DG formulation and an \emph{update} step to apply a time integration method.
-For this work, apply the weight-adjusted mass matrix inverse within the update kernel as well.
+Finally, we present a three-dimensional example of elastic wave propagation in heterogeneous media with sub-element variations and a discontinuity across an interface.
+We consider isotropic elastic wave propagation on the cube $[-.5,.5]^3$ with a discontinuity in material coefficients across $z = 0$
+\[
+\rho = 1, \qquad \mu(\bm{x}) = \begin{cases}
+2 + w(\bm{x}), & z < 0\\
+1 + w(\bm{x}), & z > 0
+\end{cases},
+\qquad
+\lambda(\bm{x}) = \begin{cases}
+2, & z < 0\\
+1, & z > 0
+\end{cases}
+\]
+where $w(\bm{x})$ is a spatially varying weight.  
+%isotropic media for $z < 0$ and a transversely isotropic media for $z > 0$.
+For isotropic media, we take $\mu(\bm{x}) = w(\bm{x})$ and $\lambda(\bm{x}) = 2w(\bm{x})$, where $w(\bm{x})$ is a spatially varying weight function to be specified.
+For $z > 0$, we take $\bm{C}$ to be
+%\[
+%\bm{C} &= w(\bm{x}) \LRp{
+%\begin{array}{cccccc}
+%c_{11} & c_{12} & \lambda & & &\\
+%c_{12} & c_{11} & \lambda & & &\\
+%\lambda & \lambda & 2\mu+\lambda & & &\\
+%& & & \mu & &\\
+%& & & & \mu &\\
+%& & & & & (c_{11}-c_{12})/2\\
+%\end{array}
+%}
+%\]
+%with $c_{11} = 1$ and $c_{12} = .5$.  
+Forcing is applied to the $x$-component of velocity through a smoothed point source and Ricker wavelet
+\[
+f(\bm{x},t) = \LRp{1 - 2(\pi f_0 (t-t_0))^2} e^{-(\pi f_0 (t-t_0))^2}e^{- \LRp{a \nor{\bm{x}-\bm{x}_0}}^2}
+\]
+where $\bm{x}_0 = (0,0,.1)^T$, $a = 100$, $f_0 = 10$, and $t_0 = 1/f_0$.  
+\begin{figure}
+\centering
+\subfloat[Computational mesh]{\includegraphics[width=.28\textwidth]{../figs/cubeSplitFine.png}\label{subfig:mesh}}
+\hspace{.5em}
+\subfloat[$w(\bm{x}) = 0$]{\includegraphics[width=.35\textwidth]{../figs/pplane0.png}}
+%\hspace{.25em}
+\subfloat[$w(\bm{x})  = .5 \cos(3\pi x)\cos(3\pi y)\cos(3\pi z) $]{\includegraphics[width=.35\textwidth]{../figs/pplanew0.png}}
+\caption{Mesh and $xz$, $xy$ slices of $\bm{v}_1$ at $T = .5$.  The order of approximation is taken to be $N=5$.  }
+\label{fig:3d}
+\end{figure}
+Figure~\ref{fig:3d} shows the $x$-velocity of the computed solution at $T = .5$ with
+\[
+w(\bm{x}) = .5 \cos(3\pi x)\cos(3\pi y)\cos(3\pi z), 
+\]  
+as well as $w(\bm{x}) = 0$ for reference.  Figure~\ref{subfig:mesh} shows the unstructured mesh of 222824 tetrahedral elements of degree $N=5$ used to compute both solutions.
+
+ These computations are performed on an AMD Tahiti GPU, following the implementation of GPU-accelerated DG methods outlined in \cite{klockner2009nodal}.
+This approach breaks the computational work for each time-step into \emph{volume} and \emph{surface} kernels (for the evaluation of the DG formulation) and an \emph{update} kernel (for the application of a time integration method).
+In this implementation, we apply the weight-adjusted mass matrix inverse within the update kernel as well.
 Strategies for volume and surface kernels follow \cite{klockner2009nodal}, while computational approaches for WADG are outlined in \cite{chan2016weight2}.
-More detailed computational performance analyses will be performed in future work.
-All three-dimensional computations are performed on an AMD Tahiti GPU.
 
@@
 Numerical examples confirm the accuracy of this method for analytic solutions in of the elastic wave equations, as well as its high order accuracy with respect to a reference solution for smoothly varying heterogeneous media.
-The behavior of this method for several application examples in also shows good agreement with existing results in the literature for both isotropic and anisotropic heterogeneous media.
+Results obtained using this method also show good agreement with existing results in the literature for both problems involving both isotropic and anisotropic heterogeneous media.
 
 We note that the implementation of this method reduces to the application of the weight-adjusted mass matrix inverse and the evaluation of constant-coefficient terms in the DG formulation.
-Fast methods based on Bernstein-Bezier bases for the application of derivative and lift matrices for constant-coefficient terms can reduce this latter cost, especially at high orders of approximation \cite{chan2015bbdg}.
-Finally, we note that the approach presented here is more generally applicable to problems involving a weighted mass matrix with a matrix-valued weight, which we will explore in the future.
+The cost of the latter step can be reduced (especially at high orders of approximation) by using fast methods based on Bernstein-Bezier bases for the application of derivative and lift matrices for constant-coefficient terms \cite{chan2015bbdg}.
+Finally, the approach presented here is more generally applicable to problems involving a weighted mass matrix with a matrix-valued weight, which we will explore in the future.
 These include hyperbolic systems of PDEs with a right hand side of the form $\bm{A}_{0}(\bm{x})\pd{\bm{U}}{t}$, such as Maxwell's equations for cloaking problems \cite{li2012time}.
@@
 
-The author gratefully thanks Thomas Hagstrom, Tim Warburton, Axel Modave, Ruichao Ye, Zheng Wang, and Mario Bencomo for helpful and informative discussions.
+The author gratefully thanks Thomas Hagstrom, Tim Warburton, Axel Modave, Ruichao Ye, and Mario Bencomo for helpful and informative discussions.
 

### 2017-01-17 docs/elasticity_notes/main.tex
adding old elas files + some euler experiments

@@
 \usepackage{fullpage}
-\usepackage{hyperref}
-%\let\proof\relax
-%\let\endproof\relax
-\usepackage{amsmath,amssymb,amsfonts,mathrsfs}%,amsthm}
+\usepackage{amsmath,amssymb,amsfonts}
 \usepackage[titletoc,toc,title]{appendix}
 
-%\usepackage{lineno}
 \usepackage{array} 
-\usepackage[utf8]{inputenc}
 \usepackage{listings}
@@
 \definecolor{markercolor}{RGB}{124.9, 255, 160.65}
-\pgfplotsset{width=10cm,compat=1.9}
+\pgfplotsset{width=10cm,compat=1.3}
 \pgfplotsset{
@@
 }
-
-\usetikzlibrary{calc}
 
@@
 \lambda &       2\mu+\lambda &      0\\
-          0 &      0           &     {\mu}
+0 & 0 & \mu
 \end{array}}, 
@@
 \[
-%\bm{C}= 
-%\LRp{
-%\begin{array}{cccccc}
-%2\mu+\lambda  &  \lambda  &     \lambda  & & & \\
-%\lambda &   2\mu+\lambda & \lambda   & & & \\
-%\lambda & \lambda &   2\mu+\lambda &    & &\\
-% &   &    & {\mu} & &\\
-%&  &   &    & {\mu} &\\
-%& &  &   &    & {\mu} \\
-%\end{array}}.  
 \LRp{\begin{array}{cccc}
@@
 \begin{align}
-\sum_{D^k\in \Oh} \LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= 
-\sum_{D^k\in \Oh} \LRp{\LRp{ \sum_{i=1}^d \bm{A}_i^T\pd{\bm{\sigma}}{\bm{x}_i} + \bm{f},\bm{w}}_{\Lk}  + \LRa{\frac{1}{2}\bm{A}_n^T\jump{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk}} \nonumber\\
-\sum_{D^k\in \Oh}\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &=
-\sum_{D^k\in \Oh}\LRp{\LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk}},
+&\sum_{D^k\in \Oh} \LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} = \sum_{D^k\in \Oh} \LRp{\LRp{ \sum_{i=1}^d \bm{A}_i^T\pd{\bm{\sigma}}{\bm{x}_i} + \bm{f},\bm{w}}_{\Lk}  + \LRa{\frac{1}{2}\bm{A}_n^T\jump{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk}} \nonumber\\
+&\sum_{D^k\in \Oh} \LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} = \sum_{D^k\in \Oh}\LRp{\LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk}},
 \label{eq:dgform}
@@
 \[
-\nor{\bm{U}}_{W^{N+1,2}\LRp{\Oh}}} < \infty, \qquad \nor{\pd{\bm{U}}{t}}_{W^{N+1,2}\LRp{\Oh}} < \infty,
-\]
-where we define $\nor{\bm{U}}_{W^{N+1,2}\LRp{\Oh}}}^2 = \sum_k \nor{\bm{U}}_{W^{N+1,2}\LRp{D^k}}}^2$.
+\nor{\bm{U}}_{W^{N+1,2}\LRp{\Oh}} < \infty, \qquad \nor{\pd{\bm{U}}{t}}_{W^{N+1,2}\LRp{\Oh}} < \infty,
+\]
+where we define $\nor{\bm{U}}_{W^{N+1,2}\LRp{\Oh}}^2 = \sum_k \nor{\bm{U}}_{W^{N+1,2}\LRp{D^k}}^2$.
 
@@
 \]
-The right hand side term involving $\bm{\epsilon}$ can be bounded using interpolation estimates and a standard $hp$ trace inequality $\nor{u}_{\Ldk} \leq Ch^{-1/2} \nor{u}_{\Lk}$ \cite{warburton2003constants,chan2015gpu}
+We eliminate factors of $b(\bm{\eta},\bm{\eta})$ on both sides and bound right hand side terms.
+The trace term is bounded using a standard $hp$ trace inequality \cite{warburton2003constants} and an interpolation estimate
 \[
@@
 \]
-We apply this estimate and eliminate factors of $b(\bm{\eta},\bm{\eta})$ on both sides.  Then, integrating from $[0,T]$ and using Lemma~\ref{lemma:props} yields
+Since $\bm{A}_0$ is independent of $t$, Theorem~\ref{thm:west} gives
+\[
+\nor{\pd{\bm{\delta}}{t}}_{\L} = \nor{\bm{A}_0 \pd{\bm{U}}{t} - T_{\bm{A}_0^{-1}}^{-1}\pd{\bm{U}}{t}}_{\L} \leq C h^{N+1} {A}_{\max} \nor{\bm{A}_0}_{W^{N+1,\infty}(\Oh)}\nor{\pd{\bm{U}}{t}}_{\L}.
+\]
+where $A_{\max} = \max\LRp{\frac{\rho_{\max}}{\rho_{\min}}, c_{\max} \tilde{c}_{\max}}$.  Then, integrating from $[0,T]$ and using Lemma~\ref{lemma:props} yields 
 \begin{equation}
-A_{\min}^2 \nor{\bm{\eta}}^2_{\L} \leq \int_0^T \nor{\pd{\bm{\delta}}{t}}_{\L} \nor{\bm{\eta}}_{\L} + C h^{2N+1} \nor{\bm{U}}^2_{W^{N+1,2}(\Omega)}.
+A_{\min} \nor{\bm{\eta}}^2_{\L} \leq C\int_0^T  h^{N+1}A_{\max}\nor{\bm{A}_0}_{W^{N+1,\infty}(\Oh)}\nor{\pd{\bm{U}}{t}}_{\L} \nor{\bm{\eta}}_{\L} + h^{2N+1} \nor{\bm{U}}^2_{W^{N+1,2}(\Omega)}.
 \label{eq:intermed}
 \end{equation}
-%where $C$ depends on $\tau_v,\tau_\sigma, N$, and 
-where $A_{\min} =\min\LRp{ \rho_{\min}, \tilde{c}_{\min}}$.  Since $\bm{A}_0$ is independent of $t$, Theorem~\ref{thm:west} gives
-\[
-\nor{\pd{\bm{\delta}}{t}}_{\L} = \nor{\bm{A}_0 \pd{\bm{U}}{t} - T_{\bm{A}_0^{-1}}^{-1}\pd{\bm{U}}{t}}_{\L} \leq C h^{N+1} {A}_{\max} \nor{\bm{A}_0}_{W^{N+1,\infty}(\Oh)}\nor{\pd{\bm{U}}{t}}_{\L}.
-\]
-where $A_{\max} = \max\LRp{\frac{\rho_{\max}}{\rho_{\min}}, c_{\max} \tilde{c}_{\max}}$.  
-Using this estimate and Young's inequality on $\nor{\pd{\bm{\delta}}{t}}_{\L} \nor{\bm{\eta}}_{\L}$ gives
-\[
-\nor{\bm{\eta}}_{\L}^2 \leq \int_0^T \nor{\bm{\eta}}^2_{\L} + C\LRp{h^{2N+2} \LRp{\frac{A_{\max}}{A_{\min}}}^2\nor{\bm{A}_0}_{W^{N+1,\infty}(\Oh)}^2\nor{\pd{\bm{U}}{t}}_{\L}^2  + h^{2N+1} \nor{\bm{U}}^2_{W^{N+1,2}(\Oh)}}.
-\]
-An application of Gronwall's inequality \cite{riviere2008discontinuous} and the arithmetic-geometric mean inequality then gives 
-\[
-\nor{\bm{\eta}}_{\L} \leq C e^{T} h^{N+1/2} \sup_{t \in [0,T]}\LRp{h^{1/2} \LRp{\frac{A_{\max}}{A_{\min}}}\nor{\bm{A}_0}_{W^{N+1,\infty}(\Oh)}\nor{\pd{\bm{U}}{t}}_{\L}  +  \nor{\bm{U}}_{W^{N+1,2}(\Oh)}}.
-\]
+where $A_{\min} =\min\LRp{ \rho_{\min}, \tilde{c}_{\min}}$.
+Applying the modified Gronwall's inequality (Lemma 1.10 in \cite{dolejvsi2015discontinuous}; see also \cite{chan2016weight1}) to (\ref{eq:intermed}) then yields
+\begin{align*}
+\nor{\bm{\eta}}_{\L} &\leq \frac{1}{A_{\min}} \int_0^T CA_{\max} h^{N+1}\nor{\bm{A}_0}_{W^{N+1,\infty}(\Oh)}\nor{\pd{\bm{U}}{t}}_{\L} + \sup_{t\in [0,T]} \sqrt{\int_0^T C h^{2N+1} \nor{\bm{U}}^2_{W^{N+1,2}(\Omega)}}\\
+&\leq \frac{C T h^{N+1/2}}{A_{\min}} \sup_{t \in [0,T]}\LRp{h^{1/2} A_{\max}\nor{\bm{A}_0}_{W^{N+1,\infty}(\Oh)}\nor{\pd{\bm{U}}{t}}_{\L}  +  \nor{\bm{U}}_{W^{N+1,2}(\Oh)}}.
+\end{align*}
 The triangle inequality gives the final estimate
 \[
-\nor{\bm{U}-\bm{U}_h} \leq C\LRp{1 + e^{T}} h^{N + 1/2}\sup_{t \in [0,T]}\LRp{ h^{1/2}C_A\nor{\pd{\bm{U}}{t}}_{W^{N+1,2}(\Oh)}+\nor{\bm{U}}_{W^{N+1,2}(\Oh)}}.
-\]
-where $C_A = \LRp{\frac{A_{\max}}{A_{\min}}}\nor{\bm{A}_0}_{W^{N+1,\infty}(\Oh)}$.
+\nor{\bm{U}-\bm{U}_h} \leq \LRp{C_1 + C_2 T} h^{N + 1/2}\sup_{t \in [0,T]}\LRp{ h^{1/2}\nor{\bm{A}_0}_{W^{N+1,\infty}(\Oh)}\nor{\pd{\bm{U}}{t}}_{W^{N+1,2}(\Oh)}+\nor{\bm{U}}_{W^{N+1,2}(\Oh)}}.
+\]
+where $C_2$ depends on $A_{\min}, A_{\max}$.
 
@@
 \centering
-\subfloat[$\tau_v = \tau_\sigma = 0$]{\includegraphics[width=.475\textwidth]{../figs/spectra_tau0.png}}
+\subfloat[$\tau_v = \tau_\sigma = 0$]{\includegraphics[width=.475\textwidth]{figs/spectra_tau0.png}}
 %\hspace{1em}
-\subfloat[$\tau_v = \tau_\sigma = 1$]{\includegraphics[width=.475\textwidth]{../figs/spectra_tau1.png}}
+\subfloat[$\tau_v = \tau_\sigma = 1$]{\includegraphics[width=.475\textwidth]{figs/spectra_tau1.png}}
 \caption{Spectra for $N=4$ and $h = 1/4$, using randomly chosen $\bm{C}(\bm{x})$ with eigenvalues between $[d_{\min}, 1]$ at each quadrature point.
@@
 \centering
-\subfloat[$\LRb{\bm{\sigma}_{xx}+\bm{\sigma}_{yy}}$]{\includegraphics[width=.475\textwidth]{../figs/inclusion_aligned.png}}
+\subfloat[$\LRb{\bm{\sigma}_{xx}+\bm{\sigma}_{yy}}$]{\includegraphics[width=.475\textwidth]{figs/inclusion_aligned.png}}
 \hspace{2em}
-\subfloat[$\LRb{\bm{\sigma}_{xy}}$]{\includegraphics[width=.475\textwidth]{../figs/inclusion_aligned_shear.png}}
-%\subfloat[Aligned mesh (zoom)]{\includegraphics[height=.24\textwidth]{../figs/inclusion_aligned_zoom.png}}
+\subfloat[$\LRb{\bm{\sigma}_{xy}}$]{\includegraphics[width=.475\textwidth]{figs/inclusion_aligned_shear.png}}
+%\subfloat[Aligned mesh (zoom)]{\includegraphics[height=.24\textwidth]{figs/inclusion_aligned_zoom.png}}
 %\\
-%\subfloat[Un-aligned mesh]{\includegraphics[height=.24\textwidth]{../figs/inclusion_unaligned.png}}
+%\subfloat[Un-aligned mesh]{\includegraphics[height=.24\textwidth]{figs/inclusion_unaligned.png}}
 %\hspace{2em}
-%\subfloat[Un-aligned mesh (zoom)]{\includegraphics[height=.24\textwidth]{../figs/inclusion_unaligned_zoom.png}}
+%\subfloat[Un-aligned mesh (zoom)]{\includegraphics[height=.24\textwidth]{figs/inclusion_unaligned_zoom.png}}
 \caption{Values of different solution fields at $T=.4$ for the stiff inclusion problem of Leveque \cite{leveque2002finite} with order of approximation $N=5$.}
@@
 %\subfloat[$\bm{v}_2$]{\includegraphics[width=.45\textwidth]{}}
-\subfloat[$T=30 \mu s$ (zoomed in)]{\includegraphics[height=18em]{../figs/aniso1.png}}
+\subfloat[$T=30 \mu s$ (zoomed in)]{\includegraphics[height=18em]{figs/aniso1.png}}
 \hspace{3em}
-\subfloat[$T=60 \mu s$]{\includegraphics[height=18.25em]{../figs/aniso2.png}}
+\subfloat[$T=60 \mu s$]{\includegraphics[height=18.25em]{figs/aniso2.png}}
 \caption{An example of wave propagation in heterogeneous anisotropic media.  The vertical component of the velocity $\bm{v}_2$ is shown at $T = 30$ and $T=60$ microseconds. }
@@
 \centering
-\subfloat[Computational mesh]{\includegraphics[width=.28\textwidth]{../figs/cubeSplitFine.png}\label{subfig:mesh}}
+\subfloat[Computational mesh]{\includegraphics[width=.28\textwidth]{figs/cubeSplitFine.png}\label{subfig:mesh}}
 \hspace{.5em}
-\subfloat[$w(\bm{x}) = 0$]{\includegraphics[width=.35\textwidth]{../figs/pplane0.png}}
+\subfloat[$w(\bm{x}) = 0$]{\includegraphics[width=.35\textwidth]{figs/pplane0.png}}
 %\hspace{.25em}
-\subfloat[$w(\bm{x})  = .5 \cos(3\pi x)\cos(3\pi y)\cos(3\pi z) $]{\includegraphics[width=.35\textwidth]{../figs/pplanew0.png}}
+\subfloat[$w(\bm{x})  = .5 \cos(3\pi x)\cos(3\pi y)\cos(3\pi z) $]{\includegraphics[width=.35\textwidth]{figs/pplanew0.png}}
 \caption{Mesh and $xz$, $xy$ slices of $\bm{v}_1$ at $T = .5$.  The order of approximation is taken to be $N=5$.  }
@@
 The cost of the latter step can be reduced (especially at high orders of approximation) by using fast methods based on Bernstein-Bezier bases for the application of derivative and lift matrices for constant-coefficient terms \cite{chan2015bbdg}.
-Finally, the approach presented here is more generally applicable to problems involving a weighted mass matrix with a matrix-valued weight, which we will explore in the future.
-These include hyperbolic systems of PDEs with a right hand side of the form $\bm{A}_{0}(\bm{x})\pd{\bm{U}}{t}$, such as Maxwell's equations for cloaking problems \cite{li2012time}.
+%Finally, the approach presented here is more generally applicable to problems involving a weighted mass matrix with a matrix-valued weight, including hyperbolic systems of PDEs with a left hand side of the form $\bm{A}_{0}(\bm{x})\pd{\bm{U}}{t}$, such as Maxwell's equations for cloaking problems \cite{li2012time}.
 

### 2017-01-17 matlab/ElasticityPtsrc2D.m

@@ -1,21 +1,20 @@
-function ElasticityPulse2D
+function ElasticityAniso2D_sem
 
 % clear all, clear
 clear -global *
 
 Globals2D
 
-K1D = 16;
-N = 4;
+K1D = 32;
+N = 5;
 c_flag = 0;
-FinalTime = 2; % milliseconds
+FinalTime = .5; % microseconds
 
 filename = 'Grid/Other/block2.neu';
 % [Nv, VX, VY, K, EToV] = MeshReaderGambit2D(filename);
 [Nv, VX, VY, K, EToV] = unif_tri_mesh(K1D,K1D);
-% VX = .325*VX; VY = .325*VY;
-% VX = .17*VX;
-% VY = .17*VY;
+% VX = .16*VX; VY = .16*VY;
+% VX = .3*VX; VY = .3*VY;
 
 % VX = (VX+1)/2; 
 % VY = (VY+1)/2;
@@ -68,25 +67,37 @@ mapBL = fbids;
 vmapBL = vmapP(mapBL);
 
 %%
-global Nfld mu lambda rho Vq Pq tau useWADG
+global Nfld rho Vq Pq tau useWADG
 Nfld = 5; %(u1,u2,sxx,syy,sxy)
 
-% rho = ones(size(xq))*sqrt(7100);
-rho = ones(size(xq)); % convert to milliseconds
+% rho = ones(size(xq));
+% keyboard
+rho = ones(size(xq)); % convert to microsec
 
 global C11 C12 C13 C22 C23 C33
 
 useWADG = 0;
 
-mu = ones(size(xq));
-lambda = 2*ones(size(xq));
+% isotropic media
+% mu = 3.96; lambda = 8.58;
+mu = 1; lambda = 2;
+C11b = 2*mu + lambda ;
+C12b = lambda ;
+C22b = 2*mu + lambda ;
+C33b = mu ;
 
-C11 = 2*mu+lambda;
-C12 = lambda;
-C13 = 0*ones(size(xq));
-C22 = 2*mu+lambda;
-C23 = 0*ones(size(xq));
-C33 = mu;
+% anisotropic
+C11a = C11b ;
+C12a = 0.5 * C12b ;
+C22a = .33 * C22b ;
+C33a = C33b;
+
+% keyboard
+
+C11 = C11a*(xq < 0) + C11b*(xq > 0);
+C12 = C12a*(xq < 0) + C12b*(xq > 0);
[…]

### 2017-01-17 matlab/ElasticityPulse2D.m

@@ -6,9 +6,9 @@ clear -global *
 Globals2D
 
 K1D = 16;
-N = 4;
+N = 5;
 c_flag = 0;
-FinalTime = 4;
+FinalTime = 1;
 
 filename = 'Grid/Other/block2.neu';
 % [Nv, VX, VY, K, EToV] = MeshReaderGambit2D(filename);
@@ -52,7 +52,7 @@ global mapBL vmapBL t0
 
 t0 = .0;
 % t0 = .075;
-t0 = .05;
+t0 = .1;
 
 % find x = 0 faces 
 fbids = reshape(find(vmapP==vmapM),Nfp,nnz(vmapP==vmapM)/Nfp);
@@ -77,19 +77,20 @@ lambda0 = lambda(1);
 
 % C = [2*mu0+lambda0       lambda0       0
 %      lambda0       2*mu0+lambda0       0
-%           0       0                mu0/2];
+%           0       0                mu0];
 
 % mu = 1 + .5*sin(2*pi*x).*sin(2*pi*y);
 % lambda = 1 + .5*sin(2*pi*x).*sin(2*pi*y);
 
 % ids =  mean(y) > .375 & mean(y) < .625 & mean(x) > .5 & mean(x) < .75;
-% mu(:,ids) = 10*mu(:,ids);
-% lambda(:,ids) = 10*lambda(:,ids);
+ids = mean(y) < .5;
+mu(:,ids) = 0*mu(:,ids);
+% lambda(:,ids) = .5*lambda(:,ids);
 
 
 global C11 C12 C13 C22 C23 C33
 
-useWADG = 1;
+useWADG = 0;
 if useWADG
     mu = Vq*mu;
     lambda = Vq*lambda;
@@ -116,13 +117,13 @@ end
 %% params setup
 
 x0 = .5;
-y0 = .5;
-p = exp(-50^2*((x-x0).^2 + (y-y0).^2));
+y0 = .65;
+p = exp(-25^2*((x-x0).^2 + (y-y0).^2));
 % p = exp(-50^2*((x-x0).^2));
 u = zeros(Np, K);
 
 U{1} = u;
-U{2} = u;
+U{2} = p;
 U{3} = u;
 U{4} = u;
 U{5} = u;
[…]

### 2017-01-17 matlab/Euler1D.m

@@ -7,8 +7,8 @@ function Euler1D
 Globals1D;
 
 % Order of polymomials used for approximation
-N = 7;
-K1D = 8;
+N = 2;
+K1D = 32;
 FinalTime = 25;
 
 % Generate simple mesh
@@ -33,7 +33,7 @@ VBe = bern_basis_1D(N,re);
 xe = Vandermonde1D(N,re)/V * x;
 
 W = get_BB_smoother(N);
-% W = VBe;
+W = VBe;
 W = VB*W/VB;
 
 % Set initial conditions
@@ -63,7 +63,7 @@ resE = zeros(Np,K);
 % compute time step size
 xmin = min(abs(x(1,:)-x(2,:)));
 
-dt   = .1*xmin;
+dt   = .2*xmin;
 Nsteps = ceil(FinalTime/dt); dt = FinalTime/Nsteps;
 
 M = inv(V*V');
@@ -125,6 +125,12 @@ rho(:,ids) = W*rho(:,ids);
 m(:,ids)   = W*m(:,ids);
 E(:,ids)   = W*E(:,ids);
 
+function TV = TV1D(N,u)
+
+TV = 0;
+for i = 1:N
+    TV = TV + abs(u(i,:) - u(i+1,:));
+end
 
 function [rho m E] = vortexSolution(x,t)
 
@@ -144,15 +150,18 @@ x0 = 5;
 % m = rho.*u;
 % E = p/(gamma-1) + .5*rho.*u.^2;
 
-
 rho = 1+exp(-(x-x0).^2); 
 % rho = 1 + (x > x0);
 p = rho.^gamma;
-
-% rho = 1+.2*sin(pi*x); p = ones(size(rho)); 
 u = 0*rho;
 E = p/(gamma-1) + .5*rho.*u.^2;
 m = rho.*u;
+
+% % steady sine solution
+% rho = 1+.2*sin(pi*x); p = ones(size(rho)); 
+% u = 0*rho;
+% E = p/(gamma-1) + .5*rho.*u.^2;
+% m = rho.*u;
 
 % % sine solution
 % rho = 2 + sin(pi*x);
@@ -160,6 +169,7 @@ m = rho.*u;
 % p = ones(size(x));
 % m = rho;
 % E = p/(gamma-1) + .5*rho.*u.^2;
+
 
 
 function [rhsrho, rhsm, rhsE] = EulerRHS1D(rho, m ,E)
[…]

### 2017-01-17 matlab/Wave2D_PML.m

@@ -7,7 +7,7 @@ Globals2D
 
 if nargin==0
     %     Nin = 4; K1D = 16;
-    Nin = 3; K1D = 8; c_flag = 0;
+    Nin = 4; K1D = 16; c_flag = 0;
     cfun = @(x,y) ones(size(x));
     % cfun = @(x,y) 1 + .5*sin(pi*x).*sin(pi*y); % smooth velocity
 %     cfun = @(x,y) (1 + .5*sin(2*pi*x).*sin(2*pi*y) + (y > 0)); % piecewise smooth velocity
@@ -18,7 +18,7 @@ N = Nin;
 
 StartUp2D;
 
-[rp sp] = EquiNodes2D(25); [rp sp] = xytors(rp,sp);
+[rp sp] = EquiNodes2D(50); [rp sp] = xytors(rp,sp);
 Vp = Vandermonde2D(N,rp,sp)/V;
 xp = Vp*x; yp = Vp*y;
 
@@ -52,37 +52,27 @@ u = zeros(Np, K); v = zeros(Np, K);
 
 showSol = 1;
 delta = .5;
-smax = 100;
+smax = 300;
 
 M = 3;
 L = 1-delta;
 sigmax = (x > L).*(x-L).^M - (x < -L).*(x+L).^M;
 sigmay = (y > L).*(y-L).^M - (y < -L).*(y+L).^M;
 
-Q = zeros(Np,K);
-R = zeros(Np,K);
-resQ = zeros(Np,K); 
-resR = zeros(Np,K);
+Q = zeros(Np,K); R = zeros(Np,K);
+resQ = zeros(Np,K); resR = zeros(Np,K);
 
 P1 = eye(Np);
 for e = 1:K
     %     sigmax(:,e) = max(sigmax(:,e));
     %     sigmay(:,e) = max(sigmay(:,e));
-    sigmax(:,e) = mean(sigmax(:,e));
-    sigmay(:,e) = mean(sigmay(:,e));
+%     sigmax(:,e) = mean(sigmax(:,e));
+%     sigmay(:,e) = mean(sigmay(:,e));
 end
 % P1 = V*diag([1;0;0;zeros(Np-3,1)])*inv(V); % reduce to const or linear modes
-% 
-sk = 1;
-for e = 1:K
-    if (max(sigmax(:,e)) > 1e-8 || max(sigmay(:,e))> 1e-8)
-        pmlK(sk) = e;
-        sk = sk + 1;
-    end
-end
 
 
[…]

### 2017-02-15 docs/IGA-DG/main.tex
adding notes on spline DG

--- 2016-12-21 docs/elasticity_notes/main.tex
@@
 \date{}
-\author{Jesse Chan}
-\title{Weight-adjusted discontinuous Galerkin methods: matrix-valued weights and elastic wave propagation in heterogeneous media}
+\author{Jesse Chan, John Evans}
+\title{B-spline DG notes}
 
@@
 \begin{abstract}
-Weight-adjusted inner products \cite{chan2016weight1,chan2016weight2} replace weighted $L^2$ inner products with easily invertible weight-adjusted approximations.
-When paired with discontinuous Galerkin (DG) methods, the result is a low storage, energy stable, and (for sufficiently regular weights) high order accurate method for wave propagation in the presence of general heterogeneous media and curvilinear meshes.
-In this work, we extend weight-adjusted DG (WADG) methods to weighted inner products with matrix-valued weights, focusing on the linear elastic wave equation as an application.
-We present a unified DG formulation for isotropic and anisotropic media based on the discretization of the symmetric form of the linear elasticity wave equation, with upwind-like dissipation incorporated through simple penalty terms in the numerical fluxes.
-This formulation is provably stable for arbitrary stiffness tensors and curvilinear meshes.
-A convergence analysis is given, and numerical results confirm the stability and high order accuracy of WADG for several problems in elastic wave propagation.
+Achieve extra $h$-resolution for the same CFL condition by using high order B-splines as a local approximation space on tensor product elements.
+
+Achieve a spectral element-type implementation using quadrature nodes for B-splines. 
 \end{abstract}
 
-%\tableofcontents
+\tableofcontents
 
@@
 
-Efficient and accurate simulations of elastic wave propagation form the foundation for various seismic applications, including imaging and inversion techniques, as well as rupture and earthquake simulations.
-The most common methods for the simulation of waves are finite difference methods \cite{virieux1986p}; however, finite element methods have garnered interest due to their low numerical dispersion and ability to accomodate geometrically flexible unstructured meshes.
+\section{Stable timestep restrictions for high order DG methods}
 
-Typical methods for time-domain wave propagation utilize explicit time stepping methods, since the hyperbolic partial differential equations (PDEs) governing wave propagation typically admit less restrictive stable timestep restrictions than parabolic equations.
-However, finite element methods require the inversion of a global mass matrix when paired with explicit time integrators.
-On hexahedral meshes, this can be avoided by using spectral element methods (SEM), which utilize mass lumping to produce a diagonal mass matrix \cite{komatitsch1998spectral}.
-This global inversion can also be avoided through the use of discontinuous Galerkin (DG) methods, which result in a locally invertible block diagonal mass matrix.
-Due to difficulties in extending mass-lumping techniques to tetrahedra, high order DG methods are often employed for seismic simulations which use simplicial meshes \cite{kaser2006arbitrary, dumbser2006arbitrary, de2007arbitrary,delcourte2009high,delcourte2015analysis,ye2016discontinuous}.
-Such methods also lend themselves well to efficient implementations using Graphics Processing Units (GPUs) \cite{klockner2009nodal, modave2015nodal, modave2016gpu, chan2015gpu}.
+We illustrate this using the advection equation.  
+\begin{align*}
+V^T A_h U &= \sum_{k} \LRp{\beta\cdot\Grad u,v}_{\Lk} + \LRa{f^*(u),v}_{\Ldk} \\
+&\leq \nor{\beta}_{L^\infty}\sum_{k} \nor{\Grad u}\nor{v} + \nor{f^*(u)}_{\Ldk}\nor{v}_{\Ldk}\\
+&\lesssim \nor{\beta}_{L^\infty}\sum_{k} \LRp{C_I \nor{J\bm{G}}_{L^\infty\LRp{\Dhat}}  + C_N^2 \nor{J^f}_{L^\infty\LRp{\Dhat}}}\nor{1/J}_{L^\infty\LRp{\Dhat}}\nor{u}_{\Lk}\nor{v}_{\Lk}
+\end{align*}
+The CFL condition is then determined by mesh-dependent geometric factors of $O(h)$ and order-dependent constants in the inverse and trace inequalities $C_I, C_N$.
 
-Most high order DG methods on simplicial meshes assume that material coefficients are constant over each element, which can result in spurious reflections and the loss of high order accuracy \cite{castro2010seismic,mercerat2015nodal,bencomo2015discontinuous}.
-This limitation can be overcome for DG methods by incorporating sub-element variation in media through weighted mass matrices, which results in an energy stable method \cite{mercerat2015nodal, bencomo2015discontinuous}.
-For DG methods on hexahedral meshes, mass lumping reduces the construction of weighted mass matrices to entry-wise scalings of a diagonal mass matrix.
-However, on tetrahedral meshes, this approach necessitates the pre-computation and storage of factorizations for each local mass matrix, which greatly increases both storage costs and data transferred at high orders of approximation.
-These costs are especially problematic for accelerator architectures such as GPUs, which possess limited memory.
-%Additionally, excessive storage costs can require reduced problem sizes and under-utilization of accelerator cards.
+\subsection{Explicit constants in trace and inverse inequalities for polynomial and B-spline approximations}
 
-Storage costs associated with weighted mass matrices can be avoided by approximating weighted $L^2$ inner products using weight-adjusted inner products, which result in easily invertible approximations to weighted mass matrices \cite{chan2016weight1,chan2016weight2}.
-For sufficiently regular weights, high order accuracy is also retained.
-When paired with an energy stable DG formulation, weight-adjusted inner products result in weight-adjusted DG methods (WADG), which preserve energy stability and high order accuracy while retaining a low asymptotic storage cost.
-Additionally, unlike mass-lumping techniques, WADG does not rely on the use of inexact quadrature rules, and reduces to the exact inversion of mass matrices for constant weights.
+For polynomials, $C_I, C_N = O(N^2)$ on the reference element (see, for example \cite{warburton2003constants, ozisik2010constants, chan2015gpu}).
 
-Weight-adjusted DG methods have been applied to both acoustic wave propagation in heterogeneous media and on curvilinear meshes \cite{chan2016weight1,chan2016weight2}.
-Both cases involve scalar weighting functions.
-In this work, we extend weight-adjusted inner products to matrix-valued weights, and demonstrate the usefulness of this generalization to elastic wave propagation in arbitrary media.
-A new energy stable DG formulation is introduced based on the symmetric form of the elastic wave equations, with upwind-like numerical dissipation introduced through simple penalty fluxes \cite{warburton2013low,ye2016discontinuous}.
+For B-splines with $K_{\rm sub} = N$, $C_N = O(N)$ instead of $O(N^2)$.
 
-This work proceeds as follows.
-Sections~\ref{sec:symelas} and \ref{sec:dgform} begin by presenting an energy stable DG formulation with simple penalty fluxes for the elastic wave equation in symmetric hyperbolic form, and discusses issues related to storage and inversion of local mass matrices for material coefficients with sub-element variations.
-Section~\ref{sec:mwadg} extends weight-adjusted approximations to weighted $L^2$ inner products and mass matrices to the case of matrix-valued weights, and proves interpolation estimates which account for the regularity of the matrix weight.
-These results are incorporated into a weight-adjusted DG method for the linear elastic wave equations in Section~\ref{sec:wadgelas}.
-Finally, numerical results in Section~\ref{sec:numerical} demonstrate the accuracy of this method for several analytic solutions in linear elasticity.
+For sufficiently regular solutions, convergence is in $L^2$ as $C (h/N)^{N+1}$ while still maintaining an $O(h/N^2)$ CFL.
+Or, if we measure by $h_N = h/N$, CFL is $O(h_N/N)$.
+This results from interpolation estimates for one-dimensional B-splines \cite{bazilevs2006isogeometric}
 
-\section{Symmetric form of the elastic wave equation }
-\label{sec:symelas}
-We begin with the linear elastic wave equation under small deformations in a domain $\Omega \in \mathbb{R}^d$.
-These equations can be written a first order velocity-stress system for velocity $\bm{v}$ and symmetric stress tensor $\bm{S}$
-\begin{align*}
-\rho \pd{\bm{v}}{t} &= \Div{\bm{S}} + \bm{f}\\
-\pd{\bm{S}}{t} &= \frac{1}{2} \tensor{\bm{C}}\LRp{\Grad\bm{v} + \Grad\bm{v}^T},
-\end{align*}
-where $\bm{f}$ is the body force per unit volume, $\rho$ is density, and $\tensor{\bm{C}}$ is the symmetric constitutive stiffness tensor relating stress and strain.
-We rewrite these equations as a symmetric hyperbolic system of PDEs \cite{hughes1978classical}
-\begin{align}
-\rho \pd{\bm{v}}{t} &= \sum_{i=1}^d \bm{A}_i^T \pd{\bm{\sigma}}{\bm{x}_i}\nonumber + \bm{f}\\
-\bm{C}^{-1} \pd{\bm{\sigma}}{t} &= \sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i}.
-\label{eq:symelas}
-\end{align}
-where $\bm{C}$ is the symmetric matrix form of the constitutive tensor $\tensor{\bm{C}}$ and $\sigma$ is a vector of length $N_d = \frac{d(d+1)}{2}$, the number of unique entries of the stress tensor $\bm{S}$ in $d$ dimensions.
-We note that the matrices $\bm{A}_i$ are constant, while $\rho$ and $\bm{C}$ can vary spatially.
-Furthermore, we will assume that $\rho$ and $\bm{C}$ are positive-definite and bounded pointwise such that
-\begin{align*}
-0 < \rho_{\min} &\leq \rho(\bm{x}) \leq \rho_{\max} < \infty, \qquad \forall \bm{x} \in \mathbb{R}^d,\\
-0 < c_{\min} \leq \bm{u}^T\bm{C}(\bm{x})\bm{u} &\leq c_{\max} < \infty, \qquad \forall \bm{u} \in \mathbb{R}^{N_d}, \quad \forall \bm{x} \in \mathbb{R}^d,
-\end{align*}
-which also implies that $\bm{C}^{-1}$ is pointwise positive-definite and bounded.
+\section{Quadrature }
 
-In two dimensions, $\bm{v} = (\bm{v}_1, \bm{v}_2)^T$ and $\bm{\sigma}= (\sigma_{xx},\sigma_{yy},\sigma_{xy})^T$ 
-\[
-\bm{S} = \LRp{\begin{array}{cc}
-\sigma_{xx} & \sigma_{xy}\\
-\sigma_{xy} & \sigma_{yy}
-\end{array}},
-\]
-while the matrices $\bm{A}_i$ are 
-\[
-\bm{A}_1 = \left(\begin{array}{cc}
-1 & 0\\
-0 & 0\\
-0 & 1
-\end{array}\right), \qquad 
-\bm{A}_2 = \left(\begin{array}{cc}
-0 & 0\\
-0 & 1\\
-1 & 0
-\end{array}\right).
-\]
-In three dimensions, the velocity is $\bm{v} = (\bm{v}_1,\bm{v}_2,\bm{v}_3)^T$, while $\bm{\sigma}= (\sigma_{xx},\sigma_{yy},\sigma_{zz},\sigma_{yz},\sigma_{xy},\sigma_{xz})^T$ denotes the unique entries of the stress tensor $\bm{S}$
-\[
-\bm{S} = \LRp{\begin{array}{ccc}
-\sigma_{xx} & \sigma_{xy} & \sigma_{xz}\\
-\sigma_{xy} & \sigma_{yy}& \sigma_{yz}\\
-\sigma_{xz} & \sigma_{yz}& \sigma_{zz}
-\end{array}}.
-\]
-The matrices $\bm{A}_i$ are then
-\[
-\bm{A}_1 = \left(\begin{array}{ccc}
-1 & 0 & 0\\
-0 & 0& 0\\
-0 & 0& 0\\
-0 & 0& 0\\
-0 & 0 & 1\\
-0 & 1 & 0
-\end{array}\right), \qquad 
-\bm{A}_2 = \left(\begin{array}{ccc}
-0 & 0 & 0\\
-0 & 1 & 0\\
-0 & 0 & 0\\
-0 & 0 & 1\\
-0 & 0 & 0\\
-1 & 0 & 0
-\end{array}\right), \qquad
-\bm{A}_3 = \left(\begin{array}{ccc}
-0 & 0 & 0\\
-0 & 0 & 0\\
-0 & 0 & 1\\
-0 & 1 & 0\\
-1 & 0 & 0\\
-0 & 0 & 0
-\end{array}\right)
-\]
+Reduce dimension of the search space by treating the weights as dependent variables and assuming symmetry of the rule.
+Initialize guess for quadrature rule using the previous order's rule to generate a polynomial interpolant from equispaced nodes to quadrature nodes \cite{warburton2006explicit,chan2015comparison}.
 
-In general anisotropic media, $\bm{C}$ is symmetric and positive-definite.  For two-dimensional isotropic media, $\bm{C}$ is given as
-\[
-\bm{C}= \LRp{
-\begin{array}{ccc}
-2\mu+\lambda  &  \lambda  &     0\\
-\lambda &       2\mu+\lambda &      0\\
-          0 &      0           &     {\mu}
-\end{array}}, 
-\]
-where $\lambda,\mu$ are Lame parameters.  For three-dimensional isotropic media, $\bm{C}$ is given instead by
-\[
-%\bm{C}= 
-%\LRp{
-%\begin{array}{cccccc}
-%2\mu+\lambda  &  \lambda  &     \lambda  & & & \\
-%\lambda &   2\mu+\lambda & \lambda   & & & \\
-%\lambda & \lambda &   2\mu+\lambda &    & &\\
-% &   &    & {\mu} & &\\
-%&  &   &    & {\mu} &\\
-%& &  &   &    & {\mu} \\
-%\end{array}}.  
-\LRp{\begin{array}{cccc}
-2\mu+\lambda  &  \lambda  &     \lambda   & \\
-\lambda &   2\mu+\lambda & \lambda   &  \\
-\lambda & \lambda &   2\mu+\lambda &    \\
- &   &    & {\mu}\bm{I}^{3\times3} \\
-\end{array}}.  
-\]
-We will consider both spatially varying isotropic and anisotropic media in this work.
+We note that these rules are generated by choosing the number of points $N_q$ such that ${\rm dim}(S_{\rm target}) \leq 2N_q$, and may not be optimal.
 
-\section{An energy stable discontinuous Galerkin formulation for elastic wave propagation}
-\label{sec:dgform}
-The advantage of the symmetric first order formulation of the elastic wave equations (\ref{eq:symelas}) is that it is straightforward to derive an energy stable discontinuous Galerkin formulation.
-We assume that the domain $\Omega$ is Lipschitz and exactly triangulated by a mesh $\Omega_h$, which consists of elements $D^k$.
-We further assume that each element $D^k$ is the image of a reference element $\widehat{D}$ under the local elemental mapping
-\[
-\bm{x}^k = \bm{\Phi}^k \widehat{\bm{x}},
-\]
-where $\bm{x}^k = \LRc{x^k,y^k,z^k}$ denote physical coordinates on $D^k$ and $\widehat{\bm{x}} = \LRc{\widehat{x},\widehat{y},\widehat{z}}$ denote coordinates on the reference element.
-We denote the determinant of the Jacobian of $\bm{\Phi}^k$ as $J$, and refer to it as the Jacobian for the remainder of this work.
+\subsection{Reduced quadrature using Gauss-Legendre-Lobatto rules}
 
-We will approximate components over each element $D^k$ from an approximation space $V_h\LRp{D^k}$, which we define as the composition of the mapping $\bm{\Phi}^k$ and a reference approximation space $V_h\LRp{\widehat{D}}$
-\[
-V_h\LRp{D^k} = \bm{\Phi}^k \circ V_h\LRp{\widehat{D}}.
-\]
-For the remainder of this work, we will take $V_h\LRp{\widehat{D}} = P^N\LRp{\widehat{D}}$, where $P^N\LRp{\widehat{D}}$ is the polynomial space of total degree $N$ on the reference simplex.
-In two dimensions,
-\[
-P^N\LRp{\widehat{D}} = \LRc{ \widehat{x}^i \widehat{y}^j, \quad 0 \leq i + j \leq N},
-\] 
-and in three dimensions, 
-\[
-P^N\LRp{\widehat{D}} = \LRc{ \widehat{x}^i \widehat{y}^j \widehat{z}^k, \quad 0 \leq i + j +k \leq N}.  
-\]
-We denote the $L^2$ inner product and norm over $D^k$ by $\LRp{\cdot,\cdot}_{\Lk}$, such that
-\[
-\LRp{\bm{f},\bm{g}}_{\Lk} = \int_{D^k} \bm{f}\cdot\bm{g} \diff x = \int_{\widehat{D}} \bm{f}\cdot\bm{g} J, \qquad \nor{\bm{f}}_{\Lk}^2 = \LRp{\bm{f},\bm{f}}_{\Lk},
-\]
-where $\bm{f},\bm{g}$ are vector-valued functions.  Global $L^2$ inner products and norms are using local $L^2$ inner products and norms
-\[
-\LRp{\bm{f},\bm{g}}_{\Oh} = \sum_{D^k\in \Oh} \int_{D^k} \LRp{\bm{f},\bm{g}}_{\Lk}, \qquad \nor{\bm{f}}_{\Oh}^2 = \sum_{D^k\in \Oh} \nor{\bm{f}}^2_{\Lk}.
-\]
-We define also the $L^2$ inner product and norm over the boundary $\partial D^k$ of an element
-\[
-\LRa{u,v}_{\Ldk} = \int_{\partial D^k}uv\diff{x} = \sum_{f\in \partial D^k}\int_{\widehat{f}}uvJ^f\diff{\widehat{x}}, \qquad \nor{u}_{\Ldk}^2 = \LRa{u,u},
-\]
-where $J^f$ is the Jacobian of the mapping from a reference face (triangle) $\widehat{f}$ to a physical face $f$ of an element.
+Can cast into collocation style using skew-symmetric formulation by using a degree $N + K_{\rm sub} - 1$ Gauss-Legendre or Gauss-Lobatto rule.
+Show plots of GLL nodal bases for polynomials vs for B-spline bases - less oscillation for basis functions at nodes on/near the boundary.
 
-Let $f$ be a face of an element $D^{k}$ with neighboring element $D^{k,+}$ and unit outward normal $\bm{n}$.
-We define $u^-$ and $u^+$ denote interior and exterior values of a discontinuous function $u$ such that
-\[
-{u}^- = \LRl{u}_{f \cap D^k}, \qquad {u}^+ = \LRl{u}_{f \cap  \partial D^{k,+}}.  
-\]
-The jump and average of a scalar function $u\in V_h\LRp{\Omega_h}$ over $f$ are then defined as
-\[
-\jump{u} = u^+ - u^-, \qquad \avg{u} = \frac{u^+ + u^-}{2}.
-\]
-Jumps and averages of a vector-valued functions $\bm{u}\in \mathbb{R}^m$ and $\bm{S}\in \mathbb{R}^{m\times n}$ are then defined component-wise 
-\[
-\LRp{\jump{\bm{u}}}_i = \jump{\bm{u}_i}, \qquad 1\leq i \leq m, \qquad \LRp{\jump{\bm{S}}}_{ij} = \jump{\bm{S}_i}, \qquad 1\leq i \leq m, \quad 1\leq j \leq n.
-\]
+Problems here - high order accuracy is preserved if the mass matrix is under-integrated, but not if an isoparametric mapping is used.
 
-We can now specify a unified DG formulation for the linear elastic wave equation (\ref{eq:symelas}) in either isotropic or anisotropic media.
-Symmetric hyperbolic systems readily admit a DG formulation based on penalty fluxes \cite{chan2016short}.
-For the linear elastic wave equation in symmetric first order form, this formulation is given as
-\begin{align}
-\sum_{D^k\in \Oh} \LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= 
-\sum_{D^k\in \Oh} \LRp{\LRp{ \sum_{i=1}^d \bm{A}_i^T\pd{\bm{\sigma}}{\bm{x}_i} + \bm{f},\bm{w}}_{\Lk}  + \LRa{\frac{1}{2}\bm{A}_n^T\jump{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk}} \nonumber\\
-\sum_{D^k\in \Oh}\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &=
-\sum_{D^k\in \Oh}\LRp{\LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk}},
-\label{eq:dgform}
-\end{align}
-for all $\bm{w},\bm{q}\in V_h\LRp{\Omega_h}$.
-Here, $\tau_v,\tau_\sigma \geq 0$ are penalty parameters and $\bm{A}_n$ is the normal matrix defined on a face $f$ as $\bm{A}_n = \sum_{i=1}^d \bm{n}_i \bm{A}_i$.
-In two dimensions, $\bm{A}_n$ is
-\[
-\bm{A}_n =  \LRp{\begin{array}{cc}
-{n}_x & 0\\
- 0 & {n}_y\\
-{n}_y & {n}_x
-\end{array}}.
-\]
-while in three dimensions, $\bm{A}_n$ is
-\[
-\bm{A}_n =  \LRp{\begin{array}{ccc}
-{n}_x & 0 & 0\\
- 0 & {n}_y & 0\\
-0&  0 & {n}_z \\
-0 &{n}_z & {n}_y\\
-{n}_z & 0 & {n}_x\\
-{n}_y & {n}_z & 0\\
-\end{array}}.  
-\]
+\section{Future work}
 
-\subsection{Boundary conditions}
-
-In this work, we assume boundary conditions on velocity and traction of the form
-\[
-\bm{v} = \bm{v}_{\rm bc}, \qquad \bm{S}\bm{n} = \bm{t}_{\rm bc}
-\]
-where $\bm{v}_{\rm bc}$ and $\bm{t}_{\rm bc}$ are given values.
-Traction boundary conditions where $\bm{t}_{\rm bc} = 0$ are referred to as free-surface boundary conditions.
-We follow \cite{leveque2002finite,wilcox2010high} and impose boundary conditions on the DG formulation through exterior values and jumps of the solution.
-Boundary conditions on the normal component of the stress can be imposed by noting that the numerical flux contains the term $\jump{\bm{A}_n^T\bm{\sigma}} = \bm{S}\bm{n}$.
-
-For a face which lies on a boundary, velocity boundary conditions are imposed by setting 
-\begin{align*}
-\jump{\bm{v}} &= 2\LRp{\bm{v}_{\rm bc} - \bm{v}^- }, \qquad \jump{\bm{A}_n^T\bm{\sigma}}=\jump{\bm{S}\bm{n}} = 0, 
-\end{align*}
-while traction boundary conditions are imposed by
-\begin{align*}
-\jump{\bm{A}_n^T\bm{\sigma}} = \jump{\bm{S}\bm{n}} = 2\LRp{\bm{t}_{\rm bc} - \bm{S}^-\bm{n} }= 2\LRp{\bm{t}_{\rm bc} - \bm{A}_n^T\bm{\sigma}^- }, \qquad \jump{\bm{v}} = 0.
-\end{align*}
-For problems which involve the truncation of infinite or large domains, absorbing boundary conditions are required.
-For such cases, we impose simple non-reflective boundary conditions \cite{leveque2002finite} through jumps
-\begin{align*}
-\jump{\bm{A}_n^T\bm{\sigma}} = \jump{\bm{S}\bm{n}} = - \bm{S}^-\bm{n} = -\bm{A}_n^T\bm{\sigma}^- , \qquad \jump{\bm{v}} = - \bm{v}^-. 
-\end{align*}
-We note that more accurate absorbing conditions can be imposed using, for example, perfectly matched layers \cite{berenger1994perfectly} or high order absorbing boundary conditions \cite{hagstrom2004new,modave2016high}.
-
-In all cases, boundary conditions are imposed by computing numerical fluxes using these modified jumps.  This imposition guarantees energy stability for all three sets of boundary conditions.
-
-%Easiest way to impose BCs: set $U^+$ values, compute central fluxes and then compute penalty fluxes via $\bm{A}_n^T\bm{A}_n \jump{U} = \bm{A}_n \jump{F}$.
-
-\subsection{Energy stability}
-
-It is straightforward to show that the formulation (\ref{eq:dgform}) is energy stable for zero velocity and traction boundary conditions, as well as non-reflective boundary conditions.
-Integrating by parts the velocity equations gives
-\begin{align*}
-\LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= -\LRp{ \sum_{i=1}^d \bm{\sigma},\bm{A}_i \pd{\bm{w}}{\bm{x}_i}}_{\Lk}  + \LRa{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk} \nonumber\\
-\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk}
-\end{align*}
-Taking $(\bm{w},\bm{q}) = (\bm{v},\bm{\sigma})$ and adding both equations together yields
-\begin{align*}
-\sum_{D^k\in \Oh}\frac{1}{2}\pd{}{t}&\LRp{\LRp{\rho \bm{v},\bm{v}}_{\Lk} + \LRp{\bm{C}^{-1} \bm{\sigma},\bm{\sigma}}_{\Lk}} =  \\
-&\sum_{D^k\in \Omega_h}\LRa{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{v}}_{\partial D^k}+ \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{\sigma}}_{\partial D^k}  \\
-&\sum_{D^k\in \Omega_h}\sum_{f\in \partial D^k}\int_{f}{\bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v}} + {\frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma}}.
-\end{align*}
-Let $\Gamma_h$ denote the set of unique faces in $\Oh$, and let $\Gamma_v,\Gamma_\sigma,\Gamma_{\rm abc}$ denote the parts of the boundary where velocity, traction, and non-reflective boundary conditions are imposed, respectively.
-We separate surface terms into contributions from interior shared faces and from boundary faces.
-On an interior shared face, we sum contributions from the two adjacent elements to yield
-\begin{align*}
-&\sum_{f\in \Gamma_h \setminus \partial \Omega} \int_{f}{\bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v}} + {\frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma}}\\
-&= - \sum_{f \in\Gamma_h \setminus \partial \Omega} \int_{f} \frac{\tau_v}{2}\LRb{\bm{A}_n\jump{\bm{v}}}^2 + \frac{\tau_\sigma}{2}\LRb{\bm{A}_n^T\jump{\bm{\sigma}}}^2.  
-%\label{thm:stability}
-\end{align*}
-For faces which lie on the boundary $\Gamma_v$ where velocity boundary conditions are imposed, $\jump{\bm{v}} = -2\bm{v}^-$, $\jump{\bm{A}_n^T\bm{\sigma}} = 0$, and $\bm{A}_n^T\avg{\bm{\sigma}} = \bm{Sn}^-$, implying that
-\begin{align*}
-&\sum_{f\in \Gamma_v} \int_f \bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v} + \frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma}\\
-&= \sum_{f\in \Gamma_v} \int_f \bm{A}_n^T \bm{\sigma}^-\bm{v}^- - \bm{A}_n \bm{v}^- \bm{\sigma}^- - \tau_v \bm{A}_n^T\bm{A}_n \LRb{\bm{v}^-}^2 = - \sum_{f\in \Gamma_v} \int_f \tau_v \bm{A}_n^T\bm{A}_n \LRb{\bm{v}^-}^2 .
-\end{align*}
-For faces which lie on $\Gamma_\sigma$, $\bm{A}_n^T\jump{\bm{\sigma}} = -2\bm{A}_n^T\bm{\sigma}^-$, $\bm{A}_n^T\avg{\bm{\sigma}} = 0$, and $\jump{\bm{v}} = 0$, yielding a similar contribution
-\[
-\sum_{f\in \Gamma_\sigma} \int_f \bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v} + \frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma} = - \sum_{f\in \Gamma_\sigma} \int_f \tau_\sigma \bm{A}_n\bm{A}_n^T \LRb{\bm{\sigma}^-}^2.
-\]
-Finally, for faces in $\Gamma_{\rm abc}$ we have $\bm{A}_n^T\avg{\bm{\sigma}} =\frac{1}{2} \bm{A}_n^T\bm{\sigma}^- $, $\bm{A}_n^T\jump{\bm{\sigma}} = -\bm{A}_n^T\bm{\sigma}^-$, and $\jump{\bm{v}} = -\bm{v}^-$, yielding
-\begin{align*}
-\sum_{f\in \Gamma_{\rm abc}} &\int_f \bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v} + \frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma} =\\
-& - \sum_{f\in \Gamma_{\rm abc}} \int_f \frac{\tau_v}{2} \bm{A}_n^T\bm{A}_n \LRb{\bm{v}^-}^2 + \frac{\tau_\sigma}{2} \bm{A}_n\bm{A}_n^T \LRb{\bm{\sigma}^-}^2.  
-\end{align*}
-Combining all face contributions together gives the following result:
-\begin{theorem}
-The DG formulation (\ref{eq:dgform}) is energy stable for $\tau_v,\tau_\sigma \geq 0$, in the sense that
-\begin{align}
-&\sum_{D^k\in \Oh} \frac{1}{2}\pd{}{t}\LRp{\LRp{\rho \bm{v},\bm{v}}_{\Lk} + \LRp{\bm{C}^{-1} \bm{\sigma},\bm{\sigma}}_{\Lk}} = - \sum_{f \in\Gamma_h \setminus \partial \Omega} \int_{f} \frac{\tau_v}{2}\LRb{\bm{A}_n\jump{\bm{v}}}^2 + \frac{\tau_\sigma}{2}\LRb{\bm{A}_n^T\jump{\bm{\sigma}}}^2\nonumber\\
-& - \sum_{f\in \Gamma_v} \int_f \tau_v  \LRb{\bm{A}_n\bm{v}^-}^2  - \sum_{f\in \Gamma_\sigma} \int_f \tau_\sigma  \LRb{\bm{A}_n^T\bm{\sigma}^-}^2- \sum_{f\in \Gamma_{\rm abc}} \int_f \frac{\tau_v}{2} \LRb{\bm{A}_n\bm{v}^-}^2 + \frac{\tau_\sigma}{2} \LRb{\bm{A}_n^T\bm{\sigma}^-}^2 \leq 0.
-\label{thm:stability}
-\end{align}
-\end{theorem}
-
-Since $\rho$ and $\bm{C}^{-1}$ are positive definite, the left hand side of (\ref{thm:stability}) is an $L^2$-equivalent norm on $(\bm{u},\bm{\sigma})$, and Theorem~\ref{thm:stability} implies that the magnitude of the DG solution $(\bm{u},\bm{\sigma})$ is non-increasing in time.
-This also shows that dissipation present for positive penalization constants $\tau_v,\tau_{\sigma}$ acts on non-conforming components with non-zero jumps $\bm{A}_n^T\jump{\bm{v}}$ and $\bm{A}_n\jump{\bm{\sigma}}$.
-In fact, it was shown in \cite{chan2016short} that, in the limit as $\tau_v,\tau_\sigma \rightarrow\infty$, the eigenspaces of DG discretizations split into a conforming part consisting of $\bm{u},\bm{\sigma}$ which satisfy
-\[
-\bm{A}_n\jump{\bm{u}} = 0, \qquad \bm{A}_n^T\jump{\bm{\sigma}} = 0
-\]
-and an non-conforming part (defined through the $L^2$ orthogonal complement) corresponding to eigenvalues contain real parts which approach $-\infty$.
-For the linear elastic wave equations, these conditions are equivalent to the requirement of $C^0$ continuity for $\bm{u}$ and normal continuity of the stress tensor $\jump{\bm{S}}\bm{n} = 0$.
-%\begin{align*}
-%\jump{\bm{S} \bm{n}} = 0.
-%\jump{\bm{\sigma}_{xx} \bm{n}_x + \bm{\sigma}_{xy} \bm{n}_y + \bm{\sigma}_{xz} \bm{n}_z} &= 0\\
-%\jump{\bm{\sigma}_{xy} \bm{n}_x + \bm{\sigma}_{yy} \bm{n}_y + \bm{\sigma}_{yz} \bm{n}_z} &= 0\\
-%\jump{\bm{\sigma}_{xz} \bm{n}_x + \bm{\sigma}_{yz} \bm{n}_y + \bm{\sigma}_{zz} \bm{n}_z} &= 0.
-%\end{align*}
-
-\subsection{The semi-discrete matrix system for DG}
-
-The solution to (\ref{eq:dgform}) can be approximated by discretizing in space and using an explicit time integrator, which requires only evaluations of local contributions over $D^k$ to the DG formulation
-\begin{align}
-\LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= {\LRp{ \sum_{i=1}^d \bm{A}_i^T\pd{\bm{\sigma}}{\bm{x}_i},\bm{w}}_{\Lk}  + \LRa{\frac{1}{2}\bm{A}_n^T\jump{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk}} \nonumber\\
-\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= {\LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk}}.
-\label{eq:dgloc} 
-\end{align}
-
-Let $\LRc{\phi_i}_{i=1}^{N_p}$ be a basis for $P^N\LRp{\widehat{D}}$.  We define the reference mass matrix $\widehat{\bm{M}}$ and the physical mass matrix $\bm{M}$ for an element $D^k$ as
-\[
-\LRp{\widehat{\bm{M}}}_{ij} = \int_{\widehat{D}}\phi_j\phi_i \diff{\widehat{\bm{x}}}, \qquad \LRp{\bm{M}}_{ij} = \int_{D^k} \phi_j\phi_i \diff{\bm{x}} = \int_{\widehat{D}}\phi_j\phi_i J\diff{ \widehat{\bm{x}}}.
-\]
-For affine mappings, $J$ is constant and $\bm{M} = J\widehat{\bm{M}}$.  We also define weak differentiation matrices $\bm{S}_{k}$ and face mass matrix $\bm{M}_f$ such that
-\[
-\LRp{\bm{S}_{k}}_{ij} =  \int_{D^k} \pd{\phi_j}{\bm{x}_k}\phi_i \diff{\bm{x}}, \qquad \LRp{\bm{M}_{f}}_{ij} =  \int_{f} \phi_j \phi_i \diff{\bm{x}} = \int_{\widehat{f}} \phi_j \phi_i J^f\diff{\widehat{\bm{x}}},
-\]
-where $J^f$ is the Jacobian of the mapping from a reference face $\widehat{f}$ to $f$.
-For affinely mapped simplices, $J^f$ is also constant and $\bm{M}_f = J^f \widehat{\bm{M}}_f$, where the definition of the reference face mass matrix $\widehat{\bm{M}}_f$ is analogous to the definition of the reference mass matrix $\widehat{\bm{M}}$.
-
-Finally, we define weighted mass matrices.
-Let $w(\bm{x})\in \mathbb{R}$ and $\bm{W}(\bm{x})\in \mathbb{R}^{m\times n}$.
-Then, scalar and matrix-weighted mass matrices $\bm{M}_w$ and $\bm{M}_{\bm{W}}$ are defined through
-\[
-\LRp{\bm{M}_w}_{ij} = \int_{D^k} w(\bm{x}) \phi_j(\bm{x})  \phi_i(\bm{x}) \diff x, \qquad \bm{M}_{\bm{W}} = \LRp{\begin{array}{ccc}
-\bm{M}_{\bm{W}_{1,1}}& \ldots & \bm{M}_{\bm{W}_{1,n}}\\
-\vdots & \ddots & \vdots\\
-\bm{M}_{\bm{W}_{m,1}} & \ldots & \bm{M}_{\bm{W}_{m,n}}\\
-\end{array}},
-\]
-where $\bm{M}_{\bm{W}_{i,j}}$ is the scalar weighted mass matrix weighted by the $(i,j)$ entry of $\bm{W}$.
-
-Local contributions to the DG variational form may then be evaluated in a quadrature-free manner using these matrices.
-Let $\bm{\Sigma}_i,\bm{V}_i$ denote vectors containing degrees of freedom for solution components $\bm{\sigma}_i, \bm{v}_i$ such that
-\begin{align*}
-\bm{\sigma}_{i}(\bm{x},t) &= \sum_{j=1}^{N_p} \bm{\Sigma}_i(t) \phi_j(\bm{x}), \qquad 1 \leq i \leq N_d\\ 
-\bm{v}_{i}(\bm{x},t) &= \sum_{j=1}^{N_p} \bm{V}_i(t) \phi_j(\bm{x}), \qquad 1 \leq i \leq d. 
-\end{align*}
-Then, the local DG formulation can be written as a block system of ordinary differential equations (ODEs)
-%\begin{align}
-%\bm{M}_{\rho} \pd{\bm{V}_n}{t} &= \sum_{i=1}^d \sum_{j=1}^{N_d}\LRp{\bm{A}_i}_{mn} \bm{S}_{i} \bm{\Sigma}_m + \sum_{f\in \partial D^k}\bm{M}_f\bm{F}_{v,n} , \qquad n = 1,\ldots, d\nonumber\\
-%\sum_{n=1}^{N_d} \bm{M}_{\bm{C}^{-1}_{mn}} \pd{\bm{\Sigma}_n}{t} &= \sum_{i=1}^d \sum_{n=1}^{d} \LRp{\bm{A}_i}_{mn}\bm{S}_{i}\bm{V}_n  +  \sum_{f\in \partial D^k}\bm{M}_f\bm{F}_{\sigma,m}, \qquad m = 1,\ldots,N_d,
-%\label{eq:dgmat}
-%\end{align}
-%where $\bm{F}_{v,n}$ is a vector representing the $n$th component of the velocity numerical flux, $\bm{F}_{\sigma,m}$ is a vector representing the $m$th component of the stress numerical flux, and $\bm{M}_{\rho},\bm{M}_{\bm{C}^{-1}_{mn}}$ are mass matrices weighted with respect to density and the $(m,n)$ entry of $\bm{C}$, respectively.
-We can rewrite this in
-by concatenating $\bm{\Sigma}_i,\bm{V}_i$ into single vectors $\bm{\Sigma},\bm{V}$ and using the Kronecker product $\otimes$
-\begin{align}
-\bm{M}_{\rho \bm{I}} \pd{\bm{V}}{t} &= \sum_{i=1}^d  \LRp{\bm{A}_i^T\otimes \bm{S}_{i}} \bm{\Sigma} + \sum_{f\in \partial D^k}\LRp{\bm{I}\otimes \bm{M}_f}\bm{F}_{v}\nonumber\\
-\bm{M}_{\bm{C}^{-1}} \pd{\bm{\Sigma}}{t} &= \sum_{i=1}^d \LRp{\bm{A}_i\otimes \bm{S}_{i}}\bm{V}  +  \sum_{f\in \partial D^k}\LRp{\bm{I}\otimes\bm{M}_f}\bm{F}_{\sigma}. 
-\label{eq:dgmat}
-\end{align}
-where $\bm{M}_f\bm{F}_{v}$ and $\bm{M}_f\bm{F}_{\sigma}$ represent contributions to the velocity and stress equations from surface integrals.
-
-In order to apply standard time integration methods, we must invert $\bm{M}_{\rho\bm{I}}$ and $\bm{M}_{\bm{C}^{-1}}$ to isolate $\pd{\bm{v}}{t}$ and $\pd{\bm{\sigma}}{t}$ on the left hand side.
-While the inversion of $\bm{M}_{\rho\bm{I}}$ and $\bm{M}_{\bm{C}^{-1}}$ can be parallelized from element to element, doing so typically requires either the precomputation and storage of large dense matrix inverses or the on-the-fly construction and solution of a large dense matrix at every timestep.
-The former option requires a large amount of storage, while the latter option is computationally expensive and difficult to parallelize.
-This cost can be overcome for $\rho, \bm{C}$ which are constant over an element $D^k$, in which case $\bm{M}_{\rho\bm{I}}$ is block diagonal with identical blocks $\bm{M}_\rho = \rho\bm{M}$, while $\bm{M}_{\bm{C}^{-1}}$ reduces to
-\[
-\bm{M}_{\bm{C}^{-1}} = \LRp{\begin{array}{ccc}
-\bm{C}_{1,1}^{-1}\bm{M}& \ldots & \bm{C}^{-1}_{1,N_d}\bm{M}\\
-\vdots & \ddots & \vdots\\
-\bm{C}_{N_d,1}^{-1}\bm{M} & \ldots & \bm{C}^{-1}_{N_d,N_d}\bm{M}\\
-\end{array}} 
-= \LRp{\bm{C}^{-1}\otimes \bm{M}}.
-\]
-Then, $\bm{M}^{-1}_\rho = \frac{1}{\rho}\bm{M}^{-1} = \frac{1}{J\rho}\widehat{\bm{M}}^{-1}$, and $\bm{M}^{-1}_{\bm{C}^{-1}} = \bm{C}\otimes \bm{M}^{-1} = \bm{C}\otimes \LRp{\frac{1}{J}\widehat{\bm{M}}^{-1}}$, and each matrix inverse can be applied using the inverse of the reference mass matrix $\widehat{\bm{M}}^{-1}$ and the values of $\rho, \bm{C}$, and $J$ on each element.
-Applying this observation to \ref{eq:dgmat} then yields the following local system of ODEs
-\begin{align*}
-\pd{\bm{V}}{t} &= \sum_{i=1}^d  \LRp{\frac{1}{\rho}\bm{A}_i^T\otimes \bm{D}_{i}} \bm{\Sigma} + \sum_{f\in \partial D^k}\LRp{\frac{1}{\rho}\bm{I}\otimes \bm{L}_f}\bm{F}_{v}\\
-\pd{\bm{\Sigma}}{t} &= \sum_{i=1}^d \LRp{\bm{C}\bm{A}_i\otimes \bm{D}_{i}}\bm{V}  +  \sum_{f\in \partial D^k}\LRp{\bm{C}\otimes\bm{L}_f}\bm{F}_{\sigma},
-\end{align*}
-where we have introduced the differentiation matrix $\bm{D}_i = \bm{M}^{-1}\bm{S}_i$ and lift matrix $\bm{L}_f = \bm{M}^{-1}\bm{M}_f$.
-For affine elements, both derivative and lift matrices may be applied using only geometric factors and reference derivative and lift matrices.
-
-Unfortunately, if $\rho$ and $\bm{C}$ vary spatially within an element, the above approach can no longer be used to invert $\bm{M}_\rho$ and $\bm{M}_{\bm{C}^{-1}}$ in an efficient and low-storage manner.
-%One can avoid these issues by using mass lumping techniques, where nodal points for Lagrange basis functions are co-located at quadrature points on simplices \cite{chin1999higher, cohen2001higher}; however, these points are only available up to order $N\leq 4$ in three dimensions.
-Additionally, these techniques can require the use of enriched polynomial spaces due to a mismatch between the number of points and dimension of $P^N$.
-Mass-lumping can also be done using interior quadratures with $N_p$ points \cite{williams2014symmetric}, though these are defined only up to $N=6$ in three dimensions, and proving high order accuracy.
-For isotropic media, one way to address sub-element variations in material parameters is to diagonalize the matrix $\bm{C}$ through a change of variables \cite{etienne2010hp}.
-This results in a local system of ODEs with only scalar weighted mass matrices \cite{mercerat2015nodal}, which can be treated using scalar weight-adjusted approximations.
-We take a different approach in addressing these issues by approximating the matrix-weighted mass matrix $\bm{M}_{\bm{C}^{-1}}$ and corresponding matrix-weighted $L^2$ inner product with a weight-adjusted approximation which is low storage, simple to invert, energy stable, and provably high order accurate for spatially varying weights $\rho, \bm{C}$ with sufficiently regularity.
-
-%This implies that the RHS terms are 
-%\[
-%\bm{A}_1 \pd{\bm{v}}{x} + \bm{A}_2 \pd{\bm{v}}{y}  = \LRp{\begin{array}{c}
-%\Div \bm{\sigma}_x \\
-%\Div \bm{\sigma}_y \\
-%\pd{u_1}{x} \\
-%\pd{u_2}{y} \\
-%\pd{u_2}{x} + \pd{u_1}{y}
-%\end{array}}
-%\]
-
-%The central flux terms for elasticity are then given as
-%\[
-%\bm{A}_n\avg{\bm{v}} = \LRp{\begin{array}{c}
-%\avg{\bm{\sigma}_x} \cdot \bm{n}\\
-%\avg{\bm{\sigma}_y} \cdot \bm{n}\\
-%\avg{u_1}{n}_x\\
-%\avg{u_2}{n}_y\\
-%\avg{u_2}{n}_x + \avg{u_1}{n}_y
-%\end{array}}
-%\]
-%while the penalty flux term is
-%\[
-%\bm{A}_n^T\bm{A}_n\jump{\bm{v}} = 
-%\LRp{\begin{array}{ccccc}
-%1 & n_xn_y & & &\\
-%n_xn_y & 1 & & &\\
-% & & n_x^2 & 0 & n_xn_y \\
-%  & & 0 & n_y^2 & n_xn_y \\
-%   & & n_xn_y & n_xn_y & 1 
-%\end{array}}
-%\LRp{\begin{array}{c}
-%\jump{u_1}\\
-%\jump{u_2}\\
-%\jump{\sigma_{xx}}\\
-%\jump{\sigma_{yy}}\\
-%\jump{\sigma_{xy}}
-%\end{array}}
-%=
-%\LRp{\begin{array}{c}
-%\jump{u_1} + \jump{u_2}n_xn_y\\
-%n_xn_y\jump{u_1} + \jump{u_2}\\
-%\LRp{\jump{\bm{\sigma}_x}\cdot\bm{n}}n_x\\
-%\LRp{\jump{\bm{\sigma}_y}\cdot\bm{n}}n_y\\
-%\jump{\sigma_{xy}} + n_x n_y \LRp{\jump{\sigma_{xx}} + \jump{\sigma_{yy}}}
-%\end{array}}.
-%\]
-%We note that $\bm{A}_n^T\bm{A}_n$ has eigenvalues $1 \pm n_x n_y$ and thus has norm less than $3/2$.  
-%We choose the penalty parameter $\bm{\tau}$ as a diagonal matrix such that $\bm{A}_0^{-1}\bm{\tau}\bm{A}_n^T\bm{A}_n = O(1)$.
-In this case, $\tau_v = O\LRp{\frac{1}{\rho}}$, while $\tau_{\sigma} = O\LRp{\frac{1}{2\mu + \lambda}}$.
-
-%This implies that the RHS terms are 
-%\[
-%\bm{A}_1 \pd{\bm{v}}{x} + \bm{A}_2 \pd{\bm{v}}{y} +\bm{A}_3 \pd{\bm{v}}{z}  = \LRp{\begin{array}{c}
-%\Div \bm{\sigma}_x \\
-%\Div \bm{\sigma}_y \\
-%\Div \bm{\sigma}_z \\
-%\pd{u_1}{x} \\
-%\pd{u_2}{y} \\
-%\pd{u_3}{z} \\
-%\pd{u_3}{y} + \pd{u_2}{z}\\
-%\pd{u_3}{x} + \pd{u_1}{z}\\
-%\pd{u_2}{x} + \pd{u_1}{y}
-%\end{array}}.
-%\]
-%The central flux terms for elasticity are then given as
-%\[
-%\bm{A}_n\avg{\bm{v}} = \LRp{\begin{array}{c}
-%\avg{\bm{\sigma}_x} \cdot \bm{n}\\
-%\avg{\bm{\sigma}_y} \cdot \bm{n}\\
-%\avg{\bm{\sigma}_z} \cdot \bm{n}\\
-%\avg{u_1}{n}_x\\
-%\avg{u_2}{n}_y\\
-%\avg{u_3}{n}_z\\
-%\avg{u_3}{n}_y + \avg{u_2}{n}_z\\
-%\avg{u_3}{n}_x + \avg{u_1}{n}_z\\
-%\avg{u_2}{n}_x + \avg{u_1}{n}_y
-%\end{array}}
-%\]
-%The penalty flux term is evaluated simply by re-applying $\bm{A}_n = \bm{A}_n^T$.  Expanded out, it yields
-%\begin{align*}
-%\bm{A}_n^T\bm{A}_n\jump{\bm{v}} &= 
-%\LRp{\begin{array}{ccc|ccc|ccc}
-%1 & n_xn_y & n_xn_z   &&&&&&\\
-%n_xn_y & 1 & n_yn_z   &&&&&&\\
-%n_xn_z & n_yn_z & 1   &&&&&&\\
-%\hline\\
-%&&&  n_x^2&&&&n_xn_z&n_xn_y\\
-%&&&  &n_y^2&&n_xn_z&&n_xn_y\\
-%&&&  &&n_z^2&n_xn_z&n_xn_y&\\
-%\hline\\
-%&&&  &n_xn_z& n_xn_y& n_y^2+n_z^2&n_xn_y&n_xn_z\\
-%&&&  n_xn_z && n_xn_y& n_xn_y& n_x^2+n_z^2&n_yn_z\\
-%&&&  n_xn_z &n_xn_y && n_xn_z& n_yn_z& n_x^2+n_y^2\\
-%\end{array}}
-%\LRp{\begin{array}{c}
-%\jump{u_1}\\
-%\jump{u_2}\\
-%\jump{u_3}\\
-%\jump{\sigma_{xx}}\\
-%\jump{\sigma_{yy}}\\
-%\jump{\sigma_{zz}}\\
-%\jump{\sigma_{yz}}\\
-%\jump{\sigma_{xz}}\\
-%\jump{\sigma_{xy}}
-%\end{array}}\\
-%&=
-%\LRp{\begin{array}{c}
-%\jump{u_1} + \jump{u_2}n_xn_y + \jump{u_3}n_xn_z\\
-%n_xn_y\jump{u_1} + \jump{u_2} + n_yn_z\jump{u_3}\\
-%n_xn_z\jump{u_1} + n_yn_z\jump{u_2} + \jump{u_3}\\
-%\LRp{\jump{\bm{\sigma}_x}\cdot\bm{n}}n_x\\
-%\LRp{\jump{\bm{\sigma}_y}\cdot\bm{n}}n_y\\
-%\LRp{\jump{\bm{\sigma}_z}\cdot\bm{n}}n_z\\
-%\jump{\sigma_{yz}}(n_y^2+n_z^2) + \LRp{\jump{\sigma_{xz}}n_y + \jump{\sigma_{xy}}n_z}n_x + \LRp{\jump{\sigma_{yy}}n_y + \jump{\sigma_{zz}}n_y}n_z\\
-%\jump{\sigma_{xz}}(n_x^2+n_z^2) + \LRp{\jump{\sigma_{xx}}n_z + \jump{\sigma_{yz}}n_y}n_x + \LRp{\jump{\sigma_{xy}}n_y + \jump{\sigma_{zz}}n_x}n_z\\
-%\jump{\sigma_{xy}}(n_x^2+n_y^2) + \LRp{\jump{\sigma_{xx}}n_y + \jump{\sigma_{yy}}n_y}n_x + \LRp{\jump{\sigma_{xz}}n_y + \jump{\sigma_{yz}}n_x}n_z\\
-%\end{array}}.
-%\end{align*}
-%
-%We choose the penalty parameter $\bm{\tau}$ as a diagonal matrix such that $\bm{A}_0^{-1}\bm{\tau}\bm{A}_n^T\bm{A}_n = O(1)$.
-In this case, $\tau_v = O\LRp{\frac{1}{\rho}}$, while $\tau_{\sigma} = O\LRp{\frac{1}{2\mu + \lambda}}$.
-
-\section{Weight-adjusted inner products for matrix-valued weights}
-\label{sec:mwadg}
-Weight-adjusted inner products are high order accurate approximations of weighted $L^2$ inner products.
-These can be interpreted as generalizations of mass lumping techniques, reducing to mass lumping when integrals are evaluated with appropriate quadrature rules.
-These weight-adjusted inner products result in weight-adjusted mass matrices, whose inverses approximate the inverse of a weighted $L^2$ mass matrix.
-
-We wish to apply weight-adjusted approximations to avoid the inversion of $\bm{M}_\rho$ and $\bm{M}_{\bm{C}^{-1}}$.
-Approximating the inverse of $\bm{M}_\rho$ can be done using weight-adjusted approximations for scalar weights \cite{chan2016weight1,chan2016weight2}, which we review in Section~\ref{sec:swadg}.
-We then extend scalar weight-adjusted approximations to matrix-valued weights in Section~\ref{sec:mwadg} to approximate the inverse of $\bm{M}_{\bm{C}^{-1}}$.
-
-\subsection{Scalar weight adjusted inner products}
-\label{sec:swadg}
-
-We introduce standard Lebesgue $L^p$ norms and their associated $L^p$ spaces over a general domain $\Omega$
-\begin{align*}
-\nor{u}_{L^p\LRp{\Omega}} = \LRp{\int_{\Omega} u^p}^{1/p}, \qquad  
-L^p\LRp{\Omega} = \LRc{u: \Omega\rightarrow \mathbb{R}, \quad \nor{u}_{L^p\LRp{\Omega}} < \infty} 
-\end{align*}
-for $1 \leq p < \infty$.  For $p = \infty$, these are defined as
-\begin{align*}
-\nor{u}_{L^{\infty}\LRp{\Omega}} = \inf\LRc{C \geq 0: \LRb{u\LRp{\bm{x}}} \leq C \quad \forall \bm{x}\in \Omega}, \qquad L^{\infty}\LRp{\Omega} &= \LRc{u: \Omega\rightarrow \mathbb{R}, \quad \nor{u}_{L^{\infty}\LRp{\Omega}} < \infty}.
-\end{align*}
-These induce $L^p$ Sobolev seminorms and norms of degree $s$ 
-\begin{align*}
-\LRb{u}_{W^{s,p}\LRp{\Omega}} &= \LRp{\sum_{\LRb{\alpha}= s} \nor{ D^{\alpha} u}_{L^p\LRp{\Omega}}^p}^{1/p}, \qquad \LRb{u}_{W^{s,\infty}\LRp{\Omega}} = \max_{\LRb{\alpha}= s} \nor{D^{\alpha}u}_{L^{\infty}\LRp{\Omega}}\\
-\nor{u}_{W^{s,p}\LRp{\Omega}} &= \LRp{\sum_{\LRb{\alpha}\leq s} \nor{ D^{\alpha} u}_{L^p\LRp{\Omega}}^p}^{1/p}, \qquad \nor{u}_{W^{s,\infty}\LRp{\Omega}} = \max_{\LRb{\alpha}\leq s} \nor{D^{\alpha}u}_{L^{\infty}\LRp{\Omega}}.
-\end{align*}
-where $\alpha = \LRc{\alpha_1,\alpha_2,\alpha_3}$ is a multi-index such that
-\[
-D^{\alpha}u = \pd{^{\alpha_1}}{x^{\alpha_1}}\pd{^{\alpha_2}}{y^{\alpha_2}}\pd{^{\alpha_3}}{z^{\alpha_3}} u,
-\]
-
-We define two operators $T_w: L^2\LRp{D^k} \rightarrow P^N\LRp{D^k}$ and $T^{-1}_w: L^2\LRp{D^k} \rightarrow P^N\LRp{D^k}$ such that
-\[
-T_wu = \Pi_N(wu), \qquad \LRp{wT^{-1}_wu,v}_{\Lk} = \LRp{u,v}_{\Lk}.  
-\]
-A weighted $L^2$ inner product $\LRp{wu,v}_{\Lk}$ can be approximated by a weight-adjusted inner product
-\[
-\LRp{wu,v}_{\Lk} = \LRp{T_wu,v}_{\Lk}\approx \LRp{T^{-1}_{1/w}u,v}_{\Lk}.  
-\]
-based on the observation that $T^{-1}_{1/w}u \approx uw$.
-This approximation is made precise by the following estimates for approximations of the product $uw$ and weighted moments on affinely mapped elements:
-\begin{theorem}[Theorem 5 in \cite{chan2016weight1}]
-Let $D^k$ be quasi-regular with representative size $h = {\rm diam}\LRp{D^k}$.  For $N \geq 0$, $w\in W^{N+1,\infty}\LRp{D^k}$, and $u \in W^{N+1,2}\LRp{D^k}$, 
-\begin{align}
-\nor{{u}{w}-T_{w}u}_{L^2\LRp{D^k}} &\leq C_wh^{N+1} \nor{u}_{W^{N+1,2}\LRp{D^k}},\\
-\nor{{u}{w}-T^{-1}_{1/w}u}_{L^2\LRp{D^k}} &\leq C_wh^{N+1}   \nor{u}_{W^{N+1,2}\LRp{D^k}}.
-\end{align}
-where $C_w$ depends on $w$ as follows:
-\[
-C_w = C\nor{w}_{L^{\infty}\LRp{D^k}}\nor{\frac{1}{w}}_{L^{\infty}\LRp{D^k}} \nor{w}_{W^{N+1,\infty}\LRp{D^k}}.
-\]
-\end{theorem}
-\begin{theorem}[Theorem 6 in \cite{chan2016weight1}]
-Let $u\in W^{N+1,2}\LRp{D^k}$, $w\in W^{N+1,\infty}\LRp{D^k}$, and $v \in P^M\LRp{D^k}$ for $0 \leq M\leq N$; then
-\begin{align*}
-&\LRb{\LRp{{w}u,v}_{L^2\LRp{D^k}} - \LRp{T^{-1}_{1/w}u,v}} \\
-&\leq Ch^{2N+2 - M} \nor{{w}}_{L^{\infty}\LRp{D^k}}\nor{\frac{1}{w}}_{L^{\infty}\LRp{D^k}}^2 \nor{w}^2_{W^{N+1,\infty}\LRp{D^k}}  \nor{u}_{W^{N+1,2}\LRp{D^k}} \nor{v}_{L^{\infty}\LRp{D^k}}.
-\end{align*}
-\end{theorem}
-Both theorems rely on a scalar weighted interpolation estimate derived in \cite{warburton2013low,chan2016weight1} for a general non-affine element $D^k$.  
-\begin{theorem}[Theorem 3.1 in \cite{chan2016weight1}]
-Let $D^k$ be a quasi-regular element with representative size $h = {\rm diam}\LRp{D^k}$. For $N \geq 0$, $w\in W^{N+1,\infty}$, and $u\in W^{N+1,2}$, 
-\[
-\nor{u - \frac{1}{w}\Pi_N(wu)}_{\Lk} \leq C_N h^{N+1} \nor{{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{w}}_{L^\infty}\nor{w}_{W^{N+1,\infty}}\nor{u}_{W^{N+1,2}}.
-\]
-\label{thm:wproj}
-\end{theorem}
-
-\subsection{Extension to matrix weights}
-
-We now aim to generalize weight-adjusted inner products to the case of matrix-valued weights.
-We first define appropriate generalizations of norms used in Section~\ref{sec:swadg} to vector-valued functions.
-Let $D^{\bm{\alpha}}\bm{v}$ denote component-wise differentiation of $\bm{v}$ with respect to a $d$-dimensional multi-index $\bm{\alpha}$.
-Then, vector $L^p$ Sobolev norms for $\bm{v}(\bm{x}) \in \mathbb{R}^m$ can be defined as
-\begin{align*}
-\nor{\bm{v}}_{W^{k,p}}^p &= {\sum_{i=1}^m \nor{\bm{v}_i}_{W^{k,2}}^p} = \sum_{\LRb{\alpha} \leq k} \nor{D^{\alpha}\bm{v}}^p_{L^p}, \qquad 1 \leq p < \infty,\\
-\nor{\bm{v}}_{W^{k,\infty}} &= \max_i \nor{\bm{v}_i}_{W^{k,\infty}}.
-\end{align*}
-The corresponding Sobolev spaces $W^{k,p}$ and $W^{k,\infty}$ are defined similar to the scalar case.
-
-Let $\bm{W}(\bm{x})$ be a matrix-valued weight function which is pointwise symmetric positive-definite matrix
-\[
-0 < w_{\min} \leq  \nor{\bm{W}(\bm{x})}_2 \leq w_{\max} < \infty, \qquad 0 < \tilde{w}_{\min} \leq  \nor{\bm{W}^{-1}(\bm{x})}_2 \leq \tilde{w}_{\max} < \infty, \qquad \forall \bm{x} \in \Omega.
-\]
-We define a $k$th order Sobolev norm for $\bm{W}(\bm{x})$ in terms of the induced $p$-norm
-\begin{align*}
-\nor{\bm{W}(\bm{x})}_{k,p,\infty}^p &= \sum_{\LRb{\bm{\alpha}} \leq k} \sup_{\bm{x}} \nor{D^{\bm{\alpha}}\bm{W}(\bm{x})}^p_p
-\end{align*}
-While this norm is not sub-multiplicative, the following bound holds
-\begin{align*}
-\nor{\bm{W}{\bm{v}}}_{W^{k,p}}^p &= \sum_{\LRb{\alpha}\leq k} \nor{D^{\alpha}\LRp{\bm{W}\bm{v}}}_{L^p}^p 
-%\leq C_N \sum_{\LRb{\alpha}\leq k} \nor{\sum_{\LRb{\beta}\leq \alpha} \LRp{D^{\beta}\bm{W}}\LRp{D^{\alpha-\beta}\bm{v}} }_{L^p}^p\\
-\leq C_N \int  \sum_{\LRb{\alpha}\leq k} \sum_{\LRb{\beta}\leq \LRb{\alpha}} \nor{\LRp{D^{\beta}\bm{W}}\LRp{D^{\alpha-\beta}\bm{v}} }_{p}^p \\
-&\leq C_N\int \LRp{\sum_{\LRb{\alpha}\leq k} \nor{\LRp{D^{\alpha}\bm{W}}}}^p \LRp{\sum_{\LRb{\alpha}\leq k} \nor{D^{\alpha}\bm{v}}_{p} }^p\\
-&\leq C_N \nor{\bm{W}}_{k,p,\infty}^p \nor{\bm{v}}_{k,p}^p,
-\end{align*}
-where we have used Leibniz's rule, Cauchy-Schwarz, and the arithmetic-geometric mean inequality.
-
-The following theorem extends Theorem~\ref{thm:wproj} to matrix weights by computing weighted interpolation estimates for the quantity $\bm{W}^{-1}\Pi_N\LRp{\bm{W}\bm{v}}$.  
-\begin{theorem}
-Let $D^k$ be a quasi-regular element with representative size $h = {\rm diam}\LRp{D^k}$. For $N \geq 0$, $\bm{W} \in \LRp{W^{N+1,\infty}}^{d\times d}$, and $\bm{v}\in \LRp{W^{N+1,2}}^d$, 
-\[
-\nor{\bm{v} - \bm{W}^{-1}\Pi_N\LRp{\bm{W}\bm{v}}}_{L^2} \leq Ch^{N+1} \nor{{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk}\tilde{w}_{\max}\nor{\bm{W}}_{N+1,2,\infty} \nor{\bm{v}}_{W^{N+1,2}}
-\]
-\label{thm:wwproj}
-\end{theorem}
-\begin{proof}
-The proof is the similar to the scalar case.  Using vector-valued versions of Bramble-Hilbert and a scaling argument for quasi-regular elements yields
-\begin{align*}
-\nor{\bm{v} - \bm{W}^{-1}\Pi_N\LRp{\bm{W}\bm{v}}}_{L^2} &\leq C_1 \nor{\sqrt{J}}_{\LinfDk} \sup_{\bm{x}}\nor{\bm{W}^{-1}}_{2} \nor{\bm{W}\bm{v} - \Pi_N\LRp{\bm{W}\bm{v}}}_{L^2}\\
-&\leq C_1 \nor{{\sqrt{J}}}_{\LinfDk} \sup_{\bm{x}}\nor{\bm{W}^{-1}}_{2} \LRb{\bm{W}\bm{v}}_{W^{N+1,2}\LRp{\Dhat}}\\
-&\leq C_2 h^{N+1}\nor{{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk} \sup_{\bm{x}}\nor{\bm{W}^{-1}}_{2} \nor{\bm{W}\bm{v}}_{W^{N+1,2}\LRp{D^k}}\\
-&\leq C_3 h^{N+1}\nor{{\sqrt{J}}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk} \tilde{w}_{\max} \nor{\bm{W}}_{{N+1,2,\infty}}\nor{\bm{v}}_{W^{N+1,2}\LRp{D^k}}.
-\end{align*}
-\end{proof}
-
-\subsubsection{Weight-adjusted approximations with matrix weights}
-\label{sec:mwadgprop}
-
-Let $\Pi_N\bm{u}$ be defined as the $L^2$ projection applied to each component of the vector-valued function $\bm{u}$.  We then define the operator $T_{\bm{W}}$ analogously to the scalar case
-\[
-T_{\bm{W}}\bm{v} = \Pi_N\LRp{\bm{W}\bm{v}}.  
-\]
-The inverse operator $T^{-1}_{\bm{W}}$ is defined implicitly via
-\[
-\LRp{\bm{W}T^{-1}_{\bm{W}}\bm{v},\bm{w}} = \LRp{\bm{v},\bm{w}} .
-\]
-This definition is a straightforward generalization of $T^{-1}_{w}$ to matrix-valued weights $\bm{W}$.
-Conveniently, all properties of $T_{w},T^{-1}_{w}$ for scalar $w(\bm{x})$ carry over to matrix weights $\bm{W}(\bm{x})$ as well \cite{chan2016weight1}.
-\begin{lemma}
-Let $\Pi_N$ denote the component-wise $L^2$ projection, and let $\bm{W} \in \LRp{L^{\infty}}^m$.  Then, $T_{\bm{W}}$ satisfies the following properties:
-\begin{enumerate}
-\vspace{.5em}
-\item $T_{\bm{W}}^{-1}T_{\bm{W}} = \Pi_N$ \label{p1}
-\vspace{.5em}
-\item $\Pi_N T_{\bm{W}}^{-1} = T_{\bm{W}}^{-1}\Pi_N  =  T_{\bm{W}}^{-1}$ \label{p2}
-\vspace{.5em}
-\item $\nor{T^{-1}_{\bm{W}}}_{\Lk} \leq \tilde{w}_{\max}$. \label{p4}
-\vspace{.5em}
-\item $\LRp{T^{-1}_{\bm{W}} \bm{v},\bm{w}}$ forms an inner product on $\LRp{P^N}^m \times \LRp{P^N}^m$, which is equivalent to the $L^2$ inner product with equivalence constants $C_1 = \tilde{w}_{\min}, C_2 = \tilde{w}_{\max}$.
-\label{p3}
-\vspace{.5em}
-\end{enumerate}
-\label{lemma:props}
-\end{lemma}
-\begin{proof}
-The proofs of properties~\ref{p1} and \ref{p2} are consequences of the definition of $T_{\bm{W}}$, and are identical to proofs for the scalar case.
-Property~\ref{p4} is a straightforward extension from the scalar case.
-Let $\bm{v}\in \LRp{P^N}^m$.
-Then,
-\[
-\LRp{T^{-1}_{\bm{W}} \bm{v},\bm{v}} = \LRp{\bm{W}^{-1}\bm{W}T^{-1}_{\bm{W}} \bm{v},\bm{v}} \leq \sup_{\bm{x}}\nor{\bm{W}^{-1}(\bm{x})}\LRp{\bm{W}T^{-1}_{\bm{W}} \bm{v},\bm{v}} = \tilde{w}_{\max} \nor{\bm{v}}_{L^2}^2.
-\]
-Property~\ref{p3} then simply requires the lower bound
-\[
-\LRp{T^{-1}_{\bm{W}} \bm{v},\bm{v}} = \LRp{\bm{W}^{-1}\bm{W}T^{-1}_{\bm{W}} \bm{v},\bm{v}} \geq \inf_{\bm{x}}\nor{\bm{W}^{-1}(\bm{x})}\nor{\bm{v}}_{L^2}^2 = \tilde{w}_{\min}\nor{\bm{v}}_{L^2}^2 .  
-\]
-\end{proof}
-
-Using Theorem~\ref{thm:wwproj}, we may also show that the matrix-valued weight-adjusted inner product is also high order accurate for sufficiently regular $\bm{W}$.
-\begin{theorem}
-Let $D^k$ be a quasi-regular element with representative size $h = {\rm diam}\LRp{D^k}$. For $N > 0$
-\[
-\nor{ \bm{W}\bm{v} - {T}^{-1}_{\bm{W}^{-1}}\bm{v} }_{\Lk } \leq C_{\bm{W}} h^{N+1} \nor{\bm{v}}_{W^{N+1,2}\LRp{D^k}}
-\]
-with constant $C_{\bm{W}}$ depending on $\bm{W}$ and $N$
-\[
-%C_{\bm{W}} = C_N \cond{\bm{W}}{2,\infty} \nor{\bm{W}}_{W^{N+1,2}\LRp{D^k}}.  
-C_{\bm{W}} = C_N \nor{\sqrt{J}}_{\LinfDk}\nor{\frac{1}{\sqrt{J}}}_{\LinfDk} \tilde{w}_{\max}{w}_{\max} \nor{\bm{W}}_{{N+1,2,\infty}}.  
-\]
-\label{thm:west}
-\end{theorem}
-\begin{proof}
-The proof is again similar to the scalar case.  The triangle inequality gives
-\[
-\nor{ \bm{W}\bm{v} - {T}^{-1}_{\bm{W}^{-1}}\bm{v} }_{\Lk } \leq \nor{ \bm{W}\bm{v} - \Pi_N(\bm{W}\bm{v})}_{\Lk} + \nor{\Pi_N(\bm{W}\bm{v}) - T^{-1}_{\bm{W}^{-1}}\bm{v} }_{\Lk }.  
-\]
-The former term is bounded by interpolation estimates and by arguments used in the proof of Theorem~\ref{thm:wwproj}.  The latter term is bounded 
-\begin{align*}
-\nor{\Pi_N(\bm{W}\bm{v})- T^{-1}_{\bm{W}^{-1}}\bm{v} }_{\Lk} &= \nor{T^{-1}_{\bm{W}^{-1}} T_{\bm{W}^{-1}}\Pi_N(\bm{W}\bm{v})- T^{-1}_{\bm{W}^{-1}}\Pi_N\bm{v} }_{\Lk } \\
-&= \nor{T^{-1}_{\bm{W}^{-1}} \Pi_N \LRp{T_{\bm{W}^{-1}}\Pi_N(\bm{W}\bm{v})}- T^{-1}_{\bm{W}^{-1}}\Pi_N\bm{v} }_{\Lk } \\
-&\leq C_N \nor{T^{-1}_{\bm{W}^{-1}}} \nor{ T_{\bm{W}^{-1}}\Pi_N(\bm{W}\bm{v})-\Pi_N\bm{v} }_{\Lk }\\
-&\leq C_N \nor{T^{-1}_{\bm{W}^{-1}}} \nor{ \Pi_N\LRp{\bm{W}^{-1}\Pi_N(\bm{W}\bm{v})} - \Pi_N\bm{v} }_{\Lk }\\
-&\leq C_N {w}_{\max} \nor{ \bm{W}^{-1}\Pi_N(\bm{W}\bm{v}) - \bm{v} }_{\Lk }.
-\end{align*}
-where we have used $\nor{T_{\bm{W}^{-1}}^{-1}} \leq {w}_{\max}$ (Lemma~\ref{lemma:props}) and $\nor{\Pi_N}_{\Lk} = 1$.  
-An application of Theorem~\ref{thm:wwproj} to bound $\nor{ \bm{W}^{-1}\Pi_N(\bm{W}\bm{v}) - \bm{v} }_{\Lk }$ completes the proof.
-\end{proof}
-
-We note that these estimates are tight, in the sense that we recover estimates for scalar weight-adjusted inner products when the matrix weight is taken to be $\bm{W} = w\bm{I}$.  
-%\note{This is a good generalization because defining $T^{-1}_w \bm{v} = T^{-1}_{w\bm{I}}\bm{v}$ recovers estimates for scalar weight-adjusted inner products.}
-
-%We note that the local conservation property of DG is lost when using weight-adjusted inner products.  However,
-
-\subsubsection{Approximation of weighted mass matrix inverses}
-
-The advantage of using weight-adjusted inner products is that the corresponding weight-adjusted mass matrices are straightforward to invert.
-For scalar weights, the weight-adjusted mass matrix approximates the weighted $L^2$ mass matrix and its inverse by
-\[
-\bm{M}_w \approx \bm{M}\bm{M}^{-1}_{1/w}\bm{M}, \qquad \bm{M}^{-1}_w \approx \bm{M}^{-1}\bm{M}_{1/w}\bm{M}^{-1}.
-\]
-By evaluating $\bm{M}_{1/w}$ in a matrix-free fashion (for example, using a sufficiently accurate quadrature rule), the inverse of the weight-adjusted mass matrix $\bm{M}^{-1}\bm{M}_{1/w}\bm{M}^{-1}$ can be applied in a low storage fashion.
-
-For weight-adjusted inner products with matrix-valued weights, the corresponding weight-adjusted mass matrices approximate weighted $L^2$ mass matrices and inverses in a similar fashion
-\begin{align}
-\bm{M}_{\bm{W}} \approx \LRp{\bm{I} \otimes \bm{M}} \bm{M}^{-1}_{\bm{W}^{-1}} \LRp{\bm{I} \otimes \bm{M}} \nonumber\\
-\bm{M}_{\bm{W}}^{-1} \approx \LRp{\bm{I} \otimes \bm{M}^{-1}} \bm{M}_{\bm{W}^{-1}} \LRp{\bm{I} \otimes \bm{M}^{-1}}.
-\label{eq:wadgmc}
-\end{align}
-
-We note that, when $\bm{W}$ is constant over $D^k$, $\bm{M}_{\bm{W}}$ reduces to the Kronecker product of the inverse stiffness tensor and the local mass matrix 
-\[
-\bm{M}_{\bm{W}} = \bm{W}\otimes \bm{M}, \qquad \bm{M}_{\bm{W}}^{-1} = \bm{W}^{-1}\otimes \bm{M}^{-1}.
-\] 
-In this case, we also have $\bm{M}_{\bm{W}^{-1}} = \bm{W}^{-1}\otimes \bm{M}$, and substituting this explicit inverse into the weight adjusted mass matrix inverse $\LRp{\bm{I} \otimes \bm{M}^{-1}} \bm{M}_{\bm{W}^{-1}} \LRp{\bm{I} \otimes \bm{M}^{-1}}$ in (\ref{eq:wadgmc}) recovers the exact inversion of $\bm{M}_{\bm{W}}$.
-
-\section{An energy stable weight-adjusted discontinuous Galerkin formulation for elastic wave propagation}
-\label{sec:wadgelas}
-%Recall the energy stable formulation for DG 
-%\begin{align*}
-%\LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= \LRp{ \sum_{i=1}^d \bm{A}_i^T\pd{\bm{\sigma}}{\bm{x}_i},\bm{w}}_{\Lk}  + \LRa{\frac{1}{2}\bm{A}_n^T\jump{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{v}},\bm{w}}_{\Ldk}\\
-%\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{\sigma}},\bm{q}}_{\Ldk}.
-%\end{align*}
-We construct a weight-adjusted DG method by simply replacing the weighted $L^2$ inner products appearing in the left hand side of the local DG formulation (\ref{eq:dgloc}) with weight-adjusted approximations
-\begin{align*}
-\LRp{T^{-1}_{1/\rho} \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= \LRp{ \sum_{i=1}^d \bm{A}_i^T\pd{\bm{\sigma}}{\bm{x}_i},\bm{w}}_{\Lk}  + \LRa{\frac{1}{2}\bm{A}_n^T\jump{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{v}},\bm{w}}_{\Ldk}\\
-\LRp{T^{-1}_{\bm{C}} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{\sigma}},\bm{q}}_{\Ldk}.
-\end{align*}
-Since the right hand side of the WADG formulation is identical to the right hand side of the DG formulation (\ref{eq:dgform}), WADG preserves a variant of the energy stability in Theorem~\ref{thm:stability}
-\begin{align*}
-&\sum_{D^k\in \Oh} \frac{1}{2}\pd{}{t}\LRp{\LRp{T^{-1}_{1/\rho} \bm{v},\bm{v}}_{\Lk} + \LRp{T^{-1}_{\bm{C}} \bm{\sigma},\bm{\sigma}}_{\Lk}} \leq - \sum_{f \in\Gamma_h \setminus \partial \Omega} \int_{f} \frac{\tau_v}{2}\LRb{\bm{A}_n\jump{\bm{v}}}^2 + \frac{\tau_\sigma}{2}\LRb{\bm{A}_n^T\jump{\bm{\sigma}}}^2\nonumber\\
-& - \sum_{f\in \Gamma_v} \int_f \tau_v  \LRb{\bm{A}_n\bm{v}^-}^2  - \sum_{f\in \Gamma_\sigma} \int_f \tau_\sigma  \LRb{\bm{A}_n^T\bm{\sigma}^-}^2- \sum_{f\in \Gamma_{\rm abc}} \int_f \frac{\tau_v}{2} \LRb{\bm{A}_n\bm{v}^-}^2 + \frac{\tau_\sigma}{2} \LRb{\bm{A}_n^T\bm{\sigma}^-}^2 \leq 0.
-\end{align*}
-
-The substitution of weight-adjusted inner products for weighted $L^2$ inner products replaces the weighted $L^2$ mass matrices in (\ref{eq:dgmat}) by their weight-adjusted approximations.  
-%\begin{align*}
-%\bm{M}_{\rho \bm{I}} \pd{\bm{V}}{t} &= \sum_{i=1}^d  \LRp{\bm{A}_i^T\otimes \bm{S}_{i}} \bm{\Sigma} + \sum_{f\in \partial D^k}\LRp{\bm{I}\otimes \bm{M}_f}\bm{F}_{v}\nonumber\\
-%\bm{M}_{\bm{C}^{-1}} \pd{\bm{\Sigma}}{t} &= \sum_{i=1}^d \LRp{\bm{A}_i\otimes \bm{S}_{i}}\bm{V}  +  \sum_{f\in \partial D^k}\LRp{\bm{I}\otimes\bm{M}_f}\bm{F}_{\sigma}. 
-%\end{align*}
-Inverting these weight-adjusted mass matrices yields the following local system of ODEs for $\bm{V},\bm{\Sigma}$
-\begin{align*}
-\pd{\bm{V}}{t} &= \LRp{\bm{I}\otimes\bm{M}^{-1}} \bm{M}_{\LRp{\rho^{-1}\bm{I}}} \LRp{\sum_{i=1}^d  \LRp{\bm{A}_i^T\otimes \bm{D}_{i}} \bm{\Sigma} + \sum_{f\in \partial D^k}\LRp{\bm{I}\otimes \bm{L}_f}\bm{F}_{v}} \nonumber\\
-\pd{\bm{\Sigma}}{t} &= \LRp{\bm{I}\otimes\bm{M}^{-1}} \bm{M}_{\bm{C}}  \LRp{\sum_{i=1}^d \LRp{\bm{A}_i\otimes \bm{D}_{i}}\bm{V}  +  \sum_{f\in \partial D^k}\LRp{\bm{I}\otimes\bm{L}_f}\bm{F}_{\sigma}}.
-\label{eq:wadgmat}
-\end{align*}
-In practice, the matrices $\LRp{\bm{I}\otimes\bm{M}^{-1}} \bm{M}_{\LRp{\rho^{-1}\bm{I}}}$ and $\LRp{\bm{I}\otimes\bm{M}^{-1}} \bm{M}_{\bm{C}}$ are applied in a matrix-free fashion using reference element matrices and values of $\rho,\bm{C}$ at quadrature points.
-After fusing operations together, this procedure can be boiled down to multiplication by two rectangular matrices \cite{chan2016weight2}.
-
-One drawback of the analysis presented in this work is that accuracy is not guaranteed in the isotropic incompressible limit $\mu / \lambda \rightarrow 0$.
-In this case, the stiffness matrix $\bm{C}$ becomes singular, and the constant $\tilde{w}_{\max}$ in the upper bound on $\nor{\bm{C}^{-1}}$ blows up.
-However, numerical experiments in Section~\ref{sec:incomp} suggest that, while taking $\mu/\lambda \approx 0$ (or $\bm{C}$ nearly singular) results in larger relative errors for $\bm{\sigma}$, the accuracy of the WADG solution for $\bm{v}$ does not degrade significantly for near-incompressible materials.
-
-\subsection{Energy stability on curvilinear meshes}
-
-We have shown the energy stability of the DG formulation (\ref{eq:dgform}) on meshes of affine elements by assuming exact integration of all terms.
-However, energy stability can still be guaranteed if integrals are evaluated inexactly using quadrature.
-Instead of discretizing the ``strong'' DG formulation (\ref{eq:dgform}), we discretize the ``skew-symmetric'' formulation, where the right hand side of the velocity equations is integrated by parts,\footnote{The choice of which equation to integrate by parts is arbitrary, since integrating the stress equations by parts also results in an energy stable formulation.
-} resulting in the formulation
-\begin{align*}
-\int_{\Dhat} T^{-1}_{(\rho J)^{-1}} \pd{\bm{v}}{t}\bm{w} &= -\int_{\Dhat}\LRp{\sum_{i=1}^d \bm{\sigma},\bm{A}_i \pd{\bm{w}}{\bm{x}_i}}J  + \sum_{f \in \partial D^k} \int_{\widehat{f}}\LRp{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}}\bm{w}J^f \nonumber\\
-\int_{\Dhat}  T^{-1}_{\LRp{J^{-1}\bm{C}}} \pd{\bm{\sigma}}{t}\bm{q} &= \int_{\Dhat} \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}J + \sum_{f\in \partial D^k}\int_{\widehat{f}}\LRp{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}}\bm{q}J^f
-%\LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= -\LRp{ \sum_{i=1}^d \bm{\sigma},\bm{A}_i \pd{\bm{w}}{\bm{x}_i}}_{\Lk}  + \LRa{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk} \nonumber\\
-%\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk}.
-\end{align*}
-where we have incorporated spatial variations of $J$ into the definition of the weights on the left hand side.
-
-Since the proof of stability in Theorem~\ref{thm:stability} requires only algebraic manipulations of this formulation, the same energy stability result holds for this formulation.
-More importantly, a discrete version of energy stability is still guaranteed in the presence of inexact quadrature, where integrals in Theorem~\ref{thm:stability} are replaced with quadrature approximations.
-This is especially important for discretizations on curvilinear meshes, where the exact integration of spatially varying geometric factors and Jacobians can be either inhibitively expensive for high order curvilinear mappings or impossible for rational mappings \cite{engvall2016isogeometric, michoski2015foundations}.
-
-We note that, to ensure energy stability on curved and non-affine elements, the ``skew-symmetric'' formulation must be evaluated explicitly using quadrature, which is typically more expensive than quadrature-free evaluations used to evaluate the ``strong'' DG formulation (\ref{eq:dgform}).
-These costs can be slightly reduced for curvilinear meshes by evaluating the DG formulation using the ``skew-symmetric'' formulation on curvilinear elements and the more efficient ``strong'' DG formulation on affine elements \cite{chan2016weight2}.
-
-\subsection{Convergence analysis}
-
-Using estimates from Section~\ref{sec:mwadgprop}, we can extend the semi-discrete convergence analysis in \cite{houston2002discontinuous,warburton2013low,chan2016weight1} to linear elastic wave propagation on meshes of affine elements.
-Techniques in \cite{warburton2013low} can be used to extend this analysis to curvilinear elements.
-
-In terms of group variables $\bm{U} = \LRp{\bm{v},\bm{\sigma}}$ and $\bm{V} = \LRp{\bm{w},\bm{q}} \in \LRp{V_h\LRp{\Oh}}^d\times\LRp{V_h\LRp{\Oh}}^{N_d}$, the variational formulation can be written as
-\begin{align*}
-&\LRp{T^{-1}_{\bm{A}_0^{-1}}\pd{\bm{U}}{t},\bm{V}}_{\Omega} + a(\bm{U},\bm{V}) + b(\bm{U},\bm{V}) = (\bm{f},\bm{V})\\
-a(\bm{U},\bm{V}) &= \sum_{D^k\in \Oh} \LRp{-\LRp{ \sum_{i=1}^d \bm{\sigma},\bm{A}_i \pd{\bm{w}}{\bm{x}_i}}_{\Lk}  + \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk}} \\
-b(\bm{U},\bm{V}) &= \sum_{D^k\in \Oh}\LRp{ \LRa{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk} },
-%a(\bm{U},\bm{V}) &= \sum_{D^k\in \Oh} \LRp{\LRp{ \sum_{i=1}^d \bm{A}_i^T\pd{\bm{\sigma}}{\bm{x}_i},\bm{w}}_{\Lk}  + \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk}} \\
-%b(\bm{U},\bm{V}) &= \sum_{D^k\in \Oh}\LRp{ \LRa{\bm{A}_n^T\jump{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk} },
-\end{align*}
-where $\bm{A}_0 = {\rm diag}\LRp{\rho\bm{I},\bm{C}^{-1}}$.  Energy stability implies that
-\begin{align*}
-&b(\bm{U},\bm{U}) = \frac{1}{2}\sum_{D^k\in\Oh} \LRp{\tau_v\nor{\jump{\bm{A}_n\bm{v}}}_{\Ldk}^2 + \tau_\sigma\nor{\jump{\bm{A}^T_n\bm{\sigma}}}_{\Ldk}^2}\\
-&\LRp{T^{-1}_{\bm{A}_0^{-1}}\pd{\bm{U}}{t},\bm{U}}_{\Omega} + b(\bm{U},\bm{U}) =  (\bm{f},\bm{U}).
-\end{align*}
-Let $\bm{U},\bm{U}_h$ denote the exact and discrete WADG solutions, respectively.  Since the DG formulation (\ref{eq:dgform}) is consistent, these solutions satisfy
-\begin{align}
-\LRp{\bm{A}_0\pd{\bm{U}}{t},\bm{V}} + a(\bm{U},\bm{V}) + b(\bm{U},\bm{V}) &= (\bm{f},\bm{V}) \nonumber\\
-\LRp{T_{\bm{A}_0^{-1}}^{-1}\pd{\bm{U}_h}{t},\bm{V}} + a(\bm{U}_h,\bm{V}) + b(\bm{U}_h,\bm{V}) &= (\bm{f},\bm{V})
-\label{eq:dgvswadg}
-\end{align}
-for all $\bm{V} \in \LRp{V_h\LRp{\Oh}}^d\times\LRp{V_h\LRp{\Oh}}^{N_d}$.  We decompose the error $\bm{U}-\bm{U}_h$ into a projection error $\bm{\epsilon}$ and discretization error $\bm{\eta}$.
-\[
-\bm{U}-\bm{U}_h  = \LRp{\Pi_N\bm{U}-\bm{U}_h}-\LRp{\Pi_N\bm{U}-\bm{U}} = \bm{\eta} - \bm{\epsilon}.
-\]
-We also introduce a consistency error $\bm{\delta} = \bm{A}_0\bm{U} - T_{\bm{A}_0^{-1}}^{-1}\bm{U}$ resulting from the approximation of $\bm{A}_0\bm{U}$ by a weight-adjusted inner product
-\[
-\bm{A}_0\pd{\bm{U}}{t} - T_{\bm{A}_0^{-1}}^{-1}\pd{\bm{U}_h}{t} = \pd{}{t}\LRp{\bm{A}_0\bm{U} - T_{\bm{A}_0^{-1}}^{-1}\bm{U}} + \pd{}{t}T_{\bm{A}_0^{-1}}^{-1}\LRp{\Pi_N\bm{U}-\bm{U}_h} = \pd{\bm{\delta}}{t} + \pd{}{t}\LRp{T_{\bm{A}_0^{-1}}^{-1}\bm{\eta}}
-\]  
-where we have used that $\bm{T}_{\bm{A}_0^{-1}}^{-1} = \bm{T}_{\bm{A}_0^{-1}}^{-1}\Pi_N$.  Subtracting the DG and WADG formulations in (\ref{eq:dgvswadg}) and setting $\bm{V} = \bm{\eta}$ then yields 
-\begin{align}
-\frac{1}{2}\pd{}{t}\LRp{T^{-1}_{\bm{A}_0^{-1}}\bm{\eta},\bm{\eta}}_{\Omega} + b(\bm{\eta},\bm{\eta}) &= \LRp{\pd{\bm{\delta}}{t},\bm{\eta}} + a(\bm{\epsilon},\bm{\eta}) + b(\bm{\epsilon},\bm{\eta}),
-\label{eq:erreqn}
-\end{align}
-where we have used that $a(\bm{\eta},\bm{\eta}) = 0$ by skew symmetry.
-
-We bound $a(\bm{\epsilon},\bm{\eta}) + b(\bm{\epsilon},\bm{\eta})$ on the right hand side by integrating by parts the stress equation and using the component-wise $L^2$ orthogonality of $\bm{\epsilon}$ to derivatives of $\bm{\eta}$.
-This reduces the term $a(\bm{\epsilon},\bm{\eta}) + b(\bm{v},\bm{\eta})$ to surface contributions over each element, which can be combined with contributions from neighboring elements
-\begin{align*}
- \sum_{D^k\in \Oh}&\LRp{ \LRa{\bm{A}_n^T\avg{\bm{\epsilon}_\sigma} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{\epsilon}_v},\bm{\eta}_v}_{\Ldk} + \LRa{\bm{A}_n\avg{\bm{\epsilon}_v} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\epsilon}_\sigma},\bm{\eta}_\sigma}_{\Ldk}}\\
-&= \sum_{D^k\in \Oh}\LRp{ 
-\LRa{\avg{\bm{\epsilon}_\sigma} - \frac{\tau_v}{2}\bm{A}_n\jump{\bm{\epsilon}_v},\bm{A}_n\jump{\bm{\eta}_v}}_{\Ldk} + \LRa{\avg{\bm{\epsilon}_v} - \frac{\tau_\sigma}{2}\bm{A}_n^T\jump{\bm{\epsilon}_\sigma},\bm{A}_n^T\jump{\bm{\eta}_{\sigma}}}_{\Ldk}}\\ %- \LRa{\bm{A}_n^T\jump{\bm{\epsilon}_\sigma},\frac{\tau_{\sigma}}{2}\bm{A}_n^T\jump{\bm{\eta}}_\sigma}_{\Ldk} }\\
-&\leq \sum_{D^k\in \Oh}\nor{\avg{\bm{\epsilon}_\sigma} - \frac{\tau_v}{2}\bm{A}_n\jump{\bm{\epsilon}_v}}_{\Ldk} \nor{\bm{A}_n\jump{\bm{\eta}_v}}_{\Ldk} + \nor{\avg{\bm{\epsilon}_v} - \frac{\tau_\sigma}{2}\bm{A}_n^T\jump{\bm{\epsilon}_\sigma}}_{\Ldk}\nor{\bm{A}_n^T\jump{\bm{\eta}_{\sigma}}}_{\Ldk}\\
-%&\leq \LRp{\nor{\avg{\bm{\epsilon}_\sigma} - \frac{\tau_v}{2}\bm{A}_n\jump{\bm{\epsilon}_v}}_{\Ldk}  + \nor{\avg{\bm{\epsilon}_v} - \frac{\tau_\sigma}{2}\bm{A}_n^T\jump{\bm{\epsilon}_\sigma}}_{\LDk}} +  \LRp{\nor{\bm{A}_n\jump{\bm{\eta}_v}}_{\Ldk}^2 + \nor{\bm{A}_n^T\jump{\bm{\eta}_{\sigma}}}_{\Ldk}^2}\\
-&\leq C_\tau \sum_{D^k\in \Oh}\nor{\bm{\epsilon}}_{\Ldk}\frac{1}{2}\LRp{\tau_v\nor{\bm{A}_n\jump{\bm{\eta}_v}}_{\Ldk}^2 + \tau_\sigma\nor{\bm{A}_n^T\jump{\bm{\eta}_{\sigma}}}_{\Ldk}^2}^{1/2}
-\end{align*}
-where $C_\tau$ is proportional to $\max\LRp{\tau_v, \tau_\sigma}$.  Using Young's inequality with $\alpha \geq 1/\max\LRp{\tau_v,\tau_\sigma,1}$ sufficiently large yields the following bound 
-\[
-\LRb{a(\bm{\epsilon},\bm{\eta}) + b(\bm{\epsilon},\bm{\eta})} \leq \frac{C}{4\alpha}{b(\bm{\eta},\bm{\eta})} + \frac{C\alpha}{4} \sum_{D^k\in \Oh}\nor{\bm{\epsilon}}_{\Ldk}^2.
-\]
-%where $\bm{E}_v,\bm{E}_\sigma,\bm{\eta}_v,\bm{\eta}_\sigma$ denote respective components of $\bm{E},\bm{\eta}$.  
-This can be used to bound (\ref{eq:erreqn}).  Using Young's inequality on the remaining terms and absorbing terms involving $\bm{\eta}$ into the left hand side yields
-\[
-\frac{1}{2}\pd{}{t}\LRp{T^{-1}_{\bm{A}_0^{-1}}\bm{\eta},\bm{\eta}}_{\Omega} + \LRp{1-\frac{C}{4\alpha}} b(\bm{\eta},\bm{\eta}) \leq \nor{\pd{\bm{\delta}}{t}}_{\Lk} + \sum_{D^k \in \Oh}\frac{ C\alpha}{4} \nor{\bm{\epsilon}}_{\Ldk}^2.
-\]
-Integrating from $[0,T]$, and using the arithmetic-geometric mean inequality yields
-\[
-\frac{1}{2}\LRp{T^{-1}_{\bm{A}_0^{-1}}\bm{\eta},\bm{\eta}}_{\Omega} + \LRp{1-\frac{C}{4\alpha}}\int_0^T  b(\bm{\eta},\bm{\eta}) \leq \int_0^T\nor{\pd{\bm{\delta}}{t}}_{\Lk} + \int_0^T\sum_{D^k \in \Oh}C\nor{\bm{\epsilon}}_{\Ldk}.
-\]
-Using Theorem~\ref{thm:west} and a standard $hp$ trace inequality $\nor{u}_{\Ldk} \leq Ch^{-1/2} \nor{u}_{\L}$ \cite{chan2015gpu} then yields
-\[
-\nor{\bm{\eta}} \leq C_{\bm{W}} T \sup_{t\in [0,T]}\LRp{\nor{\pd{\bm{\delta}}{t}} + \nor{\bm{\epsilon}}h^{-1/2}}.
-\]
-Bounding each term on the right using interpolation estimates for sufficiently regular $\bm{U}$ yields the final result
-\[
-\nor{\bm{U}-\bm{U}_h} \leq (C_1 + C_2 T) h^{N + 1/2}\LRp{\nor{\bm{U}}_{W^{N+1,2}} + C_3 h^{1/2}\nor{\pd{\bm{U}}{t}}_{W^{N+1,2}}}.
-\]
-where $C_2$ depends on $N$ and $C_3$ depends on $\nor{\bm{W}}_{{N+1,2,\infty}}$.
-\note{Todo - finish}
-
-\section{Numerical experiments}
-\label{sec:numerical}
-The following sections present several numerical experiments validating the stability and accuracy of the proposed method in two and three dimensions.
-The energy stability of the method is also confirmed for examples with sub-cell variations in heterogeneous media and curvilinear meshes.
-The convergence of the new DG formulation in piecewise constant isotropic media is confirmed using analytic solutions, while the convergence of the method for high order approximations of heterogeneous media is confirmed using a fine grid reference solution.
-Finally, the method is applied to problems with anisotropy and stiffness matrices $\bm{C}$ with sub-element variations.
-
-In all experiments, we follow \cite{chan2016weight1} and compute application of weight-adjusted mass matrices using a quadrature exact for polynomials of degree $(2N+1)$.
-Time integration is performed using the low-storage 4th order five-stage Runge-Kutta scheme of Carpenter and Kennedy \cite{carpenter1994fourth}, and the timestep is chosen based on the global restriction
-\[
-dt = \min_{f} \frac{C_{\rm CFL}}{  \nor{\bm{C}(\bm{x})}_{2,\infty,\Omega} C_N \nor{J^f}_{L^{\infty}(f)}\nor{J^{-1}}_{L^{\infty}(f)}}
-\]
-where $C_N = O(N^2)$ is the order-dependent constant in the surface polynomial trace inequality \cite{chan2015gpu}, $C_{\rm CFL}$ is a tunable global CFL constant, and $\nor{\bm{C}}_{2,\infty,\Omega}$ denotes the spatial supremum of the $2$-norm of the stiffness matrix $\bm{C}(\bm{x})$.
-\[
-\nor{\bm{C}}_{2,\infty,\Omega} = \sup_{\bm{x}\in\Omega}\nor{\bm{C}(\bm{x})}_{2}.  
-\]
-
-\subsection{Spectra and choice of penalty parameter}
-
-We first verify the energy stability of the WADG method for arbitrary heterogeneous media.
-We use a stiffness matrix constructed using random numbers and similarity transforms, such that at every quadrature point, $\bm{C}(\bm{x}) = \bm{U}\bm{D}\bm{U}^{T}$, where $\bm{D}$ is a diagonal matrix with positive entries $d_{\min} \leq \bm{D}_{ii} \leq d_{\max}$ and $\bm{U}$ is a random unitary matrix.
-Let $\bm{A}_h$ denote the matrix induced by the global semi-discrete DG formulation, such that the time evolution of the solution $\bm{v},\bm{\sigma}$ is governed by
-\[
-\pd{\bm{Q}}{t} = \bm{A}_h\bm{Q},
-\]
-with $\bm{Q}$ denotes a vector of degrees of freedom for $(\bm{v},\bm{\sigma})$.
-Figure~\ref{fig:spectra} shows computed eigenvalues of $\bm{A}_h$ for $[d_{\min},d_{\max}] = [1/10,1]$ and $[10^{-4},1]$, $\tau_v=\tau_\sigma = 0$ and $\tau_v=\tau_\sigma = 1$ under discretization parameters $N=4$ and $h = 1/4$.
-
-\begin{figure}
-\centering
-\subfloat[$\tau_v = \tau_\sigma = 0$]{\includegraphics[width=.475\textwidth]{../figs/spectra_tau0.png}}
-%\hspace{1em}
-\subfloat[$\tau_v = \tau_\sigma = 1$]{\includegraphics[width=.475\textwidth]{../figs/spectra_tau1.png}}
-\caption{Spectra for $N=4$ and $h = 1/4$, using randomly chosen $\bm{C}(\bm{x})$ with eigenvalues between $[d_{\min}, 1]$ at each quadrature point.
-For both $\tau_v = \tau_\sigma = 0$ and $\tau_v = \tau_\sigma = 1$, the largest real part of the spectra is $O(10^{-14})$.}
-\label{fig:spectra}
-\end{figure}
-
-We note that, for $\tau_v = \tau_\sigma = 1$, $\nor{\bm{A}_h}$ is approximately twice as large compared to when $\tau_v = \tau_\sigma = 0$.
-This can result in a more restrictive stability condition on the timestep.
-We observe that the growth in $\nor{\bm{A}_h}$ for $\tau_v,\tau_\sigma = 1$ is due to the presence of eigenvalues with large negative real parts, which implies that larger timesteps may be used under time integrators with extended stability regions along the real axis \cite{abdulle2002fourth}.
-
-Alternatively, we can avoid growth in $\nor{\bm{A}_h}$ by scaling $\tau$.
-To avoid artificially increasing the stiffness of (\ref{eq:wadgmat}), it was argued in \cite{chan2016short} that the magnitude of the penalty parameter should depend on the largest singular value of $\bm{A}_n$, which is $O(1)$ in two and three dimensions.
-However, this does not factor in the effect of $\rho$ and $\bm{C}^{-1}$ appearing in the left hand side.
-These coefficients can be taken into account by scaling only the stress penalization parameter with the average of local max norms over a face $f$
-\[
-\tau_{\sigma,f} \approx \frac{2}{\nor{\bm{C}^+(\bm{x})}_{2,\infty,f}+\nor{\bm{C}^-(\bm{x})}_{2,\infty,f}},
-\]
-where $\bm{C}^+,\bm{C}^-$ denote values of the stiffness matrix on elements which share the face $f$.
-It may also be advantageous to rescale $\tau_\sigma$ for each individual component of stress based on the rows of $\bm{C}$.
-Under this rescaling of the stress penalty parameter, we observe numerically that the norm of $\bm{A}_h$ remains roughly the same magnitude as for $\tau_p = \tau_\sigma = 0$.
-%Since we do not observe significant changes in computed $L^2$, we adopt this rescaling of $\tau_\sigma$ in all reported results.
-%We note that this rescaling of the penalty parameters also avoids other issues associated with over-penalization, such as the reintroduction of spurious modes which would otherwise be damped under less aggressive penalties \cite{chan2016short}.
-
-In the following experiments, we use $\tau_v = \tau_\sigma = 1$.  However, we do not observe significant changes in computed $L^2$ errors under the aforementioned rescaling of $\tau_\sigma$.
-
-\subsection{Analytic solutions}
-
-Next, we study the accuracy and convergence of weight-adjusted DG method for several analytical solutions in linear elasticity.
-In all cases, the solution in expressed terms of the displacement vector $\bm{u}(\bm{x},t) = (u_1,\ldots,u_d)$.
-Initial conditions for velocity and stress are computed through
-\[
-\bm{v}(\bm{x},t) = \pd{\bm{u}}{t}, \qquad \bm{\sigma} = \bm{C}\bm{E} = \bm{C}\frac{1}{2}\LRp{\Grad\bm{u} + \Grad\bm{u}^T}.  
-\]
-Unless otherwise stated, we report relative $L^2$ errors for all components of the solution $\bm{U} = (\bm{v},\bm{\sigma})$ 
-\[
-\frac{\nor{\bm{U}-\bm{U}_h}_{L^2\LRp{\Omega}}}{\nor{\bm{U}}_{L^2\LRp{\Omega}}} = \frac{\LRp{\sum_{i=1}^m \nor{\bm{U}_i-\bm{U}_{i,h}}^2_{L^2\LRp{\Omega}}}^{1/2}}{\LRp{\sum_{i=1}^m\nor{\bm{U}_i}^2_{L^2\LRp{\Omega}}}^{1/2}}.
-\]
-
-\subsubsection{Harmonic oscillation of a square}
-
-We first examine convergence  on a unit square domain with  $\lambda = \mu = \rho = 1$.   The components of the displacement vector is given by
-\begin{align*}
-u_1(x,y,t) &= \cos(\omega\pi t)\cos(\pi x)\sin(\pi y)\\
-u_2(x,y,t) &= -\cos(\omega\pi t)\sin(\pi x)\cos(\pi y),
-\end{align*}
-where $\omega = \sqrt{2\mu}$.
-Figure~\ref{fig:harmonic_osc} shows $L^2$ errors computed at time $T = 1$, using uniform triangular meshes constructed by bisecting a uniform mesh of quadrilaterals along the diagonal.
-For $N = 1,\ldots, 5$, optimal $O(h^{N+1})$ rates of convergence are observed.
-
-\begin{figure}
-\centering
-\begin{tikzpicture}
-\begin{loglogaxis}[
-	legend cell align=left,
-	width=.49\textwidth,
-    xlabel={Mesh size $h$},
-       ylabel={$L^2$ errors}, 
-    xmin=.025, xmax=1,
-    ymin=1e-12, ymax=1,
-    legend pos=south east,
-    xmajorgrids=true,
-    ymajorgrids=true,
-    grid style=dashed,
-]
-
-\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.162198)(0.25,0.0389974)(0.125,0.0095562)(0.0625,0.00227802)};
-\logLogSlopeTriangleFlip{0.4}{0.125}{0.81}{2.05}{blue}
-
-\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.0128844)(0.25,0.00156487)(0.125,0.000197047)(0.0625,2.42153e-05)};
-\logLogSlopeTriangleFlip{0.4}{0.125}{0.65}{3.01}{red}
-
-\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.00110186)(0.25,6.9321e-05)(0.125,4.31386e-06)(0.0625,2.49229e-07)};
-\logLogSlopeTriangleFlip{0.4}{0.125}{0.49}{4.06}{black}
-
-\addplot[color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,8.42986e-05)(0.25,2.47716e-06)(0.125,7.21337e-08)(0.0625,2.17472e-09)};
-\logLogSlopeTriangleFlip{0.4}{0.125}{0.325}{5.08}{magenta}
-
-\addplot[color=violet,mark=pentagon*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,4.91233e-06)(0.25,7.57947e-08)(0.125,1.10789e-09)(0.0625,1.63538e-11)};
-\logLogSlopeTriangleFlip{0.4}{0.125}{0.15}{6.09}{violet}
-
-\legend{$N=1$,$N=2$,$N=3$,$N=4$,$N=5$}
-\end{loglogaxis}
-\end{tikzpicture}
-\caption{Convergence of $L^2$ errors for harmonic oscillation.}
-\label{fig:harmonic_osc}
-\end{figure}
-
-\subsubsection{Rayleigh and Lamb waves}
-
-Next, we examine the convergence of WADG for Rayleigh and Lamb waves, both of which test the imposition of traction-free boundary conditions.
-
-Rayleigh waves are elastic surface waves which decay exponentially away from the surface.  These waves are given by the displacement vector
-\begin{align*}
-\bm{u}\LRp{x,y,t} &= e^{-\omega x \sqrt{1-\xi^2}}
-\LRp{
-\begin{array}{c}
-\cos(\omega(y+c_r t)) \\
-\sqrt{1-\xi^2}\sin(\omega(y+c_r t))
-\end{array}
-}\\
-&+ \LRp{\frac{\xi^2}{2}-1}e^{-\omega x\sqrt{1-\frac{\xi^2\mu}{2\mu + \lambda}}}
-\LRp{
-\begin{array}{c}
-\cos(\omega(y + c_r t))\\
-{\sin(\omega(y+c_r t))}/{\sqrt{1-\frac{\xi^2\mu}{2\mu + \lambda}}}
-\end{array}
-},
-\end{align*}
-where $\omega$ is the wavespeed, $c_r$ is the Rayleigh phase velocity $c_r = \xi \sqrt{\mu}$, and $\xi$ satisfies
-\[
-\sqrt{1-\xi^2}\sqrt{1 - \frac{\xi^2\mu}{2\mu+\lambda}} - \LRp{\frac{\xi^2}{2}-1}^2 = 0.
-\]
-In our computations, we use $\rho = \mu = \lambda = 1$, $\xi = 0.949554083888034$, and $\omega = 2\pi$.
-We solve on the domain $[0,2]\times[0,1]$ using a sequence of uniform triangular meshes, and enforce traction-free boundary conditions at $x=0$ and exact Dirichlet boundary conditions at $x=2$.
-Periodic boundary conditions are applied at $y = 0$ and $y = 1$.
-
-Lamb waves are supported by elastic waveguides with traction-free (free surface) boundary conditions at the top and bottom of the domain.  The displacement of these waves is given by
-\begin{align*}
-{u}_1\LRp{x,y,t} &= \LRp{-kB_1 \cos(p y)- qB_2 \cos(q y)} \sin(k x - \omega t)\\
-{u}_2\LRp{x,y,t} &= \LRp{-pB_1 \sin(py) + kB_2 \sin(qy)} \cos(kx - \omega t)
-\end{align*}
-where $k$ is the wavenumber and $\omega$ is the frequency, and the constants $p$ and $q$ are defined as
-\[
-p^2 = \frac{\omega^2}{2\mu + \lambda} - k^2, \qquad q^2 = \frac{\omega^2}{\mu} - k^2.
-\]
-The wavenumber $k$ and frequency $\omega$ are related through a dispersion relation.
-The ratio of the amplitudes $B_1/ B_2$ can be determined using other parameters, implying that $B_1, B_2$ are unique up to a scaling constant.
-In our experiments, we use $\rho = \mu = 1$, $\lambda = 2$, $k = 2\pi$.
-For these values, $\omega = 13.137063197233$, $B_1 = 126.1992721468$ and $B_2 = 53.88807700007$.
-We solve on the domain $[-1,1]\times[-1/2,1/2]$, with traction-free boundary conditions at $y = \pm 1/2$ and periodic boundary conditions at $x = \pm 1$.
-
-Figure~\ref{fig:rayleigh} shows $L^2$ errors for both waves at time $T = 1$.
-Computed convergence rates fall between the optimal rate of $O(h^{N+1})$ and theoretical rate of $O(h^{N+1/2})$ \cite{johnson1986analysis}.
-
-\begin{figure}
-\centering
-\subfloat[Rayleigh wave]{
-\begin{tikzpicture}
-\begin{loglogaxis}[
-    legend cell align=left,
-    width=.475\textwidth,
-    xlabel={Mesh size $h$},
-    ylabel={$L^2$ errors}, 
-    xmin=.025, xmax=1,
-    ymin=1e-8, ymax=1.5,
-    legend pos=south east,
-    xmajorgrids=true,
-    ymajorgrids=true,
-    grid style=dashed,
-]
-
-\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,1.04483)(0.25,0.342112)(0.125,0.102275)(0.0625,0.0277191)};    
-\logLogSlopeTriangleFlip{.39}{0.11}{0.815}{1.81}{blue}
-
-\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.287061)(0.25,0.0563744)(0.125,0.00785962)(0.0625,0.000992348)};    
-\logLogSlopeTriangleFlip{.38}{0.1}{0.64}{2.91}{red}
-
-\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.0736249)(0.25,0.00747192)(0.125,0.000529189)(0.0625,3.53252e-05)};   
-\logLogSlopeTriangleFlip{.38}{0.1}{0.475}{3.86}{black}
-
-\addplot[color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.0199958)(0.25,0.000774494)(0.125,2.70714e-05)(0.0625,8.78659e-07)};   
-\logLogSlopeTriangleFlip{.38}{0.1}{0.285}{4.89}{magenta}
-
-\addplot[color=violet,mark=pentagon*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.00304235)(0.25,6.78379e-05)(0.125,1.18747e-06)(0.0625,1.92431e-08)};    
-\logLogSlopeTriangleFlip{.38}{0.1}{0.09}{5.89}{violet}
-
-%\node at (axis cs:.03,6.8e-10) {$u$ discontinuous};
-\legend{$N=1$,$N=2$,$N=3$,$N=4$,$N=5$}
-\end{loglogaxis}
-\end{tikzpicture}
-}
-\subfloat[Lamb wave]{
-\begin{tikzpicture}
-\begin{loglogaxis}[
-	legend cell align=left,
-	width=.475\textwidth,
-    xlabel={Mesh size $h$},
-%       ylabel={$L^2$ errors}, 
-    xmin=.025, xmax=1,
-    ymin=2e-7, ymax=1.5,
-    legend pos=south east,
-    xmajorgrids=true,
-    ymajorgrids=true,
-    grid style=dashed,
-]
-
-\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.97515)(0.25,0.667267)(0.125,0.26441)(0.0625,0.0643182)};   
-\logLogSlopeTriangleFlip{.39}{0.11}{0.835}{2.04}{blue}
-
-\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.659983)(0.25,0.191328)(0.125,0.0172492)(0.0625,0.00168533)};   
-\logLogSlopeTriangleFlip{.38}{0.1}{0.62}{3.41}{red}
-
-\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.366154)(0.25,0.024884)(0.125,0.0014833)(0.0625,0.000103456)};    
-\logLogSlopeTriangleFlip{.38}{0.1}{0.45}{3.96}{black}
-\addplot[color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.141902)(0.25,0.00437419)(0.125,0.000154364)(0.0625,5.20702e-06)};   
-\logLogSlopeTriangleFlip{.38}{0.1}{0.26}{4.86}{magenta}
-
-\addplot[color=violet,mark=pentagon*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.0327909)(0.25,0.000752122)(0.125,1.34499e-05)(0.0625,2.18614e-07)};   
-\logLogSlopeTriangleFlip{0.38}{0.1}{0.075}{5.87}{violet}
-
-\legend{$N=1$,$N=2$,$N=3$,$N=4$,$N=5$}
-\end{loglogaxis}
-\end{tikzpicture}
-\label{fig:rayleigh}
-}
-\caption{Convergence of $L^2$ errors for Rayleigh and Lamb wave solutions.}
-\end{figure}
-
-\subsubsection{Rayleigh waves in near-incompressible materials}
-\label{sec:incomp}
-
-As noted in Section~\ref{sec:wadgelas}, error estimates for isotropic elasticity no longer hold in the incompressible limit with $\mu / \lambda \rightarrow 0$ limit due to the fact that $\bm{C}$ becomes singular.
-We use the propagation of Rayleigh waves to examine the behavior of WADG for near-incompressible materials.
-We follow \cite{sjogreen2012fourth,appelo2015energy} and fix $\lambda = 1$ and set $\mu = 1, .1, .01, .001, .0001$.
-Since the Rayleigh wave propagates with speed proportional to $\sqrt{\mu}$, we compute $L^2$ errors at the final time $1 / (4\sqrt{\mu})$ to ensure a fair comparison between solutions at different values of $\mu$.
-
-%Table~\ref{tab:incomp} shows total relative errors and relative errors for $\bm{v}$, $\bm{\sigma}$ individually at different orders and mesh sizes.
-While the relative errors for $\bm{\sigma}$ grow as $\mu/\lambda \rightarrow 0$, total errors and errors for $\bm{v}$ remain roughly the same magnitude for all values of $\mu$.
-Table~\ref{tab:incomp} shows relative errors for $\bm{v}$, $\bm{\sigma}$ at different orders and mesh sizes.
-The relative errors for $\bm{\sigma}$ grow as $\mu/\lambda \rightarrow 0$.
-This is not surprising, as the constant in the error estimates of Section~\ref{sec:mwadgprop} blows up in the incompressible limit.
-However, because the magnitude of $\bm{\sigma}$ also decreases as $\mu/\lambda\rightarrow 0$, relative errors for $\bm{v}$ remain roughly the same magnitude as the material approaches incompressibility.
-Decreasing $\mu$ by four orders of magnitude results in up to a ten-fold increase in relative error for $\bm{\sigma}$, but less than a two-fold increase in error for $\bm{v}$.
-
-\begin{table}
-\centering
-\subfloat[Relative error in $\bm{v}$]{
-\begin{tabular}{|c||c|c|c|c|c|}
-\hline
- & $\mu = 1$ & $\mu = .1$ & $\mu = .01$ & $\mu = .001$ & $\mu = .0001$  \\
-\hhline{|=|=|=|=|=|=|}
-$N = 2$, $h = 1/4$ &  3.1063e-02  &  2.6972e-02  & 3.0858e-02 &  4.0103e-02  & 5.6835e-02\\
-\hline
-$N = 3$, $h = 1/4$ & 3.1677e-03 &  2.6848e-03 &  2.9854e-03 &  3.6960e-03 &  4.9362e-03\\
-\hline
-$N = 4$, $h = 1/4$ & 2.8726e-04 &  2.4990e-04 &  2.9142e-04 &  3.9774e-04 &  5.2469e-04\\
-
-\hhline{|=|=|=|=|=|=|}
-$N = 2$, $h = 1/8$ & 3.1819e-03  & 2.5476e-03 &  2.8877e-03 &  3.5608e-03  & 4.7010e-03 \\
-\hline
-$N = 3$, $h = 1/8$ & 1.8867e-04  & 1.6925e-04  & 1.8509e-04  & 2.2520e-04  & 2.8301e-04\\
-\hline
-$N = 4$, $h = 1/8$ & 8.4999e-06  & 7.5750e-06 &  8.1094e-06  & 1.0760e-05  & 1.4886e-05\\
-\hline         
-\end{tabular}
-}\\
-\subfloat[Relative error in $\bm{\sigma}$]{
-\begin{tabular}{|c||c|c|c|c|c|}
-\hline
- & $\mu = 1$ & $\mu = .1$ & $\mu = .01$ & $\mu = .001$& $\mu = .0001$ \\
-\hhline{|=|=|=|=|=|=|}
-$N = 2$, $h = 1/4$ & 6.8250e-02  & 7.4130e-02 &  1.2104e-01 &  2.1333e-01 &  4.1451e-01\\
-\hline
-$N = 3$, $h = 1/4$ & 9.2980e-03  & 1.0685e-02 &  1.8109e-02 &  3.2046e-02 &  5.7612e-02\\
-\hline
-$N = 4$, $h = 1/4$ & 9.7251e-04  & 1.1687e-03  & 2.1941e-03 &  4.1378e-03 &  6.8404e-03\\
-\hhline{|=|=|=|=|=|=|}
-$N = 2$, $h = 1/8$ & 9.8138e-03  & 1.1429e-02  & 2.0331e-02 &  3.7889e-02 & 7.6516e-02\\
-\hline
-$N = 3$, $h = 1/8$ & 6.6596e-04  & 7.8157e-04  & 1.5353e-03  & 3.1239e-03 &  5.6341e-03\\
-\hline
-$N = 4$, $h = 1/8$ & 3.4272e-05  & 4.2639e-05 &  8.8393e-05  & 1.9764e-04 &  3.6052e-04 \\
-\hline
-\end{tabular}
-}
-%\\
-%\subfloat[Total relative error ]{
-%\begin{tabular}{|c||c|c|c|c|c|}
-%\hline
-% & $\mu = 1$ & $\mu = .1$ & $\mu = .01$ & $\mu = .001$ & \mu = .0001$ \\
-%\hhline{|=|=|=|=|=|=|}
-%$N = 2$, $h = 1/4$ & 0.0539 & 0.0369  &  0.0339 & 0.0409 & 0.0571\\
-%\hline
-%$N = 3$, $h = 1/4$ & 0.0073 & 0.0046 &0.0037 &0.0039 & 0.0050\\
-%\hline
-%$N = 4$, $h = 1/4$ & 7.7419e-04 & 4.8735e-04 & 3.9193e-04 &4.2768e-04 &5.3111e-04\\
-%\hhline{|=|=|=|=|=|=|}
-%$N = 2$, $h = 1/8$ & 0.0074 & 0.0048  & 0.0038 &  0.0038 & 0.0048  \\
-%\hline
-%$N = 3$, $h = 1/8$ & 5.4120e-04 & 3.2760e-04 &2.6070e-04 & 2.5465e-04 & 2.9104e-04\\
-%\hline
-%$N = 4$, $h = 1/8$ & 2.7006e-05 & 1.7103e-05 &1.3344e-05 &1.3133e-05&1.5507e-05\\
-%\hline
-%\end{tabular}
-%}
-\caption{Behavior of WADG for linear elastic wave propagation in the incompressible limit as $\mu / \lambda \rightarrow 0$.  Errors are shown for various orders and mesh resolutions.  }
-\label{tab:incomp}
-\end{table}
-
-%\note{Examine energy stability as $\mu/\lambda\rightarrow 0$.  Matrix $\bm{C}$ becomes singular.}
-
-\subsubsection{Stoneley waves}
-
-A Stoneley wave is supported along the interface between two solids \cite{stoneley1924elastic}.
-Like Rayleigh waves, Stoneley waves decay exponentially away from the interface, and test the effectiveness of numerical fluxes across interfaces.
-We follow \cite{wilcox2010high,appelo2015energy} and use discontinuous media defined by
-\[
-(\rho,\lambda,\mu) = \begin{cases}
-(10 ,3,3) & y > 0\\
-(1,1,1) & y < 0.
-\end{cases}.
-\]
-The displacement vector for a Stoneley wave is then given by
-\begin{align*}
-u_1(x,y,t) &= \begin{cases}
-{\rm Re}\LRp{\LRp{ikB_1e^{-kb_{1p}y} + kb_{1s}B_2 e^{-kb_{1s}y}}e^{i(ky-\omega t)}}, & y > 0\\
-{\rm Re}\LRp{\LRp{-k b_{1p}B_1e^{-k b_{1p} y} + i kB_2e^{-kb_{1s}y}}e^{i(kx-\omega t)}}, & y < 0
-\end{cases}\\
-u_2(x,y,t) &= \begin{cases}
-{\rm Re}\LRp{\LRp{ikB_3 e^{k b_{2p} y} - kb_{2s}B_4 e^{k b_{2s}y }} e^{i(kx-\omega t)}}, & y > 0\\
-{\rm Re}\LRp{\LRp{k b_{2p} B_3 e^{k b_{2p}y} + ikB_4 e^{k b_{2s}y}}e^{i(kx-\omega t)}}, & y < 0
-\end{cases},
-\end{align*}
-where $c_{st}$ is the Stoneley wave speed, and 
-\[
-k = \omega/ c_{st}, \qquad b_{jp} = \sqrt{1 - \frac{c_{st}^2}{(2\mu_j + \lambda_j)/\rho_j}}, \qquad  b_{js} = \sqrt{1 - \frac{c_{st}^2}{(\mu_j)/\rho_j}}, \qquad j = 1,2.
-\]
-The Stoneley wave speed $c_{st}$ can be determined based material parameters and interface conditions, and the amplitudes $B_1,B_2,B_3,B_4$ are determined from $c_{st}$ up to scaling by a constant.
-For the parameters used in this study, we take $c_{st} = 0.546981324213884$, $B_1 = i0.2952173626624, B_2 = -0.6798795208473, B_3 = i0.5220044931212$, and $B_4 = -0.9339639688697$.
-We assume $k =1$, which gives $\omega=c_{st}$.
-
-We solve on the domain $[-1,1]\times [-5,5]$, and enforce Dirichlet boundary conditions at all boundaries using the exact solution.
-Figure~\ref{fig:stoneley} shows $L^2$ errors for two uniform meshes of triangles constructed by bisecting a quadrilateral mesh of $K_{\rm 1D} \times 5K_{\rm 1D}$ elements.
-Figure~\ref{subfig:stoneley1} shows errors when $K_{\rm 1D}$ is even and the mesh is fitted to the interface at $y = 0$, while Figure~\ref{subfig:stoneley2} shows errors when $K_{\rm 1D}$ is odd and the interface cuts through element interiors.
-When the mesh is fitted to the interface, computed convergence rates are between $O(h^{N+1/2})$ and $O(h^{N+1})$.
-
-When the mesh is not fitted to the interface exactly, we compute the application of the weight-adjusted mass matrix inverse using a quadrature rule from Xiao and Gimbutas \cite{xiao2010quadrature} which is exact for degree $2N+1$ polynomials.
-Since the values of $\rho,\mu$, and $\lambda$ are positive at all quadrature points, the method is energy stable.
-However, since the exact solution is discontinuous, the error in elements cut by the interface is $O(1)$, resulting in $L^2$ errors which converge at rate $O(h^{1/2})$.
-We note that, when using piecewise constant approximations of $\mu$ and $\lambda$, we observe the same $O(h^{1/2})$ convergence rate, though errors are roughly twice as large in magnitude.
-
-\begin{figure}
-\centering
-\subfloat[Fitted interface]{
-\begin{tikzpicture}
-\begin{loglogaxis}[
-    legend cell align=left,
-    legend style={legend pos=south east, font=\tiny},
-    width=.475\textwidth,
-    xlabel={Mesh size $h$},
-    ylabel={$L^2$ errors}, 
-    xmin=.025, xmax=1,
-    ymin=.75e-11, ymax=.25,
-    xmajorgrids=true,
-    ymajorgrids=true,
-    grid style=dashed,
-]
-
-\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.108456)(0.25,0.0385149)(0.125,0.0138288)(0.0625,0.0043988)};    
-\logLogSlopeTriangleFlip{.4}{0.125}{0.86}{1.65}{blue}
-
-\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.00887301)(0.25,0.00128102)(0.125,0.000186756)(0.0625,2.73376e-05)};    
-\logLogSlopeTriangleFlip{.4}{0.125}{0.65}{2.77}{red}
-
-\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.000930299)(0.25,7.92165e-05)(0.125,5.2534e-06)(0.0625,3.67676e-07)};    
-\logLogSlopeTriangleFlip{.4}{0.125}{0.475}{3.84}{black}
-
-\addplot[color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,4.34823e-05)(0.25,1.45784e-06)(0.125,4.92754e-08)(0.0625,1.74105e-09)};   
-\logLogSlopeTriangleFlip{.4}{0.125}{0.26}{4.82}{magenta}
-
-\addplot[color=violet,mark=pentagon*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,2.97005e-06)(0.25,4.94512e-08)(0.125,8.48451e-10)(0.0625,1.47109e-11)};  
-\logLogSlopeTriangleFlip{0.4}{0.125}{0.075}{ 5.85}{violet}
-
-%\node at (axis cs:.03,6.8e-10) {$u$ discontinuous};
-\legend{$N=1$,$N=2$,$N=3$,$N=4$,$N=5$}
-\end{loglogaxis}
-\end{tikzpicture}
-\label{subfig:stoneley1}
-}
-\subfloat[Non-fitted interface]{
-\begin{tikzpicture}
-\begin{loglogaxis}[
-    legend cell align=left,
-    legend style={legend pos=south east, font=\tiny},
-    width=.475\textwidth,
-    xlabel={Mesh size $h$},
-%    ylabel={$L^2$ errors}, 
-    xmin=.025, xmax=1,
-    ymin=1e-2, ymax=.25,
-    xmajorgrids=true,
-    ymajorgrids=true,
-    grid style=dashed,
-]
-
-\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.1507)(0.25,0.0796)(0.125,0.0495)(0.0625,0.0331)}; 
-%\logLogSlopeTriangleFlip{0.32}{0.125}{0.84}{1}{blue}
-
-\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.0820)(0.25,0.0594)(0.125,0.0427)(0.0625, 0.0307)};
-%\logLogSlopeTriangleFlip{0.32}{0.125}{0.62}{3}{red}
-
-\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5, 0.0640)(0.25,0.0435)(0.125,0.0306)(0.0625,0.0216)};
-%\logLogSlopeTriangleFlip{0.32}{0.125}{0.41}{4}{black}
-
-\addplot[color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5, 0.0593)(0.25, 0.0436)(0.125, 0.0318)(0.0625,0.0229)};
-\logLogSlopeTriangle{0.425}{0.125}{0.225}{1/2}{black}
-
-%\node at (axis cs:.03,6.8e-10) {$u$ discontinuous};
-\legend{$N=1$,$N=2$,$N=3$,$N=4$}
-\end{loglogaxis}
-\end{tikzpicture}
-\label{subfig:stoneley2}
-}
-\caption{Convergence of WADG for a Stoneley wave using a fitted mesh aligned with the interface and a non-fitted mesh where the interface does not lie exactly on an element boundary. }
-\label{fig:stoneley}
-\end{figure}
-
-\subsubsection{Convergence to a reference solution}
-
-Finally, we examine the accuracy of the WADG method for smoothly varying heterogeneous media by comparing against a reference spectral method solution with order $N = 50$ on a unit square $[-1,1]^2$.
-Boundary conditions are enforced weakly through DG-like fluxes \cite{wilcox2010high}.
-We use a heterogeneous isotropic medium with $\rho,\lambda, \mu$ set to
-\[
-\rho(\bm{x}) = 1, \qquad \lambda(\bm{x}) = 1 + .25\sin(\pi x)\sin(\pi y), \qquad \mu(\bm{x}) = 1 + .25\cos(\pi x) \cos(\pi y).  
-\]
-Initial stresses are set to zero, while the initial velocity is set to 
-\[
-v_1(\bm{x}) = \cos(\pi x)\sin(\pi y), \qquad v_2(\bm{x}) = -\sin(\pi x)\cos(\pi y).
-\]
-Zero traction boundary conditions are enforced on all domain boundaries.
-Figure~\ref{fig:refsol} shows $L^2$ errors with respect to the reference solution at time $T=1/2$ for different mesh sizes and orders of approximation.
-Computed convergence rates fall between the optimal $O(h^{N+1})$ and predicted $O(h^{N+1/2})$.
-\begin{figure}
-\centering
-\begin{tikzpicture}
-\begin{loglogaxis}[
-    legend cell align=left,
-    legend style={legend pos=south east, font=\tiny},
-    width=.475\textwidth,
-    xlabel={Mesh size $h$},
-%    ylabel={$L^2$ errors}, 
-    xmin=.025, xmax=1,
-    ymin=1e-7, ymax=1.25,
-    xmajorgrids=true,
-    ymajorgrids=true,
-    grid style=dashed,
-]
-
-\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,1)(0.25,0.473741)(0.125,0.141342)(0.0625,0.0355409)};    
-\logLogSlopeTriangleFlip{0.375}{0.11}{0.81}{1.99}{blue}
-
-\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.484875)(0.25,0.0686547)(0.125,0.00767686)(0.0625,0.000933392)};   
-\logLogSlopeTriangleFlip{0.375}{0.11}{0.59}{3.04}{red}
-
-\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.141415)(0.25,0.0126263)(0.125,0.0008361)(0.0625,5.57276e-05)};    
-\logLogSlopeTriangleFlip{0.35}{0.09}{0.42}{3.91}{black}
-
-\addplot[color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.0452398)(0.25,0.00182005)(0.125,9.0346e-05)(0.0625,3.08518e-06)};   
-\logLogSlopeTriangleFlip{0.35}{0.08}{0.255}{4.87}{magenta}
-
-\addplot[color=violet,mark=diamond*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.0186937)(0.25,0.000684602)(0.125,1.12013e-05)(0.0625,2.13796e-07)};    
-\logLogSlopeTriangleFlip{0.35}{0.075}{0.1}{5.71}{violet}
-
-\legend{$N=1$,$N=2$,$N=3$,$N=4$,$N=5$}
-\end{loglogaxis}
-\end{tikzpicture}
-\caption{Convergence of WADG under mesh refinement to a reference $N=50$ spectral method solution with smoothly varying heterogeneous media for $N=1,\ldots,5$.}
-\label{fig:refsol}
-\end{figure}
-
-%\subsubsection{Harmonic oscillation of an annulus}
-%
-%This example tests the convergence of WADG for curvilinear mappings.
-
-\subsection{Application examples}
-
-Next, we investigate the behavior of WADG for linear elasticity for several application examples.
-
-\subsubsection{Stiff inclusion}
-
-A common test of methods for linear elastic wave propagation is the stiff inclusion problem \cite{leveque2002finite,kaser2006arbitrary,appelo2015energy}.
-We solve on a domain $[-1,1]\times [-.5, .5]$ with a rectangular inclusion located at $[-.5,.5]\times[-.1,.1]$.
-Outside of the inclusion, material parameters are taken to be
-\[
-\rho = 1, \qquad \mu = 1,\qquad \lambda = 2.
-\]
-Within the inclusion, material parameters are taken to be
-\[
-\rho = 1, \qquad \mu = 100,\qquad \lambda = 200,
-\]
-such that the wave speed in the rectangular inclusion is ten times that of the wave speed outside.  A pulse is generated through velocity boundary conditions at $x = -1$
-\[
-v_1(x,y,t) = \begin{cases}
-\sin(\pi t / t_0), & t < t_0\\
-0, & t \geq t_0
-\end{cases}, \qquad
-v_2(x,y,t) = 0.
-\]
-In our experiments, we take $t_0 = .025$.
-Traction free boundary conditions are enforced at all other domain boundaries.
-A wave is generated when the pulse hits the inclusion, bouncing back and forth and producing reflected waves which propagate outward from the stiff region.
-
-\begin{figure}
-\centering
-\subfloat[$\LRb{\bm{\sigma}_{xx}+\bm{\sigma}_{yy}}$]{\includegraphics[width=.475\textwidth]{../figs/inclusion_aligned.png}}
-\hspace{2em}
-\subfloat[$\LRb{\bm{\sigma}_{xy}}$]{\includegraphics[width=.475\textwidth]{../figs/inclusion_aligned_shear.png}}
-%\subfloat[Aligned mesh (zoom)]{\includegraphics[height=.24\textwidth]{../figs/inclusion_aligned_zoom.png}}
-%\\
-%\subfloat[Un-aligned mesh]{\includegraphics[height=.24\textwidth]{../figs/inclusion_unaligned.png}}
-%\hspace{2em}
-%\subfloat[Un-aligned mesh (zoom)]{\includegraphics[height=.24\textwidth]{../figs/inclusion_unaligned_zoom.png}}
-\caption{Values of different solution fields at $T=.4$ for the stiff inclusion problem of Leveque \cite{leveque2002finite} with order of approximation $N=5$.}
-\label{fig:incl}
-\end{figure}
-
-%We consider both aligned meshes where the inclusion boundary coincides with interior faces and an un-aligned mesh where the inclusion cuts through element interiors.
-Both are constructed by dividing each element of a quadrilateral meshes along the diagonal to produce triangular meshes.
-The aligned mesh is constructed using $100 \times 50$ quadrilateral elements in the $x$ and $y$ coordinates, respectively, while the un-aligned mesh was constructed using $101 \times 51$ elements.
-Figure~\ref{fig:incl} shows values of $\LRb{\bm{\sigma}_{xx} + \bm{\sigma}_{yy}}$ at final time $T=.4$, along with a zoomed-in view along the interface.
-The results for the interface-aligned mesh show good agreement with results in the literature \cite{leveque2002finite,kaser2006arbitrary,appelo2015energy}.
-%The sub-element resolution of $\mu,\lambda$, an error is still incurred when the mesh is un-aligned with the interface
-%For the un-aligned mesh, the sub-element resolution of material data is clearly visible, though an interface error is still incurred by the under-resolution of the interface \cite{symes2009interface}.
-
-We construct a mesh by dividing each element of a quadrilateral meshes along the diagonal to produce triangular meshes, using $100 \times 50$ quadrilateral elements in the $x$ and $y$ coordinates, respectively.
-Figure~\ref{fig:incl} shows values of $\LRb{\bm{\sigma}_{xx} + \bm{\sigma}_{yy}}$ at final time $T=.4$.
-Following \cite{leveque2002finite}, a nonlinear color scale is used in order to distinguish small-amplitude waves.
-The results for the interface-aligned mesh show qualitatively good agreement with results in the literature \cite{leveque2002finite,kaser2006arbitrary,appelo2015energy}.
-
-%\note{Finish}
-
-\subsubsection{Heterogeneous anisotropic material}
-
-We next examine a model wave propagation problem in heterogeneous anisotropy media \cite{carcione1988wave,komatitsch2000simulation,de2007arbitrary}.
-The density $\rho = 7100$ is constant over the domain, while the entries of the stiffness matrix $\bm{C}$ are taken to be
-\begin{align*}
-\bm{C}_{11} &= .165, \quad \bm{C}_{12} = .05, \quad \bm{C}_{22} = .062, \quad \bm{C}_{33} = .0396, \qquad x < 0\\
-\bm{C}_{11} &= .165, \quad \bm{C}_{12} = .0858, \quad \bm{C}_{22} = .165, \quad \bm{C}_{33} = .0396, \qquad x > 0,
-\end{align*}
-with the remaining entries determined by symmetry or set to zero if unspecified.
-For $ x < 0$, this corresponds to an anisotropic material, while for $x > 0$, this corresponds to an isotropic material with $\mu = .0396, \lambda = .0858$.
-
-The computational domain is taken to be $[-.16, .16]^2$, and following \cite{komatitsch2000simulation,de2007arbitrary}, we use $N=5$ and a triangular mesh constructed by subdividing a grid of $32\times 32$ uniform quadrilaterals.
-This mesh corresponds to an element size of $.005$, which was used in \cite{komatitsch2000simulation,de2007arbitrary}.
-Forcing is applied to the $y$-component of the velocity by a Ricker pulse point source
-\[
-f(\bm{x},t) = \LRp{1 - 2(\pi f_0 (t-t_0))^2} e^{-(\pi f_0 (t-t_0))^2}\delta(x-x_0),
-\]
-where $x_0 = -.02$, $f_0 = .17$, and $t_0 = 1/f_0$.\footnote{All values and units are adapted from \cite{komatitsch2000simulation,de2007arbitrary}, and correspond to units of meters, kg, and microseconds.}
-
-\begin{figure}
-\centering
-%\subfloat[$\bm{v}_1$]{\includegraphics[width=.45\textwidth]{}}
-%\subfloat[$\bm{v}_2$]{\includegraphics[width=.45\textwidth]{}}
-\subfloat[$T=30 \mu s$]{\includegraphics[width=.4\textwidth]{../figs/aniso.png}}
-\hspace{2em}
-\subfloat[$T=60 \mu s$]{\includegraphics[width=.4\textwidth]{../figs/aniso.png}}
-\caption{Example of wave propagation in heterogeneous anisotropic media.  The vertical component of the velocity $\bm{v}_2$ is shown at $T = 30$ and $T=60$ microseconds. }
-\end{figure}
-
-\subsection{A three-dimensional example}
-
-Finally, we present a three-dimensional example of elastic wave propagation in heterogeneous media with sub-element variations, anisotropy, and a discontinuity across an interface.
-We consider elastic wave propagation on the bi-unit cube $[-1,1]^3$, with isotropic media for $z < 0$ and transversely isotropic media for $z > 0$.
-For isotropic media, we take $\mu(\bm{x}) = \lambda(\bm{x}) = w(\bm{x})$, where
-\[
-w(\bm{x}) = 1 + \cos(k\pi x)\cos(k\pi y)\cos(k\pi z).  
-\]
-For the transversely isotropic media, we take the entries of the stiffness matrix $\bm{C}$ to be 
-\begin{align*}
-\bm{C}_{11} &= \frac{2\mu + \lambda}{4}\\
-\bm{C}_{12} &= \frac{2\mu + \lambda}{4}\\
-\bm{C}_{13} &= \frac{2\mu + \lambda}{4}\\
-\bm{C}_{22} &= \frac{2\mu + \lambda}{4}\\
-\bm{C}_{23} &= \frac{2\mu + \lambda}{4}\\
-\bm{C}_{33} &= \frac{2\mu + \lambda}{4}
-\end{align*}
-
-We follow the implementation of GPU-accelerated DG methods outlined in \cite{klockner2009nodal}, which breaks the computational work in each timestep \emph{volume} and \emph{surface} kernels for the evaluation of the DG formulation and an \emph{update} step to apply a time integration method.
-For this work, apply the weight-adjusted mass matrix inverse within the update kernel as well.
-Strategies for volume and surface kernels follow \cite{klockner2009nodal}, while computational approaches for WADG are outlined in \cite{chan2016weight2}.
-More detailed computational performance analyses will be performed in future work.
-All three-dimensional computations are performed on an AMD Tahiti GPU.
-
-\section{Conclusions}
-\label{sec:conclusions}
-
-This work presents a weight-adjusted discontinuous Galerkin (WADG) method for the linear elastic wave equations with arbitrary heterogeneous media.
-The method is energy stable and high order accurate for arbitrary stiffness matrices, and a slight modification results in an energy stable method for curvilinear meshes as well.
-The penalty numerical fluxes for this formulation are simple to derive and implement, and their lack of dependence on the stiffness matrix allows for a unified treatment of isotropic and anisotropic media.
-Numerical examples confirm the accuracy of this method for analytic solutions in of the elastic wave equations, as well as its high order accuracy with respect to a reference solution for smoothly varying heterogeneous media.
-The behavior of this method for several application examples in also shows good agreement with existing results in the literature for both isotropic and anisotropic heterogeneous media.
-
-We note that the implementation of this method reduces to the application of the weight-adjusted mass matrix inverse and the evaluation of constant-coefficient terms in the DG formulation.
-Fast methods based on Bernstein-Bezier bases for the application of derivative and lift matrices for constant-coefficient terms can reduce this latter cost, especially at high orders of approximation \cite{chan2015bbdg}.
-Finally, we note that the approach presented here is more generally applicable to problems involving a weighted mass matrix with a matrix-valued weight, which we will explore in the future.
-These include hyperbolic systems of PDEs with a right hand side of the form $\bm{A}_{0}(\bm{x})\pd{\bm{U}}{t}$, such as Maxwell's equations for cloaking problems \cite{li2012time}.
-
-%Extension to hybrid meshes in 3D, care must be taken with pyramids \cite{chan2015orthogonal, chan2015gpu}.
-
-\section{Acknowledgments}
-
-The author gratefully thanks Thomas Hagstrom, Tim Warburton, Axel Modave, Ruichao Ye, Zheng Wang, and Mario Bencomo for helpful and informative discussions.
+GLL-based spectral elements have other advantages, notably a connection to summation by parts (SBP) operators.
+Wave propagation problems yield variational formulations which are a-priori stable under inexact quadrature.
+However, SBP properties are crucial to guaranteeing energy stability for more complex physics, such as nonlinear hyperbolic conservation laws.
 

### 2017-03-29 docs/IGA-DG/main.tex
update IGA docs (may be cruft now)

@@
 
-\newcommand{\half}{1/2}
+\newcommand{\Ksub}{K_{\rm sub}}
 
@@
 \author{Jesse Chan, John Evans}
-\title{B-spline DG notes}
+\title{Time-domain multi-patch discontinuous Galerkin methods}
 
@@
 \begin{abstract}
+
 Achieve extra $h$-resolution for the same CFL condition by using high order B-splines as a local approximation space on tensor product elements.
+Alternatively, achieve an $O(h_N/N)$ CFL with respect to the fine-grid mesh size $h_N = h/N$.
 
-Achieve a spectral element-type implementation using quadrature nodes for B-splines. 
 \end{abstract}
@@
 
-\section{Stable timestep restrictions for high order DG methods}
+Original IGA paper \cite{hughes2005isogeometric}
+
+The ``value of continuity'' for direct solvers \cite{garcia2016value} is to decrease the number of degrees of freedom while maintaining a comparable resolution.
+
+DG on tets using exact geometry \cite{michoski2015foundations}
+
+Multi-patch IGA \cite{langer2015multipatch,langer2016discontinuous}.
+
+%Exploiting tensor product \cite{mantzaflaris2016low}.
+
+Can improve work estimates \cite{chan2015bbdg}, but not CFL.
+
+Improving CFL: Dual-grid filter \cite{warburton2008taming}.
+Adding conservation constraints \cite{xu2010conservation,xu2014new}, increasing coupling between elements.
+Central DG \cite{reyna2014operator}.
+
+Hermite methods \cite{goodrich2006hermite,vargas2015variations}.
+
+The work here more closely resembles the recently introduced Galerkin difference methods \cite{banks2016galerkin}, which constructs extended-support basis functions from .
+
+\section{Stable timestep restrictions for time-domain DG methods}
+
+Under the method of lines, a PDE is transformed to a system of ordinary differential equations (ODEs) 
+\[
+\td{\bm{U}}{t} = \bm{A}_h \bm{U}
+\]
+by discretizing in space.
+This system may then be solved using an explicit time stepping, which avoids the inversion of a global matrix.
+Since such a scheme requires only matrix-vector products, the storage of a global matrix may also be avoided through a matrix-free implementation.
+Under explicit time stepping, the eigenvalues of $\bm{A}_h$ must lie within the stability region of the chosen scheme.
+
+Bounds on real and imaginary parts of eigenvalues can be estimated using theorems of Bendixson and Hirsch \cite{marcus1992survey}, as is done in \cite{chan2015gpu}.
 
@@
 &\leq \nor{\beta}_{L^\infty}\sum_{k} \nor{\Grad u}\nor{v} + \nor{f^*(u)}_{\Ldk}\nor{v}_{\Ldk}\\
-&\lesssim \nor{\beta}_{L^\infty}\sum_{k} \LRp{C_I \nor{J\bm{G}}_{L^\infty\LRp{\Dhat}}  + C_N^2 \nor{J^f}_{L^\infty\LRp{\Dhat}}}\nor{1/J}_{L^\infty\LRp{\Dhat}}\nor{u}_{\Lk}\nor{v}_{\Lk}
+&\lesssim \nor{\beta}_{L^\infty}\sum_{k} \LRp{C_I \nor{J\bm{G}}_{L^\infty\LRp{\Dhat}}  + C_N^2 \nor{J^f}_{L^\infty\LRp{\Dhat}}}\nor{1/J}_{L^\infty\LRp{\Dhat}}\nor{u}_{\Lk}\nor{v}_{\Lk}.  
 \end{align*}
@@
 
+For shape regular meshes, the ratio between volume and surface area is proportional to $h$, so 
+\[
+\nor{J^f}_{\LinfDk} \nor{J^{-1}}_{\LinfDk} = O(h).
+\]
+
+For polynomials, it can be shown that $C_I,C_N$ both depend quadratically on the degree $N$ (see, for example \cite{warburton2003constants, ozisik2010constants}).
+Thus, for DG methods based on locally polynomial approximation spaces, the spectral radius can be bounded by the mesh size, polynomial order, and maximum wavespeed
+\[
+\rho\LRp{\bm{A}_h} = O\LRp{\frac{\beta_{\max} N^2}{h}}.  
+\]
+The quadratic dependence of $\rho\LRp{\bm{A}_h}$ on $N$ results in small timesteps if $N$ is large, making the use of high polynomial degrees computationally intractable.
+
 \subsection{Explicit constants in trace and inverse inequalities for polynomial and B-spline approximations}
 
-For polynomials, $C_I, C_N = O(N^2)$ on the reference element (see, for example \cite{warburton2003constants, ozisik2010constants, chan2015gpu}).
+For polynomials, $C_I, C_N = O(N^2)$ on the reference element .
 
@@
 Or, if we measure by $h_N = h/N$, CFL is $O(h_N/N)$.
-This results from interpolation estimates for one-dimensional B-splines \cite{bazilevs2006isogeometric}
+This results from interpolation estimates for one-dimensional B-splines \cite{bazilevs2006isogeometric}.
+
+Show plots of lift functions for polynomials vs for B-spline bases - less oscillation for B-splines.
+
+\begin{figure}
+\centering
+\subfloat[$N_P = N+\Ksub$]{\includegraphics[width=.4\textwidth]{figs/lift1.png}}
+\subfloat[Zoom in]{\includegraphics[width=.4\textwidth]{figs/liftZoom1.png}}\\
+\subfloat[$N_P = N$]{\includegraphics[width=.4\textwidth]{figs/lift2.png}}
+\subfloat[Zoom in]{\includegraphics[width=.4\textwidth]{figs/liftZoom2.png}}
+\caption{Comparison of lift functions for a degree $N_P$ polynomial space and spline space with degree $N = 7$ and $\Ksub = N$ elements.
+Degree $N_P$ Gauss-Radau quadrature points are also shown for comparison.
+}
+\end{figure}
+
+\section{DG formulation}
+
+Skew-symmetric form is energy stable for acoustic, electromagnetic, and elastic wave propagation.
+Disadvantage of IGA-DG vs SEM is a lack of equivalence with strong and weak formulations \cite{kopriva2010quadrature}
+
+Penalty fluxes \cite{chan2016short} for dissipation.
+
+Avoid mass matrix inversion using weight-adjusted DG \cite{chan2016weight1,chan2016weight2}.
+Especially important for IGA-DG since it decouples the inversion of the mass matrix into a Kronecker structure.
 
@@
 We note that these rules are generated by choosing the number of points $N_q$ such that ${\rm dim}(S_{\rm target}) \leq 2N_q$, and may not be optimal.
+The problem is that, to exactly integrate the target space, $N_q = O(N^2)$.
+Reduced quadratures are thus necessary.
 
-\subsection{Reduced quadrature using Gauss-Legendre-Lobatto rules}
+\subsection{Reduced quadrature rules}
+
+The derivative of a spline does not map into the same spline space; instead, it decreases both the order and continuity.
+The derivative operator then satisfies $\pd{}{x} : S^N_{N-1}\rightarrow S^{N-1}_{N-2}$.
+To define a derivative operator which maps from $S^N_{N-1}$ to itself, we use an $L^2$ projection.
+\[
+\int_{D^k} a(x)\pd{u}{x} v = \int_{\widehat{D}} a(x)\pd{r}{x}J\pd{u}{r} v = \int_{\widehat{D}} a(x)\LRp{\pd{r}{x}J} \Pi_N\LRp{\pd{u}{r}} v.  
+\]
 
 Can cast into collocation style using skew-symmetric formulation by using a degree $N + K_{\rm sub} - 1$ Gauss-Legendre or Gauss-Lobatto rule.
-Show plots of GLL nodal bases for polynomials vs for B-spline bases - less oscillation for basis functions at nodes on/near the boundary.
 

### 2017-04-12 docs/splitform/main.tex
adding splitform DG notes

--- 2017-02-15 docs/IGA-DG/main.tex
@@
 %\documentclass[preprint,10pt]{elsarticle}
-%\documentclass[10pt]{article}
+\documentclass[preprint,10pt]{article}
 %\documentclass[review]{siamart0216}
-\documentclass{siamart0216}
+%\documentclass{siamart0216}
 
@@
 \usepackage{bm}
+\usepackage{bbm}
+
 %\usepackage{tikz}
@@
 
+\newcommand{\bbm}[1]{\mathbbm{#1}}
 \newcommand{\bs}[1]{\boldsymbol{#1}}
@@
 \newcommand{\pdd}[2]{\frac{\partial^2#1}{\partial#2^2}}
+\newcommand{\pdn}[3]{\frac{\partial^{#3}#1}{\partial#2^{#3}}}
 \newcommand{\mb}[1]{\mathbf{#1}}
@@
 \renewcommand{\L}{L^2\LRp{\Omega}}
-\newcommand{\Lk}{L^2\LRp{D^k}}
-\newcommand{\Ldk}{L^2\LRp{\partial D^k}}
-%\newcommand{\Lk}{D^k}
-%\newcommand{\Ldk}{\partial D^k}
+\newcommand{\LK}{L^2\LRp{D^k}}
+\newcommand{\LdK}{L^2\LRp{\partial D^k}}
 \newcommand{\Dhat}{\widehat{D}}
 \newcommand{\Lhat}{L^2\LRp{\Dhat}}
-
-%\newtheorem{theorem}{Theorem}[section]
-%\newtheorem{lemma}[theorem]{Lemma}
-%\newtheorem{proposition}[theorem]{Proposition}
-%\newtheorem{corollary}[theorem]{Corollary}
-
-%\newenvironment{definition}[1][Definition]{\begin{trivlist}
-%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
-%\newenvironment{example}[1][Example]{\begin{trivlist}
-%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
 
@@
 \date{}
-\author{Jesse Chan, John Evans}
-\title{B-spline DG notes}
+\author{Jesse Chan}
+\title{Split form DG methods}
 
@@
 \maketitle
+%\tableofcontents
 
-\begin{abstract}
-Achieve extra $h$-resolution for the same CFL condition by using high order B-splines as a local approximation space on tensor product elements.
+\section{Notes}
 
-Achieve a spectral element-type implementation using quadrature nodes for B-splines. 
-\end{abstract}
+Integration by parts gives
+\[
+\sum_{D^k} \LRp{\Grad u, \bm{v}}_{\LK} = \sum_{D^k} \LRp{-u, \Grad \cdot \bm{v}}_{\LK} + \LRa{u,\bm{v}\cdot\bm{n}}_{\LdK}.
+\]
+We replace the values of ${u}$ with $\avg{u}$ on each element boundary to define the global DG gradient operator $\Grad_h$ 
+\[
+\sum_{D^k} \LRp{\Grad_h u, \bm{v}}_{\LK} \coloneqq \sum_{D^k} \LRp{-u, \Grad \cdot \bm{v}}_{\LK} + \LRa{\avg{u},\bm{v}\cdot\bm{n}}_{\LdK}.
+\]
+Integrating by parts again and the introduction of the lift operator shows that
+\[
+\Grad_h u = \Grad u + \frac{1}{2}L\LRp{\jump{u}\bm{n}}
+\]
+where $L$ is the lift operator.
 
-\tableofcontents
+The DG divergence operator is similarly defined as
+\[
+\Grad_h \cdot u = \Grad\cdot u + \frac{1}{2}L\LRp{\jump{\bm{u}}\cdot\bm{n}}
+\]
+and it can be shown that, for a periodic mesh,
+\[
+\LRp{\Grad_h u, \bm{v}} = \LRp{-u, \Grad_h \cdot \bm{v}}.
+\]
+Need to incorporate boundary conditions.
 
-\section{Introduction}
+We also have that, for interior elements,
+\[
+\LRp{\Grad_h\cdot \bm{u},v\bbm{1}_{D^k}} = \LRp{\bm{u}, \Grad\cdot v\bbm{1}_{D^k}}  + \LRa{\avg{\bm{u}}\cdot\bm{n},v}_{\partial D^k}.
+\]
+and as a result when $v = 1$
+\[
+\LRp{\Grad_h\cdot \bm{u},\bbm{1}_{D^k}} = \int_{\partial D^k}\avg{\bm{u}}\cdot\bm{n}
+\]
+\section{Variable advection}
 
-\section{Stable timestep restrictions for high order DG methods}
+A split formulation for advection is 
+\[
+\LRp{\pd{u}{t},v} + \frac{1}{2}\LRp{\Grad_h\cdot \Pi_N \LRp{ \bm{\beta}u},v} + \frac{1}{2}\LRp{\bm{\beta}\cdot\Grad_h u,v} + \frac{1}{2}\LRp{\LRp{\Grad\cdot \bm{\beta}} u,v} = 0.
+\]
+Taking $v = u$ yields and using $\LRp{\Grad_h u, \bm{v}} = \LRp{-u, \Grad_h \cdot \bm{v}}$ yields the energy statement
+\[
+\frac{1}{2}\nor{u}^2 + \frac{1}{2}\LRp{\Grad_h\cdot \Pi_N \LRp{ \bm{\beta}u},u} - \frac{1}{2}\LRp{ u, \Grad_h\cdot \Pi_N\LRp{\bm{\beta}u}} = \frac{1}{2}\LRp{-\LRp{\Grad \cdot \bm{\beta}} u,u},
+\]
+implying that $\frac{1}{2}\nor{u}^2 = 0$ if $\Grad\cdot \bm{\beta} = 0$, or that the method is energy conserving.
+The only difference in this formulation is the introduction of $\Pi_N$, which can be defined at a discrete level using any quadrature scheme for which a discrete projection is well-defined.
 
-We illustrate this using the advection equation.  
-\begin{align*}
-V^T A_h U &= \sum_{k} \LRp{\beta\cdot\Grad u,v}_{\Lk} + \LRa{f^*(u),v}_{\Ldk} \\
-&\leq \nor{\beta}_{L^\infty}\sum_{k} \nor{\Grad u}\nor{v} + \nor{f^*(u)}_{\Ldk}\nor{v}_{\Ldk}\\
-&\lesssim \nor{\beta}_{L^\infty}\sum_{k} \LRp{C_I \nor{J\bm{G}}_{L^\infty\LRp{\Dhat}}  + C_N^2 \nor{J^f}_{L^\infty\LRp{\Dhat}}}\nor{1/J}_{L^\infty\LRp{\Dhat}}\nor{u}_{\Lk}\nor{v}_{\Lk}
-\end{align*}
-The CFL condition is then determined by mesh-dependent geometric factors of $O(h)$ and order-dependent constants in the inverse and trace inequalities $C_I, C_N$.
+\section{Local conservation}
 
-\subsection{Explicit constants in trace and inverse inequalities for polynomial and B-spline approximations}
+Writing this in non-conservative form raises the question of local conservation.  Integrating the original equation over $D^k$ and using Gauss' theorem gives
+\[
+\int_{D^k}\pd{u}{t} + \int_{\partial D^k} \beta_n u = 0.
+\]
+Taking $v = 1$ on $D^k$ yields
+\[
+\int_{D^k}\pd{u}{t} + \frac{1}{2}\LRp{\Grad_h\cdot \Pi_N \LRp{ \bm{\beta}u},\bbm{1}_{D^k}} + \frac{1}{2}\LRp{\bm{\beta}\cdot\Grad_h u,\bbm{1}_{D^k}} + \frac{1}{2}\LRp{\LRp{\Grad\cdot \bm{\beta}} u,\bbm{1}_{D^k}} = 0.
+\]
+The first term gives
+\[
+\LRp{\Grad_h\cdot \Pi_N \LRp{ \bm{\beta}u},\bbm{1}_{D^k}} = \int_{\partial D^k} \avg{\Pi_N\LRp{\bm{\beta}u}}\cdot\bm{n}.
+\]
+The second term gives
+\[
+\LRp{\bm{\beta}\cdot\Grad_h u,\bbm{1}_{D^k}} = \LRp{\Grad u,\bm{\beta}}_{D^k} + \frac{1}{2}\LRa{\jump{u},\bm{\beta}\cdot\bm{n}} = \LRp{u,-\Grad \cdot \bm{\beta}}_{D^k} + \LRa{\avg{u},\bm{\beta}\cdot\bm{n}}
+\]
+through integration by parts and an assumption that $\bm{\beta}\cdot \bm{n}$ is periodic.  Cancelling volume terms, we end up with the statement of local conservation 
+\[
+\int_{D^k} \pd{u}{t}  + \frac{1}{2}\int_{\partial D^k} \LRp{\avg{\Pi_N\LRp{\bm{\beta}u}} + \LRp{\Pi_N\LRp{\bm{\beta}}\avg{u}}}\cdot\bm{n} = 0
+\]
+which is a discrete version of the continuous statement of local conservation.
 
-For polynomials, $C_I, C_N = O(N^2)$ on the reference element (see, for example \cite{warburton2003constants, ozisik2010constants, chan2015gpu}).
+\section{Penalization}
 
-For B-splines with $K_{\rm sub} = N$, $C_N = O(N)$ instead of $O(N^2)$.
+Penalization can be added by adding any positive-definite stabilization term (upwind, penalty, Lax-Friedrichs) through the regular divergence flux.
 
-For sufficiently regular solutions, convergence is in $L^2$ as $C (h/N)^{N+1}$ while still maintaining an $O(h/N^2)$ CFL.
-Or, if we measure by $h_N = h/N$, CFL is $O(h_N/N)$.
-This results from interpolation estimates for one-dimensional B-splines \cite{bazilevs2006isogeometric}
+\section{Extension to other hyperbolic problems}
 
-\section{Quadrature }
+Example: acoustic wave equation, simply discretize by replacing $\Grad, \Grad\cdot$ with the DG versions.  Automatically skew symmetric and energy stable.
 
-Reduce dimension of the search space by treating the weights as dependent variables and assuming symmetry of the rule.
-Initialize guess for quadrature rule using the previous order's rule to generate a polynomial interpolant from equispaced nodes to quadrature nodes \cite{warburton2006explicit,chan2015comparison}.
+Example: Burgers' equation
 
-We note that these rules are generated by choosing the number of points $N_q$ such that ${\rm dim}(S_{\rm target}) \leq 2N_q$, and may not be optimal.
-
-\subsection{Reduced quadrature using Gauss-Legendre-Lobatto rules}
-
-Can cast into collocation style using skew-symmetric formulation by using a degree $N + K_{\rm sub} - 1$ Gauss-Legendre or Gauss-Lobatto rule.
-Show plots of GLL nodal bases for polynomials vs for B-spline bases - less oscillation for basis functions at nodes on/near the boundary.
-
-Problems here - high order accuracy is preserved if the mass matrix is under-integrated, but not if an isoparametric mapping is used.
-
-\section{Future work}
-
-GLL-based spectral elements have other advantages, notably a connection to summation by parts (SBP) operators.
-Wave propagation problems yield variational formulations which are a-priori stable under inexact quadrature.
-However, SBP properties are crucial to guaranteeing energy stability for more complex physics, such as nonlinear hyperbolic conservation laws.
+Example: Entropy splitting of Buckley-Leverett.
 

### 2017-05-03 docs/dg_penalty/revision/reviewerResponse.tex
updating docs for penalty note + elasticity

%&pdflatex

%\documentclass[preprint,10pt]{elsarticle}
\documentclass[10pt]{article}
\usepackage{fullpage}
\usepackage{hyperref}
\let\proof\relax
\let\endproof\relax
\usepackage{amsmath,amssymb,amsfonts,mathrsfs,amsthm}
\usepackage[titletoc,toc,title]{appendix}

%\usepackage{lineno}
\usepackage{array}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{mathtools}
\usepackage{pdfpages}
\usepackage[textsize=footnotesize,color=green]{todonotes}
\usepackage{bm}
%\usepackage{tikz}
\usepackage[normalem]{ulem}
\usepackage{hhline}
\usepackage{csquotes}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{algorithmicx}

\usepackage{graphicx}
%\usepackage{subfigure}
\usepackage{subfig}

\usepackage{color}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

\newcommand{\vect}[1]{\ensuremath\boldsymbol{#1}}
\newcommand{\tensor}[1]{\underline{\bm{#1}}}
\newcommand{\del}{\triangle}
\newcommand{\curl}{\grad \times}
\renewcommand{\div}{\grad \cdot}
\newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}#2}}
\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
\newcommand{\pdd}[2]{\frac{\partial^2#1}{\partial#2^2}}

\newcommand{\bs}[1]{\boldsymbol{#1}}

\newcommand{\equaldef}{\stackrel{\mathrm{def}}{=}}

\newcommand{\tablab}[1]{\label{tab:#1}}
\newcommand{\tabref}[1]{Table~\ref{tab:#1}}

\newcommand{\theolab}[1]{\label{theo:#1}}
\newcommand{\theoref}[1]{\ref{theo:#1}}
\newcommand{\eqnlab}[1]{\label{eq:#1}}
\newcommand{\eqnref}[1]{\eqref{eq:#1}}
\newcommand{\seclab}[1]{\label{sec:#1}}
\newcommand{\secref}[1]{\ref{sec:#1}}
\newcommand{\lemlab}[1]{\label{lem:#1}}
\newcommand{\lemref}[1]{\ref{lem:#1}}

\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\snor}[1]{\left| #1 \right|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRb}[1]{\left| #1 \right|}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}

\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\Div} {\ensuremath{\nabla\cdot}}
\newcommand{\Nel} {\ensuremath{{N^\text{el}}}}
\newcommand{\jump}[1] {\ensuremath{\LRs{\![#1]\!}}}
\newcommand{\avg}[1] {\ensuremath{\LRc{\!\{#1\}\!}}}
\newcommand{\uh}{\widehat{u}}
\newcommand{\Bh}{\widehat{B}}
\newcommand{\fnh}{\widehat{f}_n}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\pO}{\partial\Omega}
\newcommand{\Gh}{\Gamma_h}
\newcommand{\Gm}{\Gamma_{-}}
\newcommand{\Gp}{\Gamma_{+}}
\newcommand{\Go}{\Gamma_0}
\newcommand{\Oh}{\Omega_h}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

\newenvironment{definition}[1][Definition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newenvironment{example}[1][Example]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\newcommand{\eval}[2][\right]{\relax
  \ifx#1\right\relax \left.\fi#2#1\rvert}

\def\etal{{\it et al.~}}

\def\arr#1#2#3#4{\left[
\begin{array}{cc}
#1 & #2\\
#3 & #4\\
\end{array}
\right]}
\def\vectwo#1#2{\left[
\begin{array}{c}
#1\\
#2\\
\end{array}
\right]}
\def\vecthree#1#2#3{\left[
\begin{array}{c}
#1\\
#2\\
#3\\
\end{array}
\right]}
\def\vectfour#1#2#3#4{\left[
\begin{array}{c}
#1\\
#2\\
#3\\
#4\\
\end{array}
\right]}

\newcommand{\G} {\Gamma}
\newcommand{\Gin} {\Gamma_{in}}
\newcommand{\Gout} {\Gamma_{out}}

\newcommand{\note}[1]{{\color{blue}#1}}
\newcommand{\remark}[1]{\textbf{\color{red}#1}}

\newcommand{\tri}{{\rm tri}}
\newcommand{\sqr}{{\rm quad}}

\newcommand{\refhex}{\widehat{\mathcal{H}}}
\newcommand{\reftet}{\widehat{\mathcal{T}}}
\newcommand{\refwedg}{\widehat{\mathcal{W}}}

\newcommand{\refpyr}{\widehat{\mathcal{P}}}
\newcommand{\refpyrf}{\widehat{\mathcal{P}}^f}
\newcommand{\refpyrfq}{\widehat{\mathcal{P}}^{\sqr}}
\newcommand{\refpyrft}{\widehat{\mathcal{P}}^{\tri}}

\newcommand{\hex}{{\mathcal{H}}}
\newcommand{\tet}{{\mathcal{T}}}
\newcommand{\wedg}{{\mathcal{W}}}

\newcommand{\pyr}{{\mathcal{P}}}
\newcommand{\pyrf}{{\mathcal{P}}^f}
\newcommand{\pyrfq}{{\mathcal{P}}^{\sqr}}
\newcommand{\pyrft}{{\mathcal{P}}^{\tri}}

\newcommand{\half}{1/2}

\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

%% d in integrand
\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}}

\makeatletter
\renewcommand\d[1]{\mspace{6mu}\mathrm{d}#1\@ifnextchar\d{\mspace{-3mu}}{}}
\makeatother

\newcommand{\squareface}{[-1,1]^2}

%\journal{Journal of Computational Physics (JCP)}

\date{}
\author{Jesse Chan, T. Warburton}
\title{Response to reviewers: ``A short note on the penalty flux parameter for first order discontinuous Galerkin formulations''}

\begin{document}

\maketitle

\note{We thank both Reviewers 1 and 2 for their feedback.  We describe steps taken to address reviewer comments and suggestions, which are described in the following response.  Revisions in the manuscript are also colored in blue for ease of identification.  We hope these revisions improve the readability of this paper and its suitability for the audience of CAMWA.}

\section{Referee 1}

The authors characterize (theoretically and numerically) the spectral behavior of DG schemes for hyperbolic PDEs with penalty fluxes in terms of conforming and nonconforming parts of the approximation space.  The theoretical work in this paper is a minor modification of reference [5] to hyperbolic problems. The numerical study of the behavior of the spectral behavior appears novel. The paper does not explicitly state its claimed novelty.  The exposition is clear, easy to understand, and if the issues below are addressed, suitable for publication.

\begin{itemize}
%\item The paper consistently uses the plural form "spectra" where the singular ``spectrum'' would be appropriate. This should be fixed.  
%\note{We have  this.}
%\item References 8 and 18 appear to be identical. One copy should be removed.
%\note{This has been corrected.}
\item In the reviewer's recollection, the derivation of upwind fluxes involves Rankine-Hugoniot conditions for the construction of the numerical flux/Riemann solver after the hyperbolic system is diagonalized. The author's derivation does not reflect this fact and does not cite any sources. The authors should either cite a source outlining the correctness of the used approach or more completely justify that their derivation of the upwind fluxes is correct.  
\note{We agree with the reviewer, and have added citations and additional discussion for the statement on the formulation of upwind fluxes in terms of the diagonalization of $\bm{A}_n$. }
%\item The authors should emphasize that a \emph{joint} diagonalization of $A_n$ is required (p.3 above Sec. 3) which may not be available.
\item (p.3 above Sec. 3) The comment about the CFL restriction should be removed or justified further.
\note{We agree, and have added a justification for the CFL restriction.  We have also corrected an error in the scaling of $\tau$.}
%\item The grammar at the beginning of 2.1 should be checked.  
\item p.1: The statement $\tau=O(1)$ is difficult to understand without saying what specific dependencies/scalings do *not* exist. ($h,p$?)
\note{We apologize for the confusion.  We have added an additional statement explaining that $\tau$ does not need to be chosen with respect to discretization parameters, and have replaced $\tau = O(1)$ with the specific case of $\tau = 1$.  We also mention that, while it may be advantageous to scale $\tau$ with problem parameters, taking $\tau = 1$ is a reasonable choice for the problems considered.}
\item Equation (2) appears to be missing a derivative in the second term.
\note{We apologize: summation notation was used, but not specified.  This has been remedied in the revision.}
%\item This reviewer finds the nation $D^{k,-}$ to be a confusing mixture of global and local indexing--a true global indexing scheme would include a face index.  
\item Section 2.1 seems inconsistent in its use of $V^{-1}$ vs $V^T$.  \note{This has been corrected, and  $V^{-1}$ is now used everywhere.}
\item p.6: The notation surrounding $W^C$ and $W^{NC}$ is poor and should be checked.  This reviewer was unable to tell what the resulting shape of WC and WNC would be.
\note{We agree, and have added expanded descriptions of $\bm{W}^C, \bm{W}^{NC}$ to the revised manuscript, including explicit dimensions.  We hope these revisions adequately address these notational problems.}
\item p.6 top: If $WC, WNC$ genuinely is a matrix (the reviewer was unable to tell), then the RHS of the equation at the top of p.6 should read $(WC, WNC) \Lambda_C$ (i.e. the order is reversed--to scale the columns, the eigenvalues should come second).  
\note{We thank the reviewer for pointing out this error; we have corrected this in the revised version.}
\end{itemize}

\section{Referee 2}

The focus of this paper is to analyze and numerically demonstrate how the spectra of a DG discretization for first order linear system depends on a penalty parameter in penalty flux. The topic is interesting, yet there are some important technical details which need to be further clarified or revised.

\begin{itemize}
\item The reviewer finds the conclusion, both in introduction and in the text, regarding the behavior of the spectra as $\tau->\infty$ contradicts.

 \begin{quotation}
``The first set of eigenvalues correspond to the eigenvalues of a conforming discretization, while the second set of eigenvalues have real parts approaching $-\infty$ as $\tau-> \infty$ and correspond to damped spurious modes. Numerical experiments verify these results for the advection and acoustic wave equations in one and two dimensions. Numerical experiments also show that, while increasing $\tau$ pushes most eigenvalues further left of the imaginary axis, the real parts of certain eigenvalues converge to zero as $\tau-> \infty$.''
 \end{quotation}

\begin{enumerate}
\item are all the spurious modes in the second set, or some are in the first set? In other words, will the numerical spectra of conforming discretization all be non-spurious? In fact, based on Lemma 3.1 on page 6, it seems all in the first set are non-spurious.  
\note{We apologize for this confusion, and have clarified in the revised manuscript that conforming discretizations contain spurious modes as well.  This illustrates that DG methods with upwind or penalty fluxes address both \textit{non-conforming} and \textit{conforming} spurious modes which are otherwise present both for central fluxes ($\tau = 0$) and conforming discretizations (or equivalently, for large penalty parameters $\tau\rightarrow \infty$). }
%\item If it is yes for the previous question, that is if Lemma 3.1 is true, then there is contradiction in "while the second set of eigenvalues have real parts approaching $-\infty$ as $\tau-> \infty$" and in "the real parts of certain eigenvalues converge to zero as  $\tau->\infty."$  That is, the conclusion in Lemma 3.1 is flawed.  
\item page 6: the authors on one hand wrote: ``under normalization of $W^{C}, W^{NC}$'', on the other hand, conclude $W^{NC}\rightarrow 0$ as $\tau->\infty$, the argument contradicts.  
\note{We apologize for this typo.  The normalization should be applied to $\bm{W}$, such that the component $\bm{W}^{NC}\rightarrow 0$ as $\tau \rightarrow \infty$.  This has been corrected in the revised manuscript.}
\item page 6: it seems $S$ is only semi-negative. If so, $\lambda_{\min}^{NC}$ can be zero, and the argument right above Lemma 3.1 may not go through. The authors would want to revisit or clarify.
\note{We agree that if $S$ was constructed with respect to the entire DG space, it would be only negative semi-definite.  However, $\bm{S}$ is constructed by evaluating flux terms only for elements of $V^{NC}$, and is thus negative definite.  We have added a short section clarifying and explaining this claim.}
\end{enumerate}
\item The authors mentioned several times ``for problems with continuous coefficients''
\begin{enumerate}
\item If the coefficients are discontinuous, how is the numerical flux defined? How is $A_n$ defined at mesh interface? 
\note{We have restructured the manuscript to include more general $\bm{A}_i$, including $\bm{A}_n$ which are discontinuous across interfaces.  If coefficients are discontinuous, one can replace the penalty flux penalization term $\bm{A}_n^T\bm{A}_n\jump{\bm{U}}$ with $\bm{A}_n^T\jump{\bm{A}_n\bm{U}}$, where
\[
\jump{\bm{A}_n\bm{U}} = \bm{A}_n^+\bm{U}^+ - \bm{A}_n^-\bm{U}^-, \qquad \bm{A}_n^{\pm} = \sum_{i=1}^d \bm{n}_i \bm{A}_i^{\pm}.
\]
We have added a detailed discussion to the revised manuscript.  }
\item The authors would want to clarify the precise meaning of ``continuous coefficients'', namely for which models coefficients, the continuity is referred to in the discussions and various claims. (See page 3). Such clarification should be stated initially when ``continuous coefficients'' is first mentioned.  
\note{We agree, and have added a clarification on page 2.  In the restructured manuscript, we also begin with more general $\bm{A}_i$, and mention when specific assumptions on $\bm{A}_n$ are required.  }
\end{enumerate}
\item In the definition and discussion of numerical flux: 
\begin{enumerate}
\item If $A_n$ is symmetric, what is the need to use $A_n^T$, which is indeed $A_n$? That is, one can use $A_n^2$, instead of $A_n^TA_n$.  
\note{We have restructured the manuscript such that we do not begin with the assumption that $\bm{A}_n = \bm{A}_n^T$.  We hope this addresses the reviewer's concerns and makes it more clear that the penalty flux using $\bm{A}_n^T\bm{A}_n$ can be applied to the case when $\bm{A}_n$ is non-symmetric to yield a stabilizing contribution. }
\item Even though mathematically it is correct for symmetric matrix, when one states about eigenvalue decomposition, it is more fundamental to have $A = V\Lambda V^{-1}$, instead of $V \Lambda V^{T}$. Surely one can have $V^{-1}=V^{T}$ when $A$ is symmetric. The authors should revisit all $V^{T}$ in the text, either replace it by $V^{-1}$, or explicitly stated that $V^{-1}=V^{T} $is true in the current setting.  
\note{We agree with the reviewers assessment.  To address this, we have removed all mentions of $V^T$ within the text, and have replaced them with $V^{-1}$ instead.}
\item 7 lines above section 3, there is a typo in the symbol of pseudo-inverse. The authors may also want to include the explicit expression of pseudo-inverse.
\note{We have revised this to include a citation for the pseudo-inverse.  However, we could not find the typo within the symbol of the pseudo-inverse (we are using the ``dagger'' symbol in \LaTeX).  }
\item 1 line above section 3: it is too strong to state "CFL restrictions ... will be identical..'' unless the authors have a mathematical proof for this.
\note{We agree, and have added a justification for the CFL restriction in the revised manuscript.  We have also corrected an error in the scaling of $\tau$.}
\end{enumerate}
\item More related to flux
\begin{enumerate}
\item In (4), note that $A_n^TA_n$ and $A_n$ might have different null spaces, the authors would want to revisit, whether they want to have $[u]$ belong to the null space of $A_n^TA_n$ or $A_n$.  They also want to check the consistency in discussion, see 5 lines below (4).
\note{We have added the following clarification to the revised manuscript to explain why the null spaces of $\bm{A}_n^T\bm{A}_n$ and $\bm{A}_n$ are identical: the null space of $\bm{A}_n^T\bm{A}_n$ contains the null space of $\bm{A}_n$.  However, if $\bm{A}_n^T\bm{A}_n\bm{U} = 0$, then $\bm{U}^T\bm{A}_n^T\bm{A}_n\bm{U} = \nor{\bm{A}_n\bm{U}}^2 = 0$ as well, implying that the null space of $\bm{A}_n^T\bm{A}_n$ is also contained within the null space of $\bm{A}_n$. 
}
\item page 4: see penalty flux in line 6, and the penalty terms does not seem to be the same as the general definition $(A_n^T A_n)$
\note{We have clarified this in the revised manuscript using the above discussion.}
\end{enumerate}
\item page 5: Explain how come there is no boundary integral in $A_{mn}$, note some $\phi$ is only continuous on normal direction.
\note{We have clarified in the revised manuscript that flux terms vanish due to the definition of the conforming space.  We have also rearranged the DG formulation to make these properties clearer. }
\item page 4: ``this implies that $V_u^V$ is the H(div)-conforming Brezzi-Douglas-Marini finite element space''. The authors would want to check the definition of BDM elements, which differ greatly for rectangular or triangular meshes. On rectangular, they are not related to what the authors are using as discrete space.  
\note{We thank the reviewer for pointing this out, and have specified in the revised manuscript that this equivalence holds only for simplicial elements.}
\item Title: the meaning of "first order discontinuous Galerkin formulations" is not very clear. The authors can consider to rephrase the title.
\note{We have rearranged the title to specify the paper is on discontinuous Galerkin formulations for first order hyperbolic systems.}
\item If in (2) the summation convention is used, state this.
\note{We apologize: summation notation was used, but not specified.  We have switched to explicit sums in the revised version.}
%\item 3rd line in section 2.1, switch jump and average
%\note{These definitions have been rearranged.}
%\item In definition of $V$ on page 3: state which set $D^k$ belongs to.
%\note{This has been added.}
\end{itemize}
\end{document}

### 2017-05-03 docs/dg_penalty/revision/main.tex
updating docs for penalty note + elasticity

--- 2016-12-21 docs/dg_penalty/main.tex
@@
 \documentclass[preprint,10pt]{elsarticle}
+%\usepackage{lineno}
+
 %\documentclass[10pt]{article}
@@
 
-%\usepackage{lineno}
 \usepackage{array} 
@@
 \date{}
-\author{Jesse Chan, T. Warburton}
-\title{A short note on the penalty flux parameter for first order discontinuous Galerkin formulations}
+%\author{Jesse Chan, T. Warburton}
+%\title{A short note on the penalty flux parameter for first order discontinuous Galerkin formulations}
 
@@
 \begin{frontmatter}
-\title{A short note on the penalty flux parameter for first order discontinuous Galerkin formulations}
+\title{A short note on the penalty flux parameter for \note{discontinuous Galerkin formulations of first order hyperbolic systems}}
 
@@
 Penalty fluxes are dissipative numerical fluxes for high order discontinuous Galerkin (DG) methods which depend on a penalization parameter \cite{warburton2013low, ye2016discontinuous}.
-We investigate the dependence of the spectra of high order DG discretizations on this parameter, and show that as its value increases, the spectra of the DG discretization splits into two disjoint sets of eigenvalues.
+We investigate the dependence of the spectra of high order DG discretizations on this parameter, and show that as its value increases, the \note{spectrum} of the DG discretization splits into two disjoint sets of eigenvalues.
 One set converges to the eigenvalues of a conforming discretization, while the other set corresponds to spurious eigenvalues which are damped proportionally to $\tau$.
@@
 \end{frontmatter}
-
-%\begin{abstract}
-%Penalty fluxes are dissipative numerical fluxes for high order discontinuous Galerkin (DG) methods which depend on a penalization parameter \cite{warburton2013low, ye2016discontinuous}.
-We investigate the dependence of the spectra of high order DG discretizations on this parameter, and show that as its value increases, the spectra of the DG discretization splits into two disjoint sets of eigenvalues.
-One set converges to the eigenvalues of a conforming discretization, while the other set corresponds to spurious eigenvalues which are damped proportionally to $\tau$.
-Numerical experiments also demonstrate that undamped spurious modes present in both in the limit of zero and large penalization parameters are damped for moderate values of the upwind parameter.
-%\end{abstract}
 
@@
 The role of this penalty parameter was investigated in \cite{Warburton20063205} for discretizations of the Maxwells equations.
-There, it was shown that in the limit of $\tau\rightarrow\infty$, the spectra of the DG discretization matrix splits into two disjoint sets of eigenvalues, with one set converging to eigenvalues of a conforming discretization and one set associated with divergent spurious eigenvalues.
+There, it was shown that in the limit of $\tau\rightarrow\infty$, the \note{spectrum} of the DG discretization matrix splits into two disjoint sets of eigenvalues, with one set converging to eigenvalues of a conforming discretization and one set associated with divergent spurious eigenvalues.
 
@@
 For $\tau > 0$, penalty fluxes are related to upwind fluxes for problems with continuous coefficients.
-We extend the results of \cite{Warburton20063205} to DG formulations for first order systems with penalty fluxes and show that the spectra of the DG discretization matrix again splits into two disjoint sets of eigenvalues as $\tau\rightarrow\infty$.
-The first set of eigenvalues correspond to the eigenvalues of a conforming discretization, while the second set of eigenvalues have real parts approaching $-\infty$ as $\tau\rightarrow \infty$ and correspond to damped spurious modes.
+We extend the results of \cite{Warburton20063205} to DG formulations for first order systems with penalty fluxes and show that the \note{spectrum} of the DG discretization matrix again splits into two disjoint sets of eigenvalues as $\tau\rightarrow\infty$.
+The first set of eigenvalues correspond to the eigenvalues of a conforming discretization, while the second set of eigenvalues have real parts approaching $-\infty$ as $\tau\rightarrow \infty$ and correspond to damped spurious modes \note{which are non-conforming}.
 Numerical experiments verify these results for the advection and acoustic wave equations in one and two dimensions.
 Numerical experiments also show that, while increasing $\tau$ pushes most eigenvalues further left of the imaginary axis, the real parts of certain eigenvalues converge to zero as $\tau\rightarrow\infty$.
-This implies that taking a moderate penalty parameter $\tau = O(1)$ damps spurious modes which remain otherwise undamped in time-dependent simulations with either central fluxes ($\tau = 0$) and conforming discretizations ($\tau \rightarrow \infty$).
+This implies that taking a moderate penalty parameter $\tau = 1$ damps spurious \note{\emph{conforming}} modes which remain otherwise undamped in time-dependent simulations with conforming discretizations.
+\note{We show that penalized DG methods address both \textit{non-conforming} spurious modes present when using central fluxes ($\tau = 0$) and \textit{conforming} spurious modes which are present under conforming discretizations (or equivalently, under large penalty parameters $\tau \rightarrow \infty$).}
 
@@
 \end{equation}
-with solution components $\bm{U} = (u_1, \ldots, u_M)^T$ and symmetric and spatially constant matrices $\bm{A}_i$.
-We assume the domain $\Omega$ is triangulated exactly using a mesh $\Oh$ consisting of elements $D^k$.
-We define the $L^2$ inner product over elements $D^k$ and their boundaries $\partial D^k$ as
+with solution components $\bm{U} = (u_1, \ldots, u_M)^T$ and matrices $\bm{A}_i \in \mathbb{R}^{M\times M}$.  
+\note{
+For hyperbolic systems, any linear combination of $\bm{A}_i$ 
+\begin{equation}
+\bm{A}_{\alpha} = \sum_{i=1}^d \bm{\alpha}_i \bm{A}_i
+\label{eq:hyperbolic}
+\end{equation}
+is diagonalizable with real eigenvalues.  
+}
+
+We assume the domain $\Omega$ is triangulated exactly using a mesh $\Oh$ consisting of elements $D^k$.  We define the $L^2$ inner product over elements $D^k$ and their boundaries $\partial D^k$ as
 \[
@@
 \begin{equation}
-\sum_{D^k \in \Oh} \LRp{\pd{\bm{U}}{t},\bm{V} }_{\Lk} = \sum_{D^k \in \Oh}\LRp{ \LRp{\bm{A}_i(\bm{U}), \bm{V}_{,i}}_{\Lk} - \LRa{(\bm{A}_n\bm{U})^*,\bm{V}}_{\partial D^k}}
+\sum_{D^k \in \Oh} \LRp{\pd{\bm{U}}{t},\bm{V} }_{\Lk} = \sum_{D^k \in \Oh}\LRp{ \note{\LRp{\sum_{i=1}^d\bm{A}_i \bm{U}, \pd{\bm{V}}{\bm{x}_i}}_{\Lk}} - \LRa{(\bm{A}_n\bm{U})^*,\bm{V}}_{\partial D^k}}
 \label{eq:formulation}
@@
 
-We note that, while the assumptions on constant $\bm{A}_i$ are restrictive, the results in this work are straightforward to generalize to a slight modification of (\ref{eq:cons})
+\note{
+In the main proof of this paper, we make the assumption that the matrices $\bm{A}_i$ are symmetric and spatially constant for clarity and simplicity of presentation.
+However, the numerical fluxes used are more generally applicable (for example, when $\bm{A}_i$ is non-symmetric, spatially varying, or discontinuous), and we present generalizations beyond symmetric and constant $\bm{A}_i$ where possible.
+%More specifically, we 
+%\\
+%Specify what assumptions are made on $\bm{A}_i$ and why.
+%\begin{enumerate}
+%\item Symmetric - 
+%\item Locally constant - energy stable.
+%\item Spatially constant - for simplicity and clarity.
+%\end{enumerate}
+%For the remainder of this paper, we will also assume that $\bm{A}_i$ are piecewise constant.  
+}
+
+Finally, we note that
+%, while the assumptions on constant $\bm{A}_i$ are restrictive, 
+the results in this work are straightforward to generalize to the following modification of (\ref{eq:cons})
 \begin{equation}
@@
 %Results also follow for matrices $\bm{A}_i$ may be symmetrized through an appropriate change of variables \cite{kopriva2014energy}.
-
-%We assume that $\bm{A}_{i}$ are symmetric matrices.
+\note{Additionally, we note that many hyperbolic systems of interest are symmetrizable, and can be cast into the form of (\ref{eq:consA}) with symmetric $\bm{A}_i$ through an appropriate change of variables \cite{kopriva2014energy}.
+}
 
 \subsection{Penalization terms in numerical fluxes}
+\label{sec:scaletau}
 
-Let $f$ be a shared face between a given element $D^{k,-}$ and its neighbor $D^{k,+}$.
+Let $f$ be a face \note{shared between} an element $D^{k,-}$ and its neighbor $D^{k,+}$.
 We denote the outward normal on the face $f$ of $D^{k,-}$ as $\bm{n}$, and let $\bm{U}^+, \bm{U}^-$ be evaluations of $\bm{U}$ restricted to $D^{k,+}$ and $D^{k,-}$, respectively.
-The jump and average scalar-valued functions is defined as
+The jump and average scalar-valued functions are defined as
 \[
-\avg{u} = \frac{{u}^+ + u^-}{2}, \qquad \jump{u} = u^+ - u^-.
+\note{\jump{u} = u^+ - u^-, \qquad \avg{u} = \frac{{u}^+ + u^-}{2}.}
 \]
-Define the normal matrix 
+We define the normal matrix across a face as
 \[
@@
 \]
-The penalty numerical flux is then defined as
+\note{We first introduce the penalty flux for continuous coefficients (where the value of $\bm{A}_n$ is single valued across shared faces) as}
 \[
-(\bm{A}_n\bm{U})^* = \bm{A}_n\avg{\bm{U}} - \frac{\tau}{2} \bm{A}_n^T \bm{A}_n\jump{\bm{U}}.
+(\bm{A}_n\bm{U})^* = \bm{A}_n\avg{\bm{U}} - \frac{\tau}{2} \bm{A}_n^T \bm{A}_n\jump{\bm{U}}, \qquad \tau \geq 0.
 \]
-where $\tau \geq 0$.
-We compare the penalty flux to the well-known upwind numerical flux for the case of continuous coefficients.
-By symmetry of $\bm{A}_i$, ${\bm{A}}_n$ contains real eigenvalues, and admits an eigenvalue decomposition
+We compare the penalty flux to the well-known upwind numerical flux.
+For \note{continuous coefficients, the upwind flux can be written using an eigenvalue decomposition of $\bm{A}_n$ (see, for example, \cite{ye2016discontinuous} and Chapter 2.4 of \cite{hesthaven2007nodal})}.
+%By symmetry of $\bm{A}_i$, ${\bm{A}}_n$ contains real eigenvalues, and admits an eigenvalue decomposition
+Under assumption (\ref{eq:hyperbolic}), ${\bm{A}}_n$ has real eigenvalues, and admits an eigenvalue decomposition
 \[
-\bm{A}_n = \bm{V}{\bm{\Lambda}}\bm{V}^T, \qquad \bm{\Lambda} = 
+\bm{A}_n = \bm{V}{\bm{\Lambda}}\note{\bm{V}^{-1}}, \qquad \bm{\Lambda} = 
 \left(\begin{array}{ccc}
@@
 & & \lambda_d
-\end{array}\right).
+\end{array}\right)
 \]
@@
 \begin{align*}
-\bm{A}_n^+ &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} + \LRb{\bm{\Lambda}}} \bm{V}^T\\
-\bm{A}_n^- &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} - \LRb{\bm{\Lambda}}} \bm{V}^T,
+\bm{A}_n^+ &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} + \LRb{\bm{\Lambda}}} \note{\bm{V}^{-1}}\\
+\bm{A}_n^- &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} - \LRb{\bm{\Lambda}}} \note{\bm{V}^{-1}},
 \end{align*}
@@
 \[
-(\bm{A}_n\bm{U})^* = \bm{A}_n\avg{\bm{U}} - \frac{1}{2}\bm{V}\LRb{\bm{\Lambda}}\bm{V}^T \jump{\bm{U}}.  
+(\bm{A}_n\bm{U})^* = \bm{A}_n\avg{\bm{U}} - \frac{1}{2}\bm{V}\LRb{\bm{\Lambda}}\note{\bm{V}^{-1}} \jump{\bm{U}}.  
 \]
-%For problems with discontinuous coefficients, this may be generalized by solving a Riemann problem to yield
-%\[
-%\bm{F^*}_n = \avg{\bm{A}_n\bm{U}} - \jump{{\bm{V}\LRb{\bm{\Lambda}}\bm{V}^T\bm{U}}}.  
-%\]
+\note{
+Assuming constant, symmetric $\bm{A}_i$ and a penalty numerical flux, one can show the semi-discrete energy estimate \cite{hesthaven2007nodal, warburton2013low, kopriva2014energy, chan2016weight1} for (\ref{eq:cons})
+\[
+\sum_{D^k\in \Oh}\pd{}{t}\nor{\bm{U}}^2_{L^2\LRp{D^k}} = -\sum_{D^k \in \Oh} \frac{\tau}{2}\nor{\bm{A}_n\jump{\bm{U}}}^2_{L^2\LRp{\partial D^k}} \leq 0.
+\]
+The penalization term can thus be interpreted as adding a stabilizing contribution which dissipates components of $\bm{U}$ for which $\bm{A}_n\jump{\bm{U}}$ is large, with larger values of $\tau$ increasing the rate of dissipation (this is true even when $\bm{A}_n$ is not symmetric).
+We will relate this dissipation to the spectrum of the DG discretization matrix in Section~\ref{sec:spec}.
+In particular, we show that the real parts of eigenvalues corresponding to components of eigenvectors with $\bm{A}_n\jump{\bm{U}}\neq 0$ approach $-\infty$ as $\tau \rightarrow \infty$.
+
+When $\bm{A}_n$ is not continuous across faces, the penalty numerical flux can be generalized as follows \cite{ye2016discontinuous}
+\[
+(\bm{A}_n\bm{U})^* = \avg{\bm{A}_n\bm{U}} - \frac{\tau}{2} \bm{A}_n^T \jump{\bm{A}_n\bm{U}},
+\]
+where we define the normal jump and average
+\[
+\jump{\bm{A}_n\bm{U}}\coloneqq \sum_{i=1}^d\bm{n}_i \jump{ \bm{A}_i \bm{U}}, \qquad \avg{\bm{A}_n\bm{U}} \coloneqq \sum_{i=1}^d\bm{n}_i\avg{ \bm{A}_i \bm{U}}.
+\]
+If $\bm{A}_i$ is constant over each element $D^k$, we again have an energy estimate with a dissipative contribution from the penalty flux
+\[
+\sum_{D^k\in \Oh}\pd{}{t}\nor{\bm{U}}^2_{L^2\LRp{D^k}} = -\sum_{D^k \in \Oh} \frac{\tau}{2}\nor{\jump{\bm{A}_n\bm{U}}}^2_{L^2\LRp{\partial D^k}} \leq 0.
+\]
+The upwind flux can similarly be extended for discontinuous coefficients
+\[
+(\bm{A}_n\bm{U})^* = \avg{\bm{A}_n\bm{U}} - \frac{1}{2} \jump{\bm{V}\LRb{\Lambda}\bm{V}^{-1}\bm{U}},
+\]
+where the eigenvalue decompositions inside the jump is evaluated as 
+\begin{align*}
+&\jump{\bm{V}\LRb{\Lambda}\bm{V}^{-1}\bm{U}} = {\bm{V}^+\LRb{\Lambda}^+\LRp{\bm{V}^+}^{-1}\bm{U}^+} - {\bm{V}^-\LRb{\Lambda}^-\LRp{\bm{V}^-}^{-1}\bm{U}^-}\\
+&\bm{A}_n^+ = \sum_{i=1}^d \bm{n}_i \bm{A}_i^+ = \bm{V}^+{\Lambda}^+\LRp{\bm{V}^+}^{-1}, \qquad \bm{A}_n^- = \sum_{i=1}^d \bm{n}_i \bm{A}_i^- = \bm{V}^-{\Lambda}^-\LRp{\bm{V}^-}^{-1}.
+\end{align*} 
+Finally, we note that the energy estimate changes when $\bm{A}_i$ varies over an element, though the penalty flux still adds a dissipative contribution to the right hand side \cite{kopriva2014energy}.
+}
+
+\subsection{Scaling of $\tau$}
+
 We note that including a positive-definite weight matrix $\bm{W}$ into the penalty term
@@
 \]
-recovers the upwind fluxes for $\tau = 1$ and $\bm{W} = \bm{V}\LRb{\bm{\Lambda}}^{\dagger}\bm{V}^{-1}$, where $\bm{A}^{\dagger}$ is the pseudo-inverse of $\bm{A}$.
-However, the advantage of penalty fluxes is that, unlike upwind fluxes, they can be used when explicit knowledge of a diagonalization of $\bm{A}_n$ is not known, such as for linear elasticity in anisotropic media \cite{ye2016discontinuous}.
-Thus, we assume $\bm{W} = \bm{I}$ and require that $\bm{V}\LRb{\bm{\Lambda}}\bm{V}^T$ and $\tau\bm{A}_n^T\bm{A}_n$ are of the same magnitude, which implies that $\tau$ should be proportional to $1/\LRp{\max_i\LRb{\lambda_i}}$.
-This ensures that the penalty term is of the same magnitude as the upwind penalization term, and that CFL restrictions in explicit time-stepping methods will be identical for both upwind and penalty fluxes.
+recovers the upwind fluxes for $\tau = 1$ and $\bm{W} = \bm{V}\LRb{\bm{\Lambda}}^{\dagger}\bm{V}^{-1}$, where $\bm{A}^{\dagger}$ is the pseudo-inverse of $\bm{A}$ \cite{horn2012matrix}.
+However, the advantage of penalty fluxes is that, unlike upwind fluxes, they can be used when explicit knowledge of a \note{joint} diagonalization of $\bm{A}_n$ is not known, such as for linear elasticity in anisotropic media \cite{ye2016discontinuous}.
+Thus, we assume $\bm{W} = \bm{I}$ in this work.
+% instead seek an appropriate scaling $\tau$ such that $\bm{V}\LRb{\bm{\Lambda}}\note{\bm{V}^{-1}}$ and $\tau\bm{A}_n^T\bm{A}_n$ are of the same magnitude.  
+%For $\bm{A}_i$ symmetric, this implies that $\tau$ should be proportional to $1/\LRp{\max_i\LRb{\lambda_i}}$.  
+%This ensures that the penalty term is of the same magnitude as the upwind penalization term, and that CFL restrictions in explicit time-stepping methods will be \note{similar} for both upwind and penalty fluxes.
 
-\section{Dependence of spectra on the penalty parameter}
+\note{It remains to specify an appropriate value of $\tau$.
+We first note that the timestep is proportional to the spectral radius of the DG matrix, which increases proportionally to $\tau$ \cite{chan2015gpu}.
+Since both upwind and penalty fluxes are identical apart from the penalization term, we choose $\tau$ such that the penalty flux and upwind penalization terms are of the same magnitude
+\[
+\nor{\frac{\tau}{2} \bm{A}_n^T\bm{A}_n} = \nor{\frac{1}{2}\bm{V}\LRb{\bm{\Lambda}}{\bm{V}^{-1}}}.
+\]
+This is sufficient to ensure that the spectral radius of the DG matrix is similar when using either upwind or penalty fluxes.
 
-The remainder of this work focuses on the influence of $\tau$ on the spectra of DG discretization matrices.
-We will show that, as $\tau$ increases, the spectra splits into two sets of eigenvalues corresponding respectively to conforming and spurious modes, where conformity is defined based on $\bm{A}_n$.
+The penalty flux and upwind penalization terms can both be bounded from above in terms of the largest magnitude eigenvalue $\max_i\LRb{\lambda_i}$ and the condition number $\kappa\LRp{\bm{V}}$ of the eigenvector matrix $\bm{V}$
+\[
+\nor{\frac{1}{2}\bm{V}\LRb{\bm{\Lambda}}{\bm{V}^{-1}}} \leq \max_i \LRb{\lambda_i} \kappa\LRp{\bm{V}}, \qquad \nor{\bm{A}_n^T\bm{A}_n} = \nor{\bm{V}^{-T}\bm{\Lambda}\bm{V}^T \bm{V}\bm{\Lambda}\bm{V}^{-1}} \leq \max_i\LRb{\lambda_i}^2\kappa\LRp{\bm{V}}^2.
+\]
+Thus, choosing $\tau = 1/\LRp{\max_i\LRb{\lambda_i}\kappa\LRp{\bm{V}}}$ results in a penalization term whose magnitude is the same as that of the upwind penalization term.
+When $\bm{A}_n$ is symmetric (or symmetrizable through a change of variables), $\bm{V}$ is unitary and $\kappa\LRp{\bm{V}} = 1$, simplifying the selection of $\tau$ to $\tau = 1/\max_i\LRb{\lambda_i}$.
+%Since $\bm{A}_n$ is symmetric, $\bm{V}$ is unitary and $\nor{\bm{V}\LRb{\bm{\Lambda}}\bm{V}^{-1}} = \nor{{\bm{\Lambda}}}$.
+Thus, the magnitude of the upwind penalization term is $\max_i\LRb{\lambda_i}$.
+For penalty fluxes, the penalization term is
+%\[
+%\nor{\bm{A}_n^T\bm{A}_n} = \nor{\bm{V}{\bm{\Lambda}}^2\bm{V}^{-1}} = \nor{{\bm{\Lambda}}^2} = \max_{i} \lambda_i^2,
+%\]
+%where we have again used symmetry of $\bm{A}_n$.
+Thus, choosing $\tau = 1/\LRp{\max_i\LRb{\lambda_i}}$ results in a penalization term whose magnitude is the same as that of the upwind penalization term.
+For non-symmetric $\bm{A}_n$, an appropriate scaling of $\tau$ can be determined based on the singular values of $\bm{A}_n$.
+}
+
+\section{Dependence of \note{spectrum} on the penalty parameter}
+\label{sec:spec}
+The remainder of this work focuses on the influence of $\tau$ on the \note{spectrum} of DG discretization matrices.
+We will show that, as $\tau$ increases, the \note{spectrum} splits into two sets of eigenvalues corresponding respectively to conforming and spurious modes, where conformity is defined based on $\bm{A}_n$.
+\note{For clarity and simplicity of presentation, we will assume that $\bm{A}_i$ are symmetric and spatially constant for the remainder of this paper.
+}%We will also assume that all elements $D^k$ are affine mappings of a reference element $\widehat{D}$, which will ensure that }
 
@@
 
-Let $V$ denote the piecewise polynomial approximation space 
+Let $V$ denote the piecewise polynomial approximation space \note{over an element $D^k$}
 \[
-V = \LRc{ \bm{U} \in\LRp{ L^2(\Omega)}^d : \left.\bm{U}\right|_{D^k} \in \LRp{P^N(D^k)}^d}.  
+V = \LRc{ \bm{U} \in\LRp{ L^2(\Omega)}^d : \left.\bm{U}\right|_{D^k} \in \LRp{P^N(D^k)}^d, \quad \note{\forall D^k \in \Oh}}.  
 \]
@@
 The conforming space is defined to be elements of $V$ for which the penalization term vanishes over all faces of all elements.
-For penalty fluxes, this implies $V^C$ is
+For penalty fluxes, \note{this term is $\bm{A}_n^T \bm{A}_n\jump{\bm{U}}$, such that} $V^C$ is
 \begin{equation}
@@
 \end{equation}
+\note{This simplification can be made due to the fact that the null space of $\bm{A}_n^T\bm{A}_n$ is identical to the null space of $\bm{A}_n$.\footnote{\note{The null space of $\bm{A}_n^T\bm{A}_n$ contains the null space of $\bm{A}_n$.
+However, if $\bm{A}_n^T\bm{A}_n\bm{U} = 0$, then $\bm{U}^T\bm{A}_n^T\bm{A}_n\bm{U} = \nor{\bm{A}_n\bm{U}}^2 = 0$ as well, implying that the null space of $\bm{A}_n$ also contains the null space of $\bm{A}_n^T\bm{A}_n$.
+}}
+}  
 The non-conforming approximation space $V^{NC}$ is defined as $L^2$ orthogonal complement of $V^{C}$ in $V$
@@
 \]
-A conforming space could also be defined using the upwind flux by seeking $\bm{U}$ such that $\bm{V}\LRb{\bm{\Lambda}}\bm{V}^T \jump{\bm{U}} = 0$.
-When the coefficient matrices $\bm{A}_i$ are spatially continuous, the condition $\bm{V}\LRb{\bm{\Lambda}}\bm{V}^T \jump{\bm{U}} = 0$ and the penalty flux condition $\bm{A}_n^T \bm{A}_n\jump{\bm{U}} = 0$ are both equivalent to
+A conforming space could also be defined using the upwind flux by seeking $\bm{U}$ such that $\bm{V}\LRb{\bm{\Lambda}}\note{\bm{V}^{-1}} \jump{\bm{U}} = 0$.
+When the coefficient matrices $\bm{A}_i$ are spatially continuous, the condition $\bm{V}\LRb{\bm{\Lambda}}\note{\bm{V}^{-1}} \jump{\bm{U}} = 0$ and the penalty flux condition $\bm{A}_n^T \bm{A}_n\jump{\bm{U}} = 0$ are both equivalent to
 \[
@@
 \]
-where $\bm{\beta}$ is the direction of advection.  For continuous advection vectors such that $\bm{\beta}(\bm{x}) \neq 0$, the penalty flux is 
+where $\bm{\beta}$ is the direction of advection.
+\note{We also assume $\bm{\beta}$ is constant, such that the resulting equation can be rearranged into the form (\ref{eq:cons}).}  The penalty flux is
 \[
-(\bm{A}_n\bm{U})^* = \bm{\beta}_n\avg{ u} - \tau\LRb{\bm{\beta}_n}\jump{u}.
+(\bm{A}_n\bm{U})^* = \bm{\beta}_n\avg{ u} - \tau\frac{\LRp{\bm{\beta}_n}^2}{2}\jump{u}.
 \]
+\note{Setting $\tau = \LRb{\bm{\beta}_n}$ recovers the standard upwind flux.}  
 The conforming space $V^C$ induced by this flux is then
@@
 \begin{align*}
-\frac{1}{c^2}\pd{p}{t} &= \Div \bm{u}\\
-\pd{\bm{u}}{t} &= \Grad p,
+\frac{1}{c^2}\pd{p}{t} + \Div \bm{u} &= 0\\
+\pd{\bm{u}}{t} + \Grad p &= 0,
 \end{align*}
@@
 For the velocity component $V_{\bm{u}}^C$, the penalty fluxes enforce normal continuity, such that $\jump{\bm{u}}\cdot \bm{n} = 0$ over all faces $f \in \Gh$.
-Since each component of $\bm{u}$ is approximated from $P^N(D^k)$, this implies that $V^C_{\bm{u}}$ is the $H({\rm div})$-conforming Brezzi-Douglas-Marini finite element space \cite{brezzi1985two,boffi2013mixed}.\footnote{It is possible to recover $V^C_{\bm{u}}$ as the Raviart-Thomas finite element space by approximating each component of $\bm{u}$ from a different polynomial space \cite{kirby2004algorithm}.}   These reflect minimal continuity conditions which result from requiring $\Div \bm{u} \in \LRp{L^2\LRp{\Oh}}^d$.
+Since each component of $\bm{u}$ is approximated from $P^N(D^k)$, this implies that $V^C_{\bm{u}}$ is the $H({\rm div})$-conforming Brezzi-Douglas-Marini finite element space \cite{brezzi1985two,boffi2013mixed} \note{on triangular and tetrahedral elements}.\footnote{It is possible to recover $V^C_{\bm{u}}$ as the Raviart-Thomas finite element space by approximating each component of $\bm{u}$ from a different polynomial space \cite{kirby2004algorithm}.}   These reflect minimal continuity conditions which result from requiring $\Div \bm{u} \in \LRp{L^2\LRp{\Oh}}^d$.
 
@@
 where $\bm{M}$ is the $L^2$ mass matrix.
-We adapt the approach of Warburton and Embree in \cite{Warburton20063205} to (\ref{eq:block}) to show that the spectra of $\bm{K}$ splits into two sets of eigenvalues as $\tau\rightarrow \infty$, the first of which approaches the eigenvalues of $\bm{A}$, and the second of which diverges with real part approaching $-\infty$ as $\tau\rightarrow \infty$.
+We adapt the approach of Warburton and Embree in \cite{Warburton20063205} to (\ref{eq:block}) to show that the \note{spectrum} of $\bm{K}$ splits into two sets of eigenvalues as $\tau\rightarrow \infty$, the first of which approaches the eigenvalues of $\bm{A}$, and the second of which diverges with real part approaching $-\infty$ as $\tau\rightarrow \infty$.
 
-Let $\LRc{\bm{\Phi}_i }_{i=1}^{N^{C}}$ and $\LRc{\bm{\Psi}_i }_{i=1}^{N^{NC}}$ be $L^2$-orthogonal bases for $\bm{V}^C$ and $\bm{V}^{NC}$, respectively.
-These spaces then induce a block decomposition of the DG discretization matrix $\bm{K}$
+\note{We first note that integrating by parts the right hand side of (\ref{eq:formulation}) and manipulating interface terms yields the following rearrangement of the DG formulation
+\begin{align}
+b(\bm{U},\bm{V}) &=\sum_{D^k\in \Oh} \frac{1}{2}\LRp{ \LRp{\sum_{i=1}^d\bm{A}_i\bm{U}, \pd{\bm{V}}{\bm{x}_i}}_{D^k} - \LRp{\sum_{i=1}^d\bm{A}_i \pd{\bm{U}}{\bm{x}_i}, {\bm{V}}}_{D^k}} \label{eq:dgformskew}\\
+&+ \sum_{D^k\in \Oh}\frac{1}{2}\LRp{\LRa{\bm{A}_n\jump{\bm{U}}, \avg{\bm{V}}}_{\partial D^k}-\LRa{\avg{\bm{U}}, \bm{A}_n\jump{\bm{V}}}_{\partial D^k}} - \frac{\tau}{2}\LRa{\bm{A}_n\jump{\bm{U}},\bm{A}_n\jump{\bm{V}}}_{\partial D^k}.
+\nonumber
+\end{align}
+The formulation is skew-symmetric (assuming $\bm{A}_i$ are constant and symmetric), with the exception of the symmetric semi-definite $\tau$-dependent penalization term.}
+
+Let $\bm{K}$ denote the DG discretization matrix resulting from (\ref{eq:dgformskew}).
+Let $\LRc{\bm{\Phi}_i }_{i=1}^{N^{C}}$ and $\LRc{\bm{\Psi}_i }_{i=1}^{N^{NC}}$ now be $L^2$-orthogonal bases for $\bm{V}^C$ and $\bm{V}^{NC}$, respectively.
+These spaces induce a block decomposition of $\bm{K}$
 \begin{equation}
@@
 \bm{A} & \bm{B}\\
-\tilde{\bm{B}}^T & \bm{C} + \tau \bm{S}
+-{\bm{B}}^T & \bm{C} + \tau \bm{S}
 \end{array}\right),
@@
 \end{equation}
+\note{
 where the blocks $\bm{A},\bm{B},\bm{C}$ are defined as
 \begin{align*}
-\bm{A}_{mn} &= \sum_{D^k\in \Oh} \LRp{\bm{A}_i\bm{\Phi}_n, \bm{\Phi}_{m,i}}_{D^k}, \qquad 1\leq m,n \leq N^C\\
-\bm{B}_{mn} &= \sum_{D^k\in \Oh} \LRp{\bm{A}_i\bm{\Psi}_n, \bm{\Phi}_{m,i}}_{D^k} - \LRa{\avg{\bm{A}_n\bm{\Psi}_i}, {\bm{\Phi}_i}}_{\partial D^k} \qquad 1\leq m \leq N^C, \quad 1\leq n \leq N^{NC} \\
-\bm{C}_{mn}  &= -\sum_{D^k\in \Oh} \LRa{\avg{\bm{A}_n\bm{\Psi}_i}, {\bm{\Psi}_i}}_{\partial D^k}, \qquad 1 \leq m,n \leq N^{NC}\\
-\bm{S}_{mn}  &= \frac{1}{2}\sum_{D^k\in \Oh} \LRa{\bm{A}_n^T\jump{\bm{A}_n \bm{\Psi_n}}, {\bm{\Psi}_m}}_{\partial D^k}, \qquad 1 \leq m,n \leq N^{NC}.  
+\bm{A}_{mn} &= \frac{1}{2}\sum_{D^k\in \Oh} \LRp{ \sum_{i=1}^d\LRp{\LRp{\bm{A}_i\bm{\Phi}_n, \pd{\bm{\Phi}_m}{\bm{x}_i}}_{D^k} - \LRp{\bm{A}_i \pd{\bm{\Phi}_n}{\bm{x}_i}, {\bm{\Phi}_m}}_{D^k}}}, \qquad 1\leq m,n \leq N^C\\
+\bm{B}_{mn} &= \frac{1}{2}\sum_{D^k\in \Oh} \LRp{ \sum_{i=1}^d\LRp{\LRp{\bm{A}_i\bm{\Phi}_n, \pd{\bm{\Psi}_m}{\bm{x}_i}}_{D^k} - \LRp{\bm{A}_i \pd{\bm{\Phi}_n}{\bm{x}_i}, {\bm{\Psi}_m}}_{D^k}} - {\LRa{\avg{\bm{\Phi}_n}, \bm{A}_n\jump{\bm{\Psi}_m}}_{\partial D^k}} }, \quad 1\leq n \leq N^{NC} \\
+\bm{C}_{mn}  &=  \frac{1}{2}\sum_{D^k\in \Oh}\LRp{
+\sum_{i=1}^d\LRp{\LRp{\bm{A}_i\bm{\Psi}_n, \pd{\bm{\Psi}_m}{\bm{x}_i}}_{D^k} - \LRp{\bm{A}_i \pd{\bm{\Psi}_n}{\bm{x}_i}, {\bm{\Psi}_m}}_{D^k}}}\\
+&\qquad + \frac{1}{2}\sum_{D^k\in \Oh}\LRp{\LRa{\bm{A}_n\jump{\bm{\Psi}_n}, \avg{\bm{\Psi}_m}}_{\partial D^k}-\LRa{\avg{\bm{\Psi}_n}, \bm{A}_n\jump{\bm{\Psi}_m}}_{\partial D^k}}, \qquad 1 \leq m,n \leq N^{NC}\\
+\bm{S}_{mn}  &= -\sum_{D^k\in \Oh}  \frac{1}{2}\LRa{\bm{A}_n\jump{\bm{\Psi}_n},\bm{A}_n\jump{\bm{\Psi}_m}}_{\partial D^k}, \qquad 1 \leq m,n \leq N^{NC}.  
 \end{align*}
-For exact quadrature, constant $\bm{A}_i$, and $\tau = 0$, the DG formulation is skew-symmetric \cite{warburton2013low, chan2015gpu, chan2016weight2, kopriva2014energy}, implying that $\bm{A},\bm{C}$ are skew-symmetric and that $\tilde{\bm{B}} = -\bm{B}$.
-Additionally, the entries of the penalization matrix $\bm{S}$ can be rewritten as a sum over all unique faces of elements in the mesh
+where we have used that, because $\bm{\Phi}_n\in V^{C}$, $\bm{A}_n\jump{\bm{\Phi}_n} = 0$.}
+%For exact quadrature, constant $\bm{A}_i$, and $\tau = 0$, the DG formulation is skew-symmetric \cite{warburton2013low, chan2015gpu, chan2016weight2, kopriva2014energy}, implying that $\bm{A},\bm{C}$ are skew-symmetric and that $\tilde{\bm{B}} = -\bm{B}$.
+
+%The entries of the penalization matrix $\bm{S}$ can be rewritten as %a sum over all unique faces of elements in the mesh
+%\[
+%%\bm{S}_{mn}  = \frac{1}{2}\sum_{D^k\in \Oh} \LRa{\bm{A}_n\jump{ \bm{\Psi}_n}, \bm{A}_n{\bm{\Psi}_m}}_{\partial D^k} = -\frac{1}{2}\sum_{f \in \Gh} \LRa{\jump{\bm{A}_n \bm{\Psi}_n}, \jump{\bm{A}_n{\bm{\Psi}_m}}}_{f},
+%\note{\bm{S}_{mn}  = \frac{1}{2}\sum_{D^k\in \Oh} \LRa{\bm{A}_n\jump{ \bm{\Psi}_n}, \bm{A}_n{\bm{\Psi}_m}}_{\partial D^k} = -\frac{1}{4}\sum_{D^k\in \Oh} \LRa{\bm{A}_n\jump{ \bm{\Psi}_n}, \bm{A}_n\jump{{\bm{\Psi}_m}}}_{f},}
+%\]
+%for consistently defined jumps and normal vectors $\bm{n}$  over each unique face.  
+\note{We note that $\bm{A}$ is skew-symmetric, and $\bm{S}$ is symmetric and negative semi-definite.
+To show that $\bm{S}$ is also negative definite, assume that there exists some $\bm{v} \in V^{NC}$ with degrees of freedom $\bm{V}$ such that $\bm{S}\bm{V} = 0$ and
 \[
-\bm{S}_{mn}  &= \frac{1}{2}\sum_{D^k\in \Oh} \LRa{\jump{\bm{A}_n \bm{\Psi}_n}, \bm{A}_n{\bm{\Psi}_m}}_{\partial D^k} = -\frac{1}{2}\sum_{f \in \Gh} \LRa{\jump{\bm{A}_n \bm{\Psi}_n}, \jump{\bm{A}_n{\bm{\Psi}_m}}}_{f},
+\bm{V}^T\bm{S}\bm{V} = -\frac{1}{4}\sum_{D^k \in \Oh} \int_{\partial D^k} \LRp{\bm{A}_n\jump{ \bm{v}_n}}^2 = 0.
 \]
-for consistently defined jumps and normal vectors $\bm{n}$  over each unique face.  This implies $\bm{S}$ is symmetric and negative-definite.
+However, because each term in the summation on the right hand side is negative definite, we conclude that $\bm{A}_n\jump{ \bm{v}_n} = 0$ across all mesh interfaces.
+This would imply that $\bm{v}\in V^C$, contradicting the assumption that $\bm{v} \in V^{NC}$.}  %From this, we conclude that} $\bm{S}$ is symmetric and negative-definite.
 
-%The generalized eigenvalue problem $\bm{K}\bm{u}=\lambda_i\bm{M}\bm{u}$ is then equivalent to the eigenvalue problem $\bm{K}\bm{u} = \lambda_i\bm{u}$.  
-%We begin by considering the spectra of $\bm{C} + \tau\bm{S}$.  
-%Let $\lambda, \bm{u}$ be an eigenpair of $\bm{C} + \tau{\bm{S}}$, such that $\lambda = \alpha + i\beta$ and $\bm{u} = \bm{v} + i\bm{w}$, where $\alpha,\beta$ and $\bm{v},\bm{w}$ are the real and imaginary parts of $\lambda$ and $\bm{u}$, respectively.
-Then, expanding and grouping terms in
-%\[
-%(\bm{C}+\tau\bm{S})(\bm{v} + i\bm{w}) = (\alpha + i\beta) (\bm{v} + i\bm{w})
-%\]
-%and using the fact that $\bm{C} + \tau\bm{S}$ is real-valued, we have that
-%\begin{align*}
-%(\bm{C}+\tau\bm{S})\bm{v} &= \alpha\bm{v}-\beta\bm{w}\\
-%(\bm{C}+\tau\bm{S})\bm{w} &= \beta\bm{v}+\alpha\bm{w}.
-%\end{align*}
-%%Assuming that $\nor{\bm{u}}^2 = \nor{\bm{v}}^2 + \nor{\bm{w}}^2 = 1$, multiplying both sides by $\bm{v}^T,\bm{w}^T$ and straightforward manipulations %yield
-%%\begin{align*}
-%%\bm{v}^T\bm{A}\bm{v} &= \alpha\nor{\bm{v}}^2-\beta\bm{v}^T\bm{w}\\
-%%\bm{w}^T\bm{A}\bm{w} &= \beta\bm{w}^T{\bm{v}}+\alpha\nor{\bm{w}}^2.  
-%%\end{align*}
-%%\begin{align*}
-%%\bm{v}^T\bm{A}\bm{v} + \bm{w}^T\bm{A}\bm{w} &= \alpha\\
-%%\bm{v}^T\bm{A}\bm{w} - \bm{w}^T\bm{A}\bm{w} &= \beta.
-%%\end{align*}
-%The symmetry of $\bm{S}$ then yields
-%\begin{align*}
-%\bm{v}^T\bm{C}\bm{w} - \bm{w}^T\bm{C}\bm{w} &= \beta\\
-%%\LRp{\bm{v}^T\bm{C}\bm{v} + \bm{w}^T\bm{C}\bm{w}} + 
-%\tau\LRp{\bm{v}^T\bm{S}\bm{v} + \bm{w}^T\bm{S}\bm{w}} &= \alpha.
-%\end{align*}
-%Since $\bm{v}^T\bm{C}\bm{w} - \bm{w}^T\bm{C}\bm{w}= \beta$ is independent of $\tau$ and $\bm{v},\bm{w}$ are normalized, $\beta$ must remain bounded as $\tau\rightarrow \infty$.
-Similarly, since $\bm{S}$ is symmetric, negative-definite, and independent of $\tau$, the quantity
-%\[
-%\LRb{\bm{v}^T\bm{S}\bm{v} + \bm{w}^T\bm{S}\bm{w}} = \LRb{\frac{\alpha}{\tau}}
-%\]
-%is bounded independently of $\tau$, and we conclude that $\LRb{\alpha} = O(\tau)$.
-Thus, as $\tau$ increases, the eigenvalues of $\bm{C} + \tau\bm{S}$ are shifted further and further left of the imaginary axis.
-
-We may now show how the spectra of $\bm{K}$ behaves as $\tau \rightarrow \infty$.
+We may now show how the \note{spectrum} of $\bm{K}$ behaves as $\tau \rightarrow \infty$.
 Since $\bm{A}$ is skew-symmetric and $\bm{S}$ is symmetric, they are diagonalizable under unitary matrices $\bm{U}$ and $\bm{Q}$, whose columns contain the eigenvectors of $\bm{A}$ and $\bm{S}$, respectively.
@@
 For sufficiently large $\tau$, these disks must become disjoint, implying that the union of disks centered around  $\lambda^C_i$ contains exactly $N^C$ eigenvalues, while the union of disks centered around  $\lambda^{NC}_i$ contain exactly $N^{NC}$ eigenvalues \cite{horn2012matrix}.
-In other words, the spectra of $\bm{K}$ diverges into two distinct sets of $N^C$ and $N^{NC}$ eigenvalues each as $\tau \rightarrow \infty$, with the set of $N^{NC}$ eigenvalues containing negative real parts with magnitude $O(\tau)$.
+In other words, the \note{spectrum} of $\bm{K}$ diverges into two distinct sets of $N^C$ and $N^{NC}$ eigenvalues each as $\tau \rightarrow \infty$, with the set of $N^{NC}$ eigenvalues containing negative real parts with magnitude $O(\tau)$.
 
-Consider now the $N^C$ eigenvalues with the smallest magnitude.
+Consider now the $N^C$ eigenvalues \note{ $\lambda_1, \ldots, \lambda_{N^C}$} of smallest magnitude.
 By the Gerschgorin argument, these must remain bounded as $\tau \rightarrow \infty$.
-Let $\bm{W} = (\bm{W}^C,\bm{W}^{NC})$ and $\bm{\Lambda}^C$ be the matrix of eigenvectors and eigenvalues corresponding to these $N^C$ smallest eigenvalues, such that
+\note{Let $\bm{W}$ denote the matrix of eigenvectors and $\bm{\Lambda}^C$ the matrix of eigenvalues corresponding to these $N^C$ smallest eigenvalues, such that
+\[
+\bm{W} = \LRp{\begin{array}{c}
+\bm{W}^C\\
+\bm{W}^{NC}
+\end{array}}, \qquad 
+\bm{\Lambda}^C = \LRp{\begin{array}{cccc}
+\lambda_1 & & &\\
+& \lambda_2 & &\\
+& & \ddots &\\
+& & & \lambda_{N^C}
+\end{array}
+}.
+ \]
+Here, $\bm{W}^C, \bm{W}^{NC}$ refer to components of eigenvectors from the conforming and non-conforming spaces $V^C, V^{NC}$, respectively.
+These eigenvector matrices have $N^C$ columns, and $\bm{W}^C, \bm{W}^{NC}$ have $N^{C}$ and $N^{NC}$ rows, respectively.
+By definition, $\bm{W}^C, \bm{W}^{NC}$ satisfy}
 \[
@@
 = 
-\bm{\Lambda}_C
 \left(\begin{array}{c}
@@
 \end{array}\right) 
+\note{\bm{\Lambda}_C}.
 \]
@@
 \]
-Since $\bm{\Lambda}^C$ remains bounded as $\tau\rightarrow \infty$, the right hand side is bounded independently of $\tau$ under normalization of $\bm{W}^C,\bm{W}^{NC}$, implying that the non-conforming component $\bm{W}^{NC}$ satisfies
+Since $\bm{\Lambda}^C$ remains bounded as $\tau\rightarrow \infty$, the right hand side is bounded independently of $\tau$ \note{(assuming $\bm{W}$ is normalized)}, implying that the non-conforming component $\bm{W}^{NC}$ satisfies
 \[
@@
 \end{lemma}
-We note that, while we consider skew-symmetric DG formulations in this work, Lemma~\ref{lemma:eig} is straightforward to generalize to any DG formulations where $\bm{S}$ is symmetric negative definite and $\bm{A}$ is diagonalizable.
+We note that, while we consider skew-symmetric DG formulations in this work, Lemma~\ref{lemma:eig} is straightforward to generalize to any DG formulations where $\bm{S}$ is symmetric negative definite and $\bm{A}$ is diagonalizable.%, \note{including the case when the coefficient matrices $\bm{A}_i$ are not symmetric.}
 
@@
 
-Lemma~\ref{lemma:eig} illustrates the behavior of the spectra of the DG discretization for the asymptotic cases when $\tau = 0$ or $\tau \rightarrow \infty$.
-However, it is less clear how the spectra of $\bm{K}$ behaves for $\tau \approx 1$.
+Lemma~\ref{lemma:eig} illustrates the behavior of the \note{spectrum} of the DG discretization for the asymptotic cases when $\tau = 0$ or $\tau \rightarrow \infty$.
+However, it is less clear how the \note{spectrum} of $\bm{K}$ behaves for $\tau \approx 1$.
 We rely instead on numerical experiments in one and two space dimensions to illustrate behaviors for the advection and acoustic wave equations.
-In particular, these experiments indicate that certain modes with negative real part for $\tau = O(1)$ return to the imaginary axis as $\tau \rightarrow \infty$.
+{\note{Since $\tau$ can be chosen to be any arbitrary non-negative value, we consider three cases: $\tau = 0$, $\tau = 1$, and $\tau \rightarrow \infty$.
+The choice of $\tau = 1$ is somewhat arbitrary, and (as mentioned in Section~\ref{sec:scaletau}) it may be beneficial to scale $\tau$ with problem parameters.
+However, for the simple model problems considered in this section, taking $\tau = 1$ recovers an upwind flux, and the resulting spectrum of the DG operator is clearly distinct from the cases when $\tau = 0$ or $\tau \rightarrow \infty$.}  }
+
+These experiments indicate that certain modes with negative real part for $\tau = 1$ return to the imaginary axis as $\tau \rightarrow \infty$.
 Since eigenvalues with a negative real part correspond to dissipation in time-domain simulations, this implies that these under-resolved modes become undamped as $\tau$ grows sufficiently large.
@@
 
-We consider first the scalar advection equation in 1D with periodic boundary conditions.
+We consider first the scalar advection equation in 1D with $\bm{\beta} = 1$ and periodic boundary conditions.
 For $\tau = 0$, the eigenvalues of the DG discretization matrix are purely imaginary.
-Paths taken by eigenvalues as $\tau$ increases are determined by sampling spectra over a sufficiently fine set of $\tau$ and using a particle tracking method \cite{simpletracker}.
+Paths taken by eigenvalues as $\tau$ increases are determined by sampling \note{spectrum} over a sufficiently fine set of $\tau$ and using a particle tracking method \cite{simpletracker}.
 Figure~\ref{fig:track1D} shows the paths taken by these eigenvalues as $\tau$ increases from zero to $\tau = 4$.
@@
 
-Figure~\ref{fig:spurious} also indicates that taking $\tau = O(1)$ suppresses spurious non-dissipative modes associated with both $\tau = 0$ (central fluxes) and $\tau\rightarrow \infty$ (conforming discretizations).
+Figure~\ref{fig:spurious} also indicates that taking $\tau = 1$ suppresses spurious non-dissipative modes associated with both $\tau = 0$ (central fluxes) and $\tau\rightarrow \infty$ (conforming discretizations).
 Let $\lambda^{\tau}_i, \bm{v}^{\tau}_i$ be eigenvalues and eigenvectors of the DG discretization matrix $\bm{K}$ with penalty $\tau$.
@@
 
-Figure~\ref{fig:waveeigs} shows the spectra of the DG discretization matrix for the acoustic wave equation for $\tau = 1, 10, 50$.
+Figure~\ref{fig:waveeigs} shows the \note{spectrum} of the DG discretization matrix for the acoustic wave equation for $\tau = 1, 10, 50$.
 Like the one-dimensional case, a subset of eigenvalues diverge towards the left half plane as $\tau$ increases, though unlike the one-dimensional case, these eigenvalues collapse towards the real line as $\tau$ increases.
@@
 
-For the advection equation with periodic boundary conditions, the spectra of the DG discretization behaves similarly.
+For the advection equation with periodic boundary conditions, the \note{spectrum} of the DG discretization behaves similarly.
 In all following experiments, we use advection vector $\beta = (1,0)$.
-For $\tau=0$, all eigenvalues lie on the imaginary axis, and as $\tau$ increases, the spectra splits into eigenvalues which diverge towards the left half plane and eigenvalues which return to the imaginary axis.
-Figure~\ref{fig:trackModesAdvec} shows the spectra of the DG matrix for $\tau = 0,1,50$, with eigenvalue paths for increasing values of $\tau$ overlaid.
+For $\tau=0$, all eigenvalues lie on the imaginary axis, and as $\tau$ increases, the \note{spectrum} splits into eigenvalues which diverge towards the left half plane and eigenvalues which return to the imaginary axis.
+Figure~\ref{fig:trackModesAdvec} shows the \note{spectrum} of the DG matrix for $\tau = 0,1,50$, with eigenvalue paths for increasing values of $\tau$ overlaid.
 Unlike spurious conforming modes present for the acoustic wave equation, spurious conforming modes for the advection equation satisfy $\LRb{\bm{\beta}_n}\jump{ u} = 0$, and allow discontinuities along faces which lie tangential to flow directions (Figure~\ref{fig:trackModesAdvecU}).
@@
 
-Numerical experiments also demonstrate that, for $\tau = O(1)$, there exist eigenvalues with negative real part which return to the imaginary axis as $\tau \rightarrow 0$ or $\tau\rightarrow \infty$.
+Numerical experiments also demonstrate that, for $\tau = 1$, there exist eigenvalues with negative real part which return to the imaginary axis as $\tau \rightarrow 0$ or $\tau\rightarrow \infty$.
 The corresponding eigenmodes of such eigenvalues can be interpreted as spurious under-resolved modes which persist in time-domain simulations.

### 2017-05-03 docs/dg_penalty/test.tex
updating docs for penalty note + elasticity

--- 2016-10-23 docs/dg_penalty/main.tex
@@
 
-\usepackage{fullpage}
-\usepackage{hyperref}
-%\let\proof\relax
-%\let\endproof\relax
-\usepackage{amsmath,amssymb,amsfonts,mathrsfs}%,amsthm}
-\usepackage[titletoc,toc,title]{appendix}
-
-%\usepackage{lineno}
-\usepackage{array} 
-\usepackage[utf8]{inputenc}
-\usepackage{listings}
-\usepackage{mathtools}
-\usepackage{pdfpages}
-\usepackage[textsize=footnotesize,color=green]{todonotes}
-\usepackage{bm}
-%\usepackage{tikz}
-\usepackage[normalem]{ulem}
-\usepackage{hhline}
-
-%% ====================================== alg package
-\usepackage{algorithm}
-\usepackage[noend]{algpseudocode}
-\usepackage{algorithmicx}
-\algblock{ParFor}{EndParFor}
-% customising the new block
-\algnewcommand\algorithmicparfor{\textbf{parfor}}
-\algnewcommand\algorithmicpardo{\textbf{do}}
-\algnewcommand\algorithmicendparfor{\textbf{end\ parfor}}
-\algrenewtext{ParFor}[1]{\algorithmicparfor\ #1\ \algorithmicpardo}
-\algrenewtext{EndParFor}{\algorithmicendparfor}
-%% ====================================== end alg package
-
-\usepackage{graphicx}
-\usepackage{subfig}
-\usepackage{color}
-
-%% ====================================== graphics
-
-\usepackage{pgfplots}
-\usepackage{pgfplotstable}
-\definecolor{markercolor}{RGB}{124.9, 255, 160.65}
-\pgfplotsset{width=10cm,compat=1.9}
-\pgfplotsset{
-tick label style={font=\small},
-label style={font=\small},
-legend style={font=\small}
-}
-
-\usetikzlibrary{calc}
-\usepackage{stmaryrd}
-
-\renewcommand{\topfraction}{0.85}
-\renewcommand{\textfraction}{0.1}
-\renewcommand{\floatpagefraction}{0.75}
-
-\newcommand{\vect}[1]{\ensuremath\boldsymbol{#1}}
-\newcommand{\tensor}[1]{\underline{\bm{#1}}}
-\newcommand{\del}{\triangle}
-\newcommand{\curl}{\grad \times}
-\renewcommand{\div}{\grad \cdot}
-\newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}{\rm #2}}}
-\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
-\newcommand{\pdd}[2]{\frac{\partial^2#1}{\partial#2^2}}
-
-\newcommand{\bs}[1]{\boldsymbol{#1}}
-
-\newcommand{\equaldef}{\stackrel{\mathrm{def}}{=}}
-
-\newcommand{\tablab}[1]{\label{tab:#1}}
-\newcommand{\tabref}[1]{Table~\ref{tab:#1}}
-
-\newcommand{\theolab}[1]{\label{theo:#1}}
-\newcommand{\theoref}[1]{\ref{theo:#1}}
-\newcommand{\eqnlab}[1]{\label{eq:#1}}
-\newcommand{\eqnref}[1]{\eqref{eq:#1}}
-\newcommand{\seclab}[1]{\label{sec:#1}}
-\newcommand{\secref}[1]{\ref{sec:#1}}
-\newcommand{\lemlab}[1]{\label{lem:#1}}
-\newcommand{\lemref}[1]{\ref{lem:#1}}
-
-\newcommand{\mi}{\mathrm{i}} %% roman "i"
-\newcommand{\mb}[1]{\mathbf{#1}}
-\newcommand{\mbb}[1]{\mathbb{#1}}
-\newcommand{\mc}[1]{\mathcal{#1}}
-\newcommand{\nor}[1]{\left\| #1 \right\|}
-\newcommand{\snor}[1]{\left| #1 \right|}
-\newcommand{\LRp}[1]{\left( #1 \right)}
-\newcommand{\LRs}[1]{\left[ #1 \right]}
-\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
-\newcommand{\LRb}[1]{\left| #1 \right|}
-\newcommand{\LRc}[1]{\left\{ #1 \right\}}
-\newcommand{\LRceil}[1]{\left\lceil #1 \right\rceil}
-
-\newcommand{\Grad} {\ensuremath{\nabla}}
-\newcommand{\Div} {\ensuremath{\nabla\cdot}}
-\newcommand{\Nel} {\ensuremath{{N^\text{el}}}}
-\newcommand{\jump}[1] {\ensuremath{\llbracket#1\rrbracket}}
-\newcommand{\avg}[1] {\ensuremath{\LRc{\!\{#1\}\!}}}
-\newcommand{\uh}{\widehat{u}}
-\newcommand{\Bh}{\widehat{B}}
-\newcommand{\fnh}{\widehat{f}_n}
-\renewcommand{\L}{L^2\LRp{\Omega}}
-\newcommand{\Lk}{L^2\LRp{D^k}}
-\newcommand{\Dhat}{\widehat{D}}
-\newcommand{\Lhat}{L^2\LRp{\Dhat}}
-\newcommand{\pO}{\partial\Omega}
-\newcommand{\Gh}{\Gamma_h}
-\newcommand{\Gm}{\Gamma_{-}}
-\newcommand{\Gp}{\Gamma_{+}}
-\newcommand{\Go}{\Gamma_0}
-\newcommand{\Oh}{\Omega_h}
-
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-%\newtheorem{proposition}[theorem]{Proposition}
-%\newtheorem{corollary}[theorem]{Corollary}
-
-%\newenvironment{definition}[1][Definition]{\begin{trivlist}
-%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
-%\newenvironment{example}[1][Example]{\begin{trivlist}
-%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
-
-\newcommand{\eval}[2][\right]{\relax
-  \ifx#1\right\relax \left.\fi#2#1\rvert}
-
-\def\etal{{\it et al.~}}
-
-\def\arr#1#2#3#4{\left[
-\begin{array}{cc}
-#1 & #2\\
-#3 & #4\\
-\end{array}
-\right]}
-\def\vectwo#1#2{\left[
-\begin{array}{c}
-#1\\
-#2\\
-\end{array}
-\right]}
-\def\vecthree#1#2#3{\left[
-\begin{array}{c}
-#1\\
-#2\\
-#3\\
-\end{array}
-\right]}
-\def\vectfour#1#2#3#4{\left[
-\begin{array}{c}
-#1\\
-#2\\
-#3\\
-#4\\
-\end{array}
-\right]}
-
-\newcommand{\G} {\Gamma}
-\newcommand{\Gin} {\Gamma_{in}}
-\newcommand{\Gout} {\Gamma_{out}}
-
-\newcommand{\note}[1]{{\color{blue}#1}}
-\newcommand{\remark}[1]{\textbf{\color{red}#1}}
-
-\newcommand{\tri}{{\rm tri}}
-\newcommand{\sqr}{{\rm quad}}
-
-\newcommand{\refhex}{\widehat{\mathcal{H}}}
-\newcommand{\reftet}{\widehat{\mathcal{T}}}
-\newcommand{\refwedg}{\widehat{\mathcal{W}}}
-
-\newcommand{\refpyr}{\widehat{\mathcal{P}}}
-\newcommand{\refpyrf}{\widehat{\mathcal{P}}^f}
-\newcommand{\refpyrfq}{\widehat{\mathcal{P}}^{\sqr}}
-\newcommand{\refpyrft}{\widehat{\mathcal{P}}^{\tri}}
-
-\newcommand{\hex}{{\mathcal{H}}}
-\newcommand{\tet}{{\mathcal{T}}}
-\newcommand{\wedg}{{\mathcal{W}}}
-
-\newcommand{\pyr}{{\mathcal{P}}}
-\newcommand{\pyrf}{{\mathcal{P}}^f}
-\newcommand{\pyrfq}{{\mathcal{P}}^{\sqr}}
-\newcommand{\pyrft}{{\mathcal{P}}^{\tri}}
-
-\newcommand{\LinfDk}{L^{\infty}\LRp{D^k}}
-
-\newcommand{\diag}[1]{{\rm diag}\LRp{#1}}
-
-\newcommand{\half}{1/2}
-
-\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-
-%% d in integrand
-\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}}
-
-\makeatletter
-\renewcommand\d[1]{\mspace{6mu}\mathrm{d}#1\@ifnextchar\d{\mspace{-3mu}}{}}
-\makeatother
-
-\newcommand{\squareface}{[-1,1]^2}
-
 \date{}
-\author{Jesse Chan, T. Warburton}
-\title{On penalty flux parameters in first order DG formulations}
 
@@
 \maketitle
-%\tableofcontents
 
-\begin{abstract}
-Abstract goes here.  Mention assumptions: continuously varying material data, skew-symmetric formulations.  
-\end{abstract}
-
-\section{Introduction}
-
-\note{Intro}
-
-\section{Semi-discrete discontinuous Galerkin formulation}
-
-We consider linear first order hyperbolic systems of PDEs in $\mathbb{R}^d$
-\begin{equation}
-\bm{A}_0(\bm{x})\pd{\bm{U}}{t} + \sum_{i=1}^d \pd{\LRp{\bm{A}_{i}\bm{U}}}{\bm{x}_i} = 0
-\label{eq:cons}
-\end{equation}
-with solution components $\bm{U} = (u_1, \ldots, u_M)^T$.
-In this work, we assume that $\bm{A}_0$ is symmetric and positive-definite, and that the matrices $\bm{A}_i$ are symmetric or symmetrizable.
-In this work, we assume that $\bm{A}_i$ are constant, though $\bm{A}_0(\bm{x})$ may vary spatially.
-These assumptions are true for many PDEs governing wave propagation in heterogeneous media, including such as the acoustic wave equation \cite{chan2016weight1}, linear elasticity \cite{de2008interior,wilcox2010high,ye2016discontinuous}, and Maxwell's equations \cite{hesthaven2002nodal,grote2007interior,warburton2013low}.
-%\note{Restrict to constant coeff $\bm{A}_i$?}
-
-We assume the domain $\Omega$ is triangulated exactly using a mesh $\Oh$ consisting of elements $D^k$.  We define the $L^2$ inner product over elements $D^k$ and their boundaries $\partial D^k$ as
-\[
-\LRp{u,v}_{D^k} = \int_{D^k} u(\bm{x})v(\bm{x}) \diff x, \qquad \LRa{u,v}_{\partial D^k} = \int_{\partial D^k} u(\bm{x})v(\bm{x}) \diff x.
-\]
-A semi-discrete DG formulation for (\ref{eq:cons}) may be written as 
-\begin{equation}
-\sum_{D^k \in \Oh} \LRp{\bm{A}_0\pd{\bm{U}}{t},\bm{V} }_{\Lk} = \sum_{D^k \in \Oh}\LRp{ \LRp{\bm{A}_i(\bm{U}), \bm{V}_{,i}}_{\Lk} - \LRa{(\bm{A}_n\bm{U})^*,\bm{V}}_{\partial D^k}}
-\label{eq:formulation}
-\end{equation}
-where $(\bm{A}_n\bm{U})^*$ is a numerical flux defined on boundary faces and shared faces between elements.
-
-%We assume that $\bm{A}_{i}$ are symmetric matrices.
-
-\subsection{Penalization terms in numerical fluxes}
-
-Let $f$ be a shared face between two elements $D^{k,+}$ and $D^{k,-}$ with outward normal $\bm{n}^-$ on $D^{k,-}$, and let $\bm{U}^+, \bm{U}^-$ be evaluations of $\bm{U}$ restricted to $D^{k,+}$ and $D^{k,-}$, respectively.
-The average of both scalar and vector-valued functions is defined as
-\[
-\avg{u} = \frac{{u}^+ + u^-}{2},
-\]
-and the jump of scalar and vector-valued functions over $f$ are defined as
-\[
-\jump{u} = u^+ \bm{n}^+ + u^-\bm{n}^-, \qquad \jump{\bm{u}} = \bm{u}^+\cdot \bm{n}^+ + \bm{u}^-\cdot\bm{n}^-.  
-\]
-Typical numerical fluxes are defined as the sum of the average of $\bm{U}^+$ and $\bm{U}^-$ and a penalization term $\bm{G(\bm{U}^+,\bm{U}^-)}$
-\begin{equation}
-(\bm{A}_n\bm{U})^* = \avg{\bm{A}_n\bm{U}}\cdot \bm{n}^- - \bm{G(\bm{U}^+,\bm{U}^-)}.
-\label{eq:flux}
-\end{equation}
-\note{Clean up jump notation here.}
-The upwind numerical flux is a well-known flux of the form (\ref{eq:flux}).
-For an outward normal vector $\bm{n}^-$, let ${\bm{A}}_n = \sum_{i=1}^d {\bm{A}_i\bm{n}^-_i}$.
-By symmetry of $\bm{A}_i$, ${\bm{A}}_n$ contains real eigenvalues, and admits an eigenvalue decomposition
-\[
-\bm{A}_n = \bm{V}{\bm{\Lambda}}\bm{V}^T, \qquad \bm{\Lambda} = 
-\left(\begin{array}{ccc}
-\lambda_1 & & \\
-& \ddots & \\
-& & \lambda_d
-\end{array}\right).
-\]
-For problems with continuous coefficients, the upwind numerical flux over a face $f \in \Gh$ can be defined  as
-\[
-(\bm{A}_n\bm{U})^* = \bm{A}_n^+\bm{U}^- + \bm{A_n}^- \bm{U}^+,
-\]
-where the matrices $\bm{A_n}^+,\bm{A_n}^-$ are constructed from the positive and negative eigenvalues 
-\begin{align*}
-\bm{A}_n^+ &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} + \LRb{\bm{\Lambda}}} \bm{V}^T\\
-\bm{A}_n^- &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} - \LRb{\bm{\Lambda}}} \bm{V}^T,
-\end{align*}
-and $\LRb{\bm{\Lambda}}$ is the diagonal matrix whose entries consist of the absolute values of the eigenvalues $\LRb{\lambda_i}$.
-This can be rewritten as the sum of the central flux and a penalization
-\[
-(\bm{A}_n\bm{U})^* = \avg{\bm{A}_n\bm{U}} - \bm{V}\LRb{\bm{\Lambda}}\bm{V}^T \jump{\bm{U}}.  
-\]
-%For problems with discontinuous coefficients, this may be generalized by solving a Riemann problem to yield
-%\[
-%\bm{F^*}_n = \avg{\bm{A}_n\bm{U}} - \jump{{\bm{V}\LRb{\bm{\Lambda}}\bm{V}^T\bm{U}}}.  
-%\]
-We consider penalty fluxes as an alternative to upwind fluxes, which penalize appropriately defined jumps of the solution.  One such penalty flux is 
-\[
-(\bm{A}_n\bm{U})^* = \avg{\bm{A}_n\bm{U}} - \frac{\tau}{2} \bm{A}_n^T \jump{\bm{A}_n\bm{U}}.  
-\]
-where $\tau$ is proportional to $\lambda_{\max}^{-1}$.
-This choice of $\tau$ ensures that the penalty term is of the same order of magnitude as the upwind penalization, and that the CFL restriction for both upwind and penalty fluxes are identical.
-Unlike upwind fluxes, penalty fluxes can be used when a diagonalization of $\bm{A}_n$ is not explicitly known, such as linear elasticity in anisotropic media \cite{ye2016discontinuous}.
-
-\section{Dependence of spectra on the penalty parameter}
-
-The remainder of this work focuses on the influence of $\tau$ on the spectra of DG discretization matrices.
-
-\subsection{Conforming and non-conforming spaces}
-
-Let $V$ denote the piecewise polynomial approximation space 
-\[
-V = \LRc{ \bm{U} \in\LRp{ L^2(\Omega)}^d : \left.\bm{U}\right|_{D^k} \in \LRp{P^N(D^k)}^d}.  
-\]
-We define a decomposition of $V$ into a conforming part $V^C$ and a non-conforming part $V^{NC}$ based on the penalization term $\bm{G}(\bm{U}^+,\bm{U}^-)$.
-The conforming part is defined as the null space of $\bm{G}(\bm{U}^+,\bm{U}^-)$
-\begin{equation}
-V^{C} = \LRc{ \bm{U}(\bm{x}) \in V: \quad  \bm{G}(\bm{U}^+,\bm{U}^-) = 0, \quad \forall f \in \Gh},
-\label{eq:conf}
-\end{equation}
-while the non-conforming approximation space $V^{NC}$ is defined as $L^2$ orthogonal complement of $V^{C}$ in $V$
-\[
-V^{NC} = \LRc{ \bm{U}(\bm{x}) \in V: \LRp{\bm{u},\bm{v}}_{L^2(\Omega)} = 0, \quad \forall \bm{v}\in V^C}.
-\]
-When the coefficient matrices $\bm{A}_i$ are spatially continuous, the upwind flux condition $\bm{V}\LRb{\bm{\Lambda}}\bm{V}^T \jump{\bm{U}} = 0$ and the penalty flux condition $\bm{A}_n^T \jump{\bm{A}_n\bm{U}} = 0$ are both equivalent to
-\[
-\bm{v}_i^T \jump{\bm{U}} = 0, \qquad \lambda_i \neq 0,
-\]
-where $\lambda_i, \bm{v}_i$ are an eigenvalue and eigenvector of $\bm{A}_n$.  As a result, the conforming and non-conforming spaces induced by upwind and penalty fluxes are identical.
-
-For problems with discontinuous material data, it is less straightforward to show equivalence between the conforming spaces $V^C$ resulting from upwind and penalty fluxes.
-However, for many practical wave propagation settings, spatially varying coefficients are incorporated into the matrix $\bm{A_0}$, resulting in constant $\bm{A}_i$.
-For both smoothly varying and discontinuous $\bm{A}_0$, it is possible to formulate energy stable DG methods with penalty fluxes \cite{mercerat2015nodal, chan2016weight1}.
-Additionally, DG methods using penalty fluxes are observed to perform similarly to DG methods with upwind fluxes for  wave propagation problems with discontinuous coefficients \cite{warburton2013low, chan2016weight1, ye2016discontinuous}.
-
-\subsection{Scalar advection and acoustic wave propagation}
-\label{sec:confexamples}
-We illustrate the conforming spaces induced by penalty fluxes using the scalar advection equation and acoustic wave equation as concrete examples.  The scalar advection equation is given as
-\[
-\pd{u}{t} + \div\LRp{\bm{\beta}u} = 0.
-\]
-where $\bm{\beta}$ is the direction of advection.
-For continuous advection vectors such that $\bm{\beta}(\bm{x}) \neq 0$, the penalty flux is equivalent to a parametrized upwind flux \cite{hesthaven2007nodal}
-\[
-(\bm{A}_n\bm{U})^* = \avg{\bm{\beta}_n u} - \tau\LRb{\bm{\beta}_n}\jump{u}.
-\]
-The conforming space $V^C$ induced by this flux is then
-\[
-V^C = \LRc{ u \in L^2\LRp{\Omega} : \quad \left.u\right|_{D^k} \in P^N(D^k), \quad \LRb{\bm{\beta}_n}\jump{u} = 0}.
-\]
-In one dimension, this simply implies that $u$ is $C^0$-continuous across element boundaries.
-In higher dimensions, this implies that $u$ is continuous along streamlines or directions $\bm{d}(\bm{x})$ where $\bm{\beta}\cdot \bm{d} \neq 0$.
-
-Next, we consider the acoustic wave equation in pressure-velocity form
-\begin{align*}
-\frac{1}{c^2}\pd{p}{t} &= \Div \bm{u}\\
-\pd{\bm{u}}{t} &= \Grad p,
-\end{align*}
-where $c^2(\bm{x})$ is the wavespeed.
-Let $\bm{U}$ denote the group variable $\bm{U} = (p,u,v)$, where $u$ and $v$ are the $x$ and $y$ components of velocity.
-Then, in two dimensions, the isotropic wave equation is given as
-\[
-\pd{\bm{U}}{t} + \pd{\bm{A}_x\bm{U}}{x} + \pd{\bm{A}_y\bm{U}}{y} = 0, \qquad \bm{A}_x = 
-\left(\begin{array}{ccc}
-0 & 1 & 0\\
-1 & 0 & 0\\
-0 & 0 & 0
-\end{array}
-\right), \qquad 
-\bm{A}_y = 
-\left(\begin{array}{ccc}
-0 & 0 & 1\\
-0 & 0 & 0\\
-1 & 0 & 0
-\end{array}
-\right).
-\]
-The normal flux matrix $\bm{A}_n$ is then
-\[
-\bm{A}_n = 
-\left(\begin{array}{ccc}
-0 & \bm{n}^-_x & \bm{n}^-_y\\
-\bm{n}^-_x & 0 & 0\\
-\bm{n}^-_y & 0 & 0
-\end{array}
-\right)
-\]
-implying that the penalty fluxes are 
-\[
-\avg{\bm{A}_n \bm{U}} - \frac{\tau}{2}\bm{A}_n^T \jump{\bm{A}_n\bm{U}} = \left(\begin{array}{c}
-\avg{\bm{u}_n}\\
-\avg{p }\bm{n}^-_x\\
-\avg{p }\bm{n}^-_y
-\end{array}
-\right) - 
-\frac{\tau}{2}\left(\begin{array}{c}
-\jump{p}\\
-\jump{\bm{u_n}}\bm{n}^-_x\\
-\jump{\bm{u_n}}\bm{n}^-_y
-\end{array}
-\right)
-\]
-For $\tau = 1$, these fluxes coincide with the upwind fluxes for continuously varying media.
-In this case, the conforming subspace $V^{C}$ induced by the penalty flux consists of a scalar (pressure) component $V^C_p$ and a vector (velocity) component $V^C_{\bm{u}}$.
-The pressure component $p$ satisfies $\jump{p} = 0$ over all faces $f\in \Gh$.
-For polynomial approximation spaces, this implies that $p$ is continuous across faces, edges, and vertices, and $V^C_p$ is the standard $C^0$ continuous piecewise polynomial finite element space.
-
-For the velocity component $V_{\bm{u}}^C$, the penalty fluxes enforce normal continuity, such that $\jump{\bm{u}}\cdot \bm{n}^- = 0$ over all faces $f \in \Gh$.
-Since each component of $\bm{u}$ is approximated from $P^N(D^k)$, this implies that $V^C_{\bm{u}}$ is the $H({\rm div})$-conforming Brezzi-Douglas-Marini finite element space \cite{brezzi1985two,boffi2013mixed}.\footnote{It is possible to recover $V^C_{\bm{u}}$ as the Raviart-Thomas finite element space by approximating each component of $\bm{u}$ from a smaller polynomial space \cite{kirby2004algorithm}.}
-
-Penalty fluxes can also be interpreted as Lax-Friedrichs fluxes applied to scalar and vector variables separately.  In contrast, for a component-wise Lax-Friedrichs flux 
-\[
-(\bm{A}_n\bm{U})^* = \left(\begin{array}{c}
-\avg{\bm{u}_n}\\
-\avg{p }\bm{n}^-_x\\
-\avg{p }\bm{n}^-_y
-\end{array}
-\right) - 
-\frac{\tau}{2}\left(\begin{array}{c}
-\jump{p}\\
-\jump{\bm{u}_x}\\
-\jump{\bm{u}_y}
-\end{array}
-\right),
-\]
-the conforming velocity space $V^C_{\bm{u}}$ becomes the space of vector piecewise polynomial functions which are $C^0$ continuous in both $x$ and $y$ components, resulting in over-constrained continuity conditions for $\bm{u}$.
-
-\subsection{Behavior as $\tau\rightarrow\infty$}
-
-Let $\bm{K}$ be the matrix resulting from the discretization of the DG formulation (\ref{eq:formulation}).  We are interested in eigenvalues and eigenvectors of 
-\[
-\bm{K}\bm{u} = \lambda_i\bm{M}\bm{u}
-\]
-where $\bm{M}$ is the $L^2$ mass matrix.
-We adapt the approach of Warburton and Embree in \cite{Warburton20063205} to (\ref{eq:block}) to show that the spectra of $\bm{K}$ splits into two sets of eigenvalues as $\tau\rightarrow \infty$, the first of which approaches the eigenvalues of $\bm{A}$, and the second of which diverges with real part approaching $-\infty$ as $\tau\rightarrow \infty$.
-
-Let $\LRc{\bm{\Phi}_i }_{i=1}^{N^{C}}$ and $\LRc{\bm{\Psi}_i }_{i=1}^{N^{NC}}$ be $L^2$-orthogonal bases for $\bm{V}^C$ and $\bm{V}^{NC}$, respectively.
-These spaces then induce a block decomposition of the DG discretization matrix $\bm{K}$
-\begin{equation}
-\bm{K} = \left(\begin{array}{cc}
-\bm{A} & \bm{B}\\
-\tilde{\bm{B}}^T & \bm{C} + \tau \bm{S}
-\end{array}\right),
-\label{eq:block}
-\end{equation}
-where the blocks $\bm{A},\bm{B},\bm{C}$ are defined as
-\begin{align*}
-\bm{A}_{mn} &= \sum_{D^k\in \Oh} \LRp{\bm{A}_i\bm{\Phi}_n, \bm{\Phi}_{m,i}}_{D^k}, \qquad 1\leq m,n \leq N^C\\
-\bm{B}_{mn} &= \sum_{D^k\in \Oh} \LRp{\bm{A}_i\bm{\Psi}_n, \bm{\Phi}_{m,i}}_{D^k} - \LRa{\avg{\bm{A}_n\bm{\Psi}_i}, {\bm{\Phi}_i}}_{\partial D^k} \qquad 1\leq m \leq N^C, \quad 1\leq n \leq N^{NC} \\
-\bm{C}_{mn}  &= -\sum_{D^k\in \Oh} \LRa{\avg{\bm{A}_n\bm{\Psi}_i}, {\bm{\Psi}_i}}_{\partial D^k}, \qquad 1 \leq m,n \leq N^{NC}\\
-\bm{S}_{mn}  &= \frac{1}{2}\sum_{D^k\in \Oh} \LRa{\bm{A}_n^T\jump{\bm{A}_n \bm{\Psi_n}}, {\bm{\Psi}_m}}_{\partial D^k}, \qquad 1 \leq m,n \leq N^{NC}.  
-\end{align*}
-For exact quadrature, constant $\bm{A}_i$, and $\tau = 0$, the DG formulation is skew-symmetric \cite{warburton2013low, chan2015gpu, kopriva2014energy}, implying that $\bm{A},\bm{C}$ are skew-symmetric and that $\tilde{\bm{B}} = -\bm{B}$.
-Additionally, the entries of the penalization matrix $\bm{S}$ can be rewritten as
-\[
-\bm{S}_{mn}  &= \frac{1}{2}\sum_{D^k\in \Oh} \LRa{\jump{\bm{A}_n \bm{\Psi}_n}, \bm{A}_n{\bm{\Psi}_m}}_{\partial D^k} = -\frac{1}{2}\sum_{f \in \Gh} \LRa{\jump{\bm{A}_n \bm{\Psi}_n}, \jump{\bm{A}_n{\bm{\Psi}_m}}}_{f},
-\]
-implying that $\bm{S}$ is symmetric and negative-definite.
-
-%The generalized eigenvalue problem $\bm{K}\bm{u}=\lambda_i\bm{M}\bm{u}$ is then equivalent to the eigenvalue problem $\bm{K}\bm{u} = \lambda_i\bm{u}$.  
-%We begin by considering the spectra of $\bm{C} + \tau\bm{S}$.  
-%Let $\lambda, \bm{u}$ be an eigenpair of $\bm{C} + \tau{\bm{S}}$, such that $\lambda = \alpha + i\beta$ and $\bm{u} = \bm{v} + i\bm{w}$, where $\alpha,\beta$ and $\bm{v},\bm{w}$ are the real and imaginary parts of $\lambda$ and $\bm{u}$, respectively.
-Then, expanding and grouping terms in
-%\[
-%(\bm{C}+\tau\bm{S})(\bm{v} + i\bm{w}) = (\alpha + i\beta) (\bm{v} + i\bm{w})
-%\]
-%and using the fact that $\bm{C} + \tau\bm{S}$ is real-valued, we have that
-%\begin{align*}
-%(\bm{C}+\tau\bm{S})\bm{v} &= \alpha\bm{v}-\beta\bm{w}\\
-%(\bm{C}+\tau\bm{S})\bm{w} &= \beta\bm{v}+\alpha\bm{w}.
-%\end{align*}
-%%Assuming that $\nor{\bm{u}}^2 = \nor{\bm{v}}^2 + \nor{\bm{w}}^2 = 1$, multiplying both sides by $\bm{v}^T,\bm{w}^T$ and straightforward manipulations %yield
-%%\begin{align*}
-%%\bm{v}^T\bm{A}\bm{v} &= \alpha\nor{\bm{v}}^2-\beta\bm{v}^T\bm{w}\\
-%%\bm{w}^T\bm{A}\bm{w} &= \beta\bm{w}^T{\bm{v}}+\alpha\nor{\bm{w}}^2.  
-%%\end{align*}
-%%\begin{align*}
-%%\bm{v}^T\bm{A}\bm{v} + \bm{w}^T\bm{A}\bm{w} &= \alpha\\
-%%\bm{v}^T\bm{A}\bm{w} - \bm{w}^T\bm{A}\bm{w} &= \beta.
-%%\end{align*}
-%The symmetry of $\bm{S}$ then yields
-%\begin{align*}
-%\bm{v}^T\bm{C}\bm{w} - \bm{w}^T\bm{C}\bm{w} &= \beta\\
-%%\LRp{\bm{v}^T\bm{C}\bm{v} + \bm{w}^T\bm{C}\bm{w}} + 
-%\tau\LRp{\bm{v}^T\bm{S}\bm{v} + \bm{w}^T\bm{S}\bm{w}} &= \alpha.
-%\end{align*}
-%Since $\bm{v}^T\bm{C}\bm{w} - \bm{w}^T\bm{C}\bm{w}= \beta$ is independent of $\tau$ and $\bm{v},\bm{w}$ are normalized, $\beta$ must remain bounded as $\tau\rightarrow \infty$.
-Similarly, since $\bm{S}$ is symmetric, negative-definite, and independent of $\tau$, the quantity
-%\[
-%\LRb{\bm{v}^T\bm{S}\bm{v} + \bm{w}^T\bm{S}\bm{w}} = \LRb{\frac{\alpha}{\tau}}
-%\]
-%is bounded independently of $\tau$, and we conclude that $\LRb{\alpha} = O(\tau)$.
-Thus, as $\tau$ increases, the eigenvalues of $\bm{C} + \tau\bm{S}$ are shifted further and further left of the imaginary axis.
-
-We may now show how the spectra of $\bm{K}$ behaves as $\tau \rightarrow \infty$.
-Since $\bm{A}$ is skew-symmetric and $\bm{S}$ is symmetric, they are diagonalizable under unitary matrices $\bm{U}$ and $\bm{Q}$, whose columns contain the eigenvectors of $\bm{A}$ and $\bm{S}$, respectively.
-%\note{FIX $C + \tau S$ Gerschgorin part}
-Following \cite{Warburton20063205}, we apply a block diagonal similarity transform to yield
-\[
-\tilde{\bm{K}} = \left(\begin{array}{cc}
-\bm{U}^* & \\
-& \bm{Q}^*
-\end{array}\right)
-\left(\begin{array}{cc}
-\bm{A} & \bm{B}\\
--\bm{B}^T & \bm{C} + \tau \bm{S}
-\end{array}\right)
-\left(\begin{array}{cc}
-\bm{U} & \\
-& \bm{Q}
-\end{array}\right)
- = \left(\begin{array}{cc}
-\bm{\Lambda}^C & \bm{U}^*\bm{B}\bm{Q}\\
--\bm{Q}^*\bm{B}^T\bm{U} & \bm{Q}^*\bm{C}\bm{Q} + \tau \bm{\Lambda}^{S}
-\end{array}\right),
-\]
-where $\bm{\Lambda}^C,\bm{\Lambda}^{S}$ are diagonal matrices whose entries consist of eigenvalues of $\bm{A}$ and $\bm{S}$.
-
-Since $\bm{Q}^*\bm{C}\bm{Q}$ is skew-symmetric and its diagonal has zero real part, $\bm{Q}^*\bm{C}\bm{Q} + \tau \bm{\Lambda}^{S}$ has diagonal entries $\tau \lambda^S_j + \mi\gamma_j$, where $\lambda^S_j$ is real and negative and $\gamma_j$ is independent of $\tau$.
-Thus, the diagonal of $\tilde{\bm{K}}$ consists of $N^C$ purely imaginary values and $N^{NC}$ values of the form $\tau \lambda^S_j + \mi\gamma_j$.
-Since $\bm{B}$ is independent of $\tau$, the entries of $\bm{U}^*\bm{B}\bm{Q}$ are independent of $\tau$, assuming that $\bm{U}$ and $\bm{Q}$ are normalized.
-Gerschgorin's theorem applied to $\tilde{\bm{K}}$ then implies that the eigenvalues of $\bm{K}$ are contained in two sets of discs with radii independent of $\tau$.
-The first set of discs are centered around $\lambda^C_j$ for $j = 1,\ldots,N^C$, while the second set of discs are centered around $\tau\lambda^S_j + \mi\gamma_j$ for $j = 1,\ldots,N^{NC}$.
-For sufficiently large $\tau$, these disks must become disjoint, implying that the union of disks centered around  $\lambda^C_i$ contains exactly $N^C$ eigenvalues, while the union of disks centered around  $\lambda^{NC}_i$ contain exactly $N^{NC}$ eigenvalues \cite{horn2012matrix}.
-In other words, the spectra of $\bm{K}$ diverges into two distinct sets of $N^C$ and $N^{NC}$ eigenvalues each as $\tau \rightarrow \infty$, with the set of $N^{NC}$ eigenvalues containing negative real parts with magnitude $O(\tau)$.
-
-Consider now the $N^C$ eigenvalues with the smallest magnitude.
-By the Gerschgorin argument, these must remain bounded as $\tau \rightarrow \infty$.
-Let $\bm{W} = (\bm{W}^C,\bm{W}^{NC})$ and $\bm{\Lambda}^C$ be the matrix of eigenvectors and eigenvalues corresponding to these $N^C$ smallest eigenvalues, such that
-\[
-\left(\begin{array}{cc}
-\bm{A} & \bm{B}\\
--\bm{B}^T & \bm{C} + \tau \bm{S}
-\end{array}\right)
-\left(\begin{array}{c}
-\bm{W}^C
-\\
-\bm{W}^{NC}
-\end{array}\right) = 
-\left(\begin{array}{c}
-\bm{A}\bm{W}^C + \bm{B}\bm{W}^{NC}\\
--\bm{B}^T\bm{W}^C + \bm{C}\bm{W}^{NC} + \tau \bm{S}\bm{W}^{NC}
-\end{array}\right)
-= 
-\bm{\Lambda}_C
-\left(\begin{array}{c}
-\bm{W}^C
-\\
-\bm{W}^{NC}
-\end{array}\right) 
-\]
-This implies 
-\[
-\tau \nor{\bm{S}\bm{W}^{NC}} = \nor{\bm{\Lambda}^C \bm{W}^{NC} + \bm{B}^T\bm{W}^C  - \bm{C}\bm{W}^{NC}}.
-\]
-Since $\bm{\Lambda}^C$ remains bounded as $\tau\rightarrow \infty$, the right hand side is bounded independently of $\tau$ under normalization of $\bm{W}^C,\bm{W}^{NC}$, implying that the non-conforming component $\bm{W}^{NC}$ satisfies
-\[
-\sqrt{\lambda^{NC}_{\min}}\nor{\bm{W}^{NC}} \leq \nor{\bm{S}\bm{W}^{NC}} = O(1/\tau)
-\]
-where $\lambda^{NC}_{\min} > 0$ is the smallest eigenvalue of $\bm{S}$.
-As a consequence, $\bm{W}^{NC} \rightarrow 0$ as $\tau\rightarrow \infty$, and the smallest $N^C$ eigenvalues of $\bm{K}$ converge to the eigenvalues of $\bm{A}$ at a rate of $O(1/\tau)$.
-These correspond to a discretization using the conforming approximation space (\ref{eq:conf}).
-
-These results are summarized in the following lemma:
-\begin{lemma}
-As $\tau \rightarrow \infty$, the spectrum of the DG discretization matrix $\bm{K}$ decouples into two sets of eigenvalues $\LRc{\lambda^C_1,\ldots,\lambda^C_{N^C}}$ and $\LRc{\lambda^{NC}_1,\ldots,\lambda^{NC}_{N^{NC}}}$.
-The eigenvalues $\LRc{\lambda^{NC}_1,\ldots,\lambda^{NC}_{N^{NC}}}$ diverge towards the left half plane, with $\LRb{{\rm Re}(\lambda_i)} = O(\tau)$, while  the eigenvalues $\LRc{\lambda^C_1,\ldots,\lambda^C_{N^C}}$ converge to the eigenvalues of $\bm{A}$ at the rate $O(1/\tau)$.
-\label{lemma:eig}
-\end{lemma}
-We note that, while we consider skew-symmetric DG formulations in this work, Lemma~\ref{lemma:eig} is straightforward to generalize to any DG formulations where $\bm{S}$ is symmetric negative definite and $\bm{A}$ is diagonalizable.
-
-\section{Numerical experiments}
-
-The above analysis illustrates the behavior of the spectra of the DG discretization for the asymptotic cases when $\tau = 0$ or $\tau \rightarrow \infty$.
-However, it is less clear how the spectra of $\bm{K}$ behaves for $\tau \approx 1$.
-We rely instead on numerical experiments in one and two space dimensions to illustrate behaviors for the advection and acoustic wave equations.
-In particular, these experiments indicate that certain modes with negative real part for $\tau = O(1)$ return to the imaginary axis as $\tau \rightarrow \infty$.
-Since eigenvalues with a negative real part correspond to dissipation in time-domain simulations, this implies that these under-resolved modes become undamped as $\tau$ grows sufficiently large.
-
-\subsection{1D experiments}
-
-We consider first the scalar advection equation in 1D with periodic boundary conditions.
-For $\tau = 0$, the eigenvalues of the DG discretization matrix are purely imaginary.
-Paths taken by eigenvalues as $\tau$ increases are determined by sampling spectra over a sufficiently fine set of $\tau$ and using a particle tracking method \cite{simpletracker}.
-Figure~\ref{fig:track1D} shows the paths taken by these eigenvalues as $\tau$ increases from zero to $\tau = 4$.
-As predicted, a subset of \emph{divergent} eigenvalues move further left of the imaginary axis as $\tau$ increases.
-The corresponding eigenmodes are shown in Figure~\ref{fig:trackmodes1D}, with inter-element jumps of these eigenfunctions increasing as $\tau$ increases.
-
-\begin{figure}
-\centering
-\subfloat[Paths of eigenvalues as $\tau$ increases]{\includegraphics[width=.46\textwidth]{figs/trackedEigs.png}}
-\hspace{1em}
-\subfloat[Zoom of eigenvalues near imaginary axis]{\includegraphics[width=.44\textwidth]{figs/trackedEigsZoom.png}}
-\caption{Eigenvalue paths for DG advection with $\tau \in [0,4]$ on a mesh of 8 elements of degree $N=3$.
-Eigenvalues are overlaid on these paths for $\tau = 0$, $\tau = 1$, and $\tau = 4$.
-The zoomed in view near the imaginary axis shows the return of spurious modes to the imaginary axis for $\tau$ sufficiently large.
-}
-\label{fig:track1D}
-\end{figure}
-
-\begin{figure}
-\centering
-\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigsDiverge1.png}}
-\hspace{.5em}
-\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigsDiverge2.png}}
-\hspace{.5em}
-\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigsDiverge3.png}}\\
-\subfloat[$\tau = 0$]{\includegraphics[width=.32\textwidth]{figs/tauModesDiverge1.png}}
-\hspace{.5em}
-\subfloat[$\tau = 1$]{\includegraphics[width=.32\textwidth]{figs/tauModesDiverge2.png}}
-\hspace{.5em}
-\subfloat[$\tau = 10$]{\includegraphics[width=.32\textwidth]{figs/tauModesDiverge3.png}}
-\caption{Behavior of divergent eigenvalues for advection as $\tau$ increases.
-The order of approximation is $N=3$ on a mesh of 4 elements.} %The displayed eigenfunctions correspond to the highlighted eigenvalue.
-}
-\label{fig:trackmodes1D}
-\end{figure}
-
-However, for sufficiently large $\tau$, a subset of eigenvalues return to the imaginary axis.
-Figure~\ref{fig:spurious} illustrates that, as $\tau$ increases, the magnitude of the inter-element jumps present in these modes decreases, and the mode approaches a $C^0$ continuous function.
-These eigenvalues which move right to return to the imaginary axis correspond to a second set of \emph{spurious} modes.
-These modes are related to spurious modes observed for high order $C^0$ finite element methods \cite{ainsworth2014dispersive}, which consist of higher frequency components with sharp peaks.
-Since the real part of these eigenvalues approaches zero as $\tau\rightarrow \infty$, these modes become undamped and persist as spurious solution components in time-domain simulations \cite{hughes2014finite}
-
-\begin{figure}
-\centering
-\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigs1.png}}
-\hspace{.5em}
-\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigs2.png}}
-\hspace{.5em}
-\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigs3.png}}\\
-\subfloat[$\tau = 0$]{\includegraphics[width=.32\textwidth]{figs/tauModes1.png}}
-\hspace{.5em}
-\subfloat[$\tau = 1$]{\includegraphics[width=.32\textwidth]{figs/tauModes2.png}}
-\hspace{.5em}
-\subfloat[$\tau = 10$]{\includegraphics[width=.32\textwidth]{figs/tauModes3.png}}
-\caption{Return behavior of DG advection eigenvalues for sufficiently large $\tau$.
-For $\tau = 1$ certain eigenvalues have negative real part, and the corresponding eigenmodes are damped.
-For $\tau \gg 1$, certain eigenvalues return to the imaginary axis as spurious modes in conforming discretizations.
-}
-\label{fig:spurious}
-\end{figure}
-
-We repeat the above experiment for the acoustic wave equation in 1D with reflection boundary conditions on $[-1,1]$, and compare against the exact eigenvalues $\lambda_i = i \frac{\pi}{2}$ for $i = 1,2,\ldots$.
-The behavior of the eigenvalues as $\tau$ increases (Figure~\ref{fig:trackModesWave}) is similar to the behavior observed for the scalar advection equation.
-An examination of the eigenmodes corresponding to divergent and spurious eigenvalues as $\tau \rightarrow \infty$ shows that these modes behave similarly to the spurious modes observed for the scalar advection equation, with individual pressure and velocity components converging to continuous functions.
-
-%\begin{figure}
-%\centering
-%\subfloat{\includegraphics[width=.475\textwidth]{figs/tauEigs1.png}}
-%\hspace{.5em}
-%\subfloat{\includegraphics[width=.475\textwidth]{figs/tauEigs2.png}}
-%\caption{ }
-%\label{fig:trackModesWave}
-%\end{figure}
-
-\subsection{2D experiments}
-
-In two space dimensions, taking $\tau\rightarrow \infty$ again results in spurious modes which return to the imaginary axis, with the corresponding eigenmodes approaching conforming functions.
-We illustrate this through numerical experiments for the acoustic wave equation and advection in two space dimensions.
-In all numerical experiments, we use $N=3$ and a uniform triangular mesh resulting from bisecting each element of a $2\times 2$ uniform quadrilateral mesh.
-
-Figure~\ref{fig:waveeigs} shows the spectra of the DG discretization matrix for the acoustic wave equation for $\tau = 1, 10, 50$.
-Like the one-dimensional case, a subset of eigenvalues diverge towards the left half plane as $\tau$ increases, though unlike the one-dimensional case, these eigenvalues collapse towards the real line as $\tau$ increases.
-Similarly, as predicted, a subset of eigenvalues returns towards the imaginary axis as $\tau \rightarrow \infty$.
-These are shown in Figure~\ref{fig:trackEigsWave}, along with traced paths taken by each eigenvalue as $\tau$ increases.
-
-\begin{figure}
-\centering
-\subfloat[$\tau = 1$]{\includegraphics[width=.3\textwidth]{figs/waveEigs1.png}}
-\hspace{.5em}
-\subfloat[$\tau = 10$]{\includegraphics[width=.3\textwidth]{figs/waveEigs2.png}}
-\hspace{.5em}
-\subfloat[$\tau = 100$]{\includegraphics[width=.3\textwidth]{figs/waveEigs3.png}}
-\caption{Behavior of eigenvalues for the acoustic wave equation in two dimensions. Note the changing scale of the real axis.}
-\label{fig:waveeigs}
-\end{figure}
-
-\begin{figure}
-\centering
-\subfloat[$\tau = 0,1, 50$, zoomed in view]{\includegraphics[height=.3\textheight]{figs/waveEigs2Dbox.png}\label{subfig:big}}
-\hspace{2em}
-\subfloat[$\tau= 0,1,50$, returning eigenvalue paths]{\includegraphics[height=.3\textheight]{figs/trackedWaveEigs2D.png}\label{subfig:box}}
-\caption{Behavior of eigenvalues for the acoustic wave equation in two dimensions.
-Figure~\ref{subfig:box} shows a zoom of the boxed region in Figure~\ref{subfig:big}, with overlaid eigenvalue paths as $\tau$ increases.
-Divergent eigenvalues in the far left half plane are not shown.
-}
-\label{fig:trackEigsWave}
-\end{figure}
-
-The pressure and velocity components of the corresponding acoustic eigenmodes converge to conforming approximations.
-As noted in Section~\ref{sec:confexamples}, this implies that pressure components lie in $H^1(\Oh)$ and are continuous across faces, edges, and vertices.
-Figure~\ref{fig:trackModesWave} shows the behavior of the pressure component of an eigenmode corresponding to a spurious returning eigenvalue for $\tau = .1, 1, 100$.
-As $\tau$ increases and the eigenvalue approaches the imaginary axis, the eigenmode approaches a $C^0$ continuous function with a high frequencies and sharp peak, similar to the spurious eigenmodes observed for the one-dimensional case in Figure~\ref{fig:spurious}.
-Likewise, the non-zero velocity components of spurious eigenmodes converge to approximations in $H({\rm div}; \Oh)$, such that only the normal component of velocity is continuous across faces.
-Figure~\ref{fig:trackModesWaveU} shows visualizations of these eigenmodes.
-For large $\tau$, normal jumps of the velocity component vanish.
-% approach an $H({\rm div};\Oh)$-conforming approximation.
-
-\begin{figure}
-\centering
-\subfloat[$\tau = .1$, $\lambda = -0.2379 + 8.7528i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_p1.png}}
-\hspace{.5em}
-\subfloat[$\tau = 1$, $\lambda = -1.0145 + 8.9270i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_p2.png}}
-\hspace{.5em}
-\subfloat[$\tau = 100$, $\lambda = -0.0437 + 7.6167i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_p3.png}}
-\caption{Behavior of the real part of the pressure component of spurious eigenmodes for the acoustic wave equation in two dimensions.  }
-\label{fig:trackModesWave}
-\end{figure}
-
-\begin{figure}
-\centering
-\subfloat[$\tau = .1$, $\lambda = -0.6009 +11.3209i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_u1.png}}
-\hspace{.5em}
-\subfloat[$\tau = 1$, $\lambda =  -1.9601 + 9.5629i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_u2.png}}
-\hspace{.5em}
-\subfloat[$\tau = 100$, $\lambda = -0.0634 + 7.7670i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_u3.png}}
-\caption{Behavior of the real part of the velocity component of spurious eigenmodes for the acoustic wave equation in two dimensions.  }
-\label{fig:trackModesWaveU}
-\end{figure}
-
-For the advection equation with periodic boundary conditions, the spectra of the DG discretization behaves similarly.
-In all following experiments, we use advection vector $\beta = (1,0)$.
-For $\tau=0$, all eigenvalues lie on the imaginary axis, and as $\tau$ increases, the spectra splits into eigenvalues which diverge towards the left half plane and eigenvalues which return to the imaginary axis.
-Figure~\ref{fig:trackModesAdvec} shows the spectra of the DG matrix for $\tau = 0,1,50$, with eigenvalue paths for increasing values of $\tau$ overlaid.
-
-\begin{figure}
-\centering
-\subfloat{\includegraphics[width=.475\textwidth]{figs/trackedAdvecEigs.png}}
-\hspace{.5em}
-\subfloat{\includegraphics[width=.465\textwidth]{figs/trackedAdvecEigsZoom1.png}}
-\caption{Eigenvalue paths for the DG discretization of advection as $\tau$ increases.  Divergent eigenvalues in the far left half plane are not shown.  }
-\label{fig:trackModesAdvec}
-\end{figure}
-
-Unlike spurious conforming modes present for the acoustic wave equation, spurious conforming modes for the advection equation satisfy $\jump{\beta_n u} = 0$, and allow discontinuities along faces which are tangential to flow directions, as shown in Figure~\ref{fig:trackModesAdvecU}.
-%\note{Find cite on regularity of advection solutions.}
-
-\begin{figure}
-\centering
-\subfloat[$\tau = 0$, $\lambda = 0 + 7.9246i$]{\includegraphics[width=.3\textwidth]{figs/advecEig2D_u1.png}}
-\hspace{.5em}
-\subfloat[$\tau = 1$, $\lambda =  -0.8278 + 6.7831i$]{\includegraphics[width=.3\textwidth]{figs/advecEig2D_u2.png}}
-\hspace{.5em}
-\subfloat[$\tau = 100$, $\lambda = -0.0239 + 5.9282i$]{\includegraphics[width=.3\textwidth]{figs/advecEig2D_u3.png}}
-\caption{Behavior of the real part of spurious eigenmodes for the advection equation with $\beta = (1,0)$. }
-\label{fig:trackModesAdvecU}
-\end{figure}
-
-\section{Conclusions}
-
-For symmetric linear hyperbolic systems of PDEs, penalty fluxes are an alternative to upwind fluxes for DG methods.
-Both penalty and upwind fluxes can be interpreted as weakly enforcing continuity conditions implied by the underlying PDE.
-As the penalty parameter $\tau$ increases, we show that the eigenvalues of the DG discretization matrix split into two sets, with one set of eigenvalues converging to eigenvalues of a conforming discretization and the other set diverging with real part approaching $-\infty$.
-
-Numerical experiments show that, for $\tau = O(1)$, there exist eigenvalues with negative real part which return to the imaginary axis as $\tau \rightarrow 0$ or $\tau\rightarrow \infty$.
-The corresponding eigenmodes of such eigenvalues can be interpreted as spurious under-resolved modes which persist in time-domain simulations.
-These results suggest that the use of upwind and penalty fluxes damps spurious modes present not only for $\tau = 0$ (central fluxes), but also for $\tau \rightarrow\infty$ (a conforming discretization).
-
-\bibliographystyle{unsrt}
-\bibliography{dgpenalty}
-
+One set converges to the eigenvalues of a conforming discretization, while the other set corresponds to spurious eigenvalues which are damped proportionally to $\tau$.
+Numerical experiments also demonstrate that undamped spurious modes present in both in the limit of zero and large penalization parameters are damped for moderate values of the upwind parameter.
 \end{document}

### 2017-05-03 docs/elasticity_notes/letter.tex
updating docs for penalty note + elasticity

\documentclass{letter}
\usepackage{fullpage}
\signature{Jesse Chan}
\address{6100 Main St,\\
Houston, TX 77005}

\begin{document}
\begin{letter}{Dr.\ Hans De Sterck\\SIAM J. Sci. Comp. Section Editor\\Methods and Algorithms for Scientific Computing}

\opening{Dear Dr.\ Hans De Sterck}

Please find enclosed the manuscript, 
\begin{center}
Jesse Chan\\
\textit{Weight adjusted discontinuous Galerkin methods: matrix-valued weights and elastic wave propagation in heterogeneous media},
\end{center}
which we would like to submit for publication as an original research article in the SIAM Journal on Scientific Computing.  This work extends weight-adjusted inner products, which provide a low-storage approximation of the inverse of a mass matrix with respect to an $L^2$ inner product with a spatially varying weight, to the case of spatially varying matrix-valued weights.  These approximations are then applied to the simulation of elastic wave propagation in arbitrary heterogeneous media, including anisotropy and media with sub-element variations.  We also introduce a new discontinuous Galerkin method which simplifies numerical fluxes while retaining energy stability and high order accuracy.

We hope that the method and results discussed in this manuscript will appeal to the readership of SIAM Journal on Scientific Computing.  We confirm that this manuscript has not been published elsewhere and is not under consideration by another journal.  We look forward to hearing from you at your earliest convenience.

\closing{Best regards}

\end{letter}

\end{document}

### 2017-05-03 docs/splitform/main.tex
clearnup

@@
 \[
+\sum_{D^k} \LRp{\Grad_h \cdot \bm{u}, v}_{\LK} \coloneqq \sum_{D^k} \LRp{-\bm{u}, \Grad v}_{\LK} + \LRa{\avg{\bm{u}}\cdot\bm{n},v}_{\LdK}.
+\]
+and
+\[
 \Grad_h \cdot u = \Grad\cdot u + \frac{1}{2}L\LRp{\jump{\bm{u}}\cdot\bm{n}}
@@
 \]
-Need to incorporate boundary conditions.
+We will incorporate boundary conditions in a stable way in the following sections.
 
-We also have that, for interior elements,
+When the support of $v$ is limited to a single element, we have
 \[
@@
 
-\section{Penalization}
+Penalization can be added by adding any positive-definite stabilization term (upwind, penalty, Lax-Friedrichs) through the regular divergence flux.
 
-Penalization can be added by adding any positive-definite stabilization term (upwind, penalty, Lax-Friedrichs) through the regular divergence flux.
+It's probably better to formulate this using continuous DG derivatives, recover flux terms, then discretize that - the flux terms \textit{should} still cancel out after discretization, right?
+
+\section{Discrete DG derivatives}
+
+Methods based on discrete DG derivatives also work.
+
+The discrete DG derivative-based method is not consistent in the sense that Galerkin orthogonality does not hold exactly.  The difference lies in the flux terms.  Assume $\Div \bm{\beta} = 0$, then
+\[
+\LRp{\pd{u}{t},v} + \frac{1}{2}\LRp{-{ \bm{\beta}u},\Grad_h v} + \frac{1}{2}\LRp{\bm{\beta}\cdot\Grad_h u,v} = 0.
+\]
+\begin{align*}
+\LRp{\Grad_h\cdot \Pi_N\LRp{ \bm{\beta}u},v} &= \sum_{D^k} \LRp{-\bm{\beta}u, \Grad v}_{\LK} + \LRa{\avg{\Pi_N\LRp{\bm{\beta} u}}\cdot \bm{n},v}_{\LdK}\\
+\LRp{\Grad_h u,\bm{\beta}v} &= \sum_{D^k} \LRp{-u, \Grad \cdot \LRp{\bm{\beta}v}}_{\LK} + \LRa{\bm{\beta}\cdot\bm{n}\avg{u},v}_{\LdK}.
+\end{align*}
+The latter term is consistent; the former is not due to the presence of $\avg{\Pi_N\LRp{\bm{\beta} u}}\cdot \bm{n}$ in the flux term.
+The consistency error should then be $O(h^{N+1/2})$ using a trace inequality for $L^2$ projections.
+
+Can also write the inconsistent term as 
+\[
+\LRp{-{ \bm{\beta}u},\Grad_h v} %= -\LRp{\Grad_h v,\bm{\beta}u} = \sum_{D^k} \LRp{ v,\Grad\cdot \LRp{\bm{\beta}u}}_{D^k} - \LRa{\avg{v},\bm{\beta}\cdot \bm{n} u}
+\]
+but this results in the same inconsistent flux term due to the fact that $\Grad_h$ is only defined for test functions in the polynomial space.
+
+Note: can also use interpolants in a stable manner if using $D_h$.
+Unlike SEM, this still requires an extra matvec per RHS evaluation because of the lack of diagonality of the mass matrix.
+Reduces number of steps by one (no interpolation to quadrature points) but does not reduce number of total matvecs.
+
+For curvilinear coordinates,
 
@@
 
-Example: acoustic wave equation, simply discretize by replacing $\Grad, \Grad\cdot$ with the DG versions.  Automatically skew symmetric and energy stable.
+Example: acoustic wave equation, simply discretize by replacing $\Grad, \Grad\cdot$ with discrete versions.
+Automatically skew symmetric and energy stable via integration by parts.
+Also, can show why WADG works: discretize based on discrete divergence, then test with $T_{c^2}^{-1}p$ and use identities.
+Note - I think this requires the use of the strictly discrete version.
 
@@
 
-Example: Entropy splitting of Buckley-Leverett.
+Example: Kinetic energy preserving splitting of Euler (assumes exact time discretization).  Doesn't seem to help much without extra viscosity?
+
+Example: Entropy splitting of Buckley-Leverett?
+
+Example: Entropy splitting of Euler (note - cannot extend to Navier-Stokes in an entropy-stable fashion due to fact that heat flux matrix is not symmetrizable w.r.t.\ homogeneous flux function, though viscous terms are.
+This impacts only boundary conditions.)
+
+\section{Standard entropy stability estimates}
+
+Given a nonlinear conservation law
+\begin{align*}
+\pd{u}{t} + \pd{f(u)}{x} = 0
+\end{align*}
+DG formulation is usually
+\[
+\LRp{\pd{u}{t},v} + \LRp{D_h f,v} = 0.
+\]
+Taking $v = u$ gives
+\[
+\frac{1}{2}\nor{u}^2 + \int_{D^k}{u D_h f} = 0.
+\]
+Continuous entropy stability relies on the introduction of $F,G$ such that
+\[
+\pd{F}{u} = u\pd{f}{u}, \qquad \pd{G}{u} = f, \qquad F = uf - G.
+\]
+Then, relying on product and chain rules
+\[
+\int_{D^k}{u \pd{f}{x}} = \int_{D^k}{\pd{(uf)}{x} - f\pd{u}{x}} = \int_{D^k}{\pd{(uf)}{x} - \pd{G}{u}\pd{u}{x}} =  \int_{D^k}{\pd{\LRp{uf-G}}{x}} = \int_{D^k}{\pd{F}{x}} = \int_{\partial D^k} F.
+\]
+These boundary terms can be made to cancel with appropriately defined numerical fluxes, resulting in entropy conservative schemes.
+The challenge in reproducing this in the discrete case is the lack of a product and chain rule for inexact quadratures.
+Jameson deals with this by introducing an integrated flux
+\[
+G = \frac{1}{2}\int_{-1}^1 f \LRp{ \avg{u} + \theta \jump{u}} \diff \theta.
+\]
+for which finite volume schemes satisfy (in a rough sense)
+\[
+\jump{G} = \avg{G} \jump{u}.
+\]
+Fischer and Carpenter (also Gassner and co-workers) use a similar idea but combine a symmetric two-point flux approximation with properties of SBP matrices to get 
+\[
+2\LRp{W,DF} = \LRp{W,DF} - \LRp{DW,F} + \LRa{W,F} =  \sum_i \sum_j (W_i-W_j) (MD)_{ij} F(U_i,U_j) + \LRa{W,F}.
+\]
+
+\section{Entropy splitting}
+
+It's currently unclear how to extend the symmetric two-point flux approximation to quadrature-based DG methods.
+
+%Sandham and Yee use the entropy splitting 
+%\[
+%\eta(U) =  -\beta p^*, \qquad p^* = - \LRp{\frac{p}{\rho^\gamma}}^{\frac{1}{\beta(1-\gamma)}},
+%\]
+%which generates the entropy variables
+%\[
+%W(U) = p^*\LRp{\begin{array}{ccc}
+%\frac{E}{p} - C, &\frac{-\rho u}{p}, &\frac{\rho}{p}
+%\end{array}}, \qquad C = \frac{2}{\gamma-1} + (1+\beta).
+%\]
+%Unfortunately, the inverse mapping $U(W)$ is not explicitly known, which will greatly hamper computational efforts.
+For example, this implies we cannot evolve the entropy variables explicitly, which would yield an energy estimate.
+%The Jacobian
+%\[
+%\pd{U}{W} = \LRs{
+%d
+%}, 
+%\qquad \pd{W}{U} = \pd{U}{W}^{-1} = \LRs{
+%}
+%\]
+%The flux Jacobian
+%\[
+%\pd{F}{U} = \LRs{\begin{array}{ccc}
+%0 & u \frac{(\gamma-3)}{2} & \\
+%1 & -u(\gamma-3) & 
+%\end{array}}
+%\]
 

### 2017-05-03 matlab/Advec2D_cfl.m

@@ -1,198 +1,0 @@
-function Advec2D_cfl
-
-Globals2D
-global bx by Vq Pq Vfq Pfq Vrq Vsq Prq Psq
-
-
-% Polynomial order used for approximation
-N = 7;
-
-% Read in Mesh
-[Nv, VX, VY, K, EToV] = MeshReaderGambit2D('Grid/Other/squarereg.neu');
-% [Nv, VX, VY, K, EToV] = MeshReaderGambit2D('Grid/Other/squareireg.neu');
-% [Nv, VX, VY, K, EToV] = MeshReaderGambit2D('Grid/Other/block2.neu');
-
-% K1D = 2;
-% [Nv, VX, VY, K, EToV] = unif_tri_mesh(K1D);
-
-% Initialize solver and construct grid and metric
-StartUp2D;
-
-% rebuild maps for periodic
-BuildPeriodicMaps2D(2,2);
-
-[re se] = EquiNodes2D(N); [re se] = xytors(re,se);
-Ve = Vandermonde2D(N,re,se)/V;
-xe = Ve*x; ye = Ve*y;
-
-[rp sp] = EquiNodes2D(50); [rp sp] = xytors(rp,sp);
-Vp = Vandermonde2D(N,rp,sp)/V;
-xp = Vp*x; yp = Vp*y;
-
-[rq sq wq] = Cubature2D(3*N);
-% [rq sq wq] = QNodes2D(N); [rq sq] = xytors(rq,sq);
-Vq = Vandermonde2D(N,rq,sq)/V;
-[Vrq Vsq] = GradVandermonde2D(N,rq,sq);
-Vrq = Vrq/V; Vsq = Vsq/V;
-xq = Vq*x; yq = Vq*y;
-M = Vq'*diag(wq)*Vq;
-Pq = M\(Vq'*diag(wq));
-Prq = M\(Vrq'*diag(wq));
-Psq = M\(Vsq'*diag(wq));
-
-
-[r1D, w1D] = JacobiGQ(0,0,2*N);
-% rfq = [r1D; r1D; -ones(size(r1D))];
-% sfq = [-ones(size(r1D)); -r1D; r1D];
-wfq = [w1D; w1D; w1D];
-
-Vfq1 = Vandermonde1D(N,r1D)/Vandermonde1D(N,r(Fmask(:,1)));
[…]

### 2017-05-03 matlab/BuildNodeMaps2D.m

@@ -5,23 +5,24 @@ Nfaces = 3;
 NODETOL = 1e-5;
 
 % number volume nodes consecutively
-Nfp = size(xf,1) / Nfaces; 
-mapM    = (1:K*Nfp*Nfaces)';     
-mapP    = reshape(mapM, Nfp, Nfaces, K);
+Nfp = size(xf,1) / Nfaces;
+mapM    = reshape(1:K*Nfp*Nfaces, Nfp, Nfaces, K);
+mapP    = mapM;
 
 one = ones(1, Nfp);
+
 for k1=1:K
     for f1=1:Nfaces
         % find neighbor
-        k2 = EToE(k1,f1); f2 = EToF(k1,f1);
+        k2 = EToE(k1,f1); f2 = EToF(k1,f1);        
         
         if (k1==k2)
-            mapP(idM,f1,k1) = mapM(idM,f1,k1);
+            mapP(:,f1,k1) = mapM(:,f1,k1);
         else
-            
-            % find find volume node numbers of left and right nodes
             idM = (1:Nfp) + (f1-1)*Nfp;
             idP = (1:Nfp) + (f2-1)*Nfp;
+            
+            % find find volume node numbers of left and right nodes
             x1 = xf(idM,k1); y1 = yf(idM,k1);
             x2 = xf(idP,k2); y2 = yf(idP,k2);
             x1 = x1*one;  y1 = y1*one;

### 2017-05-03 matlab/acousticCurved.m

@@ -1,614 +1,0 @@
-function  Wave2D_curved
-
-% clear -global *
-% clear
-
-Globals2D;
-
-N = 3;
-nref = 0;
-useJprojection = 1;
-FinalTime = 5;
-
-Nq = 2*N+1;
-
-filename = 'Grid/Other/circA01.neu';
-[Nv, VX, VY, K, EToV, BCType] = MeshReaderGambitBC2D(filename);
-
-% K1D = 2;
-% [Nv, VX, VY, K, EToV] = unif_tri_mesh(K1D);
-
-% This builds the nodal DG stuff
-StartUp2D;
-
-BCType = zeros(size(EToV));
-for e = 1:K
-    BCType(e,EToE(e,:)==e) = 6;
-end
-
-BuildBCMaps2D;
-
-for ref = 1:nref
-    Refine2D(ones(size(EToV)));
-    StartUp2D;
-    BuildBCMaps2D;
-end
-
-% Store the domain boundary edges
-[k,f] = find(BCType);
-
-% Push all boundary faces to unit cylinder
-if (useJprojection >= 0)
-    MakeCylinder2D([k,f], 1, 0, 0);
-end
-
-% VB = bern_basis_tri(N,r,s);
-% [re se] = EquiNodes2D(N); [re se] = xytors(re,se);
-% Ve = Vandermonde2D(N,re,se)/V; 
-% xe = Ve*x; ye = Ve*y;
-% plot3(xe,ye,VB\x-x,'o')
-% keyboard
-
-cinfo = BuildCurvedOPS2D_opt(Nq,useJprojection);
-
-% PlotMesh2D
-
-% jesse custom curved driver
-global Vq wq Vrq Vsq
-global Vfq wfq VfqFace
-global rxq sxq ryq syq Jq Jfq nxq nyq sJq
-global Pq Pfq
-
-
-[rp sp] = EquiNodes2D(35); [rp sp] = xytors(rp,sp);
-Vp = Vandermonde2D(N,rp,sp)/V;
-xp = Vp*x; yp = Vp*y;
-
-[rq sq wq] = Cubature2D(Nq);
-Vq = Vandermonde2D(N,rq,sq)/V;
-[Vrq Vsq] = GradVandermonde2D(N,rq,sq);
-Vrq = Vrq/V; Vsq = Vsq/V;
-
-% interp nodal to quadrature
[…]

### 2017-05-03 matlab/check_conservation_loss.m

@@ -1,170 +1,0 @@
-clear
-Globals2D
-
-N = 3
-avec = [1e-1 1e-2 1e-3 1e-4];
-% avec = 1e-4; % smooth case
-for a = avec
-    sk = 1;
-    for K1D = [4 8 16 32]
-        if K1D==1
-            [VX VY] = Nodes2D(1); K = 1; EToV = 1:3;
-        else
-            [Nv, VX, VY, K, EToV] = unif_tri_mesh(K1D);
-        end
-        
-        StartUp2D;
-        
-        [rq sq w] = Cubature2D(40); % integrate u*v*c
-        Vq = Vandermonde2D(N,rq,sq)/V;
-        xq = Vq*x; yq = Vq*y;
-                
-        cfun = @(x,y) 1 + .5*sin(a*pi*x).*sin(a*pi*y); % rough c
-%         cfun = @(x,y) 1 + sqrt((x+.1).^2 + (y+.1).^2 + a);
-        cfun = @(x,y) exp(x+y);
-        
-        cq = cfun(xq,yq);
-        Mref = Vq'*diag(w)*Vq;
-        
-        u = @(x,y) exp(x+y);
-        u = @(x,y) 1 + sqrt(x.^2 + y.^2 + a) + x;
-        %         u = @(x,y) sin(pi*x).*sin(pi*y); % smooth u
-        %     a = 16; u = @(x,y) sin(a*pi*x).*sin(a*pi*y); % rough u
-%             u = @(x,y) (x > 2*y);%+ (x > -2*y);% + (y>0) + (x > 0);
-        %     keyboard
-        
-        uq = u(xq,yq);
-        b = (Vq'*(diag(w)*uq));
-        uproj = Mref\b;
-        
-        %     plot3(x,y,uproj,'o');keyboard
-        
-        local_cons_err = 0.0;
-        fixed_cons_err = 0.0;
-        proj_err = 0.0;
-        proj_err_weight = 0.0;
-        proj_err_cons = 0.0;
-        proj_diff = 0.0;
-        vmax = 0;
-        for e = 1:K
-            Minvcu = J(1,e)*Vq'*(w.*1./cq(:,e).*uq(:,e)); % true 1/c mass matrix
-            MinvMcMu = J(1,e)*Mref*((Vq'*diag(w.*cq(:,e))*Vq)\(Mref*uproj(:,e))); % weighted DG mass matrix
-            
[…]

### 2017-05-03 matlab/compute_circular_eigs.m

@@ -1,23 +1,0 @@
-clear lamEx
-load WADG_eigs
-x0 = sort(imag(lamWADG));
-x0 = x0(x0 > -1e-14);
-x0 = uniquetol(x0);
-% x0 = x0(1:50);
-% x0 = linspace(0,10,100);
-% x = linspace(0,10,200);
-% plot(x,besselj(n,x))
-% hold on
-for n = 0:5 % frequency    
-    for i = 1:length(x0)       
-        lamEx(i,1+n) = fzero(@(x)besselj(n,x),x0(i),optimset('TolX',1e-9));
-    end
-end
-tol = 1e-8;
-lamEx = lamEx(lamEx(:) > -tol);
-lamEx = uniquetol(lamEx(:),tol);
-% plot(0*lam(:),lam(:),'o')
-plot(lamEx,'o')
-hold on
-plot(sort(x0),'x')
-% lamDGh = lamDGi(lamDGi(:)>1e-4);

### 2017-05-03 matlab/massPrecon1D.m

@@ -1,82 +1,0 @@
-clear
-Globals1D;
-
-N = 5;
-
-K1D = 2;
-[Nv, VX, K, EToV] = MeshGen1D(-1,1,K1D);
-
-StartUp1D;
-
-R = zeros(N*K+1,(N+1)*K);
-gids = 1:N+1;
-for e = 1:K    
-    for i = 1:N+1
-        R(gids(i),i + (N+1)*(e-1)) = 1;
-    end
-    gids = gids + N;
-end
-
-re = linspace(-1,1,N+1)';
-[rq wq] = JacobiGQ(0,0,N);
-% Vq = bern_basis_1D(N,rq);
-% Ve = Vandermonde1D(N,re); Vq = Vandermonde1D(N,rq)/Ve;
-Vq = Vandermonde1D(N,rq)/V;
-Mref = Vq'*diag(wq)*Vq;
-
-MK = kron(spdiag(J(1,:)),Mref);
-PK = kron(spdiag(J(1,:)),inv(Mref));
-M = R*MK*R';
-P = R*PK*R';
-
-% cond(diag(1./sum(M,2))*M)
-
-M1ref = [2 1; 1 2]/3;
-Jx = diff(diag(1./sum(R,2))*R*x(:));
-
-Kx = length(Jx);
-ids = 1:2;
-M1 = zeros(size(M));
-for e = 1:Kx
-    M1(ids,ids) = M1(ids,ids) + M1ref * Jx(e);
-    ids = ids + 1;
-end
-clc
-cond(M1\M)
-cond(M)
-cond(diag(1./sum(M,2))*M) % lumped
-cond(diag(1./diag(M))*M)  % 
-
-rp = linspace(-1,1,100);
-Vp = Vandermonde1D(N,rp)/V;
-xp = Vp*x;
-
-invM = inv(M);
-
-for i = 1:Np*K
-    clf
-    plot(xp,Vp*reshape(R'*invM(:,i),Np,K))
-    pause
-end
-
-return
-% M1ref
-
-e = ones(N*K+1,1);
-h = 1/(N*K);
-M1 = 4*diag(e) + diag(e(2:end),1)+ diag(e(2:end),-1);
-M1(1,1) = M1(1,1)/2; M1(end,end) = M1(end,end)/2;
-M1 = h*M1/3;
-
-cond(M1\M)
-
-
-return
-
-% cond(M1\M)
-iM = R'*inv(M);
-
[…]

### 2017-05-03 matlab/plot_Wave2D_spectra.m

@@ -1,390 +1,0 @@
-function Wave2D_spectra
-
-Globals2D
-global bx by Vq Pq Vfq Pfq Vrq Vsq Prq Psq
-
-
-% Polynomial order used for approximation
-N = 3;
-
-% Read in Mesh
-% [Nv, VX, VY, K, EToV] = MeshReaderGambit2D('Grid/Other/squarereg.neu');
-% [Nv, VX, VY, K, EToV] = MeshReaderGambit2D('Grid/Other/squareireg.neu');
-% [Nv, VX, VY, K, EToV] = MeshReaderGambit2D('Grid/Other/block2.neu');
-
-K1D = 2;
-[Nv, VX, VY, K, EToV] = unif_tri_mesh(K1D);
-
-% Initialize solver and construct grid and metric
-StartUp2D;
-
-% % rebuild maps for periodic
-% BuildPeriodicMaps2D(2,2);
-
-Ne = 25;
-[re se] = EquiNodes2D(Ne); [re se] = xytors(re,se);
-re = re*(1-2/(Ne+1)); se = se*(1-2/(Ne+1));
-Ve = Vandermonde2D(N,re,se)/V;
-xe = Ve*x; ye = Ve*y;
-
-[rp sp] = EquiNodes2D(50); [rp sp] = xytors(rp,sp);
-Vp = Vandermonde2D(N,rp,sp)/V;
-xp = Vp*x; yp = Vp*y;
-
-[rq sq wq] = Cubature2D(3*N);
-% [rq sq wq] = QNodes2D(N); [rq sq] = xytors(rq,sq);
-Vq = Vandermonde2D(N,rq,sq)/V;
-[Vrq Vsq] = GradVandermonde2D(N,rq,sq);
-Vrq = Vrq/V; Vsq = Vsq/V;
-xq = Vq*x; yq = Vq*y;
-M = Vq'*diag(wq)*Vq;
-Pq = M\(Vq'*diag(wq));
-Prq = M\(Vrq'*diag(wq));
-Psq = M\(Vsq'*diag(wq));
-
-
-[r1D, w1D] = JacobiGQ(0,0,2*N);
-% rfq = [r1D; r1D; -ones(size(r1D))];
-% sfq = [-ones(size(r1D)); -r1D; r1D];
-wfq = [w1D; w1D; w1D];
-
[…]

### 2017-05-03 matlab/spline_quadrature.m

@@ -1,217 +1,0 @@
-% rules for Ksub = NB
-function [r w] = spline_quadrature(NB)
-
-
-switch NB
-    case 1
-        [r w] = JacobiGQ(0,0,NB);
-    case 2
-        r = [-0.831996808519005
-            -0.292665127127377
-            0.292665127127377
-            0.831996808519005];
-        
-        w = [0.408332371345181
-            0.591667628654818
-            0.591667628654819
-            0.408332371345180];
-    case 3
-        % NB = 3
-        r = [  -0.934981575335309
-            -0.698764027949677
-            -0.410170199830571
-            -0.141612248222709
-            0.141612248222709
-            0.410170199830571
-            0.698764027949677
-            0.934981575335309];
-        
-        w = [   0.161871006468693
-            0.286995338077702
-            0.274272539631319
-            0.276861115822286
-            0.276861115822286
-            0.274272539631319
-            0.286995338077702
-            0.161871006468692];
-    case 4
-        r = [-0.967919350846870
-            -0.844229387499850
-            -0.669592537179010
-            -0.491244208100323
-            -0.299443775577456
-            -0.096318521198056
-            0.096318521198056
-            0.299443775577456
-            0.491244208100323
-            0.669592537179010
-            0.844229387499850
-            0.967919350846870];
-        
-        w = [   0.080760998520667
-            0.158734694529584
-            0.180942006870548
-            0.179479067664020
-            0.203848193451594
-            0.196235038963586
-            0.196235038963586
-            0.203848193451594
-            0.179479067664020
-            0.180942006870548
-            0.158734694529584
-            0.080760998520667];
-    case 5
-        r = [-0.982020491016730
-            -0.910482312196710
-            -0.801441945385562
-            -0.681282034366054
-            -0.563508354274513
-            -0.434795577152953
-            -0.300911662692229
-            -0.180361816735623
-            -0.062222604467063
-            0.062222604467063
-            0.180361816735623
-            0.300911662692229
-            0.434795577152953
-            0.563508354274513
-            0.681282034366054
-            0.801441945385562
-            0.910482312196710
-            0.982020491016730];
-        w = [0.045531304561821
-            0.094378779801463
-            0.119016437564223
-            0.118341304234286
-            0.120964363010676
-            0.135072157545729
-            0.128582282017000
-            0.115244368800790
-            0.122869002464014
-            0.122869002464014
-            0.115244368800790
-            0.128582282017000
[…]

### 2017-05-03 matlab/test_lsc_hex.m

@@ -1,181 +1,0 @@
-function test_lsc_hex
-
-N = 3;
-Nq = N+1;
-[rq1D w1D] = JacobiGQ(0,0,Nq);
-[rq sq tq] = meshgrid(rq1D);
-rq = rq(:); sq = sq(:); tq = tq(:);
-[wr ws wt] = meshgrid(w1D); w = wr(:).*ws(:).*wt(:);
-
-for ref = 1:4
-    K1D = 2^(ref-1);
-    [VY VX VZ] = meshgrid(linspace(-1,1,K1D+1));
-    VX = VX(:); VY = VY(:); VZ = VZ(:);
-    
-    EToV = zeros(K1D^3,8);
-    sk = 1;
-    for k = 0:K1D-1
-        for j = 0:K1D-1
-            for i = 0:K1D-1
-                quad_ids = [1 2 K1D+3 K1D+2];
-                hex_ids = [quad_ids quad_ids+(K1D+1)^2];
-                EToV(sk,:) = hex_ids + i + j*(K1D+1) + k*(K1D+1)^2;
-                sk = sk + 1;
-            end
-        end
-    end
-    
-    % % arnold type mesh
-    h = (.5.^(ref-1));
-    nlev = 2^(ref-1);
-    t = linspace(-1,1,nlev+1);
-    for lev = 2:nlev
-        ids = find(abs(VZ-t(lev))<1e-8);
-        
-        [VYlev yids] = sort(VY(ids));
-        VXlev = reshape(VX(ids(yids)),nlev+1,nlev+1);
-        xids = zeros(nlev+1);
-        for j = 1:nlev+1
-            [~,xidj] = sort(VXlev(:,j));
-            xids(:,j) = xidj + (j-1)*(nlev+1);
-        end
-        xids = xids(:);
-        p = ids(yids(xids));
-        p = reshape(p,nlev+1,nlev+1);
-        toggle = [1 0 -1 0]; toggle = toggle(mod(lev-2,4)+1); % switch between add, don't move, subtract
-        
-        for line = 1:nlev+1
-            perturb = repmat([-1 1],1,floor((nlev+1)/2));
-            if mod(length(VZ(p)),2)==1
-                perturb = [perturb -perturb(end)];
-            end
-            if mod(line,2)==1
[…]

### 2017-05-03 matlab/test_spline_convergence.m

@@ -1,104 +1,0 @@
-clear
-Globals1D;
-
-smoothKnots = 20;
-
-NB = 3;
-Ksub = 1;
-K1D = 1;
-
-uex = @(x) exp(x).*sin(1+2*pi*x);
-k = 5;
-uex = @(x) sin(.1+k*pi*x); 
-% uex = @(x) exp(.5*x);
-
-for i = 2:8
-    
-    NB = i;
-    Ksub = 2*NB;
-    
-    N = NB+Ksub-1; % number of sub-mesh splines
-    
-    % Generate simple mesh
-    [Nv, VX, K, EToV] = MeshGen1D(-1,1,K1D);
-    
-    ndofs = K*(N+1); % total # dofs
-    
-    % Initialize solver and construct grid and metric
-    StartUp1D;
-    
-    rp = linspace(-1,1,50)';
-    Vp = Vandermonde1D(N,rp)/V;
-    xp = Vp*x;
-    
-    % make splines
-    rB = JacobiGL(0,0,N);
-    xB = Vandermonde1D(N,rB)/V * x;            
-    
-    % [rq wq] = JacobiGQ(0,0,N);
-    [rgq wgq] = JacobiGQ(0,0,2*N+1);    
-    if Ksub> 1
-        [~, ~, ~, ~, ~, ~, ~, ~, VX] = bsplineVDM(NB,Ksub,rB,smoothKnots); % VDM for interp, mass, M\S
-        h = @(r) repmat(diff(VX(:)'),length(r),1);
-        map = @(r) reshape(h(r).*(repmat(r,1,Ksub)+1)/2 + repmat(VX(1:end-1),length(r),1),length(r)*Ksub,1);
-        
-        rq = map(rgq);
-        wq = repmat(wgq,1,Ksub).*h(rgq)/2;
-        rq = rq(:); wq = wq(:);
-%         keyboard
-    else
-        rq = rgq(:); wq = wgq(:);
-    end
-            
-    
-    xq = Vandermonde1D(N,rq)/V * x;
-    
-    h = 2/(K1D);
-    hB = h/Ksub;
-    
-    % spline projection
-    [Bq] = bsplineVDM(NB,Ksub,rq,smoothKnots); % VDM for interp, mass, M\S
-    M = Bq'*diag(wq)*Bq;
-    PBq = M\(Bq'*diag(wq));
-    
-    % polynomial projection
[…]

### 2017-05-06 docs/dg_penalty/revision/main.tex
adding more notes to split form

@@
 \begin{frontmatter}
-\title{A short note on the penalty flux parameter for \note{discontinuous Galerkin formulations of first order hyperbolic systems}}
+\title{On the \note{penalty stabilization mechanism for upwind discontinuous Galerkin formulations of first order hyperbolic systems}}
 
@@
 {\note{Since $\tau$ can be chosen to be any arbitrary non-negative value, we consider three cases: $\tau = 0$, $\tau = 1$, and $\tau \rightarrow \infty$.
-The choice of $\tau = 1$ is somewhat arbitrary, and (as mentioned in Section~\ref{sec:scaletau}) it may be beneficial to scale $\tau$ with problem parameters.
+As mentioned in Section~\ref{sec:scaletau}, it may be beneficial to scale $\tau$ with problem parameters.
 However, for the simple model problems considered in this section, taking $\tau = 1$ recovers an upwind flux, and the resulting spectrum of the DG operator is clearly distinct from the cases when $\tau = 0$ or $\tau \rightarrow \infty$.}  }
@@
 \centering
-\subfloat[Paths of eigenvalues as $\tau$ increases]{\includegraphics[width=.48\textwidth]{figs/trackedEigs.png}}
+\subfloat[Paths of eigenvalues as $\tau$ increases]{\includegraphics[width=.48\textwidth]{trackedEigs.png}}
 \hspace{1em}
-\subfloat[Zoom of eigenvalues near imaginary axis]{\includegraphics[width=.435\textwidth]{figs/trackedEigsZoom.png}}
+\subfloat[Zoom of eigenvalues near imaginary axis]{\includegraphics[width=.435\textwidth]{trackedEigsZoom.png}}
 \caption{Eigenvalue paths for DG advection with $\tau \in [0,4]$ on a mesh of 8 elements of degree $N=3$.
@@
 \centering
-\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigsDiverge1.png}}
+\subfloat{\includegraphics[width=.32\textwidth]{tauEigsDiverge1.png}}
 \hspace{.5em}
-\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigsDiverge2.png}}
+\subfloat{\includegraphics[width=.32\textwidth]{tauEigsDiverge2.png}}
 \hspace{.5em}
-\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigsDiverge3.png}}\\
-\subfloat[$\tau = 0$]{\includegraphics[width=.32\textwidth]{figs/tauModesDiverge1.png}}
+\subfloat{\includegraphics[width=.32\textwidth]{tauEigsDiverge3.png}}\\
+\subfloat[$\tau = 0$]{\includegraphics[width=.32\textwidth]{tauModesDiverge1.png}}
 \hspace{.5em}
-\subfloat[$\tau = 1$]{\includegraphics[width=.32\textwidth]{figs/tauModesDiverge2.png}}
+\subfloat[$\tau = 1$]{\includegraphics[width=.32\textwidth]{tauModesDiverge2.png}}
 \hspace{.5em}
-\subfloat[$\tau = 10$]{\includegraphics[width=.32\textwidth]{figs/tauModesDiverge3.png}}
+\subfloat[$\tau = 10$]{\includegraphics[width=.32\textwidth]{tauModesDiverge3.png}}
 \caption{Behavior of a mode for the advection equation corresponding to a divergent eigenvalue (indicated by the red star) as $\tau$ increases.
@@
 These eigenvalues which move right to return to the imaginary axis correspond to a second set of \emph{spurious} modes.
-These modes closely resemble back-propagating spurious modes observed for high order $C^0$ finite element methods \cite{ainsworth2014dispersive}, which consist of higher frequency components with sharp peaks.
-Since the real part of these eigenvalues approaches zero as $\tau\rightarrow \infty$, these modes become undamped and persist as spurious solution components in time-domain simulations \cite{hughes2014finite}
+These modes closely resemble spurious modes observed for high order $C^0$ finite element methods \cite{boffi2000problem}, which consist of back-propagating high frequency components with sharp peaks \cite{ainsworth2014dispersive}.
+Since the real part of these eigenvalues approaches zero as $\tau\rightarrow \infty$, these modes become undamped and persist as spurious solution components in time-domain simulations \cite{hughes2014finite}.
+We note that that a similar phenomena has been observed for the solution of steady state problems in \cite{burman2010interior}, where the authors also point out that large penalty parameters result in decreased accuracy of the DG solution.
 
@@
 \centering
-\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigs1.png}}
+\subfloat{\includegraphics[width=.32\textwidth]{tauEigs1.png}}
 \hspace{.5em}
-\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigs2.png}}
+\subfloat{\includegraphics[width=.32\textwidth]{tauEigs2.png}}
 \hspace{.5em}
-\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigs3.png}}\\
-\subfloat[$\tau = 0$]{\includegraphics[width=.32\textwidth]{figs/tauModes1.png}}
+\subfloat{\includegraphics[width=.32\textwidth]{tauEigs3.png}}\\
+\subfloat[$\tau = 0$]{\includegraphics[width=.32\textwidth]{tauModes1.png}}
 \hspace{.5em}
-\subfloat[$\tau = 1$]{\includegraphics[width=.32\textwidth]{figs/tauModes2.png}}
+\subfloat[$\tau = 1$]{\includegraphics[width=.32\textwidth]{tauModes2.png}}
 \hspace{.5em}
-\subfloat[$\tau = 10$]{\includegraphics[width=.32\textwidth]{figs/tauModes3.png}}
+\subfloat[$\tau = 10$]{\includegraphics[width=.32\textwidth]{tauModes3.png}}
 \caption{Return behavior of DG advection eigenvalues for sufficiently large $\tau$.
@@
 \centering
-\subfloat[$c_j^0$ (spurious modes for  $\tau = 0$)]{\includegraphics[width=.45\textwidth]{figs/ceigs0.png}}
+\subfloat[$c_j^0$ (spurious modes for  $\tau = 0$)]{\includegraphics[width=.45\textwidth]{ceigs0.png}}
 \hspace{1em}
-\subfloat[$c_j^{100}$ (spurious modes for $\tau = 100$)]{\includegraphics[width=.45\textwidth]{figs/ceigs100.png}}
+\subfloat[$c_j^{100}$ (spurious modes for $\tau = 100$)]{\includegraphics[width=.45\textwidth]{ceigs100.png}}
 \caption{Coefficients $c^0_j$ and $c^{100}_j$ of the spurious mode in Figure~\ref{fig:spurious} when expanded in the modes of $\bm{K}$ for $\tau=1$.
@@
 %\centering
-%\subfloat{\includegraphics[width=.475\textwidth]{figs/tauEigs1.png}}
+%\subfloat{\includegraphics[width=.475\textwidth]{tauEigs1.png}}
 %\hspace{.5em}
-%\subfloat{\includegraphics[width=.475\textwidth]{figs/tauEigs2.png}}
+%\subfloat{\includegraphics[width=.475\textwidth]{tauEigs2.png}}
 %\caption{ }
@@
 \centering
-\subfloat[$\tau = 1$]{\includegraphics[width=.3\textwidth]{figs/waveEigs1.png}}
+\subfloat[$\tau = 1$]{\includegraphics[width=.3\textwidth]{waveEigs1.png}}
 \hspace{.5em}
-\subfloat[$\tau = 10$]{\includegraphics[width=.3\textwidth]{figs/waveEigs2.png}}
+\subfloat[$\tau = 10$]{\includegraphics[width=.3\textwidth]{waveEigs2.png}}
 \hspace{.5em}
-\subfloat[$\tau = 100$]{\includegraphics[width=.3\textwidth]{figs/waveEigs3.png}}
+\subfloat[$\tau = 100$]{\includegraphics[width=.3\textwidth]{waveEigs3.png}}
 \caption{Behavior of eigenvalues for the acoustic wave equation in two dimensions. Note the changing scale of the real axis.}
@@
 \centering
-\subfloat[$\tau = 0,1, 50$, zoomed in view]{\includegraphics[height=.3\textheight]{figs/waveEigs2Dbox.png}\label{subfig:big}}
+\subfloat[$\tau = 0,1, 50$, zoomed in view]{\includegraphics[height=.3\textheight]{waveEigs2Dbox.png}\label{subfig:big}}
 \hspace{2em}
-\subfloat[$\tau= 0,1,50$, returning eigenvalue paths]{\includegraphics[height=.3\textheight]{figs/trackedWaveEigs2D.png}\label{subfig:box}}
+\subfloat[$\tau= 0,1,50$, returning eigenvalue paths]{\includegraphics[height=.3\textheight]{trackedWaveEigs2D.png}\label{subfig:box}}
 \caption{Behavior of eigenvalues for the acoustic wave equation in two dimensions.
@@
 \centering
-\subfloat[$\tau = .1$, $\lambda = -0.2379 + 8.7528i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_p1.png}}
+\subfloat[$\tau = .1$, $\lambda = -0.2379 + 8.7528i$]{\includegraphics[width=.3\textwidth]{waveEig2D_p1.png}}
 \hspace{.5em}
-\subfloat[$\tau = 1$, $\lambda = -1.0145 + 8.9270i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_p2.png}}
+\subfloat[$\tau = 1$, $\lambda = -1.0145 + 8.9270i$]{\includegraphics[width=.3\textwidth]{waveEig2D_p2.png}}
 \hspace{.5em}
-\subfloat[$\tau = 100$, $\lambda = -0.0437 + 7.6167i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_p3.png}}
+\subfloat[$\tau = 100$, $\lambda = -0.0437 + 7.6167i$]{\includegraphics[width=.3\textwidth]{waveEig2D_p3.png}}
 \caption{Behavior of the real part of the pressure component of spurious eigenmodes for the acoustic wave equation in two dimensions.  }
@@
 %\centering
-%\subfloat[$\tau = .1$, $\lambda = -0.6009 +11.3209i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_u1.png}}
+%\subfloat[$\tau = .1$, $\lambda = -0.6009 +11.3209i$]{\includegraphics[width=.3\textwidth]{waveEig2D_u1.png}}
 %\hspace{.5em}
-%\subfloat[$\tau = 1$, $\lambda =  -1.9601 + 9.5629i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_u2.png}}
+%\subfloat[$\tau = 1$, $\lambda =  -1.9601 + 9.5629i$]{\includegraphics[width=.3\textwidth]{waveEig2D_u2.png}}
 %\hspace{.5em}
-%\subfloat[$\tau = 100$, $\lambda = -0.0634 + 7.7670i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_u3.png}}
+%\subfloat[$\tau = 100$, $\lambda = -0.0634 + 7.7670i$]{\includegraphics[width=.3\textwidth]{waveEig2D_u3.png}}
 %\caption{Behavior of the real part of the velocity component of spurious eigenmodes for the acoustic wave equation in two dimensions.  }
@@
 \centering
-\subfloat{\includegraphics[width=.475\textwidth]{figs/trackedAdvecEigs.png}}
+\subfloat{\includegraphics[width=.475\textwidth]{trackedAdvecEigs.png}}
 \hspace{.5em}
-\subfloat{\includegraphics[width=.465\textwidth]{figs/trackedAdvecEigsZoom1.png}}
+\subfloat{\includegraphics[width=.465\textwidth]{trackedAdvecEigsZoom1.png}}
 \caption{Eigenvalue paths for the DG discretization of advection as $\tau$ increases.  Divergent eigenvalues in the far left half plane are not shown.  }
@@
 \centering
-\subfloat[$\tau = 0$, $\lambda = 0 + 7.9246i$]{\includegraphics[width=.3\textwidth]{figs/advecEig2D_u1.png}}
+\subfloat[$\tau = 0$, $\lambda = 0 + 7.9246i$]{\includegraphics[width=.3\textwidth]{advecEig2D_u1.png}}
 \hspace{.5em}
-\subfloat[$\tau = 1$, $\lambda =  -0.8278 + 6.7831i$]{\includegraphics[width=.3\textwidth]{figs/advecEig2D_u2.png}}
+\subfloat[$\tau = 1$, $\lambda =  -0.8278 + 6.7831i$]{\includegraphics[width=.3\textwidth]{advecEig2D_u2.png}}
 \hspace{.5em}
-\subfloat[$\tau = 100$, $\lambda = -0.0239 + 5.9282i$]{\includegraphics[width=.3\textwidth]{figs/advecEig2D_u3.png}}
+\subfloat[$\tau = 100$, $\lambda = -0.0239 + 5.9282i$]{\includegraphics[width=.3\textwidth]{advecEig2D_u3.png}}
 \caption{Behavior of the real part of spurious eigenmodes for the advection equation with $\beta = (1,0)$. }

### 2017-05-06 docs/splitform/main.tex
adding more notes to split form

@@
 \usepackage{fullpage}
-\usepackage{amsmath,amssymb,amsfonts}
+\usepackage{amsmath,amssymb,amsfonts,amsthm}
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}[section]
+\newtheorem*{remark}{Remark}
+%\usepackage{thmtools}
+%\declaretheorem[style=definition,qed=$\blacksquare$,numberwithin=chapter]{definition}
+
 \usepackage[titletoc,toc,title]{appendix}
@@
 \newcommand{\note}[1]{{\color{blue}{#1}}}
-\newcommand{\remark}[1]{\textbf{\color{red}#1}}
 
@@
 \author{Jesse Chan}
-\title{Split form DG methods}
+\title{Energy stable discontinuous Galerkin formulations and discrete differential operators}
 
@@
 \maketitle
-%\tableofcontents
+
+\begin{abstract}
+We show that, for non-linear hyperbolic conservation laws which admit a skew-symmetric splitting, energy stable discontinuous Galerkin (DG) methods can be constructed in a straightforward manner based on discrete DG gradient and divergence operators.
+In particular, we show that these formulations remain energy stable for both curvilinear geometries and inexact quadrature.
+Examples of energy stable formulations are given for variable advection, Burgers' equation, and a Burgers'-like system, and it is shown that the construction of DG methods based on discrete differential operators recover known entropy-conservative fluxes.
+\end{abstract}
+
+\section{Introduction}
+
+\cite{feng2016discontinuous}
+
+Lifting operators  and discrete gradients \cite{bassi1997high, di2011mathematical}
+
+\section{Discrete differential operators}
+
+We assume that the domain $\Omega$ is decomposed into non-overlapping elements $D^k$.
+We define the mesh $\Omega_h = \cup D^k$ and the corresponding global approximation space $V_h(\Omega_h) = \bigoplus V_h\LRp{D^k}$, where $V_h\LRp{D^k}$ is the approximation space over a single patch.
+Furthermore, we introduce the jump and average of discontinuous functions across element interfaces.
+Let $D^{k,+}$ denote the neighboring element of across a face $f$ of $D^k$, and let $u^+,u^-$ denote the values of $u$ on $D^{k,+}$ and $D^k$, respectively.
+The jump of $u$ across $f$ is then defined as
+\[
+\jump{u} = u^+ - u^-, \qquad \avg{u} = \frac{u^+ + u^-}{2}.
+\]
+The jump and average of vector fields are defined component-wise using the jumps and averages of components.
+
+\subsection{Discrete inner products}
+
+In order to define discrete differential operators, we first introduce the $L^2$ inner product on $V_h\LRp{\Oh}$
+\[
+\LRp{u,v}_{\Omega} = \sum_{D^k} \int_{D^k} uv \diff{x} = \sum_{D^k} \int_{\widehat{D}} uv J \diff{\widehat{x}}.  
+\]
+In practice, these integrals are computed using quadrature, such that 
+\[
+\int_{\widehat{D}}uv = \sum_{i=1}^{N_q} u(\bm{x}_i)v(\bm{x}_i) w_i,
+\]
+where $N_q$ is the number of quadrature points.  The only assumptions we make upon this quadrature is that it is sufficiently accurate such that 
+\begin{enumerate}
+\item The quadrature induces an equivalent inner product over the reference element $\widehat{D}$
+\item Integration by parts holds (with respect to the reference coordinates $\widehat{x}$).
+\end{enumerate}
+
+We also introduce the $L^2$ projection $\Pi_N: L^2\LRp{\Omega}\rightarrow P^N$ such that 
+\[
+\LRp{\Pi_N u,v}_{\Omega} = \LRp{u,v}_{\Omega}.
+\]
+
+\subsection{Discrete derivatives}
+We introduce two discrete DG derivatives in this section.  The first is the implicit definition used in \cite{di2011mathematical}
+
+\begin{definition}
+The discontinuous Galerkin differentiation operator ${D}^i: L^2\LRp{\Omega}\rightarrow V_h$ is defined implicitly as follows:
+\[
+\LRp{{D}^i u,v}_{\Oh} = \sum_{D^k} \LRp{\LRp{-u,\pd{v}{\bm{x}_i}}_{\LK} + \LRa{\avg{u},v\bm{n}_i}_{\LdK}}.
+\]
+\end{definition}
+
+We note that, for $u,v \in V_h$, integration by parts yields an equivalent definition of $D^i$ found in \cite{hesthaven2004high,Warburton20063205}
+\[
+\LRp{{D}^i u,v}_{\Oh} = \sum_{D^k} \LRp{\LRp{\pd{u}{\bm{x}_i},v}_{\LK} + \frac{1}{2}\LRa{\jump{u},v\bm{n}_i}_{\LdK}}.
+\]
+However, for $u \not\in V_h$, these two definitions are not equivalent.  This motivates the definition of a \textit{discrete} DG differentiation operator:
+\begin{definition}
+The discrete discontinuous Galerkin differentiation operator ${D}^i_h: L^2\LRp{\Omega} \rightarrow V_h$ is defined as $D^i_h = D^i \Pi_N$.
+\end{definition}
+
+The main property of
 
 \section{Notes}
-
-Integration by parts gives
+Begin with the gradient.  Multiplication by $\bm{v}$ and integration by parts gives
 \[
@@
 \]
-and it can be shown that, for a periodic mesh,
+and it can be shown that 
 \[
@@
 \bibliographystyle{unsrt}
-\bibliography{dgpenalty}
+\bibliography{dg}
 

### 2017-05-09 docs/elasticity_notes/main.tex
updating docs
updates to elasticity notes (sisc rejected reviewer comments)

@@
 %\newcommand{\cond}[1]{\kappa\LRp{#1}}
-\newcommand{\cond}[2]{\nor{#1}_{#2}\nor{{#1}^{-1}}_{#2}}
+%\newcommand{\cond}[2]{\nor{#1}_{#2}\nor{{#1}^{-1}}_{#2}}
 
@@
 A new energy stable DG formulation is introduced based on the symmetric form of the elastic wave equations, with upwind-like numerical dissipation introduced through simple penalty fluxes \cite{warburton2013low}.
-In contrast to the fluxes proposed in \cite{ye2016discontinuous}, the penalty fluxes used here are independent of material coefficients.
+In contrast to the fluxes proposed in \cite{ye2016discontinuous}, the penalty fluxes used here can be made to be independent of material coefficients.
 
@@
 We begin with the linear elastic wave equation in a domain $\Omega \in \mathbb{R}^d$.
-These equations can be written a first order velocity-stress system for velocity $\bm{v}$ and symmetric stress tensor $\bm{S}$
+These equations can be written as a first order velocity-stress system for velocity $\bm{v}$ and symmetric stress tensor $\bm{S}$
 \begin{align*}
@@
 \end{align}
-where $\bm{C}$ is the symmetric matrix form of the constitutive tensor $\tensor{\bm{C}}$ and $\sigma$ is a vector of length $N_d = \frac{d(d+1)}{2}$, the number of unique entries of the stress tensor $\bm{S}$ in $d$ dimensions.
+where $\bm{C}$ is the symmetric matrix form of the constitutive tensor $\tensor{\bm{C}}$ and $\bm{\sigma}$ is a vector of length $N_d = \frac{d(d+1)}{2}$, the number of unique entries of the stress tensor $\bm{S}$ in $d$ dimensions.
 We note that the matrices $\bm{A}_i$ are spatially constant, while $\rho,\bm{C}$, and $\bm{C}^{-1}$ can vary spatially.
@@
 \]
-where $\bm{x}^k = \LRc{x^k,y^k,z^k}$ denote physical coordinates on $D^k$ and $\widehat{\bm{x}} = \LRc{\widehat{x},\widehat{y},\widehat{z}}$ denote coordinates on the reference element.
+where $\bm{x}^k = \LRc{x^k,y^k}$ for $d=2$ and $\bm{x}^k = \LRc{x^k,y^k,z^k}$ for $d=3$ denote physical coordinates on $D^k$ and $\widehat{\bm{x}} = \LRc{\widehat{x},\widehat{y}}$  for $d = 2$ and $\widehat{\bm{x}} = \LRc{\widehat{x},\widehat{y},\widehat{z}}$ for $d = 3$ denote coordinates on the reference element.
 We denote the determinant of the Jacobian of $\bm{\Phi}^k$ as $J$, and refer to it as the Jacobian for the remainder of this work.
@@
 \[
-\LRp{\bm{f},\bm{g}}_{\Lk} = \int_{D^k} \bm{f}\cdot\bm{g} \diff x = \int_{\widehat{D}} \bm{f}\cdot\bm{g} J, \qquad \nor{\bm{f}}_{\Lk}^2 = \LRp{\bm{f},\bm{f}}_{\Lk},
+\LRp{\bm{f},\bm{g}}_{\Lk} = \int_{D^k} \bm{f}\cdot\bm{g} \diff{\bm{x}} = \int_{\widehat{D}} \bm{f}\cdot\bm{g} J\diff{\widehat{\bm{x}}}, \qquad \nor{\bm{f}}_{\Lk}^2 = \LRp{\bm{f},\bm{f}}_{\Lk},
 \]
@@
 \[
-\LRp{\bm{f},\bm{g}}_{\L} = \sum_{D^k\in \Oh} \int_{D^k} \LRp{\bm{f},\bm{g}}_{\Lk}, \qquad \nor{\bm{f}}_{\L}^2 = \sum_{D^k\in \Oh} \nor{\bm{f}}^2_{\Lk}.
+\LRp{\bm{f},\bm{g}}_{\L} = \sum_{D^k\in \Oh} \LRp{\bm{f},\bm{g}}_{\Lk}, \qquad \nor{\bm{f}}_{\L}^2 = \sum_{D^k\in \Oh} \nor{\bm{f}}^2_{\Lk}.
 \]
@@
 \[
-\LRa{u,v}_{\Ldk} = \int_{\partial D^k}uv\diff{x} = \sum_{f\in \partial D^k}\int_{\widehat{f}}uvJ^f\diff{\widehat{x}}, \qquad \nor{u}_{\Ldk}^2 = \LRa{u,u}_{\Ldk},
+\LRa{\bm{u},\bm{v}}_{\Ldk} = \int_{\partial D^k}\bm{u}\cdot\bm{v}\diff{\bm{x}} = \sum_{f\in \partial D^k}\int_{\widehat{f}}\bm{u}\cdot\bm{v}J^f\diff{\widehat{\bm{x}}}, \qquad \nor{\bm{u}}_{\Ldk}^2 = \LRa{\bm{u},\bm{u}}_{\Ldk},
 \]
@@
 \[
-\LRp{\jump{\bm{u}}}_i = \jump{\bm{u}_i}, \qquad 1\leq i \leq m, \qquad \LRp{\jump{\bm{S}}}_{ij} = \jump{\bm{S}_i}, \qquad 1\leq i \leq m, \quad 1\leq j \leq n.
+\LRp{\jump{\bm{u}}}_i = \jump{\bm{u}_i}, \qquad 1\leq i \leq m, \qquad \LRp{\jump{\bm{S}}}_{ij} = \jump{\bm{S}_{ij}}, \qquad 1\leq i \leq m, \quad 1\leq j \leq n.
 \]
@@
 \begin{align*}
-\LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= -\LRp{ \sum_{i=1}^d \bm{\sigma},\bm{A}_i \pd{\bm{w}}{\bm{x}_i}}_{\Lk}  + \LRa{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk} \nonumber\\
-\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk}
+\sum_{D^k \in \Oh}\LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= \sum_{D^k \in \Oh}-\LRp{ \sum_{i=1}^d \bm{\sigma},\bm{A}_i \pd{\bm{w}}{\bm{x}_i}}_{\Lk}  + \LRa{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk} \nonumber\\
+\sum_{D^k \in \Oh}\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \sum_{D^k \in \Oh}\LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk}
 \end{align*}
@@
 &=\sum_{D^k\in \Omega_h}\LRa{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{v}}_{\partial D^k}+ \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{\sigma}}_{\partial D^k}  \\
-&=\sum_{D^k\in \Omega_h}\sum_{f\in \partial D^k}\int_{f}{\bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v}} + {\frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma}}.
+&=\sum_{D^k\in \Omega_h}\sum_{f\in \partial D^k}\int_{f}\LRp{\bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v} + \frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma}}\diff{\bm{x}}.
 \end{align*}
@@
 \begin{align*}
-&\sum_{f\in \Gamma_h \setminus \partial \Omega} \int_{f}{\bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v}} + {\frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma}}\\
-&= - \sum_{f \in\Gamma_h \setminus \partial \Omega} \int_{f} \frac{\tau_v}{2}\LRb{\bm{A}_n\jump{\bm{v}}}^2 + \frac{\tau_\sigma}{2}\LRb{\bm{A}_n^T\jump{\bm{\sigma}}}^2.  
+&\sum_{f\in \Gamma_h \setminus \partial \Omega} \int_{f}\LRp{\bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v} + \frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma}}\diff{\bm{x}}\\
+&= - \sum_{f \in\Gamma_h \setminus \partial \Omega} \int_{f} \LRp{\frac{\tau_v}{2}\LRb{\bm{A}_n\jump{\bm{v}}}^2 + \frac{\tau_\sigma}{2}\LRb{\bm{A}_n^T\jump{\bm{\sigma}}}^2}\diff{\bm{x}}.  
 %\label{thm:stability}
@@
 \begin{align*}
-&\sum_{f\in \Gamma_v} \int_f \bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v} + \frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma}\\
-&= \sum_{f\in \Gamma_v} \int_f \bm{A}_n^T \bm{\sigma}^-\bm{v}^- - \bm{A}_n \bm{v}^- \bm{\sigma}^- - \tau_v \bm{A}_n^T\bm{A}_n \LRb{\bm{v}^-}^2 = - \sum_{f\in \Gamma_v} \int_f \tau_v \bm{A}_n^T\bm{A}_n \LRb{\bm{v}^-}^2 .
+&\sum_{f\in \Gamma_v} \int_f\LRp{ \bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v} + \frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma}}\diff{\bm{x}}\\
+&= \sum_{f\in \Gamma_v} \int_f \LRp{\bm{A}_n^T \bm{\sigma}^-\bm{v}^- - \bm{A}_n \bm{v}^- \bm{\sigma}^- - \tau_v \bm{A}_n^T\bm{A}_n \LRb{\bm{v}^-}^2}\diff{\bm{x}} = - \sum_{f\in \Gamma_v} \int_f \LRp{\tau_v \bm{A}_n^T\bm{A}_n \LRb{\bm{v}^-}^2}\diff{\bm{x}}.
 \end{align*}
@@
 \[
-\sum_{f\in \Gamma_\sigma} \int_f \bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v} + \frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma} = - \sum_{f\in \Gamma_\sigma} \int_f \tau_\sigma \bm{A}_n\bm{A}_n^T \LRb{\bm{\sigma}^-}^2.
+\sum_{f\in \Gamma_\sigma} \int_f \LRp{\bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v} + \frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma}}\diff{\bm{x}} = - \sum_{f\in \Gamma_\sigma} \int_f \LRp{\tau_\sigma \bm{A}_n\bm{A}_n^T \LRb{\bm{\sigma}^-}^2}\diff{\bm{x}}.
 \]
@@
 \begin{align*}
-\sum_{f\in \Gamma_{\rm abc}} &\int_f \bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v} + \frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma} =\\
-& - \sum_{f\in \Gamma_{\rm abc}} \int_f \frac{\tau_v}{2} \bm{A}_n^T\bm{A}_n \LRb{\bm{v}^-}^2 + \frac{\tau_\sigma}{2} \bm{A}_n\bm{A}_n^T \LRb{\bm{\sigma}^-}^2.  
+\sum_{f\in \Gamma_{\rm abc}} &\int_f \LRp{\bm{A}_n^T\avg{\bm{\sigma}}\bm{v} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}\bm{v} + \frac{1}{2}\bm{A}_n\jump{\bm{v}}\bm{\sigma} + \frac{\tau_\sigma}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}\bm{\sigma}}\diff{\bm{x}} =\\
+& - \sum_{f\in \Gamma_{\rm abc}} \int_f \LRp{ \frac{\tau_v}{2} \bm{A}_n^T\bm{A}_n \LRb{\bm{v}^-}^2 + \frac{\tau_\sigma}{2} \bm{A}_n\bm{A}_n^T \LRb{\bm{\sigma}^-}^2}\diff{\bm{x}}.  
 \end{align*}
@@
 \begin{align}
-&\sum_{D^k\in \Oh} \frac{1}{2}\pd{}{t}\LRp{\LRp{\rho \bm{v},\bm{v}}_{\Lk} + \LRp{\bm{C}^{-1} \bm{\sigma},\bm{\sigma}}_{\Lk}} = - \sum_{f \in\Gamma_h \setminus \partial \Omega} \int_{f} \frac{\tau_v}{2}\LRb{\bm{A}_n\jump{\bm{v}}}^2 + \frac{\tau_\sigma}{2}\LRb{\bm{A}_n^T\jump{\bm{\sigma}}}^2\nonumber\\
-& - \sum_{f\in \Gamma_v} \int_f \tau_v  \LRb{\bm{A}_n\bm{v}^-}^2  - \sum_{f\in \Gamma_\sigma} \int_f \tau_\sigma  \LRb{\bm{A}_n^T\bm{\sigma}^-}^2- \sum_{f\in \Gamma_{\rm abc}} \int_f \frac{\tau_v}{2} \LRb{\bm{A}_n\bm{v}^-}^2 + \frac{\tau_\sigma}{2} \LRb{\bm{A}_n^T\bm{\sigma}^-}^2 \leq 0.
+&\sum_{D^k\in \Oh} \frac{1}{2}\pd{}{t}\LRp{\LRp{\rho \bm{v},\bm{v}}_{\Lk} + \LRp{\bm{C}^{-1} \bm{\sigma},\bm{\sigma}}_{\Lk}} = - \sum_{f \in\Gamma_h \setminus \partial \Omega} \int_{f}\LRp{ \frac{\tau_v}{2}\LRb{\bm{A}_n\jump{\bm{v}}}^2 + \frac{\tau_\sigma}{2}\LRb{\bm{A}_n^T\jump{\bm{\sigma}}}^2}\diff{\bm{x}}\nonumber\\
+& - \sum_{f\in \Gamma_v} \int_f \LRp{\tau_v  \LRb{\bm{A}_n\bm{v}^-}^2}\diff{\bm{x}}  - \sum_{f\in \Gamma_\sigma} \int_f \LRp{\tau_\sigma  \LRb{\bm{A}_n^T\bm{\sigma}^-}^2}\diff{\bm{x}} - \sum_{f\in \Gamma_{\rm abc}} \int_f \LRp{\frac{\tau_v}{2} \LRb{\bm{A}_n\bm{v}^-}^2 + \frac{\tau_\sigma}{2} \LRb{\bm{A}_n^T\bm{\sigma}^-}^2}\diff{\bm{x}} \leq 0.
 \label{eq:thm:stability}
@@
 \[
-\LRp{\bm{M}_w}_{ij} = \int_{D^k} w(\bm{x}) \phi_j(\bm{x})  \phi_i(\bm{x}) \diff x, \qquad \bm{M}_{\bm{W}} = \LRp{\begin{array}{ccc}
+\LRp{\bm{M}_w}_{ij} = \int_{D^k} w(\bm{x}) \phi_j(\bm{x})  \phi_i(\bm{x})\diff{\bm{x}}, \qquad \bm{M}_{\bm{W}} = \LRp{\begin{array}{ccc}
 \bm{M}_{\bm{W}_{1,1}}& \ldots & \bm{M}_{\bm{W}_{1,n}}\\
@@
 \begin{align*}
-\nor{u}_{L^p\LRp{\Omega}} = \LRp{\int_{\Omega} u^p}^{1/p}, \qquad  
+\nor{u}_{L^p\LRp{\Omega}} = \LRp{\int_{\Omega} u^p \diff{\bm{x}}}^{1/p}, \qquad  
 L^p\LRp{\Omega} = \LRc{u: \Omega\rightarrow \mathbb{R}, \quad \nor{u}_{L^p\LRp{\Omega}} < \infty} 
@@
 \end{align*}
-where $\alpha = \LRc{\alpha_1,\alpha_2,\alpha_3}$ is a multi-index such that
-\[
-D^{\alpha}u = \pd{^{\alpha_1}}{x^{\alpha_1}}\pd{^{\alpha_2}}{y^{\alpha_2}}\pd{^{\alpha_3}}{z^{\alpha_3}} u,
-\]
-
+where $\alpha = \LRc{\alpha_1,\ldots,\alpha_d}$ is a multi-index such that
+\begin{align*}
+D^{\alpha}u &= \pd{^{\alpha_1}}{x^{\alpha_1}}\pd{^{\alpha_2}}{y^{\alpha_2}} u, \qquad d = 2\\
+D^{\alpha}u &= \pd{^{\alpha_1}}{x^{\alpha_1}}\pd{^{\alpha_2}}{y^{\alpha_2}}\pd{^{\alpha_3}}{z^{\alpha_3}} u, \qquad d = 3.
+\end{align*}
+
+Let $\Pi_N$ denote the $L^2$ projection on the reference element $\widehat{D}$.
 We define two operators $T_w: L^2\LRp{D^k} \rightarrow P^N\LRp{D^k}$ and $T^{-1}_w: L^2\LRp{D^k} \rightarrow P^N\LRp{D^k}$ such that
-\[
-T_wu = \Pi_N(wu), \qquad \LRp{wT^{-1}_wu,v}_{\Lk} = \LRp{u,v}_{\Lk}.  
-\]
+\begin{align*}
+T_wu &= \Pi_N(wu)\\
+\LRp{wT^{-1}_wu,v}_{\Lk} &= \LRp{u,v}_{\Lk}, \qquad \forall v\in P^N\LRp{D^k}.  
+\end{align*}
 A weighted $L^2$ inner product $\LRp{wu,v}_{\Lk}$ can be approximated by a weight-adjusted inner product
@@
 This theorem relies on a scalar weighted interpolation estimate derived in \cite{warburton2013low,chan2016weight1} for a general non-affine element $D^k$.  
-\begin{theorem}[Theorem 3.1 in \cite{chan2016weight1}]
+\begin{theorem}[Theorem 4 in \cite{chan2016weight1}]
 Let $D^k$ be a quasi-regular element with representative size $h = {\rm diam}\LRp{D^k}$. For $N \geq 0$, $w\in W^{N+1,\infty}\LRp{D^k}$, and $u\in W^{N+1,2}\LRp{D^k}$, 
@@
 %\leq C_N \sum_{\LRb{\alpha}\leq k} \nor{\sum_{\LRb{\beta}\leq \alpha} \LRp{D^{\beta}\bm{W}}\LRp{D^{\alpha-\beta}\bm{v}} }_{L^p}^p\\
-\leq C_N \int  \sum_{\LRb{\alpha}\leq k} \sum_{\LRb{\beta}\leq \LRb{\alpha}} \nor{\LRp{D^{\beta}\bm{W}}\LRp{D^{\alpha-\beta}\bm{v}} }_{p}^p \\
-&\leq C_N\int \LRp{\sum_{\LRb{\alpha}\leq k} \nor{\LRp{D^{\alpha}\bm{W}}}_p}^p \LRp{\sum_{\LRb{\alpha}\leq k} \nor{D^{\alpha}\bm{v}}_{p} }^p\\
+\leq C_N \int  \sum_{\LRb{\alpha}\leq k} \sum_{\LRb{\beta}\leq \LRb{\alpha}} \nor{\LRp{D^{\beta}\bm{W}}\LRp{D^{\alpha-\beta}\bm{v}} }_{p}^p \diff{\bm{x}}\\
+&\leq C_N\int \LRp{\sum_{\LRb{\alpha}\leq k} \nor{\LRp{D^{\alpha}\bm{W}}}_p}^p \LRp{\sum_{\LRb{\alpha}\leq k} \nor{D^{\alpha}\bm{v}}_{p} }^p\diff{\bm{x}}\\
 &\leq C_N \nor{\bm{W}}_{k,p,\infty}^p \nor{\bm{v}}_{k,p}^p,
@@
 \begin{proof}
-The proof is the similar to the scalar case.  Using vector-valued versions of Bramble-Hilbert and a scaling argument for quasi-regular elements yields
+The proof is similar to the scalar case.  Using vector-valued versions of Bramble-Hilbert and a scaling argument for quasi-regular elements yields
 \begin{align*}
@@
 \begin{align*}
-&\sum_{D^k\in \Oh} \frac{1}{2}\pd{}{t}\LRp{\LRp{T^{-1}_{1/\rho} \bm{v},\bm{v}}_{\Lk} + \LRp{T^{-1}_{\bm{C}} \bm{\sigma},\bm{\sigma}}_{\Lk}} \leq - \sum_{f \in\Gamma_h \setminus \partial \Omega} \int_{f} \frac{\tau_v}{2}\LRb{\bm{A}_n\jump{\bm{v}}}^2 + \frac{\tau_\sigma}{2}\LRb{\bm{A}_n^T\jump{\bm{\sigma}}}^2\nonumber\\
-& - \sum_{f\in \Gamma_v} \int_f \tau_v  \LRb{\bm{A}_n\bm{v}^-}^2  - \sum_{f\in \Gamma_\sigma} \int_f \tau_\sigma  \LRb{\bm{A}_n^T\bm{\sigma}^-}^2- \sum_{f\in \Gamma_{\rm abc}} \int_f \frac{\tau_v}{2} \LRb{\bm{A}_n\bm{v}^-}^2 + \frac{\tau_\sigma}{2} \LRb{\bm{A}_n^T\bm{\sigma}^-}^2 \leq 0.
+&\sum_{D^k\in \Oh} \frac{1}{2}\pd{}{t}\LRp{\LRp{T^{-1}_{1/\rho} \bm{v},\bm{v}}_{\Lk} + \LRp{T^{-1}_{\bm{C}} \bm{\sigma},\bm{\sigma}}_{\Lk}} \leq - \sum_{f \in\Gamma_h \setminus \partial \Omega} \int_{f} \LRp{\frac{\tau_v}{2}\LRb{\bm{A}_n\jump{\bm{v}}}^2 + \frac{\tau_\sigma}{2}\LRb{\bm{A}_n^T\jump{\bm{\sigma}}}^2}\diff{\bm{x}}\nonumber\\
+& - \sum_{f\in \Gamma_v} \int_f \LRp{\tau_v  \LRb{\bm{A}_n\bm{v}^-}^2}\diff{\bm{x}}  - \sum_{f\in \Gamma_\sigma} \int_f \LRp{\tau_\sigma  \LRb{\bm{A}_n^T\bm{\sigma}^-}^2}\diff{\bm{x}} - \sum_{f\in \Gamma_{\rm abc}} \int_f \LRp{\frac{\tau_v}{2} \LRb{\bm{A}_n\bm{v}^-}^2 + \frac{\tau_\sigma}{2} \LRb{\bm{A}_n^T\bm{\sigma}^-}^2 }\diff{\bm{x}}\leq 0.
 \end{align*}
@@
 Instead of discretizing the ``strong'' DG formulation (\ref{eq:dgform}), we discretize the ``skew-symmetric'' formulation, where the right hand side of the velocity equations is integrated by parts,\footnote{The choice of which equation to integrate by parts is arbitrary, since integrating the stress equations by parts also results in an energy stable formulation.
-} resulting in the formulation
-\begin{align*}
-\int_{\Dhat} T^{-1}_{(\rho J)^{-1}} \pd{\bm{v}}{t}\bm{w} &= -\int_{\Dhat}\LRp{\sum_{i=1}^d \bm{\sigma},\bm{A}_i \pd{\bm{w}}{\bm{x}_i}}J  + \sum_{f \in \partial D^k} \int_{\widehat{f}}\LRp{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}}\bm{w}J^f \nonumber\\
-\int_{\Dhat}  T^{-1}_{\LRp{J^{-1}\bm{C}}} \pd{\bm{\sigma}}{t}\bm{q} &= \int_{\Dhat} \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}J + \sum_{f\in \partial D^k}\int_{\widehat{f}}\LRp{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}}\bm{q}J^f
-%\LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= -\LRp{ \sum_{i=1}^d \bm{\sigma},\bm{A}_i \pd{\bm{w}}{\bm{x}_i}}_{\Lk}  + \LRa{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk} \nonumber\\
-%\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk}.
+} resulting in the (local) formulation
+\begin{align*}
+\int_{\Dhat} T^{-1}_{(\rho J)^{-1}} \pd{\bm{v}}{t}\bm{w} \diff{\widehat{\bm{x}}}&= -\int_{\Dhat}\LRp{\sum_{i=1}^d \bm{\sigma},\bm{A}_i \pd{\bm{w}}{\bm{x}_i}}J\diff{\widehat{\bm{x}}}  + \sum_{f \in \partial D^k} \int_{\widehat{f}}\LRp{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}}}\bm{w}J^f\diff{\widehat{\bm{x}}} \nonumber\\
+\int_{\Dhat}  T^{-1}_{\LRp{J^{-1}\bm{C}}} \pd{\bm{\sigma}}{t}\bm{q} \diff{\widehat{\bm{x}}}&= \int_{\Dhat} \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}J \diff{\widehat{\bm{x}}} + \sum_{f\in \partial D^k}\int_{\widehat{f}}\LRp{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}}}\bm{q}J^f\diff{\widehat{\bm{x}}}
 \end{align*}
@@
 \begin{equation}
-A_{\min} \nor{\bm{\eta}}^2_{\L} \leq C\int_0^T  h^{N+1}A_{\max}\nor{\bm{A}_0}_{W^{N+1,\infty}(\Oh)}\nor{\pd{\bm{U}}{t}}_{\L} \nor{\bm{\eta}}_{\L} + h^{2N+1} \nor{\bm{U}}^2_{W^{N+1,2}(\Omega)}.
+A_{\min} \nor{\bm{\eta}}^2_{\L} \leq C\int_0^T  h^{N+1}A_{\max}\nor{\bm{A}_0}_{W^{N+1,\infty}(\Oh)}\nor{\pd{\bm{U}}{t}}_{\L} \nor{\bm{\eta}}_{\L} + h^{2N+1} \nor{\bm{U}}^2_{W^{N+1,2}(\Omega)} \diff{t}.
 \label{eq:intermed}
@@
 \begin{align*}
-\nor{\bm{\eta}}_{\L} &\leq \frac{1}{A_{\min}} \int_0^T CA_{\max} h^{N+1}\nor{\bm{A}_0}_{W^{N+1,\infty}(\Oh)}\nor{\pd{\bm{U}}{t}}_{\L} + \sup_{t\in [0,T]} \sqrt{\int_0^T C h^{2N+1} \nor{\bm{U}}^2_{W^{N+1,2}(\Omega)}}\\
+\nor{\bm{\eta}}_{\L} &\leq \frac{1}{A_{\min}} \int_0^T CA_{\max} h^{N+1}\nor{\bm{A}_0}_{W^{N+1,\infty}(\Oh)}\nor{\pd{\bm{U}}{t}}_{\L}\diff{t} + \sup_{t\in [0,T]} \sqrt{\int_0^T C h^{2N+1} \nor{\bm{U}}^2_{W^{N+1,2}(\Omega)}\diff{t}}\\
 &\leq \frac{C T h^{N+1/2}}{A_{\min}} \sup_{t \in [0,T]}\LRp{h^{1/2} A_{\max}\nor{\bm{A}_0}_{W^{N+1,\infty}(\Oh)}\nor{\pd{\bm{U}}{t}}_{\L}  +  \nor{\bm{U}}_{W^{N+1,2}(\Oh)}}.
@@
 where $\omega = \sqrt{2\mu}$.
-Figure~\ref{fig:harmonic_osc} shows $L^2$ errors computed at time $T = 1$, using uniform triangular meshes constructed by bisecting a uniform mesh of quadrilaterals along the diagonal.
-For $N = 1,\ldots, 5$, optimal $O(h^{N+1})$ rates of convergence are observed.
+Zero traction boundary conditions are imposed.
+Figure~\ref{fig:harmonic_osc} shows $L^2$ errors computed at time $T = 5$, using uniform triangular meshes constructed by bisecting a uniform mesh of quadrilaterals along the diagonal.
+For $N = 1,\ldots, 5$, rates near the optimal $O(h^{N+1})$ rates of convergence are observed.
 
@@
 \addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.162198)(0.25,0.0389974)(0.125,0.0095562)(0.0625,0.00227802)};
-\logLogSlopeTriangleFlip{0.4}{0.125}{0.81}{2.05}{blue}
+coordinates{(0.5,0.281186)(0.25,0.0607224)(0.125,0.0114231)(0.0625,0.00232631)};
+\logLogSlopeTriangleFlip{0.4}{0.125}{0.805}{2.30}{blue}
 
 \addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.0128844)(0.25,0.00156487)(0.125,0.000197047)(0.0625,2.42153e-05)};
-\logLogSlopeTriangleFlip{0.4}{0.125}{0.65}{3.01}{red}
+coordinates{(0.5,0.010807)(0.25,0.00109063)(0.125,0.000126864)(0.0625,1.53214e-05)};
+\logLogSlopeTriangleFlip{0.4}{0.125}{0.63}{3.05}{red}
 
 \addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.00110186)(0.25,6.9321e-05)(0.125,4.31386e-06)(0.0625,2.49229e-07)};
-\logLogSlopeTriangleFlip{0.4}{0.125}{0.49}{4.06}{black}
+coordinates{(0.5,0.000760654)(0.25,4.44169e-05)(0.125,2.63258e-06)(0.0625,1.62546e-07)};
+\logLogSlopeTriangleFlip{0.4}{0.125}{0.47}{4.02}{black}
 
 \addplot[color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,8.42986e-05)(0.25,2.47716e-06)(0.125,7.21337e-08)(0.0625,2.17472e-09)};
-\logLogSlopeTriangleFlip{0.4}{0.125}{0.325}{5.08}{magenta}
+coordinates{(0.5,4.68589e-05)(0.25,1.48545e-06)(0.125,4.55009e-08)(0.0625,1.39823e-09)};
+\logLogSlopeTriangleFlip{0.4}{0.125}{0.305}{5.02}{magenta}
 
 \addplot[color=violet,mark=pentagon*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,4.91233e-06)(0.25,7.57947e-08)(0.125,1.10789e-09)(0.0625,1.63538e-11)};
-\logLogSlopeTriangleFlip{0.4}{0.125}{0.15}{6.09}{violet}
+coordinates{(0.5,3.03577e-06)(0.25,4.50412e-08)(0.125,6.98053e-10)(0.0625,1.17913e-11)};
+\logLogSlopeTriangleFlip{0.4}{0.125}{0.126}{5.89}{violet}
 
@@
 
-Figure~\ref{fig:rayleigh} shows $L^2$ errors for both waves at time $T = 1$.
+Figure~\ref{fig:rayleigh} shows $L^2$ errors for both waves at time $T = 5$.
 Computed convergence rates fall between the optimal rate of $O(h^{N+1})$ and theoretical rate of $O(h^{N+1/2})$ \cite{johnson1986analysis}.
@@
     xmin=.025, xmax=1,
-    ymin=1e-8, ymax=1.5,
+    ymin=5e-9, ymax=1.5,
     legend pos=south east,
@@
 \addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,1.04483)(0.25,0.342112)(0.125,0.102275)(0.0625,0.0277191)};    
-\logLogSlopeTriangleFlip{.39}{0.11}{0.815}{1.81}{blue}
+coordinates{(0.5,0.987525)(0.25,0.856084)(0.125,0.272343)(0.0625,0.0495455)};
+\logLogSlopeTriangleFlip{.39}{0.11}{0.855}{2.46}{blue}
 
 \addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.287061)(0.25,0.0563744)(0.125,0.00785962)(0.0625,0.000992348)};    
-\logLogSlopeTriangleFlip{.38}{0.1}{0.64}{2.91}{red}
+coordinates{(0.5,0.762022)(0.25,0.0822137)(0.125,0.00510445)(0.0625,0.000517961)};
+\logLogSlopeTriangleFlip{.38}{0.1}{0.625}{3.30}{red}
 
 \addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.0736249)(0.25,0.00747192)(0.125,0.000529189)(0.0625,3.53252e-05)};   
-\logLogSlopeTriangleFlip{.38}{0.1}{0.475}{3.86}{black}
+coordinates{(0.5,0.133471)(0.25,0.00400974)(0.125,0.000280342)(0.0625,1.93442e-05)};
+\logLogSlopeTriangleFlip{.38}{0.1}{0.47}{3.86}{black}
 
 \addplot[color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.0199958)(0.25,0.000774494)(0.125,2.70714e-05)(0.0625,8.78659e-07)};   
-\logLogSlopeTriangleFlip{.38}{0.1}{0.285}{4.89}{magenta}
+coordinates{(0.5,0.0130378)(0.25,0.000405004)(0.125,1.36821e-05)(0.0625,4.57998e-07)};
+\logLogSlopeTriangleFlip{.38}{0.1}{0.28}{4.90}{magenta}
 
 \addplot[color=violet,mark=pentagon*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.00304235)(0.25,6.78379e-05)(0.125,1.18747e-06)(0.0625,1.92431e-08)};    
-\logLogSlopeTriangleFlip{.38}{0.1}{0.09}{5.89}{violet}
+coordinates{(0.5,0.00157091)(0.25,3.42541e-05)(0.125,5.7373e-07)(0.0625,9.09776e-09)};
+\logLogSlopeTriangleFlip{.38}{0.1}{0.085}{5.98}{violet}
 
@@
     xmin=.025, xmax=1,
-    ymin=2e-7, ymax=1.5,
+    ymin=1e-7, ymax=2.5,
     legend pos=south east,
@@
 \addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.97515)(0.25,0.667267)(0.125,0.26441)(0.0625,0.0643182)};   
-\logLogSlopeTriangleFlip{.39}{0.11}{0.835}{2.04}{blue}
+coordinates{(0.5,1.00019)(0.25,1.00515)(0.125,0.633842)(0.0625,0.161762)};
+\logLogSlopeTriangleFlip{.39}{0.11}{0.875}{1.97}{blue}
 
 \addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.659983)(0.25,0.191328)(0.125,0.0172492)(0.0625,0.00168533)};   
-\logLogSlopeTriangleFlip{.38}{0.1}{0.62}{3.41}{red}
+coordinates{(0.5,1.01296)(0.25,0.503696)(0.125,0.0316637)(0.0625,0.00188428)};
+\logLogSlopeTriangleFlip{.38}{0.1}{0.63}{4.07}{red}
 
 \addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.366154)(0.25,0.024884)(0.125,0.0014833)(0.0625,0.000103456)};    
-\logLogSlopeTriangleFlip{.38}{0.1}{0.45}{3.96}{black}
+coordinates{(0.5,0.865781)(0.25,0.0445608)(0.125,0.00158453)(0.0625,0.000113704)};
+\logLogSlopeTriangleFlip{.38}{0.1}{0.465}{3.80}{black}
+
 \addplot[color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.141902)(0.25,0.00437419)(0.125,0.000154364)(0.0625,5.20702e-06)};   
-\logLogSlopeTriangleFlip{.38}{0.1}{0.26}{4.86}{magenta}
+coordinates{(0.5,0.343325)(0.25,0.00491357)(0.125,0.000156595)(0.0625,5.13222e-06)};
+\logLogSlopeTriangleFlip{.38}{0.1}{0.28}{4.93}{magenta}
 
 \addplot[color=violet,mark=pentagon*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.0327909)(0.25,0.000752122)(0.125,1.34499e-05)(0.0625,2.18614e-07)};   
-\logLogSlopeTriangleFlip{0.38}{0.1}{0.075}{5.87}{violet}
+coordinates{(0.5,0.0682545)(0.25,0.000718077)(0.125,1.37715e-05)(0.0625,2.30844e-07)};
+\logLogSlopeTriangleFlip{0.37}{0.1}{0.0985}{5.90}{violet}
 
@@
 }
+\caption{Convergence of $L^2$ errors for Rayleigh and Lamb wave solutions.}
 \label{fig:rayleigh}
-\caption{Convergence of $L^2$ errors for Rayleigh and Lamb wave solutions.}
 \end{figure}
@@
 We use the propagation of Rayleigh waves to examine the behavior of WADG for near-incompressible materials.
-We follow \cite{sjogreen2012fourth,appelo2015energy} and fix $\lambda = 1$ and set $\mu = 1, .1, .01, .001, .0001$.
+We follow \cite{sjogreen2012fourth, appelo2015energy} and fix $\lambda = 1$ and set $\mu = 1, .1, .01, .001, .0001$.
 Since the Rayleigh wave propagates with speed proportional to $\sqrt{\mu}$, we compute $L^2$ errors at the final time $1 / (4\sqrt{\mu})$ to ensure a fair comparison between solutions at different values of $\mu$.
@@
 The relative errors for $\bm{\sigma}$ grow as $\mu/\lambda \rightarrow 0$.
-This is not surprising, as the constant in the error estimates of Section~\ref{sec:mwadgprop} blows up in the incompressible limit.
+This is not surprising, as the constant in the error estimates of Section~\ref{sec:mwadgprop} depends on $\nor{\bm{C}}$, which blows up in the incompressible limit.
 However, because the magnitude of $\bm{\sigma}$ also decreases as $\mu/\lambda\rightarrow 0$, relative errors for $\bm{v}$ remain roughly the same magnitude for near-incompressible materials.
@@
 Figure~\ref{fig:stoneley} shows $L^2$ errors for two uniform meshes of triangles constructed by bisecting a quadrilateral mesh of $K_{\rm 1D} \times 5K_{\rm 1D}$ elements.
-Figure~\ref{subfig:stoneley1} shows errors when $K_{\rm 1D}$ is even and the mesh is fitted to the interface at $y = 0$, while Figure~\ref{subfig:stoneley2} shows errors when $K_{\rm 1D}$ is odd and the interface cuts through element interiors.
-
-When the mesh is fitted to the interface, computed convergence rates are between $O(h^{N+1/2})$ and $O(h^{N+1})$.
+Figure~\ref{subfig:stoneley1} shows errors at time $T=5$ when $K_{\rm 1D}$ is even and the mesh is fitted to the interface at $y = 0$, while Figure~\ref{subfig:stoneley2} shows errors when $K_{\rm 1D}$ is odd and the interface cuts through element interiors.
+
+When the mesh is fitted to the interface, computed convergence rates match the estimated $O(h^{N+1/2})$ rate.
 When the mesh is not fitted to the interface exactly, we compute the application of the weight-adjusted mass matrix inverse using a quadrature rule from Xiao and Gimbutas \cite{xiao2010quadrature} which is exact for degree $2N+1$ polynomials.
@@
     xmin=.025, xmax=1,
-    ymin=.75e-11, ymax=.25,
+    ymin=1e-11, ymax=1.5,
     xmajorgrids=true,
@@
 \addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.108456)(0.25,0.0385149)(0.125,0.0138288)(0.0625,0.0043988)};    
-\logLogSlopeTriangleFlip{.4}{0.125}{0.86}{1.65}{blue}
+coordinates{(0.5,0.291816)(0.25,0.110431)(0.125,0.0409099)(0.0625,0.0140493)};
+\logLogSlopeTriangleFlip{.4}{0.125}{0.85}{1.54}{blue}
 
 \addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.00887301)(0.25,0.00128102)(0.125,0.000186756)(0.0625,2.73376e-05)};    
-\logLogSlopeTriangleFlip{.4}{0.125}{0.65}{2.77}{red}
+coordinates{(0.5,0.0360738)(0.25,0.00627358)(0.125,0.00107612)(0.0625,0.0001848)};
+\logLogSlopeTriangleFlip{.4}{0.125}{0.675}{2.54}{red}
 
 \addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,0.000930299)(0.25,7.92165e-05)(0.125,5.2534e-06)(0.0625,3.67676e-07)};    
-\logLogSlopeTriangleFlip{.4}{0.125}{0.475}{3.84}{black}
+coordinates{(0.5,0.00234802)(0.25,0.000185264)(0.125,1.51976e-05)(0.0625,1.28801e-06)};
+\logLogSlopeTriangleFlip{.4}{0.125}{0.485}{3.56}{black}
 
 \addplot[color=magenta,mark=diamond*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,4.34823e-05)(0.25,1.45784e-06)(0.125,4.92754e-08)(0.0625,1.74105e-09)};   
-\logLogSlopeTriangleFlip{.4}{0.125}{0.26}{4.82}{magenta}
+coordinates{(0.5,0.000142893)(0.25,5.99393e-06)(0.125,2.55178e-07)(0.0625,1.10265e-08)};
+\logLogSlopeTriangleFlip{.4}{0.125}{0.31}{4.53}{magenta}
 
 \addplot[color=violet,mark=pentagon*,semithick, mark options={fill=markercolor}]
-coordinates{(0.5,2.97005e-06)(0.25,4.94512e-08)(0.125,8.48451e-10)(0.0625,1.47109e-11)};  
-\logLogSlopeTriangleFlip{0.4}{0.125}{0.075}{ 5.85}{violet}
-
+coordinates{(0.5,6.55362e-06)(0.25,1.31822e-07)(0.125,2.7319e-09)(0.0625,5.82544e-11)};
+\logLogSlopeTriangleFlip{0.4}{0.125}{0.11}{ 5.55}{violet}
+%1.5419    2.5418    3.5606    4.5325    5.5514
 %\node at (axis cs:.03,6.8e-10) {$u$ discontinuous};

### 2017-05-09 docs/splitform/main.tex
updates to elasticity notes (sisc rejected reviewer comments)

@@
 \theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
+\newtheorem{definition}{Definition}
+\theoremstyle{lemma}
+\newtheorem{lemma}{Lemma}
 \newtheorem*{remark}{Remark}
@@
 \author{Jesse Chan}
-\title{Energy stable discontinuous Galerkin formulations and discrete differential operators}
+\title{Energy stable discontinuous Galerkin formulations using discrete differential operators}
 
@@
 \]
+On faces which coincide with the boundary $\partial \Omega$, the average and jump are defined as
+\[
+\jump{u} = 0, \qquad \avg{u} = u.
+\]
 The jump and average of vector fields are defined component-wise using the jumps and averages of components.
+
+We will assume for now that all elements are affine mappings of a reference element $\widehat{D}$.
 
@@
 \[
-\LRp{u,v}_{\Omega} = \sum_{D^k} \int_{D^k} uv \diff{x} = \sum_{D^k} \int_{\widehat{D}} uv J \diff{\widehat{x}}.  
+\LRp{u,v}_{\Omega} = \sum_{D^k} \int_{D^k} uv \diff{x} = \sum_{D^k} \int_{\widehat{D}} uv J \diff{\widehat{x}},
 \]
-In practice, these integrals are computed using quadrature, such that 
-\[
-\int_{\widehat{D}}uv = \sum_{i=1}^{N_q} u(\bm{x}_i)v(\bm{x}_i) w_i,
-\]
-where $N_q$ is the number of quadrature points.  The only assumptions we make upon this quadrature is that it is sufficiently accurate such that 
-\begin{enumerate}
-\item The quadrature induces an equivalent inner product over the reference element $\widehat{D}$
-\item Integration by parts holds (with respect to the reference coordinates $\widehat{x}$).
-\end{enumerate}
-
-We also introduce the $L^2$ projection $\Pi_N: L^2\LRp{\Omega}\rightarrow P^N$ such that 
+and the associated $L^2$ projection $\Pi_N: L^2\LRp{\Omega}\rightarrow P^N$
 \[
@@
 
+In practice, these integrals are computed using quadrature, such that 
+\begin{equation}
+\int_{\widehat{D}}uv = \sum_{i=1}^{N_q} u(\bm{x}_i)v(\bm{x}_i) w_i,
+\label{eq:quad}
+\end{equation}
+where $N_q$ is the number of quadrature points.  The only assumptions we make upon this quadrature is that it is sufficiently accurate such that 
+\begin{enumerate}
+\item The quadrature induces an $L^2$-equivalent inner product over the reference element $\widehat{D}$.
+\item The quadrature is sufficiently accurate such that integration by parts holds with respect to the reference coordinates $\widehat{x}$.
+\end{enumerate}
+
 \subsection{Discrete derivatives}
-We introduce two discrete DG derivatives in this section.  The first is the implicit definition used in \cite{di2011mathematical}
-
+We introduce two discrete DG derivatives in this section.  The first is the implicit definition used in \cite{di2011mathematical}:
 \begin{definition}
@@
 \[
-\LRp{{D}^i u,v}_{\Oh} = \sum_{D^k} \LRp{\LRp{-u,\pd{v}{\bm{x}_i}}_{\LK} + \LRa{\avg{u},v\bm{n}_i}_{\LdK}}.
+\LRp{{D}^i u,v}_{\Oh} = \sum_{D^k} \LRp{\LRp{-u,\pd{v}{\bm{x}_i}}_{\LK} + \LRa{\avg{u},v\bm{n}_i}_{\LdK}}, \qquad \forall v \in V_h.
 \]
@@
 
-We note that, for $u,v \in V_h$, integration by parts yields an equivalent definition of $D^i$ found in \cite{hesthaven2004high,Warburton20063205}
+We note that, for $u,v \in V_h$, (discrete) integration by parts yields an equivalent definition of $D^i$ found in \cite{hesthaven2004high,Warburton20063205}
 \[
-\LRp{{D}^i u,v}_{\Oh} = \sum_{D^k} \LRp{\LRp{\pd{u}{\bm{x}_i},v}_{\LK} + \frac{1}{2}\LRa{\jump{u},v\bm{n}_i}_{\LdK}}.
+\LRp{{D}^i u,v}_{\Oh} = \sum_{D^k} \LRp{\LRp{\pd{u}{\bm{x}_i},v}_{\LK} - \frac{1}{2}\LRa{\jump{u},v\bm{n}_i}_{\LdK}}, \qquad \forall v\in V_h.
 \]
-However, for $u \not\in V_h$, these two definitions are not equivalent.  This motivates the definition of a \textit{discrete} DG differentiation operator:
+The main property of the discrete derivatives utilized in this work is a discrete integration-by-parts formula.  
+\begin{lemma}
+Let $u,v\in V_h$.  Then, 
+\begin{align*}
+\LRp{D^i u,v} &= \LRp{-u,D^i_h v} + \int_{\partial \Omega} uv \bm{n}_i\\
+\LRp{D_h^i u,v} &= \LRp{-u,D^i_h v} + \int_{\partial \Omega} uv \bm{n}_i.
+\end{align*}
+\label{lemma:ibp}
+\end{lemma}
+\begin{proof}
+Ez
+\end{proof}
+
+For $u \not\in V_h$, due to the fact that the quadrature rule (\ref{eq:quad}) may not be exact, and these two definitions are not equivalent and Lemma~\ref{lemma:ibp} does not hold.
+This motivates the definition of a \textit{discrete} DG differentiation operator:
 \begin{definition}
@@
 
-The main property of
+Using these discrete derivatives, we can define discrete gradient and divergence operators
+\begin{definition}
+Let $u \in V_h$ and $\bm{u}\in \LRp{V_h}^d$.  Then, discrete gradient and divergence operators are defined as follows:
+\begin{align*}
+\LRp{\Grad^{\rm DG} u}_i &= D^i u, \qquad \Grad^{\rm DG}\cdot \bm{u} = \sum_{i=1}^d D^i \bm{u}_i\\
+\LRp{\Grad^{\rm DG}_h u}_i &= D_h^i u, \qquad \Grad^{\rm DG}_h\cdot \bm{u} = \sum_{i=1}^d D_h^i \bm{u}_i.
+\end{align*}
+\end{definition}
+
+\subsection{Examples}
+
+\begin{itemize}
+\item Tensor product elements: inner product computed using GLL nodes, $\Pi_N$ reduces to nodal interpolation
+\item General elements: $\Pi_N$ is just quadrature-based projection.  For accuracy, degree $2N$ or higher?  
+\item General elements: $\Pi_N$ can be taken to be nodal interpolation.  
+\item Pyramids?  
+\end{itemize}
 
@@
 \]
-\section{Variable advection}
 
-A split formulation for advection is 
-\[
-\LRp{\pd{u}{t},v} + \frac{1}{2}\LRp{\Grad_h\cdot \Pi_N \LRp{ \bm{\beta}u},v} + \frac{1}{2}\LRp{\bm{\beta}\cdot\Grad_h u,v} + \frac{1}{2}\LRp{\LRp{\Grad\cdot \bm{\beta}} u,v} = 0.
-\]
-Taking $v = u$ yields and using $\LRp{\Grad_h u, \bm{v}} = \LRp{-u, \Grad_h \cdot \bm{v}}$ yields the energy statement
-\[
-\frac{1}{2}\nor{u}^2 + \frac{1}{2}\LRp{\Grad_h\cdot \Pi_N \LRp{ \bm{\beta}u},u} - \frac{1}{2}\LRp{ u, \Grad_h\cdot \Pi_N\LRp{\bm{\beta}u}} = \frac{1}{2}\LRp{-\LRp{\Grad \cdot \bm{\beta}} u,u},
-\]
-implying that $\frac{1}{2}\nor{u}^2 = 0$ if $\Grad\cdot \bm{\beta} = 0$, or that the method is energy conserving.
-The only difference in this formulation is the introduction of $\Pi_N$, which can be defined at a discrete level using any quadrature scheme for which a discrete projection is well-defined.
-
-\section{Local conservation}
+\subsection{Local conservation}
 
@@
 
+\section{Variable advection}
+
+A split formulation for advection is 
+\[
+\LRp{\pd{u}{t},v} + \frac{1}{2}\LRp{\Grad_h\cdot \Pi_N \LRp{ \bm{\beta}u},v} + \frac{1}{2}\LRp{\bm{\beta}\cdot\Grad_h u,v} + \frac{1}{2}\LRp{\LRp{\Grad\cdot \bm{\beta}} u,v} = 0.
+\]
+Taking $v = u$ yields and using $\LRp{\Grad_h u, \bm{v}} = \LRp{-u, \Grad_h \cdot \bm{v}}$ yields the energy statement
+\[
+\frac{1}{2}\nor{u}^2 + \frac{1}{2}\LRp{\Grad_h\cdot \Pi_N \LRp{ \bm{\beta}u},u} - \frac{1}{2}\LRp{ u, \Grad_h\cdot \Pi_N\LRp{\bm{\beta}u}} = \frac{1}{2}\LRp{-\LRp{\Grad \cdot \bm{\beta}} u,u},
+\]
+implying that $\frac{1}{2}\nor{u}^2 = 0$ if $\Grad\cdot \bm{\beta} = 0$, or that the method is energy conserving.
+The only difference in this formulation is the introduction of $\Pi_N$, which can be defined at a discrete level using any quadrature scheme for which a discrete projection is well-defined.
+
 \section{Discrete DG derivatives}
@@
 
-Can also write the inconsistent term as 
-\[
-\LRp{-{ \bm{\beta}u},\Grad_h v} %= -\LRp{\Grad_h v,\bm{\beta}u} = \sum_{D^k} \LRp{ v,\Grad\cdot \LRp{\bm{\beta}u}}_{D^k} - \LRa{\avg{v},\bm{\beta}\cdot \bm{n} u}
-\]
-but this results in the same inconsistent flux term due to the fact that $\Grad_h$ is only defined for test functions in the polynomial space.
-
 Note: can also use interpolants in a stable manner if using $D_h$.
@@
 
-For curvilinear coordinates,
+\section{Curvilinear coordinates}
+
+For curvilinear meshes, $J$ and $\bm{G}$ are no longer constant over each element.
+However, for isoparametric mappings (and special polynomial interpolations of general curvilinear mappings), metric identities hold \cite{kopriva2006metric} such that $\widehat{\Grad} \cdot \LRp{J\bm{G}} = 0$.
+\[
+\LRp{\Grad u,\bm{v}}_{D^k} = \LRp{J\bm{G}\widehat{\Grad} u,\bm{v}}_{\widehat{D}} = \LRp{-u,\widehat{\Grad} \cdot \LRp{J\bm{G}\bm{v}}}  + \LRa{J\bm{G}\widehat{\bm{n}} u,\bm{v}} 
+\]
+Noting that $J\bm{G}\widehat{\bm{n}} = \bm{n} J^f$ \cite{hesthaven2007nodal}, we recover integration by parts over the physical element.
 
@@
 
-\section{Standard entropy stability estimates}
-
-Given a nonlinear conservation law
-\begin{align*}
-\pd{u}{t} + \pd{f(u)}{x} = 0
-\end{align*}
-DG formulation is usually
-\[
-\LRp{\pd{u}{t},v} + \LRp{D_h f,v} = 0.
-\]
-Taking $v = u$ gives
-\[
-\frac{1}{2}\nor{u}^2 + \int_{D^k}{u D_h f} = 0.
-\]
-Continuous entropy stability relies on the introduction of $F,G$ such that
-\[
-\pd{F}{u} = u\pd{f}{u}, \qquad \pd{G}{u} = f, \qquad F = uf - G.
-\]
-Then, relying on product and chain rules
-\[
-\int_{D^k}{u \pd{f}{x}} = \int_{D^k}{\pd{(uf)}{x} - f\pd{u}{x}} = \int_{D^k}{\pd{(uf)}{x} - \pd{G}{u}\pd{u}{x}} =  \int_{D^k}{\pd{\LRp{uf-G}}{x}} = \int_{D^k}{\pd{F}{x}} = \int_{\partial D^k} F.
-\]
-These boundary terms can be made to cancel with appropriately defined numerical fluxes, resulting in entropy conservative schemes.
+%\section{Standard entropy stability estimates}
+%
+%Given a nonlinear conservation law
+%\begin{align*}
+%\pd{u}{t} + \pd{f(u)}{x} = 0
+%\end{align*}
+%DG formulation is usually
+%\[
+%\LRp{\pd{u}{t},v} + \LRp{D_h f,v} = 0.
+%\]
+%Taking $v = u$ gives
+%\[
+%\frac{1}{2}\nor{u}^2 + \int_{D^k}{u D_h f} = 0.
+%\]
+%Continuous entropy stability relies on the introduction of $F,G$ such that
+%\[
+%\pd{F}{u} = u\pd{f}{u}, \qquad \pd{G}{u} = f, \qquad F = uf - G.
+%\]
+%Then, relying on product and chain rules
+%\[
+%\int_{D^k}{u \pd{f}{x}} = \int_{D^k}{\pd{(uf)}{x} - f\pd{u}{x}} = \int_{D^k}{\pd{(uf)}{x} - \pd{G}{u}\pd{u}{x}} =  \int_{D^k}{\pd{\LRp{uf-G}}{x}} = \int_{D^k}{\pd{F}{x}} = \int_{\partial D^k} F.
+%\]
+%These boundary terms can be made to cancel with appropriately defined numerical fluxes, resulting in entropy conservative schemes.
 The challenge in reproducing this in the discrete case is the lack of a product and chain rule for inexact quadratures.
 Jameson deals with this by introducing an integrated flux
-\[
-G = \frac{1}{2}\int_{-1}^1 f \LRp{ \avg{u} + \theta \jump{u}} \diff \theta.
-\]
-for which finite volume schemes satisfy (in a rough sense)
-\[
-\jump{G} = \avg{G} \jump{u}.
-\]
-Fischer and Carpenter (also Gassner and co-workers) use a similar idea but combine a symmetric two-point flux approximation with properties of SBP matrices to get 
-\[
-2\LRp{W,DF} = \LRp{W,DF} - \LRp{DW,F} + \LRa{W,F} =  \sum_i \sum_j (W_i-W_j) (MD)_{ij} F(U_i,U_j) + \LRa{W,F}.
-\]
-
-\section{Entropy splitting}
-
-It's currently unclear how to extend the symmetric two-point flux approximation to quadrature-based DG methods.
+%\[
+%G = \frac{1}{2}\int_{-1}^1 f \LRp{ \avg{u} + \theta \jump{u}} \diff \theta.
+%\]
+%for which finite volume schemes satisfy (in a rough sense)
+%\[
+%\jump{G} = \avg{G} \jump{u}.
+%\]
+%Fischer and Carpenter (also Gassner and co-workers) use a similar idea but combine a symmetric two-point flux approximation with properties of SBP matrices to get 
+%\[
+%2\LRp{W,DF} = \LRp{W,DF} - \LRp{DW,F} + \LRa{W,F} =  \sum_i \sum_j (W_i-W_j) (MD)_{ij} F(U_i,U_j) + \LRa{W,F}.
+%\]
+%
+%\section{Entropy splitting}
+%
+%It's currently unclear how to extend the symmetric two-point flux approximation to quadrature-based DG methods.
 

### 2017-05-09 matlab/cfd/Euler1D_hughesEntropySplit.m

@@ -7,7 +7,7 @@ function Euler1D
 Globals1D;
 
 % Order of polymomials used for approximation
-N = 3;
+N = 4;
 K1D = 32;
 FinalTime = 2;
 
@@ -28,7 +28,7 @@ xp=  Vp*x;
 global gamma Vq Pq
 gamma = 1.4;
 
-Nq = 2*N+2;
+Nq = 2*N+1;
 % Nq = N;
 [rq wq] = JacobiGQ(0,0,Nq);
 Vq = Vandermonde1D(N,rq)/V;
@@ -69,10 +69,14 @@ dfdW33 = @(W1,W2,W3)W2.*1.0./W3.^4.*exp((W1.*W3-W3.*gamma-W2.^2.*(1.0./2.0))./(W
 W1 = Pq*W1q;
 W2 = Pq*W2q;
 W3 = Pq*W3q;
+
+% W1 = limit1D(W1);
+% W2 = limit1D(W2);
+% W3 = limit1D(W3);
 % keyboard
 % [rho m E] = WToU(W1,W2,W3);
 % hold on
-% plot(xq,E,'x')
+% plot(x,W1,'o-');return
 
 
 time = 0;
@@ -107,11 +111,17 @@ for tstep=1:Nsteps
         [rhs1, rhs2, rhs3] = EulerRHS1D(W1tmp,W2tmp,W3tmp);
         W1tmp = (1-ssprk(i))*W1 + ssprk(i)*(W1tmp + dt*rhs1);
         W2tmp = (1-ssprk(i))*W2 + ssprk(i)*(W2tmp + dt*rhs2);
-        W3tmp = (1-ssprk(i))*W3 + ssprk(i)*(W3tmp + dt*rhs3);        
+        W3tmp = (1-ssprk(i))*W3 + ssprk(i)*(W3tmp + dt*rhs3);
+        
+%         W1tmp = limit1D(W1tmp);
+%         W2tmp = limit1D(W2tmp);
+%         W3tmp = limit1D(W3tmp);
+
     end
     W1 = W1tmp;
     W2 = W2tmp;
     W3 = W3tmp;
+    
     
 %     % low storage RK
 %     for INTRK = 1:5
@@ -256,12 +266,21 @@ if opt==1 % plain entropy variables
     %         global wJq
     %     sum(sum(wJq.*(W1.*r1 + W2.*r2 + W3.*r3)))
     
+    r1 = Pq*r1;
[…]

### 2017-05-09 matlab/cfd/Euler1D_split.m

@@ -7,9 +7,9 @@ function Euler1D
 Globals1D;
 
 % Order of polymomials used for approximation
-N = 7;
+N = 5;
 K1D = 16;
-FinalTime = 2;
+FinalTime = 1;
 
 % Generate simple mesh
 [Nv, VX, K, EToV] = MeshGen1D(-1,1,K1D);
@@ -28,8 +28,8 @@ xp=  Vp*x;
 global gamma Vq Pq
 gamma = 1.4;
 
-Nq = 2*N+2;
-Nq = N;
+Nq = 2*N+1;
+% Nq = N;
 [rq wq] = JacobiGQ(0,0,Nq);
 Vq = Vandermonde1D(N,rq)/V;
 Pq = (Vq'*diag(wq)*Vq)\(Vq'*diag(wq));
@@ -89,6 +89,7 @@ for tstep=1:Nsteps
         hold off
         
 %         axis([-.5 .5 -1 3])
+        axis([-1 1 0 4])
         title(sprintf('Time = %f\n',time))
         drawnow
     end
@@ -97,13 +98,10 @@ for tstep=1:Nsteps
     mq = Vq*m;
     rhoq = Vq*rho;
     Eq = Vq*E;
-    uq = mq./rhoq;    
+    uq = mq./rhoq;        
+    pq = (gamma-1.0)*(Eq - 0.5*rhoq.*uq.^2);
     
-    u = mq./rhoq;
-    p = (gamma-1.0)*(Eq - 0.5*rhoq.*u.^2);
-    
-    S(tstep) = sum(sum(wJq.*real(log(p.*rhoq.^(-gamma)))));
-    
+    S(tstep) = sum(sum(wJq.*real(log(pq.*rhoq.^(-gamma)))));    
     
     if mod(tstep,1000)==0
         disp(sprintf('tstep = %d out of %d\n',tstep,Nsteps))
@@ -195,7 +193,7 @@ if opt==1
     rhsm   = rhsm   + .5*LIFT*(Fscale.*LFc.*jump(m_in));
     rhsE   = rhsE   + .5*LIFT*(Fscale.*LFc.*jump(E_in));
 
-elseif opt==2
+elseif opt==2 % KE conserving split form
     
     rhsrho  = -Dh(Pq*(rho.*u));
[…]

### 2017-05-09 matlab/elas/Elasticity2D.m

@@ -10,7 +10,7 @@ if nargin==0
     N = 5;
 end
 c_flag = 0;
-FinalTime = 1;
+FinalTime = 5;
 
 % filename = 'Grid/Other/block2.neu';
 % [Nv, VX, VY, K, EToV] = MeshReaderGambit2D(filename);
@@ -160,14 +160,14 @@ U{3} = sxx(x,y,0);
 U{4} = syy(x,y,0);
 U{5} = sxy(x,y,0);
 
-x0 = .5; y0 = .5;
-p = exp(-25^2*((x-x0).^2 + (y-y0).^2));
-u = zeros(Np, K);
-U{1} = p;
-U{2} = u;
-U{3} = u;
-U{4} = u;
-U{5} = u;
+% x0 = .5; y0 = .5;
+% p = exp(-25^2*((x-x0).^2 + (y-y0).^2));
+% u = zeros(Np, K);
+% U{1} = p;
+% U{2} = u;
+% U{3} = u;
+% U{4} = u;
+% U{5} = u;
 
 
 %%
@@ -214,7 +214,7 @@ while (time<FinalTime)
         clf
         
         p = U{3}+U{4}; % trace(S)        
-%         p = U{1};
+        p = U{1};
 
         vv = Vp*p;        
 %         vv = abs(vv);

### 2017-05-09 matlab/elas/elas_convergence.m

@@ -1,25 +1,44 @@
 clc
-clear
+% clear
 
 % L2 errors
 Kvec = [2 4 8 16];
-for N = 1:6
+for N = 5;1:5
     for i = 1:length(Kvec);
         K1D = Kvec(i);
         h(i) = 1/K1D;
         disp(sprintf('======== on N = %d, K1D = %d\n',N,K1D));
         
-        harmonic_err{N}(i) = Elasticity2D(N,K1D); % harmonic sol
+        harmonic_err{N}(i) = Elasticity2D(N,K1D) % harmonic sol
         
-        rayleigh_err{N}(i) = Elasticity2D_RayleighWave(N,K1D,1); % mu = 1
+        rayleigh_err{N}(i) = Elasticity2D_RayleighWave(N,K1D) % mu = 1
         
-        lamb_err{N}(i) = Elasticity2D_Lamb(N,K1D);        
-        stonely_err{N}(i) = Elasticity2D_Stonely(N,K1D);
+        lamb_err{N}(i) = Elasticity2D_Lamb(N,K1D)        
+
+        stonely_err{N}(i) = Elasticity2D_Stonely(N,K1D)
         
-%         ref_err{N}(i) = ElasticityReferenceSolution2D(N,K1D);        
-        
-    end    
+%         ref_err{N}(i) = ElasticityReferenceSolution2D(N,K1D);                
+    end            
 end
+%%
+for N = 1:5
+%     err = harmonic_err{N};
+%     err = rayleigh_err{N};    
+%     err = lamb_err{N};
+    err = stonely_err{N};
+    
+    loglog(h,err,'o--')
+    hold on
+    r = N+1/2;
+    scale = err(1)/h(1)^(r);
+    loglog(h,scale*h.^(r),'k--')
+    C(N) = (log(err(end))-log(err(end-1))) / (log(h(end))-log(h(end-1)));
+    a = 1;
+    C2=[ones(a+1,1) log(h(end-a:end))']\log(err(end-a:end)'); C(N) = C2(2);
+    print_pgf_coordinates(h,err)    
+end
+C
+
 %%
 % clc
 

### 2017-05-11 docs/splitform/main.tex
updating splitform docs and code with curvilnear

@@
 
-\cite{feng2016discontinuous}
+Non-conservative discretizations \cite{rhebergen2008discontinuous}.
+
+DG calculus \cite{feng2016discontinuous} usually assumes exact integration.
+As pointed out in \cite{gassner2016split}, this can be either be very expensive (curvilinear meshes or polynomial nonlinearities) or impossible (in the presence of rational integrands, such as for compressible flow).
 
 Lifting operators and discrete gradients \cite{bassi1997high, di2011mathematical}
+
+Goal: reduce aliasing in advective problems caused by variable coefficients and curvilinear or non-affine mappings.
 
@@
 We will assume for now that all elements are affine mappings of a reference element $\widehat{D}$.
+Extensions to curvilinear meshes and non-affine mappings will be discussed in Section~\ref{sec:curv}.
 
@@
 \[
-\LRp{u,v}_{\Omega} = \sum_{D^k} \int_{D^k} uv \diff{x} = \sum_{D^k} \int_{\widehat{D}} uv J \diff{\widehat{x}},
+\LRp{u,v}_{\Omega} = \sum_{k} \int_{D^k} uv \diff{x} = \sum_{k} \int_{\widehat{D}} uv J \diff{\widehat{x}},
 \]
@@
 \subsection{Discrete derivatives}
-We introduce two discrete DG derivatives in this section.  The first is the implicit definition used in \cite{di2011mathematical}:
+We introduce two discrete DG derivatives in this section, based on definitions used in \cite{di2011mathematical} and \cite{hesthaven2004high,Warburton20063205}:
 \begin{definition}
-The discontinuous Galerkin differentiation operator ${D}^i: L^2\LRp{\Omega}\rightarrow V_h$ is defined implicitly as follows:
-\[
-\LRp{{D}^i u,v}_{\Oh} = \sum_{D^k} \LRp{\LRp{-u,\pd{v}{\bm{x}_i}}_{\LK} + \LRa{\avg{u},v\bm{n}_i}_{\LdK}}, \qquad \forall v \in V_h.
-\]
+We define two discontinuous Galerkin differentiation operators ${D}^i: L^2\LRp{\Omega}\rightarrow V_h$ and $\tilde{D}^i: L^2\LRp{\Omega}\rightarrow V_h$ is defined implicitly as follows:
+\begin{align}
+\LRp{{D}^i u,v}_{\Oh} &= \sum_{k} \LRp{\LRp{-u,\pd{v}{\bm{x}_i}}_{\LK} + \LRa{\avg{u},v\bm{n}_i}_{\LdK}}, \qquad \forall v \in V_h\\
+\LRp{\tilde{D}^i u,v}_{\Oh} &= \sum_{k} \LRp{\LRp{\pd{u}{\bm{x}_i},v}_{\LK} - \frac{1}{2}\LRa{\jump{u},v\bm{n}_i}_{\LdK}}, \qquad \forall v\in V_h.
+\label{eq:dgd1}
+\end{align}
 \end{definition}
 
-We note that, for $u,v \in V_h$, (discrete) integration by parts yields an equivalent definition of $D^i$ found in \cite{hesthaven2004high,Warburton20063205}
-\[
-\LRp{{D}^i u,v}_{\Oh} = \sum_{D^k} \LRp{\LRp{\pd{u}{\bm{x}_i},v}_{\LK} - \frac{1}{2}\LRa{\jump{u},v\bm{n}_i}_{\LdK}}, \qquad \forall v\in V_h.
-\]
+We note that, assuming $u,v \in V_h$ and sufficiently accurate quadrature, $D^i$ and $\tilde{D}^i$ are equivalent.
 The main property of the discrete derivatives utilized in this work is a discrete integration-by-parts formula.
@@
 \begin{align*}
-\LRp{D^i u,v} &= \LRp{-u,D^i_h v} + \int_{\partial \Omega} uv \bm{n}_i\\
-\LRp{D_h^i u,v} &= \LRp{-u,D^i_h v} + \int_{\partial \Omega} uv \bm{n}_i.
+\LRp{D^i u,v} &= \LRp{-u,D^i_h v} + \int_{\partial \Omega} uv \bm{n}_i.
+%\LRp{D_h^i u,v} &= \LRp{-u,D^i_h v} + \int_{\partial \Omega} uv \bm{n}_i.
 \end{align*}
@@
 \begin{proof}
-Ez
+Assuming that quadrature is sufficiently accurate for $u, v\in V_h$, integration by parts gives the equivalence. 
 \end{proof}
@@
 \begin{definition}
-Let $u \in V_h$ and $\bm{u}\in \LRp{V_h}^d$.  Then, discrete gradient and divergence operators are defined as follows:
+Let $u \in V_h$ and $\bm{u}\in \LRp{V_h}^d$.  Then, discrete gradient and divergence operators are defined as
 \begin{align*}
@@
 \end{definition}
-
-\subsection{Examples}
-
-\begin{itemize}
-\item Tensor product elements: inner product computed using GLL nodes, $\Pi_N$ reduces to nodal interpolation
-\item General elements: $\Pi_N$ is just quadrature-based projection.  For accuracy, degree $2N$ or higher?  
-\item General elements: $\Pi_N$ can be taken to be nodal interpolation.  
-\item Pyramids?  
-\end{itemize}
-
-\section{Notes}
-Begin with the gradient.  Multiplication by $\bm{v}$ and integration by parts gives
-\[
-\sum_{D^k} \LRp{\Grad u, \bm{v}}_{\LK} = \sum_{D^k} \LRp{-u, \Grad \cdot \bm{v}}_{\LK} + \LRa{u,\bm{v}\cdot\bm{n}}_{\LdK}.
-\]
-We replace the values of ${u}$ with $\avg{u}$ on each element boundary to define the global DG gradient operator $\Grad_h$ 
-\[
-\sum_{D^k} \LRp{\Grad_h u, \bm{v}}_{\LK} \coloneqq \sum_{D^k} \LRp{-u, \Grad \cdot \bm{v}}_{\LK} + \LRa{\avg{u},\bm{v}\cdot\bm{n}}_{\LdK}.
-\]
-Integrating by parts again and the introduction of the lift operator shows that
-\[
-\Grad_h u = \Grad u + \frac{1}{2}L\LRp{\jump{u}\bm{n}}
-\]
-where $L$ is the lift operator.
-
-The DG divergence operator is similarly defined as
-\[
-\sum_{D^k} \LRp{\Grad_h \cdot \bm{u}, v}_{\LK} \coloneqq \sum_{D^k} \LRp{-\bm{u}, \Grad v}_{\LK} + \LRa{\avg{\bm{u}}\cdot\bm{n},v}_{\LdK}.
-\]
-and
-\[
-\Grad_h \cdot u = \Grad\cdot u + \frac{1}{2}L\LRp{\jump{\bm{u}}\cdot\bm{n}}
-\]
-and it can be shown that 
-\[
-\LRp{\Grad_h u, \bm{v}} = \LRp{-u, \Grad_h \cdot \bm{v}}.
-\]
-We will incorporate boundary conditions in a stable way in the following sections.
 
@@
 \]
+
+\section{Examples}
+
+\begin{itemize}
+\item Tensor product elements: inner product computed using GLL nodes, $\Pi_N$ reduces to nodal interpolation
+\item General elements: $\Pi_N$ is just quadrature-based projection.  For accuracy, degree $2N$ or higher?  
+\item Triangles and tetrahedra: lumped quadrature nodes.
+Implies that mass lumping on triangles requires sufficiently accurate quadrature such that IBP holds.
+This conclusion was first reached in the study of multi-dimensional SBP operators \cite{hicken2016multidimensional}.
+\item General elements: $\Pi_N$ can be taken to be nodal interpolation.  
+\item Pyramids: need to account for non-affine mappings.  
+\end{itemize}
+
+\subsection{Variable advection}
+
+A split formulation for advection is 
+\[
+\LRp{\pd{u}{t},v} + \frac{1}{2}\LRp{\Grad_h\cdot \Pi_N \LRp{ \bm{\beta}u},v} + \frac{1}{2}\LRp{\bm{\beta}\cdot\Grad_h u,v} + \frac{1}{2}\LRp{\LRp{\Grad\cdot \bm{\beta}} u,v} = 0.
+\]
+Taking $v = u$ yields and using $\LRp{\Grad_h u, \bm{v}} = \LRp{-u, \Grad_h \cdot \bm{v}}$ yields the energy statement
+\[
+\frac{1}{2}\nor{u}^2 + \frac{1}{2}\LRp{\Grad_h\cdot \Pi_N \LRp{ \bm{\beta}u},u} - \frac{1}{2}\LRp{ u, \Grad_h\cdot \Pi_N\LRp{\bm{\beta}u}} = \frac{1}{2}\LRp{-\LRp{\Grad \cdot \bm{\beta}} u,u},
+\]
+implying that $\frac{1}{2}\nor{u}^2 = 0$ if $\Grad\cdot \bm{\beta} = 0$, or that the method is energy conserving.
+The only difference in this formulation is the introduction of $\Pi_N$, which can be defined at a discrete level using any quadrature scheme for which a discrete projection is well-defined.
+
+Penalization can be added by adding any positive-definite stabilization term (upwind, penalty, Lax-Friedrichs) through the regular divergence flux.
+
+\section{Discrete DG derivatives}
+
+Methods based on discrete DG derivatives also work.
+
+The discrete DG derivative-based method is not consistent in the sense that Galerkin orthogonality does not hold exactly.  The difference lies in the flux terms.  Assume $\Div \bm{\beta} = 0$, then
+\[
+\LRp{\pd{u}{t},v} + \frac{1}{2}\LRp{-{ \bm{\beta}u},\Grad_h v} + \frac{1}{2}\LRp{\bm{\beta}\cdot\Grad_h u,v} = 0.
+\]
+\begin{align*}
+\LRp{\Grad_h\cdot \Pi_N\LRp{ \bm{\beta}u},v} &= \sum_{k} \LRp{-\bm{\beta}u, \Grad v}_{\LK} + \LRa{\avg{\Pi_N\LRp{\bm{\beta} u}}\cdot \bm{n},v}_{\LdK}\\
+\LRp{\Grad_h u,\bm{\beta}v} &= \sum_{k} \LRp{-u, \Grad \cdot \LRp{\bm{\beta}v}}_{\LK} + \LRa{\bm{\beta}\cdot\bm{n}\avg{u},v}_{\LdK}.
+\end{align*}
+The latter term is consistent; the former is not due to the presence of $\avg{\Pi_N\LRp{\bm{\beta} u}}\cdot \bm{n}$ in the flux term.
+The consistency error should then be $O(h^{N+1/2})$ using a trace inequality for $L^2$ projections.
+
+Note: can also use interpolants in a stable manner if using $D_h$.
+Unlike SEM, this still requires an extra matvec per RHS evaluation because of the lack of diagonality of the mass matrix.
+Reduces number of steps by one (no interpolation to quadrature points) but does not reduce number of total matvecs.
 
@@
 
-Penalization can be added by adding any positive-definite stabilization term (upwind, penalty, Lax-Friedrichs) through the regular divergence flux.
+\section{Non-affine mappings}
+\label{sec:curv}
 
-It's probably better to formulate this using continuous DG derivatives, recover flux terms, then discretize that - the flux terms \textit{should} still cancel out after discretization, right?
+For general elements and curvilinear meshes, $J$ and $\bm{G}$ are no longer constant over each element.
+However, for isoparametric mappings (and appropriate polynomial interpolations of general curvilinear mappings), metric identities hold \cite{kopriva2006metric} such that $\widehat{\Grad} \cdot \LRp{J\bm{G}^T} = 0$.
+Thus, we can write
+\[
+\LRp{\Grad u,\bm{v}}_{D^k} = \LRp{J\bm{G}\widehat{\Grad} u,\bm{v}}_{\widehat{D}} = \LRp{-u,\widehat{\Grad} \cdot \LRp{J\bm{G}\bm{v}}}  + \LRa{J\bm{G}^T\widehat{\bm{n}} u,\bm{v}}.   
+\]
+Noting that $J\bm{G}\widehat{\bm{n}} = J^f \bm{n} $ \cite{hesthaven2007nodal}, we recover integration by parts over the physical element.
 
-\section{Variable advection}
+Thus, to extend the discrete discontinuous Galerkin gradient and divergence to non-affine mappings, we simply define
+\begin{align*}
+\LRp{\widehat{\Grad}_h u,\bm{v}}_{\Omega} &= \sum_{k} \LRp{\widehat{\Grad}u,\bm{v}}_{\widehat{D}}  - \frac{1}{2}\LRa{\widehat{\bm{n}} \jump{u},\bm{v}} \\
+\LRp{\widehat{\Grad}_h \cdot\bm{u},{v}}_{\Omega} &= \sum_{k} \LRp{\widehat{\Grad}\cdot \bm{u},{v}}_{\widehat{D}}  - \frac{1}{2}\LRa{\jump{\bm{u}},v\widehat{\bm{n}}} 
+%\LRp{\Grad_h \cdot \bm{u},v}_{\Omega} &= \sum_{k} \LRp{\widehat{\Grad}\cdot{\Pi_N\LRp{J\bm{G}\bm{u}}},v}  - \frac{1}{2}\LRa{\widehat{\bm{n}} \jump{J\bm{G}\bm{u}},v} \\
+%\LRp{\Grad_h u,\bm{v}}_{\Omega} &= \sum_{k} \LRp{\Pi_N\LRp{J\bm{G}\widehat{\Grad} u},\bm{v}}  - \frac{1}{2}\LRa{\widehat{\bm{n}} \jump{J\bm{G}u},v}.
+\end{align*}
+and define the discrete physical gradient and divergence as
+\begin{align*}
+\LRp{\Grad_h u,v}_{\Omega} &= \LRp{J\bm{G}\widehat{\Grad}_h u,\bm{v}}_{\Omega} = \sum_{k}\LRp{J\bm{G}\widehat{\Grad}u,\bm{v}}_{\widehat{D}}  - \frac{1}{2}\LRa{J\bm{G}\widehat{\bm{n}} \jump{u},\bm{v}} \\
+\LRp{\Grad_h \cdot \bm{u},v}_{\Omega} &= \LRp{\widehat{\Grad}_h\cdot\LRp{J\bm{G}^T\bm{u}}}_{\Omega} = \sum_{k} \LRp{\widehat{\Grad}\cdot \LRp{J\bm{G}^T\bm{u}},{v}}_{\widehat{D}}  - \frac{1}{2}\LRa{\jump{J\bm{G}^T\bm{u}},v\widehat{\bm{n}}}.
+\end{align*}
+The discrete gradient is consistent with the physical gradient.
+The discrete divergence is also consistent with the physical divergence, provided that the jump term $\LRa{\jump{J\bm{G}^T\bm{u}},v\widehat{\bm{n}}}$ is computed via
+\[
+\LRa{\jump{J\bm{G}^T\bm{u}},v\widehat{\bm{n}}} = \LRa{\jump{J\LRp{\bm{G}\widehat{\bm{n}}}^T\bm{u}},v} = \LRa{\LRp{\widehat{\bm{n}}^T\LRp{J{\bm{G}}^T\bm{u}}}^+ + \LRp{\widehat{\bm{n}}^T\LRp{J{\bm{G}}^T\bm{u}}}^-,v}.
+\]
 
-A split formulation for advection is 
+The advantage of this projected geometric factor approach is local conservation; taking $v = \bbm{1}_{D^k}$ recovers (after integration by parts) the flux integral
 \[
-\LRp{\pd{u}{t},v} + \frac{1}{2}\LRp{\Grad_h\cdot \Pi_N \LRp{ \bm{\beta}u},v} + \frac{1}{2}\LRp{\bm{\beta}\cdot\Grad_h u,v} + \frac{1}{2}\LRp{\LRp{\Grad\cdot \bm{\beta}} u,v} = 0.
+\LRp{\Grad_h \cdot \bm{u},\bbm{1}_{D^k}}_{\Omega} = \int_{\partial D^k}\widehat{\bm{n}}^T\avg{J\bm{G}^T\bm{u}}.
 \]
-Taking $v = u$ yields and using $\LRp{\Grad_h u, \bm{v}} = \LRp{-u, \Grad_h \cdot \bm{v}}$ yields the energy statement
-\[
-\frac{1}{2}\nor{u}^2 + \frac{1}{2}\LRp{\Grad_h\cdot \Pi_N \LRp{ \bm{\beta}u},u} - \frac{1}{2}\LRp{ u, \Grad_h\cdot \Pi_N\LRp{\bm{\beta}u}} = \frac{1}{2}\LRp{-\LRp{\Grad \cdot \bm{\beta}} u,u},
-\]
-implying that $\frac{1}{2}\nor{u}^2 = 0$ if $\Grad\cdot \bm{\beta} = 0$, or that the method is energy conserving.
-The only difference in this formulation is the introduction of $\Pi_N$, which can be defined at a discrete level using any quadrature scheme for which a discrete projection is well-defined.
-
-\section{Discrete DG derivatives}
-
-Methods based on discrete DG derivatives also work.
-
-The discrete DG derivative-based method is not consistent in the sense that Galerkin orthogonality does not hold exactly.  The difference lies in the flux terms.  Assume $\Div \bm{\beta} = 0$, then
-\[
-\LRp{\pd{u}{t},v} + \frac{1}{2}\LRp{-{ \bm{\beta}u},\Grad_h v} + \frac{1}{2}\LRp{\bm{\beta}\cdot\Grad_h u,v} = 0.
-\]
-\begin{align*}
-\LRp{\Grad_h\cdot \Pi_N\LRp{ \bm{\beta}u},v} &= \sum_{D^k} \LRp{-\bm{\beta}u, \Grad v}_{\LK} + \LRa{\avg{\Pi_N\LRp{\bm{\beta} u}}\cdot \bm{n},v}_{\LdK}\\
-\LRp{\Grad_h u,\bm{\beta}v} &= \sum_{D^k} \LRp{-u, \Grad \cdot \LRp{\bm{\beta}v}}_{\LK} + \LRa{\bm{\beta}\cdot\bm{n}\avg{u},v}_{\LdK}.
-\end{align*}
-The latter term is consistent; the former is not due to the presence of $\avg{\Pi_N\LRp{\bm{\beta} u}}\cdot \bm{n}$ in the flux term.
-The consistency error should then be $O(h^{N+1/2})$ using a trace inequality for $L^2$ projections.
-
-Note: can also use interpolants in a stable manner if using $D_h$.
-Unlike SEM, this still requires an extra matvec per RHS evaluation because of the lack of diagonality of the mass matrix.
-Reduces number of steps by one (no interpolation to quadrature points) but does not reduce number of total matvecs.
-
-\section{Curvilinear coordinates}
-
-For curvilinear meshes, $J$ and $\bm{G}$ are no longer constant over each element.
-However, for isoparametric mappings (and special polynomial interpolations of general curvilinear mappings), metric identities hold \cite{kopriva2006metric} such that $\widehat{\Grad} \cdot \LRp{J\bm{G}} = 0$.
-\[
-\LRp{\Grad u,\bm{v}}_{D^k} = \LRp{J\bm{G}\widehat{\Grad} u,\bm{v}}_{\widehat{D}} = \LRp{-u,\widehat{\Grad} \cdot \LRp{J\bm{G}\bm{v}}}  + \LRa{J\bm{G}\widehat{\bm{n}} u,\bm{v}} 
-\]
-Noting that $J\bm{G}\widehat{\bm{n}} = \bm{n} J^f$ \cite{hesthaven2007nodal}, we recover integration by parts over the physical element.
 

### 2017-05-16 docs/splitform/main.tex
updating splitform

@@
 
-Lifting operators and discrete gradients \cite{bassi1997high, di2011mathematical}
+Lifting operators and discrete gradients \cite{bassi1997high, di2011mathematical}.
 
 Goal: reduce aliasing in advective problems caused by variable coefficients and curvilinear or non-affine mappings.
+Resembles ideas from summation by parts \cite{olsson1995summation1, olsson1995summation2, hicken2016multidimensional}, but based on global DG differentiation operators.
 

### 2017-05-16 matlab/cfd/checkMetricIdentities.m

@@ -8,6 +8,7 @@ K1D = 8;
 
 % Initialize solver and construct grid and metric
 StartUp2D; 
+
 vmapM = reshape(vmapM,Nfp*Nfaces,K);
 vmapP = reshape(vmapP,Nfp*Nfaces,K);
 mapM = reshape(mapM,Nfp*Nfaces,K);
@@ -73,6 +74,7 @@ ns = repmat(ns,1,K);
 
 norm(nxJ - (rxJ.*nr + sxJ.*ns),'fro')
 norm(nyJ - (ryJ.*nr + syJ.*ns),'fro')
+% return
 
 % quadrature arrays
 rxJ = Vq*(rx.*J); sxJ = Vq*(sx.*J);

### 2017-05-23 docs/splitform/main.tex
updating IGA

@@
 \]
-through integration by parts and an assumption that $\bm{\beta}\cdot \bm{n}$ is periodic.  Cancelling volume terms, we end up with the statement of local conservation 
+through integration by parts and an assumption that $\bm{\beta}\cdot \bm{n}$ is periodic.  Cancelling volume terms involving $\Grad\cdot \beta$, we end up with the statement of local conservation 
 \[
@@
 
+\subsection{Specific discrete cases}
+
+Spectral element methods and mass-lumpable tets: $\Pi_N = I_N$, and the lift operator is just the $\bm{B}$ boundary operator, which you can commute terms such that the scaling of the lift is the same as the lift of the scaling.
+
+CPR GL nodes: conservation for Burgers', but is this just because of the special nature of the flux function cancelling out other terms?
+Skew-symmetric variable advection doesn't seem to be conservative.
+
 \section{Extension to other hyperbolic problems}
@@
 This impacts only boundary conditions.)
+
+\section{Numerical experiments}
+
+\begin{itemize}
+\item Check ``true'' conservation of discrete DG differentiation formulations - does it converge at $h^{N+2}$ or greater?
+\item Copy Ranocha's experiments.  
+\item Convergence of 2D advection and (pre-shock) Burgers equation on curvilinear grids: conservative and Lax-Friedrichs fluxes.  
+\item Stability (energy/entropy evolution) of 2D advection and Burgers equation on curvilinear grids: conservative and Lax-Friedrichs fluxes.  
+\end{itemize}
 

### 2017-05-23 matlab/Codes2D/GeometricFactors2D.m

@@ -4,6 +4,7 @@ function [rx,sx,ry,sy,J] = GeometricFactors2D(x,y,Dr,Ds)
 % Purpose  : Compute the metric elements for the local mappings of the elements
 
 % Calculate geometric factors
-xr = Dr*x; xs = Ds*x; yr = Dr*y; ys = Ds*y; J = -xs.*yr + xr.*ys;
+xr = Dr*x; xs = Ds*x; yr = Dr*y; ys = Ds*y; 
+J = -xs.*yr + xr.*ys;
 rx = ys./J; sx =-yr./J; ry =-xs./J; sy = xr./J;
 return;

### 2017-05-23 matlab/IGA_DG/Elbow3Dgeo.m

@@ -1,7 +1,7 @@
 function [x y z geofacs] = Elbow3Dgeo(r,s,t)
 % function Elbow3Dgeo(r,s,t)
 
-% h = 1; w = 1; r = .5; R = 1; t = r/2;
+% h = 1; w = 1; rad = .5; R = 1; T = rad/2;
 h = 2; w = 2; rad = 0.5; R = 2; T = 0.2; 
 
 [N,N_2,N_3,n_1,n_2,n_3,Xi_1,Xi_2,Xi_3,P,W] = NURBS_Pipe_Elbow_Volume(h,w,rad,R,T);

### 2017-05-23 matlab/IGA_DG/Elbow3Dtest.m

@@ -77,9 +77,6 @@ w = V*cw;
 x = (V*(cx.*cw))./w; 
 y = (V*(cy.*cw))./w;
 z = (V*(cz.*cw))./w;
-% x = ((V*cx).*(V*cw))./w; 
-% y = ((V*cy).*(V*cw))./w; 
-% z = ((V*cz).*(V*cw))./w; 
 
 wr = Vr*cw; 
 ws = Vs*cw; 

### 2017-05-23 matlab/IGA_DG/tiConnect3DHex.m

@@ -8,14 +8,32 @@ K = size(EToV,1);
 Nnodes = max(max(EToV));
 
 % create list of all faces
-f1 = [4,1,8,5];
-f2 = [2,3,6,7];
 
-f3 = [1,2,5,6];
-f4 = [3,4,7,8];
+% f5 = [1 5 7 3];
+% f6 = [2 6 8 4];
+% f1 = [1 2 4 3];
+% f2 = [5 6 8 7];
+% f3 = [1 2 6 5];
+% f4 = [3 4 8 7];
 
-f5 = [1,2,4,3];
-f6 = [5,6,8,7];
+f1 = [1     2     5     6];
+f2 = [3     4     7     8];
+f3 = [1     3     5     7];
+f4 = [2     4     6     8];
+f5 = [1     2     3     4];
+f6 = [5     6     7     8];
+
+% r = -/+1
+% f1 = [4,1,8,5];
+% f2 = [2,3,6,7];
+% 
+% % s = -/+1
+% f3 = [1,2,5,6];
+% f4 = [3,4,7,8];
+% 
+% % t = -/+1
+% f5 = [1,2,4,3];
+% f6 = [5,6,8,7];
 
 fnodes = [EToV(:,f1);EToV(:,f2);EToV(:,f3);...
     EToV(:,f4);EToV(:,f5);EToV(:,f6)];

### 2017-05-23 matlab/hybrid/hybridgStartUp.m

@@ -188,6 +188,7 @@ for e = 1:K
         % normalize so we can scale by sJ after
         wf(fids{f}) = wf(fids{f})/sum(wf(fids{f}));
         
+        nc
     end
     
     sJ = sqrt(nx(:,e).^2 + ny(:,e).^2 + nz(:,e).^2);

### 2017-05-25 docs/splitform/main.tex
fixing pipe elbow experiments, adding 3D IGA-IPDG driver

@@
 
+One downside of this approach is that stabilization must be applied separately from the discrete gradient.
+We consider stabilization in the form of penalty fluxes.
+For jump terms without normals, such as $\jump{p}$, these can be stabilized with no additional cost.
+For terms involving normals, we often have $\jump{\bm{u}\cdot\bm{n}}\bm{n}$.
+Since normals are rational functions for high order curvilinear mappings, these require quadrature to compute, and can be expensive.
+An alternative approach would just be to use averaged normals instead, and combine them into the lifting within the reference element.
+
 \subsection{Specific discrete cases}
@@
 
-Example: Entropy splitting of Euler (note - cannot extend to Navier-Stokes in an entropy-stable fashion due to fact that heat flux matrix is not symmetrizable w.r.t.\ homogeneous flux function, though viscous terms are.
-This impacts only boundary conditions.)
+Example: Entropy splitting of Euler and Navier-Stokes - can do ``formally'', but the fluxes are singular and non-integrable.
 

### 2017-05-25 matlab/IGA_DG/IPDG_IGA1D.m

@@ -54,7 +54,8 @@ xp = Vp*x;
 
 %% set init condition
 
-uex = @(x,t) cos(3*pi*x/2).*cos(3*pi*t/2);
+k = 3;
+uex = @(x,t) cos(k*pi*x/2).*cos(k*pi*t/2);
 % uex = @(x,t) exp(-5^2*x.^2);
 u = uex(x,0);
 

### 2017-05-25 matlab/IGA_DG/Wave3D_IPDG_IGA.m

@@ -1,427 +1,0 @@
-function Wave3D_IGA
-
-Globals3D
-
-NB = 1;
-Ksub = 2;
-K1D = 1;
-smoothKnots = 0;
-useQuadrature = 1;
-
-if Ksub==1
-    CN = (NB+1)^2/2;
-    dt = 2/CN; % size of domain
-else
-    CN = (NB+2)*Ksub;
-    dt = 2/CN;
-end
-
-N = NB+Ksub-1;
-
-Np = (N+1)^3;
-dofs = Np*K1D^3;
-
-FinalTime = .53;
-
-% return
-
-%% mesh stuff
-
-global rxJ sxJ txJ ryJ syJ tyJ rzJ szJ tzJ
-global nx ny nz Jq sJ
-global DPq1D DVq1D Vq1D Pq1D D1D D1Dt LIFT Fmask M1D
-
-if Ksub==1
-    [r1D w1D] = JacobiGL(0,0,N);
-else
-    VX = linspace(-1,1,Ksub+1);
-    t = [VX(1)*ones(1,NB) VX VX(end)*ones(1,NB)]; % open knot vec
-    for i = 1:N+1
-        r1D(i,1) = mean(t((i+1):(i+NB))); % greville
-    end
-%     plot(r1D,r1D*0,'o')
-end
-
-[r s t] = meshgrid(r1D);
-r = r(:); s = s(:); t = t(:);
-
-Nfaces = 6;
-Fmask1 = find(abs(r+1)<1e-8);
-Fmask2 = find(abs(r-1)<1e-8);
-Fmask3 = find(abs(s+1)<1e-8);
-Fmask4 = find(abs(s-1)<1e-8);
-Fmask5 = find(abs(t+1)<1e-8);
-Fmask6 = find(abs(t-1)<1e-8);
-Fmask = [Fmask1(:) Fmask2(:) Fmask3(:) Fmask4(:) Fmask5(:) Fmask6(:)];
-
-rp1D = linspace(-1,1,100);
-[rp sp tp] = meshgrid(rp1D);
-rp = rp(:); sp = sp(:); tp = tp(:);
-
-% initialize TP operators
-[BVDM M1D D1D R rq1D wq1D Vq1D DVq1D] = bsplineVDM(NB,Ksub,r1D,smoothKnots); % VDM for interp, mass, M\S
-Pq1D = M1D\(Vq1D'*diag(wq1D));
[…]

### 2017-05-25 matlab/IGA_DG/advec1D_dispersion.m

@@ -1,7 +1,7 @@
 % function advec1D_dispersion
 clear
-NB = 3;
-Ksub = 32;
+NB = 2;
+Ksub = 2;
 K1D = 1;
 N = NB+Ksub-1;
 smoothKnots = 0;

### 2017-07-13 docs/splitform/main.tex
updating with new paper

--- 2017-05-16 docs/splitform/main.tex
@@
 \newtheorem*{remark}{Remark}
+\theoremstyle{theorem}
+\newtheorem{theorem}{Theorem}
 %\usepackage{thmtools}
@@
 \author{Jesse Chan}
-\title{Energy stable discontinuous Galerkin formulations using discrete differential operators}
+\title{On discrete entropy conservation for general high order discontinuous Galerkin methods}
 
@@
 \begin{abstract}
-We show that, for non-linear hyperbolic conservation laws which admit a skew-symmetric splitting, energy stable discontinuous Galerkin (DG) methods can be constructed in a straightforward manner based on discrete DG gradient and divergence operators.
-In particular, we show that these formulations remain energy stable for both curvilinear geometries and inexact quadrature.
-Examples of energy stable formulations are given for variable advection, Burgers' equation, and a Burgers'-like system, and it is shown that the construction of DG methods based on discrete differential operators recover known entropy-conservative fluxes.
+High order methods based on diagonal-norm summation by parts operators can be shown to semi-discretely conserve entropy for nonlinear systems of hyperbolic PDEs \cite{fisher2013high,carpenter2014entropy}.
+These include discontinuous Galerkin spectral element methods (DG-SEM) \cite{gassner2016split,gassner2016well,wintermeyer2017entropy, chen2017entropy} based on mass-lumping.
+In this work, we describe how to extend discretely entropy conservative schemes for a more general class of DG methods using quadrature-based projections and global differentiation operators.
+The construct of these methods is independent of the choice of basis and accuracy of quadrature rule used, and includes DG methods involving generalized SBP operators without boundary nodes and certain dense-norm SBP operators.
+Numerical experiments confirm the stability and high order accuracy of the proposed methods for the Burgers, shallow water, and compressible Euler equations.
 \end{abstract}
+
+\tableofcontents
 
@@
 
-Non-conservative discretizations \cite{rhebergen2008discontinuous}.
+\note{High order methods are good boilerplate}
 
-DG calculus \cite{feng2016discontinuous} usually assumes exact integration.
-As pointed out in \cite{gassner2016split}, this can be either be very expensive (curvilinear meshes or polynomial nonlinearities) or impossible (in the presence of rational integrands, such as for compressible flow).
+\note{High order methods are not stable for nonlinear problems.  Stability of discretizations is a hot topic: artificial viscosity, polynomial de-aliasing,  \cite{karniadakis2013spectral}}
 
-Lifting operators and discrete gradients \cite{bassi1997high, di2011mathematical}.
+\cite{tadmor1987numerical, tadmor2003entropy}
 
-Goal: reduce aliasing in advective problems caused by variable coefficients and curvilinear or non-affine mappings.
-Resembles ideas from summation by parts \cite{olsson1995summation1, olsson1995summation2, hicken2016multidimensional}, but based on global DG differentiation operators.
+Hughes entropy variables \cite{hughes1986new}
 
-\section{Discrete differential operators}
+\cite{hicken2016multidimensional}
 
-We assume that the domain $\Omega$ is decomposed into non-overlapping elements $D^k$.
-We define the mesh $\Omega_h = \cup D^k$ and the corresponding global approximation space $V_h(\Omega_h) = \bigoplus V_h\LRp{D^k}$, where $V_h\LRp{D^k}$ is the approximation space over a single patch.
-Furthermore, we introduce the jump and average of discontinuous functions across element interfaces.
-Let $D^{k,+}$ denote the neighboring element of across a face $f$ of $D^k$, and let $u^+,u^-$ denote the values of $u$ on $D^{k,+}$ and $D^k$, respectively.
-The jump of $u$ across $f$ is then defined as
+\cite{gassner2016split,gassner2016well,wintermeyer2017entropy}
+
+\cite{chen2017entropy}
+
+While DG-SEM based on flux differencing is entropy conservative, this is tied to a specific choice of basis and quadrature.
+Furthermore, GLL quadratures and diagonal-norm SBP operators do not naturally generalize to triangles or tetrahedra \cite{helenbrook2009existence}.
+Diagonal-norm SBP operators can be constructed for triangles and tetrahedra \cite{chin1999higher, cohen2001higher, hicken2016multidimensional, chen2017entropy}; however, the number of nodal points for such operators is greater than the dimension of the natural polynomial approximation space.
+Furthermore, appropriate point sets have only been constructed for $N \leq 4$ in three dimensions \cite{zhebel2014comparison}.
+
+\section{Entropy conservative collocation DG methods}
+
+\subsection{Entropy stability for systems of hyperbolic PDEs}
+
+We consider systems of nonlinear conservation laws in $\mathbb{R}^d$ with $n$ variables
+\begin{align}
+\pd{\bm{u}}{t} + \sum_{k=1}^d \pd{\bm{f_k(\bm{u})}}{\bm{x}_k} &= 0, \qquad \bm{u}(\bm{x},t) = (u_1(\bm{x},t),\ldots,u_n(\bm{x},t)).
+\label{eq:pde}
+\end{align}
+where the fluxes $f_k(\bm{u})$ are smooth functions of the vector of conservative variables $\bm{u}(\bm{x},t)$.
+The Jacobian matrix $\bm{A}_k(\bm{u})$ is defined entrywise as
 \[
-\jump{u} = u^+ - u^-, \qquad \avg{u} = \frac{u^+ + u^-}{2}.
+\LRp{\bm{A}_k(\bm{u})}_{ij} = \LRp{\pd{\bm{f}_k(\bm{u})}{\bm{u}_j}}_i.
 \]
-On faces which coincide with the boundary $\partial \Omega$, the average and jump are defined as
+We are interested in systems for which there exists a convex entropy function $U(\bm{u})$ such that  
+\begin{equation}
+U''(\bm{u})\bm{A}_k(\bm{u}) = \LRp{U''(\bm{u}) \bm{A}_k(\bm{u})}^T.
+\label{eq:entropysym}
+\end{equation}
+For systems with convex entropy functions, one can define entropy variables $\bm{v} = U'(\bm{u})$.
+The convexity of the $U(\bm{u})$ guarantees that the mapping between conservative and entropy variables is invertible.
+
+It can be shown (see, for example, \cite{mock1980systems}) that (\ref{eq:entropysym}) is equivalent to the existence of an entropy flux functions $F_k(\bm{u})$ such that
 \[
-\jump{u} = 0, \qquad \avg{u} = u.
+\bm{v}^T \pd{\bm{f}_k}{\bm{u}} = \pd{F_k(\bm{u})}{\bm{u}}^T.
 \]
-The jump and average of vector fields are defined component-wise using the jumps and averages of components.
-
-We will assume for now that all elements are affine mappings of a reference element $\widehat{D}$.
-Extensions to curvilinear meshes and non-affine mappings will be discussed in Section~\ref{sec:curv}.
-
-\subsection{Discrete inner products}
-
-In order to define discrete differential operators, we first introduce the $L^2$ inner product on $V_h\LRp{\Oh}$
+The flux function and entropy flux function are further related through the so-called entropy potential $\psi_k$ 
 \[
-\LRp{u,v}_{\Omega} = \sum_{k} \int_{D^k} uv \diff{x} = \sum_{k} \int_{\widehat{D}} uv J \diff{\widehat{x}},
-\]
-and the associated $L^2$ projection $\Pi_N: L^2\LRp{\Omega}\rightarrow P^N$
-\[
-\LRp{\Pi_N u,v}_{\Omega} = \LRp{u,v}_{\Omega}.  
+\psi_k(\bm{v}) = \bm{v}^T\bm{f}_k(\bm{u}(\bm{v})) - F_k(\bm{u}(\bm{v})), \qquad \psi_k'(\bm{v}) = \bm{f}_k(\bm{u}(\bm{v})).
 \]
 
-In practice, these integrals are computed using quadrature, such that 
+When $\bm{u}$ is smooth, multiplying (\ref{eq:pde}) on the left by $\bm{v}^T = U'(\bm{u})^T$, applying the definition of the entropy flux and using the chain rule yields the conservation of entropy
+\[
+\pd{U(\bm{u})}{t} + \sum_{k=1}^d \pd{F_k(\bm{u})}{\bm{x}_k} = 0.
+\]
+More generally, it can be shown that physically relevant solutions of (\ref{eq:pde}) (defined as the limiting solution for an appropriately defined vanishing viscosity) satisfy the entropy inequality
+\[
+\pd{U(\bm{u})}{t} + \sum_{k=1}^d \pd{F_k(\bm{u})}{\bm{x}_k} \leq 0.  
+\]
+%Integration over $\mathbb{R}^d$ then shows that the entropy of the solution decreases in time.  
+When $d = 1$, integrating the entropy inequality over $[-1,1]$ and using the definition of the entropy potential yields
 \begin{equation}
-\int_{\widehat{D}}uv = \sum_{i=1}^{N_q} u(\bm{x}_i)v(\bm{x}_i) w_i,
-\label{eq:quad}
+\int_{-1}^1 \pd{U(\bm{u})}{t} + \left.\LRp{\bm{v}^T\bm{f}_k(\bm{u}(\bm{v}))-\psi(\bm{v})}\right|_{-1}^1 \leq 0.  
+\label{eq:consentropy}
 \end{equation}
-where $N_q$ is the number of quadrature points.  The only assumptions we make upon this quadrature is that it is sufficiently accurate such that 
+
+%\note{Wave equation example}
+
+\subsection{Entropy conservative collocation methods based on SBP operators}
+
+Let $\widehat{x}_i \in [-1,1]$ denote the $(N+1)$ point Gauss-Lobatto-Legendre (GLL) quadrature rule with positive weights $w_1,\ldots,w_{N+1}$.
+Let $\ell_i(x)$ denote the nodal Lagrange basis defined at GLL points, and define the differentiation matrix $\bm{D}$ be defined as
+\[
+\bm{D}_{ij} = \pd{\ell_j(x_i)}{x}.
+\]
+The matrix $\bm{D}$ maps from nodal values of a polynomial to nodal values of its derivative.  We also define the mass matrix $\bm{M}$
+\[
+\bm{M}_{ij} = \int_{-1}^1 \ell_i(x)\ell_j(x)\diff{x} \approx \delta_{ij} w_i,
+\]
+where the approximation results evaluating the integral using GLL quadrature.
+For this choice of quadrature, the Kronecker property $\ell_j(x_i) = \delta_{ij}$ gives that $\bm{M}$ is a diagonal matrix.
+
+The mass and derivative matrices may be used to define the diagonal norm SBP operator $\bm{S} = \bm{M}\bm{D}$, with the summation-by-parts property 
+\[
+\bm{S} = \bm{B} - \bm{S}^T, \qquad \bm{B}_{ij} = \begin{cases}
+1 &i = j = 1 \\
+1 &i = j = N+1 \\
+0 &\text{otherwise}
+\end{cases}.
+\]
+This property is simply a restatement of integration by parts, using that the derivative of $\ell_j(x) \in P^{N-1}$ and that GLL quadrature is exact for polynomials of degree $2N-1$.
+
+%In order to simplify notation for flux differencing, we define the following bilinear matrix product 
+%\[
+%\LRp{\bm{A} \odot \bm{f}_S(\bm{x},\bm{y})}_i = \sum_j \bm{A}_{ij} \bm{f}_S(\bm{x}_j,\bm{y}_i).  
+%\]
+We seek a nodal approximation $\bm{u}_i \approx \bm{u}(x_i)$ to the solution $\bm{u}(x)$ of (\ref{eq:pde}).  A typical nodal collocation method in conservation form is given as follows 
+\[
+\pd{\bm{u}_i}{t} + \sum_{j=1}^{N+1}\LRp{\LRp{\bm{D}\otimes \bm{I}}_{ij}\bm{f}(\bm{u}_j) +\LRp{\bm{M}^{-1}\bm{B}}_{ij}(\bm{f}_i^* - \bm{f}(\bm{u}_i))} = 0.
+\]
+Here, $\bm{f}^*$ is a numerical flux used for the weak imposition of boundary or continuity conditions \cite{hesthaven2007nodal}.
+This method can be derived from an appropriate DG formulation under a specific choice of quadrature.
+
+This formulation conserves the quantities $\bm{u}_i$, but (except in specific cases) does not satisfy a discrete analogue of the statement of entropy conservation (\ref{eq:consentropy}).
+Entropy conservative schemes may be constructed based on the SBP property and a ``flux differencing'' approach \cite{gassner2016split} involving a definition of entropy conservative/stable fluxes given by Tadmor \cite{tadmor1987numerical, tadmor2003entropy}.
+\begin{definition}
+Let $\bm{f}_S(\bm{u}_L,\bm{u}_R)$ be a bivariate function such that it is symmetric and consistent with the flux function $\bm{f}(\bm{u})$
+\[
+\bm{f}_S(\bm{u}_L,\bm{u}_R) = \bm{f}_S(\bm{u}_R,\bm{u}_L), \qquad \bm{f}_S(\bm{u},\bm{u}) = \bm{f}(\bm{u})
+\]
+The numerical flux $\bm{f}_S(\bm{u}_L, \bm{u}_R)$ is entropy conservative if, for entropy variables $\bm{v}_L = \bm{v}(\bm{u}_L), \bm{v}_R = \bm{v}(\bm{u}_R)$
+\[
+\LRp{\bm{v}_L - \bm{v}_R}^T \bm{f}_S(\bm{u}_L,\bm{u}_R) = (\psi_L - \psi_R), \qquad \psi_L = \psi(\bm{v}(\bm{u}_L)), \quad \psi_R = \psi(\bm{v}(\bm{u}_R)).  
+\]
+Similarly, a flux $\bm{f}_S(\bm{u}_L, \bm{u}_R)$ is referred to as entropy stable if $\LRp{\bm{v}_L - \bm{v}_R}^T \bm{f}_S(\bm{u}_L,\bm{u}_R) \leq (\psi_L - \psi_R)$.
+\label{def:tadmor}
+\end{definition}
+An entropy conservative nodal collocation method can then be derived by modifying the flux derivative as follows
+\begin{equation}
+\pd{\bm{u}_i}{t} + \sum_{j=1}^{N+1}\LRp{\LRp{2\bm{D}\otimes \bm{I}}_{ij}\bm{f}_S(\bm{u}_i,\bm{u}_j) +\LRp{\bm{M}^{-1}\bm{B}}_{ij}(\bm{f}_i^* - \bm{f}(\bm{u}_i))}= 0.
+\label{eq:dgsem}
+\end{equation}
+The replacement of the derivative of the flux function with the flux differencing allows one to prove that (\ref{eq:dgsem}) conserves a discrete entropy \cite{gassner2017br1,chen2017entropy}.
+We reproduce these proofs here, as we will refer to them in following sections.
+\begin{theorem}
+The collocation method (\ref{eq:dgsem}) is discretely entropy conservative in the sense that
+\[
+\bm{1}^T\LRp{\bm{M}\pd{U(\bm{u})}{t} + \bm{B}\LRp{\bm{v}^T\bm{f}^* - \bm{\psi}}} = 0,
+\]
+where $\bm{1}$ denotes the vector of ones and $\bm{\psi}_i = \psi(\bm{v}_i)$ is the entropy potential in terms of the nodal value of $\bm{u}_i$.  
+\label{thm:dgsem}
+\end{theorem}
+\begin{proof}
+Let $\bm{v}_i = \bm{v}(\bm{u}_i)$ denote the entropy variables, defined in terms of the nodal values $\bm{u}_i$.  We multiply (\ref{eq:dgsem}) by $\LRp{\bm{M}\bm{v}}^T$.  
+Using the fact that $\bm{M}$ is diagonal, the time derivative term can be manipulated to 
+\[
+\bm{v}^T\bm{M}\pd{\bm{u}}{t} = \sum_{i=1}^{N+1} \bm{M}_{ii} \bm{v}_i^T \pd{\bm{u}_i}{t} = \sum_{i=1}^{N+1} \bm{M}_{ii} \pd{S_i}{t} = \bm{1}^T\bm{M}\pd{S_i}{t},
+\]
+where we have used continuity in time to arrive at the pointwise relation 
+\[
+\bm{v}^T \pd{\bm{u}}{t} = \pd{U(\bm{u})}{\bm{u}}^T \pd{\bm{u}}{t} = \pd{U(\bm{u})}{t}.
+\]
+
+The spatial derivative term can be manipulated using the summation by parts property $\bm{M}\bm{D} = \bm{S} = \bm{B}-\bm{S}^T$
+\begin{align}
+\sum_{i,j=1}^{N+1} \bm{v}_i^T 2 \bm{M}\bm{D}_{ij}\bm{f}_S(\bm{u}_i,\bm{u}_j) &= \sum_{i,j=1}^{N+1} \bm{v}_i^T (\bm{S} + \bm{B} - \bm{S}^T)_{ij}\bm{f}_S(\bm{u}_i,\bm{u}_j) \nonumber\\
+&= \sum_{i,j=1}^{N+1} \bm{S}_{ij}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j) - \bm{S}_{ji}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j) + \bm{B}_{ij}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j). \label{eq:sbpentropy}
+\end{align}
+Rearranging indices and using symmetry of $\bm{f}_S(\bm{u}_i,\bm{u}_j)$ then yields that
+\begin{align*}
+\sum_{i,j=1}^{N+1} \bm{S}_{ij}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j) - \bm{S}_{ji}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j) &= \sum_{i,j=1}^{N+1} \bm{S}_{ij} \LRp{\bm{v}_i - \bm{v}_j}^T\bm{f}_S(\bm{u}_i,\bm{u}_j)\\
+&= \sum_{i,j=1}^{N+1} \bm{S}_{ij} \LRp{\bm{\psi}_i-\bm{\psi}_j} = -\bm{1}^T \bm{S}\bm{\psi} = -\bm{1}^T\bm{B}\bm{\psi},
+\end{align*}
+where we have used that $\bm{S}\bm{1} = 0$ and the summation by parts property in the last line.
+
+Using that $\bm{B}$ is diagonal, combining the latter term in (\ref{eq:sbpentropy}) with boundary terms involving the numerical flux yields
+\[
+\sum_{i,j = 1}^{N+1} \bm{B}_{ij}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j) + \bm{v}_i^T\bm{B}_{ij}\LRp{\bm{f}_i^* - \bm{f}(\bm{u}_i)} = \bm{1}^T\bm{B} \LRp{\bm{v}^T \bm{f}^*},
+\]
+where we have used the consistency of $\bm{f}_S(\bm{u},\bm{u}) = \bm{f}(\bm{u})$.  Combining everything results in a discrete statement of the conservation of entropy (\ref{eq:consentropy}).   
+%Combining everything results in a discrete statement of the conservation of entropy (\ref{eq:consentropy}) 
+%\[
+%\bm{1}^T\LRp{\bm{M}\pd{U(\bm{u})}{t} + \bm{B}\LRp{\bm{v}^T\bm{f}^* - \bm{\psi}}} = 0.
+%\]
+\end{proof}
+
+\begin{remark}
+The proof of entropy conservation can be extended beyond a single domain by specifying appropriate simultaneous approximation terms (SBP-SAT).
+This can be done by taking $\bm{f}_i^*$ to be the entropy conservative flux $f_S(u_L,u_R)$ at the interface between two elements \cite{carpenter2014entropy,gassner2017br1,chen2017entropy}.
+Similarly, employing an entropy stable flux at element interfaces results in an entropy stable method which satisfies a global entropy inequality.
+\end{remark}
+
+\section{Entropy conservative general DG methods}
+
+In this section, we focus on constructing entropy conservative DG schemes with arbitrary bases and quadrature rules.
+These include dense norm SBP operators and generalized SBP (GSBP) operators which do not contain boundary points, which introduce additional difficulties in constructing entropy conservative DG schemes.
+These difficulties present themselves primarily in three areas:
 \begin{enumerate}
-\item The quadrature induces an $L^2$-equivalent inner product over the reference element $\widehat{D}$.
-\item The quadrature is sufficiently accurate such that integration by parts holds with respect to the reference coordinates $\widehat{x}$.
+\item It is unclear how to generalize the flux differencing approach to arbitrary choices of quadrature.  
+\item The proof of entropy stability relies on a diagonal norm mass matrix, which commutes with pointwise multiplication by the (nodal values of) the entropy variables.  
+\item Entropy stable formulations based on dense norm or GSBP operators require non-trivial SBP-SAT terms \cite{ranocha2017extended}.
+These can be derived by hand for simple problems such as Burgers' equation, but can be difficult to determine for more complex systems of nonlinear PDEs.
 \end{enumerate}
 
-\subsection{Discrete derivatives}
-We introduce two discrete DG derivatives in this section, based on definitions used in \cite{di2011mathematical} and \cite{hesthaven2004high,Warburton20063205}:
-\begin{definition}
-We define two discontinuous Galerkin differentiation operators ${D}^i: L^2\LRp{\Omega}\rightarrow V_h$ and $\tilde{D}^i: L^2\LRp{\Omega}\rightarrow V_h$ is defined implicitly as follows:
-\begin{align}
-\LRp{{D}^i u,v}_{\Oh} &= \sum_{k} \LRp{\LRp{-u,\pd{v}{\bm{x}_i}}_{\LK} + \LRa{\avg{u},v\bm{n}_i}_{\LdK}}, \qquad \forall v \in V_h\\
-\LRp{\tilde{D}^i u,v}_{\Oh} &= \sum_{k} \LRp{\LRp{\pd{u}{\bm{x}_i},v}_{\LK} - \frac{1}{2}\LRa{\jump{u},v\bm{n}_i}_{\LdK}}, \qquad \forall v\in V_h.
-\label{eq:dgd1}
-\end{align}
-\end{definition}
+We simultaneously address all three areas of difficulty by introducing a generalization of the entropy conservative DG-SEM scheme (\ref{eq:dgsem}) to arbitrary quadratures and choices of basis.
+Additionally, we prove that the generalized scheme is entropy conservative (stable) on multiple elements in the following sections.
 
-We note that, assuming $u,v \in V_h$ and sufficiently accurate quadrature, $D^i$ and $\tilde{D}^i$ are equivalent.
-The main property of the discrete derivatives utilized in this work is a discrete integration-by-parts formula.
-\begin{lemma}
-Let $u,v\in V_h$.  Then, 
+%We will get around these difficulties for dense-norm DG methods by satisfying three requirements for entropy stability
+%\begin{itemize}
+%\item $\bm{D}\bm{1} = 0$ for use of the flux formulation.  
+%\item For some norm matrix $\bm{W}$, $\bm{W}\bm{D} = \bm{B}-\bm{D}^T\bm{W}$ where $\bm{B}$ is a suitably defined boundary operator.
+%\item Contraction with the \textit{projection} of the entropy variables.  
+%\end{itemize}
+
+\subsection{A continuous interpretation of flux differencing}
+
+We define the quadrature-based $L^2$ projection $\Pi_N$ as the $L^2$ projection where integrals are evaluated using some (inexact) quadrature.
+We also define the discrete approximation space $V_h$, and assume that the quadrature is sufficiently accurate such that integration by parts holds for any two elements of $V_h$.
+Finally, we assume that the discrete derivative (to be defined) maps from $V_h$ to $V_h$.
+
+The split-form methodology of Gassner et al.\ can be interpreted as the following: let $f_S(u_L,u_R)$ be a symmetric and consistent two-point flux (i.e.
+$f_S(u,u) = f(u)$).
+Split forms can be rewritten at the continuous level as replacing the flux derivative with differentiation along one argument of $f_S(u_L,u_R)$, then evaluation of the other argument at $u$
+\[
+\pd{f(u)}{x} \Longrightarrow 2\left.\pd{f_S(u(x),u(y))}{x}\right|_{y=x}.
+\]
+This allows for the recovery of different split-forms.  For example, for Burgers equation, choosing $f_S(u_L,u_R) = (u_L^2 + u_Lu_R + u_R^2)/6$ gives
+\[
+2\left.\pd{f_S(u(x),u(y))}{x}\right|_{y=x} = \frac{1}{3}\left.\pd{\LRp{u(x)^2 + u(x)u(y) + u(y)^2}}{x}\right|_{y=x} = \frac{1}{3}\LRp{\pd{u^2}{x} + u\pd{u}{x}}
+\]
+due to the fact that $\pd{u(y)^2}{x} = 0$.
+The resulting form of Burgers equation appears in the intermediate steps of energy stability proofs, and is the basis for discretely stable DG methods.
+When applied discretely (using quadrature-based projections), the resulting right hand side for Burgers' equation becomes
+\[
+\LRp{\frac{1}{3}\LRp{\pd{\LRp{\Pi_N u^2}}{x} + \Pi_N \LRp{u\pd{u}{x}}}, v}.
+\]
+Using the fact integration by parts holds when considering two elements of $V_h$, and that $\LRp{\Pi_N u^2, v} = \LRp{u^2,v}$ for any $v \in V_h$, it can be shown that the volume terms cancel out for $v = u$, and we're left with boundary terms (which is a prerequisite of energy/entropy stability).
+
+We note that the recovery of the split-form of Burgers' equation in the above example relies only on the property that 
+\[
+\pd{f(u(y))}{x} = f(u(y)) \pd{1}{x} = 0.
+\]
+As such, replacing $\pd{}{x}$ by any differential operator $D$ such that $D 1 = 0$ will recover a similar formulation.
+
+This approach yields the following evaluation (in Matlab for lack of better notation): let 
+\[
+\verb|fS = @(uL,uR) uL.^2+uL.*uR+uR.^2|
+\]
+be the flux function.
+Let \verb+uq = Vq * u+ denote the interpolation of $u$ to quadrature points, and let \verb+Dx+ denote the differentiation matrix.
+Define \verb+Pq = M\(Vq' * diag(wq))+, where \verb+wq+ are quadrature weights, and let \verb+[ux uy] = meshgrid(uq)+.
+Then, assuming element-constant Jacobians, the split form can be evaluated via
+\[
+\verb+Pq * diag((Vq*Dx*Pq) * fS(ux, uy)) = Pq * sum((Vq* Dx* Pq) .* fS(ux, uy), 2).+
+\]
+The latter expression can be performed by computing \verb+fS(ux,uy)+ on the fly, avoiding the need to store it as a matrix.
+It can also be generalized to handle non-square matrices in a very similar way.
+
+\subsection{Discretely entropy conservative general DG methods on a single domain}
+
+Global DG operators reveal how to construct SBP-SAT terms for any collocation-type DG methods (where $L^2$ projection and interpolation are equivalent).
+However, for quadrature-based DG methods, it is not possible to test with the entropy variables directly because the test space is polynomial.
+However, we may test with the projections of the entropy variables.
+On the left hand side, we may use the fact that, under a method of lines discretization,
+\[
+u(x,t) = \sum_{j=1}^{N_p} u_j(t) \phi_j(x), \qquad \pd{u}{t} \in P^N.
+\]
+Thus, testing with $\Pi_N v$ yields
+\[
+\int_{D^k} \pd{u}{t}\Pi_N v = \int_{D^k} \pd{u}{t} v   
+\]
+which yields a pointwise relationship which can be manipulated into the time derivative of entropy.
+
+The right-hand side is more difficult.
+For collocation DG methods, interpolation and quadrature-based $L^2$ projection are identical.
+However, when using a quadrature rule with $N_q > N_p$, this is not the case.
+The solution is to evaluate the flux with $u(\Pi_N v)$ - in other words, evaluate the flux variables in terms of the projections of the entropy variables.
+
+\subsection{SBP-SAT terms by global DG differentiation operators}
+
+We can define a global discontinuous Galerkin operator $\bm{D}^x_h: P^N\rightarrow P^N$ such that
+\[
+\sum_k \LRp{D^x_h u,v}_{D^k} = \sum_k \LRp{\pd{u}{x},v}_{D^k} + \LRa{\frac{1}{2}\jump{u}\bm{n}_x,v}_{\partial D^k}.
+\]
+Then, composing $D^x_h \Pi_N$ gives global skew-symmetry $\LRp{D^x_h\Pi_N u, v}_{\L} = \LRp{- D^x_h\Pi_N v, u}_{\L}$ for all $u,v\in \L$.
+The downside to this simple construction is that the stencil of the DG scheme now incorporates all degrees of freedom on neighboring elements due to the composition with the projection.
+
+We adapt the construction of local differentiation operators by Shi and Shu to construct a global DG differentiation operator.  
+\[
+2\LRp{D^x_h u,v} = \sum_k \LRp{2\pd{ \Pi_N u}{x},v}_{D^k} + \LRa{\LRp{u_P - \Pi_N u_M}\bm{n}_x, v} + \LRa{\LRp{u_M - \Pi_Nu_M }\bm{n}_x,\Pi_Nv}.
+\]
+In skew-symmetric form, this gives 
+\[
+2\LRp{D^x_h u,v} = \sum_k \LRp{\pd{ \Pi_N u}{x},v}_{D^k} - \LRp{u,\pd{\Pi_N v}{x}}_{D^k} + \LRa{\LRp{u_P - \Pi_N u_M}\bm{n}_x, v} + \LRa{{u_M }\bm{n}_x,\Pi_Nv}.
+\]
+Taking $v = u_M$ and cancelling $u_P u_M \bm{n}_x$ terms shows that the differentiation operator has the global SBP property and is skew-symmetric  w.r.t.\ a discrete $L^2$ inner product.
+The coupling requires only traces of neighboring elements.
+
+It's currently unclear how to distinguish whether different operators are high order accurate.
+
+\subsection{Conservation of secondary properties}
+
+It holds but with respect to specifically defined fluxes.
+
+\section{Numerical experiments}
+
+\subsection{Burgers}
+
+\subsection{Shallow water}
+
+We describe in this section the shallow water equations in one dimension.
+We use this model problem to illustrate the proposed method because (unlike Burgers' equation) it highlights some non-trivial details in proving entropy stability associated with the nonlinear equations of compressible flow.
+
+The one-dimensional shallow water equations are 
 \begin{align*}
-\LRp{D^i u,v} &= \LRp{-u,D^i_h v} + \int_{\partial \Omega} uv \bm{n}_i.
-%\LRp{D_h^i u,v} &= \LRp{-u,D^i_h v} + \int_{\partial \Omega} uv \bm{n}_i.
+\pd{h}{t} + \pd{hv}{x} &= 0\\
+\pd{hv}{t} + \pd{\LRp{hv^2 + gh^2/2}}{x} + gh\pd{b}{x}&= 0.
 \end{align*}
-\label{lemma:ibp}
-\end{lemma}
-\begin{proof}
-Assuming that quadrature is sufficiently accurate for $u, v\in V_h$, integration by parts gives the equivalence. 
-\end{proof}
+where $h$ is the water height, $v$ is the velocity, $g$ is the gravitational constant, and $b$ is bottom topography.
 
-For $u \not\in V_h$, due to the fact that the quadrature rule (\ref{eq:quad}) may not be exact, and these two definitions are not equivalent and Lemma~\ref{lemma:ibp} does not hold.
-This motivates the definition of a \textit{discrete} DG differentiation operator:
-\begin{definition}
-The discrete discontinuous Galerkin differentiation operator ${D}^i_h: L^2\LRp{\Omega} \rightarrow V_h$ is defined as $D^i_h = D^i \Pi_N$.
-\end{definition}
-
-Using these discrete derivatives, we can define discrete gradient and divergence operators
-\begin{definition}
-Let $u \in V_h$ and $\bm{u}\in \LRp{V_h}^d$.  Then, discrete gradient and divergence operators are defined as
-\begin{align*}
-\LRp{\Grad^{\rm DG} u}_i &= D^i u, \qquad \Grad^{\rm DG}\cdot \bm{u} = \sum_{i=1}^d D^i \bm{u}_i\\
-\LRp{\Grad^{\rm DG}_h u}_i &= D_h^i u, \qquad \Grad^{\rm DG}_h\cdot \bm{u} = \sum_{i=1}^d D_h^i \bm{u}_i.
-\end{align*}
-\end{definition}
-
-When the support of $v$ is limited to a single element, we have
+The entropy associated with this system is the total energy $e = k + p$, where 
 \[
-\LRp{\Grad_h\cdot \bm{u},v\bbm{1}_{D^k}} = \LRp{\bm{u}, \Grad\cdot v\bbm{1}_{D^k}}  + \LRa{\avg{\bm{u}}\cdot\bm{n},v}_{\partial D^k}.
+k = hv^2/2, \qquad p = gh^2/2 + ghb.
 \]
-and as a result when $v = 1$
+Testing with the entropy variables 
 \[
-\LRp{\Grad_h\cdot \bm{u},\bbm{1}_{D^k}} = \int_{\partial D^k}\avg{\bm{u}}\cdot\bm{n}
+q_1 = \pd{e}{h} = g(h+b)-v^2/2, \qquad q_2 = \pd{e}{hv} = v
+\]
+recovers a statement of conservation of energy
+\[
+\pd{e}{t} + \pd{}{x}\LRp{hv^3/2 + g(hv)(h+b)} = 0.  
 \]
 
-\section{Examples}
+This conservation of energy does not hold at the discrete level for two reasons.
+The first reason is because, under inexact quadrature, calculus identities such as the chain or product rule do not hold, and naive spatial formulations can not be manipulated into a form which is amenable to a proof of discrete entropy stability.
+This issue can be remedied by the use of a split formulation (as shown in Gassner et al.) which is provably entropy stable under inexact quadrature.
 
-\begin{itemize}
-\item Tensor product elements: inner product computed using GLL nodes, $\Pi_N$ reduces to nodal interpolation
-\item General elements: $\Pi_N$ is just quadrature-based projection.  For accuracy, degree $2N$ or higher?  
-\item Triangles and tetrahedra: lumped quadrature nodes.
-Implies that mass lumping on triangles requires sufficiently accurate quadrature such that IBP holds.
-This conclusion was first reached in the study of multi-dimensional SBP operators \cite{hicken2016multidimensional}.
-\item General elements: $\Pi_N$ can be taken to be nodal interpolation.  
-\item Pyramids: need to account for non-affine mappings.  
-\end{itemize}
+The second more subtle reason why entropy conservation does not hold at the discrete level is because the entropy variables are not contained within the test space.
+Moreover, the spatial part of the proposed discretization requires one discrete definition of the entropy variables for entropy stability, while the temporal part of the discretization requires another discrete definition of the entropy variables.
+We address this by adopting a rescaling in the mass matrix on the left hand side involving both definitions of the discrete entropy variables.
+Under an appropriate choice of basis and quadrature, the proposed approach recovers existing DG-SEM discretizations.
 
-\subsection{Variable advection}
+\subsection{Compressible Euler equations}
 
-A split formulation for advection is 
-\[
-\LRp{\pd{u}{t},v} + \frac{1}{2}\LRp{\Grad_h\cdot \Pi_N \LRp{ \bm{\beta}u},v} + \frac{1}{2}\LRp{\bm{\beta}\cdot\Grad_h u,v} + \frac{1}{2}\LRp{\LRp{\Grad\cdot \bm{\beta}} u,v} = 0.
-\]
-Taking $v = u$ yields and using $\LRp{\Grad_h u, \bm{v}} = \LRp{-u, \Grad_h \cdot \bm{v}}$ yields the energy statement
-\[
-\frac{1}{2}\nor{u}^2 + \frac{1}{2}\LRp{\Grad_h\cdot \Pi_N \LRp{ \bm{\beta}u},u} - \frac{1}{2}\LRp{ u, \Grad_h\cdot \Pi_N\LRp{\bm{\beta}u}} = \frac{1}{2}\LRp{-\LRp{\Grad \cdot \bm{\beta}} u,u},
-\]
-implying that $\frac{1}{2}\nor{u}^2 = 0$ if $\Grad\cdot \bm{\beta} = 0$, or that the method is energy conserving.
-The only difference in this formulation is the introduction of $\Pi_N$, which can be defined at a discrete level using any quadrature scheme for which a discrete projection is well-defined.
+\section{Conclusions}
 
-Penalization can be added by adding any positive-definite stabilization term (upwind, penalty, Lax-Friedrichs) through the regular divergence flux.
-
-\section{Discrete DG derivatives}
-
-Methods based on discrete DG derivatives also work.
-
-The discrete DG derivative-based method is not consistent in the sense that Galerkin orthogonality does not hold exactly.  The difference lies in the flux terms.  Assume $\Div \bm{\beta} = 0$, then
-\[
-\LRp{\pd{u}{t},v} + \frac{1}{2}\LRp{-{ \bm{\beta}u},\Grad_h v} + \frac{1}{2}\LRp{\bm{\beta}\cdot\Grad_h u,v} = 0.
-\]
-\begin{align*}
-\LRp{\Grad_h\cdot \Pi_N\LRp{ \bm{\beta}u},v} &= \sum_{k} \LRp{-\bm{\beta}u, \Grad v}_{\LK} + \LRa{\avg{\Pi_N\LRp{\bm{\beta} u}}\cdot \bm{n},v}_{\LdK}\\
-\LRp{\Grad_h u,\bm{\beta}v} &= \sum_{k} \LRp{-u, \Grad \cdot \LRp{\bm{\beta}v}}_{\LK} + \LRa{\bm{\beta}\cdot\bm{n}\avg{u},v}_{\LdK}.
-\end{align*}
-The latter term is consistent; the former is not due to the presence of $\avg{\Pi_N\LRp{\bm{\beta} u}}\cdot \bm{n}$ in the flux term.
-The consistency error should then be $O(h^{N+1/2})$ using a trace inequality for $L^2$ projections.
-
-Note: can also use interpolants in a stable manner if using $D_h$.
-Unlike SEM, this still requires an extra matvec per RHS evaluation because of the lack of diagonality of the mass matrix.
-Reduces number of steps by one (no interpolation to quadrature points) but does not reduce number of total matvecs.
-
-\subsection{Local conservation}
-
-Writing this in non-conservative form raises the question of local conservation.  Integrating the original equation over $D^k$ and using Gauss' theorem gives
-\[
-\int_{D^k}\pd{u}{t} + \int_{\partial D^k} \beta_n u = 0.
-\]
-Taking $v = 1$ on $D^k$ yields
-\[
-\int_{D^k}\pd{u}{t} + \frac{1}{2}\LRp{\Grad_h\cdot \Pi_N \LRp{ \bm{\beta}u},\bbm{1}_{D^k}} + \frac{1}{2}\LRp{\bm{\beta}\cdot\Grad_h u,\bbm{1}_{D^k}} + \frac{1}{2}\LRp{\LRp{\Grad\cdot \bm{\beta}} u,\bbm{1}_{D^k}} = 0.
-\]
-The first term gives
-\[
-\LRp{\Grad_h\cdot \Pi_N \LRp{ \bm{\beta}u},\bbm{1}_{D^k}} = \int_{\partial D^k} \avg{\Pi_N\LRp{\bm{\beta}u}}\cdot\bm{n}.
-\]
-The second term gives
-\[
-\LRp{\bm{\beta}\cdot\Grad_h u,\bbm{1}_{D^k}} = \LRp{\Grad u,\bm{\beta}}_{D^k} + \frac{1}{2}\LRa{\jump{u},\bm{\beta}\cdot\bm{n}} = \LRp{u,-\Grad \cdot \bm{\beta}}_{D^k} + \LRa{\avg{u},\bm{\beta}\cdot\bm{n}}
-\]
-through integration by parts and an assumption that $\bm{\beta}\cdot \bm{n}$ is periodic.  Cancelling volume terms, we end up with the statement of local conservation 
-\[
-\int_{D^k} \pd{u}{t}  + \frac{1}{2}\int_{\partial D^k} \LRp{\avg{\Pi_N\LRp{\bm{\beta}u}} + \LRp{\Pi_N\LRp{\bm{\beta}}\avg{u}}}\cdot\bm{n} = 0
-\]
-which is a discrete version of the continuous statement of local conservation.
-
-\section{Non-affine mappings}
-\label{sec:curv}
-
-For general elements and curvilinear meshes, $J$ and $\bm{G}$ are no longer constant over each element.
-However, for isoparametric mappings (and appropriate polynomial interpolations of general curvilinear mappings), metric identities hold \cite{kopriva2006metric} such that $\widehat{\Grad} \cdot \LRp{J\bm{G}^T} = 0$.
-Thus, we can write
-\[
-\LRp{\Grad u,\bm{v}}_{D^k} = \LRp{J\bm{G}\widehat{\Grad} u,\bm{v}}_{\widehat{D}} = \LRp{-u,\widehat{\Grad} \cdot \LRp{J\bm{G}\bm{v}}}  + \LRa{J\bm{G}^T\widehat{\bm{n}} u,\bm{v}}.   
-\]
-Noting that $J\bm{G}\widehat{\bm{n}} = J^f \bm{n} $ \cite{hesthaven2007nodal}, we recover integration by parts over the physical element.
-
-Thus, to extend the discrete discontinuous Galerkin gradient and divergence to non-affine mappings, we simply define
-\begin{align*}
-\LRp{\widehat{\Grad}_h u,\bm{v}}_{\Omega} &= \sum_{k} \LRp{\widehat{\Grad}u,\bm{v}}_{\widehat{D}}  - \frac{1}{2}\LRa{\widehat{\bm{n}} \jump{u},\bm{v}} \\
-\LRp{\widehat{\Grad}_h \cdot\bm{u},{v}}_{\Omega} &= \sum_{k} \LRp{\widehat{\Grad}\cdot \bm{u},{v}}_{\widehat{D}}  - \frac{1}{2}\LRa{\jump{\bm{u}},v\widehat{\bm{n}}} 
-%\LRp{\Grad_h \cdot \bm{u},v}_{\Omega} &= \sum_{k} \LRp{\widehat{\Grad}\cdot{\Pi_N\LRp{J\bm{G}\bm{u}}},v}  - \frac{1}{2}\LRa{\widehat{\bm{n}} \jump{J\bm{G}\bm{u}},v} \\
-%\LRp{\Grad_h u,\bm{v}}_{\Omega} &= \sum_{k} \LRp{\Pi_N\LRp{J\bm{G}\widehat{\Grad} u},\bm{v}}  - \frac{1}{2}\LRa{\widehat{\bm{n}} \jump{J\bm{G}u},v}.
-\end{align*}
-and define the discrete physical gradient and divergence as
-\begin{align*}
-\LRp{\Grad_h u,v}_{\Omega} &= \LRp{J\bm{G}\widehat{\Grad}_h u,\bm{v}}_{\Omega} = \sum_{k}\LRp{J\bm{G}\widehat{\Grad}u,\bm{v}}_{\widehat{D}}  - \frac{1}{2}\LRa{J\bm{G}\widehat{\bm{n}} \jump{u},\bm{v}} \\
-\LRp{\Grad_h \cdot \bm{u},v}_{\Omega} &= \LRp{\widehat{\Grad}_h\cdot\LRp{J\bm{G}^T\bm{u}}}_{\Omega} = \sum_{k} \LRp{\widehat{\Grad}\cdot \LRp{J\bm{G}^T\bm{u}},{v}}_{\widehat{D}}  - \frac{1}{2}\LRa{\jump{J\bm{G}^T\bm{u}},v\widehat{\bm{n}}}.
-\end{align*}
-The discrete gradient is consistent with the physical gradient.
-The discrete divergence is also consistent with the physical divergence, provided that the jump term $\LRa{\jump{J\bm{G}^T\bm{u}},v\widehat{\bm{n}}}$ is computed via
-\[
-\LRa{\jump{J\bm{G}^T\bm{u}},v\widehat{\bm{n}}} = \LRa{\jump{J\LRp{\bm{G}\widehat{\bm{n}}}^T\bm{u}},v} = \LRa{\LRp{\widehat{\bm{n}}^T\LRp{J{\bm{G}}^T\bm{u}}}^+ + \LRp{\widehat{\bm{n}}^T\LRp{J{\bm{G}}^T\bm{u}}}^-,v}.
-\]
-
-The advantage of this projected geometric factor approach is local conservation; taking $v = \bbm{1}_{D^k}$ recovers (after integration by parts) the flux integral
-\[
-\LRp{\Grad_h \cdot \bm{u},\bbm{1}_{D^k}}_{\Omega} = \int_{\partial D^k}\widehat{\bm{n}}^T\avg{J\bm{G}^T\bm{u}}.
-\]
-
-\section{Extension to other hyperbolic problems}
-
-Example: acoustic wave equation, simply discretize by replacing $\Grad, \Grad\cdot$ with discrete versions.
-Automatically skew symmetric and energy stable via integration by parts.
-Also, can show why WADG works: discretize based on discrete divergence, then test with $T_{c^2}^{-1}p$ and use identities.
-Note - I think this requires the use of the strictly discrete version.
-
-Example: Burgers' equation
-
-Example: Kinetic energy preserving splitting of Euler (assumes exact time discretization).  Doesn't seem to help much without extra viscosity?
-
-Example: Entropy splitting of Buckley-Leverett?
-
-Example: Entropy splitting of Euler (note - cannot extend to Navier-Stokes in an entropy-stable fashion due to fact that heat flux matrix is not symmetrizable w.r.t.\ homogeneous flux function, though viscous terms are.
-This impacts only boundary conditions.)
-
-%\section{Standard entropy stability estimates}
-%
-%Given a nonlinear conservation law
-%\begin{align*}
-%\pd{u}{t} + \pd{f(u)}{x} = 0
-%\end{align*}
-%DG formulation is usually
-%\[
-%\LRp{\pd{u}{t},v} + \LRp{D_h f,v} = 0.
-%\]
-%Taking $v = u$ gives
-%\[
-%\frac{1}{2}\nor{u}^2 + \int_{D^k}{u D_h f} = 0.
-%\]
-%Continuous entropy stability relies on the introduction of $F,G$ such that
-%\[
-%\pd{F}{u} = u\pd{f}{u}, \qquad \pd{G}{u} = f, \qquad F = uf - G.
-%\]
-%Then, relying on product and chain rules
-%\[
-%\int_{D^k}{u \pd{f}{x}} = \int_{D^k}{\pd{(uf)}{x} - f\pd{u}{x}} = \int_{D^k}{\pd{(uf)}{x} - \pd{G}{u}\pd{u}{x}} =  \int_{D^k}{\pd{\LRp{uf-G}}{x}} = \int_{D^k}{\pd{F}{x}} = \int_{\partial D^k} F.
-%\]
-%These boundary terms can be made to cancel with appropriately defined numerical fluxes, resulting in entropy conservative schemes.
-The challenge in reproducing this in the discrete case is the lack of a product and chain rule for inexact quadratures.
-Jameson deals with this by introducing an integrated flux
-%\[
-%G = \frac{1}{2}\int_{-1}^1 f \LRp{ \avg{u} + \theta \jump{u}} \diff \theta.
-%\]
-%for which finite volume schemes satisfy (in a rough sense)
-%\[
-%\jump{G} = \avg{G} \jump{u}.
-%\]
-%Fischer and Carpenter (also Gassner and co-workers) use a similar idea but combine a symmetric two-point flux approximation with properties of SBP matrices to get 
-%\[
-%2\LRp{W,DF} = \LRp{W,DF} - \LRp{DW,F} + \LRa{W,F} =  \sum_i \sum_j (W_i-W_j) (MD)_{ij} F(U_i,U_j) + \LRa{W,F}.
-%\]
-%
-%\section{Entropy splitting}
-%
-%It's currently unclear how to extend the symmetric two-point flux approximation to quadrature-based DG methods.
-
-%Sandham and Yee use the entropy splitting 
-%\[
-%\eta(U) =  -\beta p^*, \qquad p^* = - \LRp{\frac{p}{\rho^\gamma}}^{\frac{1}{\beta(1-\gamma)}},
-%\]
-%which generates the entropy variables
-%\[
-%W(U) = p^*\LRp{\begin{array}{ccc}
-%\frac{E}{p} - C, &\frac{-\rho u}{p}, &\frac{\rho}{p}
-%\end{array}}, \qquad C = \frac{2}{\gamma-1} + (1+\beta).
-%\]
-%Unfortunately, the inverse mapping $U(W)$ is not explicitly known, which will greatly hamper computational efforts.
-For example, this implies we cannot evolve the entropy variables explicitly, which would yield an energy estimate.
-%The Jacobian
-%\[
-%\pd{U}{W} = \LRs{
-%d
-%}, 
-%\qquad \pd{W}{U} = \pd{U}{W}^{-1} = \LRs{
-%}
-%\]
-%The flux Jacobian
-%\[
-%\pd{F}{U} = \LRs{\begin{array}{ccc}
-%0 & u \frac{(\gamma-3)}{2} & \\
-%1 & -u(\gamma-3) & 
-%\end{array}}
-%\]
+\section{Acknowledgments}
 

### 2017-07-13 docs/splitform/main.tex
more paper updates

@@
 \author{Jesse Chan}
-\title{On discrete entropy conservation for general high order discontinuous Galerkin methods}
+\title{On discrete entropy conservation for a more general class of discontinuous Galerkin methods}
 
@@
 High order methods based on diagonal-norm summation by parts operators can be shown to semi-discretely conserve entropy for nonlinear systems of hyperbolic PDEs \cite{fisher2013high,carpenter2014entropy}.
-These include discontinuous Galerkin spectral element methods (DG-SEM) \cite{gassner2016split,gassner2016well,wintermeyer2017entropy, chen2017entropy} based on mass-lumping.
-In this work, we describe how to extend discretely entropy conservative schemes for a more general class of DG methods using quadrature-based projections and global differentiation operators.
+These include discontinuous Galerkin spectral element methods (DG-SEM) \cite{gassner2016split,gassner2016well,wintermeyer2017entropy, chen2017entropy} based on mass lumping and nodal collocation.
+In this work, we describe how to extend discretely entropy conservative schemes for a more general class of DG methods using quadrature-based projections and global DG differentiation operators.
 The construct of these methods is independent of the choice of basis and accuracy of quadrature rule used, and includes DG methods involving generalized SBP operators without boundary nodes and certain dense-norm SBP operators.
@@
 
-\cite{hicken2016multidimensional}
-
-\cite{gassner2016split,gassner2016well,wintermeyer2017entropy}
-
-\cite{chen2017entropy}
+\note{Specify that entropy stability has been shown for mass lumped collocation DG discretizations.
+Requirements are a diagonal norm SBP operators (or diagonal mass matrix) and collocated differentiation.
+}  \cite{gassner2016split,gassner2016well,wintermeyer2017entropy}
+\note{The construction of such operators on non-tensor product elements (such as simplices or pyramids) is non-trivial.} \cite{hicken2016multidimensional, chen2017entropy}.
+\note{Constructing such operators for non-polynomial spaces (such as splines) is also difficult within the current framework.}
 
@@
 
-\section{Entropy conservative collocation DG methods}
+\section{Entropy conservative diagonal norm collocation DG methods}
+
+\note{Specify what we mean by diagonal norm collocation DG method.}
 
@@
 
-%In order to simplify notation for flux differencing, we define the following bilinear matrix product 
-%\[
-%\LRp{\bm{A} \odot \bm{f}_S(\bm{x},\bm{y})}_i = \sum_j \bm{A}_{ij} \bm{f}_S(\bm{x}_j,\bm{y}_i).  
-%\]
-We seek a nodal approximation $\bm{u}_i \approx \bm{u}(x_i)$ to the solution $\bm{u}(x)$ of (\ref{eq:pde}).  A typical nodal collocation method in conservation form is given as follows 
+We seek a polynomial approximation to the solution $\bm{u}(x)$ of (\ref{eq:pde}).  A DG formulation in the ``strong form'' \cite{hesthaven2007nodal} is given as
+\[
+\int_{-1}^1\LRp{ \pd{\bm{u}}{t} + \pd{\bm{f}(\bm{u})}{x}}\bm{w} + \left.\LRp{\bm{f}^* - \bm{f}(\bm{u})} \bm{w}\right|_{-1}^1 = 0, \qquad \forall \bm{w}\in V_h.
+\]
+Here, $\bm{f}^*$ is a numerical flux used for the weak imposition of boundary or continuity conditions.
+
+We derive a DG-SEM approximation by evaluating the above integrals using GLL quadrature.  We furthermore assume that the solution is approximated using a nodal basis at GLL points 
+\[
+\bm{u}(\bm{x},t) = \sum_{j=1}^{N+1} \bm{u}_j(t)\ell_j(x).  
+\]
+The resulting semi-discrete system yields a nodal collocation scheme for $\bm{u}_i$ 
 \[
@@
 \]
-Here, $\bm{f}^*$ is a numerical flux used for the weak imposition of boundary or continuity conditions \cite{hesthaven2007nodal}.
-This method can be derived from an appropriate DG formulation under a specific choice of quadrature.
 
@@
 
+\subsection{Notation}
+
+We assume that the domain $\Omega \in \mathbb{R}^d$ is exactly triangulated by a mesh which consists of the union of non-overlapping elements $D^k$.
+We further assume that each element $D^k$ is the image of a reference element $\widehat{D}$ under the local elemental mapping
+\[
+\bm{x}^k = \bm{\Phi}^k \widehat{\bm{x}},
+\]
+where $\bm{x}^k$ denote physical coordinates on $D^k$ and $\widehat{\bm{x}}$ denote coordinates on the reference element.
+Finally, we identify $J$ as the determinant of the Jacobian of $\bm{\Phi}^k$, and refer to it as the Jacobian for the remainder of this work.
+
+We will approximate solution components over each element $D^k$ from an approximation space $V_h\LRp{D^k}$, which we define as the composition of the mapping $\bm{\Phi}^k$ and a reference approximation space $V_h\LRp{\widehat{D}}$
+\[
+V_h\LRp{D^k} = \bm{\Phi}^k \circ V_h\LRp{\widehat{D}}.
+\]
+The global approximation space $V_h\LRp{\Omega}$ is then defined as the direct sum of local approximation spaces
+\[
+V_h\LRp{\Omega} = \bigoplus_{D^k}V_h\LRp{D^k}.  
+\]
+We assume that $V_h(D^k) \subset L^2(D^k)$; in other words, that all elements of each local approximation space are square integrable.
+We also define the local $L^2$ projection operator $\Pi_N: L^2(D^k)\rightarrow V_h(D^k)$ such that
+\[
+\LRp{\Pi_N u,v}_{D^k} = \LRp{u,v}_{D^k}, \qquad \forall v\in V_h\LRp{D^k}.  
+\]
+Because the global approximation space $V_h\LRp{\Omega}$ is the direct sum of independent local approximation spaces, the global $L^2$ projection is simply the composition of local $L^2$ projections.
+
+%\note{We define the quadrature-based $L^2$ projection $\Pi_N$ as the $L^2$ projection where integrals are evaluated using some (inexact) quadrature.
+We also define the discrete approximation space $V_h$, and assume that the quadrature is sufficiently accurate such that integration by parts holds for any two elements of $V_h$.
+Finally, we assume that the discrete derivative (to be defined) maps from $V_h$ to $V_h$.
+}
+
+Finally, we define the jump and average of discontinuous functions across element interfaces.
+Let $D^{k,+}$ denote the neighboring element of across a face $f$ of $D^k$, and let $u^+,u^-$ denote the values of $u$ on $D^{k,+}$ and $D^k$, respectively.
+The jump of $u$ across $f$ is then defined as
+\[
+\jump{u} = u^+ - u^-, \qquad \avg{u} = \frac{u^+ + u^-}{2}.
+\]
+The jump and average of vector fields are defined component-wise using the jumps and averages of components.
+
+%\note{Specify assumption that differentiation operators $\bm{D}:V_h\rightarrow V_h$}
+
 \subsection{A continuous interpretation of flux differencing}
 
-We define the quadrature-based $L^2$ projection $\Pi_N$ as the $L^2$ projection where integrals are evaluated using some (inexact) quadrature.
-We also define the discrete approximation space $V_h$, and assume that the quadrature is sufficiently accurate such that integration by parts holds for any two elements of $V_h$.
-Finally, we assume that the discrete derivative (to be defined) maps from $V_h$ to $V_h$.
+In this section, we offer a continuous interpretation of flux differencing \cite{fisher2013high,gassner2017br1,chen2017entropy}, which also encompasses the split form methodology proposed in \cite{gassner2016split}.
+This interpretation will guide the construction of entropy conservative DG schemes.
 
-The split-form methodology of Gassner et al.\ can be interpreted as the following: let $f_S(u_L,u_R)$ be a symmetric and consistent two-point flux (i.e.
-$f_S(u,u) = f(u)$).
-Split forms can be rewritten at the continuous level as replacing the flux derivative with differentiation along one argument of $f_S(u_L,u_R)$, then evaluation of the other argument at $u$
+Assume that $\bm{f}_S(\bm{u}_L,\bm{u}_R)$ is the symmetric and consistent two-point flux defined in (\ref{def:tadmor}).  Consider the bilinear function
 \[
-\pd{f(u)}{x} \Longrightarrow 2\left.\pd{f_S(u(x),u(y))}{x}\right|_{y=x}.
+\bm{f}_S\LRp{\bm{u}(\bm{x}),\bm{u}(\bm{y})}, \qquad \bm{x}, \bm{y} \in \mathbb{R}^d.  
 \]
-This allows for the recovery of different split-forms.  For example, for Burgers equation, choosing $f_S(u_L,u_R) = (u_L^2 + u_Lu_R + u_R^2)/6$ gives
+Flux differencing can be interpreted as differentiating $\bm{f}_S\LRp{\bm{u}(\bm{x}),\bm{u}(\bm{y})}$ along one coordinate $\bm{x}_k$, then evaluating the resulting derivative with $\bm{y}=\bm{x}$.
+Since the two-point flux is consistent, this results in the following approximation of the flux derivative
+\begin{equation}
+\pd{\bm{f}(\bm{u}(\bm{x}))}{\bm{x}_k} \Longrightarrow 2\left.\pd{\bm{f}_S\LRp{\bm{u}(\bm{x}),\bm{u}(\bm{y})}}{\bm{x}_k}\right|_{\bm{y}=\bm{x}}.
+\label{eq:fluxdiff}
+\end{equation}
+
+\subsubsection{Illustration using Burgers' equation}
+
+We note that flux differencing allows for the recovery of split formulations.  We illustrate this using the one-dimensional Burgers equation, which is given in conservative form as
 \[
-2\left.\pd{f_S(u(x),u(y))}{x}\right|_{y=x} = \frac{1}{3}\left.\pd{\LRp{u(x)^2 + u(x)u(y) + u(y)^2}}{x}\right|_{y=x} = \frac{1}{3}\LRp{\pd{u^2}{x} + u\pd{u}{x}}
+\pd{u}{t} + \pd{f(u)}{x} = 0, \qquad f(u) = \frac{1}{2}u^2.
+\]
+It is well known that DG formulations based on the split form
+\[
+\pd{u}{t} + \frac{1}{3}\LRp{\pd{u^2}{x} + u\pd{u}{x}} = 0
+\]
+can be shown to be discretely energy stable under appropriately imposed boundary conditions \cite{gassner2013skew, ranocha2017extended}.
+This form may be recovered using flux differencing by choosing the two-point flux
+\[
+f_S(u_L,u_R) = \frac{1}{6}(u_L^2 + u_Lu_R + u_R^2).
+\]
+Then, applying (\ref{eq:fluxdiff}) yields
+\[
+2\left.\pd{f_S(u(x),u(y))}{x}\right|_{y=x} = \frac{1}{3}\left.\pd{\LRp{u(x)^2 + u(x)u(y) + u(y)^2}}{x}\right|_{y=x} = \frac{1}{3}\LRp{\pd{u^2}{x} + u\pd{u}{x}},
 \]
 due to the fact that $\pd{u(y)^2}{x} = 0$.
-The resulting form of Burgers equation appears in the intermediate steps of energy stability proofs, and is the basis for discretely stable DG methods.
-When applied discretely (using quadrature-based projections), the resulting right hand side for Burgers' equation becomes
+
+Both the conservative and split forms of Burgers' equation are equivalent at the continuous level (assuming continuity of $u$ and exact differentiation); however, this is no longer true at the discrete level.
+Suppose now that the solution is approximated by $u \in V_h(\Omega)$ on a single element.
+Since $f(u) = u^2/2 \not\in V_h(\Omega)$, it is not possible to differentiate $f(u)$ exactly without prior knowledge of the form of the nonlinearity.
+Instead, it is more common to first project $f(u)$ to $V_h$, then differentiate, such that the discrete scheme is equivalent to solving for $u \in V_h(\Omega)$ such that
 \[
-\LRp{\frac{1}{3}\LRp{\pd{\LRp{\Pi_N u^2}}{x} + \Pi_N \LRp{u\pd{u}{x}}}, v}.
+\int_{-1}^1 \LRp{\pd{u}{t} + \pd{\Pi_N f(u)}{x}}w = 0, \qquad \forall w \in V_h(\Omega).  
 \]
-Using the fact integration by parts holds when considering two elements of $V_h$, and that $\LRp{\Pi_N u^2, v} = \LRp{u^2,v}$ for any $v \in V_h$, it can be shown that the volume terms cancel out for $v = u$, and we're left with boundary terms (which is a prerequisite of energy/entropy stability).
+This scheme is energy stable only for projections which are computed using a quadrature of degree $3N-1$, which is sufficiently accurate to integrate the nonlinear DG weak form exactly.
+
+In comparison, when applying (\ref{eq:fluxdiff}) using $\pd{\Pi_N (\cdot)}{x}$, the resulting right hand side for Burgers' equation becomes 
+\[
+\int_{-1}^1\LRp{\pd{u}{t} + \frac{1}{3}\LRp{\pd{\LRp{\Pi_N u^2}}{x} + \Pi_N \LRp{u\pd{u}{x}}}}w = 0, \qquad \forall w \in V_h(\Omega).  
+\]
+Using the fact integration by parts holds when considering two elements of $V_h$, and that $\LRp{\Pi_N u^2, v} = \LRp{u^2,v}$ for any $v \in V_h$, it can be shown that the volume terms cancel out for $v = u$, leaving only boundary terms (which is a prerequisite of energy/entropy stability).
+
+\note{De-aliasing by increasing the degree of quadrature hits a ``bottleneck'': in the end, you are only differentiating a polynomial of degree $N$ \cite{kirby2003aliasing}.}
 
@@
 As such, replacing $\pd{}{x}$ by any differential operator $D$ such that $D 1 = 0$ will recover a similar formulation.
+
+\subsubsection{Recovering the algebraic form of flux differencing}
 
@@
 
-\subsection{Discretely entropy conservative general DG methods on a single domain}
+When using nodal collocation using either GLL or GQ points, \verb+Vq,Pq+ are the identity matrix and we recover the algebraic form of flux differencing described in \cite{gassner2016split}.
+
+\note{Introduce definition of $\bm{f}(\bm{u}_q)$ as $f(\bm{u}_q)$.}
+
+\subsection{Global DG differentiation operators}
+
+The construction of entropy conservative DG methods
+
+Global DG methods \cite{di2011mathematical}
+
+We adapt the construction of local differentiation operators by Chen and Shu \cite{chen2017entropy} to construct a global DG differentiation operator
+\begin{equation}
+\LRp{D^x_h u,v} = \sum_k \LRp{\pd{ \Pi_N u}{x},v}_{D^k} + \frac{1}{2}{\LRa{{u_P - \Pi_N u_M}, v\bm{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{{u_M - \Pi_Nu_M },\Pi_Nv \bm{n}_x}_{\partial D^k}}.
+\label{eq:dgd}
+\end{equation}
+Integrating by parts then yields the skew-symmetric form of $\bm{D}^x_h$
+\[
+\LRp{D^x_h u,v} =\frac{1}{2} \sum_k \LRp{\pd{ \Pi_N u}{x},v}_{D^k} - \LRp{u,\pd{\Pi_N v}{x}}_{D^k} + \LRa{\LRp{u_P - \Pi_N u_M}\bm{n}_x, v} + \LRa{{u_M }\bm{n}_x,\Pi_Nv}.
+\]
+Taking $v = u_M$ and cancelling $u_P u_M \bm{n}_x$ terms shows that the differentiation operator has the global SBP property and is skew-symmetric  w.r.t.\ a discrete $L^2$ inner product.
+The coupling requires only traces of neighboring elements.
+
+It's currently unclear how to distinguish whether different operators are high order accurate.
+
+\begin{remark}
+It is possible to define a simpler global DG differentiation operator based on the composition of a weak polynomial differentiation operator and the $L^2$ projection.
+Let $\bm{D}^x_h: V_h\rightarrow V_h$ be defined as
+\[
+\sum_k \LRp{D^x_h u,v}_{D^k} = \sum_k \LRp{\pd{u}{x},v}_{D^k} + \frac{1}{2}\LRa{\jump{u},v\bm{n}_x}_{\partial D^k}, \qquad \forall v \in V_h.
+\]
+Then, it is straightforward to show that $D^x_h \Pi_N$ possesses a global SBP property with respect to the $L^2$ norm, such that
+\[
+\LRp{D^x_h\Pi_N u, v}_{\Omega} = \LRa{\Pi_N u,v}_{\partial \Omega} - \LRp{D^x_h\Pi_N v, u}_{\Omega}, \qquad \forall u,v\in \L.
+\]
+Numerical results using this global differentiation operator result in slightly more accurate results than when using the DG operator (\ref{eq:dgd}).
+Unfortunately, compared to  (\ref{eq:dgd}), this construction results in a wide stencil involving all degrees of freedom on neighboring elements due to the presence of $\jump{u}$ and the $L^2$ projection.
+\end{remark}
+
+\subsection{Discretely entropy conservative general DG methods}
+
+We introduce 
+\[
+\int_{D^k}\pd{\bm{u}}{t} \bm{w} + \left.\LRp{D^x_h\bm{f}_S\LRp{\bm{u}(\Pi_N\bm{v}(\bm{x})),\bm{u}(\Pi_N\bm{v}(\bm{y}))}}\right|_{\bm{y}=\bm{x}}\bm{w} = 0, \qquad \forall \bm{w}\in V_h,
+\]
+where the global operator $D^x_h$ is understood to be applied to each component of $\bm{f}_S\LRp{\bm{u}(x),\bm{u}(y)}$.
 
@@
 \[
-u(x,t) = \sum_{j=1}^{N_p} u_j(t) \phi_j(x), \qquad \pd{u}{t} \in P^N.
+\bm{u}(x,t) = \sum_{j=1}^{N_p} \bm{u}_j(t) \phi_j(x), \qquad \pd{\bm{u}}{t} \in P^N.
 \]
-Thus, testing with $\Pi_N v$ yields
+Thus, testing with $\Pi_N \bm{v}$ over each element $D^k$ yields  
 \[
-\int_{D^k} \pd{u}{t}\Pi_N v = \int_{D^k} \pd{u}{t} v   
+\int_{D^k} \pd{\bm{u}}{t}\Pi_N \bm{v} = \int_{D^k} \pd{\bm{u}}{t}\bm{v} = \int_{D^k} \pd{U(\bm{u})}{t}.  
 \]
-which yields a pointwise relationship which can be manipulated into the time derivative of entropy.
+In matrix notation, this can be shown by multiplying with $\LRp{\bm{P}_q\bm{v}_q}^T$ 
+\[
+\LRp{\bm{P}_q\bm{v}_q}^T\bm{M}\pd{\bm{u}}{t} = \bm{v}_q^T\bm{W}\bm{V}_q\bm{M}^{-1}\bm{M}\pd{\bm{u}}{t} = \bm{1}^T \bm{W} {\rm diag}(\bm{v}_q) \pd{\bm{u}_q}{t} = \bm{1}^T \bm{W} \pd{U(\bm{u}_q)}{t},
+\]
+where we have assumed continuity in time, and have used that $\bm{W}$ is diagonal.
+Summing contributions from each element $D^k$ yields the time derivative of the entropy over the entire domain $\Omega$.
 
-The right-hand side is more difficult.
+%which yields a pointwise relationship which can be manipulated into the time derivative of entropy.
+
 For collocation DG methods, interpolation and quadrature-based $L^2$ projection are identical.
@@
 
-\subsection{SBP-SAT terms by global DG differentiation operators}
+Not easy to deal with SBP-SAT terms for more complex systems \cite{ranocha2017extended,ranocha2017comparison}.
 
-We can define a global discontinuous Galerkin operator $\bm{D}^x_h: P^N\rightarrow P^N$ such that
+We summarize this analysis in the following theorem:
+\begin{theorem}
+\note{Blah is entropy conservative.  Finish up proof + statement of entropy conservation.}
+\end{theorem}
+
+\begin{remark}
+The analysis in this section has focused on the construction of entropy conservative schemes.
+However, for problems with shocks, entropy is not conserved and should be dissipated away.
+To this end, one can construct entropy stable schemes by adding additional stabilization terms (for example, by adding Lax-Friedrichs penalization or matrix dissipation terms \cite{chandrashekar2013kinetic,winters2017uniquely}).
+\end{remark}
+
+\subsubsection{On implementation}
+
+Let $V_h^f$ denote the trace space of the approximation space $V_h$ 
 \[
-\sum_k \LRp{D^x_h u,v}_{D^k} = \sum_k \LRp{\pd{u}{x},v}_{D^k} + \LRa{\frac{1}{2}\jump{u}\bm{n}_x,v}_{\partial D^k}.
+V_h^f = \LRc{ \left.u\right|_{\partial D^k}, \quad u \in V_h}.  
 \]
-Then, composing $D^x_h \Pi_N$ gives global skew-symmetry $\LRp{D^x_h\Pi_N u, v}_{\L} = \LRp{- D^x_h\Pi_N v, u}_{\L}$ for all $u,v\in \L$.
-The downside to this simple construction is that the stencil of the DG scheme now incorporates all degrees of freedom on neighboring elements due to the composition with the projection.
+We introduce the lifting operator ${L}: V_h^f\rightarrow V_h$.  For $u_f \in V^f_h$, ${L}$ is defined as 
+\[
+\LRp{{L} u_f,v}_{D^k} = \LRa{u_f,v}_{\partial D^k}, \qquad \forall v\in V_h.
+\]
 
-We adapt the construction of local differentiation operators by Shi and Shu to construct a global DG differentiation operator.  
+\subsubsection{Example: Burgers equation}
+
+We illustrate the application of the global flux differencing approach to Burgers' equation using the energy conserving flux
 \[
-2\LRp{D^x_h u,v} = \sum_k \LRp{2\pd{ \Pi_N u}{x},v}_{D^k} + \LRa{\LRp{u_P - \Pi_N u_M}\bm{n}_x, v} + \LRa{\LRp{u_M - \Pi_Nu_M }\bm{n}_x,\Pi_Nv}.
+f_S(u_L,u_R) = \frac{1}{6}\LRp{u_L^2 + u_Lu_R u_R^2}.  
 \]
-In skew-symmetric form, this gives 
+We will derive 
 \[
-2\LRp{D^x_h u,v} = \sum_k \LRp{\pd{ \Pi_N u}{x},v}_{D^k} - \LRp{u,\pd{\Pi_N v}{x}}_{D^k} + \LRa{\LRp{u_P - \Pi_N u_M}\bm{n}_x, v} + \LRa{{u_M }\bm{n}_x,\Pi_Nv}.
+\LRp{\left.\bm{D}^x_h f_S(u(x),u(y))\right|_{y=x}, w} = 
 \]
-Taking $v = u_M$ and cancelling $u_P u_M \bm{n}_x$ terms shows that the differentiation operator has the global SBP property and is skew-symmetric  w.r.t.\ a discrete $L^2$ inner product.
-The coupling requires only traces of neighboring elements.
 
-It's currently unclear how to distinguish whether different operators are high order accurate.
+We note that the global flux differencing approach with the DG operator (\ref{eq:dgd}) recovers the formulation and correction terms introduced in \cite{ranocha2017extended}.
 
@@
 
+\note{We illustrate the entropy conservation and accuracy of the proposed schemes for three equations.  These gradually increase in complexity.  }
+
 \subsection{Burgers}
+
+We begin by computing $L^2$ errors and rates of convergence for smooth solutions of Burgers' equation.
 
@@
 
-We describe in this section the shallow water equations in one dimension.
-We use this model problem to illustrate the proposed method because (unlike Burgers' equation) it highlights some non-trivial details in proving entropy stability associated with the nonlinear equations of compressible flow.
-
-The one-dimensional shallow water equations are 
+We now move on to the one-dimensional shallow water equations, which consist of a conservation equation and momentum equation
 \begin{align*}
@@
 where $h$ is the water height, $v$ is the velocity, $g$ is the gravitational constant, and $b$ is bottom topography.
-
-The entropy associated with this system is the total energy $e = k + p$, where 
+The entropy associated with this system is the total energy $e = k + p$, where $k, p$ are the kinetic and potential energies
 \[
@@
 \]
-Testing with the entropy variables 
+The entropy variables can be derived from the entropy $e$
 \[
-q_1 = \pd{e}{h} = g(h+b)-v^2/2, \qquad q_2 = \pd{e}{hv} = v
+q_1 = \pd{e}{h} = g(h+b)-v^2/2, \qquad q_2 = \pd{e}{hv} = v.
 \]
-recovers a statement of conservation of energy
+Multiplying the continuity equation by $q_1$ and the momentum equation by $q_2$ recovers a statement of conservation of energy
 \[
-\pd{e}{t} + \pd{}{x}\LRp{hv^3/2 + g(hv)(h+b)} = 0.  
+\pd{e}{t} + \pd{}{x}\LRp{\frac{hv^3}{2} + g(hv)(h+b)} = 0.  
 \]
+Like the Burgers' equation, the shallow water equations are stable under an appropriate split formulation\cite{gassner2016well, chen2017entropy}.
+However, unlike Burgers' equation, the shallow water equations are not entropy stable with respect to the square entropy.
+Additionally, a key difference between the shallow water equations and Burgers' equation is that, when discretizing with respect to the conservative variables $h, hv$, the entropy variables $v_1,v_2$ are not contained within the space $V_h$.
 
-This conservation of energy does not hold at the discrete level for two reasons.
+%This conservation of energy does not hold at the discrete level for two reasons.
 The first reason is because, under inexact quadrature, calculus identities such as the chain or product rule do not hold, and naive spatial formulations can not be manipulated into a form which is amenable to a proof of discrete entropy stability.
 This issue can be remedied by the use of a split formulation (as shown in Gassner et al.) which is provably entropy stable under inexact quadrature.
-
-The second more subtle reason why entropy conservation does not hold at the discrete level is because the entropy variables are not contained within the test space.
+%
+%The second more subtle reason why entropy conservation does not hold at the discrete level is because the entropy variables are not contained within the test space.
 Moreover, the spatial part of the proposed discretization requires one discrete definition of the entropy variables for entropy stability, while the temporal part of the discretization requires another discrete definition of the entropy variables.
@@
 
+We now test the proposed entropy conservative schemes on the one-dimensional compressible Euler equations, which correspond to the inviscid limit of the compressible Navier-Stokes equations
+\begin{align*}
+\pd{\rho}{t} + \pd{\LRp{\rho u}}{x} &= 0,\\
+\pd{\rho u}{t} + \pd{\LRp{\rho u^2 + p }}{x} &= 0,\\
+\pd{E}{t} + \pd{\LRp{u(E+p)}}{x} &= 0.
+\end{align*}
+We assume an ideal gas, such that the pressure satisfies the constitutive relation $p = (\gamma-1)\LRp{E - \frac{1}{2}\rho u^2}$, where $\gamma = 1.4$ is the ratio of specific heat for a diatomic gas.
+
+The choice of convex entropy for the Euler equations is non-unique \cite{harten1983symmetric}.
+However, a unique entropy can be chosen by restricting to choices of entropy variables which symmetrize the viscous heat conduction term in the compressible Navier-Stokes equations \cite{hughes1986new}.
+The entropy is given by
+\[
+U(\bm{u}) = -\frac{\rho s}{\gamma-1},
+\]
+where $s = \log\LRp{\frac{p}{\rho^\gamma}}$ is the physical specific entropy.  The entropy variables under this choice of entropy are then
+\begin{align*}
+q_1 = \frac{\gamma-s}{\gamma-1} - \frac{\rho u^2}{2p}, \qquad q_2 = \frac{\rho u}{p}, \qquad q_3 = -\frac{\rho}{p}.
+\end{align*}
+The reverse mapping is given by 
+\[
+d
+\]
+
+\note{Add Chandreshekar's fluxes \cite{chandrashekar2013kinetic}.}
+
+Unlike the shallow water equations, these entropy conservative fluxes do not correspond to stable split formulations of the Euler equations.
+Thus, the compressible Euler equations serve as a test of the flux differencing formulation, as entropy stability cannot be achieved through skew-symmetry.
+
 \section{Conclusions}
@@
 \section{Acknowledgments}
+
+The author thanks Andrew Winters for several informative discussions.  Jesse Chan is supported by NSF awards DMS-1719818 and DMS-1712639.
 

### 2017-07-25 docs/splitform/main.tex
updating paper

@@
 \newtheorem{theorem}{Theorem}
+\theoremstyle{assumption}
+\newtheorem{assumption}{Assumption}
 %\usepackage{thmtools}
@@
 \author{Jesse Chan}
-\title{On discrete entropy conservation for a more general class of discontinuous Galerkin methods}
+\title{On a general class of discretely entropy conservative discontinuous Galerkin methods}
 
@@
 High order methods based on diagonal-norm summation by parts operators can be shown to semi-discretely conserve entropy for nonlinear systems of hyperbolic PDEs \cite{fisher2013high,carpenter2014entropy}.
-These include discontinuous Galerkin spectral element methods (DG-SEM) \cite{gassner2016split,gassner2016well,wintermeyer2017entropy, chen2017entropy} based on mass lumping and nodal collocation.
-In this work, we describe how to extend discretely entropy conservative schemes for a more general class of DG methods using quadrature-based projections and global DG differentiation operators.
-The construct of these methods is independent of the choice of basis and accuracy of quadrature rule used, and includes DG methods involving generalized SBP operators without boundary nodes and certain dense-norm SBP operators.
-Numerical experiments confirm the stability and high order accuracy of the proposed methods for the Burgers, shallow water, and compressible Euler equations.
+These methods can be interpreted as mass lumped discontinuous Galerkin spectral element methods (DG-SEM) \cite{gassner2016split,gassner2016well,wintermeyer2017entropy, chen2017entropy}.
+In this work, we describe how to extend discretely entropy conservative schemes for a more general class of DG methods using quadrature-based $L^2$ projections, flux differencing, and global DG differentiation operators.
+This approach also recovers existing methods for Burgers' equation involving generalized SBP operators without boundary nodes and dense norm SBP operators.
+Numerical experiments confirm the stability and high order accuracy of the proposed methods for the shallow water and compressible Euler equations.
 \end{abstract}
@@
 
-\note{High order methods are good boilerplate}
-
-\note{High order methods are not stable for nonlinear problems.  Stability of discretizations is a hot topic: artificial viscosity, polynomial de-aliasing,  \cite{karniadakis2013spectral}}
-
-\cite{tadmor1987numerical, tadmor2003entropy}
-
-Hughes entropy variables \cite{hughes1986new}
-
-\note{Specify that entropy stability has been shown for mass lumped collocation DG discretizations.
-Requirements are a diagonal norm SBP operators (or diagonal mass matrix) and collocated differentiation.
-}  \cite{gassner2016split,gassner2016well,wintermeyer2017entropy}
-\note{The construction of such operators on non-tensor product elements (such as simplices or pyramids) is non-trivial.} \cite{hicken2016multidimensional, chen2017entropy}.
-\note{Constructing such operators for non-polynomial spaces (such as splines) is also difficult within the current framework.}
-
-While DG-SEM based on flux differencing is entropy conservative, this is tied to a specific choice of basis and quadrature.
-Furthermore, GLL quadratures and diagonal-norm SBP operators do not naturally generalize to triangles or tetrahedra \cite{helenbrook2009existence}.
+Numerical simulations in engineering increasingly require higher accuracy without sacrificing computational efficiency.
+High order methods are more accurate than low order methods per degree of freedom for sufficiently regular solutions, and provide one avenue towards improving fidelity in numerical simulations while maintaining reasonable computational costs.
+High order methods which can accomodate unstructured meshes are particularly well-suited for problems with complex geometries.
+Among such methods, high order discontinuous Galerkin (DG) methods are particularly well-suited to the solution of time-dependent hyperbolic problems on modern computing architectures \cite{hesthaven2007nodal, klockner2009nodal}.
+
+The accuracy observed for high order methods can be attributed in part to their low numerical dissipation and dispersion properties \cite{ainsworth2004dispersive}.
+While low numerical dissipation is advantageous for problems in linear wave propagation \cite{komatitsch1998spectral}, the high numerical dissipation present in low order schemes adds a stabilizing effect for nonlinear hyperbolic problems \cite{wang2013high}.
+As a result, high order schemes for nonlinear conservation laws tend to require additional stabilization, including (but not limited to) filtering, slope limiting, artificial viscosity, and polynomial de-aliasing \cite{persson2006sub, krivodonova2007limiters, karniadakis2013spectral, hesthaven2007nodal, kirby2003aliasing}.
+Moreover, stabilized numerical methods can still fail, requiring user intervention or heuristic modifications to achieve non-divergent solutions.
+
+For linear wave propagation problems, numerical methods can be shown to be semi-discretely energy stable, even in the presence of curvilinear coordinates or variable coefficients \cite{warburton2013low, chan2016weight1, chan2016weight2, chan2017weight}.
+This semi-discrete stability can be used to show that, under a stable timestep restriction (CFL), discrete solutions obtained by explicit time-stepping do not suffer from un-physical growth in time.
+For nonlinear systems of conservation laws, typical methods for fluid dynamics do not admit formal proofs of stability.
+However, progress has been made in designing provably discretely stable methods, beginning with the introduction of entropy conservative and entropy stable finite volume schemes by Tadmor \cite{tadmor1987numerical, tadmor2003entropy}.
+These schemes rely on a specific entropy conservative flux which satisfies a condition involving the entropy variables and entropy potential, and have been extended to low order finite volume methods on unstructured grids \cite{ray2016entropy}.
+
+The construction of entropy stable high order schemes was done much more recently for the compressible Euler and Navier-Stokes equations in \cite{fisher2013high, carpenter2014entropy} based on a spectral collocation approach on tensor product elements, which can be interpreted as a mass-lumped DG spectral element (DG-SEM) scheme.
+The proof of entropy stability relies on the presence of a diagonal mass matrix, as well as the summation by parts property \cite{gassner2013skew}.
+Similar entropy-stable schemes were constructed for the shallow water and MHD equations \cite{derigs2016novel, gassner2016well, wintermeyer2017entropy}.
+More recently, entropy stable schemes were extended to unstructured triangular meshes in \cite{chen2017entropy} using the split-form methodology of \cite{gassner2016split} and multi-dimensional SBP operators \cite{hicken2016multidimensional}.
+
+To the authors knowledge, the construction of entropy conservative and entropy stable schemes have relied on the use of lumped diagonal mass matrices and nodal collocation (diagonal norm SBP operators).
+Entropy stable high order finite element and DG methods have been proposed \cite{hughes1986new}, but the proofs are often given at the continuous level, relying on exact integration or the chain rule, neither of which holds at the discrete level.
+Thus, while DG-SEM based on flux differencing is entropy conservative, this is tied to a specific choice of basis and quadrature.
 Diagonal-norm SBP operators can be constructed for triangles and tetrahedra \cite{chin1999higher, cohen2001higher, hicken2016multidimensional, chen2017entropy}; however, the number of nodal points for such operators is greater than the dimension of the natural polynomial approximation space.
 Furthermore, appropriate point sets have only been constructed for $N \leq 4$ in three dimensions \cite{zhebel2014comparison}.
+%Furthermore, diagonal-norm SBP operators do not naturally generalize to triangles or tetrahedra \cite{helenbrook2009existence}.
+
+This work focuses on the construction of entropy conservative high order DG schemes for systems of conservation laws.
+In order to generalize beyond collocation DG and diagonal norm SBP operators, we will consider dense (non-lumped) mass matrices and quadrature rules of arbitrarily high accuracy (over-integration).
+The outline of the paper is as follows: Section~\ref{sec:intro} will briefly review the construction of entropy conservative collocation methods on a single domain.
+Section~\ref{sec:ecdg} will describe how to construct analogous entropy conservative methods on single domains.
+Section~\ref{sec:ecdg2} will discuss extensions to multiple domains, including comparisons of different coupling terms and entropy stable fluxes.
+We note that, for simplicity and clarity of presentation, the proposed methods and proofs are given in one dimension.
+However, the extension to multiple dimensions (including simplicial elements) is straightforward and will be described in an upcoming manuscript.
+Finally, Section~\ref{sec:num} presents numerical results which verify the high order accuracy and discrete entropy conservation of the proposed methods.
+
+%\note{The construction of such operators on non-tensor product elements (such as simplices or pyramids) is non-trivial.} \cite{hicken2016multidimensional, chen2017entropy}.
+\note{Constructing such operators for non-polynomial spaces (such as splines) is also difficult within the current framework.}
 
 \section{Entropy conservative diagonal norm collocation DG methods}
-
-\note{Specify what we mean by diagonal norm collocation DG method.}
+\label{sec:intro}
+
+We will begin by reviewing continuous entropy theory, as well as existing high order methods which are provably entropy conservative at the discrete level.
+These methods were introduced in \cite{fisher2013high, carpenter2014entropy} for tensor product elements, though we adopt methods of proof introduced more recently in \cite{gassner2017br1, chen2017entropy}.
+
+%\note{Specify what we mean by diagonal norm collocation DG method.}
 
@@
 
-We consider systems of nonlinear conservation laws in $\mathbb{R}^d$ with $n$ variables
+%\note{Specify that all analysis is done in 1D but is extendable to higher dimensions!!  For simplicity and clarity of presentation, we will begin by focusing on the case $d = 1$.  }
+
+We consider systems of nonlinear conservation laws in one dimension with $n$ variables
 \begin{align}
-\pd{\bm{u}}{t} + \sum_{k=1}^d \pd{\bm{f_k(\bm{u})}}{\bm{x}_k} &= 0, \qquad \bm{u}(\bm{x},t) = (u_1(\bm{x},t),\ldots,u_n(\bm{x},t)).
+%\pd{\bm{u}}{t} + \sum_{k=1}^d \pd{\bm{f_k(\bm{u})}}{x} &= 0, \qquad \bm{u}(x,t) = (u_1(x,t),\ldots,u_n(x,t)).
+\pd{\bm{u}}{t} + \pd{\bm{f(\bm{u})}}{x} &= 0, \qquad \bm{u}(x,t) = (u_1(x,t),\ldots,u_n(x,t)).
 \label{eq:pde}
 \end{align}
-where the fluxes $f_k(\bm{u})$ are smooth functions of the vector of conservative variables $\bm{u}(\bm{x},t)$.
-The Jacobian matrix $\bm{A}_k(\bm{u})$ is defined entrywise as
-\[
-\LRp{\bm{A}_k(\bm{u})}_{ij} = \LRp{\pd{\bm{f}_k(\bm{u})}{\bm{u}_j}}_i.
+where the fluxes $f(\bm{u})$ are smooth functions of the vector of conservative variables $\bm{u}({x},t)$.%$\bm{u}(x,t)$.
+The Jacobian matrix $\bm{A}(\bm{u})$ is defined entrywise as
+\[
+%\LRp{\bm{A}_k(\bm{u})}_{ij} = \LRp{\pd{\bm{f}_k(\bm{u})}{\bm{u}_j}}_i.
+\LRp{\bm{A}(\bm{u})}_{ij} = \LRp{\pd{\bm{f}(\bm{u})}{\bm{u}_j}}_i.
 \]
@@
 \begin{equation}
-U''(\bm{u})\bm{A}_k(\bm{u}) = \LRp{U''(\bm{u}) \bm{A}_k(\bm{u})}^T.
+%U''(\bm{u})\bm{A}_k(\bm{u}) = \LRp{U''(\bm{u}) \bm{A}_k(\bm{u})}^T.
+U''(\bm{u})\bm{A}(\bm{u}) = \LRp{U''(\bm{u}) \bm{A}(\bm{u})}^T.
 \label{eq:entropysym}
@@
 
-It can be shown (see, for example, \cite{mock1980systems}) that (\ref{eq:entropysym}) is equivalent to the existence of an entropy flux functions $F_k(\bm{u})$ such that
-\[
-\bm{v}^T \pd{\bm{f}_k}{\bm{u}} = \pd{F_k(\bm{u})}{\bm{u}}^T.
+It can be shown (see, for example, \cite{mock1980systems}) that (\ref{eq:entropysym}) is equivalent to the existence of an entropy flux functions $F(\bm{u})$ such that
+\[
+\bm{v}^T \pd{\bm{f}}{\bm{u}} = \pd{F(\bm{u})}{\bm{u}}^T.
 \]
@@
 \[
-\psi_k(\bm{v}) = \bm{v}^T\bm{f}_k(\bm{u}(\bm{v})) - F_k(\bm{u}(\bm{v})), \qquad \psi_k'(\bm{v}) = \bm{f}_k(\bm{u}(\bm{v})).
+%\psi_k(\bm{v}) = \bm{v}^T\bm{f}_k(\bm{u}(\bm{v})) - F_k(\bm{u}(\bm{v})), \qquad \psi_k'(\bm{v}) = \bm{f}_k(\bm{u}(\bm{v})).
+\psi(\bm{v}) = \bm{v}^T\bm{f}(\bm{u}(\bm{v})) - F(\bm{u}(\bm{v})), \qquad \psi'(\bm{v}) = \bm{f}(\bm{u}(\bm{v})).
 \]
@@
 \[
-\pd{U(\bm{u})}{t} + \sum_{k=1}^d \pd{F_k(\bm{u})}{\bm{x}_k} = 0.
+\pd{U(\bm{u})}{t} + \pd{F(\bm{u})}{{x}} = 0.
 \]
@@
 \[
-\pd{U(\bm{u})}{t} + \sum_{k=1}^d \pd{F_k(\bm{u})}{\bm{x}_k} \leq 0.  
-\]
-%Integration over $\mathbb{R}^d$ then shows that the entropy of the solution decreases in time.  
-When $d = 1$, integrating the entropy inequality over $[-1,1]$ and using the definition of the entropy potential yields
+\pd{U(\bm{u})}{t} + \pd{F(\bm{u})}{{x}} \leq 0.  
+\]
+
+We assume now that the domain is the interval $[-1,1]$.  Integrating the entropy inequality over this interval and using the definition of the entropy potential then yields
 \begin{equation}
-\int_{-1}^1 \pd{U(\bm{u})}{t} + \left.\LRp{\bm{v}^T\bm{f}_k(\bm{u}(\bm{v}))-\psi(\bm{v})}\right|_{-1}^1 \leq 0.  
+\int_{-1}^1 \pd{U(\bm{u})}{t} + \left.\LRp{\bm{v}^T\bm{f}(\bm{u}(\bm{v}))-\psi(\bm{v})}\right|_{-1}^1 \leq 0.  
 \label{eq:consentropy}
 \end{equation}
-
-%\note{Wave equation example}
+The focus of this work is the construction of methods which satisfy a discrete analogue of this entropy inequality.
+We begin by reviewing existing discretely entropy conservative methods based on collocation and diagonal norm summation by parts (SBP) operators.
 
@@
 
-Let $\widehat{x}_i \in [-1,1]$ denote the $(N+1)$ point Gauss-Lobatto-Legendre (GLL) quadrature rule with positive weights $w_1,\ldots,w_{N+1}$.
+Let ${x}_i \in [-1,1]$ denote the $(N+1)$ point Gauss-Lobatto-Legendre (GLL) quadrature rule with positive weights $w_1,\ldots,w_{N+1}$.
 Let $\ell_i(x)$ denote the nodal Lagrange basis defined at GLL points, and define the differentiation matrix $\bm{D}$ be defined as
@@
 \]
-where the approximation results evaluating the integral using GLL quadrature.
-For this choice of quadrature, the Kronecker property $\ell_j(x_i) = \delta_{ij}$ gives that $\bm{M}$ is a diagonal matrix.
+where the approximation results from evaluating the integral using GLL quadrature.
+For this choice of quadrature, the Kronecker property $\ell_j(x_i) = \delta_{ij}$ implies that $\bm{M}$ is a diagonal matrix whose entries are simply the quadrature weights.
 
@@
 \[
-\bm{u}(\bm{x},t) = \sum_{j=1}^{N+1} \bm{u}_j(t)\ell_j(x).  
+\bm{u}(x,t) = \sum_{j=1}^{N+1} \bm{u}_j(t)\ell_j(x).  
 \]
@@
 \end{equation}
-The replacement of the derivative of the flux function with the flux differencing allows one to prove that (\ref{eq:dgsem}) conserves a discrete entropy \cite{gassner2017br1,chen2017entropy}.
+The replacement of the derivative of the flux function with flux differencing allows one to prove that (\ref{eq:dgsem}) conserves a discrete entropy \cite{gassner2017br1,chen2017entropy}.
 We reproduce these proofs here, as we will refer to them in following sections.
@@
 \]
-where we have used continuity in time to arrive at the pointwise relation 
+where we have used continuity in time to arrive at the point-wise relation 
 \[
@@
 \end{align*}
-where we have used that $\bm{S}\bm{1} = 0$ and the summation by parts property in the last line.
+where we have used $\bm{S}\bm{1} = 0$ and the summation by parts property in the last line.
 
@@
 \end{proof}
-
-\begin{remark}
-The proof of entropy conservation can be extended beyond a single domain by specifying appropriate simultaneous approximation terms (SBP-SAT).
-This can be done by taking $\bm{f}_i^*$ to be the entropy conservative flux $f_S(u_L,u_R)$ at the interface between two elements \cite{carpenter2014entropy,gassner2017br1,chen2017entropy}.
+We note that, unlike the proof of continuous entropy conservation, the proof of discrete entropy conservation does not require the use of the chain rule, which may not hold at the discrete level.
+
+The proof of entropy conservation can also be extended beyond a single domain by specifying appropriate simultaneous approximation terms (SBP-SAT).
+This can be done by taking $\bm{f}_i^*$ to be the entropy conservative flux $f_S(u_L,u_R)$ at the interface between two elements \cite{carpenter2014entropy, gassner2017br1, chen2017entropy}.
 Similarly, employing an entropy stable flux at element interfaces results in an entropy stable method which satisfies a global entropy inequality.
-\end{remark}
-
-\section{Entropy conservative general DG methods}
-
-In this section, we focus on constructing entropy conservative DG schemes with arbitrary bases and quadrature rules.
-These include dense norm SBP operators and generalized SBP (GSBP) operators which do not contain boundary points, which introduce additional difficulties in constructing entropy conservative DG schemes.
-These difficulties present themselves primarily in three areas:
+%\end{remark}
+
+\section{Entropy conservative general DG methods on a single domain}
+\label{sec:ecdg}
+
+In this section, we focus on constructing entropy conservative DG schemes for general choices of basis and quadrature.
+These include (as special cases) dense norm SBP operators and generalized SBP (GSBP) operators which do not contain boundary points, which introduce additional difficulties in constructing entropy conservative DG schemes.
+These difficulties present themselves primarily in two areas:
 \begin{enumerate}
-\item It is unclear how to generalize the flux differencing approach to arbitrary choices of quadrature.  
-\item The proof of entropy stability relies on a diagonal norm mass matrix, which commutes with pointwise multiplication by the (nodal values of) the entropy variables.  
-\item Entropy stable formulations based on dense norm or GSBP operators require non-trivial SBP-SAT terms \cite{ranocha2017extended}.
-These can be derived by hand for simple problems such as Burgers' equation, but can be difficult to determine for more complex systems of nonlinear PDEs.
+\item the generalization of flux differencing to a continuous formulation.  
+\item the proof of entropy stability assumes a diagonal norm mass matrix, which commutes with point-wise multiplication by the (nodal values of) the entropy variables.  
 \end{enumerate}
 
-We simultaneously address all three areas of difficulty by introducing a generalization of the entropy conservative DG-SEM scheme (\ref{eq:dgsem}) to arbitrary quadratures and choices of basis.
-Additionally, we prove that the generalized scheme is entropy conservative (stable) on multiple elements in the following sections.
+We simultaneously address both areas of difficulty by generalizing the entropy conservative DG-SEM scheme (\ref{eq:dgsem}).
+In particular, we give a continuous DG formulation with flux differencing, and prove that the generalized scheme is entropy conservative (stable) on a single element in the following sections.
+As an example, we illustrate how to apply the proposed methods to produce energy conservative formulations of Burgers' equation.
+
+% to a larger class of quadratures and choices of basis.  Additionally, we prove that the generalized scheme is entropy conservative (stable) on multiple elements in the following sections.
 
@@
 \[
-\bm{x}^k = \bm{\Phi}^k \widehat{\bm{x}},
-\]
-where $\bm{x}^k$ denote physical coordinates on $D^k$ and $\widehat{\bm{x}}$ denote coordinates on the reference element.
+x^k = \bm{\Phi}^k \widehat{x},
+\]
+where $x^k$ denote physical coordinates on $D^k$ and $\widehat{x}$ denote coordinates on the reference element.
 Finally, we identify $J$ as the determinant of the Jacobian of $\bm{\Phi}^k$, and refer to it as the Jacobian for the remainder of this work.
+For clarity and simplicity of presentation, we assume $\bm{\Phi}^k$ is affine in this case, such that $J$ is constant over each element $D^k$.
 
@@
 \]
-We assume that $V_h(D^k) \subset L^2(D^k)$; in other words, that all elements of each local approximation space are square integrable.
+We assume that $V_h(D^k) \subset H^1(D^k)$; in other words, that all elements of each local approximation space are square integrable and admit well-defined traces.
+
+We define the $L^2$ norm and inner products over the element $D^k$ and the surface of the element $\partial D^k$\footnote{In one dimension, $D^k$ is some interval $[x_L,x_R]$ and the $L^2$ inner product over the surface reduces to
+\[
+\LRa{u,v}_{\partial D^k} = u(x_L)v(x_L) + u(x_R)v(x_R).
+\]
+However, the presented proofs in this work will not utilize this simplification, and are thus generalizable to higher dimensions in a straightforward manner.}
+\[
+\LRp{u,v}_{D^k} = \int_{D^k} u v\diff{x}, \qquad \nor{u}^2_{D^k} = (u,u)_{D^k}, \qquad \LRa{u,v}_{\partial D^k} = \int_{\partial D^k} u v \diff{x}.
+\]
 We also define the local $L^2$ projection operator $\Pi_N: L^2(D^k)\rightarrow V_h(D^k)$ such that
@@
 \]
-Because the global approximation space $V_h\LRp{\Omega}$ is the direct sum of independent local approximation spaces, the global $L^2$ projection is simply the composition of local $L^2$ projections.
-
-%\note{We define the quadrature-based $L^2$ projection $\Pi_N$ as the $L^2$ projection where integrals are evaluated using some (inexact) quadrature.
-We also define the discrete approximation space $V_h$, and assume that the quadrature is sufficiently accurate such that integration by parts holds for any two elements of $V_h$.
-Finally, we assume that the discrete derivative (to be defined) maps from $V_h$ to $V_h$.
+Throughout this work, we will use that this definition of the $L^2$ projection holds even when integrals are approximated using quadrature.
+We will make no assumptions on quadrature except that \note{specify assumptions - integrate polynomials of degree $N$ on the reference element exactly, such that integration by parts holds.
+Add footnote that we can also generalize to the under-integrated case.
 }
@@
 Finally, we define the jump and average of discontinuous functions across element interfaces.
-Let $D^{k,+}$ denote the neighboring element of across a face $f$ of $D^k$, and let $u^+,u^-$ denote the values of $u$ on $D^{k,+}$ and $D^k$, respectively.
-The jump of $u$ across $f$ is then defined as
-\[
-\jump{u} = u^+ - u^-, \qquad \avg{u} = \frac{u^+ + u^-}{2}.
+Let $D^{k,+}$ denote the neighboring element of across a face $f$ of $D^k$, and let $u^+$ denote the values of $u$ on $D^{k,+}$.
+The jump of $u$ across $f$ on $D^k$ is then defined as
+\[
+\jump{u} = u^+ - u, \qquad \avg{u} = \frac{u^+ + u}{2}.
 \]
 The jump and average of vector fields are defined component-wise using the jumps and averages of components.
-
-%\note{Specify assumption that differentiation operators $\bm{D}:V_h\rightarrow V_h$}
+When the face $f$ lies on the boundary, we define $u^+ = u$, such that $\jump{u} = 0$ and $\avg{u} = u$ on $\partial \Omega$.
 
@@
 \[
-\bm{f}_S\LRp{\bm{u}(\bm{x}),\bm{u}(\bm{y})}, \qquad \bm{x}, \bm{y} \in \mathbb{R}^d.  
-\]
-Flux differencing can be interpreted as differentiating $\bm{f}_S\LRp{\bm{u}(\bm{x}),\bm{u}(\bm{y})}$ along one coordinate $\bm{x}_k$, then evaluating the resulting derivative with $\bm{y}=\bm{x}$.
-Since the two-point flux is consistent, this results in the following approximation of the flux derivative
+\bm{f}_S\LRp{\bm{u}(x),\bm{u}(y)}, \qquad x, y \in \mathbb{R}.  
+\]
+Flux differencing can be interpreted as differentiating $\bm{f}_S\LRp{\bm{u}(x),\bm{u}(y)}$ along one coordinate $x$, then evaluating the result with $y=x$
 \begin{equation}
-\pd{\bm{f}(\bm{u}(\bm{x}))}{\bm{x}_k} \Longrightarrow 2\left.\pd{\bm{f}_S\LRp{\bm{u}(\bm{x}),\bm{u}(\bm{y})}}{\bm{x}_k}\right|_{\bm{y}=\bm{x}}.
+\pd{\bm{f}(\bm{u}(x))}{x} \Longrightarrow 2\left.\pd{\bm{f}_S\LRp{\bm{u}(x),\bm{u}(y)}}{x}\right|_{y=x}.
 \label{eq:fluxdiff}
 \end{equation}
-
-\subsubsection{Illustration using Burgers' equation}
-
-We note that flux differencing allows for the recovery of split formulations.  We illustrate this using the one-dimensional Burgers equation, which is given in conservative form as
+It was shown in \cite{chen2017entropy} that flux differencing is a consistent approximation of the flux derivative.  
+%Since the two-point flux is consistent,
+
+%\subsubsection{Illustration using Burgers' equation}
+
+It is helpful to first understand flux differencing as a method for recovering split formulations.
+We illustrate this using the one-dimensional Burgers equation on domain $ [-1,1]$, which is given in conservative form as
 \[
@@
 \]
-It is well known that DG formulations based on the split form
-\[
-\pd{u}{t} + \frac{1}{3}\LRp{\pd{u^2}{x} + u\pd{u}{x}} = 0
-\]
-can be shown to be discretely energy stable under appropriately imposed boundary conditions \cite{gassner2013skew, ranocha2017extended}.
-This form may be recovered using flux differencing by choosing the two-point flux
+%An alternative form of the Burgers' equation can be derived by taking a convex combination of the conservative and non-conservative forms of the spatial term
+%\begin{equation}
+%\pd{u}{t} + \frac{1}{3}\LRp{\pd{u^2}{x} + u\pd{u}{x}} = 0.
+%\label{eq:burgerssplit}
+%\end{equation}
+Assuming smooth $u$, continuous energy conservation can be shown by multiplying by $u$, integrating over $[-1,1]$, and using $\pd{u}{t}u = \frac{1}{2}\pd{u^2}{t}$
+\begin{equation}
+\frac{1}{2}\pd{}{t}\int_{-1}^1{u}^2 = -\left.\frac{u^3}{3}\right|_{-1}^1.
+\label{eq:burgersconservation}
+\end{equation}
+
+It is well known that DG formulations based on a split formulation can be shown to conserve the square entropy $u^2$ \cite{gassner2013skew, ranocha2017extended}.
+This form may be also be recovered using flux differencing under the two-point flux
 \[
@@
 \]
-Then, applying (\ref{eq:fluxdiff}) yields
-\[
-2\left.\pd{f_S(u(x),u(y))}{x}\right|_{y=x} = \frac{1}{3}\left.\pd{\LRp{u(x)^2 + u(x)u(y) + u(y)^2}}{x}\right|_{y=x} = \frac{1}{3}\LRp{\pd{u^2}{x} + u\pd{u}{x}},
-\]
-due to the fact that $\pd{u(y)^2}{x} = 0$.
+Applying (\ref{eq:fluxdiff}) yields the split form of Burgers' equation
+\[
+2\left.\pd{f_S(u(x),u(y))}{x}\right|_{y=x} = \frac{1}{3}\left.\pd{\LRp{u(x)^2 + u(x)u(y) + u(y)^2}}{x}\right|_{y=x} = \frac{1}{3}\LRp{\pd{u^2}{x} + u\pd{u}{x}}.
+\]
+We note that recovering the split-form of Burgers' equation relies on the property
+%\footnote{This is also the reason why the factor of $2$ is necessary when using flux differentiation (\ref{eq:fluxdiff}); the differentiation of terms which depend only on $y$ vanish, \note{finish...}.}
+that 
+\[
+\pd{\LRp{u(y)^2}}{x} = u(y)^2\pd{\LRp{1}}{x} = 0.  
+\]
+Thus, recovery of the split form can be achieved by using flux differencing in combination with any differential operator $D$ such that $D1 = 0$.  
+%\[
+%D:L^2 \rightarrow V_h, \qquad D1 = 0.
+%\]
+%For Burgers' equation (as well as any skew-symmetric splitting), replacing the exact derivative $\pd{}{x}$ by $D$ will again recover a statement of the split formulation (\ref{eq:burgerssplit}) 
+%\[
+%\pd{u}{t} + \frac{1}{3}\LRp{Du^2 + uDu} = 0.
+%\]
+%The introduction of $D$ is necessary when adapting the flux differentiation procedure in the context of numerical discretizations.
+
+\subsection{Discrete differential operators and weighted integration by parts}
 
@@
 Suppose now that the solution is approximated by $u \in V_h(\Omega)$ on a single element.
-Since $f(u) = u^2/2 \not\in V_h(\Omega)$, it is not possible to differentiate $f(u)$ exactly without prior knowledge of the form of the nonlinearity.
-Instead, it is more common to first project $f(u)$ to $V_h$, then differentiate, such that the discrete scheme is equivalent to solving for $u \in V_h(\Omega)$ such that
-\[
-\int_{-1}^1 \LRp{\pd{u}{t} + \pd{\Pi_N f(u)}{x}}w = 0, \qquad \forall w \in V_h(\Omega).  
-\]
-This scheme is energy stable only for projections which are computed using a quadrature of degree $3N-1$, which is sufficiently accurate to integrate the nonlinear DG weak form exactly.
-
-In comparison, when applying (\ref{eq:fluxdiff}) using $\pd{\Pi_N (\cdot)}{x}$, the resulting right hand side for Burgers' equation becomes 
-\[
-\int_{-1}^1\LRp{\pd{u}{t} + \frac{1}{3}\LRp{\pd{\LRp{\Pi_N u^2}}{x} + \Pi_N \LRp{u\pd{u}{x}}}}w = 0, \qquad \forall w \in V_h(\Omega).  
-\]
-Using the fact integration by parts holds when considering two elements of $V_h$, and that $\LRp{\Pi_N u^2, v} = \LRp{u^2,v}$ for any $v \in V_h$, it can be shown that the volume terms cancel out for $v = u$, leaving only boundary terms (which is a prerequisite of energy/entropy stability).
-
-\note{De-aliasing by increasing the degree of quadrature hits a ``bottleneck'': in the end, you are only differentiating a polynomial of degree $N$ \cite{kirby2003aliasing}.}
-
-We note that the recovery of the split-form of Burgers' equation in the above example relies only on the property that 
-\[
-\pd{f(u(y))}{x} = f(u(y)) \pd{1}{x} = 0.
-\]
-As such, replacing $\pd{}{x}$ by any differential operator $D$ such that $D 1 = 0$ will recover a similar formulation.
-
-\subsubsection{Recovering the algebraic form of flux differencing}
-
-This approach yields the following evaluation (in Matlab for lack of better notation): let 
-\[
-\verb|fS = @(uL,uR) uL.^2+uL.*uR+uR.^2|
-\]
-be the flux function.
+Since $f(u) = u^2/2 \not\in V_h(\Omega)$ for general choices of $u, V_h$, it is not possible to differentiate $f(u)$ exactly without prior knowledge of the form of the nonlinearity.
+Indeed, it is more common to first project $f(u)$ to $V_h$, then differentiate, such that the discrete scheme is equivalent to solving for $u \in V_h(\Omega)$ such that
+\[
+\int_{-1}^1 \LRp{\pd{u}{t} + \pd{\Pi_N f(u)}{x}}w \diff{x}  
+%\int_{-1}^1 \LRp{\pd{u}{t} +D_N f(u)}w \diff{x} 
+=0, \qquad \forall w \in V_h(\Omega).  
+\]
+This scheme does not, in general, conserve energy.
+%Moreover, if $\Pi_N$ is computed using quadrature, rewriting the scheme into this form suggests that de-aliasing techniques based on over-integration hit a ``bottleneck'', due to the fact that increasing quadrature accuracy only increases the accuracy of the projection \cite{kirby2003aliasing}.
+
+We can derive a semi-discrete Galerkin scheme based which conserves energy discretely using flux differencing and a different choice of discrete differentiation operator $D_N$.
+Let $w(x)$ be some bounded real-valued function.
+Then, we can define the discrete differential operator $D_N: H^1\rightarrow V_h$ such that, for $u\in H^1$,
+\begin{align}
+\LRp{D_N u, vw}_{\widehat{D}} = \LRp{\pd{ \Pi_N u}{x},vw}_{\widehat{D}} + \frac{1}{2}{\LRa{{u - \Pi_N u}, vw\bm{n}_x}}_{\partial \widehat{D}} + \frac{1}{2}\LRa{u - \Pi_Nu ,\Pi_N\LRp{vw}\bm{n}_x }_{\widehat{D}}, \qquad \forall v\in V_h.
+\label{eq:dgd}
+\end{align}
+This operator satisfies a ``weighted'' integration by parts property, which is summarized by the following lemma:
+\begin{lemma}
+The linear operator $D_N$ is symmetric with respect the $L^2$ inner product and any bounded real-valued function $w(x)$ 
+\[
+\LRp{D_N u, wv}_{\widehat{D}} = \LRa{u,wv\bm{n}_x}_{\partial \widehat{D}} - \LRp{u,D_N (w v)}_{\widehat{D}}.
+\]
+\label{lemma:dgd_local}
+\end{lemma}
+\begin{proof}
+Integrating the left hand side by parts gives
+\[
+-\LRp{\pd{ \Pi_N (vw)}{x},u}_{\widehat{D}} + \frac{1}{2}{\LRa{{u - \Pi_N u}, vw\bm{n}_x}}_{\partial \widehat{D}} + \frac{1}{2}\LRa{u + \Pi_Nu ,\Pi_N\LRp{vw}\bm{n}_x }_{\partial \widehat{D}}, \qquad \forall v\in V_h.
+\]
+Rearranging boundary terms yields
+\[
+-\LRp{\pd{ \Pi_N (vw)}{x},u}_{\widehat{D}} + \frac{1}{2}\LRa{vw + \Pi_N(vw) , u\bm{n}_x }_{\partial \widehat{D}}  - \frac{1}{2}\LRa{vw - \Pi_N (vw), \Pi_N u \bm{n}_x}_{\partial \widehat{D}}.  
+\]
+Adding and subtracting $\LRa{vw,u\bm{n}_x}_{\partial \widehat{D}}$ completes the proof.
+\end{proof}
+
+Direct computations also show that $\LRp{D_N 1,vw}_{\widehat{D}} = 0$ for all bounded weights $w(x)$ and $\forall v \in V_h$.
+Thus, we can apply (\ref{eq:fluxdiff}) with $D_N$ to construct a discrete formulation for Burgers' equation
+\begin{align}
+\LRp{\pd{u}{t},v} + \frac{1}{3}\LRp{D_N u^2 + uD_Nu,v} &= \LRp{\pd{u}{t},v}_{\widehat{D}} + \frac{1}{3}\LRp{D_N u^2,v}_{\widehat{D}} + \frac{1}{3}\LRp{D_Nu,uv}_{\widehat{D}},
+%\LRp{\pd{u}{t},v}_{\widehat{D}} + \frac{1}{3}\LRp{\pd{\Pi_N u^2}{x},v}_{\widehat{D}} + \frac{1}{3}\LRp{D_Nu,uv}_{\widehat{D}}
+\qquad \forall v\in V_h.
+\label{eq:burgersdiscrete}
+\end{align}
+To show this formulation is discretely energy conservative, we take $v=u$ and use Lemma~\ref{lemma:dgd_local} to cancel volume terms, yielding
+\[
+\frac{1}{2}\pd{}{t}\LRp{u,u}_{\widehat{D}} = -\frac{1}{3}\LRa{u^2,u}_{\partial \widehat{D}},
+\]
+which is a discrete analogue of the conservation condition (\ref{eq:burgersconservation}).
+This is equivalent to the skew-symmetric split formulation described in \cite{gassner2013skew}; however, as noted in \cite{chen2017entropy}, it can be derived from flux differencing without explicit knowledge of the skew form.
+We will show later in Section~\ref{sec:recovery} that a natural extension of this formulation to multiple elements recovers the weak form of the method introduced in \cite{ranocha2017extended}.
+
+%We note that the $L^2$ projection appears within the boundary terms.
+While this may not pose a significant problem on one element, in the case of multiple elements, it introduces an all-to-all coupling between degrees of freedom on elements which share a face \cite{fernandez2016simultaneous}.
+In the following section, we describe an alternative choice of discrete differential operator $D$ inspired by Chen and Shu \cite{chen2017entropy} which avoids this problem of increased bandwidth while maintaining properties which are necessary to guarantee provably discrete entropy conservation.
+
+%Using the fact integration by parts holds when considering two elements of $V_h$, and that $\LRp{\Pi_N u^2, v} = \LRp{u^2,v}$ for any $v \in V_h$, it can be shown that the volume terms cancel out for $v = u$, leaving only boundary terms (which is a prerequisite of energy/entropy stability).
+
+\subsection{Entropy conservation on one element}
+
+While flux differencing can be used to recover split formulations, it can also be used to construct entropy stable formulations which do not correspond to skew-symmetric splittings.
+In this section we will show that, for any sufficiently accurate quadrature rule, it is possible to construct discretizations which satisfy a discrete equivalent of conservation of entropy (\ref{eq:consentropy}).
+
+For the proof, we will need the following assumption:
+\begin{assumption}
+The numerical flux $\bm{f}_S(x,y)$ admits an absolutely convergent expansion of the form
+\[
+\bm{f}_S(x,y) = \sum_{i=1}^{\infty} \bm{f}_i(x) \bm{g}_i(y), 
+\]
+where $\bm{f}_i, \bm{g}_i$ are differentiable.  
+\label{eq:assumption}
+\end{assumption}
+This assumption is satisfied, for example, by $\bm{f}_S$ which are analytic and admit a convergent Taylor expansion in $x$ and $y$.
+In Appendix~\ref{appendix:A}, we show that this assumption also holds for the choices of flux functions $\bm{f}_S$ considered in this work, assuming the solution satisfies physical bound constraints.
+However, we do need this assumption in computation implementations of the proposed methods.
+
+We now introduce a discretization for the nonlinear system of conservation laws (\ref{eq:pde}) on $[-1,1]$: find $\bm{u}\in V_h$ such that
+\begin{equation}
+\LRp{\pd{\bm{u}}{t},\bm{w}}_{\widehat{D}} + \LRp{\left.\LRp{2D_N \bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x},\bm{w}}_{\widehat{D}}= \LRa{\bm{f}^*-\bm{f}(\bm{u}_x),\bm{w}}_{\partial \widehat{D}}, \qquad \bm{w}\in V_h,
+\label{eq:ecdg_oneelem}
+\end{equation}
+where $\bm{f}^*$ is a numerical flux used to impose boundary conditions weakly, and $\bm{u}_x,\bm{u}_y$ are evaluations of $\bm{u}$ in terms of the projections of entropy variables $\bm{v}(\bm{u}(x))$
+\begin{equation}
+%\bm{u}_x = \bm{u}(\Pi_N\bm{v}(\bm{u}(x))), \qquad \bm{u}_y = \bm{u}(\Pi_N\bm{v}(\bm{u}(y))).
+\bm{u}_x(x) = \bm{u}\LRp{(\Pi_N\bm{v})(x)}, \qquad \bm{u}_y(y) = \bm{u}\LRp{(\Pi_N\bm{v})(y)}.
+\label{eq:projectedentropyvars}
+\end{equation}
+
+%We first introduce matrices defined over $[-1,1]$.
+We assume some discrete quadrature rule on the reference element consisting of $N_q$ points ${\bm{x}}_q = \LRc{{x}_1,\ldots,{x}_{N_q}}$ and quadrature weights $\bm{w}_q = \LRc{w_1,\ldots,w_{N_q}}$.
+Let $\bm{W}$ denote the diagonal matrix whose entries are quadrature weights.
+Then, assuming that each component of $\bm{u}$ is represented in some basis $\phi_j(x)$ of dimension $N_p$ which spans $V_h$
+%\[
+%V_h = {\rm span}\LRc{\phi_j(\widehat{x})}_{j=1}^{N_p},
+%\]
+%we can define the quadrature interpolation matrix $\bm{V}_q$
+%\[
+%\LRp{\bm{V}_q}_{ij} = \phi_j(\widehat{x}_i), \qquad 1 \leq j \leq N_p, \qquad 1 \leq i \leq N_q.  
+%\]
+%Given $\bm{V}_q$, we introduce the mass matrix $\bm{M} = \bm{V}_q^T\bm{W}\bm{V}_q$, as well as the quadrature-based $L^2$ projection operator $\bm{P}_q$
+%\[
+%\bm{P}_q = \bm{M}^{-1}\bm{V}_q^T\bm{W}.
+%\]
+%Let $\bm{D}_x$ denote the differentiation matrix defined implicitly through
+%\[
+%u(x) = \sum_{j=1}^{N_p}\bm{u} \phi_j(\widehat{x}), \qquad \pd{u}{\widehat{x}} = \sum_{j=1}^{N_p} \LRp{\bm{D}_x \bm{u}}_j\phi_j(\widehat{x})
+%\]
+%In other words, $\bm{D}_x$ maps basis coefficients of some function $u \in V_h$ to coefficients of its derivative on $[-1,1]$.  
+%
+%Finally, we introduce matrices used to compute boundary terms.
+Let $\bm{V}_f$ denote the matrix which interpolates to boundary values\footnote{In generalizing to higher dimensions, $\bm{V}_f$ would be the matrix which interpolates to surface quadrature points.}
+%\[
+%\LRp{\bm{V}_f}_{1j} = \phi_j(-1), \qquad \LRp{\bm{V}_f}_{2j} = \phi_j(1), \qquad 1 \leq j \leq N_p.  
+%\]
+%Furthermore, let $\bm{W}_f$ denote the diagonal matrix of surface quadrature weights, which is simply the $2\times 2$ identity matrix in one dimension.  
+%
+%Together, these matrices can be used to define the quadrature-based differentiation matrix $\bm{D}_q$
+%\[
+%\bm{D}_q = \bm{V}_q \bm{D}_x \bm{P}_q,
+%\]
+%which is a discrete analogue to $D_N$.  It is straightforward then to show that this matrix satisfies properties similar to summation-by-parts:
+%\begin{lemma}
+%Let $\bm{n}_x = [-1,1]$ be the vector of unit normals in one dimension.  Then, 
+%\[
+%\bm{W}\bm{D}_q = \bm{P}_q^T\bm{V}_f^T\bm{W}_f {\rm diag}(\bm{n}_x) \bm{V}_f \bm{P}_q  - \bm{D}_q^T \bm{W}, \qquad \bm{1}^T\bm{W}\bm{D}_q =  \bm{1}^T\bm{W}_f {\rm diag}(\bm{n}_x) \bm{V}_f \bm{P}_q.  
+%\]
+%\label{lemma:sbp}
+%\end{lemma}
+%\begin{proof}
+%\note{Finish}
+%\end{proof}
+
+%Let $\bm{u}^k$ now denote the vector of coefficients for the $k$th component of $\bm{u}(x)$.  The formulation (\ref{eq:ecdg_oneelem}) is equivalent to the semi-discrete system:
+%\begin{equation}
+%\bm{M} \td{\bm{u}^k}{t} + \bm{V}_q^T\bm{W}{\rm diag}\LRp{\bm{V}_q\bm{D}_x\bm{P}_q \bm{F}^k_S} = 0, \qquad k = 1,\ldots,n.
+%\label{eq:ecdg_oneelem_mat}
+%\end{equation}
+%Here, $\bm{F}^k_S$ denotes the matrix
+%\[
+%\LRp{\bm{F}^k_S}_{ij} = \bm{f}^k_S\LRp{\bm{u}_{x,i},\bm{u}_{y,j}}, \qquad 1 \leq i,j \leq N_q,
+%\]
+%where $\bm{f}^k_S$ is the $k$th component of the vector flux $\bm{f}_S$, and
+%\begin{equation}
+%\bm{u}_{x,i} = \bm{u}\LRp{\Pi_N \bm{v}(\bm{u}(x_i))}, \qquad \bm{u}_{y,j} = \bm{u}\LRp{\Pi_N \bm{v}(\bm{u}(x_j))}.  
+%\label{eq:projectedentropyvars}
+%\end{equation}
+
+\begin{theorem}
+The formulation (\ref{eq:ecdg_oneelem})
+ %and (\ref{eq:ecdg_oneelem_mat}) are
+ is semi-discretely entropy conservative in the sense that
+\[
+\LRp{\pd{U(\bm{u})}{t},1}_{\widehat{D}} =  \LRa{\psi,1}_{\partial \widehat{D}}-\LRa{{\bm{f}^*, \Pi_N\bm{v}}}_{\partial \widehat{D}}.
+\]
+\label{lemma:ec1}
+\end{theorem}
+\begin{proof}
+We take the test function $\bm{w}$ to be the projection of the entropy variables.  On the left hand side, we may use the fact that, under a method of lines discretization, 
+\[
+\bm{u}(x,t) = \sum_{j=1}^{N_p} \bm{u}_j(t) \phi_j(x), \qquad \pd{\bm{u}(x,t)}{t} \in V_h.
+\]
+Thus, testing with $\Pi_N \bm{v}$ yields 
+\[
+\LRp{\pd{\bm{u}}{t}, \Pi_N \bm{v}(\bm{u})}_{\widehat{D}} = \LRp{\pd{\bm{u}}{t},\bm{v}(\bm{u})}_{\widehat{D}} = \LRp{\pd{U(\bm{u})}{t},1}_{\widehat{D}}, \qquad \forall \bm{u}\in V_h.
+\]
+%Let $\bm{v}_q$ denote the entropy variables evaluated in terms of the conservative variables at quadrature points.
+Then, the same argument involving the time term can be shown using matrix notation by testing with $\bm{v} = {\bm{P}_q\bm{v}_q}$,
+%\[
+%\LRp{\bm{P}_q\bm{v}_q}^T\bm{M}\pd{\bm{u}}{t} = \bm{v}_q^T\bm{W}\bm{V}_q\bm{M}^{-1}\bm{M}\pd{\bm{u}}{t} = \bm{1}^T \bm{W} {\rm diag}(\bm{v}_q) \pd{\bm{u}_q}{t} = \bm{1}^T \bm{W} \pd{U(\bm{u}_q)}{t},
+%\]
+%where we have assumed continuity in time, and have used that $\bm{W}$ is diagonal.  
+For the spatial derivative, we expand $\bm{f}_S$ out using Assumption~\ref{eq:assumption} and use Lemma~\ref{lemma:dgd_local}
+\begin{align*}
+&\LRp{\left.\LRp{D_N \sum_{i=1}^{\infty}\bm{f}_i(\bm{u}_x)\bm{g}_i(\bm{u}_y)}\right|_{y=x}, \Pi_N\bm{v}(x)}_{\widehat{D}} = \sum_{i=1}^{\infty} \LRp{{D_N \bm{f}_i(\bm{u}_x)}, {\bm{g}_i(\bm{u}_x)\Pi_N\bm{v} }}_{\widehat{D}}\\
+&= \sum_{i=1}^{\infty}\LRa{\bm{f}_i(\bm{u}_x) \bm{g}_i(\bm{u}_x),\Pi_N\bm{v}}_{\partial \widehat{D}}  - \LRp{ \bm{f}_i(\bm{u}_x), D_N \LRp{\bm{g}_i\Pi_N\bm{v} }}_{\widehat{D}}\\
+&= \LRa{f_S(\bm{u}_x,\bm{u}_x), \Pi_N\bm{v}(x)}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N \Pi_N\LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y)\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}}\\
+&= \LRa{f(\bm{u}_x), \Pi_N\bm{v}(x)}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N \Pi_N\LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y)\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}},
+\end{align*}
+where we have used the consistency of $\bm{f}_S$ in the last step.  
+%where $\Pi_N^x, \Pi_N^y$ denote $L^2$ projections taken with respect to the $x$ and $y$ coordinates, respectively.  
+The volume terms can be manipulated to yield
+\begin{align*}
+%&\LRp{\left.\LRp{D_N \LRp{\bm{f}_S(\bm{u}(\Pi_N\bm{v}(x)),\bm{u}(\Pi_N\bm{v}(y)))\LRp{\Pi_N\bm{v}(y)-\Pi_N\bm{v}(x)} }}\right|_{y=x},1}_{\widehat{D}} \\
+&\LRp{\left.\LRp{D_N \LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y)\LRp{\Pi_N\bm{v}(y)-\Pi_N\bm{v}(x)} }}\right|_{y=x},1}_{\widehat{D}} \\
+&= \LRp{ \left.\LRp{D_N\LRp{\psi(y)-\psi(x) }}\right|_{y=x},1}_{\widehat{D}} = \LRp{-D_N \psi(x),1}_{\widehat{D}} = \LRa{-\psi,1}_{\partial \widehat{D}},
+\end{align*}
+where we have used Definition~\ref{def:tadmor} in going from the first to second integral and Lemma~\ref{lemma:dgd_local} in the final step.  
+%\[
+%{\rm diag}\LRp{\bm{V}_q\bm{D}_x\bm{P}_q \bm{F}^k_S} \not\in V_h
+%\]
+%for general flux functions.  However, we may use Lemma~\ref{lemma:sbp} to expand out terms
+%\begin{align*}
+%\bm{v}^T\bm{V}_q^T\bm{W}{\rm diag}\LRp{\bm{V}_q\bm{D}_x\bm{P}_q \bm{F}^k_S} &= \bm{v}^T\bm{V}_q^T{\rm diag}\LRp{\LRp{\bm{P}_q^T\bm{V}_f^T\bm{W}_f {\rm diag}(\bm{n}_x) \bm{V}_f \bm{P}_q  + \bm{W}\bm{D}_q - \bm{D}_q^T\bm{W} } \bm{F}^k_S}.
+%\end{align*}
+%Because $\bm{F}^k_S$ is evaluated in terms of the projection of the entropy variables, we may apply Definition~\ref{def:tadmor} to simplify the latter terms
+%\begin{align*}
+%\bm{v}^T\bm{V}_q^T{\rm diag}\LRp{\LRp{\bm{W}\bm{D}_q - \bm{D}_q^T\bm{W} }\bm{F}^k_S} &=\sum_{i,j=1}^{N_q} \LRp{\bm{V}_q\bm{v}}_i \LRp{\bm{W}\bm{D}_q - \bm{D}_q^T\bm{W} }_{ij}\LRp{\bm{F}^k_S}_{ij}\\
+%&=\sum_{i,j=1}^{N_q} \LRp{\bm{W}\bm{D}_q}_{ij} \LRp{\psi_i - \psi_j}.  
+%\end{align*}
+%
+%The spatial term does not 
+%\[
+%\bm{v}\bm{V}_q^T\bm{W}{\rm diag}\LRp{\bm{V}_q\bm{D}_x\bm{P}_q \bm{F}^k_S}
+%\]
+\end{proof}
+
+We see that, in the proof of Theorem~\ref{lemma:ec1}, recovering the time derivative of entropy on the left hand side requires testing with the projection of the entropy variables.
+For collocation-type DG methods, $L^2$ projection and interpolation are equivalent, which implies that evaluating the flux $\bm{f}_S$ using point-wise nodal values of the solution satisfies the entropy conservation condition given in Definition~\ref{def:tadmor}.
+For collocation-type DG methods, the method can also be simplified by noting that $\bm{V}_q,\bm{P}_q$ are identity matrices.
+
+In general, the projection of the entropy variables does not agree with their point-wise evaluation.
+However, the entropy conservation condition given in Definition~\ref{def:tadmor} holds so long as the flux is evaluated in a manner consistent with the entropy variables.
+This motivates (\ref{eq:projectedentropyvars}) as a non-standard way of evaluating the flux, which is the primary difference between the construction of entropy conservative schemes for collocation-type DG methods and projection-based DG methods.
+
+\section{Entropy conservative general DG methods on multiple elements}
+\label{sec:ecdg2}
+
+In this section, we discuss the extension of entropy conservative DG schemes to multiple elements.
+The primary difficulty in this case is the construction of appropriate interface (SBP-SAT) terms.
+Entropy stable formulations based on dense norm or GSBP operators require SBP-SAT terms which are either non-trivial to derive or have a wide stencil \cite{fernandez2016simultaneous, ranocha2017extended}.
+
+Extending from
+
+We present an approach which utilizes a global approach to flux differencing to derive entropy conservative interface terms.
+
+\subsection{Global DG differentiation operators}
+
+The construction of energy and entropy conservative DG-SBP methods requires the specification of appropriate SBP-SAT terms, which are often derived by reverse-engineering  boundary terms such that interface contributions cancel \cite{fernandez2016simultaneous, ranocha2017comparison, ranocha2017extended}.
+We present an alternative systematic approach to the derivation of interface terms for DG formulations by combining flux differencing with global discrete DG differential operators, which have been used to simplify the notation and analysis of DG formulations for elliptic equations \cite{di2011mathematical, wang2013weak}.
+This approach does not require the specification of appropriate SBP-SAT terms in order to prove entropy conservation; rather, explicit expressions for SBP-SAT terms may be derived in a systematic manner using flux differencing and the definitions of global DG operators.
+
+%Global DG differential operators have been used to simplify the notation and analysis of DG formulations for elliptic equations \cite{di2011mathematical, wang2013weak}.
+Here, we will use their properties to derive entropy conservative DG formulations based on a global approach to flux differencing.
+We note that this approach does not require the specification of appropriate SBP-SAT terms in order to prove entropy conservation; rather, explicit expressions for SBP-SAT terms may be derived in a systematic manner using flux differencing and the definitions of global DG operators.
+
+We adapt the construction of local differentiation operators by Chen and Shu \cite{chen2017entropy} to construct a global DG differentiation operator
+\begin{equation}
+\LRp{D^x_h u,v}_{\Omega} = \sum_k \LRp{\pd{ \Pi_N u}{x},v}_{D^k} + \frac{1}{2}{\LRa{{u^+ - \Pi_N u}, v\bm{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{{u - \Pi_Nu },\Pi_Nv \bm{n}_x}_{\partial D^k}},
+\label{eq:dgd_global}
+\end{equation}
+where $\LRp{D^x_h u,v}_{\Omega} = \sum_k \LRp{D^x_h u,v}_{D^k}$.  
+Entropy conservative DG schemes can be constructed using the fact that $D^x_h$ satisfies a \textit{global} analogue of integration by parts:
+\begin{lemma}
+Let $D^x_h$ be defined as in (\ref{eq:dgd_global}), and define $v^+ = v$ on $\partial \Omega$.  Then, 
+\[
+\LRp{D^x_h u,vw}_{\Omega} = \LRa{u,vw\bm{n}_x}_{\partial \Omega} - \LRp{u,D^x_h (vw)}_{\Omega}.
+\]
+for $u \in H^1(\Oh)$ and $\forall v\in V_h(\Omega_h)$.  
+\label{lemma:ibp}
+\end{lemma}
+\begin{proof}
+We may rewrite $\LRp{D^x_h u,vw}_{\Omega}$ using the definition of $D_N$ in (\ref{eq:dgd})
+\[
+\LRp{D^x_h u,vw}_{\Omega} = \sum_k \LRp{D_N u,vw}_{D^k} + \frac{1}{2}\LRa{\jump{u},vw\bm{n}_x}_{\partial D^k \setminus \partial \Omega}.  
+\]
+Applying Lemma~\ref{lemma:dgd_local} over each element yields 
+\begin{align*}
+\LRp{D^x_h u,vw}_{\Omega} &= \sum_k \LRp{-u,D_N (vw)}_{D^k} + \LRa{u + \frac{1}{2}\jump{u},vw\bm{n}_x}_{\partial D^k}\\
+&= \sum_k \LRp{-u,D_N (vw)}_{D^k} + \LRa{\avg{u},vw\bm{n}_x}_{\partial D^k}\\
+&= \sum_k \LRp{-u,D_N (vw)}_{D^k} - \LRa{\jump{vw},u\bm{n}_x}_{\partial D^k \setminus \partial \Omega} + \LRa{vw,u\bm{n}_x}_{\partial \Omega} \\
+&= \LRa{vw,u\bm{n}_x}_{\partial \Omega}-\LRp{u,D^x_h (vw)}_{\Omega}.    
+\end{align*}
+%Since $\pd{ \Pi_N u}{x}\in V_h$ by assumption, we can integrate by parts the local volume term of (\ref{eq:dgd_global}) on $D^k$.  Using properties of the $L^2$ projection then yields
+%\begin{align*}
+%\LRp{\pd{ \Pi_N u}{x},v}_{D^k} &= \LRp{\pd{ \Pi_N u}{x}, \Pi_N v}_{D^k} = \LRa{\Pi_N u,\Pi_N v}_{\partial D^k} - \LRp{\Pi_N u, \pd{\Pi_N v}{x}}_{D^k}\\
+%&= \LRa{\Pi_N u,\Pi_N v}_{\partial D^k} - \LRp{u, \pd{\Pi_N v}{x}}_{D^k}.  
+%\end{align*}
+%Substituting this into (\ref{eq:dgd_global}) yields
+%\[
+%\LRp{D^x_h u,v} = \sum_k  - \LRp{u, \pd{\Pi_N v}{x}}_{D^k} + \frac{1}{2}{\LRa{{u^+ - \Pi_N u}, v\bm{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{{u},\Pi_Nv \bm{n}_x}_{\partial D^k}}.
+%\]
+%For elements which do not intersect with the boundary, the term $\LRa{u^+,v\bm{n}_x}_{\partial D^k}$ can be exchanged with terms from neighboring elements to show
+%\[
+%\sum_k \frac{1}{2}\LRa{u^+,v\bm{n}_x}_{\partial D^k} = \sum_k \frac{1}{2}\LRa{u,v^+\bm{n}^+_x}_{\partial D^k} = 
+%\sum_k - \frac{1}{2}\LRa{u,v^+\bm{n}_x}_{\partial D^k}.  
+%\]
+%On the boundary $\partial \Omega$, the interface terms satisfy
+%\[
+%\frac{1}{2}\LRa{u^+,v\bm{n}_x}_{\partial \Omega} = \frac{1}{2}\LRa{u,v\bm{n}_x}_{\partial \Omega} = \LRa{u,v\bm{n}_x}_{\partial \Omega}-\frac{1}{2}\LRa{v^+,u\bm{n}_x}_{\partial \Omega}.
+%\]
+%Combining these yields the following statement
+%\begin{align*}
+%\LRp{D^x_h u,v} &= \LRa{u,v\bm{n}_x}_{\partial \Omega} - \LRp{\sum_k  \LRp{\pd{\Pi_N v}{x},u}_{D^k} + \frac{1}{2}\LRa{{v^+ - \Pi_N v}, u\bm{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{{v},\Pi_Nu \bm{n}_x}_{\partial D^k}} \\
+%&= \LRa{u,v\bm{n}_x}_{\partial \Omega}-\LRp{u,D^x_h v} 
+%\end{align*}
+%which is the desired analogue of integration by parts.  
+\end{proof}
+%Integration by parts can also be shown to yield the skew-symmetric form of $\bm{D}^x_h$
+%\[
+%\LRp{D^x_h u,v} =\frac{1}{2} \sum_k \LRp{\pd{ \Pi_N u}{x},v}_{D^k} - \LRp{u,\pd{\Pi_N v}{x}}_{D^k} + \LRa{\LRp{u_P - \Pi_N u_M}\bm{n}_x, v} + \LRa{{u_M }\bm{n}_x,\Pi_Nv}.
+%\]
+%Taking $v = u_M$ and cancelling $u_P u_M \bm{n}_x$ terms shows that the differentiation operator is skew-symmetric w.r.t.\ a discrete $L^2$ inner product.
+
+%Lemma~\ref{lemma:ibp} is equivalent to showing that the matrix representation $\bm{D}^x_h$ of the global DG differential operator satisfies a global SBP property with respect to a discrete $L^2$ norm defined by quadrature weights.
+%\begin{lemma}
+%Let $\bm{u}$ be degrees of freedom for $u,v \in V_h$.  Define the matrix $\bm{D}^x_h$ such that 
+%\[
+%\bm{W}
+%\]
+%\end{lemma}
+\begin{remark}
+It is possible to define a simpler global DG differentiation operator based on the composition of a discrete differentiation operator and the $L^2$ projection.
+Let ${D}^x_h: V_h\rightarrow V_h$ be defined as
+\[
+%\sum_k \LRp{D^x_h u,v}_{D^k} = \sum_k \LRp{\pd{u}{x},v}_{D^k} + \frac{1}{2}\LRa{\jump{u},v\bm{n}_x}_{\partial D^k}, \qquad \forall v \in V_h.
+\LRp{D^x_h u,v}_{\Omega} = \sum_k \LRp{\pd{u}{x},v}_{D^k} + \frac{1}{2}\LRa{\jump{u},v\bm{n}_x}_{\partial D^k}, \qquad \forall v \in V_h.
+\]
+Then, using that the range of $D^x_h$ is $V_h$, it is straightforward to show that $D^x_h \Pi_N$ satisfies an analogue of integration by parts, such that
+\[
+\LRp{D^x_h\Pi_N u, v}_{\Omega} = \LRa{\Pi_N u,v}_{\partial \Omega} - \LRp{D^x_h\Pi_N v, u}_{\Omega}, \qquad \forall u,v\in \L.
+\]
+%Numerical results using this global differentiation operator result in slightly more accurate results than when using the DG operator (\ref{eq:dgd_global}).  
+Unfortunately, while (\ref{eq:dgd_global}) requires only traces of neighboring elements, this construction results in a wide stencil involving all degrees of freedom on neighboring elements due to the presence of $\jump{u}$ and the $L^2$ projection.
+This choice of global DG differentiation operator is analogous to the choice of SBP-SAT terms specified in \cite{fernandez2016simultaneous} for SBP operators on simplicial elements without boundary nodes.
+\end{remark}
+
+\note{It's currently unclear how to distinguish whether different operators are high order accurate. }
+
+\note{Kopriva's note ``Stability of Overintegration Methods for Nodal Discontinuous Galerkin Spectral Element Methods'' does not apply here, because the formulation is entropy stable for any quadratures which evaluate $P^N$ exactly (assuming affine meshes!).
+Certain terms are under-integrated, but they do not contribute to the interplay between volume and surface terms described by Kopriva.}
+
+\subsection{Discretely entropy conservative general DG methods}
+
+Extending the entropy stable DG formulation (\ref{eq:ecdg_oneelem}) is done by combining the flux differencing with the global DG differential operator $D^x_h$
+\begin{equation}
+\sum_{k}\LRp{\pd{\bm{u}}{t},\bm{w}}_{D^k} + \LRp{\left.\LRp{2D^x_h\bm{f}(\bm{u}_x,\bm{u}_y)}\right|_{y=x},\bm{w}}_{D^k} = \LRa{\bm{f}(\bm{u}_x)-\bm{f}^*,\bm{w}}_{\partial \Omega},
+%\int_{D^k}\pd{\bm{u}}{t} \bm{w} + \left.\LRp{2D^x_h\bm{f}_S\LRp{\bm{u}(\Pi_N\bm{v}({x})),\bm{u}(\Pi_N\bm{v}({y}))}}\right|_{{y}={x}}\bm{w} + \LRa{\bm{f}^*-\bm{f}(\bm{u}),\bm{w}}_{\partial \Omega}= 0, \qquad \forall \bm{w}\in V_h^n,
+\label{eq:ecdg}
+\end{equation}
+where $\bm{f}^*$ is some numerical flux used in the imposition of boundary conditions, and the global operator $D^x_h$ is understood to be applied to each component of $\bm{f}_S\LRp{\bm{u}(x),\bm{u}(y)}$.
+
+\note{Global DG operators reveal how to construct SBP-SAT terms for any collocation-type DG methods (where $L^2$ projection and interpolation are equivalent).
+Not easy to deal with SBP-SAT terms for more complex systems \cite{ranocha2017extended, ranocha2017comparison}.
+}
+
+\begin{theorem}
+The formulation (\ref{eq:ecdg}) is entropy conservative in the sense that
+\[
+\LRp{\pd{U(\bm{u})}{t},1}_{\Omega} =  \LRa{\psi,1}_{\partial \Omega}-\LRa{{\bm{f}^*, \Pi_N\bm{v}}}_{\partial \Omega}.
+\]
+\end{theorem}
+\begin{proof}
+The proof is identical to the proof of Theorem~\ref{lemma:ec1}, except that Lemma~\ref{lemma:ibp} is used in place of Lemma~\ref{lemma:dgd_local}.
+\end{proof}
+
+\begin{remark}
+The analysis in this section has focused on the construction of entropy conservative schemes.
+However, in general, entropy is not conserved and should be dissipated away at discontinuities and shocks.
+To this end, one can construct discretely entropy stable schemes by adding additional dissipative terms (for example, by adding HLL or Lax-Friedrichs penalization \cite{chen2017entropy} or matrix dissipation terms \cite{chandrashekar2013kinetic, winters2017uniquely}).
+\end{remark}
+
+\begin{remark}
+The presented analysis focuses on discrete entropy conservation in one dimension.
+However, the extension to two and three dimensions is straightforward (as shown in \cite{chen2017entropy}), based on analogous definitions of global DG operators in higher dimensions.
+The construction of higher dimensional DG operators differ from the one-dimensional case in that the evaluation of surface integrals also requires appropriately accurate quadrature rules over the reference element.
+\end{remark}
+
+\subsubsection{Recovery of known SBP-SATs for Burgers' equation}
+\label{sec:recovery}
+
+To illustrate the application of the global flux differencing approach, we derive explicit expressions for an energy stable formulation for Burgers' equation.
+We will assume periodic boundary conditions for simplicity and use the energy conserving flux
+\[
+f_S(u_L,u_R) = \frac{1}{6}\LRp{u_L^2 + u_Lu_R + u_R^2}.  
+\]
+We focus on the discretization of the spatial terms.  Applying flux differencing to the flux to yield the following contributions
+\[
+\LRp{\left.2{D}^x_h f_S(u(x),u(y))\right|_{y=x}, v} = \frac{1}{3}\LRp{\left.{D}^x_h\LRp{u(x)^2 + u(x)u(y) + u(y)^2 }\right|_{y=x}, v}= \frac{1}{3}\LRp{D^x_h\LRp{u^2} + u D^x_h u, v}.
+\]
+Expanding out the definition of $D^x_h\LRp{u^2}$ gives
+\begin{align*}
+\LRp{D^x_h\LRp{u^2} , v}_{\Omega} &= \sum_k \LRp{\pd{\Pi_N u^2}{x} , v}_{D^k} + \frac{1}{2}\LRa{\LRp{u^+}^2-\Pi_N u^2,v\bm{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{u^2 - \Pi_N u^2, \Pi_N v \bm{n}_x}_{\partial D^k}\\
+&= \sum_k \LRp{\pd{\Pi_N u^2}{x} , v}_{D^k} + \LRa{\frac{1}{2}\LRp{(u^+)^2 + u^2}-\Pi_N u^2,v\bm{n}_x}_{\partial D^k}.
+\end{align*}
+Similarly, expanding out $\LRp{u D^x_h u, v}_{\Omega}$ gives
+\begin{align*}
+\LRp{u D^x_h u, v}_{\Omega} &= \LRp{D^x_h u, uv}_{\Omega} = \sum_k \LRp{\pd{u}{x},uv}_{D^k} + \frac{1}{2}\LRa{{u^+}-\Pi_N u, uv\bm{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{u - \Pi_N u, \Pi_N (uv)}_{\partial D^k}\\
+&= \sum_k \LRp{\pd{u}{x},uv}_{D^k} + \frac{1}{2}\LRa{{u^+}u- u^2, v\bm{n}_x}_{\partial D^k}.
+\end{align*}
+Summing these together yields explicit expressions for the DG weak formulation
+\[
+%\sum_k \LRp{\pd{u}{t} + \frac{1}{3}\LRp{\pd{\Pi_N u^2}{x} + u\pd{u}{x}} , v}_{D^k} + \LRa{\frac{1}{6}\LRp{(u^+)^2 + u^+u + u^2}- \frac{1}{3}\Pi_N u^2 - \frac{1}{6}u^2,v\bm{n}_x}_{\partial D^k} = 0.
+\sum_k \LRp{\pd{u}{t} + \frac{1}{3}\LRp{\pd{\Pi_N u^2}{x} + u\pd{u}{x}} , v}_{D^k} + \LRa{f_S(u^+,u)- \frac{1}{3}\Pi_N u^2 - \frac{1}{6}u^2,v\bm{n}_x}_{\partial D^k} = 0.
+\]
+This formulation (which was derived through the use of flux differencing with the global DG operator (\ref{eq:dgd_global})) recovers the equivalent split formulation and correction terms introduced in \cite{ranocha2017extended} for Burgers' equation using generalized and dense norm  SBP operators.
+
+\subsubsection{On implementation for general flux functions}
+
+While it is possible to use (\ref{eq:ecdg}) to derive explicit DG weak formulations, this can be difficult for complicated flux functions.
+Instead, we follow \cite{gassner2016split, chen2017entropy} and seek an implementation of flux differencing and (\ref{eq:ecdg}) which requires only the specification of $\bm{f}_S(\bm{u}_L,\bm{u}_R)$ and local matrices.
+
+We begin by rewriting the global DG differentiation operator $D^x_h$ using the lifting operator  ${L}: V_h^f\rightarrow V_h$
+\[
+\LRp{{L} u_f,v}_{D^k} = \LRa{u_f,v}_{\partial D^k}, \qquad u_f \in V_h^f, \qquad \forall v\in V_h.
+\]
+where $V_h^f$ denotes the trace space of the approximation space $V_h$
+\[
+V_h^f = \LRc{ \left.u\right|_{\partial D^k}, \quad u \in V_h}.
+\]
+The lift operator can be used to rewrite surface terms which involve the volume $L^2$ projection
+\[
+\LRa{u,\Pi_Nv}_{\partial D^k} = \LRp{Lu,\Pi_N v}_{D^k} = \LRp{Lu,v}_{D^k}, \qquad u\in V_h, \qquad v\not\in V_h.  
+\]
+Then, the global DG differentiation operator can be rewritten as 
+\begin{equation}
+\LRp{D^x_h u,v}_{\Omega} = \sum_k \LRp{\pd{ \Pi_N u}{x} + \frac{1}{2}L\LRp{u-\Pi_N u}\bm{n}_x,v}_{D^k} + \frac{1}{2}{\LRa{{u^+ - \Pi_N u}, v\bm{n}_x}_{\partial D^k}}.
+\label{eq:dgd_globalimplement}
+\end{equation}
+
+%\subsection{Matrix and algebraic forms of entropy conservative DG methods}
+%
+%We note that it also possible to show that the proposed formulation (\ref{eq:ecdg}) is entropy conservative using matrix notation.  
+%
+%\note{Introduce global operators.}
+%
+%
+%This approach yields the following evaluation (in Matlab for lack of better notation): let 
+%\[
+%\verb|fS = @(uL,uR) uL.^2+uL.*uR+uR.^2|
+%\]
+%be the flux function.
 Let \verb+uq = Vq * u+ denote the interpolation of $u$ to quadrature points, and let \verb+Dx+ denote the differentiation matrix.
@@
 Then, assuming element-constant Jacobians, the split form can be evaluated via
-\[
-\verb+Pq * diag((Vq*Dx*Pq) * fS(ux, uy)) = Pq * sum((Vq* Dx* Pq) .* fS(ux, uy), 2).+
-\]
-The latter expression can be performed by computing \verb+fS(ux,uy)+ on the fly, avoiding the need to store it as a matrix.
+%\[
+%\verb+Pq * diag((Vq*Dx*Pq) * fS(ux, uy)) = Pq * sum((Vq* Dx* Pq) .* fS(ux, uy), 2).+
+%\]
+%The latter expression can be performed by computing \verb+fS(ux,uy)+ on the fly, avoiding the need to store it as a matrix.
 It can also be generalized to handle non-square matrices in a very similar way.
-
-When using nodal collocation using either GLL or GQ points, \verb+Vq,Pq+ are the identity matrix and we recover the algebraic form of flux differencing described in \cite{gassner2016split}.
-
-\note{Introduce definition of $\bm{f}(\bm{u}_q)$ as $f(\bm{u}_q)$.}
-
-\subsection{Global DG differentiation operators}
-
-The construction of entropy conservative DG methods
-
-Global DG methods \cite{di2011mathematical}
-
-We adapt the construction of local differentiation operators by Chen and Shu \cite{chen2017entropy} to construct a global DG differentiation operator
-\begin{equation}
-\LRp{D^x_h u,v} = \sum_k \LRp{\pd{ \Pi_N u}{x},v}_{D^k} + \frac{1}{2}{\LRa{{u_P - \Pi_N u_M}, v\bm{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{{u_M - \Pi_Nu_M },\Pi_Nv \bm{n}_x}_{\partial D^k}}.
-\label{eq:dgd}
-\end{equation}
-Integrating by parts then yields the skew-symmetric form of $\bm{D}^x_h$
-\[
-\LRp{D^x_h u,v} =\frac{1}{2} \sum_k \LRp{\pd{ \Pi_N u}{x},v}_{D^k} - \LRp{u,\pd{\Pi_N v}{x}}_{D^k} + \LRa{\LRp{u_P - \Pi_N u_M}\bm{n}_x, v} + \LRa{{u_M }\bm{n}_x,\Pi_Nv}.
-\]
-Taking $v = u_M$ and cancelling $u_P u_M \bm{n}_x$ terms shows that the differentiation operator has the global SBP property and is skew-symmetric  w.r.t.\ a discrete $L^2$ inner product.
-The coupling requires only traces of neighboring elements.
-
-It's currently unclear how to distinguish whether different operators are high order accurate.
-
-\begin{remark}
-It is possible to define a simpler global DG differentiation operator based on the composition of a weak polynomial differentiation operator and the $L^2$ projection.
-Let $\bm{D}^x_h: V_h\rightarrow V_h$ be defined as
-\[
-\sum_k \LRp{D^x_h u,v}_{D^k} = \sum_k \LRp{\pd{u}{x},v}_{D^k} + \frac{1}{2}\LRa{\jump{u},v\bm{n}_x}_{\partial D^k}, \qquad \forall v \in V_h.
-\]
-Then, it is straightforward to show that $D^x_h \Pi_N$ possesses a global SBP property with respect to the $L^2$ norm, such that
-\[
-\LRp{D^x_h\Pi_N u, v}_{\Omega} = \LRa{\Pi_N u,v}_{\partial \Omega} - \LRp{D^x_h\Pi_N v, u}_{\Omega}, \qquad \forall u,v\in \L.
-\]
-Numerical results using this global differentiation operator result in slightly more accurate results than when using the DG operator (\ref{eq:dgd}).
-Unfortunately, compared to  (\ref{eq:dgd}), this construction results in a wide stencil involving all degrees of freedom on neighboring elements due to the presence of $\jump{u}$ and the $L^2$ projection.
-\end{remark}
-
-\subsection{Discretely entropy conservative general DG methods}
-
-We introduce 
-\[
-\int_{D^k}\pd{\bm{u}}{t} \bm{w} + \left.\LRp{D^x_h\bm{f}_S\LRp{\bm{u}(\Pi_N\bm{v}(\bm{x})),\bm{u}(\Pi_N\bm{v}(\bm{y}))}}\right|_{\bm{y}=\bm{x}}\bm{w} = 0, \qquad \forall \bm{w}\in V_h,
-\]
-where the global operator $D^x_h$ is understood to be applied to each component of $\bm{f}_S\LRp{\bm{u}(x),\bm{u}(y)}$.
-
-Global DG operators reveal how to construct SBP-SAT terms for any collocation-type DG methods (where $L^2$ projection and interpolation are equivalent).
-However, for quadrature-based DG methods, it is not possible to test with the entropy variables directly because the test space is polynomial.
-However, we may test with the projections of the entropy variables.
-On the left hand side, we may use the fact that, under a method of lines discretization,
-\[
-\bm{u}(x,t) = \sum_{j=1}^{N_p} \bm{u}_j(t) \phi_j(x), \qquad \pd{\bm{u}}{t} \in P^N.
-\]
-Thus, testing with $\Pi_N \bm{v}$ over each element $D^k$ yields  
-\[
-\int_{D^k} \pd{\bm{u}}{t}\Pi_N \bm{v} = \int_{D^k} \pd{\bm{u}}{t}\bm{v} = \int_{D^k} \pd{U(\bm{u})}{t}.  
-\]
-In matrix notation, this can be shown by multiplying with $\LRp{\bm{P}_q\bm{v}_q}^T$ 
-\[
-\LRp{\bm{P}_q\bm{v}_q}^T\bm{M}\pd{\bm{u}}{t} = \bm{v}_q^T\bm{W}\bm{V}_q\bm{M}^{-1}\bm{M}\pd{\bm{u}}{t} = \bm{1}^T \bm{W} {\rm diag}(\bm{v}_q) \pd{\bm{u}_q}{t} = \bm{1}^T \bm{W} \pd{U(\bm{u}_q)}{t},
-\]
-where we have assumed continuity in time, and have used that $\bm{W}$ is diagonal.
-Summing contributions from each element $D^k$ yields the time derivative of the entropy over the entire domain $\Omega$.
-
-%which yields a pointwise relationship which can be manipulated into the time derivative of entropy.
-
-For collocation DG methods, interpolation and quadrature-based $L^2$ projection are identical.
-However, when using a quadrature rule with $N_q > N_p$, this is not the case.
-The solution is to evaluate the flux with $u(\Pi_N v)$ - in other words, evaluate the flux variables in terms of the projections of the entropy variables.
-
-Not easy to deal with SBP-SAT terms for more complex systems \cite{ranocha2017extended,ranocha2017comparison}.
-
-We summarize this analysis in the following theorem:
-\begin{theorem}
-\note{Blah is entropy conservative.  Finish up proof + statement of entropy conservation.}
-\end{theorem}
-
-\begin{remark}
-The analysis in this section has focused on the construction of entropy conservative schemes.
-However, for problems with shocks, entropy is not conserved and should be dissipated away.
-To this end, one can construct entropy stable schemes by adding additional stabilization terms (for example, by adding Lax-Friedrichs penalization or matrix dissipation terms \cite{chandrashekar2013kinetic,winters2017uniquely}).
-\end{remark}
-
-\subsubsection{On implementation}
-
-Let $V_h^f$ denote the trace space of the approximation space $V_h$ 
-\[
-V_h^f = \LRc{ \left.u\right|_{\partial D^k}, \quad u \in V_h}.  
-\]
-We introduce the lifting operator ${L}: V_h^f\rightarrow V_h$.  For $u_f \in V^f_h$, ${L}$ is defined as 
-\[
-\LRp{{L} u_f,v}_{D^k} = \LRa{u_f,v}_{\partial D^k}, \qquad \forall v\in V_h.
-\]
-
-\subsubsection{Example: Burgers equation}
-
-We illustrate the application of the global flux differencing approach to Burgers' equation using the energy conserving flux
-\[
-f_S(u_L,u_R) = \frac{1}{6}\LRp{u_L^2 + u_Lu_R u_R^2}.  
-\]
-We will derive 
-\[
-\LRp{\left.\bm{D}^x_h f_S(u(x),u(y))\right|_{y=x}, w} = 
-\]
-
-We note that the global flux differencing approach with the DG operator (\ref{eq:dgd}) recovers the formulation and correction terms introduced in \cite{ranocha2017extended}.
-
-\subsection{Conservation of secondary properties}
-
-It holds but with respect to specifically defined fluxes.
+%
+%
+%\note{Introduce definition of $\bm{f}(\bm{u}_q)$ as $f(\bm{u}_q)$.}
+
+\subsection{Entropy stability}
+
+\note{Unlike collocative DG, entropy stability must be in terms of $\jump{\bm{u}\LRp{\Pi_N \bm{v}}}$ to guarantee discrete entropy decay.  }
+
+\subsection{Local conservation}
+
+It holds but with respect to specifically defined fluxes.  Need to use $v = 1$ on $D^k$ and then symmetry of the global DG operator $D^x_h$.
 
 \section{Numerical experiments}
-
-\note{We illustrate the entropy conservation and accuracy of the proposed schemes for three equations.  These gradually increase in complexity.  }
-
-\subsection{Burgers}
-
-We begin by computing $L^2$ errors and rates of convergence for smooth solutions of Burgers' equation.
+\label{sec:num}
+\note{We illustrate the entropy conservation and accuracy of the proposed schemes for three equations.
+These gradually increase in complexity - Burgers has entropy variables in test space and skew symmetric form, shallow water has only skew symmetric form, and compressible Euler has neither.
+}
+
+%\subsection{Burgers}
+%
+%We begin by computing $L^2$ errors and rates of convergence for smooth solutions of Burgers' equation.  
+%\subsection{Accuracy of the global DG differentiation operator}
+%
+%\note{Show that both ``correction'' and ``coupling'' terms are necessary for full $h^{N+1}$ convergence.  Removing either degrades convergence down to $h^N$.}
+
+%\subsection{Burgers' equation}
 
@@
 \]
-Like the Burgers' equation, the shallow water equations are stable under an appropriate split formulation\cite{gassner2016well, chen2017entropy}.
+Like the Burgers' equation, the shallow water equations are stable under an appropriate split formulation \cite{gassner2016well, chen2017entropy}.
 However, unlike Burgers' equation, the shallow water equations are not entropy stable with respect to the square entropy.
 Additionally, a key difference between the shallow water equations and Burgers' equation is that, when discretizing with respect to the conservative variables $h, hv$, the entropy variables $v_1,v_2$ are not contained within the space $V_h$.
-
-%This conservation of energy does not hold at the discrete level for two reasons.
-The first reason is because, under inexact quadrature, calculus identities such as the chain or product rule do not hold, and naive spatial formulations can not be manipulated into a form which is amenable to a proof of discrete entropy stability.
-This issue can be remedied by the use of a split formulation (as shown in Gassner et al.) which is provably entropy stable under inexact quadrature.
-%
-%The second more subtle reason why entropy conservation does not hold at the discrete level is because the entropy variables are not contained within the test space.
-Moreover, the spatial part of the proposed discretization requires one discrete definition of the entropy variables for entropy stability, while the temporal part of the discretization requires another discrete definition of the entropy variables.
-We address this by adopting a rescaling in the mass matrix on the left hand side involving both definitions of the discrete entropy variables.
-Under an appropriate choice of basis and quadrature, the proposed approach recovers existing DG-SEM discretizations.
 
@@
 The choice of convex entropy for the Euler equations is non-unique \cite{harten1983symmetric}.
-However, a unique entropy can be chosen by restricting to choices of entropy variables which symmetrize the viscous heat conduction term in the compressible Navier-Stokes equations \cite{hughes1986new}.
-The entropy is given by
+However, a unique entropy can be chosen by restricting to choices of entropy variables which symmetrize the viscous heat conduction term in the compressible Navier-Stokes equations \cite{hughes1986new}, which leads to $U(\bm{u})$ of the form
 \[
@@
 \end{align*}
-The reverse mapping is given by 
-\[
-d
-\]
-
-\note{Add Chandreshekar's fluxes \cite{chandrashekar2013kinetic}.}
+The inverse mapping is given by 
+\[
+\rho = -(\rho e) q_3, \qquad m = (\rho e) v_2, \qquad E = (\rho e)\LRp{1 - \frac{q_2^2}{2 q_3}},
+\]
+where $\rho e$ and $s$ in terms of the entropy variables are 
+\[
+\rho e = \LRp{\frac{(\gamma-1)}{\LRp{-q_3}^{\gamma}}}^{1/(\gamma-1)}e^{\frac{-s}{\gamma-1}}, \qquad s = \gamma - q_1 + \frac{q_2^2}{2q_3}.
+\]
+
+Examples of entropy conservative flux functions can be found in \cite{ismail2009affordable, chandrashekar2013kinetic}.
+In this work, we utilize the flux function $\bm{f}_S(\bm{u}_L,\bm{u}_R)$ introduced by Chandreshekar \cite{chandrashekar2013kinetic}, whose components are given as
+\begin{align*}
+f^1_S &= \avg{\rho}^{\log} \avg{u}\\
+f^2_S &= \frac{\avg{\rho}}{2\avg{\beta}} + \avg{u}f^1_S\\
+f^3_S &= f^1_S\LRp{\frac{1}{2(\gamma-1)\avg{\beta}^{\log}} - \frac{1}{2}\avg{u^2}} + \avg{u}f^2_S,
+\end{align*}
+where we have introduced the inverse temperature $\beta$
+\[
+\beta = \frac{\rho}{2p}.
+\]
+These fluxes are both entropy conservative and kinetic energy preserving.
 
@@
 
+\note{Show how RHS residual depends on tolerance on the log mean}
+
 \section{Conclusions}
@@
 
-The author thanks Andrew Winters for several informative discussions.  Jesse Chan is supported by NSF awards DMS-1719818 and DMS-1712639.
+The author thanks Andrew Winters and David M.\ Williams for informative discussions.  Jesse Chan is supported by NSF awards DMS-1719818 and DMS-1712639.
+
+\appendix
+\section{Expansion of the flux function}
+\label{appendix:A}
+
+\note{For Burgers and shallow water, this sum is finite.
+For the Chandreshekar flux, we consider individual components: the logarithmic mean $L(x,y)$ admits a convergent expansion and is analytic in $x$ \cite{mustonen2002logarithmic}.
+CH's flux introduces rational denominators involving the mean and log-mean of $\beta = \rho/(2p)$, which is bounded away from zero if the density and temperature are bounded away from zero.
+By composition and products of analytic functions, the CH flux is also analytic.
+}
 

### 2017-07-25 docs/splitform/main.tex
updating paper

@@
 \author{Jesse Chan}
-\title{On a general class of discretely entropy conservative discontinuous Galerkin methods}
+\title{On discretely entropy conservative discontinuous Galerkin methods}
 
@@
 These methods can be interpreted as mass lumped discontinuous Galerkin spectral element methods (DG-SEM) \cite{gassner2016split,gassner2016well,wintermeyer2017entropy, chen2017entropy}.
-In this work, we describe how to extend discretely entropy conservative schemes for a more general class of DG methods using quadrature-based $L^2$ projections, flux differencing, and global DG differentiation operators.
+In this work, we describe how to extend discretely entropy conservative schemes for a more general class of DG methods using quadrature-based $L^2$ projections, flux differencing, and specific DG differentiation operators.
 This approach also recovers existing methods for Burgers' equation involving generalized SBP operators without boundary nodes and dense norm SBP operators.
-Numerical experiments confirm the stability and high order accuracy of the proposed methods for the shallow water and compressible Euler equations.
+Numerical experiments confirm the stability and high order accuracy of the proposed methods for the one-dimensional shallow water and compressible Euler equations.
 \end{abstract}
 
-\tableofcontents
+%\tableofcontents
 
@@
 While low numerical dissipation is advantageous for problems in linear wave propagation \cite{komatitsch1998spectral}, the high numerical dissipation present in low order schemes adds a stabilizing effect for nonlinear hyperbolic problems \cite{wang2013high}.
-As a result, high order schemes for nonlinear conservation laws tend to require additional stabilization, including (but not limited to) filtering, slope limiting, artificial viscosity, and polynomial de-aliasing \cite{persson2006sub, krivodonova2007limiters, karniadakis2013spectral, hesthaven2007nodal, kirby2003aliasing}.
+As a result, high order schemes for nonlinear conservation laws tend to require additional stabilization, including (but not limited to) filtering \cite{hesthaven2007nodal}, slope limiting \cite{krivodonova2007limiters}, artificial viscosity \cite{persson2006sub}, and polynomial de-aliasing \cite{ karniadakis2013spectral, kirby2003aliasing}.
 Moreover, stabilized numerical methods can still fail, requiring user intervention or heuristic modifications to achieve non-divergent solutions.
@@
 Thus, while DG-SEM based on flux differencing is entropy conservative, this is tied to a specific choice of basis and quadrature.
-Diagonal-norm SBP operators can be constructed for triangles and tetrahedra \cite{chin1999higher, cohen2001higher, hicken2016multidimensional, chen2017entropy}; however, the number of nodal points for such operators is greater than the dimension of the natural polynomial approximation space.
+Diagonal-norm SBP operators can be constructed for triangles and tetrahedra \cite{chin1999higher,  hicken2016multidimensional, chen2017entropy}; however, the number of nodal points for such operators is greater than the dimension of the natural polynomial approximation space.
 Furthermore, appropriate point sets have only been constructed for $N \leq 4$ in three dimensions \cite{zhebel2014comparison}.
@@
 In order to generalize beyond collocation DG and diagonal norm SBP operators, we will consider dense (non-lumped) mass matrices and quadrature rules of arbitrarily high accuracy (over-integration).
-The outline of the paper is as follows: Section~\ref{sec:intro} will briefly review the construction of entropy conservative collocation methods on a single domain.
-Section~\ref{sec:ecdg} will describe how to construct analogous entropy conservative methods on single domains.
-Section~\ref{sec:ecdg2} will discuss extensions to multiple domains, including comparisons of different coupling terms and entropy stable fluxes.
+The outline of the paper is as follows: Section~\ref{sec:intro} will briefly review the construction of entropy conservative collocation methods on a single element.
+Section~\ref{sec:ecdg} will describe how to construct analogous entropy conservative methods on single element.
+Section~\ref{sec:ecdg2} will discuss extensions to multiple elements, including comparisons of different coupling terms and entropy stable fluxes.
 We note that, for simplicity and clarity of presentation, the proposed methods and proofs are given in one dimension.
@@
 
-The proof of entropy conservation can also be extended beyond a single domain by specifying appropriate simultaneous approximation terms (SBP-SAT).
+The proof of entropy conservation can also be extended beyond a single element by specifying appropriate simultaneous approximation terms (SBP-SAT).
 This can be done by taking $\bm{f}_i^*$ to be the entropy conservative flux $f_S(u_L,u_R)$ at the interface between two elements \cite{carpenter2014entropy, gassner2017br1, chen2017entropy}.
@@
 
-\section{Entropy conservative general DG methods on a single domain}
+\section{Entropy conservative DG methods on a single element}
 \label{sec:ecdg}
@@
 
-We simultaneously address both areas of difficulty by generalizing the entropy conservative DG-SEM scheme (\ref{eq:dgsem}).
+We address both areas of difficulty in generalizing the entropy conservative DG-SEM scheme (\ref{eq:dgsem}).
 In particular, we give a continuous DG formulation with flux differencing, and prove that the generalized scheme is entropy conservative (stable) on a single element in the following sections.
@@
 
-We can derive a semi-discrete Galerkin scheme based which conserves energy discretely using flux differencing and a different choice of discrete differentiation operator $D_N$.
-Let $w(x)$ be some bounded real-valued function.
-Then, we can define the discrete differential operator $D_N: H^1\rightarrow V_h$ such that, for $u\in H^1$,
+We can derive a semi-discrete Galerkin scheme based which conserves energy discretely using flux differencing and a different choice of discrete differentiation operator $D_N$, which was motivated by the SBP operators introduced in \cite{chen2017entropy}.
+Let $w(x)$ be some bounded real-valued function; then, we can define the discrete differential operator $D_N: H^1\rightarrow V_h$ such that, for $u\in H^1$,
 \begin{align}
@@
 \end{align}
-This operator satisfies a ``weighted'' integration by parts property, which is summarized by the following lemma:
+This operator satisfies a form of integration by parts, which is summarized by the following lemma:
 \begin{lemma}
-The linear operator $D_N$ is symmetric with respect the $L^2$ inner product and any bounded real-valued function $w(x)$ 
+\label{lemma:dgd_local}
+The linear operator $D_N$ satisfies a ``weighted'' integration by parts property with respect the $L^2$ inner product and any bounded real-valued function $w(x)$ 
 \[
@@
 \]
-\label{lemma:dgd_local}
 \end{lemma}
@@
 
-\subsection{Entropy conservation on one element}
+\subsection{Conservation of entropy on one element}
 
@@
 \begin{theorem}
+\label{lemma:ec1}
 The formulation (\ref{eq:ecdg_oneelem})
@@
 \]
-\label{lemma:ec1}
 \end{theorem}
@@
 \]
-Thus, testing with $\Pi_N \bm{v}$ yields 
+Thus, testing with $\Pi_N \bm{v}$ and assuming continuity in time yields 
 \[
@@
 \]
+
 %Let $\bm{v}_q$ denote the entropy variables evaluated in terms of the conservative variables at quadrature points.
@@
 
-\section{Entropy conservative general DG methods on multiple elements}
+\section{Entropy conservative DG methods on multiple elements}
 \label{sec:ecdg2}
@@
 Entropy stable formulations based on dense norm or GSBP operators require SBP-SAT terms which are either non-trivial to derive or have a wide stencil \cite{fernandez2016simultaneous, ranocha2017extended}.
-
-Extending from
-
-We present an approach which utilizes a global approach to flux differencing to derive entropy conservative interface terms.
+We present an approach which instead utilizes a global approach to flux differencing to derive entropy conservative interface terms, and show that it recovers existing energy stable SBP-SAT terms for Burgers' equation.
 
@@
 
-We adapt the construction of local differentiation operators by Chen and Shu \cite{chen2017entropy} to construct a global DG differentiation operator
+We adapt the construction of local differentiation operators introduced previously to construct a global DG differentiation operator
 \begin{equation}
-\LRp{D^x_h u,v}_{\Omega} = \sum_k \LRp{\pd{ \Pi_N u}{x},v}_{D^k} + \frac{1}{2}{\LRa{{u^+ - \Pi_N u}, v\bm{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{{u - \Pi_Nu },\Pi_Nv \bm{n}_x}_{\partial D^k}},
+\LRp{D^x_h u,vw}_{\Omega} = \sum_k \LRp{\pd{ \Pi_N u}{x},vw}_{D^k} + \frac{1}{2}{\LRa{{u^+ - \Pi_N u}, vw\bm{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{{u - \Pi_Nu },\Pi_N(vw) \bm{n}_x}_{\partial D^k}},
 \label{eq:dgd_global}
 \end{equation}
-where $\LRp{D^x_h u,v}_{\Omega} = \sum_k \LRp{D^x_h u,v}_{D^k}$.  
+where $w(x)$ is a bounded smooth function over each element and $\LRp{D^x_h u,v}_{\Omega} = \sum_k \LRp{D^x_h u,v}_{D^k}$.  
 Entropy conservative DG schemes can be constructed using the fact that $D^x_h$ satisfies a \textit{global} analogue of integration by parts:
 \begin{lemma}
+\label{lemma:ibp}
 Let $D^x_h$ be defined as in (\ref{eq:dgd_global}), and define $v^+ = v$ on $\partial \Omega$.  Then, 
@@
 for $u \in H^1(\Oh)$ and $\forall v\in V_h(\Omega_h)$.  
-\label{lemma:ibp}
 \end{lemma}
@@
 
-\subsection{Discretely entropy conservative general DG methods}
+\subsection{Conservation of entropy on multiple elements}
 
@@
 
-\note{Global DG operators reveal how to construct SBP-SAT terms for any collocation-type DG methods (where $L^2$ projection and interpolation are equivalent).
+\note{Add assumptions about $\bm{u}(x)$ - needs to be locally analytic.}
+
+%\note{Global DG operators reveal how to construct SBP-SAT terms for any collocation-type DG methods (where $L^2$ projection and interpolation are equivalent).
 Not easy to deal with SBP-SAT terms for more complex systems \cite{ranocha2017extended, ranocha2017comparison}.
 }
-
+This formulation  yields a theorem analogous to Theorem~\ref{lemma:ec1}:
 \begin{theorem}
-The formulation (\ref{eq:ecdg}) is entropy conservative in the sense that
+The formulation (\ref{eq:ecdg}) is globally entropy conservative in the sense that
 \[
@@
 \begin{remark}
-The analysis in this section has focused on the construction of entropy conservative schemes.
-However, in general, entropy is not conserved and should be dissipated away at discontinuities and shocks.
-To this end, one can construct discretely entropy stable schemes by adding additional dissipative terms (for example, by adding HLL or Lax-Friedrichs penalization \cite{chen2017entropy} or matrix dissipation terms \cite{chandrashekar2013kinetic, winters2017uniquely}).
-\end{remark}
-
-\begin{remark}
 The presented analysis focuses on discrete entropy conservation in one dimension.
 However, the extension to two and three dimensions is straightforward (as shown in \cite{chen2017entropy}), based on analogous definitions of global DG operators in higher dimensions.
-The construction of higher dimensional DG operators differ from the one-dimensional case in that the evaluation of surface integrals also requires appropriately accurate quadrature rules over the reference element.
+The construction of higher dimensional DG operators differ from the one-dimensional case in that the evaluation of surface integrals also requires appropriately accurate quadrature rules over the reference element, such that integration by parts holds for elements of the local approximation space.
 \end{remark}
+
+\subsubsection{Local conservation}
+
+It is also straightforward to show that the formulation (\ref{eq:ecdg}) is locally conservative.  Taking $v = \mathbbm{1}_{D^k}$ in the formulation (\ref{eq:ecdg}) yields
+\[
+d
+\]
+
+\note{It holds but with respect to specifically defined fluxes.  Need to use $v = 1$ on $D^k$ and then symmetry of the global DG operator $D^x_h$.  }
 
@@
 \]
-This formulation (which was derived through the use of flux differencing with the global DG operator (\ref{eq:dgd_global})) recovers the equivalent split formulation and correction terms introduced in \cite{ranocha2017extended} for Burgers' equation using generalized and dense norm  SBP operators.
-
-\subsubsection{On implementation for general flux functions}
+This formulation, which was derived through the use of flux differencing with the global DG operator (\ref{eq:dgd_global}), recovers the weak form of the split formulation and the correction terms introduced in \cite{ranocha2017extended} for Burgers' equation.
+
+\subsection{On implementation for general flux functions}
 
@@
 \end{equation}
+
+Applying Assumption~\ref{eq:assumption}, we can rewrite 
+\begin{align*}
+\LRp{\left.\LRp{D^x_h \bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x},\bm{w}}_{\Omega} &= \sum_k \sum_{i=1}^{\infty} \LRp{\pd{\Pi_N \bm{f}_i(\bm{u})}{x},\bm{g}_i(\bm{u})\bm{w}}_{D^k}\\ 
+&+  \sum_k \sum_{i=1}^{\infty}\frac{1}{2}\LRa{\bm{f}_i(\bm{u}^+)\bm{g}_i(\bm{u}) - \bm{g}_i(\bm{u})\Pi_N\bm{f}_i(\bm{u}),\bm{w}\bm{n}_x}_{\partial D^k}\\
+&+  \sum_k \sum_{i=1}^{\infty}\frac{1}{2}\LRp{\bm{g}_i(\bm{u})L\LRp{\bm{f}_i(\bm{u}^+) - \Pi_N\bm{f}_i(\bm{u})}\bm{n}_x,\bm{w}}_{D^k},
+\end{align*}
+where $\bm{w}\bm{n}_x$ denotes component-wise multiplication of $\bm{w}$ by $\bm{n}_x$, the $x$-component of the unit outward normal.
+Moving the sum inside then allows us to re-write the resulting expression as
+\begin{align}
+\LRp{\left.\LRp{D^x_h \bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x},\bm{w}}_{\Omega} &= \sum_k \LRp{\left.\pd{\Pi_N \bm{f}_S(\bm{u}_x,\bm{u}_y)}{x}\right|_{y=x},\bm{w}}_{D^k}\nonumber\\ 
+&+ \frac{1}{2}\LRa{\bm{f}_S(\bm{u}^+,\bm{u}) - \left.\Pi_N\bm{f}_S(\bm{u}_x,\bm{u}_y)\right|_{y=x},\bm{w}\bm{n}_x}_{\partial D^k} \label{eq:eval}\\
+&+ \frac{1}{2}\LRp{\left.\LRp{L\LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y) - \Pi_N\bm{f}_S(\bm{u}_x,\bm{u}_y)}\bm{n}_x}\right|_{y=x},\bm{w}}_{D^k},\nonumber
+\end{align}
+This can be used to practically evaluate $\left.\LRp{\bm{D}^x_h \bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x}$ using only reference operators and evaluations of $\bm{f}_S$.
+However, unlike previous entropy conservative collocation DG methods, the boundary terms in (\ref{eq:eval}) interact with volume terms through the $L^2$ projection and the lift operator.
+This can be simplified for diagonal norm SBP operators with boundary nodes, using equivalences between $L^2$ projection and interpolation at nodal points in those cases.
 
@@
 
+The analysis in previous sections has focused on the construction of entropy conservative schemes.
+However, entropy is only conserved for smooth solutions and should be dissipated away in the presence of discontinuities and shocks.
+To this end, one can construct discretely entropy stable schemes by adding additional dissipative terms (for example, by adding HLL or Lax-Friedrichs penalization \cite{chen2017entropy} or matrix dissipation terms \cite{chandrashekar2013kinetic, winters2017uniquely}).
+
 \note{Unlike collocative DG, entropy stability must be in terms of $\jump{\bm{u}\LRp{\Pi_N \bm{v}}}$ to guarantee discrete entropy decay.  }
-
-\subsection{Local conservation}
-
-It holds but with respect to specifically defined fluxes.  Need to use $v = 1$ on $D^k$ and then symmetry of the global DG operator $D^x_h$.
 
@@
 \note{We illustrate the entropy conservation and accuracy of the proposed schemes for three equations.
-These gradually increase in complexity - Burgers has entropy variables in test space and skew symmetric form, shallow water has only skew symmetric form, and compressible Euler has neither.
+We focus on shallow water and compressible Euler.
+Compared to Burgers' equation, shallow water has a skew symmetric form but the entropy variables are not contained in the test space, and compressible Euler has neither.
 }
@@
 %\subsection{Burgers' equation}
+%
+%As mentioned previously, the Burgers' equation and and an entropy conservative flux are given by
+%\[
+%\pd{u}{t} + \frac{1}{2}\pd{u^2}{2} = 0, \qquad f_S(u_L,u_R) = \frac{1}{6}\LRp{u_L^2 + u_Lu_R + u_R^2}.
+%\]
 
@@
 \[
-q_1 = \pd{e}{h} = g(h+b)-v^2/2, \qquad q_2 = \pd{e}{hv} = v.
+q_1 = \pd{e}{h} = g(h+b)-v^2/2, \qquad q_2 = \pd{e}{hv} = v,
+\]
+with an inverse mapping
+\[
+h = \frac{1}{g}\LRp{q_1 + \frac{q_2^2}{2}}, \qquad v = q_2.
 \]
@@
 Additionally, a key difference between the shallow water equations and Burgers' equation is that, when discretizing with respect to the conservative variables $h, hv$, the entropy variables $v_1,v_2$ are not contained within the space $V_h$.
+
+The entropy conservative flux functions for the shallow water equations are given as follows:
+\begin{align*}
+\bm{f}^1_S(\bm{u}_L,\bm{u}_R)&= \avg{h v}\\
+\bm{f}^2_S(\bm{u}_L,\bm{u}_R) &= \avg{hv}\avg{v} + \frac{1}{2}g h_L h_R.  
+\end{align*}
+These correspond also to the skew-symmetric splitting introduced in \cite{gassner2016well}.
 
@@
 \begin{align*}
-f^1_S &= \avg{\rho}^{\log} \avg{u}\\
-f^2_S &= \frac{\avg{\rho}}{2\avg{\beta}} + \avg{u}f^1_S\\
-f^3_S &= f^1_S\LRp{\frac{1}{2(\gamma-1)\avg{\beta}^{\log}} - \frac{1}{2}\avg{u^2}} + \avg{u}f^2_S,
+f^1_S(\bm{u}_L,\bm{u}_R) &= \avg{\rho}^{\log} \avg{u}\\
+f^2_S(\bm{u}_L,\bm{u}_R) &= \frac{\avg{\rho}}{2\avg{\beta}} + \avg{u}f^1_S\\
+f^3_S(\bm{u}_L,\bm{u}_R) &= f^1_S\LRp{\frac{1}{2(\gamma-1)\avg{\beta}^{\log}} - \frac{1}{2}\avg{u^2}} + \avg{u}f^2_S,
 \end{align*}
@@
 
+In evaluating the logarithmic mean, we utilize the procedure described by Ismail and Roe in \cite{ismail2009affordable}.  
 \note{Show how RHS residual depends on tolerance on the log mean}

### 2017-07-25 docs/splitform/main.tex
updating paper

@@
 \]
-
 %Let $\bm{v}_q$ denote the entropy variables evaluated in terms of the conservative variables at quadrature points.
@@
 &= \sum_{i=1}^{\infty}\LRa{\bm{f}_i(\bm{u}_x) \bm{g}_i(\bm{u}_x),\Pi_N\bm{v}}_{\partial \widehat{D}}  - \LRp{ \bm{f}_i(\bm{u}_x), D_N \LRp{\bm{g}_i\Pi_N\bm{v} }}_{\widehat{D}}\\
-&= \LRa{f_S(\bm{u}_x,\bm{u}_x), \Pi_N\bm{v}(x)}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N \Pi_N\LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y)\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}}\\
-&= \LRa{f(\bm{u}_x), \Pi_N\bm{v}(x)}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N \Pi_N\LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y)\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}},
+&= \LRa{\bm{f}_S(\bm{u}_x,\bm{u}_x), \Pi_N\bm{v}(x)}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N \Pi_N\LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y)\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}}\\
+&= \LRa{\bm{f}(\bm{u}_x), \Pi_N\bm{v}(x)}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N \Pi_N\LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y)\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}},
 \end{align*}
@@
 \end{align*}
-%Since $\pd{ \Pi_N u}{x}\in V_h$ by assumption, we can integrate by parts the local volume term of (\ref{eq:dgd_global}) on $D^k$.  Using properties of the $L^2$ projection then yields
-%\begin{align*}
-%\LRp{\pd{ \Pi_N u}{x},v}_{D^k} &= \LRp{\pd{ \Pi_N u}{x}, \Pi_N v}_{D^k} = \LRa{\Pi_N u,\Pi_N v}_{\partial D^k} - \LRp{\Pi_N u, \pd{\Pi_N v}{x}}_{D^k}\\
-%&= \LRa{\Pi_N u,\Pi_N v}_{\partial D^k} - \LRp{u, \pd{\Pi_N v}{x}}_{D^k}.  
-%\end{align*}
-%Substituting this into (\ref{eq:dgd_global}) yields
-%\[
-%\LRp{D^x_h u,v} = \sum_k  - \LRp{u, \pd{\Pi_N v}{x}}_{D^k} + \frac{1}{2}{\LRa{{u^+ - \Pi_N u}, v\bm{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{{u},\Pi_Nv \bm{n}_x}_{\partial D^k}}.
-%\]
-%For elements which do not intersect with the boundary, the term $\LRa{u^+,v\bm{n}_x}_{\partial D^k}$ can be exchanged with terms from neighboring elements to show
-%\[
-%\sum_k \frac{1}{2}\LRa{u^+,v\bm{n}_x}_{\partial D^k} = \sum_k \frac{1}{2}\LRa{u,v^+\bm{n}^+_x}_{\partial D^k} = 
-%\sum_k - \frac{1}{2}\LRa{u,v^+\bm{n}_x}_{\partial D^k}.  
-%\]
-%On the boundary $\partial \Omega$, the interface terms satisfy
-%\[
-%\frac{1}{2}\LRa{u^+,v\bm{n}_x}_{\partial \Omega} = \frac{1}{2}\LRa{u,v\bm{n}_x}_{\partial \Omega} = \LRa{u,v\bm{n}_x}_{\partial \Omega}-\frac{1}{2}\LRa{v^+,u\bm{n}_x}_{\partial \Omega}.
-%\]
-%Combining these yields the following statement
-%\begin{align*}
-%\LRp{D^x_h u,v} &= \LRa{u,v\bm{n}_x}_{\partial \Omega} - \LRp{\sum_k  \LRp{\pd{\Pi_N v}{x},u}_{D^k} + \frac{1}{2}\LRa{{v^+ - \Pi_N v}, u\bm{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{{v},\Pi_Nu \bm{n}_x}_{\partial D^k}} \\
-%&= \LRa{u,v\bm{n}_x}_{\partial \Omega}-\LRp{u,D^x_h v} 
-%\end{align*}
-%which is the desired analogue of integration by parts.  
 \end{proof}
-%Integration by parts can also be shown to yield the skew-symmetric form of $\bm{D}^x_h$
-%\[
-%\LRp{D^x_h u,v} =\frac{1}{2} \sum_k \LRp{\pd{ \Pi_N u}{x},v}_{D^k} - \LRp{u,\pd{\Pi_N v}{x}}_{D^k} + \LRa{\LRp{u_P - \Pi_N u_M}\bm{n}_x, v} + \LRa{{u_M }\bm{n}_x,\Pi_Nv}.
-%\]
-%Taking $v = u_M$ and cancelling $u_P u_M \bm{n}_x$ terms shows that the differentiation operator is skew-symmetric w.r.t.\ a discrete $L^2$ inner product.
-
-%Lemma~\ref{lemma:ibp} is equivalent to showing that the matrix representation $\bm{D}^x_h$ of the global DG differential operator satisfies a global SBP property with respect to a discrete $L^2$ norm defined by quadrature weights.
-%\begin{lemma}
-%Let $\bm{u}$ be degrees of freedom for $u,v \in V_h$.  Define the matrix $\bm{D}^x_h$ such that 
-%\[
-%\bm{W}
-%\]
-%\end{lemma}
+
 \begin{remark}
@@
 \begin{equation}
-\sum_{k}\LRp{\pd{\bm{u}}{t},\bm{w}}_{D^k} + \LRp{\left.\LRp{2D^x_h\bm{f}(\bm{u}_x,\bm{u}_y)}\right|_{y=x},\bm{w}}_{D^k} = \LRa{\bm{f}(\bm{u}_x)-\bm{f}^*,\bm{w}}_{\partial \Omega},
+\sum_{k}\LRp{\pd{\bm{u}}{t},\bm{w}}_{D^k} + \LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x},\bm{w}}_{D^k} = \LRa{\bm{f}(\bm{u}_x)-\bm{f}^*,\bm{w}}_{\partial \Omega},
 %\int_{D^k}\pd{\bm{u}}{t} \bm{w} + \left.\LRp{2D^x_h\bm{f}_S\LRp{\bm{u}(\Pi_N\bm{v}({x})),\bm{u}(\Pi_N\bm{v}({y}))}}\right|_{{y}={x}}\bm{w} + \LRa{\bm{f}^*-\bm{f}(\bm{u}),\bm{w}}_{\partial \Omega}= 0, \qquad \forall \bm{w}\in V_h^n,
@@
 
-It is also straightforward to show that the formulation (\ref{eq:ecdg}) is locally conservative.  Taking $v = \mathbbm{1}_{D^k}$ in the formulation (\ref{eq:ecdg}) yields
-\[
-d
-\]
-
-\note{It holds but with respect to specifically defined fluxes.  Need to use $v = 1$ on $D^k$ and then symmetry of the global DG operator $D^x_h$.  }
+It is also straightforward to show that the formulation (\ref{eq:ecdg}) is locally conservative.
+We follow \cite{chen2017entropy} and take $\bm{w} = \mathbbm{1}_{D^k}$ in the formulation (\ref{eq:ecdg}) yields
+\[
+\sum_{k}\LRp{\pd{\bm{u}}{t},1}_{D^k} + \LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x},1}_{D^k} = \LRa{\bm{f}(\bm{u}_x)-\bm{f}^*,1}_{\partial \Omega \cap D^k}.  
+\]
+Using Lemma~\ref{lemma:dgd_local} and the symmetry of $\bm{f}_S$ yields that
+\[
+\LRp{\left.\LRp{D^x_h\bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x},1}_{D^k} = \LRa{\bm{f}(\bm{u}_x) ,1}_{\partial D^k} - \LRp{\left.\LRp{D^x_h\bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x},1}_{D^k}.
+\]
+Substituting this back into the first equation cancels volume terms, leaving
+\[
+\sum_{k}\LRp{\pd{\bm{u}}{t},1}_{D^k} + \LRa{\bm{f}(\bm{u}_x) ,1}_{\partial D^k \setminus \partial \Omega} + \LRa{\bm{f}^*,1}_{\partial \Omega \cap D^k} = 0.  
+\]
+We note that this definition of local conservation is given in terms of the conservation variables evaluated as a function of the projected entropy variables
+\[
+\bm{u}_x(x) = \bm{u}\LRp{\LRp{\Pi_N \bm{v}}(x)}.
+\]
+It is not immediately clear whether this statement of local conservation satisfies the conditions of the classic Lax-Wendroff theorem.
+However, this form of local conservation does satisfy a generalized definition of local conservation given in \cite{shi2017local}, and is sufficient to guarantee convergence to the weak solution.
 
@@
 
-\subsection{On implementation for general flux functions}
+\subsection{On implementation in terms of the flux function}
 
@@
 
-We begin by rewriting the global DG differentiation operator $D^x_h$ using the lifting operator  ${L}: V_h^f\rightarrow V_h$
-\[
-\LRp{{L} u_f,v}_{D^k} = \LRa{u_f,v}_{\partial D^k}, \qquad u_f \in V_h^f, \qquad \forall v\in V_h.
-\]
-where $V_h^f$ denotes the trace space of the approximation space $V_h$
-\[
-V_h^f = \LRc{ \left.u\right|_{\partial D^k}, \quad u \in V_h}.
+We begin by rewriting the global DG differentiation operator $D^x_h$ using the lifting operator  ${L}: V_h \rightarrow V_h$ \cite{hesthaven2007nodal, di2011mathematical}
+\[
+\LRp{{L} u,v}_{D^k} = \LRa{u,v}_{\partial D^k}, \qquad u \in V_h, \qquad \forall v\in V_h.
 \]
@@
 where $\bm{w}\bm{n}_x$ denotes component-wise multiplication of $\bm{w}$ by $\bm{n}_x$, the $x$-component of the unit outward normal.
-Moving the sum inside then allows us to re-write the resulting expression as
+Moving the sum inside and simplifying then allows us to re-write the resulting expression as
 \begin{align}
 \LRp{\left.\LRp{D^x_h \bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x},\bm{w}}_{\Omega} &= \sum_k \LRp{\left.\pd{\Pi_N \bm{f}_S(\bm{u}_x,\bm{u}_y)}{x}\right|_{y=x},\bm{w}}_{D^k}\nonumber\\ 
-&+ \frac{1}{2}\LRa{\bm{f}_S(\bm{u}^+,\bm{u}) - \left.\Pi_N\bm{f}_S(\bm{u}_x,\bm{u}_y)\right|_{y=x},\bm{w}\bm{n}_x}_{\partial D^k} \label{eq:eval}\\
-&+ \frac{1}{2}\LRp{\left.\LRp{L\LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y) - \Pi_N\bm{f}_S(\bm{u}_x,\bm{u}_y)}\bm{n}_x}\right|_{y=x},\bm{w}}_{D^k},\nonumber
+&+ \frac{1}{2}\LRa{\bm{f}_S(\bm{u}^+,\bm{u}) - \left.\Pi_N^x\bm{f}_S(\bm{u}_x,\bm{u}_y)\right|_{y=x},\bm{w}\bm{n}_x}_{\partial D^k} \label{eq:eval}\\
+&+ \frac{1}{2}\LRp{\left.\LRp{L^x\LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y) - \Pi_N^x\bm{f}_S(\bm{u}_x,\bm{u}_y)}\bm{n}_x}\right|_{y=x},\bm{w}}_{D^k},\nonumber
 \end{align}
+where the superscripts $L^x, \Pi_N^x$ are used to emphasize that the lift and projection operators are applied in $x$ and not in $y$.  
 This can be used to practically evaluate $\left.\LRp{\bm{D}^x_h \bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x}$ using only reference operators and evaluations of $\bm{f}_S$.
@@
 
+Finally, we show how to recover a computational form of flux differencing analogous to those presented in \cite{gassner2016split, chen2017entropy}.
+We will describe how to compute the three non-standard terms which appear within (\ref{eq:eval}):
+\begin{align}
+\LRp{\left.\pd{\Pi_N {f}_S({u}_x,{u}_y)}{x}\right|_{y=x},{w}}_{D^k}\label{eq:voleval}\\
+\LRa{\left.\Pi_N^x{f}_S({u}_x,{u}_y)\right|_{y=x},{w}\bm{n}_x}_{\partial D^k}\label{eq:surfeval}\\
+\LRp{\left.\LRp{L^x\LRp{{f}_S({u}_x,{u}_y) - \Pi_N^x{f}_S({u}_x,{u}_y)}\bm{n}_x}\right|_{y=x},{w}}_{D^k}\label{eq:volsurfeval}
+\end{align}
+where we have assumed a scalar variable $u(x)$ and flux function $f_S$ for simplicity.
+
+We first introduce matrices defined over $[-1,1]$.
+We assume some discrete quadrature rule on the reference element consisting of $N_q$ points ${\bm{x}}_q = \LRc{{x}_1,\ldots,{x}_{N_q}}$ and quadrature weights $\bm{w}_q = \LRc{w_1,\ldots,w_{N_q}}$.
+Let $\bm{W}$ denote the diagonal matrix whose entries are quadrature weights.
+Then, let $u(x)$ be represented in some basis $\phi_j(x)$ of dimension $N_p$ which spans $V_h$
+\[
+V_h = {\rm span}\LRc{\phi_j(\widehat{x})}_{j=1}^{N_p},
+\]
+we can define the quadrature interpolation matrix $\bm{V}_q$
+\[
+\LRp{\bm{V}_q}_{ij} = \phi_j(\widehat{x}_i), \qquad 1 \leq j \leq N_p, \qquad 1 \leq i \leq N_q.  
+\]
+Given $\bm{V}_q$, we introduce the mass matrix $\bm{M} = \bm{V}_q^T\bm{W}\bm{V}_q$, as well as the quadrature-based $L^2$ projection operator $\bm{P}_q$
+\[
+\bm{P}_q = \bm{M}^{-1}\bm{V}_q^T\bm{W}.
+\]
+Let $\bm{D}_x$ denote the differentiation matrix defined implicitly through
+\[
+u(x) = \sum_{j=1}^{N_p} \bm{u}_j \phi_j(\widehat{x}), \qquad \pd{u}{\widehat{x}} = \sum_{j=1}^{N_p} \LRp{\bm{D}_x \bm{u}}_j\phi_j(\widehat{x})
+\]
+In other words, $\bm{D}_x$ maps basis coefficients of some function $u \in V_h$ to coefficients of its derivative on $[-1,1]$.
+
+We also introduce matrices used to compute boundary terms.
+Let $\bm{V}_f$ denote the matrix which interpolates to boundary values\footnote{In generalizing to higher dimensions, $\bm{V}_f$ would be the matrix which interpolates to surface quadrature points.}
+\[
+\LRp{\bm{V}_f}_{1j} = \phi_j(-1), \qquad \LRp{\bm{V}_f}_{2j} = \phi_j(1), \qquad 1 \leq j \leq N_p.  
+\]
+Furthermore, let $\bm{W}_f$ denote the diagonal matrix of surface quadrature weights.
+In one dimension, this is simply the $2\times 2$ identity matrix.
+This allows us to define the quadrature-based lift matrix
+\[
+\bm{L} = \bm{M}^{-1}\bm{V}_f^T \bm{W}_f.
+\]
+Together, these matrices can be used to interpolate the output of the differentiation and lift matrices to volume quadrature points
+\[
+\bm{D}_q = \bm{V}_q \bm{D}_x \bm{P}_q, \qquad \bm{L}_q = \bm{V}_q\bm{L}_q.
+\]
+
+Hadamard product \cite{horn2012matrix}
 %\subsection{Matrix and algebraic forms of entropy conservative DG methods}
@@
 
+\note{Analysis and experiments are presented in one dimension for clarity; however, they are directly extendable to higher dimensions, including on simplicial elements.
+This will be the focus of future work.
+}
+
 \section{Acknowledgments}
 
-The author thanks Andrew Winters and David M.\ Williams for informative discussions.  Jesse Chan is supported by NSF awards DMS-1719818 and DMS-1712639.
+The author thanks Andrew Winters, David M.\ Williams, and Paul Hand for informative discussions.  Jesse Chan is supported by NSF awards DMS-1719818 and DMS-1712639.
 

### 2017-07-29 docs/splitform/main.tex
committing updates to entropy stable DG paper

@@
         % Draw slope triangle.
-        \draw[#5]   (A)-- node[pos=0.5,anchor=north] {1}
+        \draw[#5]   (A)-- node[pos=0.5,anchor=north] {}
                     (B)-- 
@@
         % Draw slope triangle.
-        \draw[#5]   (A)-- node[pos=.5,anchor=south] {1}
+        \draw[#5]   (A)-- node[pos=.5,anchor=south] {}
                     (B)-- 
@@
                     (B)-- 
-                    (C)-- node[pos=0.5,anchor=south] {1}
+                    (C)-- node[pos=0.5,anchor=south] {}
                     cycle;
@@
 The accuracy observed for high order methods can be attributed in part to their low numerical dissipation and dispersion properties \cite{ainsworth2004dispersive}.
-While low numerical dissipation is advantageous for problems in linear wave propagation \cite{komatitsch1998spectral}, the high numerical dissipation present in low order schemes adds a stabilizing effect for nonlinear hyperbolic problems \cite{wang2013high}.
+While low numerical dissipation is advantageous for problems in linear wave propagation \cite{chan2015gpu}, the high numerical dissipation present in low order schemes adds a stabilizing effect for nonlinear hyperbolic problems \cite{wang2013high}.
 As a result, high order schemes for nonlinear conservation laws tend to require additional stabilization, including (but not limited to) filtering \cite{hesthaven2007nodal}, slope limiting \cite{krivodonova2007limiters}, artificial viscosity \cite{persson2006sub}, and polynomial de-aliasing \cite{ karniadakis2013spectral, kirby2003aliasing}.
@@
 
-The construction of entropy stable high order schemes was done much more recently for the compressible Euler and Navier-Stokes equations in \cite{fisher2013high, carpenter2014entropy} based on a spectral collocation approach on tensor product elements, which can be interpreted as a mass-lumped DG spectral element (DG-SEM) scheme.
-The proof of entropy stability relies on the presence of a diagonal mass matrix, as well as the summation by parts property \cite{gassner2013skew}.
-Similar entropy-stable schemes were constructed for the shallow water and MHD equations \cite{derigs2016novel, gassner2016well, wintermeyer2017entropy}.
-More recently, entropy stable schemes were extended to unstructured triangular meshes in \cite{chen2017entropy} using the split-form methodology of \cite{gassner2016split} and multi-dimensional SBP operators \cite{hicken2016multidimensional}.
+The construction of entropy conservative high order schemes was done much more recently for the compressible Euler and Navier-Stokes equations in \cite{fisher2013high, carpenter2014entropy} based on a spectral collocation approach on tensor product elements, which can be interpreted as a mass-lumped DG spectral element (DG-SEM) scheme.
+The proof of entropy conservation relies on the presence of a diagonal mass matrix, as well as the summation by parts property \cite{gassner2013skew}.
+Similar entropy-stable schemes were constructed for the shallow water and MHD equations \cite{gassner2016well, wintermeyer2017entropy,winters2017uniquely}.
+More recently, entropy conservative and entropy stable schemes were extended to unstructured triangular meshes in \cite{chen2017entropy} using the split-form methodology of \cite{gassner2016split} and multi-dimensional SBP operators \cite{hicken2016multidimensional}.
 
@@
 This work focuses on the construction of entropy conservative high order DG schemes for systems of conservation laws.
-In order to generalize beyond collocation DG and diagonal norm SBP operators, we will consider dense (non-lumped) mass matrices and quadrature rules of arbitrarily high accuracy (over-integration).
+In order to generalize beyond collocation DG and diagonal norm SBP operators, we will consider dense (non-lumped) mass matrices and more general quadrature rules (e.g.\ rules without boundary points, rules with more points than the dimension of the corresponding approximation space).
+We will present proofs of discrete entropy stability in terms of integrals and $L^2$ projections, which we assume are computed using an appropriate quadrature rule.
+We stress that, while we utilize a continuous notation involving integrals rather than discrete sums, the proofs rely only on properties of quadrature-based integration and $L^2$ projection.
+
 The outline of the paper is as follows: Section~\ref{sec:intro} will briefly review the construction of entropy conservative collocation methods on a single element.
@@
 These methods were introduced in \cite{fisher2013high, carpenter2014entropy} for tensor product elements, though we adopt methods of proof introduced more recently in \cite{gassner2017br1, chen2017entropy}.
-
-%\note{Specify what we mean by diagonal norm collocation DG method.}
+These methods all share a basis in diagonal norm summation by parts operators.
+Because they can be derived from a nodal discontinuous Galerkin framework through the collocation of nodal points and quadrature points, we refer to them as ``diagonal norm collocation DG methods'' in this work.
 
@@
 We note that, unlike the proof of continuous entropy conservation, the proof of discrete entropy conservation does not require the use of the chain rule, which may not hold at the discrete level.
-
-The proof of entropy conservation can also be extended beyond a single element by specifying appropriate simultaneous approximation terms (SBP-SAT).
+This proof of entropy conservation can also be extended beyond a single element by specifying appropriate simultaneous approximation terms (SBP-SAT).
 This can be done by taking $\bm{f}_i^*$ to be the entropy conservative flux $f_S(u_L,u_R)$ at the interface between two elements \cite{carpenter2014entropy, gassner2017br1, chen2017entropy}.
 Similarly, employing an entropy stable flux at element interfaces results in an entropy stable method which satisfies a global entropy inequality.
-%\end{remark}
+
+It is important to emphasize that conservation of entropy does not, in general, guarantee stability of the numerical scheme unless the solution satisfies additional constraints.
+For example, the convex entropy associated with the compressible Euler and Navier-Stokes equations is well-defined only if the density and pressure are positive, and steps must be taken in any numerical scheme to guarantee that the discrete solution of such constraints.
+For DG methods, this is most commonly done using bound and positivity-preserving limiters, such as those introduced in \cite{zhang2010positivity, zhang2012maximum}.
 
@@
 
+\note{Fix notation: switch $V_h$ to $P^N$}
 We assume that the domain $\Omega \in \mathbb{R}^d$ is exactly triangulated by a mesh which consists of the union of non-overlapping elements $D^k$.
@@
 %\end{equation}
-Assuming smooth $u$, continuous energy conservation can be shown by multiplying by $u$, integrating over $[-1,1]$, and using $\pd{u}{t}u = \frac{1}{2}\pd{u^2}{t}$
+Assuming smooth $u$, Burgers' equation conserves a square entropy $U(\bm{u}) = u^2/2$.
+This can be shown by multiplying by $u$, integrating over $[-1,1]$, and using $\pd{u}{t}u = \frac{1}{2}\pd{u^2}{t}$
 \begin{equation}
@@
 This is equivalent to the skew-symmetric split formulation described in \cite{gassner2013skew}; however, as noted in \cite{chen2017entropy}, it can be derived from flux differencing without explicit knowledge of the skew form.
-We will show later in Section~\ref{sec:recovery} that a natural extension of this formulation to multiple elements recovers the weak form of the method introduced in \cite{ranocha2017extended}.
+We will show later in Section~\ref{sec:recovery} that a natural extension of this formulation to multiple elements recovers the weak form of the scheme introduced in \cite{ranocha2017extended}.
 
@@
 \begin{assumption}
-The numerical flux $\bm{f}_S(x,y)$ admits an absolutely convergent expansion of the form
+The numerical flux function $\bm{f}_S$ admits an absolutely convergent expansion of the form 
 \[
@@
 \]
-where $\bm{f}_i, \bm{g}_i$ are differentiable.  
+where $\bm{f}_i, \bm{g}_i$ are real and bounded over each element $D^k$.  
 \label{eq:assumption}
 \end{assumption}
-This assumption is satisfied, for example, by $\bm{f}_S$ which are analytic and admit a convergent Taylor expansion in $x$ and $y$.
+This assumption is satisfied, for example, by $\bm{f}_S$ which are real analytic and thus admit a Taylor expansion in $x$ and $y$.
 In Appendix~\ref{appendix:A}, we show that this assumption also holds for the choices of flux functions $\bm{f}_S$ considered in this work, assuming the solution satisfies physical bound constraints.
-However, we do need this assumption in computation implementations of the proposed methods.
+However, we note that this assumption is not required in the computation implementations of the proposed methods, and may be an overly restrictive condition on $\bm{f}_S$.
 
@@
 \begin{equation}
-\LRp{\pd{\bm{u}}{t},\bm{w}}_{\widehat{D}} + \LRp{\left.\LRp{2D_N \bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x},\bm{w}}_{\widehat{D}}= \LRa{\bm{f}^*-\bm{f}(\bm{u}_x),\bm{w}}_{\partial \widehat{D}}, \qquad \bm{w}\in V_h,
+\LRp{\pd{\bm{u}}{t},\bm{w}}_{\widehat{D}} + \LRp{\left.\LRp{2D_N \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},\bm{w}}_{\widehat{D}}= \LRa{\bm{f}^*-\bm{f}(\bm{u}_{\bm{v}}),\bm{w}}_{\partial \widehat{D}}, \qquad \bm{w}\in V_h,
 \label{eq:ecdg_oneelem}
 \end{equation}
-where $\bm{f}^*$ is a numerical flux used to impose boundary conditions weakly, and $\bm{u}_x,\bm{u}_y$ are evaluations of $\bm{u}$ in terms of the projections of entropy variables $\bm{v}(\bm{u}(x))$
+where $\bm{f}^*$ is a numerical flux used to impose boundary conditions weakly, and $\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y)$ are evaluations of $\bm{u}$ in terms of the projections of entropy variables $\bm{v}(\bm{u}(x))$
 \begin{equation}
-%\bm{u}_x = \bm{u}(\Pi_N\bm{v}(\bm{u}(x))), \qquad \bm{u}_y = \bm{u}(\Pi_N\bm{v}(\bm{u}(y))).
-\bm{u}_x(x) = \bm{u}\LRp{(\Pi_N\bm{v})(x)}, \qquad \bm{u}_y(y) = \bm{u}\LRp{(\Pi_N\bm{v})(y)}.
+%\bm{u}_{\bm{v}}(x) = \bm{u}(\Pi_N\bm{v}(\bm{u}(x))), \qquad \bm{u}_{\bm{v}}(y) = \bm{u}(\Pi_N\bm{v}(\bm{u}(y))).
+\bm{u}_{\bm{v}}(x) = \bm{u}\LRp{(\Pi_N\bm{v})(x)}.
 \label{eq:projectedentropyvars}
@@
 \begin{align*}
-&\LRp{\left.\LRp{D_N \sum_{i=1}^{\infty}\bm{f}_i(\bm{u}_x)\bm{g}_i(\bm{u}_y)}\right|_{y=x}, \Pi_N\bm{v}(x)}_{\widehat{D}} = \sum_{i=1}^{\infty} \LRp{{D_N \bm{f}_i(\bm{u}_x)}, {\bm{g}_i(\bm{u}_x)\Pi_N\bm{v} }}_{\widehat{D}}\\
-&= \sum_{i=1}^{\infty}\LRa{\bm{f}_i(\bm{u}_x) \bm{g}_i(\bm{u}_x),\Pi_N\bm{v}}_{\partial \widehat{D}}  - \LRp{ \bm{f}_i(\bm{u}_x), D_N \LRp{\bm{g}_i\Pi_N\bm{v} }}_{\widehat{D}}\\
-&= \LRa{\bm{f}_S(\bm{u}_x,\bm{u}_x), \Pi_N\bm{v}(x)}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N \Pi_N\LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y)\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}}\\
-&= \LRa{\bm{f}(\bm{u}_x), \Pi_N\bm{v}(x)}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N \Pi_N\LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y)\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}},
+&\LRp{\left.\LRp{D_N \sum_{i=1}^{\infty}\bm{f}_i(\bm{u}_{\bm{v}}(x))\bm{g}_i(\bm{u}_{\bm{v}}(y))}\right|_{y=x}, \Pi_N\bm{v}(x)}_{\widehat{D}} = \sum_{i=1}^{\infty} \LRp{{D_N \bm{f}_i(\bm{u}_{\bm{v}})}, {\bm{g}_i(\bm{u}_{\bm{v}})\Pi_N\bm{v} }}_{\widehat{D}}\\
+&= \sum_{i=1}^{\infty}\LRa{\bm{f}_i(\bm{u}_{\bm{v}}) \bm{g}_i(\bm{u}_{\bm{v}}),\Pi_N\bm{v}}_{\partial \widehat{D}}  - \LRp{ \bm{f}_i(\bm{u}_{\bm{v}}), D_N \LRp{\bm{g}_i\Pi_N\bm{v} }}_{\widehat{D}}\\
+&= \LRa{\bm{f}_S(\bm{u}_{\bm{v}},\bm{u}_{\bm{v}}), \Pi_N\bm{v}(x)}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N \LRp{\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))\Pi_N\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}}\\
+&= \LRa{\bm{f}(\bm{u}_{\bm{v}}), \Pi_N\bm{v}(x)}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N\LRp{\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))\Pi_N\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}},
 \end{align*}
@@
 %where $\Pi_N^x, \Pi_N^y$ denote $L^2$ projections taken with respect to the $x$ and $y$ coordinates, respectively.  
-The volume terms can be manipulated to yield
+Noting that 
+\[
+\left.\LRp{D_N \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x}\bm{v}(x) = \left.\LRp{D_N \LRp{\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))\bm{v}(y)}}\right|_{y=x},
+\]
+the volume terms can be manipulated to yield
 \begin{align*}
 %&\LRp{\left.\LRp{D_N \LRp{\bm{f}_S(\bm{u}(\Pi_N\bm{v}(x)),\bm{u}(\Pi_N\bm{v}(y)))\LRp{\Pi_N\bm{v}(y)-\Pi_N\bm{v}(x)} }}\right|_{y=x},1}_{\widehat{D}} \\
-&\LRp{\left.\LRp{D_N \LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y)\LRp{\Pi_N\bm{v}(y)-\Pi_N\bm{v}(x)} }}\right|_{y=x},1}_{\widehat{D}} \\
+&\LRp{\left.\LRp{D_N \LRp{\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))\LRp{\Pi_N\bm{v}(y)-\Pi_N\bm{v}(x)} }}\right|_{y=x},1}_{\widehat{D}} \\
 &= \LRp{ \left.\LRp{D_N\LRp{\psi(y)-\psi(x) }}\right|_{y=x},1}_{\widehat{D}} = \LRp{-D_N \psi(x),1}_{\widehat{D}} = \LRa{-\psi,1}_{\partial \widehat{D}},
@@
 For collocation-type DG methods, $L^2$ projection and interpolation are equivalent, which implies that evaluating the flux $\bm{f}_S$ using point-wise nodal values of the solution satisfies the entropy conservation condition given in Definition~\ref{def:tadmor}.
-For collocation-type DG methods, the method can also be simplified by noting that $\bm{V}_q,\bm{P}_q$ are identity matrices.
+For diagonal norm collocation-type DG methods, the method can also be simplified by noting that the projection operator reduces to a nodal interpolation operator.
 
@@
 However, the entropy conservation condition given in Definition~\ref{def:tadmor} holds so long as the flux is evaluated in a manner consistent with the entropy variables.
-This motivates (\ref{eq:projectedentropyvars}) as a non-standard way of evaluating the flux, which is the primary difference between the construction of entropy conservative schemes for collocation-type DG methods and projection-based DG methods.
+This motivates (\ref{eq:projectedentropyvars}) as a non-standard way of evaluating the flux, which is the primary difference between the construction of entropy conservative schemes for diagonal norm collocation-type DG methods and projection-based DG methods.
+
+\begin{remark}
+In the proof of Theorem~\ref{lemma:ec1}, Assumption~\ref{eq:assumption} is used only to formally apply Lemma~\ref{lemma:dgd_local}.
+The use of this assumption can be avoided by proving instead a discrete conservation of entropy in terms of a fixed quadrature rule and the quadrature-based matrices introduced in Section~\ref{sec:implementation}.
+\end{remark}
 
@@
 
-\note{It's currently unclear how to distinguish whether different operators are high order accurate. }
-
-\note{Kopriva's note ``Stability of Overintegration Methods for Nodal Discontinuous Galerkin Spectral Element Methods'' does not apply here, because the formulation is entropy stable for any quadratures which evaluate $P^N$ exactly (assuming affine meshes!).
+%\note{It's currently unclear how to distinguish whether different operators are high order accurate. }
+%\note{Kopriva's note ``Stability of Overintegration Methods for Nodal Discontinuous Galerkin Spectral Element Methods'' does not apply here, because the formulation is entropy stable for any quadratures which evaluate $P^N$ exactly (assuming affine meshes!).
 Certain terms are under-integrated, but they do not contribute to the interplay between volume and surface terms described by Kopriva.}
 
-\subsection{Conservation of entropy on multiple elements}
-
+\subsection{Conservation of entropy and entropy stability on multiple elements}
+\label{sec:multipleelems}
 Extending the entropy stable DG formulation (\ref{eq:ecdg_oneelem}) is done by combining the flux differencing with the global DG differential operator $D^x_h$
 \begin{equation}
-\sum_{k}\LRp{\pd{\bm{u}}{t},\bm{w}}_{D^k} + \LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x},\bm{w}}_{D^k} = \LRa{\bm{f}(\bm{u}_x)-\bm{f}^*,\bm{w}}_{\partial \Omega},
+\sum_{k}\LRp{\pd{\bm{u}}{t},\bm{w}}_{D^k} + \LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},\bm{w}}_{D^k} = \LRa{\bm{f}(\bm{u}_{\bm{v}})-\bm{f}^*,\bm{w}}_{\partial \Omega},
 %\int_{D^k}\pd{\bm{u}}{t} \bm{w} + \left.\LRp{2D^x_h\bm{f}_S\LRp{\bm{u}(\Pi_N\bm{v}({x})),\bm{u}(\Pi_N\bm{v}({y}))}}\right|_{{y}={x}}\bm{w} + \LRa{\bm{f}^*-\bm{f}(\bm{u}),\bm{w}}_{\partial \Omega}= 0, \qquad \forall \bm{w}\in V_h^n,
@@
 
-\note{Add assumptions about $\bm{u}(x)$ - needs to be locally analytic.}
+%\note{Add assumptions about $\bm{u}(x)$ - needs to be locally analytic.}
 
@@
 \begin{theorem}
+\label{thm:ecdg_global}
 The formulation (\ref{eq:ecdg}) is globally entropy conservative in the sense that
@@
 
+Finally, we note that the analysis in previous sections has focused on the construction of entropy conservative schemes.
+However, entropy is only conserved for smooth solutions and should be dissipated away in the presence of discontinuities and shocks.
+To this end, one can construct discretely entropy stable schemes by adding additional dissipative terms (for example, by adding HLL or Lax-Friedrichs penalization \cite{chen2017entropy} or matrix dissipation terms \cite{chandrashekar2013kinetic, winters2017uniquely}).
+For the numerical experiments presented here, we apply a Lax-Friedrichs penalization, augmenting the flux function at element interfaces with the additional term
+\[
+ \bm{f}_S\LRp{\bm{u}_L,\bm{u}_R}\Rightarrow \bm{f}_S\LRp{\bm{u}_L,\bm{u}_R} - \frac{\lambda}{2}\jump{\bm{u}_{\bm{v}}},
+\]
+where $\lambda$ is an estimate of the maximum eigenvalue of the flux Jacobian.
+For an appropriate approximation of $\lambda$, Lax-Friedrichs penalization dissipates entropy \cite{chen2017entropy}.
+We note that, since conservation of entropy requires testing with the projection of the entropy variables, the jump term should involve the evaluations of the conservation variables in terms of the projected entropy variables $\bm{u}_{\bm{v}} = \bm{u}\LRp{\Pi_N \bm{v}}$ in order to guarantee entropy dissipation.
+
+%\note{Unlike collocative DG, entropy stability must be in terms of $\jump{\bm{u}\LRp{\Pi_N \bm{v}}}$ to guarantee discrete entropy decay.  }
+
 \subsubsection{Local conservation}
@@
 It is also straightforward to show that the formulation (\ref{eq:ecdg}) is locally conservative.
-We follow \cite{chen2017entropy} and take $\bm{w} = \mathbbm{1}_{D^k}$ in the formulation (\ref{eq:ecdg}) yields
-\[
-\sum_{k}\LRp{\pd{\bm{u}}{t},1}_{D^k} + \LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x},1}_{D^k} = \LRa{\bm{f}(\bm{u}_x)-\bm{f}^*,1}_{\partial \Omega \cap D^k}.  
-\]
-Using Lemma~\ref{lemma:dgd_local} and the symmetry of $\bm{f}_S$ yields that
-\[
-\LRp{\left.\LRp{D^x_h\bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x},1}_{D^k} = \LRa{\bm{f}(\bm{u}_x) ,1}_{\partial D^k} - \LRp{\left.\LRp{D^x_h\bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x},1}_{D^k}.
-\]
-Substituting this back into the first equation cancels volume terms, leaving
-\[
-\sum_{k}\LRp{\pd{\bm{u}}{t},1}_{D^k} + \LRa{\bm{f}(\bm{u}_x) ,1}_{\partial D^k \setminus \partial \Omega} + \LRa{\bm{f}^*,1}_{\partial \Omega \cap D^k} = 0.  
-\]
+We follow \cite{chen2017entropy} and take $\bm{w} = \mathbbm{1}_{D^k}$ in the formulation (\ref{eq:ecdg})
+\[
+\sum_{k}\LRp{\pd{\bm{u}}{t},1}_{D^k} + \LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},1}_{D^k} = \LRa{\bm{f}(\bm{u}_{\bm{v}})-\bm{f}^*,1}_{\partial \Omega \cap D^k}.
+\]
+Using Assumption~\ref{eq:assumption}, Lemmas~\ref{lemma:dgd_local} and \ref{lemma:ibp}, and the symmetry of $\bm{f}_S$, one can show that
+\begin{align*}
+\LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},1}_{D^k} =& \sum_{i=1}^{\infty} \LRp{2D_N \bm{f}_i(\bm{u}_{\bm{v}}),\bm{g}_i(\bm{u}_{\bm{v}})}_{D^k} + \LRa{\jump{\bm{f}_i(\bm{u}_{\bm{v}})},\bm{g}_i(\bm{u}_{\bm{v}})}_{\partial D^k}\\
+=& \sum_{i=1}^{\infty} \LRp{D_N \bm{f}_i(\bm{u}_{\bm{v}}),\bm{g}_i(\bm{u}_{\bm{v}})}_{D^k} - \LRp{D_N \bm{f}_i(\bm{u}_{\bm{v}}),\bm{g}_i(\bm{u}_{\bm{v}})}_{D^k} \\
+&+ \LRa{\bm{f}_i(\bm{u}_{\bm{v}}),\bm{g}_i(\bm{u}_{\bm{v}})}_{\partial D^k} + \LRa{\bm{f}_i(\bm{u}_{\bm{v}}^+)-\bm{f}_i(\bm{u}_{\bm{v}}),\bm{g}_i(\bm{u}_{\bm{v}})}_{\partial D^k}\\
+=& \LRp{D_N \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y)),1}_{D^k} -  \LRp{D_N \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y)),1}_{D^k}\\
+&+ \LRa{\bm{f}_S(\bm{u}_{\bm{v}}^+,\bm{u}_{\bm{v}}),1}_{\partial D^k}.
+\end{align*}
+The volume terms cancel, leaving 
+\[
+\sum_{k}\LRp{\pd{\bm{u}}{t},1}_{D^k} + \LRa{\bm{f}_S(\bm{u}_{\bm{v}}^+,\bm{u}_{\bm{v}}),1}_{\partial D^k} = \LRa{\bm{f}(\bm{u}_{\bm{v}})-\bm{f}^*,1}_{\partial \Omega \cap D^k}.  
+\]
+On the domain boundary $\partial \Omega$, $\bm{u}_{\bm{v}}^+ = \bm{u}_{\bm{v}}$ and $\bm{f}_S(\bm{u}_{\bm{v}}^+,\bm{u}_{\bm{v}}) = \bm{f}(\bm{u}_{\bm{v}})$.
+This implies that the statement of local conservation replaces $\bm{f}_S$ with the boundary condition numerical flux $\bm{f}^*$ on element boundaries which intersect with the domain boundary.
+%\[
+%\sum_{k}\LRp{\pd{\bm{u}}{t},1}_{D^k} + \LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},1}_{D^k} = \LRa{\bm{f}(\bm{u}_{\bm{v}}(x))-\bm{f}^*,1}_{\partial \Omega \cap D^k}.
+%\]
+
 We note that this definition of local conservation is given in terms of the conservation variables evaluated as a function of the projected entropy variables
 \[
-\bm{u}_x(x) = \bm{u}\LRp{\LRp{\Pi_N \bm{v}}(x)}.
+\bm{u}_{\bm{v}}(x)(x) = \bm{u}\LRp{\LRp{\Pi_N \bm{v}}(x)}.
 \]
 It is not immediately clear whether this statement of local conservation satisfies the conditions of the classic Lax-Wendroff theorem.
-However, this form of local conservation does satisfy a generalized definition of local conservation given in \cite{shi2017local}, and is sufficient to guarantee convergence to the weak solution.
+However, this form of local conservation does satisfy the generalized definition of local conservation given in \cite{shi2017local}, which is sufficient to guarantee convergence to the weak solution.
 
@@
 \subsection{On implementation in terms of the flux function}
-
+\label{sec:implementation}
 While it is possible to use (\ref{eq:ecdg}) to derive explicit DG weak formulations, this can be difficult for complicated flux functions.
@@
 \begin{align*}
-\LRp{\left.\LRp{D^x_h \bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x},\bm{w}}_{\Omega} &= \sum_k \sum_{i=1}^{\infty} \LRp{\pd{\Pi_N \bm{f}_i(\bm{u})}{x},\bm{g}_i(\bm{u})\bm{w}}_{D^k}\\ 
-&+  \sum_k \sum_{i=1}^{\infty}\frac{1}{2}\LRa{\bm{f}_i(\bm{u}^+)\bm{g}_i(\bm{u}) - \bm{g}_i(\bm{u})\Pi_N\bm{f}_i(\bm{u}),\bm{w}\bm{n}_x}_{\partial D^k}\\
-&+  \sum_k \sum_{i=1}^{\infty}\frac{1}{2}\LRp{\bm{g}_i(\bm{u})L\LRp{\bm{f}_i(\bm{u}^+) - \Pi_N\bm{f}_i(\bm{u})}\bm{n}_x,\bm{w}}_{D^k},
+\LRp{\left.\LRp{D^x_h \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},\bm{w}}_{\Omega} &= \sum_k \sum_{i=1}^{\infty} \LRp{\pd{\Pi_N \bm{f}_i(\bm{u}_{\bm{v}})}{x},\bm{g}_i(\bm{u}_{\bm{v}})\bm{w}}_{D^k}\\
+&+  \sum_k \sum_{i=1}^{\infty}\frac{1}{2}\LRa{\bm{f}_i(\bm{u}^+_{\bm{v}})\bm{g}_i(\bm{u}_{\bm{v}}) - \bm{g}_i(\bm{u}_{\bm{v}})\Pi_N\bm{f}_i(\bm{u}_{\bm{v}}),\bm{w}\bm{n}_x}_{\partial D^k}\\
+&+  \sum_k \sum_{i=1}^{\infty}\frac{1}{2}\LRp{\bm{g}_i(\bm{u}_{\bm{v}})L\LRp{\bm{f}_i(\bm{u}^+_{\bm{v}}) - \Pi_N\bm{f}_i(\bm{u}_{\bm{v}})}\bm{n}_x,\bm{w}}_{D^k},
 \end{align*}
@@
 \begin{align}
-\LRp{\left.\LRp{D^x_h \bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x},\bm{w}}_{\Omega} &= \sum_k \LRp{\left.\pd{\Pi_N \bm{f}_S(\bm{u}_x,\bm{u}_y)}{x}\right|_{y=x},\bm{w}}_{D^k}\nonumber\\ 
-&+ \frac{1}{2}\LRa{\bm{f}_S(\bm{u}^+,\bm{u}) - \left.\Pi_N^x\bm{f}_S(\bm{u}_x,\bm{u}_y)\right|_{y=x},\bm{w}\bm{n}_x}_{\partial D^k} \label{eq:eval}\\
-&+ \frac{1}{2}\LRp{\left.\LRp{L^x\LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y) - \Pi_N^x\bm{f}_S(\bm{u}_x,\bm{u}_y)}\bm{n}_x}\right|_{y=x},\bm{w}}_{D^k},\nonumber
+\LRp{\left.\LRp{D^x_h \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},\bm{w}}_{\Omega} &= \sum_k \LRp{\left.\pd{\Pi_N \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}{x}\right|_{y=x},\bm{w}}_{D^k}\nonumber\\
+&+ \frac{1}{2}\LRa{\bm{f}_S(\bm{u}^+_{\bm{v}},\bm{u}_{\bm{v}}) - \left.\Pi_N^x\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))\right|_{y=x},\bm{w}\bm{n}_x}_{\partial D^k} \label{eq:eval}\\
+&+ \frac{1}{2}\LRp{\left.\LRp{L^x\LRp{\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y)) - \Pi_N^x\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\bm{n}_x}\right|_{y=x},\bm{w}}_{D^k},\nonumber
 \end{align}
 where the superscripts $L^x, \Pi_N^x$ are used to emphasize that the lift and projection operators are applied in $x$ and not in $y$.  
-This can be used to practically evaluate $\left.\LRp{\bm{D}^x_h \bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x}$ using only reference operators and evaluations of $\bm{f}_S$.
+This can be used to practically evaluate $\left.\LRp{\bm{D}^x_h \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x}$ using only reference operators and evaluations of $\bm{f}_S$.
 However, unlike previous entropy conservative collocation DG methods, the boundary terms in (\ref{eq:eval}) interact with volume terms through the $L^2$ projection and the lift operator.
@@
 
-Finally, we show how to recover a computational form of flux differencing analogous to those presented in \cite{gassner2016split, chen2017entropy}.
+Finally, we discuss how to recover a computational form of flux differencing analogous to the approaches presented in \cite{gassner2016split, chen2017entropy}.
 We will describe how to compute the three non-standard terms which appear within (\ref{eq:eval}):
 \begin{align}
-\LRp{\left.\pd{\Pi_N {f}_S({u}_x,{u}_y)}{x}\right|_{y=x},{w}}_{D^k}\label{eq:voleval}\\
-\LRa{\left.\Pi_N^x{f}_S({u}_x,{u}_y)\right|_{y=x},{w}\bm{n}_x}_{\partial D^k}\label{eq:surfeval}\\
-\LRp{\left.\LRp{L^x\LRp{{f}_S({u}_x,{u}_y) - \Pi_N^x{f}_S({u}_x,{u}_y)}\bm{n}_x}\right|_{y=x},{w}}_{D^k}\label{eq:volsurfeval}
+\LRp{\left.\pd{\Pi_N {f}_S({u}(x),{u}(y))}{x}\right|_{y=x},{w}}_{D^k}\label{eq:voleval}\\
+\LRa{{f}_S({u}^+,{u})-\left.\Pi_N^x{f}_S({u}(x),{u}(y))\right|_{y=x},{w}\bm{n}_x}_{\partial D^k}\label{eq:surfeval}\\
+\LRp{\left.\LRp{L^x\LRp{{f}_S({u}(x),{u}(y)) - \Pi_N^x{f}_S({u}(x),{u}(y))}\bm{n}_x}\right|_{y=x},{w}}_{D^k}\label{eq:volsurfeval}
 \end{align}
@@
 \[
-V_h = {\rm span}\LRc{\phi_j(\widehat{x})}_{j=1}^{N_p},
+V_h = {\rm span}\LRc{\phi_i(\widehat{x})}_{i=1}^{N_p},
 \]
@@
 \[
-\bm{P}_q = \bm{M}^{-1}\bm{V}_q^T\bm{W}.
-\]
-Let $\bm{D}_x$ denote the differentiation matrix defined implicitly through
+\bm{P}_q = \bm{M}^{-1}\bm{V}_q^T\bm{W},
+\]
+which maps a function (in terms of its evaluation at quadrature points) to coefficients of the basis $\phi_j(x)$.  Next, let $\bm{D}_x$ denote the differentiation matrix defined implicitly through
 \[
@@
 
-We also introduce matrices used to compute boundary terms.
-Let $\bm{V}_f$ denote the matrix which interpolates to boundary values\footnote{In generalizing to higher dimensions, $\bm{V}_f$ would be the matrix which interpolates to surface quadrature points.}
+We first describe how to evaluate the volume term (\ref{eq:voleval}) using these matrices.  Let $\bm{F}_S$ denote the $N_q\times N_q$ matrix constructed by evaluating $f_S$ at volume quadrature points
+\[
+\LRp{\bm{F}_S}_{ij} = f_S(u(x_i),u(x_j)).  
+\]
+Then, integrating (\ref{eq:voleval}) using quadrature is equivalent to computing
+\[
+\LRp{\left.\pd{\Pi_N {f}_S({u}(x),{u}(y))}{x}\right|_{y=x},{w}}_{D^k} = \bm{w}^T \bm{V}_q^T\bm{W} {\rm diag}\LRp{\bm{V}_q\bm{D}_x\bm{P}_q \bm{F}_S},  
+\]
+where $\bm{w}$ denote the vector of coefficients for the weighting function $w(x)$.
+
+The diagonal operation can be replaced by a more convenient evaluation.  Let $\bm{A}\circ\bm{B}$ denote the Hadamard product of two matrices $\bm{A},\bm{B} \in \mathbb{R}^{N_q\times N_q}$, such that 
+\[
+\LRp{\bm{A}\circ\bm{B}}_{ij} = \bm{A}_{ij}\bm{B}_{ij}, \qquad 1\leq i,j \leq N_q.
+\]
+The Hadamard product of two matrices can also be used to conveniently represent the diagonal of a product of matrices \cite{styan1973hadamard}:
+\[
+{\rm diag}\LRp{\bm{A}\bm{B}^T} = \LRp{\bm{A}\circ \bm{B}}\bm{1}.  
+\]
+Using this property and the symmetry of $\bm{F}_S$, we can rewrite the diagonal operation as follows
+\[
+{\rm diag}\LRp{\bm{V}_q \bm{D}_q \bm{P}_q \bm{F}_S} = \LRp{\LRp{\bm{V}_q \bm{D}_q \bm{P}_q} \circ \bm{F}_S}\bm{1}.
+\]
+A computational advantage of evaluation in terms of the Hadamard product is that $\bm{F}_S$ does not need to be formed explicitly and stored.
+
+The same approach may also be applied to (\ref{eq:surfeval}) and (\ref{eq:volsurfeval}).
+We focus first on (\ref{eq:surfeval}) and introduce matrices used to compute boundary terms.
+Let $\bm{V}_f$ denote the matrix which interpolates to boundary values\footnote{In generalizing to higher dimensions, $\bm{V}_f$ would be the matrix which interpolates to boundary points.}
 \[
@@
 In one dimension, this is simply the $2\times 2$ identity matrix.
-This allows us to define the quadrature-based lift matrix
-\[
-\bm{L} = \bm{M}^{-1}\bm{V}_f^T \bm{W}_f.
-\]
-Together, these matrices can be used to interpolate the output of the differentiation and lift matrices to volume quadrature points
-\[
-\bm{D}_q = \bm{V}_q \bm{D}_x \bm{P}_q, \qquad \bm{L}_q = \bm{V}_q\bm{L}_q.
-\]
-
-Hadamard product \cite{horn2012matrix}
-%\subsection{Matrix and algebraic forms of entropy conservative DG methods}
-%
-%We note that it also possible to show that the proposed formulation (\ref{eq:ecdg}) is entropy conservative using matrix notation.  
-%
-%\note{Introduce global operators.}
-%
-%
-%This approach yields the following evaluation (in Matlab for lack of better notation): let 
-%\[
-%\verb|fS = @(uL,uR) uL.^2+uL.*uR+uR.^2|
-%\]
-%be the flux function.
-Let \verb+uq = Vq * u+ denote the interpolation of $u$ to quadrature points, and let \verb+Dx+ denote the differentiation matrix.
-Define \verb+Pq = M\(Vq' * diag(wq))+, where \verb+wq+ are quadrature weights, and let \verb+[ux uy] = meshgrid(uq)+.
-Then, assuming element-constant Jacobians, the split form can be evaluated via
-%\[
-%\verb+Pq * diag((Vq*Dx*Pq) * fS(ux, uy)) = Pq * sum((Vq* Dx* Pq) .* fS(ux, uy), 2).+
-%\]
-%The latter expression can be performed by computing \verb+fS(ux,uy)+ on the fly, avoiding the need to store it as a matrix.
-It can also be generalized to handle non-square matrices in a very similar way.
-%
-%
-%\note{Introduce definition of $\bm{f}(\bm{u}_q)$ as $f(\bm{u}_q)$.}
-
-\subsection{Entropy stability}
-
-The analysis in previous sections has focused on the construction of entropy conservative schemes.
-However, entropy is only conserved for smooth solutions and should be dissipated away in the presence of discontinuities and shocks.
-To this end, one can construct discretely entropy stable schemes by adding additional dissipative terms (for example, by adding HLL or Lax-Friedrichs penalization \cite{chen2017entropy} or matrix dissipation terms \cite{chandrashekar2013kinetic, winters2017uniquely}).
-
-\note{Unlike collocative DG, entropy stability must be in terms of $\jump{\bm{u}\LRp{\Pi_N \bm{v}}}$ to guarantee discrete entropy decay.  }
+We now define $\bm{F}_S^f$ to be the matrix constructed by evaluating $\bm{F}_S$ at both volume quadrature and boundary points
+\[
+\LRp{\bm{F}_S^f}_{ij} = f_S\LRp{u(x_i),u\LRp{x^f_j}}, \qquad 1 \leq i \leq N_q, \quad j = 1,2.  
+\]
+We also define $\bm{F}^+_S$ to be the vector of evaluations of the flux function $f_S\LRp{u^+,u}$ at the boundary points of $D^k$.  Using these matrices, we can then evaluate 
+\begin{align*}
+\LRa{{f}_S({u}^+,{u})-\left.\Pi_N^x{f}_S({u}(x),{u}(y))\right|_{y=x},{w}\bm{n}_x}_{\partial D^k} &= \bm{w}^T\bm{V}_f^T \bm{W}_f {\rm diag}(\bm{n}_x)\LRp{\bm{F}_S^+ - {\rm diag}\LRp{\bm{V}_f\bm{P}_q \bm{F}^f_S}} \\
+&= \bm{w}^T\bm{V}_f^T \bm{W}_f {\rm diag}(\bm{n}_x)\LRp{\bm{F}_S^+ - \LRp{\LRp{\bm{V}_f\bm{P}_q} \circ \LRp{\bm{F}^f_S}^T}\bm{1}}.
+\end{align*}
+This step can be interpreted as coupling together evaluations of the flux function at volume and boundary points.
+We note that this is the only step in evaluating (\ref{eq:eval}) which requires non-local data (to compute $\bm{F}^+_S$).
+
+In order to evaluate the final term (\ref{eq:volsurfeval}), we need to introduce the quadrature-based lift matrix 
+\[
+\bm{L} = \bm{M}^{-1}\bm{V}_f^T \bm{W}_f,
+\]
+which (as the name implies) ``lifts'' a function from the boundary (in terms of its evaluation at surface quadrature points) to coefficients of a basis defined on the interior of the element.
+Then, we have that
+\begin{align*}
+&\LRp{\left.\LRp{L^x\LRp{{f}_S({u}(x),{u}(y)) - \Pi_N^x{f}_S({u}(x),{u}(y))}\bm{n}_x}\right|_{y=x},{w}}_{D^k} \\
+&= \bm{w}^T\bm{V}_q^T\bm{W}{\rm diag}\LRp{\bm{V}_q\bm{L}{\rm diag}(\bm{n}_x) \LRp{\LRp{\bm{F}_S^f}^T - \bm{V}_f\bm{P}_q \bm{F}_S^f}}\\
+&= \bm{w}^T\bm{V}_q^T\bm{W}\LRp{ \LRp{\bm{V}_q\bm{L}{\rm diag}(\bm{n}_x)} \circ {\bm{F}_S^f} - \LRp{\bm{V}_q\bm{L}{\rm diag}(\bm{n}_x)\bm{V}_f\bm{P}_q}\circ {\bm{F}_S}}\bm{1}.
+\end{align*}
+This expression is (in a loose sense) analogous to the evaluation of (\ref{eq:surfeval}) in that it couples together volume and boundary evaluations of the flux function, but in the reverse order.
+
+It is worth pointing out that these sets of expressions can be simplified; for example, straightforward computations show that $\bm{W}\bm{V}_q\bm{L} = \LRp{\bm{W}_f{\rm diag}(\bm{n}_x)\bm{V}_f\bm{P}_q}^T$, and the computation of the volume term $\LRp{\LRp{\bm{V}_q \bm{D}_q \bm{P}_q} \circ \bm{F}_S}\bm{1}$ and the surface term $\LRp{\LRp{\bm{V}_q\bm{L}{\rm diag}(\bm{n}_x)\bm{V}_f\bm{P}_q}\circ {\bm{F}_S}}\bm{1}$ can be combined into one single operation.
+Furthermore, for specific choices of basis and quadrature, one can eliminate certain terms altogether.
+For example, the co-location of nodal and quadrature points (assuming that the dimension of the approximation space is identical to the number of quadrature points) reduces $\bm{V}_q, \bm{P}_q$ to identity matrices.
+Furthermore, if quadrature points coincide with boundary points (as with DG-SEM methods), the lift matrix $\bm{L}$ is zero except for entries which correspond to those boundary points.
+These assumptions greatly simplify both the formulation and implementation of entropy conservative/stable DG methods, and will be discussed more thoroughly in a future manuscript.
 
@@
 \label{sec:num}
-\note{We illustrate the entropy conservation and accuracy of the proposed schemes for three equations.
-We focus on shallow water and compressible Euler.
-Compared to Burgers' equation, shallow water has a skew symmetric form but the entropy variables are not contained in the test space, and compressible Euler has neither.
-}
-
-%\subsection{Burgers}
+
+In this section, we illustrate the entropy conservation and accuracy of the proposed schemes for the one dimensional compressible Euler equations.  
+All numerical experiments utilize the fourth order five-stage low-storage Runge-Kutta method Carpenter and Kennedy \cite{carpenter1994fourth}.
+Finally, following the derivation of stable timestep restrictions in \cite{chan2015gpu}, we define the timestep $\Delta t$ to be
+\[
+\Delta t = {\rm CFL} \times \frac{h}{C_N}, \qquad C_N = \frac{(N+1)^2}{2},
+\] 
+where $C_N$ is the one-dimensional constant in the trace inequality \cite{warburton2003constants}, and ${\rm CFL}$ is a user-defined constant.
+
+%Compared to Burgers' equation, shallow water has a skew symmetric form but the entropy variables are not contained in the test space, and compressible Euler has neither.  }
+
+%\subsection{Shallow water}
 %
-%We begin by computing $L^2$ errors and rates of convergence for smooth solutions of Burgers' equation.  
-%\subsection{Accuracy of the global DG differentiation operator}
-%
-%\note{Show that both ``correction'' and ``coupling'' terms are necessary for full $h^{N+1}$ convergence.  Removing either degrades convergence down to $h^N$.}
-
-%\subsection{Burgers' equation}
-%
-%As mentioned previously, the Burgers' equation and and an entropy conservative flux are given by
-%\[
-%\pd{u}{t} + \frac{1}{2}\pd{u^2}{2} = 0, \qquad f_S(u_L,u_R) = \frac{1}{6}\LRp{u_L^2 + u_Lu_R + u_R^2}.
-%\]
-
-\subsection{Shallow water}
-
-We now move on to the one-dimensional shallow water equations, which consist of a conservation equation and momentum equation
-\begin{align*}
-\pd{h}{t} + \pd{hv}{x} &= 0\\
-\pd{hv}{t} + \pd{\LRp{hv^2 + gh^2/2}}{x} + gh\pd{b}{x}&= 0.
-\end{align*}
-where $h$ is the water height, $v$ is the velocity, $g$ is the gravitational constant, and $b$ is bottom topography.
+%We begin with the one-dimensional shallow water equations, which consist of a conservation equation and momentum equation
+%\begin{align*}
+%\pd{h}{t} + \pd{hv}{x} &= 0\\
+%\pd{hv}{t} + \pd{\LRp{hv^2 + gh^2/2}}{x} + gh\pd{b}{x}&= 0.
+%\end{align*}
+%where $h$ is the water height, $v$ is the velocity, $g$ is the gravitational constant, and $b$ is bottom topography.
 The entropy associated with this system is the total energy $e = k + p$, where $k, p$ are the kinetic and potential energies
-\[
-k = hv^2/2, \qquad p = gh^2/2 + ghb.
-\]
-The entropy variables can be derived from the entropy $e$
-\[
-q_1 = \pd{e}{h} = g(h+b)-v^2/2, \qquad q_2 = \pd{e}{hv} = v,
-\]
-with an inverse mapping
-\[
-h = \frac{1}{g}\LRp{q_1 + \frac{q_2^2}{2}}, \qquad v = q_2.
-\]
-Multiplying the continuity equation by $q_1$ and the momentum equation by $q_2$ recovers a statement of conservation of energy
-\[
-\pd{e}{t} + \pd{}{x}\LRp{\frac{hv^3}{2} + g(hv)(h+b)} = 0.  
-\]
-Like the Burgers' equation, the shallow water equations are stable under an appropriate split formulation \cite{gassner2016well, chen2017entropy}.
+%\[
+%k = hv^2/2, \qquad p = gh^2/2 + ghb.
+%\]
+%The entropy variables can be derived from the entropy $e$
+%\[
+%q_1 = \pd{e}{h} = g(h+b)-v^2/2, \qquad q_2 = \pd{e}{hv} = v,
+%\]
+%with an inverse mapping
+%\[
+%h = \frac{1}{g}\LRp{q_1 + \frac{q_2^2}{2}}, \qquad v = q_2.
+%\]
+%Multiplying the continuity equation by $q_1$ and the momentum equation by $q_2$ recovers a statement of conservation of energy
+%\[
+%\pd{e}{t} + \pd{}{x}\LRp{\frac{hv^3}{2} + g(hv)(h+b)} = 0.  
+%\]
+%Like the Burgers' equation, the shallow water equations are stable under an appropriate split formulation \cite{gassner2016well, chen2017entropy}.
 However, unlike Burgers' equation, the shallow water equations are not entropy stable with respect to the square entropy.
 Additionally, a key difference between the shallow water equations and Burgers' equation is that, when discretizing with respect to the conservative variables $h, hv$, the entropy variables $v_1,v_2$ are not contained within the space $V_h$.
-
-The entropy conservative flux functions for the shallow water equations are given as follows:
-\begin{align*}
-\bm{f}^1_S(\bm{u}_L,\bm{u}_R)&= \avg{h v}\\
-\bm{f}^2_S(\bm{u}_L,\bm{u}_R) &= \avg{hv}\avg{v} + \frac{1}{2}g h_L h_R.  
-\end{align*}
-These correspond also to the skew-symmetric splitting introduced in \cite{gassner2016well}.
+%
+%The entropy conservative flux functions for the shallow water equations are given as follows:
+%\begin{align*}
+%\bm{f}^1_S(\bm{u}_L,\bm{u}_R)&= \avg{h v}\\
+%\bm{f}^2_S(\bm{u}_L,\bm{u}_R) &= \avg{hv}\avg{v} + \frac{1}{2}g h_L h_R.  
+%\end{align*}
+%These correspond also to the skew-symmetric splitting introduced in \cite{gassner2016well}.  
+%
+%\note{Add some analytic solution}
+%We note that important
 
@@
 
-We now test the proposed entropy conservative schemes on the one-dimensional compressible Euler equations, which correspond to the inviscid limit of the compressible Navier-Stokes equations
+The one-dimensional compressible Euler equations correspond to the inviscid limit of the compressible Navier-Stokes equations and are given as follows:
 \begin{align*}
@@
 The choice of convex entropy for the Euler equations is non-unique \cite{harten1983symmetric}.
-However, a unique entropy can be chosen by restricting to choices of entropy variables which symmetrize the viscous heat conduction term in the compressible Navier-Stokes equations \cite{hughes1986new}, which leads to $U(\bm{u})$ of the form
+However, a unique entropy can be chosen by restricting to choices of entropy variables which symmetrize the viscous heat conduction term in the compressible Navier-Stokes equations \cite{hughes1986new}.
+This leads to $U(\bm{u})$ of the form
 \[
@@
 \]
+In order for the entropy $U(\bm{u})$ to be well defined, we require the assumption that the discrete density and pressure solutions are bounded away from zero
+\begin{equation}
+\rho \geq \rho_0 > 0,  \qquad p \geq p_0 > 0.  \label{eq:assumption2}
+\end{equation}
+This can be achieved using positivity-preserving limiters \cite{zhang2010positivity, zhang2012maximum}.
+These were not found to be necessary for the experiments in this work, and have thus not been implemented in our numerical simulations.
 
@@
 \[
-\beta = \frac{\rho}{2p}.
-\]
+\beta = \frac{\rho}{2p}
+\]
+and the logarithmic mean
+\[
+\avg{u}^{\log} = \frac{u_L - u_R}{\log{u_L}- \log{u_R}}.  
+\]
+We note that, because the direct evaluation of the logarithmic mean is numerically sensitive for $u_L\approx u_R$, we switch to evaluation using a high order accurate expansion introduced by Ismail and Roe \cite{ismail2009affordable} when $\LRb{u_L-u_R}<\epsilon$.
+
 These fluxes are both entropy conservative and kinetic energy preserving.
-
 Unlike the shallow water equations, these entropy conservative fluxes do not correspond to stable split formulations of the Euler equations.
@@
 
-In evaluating the logarithmic mean, we utilize the procedure described by Ismail and Roe in \cite{ismail2009affordable}.  
-\note{Show how RHS residual depends on tolerance on the log mean}
+We also present results which utilize the dissipative local Lax-Friedrichs flux described in Section~\ref{sec:multipleelems}, where the value of $\lambda$ is estimated by
+\[
+\lambda = \max_{\bm{u}^+, \bm{u}} \LRc{\LRb{u} + c}, 
+%\frac{1}{2}\LRp{\LRb{u}_R + c_R + \LRb{u}_L + c_L}, 
+\qquad c = \sqrt{\frac{\gamma p}{\rho}}.
+\]
+
+\begin{figure}
+\centering
+\subfloat[Entropy conservative flux]{
+\begin{tikzpicture}
+\begin{loglogaxis}[
+    width=.475\textwidth,
+    xlabel={Mesh size $h$},
+    ylabel={$L^2$ errors}, 
+    xmin=.0125, xmax=.75,
+    ymin=1e-11, ymax=1.5,
+    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
+    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
+    legend entries={GLL,GQ-$(N+2)$}    
+]
+\pgfplotsset{
+cycle list={
+{blue, mark=*}, {red, mark=square*}
+}
+}
+%\addlegendimage{no markers,blue}
+%\addlegendimage{no markers,red}
+
+\addplot+[semithick, mark options={fill=markercolor}]
+% N = 1, tau = 0.000000 =======================
+coordinates{(0.5,1)(0.25,0.485059)(0.125,0.203599)(0.0625,0.0947163)(0.03125,0.0463705)};
+\addplot+[semithick, mark options={fill=markercolor}]
+%N = 1, tau = 0.000000 =======================
+coordinates{(0.5,0.402314)(0.25,0.167917)(0.125,0.106574)(0.0625,0.058359)(0.03125,0.0298728)}
+[yshift=2pt] node[left, pos=1.025, color=black] {$N = 1$};
+
+\addplot+[semithick, mark options={fill=markercolor}]
+% N = 2, tau = 0.000000 =======================
+coordinates{(0.5,0.746606)(0.25,0.156701)(0.125,0.0137392)(0.0625,0.000701926)(0.03125,8.64531e-05)};
+\addplot+[semithick, mark options={fill=markercolor}]
+%N = 2, tau = 0.000000 =======================
+coordinates{(0.5,0.993771)(0.25,0.0219437)(0.125,0.00180028)(0.0625,0.000194939)(0.03125,2.4045e-05)}
+[yshift=7pt] node[left, pos=1.025, color=black] {$N = 2$};
+
+\addplot+[semithick, mark options={fill=markercolor}]
+% N = 3, tau = 0.000000 =======================
+coordinates{(0.5,0.103299)(0.25,0.00829887)(0.125,0.00073573)(0.0625,9.05975e-05)(0.03125,1.13596e-05)};
+\addplot+[semithick, mark options={fill=markercolor}]
+%N = 3, tau = 0.000000 =======================
+coordinates{(0.5,0.0154054)(0.25,0.00167426)(0.125,0.000260859)(0.0625,3.76182e-05)(0.03125,3.86238e-06)}
+[yshift=3pt] node[left, pos=1.025, color=black] {$N = 3$};
+
+\addplot+[semithick, mark options={fill=markercolor}]
+% N = 4, tau = 0.000000 =======================
+coordinates{(0.5,0.0385542)(0.25,0.00133048)(0.125,0.000176663)(0.0625,1.64135e-06)(0.03125,2.66024e-08)};
+\addplot+[semithick, mark options={fill=markercolor}]
+%N = 4, tau = 0.000000 =======================
+coordinates{(0.5,0.0367592)(0.25,0.000202817)(0.125,3.57758e-06)(0.0625,9.58294e-08)(0.03125,2.94985e-09)}[yshift=8pt] node[left, pos=1.025, color=black] {$N = 4$};
+
+\addplot+[semithick, mark options={fill=markercolor}]
+% N = 5, tau = 0.000000 =======================
+coordinates{(0.5,0.00436131)(0.25,0.00039846)(0.125,2.1282e-06)(0.0625,4.49046e-08)(0.03125,9.99912e-10)};
+\addplot+[semithick, mark options={fill=markercolor}]
+%N = 5, tau = 0.000000 =======================
+coordinates{(0.5,0.000390565)(0.25,1.31188e-05)(0.125,3.6544e-07)(0.0625,4.95271e-09)(0.03125,2.42763e-10)}
+[yshift=5pt] node[left, pos=1.025, color=black] {$N = 5$};
+
+%\legend{$N=1$,$N=2$,$N=3$,$N=4$,$N=5$}
+\end{loglogaxis}
+\end{tikzpicture}
+}
+\subfloat[With Lax-Friedrichs penalization]{
+\begin{tikzpicture}
+\begin{loglogaxis}[
+    width=.475\textwidth,
+    xlabel={Mesh size $h$},  ylabel={$L^2$ errors}, 
+    xmin=.0125, xmax=.75,
+    ymin=1e-11, ymax=1.5,
+    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
+    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
+    legend entries={GLL,GQ-$(N+2)$}
+] 
+\pgfplotsset{
+cycle list={
+{blue, mark=*}, {red, mark=square*}
+}
+}
+%\addlegendimage{no markers,blue}
+%\addlegendimage{no markers,red}
+
+\addplot+[semithick, mark options={fill=markercolor}]
+% N = 1, tau = 0.500000 =======================
+coordinates{(0.5,1)(0.25,0.4932)(0.125,0.183839)(0.0625,0.0562398)(0.03125,0.0151873)};
+\addplot+[semithick, mark options={fill=markercolor}]
+%N = 1, tau = 0.500000 =======================
+coordinates{(0.5,0.547558)(0.25,0.148981)(0.125,0.0384647)(0.0625,0.00974763)(0.03125,0.00244539)}
+[yshift=5pt] node[left, pos=1.025, color=black] {$N = 1$};
+
+\addplot+[semithick, mark options={fill=markercolor}]	
+% N = 2, tau = 0.500000 =======================
+coordinates{(0.5,0.336817)(0.25,0.04941)(0.125,0.00605428)(0.0625,0.000748842)(0.03125,9.25456e-05)};
+\addplot+[semithick, mark options={fill=markercolor}]
+%N = 2, tau = 0.500000 =======================
+coordinates{(0.5,0.165807)(0.25,0.0190013)(0.125,0.00227903)(0.0625,0.00028425)(0.03125,3.54865e-05)}
+[yshift=5pt] node[left, pos=1.025, color=black] {$N = 2$};
+
+% N = 3, tau = 0.500000 =======================
+\addplot+[semithick, mark options={fill=markercolor}]
+coordinates{(0.5,0.0463242)(0.25,0.00408748)(0.125,0.000346831)(0.0625,1.99064e-05)(0.03125,1.22357e-06)};
+\addplot+[semithick, mark options={fill=markercolor}]
+%N = 3, tau = 0.500000 =======================
+coordinates{(0.5,0.0174194)(0.25,0.00182234)(0.125,0.000116147)(0.0625,7.39839e-06)(0.03125,4.6305e-07)}
+[yshift=5pt] node[left, pos=1.025, color=black] {$N = 3$};
+
+\addplot+[semithick, mark options={fill=markercolor}]
+% N = 4, tau = 0.500000 =======================
+coordinates{(0.5,0.0100716)(0.25,0.000625923)(0.125,1.89866e-05)(0.0625,7.03865e-07)(0.03125,2.10265e-08)};
+\addplot+[semithick, mark options={fill=markercolor}]
+%N = 4, tau = 0.500000 =======================
+coordinates{(0.5,0.00556743)(0.25,0.000144595)(0.125,4.33972e-06)(0.0625,1.37151e-07)(0.03125,4.16335e-09)}
+[yshift=5pt] node[left, pos=1.025, color=black] {$N = 4$};
+
+\addplot+[semithick, mark options={fill=markercolor}]
+% N = 5, tau = 0.500000 =======================
+coordinates{(0.5,0.00356628)(0.25,8.73125e-05)(0.125,2.20528e-06)(0.0625,3.20127e-08)(0.03125,4.63639e-10)};
+\addplot+[semithick, mark options={fill=markercolor}]
+%N = 5, tau = 0.500000 =======================
+coordinates{(0.5,0.000547621)(0.25,8.7194e-06)(0.125,1.47105e-07)(0.0625,2.34345e-09)(0.03125,3.65306e-11)}
+[yshift=10pt] node[left, pos=1.025, color=black] {$N = 5$};
+
+\end{loglogaxis}
+\end{tikzpicture}
+}
+\caption{$L^2$ errors under mesh refinement for entropy conservative and Lax-Friedrichs fluxes under both Gauss-Legendre-Lobatto (GLL) and over-integrated $(N+2)$ point Gauss quadrature (GQ-$(N+2)$).}
+\label{fig:convergence}
+\end{figure}
+We begin by verifying the high order accuracy of the proposed methods using a periodic entropy wave solution
+\[
+\rho(x,t) = 2 + \sin\LRp{\pi (x - t)}, \qquad u(x,t) = 1, \qquad p(x,t) = 1.
+\]
+We compute the $L^2$ error in the conservation variables 
+\[
+\nor{\bm{u} - \bm{u}_h}_{L^2}^2 = \nor{\rho - \rho_h}_{L^2}^2 + \nor{m - m_h}_{L^2}^2 + \nor{E - E_h}_{L^2}^2
+\]
+at final time $T = .7$ using both the entropy conservative and local Lax-Friedrichs fluxes.
+For these experiments, we compare two quadrature rules: the $(N+1)$ point Gauss-Legendre-Lobatto rule (which induces a DG-SEM discretization and is referred to in Figure~\ref{fig:convergence} as ``GLL'') and an $(N+2)$ point Gauss quadrature rule (referred to in the figure as ``GQ-$(N+2)$'').
+We do not compare against the $(N+1)$ point Gauss quadrature rule, which behaves very similarly to the GQ-$(N+2)$ case.
+The $L^2$ error is evaluated using a more accurate $N+5$ point Gauss quadrature rule, and a CFL of $.125$ is used.
+
+\begin{table}[h!]
+\centering
+ \begin{tabular}{||c | c |c |c |c | c|} 
+ \hline
+ &  $N=1$ &  $N=2$ &  $N=3$ &  $N=4$ &  $N=5$\\
+ \hline
+ GLL (entropy conservative) & 1.0672 & 3.6561  &  3.0086 &   6.3486 &   5.5278\\ 
+ \hline
+ GQ-$(N+2)$ (entropy conservative) &  0.9175  &  3.1132  &  3.0388 &   5.1221 &   5.2779 \\ 
+ \hline
+ GLL (Lax-Friedrichs) & 1.8887  &  3.0164  &  4.0241 &   5.0650 &   6.1095\\ 
+ \hline
+ GQ-$(N+2)$ (Lax-Friedrichs) & 1.9950  &  3.0018  &  3.9980  &  5.0419  &  6.0034\\ 
+ \hline
+ \end{tabular}
+ \caption{Computed asymptotic convergence rates of $L^2$ errors under mesh refinement.}
+ \label{tab:rates}
+\end{table}
+
+Figure~\ref{fig:convergence} shows computed $L^2$ errors under mesh refinement for both GLL and GQ-$(N+2)$ quadrature rules.
+All experiments utilize the entropy conservative flux of Chandreshekar, and we compare errors with and without Lax-Friedrichs penalization.
+Computed asymptotic convergence rates are reported in Table~\ref{tab:rates}.
+It can be observed that, for both GLL and GQ-$(N+2)$ quadratures, the entropy conservative flux exhibits suboptimal convergence rates for odd orders, while the inclusion of Lax-Friedrichs penalization restores exhibits the optimal $O(h^{N+1})$ convergence rate for both quadrature choices.
+
+Next, we examine the discrete evolution of entropy by evolving a discontinuous initial profile to final time $T=4$.  We initialize the density and velocity to be 
+\begin{equation}
+\rho(x,t) = \begin{cases}
+3 & \LRb{x} < 1/2\\
+2 & \text{otherwise},
+\end{cases} \qquad 
+u(x,t) = 0, \qquad
+p(x,t) = \rho^\gamma.
+\label{eq:discontin}
+\end{equation}
+Non-physical periodic boundary conditions are enforced in order to examine the evolution of entropy over longer time periods.
+Figure~\ref{subfig:sol} shows $\rho, u$ at time $T = 1/10$ using both entropy conservative and Lax-Friedrichs fluxes (referred to as ``EC'' and ``LF'', respectively in the figure) and GQ-$(N+2)$ quadrature.
+As expected, the entropy conservative flux results in spurious high frequency oscillations, which are significantly damped under Lax-Friedrichs penalization.
+
+We note that, while the proof of conservation of entropy holds at the semi-discrete level, it does not take into account the time discretization and thus does not hold at the fully discrete level.
+This is reflected into numerical observation that the change in entropy $\Delta U(t) = U\LRp{\bm{u}(t}) - U\LRp{\bm{u}(t)}$ increases as $t$ increases.
+However, numerical experiments in \cite{gassner2016well} suggest that the discrete change in entropy over time should converge to zero as the timestep decreases, which (under appropriate assumptions) implies that the fully discrete scheme converges to the semi-discrete scheme.
+
+Figure~\ref{subfig:dS} shows the evolution of $\Delta U(t)$ to final time $T=2$ for both entropy conservative and Lax-Friedrichs fluxes at various CFL numbers.
+For the entropy conservative flux, we observe that $\Delta U(t)$ to decreases as the CFL and $\Delta t$ decrease.
+This is expected, since the discrete time problem should converge to the continuous semi-discrete problem (for which $\Delta U(t) = 0$) as $\Delta t\rightarrow 0$.
+We also observe that, as expected, $\Delta U(t)$ does not change significantly with CFL for the Lax-Friedrichs flux, indicating that the change in entropy is due to the effect of the dissipative flux rather than the time discretization.
+
+\begin{figure}
+\centering
+\subfloat[Solution at time $T = .1$]{
+\includegraphics[width=.475\textwidth]{figs/sol_ECLF.png}
+\label{subfig:sol}}
+\hspace{.5em}
+\subfloat[$\Delta U(t)$]{
+\includegraphics[width=.475\textwidth]{figs/dS_ECLF.png}
+\label{subfig:dS}}
+\caption{Solution snapshot and change in entropy $\Delta U(t)$ for both entropy conservative (EC) and Lax-Friedrichs (LF) fluxes.}
+\label{fig:dS}
+\end{figure}
+
+We also numerically evaluate the spatial formulation tested against the projected the entropy variables
+\[
+\delta(t) = \LRb{\LRp{\left.\LRp{D^x_h \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},\Pi_N \bm{v}}_{\Omega}}, \qquad  \qquad 0\leq t \leq T.
+\]
+We utilize the non-dissipative entropy conservative flux and evolve the initial profile (\ref{eq:discontin}) until time $T = 1$ using a GQ-$(N+2)$ quadrature rule.
+From Theorem~\ref{thm:ecdg_global}, we expect $\delta_{\max} = \max_{t\in (0,T)}{\delta}$ to be machine zero.
+In practice, we have found that $\delta_{\max}$ depends on the tolerance $\epsilon$ used in evaluation of the logarithmic mean.
+For a simulation to time $T=4$ using $N=4$ and $K = 16$ with a CFL of $1/2$, we observe that using $\epsilon = 10^{-2}$ (as recommended in \cite{ismail2009affordable} results in $\delta_{\max} = O\LRp{10^{-10}}$.
+Decreasing $\epsilon$ to $10^{-3}$ reduces $\delta_{\max}$ to $O\LRp{10^{-14}}$, and decreasing $\epsilon$ further to $1\times 10^{-4}$ reduces $\delta_{\max}$ to $10^{-15}$.
+Smaller values of $\epsilon$ do result in observable changes to $\delta_{\max}$, and we do not observe any significant dependence of $\delta_{\max}$ on other discretization parameters.
+
+\begin{figure}
+\centering
+\includegraphics[width=.5\textwidth]{figs/rhstest.png}
+\caption{Comparison of $\delta(t)$ when evaluating the flux function $\bm{f}_S$ using conservative variables and projected entropy variables.}
+\label{fig:rhstest}
+\end{figure}
+Finally, in order to illustrate the importance of evaluating the flux in terms of the projected entropy variables $\Pi_N \bm{v}$, we compute $\delta(t)$ while evaluating the flux function $\bm{f}_S$ directly in terms of the conservative variables $\bm{u}(x)$ and in terms of $\bm{u}\LRp{(\Pi_N \bm{v})(x)}$.
+It can be observed from Figure~\ref{fig:rhstest} that $\delta(t)$ is near machine precision when evaluating the flux in terms of the projected entropy variables.
+When evaluating the flux function using the conservative variables directly, $\delta(t)$ begins near $10^{-4}$ but grows steadily, blowing up exponentially near $t = 1$.
 
@@
 
-\note{Analysis and experiments are presented in one dimension for clarity; however, they are directly extendable to higher dimensions, including on simplicial elements.
-This will be the focus of future work.
-}
+This work presents a generalization of discretely entropy conservative methods from diagonal-norm SBP-DG methods to a more general class of high order DG methods which include dense norm SBP methods and over-integrated quadrature rules.
+The resulting DG methods satisfy a discrete conservation of entropy while maintaining a reasonable computational cost and small stencil.
+Numerical experiments indicate the resulting methods deliver high order accuracy for smooth solutions while displaying greatly improved stability for under-resolved and shock solutions.
+
+While entropy conservative and stability improve the robustness of solvers for nonlinear hyperbolic conservation laws, they do not address problems such as spurious oscillations in high order approximations of shock solutions or positivity preservation of density and pressure variables \cite{chen2017entropy}.
+These issues can be addressed through the use of regularization (e.g.\ filtering, artificial viscosity) and/or limiting; however, the goal in constructing discretely entropy stable methods is to avoid using excessive regularization and limiting in the pursuit of stability.
+
+Finally, we note that the analysis and numerical experiments in this work are presented in one dimension for clarity and conciseness of presentation.
+The main concepts are directly extendable to higher dimensions and simplicial and pyramidal elements, which will be the focus of future work.
+
+%\note{
+%Open question: it's consistent with existing entropy stable collocation DG discretizations, but is the use of $\bm{u}\LRp{\Pi_N \bm{v}}$; is this consistent, can we show this converges?  
+%}
 
@@
 \appendix
-\section{Expansion of the flux function}
+\section{Assumptions on the flux function}
 \label{appendix:A}
 
-\note{For Burgers and shallow water, this sum is finite.
-For the Chandreshekar flux, we consider individual components: the logarithmic mean $L(x,y)$ admits a convergent expansion and is analytic in $x$ \cite{mustonen2002logarithmic}.
+In this work, we assume that the flux function $\bm{f}_S$ admits an absolutely convergent expansion of the form
+\[
+\bm{f}_S(x,y) = \sum_{i=1}^{\infty} \bm{f}_i(x) \bm{g}_i(y), 
+\]
+where $\bm{f}_i, \bm{g}_i$ are real and bounded over each element $D^k$.
+This holds, for example, for analytic $\bm{f}_S$, for which the Taylor series can be rearranged to yield the form of the desired expansion.
+This assumption is trivially satisfied for the Burgers' and shallow water equations, as the flux functions consist only of finite sums of products of variables.
+However, for the compressible Euler equations, this expansion is less clear due to the presence of the logarithmic mean in the entropy conservative Chandreshekar flux and the presence of inverses of averages and logarithmic means of the variable $\beta = \rho/(2p)$.
+
+%\note{For Burgers and shallow water, this sum is finite.  
+%For the Chandreshekar flux, we consider individual components: the logarithmic mean $L(x,y)$ admits a convergent expansion and is analytic in $x$ \cite{mustonen2002logarithmic}.
 CH's flux introduces rational denominators involving the mean and log-mean of $\beta = \rho/(2p)$, which is bounded away from zero if the density and temperature are bounded away from zero.
@@
 
+We first address the logarithmic mean by noting that it can be expanded out into series form using multi-index notation \cite{mustonen2002logarithmic}.
+Let $\mu= (\mu_1,\mu_2)$ be positive integers and define $\LRb{\mu} = \mu_1 + \mu_2$; then, the logarithmic mean can be rewritten as
+\[
+\frac{x - y}{\log(x)-\log(y)} = \sum_{n=0}^{\infty} \frac{x_n}{n!}, \qquad x_n = \sum_{\LRb{\mu} = n} \frac{\log(x)^{\mu_1}\log(y)^{\mu_2}}{n+1},
+\]
+where division by $(n+1)$ averages over the $(n+1)$ different permutations of $\mu_1,\mu_2 \geq 0$ which sum to $n$.  
+Since $x_n$ is the average of degree $n$ powers of $\log(x),\log(y)$, we have that
+\[
+%\LRb{x_n} \leq \max{\LRb{\log(x)^n,\log(y)^n}}, \qquad 
+\LRb{\frac{x_n}{n!}} \leq \frac{\max{\LRb{\log(x),\log(y)}}^n}{n!}. 
+\]
+The latter term is simply the $n$th term of the exponential Taylor series evaluated with the argument $\max{\LRb{\log(x),\log(y)}}$.
+Because the exponential Taylor series converges absolutely everywhere, the series for the logarithmic mean also converges absolutely for $x,y$ bounded away from $0$ and $\infty$ (i.e.\ $x,y$ are in the logarithmic interval of convergence).
+
+The entropy conservative Chandreshekar flux also requires multiplication by the terms
+\[
+\frac{1}{\avg{\beta}}, \qquad \frac{1}{\avg{\beta}^{\log}}, \qquad \beta = \frac{\rho}{2p}.
+\]
+These terms are analytic if $\avg{\beta}, \avg{\beta}^{\log}$ are bounded away from zero, which is satisfied if the density and pressure are bounded from below as assumed in (\ref{eq:assumption2}).
+Because the composition, sum, and product of analytic functions are also analytic, the Chandreshekar flux also is analytic under positivity assumptions.
+
+Finally, we note that the application of the operator $D_N$ to convergent infinite sums should be rigorously justified.  
+This requires $D_N$ to be a bounded operator from $H^1$ to $P^N$, which can be shown using boundedness of the $L^2$ projection operator and discrete inverse and trace inequalities \cite{chan2015gpu}.
+
 \bibliographystyle{unsrt}

### 2017-07-30 docs/dg_penalty/revision/main.tex

@@
 
-\newcommand{\note}[1]{{\color{blue}#1}}
+%\newcommand{\note}[1]{{\color{blue}#1}}
+\newcommand{\note}[1]{#1}
 \newcommand{\remark}[1]{\textbf{\color{red}#1}}
@@
 %\title{A short note on the penalty flux parameter for first order discontinuous Galerkin formulations}
+\graphicspath{{./figs/}}
 
@@
 \]
-The upwind flux can similarly be extended for discontinuous coefficients
-\[
-(\bm{A}_n\bm{U})^* = \avg{\bm{A}_n\bm{U}} - \frac{1}{2} \jump{\bm{V}\LRb{\Lambda}\bm{V}^{-1}\bm{U}},
-\]
-where the eigenvalue decompositions inside the jump is evaluated as 
-\begin{align*}
-&\jump{\bm{V}\LRb{\Lambda}\bm{V}^{-1}\bm{U}} = {\bm{V}^+\LRb{\Lambda}^+\LRp{\bm{V}^+}^{-1}\bm{U}^+} - {\bm{V}^-\LRb{\Lambda}^-\LRp{\bm{V}^-}^{-1}\bm{U}^-}\\
-&\bm{A}_n^+ = \sum_{i=1}^d \bm{n}_i \bm{A}_i^+ = \bm{V}^+{\Lambda}^+\LRp{\bm{V}^+}^{-1}, \qquad \bm{A}_n^- = \sum_{i=1}^d \bm{n}_i \bm{A}_i^- = \bm{V}^-{\Lambda}^-\LRp{\bm{V}^-}^{-1}.
-\end{align*} 
+%The upwind flux can similarly be extended for discontinuous coefficients
+%\[
+%(\bm{A}_n\bm{U})^* = \avg{\bm{A}_n\bm{U}} - \frac{1}{2} \jump{\bm{V}\LRb{\Lambda}\bm{V}^{-1}\bm{U}},
+%\]
+%where the eigenvalue decompositions inside the jump is evaluated as 
+%\begin{align*}
+%&\jump{\bm{V}\LRb{\Lambda}\bm{V}^{-1}\bm{U}} = {\bm{V}^+\LRb{\Lambda}^+\LRp{\bm{V}^+}^{-1}\bm{U}^+} - {\bm{V}^-\LRb{\Lambda}^-\LRp{\bm{V}^-}^{-1}\bm{U}^-}\\
+%&\bm{A}_n^+ = \sum_{i=1}^d \bm{n}_i \bm{A}_i^+ = \bm{V}^+{\Lambda}^+\LRp{\bm{V}^+}^{-1}, \qquad \bm{A}_n^- = \sum_{i=1}^d \bm{n}_i \bm{A}_i^- = \bm{V}^-{\Lambda}^-\LRp{\bm{V}^-}^{-1}.
+%\end{align*} 
 Finally, we note that the energy estimate changes when $\bm{A}_i$ varies over an element, though the penalty flux still adds a dissipative contribution to the right hand side \cite{kopriva2014energy}.

### 2017-07-30 docs/splitform/main.tex

@@
 
-To the authors knowledge, the construction of entropy conservative and entropy stable schemes have relied on the use of lumped diagonal mass matrices and nodal collocation (diagonal norm SBP operators).
-Entropy stable high order finite element and DG methods have been proposed \cite{hughes1986new}, but the proofs are often given at the continuous level, relying on exact integration or the chain rule, neither of which holds at the discrete level.
-Thus, while DG-SEM based on flux differencing is entropy conservative, this is tied to a specific choice of basis and quadrature.
-Diagonal-norm SBP operators can be constructed for triangles and tetrahedra \cite{chin1999higher,  hicken2016multidimensional, chen2017entropy}; however, the number of nodal points for such operators is greater than the dimension of the natural polynomial approximation space.
-Furthermore, appropriate point sets have only been constructed for $N \leq 4$ in three dimensions \cite{zhebel2014comparison}.
+To the authors knowledge, the construction of entropy conservative and entropy stable schemes have relied on the use of lumped diagonal mass matrices and nodal collocation.
+Since these can be reinterpreted as diagonal norm SBP operators, we refer to DG methods with these properties as diagonal norm SBP-DG methods.
+Entropy stable high order finite element and DG methods which do not fall under the diagonal norm SBP-DG category have been proposed \cite{hughes1986new}, but the proofs are often given at the continuous level, relying on exact integration or the chain rule, neither of which holds at the discrete level.
+
+On the other hand, while entropy conservative schemes can be constructed by combining flux differencing with diagonal norm SBP operators, this can restrict one to specific choices of basis and quadrature.
+Appropriate diagonal norm SBP operators can be constructed on tensor product elements based on a DG-SEM discretization.
+Diagonal-norm SBP operators can also be constructed for triangles and tetrahedra \cite{chin1999higher,  hicken2016multidimensional, chen2017entropy}; however, the number of nodal points for such operators is greater than the dimension of the natural polynomial approximation space.
+Furthermore, appropriate point sets have only been constructed for $N \leq 4$ in three dimensions \cite{zhebel2014comparison}, and (to the author's knowledge) the construction of diagonal norm SBP-DG methods has not yet been performed for less standard elements such as pyramids \cite{chan2016orthogonal}.
 %Furthermore, diagonal-norm SBP operators do not naturally generalize to triangles or tetrahedra \cite{helenbrook2009existence}.
@@
 This work focuses on the construction of entropy conservative high order DG schemes for systems of conservation laws.
-In order to generalize beyond collocation DG and diagonal norm SBP operators, we will consider dense (non-lumped) mass matrices and more general quadrature rules (e.g.\ rules without boundary points, rules with more points than the dimension of the corresponding approximation space).
+In order to generalize beyond diagonal norm SBP-DG methods, we will consider dense (non-lumped) mass matrices and more general quadrature rules (e.g.\ rules without boundary points, rules with more points than the dimension of the corresponding approximation space).
 We will present proofs of discrete entropy stability in terms of integrals and $L^2$ projections, which we assume are computed using an appropriate quadrature rule.
@@
 
-The outline of the paper is as follows: Section~\ref{sec:intro} will briefly review the construction of entropy conservative collocation methods on a single element.
+The outline of the paper is as follows: Section~\ref{sec:intro} will briefly review the construction of entropy conservative diagonal norm SBP-DG methods on a single element.
 Section~\ref{sec:ecdg} will describe how to construct analogous entropy conservative methods on single element.
@@
 
-\section{Entropy conservative diagonal norm collocation DG methods}
+\section{Entropy conservative diagonal norm SBP-DG methods}
 \label{sec:intro}
 
-We will begin by reviewing continuous entropy theory, as well as existing high order methods which are provably entropy conservative at the discrete level.
+We will begin by reviewing continuous entropy theory and existing high order methods which are provably entropy conservative at the discrete level.
 These methods were introduced in \cite{fisher2013high, carpenter2014entropy} for tensor product elements, though we adopt methods of proof introduced more recently in \cite{gassner2017br1, chen2017entropy}.
-These methods all share a basis in diagonal norm summation by parts operators.
-Because they can be derived from a nodal discontinuous Galerkin framework through the collocation of nodal points and quadrature points, we refer to them as ``diagonal norm collocation DG methods'' in this work.
+These methods all share a basis in diagonal norm summation by parts operators, and can be derived from a nodal discontinuous Galerkin framework through the collocation of nodal points and quadrature points.
 
@@
 The focus of this work is the construction of methods which satisfy a discrete analogue of this entropy inequality.
-We begin by reviewing existing discretely entropy conservative methods based on collocation and diagonal norm summation by parts (SBP) operators.
-
-\subsection{Entropy conservative collocation methods based on SBP operators}
+We begin by reviewing existing discretely entropy conservative methods based on diagonal norm summation by parts (SBP) operators.
+
+\subsection{Entropy conservative methods based on diagonal-norm SBP operators}
 
@@
 \end{definition}
-An entropy conservative nodal collocation method can then be derived by modifying the flux derivative as follows
+An entropy conservative diagonal norm SBP-DG method can then be derived by modifying the flux derivative as follows
 \begin{equation}
@@
 \begin{theorem}
-The collocation method (\ref{eq:dgsem}) is discretely entropy conservative in the sense that
+The nodal collocation method (\ref{eq:dgsem}) is discretely entropy conservative in the sense that
 \[
@@
 
-\note{Fix notation: switch $V_h$ to $P^N$}
-We assume that the domain $\Omega \in \mathbb{R}^d$ is exactly triangulated by a mesh which consists of the union of non-overlapping elements $D^k$.
-We further assume that each element $D^k$ is the image of a reference element $\widehat{D}$ under the local elemental mapping
-\[
-x^k = \bm{\Phi}^k \widehat{x},
-\]
-where $x^k$ denote physical coordinates on $D^k$ and $\widehat{x}$ denote coordinates on the reference element.
-Finally, we identify $J$ as the determinant of the Jacobian of $\bm{\Phi}^k$, and refer to it as the Jacobian for the remainder of this work.
-For clarity and simplicity of presentation, we assume $\bm{\Phi}^k$ is affine in this case, such that $J$ is constant over each element $D^k$.
-
-We will approximate solution components over each element $D^k$ from an approximation space $V_h\LRp{D^k}$, which we define as the composition of the mapping $\bm{\Phi}^k$ and a reference approximation space $V_h\LRp{\widehat{D}}$
-\[
-V_h\LRp{D^k} = \bm{\Phi}^k \circ V_h\LRp{\widehat{D}}.
-\]
-The global approximation space $V_h\LRp{\Omega}$ is then defined as the direct sum of local approximation spaces
-\[
-V_h\LRp{\Omega} = \bigoplus_{D^k}V_h\LRp{D^k}.  
-\]
-We assume that $V_h(D^k) \subset H^1(D^k)$; in other words, that all elements of each local approximation space are square integrable and admit well-defined traces.
-
-We define the $L^2$ norm and inner products over the element $D^k$ and the surface of the element $\partial D^k$\footnote{In one dimension, $D^k$ is some interval $[x_L,x_R]$ and the $L^2$ inner product over the surface reduces to
+%\note{Fix}
+
+We assume that the domain $\Omega$ is an interval, which is decomposed into non-overlapping intervals (elements) $D^k$ of size $h$.
+We define the reference element $\widehat{D} = [-1,1]$, such that $D^k$ is the image of $\widehat{D}$ under an affine mapping $x = \Phi^k(\widehat{x})$.
+We approximate the solution over the reference element using degree $N$ polynomials
+\[
+P^N\LRp{\widehat{D}} = \LRc{\widehat{x}^i, \quad \widehat{x} \in \widehat{D}, \quad 0\leq i \leq N}.
+\]
+%We define an approximation space over the reference element $\widehat{D}$ as the space of degree $N$ polynomials 
+%\[
+%P^N\LRp{\widehat{D}} = \LRc{\widehat{x}^i, \qquad 0\leq i \leq N}.
+%\]
+%We will approximate solution components over each element $D^k$ using 
+% $V_h\LRp{D^k}$, which we define as the composition of the mapping $\bm{\Phi}^k$ and $P^N\LRp{\widehat{D}}$
+%\[
+%V_h\LRp{D^k} = \bm{\Phi}^k \circ V_h\LRp{\widehat{D}}.
+%\]
+We approximate the solution over a physical element by mapping $P^N\LRp{\widehat{D}}$ to $D^k$ under $\Phi^k$, and define the global approximation space $V_h$ as the direct sum of local approximation spaces
+\[
+P^N\LRp{D^k} =  \bm{\Phi}^k \circ P^N\LRp{\widehat{D}}, \qquad V_h = \bigoplus_{k}P^N\LRp{D^k}.
+\]
+We define the $L^2$ norm and inner products over the element $D^k$ and the surface of the element $\partial D^k$
+\[
+\LRp{u,v}_{D^k} = \int_{D^k} u v\diff{x} =  \int_{D^k} u v J^k \diff{\widehat{x}}, \qquad \nor{u}^2_{D^k} = (u,u)_{D^k}, \qquad \LRa{u,v}_{\partial D^k} = \int_{\partial D^k} u v \diff{x},
+\]
+where $J^k$ is the determinant of the Jacobian of $\Phi^k$.  
+We also define the local $L^2$ projection operator $\Pi_N: L^2(D^k)\rightarrow P^N(D^k)$ such that
+\[
+\LRp{\Pi_N u,v}_{D^k} = \LRp{u,v}_{D^k}, \qquad \forall v\in V_h\LRp{D^k}.  
+\]
+In one dimension, $D^k$ is some interval and the $L^2$ inner product over the surface reduces to 
 \[
@@
 \]  
-However, the presented proofs in this work will not utilize this simplification, and are thus generalizable to higher dimensions in a straightforward manner.}
-\[
-\LRp{u,v}_{D^k} = \int_{D^k} u v\diff{x}, \qquad \nor{u}^2_{D^k} = (u,u)_{D^k}, \qquad \LRa{u,v}_{\partial D^k} = \int_{\partial D^k} u v \diff{x}.
-\]
-We also define the local $L^2$ projection operator $\Pi_N: L^2(D^k)\rightarrow V_h(D^k)$ such that
-\[
-\LRp{\Pi_N u,v}_{D^k} = \LRp{u,v}_{D^k}, \qquad \forall v\in V_h\LRp{D^k}.  
-\]
-Throughout this work, we will use that this definition of the $L^2$ projection holds even when integrals are approximated using quadrature.
-We will make no assumptions on quadrature except that \note{specify assumptions - integrate polynomials of degree $N$ on the reference element exactly, such that integration by parts holds.
+However, the presented proofs in this work will not directly utilize this simplification in order to remain generalizable to higher dimensions in a straightforward manner.
+
+Throughout this work, we will assume that integrals and inner products are approximated using some quadrature rule consisting of $N_q$ points $\widehat{x}_i$ and weights $w_i$ over the reference element $\widehat{D}$.
+\[
+\LRp{u,v}_{D^k} \coloneqq \sum_{i=1}^{N_q} u(\widehat{x}_i)v(\widehat{x}_i)w_i J^k.     
+\]
+This also defines a quadrature-based $L^2$ projection operator.
+We emphasize that, throughout this work, we will only use properties of inner products and projection operators which hold when integrals are approximated using quadrature.
+We assume only that the quadrature is sufficiently accurate to integrate degree $(2N-1)$ polynomials, such that integration by parts holds on the reference element.
+However, we will also allow for quadrature with an arbitrary number of points and arbitrary accuracy.
+
+%\note{specify assumptions - integrate polynomials of degree $N$ on the reference element exactly, such that integration by parts holds.
 Add footnote that we can also generalize to the under-integrated case.
@@
 The jump and average of vector fields are defined component-wise using the jumps and averages of components.
-When the face $f$ lies on the boundary, we define $u^+ = u$, such that $\jump{u} = 0$ and $\avg{u} = u$ on $\partial \Omega$.
+When the face $f$ lies on the boundary, we define $u^+ = u$ such that $\jump{u} = 0$ and $\avg{u} = u$ on $\partial \Omega$.
 
@@
 Thus, recovery of the split form can be achieved by using flux differencing in combination with any differential operator $D$ such that $D1 = 0$.
-%\[
-%D:L^2 \rightarrow V_h, \qquad D1 = 0.
-%\]
-%For Burgers' equation (as well as any skew-symmetric splitting), replacing the exact derivative $\pd{}{x}$ by $D$ will again recover a statement of the split formulation (\ref{eq:burgerssplit}) 
-%\[
-%\pd{u}{t} + \frac{1}{3}\LRp{Du^2 + uDu} = 0.
-%\]
-%The introduction of $D$ is necessary when adapting the flux differentiation procedure in the context of numerical discretizations.
 
@@
 Both the conservative and split forms of Burgers' equation are equivalent at the continuous level (assuming continuity of $u$ and exact differentiation); however, this is no longer true at the discrete level.
-Suppose now that the solution is approximated by $u \in V_h(\Omega)$ on a single element.
-Since $f(u) = u^2/2 \not\in V_h(\Omega)$ for general choices of $u, V_h$, it is not possible to differentiate $f(u)$ exactly without prior knowledge of the form of the nonlinearity.
-Indeed, it is more common to first project $f(u)$ to $V_h$, then differentiate, such that the discrete scheme is equivalent to solving for $u \in V_h(\Omega)$ such that
+Suppose now that the solution is approximated by $u \in P^N\LRp{\widehat{D}}$ on the reference element.
+Since $f(u) = u^2/2 \not\in P^N\LRp{\widehat{D}}$ for general choices of $u$, it is not possible to differentiate $f(u)$ exactly without prior knowledge of the form of the nonlinearity.
+Indeed, it is more common to first project $f(u)$ to $P^N\LRp{\widehat{D}}$, then differentiate, such that the discrete scheme is equivalent to solving for $u \in P^N\LRp{\widehat{D}}$ such that
 \[
@@
 %\int_{-1}^1 \LRp{\pd{u}{t} +D_N f(u)}w \diff{x} 
-=0, \qquad \forall w \in V_h(\Omega).  
+=0, \qquad \forall w \in P^N\LRp{\widehat{D}}.  
 \]
@@
 We can derive a semi-discrete Galerkin scheme based which conserves energy discretely using flux differencing and a different choice of discrete differentiation operator $D_N$, which was motivated by the SBP operators introduced in \cite{chen2017entropy}.
-Let $w(x)$ be some bounded real-valued function; then, we can define the discrete differential operator $D_N: H^1\rightarrow V_h$ such that, for $u\in H^1$,
+Define $H^1\LRp{\widehat{D}}$ as the Sobolev space
+\[
+\nor{u}_{H^1\LRp{\widehat{D}}}^2 = \nor{u}_{L^2\LRp{\widehat{D}}}^2 + \nor{\pd{u}{x}}_{L^2\LRp{\widehat{D}}}^2, \qquad H^1\LRp{\widehat{D}} = \LRc{ u \in L^2\LRp{\widehat{D}}, \qquad \nor{u}_{H^1\LRp{\widehat{D}}} < \infty}.
+\]
+Because functions $u \in H^1\LRp{\widehat{D}}$ admit a well-defined trace \cite{brenner2007mathematical}, we can define the following discrete differential operator as a map from $H^1$ to $P^N$:
+\begin{definition}
+Let $w(x)$ be a bounded function on $\widehat{D}$.  Define $D_N: H^1\LRp{\widehat{D}}\rightarrow P^N\LRp{\widehat{D}}$ such that, for $u\in H^1\LRp{\widehat{D}}$, 
 \begin{align}
-\LRp{D_N u, vw}_{\widehat{D}} = \LRp{\pd{ \Pi_N u}{x},vw}_{\widehat{D}} + \frac{1}{2}{\LRa{{u - \Pi_N u}, vw\bm{n}_x}}_{\partial \widehat{D}} + \frac{1}{2}\LRa{u - \Pi_Nu ,\Pi_N\LRp{vw}\bm{n}_x }_{\widehat{D}}, \qquad \forall v\in V_h.
+&\LRp{D_N u, vw}_{\widehat{D}} = \LRp{\pd{ \Pi_N u}{x},vw}_{\widehat{D}}\nonumber \\
+&+ \frac{1}{2}{\LRa{{u - \Pi_N u}, vw\bm{n}_x}}_{\partial \widehat{D}} + \frac{1}{2}\LRa{u - \Pi_Nu ,\Pi_N\LRp{vw}\bm{n}_x }_{\widehat{D}}, \qquad \forall v\in P^N\LRp{\widehat{D}}.
 \label{eq:dgd}
 \end{align}
+\end{definition}
 This operator satisfies a form of integration by parts, which is summarized by the following lemma:
@@
 \[
--\LRp{\pd{ \Pi_N (vw)}{x},u}_{\widehat{D}} + \frac{1}{2}{\LRa{{u - \Pi_N u}, vw\bm{n}_x}}_{\partial \widehat{D}} + \frac{1}{2}\LRa{u + \Pi_Nu ,\Pi_N\LRp{vw}\bm{n}_x }_{\partial \widehat{D}}, \qquad \forall v\in V_h.
+-\LRp{\pd{ \Pi_N (vw)}{x},u}_{\widehat{D}} + \frac{1}{2}{\LRa{{u - \Pi_N u}, vw\bm{n}_x}}_{\partial \widehat{D}} + \frac{1}{2}\LRa{u + \Pi_Nu ,\Pi_N\LRp{vw}\bm{n}_x }_{\partial \widehat{D}}, \qquad \forall v\in P^N\LRp{\widehat{D}}.
 \]
@@
 
-Direct computations also show that $\LRp{D_N 1,vw}_{\widehat{D}} = 0$ for all bounded weights $w(x)$ and $\forall v \in V_h$.
+Direct computations also show that $\LRp{D_N 1,vw}_{\widehat{D}} = 0$ for all bounded weights $w(x)$ and $\forall v \in P^N\LRp{\widehat{D}}$.
 Thus, we can apply (\ref{eq:fluxdiff}) with $D_N$ to construct a discrete formulation for Burgers' equation
@@
 %\LRp{\pd{u}{t},v}_{\widehat{D}} + \frac{1}{3}\LRp{\pd{\Pi_N u^2}{x},v}_{\widehat{D}} + \frac{1}{3}\LRp{D_Nu,uv}_{\widehat{D}}
-\qquad \forall v\in V_h.
+\qquad \forall v\in P^N\LRp{\widehat{D}}.
 \label{eq:burgersdiscrete}
@@
 
-%Using the fact integration by parts holds when considering two elements of $V_h$, and that $\LRp{\Pi_N u^2, v} = \LRp{u^2,v}$ for any $v \in V_h$, it can be shown that the volume terms cancel out for $v = u$, leaving only boundary terms (which is a prerequisite of energy/entropy stability).
-
 \subsection{Conservation of entropy on one element}
@@
 \]
-where $\bm{f}_i, \bm{g}_i$ are real and bounded over each element $D^k$.  
+where $\bm{f}_i, \bm{g}_i$ are real and bounded for appropriate values of $x,y$.  
 \label{eq:assumption}
 \end{assumption}
-This assumption is satisfied, for example, by $\bm{f}_S$ which are real analytic and thus admit a Taylor expansion in $x$ and $y$.
-In Appendix~\ref{appendix:A}, we show that this assumption also holds for the choices of flux functions $\bm{f}_S$ considered in this work, assuming the solution satisfies physical bound constraints.
-However, we note that this assumption is not required in the computation implementations of the proposed methods, and may be an overly restrictive condition on $\bm{f}_S$.
-
-We now introduce a discretization for the nonlinear system of conservation laws (\ref{eq:pde}) on $[-1,1]$: find $\bm{u}\in V_h$ such that
+The precise definition of what constitutes ``appropriate values of $x,y$'' depends on the physical problem.
+For example, the flux functions for the compressible Euler equations require positivity constraints, which are outlined in more detail in Section~\ref{sec:num}.
+This assumption is satisfied, for example, by $\bm{f}_S$ which are real analytic and thus admit a Taylor expansion in the two arguments.
+In Appendix~\ref{appendix:A}, we argue that this assumption also holds for the choices of flux functions $\bm{f}_S$ considered in this work for the Burgers' and compressible Euler equations, assuming the solution satisfies physical bound constraints.
+However, we emphasize that this assumption is not required in the computation implementations of the proposed methods, and that it may be possible to relax these conditions on $\bm{f}_S$.
+
+We now introduce a discretization for the nonlinear system of conservation laws (\ref{eq:pde}) on the reference element $\widehat{D}$: find $\bm{u}\in P^N\LRp{\widehat{D}}$ such that
 \begin{equation}
-\LRp{\pd{\bm{u}}{t},\bm{w}}_{\widehat{D}} + \LRp{\left.\LRp{2D_N \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},\bm{w}}_{\widehat{D}}= \LRa{\bm{f}^*-\bm{f}(\bm{u}_{\bm{v}}),\bm{w}}_{\partial \widehat{D}}, \qquad \bm{w}\in V_h,
+\LRp{\pd{\bm{u}}{t},\bm{w}}_{\widehat{D}} + \LRp{\left.\LRp{2D_N \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},\bm{w}}_{\widehat{D}}= \LRa{\bm{f}^*-\bm{f}(\bm{u}_{\bm{v}}),\bm{w}}_{\partial \widehat{D}}, \qquad \bm{w}\in P^N\LRp{\widehat{D}},
 \label{eq:ecdg_oneelem}
@@
 
-%We first introduce matrices defined over $[-1,1]$.
-We assume some discrete quadrature rule on the reference element consisting of $N_q$ points ${\bm{x}}_q = \LRc{{x}_1,\ldots,{x}_{N_q}}$ and quadrature weights $\bm{w}_q = \LRc{w_1,\ldots,w_{N_q}}$.
-Let $\bm{W}$ denote the diagonal matrix whose entries are quadrature weights.
-Then, assuming that each component of $\bm{u}$ is represented in some basis $\phi_j(x)$ of dimension $N_p$ which spans $V_h$
-%\[
-%V_h = {\rm span}\LRc{\phi_j(\widehat{x})}_{j=1}^{N_p},
-%\]
-%we can define the quadrature interpolation matrix $\bm{V}_q$
-%\[
-%\LRp{\bm{V}_q}_{ij} = \phi_j(\widehat{x}_i), \qquad 1 \leq j \leq N_p, \qquad 1 \leq i \leq N_q.  
-%\]
-%Given $\bm{V}_q$, we introduce the mass matrix $\bm{M} = \bm{V}_q^T\bm{W}\bm{V}_q$, as well as the quadrature-based $L^2$ projection operator $\bm{P}_q$
-%\[
-%\bm{P}_q = \bm{M}^{-1}\bm{V}_q^T\bm{W}.
-%\]
-%Let $\bm{D}_x$ denote the differentiation matrix defined implicitly through
-%\[
-%u(x) = \sum_{j=1}^{N_p}\bm{u} \phi_j(\widehat{x}), \qquad \pd{u}{\widehat{x}} = \sum_{j=1}^{N_p} \LRp{\bm{D}_x \bm{u}}_j\phi_j(\widehat{x})
-%\]
-%In other words, $\bm{D}_x$ maps basis coefficients of some function $u \in V_h$ to coefficients of its derivative on $[-1,1]$.  
-%
-%Finally, we introduce matrices used to compute boundary terms.
-Let $\bm{V}_f$ denote the matrix which interpolates to boundary values\footnote{In generalizing to higher dimensions, $\bm{V}_f$ would be the matrix which interpolates to surface quadrature points.}
-%\[
-%\LRp{\bm{V}_f}_{1j} = \phi_j(-1), \qquad \LRp{\bm{V}_f}_{2j} = \phi_j(1), \qquad 1 \leq j \leq N_p.  
-%\]
-%Furthermore, let $\bm{W}_f$ denote the diagonal matrix of surface quadrature weights, which is simply the $2\times 2$ identity matrix in one dimension.  
-%
-%Together, these matrices can be used to define the quadrature-based differentiation matrix $\bm{D}_q$
-%\[
-%\bm{D}_q = \bm{V}_q \bm{D}_x \bm{P}_q,
-%\]
-%which is a discrete analogue to $D_N$.  It is straightforward then to show that this matrix satisfies properties similar to summation-by-parts:
-%\begin{lemma}
-%Let $\bm{n}_x = [-1,1]$ be the vector of unit normals in one dimension.  Then, 
-%\[
-%\bm{W}\bm{D}_q = \bm{P}_q^T\bm{V}_f^T\bm{W}_f {\rm diag}(\bm{n}_x) \bm{V}_f \bm{P}_q  - \bm{D}_q^T \bm{W}, \qquad \bm{1}^T\bm{W}\bm{D}_q =  \bm{1}^T\bm{W}_f {\rm diag}(\bm{n}_x) \bm{V}_f \bm{P}_q.  
-%\]
-%\label{lemma:sbp}
-%\end{lemma}
-%\begin{proof}
-%\note{Finish}
-%\end{proof}
-
-%Let $\bm{u}^k$ now denote the vector of coefficients for the $k$th component of $\bm{u}(x)$.  The formulation (\ref{eq:ecdg_oneelem}) is equivalent to the semi-discrete system:
-%\begin{equation}
-%\bm{M} \td{\bm{u}^k}{t} + \bm{V}_q^T\bm{W}{\rm diag}\LRp{\bm{V}_q\bm{D}_x\bm{P}_q \bm{F}^k_S} = 0, \qquad k = 1,\ldots,n.
-%\label{eq:ecdg_oneelem_mat}
-%\end{equation}
-%Here, $\bm{F}^k_S$ denotes the matrix
-%\[
-%\LRp{\bm{F}^k_S}_{ij} = \bm{f}^k_S\LRp{\bm{u}_{x,i},\bm{u}_{y,j}}, \qquad 1 \leq i,j \leq N_q,
-%\]
-%where $\bm{f}^k_S$ is the $k$th component of the vector flux $\bm{f}_S$, and
-%\begin{equation}
-%\bm{u}_{x,i} = \bm{u}\LRp{\Pi_N \bm{v}(\bm{u}(x_i))}, \qquad \bm{u}_{y,j} = \bm{u}\LRp{\Pi_N \bm{v}(\bm{u}(x_j))}.  
-%\label{eq:projectedentropyvars}
-%\end{equation}
-
 \begin{theorem}
@@
 \[
-\bm{u}(x,t) = \sum_{j=1}^{N_p} \bm{u}_j(t) \phi_j(x), \qquad \pd{\bm{u}(x,t)}{t} \in V_h.
+\bm{u}(x,t) = \sum_{j=1}^{N_p} \bm{u}_j(t) \phi_j(x), \qquad \pd{\bm{u}(x,t)}{t} \in P^N\LRp{\widehat{D}}.
 \]
@@
 \[
-\LRp{\pd{\bm{u}}{t}, \Pi_N \bm{v}(\bm{u})}_{\widehat{D}} = \LRp{\pd{\bm{u}}{t},\bm{v}(\bm{u})}_{\widehat{D}} = \LRp{\pd{U(\bm{u})}{t},1}_{\widehat{D}}, \qquad \forall \bm{u}\in V_h.
+\LRp{\pd{\bm{u}}{t}, \Pi_N \bm{v}(\bm{u})}_{\widehat{D}} = \LRp{\pd{\bm{u}}{t},\bm{v}(\bm{u})}_{\widehat{D}} = \LRp{\pd{U(\bm{u})}{t},1}_{\widehat{D}}, \qquad \forall \bm{u}\in P^N\LRp{\widehat{D}}.
 \]
@@
 where we have used Definition~\ref{def:tadmor} in going from the first to second integral and Lemma~\ref{lemma:dgd_local} in the final step.  
-%\[
-%{\rm diag}\LRp{\bm{V}_q\bm{D}_x\bm{P}_q \bm{F}^k_S} \not\in V_h
-%\]
-%for general flux functions.  However, we may use Lemma~\ref{lemma:sbp} to expand out terms
-%\begin{align*}
-%\bm{v}^T\bm{V}_q^T\bm{W}{\rm diag}\LRp{\bm{V}_q\bm{D}_x\bm{P}_q \bm{F}^k_S} &= \bm{v}^T\bm{V}_q^T{\rm diag}\LRp{\LRp{\bm{P}_q^T\bm{V}_f^T\bm{W}_f {\rm diag}(\bm{n}_x) \bm{V}_f \bm{P}_q  + \bm{W}\bm{D}_q - \bm{D}_q^T\bm{W} } \bm{F}^k_S}.
-%\end{align*}
-%Because $\bm{F}^k_S$ is evaluated in terms of the projection of the entropy variables, we may apply Definition~\ref{def:tadmor} to simplify the latter terms
-%\begin{align*}
-%\bm{v}^T\bm{V}_q^T{\rm diag}\LRp{\LRp{\bm{W}\bm{D}_q - \bm{D}_q^T\bm{W} }\bm{F}^k_S} &=\sum_{i,j=1}^{N_q} \LRp{\bm{V}_q\bm{v}}_i \LRp{\bm{W}\bm{D}_q - \bm{D}_q^T\bm{W} }_{ij}\LRp{\bm{F}^k_S}_{ij}\\
-%&=\sum_{i,j=1}^{N_q} \LRp{\bm{W}\bm{D}_q}_{ij} \LRp{\psi_i - \psi_j}.  
-%\end{align*}
-%
-%The spatial term does not 
-%\[
-%\bm{v}\bm{V}_q^T\bm{W}{\rm diag}\LRp{\bm{V}_q\bm{D}_x\bm{P}_q \bm{F}^k_S}
-%\]
 \end{proof}
@@
 We see that, in the proof of Theorem~\ref{lemma:ec1}, recovering the time derivative of entropy on the left hand side requires testing with the projection of the entropy variables.
-For collocation-type DG methods, $L^2$ projection and interpolation are equivalent, which implies that evaluating the flux $\bm{f}_S$ using point-wise nodal values of the solution satisfies the entropy conservation condition given in Definition~\ref{def:tadmor}.
-For diagonal norm collocation-type DG methods, the method can also be simplified by noting that the projection operator reduces to a nodal interpolation operator.
+For diagonal norm SBP-DG methods, $L^2$ projection and interpolation are equivalent, which implies that evaluating the flux $\bm{f}_S$ using point-wise nodal values of the solution satisfies the entropy conservation condition given in Definition~\ref{def:tadmor}.
+Diagonal norm SBP-DG methods can also be simplified by noting that the projection operator reduces to a nodal interpolation operator.
 
@@
 However, the entropy conservation condition given in Definition~\ref{def:tadmor} holds so long as the flux is evaluated in a manner consistent with the entropy variables.
-This motivates (\ref{eq:projectedentropyvars}) as a non-standard way of evaluating the flux, which is the primary difference between the construction of entropy conservative schemes for diagonal norm collocation-type DG methods and projection-based DG methods.
+This motivates (\ref{eq:projectedentropyvars}) as a non-standard way of evaluating the flux, which is the primary difference between the construction of entropy conservative schemes for diagonal norm SBP-DG methods and projection-based DG methods.
 
@@
 
-%Global DG differential operators have been used to simplify the notation and analysis of DG formulations for elliptic equations \cite{di2011mathematical, wang2013weak}.
-Here, we will use their properties to derive entropy conservative DG formulations based on a global approach to flux differencing.
-We note that this approach does not require the specification of appropriate SBP-SAT terms in order to prove entropy conservation; rather, explicit expressions for SBP-SAT terms may be derived in a systematic manner using flux differencing and the definitions of global DG operators.
-
-We adapt the construction of local differentiation operators introduced previously to construct a global DG differentiation operator
-\begin{equation}
-\LRp{D^x_h u,vw}_{\Omega} = \sum_k \LRp{\pd{ \Pi_N u}{x},vw}_{D^k} + \frac{1}{2}{\LRa{{u^+ - \Pi_N u}, vw\bm{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{{u - \Pi_Nu },\Pi_N(vw) \bm{n}_x}_{\partial D^k}},
+We introduce the global $L^2$ inner product and the broken Sobolev space $H^1\LRp{\Oh}$ 
+\[
+\LRp{u,v}_{\Omega} = \sum_k \LRp{u,v}_{D^k}, \qquad H^1\LRp{\Oh} = \bigoplus_{k} H^1\LRp{D^k}.
+\]
+We adapt the construction of local differentiation operators introduced previously to construct a global DG differentiation operator:
+\begin{definition}
+Let $w(x)$ is a bounded function over each element $D^k$.  Define $D^x_h: H^1\LRp{\Oh}\rightarrow V_h$ be as the operator which satisfies
+\begin{align}
+&\LRp{D^x_h u,vw}_{\Omega} = \sum_k \LRp{\pd{ \Pi_N u}{x},vw}_{D^k} \nonumber\\
+&+ \frac{1}{2}{\LRa{{u^+ - \Pi_N u}, vw\bm{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{{u - \Pi_Nu },\Pi_N(vw) \bm{n}_x}_{\partial D^k}}, \qquad \forall v \in V_h.
 \label{eq:dgd_global}
-\end{equation}
-where $w(x)$ is a bounded smooth function over each element and $\LRp{D^x_h u,v}_{\Omega} = \sum_k \LRp{D^x_h u,v}_{D^k}$.  
+\end{align}
+\end{definition}
 Entropy conservative DG schemes can be constructed using the fact that $D^x_h$ satisfies a \textit{global} analogue of integration by parts:
@@
 \]
-for $u \in H^1(\Oh)$ and $\forall v\in V_h(\Omega_h)$.  
+for $u \in H^1(\Oh)$ and $\forall v\in V_h$.  
 \end{lemma}
 \begin{proof}
-We may rewrite $\LRp{D^x_h u,vw}_{\Omega}$ using the definition of $D_N$ in (\ref{eq:dgd})
+We may rewrite $\LRp{D^x_h u,vw}_{\Omega}$ using the Definition~\ref{eq:dgd} of $D_N$
 \[
@@
 \begin{equation}
-\sum_{k}\LRp{\pd{\bm{u}}{t},\bm{w}}_{D^k} + \LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},\bm{w}}_{D^k} = \LRa{\bm{f}(\bm{u}_{\bm{v}})-\bm{f}^*,\bm{w}}_{\partial \Omega},
+\LRp{\pd{\bm{u}}{t},\bm{w}}_{\Omega} + \LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},\bm{w}}_{\Omega} = \LRa{\bm{f}(\bm{u}_{\bm{v}})-\bm{f}^*,\bm{w}}_{\partial \Omega},
 %\int_{D^k}\pd{\bm{u}}{t} \bm{w} + \left.\LRp{2D^x_h\bm{f}_S\LRp{\bm{u}(\Pi_N\bm{v}({x})),\bm{u}(\Pi_N\bm{v}({y}))}}\right|_{{y}={x}}\bm{w} + \LRa{\bm{f}^*-\bm{f}(\bm{u}),\bm{w}}_{\partial \Omega}= 0, \qquad \forall \bm{w}\in V_h^n,
@@
 
-%\note{Add assumptions about $\bm{u}(x)$ - needs to be locally analytic.}
-
-%\note{Global DG operators reveal how to construct SBP-SAT terms for any collocation-type DG methods (where $L^2$ projection and interpolation are equivalent).
-Not easy to deal with SBP-SAT terms for more complex systems \cite{ranocha2017extended, ranocha2017comparison}.
-}
 This formulation yields a theorem analogous to Theorem~\ref{lemma:ec1}:
@@
 It is also straightforward to show that the formulation (\ref{eq:ecdg}) is locally conservative.
-We follow \cite{chen2017entropy} and take $\bm{w} = \mathbbm{1}_{D^k}$ in the formulation (\ref{eq:ecdg})
-\[
-\sum_{k}\LRp{\pd{\bm{u}}{t},1}_{D^k} + \LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},1}_{D^k} = \LRa{\bm{f}(\bm{u}_{\bm{v}})-\bm{f}^*,1}_{\partial \Omega \cap D^k}.
+We follow \cite{chen2017entropy} and take $\bm{w} = \mathbbm{1}_{D^k}$ in the formulation (\ref{eq:ecdg}) to yield
+\[
+\LRp{\pd{\bm{u}}{t},1}_{D^k} + \LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},1}_{D^k} = \LRa{\bm{f}(\bm{u}_{\bm{v}})-\bm{f}^*,1}_{\partial \Omega \cap D^k}.  
 \]
@@
 This can be used to practically evaluate $\left.\LRp{\bm{D}^x_h \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x}$ using only reference operators and evaluations of $\bm{f}_S$.
-However, unlike previous entropy conservative collocation DG methods, the boundary terms in (\ref{eq:eval}) interact with volume terms through the $L^2$ projection and the lift operator.
+However, unlike previous entropy conservative diagonal norm SBP-DG methods, the boundary terms in (\ref{eq:eval}) interact with volume terms through the $L^2$ projection and the lift operator.
 This can be simplified for diagonal norm SBP operators with boundary nodes, using equivalences between $L^2$ projection and interpolation at nodal points in those cases.
@@
 
-We first introduce matrices defined over $[-1,1]$.
-We assume some discrete quadrature rule on the reference element consisting of $N_q$ points ${\bm{x}}_q = \LRc{{x}_1,\ldots,{x}_{N_q}}$ and quadrature weights $\bm{w}_q = \LRc{w_1,\ldots,w_{N_q}}$.
+We first introduce matrices defined over the reference element $\widehat{D} = [-1,1]$.
 Let $\bm{W}$ denote the diagonal matrix whose entries are quadrature weights.
-Then, let $u(x)$ be represented in some basis $\phi_j(x)$ of dimension $N_p$ which spans $V_h$
-\[
-V_h = {\rm span}\LRc{\phi_i(\widehat{x})}_{i=1}^{N_p},
-\]
-we can define the quadrature interpolation matrix $\bm{V}_q$
+Then, assuming $u(x) \in P^N\LRp{\widehat{D}}$, it can be represented in some polynomial basis $\phi_j(\widehat{x})$ of degree $N$ and dimension $N_p$
+\[
+V_h = {\rm span}\LRc{\phi_i(\widehat{x})}_{i=1}^{N_p}.
+\]
+We can define the quadrature interpolation matrix $\bm{V}_q$
 \[
@@
 
-We first describe how to evaluate the volume term (\ref{eq:voleval}) using these matrices.  Let $\bm{F}_S$ denote the $N_q\times N_q$ matrix constructed by evaluating $f_S$ at volume quadrature points
+We now describe how to evaluate the volume term (\ref{eq:voleval}) using these matrices.\footnote{We can also use this matrix notation to directly show conservation of entropy.
+For example, the time term in the statement of discrete entropy conservation or dissipation can be derived as follows: let $\bm{u}_q, \bm{v}_q$ denote the conservation and entropy variables evaluated at quadrature points.
+Then, multiplying with the transpose of $\bm{v} = {\bm{P}_q\bm{v}_q}$ yields
+\[
+\LRp{\bm{P}_q\bm{v}_q}^T\bm{M}\pd{\bm{u}}{t} = \bm{v}_q^T\bm{W}\bm{V}_q\bm{M}^{-1}\bm{M}\pd{\bm{u}}{t} = \bm{1}^T \bm{W} {\rm diag}(\bm{v}_q) \pd{\bm{u}_q}{t} = \bm{1}^T \bm{W} \pd{U(\bm{u}_q)}{t},
+\]
+where we have assumed continuity in time, and have used that $\bm{W}$ is diagonal.  }  Let $\bm{F}_S$ denote the $N_q\times N_q$ matrix constructed by evaluating $f_S$ at volume quadrature points
 \[
@@
 
-%\note{
-%Open question: it's consistent with existing entropy stable collocation DG discretizations, but is the use of $\bm{u}\LRp{\Pi_N \bm{v}}$; is this consistent, can we show this converges?  
-%}
-
 \section{Acknowledgments}
@@
 \]
-where $\bm{f}_i, \bm{g}_i$ are real and bounded over each element $D^k$.
+where $\bm{f}_i, \bm{g}_i$ are real and bounded for appropriate values of $x,y$.
 This holds, for example, for analytic $\bm{f}_S$, for which the Taylor series can be rearranged to yield the form of the desired expansion.

### 2017-07-30 talks/USNCCM17/main.tex
adding USNCCM talk

\documentclass[compress]{beamer}
\mode<presentation>
\usetheme{Warsaw}
\usecolortheme{seagull}
%\useoutertheme[subsection=false]{smoothbars}

\usepackage{stackengine}
%\setbeamertemplate{caption}{\raggedright\insertcaption\par}
\setbeamertemplate{caption}{\insertcaption} 
%\usepackage{caption}

%% ====================================== graphics
\newcommand{\smiley}{\tikz[baseline=-0.75ex,black]{
    \draw circle (2mm);
\node[fill,circle,inner sep=0.5pt] (left eye) at (135:0.8mm) {};
\node[fill,circle,inner sep=0.5pt] (right eye) at (45:0.8mm) {};
\draw (-145:0.9mm) arc (-120:-60:1.5mm);
    }
}

\newcommand{\frownie}{\tikz[baseline=-0.75ex,black]{
    \draw circle (2mm);
\node[fill,circle,inner sep=0.5pt] (left eye) at (135:0.8mm) {};
\node[fill,circle,inner sep=0.5pt] (right eye) at (45:0.8mm) {};
\draw (-145:0.9mm) arc (120:60:1.5mm);
    }
}

\newcommand{\neutralie}{\tikz[baseline=-0.75ex,black]{
    \draw circle (2mm);
\node[fill,circle,inner sep=0.5pt] (left eye) at (135:0.8mm) {};
\node[fill,circle,inner sep=0.5pt] (right eye) at (45:0.8mm) {};
\draw (-135:0.9mm) -- (-45:0.9mm);
    }
}

\usepackage{pgfplots}
\pgfplotsset{width=10cm,compat=1.9}
%\usepgfplotslibrary{external}
%\tikzexternalize
 \usepackage{pgfplotstable}

%\definecolor{markercolor}{RGB}{.49, 1, .63}
\definecolor{markercolor}{RGB}{124.9, 255, 160.65}

\pgfplotsset{
tick label style={font=\scriptsize},
label style={font=\scriptsize},
legend style={font=\scriptsize},
title style={font=\footnotesize}
}

%% ===========================================

\usetikzlibrary{calc}

%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
\newcommand{\logLogSlopeTriangle}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        \pgfmathsetmacro{\xArel}{#1}
        \pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{\yArel}
        \pgfmathsetmacro{\xCrel}{\xArel}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- %node[pos=0.5,anchor=north] {1}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=west] {#4}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
\newcommand{\logLogSlopeTriangleFlip}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        %\pgfmathsetmacro{\xArel}{#1}
        %\pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{#3}
        \pgfmathsetmacro{\xCrel}{#1}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

        \pgfmathsetmacro{\xArel}{\xBrel}
        \pgfmathsetmacro{\yArel}{\yCrel}

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=east] {#4}
                    (B)-- 
                    (C)-- %node[pos=0.5,anchor=south] {1}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

\useoutertheme{infolines}
\useinnertheme{rectangles}
\usepackage{hhline}
\setbeamercovered{dynamic}

\usepackage{soul}

\usepackage{array}
\usepackage{amsmath,amssymb,amsfonts,mathrsfs,amsthm}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{mathtools}
\usepackage{dsfont}
\usepackage{pdfpages}
%\usepackage[textsize=footnotesize,color=green]{todonotes}
\usepackage{algorithm, algorithmic}
\usepackage{bm}
\usepackage{bbm}

\usepackage{tikz}
\usepackage[normalem]{ulem}

\usepackage{graphicx}
%\usepackage{subfigure}
\usepackage{subfig}
%\usepackage{caption}
%\usepackage{subcaption}

\usepackage{color}
\usepackage{pdflscape}
\usepackage{pifont}

\usepackage{bibentry}
\nobibliography*

%\usepackage[osf]{mathpazo}
%\usepackage{mathpazo}
%\renewcommand\rmdefault{ptm}
\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}}
\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

\newcommand{\vect}[1]{\ensuremath\boldsymbol{#1}}
\newcommand{\tensor}[1]{\underline{\vect{#1}}}
\newcommand{\del}{\triangle}
\newcommand{\grad}{\nabla}
\newcommand{\curl}{\grad \times}
\renewcommand{\div}{\grad \cdot}
\newcommand{\ip}[1]{\left\langle #1 \right\rangle}
\newcommand{\eip}[1]{a\left( #1 \right)}
\newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}#2}}
\newcommand{\pd}[3]{\frac{\partial^{#3} #1}{\partial#2^{#3}}}
\newcommand{\pdd}[2]{\frac{\partial^2#1}{\partial#2^2}}

\newcommand{\circone}{\ding{192}}
\newcommand{\circtwo}{\ding{193}}
\newcommand{\circthree}{\ding{194}}
\newcommand{\circfour}{\ding{195}}
\newcommand{\circfive}{\ding{196}}

\newcommand{\Reyn}{\rm Re}

\newcommand{\bs}[1]{\boldsymbol{#1}}
\DeclareMathOperator{\diag}{diag}

\newcommand{\equaldef}{\stackrel{\mathrm{def}}{=}}

\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\snor}[1]{\left| #1 \right|}
\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\Div} {\ensuremath{\nabla\cdot}}
\newcommand{\Nel} {\ensuremath{{N^\text{el}}}}
\newcommand{\jump}[1] {\ensuremath{\LRs{\![#1]\!}}}
\newcommand{\avg}[1] {\ensuremath{\LRc{\!\{#1\}\!}}}

\newcommand{\uh}{\widehat{u}}
\newcommand{\fnh}{\widehat{f}_n}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\pO}{\partial\Omega}
\newcommand{\Gh}{\Gamma_h}
\newcommand{\Gm}{\Gamma_{-}}
\newcommand{\Gp}{\Gamma_{+}}
\newcommand{\Go}{\Gamma_0}
\newcommand{\Oh}{\Omega_h}

%\newcommand{\nor}[1]{\left\| #1 \right\|}
%\newcommand{\snor}[1]{\left| #1 \right|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRb}[1]{\left| #1 \right|}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}

\newcommand{\bibfoot}[1]{\footnote{\tiny\bibentry{#1}}}
\renewcommand{\note}[1]{\textcolor{red}{{#1}}}

\newcommand{\eval}[2][\right]{\relax
  \ifx#1\right\relax \left.\fi#2#1\rvert}

\def\etal{{\it et al.~}}

\def\arr#1#2#3#4{\left[
\begin{array}{cc}
#1 & #2\\
#3 & #4\\
\end{array}
\right]}
\def\vecttwo#1#2{\left[
\begin{array}{c}
#1\\
#2\\
\end{array}
\right]}
\def\vectthree#1#2#3{\left[
\begin{array}{c}
#1\\
#2\\
#3\\
\end{array}
\right]}
\def\vectfour#1#2#3#4{\left[
\begin{array}{c}
#1\\
#2\\
#3\\
#4\\
\end{array}
\right]}

\newcommand{\G} {\Gamma}
\newcommand{\Gin} {\Gamma_{in}}
\newcommand{\Gout} {\Gamma_{out}}

%\newcommand\blfootnote[1]{%
%  \begingroup
%  \renewcommand\thefootnote{}\footnote{#1}%
%  \addtocounter{footnote}{-1}%
%  \endgroup
%}

% removes nav symbols
\beamertemplatenavigationsymbolsempty
%\setbeamertemplate{caption}{\raggedright\insertcaption\par}

% defines newblock as null, giving compile issues otherwise
\let\newblock\relax

\title[Time-domain IGA]{Time-domain multi-patch\\discontinuous Galerkin methods }
\date[May 19, 2017]{USNCCM 2017\\July 18, 2017}
\author[Chan, Evans]{Jesse Chan\inst{1}, J. A. Evans\inst{2}}
\institute[Rice CAAM]{\inst{1}Department of Computational and Applied Math, Rice University\\ \inst{2} Ann and H.J. Smead Aerospace Engineering Sciences, UC Boulder}

\begin{document}

\begin{frame}
\maketitle
\end{frame}

\frame{
\frametitle{High order Galerkin methods for hyperbolic PDEs}
\setcounter{subfigure}{0}
\vspace{-1em}
\begin{columns}
\begin{column}{.5\textwidth}
\begin{itemize}
\item<1-> Time-dependent solutions of hyperbolic equations. 
\vspace{.5em}
\item<2-> Low numerical dissipation and dispersion.
\vspace{.5em}
\item<4-> High order approximations: more accurate per unknown.
\vspace{.5em}
\item<5-> High performance on many-core (explicit time-stepping).
\end{itemize}
\end{column}
\begin{column}{.475\textwidth}
\begin{figure}
\centering
\begin{overlayarea}{\textwidth}{.5\textheight}
\only<1>{
\vspace{-2em}\includegraphics[width=.95\textwidth]{figs/wave.png}
\caption*{Figure courtesy of Axel Modave.}
}
%\only<2>{
%\vspace{1em}
%\includegraphics[width=.95\textwidth]{figs/turbulent1.png}\\
%\vspace{1em}
%\includegraphics[width=.95\textwidth]{figs/turbulent2.png}
%\caption*{Figure courtesy of Per-Olof Persson.}
%}
\only<2>{
\includegraphics[width=.95\textwidth]{figs/wave_N1.eps}
\caption*{\textbf{Fine} linear approximation.}
}
\only<3-4>{
\includegraphics[width=.95\textwidth]{figs/wave_N2.eps}
\caption*{\textbf{Coarse} quadratic approximation.}
}
\only<5>{
\includegraphics[width=.975\textwidth]{figs/gpu.pdf}
\caption*{A graphics processing unit (GPU).}
}
%\caption*{Image courtesy of Axel Modave.}
\end{overlayarea}
\end{figure}
\end{column}
\end{columns}
}

\frame{
\frametitle{Multi-patch discontinuous Galerkin formulations}
\begin{columns}
\begin{column}{.5\textwidth}
\begin{itemize}
\item Model problem: (second order) acoustic wave equation 
\begin{align*}
%\only<1>{
\frac{1}{c^2}\pdd{p}{t}{} - \Delta p &= f.
%}
%\only<2>{
%\pd{v}{t}{} &= \sum_{D^k}\LRp{\Grad p,\grad w}_{D^k} + \LRa{\jump{p},w}\\
%\pd{p}{t}{} &= v.
%}
\end{align*}
\item Pressure-velocity system (first order formulation)
\begin{align*}
\frac{1}{c^2}\pd{p}{t}{} + \Div \bm{u} &= f\\
\pd{\bm{u}}{t}{} + \Grad p &= 0.
\end{align*}
\end{itemize}
\end{column}

\begin{column}{.5\textwidth}
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figs/multipatch_diagram.pdf}
\end{figure}
\end{column}

\end{columns}

\vspace{1em}
\begin{itemize}
\item Weak patch coupling through DG numerical flux (SIPG, upwind).  
\end{itemize}

\let\thefootnote\relax\footnotetext{\tiny Langer et al 2014. Multipatch discontinuous Galerkin isogeometric analysis.}
\let\thefootnote\relax\footnotetext{\tiny Wilcox et al 2010. A high-order DG method for wave propagation through coupled elastic-acoustic media.}
}

\frame{
\frametitle{Method of lines multi-patch DG discretization}
\vspace{-1em}
\begin{itemize}
\item Semi-discrete system:
\[
\bm{M}_h\td{\bm{u}}{t} = \bm{A}_h\bm{u} \quad \Rightarrow \quad \td{\bm{u}}{t} = \bm{M}_h^{-1}\bm{A}_h\bm{u}.
\]
\item Global mass matrix $\bm{M}_h$ is (patch) block diagonal.  
\vspace{1em}
\item Tailored patch discretizations: explicit time-stepping, curved domains. 
%\vspace{.25em}
\begin{itemize}
\item Efficient mass matrix inversion over each patch.
\vspace{.5em}
\item High order accuracy, energy stability on curved geometries.
\vspace{.5em}
\item Improved stable timestep restriction (CFL).
\end{itemize}
\end{itemize}
}

\section{Spline spaces and optimal knot vectors }
\frame[noframenumbering]{
\frametitle{Outline}
\only<1>{
\tableofcontents
}
\only<2>{
\tableofcontents[currentsection]
}
}

%\frame{
%\frametitle{Spline spaces and sup-infs}
%
%\begin{itemize}
%\item Let $X$ be Banach, and $X_n$ be some $n$-dimensional subspace of $X$
%\item 
%\item The $n$-width of $X$ is then defined to be the infimum over all subspaces $X_n$ with ${\rm dim}\LRp{X_n} = n$, and a specific subspace $X_n$ is referred to as an \textit{optimal} subspace if its sup-inf attains the $n$-width.
%\item \note{FINISH}
%\end{itemize}
%}
\frame{
\frametitle{B-spline basics}

\begin{itemize}
\item Standard 1D B-splines: $B^0_i(x) = \mathbbm{1}_{\xi_i \leq x \leq \xi_{i+1}}$, 
\[
B^k_i(x) = \frac{x - \xi_i}{\xi_{i+p}-\xi_i}B^{k-1}_i(x) + \frac{\xi_{i+p+1}-x}{\xi_{i+p+1}-\xi_{i+1}}B^{k-1}_{i+1}(x).
\]
\item Assume patches are mapped quads/cubes: tensor product basis
\[
B^p_{ijk}(x,y,z) = B^p_{i}(x)B^p_{j}(y)B^p_{k}(z).
\]
\item Assume maximally continuous, open knot vectors
\begin{align*}
%\bm{\Xi} &= \LRc{-1 = \xi_1,\ldots,\xi_{{2p}+K+1} = 1}, \qquad \xi_i \leq \xi_{i+1},\\
\xi_{p+1} &< \ldots < \xi_{p+1+K},\\
\xi_1 &= \ldots = \xi_{p+1},\\
\xi_{p+1 + K} &= \ldots = \xi_{2p+1+K}.\\
\end{align*}
\end{itemize}
}
\frame{
\frametitle{B-spline bases and optimal spline spaces}
\vspace{-1em}
\begin{figure}
\centering
\subfloat[Uniform knots]{\includegraphics[width=.32\textwidth]{figs/unifknots.png}}
\hspace{2.5em}
\subfloat[Optimal knots]{\includegraphics[width=.32\textwidth]{figs/optknots.png}}
%\caption{B-spline bases with uniform and optimal knot vectors ($p = 4, K = 4$).}
\end{figure}

\begin{itemize}
\item Sup-inf: ``worst best approximation'' in $X$ from $X_n$ 
\[
d_n(X;X_n) =  \sup_{x\in X} \inf_{y\in X_n} \nor{x - y}, \qquad {\rm dim}\LRp{X_n} = n.
\]
\item Spline spaces with optimal knot vectors: minimal sup-inf for
\[
X = \LRc{f \in L^2([-1,1]): \pd{f}{x}{p-1} \text{ continuous}, \quad \nor{f}_{L^2} \leq 1},
\]
\end{itemize}

\let\thefootnote\relax\footnotetext{\tiny Melkman and Micchelli 1978. Spline spaces are optimal for $L^2$ $n$-width.}
}

\frame{
\frametitle{Optimal knot vectors: roots of eigenfunctions}
\setcounter{subfigure}{0}
\vspace{-1em}
\begin{figure}
\centering
\subfloat[$p = 2$]{\includegraphics[width=.33\textwidth]{figs/fnwidth_N2_K8.png}}
\subfloat[$p = 3$]{\includegraphics[width=.33\textwidth]{figs/fnwidth_N3_K8.png}}
\subfloat[$p = 4$]{\includegraphics[width=.33\textwidth]{figs/fnwidth_N4_K8.png}}
\caption{Eigenfunctions $y_{K+1,p}(x)$ for $K=8$ and various $p$.}
\end{figure}
\vspace{-1em}
\begin{itemize}
\item Optimal knots are roots of eigenfunctions $y_{K+1,p}(x)$.
\[
(-1)^p \frac{\partial^{2p}y}{\partial x^{2p}} = \lambda y(x), \qquad \frac{\partial^k y}{\partial x^k}(-1) = \frac{\partial^k y}{\partial x^k}(1) = 0, \quad 1\leq k \leq p-1.
\]
\item Approximate $y_{K+1,p}(x)$ using fine spline space; difficult for high $K, p$!
\end{itemize}
}

\frame{
\frametitle{Knot smoothing: approximating optimal knots}
\vspace{-.25em}
\begin{figure}
\begin{overlayarea}{\textwidth}{.5\textheight}
\centering
\only<1-2>{
\setcounter{subfigure}{0}
\subfloat[Knots $\xi_i$ (optimal in red)]{\includegraphics[width=.375\textwidth]{figs/knots.png}}
\hspace{2em}
\subfloat[Greville abscissae $\tau_j$]{\includegraphics[width=.375\textwidth]{figs/greville.png}}
}
\only<3>{
\setcounter{subfigure}{0}
\subfloat{
\begin{tikzpicture}
\begin{axis}[
width=.425\textwidth,
    xlabel={Number of elements $K$},   
        ylabel={Max knot error},
    xmin=2, xmax=66,
    ymax=0,ymax=0.0125,    
    legend pos=north east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
]

\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
coordinates{(4,0.0121)(8,0.0075)(16,0.0046)(32,0.0023)(64,0.0012)};

\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
coordinates{(4,0.009)(8,0.0041)(16,0.0028)(32,0.002)(64,0.0012)};

\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
coordinates{(4,0.0016)(8,0.0101)(16,0.0074)(32,0.0048)(64,0.0028)};

\legend{$p = 2$, $p=3$, $p=4$}
\end{axis}
\end{tikzpicture}
}
\hspace{2em}
\subfloat{
\begin{tikzpicture}
\begin{axis}[
width=.425\textwidth,
    xlabel={Number of elements $K$},   
    ylabel={Max Greville error},
    xmin=2, xmax=66,
    ymax=0,ymax=0.0125,    
    legend pos=north east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
]

\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
coordinates{(4,0.006)(8,0.0038)(16,0.0024)(32,0.0011)(64,0.00058806)};

\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
coordinates{(4,0.003)(8,0.002)(16,0.0018)(32,0.0012)(64,0.00067857)};

\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
coordinates{(4,0.000401)(8,0.0025)(16,0.0016)(32,0.00089352)(64,0.00063901)};

\legend{$p = 2$, $p=3$, $p=4$}
\end{axis}
\end{tikzpicture}
}
}
\end{overlayarea}
%\caption{Knot locations for iterations $k = 0, \ldots, 10$ for a spline space with $p = 3, K = 8$.  Computed optimal knot positions and Greville abscissae are overlaid in \textcolor{red}{red}.}
\label{fig:smoothedknots}
\end{figure}
\vspace{-1.25em}
\begin{itemize}
\item Greville abscissae $\tau_j$: coefficients for linear coordinate $x$.
\only<1>{
\[
x = \sum_{1\leq j \leq p+K} \tau_j B^p_j(x), \qquad \tau_j = \frac{1}{p}\sum_{1\leq i \leq p} \xi_{i+j-1}, \quad j = 1,\ldots,p.
\]  
}
\only<2->{
\[
\textcolor{red}{\xi_i} = \sum_{1\leq j \leq p+K} \tau_j B^p_j(\textcolor{red}{\xi_i}), \qquad \tau_j = \frac{1}{p}\sum_{1\leq i \leq p} \xi_{i+j-1}, \quad j = 1,\ldots,p.
\]  
}
\item Replace Greville abscissae with equispaced points $\widehat{x}_i$ and iterate 
\[
\tilde{\xi}^{k+1}_i = \sum_{1\leq j\leq p+K} \widehat{x}_i B^p_j(\xi_i;  \tilde{\xi}^k), \qquad \tilde{\xi}^0_i = \xi_i,
\]
\end{itemize}
}

\frame{
\frametitle{Approximation properties: oscillatory functions}
\setcounter{subfigure}{0}

\vspace{-1em}
\begin{figure}
\centering
\begin{overlayarea}{\textwidth}{.7\textheight}
\only<1>{
\subfloat[$L^2$ error vs $k/N_{\rm dofs}$ ($\cos\LRp{\frac{k\pi x}{2}}$)]{\includegraphics[width=.49\textwidth]{figs/splineapprox.png}}
\hspace{.25em}
\subfloat[Semi-log scale]{\includegraphics[width=.49\textwidth]{figs/splineapproxlog.png}}
%\caption{}
}
\only<2>{
\subfloat[First order formulation]{
\begin{tikzpicture}
\begin{semilogyaxis}[
width=.45\textwidth,
    xlabel={Wavelength $k$},   
    ylabel={$L^2$ error},
    ymin=1e-5, ymax=2,    
    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
] 
\addplot[color=blue,mark=*, dashdotted, semithick, mark options={solid,fill=markercolor}]
coordinates{(1,1.86794e-05)(3,0.00199192)(5,0.0632619)(7,0.616786)(9,0.712371)(11,0.768705)};
\addplot[color=red,mark=*, dashdotted, semithick, mark options={solid,fill=markercolor}]
coordinates{(1,5.56475e-05)(3,0.00612773)(5,0.0351819)(7,0.405041)(9,0.772862)(11,0.731632)};
\legend{Uniform knots, Smoothed knots}
\end{semilogyaxis}
\end{tikzpicture}
}
\subfloat[Second order formulation]{
\begin{tikzpicture}
\begin{semilogyaxis}[
width=.45\textwidth,
    xlabel={Wavelength $k$},   
    ylabel={$L^2$ error},
    ymin=1e-5, ymax=2,    
    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
] 
\addplot[color=blue,mark=*, dashdotted, semithick, mark options={solid,fill=markercolor}]
coordinates{(1,1.77384e-05)(3,0.00244169)(5,0.0592546)(7,1.07406)(9,0.662767)(11,0.872701)};
\addplot[color=red,mark=*, dashdotted, semithick, mark options={solid,fill=markercolor}]
coordinates{(1,1.78285e-05)(3,0.00465007)(5,0.107561)(7,0.751711)(9,0.864029)(11,0.773654)};

\legend{Uniform knots, Smoothed knots}
\end{semilogyaxis}
\end{tikzpicture}
}
%\caption{Comparison of $L^2$ errors for uniform and smooth knots as a function of wavenumber $k$ using $p = 3, K= 8$ splines.}
}
\end{overlayarea}
\end{figure}
\vspace{-3em}
\begin{itemize}
\item<1-> $L^2$ best approximations with smoothed knots (fixed $p,K$): decreased high frequency errors, increased low frequency errors.   
\item<2-> Similar errors approximating solution $\cos\LRp{\frac{k\pi x}{2}}\cos\LRp{\frac{k\pi t}{2}}$ ($k$ odd).  
\end{itemize}

}

\frame{
\frametitle{Spectral properties of $\bm{A}_h$}
\setcounter{subfigure}{0}
\vspace{.5em}
\only<2>{
Constant 1D advection: numerical dispersion error $\LRb{{\rm Re}(\omega) - {\rm Re}(\omega_h)}$
\[
\pd{u}{t}{} + \pd{u}{x}{} = 0, \qquad u(x,t) = e^{i(kx - \omega t)}, \quad u_h(x,t) = e^{i(kx-\omega_h t)}.
\]
\vspace{-2em}
\begin{figure}
\centering
\subfloat[Dispersion errors (uniform knots)]{
\includegraphics[width=.45\textwidth]{figs/dispersionErrUnif.png}
}
\subfloat[Dispersion errors (smoothed knots)]{
\includegraphics[width=.45\textwidth]{figs/dispersionErrSmooth.png}
}
\caption{Dispersion errors using full upwinding, $p = 4$, and $K = 1, 4, 8$. } %The exact dispersion relation is given as a black dotted line.}% in Figures~\ref{subfig:disp1}, \ref{subfig:disp2}.  }
%\label{fig:eigerr1}
\end{figure}

}
\only<1>{
Eigenvalues, eigenfunctions of 1D Laplacian: uniform vs smoothed knots.
\[
\LRb{\lambda_k-\lambda_{h,k}}, \qquad \nor{w_k(x)-w_{h,k}(x)}_{L^2}^2
\]
\vspace{-1.5em}
%\[
%-\pd{u}{x}{2} = \lambda u, \qquad \LRp{\Grad u, \Grad v}_{\Omega} = \lambda_h \LRp{u,v}_{\Omega}, \quad \forall v\in V_h.
%\]
\begin{figure}
\centering
\subfloat[Eigenvalue errors]{
\includegraphics[width=.45\textwidth]{figs/eigvals2ndorder.png}
}
\subfloat[Eigenvector errors]{
\includegraphics[width=.45\textwidth]{figs/eigfunc2ndorder.png}
}
\caption{Eigenvalue and eigenvector errors for $p = 4, K = 32$ splines. }
\end{figure}
}
}

\section{Curved domains and weight-adjusted mass matrices}
\frame[noframenumbering]{
\tableofcontents[currentsection]
}

\frame{
\frametitle{Approximation properties: curvilinear domains}

\begin{overlayarea}{\textwidth}{\textheight}

\begin{itemize}
\item<1-> Smoothed knot vectors: more accurate on curved domains.
\item<2-> Differences between first, second order forms ($L^2$ vs energy norm?).
\end{itemize}
\vspace{-1.5em}
%\begin{overlayarea}{\textwidth}{.725\textheight}
\only<1>{
\begin{figure}
\centering
\subfloat[Warped mesh, $\alpha = 1/8$]{\includegraphics[width=.51\textwidth]{figs/mapped.png}}
\subfloat[$L^2$ errors ($\alpha = 1/64$) ]{
\begin{tikzpicture}
\begin{semilogyaxis}[
    legend cell align=left,
    legend style={legend pos=north east, font=\tiny},
    width=.48\textwidth,
    xlabel={Degrees of freedom},
    ylabel={$L^2$ error}, 
    grid style=dashed,
] 
\addplot[color=blue,mark=*,semithick, mark options={solid,fill=markercolor}]
coordinates{(16,0.0345598)(36,0.00291687)(64,0.000910808)(100,0.000228283)(144,4.68154e-05)(196,1.08049e-05)(256,2.35056e-06)};
%[yshift=4pt] node[above, pos=.8, color=black] {$\alpha = 1/64$};
\addplot[color=red,mark=square*,semithick, mark options={solid,fill=markercolor}]
coordinates{(16,0.0345598)(36,0.0018151)(64,0.000489718)(100,8.29769e-05)(144,1.29416e-05)};
\addplot[color=black,mark=triangle*,semithick, mark options={solid,fill=markercolor}]
coordinates{(16,0.0345598)(36,0.00184734)(64,0.000495297)(100,8.68887e-05)(144,1.63983e-05)(196,1.69855e-06)(256,4.46827e-07)};
\addplot[color=violet,mark=diamond*,semithick, mark options={solid,fill=markercolor}]
coordinates{(9,0.0353096)(16,0.0346236)(25,0.00282808)(36,0.00233183)(49,0.00142429)(64,0.000710684)(81,0.000305903)(100,0.000123394)(121,4.1829e-05)(144,2.40683e-05)(169,6.2327e-06)(196,5.06423e-06)(225,1.05795e-06)(256,8.08652e-07)};
\legend{Uniform, Optimal, Smoothed, Polynomial}
%\legend{Uniform, Optimal, Smoothed}
\end{semilogyaxis}
\end{tikzpicture}
}
\caption{$L^2$ approx.\ errors: $\cos\LRp{\frac{\pi x}{2}}\cos\LRp{\frac{\pi y}{2}}$, $p = 2,\ldots, 8$ and $K = p$. }
\end{figure}
}

\only<2>{
\vspace{.5em}
\begin{figure}
\centering
\subfloat[First order formulation]{
\begin{tikzpicture}
\begin{semilogyaxis}[
width=.45\textwidth,
    xlabel={Warping factor $\alpha$},   
    ylabel={$L^2$ error},
    ymin=1e-3, ymax=.1,    
    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
] 
\addplot[color=blue,mark=*, dashdotted, semithick, mark options={solid,fill=markercolor}]
coordinates{(0,0.00462693)(0.0125,0.007073)(0.025,0.0116375)(0.0375,0.0166377)(0.05,0.021786)(0.0625,0.0270765)(0.075,0.0326103)(0.0875,0.0385547)(0.1,0.0451154)(0.1125,0.0525108)(0.125,0.0609548)};
\addplot[color=red,mark=*, dashdotted, semithick, mark options={solid,fill=markercolor}]
coordinates{(0,0.00458676)(0.0125,0.0055373)(0.025,0.00775311)(0.0375,0.0105429)(0.05,0.0136791)(0.0625,0.0171546)(0.075,0.0210896)(0.0875,0.0257146)(0.1,0.0313525)(0.1125,0.0383814)(0.125,0.0471834)};

\legend{Uniform knots, Smoothed knots}
\end{semilogyaxis}
\end{tikzpicture}
}
\subfloat[Second order formulation]{
\begin{tikzpicture}
\begin{semilogyaxis}[
width=.45\textwidth,
    xlabel={Warping factor $\alpha$},   
    ylabel={$L^2$ error},
    ymin=1e-3, ymax=.1,    
    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
] 
\addplot[color=blue,mark=*, dashdotted, semithick, mark options={solid,fill=markercolor}]
coordinates{(0,0.00504237)(0.0125,0.00834838)(0.025,0.0142055)(0.0375,0.0205381)(0.05,0.0269995)(0.0625,0.0335261)(0.075,0.0400686)(0.0875,0.0466709)(0.1,0.0536183)(0.1125,0.0616203)(0.125,0.0708213)};
\addplot[color=red,mark=*, dashdotted, semithick, mark options={solid,fill=markercolor}]
coordinates{(0,0.0113237)(0.0125,0.0118539)(0.025,0.013336)(0.0375,0.0157542)(0.05,0.0189697)(0.0625,0.0229092)(0.075,0.027875)(0.0875,0.0339693)(0.1,0.0406825)(0.1125,0.0473484)(0.125,0.0549665)};

\legend{Uniform knots, Smoothed knots}
\end{semilogyaxis}
\end{tikzpicture}
}
\caption{$L^2$ errors for increasing $\alpha$ ($p = 3, K= 8$ splines on a single patch).}
\label{fig:warpingconverge}
\end{figure}
}
\end{overlayarea}
}

\frame{
\frametitle{Approximate mass matrix inversion}

\begin{itemize}
\item Patch mass matrices $\bm{M}_J$: tensor product basis but no Kronecker structure due to $J(\widehat{\bm{x}})$, and mass lumping is inaccurate.
\[
\LRp{\bm{M}_J}_{ijk,lmn} = \int_{\widehat{D}} B^p_{ijk}(\widehat{\bm{x}})B^p_{lmn}(\widehat{\bm{x}}) \textcolor{red}{J(\widehat{\bm{x}})}\diff{\widehat{\bm{x}}}.
\]
%\vspace{.25em}
\item Preconditioning: approximating $\bm{M}_J^{-1}$ impacts semi-discrete stability.  Note - Krylov methods approximate $\bm{M}_J^{-1}$ as a non-linear operator!
\[
\frac{1}{2}\td{}{t}\nor{u}_{L^2}^2 = \bm{u}^T\bm{M}_J\td{\bm{u}}{t} \leq 0.%= \bm{u}^T\bm{A_h}\bm{u} \leq 0.
\]
%\vspace{.25em}
\item Isogeometric collocation: restores tensor product structure, but semi-discrete stability is more difficult to prove.  
\end{itemize}

\let\thefootnote\relax\footnotetext{\tiny Gao and Calo 2014. Fast isogeometric solvers for explicit dynamics.}
%\let\thefootnote\relax\footnotetext{\tiny Evans et al 2017.  Explicit Isogeometric Collocation Methods for Structural Dynamics by Means of Predictor-Multicorrector Schemes}
\let\thefootnote\relax\footnotetext{\tiny Wathen and Rees 2009. Chebyshev semi-iteration in preconditioning for problems including the mass matrix.}
\let\thefootnote\relax\footnotetext{\tiny Auricchio et al 2012. Isogeometric collocation for elastostatics and explicit dynamics.}
}

\frame{
\frametitle{Restoring Kronecker structure to $\bm{M}^{-1}$}

\begin{itemize}
\item Replace $\bm{M}_J$ with ``weight-adjusted'' approximation:
\[
\td{}{t} \bm{M}_J\bm{u} \Rightarrow \td{}{t}\textcolor{red}{\widehat{\bm{M}}\bm{M}_{1/J}^{-1}\widehat{\bm{M}}} \bm{u}, \qquad \LRp{\widehat{\bm{M}}}_{ijk,lmn} = \int_{\widehat{D}} B^p_{ijk}(\widehat{\bm{x}})B^p_{lmn}(\widehat{\bm{x}})\diff{\widehat{\bm{x}}}.
\]
%\vspace{.5em}
\item Weight-adjusted inverse: tensor product + matrix-free eval.\ of $\bm{M}_{1/J}$
\begin{align*}
\bm{M}^{-1}_J &\approx \LRp{\widehat{\bm{M}}\bm{M}_{1/J}^{-1}\widehat{\bm{M}}}^{-1}  = \widehat{\bm{M}}^{-1} \bm{M}_{1/J} \widehat{\bm{M}}^{-1}\\
\widehat{\bm{M}}^{-1} &= \widehat{\bm{M}}^{-1}_{\rm 1D} \otimes \widehat{\bm{M}}^{-1}_{\rm 1D} \otimes \widehat{\bm{M}}^{-1}_{\rm 1D}.
\end{align*}
\item Energy stability with respect to equivalent norm 
\[
C_1(J) \nor{\bm{u}}_{\widehat{\bm{M}}\bm{M}_{1/J}^{-1}\widehat{\bm{M}}} \leq \nor{\bm{u}}_{\bm{M}_J} \leq C_2 \nor{\bm{u}}_{\widehat{\bm{M}}\bm{M}_{1/J}^{-1}\widehat{\bm{M}}}.
\]
%\vspace{.5em}
%\item The same approach has also been successfully applied to Galerkin difference methods.  
\end{itemize}

%\let\thefootnote\relax\footnotetext{\tiny Mantzaflaris et al 2016. Low rank tensor methods in Galerkin-based IGA.}
%\let\thefootnote\relax\footnotetext{\tiny Kirby 2017. Fast inversion of the simplicial Bernstein mass matrix.}
\let\thefootnote\relax\footnotetext{\tiny Chan, et al.\ 2016.  Weight-adjusted DG methods: wave prop.\ in heterogeneous media. (SISC).  }
\let\thefootnote\relax\footnotetext{\tiny Chan, et al.\ 2016.  {Weight-adjusted DG methods: curvilinear meshes} (arXiv).  }
\let\thefootnote\relax\footnotetext{\tiny Also applied to GD methods (Banks and Hagstrom 2016. On Galerkin difference methods).}
}

\frame{
\frametitle{Weight-adjusted DG for curvilinear meshes}

\begin{itemize}
\item Weight-adjusted projection $\tilde{P}_h$ on curved domains
\[
\tilde{P}_h(u) \coloneqq \widehat{P}_h\LRp{\frac{1}{J}\widehat{P}_h\LRp{uJ}}.
\]
where $\widehat{P}_h$ is the $L^2$ projection onto the reference element.
\vspace{1em}
\item $L^2$ estimates for weight-adjusted projection:
\begin{align*}
\nor{u - \tilde{P}_h u}_{L^2\LRp{D^k}} &\lesssim \nor{\frac{1}{\sqrt{J}}}_{L^{\infty}}^2 \textcolor{red}{\nor{J}_{W^{N+1,\infty}\LRp{D^k}}} h^{N+1}\nor{u}_{W^{N+1,2}\LRp{D^k}}.
\end{align*}
High order Sobelev norm of $J$ - can lose accuracy if the geometric mapping is not sufficiently regular!
\end{itemize}
}

\frame{
\frametitle{$L^2$ vs weight-adjusted projection}

\vspace{-1em}
\begin{figure}
\centering
\subfloat[Warped mesh, $\alpha = 1/8$]{\includegraphics[width=.51\textwidth]{figs/mapped.png}}
\subfloat[$L^2$ errors ]{
\begin{tikzpicture}
\begin{loglogaxis}[
    legend cell align=left,
    legend style={legend pos=south east, font=\tiny},
    width=.475\textwidth,
    xlabel={Mesh size $h$},
    ylabel={$L^2$ error}, 
    grid style=dashed,
]

\addplot[color=blue,mark=*,semithick, mark options={solid,fill=markercolor}]
coordinates{(1,0.00192384)(0.5,0.00050273)(0.25,1.98999e-05)(0.125,3.15657e-07)};
\addplot[color=red,mark=x,semithick, mark options={solid,fill=markercolor}]
coordinates{(1,0.0019422)(0.5,0.000506873)(0.25,1.9966e-05)(0.125,3.15663e-07)};
\addplot[color=black,mark=triangle*,semithick, mark options={solid,fill=markercolor}]
coordinates{(1,1.836e-05)(0.5,4.14359e-06)(0.25,6.60985e-08)(0.125,5.88659e-12)};

\legend{$L^2$ projection,Weight-adjusted,Difference}
%\legend{Uniform, Optimal, Smoothed}
\end{loglogaxis}
\end{tikzpicture}
}
\caption{$L^2$ vs weight-adjusted projection: $\cos\LRp{\frac{\pi x}{2}}\cos\LRp{\frac{\pi y}{2}}$, $p = 4$. }
%\caption{$L^2$ vs weight-adjusted projection. }
\end{figure}

}

\section{Stable timestep restrictions}
\frame[noframenumbering]{
\tableofcontents[currentsection]
}

\frame{
\frametitle{Stable timestep restrictions: $h, p$ scaling}
\setcounter{subfigure}{0}

\begin{itemize}
\item Estimate $dt \leq 1/\rho(\bm{M}_h^{-1}\bm{A}_h)$.  Bound $\rho(\bm{M}_h^{-1}\bm{A}_h)$ using Rayleigh quotients and Bendixon-Hirsch; depends on $h$ and \textbf{constants} $C_T, C_I$.
\[
\nor{u}_{L^2\LRp{\partial \widehat{D}}} \leq C_T \nor{u}_{L^2\LRp{\widehat{D}}}, \qquad \nor{\Grad u}_{L^2\LRp{\widehat{D}}} \leq C_I \nor{u}_{L^2\LRp{\widehat{D}}}.
\]
\item CFL: $O(h/ p^2)$ for polynomials, $\textcolor{red}{O(h/ p)}$ for splines if $h \leq O(1/p)$.  
\end{itemize}
\begin{overlayarea}{\textwidth}{.65\textheight}
\only<1>{
\vspace{-.5em}
\begin{figure}
\centering
\subfloat[Inverse inequality, $K = 2p$]{
\begin{tikzpicture}
\begin{axis}[
width=.45\textwidth,
    xlabel={Degree of approximation $p$},   
        ylabel={$C_I(p)$},
    xmin=.5, xmax=12.5,
    ymax=0,ymax=100,    
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
    xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
]

\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
coordinates{(1,2)(2,4.5)(3,8)(4,12.5)(5,18)(6,24.5)(7,32)(8,40.5)(9,50)(10,60.5)(11,72)(12,84.5)};
\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
coordinates{(1,2)(2,3.42857)(3,5.27022)(4,7.3491)(5,9.61122)(6,12.0258)(7,14.5716)(8,17.233)(9,19.9977)(10,22.8559)(11,25.7995)(12,28.8217)};
\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
coordinates{(1,2)(2,3.15356)(3,4.41024)(4,5.72791)(5,7.09105)(6,8.48919)(7,9.91299)(8,11.3583)(9,12.8221)(10,14.3019)(11,15.796)(12,17.3028)};

\legend{Polynomial $C_I$, Uniform $C_I/K$, Smoothed $C_I/K$ }
\end{axis}
\end{tikzpicture}
}
\hspace{1em}
\subfloat[Trace inequality, $K = 2p$]{
\begin{tikzpicture}
\begin{axis}[
width=.45\textwidth,
    xlabel={Degree of approximation $p$},   
    ylabel={$C_T(p)$},
    xmin=.5, xmax=12.5,
    ymax=0,ymax=80,    
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
    xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
]

\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
coordinates{(1,1.73205)(2,3.87298)(3,6.5216)(4,9.74981)(5,13.5914)(6,18.0596)(7,23.1598)(8,28.894)(9,35.2634)(10,42.2685)(11,49.9096)(12,58.187)};
\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
coordinates{(1,1.73205)(2,2.43668)(3,3.48657)(4,4.75661)(5,6.16327)(6,7.6751)(7,9.27477)(8,10.9506)(9,12.694)(10,14.4983)(11,16.3578)(12,18.2682)};
\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
coordinates{(1,1.73205)(2,2.32019)(3,2.9916)(4,3.77273)(5,4.61001)(6,5.4806)(7,6.37331)(8,7.28317)(9,8.20701)(10,9.14265)(11,10.0885)(12,11.0432)};

\legend{Polynomial $C_T$, Uniform $C_T / K $, Smoothed $C_T/K$ }
\end{axis}
\end{tikzpicture}
}
\end{figure}
}
\only<2>{
\vspace{-1em}
\begin{figure}
\centering
\subfloat[Uniform knots]{
\begin{tikzpicture}
\begin{axis}[
width=.45\textwidth,
    xlabel={Number of elements $K$},   
    ylabel={$\rho\LRp{\bm{M}_h^{-1}\bm{A}_h}$},       
    xmin=0, xmax=300,
    ymax=0,ymax=1600,    
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
] 
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,66.4194)(64,132.839)(96,199.258)(128,265.678)(160,332.097)(192,398.517)(224,464.936)(256,531.355)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,101.838)(64,203.677)(96,305.515)(128,407.354)(160,509.192)(192,611.031)(224,712.869)(256,814.707)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,141.497)(64,282.993)(96,424.49)(128,565.987)(160,707.484)(192,848.98)(224,990.477)(256,1131.97)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,184.614)(64,369.228)(96,553.842)(128,738.456)(160,923.07)(192,1107.68)(224,1292.3)(256,1476.91)};

\legend{$p = 2$, $p=3$, $p=4$, $p = 5$}
\end{axis}
\end{tikzpicture}
}
\subfloat[Smoothed knots]{
\begin{tikzpicture}
\begin{axis}[
width=.45\textwidth,
    xlabel={Number of elements $K$},   
%    ylabel={$\rho\LRp{\bm{A}_h}$},   
    xmin=0, xmax=300,
    ymax=0,ymax=1600,    
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
]

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,55.5039)(64,110.088)(96,164.669)(128,219.248)(160,273.827)(192,328.405)(224,382.984)(256,437.563)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,75.8822)(64,149.237)(96,222.58)(128,295.92)(160,369.26)(192,442.599)(224,515.938)(256,589.276)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,97.8724)(64,190.915)(96,283.926)(128,376.931)(160,469.932)(192,562.932)(224,655.931)(256,748.929)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,121.298)(64,234.744)(96,348.126)(128,461.491)(160,574.85)(192,688.206)(224,801.56)(256,914.912)};

\legend{$p = 2$, $p=3$, $p=4$, $p = 5$}
\end{axis}
\end{tikzpicture}
}
%\caption{Growth of $\rho\LRp{\bm{A}_h}$ for advection (partial upwinding) and $p = 2, 3, 4, 5$. }
%\label{fig:rhoA}
\end{figure}
}
\end{overlayarea}
}

%\section{Numerical experiments}
%\frame[noframenumbering]{
%\frametitle{Outline}
%\tableofcontents[currentsection]
%}
%

\frame{
\frametitle{A 3D multi-patch example}
\setcounter{subfigure}{0}
\vspace{-1em}
\begin{figure}
\centering
\subfloat{
\includegraphics[width=.285\textwidth]{figs/pipe_crop.png}
}
\subfloat{
\includegraphics[width=.285\textwidth]{figs/pipeDG_crop.png}
}
\subfloat{
\includegraphics[width=.375\textwidth]{figs/pipeDG_cut.png}
}
%\caption{Multi-patch 3D pipe, .}
\vspace{1em}
\end{figure}
\begin{itemize}
\item 12 patch pipe model, first order formulation, pulse inflow condition.  
\vspace{.5em}
\item Isotropic $p=6, K=16$ splines, smoothed knots on each patch.
\end{itemize}
\let\thefootnote\relax\footnotetext{\tiny Chan, Evans.\ 2017.  {Multi-patch discontinuous Galerkin spline FEM for time-domain wave propagation} (in preparation).}
}

\frame{
\frametitle{Summary and acknowledgements}

\begin{itemize}
\item Optimal/smoothed knot vectors for time-domain simulations.  
%\vspace{1em}
\item Restore Kronecker product with weight-adjusted mass matrix.  
\vspace{1em}
\item Future directions:
\vspace{.5em}
\begin{itemize}
\item Conforming spaces for first order formulations.
\vspace{.5em}
\item Nonlinear systems (compressible flow).
\end{itemize}
\end{itemize}
\vspace{1em}
\begin{center}
This research is supported by DMS-1712639 and TOTAL E\&P Research and Technology USA.
\end{center}

\let\thefootnote\relax\footnotetext{\tiny Chan, Evans.\ 2017.  {Multi-patch discontinuous Galerkin spline FEM for time-domain wave propagation} (in preparation).}
\let\thefootnote\relax\footnotetext{\tiny Chan, et al.\ 2016.  {Weight-adjusted DG methods: curvilinear meshes} (arXiv).  }
%\let\thefootnote\relax\footnotetext{\tiny Chan 2017.  Weight-adjusted DG methods: matrix-valued weights and elastic wave prop.\ in heterogeneous media (arXiv).}
%\let\thefootnote\relax\footnotetext{\tiny Chan, Warburton 2015. {GPU-accelerated Bernstein-Bezier DG methods for wave problems} (SISC).}
}

%\frame{
%\begin{center}
%The authors thank TOTAL E\&P Research and Technology USA\\
%for their generous support of this work.
%\end{center}
%\let\thefootnote\relax\footnotetext{\tiny Chan, Warburton 2015. {GPU-accelerated Bernstein-Bezier DG methods for wave problems}.}
%\let\thefootnote\relax\footnotetext{\tiny Chan, Warburton 2015.  A short note on a Bernstein-Bezier basis for the pyramid.  (SISC, accepted)}
%\let\thefootnote\relax\footnotetext{\tiny Chan, et al.\ 2016.  {WADG methods I: wave propagation in heterogeneous media}.  In preparation.}
%\let\thefootnote\relax\footnotetext{\tiny Chan, et al.\ 2016.  {WADG methods II: curvilinear meshes}.  In preparation.}
%}
%% =================== extra slides =======================

\begin{frame}[noframenumbering]
\frametitle{Additional slides }
\end{frame}

\frame[noframenumbering]{
\frametitle{Patch refinement vs knot insertion (uniform knots)}
\setcounter{subfigure}{0}
\begin{itemize}
\item Patch size $H$, number of sub-elements $K$: $h = H/K$.
\vspace{.125em}
\item Optimal $O(h^{p+1})$ $L^2$ error for both patch refinement, knot insertion.
%\item Time integration: low storage 5-stage 4th order Runge-Kutta. 
\end{itemize}
\vspace{-1em}
\begin{figure}
\centering
\subfloat[First order (uniform knots)]{
\begin{tikzpicture}
\begin{loglogaxis}[
width=.5\textwidth,
    xlabel={Mesh size },
    ylabel={$L^2$ error},
    xmin=1e-2, xmax=.3,
    ymin=1e-10, ymax=.025,    
    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
    legend entries={$p = 2$, $p=3$, $p=4$},
] 
\pgfplotsset{
cycle list={
{blue, mark=*}, {red,mark=square*},
{blue, mark=*,dashed}, {red,mark=square*,dashed},
{blue, mark=*,dotted}, {red,mark=square*,dotted},
legend style={mark options={no markers}}
}}
\addlegendimage{no markers,black}
\addlegendimage{no markers,black,dashed}
\addlegendimage{no markers,black,dotted}

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.0106673)(0.125,0.000954815)(0.0625,0.000110851)(0.03125,1.37364e-05)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.00629266)(0.125,0.000476247)(0.0625,7.45318e-05)(0.03125,1.11051e-05)};
\logLogSlopeTriangleFlip{.5}{.125}{.65}{3}{}

    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).

%    4.0252
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.00183485)(0.125,8.58892e-05)(0.0625,4.97075e-06)(0.03125,3.0529e-07)};
%    3.8576
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.000679534)(0.125,3.39478e-05)(0.0625,2.43527e-06)(0.03125,1.67989e-07)};
\logLogSlopeTriangleFlip{.5}{.125}{.475}{4}{}

%    5.0305
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.000412175)(0.125,1.05591e-05)(0.0625,2.98645e-07)(0.03125,9.13748e-09)};
%    4.8741
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,7.04497e-05)(0.125,1.74612e-06)(0.0625,6.0791e-08)(0.03125,2.07289e-09)};
\logLogSlopeTriangle{.5}{.15}{.125}{5}{}

\end{loglogaxis}
\end{tikzpicture}
}
\subfloat[Second order (uniform knots)]{
\begin{tikzpicture}
\begin{loglogaxis}[
width=.5\textwidth,
    xlabel={Mesh size}, 
%    ylabel={$L^2$ error},
    xmin=1e-2, xmax=.3,
    ymin=1e-10, ymax=.025,    
    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
%    legend entries={$p = 2$, $p=3$, $p=4$},
legend entries={Knot insert, Patch refine}
] 
\pgfplotsset{
cycle list={
{blue, mark=*}, {red,mark=square*},
{blue, mark=*,dashed}, {red,mark=square*,dashed},
{blue, mark=*,dotted}, {red,mark=square*,dotted},
legend style={mark options={no markers}}
}}
\addlegendimage{mark=*,black,mark options={solid,fill=markercolor}}
\addlegendimage{mark=square*,black,mark options={solid,fill=markercolor}}

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.0150916)(0.125,0.00135624)(0.0625,0.000126749)(0.03125,1.27992e-05)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.00727599)(0.125,0.000820116)(0.0625,9.52161e-05)(0.03125,1.14411e-05)};
\logLogSlopeTriangleFlip{.5}{.125}{.65}{3}{}

    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.00274207)(0.125,8.37621e-05)(0.0625,4.79288e-06)(0.03125,2.97136e-07)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.001341)(0.125,8.99204e-05)(0.0625,4.17057e-06)(0.03125,3.6443e-07)};
\logLogSlopeTriangleFlip{.5}{.125}{.475}{4}{}

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.000623337)(0.125,9.84177e-06)(0.0625,2.93926e-07)(0.03125,9.019e-09)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.000128353)(0.125,3.87921e-06)(0.0625,6.10617e-08)(0.03125,3.57278e-09)};
\logLogSlopeTriangleFlip{.5}{.125}{.285}{5}{}

\end{loglogaxis}
\end{tikzpicture}
}
%\caption{Knot insertion vs patch refinement (uniform knots).  Results for knot insertion denoted by circles, results for patch refinement denoted by squares.}
\label{fig:hconvergence1D}
\end{figure}
}

\frame[noframenumbering]{
\frametitle{Convergence of WADG on curvilinear meshes}

\begin{figure}
\centering
\subfloat{\includegraphics[width=.375\textwidth]{figs/curved_mesh.png}}
\hspace{.5em}
\setcounter{subfigure}{0}
\subfloat[$L^2$ errors for $N = 2,3,4$]{

\begin{tikzpicture}
\begin{loglogaxis}[
	legend cell align=left,
	width=.5\textwidth,
%    title={Convergence},
    xlabel={Mesh size $h$},
    ylabel={$L^2$ error},
    xmin=.01, xmax=10,
    ymin=1e-9, ymax=1,        
    legend pos=south east,
    xmajorgrids=true,
    ymajorgrids=true,
    grid style=dashed,
] 
\addplot+[color=blue,mark=*,mark options={fill=markercolor},semithick]
coordinates{(0.5,0.026045)(0.25,0.00661508)(0.125,0.000722446)(0.0625,9.5217e-05)};
\addplot+[color=blue,mark=x,mark options={solid,fill=markercolor},semithick]
coordinates{(0.5,0.0260431)(0.25,0.00661506)(0.125,0.000722445)(0.0625,9.5217e-05)};

\addplot+[color=red,mark=square*,mark options={fill=markercolor},semithick]
coordinates{(0.5,0.0101239)(0.25,0.000387682)(0.125,2.88879e-05)(0.0625,1.78875e-06)};
\addplot+[color=red,mark=x,mark options={solid,fill=markercolor},semithick]
coordinates{(0.5,0.0101238)(0.25,0.000387682)(0.125,2.88879e-05)(0.0625,1.78875e-06)};

\addplot+[color=black,mark=triangle*,mark options={fill=markercolor},semithick]
coordinates{(0.5,0.000483144)(0.25,4.84164e-05)(0.125,1.38952e-06)(0.0625,4.71587e-08)};
\addplot+[color=black,mark=x,mark options={solid,fill=markercolor},semithick]
coordinates{(0.5,0.000483096)(0.25,4.84162e-05)(0.125,1.38952e-06)(0.0625,4.71587e-08)};
\legend{DG,WADG}
\end{loglogaxis}
\end{tikzpicture}
}
\caption{Optimal $L^2$ convergence rates observed for curvilinear meshes.  }
\end{figure}
}

\frame[noframenumbering]{
\frametitle{Curvilinear meshes: DG eigenvalues (circular domain)}
\vspace{-1.5em}
\begin{figure}
\centering
\subfloat[Central fluxes]{
\includegraphics[width=.38\textwidth]{figs/eigstau0.png}
}
\subfloat[${\rm Im}\LRp{\lambda_i}$ for central fluxes]{
\includegraphics[width=.38\textwidth]{figs/eigstau0imag.png}
}\\
\vspace{-.75em}
\subfloat[Upwind fluxes]{
\includegraphics[width=.38\textwidth]{figs/eigstau1.png}
}
%\caption{Spectra of discretized wave equation using DG and WADG for $N=3$.  Both central ($\tau = 0$) and dissipative ($\tau = 1$) numerical fluxes are shown.}
\end{figure}
%\begin{itemize}
%\item Energy stability requires quadrature, skew-symmetric formulation
%\begin{align*}
%\int_{D^k}\pd{p}{t}{} q &= \int_{D^k} -\bm{u}\cdot \Grad q  + \frac{1}{2}\int_{\partial D^k} \LRp{2\avg{\bm{u}}\cdot{\bm{n}} + \tau_p\jump{p}} q\\
%\int_{D^k}\pd{\bm{u}}{t}{} \bm{v} &= \int_{D^k} \Grad p \cdot \bm{v} + \frac{1}{2}\int_{\partial D^k} \LRp{\jump{p} + \tau_u\jump{\bm{u}}\cdot\bm{n}} \bm{v}
%\end{align*}
%\end{itemize}
}

\frame[noframenumbering]{
\frametitle{Behavior of weight-adjusted $L^2$ projection}

\only<1-3>{
Comparison with $L^2$ projection and Low-Storage Curvilinear DG
\[
\tilde{\phi}_i = \frac{\phi_i}{\sqrt{J}}, \qquad \bm{M}_{ij} = \int_{D^k} \tilde{\phi}_j\tilde{\phi}_i J = \int_{\widehat{D}} \phi_j\phi_i = \widehat{\bm{M}}_{ij}.
\]
\vspace{-2em}
}
\begin{figure}
\begin{overlayarea}{\textwidth}{.6\textheight}
\centering
\only<1>{
\subfloat{
\includegraphics[width=.37\textwidth]{figs/arnoldmesh1.png}
}
\subfloat{
\begin{tikzpicture}
\begin{loglogaxis}[
	legend cell align=left,
	        legend style={font=\tiny},	
	width=.49\textwidth,
    xlabel={Mesh size $h$},
    ylabel={$L^2$ error},
    xmin=.005, xmax=1,
    ymin=1e-10, ymax=1e-1,
    legend pos=south east,
    xmajorgrids=true,
    ymajorgrids=true,
    grid style=dashed,
] 
% adding this 2x because pgfplots dashes the line for some reason...
\addplot[color=blue,mark=*,mark size=3,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.00145465)(0.25,0.000140693)(0.125,1.17949e-05)(0.0625,8.58138e-07)(0.03125,5.78543e-08)(0.015625,3.75483e-09)};
\addplot[color=black,mark=diamond*,mark size=4,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.00159198)(0.25,0.000658818)(0.125,0.000619747)(0.0625,0.000677173)(0.03125,0.00073964)(0.015625,0.000782687)};
\addplot[color=red,mark=square*,mark size=3,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.00148852)(0.25,0.000176325)(0.125,2.60169e-05)(0.0625,3.94602e-06)(0.03125,5.58752e-07)(0.015625,7.47639e-08)};
\logLogSlopeTriangleFlip{0.325}{0.125}{0.37}{3}{red};
\logLogSlopeTriangle{0.35}{0.125}{0.12}{4}{blue};

\legend{$L^2$ projection, LSC-DG, WADG}
\end{loglogaxis}
\end{tikzpicture}
}
\caption{Arnold-type mesh with $\nor{J}_{W^{N+1,\infty}} = O(h^{-1})$ for $N = 3$.}
}
\only<2>{
\subfloat{
\includegraphics[width=.37\textwidth]{figs/arnoldmesh2.png}
}
\subfloat{
\begin{tikzpicture}
\begin{loglogaxis}[
	legend cell align=left,
	        legend style={font=\tiny},	
	width=.49\textwidth,
    xlabel={Mesh size $h$},
    ylabel={$L^2$ error},
    xmin=.005, xmax=1,
    ymin=1e-10, ymax=1e-1,
    legend pos=south east,
    xmajorgrids=true,
    ymajorgrids=true,
    grid style=dashed,
] 
% adding this 2x because pgfplots dashes the line for some reason...
\addplot[color=blue,mark=*,mark size=3,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.00145465)(0.25,0.000140693)(0.125,1.17949e-05)(0.0625,8.58138e-07)(0.03125,5.78543e-08)(0.015625,3.75483e-09)};
\addplot[color=black,mark=diamond*,mark size=4,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.00159198)(0.25,0.000658818)(0.125,0.000619747)(0.0625,0.000677173)(0.03125,0.00073964)(0.015625,0.000782687)};
\addplot[color=red,mark=square*,mark size=3,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.00148852)(0.25,0.000176325)(0.125,2.60169e-05)(0.0625,3.94602e-06)(0.03125,5.58752e-07)(0.015625,7.47639e-08)};
\logLogSlopeTriangleFlip{0.325}{0.125}{0.37}{3}{red};
\logLogSlopeTriangle{0.35}{0.125}{0.12}{4}{blue};

\legend{$L^2$ projection, LSC-DG, WADG}
\end{loglogaxis}
\end{tikzpicture}
}
\caption{Arnold-type mesh with $\nor{J}_{W^{N+1,\infty}} = O(h^{-1})$ for $N = 3$.}
}

\only<3>{
\subfloat{
\includegraphics[width=.37\textwidth]{figs/randunif1.png}
}
\subfloat{
\begin{tikzpicture}
\begin{loglogaxis}[
	legend cell align=left,
	        legend style={font=\tiny},
	width=.49\textwidth,
    xlabel={Mesh size $h$},
    ylabel={$L^2$ error},
    xmin=.005, xmax=1,
    ymin=1e-10, ymax=1e-1,
    legend pos=south east,
    xmajorgrids=true,
    ymajorgrids=true,
    grid style=dashed,
] 
\addplot[color=blue,mark=*,mark size=3,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.00943375)(0.25,0.00276313)(0.125,0.000415615)(0.0625,6.56974e-05)(0.03125,8.95162e-06)(0.015625,1.14109e-06)};
\addplot[color=black,mark=diamond*,mark size=4,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.0134065)(0.25,0.012209)(0.125,0.004612)(0.0625,0.00163555)(0.03125,0.000458367)(0.015625,0.00011822)};
\addplot[color=red,mark=square*,mark size=3,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.00990795)(0.25,0.00389737)(0.125,0.00064227)(0.0625,0.000123078)(0.03125,1.80455e-05)(0.015625,2.28866e-06)};
\logLogSlopeTriangleFlip{0.325}{0.125}{0.525}{3}{red};
\logLogSlopeTriangle{0.325}{0.125}{0.375}{3}{blue};

\legend{$L^2$ projection, LSC-DG, WADG}
\end{loglogaxis}
\end{tikzpicture}
}
\caption{Curvilinear mesh constructed through random perturbation for $N = 3$.}
}
\only<4>{
\vspace{-2em}
High order convergence \textcolor{red}{slowed} by growth of $\nor{J}_{W^{N+1,\infty}} = O(h^N)$.
\vspace{1em}

\subfloat{
\includegraphics[width=.37\textwidth]{figs/curvarnold2ref.png}
}
\subfloat{
\begin{tikzpicture}
\begin{loglogaxis}[
legend style={font=\tiny},
	legend cell align=left,
	width=.49\textwidth,
    xlabel={Mesh size $h$},
    ylabel={$L^2$ error},
    xmin=.0025, xmax=1,
    ymin=1e-7, ymax=1e-0,
    legend pos=south east,
    xmajorgrids=true,
    ymajorgrids=true,
    grid style=dashed,
] 
% adding this 2x because pgfplots dashes the line for some reason...
\addplot[color=blue,mark=*,mark size=3,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.0149174)(0.25,0.00576658)(0.125,0.00165616)(0.0625,0.000433839)(0.03125,0.000110389)(0.015625,2.78024e-05)};
%\addplot[color=black,mark=diamond*,mark size=4,semithick, mark options={fill=markercolor}]
%coordinates{(0.5,0.018737)(0.25,0.0147278)(0.125,0.0117761)(0.0625,0.0110991)(0.03125,0.0112431)(0.015625,0.0114663)};
\addplot[color=red,mark=square*,mark size=3,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.0165783)(0.25,0.00872368)(0.125,0.00464969)(0.0625,0.00253621)(0.03125,0.00134518)(0.015625,0.000695124)};
\logLogSlopeTriangleFlip{0.41}{0.15}{0.6}{1}{red};
\logLogSlopeTriangle{0.41}{0.15}{0.25}{2}{blue};

%\node at (axis cs:.03,2.1e-07) {$a = 10^{-1}$};

%\legend{$L^2$ proj, LSC-DG, WADG}
\legend{$L^2$ proj, WADG}
\end{loglogaxis}
\end{tikzpicture}
}
\caption{Moderately warped curved Arnold-type mesh for $N = 3$.}
}
\only<5>{
\vspace{-2em}
High order convergence is \textcolor{red}{stalled} by growth of $\nor{J}_{W^{N+1,\infty}} = O(h^{N+1})$.  
\vspace{1em}

\subfloat{
\includegraphics[width=.37\textwidth]{figs/curvarnold1ref.png}
}
\subfloat{
\begin{tikzpicture}
\begin{loglogaxis}[
legend style={font=\tiny},
	legend cell align=left,
	width=.49\textwidth,
    xlabel={Mesh size $h$},
    ylabel={$L^2$ error},
    xmin=.0025, xmax=1,
    ymin=1e-7, ymax=1e-0,
    legend pos=south east,
    xmajorgrids=true,
    ymajorgrids=true,
    grid style=dashed,
] 
\addplot[color=blue,mark=*,mark size=3,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.038232)(0.25,0.0144542)(0.125,0.00400233)(0.0625,0.000991275)(0.03125,0.000234827)(0.015625,5.4896e-05)};
%\addplot[color=black,mark=diamond*,mark size=4,semithick, mark options={fill=markercolor}]
%coordinates{(0.5,0.0579895)(0.25,0.052133)(0.125,0.0462352)(0.0625,0.0434683)(0.03125,0.0394193)(0.015625,0.032696)};
\addplot[color=red,mark=square*,mark size=3,semithick, mark options={fill=markercolor}]
coordinates{(0.5,0.0681969)(0.25,0.0751338)(0.125,0.0744307)(0.0625,0.0695447)(0.03125,0.0656228)(0.015625,0.0653247)};
%\logLogSlopeTriangleFlip{0.3}{0.125}{0.475}{1}{red};
\logLogSlopeTriangleFlip{0.425}{0.15}{0.45}{2}{blue};

%\node at (axis cs:.03,2.1e-07) {$a = 10^{-1}$};

%\legend{$L^2$ proj, LSC-DG, WADG}
\legend{$L^2$ proj, WADG}
\end{loglogaxis}
\end{tikzpicture}
}
\caption{Heavily warped curved Arnold-type mesh for $N = 3$.}
}
\end{overlayarea}

\end{figure}
}

\frame[noframenumbering]{
\frametitle{Weight-adjusted DG: not locally conservative}

\begin{itemize}
\item \textcolor{red}{Con:} loss of local conservation for $w(x) \not\in P^N$!
\vspace{1em}
\item \textcolor{blue}{Pro:} superconvergence of conservation error
\begin{align*}
%&\LRb{\int_{\widehat{D}} w p(\bm{x},t) - \int_{\widehat{D}} T_{1/w}^{-1} p(\bm{x},t)} \\
\text{Conservation error}
&\leq C \textcolor{red}{h^{2N+2}} \nor{w}_{W^{N+1,\infty}}\nor{p}_{W^{N+1,2}}
\end{align*}
where $C$ depends on mesh quality and max/min values of $w$.
\vspace{1em}
\item \textcolor{blue}{Pro:} can restore local conservation with rank-1 update (Shermann-Morrison).
\end{itemize}
}

\frame[noframenumbering]{
\frametitle{Effect of conservation on shock speeds}

\begin{itemize}
\item Weighted Burgers' equation, $w(x)$ curves characteristic lines.
\[
w(x)\pd{u}{t}{} + \frac{1}{2}\pd{u^2}{x}{} = 0.
\]
\item WADG yields high order convergence, correct shock speed for both $w(x)$ smooth, discontinuous (within an element).
\end{itemize}
%\vspace{-1em}
\begin{overlayarea}{\textwidth}{.55\textheight}
\only<1>{
\begin{figure}
\centering
\subfloat[Smooth solution]{\includegraphics[width=.4\textwidth]{figs/burgersSmooth2.png}}
\hspace{1em}
\subfloat[Shock solution]{\includegraphics[width=.39\textwidth]{figs/burgersShock2.png}}
\end{figure}
}
\only<2>{
\vspace{2em}
\begin{center}
Best guess: {where} and {what} is locally conserved matters;\\ non-conservation of \textit{nonlinear flux} results in incorrect shock speeds.  
\end{center}
}
\end{overlayarea}
}

\bibliographystyle{plain}
{\scriptsize
\bibliography{pyramids}
}

\end{document}

### 2017-08-23 docs/splitform/main.tex
updating old routines

@@
 
-%\documentclass[preprint,10pt]{elsarticle}
-\documentclass[preprint,10pt]{article}
+\documentclass[preprint,10pt]{elsarticle}
+%\documentclass[preprint,10pt]{article}
 %\documentclass[review]{siamart0216}
@@
 \usepackage{fullpage}
+\usepackage[colorlinks=true]{hyperref}
+
 \usepackage{amsmath,amssymb,amsfonts,amsthm}
@@
 \newtheorem{assumption}{Assumption}
-%\usepackage{thmtools}
-%\declaretheorem[style=definition,qed=$\blacksquare$,numberwithin=chapter]{definition}
 
@@
 
-%\usepackage{tikz}
+\usepackage{tikz}
 \usepackage[normalem]{ulem}
 \usepackage{hhline}
-
-%% ====================================== alg package
-\usepackage{algorithm}
-\usepackage[noend]{algpseudocode}
-\usepackage{algorithmicx}
-\algblock{ParFor}{EndParFor}
-% customising the new block
-\algnewcommand\algorithmicparfor{\textbf{parfor}}
-\algnewcommand\algorithmicpardo{\textbf{do}}
-\algnewcommand\algorithmicendparfor{\textbf{end\ parfor}}
-\algrenewtext{ParFor}[1]{\algorithmicparfor\ #1\ \algorithmicpardo}
-\algrenewtext{EndParFor}{\algorithmicendparfor}
-%% ====================================== end alg package
 
@@
 \definecolor{markercolor}{RGB}{124.9, 255, 160.65}
-\pgfplotsset{width=10cm,compat=1.3}
 \pgfplotsset{
+compat=1.3,
+width=10cm,
 tick label style={font=\small},
@@
 \usetikzlibrary{calc}
+\usetikzlibrary{intersections}
 
@@
 
-\date{}
-\author{Jesse Chan}
-\title{On discretely entropy conservative discontinuous Galerkin methods}
+%\date{}
+%\author{Jesse Chan}
+%\title{On discretely entropy conservative discontinuous Galerkin methods}
+\graphicspath{{./figs/}}
 
@@
 
-\maketitle
+%\maketitle
+
+\begin{frontmatter}
+\title{On discretely entropy conservative and entropy stable discontinuous Galerkin methods}
+
+\author[rice]{Jesse Chan\corref{cor1}}
+\ead{Jesse.Chan@caam.rice.edu}
+\address[rice]{Department of Computational and Applied Mathematics, Rice University, 6100 Main St, Houston, TX, 77005}
 
 \begin{abstract}
-High order methods based on diagonal-norm summation by parts operators can be shown to semi-discretely conserve entropy for nonlinear systems of hyperbolic PDEs \cite{fisher2013high,carpenter2014entropy}.
-These methods can be interpreted as mass lumped discontinuous Galerkin spectral element methods (DG-SEM) \cite{gassner2016split,gassner2016well,wintermeyer2017entropy, chen2017entropy}.
-In this work, we describe how to extend discretely entropy conservative schemes for a more general class of DG methods using quadrature-based $L^2$ projections, flux differencing, and specific DG differentiation operators.
-This approach also recovers existing methods for Burgers' equation involving generalized SBP operators without boundary nodes and dense norm SBP operators.
-Numerical experiments confirm the stability and high order accuracy of the proposed methods for the one-dimensional shallow water and compressible Euler equations.
+High order methods based on diagonal-norm summation by parts operators can be shown to satisfy a discrete conservation or dissipation of entropy for nonlinear systems of hyperbolic PDEs \cite{fisher2013high, carpenter2014entropy}.
+These methods can also be interpreted as nodal discontinuous Galerkin methods with diagonal mass matrices \cite{gassner2016split, gassner2016well, wintermeyer2017entropy, chen2017entropy}.
+In this work, we describe how to construct discretely entropy conservative schemes to a more general class of DG methods using flux differencing, quadrature-based projections, and specific DG differentiation operators.
+This approach also recovers existing methods for Burgers' equation involving dense norm and generalized SBP operators without boundary nodes \cite{ranocha2016summation, ranocha2017extended}.
+Numerical experiments confirm the stability and high order accuracy of the proposed methods for the one-dimensional compressible Euler equations.
 \end{abstract}
+\end{frontmatter}
 
@@
 Numerical simulations in engineering increasingly require higher accuracy without sacrificing computational efficiency.
-High order methods are more accurate than low order methods per degree of freedom for sufficiently regular solutions, and provide one avenue towards improving fidelity in numerical simulations while maintaining reasonable computational costs.
-High order methods which can accomodate unstructured meshes are particularly well-suited for problems with complex geometries.
-Among such methods, high order discontinuous Galerkin (DG) methods are particularly well-suited to the solution of time-dependent hyperbolic problems on modern computing architectures \cite{hesthaven2007nodal, klockner2009nodal}.
-
-The accuracy observed for high order methods can be attributed in part to their low numerical dissipation and dispersion properties \cite{ainsworth2004dispersive}.
-While low numerical dissipation is advantageous for problems in linear wave propagation \cite{chan2015gpu}, the high numerical dissipation present in low order schemes adds a stabilizing effect for nonlinear hyperbolic problems \cite{wang2013high}.
-As a result, high order schemes for nonlinear conservation laws tend to require additional stabilization, including (but not limited to) filtering \cite{hesthaven2007nodal}, slope limiting \cite{krivodonova2007limiters}, artificial viscosity \cite{persson2006sub}, and polynomial de-aliasing \cite{ karniadakis2013spectral, kirby2003aliasing}.
+Because they are more accurate than low order methods per degree of freedom for sufficiently regular solutions, high order methods provide one avenue towards improving fidelity in numerical simulations while maintaining reasonable computational costs.
+High order methods which can accomodate unstructured meshes are desirable for problems with complex geometries, and among such methods, high order discontinuous Galerkin (DG) methods are particularly well-suited to the solution of time-dependent hyperbolic problems on modern computing architectures \cite{hesthaven2007nodal, klockner2009nodal}.
+
+The accuracy of high order methods can be attributed in part to their low numerical dissipation and dispersion compared to low order schemes \cite{ainsworth2004dispersive}.
+This accuracy has made them advantageous for the simulation of wave propagation \cite{hesthaven2007nodal, wilcox2010high}.
+However, while high order methods can be applied in a stable manner to linear wave propagation problems, instabilities are observed when applying them to nonlinear hyperbolic problems.
+This is contrast to low order schemes, whose high numerical dissipation tends to apply a stabilizing effect \cite{wang2013high}.
+As a result, most high order schemes for nonlinear conservation laws typically require additional stabilization procedures, including  filtering \cite{hesthaven2007nodal}, slope limiting \cite{krivodonova2007limiters}, artificial viscosity \cite{persson2006sub}, and polynomial de-aliasing through over-integration \cite{kirby2003aliasing}.
 Moreover, stabilized numerical methods can still fail, requiring user intervention or heuristic modifications to achieve non-divergent solutions.
 
-For linear wave propagation problems, numerical methods can be shown to be semi-discretely energy stable, even in the presence of curvilinear coordinates or variable coefficients \cite{warburton2013low, chan2016weight1, chan2016weight2, chan2017weight}.
-This semi-discrete stability can be used to show that, under a stable timestep restriction (CFL), discrete solutions obtained by explicit time-stepping do not suffer from un-physical growth in time.
-For nonlinear systems of conservation laws, typical methods for fluid dynamics do not admit formal proofs of stability.
-However, progress has been made in designing provably discretely stable methods, beginning with the introduction of entropy conservative and entropy stable finite volume schemes by Tadmor \cite{tadmor1987numerical, tadmor2003entropy}.
-These schemes rely on a specific entropy conservative flux which satisfies a condition involving the entropy variables and entropy potential, and have been extended to low order finite volume methods on unstructured grids \cite{ray2016entropy}.
-
-The construction of entropy conservative high order schemes was done much more recently for the compressible Euler and Navier-Stokes equations in \cite{fisher2013high, carpenter2014entropy} based on a spectral collocation approach on tensor product elements, which can be interpreted as a mass-lumped DG spectral element (DG-SEM) scheme.
-The proof of entropy conservation relies on the presence of a diagonal mass matrix, as well as the summation by parts property \cite{gassner2013skew}.
-Similar entropy-stable schemes were constructed for the shallow water and MHD equations \cite{gassner2016well, wintermeyer2017entropy,winters2017uniquely}.
-More recently, entropy conservative and entropy stable schemes were extended to unstructured triangular meshes in \cite{chen2017entropy} using the split-form methodology of \cite{gassner2016split} and multi-dimensional SBP operators \cite{hicken2016multidimensional}.
-
-To the authors knowledge, the construction of entropy conservative and entropy stable schemes have relied on the use of lumped diagonal mass matrices and nodal collocation.
-Since these can be reinterpreted as diagonal norm SBP operators, we refer to DG methods with these properties as diagonal norm SBP-DG methods.
-Entropy stable high order finite element and DG methods which do not fall under the diagonal norm SBP-DG category have been proposed \cite{hughes1986new}, but the proofs are often given at the continuous level, relying on exact integration or the chain rule, neither of which holds at the discrete level.
-
-On the other hand, while entropy conservative schemes can be constructed by combining flux differencing with diagonal norm SBP operators, this can restrict one to specific choices of basis and quadrature.
-Appropriate diagonal norm SBP operators can be constructed on tensor product elements based on a DG-SEM discretization.
-Diagonal-norm SBP operators can also be constructed for triangles and tetrahedra \cite{chin1999higher,  hicken2016multidimensional, chen2017entropy}; however, the number of nodal points for such operators is greater than the dimension of the natural polynomial approximation space.
-Furthermore, appropriate point sets have only been constructed for $N \leq 4$ in three dimensions \cite{zhebel2014comparison}, and (to the author's knowledge) the construction of diagonal norm SBP-DG methods has not yet been performed for less standard elements such as pyramids \cite{chan2016orthogonal}.
-%Furthermore, diagonal-norm SBP operators do not naturally generalize to triangles or tetrahedra \cite{helenbrook2009existence}.
+For linear wave propagation problems, semi-discretely energy stable numerical methods can be constructed, even in the presence of curvilinear coordinates or variable coefficients \cite{warburton2013low, chan2016weight1, chan2016weight2, chan2017weight}.
+This semi-discrete stability implies that, under a stable timestep restriction (CFL condition), discrete solutions do not suffer from non-physical growth in time.
+However, for nonlinear systems of conservation laws, traditional methods do not admit theoretical proofs of semi-discrete stability.
+This was addressed for low order methods with the introduction of discretely entropy conservative and entropy stable finite volume schemes by Tadmor \cite{tadmor1987numerical}.
+These schemes rely on a specific entropy conservative flux which satisfies a condition involving the entropy variables and entropy potential, and were extended to low order finite volume methods on unstructured grids in \cite{ray2016entropy}.
+High order entropy stable methods were also developed for structured grids in \cite{fjordholm2012arbitrarily} based on an entropy conservative essentially non-oscillatory (ENO) reconstruction.
+
+The extension of entropy conservative schemes to unstructured high order methods was done much more recently for the compressible Euler and Navier-Stokes equations in \cite{fisher2013high, carpenter2014entropy} based on a spectral collocation approach on tensor product elements, which can also be interpreted as a mass-lumped DG spectral element (DG-SEM) scheme.
+The proof of entropy conservation relies on the presence of a diagonal mass matrix, the summation by parts (SBP) property \cite{gassner2013skew}, and  a concept referred to as flux differencing.
+Similar entropy-stable schemes have been constructed for the shallow water and MHD equations \cite{gassner2016well,  wintermeyer2017entropy, winters2017uniquely}.
+More recently, high order entropy conservative and entropy stable schemes were extended to unstructured triangular meshes in \cite{chen2017entropy} using the split-form methodology of \cite{gassner2016split} and multi-dimensional SBP operators \cite{hicken2016multidimensional}.
+
+To the authors knowledge, the construction of unstructured high order entropy conservative and entropy stable schemes  has required diagonal norm SBP operators.
+We refer to DG methods with these properties as diagonal norm SBP-DG methods.
+Entropy stable high order finite element and DG methods which do not fall under the diagonal norm SBP-DG category have been proposed \cite{hughes1986new}, but the proofs are often given at the continuous level, relying on exact integration or the chain rule, neither of which holds discretely.
+
+On the other hand, while entropy conservative schemes can be constructed by combining flux differencing with diagonal norm SBP operators, this restricts methods to specific choices of basis and quadrature.
+Appropriate diagonal norm SBP operators are straightforward to construct on tensor product elements based on a DG-SEM discretization.
+Diagonal-norm SBP operators can also be constructed for triangles and tetrahedra \cite{chin1999higher,  hicken2016multidimensional, chen2017entropy}; however, the number of nodal points for such operators is typically greater than the dimension of the natural polynomial approximation space.
+Furthermore, to the author's knowledge, appropriate point sets have only been constructed for $N \leq 4$ in three dimensions \cite{zhebel2014comparison}, and the construction of high order diagonal norm SBP-DG methods has not yet been performed for uncommon elements such as pyramids \cite{chan2016orthogonal}.
 
 This work focuses on the construction of entropy conservative high order DG schemes for systems of conservation laws.
-In order to generalize beyond diagonal norm SBP-DG methods, we will consider dense (non-lumped) mass matrices and more general quadrature rules (e.g.\ rules without boundary points, rules with more points than the dimension of the corresponding approximation space).
+In order to generalize beyond diagonal norm SBP-DG methods, we will consider non-lumped (dense) mass matrices and more general quadrature rules (e.g.\ rules without boundary points, rules with more points than the dimension of the corresponding approximation space), which are also related to dense norm and generalized SBP operators \cite{ranocha2016summation, ranocha2017extended, ranocha2017comparison}.
+The work presented in this manuscript is intended to serve as a stepping stone to the construction of entropy stable DG methods on general elements in two and three dimensions.
 We will present proofs of discrete entropy stability in terms of integrals and $L^2$ projections, which we assume are computed using an appropriate quadrature rule.
-We stress that, while we utilize a continuous notation involving integrals rather than discrete sums, the proofs rely only on properties of quadrature-based integration and $L^2$ projection.
+We stress that, while this work adopts a continuous notation involving integrals rather than discrete sums, the proofs rely only on properties of quadrature-based integration and quadrature-based $L^2$ projection.
 
 The outline of the paper is as follows: Section~\ref{sec:intro} will briefly review the construction of entropy conservative diagonal norm SBP-DG methods on a single element.
-Section~\ref{sec:ecdg} will describe how to construct analogous entropy conservative methods on single element.
+Section~\ref{sec:ecdg} will describe how to construct analogous entropy conservative methods on single element in a continuous setting.
 Section~\ref{sec:ecdg2} will discuss extensions to multiple elements, including comparisons of different coupling terms and entropy stable fluxes.
-We note that, for simplicity and clarity of presentation, the proposed methods and proofs are given in one dimension.
+Finally, Section~\ref{sec:num} presents numerical results which verify the high order accuracy and discrete entropy conservation of the proposed methods.
+We note that, for simplicity and clarity of presentation, the proposed methods, proofs, and numerical experiments are given in one dimension.
 However, the extension to multiple dimensions (including simplicial elements) is straightforward and will be described in an upcoming manuscript.
-Finally, Section~\ref{sec:num} presents numerical results which verify the high order accuracy and discrete entropy conservation of the proposed methods.
 
@@
 We will begin by reviewing continuous entropy theory and existing high order methods which are provably entropy conservative at the discrete level.
-These methods were introduced in \cite{fisher2013high, carpenter2014entropy} for tensor product elements, though we adopt methods of proof introduced more recently in \cite{gassner2017br1, chen2017entropy}.
-These methods all share a basis in diagonal norm summation by parts operators, and can be derived from a nodal discontinuous Galerkin framework through the collocation of nodal points and quadrature points.
+These methods were introduced in \cite{fisher2013high, carpenter2014entropy} for tensor product elements, though we adopt more general methods of proof introduced recently in \cite{gassner2017br1, chen2017entropy}.
+The proofs are based on diagonal norm SBP operators, which can be derived from a nodal discontinuous Galerkin framework through the collocation of nodal points and quadrature points \cite{gassner2013skew}.
 
@@
 \end{align}
-where the fluxes $f(\bm{u})$ are smooth functions of the vector of conservative variables $\bm{u}({x},t)$.%$\bm{u}(x,t)$.
-The Jacobian matrix $\bm{A}(\bm{u})$ is defined entrywise as
+where the fluxes $f(\bm{u})$ are smooth functions of the vector of conservative variables $\bm{u}({x},t)$.  The Jacobian matrix $\bm{A}(\bm{u})$ is defined entrywise as
 \[
 %\LRp{\bm{A}_k(\bm{u})}_{ij} = \LRp{\pd{\bm{f}_k(\bm{u})}{\bm{u}_j}}_i.
-\LRp{\bm{A}(\bm{u})}_{ij} = \LRp{\pd{\bm{f}(\bm{u})}{\bm{u}_j}}_i.
+\LRp{\bm{A}(\bm{u})}_{ij} = \LRp{\pd{\bm{f}(\bm{u})}{{u}_j}}_i.
 \]
@@
 
-It can be shown (see, for example, \cite{mock1980systems}) that (\ref{eq:entropysym}) is equivalent to the existence of an entropy flux functions $F(\bm{u})$ such that
+It can be shown (see, for example, \cite{mock1980systems}) that (\ref{eq:entropysym}) is equivalent to the existence of an entropy flux function $F(\bm{u})$ such that
 \[
@@
 \]
-The flux function and entropy flux function are further related through the so-called entropy potential $\psi_k$ 
+The flux function and entropy flux function are further related through the so-called entropy potential $\psi$ 
 \[
@@
 The focus of this work is the construction of methods which satisfy a discrete analogue of this entropy inequality.
-We begin by reviewing existing discretely entropy conservative methods based on diagonal norm summation by parts (SBP) operators.
+We begin by reviewing existing discretely entropy conservative methods based on diagonal norm SBP operators.
 
@@
 
-Let ${x}_i \in [-1,1]$ denote the $(N+1)$ point Gauss-Lobatto-Legendre (GLL) quadrature rule with positive weights $w_1,\ldots,w_{N+1}$.
+Let ${x}_i \in [-1,1]$ denote the $(N+1)$ point Gauss-Lobatto-Legendre (GLL) quadrature rule with positive weights $w_1,\ldots,w_{N+1}$, which integrates exactly polynomials of degree $(2N-1)$.
 Let $\ell_i(x)$ denote the nodal Lagrange basis defined at GLL points, and define the differentiation matrix $\bm{D}$ be defined as
@@
 \[
-\bm{M}_{ij} = \int_{-1}^1 \ell_i(x)\ell_j(x)\diff{x} \approx \delta_{ij} w_i,
-\]
-where the approximation results from evaluating the integral using GLL quadrature.
+\bm{M}_{ij} = \int_{-1}^1 \ell_i(x)\ell_j(x)\diff{x} \approx \sum_{k} \ell_i(x_k)\ell_j(x_k) w_k = \delta_{ij} w_i,
+\]
+where the equality holds only approximately due to the inexactness of GLL quadrature for polynomials of degree $2N$.
 For this choice of quadrature, the Kronecker property $\ell_j(x_i) = \delta_{ij}$ implies that $\bm{M}$ is a diagonal matrix whose entries are simply the quadrature weights.
@@
 \bm{S} = \bm{B} - \bm{S}^T, \qquad \bm{B}_{ij} = \begin{cases}
-1 &i = j = 1 \\
+-1 &i = j = 1 \\
 1 &i = j = N+1 \\
@@
 \]
-This property is simply a restatement of integration by parts, using that the derivative of $\ell_j(x) \in P^{N-1}$ and that GLL quadrature is exact for polynomials of degree $2N-1$.
+This property is simply a restatement of integration by parts, using that the derivative of $\ell_j(x)$ is a polynomial of degree $(N-1)$ and that GLL quadrature is exact for polynomials of degree $(2N-1)$.
 
@@
 \[
-\int_{-1}^1\LRp{ \pd{\bm{u}}{t} + \pd{\bm{f}(\bm{u})}{x}}\bm{w} + \left.\LRp{\bm{f}^* - \bm{f}(\bm{u})} \bm{w}\right|_{-1}^1 = 0, \qquad \forall \bm{w}\in V_h.
-\]
-Here, $\bm{f}^*$ is a numerical flux used for the weak imposition of boundary or continuity conditions.
-
-We derive a DG-SEM approximation by evaluating the above integrals using GLL quadrature.  We furthermore assume that the solution is approximated using a nodal basis at GLL points 
+\int_{-1}^1\LRp{ \pd{\bm{u}}{t} + \pd{\bm{f}(\bm{u})}{x}}\bm{w} + \left.\LRp{\bm{f}^* - \bm{f}(\bm{u})} \bm{w}\right|_{-1}^1 = 0, \qquad \forall \bm{w}\in V_h,
+\]
+where $\bm{w}(x)$ is a vector-valued test function.
+Here, $\bm{f}^*$ is some numerical flux used for the weak imposition of boundary or continuity conditions.
+
+We derive a DG-SEM approximation by evaluating the above integrals using GLL quadrature.  We furthermore assume that the vector solution is approximated using a nodal basis at GLL points 
 \[
@@
 \]
-The resulting semi-discrete system yields a nodal collocation scheme for $\bm{u}_i$ 
-\[
-\pd{\bm{u}_i}{t} + \sum_{j=1}^{N+1}\LRp{\LRp{\bm{D}\otimes \bm{I}}_{ij}\bm{f}(\bm{u}_j) +\LRp{\bm{M}^{-1}\bm{B}}_{ij}(\bm{f}_i^* - \bm{f}(\bm{u}_i))} = 0.
-\]
-
-This formulation conserves the quantities $\bm{u}_i$, but (except in specific cases) does not satisfy a discrete analogue of the statement of entropy conservation (\ref{eq:consentropy}).
-Entropy conservative schemes may be constructed based on the SBP property and a ``flux differencing'' approach \cite{gassner2016split} involving a definition of entropy conservative/stable fluxes given by Tadmor \cite{tadmor1987numerical, tadmor2003entropy}.
+The resulting semi-discrete system yields a nodal collocation scheme for each (vector-valued) degree of freedom $\bm{u}_i$ 
+\[
+%\pd{\bm{u}_i}{t} + \sum_{j=1}^{N+1}\LRp{\LRp{\bm{D}\otimes \bm{I}}_{ij}\bm{f}(\bm{u}_j) +\LRp{\LRp{\bm{M}^{-1}\bm{B}} \otimes \bm{I}}_{ij}(\bm{f}_i^* - \bm{f}(\bm{u}_i))} = 0, \qquad i = 1,\ldots,N+1.
+\pd{\bm{u}_i}{t} + \sum_{j=1}^{N+1}\LRp{\LRp{\bm{D}_{ij}\otimes \bm{I}}\bm{f}(\bm{u}_j) +\LRp{\LRp{\bm{M}^{-1}\bm{B}}_{ij}\otimes \bm{I} }(\bm{f}_j^* - \bm{f}(\bm{u}_j))} = 0, \qquad i = 1,\ldots,N+1,
+\]
+where $\bm{f}^*$ is the vector whose entries consist of evaluations of the numerical flux at boundary nodes and zeros at interior nodes.
+
+This formulation conserves the quantities $\bm{u}_i$, but does not satisfy a discrete analogue of the statement of entropy conservation (\ref{eq:consentropy}) except in specific cases.
+Entropy conservative schemes may be constructed based on the SBP property and a ``flux differencing'' approach \cite{gassner2016split} involving a definition of entropy conservative/stable fluxes given by Tadmor \cite{tadmor1987numerical}.
 \begin{definition}
-Let $\bm{f}_S(\bm{u}_L,\bm{u}_R)$ be a bivariate function such that it is symmetric and consistent with the flux function $\bm{f}(\bm{u})$
+Let $\bm{f}_S(\bm{u}_L,\bm{u}_R)$ be a bivariate function which is symmetric and consistent with the flux function $\bm{f}(\bm{u})$
 \[
@@
 \begin{equation}
-\pd{\bm{u}_i}{t} + \sum_{j=1}^{N+1}\LRp{\LRp{2\bm{D}\otimes \bm{I}}_{ij}\bm{f}_S(\bm{u}_i,\bm{u}_j) +\LRp{\bm{M}^{-1}\bm{B}}_{ij}(\bm{f}_i^* - \bm{f}(\bm{u}_i))}= 0.
+\pd{\bm{u}_i}{t} + \sum_{j=1}^{N+1}\LRp{\LRp{2\bm{D}_{ij}\otimes \bm{I}}\bm{f}_S(\bm{u}_i,\bm{u}_j) +\LRp{\LRp{\bm{M}^{-1}\bm{B}}_{ij}\otimes \bm{I}}(\bm{f}_j^* - \bm{f}(\bm{u}_j))}= 0.
 \label{eq:dgsem}
@@
 \]
-where $\bm{1}$ denotes the vector of ones and $\bm{\psi}_i = \psi(\bm{v}_i)$ is the entropy potential in terms of the nodal value of $\bm{u}_i$.  
+where $\bm{1}$ denotes the vector of ones and $\bm{\psi}_i = \psi(\bm{v}_i)$ is the entropy potential evaluated in terms of the nodal values $\bm{u}_i$.  
 \label{thm:dgsem}
@@
 \begin{proof}
-Let $\bm{v}_i = \bm{v}(\bm{u}_i)$ denote the entropy variables, defined in terms of the nodal values $\bm{u}_i$.  We multiply (\ref{eq:dgsem}) by $\LRp{\bm{M}\bm{v}}^T$.  
+In this proof, we assume the matrices $\bm{M}$, $\bm{D}$, $\bm{B}$ are understood to act component-wise on the vector of solution degrees of freedom $\bm{u}$ and drop the explicit Kronecker product for simplicity of notation.
+Let $\bm{v}_i = \bm{v}(\bm{u}_i)$ denote nodal values of the entropy variables.
+We multiply (\ref{eq:dgsem}) by $\LRp{\bm{M}\bm{v}}^T$.
 Using the fact that $\bm{M}$ is diagonal, the time derivative term can be manipulated to 
 \[
-\bm{v}^T\bm{M}\pd{\bm{u}}{t} = \sum_{i=1}^{N+1} \bm{M}_{ii} \bm{v}_i^T \pd{\bm{u}_i}{t} = \sum_{i=1}^{N+1} \bm{M}_{ii} \pd{S_i}{t} = \bm{1}^T\bm{M}\pd{S_i}{t},
+\bm{v}^T\bm{M}\pd{\bm{u}}{t} = \sum_{i=1}^{N+1} \bm{M}_{ii} \bm{v}_i^T \pd{\bm{u}_i}{t} = \sum_{i=1}^{N+1} \bm{M}_{ii} \pd{U(\bm{u}_i)}{t} = \bm{1}^T\bm{M}\pd{U(\bm{u}_i)}{t},
 \]
@@
 \[
-\sum_{i,j = 1}^{N+1} \bm{B}_{ij}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j) + \bm{v}_i^T\bm{B}_{ij}\LRp{\bm{f}_i^* - \bm{f}(\bm{u}_i)} = \bm{1}^T\bm{B} \LRp{\bm{v}^T \bm{f}^*},
-\]
-where we have used the consistency of $\bm{f}_S(\bm{u},\bm{u}) = \bm{f}(\bm{u})$.  Combining everything results in a discrete statement of the conservation of entropy (\ref{eq:consentropy}).   
+\sum_{i,j = 1}^{N+1} \bm{B}_{ij}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j) + \bm{v}_i^T\bm{B}_{ij}\LRp{\bm{f}_j^* - \bm{f}(\bm{u}_j)} = \bm{1}^T\bm{B} \LRp{\bm{v}^T \bm{f}^*},
+\]
+where we have used the consistency of $\bm{f}_S(\bm{u},\bm{u}) = \bm{f}(\bm{u})$.
+Putting these components together results in a discrete statement of the conservation of entropy (\ref{eq:consentropy}).
 %Combining everything results in a discrete statement of the conservation of entropy (\ref{eq:consentropy}) 
@@
 \end{proof}
-We note that, unlike the proof of continuous entropy conservation, the proof of discrete entropy conservation does not require the use of the chain rule, which may not hold at the discrete level.
-This proof of entropy conservation can also be extended beyond a single element by specifying appropriate simultaneous approximation terms (SBP-SAT).
-This can be done by taking $\bm{f}_i^*$ to be the entropy conservative flux $f_S(u_L,u_R)$ at the interface between two elements \cite{carpenter2014entropy, gassner2017br1, chen2017entropy}.
+We note that, unlike the continuous proof of entropy conservation, the proof of discrete entropy conservation avoids the use of the chain rule, which does not hold in general at the discrete level.
+This proof of entropy conservation can also be extended beyond a single element by by taking $\bm{f}^*$ to be the entropy conservative flux $f_S(u_L,u_R)$ at the interface between two elements \cite{carpenter2014entropy, gassner2017br1, chen2017entropy}.
+%specifying appropriate simultaneous approximation terms (SBP-SAT).  
 Similarly, employing an entropy stable flux at element interfaces results in an entropy stable method which satisfies a global entropy inequality.
 
-It is important to emphasize that conservation of entropy does not, in general, guarantee stability of the numerical scheme unless the solution satisfies additional constraints.
-For example, the convex entropy associated with the compressible Euler and Navier-Stokes equations is well-defined only if the density and pressure are positive, and steps must be taken in any numerical scheme to guarantee that the discrete solution of such constraints.
-For DG methods, this is most commonly done using bound and positivity-preserving limiters, such as those introduced in \cite{zhang2010positivity, zhang2012maximum}.
+It is important to emphasize that conservation of entropy does not, in general, imply stability of the numerical scheme unless the solution satisfies additional constraints.
+For example, the transformation between conservative and entropy variables for the compressible Euler and Navier-Stokes equations is well-defined only if the density and pressure are positive, and steps must be taken in any numerical scheme to guarantee that the discrete solution satisfies such constraints.
+For DG methods, this is most commonly done using bound and positivity-preserving limiters \cite{zhang2010positivity, zhang2012maximum}.
 
@@
 In this section, we focus on constructing entropy conservative DG schemes for general choices of basis and quadrature.
-These include (as special cases) dense norm SBP operators and generalized SBP (GSBP) operators which do not contain boundary points, which introduce additional difficulties in constructing entropy conservative DG schemes.
-These difficulties present themselves primarily in two areas:
-\begin{enumerate}
-\item the generalization of flux differencing to a continuous formulation.  
-\item the proof of entropy stability assumes a diagonal norm mass matrix, which commutes with point-wise multiplication by the (nodal values of) the entropy variables.  
-\end{enumerate}
-
-We address both areas of difficulty in generalizing the entropy conservative DG-SEM scheme (\ref{eq:dgsem}).
-In particular, we give a continuous DG formulation with flux differencing, and prove that the generalized scheme is entropy conservative (stable) on a single element in the following sections.
-As an example, we illustrate how to apply the proposed methods to produce energy conservative formulations of Burgers' equation.
+These include (as special cases) dense norm SBP operators and generalized SBP (GSBP) operators which do not contain boundary points.
+Difficulties in generalizing entropy conservative SBP-DG schemes present themselves primarily in two areas: the assumption of a diagonal mass matrix, which is used in the proof of entropy stability, and the generalization of flux differencing to a continuous formulation.
+We address both areas of difficulty by extending the entropy conservative DG-SEM scheme (\ref{eq:dgsem}) from a discrete formulation to a continuous formulation involving flux differencing.
+We prove that the continuous scheme is entropy conservative (stable) on a single element in the following sections, and describe how to apply the proposed methods to produce energy conservative formulations of Burgers' equation as an illustrative example.
 
@@
 
-We assume that the domain $\Omega$ is an interval, which is decomposed into non-overlapping intervals (elements) $D^k$ of size $h$.
+We assume that the domain $\Omega$ is a finite interval, which is decomposed into non-overlapping elements $D^k$ of size $h$.
 We define the reference element $\widehat{D} = [-1,1]$, such that $D^k$ is the image of $\widehat{D}$ under an affine mapping $x = \Phi^k(\widehat{x})$.
-We approximate the solution over the reference element using degree $N$ polynomials
+We define an approximation space using degree $N$ polynomials on the reference element
 \[
@@
 \]
+where we define $N_p = {\rm dim}\LRp{P^N\LRp{\widehat{D}}}$ (with $N_p = N+1$ in one dimension).  
 %We define an approximation space over the reference element $\widehat{D}$ as the space of degree $N$ polynomials 
@@
 \[
-P^N\LRp{D^k} =  \bm{\Phi}^k \circ P^N\LRp{\widehat{D}}, \qquad V_h = \bigoplus_{k}P^N\LRp{D^k}.
+P^N\LRp{D^k} =  {\Phi}^k \circ P^N\LRp{\widehat{D}}, \qquad V_h = \bigoplus_{k}P^N\LRp{D^k}.
 \]
@@
 \[
-\LRp{u,v}_{D^k} = \int_{D^k} u v\diff{x} =  \int_{D^k} u v J^k \diff{\widehat{x}}, \qquad \nor{u}^2_{D^k} = (u,u)_{D^k}, \qquad \LRa{u,v}_{\partial D^k} = \int_{\partial D^k} u v \diff{x},
+\LRp{\bm{u},\bm{v}}_{D^k} = \int_{D^k} \bm{u}\cdot\bm{v}\diff{x} =  \int_{\widehat{D}} \bm{u}\cdot\bm{v} J^k \diff{\widehat{x}}, \qquad \nor{\bm{u}}^2_{D^k} = (\bm{u},\bm{u})_{D^k}, \qquad \LRa{\bm{u},\bm{v}}_{\partial D^k} = \int_{\partial D^k} \bm{u} \cdot \bm{v} \diff{x},
 \]
@@
 \[
-\LRp{\Pi_N u,v}_{D^k} = \LRp{u,v}_{D^k}, \qquad \forall v\in V_h\LRp{D^k}.  
-\]
-In one dimension, $D^k$ is some interval and the $L^2$ inner product over the surface reduces to 
+\LRp{\Pi_N u,v}_{D^k} = \LRp{u,v}_{D^k}, \qquad \forall v\in P^N\LRp{D^k}.  
+\]
+In one dimension, $D^k$ is some interval and the $L^2$ inner product over the surface $\LRc{-1,1}$ reduces to 
 \[
@@
 
-Throughout this work, we will assume that integrals and inner products are approximated using some quadrature rule consisting of $N_q$ points $\widehat{x}_i$ and weights $w_i$ over the reference element $\widehat{D}$.
+In typical implementations, integrals and inner products are  approximated using a quadrature rule consisting of $N_q \geq N_p$ points $\widehat{x}_i$ and weights $w_i$ over the reference element $\widehat{D}$, such that
 \[
@@
 \]
-This also defines a quadrature-based $L^2$ projection operator.
-We emphasize that, throughout this work, we will only use properties of inner products and projection operators which hold when integrals are approximated using quadrature.
-We assume only that the quadrature is sufficiently accurate to integrate degree $(2N-1)$ polynomials, such that integration by parts holds on the reference element.
-However, we will also allow for quadrature with an arbitrary number of points and arbitrary accuracy.
+This also defines an analogous quadrature-based $L^2$ projection operator.
+We emphasize that, throughout this work, we will only use properties of quadrature-based inner products and projection operators.
+We assume only that the quadrature is sufficiently accurate to integrate degree $(2N-1)$ polynomials, such that integration by parts holds for any two polynomials on the reference element, while allowing for quadratures with an arbitrary number of points.
 
@@
 \end{equation}
-It was shown in \cite{chen2017entropy} that flux differencing is a consistent approximation of the flux derivative.  
+It was shown in \cite{chen2017entropy} that flux differencing provides a consistent approximation of the flux derivative $\pd{\bm{f}(\bm{u})}{x}$.  
 %Since the two-point flux is consistent,
@@
 Assuming smooth $u$, Burgers' equation conserves a square entropy $U(\bm{u}) = u^2/2$.
-This can be shown by multiplying by $u$, integrating over $[-1,1]$, and using $\pd{u}{t}u = \frac{1}{2}\pd{u^2}{t}$
+This can be shown by multiplying by $u$, integrating over $[-1,1]$, using the chain rule in time and space, and integrating by parts to arrive at
 \begin{equation}
@@
 
-It is well known that DG formulations based on a split formulation can be shown to conserve the square entropy $u^2$ \cite{gassner2013skew, ranocha2017extended}.
-This form may be also be recovered using flux differencing under the two-point flux
+An alternative rewriting of the Burgers' equation is the split formulation \cite{gassner2013skew, ranocha2017extended}, which may be also be recovered using flux differencing under the following two-point flux
 \[
@@
 
-Both the conservative and split forms of Burgers' equation are equivalent at the continuous level (assuming continuity of $u$ and exact differentiation); however, this is no longer true at the discrete level.
+While both the conservative and split forms of Burgers' equation are equivalent at the continuous level (assuming smoothness of $u$), they are not equivalent at the discrete level.
 Suppose now that the solution is approximated by $u \in P^N\LRp{\widehat{D}}$ on the reference element.
@@
 \]
-This scheme does not, in general, conserve energy.
+This scheme does not, in general, conserve the square entropy $u^2/2$.
 %Moreover, if $\Pi_N$ is computed using quadrature, rewriting the scheme into this form suggests that de-aliasing techniques based on over-integration hit a ``bottleneck'', due to the fact that increasing quadrature accuracy only increases the accuracy of the projection \cite{kirby2003aliasing}.
 
-We can derive a semi-discrete Galerkin scheme based which conserves energy discretely using flux differencing and a different choice of discrete differentiation operator $D_N$, which was motivated by the SBP operators introduced in \cite{chen2017entropy}.
-Define $H^1\LRp{\widehat{D}}$ as the Sobolev space
-\[
-\nor{u}_{H^1\LRp{\widehat{D}}}^2 = \nor{u}_{L^2\LRp{\widehat{D}}}^2 + \nor{\pd{u}{x}}_{L^2\LRp{\widehat{D}}}^2, \qquad H^1\LRp{\widehat{D}} = \LRc{ u \in L^2\LRp{\widehat{D}}, \qquad \nor{u}_{H^1\LRp{\widehat{D}}} < \infty}.
-\]
-Because functions $u \in H^1\LRp{\widehat{D}}$ admit a well-defined trace \cite{brenner2007mathematical}, we can define the following discrete differential operator as a map from $H^1$ to $P^N$:
+We can derive a semi-discrete Galerkin scheme which conserves energy using flux differencing and a discrete differentiation operator $D_N$, which is inspired by the SBP operators introduced in \cite{chen2017entropy}.
+Define $H^1\LRp{\widehat{D}}$ as the Sobolev space over $\widehat{D}$
+\begin{equation}
+\nor{u}_{H^1\LRp{\widehat{D}}}^2 = \nor{u}_{L^2\LRp{\widehat{D}}}^2 + \nor{\pd{u}{x}}_{L^2\LRp{\widehat{D}}}^2, \qquad H^1\LRp{\widehat{D}} = \LRc{ u \in L^2\LRp{\widehat{D}}, \qquad \nor{u}_{H^1\LRp{\widehat{D}}} < \infty}.\label{eq:sobolev}
+\end{equation}
+Because functions $u \in H^1\LRp{\widehat{D}}$ admit a well-defined boundary trace \cite{brenner2007mathematical}, we can define the following discrete differential operator as a map from $H^1$ to $P^N$:
 \begin{definition}
-Let $w(x)$ be a bounded function on $\widehat{D}$.  Define $D_N: H^1\LRp{\widehat{D}}\rightarrow P^N\LRp{\widehat{D}}$ such that, for $u\in H^1\LRp{\widehat{D}}$, 
+\label{def:dgd}
+Let $w(x)$ be a continuous and bounded function on $\widehat{D}$.  Define $D_N: H^1\LRp{\widehat{D}}\rightarrow P^N\LRp{\widehat{D}}$ such that, for $u\in H^1\LRp{\widehat{D}}$, 
 \begin{align}
 &\LRp{D_N u, vw}_{\widehat{D}} = \LRp{\pd{ \Pi_N u}{x},vw}_{\widehat{D}}\nonumber \\
-&+ \frac{1}{2}{\LRa{{u - \Pi_N u}, vw\bm{n}_x}}_{\partial \widehat{D}} + \frac{1}{2}\LRa{u - \Pi_Nu ,\Pi_N\LRp{vw}\bm{n}_x }_{\widehat{D}}, \qquad \forall v\in P^N\LRp{\widehat{D}}.
-\label{eq:dgd}
+&+ \frac{1}{2}{\LRa{{u - \Pi_N u}, vw{n}_x}}_{\partial \widehat{D}} + \frac{1}{2}\LRa{u - \Pi_Nu ,\Pi_N\LRp{vw}{n}_x }_{\widehat{D}}, \qquad \forall v\in P^N\LRp{\widehat{D}},
 \end{align}
+where ${n}_x$ is the $x$-component of the outward normal vector.  
 \end{definition}
@@
 \[
-\LRp{D_N u, wv}_{\widehat{D}} = \LRa{u,wv\bm{n}_x}_{\partial \widehat{D}} - \LRp{u,D_N (w v)}_{\widehat{D}}.
+\LRp{D_N u, wv}_{\widehat{D}} = \LRa{u,wv{n}_x}_{\partial \widehat{D}} - \LRp{u,D_N (w v)}_{\widehat{D}}.
 \]
@@
 \[
--\LRp{\pd{ \Pi_N (vw)}{x},u}_{\widehat{D}} + \frac{1}{2}{\LRa{{u - \Pi_N u}, vw\bm{n}_x}}_{\partial \widehat{D}} + \frac{1}{2}\LRa{u + \Pi_Nu ,\Pi_N\LRp{vw}\bm{n}_x }_{\partial \widehat{D}}, \qquad \forall v\in P^N\LRp{\widehat{D}}.
+-\LRp{\pd{ \Pi_N (vw)}{x},u}_{\widehat{D}} + \frac{1}{2}{\LRa{{u - \Pi_N u}, vw{n}_x}}_{\partial \widehat{D}} + \frac{1}{2}\LRa{u + \Pi_Nu ,\Pi_N\LRp{vw}{n}_x }_{\partial \widehat{D}}, \qquad \forall v\in P^N\LRp{\widehat{D}}.
 \]
@@
 \[
--\LRp{\pd{ \Pi_N (vw)}{x},u}_{\widehat{D}} + \frac{1}{2}\LRa{vw + \Pi_N(vw) , u\bm{n}_x }_{\partial \widehat{D}}  - \frac{1}{2}\LRa{vw - \Pi_N (vw), \Pi_N u \bm{n}_x}_{\partial \widehat{D}}.  
-\]
-Adding and subtracting $\LRa{vw,u\bm{n}_x}_{\partial \widehat{D}}$ completes the proof.
+-\LRp{\pd{ \Pi_N (vw)}{x},u}_{\widehat{D}} + \frac{1}{2}\LRa{vw + \Pi_N(vw) , u{n}_x }_{\partial \widehat{D}}  - \frac{1}{2}\LRa{vw - \Pi_N (vw), \Pi_N u {n}_x}_{\partial \widehat{D}}.  
+\]
+Adding and subtracting $\LRa{vw,u{n}_x}_{\partial \widehat{D}}$ completes the proof.
 \end{proof}
@@
 \[
-\frac{1}{2}\pd{}{t}\LRp{u,u}_{\widehat{D}} = -\frac{1}{3}\LRa{u^2,u}_{\partial \widehat{D}},
-\]
-which is a discrete analogue of the conservation condition (\ref{eq:burgersconservation}).
+\frac{1}{2}\pd{}{t}\LRp{u,u}_{\widehat{D}} = \frac{1}{2}\pd{}{t}\nor{u}^2_{\widehat{D}} = -\frac{1}{3}\LRa{u^2,u}_{\partial \widehat{D}},
+\]
+which is a discrete analogue of the conservation of energy (\ref{eq:burgersconservation}) for smooth solutions.
 This is equivalent to the skew-symmetric split formulation described in \cite{gassner2013skew}; however, as noted in \cite{chen2017entropy}, it can be derived from flux differencing without explicit knowledge of the skew form.
@@
 \subsection{Conservation of entropy on one element}
-
+\label{sec:ec}
 While flux differencing can be used to recover split formulations, it can also be used to construct entropy stable formulations which do not correspond to skew-symmetric splittings.
-In this section we will show that, for any sufficiently accurate quadrature rule, it is possible to construct discretizations which satisfy a discrete equivalent of conservation of entropy (\ref{eq:consentropy}).
+In this section we will show that, for any sufficiently accurate quadrature rule and appropriate numerical flux function, it is possible to construct discretizations which satisfy a discrete equivalent of conservation of entropy (\ref{eq:consentropy}).
 
@@
 \]
-where $\bm{f}_i, \bm{g}_i$ are real and bounded for appropriate values of $x,y$.  
+where $\bm{f}_i, \bm{g}_i$ are real and bounded for appropriate values of $x,y$ over each element.  
 \label{eq:assumption}
@@
 The precise definition of what constitutes ``appropriate values of $x,y$'' depends on the physical problem.
-For example, the flux functions for the compressible Euler equations require positivity constraints, which are outlined in more detail in Section~\ref{sec:num}.
+For example, the flux functions for the compressible Euler equations require positivity constraints on the inputs, which are outlined in more detail in Section~\ref{sec:num}.
 This assumption is satisfied, for example, by $\bm{f}_S$ which are real analytic and thus admit a Taylor expansion in the two arguments.
-In Appendix~\ref{appendix:A}, we argue that this assumption also holds for the choices of flux functions $\bm{f}_S$ considered in this work for the Burgers' and compressible Euler equations, assuming the solution satisfies physical bound constraints.
-However, we emphasize that this assumption is not required in the computation implementations of the proposed methods, and that it may be possible to relax these conditions on $\bm{f}_S$.
+In \ref{appendix:A}, we argue that this assumption also holds for the flux functions $\bm{f}_S$ considered in this work for the Burgers' and compressible Euler equations, assuming the solution satisfies physical bound constraints.
+However, we emphasize that this assumption is not required in the computation implementations of the proposed methods, and that it may also be possible to relax these conditions on $\bm{f}_S$.
 
@@
 \begin{equation}
-\LRp{\pd{\bm{u}}{t},\bm{w}}_{\widehat{D}} + \LRp{\left.\LRp{2D_N \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},\bm{w}}_{\widehat{D}}= \LRa{\bm{f}^*-\bm{f}(\bm{u}_{\bm{v}}),\bm{w}}_{\partial \widehat{D}}, \qquad \bm{w}\in P^N\LRp{\widehat{D}},
+\LRp{\pd{\bm{u}}{t},\bm{w}}_{\widehat{D}} + \LRp{\left.\LRp{2D_N \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},\bm{w}}_{\widehat{D}}+ \LRa{\bm{f}^*-\bm{f}(\bm{u}_{\bm{v}}),\bm{w}{n}_x}_{\partial \widehat{D}} = 0, \qquad \bm{w}\in P^N\LRp{\widehat{D}},
 \label{eq:ecdg_oneelem}
@@
 \label{lemma:ec1}
-The formulation (\ref{eq:ecdg_oneelem})
+Let Assumption~\ref{eq:assumption} hold on $\widehat{D}$.  The formulation (\ref{eq:ecdg_oneelem})
  %and (\ref{eq:ecdg_oneelem_mat}) are
@@
 \[
-\LRp{\pd{U(\bm{u})}{t},1}_{\widehat{D}} =  \LRa{\psi,1}_{\partial \widehat{D}}-\LRa{{\bm{f}^*, \Pi_N\bm{v}}}_{\partial \widehat{D}}.
+\LRp{\pd{U(\bm{u})}{t},1}_{\widehat{D}} =  \LRa{\psi,1{n}_x}_{\partial \widehat{D}}-\LRa{{\bm{f}^*, \Pi_N\bm{v}}{n}_x}_{\partial \widehat{D}}.
 \]
@@
 %where we have assumed continuity in time, and have used that $\bm{W}$ is diagonal.  
-For the spatial derivative, we expand $\bm{f}_S$ out using Assumption~\ref{eq:assumption} and use Lemma~\ref{lemma:dgd_local}
+For the spatial derivative, we expand $\bm{f}_S$ out using Assumption~\ref{eq:assumption} and apply Lemma~\ref{lemma:dgd_local}
 \begin{align*}
 &\LRp{\left.\LRp{D_N \sum_{i=1}^{\infty}\bm{f}_i(\bm{u}_{\bm{v}}(x))\bm{g}_i(\bm{u}_{\bm{v}}(y))}\right|_{y=x}, \Pi_N\bm{v}(x)}_{\widehat{D}} = \sum_{i=1}^{\infty} \LRp{{D_N \bm{f}_i(\bm{u}_{\bm{v}})}, {\bm{g}_i(\bm{u}_{\bm{v}})\Pi_N\bm{v} }}_{\widehat{D}}\\
-&= \sum_{i=1}^{\infty}\LRa{\bm{f}_i(\bm{u}_{\bm{v}}) \bm{g}_i(\bm{u}_{\bm{v}}),\Pi_N\bm{v}}_{\partial \widehat{D}}  - \LRp{ \bm{f}_i(\bm{u}_{\bm{v}}), D_N \LRp{\bm{g}_i\Pi_N\bm{v} }}_{\widehat{D}}\\
-&= \LRa{\bm{f}_S(\bm{u}_{\bm{v}},\bm{u}_{\bm{v}}), \Pi_N\bm{v}(x)}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N \LRp{\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))\Pi_N\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}}\\
-&= \LRa{\bm{f}(\bm{u}_{\bm{v}}), \Pi_N\bm{v}(x)}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N\LRp{\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))\Pi_N\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}},
+&= \sum_{i=1}^{\infty}\LRa{\bm{f}_i(\bm{u}_{\bm{v}}) \bm{g}_i(\bm{u}_{\bm{v}}),\Pi_N\bm{v}{n}_x}_{\partial \widehat{D}}  - \LRp{ \bm{f}_i(\bm{u}_{\bm{v}}), D_N \LRp{\bm{g}_i\Pi_N\bm{v} }}_{\widehat{D}}\\
+&= \LRa{\bm{f}_S(\bm{u}_{\bm{v}},\bm{u}_{\bm{v}}), \Pi_N\bm{v}(x){n}_x}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N \LRp{\bm{f}_S(\bm{u}_{\bm{v}}(y),\bm{u}_{\bm{v}}(x))\Pi_N\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}}\\
+%&= \LRa{\bm{f}_S(\bm{u}_{\bm{v}},\bm{u}_{\bm{v}}), \Pi_N\bm{v}(x)}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N \LRp{\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))\Pi_N\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}}\\
+&= \LRa{\bm{f}(\bm{u}_{\bm{v}}), \Pi_N\bm{v}(x){n}_x}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N\LRp{\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))\Pi_N\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}},
 \end{align*}
-where we have used the consistency of $\bm{f}_S$ in the last step.  
+where we have used the linearity of $D_N$ in the first step and the symmetry and consistency of $\bm{f}_S$ in the third and last steps.  
 %where $\Pi_N^x, \Pi_N^y$ denote $L^2$ projections taken with respect to the $x$ and $y$ coordinates, respectively.  
@@
 We see that, in the proof of Theorem~\ref{lemma:ec1}, recovering the time derivative of entropy on the left hand side requires testing with the projection of the entropy variables.
-For diagonal norm SBP-DG methods, $L^2$ projection and interpolation are equivalent, which implies that evaluating the flux $\bm{f}_S$ using point-wise nodal values of the solution satisfies the entropy conservation condition given in Definition~\ref{def:tadmor}.
-Diagonal norm SBP-DG methods can also be simplified by noting that the projection operator reduces to a nodal interpolation operator.
+For diagonal norm SBP-DG methods, $L^2$ projection and nodal interpolation are equivalent, which implies that evaluating the flux $\bm{f}_S$ using point-wise nodal values of the solution satisfies the entropy conservation condition given in Definition~\ref{def:tadmor}.
+%Diagonal norm SBP-DG methods can be further simplified by noting that the projection operator reduces to a nodal interpolation operator.
 
 In general, the projection of the entropy variables does not agree with their point-wise evaluation.
-However, the entropy conservation condition given in Definition~\ref{def:tadmor} holds so long as the flux is evaluated in a manner consistent with the entropy variables.
-This motivates (\ref{eq:projectedentropyvars}) as a non-standard way of evaluating the flux, which is the primary difference between the construction of entropy conservative schemes for diagonal norm SBP-DG methods and projection-based DG methods.
+However, the entropy conservation condition given in Definition~\ref{def:tadmor} holds so long as the flux is evaluated in a manner consistent with the projected entropy variables used to recover the time derivative of entropy.
+This motivates (\ref{eq:projectedentropyvars}) as a non-standard way of evaluating the flux, which is the primary difference between the construction of entropy conservative schemes for diagonal norm SBP-DG methods and more general DG methods.
+
+Finally, we note that, while projected entropy variables can be used to build entropy conservative DG methods, the evaluation $\bm{u}\LRp{\Pi_N \bm{v}}$ can be sensitive to differences between $\bm{v}$ and $\Pi_N \bm{v}$, which are illustrated numerically and described in more detail in Section~\ref{sec:instab}.
 
@@
 In this section, we discuss the extension of entropy conservative DG schemes to multiple elements.
-The primary difficulty in this case is the construction of appropriate interface (SBP-SAT) terms.
+The primary difficulty is the construction of appropriate interface (SBP-SAT) terms.
 Entropy stable formulations based on dense norm or GSBP operators require SBP-SAT terms which are either non-trivial to derive or have a wide stencil \cite{fernandez2016simultaneous, ranocha2017extended}.
-We present an approach which instead utilizes a global approach to flux differencing to derive entropy conservative interface terms, and show that it recovers existing energy stable SBP-SAT terms for Burgers' equation.
+We present an approach which instead utilizes a global approach to flux differencing to derive entropy conservative interface terms, and show that we can recover existing energy stable SBP-SAT terms for Burgers' equation.
 
@@
 The construction of energy and entropy conservative DG-SBP methods requires the specification of appropriate SBP-SAT terms, which are often derived by reverse-engineering  boundary terms such that interface contributions cancel \cite{fernandez2016simultaneous, ranocha2017comparison, ranocha2017extended}.
-We present an alternative systematic approach to the derivation of interface terms for DG formulations by combining flux differencing with global discrete DG differential operators, which have been used to simplify the notation and analysis of DG formulations for elliptic equations \cite{di2011mathematical, wang2013weak}.
+We present an alternative systematic approach to the derivation of interface terms for DG formulations by combining flux differencing with global DG differential operators, which have been used to simplify the notation and analysis of DG formulations \cite{hesthaven2004high, di2011mathematical, wang2013weak}.
 This approach does not require the specification of appropriate SBP-SAT terms in order to prove entropy conservation; rather, explicit expressions for SBP-SAT terms may be derived in a systematic manner using flux differencing and the definitions of global DG operators.
@@
 \[
-\LRp{u,v}_{\Omega} = \sum_k \LRp{u,v}_{D^k}, \qquad H^1\LRp{\Oh} = \bigoplus_{k} H^1\LRp{D^k}.
-\]
-We adapt the construction of local differentiation operators introduced previously to construct a global DG differentiation operator:
+\LRp{u,v}_{\Omega} = \sum_k \LRp{u,v}_{D^k}, \qquad H^1\LRp{\Oh} = \bigoplus_{k} H^1\LRp{D^k},
+\]
+where the local Sobolev spaces $H^1\LRp{D^k}$ are defined analogously to (\ref{eq:sobolev}).  
+We adapt the construction of local differentiation operators introduced in Section~\ref{sec:ec} and \cite{chen2017entropy} to construct a global DG differentiation operator:
 \begin{definition}
-Let $w(x)$ is a bounded function over each element $D^k$.  Define $D^x_h: H^1\LRp{\Oh}\rightarrow V_h$ be as the operator which satisfies
+Let $w(x)$ be a bounded function over each element $D^k$.  Define $D^x_h: H^1\LRp{\Oh}\rightarrow V_h$ as the operator which satisfies
 \begin{align}
 &\LRp{D^x_h u,vw}_{\Omega} = \sum_k \LRp{\pd{ \Pi_N u}{x},vw}_{D^k} \nonumber\\
-&+ \frac{1}{2}{\LRa{{u^+ - \Pi_N u}, vw\bm{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{{u - \Pi_Nu },\Pi_N(vw) \bm{n}_x}_{\partial D^k}}, \qquad \forall v \in V_h.
+&+ \frac{1}{2}{\LRa{{u^+ - \Pi_N u}, vw{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{{u - \Pi_Nu },\Pi_N(vw) {n}_x}_{\partial D^k}}, \qquad \forall v \in V_h.
 \label{eq:dgd_global}
@@
 \[
-\LRp{D^x_h u,vw}_{\Omega} = \LRa{u,vw\bm{n}_x}_{\partial \Omega} - \LRp{u,D^x_h (vw)}_{\Omega}.
+\LRp{D^x_h u,vw}_{\Omega} = \LRa{u,vw{n}_x}_{\partial \Omega} - \LRp{u,D^x_h (vw)}_{\Omega}.
 \]
@@
 \begin{proof}
-We may rewrite $\LRp{D^x_h u,vw}_{\Omega}$ using the Definition~\ref{eq:dgd} of $D_N$
-\[
-\LRp{D^x_h u,vw}_{\Omega} = \sum_k \LRp{D_N u,vw}_{D^k} + \frac{1}{2}\LRa{\jump{u},vw\bm{n}_x}_{\partial D^k \setminus \partial \Omega}.  
-\]
-Applying Lemma~\ref{lemma:dgd_local} over each element yields 
+We may rewrite $\LRp{D^x_h u,vw}_{\Omega}$ using Definition~\ref{def:dgd} of $D_N$
+\[
+\LRp{D^x_h u,vw}_{\Omega} = \sum_k \LRp{D_N u,vw}_{D^k} + \frac{1}{2}\LRa{\jump{u},vw{n}_x}_{\partial D^k \setminus \partial \Omega}.  
+\]
+Assuming $D^k$ are affine mappings of $\widehat{D}$ and applying Lemma~\ref{lemma:dgd_local} over each element yields 
 \begin{align*}
-\LRp{D^x_h u,vw}_{\Omega} &= \sum_k \LRp{-u,D_N (vw)}_{D^k} + \LRa{u + \frac{1}{2}\jump{u},vw\bm{n}_x}_{\partial D^k}\\
-&= \sum_k \LRp{-u,D_N (vw)}_{D^k} + \LRa{\avg{u},vw\bm{n}_x}_{\partial D^k}\\
-&= \sum_k \LRp{-u,D_N (vw)}_{D^k} - \LRa{\jump{vw},u\bm{n}_x}_{\partial D^k \setminus \partial \Omega} + \LRa{vw,u\bm{n}_x}_{\partial \Omega} \\
-&= \LRa{vw,u\bm{n}_x}_{\partial \Omega}-\LRp{u,D^x_h (vw)}_{\Omega}.    
+\LRp{D^x_h u,vw}_{\Omega} &= \sum_k \LRp{-u,D_N (vw)}_{D^k} + \LRa{u + \frac{1}{2}\jump{u},vw{n}_x}_{\partial D^k}\\
+&= \sum_k \LRp{-u,D_N (vw)}_{D^k} + \LRa{\avg{u},vw{n}_x}_{\partial D^k}\\
+&= \sum_k \LRp{-u,D_N (vw)}_{D^k} - \LRa{\jump{vw},u{n}_x}_{\partial D^k \setminus \partial \Omega} + \LRa{vw,u{n}_x}_{\partial \Omega} \\
+&= \LRa{vw,u{n}_x}_{\partial \Omega}-\LRp{u,D^x_h (vw)}_{\Omega}.    
 \end{align*}
@@
 \[
-%\sum_k \LRp{D^x_h u,v}_{D^k} = \sum_k \LRp{\pd{u}{x},v}_{D^k} + \frac{1}{2}\LRa{\jump{u},v\bm{n}_x}_{\partial D^k}, \qquad \forall v \in V_h.
-\LRp{D^x_h u,v}_{\Omega} = \sum_k \LRp{\pd{u}{x},v}_{D^k} + \frac{1}{2}\LRa{\jump{u},v\bm{n}_x}_{\partial D^k}, \qquad \forall v \in V_h.
-\]
-Then, using that the range of $D^x_h$ is $V_h$, it is straightforward to show that $D^x_h \Pi_N$ satisfies an analogue of integration by parts, such that
-\[
-\LRp{D^x_h\Pi_N u, v}_{\Omega} = \LRa{\Pi_N u,v}_{\partial \Omega} - \LRp{D^x_h\Pi_N v, u}_{\Omega}, \qquad \forall u,v\in \L.
+%\sum_k \LRp{D^x_h u,v}_{D^k} = \sum_k \LRp{\pd{u}{x},v}_{D^k} + \frac{1}{2}\LRa{\jump{u},v{n}_x}_{\partial D^k}, \qquad \forall v \in V_h.
+\LRp{D^x_h u,v}_{\Omega} = \sum_k \LRp{\pd{u}{x},v}_{D^k} + \frac{1}{2}\LRa{\jump{u},v{n}_x}_{\partial D^k}, \qquad \forall v \in V_h.
+\]
+Then, using that the range of $D^x_h$ is $V_h$, it is straightforward to show that $D^x_h \Pi_N$ satisfies a different discrete analogue of integration by parts
+\[
+\LRp{D^x_h\Pi_N u, v}_{\Omega} = \LRa{\Pi_N u,v{n}_x}_{\partial \Omega} - \LRp{D^x_h\Pi_N v, u}_{\Omega}, \qquad \forall u,v\in \L.
 \]
 %Numerical results using this global differentiation operator result in slightly more accurate results than when using the DG operator (\ref{eq:dgd_global}).  
-Unfortunately, while (\ref{eq:dgd_global}) requires only traces of neighboring elements, this construction results in a wide stencil involving all degrees of freedom on neighboring elements due to the presence of $\jump{u}$ and the $L^2$ projection.
+Unfortunately, while (\ref{eq:dgd_global}) requires only boundary values of solutions on neighboring elements, this construction results in a wide stencil involving all degrees of freedom on neighboring elements due to the combination of $\jump{u}$ and the $L^2$ projection.
 This choice of global DG differentiation operator is analogous to the choice of SBP-SAT terms specified in \cite{fernandez2016simultaneous} for SBP operators on simplicial elements without boundary nodes.
@@
 \label{sec:multipleelems}
-Extending the entropy stable DG formulation (\ref{eq:ecdg_oneelem}) is done by combining the flux differencing with the global DG differential operator $D^x_h$
+Extending the entropy stable DG formulation (\ref{eq:ecdg_oneelem}) to multiple elements is done by combining flux differencing with the global DG differential operator $D^x_h$
 \begin{equation}
-\LRp{\pd{\bm{u}}{t},\bm{w}}_{\Omega} + \LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},\bm{w}}_{\Omega} = \LRa{\bm{f}(\bm{u}_{\bm{v}})-\bm{f}^*,\bm{w}}_{\partial \Omega},
+\LRp{\pd{\bm{u}}{t},\bm{w}}_{\Omega} + \LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},\bm{w}}_{\Omega} + \LRa{\bm{f}^*-\bm{f}(\bm{u}_{\bm{v}}),\bm{w}{n}_x}_{\partial \Omega},
 %\int_{D^k}\pd{\bm{u}}{t} \bm{w} + \left.\LRp{2D^x_h\bm{f}_S\LRp{\bm{u}(\Pi_N\bm{v}({x})),\bm{u}(\Pi_N\bm{v}({y}))}}\right|_{{y}={x}}\bm{w} + \LRa{\bm{f}^*-\bm{f}(\bm{u}),\bm{w}}_{\partial \Omega}= 0, \qquad \forall \bm{w}\in V_h^n,
@@
 where $\bm{f}^*$ is some numerical flux used in the imposition of boundary conditions, and the global operator $D^x_h$ is understood to be applied to each component of $\bm{f}_S\LRp{\bm{u}(x),\bm{u}(y)}$.
-
 This formulation yields a theorem analogous to Theorem~\ref{lemma:ec1}:
@@
 \label{thm:ecdg_global}
-The formulation (\ref{eq:ecdg}) is globally entropy conservative in the sense that
-\[
-\LRp{\pd{U(\bm{u})}{t},1}_{\Omega} =  \LRa{\psi,1}_{\partial \Omega}-\LRa{{\bm{f}^*, \Pi_N\bm{v}}}_{\partial \Omega}.
+Let Assumption~\ref{eq:assumption} hold over each element $D^k$.  The formulation (\ref{eq:ecdg}) is globally entropy conservative in the sense that
+\[
+\LRp{\pd{U(\bm{u})}{t},1}_{\Omega} =  \LRa{\psi,1{n}_x}_{\partial \Omega}-\LRa{{\bm{f}^*, \Pi_N\bm{v}}{n}_x}_{\partial \Omega}.
 \]
@@
 For an appropriate approximation of $\lambda$, Lax-Friedrichs penalization dissipates entropy \cite{chen2017entropy}.
-We note that, since conservation of entropy requires testing with the projection of the entropy variables, the jump term should involve the evaluations of the conservation variables in terms of the projected entropy variables $\bm{u}_{\bm{v}} = \bm{u}\LRp{\Pi_N \bm{v}}$ in order to guarantee entropy dissipation.
+We note that, since conservation of entropy requires testing with the projection of the entropy variables, the jump term should involve the evaluations of the conservative variables in terms of the projected entropy variables $\bm{u}_{\bm{v}} = \bm{u}\LRp{\Pi_N \bm{v}}$ in order to guarantee entropy dissipation.
 
@@
 \[
-\LRp{\pd{\bm{u}}{t},1}_{D^k} + \LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},1}_{D^k} = \LRa{\bm{f}(\bm{u}_{\bm{v}})-\bm{f}^*,1}_{\partial \Omega \cap D^k}.  
+\LRp{\pd{\bm{u}}{t},1}_{D^k} + \LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},1}_{D^k} + \LRa{\bm{f}^*-\bm{f}(\bm{u}_{\bm{v}}),1{n}_x}_{\partial \Omega \cap D^k} = 0.
 \]
@@
 \begin{align*}
-\LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},1}_{D^k} =& \sum_{i=1}^{\infty} \LRp{2D_N \bm{f}_i(\bm{u}_{\bm{v}}),\bm{g}_i(\bm{u}_{\bm{v}})}_{D^k} + \LRa{\jump{\bm{f}_i(\bm{u}_{\bm{v}})},\bm{g}_i(\bm{u}_{\bm{v}})}_{\partial D^k}\\
+\LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},1}_{D^k} =& \sum_{i=1}^{\infty} \LRp{2D_N \bm{f}_i(\bm{u}_{\bm{v}}),\bm{g}_i(\bm{u}_{\bm{v}})}_{D^k} + \LRa{\jump{\bm{f}_i(\bm{u}_{\bm{v}})},\bm{g}_i(\bm{u}_{\bm{v}}){n}_x}_{\partial D^k}\\
 =& \sum_{i=1}^{\infty} \LRp{D_N \bm{f}_i(\bm{u}_{\bm{v}}),\bm{g}_i(\bm{u}_{\bm{v}})}_{D^k} - \LRp{D_N \bm{f}_i(\bm{u}_{\bm{v}}),\bm{g}_i(\bm{u}_{\bm{v}})}_{D^k} \\
-&+ \LRa{\bm{f}_i(\bm{u}_{\bm{v}}),\bm{g}_i(\bm{u}_{\bm{v}})}_{\partial D^k} + \LRa{\bm{f}_i(\bm{u}_{\bm{v}}^+)-\bm{f}_i(\bm{u}_{\bm{v}}),\bm{g}_i(\bm{u}_{\bm{v}})}_{\partial D^k}\\
+&+ \LRa{\bm{f}_i(\bm{u}_{\bm{v}}),\bm{g}_i(\bm{u}_{\bm{v}}){n}_x}_{\partial D^k} + \LRa{\bm{f}_i(\bm{u}_{\bm{v}}^+)-\bm{f}_i(\bm{u}_{\bm{v}}),\bm{g}_i(\bm{u}_{\bm{v}}){n}_x}_{\partial D^k}\\
 =& \LRp{D_N \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y)),1}_{D^k} -  \LRp{D_N \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y)),1}_{D^k}\\
-&+ \LRa{\bm{f}_S(\bm{u}_{\bm{v}}^+,\bm{u}_{\bm{v}}),1}_{\partial D^k}.
+&+ \LRa{\bm{f}_S(\bm{u}_{\bm{v}}^+,\bm{u}_{\bm{v}}),1{n}_x}_{\partial D^k}.
 \end{align*}
@@
 \[
-\sum_{k}\LRp{\pd{\bm{u}}{t},1}_{D^k} + \LRa{\bm{f}_S(\bm{u}_{\bm{v}}^+,\bm{u}_{\bm{v}}),1}_{\partial D^k} = \LRa{\bm{f}(\bm{u}_{\bm{v}})-\bm{f}^*,1}_{\partial \Omega \cap D^k}.  
-\]
-On the domain boundary $\partial \Omega$, $\bm{u}_{\bm{v}}^+ = \bm{u}_{\bm{v}}$ and $\bm{f}_S(\bm{u}_{\bm{v}}^+,\bm{u}_{\bm{v}}) = \bm{f}(\bm{u}_{\bm{v}})$.
+\sum_{k}\LRp{\pd{\bm{u}}{t},1}_{D^k} + \LRa{\bm{f}_S(\bm{u}_{\bm{v}}^+,\bm{u}_{\bm{v}}),1{n}_x}_{\partial D^k} + \LRa{\bm{f}^* - \bm{f}(\bm{u}_{\bm{v}}),1{n}_x}_{\partial \Omega \cap D^k} = 0.  
+\]
+On the domain boundary $\partial \Omega$, $\bm{u}_{\bm{v}}^+ = \bm{u}_{\bm{v}}$ and $\bm{f}_S(\bm{u}_{\bm{v}}^+,\bm{u}_{\bm{v}}) = \bm{f}(\bm{u}_{\bm{v}})$ using the consistency of $\bm{f}_S$.
 This implies that the statement of local conservation replaces $\bm{f}_S$ with the boundary condition numerical flux $\bm{f}^*$ on element boundaries which intersect with the domain boundary.
@@
 
-We note that this definition of local conservation is given in terms of the conservation variables evaluated as a function of the projected entropy variables
-\[
-\bm{u}_{\bm{v}}(x)(x) = \bm{u}\LRp{\LRp{\Pi_N \bm{v}}(x)}.
-\]
-It is not immediately clear whether this statement of local conservation satisfies the conditions of the classic Lax-Wendroff theorem.
-However, this form of local conservation does satisfy the generalized definition of local conservation given in \cite{shi2017local}, which is sufficient to guarantee convergence to the weak solution.
+We note that this definition of local conservation is given in terms of the conservative variables evaluated as a function of the projected entropy variables $\bm{u}\LRp{{\Pi_N \bm{v}}}$.  
+%\[
+%\bm{u}_{\bm{v}}(x)(x) = \bm{u}\LRp{\LRp{\Pi_N \bm{v}}(x)}.
+%\]
+It is not immediately clear that this statement of local conservation satisfies the conditions of the classic Lax-Wendroff theorem.
+However, this form of local conservation does satisfy a generalized definition of local conservation, which is sufficient to guarantee convergence to a weak solution under mesh refinement \cite{shi2017local}.
 
@@
 \]
-We focus on the discretization of the spatial terms.  Applying flux differencing to the flux to yield the following contributions
+We focus on the discretization of the spatial terms.
+Since the solution is itself the entropy variable for Burgers equation, we can apply flux differencing directly to the flux evaluated with $u$ to yield the following contributions
 \[
@@
 \begin{align*}
-\LRp{D^x_h\LRp{u^2} , v}_{\Omega} &= \sum_k \LRp{\pd{\Pi_N u^2}{x} , v}_{D^k} + \frac{1}{2}\LRa{\LRp{u^+}^2-\Pi_N u^2,v\bm{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{u^2 - \Pi_N u^2, \Pi_N v \bm{n}_x}_{\partial D^k}\\
-&= \sum_k \LRp{\pd{\Pi_N u^2}{x} , v}_{D^k} + \LRa{\frac{1}{2}\LRp{(u^+)^2 + u^2}-\Pi_N u^2,v\bm{n}_x}_{\partial D^k}.
+\LRp{D^x_h\LRp{u^2} , v}_{\Omega} &= \sum_k \LRp{\pd{\Pi_N u^2}{x} , v}_{D^k} + \frac{1}{2}\LRa{\LRp{u^+}^2-\Pi_N u^2,v{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{u^2 - \Pi_N u^2, \Pi_N v {n}_x}_{\partial D^k}\\
+&= \sum_k \LRp{\pd{\Pi_N u^2}{x} , v}_{D^k} + \LRa{\frac{1}{2}\LRp{(u^+)^2 + u^2}-\Pi_N u^2,v{n}_x}_{\partial D^k}.
 \end{align*}
@@
 \begin{align*}
-\LRp{u D^x_h u, v}_{\Omega} &= \LRp{D^x_h u, uv}_{\Omega} = \sum_k \LRp{\pd{u}{x},uv}_{D^k} + \frac{1}{2}\LRa{{u^+}-\Pi_N u, uv\bm{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{u - \Pi_N u, \Pi_N (uv)}_{\partial D^k}\\
-&= \sum_k \LRp{\pd{u}{x},uv}_{D^k} + \frac{1}{2}\LRa{{u^+}u- u^2, v\bm{n}_x}_{\partial D^k}.
+\LRp{u D^x_h u, v}_{\Omega} &= \LRp{D^x_h u, uv}_{\Omega} = \sum_k \LRp{\pd{u}{x},uv}_{D^k} + \frac{1}{2}\LRa{{u^+}-\Pi_N u, uv{n}_x}_{\partial D^k} + \frac{1}{2}\LRa{u - \Pi_N u, \Pi_N (uv)}_{\partial D^k}\\
+&= \sum_k \LRp{\pd{u}{x},uv}_{D^k} + \frac{1}{2}\LRa{{u^+}u- u^2, v{n}_x}_{\partial D^k}.
 \end{align*}
@@
 \[
-%\sum_k \LRp{\pd{u}{t} + \frac{1}{3}\LRp{\pd{\Pi_N u^2}{x} + u\pd{u}{x}} , v}_{D^k} + \LRa{\frac{1}{6}\LRp{(u^+)^2 + u^+u + u^2}- \frac{1}{3}\Pi_N u^2 - \frac{1}{6}u^2,v\bm{n}_x}_{\partial D^k} = 0.
-\sum_k \LRp{\pd{u}{t} + \frac{1}{3}\LRp{\pd{\Pi_N u^2}{x} + u\pd{u}{x}} , v}_{D^k} + \LRa{f_S(u^+,u)- \frac{1}{3}\Pi_N u^2 - \frac{1}{6}u^2,v\bm{n}_x}_{\partial D^k} = 0.
-\]
-This formulation, which was derived through the use of flux differencing with the global DG operator (\ref{eq:dgd_global}), recovers the weak form of the split formulation and the correction terms introduced in \cite{ranocha2017extended} for Burgers' equation.
+%\sum_k \LRp{\pd{u}{t} + \frac{1}{3}\LRp{\pd{\Pi_N u^2}{x} + u\pd{u}{x}} , v}_{D^k} + \LRa{\frac{1}{6}\LRp{(u^+)^2 + u^+u + u^2}- \frac{1}{3}\Pi_N u^2 - \frac{1}{6}u^2,v{n}_x}_{\partial D^k} = 0.
+\sum_k \LRp{\pd{u}{t} + \frac{1}{3}\LRp{\pd{\Pi_N u^2}{x} + u\pd{u}{x}} , v}_{D^k} + \LRa{f_S(u^+,u)- \frac{1}{3}\Pi_N u^2 - \frac{1}{6}u^2,v{n}_x}_{\partial D^k} = 0.
+\]
+This formulation, which was derived through the use of flux differencing with the global DG operator (\ref{eq:dgd_global}), recovers the weak form of the split formulation and the correction terms introduced in \cite{ranocha2017extended} for Burgers' equation.\footnote{Entropy conservative SBP-SAT were introduced for generalized SBP operators and the compressible Euler equations in \cite{ranocha2017comparison}; however, we have not yet determined if the combination of flux differencing with (\ref{eq:dgd_global}) recovers these fluxes as well.
+}
 
@@
 \label{sec:implementation}
-While it is possible to use (\ref{eq:ecdg}) to derive explicit DG weak formulations, this can be difficult for complicated flux functions.
+While it is possible to use (\ref{eq:ecdg}) to derive explicit DG weak formulations, this becomes impractical for complicated flux functions.
 Instead, we follow \cite{gassner2016split, chen2017entropy} and seek an implementation of flux differencing and (\ref{eq:ecdg}) which requires only the specification of $\bm{f}_S(\bm{u}_L,\bm{u}_R)$ and local matrices.
@@
 \[
-\LRa{u,\Pi_Nv}_{\partial D^k} = \LRp{Lu,\Pi_N v}_{D^k} = \LRp{Lu,v}_{D^k}, \qquad u\in V_h, \qquad v\not\in V_h.  
-\]
-Then, the global DG differentiation operator can be rewritten as 
+\LRa{u,\Pi_N(vw)}_{\partial D^k} = \LRp{Lu,\Pi_N (vw)}_{D^k} = \LRp{Lu,vw}_{D^k}, \qquad u,v\in V_h, 
+\]
+where $w(x)$ is smooth and bounded over $D^k$.  Then, the global DG operator $D^x_h$ can be rewritten as 
 \begin{equation}
-\LRp{D^x_h u,v}_{\Omega} = \sum_k \LRp{\pd{ \Pi_N u}{x} + \frac{1}{2}L\LRp{u-\Pi_N u}\bm{n}_x,v}_{D^k} + \frac{1}{2}{\LRa{{u^+ - \Pi_N u}, v\bm{n}_x}_{\partial D^k}}.
+\LRp{D^x_h u,v}_{\Omega} = \sum_k \LRp{\pd{ \Pi_N u}{x} + \frac{1}{2}L\LRp{u-\Pi_N u}{n}_x,v}_{D^k} + \frac{1}{2}{\LRa{{u^+ - \Pi_N u}, v{n}_x}_{\partial D^k}}.
 \label{eq:dgd_globalimplement}
 \end{equation}
-
 Applying Assumption~\ref{eq:assumption}, we can rewrite 
@@
 \LRp{\left.\LRp{D^x_h \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},\bm{w}}_{\Omega} &= \sum_k \sum_{i=1}^{\infty} \LRp{\pd{\Pi_N \bm{f}_i(\bm{u}_{\bm{v}})}{x},\bm{g}_i(\bm{u}_{\bm{v}})\bm{w}}_{D^k}\\
-&+  \sum_k \sum_{i=1}^{\infty}\frac{1}{2}\LRa{\bm{f}_i(\bm{u}^+_{\bm{v}})\bm{g}_i(\bm{u}_{\bm{v}}) - \bm{g}_i(\bm{u}_{\bm{v}})\Pi_N\bm{f}_i(\bm{u}_{\bm{v}}),\bm{w}\bm{n}_x}_{\partial D^k}\\
-&+  \sum_k \sum_{i=1}^{\infty}\frac{1}{2}\LRp{\bm{g}_i(\bm{u}_{\bm{v}})L\LRp{\bm{f}_i(\bm{u}^+_{\bm{v}}) - \Pi_N\bm{f}_i(\bm{u}_{\bm{v}})}\bm{n}_x,\bm{w}}_{D^k},
+&+  \sum_k \sum_{i=1}^{\infty}\frac{1}{2}\LRa{\bm{f}_i(\bm{u}^+_{\bm{v}})\bm{g}_i(\bm{u}_{\bm{v}}) - \bm{g}_i(\bm{u}_{\bm{v}})\Pi_N\bm{f}_i(\bm{u}_{\bm{v}}),\bm{w}{n}_x}_{\partial D^k}\\
+&+  \sum_k \sum_{i=1}^{\infty}\frac{1}{2}\LRp{\bm{g}_i(\bm{u}_{\bm{v}})L\LRp{\bm{f}_i(\bm{u}^+_{\bm{v}}) - \Pi_N\bm{f}_i(\bm{u}_{\bm{v}})}{n}_x,\bm{w}}_{D^k},
 \end{align*}
-where $\bm{w}\bm{n}_x$ denotes component-wise multiplication of $\bm{w}$ by $\bm{n}_x$, the $x$-component of the unit outward normal.
-Moving the sum inside and simplifying then allows us to re-write the resulting expression as
+where $\bm{w}{n}_x$ denotes component-wise multiplication of $\bm{w}$ by ${n}_x$, the $x$-component of the unit outward normal.  Moving the sum inside and simplifying then reduces this expression to
 \begin{align}
 \LRp{\left.\LRp{D^x_h \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},\bm{w}}_{\Omega} &= \sum_k \LRp{\left.\pd{\Pi_N \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}{x}\right|_{y=x},\bm{w}}_{D^k}\nonumber\\
-&+ \frac{1}{2}\LRa{\bm{f}_S(\bm{u}^+_{\bm{v}},\bm{u}_{\bm{v}}) - \left.\Pi_N^x\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))\right|_{y=x},\bm{w}\bm{n}_x}_{\partial D^k} \label{eq:eval}\\
-&+ \frac{1}{2}\LRp{\left.\LRp{L^x\LRp{\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y)) - \Pi_N^x\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\bm{n}_x}\right|_{y=x},\bm{w}}_{D^k},\nonumber
+&+ \frac{1}{2}\LRa{\bm{f}_S(\bm{u}^+_{\bm{v}},\bm{u}_{\bm{v}}) - \left.\Pi_N^x\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))\right|_{y=x},\bm{w}{n}_x}_{\partial D^k} \label{eq:eval}\\
+&+ \frac{1}{2}\LRp{\left.\LRp{L^x\LRp{\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y)) - \Pi_N^x\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}{n}_x}\right|_{y=x},\bm{w}}_{D^k},\nonumber
 \end{align}
-where the superscripts $L^x, \Pi_N^x$ are used to emphasize that the lift and projection operators are applied in $x$ and not in $y$.  
+where the superscripts $L^x, \Pi_N^x$ are used to emphasize that the lift and volume projection operators are applied in $x$ but not in $y$.  
 This can be used to practically evaluate $\left.\LRp{\bm{D}^x_h \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x}$ using only reference operators and evaluations of $\bm{f}_S$.
-However, unlike previous entropy conservative diagonal norm SBP-DG methods, the boundary terms in (\ref{eq:eval}) interact with volume terms through the $L^2$ projection and the lift operator.
-This can be simplified for diagonal norm SBP operators with boundary nodes, using equivalences between $L^2$ projection and interpolation at nodal points in those cases.
-
-Finally, we discuss how to recover a computational form of flux differencing analogous to the approaches presented in \cite{gassner2016split, chen2017entropy}.
+We note that in general, the boundary terms in (\ref{eq:eval}) interact with volume terms through the $L^2$ projection and the lift operator.
+This can be simplified for diagonal norm SBP operators with boundary nodes, using the equivalence between $L^2$ projection and interpolation at nodal points for diagonal norm nodal SBP operators.
+
+We now discuss how to recover a computational form of flux differencing analogous to the approaches presented in \cite{gassner2016split, chen2017entropy}.
 We will describe how to compute the three non-standard terms which appear within (\ref{eq:eval}):
 \begin{align}
-\LRp{\left.\pd{\Pi_N {f}_S({u}(x),{u}(y))}{x}\right|_{y=x},{w}}_{D^k}\label{eq:voleval}\\
-\LRa{{f}_S({u}^+,{u})-\left.\Pi_N^x{f}_S({u}(x),{u}(y))\right|_{y=x},{w}\bm{n}_x}_{\partial D^k}\label{eq:surfeval}\\
-\LRp{\left.\LRp{L^x\LRp{{f}_S({u}(x),{u}(y)) - \Pi_N^x{f}_S({u}(x),{u}(y))}\bm{n}_x}\right|_{y=x},{w}}_{D^k}\label{eq:volsurfeval}
+\LRp{\left.\pd{\Pi_N {f}_S({u}(x),{u}(y))}{x}\right|_{y=x},{w}}_{\widehat{D}}\label{eq:voleval}\\
+\LRa{{f}_S({u}^+,{u})-\left.\Pi_N^x{f}_S({u}(x),{u}(y))\right|_{y=x},{w}{n}_x}_{\partial \widehat{D}}\label{eq:surfeval}\\
+\LRp{\left.\LRp{L^x\LRp{{f}_S({u}(x),{u}(y)) - \Pi_N^x{f}_S({u}(x),{u}(y))}{n}_x}\right|_{y=x},{w}}_{\widehat{D}} \label{eq:volsurfeval}.
 \end{align}
-where we have assumed a scalar variable $u(x)$ and flux function $f_S$ for simplicity.
-
-We first introduce matrices defined over the reference element $\widehat{D} = [-1,1]$.
-Let $\bm{W}$ denote the diagonal matrix whose entries are quadrature weights.
+where we have restricted ourselves to the reference element $\widehat{D}$ and a scalar variable $u(x)$ and flux function $f_S$ for simplicity.
+Extending to vector variables and flux functions requires repeating the same procedures component-by-component, and extending to affinely mapped elements $D^k$ involves scaling by constant geometric factors.
+
+We first introduce matrices defined over the reference element $\widehat{D} = [-1,1]$.  Let $\bm{W} \in \mathbb{R}^{N_q\times N_q}$ denote the diagonal matrix whose entries are quadrature weights
+\[
+\bm{W}_{ij} = \begin{cases}
+w_i & i=j\\
+0 & \text{otherwise}.
+\end{cases}
+\]
 Then, assuming $u(x) \in P^N\LRp{\widehat{D}}$, it can be represented in some polynomial basis $\phi_j(\widehat{x})$ of degree $N$ and dimension $N_p$ 
 \[
-V_h = {\rm span}\LRc{\phi_i(\widehat{x})}_{i=1}^{N_p}.
+P^N\LRp{\widehat{D}} = {\rm span}\LRc{\phi_i(\widehat{x})}_{i=1}^{N_p}.
 \]
@@
 \]
-which maps a function (in terms of its evaluation at quadrature points) to coefficients of the basis $\phi_j(x)$.  Next, let $\bm{D}_x$ denote the differentiation matrix defined implicitly through
-\[
-u(x) = \sum_{j=1}^{N_p} \bm{u}_j \phi_j(\widehat{x}), \qquad \pd{u}{\widehat{x}} = \sum_{j=1}^{N_p} \LRp{\bm{D}_x \bm{u}}_j\phi_j(\widehat{x})
-\]
-In other words, $\bm{D}_x$ maps basis coefficients of some function $u \in V_h$ to coefficients of its derivative on $[-1,1]$.
+which maps a function (in terms of its evaluation at quadrature points) to coefficients of the basis $\phi_j(x)$.
+Next, let $\widehat{\bm{D}}$ denote the differentiation matrix defined implicitly through
+\[
+u(x) = \sum_{j=1}^{N_p} \bm{u}_j \phi_j(\widehat{x}), \qquad \pd{u}{\widehat{x}} = \sum_{j=1}^{N_p} \LRp{\widehat{\bm{D}} \bm{u}}_j\phi_j(\widehat{x})
+\]
+In other words, $\widehat{\bm{D}}$ maps basis coefficients of some function $u \in V_h$ to coefficients of its derivative on $[-1,1]$.
 
@@
 \]
-where we have assumed continuity in time, and have used that $\bm{W}$ is diagonal.  }  Let $\bm{F}_S$ denote the $N_q\times N_q$ matrix constructed by evaluating $f_S$ at volume quadrature points
+where we have assumed continuity in time, and have used that $\bm{W}$ is diagonal.
+}  We note that this procedure only evaluates the variational formulation.
+To use this evaluation within an explicit time-stepping scheme, the additional inversion of a local mass matrix is required to compute the right hand side contribution over each element.
+
+Let $\bm{F}_S$ denote the $N_q\times N_q$ matrix constructed by evaluating $f_S$ at volume quadrature points
 \[
@@
 \[
-\LRp{\left.\pd{\Pi_N {f}_S({u}(x),{u}(y))}{x}\right|_{y=x},{w}}_{D^k} = \bm{w}^T \bm{V}_q^T\bm{W} {\rm diag}\LRp{\bm{V}_q\bm{D}_x\bm{P}_q \bm{F}_S},  
+\LRp{\left.\pd{\Pi_N {f}_S({u}(x),{u}(y))}{x}\right|_{y=x},{w}}_{D^k} = \bm{w}^T \bm{V}_q^T{\rm diag}\LRp{\bm{W} \bm{V}_q\widehat{\bm{D}}\bm{P}_q \bm{F}_S},  
 \]
@@
 \[
-{\rm diag}\LRp{\bm{V}_q \bm{D}_q \bm{P}_q \bm{F}_S} = \LRp{\LRp{\bm{V}_q \bm{D}_q \bm{P}_q} \circ \bm{F}_S}\bm{1}.
-\]
-A computational advantage of evaluation in terms of the Hadamard product is that $\bm{F}_S$ does not need to be formed explicitly and stored.
+{\rm diag}\LRp{\bm{W} \bm{V}_q \bm{D}_q \bm{P}_q \bm{F}_S} = \LRp{\LRp{\bm{W} \bm{V}_q \bm{D}_q \bm{P}_q} \circ \bm{F}_S}\bm{1}.
+\]
+A computational advantage of evaluation in terms of the Hadamard product is that $\bm{F}_S$ does not need to be formed explicitly.
 
@@
 We focus first on (\ref{eq:surfeval}) and introduce matrices used to compute boundary terms.
-Let $\bm{V}_f$ denote the matrix which interpolates to boundary values\footnote{In generalizing to higher dimensions, $\bm{V}_f$ would be the matrix which interpolates to boundary points.}
+Let $\bm{V}_f$ denote the matrix which interpolates to boundary values\footnote{In generalizing to higher dimensions, $\bm{V}_f$ would be the matrix which interpolates to quadrature points on the boundary.}
 \[
@@
 In one dimension, this is simply the $2\times 2$ identity matrix.
-We now define $\bm{F}_S^f$ to be the matrix constructed by evaluating $\bm{F}_S$ at both volume quadrature and boundary points
+We now define $\bm{F}_S^f$ to be the matrix constructed by evaluating $\bm{F}_S$ at both volume quadrature points $x_i$ and boundary points $x^f_j$
 \[
@@
 \]
-We also define $\bm{F}^+_S$ to be the vector of evaluations of the flux function $f_S\LRp{u^+,u}$ at the boundary points of $D^k$.  Using these matrices, we can then evaluate 
+We also define $\bm{F}^+_S$ to be the vector of evaluations of the flux function $f_S\LRp{u^+,u}$ at the boundary points of $D^k$.
+Using these matrices and properties of Hadamard products, we can then evaluate
 \begin{align*}
-\LRa{{f}_S({u}^+,{u})-\left.\Pi_N^x{f}_S({u}(x),{u}(y))\right|_{y=x},{w}\bm{n}_x}_{\partial D^k} &= \bm{w}^T\bm{V}_f^T \bm{W}_f {\rm diag}(\bm{n}_x)\LRp{\bm{F}_S^+ - {\rm diag}\LRp{\bm{V}_f\bm{P}_q \bm{F}^f_S}} \\
-&= \bm{w}^T\bm{V}_f^T \bm{W}_f {\rm diag}(\bm{n}_x)\LRp{\bm{F}_S^+ - \LRp{\LRp{\bm{V}_f\bm{P}_q} \circ \LRp{\bm{F}^f_S}^T}\bm{1}}.
+\LRa{{f}_S({u}^+,{u})-\left.\Pi_N^x{f}_S({u}(x),{u}(y))\right|_{y=x},{w}{n}_x}_{\partial D^k} &= \bm{w}^T\bm{V}_f^T \bm{W}_f {\rm diag}({n}_x)\LRp{\bm{F}_S^+ - {\rm diag}\LRp{\bm{V}_f\bm{P}_q \bm{F}^f_S}} \\
+&= \bm{w}^T\bm{V}_f^T \bm{W}_f {\rm diag}({n}_x)\LRp{\bm{F}_S^+ - \LRp{\LRp{\bm{V}_f\bm{P}_q} \circ \LRp{\bm{F}^f_S}^T}\bm{1}}.
 \end{align*}
@@
 \begin{align*}
-&\LRp{\left.\LRp{L^x\LRp{{f}_S({u}(x),{u}(y)) - \Pi_N^x{f}_S({u}(x),{u}(y))}\bm{n}_x}\right|_{y=x},{w}}_{D^k} \\
-&= \bm{w}^T\bm{V}_q^T\bm{W}{\rm diag}\LRp{\bm{V}_q\bm{L}{\rm diag}(\bm{n}_x) \LRp{\LRp{\bm{F}_S^f}^T - \bm{V}_f\bm{P}_q \bm{F}_S^f}}\\
-&= \bm{w}^T\bm{V}_q^T\bm{W}\LRp{ \LRp{\bm{V}_q\bm{L}{\rm diag}(\bm{n}_x)} \circ {\bm{F}_S^f} - \LRp{\bm{V}_q\bm{L}{\rm diag}(\bm{n}_x)\bm{V}_f\bm{P}_q}\circ {\bm{F}_S}}\bm{1}.
+&\LRp{\left.\LRp{L^x\LRp{{f}_S({u}(x),{u}(y)) - \Pi_N^x{f}_S({u}(x),{u}(y))}{n}_x}\right|_{y=x},{w}}_{D^k} \\
+&= \bm{w}^T\bm{V}_q^T\bm{W}{\rm diag}\LRp{\bm{V}_q\bm{L}{\rm diag}({n}_x) \LRp{\LRp{\bm{F}_S^f}^T - \bm{V}_f\bm{P}_q \bm{F}_S^f}}\\
+&= \bm{w}^T\bm{V}_q^T\bm{W}\LRp{ \LRp{\bm{V}_q\bm{L}{\rm diag}({n}_x)} \circ {\bm{F}_S^f} - \LRp{\bm{V}_q\bm{L}{\rm diag}({n}_x)\bm{V}_f\bm{P}_q}\circ {\bm{F}_S}}\bm{1}.
 \end{align*}
-This expression is (in a loose sense) analogous to the evaluation of (\ref{eq:surfeval}) in that it couples together volume and boundary evaluations of the flux function, but in the reverse order.
-
-It is worth pointing out that these sets of expressions can be simplified; for example, straightforward computations show that $\bm{W}\bm{V}_q\bm{L} = \LRp{\bm{W}_f{\rm diag}(\bm{n}_x)\bm{V}_f\bm{P}_q}^T$, and the computation of the volume term $\LRp{\LRp{\bm{V}_q \bm{D}_q \bm{P}_q} \circ \bm{F}_S}\bm{1}$ and the surface term $\LRp{\LRp{\bm{V}_q\bm{L}{\rm diag}(\bm{n}_x)\bm{V}_f\bm{P}_q}\circ {\bm{F}_S}}\bm{1}$ can be combined into one single operation.
+This expression is similar to the evaluation of (\ref{eq:surfeval}) in that it couples together volume and boundary evaluations of the flux function, but in the reverse order.
+
+It is worth pointing out that these sets of expressions can be simplified; for example, straightforward computations show that the $\bm{W}\bm{V}_q\bm{L} = \LRp{\bm{W}_f{\rm diag}({n}_x)\bm{V}_f\bm{P}_q}^T$, and the computation of the volume term $\LRp{\LRp{\bm{V}_q \bm{D}_q \bm{P}_q} \circ \bm{F}_S}\bm{1}$ and the surface term $\LRp{\LRp{\bm{V}_q\bm{L}{\rm diag}({n}_x)\bm{V}_f\bm{P}_q}\circ {\bm{F}_S}}\bm{1}$ can be combined into one single operation.
 Furthermore, for specific choices of basis and quadrature, one can eliminate certain terms altogether.
@@
 All numerical experiments utilize the fourth order five-stage low-storage Runge-Kutta method Carpenter and Kennedy \cite{carpenter1994fourth}.
-Finally, following the derivation of stable timestep restrictions in \cite{chan2015gpu}, we define the timestep $\Delta t$ to be
+Following the derivation of stable timestep restrictions in \cite{chan2015gpu}, we define the timestep $\Delta t$ to be
 \[
@@
 where $C_N$ is the one-dimensional constant in the trace inequality \cite{warburton2003constants}, and ${\rm CFL}$ is a user-defined constant.
+
+We note that the numerical implementations used here are oblivious to the choice of basis, and the discretization is specified completely by the choice of quadrature.
+For example, when GLL quadratures are used, an entropy conservative/stable DG-SEM discretization is recovered \cite{carpenter2014entropy, chen2017entropy, gassner2017br1}, and when a Gauss quadrature with $(N+1)$ points is used, generalized SBP-DG methods are recovered \cite{ranocha2017extended, ranocha2017comparison}.
 
@@
 
-The one-dimensional compressible Euler equations correspond to the inviscid limit of the compressible Navier-Stokes equations and are given as follows:
+The one-dimensional compressible Euler equations, which correspond to the inviscid limit of the compressible Navier-Stokes equations, are given as follows:
 \begin{align*}
@@
 \begin{align*}
-q_1 = \frac{\gamma-s}{\gamma-1} - \frac{\rho u^2}{2p}, \qquad q_2 = \frac{\rho u}{p}, \qquad q_3 = -\frac{\rho}{p}.
+v_1 = \frac{\gamma-s}{\gamma-1} - \frac{\rho u^2}{2p}, \qquad v_2 = \frac{\rho u}{p}, \qquad v_3 = -\frac{\rho}{p}.
 \end{align*}
@@
 \[
-\rho = -(\rho e) q_3, \qquad m = (\rho e) v_2, \qquad E = (\rho e)\LRp{1 - \frac{q_2^2}{2 q_3}},
+\rho = -(\rho e) v_3, \qquad m = (\rho e) v_2, \qquad E = (\rho e)\LRp{1 - \frac{v_2^2}{2 v_3}},
 \]
@@
 \[
-\rho e = \LRp{\frac{(\gamma-1)}{\LRp{-q_3}^{\gamma}}}^{1/(\gamma-1)}e^{\frac{-s}{\gamma-1}}, \qquad s = \gamma - q_1 + \frac{q_2^2}{2q_3}.
+\rho e = \LRp{\frac{(\gamma-1)}{\LRp{-v_3}^{\gamma}}}^{1/(\gamma-1)}e^{\frac{-s}{\gamma-1}}, \qquad s = \gamma - v_1 + \frac{v_2^2}{2v_3}.
 \]
@@
 \end{equation}
-This can be achieved using positivity-preserving limiters \cite{zhang2010positivity, zhang2012maximum}.
-These were not found to be necessary for the experiments in this work, and have thus not been implemented in our numerical simulations.
+This can be achieved using positivity-preserving limiters \cite{zhang2010positivity, zhang2012maximum}.  However, these have not been implemented in our numerical simulations.
 
@@
 \]
-We note that, because the direct evaluation of the logarithmic mean is numerically sensitive for $u_L\approx u_R$, we switch to evaluation using a high order accurate expansion introduced by Ismail and Roe \cite{ismail2009affordable} when $\LRb{u_L-u_R}<\epsilon$.
+We note that, because the direct evaluation of the logarithmic mean is numerically sensitive for $u_L\approx u_R$, when $\LRb{u_L-u_R}<\epsilon$ we switch to evaluation using a high order accurate expansion introduced by Ismail and Roe \cite{ismail2009affordable}.
 
@@
 \]
+We will refer to the combination of the entropy conservative flux with Lax-Friedrichs dissipation as the ``Lax-Friedrichs'' flux.
+
+\subsubsection{Smooth entropy wave solution}
 
@@
 \pgfplotsset{
-cycle list={
-{blue, mark=*}, {red, mark=square*}
-}
+cycle list={{blue, dashed, mark=*}, {red, mark=square*}}
 }
@@
 
-\addplot+[semithick, mark options={fill=markercolor}]
+\addplot+[semithick, mark options={solid, fill=markercolor}]
 % N = 1, tau = 0.000000 =======================
 coordinates{(0.5,1)(0.25,0.485059)(0.125,0.203599)(0.0625,0.0947163)(0.03125,0.0463705)};
-\addplot+[semithick, mark options={fill=markercolor}]
+\addplot+[semithick, mark options={solid, fill=markercolor}]
 %N = 1, tau = 0.000000 =======================
@@
 
-\addplot+[semithick, mark options={fill=markercolor}]
+\addplot+[semithick, mark options={solid, fill=markercolor}]
 % N = 2, tau = 0.000000 =======================
 coordinates{(0.5,0.746606)(0.25,0.156701)(0.125,0.0137392)(0.0625,0.000701926)(0.03125,8.64531e-05)};
-\addplot+[semithick, mark options={fill=markercolor}]
+\addplot+[semithick, mark options={solid, fill=markercolor}]
 %N = 2, tau = 0.000000 =======================
@@
 
-\addplot+[semithick, mark options={fill=markercolor}]
+\addplot+[semithick, mark options={solid, fill=markercolor}]
 % N = 3, tau = 0.000000 =======================
 coordinates{(0.5,0.103299)(0.25,0.00829887)(0.125,0.00073573)(0.0625,9.05975e-05)(0.03125,1.13596e-05)};
-\addplot+[semithick, mark options={fill=markercolor}]
+\addplot+[semithick, mark options={solid, fill=markercolor}]
 %N = 3, tau = 0.000000 =======================
@@
 
-\addplot+[semithick, mark options={fill=markercolor}]
+\addplot+[semithick, mark options={solid, fill=markercolor}]
 % N = 4, tau = 0.000000 =======================
 coordinates{(0.5,0.0385542)(0.25,0.00133048)(0.125,0.000176663)(0.0625,1.64135e-06)(0.03125,2.66024e-08)};
-\addplot+[semithick, mark options={fill=markercolor}]
+\addplot+[semithick, mark options={solid, fill=markercolor}]
 %N = 4, tau = 0.000000 =======================
@@
 
-\addplot+[semithick, mark options={fill=markercolor}]
+\addplot+[semithick, mark options={solid, fill=markercolor}]
 % N = 5, tau = 0.000000 =======================
 coordinates{(0.5,0.00436131)(0.25,0.00039846)(0.125,2.1282e-06)(0.0625,4.49046e-08)(0.03125,9.99912e-10)};
-\addplot+[semithick, mark options={fill=markercolor}]
+\addplot+[semithick, mark options={solid, fill=markercolor}]
 %N = 5, tau = 0.000000 =======================
@@
 ] 
-\pgfplotsset{
-cycle list={
-{blue, mark=*}, {red, mark=square*}
-}
-}
-%\addlegendimage{no markers,blue}
-%\addlegendimage{no markers,red}
-
-\addplot+[semithick, mark options={fill=markercolor}]
+%\pgfplotsset{cycle list={{blue, dashed, mark=*}, {red, mark=square*}}}
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
 % N = 1, tau = 0.500000 =======================
 coordinates{(0.5,1)(0.25,0.4932)(0.125,0.183839)(0.0625,0.0562398)(0.03125,0.0151873)};
-\addplot+[semithick, mark options={fill=markercolor}]
+\addplot+[semithick, mark options={solid, fill=markercolor}]
 %N = 1, tau = 0.500000 =======================
@@
 
-\addplot+[semithick, mark options={fill=markercolor}]	
+\addplot+[semithick, mark options={solid, fill=markercolor}]	
 % N = 2, tau = 0.500000 =======================
 coordinates{(0.5,0.336817)(0.25,0.04941)(0.125,0.00605428)(0.0625,0.000748842)(0.03125,9.25456e-05)};
-\addplot+[semithick, mark options={fill=markercolor}]
+\addplot+[semithick, mark options={solid, fill=markercolor}]
 %N = 2, tau = 0.500000 =======================
@@
 % N = 3, tau = 0.500000 =======================
-\addplot+[semithick, mark options={fill=markercolor}]
+\addplot+[semithick, mark options={solid, fill=markercolor}]
 coordinates{(0.5,0.0463242)(0.25,0.00408748)(0.125,0.000346831)(0.0625,1.99064e-05)(0.03125,1.22357e-06)};
-\addplot+[semithick, mark options={fill=markercolor}]
+\addplot+[semithick, mark options={solid, fill=markercolor}]
 %N = 3, tau = 0.500000 =======================
@@
 
-\addplot+[semithick, mark options={fill=markercolor}]
+\addplot+[semithick, mark options={solid, fill=markercolor}]
 % N = 4, tau = 0.500000 =======================
 coordinates{(0.5,0.0100716)(0.25,0.000625923)(0.125,1.89866e-05)(0.0625,7.03865e-07)(0.03125,2.10265e-08)};
-\addplot+[semithick, mark options={fill=markercolor}]
+\addplot+[semithick, mark options={solid, fill=markercolor}]
 %N = 4, tau = 0.500000 =======================
@@
 
-\addplot+[semithick, mark options={fill=markercolor}]
+\addplot+[semithick, mark options={solid, fill=markercolor}]
 % N = 5, tau = 0.500000 =======================
 coordinates{(0.5,0.00356628)(0.25,8.73125e-05)(0.125,2.20528e-06)(0.0625,3.20127e-08)(0.03125,4.63639e-10)};
-\addplot+[semithick, mark options={fill=markercolor}]
+\addplot+[semithick, mark options={solid, fill=markercolor}]
 %N = 5, tau = 0.500000 =======================
@@
 \]
-We compute the $L^2$ error in the conservation variables 
+We compute the $L^2$ error in the conservative variables 
 \[
@@
 \]
-at final time $T = .7$ using both the entropy conservative and local Lax-Friedrichs fluxes.
-For these experiments, we compare two quadrature rules: the $(N+1)$ point Gauss-Legendre-Lobatto rule (which induces a DG-SEM discretization and is referred to in Figure~\ref{fig:convergence} as ``GLL'') and an $(N+2)$ point Gauss quadrature rule (referred to in the figure as ``GQ-$(N+2)$'').
-We do not compare against the $(N+1)$ point Gauss quadrature rule, which behaves very similarly to the GQ-$(N+2)$ case.
-The $L^2$ error is evaluated using a more accurate $N+5$ point Gauss quadrature rule, and a CFL of $.125$ is used.
+at final time $T = .7$ using both the entropy conservative and local Lax-Friedrichs fluxes and a CFL of .125.  For these experiments, we compare three quadrature rules: 
+\begin{enumerate}
+\item the $(N+1)$ point Gauss-Legendre-Lobatto rule (referred to as ``GLL''),
+\item a $(N+1)$ point Gauss quadrature rule (referred to as ``GQ-$(N+1)$''),
+\item an over-integrated $(N+2)$ point Gauss quadrature rule (referred to as ``GQ-$(N+2)$'').  
+\end{enumerate}
+The $L^2$ error is evaluated using a more accurate $N+5$ point Gauss quadrature rule.
 
@@
 
-Figure~\ref{fig:convergence} shows computed $L^2$ errors under mesh refinement for both GLL and GQ-$(N+2)$ quadrature rules.
-All experiments utilize the entropy conservative flux of Chandreshekar, and we compare errors with and without Lax-Friedrichs penalization.
+Figure~\ref{fig:convergence} shows computed $L^2$ errors under mesh refinement for both GLL and GQ-$(N+2)$ quadrature rules with and without Lax-Friedrichs penalization.
+We do not compare against the $(N+1)$ point Gauss quadrature rule or quadrature rules with $N_q > (N+2)$, as the $L^2$ errors are very similar to the GQ-$(N+2)$ case.
 Computed asymptotic convergence rates are reported in Table~\ref{tab:rates}.
-It can be observed that, for both GLL and GQ-$(N+2)$ quadratures, the entropy conservative flux exhibits suboptimal convergence rates for odd orders, while the inclusion of Lax-Friedrichs penalization restores exhibits the optimal $O(h^{N+1})$ convergence rate for both quadrature choices.
-
-Next, we examine the discrete evolution of entropy by evolving a discontinuous initial profile to final time $T=4$.  We initialize the density and velocity to be 
+It can be observed that, for both GLL and GQ-$(N+2)$ quadratures, the entropy conservative flux exhibits suboptimal convergence rates for odd orders, while the inclusion of Lax-Friedrichs penalization restores the optimal $O(h^{N+1})$ convergence rate for both quadrature choices.
+
+\subsubsection{Discontinuous profile on a periodic domain}
+
+Next, we examine the discrete evolution of entropy by evolving a discontinuous initial profile to final time $T=2$ on the domain $[-1,1]$.  We initialize the density and velocity to be 
 \begin{equation}
@@
 \end{equation}
-Non-physical periodic boundary conditions are enforced in order to examine the evolution of entropy over longer time periods.
-Figure~\ref{subfig:sol} shows $\rho, u$ at time $T = 1/10$ using both entropy conservative and Lax-Friedrichs fluxes (referred to as ``EC'' and ``LF'', respectively in the figure) and GQ-$(N+2)$ quadrature.
+Periodic boundary conditions are enforced in order to examine the evolution of entropy over longer time periods.
+Figure~\ref{subfig:sol} shows $\rho, u$ at time $T = 1/10$ using both entropy conservative and Lax-Friedrichs fluxes (referred to in the figure as ``EC'' and ``LF'', respectively) and GQ-$(N+2)$ quadrature.
 As expected, the entropy conservative flux results in spurious high frequency oscillations, which are significantly damped under Lax-Friedrichs penalization.
 
-We note that, while the proof of conservation of entropy holds at the semi-discrete level, it does not take into account the time discretization and thus does not hold at the fully discrete level.
-This is reflected into numerical observation that the change in entropy $\Delta U(t) = U\LRp{\bm{u}(t}) - U\LRp{\bm{u}(t)}$ increases as $t$ increases.
-However, numerical experiments in \cite{gassner2016well} suggest that the discrete change in entropy over time should converge to zero as the timestep decreases, which (under appropriate assumptions) implies that the fully discrete scheme converges to the semi-discrete scheme.
+We next examine the change in entropy over time.
+While the proof of conservation of entropy holds at the semi-discrete level, it does not take into account the time discretization and thus does not hold at the fully discrete level.
+This is reflected in the observation that the numerical change in entropy
+\[
+\Delta U(t) = U\LRp{\bm{u}(x,t)} - U\LRp{\bm{u}(x,0)}
+\]
+increases as $t$ increases.
+However, numerical experiments in \cite{gassner2016well} suggest that the discrete change in entropy over time should converge to zero as the timestep decreases.%, which (under appropriate assumptions) implies that the fully discrete scheme converges to the semi-discrete scheme.
 
 Figure~\ref{subfig:dS} shows the evolution of $\Delta U(t)$ to final time $T=2$ for both entropy conservative and Lax-Friedrichs fluxes at various CFL numbers.
-For the entropy conservative flux, we observe that $\Delta U(t)$ to decreases as the CFL and $\Delta t$ decrease.
+For the entropy conservative flux, we observe that $\Delta U(t)$ decreases as the CFL and timestep $\Delta t$ decrease.
 This is expected, since the discrete time problem should converge to the continuous semi-discrete problem (for which $\Delta U(t) = 0$) as $\Delta t\rightarrow 0$.
-We also observe that, as expected, $\Delta U(t)$ does not change significantly with CFL for the Lax-Friedrichs flux, indicating that the change in entropy is due to the effect of the dissipative flux rather than the time discretization.
+We also observe that $\Delta U(t)$ does not change significantly as a function of the CFL for the Lax-Friedrichs flux, indicating that the non-zero change in entropy in this case is due to the effect of the dissipative flux rather than the time discretization.
 
@@
 \subfloat[Solution at time $T = .1$]{
-\includegraphics[width=.475\textwidth]{figs/sol_ECLF.png}
+\includegraphics[width=.475\textwidth]{sol_ECLF.png}
 \label{subfig:sol}}
@@
 \subfloat[$\Delta U(t)$]{
-\includegraphics[width=.475\textwidth]{figs/dS_ECLF.png}
+\includegraphics[width=.475\textwidth]{dS_ECLF.png}
 \label{subfig:dS}}
@@
 
-We also numerically evaluate the spatial formulation tested against the projected the entropy variables
+We also numerically evaluate the spatial formulation tested against the projected entropy variables
 \[
@@
 In practice, we have found that $\delta_{\max}$ depends on the tolerance $\epsilon$ used in evaluation of the logarithmic mean.
-For a simulation to time $T=4$ using $N=4$ and $K = 16$ with a CFL of $1/2$, we observe that using $\epsilon = 10^{-2}$ (as recommended in \cite{ismail2009affordable} results in $\delta_{\max} = O\LRp{10^{-10}}$.
+For a simulation to time $T=4$ using $N=4$ and $K = 16$ with a CFL of $1/2$, we observe that using $\epsilon = 10^{-2}$ (as recommended in \cite{ismail2009affordable}) results in $\delta_{\max} = O\LRp{10^{-10}}$.
 Decreasing $\epsilon$ to $10^{-3}$ reduces $\delta_{\max}$ to $O\LRp{10^{-14}}$, and decreasing $\epsilon$ further to $1\times 10^{-4}$ reduces $\delta_{\max}$ to $10^{-15}$.
@@
 \centering
-\includegraphics[width=.5\textwidth]{figs/rhstest.png}
+\includegraphics[width=.525\textwidth]{rhstest.png}
 \caption{Comparison of $\delta(t)$ when evaluating the flux function $\bm{f}_S$ using conservative variables and projected entropy variables.}
@@
 \end{figure}
-Finally, in order to illustrate the importance of evaluating the flux in terms of the projected entropy variables $\Pi_N \bm{v}$, we compute $\delta(t)$ while evaluating the flux function $\bm{f}_S$ directly in terms of the conservative variables $\bm{u}(x)$ and in terms of $\bm{u}\LRp{(\Pi_N \bm{v})(x)}$.
+Next, in order to illustrate the importance of evaluating the flux in terms of the projected entropy variables $\Pi_N \bm{v}$, we compute $\delta(t)$ while evaluating the flux function $\bm{f}_S$ directly in terms of the conservative variables $\bm{u}(x)$ and in terms of the projected entropy variables $\bm{u}\LRp{(\Pi_N \bm{v})(x)}$.
 It can be observed from Figure~\ref{fig:rhstest} that $\delta(t)$ is near machine precision when evaluating the flux in terms of the projected entropy variables.
-When evaluating the flux function using the conservative variables directly, $\delta(t)$ begins near $10^{-4}$ but grows steadily, blowing up exponentially near $t = 1$.
+When evaluating the flux function directly in terms of the conservative variables, $\delta(t)$ begins near $10^{-4}$ but grows steadily, blowing up exponentially near $t = 1$.
+
+\subsubsection{Sod shock tube}
+
+We now examine the behavior of the proposed DG methods for some common one-dimensional test problems.
+We begin with the Sod shock tube, which is posed on the domain $[-1/2,1/2]$ with initial conditions
+\[
+\rho = \begin{cases}
+1 & x < 0\\
+.125 & x \geq 0,
+\end{cases} 
+\qquad
+u = 0, \qquad
+p = \begin{cases}
+1 & x < 0\\
+.1 & x \geq 0.
+\end{cases}
+\]
+Boundary conditions are enforced by taking the external value $\bm{u}^+$ in the numerical flux to be that of the initial condition at $x = \pm 1$.
+The solution develops a left-moving rarefaction, as well as a right moving shock wave and a contact discontinuity.
+We simulate the solution until time $T = .2$, without the use of positivity preserving or TVD-type limiters.
+For all choices of quadrature tested, the solution diverges when using the entropy conservative flux, which is a result of oscillations in the solution and density and temperature becoming negative.
+This is remedied when using the dissipative Lax-Friedrichs flux, for which we do not observe blowup of the solution.
+\begin{figure}
+\centering
+\subfloat[GLL quadrature]{\includegraphics[width=.475\textwidth]{sodGLL.png}}
+\hspace{.1em}
+\subfloat[GQ-$(N+2)$ quadrature]{\includegraphics[width=.475\textwidth]{sodGQ2.png}}
+\caption{Density and pressure solutions for the Sod shock tube at $T = .2$ for $N = 4$ and $K = 32$ elements.  Cell averages are overlaid as filled circles.  }
+\label{fig:sod}
+\end{figure}
+
+Figure~\ref{fig:sod} shows both the exact solution and the computed density and pressure along with their cell averages.
+These results were obtained using a CFL of $.125$ and GLL and GQ-$(N+2)$ quadrature rules.
+For both quadratures, the cell averages agree relatively well with the exact solution.
+Both solutions contain spurious oscillations, though the oscillations under GQ-$(N+2)$ quadrature appear qualitatively smoother and smaller.
+
+%\note{Add discussion of Figure~\ref{fig:sod} - reference solution is exact solution computed using a Riemann solver (``PRINCIPLES OF COMPUTATIONAL FLUID DYNAMICS'' by P. Wesseling).  }
+
+\subsubsection{Sine-shock interaction}
+\label{sec:sineshock}
+
+The next benchmark problem we consider is the sine-shock interaction problem, which is posed on domain $[-5,5]$ with initial conditions
+\[
+\rho(x,0) = \begin{cases}
+3.857143 & x < -4\\
+1 + .2\sin(5x) & x \geq -4,
+\end{cases} 
+\qquad
+u(x,0) = \begin{cases}
+2.629369 & x < -4\\
+0 & x \geq -4,
+\end{cases}
+\qquad
+p(x,0) = \begin{cases}
+10.3333 & x < -4\\
+1 & x \geq -4.
+\end{cases}
+\]
+We simulate the solution using different quadrature rules with $N=4$, $K=40$ elements, and a Lax-Friedrichs flux.
+No TVD or positivity-preserving limiters are applied.
+A smaller CFL of $.05$ is used, and is necessary to avoid solution divergence when using GQ-$(N+2)$ quadrature.
+It should be pointed out that, for GLL quadrature, it is possible to use a larger CFL of $.125$ without observing solution blowup.
+The reason for this discrepancy is the sensitivity of the evaluation $\bm{u}\LRp{\Pi_N \bm{v}}$ when $\Pi_N \bm{v}$ differs significantly from $\bm{v}$, and is described in more detail in Section~\ref{sec:instab}.
+\begin{figure}
+\centering
+%\subfloat[GLL quadrature]{\includegraphics[width=.475\textwidth]{sineShockGLLrhop.png}}
+\subfloat[GLL quadrature]{\includegraphics[width=.475\textwidth]{sineShockGLL.png}}
+\hspace{.1em}
+%\subfloat[GQ-$(N+2)$ quadrature]{\includegraphics[width=.475\textwidth]{sineShockGQ2rhop.png}}
+\subfloat[GQ-$(N+2)$ quadrature]{\includegraphics[width=.475\textwidth]{sineShockGQ2.png}}
+\caption{Density solutions for the sine-shock interaction problem at $T = 1.8$ for $N = 4$ and $K = 40$ elements.  Cell averages are overlaid as filled circles.  }
+\label{fig:sineshock}
+\end{figure}
+
+Figure~\ref{fig:sineshock} shows snapshots of the density at final time $T= 1.8$ for both GLL and GQ-$(N+2)$ quadrature, along with a reference solution computed using a 5th order WENO scheme with 25000 cells \cite{shu2009high}.
+For both GLL and GQ-$(N+2)$ quadrature, the cell averages are close to the reference solution, though the solutions still contain spurious oscillations resulting from the presence of shocks and discontinuities.
+However, as with the Sod shock tube, we observe that these oscillations are smoother and smaller in amplitude for the choice of GQ-$(N+2)$ quadrature.
+
+%\note{Add reference solution using 25000 cell 5th order WENO method \cite{shu2009high}.
+Simulations in Figure~\ref{fig:sineshock} use CFL of $.05$; necessary for GQ-$(N+2)$.
+GLL can still run at CFL of $.125$.
+}
+
+\subsubsection{Sensitivity of evaluation in terms of the projected entropy variables}
+\label{sec:instab}
+
+The numerical experiments in previous sections suggest that solutions computed using Gauss quadrature rules can be more accurate than those computed using GLL quadratures.
+However, we also observe that, for the sine-shock interaction problem, a larger CFL of $.125$ can be taken when using GLL quadrature, whereas a much smaller CFL of $.05$ is required to prevent solution blowup when using GQ-$(N+1)$ and GQ-$(N+2)$ quadratures.
+%The reason for this discrepancy between Gauss-Lobatto and Gauss quadratures is the presence of boundary nodes in GLL quadratures.
+For GQ-$(N+1)$ and GQ-$(N+2)$ quadratures, solution spikes can occur when evaluating the conservative variables in terms of the projected entropy variables at surface points.
+The use of GLL quadrature avoids this phenomena because of two factors: the equivalence between interpolation and projection under a $(N+1)$ point quadrature rule, and the presence of boundary points in GLL quadrature.
+
+Figure~\ref{fig:proj} shows snapshots of different variables for $N=2$ and $K = 20$ elements at the fifth Runge-Kutta stage of the first timestep (just prior to the detection of negative density and pressure values) using a GQ-$(N+2)$ quadrature rule.
+Discrepancies between the evaluated and projected values of the entropy variables $v_1, v_2, v_3$ at element boundaries are present, and while these discrepancies are not extremely large, they produce large spikes in the conservative variables due to the sensitivity of the nonlinear evaluation $\bm{u}\LRp{\Pi_N \bm{v}}$.
+In particular, because $v_3$ appears in the denominators of entropy and $\rho e$ (as functions of the entropy variables), values of $v_3$ near zero result in large values of density and energy.
+These spikes result in large oscillations in the solution if the CFL is too large, which eventually cause the density and pressure to become negative at quadrature or boundary points.
+%\footnote{For the sine-shock interaction problem, these spikes appear to be related to the large jump in the pressure initial condition.
+Instabilities were not observed for the sine-shock interaction when using an initial pressure condition with a smaller magnitude jump, or for other problems with discontinuous non-zero initial profiles, such as the modified Sod problem \cite{ranocha2017comparison}.
+}
+\begin{figure}
+\centering
+\subfloat[$v_1$ and $\Pi_N(v_1)$]{\includegraphics[width=.32\textwidth]{sineShockQ1Compare.png}}\label{subfig:v1}
+\hspace{.1em}
+\subfloat[$v_2$ and $\Pi_N(v_2)$]{\includegraphics[width=.32\textwidth]{sineShockQ2Compare.png}}\label{subfig:v2}
+\hspace{.1em}
+\subfloat[$v_3$ and $\Pi_N(v_3)$]{\includegraphics[width=.32\textwidth]{sineShockQ3Compare.png}}\label{subfig:v3}\\
+\subfloat[$\rho(x)$ and $\rho\LRp{\Pi_N \bm{v}}$]{\includegraphics[width=.32\textwidth]{sineShockDensityCompare.png}}\label{subfig:u1}
+\hspace{.1em}
+\subfloat[$u(x)$ and $u\LRp{\Pi_N \bm{v}}$]{\includegraphics[width=.32\textwidth]{sineShockVelocityCompare.png}}\label{subfig:u2}
+\hspace{.1em}
+\subfloat[$p(x)$ and $p\LRp{\Pi_N \bm{v}}$]{\includegraphics[width=.32\textwidth]{sineShockPressureCompare.png}}\label{subfig:u3}
+\caption{Top row: comparisons of entropy variables (evaluated directly as functions of conservative variables) and projected entropy variables (using $N=2$, $K=20$ elements, and a GQ-$(N+2)$ quadrature rule) for the sine-shock interaction before the detection of negative density and pressures.
+Bottom row: comparisons of the conservative variables and their evaluations in terms of the projected entropy variables.
+}
+\label{fig:proj}
+\end{figure}
+
+Qualitatively similar behavior is observed when using a GQ-$(N+1)$ quadrature rule, though negative density and pressure values occur at a slightly later time.
+In this case, the $L^2$ projection reduces to interpolation at interior Gauss points.
+This still implies that $\bm{u}\LRp{\Pi_N\bm{v}}$ does not necessarily agree with $\bm{u}$ at element boundaries, and this discrepancy again leads to spikes in the conservative variables similar to those presented in Figure~\ref{fig:proj}.
+These experiments suggest that the lack of control of boundary values in $\Pi_N \bm{v}$ leads to large oscillations or spikes in the boundary values of $\bm{u}\LRp{\Pi_N\bm{v}}$.
+The presence of boundary nodes in GLL quadratures avoids this issue: $\bm{u}\LRp{\Pi_N\bm{v}}$ and $\bm{u}$ agree at element boundaries due to the fact that quadrature-based $L^2$ projection under a GLL quadrature is equivalent to interpolation at GLL points.
+
+This scenario illustrates some of the more subtle differences between the use of GLL and GQ quadratures.
+We note that bound-preserving and TVD limiters \cite{zhang2010positivity, zhang2012maximum} are typically implemented in practice to detect and control spikes of the kind observed in these numerical experiments.
+However, slope-limiting the conservative variables $\bm{u}$ may still result in spikes being generated when evaluating $\bm{u}\LRp{\Pi_N\bm{v}}$ (unless the slope is set to zero such that $\bm{u}$ is constant).
+A more nuanced approach will likely require modifying the conservative variables $\bm{u}$ based on the projected entropy variables $\Pi_N \bm{v}$ to control oscillations while maintaining entropy conservation.
+These strategies will be the focus in future work.
 
@@
 
-This work presents a generalization of discretely entropy conservative methods from diagonal-norm SBP-DG methods to a more general class of high order DG methods which include dense norm SBP methods and over-integrated quadrature rules.
-The resulting DG methods satisfy a discrete conservation of entropy while maintaining a reasonable computational cost and small stencil.
-Numerical experiments indicate the resulting methods deliver high order accuracy for smooth solutions while displaying greatly improved stability for under-resolved and shock solutions.
-
-While entropy conservative and stability improve the robustness of solvers for nonlinear hyperbolic conservation laws, they do not address problems such as spurious oscillations in high order approximations of shock solutions or positivity preservation of density and pressure variables \cite{chen2017entropy}.
-These issues can be addressed through the use of regularization (e.g.\ filtering, artificial viscosity) and/or limiting; however, the goal in constructing discretely entropy stable methods is to avoid using excessive regularization and limiting in the pursuit of stability.
+This work presents a generalization of discretely entropy conservative methods from diagonal-norm SBP-DG methods to a more general class of high order DG methods, which include dense norm SBP operators and over-integrated quadrature rules.
+The resulting schemes satisfy a discrete conservation of entropy while maintaining a reasonable computational cost and small stencil.
+Numerical results indicate that the resulting methods deliver high order accuracy for smooth solutions while improving stability for under-resolved and shock solutions compared to non-entropy conservative and non-entropy stable solvers.
+We also describe differences between the sensitivity of methods based on Gauss-Lobatto quadrature and methods based on Gauss quadrature rules.
+
+We note that, while entropy conservative and entropy stable schemes improve the robustness of solvers for nonlinear hyperbolic conservation laws, they do not address problems such as spurious oscillations in high order approximations of shock solutions or positivity preservation of density and pressure variables \cite{chen2017entropy}.
+These issues can be addressed through the use of regularization (e.g.\ filtering, artificial viscosity) and/or limiting.
+However, this can lead to a reliance on regularization techniques as ad-hoc stabilization mechanisms.
+Addressing this issue requires constructing methods which avoid the need for excessive regularization and limiting in the pursuit of stability, and is a significant motivation for the development of entropy conservative and entropy stable discretizations.
 
 Finally, we note that the analysis and numerical experiments in this work are presented in one dimension for clarity and conciseness of presentation.
-The main concepts are directly extendable to higher dimensions and simplicial and pyramidal elements, which will be the focus of future work.
+This work is intended as a foundation on which higher dimensional schemes can be constructed.
+Future work will focus on generalizations to higher dimensions (including simplicial and pyramidal elements), as well as convergence and error analysis.
 
@@
 
-The author thanks Andrew Winters, David M.\ Williams, and Paul Hand for informative discussions.  Jesse Chan is supported by NSF awards DMS-1719818 and DMS-1712639.
+The author thanks Lucas Wilcox, Andrew Winters, David M.\ Williams, Weifeng Qiu, and Paul Hand for helpful discussions.
+Jesse Chan is supported by the National Science Foundation under awards DMS-1719818 and DMS-1712639.
 
@@
 \]
-These terms are analytic if $\avg{\beta}, \avg{\beta}^{\log}$ are bounded away from zero, which is satisfied if the density and pressure are bounded from below as assumed in (\ref{eq:assumption2}).
+These terms are analytic if $\rho, p, \avg{\beta}, \avg{\beta}^{\log}$ are bounded away from zero as assumed in (\ref{eq:assumption2}).
 Because the composition, sum, and product of analytic functions are also analytic, the Chandreshekar flux also is analytic under positivity assumptions.
 
-Finally, we note that the application of the operator $D_N$ to convergent infinite sums should be rigorously justified.  
-This requires $D_N$ to be a bounded operator from $H^1$ to $P^N$, which can be shown using boundedness of the $L^2$ projection operator and discrete inverse and trace inequalities \cite{chan2015gpu}.
+Finally, we note that the application of the operators $D_N, D^x_h$ to convergent infinite sums should be rigorously justified.  
+This requires $D_N, D^x_h$ to be bounded operators mapping from a broken Sobolev space to $P^N$ or $V_h$, which can be shown using boundedness of local $L^2$ projection operators and discrete inverse and trace inequalities \cite{chan2015gpu}.
 

### 2017-10-02 docs/dg_penalty/revision/reviewerResponse.tex

@@
 This illustrates that DG methods with upwind or penalty fluxes address both \textit{non-conforming} and \textit{conforming} spurious modes which are otherwise present both for central fluxes ($\tau = 0$) and conforming discretizations (or equivalently, for large penalty parameters $\tau\rightarrow \infty$).
+We have also added citations \cite{boffi2000problem,ainsworth2014dispersive}, which give a more in-depth description and analysis of spurious modes present for conforming discretizations.
 }
@@
 \item Title: the meaning of "first order discontinuous Galerkin formulations" is not very clear. The authors can consider to rephrase the title.
-\note{We have rearranged the title to specify the paper is on discontinuous Galerkin formulations for first order hyperbolic systems.}
+\note{We have rearranged the title to specify the paper is on the penalization stabilization mechanism in upwinded discontinuous Galerkin formulations for first order hyperbolic systems.}
 \item If in (2) the summation convention is used, state this.
@@
 \end{itemize}
+
+\bibliographystyle{unsrt}
+\bibliography{dgpenalty}
+
 \end{document}

### 2017-10-02 docs/dg_penalty/arxiv/main.tex

--- 2016-12-21 docs/dg_penalty/main.tex
@@
 \usepackage{fullpage}
-\usepackage{hyperref}
+%\usepackage{hyperref}
 %\let\proof\relax
@@
 
-High order discontinuous Galerkin (DG) methods are a popular choice of discretization for simulations of time-domain wave propagation, due to their low numerical dispersion and their ability to accomodate unstructured meshes \cite{grote2006discontinuous,hesthaven2007nodal}.
+High order discontinuous Galerkin (DG) methods are a popular choice of discretization for simulations of time-domain wave propagation, due to their low numerical dispersion and their ability to accomodate unstructured meshes \cite{riviere2008discontinuous,hesthaven2007nodal}.
 For PDEs in second order form, a common choice of DG method is the symmetric interior penalty (SIPG) discontinuous Galerkin method \cite{riviere2008discontinuous}, which involves a penalization parameter $\tau$.
@@
 \]
+Let $\bm{U}$ be the solution and $\bm{V}$ be a test function, and assume both are represented as piecewise polynomials over $\Oh$.
 A semi-discrete DG formulation for (\ref{eq:cons}) may be written as
 \begin{equation}
-\sum_{D^k \in \Oh} \LRp{\pd{\bm{U}}{t},\bm{V} }_{\Lk} = \sum_{D^k \in \Oh}\LRp{ \LRp{\bm{A}_i(\bm{U}), \bm{V}_{,i}}_{\Lk} - \LRa{(\bm{A}_n\bm{U})^*,\bm{V}}_{\partial D^k}}
+\sum_{D^k \in \Oh} \LRp{\pd{\bm{U}}{t},\bm{V} }_{\Lk} = \sum_{D^k \in \Oh}\LRp{ \LRp{\bm{A}_i(\bm{U}), \bm{V}_{,i}}_{\Lk} - \LRa{(\bm{A}_n\bm{U})^*,\bm{V}}_{\partial D^k}}, 
 \label{eq:formulation}
@@
 
-We note that, while the assumptions on constant $\bm{A}_i$ are restrictive, the results in this work are straightforward to generalize to a slight modification of (\ref{eq:cons})
+We note that, while the assumptions on constant $\bm{A}_i$ are relatively restrictive, the results in this work are straightforward to generalize to a slight modification of (\ref{eq:cons})
 \begin{equation}
@@
 We assume $\bm{A}_0 = \bm{I}$ in this work for brevity; however, all results are straightforward to extend to systems of the form for any $\bm{A}_0 = \bm{A}_0(\bm{x})$ which may vary spatially and is pointwise symmetric and positive-definite.
-Many models of wave propagation in general heterogeneous media can be represented in the form of (\ref{eq:consA}), including the acoustic wave equation \cite{chan2016weight1}, linear elasticity \cite{de2008interior,wilcox2010high,ye2016discontinuous}, and Maxwell's equations \cite{hesthaven2002nodal,hesthaven2004high,grote2007interior,warburton2013low}.
+Many models of wave propagation in general heterogeneous media can be represented in the form of (\ref{eq:consA}), including the acoustic wave equation \cite{chan2016weight1}, linear elasticity \cite{de2008interior,wilcox2010high,ye2016discontinuous}, and Maxwell's equations \cite{hesthaven2002nodal,grote2007interior,warburton2013low}.
 %Results also follow for matrices $\bm{A}_i$ may be symmetrized through an appropriate change of variables \cite{kopriva2014energy}.
@@
 \]
-Define the normal matrix 
+The normal matrix and penalty numerical flux are then given as
 \[
-{\bm{A}}_n = \sum_{i=1}^d {\bm{A}_i\bm{n}_i}.
+{\bm{A}}_n = \sum_{i=1}^d {\bm{A}_i\bm{n}_i}, \qquad 
+(\bm{A}_n\bm{U})^* = \bm{A}_n\avg{\bm{U}} - \frac{\tau}{2} \bm{A}_n^T \bm{A}_n\jump{\bm{U}}.
+%(\bm{A}_n\bm{U})^* = \frac{1}{2}\bm{A}_n\LRp{\bm{U}^+ + \bm{U}^-} - \frac{\tau}{2} \bm{A}_n^T \bm{A}_n\LRp{\bm{U}^+ - \bm{U}^-}.
 \]
-The penalty numerical flux is then defined as
-\[
-(\bm{A}_n\bm{U})^* = \bm{A}_n\avg{\bm{U}} - \frac{\tau}{2} \bm{A}_n^T \bm{A}_n\jump{\bm{U}}.
-\]
-where $\tau \geq 0$.
+where $\tau \geq 0$ and jumps and averages are taken component-wise.
 We compare the penalty flux to the well-known upwind numerical flux for the case of continuous coefficients.
@@
 \[
-\bm{A}_n = \bm{V}{\bm{\Lambda}}\bm{V}^T, \qquad \bm{\Lambda} = 
-\left(\begin{array}{ccc}
-\lambda_1 & & \\
-& \ddots & \\
-& & \lambda_d
-\end{array}\right).
+\bm{A}_n = \bm{V}{\bm{\Lambda}}\bm{V}^T, \qquad \bm{\Lambda}_{ii} = \lambda_i, \quad 1 \leq i \leq d.
 \]
@@
 where the matrices $\bm{A_n}^+,\bm{A_n}^-$ are constructed from the positive and negative eigenvalues 
-\begin{align*}
-\bm{A}_n^+ &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} + \LRb{\bm{\Lambda}}} \bm{V}^T\\
-\bm{A}_n^- &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} - \LRb{\bm{\Lambda}}} \bm{V}^T,
-\end{align*}
+\[
+\bm{A}_n^+ = \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} + \LRb{\bm{\Lambda}}} \bm{V}^T, \qquad \bm{A}_n^- = \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} - \LRb{\bm{\Lambda}}} \bm{V}^T,
+\]
 and $\LRb{\bm{\Lambda}}$ is the diagonal matrix whose entries consist of the absolute values of the eigenvalues $\LRb{\lambda_i}$.
@@
 (\bm{A}_n\bm{U})^* = \bm{A}_n\avg{\bm{U}} - \frac{1}{2}\bm{V}\LRb{\bm{\Lambda}}\bm{V}^T \jump{\bm{U}}.  
+%(\bm{A}_n\bm{U})^* = \frac{1}{2}\bm{A}_n\LRp{\bm{U}^++\bm{U}^-} - \frac{1}{2}\bm{V}\LRb{\bm{\Lambda}}\bm{V}^T \LRp{\bm{U}^+ - \bm{U}^-}.  
 \]
@@
 \]
-recovers the upwind fluxes for $\tau = 1$ and $\bm{W} = \bm{V}\LRb{\bm{\Lambda}}^{\dagger}\bm{V}^{-1}$, where $\bm{A}^{\dagger}$ is the pseudo-inverse of $\bm{A}$.
-However, the advantage of penalty fluxes is that, unlike upwind fluxes, they can be used when explicit knowledge of a diagonalization of $\bm{A}_n$ is not known, such as for linear elasticity in anisotropic media \cite{ye2016discontinuous}.
+recovers the upwind fluxes for $\tau = 1$ and $\bm{W} = \bm{V}\LRb{\bm{\Lambda}}^{\dagger}\bm{V}^{-1}$, where $\bm{\lambda}^{\dagger}$ is the pseudo-inverse.
+However, the advantage of penalty fluxes is that, unlike upwind fluxes, they can be used when explicit knowledge of a diagonalization of $\bm{A}_n$ is not known, such as for linear elasticity in general anisotropic media \cite{ye2016discontinuous}.
 Thus, we assume $\bm{W} = \bm{I}$ and require that $\bm{V}\LRb{\bm{\Lambda}}\bm{V}^T$ and $\tau\bm{A}_n^T\bm{A}_n$ are of the same magnitude, which implies that $\tau$ should be proportional to $1/\LRp{\max_i\LRb{\lambda_i}}$.
@@
 Next, we consider the acoustic wave equation in pressure-velocity form
-\begin{align*}
-\frac{1}{c^2}\pd{p}{t} &= \Div \bm{u}\\
-\pd{\bm{u}}{t} &= \Grad p,
-\end{align*}
+\[
+\frac{1}{c^2}\pd{p}{t} +\Div \bm{u}=0, \qquad \pd{\bm{u}}{t} + \Grad p = 0,
+\]
 where $c^2(\bm{x})$ is the wavespeed.
@@
 \]
-The normal flux matrix $\bm{A}_n$ is then
+The normal flux matrix $\bm{A}_n$ and penalty fluxes are then
 \[
@@
 \right)
-\]
-implying that the penalty fluxes are 
-\[
+,
+\qquad
 \bm{A}_n \avg{\bm{U}} - \frac{\tau}{2}\bm{A}_n^T\bm{A}_n \jump{\bm{U}} = \left(\begin{array}{c}
@@
 \bm{A}_{mn} &= \sum_{D^k\in \Oh} \LRp{\bm{A}_i\bm{\Phi}_n, \bm{\Phi}_{m,i}}_{D^k}, \qquad 1\leq m,n \leq N^C\\
-\bm{B}_{mn} &= \sum_{D^k\in \Oh} \LRp{\bm{A}_i\bm{\Psi}_n, \bm{\Phi}_{m,i}}_{D^k} - \LRa{\avg{\bm{A}_n\bm{\Psi}_i}, {\bm{\Phi}_i}}_{\partial D^k} \qquad 1\leq m \leq N^C, \quad 1\leq n \leq N^{NC} \\
-\bm{C}_{mn}  &= -\sum_{D^k\in \Oh} \LRa{\avg{\bm{A}_n\bm{\Psi}_i}, {\bm{\Psi}_i}}_{\partial D^k}, \qquad 1 \leq m,n \leq N^{NC}\\
-\bm{S}_{mn}  &= \frac{1}{2}\sum_{D^k\in \Oh} \LRa{\bm{A}_n^T\jump{\bm{A}_n \bm{\Psi_n}}, {\bm{\Psi}_m}}_{\partial D^k}, \qquad 1 \leq m,n \leq N^{NC}.  
+\bm{B}_{mn} &= \sum_{D^k\in \Oh} \LRp{\bm{A}_i\bm{\Psi}_n, \bm{\Phi}_{m,i}}_{D^k} - \LRa{\bm{A}_n\avg{\bm{\Psi}_i}, {\bm{\Phi}_i}}_{\partial D^k} \qquad 1\leq m \leq N^C, \quad 1\leq n \leq N^{NC} \\
+\bm{C}_{mn}  &= -\sum_{D^k\in \Oh} \LRa{\bm{A}_n\avg{\bm{\Psi}_i}, {\bm{\Psi}_i}}_{\partial D^k}, \qquad 1 \leq m,n \leq N^{NC}\\
+\bm{S}_{mn}  &= \frac{1}{2}\sum_{D^k\in \Oh} \LRa{\bm{A}_n^T\bm{A}_n\jump{ \bm{\Psi_n}}, {\bm{\Psi}_m}}_{\partial D^k}, \qquad 1 \leq m,n \leq N^{NC}.  
 \end{align*}
@@
 \[
-\bm{S}_{mn}  &= \frac{1}{2}\sum_{D^k\in \Oh} \LRa{\jump{\bm{A}_n \bm{\Psi}_n}, \bm{A}_n{\bm{\Psi}_m}}_{\partial D^k} = -\frac{1}{2}\sum_{f \in \Gh} \LRa{\jump{\bm{A}_n \bm{\Psi}_n}, \jump{\bm{A}_n{\bm{\Psi}_m}}}_{f},
+\bm{S}_{mn} = \frac{1}{2}\sum_{D^k\in \Oh} \LRa{\bm{A}_n\jump{ \bm{\Psi}_n}, \bm{A}_n{\bm{\Psi}_m}}_{\partial D^k} = -\frac{1}{2}\sum_{f \in \Gh} \LRa{\bm{A}_n\jump{ \bm{\Psi}_n}, \bm{A}_n\jump{{\bm{\Psi}_m}}}_{f},
 \]
@@
 \centering
-\subfloat[$c_j^0$ (spurious modes for  $\tau = 0$)]{\includegraphics[width=.45\textwidth]{figs/ceigs0.png}}
+\subfloat[$c_j^0$ (coefficients of a spurious mode for  $\tau = 0$)]{\includegraphics[width=.425\textwidth]{ceigs0.png}}
 \hspace{1em}
-\subfloat[$c_j^{100}$ (spurious modes for $\tau = 100$)]{\includegraphics[width=.45\textwidth]{figs/ceigs100.png}}
+\subfloat[$c_j^{100}$ (coefficients of a spurious mode for $\tau = 100$)]{\includegraphics[width=.425\textwidth]{ceigs100.png}}
 \caption{Coefficients $c^0_j$ and $c^{100}_j$ of the spurious mode in Figure~\ref{fig:spurious} when expanded in the modes of $\bm{K}$ for $\tau=1$.
@@
 \centering
-\subfloat[$\tau = 0,1, 50$, zoomed in view]{\includegraphics[height=.3\textheight]{figs/waveEigs2Dbox.png}\label{subfig:big}}
-\hspace{2em}
-\subfloat[$\tau= 0,1,50$, returning eigenvalue paths]{\includegraphics[height=.3\textheight]{figs/trackedWaveEigs2D.png}\label{subfig:box}}
+\subfloat[$\tau = 0,1, 50$, zoomed in view]{\includegraphics[height=.275\textheight]{waveEigs2Dbox.png}\label{subfig:big}}
+\hspace{3em}
+\subfloat[$\tau= 0,1,50$, returning eigenvalue paths]{\includegraphics[height=.275\textheight]{trackedWaveEigs2D.png}\label{subfig:box}}
 \caption{Behavior of eigenvalues for the acoustic wave equation in two dimensions.
@@
 \centering
-\subfloat[$\tau = .1$, $\lambda = -0.2379 + 8.7528i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_p1.png}}
+\subfloat[$\tau = .1$, $\lambda = -0.2379 + 8.7528i$]{\includegraphics[width=.3\textwidth]{waveEig2D_p1.png}}
 \hspace{.5em}
-\subfloat[$\tau = 1$, $\lambda = -1.0145 + 8.9270i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_p2.png}}
+\subfloat[$\tau = 1$, $\lambda = -1.0145 + 8.9270i$]{\includegraphics[width=.3\textwidth]{waveEig2D_p2.png}}
 \hspace{.5em}
-\subfloat[$\tau = 100$, $\lambda = -0.0437 + 7.6167i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_p3.png}}
-\caption{Behavior of the real part of the pressure component of spurious eigenmodes for the acoustic wave equation in two dimensions.  }
+\subfloat[$\tau = 100$, $\lambda = -0.0437 + 7.6167i$]{\includegraphics[width=.3\textwidth]{waveEig2D_p3.png}}
+\caption{Behavior of a spurious eigenmode for the acoustic wave equation in two dimensions.  The real part of the pressure component is shown for different values of $\tau$.}
 \label{fig:trackModesWave}
@@
 \centering
-\subfloat{\includegraphics[width=.475\textwidth]{figs/trackedAdvecEigs.png}}
-\hspace{.5em}
-\subfloat{\includegraphics[width=.465\textwidth]{figs/trackedAdvecEigsZoom1.png}}
+\subfloat{\includegraphics[width=.4275\textwidth]{trackedAdvecEigs.png}}
+\hspace{2em}
+\subfloat{\includegraphics[width=.42\textwidth]{trackedAdvecEigsZoom1.png}}
 \caption{Eigenvalue paths for the DG discretization of advection as $\tau$ increases.  Divergent eigenvalues in the far left half plane are not shown.  }
@@
 \centering
-\subfloat[$\tau = 0$, $\lambda = 0 + 7.9246i$]{\includegraphics[width=.3\textwidth]{figs/advecEig2D_u1.png}}
+\subfloat[$\tau = 0$, $\lambda = 0 + 7.9246i$]{\includegraphics[width=.3\textwidth]{advecEig2D_u1.png}}
 \hspace{.5em}
-\subfloat[$\tau = 1$, $\lambda =  -0.8278 + 6.7831i$]{\includegraphics[width=.3\textwidth]{figs/advecEig2D_u2.png}}
+\subfloat[$\tau = 1$, $\lambda =  -0.8278 + 6.7831i$]{\includegraphics[width=.3\textwidth]{advecEig2D_u2.png}}
 \hspace{.5em}
-\subfloat[$\tau = 100$, $\lambda = -0.0239 + 5.9282i$]{\includegraphics[width=.3\textwidth]{figs/advecEig2D_u3.png}}
-\caption{Behavior of the real part of spurious eigenmodes for the advection equation with $\beta = (1,0)$. }
+\subfloat[$\tau = 100$, $\lambda = -0.0239 + 5.9282i$]{\includegraphics[width=.3\textwidth]{advecEig2D_u3.png}}
+\caption{Behavior of a spurious eigenmode for the advection equation with $\beta = (1,0)$. The real part of the eigenmode is shown for different values of $\tau$.}
 \label{fig:trackModesAdvecU}

### 2017-10-02 docs/dg_penalty/letter.tex

\documentclass{letter}
\usepackage{fullpage}
\signature{Jesse Chan, T.\ Warburton}
\address{6100 Main St,\\
Houston, TX 77005}

\begin{document}
\begin{letter}{Dr Leszek Demkowicz\\CAMWA Editor-in-Chief}

\opening{Dear Dr Demkowicz,}

Please find enclosed our manuscript, 
\begin{center}
Jesse Chan, T.\ Warburton\\
\textit{A short note on the penalty flux parameter for first order discontinuous Galerkin formulations},
\end{center}
which we would like to submit for publication as a short research note in CAMWA.

The submitted manuscript investigates discontinuous Galerkin (DG) methods for symmetric first order hyperbolic systems using a dissipative penalty flux.  We show that, as with second order DG formulations, increasing the penalty parameter $\tau$ divides the spectra into eigenvalues which are bounded and eigenvalues where ${\rm Re}(\lambda_i)\rightarrow -\infty$ as the $\tau \rightarrow \infty$.  In the context of time-domain simulations, bounded eigenvalues correspond to \emph{conforming} solution components, while the latter group of eigenvalues correspond to non-conforming components of the solution and are dissipated from the solution proportionally to ${\rm Re}(\lambda_i)$.

Finally, we demonstrate that spurious solution components which are present for both $\tau = 0$ (non-dissipative central fluxes) and $\tau \rightarrow \infty$ (conforming finite element methods) are damped when $\tau = O(1)$.  This illustrates a unique advantage of dissipative DG methods for time-domain simulations.

We hope that the method and results discussed in this manuscript would appeal to the readership of CAMWA.  All authors have approved the manuscript and agree with its submission, and we confirm that this manuscript has not been published elsewhere and is not under consideration by another journal.  We look forward to hearing from you at your earliest convenience.

\closing{Best regards}

\end{letter}

\end{document}

### 2017-10-02 docs/elasticity_notes/letter.tex

@@
 \begin{document}
-\begin{letter}{Dr.\ Hans De Sterck\\SIAM J. Sci. Comp. Section Editor\\Methods and Algorithms for Scientific Computing}
+\begin{letter}{Dr.\ Charbel Farhat\\Editor of IJNME}
 
-\opening{Dear Dr.\ Hans De Sterck}
+\opening{Dear Dr.\ Farhat}
 
@@
 \end{center}
-which we would like to submit for publication as an original research article in the SIAM Journal on Scientific Computing.
+which we would like to submit for publication as an original research article in IJNME.
+
 This work extends weight-adjusted inner products, which provide a low-storage approximation of the inverse of a mass matrix with respect to an $L^2$ inner product with a spatially varying weight, to the case of spatially varying matrix-valued weights.
-These approximations are then applied to the simulation of elastic wave propagation in arbitrary heterogeneous media, including anisotropy and media with sub-element variations.
-We also introduce a new discontinuous Galerkin method which simplifies numerical fluxes while retaining energy stability and high order accuracy.
+These approximations are then applied to the time-domain simulation of elastic wave propagation in arbitrary heterogeneous media, including anisotropy and media with sub-element variations.
+The novelty of this work also includes a new energy stable and high order accurate discontinuous Galerkin method with simplified numerical fluxes for elastic wave propagation in arbitrary media.
 
-We hope that the method and results discussed in this manuscript will appeal to the readership of SIAM Journal on Scientific Computing.
+We hope that the method and results discussed in this manuscript will appeal to the readership of IJNME.
 We confirm that this manuscript has not been published elsewhere and is not under consideration by another journal.

### 2017-10-02 docs/elasticity_notes/main.tex

@@
 %\documentclass[preprint,10pt]{elsarticle}
-%\documentclass[10pt]{article}
+\documentclass[10pt]{article}
 %\documentclass[review]{siamart0216}
-\documentclass{siamart0216}
+%\documentclass{siamart0216}
+
+%\documentclass[AMA]{WileyNJD-v1}
+%\articletype{Article Type}
 
 \usepackage{fullpage}
+
+\usepackage{lineno}
+
 \usepackage{amsmath,amssymb,amsfonts}
+\usepackage{amsthm}
 \usepackage[titletoc,toc,title]{appendix}
@@
 \usepackage{mathtools}
-\usepackage{pdfpages}
-\usepackage[textsize=footnotesize,color=green]{todonotes}
+%\usepackage{pdfpages}
+%\usepackage[textsize=footnotesize,color=green]{todonotes}
 \usepackage{bm}
-%\usepackage{tikz}
+\usepackage{tikz}
 \usepackage[normalem]{ulem}
 \usepackage{hhline}
-
-%% ====================================== alg package
-\usepackage{algorithm}
-\usepackage[noend]{algpseudocode}
-\usepackage{algorithmicx}
-\algblock{ParFor}{EndParFor}
-% customising the new block
-\algnewcommand\algorithmicparfor{\textbf{parfor}}
-\algnewcommand\algorithmicpardo{\textbf{do}}
-\algnewcommand\algorithmicendparfor{\textbf{end\ parfor}}
-\algrenewtext{ParFor}[1]{\algorithmicparfor\ #1\ \algorithmicpardo}
-\algrenewtext{EndParFor}{\algorithmicendparfor}
-%% ====================================== end alg package
 
@@
 
-%\newtheorem{theorem}{Theorem}[section]
-%\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}[theorem]{Lemma}
+
 %\newtheorem{proposition}[theorem]{Proposition}
@@
 \newcommand{\note}[1]{{\color{blue}{#1}}}
-\newcommand{\remark}[1]{\textbf{\color{red}#1}}
+%\newcommand{\remark}[1]{\textbf{\color{red}#1}}
 
@@
 
+%\title{This is the sample article title\protect\thanks{This is an example for title footnote.}}
+%\author[1]{Author One*}
+%\authormark{AUTHOR ONE \textsc{et al}}
+%\address[1]{\orgdiv{Original Division}, \orgname{Original University}, \orgaddress{\state{State name}, \country{Country name}}}
+%\corres{*Corresponding author name, This is sample corresponding address. \email{authorone@gmail.com}}
+
 \maketitle
 
+\linenumbers
 \begin{abstract}
@@
 0 &< \tilde{c}_{\min} \leq \bm{u}^T\bm{C}^{-1}(\bm{x})\bm{u} \leq \tilde{c}_{\max} < \infty
-%, \qquad \forall \bm{u} \in \mathbb{R}^{N_d}, \quad \forall \bm{x} \in \mathbb{R}^d.
 \end{align*}
@@
 In this case, the stiffness matrix $\bm{C}$ becomes singular, and the constant $\tilde{w}_{\max}$ in the upper bound on $\nor{\bm{C}^{-1}}$ blows up.
-However, numerical experiments in Section~\ref{sec:incomp} suggest that, while taking $\mu/\lambda \approx 0$ (or $\bm{C}$ nearly singular) results in larger relative errors for $\bm{\sigma}$, the accuracy of the WADG solution for $\bm{v}$ does not degrade significantly for near-incompressible materials.
+However, numerical experiments in Section~\ref{sec:incomp} suggest that, while taking $\mu/\lambda \approx 0$ (or $\bm{C}$ nearly singular) results in larger relative errors for $\bm{\sigma}$, the accuracy of the WADG solution for $\bm{v}$ does not degrade as significantly for near-incompressible materials.
 

### 2017-10-02 docs/elasticity_notes/novelty.tex

\documentclass[10pt]{article}
\title{Author's novelty file}
\date{}
\author{}

\begin{document}
\maketitle
\begin{itemize}
\item Extension of weight-adjusted inner products (closed-form approximation of the inverse of a weighted mass matrix) to matrix-valued weights.  
\item A new discontinuous Galerkin (DG) method with simplified numerical fluxes for elastic wave propagation in arbitrary heterogeneous media.  
\item Proofs of energy stability and high order accuracy for the proposed DG method.  
\end{itemize}

\end{document}

### 2017-10-02 docs/elasticity_notes/wileyNJD-AMA.tex

\documentclass[AMA]{WileyNJD-v1}

\articletype{Article Type}%

\received{26 April 2016}
\revised{6 June 2016}
\accepted{6 June 2016}

\usepackage{amsmath}

\raggedbottom

\begin{document}

\title{This is the sample article title\protect\thanks{This is an example for title footnote.}}

\author[1]{Author One*}

\author[2,3]{Author Two}

\author[3]{Author Three}

\authormark{AUTHOR ONE \textsc{et al}}

\address[1]{\orgdiv{Original Division}, \orgname{Original University}, \orgaddress{\state{State name}, \country{Country name}}}

\address[2]{\orgdiv{Original Division}, \orgname{Original University}, \orgaddress{\state{State name}, \country{Country name}}}

\address[3]{\orgdiv{Original Division}, \orgname{Original University}, \orgaddress{\state{State name}, \country{Country name}}}

\corres{*Corresponding author name, This is sample corresponding address. \email{authorone@gmail.com}}

\presentaddress{This is sample for present address text this is sample for present address text}

\abstract[Summary]{This is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text this is sample abstract text.}

\keywords{keyword1, keyword2, keyword3, keyword4}

\jnlcitation{\cname{%
\author{Williams K.}, 
\author{B. Hoskins}, 
\author{R. Lee}, 
\author{G. Masato}, and 
\author{T. Woollings}} (\cyear{2016}), 
\ctitle{A regime analysis of Atlantic winter jet variability applied to evaluate HadGEM3-GC2}, \cjournal{Q.J.R. Meteorol. Soc.}, \cvol{2017;00:1--6}.}

\maketitle

\footnotetext{\textbf{Abbreviations:} ANA, anti-nuclear antibodies; APC, antigen-presenting cells; IRF, interferon regulatory factor}

\section{Sample for first level head}\label{sec1}

\begin{align*}
0 &< \rho_{\min} \leq \rho(\bm{x}) \leq \rho_{\max} < \infty, \\
0 &< c_{\min} \leq \bm{u}^T\bm{C}(\bm{x})\bm{u} \leq c_{\max} < \infty\\
0 &< \tilde{c}_{\min} \leq \bm{u}^T\bm{C}^{-1}(\bm{x})\bm{u} \leq \tilde{c}_{\max} < \infty
\end{align*}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit.\cite{Hirt-Amsden-Cook-74} Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae,
felis. Curabitur dictum gravida mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna. Donec
vehicula augue eu neque. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
Mauris ut leo. Cras viverra metus rhoncus sem. Nulla et lectus vestibulum urna fringilla ultrices. Phasellus eu tellus
sit amet tortor gravida placerat. Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. Praesent eget sem vel
leo ultrices bibendum. Aenean faucibus. Morbi dolor nulla, malesuada eu, pulvinar at, mollis ac, nulla. Curabitur
auctor semper nulla. Donec varius orci eget risus. Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

\begin{eqnarray}
s(nT_{s}) &= &s(t)\times \sum\limits_{n=0}^{N-1} \delta (t-nT_{s}) \xleftrightarrow{\mathrm{DFT}}  S \left(\frac{m}{NT_{s}}\right) \nonumber\\
&= &\frac{1}{N} \sum\limits_{n=0}^{N-1} \sum\limits_{k=-N/2}^{N/2-1} s_{k} e^{\mathrm{j}2\pi k\Delta fnT_{s}} e^{-j\frac{2\pi}{N}mn}
\end{eqnarray}

\section{Sample for another first level head}\label{sec2}

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus viverra
fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa ac
quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo.\cite{Liska_etal-Synchronized_FCT-JCP-10} Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

Example for bibliography citations cite\cite{Taylor499}, cites\cites{Knupp-Winslow_smoothing_2D_unstruct-EwC-99,ka00}

Quisque ullamcorper placerat ipsum. Cras nibh.\cites{Kucharik-Shashkov-Wendroff-JCP-paper-03,Blanchard-Loubere-MOOD_remap-techrep-15} Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
vitae risus porta vehicula.

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.

\begin{figure}[t]
%\centerline{\includegraphics[width=342pt,height=9pc,draft]{blankfig}}
\caption{This is the sample figure caption.\label{fig1}}
\end{figure}

Suspendisse vel felis. Ut lorem lorem, interdum eu, tincidunt sit amet, laoreet vitae, arcu. Aenean faucibus pede eu
ante. Praesent enim elit, rutrum at, molestie non, nonummy vel, nisl. Ut lectus eros, malesuada sit amet, fermentum
eu, sodales cursus, magna. Donec eu purus. Quisque vehicula, urna sed ultricies auctor, pede lorem egestas dui, et
convallis elit erat sed nulla. Donec luctus. Curabitur et nunc. Aliquam dolor odio, commodo pretium, ultricies non,
pharetra in, velit. Integer arcu est, nonummy in, fermentum faucibus, egestas vel, odio.

Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non
odio. Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce
sed justo eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor.
Donec pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin
et quam. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien
turpis, fermentum vel, eleifend faucibus, vehicula eu, lacus.

\begin{figure*}
%\centerline{\includegraphics[width=342pt,height=9pc,draft]{blankfig}}
\caption{This is the sample figure caption.\label{fig2}}
\end{figure*}

\subsection{Example for second level head}

Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit, dictum
in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam vestibulum
fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi. Pellentesque
eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut mi mi, lacinia
sit amet, placerat et, mollis vitae, dui. Sed ante tellus, tristique ut, iaculis eu, malesuada ac, dui. Mauris nibh leo,
facilisis non, adipiscing quis, ultrices a, dui.

Morbi luctus, wisi viverra faucibus pretium, nibh est placerat odio, nec commodo wisi enim eget quam. Quisque
libero justo, consectetuer a, feugiat vitae, porttitor eu, libero. Suspendisse sed mauris vitae elit sollicitudin malesuada.

Maecenas ultricies eros sit amet ante. Ut venenatis velit. Maecenas sed mi eget dui varius euismod. Phasellus aliquet
volutpat odio. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Pellentesque sit
amet pede ac sem eleifend consectetuer. Nullam elementum, urna vel imperdiet sodales, elit ipsum pharetra ligula,
ac pretium ante justo a nulla. Curabitur tristique arcu eu metus. Vestibulum lectus. Proin mauris. Proin eu nunc eu
urna hendrerit faucibus. Aliquam auctor, pede consequat laoreet varius, eros tellus scelerisque quam, pellentesque
hendrerit ipsum dolor sed augue. Nulla nec lacus.

\begin{quote}
This is an example\cites{proceedRef,Galera_etal-2step_hybrid-10,Kucharik20122851} for quote text. This is an example for quote text. This is an example for quote text. This is an example for quote text.\cite{FLD:FLD3996} This is an example for quote text. This is an example for quote text. This is an example for quote text. This is an example for quote text. This is an example for quote text. This is an example for quote text.\cite{Barth-97} This is an example for quote text. This is an example for quote text. This is an example for quote text. 
\end{quote}

\section{Sample for next first level head}\label{sec3}

\subsection{Example for another second level head}

Suspendisse vitae elit. Aliquam arcu neque, ornare in, ullamcorper quis, commodo eu, libero. Fusce sagittis erat at
erat tristique mollis. Maecenas sapien libero, molestie et, lobortis in, sodales eget, dui. Morbi ultrices rutrum lorem.
Nam elementum ullamcorper leo. Morbi dui. Aliquam sagittis. Nunc placerat. Pellentesque tristique sodales est.
Maecenas imperdiet lacinia velit. Cras non urna. Morbi eros pede, suscipit ac, varius vel, egestas non, eros. Praesent
malesuada, diam id pretium elementum, eros sem dictum tortor, vel consectetuer odio sem sed wisi.

Sed feugiat. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Ut pellentesque
augue sed urna. Vestibulum diam eros, fringilla et, consectetuer eu, nonummy id, sapien. Nullam at lectus. In sagittis
ultrices mauris. Curabitur malesuada erat sit amet massa. Fusce blandit. Aliquam erat volutpat. Aliquam euismod.
Aenean vel lectus. Nunc imperdiet justo nec dolor.

\subsection{Second level head text}

Etiam euismod. Fusce facilisis lacinia dui. Suspendisse potenti. In mi erat, cursus id, nonummy sed, ullamcorper
eget, sapien. Praesent pretium, magna in eleifend egestas, pede pede pretium lorem, quis consectetuer tortor sapien
facilisis magna. Mauris quis magna varius nulla scelerisque imperdiet. Aliquam non quam. Aliquam porttitor quam
a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.

\subsubsection{Third level head text}

Aliquam lectus. Vivamus leo. Quisque ornare tellus ullamcorper nulla. Mauris porttitor pharetra tortor. Sed fringilla
justo sed mauris. Mauris tellus. Sed non leo. Nullam elementum, magna in cursus sodales, augue est scelerisque
sapien, venenatis congue nulla arcu et pede. Ut suscipit enim vel sapien. Donec congue. Maecenas urna mi, suscipit
in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl.

Etiam ac leo a risus tristique nonummy. Donec dignissim tincidunt nulla. Vestibulum rhoncus molestie odio. Sed
lobortis, justo et pretium lobortis, mauris turpis condimentum augue, nec ultricies nibh arcu pretium enim. Nunc
purus neque, placerat id, imperdiet sed, pellentesque nec, nisl. Vestibulum imperdiet neque non sem accumsan laoreet.
In hac habitasse platea dictumst. Etiam condimentum facilisis libero. Suspendisse in elit quis nisl aliquam dapibus.
Pellentesque auctor sapien. Sed egestas sapien nec lectus. Pellentesque vel dui vel neque bibendum viverra. Aliquam
porttitor nisl nec pede. Proin mattis libero vel turpis. Donec rutrum mauris et libero. Proin euismod porta felis.
Nam lobortis, metus quis elementum commodo, nunc lectus elementum mauris, eget vulputate ligula tellus eu neque.
Vivamus eu dolor.

Nulla in ipsum. Praesent eros nulla, congue vitae, euismod ut, commodo a, wisi. Pellentesque habitant morbi
tristique senectus et netus et malesuada fames ac turpis egestas. Aenean nonummy magna non leo. Sed felis erat,
ullamcorper in, dictum non, ultricies ut, lectus. Proin vel arcu a odio lobortis euismod. Vestibulum ante ipsum primis
in faucibus orci luctus et ultrices posuere cubilia Curae; Proin ut est. Aliquam odio. Pellentesque massa turpis, cursus
eu, euismod nec, tempor congue, nulla. Duis viverra gravida mauris. Cras tincidunt. Curabitur eros ligula, varius ut,
pulvinar in, cursus faucibus, augue.

\begin{boxtext}
\section*{Example of Boxtext}%
This is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext this is sample for boxtext.
\end{boxtext}

\paragraph{Fourth level head text}

Sed feugiat. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Ut pellentesque
augue sed urna. Vestibulum diam eros, fringilla et, consectetuer eu, nonummy id, sapien. Nullam at lectus. In sagittis
ultrices mauris. Curabitur malesuada erat sit amet massa. Fusce blandit. Aliquam erat volutpat. Aliquam euismod.
Aenean vel lectus. Nunc imperdiet justo nec dolor.

Etiam euismod. Fusce facilisis lacinia dui. Suspendisse potenti. In mi erat, cursus id, nonummy sed, ullamcorper
eget, sapien. Praesent pretium, magna in eleifend egestas, pede pede pretium lorem, quis consectetuer tortor sapien
facilisis magna. Mauris quis magna varius nulla scelerisque imperdiet. Aliquam non quam. Aliquam porttitor quam
a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.

\subparagraph{Fifth level head text}

Aliquam lectus. Vivamus leo. Quisque ornare tellus ullamcorper nulla. Mauris porttitor pharetra
tortor. Sed fringilla justo sed mauris. Mauris tellus. Sed non leo. Nullam elementum, magna in cursus sodales, augue
est scelerisque sapien, venenatis congue nulla arcu et pede. Ut suscipit enim vel sapien. Donec congue. Maecenas
urna mi, suscipit in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl.

Etiam ac leo a risus tristique nonummy. Donec dignissim tincidunt nulla. Vestibulum rhoncus molestie odio. Sed
lobortis, justo et pretium lobortis, mauris turpis condimentum augue, nec ultricies nibh arcu pretium enim. Nunc
purus neque, placerat id, imperdiet sed, pellentesque nec, nisl. Vestibulum imperdiet neque non sem accumsan laoreet.
In hac habitasse platea dictumst. Etiam condimentum facilisis libero. Suspendisse in elit quis nisl aliquam dapibus.
Pellentesque auctor sapien. Sed egestas sapien nec lectus. Pellentesque vel dui vel neque bibendum viverra. Aliquam
porttitor nisl nec pede. Proin mattis libero vel turpis. Donec rutrum mauris et libero. Proin euismod porta felis.
Nam lobortis, metus quis elementum commodo, nunc lectus elementum mauris, eget vulputate ligula tellus eu neque.
Vivamus eu dolor.

in faucibus orci luctus et ultrices posuere cubilia Curae; Proin ut est. Aliquam odio. Pellentesque massa turpis, cursus
eu, euismod nec, tempor congue, nulla. Duis viverra gravida mauris. Cras tincidunt. Curabitur eros ligula, varius ut,
pulvinar in, cursus faucibus, augue.

Curabitur tellus magna, porttitor a, commodo a, commodo in, tortor. Donec interdum. Praesent scelerisque. Mae-
cenas posuere sodales odio. Vivamus metus lacus, varius quis, imperdiet quis, rhoncus a, turpis. Etiam ligula arcu,
elementum a, venenatis quis, sollicitudin sed, metus. Donec nunc pede, tincidunt in, venenatis vitae, faucibus vel,
nibh. Pellentesque wisi. Nullam malesuada. Morbi ut tellus ut pede tincidunt porta. Lorem ipsum dolor sit amet,
consectetuer adipiscing elit. Etiam congue neque id dolor.

Donec et nisl at wisi luctus bibendum. Nam interdum tellus ac libero. Sed sem justo, laoreet vitae, fringilla at,
adipiscing ut, nibh. Maecenas non sem quis tortor eleifend fermentum. Etiam id tortor ac mauris porta vulputate.
Integer porta neque vitae massa. Maecenas tempus libero a libero posuere dictum. Vestibulum ante ipsum primis in
faucibus orci luctus et ultrices posuere cubilia Curae; Aenean quis mauris sed elit commodo placerat. Class aptent
taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Vivamus rhoncus tincidunt libero.
Etiam elementum pretium justo. Vivamus est. Morbi a tellus eget pede tristique commodo. Nulla nisl. Vestibulum
sed nisl eu sapien cursus rutrum.

Nulla non mauris vitae wisi posuere convallis. Sed eu nulla nec eros scelerisque pharetra. Nullam varius. Etiam
dignissim elementum metus. Vestibulum faucibus, metus sit amet mattis rhoncus, sapien dui laoreet odio, nec ultricies
nibh augue a enim. Fusce in ligula. Quisque at magna et nulla commodo consequat. Proin accumsan imperdiet sem.
Nunc porta. Donec feugiat mi at justo. Phasellus facilisis ipsum quis ante. In ac elit eget ipsum pharetra faucibus.
Maecenas viverra nulla in massa.

Nulla in ipsum. Praesent eros nulla, congue vitae, euismod ut, commodo a, wisi. Pellentesque habitant morbi
tristique senectus et netus et malesuada fames ac turpis egestas. Aenean nonummy magna non leo. Sed felis erat,
ullamcorper in, dictum non, ultricies ut, lectus. Proin vel arcu a odio lobortis euismod. Vestibulum ante ipsum primis

\begin{center}
\begin{table*}[t]%
\caption{This is sample table caption.\label{tab1}}
\centering
\begin{tabular*}{500pt}{@{\extracolsep\fill}lccD{.}{.}{3}c@{\extracolsep\fill}}
\toprule
&\multicolumn{2}{@{}c@{}}{\textbf{Spanned heading\tnote{1}}} & \multicolumn{2}{@{}c@{}}{\textbf{Spanned heading\tnote{2}}} \\\cmidrule{2-3}\cmidrule{4-5}
\textbf{col1 head} & \textbf{col2 head}  & \textbf{col3 head}  & \multicolumn{1}{@{}l@{}}{\textbf{col4 head}}  & \textbf{col5 head}   \\
\midrule
col1 text & col2 text  & col3 text  & 12.34  & col5 text\tnote{1}   \\
col1 text & col2 text  & col3 text  & 1.62  & col5 text\tnote{2}   \\
col1 text & col2 text  & col3 text  & 51.809  & col5 text   \\
\bottomrule
\end{tabular*}
\begin{tablenotes}%%[341pt]
\item Source: Example for table source text.
\item[1] Example for a first table footnote.
\item[2] Example for a second table footnote.
\end{tablenotes}
\end{table*}
\end{center}

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.

\begin{center}
\begin{table}[t]%
\centering
\caption{This is sample table caption.\label{tab2}}%
\begin{tabular*}{500pt}{@{\extracolsep\fill}lcccc@{\extracolsep\fill}}
\toprule
\textbf{col1 head} & \textbf{col2 head}  & \textbf{col3 head}  & \textbf{col4 head}  & \textbf{col5 head} \\
\midrule
col1 text & col2 text  & col3 text  & col4 text  & col5 text\tnote{$\dagger$}   \\
col1 text & col2 text  & col3 text  & col4 text  & col5 text   \\
col1 text & col2 text  & col3 text  & col4 text  & col5 text\tnote{$\ddagger$}   \\
\bottomrule
\end{tabular*}
\begin{tablenotes}
\item Source: Example for table source text.
\item[$\dagger$] Example for a first table footnote.
\item[$\ddagger$] Example for a second table footnote.
\end{tablenotes}
\end{table}
\end{center}

Below is the example\cites{Liska_etal-Synchronized_FCT-JCP-10,Kucharik-Shashkov-Wendroff-JCP-paper-03,Blanchard-Loubere-MOOD_remap-techrep-15} for bulleted list. Below is the example for bulleted list. Below is the example for bulleted list. Below is the example for bulleted list. Below is the example for bulleted list. Below is the example for bulleted list\footnote{This is an example for footnote.}:
\begin{itemize}
\item bulleted list entry sample bulleted list entry.\cite{Lauritzen_Erath_Mittal-swept_exact_error_analysis-JCP-11} sample list entry text. 
\item bulleted list entry sample bulleted list entry. bulleted list entry sample bulleted list entry. bulleted list entry sample bulleted list entry.
\item bulleted list entry sample bulleted list entry.\cite{Klima-Kucharik-analysis} bulleted list entry sample bulleted list entry.\cite{Dukowicz_Baumgardner-advection-JCP-00} sample list entry text.  bulleted list entry sample bulleted list entry.
\item sample list entry text. sample list entry text.  
\end{itemize}

Suspendisse vel felis. Ut lorem lorem, interdum eu, tincidunt sit amet, laoreet vitae, arcu. Aenean faucibus pede eu
ante. Praesent enim elit, rutrum at, molestie non, nonummy vel, nisl. Ut lectus eros, malesuada sit amet, fermentum
eu, sodales cursus, magna. Donec eu purus. Quisque vehicula, urna sed ultricies auctor, pede lorem egestas dui, et
convallis elit erat sed nulla. Donec luctus. Curabitur et nunc. Aliquam dolor odio, commodo pretium, ultricies non,
pharetra in, velit. Integer arcu est, nonummy in, fermentum faucibus, egestas vel, odio.

Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non
odio. Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce
sed justo eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor. Donec pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin
et quam. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien
turpis, fermentum vel, eleifend faucibus, vehicula eu, lacus.

Below is the sample for description list. Below is the example for description list. Below is the example for description list. Below is the example for description list. Below is the example for description list. Below is the example for description list:\\[12pt]

\noindent\textbf{Description sample:}

\begin{description}
\item[first entry] description text. description text.\cites{Kucharik_Shashkov-MM_remap-FVCA-Proc-11,Kucharik_Shashkov-MM_remap-13,Loubere-Shashkov-subcell-remap-JCP-05} description text. description text. description text. description text. description text. 
\item[second long entry] description text. description text. description text. description text. description text. description text. description text. 
\item[third entry] description text. description text. description text. description text. description text. 
\item[fourth entry] description text. description text. 
\end{description}

\noindent\textbf{Numbered list items sample:}

\begin{enumerate}[1.]
\item First level numbered list entry. sample numbered list entry.

\item First numbered list entry. sample numbered list entry. Numbered list entry.\cite{Caramana-Shashkov-98} sample numbered list entry. Numbered list entry. sample numbered list entry.

\begin{enumerate}[a.]
\item Second level alpabetical list entry. Second level alpabetical list entry. Second level alpabetical list entry.\cite{Hoch-ALE-HAL-09} Second level alpabetical list entry.

\item Second level alpabetical list entry. Second level alpabetical list entry.\cites{Shashkov-book-96,Knupp-Winslow_smoothing_2D_unstruct-EwC-99,Knupp-Winslow_smoothing_2D_unstruct-EwC-99}

\begin{enumerate}[ii.]
\item Third level lowercase roman numeral list entry. Third level lowercase roman numeral list entry. Third level lowercase roman numeral list entry.

\item Third level lowercase roman numeral list entry. Third level lowercase roman numeral list entry.\cite{ka00}
\end{enumerate}

\item Second level alpabetical list entry. Second level alpabetical list entry.\cite{Taylor499}
\end{enumerate}

\item First level numbered list entry. sample numbered list entry. Numbered list entry. sample numbered list entry. Numbered list entry.

\item Another first level numbered list entry. sample numbered list entry. Numbered list entry. sample numbered list entry. Numbered list entry. 
\end{enumerate}

\noindent\textbf{un-numbered list items sample:}

\begin{enumerate}[]
\item Sample unnumberd list text..
\item Sample unnumberd list text.
\item sample unnumberd list text. 
\item Sample unnumberd list text.
\end{enumerate}

\section{Examples for enunciations}\label{sec4}

\begin{theorem}[Theorem subhead]\label{thm1}
Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text. 
\end{theorem}

Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
vitae risus porta vehicula.

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.

\begin{proposition}
Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text. 
\end{proposition}

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
vitae risus porta vehicula.

\begin{definition}[Definition sub head]
Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text.
\end{definition}

Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non
odio. Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce
sed justo eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor.
Donec pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin
et quam. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien
turpis, fermentum vel, eleifend faucibus, vehicula eu, lacus.

Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit,
dictum in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam
vestibulum fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi.
Pellentesque eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut
mi mi, lacinia sit amet, placerat et, mollis vitae, dui. Sed ante tellus, tristique ut, iaculis eu, malesuada ac, dui.
Mauris nibh leo, facilisis non, adipiscing quis, ultrices a, dui.

\begin{proof}
Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. 
\end{proof}

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

\begin{proof}[Proof of Theorem~\ref{thm1}]
Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. 
\end{proof}

Etiam euismod. Fusce facilisis lacinia dui. Suspendisse potenti. In mi erat, cursus id, nonummy sed, ullamcorper
eget, sapien. Praesent pretium, magna in eleifend egestas, pede pede pretium lorem, quis consectetuer tortor sapien
facilisis magna. Mauris quis magna varius nulla scelerisque imperdiet. Aliquam non quam. Aliquam porttitor quam
a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.
Aliquam lectus. Vivamus leo. Quisque ornare tellus ullamcorper nulla. Mauris porttitor pharetra tortor. Sed fringilla
justo sed mauris. Mauris tellus. Sed non leo. Nullam elementum, magna in cursus sodales, augue est scelerisque
sapien, venenatis congue nulla arcu et pede. Ut suscipit enim vel sapien. Donec congue. Maecenas urna mi, suscipit
in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl.

Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit,
dictum in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam
vestibulum fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi.
Pellentesque eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut Curabitur tellus magna, porttitor a, commodo a, commodo in, tortor. Donec interdum. Praesent scelerisque. Mae-
cenas posuere sodales odio. Vivamus metus lacus, varius quis, imperdiet quis, rhoncus a, turpis. Etiam ligula arcu,
elementum a, venenatis quis, sollicitudin sed, metus. Donec nunc pede, tincidunt in, venenatis vitae, faucibus vel,

\begin{sidewaystable}%[h]
\caption{Sideways table caption. For decimal alignment refer column 4 to 9 in tabular* preamble.\label{tab3}}%
\begin{tabular*}{\textheight}{@{\extracolsep\fill}lccD{.}{.}{4}D{.}{.}{4}D{.}{.}{4}D{.}{.}{4}D{.}{.}{4}D{.}{.}{4}@{\extracolsep\fill}}%
\toprule
  & \textbf{col2 head} & \textbf{col3 head} & \multicolumn{1}{c}{\textbf{10}} &\multicolumn{1}{c}{\textbf{20}} &\multicolumn{1}{c}{\textbf{30}} &\multicolumn{1}{c}{\textbf{10}} &\multicolumn{1}{c}{\textbf{20}} &\multicolumn{1}{c}{\textbf{30}} \\
\midrule
  &col2 text &col3 text &0.7568&1.0530&1.2642&0.9919&1.3541&1.6108 \\
  & &col2 text &12.5701 &19.6603&25.6809&18.0689&28.4865&37.3011 \\
3 &col2 text  & col3 text &0.7426&1.0393&1.2507&0.9095&1.2524&1.4958 \\
  & &col3 text &12.8008&19.9620&26.0324&16.6347&26.0843&34.0765 \\
  & col2 text & col3 text &0.7285&1.0257&1.2374&0.8195&1.1407&1.3691\tnote{*} \\
  & & col3 text &13.0360&20.2690&26.3895&15.0812&23.4932&30.6060\tnote{\dagger} \\
\bottomrule
\end{tabular*}
\begin{tablenotes}%%[\textheight]
\item[*] First sideways table footnote. Sideways table footnote. Sideways table footnote. Sideways table footnote.
\item[$\dagger$] Second sideways table footnote. Sideways table footnote. Sideways table footnote. Sideways table footnote.
\end{tablenotes}
\end{sidewaystable}

\begin{sidewaysfigure}
%\centerline{\includegraphics[width=542pt,height=9pc,draft]{blankfig}}
\caption{Sideways figure caption. Sideways figure caption. Sideways figure caption. Sideways figure caption. Sideways figure caption. Sideways figure caption.\label{fig3}}
\end{sidewaysfigure}

nibh. Pellentesque wisi.\cite{Kucharik20122851} Nullam malesuada. Morbi ut tellus ut pede tincidunt porta. Lorem ipsum dolor sit amet,
consectetuer adipiscing elit. Etiam congue neque id dolor.

\begin{algorithm}
\caption{Pseudocode for our algorithm}\label{alg1}
\begin{algorithmic}
  \For each frame
  \For water particles $f_{i}$
  \State compute fluid flow\cite{Hirt-Amsden-Cook-74}
  \State compute fluid--solid interaction\cite{Benson-92}
  \State apply adhesion and surface tension\cite{Margolin-Shashkov-remapping-03}
  \EndFor
   \For solid particles $s_{i}$
   \For neighboring water particles $f_{j}$
   \State compute virtual water film \\(see Section~\ref{Waterfilm})
   \EndFor
   \EndFor
   \For solid particles $s_{i}$
   \For neighboring water particles $f_{j}$
   \State compute growth direction vector \\(see Section~\ref{GDV})
   \EndFor
   \EndFor
   \For solid particles $s_{i}$
   \For neighboring water particles $f_{j}$
   \State compute $F_{\theta}$ (see Section~\ref{Freezingfactor})
   \State compute $CE(s_{i},f_{j})$ \\(see Section~\ref{Phasechange})
   \If $CE(b_{i}, f_{j})$ $>$ glaze threshold
   \State $j$th water particle's phase $\Leftarrow$ ICE
   \EndIf
   \If $CE(c_{i}, f_{j})$ $>$ icicle threshold
   \State $j$th water particle's phase $\Leftarrow$ ICE
   \EndIf
   \EndFor
   \EndFor
  \EndFor
\end{algorithmic}
\end{algorithm}

Donec et nisl at wisi luctus bibendum. Nam interdum tellus ac libero. Sed sem justo, laoreet vitae, fringilla at,
adipiscing ut, nibh. Maecenas non sem quis tortor eleifend fermentum. Etiam id tortor ac mauris porta vulputate.
Integer porta neque vitae massa.\cites{Hirt-Amsden-Cook-74,Benson-92}  Maecenas tempus libero a libero posuere dictum. Vestibulum ante ipsum primis in
faucibus orci luctus et ultrices posuere cubilia Curae; Aenean quis mauris sed elit commodo placerat. Class aptent
taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Vivamus rhoncus tincidunt libero.
Etiam elementum pretium justo. Vivamus est. Morbi a tellus eget pede tristique commodo.\cites{Benson-92} Nulla nisl. Vestibulum
sed nisl eu sapien cursus rutrum.

Pellentesque wisi. Nullam malesuada. Morbi ut tellus ut pede tincidunt porta. Lorem ipsum dolor sit amet,
consectetuer adipiscing elit. Etiam congue neque id dolor.

Donec et nisl at wisi luctus bibendum. Nam interdum tellus ac libero. Sed sem justo, laoreet vitae, fringilla at,
adipiscing ut, nibh. Maecenas non sem quis tortor eleifend fermentum. Etiam id tortor ac mauris porta vulputate.
Integer porta neque vitae massa. Maecenas tempus libero a libero posuere dictum. Vestibulum ante ipsum primis in
faucibus orci luctus et ultrices posuere cubilia Curae; Aenean quis mauris sed elit commodo placerat. Class aptent
taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Vivamus rhoncus tincidunt libero.
Etiam elementum pretium justo. Vivamus est. Morbi a tellus eget pede tristique commodo. Nulla nisl. Vestibulum
sed nisl eu sapien cursus rutrum.

\begin{equation}\label{eq23}
 \|\tilde{X}(k)\|^2
 =\frac{\left\|\sum\limits_{i=1}^{p}\tilde{Y}_i(k)+\sum\limits_{j=1}^{q}\tilde{Z}_j(k) \right\|^2}{(p+q)^2}
 \leq\frac{\sum\limits_{i=1}^{p}\left\|\tilde{Y}_i(k)\right\|^2+\sum\limits_{j=1}^{q}\left\|\tilde{Z}_j(k)\right\|^2 }{p+q}.
\end{equation}

Sed feugiat. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Ut pellentesque
augue sed urna. Vestibulum diam eros, fringilla et, consectetuer eu, nonummy id, sapien. Nullam at lectus. In sagittis
ultrices mauris. Curabitur malesuada erat sit amet massa. Fusce blandit. Aliquam erat volutpat. Aliquam euismod.
Aenean vel lectus. Nunc imperdiet justo nec dolor.

Etiam euismod. Fusce facilisis lacinia dui. Suspendisse potenti. In mi erat, cursus id, nonummy sed, ullamcorper
eget, sapien. Praesent pretium, magna in eleifend egestas, pede pede pretium lorem, quis consectetuer tortor sapien
facilisis magna. Mauris quis magna varius nulla scelerisque imperdiet. Aliquam non quam. Aliquam porttitor quam
a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.

\begin{equation}\label{eq24}
 \|\tilde{X}(k)\|^2
 =\frac{\left\|\sum\limits_{i=1}^{p}\tilde{Y}_i(k)+\sum\limits_{j=1}^{q}\tilde{Z}_j(k) \right\|^2}{(p+q)^2}
 \leq\frac{\sum\limits_{i=1}^{p}\left\|\tilde{Y}_i(k)\right\|^2+\sum\limits_{j=1}^{q}\left\|\tilde{Z}_j(k)\right\|^2 }{p+q}.
\end{equation}

Aliquam lectus. Vivamus leo. Quisque ornare tellus ullamcorper nulla. Mauris porttitor pharetra
tortor. Sed fringilla justo sed mauris. Mauris tellus. Sed non leo. Nullam elementum, magna in cursus sodales, augue
est scelerisque sapien, venenatis congue nulla arcu et pede. Ut suscipit enim vel sapien. Donec congue. Maecenas
urna mi, suscipit in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl.

Etiam ac leo a risus tristique nonummy. Donec dignissim tincidunt nulla. Vestibulum rhoncus molestie odio. Sed
lobortis, justo et pretium lobortis, mauris turpis condimentum augue, nec ultricies nibh arcu pretium enim. Nunc
purus neque, placerat id, imperdiet sed, pellentesque nec, nisl. Vestibulum imperdiet neque non sem accumsan laoreet.
In hac habitasse platea dictumst. Etiam condimentum facilisis libero. Suspendisse in elit quis nisl aliquam dapibus.
Pellentesque auctor sapien. Sed egestas sapien nec lectus. Pellentesque vel dui vel neque bibendum viverra. Aliquam
porttitor nisl nec pede. Proin mattis libero vel turpis. Donec rutrum mauris et libero. Proin euismod porta felis.
Nam lobortis, metus quis elementum commodo, nunc lectus elementum mauris, eget vulputate ligula tellus eu neque.
Vivamus eu dolor.

\section{Conclusions}\label{sec5}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae,
felis. Curabitur dictum gravida mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna. Donec
vehicula augue eu neque. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
Mauris ut leo. Cras viverra metus rhoncus sem. Nulla et lectus vestibulum urna fringilla ultrices. Phasellus eu tellus
sit amet tortor gravida placerat. Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. Praesent eget sem vel
leo ultrices bibendum. Aenean faucibus. Morbi dolor nulla, malesuada eu, pulvinar at, mollis ac, nulla. Curabitur
auctor semper nulla. Donec varius orci eget risus. Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

%\backmatter

\section*{Acknowledgments}
This is acknowledgment text.\cite{Kenamond_Burton-Exact_remap_MM_poly-Multimat_SF-Pres13} Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here. This is acknowledgment text. Provide text here.

\subsection*{Author contributions}

This is an author contribution text. This is an author contribution text. This is an author contribution text. This is an author contribution text. This is an author contribution text.

\subsection*{Financial disclosure}

None reported.

\subsection*{Conflict of interest}

The authors declare no potential conflict of interests.

\section*{Supporting information}

The following supporting information is available as part of the online article:

\noindent
\textbf{Figure S1.}
{500{\uns}hPa geopotential anomalies for GC2C calculated against the ERA Interim reanalysis. The period is 1989--2008.}

\noindent
\textbf{Figure S2.}
{The SST anomalies for GC2C calculated against the observations (OIsst).}

\appendix

\section{Section title of first appendix\label{app1}}

Use \verb+\begin{verbatim}...\end{verbatim}+ for program codes without math. Use \verb+\begin{alltt}...\end{alltt}+ for program codes with math. Based on the text provided inside the optional argument of \verb+\begin{code}[Psecode|Listing|Box|Code|+\hfill\break \verb+Specification|Procedure|Sourcecode|Program]...+ \verb+\end{code}+ tag corresponding boxed like floats are generated. Also note that \verb+\begin{code}[Code|Listing]...+ \verb+\end{code}+ tag with either Code or Listing text as optional argument text are set with computer modern typewriter font.  All other code environments are set with normal text font. Refer below example:

\begin{lstlisting}[caption={Descriptive Caption Text},label=DescriptiveLabel]
for i:=maxint to 0 do
begin
{ do nothing }
end;
Write('Case insensitive ');
WritE('Pascal keywords.');
\end{lstlisting}

\subsection{Subsection title of first appendix\label{app1.1a}}

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

\subsubsection{Subsection title of first appendix\label{app1.1.1a}}

\noindent\textbf{Unnumbered figure}

\begin{center}
%\includegraphics[width=7pc,height=8pc,draft]{blankfig}
\end{center}

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.

Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.
Suspendisse vel felis. Ut lorem lorem, interdum eu, tincidunt sit amet, laoreet vitae, arcu. Aenean faucibus pede eu
ante. Praesent enim elit, rutrum at, molestie non, nonummy vel, nisl. Ut lectus eros, malesuada sit amet, fermentum
eu, sodales cursus, magna. Donec eu purus. Quisque vehicula, urna sed ultricies auctor, pede lorem egestas dui, et
convallis elit erat sed nulla. Donec luctus. Curabitur et nunc. Aliquam dolor odio, commodo pretium, ultricies non,
pharetra in, velit. Integer arcu est, nonummy in, fermentum faucibus, egestas vel, odio.

\section{Section title of second appendix\label{app2}}%

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.

Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.
Suspendisse vel felis. Ut lorem lorem, interdum eu, tincidunt sit amet, laoreet vitae, arcu. Aenean faucibus pede eu
ante. Praesent enim elit, rutrum at, molestie non, nonummy vel, nisl. Ut lectus eros, malesuada sit amet, fermentum
eu, sodales cursus, magna. Donec eu purus. Quisque vehicula, urna sed ultricies auctor, pede lorem egestas dui, et
convallis elit erat sed nulla. Donec luctus. Curabitur et nunc. Aliquam dolor odio, commodo pretium, ultricies non,
pharetra in, velit. Integer arcu est, nonummy in, fermentum faucibus, egestas vel, odio.

%== Figure 4 ==
%% Example for figure inside appendix
\begin{figure}[t]
%\centerline{\includegraphics[height=10pc,width=78mm,draft]{blankfig}}
\caption{This is an example for appendix figure.\label{fig5}}
\end{figure}

\subsection{Subsection title of second appendix\label{app2.1a}}

Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non odio.
Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce sed justo
eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor. Donec
pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin et quam.
Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien turpis,
fermentum vel, eleifend faucibus, vehicula eu, lacus.

Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit,
dictum in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam
vestibulum fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi.
Pellentesque eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut
mi mi, lacinia sit amet, placerat et, mollis vitae, dui. Sed ante tellus, tristique ut, iaculis eu, malesuada ac, dui.
Mauris nibh leo, facilisis non, adipiscing quis, ultrices a, dui.

\subsubsection{Subsection title of second appendix\label{app2.1.1a}}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae,
felis. Curabitur dictum gravida mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna. Donec
vehicula augue eu neque. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
Mauris ut leo. Cras viverra metus rhoncus sem. Nulla et lectus vestibulum urna fringilla ultrices. Phasellus eu tellus
sit amet tortor gravida placerat. Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. Praesent eget sem vel
leo ultrices bibendum. Aenean faucibus. Morbi dolor nulla, malesuada eu, pulvinar at, mollis ac, nulla. Curabitur
auctor semper nulla. Donec varius orci eget risus. Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

\begin{center}
\begin{table}[b]%
\centering
\caption{This is an example of Appendix table showing food requirements of army, navy and airforce.\label{tab4}}%
\begin{tabular*}{300pt}{@{\extracolsep\fill}lcc@{\extracolsep\fill}}%
\toprule
\textbf{col1 head} & \textbf{col2 head} & \textbf{col3 head} \\
\midrule
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text\\
\bottomrule
\end{tabular*}
\end{table}
\end{center}

Example for an equation inside appendix
\begin{equation}
\mathcal{L}\quad \mathbf{\mathcal{L}} = i \bar{\psi} \gamma^\mu D_\mu \psi - \frac{1}{4} F_{\mu\nu}^a F^{a\mu\nu} - m \bar{\psi} \psi\label{eq25}
\end{equation}

\section{Example of another appendix section\label{app3}}%

This is sample for paragraph text this is sample for paragraph text  this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text this is sample for paragraph text

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

\begin{equation}
\mathcal{L} = i \bar{\psi} \gamma^\mu D_\mu \psi
    - \frac{1}{4} F_{\mu\nu}^a F^{a\mu\nu} - m \bar{\psi} \psi
\label{eq26}
\end{equation}

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
vitae risus porta vehicula.

\begin{center}
\begin{tabular*}{300pt}{@{\extracolsep\fill}lcc@{\extracolsep\fill}}%
\toprule
\textbf{col1 head} & \textbf{col2 head} & \textbf{col3 head} \\
\midrule
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text \\
\bottomrule
\end{tabular*}
\end{center}

Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio placerat
quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin tempus nibh sit amet nisl. Vivamus quis tortor
vitae risus porta vehicula.

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, evehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
Nulla egestas. Curabitur a leo. Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur consectetuer.

Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Donec odio elit,
dictum in, hendrerit sit amet, egestas sed, leo. Praesent feugiat sapien aliquet odio. Integer vitae justo. Aliquam
vestibulum fringilla lorem. Sed neque lectus, consectetuer at, consectetuer sed, eleifend ac, lectus. Nulla facilisi.
Pellentesque eget lectus. Proin eu metus. Sed porttitor. In hac habitasse platea dictumst. Suspendisse eu lectus. Ut
mi mi, lacinia sit amet, placerat et, mollis vitae, dui. Sed ante tellus, tristique ut, iaculis eu, malesuada ac, dui.
Mauris nibh leo, facilisis non, adipiscing quis, ultrices a, dui.

\begin{thebibliography}{55}

%==Ref1
\bib{Hirt-Amsden-Cook-74}{article}{
      author={Hirt, CW},
      author={Amsden, AA},
      author={Cook, JL},
       title={An arbitrary Lagrangian-Eulerian computing method for all flow speeds},
        date={1974},
     journal={J Comput Phys},
      volume={14},
      number={3},
       pages={227--253},
}

%==Ref2
\bib{Benson-92}{article}{
      author={Benson, DJ},
       title={Computational methods in Lagrangian and Eulerian hydrocodes},
        date={1992},
     journal={Comput Method Appl M},
      volume={99},
      number={2--3},
       pages={235--394},
}

%==Ref3
\bib{Dukowicz-Conservative_remap-JCP-84}{article}{
      author={Dukowicz, J},
       title={Conservative rezoning (remapping) for general quadrilateral meshes},
        date={1984},
     journal={J Comput Phys},
      volume={54},
      number={3},
       pages={411--424},
}

%==Ref4
\bib{Margolin-Shashkov-remapping-03}{article}{
      author={Margolin, LG},
      author={Shashkov, M},
       title={Second-order sign-preserving conservative interpolation (remapping) on general grids},
        date={2003},
     journal={J Comput Phys},
      volume={184},
      number={1},
       pages={266--298},
}

%==Ref5
\bib{Kenamond_Burton-Exact_remap_MM_poly-Multimat_SF-Pres13}{proceedings}{
      author={Kenamond, MA},
      author={Burton, DE},
       title={Exact intersection remapping of multi-material domain-decomposed polygonal meshes},
  conference={Talk at Multimat 2013, International Conference on Numerical Methods for Multi-Material Fluid Flows},
     address={San Francisco},
        date={{September 2--6, 2013}},
        note={LA-UR-13-26794},
}

%==Ref6
\bib{proceedRef}{proceedings}{
      author={Burton, DE},
      author={Kenamond, MA},
      author={Morgan, NR},
      author={Carney, TC},
      author={Shashkov, MJ},
       title={An intersection based ALE scheme (xALE) for cell centered hydrodynamics (CCH)},
         conference={Talk at Multimat 2013, International Conference on Numerical Methods for Multi-Material Fluid Flows},
         address={San Francisco},
        date={{September 2--6, 2013}},
        note={LA-UR-13-26756.2},
}

%==Ref7
\bib{Galera_etal-2step_hybrid-10}{article}{
      author={Berndt, M},
      author={Breil, J},
      author={Galera, S},
      author={Kucharik, M},
      author={Maire, PH},
      author={Shashkov, M},
       title={Two-step hybrid conservative remapping for multimaterial arbitrary Lagrangian-Eulerian methods},
        date={2011},
     journal={J Comput Phys},
      volume={230},
      number={17},
       pages={6664--6687},
}

%==Ref8
\bib{Kucharik20122851}{article}{
      author={Kucharik, M},
      author={Shashkov, M},
       title={One-step hybrid remapping algorithm for multi-material arbitrary Lagrangian-Eulerian methods},
        date={2012},
     journal={J Comput Phys},
      volume={231},
      number={7},
       pages={2851--2864},
}

%==Ref9
\bib{FLD:FLD3996}{article}{
      author={Breil, J},
      author={Alcin, H},
      author={Maire, PH},
       title={A swept intersection-based remapping method for axisymmetric ReALE computation},
        date={2015},
      journal={Int J Numer Meth Fl},
      volume={77},
      number={11},
       pages={694--706},
        note={Fld.3996},
}

%==Ref10
\bib{Barth-97}{incollection}{
      author={Barth, TJ},
       title={Numerical methods for gasdynamic systems on unstructured meshes},
   book={
   title={An Introduction to Recent Developments in Theory and Numerics for Conservation Laws, Proceedings of the International School on Theory and Numerics for Conservation Laws},
      editor={Kroner, D},
      editor={Rohde, C},
      editor={Ohlberger, M},
   publisher={Springer},
   series={Lecture Notes in Computational Science and Engineering}
     address={Berlin},}
        date={1997},
        note={ISBN 3-540-65081-4},
}

%==Ref11
\bib{Liska_etal-Synchronized_FCT-JCP-10}{article}{
      author={Liska, R},
      author={Shashkov, M},
      author={Vachal, P},
      author={Wendroff, B},
       title={Optimization-based synchronized flux-corrected conservative interpolation (remapping) of mass and momentum for arbitrary Lagrangian-Eulerian methods},
        date={2010},
     journal={J Comput Phys},
      volume={229},
      number={5},
       pages={1467--1497},
}

%==Ref12
\bib{Kucharik-Shashkov-Wendroff-JCP-paper-03}{article}{
      author={Kucharik, M},
      author={Shashkov, M},
      author={Wendroff, B},
       title={An efficient linearity-and-bound-preserving remapping method},
        date={2003},
     journal={J Comput Phys},
      volume={188},
      number={2},
       pages={462--471},
}

%==Ref13
\bib{Blanchard-Loubere-MOOD_remap-techrep-15}{misc}{
      author={Blanchard, G},
      author={Loubere, R},
       title={High-Order Conservative Remapping with a posteriori MOOD stabilization on polygonal meshes},
        date={2015},
         note={\url{https://hal.archives-ouvertes.fr/hal-01207156}, the HAL Open Archive, hal-01207156. Accessed January 13, 2016},
}

%==Ref14
\bib{Lauritzen_Erath_Mittal-swept_exact_error_analysis-JCP-11}{article}{
      author={Lauritzen, P},
      author={Erath, C},
      author={Mittal, R},
       title={On simplifying `incremental remap'-based transport schemes},
        date={2011},
     journal={J Comput Phys},
      volume={230},
      number={22},
       pages={7957--7963},
}

%==Ref15
\bib{Klima-Kucharik-analysis}{article}{
      author={Klima, M},
      author={Kucharik, M},
      author={Shashkov, M},
       title={Local error analysis and comparison of the swept- and
  intersection-based remapping methods},
        date={2017},
     journal={Commun Comput Phys},
      volume={21},
      number={2},
       pages={526--558},
}

%==Ref16
\bib{Dukowicz_Baumgardner-advection-JCP-00}{article}{
      author={Dukowicz, JK},
      author={Baumgardner, JR},
       title={Incremental remapping as a transport/advection algorithm},
        date={2000},
     journal={J Comput Phys},
      volume={160},
      number={1},
       pages={318--335},
}

%==Ref17
\bib{Kucharik_Shashkov-MM_remap-FVCA-Proc-11}{incollection}{
      author={Kucharik, M},
      author={Shashkov, M},
       title={Flux-based approach for conservative remap of multi-material
  quantities in 2D arbitrary Lagrangian-Eulerian simulations},
    book={
   title={Finite Volumes for Complex Applications VI Problems \& Perspectives},
      editor={Fo\v{r}t, J},
      editor={F\"{u}rst, J},
      editor={Halama, J},
      editor={Herbin, R},
      editor={Hubert, F},
      series={Springer Proceedings in Mathematics},
      volume={1},
   publisher={Springer},}
        date={2011},
       pages={623--631},
}

%==Ref18
\bib{Kucharik_Shashkov-MM_remap-13}{article}{
      author={Kucharik, M},
      author={Shashkov, M},
       title={Conservative multi-material remap for staggered multi-material
  arbitrary Lagrangian-Eulerian methods},
        date={2014},
     journal={J Comput Phys},
      volume={258},
       pages={268--304},
}

%==Ref19
\bib{Loubere-Shashkov-subcell-remap-JCP-05}{article}{
      author={Loubere, R},
      author={Shashkov, M},
       title={A subcell remapping method on staggered polygonal grids for arbitrary-Lagrangian-Eulerian methods},
        date={2005},
     journal={J Comput Phys},
      volume={209},
      number={1},
       pages={105--138},
}

%==Ref20
\bib{Margolin-Shashkov-remapreport-02}{report}{
      author={Margolin, LG},
      author={Shashkov, M},
       title={Second-order sign-preserving remapping on general grids},
 organization={Los Alamos National Laboratory},
        date={2002},
      number={Technical Report LA-UR-02-525.\nopunct},
}

%==Ref21
\bib{Mavriplis-AIAA-03}{proceedings}{
      author={Mavriplis, DJ},
       title={Revisiting the least-squares procedure for gradient reconstruction on unstructured meshes},
        date={{June 23--26, 2003}},
  conference={AIAA 2003-3986. 16th AIAA Computational Fluid Dynamics Conference}
        address={Orlando, Florida}
}

%==Ref22
\bib{Scovazzi-multi-scale_FE-CMAME-08}{article}{
      author={Scovazzi, G},
      author={Love, E},
      author={Shashkov, M},
       title={Multi-scale Lagrangian shock hydrodynamics on Q1/P0 finite elements: Theoretical framework and two-dimensional computations},
        date={2008},
     journal={Comput Method Appl M},
      volume={197},
      number={9--12},
       pages={1056--1079},
}

%==Ref23
\bib{Caramana-Shashkov-98}{article}{
      author={Caramana, EJ},
      author={Shashkov, MJ},
       title={Elimination of artificial grid distortion and hourglass-type motions by means of Lagrangian subzonal masses and pressures},
        date={1998},
     journal={J Comput Phys},
      volume={142},
      number={2},
       pages={521--561},
}

%==Ref24
\bib{Hoch-ALE-HAL-09}{report}{
      author={Hoch, P},
       title={An arbitrary Lagrangian-Eulerian strategy to solve compressible fluid flows},
      number={Technical Report},
organization={CEA},
        date={2009},
        note={HAL: hal-00366858. \url{https://hal.archives-ouvertes.fr/docs/00/36/68/58/PDF/ale2d.pdf}. Accessed January 13, 2016},
}

%==Ref25
\bib{Shashkov-book-96}{book}{
      author={Shashkov, M},
       title={Conservative Finite-Difference Methods on General Grids},
   publisher={CRC Press},
     address={Boca Raton, Florida},
        date={1996},
        note={ISBN 0-8493-7375-1},
}

%==Ref26
\bib{Knupp-Winslow_smoothing_2D_unstruct-EwC-99}{article}{
      author={Knupp, PM},
       title={Winslow smoothing on two-dimensional unstructured meshes},
        date={1999},
     journal={Eng Comput},
      volume={15},
       pages={263--268},
}

%==Ref27
\bib{ka00}{report}{
      author={Kamm, J},
       title={Evaluation of the Sedov-von Neumann-Taylor blast wave solution},
organization={Los Alamos National Laboratory},
        date={2000},
      number={Technical Report LA-UR-00-6055.\nopunct},
}

%==Ref28
\bib{Taylor499}{article}{
      author={Taylor, GI},
      author={Green, AE},
       title={Mechanism of the production of small eddies from large ones},
        date={1937},
     journal={P Roy Soc Lond A Mat},
      volume={158},
      number={895},
       pages={499--521},
        note={\url{https://doi.org/10.1098/rspa.1937.0036}. \url{http://rspa.royalsocietypublishing.org/content/158/895/499}},
}

\end{thebibliography}

\section*{Author Biography}

\begin{biography}{
%\includegraphics[width=66pt,height=86pt,draft]{blankfig}
}{\textbf{Author Name.} 
This is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text.}
\end{biography}

\end{document}

### 2017-10-02 docs/splitform/letter.tex

\documentclass{letter}
\usepackage{fullpage}
\signature{Jesse Chan}
\address{6100 Main St,\\
Houston, TX 77005}

\begin{document}
\begin{letter}{Dr Mikhail Shashkov\\JCP Editor}

\opening{Dear Dr Shashkov,}

Please find enclosed the manuscript, 
\begin{center}
Jesse Chan\\
\textit{On discretely entropy conservative and entropy stable discontinuous Galerkin methods},
\end{center}
which we would like to submit for publication as a research paper in JCP.

The submitted manuscript describes the construction of high order discontinuous Galerkin (DG) methods which satisfy a discrete conservation or dissipation of entropy.  Existing entropy conservative and entropy stable high order DG methods have been constructed based on diagonal norm summation-by-parts operators, or equivalently nodal collocation and under-integrated (lumped) mass matrices.  This work extends the construction of entropy conservative and entropy stable schemes to non-diagonal mass matrices (or dense norm SBP operators) and over-integrated quadrature rules.  Numerical experiments confirm the entropy conservation and high order accuracy of the new methods based on over-integrated quadrature, and comparisons are made with existing entropy stable methods based on under-integration.

In contrast to existing entropy conservative DG schemes, these new methods are formulated at the continuous level, and the proofs of entropy conservation rely only on properties of quadrature-based integration and $L^2$ projection.  This continuous formulation also makes it possible to construct entropy conservative and entropy-stable methods using non-standard bases and approximation spaces.

We hope that the method and results discussed in this manuscript would appeal to the readership of JCP.  The author confirms that this manuscript has not been published elsewhere and is not under consideration by another journal.  We look forward to hearing from you at your earliest convenience.

\closing{Best regards}

\end{letter}

\end{document}

### 2018-01-17 docs/splitform/main.tex
updating bbwadg cfd files

@@
 In this work, we describe how to construct discretely entropy conservative schemes to a more general class of DG methods using flux differencing, quadrature-based projections, and specific DG differentiation operators.
-This approach also recovers existing methods for Burgers' equation involving dense norm and generalized SBP operators without boundary nodes \cite{ranocha2016summation, ranocha2017extended}.
+This approach also recovers existing methods for Burgers' equation involving dense norm and generalized SBP operators without boundary nodes \cite{fernandez2014generalized, ranocha2016summation, ranocha2017extended}.
 Numerical experiments confirm the stability and high order accuracy of the proposed methods for the one-dimensional compressible Euler equations.
@@
 Similar entropy-stable schemes have been constructed for the shallow water and MHD equations \cite{gassner2016well,  wintermeyer2017entropy, winters2017uniquely}.
-More recently, high order entropy conservative and entropy stable schemes were extended to unstructured triangular meshes in \cite{chen2017entropy} using the split-form methodology of \cite{gassner2016split} and multi-dimensional SBP operators \cite{hicken2016multidimensional}.
+More recently, high order entropy conservative and entropy stable schemes were extended to unstructured triangular meshes in \cite{chen2017entropy} using the split-form methodology of \cite{gassner2016split} and multi-dimensional SBP operators \cite{hicken2016multidimensional, crean2017high}.
 
@@
 This work focuses on the construction of entropy conservative high order DG schemes for systems of conservation laws.
-In order to generalize beyond diagonal norm SBP-DG methods, we will consider non-lumped (dense) mass matrices and more general quadrature rules (e.g.\ rules without boundary points, rules with more points than the dimension of the corresponding approximation space), which are also related to dense norm and generalized SBP operators \cite{ranocha2016summation, ranocha2017extended, ranocha2017comparison}.
+In order to generalize beyond diagonal norm SBP-DG methods, we will consider non-lumped (dense) mass matrices and more general quadrature rules (e.g.\ rules without boundary points, rules with more points than the dimension of the corresponding approximation space), which are also related to dense norm and generalized SBP operators \cite{fernandez2014generalized, ranocha2016summation, ranocha2017extended, ranocha2017comparison}.
 The work presented in this manuscript is intended to serve as a stepping stone to the construction of entropy stable DG methods on general elements in two and three dimensions.

### 2018-01-17 matlab/Codes2D/BuildPeriodicMaps2D.m

@@ -1,7 +1,7 @@
 function BuildPeriodicMaps2D(xperiod, yperiod)
 
 % function [] = BuildPeriodicMaps2D(xperiod, yperiod);
-% Purpose: Connectivity and boundary tables for with all maps returned 
+% Purpose: Connectivity and boundary tables for with all maps returned
 %          in Globals2D assuming periodicity
 
 Globals2D;
@@ -14,75 +14,81 @@ for k1=1:K
         k2 = EToE(k1,f1); f2 = EToF(k1,f1);
         
         vidL = Fmask(:,f1) + (k1-1)*Np; vidR = Fmask(:,f2) + (k2-1)*Np;
-
+        
         fidL = (1:Nfp) + (f1-1)*Nfp + (k1-1)*Nfp*Nfaces;
         fidR = (1:Nfp) + (f2-1)*Nfp + (k2-1)*Nfp*Nfaces;
-
+        
         vmapM(fidL) = vidL; vmapP(fidL) = vidL;
-
+        
         x1  = Fx(fidL')*ones(1,Nfp);  y1  = Fy(fidL')*ones(1,Nfp);
         x2  = Fx(fidR')*ones(1,Nfp);  y2  = Fy(fidR')*ones(1,Nfp);
-
+        
         % Compute distance matrix
         D = (x1-x2').^2+(y1-y2').^2;
-
+        
         [idL, idR] = find(abs(D)<NODETOL);
         vmapP(fidL(idL)) = vidR(idR);
     end
 end
 
 for k1=1:K
-  for f1=1:Nfaces
-    cx1 = sum(x(Fmask(:,f1), k1))/Nfp;
-    cy1 = sum(y(Fmask(:,f1), k1))/Nfp;
-
-    k2 = EToE(k1,f1); f2 = EToF(k1,f1);    
-    if(k2==k1)
-      for k2=1:K
-	    if(k1~=k2)
-	      for f2=1:Nfaces
-	        if(EToE(k2,f2)==k2)
-
-	          cx2 = sum(x(Fmask(:,f2), k2))/Nfp;
-	          cy2 = sum(y(Fmask(:,f2), k2))/Nfp;
-	      
-	          dx = sqrt( (abs(cx1-cx2)-xperiod)^2 + (cy1-cy2)^2);
[…]

### 2018-01-17 matlab/IGA_DG/Wave1D_IGA.m

@@ -66,7 +66,7 @@ p = Pq*pex(xq,0);
 u = zeros(size(x));
 
 %% compute eigs
-if 0
+if 1
     U = zeros(Np*K,2);
     A = zeros(2*Np*K);
     for i = 1:2*Np*K
@@ -77,7 +77,7 @@ if 0
     end
     
     lam = eig(A);
-    rho = max(abs(lam));
+    rho = max(abs(lam))
     return
 end
 %% Solve Problem

### 2018-01-17 matlab/IGA_DG/plotAdvecDispersion1D.m

@@ -1,14 +1,14 @@
 % function advec1D_dispersion
 
 clear
-smoothKnots = 50;
+smoothKnots = 0; 50;
 for Ksub = [1 4 8 16]
     NB = 4;
     K1D = 1;
     N = NB+Ksub-1;
     
     global alpha
-    alpha = 0; % 0 = upwinding
+    alpha = 1; % 0 = upwinding
     
     ndofs = (N+1)*K1D;
     

### 2018-01-17 matlab/IGA_DG/plotLaplSpectra1D.m

@@ -7,14 +7,14 @@ for smoothKnots = [0 50];
     % Order of polymomials used for approximation
     global NB Ksub
     NB = 4;
-    Ksub = 8;
-    K1D = 8;
+    Ksub = 32;
+    K1D = 1;
     
     N = NB+Ksub-1;
     
     FinalTime = 1;
     
-    dofs = (N+1)*K1D
+    dofs = (N+1)*K1D;
     
     [Nv, VX, K, EToV] = MeshGen1D(-1,1,K1D);
     
@@ -83,27 +83,28 @@ for smoothKnots = [0 50];
     
     %% check matrix symmetry
     if 1
-        MM = kron(diag(J(1,:)),M); % global mass matrix
+%         MM = kron(diag(J(1,:)),M); % global mass matrix
+%         
+%         u = zeros(Np,K);
+%         A = zeros(Np*K);
+%         for i = 1:Np*K
+%             u(i) = 1;
+%             [rhsu] = IPDGRHS(u,0);
+%             u(i) = 0;
+%             
+%             A(:,i) = rhsu(:);
+%         end
+%         A = -MM*A; % remove implicit mass matrix inverse
+%         A(abs(A)<1e-10) = 0;
         
-        u = zeros(Np,K);
-        A = zeros(Np*K);
-        for i = 1:Np*K
-            u(i) = 1;
-            [rhsu] = IPDGRHS(u,0);
-            u(i) = 0;
-            
-            A(:,i) = rhsu(:);
-        end
-        A = -MM*A; % remove implicit mass matrix inverse
-        A(abs(A)<1e-10) = 0;
-        
-%         A = Brq'*diag(wq)*Brq;
+MM = Bq'*diag(wq)*Bq;
+        A = Brq'*diag(wq)*Brq;
 %         A(1,:) = 0; A(:,1) = 0; A(1,1) = 1;
 %         A(end,:) = 0; A(:,end) = 0; A(end,end) = 1;
-[W D] = eig(A,MM);
+%         [W D] = eig(A,MM);
         
-%         [W D] = eig(A(2:end-1,2:end-1),MM(2:end-1,2:end-1));
-%         W = [zeros(1,size(W,2));W;zeros(1,size(W,2))];
+        [W D] = eig(A(2:end-1,2:end-1),MM(2:end-1,2:end-1));
+        W = [zeros(1,size(W,2));W;zeros(1,size(W,2))];
         
         [lam p] = sort(diag(D));
[…]

### 2018-01-17 matlab/IGA_DG/printSplineConstants.m

@@ -6,8 +6,8 @@ smoothKnots = 1;'opt';
 
 % for K = [2 4 8 16];
 sk = 1;
-Nvec = 1:7;
-for N = 2:4
+Nvec = 1:5;
+for N = Nvec
         
     % Generate simple mesh
     [Nv, VX, K, EToV] = MeshGen1D(-1,1,K);

### 2018-01-17 matlab/IGA_DG/test_spline_convergence_ppw.m

@@ -1,11 +1,13 @@
 clear
 Globals1D;
 
-smoothKnots = 75;'opt';
+smoothKnots = 75;
+% smoothKnots = 'opt';
 
 K1D = 1;
 
 NB = 4;  Ksub = 16;
+% NB = 3; Ksub = 8;
 % NB = 8; Ksub = 1;
 
 N = NB+Ksub-1; % number of sub-mesh splines
@@ -55,31 +57,38 @@ PBq = M\(Bq'*spdiag(wq));
 Vq = Vandermonde1D(NB,rq)/Vandermonde1D(NB,JacobiGL(0,0,NB));
 Pq = (Vq'*spdiag(wq)*Vq)\(Vq'*spdiag(wq));
 
-kvec = linspace(1/N,N,1000);
+kvec = linspace(1,N+1,250);
+% kvec = 1:N+1;
 sk = 1;
 for k = kvec
     
     %uex = @(x) cos(k*pi*x/2);
-    uex = @(x) sin(k*pi*x);
+%     uex = @(x) sin(k*pi*x);
+uex = @(x) cos((2*k-1)/2*pi*x);
     uB = PBq*uex(xq); % initialize w/lsq fit
     err = Bq*uB-uex(xq);
     err = spdiag(wq)*(err.^2).*J(1);
+%     clf;plot(xq,Bq*uB);hold on;plot(xq,uex(xq),'--');axis([-1,1,-1,1]);title(sprintf('k= %d',k));pause(.1)
     L2errB(sk) = sqrt(sum(err(:)));
     sk = sk + 1;
 end
 
 % error / number of wavelengths
+% hold on
 if smoothKnots==0
-    semilogy(kvec/N,L2errB,'-','linewidth',2)
+    semilogy(kvec/(N+1),L2errB,'-','linewidth',2)
 elseif strcmp(smoothKnots,'opt')==1
-    semilogy(kvec/N,L2errB,'--','linewidth',2)
+    semilogy(kvec/(N+1),L2errB,'--','linewidth',2)
 else
-    semilogy(kvec/N,L2errB,'-.','linewidth',2)
+    semilogy(kvec/(N+1),L2errB,'-.','linewidth',2)
 end
 hold on
 set(gca,'fontsize',15);grid on
 ylabel('L^2 error','fontsize',15)
-axis([0 .5 10^-14 1])
+xlabel('Wavenumber (k) per dof','fontsize',15)
+%axis([0 1 10^-14 1])
[…]

### 2018-01-17 matlab/Wave2D_curved.m

@@ -5,9 +5,10 @@ function  Wave2D_curved
 
 Globals2D;
 
-N = 5;
-nref = 1;
+N = 3;
+nref = 0;
 useJprojection = 1;
+FinalTime = 1;
 
 Nq = 2*N+1;
 
@@ -39,8 +40,19 @@ end
 % Push all boundary faces to unit cylinder
 if (useJprojection >= 0)
     MakeCylinder2D([k,f], 1, 0, 0);
+
+    a = 2/16;
+    r = sqrt(x.^2+y.^2);
+    x = x + a*cos(3*pi*r/2);
+    y = y + a*cos(3*pi*r/2);
+%     plot(x(Fmask(:),:),y(Fmask(:),:),'o'); return
+% vv = cos(3*pi*r/2);
+%     color_line3(x,y,vv,vv,'.')
+% return
+
 end
 cinfo = BuildCurvedOPS2D_opt(Nq,useJprojection);
+%     return
 
 % PlotMesh2D
 
@@ -66,6 +78,7 @@ rfq = [rq1D, -rq1D, -ones(size(rq1D))];
 sfq = [-ones(size(rq1D)), rq1D, -rq1D];
 rfq = rfq(:); sfq = sfq(:);
 wfq = repmat(wq1D,Nfaces,1);
+
 Vfq = Vandermonde2D(N,rfq,sfq)/V;
 [Vrfq Vsfq] = GradVandermonde2D(N,rfq,sfq);
 Vrfq = Vrfq/V;
@@ -107,13 +120,12 @@ for e = 1:K
     Jfq(:,e) = JfqK;
 end
 
-
 global tau;
-tau = 1;
+tau = 0;
 
 
 %%
-if 0 && nargin==0
+if 1 && nargin==0
     
     e = zeros(3*Np*K,1);
     A = zeros(3*Np*K);
@@ -166,19 +178,21 @@ alpha0 = alpha(2); rad = sqrt(x.^2+y.^2);
 
 p = besselj(0, alpha0*rad);
 
-d = 25; x0 = -1/3; y0 = 1/3;
-p = exp(-d*sqrt((x-x0).^2 + (y-y0).^2)) + exp(-10*sqrt((x+x0).^2 + (y+y0).^2));
-
+% d = 25; x0 = -1/3; y0 = 1/3;
[…]

### 2018-01-17 matlab/cfd/Euler1D_local.m

@@ -278,7 +278,7 @@ for i = 1:Nsteps
         if projectV
             rhoq = U1(q1,q2,q3);
             mq   = U2(q1,q2,q3);
-            Eq   = U3(q1,q2,q3); 
+            Eq   = U3(q1,q2,q3);             
         end
         
         [rhs1 rhs2 rhs3] = rhsEuler(rhoq,mq,Eq,INTRK);
@@ -310,6 +310,7 @@ for i = 1:Nsteps
         hold on
         plot(xq,uq,'r-','linewidth',2)
         plot(xq,pq,'k-','linewidth',2)
+        
 %             rhop = Vp*rho;
 %             mp = Vp*m;
 %             Ep = Vp*E;

### 2018-01-17 matlab/cfd2D/Euler2D_affine.m

@@ -1,38 +1,45 @@
-clear
+% clear
 Globals2D
 
-N = 4;
-K1D = 8;
-FinalTime = 5;
-CFL = .4;
+N = 2;
+K1D = 1;
+FinalTime = 10;
+CFL = .25;
 global tau
 tau = 1;
 a = 0/16; % curv warping
 
 Lx = 10; Ly = 5;
-
-[Nv, VX, VY, K, EToV] = unif_tri_mesh(K1D,round(K1D*Ly/Lx));
-VX = VX/max(abs(VX));  VY = VY/max(abs(VY)); 
-VX = VX*Lx + Lx; VY = VY*Ly;
+if 0   
+    [Nv, VX, VY, K, EToV] = unif_tri_mesh(K1D,round(K1D*Ly/Lx));
+    VX = VX/max(abs(VX));  VY = VY/max(abs(VY));
+    VX = VX*Lx + Lx; VY = VY*Ly;
+else
+    [Nv, VX, VY, K, EToV] = unif_tri_mesh(K1D);    
+end
 
 StartUp2D;
 BuildPeriodicMaps2D(max(VX)-min(VX),max(VY)-min(VY));
 
 % plotting nodes
-[rp sp] = EquiNodes2D(15); [rp sp] = xytors(rp,sp);
+[rp sp] = EquiNodes2D(25); [rp sp] = xytors(rp,sp);
 Vp = Vandermonde2D(N,rp,sp)/V;
 xp = Vp*x; yp = Vp*y;
 % plot(xp,yp,'o'); return
 
-global Vq Pq Lq Lqf Vfqf Vfq Pfqf
+global Vq Pq Lq Lqf Vfqf Vfq Pfqf VqPq
 global rxJ sxJ ryJ syJ rxJf sxJf ryJf syJf
 global nxJ nyJ nrJ nsJ nrJq nsJq
 global mapPq
 Nq = 2*N+1;
 [rq sq wq] = Cubature2D(Nq); % integrate u*v*c
+
 Vq = Vandermonde2D(N,rq,sq)/V;
 M = Vq'*diag(wq)*Vq;
 Pq = M\(Vq'*diag(wq)); % J's cancel out
+
+VqPq = Vq*Pq;
+
 xq = Vq*x; yq = Vq*y;
 Jq = Vq*J;
 
@@ -49,6 +56,8 @@ wfq = [wq1D; wq1D; wq1D];
[…]

### 2018-01-17 matlab/elas/Elasticity2D_RayleighWave.m

@@ -13,6 +13,7 @@ if nargin==0
     K1D = 4;
     N = 4;
     FinalTime = 5;
+    tau0 = 1;
 end
 
 % for convergence results
@@ -20,6 +21,7 @@ if nargin==2
     mu = 1;
     lambda = 1;
     FinalTime = 5;
+    tau0 = 0
 elseif nargin==4
     mu = muin;
     lambda = lambdain;
@@ -101,7 +103,7 @@ rho = 1;
 %           0       0                mu];
 
 % plot3(x,y,mu,'.');return
-tau0 = 1;
+
 for fld = 1:5
     tau{fld} = tau0;
     if fld > 2

### 2018-01-17 matlab/elas/Elasticity2D_Stonely.m

@@ -6,17 +6,36 @@ clear -global *
 Globals2D
 
 if nargin==0
-    K1D = 2;
-    N = 4;
+    K1D = 4;
+    N = 3;
 end
 c_flag = 0;
 FinalTime = 5;
 
-% filename = 'Grid/Other/block2.neu';
-% [Nv, VX, VY, K, EToV] = MeshReaderGambit2D(filename);
-aa = 2;
-% [Nv, VX, VY, K, EToV] = unif_tri_mesh(aa*K1D,K1D);
-[Nv, VX, VY, K, EToV] = unif_tri_mesh(3*K1D,aa*K1D);
+tau0 = 0;
+
+aa = 2;    
+if 1
+    if K1D==2
+        filename = 'Grid/Maxwell2D/Maxwell1.neu';
+    elseif K1D==4
+        filename = 'Grid/Maxwell2D/Maxwell05.neu';
+    elseif K1D==8
+        filename = 'Grid/Maxwell2D/Maxwell025.neu';
+    elseif K1D==16
+        filename = 'Grid/Maxwell2D/Maxwell0125.neu';
+    elseif K1D==32
+        filename = 'Grid/Maxwell2D/Maxwell00625.neu';
+    end
+    [Nv, VX, VY, K, EToV] = MeshReaderGambit2D(filename);
+else
+    [Nv, VX, VY, K, EToV] = unif_tri_mesh(3*K1D,aa*K1D+1);
+%     a = .1/K1D;
+%     ids = find(abs(abs(VX)-1) > 1e-8 & abs(abs(VY)-1) > 1e-8);
+%     VX(ids) = VX(ids) + a*randn(size(ids));
+%     VY(ids) = VY(ids) + a*randn(size(ids));
+    
+end
 VY = aa*VY;
 VX = pi*VX;
 
@@ -122,7 +141,6 @@ lambda(ids) = lambda1;
 % rho = Pq*rho;
 % mu = Pq*mu;
 % lambda = Pq*lambda;
-tau0 = 1/2;
 % rhoF = max(rho(vmapM),rho(vmapP));
 % muF = max(mu(vmapM),mu(vmapP));
 % lambdaF = max(lambda(vmapM),lambda(vmapP));

### 2018-05-28 docs/skewform/main.tex
adding skewform paper

\pdfoutput=1

\documentclass[preprint,10pt]{elsarticle}
%\documentclass[preprint,10pt]{article}
%\documentclass[review]{siamart0216}
%\documentclass{siamart0216}

\usepackage{fullpage}
\usepackage[colorlinks=true]{hyperref}

\usepackage{amsmath,amssymb,amsfonts,amsthm}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\theoremstyle{lemma}
\newtheorem{lemma}{Lemma}
\theoremstyle{corollary}
\newtheorem{corollary}{Corollary}
\newtheorem*{remark}{Remark}
\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}
\theoremstyle{assumption}
\newtheorem{assumption}{Assumption}

\usepackage[titletoc,toc,title]{appendix}

\usepackage{array} 
\usepackage{listings}
\usepackage{mathtools}
\usepackage{pdfpages}
\usepackage[textsize=footnotesize,color=green]{todonotes}
\usepackage{bm}
\usepackage{bbm}

\usepackage{tikz}
\usepackage[normalem]{ulem}
\usepackage{hhline}

\usepackage{graphicx}
\usepackage{subfig}
\usepackage{color}

%% ====================================== graphics

\usepackage{pgfplots}
\usepackage{pgfplotstable}
\definecolor{markercolor}{RGB}{124.9, 255, 160.65}
\pgfplotsset{
compat=1.3,
width=10cm,
tick label style={font=\small},
label style={font=\small},
legend style={font=\small}
}

\usetikzlibrary{calc}
\usetikzlibrary{intersections}

%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
\newcommand{\logLogSlopeTriangle}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        \pgfmathsetmacro{\xArel}{#1}
        \pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{\yArel}
        \pgfmathsetmacro{\xCrel}{\xArel}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=north] {}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=west] {\textcolor{black}{#4}}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

\newcommand{\logLogSlopeTriangleNeg}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        \pgfmathsetmacro{\xArel}{#1}
        \pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{\yArel}
        \pgfmathsetmacro{\xCrel}{\xArel}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=.5,anchor=south] {}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=west] {\textcolor{black}{#4}}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
\newcommand{\logLogSlopeTriangleFlipNeg}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        %\pgfmathsetmacro{\xArel}{#1}
        %\pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{#3}
        \pgfmathsetmacro{\xCrel}{#1}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

	\pgfmathsetmacro{\xArel}{\xBrel}
        \pgfmathsetmacro{\yArel}{\yCrel}

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=east] {\textcolor{black}{#4}}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=north] {1}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
\newcommand{\logLogSlopeTriangleFlip}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        %\pgfmathsetmacro{\xArel}{#1}
        %\pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{#3}
        \pgfmathsetmacro{\xCrel}{#1}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

	\pgfmathsetmacro{\xArel}{\xBrel}
        \pgfmathsetmacro{\yArel}{\yCrel}

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=east] {\textcolor{black}{#4}}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=south] {}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

\usepackage{stmaryrd}

\renewcommand{\hat}[1]{\widehat{#1}}
\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

\newcommand{\vect}[1]{\ensuremath\boldsymbol{#1}}
\newcommand{\tensor}[1]{\underline{\bm{#1}}}
\newcommand{\del}{\triangle}
\newcommand{\curl}{\grad \times}
\renewcommand{\div}{\grad \cdot}

\newcommand{\bbm}[1]{\mathbbm{#1}}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\equaldef}{\stackrel{\mathrm{def}}{=}}

\newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}{\rm #2}}}
\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRb}[1]{\left| #1 \right|}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}
\newcommand{\LRceil}[1]{\left\lceil #1 \right\rceil}
\newcommand{\LRl}[1]{\left. #1 \right|}
\newcommand{\pdd}[2]{\frac{\partial^2#1}{\partial#2^2}}
\newcommand{\pdn}[3]{\frac{\partial^{#3}#1}{\partial#2^{#3}}}
\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\snor}[1]{\left| #1 \right|}

%\newcommand{\cond}[1]{\kappa\LRp{#1}}
\newcommand{\cond}[2]{\nor{#1}_{#2}\nor{{#1}^{-1}}_{#2}}

\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\Div} {\ensuremath{\nabla\cdot}}
\newcommand{\jump}[1] {\ensuremath{\llbracket#1\rrbracket}}
\newcommand{\avg}[1] {\ensuremath{\LRc{\!\{#1\}\!}}}

\newcommand{\Oh}{{\Omega_h}}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\LK}{L^2\LRp{D^k}}
\newcommand{\LdK}{L^2\LRp{\partial D^k}}
\newcommand{\Dhat}{\widehat{D}}
\newcommand{\Lhat}{L^2\LRp{\Dhat}}

\newcommand{\eval}[2][\right]{\relax
  \ifx#1\right\relax \left.\fi#2#1\rvert}

\def\etal{{\it et al.~}}

\newcommand{\note}[1]{{\color{blue}{#1}}}
%\newcommand{\noteOne}[1]{{\color{blue}{#1}}}
%\newcommand{\noteTwo}[1]{{\color{red}{#1}}}
%\newcommand{\note}[1]{#1}
%\newcommand{\noteOne}[1]{#1}
%\newcommand{\noteTwo}[1]{#1}

\newcommand{\LinfDk}{L^{\infty}\LRp{D^k}}

\newcommand{\diag}[1]{{\rm diag}\LRp{#1}}

\newcommand{\Ksub}{K_{\rm sub}}

\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

%% d in integrand
\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}}

\makeatletter
\renewcommand\d[1]{\mspace{6mu}\mathrm{d}#1\@ifnextchar\d{\mspace{-3mu}}{}}
\makeatother

%\date{}
%\author{Jesse Chan}
%\title{On discretely entropy conservative discontinuous Galerkin methods}
\graphicspath{{./figs/}}

\begin{document}

%\maketitle

\begin{frontmatter}
\title{Skew-symmetric formulations of entropy stable discontinuous Galerkin methods with applications to hybrid and non-conforming meshes. }

\author[rice]{Jesse Chan\corref{cor1}}
\ead{Jesse.Chan@caam.rice.edu}
\address[rice]{Department of Computational and Applied Mathematics, Rice University, 6100 Main St, Houston, TX, 77005}

\begin{abstract}
Entropy stable high order methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.  
\end{abstract}
\end{frontmatter}

%\tableofcontents

\section{Introduction}

\section{Entropy stability for systems of nonlinear conservation laws}

We begin by reviewing the dissipation of entropy for a $d$-dimensional system of nonlinear conservation laws on a domain $\Omega$
\begin{equation}
\pd{\bm{u}}{t}  + \sum_{j=1}^d\pd{\bm{f}_j(\bm{u})}{x_j} = 0, \qquad \bm{u}\in \mathbb{R}^n, \qquad \bm{f}:\mathbb{R}^n\rightarrow\mathbb{R}^n,
\label{eq:nonlineqs}
\end{equation}
where $\bm{u}$ are the conservative variables and $\bm{f}(\bm{u})$ is a vector-valued nonlinear flux function.  We are interested in nonlinear conservation laws for which a convex entropy function $U(\bm{u})$ exists.  For such systems, the  \emph{entropy variables} are an invertible mapping $\bm{v}(\bm{u}):\mathbb{R}^n\rightarrow \mathbb{R}^n$ defined as the derivative of the entropy function with respect to the conservative variables 
\begin{align}
\bm{v}(\bm{u}) = \pd{U}{\bm{u}}.%, \qquad  \text{ invertible}.%\bm{u}\LRp{\bm{v}(\bm{u})} = \bm{u}.
\label{eq:entropyvarsmap}
\end{align}
Several widely used equations in fluid modeling (Burgers, shallow water, compressible Euler and Navier-Stokes equations) yield convex entropy functions $U(\bm{u})$ \cite{hughes1986new, chen2017entropy}.  Let $\partial \Omega$ be the boundary of $\Omega$ with outward unit normal $\bm{n}$.  By multiplying the equation (\ref{eq:nonlineqs}) with $\bm{v}(\bm{u})^T$, the solutions $\bm{u}$ of (\ref{eq:nonlineqs}) can be shown to satisfy an entropy inequality
\begin{equation}
\int_{\Omega}\pd{U(\bm{u})}{t}\diff{x} + \int_{\partial \Omega} \sum_{j=1}^d \LRp{\bm{v}(\bm{u})^T\bm{f}_j(\bm{u}) - \psi_j\LRp{\bm{v}(\bm{u})}}n_j \diff{x} \leq 0, \qquad \bm{n} = \LRp{n_1,\ldots,n_d}, 
\label{eq:entropyineq}
\end{equation}
where $\psi_j(\bm{u})$ is a function referred to as the entropy potential.

The proof of (\ref{eq:entropyineq}) requires the use of the chain rule \cite{mock1980systems, harten1983symmetric, dafermos2005compensated}.  The instability-in-practice of high order schemes for (\ref{eq:nonlineqs}) can be attributed in part to the fact that the discrete form of the equations do not satisfy the chain rule, and thus do not satisfy (\ref{eq:entropyineq}).  As a result, discretizations of (\ref{eq:nonlineqs}) do not typically possess an underlying statement of stability.  For low order schemes, this can be offset in practice by the inherent numerical dissipation.  However, because high order discretizations possess low numerical dissipation, the lack of an underlying discrete stability has been conflated with the idea that high order methods are inherently less stable than low order methods.

\section{Approximation spaces}

We define an approximation space using degree $N$ polynomials on the reference element.  In this work, we consider either simplicial reference elements (triangles and tetrahedra) or tensor product reference elements (quadrilaterals and hexahedra).  The natural polynomial approximation space differs depending on the element type \cite{chan2015gpu}.

On a $d$-dimensional reference simplex, the natural polynomial space are total degree $N$ polynomials 
\[
P^N\LRp{\widehat{D}} = \LRc{\hat{x}_1^{i_1}\ldots\hat{x}_d^{i_d}, \quad \hat{\bm{x}} \in \widehat{D}, \quad 0\leq \sum_{k=1}^d i_k \leq N}.
\]
In contrast, the natural polynomial space on a $d$-dimensional tensor product element is 
\[
Q^N\LRp{\widehat{D}} = \LRc{\hat{x}_1^{i_1}\ldots\hat{x}_d^{i_d}, \quad \hat{\bm{x}} \in \widehat{D}, \quad 0\leq i_k \leq N, \quad k = 1,\ldots, d}.
\]
We take $V^N$ to denote the natural approximation space on a given element.  Furthermore, we denote the dimension of $V^N$  as $N_p = {\rm dim}\LRp{V^N\LRp{\widehat{D}}}$.

We also define trace spaces for each reference element.  Let $\hat{f}$ be a face of the reference element $\hat{D}$.  The trace space $V^N \LRp{\hat{f}}$ is defined as the restrictions of functions in $V^N\LRp{\hat{D}}$ to $\hat{f}$
\[
V^N \LRp{\hat{f}} = \LRc{ \left.u\right|_{\hat{f}}, \quad u \in V^N\LRp{\hat{D}}}, \qquad \hat{f}\in \partial\hat{D}.
\]
For example, the trace space for degree $N$ polynomials on a quadrilateral face $\hat{f}$ of the bi-unit hexahedral element $[-1,1]^3$ is $Q^N\LRp{\hat{f}}$, while the trace space for degree $N$ polynomials on a triangular face $f$ of the tetrahedron is $P^N\LRp{\hat{f}}$.  We denote the dimension of the trace space as ${\rm dim}\LRp{V^N\LRp{{\hat{f}}}} = N^f_p$.

We similarly define the trace space for the surface $\partial \hat{D}$ of $\hat{D}$
\[
V^N \LRp{\partial \hat{D}} = \LRc{ \left.u\right|_{\partial \hat{D}}, \quad u \in V^N\LRp{\hat{D}}}.
\]

\section{$L^2$ projection and lifting matrices and a decoupled SBP operator}

The schemes in \cite{chan2017discretely} begin on the reference element $\hat{D}$ by approximating the solution in a degree $N$ polynomial basis $\LRc{\phi_j({\bm{x}})}_{i=1}^{N_p}$.  It is assumed that volume and surface quadrature rules $({\bm{x}}_i, w_i)$, $\LRp{{\bm{x}}^f_i,w^f_i}$ are provided for $\hat{D}$ and the surface $\partial \hat{D}$ of $\hat{D}$, respectively.  We make the following assumption on the volume quadrature rule:
\begin{assumption}
The volume quadrature rule is exact for polynomials of degree $(2N-1)$.
\label{ass:quad}
\end{assumption}
In addition to being related to the positive definiteness of the mass matrix, Assumption~\ref{ass:quad} ensures that, for any $u,v \in V^N$, the following integral is exactly approximated using quadrature
\[
\int_{\hat{D}} \pd{u(\hat{\bm{x}})}{\hat{x}_i} v(\hat{\bm{x}})\diff{\hat{\bm{x}}} = \sum_{k=1}^{N_q} w_k \pd{u(\hat{\bm{x}}_k)}{\hat{x}_i} v(\hat{\bm{x}}_k).  
\]
In other words, evaluating the integral of the derivative of $u$ and $v$ using quadrature is exact because the integrand is a polynomial of degree $(2N-1)$.

\note{TODO: Standardize hat notation for operators/normals.}

Let $\bm{V}_q,\bm{V}_f$ denote interpolation matrices, and let $\bm{D}^i$ be the differentiation matrix with respect to the $i$th coordinate such that
\begin{gather}
\LRp{\bm{V}_q}_{ij} = \phi_j(\bm{x}_i), \qquad \LRp{\bm{V}_f}_{ij} = \phi_j(\bm{x}^f_i), \qquad \pd{\phi_j(\bm{x})}{x_i} = \sum_{k=1}^{N_p} \LRp{\bm{D}^i_{jk}} \phi_k(\bm{x}).
\end{gather}
The interpolation matrices $\bm{V}_q,\bm{V}_f$ map basis coefficients to evaluations at volume and surface quadrature points respectively, while the differentiation matrix ${\bm{D}}_i$ maps basis coefficients of a function to the basis coefficients of its derivative with respect to $x_k$.  The interpolation matrices are used to assemble the mass matrix $\bm{M}$, the quadrature-based projection matrix $\bm{P}_q$, and lifting matrix $\bm{L}_f$
\begin{gather}
\bm{M} = \bm{V}_q^T\bm{W}\bm{V}_q, \qquad \bm{P}_q = \bm{M}^{-1}\bm{V}_q^T\bm{W}, \qquad \bm{L}_f = \bm{M}^{-1}\bm{V}_f^T\bm{W}_f,
\end{gather}
where $\bm{W}, \bm{W}_f$ are diagonal matrices of volume and surface quadrature weights, respectively.  The matrix $\bm{P}_q$ is a quadrature-based discretization of the $L^2$ projection operator $\Pi_N$ onto degree $N$ polynomials.

Interpolation, differentiation, and $L^2$ projection matrices can be combined to construct finite difference operators.  For example, the matrix $\bm{D}_q = \bm{V}_q\bm{D}^i\bm{P}_q$ maps function values at quadrature points to approximate values of the derivative at quadrature points.  By choosing specific quadrature rules, $\bm{D}_q$ recovers high order summation-by-parts finite difference operators in \cite{gassner2013skew, fernandez2014generalized, ranocha2018generalised} and certain operators in \cite{hicken2016multidimensional}.  However, to address difficulties in designing efficient entropy stable interface terms for nonlinear conservation laws, the PI introduced in \cite{chan2017discretely} a new ``decoupled'' summation by parts matrix which builds interface terms directly into the approximation of the derivative.

Let $\bm{n}$ denote the outward unit normal $\bm{n}$ on $\hat{D}$, and let $\bm{n}_i$ denote the vector containing values of the $i$th component at all surface quadrature points.  The ``decoupled'' summation by parts operator $\bm{D}^i_N$ is defined as the block matrix involving both volume and surface quadratures
\begin{align}
\bm{D}^i_N  &= \LRs{
\begin{array}{cc}
\bm{D}^i_q - \frac{1}{2}\bm{V}_q \bm{L}_f {\rm diag}({\bm{n}}_i) \bm{V}_f\bm{P}_q &  \frac{1}{2}\bm{V}_q\bm{L}_f{\rm diag}({\bm{n}}_i)\\
-\frac{1}{2}{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q & \frac{1}{2}{\rm diag}({\bm{n}}_i)
\end{array}}.  
\label{eq:DN}
\end{align}
When combined with projection and lifting matrices, $\bm{D}^i_N$ produces a high order polynomial approximation of $f\pd{g}{x_i}$. 
Let $f, g$ be differentiable functions, and let $\bm{f}_i = f(\bm{x}_i)$, $\bm{g}_i = g(\bm{x}_i)$ denote values of $f,g$ at volume and surface quadrature points.  Then,
\begin{gather}
f\pd{g}{x_i} \approx \LRs{\begin{array}{cc}
\bm{P}_q & \bm{L}_f\end{array}} {\rm diag}\LRp{\bm{f}}\bm{D}^i_N \bm{g}.  
\label{eq:dnapprox}
\end{gather}
The approximation can also be interpreted as solving a variational problem.  Let $n_i$ be the $i$th component of the unit normal on $\hat{D}$.  Then (\ref{eq:dnapprox}) is equivalent to finding $u(\bm{x}) \approx f\pd{g}{x_i}$ such that 
\[
\int_{\hat{D}} u(\bm{x})v(\bm{x})  = \int_{\hat{D}}{g\pd{\Pi_Nf}{x_i}v} + \int_{\partial \hat{D}}{(f-\Pi_Nf)\frac{\LRp{gv + \Pi_N(gv)}}{2}}n_i, \qquad \forall v\in V^N\LRp{\hat{D}}.
\]
If the surface quadrature rule is exact for polynomials of degree $2N$, then the decoupled SBP operator $\bm{D}^i_N$ also satisfies a summation by parts property.  
\begin{gather}
\bm{Q}^i_N = \bm{W}_N \bm{D}^i_N, \qquad \bm{W}_N = \LRp{\begin{array}{cc}\bm{W}&\\ & \bm{W}_f\end{array}},\label{eq:dsbp}\\ \bm{Q}^i_N+\LRp{\bm{Q}^i_N}^T = \LRp{\begin{array}{cc}\bm{0}&\\ & \bm{W}_f {\rm diag}(\bm{n}_i)\end{array}} = \bm{B}_N\nonumber
\end{gather}
By noting that $\bm{W}\bm{V}_q \bm{L}_q = \bm{P}_q^T\bm{V}_f^T\bm{W}_f$, the entries of $\bm{Q}^i_N$ can be re-written as
\begin{align}
\bm{Q}^i_N  &= \LRs{
\begin{array}{cc}
\bm{Q}_i - \frac{1}{2}\LRp{\bm{V}_f \bm{P}_q}^T  \bm{W}_f{\rm diag}({\bm{n}}_i) \bm{V}_f\bm{P}_q &  \frac{1}{2}\LRp{\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q}^T\\
-\frac{1}{2}\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q & \frac{1}{2}\bm{W}_f{\rm diag}({\bm{n}}_i)
\end{array}}, \qquad \bm{Q}_i = \bm{W}\bm{V}_q\bm{D}^i\bm{P}_q,
\label{eq:QN}
\end{align}
from which the SBP property (\ref{eq:dsbp}) is more apparent.

%\begin{lemma}[Lemma 1 in \cite{chan2017discretely}]
%Let $\bm{D}^i_q = \bm{V}_q\bm{D}^i\bm{P}_q$.  Then, for volume quadratures which are exact for degree $2N-1$ polynomials, $\bm{D}^i_q$ satisfies the SBP property with respect to $\bm{W}$ 
%\[
%\bm{W}\bm{D}^i_q = \LRp{\bm{V}_f\bm{P}_q}^T \bm{W}_f\diag{\hat{\bm{n}}} \bm{V}_f\bm{P}_q - \LRp{\bm{W}\bm{D}^i_q}^T.  
%\]
%\label{lemma:gsbp}
%\end{lemma}

\section{Skew-symmetric entropy conservative formulations}

\subsection{Loss of the SBP property for hybrid and non-conforming meshes}
While the SBP property has been used to derive entropy stable schemes, the SBP property is difficult or impossible to enforce in certain discretization settings, such as hybrid and non-conforming meshes arises.  This difficulty is a result of the choices of surface quadrature which naturally arise for these settings.

\begin{figure}
\centering
\begingroup
\captionsetup[subfigure]{width=.43\textwidth}
\subfloat[Insufficiently accurate surface quadrature on the triangle element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D.png}\label{subfig:hybrid1}}
\endgroup
\hspace{2em}
\subfloat[Incompatible surface quadrature on the quadrilateral element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D_GQ.png}\label{subfig:hybrid2}}
\caption{Examples of interface couplings which do not result in an SBP property and are not entropy stable.  On the left, the surface quadrature is a $(N+1)$ point GLL rule, and results in a loss of the SBP property on the triangle.  On the right, the surface quadrature is a $(N+1)$ point Gauss-Legendre rule, and results in a loss of the SBP property on the quadrilateral.  }
\label{fig:hybrid}
\end{figure}

\begin{figure}
\centering
\subfloat[Hex-pyramid coupling]{\includegraphics[width=.35\textwidth]{figs/hybrid3D.png}}
\hspace{2em}
\subfloat[Hex-prism coupling]{\includegraphics[width=.375\textwidth]{figs/hybrid3D_wedge.png}}
\caption{Illustration of a 3D coupling between a GLL hexahedral element and a pyramid.  The SBP property does not hold on the pyramid due to the use of GLL quadrature on the quadrilateral face. }
\label{fig:hybrid3d}
\end{figure}

\begin{figure}[!h]
\centering
\begingroup
\captionsetup[subfigure]{width=.425\textwidth}
\subfloat[Coupling of nodes for a non-conforming interface (entropy stable but tighter coupling).]{\raisebox{0em}{\includegraphics[width=.425\textwidth]{figs/noncon.png}}\label{subfig:noncon2}}
\hspace{2em}
\subfloat[Surface quadrature rule induced by non-conforming refinement (no SBP property).]{\raisebox{-.3em}{\includegraphics[width=.45\textwidth]{figs/nonconQuad.png}}\label{subfig:noncon1}}
\endgroup
%\subfloat[3D hexahedra-pyramid coupling]{\includegraphics[width=.31\textwidth]{figs/hybrid3D.png}}
\caption{Illustration of two different entropy stable inter-element couplings for $h$ non-conforming meshes.  Each dashed red line indicates a dyadic flux computation required between two nodes.  The coupling terms in \cite{friedrich2017entropy} requires computing dyadic fluxes between \emph{each} pair of nodes on adjacent interfaces, while the proposed approach requires dyadic flux computations only between immediately adjacent nodes.   }
\label{fig:noncon}
\end{figure}

\subsection{A variational SBP property}

It can be shown that the SBP property holds for $\bm{D}^i_N$ if the surface quadrature is exact for polynomials of degree $2N$ \cite{chan2017discretely}.  The SBP property is lost if the surface quadrature is only exact for polynomial integrands of degree less than $2N$.  However, one can still show that a variational version of the SBP property is satisfied.

\begin{lemma}
Let $v\in V^N$, and let the surface trace $\LRl{v}_{\partial \hat{D}} \in V^M\LRp{\partial \hat{D}}$ for some $M \leq N$. If the surface quadrature is exact for polynomials of degree $N+M$,  then for an arbitrary vector $\bm{u}_q$, 
\[
\bm{v}_q^T\bm{Q}_i \bm{u}_q = \bm{v}_q^T\LRp{ \LRp{\bm{V}_f\bm{P}_q}^T \bm{W}_f \diag{{\bm{n}}_i}\bm{V}_f\bm{P}_q - \bm{Q}_i^T}\bm{u}_q
\]
where $\bm{v}$ denotes the vector of values of $v$ of at both volume and surface quadrature points. 
\label{lemma:vsbp}
\end{lemma}
\begin{proof}
Recall that $\bm{Q}_i = \bm{W} \bm{V}_q \bm{D}^i\bm{P}_q$.  Since the volume quadrature is exact for degree $2N-1$ polynomials, we have that
\begin{align*}
\bm{v}_q^T\bm{Q}_i \bm{u}_q &= \bm{v}_q^T\bm{W} \bm{D}^i \bm{P}_q \bm{u}_q = \int_{\hat{D}} \pd{\Pi_N u}{\hat{x}_i} v = \int_{\partial \hat{D}} (\Pi_N u) v \hat{n}_i - \int_{\hat{D}} \LRp{\Pi_N u} \pd{v}{\hat{x}_i}.
\end{align*}
Since the volume integrand $\LRp{\Pi_N u} \pd{v}{\hat{x}_i} \in V^{2N-1}$, it is exactly integrated using the volume quadrature.  Additionally, since $\Pi_N u \in V^N$ and $v\in V^M$ on the surface of $\hat{D}$, the surface quadrature is exact for the surface integral and 
\begin{align*}
\int_{\partial \hat{D}} (\Pi_N u) v \hat{n}_i -   \int_{\hat{D}} \LRp{\Pi_N u} \pd{v}{\hat{x}_i} &= \bm{v}_f^T\bm{W}_f{\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q\bm{u}_q - \LRp{\bm{V}_q\bm{D}^i\bm{P}_q\bm{v}_q}^T\bm{W} \bm{V}_q\bm{P}_q\bm{u}_q\\
&= \bm{v}_q^T \LRp{\bm{V}_f\bm{P}_q}^T\bm{W}_f{\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q\bm{u}_q - \bm{v}_q^T \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T \bm{W} \bm{V}_q\bm{P}_q\bm{u}_q,
\end{align*}
where we have used that, since $v\in V^N$, $\bm{v}_f = \bm{V}_f\bm{P}_q\bm{v}_q$.  The proof is completed by noting that 
\[
\bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T \bm{W} \bm{V}_q\bm{P}_q = \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{M}\bm{P}_q = \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T\bm{W} = \bm{Q}_i^T.
\]
\end{proof}

The variational SBP property also extends to the decoupled SBP operator.  This property, along with the exact differentiation of constants, is necessary for the proof of entropy stability.  
\begin{lemma}
Let $\bm{D}^i_N$ be a decoupled SBP operator on the reference element $\hat{D}$, and let the surface quadrature be exact for polynomials of degree $N+M$ for some $M \leq N$.  Suppose $v\in V^N$ and that $\LRl{v}_{\partial \hat{D}} \in P^M\LRp{\partial \hat{D}}$, Then, 
\[
\bm{v}^T\bm{Q}^i_N\bm{u} = \bm{v}^T\LRp{\bm{B}^i_N - \LRp{\bm{Q}^i_N}^T}\bm{u}.%, \qquad \bm{Q}^i_N \bm{1} = \bm{0}.
\]
where $\bm{v}$ denotes the values of $v$ of at both volume and surface quadrature points.  
\label{lemma:vdsbp}
\end{lemma}
\begin{proof}
For convenience, let $\bm{u}_q, \bm{u}_f$ denote evaluations of $u$ at volume and surface points, such that 
\[
\bm{u} = \begin{pmatrix} \bm{u}_q\\ \bm{u}_f\end{pmatrix}, \qquad \bm{v} = \begin{pmatrix} \bm{v}_q\\ \bm{v}_f\end{pmatrix}.  
\]
The proof of the variational summation by parts property uses the definition of $\bm{Q}^i_N$ (\ref{eq:QN}), 
\begin{align*}
\bm{v}^T\bm{Q}^i_N\bm{u} &= \bm{v}_q^T\bm{Q}_i \bm{u}_q - \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}} \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}} \bm{u}_f\\
& - \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}} \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} + \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}} \bm{u}_f.
\end{align*}
Applying Lemma~\ref{lemma:vsbp} then yields
\begin{align*}
\bm{v}^T\bm{Q}^i_N\bm{u} &= -\bm{v}_q^T\bm{Q}^T_i \bm{u}_q + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}} \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}} \bm{u}_f\\
&\qquad - \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}} \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} - \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}} \bm{u}_f + \bm{v}_f^T\bm{W}_f\hat{\bm{n}} \bm{u}_f\\
&= \begin{pmatrix} \bm{v}_q\\ \bm{v}_f\end{pmatrix}^T 
\LRp{\begin{pmatrix}
\bm{0}& \\
& \bm{W}_f\diag{\hat{\bm{n}}}
\end{pmatrix}
+
\begin{pmatrix}
-\bm{Q}_i^T + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q}^T\bm{W}_f\hat{\bm{n}} \bm{V}_f\bm{P}_q & -\frac{1}{2} \LRp{\bm{W}_f\hat{\bm{n}} \bm{V}_f\bm{P}_q}^T\\
\frac{1}{2}\bm{W}_f\hat{\bm{n}} \bm{V}_f\bm{P}_q & -\frac{1}{2}\bm{W}_f\hat{\bm{n}}
\end{pmatrix} } 
\begin{pmatrix} \bm{u}_q\\ \bm{u}_f\end{pmatrix}\\
&= \bm{v}^T\LRp{\bm{B}_N - \LRp{\bm{Q}^i_N}^T}\bm{u}.
\end{align*}
\end{proof}
Using Lemma~\ref{lemma:vdsbp}, we will determine the accuracy of the surface quadrature necessary to ensure the proof of semi-discrete entropy stability is valid.  This minimal degree of surface quadrature accuracy will depend on the nature of the reference-to-physical mapping (e.g.\ affine vs curved elements).

\begin{corollary}
\label{lemma:sbpcor}
Let the surface quadrature be exact for polynomials of degree $N$.  Then, 
\[
\bm{1}^T\bm{Q}^i_N\bm{u} = \bm{1}^T\bm{B}_N\bm{u}, \qquad \bm{Q}^i_N\bm{1} = \bm{0},
\]
where $\bm{u}$ is an arbitrary vector of appropriate size.  
\end{corollary}
\begin{proof}
The proof that $\bm{Q}^i_N \bm{1} = \bm{0}$ follows from the property that polynomials are equal to their $L^2$ projection, and is identical to that of \cite{chan2017discretely,chan2018discretely}.     The proof of the first equality follows $\bm{Q}^i_N \bm{1} = \bm{0}$ and Lemma~\ref{lemma:vdsbp} with $M=0$
\[
\bm{1}^T\LRp{\bm{Q}^i_N}\bm{u} = \bm{1}^T\LRp{\bm{B}_N - \LRp{\bm{Q}^i_N}^T}\bm{u} = \bm{1}^T{\bm{B}_N}\bm{u}.
\]
\end{proof}
\subsection{Entropy stability on affine meshes}

The high order methods in this work ensure that (\ref{eq:entropyineq}) is satisfied discretely by avoiding the use of the chain rule in the proof of entropy dissipation.  These ``entropy stable'' schemes rely on two main ingredients: an entropy stable numerical flux as defined by Tadmor \cite{tadmor1987numerical} and a concept referred to as ``flux differencing''.  % entropy stable schemes in one-dimension.  
Let $\bm{f}_S\LRp{\bm{u}_L,\bm{u}_R}$ be a numerical flux function which is a function of ``left'' and ``right'' states $\bm{u}_L,\bm{u}_R$.  
The numerical flux $\bm{f}_S$ is \textit{entropy conservative} if it satisfies the following three conditions:  
%The numerical flux $\bm{f}_S$ is an \textit{entropy stable} flux function if it satisfies the following three conditions:  
\begin{gather}
\bm{f}_S(\bm{u},\bm{u}) = \bm{f}(\bm{u}), \qquad \text{(consistency)}\\
\bm{f}_S(\bm{u}_L,\bm{u}_R) = \bm{f}_S(\bm{u}_R,\bm{u}_R), \qquad \text{(symmetry)}\nonumber\\
\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}_S(\bm{u}_L,\bm{u}_R) = \psi(\bm{u}_L) - \psi(\bm{u}_R), \qquad \text{(conservation)}\nonumber
%\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}_S(\bm{u}_L,\bm{u}_R) \leq \psi(\bm{u}_L) - \psi(\bm{u}_R), \qquad \text{(entropy dissipation)}\nonumber
\label{eq:esflux}
\end{gather}
If instead of the third condition, $\bm{f}_S$ satisfies a \emph{stability} property $\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}_S(\bm{u}_L,\bm{u}_R) \leq \psi(\bm{u}_L) - \psi(\bm{u}_R)$, the flux $\bm{f}_S$ is referred to as \textit{entropy stable}.

We begin by deriving a skew-symmetric formulation on the reference element $\hat{D}$, and showing it is entropy conservative.  This formulation can then be extended in a straightforward manner to affine elements.  The skew-symmetric formulation for (\ref{eq:nonlineqs}) on $\hat{D}$ is 
\begin{gather}
\pd{\bm{u}}{t} + \sum_{i=1}^d \LRs{\begin{array}{cc}
\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{\hat{\bm{n}}}\bm{f}_i^* = 0,\\
\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q,
\label{eq:esdgSkew}
\end{gather}
where $\bm{f}^*$ is some numerical flux.  Multiplying the formulation (\ref{eq:esdgSkew}) by $\bm{M}$ on both sides yields a weak form 
\begin{gather}
\bm{M}\pd{\bm{u}}{t} + \sum_{i=1}^d \LRs{\begin{array}{c}
\bm{V}_q \\ \bm{V}_f\end{array}}^T \begin{pmatrix}
\bm{W}&\\
&\bm{W}_f
\end{pmatrix}
\LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} +  \bm{V}_f^T\bm{W}_f\diag{\hat{\bm{n}}}\bm{f}_i^*\label{eq:esdgSkewWeak}\\
= \bm{M}\pd{\bm{u}}{t} + \sum_{i=1}^d\LRs{\begin{array}{c}
\bm{V}_q \\ \bm{V}_f\end{array}}^T 
\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{V}_f^T\bm{W}_f \diag{\hat{\bm{n}}}\bm{f}_i^* = 0.  \nonumber
\end{gather}

We can now show that the skew-symmetric formulation is semi-discretely entropy conservative.  
\begin{theorem}
The formulation (\ref{eq:esdgSkew}) is entropy conservative for any surface quadrature which is exact for polynomials of degree $N$, such that
\[
\bm{1}^T\bm{W}\pd{U(\bm{u}_q)}{t} + \sum_{i=1}^d\bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}} \LRp{\psi_i(\tilde{\bm{u}}_f) - \tilde{\bm{v}}_f^T\bm{f}_i^*} = 0, \qquad \bm{u}_q = \bm{V}_q\bm{u}.
\]
\label{thm:esdg}
\end{theorem}
\begin{proof}
Testing (\ref{eq:esdgSkewWeak}) by $\bm{v}_h = \bm{P}_q\bm{v}_q$ yields 
\begin{align}
\bm{v}_q^T\bm{W}\pd{\LRp{\bm{V}_q\bm{u}}}{t} + \sum_{i=1}^d
\tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \tilde{\bm{v}}_f^T \bm{W}_f \diag{\hat{\bm{n}}}\bm{f}_i^* = 0.
\end{align}
One can show that \cite{chan2017discretely}
\begin{align*}
\tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} &= \tilde{\bm{v}}^T \LRp{\bm{Q}^i_N \circ \bm{F}_S}\bm{1} - \tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N}^T \circ \bm{F}_S}\bm{1}\\
&= \tilde{\bm{v}}^T \LRp{\bm{Q}^i_N \circ \bm{F}_S}\bm{1} - \bm{1}^T \LRp{{\bm{Q}^i_N} \circ \bm{F}_S}\tilde{\bm{v}}.
\end{align*}
Where we have used that $\bm{F}_S$ is symmetric and that the Hadamard product commutes.  Applying the conservation condition on $\bm{f}_S$ in (\ref{eq:esflux}) then yields
\begin{align*}
\tilde{\bm{v}}^T \LRp{\bm{Q}^i_N \circ \bm{F}_S}\bm{1} - \bm{1}^T \LRp{{\bm{Q}^i_N} \circ \bm{F}_S}\tilde{\bm{v}} &= \sum_{jk} \LRp{\bm{Q}^i_N}_{jk} \LRp{\tilde{\bm{v}}_j-\tilde{\bm{v}}_k} \bm{f}_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k} \\
&= \sum_{ij} \LRp{\bm{Q}^i_N}_{jk} \LRp{\psi_i(\tilde{\bm{u}}_j) - \psi_i(\tilde{\bm{u}}_k)}\\
&= \bm{1}^T\LRp{\bm{Q}^i_N}\psi_i(\tilde{\bm{u}}) - \psi_i(\tilde{\bm{u}})^T\LRp{\bm{Q}^i_N}\bm{1} \\
&= \bm{1}^T\LRp{\bm{Q}^i_N}\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i_N\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}} \psi_i(\tilde{\bm{u}}_f)
\end{align*}
where we have used Corollary~\ref{lemma:sbpcor} in the second to last equality.
%The final step of the proof is  where we have used Lemma~\ref{lemma:vdsbp} to integrate by parts and conclude that $\bm{1}^T\LRp{\bm{Q}^i_N}^T = 0$.
\end{proof}

\begin{remark}
The only result necessary to prove Theorem~\ref{thm:esdg} is Corollary~\ref{lemma:sbpcor}.  Thus, when generalizing to curved elements, the proof of entropy stability for the skew-symmetric form will depend only on the curved version of Corollary~\ref{lemma:sbpcor}.
\end{remark}

The skew symmetric formulation can also be shown to be locally conservative in the sense of \cite{shi2017local}, which is sufficient to show the numerical solution convergences to the weak solution under mesh refinement.  
\begin{theorem}
The formulation (\ref{eq:esdgSkew}) is locally conservative such that
\begin{align}
\bm{1}^T\bm{W}\pd{\LRp{\bm{V}_q\bm{u}}}{t} + \sum_{i=1}^d\bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}_i}\bm{f}_i^* = 0. 
\end{align}
\end{theorem}
\begin{proof}
To show local conservation, we test (\ref{eq:esdgSkewWeak}) with $1$.  Let $\bm{e}$ be the coefficients of $1$ under the basis $\phi_j(\bm{x})$.
\begin{align}
\bm{1}^T\bm{W}\bm{V}_q\pd{\bm{u}}{t} + \sum_{i=1}^d
\bm{1}^T
\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}}\bm{f}_i^* = 0. 
\end{align}
Because $\bm{F}_S$ is symmetric and $\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}$ is skew-symmetric, the term $\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}$ is a skew-symmetric matrix.  The volume term vanishes
\[
\bm{1}^T\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} = 0,
\]
where we have used that, for any skew symmetric matrix $\bm{A}$, $\bm{x}^T\bm{A}\bm{x} = 0$.
% and that the Hadamard product is commutable $\LRp{\bm{A}\circ\bm{B} }^T = \bm{A}^T\circ\bm{B}^T$
%\begin{align}
%\bm{1}^T\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} &= \bm{1}^T\LRp{\bm{Q}^i_N\circ \bm{F}_S}\bm{1} - \bm{1}^T\LRp{\LRp{\bm{Q}^i_N}^T\circ \bm{F}_S}\bm{1}\\
%&= \bm{1}^T\LRp{\bm{Q}^i_N\circ \bm{F}_S}\bm{1} - 
%\bm{1}^T\LRp{\bm{Q}^i_N\circ \bm{F}_S}\bm{1} = 0,
%\end{align}
%where we have used that $\bm{F}_S$ is symmetric.
%\begin{align}
%\end{align}
\end{proof}

Derivatives with respect to physical coordinates on $D^k$ are computed in terms of a change of variables formula and derivatives with respect to reference coordinates 
\[
\pd{u}{x_i} = \sum_{ij} \bm{G}^k_{ij}\pd{u}{\hat{x}_j}, \qquad \bm{G}^k_{ij} = J^k\pd{\hat{x}_j}{{x}_i}, 
\]
where $J^k$ is the determinant of the Jacobian of the geometric mapping on the element $D^k$.
For affine meshes, the scaled geometric terms $\bm{G}^k_{ij}$ are constant over each element.  It was shown in \cite{chan2017discretely} that one can define a physical differentiation matrix as a linear combination of reference differentiation matrices.  Let $\hat{\bm{D}}^i_N$ denote the decoupled SBP operator on the reference element $\hat{D}$, and let the decoupled SBP operators with respect to the physical coordinates on $D^k$ be defined as
\begin{align}
{\bm{D}}^i_N = \sum_{j=1}^d \bm{G}_{ij}\hat{\bm{D}}^j_N.
\end{align}
%Then, $\bm{Q}^i_N = \bm{W} \bm{D}^i_N$ satisfies the SBP property on each element $D^k$
%\begin{equation}
%\bm{Q}^i_N + \LRp{\bm{Q}^i_N}^T = \bm{B}^i_N, \qquad \bm{B}^i_N = \begin{pmatrix}
%\bm{0}&\\
%& {J^k_f}\bm{W}_f \diag{{\bm{n}_i}}
%\end{pmatrix}
%\label{eq:physsbp}
%\end{equation}
%where $J^k_f$ are the Jacobian factors which result from mapping the faces of $D^k$ to reference faces.
%\end{lemma}
Then, an entropy conservative formulation can be given on $D^k$ as follows:
\begin{gather}
\pd{\bm{u}}{t} + \sum_{i=1}^d \LRs{\begin{array}{cc}
\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{\hat{\bm{n}}}\bm{f}_i^* = 0, \label{eq:skewform}\\
\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q, \nonumber\\
\bm{f}^* = \bm{f}_S(\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f), \qquad \text{ on interior interfaces,} \nonumber
\end{gather}
where $\tilde{\bm{u}}_f^+$ denotes the face value of the entropy-projected conservative variables $\tilde{\bm{u}}_f$ on the neighboring element.  The formulation can be made entropy stable by adding an appropriate penalization term, such as a Lax-Friedrichs or matrix-based dissipation \cite{winters2017uniquely, chen2017entropy, chan2017discretely}.

\subsection{Curvilinear meshes}

On affine meshes, it is possible to show entropy stability of the skew-symmetric formulation (\ref{eq:esdgSkew}) under a surface quadrature which is only exact for degree $N$ polynomials.  However, on curved meshes, a stronger surface quadrature rule is required to demonstrate entropy stability, where the strength of the surface rule depends on the order and approximation of geometric terms.

Let $\bm{J}^k\bm{G}^k_{ij}$ denote the vector of scaled geometric terms ${J}^k\bm{G}^k_{ij}$ evaluated at volume and surface quadrature points.  Decoupled SBP operators on a curved element $D^k$  can be defined as in \cite{chan2018discretely} by
\begin{equation}
\bm{D}^i_N = \sum_{j=1}^d \diag{\bm{G}^k_{ij}}\hat{\bm{D}}^j_N + \hat{\bm{D}}^j_N\diag{\bm{G}^k_{ij}}.
\label{eq:dncurved}
\end{equation}
The skew-symmetric formulation (\ref{eq:skewform}) can be extended to curvilinear meshes using the definition of $\bm{D}^i_N$ in (\ref{eq:dncurved}).  However, two additional assumptions must be satisfied in order to prove that the resulting formulation is entropy stable.

\subsection{Discrete geometric conservation law and surface quadrature accuracy}

\note{Add description of how normals factor into this and note that they're computed based on $\bm{G}_{ij}$.}

The first assumption which must be satisfied is the discrete geometric conservation law (GCL) \cite{thomas1979geometric, kopriva2006metric}.  For curved elements, Lemma~\ref{lemma:vdsbp} and Corollary~\ref{lemma:sbpcor} do not necessarily hold at the discrete level.  For example, expanding out the condition $\bm{Q}^i_N\bm{1} = \bm{0}$ in terms of (\ref{eq:dncurved}) yields
\begin{align}
\bm{D}^i_N \bm{1} = \sum_{j=1}^d \diag{\bm{G}^k_{ij}}\hat{\bm{D}}^j_N \bm{1} + \hat{\bm{D}}^j_N\diag{\bm{G}^k_{ij}}\bm{1} = \sum_{j=1}^d \hat{\bm{D}}^j_N\LRp{\bm{G}^k_{ij}} = 0,
\label{eq:dgcl}
\end{align}
where we have used that $\hat{\bm{D}}^j_N \bm{1} = 0$.  For degree $N$ isoparametric mappings, the GCL is automatically satisfied in two dimensions due to the fact that the exact geometric terms $\bm{G}^k_{ij}\in P^{N-1}$ \cite{kopriva2006metric}.  However, in three dimensions, the GCL is not automatically preserved due to the fact that $\bm{G}^k_{ij}\in P^{2N-2}$ and cannot be represented exactly using degree $N$ polynomials.  Thus, (\ref{eq:dgcl}) must be enforced through an alternative construction of $\bm{G}^k_{ij}$.

The most common approach is to rewrite the geometric terms as the curl of some quantity $\bm{r}^i$, but to interpolate $\bm{r}^i$ before applying the curl \cite{visbal2002use, kopriva2006metric, hindenlang2012explicit}:
\begin{align}
\bm{r}^i = { \pd{\bm{x}}{\hat{x}_i}\times \bm{x}}, \qquad
\LRs{\begin{array}{c}
\bm{G}^k_{i1}\\
\bm{G}^k_{i2}\\
\bm{G}^k_{i3}\end{array}} = -\frac{1}{2}\LRp{\pd{I_{R}\bm{r}^k}{\hat{x}_j}-\pd{I_{R}\bm{r}^j}{\hat{x}_k}}, 
\label{eq:iconscurl}
\end{align}
where $I_{R}$ denotes a degree ${R}$ polynomial interpolation operator with appropriate interpolation nodes.\footnote{This interpolation step must be performed using interpolation points with an appropriate number of nodes on each boundary \cite{chan2018discretely}.  These include, for example, GLL nodes on tensor product elements, as well as Warp and Blend nodes on non-tensor product elements \cite{warburton2006explicit, chan2015comparison}.}  Because the geometric terms are computed by applying the curl, both $\bm{G}^k_{ij}$ and $n_iJ^f$ are approximated using degree $(R-1)$ polynomials.

Because $\bm{D}^i_N$ are now defined through (\ref{eq:dncurved}), Corollary~\ref{lemma:sbpcor} and the proof of entropy stability may not hold for curved elements and must be modified.  The introduction of curvilinear meshes will impose slightly different conditions on the accuracy of the surface quadrature, which are summarized in the following curved version of Corollary~\ref{lemma:sbpcor}.
\begin{lemma}
Assume that the geometric terms $\bm{G}_{ij} \in V^N$ satisfy the discrete GCL (\ref{eq:dgcl}), that $\LRl{\bm{G}_{ij}}_{\partial \hat{D}}\hat{n}_j \in V^R\LRp{\partial \hat{D}}$, and that the surface quadrature is exact for polynomials of degree $N+R$.  Then, 
\[
\bm{1}^T\bm{Q}^i_N\bm{u} = \bm{1}^T\bm{B}^i_N \bm{u}, %\qquad \bm{B}^i_N = \begin{pmatrix}
%\bm{0}& \\
%& \bm{W}_f\diag{\bm{n}_i}
%\end{pmatrix}, 
\qquad \bm{Q}^i_N \bm{1} = 0, 
\]
where 
\[
\bm{B}^i_N =  \begin{pmatrix}
\bm{0}&\\
& {J^k_f}\bm{W}_f \diag{{\bm{n}_i}}
\end{pmatrix}.
\]
\label{lemma:vdsbpcurved}
\end{lemma}
\begin{proof}
The second equality $\bm{Q}^i_N \bm{1} = 0$ is a consequence of the discrete GCL (\ref{eq:dgcl}), and the proof is identical to that of \cite{chan2018discretely}.  For the first equality, if the surface quadrature is exact for polynomials of degree $2N$ on the trace space, then the stronger matrix SBP property $\bm{Q}^i_N = \bm{B}^i_N - \LRp{\bm{Q}^i_N}^T$ holds \cite{chan2018discretely}.  Combining this with $\bm{Q}^i_N\bm{1} = 0$ yields the desired result.  We focus on the proof of the variational SBP property for surface quadrature which are exact for polynomials of degree $N+R < 2N$.

Expanding $\bm{1}^T\bm{Q}^i_N\bm{u}$ yields
\[
\bm{1}^T\bm{Q}^i_N\bm{u} = \frac{1}{2}\sum_{j=1}^d \bm{1}^T \diag{\bm{G}_{ij}} \hat{\bm{Q}}^j_N \bm{u} + \bm{1}^T\hat{\bm{Q}}^j_N \diag{\bm{G}_{ij}} \bm{u}. %=  \frac{1}{2}\sum_{j=1}^d \bm{1}^T \diag{\bm{G}_{ij}} \hat{\bm{Q}}^j_N \bm{u},
\]
For the latter term in the sum, Lemma~\ref{lemma:vdsbp} holds and
\begin{equation}
\sum_{j=1}^d\bm{1}^T\hat{\bm{Q}}^j_N \diag{\bm{G}_{ij}} \bm{u} = 
\sum_{j=1}^d\bm{1}^T\LRp{\hat{\bm{B}}^j_N - \LRp{\hat{\bm{Q}}^j_N}^T} \diag{\bm{G}_{ij}} \bm{u} = \sum_{j=1}^d\bm{1}^T\hat{\bm{B}}^j_N\diag{\bm{G}_{ij}}\bm{u}, 
\label{eq:vsbp1}
\end{equation}
where we have used that $\hat{\bm{Q}}^j_N \bm{1} = \bm{0}$.  For the former term in the sum, we expand out 
\begin{gather}
\sum_{j=1}^d\bm{1}^T \diag{\bm{G}_{ij}} \hat{\bm{Q}}^j_N \bm{u} = 
\sum_{j=1}^d{\bm{G}_{ij}}^T \hat{\bm{Q}}^j_N \bm{u} \\
= \sum_{j=1}^d\begin{bmatrix}
\bm{G}^q_{ij}\\
\bm{G}^f_{ij}
\end{bmatrix}^T \begin{bmatrix} 
\hat{\bm{Q}}_i - \frac{1}{2}\LRp{\bm{V}_f \bm{P}_q}^T  \bm{W}_f{\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q &  \frac{1}{2}\LRp{\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q}^T\label{eq:ibpg}\\
-\frac{1}{2}\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q & \frac{1}{2}\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)
\end{bmatrix} \begin{bmatrix}\bm{u}_q\\
\bm{u}_f\end{bmatrix}.\nonumber
\end{gather}
where $\bm{G}^q_{ij},\bm{G}^f_{ij}$ denote the values of $\bm{G}_{ij}$ at volume and surface quadrature points.

Because $\bm{G}_{ij} \in V^N$, it is equal to its own $L^2$ projection, and the values at volume and surface quadrature points are related by $\bm{G}^f_{ij} = \bm{V}_f\bm{P}_q\bm{G}^q_{ij}$.  We can use this to simplify (\ref{eq:ibpg})
\begin{align*}
\bm{1}^T \diag{\bm{G}_{ij}} \hat{\bm{Q}}^j_N \bm{u} =&  \LRp{\bm{G}^q_{ij}}^T { \bm{Q}_i}  \bm{u}_q - \frac{1}{2} \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q+\bm{G}^f_{ij}}^T{\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q}\bm{u}_q\\
&+ \frac{1}{2} \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q+\bm{G}^f_{ij}}^T\frac{1}{2}\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{u}_f\\
=& \LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}}_i - \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q}^T{\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q}}\bm{u}_q + \LRp{\bm{G}^f_{ij}}^T\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{u}_f. %\\
%&= \LRp{\bm{G}^q_{ij}}^T \LRp{ \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q}^T{\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q}}\bm{u}_q + \LRp{\bm{G}^f_{ij}}^T\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{u}_f 
\end{align*}
The first term can be further simplified using Lemma~\ref{lemma:vsbp}
\[
\LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}}_i - \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q}^T{\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q}}\bm{u}_q = \LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}}_i}^T\bm{u}_q.
\]
The latter term can be combined with the surface term of (\ref{eq:vsbp1}) by noting that 
\[
\LRp{\bm{G}^f_{ij}}^T\bm{W}_f \diag{\hat{\bm{n}}_i}\bm{u}_f =\bm{1}^T\hat{\bm{B}}^j_N\diag{\bm{G}_{ij}}\bm{u} = \bm{1}^T\diag{\bm{G}_{ij}}\hat{\bm{B}}^j_N\bm{u},
\]
where we have used that $\hat{\bm{B}}^j_N$ is diagonal and commutes with diagonal scaling by $\bm{G}_{ij}$.
%as $\LRp{\bm{G}^f_{ij}}^T\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{u}_f = \bm{1}^T\hat{\bm{B}}^j_N \bm{u}$.  
We can use polynomial exactness and the fact that the geometric terms satisfy the continuous GCL by construction \cite{chan2018discretely} to show that $\sum_{j=1}^d  \hat{\bm{Q}}_j {\bm{G}^q_{ij}} = 0$.
Combining this with (\ref{eq:vsbp1}) and expanding out $\hat{\bm{B}}^j_N$ yields that
\begin{align*}
\bm{1}^T\bm{Q}^i_N\bm{u} &= \sum_{j=1}^d \LRp{
\frac{1}{2}\LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}_i}}^T \bm{u}_q + \bm{1}^T\begin{pmatrix}
\bm{0} &\\
& \bm{W}_f \diag{\bm{G}_{ij} \circ \hat{\bm{n}}_j}
\end{pmatrix}
\bm{u}} \\
&=  \bm{1}^T\begin{pmatrix}
\bm{0} &\\
& \bm{W}_f \diag{\sum_{j=1}^d \bm{G}_{ij} \circ \hat{\bm{n}}_j}
\end{pmatrix}
\bm{u} = \bm{1}^T\bm{B}^i_N\bm{u}.
\end{align*}
where we have used in the final step that $\sum_{j=1}^d\bm{G}_{ij}\hat{n}_j = \bm{n}_i J^k_f$ \cite{ciarlet1978finite, chan2018discretely}

\end{proof}

Ensuring that Lemma~\ref{lemma:vdsbpcurved} holds to take $R$ as large as possible, which maximizes the degree to which geometric terms are approximated.  However,

%\begin{theorem}
%Let the geometric terms be approximated using (\ref{eq:iconscurl}) with $1\leq R \leq (N+1)$.  Then, the skew-symmetric formulation is stable on non-conforming meshes using any composite surface quadrature which is exact for degree $(N+R-1)$ polynomials on $\partial \hat{D}$.
%\end{theorem}
%\begin{proof}
%
%\note{Trick here is to note effect of geometric mapping: surface quadrature must be tied to degree of geometric factors.  For Kopriva geofacs where $J\bm{G}_{ij}\in P^{N-1}$, it suffices to take surface quadrature to be exact for degree $(2N-1)$ polynomials.  For general geofacs $J\bm{G}_{ij}\in V^N$, surface quadrature which is exact for degree $2N$ polynomials must be used.  Note that this is still more flexible than the strong form: you can still use different surface quadratures on each face. }
%
%\end{proof}
Thus, if geometric terms are approximated using degree $N$ polynomials as in \cite{kopriva2006metric}, then the skew-symmetric formulation with composite GLL quadrature can be used

\subsection{Simplicial vs tensor product elements}

It was shown in \cite{chan2018discretely} that the approximation error is $O\LRp{h^{R+1}}$, and that taking $R = (N+1)$ approximates the geometric terms with an error of $O(h^{N+2})$.

\note{Need to note image of curl is in different spaces for simplicial vs tensor product elements.  Influences proofs that geometric terms satisfy appropriate assumptions in those cases.}
%\subsection{On implementation and computational cost}

\section{Numerical experiments}

In this section, we present two-dimensional experiments on hybrid meshes consisting of quadrilaterals and triangles, as well as on non-conforming meshes of quadrilateral elements.

\subsection{Reduced surface quadrature}

\note{Affine and curved triangles with GLL surface quadrature (under-integrated).}

\subsection{Hybrid meshes}

\subsection{Non-conforming meshes}

\appendix

\section{An explicit skew-symmetric entropy stable formulation for DG-SEM}

\note{$\bm{V}_q,\bm{P}_q = \bm{I}$, while $\bm{V}_f$ reduces to a permutation matrix.  }

\bibliographystyle{unsrt}
\bibliography{dg}

\end{document}

### 2018-06-07 talks/ECCM2018/iga/revision/reviewerResponse.tex
updating ECCM talks

%&pdflatex

%\documentclass[preprint,10pt]{elsarticle}
\documentclass[10pt]{article}
\usepackage{fullpage}
\usepackage{hyperref}
\let\proof\relax
\let\endproof\relax
\usepackage{amsmath,amssymb,amsfonts,mathrsfs,amsthm}
\usepackage[titletoc,toc,title]{appendix}

%\usepackage{lineno}
\usepackage{array}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{mathtools}
\usepackage{pdfpages}
\usepackage[textsize=footnotesize,color=green]{todonotes}
\usepackage{bm}
%\usepackage{tikz}
\usepackage[normalem]{ulem}
\usepackage{hhline}
\usepackage{csquotes}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{algorithmicx}

\usepackage{graphicx}
%\usepackage{subfigure}
\usepackage{subfig}

\usepackage{color}
\definecolor{forestgreen}{rgb}{0, 0.5, 0}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

\newcommand{\vect}[1]{\ensuremath\boldsymbol{#1}}
\newcommand{\tensor}[1]{\underline{\bm{#1}}}
\newcommand{\del}{\triangle}
\newcommand{\curl}{\grad \times}
\renewcommand{\div}{\grad \cdot}
\newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}#2}}
\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
\newcommand{\pdd}[2]{\frac{\partial^2#1}{\partial#2^2}}

\newcommand{\bs}[1]{\boldsymbol{#1}}

\newcommand{\equaldef}{\stackrel{\mathrm{def}}{=}}

\newcommand{\tablab}[1]{\label{tab:#1}}
\newcommand{\tabref}[1]{Table~\ref{tab:#1}}

\newcommand{\theolab}[1]{\label{theo:#1}}
\newcommand{\theoref}[1]{\ref{theo:#1}}
\newcommand{\eqnlab}[1]{\label{eq:#1}}
\newcommand{\eqnref}[1]{\eqref{eq:#1}}
\newcommand{\seclab}[1]{\label{sec:#1}}
\newcommand{\secref}[1]{\ref{sec:#1}}
\newcommand{\lemlab}[1]{\label{lem:#1}}
\newcommand{\lemref}[1]{\ref{lem:#1}}

\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\snor}[1]{\left| #1 \right|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRb}[1]{\left| #1 \right|}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}

\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\Div} {\ensuremath{\nabla\cdot}}
\newcommand{\Nel} {\ensuremath{{N^\text{el}}}}
\newcommand{\jump}[1] {\ensuremath{\LRs{\![#1]\!}}}
\newcommand{\avg}[1] {\ensuremath{\LRc{\!\{#1\}\!}}}
\newcommand{\uh}{\widehat{u}}
\newcommand{\Bh}{\widehat{B}}
\newcommand{\fnh}{\widehat{f}_n}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\pO}{\partial\Omega}
\newcommand{\Gh}{\Gamma_h}
\newcommand{\Gm}{\Gamma_{-}}
\newcommand{\Gp}{\Gamma_{+}}
\newcommand{\Go}{\Gamma_0}
\newcommand{\Oh}{\Omega_h}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

\newenvironment{definition}[1][Definition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newenvironment{example}[1][Example]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\newcommand{\eval}[2][\right]{\relax
  \ifx#1\right\relax \left.\fi#2#1\rvert}

\def\etal{{\it et al.~}}

\def\arr#1#2#3#4{\left[
\begin{array}{cc}
#1 & #2\\
#3 & #4\\
\end{array}
\right]}
\def\vectwo#1#2{\left[
\begin{array}{c}
#1\\
#2\\
\end{array}
\right]}
\def\vecthree#1#2#3{\left[
\begin{array}{c}
#1\\
#2\\
#3\\
\end{array}
\right]}
\def\vectfour#1#2#3#4{\left[
\begin{array}{c}
#1\\
#2\\
#3\\
#4\\
\end{array}
\right]}

\newcommand{\G} {\Gamma}
\newcommand{\Gin} {\Gamma_{in}}
\newcommand{\Gout} {\Gamma_{out}}

\newcommand{\note}[1]{{\color{violet}#1}}
\newcommand{\reviewerOne}[1]{{\color{blue}#1}}
\newcommand{\reviewerTwo}[1]{{\color{red}#1}}
%\newcommand{\remark}[1]{\textbf{\color{red}#1}}

\newcommand{\tri}{{\rm tri}}
\newcommand{\sqr}{{\rm quad}}

\newcommand{\refhex}{\widehat{\mathcal{H}}}
\newcommand{\reftet}{\widehat{\mathcal{T}}}
\newcommand{\refwedg}{\widehat{\mathcal{W}}}

\newcommand{\refpyr}{\widehat{\mathcal{P}}}
\newcommand{\refpyrf}{\widehat{\mathcal{P}}^f}
\newcommand{\refpyrfq}{\widehat{\mathcal{P}}^{\sqr}}
\newcommand{\refpyrft}{\widehat{\mathcal{P}}^{\tri}}

\newcommand{\hex}{{\mathcal{H}}}
\newcommand{\tet}{{\mathcal{T}}}
\newcommand{\wedg}{{\mathcal{W}}}

\newcommand{\pyr}{{\mathcal{P}}}
\newcommand{\pyrf}{{\mathcal{P}}^f}
\newcommand{\pyrfq}{{\mathcal{P}}^{\sqr}}
\newcommand{\pyrft}{{\mathcal{P}}^{\tri}}

\newcommand{\half}{1/2}

\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

%% d in integrand
\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}}

\makeatletter
\renewcommand\d[1]{\mspace{6mu}\mathrm{d}#1\@ifnextchar\d{\mspace{-3mu}}{}}
\makeatother

\newcommand{\squareface}{[-1,1]^2}

%\journal{Journal of Computational Physics (JCP)}

\date{}
\author{Jesse Chan, John A.\ Evans}
\title{Response to reviewers: ``Multi-patch discontinuous Galerkin spline finite element methods for time-domain wave propagation''}

\graphicspath{{./figs/}}

\begin{document}

\maketitle

\note{We thank the two anonymous reviewers for their feedback.  We describe steps taken to address reviewer comments and suggestions, which are described in the following response.  Revisions in the manuscript are also colored \reviewerOne{for Reviewer 1} and \reviewerTwo{for Reviewer 2} for ease of identification.  We have modified the title to ``Multi-patch discontinuous Galerkin isogeometric analysis for wave propagation: explicit time-stepping and efficient mass matrix inversion'' in order to provide more description of the material in the paper.  We have also made minor changes in nomenclature, referring to micro-elements simply as ``elements'' and referring to subdomains $D^k$ as ``patches'' and correcting an inconsistent definition of the trace and inverse inequality constants.  A figure has been added to further clarify these notational changes.  These changes are also highlighted in {\color{forestgreen} green} in the revised manuscript.   We hope these revisions improve the readability of this paper and its suitability for the audience of CMAME.}

\section{Reviewer 1}

The paper describes a new approach to handling the patch dependent mass matrix in for spline based finite element methods in order to improve computational efficiency. The author also introduce a novel approach to determining improved knot locations for B-splines. Finally the authors explore the time step restriction for B-spline methods.

I believe the work is worthy of publication once the comments below (particularly those concerning the time step restriction) are resolved.

\begin{enumerate}
\item One thing that I got a bit turned around in the paper is the meaning of $h$ throughout the text. *I think* that in some places, $h$ is the 'micro-element' size and in other places it is the 'patch size'; sometimes the patch size is denoted $H$ as well. This was particularly confusing when talking about time step restrictions, since on page 13 $h$ means patch size and the micro-element size shows up $C_{I}$ and $C_{T}$. Similarly, the $h$ which shows up in the introduction when talking about time step size (page 2) is the patch size when talking about the finite element methods (since patch size is element size in the this case) but micro-element size when talking about spline methods (*I think*).

\note{We have modified the revision in a number of ways to address this comment.  First of all, we have removed all mention of ``micro-element'' and instead refer only to ``patches'' and ``elements'' (which previously were referred to as micro-elements).  This is consistent with the standard terminology in the isogeometric analysis community.  Second, we have modified the revision to ensure that $H$ always refers to the patch size and $h$ refers to the element size.  Third, we have included a figure to clearly demonstrate what $H$ and $h$ are in a visual manner.  Fourth, we have included a remark noting that $H = h$ for a $C^0$-continuous or discontinuous finite element method, and hence the terms patch and element are interchangeable for this setting.  We hope these modification clarify the manuscript.  }

\item I'm not completely convinced that the time step for the B-spline methods is that much better than for polynomials. If we use a B-spline approximation that if $p$-th order accurate with $K$ micro-elements, this results in roughly $K$ degrees of freedom. Assume that the domain is of length $H$, this results in a time step that if proportional to $H / K p$. Now, presumably to get a finite element solution of roughly the same order of accuracy, I need roughly the same number of degrees of freedom. To do this, we would need our finite element mesh to have $E = K / p$ elements (so that the total degrees of freedom is roughly $E p = K$).  Each of these element is of size $H / E = p H / K$, and thus the time step restriction if $(H / E) / p^2 = H / K p$ which is the same as for the B-spline method. This back of the envelop calculation suggests that they have roughly the same time step restriction, but does not consider the scaling constant that comes into both of these. I think that by dividing by $K$ in the Figures 7 and 8 there is a factor of $p$ that is being lost since $K$ scales with $p$. (For splines $H / K$ is similar to $h / p$ polynomials, since both are approximately an average node spacing.)

\note{We thank the reviewer for their comments.  We believe the discrepancy lies is in the statement ``presumably to get a finite element solution of roughly the same order of accuracy, I need roughly the same number of degrees of freedom''.  Numerical experiments in the spline literature show that instead, for a fixed order $p$ and similar mesh size $h$, splines are more efficient per degree of freedom than $C^0$ finite elements \cite{evans2009n} (with the caveat that those degrees of freedom are more tightly coupled).  
\\
\\
Thus, assuming a physical patch (element) of size $O(1)$, a degree $p$ polynomial method would have a timestep restriction of $O(1/p^2)$.  A spline method would have a timestep restriction of $O(1/(pK))$, and with $K = O(p)$, this results in the reviewer's observation that a spline method would have an $O(1/p^2)$ timestep restriction as well.  However, what this does not account for is the fact that the spline also increases the mesh resolution of by a factor of $O(1/K) = O(1/p)$.  Thus, to achieve the same resolution as a spline space of degree $p$ and $K$ elements, a polynomial finite element method would require a mesh of size $O(1/p)$, which would decrease the timestep restriction to $O(h/p^2) = O(1/p^3)$.  Thus, splines allow for a factor of $O(p)$ increase in the time-step over a $C^0$ finite element method of equivalent $h$ and $p$ resolution.  
\\
\\
We have added a section explaining this in the revised manuscript.
}

\item In section 3 (definition of B-spline basis) it could be helpful to readers who have minimal exposure to B-splines to add a few more broad details. In particular, I think that it would be useful to write down the final approximation once the degrees of freedom have been set and point out/remind the reader B-splines is not an interpolant of the function values at the knots (i.e., the approximation does not pass through the degrees of freedom and hence there is no problem with having multiple values at the same physical point).

\note{We have added a brief discussion of B-spline coefficients to the revised manuscript, and have noted that they should not be treated as interpolatory basis functions.}

\item I was a little confused by the iterative smoothing procedure outlined on page 8 for the knot locations. In particular, when (2) is used what exactly are the $\hat{n}_{i}$ values. At first I thought that $\hat{x}_{i} = \frac{2}{K}(i-1)-1$ (i.e., equally spaced between $[-1,1]$), but then footnote 3 seems to contradict this, since the $\tau_{i}$ cannot be equally spaced.

\note{We apologize; there is an error in (2), and the $\hat{x}_i$ should be $\hat{x}_j$.  We have corrected this in the revised manuscript.
\\
\\
The $\hat{x}_i$ points are in fact equally spaced points on $[-1,1]$.  We have noted in the revision that (2) can be interpreted as an attempt to reverse-engineer a knot vector which produces equispaced Greville abscissae. The goal is to produce new knot locations $\tilde{\xi}_i$ for which the Greville abscissae are less clustered towards the endpoints.  
}

\item Specific Comments:
\begin{enumerate}
\item [page 2, lines 20-29] As hinted to above, I think that you should spell out what $h$ is in each of the cases in this paragraph (since I think that the mean different things).

\note{We agree, and have added descriptions of both patch size $H$ and mesh size $h$ to this paragraph.}

\item [page 3, line 6] Not necessary, but you may want to mention what $\phi$ is.

\note{We have specified that it represents some concentration.}

\item [page 3, line 14] You may want to mention this assumes that $c$ is a constant wave speed as well as define $p$ and $\vec{u}$ (and state the dimension of $\vec{u}$)

\note{We have added a description of $p$ and $\bm{u}$.  However, to the authors' understanding, $c$ does not necessarily need to be constant for this formulation to be applicable (see, for example \cite{chan2017weight1,mercerat2015nodal}). }

\item [page 3, line 26] This equations should still include $c$ I believe, and you may want to state that the $f$ here and the $f$ above are different.

\note{We thank the reviewer for their careful reading; we have corrected this in the revision.}

\item [page 3, line 33] Is it true that NURBS is always used, or is it just often used? (I don't know for certain)

\note{We thank the reviewer for their question.  For spline discretizations, NURBS are the most commonly used type of geometric mapping.  Additionally, NURBS mappings also include isoparametric mappings.  We have removed the NURBS reference in the revision, since the method can be used for more general geometric mappings as well.}

\item [page 4, line 42] Might be worth stating that energy stable in this case is that $d/dt\sum_{k}(v, v) \le 0$.

\note{We have added this to the revision.}

\item [page 7, line 45] It is hard to see in the figure that the points are clustering. Not sure how this could be improved, perhaps plotting the optimal points all on the same graph?

\note{We note that ``clustering'' may not be an appropriate description of the point behavior, and have modified this in the revision to read: ``the gap between each of the boundary knots and the closest interior knot grows'' as $r$ increases.}

\item [Page 8, line 9] should $j$ run from $1$ to $K$?

\note{We believe $j$ should run to $p+K$, which is the total number of degree $p$ B-spline basis functions for a knot vector containing $K$ elements.}

\item [page 12, line 48] reference 11 describes the Galerkin difference methods in 1-D, and does not consider mapped domains. Perhaps the reference should be moved a few words earlier?

\note{We have removed this reference, as the cited work references a paper in progress.}

\item [page 13, lines 23, 28, 40] Should there be absolute values around the eigenvalues ($A_{h}$ have non-positive real components) for stability.  [page 13, line 34] Perhaps the sentence on line 44 "For illustation..." should be moved up to before the advection example begins (line 34)?

\note{We thank the reviewer for their careful reading; this has been corrected in the revision.}

\item [page 13, line 37] Should $a_{\mbox{sym}}$ be included too?

\note{We specify $a_{\mbox{sym}}$ in terms of $a(u,v)$ in the previous few lines; we have only given an explicit expression for $a_{\mbox{skew}}$ since the complex conjugate is introduced (which was not present in the original form $a(u,v)$).}

\item [page 13, line 56] Perhaps $H$ should be used instead of $h$ so as not to be confused with previously defined $h = H / K$?  [page 14, line 13] Should the $\beta$ here be a $c^2$?  [page 14, line 13] Perhaps $C_{p}$ should be defined earlier, and used in the other expressions?

\note{Yes, and we thank the reviewer for pointing these issues out.  We have also modified the revision so that $C_p$ is introduced earlier, and have mentioned that the trace and inverse inequality constants depend on $h$.}

\item [page 14, line 13] Strictly speaking $\Delta t$ should also have a constant multiplying due to the stability region of the time stepping method.

\note{We have added a constant describing the impact of the stability region of the time-stepping method.}

\item [page 15, line 35] You may want to point out that the numerical stability issues are in the calculation of the points not the PDE approximation using the points.

\note{We thank the reviewer for their suggestion, and have clarified this in the revision.}

\item [page 15, line 53] I had to read the sentence "In contrast, not only..." several times due to all the negations. Perhaps this sentence could be reworeded.

\note{We have broken this sentence into two: ``In contrast, knot smoothing does not significantly decrease accuracy.  Additionally, the approximation power of spline spaces under knot smoothing is actually improved for certain functions''.  We hope this improves readability.}

\item [page 16, line 26, 35] In the table captions I think that a scaling by $K$ is missing.

\note{We thank the reviewer for their careful reading; this has been corrected in the revision.}

\item [page 17, line 47] Should you define how you are computing the approximation? I assume that you mean that you are doing an $L^2$ projection of the function to the spline representation (but this is never stated). You may also want to state how you are computing the $L^2$ error (presumable integration using high order quadrature?)

\note{We agree with the reviewer, and have clarified how the approximation is computed, including a description of the composite Gaussian quadrature rule used.}

\item [page 19, line 19] You may want to consider adding labels to the x-axis.

\note{We have added $x$-axis labels to the revision.}

\item [page 19, figs 10 and 11] Would it be useful showing both Figures 10 and 11 with the same $p$ and $K$ and using the same x-axis (one use $k$ and the other $k / dof$) so that the effects of the approximation can be compared between the two?

\note{We have modified both figures to have the same axes.  Additionally, in order to facilitate a comparison between the two, we have switched to approximating the same function in both Figures 10 and 11.  }

%\item [page 19, line 52] As above you may want to explicit about how you are approximating the function in the approximation space.

\item [page 19, line 53] Should the word "elements", really be "micro-elements"?  (same in caption of Fig. 13). You may also want to make clear that you are using a single polynomial "element" not a mesh of elements (at least this is what I gather you are doing).

\note{As previously mentioned, we have replaced the term ``micro-element'' with ``element'' throughout the paper, so this comment is no longer applicable.}

\item [page 20, line 57] Might be better to say "are closer than in the one-dimensional case" (I found the wording confusing since you are not comparing the one-dimensional and two-dimensional cases, but the two knot distributions in each case).

\note{We agree, and have corrected this issue in the revision.}

\item [page 20, line 58] I had to reread the sentence ``This suggests...'' several times to understand what was being said. My understanding of the logic is that since the errors are same AND the timestep for the smooth knots is larger, a larger timestep can be taken to get the same spatial error. And that this is not true in the one-dimensional case because you would need to increase the number of knots for the smoothed distribution to get the same error and thus needed a smaller timestep.

\note{We apologize for the confusion, and have restructured this sentence in the revision: ``This suggests that the use of smoothed knots instead of uniform knots can improve solver efficiency, since it is possible to take a larger maximum stable timestep without reducing accuracy.  A more rigorous study of this claim is still required.''  Additionally, we have clarified that the improvement in efficiency is not due to the dimension of the problem, but the presence of a curvilinear warping (which does not naturally appear in 1D problems).}

\item [page 21, fig 14] The placement of the $10^4$ in the figures looks like its in the axis label, but I think that it is really scaling the axis.  [page 21, line 40] You may want to put the $\alpha$ factor in the caption so it is more self contained.

\note{We thank the reviewer for their suggestions.  We have addressed both of these points in the revision by switching to a log-log axis and modifying the caption to include $\alpha$.}

\item [page 22, sec 5.4.1] It should be pointed out somewhere in this section that you are considering a 1-D problem. Also, is it worth showing the spectrum in any of these cases? Wondering since RK stability regions are not circles, but a peanut-shaped region so spectral radius is not the full story for a stable time step.

\note{We agree, and have added plots of the spectra for the advection equation.  These plots further and illustrate how the spectra depends on both the penalty parameter $\tau$ and the knot distribution.}
%\textcolor{red}{TODO}

\item [page 23, line 48] Does this suggest that upwind fluxes should be used instead of the penalty flux? Is similar behavior seen for the wave equation?

\note{For the constant coefficient advection equation, the penalty and upwind fluxes coincide, so we cannot draw a conclusion on the use of penalty or upwind fluxes from this specific experiment.  Additionally, it is unclear whether this phenomena holds for more complex scenarios where the upwind and penalty fluxes differ (for example, curvilinear meshes or discontinuous wave speeds).  
\\
\\
Similar behavior is observed for the wave equation, and we have mentioned this in the revised manuscript.}

\item [page 24, Fig 18c,d] What is the black line in these figures? What is $h^{11}$ and what does it represent?

\note{We have clarified in the caption that the black line represents the exact dispersion relation in Figure 18c, while in Figure 19c, the black dotted line is the slope $h^{11}$, which corresponds to the theoretical rate of convergence of the dispersion error for $p=4$.}

\item [page 24, line 43] From the figure, it seems that changing $K$ does more than just shifting things by a constant.

\note{We have clarified this in the revision to specify that $h$-refinement shifts the dispersion error curve by a constant number of wavelengths per degree of freedom. }

\item [page 24, line 50] How does $\tau$ effect the dispersion error results?

\note{As $\tau\rightarrow 0$, we observe that the dispersion error increases for a larger number of wavelengths per degree of freedom.  We have included plots for reference here.  In the revision, we have added a discussion of this in a footnote.}
\begin{figure}
\centering
\includegraphics[width=.4\textwidth]{tau0dispersionRelation.png}
\includegraphics[width=.4\textwidth]{tau0dispersionError.png}
\caption{Dispersion relation and dispersion errors for $p=4$ and $\tau = 0$.  }
\end{figure}

\item [page 24, sec 5.5] Are the eigenvalues just sorted to determine which discrete eigenvalues correspond to the exact eigenvalues? Would this explain the ``blip'' in the error for the uniform mesh? Also, how are the eigenvectors normalized in computing the error?

\note{Yes, the eigenvalues are sorted by magnitude.  The eigenvectors are sorted by the magnitude of the corresponding eigenvalue, and the normalization constant for a discrete eigenvector is computed by projecting the exact eigenvector onto the span of that single discrete eigenvector.  
\\
\\
The ``blip'' in the error corresponds to a very fortuitous aliasing error for $p=4$, where the discrete eigenvector on the uniform mesh matches the exact eigenvector almost exactly due to the matching of the uniformly spaced knot locations with zeros of the exact eigenvector.  It does not appear for smoothed knots since the knot locations are no longer uniformly spaced. }

\end{enumerate}
\end{enumerate}

\section{Reviewer 2}

This work presents efficient explicit time integration techniques in the context of isogeometric analysis applied to wave propagation problems. The method couples B-spline patches weakly by means of standard Discontinuous Galerkin techniques. The consistent mass matrix is block-diagonal and can be inverted patch by patch separately. The main contribution of the paper is an efficient strategy to invert the mass matrix in the context of curved geometries. The approach utilizes the tensor structure of the B-spline spaces and maintains higher order accuracy, which is critical. A secondary contribution is a study, based on the theory of n-widths, in which the spline knot vector is optimized to obtain optimality in the $L^2$-norm. It is shown that the spline spaces based on such a knot vector have slightly improved approximation properties of the medium to high modes in the spectrum, although the approximation of the low modes is slightly decreased. The improved approximation
properties of the high modes additionally lead to improved critical time step size. It is furthermore shown that the weak treatment of patch continuity can have a positive effect on the critical time step.

The paper is well written and is tackling an important topic in finite element analysis. I approve the manuscript for publication if some of the following matters are taken into consideration.

\begin{enumerate}
\item The main contribution, which is the efficient inversion of the mass matrix (approximately) is undersold, however. It would be good to show more numerical evidence, such as in Figure 12b, that shows that high order accuracy is maintained.

\note{We have added another 2D experiment comparing approximation errors, and have also added two numerical experiments comparing errors for the 2D wave equation using both the inversion of the curvilinear mass matrix and the inversion of the weight-adjusted approximation.}

\item A lot of attention is spend on the improvement of the knot vector while little gain is achieved w.r.t accuracy or robustness. Perhaps the greatest efficiency here is w.r.t the gain in critical time step. The authors study the spectral radius and the constants in trace inequalities. Although this provides some evidence; why not directly estimate the critical time step by solving an eigenvalue problem and check if is tight. Then one can show using non-dimensional plots how the the critical time-step behaves.

\note{We agree that the greatest efficiency gain here is likely with respect to the gain in critical time step.  For the advection equation and first-order wave equation, the maximum stable timestep size is directly tied to the spectral radius, with a constant of proportionality associated with the chosen explicit timestepper.  For the second-order wave equation, the maximum timestep size may indeed be attained from an eigenproblem, but we note that it may be equivalently attained from the square root of the spectral radius.  In this direction, we have added a paragraph to the paper noting the relationship between the maximum stable timestep and the spectral radius for the second-order wave equation.  We have further added plots of the square root of the spectral radius for the second-order wave equation to explicitly study the impact of polynomial degree $p$ on the maximum stable timestep size.}

\item Please introduce some equation numbering of the most important equations. Please proofread all equations again. I found several typo's but there may be more. Here is a list with more specific comments, fixes, etc.

\begin{enumerate}
\item Combined comments: [page 2, line 5]  geo-metrically i.s.o  ge-ometrically.  [page 2, line 12] requires the inversion of a 'sparse' matrix.  [page 3, line 25 and 26] The equation is different. There is a time derivative on the righthand side and a factor $(1/c^2)$ in front of the time term.  [page 3, line 50]                  denote the neighboring element 'sharing' a face.  [page 3, line 51] The jump 'and average' of .  [page 3, line 58] remove 'rule'

\note{These points have been corrected in the revision.}

\item page 3, line 45 -                [16,18]

\note{References [16, 18] refer to the inversion of a weighted mass matrix.  This line refers instead to constants in trace and inverse inequalities, which are first described in [14].  We have also added two additional references on explicit expressions for constants in trace and inverse inequalities to further clarify this.  }

\item page 4 and 5 -                  section 2.1 and 2.3 - complete the formulations. This is sloppy. 'Find ??.. such that for all??'

\note{The three formulations have been expanded as recommended in the revision.}

\item page 5                               define $x$ and $\hat{x}$

\note{The coordinates $x$ and $\hat{x}$ have been defined on page 3, line 35 of the original submission.}

\item page 5, line 6 -                  'pulled back' i.s.o 'mapped'

\note{We have added ``pulled back'' to clarify this sentence. }

\item page 5, line 14 -                where $G_{ij}$ is the inverse of the Jacobian matrix. 'a matrix of geometric change of variables' ?

\note{We have added the description of $G_{ij}$ as the inverse of the Jacobian matrix of the mapping, and have added two references showing that this matrix is also referred to as ``metric terms'' or ``geometric factors'' in the literature.}

\item page 6, line 47 -               [3, 12, 24] add the reference of the spline book by schumaker.

\note{We have added the reference to ``Spline functions: basic theory'' by Larry Schumaker.}

\item page 11, line 9-10             Remove the line "The extension of this work to NURBS approximations ??.". What you say in this sentence                                            is false. The mass matrix is not tensorizable in the case of NURBS. Obviously you can extend the approach given here, but in that case show how that is done.

\note{We have added a section describing how to generalize the weight-adjusted mass matrix inverse to rational NURBS basis functions.  }

\item page 11, line 40                Poor notation of this equation. Firstly, you have not defined multi-dimensional B-spline basis functions. Secondly, the pull back in combination with the same representation of the basis functions in physical and reference domain is poor notation.

\note{Multi-dimensional B-spline basis functions are defined in Section 3.3 of the original submitted manuscript.  We have moved this section such that it is now right after the section defining univariate B-spline basis functions.  We have also corrected the notation by introducing the inverse of the reference-to-physical geometric mapping.  }

\item page 12, line 33                Typo in equation. I think you mean ``$M_{1/J} = \ldots$'' i.s.o. ``$M^{-1}_{1/J} = \ldots$''.

\note{We thank the reviewer for their careful reading; this has been corrected in the revised manuscript.}

\item Combined comments: [page 12, line 40]                "The application of $M_{1/J}$ can be performed efficiently in a matrix free fashion". Just precompute it once and  use it repeatedly in the time integration algorithm.  [page 12, line 37]           You can write $\hat{M}$ using Cholesky as $L^T L$ and write $M^{-1} \approx L^{-T} W L^{-1}$ where $W = (L^{-1}  M_{1/J} L^{-T})$.   You can precompute $W$.

\note{We have added a note that one can also precompute and store this matrix.  However, $M_{1/J}$ does not admit a Kronecker product structure due to the $1/J$ weighting.  
%However, we note that this can increase storage costs and require more data movement, which can lead to slower runtimes on many-core architectures such as GPUs .  Additionally, the cost of the matrix-free application of $M_{1/J}$ can be combined with existing steps in explicit solvers to further reduce costs \cite{chan2016weight1}. 
This is not as problematic if only $M_{1/J}$ is stored, since it is sparse.  However, computing $W = (L^{-1}  M_{1/J} L^{-T})$ produces a three-dimensional dense matrix, and requires the same storage and computational work as explicitly pre-computing and storing the original weighted mass matrix. }

\item Combined comments: [page 12, line 57]     For reduced quadrature reference [Hiemstra, 2017, Schillinger, 2014].  [page 13, line 5]                 [37] is not a quadrature free method.  It proposes a different way of performing quadrature, which scales much better with regard to polynomial degree $p$.

\note{We thank the reviewer for their comments, and have fixed both of these points in the revised manuscript.}

\item page 13, line 15                This sentence is not true. Although a Cholesky solve is not scalable, you have to perform $3n^2$ ($n$ is dim of 1D space) of them per patch since you are using the univariate operators. This is very scalable!

\note{We thank the reviewer for pointing this out.  This is true; however, parallelizing over each of the $3n^2$ Cholesky solves also requires significantly more memory (due to the fact that the Cholesky backsolve requires storage of previous).  For accelerator and many-core architectures such as GPUs, this can lead to slower runtimes \cite{chan2017weight}.  We have added a footnote to the revised manuscript mentioning this.  }

\item page 13, line 31                Typo. $A_h - A_h^T$ i.s.o. $A_h - A_h$. Same typo as in line 32.

\note{We have corrected this in the revision.}

\item page 20, Figure 13           What does polynomial refer to here? Also, why not use a log-log scales for these plots (on x-axis use log of univariate number of dofs)

\note{We have specified that ``polynomial'' refers to a high order polynomial approximation.  We have also changed the semi-log scales to log-log scales in the revision.}

\item page 20, Figure 12b         This is a good Figure. I would show this for all the computations. For the 3D problem you could look at the relative difference in solution (in the $L^2$ norm) with respect to the consistent collocation mass.

\note{We have added a few more experiments comparing approximation errors, and have added one plot comparing multi-patch DG errors using both the full weighted mass matrix inverse and the weight-adjusted approximate mass matrix inverse.  The high computational cost of computing fully dense inverses of 3D curvilinear mass matrices makes a detailed comparison in 3D infeasible.  Additionally, the comparison to collocation spline methods is outside of the scope of this work, and will be investigated in future work. }

\item Combined comments: [page 21, Figure 14]           Use a log-log scales for these plots (on $x$-axis use log of univariate number of dofs), [page 26, line 35]     integrate spline spaces to sufficient accuracy.

\note{These have been corrected in the revision.}

\end{enumerate}
\end{enumerate}

\bibliographystyle{unsrt}
\bibliography{dgpenalty}

\end{document}

### 2018-06-07 talks/ECCM2018/abstract.tex
updating ECCM talks

\documentclass[a4paper]{article} %
\usepackage{graphicx,amssymb} %

\textwidth=15cm \hoffset=-1.2cm %
\textheight=25cm \voffset=-2cm %

\pagestyle{empty} %

\date{} %

\def\keywords#1{\begin{center}{\bf Keywords}\\{#1}\end{center}} %

% Please, do not change any of the above lines

\begin{document}

% Type down your paper title
\title{Entropy stable discontinuous Galerkin methods using arbitrary bases and quadratures}

% Authors
\author{Jesse Chan, \\ %
       Rice University (USA) \\ \\ % Affiliation 1
       \tt{jesse.chan@rice.edu} % Only one corresponding e-mail
       }%

\maketitle

\thispagestyle{empty}

% The abstract

\begin{abstract}
High order discontinuous Galerkin (DG) methods offer several advantages in the approximation of solutions of nonlinear conservation laws, such as improved accuracy and low numerical dispersion/dissipation.  However, these methods also tend to suffer from instability in practice, requiring filtering, limiting, or artificial dissipation to prevent solution blow up.  Provably stable DG methods based on summation-by-parts (SBP) operators and flux differencing address this inherent instability by ensuring that the solution satisfies a semi-discrete entropy inequality.

Under specific combinations of basis functions or quadratures, nodal DG methods can be interpreted within an SBP framework.  In this talk, we discuss how to construct discretely entropy stable discontinuous Galerkin methods for general choices of basis and quadrature using discrete $L^2$ projection operators and ``decoupled'' SBP operators.  Extensions to curvilinear meshes and implementational aspects on Graphics Processing Units will be also discussed.  
\end{abstract}

\keywords{High order, discontinuous Galerkin, entropy stability} % Write down at least 3 Keywords

% \section{Introduction}

\end{document}

### 2018-06-07 talks/ECCM2018/abstract2.tex
updating ECCM talks

\documentclass[a4paper]{article} %
\usepackage{graphicx,amssymb} %

\textwidth=15cm \hoffset=-1.2cm %
\textheight=25cm \voffset=-2cm %

\pagestyle{empty} %

\date{} %

\def\keywords#1{\begin{center}{\bf Keywords}\\{#1}\end{center}} %

% Please, do not change any of the above lines

\begin{document}

% Type down your paper title
\title{Efficient explicit solvers for multipatch discontinuous Galerkin isogeometric analysis}

% Authors
\author{Jesse Chan, \\ %
       Rice University (USA) \\ \\ % Affiliation 1
       John A. Evans, \\ %
       UC Boulder (USA) \\ \\ % Affiliation 1
       \tt{jesse.chan@rice.edu} % Only one corresponding e-mail
       }%

\maketitle

\thispagestyle{empty}

% The abstract

\begin{abstract}
In this talk, we present explicit solvers for time-dependent hyperbolic problems on curved geometries using isogeometric analysis and spline-based finite element methods.  Explicit finite element solvers require mass matrix inversions at each time-step.  For affine geometries in two and three dimensions, isogeometric mass matrices possess a Kronecker product structure, which can be exploited to reduce the application of the inverse mass matrix to the application of one-dimensional inverse mass matrices.  However, this Kronecker structure is lost for curved domains due to the introduction of a spatially varying weight into the isogeometric mass matrix.  The inversion of a large dense mass matrix can be avoided by approximating the inverse mass matrix using techniques such as mass lumping; however, it can be difficult to maintain both energy stability and high order accuracy under such approaches.

We present a simple ``weight-adjusted'' approximation for the inverse of isogeometric mass matrices on a single curved patch.  This approximation can be applied in a Kronecker-like manner using one-dimensional spline matrices, while maintaining provable energy stability and high order accuracy.  We extend this approach to multiple geometric patches using a multi-patch discontinuous Galerkin isogeometric analysis (DG-IGA) approach.  We also describe advantages offered by isogeometric methods over $C^0$ and discontinuous high order finite element discretizations when paired with explicit time integration on curvilinear geometries.  Numerical results in one, two, and three dimensions confirm theoretical results and illustrate the advantages of the proposed methods.

%We present a class of spline finite element methods for time-domain wave propagation which are particularly amenable to explicit time-stepping. The proposed methods utilize a discontinuous Galerkin discretization to enforce continuity of the solution field across geometric patches in a multi-patch setting, which yields a mass matrix with convenient block diagonal structure. Over each patch, we show how to accurately and efficiently invert mass matrices in the presence of curved geometries by using a weight-adjusted approximation of the mass matrix inverse. This approximation restores a tensor product structure while retaining provable high order accuracy and semi-discrete energy stability. We also estimate the maximum stable timestep for spline-based finite elements and show that the use of spline spaces result in less stringent CFL restrictions than equivalent piecewise continuous or discontinuous finite element spaces. Finally, we explore the use of optimal knot vectors based on $L^2$ $n$-widths. We show how the use of optimal knot vectors can improve both approximation properties and the maximum stable timestep, and present a simple heuristic method for approximating optimal knot positions. Numerical experiments confirm the accuracy and stability of the proposed methods.

\end{abstract}

\keywords{Isogeometric analysis, high order, mass matrix, explicit time-stepping} % Write down at least 3 Keywords

\begin{thebibliography}{1}
\bibitem{key} Chan, Jesse, and John A. Evans. \emph{Multi-patch discontinuous Galerkin spline finite element methods for time-domain wave propagation.} arXiv preprint arXiv:1708.02972 (2017).  Submitted to CMAME, in revision.
\bibitem{mpiga} Langer, Ulrich, Angelos Mantzaflaris, Stephen E. Moore, and Ioannis Toulopoulos. \emph{Multipatch discontinuous Galerkin isogeometric analysis.} In Isogeometric Analysis and Applications 2014, pp. 1-32. Springer, Cham, 2015.  	
\end{thebibliography}

% \section{Introduction}

\end{document}

### 2018-06-07 talks/ECCM2018/eccm-ecfd_Abract.tex
updating ECCM talks

\documentclass[12pt]{eccm-ecfd_abstract}

%\usepackage{graphicx}
%\usepackage{amsmath}
%\usepackage{amsfonts}
%\usepackage{amssymb}

\title{INSTRUCTIONS TO PREPARE A ONE PAGE ABSTRACT FOR THE
ECCM 6 AND FOR THE ECFD 7}

\author{First A. Author$^{1}$, Second B. Author$^{2}$ and Third C. Author$^{3}$}

\address{$^{1}$ Affiliation, Postal Address, E-mail address and URL
\and
$^{2}$ Affiliation, Postal Address, E-mail address and URL
\and
$^{3}$ Affiliation, Postal Address, E-mail address and URL}

\begin{document}

\noindent {\bf Keywords}: {\it Instructions, Multiphysics Problems, Applications, Computing Methods}
\vskip0.5cm

Authors are invited to submit electronically a one page abstract through the Congress web site before December 15, 2017. Abstracts should outline the main features, results and conclusions as well as their general significance, and contain relevant references.

The Abstract can be submitted directly in its final form. Authors will have the possibility of replacing the file by an updated version after the acceptance notification.

The Abstract should be written following the format. The file must be converted to Portable Document Format (PDF) before submission through the Congress web site. Other formats are not accepted by the system.

The Abstract must be written in English following this format template. It must contain the full name and full address of author/s. In the case of joint authorships, the name of the author who will actually present the paper at the Conference should be indicated with an asterisk. Papers can only be accepted on the understanding that they will be presented at the Conference.

Preliminary acceptance of the contribution will be communicated to the corresponding author by January 22, 2018.

Final acceptance of papers for presentation is conditional to receiving the final Abstract in format for publication and the payment of the presenting author's conference registration fee before March 15, 2018. Submission of full papers is not mandatory. Only one presentation per delegate is allowed.

For any question, please contact the ECCM-ECFD 2018 Conference Secretariat.
E-mail: eccm-ecfd2018\_sec@cimne.upc.edu.

\begin{thebibliography}{99}
\bibitem{Zienkiewicz}  O.C. Zienkiewicz, O.C. and  R.L. Taylor,  \textit{The finite element method}. 6th Edition, Elsevier, 2005.
\bibitem{OnCe93} E. O\~{n}ate and M. Cervera, Derivation of thin plate bending elements with one degree of freedom per node.
\textit{Engng. Comput.}, Vol. \textbf{10}, pp. 543--561, 1993. 
\end{thebibliography}

\end{document}

### 2018-06-07 talks/ECCM2018/eccm1.tex
updating ECCM talks

--- 2018-06-07 talks/ECCM2018/abstract.tex
@@
-\documentclass[a4paper]{article} %
-\usepackage{graphicx,amssymb} %
+\documentclass[12pt]{eccm-ecfd_abstract}
 
-\textwidth=15cm \hoffset=-1.2cm %
-\textheight=25cm \voffset=-2cm %
+\usepackage{graphicx}
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{amssymb}
+\usepackage{hyperref}
 
-\pagestyle{empty} %
+\title{Entropy stable discontinuous Galerkin methods using arbitrary bases and quadratures}
 
-\date{} %
+\author{Jesse Chan$^{1}$}
 
-\def\keywords#1{\begin{center}{\bf Keywords}\\{#1}\end{center}} %
-
-% Please, do not change any of the above lines
+\address{$^{1}$ Rice University, 6100 Main St MS-134, Houston, TX 77005,\\{jesse.chan@rice.edu}, http://www.caam.rice.edu/\textasciitilde jchan985/}
+%\and
+%$^{2}$ Affiliation, Postal Address, E-mail address and URL
+%\and
+%$^{3}$ Affiliation, Postal Address, E-mail address and URL}
 
@@
 
-% Type down your paper title
-\title{Entropy stable discontinuous Galerkin methods using arbitrary bases and quadratures}
+\noindent {\bf Keywords}: {\it High order, discontinuous Galerkin, entropy stability}
+\vskip0.5cm
 
-% Authors
-\author{Jesse Chan, \\ %
-       Rice University (USA) \\ \\ % Affiliation 1
-       \tt{jesse.chan@rice.edu} % Only one corresponding e-mail
-       }%
-
-\maketitle
-
-\thispagestyle{empty}
-
-% The abstract
-
-\begin{abstract}
 High order discontinuous Galerkin (DG) methods offer several advantages in the approximation of solutions of nonlinear conservation laws, such as improved accuracy and low numerical dispersion/dissipation.
@@
 In this talk, we discuss how to construct discretely entropy stable discontinuous Galerkin methods for general choices of basis and quadrature using discrete $L^2$ projection operators and ``decoupled'' SBP operators.
-Extensions to curvilinear meshes and implementational aspects on Graphics Processing Units will be also discussed.
-\end{abstract}
+%Extensions to curvilinear meshes  and implementational aspects on Graphics Processing Units will be also discussed.
 
-\keywords{High order, discontinuous Galerkin, entropy stability} % Write down at least 3 Keywords
-
-% \section{Introduction}
+\begin{thebibliography}{99}
+\bibitem{Chan}  Jesse Chan, 
+On discretely entropy conservative and entropy stable discontinuous Galerkin methods, Journal of Computational Physics, Volume 362, 2018, Pages 346-374.
+\end{thebibliography}
 

### 2018-06-07 talks/ECCM2018/eccm2.tex
updating ECCM talks

--- 2018-06-07 talks/ECCM2018/abstract2.tex
@@
-\documentclass[a4paper]{article} %
-\usepackage{graphicx,amssymb} %
+\documentclass[12pt]{eccm-ecfd_abstract}
 
-\textwidth=15cm \hoffset=-1.2cm %
-\textheight=25cm \voffset=-2cm %
+\usepackage{graphicx}
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{amssymb}
+\usepackage{verbatim}
 
-\pagestyle{empty} %
+\title{Efficient explicit solvers for multipatch discontinuous Galerkin isogeometric analysis}
 
-\date{} %
+\author{Jesse Chan$^{1}$ and John A. Evans$^{2}$}
 
-\def\keywords#1{\begin{center}{\bf Keywords}\\{#1}\end{center}} %
-
-% Please, do not change any of the above lines
+\address{$^{1}$ Rice University, 6100 Main St MS-134, Houston, TX 77005,\\{jesse.chan@rice.edu}, http://www.caam.rice.edu/\textasciitilde jchan985
+\and $^{2}$ University of Colorado Boulder, 1111 Engineering Drive, Boulder, CO 80309,\\john.a.evans@colorado.edu, {spot.colorado.edu/\textasciitilde joev9679}}
+%\and
+%$^{3}$ Affiliation, Postal Address, E-mail address and URL}
 
@@
 
-% Type down your paper title
-\title{Efficient explicit solvers for multipatch discontinuous Galerkin isogeometric analysis}
+\noindent {\bf Keywords}: {\it Isogeometric analysis, high order, mass matrix, explicit time-stepping}
+\vskip0.5cm
 
-% Authors
-\author{Jesse Chan, \\ %
-       Rice University (USA) \\ \\ % Affiliation 1
-       John A. Evans, \\ %
-       UC Boulder (USA) \\ \\ % Affiliation 1
-       \tt{jesse.chan@rice.edu} % Only one corresponding e-mail
-       }%
-
-\maketitle
-
-\thispagestyle{empty}
-
-% The abstract
-
-\begin{abstract}
 In this talk, we present explicit solvers for time-dependent hyperbolic problems on curved geometries using isogeometric analysis and spline-based finite element methods.
@@
 
-%We present a class of spline finite element methods for time-domain wave propagation which are particularly amenable to explicit time-stepping.
-The proposed methods utilize a discontinuous Galerkin discretization to enforce continuity of the solution field across geometric patches in a multi-patch setting, which yields a mass matrix with convenient block diagonal structure.
-Over each patch, we show how to accurately and efficiently invert mass matrices in the presence of curved geometries by using a weight-adjusted approximation of the mass matrix inverse.
-This approximation restores a tensor product structure while retaining provable high order accuracy and semi-discrete energy stability.
-We also estimate the maximum stable timestep for spline-based finite elements and show that the use of spline spaces result in less stringent CFL restrictions than equivalent piecewise continuous or discontinuous finite element spaces.
-Finally, we explore the use of optimal knot vectors based on $L^2$ $n$-widths.
-We show how the use of optimal knot vectors can improve both approximation properties and the maximum stable timestep, and present a simple heuristic method for approximating optimal knot positions.
-Numerical experiments confirm the accuracy and stability of the proposed methods.
-
-\end{abstract}
-
-\keywords{Isogeometric analysis, high order, mass matrix, explicit time-stepping} % Write down at least 3 Keywords
-
-\begin{thebibliography}{1}
+\begin{thebibliography}{99}
 \bibitem{key} Chan, Jesse, and John A.
 Evans.
-\emph{Multi-patch discontinuous Galerkin spline finite element methods for time-domain wave propagation.} arXiv preprint arXiv:1708.02972 (2017).
-Submitted to CMAME, in revision.
-\bibitem{mpiga} Langer, Ulrich, Angelos Mantzaflaris, Stephen E.
+"Multi-patch discontinuous Galerkin isogeometric analysis for wave propagation: Explicit time-stepping and efficient mass matrix inversion." Computer Methods in Applied Mechanics and Engineering 333 (2018): 22-54.
+%\bibitem{mpiga} Langer, Ulrich, Angelos Mantzaflaris, Stephen E.
 Moore, and Ioannis Toulopoulos.
@@
 
-% \section{Introduction}
+%\begin{thebibliography}{99}
+%\bibitem{Chan}  
+%\end{thebibliography}
 

### 2018-06-07 talks/ECCM2018/iga/revision/main.tex
updating ECCM talks

\pdfoutput=1

\documentclass[preprint,10pt]{elsarticle}
%\documentclass[preprint,10pt]{article}
%\documentclass[review]{siamart0216}
%\documentclass{siamart0216}

\usepackage{fullpage}
\usepackage[a-1b]{pdfx}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage[titletoc,toc,title]{appendix}

\usepackage{array} 
\usepackage{listings}
\usepackage{mathtools}
\usepackage{pdfpages}
\usepackage[textsize=footnotesize,color=green]{todonotes}
\usepackage{bm}
%\usepackage{tikz}
\usepackage[normalem]{ulem}
\usepackage{hhline}

%% ====================================== alg package
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{algorithmicx}
\algblock{ParFor}{EndParFor}
% customising the new block
\algnewcommand\algorithmicparfor{\textbf{parfor}}
\algnewcommand\algorithmicpardo{\textbf{do}}
\algnewcommand\algorithmicendparfor{\textbf{end\ parfor}}
\algrenewtext{ParFor}[1]{\algorithmicparfor\ #1\ \algorithmicpardo}
\algrenewtext{EndParFor}{\algorithmicendparfor}
%% ====================================== end alg package

\usepackage{graphicx}
\usepackage{subfig}
\usepackage{color}

\definecolor{forestgreen}{rgb}{0, 0.5, 0}

%% ====================================== graphics

\usepackage{pgfplots}
\usepackage{pgfplotstable}
\definecolor{markercolor}{RGB}{124.9, 255, 160.65}
\pgfplotsset{width=10cm}
\pgfplotsset{
tick label style={font=\small},
label style={font=\small},
legend style={font=\small}
}

\usetikzlibrary{calc}

%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
\newcommand{\logLogSlopeTriangle}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        \pgfmathsetmacro{\xArel}{#1}
        \pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{\yArel}
        \pgfmathsetmacro{\xCrel}{\xArel}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=north] {1}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=west] {#4}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

\newcommand{\logLogSlopeTriangleNeg}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        \pgfmathsetmacro{\xArel}{#1}
        \pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{\yArel}
        \pgfmathsetmacro{\xCrel}{\xArel}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=.5,anchor=south] {1}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=west] {#4}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
\newcommand{\logLogSlopeTriangleFlipNeg}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        %\pgfmathsetmacro{\xArel}{#1}
        %\pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{#3}
        \pgfmathsetmacro{\xCrel}{#1}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

	\pgfmathsetmacro{\xArel}{\xBrel}
        \pgfmathsetmacro{\yArel}{\yCrel}

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=east] {#4}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=north] {1}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
\newcommand{\logLogSlopeTriangleFlip}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        %\pgfmathsetmacro{\xArel}{#1}
        %\pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{#3}
        \pgfmathsetmacro{\xCrel}{#1}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

	\pgfmathsetmacro{\xArel}{\xBrel}
        \pgfmathsetmacro{\yArel}{\yCrel}

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=east] {#4}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=south] {1}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

\usepackage{stmaryrd}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

\newcommand{\vect}[1]{\ensuremath\boldsymbol{#1}}
\newcommand{\tensor}[1]{\underline{\bm{#1}}}
\newcommand{\del}{\triangle}
\newcommand{\curl}{\grad \times}
\renewcommand{\div}{\grad \cdot}

\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\equaldef}{\stackrel{\mathrm{def}}{=}}

\newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}{\rm #2}}}
\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
\newcommand{\pdd}[2]{\frac{\partial^2#1}{\partial#2^2}}
\newcommand{\pdn}[3]{\frac{\partial^{#3}#1}{\partial#2^{#3}}}
\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\snor}[1]{\left| #1 \right|}
\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRb}[1]{\left| #1 \right|}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}
\newcommand{\LRceil}[1]{\left\lceil #1 \right\rceil}
\newcommand{\LRl}[1]{\left. #1 \right|}

%\newcommand{\cond}[1]{\kappa\LRp{#1}}
\newcommand{\cond}[2]{\nor{#1}_{#2}\nor{{#1}^{-1}}_{#2}}

\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\Div} {\ensuremath{\nabla\cdot}}
\newcommand{\jump}[1] {\ensuremath{\llbracket#1\rrbracket}}
\newcommand{\avg}[1] {\ensuremath{\LRc{\!\{#1\}\!}}}

\newcommand{\Oh}{{\Omega_h}}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\Lk}{L^2\LRp{D^k}}
\newcommand{\Ldk}{L^2\LRp{\partial D^k}}
%\newcommand{\Lk}{D^k}
%\newcommand{\Ldk}{\partial D^k}
\newcommand{\Dhat}{\widehat{D}}
\newcommand{\Lhat}{L^2\LRp{\Dhat}}

%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{lemma}[theorem]{Lemma}
%\newtheorem{proposition}[theorem]{Proposition}
%\newtheorem{corollary}[theorem]{Corollary}

%\newenvironment{definition}[1][Definition]{\begin{trivlist}
%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
%\newenvironment{example}[1][Example]{\begin{trivlist}
%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\newcommand{\eval}[2][\right]{\relax
  \ifx#1\right\relax \left.\fi#2#1\rvert}

\def\etal{{\it et al.~}}

%\newcommand{\reviewerOne}[1]{{\color{blue}{#1}}}
%\newcommand{\reviewerTwo}[1]{{\color{red}{#1}}}
\newcommand{\reviewerOne}[1]{#1}
\newcommand{\reviewerTwo}[1]{#1}
\newcommand{\note}[1]{{\color{blue}{#1}}}
\newcommand{\remark}[1]{\textbf{\color{red}#1}}

\newcommand{\LinfDk}{L^{\infty}\LRp{D^k}}

\newcommand{\diag}[1]{{\rm diag}\LRp{#1}}

\newcommand{\Ksub}{K_{\rm sub}}

\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

%% d in integrand
\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}}

\makeatletter
\renewcommand\d[1]{\mspace{6mu}\mathrm{d}#1\@ifnextchar\d{\mspace{-3mu}}{}}
\makeatother

\date{}
%\author{Jesse Chan, John Evans}
%\title{Multi-patch discontinuous Galerkin spline finite element methods for time-domain wave propagation}
\graphicspath{{./figs/}}

\begin{document}

%\maketitle
%\tableofcontents
\begin{frontmatter}
\author[rice]{Jesse Chan\corref{cor1}}
\ead{Jesse.Chan@caam.rice.edu}
\cortext[cor1]{Principal Corresponding author}
\author[vt]{John A.\ Evans}
\ead{John.A.Evans@colorado.edu}
\address[rice]{Department of Computational and Applied Mathematics, Rice University, 6100 Main St, Houston, TX, 77005}
\address[vt]{Ann and H.J. Smead Aerospace Engineering Sciences, University of Colorado Boulder, 429 UCB, Boulder, Colorado, 80309}

%\title{Multi-patch discontinuous Galerkin spline finite element methods for time-domain wave propagation}
\title{Multi-patch discontinuous Galerkin isogeometric analysis for wave propagation: explicit time-stepping and efficient mass matrix inversion}

\begin{abstract}
We present a class of spline finite element methods for time-domain wave propagation which are particularly amenable to explicit time-stepping.  The proposed methods utilize a discontinuous Galerkin discretization to enforce continuity of the solution field across geometric patches in a multi-patch setting, which yields a mass matrix with convenient block diagonal structure.  Over each patch, we show how to accurately and efficiently invert mass matrices in the presence of curved geometries by using a weight-adjusted approximation of the mass matrix inverse.  This approximation restores a tensor product structure while retaining provable high order accuracy and semi-discrete energy stability.  We also estimate the maximum stable timestep for spline-based finite elements and show that the use of spline spaces result in less stringent CFL restrictions than equivalent $C^0$ or discontinuous finite element spaces.  Finally, we explore the use of optimal knot vectors based on $L^2$ $n$-widths.  We show how the use of optimal knot vectors can improve both approximation properties and the maximum stable timestep, and present a simple heuristic method for approximating optimal knot positions.  Numerical experiments confirm the accuracy and stability of the proposed methods.  
\end{abstract}
\end{frontmatter}
\section{Introduction}

The key concept behind isogeometric analysis is the integration of computer representations of geometry in design and numerical simulations.  
%computer representations of geometry (i.e.\ CAD) and numerical simulations.  
This integration aims to address the engineering bottleneck in converting engineering designs into a form appropriate for analysis \cite{hughes2005isogeometric} by {representing solutions of PDEs using the spline representations underlying geometric parametrizations}.  While the main advantage of isogeometric analysis is the elimination of errors in the approximation of curved geometries, it has been observed that discretizations using high order B-spline and NURBS possess advantages over traditional high order $C^0$ finite element methods, including greater efficiency per degree of freedom, efficient high order accurate approximations of eigenvalues and eigenfunctions of continuous differential operators, and {larger maximum stable time-step sizes for explicit dynamics} \cite{hughes2008duality,evans2009n, hughes2014finite}.

In \cite{hughes2008duality}, NURBS-based discretizations were compared to traditional $C^0$ finite element methods for steady-state wave propagation problems such as time-harmonic wave propagation and eigenvalue problems in structural vibrations.  The authors noted that, compared to $C^0$ finite element discretizations, B-spline and NURBS discretizations approximated high frequency modes more accurately relative to the number of degrees of freedom.  These properties of spline and NURBS were also shown to translate into improved numerical simulations of one-dimensional time-dependent wave propagation in \cite{hughes2014finite}.  However, {the application of NURBS-based finite element discretizations to higher dimensional time-dependent wave propagation problems faces several computational challenges, such as the cost of inverting spline mass matrices and dealing with multi-patch geometries}.
%\footnote{Spline collocation methods have been applied to higher dimensional time-dependent wave propagation problems \cite{auricchio2012isogeometric}; however, these are not equivalent in general to finite element discretiations.}  
The goal of this work is to make the advances of isogeometric analysis simpler to realize through the use of multi-patch discontinuous Galerkin discretizations of time-dependent wave propagation problems in two and three dimensions, focusing in particular on explicit time integration.

We consider geometries which are constructed using multiple geometric ``patches'', which are mappings of a reference patch to physical space.  While a single NURBS patch can exactly represent conic sections, \reviewerTwo{geo-metrically} and topologically complex objects require representations in terms of multiple NURBS patches.  We discretize these geometries using a multi-patch discontinuous Galerkin method \cite{langer2015multipatch}, which imposes continuity conditions weakly across patch interfaces through a numerical flux \cite{nguyen2014nitsche}.  These methods are analogous to multi-domain spectral methods \cite{canuto2012spectral}, which combine the accuracy of spectral methods with the use of multiple domains (patches) for geometric flexibility.  %We study multi-patch DG discretizations of model problems in wave propagation in both first and second order form.

When paired with explicit time integration, multi-patch DG methods require the inversion of a \reviewerTwo{sparse but large} mass matrix over each patch at every timestep.   
%Many isogeometric methods for time-dependent problems have traditionally utilized either implicit time integration or space-time formulations \cite{cottrell2009isogeometric}.  
%The combination of NURBS with explicit time-stepping is uncommon due to the fact that, 
For geometric patches which are affine transformations of reference patches, the tensor product structure of NURBS bases result in mass matrices which can be inverted by applying inverses of one-dimensional matrices.  However, for curved geometries of practical engineering interest, the Kronecker structure of the mass matrix inverse is lost.  We address the issue of mass matrix inversion by using a weight-adjusted approximation to the mass matrix, which restores a tensor product structure to the approximate mass inverse while maintaining provable high order accuracy and energy stability.  %These properties address existing approaches to efficient explicit time-integration, such as isogeometric collocation and preconditioned iterative methods \cite{auricchio2012isogeometric, gao2014fast}.

This work also studies timestep restrictions for explicit methods under high order NURBS finite element discretizations.  High order $C^0$ and DG finite element discretizations possess maximum stable timesteps which scale as $O(h/p^2)$, where $p$ is the order of approximation {\color{forestgreen} and $h$ is the mesh size} \cite{warburton2008taming}.  By using special filters or discretizations involving dual grids, the maximum stable timestep can be made to scale as $O(h/p)$ instead \cite{warburton2008taming, reyna2014operator}, though these methods can be difficult to generalize to unstructured meshes.  In this work, we derive estimates for the maximum stable timestep based on a modification of arguments in \cite{chan2015gpu}.  These estimates depend on constants in $hp$ inverse and trace inequalities.  
%{\color{forestgreen} \sout{Multi-patch spline discretizations introduce two meshes: a macro-element mesh consisting of geometric ``patches'' of size $H$, and a micro-element mesh containing $K$ micro-elements {\color{forestgreen} per-side} over which a spline approximation is defined.  While explicit estimates for these constants are not currently available except in special cases, numerical experiments show that these constants (and the resulting spectral radius of the NURBS discretization matrix) grow as $O(hH/p)$ for a micro-element mesh size $h$ which is sufficient small with respect to $p$.} 
While explicit values for these constants are not currently available except in special cases, numerical experiments show that the maximum stable timestep for multi-patch discontinuous Galerkin discretizations scales like $O(h/p)$ provided the number of elements per patch is sufficiently large with respect to $p$.\footnote{We note that the slow growth of the maximum stable timestep with $p$ appears to be correlated with the use of extended-support basis functions.  For example, finite element discretizations using extended-support $C^0$ continuous piecewise polynomial bases also possess a maximum stable timestep which scales more slowly in $p$ than traditional high order $C^0$ and DG finite element methods \cite{banks2016galerkin}.}

Finally, we investigate the use of optimal spline spaces in NURBS discretizations, which are defined using knot locations at the roots of a specific eigenfunction \cite{melkman1978spline}.  These optimal knot positions bear some resemblance to the nonlinear parametrization used in \cite{cottrell2006isogeometric} to eliminate spurious high frequency ``outlier'' modes present in isogeometric discretizations with uniform knots.  Numerical experiments show that, under certain circumstances, spline spaces defined using optimal knot vectors are more efficient than splines defined using uniform knot vectors in terms of degrees of freedom required to reach a certain error.  Moreover, for time-dependent problems, the use of optimal knot vectors increases the maximum stable timestep while retaining reasonable approximation properties.

The structure of the paper is as follows: Section~\ref{sec:form2} discusses Galerkin formulations of hyperbolic PDEs on multi-patch geometries.  Section~\ref{sec:basis} introduces B-spline bases and optimal knot distributions, and introduces an inexpensive heuristic iteration to approximate optimal knots locations.  Section~\ref{sec:mol} discusses the efficient solution of the semi-discrete finite element system using explicit time integration, including weight-adjusted approximations to the inverse mass matrix and estimates for stable timestep restrictions based on constants in $hp$ inverse and trace inequalities.  Finally, Section~\ref{sec:num} presents numerical experiments illustrating salient properties of the proposed methods, and Section~\ref{sec:conc} concludes with some future directions.

\section{Multi-patch discontinuous Galerkin formulations}
\label{sec:form2}

In this work, we are interested in the solution of the hyperbolic PDEs of acoustic wave propagation and linear advection over a domain $\Omega \in \mathbb{R}^d$ with boundary $\partial \Omega$.  We begin the model problem of time-dependent advection \reviewerOne{of some concentration $\phi$} (with a divergence-free velocity field $\bm{\beta}$) 
\begin{align*}
\pd{\phi}{t} + \Div \LRp{\bm{\beta}\phi} &= 0, \qquad \Div\bm{\beta} = 0\\
\phi(\bm{x},0) &= \phi_0(\bm{x}).
\end{align*}
Since the advection equation will primarily be used to study dispersive and dissipative properties of the proposed discretizations, periodic boundary conditions will be imposed on the boundary $\partial \Omega$.

We consider also the acoustic wave equation as a first order pressure-velocity system
\begin{align}
\frac{1}{c^2}\pd{p}{t} + \Div \bm{u} &= f \label{eq:waveeq}\\
\pd{\bm{u}}{t} + \Grad p &= 0. \nonumber
\end{align} 
\reviewerOne{where $p$ and $\bm{u}$ represent the pressure and velocity variables, respectively, and $c$ is the wavespeed.}  
This problem is well-posed when paired with either Dirichlet or Neumann boundary conditions on $\partial \Omega$ and initial conditions 
\[
p(\bm{x},0) = p_0(\bm{x}), \qquad \bm{u}(\bm{x},0) = \bm{u}_0(\bm{x}).
\]
The second order form of the acoustic wave equation can be derived from the first order form.  Taking the derivative in time of the pressure equation and the divergence of the velocity equation and summing the two equations yields
\[
\reviewerOne{\frac{1}{c^2}}\pdd{p}{t} - \Delta p = f,  
\]
\reviewerOne{where $f$ is related to, but not necessarily the same as the $f$ in (\ref{eq:waveeq}).}
The second order formulation is well posed for appropriate boundary conditions and initial conditions 
\[
p(\bm{x},0) = p_0(\bm{x}), \qquad \pd{p}{t}(\bm{x},0) = v_0(\bm{x}).  
\]

%Let $\bm{n}$ be the outward unit normal on $\Omega$; the inflow and outflow boundaries $\Gamma_{\rm in}, \Gamma_{\rm out} \subset \partial \Omega$ are then defined as
%\[
%\Gamma_{\rm in} = \LRc{\bm{x}: \bm{\beta}\cdot \bm{n} < 0}, \qquad \Gamma_{\rm out} = \LRc{\bm{x}: \bm{\beta}\cdot \bm{n} \geq 0}
%\]
%Boundary conditions are imposed on the inflow domain $\left.u\right|_{\Gamma_{\rm in}} = u_0$.

In practical applications, sufficiently complex geometries are represented using multiple curved geometric patches.  We assume that $\Omega$ is decomposed into non-overlapping patches $D^k$, each of which is the image of the reference patch $\Dhat$ with coordinates $\widehat{\bm{x}}$ under a geometric mapping $\bm{x} = \bm{\Phi}_k(\bm{\widehat{x}})$, as shown in Figure~\ref{fig:multipatch_diagram}.   We define the mesh $\Omega_h = \cup D^k$ and the corresponding global approximation space $V_h(\Omega_h) = \bigoplus V_h\LRp{D^k}$, where $V_h\LRp{D^k}$ is the local approximation space over a single patch.  We further assume that the local approximation space over $D^k$ is the image of a reference approximation space $V_h\LRp{\widehat{D}}$ under $\bm{\Phi}_k$, and  that trace and inverse inequalities hold on $\Dhat$ such that 
\begin{equation}
\nor{u}_{L^2\LRp{\partial \Dhat}} \leq \sqrt{C_T} \nor{u}_{L^2\LRp{\Dhat}}, \qquad \nor{\widehat{\Grad} u}_{L^2\LRp{\Dhat}} \leq C_I \nor{u}_{L^2\LRp{\Dhat}}, \qquad \forall v\in V_h\LRp{\Dhat},
%\nor{u}^2_{L^2\LRp{\partial \Dhat}} \leq C_T \nor{u}^2_{L^2\LRp{\Dhat}}, \qquad \nor{\widehat{\Grad} u}^2_{L^2\LRp{\Dhat}} \leq C_I^2 \nor{u}^2_{L^2\LRp{\Dhat}}, \qquad \forall v\in V_h\LRp{\Dhat}
\label{eq:ineqs}
\end{equation}
where $C_T, C_I$ are constants which depend on the space $V_h\LRp{\Dhat}$.  We will describe specific approximation spaces in more detail in Section~\ref{sec:basis}.  %We note that we have used the norm squared for the trace constant, but only the norm for the inverse inequality constant.  This is to ensure that the scaling
{\color{forestgreen} We emphasize that the constant $C_I$ appears in the inverse inequality in (\ref{eq:ineqs}), but the constant $\sqrt{C_T}$ appears in the corresponding trace inequality.  The trace inequality $C_T$ in this work is thus the square of the constant typically presented $hp$ finite element trace and inverse inequalities \cite{ciarlet1978finite, warburton2003constants}.  This definition ensures that $C_T, C_I$ both scale similarly (as shown in Section~\ref{sec:numconsts}) and simplifies expressions for bounds involved in estimating stable timestep restrictions in Section~\ref{sec:cfl}.  }

\begin{figure}
\centering
\includegraphics[width=.4\textwidth]{multipatch_diagram.pdf}
\caption{An example of a multi-patch geometry.}
\label{fig:multipatch_diagram}
\end{figure}

We will discretize the equations of interest based on a discontinuous Galerkin (DG) approach \cite{hesthaven2007nodal,langer2015multipatch}, and will explore first order variational formulations of the acoustic wave equation and linear advection, as well as second order formulations of acoustic wave propagation.  We first introduce the jump and average of discontinuous functions across {\color{forestgreen} patch} interfaces.  Let $D^{k,+}$ denote the neighboring {\color{forestgreen}  patch} \reviewerTwo{sharing} a face $f$ of $D^k$, and let $u^+,u^-$ denote the values of $u$ on $D^{k,+}$ and $D^k$, respectively.  The jump \reviewerTwo{and average} of $u$ across $f$ is then defined as
\[
\jump{u} = u^+ - u^-, \qquad \avg{u} = \frac{u^+ + u^-}{2}.
\]
The jump and average of vector fields are defined component-wise using the jumps and averages of components.

We note that all formulations presented in the following sections are constructed to be \textit{a-priori} energy stable so long as integrals are explicitly approximated using quadrature (as opposed to quadrature-free implementations).  Additionally, high order accuracy is observed in all numerical experiments where the quadrature is sufficiently accurate \cite{chan2016weight2}.  These formulations are especially important in context of curvilinear meshes, where exact evaluation integrals is inefficient or impossible due to the high order or rational nature of the determinant of the mapping Jacobian $J$.

\subsection{Formulation for linear advection}

We begin with the advection equation, for which we can construct a discontinuous Galerkin formulation based on \cite{kopriva2014energy}: \reviewerTwo{find $\phi \in V_h(\Omega_h)$ such that}
\begin{align}
\sum_k\LRp{\pd{\phi}{t},v}_{L^2\LRp{D^k}} + a(\phi,v) &= 0, \qquad \reviewerTwo{\forall v\in V_h(\Omega_h),}\label{eq:varformadvec}\\
a(\phi,v) &= a_{\rm skew}(\phi,v) + a_{\rm sym}(\phi,v) \nonumber\\
a_{\rm skew}(\phi,v) &= \frac{1}{2}\sum_k\int_{\Omega} \LRp{-\phi\bm{\beta}\cdot \Grad v + \bm{\beta}\cdot \Grad \phi v}\diff{{\bm{x}}}\nonumber\\
a_{\rm sym}(\phi,v) &= \frac{1}{2}\sum_{D^k}\int_{\partial D^k} \LRp{\avg{\bm{\beta}}\cdot\bm{n} - \tau\LRb{\avg{\bm{\beta}}\cdot\bm{n}}} \jump{\phi} v \diff{{\bm{x}}}.\nonumber  
\end{align}
where $0 \leq \tau$ is a parameter which switches between a dissipative flux ($\tau > 0$) and a central flux ($\tau = 0$).  Assuming appropriate boundary conditions (homogeneous or periodic), taking $v = \phi$ shows the formulation to be energy stable, such that
\[
\reviewerOne{\pd{}{t}\sum_{k}\LRp{\phi,\phi}_{D^k} \leq 0.}
\]

\subsection{First order formulation of the acoustic wave equation}

We now consider the acoustic wave equation, and follow approaches outlined in \cite{chan2016weight1, chan2016weight2} to construct an energy-stable discretization using a skew-symmetric formulation and penalty flux.  This formulation is given for $U = \LRc{p,\bm{u}}, V = \LRc{q,\bm{v}}$ as follows: \reviewerTwo{find $U \in V_h(\Omega_h) \times V_h(\Omega_h)^d$ such that}
\begin{align}
\sum_k\LRp{\frac{1}{c^2}\pd{p}{t},q}_{L^2\LRp{D^k}} + \LRp{\pd{\bm{u}}{t},\bm{v}}_{L^2\LRp{D^k}} + a(U,V) &= \sum_k\LRp{f,q}_{L^2\LRp{D^k}}, \qquad \reviewerTwo{V \in V_h(\Omega_h) \times V_h(\Omega_h)^d,} \label{eq:varformwave1}
\end{align}
where we define $a(U,V)$ as
\begin{align*}
a(U,V) &= \sum_k\LRp{-\LRp{\bm{u},\Grad q}_{L^2\LRp{D^k}} + \frac{1}{2}\LRa{2\avg{\bm{u}}\cdot\bm{n} - \tau_p \jump{p},q}_{\partial D^k}} \\
 &+ \sum_k\LRp{\LRp{\Grad p,\bm{v}}_{L^2\LRp{D^k}} + \frac{1}{2}\LRa{\jump{p} - \tau_u \jump{\bm{u}}\cdot\bm{n},\bm{v}\cdot\bm{n}}_{\partial D^k}},
\end{align*}
where $\tau_p, \tau_u$ are positive penalization parameters, and both the pressure $p$ and each component of the velocity $\bm{u}$ are approximated locally from $V_h\LRp{D^k}$.  The variational formulation over each patch can be mapped \reviewerTwo{(or pulled back)} to $\Dhat$ as
\begin{align*}
\int_{\Dhat} \pd{p}{t} q J\diff{\bm{\widehat{x}}} + \int_{\Dhat} \widehat{\Grad}\cdot \LRp{\bm{G}^T\bm{u}}qJ\diff{\bm{\widehat{x}}} + \frac{1}{2}\int_{\partial \widehat{D}}\LRp{2\avg{\bm{u}}\cdot\bm{n} - \tau_p \jump{p}}J^s &= \int_{\Dhat} f q J\diff{\bm{\widehat{x}}}\\
\int_{\Dhat} \pd{\bm{u}}{t} \bm{v} J\diff{\bm{\widehat{x}}} + \int_{\Dhat} \bm{G}\widehat{\Grad} {p}\bm{v}J\diff{\bm{\widehat{x}}} + \frac{1}{2}\int_{\partial \widehat{D}}\LRp{\jump{p} - \tau_u \jump{\bm{u}}\cdot\bm{n}}\bm{v}\cdot\bm{n} J^s &=0, 
\end{align*}
where $\bm{G}_{ij} = \pd{\widehat{\bm{x}}_i}{\bm{x}_j}$ \reviewerTwo{is the inverse of the Jacobian matrix of $\bm{\Phi}_k(\hat{x})$ (also referred to as the metric terms \cite{kopriva2006metric} or geometric factors \cite{hesthaven2007nodal})}, and $J, J^s$ denote the determinants of volume and surface geometric Jacobians, respectively.

The global formulation results from the summation of local formulations over each {\color{forestgreen}  patch}.  For constant wavespeed $c^2$  and $\tau_p, \tau_u = 1$, the formulation reduces to the standard upwind DG method.  A standard energy argument gives that the method is dissipative for $\tau_p, \tau_u > 0$, in the sense that 
\[
\pd{}{t}\sum_k \int_{D^k}\LRp{\frac{p^2}{c^2} + \LRb{\bm{u}}^2}\diff{{\bm{x}}} \leq -\sum_k \int_{\partial D^k} \LRp{\tau_p\jump{p}^2 + \tau_u\LRb{\jump{\bm{u}}\cdot \bm{n}}^2}\diff{{\bm{x}}}.
\]
The dissipation resulting from the stabilization terms provides a natural way to damp spurious non-conforming components of the solution in time-domain simulations \cite{chan2016short}.  Similar energy stable skew-symmetric formulations can be constructed for electromagnetics and elastodynamics \cite{warburton2013low, chan2017weight}.

Dirichlet boundary conditions on pressure are enforced in an energy-stable fashion by specifying the jump of $p$ on boundary faces 
\[
\jump{p} = 2 (p_D - p),
\]
where $p_D$ is the prescribed value of $p$ on the boundary.  Similarly, Neumann boundary conditions are enforced by specifying the jump of $\bm{u}$ on boundary faces
\[
\jump{\bm{u}}\cdot \bm{n} = 2 (u_N - \bm{u}\cdot\bm{n}),
\]
where $u_N$ is the prescribed value of the normal flux on the boundary.

\subsection{Second order formulation of the acoustic wave equation}

We use the acoustic wave equation as a prototypical example of a second order formulation.  We discretize the Laplacian using a symmetric interior penalty DG method \cite{arnold1982interior, langer2015multipatch}.  The resulting formulation is given as follows: \reviewerTwo{find $p\in V_h(\Omega_h)$ such that}
\begin{align}
\sum_k\LRp{\pdd{p}{t},v}_{L^2\LRp{D^k}} + a(p,v) &= \sum_k\LRp{f,v}_{L^2\LRp{D^k}}, \qquad \reviewerTwo{\forall v \in V_h(\Omega_h)}\\
a(p,v) &= \sum_k\LRp{\Grad p,\Grad v}_{L^2\LRp{D^k}}  -\LRa{\avg{\Grad p}\cdot\bm{n}, v}_{L^2\LRp{\partial D^k}} \label{eq:varformwave2}\\
&+ \sum_k \LRa{\jump{p}, \Grad v\cdot\bm{n}}_{L^2\LRp{\partial D^k}} -\tau\LRa{\jump{p},v}_{L^2\LRp{\partial D^k}}\nonumber.
\end{align}
where $\tau$ is a penalization parameter.  In contrast with first order systems, the second order system achieves optimal rates of convergence in the $L^2$ norm and is non-dissipative and energy conserving.  % (though dissipative second order formulations have been introduced in \cite{appelonew}).

We note that, in order to ensure coercivity and stability, the penalization constant $\tau$ must be sufficiently large with respect to discretization parameters.  We follow the approach in \cite{Shahbazi:2005:SNE:1083044.1083046}, where a straightforward generalization of the proof to curved patches yields the following sufficient global condition for coercivity
\[
%\tau \geq \max_{D^k} {C_T^2}{\nor{J^s}_{L^\infty\LRp{\partial D^k}}\nor{\frac{1}{J}}_{L^\infty\LRp{D^k}}}.
\tau \geq \max_{D^k} {C_T}{\nor{J^s}_{L^\infty\LRp{\partial D^k}}\nor{\frac{1}{J}}_{L^\infty\LRp{D^k}}}.
\]
Here, $C_T$ is the trace constant over the reference patch, while $J, J^s$ are Jacobian factors for the volume and surface (face) of a physical patch, respectively.\footnote{One may also define the penalization constant locally over each patch face; this may be useful for local time-stepping, as the maximum stable timestep is dependent on the magnitude of $\tau$.}

\section{B-spline bases}
\label{sec:basis}

We are interested in approximating solutions over each patch $D^k$ from spline spaces, which are piecewise polynomials of degree $p$ with up to $p-1$ continuous global derivatives.  The following sections briefly review the construction of B-spline basis functions, and introduce choices of knot vectors which modify the approximation properties of the corresponding spline space.

Univariate spline spaces of degree $p$ are constructed given an ordered knot vector $\bm{\Xi}$ 
\[
\bm{\Xi} = \LRc{-1 = \xi_1,\ldots,\xi_{{2p}+K+1} = 1}, \qquad \xi_i \leq \xi_{i+1}.  
\]
The unique (i.e. non-repeating) knots in $\bm{\Xi}$ define $K$ non-overlapping sub-intervals (elements) over which the spline space is defined.  B-spline basis functions can then be constructed recursively as follows
\[
B^0_i(\widehat{x}) = \begin{cases}
1, & \xi_i \leq \widehat{x} \leq \xi_{i+1}\\
0, & {\rm otherwise},
\end{cases}
\qquad 
B^k_i(\widehat{x}) = \frac{\widehat{x} - \xi_i}{\xi_{i+p}-\xi_i}B^{k-1}_i(\widehat{x}) + \frac{\xi_{i+p+1}-\widehat{x}}{\xi_{i+p+1}-\xi_{i+1}}B^{k-1}_{i+1}(\widehat{x}).
\]
In the case when $\xi_{i+p}-\xi_i = 0$, the ratio $\frac{\widehat{x} - \xi_i}{\xi_{i+p}-\xi_i}$ is also set to zero.  Similarly, if $\xi_{i+p+1}-\xi_{i+1} = 0$, $ \frac{\xi_{i+p+1}-\widehat{x}}{\xi_{i+p+1}-\xi_{i+1}}$ is also set to zero.  The resulting basis functions $B^p_i(\widehat{x})$ have local support and span a piecewise polynomial space of degree $p$.  In this work, we focus on \textit{open} knot vectors $\bm{\Xi}$, where the first and last knots are repeated $p$ times
\[
\xi_1 = \ldots = \xi_{p+1}, \qquad \xi_{p+1 + K} = \ldots = \xi_{2p+K+1}.
\]
The resulting bases contain $p+K$ basis functions, are interpolatory at the endpoints, and contain the space of degree $p$ polynomials $P^p([-1,1])$ as a proper subset.  For $K=1$, this basis recovers the degree $p$ Bernstein polynomial basis.  Finally, we define the one-dimensional degree $p$ approximation space over the reference interval $V_h([-1,1])$ as the span of B-splines 
\[
V_h([-1,1]) = {\rm span}\LRc{B^p_i(\widehat{x})}_{i=1}^{p+K}.
\]
For the remainder of this work, we focus on spline spaces of maximal continuity, such that the resulting degree $p$ splines possess $p-1$ continuous derivatives.  {\color{forestgreen} Moreover, we focus on uniform knot vectors, that is, knot vectors with sub-integrals of equal length.} %Spline spaces of reduced continuity may also be used, but are not the current focus of this work.

\reviewerOne{Any function in $V_h([-1,1])$ is represented as a linear combination of B-spline basis functions
\[
u(\hat{x}) = \sum_{j=1}^{p+K} c_j B^p_j(\hat{x}), \qquad \forall u \in V_h([-1,1]),
\]
where $c_j$ are the coefficients or ``control points'' of the curve defined by $u(x)$.  We note that $c_1, c_{p+K}$ interpolate the values of $u(-1)$ and $u(1)$, respectively.  However, since interior basis functions are not interpolatory, the solution does not necessarily coincide with the value of the interior coefficients/control points. }

\subsection{Higher dimensional splines and approximation spaces}

Given a one-dimensional spline basis, we can extend spline bases to $d$ dimensions through a tensor product construction. For example, B-spline basis functions $B^{p}_{ij}(\hat{x},\hat{y})$ in two dimensions and $B^{p}_{ijk}(\hat{x},\hat{y},\hat{z})$ in three dimensions can be constructed as
\[
B^{p}_{ij}(\widehat{x},\widehat{y}) = B^p_i(\widehat{x}) B^p_j(\widehat{y}), \qquad B^{p}_{ijk}(\widehat{x},\widehat{y},\widehat{z}) = B^p_i(\widehat{x}) B^p_j(\widehat{y})B^p_k(\widehat{z}), \qquad 1 \leq i,j,k\leq p+K.
\]
We can then define the local spline approximation space of order $p$ over the reference patch $\widehat{D}$ as 
\[
V_h\LRp{\widehat{D}} = {\rm span}\LRc{B^p_{i}\LRp{\bm{\widehat{x}}}}_{i=1}^{(p+K)^d},
\]
where $B^p_i\LRp{\bm{\widehat{x}}}$ denotes the $i^{\rm th}$ tensor product spline in $d$ dimensions \reviewerTwo{(here, we have switched from a multi-index $ijk$ to a single index $i$).}

\begin{figure}[t]
\centering
\includegraphics[width=.85\textwidth]{Definitions.pdf}
\caption{{\color{forestgreen} Illustrations of the patch size $H$, the mesh size $h$, and the number of elements per side $K$ for a particular patch $D^k$.  The dashed lines represented the knot lines both on the parametric patch $\hat{D}$ and the physical patch $D^k$.}}
\label{fig:definitions}
\end{figure}

For isotropic spline approximation spaces, we define the patch size $H$ as the diameter {\color{forestgreen}(which we take to mean the longest side)} of $D^k$.  We will refer to subdivision of patches into patches of size $H/2$ as patch refinement, while mesh refinement will refer to the generation of new knot vectors with $2K$ elements each. Finally, we refer to the physical mesh size as the patch size divided by the number of elements $h = H/K$.  We note that this definition assumes quasi-uniform patches and knot vectors.  {\color{forestgreen} We have graphically depicted all of the aforementioned geometric quantities in Figure \ref{fig:definitions}. Note that when only one element is included per patch, as is the case with the standard discontinuous Galerkin method using discontinuous piecewise polynomial approximations, then $K = 1$ and hence $h = H$.  Thus, the terms element and patch are interchangeable for this setting.}  Because the formulations presented in Section~\ref{sec:form2} are standard Galerkin formulations using spline basis functions, they are consistent and stable in the finite element sense, and the $h$-convergence of these methods on curved geometries can be proven using techniques adapted from \cite{warburton2013low} and interpolation estimates and inverse inequalities established in \cite{bazilevs2006isogeometric, evans2013explicit}.

Finally, we note that in engineering applications, the geometric mapping $\bm{\Phi}_k$ is often defined using a rational NURBS mapping.  An approach popularized in isogeometric analysis is the use of a NURBS space (consisting of rational B-splines) as an approximation space as well.  For the numerical experiments presented in this work, we do not utilize NURBS approximation spaces, and instead map B-splines defined on $\Dhat$ to the physical domain using the mapping $\bm{\Phi}_k$.  \reviewerTwo{However, the extension to NURBS is straightforward, and Section~\ref{sec:wadgnurbs} briefly describes how to extend the techniques of Section~\ref{sec:wadg} to NURBS approximations.}

\subsection{Approximation properties and optimal knot vectors}

In this work, we restrict ourselves to open knot vectors, and seek interior knot locations which improve approximation properties of the underlying spline space.  It is clear that the span of B-splines contains polynomials of degree $p$ over $[-1,1]$; however, the relationship between approximation properties and interior knot locations is more complex.  This relationship has been analyzed using sup-infs and $n$-widths \cite{kolmogoroffOber1936, melkman1978spline, schumaker2007spline, evans2009n}.  Let $W$ be a normed linear space, let $X \subset W$, and $X_n$ be some $n$-dimensional subspace of $X$; then, the sup-inf $E(X;X_n)$ of $X$ relative to $X_n$ is defined as follows: 
\[
E(X;X_n) =  \sup_{x\in X} \inf_{y\in X_n} \nor{x - y},
\]
and can be interpreted as measuring the worst possible best approximation error in approximating $X$ from $X_n$.  The $n$-width of $X$ is then defined to be the infimum of $E(X;X_n)$ over all subspaces $X_n$ with ${\rm dim}\LRp{X_n} = n$, and a specific subspace $X_n$ is referred to as an \textit{optimal} subspace if its sup-inf attains the $n$-width.

Melkman and Michelli showed in \cite{melkman1978spline} that, for $W = L^2([-1,1])$ and 
\[
X = \LRc{f \in W: \pdn{f}{x}{r} \text{ is absolutely continuous on } [-1,1], \quad \nor{f}_{L^2([-1,1])} \leq 1},
\]
an optimal subspace $X_n$ is the spline space of degree $r$ with knot locations at the roots of $y_{n+1,r}(x)$, where $y_{n+1,r}(x)$ is the $(n+1)$st eigenfunction of the boundary value problem
\begin{equation}
(-1)^r\pdn{y}{x}{2r} = \lambda y(x), \qquad \pdn{y}{x}{k}(-1) = \pdn{y}{x}{k}(1) = 0, \quad k = 0, 1, \ldots, r-1.
\label{eq:nwidtheig}
\end{equation}
We note that, for $r=1$, this reduces down the standard Laplacian eigenvalue problem with Dirichlet boundary conditions, whose eigenfunctions $y_{j,1}(x)$ are 
\[
y_{j,1}(x) = \sin\LRp{j \pi \LRp{\frac{1+x}{2}}}, \qquad j = 1,2,\ldots, \qquad x \in (-1,1).
\]
Thus, the optimal $n$-dimensional space for $r = 1$ consists of piecewise linear splines with $n$ equispaced knots along $[-1,1]$, which are simply piecewise linear $C^0$ polynomials on a mesh with $K = n$ elements.  Similarly, for $r = 2$, (\ref{eq:nwidtheig}) reduces to the biharmonic eigenvalue problem with clamped plate conditions \cite{brenner2007mathematical}.

\begin{figure}
\centering
\subfloat[$r = 2$]{\includegraphics[width=.32\textwidth]{fnwidth_N2_K8.png}}
\hspace{.5em}
\subfloat[$r = 3$]{\includegraphics[width=.32\textwidth]{fnwidth_N3_K8.png}}
\hspace{.5em}
\subfloat[$r = 4$]{\includegraphics[width=.32\textwidth]{fnwidth_N4_K8.png}}
\caption{Eigenfunctions $y_{r,n}(x)$ and their roots for $n = 8$ and various $r$.  All eigenfunctions are approximated using a symmetric Galerkin formulation with degree $p = 8$ spline spaces of maximal continuity and $K = 64$ elements. }
\label{fig:eigoptknots}
\end{figure}

For $r > 2$, optimal knot locations are not explicitly known.  However, following \cite{evans2009n}, we can approximate knot locations using a Galerkin approximation of (\ref{eq:nwidtheig}) with $C^{r-1}$ approximation spaces.  Figure~\ref{fig:eigoptknots} shows approximations of eigenfunctions $y_{n+1,r}(x)$ for various $n,r$ using a fine spline space, with knot locations overlaid.  These knot locations are determined by approximating the roots of eigenfunctions numerically up to a tolerance of $10^{-12}$.  We observe that, as $r$ increases, 
%these knot locations cluster towards the interior relative to equispaced knot locations 
\reviewerOne{the gap between each of the boundary knots and the closest interior knot widens}
due to the presence of the homogeneous boundary conditions on derivatives up to degree $(r-1)$.

\subsection{Knot smoothing}

It is possible to pre-compute and store optimal knot locations for small to moderate $r$ and $n$ for use in practical applications.  However, for large $n$, optimal knot locations become difficult to approximate due to the resolution required to accurately resolve $y_{n+1,r}(x)$.  For large $r$, symmetric Galerkin discretizations of (\ref{eq:nwidtheig}) also become ill-conditioned due to the presence of high order derivatives.  The ill-conditioning of (\ref{eq:nwidtheig}) for $r \gg 1$ may potentially be addressed using a mixed formulation of the eigenvalue problem; however, the construction of stable and convergent mixed methods for higher order elliptic operators is non-trivial \cite{gudi2011interior}. Thus, we seek a heuristic alternative to approximate optimal knot distributions without the computational cost and conditioning issues associated with the discretization and solution of (\ref{eq:nwidtheig}).  We propose an alternative approach to determining a ``good'' knot vector based on an iterative redistribution of knot locations based on the associated Greville abscissae.

The Greville abscissae are a collection of points associated with a B-spline knot vector \cite{johnson2005higher}, and are defined as knot averages 
\[
\tau_j = \frac{1}{p}\sum_{i=1}^p \xi_{i+j-1}, \qquad j = 1,\ldots,p.
\]  
It can also be shown that these points satisfy
\[
\sum_{j=1}^{p+K} \tau_j B^p_j(\widehat{x}) = \widehat{x}.
\]
In other words, the Greville abscissae are the coefficients which reproduce the coordinate $\widehat{x}$ under a one-dimensional B-spline basis.  Because the Greville points $\tau_j$ reproduce $\widehat{x}$, the knot locations $\xi_i$ of a knot vector satisfy 
\[
\xi_i = \sum_{j=1}^{p+K} \tau_j B^p_j(\xi_i).  
\]
For an open knot vector, the distribution of the Greville abscissae clusters near the endpoints of the interval (similarly to stable polynomial interpolation nodes, such as Chebyshev or Gauss-Lobatto points \cite{canuto2012spectral}), as shown in Figure~\ref{fig:grevilleComparison}.

In \cite{hughes2008duality}, the authors introduce a nonlinear parametrization of the interval $[-1,1]$, which is defined implicitly as the mapping which takes Greville abscissae to uniformly spaced points.  Numerical experiments showed that, by using the image of the spline space $V_h$ under this nonlinear mapping in a Galerkin discretization of the Laplacian, the so-called ``outlier frequencies'' of spurious high frequency eigenmodes present in isogeometric discretizations are removed completely.  However, under this mapping, the spline approximation space no longer contains polynomials of degree $p$, and approximation properties are not immediately clear.  Motivated by the results in \cite{hughes2008duality}, we seek instead a heuristic modification of a uniform open knot vector such that the resulting Greville abscissae are closer to equispaced points.  Moreover, we will show that this heuristic modification results in a reasonable approximation of optimal knot locations.   %In contrast, spline spaces defined by smoothed knot vectors are still polynomially complete, and (as demonstrated by the following numerical experiments) appear to produce good approximations of optimal knot locations.

We wish to define a new knot vector $\tilde{\bm{\Xi}}$ such that the resulting Greville abscissae more closely resemble equispaced points.  We can introduce new knot positions $\tilde{\xi}_i$ by defining a linear parametrization where the Greville abscissae $\tau_j$ are \reviewerOne{arbitrarily} replaced with equispaced points $\widehat{x}_i$ along $[-1,1]$
\begin{equation}
\tilde{\xi}_i = \sum_{j=1}^{p+K} \reviewerOne{\widehat{x}_j} B^p_j(\xi_i).  
\label{eq:newKnots}
\end{equation}
\reviewerOne{One can interpret (\ref{eq:newKnots}) as attempting to reverse-engineer a knot vector for which the Greville abscissae are equispaced.  The goal is to produce new knot locations $\tilde{\xi}_i$ for which the Greville abscissae are less clustered towards the endpoints. }  The positions of the first and last knots produced by (\ref{eq:newKnots}) remains the same because the B-spline basis is interpolatory at the endpoints.  However, the locations of interior knots and Greville abscissae are squeezed closer towards the center of the interval, as shown in Figure~\ref{fig:grevilleComparison}.\footnote{We note that one might seek to further redistribute knots to produce Greville abscissae which are equispaced; however, because Greville abscissae are averages of knot positions and because open knot vectors contain repeating knots at endpoints, producing equispaced Greville points by simply redistributing knot positions is impossible without violating the assumption that knot positions $\xi_i$ must be monotonically increasing.}

\begin{figure}
\centering
\includegraphics[width=.5\textwidth]{grevilleComparison.png}
\caption{Comparison of Greville abscissae for an $p=3, K = 8$ spline space when using uniform knots (shown in \textcolor{blue}{blue}) and knot positions $\tilde{\xi}_i$ defined by equation (\ref{eq:newKnots}) (shown in \textcolor{red}{red}).}
\label{fig:grevilleComparison}
\end{figure}

Because the spline basis depends on the knot vector, this process can be repeated 
\begin{equation}
\tilde{\xi}^{k+1}_i = \sum_{j=1}^{p+K} \widehat{x}_j B^p_j(\xi_i;  \tilde{\xi}^k), \qquad \tilde{\xi}^0_i = \xi_i,
\label{eq:knotiter}
\end{equation}
where we have used the notation $B^p_j(\xi_i;  \tilde{\xi}^k)$ to emphasize that the spline basis is evaluated at the original equispaced (i.e. uniform open) knots $\xi_i$, but defined using the knot vector $\tilde{\xi}^k$.\footnote{
We have also tested a knot smoothing iteration based on
\begin{equation}
\tilde{\xi}^{k+1}_i = \sum_{j=1}^{p+K} \widehat{x}_j B^p_j( \tilde{\xi}^k_i;  \tilde{\xi}^k), \qquad \tilde{\xi}^0_i = \xi_i,
\end{equation}
where splines are evaluated at the iterated knots $\tilde{\xi}^k_i$ instead of equispaced knots $\xi_i$.  However, we observe that this iteration produces collapsed knot locations, resulting in spline spaces which are not maximally continuous.  
}
Numerical experiments show that $\tilde{\xi}^k$ converges to an equilibrium distribution as $k\rightarrow \infty$, which can also be determined through the solution of a nonlinear system of equations.  Figure~\ref{fig:smoothedknots} shows the distribution of the knot vectors and Greville abscissae for each iteration $k$, along with optimal knot positions overlaid in red at the final iteration.  Rather than squeezing knots further towards the center, taking $k \rightarrow \infty$ results in knot locations which are close to the optimal knot positions.  
\begin{figure}
\centering
\subfloat[Knot positions]{\includegraphics[width=.4\textwidth]{knots.png}}
\hspace{1em}
\subfloat[Greville abscissae]{\includegraphics[width=.4\textwidth]{greville.png}}
\caption{Knot locations for iterations $k = 0, \ldots, 10$ for a spline space with $p = 3, K = 8$.  Computed optimal knot positions and Greville abscissae are overlaid in \textcolor{red}{red}.}
\label{fig:smoothedknots}
\end{figure}

The results in Figure~\ref{fig:smoothedknots} suggest that this simple knot smoothing produces reasonable approximations to optimal knot positions.  Further numerical experiments show that these approximations do not worsen as $K$ increases.  Let $\widehat{\xi}_i$, $\widehat{\tau}_i$ denote knot locations and Greville abscissae for a uniform open knot vector.  Figure~\ref{fig:smoothedknoterror} shows for $p = 2, 3, 4$ the maximum difference between positions of optimal and smoothed knots and Greville abscissae
\[
\delta_{\rm knot} = \frac{\underset{i}{\max} \LRb{\tilde{\xi}_i - \xi^{\rm opt}_i}}{\underset{i}{\max} \LRb{\widehat{\xi}_i - \xi^{\rm opt}_i}}, \qquad \delta_{\rm Greville} = \frac{\underset{i}{\max} \LRb{\tilde{\tau}_i - \tau^{\rm opt}_i}}{\underset{i}{\max} \LRb{\widehat{\tau}_i - \tau^{\rm opt}_i}}, 
\]
where we have normalized by the maximum difference between optimal and uniform knot and Greville positions.  The smoothed knot positions are computed by iterating (\ref{eq:knotiter}) until the change in the knot vector $\nor{\tilde{\bm{\Xi}}^{k} - \tilde{\bm{\Xi}}^{k-1}}_2$ is less than $10^{-8}$.  We observe that the normalized difference between smoothed and optimal knot positions approaches a fixed value around $.1$ as $K$ increases, with the asymptotic value depending irregularly on $p$.  In contrast, the normalized difference between Greville abscissae for smoothed knots and Greville abscissae for optimal knots approaches a $p$-dependent value as $K$ increases, with this asymptotic value decreasing as $p$ increases.  These results quantify how much more accurately smoothed knot vectors approximate optimal knot distributions compared to uniform open knot vectors.

\begin{figure}
\centering
\subfloat[Normalized knot error $\delta_{\rm knot}$ ]{
\begin{tikzpicture}
\begin{axis}[
width=.425\textwidth,
    xlabel={Number of elements $K$},   
    xmin=2, xmax=66,
    ymax=0,ymax=0.25,    
    legend pos=north east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
]

\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
%coordinates{(4,0.0121)(8,0.0075)(16,0.0046)(32,0.0023)(64,0.0012)};
coordinates{(4,0.210397)(8,0.165792)(16,0.156554)(32,0.15437)(64,0.153806)};

\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
%coordinates{(4,0.009)(8,0.0041)(16,0.0028)(32,0.002)(64,0.0012)};
coordinates{(4,0.0887031)(8,0.0485602)(16,0.0506855)(32,0.0693288)(64,0.0798586)};

\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
%coordinates{(4,0.0016)(8,0.0101)(16,0.0074)(32,0.0048)(64,0.0028)};
coordinates{(4,0.0290577)(8,0.0819356)(16,0.099327)(32,0.115575)(64,0.12577)};

\legend{$p = 2$, $p=3$, $p=4$}
\end{axis}
\end{tikzpicture}
}
\hspace{2em}
\subfloat[Normalized Greville error $\delta_{\rm Greville}$]{
\begin{tikzpicture}
\begin{axis}[
width=.425\textwidth,
    xlabel={Number of elements $K$},   
    xmin=2, xmax=66,
    ymax=0,ymax=0.25,    
    legend pos=north east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
]

\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
%coordinates{(4,0.006)(8,0.0038)(16,0.0024)(32,0.0011)(64,0.00058806)};
coordinates{(4,0.21038)(8,0.100464)(16,0.0851055)(32,0.0805913)(64,0.0788944)};

\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
%coordinates{(4,0.003)(8,0.002)(16,0.0018)(32,0.0012)(64,0.00067857)};
coordinates{(4,0.0887023)(8,0.0365552)(16,0.0390282)(32,0.043861)(64,0.0468188)};

\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
%coordinates{(4,0.000401)(8,0.0025)(16,0.0016)(32,0.00089352)(64,0.00063901)};
coordinates{(4,0.0290725)(8,0.040789)(16,0.027371)(32,0.0237124)(64,0.0301967)};

\legend{$p = 2$, $p=3$, $p=4$}
\end{axis}
\end{tikzpicture}
}
\caption{Normalized difference between the locations of smoothed and optimal knots for $p = 2, 3, 4$ and $K = 4, 8, 16, 32, 64$. }
\label{fig:smoothedknoterror}
\end{figure}

Finally, we note that while the optimal knot positions minimize the $n$-width of spline spaces over the $L^2$ unit ball of $C^{r-1}$ functions, this does not imply that their approximation properties are uniformly better than those of spline spaces with equispaced knots.  Numerical experiments in Section~\ref{sec:approx} illustrate that spline spaces with optimal knot positions are slightly less accurate than spline spaces with uniform knot vectors in approximating non-oscillatory functions, but are more efficient for the approximation of oscillatory functions.  We will also show that optimal knot positions possess additional advantages over uniform open knot vectors in Section~\ref{sec:cfl}, such as a larger maximum stable timestep and improved approximation properties in the presence of non-affine geometric mappings.

\section{Method of lines discretization in time}
\label{sec:mol}

\reviewerTwo{In this section, we introduce the semi-discrete system which results from a finite element discretization of time-dependent variational formulations such as (\ref{eq:varformadvec}), (\ref{eq:varformwave1}), and (\ref{eq:varformwave2}).}  Let $\bm{U},\bm{V}$ denote vectors of degrees of freedom for $U,V$, and define the global mass and finite element matrices $\bm{M}_h, \bm{S}_h$ such that 
\[
\bm{V}^T\bm{M}_h\bm{U} = \LRp{ U,V}_{L^2\LRp{\Omega}}, \qquad \bm{V}^T\bm{S}_h\bm{U} = a(U,V),
\]
where $U,V$ are understood to represent either scalar or group variables whose components lie in $V_h$.  
Under a method of lines discretization, a PDE is transformed to a system of ordinary differential equations (ODEs) 
\begin{equation}
\td{\bm{U}}{t} = \bm{A}_h \bm{U}, \qquad \bm{A}_h = -\bm{M}_h^{-1}\bm{S}_h
\label{eq:mol}
\end{equation}
\reviewerTwo{for first-order (in time) formulations and
\begin{equation}
\frac{d^2\bm{U}}{dt^2} = \bm{A}_h \bm{U}, \qquad \bm{A}_h = -\bm{M}_h^{-1}\bm{S}_h
\label{eq:mol_second}
\end{equation} for second-order formulations} by discretizing in space.  The resulting ODE system can then be solved with explicit time stepping for some step size $\Delta t$.

The following sections discuss efficient strategies for solving (\ref{eq:mol}), focusing on the inversion of $\bm{M}_h^{-1}$ and the choice of timestep $\Delta t$.  Efficient implementations of method of lines solvers typically apply $\bm{A}_h$ in a matrix-free fashion.  However, this requires the computation of $\bm{M}_h^{-1}$ for each application of $\bm{A}_h$.  We discuss in Section~\ref{sec:wadg} the efficient inversion of $\bm{M}_h^{-1}$ using a weight-adjusted approximation to the mass matrix.  Under explicit time stepping, $\Delta t$ must also be chosen such that the eigenvalues of $\Delta t\bm{A}_h$ lie within the stability region of the chosen scheme.  %If an explicit time-stepper is used, the timestep must be sufficiently small to satisfy a stability condition.  
We discuss this condition further in Section~\ref{sec:cfl}.

%The spatial formulations defining $\bm{S}_h$ are constructed such that energy stability holds under inexact quadrature.  As such, the eigenvalues of $\bm{A}_h$ lie in the left half plane

%In order to ensure the eigenvalues of $\bm{A}_h$ lie in the left half plane under inexact quadrature, we adopt a discretization which is \textit{a-priori} energy stable for any arbitrary quadrature with positive weights.  The matrix $\bm{S}_h$ can be written as the sum of skew-symmetric and symmetric parts.   If the symmetric part $\frac{1}{2}\LRp{\bm{S}_h+\bm{S}_h^T}$ is positive semi-definite, all eigenvalues lie in the left half plane and the semi-discrete problem is energy stable \cite{chan2016short}.  We construct a formulation which is stable under inexact numerical integration by splitting $a(u,v)$ into symmetric positive semi-definite and skew-symmetric parts
%\[
%a(u,v) = a_{\rm sym}(u,v) + a_{\rm skew}(u,v), \qquad a_{\rm sym}(u,v) = a_{\rm sym}(v,u), \qquad a_{\rm skew}(u,v) = -a_{\rm skew}(v,u).
%\]
%For advection, integration by parts gives that 
%\[
%a_{\rm sym}(u,v) = \frac{1}{2}\int_{\partial \Omega} \LRb{\bm{\beta}_n} uv\diff{\bm{{x}}}, \qquad {a_{\rm skew}(u,v)} = \frac{1}{2}\int_{\Omega} \LRp{-u\bm{\beta}\cdot \Grad v + \bm{\beta}\cdot \Grad u v}\diff{\bm{{x}}}.
%\]
%Similar approaches have been applied to ensure discrete energy stability for wave propagation problems in a variety of different settings \cite{warburton2013low, chan2015gpu, banks2016galerkin, chan2016weight2, chan2017weight}.

\subsection{Efficient inversion of mass matrices}
\label{sec:wadg}

Recall that $\bm{\Phi}_k: \Dhat \rightarrow D^k$ denotes the mapping from the reference patch $\Dhat$ to the physical patch $D^k$, such that $\bm{\Phi}_k(\widehat{\bm{x}}) = \bm{x} \in D^k$. Then, because solutions are approximated independently over each patch, the global mass matrix $\bm{M}_h$ is block diagonal, with each block corresponding to a patch mass matrix.  The entries of each patch mass matrix are then 
\[
\LRp{\bm{M}}_{ij} =\reviewerTwo{\int_{\Omega} B^p_i(\bm{\Phi}^{-1}_k\LRp{\bm{x}})B^p_j(\bm{\Phi}^{-1}_k\LRp{\bm{x}})\diff{{\bm{x}}}} = \int_{\Dhat} B^p_i(\bm{\widehat{x}})B^p_j(\bm{\widehat{x}}) J \diff{\widehat{\bm{x}}}.  
\]
If $\bm{\Phi}_k$ is affine, then $J$  (the determinant of the Jacobian of the mapping) is constant, and (assuming a tensor product basis) the mass matrix satisfies
\begin{align*}
\bm{M} &= \widehat{\bm{M}} J = J \LRp{\widehat{\bm{M}}_{\rm 1D} \otimes \ldots \otimes \widehat{\bm{M}}_{\rm 1D}}\\
\LRp{\widehat{\bm{M}}_{\rm 1D}}_{ij} &= \int_{-1}^1 B^p_i({\widehat{x}})B^p_j({\widehat{x}}) \diff{\widehat{{x}}}.  
\end{align*}
Thus, for affine transformations, the inverse of the mass matrix possesses a Kronecker product structure
\[
\bm{M}^{-1} = \LRp{\widehat{\bm{M}}J}^{-1}  = \frac{1}{J} \LRp{\widehat{\bm{M}}^{-1}_{\rm 1D} \otimes \ldots \otimes \widehat{\bm{M}}^{-1}_{\rm 1D}},
\]
which greatly reduces the cost of application in higher dimensions.  However, for non-affine mappings, $J$ is not constant, and the Kronecker product structure of the inverse mass matrix is lost.

Several approaches have been proposed to address this difficulty in the inversion of NURBS mass matrices.  Spline collocation methods for explicit dynamics circumvent the construction of weighted $L^2$ mass matrices over each patch \cite{auricchio2012isogeometric}.  These formulations discretize the underlying PDE in its strong form, and require only the inversion of an un-weighted generalized Vandermonde matrix, which always admits a Kronecker product structure.  However, in contrast to methods based the Galerkin framework, it can be difficult to prove energy stability of the resulting discretizations.

Another approach is to replace the exact inversion of the mass matrix with an iterative solver, using a preconditioner possessing a tensor product structure \cite{gao2014fast}.  However, this adds significant complexity and computational cost to solvers based on explicit timestepping, and the convergence of the iteration can depend on the specific geometry.  A more subtle concern associated with preconditioned iterative solvers is energy stability.  The use of an exact mass matrix induces a semi-discrete $L^2$ energy stability, and replacing the mass matrix with a symmetric positive-definite approximation maintains energy stability but changes the norm in which energy is measured.  As noted in \cite{wathen2009chebyshev}, Krylov methods result in a non-linear approximation of the mass matrix inverse, which depends on the initial guess and right hand side vector.  Consequentially, a discrete energy stability no longer holds, as an iterative scheme effectively inverts a slightly different mass matrix each time-step.

A related iterative approach is the combination of lumped spline mass matrices with predictor multicorrector iterations \cite{auricchio2012isogeometric}.  $C^0$ finite element discretizations can employ ``lumped'' diagonal approximations to the mass matrix, giving rise to the spectral element method when lumping is performed through collocation at Gauss-Lobatto quadrature nodes \cite{patera1984spectral}.  %The relationship of collocation nodes to quadrature also provides a way to incorporate curved geometries into spectral element discretizations in an provably energy stable manner \cite{kopriva2006metric, kopriva2010quadrature}.  
However, the extension of mass lumping to NURBS is not straightforward, as collocation nodes (Greville abscissae) can not be used to construct a sufficiently accurate quadrature for spline spaces, and alternative row-sum mass lumping techniques have been observed to reduce accuracy to second order \cite{cottrell2006isogeometric}.  However, high order accuracy can be restored at the cost of additional corrector iterations \cite{auricchio2012isogeometric}.  These predictor-multicorrector iterations were recently shown to be equivalent to a linear operator, thus avoiding problems of non-linearity present for Krylov iterations \cite{evans2017explicit}.

We seek an alternative approach to restoring the Kronecker product structure by replacing the weighted $L^2$ mass matrix with a \textit{weight-adjusted} approximation.  Let $B^p_i(\widehat{\bm{x}})$ denote a basis for the $d$-dimensional spline space over the reference patch.  Then, the mass matrix may be approximated by a weight-adjusted mass matrix
\[
\bm{M} \approx \widehat{\bm{M}} \bm{M}^{-1}_{1/J} \widehat{\bm{M}}, \qquad \reviewerTwo{ \LRp{\bm{M}_{1/J}}_{ij} }= \int_{\Dhat} B^p_j(\widehat{\bm{x}})B^p_i(\widehat{\bm{x}}) \frac{1}{J} \diff{\widehat{\bm{x}}},
\]
such that the inverse of the weight-adjusted mass matrix is 
\begin{align*}
\bm{M}^{-1} \approx \widehat{\bm{M}}^{-1} \bm{M}_{1/J} \widehat{\bm{M}}^{-1}.
\end{align*}
This approximation requires only the application of a weighted mass matrix $\bm{M}_{1/J}$ and the inversion of reference mass matrices $\widehat{\bm{M}}^{-1}$.  
\reviewerTwo{Assuming a fixed geometry, the matrix $\bm{M}_{1/J}$ is constant in time, and can thus be pre-computed and reused at each time-step.  
%Moreover, assuming a Cholesky factorization of $\widehat{\bm{M}} = \bm{L}\bm{L}^T$, the inverse of the reference mass matrix is $\widehat{\bm{M}}^{-1} = \bm{L}^{-T}\bm{L}^{-1}$, and the application of the weight-adjusted mass matrix inverse can be written as 
%\[
%\widehat{\bm{M}}^{-1} \bm{M}_{1/J} \widehat{\bm{M}}^{-1} = \bm{L}^{-T}\bm{L}^{-1}\bm{M}_{1/J}\bm{L}^{-T}\bm{L}^{-1} = \bm{L}^{-T}\bm{L}^{-1}\bm{M}_{1/J}\bm{L}^{-T}\bm{L}^{-1}.
%\]
}
Alternatively, the application of $\bm{M}_{1/J}$ can be performed efficiently in a matrix-free fashion using quadrature and the tensor-product structure of the spline space, while the inversion of $\widehat{\bm{M}}$ can be made efficient by taking advantage of the Kronecker structure present on the reference patch
\[
\widehat{\bm{M}}^{-1} = \widehat{\bm{M}}_{\rm 1D}^{-1} \otimes \ldots  \widehat{\bm{M}}_{\rm 1D}^{-1}.  
\]
The computational cost of this approach can be further reduced by fusing products of reference element matrices to yield an implementation in terms of ``projection'' and ``interpolation'' matrices, as described \cite{chan2016weight1, chan2016weight2}.  
%\reviewerTwo{We note that the relative efficiency of each strategy will likely depend on the choice of computational architecture used.  For example, the low storage cost of the matrix-free application is attractive for accelerator architecture such as Graphics Processing Units, which  \cite{chan2017weight} }

For sufficiently regular $J$, this approximation is high order accurate \cite{chan2016weight1,chan2016weight2}.  Additionally, since the weight-adjusted approximation is symmetric-positive definite, it preserves the energy stability of the underlying Galerkin formulation.  Moreover, the weighed and weight-adjusted mass matrices induce discrete $L^2$-equivalent norms with the same equivalence constants \cite{chan2016weight1}.  As a result, the substitution of the weight adjusted mass matrix does not decrease the maximum stable time-step.

The matrix-free application of the weight-adjusted mass matrix inverse also requires the specification of a sufficiently accurate quadrature.  In this work, we have used a $(p+1)$ point Gaussian quadrature over each knot interval, resulting in composite quadrature rules with a relatively large number of points.  Future work will investigate optimal and reduced quadrature rules tailored to splines \cite{hughes2010efficient, auricchio2012simple, schillinger2014reduced, hiemstra2017optimal}, which greatly reduce the number of quadrature points relative to the number of degrees of freedom in a single patch.  Another method to reduce costs is the use of \reviewerTwo{``weighted quadrature'' methods for applying NURBS mass matrices \cite{calabro2016fast}.  However, we note that energy stability can be lost under this approach, since the curvilinear mass matrices computed using weighted quadrature are not necessarily symmetric}.

Finally, we note that, while the efficient computational implementation of the proposed methods is outside of the scope of this work, different computational implementations may be preferable depending on computational parameters.  For example, the inverse mass matrix may be pre-multiplied into derivative matrices, resulting in dense size $p+K$ operators in each dimension.  This approach may be useful if $K = O(p)$, and is trivially parallelizable down to individual degrees of freedom.  However, for $K \gg p$, the matrices $\bm{S}_i$ and $\bm{M}$ possess a sparse banded structure, which can be exploited to improve computational efficiency on some architectures.  A Cholesky factorization may be used to apply $\bm{M}^{-1}$ in linear complexity with respect to the size of $\bm{M}$, though the constant will depend on $p$ through the matrix bandwidth.  A disadvantage of this approach is that the triangular solve necessary in Cholesky is largely sequential over each patch.\footnote{\reviewerTwo{We note that, in higher dimensions, applying $\widehat{\bm{M}}^{-1}$ in a Kronecker product fashion involves multiple one-dimensional mass matrix inversions.  It is possible to offset the sequential nature of the Cholesky backsolve by parallelizing over each of these one-dimensional mass inversions.  However, this approach also requires significantly more storage, which can be problematic on many-core architectures such as Graphics Processing Units \cite{chan2017weight}.}}

\subsubsection{NURBS basis functions and mass matrices}
\label{sec:wadgnurbs}

\reviewerTwo{The use of rational NURBS basis functions has been popularized in isogeometric analysis \cite{hughes2005isogeometric}.  While the approximate weight-adjusted mass matrix inverse described in Section~\ref{sec:wadg} is formulated for mapped B-spline basis functions, it is straightforward to extend the approach to approximate the inverse of NURBS mass matrices.

NURBS basis functions are rational B-splines, defined as follows 
\begin{equation*}
R^p_i(\widehat{\bm{x}}) = \frac{B^p_i(\widehat{\bm{x}}) w_i}{\sum_{j=1}^{p+K}B^p_j(\widehat{\bm{x}}) w_j} = \frac{B^p_i(\widehat{\bm{x}}) w_i}{w_R(\widehat{\bm{x}})},
\end{equation*}
where we have introduced the rational weight $w_R(\widehat{\bm{x}}) = \sum_{j=1}^{p+K}B^p_j(\widehat{\bm{x}}) w_j$.  We note that the weight $w_R$ varies from patch to patch.  The NURBS mass matrix $\bm{M}_R$ for a curvilinear geometric patch $D^k$ is computed by mapping/pulling back to the reference element
\[
(\bm{M}_R)_{ij} = \int_{\Dhat} R^p_i(\widehat{\bm{x}}) R^p_j(\widehat{\bm{x}}) J\diff{\widehat{\bm{x}}} = \int_{\Dhat} B^p_i(\widehat{\bm{x}}) B^p_j(\widehat{\bm{x}}) \frac{J}{w_R^2}\diff{\widehat{\bm{x}}}.  
\]
Thus, the NURBS mass matrix $\bm{M}_R$ can be interpreted as a weighted B-spline mass matrix with weight $J/w_R^2$, and can be approximated by a weight-adjusted mass matrix inverse 
\[
\bm{M}_R^{-1} \approx \widehat{\bm{M}}^{-1} \bm{M}_{w_R^2/J}  \widehat{\bm{M}}^{-1}, \qquad  \LRp{\bm{M}_{w_R^2/J}}_{ij} = \int_{\Dhat} B^p_i(\widehat{\bm{x}})B^p_j(\widehat{\bm{x}}) \frac{w_R^2}{J}\diff{\widehat{\bm{x}}}.
\]
Approximation results in \cite{bazilevs2006isogeometric, evans2013explicit} imply that the weight-adjusted approximation to the NURBS mass matrix will be high order accurate if $w_R^2/J$ is sufficiently regular.  For example, the weight-adjusted approximation of $\bm{M}_R^{-1}$ is expected to maintain high order accuracy if $J, w_R$ vary smoothly and $w_R$ is bounded away from zero.  
}

%\note{Equivalence of WADG with weighted $L^2$ norm:
%\[
%C_1 \nor{u}_{T^{-1}_{1/w}} \leq \nor{u}_w \leq C_2 \nor{u}_{T^{-1}_{1/w}}. 
%\]
%Can I prove that $C_1 = 1$ and bound $C_2$ based on $w(x)$?
%
%We first note that $C_1, C_2$ are extrema of generalized Rayleigh quotients, or minimum and maximum generalized eigenvalues.  Assume $w(x) \in P^M$, $M < N$, or let integrals be computed with quadrature exact for degree $2N+M$ polynomials (these conditions are equivalent), then $N-M$ eigenvalues are unity - note that for $wu \in P^N$, 
%\[
%\nor{u}_{T^{-1}_{1/w}}^2 = \LRp{T_{1/w}^{-1} u, u} = \LRp{\frac{1}{w}T_{1/w}^{-1} u, wu} = \LRp{u,wu} = \nor{u}_w^2.
%\]
%This implies $u \in P^{N-M}$, from which we can construct $N-M$ weighted-orthogonal polynomials.  These are generalized eigenfunctions.  
%%Max extremal eigenvalues are $\Pi_N\LRp{P_N / w}$ where $P_N$ are $L^2$ orthogonal bases.  
%}

\subsection{Stable timestep restrictions}
\label{sec:cfl}
%Let $\bm{U},\bm{V}$ denote vectors of degrees of freedom for $u,v \in V_h$, and define the mass and advection matrices $\bm{M}, \bm{S}_h$ such that $\bm{V}^T\bm{M}\bm{U} = \int_\Omega u v$ and $\bm{V}^T\bm{S}\bm{U} = a(u,v)$.  Under the method of lines, a PDE is transformed to a system of ordinary differential equations (ODEs) 
%\[
%\td{\bm{U}}{t} = \bm{A}_h \bm{U}, \qquad \bm{A}_h = \bm{M}^{-1}\bm{S}_h
%\]
%by discretizing in space.  This system may then be solved using explicit time stepping with some timestep $\Delta t$, such that the eigenvalues of $\Delta t\bm{A}_h$ lie within the stability region of the chosen scheme.  
%If an explicit time-stepper is used, a stability restriction on the timestep must also be imposed.

\reviewerTwo{For first-order systems,} a rough estimate of the maximum stable timestep is that it is inversely proportional to the spectral radius $\rho(\bm{A}_h)$, which can be estimated by bounding the real and imaginary parts of eigenvalues of $\bm{A}_h$.  The real and imaginary parts of the spectra of $\bm{A}_h$ can be estimated using the symmetric and skew-symmetric parts of $\bm{A}_h$ and theorems from Bendixson and Hirsch \cite{marcus1992survey}, as is done in \cite{chan2015gpu}.  
\[
\max_j \reviewerOne{\LRb{{\rm Re}\LRp{\lambda_j}}} \leq \rho\LRp{\frac{\bm{A}_h+\bm{A}_h^T}{2}}, \qquad \max_j \reviewerOne{\LRb{{\rm Im}\LRp{\lambda_j}}} \leq \rho\LRp{\frac{\bm{A}_h-\bm{A}_h^T}{2}}
\]
%Let $\bm{U},\bm{V}$ denote vectors of degrees of freedom for $u,v \in V_h$.  Since $\bm{A}_h = \bm{M}^{-1}\bm{S}_h$ where $\bm{V}^T\bm{S}\bm{U} = a(u,v)$, 
We estimate the spectral radius of the symmetric part of $\bm{A}_h$ through a generalized Rayleigh quotient
\[
\max_j \reviewerOne{\LRb{{\rm Re}\LRp{\lambda_j}}} \leq \sup_{u \in V_h} \frac{\LRb{a_{\rm sym}(u,u)}}{\nor{u}^2_{\L}}, \qquad a_{\rm sym}(u,v) = \frac{a(u,v) + a(v,u)}{2}.  
\]
The spectral radius of the skew-symmetric part \reviewerTwo{$\rho\LRp{\frac{\bm{A}_h-\bm{A}_h^T}{2}}$} can be bounded using properties of normal matrices \cite{chan2015gpu}; however, it is more straightforward to estimate this term by noting that \reviewerTwo{$\frac{i}{2} \LRp{\bm{A}_h-\bm{A}_h^T}$} is Hermitian and using the generalized Rayleigh quotient.

\reviewerOne{For illustration, we will derive these bounds for scalar advection.}  We define the skew-symmetric part $a_{\rm skew}(u,v)$ to be sesquilinear such that 
\[
a_{\rm skew}(u,v) = \frac{1}{2}\int_{\Omega} \LRp{-u\bm{\beta}\cdot \overline{\Grad v} + \bm{\beta}\cdot \Grad u \overline{v}}\diff{{\bm{x}}}, \qquad a_{\rm skew}(u,vi) = \bm{V}^T\frac{i}{2} \LRp{\bm{A}_h-\bm{A}_h}\bm{U} = a_{\rm skew}(v,ui).
\]
Then, the Bendixson-Hirsch theorems give that
\[
\max_j \reviewerOne{\LRb{{\rm Im}\LRp{\lambda_j}}} \leq \sup_{u \in V_h} \frac{\LRb{a_{\rm skew}(u,ui)}}{\nor{u}^2_{\L}}, \qquad \LRb{a_{\rm skew}(u,ui)} = \int_{\Omega} \LRp{u\bm{\beta}\cdot \Grad u}\diff{{\bm{x}}} \leq \nor{\bm{\beta}}_{L^\infty}\nor{u}_{\L}\nor{ \Grad u}_{\L}.
\]
%For quasi-uniform patches, 
%\[
%\nor{ \Grad u}_{\L} \leq \nor{\bm{G}}_{L^{\infty}} \nor{\widehat{\Grad}u}_{\L} \leq C \frac{\nor{\widehat{\Grad}u}_{\L}}{H} \leq C\nor{\widehat{\Grad}u}_{L^2\LRp{\widehat{D}}} \leq C C_I \nor{u}_{L^2\LRp{\widehat{D}}} \leq C C_I \nor{u}_{\L}.
%\]
Mapping the formulation for advection to and from the reference patch $\Dhat$ and applying inverse and trace inequalities then yields
\begin{align*}
\LRb{a_{\rm sym}(u,u)} 
&\leq \max_k \LRp{\frac{\nor{\bm{\beta}}_{L^\infty}}{2} C_T \nor{J^s}_{L^\infty}\nor{J^{-1}}_{L^\infty} \nor{u}^2_{L^2\LRp{D^k}}} \\
\LRb{a_{\rm skew}(u,u)} &\leq \max_k \LRp{{\nor{\bm{\beta}}_{L^\infty}} C_I \nor{J\bm{G}}_{L^\infty}\nor{J^{-1}}_{L^\infty} \nor{u}^2_{L^2\LRp{D^k}}}.
\end{align*}
These bounds together imply that 
\begin{align*}
\rho\LRp{\bm{A}_h} &\leq \max_k \LRp{\nor{\bm{\beta}}_{L^\infty}\max\LRc{\frac{1}{2} C_T \nor{J^s}_{L^\infty}, C_I \nor{J\bm{G}}_{L^\infty}}\nor{J^{-1}}_{L^\infty}}.
%\max_j \reviewerOne{\LRb{{\rm Im}\LRp{\lambda_j}}} &\leq \frac{\nor{\bm{\beta}}_{L^\infty}}{2} .
\end{align*}

In one dimension, on a uniform domain subdivided into patches of size \reviewerOne{$H$}, this reduces to the expression
\[
\rho\LRp{\bm{A}_h} \leq \max_k  \LRp{\frac{\nor{\bm{\beta}}_{L^\infty}}{\reviewerOne{H}} C_p}, \qquad \qquad \reviewerOne{C_p = \max\LRc{\frac{1}{2} C_T, C_I}},
\]
which can be used to determine a stable timestep by setting $\Delta t = 1/\rho\LRp{\bm{A}_h}$.

For the acoustic wave equation, a similar argument can be applied to bound the spectral radius via
\[
\rho\LRp{\bm{A}_h} \leq \max_k\LRp{ \frac{\nor{c}_{L^\infty}}{\reviewerOne{H}} C_p},
\]
where $c$ is the wavespeed.  Setting $\Delta t = O(1/\rho\LRp{\bm{A}_h})$ then yields an estimate of the maximum stable timestep 
\[
\Delta t \leq C_t\frac{\reviewerOne{H}}{C_p\nor{c}_{L^\infty} },
\]
\reviewerOne{where $C_t$ is some constant which depends on the stability region of the explicit time-stepper.}
Thus, for high order finite element discretizations, the stable timestep is determined not only by the \reviewerOne{patch} size $\reviewerOne{H}$ and maximum wavespeed ($\nor{\bm{\beta}}_{L^{\infty}}$ or $\nor{c}_{L^\infty}$, but also by trace and inverse inequality constants $C_T, C_I$ \reviewerOne{which depend on the number of elements $K$ per patch and the order $p$}.

\reviewerTwo{For second-order systems, a rough estimate of the maximum stable timestep is that it is inversely proportional to the square root of the spectral radius $\rho(\bm{A}_h)$.  For the second-order order formulation of the acoustic wave equation, the spectral radius $\rho(\bm{A}_h)$ is equivalent to the Rayleigh quotient
\begin{align*}
\rho(\bm{A}_h) = \sup_{u \in V_h} \frac{a(u,u)}{\nor{u}^2_{\L}}
\end{align*}
provided the constant $\tau$ is sufficiently large to guarantee positive-definiteness of $\bm{A}_h$.  If $\tau$ scales linearly with the trace constant $C_T$, then an application of inverse and trace inequalities yields
\begin{align*}
\rho(\bm{A}_h) \leq C_{\tau} \frac{\left( C_T^2 + C_I^2 \right) \nor{c}^2_{L^\infty}}{H^2}
\end{align*}
where $C_{\tau}$ is a constant of proportionality between $\tau$ and $C_T$.  Thus, we arrive at a maximum stable timestep estimate of
\[
\Delta t \leq C_t C_\tau\frac{H}{ \sqrt{C_T^2 + C_I^2} \nor{c}_{L^\infty} },
\]
where again $C_t$ is some constant which depends on the stability region of the explicit time-stepper.}

\subsubsection{Constants in trace and inverse inequalities for spline spaces}

It is possible to derive explicit formulas for the constants in trace inequalities for polynomials \cite{warburton2003constants, hillewaert2011sharp, chan2015hp}.  While one can repeat this process for inverse inequalities or spline spaces, it is more difficult to derive tight, explicit expressions for such constants in terms of $p$.  However, because these constants are given as extrema of generalized Rayleigh quotients, one can also compute constants in trace or inverse inequality through the solution of a generalized eigenvalue problem over the reference patch.  In one dimension, these problems are given as
\begin{equation}
\widehat{\bm{M}}^fu = \lambda_T\widehat{\bm{M}}u, \qquad \widehat{\bm{K}}u = \lambda_I\widehat{\bm{M}}u,
\label{eq:eigconst}
\end{equation}
where $\widehat{\bm{M}}^f$ is the reference boundary mass matrix
\[
\LRp{\widehat{\bm{M}}^f}_{ij} = B^p_i(-1) B^p_j(-1) + B^p_i(1) B^p_j(1),
\]
and $\widehat{\bm{K}}$ is the reference stiffness matrix
\[
\LRp{\widehat{\bm{K}}}_{ij} = \int_{-1}^1 \pd{B^p_i}{\widehat{x}}\pd{B^p_j}{\widehat{x}} \diff{\widehat{x}}.
\]
{\color{forestgreen} Then, the constant $C_T$ is the largest generalized eigenvalue of the boundary mass matrix, and the constant $C_I$ is the square root of the largest generalized eigenvalue of the stiffness matrix.}

Assuming a tensor product construction of $V_h$, constants in higher dimensional trace inequalities can be determined in terms of the constants in one-dimensional trace inequalities.  For example, in two dimensions, the mass, face mass, and stiffness matrices are given as
\begin{align*}
\widehat{\bm{M}} &= \widehat{\bm{M}}_{\rm 1D}\otimes \widehat{\bm{M}}_{\rm 1D}, \qquad \widehat{\bm{M}}^f = \widehat{\bm{M}}_{\rm 1D}\otimes \widehat{\bm{M}}^f_{\rm 1D}\\
\widehat{\bm{K}} &= \widehat{\bm{M}}_{\rm 1D}\otimes \widehat{\bm{K}}_{\rm 1D} + \widehat{\bm{K}}_{\rm 1D} \otimes \widehat{\bm{M}}_{\rm 1D}.
\end{align*}
Substituting these relations into the generalized eigenvalue problem (\ref{eq:eigconst}) for the trace constant and reducing gives
\[
\bm{I}\otimes \widehat{\bm{M}}^f_{\rm 1D} u = \lambda_T\LRp{\bm{I}\otimes \widehat{\bm{M}}}u.  
\]
Because the spectra of $\bm{I}\otimes \widehat{\bm{M}}^f_{\rm 1D}$ consists only of multiples of one-dimensional eigenvalues $\lambda_T$, constants in higher dimensional trace inequalities are identical to the constants in one spatial dimension.

In comparison, constants in $d$-dimensional inverse inequalities scale with $\sqrt{d}$, which can be seen by diagonalizing (\ref{eq:eigconst}).  Let $\bm{W}_{\rm 1D}$ denote the coefficient matrix for eigenvectors of (\ref{eq:eigconst}) in one dimension, such that $\bm{W}_{\rm 1D}^* \widehat{\bm{K}}_{\rm 1D} \bm{W}_{\rm 1D} = \bm{\Lambda}_{\rm 1D}$ and $\bm{W}_{\rm 1D}^* \widehat{\bm{M}}_{\rm 1D} \bm{W}_{\rm 1D} = \bm{I}$.  Then, the change of basis $\bm{W}_{\rm 1D}\otimes \bm{W}_{\rm 1D}$ reduces the eigenvalue problem for constants in the two-dimensional inverse inequality to 
\[
\LRp{\bm{I}_{\rm 1D}\otimes \bm{\Lambda}_{\rm 1D} + \bm{\Lambda}_{\rm 1D}\otimes\bm{I}_{\rm 1D}} u = \lambda u,
\]
Let $\mu_j$ denote one-dimensional generalized eigenvalues for the stiffness matrix, and let the two dimensional eigenvalues be indexed by $ij$.  Then, $\lambda_{ij} = \mu_i + \mu_j$ and $\max_{ij} \lambda_{ij} = 2\max_j \mu_j$.  Assuming a indexing by $i_1, \ldots, i_d$ in $d$ dimensions, the result generalizes to higher dimensions as $\max_{i_1,\ldots,i_d} \lambda_{i_1,\ldots,i_d} = d\max_j \mu_j$.

For polynomials, it can be shown that the constants $C_I,C_T$ depend quadratically on the degree of approximation $p$ (see, for example \cite{warburton2003constants, ozisik2010constants,hillewaert2011sharp}).  For standard $C^0$-continuous {\color{forestgreen} and discontinuous Galerkin discretizations} of the advection equation {\color{forestgreen} using a polynomial approximation over each patch}, it follows that $\rho(\bm{A}_h) = O(p^2 / h)$ {\color{forestgreen} since $H = h$}.  {\color{forestgreen} In contrast, for spline spaces of maximal smoothness, numerical experiments in the following sections and analytic results for special cases \cite{takacs2016approximation} suggest that the constants $C_I, C_T$ grow as $O(pK)$ when the number elements $K$ in a patch is sufficiently large with respect to $p$, and thus $\rho(\bm{A}_h)$ grows as $O(p / h)$.}  In the special case of periodic splines, the stronger result can be shown that $C_I, C_T$ are independent of $p$ \cite{takacs2016approximation}.

\subsubsection{Numerically computed constants}
\label{sec:numconsts}

We consider spline bases with uniform, optimal, and smoothed knot vectors, though we limit our experiments using optimal knot vectors to $p \leq 5$ due to issues of numerical stability \reviewerOne{in the computation of optimal knot locations}.  For each basis, we compute constants $C_I,C_T$ in trace and inverse inequalities through the solution of a generalized eigenvalue problem.  The number of elements is increased with $p$ as $K = \lceil p/2 \rceil, p, 2p$.  %We consider two scenarios for the number of elements $K$: in the first, we fix the number of elements $K = 4, 8, 16$ while increasing the degree $p$.  In the second scenario, we increase the number of elements with $p$ as $K = \lceil p/2 \rceil, p, 2p$.

Figures~\ref{fig:iconsts} and \ref{fig:tconsts} show the growth of $C_I /K, C_T/K$ for one dimensional spline bases, where we have scaled the computed trace and inverse inequality constants by $1/K$ to isolate the growth of $C_I, C_T$ with respect to $p$.  We observe that, for $K \geq p$, the growth of $C_I, C_T$ for all spline spaces is roughly linear in {\color{forestgreen} both} $p$ {\color{forestgreen} and $K$}.  Additionally, while knot smoothing does not decrease the asymptotic rate of growth of $C_I, C_T$, we observe that it results in noticeable reductions in each constant.  Moreover, the reduction in $C_I, C_T$ from knot smoothing increases the larger $K$ is relative to $p$.

For a uniform open knot vector, the distribution of the Greville abscissae clusters near the endpoints of the interval.  Because knot smoothing reduces the clustering of Greville abscissae near the boundaries, the reduction in $C_I, C_T$ is to be expected.  For example, the clustering of stable interpolation points has been identified as contributing to the $O(1/p^2)$ CFL condition in high order pseudospectral methods: for Gauss-Legendre-Lobatto or Chebyshev nodes, the minimum spacing between nodes decreases as $O(1/p^2)$.  Kosloff and Tal-Ezer proposed a mapping of the spatial domain \cite{kosloff1993modified} to address this timestep limitation, and showed that the mapping improved the CFL condition from $O(1/p^2)$ to $O(1/p)$.  However, this mapping also reduces accuracy (especially at lower degrees of approximation \cite{atcheson2013explicit}).  \reviewerOne{By contrast, knot smoothing does not significantly decrease accuracy.  In fact, the approximation power of spline spaces under knot smoothing is actually improved for certain functions.    }

\begin{figure}
\centering
\subfloat[$K = \lceil \frac{p}{2}\rceil$]{
\begin{tikzpicture}
\begin{axis}[
width=.35\textwidth,
    xlabel={Degree of approximation $p$},   
        ylabel={$C_I / K$},
    xmin=.5, xmax=12.5,
    ymax=0,ymax=90,        
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
    xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
]

\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
coordinates{(1,2)(2,4.5)(3,8)(4,12.5)(5,18)(6,24.5)(7,32)(8,40.5)(9,50)(10,60.5)(11,72)(12,84.5)};
\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
coordinates{(1,3)(2,6)(3,6.5)(4,9.5)(5,11.0172)(6,14.178)(7,16.171)(8,19.4951)(9,21.7992)(10,25.2711)(11,27.8076)(12,31.4113)};
\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
coordinates{(1,3)(2,6)(3,6.5)(4,9.5)(5,10.5251)(6,13.6497)(7,14.7755)(8,17.9829)(9,19.1754)(10,22.4404)(11,23.6889)(12,26.9988)};

\legend{Polynomial $C_I$, Uniform $C_I/K$, Smoothed $C_I/K$ }
\end{axis}
\end{tikzpicture}
}
\subfloat[$K = p$]{
\begin{tikzpicture}
\begin{axis}[
width=.35\textwidth,
    xlabel={Degree of approximation $p$},   
%        ylabel={$C_I / K$},
    xmin=.5, xmax=12.5,
    ymax=0,ymax=90,    
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
    xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
]

\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
coordinates{(1,2)(2,4.5)(3,8)(4,12.5)(5,18)(6,24.5)(7,32)(8,40.5)(9,50)(10,60.5)(11,72)(12,84.5)};
\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
coordinates{(1,3)(2,4)(3,5.70747)(4,7.72711)(5,9.95857)(6,12.3561)(7,14.8924)(8,17.5488)(9,20.3115)(10,23.1696)(11,26.1146)(12,29.1391)};
\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
coordinates{(1,3)(2,4)(3,5.34993)(4,6.8709)(5,8.45922)(6,10.0982)(7,11.7767)(8,13.4862)(9,15.2211)(10,16.9775)(11,18.7525)(12,20.5439)};

\legend{Polynomial $C_I$, Uniform $C_I/K$, Smoothed $C_I/K$ }
\end{axis}
\end{tikzpicture}
}
\subfloat[$K = 2p$]{
\begin{tikzpicture}
\begin{axis}[
width=.35\textwidth,
    xlabel={Degree of approximation $p$},   
%    ylabel={$C_I / K$},
    xmin=.5, xmax=12.5,
    ymax=0,ymax=90,    
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
    xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
]

\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
coordinates{(1,2)(2,4.5)(3,8)(4,12.5)(5,18)(6,24.5)(7,32)(8,40.5)(9,50)(10,60.5)(11,72)(12,84.5)};
\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
coordinates{(1,2)(2,3.42857)(3,5.27022)(4,7.3491)(5,9.61122)(6,12.0258)(7,14.5716)(8,17.233)(9,19.9977)(10,22.8559)(11,25.7995)(12,28.8217)};
\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
coordinates{(1,2)(2,3.15356)(3,4.41024)(4,5.72791)(5,7.09105)(6,8.48919)(7,9.91299)(8,11.3583)(9,12.8221)(10,14.3019)(11,15.796)(12,17.3028)};

\legend{Polynomial $C_I$, Uniform $C_I/K$, Smoothed $C_I/K$ }
\end{axis}
\end{tikzpicture}
}
\caption{Scaled constants in inverse inequalities for splines with various choices of $K$, along with corresponding constants for polynomials. }
\label{fig:iconsts}
\end{figure}

%% trace inequalities
\begin{figure}
\centering
\subfloat[$K = \lceil \frac{p}{2}\rceil$]{
\begin{tikzpicture}
\begin{axis}[
width=.35\textwidth,
    xlabel={Degree of approximation $p$},   
        ylabel={$C_T / K$},
    xmin=.5, xmax=12.5,
    ymax=0,ymax=65,        
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
    xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
]

\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
coordinates{(1,1.73205)(2,3.87298)(3,6.5216)(4,9.74981)(5,13.5914)(6,18.0596)(7,23.1598)(8,28.894)(9,35.2634)(10,42.2685)(11,49.9096)(12,58.187)};
\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
coordinates{(1,1.73205)(2,3.87298)(3,4.37379)(4,6.25368)(5,7.1462)(6,9.14314)(7,10.3601)(8,12.4668)(9,13.8967)(10,16.0995)(11,17.6856)(12,19.973)};
\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
coordinates{(1,1.73205)(2,3.87298)(3,4.37379)(4,6.25368)(5,6.85572)(6,8.8263)(7,9.51337)(8,11.5444)(9,12.2862)(10,14.3577)(11,15.1409)(12,17.2432)};

\legend{Polynomial $C_T$, Uniform $C_T/K$, Smoothed $C_T/K$ }
\end{axis}
\end{tikzpicture}
}
\subfloat[$K = p$]{
\begin{tikzpicture}
\begin{axis}[
width=.35\textwidth,
    xlabel={Degree of approximation $p$},   
    xmin=.5, xmax=12.5,
    ymax=0,ymax=65,    
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
    xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
]

\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
coordinates{(1,1.73205)(2,3.87298)(3,6.5216)(4,9.74981)(5,13.5914)(6,18.0596)(7,23.1598)(8,28.894)(9,35.2634)(10,42.2685)(11,49.9096)(12,58.187)};
\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
coordinates{(1,1.73205)(2,2.8364)(3,3.82963)(4,5.04001)(5,6.41327)(6,7.90642)(7,9.49532)(8,11.165)(9,12.9051)(10,14.7078)(11,16.5672)(12,18.4783)};
\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
coordinates{(1,1.73205)(2,2.8364)(3,3.6403)(4,4.54195)(5,5.51171)(6,6.52839)(7,7.57837)(8,8.65295)(9,9.74686)(10,10.8566)(11,11.9797)(12,13.1144)};

\legend{Polynomial $C_T$, Uniform $C_T/K$, Smoothed $C_T/K$ }
\end{axis}
\end{tikzpicture}
}
\subfloat[$K = 2p$]{
\begin{tikzpicture}
\begin{axis}[
width=.35\textwidth,
    xlabel={Degree of approximation $p$},   
    xmin=.5, xmax=12.5,
    ymax=0,ymax=65,    
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
    xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
]

\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
coordinates{(1,1.73205)(2,3.87298)(3,6.5216)(4,9.74981)(5,13.5914)(6,18.0596)(7,23.1598)(8,28.894)(9,35.2634)(10,42.2685)(11,49.9096)(12,58.187)};
\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
coordinates{(1,1.73205)(2,2.43668)(3,3.48657)(4,4.75661)(5,6.16327)(6,7.6751)(7,9.27477)(8,10.9506)(9,12.694)(10,14.4983)(11,16.3578)(12,18.2682)};
\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
coordinates{(1,1.73205)(2,2.32019)(3,2.9916)(4,3.77273)(5,4.61001)(6,5.4806)(7,6.37331)(8,7.28317)(9,8.20701)(10,9.14265)(11,10.0885)(12,11.0432)};

\legend{Polynomial $C_T$, Uniform $C_T/K$, Smoothed $C_T/K$ }
\end{axis}
\end{tikzpicture}
}
\caption{Scaled constants in trace inequalities for splines with various choices of $K$, along with corresponding constants for polynomials. }
\label{fig:tconsts}
\end{figure}

Finally, we examine the difference between $C_I, C_T$ for smoothed and optimal knot vectors.  Table~\ref{tab:consts} shows constants $C_I, C_T$ in inverse and trace inequalities for smoothed and optimal spline spaces of degree $p = 2,\ldots, 5$ with $K = p, 2p$ elements.  We observe that the values of $C_I, C_T$ do not differ greatly between smoothed and optimal knot vectors.

\begin{table}
\centering
\subfloat[Scaled trace constants \reviewerOne{$C_T/K$}]{
\begin{tabular}{|c||c|c|c|c|c|}
\hline
 & $p = 2$ & $p = 3$ & $p = 4$ & $p = 5$ \\
\hhline{|=|=|=|=|=|=|}
%$K = p$ (uniform knots) &  2.8364  &  3.8296&    5.0400&    6.4133\\
\hline
$K = p$ (smoothed knots) & 2.8364 &   3.6403&    4.5420&    5.5117\\
\hline
$K = p$ (optimal knots) & 2.8364  &  3.6604&    4.5303&    5.4323\\
\hhline{|=|=|=|=|=|=|}
%$K = 2p$ (uniform knots) &  2.4367    &3.4866    &4.7566  &  6.1633\\
\hline
$K = 2p$ (smoothed knots) &  2.3202   & 2.9916&    3.7727&    4.6100\\
\hline
$K = 2p$  (optimal knots)  & 2.3215   & 3.0011   & 3.7515  &  4.5251\\
\hline
\end{tabular}
}\\
\subfloat[Scaled inverse inequality constants \reviewerOne{$C_I/K$}]{
\begin{tabular}{|c||c|c|c|c|c|}
\hline
 & $p = 2$ & $p = 3$ & $p = 4$ & $p = 5$ \\
\hhline{|=|=|=|=|=|=|}
%$K = p$ (uniform knots) & 4.0000   & 5.7075&    7.7271&    9.9586\\
\hline
$K = p$ (smoothed knots) &     4.0000    &5.3499    &6.8709&    8.4592\\
\hline
$K = p$ (optimal knots) & 4.0000    &5.3907&    6.8514&    8.3260\\
\hhline{|=|=|=|=|=|=|}
%$K = 2p$ (uniform knots) & 3.4286    &5.2702    &7.3491    &9.6112 \\
\hline
$K = 2p$ (smoothed knots) & 3.1536   & 4.4102&    5.7279&    7.0910\\
\hline
$K = 2p$ (optimal knots) & 3.1768   & 4.4311   & 5.6895   & 6.9474\\
\hline
\end{tabular}
}
\caption{Comparison of constants in spline trace and inverse inequalities using smoothed and optimal knot vectors.}
\label{tab:consts}
\end{table}

\subsection{On the CFL restriction for piecewise polynomial and spline finite elements: $h$ vs $p$ methods}

\reviewerOne{As mentioned previously, $C^0$ and discontinuous finite element methods result in a stable timestep restriction of $O(h/p^2)$ for a given $h$ and $p$.  Assuming an {$O(pK)$} scaling of inverse and trace constants $C_I, C_T$ for spline approximations, spline discretizations yield a stable time-step restriction of $O(h/p)$.  This $O(h/p)$ stable timestep restriction for splines applies when interpreting spline finite elements as $h$-methods, where $h, K$ are assumed to be independent of $p$ with $K \gg p$.

However, when interpreting spline discretizations as $p$-methods \cite{babuska1981p}, resolution is achieved by increasing the order rather than by mesh refinement, and the assumption $K \gg p$ may not be appropriate.  Consider { then the case of} a single patch of size $H=O(1)$ { with $K = O(p)$.  Since $h = O(1/K)$ and $C_I, C_T$ scale like $O(pK)$ when $K = O(p)$}, we have that the stable timestep restriction for splines on a single patch is $O(h/p) = O(1/p^2)$.  This analysis shows {spline discretizations with $K = O(p)$ exhibit stable timestep sizes with the same asymptotic dependence on $p$ as those of $C^0$ and discontinuous finite element methods}.  However, we note that this does not account for the fact that for the same stable timestep, spline discretizations achieve an extra factor $O(1/K) = O(1/p)$ of additional mesh resolution compared to a polynomial discretization.  Thus splines can achieve the same stable timestep restriction as $C^0$ finite elements while providing a higher resolution.

An alternative way to compare the stable timestep restriction for splines and $C^0$ finite elements is in terms of degrees of freedom.  Up to this point, our discussion has related the time-step size to the mesh size $h$ and the polynomial degree $p$, but for the same mesh size $h$ and polynomial degree $p$, a piecewise polynomial finite element basis has a larger number of degrees of freedom $N_{\rm dofs}$ than a corresponding spline basis.  Thus, it may seem appropriate instead to examine the maximum stable timestep in terms of the number of degrees of freedom.  In this direction, we note that $N_{\rm dofs} \sim p/h$ for $C^0$ or discontinuous piecewise polynomial finite elements in the one-dimensional setting, so the maximum stable timestep scales like $1/(pN_{\rm dofs})$.  For one-dimensional spline finite elements, $N_{\rm dofs} \sim 1/h$, so the maximum stable timestep also scales like $1/(pN_{\rm dofs})$.  This analysis seems to suggest spline finite elements do not exhibit any improvement over $C^0$ or discontinuous piecewise polynomial finite elements, at least with respect to timestep size.  However, for the same polynomial degree $p$ and number of degrees of freedom $N_{\rm dofs}$, a spline basis is generally more accurate than a piecewise polynomial finite element basis \cite{evans2009n}.  Thus, to attain the same level of accuracy for a fixed $p$, one can utilize a larger timestep size with splines than with $C^0$ or discontinuous finite elements.  %In fact, the same level of accuracy is typically attained when the mesh sizes are comparable, so an $O(p)$ times larger timestep may be employed for splines versus $C^0$ or discontinuous finite elements.  
%To clarify the behavior of spline finite element methods under $p$-refinement, we can interpret the stable time step restriction in terms of the number of degrees of freedom by examining the product of the the maximum stable timestep $\Delta t_{\max}$ and the number of degrees of freedom $N_{\rm dofs}$.  The product $\Delta t_{\max} N_{\rm dofs}$ is a measure of the efficiency of a method for a fixed mesh resolution $h$ and order $p$.  A larger value of $\Delta t_{\max} N_{\rm dofs}$ implies either that a larger time-step can be taken for a fixed number of degrees of freedom, or that a larger number of degrees of freedom (higher resolution) can be used without decreasing the maximum stable time-step.  Additionally, since spline and high order polynomial approximations provide approximately the same resolution for the same number of degrees of freedom on a curvilinear mesh (based on numerical experiments in Section~\ref{sec:twodimexp}), this product provides an alternative way to compare the two methods.  
%
%We now consider a degree $p$ approximation with $K$ elements or micro-elements in one spatial dimension.  For piecewise polynomial finite elements, $N_{\rm dofs} = O\LRp{K(p+1)}$, while  $\Delta t_{\max} = O(h/p^2) = O(1/(Kp^2))$.  Thus, the product $\Delta t_{\max} N_{\rm dofs}$ results in the  $K$-independent quantity
%\[
%\Delta t_{\max} N_{\rm dofs} = O\LRp{\frac{1}{p}}.    
%\]
%In comparison, the number of degrees of freedom for a spline space of degree $p$ and $K$ micro-elements is $p+K$.  If the number of elements $K$ is related to the order as $K = L p$ for some fixed integer $L = O(1)$ (as in Figures~\ref{fig:iconsts} and \ref{fig:tconsts}), then the number of degrees of freedom for a spline space is $O((L+1)p)$.  For spline finite element methods with the number of elements $K = Lp$, the scaling of the inverse and trace inequality constants suggest that $\Delta t_{\max} = O(h/p) = O(1/(Kp)) = O(1/(Lp^2))$.  Then, $N_{\rm dofs} = p+K = (L+1)p$, and the product $\Delta t_{\max} N_{\rm dofs}$ is
%\[
%\Delta t_{\max} N_{\rm dofs} = O\LRp{\frac{(L+1)}{L}\frac{1}{p}}.  
%\]
%This suggests that, while splines may not improve the asymptotic scaling of $\Delta t_{\max}$ as a $p$-method, the time-step is improved by some constant for $L= O(1)$.  We note that this analysis does not take into account the work per degree of freedom, which varies depending on the implementation for splines and piecewise polynomial methods.  A detailed computational comparison is beyond the scope of this paper, and will be investigated in future work.  
}

\section{Numerical experiments}
\label{sec:num}

In the following sections, we present numerical experiments which verify properties of the proposed approximation spaces and discretizations.  All computations use Gaussian quadratures of degree $p$ over each knot interval, and all time-dependent simulations use a 4th order 5-stage low-storage Runge-Kutta method for time integration \cite{carpenter1994fourth} (with the second order formulation rewritten as a first order system).  All experiments use first or second order multi-patch DG formulations for the acoustic wave equation, except for the computation of discrete dispersion relations (which use the multi-patch DG formulation for the advection equation).

\subsection{One dimensional experiments}

First, we compute convergence rates of $L^2$ errors under mesh refinement.  We begin with the acoustic wave equation in 1D, using the exact solution $p(x,t) = \cos\LRp{\frac{3\pi x}{2}}\cos\LRp{\frac{3\pi t}{2}}$ evaluated at final time $T=1/2$..  Figure~\ref{fig:hconvergence1D} shows the convergence of one-dimensional multi-patch discretizations.  Optimal $O(h^{p+1})$ rates of convergence are observed under both mesh refinement (knot insertion) and patch refinement (subdivision of patches) for first and second order formulations.  We also observe that, while optimal rates are still observed, $L^2$ errors for knot insertion are larger than $L^2$ errors for patch refinement, and that this effect grows as the degree $p$ increases.  However, we will demonstrate that, as suggested in the previous section, spline spaces achieve a lower $L^2$ error per degree of freedom.  We also observe that $L^2$ errors are larger when using smoothed knots, though (as indicated by previous experiments) this may be due to the fact that the solution is relatively smooth and non-oscillatory.

\begin{figure}
\centering
\subfloat[First order formulation]{
\begin{tikzpicture}
\begin{loglogaxis}[
width=.45\textwidth,
    xlabel={Mesh size $h$},
    ylabel={$L^2$ error},
    xmin=1e-2, xmax=.3,
    ymin=1e-10, ymax=.1,    
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
    legend entries={$p = 2$, $p=3$, $p=4$},
] 
\pgfplotsset{
cycle list={
{blue, mark=*}, {red,mark=square*},
{blue, mark=*,dashed}, {red,mark=square*,dashed},
{blue, mark=*,dotted}, {red,mark=square*,dotted},
legend style={mark options={no markers}}
}}
\addlegendimage{no markers,black}
\addlegendimage{no markers,black,dashed}
\addlegendimage{no markers,black,dotted}

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.0106673)(0.125,0.000954815)(0.0625,0.000110851)(0.03125,1.37364e-05)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.00629266)(0.125,0.000476247)(0.0625,7.45318e-05)(0.03125,1.11051e-05)};
\logLogSlopeTriangleFlip{.5}{.125}{.65}{3}{}

    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).

%    4.0252
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.00183485)(0.125,8.58892e-05)(0.0625,4.97075e-06)(0.03125,3.0529e-07)};
%    3.8576
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.000679534)(0.125,3.39478e-05)(0.0625,2.43527e-06)(0.03125,1.67989e-07)};
\logLogSlopeTriangleFlip{.5}{.125}{.475}{4}{}

%    5.0305
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.000412175)(0.125,1.05591e-05)(0.0625,2.98645e-07)(0.03125,9.13748e-09)};
%    4.8741
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,7.04497e-05)(0.125,1.74612e-06)(0.0625,6.0791e-08)(0.03125,2.07289e-09)};
\logLogSlopeTriangle{.5}{.15}{.125}{5}{}

\end{loglogaxis}
\end{tikzpicture}
}
\subfloat[Second order formulation]{
\begin{tikzpicture}
\begin{loglogaxis}[
width=.45\textwidth,
    xlabel={Mesh size $h$}, 
    ylabel={$L^2$ error},
    xmin=1e-2, xmax=.3,
    ymin=1e-10, ymax=.1,    
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
    legend entries={$p = 2$, $p=3$, $p=4$},
] 
\pgfplotsset{
cycle list={
{blue, mark=*}, {red,mark=square*},
{blue, mark=*,dashed}, {red,mark=square*,dashed},
{blue, mark=*,dotted}, {red,mark=square*,dotted},
legend style={mark options={no markers}}
}}
\addlegendimage{no markers,black}
\addlegendimage{no markers,black,dashed}
\addlegendimage{no markers,black,dotted}

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.0150916)(0.125,0.00135624)(0.0625,0.000126749)(0.03125,1.27992e-05)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.00727599)(0.125,0.000820116)(0.0625,9.52161e-05)(0.03125,1.14411e-05)};
\logLogSlopeTriangleFlip{.5}{.125}{.65}{3}{}

    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.00274207)(0.125,8.37621e-05)(0.0625,4.79288e-06)(0.03125,2.97136e-07)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.001341)(0.125,8.99204e-05)(0.0625,4.17057e-06)(0.03125,3.6443e-07)};
\logLogSlopeTriangleFlip{.5}{.125}{.475}{4}{}

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.000623337)(0.125,9.84177e-06)(0.0625,2.93926e-07)(0.03125,9.019e-09)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(0.25,0.000128353)(0.125,3.87921e-06)(0.0625,6.10617e-08)(0.03125,3.57278e-09)};
\logLogSlopeTriangle{.5}{.15}{.125}{5}{}

\end{loglogaxis}
\end{tikzpicture}
}
\caption{One-dimensional convergence of multi-patch DG methods under knot insertion and patch refinement.  Uniform knots are used for all experiments.  Results for refinement by knot insertion are denoted by circles, while results for patch refinement are denoted by squares.}
\label{fig:hconvergence1D}
\end{figure}

We now compare per-dof efficiency between uniform/smoothed knot vectors and mesh/patch refinement.  Figure~\ref{fig:convergence1D} shows the convergence of $L^2$ errors for uniform and smoothed knots at various orders of approximation.  For uniform knots, mesh refinement is more efficient than patch refinement in terms of the number of degrees of freedom required to reach a certain $L^2$ error.  The same behavior is observed for smoothed knots at higher orders ($p>3$), though mesh refinement becomes less efficient than patch refinement at lower orders ($p=2$) for the first order formulation of DG.  This is consistent with observations in Section~\ref{sec:approx} that  (in one dimension) smoothed knots provide an advantage over uniform knot vectors only in the pre-asymptotic under-resolved regime.  For both the first and second order formulations, $L^2$ errors are slightly higher when using smoothed knots, though the difference between uniform and smoothed knot vectors becomes smaller as the number of patches increases.

\begin{figure}
\centering
\subfloat[First order formulation, uniform knots]{
\begin{tikzpicture}
\begin{loglogaxis}[
width=.45\textwidth,
    xlabel={Number of degrees of freedom},   
    ylabel={$L^2$ error},
    xmin=0, xmax=150,
    ymin=1e-10, ymax=.1,    
    legend pos=north east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
    legend entries={$p = 2$, $p=3$, $p=4$, $p=5$},
] 
\pgfplotsset{
cycle list={
{blue, mark=*}, {red,mark=square*}, {black,mark=triangle*},
{blue, mark=*,dashed}, {red,mark=square*,dashed}, {black,mark=triangle*,dashed},
{blue, mark=*,dotted}, {red,mark=square*,dotted}, {black,mark=triangle*,dotted},
{blue, mark=*,dashdotted}, {red,mark=square*,dashdotted}, {black,mark=triangle*,dashdotted}},
legend style={mark options={no markers}}
}
\addlegendimage{no markers,black}
\addlegendimage{no markers,black,dashed}
\addlegendimage{no markers,black,dotted}
\addlegendimage{no markers,black,dashdotted}

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(12,0.0106631)(20,0.00095481)(36,0.000110851)(68,1.37364e-05)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(24,0.000855511)(40,0.00010137)(72,1.2405e-05)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(48,7.68373e-05)(80,1.02342e-05)};

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(14,0.00183495)(22,8.58897e-05)(38,4.97075e-06)(70,3.0529e-07)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(28,7.94928e-05)(44,4.72941e-06)(76,2.92657e-07)};
\addplot+[semithick,  mark options={solid,fill=markercolor}]
coordinates{(56,4.30145e-06)(88,2.845e-07)};

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(16,0.00041217)(24,1.05595e-05)(40,2.98648e-07)(72,9.13749e-09)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,8.10195e-06)(48,2.39798e-07)(80,7.44693e-09)};
\addplot+[semithick,  mark options={solid,fill=markercolor}]
coordinates{(64,2.19131e-07)(96,7.01346e-09)};

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(18,7.95525e-05)(26,1.04357e-06)(42,1.41959e-08)(74,2.14126e-10)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(36,7.59324e-07)(52,1.17281e-08)(84,1.84885e-10)};
\addplot+[semithick,  mark options={solid,fill=markercolor}]
coordinates{(72,1.08769e-08)(104,2.53553e-10)};
\end{loglogaxis}
\end{tikzpicture}
}
\subfloat[First order formulation, smoothed knots]{
\begin{tikzpicture}
\begin{loglogaxis}[
width=.45\textwidth,
    xlabel={Number of degrees of freedom},   
    ylabel={$L^2$ error},
    xmin=0, xmax=150,
    ymin=1e-10, ymax=.1,    
    legend pos=north east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
    legend entries={$p = 2$, $p=3$, $p=4$, $p=5$},
] 
\pgfplotsset{
cycle list={
{blue, mark=*}, {red,mark=square*}, {black,mark=triangle*},
{blue, mark=*,dashed}, {red,mark=square*,dashed}, {black,mark=triangle*,dashed},
{blue, mark=*,dotted}, {red,mark=square*,dotted}, {black,mark=triangle*,dotted},
{blue, mark=*,dashdotted}, {red,mark=square*,dashdotted}, {black,mark=triangle*,dashdotted}},
legend style={mark options={no markers}}
}
\addlegendimage{no markers,black}
\addlegendimage{no markers,black,dashed}
\addlegendimage{no markers,black,dotted}
\addlegendimage{no markers,black,dashdotted}

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(12,0.0210966)(20,0.00277327)(36,0.000377052)(68,4.89405e-05)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(24,0.00122621)(40,0.000191822)(72,2.82418e-05)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(48,8.23467e-05)(80,1.233e-05)};

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(14,0.00350747)(22,0.000280473)(38,2.07193e-05)(70,1.3648e-06)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(28,9.7771e-05)(44,9.29359e-06)(76,7.0952e-07)};
\addplot+[semithick,  mark options={solid,fill=markercolor}]
coordinates{(56,4.41167e-06)(88,3.45481e-07)};

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(16,0.000444765)(24,2.14376e-05)(40,8.86832e-07)(72,3.23075e-08)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,6.77359e-06)(48,3.72033e-07)(80,1.57506e-08)};
\addplot+[semithick,  mark options={solid,fill=markercolor}]
coordinates{(64,1.72723e-07)(96,7.66514e-09)};

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(18,4.94071e-05)(26,1.49166e-06)(42,3.28811e-08)(74,5.88541e-10)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(36,4.05142e-07)(52,1.42633e-08)(84,3.36761e-10)};
\addplot+[semithick,  mark options={solid,fill=markercolor}]
coordinates{(72,6.53937e-09)(104,2.53727e-10)};
\end{loglogaxis}
\end{tikzpicture}
}\\
\subfloat[Second order formulation, uniform knots]{
\begin{tikzpicture}
\begin{loglogaxis}[
width=.45\textwidth,
    xlabel={Number of degrees of freedom},   
    ylabel={$L^2$ error},
    xmin=0, xmax=150,
    ymin=1e-10, ymax=.1,    
    legend pos=north east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
    legend entries={$p = 2$, $p=3$, $p=4$, $p=5$},
] 
\pgfplotsset{
cycle list={
{blue, mark=*}, {red,mark=square*}, {black,mark=triangle*},
{blue, mark=*,dashed}, {red,mark=square*,dashed}, {black,mark=triangle*,dashed},
{blue, mark=*,dotted}, {red,mark=square*,dotted}, {black,mark=triangle*,dotted},
{blue, mark=*,dashdotted}, {red,mark=square*,dashdotted}, {black,mark=triangle*,dashdotted}},
legend style={mark options={no markers}}
}
\addlegendimage{no markers,black}
\addlegendimage{no markers,black,dashed}
\addlegendimage{no markers,black,dotted}
\addlegendimage{no markers,black,dashdotted}

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(12,0.0162082)(20,0.00141823)(36,0.000125808)(68,1.26424e-05)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(24,0.000703955)(40,0.0001003)(72,1.20226e-05)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(48,0.000125593)(80,1.30507e-05)};

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(14,0.00271749)(22,8.35594e-05)(38,4.79948e-06)(70,2.97146e-07)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(28,7.90339e-05)(44,4.66277e-06)(76,2.96322e-07)};
\addplot+[semithick,  mark options={solid,fill=markercolor}]
coordinates{(56,4.37416e-06)(88,2.86356e-07)};

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(16,0.000623315)(24,9.82538e-06)(40,2.93292e-07)(72,9.02485e-09)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,9.57229e-06)(48,2.79274e-07)(80,8.95717e-09)};
\addplot+[semithick,  mark options={solid,fill=markercolor}]
coordinates{(64,2.29205e-07)(96,8.78945e-09)};

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(18,0.000116261)(26,9.33596e-07)(42,1.13206e-08)(74,1.79528e-10)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(36,9.69517e-07)(52,1.10416e-08)(84,1.89631e-10)};
\addplot+[semithick,  mark options={solid,fill=markercolor}]
coordinates{(72,1.35402e-08)(104,1.86444e-10)};
\end{loglogaxis}
\end{tikzpicture}
}
\subfloat[Second order formulation, smoothed knots]{
\begin{tikzpicture}
\begin{loglogaxis}[
width=.45\textwidth,
    xlabel={Number of degrees of freedom},   
    ylabel={$L^2$ error},
    xmin=0, xmax=150,
    ymin=1e-10, ymax=.1,    
    legend pos=north east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
    legend entries={$p = 2$, $p=3$, $p=4$, $p=5$},
] 
\pgfplotsset{
cycle list={
{blue, mark=*}, {red,mark=square*}, {black,mark=triangle*},
{blue, mark=*,dashed}, {red,mark=square*,dashed}, {black,mark=triangle*,dashed},
{blue, mark=*,dotted}, {red,mark=square*,dotted}, {black,mark=triangle*,dotted},
{blue, mark=*,dashdotted}, {red,mark=square*,dashdotted}, {black,mark=triangle*,dashdotted}},
legend style={mark options={no markers}}
}
\addlegendimage{no markers,black}
\addlegendimage{no markers,black,dashed}
\addlegendimage{no markers,black,dotted}
\addlegendimage{no markers,black,dashdotted}

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(12,0.0100998)(20,0.00183953)(36,0.000334866)(68,3.38972e-05)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(24,0.00148053)(40,0.000351277)(72,2.61946e-05)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(48,0.000243683)(80,4.22802e-05)};

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(14,0.00341013)(22,0.000116041)(38,8.20099e-06)(70,9.12238e-07)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(28,0.000166115)(44,1.59664e-05)(76,6.36059e-07)};
\addplot+[semithick,  mark options={solid,fill=markercolor}]
coordinates{(56,1.36033e-05)(88,9.49495e-07)};

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(16,0.00036462)(24,9.40051e-06)(40,1.13518e-06)(72,3.45277e-08)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,1.08149e-05)(48,5.82473e-07)(80,1.9993e-08)};
\addplot+[semithick,  mark options={solid,fill=markercolor}]
coordinates{(64,5.05283e-07)(96,2.09237e-08)};

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(18,4.74162e-05)(26,1.99624e-06)(42,3.33867e-08)(74,3.99503e-10)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(36,9.26336e-07)(52,2.22652e-08)(84,3.71367e-10)};
\addplot+[semithick,  mark options={solid,fill=markercolor}]
coordinates{(72,8.35937e-09)(104,2.87484e-10)};

\end{loglogaxis}
\end{tikzpicture}
}
\caption{Convergence under mesh refinement (knot insertion) of first and second order formulations of the 1D acoustic wave equation using multi-patch DG for $p = 2,\ldots,5$.  Different markers correspond to patches refinement: two patches (circle), four patches (square), eight patches (triangle).  }
\label{fig:convergence1D}
\end{figure}

\subsection{Approximation of oscillatory functions using uniform, optimal, and smoothed knot vectors}
\label{sec:approx}

We now compare the approximation properties of splines under uniform, optimal, and smoothed knot vectors.  Recall that optimal knot vectors are defined such that the corresponding spline space minimizes the ``worst best approximation error'' in the sense of $n$-widths.  This implies that optimal knot vectors (and their approximations using smoothed knot vectors) will be less accurate than spline spaces using uniform knot vectors for certain functions.

Figure~\ref{fig:ppw} shows $L^2$ errors in approximating $\cos\LRp{\frac{(2k-1)\pi x}{2}}$ for various $k$ under uniform, optimal, and smoothed knots over a single patch.  \reviewerOne{Here, we approximate the function using an $L^2$ projection, using a composite Gaussian quadrature rule which is exact for the product of two degree $p$ B-splines.}  The number of wavelengths is taken to be $k = 1/{N_{\rm dofs}},\ldots, N_{\rm dofs}/2$ (where $N_{\rm dofs}$ denotes the dimension of the spline space).  We observe that, while $L^2$ errors grow with $k$ for all knot distributions, this growth of error is delayed until a slightly higher $k$ when using optimal and smoothed knot distributions (compared to uniform knot distributions).  These results also indicate that smoothed and optimal knot distributions behave similarly to uniform knots for smaller values of $k$.

\begin{figure}
\centering
\subfloat[$L^2$ error vs wavelengths ($k$) per dof]{\includegraphics[width=.4675\textwidth]{splineapprox.png}}
\hspace{.5em}
\subfloat[Semi-log scale]{\includegraphics[width=.475\textwidth]{splineapproxlog.png}}
\caption{$L^2$ error in approximating $u(x) = \cos\LRp{\frac{(2k-1)\pi x}{2}}$ at various $k$ with $p = 4, K = 16$.}
\label{fig:ppw}
\end{figure}

Finally, we compare DG solution errors for both uniform and smoothed knot vectors.  We use the exact pressure solution $p(x,t) = \cos\LRp{\frac{2k-1}{2}\pi x}\cos\LRp{\frac{2k-1}{2}\pi t}$  at final time $T=1/2$, and compute $L^2$ errors on a single patch domain.  Errors are computed for various wavenumbers $k$ but for a fixed order $p$ and number of elements $K$.  The $L^2$ errors (shown in Figure~\ref{fig:wavelengthconverge}) show qualitatively similarity to the approximation results of Figure~\ref{fig:ppw}, with uniform knots performing better at small $k$.  Smoothed knots result in lower errors for higher $k$, though this effect is more pronounced for first order formulations of the acoustic wave equation and less noticeable for the second order formulation.  However, in all cases, the $L^2$ errors for both uniform and smoothed knots are roughly of the same order of magnitude.  We note that the value of $k$ at which smoothed knots deliver lower error than uniform knots depends on $p$ and $K$; however, we observe in numerical experiments that this ``crossover point'' grows roughly proportionally to $K$ but depends only weakly on $p$.

\begin{figure}
\centering
\subfloat[First order formulation]{
\begin{tikzpicture}
\begin{semilogyaxis}[
width=.45\textwidth,
%    xlabel={Wavenumber $k$},   
    xlabel={Wavelengths $k$ per dof},   
    ylabel={$L^2$ error},
    ymin=1e-5, ymax=2,    
    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
] 
\addplot[color=blue,mark=*, semithick, mark options={solid,fill=markercolor}]
%coordinates{(1,1.86794e-05)(3,0.00199192)(5,0.0632619)(7,0.616786)(9,0.712371)(11,0.768705)};
coordinates{(1/11,1.86794e-05)(3/11,0.00199192)(5/11,0.0632619)(7/11,0.616786)(9/11,0.712371)(11/11,0.768705)};
\addplot[color=red,mark=*, dashed, semithick, mark options={solid,fill=markercolor}]
%coordinates{(1,5.56475e-05)(3,0.00612773)(5,0.0351819)(7,0.405041)(9,0.772862)(11,0.731632)};
coordinates{(1/11,5.56475e-05)(3/11,0.00612773)(5/11,0.0351819)(7/11,0.405041)(9/11,0.772862)(11/11,0.731632)};
\legend{Uniform knots, Smoothed knots}
\end{semilogyaxis}
\end{tikzpicture}
}
\subfloat[Second order formulation]{
\begin{tikzpicture}
\begin{semilogyaxis}[
width=.45\textwidth,
%    xlabel={Wavenumber $k$},   
    xlabel={Wavelengths $k$ per dof},   
    ylabel={$L^2$ error},
    ymin=1e-5, ymax=2,    
    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
] 
\addplot[color=blue,mark=*, semithick, mark options={solid,fill=markercolor}]
%coordinates{(1,1.77384e-05)(3,0.00244169)(5,0.0592546)(7,1.07406)(9,0.662767)(11,0.872701)};
coordinates{(1/11,1.77384e-05)(3/11,0.00244169)(5/11,0.0592546)(7/11,1.07406)(9/11,0.662767)(11/11,0.872701)};
\addplot[color=red,mark=*, dashed, semithick, mark options={solid,fill=markercolor}]
%coordinates{(1,1.78285e-05)(3,0.00465007)(5,0.107561)(7,0.751711)(9,0.864029)(11,0.773654)};
coordinates{(1/11,1.78285e-05)(3/11,0.00465007)(5/11,0.107561)(7/11,0.751711)(9/11,0.864029)(11/11,0.773654)};

\legend{Uniform knots, Smoothed knots}
\end{semilogyaxis}
\end{tikzpicture}
}
\caption{Comparison of $L^2$ errors for uniform and smooth knots as a function of wavenumber $k$ using $p = 3, K= 8$ splines.}
\label{fig:wavelengthconverge}
\end{figure}

\subsection{Two-dimensional experiments}
\label{sec:twodimexp}
%In higher spatial dimensions, we observe similar approximation errors as reported in Figures~\ref{fig:href1D}, \ref{fig:pref1D}, and \ref{fig:ppw}.  
In higher spatial dimensions, we observe similar approximation behavior as reported in Figure~\ref{fig:ppw}.  We consider the curvilinear mapping $(\tilde{x},\tilde{y})$ of a bi-unit square $[-1,1]^2$ shown in Figure~\ref{subfig:mapped} and defined by 
\[
\tilde{x} = x + \alpha\cos\LRp{\frac{3 \pi y}{2}} \cos\LRp{\frac{\pi x}{2}}, \qquad \tilde{y} = y + \alpha \sin\LRp{\frac{3 \pi x}{2}}\cos\LRp{\frac{\pi y}{2}}.
\]
We begin by verifying that the weight-adjusted approximation to the mass matrix inverse is sufficiently accurate approximation.  Figure~\ref{fig:l2vswadg} shows $L^2$ errors in approximating the smooth function $\cos\LRp{\frac{k\pi x}{2}}\cos\LRp{\frac{k\pi y}{2}}$ from a uniform $p=4$ spline space.  The errors for the $L^2$ projection computed using the true inverse mass matrix and using the weight-adjusted mass matrix inverse are nearly identical, and are indistinguishable visually for both $k=1$ (corresponding to a smooth function) and for $k=10$ (corresponding to an oscillatory function).

We refer to the $L^2$ projection computed using the weight adjusted mass matrix inverse as a weight-adjusted projection.  igure~\ref{fig:l2vswadg} also shows the difference between the $L^2$ and weight-adjusted projections, which we observe to be at least one order of magnitude smaller than the approximation error for both $k=1$ and $k=10$.  Additionally, this difference appears to converge more rapidly than the error under mesh refinement.

\begin{figure}
\centering
\subfloat[Warped mesh, $\alpha = 1/8$]{\includegraphics[width=.275\textwidth]{mapped.png}\label{subfig:mapped}}
\subfloat[$L^2$ errors for $k = 1$]{
\begin{tikzpicture}
\begin{loglogaxis}[
    legend cell align=left,
    legend style={legend pos=south east, font=\tiny},
    width=.36\textwidth,
    xlabel={Mesh size $h$},
    ylabel={$L^2$ error}, 
             ymin=1e-10, ymax=2,    
    grid style=dashed,
] 
\addplot[color=blue,mark=*,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.0110908)(0.25,0.00138298)(0.125,2.29018e-05)(0.0625,3.18561e-07)};
\addplot[color=red,mark=x,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.0111346)(0.25,0.00138582)(0.125,2.30061e-05)(0.0625,3.18565e-07)};
\addplot[color=black,mark=triangle*,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.000986363)(0.25,8.87776e-05)(0.125,2.18816e-06)(0.0625,1.65013e-09)};

\legend{$L^2$ projection,Weight-adjusted,Difference}
%\legend{Uniform, Optimal, Smoothed}
\end{loglogaxis}
\end{tikzpicture}
}
\subfloat[$L^2$ errors for $k = 10$]{
\begin{tikzpicture}
\begin{loglogaxis}[
    legend cell align=left,
    legend style={legend pos=south east, font=\tiny},
    width=.36\textwidth,
         ymin=1e-10, ymax=2,    
    xlabel={Mesh size $h$},
    ylabel={$L^2$ error}, 
    grid style=dashed,
] 
\addplot[color=blue,mark=*,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.999801)(0.25,0.903719)(0.125,0.218432)(0.0625,0.00375629)};
\addplot[color=red,mark=x,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.999803)(0.25,0.90373)(0.125,0.218442)(0.0625,0.00375629)};
\addplot[color=black,mark=triangle*,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.00163151)(0.25,0.00435668)(0.125,0.00205162)(0.0625,5.51212e-06)};
\legend{$L^2$ projection, Weight-adjusted, Difference}
%\legend{Uniform, Optimal, Smoothed}
\end{loglogaxis}
\end{tikzpicture}
}
\caption{Errors for $L^2$ and weight-adjusted projection when approximating $\cos\LRp{\frac{k\pi x}{2}}\cos\LRp{\frac{k\pi y}{2}}$ on a warped curvilinear mesh with $p = 4$ splines for $k = 1$ and $k=10$. }
%\caption{$L^2$ vs weight-adjusted projection. }
\label{fig:l2vswadg}
\end{figure}

\reviewerTwo{Theory in \cite{chan2016weight1} suggests that the weight-adjusted approximation becomes less accurate as the geometric mapping becomes less regular.  We test this by considering a more aggressively warped mapping with $\alpha = .28$.  This value is chosen to maximize the irregularity of $J$ without losing invertibility of the geometric mapping (for a slightly larger $\alpha = .285$, negative values of $J$ are detected, indicating that the mapping becomes non-invertible).  Figure~\ref{fig:l2vswadg2} shows a visualization of the warped mesh, along with  $L^2$ errors for both the  $L^2$ and weight-adjusted projections and the $L^2$ difference between the $L^2$ and weight-adjusted projection.

We consider again the function $\cos\LRp{\frac{k\pi x}{2}}\cos\LRp{\frac{k\pi y}{2}}$ for $k=1$ and $k = 10$.  For $k=10$, the approximated function is oscillatory and less well-resolved, and we do not observe any significant differences between the $L^2$ and weight-adjusted projections.  However, for $k=1$, the approximated function is smooth, and we observe a noticeable difference between the $L^2$ error for the weight-adjusted and $L^2$ projections on the coarsest mesh.  This confirms the fact that the weight-adjusted projection becomes less accurate as the geometric mapping approaches non-invertibility.  However, even for this more irregular mapping, the difference between the weight-adjusted and $L^2$ projection decreases rapidly under mesh refinement.

Finally, we note that it was shown in \cite{chan2016weight2} that computing the weight-adjusted projection of some function $u$ is equivalent to computing 
\[
\widehat{\Pi}_h\LRp{\frac{1}{J}\widehat{\Pi}_h \LRp{uJ}},
\]
where $\widehat{\Pi}_h$ denotes the $L^2$ projection operator onto $V_h\LRp{\widehat{D}}$.  This suggests that the weight-adjusted inverse mass matrix is a highly accurate approximation to the inverse curvilinear mass matrix, so long as the geometric mapping is sufficiently well-resolved approximated by the spline approximation space on the reference element.
%These experiments suggest that under mesh refinement, 
}

\begin{figure}
\centering
\subfloat[Warped mesh, $\alpha = .28$]{\includegraphics[width=.275\textwidth]{mapped2.png}\label{subfig:mapped2}}
\subfloat[$L^2$ errors for $k = 1$]{
\begin{tikzpicture}
\begin{loglogaxis}[
    legend cell align=left,
    legend style={legend pos=south east, font=\tiny},
    width=.36\textwidth,
     ymin=1e-8, ymax=2,    
    xlabel={Mesh size $h$},
    ylabel={$L^2$ error}, 
    grid style=dashed,
] 
\addplot[color=blue,mark=*,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.0444376)(0.25,0.0144725)(0.125,0.000702933)(0.0625,6.79061e-06)};
\addplot[color=red,mark=x,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.1182)(0.25,0.017419)(0.125,0.000745669)(0.0625,6.79513e-06)};
\addplot[color=black,mark=triangle*,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.109529)(0.25,0.00969371)(0.125,0.000248811)(0.0625,2.47881e-07)};

\legend{$L^2$ projection,Weight-adjusted,Difference}
%\legend{Uniform, Optimal, Smoothed}
\end{loglogaxis}
\end{tikzpicture}
}
\subfloat[$L^2$ errors for $k = 10$]{
\begin{tikzpicture}
\begin{loglogaxis}[
    legend cell align=left,
    legend style={legend pos=south east, font=\tiny},
    width=.36\textwidth,
         ymin=1e-8, ymax=2,    
    xlabel={Mesh size $h$},
    ylabel={$L^2$ error}, 
    grid style=dashed,
] 
\addplot[color=blue,mark=*,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.952758)(0.25,0.89781)(0.125,0.4989)(0.0625,0.121744)};
\addplot[color=red,mark=x,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.954139)(0.25,0.898161)(0.125,0.499104)(0.0625,0.121753)};
\addplot[color=black,mark=triangle*,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.0513167)(0.25,0.0251125)(0.125,0.0142617)(0.0625,0.00150984)};

\legend{$L^2$ projection, Weight-adjusted, Difference}
%\legend{Uniform, Optimal, Smoothed}
\end{loglogaxis}
\end{tikzpicture}
}
\caption{Errors for $L^2$ and weight-adjusted projection when approximating $\cos\LRp{\frac{k\pi x}{2}}\cos\LRp{\frac{k\pi y}{2}}$ on a more aggressively warped curvilinear mesh with $p = 4$ splines for $k = 1$ and $k=10$. }
%\caption{$L^2$ vs weight-adjusted projection. }
\label{fig:l2vswadg2}
\end{figure}

\subsubsection{Approximation errors under degree refinement}

We now investigate approximation errors under degree refinement on curved meshes.  These experiments suggest that splines under optimal and smoothed knot vectors possess some advantages over uniform knot distributions in the presence of non-affine mappings.  Figure~\ref{fig:mapped} shows computed $L^2$ errors in approximating $\cos\LRp{\frac{3\pi x}{2}}\cos\LRp{\frac{3\pi y}{2}}$ on a curvilinear mesh.  We examine errors under degree refinement for both tensor-product polynomial and spline spaces (where we fix the number of  elements $K=p$).  \reviewerOne{We compute approximations using quadrature-based weight-adjusted projection, where the inverse curvilinear mass matrix is approximated using a weight-adjusted approximation.}

In contrast to the affine case, we observe that both polynomials and spline spaces with uniform knot vectors produce $L^2$ errors of similar magnitude per degree of freedom, even for low values of $\alpha$ (lightly warped curvilinear meshes).  We also observe that splines spaces using optimal and smoothed knot vectors produce lower $L^2$ errors than both polynomials and uniform splines.  These numerical results suggest that $L^2$ approximation errors under spline spaces are comparable to polynomials not only for oscillatory functions (as was observed in 1D), but also for curvilinear mappings.  Additionally, spline spaces with optimal or smoothed knot vectors produce better approximations for both oscillatory functions and curved mappings.

\begin{figure}
\centering
%\subfloat[Warped mesh, $\alpha = 1/8$]{\includegraphics[width=.5\textwidth]{mapped.png}}
\subfloat[$L^2$ errors ]{
\begin{tikzpicture}
\begin{semilogyaxis}[
    legend cell align=left,
    legend style={legend pos=north east, font=\tiny},
    width=.5\textwidth,
    xlabel={Degrees of freedom},
    ylabel={$L^2$ error}, 
    grid style=dashed,
]

\addplot[color=blue,mark=*,semithick, mark options={solid,fill=markercolor}]
coordinates{(16,0.0345598)(36,0.00291687)(64,0.000910808)(100,0.000228283)(144,4.68154e-05)(196,1.08049e-05)(256,2.35056e-06)}
[yshift=4pt] node[above, pos=.8, color=black] {$\alpha = 1/64$};
\addplot[color=red,mark=square*,semithick, mark options={solid,fill=markercolor}]
coordinates{(16,0.0345598)(36,0.0018151)(64,0.000489718)(100,8.29769e-05)(144,1.29416e-05)};
\addplot[color=black,mark=triangle*,semithick, mark options={solid,fill=markercolor}]
coordinates{(16,0.0345598)(36,0.00184734)(64,0.000495297)(100,8.68887e-05)(144,1.63983e-05)(196,1.69855e-06)(256,4.46827e-07)};
\addplot[color=violet,mark=diamond*,semithick, mark options={solid,fill=markercolor}]
coordinates{(9,0.0353096)(16,0.0346236)(25,0.00282808)(36,0.00233183)(49,0.00142429)(64,0.000710684)(81,0.000305903)(100,0.000123394)(121,4.1829e-05)(144,2.40683e-05)(169,6.2327e-06)(196,5.06423e-06)(225,1.05795e-06)(256,8.08652e-07)};

\addplot[color=blue,mark=*,dashed,semithick, mark options={solid,fill=markercolor}]
coordinates{(16,0.0367809)(36,0.0169805)(64,0.0111257)(100,0.00520972)(144,0.00196454)(196,0.000609694)(256,0.000194924)}
[yshift=4pt] node[above, pos=.85,color=black] {$\alpha = 1/8$};
\addplot[color=red,mark=square*,dashed,semithick, mark options={solid,fill=markercolor}]
coordinates{(16,0.0367809)(36,0.014593)(64,0.00828577)(100,0.00316138)(144,0.000836099)};
\addplot[color=black,mark=triangle*,dashed,semithick, mark options={solid,fill=markercolor}]
coordinates{(16,0.0367809)(36,0.014438)(64,0.008333)(100,0.00330281)(144,0.000899406)(196,0.000319152)(256,0.000106414)};
\addplot[color=violet,mark=diamond*,dashed,semithick, mark options={solid,fill=markercolor}]
coordinates{(9,0.0744743)(16,0.0402554)(25,0.0332601)(36,0.0185583)(49,0.01937)(64,0.00949402)(81,0.00704509)(100,0.00431925)(121,0.00311497)(144,0.00145279)(169,0.00121632)(196,0.000388163)(225,0.00042613)};

\legend{Uniform, Optimal, Smoothed,Polynomial}
%\legend{Uniform, Optimal, Smoothed}
\end{semilogyaxis}
\end{tikzpicture}
}
\caption{$L^2$ best approximation errors for $\cos\LRp{\frac{k\pi x}{2}}\cos\LRp{\frac{k\pi y}{2}}, k = 1$ under different mesh warpings and degree refinement $p = 2,\ldots, 8$.  All results assume that spline spaces contain $K = p$ elements, \reviewerTwo{except for the ``Polynomial'' curve, which assumes degree $p = 2,\ldots, 12$ polynomial approximations over a single element}. }
\label{fig:mapped}
\end{figure}

\subsubsection{$L^2$ errors for multi-patch DG solutions }

Next, we examine the behavior of multi-patch DG methods for the acoustic wave equation in two dimensions using both uniform and smoothed knot vectors.  We compute $L^2$ errors for the exact solution $p(x,y,t) = \cos\LRp{\frac{3\pi x}{2}}\cos\LRp{\frac{3\pi y}{2}}\cos\LRp{\frac{3\sqrt{2} \pi t}{2}}$ at $T = 1/2$.  For affine patches, the convergence of $L^2$ errors is similar to the 1D case; thus, we consider a curved interior warping of a square (Figure~\ref{fig:mapped}) in order to account for the effect of geometric mappings.

\reviewerTwo{We first compare $L^2$ pressure errors when using the curvilinear mass matrix inverse and the weight adjusted mass matrix inverse.  In both cases, uniform splines of degree $p=4$ are used.  Figure~\ref{fig:l2vswadgerror} shows $L^2$ errors for both first and second order formulations for both a lightly warped curvilinear mesh ($\alpha = 1/8$) and a heavily warped curvilinear mesh ($\alpha = .28$).   We observe that the $L^2$ errors are visually indistinguishable, and that the $L^2$ difference between the two solutions is several orders of magnitude smaller than the $L^2$ discretization error.  This mirrors the qualitative behavior observed in Figures~\ref{fig:l2vswadg} and \ref{fig:l2vswadg2}; however, the $L^2$ difference between the solutions obtained using the curvilinear and weight-adjusted mass matrix inverses is significantly smaller than the difference between $L^2$ and weight-adjusted projections. }  $L^2$ errors are shown in Figure~\ref{fig:convergence2D}, where, for clarity, we show results only for $p=2$ and $p=5$.

\begin{figure}
\subfloat[First order formulation, $\alpha = 1/8$]{
\begin{tikzpicture}
\begin{loglogaxis}[
    legend cell align=left,
    legend style={legend pos=south east, font=\tiny},
    width=.4\textwidth,
     ymin=1e-17, ymax=2,    
    xlabel={Mesh size $h$},
    ylabel={$L^2$ error}, 
    grid style=dashed,
] 
\addplot[color=blue,mark=*,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.154255)(0.25,0.0280139)(0.125,0.000775805)(0.0625,7.51259e-06)};
\addplot[color=red,mark=x,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.154523)(0.25,0.0280342)(0.125,0.000775107)(0.0625,7.51217e-06)};
\addplot[color=black,mark=triangle*,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,7.56052e-06)(0.25,1.32834e-07)(0.125,1.51285e-11)(0.0625,1.69165e-17)};

\legend{$L^2$ projection,Weight-adjusted,Difference}
%\legend{Uniform, Optimal, Smoothed}
\end{loglogaxis}
\end{tikzpicture}
}
\subfloat[Second order formulation, $\alpha = 1/8$]{
\begin{tikzpicture}
\begin{loglogaxis}[
    legend cell align=left,
    legend style={legend pos=south east, font=\tiny},
    width=.4\textwidth,
     ymin=1e-17, ymax=2,    
    xlabel={Mesh size $h$},
    ylabel={$L^2$ error}, 
    grid style=dashed,
] 
\addplot[color=blue,mark=*,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.270044)(0.25,0.0483258)(0.125,0.00134115)(0.0625,1.00287e-05)};
\addplot[color=red,mark=x,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.272282)(0.25,0.0483238)(0.125,0.00133867)(0.0625,1.0026e-05)};
\addplot[color=black,mark=triangle*,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,4.77833e-05)(0.25,1.1668e-06)(0.125,3.6017e-11)(0.0625,2.65088e-16)};

\legend{$L^2$ projection,Weight-adjusted,Difference}
%\legend{Uniform, Optimal, Smoothed}
\end{loglogaxis}
\end{tikzpicture}
}\\
\subfloat[First order formulation, $\alpha = .28$]{
\begin{tikzpicture}
\begin{loglogaxis}[
    legend cell align=left,
    legend style={legend pos=south east, font=\tiny},
    width=.4\textwidth,
     ymin=1e-17, ymax=2,    
    xlabel={Mesh size $h$},
    ylabel={$L^2$ error}, 
    grid style=dashed,
] 
\addplot[color=blue,mark=*,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.338603)(0.25,0.143169)(0.125,0.0234825)(0.0625,0.00033695)};
\addplot[color=red,mark=x,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.388935)(0.25,0.146233)(0.125,0.0235623)(0.0625,0.000336824)};
\addplot[color=black,mark=triangle*,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.00977253)(0.25,0.000187168)(0.125,3.16763e-07)(0.0625,9.44863e-12)};

\legend{$L^2$ projection,Weight-adjusted,Difference}
%\legend{Uniform, Optimal, Smoothed}
\end{loglogaxis}
\end{tikzpicture}
}
\subfloat[Second order formulation, $\alpha = .28$]{
\begin{tikzpicture}
\begin{loglogaxis}[
    legend cell align=left,
    legend style={legend pos=south east, font=\tiny},
    width=.4\textwidth,
     ymin=1e-17, ymax=2,  
           xlabel={Mesh size $h$},
    ylabel={$L^2$ error}, 
    grid style=dashed,
] 
\addplot[color=blue,mark=*,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.660221)(0.25,0.272201)(0.125,0.0336296)(0.0625,0.000337592)};
\addplot[color=red,mark=x,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.68491)(0.25,0.280896)(0.125,0.0335946)(0.0625,0.000337711)};
\addplot[color=black,mark=triangle*,semithick, mark options={solid,fill=markercolor}]
coordinates{(0.5,0.0580108)(0.25,0.000653681)(0.125,7.91095e-07)(0.0625,1.2372e-11)};

\legend{$L^2$ projection,Weight-adjusted,Difference}
%\legend{Uniform, Optimal, Smoothed}
\end{loglogaxis}
\end{tikzpicture}
}

\caption{$L^2$ errors for the acoustic wave equation when inverting the curvilinear mass matrix and the weight-adjusted mass matrix.   Uniform $p = 4$ splines are used, and the meshes are the two warped curvilinear meshes shown in Figures~\ref{fig:l2vswadg} and \ref{fig:l2vswadg2}. }
\label{fig:l2vswadgerror}
\end{figure}

For first order DG formulations, we observe that $L^2$ errors for uniform and smoothed knot vectors are closer \reviewerOne{than in} the one-dimensional case, which appears to be due to \reviewerOne{the presence of the non-affine (curvilinear) geometric mapping}.  \reviewerOne{This suggests that the use of smoothed knots instead of uniform knots can improve solver efficiency, since it is possible to take a larger maximum stable timestep without reducing accuracy.  A more rigorous study of this claim is still required.}  Under the second order DG formulation, $L^2$ errors for smoothed knot vectors behave less consistently.  As with the first order formulation, the use of smoothed knot vectors reduces errors in the pre-asymptotic range for the second order formulation with $p > 2$.  However, in contrast with the first order formulation, $L^2$ errors for smoothed knots become larger than errors for uniform knot vectors under continued patch or mesh refinement.\footnote{Smoothed knot vectors tend to offer less of an advantage over uniform knot vectors for second order formulations.  This may be due to the fact that smoothed knot vectors are constructed to optimize the $L^2$ sup-inf in which first order DG formulations are more naturally convergent  \cite{warburton2013low, chan2016weight1}, while second order DG formulations converge in a stronger energy norm.  However, the precise reasons for this difference are not immediately clear, and will be investigated in future work.}

\begin{figure}
\centering
\subfloat[First order formulation, uniform knots]{
\begin{tikzpicture}
\begin{loglogaxis}[
width=.45\textwidth,
    xlabel={Number of degrees of freedom},   
    ylabel={$L^2$ error},
    xmin=0, xmax=15000,
    ymin=1e-8, ymax=1,    
    legend pos=north east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
%    legend entries={$p = 2$, $p=3$, $p=4$, $p=5$},
    legend entries={$p = 2$, $p=5$},
] 
\pgfplotsset{
cycle list={
{blue, mark=*}, {red,mark=square*}, {black,mark=triangle*},
{blue, mark=*,dashed}, {red,mark=square*,dashed}, {black,mark=triangle*,dashed},
{blue, mark=*,dotted}, {red,mark=square*,dotted}, {black,mark=triangle*,dotted},
{blue, mark=*,dashdotted}, {red,mark=square*,dashdotted}, {black,mark=triangle*,dashdotted}},
legend style={mark options={no markers}}
}
\addlegendimage{no markers,black}
\addlegendimage{no markers,black,dashed}
\addlegendimage{no markers,black,dotted}
\addlegendimage{no markers,black,dashdotted}

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(64,0.242695)(144,0.0715275)(400,0.00658214)(1296,0.000525301)(4624,6.03423e-05)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(256,0.0357164)(576,0.00686446)(1600,0.000545882)(5184,6.20872e-05)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(1024,0.00435721)(2304,0.000488199)(6400,5.83311e-05)};

%\addplot+[semithick, mark options={solid,fill=markercolor}]
%coordinates{(100,0.132165)(196,0.0334468)(484,0.00222776)(1444,6.0903e-05)(4900,3.07891e-06)};
%\addplot+[semithick, mark options={solid,fill=markercolor}]
%coordinates{(400,0.00927583)(784,0.00166601)(1936,5.76982e-05)(5776,3.01869e-06)};
%\addplot+[semithick,  mark options={solid,fill=markercolor}]
%coordinates{(1600,0.000542582)(3136,5.61886e-05)(7744,2.98486e-06)};
%
%\addplot+[semithick, mark options={solid,fill=markercolor}]
%coordinates{(144,0.0619962)(256,0.017935)(576,0.000867235)(1600,9.30881e-06)(5184,2.06241e-07)};
%\addplot+[semithick, mark options={solid,fill=markercolor}]
%coordinates{(576,0.00259989)(1024,0.000491411)(2304,9.22672e-06)(6400,2.10835e-07)};
%\addplot+[semithick,  mark options={solid,fill=markercolor}]
%coordinates{(2304,7.22065e-05)(4096,8.75141e-06)(9216,1.9849e-07)};

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(196,0.0288428)(324,0.007685)(676,0.000482686)(1764,1.86168e-06)(5476,1.65548e-08)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(784,0.000691368)(1296,0.000135871)(2704,1.79719e-06)(7056,1.65087e-08)};
\addplot+[semithick,  mark options={solid,fill=markercolor}]
coordinates{(3136,9.71515e-06)(5184,1.35491e-06)(10816,1.51573e-08)};

\end{loglogaxis}
\end{tikzpicture}
}
\subfloat[First order formulation, smoothed knots]{
\begin{tikzpicture}
\begin{loglogaxis}[
width=.45\textwidth,
    xlabel={Number of degrees of freedom},   
    ylabel={$L^2$ error},
    xmin=0, xmax=15000,
    ymin=1e-8, ymax=1,    
    legend pos=north east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
%    legend entries={$p = 2$, $p=3$, $p=4$, $p=5$},
        legend entries={$p = 2$, $p=5$},
] 
\pgfplotsset{
cycle list={
{blue, mark=*}, {red,mark=square*}, {black,mark=triangle*},
{blue, mark=*,dashed}, {red,mark=square*,dashed}, {black,mark=triangle*,dashed},
{blue, mark=*,dotted}, {red,mark=square*,dotted}, {black,mark=triangle*,dotted},
{blue, mark=*,dashdotted}, {red,mark=square*,dashdotted}, {black,mark=triangle*,dashdotted}},
legend style={mark options={no markers}}
}
\addlegendimage{no markers,black}
\addlegendimage{no markers,black,dashed}
\addlegendimage{no markers,black,dotted}
\addlegendimage{no markers,black,dashdotted}

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(64,0.153639)(144,0.0418493)(400,0.00442582)(1296,0.000564782)(4624,7.79315e-05)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(256,0.0256049)(576,0.00420403)(1600,0.000586667)(5184,8.83518e-05)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(1024,0.00388923)(2304,0.000639539)(6400,9.87198e-05)};

%\addplot+[semithick, mark options={solid,fill=markercolor}]
%coordinates{(100,0.0930522)(196,0.0171325)(484,0.00138355)(1444,8.64099e-05)(4900,4.83299e-06)};
%\addplot+[semithick, mark options={solid,fill=markercolor}]
%coordinates{(400,0.0072846)(784,0.000906647)(1936,8.32599e-05)(5776,5.96369e-06)};
%\addplot+[semithick,  mark options={solid,fill=markercolor}]
%coordinates{(1600,0.000530081)(3136,6.26226e-05)(7744,4.79183e-06)};
%
%\addplot+[semithick, mark options={solid,fill=markercolor}]
%coordinates{(144,0.0438962)(256,0.00927686)(576,0.000397798)(1600,1.40392e-05)(5184,3.60713e-07)};
%\addplot+[semithick, mark options={solid,fill=markercolor}]
%coordinates{(576,0.002197)(1024,0.000194099)(2304,1.36422e-05)(6400,4.07817e-07)};
%\addplot+[semithick,  mark options={solid,fill=markercolor}]
%coordinates{(2304,6.84186e-05)(4096,5.84491e-06)(9216,2.17164e-07)};

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(196,0.0209849)(324,0.00383302)(676,0.00020681)(1764,2.96835e-06)(5476,3.21437e-08)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(784,0.000588702)(1296,5.06718e-05)(2704,1.96211e-06)(7056,2.59168e-08)};
\addplot+[semithick,  mark options={solid,fill=markercolor}]
coordinates{(3136,8.87494e-06)(5184,4.8802e-07)(10816,1.50029e-08)};

\end{loglogaxis}
\end{tikzpicture}
}\\
\subfloat[Second order formulation, uniform knots]{
\begin{tikzpicture}
\begin{loglogaxis}[
width=.45\textwidth,
    xlabel={Number of degrees of freedom},   
    ylabel={$L^2$ error},
    xmin=0, xmax=15000,
    ymin=1e-8, ymax=1,    
    legend pos=north east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
%    legend entries={$p = 2$, $p=3$, $p=4$, $p=5$},
    legend entries={$p = 2$, $p=5$},
] 
\pgfplotsset{
cycle list={
{blue, mark=*}, {red,mark=square*}, {black,mark=triangle*},
{blue, mark=*,dashed}, {red,mark=square*,dashed}, {black,mark=triangle*,dashed},
{blue, mark=*,dotted}, {red,mark=square*,dotted}, {black,mark=triangle*,dotted},
{blue, mark=*,dashdotted}, {red,mark=square*,dashdotted}, {black,mark=triangle*,dashdotted}},
legend style={mark options={no markers}}
}
\addlegendimage{no markers,black}
\addlegendimage{no markers,black,dashed}
\addlegendimage{no markers,black,dotted}
\addlegendimage{no markers,black,dashdotted}

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(64,0.359903)(144,0.0988039)(400,0.00700231)(1296,0.000596158)(4624,6.44021e-05)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(256,0.0647089)(576,0.00662372)(1600,0.000637467)(5184,6.86903e-05)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(1024,0.00635845)(2304,0.000602189)(6400,6.75732e-05)};

%\addplot+[semithick, mark options={solid,fill=markercolor}]
%coordinates{(100,0.190103)(196,0.0415714)(484,0.00204923)(1444,6.0709e-05)(4900,3.1327e-06)};
%\addplot+[semithick, mark options={solid,fill=markercolor}]
%coordinates{(400,0.0181315)(784,0.0017007)(1936,5.89272e-05)(5776,3.07284e-06)};
%\addplot+[semithick,  mark options={solid,fill=markercolor}]
%coordinates{(1600,0.00077351)(3136,5.6308e-05)(7744,3.01529e-06)};
%
%\addplot+[semithick, mark options={solid,fill=markercolor}]
%coordinates{(144,0.113336)(256,0.0242542)(576,0.000803146)(1600,1.00835e-05)(5184,2.17046e-07)};
%\addplot+[semithick, mark options={solid,fill=markercolor}]
%coordinates{(576,0.00480314)(1024,0.000546509)(2304,9.80952e-06)(6400,2.23621e-07)};
%\addplot+[semithick,  mark options={solid,fill=markercolor}]
%coordinates{(2304,0.000104349)(4096,9.11874e-06)(9216,2.13889e-07)};

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(196,0.0419686)(324,0.0108523)(676,0.00046736)(1764,1.88401e-06)(5476,1.74519e-08)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(784,0.00112145)(1296,0.000171121)(2704,1.86063e-06)(7056,1.87817e-08)};
\addplot+[semithick,  mark options={solid,fill=markercolor}]
coordinates{(3136,1.45546e-05)(5184,1.52434e-06)(10816,1.79508e-08)};
\end{loglogaxis}
\end{tikzpicture}
}
\subfloat[Second order formulation, smoothed knots]{
\begin{tikzpicture}
\begin{loglogaxis}[
width=.45\textwidth,
    xlabel={Number of degrees of freedom},   
    ylabel={$L^2$ error},
    xmin=0, xmax=15000,
    ymin=1e-8, ymax=1,    
    legend pos=north east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
%    legend entries={$p = 2$, $p=3$, $p=4$, $p=5$},
        legend entries={$p = 2$, $p=5$},
] 
\pgfplotsset{
cycle list={
{blue, mark=*}, {red,mark=square*}, {black,mark=triangle*},
{blue, mark=*,dashed}, {red,mark=square*,dashed}, {black,mark=triangle*,dashed},
{blue, mark=*,dotted}, {red,mark=square*,dotted}, {black,mark=triangle*,dotted},
{blue, mark=*,dashdotted}, {red,mark=square*,dashdotted}, {black,mark=triangle*,dashdotted}},
legend style={mark options={no markers}}
}
\addlegendimage{no markers,black}
\addlegendimage{no markers,black,dashed}
\addlegendimage{no markers,black,dotted}
\addlegendimage{no markers,black,dashdotted}

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(64,0.354708)(144,0.0897704)(400,0.00913605)(1296,0.0010847)(4624,0.000122353)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(256,0.0635767)(576,0.00668419)(1600,0.00106006)(5184,0.000137968)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(1024,0.00603353)(2304,0.000890171)(6400,0.000122191)};

%\addplot+[semithick, mark options={solid,fill=markercolor}]
%coordinates{(100,0.189564)(196,0.0341518)(484,0.00200879)(1444,0.000127476)(4900,9.41605e-06)};
%\addplot+[semithick, mark options={solid,fill=markercolor}]
%coordinates{(400,0.0180879)(784,0.0014787)(1936,0.000118006)(5776,1.07647e-05)};
%\addplot+[semithick,  mark options={solid,fill=markercolor}]
%coordinates{(1600,0.00075651)(3136,8.13628e-05)(7744,8.37923e-06)};
%
%\addplot+[semithick, mark options={solid,fill=markercolor}]
%coordinates{(144,0.113363)(256,0.0163238)(576,0.00056338)(1600,2.35036e-05)(5184,5.80131e-07)};
%\addplot+[semithick, mark options={solid,fill=markercolor}]
%coordinates{(576,0.0047926)(1024,0.000296857)(2304,1.63539e-05)(6400,5.98945e-07)};
%\addplot+[semithick,  mark options={solid,fill=markercolor}]
%coordinates{(2304,0.000103714)(4096,8.61181e-06)(9216,4.79104e-07)};

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(196,0.0420621)(324,0.00671924)(676,0.000284211)(1764,3.67017e-06)(5476,4.51605e-08)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(784,0.00111957)(1296,6.0986e-05)(2704,2.161e-06)(7056,5.66052e-08)};
\addplot+[semithick,  mark options={solid,fill=markercolor}]
coordinates{(3136,1.45279e-05)(5184,8.23008e-07)(10816,9.14345e-08)};

\end{loglogaxis}
\end{tikzpicture}
}
\caption{Convergence under mesh refinement (knot insertion) of first and second order formulations of the 2D acoustic wave equation using multi-patch DG for $p = 2$ and $p=5$ on \reviewerOne{a warped curved square domain with $\alpha = 1/8$}.  Different markers correspond to patches refinement: two patches (circle), four patches (square), eight patches (triangle).  }
\label{fig:convergence2D}
\end{figure}

Finally, examine the effect of mesh warping on multi-patch DG solutions using both uniform and smoothed knot vectors.  We use the mapping shown in Figure~\ref{fig:mapped}, and compute $L^2$ errors as the warping factor $\alpha$ increases from $0$ to $1/4$.  We fix $K = 8$ and compute errors for $p = 3, 4, 5$ on a single patch domain.  For both first and second order formulations of the acoustic wave equation, smooth knots deliver slightly lower $L^2$ errors than uniform knots.  As with experiments in Figure~\ref{fig:wavelengthconverge}, increasing $K$ increases the value of $\alpha$ (the ``crossover point'') at which smoothed knots become more accurate than uniform knots.

\begin{figure}
\centering
\subfloat[First order formulation]{
\begin{tikzpicture}
\begin{semilogyaxis}[
width=.45\textwidth,
    xlabel={Warping factor $\alpha$},   
    ylabel={$L^2$ error},
    ymin=1e-3, ymax=.1,    
    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
] 
\addplot[color=blue,mark=*, semithick, mark options={solid,fill=markercolor}]
coordinates{(0,0.00462693)(0.0125,0.007073)(0.025,0.0116375)(0.0375,0.0166377)(0.05,0.021786)(0.0625,0.0270765)(0.075,0.0326103)(0.0875,0.0385547)(0.1,0.0451154)(0.1125,0.0525108)(0.125,0.0609548)};
\addplot[color=red,mark=*, dashed, semithick, mark options={solid,fill=markercolor}]
coordinates{(0,0.00458676)(0.0125,0.0055373)(0.025,0.00775311)(0.0375,0.0105429)(0.05,0.0136791)(0.0625,0.0171546)(0.075,0.0210896)(0.0875,0.0257146)(0.1,0.0313525)(0.1125,0.0383814)(0.125,0.0471834)};

\legend{Uniform knots, Smoothed knots}
\end{semilogyaxis}
\end{tikzpicture}
}
\subfloat[Second order formulation]{
\begin{tikzpicture}
\begin{semilogyaxis}[
width=.45\textwidth,
    xlabel={Warping factor $\alpha$},   
    ylabel={$L^2$ error},
    ymin=1e-3, ymax=.1,    
    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
] 
\addplot[color=blue,mark=*, semithick, mark options={solid,fill=markercolor}]
coordinates{(0,0.00504237)(0.0125,0.00834838)(0.025,0.0142055)(0.0375,0.0205381)(0.05,0.0269995)(0.0625,0.0335261)(0.075,0.0400686)(0.0875,0.0466709)(0.1,0.0536183)(0.1125,0.0616203)(0.125,0.0708213)};
\addplot[color=red,mark=*, dashed, semithick, mark options={solid,fill=markercolor}]
coordinates{(0,0.0113237)(0.0125,0.0118539)(0.025,0.013336)(0.0375,0.0157542)(0.05,0.0189697)(0.0625,0.0229092)(0.075,0.027875)(0.0875,0.0339693)(0.1,0.0406825)(0.1125,0.0473484)(0.125,0.0549665)};

\legend{Uniform knots, Smoothed knots}
\end{semilogyaxis}
\end{tikzpicture}
}
\caption{Comparison of $L^2$ errors for uniform and smooth knots as a function of the warping parameter $\alpha$ using $p = 3, K= 8$ splines on a single patch.}
\label{fig:warpingconverge}
\end{figure}

%\subsection{Spectral properties of $\bm{A}_h$ for the advection equation}
\subsection{Spectral properties of spline discretizations of the advection and first-order acoustic wave equation}

In the next few sections, we examine the spectral properties of the DG discretization matrix $\bm{A}_h$ for the \reviewerOne{one-dimensional} advection equation, such as the spectral radius (which can be used to estimate a maximum stable timestep) and discrete dispersion relations.  %Sections~\ref{sec:specadv} and \ref{sec:disp} focus on the advection equation, while Section~\ref{sec:eigerr} focuses on the second order discretization of the Laplacian.

%\subsubsection{Spectral radius of $\bm{A}_h$}
\subsubsection{Spectral radius of the discretization matrix}
\label{sec:specadv}
We begin by verifying that the growth of the spectral radius of the DG discretization matrix $\rho\LRp{\bm{A}_h}$ is $O(p/h)$.  Figure~\ref{fig:rhoA} shows the growth of the spectral radius for the advection equation (with weakly enforced periodic boundary conditions) using $\tau = 1/2$ and both uniform and smoothed knots.  We observe that the growth of the spectral radius matches very closely with $O(p/h)$, and constants are estimated in Table~\ref{table:rhoA}.  The use of smoothed knots reduces this growth by a factor slightly less than two.  The growth of $\rho\LRp{\bm{A}_h}$ for the acoustic wave equation is virtually identical, and we do not show it.  Similarly, the growth of $\rho\LRp{\bm{A}_h}$ for the non-dissipative case when $\tau = 0$ is very similar, and is also not shown.

\begin{figure}
\centering
\subfloat[Uniform knots]{
\begin{tikzpicture}
\begin{axis}[
width=.425\textwidth,
    xlabel={Number of elements $K$},   
    ylabel={$\rho\LRp{\bm{A}_h}$},       
    xmin=0, xmax=300,
    ymax=0,ymax=2100,    
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
] 
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,66.4194)(64,132.839)(96,199.258)(128,265.678)(160,332.097)(192,398.517)(224,464.936)(256,531.355)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,101.838)(64,203.677)(96,305.515)(128,407.354)(160,509.192)(192,611.031)(224,712.869)(256,814.707)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,141.497)(64,282.993)(96,424.49)(128,565.987)(160,707.484)(192,848.98)(224,990.477)(256,1131.97)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,184.614)(64,369.228)(96,553.842)(128,738.456)(160,923.07)(192,1107.68)(224,1292.3)(256,1476.91)};

\legend{$p = 2$, $p=3$, $p=4$, $p = 5$}
\end{axis}
\end{tikzpicture}
}
\subfloat[Smoothed knots]{
\begin{tikzpicture}
\begin{axis}[
width=.425\textwidth,
    xlabel={Number of elements $K$},   
    ylabel={$\rho\LRp{\bm{A}_h}$},   
    xmin=0, xmax=300,
    ymax=0,ymax=2100,    
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
]

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,55.5039)(64,110.088)(96,164.669)(128,219.248)(160,273.827)(192,328.405)(224,382.984)(256,437.563)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,75.8822)(64,149.237)(96,222.58)(128,295.92)(160,369.26)(192,442.599)(224,515.938)(256,589.276)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,97.8724)(64,190.915)(96,283.926)(128,376.931)(160,469.932)(192,562.932)(224,655.931)(256,748.929)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,121.298)(64,234.744)(96,348.126)(128,461.491)(160,574.85)(192,688.206)(224,801.56)(256,914.912)};

\legend{$p = 2$, $p=3$, $p=4$, $p = 5$}
\end{axis}
\end{tikzpicture}
}
\caption{Growth of $\rho\LRp{\bm{A}_h}$ for the advection equation using $\tau = .5$ and spline spaces of degree $p = 2, \ldots, 5$. }
\label{fig:rhoA}
\end{figure}

\begin{table}[!h]
\centering
\begin{tabular}{|c||c|c|c|c|}
\hline
& $p = 2$ & $p = 3$ & $p=4$ & $p=5$\\
\hhline{|=|=|=|=|=|}
Uniform knots & 2.0756 &   3.1824 &   4.4218 &   5.7692\\
\hline
Smoothed knots & 1.7092  &  2.3019 &   2.9255 &   3.5739\\
\hline
\end{tabular}
\caption{Estimated rate of growth of $\rho\LRp{\bm{A}_h}$ for the advection equation with respect to $1/h$.}
\label{table:rhoA}
\end{table}

The case when $\tau = 1$ is special.  When taking a uniform wavespeed and $\tau = 1$, the penalty flux is identical to the upwind flux derived from Riemann problems \cite{hesthaven2007nodal}.  For first order formulations using an exact upwind flux, the dependence of $\rho\LRp{\bm{A}_h}$ on $h$ changes significantly for $h$ small and $p \leq 5$.  Figure~\ref{fig:rhoAupwind} shows the growth of the spectral radius with the number of elements on a single patch for uniform knot vectors and $p = 2, \ldots, 7$.  The rate of growth of $\rho\LRp{\bm{A}_h}$ with $1/h$ does not change significantly for $p \leq 5$, increasing at the expected rate of $O(p/h)$ only as $p > 5$.   %The same behavior is observed when uniformly refining multiple spline patches.  
If smoothed knots are used in lieu of uniform knot vectors when $\tau = 1$, we observe that $\rho\LRp{\bm{A}_h}$ is significantly smaller in the pre-asymptotic range of $1/h$ and slightly larger in the asymptotic range of $1/h \rightarrow \infty$, though the rate of growth of the spectral radius is unchanged for $1/h$ sufficiently large.  When using smoothed knots and exact upwind fluxes, we also observe that the rate of growth $\rho\LRp{\bm{A}_h}$ with $1/h$ is slower than the estimated $O(p/h)$ rate up to degree 8.

This phenomena appears to be restricted to the specific case of \reviewerOne{$\tau=1$, where the penalty flux coincides with the upwind flux}, and the growth of $\rho\LRp{\bm{A}_h}$ returns to $O(p/h)$ when the penalty parameter deviates by more than $O(1\%)$ from $\tau = 1$.  \reviewerOne{However, numerical experiments also indicate that this phenomena is present for the acoustic wave equation, and persists when using multiple patches.  
%even discontinuous advection $a(x)$, provided that the upwind numerical flux is used
%\[
%\avg{a \phi} - \frac{1}{2}\LRp{|a|^+\phi^+ - |a|^-\phi^-}
%\]
Future work will study whether this phenomena is present for more complex settings (such as curvilinear meshes and discontinuous wavespeeds).}

\begin{figure}
\centering
\subfloat[Uniform knots]{
\begin{tikzpicture}
\begin{axis}[
width=.425\textwidth,
    xlabel={Number of elements $K$},   
        ylabel={$\rho\LRp{\bm{A}_h}$},       
    xmin=0, xmax=300,
    ymax=0,ymax=700,    
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
] 
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,32.9239)(64,66.339)(96,99.6489)(128,132.926)(160,166.191)(192,199.449)(224,232.703)(256,265.956)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,36.0829)(64,72.8362)(96,109.357)(128,145.911)(160,182.425)(192,218.937)(224,255.45)(256,291.953)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,51.525)(64,77.0896)(96,115.797)(128,154.475)(160,193.14)(192,231.798)(224,270.454)(256,309.109)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,69.1513)(64,91.7232)(96,120.356)(128,160.533)(160,200.771)(192,240.95)(224,281.151)(256,321.333)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,92.592)(64,121.947)(96,154.335)(128,198.112)(160,246.867)(192,296.187)(224,345.548)(256,394.912)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,112.952)(64,150.924)(96,193.087)(128,254.178)(160,317.703)(192,381.243)(224,444.784)(256,508.325)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,144.171)(64,189.195)(96,241.188)(128,311.773)(160,388.78)(192,466.469)(224,544.209)(256,621.953)};

%\addplot[color=black,dashed,semithick, mark options={solid,fill=markercolor}]
%coordinates{(32,32)(64,64)(96,96)(128,128)(160,160)(192,192)(224,224)(256,256)}; % O(2/h) line
\legend{$p = 2$, $p=3$, $p=4$, $p = 5$, $p = 6$, $p = 7$, $p=8$}
%\legend{$\frac{1}{h}$, $\frac{1.5}{h}$}
\end{axis}
\end{tikzpicture}
}
\subfloat[Smoothed knots]{
\begin{tikzpicture}
\begin{axis}[
width=.425\textwidth,
    xlabel={Number of elements $K$},   
        ylabel={$\rho\LRp{\bm{A}_h}$},       
    xmin=0, xmax=300,
    ymax=0,ymax=700,    
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
]

%\addplot[color=black,semithick, mark options={solid,fill=markercolor}]
%coordinates{(32,32)(64,64)(96,96)(128,128)(160,160)(192,192)(224,224)(256,256)};% [yshift=-3pt] node[below, pos=.95,color=black] {$y = 1/h$};% O(1/h) line
%\addplot[color=red,semithick, mark options={solid,fill=markercolor}]
%coordinates{(32,48)(64,96)(96,144)(128,192)(160,240)(192,288)(224,336)(256,384)}; %[yshift=3pt] node[above, pos=.95,color=black] {$y = 1.5/h$};
%\pgfplotsset{cycle list set=0}
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,33.8023)(64,67.3347)(96,100.665)(128,133.949)(160,167.218)(192,200.479)(224,233.737)(256,266.992)}; %p = 2
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,37.5425)(64,74.8178)(96,111.545)(128,148.129)(160,184.671)(192,221.196)(224,257.712)(256,294.223)}; % p = 3
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,40.4264)(64,80.3236)(96,119.264)(128,158.017)(160,196.716)(192,235.393)(224,274.059)(256,312.72)}; % p = 4
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,50.3593)(64,84.3661)(96,125.088)(128,165.445)(160,205.705)(192,245.927)(224,286.133)(256,326.328)}; % p = 5
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,64.6167)(64,86.4858)(96,129.798)(128,173.13)(160,216.469)(192,259.811)(224,303.155)(256,346.5)}; % p = 6
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,76.7086)(64,97.5475)(96,133.384)(128,177.856)(160,222.354)(192,266.858)(224,311.365)(256,355.876)}; %p = 7
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,94.4519)(64,116.577)(96,136.311)(128,181.68)(160,227.101)(192,272.535)(224,317.975)(256,363.42)};  % p =8
%\addplot[color=black,dashed,semithick, mark options={solid,fill=markercolor}]
%coordinates{(32,32)(64,64)(96,96)(128,128)(160,160)(192,192)(224,224)(256,256)}; % O(2/h) line
\legend{$p = 2$, $p=3$, $p=4$, $p = 5$, $p = 6$, $p = 7$, $p=8$}
%\legend{${1}/{h}$, ${1.5}/{h}$}
\end{axis}
\end{tikzpicture}
}
\caption{Growth of $\rho\LRp{\bm{A}_h}$ for the advection equation using an exact upwind flux ($\tau = 1$) and spline spaces of degree $p = 2, \ldots, 8$. }
\label{fig:rhoAupwind}
\end{figure}

\reviewerOne{A closer examination of the spectra of $\bm{A}_h$ offers some insight into the growth of $\rho(\bm{A}_h)$ and its dependence on $\tau$ and knot smoothing.  We note that the distribution of the spectra can also impact the maximum stable timestep, which depends on the choice of time-stepping method and its corresponding region of stability.  
Figure~\ref{fig:spectra} shows various spectra for the advection equation for a $p=7$, $K=32$ spline space.  It can be observed that $\rho(\bm{A}_h)$ is much larger for $\tau = 0, .5$ than for $\tau = 1$ due to the presence of two outlying eigenvalues with large imaginary part.  For $\tau = 1$, these extremal eigenvalues fall into a tight semi-circular distribution in the left half plane, resulting in a significant reduction in the spectral radius.  When knot smoothing is applied, the eigenvalues contract into a semi-circular distribution of slightly smaller radius.}
\begin{figure}
\centering
\subfloat[Uniform knots, $\tau = 0,.5,1$]{\includegraphics[height=.3\textheight]{spectraAdvec.png}}
\hspace{2em}
\subfloat[Uniform vs smoothed knots, $\tau = 1$]{\includegraphics[height=.3\textheight]{spectraAdvecSmoothed.png}}
\caption{Spectra for the advection equation using a spline space with $p=7$, $K = 32$. }
\label{fig:spectra}
\end{figure}

\subsubsection{Discrete dispersion relations }
\label{sec:disp}
We compute numerical dissipation and dispersion relations for the periodic advection equation using an upwind flux ($\tau = 1$).\footnote{\reviewerOne{For $\tau = 0$ (which corresponds to a central flux), dispersion errors are similar to the $\tau =1$ case for a small number of wavelengths per degree of freedom.  For a larger number of wavelengths per degree of freedom, dispersion errors are larger for $\tau = 0$ than for $\tau = 1$.  The qualitative behavior of the discrete dispersion relation is similar to that of polynomial DG discretizations \cite{ainsworth2004dispersive}.}}  Assuming a discrete solution of the form $u_h(x,t) = e^{i (kx - \omega_h t)}$, we seek a relation between the discrete frequency $k$ and wavenumber $\omega$.  Inserting this ansatz in the semi-discrete variational formulation for advection over a single patch yields a generalized eigenvalue problem, which can be solved for $\omega_h$ \cite{hesthaven2007nodal}.  While the exact relation between frequency and wavenumber is $\omega = k$, the discrete wavenumber $\omega_h$ will differ due to discretization errors.  Figure~\ref{fig:eigerr1} compares ${\rm Re}(\omega_h)$ (which corresponds to numerically introduced dispersion) with the exact relation $\omega = k$.

We note that increasing $p$ improves the rate of convergence of the dispersion error with the number of wavelengths per degree of freedom, while increasing $K$ mainly shifts the dispersion error curve \reviewerOne{by a constant number of wavelengths per degree of freedom}.  The dispersion error converges at a rate of $(2p+3)$ for $p$ even ($(2p+2)$ for $p$ odd) with respect to the number of wavelengths per degree of freedom \cite{ainsworth2004dispersive}.  As expected, since spline spaces contain polynomials of degree $p$, we observe this rate for both uniform and smoothed knots.  Both uniform and smoothed knot vectors result in comparable dispersion errors, though smoothed knot vectors result in slightly decreased dispersion errors at a higher number of wavelengths per degree of freedom.  We note that these observations are in line with the approximation results of Figure~\ref{fig:ppw}.

\begin{figure}
\centering
\subfloat[Uniform knots]{
\includegraphics[width=.45\textwidth]{dispersionUnif.png}
\label{subfig:disp1}
}
\subfloat[Smoothed knots]{
\includegraphics[width=.45\textwidth]{dispersionSmooth.png}
\label{subfig:disp2}
}\\
\subfloat[Dispersion errors (uniform knots)]{
\includegraphics[width=.45\textwidth]{dispersionErrUnif.png}
}
\subfloat[Dispersion errors (smoothed knots)]{
\includegraphics[width=.45\textwidth]{dispersionErrSmooth.png}
}
\caption{Numerical dispersion relations and dispersion error $\LRb{{\rm Re}(\omega) - {\rm Re}(\omega_h)}$ for spline spaces with $p = 4, K = 1, 4, 8, 16$ and $\tau = 1$.  The exact dispersion relation is given as a black dotted line in Figure~\ref{subfig:disp1}, \reviewerOne{ and the theoretical rate of convergence of the dispersion error (for $p=4$, this is $O(h^{11})$) is plotted as a black dotted line in Figure~\ref{subfig:disp2}}.  }
\label{fig:eigerr1}
\end{figure}

%\subsection{Eigenvalue and eigenvector errors for the second order Laplacian}
\subsection{Spectral properties of spline discretizations of the second-order acoustic wave equation}
\label{sec:eigerr}

\reviewerTwo{We now turn our attention to the second order formulation of the acoustic wave equation.  
Recall that the stable timestep restriction for the second order formulation is estimated by the square root of the spectral radius $\sqrt{\rho\LRp{\bm{A}_h}}$.  Figure~\ref{fig:rhoA2} shows the value of $\sqrt{\rho\LRp{\bm{A}_h}}$ for various $p$ and $K$, and it can be seen that the values of $\sqrt{\rho\LRp{\bm{A}_h}}$ for the second order formulation are slightly larger than the values of $\rho\LRp{\bm{A}_h}$ observed for the first order formulation in Figure~\ref{fig:rhoA}.  Table~\ref{table:rhoA2} also plots the estimated slope of the growth of $\sqrt{\rho\LRp{\bm{A}_h}}$ with respect to $1/h$.  The slopes are roughly $O(p)$, though the values of the estimated slopes are also slightly higher than those shown in Table~\ref{table:rhoA} for the first order formulation.
}

\begin{figure}
\centering
\subfloat[Uniform knots]{
\begin{tikzpicture}
\begin{axis}[
width=.425\textwidth,
    xlabel={Number of elements $K$},   
    ylabel={$\sqrt{\rho\LRp{\bm{A}_h}}$},       
    xmin=0, xmax=300,
    ymax=0,ymax=2100,    
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
] 
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,83.0595)(64,171.538)(96,259.969)(128,348.389)(160,436.804)(192,525.216)(224,613.628)(256,702.038)};

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,129.151)(64,267.997)(96,406.744)(128,545.468)(160,684.182)(192,822.892)(224,961.599)(256,1100.3)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,179.72)(64,374.676)(96,569.456)(128,764.195)(160,958.918)(192,1153.63)(224,1348.34)(256,1543.05)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,233.735)(64,489.525)(96,745.037)(128,1000.48)(160,1255.9)(192,1511.31)(224,1766.7)(256,2022.1)};

\legend{$p = 2$, $p=3$, $p=4$, $p = 5$}
\end{axis}
\end{tikzpicture}
}
\subfloat[Smoothed knots]{
\begin{tikzpicture}
\begin{axis}[
width=.425\textwidth,
    xlabel={Number of elements $K$},   
    ylabel={$\sqrt{\rho\LRp{\bm{A}_h}}$},   
    xmin=0, xmax=300,
    ymax=0,ymax=2100,    
    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
]

\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,66.9105)(64,138.075)(96,209.18)(128,280.27)(160,351.355)(192,422.436)(224,493.516)(256,564.595)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,91.7975)(64,190.07)(96,288.195)(128,386.284)(160,484.359)(192,582.426)(224,680.489)(256,778.55)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,117.432)(64,243.924)(96,370.121)(128,496.246)(160,622.341)(192,748.422)(224,874.495)(256,1000.56)};
\addplot+[semithick, mark options={solid,fill=markercolor}]
coordinates{(32,143.557)(64,299.146)(96,454.218)(128,609.163)(160,764.056)(192,918.925)(224,1073.78)(256,1228.62)};

\legend{$p = 2$, $p=3$, $p=4$, $p = 5$}
\end{axis}
\end{tikzpicture}
}
\caption{Growth of $\sqrt{\rho\LRp{\bm{A}_h}}$ for the second order wave equation using spline spaces of degree $p = 2, \ldots, 5$. }
\label{fig:rhoA2}
\end{figure}

\begin{table}[!h]
\centering
\begin{tabular}{|c||c|c|c|c|}
\hline
& $p = 2$ & $p = 3$ & $p=4$ & $p=5$\\
\hhline{|=|=|=|=|=|}
Uniform knots & 2.7632  &  4.3353  &  6.0859  &  7.9831\\
\hline
Smoothed knots & 2.2217     &3.0655     & 3.9419    &4.8429\\
\hline
\end{tabular}
\caption{Estimated rate of growth of $\sqrt{\rho\LRp{\bm{A}_h}}$ for the second order wave equation with respect to $1/h$.}
\label{table:rhoA2}
\end{table}

%

\reviewerTwo{We next examine eigenvalue and eigenvector errors for spline discretizations of the following one-dimensional generalized eigenproblem:
\[
-\frac{d^2p_k}{dx^2} = \lambda_k p_k
\]
over the unit interval $(0,1)$ subject to the boundary conditions $p_k(0) = p_k(1) = 0$.  The above eigenproblem admits an infinite number of eigenvalue and eigenvector solutions $\left(\lambda_k,p_k(x) \right) = \left(k^2\pi^2, \sin(k\pi x) \right)$ for $k \in \mathbb{N}$.  A finite number of discrete eigenvalues and eigenvectors $\left(\lambda_{h,k}, u_{h,k}(x)\right)$ with $0 < \lambda_{h,1} \leq \lambda_{h,2} \leq \ldots \leq \lambda_{h,N_{\rm dofs}}$ are attained from a standard Galerkin discretization on a single patch.  Homogeneous Dirichlet boundary conditions are imposed in a strong fashion as to allow for a more direct comparison to earlier works in the literature \cite{hughes2008duality, hughes2014finite}.  We have also examined the impact of obtaining boundary conditions in a weak manner and found that this did not significantly affect solution quality.  In prior work, it has been established that the error associated with a standard Galerkin discretization of the second order wave equation is directly tied to the eigenvalue and eigenvector errors associated with the above generalized eigenproblem \cite{hughes2014finite}, inspiring us to examine the errors associated with each of the eigenvalues and eigenvectors for a given discretization.}

%We now examine eigenvalue and eigenvector errors for the second order discretization of the Laplacian.  We compute discrete eigenvalues and eigenvectors $\lambda_{h,k}, u_{h,k}(x)$ from a standard Galerkin discretization on a single patch, where homogeneous Dirichlet boundary conditions are imposed in a strong fashion.  We note that the choice of strong imposition is to allow for a more direct comparison to earlier works in the literature \cite{hughes2008duality, hughes2014finite}, and that boundary conditions can also be imposed in a weak manner without significantly affecting these results.

Figure~\ref{fig:eigerr2} shows errors in eigenvalues $\LRb{\lambda_k-\lambda_{h,k}}$ { and eigenfunctions $\LRb{u_k-u_{h,k}}$} using splines with both uniform and smoothed knots { for the specific case of $p = 4$ and $K = 32$}.  We observe that smoothed knots are slightly less accurate for low $k$ and slightly more accurate for high $k$.  { Note furthermore that the last two eigenvalues are poorly approximated using uniform knots.  This is due to the fact that the last two discrete eigenvalues correspond to outlier frequencies \cite{hughes2008duality}.  The last two eigenvalues are slightly better approximated and smaller in magnitude for smoothed knots, and are the reason why the maximum stable timestep is larger with smoothed knots than with uniform knots.  
}
\begin{figure}
\centering
\subfloat[Eigenvalue errors]{
\includegraphics[width=.45\textwidth]{eigvals2ndorder.png}
}
\subfloat[Eigenvector errors]{
\includegraphics[width=.45\textwidth]{eigfunc2ndorder.png}
}
\caption{Eigenvalue errors $\LRb{\lambda_k-\lambda_{h,k}}$ and eigenvector errors $\nor{p_k-p_{h,k}}_{L^2}^2$ for an $p = 4, K = 32$ spline space using uniform and smoothed knots.  }
\label{fig:eigerr2}
\end{figure}

\subsection{A three-dimensional problem on a multi-patch geometry}

Finally, we show solutions to a model wave propagation problem using a non-trivial geometry.  Figure~\ref{fig:pipe3D} shows a multi-patch discretization of a three-dimensional pipe elbow, constructed using 12 different patches, as well as a first order DG simulation of acoustic wave propagation through the pipe.  We assume $c = 1$ and a zero initial condition, and impose a pulse velocity boundary condition at $x=2$
\[
u(\bm{x},0) = \begin{cases}
1-\cos(2\pi t /t_0), &t < t_0\\
0, &\text{ otherwise}
\end{cases}
\]
where $u$ is the velocity in the $x$-direction, and we take $t_0 = 2$.  We impose zero Neumann boundary conditions at all other boundary faces.  
%\note{Talk about pipe geometry and setup.  Show plots.}

\begin{figure}
\centering
\subfloat[Multi-patch model]{\includegraphics[width=.35\textwidth]{pipe.png}}
\subfloat[Pressure]{\includegraphics[width=.33\textwidth]{pipeDG.png}}
\subfloat[Pressure (cut view)]{\includegraphics[width=.33\textwidth]{pipeDG_cut.png}}
\caption{A multi-patch model of a 3D pipe elbow, and pressure solution at time $T = 4$. }
\label{fig:pipe3D}
\end{figure}

We discretize each patch using a degree $p=6$ spline basis (using a smoothed knot vector) and $K = 16$ elements in each coordinate direction.  We note that, due to the stretched nature of some of the patches, additional efficiency might be gained by using an anisotropic discretization with a varying number of elements in each coordinate direction, which we will explore in a future manuscript.

\section{Conclusions}
\label{sec:conc}

This work presents a strategy for applying NURBS-based finite element discretizations to transient hyperbolic problems using explicit time-stepping.  We utilize a multi-patch DG discretization, and apply a weight-adjusted approximation to the mass matrix inverse over each patch.  Additionally, the approximation is efficient to invert, involving only one-dimensional operations due to the tensor product structure in approximation of the mass matrix inverse.  The resulting methods are energy stable and high order accurate under assumptions on the regularity of the geometric mapping, and numerical experiments show that the approximate weight-adjusted mass matrix inverse delivers $L^2$ errors which are virtually identical to those using the full mass matrix inverse.

We also investigate the timestep restriction associated with NURBS-based discretizations, and show numerically that (for patches where $h$ is sufficiently small with respect to $p$) the maximum stable timestep decreases as $O(h/p)$ instead of the $O(h/p^2)$ associated with $C^0$ and DG finite element discretizations.  Finally, we investigate the use of smoothed knot vectors (which approximate $n$-width optimal knot vectors) in spline discretizations.  Numerical experiments show that spline spaces under smoothed knot vectors are slightly more accurate than splines with uniform knot vectors for approximations involving oscillatory functions or warped geometric mappings.  We remark that, while the use of smoothed knots does not result in drastic decreases in $L^2$ error for high frequencies or curved mappings, their improved accuracy and larger stable time-step make their use attractive for time-domain solvers based on explicit time integration.

We note that several areas related to this work remain to be explored.  For example, it is not immediately clear when splines can achieve a computational advantage over very high order polynomials.  Numerical experiments suggest that in certain situations (for example, on curved domains or approximating highly oscillatory functions) splines yield lower $L^2$ errors than polynomials for a similar number of degrees of freedom.  Additionally, for the same resolution, splines can take larger time steps than an equivalent finite element $p$-method, making it possible to achieve the same level of error with a reduced number of time steps, assuming a sufficiently accurate time stepping scheme.  However, these advantages must be balanced with the fact that, in order to integrate spline spaces \reviewerTwo{to} sufficient accuracy, the number of quadrature points per degree of freedom is larger for splines than for polynomials.  This additional cost can be significantly reduced through the use of optimal and near-optimal spline quadrature rules \cite{hughes2010efficient, auricchio2012simple} or by specialized techniques for the assembly and application of spline finite element matrices \cite{calabro2016fast}.  Finally, we will explore ways to fully remove any inverse scaling of the maximum stable timestep by $p$.  It may be advantageous to combine the nonlinear mapping used in \cite{hughes2008duality} with optimal or smoothed knot vectors.  Additionally, preliminary results in this work suggest that, when using smoothed knot vectors, a first order formulation combined with an appropriately chosen dissipative flux result in discretization matrices whose spectral radius is nearly independent of $p$ for $K$ sufficiently large.

The conformity requirements of multi-patch DG methods also remain to be explored.  In this work, all solution components are approximated using the same spline space; however, it is well-known that such a choice of approximation space can introduce problems such as locking or spurious modes to finite element discretizations of problems such as Maxwell's equations \cite{buffa2010isogeometric, ratnani2012arbitrary}.  DG methods sidestep these issues by using numerical fluxes to penalize or dissipate away non-conforming components of the solution \cite{Warburton20063205, chan2016short}.  However, for multi-patch DG methods, it may also be necessary to utilize locally conforming approximation spaces within each patch, which we will consider in future work.

\section{Acknowledgments}

The authors thank Joseph Benzaken help in constructing the three-dimensional pipe elbow geometry.  Jesse Chan is supported by the National Science Foundation under awards DMS-1719818 and DMS-1712639.  John A. Evans was partially supported by the Air Force Office of Scientific Research under Grant No. FA9550-14-1-0113.

\bibliographystyle{unsrt}
\bibliography{dgpenalty}

\end{document}

### 2018-06-20 docs/skewform/main.tex
removing old okl, adding okl kernel building

@@
 
-\documentclass[preprint,10pt]{elsarticle}
+%\documentclass[preprint,10pt]{elsarticle}
 %\documentclass[preprint,10pt]{article}
-%\documentclass[review]{siamart0216}
+\documentclass[review]{siamart0216}
 %\documentclass{siamart0216}
 
-\usepackage{fullpage}
-\usepackage[colorlinks=true]{hyperref}
+%\usepackage{fullpage}
+%\usepackage[colorlinks=true]{hyperref}
 
-\usepackage{amsmath,amssymb,amsfonts,amsthm}
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}
-\theoremstyle{lemma}
-\newtheorem{lemma}{Lemma}
-\theoremstyle{corollary}
-\newtheorem{corollary}{Corollary}
+\usepackage{amsmath,amssymb,amsfonts}
+%\usepackage{amsthm}
+%\theoremstyle{definition}
+%\newtheorem{definition}{Definition}
+%\theoremstyle{lemma}
+%\newtheorem{lemma}{Lemma}
+%\theoremstyle{corollary}
+%\newtheorem{corollary}{Corollary}
+%\theoremstyle{theorem}
+%\newtheorem{theorem}{Theorem}
 \newtheorem*{remark}{Remark}
-\theoremstyle{theorem}
-\newtheorem{theorem}{Theorem}
 \theoremstyle{assumption}
@@
 
-%\date{}
-%\author{Jesse Chan}
-%\title{On discretely entropy conservative discontinuous Galerkin methods}
+\date{}
+\author{Jesse Chan}
+\title{Skew-symmetric entropy stable discontinuous Galerkin formulations with application to hybrid meshes }
+
 \graphicspath{{./figs/}}
@@
 
-%\maketitle
+\maketitle
 
-\begin{frontmatter}
-\title{Skew-symmetric formulations of entropy stable discontinuous Galerkin methods with applications to hybrid and non-conforming meshes. }
-
-\author[rice]{Jesse Chan\corref{cor1}}
-\ead{Jesse.Chan@caam.rice.edu}
-\address[rice]{Department of Computational and Applied Mathematics, Rice University, 6100 Main St, Houston, TX, 77005}
-
-\begin{abstract}
-Entropy stable high order methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.  
-\end{abstract}
-\end{frontmatter}
+%\begin{frontmatter}
+%\title{Entropy stable skew-symmetric discontinuous Galerkin formulations with applications to hybrid and non-conforming meshes. }
+%
+%\author[rice]{Jesse Chan\corref{cor1}}
+%\ead{Jesse.Chan@caam.rice.edu}
+%\address[rice]{Department of Computational and Applied Mathematics, Rice University, 6100 Main St, Houston, TX, 77005}
+%
+%\begin{abstract}
+%Entropy stable high order methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.  
+%\end{abstract}
+%\end{frontmatter}
 
@@
 \section{Introduction}
+
+High order methods for the simulation of time-dependent compressible flow have the potential to achieve higher levels of accuracy at lower costs compared to current low order schemes \cite{wang2013high}.
+In addition to superior accuracy and efficiency on modern computing architectures, the low numerical dispersion and dissipation of high order methods \cite{ainsworth2004dispersive} enables the accurate propagation of waves over long distances and time scales.
+The same properties also make high order methods attractive for the resolution of unsteady phenomena such as vorticular and turbulent flows, which are sensitive to numerical dissipation \cite{visbal1999high, wang2013high}.
+However, while high order methods are often provably stable for wave problems, high order schemes for nonlinear conservation laws have consistently been hampered by problems of instability.
+
+When applied to nonlinear conservation laws, high order methods can experience artificial growth and blow-up near under-resolved features such as shocks, turbulence, or boundary layers.
+In practice, the application of high order methods to practical problems requires shock capturing and stabilization techniques (such as artificial viscosity) or solution regularization (such as filtering or limiting) to prevent solution blow-up.
+The resulting schemes for nonlinear conservation laws walk a fine line between stability, robustness, and accuracy.
+Aggressive stabilization or regularization can result in the loss of high order accuracy, while too little can result in instability \cite{wang2013high}.
+Moreover, it can be difficult to determine robust expressions for stabilization paramaters, as parameters which work for one simulation can fail when applied to a different physical regime or discretization setting.
+
+These issues have motivated the introduction of high order \textit{entropy stable} discretizations, which satisfy a semi-discrete entropy inequality while maintaining high order accuracy in smooth regions.
+Proofs of continuous entropy inequalities rely on the chain rule.
+In contrast, these discrete entropy inequalities account for the loss of the chain rule due to effects such as quadrature errors, which are incurred when applying polynomially exact quadrature rules to nonlinear and rational integrands within the DG formulation.
+These schemes were first introduced as high order collocation methods on tensor product elements in \cite{fisher2013high, carpenter2014entropy, gassner2016split, gassner2017br1}, and were extended to simplicial elements in \cite{crean2017high, chen2017entropy, crean2018entropy, chan2017discretely}.
+These methods have also been extended to staggered grid methods \cite{parsani2016entropy} and non-conforming meshes \cite{friedrich2017entropy}.
+Entropy stable boundary conditions have also been determined for the compressible Euler and Navier-Stokes equations \cite{parsani2015entropy, svard2018entropy}.
 
@@
 \begin{assumption}
-The volume quadrature rule is exact for polynomials of degree $(2N-1)$.
+The volume quadrature rule integrates exactly all polynomials up to degree $(2N-1)$.
 \label{ass:quad}
@@
 
-\note{TODO: Standardize hat notation for operators/normals.}
+\note{TODO: Standardize hat notation for operators/normals: can refer to operators on reference element and do $k$ superscript for physical elements, or do hats on reference element.}
 
@@
 \begin{align}
+\bm{Q}_i &= \bm{W}\bm{V}_q\bm{D}^i\bm{P}_q,\\
 \bm{Q}^i_N  &= \LRs{
@@
 -\frac{1}{2}\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q & \frac{1}{2}\bm{W}_f{\rm diag}({\bm{n}}_i)
-\end{array}}, \qquad \bm{Q}_i = \bm{W}\bm{V}_q\bm{D}^i\bm{P}_q,
+\end{array}}.
 \label{eq:QN}
@@
 \subsection{Loss of the SBP property for hybrid and non-conforming meshes}
+
 While the SBP property has been used to derive entropy stable schemes, the SBP property is difficult or impossible to enforce in certain discretization settings, such as hybrid and non-conforming meshes arises.
@@
 \begingroup
-\captionsetup[subfigure]{width=.43\textwidth}
+\captionsetup[subfigure]{width=.425\textwidth}
 \subfloat[Insufficiently accurate surface quadrature on the triangle element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D.png}\label{subfig:hybrid1}}
@@
 \begingroup
-\captionsetup[subfigure]{width=.425\textwidth}
-\subfloat[Coupling of nodes for a non-conforming interface (entropy stable but tighter coupling).]{\raisebox{0em}{\includegraphics[width=.425\textwidth]{figs/noncon.png}}\label{subfig:noncon2}}
+\captionsetup[subfigure]{width=.45\textwidth}
+\subfloat[Entropy stable inter-element coupling in \cite{friedrich2017entropy} for a non-conforming interface.]{\raisebox{0em}{\includegraphics[width=.45\textwidth]{figs/nonconSBP.png}}\label{subfig:noncon2}}
 \hspace{2em}
-\subfloat[Surface quadrature rule induced by non-conforming refinement (no SBP property).]{\raisebox{-.3em}{\includegraphics[width=.45\textwidth]{figs/nonconQuad.png}}\label{subfig:noncon1}}
+\subfloat[Entropy stable inter-element coupling in this work for a non-conforming interface.
+The dotted black lines denote communication between neighboring elements.]{\raisebox{-0em}{\includegraphics[width=.45\textwidth]{figs/nonconQuad.png}}\label{subfig:noncon1}}
 \endgroup
@@
 Each dashed red line indicates a dyadic flux computation required between two nodes.
-The coupling terms in \cite{friedrich2017entropy} requires computing dyadic fluxes between \emph{each} pair of nodes on adjacent interfaces, while the proposed approach requires dyadic flux computations only between immediately adjacent nodes.
+} %The coupling terms in \cite{friedrich2017entropy} requires computing dyadic fluxes between \emph{each} pair of nodes on adjacent interfaces.
+The new approach results in a simpler communication pattern between elements.
 }
@@
 \begin{align*}
-\int_{\partial \hat{D}} (\Pi_N u) v \hat{n}_i -   \int_{\hat{D}} \LRp{\Pi_N u} \pd{v}{\hat{x}_i} &= \bm{v}_f^T\bm{W}_f{\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q\bm{u}_q - \LRp{\bm{V}_q\bm{D}^i\bm{P}_q\bm{v}_q}^T\bm{W} \bm{V}_q\bm{P}_q\bm{u}_q\\
-&= \bm{v}_q^T \LRp{\bm{V}_f\bm{P}_q}^T\bm{W}_f{\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q\bm{u}_q - \bm{v}_q^T \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T \bm{W} \bm{V}_q\bm{P}_q\bm{u}_q,
+&\int_{\partial \hat{D}} (\Pi_N u) v \hat{n}_i -   \int_{\hat{D}} \LRp{\Pi_N u} \pd{v}{\hat{x}_i} =\\
+ &\bm{v}_f^T\bm{W}_f{\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q\bm{u}_q - \LRp{\bm{V}_q\bm{D}^i\bm{P}_q\bm{v}_q}^T\bm{W} \bm{V}_q\bm{P}_q\bm{u}_q\\
+=& \bm{v}_q^T \LRp{\bm{V}_f\bm{P}_q}^T\bm{W}_f{\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q\bm{u}_q - \bm{v}_q^T \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T \bm{W} \bm{V}_q\bm{P}_q\bm{u}_q,
 \end{align*}
@@
 \begin{align*}
-\bm{v}^T\bm{Q}^i_N\bm{u} &= -\bm{v}_q^T\bm{Q}^T_i \bm{u}_q + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}} \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}} \bm{u}_f\\
+\bm{v}^T\bm{Q}^i_N\bm{u} =& -\bm{v}_q^T\bm{Q}^T_i \bm{u}_q + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}} \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}} \bm{u}_f\\
 &\qquad - \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}} \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} - \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}} \bm{u}_f + \bm{v}_f^T\bm{W}_f\hat{\bm{n}} \bm{u}_f\\
-&= \begin{pmatrix} \bm{v}_q\\ \bm{v}_f\end{pmatrix}^T 
-\LRp{\begin{pmatrix}
+=& \begin{pmatrix} \bm{v}_q\\ \bm{v}_f\end{pmatrix}^T 
+\left(\begin{pmatrix}
 \bm{0}& \\
 & \bm{W}_f\diag{\hat{\bm{n}}}
-\end{pmatrix}
-+
-\begin{pmatrix}
+\end{pmatrix}\right.\\
+&+
+\left.\begin{pmatrix}
 -\bm{Q}_i^T + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q}^T\bm{W}_f\hat{\bm{n}} \bm{V}_f\bm{P}_q & -\frac{1}{2} \LRp{\bm{W}_f\hat{\bm{n}} \bm{V}_f\bm{P}_q}^T\\
 \frac{1}{2}\bm{W}_f\hat{\bm{n}} \bm{V}_f\bm{P}_q & -\frac{1}{2}\bm{W}_f\hat{\bm{n}}
-\end{pmatrix} } 
+\end{pmatrix}  \right)
 \begin{pmatrix} \bm{u}_q\\ \bm{u}_f\end{pmatrix}\\
-&= \bm{v}^T\LRp{\bm{B}_N - \LRp{\bm{Q}^i_N}^T}\bm{u}.
+=& \bm{v}^T\LRp{\bm{B}_N - \LRp{\bm{Q}^i_N}^T}\bm{u}.
 \end{align*}
@@
 \]
-where $\bm{u}$ is an arbitrary vector of appropriate size.  
+where $\bm{u}$ is a vector of values of some function at volume and surface quadrature points. %is an arbitrary vector of appropriate size.  
 \end{corollary}
@@
 The proof that $\bm{Q}^i_N \bm{1} = \bm{0}$ follows from the property that polynomials are equal to their $L^2$ projection, and is identical to that of \cite{chan2017discretely,chan2018discretely}.
-The proof of the first equality follows $\bm{Q}^i_N \bm{1} = \bm{0}$ and Lemma~\ref{lemma:vdsbp} with $M=0$
+The proof of the first equality follows from $\bm{Q}^i_N \bm{1} = \bm{0}$ and Lemma~\ref{lemma:vdsbp} with $M=0$
 \[
@@
 \end{gather}
-If instead of the third condition, $\bm{f}_S$ satisfies a \emph{stability} property $\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}_S(\bm{u}_L,\bm{u}_R) \leq \psi(\bm{u}_L) - \psi(\bm{u}_R)$, the flux $\bm{f}_S$ is referred to as \textit{entropy stable}.
+%If instead of the third condition, $\bm{f}_S$ satisfies a \emph{stability} property 
+%\[
+%\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}_S(\bm{u}_L,\bm{u}_R) \leq \psi(\bm{u}_L) - \psi(\bm{u}_R)$, the flux $\bm{f}_S$ is referred to as \textit{entropy stable}.
 
 We begin by deriving a skew-symmetric formulation on the reference element $\hat{D}$, and showing it is entropy conservative.
-This formulation can then be extended in a straightforward manner to affine elements.
+This formulation is then made entropy stable by adding interface dissipation, and can extended in a straightforward manner to affine elements.
 The skew-symmetric formulation for (\ref{eq:nonlineqs}) on $\hat{D}$ is
@@
 \begin{gather}
-\bm{M}\pd{\bm{u}}{t} + \sum_{i=1}^d \LRs{\begin{array}{c}
-\bm{V}_q \\ \bm{V}_f\end{array}}^T \begin{pmatrix}
-\bm{W}&\\
-&\bm{W}_f
-\end{pmatrix}
-\LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} +  \bm{V}_f^T\bm{W}_f\diag{\hat{\bm{n}}}\bm{f}_i^*\label{eq:esdgSkewWeak}\\
-= \bm{M}\pd{\bm{u}}{t} + \sum_{i=1}^d\LRs{\begin{array}{c}
+%\bm{M}\pd{\bm{u}}{t} + \sum_{i=1}^d \LRs{\begin{array}{c}
+%\bm{V}_q \\ \bm{V}_f\end{array}}^T \begin{pmatrix}
+%\bm{W}&\\
+%&\bm{W}_f
+%\end{pmatrix}
+%\LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} +  \bm{V}_f^T\bm{W}_f\diag{\hat{\bm{n}}}\bm{f}_i^*\label{eq:esdgSkewWeak}\\
+%= 
+\bm{M}\pd{\bm{u}}{t} + \sum_{i=1}^d\LRs{\begin{array}{c}
 \bm{V}_q \\ \bm{V}_f\end{array}}^T 
-\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{V}_f^T\bm{W}_f \diag{\hat{\bm{n}}}\bm{f}_i^* = 0.  \nonumber
+\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{V}_f^T\bm{W}_f \diag{\hat{\bm{n}}}\bm{f}_i^* = 0.  \label{eq:esdgSkewWeak}
 \end{gather}
+where the matrix $\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}$ possesses the following structure
+\[
+\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} = \begin{pmatrix}
+\bm{Q}_i-\bm{Q}_i^T & \LRp{\bm{V}_f \bm{P}_q}^T\bm{W}_f\diag{\bm{n}_i}\\
+\bm{W}_f\diag{\bm{n}_i}{\bm{V}_f \bm{P}_q} & \bm{0}
+\end{pmatrix}.
+\]
 
@@
 \begin{theorem}
-The formulation (\ref{eq:esdgSkew}) is entropy conservative for any surface quadrature which is exact for polynomials of degree $N$, such that
+Let the surface quadrature be exact for polynomials of degree $N$.  Then, the formulation (\ref{eq:esdgSkew}) is entropy conservative such that
 \[
@@
 &= \bm{1}^T\LRp{\bm{Q}^i_N}\psi_i(\tilde{\bm{u}}) - \psi_i(\tilde{\bm{u}})^T\LRp{\bm{Q}^i_N}\bm{1} \\
-&= \bm{1}^T\LRp{\bm{Q}^i_N}\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i_N\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}} \psi_i(\tilde{\bm{u}}_f)
+&= \bm{1}^T\LRp{\bm{Q}^i_N}\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i_N\psi_i(\tilde{\bm{u}}) \\
+&= \bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}} \psi_i(\tilde{\bm{u}}_f)
 \end{align*}
@@
 \end{align}
-Because $\bm{F}_S$ is symmetric and $\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}$ is skew-symmetric, the term $\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}$ is a skew-symmetric matrix.
-The volume term vanishes
+Because $\bm{F}_S$ is symmetric and $\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}$ is skew-symmetric, the term 
 \[
-\bm{1}^T\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} = 0,
+\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}
 \]
-where we have used that, for any skew symmetric matrix $\bm{A}$, $\bm{x}^T\bm{A}\bm{x} = 0$.
+is a skew-symmetric matrix.  Using that $\bm{A}$, $\bm{x}^T\bm{A}\bm{x} = 0$ for any skew symmetric matrix $\bm{A}$, the volume term vanishes
+\[
+\bm{1}^T\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} = 0.
+\]
 % and that the Hadamard product is commutable $\LRp{\bm{A}\circ\bm{B} }^T = \bm{A}^T\circ\bm{B}^T$
@@
 
+Next, we extend this formulation to affinely mapped elements.
 Derivatives with respect to physical coordinates on $D^k$ are computed in terms of a change of variables formula and derivatives with respect to reference coordinates
@@
 \pd{\bm{u}}{t} + \sum_{i=1}^d \LRs{\begin{array}{cc}
-\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{\hat{\bm{n}}}\bm{f}_i^* = 0, \label{eq:skewform}\\
+\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{{\bm{n}_i}}\bm{f}_i^* = 0, \label{eq:skewform}\\
 \LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q, \nonumber\\
@@
 The skew-symmetric formulation (\ref{eq:skewform}) can be extended to curvilinear meshes using the definition of $\bm{D}^i_N$ in (\ref{eq:dncurved}).
-However, two additional assumptions must be satisfied in order to prove that the resulting formulation is entropy stable.
+However, additional assumptions must be satisfied in order to prove that the resulting formulation is entropy stable.
 
@@
 For degree $N$ isoparametric mappings, the GCL is automatically satisfied in two dimensions due to the fact that the exact geometric terms $\bm{G}^k_{ij}\in P^{N-1}$ \cite{kopriva2006metric}.
-However, in three dimensions, the GCL is not automatically preserved due to the fact that $\bm{G}^k_{ij}\in P^{2N-2}$ and cannot be represented exactly using degree $N$ polynomials.
-Thus, (\ref{eq:dgcl}) must be enforced through an alternative construction of $\bm{G}^k_{ij}$.
+However, in three dimensions, the GCL is not automatically preserved due to the fact that $\bm{G}^k_{ij}\in P^{2N-2}$, and thus cannot be represented exactly using degree $N$ polynomials.
+As a consequence, (\ref{eq:dgcl}) must be enforced through an alternative construction of $\bm{G}^k_{ij}$.
 
-The most common approach is to rewrite the geometric terms as the curl of some quantity $\bm{r}^i$, but to interpolate $\bm{r}^i$ before applying the curl \cite{visbal2002use, kopriva2006metric, hindenlang2012explicit}:
+A common approach is to rewrite the geometric terms as the curl of some quantity $\bm{r}^i$, but to interpolate $\bm{r}^i$ before applying the curl \cite{visbal2002use, kopriva2006metric, hindenlang2012explicit}:
 \begin{align}
@@
 For the latter term in the sum, Lemma~\ref{lemma:vdsbp} holds and
-\begin{equation}
-\sum_{j=1}^d\bm{1}^T\hat{\bm{Q}}^j_N \diag{\bm{G}_{ij}} \bm{u} = 
-\sum_{j=1}^d\bm{1}^T\LRp{\hat{\bm{B}}^j_N - \LRp{\hat{\bm{Q}}^j_N}^T} \diag{\bm{G}_{ij}} \bm{u} = \sum_{j=1}^d\bm{1}^T\hat{\bm{B}}^j_N\diag{\bm{G}_{ij}}\bm{u}, 
+\begin{align}
+\sum_{j=1}^d\bm{1}^T\hat{\bm{Q}}^j_N \diag{\bm{G}_{ij}} \bm{u} &= 
+\sum_{j=1}^d\bm{1}^T\LRp{\hat{\bm{B}}^j_N - \LRp{\hat{\bm{Q}}^j_N}^T} \diag{\bm{G}_{ij}} \bm{u} \\
+&= \sum_{j=1}^d\bm{1}^T\hat{\bm{B}}^j_N\diag{\bm{G}_{ij}}\bm{u}, \nonumber
 \label{eq:vsbp1}
-\end{equation}
+\end{align}
 where we have used that $\hat{\bm{Q}}^j_N \bm{1} = \bm{0}$.  For the former term in the sum, we expand out 
@@
 &+ \frac{1}{2} \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q+\bm{G}^f_{ij}}^T\frac{1}{2}\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{u}_f\\
-=& \LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}}_i - \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q}^T{\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q}}\bm{u}_q + \LRp{\bm{G}^f_{ij}}^T\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{u}_f.
-%\\
+=& \LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}}_i - \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q}^T{\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q}}\bm{u}_q \\
+&+ \LRp{\bm{G}^f_{ij}}^T\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{u}_f. %\\
 %&= \LRp{\bm{G}^q_{ij}}^T \LRp{ \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q}^T{\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q}}\bm{u}_q + \LRp{\bm{G}^f_{ij}}^T\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{u}_f 
@@
 where we have used in the final step that $\sum_{j=1}^d\bm{G}_{ij}\hat{n}_j = \bm{n}_i J^k_f$ \cite{ciarlet1978finite, chan2018discretely}
-
 \end{proof}
 
-Ensuring that Lemma~\ref{lemma:vdsbpcurved} holds to take $R$ as large as possible, which maximizes the degree to which geometric terms are approximated.  However,
+Lemma~\ref{lemma:vdsbpcurved} implies that the polynomial degree $R$ of the surface geometric terms must be compatible with the accuracy of the surface quadrature.
+This, in turn, will depend on the way in which geometric terms are approximated.
+\note{Finish.}
 
-%\begin{theorem}
-%Let the geometric terms be approximated using (\ref{eq:iconscurl}) with $1\leq R \leq (N+1)$.
-Then, the skew-symmetric formulation is stable on non-conforming meshes using any composite surface quadrature which is exact for degree $(N+R-1)$ polynomials on $\partial \hat{D}$.
-%\end{theorem}
-%\begin{proof}
-%
-%\note{Trick here is to note effect of geometric mapping: surface quadrature must be tied to degree of geometric factors.
-For Kopriva geofacs where $J\bm{G}_{ij}\in P^{N-1}$, it suffices to take surface quadrature to be exact for degree $(2N-1)$ polynomials.
-For general geofacs $J\bm{G}_{ij}\in V^N$, surface quadrature which is exact for degree $2N$ polynomials must be used.
-Note that this is still more flexible than the strong form: you can still use different surface quadratures on each face.
-}
-%
 %\end{proof}
-Thus, if geometric terms are approximated using degree $N$ polynomials as in \cite{kopriva2006metric}, then the skew-symmetric formulation with composite GLL quadrature can be used
+%Thus, if geometric terms are approximated using degree $N$ polynomials as in \cite{kopriva2006metric}, then the skew-symmetric formulation with composite GLL quadrature can be used.
 
@@
 %\subsection{On implementation and computational cost}
+
+\note{Kopriva trick to enforcing GCL still results in $Q^N$ normals 3D.
+May not satisfy requirement that $\bm{1}^T\bm{Q}^i_N \psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i_N \psi_i(\tilde{\bm{u}})$ because of quadrature inaccuracy.
+Note - this still works for conforming hexes because of the full matrix SBP property.}
 
@@
 
-\appendix
-
-\section{An explicit skew-symmetric entropy stable formulation for DG-SEM}
-
-\note{$\bm{V}_q,\bm{P}_q = \bm{I}$, while $\bm{V}_f$ reduces to a permutation matrix.  }
+%\appendix
+%
+%\section{An explicit skew-symmetric entropy stable formulation for DG-SEM} 
+%
+%\note{$\bm{V}_q,\bm{P}_q = \bm{I}$, while $\bm{V}_f$ reduces to a permutation matrix.  }
+%
+%\note{For non-conforming faces, need to interpolate entropy variables to form fluxes.  Given this info, can rotate from skew form to strong form using mortar variables.  }
 

### 2018-06-20 talks/ECCM2018/iga/main.tex
updating ECCM talks
removing old okl, adding okl kernel building

--- 2017-07-30 talks/USNCCM17/main.tex
@@
 %\useoutertheme[subsection=false]{smoothbars}
+\usepackage{fancybox}
 
@@
 \newcommand{\fnh}{\widehat{f}_n}
+\renewcommand{\hat}{\widehat}
 \renewcommand{\L}{L^2\LRp{\Omega}}
@@
 
-\title[Time-domain IGA]{Time-domain multi-patch\\discontinuous Galerkin methods }
-\date[May 19, 2017]{USNCCM 2017\\July 18, 2017}
+\title[Multi-patch IGA]{Efficient explicit solvers for multipatch discontinuous Galerkin isogeometric analysis}
+\date[June 9-15, 2018]{ECCM-ECFD 2018\\June 9-15, 2018}
 \author[Chan, Evans]{Jesse Chan\inst{1}, J. A. Evans\inst{2}}
@@
 \frame{
-\frametitle{High order Galerkin methods for hyperbolic PDEs}
+\frametitle{High order methods for hyperbolic PDEs}
 \setcounter{subfigure}{0}
@@
 \begin{itemize}
-\item Model problem: (second order) acoustic wave equation 
-\begin{align*}
-%\only<1>{
-\frac{1}{c^2}\pdd{p}{t}{} - \Delta p &= f.
-%}
-%\only<2>{
-%\pd{v}{t}{} &= \sum_{D^k}\LRp{\Grad p,\grad w}_{D^k} + \LRa{\jump{p},w}\\
-%\pd{p}{t}{} &= v.
-%}
-\end{align*}
-\item Pressure-velocity system (first order formulation)
+\item Wave propagation problems (acoustics, Maxwells, elasticity).
+\vspace{.1em}
+%\begin{align*}
+%%\only<1>{
+%\frac{1}{c^2}\pdd{p}{t}{} - \Delta p &= f.
+%%}
+%%\only<2>{
+%%\pd{v}{t}{} &= \sum_{D^k}\LRp{\Grad p,\grad w}_{D^k} + \LRa{\jump{p},w}\\
+%%\pd{p}{t}{} &= v.
+%%}
+%\end{align*}
+\item Model problem: acoustic wave equation (both first and second order formulations)
 \begin{align*}
@@
 \end{align*}
+%\item Results extend to second order formulation. 
+%\item Multiple geometric patches,\\IGA on each patch.  
 \end{itemize}
@@
 \begin{itemize}
-\item Weak patch coupling through DG numerical flux (SIPG, upwind).  
-\end{itemize}
-
-\let\thefootnote\relax\footnotetext{\tiny Langer et al 2014. Multipatch discontinuous Galerkin isogeometric analysis.}
-\let\thefootnote\relax\footnotetext{\tiny Wilcox et al 2010. A high-order DG method for wave propagation through coupled elastic-acoustic media.}
-}
-
-\frame{
-\frametitle{Method of lines multi-patch DG discretization}
-\vspace{-1em}
+\item Multiple geometric patches, weak patch coupling through DG-like numerical interface flux (SIPG, upwind).  
+\end{itemize}
+
+\let\thefootnote\relax\footnotetext{\tiny Langer et al (2014). \textit{Multipatch discontinuous Galerkin isogeometric analysis}.}
+\let\thefootnote\relax\footnotetext{\tiny Wilcox et al (2010). \textit{A high-order DG method for wave propagation through coupled elastic-acoustic media}.}
+}
+
+\frame{
+\frametitle{Explicit solvers for multi-patch DG-IGA}
+%\vspace{-1em}
 \begin{itemize}
@@
 \]
+\vspace{.125em}
 \item Global mass matrix $\bm{M}_h$ is (patch) block diagonal.  
-\vspace{1em}
-\item Tailored patch discretizations: explicit time-stepping, curved domains. 
-%\vspace{.25em}
-\begin{itemize}
-\item Efficient mass matrix inversion over each patch.
+\vspace{1.5em}
+\item Challenges and questions: 
 \vspace{.5em}
-\item High order accuracy, energy stability on curved geometries.
+\begin{itemize}
+\item<2-> How to efficiently invert 3D patch mass matrices while guaranteeing accuracy and \textcolor{red}{stability} (especially for explicit methods)?  
 \vspace{.5em}
-\item Improved stable timestep restriction (CFL).
-\end{itemize}
-\end{itemize}
-}
-
-\section{Spline spaces and optimal knot vectors }
+\item<3-> Do splines/IGA offer advantages over $C^0$-FEM/DG for explicit solvers?  
+\vspace{.5em}
+\item<4-> Can we tailor IGA discretizations towards explicit time-stepping?
+\end{itemize}
+
+%\vspace{1em}
+%\item Explicit time-stepping, curved domains. 
+%%\vspace{.25em}
+%\begin{itemize}
+%\item Efficient mass matrix inversion over each patch.
+%\vspace{.5em}
+%\item High order accuracy, energy stability on curved geometries.
+%\vspace{.5em}
+%\item Improved stable timestep restriction (CFL).
+%\end{itemize}
+\end{itemize}
+}
+
+\section{Inverting IGA mass matrices }
 \frame[noframenumbering]{
@@
 
-%\frame{
-%\frametitle{Spline spaces and sup-infs}
-%
-%\begin{itemize}
-%\item Let $X$ be Banach, and $X_n$ be some $n$-dimensional subspace of $X$
-%\item 
-%\item The $n$-width of $X$ is then defined to be the infimum over all subspaces $X_n$ with ${\rm dim}\LRp{X_n} = n$, and a specific subspace $X_n$ is referred to as an \textit{optimal} subspace if its sup-inf attains the $n$-width.
-%\item \note{FINISH}
-%\end{itemize}
-%}
-\frame{
-\frametitle{B-spline basics}
-
-\begin{itemize}
-\item Standard 1D B-splines: $B^0_i(x) = \mathbbm{1}_{\xi_i \leq x \leq \xi_{i+1}}$, 
-\[
-B^k_i(x) = \frac{x - \xi_i}{\xi_{i+p}-\xi_i}B^{k-1}_i(x) + \frac{\xi_{i+p+1}-x}{\xi_{i+p+1}-\xi_{i+1}}B^{k-1}_{i+1}(x).
-\]
-\item Assume patches are mapped quads/cubes: tensor product basis
-\[
-B^p_{ijk}(x,y,z) = B^p_{i}(x)B^p_{j}(y)B^p_{k}(z).
-\]
+\frame{
+\frametitle{B-splines: assumptions for this talk}
+
+\begin{itemize}
+\item Reference coordinates $\hat{\bm{x}} \in \hat{D}$, physical coordinates $\bm{x} \in D^k$.  
+\vspace{.5em}
+\item Standard 1D B-splines: $B^0_i(\hat{x}) = \mathbbm{1}_{\xi_i \leq \hat{x} \leq \xi_{i+1}}$, 
+\[
+B^k_i(\hat{x}) = \frac{\hat{x} - \xi_i}{\xi_{i+p}-\xi_i}B^{k-1}_i(\hat{x}) + \frac{\xi_{i+p+1}-\hat{x}}{\xi_{i+p+1}-\xi_{i+1}}B^{k-1}_{i+1}(\hat{x}).
+\]
+\item Physical basis: mapping of tensor product basis on reference domain
+%\begin{overlayarea}{\textwidth}{.17\textheight}
+\[
+%\only<1>{
+B^p_{ijk}(\hat{\bm{x}}) = B^p_{i}(\hat{x})B^p_{j}(\hat{y})B^p_{k}(\hat{z}).%, \qquad \text{(mapped B-splines)}}
+%\only<2>{B^p_{ijk}({\bm{x}}) = \frac{B^p_{i}(\hat{x})B^p_{j}(\hat{y})B^p_{k}(\hat{z})}{w_R(\hat{\bm{x}})}, \qquad \text{(NURBS)}}
+\]
+%\end{overlayarea}
 \item Assume maximally continuous, open knot vectors
@@
 }
+
+\frame{
+\frametitle{IGA mass matrices: stability/accuracy vs efficiency}
+
+\begin{itemize}
+\item Energy stability: if \note{$\bm{u}^T\bm{A}_h\bm{u} \leq 0$}, semi-discrete solution won't blow up
+\[
+\bm{M}_h\td{\bm{u}}{t} = \bm{A}_h\bm{u} \quad \Longrightarrow \quad \td{}{t}\LRp{\bm{u}^T\bm{M}_h \bm{u}} =  \frac{1}{2}\pd{}{t}{}\nor{u}_{L^2}^2 \leq 0.
+\]
+\item Approximating $\bm{M}_h^{-1}$ impacts semi-discrete \textcolor{red}{stability} and \textcolor{red}{accuracy}.  
+\vspace{2em}
+\item<2-> Curved patch mass matrices $\bm{M}_J$: tensor product basis $B^p_{ijk}(\hat{\bm{x}})$
+\begin{overlayarea}{\textwidth}{.185\textheight}
+\[
+\only<1-2>{\LRp{\bm{M}_J}_{ijk,lmn} = \int_{\widehat{D}} B^p_{ijk}(\widehat{\bm{x}})B^p_{lmn}(\widehat{\bm{x}}) \textcolor{red}{J(\widehat{\bm{x}})}\diff{\widehat{\bm{x}}}, \qquad \text{(B-splines)}}
+\only<3>{\LRp{\bm{M}_J}_{ijk,lmn} = \int_{\widehat{D}} \frac{B^p_{ijk}(\widehat{\bm{x}})}{\textcolor{red}{w_R\LRp{\hat{\bm{x}}}}}\frac{B^p_{lmn}(\widehat{\bm{x}})}{\textcolor{red}{w_R\LRp{\hat{\bm{x}}}}} \textcolor{red}{{J(\widehat{\bm{x}})}}\diff{\widehat{\bm{x}}} \qquad \text{(NURBS)}}
+\only<4>{\LRp{\bm{M}_J}_{ijk,lmn} = \int_{\widehat{D}} B^p_{ijk}(\widehat{\bm{x}})B^p_{lmn}(\widehat{\bm{x}}) \textcolor{red}{\frac{J(\widehat{\bm{x}})}{w^2_R\LRp{\hat{\bm{x}}}}}\diff{\widehat{\bm{x}}} \qquad \text{(NURBS)}}
+\]
+\end{overlayarea}
+\item<2-> No Kronecker structure due to $J(\widehat{\bm{x}})$: $\bm{M}_h$ \textcolor{red}{expensive} to invert in 3D!
+
+\end{itemize}
+}
+
+\frame{
+\frametitle{Approximate mass matrix inversion}
+
+\begin{itemize}
+\item Mass lumping: loss of high order accuracy for IGA.
+\vspace{.75em}
+\item Preconditioning: 
+\vspace{.35em}
+\begin{itemize}
+\item Additional cost and complexity for a time-domain code.
+\vspace{.35em}
+\item Semi-discrete stability requires approximation of $\bm{M}_J^{-1}$ to induce a norm on $\bm{u}$ (e.g.\ a fixed symmetric positive-definite linear operator).
+\vspace{.35em}
+\item Example: Krylov methods approximate $\bm{M}_J^{-1}$ as a non-linear operator!
+\end{itemize}
+\vspace{.75em}
+\item Isogeometric collocation: restores tensor product structure, but semi-discrete stability is more difficult to prove.  
+\end{itemize}
+
+\let\thefootnote\relax\footnotetext{\tiny Gao and Calo (2014). \textit{Fast isogeometric solvers for explicit dynamics}.}
+\let\thefootnote\relax\footnotetext{\tiny Evans, Hiemstra, Hughes, Reali (2017).  \textit{Explicit higher-order accurate IG collocation methods for structural dynamics}.}
+\let\thefootnote\relax\footnotetext{\tiny Wathen and Rees (2009). \textit{Chebyshev semi-iteration in preconditioning for problems including the mass matrix}.}
+\let\thefootnote\relax\footnotetext{\tiny Auricchio et al (2012). \textit{Isogeometric collocation for elastostatics and explicit dynamics}.}
+}
+
+\frame{
+\frametitle{An energy stable and efficient approximation to $\bm{M}^{-1}$}
+
+\begin{itemize}
+\item Replace $\bm{M}_J$ with symmetric pos-def ``weight-adjusted'' approximation:
+\[
+%\td{}{t}
+ \bm{M}_J\bm{u} \Rightarrow %\td{}{t}
+ \textcolor{red}{\widehat{\bm{M}}\bm{M}_{1/J}^{-1}\widehat{\bm{M}}} \bm{u}, \qquad \LRp{\widehat{\bm{M}}}_{ijk,lmn} = \int_{\widehat{D}} B^p_{ijk}(\widehat{\bm{x}})B^p_{lmn}(\widehat{\bm{x}})\diff{\widehat{\bm{x}}}.
+\]
+%\vspace{.5em}
+\item Weight-adjusted inverse: Kronecker product, matrix-free eval.\ of $\bm{M}_{1/J}$
+\begin{align*}
+\bm{M}^{-1}_J &\approx \LRp{\widehat{\bm{M}}\bm{M}_{1/J}^{-1}\widehat{\bm{M}}}^{-1}  = \widehat{\bm{M}}^{-1} \bm{M}_{1/J} \widehat{\bm{M}}^{-1}\\
+\widehat{\bm{M}}^{-1} &= \widehat{\bm{M}}^{-1}_{\rm 1D} \otimes \widehat{\bm{M}}^{-1}_{\rm 1D} \otimes \widehat{\bm{M}}^{-1}_{\rm 1D}.
+\end{align*}
+\item Energy stability with respect to an \note{equivalent} norm 
+\[
+C_1(J) \nor{\bm{u}}_{\widehat{\bm{M}}\bm{M}_{1/J}^{-1}\widehat{\bm{M}}} \leq \nor{\bm{u}}_{\bm{M}_J} \leq C_2 \nor{\bm{u}}_{\widehat{\bm{M}}\bm{M}_{1/J}^{-1}\widehat{\bm{M}}}.
+\]
+\end{itemize}
+
+\let\thefootnote\relax\footnotetext{\tiny Chan, Hewett, Warburton (2016).  \textit{Weight-adjusted DG methods: wave prop.\ in heterogeneous media}. }
+\let\thefootnote\relax\footnotetext{\tiny Chan, Hewett, Warburton (2016).  \textit{Weight-adjusted DG methods: curvilinear meshes}. }
+}
+
+\frame{
+\frametitle{Accuracy: weighted vs weight-adjusted mass matrix}
+
+\begin{figure}
+\centering
+\only<1>{
+\subfloat[Warped mesh]{\includegraphics[width=.425\textwidth]{figs/mapped.png}}
+\subfloat[Error for acoustics (first order form)]{
+\begin{tikzpicture}
+\begin{loglogaxis}[
+    legend cell align=left,
+    legend style={legend pos=south east, font=\tiny},
+    width=.5\textwidth,
+     ymin=1e-17, ymax=2,    
+    xlabel={Mesh size $h$},
+    ylabel={$L^2$ error}, 
+    grid style=dashed,
+] 
+\addplot[color=blue,mark=*,semithick, mark options={solid,fill=markercolor}]
+coordinates{(0.5,0.154255)(0.25,0.0280139)(0.125,0.000775805)(0.0625,7.51259e-06)};
+\addplot[color=red,mark=x,semithick, mark options={solid,fill=markercolor}]
+coordinates{(0.5,0.154523)(0.25,0.0280342)(0.125,0.000775107)(0.0625,7.51217e-06)};
+\addplot[color=black,mark=triangle*,semithick, mark options={solid,fill=markercolor}]
+coordinates{(0.5,7.56052e-06)(0.25,1.32834e-07)(0.125,1.51285e-11)(0.0625,1.69165e-17)};
+
+\legend{$L^2$ projection,Weight-adjusted,Difference}
+%\legend{Uniform, Optimal, Smoothed}
+\end{loglogaxis}
+\end{tikzpicture}
+}
+%\subfloat[$L^2$ errors ]{
+%\begin{tikzpicture}
+%\begin{loglogaxis}[
+%    legend cell align=left,
+%    legend style={legend pos=south east, font=\tiny},
+%    width=.5\textwidth,
+%    xlabel={Mesh size $h$},
+%    ylabel={$L^2$ error}, 
+%             ymin=1e-10, ymax=2,    
+%    grid style=dashed,
+%] 
+%\addplot[color=blue,mark=*,semithick, mark options={solid,fill=markercolor}]
+%coordinates{(0.5,0.0110908)(0.25,0.00138298)(0.125,2.29018e-05)(0.0625,3.18561e-07)};
+%\addplot[color=red,mark=x,semithick, mark options={solid,fill=markercolor}]
+%coordinates{(0.5,0.0111346)(0.25,0.00138582)(0.125,2.30061e-05)(0.0625,3.18565e-07)};
+%\addplot[color=black,mark=triangle*,semithick, mark options={solid,fill=markercolor}]
+%coordinates{(0.5,0.000986363)(0.25,8.87776e-05)(0.125,2.18816e-06)(0.0625,1.65013e-09)};
+%\legend{$L^2$ projection,Weight-adjusted,Difference}
+%%\legend{Uniform, Optimal, Smoothed}
+%\end{loglogaxis}
+%\end{tikzpicture}}
+}
+\only<2>{
+\subfloat[Warped mesh]{\includegraphics[width=.425\textwidth]{figs/mapped2.png}}
+\subfloat[Error for acoustics (first order form)]{
+\begin{tikzpicture}
+\begin{loglogaxis}[
+    legend cell align=left,
+    legend style={legend pos=south east, font=\tiny},
+    width=.5\textwidth,
+     ymin=1e-17, ymax=2,    
+    xlabel={Mesh size $h$},
+    ylabel={$L^2$ error}, 
+    grid style=dashed,
+] 
+\addplot[color=blue,mark=*,semithick, mark options={solid,fill=markercolor}]
+coordinates{(0.5,0.338603)(0.25,0.143169)(0.125,0.0234825)(0.0625,0.00033695)};
+\addplot[color=red,mark=x,semithick, mark options={solid,fill=markercolor}]
+coordinates{(0.5,0.388935)(0.25,0.146233)(0.125,0.0235623)(0.0625,0.000336824)};
+\addplot[color=black,mark=triangle*,semithick, mark options={solid,fill=markercolor}]
+coordinates{(0.5,0.00977253)(0.25,0.000187168)(0.125,3.16763e-07)(0.0625,9.44863e-12)};
+
+\legend{$L^2$ projection,Weight-adjusted,Difference}
+%\legend{Uniform, Optimal, Smoothed}
+\end{loglogaxis}
+\end{tikzpicture}
+}
+%\subfloat[$L^2$ errors ]{
+%\begin{tikzpicture}
+%\begin{loglogaxis}[
+%    legend cell align=left,
+%    legend style={legend pos=south east, font=\tiny},
+%    width=.5\textwidth,
+%     ymin=1e-10, ymax=2,    
+%    xlabel={Mesh size $h$},
+%    ylabel={$L^2$ error}, 
+%    grid style=dashed,
+%] 
+%\addplot[color=blue,mark=*,semithick, mark options={solid,fill=markercolor}]
+%coordinates{(0.5,0.0444376)(0.25,0.0144725)(0.125,0.000702933)(0.0625,6.79061e-06)};
+%\addplot[color=red,mark=x,semithick, mark options={solid,fill=markercolor}]
+%coordinates{(0.5,0.1182)(0.25,0.017419)(0.125,0.000745669)(0.0625,6.79513e-06)};
+%\addplot[color=black,mark=triangle*,semithick, mark options={solid,fill=markercolor}]
+%coordinates{(0.5,0.109529)(0.25,0.00969371)(0.125,0.000248811)(0.0625,2.47881e-07)};
+%\legend{$L^2$ projection,Weight-adjusted,Difference}
+%%\legend{Uniform, Optimal, Smoothed}
+%\end{loglogaxis}
+%\end{tikzpicture}}
+}
+\caption{$L^2$ errors for the acoustic wave equation using weighted and weight-adjusted mass matrices for tensor product $p = 4$ splines. }
+
+\end{figure}
+}
+
+\frame{
+\frametitle{Accuracy of the weight-adjusted mass matrix}
+
+\begin{itemize}
+\item Difference between the weighted $L^2$ and weight-adjusted inner products is high order accurate: for $v(\bm{x})$ of degree $q$, 
+\begin{align*}
+&\LRb{ \bm{v}^T\bm{M}_J\bm{u} - \bm{v}^T\widehat{\bm{M}}\bm{M}_{1/J}^{-1}\widehat{\bm{M}}\bm{u}}\\
+&\leq C_{J} \nor{J}_{W^{p+1,\infty}\LRp{D^k}} h^{2p+2-q} \nor{u}_{W^{p+1,2}\LRp{D^k}}.
+\end{align*}
+\vspace{.5em}
+\item Difference between $L^2$ and weight-adjusted projection is \textcolor{red}{$O(h^{p+2})$}!
+\begin{align*}
+\nor{P_h u - \tilde{P}_h u}_{L^2\LRp{D^k}} &\lesssim \nor{\frac{1}{\sqrt{J}}}_{L^{\infty}}^2 \nor{J}_{W^{p+1,\infty}\LRp{D^k}} h^{p+2}\nor{u}_{W^{p+1,2}\LRp{D^k}}.
+\end{align*}
+\end{itemize}
+
+\let\thefootnote\relax\footnotetext{\tiny Chan, Wilcox (2018). \textit{On discretely entropy stable weight-adjusted discontinuous Galerkin methods: curvilinear meshes}.}
+}
+
+\section{Why splines vs $C^0$-FEM?  CFL restrictions}
+
+\frame[noframenumbering]{
+\tableofcontents[currentsection]
+}
+
+\frame{
+\frametitle{Estimating the CFL restriction}
+
+\begin{itemize}
+%\item Maximum stable timestep (CFL) depends on $h, p$, and wavespeed.
+\item For explicit time-stepping method: estimate $dt \propto \frac{1}{\max\LRb{\lambda_j}}$
+\[
+\bm{M}_h \bm{v} = \lambda \bm{A}_h\bm{v}.
+\]
+\item Bendixon-Hirsch lemma: bound ${\rm Re}\LRp{\lambda_j}, {\rm Im}\LRp{\lambda_j}$ using the symmetric and skew-symmetric parts  of $\bm{A}_h$
+\[
+\underbrace{\frac{1}{2}\LRp{\bm{A}_h+\bm{A}_h^T}}_{\bm{A}_{\rm sym}}+ \underbrace{\frac{1}{2}\LRp{\bm{A}_h-\bm{A}_h^T}}_{\bm{A}_{\rm skew}}, \qquad
+\begin{array}{c}
+\LRb{{\rm Re}\LRp{\lambda_j}} \leq \rho\LRp{\bm{M}^{-1}_h\bm{A}_{\rm sym}},\\
+\LRb{{\rm Im}\LRp{\lambda_j}} \leq \rho\LRp{\bm{M}^{-1}_h\bm{A}_{\rm skew}}.
+\end{array}
+\]
+\item $\rho\LRp{\bm{M}_h^{-1}\bm{A}_{\rm sym}}, \rho\LRp{\bm{M}_h^{-1}\bm{A}_{\rm skew}}$: generalized Rayleigh quotients
+\[
+\rho\LRp{\bm{M}_h^{-1}\bm{A}_{\rm sym}} = \frac{\bm{u}^T\bm{A}_{\rm sym}\bm{u}}{\bm{u}^T\bm{M}_h\bm{u}}, \qquad 
+\rho\LRp{\bm{M}_h^{-1}\bm{A}_{\rm skew}} = \frac{\LRb{\bm{u}^* (i\bm{A}_{\rm skew})\bm{u}}}{\bm{u}^*\bm{M}_h\bm{u}}
+\]
+\end{itemize}
+}
+
+\frame{
+\frametitle{CFL: constants in trace and inverse inequalities}
+\begin{itemize}
+\item For hyperbolic problems (advection, acoustics), can bound
+\begin{align*}
+&\frac{\bm{u}^T\bm{A}_{\rm sym}\bm{u}}{\bm{u}^T\bm{M}_h\bm{u}} \lesssim \frac{\nor{u}_{L^2\LRp{\partial D^k}}^2}{\nor{u}_{L^2\LRp{D^k}}^2} \lesssim \frac{C_T}{h}, \\
+&\frac{\LRb{\bm{u}^* (i\bm{A}_{\rm skew})\bm{u}}}{\bm{u}^*\bm{M}_h\bm{u}} \lesssim \frac{\nor{\Grad u}_{L^2\LRp{D^k}}}{\nor{u}_{L^2\LRp{D^k}}} \lesssim \frac{C_I}{h}.
+\end{align*}
+\item $C_T, C_I$: \note{$p$-dependent constants} in trace, inverse inequalities
+\[
+\nor{\Grad u}_{L^2\LRp{\hat{D}}} \leq C_I \nor{u}_{L^2\LRp{\hat{D}}} , \qquad \nor{u}_{L^2\LRp{\partial \hat{D}}}^2 \leq C_T \nor{u}_{L^2\LRp{\hat{D}}}^2.
+\]
+\item Summary: $dt \propto \frac{1}{\max\LRb{\lambda_j}} \leq \frac{h}{\max{C_T, C_I}}$.
+What do $C_T, C_I$ look like?
+Can compute $C_T, C_I$ using a generalized eigenvalue problem.
+%For $C^0$-FEM and DG, $dt \leq O\LRp{\frac{h}{p^2}}$.
+\end{itemize}
+}
+
+\frame{
+\frametitle{Trace and inverse inequality constants: $C^0$-FEM vs splines}
+\vspace{-.5em}
+\begin{figure}
+\centering
+\only<1>{\includegraphics[width=\textwidth]{figs/Definitions.pdf}
+\caption{A parametric patch has $K$ elements per side, while a physical patch has size $H$.  The mesh resolution is $h = H/K$.}}
+\only<2->{
+\setcounter{subfigure}{0}
+\subfloat[Inverse constants, $K = 2p$]{
+\begin{tikzpicture}
+\begin{axis}[
+width=.45\textwidth,
+    xlabel={Degree of approximation $p$},   
+%    ylabel={$C_I / K$},
+    xmin=.5, xmax=12.5,
+    ymax=0,ymax=90,    
+    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
+    xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
+]
+
+\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
+coordinates{(1,2)(2,4.5)(3,8)(4,12.5)(5,18)(6,24.5)(7,32)(8,40.5)(9,50)(10,60.5)(11,72)(12,84.5)};
+\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
+coordinates{(1,2)(2,3.42857)(3,5.27022)(4,7.3491)(5,9.61122)(6,12.0258)(7,14.5716)(8,17.233)(9,19.9977)(10,22.8559)(11,25.7995)(12,28.8217)};
+%\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
+%coordinates{(1,2)(2,3.15356)(3,4.41024)(4,5.72791)(5,7.09105)(6,8.48919)(7,9.91299)(8,11.3583)(9,12.8221)(10,14.3019)(11,15.796)(12,17.3028)};
+
+\legend{Polynomial $C_I$, Spline $C_I/K$}%, Smoothed $C_I/K$ }
+\end{axis}
+\end{tikzpicture}
+}
+\hspace{1em}
+\subfloat[Trace constants, $K = 2p$]{
+\begin{tikzpicture}
+\begin{axis}[
+width=.45\textwidth,
+    xlabel={Degree of approximation $p$},   
+    xmin=.5, xmax=12.5,
+    ymax=0,ymax=65,    
+    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
+    xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
+]
+
+\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
+coordinates{(1,1.73205)(2,3.87298)(3,6.5216)(4,9.74981)(5,13.5914)(6,18.0596)(7,23.1598)(8,28.894)(9,35.2634)(10,42.2685)(11,49.9096)(12,58.187)};
+\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
+coordinates{(1,1.73205)(2,2.43668)(3,3.48657)(4,4.75661)(5,6.16327)(6,7.6751)(7,9.27477)(8,10.9506)(9,12.694)(10,14.4983)(11,16.3578)(12,18.2682)};
+%\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
+%coordinates{(1,1.73205)(2,2.32019)(3,2.9916)(4,3.77273)(5,4.61001)(6,5.4806)(7,6.37331)(8,7.28317)(9,8.20701)(10,9.14265)(11,10.0885)(12,11.0432)};
+
+\legend{Polynomial $C_T$, Spline $C_T/K$}%, Smoothed $C_T/K$ }
+\end{axis}
+\end{tikzpicture}
+}
+\caption*{Polynomial constants are $O(p^2)$, observed spline constants \textcolor{red}{$O(p)$} for $K \geq O(1/p)$.}
+\let\thefootnote\relax\footnotetext{\tiny Takacs, Takacs (2016). \textit{Approximation error estimates and inverse inequalities for B-splines of maximum smoothness}.}
+}
+\end{figure}
+\only<2->{\uncover<3->{
+\begin{center}
+\ovalbox{\textcolor{red}{$dt \propto \frac{h}{p^2}$} for $C^0$-FEM and DG, \textcolor{red}{$dt \propto \frac{h}{p}$} for IGA!}
+\end{center}}}
+}
+
+\section{Optimizing splines for explicit solvers}
+\frame[noframenumbering]{
+\tableofcontents[currentsection]
+}
+
 \frame{
@@
 
-\let\thefootnote\relax\footnotetext{\tiny Melkman and Micchelli 1978. Spline spaces are optimal for $L^2$ $n$-width.}
+\let\thefootnote\relax\footnotetext{\tiny Melkman and Micchelli (1978). \textit{Spline spaces are optimal for $L^2$ $n$-width}.}
 }
@@
 }
-\only<3>{
-\setcounter{subfigure}{0}
-\subfloat{
-\begin{tikzpicture}
-\begin{axis}[
-width=.425\textwidth,
-    xlabel={Number of elements $K$},   
-        ylabel={Max knot error},
-    xmin=2, xmax=66,
-    ymax=0,ymax=0.0125,    
-    legend pos=north east, legend cell align=left, legend style={font=\tiny},	
-       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
-]
-
-\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
-coordinates{(4,0.0121)(8,0.0075)(16,0.0046)(32,0.0023)(64,0.0012)};
-
-\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
-coordinates{(4,0.009)(8,0.0041)(16,0.0028)(32,0.002)(64,0.0012)};
-
-\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
-coordinates{(4,0.0016)(8,0.0101)(16,0.0074)(32,0.0048)(64,0.0028)};
-
-\legend{$p = 2$, $p=3$, $p=4$}
-\end{axis}
-\end{tikzpicture}
-}
-\hspace{2em}
-\subfloat{
-\begin{tikzpicture}
-\begin{axis}[
-width=.425\textwidth,
-    xlabel={Number of elements $K$},   
-    ylabel={Max Greville error},
-    xmin=2, xmax=66,
-    ymax=0,ymax=0.0125,    
-    legend pos=north east, legend cell align=left, legend style={font=\tiny},	
-       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
-]
-
-\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
-coordinates{(4,0.006)(8,0.0038)(16,0.0024)(32,0.0011)(64,0.00058806)};
-
-\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
-coordinates{(4,0.003)(8,0.002)(16,0.0018)(32,0.0012)(64,0.00067857)};
-
-\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
-coordinates{(4,0.000401)(8,0.0025)(16,0.0016)(32,0.00089352)(64,0.00063901)};
-
-\legend{$p = 2$, $p=3$, $p=4$}
-\end{axis}
-\end{tikzpicture}
-}
-}
+%\only<3>{
+%\setcounter{subfigure}{0}
+%\subfloat{
+%\begin{tikzpicture}
+%\begin{axis}[
+%width=.425\textwidth,
+%    xlabel={Number of elements $K$},   
+%        ylabel={Max knot error},
+%    xmin=2, xmax=66,
+%    ymax=0,ymax=0.0125,    
+%    legend pos=north east, legend cell align=left, legend style={font=\tiny},	
+%       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
+%] 
+%
+%\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
+%coordinates{(4,0.0121)(8,0.0075)(16,0.0046)(32,0.0023)(64,0.0012)};
+%
+%\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
+%coordinates{(4,0.009)(8,0.0041)(16,0.0028)(32,0.002)(64,0.0012)};
+%
+%\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
+%coordinates{(4,0.0016)(8,0.0101)(16,0.0074)(32,0.0048)(64,0.0028)};
+%
+%\legend{$p = 2$, $p=3$, $p=4$}
+%\end{axis}
+%\end{tikzpicture}
+%}
+%\hspace{2em}
+%\subfloat{
+%\begin{tikzpicture}
+%\begin{axis}[
+%width=.425\textwidth,
+%    xlabel={Number of elements $K$},   
+%    ylabel={Max Greville error},
+%    xmin=2, xmax=66,
+%    ymax=0,ymax=0.0125,    
+%    legend pos=north east, legend cell align=left, legend style={font=\tiny},	
+%       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
+%] 
+%
+%\addplot[color=blue,mark=*,semithick, mark options={fill=markercolor}]
+%coordinates{(4,0.006)(8,0.0038)(16,0.0024)(32,0.0011)(64,0.00058806)};
+%
+%\addplot[color=red,mark=square*,semithick, mark options={fill=markercolor}]
+%coordinates{(4,0.003)(8,0.002)(16,0.0018)(32,0.0012)(64,0.00067857)};
+%
+%\addplot[color=black,mark=triangle*,semithick, mark options={fill=markercolor}]
+%coordinates{(4,0.000401)(8,0.0025)(16,0.0016)(32,0.00089352)(64,0.00063901)};
+%
+%\legend{$p = 2$, $p=3$, $p=4$}
+%\end{axis}
+%\end{tikzpicture}
+%}
+%}
 \end{overlayarea}
@@
 \frame{
-\frametitle{Approximation properties: oscillatory functions}
+\frametitle{Approximation properties in 1D: oscillatory functions}
 \setcounter{subfigure}{0}
@@
 \centering
-\begin{overlayarea}{\textwidth}{.7\textheight}
-\only<1>{
-\subfloat[$L^2$ error vs $k/N_{\rm dofs}$ ($\cos\LRp{\frac{k\pi x}{2}}$)]{\includegraphics[width=.49\textwidth]{figs/splineapprox.png}}
-\hspace{.25em}
-\subfloat[Semi-log scale]{\includegraphics[width=.49\textwidth]{figs/splineapproxlog.png}}
-%\caption{}
-}
-\only<2>{
 \subfloat[First order formulation]{
@@
 \begin{semilogyaxis}[
-width=.45\textwidth,
+width=.475\textwidth,
     xlabel={Wavelength $k$},   
@@
 }
-%\caption{Comparison of $L^2$ errors for uniform and smooth knots as a function of wavenumber $k$ using $p = 3, K= 8$ splines.}
-}
-\end{overlayarea}
+\caption{$L^2$ errors for 1D acoustics using uniform and smoothed knot vectors: smoothed knots emphasize high frequencies over low frequencies. }
 \end{figure}
-\vspace{-3em}
-\begin{itemize}
-\item<1-> $L^2$ best approximations with smoothed knots (fixed $p,K$): decreased high frequency errors, increased low frequency errors.   
-\item<2-> Similar errors approximating solution $\cos\LRp{\frac{k\pi x}{2}}\cos\LRp{\frac{k\pi t}{2}}$ ($k$ odd).  
-\end{itemize}
-
-}
-
-\frame{
-\frametitle{Spectral properties of $\bm{A}_h$}
-\setcounter{subfigure}{0}
-\vspace{.5em}
-\only<2>{
-Constant 1D advection: numerical dispersion error $\LRb{{\rm Re}(\omega) - {\rm Re}(\omega_h)}$
-\[
-\pd{u}{t}{} + \pd{u}{x}{} = 0, \qquad u(x,t) = e^{i(kx - \omega t)}, \quad u_h(x,t) = e^{i(kx-\omega_h t)}.
-\]
-\vspace{-2em}
-\begin{figure}
-\centering
-\subfloat[Dispersion errors (uniform knots)]{
-\includegraphics[width=.45\textwidth]{figs/dispersionErrUnif.png}
-}
-\subfloat[Dispersion errors (smoothed knots)]{
-\includegraphics[width=.45\textwidth]{figs/dispersionErrSmooth.png}
-}
-\caption{Dispersion errors using full upwinding, $p = 4$, and $K = 1, 4, 8$. } %The exact dispersion relation is given as a black dotted line.}% in Figures~\ref{subfig:disp1}, \ref{subfig:disp2}.  }
-%\label{fig:eigerr1}
-\end{figure}
-
-}
-\only<1>{
-Eigenvalues, eigenfunctions of 1D Laplacian: uniform vs smoothed knots.
-\[
-\LRb{\lambda_k-\lambda_{h,k}}, \qquad \nor{w_k(x)-w_{h,k}(x)}_{L^2}^2
-\]
-\vspace{-1.5em}
+}
+
+%\frame{
+%\frametitle{Spectral properties of $\bm{A}_h$}
+%\setcounter{subfigure}{0}
+%\vspace{.5em}
+%\only<2>{
+%Constant 1D advection: numerical dispersion error $\LRb{{\rm Re}(\omega) - {\rm Re}(\omega_h)}$
 %\[
-%-\pd{u}{x}{2} = \lambda u, \qquad \LRp{\Grad u, \Grad v}_{\Omega} = \lambda_h \LRp{u,v}_{\Omega}, \quad \forall v\in V_h.
+%\pd{u}{t}{} + \pd{u}{x}{} = 0, \qquad u(x,t) = e^{i(kx - \omega t)}, \quad u_h(x,t) = e^{i(kx-\omega_h t)}.
 %\]
-\begin{figure}
-\centering
-\subfloat[Eigenvalue errors]{
-\includegraphics[width=.45\textwidth]{figs/eigvals2ndorder.png}
-}
-\subfloat[Eigenvector errors]{
-\includegraphics[width=.45\textwidth]{figs/eigfunc2ndorder.png}
-}
-\caption{Eigenvalue and eigenvector errors for $p = 4, K = 32$ splines. }
-\end{figure}
-}
-}
-
-\section{Curved domains and weight-adjusted mass matrices}
-\frame[noframenumbering]{
-\tableofcontents[currentsection]
-}
-
-\frame{
-\frametitle{Approximation properties: curvilinear domains}
+%\vspace{-2em}
+%\begin{figure}
+%\centering
+%\subfloat[Dispersion errors (uniform knots)]{
+%\includegraphics[width=.45\textwidth]{figs/dispersionErrUnif.png}
+%}
+%\subfloat[Dispersion errors (smoothed knots)]{
+%\includegraphics[width=.45\textwidth]{figs/dispersionErrSmooth.png}
+%}
+%\caption{Dispersion errors using full upwinding, $p = 4$, and $K = 1, 4, 8$. } %The exact dispersion relation is given as a black dotted line.}% in Figures~\ref{subfig:disp1}, \ref{subfig:disp2}.  }
+%%\label{fig:eigerr1}
+%\end{figure}
+%
+%}
+%\only<1>{
+%Eigenvalues, eigenfunctions of 1D Laplacian: uniform vs smoothed knots.
+%\[
+%\LRb{\lambda_k-\lambda_{h,k}}, \qquad \nor{w_k(x)-w_{h,k}(x)}_{L^2}^2
+%\]
+%\vspace{-1.5em}
+%%\[
+%%-\pd{u}{x}{2} = \lambda u, \qquad \LRp{\Grad u, \Grad v}_{\Omega} = \lambda_h \LRp{u,v}_{\Omega}, \quad \forall v\in V_h.
+%%\]
+%\begin{figure}
+%\centering
+%\subfloat[Eigenvalue errors]{
+%\includegraphics[width=.45\textwidth]{figs/eigvals2ndorder.png}
+%}
+%\subfloat[Eigenvector errors]{
+%\includegraphics[width=.45\textwidth]{figs/eigfunc2ndorder.png}
+%}
+%\caption{Eigenvalue and eigenvector errors for $p = 4, K = 32$ splines. }
+%\end{figure}
+%}
+%}
+
+\frame{
+\frametitle{Approximation properties in 2D/3D: curvilinear domains}
 
@@
 \centering
-\subfloat[Warped mesh, $\alpha = 1/8$]{\includegraphics[width=.51\textwidth]{figs/mapped.png}}
+\subfloat[Warped mesh, $\alpha = 1/8$]{\includegraphics[width=.375\textwidth]{figs/mapped.png}}
 \subfloat[$L^2$ errors ($\alpha = 1/64$) ]{
@@
     legend style={legend pos=north east, font=\tiny},
-    width=.48\textwidth,
+    width=.49\textwidth,
     xlabel={Degrees of freedom},
@@
 }
-\caption{$L^2$ errors for increasing $\alpha$ ($p = 3, K= 8$ splines on a single patch).}
+\caption{$L^2$ errors  w.r.t.\ curved warping for 2D acoustics ($p = 3, K= 8$ splines).}
 \label{fig:warpingconverge}
@@
 \frame{
-\frametitle{Approximate mass matrix inversion}
-
-\begin{itemize}
-\item Patch mass matrices $\bm{M}_J$: tensor product basis but no Kronecker structure due to $J(\widehat{\bm{x}})$, and mass lumping is inaccurate.
-\[
-\LRp{\bm{M}_J}_{ijk,lmn} = \int_{\widehat{D}} B^p_{ijk}(\widehat{\bm{x}})B^p_{lmn}(\widehat{\bm{x}}) \textcolor{red}{J(\widehat{\bm{x}})}\diff{\widehat{\bm{x}}}.
-\]
-%\vspace{.25em}
-\item Preconditioning: approximating $\bm{M}_J^{-1}$ impacts semi-discrete stability.  Note - Krylov methods approximate $\bm{M}_J^{-1}$ as a non-linear operator!
-\[
-\frac{1}{2}\td{}{t}\nor{u}_{L^2}^2 = \bm{u}^T\bm{M}_J\td{\bm{u}}{t} \leq 0.%= \bm{u}^T\bm{A_h}\bm{u} \leq 0.
-\]
-%\vspace{.25em}
-\item Isogeometric collocation: restores tensor product structure, but semi-discrete stability is more difficult to prove.  
-\end{itemize}
-
-\let\thefootnote\relax\footnotetext{\tiny Gao and Calo 2014. Fast isogeometric solvers for explicit dynamics.}
-%\let\thefootnote\relax\footnotetext{\tiny Evans et al 2017.  Explicit Isogeometric Collocation Methods for Structural Dynamics by Means of Predictor-Multicorrector Schemes}
-\let\thefootnote\relax\footnotetext{\tiny Wathen and Rees 2009. Chebyshev semi-iteration in preconditioning for problems including the mass matrix.}
-\let\thefootnote\relax\footnotetext{\tiny Auricchio et al 2012. Isogeometric collocation for elastostatics and explicit dynamics.}
-}
-
-\frame{
-\frametitle{Restoring Kronecker structure to $\bm{M}^{-1}$}
-
-\begin{itemize}
-\item Replace $\bm{M}_J$ with ``weight-adjusted'' approximation:
-\[
-\td{}{t} \bm{M}_J\bm{u} \Rightarrow \td{}{t}\textcolor{red}{\widehat{\bm{M}}\bm{M}_{1/J}^{-1}\widehat{\bm{M}}} \bm{u}, \qquad \LRp{\widehat{\bm{M}}}_{ijk,lmn} = \int_{\widehat{D}} B^p_{ijk}(\widehat{\bm{x}})B^p_{lmn}(\widehat{\bm{x}})\diff{\widehat{\bm{x}}}.
-\]
-%\vspace{.5em}
-\item Weight-adjusted inverse: tensor product + matrix-free eval.\ of $\bm{M}_{1/J}$
-\begin{align*}
-\bm{M}^{-1}_J &\approx \LRp{\widehat{\bm{M}}\bm{M}_{1/J}^{-1}\widehat{\bm{M}}}^{-1}  = \widehat{\bm{M}}^{-1} \bm{M}_{1/J} \widehat{\bm{M}}^{-1}\\
-\widehat{\bm{M}}^{-1} &= \widehat{\bm{M}}^{-1}_{\rm 1D} \otimes \widehat{\bm{M}}^{-1}_{\rm 1D} \otimes \widehat{\bm{M}}^{-1}_{\rm 1D}.
-\end{align*}
-\item Energy stability with respect to equivalent norm 
-\[
-C_1(J) \nor{\bm{u}}_{\widehat{\bm{M}}\bm{M}_{1/J}^{-1}\widehat{\bm{M}}} \leq \nor{\bm{u}}_{\bm{M}_J} \leq C_2 \nor{\bm{u}}_{\widehat{\bm{M}}\bm{M}_{1/J}^{-1}\widehat{\bm{M}}}.
-\]
-%\vspace{.5em}
-%\item The same approach has also been successfully applied to Galerkin difference methods.  
-\end{itemize}
-
-%\let\thefootnote\relax\footnotetext{\tiny Mantzaflaris et al 2016. Low rank tensor methods in Galerkin-based IGA.}
-%\let\thefootnote\relax\footnotetext{\tiny Kirby 2017. Fast inversion of the simplicial Bernstein mass matrix.}
-\let\thefootnote\relax\footnotetext{\tiny Chan, et al.\ 2016.  Weight-adjusted DG methods: wave prop.\ in heterogeneous media. (SISC).  }
-\let\thefootnote\relax\footnotetext{\tiny Chan, et al.\ 2016.  {Weight-adjusted DG methods: curvilinear meshes} (arXiv).  }
-\let\thefootnote\relax\footnotetext{\tiny Also applied to GD methods (Banks and Hagstrom 2016. On Galerkin difference methods).}
-}
-
-\frame{
-\frametitle{Weight-adjusted DG for curvilinear meshes}
-
-\begin{itemize}
-\item Weight-adjusted projection $\tilde{P}_h$ on curved domains
-\[
-\tilde{P}_h(u) \coloneqq \widehat{P}_h\LRp{\frac{1}{J}\widehat{P}_h\LRp{uJ}}.
-\]
-where $\widehat{P}_h$ is the $L^2$ projection onto the reference element.
-\vspace{1em}
-\item $L^2$ estimates for weight-adjusted projection:
-\begin{align*}
-\nor{u - \tilde{P}_h u}_{L^2\LRp{D^k}} &\lesssim \nor{\frac{1}{\sqrt{J}}}_{L^{\infty}}^2 \textcolor{red}{\nor{J}_{W^{N+1,\infty}\LRp{D^k}}} h^{N+1}\nor{u}_{W^{N+1,2}\LRp{D^k}}.
-\end{align*}
-High order Sobelev norm of $J$ - can lose accuracy if the geometric mapping is not sufficiently regular!
-\end{itemize}
-}
-
-\frame{
-\frametitle{$L^2$ vs weight-adjusted projection}
-
-\vspace{-1em}
+\frametitle{Smoothed knot vectors improve the CFL}
+
 \begin{figure}
 \centering
-\subfloat[Warped mesh, $\alpha = 1/8$]{\includegraphics[width=.51\textwidth]{figs/mapped.png}}
-\subfloat[$L^2$ errors ]{
-\begin{tikzpicture}
-\begin{loglogaxis}[
-    legend cell align=left,
-    legend style={legend pos=south east, font=\tiny},
+\only<1>{
+\setcounter{subfigure}{0}
+\subfloat[Inverse constants, $K = 2p$]{
+\begin{tikzpicture}
+\begin{axis}[
 width=.475\textwidth,
-    xlabel={Mesh size $h$},
-    ylabel={$L^2$ error}, 
-    grid style=dashed,
-]
-
-\addplot[color=blue,mark=*,semithick, mark options={solid,fill=markercolor}]
-coordinates{(1,0.00192384)(0.5,0.00050273)(0.25,1.98999e-05)(0.125,3.15657e-07)};
-\addplot[color=red,mark=x,semithick, mark options={solid,fill=markercolor}]
-coordinates{(1,0.0019422)(0.5,0.000506873)(0.25,1.9966e-05)(0.125,3.15663e-07)};
-\addplot[color=black,mark=triangle*,semithick, mark options={solid,fill=markercolor}]
-coordinates{(1,1.836e-05)(0.5,4.14359e-06)(0.25,6.60985e-08)(0.125,5.88659e-12)};
-
-\legend{$L^2$ projection,Weight-adjusted,Difference}
-%\legend{Uniform, Optimal, Smoothed}
-\end{loglogaxis}
-\end{tikzpicture}
-}
-\caption{$L^2$ vs weight-adjusted projection: $\cos\LRp{\frac{\pi x}{2}}\cos\LRp{\frac{\pi y}{2}}$, $p = 4$. }
-%\caption{$L^2$ vs weight-adjusted projection. }
-\end{figure}
-
-}
-
-\section{Stable timestep restrictions}
-\frame[noframenumbering]{
-\tableofcontents[currentsection]
-}
-
-\frame{
-\frametitle{Stable timestep restrictions: $h, p$ scaling}
-\setcounter{subfigure}{0}
-
-\begin{itemize}
-\item Estimate $dt \leq 1/\rho(\bm{M}_h^{-1}\bm{A}_h)$.  Bound $\rho(\bm{M}_h^{-1}\bm{A}_h)$ using Rayleigh quotients and Bendixon-Hirsch; depends on $h$ and \textbf{constants} $C_T, C_I$.
-\[
-\nor{u}_{L^2\LRp{\partial \widehat{D}}} \leq C_T \nor{u}_{L^2\LRp{\widehat{D}}}, \qquad \nor{\Grad u}_{L^2\LRp{\widehat{D}}} \leq C_I \nor{u}_{L^2\LRp{\widehat{D}}}.
-\]
-\item CFL: $O(h/ p^2)$ for polynomials, $\textcolor{red}{O(h/ p)}$ for splines if $h \leq O(1/p)$.  
-\end{itemize}
-\begin{overlayarea}{\textwidth}{.65\textheight}
-\only<1>{
-\vspace{-.5em}
-\begin{figure}
-\centering
-\subfloat[Inverse inequality, $K = 2p$]{
-\begin{tikzpicture}
-\begin{axis}[
-width=.45\textwidth,
     xlabel={Degree of approximation $p$},   
-        ylabel={$C_I(p)$},
+%    ylabel={$C_I / K$},
     xmin=.5, xmax=12.5,
-    ymax=0,ymax=100,    
+    ymax=0,ymax=90,    
     legend pos=north west, legend cell align=left, legend style={font=\tiny},	
@@
 
-\legend{Polynomial $C_I$, Uniform $C_I/K$, Smoothed $C_I/K$ }
+\legend{Polynomial $C_I$, Spline $C_I/K$, Smoothed $C_I/K$ }
 \end{axis}
@@
 \hspace{1em}
-\subfloat[Trace inequality, $K = 2p$]{
+\subfloat[Trace constants, $K = 2p$]{
 \begin{tikzpicture}
 \begin{axis}[
-width=.45\textwidth,
+width=.475\textwidth,
     xlabel={Degree of approximation $p$},   
-    ylabel={$C_T(p)$},
     xmin=.5, xmax=12.5,
-    ymax=0,ymax=80,    
+    ymax=0,ymax=65,    
     legend pos=north west, legend cell align=left, legend style={font=\tiny},	
@@
 
-\legend{Polynomial $C_T$, Uniform $C_T / K $, Smoothed $C_T/K$ }
+\legend{Polynomial $C_T$, Spline $C_T/K$, Smoothed $C_T/K$ }
 \end{axis}
-\end{tikzpicture}
-}
-\end{figure}
+\end{tikzpicture}}
+\caption{Knot smoothing results in roughly $2\times$ smaller trace, inverse constants.}
 }
 \only<2>{
-\vspace{-1em}
-\begin{figure}
-\centering
+\setcounter{subfigure}{0}
 \subfloat[Uniform knots]{
@@
 \begin{axis}[
-width=.45\textwidth,
+width=.475\textwidth,
     xlabel={Number of elements $K$},   
-    ylabel={$\rho\LRp{\bm{M}_h^{-1}\bm{A}_h}$},       
+    ylabel={$\rho\LRp{\bm{A}_h}$},       
     xmin=0, xmax=300,
-    ymax=0,ymax=1600,    
+    ymax=0,ymax=2100,    
     legend pos=north west, legend cell align=left, legend style={font=\tiny},	
@@
 \begin{axis}[
-width=.45\textwidth,
+width=.475\textwidth,
     xlabel={Number of elements $K$},   
@@
     xmin=0, xmax=300,
-    ymax=0,ymax=1600,    
+    ymax=0,ymax=2100,    
     legend pos=north west, legend cell align=left, legend style={font=\tiny},	
@@
 }
-%\caption{Growth of $\rho\LRp{\bm{A}_h}$ for advection (partial upwinding) and $p = 2, 3, 4, 5$. }
-%\label{fig:rhoA}
+\caption{Growth of $\rho\LRp{\bm{A}_h}$ for advection using spline spaces of degree $p = 2, \ldots, 5$. }
+}
+\only<3>{
+\setcounter{subfigure}{0}
+\subfloat[Uniform knots]{
+\begin{tikzpicture}
+\begin{axis}[
+width=.475\textwidth,
+    xlabel={Number of elements $K$},   
+        ylabel={$\rho\LRp{\bm{A}_h}$},       
+    xmin=0, xmax=300,
+    ymax=0,ymax=700,    
+    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
+       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
+] 
+\addplot+[semithick, mark options={solid,fill=markercolor}]
+coordinates{(32,32.9239)(64,66.339)(96,99.6489)(128,132.926)(160,166.191)(192,199.449)(224,232.703)(256,265.956)};
+\addplot+[semithick, mark options={solid,fill=markercolor}]
+coordinates{(32,36.0829)(64,72.8362)(96,109.357)(128,145.911)(160,182.425)(192,218.937)(224,255.45)(256,291.953)};
+\addplot+[semithick, mark options={solid,fill=markercolor}]
+coordinates{(32,51.525)(64,77.0896)(96,115.797)(128,154.475)(160,193.14)(192,231.798)(224,270.454)(256,309.109)};
+\addplot+[semithick, mark options={solid,fill=markercolor}]
+coordinates{(32,69.1513)(64,91.7232)(96,120.356)(128,160.533)(160,200.771)(192,240.95)(224,281.151)(256,321.333)};
+\addplot+[semithick, mark options={solid,fill=markercolor}]
+coordinates{(32,92.592)(64,121.947)(96,154.335)(128,198.112)(160,246.867)(192,296.187)(224,345.548)(256,394.912)};
+\addplot+[semithick, mark options={solid,fill=markercolor}]
+coordinates{(32,112.952)(64,150.924)(96,193.087)(128,254.178)(160,317.703)(192,381.243)(224,444.784)(256,508.325)};
+\addplot+[semithick, mark options={solid,fill=markercolor}]
+coordinates{(32,144.171)(64,189.195)(96,241.188)(128,311.773)(160,388.78)(192,466.469)(224,544.209)(256,621.953)};
+
+%\addplot[color=black,dashed,semithick, mark options={solid,fill=markercolor}]
+%coordinates{(32,32)(64,64)(96,96)(128,128)(160,160)(192,192)(224,224)(256,256)}; % O(2/h) line
+%\legend{$p = 2$, $p=3$, $p=4$, $p = 5$, $p = 6$, $p = 7$, $p=8$}
+%\legend{$\frac{1}{h}$, $\frac{1.5}{h}$}
+\end{axis}
+\end{tikzpicture}
+}
+\subfloat[Smoothed knots]{
+\begin{tikzpicture}
+\begin{axis}[
+width=.475\textwidth,
+    xlabel={Number of elements $K$},   
+%        ylabel={$\rho\LRp{\bm{A}_h}$},       
+    xmin=0, xmax=300,
+    ymax=0,ymax=700,    
+    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
+       xmajorgrids=true,  ymajorgrids=true, grid style=dashed,
+]
+
+%\addplot[color=black,semithick, mark options={solid,fill=markercolor}]
+%coordinates{(32,32)(64,64)(96,96)(128,128)(160,160)(192,192)(224,224)(256,256)};% [yshift=-3pt] node[below, pos=.95,color=black] {$y = 1/h$};% O(1/h) line
+%\addplot[color=red,semithick, mark options={solid,fill=markercolor}]
+%coordinates{(32,48)(64,96)(96,144)(128,192)(160,240)(192,288)(224,336)(256,384)}; %[yshift=3pt] node[above, pos=.95,color=black] {$y = 1.5/h$};
+%\pgfplotsset{cycle list set=0}
+\addplot+[semithick, mark options={solid,fill=markercolor}]
+coordinates{(32,33.8023)(64,67.3347)(96,100.665)(128,133.949)(160,167.218)(192,200.479)(224,233.737)(256,266.992)}; %p = 2
+\addplot+[semithick, mark options={solid,fill=markercolor}]
+coordinates{(32,37.5425)(64,74.8178)(96,111.545)(128,148.129)(160,184.671)(192,221.196)(224,257.712)(256,294.223)}; % p = 3
+\addplot+[semithick, mark options={solid,fill=markercolor}]
+coordinates{(32,40.4264)(64,80.3236)(96,119.264)(128,158.017)(160,196.716)(192,235.393)(224,274.059)(256,312.72)}; % p = 4
+\addplot+[semithick, mark options={solid,fill=markercolor}]
+coordinates{(32,50.3593)(64,84.3661)(96,125.088)(128,165.445)(160,205.705)(192,245.927)(224,286.133)(256,326.328)}; % p = 5
+\addplot+[semithick, mark options={solid,fill=markercolor}]
+coordinates{(32,64.6167)(64,86.4858)(96,129.798)(128,173.13)(160,216.469)(192,259.811)(224,303.155)(256,346.5)}; % p = 6
+\addplot+[semithick, mark options={solid,fill=markercolor}]
+coordinates{(32,76.7086)(64,97.5475)(96,133.384)(128,177.856)(160,222.354)(192,266.858)(224,311.365)(256,355.876)}; %p = 7
+\addplot+[semithick, mark options={solid,fill=markercolor}]
+coordinates{(32,94.4519)(64,116.577)(96,136.311)(128,181.68)(160,227.101)(192,272.535)(224,317.975)(256,363.42)};  % p =8
+%\addplot[color=black,dashed,semithick, mark options={solid,fill=markercolor}]
+%coordinates{(32,32)(64,64)(96,96)(128,128)(160,160)(192,192)(224,224)(256,256)}; % O(2/h) line
+%\legend{$p = 2$, $p=3$, $p=4$, $p = 5$, $p = 6$, $p = 7$, $p=8$}
+%\legend{${1}/{h}$, ${1.5}/{h}$}
+\end{axis}
+\end{tikzpicture}
+}
+\caption{Growth of $\rho\LRp{\bm{A}_h}$ using an \textcolor{red}{upwind flux} and spline spaces of degree $p = 2, \ldots, 8$.
+\textcolor{red}{Nearly $p$-independent CFL observed for advection, acoustics}.}
+}
 \end{figure}
-}
-\end{overlayarea}
 }
@@
 \frame{
-\frametitle{A 3D multi-patch example}
+\frametitle{Acoustics: a 3D multi-patch example}
 \setcounter{subfigure}{0}
@@
 \end{itemize}
-\let\thefootnote\relax\footnotetext{\tiny Chan, Evans.\ 2017.  {Multi-patch discontinuous Galerkin spline FEM for time-domain wave propagation} (in preparation).}
+%\let\thefootnote\relax\footnotetext{\tiny Chan, Evans.\ 2017.  {Multi-patch discontinuous Galerkin spline FEM for time-domain wave propagation} (in preparation).}
 }
@@
 \begin{itemize}
-\item Optimal/smoothed knot vectors for time-domain simulations.  
-%\vspace{1em}
-\item Restore Kronecker product with weight-adjusted mass matrix.  
-\vspace{1em}
-\item Future directions:
+\item Weight-adjusted mass matrix: restore Kronecker structure while retaining energy stability and high order accuracy.  
+\vspace{.25em}
+\item Improved $O(h/p)$ CFL scaling for IGA, optimal $L^2$ convergence rates.  
+\vspace{.25em}
+\item Smoothed knots: improved CFL, better curved approximations.
+\vspace{.25em}
+\item Future directions: curl-conforming spline spaces (Maxwells).
+\vspace{.25em}
+\item This research is supported by DMS-1719818 and DMS-1712639.  
+\end{itemize}
 \vspace{.5em}
-\begin{itemize}
-\item Conforming spaces for first order formulations.
-\vspace{.5em}
-\item Nonlinear systems (compressible flow).
-\end{itemize}
-\end{itemize}
-\vspace{1em}
 \begin{center}
-This research is supported by DMS-1712639 and TOTAL E\&P Research and Technology USA.
+Thank you!  Questions?
+\vspace{.25em}
+
+{\includegraphics[width=.15\textwidth]{figs/nsf.jpg}}
 \end{center}
 
-\let\thefootnote\relax\footnotetext{\tiny Chan, Evans.\ 2017.  {Multi-patch discontinuous Galerkin spline FEM for time-domain wave propagation} (in preparation).}
-\let\thefootnote\relax\footnotetext{\tiny Chan, et al.\ 2016.  {Weight-adjusted DG methods: curvilinear meshes} (arXiv).  }
+\let\thefootnote\relax\footnotetext{\tiny Chan, Evans (2018).
+\textit{Multi-patch discontinuous Galerkin isogeometric analysis for wave propagation: explicit time-stepping and efficient mass matrix inversion}.}
+%\let\thefootnote\relax\footnotetext{\tiny Chan, et al.\ 2016.  {Weight-adjusted DG methods: curvilinear meshes} (arXiv).  }
 %\let\thefootnote\relax\footnotetext{\tiny Chan 2017.  Weight-adjusted DG methods: matrix-valued weights and elastic wave prop.\ in heterogeneous media (arXiv).}
@@
 \end{figure}
-}
-
-\frame[noframenumbering]{
-\frametitle{Convergence of WADG on curvilinear meshes}
-
-\begin{figure}
-\centering
-\subfloat{\includegraphics[width=.375\textwidth]{figs/curved_mesh.png}}
-\hspace{.5em}
-\setcounter{subfigure}{0}
-\subfloat[$L^2$ errors for $N = 2,3,4$]{
-
-\begin{tikzpicture}
-\begin{loglogaxis}[
-	legend cell align=left,
-	width=.5\textwidth,
-%    title={Convergence},
-    xlabel={Mesh size $h$},
-    ylabel={$L^2$ error},
-    xmin=.01, xmax=10,
-    ymin=1e-9, ymax=1,        
-    legend pos=south east,
-    xmajorgrids=true,
-    ymajorgrids=true,
-    grid style=dashed,
-] 
-\addplot+[color=blue,mark=*,mark options={fill=markercolor},semithick]
-coordinates{(0.5,0.026045)(0.25,0.00661508)(0.125,0.000722446)(0.0625,9.5217e-05)};
-\addplot+[color=blue,mark=x,mark options={solid,fill=markercolor},semithick]
-coordinates{(0.5,0.0260431)(0.25,0.00661506)(0.125,0.000722445)(0.0625,9.5217e-05)};
-
-\addplot+[color=red,mark=square*,mark options={fill=markercolor},semithick]
-coordinates{(0.5,0.0101239)(0.25,0.000387682)(0.125,2.88879e-05)(0.0625,1.78875e-06)};
-\addplot+[color=red,mark=x,mark options={solid,fill=markercolor},semithick]
-coordinates{(0.5,0.0101238)(0.25,0.000387682)(0.125,2.88879e-05)(0.0625,1.78875e-06)};
-
-\addplot+[color=black,mark=triangle*,mark options={fill=markercolor},semithick]
-coordinates{(0.5,0.000483144)(0.25,4.84164e-05)(0.125,1.38952e-06)(0.0625,4.71587e-08)};
-\addplot+[color=black,mark=x,mark options={solid,fill=markercolor},semithick]
-coordinates{(0.5,0.000483096)(0.25,4.84162e-05)(0.125,1.38952e-06)(0.0625,4.71587e-08)};
-\legend{DG,WADG}
-\end{loglogaxis}
-\end{tikzpicture}
-}
-\caption{Optimal $L^2$ convergence rates observed for curvilinear meshes.  }
-\end{figure}
-}
-
-\frame[noframenumbering]{
-\frametitle{Curvilinear meshes: DG eigenvalues (circular domain)}
-\vspace{-1.5em}
-\begin{figure}
-\centering
-\subfloat[Central fluxes]{
-\includegraphics[width=.38\textwidth]{figs/eigstau0.png}
-}
-\subfloat[${\rm Im}\LRp{\lambda_i}$ for central fluxes]{
-\includegraphics[width=.38\textwidth]{figs/eigstau0imag.png}
-}\\
-\vspace{-.75em}
-\subfloat[Upwind fluxes]{
-\includegraphics[width=.38\textwidth]{figs/eigstau1.png}
-}
-%\caption{Spectra of discretized wave equation using DG and WADG for $N=3$.  Both central ($\tau = 0$) and dissipative ($\tau = 1$) numerical fluxes are shown.}
-\end{figure}
-%\begin{itemize}
-%\item Energy stability requires quadrature, skew-symmetric formulation
-%\begin{align*}
-%\int_{D^k}\pd{p}{t}{} q &= \int_{D^k} -\bm{u}\cdot \Grad q  + \frac{1}{2}\int_{\partial D^k} \LRp{2\avg{\bm{u}}\cdot{\bm{n}} + \tau_p\jump{p}} q\\
-%\int_{D^k}\pd{\bm{u}}{t}{} \bm{v} &= \int_{D^k} \Grad p \cdot \bm{v} + \frac{1}{2}\int_{\partial D^k} \LRp{\jump{p} + \tau_u\jump{\bm{u}}\cdot\bm{n}} \bm{v}
-%\end{align*}
-%\end{itemize}
 }

### 2018-06-20 esdg_quadhex_cpp/include/Mesh.h

@@ -43,17 +43,6 @@ typedef struct foo {
   MatrixXi EToE;
   MatrixXi EToF;
   
-  VectorXi EToGmshE;
-
-  int **EToVC; /* element to global vertex list  */
-  int **EToEC; /* element to neighbor element (elements numbered by their proc) */
-  int **EToFC; /* element to neighbor face    (element local number 0,1,2) */
-  //dfloat **GX; /* x-coordinates of element vertices */
-  //dfloat **GY; /* y-coordinates of element vertices */
-  //dfloat **GZ; /* z-coordinates of element vertices (3d) */
-  MatrixXd GX,GY,GZ;
-
-
   // =============== reference elem stuff ===================
 
   int Nverts; /* number of vertices per element */
@@ -68,8 +57,9 @@ typedef struct foo {
   MatrixXi Fmask;
   int   **FmaskC; /* face node numbers in element volume data */
 
-  // 1D operators (tensor product)
-  VectorXd V1D, D1D, Vf1D;
+  // 1D operators (for tensor product)
+  VectorXd D1D, Vf1D, Lf1D;
+  VectorXd wq1D;
   
   // nodal points (GLL for quads/hexes)
   VectorXd r,s,t;
@@ -87,14 +77,34 @@ typedef struct foo {
   MatrixXd rxJ,ryJ,rzJ,sxJ,syJ,szJ,txJ,tyJ,tzJ,J; // geofacs
   MatrixXd nxJ,nyJ,nzJ,sJ; // surface geofacs
 
+  MatrixXi mapPq; // node map
+  
   // time stepping constants
-  dfloat *rk4a, *rk4b, *rk4c;
+  VectorXd rk4a, rk4b, rk4c;  
 
   // mesh size for dt - computed  
   double hMax; 
   double hMin;
-  
+ 
 }Mesh;
 
+
+// struct for occa arrays, kernels, etc
+typedef struct bar {
+
+  occa::device device;
+  occa::properties props; 
+  occa::kernel volume, surface, update; 
+
+  occa::memory o_D, o_Vf, o_Lf; // operators
+  occa::memory o_wq1D; // 1D weights (for lifting ops)
+
+  occa::memory o_Q, o_Qf; // solution and flux vals
+  occa::memory o_vgeo, o_fgeo; // geometric terms
+  occa::memory o_rhs, o_res; // rhs and RK residual
+  occa::memory o_mapPq; // node map
+
+  
+}App;
 
 #endif

### 2018-06-20 esdg_quadhex_cpp/src/StartUp.cpp

@@ -14,8 +14,8 @@ void InitRefData2d(Mesh *mesh, int N, int Nfields){
   meshgrid(r1D,r1D, smat,rmat);
   VectorXd r = Map<VectorXd>(rmat.data(), rmat.cols()*rmat.rows());
   VectorXd s = Map<VectorXd>(smat.data(), smat.cols()*smat.rows());
-  cout << "r = " << r << endl;
-  cout << "s = " << s << endl;  
+  //  cout << "r = " << r << endl;
+  //  cout << "s = " << s << endl;  
 
   // quad points 
   VectorXd rq1D, wq1D;
@@ -82,9 +82,35 @@ void InitRefData2d(Mesh *mesh, int N, int Nfields){
   mesh->Vf = Vf;  
 
   // GQ nodes
-  mesh->V1D = Vq1D;
   mesh->D1D = Dq1D;
-  mesh->Vf1D = Vf1D;    
+  mesh->Vf1D = Vf1D;
+  VectorXd wqInv = wq1D.array().inverse();
+  mesh->Lf1D = wqInv.asDiagonal()*Vf1D; // 1D lift - may not need
+  mesh->wq1D = wq1D;
+
+  // LSRK-45 coefficients
+  mesh->rk4a.resize(5);
+  mesh->rk4a(0) =              0.0;
+  mesh->rk4a(1) =  -567301805773.0 / 1357537059087.0;
+  mesh->rk4a(2) = -2404267990393.0 / 2016746695238.0;
+  mesh->rk4a(3) = -3550918686646.0 / 2091501179385.0;
+  mesh->rk4a(4) = -1275806237668.0 /  842570457699.0;
+
+  mesh->rk4b.resize(5);
+  mesh->rk4b(0) =  1432997174477.0 /  9575080441755.0;
+  mesh->rk4b(1) =  5161836677717.0 / 13612068292357.0;
+  mesh->rk4b(2) =  1720146321549.0 /  2090206949498.0;
+  mesh->rk4b(3) =  3134564353537.0 /  4481467310338.0;
+  mesh->rk4b(4) =  2277821191437.0 / 14882151754819.0;
+
+  mesh->rk4c.resize(5);
+  mesh->rk4c(0) =              0.0;
+  mesh->rk4c(1) =  1432997174477.0 / 9575080441755.0;
+  mesh->rk4c(2) =  2526269341429.0 / 6820363962896.0;
[…]

### 2018-06-20 matlab/cfd2D/Euler2D_ext_decoupled.m

@@ -5,14 +5,14 @@ Globals2D
 a = 0/8; % warping factor
 % FinalTime = 10;
 
-N = 2;
-K1D = 16;
+N = 4;
+K1D = 8;
 FinalTime = 1;
 
 wadgProjEntropyVars = abs(a)>1e-8;
-CFL = .5;
+CFL = .25;
 global tau
-tau = 1;
+tau = 0;
 
 % Lx = 7.5; Ly = 5; ratiox = 3/4; ratioy = .5;
 Lx = 10; Ly = 5; ratiox = 1; ratioy = Ly/Lx;
@@ -33,7 +33,7 @@ BuildPeriodicMaps2D(max(VX)-min(VX),max(VY)-min(VY));
 % PlotMesh2D;axis on;return
 
 % plotting nodes
-[rp sp] = EquiNodes2D(15); [rp sp] = xytors(rp,sp);
+[rp sp] = EquiNodes2D(10); [rp sp] = xytors(rp,sp);
 Vp = Vandermonde2D(N,rp,sp)/V;
 xp = Vp*x; yp = Vp*y;
 % PlotMesh2D; axis on;return
@@ -57,7 +57,9 @@ Pq = M\(Vq'*diag(wq)); % J's cancel out
 xq = Vq*x; yq = Vq*y;
 
 % face nodes
-[rq1D wq1D] = JacobiGQ(0,0,N+1);
+% [rq1D wq1D] = JacobiGQ(0,0,N+1);
+rq1D = [-1 + (1+rq1D)/2; (1+rq1D)/2];
+wq1D = .5*[wq1D;wq1D];
 Nfq = length(rq1D);
 e = ones(size(rq1D));
 rfq = [rq1D; e; -rq1D; -e];
@@ -649,7 +651,7 @@ p = rho.^gamma;
 % v = zeros(size(x));
 % p = ones(size(x));
 
-if 0
+if 1
     % pulse condition
     x0 = 5;
     rho = 2 + (abs(x-x0) < 5);

### 2018-06-20 matlab/cfd2D/Euler2D_noncon.m

@@ -4,8 +4,8 @@ clear
 % useQuads = 1;mypath
 
 Globals2D;
-N = 3;
-K1D = 16;
+N = 1;
+K1D = 2;
 useSkew = 1;
 CFL = .25;
 FinalTime = 1;
@@ -21,7 +21,7 @@ VX = (VX+1)*Lx; VY = VY*Ly;
 
 % [Nv, VX, VY, K, EToV] = QuadMesh2D(K1D,K1D);
 
-a = .0/K1D;
+a = 1/K1D;
 wadgProjEntropyVars = abs(a)>1e-8;
 
 ids = find(abs(abs(VX)-1) > 1e-8 & abs(abs(VY)-1) > 1e-8);
@@ -75,7 +75,7 @@ rxJ = Vq*rxJ; sxJ = Vq*sxJ;
 ryJ = Vq*ryJ; syJ = Vq*syJ;
 J = Vq*J;
 
-[rq1D wq1D] = JacobiGL(0,0,N);
+[rq1D wq1D] = JacobiGQ(0,0,N);
 % [rq1D wq1D] = JacobiGQ(0,0,N); % exact for 2N-1
 % [rq1D2 wq1D2] = JacobiGQ(0,0,N);
 % [rq1D wq1D] = JacobiGQ(0,0,N-1); % exact for 2N-1
@@ -361,32 +361,40 @@ if 0
     return
 end
 
-rxJ = zeros(Nq,K); sxJ = zeros(Nq,K);
-ryJ = zeros(Nq,K); syJ = zeros(Nq,K);
-J = zeros(Nq,K);
-rxJf = zeros(Nfq*Nfaces,K); sxJf = zeros(Nfq*Nfaces,K);
-ryJf = zeros(Nfq*Nfaces,K); syJf = zeros(Nfq*Nfaces,K);
-Jf = zeros(Nfq*Nfaces,K);
+[rxk,sxk,ryk,syk,J] = GeometricFactorsQuad2D(x,y,Vq*Dr,Vq*Ds);
+rxJ = rxk.*J;
+ryJ = ryk.*J;
+sxJ = sxk.*J;
+syJ = syk.*J;
+
+[rxk,sxk,ryk,syk,Jk] = GeometricFactorsQuad2D(x,y,Vfq*Dr,Vfq*Ds);
+rxJf = rxk.*Jk;    sxJf = sxk.*Jk;
[…]

### 2018-07-19 docs/ESDG_GQ/esdg_gq.tex
updating ESDG GQ paper

\pdfoutput=1

%\documentclass[preprint,10pt]{elsarticle}
%\documentclass[preprint,10pt]{amsart}
%\documentclass[review, 10pt]{siamart0216}
%\documentclass{siamart0216}
\documentclass[review,onefignum,onetabnum,final]{siamart171218}

%\usepackage{fullpage}
%\usepackage[colorlinks=true]{hyperref}

\usepackage{amsmath,amssymb,amsfonts}

%\usepackage{pdfpages}
\usepackage[titletoc,toc,title]{appendix}
\usepackage{listings}
\usepackage{array} 
\usepackage{mathtools}
\usepackage{bm}
\usepackage{bbm}

\usepackage{tikz}
\usepackage[normalem]{ulem}
\usepackage{hhline}

\usepackage{graphicx}
\usepackage{subfig}
\usepackage{color}

%% ====================================== graphics

\usepackage{pgfplots}
\usepackage{pgfplotstable}
\definecolor{markercolor}{RGB}{124.9, 255, 160.65}
\pgfplotsset{
compat=1.3,
width=10cm,
tick label style={font=\small},
label style={font=\small},
legend style={font=\small}
}

\usetikzlibrary{calc}
\usetikzlibrary{intersections}

%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
\newcommand{\logLogSlopeTriangle}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        \pgfmathsetmacro{\xArel}{#1}
        \pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{\yArel}
        \pgfmathsetmacro{\xCrel}{\xArel}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=north] {}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=west] {\textcolor{black}{#4}}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

\newcommand{\logLogSlopeTriangleNeg}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        \pgfmathsetmacro{\xArel}{#1}
        \pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{\yArel}
        \pgfmathsetmacro{\xCrel}{\xArel}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=.5,anchor=south] {}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=west] {\textcolor{black}{#4}}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
\newcommand{\logLogSlopeTriangleFlipNeg}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        %\pgfmathsetmacro{\xArel}{#1}
        %\pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{#3}
        \pgfmathsetmacro{\xCrel}{#1}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

	\pgfmathsetmacro{\xArel}{\xBrel}
        \pgfmathsetmacro{\yArel}{\yCrel}

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=east] {\textcolor{black}{#4}}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=north] {1}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
\newcommand{\logLogSlopeTriangleFlip}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        %\pgfmathsetmacro{\xArel}{#1}
        %\pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{#3}
        \pgfmathsetmacro{\xCrel}{#1}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

	\pgfmathsetmacro{\xArel}{\xBrel}
        \pgfmathsetmacro{\yArel}{\yCrel}

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=east] {\textcolor{black}{#4}}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=south] {}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

\usepackage{stmaryrd}

\renewcommand{\tilde}{\widetilde}
\renewcommand{\hat}{\widehat}
%\renewcommand{\topfraction}{0.85}
%\renewcommand{\textfraction}{0.1}
%\renewcommand{\floatpagefraction}{0.75}

\newcommand{\vect}[1]{\ensuremath\boldsymbol{#1}}
\newcommand{\tensor}[1]{\underline{\bm{#1}}}
\newcommand{\del}{\triangle}
\newcommand{\curl}{\grad \times}
\renewcommand{\div}{\grad \cdot}

\newcommand{\bbm}[1]{\mathbbm{#1}}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\equaldef}{\stackrel{\mathrm{def}}{=}}

\newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}{\rm #2}}}
\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRb}[1]{\left| #1 \right|}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}
\newcommand{\LRceil}[1]{\left\lceil #1 \right\rceil}
\newcommand{\LRl}[1]{\left. #1 \right|}
\newcommand{\pdd}[2]{\frac{\partial^2#1}{\partial#2^2}}
\newcommand{\pdn}[3]{\frac{\partial^{#3}#1}{\partial#2^{#3}}}
\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\snor}[1]{\left| #1 \right|}

%\newcommand{\cond}[1]{\kappa\LRp{#1}}
\newcommand{\cond}[2]{\nor{#1}_{#2}\nor{{#1}^{-1}}_{#2}}

\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\Div} {\ensuremath{\nabla\cdot}}
\newcommand{\jump}[1] {\ensuremath{\llbracket#1\rrbracket}}
\newcommand{\avg}[1] {\ensuremath{\LRc{\!\{#1\}\!}}}

\newcommand{\Oh}{{\Omega_h}}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\LK}{L^2\LRp{D^k}}
\newcommand{\LdK}{L^2\LRp{\partial D^k}}
\newcommand{\Dhat}{\widehat{D}}
\newcommand{\Lhat}{L^2\LRp{\Dhat}}

\newcommand{\eval}[2][\right]{\relax
  \ifx#1\right\relax \left.\fi#2#1\rvert}

\def\etal{{\it et al.~}}

\newcommand{\note}[1]{{\color{blue}{#1}}}
%\newcommand{\noteOne}[1]{{\color{blue}{#1}}}
%\newcommand{\noteTwo}[1]{{\color{red}{#1}}}
%\newcommand{\note}[1]{#1}
%\newcommand{\noteOne}[1]{#1}
%\newcommand{\noteTwo}[1]{#1}

\newcommand{\LinfDk}{L^{\infty}\LRp{D^k}}

\newcommand{\diag}[1]{{\rm diag}\LRp{#1}}

\newcommand{\Ksub}{K_{\rm sub}}

\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

%% d in integrand
\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}}

\makeatletter
\renewcommand\d[1]{\mspace{6mu}\mathrm{d}#1\@ifnextchar\d{\mspace{-3mu}}{}}
\makeatother

%\date{}
\author{Jesse Chan, David C.\ Del Rey Fernandez, Mark H.\ Carpenter}
\title{Entropy stable Gauss-Legendre collocation methods}
\graphicspath{{./figs/}}

\begin{document}

\maketitle

\begin{abstract}
Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.  Here is the abstract.

\end{abstract}

\section{Introduction}

Time dependent nonlinear conservation laws are ubiquitous in computational fluid dynamics (CFD).  These conservation laws include the compressible Euler and Navier-Stokes equations, which govern compressible fluid behavior within aerospace applications, as well as the equations governing coastal flooding and geophysical flows (shallow water), and those describing astrophysical phenomena (magnetohydrodynamics).  The utility of these simulations depends both on the reliability and accuracy of numerical methods for the solution of the associated nonlinear partial differential equations (PDEs).

High order methods are increasingly of interest for these applications.  Such methods are more accurate per degree of freedom than low order methods, while also possessing much smaller numerical dispersion and dissipation errors.  This makes high order methods especially well suited to time-dependent simulations.  In this work, we focus specifically on discontinuous Galerkin methods on unstructured quadrilateral and hexahedral meshes.  These methods combine properties of high order approximations with the geometric flexibility of unstructured meshing.  %When paired with explicit time-stepping methods, high order DG and finite element methods are also well-suited for many-core and GPU computing \cite{klockner2009nodal, swirydowicz2017acceleration, wintermeyer2018entropy}.

However, high order methods are notorious for being more prone to instability compared to low order methods \cite{wang2013high}.  This instability is addressed through various stabilization techniques (e.g.\ artificial viscosity, filtering, slope limiting).  However, these techniques often reduce accuracy to first or second order, and can prevent solvers from realizing the advantages of high order approximations.  Moreover, it is often not possible to prove that a high order scheme does not blow up even in the presence of stabilization.  This ambiguity can necessitate the re-tuning of stabilization parameters, as parameters which are both stable and accurate for one problem or discretization setting  may provide either too little or too much numerical dissipation for another.

The instability of high order methods is rooted in the fact that discretizations of nonlinear conservation laws do not typically satisfy a discrete analogue of the conservation or dissipation of energy (entropy).  While the numerical dissipation present for low order methods serves as a stabilizing mechanism, the low of numerical dissipation in high order methods reveals the lack of an inherent statement of stability.  The dissipation of entropy serves as an energetic principle for nonlinear conservation laws \cite{dafermos2005compensated}, and requires the use of the chain rule in its proof.  The instability of high order methods is tied to the fact that, when discretizing systems of nonlinear PDEs, the chain rule does not typically hold at the discrete level.  The lack of a chain rule was circumvented by using a non-standard ``flux differencing'' formulation \cite{fisher2013high, carpenter2014entropy, gassner2016split, gassner2017br1}, which is key to constructing semi-discretely entropy stable high order schemes on unstructured quadrilateral and hexahedral meshes.  Flux differencing replaces the derivative of the nonlinear flux with the discrete differentiation of an auxiliary quantity.  This auxiliary quantity is computed through the evaluation of a two-point entropy conservative flux \cite{tadmor1987numerical} using pairs of solution values at quadrature points.  These entropy stable schemes were later extended to non-tensor product elements using GLL-like quadrature points on triangles and tetrahedra \cite{chen2017entropy, crean2018entropy}.  More recently, the construction of efficient entropy stable schemes was extended to more arbitrary choices of basis and quadrature \cite{chan2017discretely, chan2018discretely}.

This paper focuses on the construction of efficient semi-discretely entropy stable collocation schemes based on Gauss nodes, using the framework introduced in \cite{chan2017discretely, chan2018discretely}.  While entropy stable collocation schemes have been constructed on Gauss-like quadrature points without boundary nodes \cite{crean2017high}, the inter-element coupling terms for such schemes introduce significantly more communication and computation than collocation points which contain boundary nodes.  This work introduces new inter-element coupling terms which avoid this all-to-all coupling, requiring only communication of face values between neighboring elements.

In Section~\ref{sec:0}, we briefly review the derivation of continuous entropy inequalities for systems of nonlinear conservation laws.  In Section~\ref{sec:1}, we describe how to construct entropy stable discretizations of nonlinear conservation laws using different quadrature points on affine tensor product elements.  In Section~\ref{sec:2}, we describe how to extend this construction to curvilinear elements, and Section~\ref{sec:3} presents numerical results which confirm the high order accuracy and stability of the proposed method for smooth, discontinuous, and under-resolved (turbulent) solutions of the compressible Euler equations in two and three dimensions.

\section{A brief review of entropy stability theory}
\label{sec:0}

We are interested in methods for the numerical solution of systems of conservation laws in $d$ dimensions
\begin{equation}
\pd{\bm{u}}{t} + \sum_{i=1}^d \pd{\bm{f}_i\LRp{\bm{u}}}{x_i} = 0,
\label{eq:nonlinpde}
\end{equation}
where $\bm{u}$ denotes the conservative variables, $\bm{f}_i(\bm{u})$ are nonlinear fluxes, and $x_i$ denotes the $i$th coordinate.  Many physically motivated conservation laws admit a statement of stability involving a convex scalar entropy $S(\bm{u})$.  We first define the entropy variables $\bm{v}(\bm{u})$ to be the gradient of the entropy $S(\bm{u})$ with respect to the conservative variables 
\[
\bm{v} = \pd{S(\bm{u})}{\bm{u}}.  
\]
For a convex entropy, $\bm{v}(\bm{u})$ defines an invertible mapping from conservative to entropy variables.  We denote the inverse of this mapping (from entropy to conservative variables) by $\bm{u}(\bm{v})$.

At the continuous level, it can be shown (for example, in \cite{dafermos2005compensated}) that vanishing viscosity solutions to (\ref{eq:nonlinpde}) satisfy the strong form of an entropy inequality 
\begin{gather}
\pd{S(\bm{u})}{t} + \sum_{i=1}^d \pd{F_i(\bm{u})}{x_i} \leq 0,
\label{eq:entropyineq}
\end{gather}
where $F_i$ denotes the $i$th scalar entropy flux function and $\psi_i(\bm{u})$ denotes the $i$th entropy potential
\[
F_i(\bm{u}) = \bm{v}^T\pd{\bm{f}_i}{x_i}, \qquad \psi_i(\bm{u}) = \bm{v}^T\bm{f}_i(\bm{u}) - F_i(\bm{u}). 
\]
Integrating (\ref{eq:entropyineq}) over a domain $\Omega$ and applying the divergence theorem yields an averaged entropy inequality
\begin{equation}
\int_{\Omega} \pd{S(\bm{u})}{t} + \int_{\partial \Omega} \sum_{i=1}^d \bm{n}_i \LRp{\bm{v}^T\bm{f}_i(\bm{u}) - \psi_i(\bm{u})} \leq 0,
\label{eq:weakentropyineq}
\end{equation}
where $\partial \Omega$ denotes the boundary of $\Omega$ and $\bm{n}_i$ denotes the $i$th component of the outward normal on $\partial \Omega$.  Roughly speaking, this implies that the time rate of change of entropy is less than or equal to the flux of entropy through the boundary.

\section{Entropy stable Gauss and Gauss-Legendre-Lobatto collocation methods}
\label{sec:1}

This focus of this paper is on entropy stable high order collocation methods which satisfy a semi-discrete version of the entropy inequality (\ref{eq:weakentropyineq}).  
These methods collocate the solution at some choice of collocation nodes, and are applicable to tensor product meshes consisting of quadrilateral and hexahedral elements.

Entropy stable collocation methods have largely utilized Gauss-Legendre-Lobatto (GLL) nodes \cite{fisher2013high, carpenter2014entropy, gassner2016split, gassner2017br1}, which contain points on the boundary.  The popularity of GLL nodes can be attributed in part to a connection made in \cite{gassner2013skew}, where it was shown by Gassner that collocation DG discretizations based on GLL nodes could be recast in terms of summation-by-parts (SBP) operators.  This equivalence allowed Gassner to leverage existing finite difference formulations to produce stable high order discretizations of the nonlinear Burgers' equation.

GLL quadratures contain boundary points, which greatly simplifies the construction of inter-element coupling terms for entropy stable collocation schemes.  However, it is also known that the use of GLL quadrature within DG methods under-integrates the mass matrix, which can lead to solution ``aliasing'' and lower accuracy \cite{parsani2016entropy}.  In this work, we explore entropy stable collocation schemes based on Gauss quadrature points instead of GLL points.

This comparison is motivated by the accuracy of each respective quadrature rule.  While $(N+1)$-point GLL quadrature rules are exact for polynomial integrands of degree $(2N-1)$, $(N+1)$-point Gauss quadrature rules are exact for polynomials of degree $(2N+1)$.  This higher accuracy of Gauss quadrature has been shown to translate to lower errors and slightly improved rates of convergence in simulations of wave propagation and fluid flow \cite{kopriva2010quadrature, hindenlang2012explicit, chan2015gpu}.  However, Gauss points have not been widely used to construct entropy stable discretizations due to the lack of efficient, stable, and high order accurate inter-element coupling terms, known as simultaneous approximation terms (SBP-SAT) in the finite difference literature \cite{fernandez2014review, crean2017high, fernandez2018simultaneous}.  SBP-SATs for Gauss points are non-compact, in the sense that they introduce all-to-all coupling between degrees of freedom on neighboring elements.  This results in greater communication between elements, as well as a significantly larger number of two-point flux evaluations and floating point operations.

It is possible to realize the improved accuracy of Gauss points while avoiding non-compact SBP-SATs through a staggered grid formulation, where the solution is stored at Gauss nodes but interpolated to a set of higher degree $(N+2)$ GLL ``flux'' points for computation \cite{parsani2016entropy}.  Because GLL nodes include boundary points, compact and high order accurate SBP-SAT terms can be constructed for the flux points.  After performing computations on the flux points, the results are interpolated back to Gauss points and used to evolve the solution forward in time.  Figure~\ref{fig:nodesets} shows an illustration of GLL, staggered grid, and Gauss point sets for a 2D quadrilateral element.

\begin{figure}
\centering
\subfloat[GLL nodes]{\includegraphics[width=.3\textwidth]{figs/sbp.png}}
\hspace{1em}
\subfloat[Staggered grid nodes]{\includegraphics[width=.3\textwidth]{figs/staggered.png}}
\hspace{1em}
\subfloat[Gauss nodes]{\includegraphics[width=.3\textwidth]{figs/gsbp.png}}
\caption{Examples of nodal sets under which efficient entropy stable schemes can be constructed.  This work focuses on the construction of efficient and accurate SBP-SAT terms for Gauss nodal sets.}
\label{fig:nodesets}
\end{figure}

The following sections will describe how to construct efficient high order entropy stable schemes using Gauss points.  These schemes are based on ``decoupled'' SBP operators introduced in \cite{chan2017discretely, chan2018discretely}, which are applicable to general choices of basis and quadrature.  By choosing a tensor product Lagrange polynomial basis and $(N+1)$ point Gauss quadrature rules, we recover a Gauss collocation scheme.  The high order accuracy and entropy stability of this scheme are direct results of theorems presented in  \cite{chan2017discretely, chan2018discretely}.  However, we will also present a different proof of entropy stability in one dimension for completeness.

\subsection{Gauss nodes and generalized summation by parts operators}

We assume the solution is collocated at $(N+1)$ quadrature points $x_i$ with associated quadrature weights $w_i$.  We do not make any assumptions on the points, in order to accommodate both GLL and Gauss nodes using this notation.  The collocation assumption is equivalent to approximating the solution using a degree $N$ Lagrange basis $\ell_j(x)$ defined over the $(N+1)$ quadrature points.

Let $\bm{D}$ denote the nodal differentiation matrix, and let $\bm{V}_f$ denote the $(N+1)\times 2$ matrix which interpolates polynomials at Gauss nodes to values at endpoints.  These two matrices are defined entrywise as
\[
\bm{D}_{ij} = \LRl{\pd{\ell_j}{x}}_{x = x_i}, \qquad \LRp{\bm{V}_f}_{i1} = \ell_i(-1), \qquad  \LRp{\bm{V}_f}_{i2} = \ell_i(1).
\]
We also introduce the diagonal matrix of quadrature weights $\bm{W}_{ij} = \delta_{ij} w_i$.  In this work, $\bm{W}$ is identical to the mass matrix, whose entries are evaluated as $L^2$ inner products of basis functions.  This equality holds so long as the basis functions are Lagrange polynomials and integrals are approximated using quadrature (either GLL or Gauss).  
%\[
%\bm{M}_{ij} = \int_{-1}^{1} \ell_j(x)\ell_i(x) = \sum_{k=1}^{N+1}  w_k\ell_j(x_k)\ell_i(x_k)  = \delta_{ij} w_i.  
%\]
It can be show that the mass and differentiation matrices for Gauss nodes fall under the class of generalized SBP (GSBP) operators \cite{fernandez2014generalized}.  
\begin{lemma}
$\bm{Q} = \bm{W}\bm{D}$ satisfies the generalized summation by parts property
\[
\bm{Q} = \bm{V}_f^T \bm{B} \bm{V}_f - \bm{Q}^T, \qquad \bm{B} = \begin{bmatrix}-1 & \\ & 1\end{bmatrix}.
\]
\label{lemma:sbp}
\end{lemma}
\begin{proof}
The proof is a direct restatement of integration by parts, and can be found in \cite{ortleb2016kinetic, ortleb2017kinetic, ranocha2018generalised}.  We reproduce it here for completeness.  Let $\bm{u},\bm{v}$ be two arbitrary vectors, and let $u(x), v(x)$ denote the polynomials whose nodal values are given by $\bm{u},\bm{v}$.  Then, 
\begin{align*}
\bm{v}^T\bm{Q}\bm{u} &= \bm{v}^T\bm{W}\bm{D}\bm{u} = \int_{-1}^1 \pd{u}{x}v = \LRl{uv}_{-1}^1 - \int_{-1}^1 u\pd{v}{x} \\
&= \LRp{\bm{V}_f \bm{v}}^T\bm{B}\bm{V}_f\bm{u} - \LRp{\bm{D}\bm{v}}^T\bm{W}\bm{u} = \bm{v}^T\bm{V}_f^T\bm{B}\bm{V}_f \bm{u} - \bm{v}^T\bm{Q}^T\bm{u},
\end{align*}
where we have used that $\bm{W}$ is diagonal and that $(N+1)$-point Gauss quadrature is exact for the above integrands.  Taking $\bm{v} = \bm{e}_i$ and $\bm{u} = \bm{e}_j$ for $i,j = 1,\ldots, N+1$ (where $\bm{e}_i$ denotes the $i$th canonical vector) completes the proof.  
\end{proof}

Lemma~\ref{lemma:sbp} holds for both GLL and Gauss nodes, and switching between these two nodal sets simply results in a redefinition of the matrices $\bm{D}, \bm{V}_f$.  For example, because GLL nodes include boundary points, the interpolation matrix $\bm{V}_f$ reduces to a generalized permutation matrix which extracts the nodal values associated with the left and right endpoints.  %Lemma~\ref{lemma:sbp} can be extended directly to higher dimensions through a tensor product construction.

%However, there are some properties which are exclusive to Gauss nodes.  
%\begin{lemma}
%Let $p(x)$ be a linear function with Gauss nodal values $\bm{p}$.  Then, the Gauss node operators satisfy the following summation by parts property
%\[
%\diag{\bm{p}} \bm{Q} = \bm{V}_f^T \bm{B}\diag{\bm{V}_f\bm{p}} \bm{V}_f - \bm{Q}^T\diag{\bm{p}},
%\]
%where $\bm{B}$ is the boundary matrix introduced in Lemma~\ref{lemma:sbp}.
%\label{lemma:vertexmapped}
%\end{lemma}
%\begin{proof}
%Due to the fact that Gauss quadrature is exact for polynomial integrands of degree $(2N+1)$, the proof is very similar to that of Lemma~\ref{lemma:sbp}.  Let $\bm{u}, \bm{v}$ be vectors and let $u(x),v(x)$ denote polynomials with associated nodal values.  Then, 
%\begin{align*}
%\bm{v}^T\diag{\bm{p}} \bm{Q}\bm{u} &= \int_{-1}^1 \pd{u}{x} pv = 
%\LRl{ u pv}_{-1}^1 - \int_{-1}^1 u \pd{(pv)}{x}.
%\end{align*}
%The latter integrand is a polynomial of degree $2N$
%\begin{align*}
%&= \LRp{\bm{V}_f\bm{v}}^T \bm{B} \diag{\bm{V}_f\bm{p}} \bm{V}_f\bm{u}  - \LRp{\bm{D}\diag{\bm{p}}\bm{v}}^T\bm{W}\bm{u}.
%\end{align*}
%\end{proof}
%
%Both Lemma~\ref{lemma:sbp} and \ref{lemma:vertexmapped} extend directly to higher dimensions through tensor products.

\subsection{Existing entropy stable SBP-SATs for generalized SBP operators}
\label{sec:gsbpsat}

In this section, we will review the construction of semi-discretely entropy stable discretizations.  Entropy stable discretizations can be constructed by first introducing an entropy conservative scheme, then adding appropriate interface dissipation to produce an entropy inequality.  The construction of entropy conservative schemes relies on the existence of an two-point (dyadic) entropy conservative flux \cite{tadmor1987numerical}.  
\begin{definition}
Let $\bm{f}_S(\bm{u}_L,\bm{u}_R)$ be a bivariate function which is symmetric and consistent with the flux function $\bm{f}(\bm{u})$
\begin{align*}
\bm{f}_S(\bm{u}_L,\bm{u}_R) = \bm{f}_S(\bm{u}_R,\bm{u}_L), \qquad \bm{f}_S(\bm{u},\bm{u}) = \bm{f}(\bm{u})
\end{align*}
The numerical flux $\bm{f}_S(\bm{u}_L, \bm{u}_R)$ is entropy conservative if, for entropy variables $\bm{v}_L = \bm{v}(\bm{u}_L), \bm{v}_R = \bm{v}(\bm{u}_R)$, the Tadmor ``shuffle'' condition is satisfied
\begin{align*}
\LRp{\bm{v}_L - \bm{v}_R}^T \bm{f}_S(\bm{u}_L,\bm{u}_R) = (\psi_L - \psi_R), \qquad \psi_L = \psi(\bm{v}(\bm{u}_L)), \quad \psi_R = \psi(\bm{v}(\bm{u}_R)).  
\end{align*}
\label{def:tadmor}
\end{definition}
Here, $\psi$ is the entropy potential.

\note{Ask DCDR to fill this in \cite{fernandez2014review, crean2017high, fernandez2018simultaneous}.}

\subsection{Decoupled SBP operators}

The main goal of this work is to circumvent the all-to-all coupling of degrees of freedom introduced by the SBP-SAT terms described in Section~\ref{sec:gsbpsat}, and will do so through the use of ``decoupled'' SBP operators.  Decoupled SBP operators were first introduced in \cite{chan2017discretely} and used to construct entropy stable schemes on simplicial elements.  These operators (and simplifications under a collocation assumption) are presented in a more general setting in Appendix~\ref{app:decoupled}.  In this section, decoupled SBP operators are introduced in one dimension for either GLL or Gauss nodal sets.

Let $\bm{Q} = \bm{W}\bm{D}$.  The decoupled SBP operator $\bm{Q}_N$ is defined as the block matrix 
\begin{equation}
\bm{Q}_N  = \begin{bmatrix}
\bm{Q} - \frac{1}{2}\bm{V}_f^T\bm{B}\bm{V}_f & \frac{1}{2}\bm{V}_f^T\bm{B}\\
-\frac{1}{2}\bm{B}\bm{V}_f & \frac{1}{2}\bm{B}
\end{bmatrix}.
\label{eq:qndef}
\end{equation}
Lemma~\ref{lemma:sbp} and straightforward computations show that $\bm{Q}_N$ also satisfies the following SBP property
\begin{lemma}
\label{lemma:dsbp}
Let $\bm{Q}_N$ be defined through (\ref{eq:qndef}).  Then,
\[
\bm{Q}_N + \bm{Q}_N^T = \begin{bmatrix}
\bm{0} &\\
& \bm{B}
\end{bmatrix}.
\]
\end{lemma}

The matrix $\bm{Q}_N$ acts not only on volume nodes, but on both volume and surface nodes.  However, $\bm{Q}_N$ does not directly define an approximation to the derivative.  Let $f(x), g(x)$ denote two functions, and let $\bm{f}, \bm{g}$ denote the values of $f,g$ at interior nodal points.  We also define vectors $\bm{f}_N, \bm{g}_N$ denoting the values of $f,g$ at both interior and boundary points 
\begin{equation}
\bm{f}_N = \begin{bmatrix}
f\LRp{x_1}\\
\vdots\\
f\LRp{x_{N+1}}\\
f(-1)\\
f(1)
\end{bmatrix} = \begin{bmatrix}
\bm{f} \\
\bm{f}_f
\end{bmatrix}, \qquad
\bm{g}_N = \begin{bmatrix}
g\LRp{x_1}\\
\vdots\\
g\LRp{x_{N+1}}\\
g(-1)\\
g(1)
\end{bmatrix} = \begin{bmatrix}
\bm{g} \\
\bm{g}_f
\end{bmatrix}.
\label{eq:fg}
\end{equation}
Then, a polynomial approximation to $f\pd{g}{x}$ can be computed using $\bm{Q}_N$.  Let $\bm{u}$ denote the nodal values of the polynomial $u(x) \approx f\pd{g}{x}$.  These coefficients are computed via
\begin{equation}
\bm{W}\bm{u} = \begin{bmatrix}
\bm{I}\\
\bm{V}_f
\end{bmatrix}^T \diag{\bm{f}_N} \bm{Q}_N \bm{g}_N.
\label{eq:qn}
\end{equation}
The expression (\ref{eq:qn}) can be rewritten in ``strong'' form as follows
\begin{align*}
\bm{u} = \bm{W}^{-1}\begin{bmatrix}
\bm{I}\\
\bm{V}_f
\end{bmatrix}^T& \diag{\bm{f}_N} \bm{Q}_N \bm{g}_N \\
= \diag{\bm{f}}\bm{D}\bm{g} &+ \frac{1}{2}\diag{\bm{f}} \bm{W}^{-1} \bm{V}_f^T \bm{B}\LRp{\bm{g}_f - \bm{V}_f\bm{g}} \\
&+ \frac{1}{2}\bm{W}^{-1} \bm{V}_f^T\bm{B}\diag{\bm{f}_f} \LRp{\bm{g}_f - \bm{V}_f\bm{g}},
\end{align*}
where we have used the fact that diagonal matrices commute to simplify expressions.  Thus, the decoupled SBP operator $\bm{Q}_N$ can be interpreted as adding boundary corrections to the GSBP operator $\bm{D}$ in a skew-symmetric fashion.

It was shown in \cite{chan2017discretely} that $\bm{u}$ is a high order accurate approximation to the quantity $f\pd{g}{x}$.  Both the generalized SBP operator $\bm{D}$ and the expression in (\ref{eq:qn}) involving the decoupled SBP operator recover exact derivatives of high order polynomials.  However, when applied to non-polynomial functions, the decoupled SBP operator $\bm{Q}_N$ improves accuracy near the boundaries.  Figure~\ref{fig:dsbpcorrect} illustrates this by using both operators to approximate the derivative of a Gaussian $e^{-4x^2}$.  The decoupled SBP operator results in an improved approximation near the boundary.  %This can be interpreted as  \cite{brenner2007mathematical}

\begin{figure}
\centering
\begin{tikzpicture}
\begin{axis}[
    width=.6\textwidth,
    xlabel={$x$ coordinate},
%    ylabel={$L^2$ errors}, 
%    xmin=.0125, xmax=.75,
    ymin=-2, ymax=2.25,
    legend pos=north east, legend cell align=left, legend style={font=\tiny},	
    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
    legend entries={GSBP, Decoupled SBP, Exact}    
]
\pgfplotsset{
cycle list={{blue, only marks, mark=*}, {blue}, {red, only marks,mark=square*},{red},{black,dashed}}
}
\addlegendimage{blue, mark=*}
\addlegendimage{red, mark=square*}
\addlegendimage{black, dashed}

\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(-0.93247,-0.677965)(-0.661209,1.35632)(-0.238619,1.07263)(0.238619,-1.07263)(0.661209,-1.35632)(0.93247,0.677965)};

\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(-1,-1.56577)(-0.959184,-1.00891)(-0.918367,-0.51366)(-0.877551,-0.0774098)(-0.836735,0.302464)(-0.795918,0.628585)(-0.755102,0.903575)(-0.714286,1.13006)(-0.673469,1.31065)(-0.632653,1.44798)(-0.591837,1.54466)(-0.55102,1.60333)(-0.510204,1.62659)(-0.469388,1.61708)(-0.428571,1.57742)(-0.387755,1.51022)(-0.346939,1.41812)(-0.306122,1.30372)(-0.265306,1.16966)(-0.22449,1.01856)(-0.183673,0.85303)(-0.142857,0.675704)(-0.102041,0.489201)(-0.0612245,0.296143)(-0.0204082,0.0991513)(0.0204082,-0.0991513)(0.0612245,-0.296143)(0.102041,-0.489201)(0.142857,-0.675704)(0.183673,-0.85303)(0.22449,-1.01856)(0.265306,-1.16966)(0.306122,-1.30372)(0.346939,-1.41812)(0.387755,-1.51022)(0.428571,-1.57742)(0.469388,-1.61708)(0.510204,-1.62659)(0.55102,-1.60333)(0.591837,-1.54466)(0.632653,-1.44798)(0.673469,-1.31065)(0.714286,-1.13006)(0.755102,-0.903575)(0.795918,-0.628585)(0.836735,-0.302464)(0.877551,0.0774098)(0.918367,0.51366)(0.959184,1.00891)(1,1.56577)};

\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(-0.93247,-0.194074)(-0.661209,1.16978)(-0.238619,1.20913)(0.238619,-1.20913)(0.661209,-1.16978)(0.93247,0.194074)};

\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(-1,-0.434491)(-0.959184,-0.303266)(-0.918367,-0.130624)(-0.877551,0.0698403)(-0.836735,0.286078)(-0.795918,0.507515)(-0.755102,0.724981)(-0.714286,0.930644)(-0.673469,1.11794)(-0.632653,1.28151)(-0.591837,1.41712)(-0.55102,1.5216)(-0.510204,1.59278)(-0.469388,1.62942)(-0.428571,1.63111)(-0.387755,1.59825)(-0.346939,1.53198)(-0.306122,1.43404)(-0.265306,1.30681)(-0.22449,1.15314)(-0.183673,0.976347)(-0.142857,0.78012)(-0.102041,0.568459)(-0.0612245,0.345602)(-0.0204082,0.115959)(0.0204082,-0.115959)(0.0612245,-0.345602)(0.102041,-0.568459)(0.142857,-0.78012)(0.183673,-0.976347)(0.22449,-1.15314)(0.265306,-1.30681)(0.306122,-1.43404)(0.346939,-1.53198)(0.387755,-1.59825)(0.428571,-1.63111)(0.469388,-1.62942)(0.510204,-1.59278)(0.55102,-1.5216)(0.591837,-1.41712)(0.632653,-1.28151)(0.673469,-1.11794)(0.714286,-0.930644)(0.755102,-0.724981)(0.795918,-0.507515)(0.836735,-0.286078)(0.877551,-0.0698403)(0.918367,0.130624)(0.959184,0.303266)(1,0.434491)};

\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(-1,0.146525)(-0.959184,0.193522)(-0.918367,0.251752)(-0.877551,0.322529)(-0.836735,0.406852)(-0.795918,0.50522)(-0.755102,0.617438)(-0.714286,0.742415)(-0.673469,0.877994)(-0.632653,1.02082)(-0.591837,1.1663)(-0.55102,1.30861)(-0.510204,1.44089)(-0.469388,1.55553)(-0.428571,1.64452)(-0.387755,1.70003)(-0.346939,1.71492)(-0.306122,1.68342)(-0.265306,1.60163)(-0.22449,1.46805)(-0.183673,1.2839)(-0.142857,1.05327)(-0.102041,0.783025)(-0.0612245,0.482507)(-0.0204082,0.162994)(0.0204082,-0.162994)(0.0612245,-0.482507)(0.102041,-0.783025)(0.142857,-1.05327)(0.183673,-1.2839)(0.22449,-1.46805)(0.265306,-1.60163)(0.306122,-1.68342)(0.346939,-1.71492)(0.387755,-1.70003)(0.428571,-1.64452)(0.469388,-1.55553)(0.510204,-1.44089)(0.55102,-1.30861)(0.591837,-1.1663)(0.632653,-1.02082)(0.673469,-0.877994)(0.714286,-0.742415)(0.755102,-0.617438)(0.795918,-0.50522)(0.836735,-0.406852)(0.877551,-0.322529)(0.918367,-0.251752)(0.959184,-0.193522)(1,-0.146525)};
\end{axis}
\end{tikzpicture}

\caption{Degree $N= 5$ approximations of derivatives of a Gaussian $e^{-4x^2}$ using the generalized SBP operator $\bm{D}$ and the decoupled SBP operator $\bm{Q}_N$ via (\ref{eq:qn}). }
\label{fig:dsbpcorrect}
\end{figure}

\subsection{An entropy stable Gauss collocation scheme based on decoupled SBP operators}

We can now construct an entropy conservative Gauss collocation scheme with compact SBP-SATs using decoupled SBP operators.  Let $\bm{u}_N$ and $\bm{v}(\bm{u}_N)$ denote the values of the conservative variables and entropy variables at Gauss points, respectively.  We define approximate values $\tilde{\bm{u}}_f$ of the conservative variables by interpolating the values of the entropy variables to element boundaries 
\[
\tilde{\bm{u}}_f = \bm{u}(\bm{v}_f), \qquad \bm{v}_f = \bm{V}_f\bm{v}(\bm{u}_N).  
\]
We now define $\tilde{\bm{u}}$ as the concatenated vector of solution values at Gauss points and the ``entropy-projected conservative variables'' $\tilde{\bm{u}}_f$.  We also define $\bm{F}_S$ as the matrix of evaluations of the two-point flux $\bm{f}_S$ at combinations of values of $\tilde{\bm{u}}$
\[
\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad \tilde{\bm{u}} = \begin{bmatrix}
\bm{u}_N\\
\tilde{\bm{u}}_f
\end{bmatrix}.
\]
\begin{theorem}
\label{thm:consentropy}
Let $\circ$ denote the Hadamard product, and let $\tilde{\bm{u}}_f^+$ denote values of the entropy-projected conservative variables on neighboring elements.  Then, the formulation
\begin{equation}
\label{eq:form}
\bm{W}\td{\bm{u}_N}{t} + 2 \begin{bmatrix}
\bm{I}\\
\bm{V}_f
\end{bmatrix}^T\LRp{\bm{Q}_N \circ \bm{F}_S}\bm{1} + \bm{V}_f^T \bm{B} \LRp{\bm{f}_S\LRp{\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f} - \bm{f}(\bm{u})} = 0
\end{equation}
satisfies the quadrature form of the semi-discrete conservation of entropy
\[
\bm{1}^T\bm{W}\td{S(\bm{u}_N)}{t} + \bm{1}^T\bm{B} \LRp{\bm{v}_f^T\bm{f}(\tilde{\bm{u}}_f) - \psi_i\LRp{\tilde{\bm{u}}_f}} = 0.
\]
\end{theorem}
\begin{proof}
Theorem~\ref{thm:consentropy} follows from choosing either GLL or Gauss quadratures in Theorem 4 of \cite{chan2017discretely}.  We provide a separate illustrative proof on a two-element mesh.  The extension to multiple elements is straightforward.

First, we rewrite formulation (\ref{eq:form}) in a skew-symmetric form.  Using Lemma~\ref{lemma:dsbp} and the diagonal nature of $\bm{B}$, (\ref{eq:form}) is equivalent to solving
\begin{equation}
\bm{W}\td{\bm{u}_N}{t} + \begin{bmatrix}
\bm{I}\\
\bm{V}_f
\end{bmatrix}^T\LRp{\LRp{\bm{Q}_N - \bm{Q}_N^T} \circ \bm{F}_S}\bm{1} + \bm{V}_f^T \bm{B} {\bm{f}_S\LRp{\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f}} = 0
\label{eq:skew}
\end{equation}

\end{proof}

\note{Add two-element case.}

A semi-discrete entropy inequality can be enforced by adding appropriate interface dissipation terms, such as Lax-Friedrichs or matrix dissipation \cite{winters2017uniquely}.  We note that these terms must be computed in terms of $\tilde{\bm{u}}_f$ in order to ensure a discrete dissipation of entropy \cite{chen2017entropy,chan2017discretely}.

\section{Extension to higher dimensions and curved meshes}
\label{sec:2}

Theorem~\ref{thm:consentropy}

\note{Describe GCL using GLL nodes \cite{chan2018discretely}}

Retaining entropy stability on curved meshes requires satisfaction of a discrete geometric conservation law (GCL) \cite{carpenter2014entropy, gassner2017br1, crean2018entropy, chan2018discretely}.  For two-dimensional isoparametric mappings, the GCL is automatically satisfied.  However, in three dimensions, the GCL is not guaranteed to be satisfied at the discrete level, due to the fact that geometric terms are polynomials of degree higher than $N$.  Modifications to the computation of geometric terms are required to ensure that the GCL is satisfied at the discrete level.

For general SBP operators, the discrete GCL can be enforced through the solution of a local least squares problem \cite{crean2018entropy}.  We take an alternative approach and  construct geometric terms using the approach of Kopriva \cite{kopriva2006metric}.  This takes advantage of the fact that the collocation methods constructed in this work correspond to polynomial discretizations.  Kopriva's construction is based on rewriting the geometric terms as the curl of some auxiliary quantity.  By interpolating this auxiliary quantity prior to applying the curl,

To summarize, entropy stability curved meshes
\begin{enumerate}
\item Construct polynomial approximations of the geometric terms using interpolation at GLL nodes \cite{kopriva2006metric}.  
\item Interpolate the geometric terms to volume and surface quadrature points.  
\item \note{Split form}.
\end{enumerate}

\section{Numerical results}
\label{sec:3}
The compressible Euler equations in $d$ dimensions are given as follows:
\begin{align*}
\pd{\rho}{t} &+ \sum_{j=1}^d \pd{\LRp{\rho \bm{u}_j}}{x_j} = 0,\\
\pd{\rho \bm{u}_i}{t} &+ \sum_{j=1}^d \pd{\LRp{\rho \bm{u}_i\bm{u}_j + p\delta_{ij} }}{x_j} = 0, \qquad i = 1,\ldots,d\\
\pd{E}{t} &+ \sum_{j=1}^d \pd{\LRp{\bm{u}_j(E+p)}}{x_j} = 0.\nonumber
\end{align*}
Here, $\rho$ is density, $\bm{u} = \LRp{\bm{u}_1, \ldots,\bm{u}_d}$ is the vector of velocities, and $E$ is the total energy.  The pressure $p$ and specific internal energy $\rho e$ are given by 
\[
p = (\gamma-1)\LRp{E - \frac{1}{2}\rho \LRb{\bm{u}}^2}, \qquad \rho e = E - \frac{1}{2}\rho \LRb{\bm{u}}^2, \qquad \LRb{\bm{u}}^2 = \LRp{\sum_{j=1}^d \bm{u}_j^2}.
\]

There exists an infinite family of suitable convex entropies for the compressible Euler equations \cite{harten1983symmetric}.  However, there is only a single unique entropy which appropriately treats the viscous heat conduction term in the compressible Navier-Stokes equations \cite{hughes1986new}.  This entropy $S(\bm{u})$ is given in $d$ dimensions by
\begin{equation*}
S(\bm{u}) = -\frac{\rho s}{\gamma-1},
\end{equation*}
where $s = \log\LRp{\frac{p}{\rho^\gamma}}$ is the physical specific entropy.  The entropy variables in $d$ dimensions are given by
\begin{align*}
v_1 &= \frac{\rho e (\gamma + 1 - s) - E}{\rho e}, \\
v_{1+i}&= \frac{\rho \bm{u}_i}{\rho e}, \qquad i = 1,\ldots, d,\\
v_{d+2} &= -\frac{\rho}{\rho e},
\end{align*}
while the conservation variables in terms of the entropy variables are given by
\begin{align*}
\rho &= -(\rho e) v_{d+2}, \\
 \rho \bm{u}_i &= (\rho e) v_{1+i}, \qquad i = 1,\ldots,d\\
 E &= (\rho e)\LRp{1 - \frac{\sum_{j=1}^d{v_{1+j}^2}}{2 v_{d+2}}},
\end{align*}
where the quantities $\rho e$ and $s$ in terms of the entropy variables are 
\begin{equation*}
\rho e = \LRp{\frac{(\gamma-1)}{\LRp{-v_{d+2}}^{\gamma}}}^{1/(\gamma-1)}e^{\frac{-s}{\gamma-1}}, \qquad s = \gamma - v_1 + \frac{\sum_{j=1}^d{v_{1+j}^2}}{2v_{d+2}}.
\end{equation*}
Explicit expressions for entropy conservative numerical fluxes in two dimensions  are given by Chandrashekar \cite{chandrashekar2013kinetic}
\begin{align*}
&f^1_{1,S}(\bm{u}_L,\bm{u}_R) = \avg{\rho}^{\log} \avg{u},& &f^1_{2,S}(\bm{u}_L,\bm{u}_R) = \avg{\rho}^{\log} \avg{v},&\\
&f^2_{1,S}(\bm{u}_L,\bm{u}_R) = f^1_{1,S} \avg{u} + p_{\rm avg},&  &f^2_{2,S}(\bm{u}_L,\bm{u}_R) = f^1_{2,S} \avg{u},&\nonumber\\
&f^3_{1,S}(\bm{u}_L,\bm{u}_R) = f^2_{2,S},& &f^3_{2,S}(\bm{u}_L,\bm{u}_R) = f^1_{2,S} \avg{v} + p_{\rm avg},&\nonumber\\
&f^4_{1,S}(\bm{u}_L,\bm{u}_R) = \LRp{E_{\rm avg} + p_{\rm avg}}\avg{u},& &f^4_{2,S}(\bm{u}_L,\bm{u}_R) = \LRp{E_{\rm avg} + p_{\rm avg} }\avg{v},& \nonumber
\end{align*}
where we have defined the auxiliary quantities 
\begin{gather*}
p_{\rm avg} = \frac{\avg{\rho}}{2\avg{\beta}}, \qquad E_{\rm avg} = \frac{\avg{\rho}^{\log}}{2\avg{\beta}^{\log}\LRp{\gamma -1}}   + \frac{\nor{\bm{u}}^2_{\rm avg}}{2}, \\
 \nor{\bm{u}}^2_{\rm avg} = 2(\avg{u}^2 + \avg{v}^2) - \LRp{\avg{u^2} +\avg{v^2}} \nonumber.  
\end{gather*}
Expressions for entropy conservative numerical fluxes for the three-dimensional compressible Euler equations can also be explicitly written as
\begin{gather*}
\bm{f}_{1,S} = \LRp{\begin{array}{c}
\avg{\rho}^{\log}\avg{u}\\
\avg{\rho}^{\log}\avg{u}^2 + p_{\rm avg}\\
\avg{\rho}^{\log}\avg{u}\avg{v}\\
\avg{\rho}^{\log}\avg{u}\avg{w}\\
(E_{\rm avg}+ p_{\rm avg})\avg{u}\\
\end{array}}, 
\qquad 
\bm{f}_{2,S} = \LRp{\begin{array}{c}
\avg{\rho}^{\log}\avg{v}\\
\avg{\rho}^{\log}\avg{u}\avg{v}\\
\avg{\rho}^{\log}\avg{v}^2 + p_{\rm avg}\\
\avg{\rho}^{\log}\avg{v}\avg{w}\\
(E_{\rm avg}+ p_{\rm avg})\avg{v}\\
\end{array}},\\
\bm{f}_{3,S} = \LRp{\begin{array}{c}
\avg{\rho}^{\log}\avg{w}\\
\avg{\rho}^{\log}\avg{u}\avg{w}\\
\avg{\rho}^{\log}\avg{v}\avg{w}\\
\avg{\rho}^{\log}\avg{w}^2 + p_{\rm avg}\\
(E_{\rm avg}+ p_{\rm avg})\avg{w}\\
\end{array}}.\nonumber
%&f^1_{1,S}(\bm{u}_L,\bm{u}_R) = \avg{\rho}^{\log} \avg{u},& &f^1_{2,S}(\bm{u}_L,\bm{u}_R) = \avg{\rho}^{\log} \avg{v},&\\
%&f^2_{1,S}(\bm{u}_L,\bm{u}_R) = f^1_{1,S} \avg{u} + p_{\rm avg},&  &f^2_{2,S}(\bm{u}_L,\bm{u}_R) = f^1_{2,S} \avg{u},&\nonumber\\
%&f^3_{1,S}(\bm{u}_L,\bm{u}_R) = f^2_{2,S},& &f^3_{2,S}(\bm{u}_L,\bm{u}_R) = f^1_{2,S} \avg{v} + p_{\rm avg},&\nonumber\\
%&f^4_{1,S}(\bm{u}_L,\bm{u}_R) = \LRp{\frac{p_{\rm avg}^{\log}}{\gamma -1} + p_{\rm avg} + \frac{\nor{\bm{u}}^2_{\rm avg}}{2}}\avg{u},& &f^4_{2,S}(\bm{u}_L,\bm{u}_R) = \LRp{\frac{p_{\rm avg}^{\log}}{\gamma -1} + p_{\rm avg} + \frac{\nor{\bm{u}}^2_{\rm avg}}{2}}\avg{v},& \nonumber
\end{gather*}
with auxiliary quantities 
\begin{align*}
p_{\rm avg} &= \frac{\avg{\rho}}{2\avg{\beta}}, \qquad E_{\rm avg} = \frac{\avg{\rho}^{\log}}{2(\gamma-1)\avg{\beta}^{\log}} + \frac{1}{2}\avg{\rho}^{\log}\nor{\bm{u}}^2_{\rm avg}\\
\nor{\bm{u}}^2_{\rm avg} &= 2(\avg{u}^2 + \avg{v}^2 + \avg{w}^2) - \LRp{\avg{u^2} +\avg{v^2} + \avg{w^2}}.\nonumber
\end{align*}

In all problems, we estimate the timestep size $dt$ using $J, J^f$, and degree-dependent $L^2$ trace constants $C_N$ 
\[
dt = C_{\rm CFL}\frac{ h}{C_N}, \qquad h = \frac{ \nor{J^f}_{L^{\infty}}}{\nor{J^{-1}}_{L^{\infty}}},
\]
where $C_{\rm CFL}$ is some user-defined CFL constant.  For isotropic elements, the ratio of $J$ to $J^f$ scales as the mesh size $h$, while $C_N$ captures the dependence of the maximum stable timestep on the polynomial degree $N$.  For hexahedral elements, $C_N$ varies depending on the choice of quadrature.  It was shown in \cite{chan2015gpu} that 
\[
C_N =\begin{cases}
 d\frac{N(N+1)}{2} & \text{for GLL nodes}\\
d\frac{(N+1)(N+2)}{2} & \text{for Gauss nodes}
\end{cases}.
\]
Thus, based on this rough estimate of the maximum stable timestep, GLL collocation schemes should be able to take a timestep which is roughly $(1 + 2/N)$ times larger than the maximum stable timestep for Gauss collocation schemes.  We do not account for this discrepancy in this work, and instead set the timestep for both GLL and Gauss collocation schemes based on the more conservative Gauss collocation estimate of $dt$.  %Unless stated otherwise, all experiments utilize $C_{\rm CFL} = .5$.

\subsection{2D isentropic vortex problem}

We begin by examining high order convergence of the proposed methods in two dimensions using the isentropic vortex problem \cite{shu1998essentially, crean2017high}.  The analytical solution is 
\begin{align}
\rho(\bm{x},t) &= \LRp{1 - \frac{\frac{1}{2}(\gamma-1)(\beta e^{1-r(\bm{x},t)^2})^2}{8\gamma \pi^2}}^{\frac{1}{\gamma-1}}, \qquad p = \rho^{\gamma},\\
u(\bm{x},t) &= 1 - \frac{\beta}{2\pi} e^{1-r(\bm{x},t)^2}(y-y_0), \qquad v(\bm{x},t) = \frac{\beta}{2\pi} e^{1-r(\bm{x},t)^2}(y-y_0),\nonumber
\end{align}
where $u, v$ are the $x$ and $y$ velocity and $r(\bm{x},t) = \sqrt{(x-x_0-t)^2 + (y-y_0)^2}$.  Here, we take $x_0 = 5, y_0 = 0$ and $\beta = 5$.

\begin{figure}
\centering
\subfloat[Moderately warped mesh]{\includegraphics[width=.45\textwidth]{figs/warp2d_a125.png}}
\hspace{1em}
\subfloat[Heavily warped mesh]{\includegraphics[width=.45\textwidth]{figs/warp2d_a25.png}}
\caption{Examples of moderate and heavily warped meshes for $N=4, K= 16$.  }
\label{fig:warp2d}
\end{figure}

We solve on a periodic rectangular domain $[0, 20] \times [-5,5]$ until final time $T=5$, and compute errors over all all solution fields.  For a degree $N$ approximation, we approximate the $L^2$ error using an $(N+2)$ point Gauss quadrature rule.  We also examine the influence of element curvature for both GLL and Gauss collocation schemes by examining $L^2$ errors on a sequence of moderately and heavily warped curvilinear meshes (see Figure~\ref{fig:warp2d}).  The following results are computed using $C_{\rm CFL} = 1/2$ and Lax-Friedrichs interface dissipation.

%\begin{center}
%\begin{tabular}{|c || c ||c ||c ||c ||c ||c |}
%\hline
%& $N=2$ & $N=3$& $N=4$& $N=5$& $N=6$& $N=7$\\
%\hline\hline
%$h = 1/4$ & $1.229e{0}$ & $4.144e{-1}$ & $1.235e{-1}$ & $3.325e{-2}$ & $1.068e{-2}$ & $3.309e{-3}$ \\
%$h = 1/8$ & $2.366e{-1}$ & $4.122e{-2}$ & $4.393e{-3}$ & $9.701e{-4}$ & $1.333e{-4}$ & $1.678e{-5}$ \\
%$h = 1/16$ & $2.589e{-2}$ & $2.930e{-3}$ & $1.269e{-4}$ & $2.103e{-5}$ & $7.970e{-7}$ & $1.327e{-7}$   
%\end{tabular}
%\end{center}
\begin{figure}
\centering
\subfloat[Affine mesh]{
\begin{tikzpicture}
\begin{loglogaxis}[
    width=.5\textwidth,
    xlabel={Mesh size $h$},
    ylabel={$L^2$ errors}, 
%    xmin=.0125, 
    ymin=5e-8, ymax=5,
    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
    legend entries={GLL, Gauss}    
]
\pgfplotsset{
cycle list={{blue, mark=*}, {red, dashed ,mark=square*}}
}
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,1.22861)(0.0625,0.236646)(0.03125,0.0258883)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.553435)(0.0625,0.0625441)(0.03125,0.0117156)};

\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.414449)(0.0625,0.0412224)(0.03125,0.00293039)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.25082)(0.0625,0.0198863)(0.03125,0.000673152)};

\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.123516)(0.0625,0.00439285)(0.03125,0.000126852)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.0768018)(0.0625,0.0020579)(0.03125,5.55322e-05)};

\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.0332509)(0.0625,0.000970148)(0.03125,2.10334e-05)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.0151231)(0.0625,0.000454595)(0.03125,1.00203e-05)};

\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.0106752)(0.0625,0.000133284)(0.03125,7.97032e-07)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.00657096)(0.0625,9.22845e-05)(0.03125,3.60025e-07)};

\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.00330917)(0.0625,1.67825e-05)(0.03125,1.32696e-07)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.00185273)(0.0625,1.21465e-05)(0.03125,1.14289e-07)};
\end{loglogaxis}
\end{tikzpicture}
}
\subfloat[Moderately warped mesh]{
\begin{tikzpicture}
\begin{loglogaxis}[
    width=.5\textwidth,
    xlabel={Mesh size $h$},
%    ylabel={$L^2$ errors}, 
%    xmin=.0125,
    ymin=5e-8, ymax=5,    
    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
    legend entries={GLL, Gauss}    
]
\pgfplotsset{
cycle list={{blue, mark=*}, {red, dashed ,mark=square*}}
}
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,1.64274)(0.0625,0.445945)(0.03125,0.0720148)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.882247)(0.0625,0.138116)(0.03125,0.0268145)};

\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.662165)(0.0625,0.146131)(0.03125,0.0234934)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.368084)(0.0625,0.0734059)(0.03125,0.00420438)};

\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.349184)(0.0625,0.0508268)(0.03125,0.00234785)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.183472)(0.0625,0.0127617)(0.03125,0.000251929)};

\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.171986)(0.0625,0.0103578)(0.03125,0.000386834)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.0924215)(0.0625,0.00276519)(0.03125,6.80528e-05)};

\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.0729963)(0.0625,0.00276104)(0.03125,4.11305e-05)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.0378013)(0.0625,0.0007649)(0.03125,4.15848e-06)};

\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.026577)(0.0625,0.000589552)(0.03125,5.3039e-06)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.0154563)(0.0625,0.000146207)(0.03125,8.83156e-07)};

\end{loglogaxis}
\end{tikzpicture}
}\\
\subfloat[Heavily warped mesh]{
\begin{tikzpicture}
\begin{loglogaxis}[
    width=.5\textwidth,
    xlabel={Mesh size $h$},
%    ylabel={$L^2$ errors}, 
%    xmin=.0125,
    ymin=5e-8, ymax=5,    
    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
    legend entries={GLL, Gauss}    
]
\pgfplotsset{
cycle list={{blue, mark=*}, {red, dashed ,mark=square*}}
}
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,2.08578)(0.0625,1.33668)(0.03125,0.433653)(0.015625,0.0721002)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,1.82833)(0.0625,0.746283)(0.03125,0.11006)(0.015625,0.0110115)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,1.74649)(0.0625,0.546097)(0.03125,0.0879841)(0.015625,0.0090851)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,1.39268)(0.0625,0.245852)(0.03125,0.0255562)(0.015625,0.00165226)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,1.06289)(0.0625,0.218361)(0.03125,0.0209471)(0.015625,0.000665425)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.740071)(0.0625,0.105075)(0.03125,0.003505)(0.015625,8.26421e-05)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.695283)(0.0625,0.0996971)(0.03125,0.00390099)(0.015625,0.000118992)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.457324)(0.0625,0.0354479)(0.03125,0.000767746)(0.015625,1.95823e-05)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.449493)(0.0625,0.0374955)(0.03125,0.000909834)(0.015625,7.70064e-06)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.274721)(0.0625,0.0120499)(0.03125,0.000157175)(0.015625,9.61243e-07)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.294238)(0.0625,0.0162153)(0.03125,0.000166829)(.015625,1.19661e-06)};
\addplot+[semithick, mark options={solid, fill=markercolor}]
coordinates{(0.125,0.150916)(0.0625,0.00538908)(0.03125,2.79182e-05)};
\end{loglogaxis}
\end{tikzpicture}
}
\caption{$L^2$ errors for the 2D isentropic vortex at time $T=5$ for degree $N = 2,\ldots,7$ GLL and Gauss collocation schemes.}
\label{fig:err2d}
\end{figure}

%\note{Investigate robustness with respect to mesh distortion: GLL vs GQ errors.  GQ is better for more curved meshes.  Note that there is little difference between bilinear/trilinear and high order curved.  }

\subsection{3D isentropic vortex problem}

As in two dimensions, we test the accuracy of the proposed scheme using an isentropic vortex solution adapted to three dimensions.  We take the solution to be the extruded 2D vortex propagating in the $x_2$ direction, whose analytic expression is derived from \cite{williams2013nodal}
\begin{align*}
\rho(\bm{x},t) &= \LRp{1-\frac{(\gamma-1)}{2}\Pi^2}^{\frac{1}{\gamma-1}}\\
\bm{u}(\bm{x},t) &= \Pi \bm{r}, \\
E(\bm{x},t) &= \frac{p_0}{\gamma-1}\LRp{1-\frac{\gamma-1}{2}\Pi^2}^{\frac{\gamma}{\gamma-1}} + \frac{\rho}{2}\LRb{\bm{u}}.
\end{align*}
where $\bm{u} = (u,v,w)^T$ is the velocity vector and % $\Pi$, $\bm{r}$ are defined as
\[
\Pi = \Pi_{\max}e^{\frac{1-\bm{r}^T\bm{r}}{2}}, \qquad \bm{r} = \begin{pmatrix}
-(x_2-c_2-t)\\
x_1-c_1\\
0
\end{pmatrix}.
\]
In this problem, we take $c_1 = c_2 = 5$, $p_0 = {1}/{\gamma}$, and $\Pi_{\max} = 0.4$.  The problem is solved on the domain $[0,10]\times [0,20]\times [0,10]$.  As in the 2D case, we also examine the effect of curvilinear mesh warping.  As in \cite{chan2018discretely}, we construct a curved warping of the initial Cartesian mesh by mapping nodes on each hexahedron to warped nodal positions $(\tilde{x_1},\tilde{x_2},\tilde{x_3})$ through the transformation
\begin{align*}
\tilde{\bm{x}} &= \bm{x} + \frac{1}{2}\sin\LRp{\pi \frac{(x_1-5)}{5}}\sin\LRp{2\pi \frac{(x_2-10)}{10}}\sin\LRp{\pi \frac{(x_3-5)}{5}}.
\end{align*}

\subsection{Shock-vortex interaction}

\subsection{Inviscid Taylor-Green vortex}

We conclude by investigating the behavior of entropy stable Gauss collocation schemes for the inviscid Taylor--Green vortex \cite{ae1937mechanism, gassner2016split, crean2018entropy}.  This problem is posed on the periodic box $[-\pi,\pi]^3$, with initial conditions 
\begin{align*}
\rho &= 1\\
u &= \sin(x_1)\cos(x_2)\cos(x_3),\\
v &= -\cos(x_1)\sin(x_2)\cos(x_3),\\ 
w &= 0,\\
p &= \frac{100}{\gamma} + \frac{1}{16} \LRp{\cos(2x_1) + \cos(2x_2)}\LRp{2+\cos(2x_3)}.
\end{align*}
The Taylor--Green vortex is used to study the transition and decay of turbulence \cite{debonis2013solutions}.  Under an inviscid assumption, the Taylor--Green vortex develops successively smaller scales.  As a result, the solution is guaranteed to contain under-resolved features after a sufficiently large time.  We study the evolution of the kinetic energy $\kappa(t)$ 
\[
\kappa(t) =\frac{1}{\LRb{\Omega}} \int_{\Omega} \rho \bm{u}\cdot\bm{u} \diff{\bm{x}},
\]
as well as the kinetic energy dissipation rate $-\pd{\kappa}{t}$, which is approximated by differencing $\kappa(t)$.

\section{Conclusion}

\note{blah}

\appendix

\section{Decoupled SBP operators}
\label{app:decoupled}

For general choices of quadrature and basis, decoupled projection operators involve a volume quadrature interpolation matrix $\bm{V}_q$, a face quadrature interpolation matrix $\bm{V}_f$, and a quadrature-based $L^2$ projection matrix $\bm{P}_q$.  Let $\LRc{\phi_j}_{j=1}^{N_p}$ denote a set of $N_p$ basis functions, and let $\LRc{\bm{x}_i, \bm{w}_i}_{i = 1}^{N_q}$ denote a set of $N_q$ volume quadrature points and weights in $d$ dimensions.  
Then, $\bm{V}_q, \bm{V}_f$ are given as 
\begin{align*}
\LRp{\bm{V}_q}_{ij} &= \phi_j(\bm{x}_i), \quad 1 \leq i,j \leq N_q,\\
\LRp{\bm{V}_f}_{ij} &= \phi_j\LRp{\bm{x}^f_i}, \quad 1 \leq i \leq N_q, \quad 1\leq j \leq N^f_q.
\end{align*}
These matrices can be used to define the quadrature-based $L^2$ projection matrix $\bm{P}_q$.  Let $\bm{W}$ denote the diagonal matrix of quadrature weights.  Then, 
\[
\bm{M} = \bm{V}_q^T\bm{W}\bm{V}_q, \qquad \bm{P}_q = \bm{M}^{-1}\bm{V}_q^T\bm{W}.  
\]
Let $\bm{D}^i$ now denote a modal differentiation matrix with respect to the $i$th coordinate, which maps coefficients in the basis $\phi_j$ to coefficients of the $i$th derivative.  By composing this matrix with interpolation and projection matrices, one can define differencing operators $\bm{D}_q^i = \bm{V}_q\bm{D}^i\bm{P}_q$ which map values at quadrature points to values of approximate derivatives at quadrature points. It was shown in \cite{chan2017discretely} that $\bm{Q}^i = \bm{W}\bm{D}_q^i$ satisfies a generalized SBP property involving the face interpolation and projection matrices $\bm{V}_f, \bm{P}_q$.

The decoupled SBP operator $\bm{Q}_N^i$ is then given as
\[
\bm{Q}_N^i = \begin{bmatrix}
\bm{Q}^i - \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q}^T\bm{W}_f \diag{\bm{n}_i}{\bm{V}_f\bm{P}_q} & \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q}^T\bm{W}_f \diag{\bm{n}_i}\\
-\frac{1}{2}\bm{W}_f \diag{\bm{n}_i}{\bm{V}_f\bm{P}_q} & \frac{1}{2}\bm{W}_f \diag{\bm{n}_i}
\end{bmatrix}.
\]
A straightforward computation shows that $\bm{Q}_N^i$ satisfies an SBP property \cite{chan2017discretely}.  
%\[
%\bm{Q}_N^i + \LRp{\bm{Q}_N^i}^T = \begin{bmatrix}\bm{0} &\\
%& \bm{W}_f \diag{\bm{n}_i}\end{bmatrix}.
%\]

It is worth noting that the form of $\bm{Q}_N^i$ does not depend on the choice of basis.  So long as the approximation space spanned by the basis $\phi_j$ does not change, the domain and range of $\bm{Q}_N^i$ depend solely on the choice of volume and surface quadrature points.

Assuming that the solution is represented using degree $N$ Lagrange polynomials at $(N+1)$ quadrature points, these matrices reduce down to
\[
\LRp{\bm{V}_q}_{ij} = \delta_{ij}, \qquad \bm{M} = \bm{W}, \qquad \bm{P}_q = \bm{M}^{-1}\bm{V}_q^T\bm{W} = \bm{I}.
\]

\note{Todo: add.}

\bibliographystyle{unsrt}
\bibliography{dg}

\end{document}

### 2018-07-19 docs/ESDG_GQ/ext_decoupled_sbp.tex
updating ESDG GQ paper

--- 2018-05-28 docs/skewform/main.tex
@@
 
-\documentclass[preprint,10pt]{elsarticle}
-%\documentclass[preprint,10pt]{article}
-%\documentclass[review]{siamart0216}
+%\documentclass[preprint,10pt]{elsarticle}
+%\documentclass[preprint,10pt]{amsart}
+\documentclass[review]{siamart0216}
 %\documentclass{siamart0216}
 
-\usepackage{fullpage}
-\usepackage[colorlinks=true]{hyperref}
+%\usepackage{fullpage}
+%\usepackage[colorlinks=true]{hyperref}
 
-\usepackage{amsmath,amssymb,amsfonts,amsthm}
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}
-\theoremstyle{lemma}
-\newtheorem{lemma}{Lemma}
-\theoremstyle{corollary}
-\newtheorem{corollary}{Corollary}
-\newtheorem*{remark}{Remark}
-\theoremstyle{theorem}
-\newtheorem{theorem}{Theorem}
-\theoremstyle{assumption}
-\newtheorem{assumption}{Assumption}
+\usepackage{amsmath,amssymb,amsfonts}
+%\usepackage{amsthm}
+%\theoremstyle{definition}
+%\newtheorem{definition}{Definition}
+%\theoremstyle{lemma}
+%\newtheorem{lemma}{Lemma}
+%\theoremstyle{corollary}
+%\newtheorem{corollary}{Corollary}
+%\newtheorem*{remark}{Remark}
+%\theoremstyle{theorem}
+%\newtheorem{theorem}{Theorem}
+%\theoremstyle{assumption}
+%\newtheorem{assumption}{Assumption}
 
@@
 
-\renewcommand{\hat}[1]{\widehat{#1}}
+\renewcommand{\tilde}{\widetilde}
+\renewcommand{\hat}{\widehat}
 \renewcommand{\topfraction}{0.85}
@@
 %\date{}
-%\author{Jesse Chan}
-%\title{On discretely entropy conservative discontinuous Galerkin methods}
+\author{Jesse Chan, David C.\ Del Rey Fernandez}
+\title{Entropy stable skew-symmetric discontinuous Galerkin formulations on non-conforming meshes}
 \graphicspath{{./figs/}}
@@
 
-%\maketitle
+\maketitle
 
-\begin{frontmatter}
-\title{Skew-symmetric formulations of entropy stable discontinuous Galerkin methods with applications to hybrid and non-conforming meshes. }
-
-\author[rice]{Jesse Chan\corref{cor1}}
-\ead{Jesse.Chan@caam.rice.edu}
-\address[rice]{Department of Computational and Applied Mathematics, Rice University, 6100 Main St, Houston, TX, 77005}
-
-\begin{abstract}
-Entropy stable high order methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.  
-\end{abstract}
-\end{frontmatter}
+%\begin{frontmatter}
+%\title{Skew-symmetric formulations for entropy stable discontinuous Galerkin methods on non-conforming meshes}
+%
+%\author[rice]{Jesse Chan\corref{cor1}}
+%\ead{Jesse.Chan@caam.rice.edu}
+%\address[rice]{Department of Computational and Applied Mathematics, Rice University, 6100 Main St, Houston, TX, 77005}
+%
+%\begin{abstract}
+%Entropy stable high order methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.  
+%\end{abstract}
+%\end{frontmatter}
 
@@
 
-\section{Entropy stability for systems of nonlinear conservation laws}
+\section{Entropy stable Gauss-Legendre collocation methods}
 
-We begin by reviewing the dissipation of entropy for a $d$-dimensional system of nonlinear conservation laws on a domain $\Omega$
-\begin{equation}
-\pd{\bm{u}}{t}  + \sum_{j=1}^d\pd{\bm{f}_j(\bm{u})}{x_j} = 0, \qquad \bm{u}\in \mathbb{R}^n, \qquad \bm{f}:\mathbb{R}^n\rightarrow\mathbb{R}^n,
-\label{eq:nonlineqs}
-\end{equation}
-where $\bm{u}$ are the conservative variables and $\bm{f}(\bm{u})$ is a vector-valued nonlinear flux function.
-We are interested in nonlinear conservation laws for which a convex entropy function $U(\bm{u})$ exists.
-For such systems, the  \emph{entropy variables} are an invertible mapping $\bm{v}(\bm{u}):\mathbb{R}^n\rightarrow \mathbb{R}^n$ defined as the derivative of the entropy function with respect to the conservative variables
-\begin{align}
-\bm{v}(\bm{u}) = \pd{U}{\bm{u}}.%, \qquad  \text{ invertible}.%\bm{u}\LRp{\bm{v}(\bm{u})} = \bm{u}.
-\label{eq:entropyvarsmap}
-\end{align}
-Several widely used equations in fluid modeling (Burgers, shallow water, compressible Euler and Navier-Stokes equations) yield convex entropy functions $U(\bm{u})$ \cite{hughes1986new, chen2017entropy}.
-Let $\partial \Omega$ be the boundary of $\Omega$ with outward unit normal $\bm{n}$.
-By multiplying the equation (\ref{eq:nonlineqs}) with $\bm{v}(\bm{u})^T$, the solutions $\bm{u}$ of (\ref{eq:nonlineqs}) can be shown to satisfy an entropy inequality
-\begin{equation}
-\int_{\Omega}\pd{U(\bm{u})}{t}\diff{x} + \int_{\partial \Omega} \sum_{j=1}^d \LRp{\bm{v}(\bm{u})^T\bm{f}_j(\bm{u}) - \psi_j\LRp{\bm{v}(\bm{u})}}n_j \diff{x} \leq 0, \qquad \bm{n} = \LRp{n_1,\ldots,n_d},
-\label{eq:entropyineq}
-\end{equation}
-where $\psi_j(\bm{u})$ is a function referred to as the entropy potential.
+\note{Idea: decoupled SBP operators allow for the construction of entropy stable schemes by chaining together different sets of nodes through extrapolation (interpolation) operators.  }
 
-The proof of (\ref{eq:entropyineq}) requires the use of the chain rule \cite{mock1980systems, harten1983symmetric, dafermos2005compensated}.
-The instability-in-practice of high order schemes for (\ref{eq:nonlineqs}) can be attributed in part to the fact that the discrete form of the equations do not satisfy the chain rule, and thus do not satisfy (\ref{eq:entropyineq}).
-As a result, discretizations of (\ref{eq:nonlineqs}) do not typically possess an underlying statement of stability.
-For low order schemes, this can be offset in practice by the inherent numerical dissipation.
-However, because high order discretizations possess low numerical dissipation, the lack of an underlying discrete stability has been conflated with the idea that high order methods are inherently less stable than low order methods.
+\section{Gauss-Legendre collocation on non-conforming meshes}
 
-\section{Approximation spaces}
+On conforming meshes, it is most efficient to utilize both Gauss quadrature for volume integrals and Gauss quadrature for face or surface integrals.
+For solutions represented in terms of their values at tensor product volume Gauss nodes, extrapolation to face Gauss nodes can be done in an efficient line-by-line manner using one-dimensional interpolation matrices.
 
-We define an approximation space using degree $N$ polynomials on the reference element.
-In this work, we consider either simplicial reference elements (triangles and tetrahedra) or tensor product reference elements (quadrilaterals and hexahedra).
-The natural polynomial approximation space differs depending on the element type \cite{chan2015gpu}.
-
-On a $d$-dimensional reference simplex, the natural polynomial space are total degree $N$ polynomials 
-\[
-P^N\LRp{\widehat{D}} = \LRc{\hat{x}_1^{i_1}\ldots\hat{x}_d^{i_d}, \quad \hat{\bm{x}} \in \widehat{D}, \quad 0\leq \sum_{k=1}^d i_k \leq N}.
-\]
-In contrast, the natural polynomial space on a $d$-dimensional tensor product element is 
-\[
-Q^N\LRp{\widehat{D}} = \LRc{\hat{x}_1^{i_1}\ldots\hat{x}_d^{i_d}, \quad \hat{\bm{x}} \in \widehat{D}, \quad 0\leq i_k \leq N, \quad k = 1,\ldots, d}.
-\]
-We take $V^N$ to denote the natural approximation space on a given element.  Furthermore, we denote the dimension of $V^N$  as $N_p = {\rm dim}\LRp{V^N\LRp{\widehat{D}}}$.
-
-We also define trace spaces for each reference element.
-Let $\hat{f}$ be a face of the reference element $\hat{D}$.
-The trace space $V^N \LRp{\hat{f}}$ is defined as the restrictions of functions in $V^N\LRp{\hat{D}}$ to $\hat{f}$
-\[
-V^N \LRp{\hat{f}} = \LRc{ \left.u\right|_{\hat{f}}, \quad u \in V^N\LRp{\hat{D}}}, \qquad \hat{f}\in \partial\hat{D}.
-\]
-For example, the trace space for degree $N$ polynomials on a quadrilateral face $\hat{f}$ of the bi-unit hexahedral element $[-1,1]^3$ is $Q^N\LRp{\hat{f}}$, while the trace space for degree $N$ polynomials on a triangular face $f$ of the tetrahedron is $P^N\LRp{\hat{f}}$.
-We denote the dimension of the trace space as ${\rm dim}\LRp{V^N\LRp{{\hat{f}}}} = N^f_p$.
-
-We similarly define the trace space for the surface $\partial \hat{D}$ of $\hat{D}$
-\[
-V^N \LRp{\partial \hat{D}} = \LRc{ \left.u\right|_{\partial \hat{D}}, \quad u \in V^N\LRp{\hat{D}}}.
-\]
-
-\section{$L^2$ projection and lifting matrices and a decoupled SBP operator}
-
-The schemes in \cite{chan2017discretely} begin on the reference element $\hat{D}$ by approximating the solution in a degree $N$ polynomial basis $\LRc{\phi_j({\bm{x}})}_{i=1}^{N_p}$.
-It is assumed that volume and surface quadrature rules $({\bm{x}}_i, w_i)$, $\LRp{{\bm{x}}^f_i,w^f_i}$ are provided for $\hat{D}$ and the surface $\partial \hat{D}$ of $\hat{D}$, respectively.
-We make the following assumption on the volume quadrature rule:
-\begin{assumption}
-The volume quadrature rule is exact for polynomials of degree $(2N-1)$.
-\label{ass:quad}
-\end{assumption}
-In addition to being related to the positive definiteness of the mass matrix, Assumption~\ref{ass:quad} ensures that, for any $u,v \in V^N$, the following integral is exactly approximated using quadrature
-\[
-\int_{\hat{D}} \pd{u(\hat{\bm{x}})}{\hat{x}_i} v(\hat{\bm{x}})\diff{\hat{\bm{x}}} = \sum_{k=1}^{N_q} w_k \pd{u(\hat{\bm{x}}_k)}{\hat{x}_i} v(\hat{\bm{x}}_k).  
-\]
-In other words, evaluating the integral of the derivative of $u$ and $v$ using quadrature is exact because the integrand is a polynomial of degree $(2N-1)$.
-
-\note{TODO: Standardize hat notation for operators/normals.}
-
-Let $\bm{V}_q,\bm{V}_f$ denote interpolation matrices, and let $\bm{D}^i$ be the differentiation matrix with respect to the $i$th coordinate such that
-\begin{gather}
-\LRp{\bm{V}_q}_{ij} = \phi_j(\bm{x}_i), \qquad \LRp{\bm{V}_f}_{ij} = \phi_j(\bm{x}^f_i), \qquad \pd{\phi_j(\bm{x})}{x_i} = \sum_{k=1}^{N_p} \LRp{\bm{D}^i_{jk}} \phi_k(\bm{x}).
-\end{gather}
-The interpolation matrices $\bm{V}_q,\bm{V}_f$ map basis coefficients to evaluations at volume and surface quadrature points respectively, while the differentiation matrix ${\bm{D}}_i$ maps basis coefficients of a function to the basis coefficients of its derivative with respect to $x_k$.
-The interpolation matrices are used to assemble the mass matrix $\bm{M}$, the quadrature-based projection matrix $\bm{P}_q$, and lifting matrix $\bm{L}_f$
-\begin{gather}
-\bm{M} = \bm{V}_q^T\bm{W}\bm{V}_q, \qquad \bm{P}_q = \bm{M}^{-1}\bm{V}_q^T\bm{W}, \qquad \bm{L}_f = \bm{M}^{-1}\bm{V}_f^T\bm{W}_f,
-\end{gather}
-where $\bm{W}, \bm{W}_f$ are diagonal matrices of volume and surface quadrature weights, respectively.
-The matrix $\bm{P}_q$ is a quadrature-based discretization of the $L^2$ projection operator $\Pi_N$ onto degree $N$ polynomials.
-
-Interpolation, differentiation, and $L^2$ projection matrices can be combined to construct finite difference operators.
-For example, the matrix $\bm{D}_q = \bm{V}_q\bm{D}^i\bm{P}_q$ maps function values at quadrature points to approximate values of the derivative at quadrature points.
-By choosing specific quadrature rules, $\bm{D}_q$ recovers high order summation-by-parts finite difference operators in \cite{gassner2013skew, fernandez2014generalized, ranocha2018generalised} and certain operators in \cite{hicken2016multidimensional}.
-However, to address difficulties in designing efficient entropy stable interface terms for nonlinear conservation laws, the PI introduced in \cite{chan2017discretely} a new ``decoupled'' summation by parts matrix which builds interface terms directly into the approximation of the derivative.
-
-Let $\bm{n}$ denote the outward unit normal $\bm{n}$ on $\hat{D}$, and let $\bm{n}_i$ denote the vector containing values of the $i$th component at all surface quadrature points.
-The ``decoupled'' summation by parts operator $\bm{D}^i_N$ is defined as the block matrix involving both volume and surface quadratures
-\begin{align}
-\bm{D}^i_N  &= \LRs{
-\begin{array}{cc}
-\bm{D}^i_q - \frac{1}{2}\bm{V}_q \bm{L}_f {\rm diag}({\bm{n}}_i) \bm{V}_f\bm{P}_q &  \frac{1}{2}\bm{V}_q\bm{L}_f{\rm diag}({\bm{n}}_i)\\
--\frac{1}{2}{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q & \frac{1}{2}{\rm diag}({\bm{n}}_i)
-\end{array}}.  
-\label{eq:DN}
-\end{align}
-When combined with projection and lifting matrices, $\bm{D}^i_N$ produces a high order polynomial approximation of $f\pd{g}{x_i}$. 
-Let $f, g$ be differentiable functions, and let $\bm{f}_i = f(\bm{x}_i)$, $\bm{g}_i = g(\bm{x}_i)$ denote values of $f,g$ at volume and surface quadrature points.  Then,
-\begin{gather}
-f\pd{g}{x_i} \approx \LRs{\begin{array}{cc}
-\bm{P}_q & \bm{L}_f\end{array}} {\rm diag}\LRp{\bm{f}}\bm{D}^i_N \bm{g}.  
-\label{eq:dnapprox}
-\end{gather}
-The approximation can also be interpreted as solving a variational problem.
-Let $n_i$ be the $i$th component of the unit normal on $\hat{D}$.
-Then (\ref{eq:dnapprox}) is equivalent to finding $u(\bm{x}) \approx f\pd{g}{x_i}$ such that
-\[
-\int_{\hat{D}} u(\bm{x})v(\bm{x})  = \int_{\hat{D}}{g\pd{\Pi_Nf}{x_i}v} + \int_{\partial \hat{D}}{(f-\Pi_Nf)\frac{\LRp{gv + \Pi_N(gv)}}{2}}n_i, \qquad \forall v\in V^N\LRp{\hat{D}}.
-\]
-If the surface quadrature rule is exact for polynomials of degree $2N$, then the decoupled SBP operator $\bm{D}^i_N$ also satisfies a summation by parts property.  
-\begin{gather}
-\bm{Q}^i_N = \bm{W}_N \bm{D}^i_N, \qquad \bm{W}_N = \LRp{\begin{array}{cc}\bm{W}&\\ & \bm{W}_f\end{array}},\label{eq:dsbp}\\ \bm{Q}^i_N+\LRp{\bm{Q}^i_N}^T = \LRp{\begin{array}{cc}\bm{0}&\\ & \bm{W}_f {\rm diag}(\bm{n}_i)\end{array}} = \bm{B}_N\nonumber
-\end{gather}
-By noting that $\bm{W}\bm{V}_q \bm{L}_q = \bm{P}_q^T\bm{V}_f^T\bm{W}_f$, the entries of $\bm{Q}^i_N$ can be re-written as
-\begin{align}
-\bm{Q}^i_N  &= \LRs{
-\begin{array}{cc}
-\bm{Q}_i - \frac{1}{2}\LRp{\bm{V}_f \bm{P}_q}^T  \bm{W}_f{\rm diag}({\bm{n}}_i) \bm{V}_f\bm{P}_q &  \frac{1}{2}\LRp{\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q}^T\\
--\frac{1}{2}\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q & \frac{1}{2}\bm{W}_f{\rm diag}({\bm{n}}_i)
-\end{array}}, \qquad \bm{Q}_i = \bm{W}\bm{V}_q\bm{D}^i\bm{P}_q,
-\label{eq:QN}
-\end{align}
-from which the SBP property (\ref{eq:dsbp}) is more apparent.
-
-%\begin{lemma}[Lemma 1 in \cite{chan2017discretely}]
-%Let $\bm{D}^i_q = \bm{V}_q\bm{D}^i\bm{P}_q$.  Then, for volume quadratures which are exact for degree $2N-1$ polynomials, $\bm{D}^i_q$ satisfies the SBP property with respect to $\bm{W}$ 
-%\[
-%\bm{W}\bm{D}^i_q = \LRp{\bm{V}_f\bm{P}_q}^T \bm{W}_f\diag{\hat{\bm{n}}} \bm{V}_f\bm{P}_q - \LRp{\bm{W}\bm{D}^i_q}^T.  
-%\]
-%\label{lemma:gsbp}
-%\end{lemma}
-
-\section{Skew-symmetric entropy conservative formulations}
-
-\subsection{Loss of the SBP property for hybrid and non-conforming meshes}
-While the SBP property has been used to derive entropy stable schemes, the SBP property is difficult or impossible to enforce in certain discretization settings, such as hybrid and non-conforming meshes arises.
-This difficulty is a result of the choices of surface quadrature which naturally arise for these settings.
+For non-conforming meshes, it can be advantageous to use composite Gauss quadratures on non-conforming interfaces.
+\note{Reference JK and LCW's paper on full-side vs split side mortars.}  However, interpolating the solution at volume Gauss nodes to split-side Gauss nodes is no longer a one-dimensional operation.
 
@@
 \begingroup
-\captionsetup[subfigure]{width=.43\textwidth}
-\subfloat[Insufficiently accurate surface quadrature on the triangle element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D.png}\label{subfig:hybrid1}}
+\captionsetup{width=.25\linewidth}
+\subfloat[Conforming faces]{\includegraphics[width=.25\textwidth]{figs/gauss1.png}}
+\hspace{4em}
+\subfloat[Non-conforming faces]{\includegraphics[width=.25\textwidth]{figs/gauss2.png}}
 \endgroup
-\hspace{2em}
-\subfloat[Incompatible surface quadrature on the quadrilateral element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D_GQ.png}\label{subfig:hybrid2}}
-\caption{Examples of interface couplings which do not result in an SBP property and are not entropy stable.
-On the left, the surface quadrature is a $(N+1)$ point GLL rule, and results in a loss of the SBP property on the triangle.
-On the right, the surface quadrature is a $(N+1)$ point Gauss-Legendre rule, and results in a loss of the SBP property on the quadrilateral.
-}
-\label{fig:hybrid}
+\caption{Examples of surface quadratures for conforming and non-conforming interfaces for Gauss collocation methods.  }
+\label{fig:gqcon_noncon}
 \end{figure}
 
-\begin{figure}
-\centering
-\subfloat[Hex-pyramid coupling]{\includegraphics[width=.35\textwidth]{figs/hybrid3D.png}}
-\hspace{2em}
-\subfloat[Hex-prism coupling]{\includegraphics[width=.375\textwidth]{figs/hybrid3D_wedge.png}}
-\caption{Illustration of a 3D coupling between a GLL hexahedral element and a pyramid.  The SBP property does not hold on the pyramid due to the use of GLL quadrature on the quadrilateral face. }
-\label{fig:hybrid3d}
-\end{figure}
+It is possible to efficiently interpolate from volume Gauss nodes to split-side Gauss nodes in two steps.
+We first interpolate to full-side Gauss nodes on the face, then interpolate from full-side  to split-side Gauss nodes.
+Each of these two operations can be done using only one-dimensional interpolation matrices, implying that the interpolation matrix from volume to split-side Gauss nodes can be decomposed into
+\[
+\bm{V}_{\rm vol-to-split} = \bm{V}_{\rm vol-to-full} \bm{V}_{\rm full-to-split}
+\]
+where each of the matrices in the decomposition possess either a sparse or Kronecker product structure.
 
-\begin{figure}[!h]
-\centering
-\begingroup
-\captionsetup[subfigure]{width=.425\textwidth}
-\subfloat[Coupling of nodes for a non-conforming interface (entropy stable but tighter coupling).]{\raisebox{0em}{\includegraphics[width=.425\textwidth]{figs/noncon.png}}\label{subfig:noncon2}}
-\hspace{2em}
-\subfloat[Surface quadrature rule induced by non-conforming refinement (no SBP property).]{\raisebox{-.3em}{\includegraphics[width=.45\textwidth]{figs/nonconQuad.png}}\label{subfig:noncon1}}
-\endgroup
-%\subfloat[3D hexahedra-pyramid coupling]{\includegraphics[width=.31\textwidth]{figs/hybrid3D.png}}
-\caption{Illustration of two different entropy stable inter-element couplings for $h$ non-conforming meshes.
-Each dashed red line indicates a dyadic flux computation required between two nodes.
-The coupling terms in \cite{friedrich2017entropy} requires computing dyadic fluxes between \emph{each} pair of nodes on adjacent interfaces, while the proposed approach requires dyadic flux computations only between immediately adjacent nodes.
-}
-\label{fig:noncon}
-\end{figure}
+ESDG schemes based on decoupled SBP operators introduced in \cite{chan2017discretely} assume a single interpolation matrix which extrapolates from volume to surface nodes.
+These schemes require dyadic flux evaluations between all volume and surface nodes connected by this interpolation matrix.
+For non-conforming interfaces with split-side Gauss nodes on the face, this necessitates dyadic flux computations between all volume nodes and all split-side Gauss nodes.
 
-\subsection{A variational SBP property}
+We show in this section how to construct an entropy stable Gauss collocation scheme based on two layers of surface nodes: full-side and split-side nodes.
+These surface nodes will be loosely through a skew-symmetric coupling term involving polynomial interpolation matrices between full-side and split-side nodes.
+This scheme requires dyadic flux computations between volume and full-side surface nodes, and between full-side and split-side surface nodes.
+Each set of dyadic flux computations can be reduced to
 
-It can be shown that the SBP property holds for $\bm{D}^i_N$ if the surface quadrature is exact for polynomials of degree $2N$ \cite{chan2017discretely}.
-The SBP property is lost if the surface quadrature is only exact for polynomial integrands of degree less than $2N$.
-However, one can still show that a variational version of the SBP property is satisfied.
+Let ${\bm{V}_f}\tilde{P}_f$ denote the matrix which projects to the trace space using split-side quadrature and evaluates at full-side quadrature points.
+Similarly, let $\tilde{\bm{V}_f}{P}_f$ denote the matrix which projects to the trace space using full-side quadrature and evaluates at split-side quadrature points
+\begin{align*}
+\bm{C}^i_f &= \begin{pmatrix}
+\bm{0} & {\rm diag}\LRp{\hat{\bm{n}}_i}{\bm{V}_f}\tilde{\bm{P}}_f\\
+-\tilde{\bm{V}_f}{\bm{P}}_f{\rm diag}\LRp{\hat{\bm{n}}_i} &\bm{0}
+\end{pmatrix}, \\
+\bm{S}^i_f &= \begin{pmatrix}
+\diag{\bm{w}_f} & \\
+& \diag{\tilde{\bm{w}}_f}
+\end{pmatrix}\bm{C}^i_f.
+\end{align*}
+It is straightforward to show that the matrix $\bm{S}_f$ is skew-symmetric.
 
-\begin{lemma}
-Let $v\in V^N$, and let the surface trace $\LRl{v}_{\partial \hat{D}} \in V^M\LRp{\partial \hat{D}}$ for some $M \leq N$.
-If the surface quadrature is exact for polynomials of degree $N+M$,  then for an arbitrary vector $\bm{u}_q$,
+%{fig:gqcon_noncon}
+Let $\tilde{\bm{F}}_S$ denote the flux matrix, then 
 \[
-\bm{v}_q^T\bm{Q}_i \bm{u}_q = \bm{v}_q^T\LRp{ \LRp{\bm{V}_f\bm{P}_q}^T \bm{W}_f \diag{{\bm{n}}_i}\bm{V}_f\bm{P}_q - \bm{Q}_i^T}\bm{u}_q
+\begin{bmatrix}\bm{L}_f & \tilde{\bm{L}}_f\end{bmatrix} \LRp{\bm{C}^i_f\circ\tilde{\bm{F}}_S}\bm{1}
 \]
-where $\bm{v}$ denotes the vector of values of $v$ of at both volume and surface quadrature points. 
-\label{lemma:vsbp}
-\end{lemma}
-\begin{proof}
-Recall that $\bm{Q}_i = \bm{W} \bm{V}_q \bm{D}^i\bm{P}_q$.  Since the volume quadrature is exact for degree $2N-1$ polynomials, we have that
-\begin{align*}
-\bm{v}_q^T\bm{Q}_i \bm{u}_q &= \bm{v}_q^T\bm{W} \bm{D}^i \bm{P}_q \bm{u}_q = \int_{\hat{D}} \pd{\Pi_N u}{\hat{x}_i} v = \int_{\partial \hat{D}} (\Pi_N u) v \hat{n}_i - \int_{\hat{D}} \LRp{\Pi_N u} \pd{v}{\hat{x}_i}.
-\end{align*}
-Since the volume integrand $\LRp{\Pi_N u} \pd{v}{\hat{x}_i} \in V^{2N-1}$, it is exactly integrated using the volume quadrature.
-Additionally, since $\Pi_N u \in V^N$ and $v\in V^M$ on the surface of $\hat{D}$, the surface quadrature is exact for the surface integral and
-\begin{align*}
-\int_{\partial \hat{D}} (\Pi_N u) v \hat{n}_i -   \int_{\hat{D}} \LRp{\Pi_N u} \pd{v}{\hat{x}_i} &= \bm{v}_f^T\bm{W}_f{\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q\bm{u}_q - \LRp{\bm{V}_q\bm{D}^i\bm{P}_q\bm{v}_q}^T\bm{W} \bm{V}_q\bm{P}_q\bm{u}_q\\
-&= \bm{v}_q^T \LRp{\bm{V}_f\bm{P}_q}^T\bm{W}_f{\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q\bm{u}_q - \bm{v}_q^T \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T \bm{W} \bm{V}_q\bm{P}_q\bm{u}_q,
-\end{align*}
-where we have used that, since $v\in V^N$, $\bm{v}_f = \bm{V}_f\bm{P}_q\bm{v}_q$.  The proof is completed by noting that 
-\[
-\bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T \bm{W} \bm{V}_q\bm{P}_q = \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{M}\bm{P}_q = \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T\bm{W} = \bm{Q}_i^T.
-\]
-\end{proof}
+is a skew-symmetric correction term which ensures entropy stability and high order accuracy (through consistency up to quadrature order).
 
-The variational SBP property also extends to the decoupled SBP operator.  This property, along with the exact differentiation of constants, is necessary for the proof of entropy stability.  
-\begin{lemma}
-Let $\bm{D}^i_N$ be a decoupled SBP operator on the reference element $\hat{D}$, and let the surface quadrature be exact for polynomials of degree $N+M$ for some $M \leq N$.
-Suppose $v\in V^N$ and that $\LRl{v}_{\partial \hat{D}} \in P^M\LRp{\partial \hat{D}}$, Then,
-\[
-\bm{v}^T\bm{Q}^i_N\bm{u} = \bm{v}^T\LRp{\bm{B}^i_N - \LRp{\bm{Q}^i_N}^T}\bm{u}.%, \qquad \bm{Q}^i_N \bm{1} = \bm{0}.
-\]
-where $\bm{v}$ denotes the values of $v$ of at both volume and surface quadrature points.  
-\label{lemma:vdsbp}
-\end{lemma}
-\begin{proof}
-For convenience, let $\bm{u}_q, \bm{u}_f$ denote evaluations of $u$ at volume and surface points, such that 
-\[
-\bm{u} = \begin{pmatrix} \bm{u}_q\\ \bm{u}_f\end{pmatrix}, \qquad \bm{v} = \begin{pmatrix} \bm{v}_q\\ \bm{v}_f\end{pmatrix}.  
-\]
-The proof of the variational summation by parts property uses the definition of $\bm{Q}^i_N$ (\ref{eq:QN}), 
-\begin{align*}
-\bm{v}^T\bm{Q}^i_N\bm{u} &= \bm{v}_q^T\bm{Q}_i \bm{u}_q - \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}} \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}} \bm{u}_f\\
-& - \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}} \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} + \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}} \bm{u}_f.
-\end{align*}
-Applying Lemma~\ref{lemma:vsbp} then yields
-\begin{align*}
-\bm{v}^T\bm{Q}^i_N\bm{u} &= -\bm{v}_q^T\bm{Q}^T_i \bm{u}_q + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}} \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}} \bm{u}_f\\
-&\qquad - \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}} \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} - \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}} \bm{u}_f + \bm{v}_f^T\bm{W}_f\hat{\bm{n}} \bm{u}_f\\
-&= \begin{pmatrix} \bm{v}_q\\ \bm{v}_f\end{pmatrix}^T 
-\LRp{\begin{pmatrix}
-\bm{0}& \\
-& \bm{W}_f\diag{\hat{\bm{n}}}
-\end{pmatrix}
-+
-\begin{pmatrix}
--\bm{Q}_i^T + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q}^T\bm{W}_f\hat{\bm{n}} \bm{V}_f\bm{P}_q & -\frac{1}{2} \LRp{\bm{W}_f\hat{\bm{n}} \bm{V}_f\bm{P}_q}^T\\
-\frac{1}{2}\bm{W}_f\hat{\bm{n}} \bm{V}_f\bm{P}_q & -\frac{1}{2}\bm{W}_f\hat{\bm{n}}
-\end{pmatrix} } 
-\begin{pmatrix} \bm{u}_q\\ \bm{u}_f\end{pmatrix}\\
-&= \bm{v}^T\LRp{\bm{B}_N - \LRp{\bm{Q}^i_N}^T}\bm{u}.
-\end{align*}
-\end{proof}
-Using Lemma~\ref{lemma:vdsbp}, we will determine the accuracy of the surface quadrature necessary to ensure the proof of semi-discrete entropy stability is valid.
-This minimal degree of surface quadrature accuracy will depend on the nature of the reference-to-physical mapping (e.g.\ affine vs curved elements).
+Since the difference is high order accurate, you can remove this correction term and still get high order accuracy.
+However, the spatial entropy RHS is no longer zero if you use inconsistent surface and coupling quadrature without the correction term (\note{prove this}).
 
-\begin{corollary}
-\label{lemma:sbpcor}
-Let the surface quadrature be exact for polynomials of degree $N$.  Then, 
-\[
-\bm{1}^T\bm{Q}^i_N\bm{u} = \bm{1}^T\bm{B}_N\bm{u}, \qquad \bm{Q}^i_N\bm{1} = \bm{0},
-\]
-where $\bm{u}$ is an arbitrary vector of appropriate size.  
-\end{corollary}
-\begin{proof}
-The proof that $\bm{Q}^i_N \bm{1} = \bm{0}$ follows from the property that polynomials are equal to their $L^2$ projection, and is identical to that of \cite{chan2017discretely,chan2018discretely}.
-The proof of the first equality follows $\bm{Q}^i_N \bm{1} = \bm{0}$ and Lemma~\ref{lemma:vdsbp} with $M=0$
-\[
-\bm{1}^T\LRp{\bm{Q}^i_N}\bm{u} = \bm{1}^T\LRp{\bm{B}_N - \LRp{\bm{Q}^i_N}^T}\bm{u} = \bm{1}^T{\bm{B}_N}\bm{u}.
-\]
-\end{proof}
-\subsection{Entropy stability on affine meshes}
-
-The high order methods in this work ensure that (\ref{eq:entropyineq}) is satisfied discretely by avoiding the use of the chain rule in the proof of entropy dissipation.
-These ``entropy stable'' schemes rely on two main ingredients: an entropy stable numerical flux as defined by Tadmor \cite{tadmor1987numerical} and a concept referred to as ``flux differencing''.
-% entropy stable schemes in one-dimension.
-Let $\bm{f}_S\LRp{\bm{u}_L,\bm{u}_R}$ be a numerical flux function which is a function of ``left'' and ``right'' states $\bm{u}_L,\bm{u}_R$.  
-The numerical flux $\bm{f}_S$ is \textit{entropy conservative} if it satisfies the following three conditions:  
-%The numerical flux $\bm{f}_S$ is an \textit{entropy stable} flux function if it satisfies the following three conditions:  
-\begin{gather}
-\bm{f}_S(\bm{u},\bm{u}) = \bm{f}(\bm{u}), \qquad \text{(consistency)}\\
-\bm{f}_S(\bm{u}_L,\bm{u}_R) = \bm{f}_S(\bm{u}_R,\bm{u}_R), \qquad \text{(symmetry)}\nonumber\\
-\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}_S(\bm{u}_L,\bm{u}_R) = \psi(\bm{u}_L) - \psi(\bm{u}_R), \qquad \text{(conservation)}\nonumber
-%\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}_S(\bm{u}_L,\bm{u}_R) \leq \psi(\bm{u}_L) - \psi(\bm{u}_R), \qquad \text{(entropy dissipation)}\nonumber
-\label{eq:esflux}
-\end{gather}
-If instead of the third condition, $\bm{f}_S$ satisfies a \emph{stability} property $\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}_S(\bm{u}_L,\bm{u}_R) \leq \psi(\bm{u}_L) - \psi(\bm{u}_R)$, the flux $\bm{f}_S$ is referred to as \textit{entropy stable}.
-
-We begin by deriving a skew-symmetric formulation on the reference element $\hat{D}$, and showing it is entropy conservative.
-This formulation can then be extended in a straightforward manner to affine elements.
-The skew-symmetric formulation for (\ref{eq:nonlineqs}) on $\hat{D}$ is
-\begin{gather}
-\pd{\bm{u}}{t} + \sum_{i=1}^d \LRs{\begin{array}{cc}
-\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{\hat{\bm{n}}}\bm{f}_i^* = 0,\\
-\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q,
-\label{eq:esdgSkew}
-\end{gather}
-where $\bm{f}^*$ is some numerical flux.  Multiplying the formulation (\ref{eq:esdgSkew}) by $\bm{M}$ on both sides yields a weak form 
-\begin{gather}
-\bm{M}\pd{\bm{u}}{t} + \sum_{i=1}^d \LRs{\begin{array}{c}
-\bm{V}_q \\ \bm{V}_f\end{array}}^T \begin{pmatrix}
-\bm{W}&\\
-&\bm{W}_f
-\end{pmatrix}
-\LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} +  \bm{V}_f^T\bm{W}_f\diag{\hat{\bm{n}}}\bm{f}_i^*\label{eq:esdgSkewWeak}\\
-= \bm{M}\pd{\bm{u}}{t} + \sum_{i=1}^d\LRs{\begin{array}{c}
-\bm{V}_q \\ \bm{V}_f\end{array}}^T 
-\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{V}_f^T\bm{W}_f \diag{\hat{\bm{n}}}\bm{f}_i^* = 0.  \nonumber
-\end{gather}
-
-We can now show that the skew-symmetric formulation is semi-discretely entropy conservative.  
-\begin{theorem}
-The formulation (\ref{eq:esdgSkew}) is entropy conservative for any surface quadrature which is exact for polynomials of degree $N$, such that
-\[
-\bm{1}^T\bm{W}\pd{U(\bm{u}_q)}{t} + \sum_{i=1}^d\bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}} \LRp{\psi_i(\tilde{\bm{u}}_f) - \tilde{\bm{v}}_f^T\bm{f}_i^*} = 0, \qquad \bm{u}_q = \bm{V}_q\bm{u}.
-\]
-\label{thm:esdg}
-\end{theorem}
-\begin{proof}
-Testing (\ref{eq:esdgSkewWeak}) by $\bm{v}_h = \bm{P}_q\bm{v}_q$ yields 
-\begin{align}
-\bm{v}_q^T\bm{W}\pd{\LRp{\bm{V}_q\bm{u}}}{t} + \sum_{i=1}^d
-\tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \tilde{\bm{v}}_f^T \bm{W}_f \diag{\hat{\bm{n}}}\bm{f}_i^* = 0.
-\end{align}
-One can show that \cite{chan2017discretely}
-\begin{align*}
-\tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} &= \tilde{\bm{v}}^T \LRp{\bm{Q}^i_N \circ \bm{F}_S}\bm{1} - \tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N}^T \circ \bm{F}_S}\bm{1}\\
-&= \tilde{\bm{v}}^T \LRp{\bm{Q}^i_N \circ \bm{F}_S}\bm{1} - \bm{1}^T \LRp{{\bm{Q}^i_N} \circ \bm{F}_S}\tilde{\bm{v}}.
-\end{align*}
-Where we have used that $\bm{F}_S$ is symmetric and that the Hadamard product commutes.  Applying the conservation condition on $\bm{f}_S$ in (\ref{eq:esflux}) then yields
-\begin{align*}
-\tilde{\bm{v}}^T \LRp{\bm{Q}^i_N \circ \bm{F}_S}\bm{1} - \bm{1}^T \LRp{{\bm{Q}^i_N} \circ \bm{F}_S}\tilde{\bm{v}} &= \sum_{jk} \LRp{\bm{Q}^i_N}_{jk} \LRp{\tilde{\bm{v}}_j-\tilde{\bm{v}}_k} \bm{f}_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k} \\
-&= \sum_{ij} \LRp{\bm{Q}^i_N}_{jk} \LRp{\psi_i(\tilde{\bm{u}}_j) - \psi_i(\tilde{\bm{u}}_k)}\\
-&= \bm{1}^T\LRp{\bm{Q}^i_N}\psi_i(\tilde{\bm{u}}) - \psi_i(\tilde{\bm{u}})^T\LRp{\bm{Q}^i_N}\bm{1} \\
-&= \bm{1}^T\LRp{\bm{Q}^i_N}\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i_N\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}} \psi_i(\tilde{\bm{u}}_f)
-\end{align*}
-where we have used Corollary~\ref{lemma:sbpcor} in the second to last equality.
-%The final step of the proof is  where we have used Lemma~\ref{lemma:vdsbp} to integrate by parts and conclude that $\bm{1}^T\LRp{\bm{Q}^i_N}^T = 0$.
-\end{proof}
-
-\begin{remark}
-The only result necessary to prove Theorem~\ref{thm:esdg} is Corollary~\ref{lemma:sbpcor}.
-Thus, when generalizing to curved elements, the proof of entropy stability for the skew-symmetric form will depend only on the curved version of Corollary~\ref{lemma:sbpcor}.
-\end{remark}
-
-The skew symmetric formulation can also be shown to be locally conservative in the sense of \cite{shi2017local}, which is sufficient to show the numerical solution convergences to the weak solution under mesh refinement.
-\begin{theorem}
-The formulation (\ref{eq:esdgSkew}) is locally conservative such that
-\begin{align}
-\bm{1}^T\bm{W}\pd{\LRp{\bm{V}_q\bm{u}}}{t} + \sum_{i=1}^d\bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}_i}\bm{f}_i^* = 0. 
-\end{align}
-\end{theorem}
-\begin{proof}
-To show local conservation, we test (\ref{eq:esdgSkewWeak}) with $1$.  Let $\bm{e}$ be the coefficients of $1$ under the basis $\phi_j(\bm{x})$.
-\begin{align}
-\bm{1}^T\bm{W}\bm{V}_q\pd{\bm{u}}{t} + \sum_{i=1}^d
-\bm{1}^T
-\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}}\bm{f}_i^* = 0. 
-\end{align}
-Because $\bm{F}_S$ is symmetric and $\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}$ is skew-symmetric, the term $\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}$ is a skew-symmetric matrix.
-The volume term vanishes
-\[
-\bm{1}^T\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} = 0,
-\]
-where we have used that, for any skew symmetric matrix $\bm{A}$, $\bm{x}^T\bm{A}\bm{x} = 0$.
-% and that the Hadamard product is commutable $\LRp{\bm{A}\circ\bm{B} }^T = \bm{A}^T\circ\bm{B}^T$
-%\begin{align}
-%\bm{1}^T\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} &= \bm{1}^T\LRp{\bm{Q}^i_N\circ \bm{F}_S}\bm{1} - \bm{1}^T\LRp{\LRp{\bm{Q}^i_N}^T\circ \bm{F}_S}\bm{1}\\
-%&= \bm{1}^T\LRp{\bm{Q}^i_N\circ \bm{F}_S}\bm{1} - 
-%\bm{1}^T\LRp{\bm{Q}^i_N\circ \bm{F}_S}\bm{1} = 0,
-%\end{align}
-%where we have used that $\bm{F}_S$ is symmetric.
-%\begin{align}
-%\end{align}
-\end{proof}
-
-Derivatives with respect to physical coordinates on $D^k$ are computed in terms of a change of variables formula and derivatives with respect to reference coordinates 
-\[
-\pd{u}{x_i} = \sum_{ij} \bm{G}^k_{ij}\pd{u}{\hat{x}_j}, \qquad \bm{G}^k_{ij} = J^k\pd{\hat{x}_j}{{x}_i}, 
-\]
-where $J^k$ is the determinant of the Jacobian of the geometric mapping on the element $D^k$.
-For affine meshes, the scaled geometric terms $\bm{G}^k_{ij}$ are constant over each element.
-It was shown in \cite{chan2017discretely} that one can define a physical differentiation matrix as a linear combination of reference differentiation matrices.
-Let $\hat{\bm{D}}^i_N$ denote the decoupled SBP operator on the reference element $\hat{D}$, and let the decoupled SBP operators with respect to the physical coordinates on $D^k$ be defined as
-\begin{align}
-{\bm{D}}^i_N = \sum_{j=1}^d \bm{G}_{ij}\hat{\bm{D}}^j_N.
-\end{align}
-%Then, $\bm{Q}^i_N = \bm{W} \bm{D}^i_N$ satisfies the SBP property on each element $D^k$
-%\begin{equation}
-%\bm{Q}^i_N + \LRp{\bm{Q}^i_N}^T = \bm{B}^i_N, \qquad \bm{B}^i_N = \begin{pmatrix}
-%\bm{0}&\\
-%& {J^k_f}\bm{W}_f \diag{{\bm{n}_i}}
-%\end{pmatrix}
-%\label{eq:physsbp}
-%\end{equation}
-%where $J^k_f$ are the Jacobian factors which result from mapping the faces of $D^k$ to reference faces.
-%\end{lemma}
-Then, an entropy conservative formulation can be given on $D^k$ as follows:
-\begin{gather}
-\pd{\bm{u}}{t} + \sum_{i=1}^d \LRs{\begin{array}{cc}
-\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{\hat{\bm{n}}}\bm{f}_i^* = 0, \label{eq:skewform}\\
-\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q, \nonumber\\
-\bm{f}^* = \bm{f}_S(\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f), \qquad \text{ on interior interfaces,} \nonumber
-\end{gather}
-where $\tilde{\bm{u}}_f^+$ denotes the face value of the entropy-projected conservative variables $\tilde{\bm{u}}_f$ on the neighboring element.
-The formulation can be made entropy stable by adding an appropriate penalization term, such as a Lax-Friedrichs or matrix-based dissipation \cite{winters2017uniquely, chen2017entropy, chan2017discretely}.
-
-\subsection{Curvilinear meshes}
-
-On affine meshes, it is possible to show entropy stability of the skew-symmetric formulation (\ref{eq:esdgSkew}) under a surface quadrature which is only exact for degree $N$ polynomials.
-However, on curved meshes, a stronger surface quadrature rule is required to demonstrate entropy stability, where the strength of the surface rule depends on the order and approximation of geometric terms.
-
-Let $\bm{J}^k\bm{G}^k_{ij}$ denote the vector of scaled geometric terms ${J}^k\bm{G}^k_{ij}$ evaluated at volume and surface quadrature points.
-Decoupled SBP operators on a curved element $D^k$  can be defined as in \cite{chan2018discretely} by
-\begin{equation}
-\bm{D}^i_N = \sum_{j=1}^d \diag{\bm{G}^k_{ij}}\hat{\bm{D}}^j_N + \hat{\bm{D}}^j_N\diag{\bm{G}^k_{ij}}.
-\label{eq:dncurved}
-\end{equation}
-The skew-symmetric formulation (\ref{eq:skewform}) can be extended to curvilinear meshes using the definition of $\bm{D}^i_N$ in (\ref{eq:dncurved}).
-However, two additional assumptions must be satisfied in order to prove that the resulting formulation is entropy stable.
-
-\subsection{Discrete geometric conservation law and surface quadrature accuracy}
-
-\note{Add description of how normals factor into this and note that they're computed based on $\bm{G}_{ij}$.}
-
-The first assumption which must be satisfied is the discrete geometric conservation law (GCL) \cite{thomas1979geometric, kopriva2006metric}.
-For curved elements, Lemma~\ref{lemma:vdsbp} and Corollary~\ref{lemma:sbpcor} do not necessarily hold at the discrete level.
-For example, expanding out the condition $\bm{Q}^i_N\bm{1} = \bm{0}$ in terms of (\ref{eq:dncurved}) yields
-\begin{align}
-\bm{D}^i_N \bm{1} = \sum_{j=1}^d \diag{\bm{G}^k_{ij}}\hat{\bm{D}}^j_N \bm{1} + \hat{\bm{D}}^j_N\diag{\bm{G}^k_{ij}}\bm{1} = \sum_{j=1}^d \hat{\bm{D}}^j_N\LRp{\bm{G}^k_{ij}} = 0,
-\label{eq:dgcl}
-\end{align}
-where we have used that $\hat{\bm{D}}^j_N \bm{1} = 0$.
-For degree $N$ isoparametric mappings, the GCL is automatically satisfied in two dimensions due to the fact that the exact geometric terms $\bm{G}^k_{ij}\in P^{N-1}$ \cite{kopriva2006metric}.
-However, in three dimensions, the GCL is not automatically preserved due to the fact that $\bm{G}^k_{ij}\in P^{2N-2}$ and cannot be represented exactly using degree $N$ polynomials.
-Thus, (\ref{eq:dgcl}) must be enforced through an alternative construction of $\bm{G}^k_{ij}$.
-
-The most common approach is to rewrite the geometric terms as the curl of some quantity $\bm{r}^i$, but to interpolate $\bm{r}^i$ before applying the curl \cite{visbal2002use, kopriva2006metric, hindenlang2012explicit}:
-\begin{align}
-\bm{r}^i = { \pd{\bm{x}}{\hat{x}_i}\times \bm{x}}, \qquad
-\LRs{\begin{array}{c}
-\bm{G}^k_{i1}\\
-\bm{G}^k_{i2}\\
-\bm{G}^k_{i3}\end{array}} = -\frac{1}{2}\LRp{\pd{I_{R}\bm{r}^k}{\hat{x}_j}-\pd{I_{R}\bm{r}^j}{\hat{x}_k}}, 
-\label{eq:iconscurl}
-\end{align}
-where $I_{R}$ denotes a degree ${R}$ polynomial interpolation operator with appropriate interpolation nodes.\footnote{This interpolation step must be performed using interpolation points with an appropriate number of nodes on each boundary \cite{chan2018discretely}.
-These include, for example, GLL nodes on tensor product elements, as well as Warp and Blend nodes on non-tensor product elements \cite{warburton2006explicit, chan2015comparison}.}  Because the geometric terms are computed by applying the curl, both $\bm{G}^k_{ij}$ and $n_iJ^f$ are approximated using degree $(R-1)$ polynomials.
-
-Because $\bm{D}^i_N$ are now defined through (\ref{eq:dncurved}), Corollary~\ref{lemma:sbpcor} and the proof of entropy stability may not hold for curved elements and must be modified.
-The introduction of curvilinear meshes will impose slightly different conditions on the accuracy of the surface quadrature, which are summarized in the following curved version of Corollary~\ref{lemma:sbpcor}.
-\begin{lemma}
-Assume that the geometric terms $\bm{G}_{ij} \in V^N$ satisfy the discrete GCL (\ref{eq:dgcl}), that $\LRl{\bm{G}_{ij}}_{\partial \hat{D}}\hat{n}_j \in V^R\LRp{\partial \hat{D}}$, and that the surface quadrature is exact for polynomials of degree $N+R$.
-Then,
-\[
-\bm{1}^T\bm{Q}^i_N\bm{u} = \bm{1}^T\bm{B}^i_N \bm{u}, %\qquad \bm{B}^i_N = \begin{pmatrix}
-%\bm{0}& \\
-%& \bm{W}_f\diag{\bm{n}_i}
-%\end{pmatrix}, 
-\qquad \bm{Q}^i_N \bm{1} = 0, 
-\]
-where 
-\[
-\bm{B}^i_N =  \begin{pmatrix}
-\bm{0}&\\
-& {J^k_f}\bm{W}_f \diag{{\bm{n}_i}}
-\end{pmatrix}.
-\]
-\label{lemma:vdsbpcurved}
-\end{lemma}
-\begin{proof}
-The second equality $\bm{Q}^i_N \bm{1} = 0$ is a consequence of the discrete GCL (\ref{eq:dgcl}), and the proof is identical to that of \cite{chan2018discretely}.
-For the first equality, if the surface quadrature is exact for polynomials of degree $2N$ on the trace space, then the stronger matrix SBP property $\bm{Q}^i_N = \bm{B}^i_N - \LRp{\bm{Q}^i_N}^T$ holds \cite{chan2018discretely}.
-Combining this with $\bm{Q}^i_N\bm{1} = 0$ yields the desired result.
-We focus on the proof of the variational SBP property for surface quadrature which are exact for polynomials of degree $N+R < 2N$.
-
-Expanding $\bm{1}^T\bm{Q}^i_N\bm{u}$ yields
-\[
-\bm{1}^T\bm{Q}^i_N\bm{u} = \frac{1}{2}\sum_{j=1}^d \bm{1}^T \diag{\bm{G}_{ij}} \hat{\bm{Q}}^j_N \bm{u} + \bm{1}^T\hat{\bm{Q}}^j_N \diag{\bm{G}_{ij}} \bm{u}.
-%=  \frac{1}{2}\sum_{j=1}^d \bm{1}^T \diag{\bm{G}_{ij}} \hat{\bm{Q}}^j_N \bm{u},
-\]
-For the latter term in the sum, Lemma~\ref{lemma:vdsbp} holds and
-\begin{equation}
-\sum_{j=1}^d\bm{1}^T\hat{\bm{Q}}^j_N \diag{\bm{G}_{ij}} \bm{u} = 
-\sum_{j=1}^d\bm{1}^T\LRp{\hat{\bm{B}}^j_N - \LRp{\hat{\bm{Q}}^j_N}^T} \diag{\bm{G}_{ij}} \bm{u} = \sum_{j=1}^d\bm{1}^T\hat{\bm{B}}^j_N\diag{\bm{G}_{ij}}\bm{u}, 
-\label{eq:vsbp1}
-\end{equation}
-where we have used that $\hat{\bm{Q}}^j_N \bm{1} = \bm{0}$.  For the former term in the sum, we expand out 
-\begin{gather}
-\sum_{j=1}^d\bm{1}^T \diag{\bm{G}_{ij}} \hat{\bm{Q}}^j_N \bm{u} = 
-\sum_{j=1}^d{\bm{G}_{ij}}^T \hat{\bm{Q}}^j_N \bm{u} \\
-= \sum_{j=1}^d\begin{bmatrix}
-\bm{G}^q_{ij}\\
-\bm{G}^f_{ij}
-\end{bmatrix}^T \begin{bmatrix} 
-\hat{\bm{Q}}_i - \frac{1}{2}\LRp{\bm{V}_f \bm{P}_q}^T  \bm{W}_f{\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q &  \frac{1}{2}\LRp{\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q}^T\label{eq:ibpg}\\
--\frac{1}{2}\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q & \frac{1}{2}\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)
-\end{bmatrix} \begin{bmatrix}\bm{u}_q\\
-\bm{u}_f\end{bmatrix}.\nonumber
-\end{gather}
-where $\bm{G}^q_{ij},\bm{G}^f_{ij}$ denote the values of $\bm{G}_{ij}$ at volume and surface quadrature points.
-
-Because $\bm{G}_{ij} \in V^N$, it is equal to its own $L^2$ projection, and the values at volume and surface quadrature points are related by $\bm{G}^f_{ij} = \bm{V}_f\bm{P}_q\bm{G}^q_{ij}$.
-We can use this to simplify (\ref{eq:ibpg})
-\begin{align*}
-\bm{1}^T \diag{\bm{G}_{ij}} \hat{\bm{Q}}^j_N \bm{u} =&  \LRp{\bm{G}^q_{ij}}^T { \bm{Q}_i}  \bm{u}_q - \frac{1}{2} \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q+\bm{G}^f_{ij}}^T{\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q}\bm{u}_q\\
-&+ \frac{1}{2} \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q+\bm{G}^f_{ij}}^T\frac{1}{2}\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{u}_f\\
-=& \LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}}_i - \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q}^T{\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q}}\bm{u}_q + \LRp{\bm{G}^f_{ij}}^T\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{u}_f.
-%\\
-%&= \LRp{\bm{G}^q_{ij}}^T \LRp{ \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q}^T{\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q}}\bm{u}_q + \LRp{\bm{G}^f_{ij}}^T\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{u}_f 
-\end{align*}
-The first term can be further simplified using Lemma~\ref{lemma:vsbp}
-\[
-\LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}}_i - \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q}^T{\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q}}\bm{u}_q = \LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}}_i}^T\bm{u}_q.
-\]
-The latter term can be combined with the surface term of (\ref{eq:vsbp1}) by noting that 
-\[
-\LRp{\bm{G}^f_{ij}}^T\bm{W}_f \diag{\hat{\bm{n}}_i}\bm{u}_f =\bm{1}^T\hat{\bm{B}}^j_N\diag{\bm{G}_{ij}}\bm{u} = \bm{1}^T\diag{\bm{G}_{ij}}\hat{\bm{B}}^j_N\bm{u},
-\]
-where we have used that $\hat{\bm{B}}^j_N$ is diagonal and commutes with diagonal scaling by $\bm{G}_{ij}$.
-%as $\LRp{\bm{G}^f_{ij}}^T\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{u}_f = \bm{1}^T\hat{\bm{B}}^j_N \bm{u}$.  
-We can use polynomial exactness and the fact that the geometric terms satisfy the continuous GCL by construction \cite{chan2018discretely} to show that $\sum_{j=1}^d  \hat{\bm{Q}}_j {\bm{G}^q_{ij}} = 0$.
-Combining this with (\ref{eq:vsbp1}) and expanding out $\hat{\bm{B}}^j_N$ yields that
-\begin{align*}
-\bm{1}^T\bm{Q}^i_N\bm{u} &= \sum_{j=1}^d \LRp{
-\frac{1}{2}\LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}_i}}^T \bm{u}_q + \bm{1}^T\begin{pmatrix}
-\bm{0} &\\
-& \bm{W}_f \diag{\bm{G}_{ij} \circ \hat{\bm{n}}_j}
-\end{pmatrix}
-\bm{u}} \\
-&=  \bm{1}^T\begin{pmatrix}
-\bm{0} &\\
-& \bm{W}_f \diag{\sum_{j=1}^d \bm{G}_{ij} \circ \hat{\bm{n}}_j}
-\end{pmatrix}
-\bm{u} = \bm{1}^T\bm{B}^i_N\bm{u}.
-\end{align*}
-where we have used in the final step that $\sum_{j=1}^d\bm{G}_{ij}\hat{n}_j = \bm{n}_i J^k_f$ \cite{ciarlet1978finite, chan2018discretely}
-
-\end{proof}
-
-Ensuring that Lemma~\ref{lemma:vdsbpcurved} holds to take $R$ as large as possible, which maximizes the degree to which geometric terms are approximated.  However,
-
-%\begin{theorem}
-%Let the geometric terms be approximated using (\ref{eq:iconscurl}) with $1\leq R \leq (N+1)$.
-Then, the skew-symmetric formulation is stable on non-conforming meshes using any composite surface quadrature which is exact for degree $(N+R-1)$ polynomials on $\partial \hat{D}$.
-%\end{theorem}
-%\begin{proof}
-%
-%\note{Trick here is to note effect of geometric mapping: surface quadrature must be tied to degree of geometric factors.
-For Kopriva geofacs where $J\bm{G}_{ij}\in P^{N-1}$, it suffices to take surface quadrature to be exact for degree $(2N-1)$ polynomials.
-For general geofacs $J\bm{G}_{ij}\in V^N$, surface quadrature which is exact for degree $2N$ polynomials must be used.
-Note that this is still more flexible than the strong form: you can still use different surface quadratures on each face.
-}
-%
-%\end{proof}
-Thus, if geometric terms are approximated using degree $N$ polynomials as in \cite{kopriva2006metric}, then the skew-symmetric formulation with composite GLL quadrature can be used
-
-\subsection{Simplicial vs tensor product elements}
-
-It was shown in \cite{chan2018discretely} that the approximation error is $O\LRp{h^{R+1}}$, and that taking $R = (N+1)$ approximates the geometric terms with an error of $O(h^{N+2})$.
-
-\note{Need to note image of curl is in different spaces for simplicial vs tensor product elements.  Influences proofs that geometric terms satisfy appropriate assumptions in those cases.}
-%\subsection{On implementation and computational cost}
-
-\section{Numerical experiments}
-
-In this section, we present two-dimensional experiments on hybrid meshes consisting of quadrilaterals and triangles, as well as on non-conforming meshes of quadrilateral elements.
-
-\subsection{Reduced surface quadrature}
-
-\note{Affine and curved triangles with GLL surface quadrature (under-integrated).}
-
-\subsection{Hybrid meshes}
-
-\subsection{Non-conforming meshes}
-
-\appendix
-
-\section{An explicit skew-symmetric entropy stable formulation for DG-SEM}
-
-\note{$\bm{V}_q,\bm{P}_q = \bm{I}$, while $\bm{V}_f$ reduces to a permutation matrix.  }
+\note{Do we need 3 layers of surface nodes in 3D?  Interpolation to isotropic split-side nodes is a Kronecker product - does that reduce things down?  }
 

### 2018-07-19 docs/ESDG_GQ/figs/plotDN.m

@@ -1,0 +1,44 @@
+N = 5;
+[rq wq] = JacobiGQ(0,0,N);
+rp = linspace(-1,1,50)';
+Vq = Vandermonde1D(N,rq);
+D = GradVandermonde1D(N,rq)/Vq;
+Vf = Vandermonde1D(N,[-1;1])/Vq;
+Vp = Vandermonde1D(N,rp)/Vq;
+
+B = [-1 0;0 1];
+QN = [D - .5*Vf'*B*Vf .5*Vf'*B;
+    -.5*B*Vf .5*B];
+PN = [eye(N+1) diag(1./wq)*Vf'];
+
+% gex = @(x) exp(x).*sin(pi*x);%exp(-25*x.^2); %
+% gex = @(x) 1./(1+25*(x-0).^2);
+a = 4;
+gex = @(x) exp(-a*x.^2);
+dgex = @(x) -2*a*x.*exp(-a*x.^2);
+fex = @(x) 1 + 0*sin(pi*x);
+f = fex(rq);
+g = gex(rq);
+fdgex = @(x) fex(x).*dgex(x);
+
+fN = fex([rq;-1;1]);
+gN = gex([rq;-1;1]);
+u1 = f.*(D*g);
+u2 = PN*(fN.*(QN*gN));
+hold on
+p1 = plot(rp,Vp*u1,'b-','linewidth',2);
+p2 = plot(rp,Vp*u2,'r-','linewidth',2);
+plot(rq,u1,'bo','linewidth',2,'markersize',15,'MarkerFaceColor',[.49 1 .63])
+plot(rq,u2,'rs','linewidth',2,'markersize',15,'MarkerFaceColor',[.49 1 .63])
+plot(rp,fdgex(rp),'k--','linewidth',2)
+% legend('GSBP','Decoupled SBP')
+set(gca,'fontsize',16)
+grid on
+
+
+print_pgf_coordinates(rq,u1)
+print_pgf_coordinates(rp,Vp*u1)
+print_pgf_coordinates(rq,u2)
+print_pgf_coordinates(rp,Vp*u2)
+print_pgf_coordinates(rp,fdgex(rp))
+

### 2018-07-27 matlab/Codes1D/JacobiGL.m
updates before taking computer in

@@ -12,7 +12,7 @@ if (N==1)
     return; 
 end;
 
-[xint,w] = JacobiGQ(alpha+1,beta+1,N-2);
+[xint,wq] = JacobiGQ(alpha+1,beta+1,N-2);
 x = [-1, xint', 1]';
 
 if nargout == 2

### 2018-07-27 matlab/cfd2D/Euler2D_skew.m

@@ -1,32 +1,27 @@
 clear
 Globals2D
 
-a = 1/8; % warping factor
+a = 0/8; % warping factor
 % FinalTime = 10;
 
-N = 4;
-K1D = 8;
-FinalTime = 1.1;
+N = 2;
+K1D = 16;
+FinalTime = 1;
 
 wadgProjEntropyVars = abs(a)>1e-8;
-CFL = .5;
+CFL = .25;
 global tau
-tau = 0;
+tau = 1;
 global useSkew;
 useSkew = 1;
-useGLL = 1;
+useGLL = 0;
 
 
 % Lx = 7.5; Ly = 5; ratiox = 3/4; ratioy = .5;
 Lx = 10; Ly = 5; ratiox = 1; ratioy = Ly/Lx;
 [Nv, VX, VY, K, EToV] = unif_tri_mesh(round(ratiox*K1D),round(K1D*ratioy));
-
-% ids = abs(abs(VX)-1)>1e-8 & abs(abs(VY)-1)>1e-8;
-% VY(ids) = VY(ids) + .05*randn(size(VX(ids)));
-
-
-VX = VX/max(abs(VX));  VY = VY/max(abs(VY));
-VX = (VX+1)*Lx; VY = VY*Ly;
+% VX = VX/max(abs(VX));  VY = VY/max(abs(VY));
+% VX = (VX+1)*Lx; VY = VY*Ly;
 
 % VX = VX + randn(size(VX));
 StartUp2D;
@@ -44,7 +39,7 @@ global M Vq Pq Lq Vfqf Vfq Pfqf VqPq VqLq
 global rxJ sxJ ryJ syJ rxJf sxJf ryJf syJf
 global nxJ nyJ nrJ nsJ nrJq nsJq wfq
 global mapPq
-Nq = 2*N+1;
+Nq = 2*N;
 [rq sq wq] = Cubature2D(Nq); % integrate u*v*c
 [rq sq wq] = QNodes2D(N); if length(rq)~=Np;  keyboard; end
 
@@ -174,6 +169,15 @@ for e = 1:K
     end
 end
 
+global mapBwall
+%mapBwall = find(abs(xf-max(x(:)))<1e-8 | abs(xf-min(x(:)))<1e-8);
+mapBwall = find(abs(yf-max(y(:)))<1e-8 | abs(yf-min(y(:)))<1e-8);
+
+% plot(xf,yf,'o')
+% hold on
[…]

### 2018-07-28 docs/ESDG_GQ/esdg_gq.tex

@@
 \usepackage{amsmath,amssymb,amsfonts}
+\newtheorem*{remark}{Remark}
 
@@
 \newcommand{\LinfDk}{L^{\infty}\LRp{D^k}}
-
 \newcommand{\diag}[1]{{\rm diag}\LRp{#1}}
-
 \newcommand{\Ksub}{K_{\rm sub}}
-
 \newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
@@
 
+\newcommand*{\vertbar}{\rule[-1ex]{0.5pt}{2.5ex}}
+\newcommand*{\horzbar}{\rule[.5ex]{2.5ex}{0.5pt}}
+
 %\date{}
 \author{Jesse Chan, David C.\ Del Rey Fernandez, Mark H.\ Carpenter}
-\title{Entropy stable Gauss-Legendre collocation methods}
+\title{Efficient entropy stable Gauss collocation methods}
 \graphicspath{{./figs/}}
@@
 \begin{abstract}
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-Here is the abstract.
-
+The construction of high order entropy stable collocation schemes on quadrilateral and hexahedral elements has relied on the use of Gauss-Legendre-Lobatto collocation points \cite{fisher2013high, carpenter2014entropy, gassner2016split} and their equivalence with summation-by-parts (SBP) finite difference operators \cite{gassner2013skew}.
+In this work, we show how to efficiently generalize the construction of semi-discretely entropy stable schemes on tensor product elements to Gauss points and generalized SBP operators.
+Numerical experiments suggest that the use of Gauss points can significant improve accuracy on curved meshes.
 \end{abstract}
@@
 
-Time dependent nonlinear conservation laws are ubiquitous in computational fluid dynamics (CFD).
+%Time dependent nonlinear conservation laws are ubiquitous in computational fluid dynamics (CFD).
 These conservation laws include the compressible Euler and Navier-Stokes equations, which govern compressible fluid behavior within aerospace applications, as well as the equations governing coastal flooding and geophysical flows (shallow water), and those describing astrophysical phenomena (magnetohydrodynamics).
 The utility of these simulations depends both on the reliability and accuracy of numerical methods for the solution of the associated nonlinear partial differential equations (PDEs).
-
-High order methods are increasingly of interest for these applications.
+Time dependent nonlinear conservation laws are ubiquitous in computational fluid dynamics (CFD), for which high order methods are increasingly of interest.
 Such methods are more accurate per degree of freedom than low order methods, while also possessing much smaller numerical dispersion and dissipation errors.
@@
 The instability of high order methods is tied to the fact that, when discretizing systems of nonlinear PDEs, the chain rule does not typically hold at the discrete level.
+
 The lack of a chain rule was circumvented by using a non-standard ``flux differencing'' formulation \cite{fisher2013high, carpenter2014entropy, gassner2016split, gassner2017br1}, which is key to constructing semi-discretely entropy stable high order schemes on unstructured quadrilateral and hexahedral meshes.
@@
 \]
-Integrating (\ref{eq:entropyineq}) over a domain $\Omega$ and applying the divergence theorem yields an averaged entropy inequality
+Integrating (\ref{eq:entropyineq}) over a domain $\Omega$ and applying the divergence theorem yields an integrated entropy inequality
 \begin{equation}
@@
 \subfloat[GLL nodes]{\includegraphics[width=.3\textwidth]{figs/sbp.png}}
-\hspace{1em}
+\hspace{.5em}
 \subfloat[Staggered grid nodes]{\includegraphics[width=.3\textwidth]{figs/staggered.png}}
-\hspace{1em}
+\hspace{.5em}
 \subfloat[Gauss nodes]{\includegraphics[width=.3\textwidth]{figs/gsbp.png}}
@@
 \subsection{Gauss nodes and generalized summation by parts operators}
-
+\label{sec:gsbp}
 We assume the solution is collocated at $(N+1)$ quadrature points $x_i$ with associated quadrature weights $w_i$.
@@
 \begin{lemma}
+\label{lemma:sbp}
 $\bm{Q} = \bm{W}\bm{D}$ satisfies the generalized summation by parts property
@@
 \]
-\label{lemma:sbp}
 \end{lemma}
-\begin{proof}
+%\begin{proof}
 The proof is a direct restatement of integration by parts, and can be found in \cite{ortleb2016kinetic, ortleb2017kinetic, ranocha2018generalised}.  
-We reproduce it here for completeness.
-Let $\bm{u},\bm{v}$ be two arbitrary vectors, and let $u(x), v(x)$ denote the polynomials whose nodal values are given by $\bm{u},\bm{v}$.
-Then,
-\begin{align*}
-\bm{v}^T\bm{Q}\bm{u} &= \bm{v}^T\bm{W}\bm{D}\bm{u} = \int_{-1}^1 \pd{u}{x}v = \LRl{uv}_{-1}^1 - \int_{-1}^1 u\pd{v}{x} \\
-&= \LRp{\bm{V}_f \bm{v}}^T\bm{B}\bm{V}_f\bm{u} - \LRp{\bm{D}\bm{v}}^T\bm{W}\bm{u} = \bm{v}^T\bm{V}_f^T\bm{B}\bm{V}_f \bm{u} - \bm{v}^T\bm{Q}^T\bm{u},
-\end{align*}
-where we have used that $\bm{W}$ is diagonal and that $(N+1)$-point Gauss quadrature is exact for the above integrands.
+%We reproduce it here for completeness.  Let $\bm{u},\bm{v}$ be two arbitrary vectors, and let $u(x), v(x)$ denote the polynomials whose nodal values are given by $\bm{u},\bm{v}$.  Then, 
+%\begin{align*}
+%\bm{v}^T\bm{Q}\bm{u} &= \bm{v}^T\bm{W}\bm{D}\bm{u} = \int_{-1}^1 \pd{u}{x}v = \LRl{uv}_{-1}^1 - \int_{-1}^1 u\pd{v}{x} \\
+%&= \LRp{\bm{V}_f \bm{v}}^T\bm{B}\bm{V}_f\bm{u} - \LRp{\bm{D}\bm{v}}^T\bm{W}\bm{u} = \bm{v}^T\bm{V}_f^T\bm{B}\bm{V}_f \bm{u} - \bm{v}^T\bm{Q}^T\bm{u},
+%\end{align*}
+%where we have used that $\bm{W}$ is diagonal and that $(N+1)$-point Gauss quadrature is exact for the above integrands.
 Taking $\bm{v} = \bm{e}_i$ and $\bm{u} = \bm{e}_j$ for $i,j = 1,\ldots, N+1$ (where $\bm{e}_i$ denotes the $i$th canonical vector) completes the proof.
-\end{proof}
-
+%\end{proof}
 Lemma~\ref{lemma:sbp} holds for both GLL and Gauss nodes, and switching between these two nodal sets simply results in a redefinition of the matrices $\bm{D}, \bm{V}_f$.
@@
 \begin{definition}
+\label{def:tadmor}
 Let $\bm{f}_S(\bm{u}_L,\bm{u}_R)$ be a bivariate function which is symmetric and consistent with the flux function $\bm{f}(\bm{u})$
@@
 \end{align*}
-\label{def:tadmor}
 \end{definition}
-Here, $\psi$ is the entropy potential.
-
-\note{Ask DCDR to fill this in \cite{fernandez2014review, crean2017high, fernandez2018simultaneous}.}
+
+For illustrative purposes, we will prove a semi-discrete entropy inequality on a one-dimensional mesh consisting of two elements of degree $N$.
+We assume both meshes are translations of a reference element  $[-1,1]$, such that derivatives with respect to physical coordinates are identical to derivatives with respect to reference coordinates.
+The extension to multiple elements and variable mesh sizes is straightforward.
+
+The construction of entropy conservative schemes relies on appropriate SBP-SATs for Gauss collocation schemes \cite{fernandez2014review, crean2017high, fernandez2018simultaneous}.
+Let the rows of $\bm{V}_f$ be denoted as $\bm{t}_L, \bm{t}_R$
+\[
+\bm{V}_f = \begin{bmatrix}
+\horzbar & \bm{t}_L & \horzbar\\
+\horzbar & \bm{t}_R & \horzbar
+\end{bmatrix}, \qquad \LRp{\bm{t}_L}_j = \ell_j(-1), \qquad \LRp{\bm{t}_R}_j = \ell_j(1).
+\]
+The inter-element coupling terms in \cite{fernandez2014review, crean2017high, fernandez2018simultaneous} utilize a decomposition of the surface mass matrix $\bm{V}_f^T\bm{B}\bm{V}_f$ as  
+\begin{equation}
+\bm{V}_f^T\bm{B}\bm{V}_f 
+= \bm{t}_R\bm{t}_R^T - \bm{t}_L\bm{t}_L^T.
+%, \qquad \bm{V}_f = \begin{bmatrix}
+%\bm{t}_L & \bm{t}_R
+%\end{bmatrix}
+\label{eq:bmatdecomp}
+\end{equation}
+The construction of entropy conservative schemes on multiple elements requires appropriate inter-element coupling terms (SBP-SATs) involving $\bm{t}_L, \bm{t}_R$.
+We consider a two element mesh, and show how SBP-SATs can be interpreted as constructing a global SBP operator.
+
+Let $\bm{u}^1_N, \bm{u}^2_N$ denote nodal degrees of freedom of the vector valued solution $\bm{u}(x)$ on the first and second element, respectively.
+To simplify notation, we assume that all following operators are defined in terms of Kronecker products, such that they are applied to each component of $\bm{u}^1_N, \bm{u}^2_N$.
+We first define the matrix 
+\begin{equation}
+\bm{S} = \bm{Q} - \frac{1}{2}\bm{V}_f^T\bm{B}\bm{V}_f.
+\label{eq:Smat}
+\end{equation}
+It is straightforward to show (using Lemma~\ref{lemma:sbp}) that $\bm{S}$ is skew-symmetric.  We can now define an SBP operator $\bm{D}_h = \bm{W}_h^{-1}\bm{Q}_h$ over two elements
+\begin{equation}
+\bm{Q}_{h} = 
+\underbrace{\begin{bmatrix}
+\bm{S}  & \frac{1}{2}\bm{t}_R\bm{t}_L^T \\
+- \frac{1}{2}\bm{t}_L\bm{t}_R^T & \bm{S}
+\end{bmatrix}}_{\bm{S}_h}
++
+\underbrace{\begin{bmatrix}
+-\frac{1}{2}\bm{t}_L\bm{t}_L^T  & \\
+ & \frac{1}{2} \bm{t}_R\bm{t}_R^T
+\end{bmatrix}}_{\frac{1}{2}\bm{B}_h}
+, \qquad 
+\bm{W}_h = \begin{bmatrix}\bm{W} &\\
+& \bm{W}
+\end{bmatrix}.
+%\bm{D}_h = \begin{bmatrix}\bm{W}^{-1} &\\
+%& \bm{W}^{-1}
+%\end{bmatrix}
+% \bm{Q}_h.
+\label{eq:gsbp}
+\end{equation}
+It can be shown that $\bm{D}_h$ is high order accurate such that, if $\bm{u}_h$ is a polynomial of degree $N$, it is differentiated exactly.
+Straightforward computations show that $\bm{Q}_h$ also satisfies an SBP property $\bm{Q}_h + \bm{Q}_h^T = \bm{B}_h$.
+%\[
+%\bm{Q}_h + \bm{Q}_h^T = \bm{B}_h.
+%\begin{bmatrix}
+%-\bm{t}_L\bm{t}_L  & \\
+% & \bm{t}_R\bm{t}_R
+%\end{bmatrix}.
+%\]
+
+Ignoring boundary conditions, an entropy conservative scheme for (\ref{eq:nonlinpde}) on two elements can then be given as
+\begin{gather}
+\td{}{t}\bm{W}_h\bm{u}_h + 2\LRp{\bm{Q}_h \circ \bm{F}_S}\bm{1} = 0, \qquad \bm{u}_h = \begin{bmatrix}
+\bm{u}^1_N\\
+\bm{u}^2_N
+\end{bmatrix} \label{eq:ESgsbp} \\
+\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\LRp{\bm{u}_h}_i,\LRp{\bm{u}_h}_j}, \qquad 1 \leq  i,j \leq 2(N+1). \nonumber
+\end{gather}
+where $\circ$ denotes the Hadamard product \cite{horn2012matrix}.  Multiplying (\ref{eq:ESgsbp}) by $\bm{v}_h^T = \bm{v}\LRp{\bm{u}_h}^T$ 
+%\begin{align*}
+%\bm{v}_h^T\td{}{t}\bm{W}_h\bm{u}_h + 2\bm{v}_h^T\LRp{\bm{Q}_h \circ \bm{F}_S}\bm{1} = 0
+%\end{align*}
+will yield a semi-discrete version of the conservation of entropy (mimicking (\ref{eq:weakentropyineq}) with the inequality replaced by an equality)
+\begin{equation}
+\label{eq:consentropyGSBP}
+\td{}{t}\bm{W}_h S(\bm{u}_h) + \bm{v}_h^T\LRp{\bm{B}_h\circ \bm{F}_S}\bm{1} -\bm{1}^T\bm{B}_h\psi\LRp{\bm{u}_h} = 0.
+\end{equation}
+We refer to \cite{crean2017high,crean2018entropy} for the proof of (\ref{eq:consentropyGSBP}).
+
+The drawback of the SBP-SATs introduced in this section lies in the nature of the off-diagonal matrices $\bm{t}_R\bm{t}_L$ and $-\bm{t}_L\bm{t}_R$.
+For Gauss nodes, these blocks are dense, which implies that inter-element coupling terms produce a non-compact stencil.
+Evaluating (\ref{eq:ESgsbp}) requires computing two-point fluxes $\bm{f}_S$ between all nodes on two neighboring elements, which significantly increases both the computational work, as well as communication between neighboring elements.
+The main goal of this work is to circumvent the all-to-all coupling of degrees of freedom introduced by the SBP-SATs described in this section, which can be done through the use of ``decoupled'' SBP operators.
 
@@
 
-The main goal of this work is to circumvent the all-to-all coupling of degrees of freedom introduced by the SBP-SAT terms described in Section~\ref{sec:gsbpsat}, and will do so through the use of ``decoupled'' SBP operators.
 Decoupled SBP operators were first introduced in \cite{chan2017discretely} and used to construct entropy stable schemes on simplicial elements.
-These operators (and simplifications under a collocation assumption) are presented in a more general setting in Appendix~\ref{app:decoupled}.
-In this section, decoupled SBP operators are introduced in one dimension for either GLL or Gauss nodal sets.
-
-Let $\bm{Q} = \bm{W}\bm{D}$.  The decoupled SBP operator $\bm{Q}_N$ is defined as the block matrix 
+These operators (and simplifications under a collocation assumption) are presented in a more general setting in \cite{chan2017discretely, chan2018discretely} and in Appendix~\ref{app:decoupled}.
+In this section, decoupled SBP operators are introduced in one dimension for GLL and Gauss nodal sets.
+
+Decoupled SBP operators build upon the GSBP matrices $\bm{W}, \bm{Q}$, interpolation matrix $\bm{V}_f$, and boundary matrix $\bm{B}$ introduced in Section~\ref{sec:gsbp}.
+The decoupled SBP operator $\bm{Q}_N$ is defined as the block matrix
 \begin{equation}
@@
 where we have used the fact that diagonal matrices commute to simplify expressions.
-Thus, the decoupled SBP operator $\bm{Q}_N$ can be interpreted as adding boundary corrections to the GSBP operator $\bm{D}$ in a skew-symmetric fashion.
+The decoupled SBP operator $\bm{Q}_N$ can thus be interpreted as adding boundary corrections to the GSBP operator $\bm{D}$ in a skew-symmetric fashion.
 
@@
 
-\caption{Degree $N= 5$ approximations of derivatives of a Gaussian $e^{-4x^2}$ using the generalized SBP operator $\bm{D}$ and the decoupled SBP operator $\bm{Q}_N$ via (\ref{eq:qn}). }
+\caption{Degree $N= 5$ approximations of derivatives of a Gaussian $e^{-4x^2}$ using the generalized SBP operator $\bm{D}$ and the decoupled SBP operator $\bm{Q}_N$ via (\ref{eq:qn}).
+The colored circles and squares denote values at Gauss points.
+}
 \label{fig:dsbpcorrect}
@@
 We can now construct an entropy conservative Gauss collocation scheme with compact SBP-SATs using decoupled SBP operators.
-Let $\bm{u}_N$ and $\bm{v}(\bm{u}_N)$ denote the values of the conservative variables and entropy variables at Gauss points, respectively.
-We define approximate values $\tilde{\bm{u}}_f$ of the conservative variables by interpolating the values of the entropy variables to element boundaries
-\[
-\tilde{\bm{u}}_f = \bm{u}(\bm{v}_f), \qquad \bm{v}_f = \bm{V}_f\bm{v}(\bm{u}_N).  
-\]
-We now define $\tilde{\bm{u}}$ as the concatenated vector of solution values at Gauss points and the ``entropy-projected conservative variables'' $\tilde{\bm{u}}_f$.
-We also define $\bm{F}_S$ as the matrix of evaluations of the two-point flux $\bm{f}_S$ at combinations of values of $\tilde{\bm{u}}$
-\[
-\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad \tilde{\bm{u}} = \begin{bmatrix}
-\bm{u}_N\\
-\tilde{\bm{u}}_f
+As in Section~\ref{sec:gsbpsat}, we will construct a Gauss collocation scheme and provide a proof of semi-discrete entropy conservation for a two-element mesh.
+
+(see Figure~\ref{fig:twoelem})
+\note{Add pics of two-element mesh}
+
+We first note that $\bm{B}$ can be trivially decomposed into the sum of two outer products as in (\ref{eq:bmatdecomp})
+\begin{equation}
+\bm{B} = \begin{bmatrix}
+-1 & 0 \\
+0 & 1
+\end{bmatrix} = \bm{e}_R\bm{e}_R^T - \bm{e}_L\bm{e}_L^T, \qquad \bm{e}_L = \begin{bmatrix} 1\\0\end{bmatrix}, \qquad \bm{e}_R = \begin{bmatrix} 0\\1\end{bmatrix}.
+\label{eq:bmatdecomp2}
+\end{equation}
+The vectors $\bm{e}_L, \bm{e}_R$ are related to $\bm{t}_L, \bm{t}_R$ through the interpolation matrix $\bm{V}_f$.
+Because $\bm{t}_L, \bm{t}_R$ are rows of $\bm{V}_f$, $\bm{t}_L\bm{1} = \bm{t}_R\bm{1} = 1$.
+This can be used to show, for example, that
+\begin{equation}
+\label{eq:et}
+\bm{B}\bm{V}_f\bm{1} = \LRp{\bm{e}_R\bm{e}_R^T - \bm{e}_L\bm{e}_L}\bm{1} = \bm{e}_R-\bm{e}_L.
+\end{equation}
+
+We can define a decoupled SBP matrix $\bm{Q}_h$ over two elements as follows
+\begin{align}
+\bm{Q}_h =& \LRs{\begin{array}{cc|cc}
+\bm{S} &  \frac{1}{2}\bm{V}_f^T\bm{B} & & \\
+ -\frac{1}{2}\bm{B}\bm{V}_f & -\frac{1}{2}\bm{e}_L\bm{e}_L^T & & \frac{1}{2}\bm{e}_R\bm{e}_L^T\\
+ \hline
+&& \bm{S} & \frac{1}{2}\bm{V}_f^T\bm{B} \\
+& -\frac{1}{2}\bm{e}_L\bm{e}_R^T & - \frac{1}{2}\bm{B}\bm{V}_f & \frac{1}{2}\bm{e}_R\bm{e}_R^T
+\end{array}} \label{eq:decoupledsbp2elem}
+\end{align}
+We can also show that $\bm{Q}_h\bm{1} = 0$.  Using (\ref{eq:et}), we have that
+\begin{align}
+\bm{Q}_h\bm{1} &= \begin{bmatrix}
+\LRp{\bm{S} + \frac{1}{2}\bm{V}_f^T\bm{B} }\bm{1}\\
+\frac{1}{2}\bm{B}\bm{V}_f\bm{1} + \frac{1}{2}\LRp{\bm{e}_R-\bm{e}_L}\\
+\LRp{\bm{S} + \frac{1}{2}\bm{V}_f^T\bm{B} }\bm{1}\\
+\frac{1}{2}\bm{B}\bm{V}_f\bm{1} + \frac{1}{2}\LRp{\bm{e}_R-\bm{e}_L}
+\end{bmatrix} 
+= 
+\begin{bmatrix}
+\LRp{\bm{Q} - \frac{1}{2}\bm{V}_f^T\bm{B}\bm{V}_f + \frac{1}{2}\bm{V}_f^T\bm{B}\bm{V}_f}\bm{1}\\
+\frac{1}{2}\bm{B}\LRp{-\bm{1} + \bm{1}}\\
+\LRp{\bm{Q} - \frac{1}{2}\bm{V}_f^T\bm{B}\bm{V}_f + \frac{1}{2}\bm{V}_f^T\bm{B}\bm{V}_f}\bm{1}\\
+\frac{1}{2}\bm{B}\LRp{-\bm{1} + \bm{1}}\\
+\end{bmatrix} = \bm{0}.
+\label{eq:Q1zero}
+\end{align}
+Here, we have used the definition of $\bm{B}$ in (\ref{eq:bmatdecomp2}), the fact that $\bm{V}_f\bm{1} = \bm{1}$ and $\bm{Q}\bm{1} = 0$ \cite{fernandez2014generalized}, and the definition of $\bm{S}$ in (\ref{eq:Smat}).
+This property (\ref{eq:Q1zero}) will be used in the proof of entropy conservation.
+
+It can be helpful to split up $\bm{Q}_h$ into two matrices
+\begin{gather*}
+\bm{Q}_h = \bm{S}_h + \frac{1}{2}\bm{B}_h, \qquad
+\bm{S}_h= \LRs{\begin{array}{cc|cc}
+\bm{S} &  \frac{1}{2}\bm{V}_f^T\bm{B} & & \\
+ - \frac{1}{2}\bm{B}\bm{V}_f &  & & \frac{1}{2}\bm{e}_R\bm{e}_L^T\\
+ \hline
+&& \bm{S} &  \frac{1}{2}\bm{V}_f^T\bm{B} \\
+& -\frac{1}{2}\bm{e}_L\bm{e}_R^T &  - \frac{1}{2}\bm{B}\bm{V}_f &\\
+\end{array}}\\
+\bm{B}_h = \LRs{\begin{array}{cc|cc}
+&  & & \\
+ & -\bm{e}_L\bm{e}_L^T & & \\
+ \hline
+&&  &  \\
+&  && \bm{e}_R\bm{e}_R^T\\
+\end{array}}.
+\end{gather*}
+The matrix $\bm{S}_h$ is skew-symmetric, while the matrix $\bm{B}_h$ functions as a boundary operator which extracts boundary values over the two-element domain (i.e.\ $\bm{B}_h$ extracts a left boundary value from element 1 and a right boundary value from element 2).
+Note that here, $\bm{B}_h$ is diagonal, unlike the boundary operator defined in Section~\ref{sec:gsbpsat}.
+
+The two-element decoupled SBP operator $\bm{Q}_h$ in (\ref{eq:decoupledsbp2elem}) is related to the GSBP operator in (\ref{eq:gsbp}) through a block interpolation matrix
+\[
+\bm{I}_h = \begin{bmatrix}
+\bm{I} &\\
+\bm{V}_f &\\
+&\bm{I}\\
+&\bm{V}_f\\
+\end{bmatrix}, \qquad \bm{I}_h^T\bm{Q}_h\bm{I}_h = {\begin{bmatrix}
+\bm{S}  & \frac{1}{2}\bm{t}_R\bm{t}_L^T \\
+- \frac{1}{2}\bm{t}_L\bm{t}_R^T & \bm{S}
+\end{bmatrix}}
++
+{\begin{bmatrix}
+-\frac{1}{2}\bm{t}_L\bm{t}_L^T  & \\
+ & \frac{1}{2} \bm{t}_R\bm{t}_R^T
+\end{bmatrix}}.
+\]
+
+Let $\bm{u}^1_N, \bm{u}^2_N$ denote the values of the conservative variables at Gauss points on elements 1 and 2, respectively, and let $\bm{u}_h$ denote their concatenation as defined in (\ref{eq:ESgsbp}).
+Let $\bm{v}(\bm{u}^i_N)$ denote the evaluation of entropy variables at Gauss points on element $i$, and define the ``entropy-projected conservative variables'' $\tilde{\bm{u}}^1_f, \tilde{\bm{u}}^2_f$ by evaluating the conservative variables in terms of the  interpolated values of the entropy variables at element boundaries
+\[
+\tilde{\bm{u}}^i_f = \bm{u}(\bm{v}^i_f), \qquad \bm{v}^i_f = \bm{V}_f\bm{v}(\bm{u}^i_N), \qquad i = 1,2. 
+\]
+We now introduce $\bm{u}_h = \begin{bmatrix}\bm{u}^1_N & \bm{u}^2_N\end{bmatrix}^T$ as the concatenated vector of solution values at Gauss points, and define $\bm{v}_h, \tilde{\bm{v}}$, and $\tilde{\bm{u}}$ as follows:
+\[
+\bm{v}_h = \bm{v}\LRp{\bm{u}_h} = \bm{v}\LRp{\begin{bmatrix}
+{\bm{u}^1_N}\\
+{\bm{u}^2_N}
+\end{bmatrix}}, \qquad \tilde{\bm{v}} = \bm{I}_h\bm{v}_h, \qquad \tilde{\bm{u}} = \bm{u}\LRp{\tilde{\bm{v}}} = 
+%\bm{u}\LRp{
+%\begin{bmatrix}
+%\bm{I}\\
+%\bm{V}_f 
+%\end{bmatrix}\bm{v}
+%}=
+\begin{bmatrix}
+\bm{u}^1_N \\
+\tilde{\bm{u}}^1_f \\
+\bm{u}^2_N \\
+\tilde{\bm{u}}^2_f
 \end{bmatrix}.  
 \]
+The term $\bm{v}_h$ corresponds to the vector of entropy variables at $\bm{u}_h$, while $\tilde{\bm{v}}$ corresponds to the concatenated vector of Gauss point values and interpolated boundary values $\bm{v}^i_f$ of the entropy variables.
+The term $\tilde{\bm{u}}$ denotes the evaluation of the conservative variables in terms of $\tilde{\bm{v}}$.
+
+Finally, we define $\bm{F}_S$ as the matrix of evaluations of the two-point flux $\bm{f}_S$ at combinations of values of $\tilde{\bm{u}}$
+\[
+\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \quad 1\leq i,j \leq 2(N+3).
+\]
+Note that, due to the consistency of $\bm{f}_S$, the diagonal of $\bm{F}_S$ reduces to flux evaluations
+\begin{equation}
+\label{eq:diagFS}
+\LRp{\bm{F}_S}_{ii} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j} = \bm{f}\LRp{\tilde{\bm{u}}_i}.  
+\end{equation}
+We can now construct a semi-discretely entropy conservative formulation based on decoupled SBP operators:
 \begin{theorem}
 \label{thm:consentropy}
-Let $\circ$ denote the Hadamard product, and let $\tilde{\bm{u}}_f^+$ denote values of the entropy-projected conservative variables on neighboring elements.  Then, the formulation
-\begin{equation}
+Let $\bm{Q}_h$ be defined by (\ref{eq:decoupledsbp2elem}), and let $\bm{u}_h$ denote the two-element solution of the following formulation
+\begin{gather}
 \label{eq:form}
-\bm{W}\td{\bm{u}_N}{t} + 2 \begin{bmatrix}
-\bm{I}\\
-\bm{V}_f
-\end{bmatrix}^T\LRp{\bm{Q}_N \circ \bm{F}_S}\bm{1} + \bm{V}_f^T \bm{B} \LRp{\bm{f}_S\LRp{\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f} - \bm{f}(\bm{u})} = 0
-\end{equation}
-satisfies the quadrature form of the semi-discrete conservation of entropy
-\[
-\bm{1}^T\bm{W}\td{S(\bm{u}_N)}{t} + \bm{1}^T\bm{B} \LRp{\bm{v}_f^T\bm{f}(\tilde{\bm{u}}_f) - \psi_i\LRp{\tilde{\bm{u}}_f}} = 0.
-\]
+\bm{W}_h \td{}{t} \bm{u}_h + \bm{I}_h^T\LRp{\bm{Q}_h \circ \bm{F}_S}\bm{1} = 0.
+\end{gather}
+Then, $\bm{u}_h$ satisfies a semi-discrete conservation of entropy
+\begin{gather*}
+\bm{1}^T\bm{W}_h\td{S(\bm{u}_h)}{t} + \bm{1}^T\bm{B}_h\LRp{\bm{v}_f^T\bm{f}(\tilde{\bm{u}}_f) - \psi\LRp{\tilde{\bm{u}}_f}} = 0\\
+\bm{v}_f = \begin{bmatrix}
+\bm{t}_L \bm{v}\LRp{\bm{u}_N^1}\\
+\bm{t}_R \bm{v}\LRp{\bm{u}_N^2}
+\end{bmatrix}, \qquad \tilde{\bm{u}}_f = \bm{u}(\bm{v}_f).
+\end{gather*}
+%Let $\circ$ denote the Hadamard product, and let $\tilde{\bm{u}}_f^+$ denote values of the entropy-projected conservative variables on neighboring elements.  Then, the formulation
+%\begin{equation}
+%\label{eq:form}
+%\bm{W}\td{\bm{u}_N}{t} + 2 \begin{bmatrix}
+%\bm{I}\\
+%\bm{V}_f
+%\end{bmatrix}^T\LRp{\bm{Q}_N \circ \bm{F}_S}\bm{1} + \bm{V}_f^T \bm{B} \LRp{\bm{f}_S\LRp{\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f} - \bm{f}(\bm{u})} = 0
+%\end{equation}
+%satisfies the quadrature form of the semi-discrete conservation of entropy
+%\[
+%\bm{1}^T\bm{W}\td{S(\bm{u}_N)}{t} + \bm{1}^T\bm{B} \LRp{\bm{v}_f^T\bm{f}(\tilde{\bm{u}}_f) - \psi_i\LRp{\tilde{\bm{u}}_f}} = 0.
+%\]
 \end{theorem}
 \begin{proof}
-Theorem~\ref{thm:consentropy} follows from choosing either GLL or Gauss quadratures in Theorem 4 of \cite{chan2017discretely}.
-We provide a separate illustrative proof on a two-element mesh.
-The extension to multiple elements is straightforward.
-
-First, we rewrite formulation (\ref{eq:form}) in a skew-symmetric form.  Using Lemma~\ref{lemma:dsbp} and the diagonal nature of $\bm{B}$, (\ref{eq:form}) is equivalent to solving
-\begin{equation}
-\bm{W}\td{\bm{u}_N}{t} + \begin{bmatrix}
-\bm{I}\\
-\bm{V}_f
-\end{bmatrix}^T\LRp{\LRp{\bm{Q}_N - \bm{Q}_N^T} \circ \bm{F}_S}\bm{1} + \bm{V}_f^T \bm{B} {\bm{f}_S\LRp{\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f}} = 0
-\label{eq:skew}
-\end{equation}
-
+The proof results from testing with $\bm{v}_h^T$.  Since $\bm{v}(\bm{u}) = \pd{S(\bm{u})}{\bm{u}}$ and $\bm{W}_h$ is diagonal, the time term yields
+\[
+\bm{v}_h^T\td{}{t}\bm{W}_h\bm{u}_h = \bm{1}^T \td{}{t}\bm{W}_h\diag{\bm{v}_h}\bm{u}_h = \bm{1}^T\td{}{t}\bm{W}_h S(\bm{u}_h).
+\]
+The spatial term can be manipulated as follows
+\begin{align*}
+2\bm{v}_h^T \bm{I}_h^T\LRp{\bm{Q}_h \circ \bm{F}_S}\bm{1} &= 2\LRp{\bm{I}_h\bm{v}_h}^T \LRp{\LRp{\bm{S}_h \circ \bm{F}_S}\bm{1} + \frac{1}{2}\LRp{\bm{B}_h \circ \bm{F}_S}\bm{1}}\\
+&= \tilde{\bm{v}}^T\LRp{\bm{B}_h \circ \bm{F}_S}\bm{1} + \tilde{\bm{v}}^T \LRp{\bm{S}_h \circ \bm{F}_S}\bm{1} - \bm{1}^T \LRp{\bm{S}_h \circ \bm{F}_S}\tilde{\bm{v}},
+\end{align*}
+where we have used the skew-symmetry of $\bm{S}_h$ in the last step.  The boundary term reduces to 
+\[
+\tilde{\bm{v}}^T\LRp{\bm{B}_h \circ \bm{F}_S}\bm{1} = \bm{1}^T\bm{B}_h\tilde{\bm{v}}_f^T\bm{f}\LRp{\tilde{\bm{u}}},
+\]
+where we have used (\ref{eq:diagFS}) and the fact that $\bm{B}_h$ is diagonal.
+Here, $\tilde{\bm{v}}_f^T\bm{f}\LRp{\tilde{\bm{u}}}$ denotes a vector whose entries are the inner product of $\tilde{\bm{v}}$ and $\bm{f}\LRp{\tilde{\bm{u}}}$ at face points.
+
+The volume terms can be manipulated using the definition of $\bm{F}_S$ and the Tadmor shuffle condition in Definition~\ref{def:tadmor}
+\begin{align*}
+\tilde{\bm{v}}^T \LRp{\bm{S}_h \circ \bm{F}_S} \bm{1} - \bm{1}^T \LRp{\bm{S}_h \circ \bm{F}_S}\tilde{\bm{v}} 
+&= \sum_{ij} \LRp{\bm{S}_h}_{ij} \LRp{\tilde{\bm{v}}_i - \tilde{\bm{v}}_j}^T\bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}\\
+&= \sum_{ij} \LRp{\bm{S}_h}_{ij} \LRp{\psi\LRp{\tilde{\bm{u}}_i} - \psi\LRp{\tilde{\bm{u}}_j}}\\
+&= \psi\LRp{\tilde{\bm{u}}}^T \bm{S}_h \bm{1}- \bm{1}^T \bm{S}_h \psi\LRp{\tilde{\bm{u}}} = 2\LRp{ \psi\LRp{\tilde{\bm{u}}}^T\bm{S}_h\bm{1}}.
+\end{align*}
+where we have again used the skew-symmetry of $\bm{S}_h$ in the last step.  Substituting $\bm{S}_h = \bm{Q}_h - \frac{1}{2}\bm{B}_h$ and using (\ref{eq:Q1zero}) yields
+\[
+2\LRp{ \psi\LRp{\tilde{\bm{u}}}^T\bm{S}_h\bm{1}} = \psi\LRp{\tilde{\bm{u}}}^T\LRp{2\bm{Q}_h - \bm{B}_h} \bm{1} =-\psi\LRp{\tilde{\bm{u}}}^T\bm{B}_h \bm{1}  = -\bm{1}^T\bm{B}_h \psi\LRp{\tilde{\bm{u}}},
+\]
+where we have used the symmetry of $\bm{B}_h$ in the last step.
 \end{proof}
 
-\note{Add two-element case.}
-
-A semi-discrete entropy inequality can be enforced by adding appropriate interface dissipation terms, such as Lax-Friedrichs or matrix dissipation \cite{winters2017uniquely}.
-We note that these terms must be computed in terms of $\tilde{\bm{u}}_f$ in order to ensure a discrete dissipation of entropy \cite{chen2017entropy,chan2017discretely}.
+\begin{remark}
+The proof of Theorem~\ref{thm:consentropy} follows directly from choosing either GLL or Gauss quadratures in Theorem 4 of \cite{chan2017discretely}.
+The proof is reproduced here for clarity, as the two-element case illuminates the skew-symmetric nature and structure of the inter-element coupling more explicitly.
+\end{remark}
+
+It is possible to convert the semi-discrete entropy equality in Theorem~\ref{thm:consentropy} to a semi-discrete entropy inequality by adding appropriate interface dissipation terms, such as Lax-Friedrichs or matrix dissipation \cite{winters2017uniquely}.
+We note that these terms must be computed in terms of $\tilde{\bm{u}}_f$ in order to ensure a discrete dissipation of entropy \cite{chen2017entropy, chan2017discretely}.
 
@@
 
-Theorem~\ref{thm:consentropy}
-
-\note{Describe GCL using GLL nodes \cite{chan2018discretely}}
-
-Retaining entropy stability on curved meshes requires satisfaction of a discrete geometric conservation law (GCL) \cite{carpenter2014entropy, gassner2017br1, crean2018entropy, chan2018discretely}.
-For two-dimensional isoparametric mappings, the GCL is automatically satisfied.
-However, in three dimensions, the GCL is not guaranteed to be satisfied at the discrete level, due to the fact that geometric terms are polynomials of degree higher than $N$.
-Modifications to the computation of geometric terms are required to ensure that the GCL is satisfied at the discrete level.
-
+The formulation in Theorem~\ref{thm:consentropy} can be naturally extended to Cartesian meshes in higher dimensions through a tensor product construction.
+We consider first the construction of higher dimensional differentiation matrices on a two-dimensional reference element $\hat{D}$, assuming a two dimensional tensor product grid of quadrature nodes (the construction of decoupled SBP operators in three dimensions is straightforward and similar to the two-dimensional case).
+We then construct physical differentiation matrices on mapped elements $D^k$, through which we construct an entropy conservative scheme.
+
+Let $\bm{D}_{\rm 1D}, \bm{W}_{\rm 1D}$ denote the 1D differentiation and mass matrices, respectively, on on the reference interval $[-1,1]$.
+Let ${\bm{W}}$ denote the 2D reference mass matrix, and let ${\bm{D}}^i$ denote the differentiation matrices with respect to the $i$th reference coordinate.
+These matrices can be expressed in terms of Kronecker products
+\[
+{\bm{D}}^1 = \bm{D}_{\rm 1D} \otimes \bm{I}_{N+1}, \qquad {\bm{D}}^2 = \bm{I}_{N+1} \otimes \bm{D}_{\rm 1D}, \qquad {\bm{W}} = \bm{W}_{\rm 1D} \otimes  \bm{W}_{\rm 1D}.
+\]
+where $\bm{I}_{N+1}$ denotes the $(N+1)\times (N+1)$ identity matrix.  %The extension to 3D is done similarly.
+
+We also construct higher dimensional face interpolation matrices.
+Let $\bm{V}_{f, {\rm 1D}}$ denote the one-dimensional interpolation matrix, and let $\bm{B}_{\rm 1D}$ denote the boundary matrix defined in Lemma~\ref{lemma:sbp}.
+For an appropriate ordering of face quadrature points, the two-dimensional face interpolation matrix $\bm{V}_f$ and reference boundary matrices ${\bm{B}}^1, {\bm{B}}^2$ can be expressed as the concatenation of Kronecker product matrices
+\[
+\bm{V}_f = \begin{bmatrix}
+\bm{V}_{f, {\rm 1D}} \otimes \bm{I}_{2}\\
+\bm{I}_{2} \otimes \bm{V}_{f, {\rm 1D}} 
+\end{bmatrix}, \qquad 
+{\bm{B}}^1 = \begin{bmatrix}
+\bm{B}_{\rm 1D} \otimes \bm{I}_{2} & \\
+& \bm{0}
+\end{bmatrix}, \qquad 
+{\bm{B}}^2 = \begin{bmatrix}
+\bm{0} &\\
+& \bm{I}_{2} \otimes \bm{B}_{\rm 1D} 
+\end{bmatrix}.
+\]
+In three dimensions, the face interpolation matrix would be expressed as the concatenation of three Kronecker products involving $\bm{V}_{f, {\rm 1D}}$.
+The higher dimensional differentiation and interpolation matrices $\bm{D}^i, \bm{V}_f$ can now be used to construct higher dimensional decoupled SBP operators.
+Let ${\bm{Q}}^i_N$ denote the decoupled SBP operator for the $i$th coordinate on the reference element, where ${\bm{Q}}^i_N$ is defined as
+\[
+{\bm{Q}}^i_N = \begin{bmatrix}
+{\bm{Q}}^i - \frac{1}{2}\bm{V}_f^T{\bm{B}}^i\bm{V}_f & \frac{1}{2}\bm{V}_f^T{\bm{B}}^i\\
+-\frac{1}{2}{\bm{B}}^i\bm{V}_f & \frac{1}{2}{\bm{B}}^i
+\end{bmatrix}, \qquad {\bm{Q}}^i = {\bm{W}}{\bm{D}}^i.
+\]
+
+Let the domain now be decomposed into non-overlapping elements $D^k$, such that $D^k$ is the image of $\hat{D}$ under a degree $N$ polynomial mapping $\bm{\Phi}^k$.
+We define geometric terms $\bm{G}^k_{ij}$ as scaled derivatives of reference coordinates $\hat{\bm{x}}$ with respect to physical coordinates $\bm{x}$
+\[
+J^k\bm{G}^k_{ij} = J^k\pd{\hat{x}_j}{\bm{x}_j},
+\]
+where $J^k$ is the determinant of the Jacobian of $\bm{\Phi}^k$.
+We also introduce the face Jacobian $J^k_f$, which is computed from the mapping of a physical face to a reference face.
+We define matrices over each physical element $D^k$
+\begin{gather*}
+\bm{W}_k = \bm{W} \diag{\bm{J}^k}, \qquad \bm{B}^i_k = \bm{B}^i_N \diag{\bm{n}_i \circ\bm{J}^f }\\
+\bm{Q}^i_k = \frac{1}{2} \sum_{j=1}^d \LRp{\diag{\tilde{\bm{G}}^k_{ij}}\bm{Q}^i_N  + \bm{Q}^i_N \diag{\tilde{\bm{G}}^k_{ij}} },
+\end{gather*}
+where we have discretized the differentiation matrix in split form \cite{nordstrom2006conservative}.  
+Here, $\bm{J}^k$ denotes the vector of values of $J^k$ at volume quadrature points, $\tilde{\bm{G}}^k_{ij}$ denotes the vector of values of $J^k\bm{G}^k_{ij}$ at volume and face quadrature points, and $\bm{n}^k_i, \bm{J}^k_f$ denote the $i$th outward unit normal and $J^k_f$ at face quadrature points.
+Then, an entropy conservative scheme in two dimensions is given as follows:
+\begin{theorem}
+\label{thm:esdg2d}
+Assume that $\bm{Q}^k\bm{1} = 0$.  Let $\bm{u}^k_N$ denote the solution on $D^k$ satisfying
+\begin{gather*}
+\sum_{k}\LRp{\bm{W}_k\td{\bm{u}^k_N}{t} + \sum_{i=1}^d \LRp{2\LRp{\bm{Q}^i_k\circ\bm{F}_S} + \bm{V}_f^T\bm{B}_k^i\bm{f}_S\LRp{{\tilde{\bm{u}}_f}^+,\tilde{\bm{u}}^k_f}}} = 0\\
+\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad \tilde{\bm{u}} = \begin{bmatrix}
+\bm{u}^k_N\\
+\tilde{\bm{u}}^k_f
+\end{bmatrix}, \qquad \tilde{\bm{u}}^k_f = \bm{u}\LRp{\bm{v}^k_f}, \qquad 
+\bm{v}^k_f = {\bm{V}_f\bm{v}\LRp{\bm{u}^k_N}},
+\end{gather*}
+where $\tilde{\bm{u}}^+$ denotes boundary values of the entropy-projected conservative variables on neighboring elements.  Then, $\bm{u}^k_N$ satisfies the discrete conservation of entropy 
+\[
+\sum_{k} \LRp{\bm{1}^T\bm{W}\td{\bm{u}^k_N}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}^i\LRp{\bm{v}_f^T\bm{f}(\tilde{\bm{u}}_f) - \psi\LRp{\tilde{\bm{u}}_f}}} = 0.
+\]
+\end{theorem}
+\begin{proof}
+The proof is a special case of the proof of Theorem 1 in \cite{chan2018discretely}.  
+\end{proof}
+
+We note that Theorem~\ref{thm:esdg2d} relies on the assumption that $\bm{Q}^k\bm{1} = 0$.
+This condition is equivalent to ensuring that the scaled geometric terms $\tilde{\bm{G}}^k_{ij}$ satisfy a discrete geometric conservation law (GCL) \cite{carpenter2014entropy, gassner2017br1, crean2018entropy, chan2018discretely}.
+For two-dimensional degree $N$ (isoparametric) mappings, the GCL is automatically satisfied.
+However, in three dimensions, the GCL is not guaranteed to be satisfied at the discrete level, due to the fact that geometric terms for isoparametric mappings are polynomials of degree higher than $N$.
+
+It is possible to ensure the satisfaction of a discrete GCL by using a sub-parametric polynomial geometric mapping.
+Let $N_{\rm geo}$ denote the degree of a polynomial geometric mapping.
+In three dimensions, the exact geometric terms for a degree $N_{\rm geo}$ polynomial mapping are polynomials of degree $(2N_{\rm geo} - 2)$ \cite{kopriva2006metric, hindenlang2012explicit, crean2018entropy}.
+If $(2N_{\rm geo}-2) \leq N$, or if $N_{\rm geo} \leq \left\lfloor\frac{N}{2}\right\rfloor + 1$, then the discrete GCL is automatically satisfied.
+
+For $N_{\rm geo} \geq \left\lfloor\frac{N}{2}\right\rfloor + 1$, modifications to the computation of geometric terms are required to ensure that the GCL is satisfied at the discrete level.
 For general SBP operators, the discrete GCL can be enforced through the solution of a local least squares problem \cite{crean2018entropy}.
 We take an alternative approach and  construct geometric terms using the approach of Kopriva \cite{kopriva2006metric}.
-This takes advantage of the fact that the collocation methods constructed in this work correspond to polynomial discretizations.
+This construction takes advantage of the fact that GLL and Gauss collocation methods correspond to polynomial discretizations.
 Kopriva's construction is based on rewriting the geometric terms as the curl of some auxiliary quantity.
-By interpolating this auxiliary quantity prior to applying the curl,
-
-To summarize, entropy stability curved meshes
+By approximating this auxiliary quantity using polynomial interpolation prior to applying the curl, the geometric terms are approximated by degree $N$ polynomials which satisfy the discrete GCL by construction.
+%We note that this approach can also be adapted to more general geometric mappings (e.g.\ rational or superparametric).
+These mappings can offer additional geometric accuracy, which can be advantageous for problems with wall boundary conditions \cite{toulorge2016optimizing, zwanenburg2017necessity}.
+
+To summarize, entropy stable Gauss collocation schemes on curved meshes require the following additional steps:
 \begin{enumerate}
-\item Construct polynomial approximations of the geometric terms using interpolation at GLL nodes \cite{kopriva2006metric}.  
-\item Interpolate the geometric terms to volume and surface quadrature points.  
-\item \note{Split form}.
+\item Construct polynomial approximations of the geometric terms at GLL nodes using (\ref{eq:curlcons}) \cite{kopriva2006metric}.  
+\item Evaluate the geometric terms at volume and surface quadrature points.  
+\item Compute normal vectors using (\ref{eq:normal}).  
 \end{enumerate}
@@
 \[
-dt = C_{\rm CFL}\frac{ h}{C_N}, \qquad h = \frac{ \nor{J^f}_{L^{\infty}}}{\nor{J^{-1}}_{L^{\infty}}},
+dt = C_{\rm CFL}\frac{ h}{C_N}, \qquad h = \frac{1}{{\nor{J^{-1}}_{L^{\infty}}}\nor{J^f}_{L^{\infty}}},
 \]
@@
 %Unless stated otherwise, all experiments utilize $C_{\rm CFL} = .5$.
+
+Numerical results in 1D are similar to those presented in \cite{chan2017discretely}.
+Thus, we focus on two and three dimensional problems and comparisons of entropy stable GLL and Gauss collocation schemes.
 
@@
 \centering
-\subfloat[Moderately warped mesh]{\includegraphics[width=.45\textwidth]{figs/warp2d_a125.png}}
+\subfloat[Lightly warped mesh]{\includegraphics[width=.475\textwidth]{figs/warp2d_a0312.png}}
 \hspace{1em}
+\subfloat[Moderately warped mesh]{\includegraphics[width=.475\textwidth]{figs/warp2d_a125.png}}\\
 \subfloat[Heavily warped mesh]{\includegraphics[width=.45\textwidth]{figs/warp2d_a25.png}}
-\caption{Examples of moderate and heavily warped meshes for $N=4, K= 16$.  }
+\caption{Lightly, moderately, and heavily warped meshes for $N=4, K= 16$.  }
 \label{fig:warp2d}
@@
 We also examine the influence of element curvature for both GLL and Gauss collocation schemes by examining $L^2$ errors on a sequence of moderately and heavily warped curvilinear meshes (see Figure~\ref{fig:warp2d}).
-The following results are computed using $C_{\rm CFL} = 1/2$ and Lax-Friedrichs interface dissipation.
+These warpings are constructed by modifying nodal positions according to the following mapping
+\begin{align*}
+\tilde{x} &= x + \alpha\cos\LRp{\frac{\pi}{20}(x-10)}\cos\LRp{\frac{3\pi}{10}y}\\
+\tilde{y} &= y + \alpha\sin\LRp{\frac{\pi}{5}(\tilde{x}-10)}\cos\LRp{\frac{\pi}{10}y}.
+\end{align*}
+The lightly warped mesh corresponds to $\alpha = 1/32$, the moderately warped mesh corresponds to $\alpha = 1/8$, and the heavily warped mesh corresponds to $\alpha = 1/4$.
+All results are computed using $C_{\rm CFL} = 1/2$ and Lax-Friedrichs interface dissipation.
 
@@
 %    xmin=.0125, 
-    ymin=5e-8, ymax=5,
+    ymin=2.5e-8, ymax=5,
     legend pos=south east, legend cell align=left, legend style={font=\tiny},	
@@
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,1.22861)(0.0625,0.236646)(0.03125,0.0258883)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.553435)(0.0625,0.0625441)(0.03125,0.0117156)};
-
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.414449)(0.0625,0.0412224)(0.03125,0.00293039)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.25082)(0.0625,0.0198863)(0.03125,0.000673152)};
-
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.123516)(0.0625,0.00439285)(0.03125,0.000126852)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.0768018)(0.0625,0.0020579)(0.03125,5.55322e-05)};
-
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.0332509)(0.0625,0.000970148)(0.03125,2.10334e-05)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.0151231)(0.0625,0.000454595)(0.03125,1.00203e-05)};
-
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.0106752)(0.0625,0.000133284)(0.03125,7.97032e-07)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.00657096)(0.0625,9.22845e-05)(0.03125,3.60025e-07)};
-
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.00330917)(0.0625,1.67825e-05)(0.03125,1.32696e-07)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.00185273)(0.0625,1.21465e-05)(0.03125,1.14289e-07)};
+coordinates{(0.125,1.22861)(0.0625,0.236646)(0.03125,0.0258883)(0.015625,0.00406122)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.553435)(0.0625,0.0625441)(0.03125,0.0117156)(0.015625,0.0022841)};
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.414449)(0.0625,0.0412224)(0.03125,0.00293039)(0.015625,0.000145701)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.25082)(0.0625,0.0198863)(0.03125,0.000673152)(0.015625,2.98263e-05)};
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.123516)(0.0625,0.00439285)(0.03125,0.000126852)(0.015625,5.61786e-06)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.0768018)(0.0625,0.0020579)(0.03125,5.55322e-05)(0.015625,2.36108e-06)};
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.0332509)(0.0625,0.000970148)(0.03125,2.10334e-05)(0.015625,2.60008e-07)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.0151231)(0.0625,0.000454595)(0.03125,1.00203e-05)(0.015625,6.57796e-08)};
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.0106752)(0.0625,0.000133284)(0.03125,7.97032e-07)(0.015625,NaN)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.00657096)(0.0625,9.22845e-05)(0.03125,3.60025e-07)(0.015625,NaN)};
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.00330917)(0.0625,1.67825e-05)(0.03125,1.32696e-07)(0.015625,NaN)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.00185273)(0.0625,1.21465e-05)(0.03125,1.14289e-07)(0.015625,NaN)};
 \end{loglogaxis}
@@
 }
-\subfloat[Moderately warped mesh]{
+\subfloat[Lightly warped mesh]{
 \begin{tikzpicture}
@@
 %    xmin=.0125, 
-    ymin=5e-8, ymax=5,    
+    ymin=2.5e-8, ymax=5,
     legend pos=south east, legend cell align=left, legend style={font=\tiny},	
@@
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,1.64274)(0.0625,0.445945)(0.03125,0.0720148)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.882247)(0.0625,0.138116)(0.03125,0.0268145)};
-
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.662165)(0.0625,0.146131)(0.03125,0.0234934)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.368084)(0.0625,0.0734059)(0.03125,0.00420438)};
-
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.349184)(0.0625,0.0508268)(0.03125,0.00234785)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.183472)(0.0625,0.0127617)(0.03125,0.000251929)};
-
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.171986)(0.0625,0.0103578)(0.03125,0.000386834)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.0924215)(0.0625,0.00276519)(0.03125,6.80528e-05)};
-
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.0729963)(0.0625,0.00276104)(0.03125,4.11305e-05)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.0378013)(0.0625,0.0007649)(0.03125,4.15848e-06)};
-
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.026577)(0.0625,0.000589552)(0.03125,5.3039e-06)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.0154563)(0.0625,0.000146207)(0.03125,8.83156e-07)};
+coordinates{(0.125,1.2082)(0.0625,0.265357)(0.03125,0.0301118)(0.015625,0.00467693)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.555586)(0.0625,0.0773534)(0.03125,0.0129171)(0.015625,0.0024363)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.448878)(0.0625,0.0485323)(0.03125,0.00385418)(0.015625,0.000225396)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.287685)(0.0625,0.0190431)(0.03125,0.000721055)(0.015625,3.19095e-05)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.122081)(0.0625,0.00694351)(0.03125,0.000224524)(0.015625,1.14555e-05)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.0731583)(0.0625,0.00213914)(0.03125,6.32128e-05)(0.015625,2.73037e-06)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.0424351)(0.0625,0.00114569)(0.03125,3.59697e-05)(0.015625,5.09643e-07)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.0215298)(0.0625,0.000523714)(0.03125,9.27521e-06)(0.015625,7.14031e-08)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.0112838)(0.0625,0.000219049)(0.03125,1.84304e-06)(0.015625,NaN)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.00809138)(0.0625,8.23415e-05)(0.03125,3.95565e-07)(0.015625,NaN)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.00388316)(0.0625,2.55074e-05)(0.03125,2.68102e-07)(0.015625,NaN)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.00222029)(0.0625,1.33111e-05)(0.03125,9.72405e-08)(0.015625,NaN)};
 
@@
 \end{tikzpicture}
-}\\
+}
+\\
+\subfloat[Moderately warped mesh]{
+\begin{tikzpicture}
+\begin{loglogaxis}[
+    width=.5\textwidth,
+    xlabel={Mesh size $h$},
+    ylabel={$L^2$ errors}, 
+%    xmin=.0125,
+    ymin=2.5e-8, ymax=5,    
+    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
+    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
+    legend entries={GLL, Gauss}    
+]
+\pgfplotsset{
+cycle list={{blue, mark=*}, {red, dashed ,mark=square*}}
+}
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,1.64274)(0.0625,0.445945)(0.03125,0.0720148)(0.015625,0.0091066)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.882247)(0.0625,0.138116)(0.03125,0.0268145)(0.015625,0.0039945)};
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.662165)(0.0625,0.146131)(0.03125,0.0234934)(0.015625,0.00105368)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.368084)(0.0625,0.0734059)(0.03125,0.00420438)(0.015625,9.19378e-05)};
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.349184)(0.0625,0.0508268)(0.03125,0.00234785)(0.015625,5.36192e-05)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.183472)(0.0625,0.0127617)(0.03125,0.000251929)(0.015625,7.6013e-06)};
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.171986)(0.0625,0.0103578)(0.03125,0.000386834)(0.015625,5.76284e-06)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.0924215)(0.0625,0.00276519)(0.03125,6.80528e-05)(0.015625,3.27246e-07)};
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.0729963)(0.0625,0.00276104)(0.03125,4.11305e-05)(0.015625,2.36125e-07)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.0378013)(0.0625,0.0007649)(0.03125,4.15848e-06)(0.015625,NaN)};
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.026577)(0.0625,0.000589552)(0.03125,5.3039e-06)(0.015625,NaN)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.0154563)(0.0625,0.000146207)(0.03125,8.83156e-07)(0.015625,NaN)};
+
+\end{loglogaxis}
+\end{tikzpicture}
+}
 \subfloat[Heavily warped mesh]{
@@
 %    xmin=.0125,
-    ymin=5e-8, ymax=5,    
+    ymin=2.5e-8, ymax=5,    
     legend pos=south east, legend cell align=left, legend style={font=\tiny},	
@@
 
+Figure~\ref{fig:err2d} shows the $L^2$ errors for affine, moderately warped, and heavily warped meshes.
+For affine meshes, Gauss collocation results in a lower errors than GLL collocation at all orders.
+However, the difference between both schemes decreases as $N$ increases.
+This is not too surprising: on a Cartesian domain, the discrete $L^2$ inner product resulting from GLL quadrature converges to exact $L^2$ inner product over the space of polynomials as $N$ increases \cite{quarteroni1994introduction}.
+However, GLL and Gauss collocation differ more significantly on curved meshes.
+For both moderately and heavily warped meshes, the errors for a degree $N$ Gauss collocation scheme are nearly identical to errors for a higher order GLL collocation scheme of degree $(N+1)$.
+These results are in line with numerical experiments in \cite{parsani2016entropy}, which show that GLL collocation schemes lose one order of convergence in the $L^2$ norm on unstructured non-uniform meshes.
+Both results show that increasing quadrature accuracy significantly reduces the effect of polynomial aliasing due to curved meshes and spatially varying geometric terms.
+
+We note that $L^2$ approximation estimates on curved meshes \cite{lenoir1986optimal, bernardi1989optimal, warburton2013low} assume that the mesh size is small enough to be in the asymptotic regime (such that asymptotic error estimates hold).
+On curved meshes, this requires that the mesh is sufficiently fine to resolve both the solution and geometric mapping.
+The numerical results in Figure~\ref{fig:err2d} suggest that, compared to Gauss collocation schemes, under-integrated GLL collocation schemes may require a much finer mesh resolution to reach the asymptotic regime.
+
 %\note{Investigate robustness with respect to mesh distortion: GLL vs GQ errors.
@@
 As in two dimensions, we test the accuracy of the proposed scheme using an isentropic vortex solution adapted to three dimensions.
-We take the solution to be the extruded 2D vortex propagating in the $x_2$ direction, whose analytic expression is derived from \cite{williams2013nodal}
+The solution is the extruded 2D vortex propagating in the $y$ direction, with an analytic expression given in \cite{williams2013nodal}
 \begin{align*}
@@
 \]
-In this problem, we take $c_1 = c_2 = 5$, $p_0 = {1}/{\gamma}$, and $\Pi_{\max} = 0.4$.
-The problem is solved on the domain $[0,10]\times [0,20]\times [0,10]$.
+We take $c_1 = c_2 = 5$, $p_0 = {1}/{\gamma}$, and $\Pi_{\max} = 0.4$, and solve on the domain $[0,10]\times [0,20]\times [0,10]$.
 As in the 2D case, we also examine the effect of curvilinear mesh warping.
-As in \cite{chan2018discretely}, we construct a curved warping of the initial Cartesian mesh by mapping nodes on each hexahedron to warped nodal positions $(\tilde{x_1},\tilde{x_2},\tilde{x_3})$ through the transformation
+Following \cite{chan2018discretely}, we construct a curved warping of the initial Cartesian mesh by mapping nodes on each hexahedron to warped nodal positions $(\tilde{x},\tilde{y},\tilde{z})$ through the transformation
 \begin{align*}
-\tilde{\bm{x}} &= \bm{x} + \frac{1}{2}\sin\LRp{\pi \frac{(x_1-5)}{5}}\sin\LRp{2\pi \frac{(x_2-10)}{10}}\sin\LRp{\pi \frac{(x_3-5)}{5}}.
+\tilde{\bm{x}} &= \bm{x} + \sin\LRp{\pi \frac{(x-5)}{5}}\sin\LRp{2\pi \frac{(y-10)}{10}}\sin\LRp{\pi \frac{(z-5)}{5}}.
 \end{align*} 
+On curved meshes, the geometric terms are constructed using the approach of Kopriva described in Section~\ref{sec:2}.  A $C_{\rm CFL} = .75$ is used for all experiments.
+
+Figure~\ref{fig:err3d} shows the $L^2$ errors for degrees $N = 2, 3, 4$.
+As in the 2D case, Gauss collocation schemes produce smaller errors than GLL collocation schemes, especially on on curved meshes.
+The difference between the two schemes on affine meshes is slightly more pronounced than in 2D.
+The difference between the two schemes on curved meshes is qualitatively similar to the 2D case, with Gauss collocation achieving higher accuracy than GLL collocation schemes at all polynomial degrees and mesh resolutions.
+%, which may be due to the specific curvilinear warping applied in this work.
+
+\begin{figure}
+\centering
+\subfloat[Affine mesh]{
+\begin{tikzpicture}
+\begin{loglogaxis}[
+    width=.5\textwidth,
+    xlabel={Mesh size $h$},
+    ylabel={$L^2$ errors}, 
+%    xmin=.0125, 
+    ymin=1e-6, ymax=5,
+    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
+    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
+    legend entries={GLL, Gauss}    
+]
+\pgfplotsset{
+cycle list={{blue, mark=*}, {red, dashed ,mark=square*}}
+}
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,2.62759)(0.125,0.728523)(0.0625,0.112225)(0.03125,0.018711)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,1.53203)(0.125,0.23283)(0.0625,0.0417033)(0.03125,0.00806707)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,0.934758)(0.125,0.191949)(0.0625,0.0119788)(0.03125,0.000558681)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,0.562488)(0.125,0.0864096)(0.0625,0.00268061)(0.03125,0.000136725)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,0.460425)(0.125,0.02445)(0.0625,0.000821249)(0.03125,4.67491e-05)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,0.239027)(0.125,0.0077433)(0.0625,0.000219705)(0.03125,2.67589e-05)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,0.138105)(0.125,0.00744055)(0.0625,0.000126057)(0.03125,NaN)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,0.0547127)(0.125,0.00242675)(0.0625,4.2962e-05)(0.03125,NaN)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,0.0489719)(0.125,0.000897467)(0.0625,2.55644e-05)(0.03125,NaN)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,0.0246519)(0.125,0.000324802)(0.0625,2.44009e-05)(0.03125,NaN)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,0.023319)(0.125,0.00021951)(0.0625,NaN)(0.03125,NaN)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,0.00801774)(0.125,7.39203e-05)(0.0625,NaN)(0.03125,NaN)};
+
+\end{loglogaxis}
+\end{tikzpicture}
+}
+%\subfloat[Sub-parametric mesh]{
+%\begin{tikzpicture}
+%\begin{loglogaxis}[
+%    width=.5\textwidth,
+%    xlabel={Mesh size $h$},
+%%    ylabel={$L^2$ errors}, 
+%%    xmin=.0125, 
+%    ymin=1e-6, ymax=5,
+%      legend pos=south east, legend cell align=left, legend style={font=\tiny},	
+%    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
+%    legend entries={GLL, Gauss}    
+%]
+%\pgfplotsset{
+%cycle list={{blue, mark=*}, {red, dashed ,mark=square*}}
+%}
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(0.25,2.80762)(0.125,0.951269)(0.0625,0.173495)(0.03125,0.0295367)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(0.25,1.89337)(0.125,0.366475)(0.0625,0.0599646)(0.03125,0.0133605)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(0.25,1.21946)(0.125,0.257914)(0.0625,0.0494569)(0.03125,0.00380144)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(0.25,0.708125)(0.125,0.139733)(0.0625,0.015052)(0.03125,0.000419943)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(0.25,0.594597)(0.125,0.100782)(0.0625,0.0051111)(0.03125,0.000186621)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(0.25,0.347316)(0.125,0.0351013)(0.0625,0.000702893)(0.03125,4.39052e-05)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(0.25,0.306622)(0.125,0.0198205)(0.0625,0.00156957)(0.03125,NaN)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(0.25,0.142279)(0.125,0.00620177)(0.0625,0.000311399)(0.03125,NaN)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(0.25,0.130865)(0.125,0.0108333)(0.0625,0.000188656)(0.03125,NaN)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(0.25,0.0518407)(0.125,0.00289859)(0.0625,3.71549e-05)(0.03125,NaN)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(0.25,0.0558827)(0.125,0.00251998)(0.0625,NaN)(0.03125,NaN)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(0.25,0.0208791)(0.125,0.000423102)(0.0625,NaN)(0.03125,NaN)};
+%\end{loglogaxis}
+%\end{tikzpicture}
+%}\\
+\subfloat[Curved mesh]{
+\begin{tikzpicture}
+\begin{loglogaxis}[
+    width=.5\textwidth,
+    xlabel={Mesh size $h$},
+    ylabel={$L^2$ errors}, 
+%    xmin=.0125, 
+    ymin=1e-6, ymax=5,
+    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
+    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
+    legend entries={GLL, Gauss}    
+]
+\pgfplotsset{
+cycle list={{blue, mark=*}, {red, dashed ,mark=square*}}
+}
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,2.93541)(0.125,1.00445)(0.0625,0.18488)(0.03125,0.0323814)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,1.99429)(0.125,0.423426)(0.0625,0.0787544)(0.03125,0.0173475)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,1.23231)(0.125,0.266638)(0.0625,0.0513328)(0.03125,0.00390276)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,0.727022)(0.125,0.148442)(0.0625,0.0158392)(0.03125,0.000446109)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,0.599851)(0.125,0.101098)(0.0625,0.00511836)(0.03125,0.00018702)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,0.351163)(0.125,0.0353206)(0.0625,0.000705351)(0.03125,4.41811e-05)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,0.310649)(0.125,0.0199095)(0.0625,0.00157312)(0.03125,NaN)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,0.144996)(0.125,0.00619848)(0.0625,0.000311659)(0.03125,NaN)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,0.130579)(0.125,0.0108408)(0.0625,0.000188745)(0.03125,NaN)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,0.0517392)(0.125,0.00290387)(0.0625,3.71606e-05)(0.03125,NaN)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,0.0558635)(0.125,0.00251764)(0.0625,NaN)(0.03125,NaN)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.25,0.020886)(0.125,0.000421991)(0.0625,NaN)(0.03125,NaN)};
+\end{loglogaxis}
+\end{tikzpicture}
+}
+\caption{$L^2$ errors for the 3D isentropic vortex for $N = 2,\ldots,7$ on sequences of Cartesian and curved meshes.}
+\label{fig:err3d}
+\end{figure}
+
+We also compared $L^2$ errors for both isoparametric and sub-parametric geometric mappings.
+For sub-parametric mappings, we chose the degree of approximation of geometry $N_{\rm geo} = \left\lfloor\frac{N}{2} \right\rfloor + 1$, such that the geometric terms are computed exactly and the GCL is satisfied by default.
+This test is intended to address the fact that the GCL-preserving interpolation of Kopriva introduces a small approximation error, as the geometric terms are no longer exact.
+For these sub-parametric mappings, the gap between GLL and Gauss collocation widens slightly at $N=2$.
+However, the results for sub-parametric mappings are nearly identical to the results in the isoparametric case for higher polynomial degrees.
 
 \subsection{Shock-vortex interaction}
+
+\note{describe shock-vortex problem}
+
+We modify the shock-vortex interaction problem such that wall boundary conditions are imposed at top and bottom boundaries.  For left and right boundaries, we impose periodic boundary conditions.
+
+Experiments using the matrix dissipation flux \cite{winters2017uniquely}.
+
+We note that entropy stable wall boundary conditions can also be imposed when using decoupled SBP operators with over-integration as well \cite{chan2017discretely}.
+As far as the authors know, the stable and high order accurate imposition of such boundary conditions for existing GSBP couplings \cite{crean2017high} remains an open problem.
 
@@
 The Taylor--Green vortex is used to study the transition and decay of turbulence \cite{debonis2013solutions}.
-Under an inviscid assumption, the Taylor--Green vortex develops successively smaller scales.
+In the absence of viscosity, the Taylor--Green vortex develops successively smaller scales as time increases.
 As a result, the solution is guaranteed to contain under-resolved features after a sufficiently large time.
@@
 
-\note{blah}
+\note{Give computational comparison: two point flux evaluations, FLOPS due to matrix operations. }
 
@@
 
-\section{Decoupled SBP operators}
+\section{Decoupled SBP operators for general choices of quadrature and basis}
 \label{app:decoupled}
@@
 %\]
-
 It is worth noting that the form of $\bm{Q}_N^i$ does not depend on the choice of basis.
@@
 
-Assuming that the solution is represented using degree $N$ Lagrange polynomials at $(N+1)$ quadrature points, these matrices reduce down to
+A collocation scheme assumes that the number of quadrature points is identical to the number of basis functions.
+If the solution is represented using degree $N$ Lagrange polynomials at quadrature points, the matrices $\bm{V}_q, \bm{P}_q$ simplify
 \[
@@
 \]
-
-\note{Todo: add.}
+In summary, the collocation assumption removes the need to interpolate to and from a polynomial space to volume quadrature points within an entropy stable scheme.
+
+%\note{Todo: add.}
 

### 2018-08-05 docs/ESDG_GQ/esdg_gq.tex
updating ESDG with EC shock vortex

@@
 
-This paper focuses on the construction of efficient semi-discretely entropy stable collocation schemes based on Gauss nodes, using the framework introduced in \cite{chan2017discretely, chan2018discretely}.
-While entropy stable collocation schemes have been constructed on Gauss-like quadrature points without boundary nodes \cite{crean2017high}, the inter-element coupling terms for such schemes introduce significantly more communication and computation than collocation points which contain boundary nodes.
-This work introduces new inter-element coupling terms which avoid this all-to-all coupling, requiring only communication of face values between neighboring elements.
+While entropy stable collocation schemes have been constructed on Gauss-like quadrature points without boundary nodes \cite{crean2017high}, the inter-element coupling terms for such schemes introduce an ``all-to-all'' coupling between degrees of freedom on two neighboring elements.
+These coupling terms require evaluating two-point fluxes between solution states at all collocation nodes on two neighboring elements, resulting in significantly more communication and computation compared to collocation schemes based on point sets containing boundary nodes.
+This work introduces efficient and entropy stable inter-element coupling terms for Gauss collocation schemes which require only communication of face values between neighboring elements.
+The construction of these terms follows the framework introduced in \cite{chan2017discretely, chan2018discretely} for triangles and tetrahedra.
+
+The main motivation for exploring tensor product (quadrilateral and hexahedral) elements is the significant reduction in the number of operations required compared to high order entropy stable schemes on simplicial meshes \cite{chan2017discretely, chan2018discretely}.
+Entropy stability schemes on simplicial elements require evaluating two-point fluxes between solution states at all quadrature points on an element.
+The number of quadrature points on a simplex scales as $O(N^d)$ in $d$ dimensions, and results in $O(N^{2d})$ two-point flux evaluations per element.
+In contrast, entropy stable schemes on quadrilateral and hexahedral elements require only the evaluation of two-point fluxes along lines of nodes due to a tensor product structure, resulting in $O(N^{d+1})$ evaluations in $d$ dimensions.
 
@@
 
-(see Figure~\ref{fig:twoelem})
-\note{Add pics of two-element mesh}
+%(see Figure~\ref{fig:twoelem})
+%\note{Add pics of two-element mesh}
 
@@
 
-\section{Extension to higher dimensions and curved meshes}
+\section{Extension to higher dimensions and non-affine meshes}
 \label{sec:2}
@@
 Let the domain now be decomposed into non-overlapping elements $D^k$, such that $D^k$ is the image of $\hat{D}$ under a degree $N$ polynomial mapping $\bm{\Phi}^k$.
-We define geometric terms $\bm{G}^k_{ij}$ as scaled derivatives of reference coordinates $\hat{\bm{x}}$ with respect to physical coordinates $\bm{x}$
-\[
-J^k\bm{G}^k_{ij} = J^k\pd{\hat{x}_j}{\bm{x}_j},
+We define geometric terms ${G}^k_{ij}$ as scaled derivatives of reference coordinates $\hat{\bm{x}}$ with respect to physical coordinates $\bm{x}$
+\[
+{G}^k_{ij} = J^k\pd{\hat{x}_j}{\bm{x}_j},
 \]
@@
 We define matrices over each physical element $D^k$
-\begin{gather*}
-\bm{W}_k = \bm{W} \diag{\bm{J}^k}, \qquad \bm{B}^i_k = \bm{B}^i_N \diag{\bm{n}_i \circ\bm{J}^f }\\
-\bm{Q}^i_k = \frac{1}{2} \sum_{j=1}^d \LRp{\diag{\tilde{\bm{G}}^k_{ij}}\bm{Q}^i_N  + \bm{Q}^i_N \diag{\tilde{\bm{G}}^k_{ij}} },
-\end{gather*}
-where we have discretized the differentiation matrix in split form \cite{nordstrom2006conservative}.  
-Here, $\bm{J}^k$ denotes the vector of values of $J^k$ at volume quadrature points, $\tilde{\bm{G}}^k_{ij}$ denotes the vector of values of $J^k\bm{G}^k_{ij}$ at volume and face quadrature points, and $\bm{n}^k_i, \bm{J}^k_f$ denote the $i$th outward unit normal and $J^k_f$ at face quadrature points.
+\begin{gather}
+\bm{W}_k = \bm{W} \diag{\bm{J}^k}, \qquad \bm{B}^i_k = \bm{B}^i_N \diag{\bm{n}_i \circ\bm{J}^f } \nonumber\\
+\bm{Q}^i_k = \frac{1}{2} \sum_{j=1}^d \LRp{\diag{{\bm{G}}^k_{ij}}\bm{Q}^i_N  + \bm{Q}^i_N \diag{{\bm{G}}^k_{ij}} } \label{eq:splitcurv},
+\end{gather}
+where we have discretized the curved differentiation matrix in split form \cite{nordstrom2006conservative}.  
+Here, $\bm{J}^k$ denotes the vector of values of $J^k$ at volume quadrature points, ${\bm{G}}^k_{ij}$ denotes the vector of values of ${G}^k_{ij}$ at volume and face quadrature points, and $\bm{n}^k_i, \bm{J}^k_f$ denote the $i$th outward unit normal and $J^k_f$ at face quadrature points.
 Then, an entropy conservative scheme in two dimensions is given as follows:
@@
 \label{thm:esdg2d}
-Assume that $\bm{Q}^k\bm{1} = 0$.  Let $\bm{u}^k_N$ denote the solution on $D^k$ satisfying
+Assume that $\bm{Q}^k\bm{1} = 0$, and that $D^k$ are mapped curvilinear elements.  Let $\bm{u}^k_N$ denote the solution on $D^k$ satisfying
 \begin{gather*}
@@
 \end{theorem}
-\begin{proof}
-The proof is a special case of the proof of Theorem 1 in \cite{chan2018discretely}.  
-\end{proof}
+The proof is a special case of the proof of Theorem 1 in \cite{chan2018discretely}, with the quadrature rule taken to be a tensor product rule with $(N+1)$ Gauss (or GLL) points in each coordinate direction.
+
+\begin{remark}
+The operator $\bm{Q}^i_k$  in (\ref{eq:splitcurv}) can be applied without needing to explicitly store geometric terms $G^k_{ij}$ at face quadrature points.
+By using the structure of the boundary matrix $\bm{B}^i$, expressions involving geometric terms on faces can be replaced by expressions involving components of unit outward normals $\bm{n}_i$ on $D^k$.
+\end{remark}
 
 We note that Theorem~\ref{thm:esdg2d} relies on the assumption that $\bm{Q}^k\bm{1} = 0$.
-This condition is equivalent to ensuring that the scaled geometric terms $\tilde{\bm{G}}^k_{ij}$ satisfy a discrete geometric conservation law (GCL) \cite{carpenter2014entropy, gassner2017br1, crean2018entropy, chan2018discretely}.
+This condition is equivalent to ensuring that the scaled geometric terms ${\bm{G}}^k_{ij}$ satisfy a discrete geometric conservation law (GCL) \cite{carpenter2014entropy, gassner2017br1, crean2018entropy, chan2018discretely}.
 For two-dimensional degree $N$ (isoparametric) mappings, the GCL is automatically satisfied.
@@
 This construction takes advantage of the fact that GLL and Gauss collocation methods correspond to polynomial discretizations.
-Kopriva's construction is based on rewriting the geometric terms as the curl of some auxiliary quantity.
-By approximating this auxiliary quantity using polynomial interpolation prior to applying the curl, the geometric terms are approximated by degree $N$ polynomials which satisfy the discrete GCL by construction.
+Kopriva's construction is based on rewriting the geometric terms as the reference curl of an interpolated auxiliary quantity
+\begin{equation}
+\label{eq:conscurl}
+\begin{bmatrix}
+\bm{G}^k_{1j}\\
+\bm{G}^k_{2j}\\
+\bm{G}^k_{3j}
+\end{bmatrix} = \begin{bmatrix}
+\LRp{-\hat{\Grad} \times I_N\LRp{z\hat{\Grad}y}}_j\\
+\LRp{\hat{\Grad} \times I_N\LRp{z\hat{\Grad}x}}_j\\
+\LRp{\hat{\Grad} \times I_N\LRp{x\hat{\Grad}y}}_j
+\end{bmatrix}, \qquad j = 1,2,3.
+\end{equation}
+Here, $I_N$ denotes the polynomial interpolation operator using GLL nodes.
+By interpolating the auxiliary quantity in (\ref{eq:conscurl}) using polynomial interpolation prior to applying the curl, the geometric terms are approximated by degree $N$ polynomials which satisfy the discrete GCL by construction.
+These GCL-satisfying geometric terms can then be used to compute normal vectors.
+For a watertight mesh, the constructed normal vectors are guaranteed to be continuous across faces \cite{chan2018discretely}.
 %We note that this approach can also be adapted to more general geometric mappings (e.g.\ rational or superparametric).
@@
 
-To summarize, entropy stable Gauss collocation schemes on curved meshes require the following additional steps:
+To summarize, extending higher dimensional entropy stable Gauss collocation schemes to curved meshes requires the following steps: 
 \begin{enumerate}
-\item Construct polynomial approximations of the geometric terms at GLL nodes using (\ref{eq:curlcons}) \cite{kopriva2006metric}.  
-\item Evaluate the geometric terms at volume and surface quadrature points.  
-\item Compute normal vectors using (\ref{eq:normal}).  
+\item Construct polynomial approximations of the geometric terms using equation (\ref{eq:conscurl}) and interpolation at GLL nodes \cite{kopriva2006metric}.  
+\item Evaluate approximate geometric terms at volume and surface points, and compute normal vectors in terms of the approximate geometric terms.  
+\item Compute physical derivatives using the split form (\ref{eq:splitcurv}).  
 \end{enumerate}
+Apart from evaluating the GCL-satisfying geometric terms at separate volume and surface points, entropy stable Gauss collocation schemes are extended to curved meshes in the same manner as GLL and staggered-grid collocation schemes \cite{carpenter2014entropy, parsani2016entropy, chan2018discretely}.
 
@@
 where we have defined the auxiliary quantities 
-\begin{gather*}
-p_{\rm avg} = \frac{\avg{\rho}}{2\avg{\beta}}, \qquad E_{\rm avg} = \frac{\avg{\rho}^{\log}}{2\avg{\beta}^{\log}\LRp{\gamma -1}}   + \frac{\nor{\bm{u}}^2_{\rm avg}}{2}, \\
+\begin{gather}
+p_{\rm avg} = \frac{\avg{\rho}}{2\avg{\beta}}, \qquad E_{\rm avg} = \frac{\avg{\rho}^{\log}}{2\avg{\beta}^{\log}\LRp{\gamma -1}}   + \frac{\nor{\bm{u}}^2_{\rm avg}}{2}, \label{eq:fluxaux} \\
  \nor{\bm{u}}^2_{\rm avg} = 2(\avg{u}^2 + \avg{v}^2) - \LRp{\avg{u^2} +\avg{v^2}} \nonumber.  
-\end{gather*}
+\end{gather}
 Expressions for entropy conservative numerical fluxes for the three-dimensional compressible Euler equations can also be explicitly written as
@@
 %    xmin=.0125, 
-    ymin=1e-6, ymax=5,
-    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
+    ymin=1e-5, ymax=5,
+    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
     xmajorgrids=true, ymajorgrids=true, grid style=dashed,
@@
 
-\note{describe shock-vortex problem}
-
-We modify the shock-vortex interaction problem such that wall boundary conditions are imposed at top and bottom boundaries.  For left and right boundaries, we impose periodic boundary conditions.
-
-Experiments using the matrix dissipation flux \cite{winters2017uniquely}.
-
-We note that entropy stable wall boundary conditions can also be imposed when using decoupled SBP operators with over-integration as well \cite{chan2017discretely}.
-As far as the authors know, the stable and high order accurate imposition of such boundary conditions for existing GSBP couplings \cite{crean2017high} remains an open problem.
+The next problem considered is the shock-vortex interaction described in \cite{shu1998essentially}.
+The domain is taken to be $[0,2]\times [0,1]$, and is triangulated with uniform quadrilateral elements.
+Wall boundary conditions are imposed on the top and bottom boundaries, and inflow boundary conditions are typically imposed on the left and right boundaries.
+We modify the problem setup such that periodic boundary conditions are imposed at the left and right boundaries.
+Wall boundary conditions are imposed using a mirror state for the normal velocity, which was shown to be entropy stable in \cite{svard2014entropy, chen2017entropy}.
+
+The initial condition is taken to be the superposition of a stationary shock and a vortex propagating towards the right.
+The stationary Mach $M_s = 1.1$ shock is positioned at $x = .5$ normal to the $x$ axis, with left state $\LRp{\rho_L, u_L, v_L, p_L} = \LRp{1, \sqrt{\gamma}, 0, 1}$.
+The right state is a scaling of the left state computed using the Rankine-Hugoniot conditions, such that the ratio of upstream and downstream states is
+%$\LRp{\rho_R, u_R, v_R, p_R} = \LRp{\rho_L/ \alpha_\rho, u_L / \alpha_u, 0, p_L/\alpha_p}$, where
+\begin{align*}
+\frac{\rho_L}{\rho_R} = \frac{u_L}{u_R} = \frac{2+ (\gamma-1) M_s^2}{(\gamma+1)M_s^2}, \qquad \frac{p_L}{p_R} = 1+ \frac{2\gamma}{\gamma+1}\LRp{M_s^2-1}, \qquad v_R = 0.
+\end{align*}
+The isentropic vortex is centered at $(x_c,y_c) = (.25, .5)$ and given in terms of velocity fluctuations $\delta u$ and $\delta v$, which are functions of the tangential velocity $v_{\theta}$
+\[
+\delta u = v_{\theta} \sin(\theta), \qquad
+\delta v = -v_{\theta} \cos(\theta), \qquad
+v_{\theta} = \epsilon \tau e^{\alpha(1-\tau^2)},
+%u = \epsilon \tau e^{\alpha(1-\tau^2)}\sin(\theta), \qquad
+%v = -\epsilon \tau e^{\alpha(1-\tau^2)}\cos(\theta), \qquad
+%\delta T = -\frac{(\gamma-1)\epsilon^2e^{2\alpha(1-\tau^2)}}{4\alpha\gamma}.
+\]
+where $r = \sqrt{(x-x_c)^2 + (y-y_c)^2}$ is the radius from the vortex center, $\tau = r/r_c$, and $\theta = \tan^{-1}\LRp{\frac{y-y_c}{x-x_c}}$.
+We follow \cite{shu1998essentially} and take $\epsilon = .3$, $\alpha = .204$, and $r_c = .05$.
+The vortex temperature is computed as a fluctuation $\delta T$ of the upstream state $T_L = p_L/\rho_L$
+\[
+\delta T =  - \frac{(\gamma-1)\epsilon^2 e^{2\alpha (1-\tau^2)}}{4\alpha\gamma}.  
+\]
+The vortex density and pressure are computed using an isentropic assumption.  To summarize, the initial condition for the shock-vortex interaction problem is 
+\begin{gather*}
+\rho = \rho_s \LRp{\frac{T_{\rm vor}}{T_L}}^{\frac{1}{\gamma-1}}, \qquad u = u_s + \delta u, \qquad v = v_s + \delta v, \qquad p = p_s \LRp{\frac{T_{\rm vor}}{T_L}}^{\frac{1}{\gamma-1}},
+\end{gather*}
+where $(\rho_s,  u_s, v_s, p_s)$ denote the discontinuous stationary shock solution given by the left and right states $\LRp{\rho_L, u_L, v_L, p_L}, \LRp{\rho_R, u_R, v_R, p_R}$.
+
+\begin{figure}
+\centering
+\subfloat[Entropy conservative flux, $T = .3$]{\includegraphics[width=.495\textwidth]{figs/shockVortexTp3_EC.png}}
+\hspace{.05em}
+\subfloat[Entropy conservative flux, $T = .7$]{\includegraphics[width=.495\textwidth]{figs/shockVortexTp7_EC.png}}\\
+\subfloat[Lax-Friedrichs flux, $T = .3$]{\includegraphics[width=.495\textwidth]{figs/shockVortexTp3_LF.png}}
+\hspace{.05em}
+\subfloat[Lax-Friedrichs flux, $T = .7$]{\includegraphics[width=.495\textwidth]{figs/shockVortexTp7_LF.png}}\\
+\subfloat[Matrix dissipation flux, $T = .3$]{\includegraphics[width=.495\textwidth]{figs/shockVortexTp3.png}}
+\hspace{.05em}
+\subfloat[Matrix dissipation flux, $T = .7$]{\includegraphics[width=.495\textwidth]{figs/shockVortexTp7.png}}
+\caption{Shock vortex solution at time $T = .7$ using entropy stable Gauss collocation schemes with $N=4, h = 1/100$.}
+\label{fig:shockvort}
+\end{figure}
+
+We compare three different entropy stable Gauss collocation schemes.
+All three utilize the entropy conservative flux of Chandrashekar \cite{chandrashekar2013kinetic}.
+For the first scheme, we do not introduce any additional interface dissipation, which produces an entropy conservative scheme.
+The second scheme introduces an entropy-dissipative interface term using Lax-Friedrichs penalization, while the third scheme utilizes the matrix dissipation flux introduced in \cite{winters2017uniquely}.
+This flux adds a dissipation of the form $\bm{R}\bm{D}\bm{R}^T\jump{{\bm{v}}^k_f}$, where $\jump{{\bm{v}}^k_f}$ denotes the jump in the entropy variables.
+In two dimensions, the matrices $\bm{R}, \bm{D}$ are
+\begin{align*}
+\bm{R} &= \begin{bmatrix}
+1 & 1 & 0 & 1\\
+\avg{u}-\bar{a} n_x & \avg{u} & n_y & \avg{u}+\bar{a} n_x\\
+\avg{v}-\bar{a} n_y & \avg{v} & -n_x & \avg{v}+\bar{a} n_y\\
+h-\bar{a} \bar{u}_n & \frac{1}{2}\nor{\bm{u}}^2_{\rm avg} & \bar{u}_n n_y - \avg{v}n_x & h + \bar{a} \bar{u}_n
+\end{bmatrix}\\
+\bm{D} &= {\rm diag}{\begin{pmatrix}
+\LRb{\bar{u}_n - a}\frac{\avg{\rho}^{\log}}{2\gamma}, & \LRb{\bar{u}_n}\frac{\avg{\rho}^{\log}(\gamma-1)}{\gamma}, & \LRb{\bar{u}_n}p_{\rm avg}, & \LRb{\bar{u}_n + a}\frac{\avg{\rho}^{\log}}{2\gamma} \end{pmatrix}},
+\end{align*}
+where $\bar{a}, \bar{u}_n$ are defined as
+\begin{gather*}
+\bar{u}_n = \avg{u}n_x + \avg{v}n_y, \qquad \bar{a} = \sqrt{\frac{\gamma p_{\rm avg}}{\avg{\rho}^{\log}}}, \qquad h = \frac{\gamma}{2(\gamma-1)\avg{\beta}^{\log}} + \frac{1}{2}\bar{u}, 
+\end{gather*}
+and $p_{\rm avg}, \nor{\bm{u}}^2_{\rm avg}$ are defined as in (\ref{eq:fluxaux}).
+
+%It is known that the Lax-Friedrichs flux is fairly dissipative compared to more nuanced fluxes \cite{harten1983upstream}, though this dissipation is less pronounced at higher polynomial degrees in practice.
+Figure~\ref{fig:shockvort} shows solutions for with $N=4$ and $h = 1/100$ Gauss collocation schemes using a non-dissipative entropy conservative flux, a dissipative Lax-Friedrichs flux, and a matrix dissipation flux.
+In all cases, the vortex passes through the shock stably without the use of additional slope limiting, filtering, or artificial viscosity.
+However, the entropy conservative scheme produces a large number of spurious oscillations in the solution.
+These are reduced away from the shock under Lax-Friedrichs dissipation, though oscillations still persist around a large neighborhood of the discontinuity.
+The Gibbs-type oscillations are most localized under the matrix dissipation flux.
+
+We note that this experiment also verifies that entropy stable decoupled SBP schemes (including the over-integrated case \cite{chan2017discretely}) are compatible with entropy stable wall boundary conditions.
+As far as the authors know, the stable and high order accurate imposition of such boundary conditions for existing GSBP couplings described in \cite{crean2017high} and Section~\ref{sec:gsbpsat} remains an open problem.
 
@@
 
+\begin{figure}
+\centering
+\subfloat[GLL collocation]{
+\begin{tikzpicture}
+\begin{axis}[
+    width=.525\textwidth,
+    xlabel={Time $t$},
+    ylabel={$-\pd{\kappa}{t}$}, 
+%    xmin=.0125, xmax=.75,
+    ymin=-.001, ymax=.016,
+    legend pos=north east, legend cell align=left, legend style={font=\tiny},	
+    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
+    legend entries={Affine, Curved}    
+]
+\pgfplotsset{
+cycle list={{blue}, {red, dashed}}
+}
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0,-0)(0.1,-0)(0.2,-1.1e-05)(0.3,-1e-05)(0.4,-1.4e-05)(0.5,-1.6e-05)(0.6,-2.5e-05)(0.7,-1.5e-05)(0.8,-2.8e-05)(0.9,-2.4e-05)(1,-2.6e-05)(1.1,-2.4e-05)(1.2,-3.2e-05)(1.3,-2.1e-05)(1.4,-2.5e-05)(1.5,-2.6e-05)(1.6,-2.1e-05)(1.7,-1.9e-05)(1.8,-2.6e-05)(1.9,-1.1e-05)(2,-2.5e-05)(2.1,-8e-06)(2.2,-2.5e-05)(2.3,-3e-06)(2.4,-2.3e-05)(2.5,-3e-06)(2.6,-1.7e-05)(2.7,-5e-06)(2.8,-1.2e-05)(2.9,-8e-06)(3,1e-06)(3.1,-1.4e-05)(3.2,1.7e-05)(3.3,2e-06)(3.4,3.8e-05)(3.5,5.3e-05)(3.6,0.000105)(3.7,0.000139)(3.8,0.000221)(3.9,0.000276)(4,0.000342)(4.1,0.000419)(4.2,0.000521)(4.3,0.000678)(4.4,0.000837)(4.5,0.000981)(4.6,0.001062)(4.7,0.001087)(4.8,0.00113)(4.9,0.001166)(5,0.001202)(5.1,0.001253)(5.2,0.001359)(5.3,0.001478)(5.4,0.001732)(5.5,0.002089)(5.6,0.002263)(5.7,0.002498)(5.8,0.002994)(5.9,0.003125)(6,0.003437)(6.1,0.004011)(6.2,0.004722)(6.3,0.005376)(6.4,0.006272)(6.5,0.007188)(6.6,0.00777)(6.7,0.007294)(6.8,0.008198)(6.9,0.009458)(7,0.009693)(7.1,0.011262)(7.2,0.010684)(7.3,0.011183)(7.4,0.012658)(7.5,0.013435)(7.6,0.012556)(7.7,0.012895)(7.8,0.012983)(7.9,0.011737)(8,0.011435)(8.1,0.01177)(8.2,0.011534)(8.3,0.011911)(8.4,0.011895)(8.5,0.011791)(8.6,0.011377)(8.7,0.009956)(8.8,0.010736)(8.9,0.011693)(9,0.009853)(9.1,0.009799)(9.2,0.009839)(9.3,0.010534)(9.4,0.010571)(9.5,0.009947)(9.6,0.010718)(9.7,0.011076)(9.8,0.01093)(9.9,0.011329)(10,0.011703)(10.1,0.01164)(10.2,0.012807)(10.3,0.012246)(10.4,0.010467)(10.5,0.010364)(10.6,0.010194)(10.7,0.008977)(10.8,0.009781)(10.9,0.010854)(11,0.010576)(11.1,0.009041)(11.2,0.008563)(11.3,0.007785)(11.4,0.007869)(11.5,0.008274)(11.6,0.008207)(11.7,0.007814)(11.8,0.007557)(11.9,0.00751)(12,0.0082)(12.1,0.007652)(12.2,0.007354)(12.3,0.007167)(12.4,0.007458)(12.5,0.007237)(12.6,0.007162)(12.7,0.006955)(12.8,0.006604)(12.9,0.006469)(13,0.006581)(13.1,0.006469)(13.2,0.00626)(13.3,0.006146)(13.4,0.006075)(13.5,0.00601)(13.6,0.005883)(13.7,0.005825)(13.8,0.005701)(13.9,0.005585)(14,0.005407)(14.1,0.005251)(14.2,0.005187)(14.3,0.005117)(14.4,0.005088)(14.5,0.00498)(14.6,0.004898)(14.7,0.004772)(14.8,0.004717)(14.9,0.004651)(15,0.004603)(15.1,0.004514)(15.2,0.004408)(15.3,0.004296)(15.4,0.004199)(15.5,0.004117)(15.6,0.004069)(15.7,0.003965)(15.8,0.003854)(15.9,0.003777)(16,0.003716)(16.1,0.003633)(16.2,0.003617)(16.3,0.003594)(16.4,0.003547)(16.5,0.003478)(16.6,0.003385)(16.7,0.003317)(16.8,0.003279)(16.9,0.003207)(17,0.003162)(17.1,0.003144)(17.2,0.003097)(17.3,0.00307)(17.4,0.003008)(17.5,0.002939)(17.6,0.00292)(17.7,0.002873)(17.8,0.002805)(17.9,0.002742)(18,0.00271)(18.1,0.00266)(18.2,0.002629)(18.3,0.002588)(18.4,0.002547)(18.5,0.002515)(18.6,0.002468)(18.7,0.002415)(18.8,0.002386)(18.9,0.002366)(19,0.002338)(19.1,0.002301)(19.2,0.002265)(19.3,0.002243)(19.4,0.002204)(19.5,0.002174)(19.6,0.002139)(19.7,0.002107)(19.8,0.002075)(19.9,0.002052)(20,0.00202)};
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0,-0)(0.1,-0)(0.2,-1.1e-05)(0.3,-1e-05)(0.4,-1.3e-05)(0.5,-1.6e-05)(0.6,-2.5e-05)(0.7,-1.6e-05)(0.8,-2.8e-05)(0.9,-2.5e-05)(1,-2.5e-05)(1.1,-2.4e-05)(1.2,-3.1e-05)(1.3,-2.2e-05)(1.4,-2.4e-05)(1.5,-2.7e-05)(1.6,-2.1e-05)(1.7,-1.9e-05)(1.8,-2.6e-05)(1.9,-1.2e-05)(2,-2.5e-05)(2.1,-8e-06)(2.2,-2.6e-05)(2.3,-2e-06)(2.4,-2.4e-05)(2.5,-3e-06)(2.6,-1.8e-05)(2.7,-4e-06)(2.8,-1.5e-05)(2.9,-5e-06)(3,-3e-06)(3.1,-1.4e-05)(3.2,1.1e-05)(3.3,-5e-06)(3.4,2.5e-05)(3.5,3.2e-05)(3.6,7.1e-05)(3.7,9e-05)(3.8,0.000156)(3.9,0.000201)(4,0.000279)(4.1,0.000375)(4.2,0.00051)(4.3,0.000691)(4.4,0.000864)(4.5,0.001022)(4.6,0.00114)(4.7,0.001224)(4.8,0.001302)(4.9,0.001404)(5,0.001526)(5.1,0.001685)(5.2,0.001869)(5.3,0.002032)(5.4,0.002409)(5.5,0.002752)(5.6,0.003089)(5.7,0.003588)(5.8,0.004093)(5.9,0.004445)(6,0.004825)(6.1,0.005375)(6.2,0.005868)(6.3,0.006183)(6.4,0.006331)(6.5,0.006797)(6.6,0.007712)(6.7,0.008319)(6.8,0.00896)(6.9,0.009326)(7,0.009777)(7.1,0.010393)(7.2,0.010403)(7.3,0.010421)(7.4,0.010724)(7.5,0.011246)(7.6,0.011484)(7.7,0.011616)(7.8,0.012531)(7.9,0.012689)(8,0.012545)(8.1,0.012473)(8.2,0.012667)(8.3,0.012824)(8.4,0.012941)(8.5,0.0127)(8.6,0.012743)(8.7,0.012718)(8.8,0.012434)(8.9,0.011995)(9,0.011842)(9.1,0.01197)(9.2,0.011787)(9.3,0.011592)(9.4,0.011345)(9.5,0.010925)(9.6,0.011148)(9.7,0.01112)(9.8,0.010876)(9.9,0.010962)(10,0.010975)(10.1,0.010899)(10.2,0.010607)(10.3,0.010373)(10.4,0.01008)(10.5,0.009831)(10.6,0.009418)(10.7,0.009322)(10.8,0.009141)(10.9,0.009067)(11,0.009091)(11.1,0.008885)(11.2,0.008732)(11.3,0.008681)(11.4,0.008587)(11.5,0.008343)(11.6,0.008185)(11.7,0.008042)(11.8,0.008009)(11.9,0.007849)(12,0.007513)(12.1,0.007208)(12.2,0.007146)(12.3,0.007062)(12.4,0.007009)(12.5,0.006919)(12.6,0.006761)(12.7,0.006594)(12.8,0.006558)(12.9,0.00642)(13,0.00625)(13.1,0.006121)(13.2,0.006015)(13.3,0.00599)(13.4,0.00586)(13.5,0.005745)(13.6,0.005636)(13.7,0.005559)(13.8,0.005431)(13.9,0.005396)(14,0.005265)(14.1,0.0052)(14.2,0.005097)(14.3,0.004995)(14.4,0.004901)(14.5,0.004814)(14.6,0.0047)(14.7,0.004622)(14.8,0.004574)(14.9,0.004471)(15,0.004354)(15.1,0.004272)(15.2,0.00426)(15.3,0.004171)(15.4,0.004091)(15.5,0.004049)(15.6,0.003979)(15.7,0.003897)(15.8,0.003842)(15.9,0.003795)(16,0.003726)(16.1,0.003677)(16.2,0.003605)(16.3,0.003528)(16.4,0.003477)(16.5,0.003454)(16.6,0.003389)(16.7,0.003346)(16.8,0.003298)(16.9,0.003229)(17,0.003182)(17.1,0.003154)(17.2,0.003085)(17.3,0.003019)(17.4,0.002948)(17.5,0.002911)(17.6,0.002874)(17.7,0.002836)(17.8,0.002812)(17.9,0.002746)(18,0.0027)(18.1,0.002667)(18.2,0.002635)(18.3,0.002613)(18.4,0.002582)(18.5,0.002542)(18.6,0.002495)(18.7,0.002458)(18.8,0.002435)(18.9,0.002405)(19,0.002376)(19.1,0.002342)(19.2,0.002314)(19.3,0.002278)(19.4,0.00225)(19.5,0.00221)(19.6,0.002181)(19.7,0.002155)(19.8,0.002114)(19.9,0.002085)(20,0.002057)};
+
+\end{axis}
+\end{tikzpicture}
+}
+\subfloat[Gauss collocation]{
+\begin{tikzpicture}
+\begin{axis}[
+    width=.525\textwidth,
+    xlabel={Time $t$},
+    ymin=-.001, ymax=.016,    
+    legend pos=north east, legend cell align=left, legend style={font=\tiny},	
+    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
+    legend entries={Affine, Curved}    
+]
+\pgfplotsset{
+cycle list={{blue}, {red, dashed}}
+}
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0,-0)(0.1,-0)(0.2,-1.1e-05)(0.3,-1e-05)(0.4,-1.4e-05)(0.5,-1.6e-05)(0.6,-2.5e-05)(0.7,-1.5e-05)(0.8,-2.8e-05)(0.9,-2.4e-05)(1,-2.6e-05)(1.1,-2.4e-05)(1.2,-3.2e-05)(1.3,-2.1e-05)(1.4,-2.5e-05)(1.5,-2.6e-05)(1.6,-2.1e-05)(1.7,-1.9e-05)(1.8,-2.6e-05)(1.9,-1.1e-05)(2,-2.5e-05)(2.1,-8e-06)(2.2,-2.5e-05)(2.3,-3e-06)(2.4,-2.3e-05)(2.5,-3e-06)(2.6,-1.7e-05)(2.7,-5e-06)(2.8,-1.2e-05)(2.9,-8e-06)(3,0)(3.1,-1.4e-05)(3.2,1.6e-05)(3.3,2e-06)(3.4,3.8e-05)(3.5,5.5e-05)(3.6,0.000113)(3.7,0.000157)(3.8,0.000259)(3.9,0.000347)(4,0.000462)(4.1,0.00061)(4.2,0.000794)(4.3,0.001034)(4.4,0.001274)(4.5,0.001479)(4.6,0.001593)(4.7,0.001646)(4.8,0.001711)(4.9,0.001762)(5,0.001796)(5.1,0.001797)(5.2,0.001863)(5.3,0.00193)(5.4,0.002103)(5.5,0.002362)(5.6,0.002622)(5.7,0.002965)(5.8,0.0034)(5.9,0.003864)(6,0.004146)(6.1,0.004136)(6.2,0.004996)(6.3,0.005703)(6.4,0.007004)(6.5,0.008301)(6.6,0.009652)(6.7,0.010951)(6.8,0.011407)(6.9,0.013516)(7,0.013684)(7.1,0.014113)(7.2,0.013742)(7.3,0.013059)(7.4,0.014071)(7.5,0.013728)(7.6,0.012821)(7.7,0.012544)(7.8,0.013137)(7.9,0.01424)(8,0.0126)(8.1,0.012763)(8.2,0.012715)(8.3,0.013327)(8.4,0.014155)(8.5,0.013513)(8.6,0.01272)(8.7,0.012531)(8.8,0.013325)(8.9,0.013302)(9,0.013029)(9.1,0.012178)(9.2,0.012225)(9.3,0.013543)(9.4,0.014134)(9.5,0.013937)(9.6,0.012882)(9.7,0.012053)(9.8,0.013117)(9.9,0.012565)(10,0.012059)(10.1,0.011629)(10.2,0.011937)(10.3,0.012583)(10.4,0.012349)(10.5,0.011896)(10.6,0.012309)(10.7,0.011838)(10.8,0.011069)(10.9,0.01059)(11,0.009978)(11.1,0.01026)(11.2,0.010219)(11.3,0.010047)(11.4,0.009645)(11.5,0.009094)(11.6,0.008931)(11.7,0.009278)(11.8,0.009438)(11.9,0.008792)(12,0.008361)(12.1,0.00806)(12.2,0.007456)(12.3,0.007163)(12.4,0.007304)(12.5,0.007684)(12.6,0.00785)(12.7,0.007487)(12.8,0.007027)(12.9,0.006911)(13,0.006836)(13.1,0.00682)(13.2,0.006961)(13.3,0.006948)(13.4,0.006935)(13.5,0.007177)(13.6,0.007091)(13.7,0.006754)(13.8,0.006527)(13.9,0.006203)(14,0.006094)(14.1,0.005863)(14.2,0.005841)(14.3,0.005897)(14.4,0.005707)(14.5,0.005541)(14.6,0.005396)(14.7,0.00524)(14.8,0.005027)(14.9,0.004758)(15,0.004552)(15.1,0.00432)(15.2,0.00422)(15.3,0.004206)(15.4,0.004258)(15.5,0.004133)(15.6,0.003853)(15.7,0.003705)(15.8,0.003694)(15.9,0.003643)(16,0.003556)(16.1,0.003453)(16.2,0.003301)(16.3,0.003159)(16.4,0.003127)(16.5,0.003173)(16.6,0.003171)(16.7,0.003147)(16.8,0.003069)(16.9,0.002928)(17,0.002782)(17.1,0.00268)(17.2,0.00265)(17.3,0.002667)(17.4,0.002667)(17.5,0.00264)(17.6,0.002643)(17.7,0.002728)(17.8,0.002793)(17.9,0.002758)(18,0.002614)(18.1,0.00245)(18.2,0.002369)(18.3,0.002311)(18.4,0.002249)(18.5,0.002216)(18.6,0.002239)(18.7,0.002253)(18.8,0.002197)(18.9,0.002097)(19,0.002026)(19.1,0.00198)(19.2,0.001938)(19.3,0.001898)(19.4,0.001831)(19.5,0.001766)(19.6,0.001727)(19.7,0.001723)(19.8,0.001719)(19.9,0.001707)(20,0.001698)};
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0,-0)(0.1,-0)(0.2,-1.1e-05)(0.3,-1e-05)(0.4,-1.3e-05)(0.5,-1.6e-05)(0.6,-2.5e-05)(0.7,-1.6e-05)(0.8,-2.8e-05)(0.9,-2.5e-05)(1,-2.5e-05)(1.1,-2.4e-05)(1.2,-3.1e-05)(1.3,-2.2e-05)(1.4,-2.4e-05)(1.5,-2.7e-05)(1.6,-2.1e-05)(1.7,-1.9e-05)(1.8,-2.6e-05)(1.9,-1.2e-05)(2,-2.5e-05)(2.1,-8e-06)(2.2,-2.6e-05)(2.3,-2e-06)(2.4,-2.4e-05)(2.5,-3e-06)(2.6,-1.8e-05)(2.7,-3e-06)(2.8,-1.4e-05)(2.9,-5e-06)(3,-2e-06)(3.1,-1.3e-05)(3.2,1.3e-05)(3.3,-2e-06)(3.4,3.1e-05)(3.5,4.2e-05)(3.6,8.7e-05)(3.7,0.000117)(3.8,0.000203)(3.9,0.000272)(4,0.000381)(4.1,0.000525)(4.2,0.000714)(4.3,0.000946)(4.4,0.001174)(4.5,0.001385)(4.6,0.001526)(4.7,0.001622)(4.8,0.001724)(4.9,0.001872)(5,0.002067)(5.1,0.002293)(5.2,0.002575)(5.3,0.002922)(5.4,0.003262)(5.5,0.003483)(5.6,0.003769)(5.7,0.004017)(5.8,0.004277)(5.9,0.004577)(6,0.005016)(6.1,0.005492)(6.2,0.006201)(6.3,0.006815)(6.4,0.007136)(6.5,0.008004)(6.6,0.008172)(6.7,0.008672)(6.8,0.009029)(6.9,0.009603)(7,0.010511)(7.1,0.010773)(7.2,0.01077)(7.3,0.011114)(7.4,0.011948)(7.5,0.012385)(7.6,0.012737)(7.7,0.012609)(7.8,0.013135)(7.9,0.0134)(8,0.01394)(8.1,0.014723)(8.2,0.014402)(8.3,0.01389)(8.4,0.013833)(8.5,0.014437)(8.6,0.014767)(8.7,0.014866)(8.8,0.014848)(8.9,0.014442)(9,0.014686)(9.1,0.014743)(9.2,0.014514)(9.3,0.013853)(9.4,0.013283)(9.5,0.013316)(9.6,0.013374)(9.7,0.013275)(9.8,0.013436)(9.9,0.013322)(10,0.012779)(10.1,0.01246)(10.2,0.012098)(10.3,0.011781)(10.4,0.011687)(10.5,0.011505)(10.6,0.011334)(10.7,0.011239)(10.8,0.010745)(10.9,0.010255)(11,0.010106)(11.1,0.009861)(11.2,0.009627)(11.3,0.009393)(11.4,0.009309)(11.5,0.009199)(11.6,0.00884)(11.7,0.00852)(11.8,0.008338)(11.9,0.008167)(12,0.008026)(12.1,0.007844)(12.2,0.007667)(12.3,0.007473)(12.4,0.007233)(12.5,0.007066)(12.6,0.006955)(12.7,0.00674)(12.8,0.006575)(12.9,0.006464)(13,0.006253)(13.1,0.005983)(13.2,0.005785)(13.3,0.005656)(13.4,0.005644)(13.5,0.005576)(13.6,0.005468)(13.7,0.005338)(13.8,0.005177)(13.9,0.004984)(14,0.004959)(14.1,0.004922)(14.2,0.004777)(14.3,0.004656)(14.4,0.004532)(14.5,0.004364)(14.6,0.004265)(14.7,0.004247)(14.8,0.004179)(14.9,0.004047)(15,0.003934)(15.1,0.003867)(15.2,0.003779)(15.3,0.003693)(15.4,0.003662)(15.5,0.003621)(15.6,0.003529)(15.7,0.003449)(15.8,0.003409)(15.9,0.003362)(16,0.003323)(16.1,0.003276)(16.2,0.003213)(16.3,0.003145)(16.4,0.003069)(16.5,0.003003)(16.6,0.00294)(16.7,0.002864)(16.8,0.002815)(16.9,0.002753)(17,0.002676)(17.1,0.002615)(17.2,0.002597)(17.3,0.002599)(17.4,0.002565)(17.5,0.002509)(17.6,0.002445)(17.7,0.002379)(17.8,0.002328)(17.9,0.002295)(18,0.002251)(18.1,0.002218)(18.2,0.002207)(18.3,0.002183)(18.4,0.002146)(18.5,0.002114)(18.6,0.002093)(18.7,0.002071)(18.8,0.002029)(18.9,0.001986)(19,0.001957)(19.1,0.001951)(19.2,0.001951)(19.3,0.001943)(19.4,0.001932)(19.5,0.001914)(19.6,0.001889)(19.7,0.001859)(19.8,0.001846)(19.9,0.001833)(20,0.001798)};
+\end{axis}
+\end{tikzpicture}
+}
+\caption{Kinetic energy dissipation rate for entropy stable GLL and Gauss collocation schemes with $N=7$ and $h = \pi/8$. }
+\label{fig:tgv}
+\end{figure}
+
+Figure~\ref{fig:tgv} shows the evolution of the kinetic energy dissipation rate from $t\in [0,20]$ for Gauss and GLL collocation schemes on affine and curved meshes.
+The curved meshes used here are constructed by modifying nodal positions through the mapping
+\[
+\tilde{\bm{x}} = \bm{x} + \frac{1}{2}\sin(x)\sin(y)\sin(z).
+\]
+All cases utilize $N=7$ and $h = \pi/8$ (corresponding to $8$ elements per side), as well as a CFL of $.25$.
+Lax-Friedrichs dissipation is used for all simulations.
+For both affine and curved meshes, the presented Gauss collocation schemes remain stable in the presence of highly under-resolved solution features.
+Kinetic energy dissipation rates for both GLL and Gauss collocation are qualitatively similar and are consistent with existing results in the literature for the inviscid Taylor-Green vortex \cite{gassner2016split,chan2018discretely}.
+
+\section{A theoretical cost comparison}
+
+While the numerical experiments presented in previous sections demonstrate several advantages of Gauss collocation methods over GLL collocation schemes, these do not account for additional costs associated with Gauss collocation schemes.
+While a detailed time-to-solution comparison is outside of the scope of this work, we can compare computational costs associated with Gauss, staggered-grid, and GLL collocation schemes.
+
+The main computational costs associated with entropy stable schemes are volume operations, which include evaluations of two-point fluxes and applications of one-dimensional differentiation and interpolation matrices.
+We do not count inter-element communication or flux computations, as these are typically less expensive than volume operations (especially for higher polynomial degrees).
+In three dimensions, a degree $N$ GLL collocation scheme contains $(N+1)^3$ nodes.
+Two-point fluxes are computed between states at one node and states at $(N+1)$ additional nodes, resulting in a total of $(N+1)^4$ two-point flux evaluations in each dimension.
+For a staggered grid scheme, two-point fluxes are computed on a degree $(N+1)$ GLL grid consisting of $(N+2)^3$ nodes, resulting in $(N+2)^4$ flux evaluations per dimension.
+
+Gauss collocation schemes compute two-point fluxes on a grid of Gauss nodes, resulting in $(N+1)^4$ flux evaluations.
+However, one must also evaluate two-point fluxes between face nodes and volume nodes, and vice versa.
+Specifically, Gauss schemes perform two evaluations of two-point fluxes between each face node and the line of $(N+1)$ volume nodes normal to that face, resulting in an extra $4(N+1)^3$ flux evaluations per dimension.
+Because these flux evaluations correspond to the computation of $\bm{Q}^i_N\circ \bm{F}_S$, one can also exploit the skew symmetric nature of the off-diagonal blocks of $\bm{Q}^i_N$ to reduce the number of two-point flux evaluations to $2(N+1)^3$.
+
+We also consider costs associated with applying operator matrices.
+GLL collocation schemes require applying a one-dimensional differentiation matrix in each dimension, resulting in $O\LRp{(N+1)^4}$ operations per dimension in 3D.
+Staggered-grid schemes require $O\LRp{(N+2)^4}$ operations per dimension (corresponding to differentiation on a degree $(N+1)$ GLL grid) as well as $O\LRp{2(N+2)(N+1)^3}$ operations required for interpolation to and from a $(N+1)$ point Gauss grid to a $(N+2)$ point GLL grid in three dimensions.
+Gauss collocation schemes require only interpolation to and from face points, resulting in an additional $O\LRp{2(N+1)^3}$ operations per dimension per element.
+
+\begin{figure}
+\centering
+\subfloat[Two-point flux evaluations]{
+\begin{tikzpicture}
+\begin{axis}[
+	width=.525\textwidth,
+	legend cell align=left,
+	xlabel={Degree $N$},
+%	ylabel={Number of flux computations},
+	xmin=.5, xmax=7.5,
+%	ymin=0,ymax=7,
+%	ybar stacked,
+             ybar=2*\pgflinewidth,
+	bar width=2.5pt,
+%	nodes near coords,
+	%xmin=.5, xmax=9.5,
+	xtick={1,2,3,4,5,6,7},
+	legend pos=north west,
+	xmajorgrids=true,
+	ymajorgrids=true,
+	grid style=dashed,
+] 
+%nodal runtimes
+\addplot coordinates{(1,48)(2,243)(3,768)(4,1875)(5,3888)(6,7203)(7,12288)};
+\addplot coordinates{(1,243)(2,768)(3,1875)(4,3888)(5,7203)(6,12288)(7,19683)};
+\addplot coordinates{(1,144)(2,567)(3,1536)(4,3375)(5,6480)(6,11319)(7,18432)};
+\addplot coordinates{(1,96)(2,405)(3,1152)(4,2625)(5,5184)(6,9261)(7,15360)};
+\legend{GLL, Staggered, Gauss, Gauss (w/sym.)}
+\end{axis}
+\end{tikzpicture}
+}
+\subfloat[Matrix computation operations]{
+\begin{tikzpicture}
+\begin{axis}[
+	width=.525\textwidth,
+	legend cell align=left,
+%	title={Node-per-thread (NPT) nodal},
+	xlabel={Degree $N$},
+%	ylabel={Runtime (s)},
+	xmin=.5, xmax=7.5,
+%	ymin=0,ymax=7,
+             ybar=2*\pgflinewidth,
+	bar width=4pt,             
+	xtick={1,2,3,4,5,6,7},
+	legend pos=north west,
+	xmajorgrids=true,
+	ymajorgrids=true,
+	grid style=dashed,
+] 
+%nodal runtimes
+\addplot coordinates{(1,96)(2,486)(3,1536)(4,3750)(5,7776)(6,14406)(7,24576)};
+\addplot coordinates{(1,582)(2,1968)(3,5030)(4,10776)(5,20454)(6,35552)(7,57798)};
+\addplot coordinates{(1,192)(2,810)(3,2304)(4,5250)(5,10368)(6,18522)(7,30720)};
+
+\legend{GLL, Staggered, Gauss}
+
+\end{axis}
+\end{tikzpicture}}
+\caption{Comparison of GLL, staggered grid, and Gauss collocation schemes in terms of number of two-point flux evaluations and operations associated with matrix computations. }
+\label{fig:cost}
+\end{figure}
+
+Figure~\ref{fig:cost} shows the estimated number of two-point flux evaluations and matrix operations for GLL, staggered grid, and Gauss collocation schemes in three dimensions.
+We observe that a straightforward implementation of Gauss collocation does not significantly reduce the number of flux evaluations compared to staggered grid schemes, though this can be improved by exploiting symmetry in the off-diagonal blocks of the decoupled SBP operator.
+Gauss collocation schemes see a much more significant gain over staggered grid schemes in terms of matrix computations.
+
+We note that the number of two-point flux evaluations and matrix operations will impact runtime differently depending on the implementation and computational architecture.
+For example, while flux evaluations typically dominate runtimes for serial CPU implementations at all orders of approximation, they do not contribute significantly to runtimes at polynomial degrees $N=1,\ldots, 8$ for implementations on Graphics Processing Units (GPUs)  \cite{wintermeyer2018entropy}.
+
 \section{Conclusion}
 
-\note{Give computational comparison: two point flux evaluations, FLOPS due to matrix operations. }
+This work shows how to construct efficient entropy stable high order Gauss collocation DG schemes on quadrilateral and hexahedral meshes.
+Key to the construction of efficient methods are decoupled SBP operators, which deliver entropy stability and high order accuracy while maintaining compact inter-element coupling terms.
+These operators are also compatible with existing entropy stable methods for applying interface dissipation \cite{winters2017uniquely} or imposing boundary conditions.
+Numerical experiments demonstrate both the stability and high order accuracy of the proposed Gauss collocation schemes on both affine and curvilinear meshes.
+
+We note that results for Gauss collocation are similar to those attained by entropy stable staggered-grid schemes \cite{parsani2016entropy}.
+However, Gauss collocation schemes result in a lower number of two-point flux evaluations and floating point operations from matrix computations compared to staggered-grid methods.
+While a rigorous computational comparison between GLL and Gauss collocation schemes remains to be done, Gauss collocation schemes show significant improvements in accuracy compared to GLL collocation schemes on non-Cartesian meshes.
+In particular, for sufficiently warped curvilinear mappings, degree $N$ Gauss collocation schemes achieve an accuracy comparable to degree $(N+1)$ GLL collocation schemes in two and three dimensions.
+
+\section{Acknowledgments}
+
+Jesse Chan is supported by NSF DMS-1719818 and DMS-1712639.
 
@@
 The decoupled SBP operator $\bm{Q}_N^i$ is then given as
-\[
+\begin{equation}
 \bm{Q}_N^i = \begin{bmatrix}
@@
 \end{bmatrix}.
-\]
+\label{eq:qni_quad}
+\end{equation}
 A straightforward computation shows that $\bm{Q}_N^i$ satisfies an SBP property \cite{chan2017discretely}.  
@@
 A collocation scheme assumes that the number of quadrature points is identical to the number of basis functions.
-If the solution is represented using degree $N$ Lagrange polynomials at quadrature points, the matrices $\bm{V}_q, \bm{P}_q$ simplify
+If the solution is represented using degree $N$ Lagrange polynomials at quadrature points, the matrices $\bm{V}_q, \bm{P}_q$ simplify to
 \[
@@
 \]
-In summary, the collocation assumption removes the need to interpolate to and from a polynomial space to volume quadrature points within an entropy stable scheme.
+Plugging these simplifications into (\ref{eq:qni_quad}) and restricting to one spatial dimension recovers the decoupled SBP operator (\ref{eq:qndef}).
+%In summary, the collocation assumption removes the need to interpolate to and from a polynomial space to volume quadrature points within an entropy stable scheme.
 

### 2018-08-05 docs/skewform/main.tex
updating ESDG with EC shock vortex

@@
 \date{}
-\author{Jesse Chan}
-\title{Skew-symmetric entropy stable discontinuous Galerkin formulations with application to hybrid meshes }
+\author{Jesse Chan}%, David C.\ Del Rey Fernandez}
+\title{Skew-symmetric entropy stable discontinuous Galerkin formulations with applications to hybrid and non-conforming meshes}
 
@@
 \begin{equation}
-\int_{\Omega}\pd{U(\bm{u})}{t}\diff{x} + \int_{\partial \Omega} \sum_{j=1}^d \LRp{\bm{v}(\bm{u})^T\bm{f}_j(\bm{u}) - \psi_j\LRp{\bm{v}(\bm{u})}}n_j \diff{x} \leq 0, \qquad \bm{n} = \LRp{n_1,\ldots,n_d},
+\int_{\Omega}\pd{U(\bm{u})}{t}\diff{x} + \int_{\partial \Omega} \sum_{j=1}^d \LRp{\bm{v}(\bm{u})^T\bm{f}_j(\bm{u}) - \psi_j\LRp{\bm{v}(\bm{u})}}n_j \diff{x} \leq 0, 
 \label{eq:entropyineq}
 \end{equation}
-where $\psi_j(\bm{u})$ is a function referred to as the entropy potential.
+where $\bm{n} = \LRp{n_1,\ldots,n_d}$ denotes the outward unit normal, and $\psi_j(\bm{u})$ is some function referred to as the entropy potential.
 
@@
 \begin{assumption}
-The volume quadrature rule integrates exactly all polynomials up to degree $(2N-1)$.
+The volume quadrature rule integrates exactly all polynomials of degree $(2N-1)$.
 \label{ass:quad}
@@
 \begin{proof}
-To show local conservation, we test (\ref{eq:esdgSkewWeak}) with $1$.  Let $\bm{e}$ be the coefficients of $1$ under the basis $\phi_j(\bm{x})$.
+To show local conservation, we test (\ref{eq:esdgSkewWeak}) with $1$  %Let $\bm{e}$ be the coefficients of $1$ under the basis $\phi_j(\bm{x})$.
 \begin{align}
@@
 
-\subsection{Simplicial vs tensor product elements}
-
-It was shown in \cite{chan2018discretely} that the approximation error is $O\LRp{h^{R+1}}$, and that taking $R = (N+1)$ approximates the geometric terms with an error of $O(h^{N+2})$.
-
-\note{Need to note image of curl is in different spaces for simplicial vs tensor product elements.  Influences proofs that geometric terms satisfy appropriate assumptions in those cases.}
-%\subsection{On implementation and computational cost}
-
+%\subsection{Simplicial vs tensor product elements}
+%
+%It was shown in \cite{chan2018discretely} that the approximation error is $O\LRp{h^{R+1}}$, and that taking $R = (N+1)$ approximates the geometric terms with an error of $O(h^{N+2})$.
+%
+%\note{Need to note image of curl is in different spaces for simplicial vs tensor product elements.  Influences proofs that geometric terms satisfy appropriate assumptions in those cases.}
+%%\subsection{On implementation and computational cost}
+%
 \note{Kopriva trick to enforcing GCL still results in $Q^N$ normals 3D.
 May not satisfy requirement that $\bm{1}^T\bm{Q}^i_N \psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i_N \psi_i(\tilde{\bm{u}})$ because of quadrature inaccuracy.
-Note - this still works for conforming hexes because of the full matrix SBP property.}
+Note - this still works regardless for conforming hexes because of the full matrix SBP property.}
 

### 2018-08-13 docs/skewform/main.tex
updating with new skew form writing

@@
 
-%\documentclass[preprint,10pt]{elsarticle}
-%\documentclass[preprint,10pt]{article}
 \documentclass[review]{siamart0216}
-%\documentclass{siamart0216}
-
-%\usepackage{fullpage}
-%\usepackage[colorlinks=true]{hyperref}
 
 \usepackage{amsmath,amssymb,amsfonts}
-%\usepackage{amsthm}
-%\theoremstyle{definition}
-%\newtheorem{definition}{Definition}
-%\theoremstyle{lemma}
-%\newtheorem{lemma}{Lemma}
-%\theoremstyle{corollary}
-%\newtheorem{corollary}{Corollary}
-%\theoremstyle{theorem}
-%\newtheorem{theorem}{Theorem}
 \newtheorem*{remark}{Remark}
@@
 
-\renewcommand{\hat}[1]{\widehat{#1}}
+\renewcommand{\hat}[1]{\hat{#1}}
 \renewcommand{\topfraction}{0.85}
@@
 \newcommand{\LdK}{L^2\LRp{\partial D^k}}
-\newcommand{\Dhat}{\widehat{D}}
+\newcommand{\Dhat}{\hat{D}}
 \newcommand{\Lhat}{L^2\LRp{\Dhat}}
+
+\renewcommand{\hat}{\widehat}
 
@@
 
-%\begin{frontmatter}
-%\title{Entropy stable skew-symmetric discontinuous Galerkin formulations with applications to hybrid and non-conforming meshes. }
-%
-%\author[rice]{Jesse Chan\corref{cor1}}
-%\ead{Jesse.Chan@caam.rice.edu}
-%\address[rice]{Department of Computational and Applied Mathematics, Rice University, 6100 Main St, Houston, TX, 77005}
-%
-%\begin{abstract}
-%Entropy stable high order methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.  
-%\end{abstract}
-%\end{frontmatter}
-
-%\tableofcontents
+\begin{abstract}
+Entropy stable high order methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.
+The construction of such schemes has relied on the use of carefully chosen collocation points \cite{gassner2013skew, fisher2013high, carpenter2014entropy, chan2018efficient} or volume and surface quadrature rules \cite{chan2017discretely, chan2018discretely} to produce operators which satisfy a summation-by-parts (SBP) property.
+In this work, we show how to generalize the construction of semi-discretely entropy stable schemes to accommodate more general pairs of volume and surface quadratures under which an SBP property does not hold.
+We conclude by discussing the applicability of these new operators for entropy stable schemes on hybrid and non-conforming meshes.
+\end{abstract}
 
@@
 In contrast, these discrete entropy inequalities account for the loss of the chain rule due to effects such as quadrature errors, which are incurred when applying polynomially exact quadrature rules to nonlinear and rational integrands within the DG formulation.
-These schemes were first introduced as high order collocation methods on tensor product elements in \cite{fisher2013high, carpenter2014entropy, gassner2016split, gassner2017br1}, and were extended to simplicial elements in \cite{crean2017high, chen2017entropy, crean2018entropy, chan2017discretely}.
-These methods have also been extended to staggered grid methods \cite{parsani2016entropy} and non-conforming meshes \cite{friedrich2017entropy}.
+These schemes were first introduced as high order collocation methods on tensor product elements in \cite{fisher2013high, carpenter2014entropy, gassner2016split, gassner2017br1}, and were extended to simplicial elements in \cite{crean2017high, chen2017entropy, crean2018entropy, chan2017discretely, chan2018discretely}.
+These methods have also been extended to staggered grid \cite{parsani2016entropy}, generalized SBP and Gauss nodes \cite{chan2018efficient} and non-conforming meshes \cite{friedrich2017entropy}.
 Entropy stable boundary conditions have also been determined for the compressible Euler and Navier-Stokes equations \cite{parsani2015entropy, svard2018entropy}.
+
+\note{Finish: add Section description.
+Talk about situations where the SBP or decoupled SBP property doesn't hold: triangles with reduced surface quadrature, GLL quads with GQ face quadratures, and hybrid couplings.}
 
@@
 
-\section{Approximation spaces}
+\section{Polynomial approximation spaces}
 
-We define an approximation space using degree $N$ polynomials on the reference element.
 In this work, we consider either simplicial reference elements (triangles and tetrahedra) or tensor product reference elements (quadrilaterals and hexahedra).
-The natural polynomial approximation space differs depending on the element type \cite{chan2015gpu}.
+We define an approximation space using degree $N$ polynomials on the reference element; however, the natural polynomial approximation space differs depending on the element type \cite{chan2015gpu}.
 
@@
 \[
-P^N\LRp{\widehat{D}} = \LRc{\hat{x}_1^{i_1}\ldots\hat{x}_d^{i_d}, \quad \hat{\bm{x}} \in \widehat{D}, \quad 0\leq \sum_{k=1}^d i_k \leq N}.
+P^N\LRp{\hat{D}} = \LRc{\hat{x}_1^{i_1}\ldots\hat{x}_d^{i_d}, \quad \hat{\bm{x}} \in \hat{D}, \quad 0\leq \sum_{k=1}^d i_k \leq N}.
 \]
-In contrast, the natural polynomial space on a $d$-dimensional tensor product element is 
+In contrast, the natural polynomial space on a $d$-dimensional tensor product element is the space of maximum degree $N$ polynomials
 \[
-Q^N\LRp{\widehat{D}} = \LRc{\hat{x}_1^{i_1}\ldots\hat{x}_d^{i_d}, \quad \hat{\bm{x}} \in \widehat{D}, \quad 0\leq i_k \leq N, \quad k = 1,\ldots, d}.
+Q^N\LRp{\hat{D}} = \LRc{\hat{x}_1^{i_1}\ldots\hat{x}_d^{i_d}, \quad \hat{\bm{x}} \in \hat{D}, \quad 0\leq i_k \leq N, \quad k = 1,\ldots, d}.
 \]
-We take $V^N$ to denote the natural approximation space on a given element.  Furthermore, we denote the dimension of $V^N$  as $N_p = {\rm dim}\LRp{V^N\LRp{\widehat{D}}}$.
+We denote the natural approximation space on a given reference element $\hat{D}$ by $V^N$.  Furthermore, we denote the dimension of $V^N$ as $N_p = {\rm dim}\LRp{V^N\LRp{\hat{D}}}$.
 
@@
 Let $\hat{f}$ be a face of the reference element $\hat{D}$.
-The trace space $V^N \LRp{\hat{f}}$ is defined as the restrictions of functions in $V^N\LRp{\hat{D}}$ to $\hat{f}$
+The trace space $V^N \LRp{\hat{f}}$ is defined as the restrictions of functions in $V^N\LRp{\hat{D}}$ to $\hat{f}$, and denote the dimension of the trace space as ${\rm dim}\LRp{V^N\LRp{{\hat{f}}}} = N^f_p$.
 \[
@@
 \]
-For example, the trace space for degree $N$ polynomials on a quadrilateral face $\hat{f}$ of the bi-unit hexahedral element $[-1,1]^3$ is $Q^N\LRp{\hat{f}}$, while the trace space for degree $N$ polynomials on a triangular face $f$ of the tetrahedron is $P^N\LRp{\hat{f}}$.
-We denote the dimension of the trace space as ${\rm dim}\LRp{V^N\LRp{{\hat{f}}}} = N^f_p$.
+For example, on a $d$-dimensional simplex, $V^N \LRp{\partial \hat{D}}$ consists of total degree $N$ polynomials on simplices of dimension $(d-1)$.
+On a $d$-dimensional tensor product element, $V^N \LRp{\partial \hat{D}}$ consists of maximum degree $N$ polynomials on a tensor product element of dimension $(d-1)$.
 
-We similarly define the trace space for the surface $\partial \hat{D}$ of $\hat{D}$
-\[
-V^N \LRp{\partial \hat{D}} = \LRc{ \left.u\right|_{\partial \hat{D}}, \quad u \in V^N\LRp{\hat{D}}}.
-\]
+%For example, the trace space for degree $N$ polynomials on a quadrilateral face $\hat{f}$ of the bi-unit hexahedral element $[-1,1]^3$ is $Q^N\LRp{\hat{f}}$, while the trace space for degree $N$ polynomials on a triangular face $f$ of the tetrahedron is $P^N\LRp{\hat{f}}$.
 
-\section{$L^2$ projection and lifting matrices and a decoupled SBP operator}
+%We similarly define the trace space for the surface $\partial \hat{D}$ of $\hat{D}$
+%\[
+%V^N \LRp{\partial \hat{D}} = \LRc{ \left.u\right|_{\partial \hat{D}}, \quad u \in V^N\LRp{\hat{D}}}.
+%\]
+%For example, on a $d$-dimensional simplex, $V^N \LRp{\partial \hat{D}}$ consists of total degree $N$ polynomials on simplices of dimension $(d-1)$.
+On a $d$-dimensional tensor product element, $V^N \LRp{\partial \hat{D}}$ consists of maximum degree $N$ polynomials on a tensor product element of dimension $(d-1)$.
 
-The schemes in \cite{chan2017discretely} begin on the reference element $\hat{D}$ by approximating the solution in a degree $N$ polynomial basis $\LRc{\phi_j({\bm{x}})}_{i=1}^{N_p}$.
-It is assumed that volume and surface quadrature rules $({\bm{x}}_i, w_i)$, $\LRp{{\bm{x}}^f_i,w^f_i}$ are provided for $\hat{D}$ and the surface $\partial \hat{D}$ of $\hat{D}$, respectively.
-We make the following assumption on the volume quadrature rule:
-\begin{assumption}
-The volume quadrature rule integrates exactly all polynomials of degree $(2N-1)$.
-\label{ass:quad}
-\end{assumption}
-In addition to being related to the positive definiteness of the mass matrix, Assumption~\ref{ass:quad} ensures that, for any $u,v \in V^N$, the following integral is exactly approximated using quadrature
-\[
-\int_{\hat{D}} \pd{u(\hat{\bm{x}})}{\hat{x}_i} v(\hat{\bm{x}})\diff{\hat{\bm{x}}} = \sum_{k=1}^{N_q} w_k \pd{u(\hat{\bm{x}}_k)}{\hat{x}_i} v(\hat{\bm{x}}_k).  
-\]
-In other words, evaluating the integral of the derivative of $u$ and $v$ using quadrature is exact because the integrand is a polynomial of degree $(2N-1)$.
+\section{Quadrature-based matrices and decoupled SBP operators}
 
 \note{TODO: Standardize hat notation for operators/normals: can refer to operators on reference element and do $k$ superscript for physical elements, or do hats on reference element.}
+Let $\hat{D} \subset\mathbb{R}^d$ denote a reference element with surface $\partial \hat{D}$.  
+The high order schemes in \cite{chan2017discretely, chan2018discretely} begin by approximating the solution in a degree $N$ polynomial basis $\LRc{\phi_j({\bm{x}})}_{i=1}^{N_p}$ on $\hat{D}$.
+These schemes also assume volume and surface quadrature rules $({\bm{x}}_i, w_i)$, $\LRp{{\bm{x}}^f_i,w^f_i}$ on $\hat{D}$.
+We will specify the accuracy of each quadrature rule later, and discuss how quadrature accuracy implies specific summation-by-parts properties.
 
@@
 Interpolation, differentiation, and $L^2$ projection matrices can be combined to construct finite difference operators.
-For example, the matrix $\bm{D}_q = \bm{V}_q\bm{D}^i\bm{P}_q$ maps function values at quadrature points to approximate values of the derivative at quadrature points.
-By choosing specific quadrature rules, $\bm{D}_q$ recovers high order summation-by-parts finite difference operators in \cite{gassner2013skew, fernandez2014generalized, ranocha2018generalised} and certain operators in \cite{hicken2016multidimensional}.
+For example, the matrix $\bm{D}^i_q = \bm{V}_q\bm{D}^i\bm{P}_q$ maps function values at quadrature points to approximate values of the derivative at quadrature points.
+By choosing specific quadrature rules, $\bm{D}^i_q$ recovers high order summation-by-parts finite difference operators in \cite{gassner2013skew, fernandez2014generalized, ranocha2018generalised} and certain operators in \cite{hicken2016multidimensional}.
 However, to address difficulties in designing efficient entropy stable interface terms for nonlinear conservation laws, the PI introduced in \cite{chan2017discretely} a new ``decoupled'' summation by parts matrix which builds interface terms directly into the approximation of the derivative.
 
-Let $\bm{n}$ denote the outward unit normal $\bm{n}$ on $\hat{D}$, and let $\bm{n}_i$ denote the vector containing values of the $i$th component at all surface quadrature points.
+Let $\hat{\bm{n}}$ denote the scaled outward normal vector $\hat{\bm{n}} = \LRc{\hat{n}_1\hat{J}_f,\ldots,\hat{n}_d\hat{J}_f}$, where $\hat{J}_f$ is the determinant of the Jacobian of the mapping of a face of $\partial \hat{D}$ to a reference face.
+Let $\hat{\bm{n}}_i$ denote the vector containing values of the $i$th component $\hat{n}_i\hat{J}f$ at all surface quadrature points, and let $\bm{Q}^i = \bm{W}\bm{D}^i_q$.
 The ``decoupled'' summation by parts operator $\bm{D}^i_N$ is defined as the block matrix involving both volume and surface quadratures
-\begin{align}
-\bm{D}^i_N  &= \LRs{
+\begin{gather}
+\bm{D}^i_N = \bm{W}_N^{-1} \bm{Q}^i_N, \qquad \bm{W}_N = \LRp{\begin{array}{cc}\bm{W}&\\ & \bm{W}_f\end{array}},\label{eq:QN}\\
+\bm{Q}^i_N  = \LRs{
 \begin{array}{cc}
-\bm{D}^i_q - \frac{1}{2}\bm{V}_q \bm{L}_f {\rm diag}({\bm{n}}_i) \bm{V}_f\bm{P}_q &  \frac{1}{2}\bm{V}_q\bm{L}_f{\rm diag}({\bm{n}}_i)\\
--\frac{1}{2}{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q & \frac{1}{2}{\rm diag}({\bm{n}}_i)
-\end{array}}.  
-\label{eq:DN}
-\end{align}
+\bm{Q}^i - \frac{1}{2}\LRp{\bm{V}_f \bm{P}_q}^T \bm{W}_f {\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q &  \frac{1}{2}\LRp{\bm{V}_f \bm{P}_q}^T \bm{W}_f {\rm diag}(\hat{\bm{n}}_i)\\
+-\frac{1}{2}\bm{W}_f{\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q & \frac{1}{2} \bm{W}_f{\rm diag}(\hat{\bm{n}}_i)
+\end{array}}.  \nonumber
+\end{gather}
 When combined with projection and lifting matrices, $\bm{D}^i_N$ produces a high order polynomial approximation of $f\pd{g}{x_i}$. 
@@
 Let $n_i$ be the $i$th component of the unit normal on $\hat{D}$.
-Then (\ref{eq:dnapprox}) is equivalent to finding $u(\bm{x}) \approx f\pd{g}{x_i}$ such that
+Then (\ref{eq:dnapprox}) is equivalent to finding $u(\bm{x}) \approx f\pd{g}{x_i}$ such that, for all $v\in V^N\LRp{\hat{D}}$,
+\begin{align*}
+\int_{\hat{D}} u(\bm{x})v(\bm{x}) = \int_{\hat{D}}{g\pd{\Pi_Nf}{x_i}v} + \int_{\partial \hat{D}}{(f-\Pi_Nf)\frac{\LRp{gv + \Pi_N(gv)}}{2}}\hat{n}_i\hat{J}_f.
+\end{align*}
+
+The matrix $\bm{Q}^i_N$ also satisfies a decoupled summation-by-parts (SBP) property, which is used to prove semi-discrete entropy stability for nonlinear conservation laws.  
+\begin{lemma}
+Assume that the volume and surface quadrature rules are sufficiently accurate such that the quantities
 \[
-\int_{\hat{D}} u(\bm{x})v(\bm{x})  = \int_{\hat{D}}{g\pd{\Pi_Nf}{x_i}v} + \int_{\partial \hat{D}}{(f-\Pi_Nf)\frac{\LRp{gv + \Pi_N(gv)}}{2}}n_i, \qquad \forall v\in V^N\LRp{\hat{D}}.
+\int_{\hat{D}} \pd{u}{x_i} v, \qquad \int_{\partial \hat{D}} u v n_i
 \]
-If the surface quadrature rule is exact for polynomials of degree $2N$, then the decoupled SBP operator $\bm{D}^i_N$ also satisfies a summation by parts property.  
+are integrated exactly for all $u,v \in V^N\LRp{\hat{D}}$ and $i = 1,\ldots, d$.  Then, the decoupled SBP operator $\bm{D}^i_N$ satisfies a summation by parts property:
 \begin{gather}
-\bm{Q}^i_N = \bm{W}_N \bm{D}^i_N, \qquad \bm{W}_N = \LRp{\begin{array}{cc}\bm{W}&\\ & \bm{W}_f\end{array}},\label{eq:dsbp}\\ \bm{Q}^i_N+\LRp{\bm{Q}^i_N}^T = \LRp{\begin{array}{cc}\bm{0}&\\ & \bm{W}_f {\rm diag}(\bm{n}_i)\end{array}} = \bm{B}_N\nonumber
+\bm{Q}^i_N+\LRp{\bm{Q}^i_N}^T = \LRp{\begin{array}{cc}\bm{0}&\\ & \bm{W}_f {\rm diag}(\bm{n}_i)\end{array}} = \bm{B}_N\label{eq:dsbp}
 \end{gather}
-By noting that $\bm{W}\bm{V}_q \bm{L}_q = \bm{P}_q^T\bm{V}_f^T\bm{W}_f$, the entries of $\bm{Q}^i_N$ can be re-written as
-\begin{align}
-\bm{Q}_i &= \bm{W}\bm{V}_q\bm{D}^i\bm{P}_q,\\
-\bm{Q}^i_N  &= \LRs{
-\begin{array}{cc}
-\bm{Q}_i - \frac{1}{2}\LRp{\bm{V}_f \bm{P}_q}^T  \bm{W}_f{\rm diag}({\bm{n}}_i) \bm{V}_f\bm{P}_q &  \frac{1}{2}\LRp{\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q}^T\\
--\frac{1}{2}\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q & \frac{1}{2}\bm{W}_f{\rm diag}({\bm{n}}_i)
-\end{array}}.
-\label{eq:QN}
-\end{align}
-from which the SBP property (\ref{eq:dsbp}) is more apparent.
+%\begin{enumerate}
+%\item The mass matrix is positive definite, 
+%\item The quadrature integrates exactly $\int_{\hat{D}} \pd{u}{x_i} v$ for all $u,v \in V^N\LRp{\hat{D}}$ and $i = 1,\ldots, d$.
+%\end{enumerate}
+\label{lemma:dsbp}
+\end{lemma}
+\begin{proof}
+The proof is a straightforward extension of Theorem 1 in \cite{chan2017discretely} to polynomial approximation spaces on non-simplicial elements.  
+\end{proof}
 
-%\begin{lemma}[Lemma 1 in \cite{chan2017discretely}]
-%Let $\bm{D}^i_q = \bm{V}_q\bm{D}^i\bm{P}_q$.  Then, for volume quadratures which are exact for degree $2N-1$ polynomials, $\bm{D}^i_q$ satisfies the SBP property with respect to $\bm{W}$ 
-%\[
-%\bm{W}\bm{D}^i_q = \LRp{\bm{V}_f\bm{P}_q}^T \bm{W}_f\diag{\hat{\bm{n}}} \bm{V}_f\bm{P}_q - \LRp{\bm{W}\bm{D}^i_q}^T.  
-%\]
-%\label{lemma:gsbp}
-%\end{lemma}
+The assumptions of Lemma~\ref{lemma:dsbp} are satisfied for sufficiently accurate volume and surface quadratures.
+For example, on simplicial elements, (\ref{eq:dsbp}) holds if the volume quadrature is exact for polynomial integrands of total degree $(2N-1)$, and the surface integral is exact for degree $2N$ polynomials on each face.
+Tensor product elements require stricter conditions: (\ref{eq:dsbp}) holds if the volume quadrature is exact for polynomial integrands of degree $2N$ in each coordinate, due to the fact that derivatives of $u\in Q^N$ are degree $(N-1)$ with respect to one coordinate and degree $N$ with respect to others.
+
+\begin{remark}
+It is worth noting that the assumptions of Lemma~\ref{lemma:dsbp} are sufficient but not necessary conditions.
+For example, suppose $\hat{D}$ is a quadrilateral element, and that the volume quadrature is a tensor product of $(N+1)$ point one-dimensional Gauss-Legendre-Lobatto (GLL) rule.
+If the surface quadrature also taken to be an $(N+1)$-point GLL rule over each face, then $\bm{Q}^i$ satisfies a traditional SBP property, which can be used to prove the decoupled SBP property for $\bm{Q}^i_N$.
+\end{remark}
 
@@
 
-\subsection{Loss of the SBP property for hybrid and non-conforming meshes}
+While the SBP property has been used to derive entropy stable schemes, the SBP property is difficult or impossible to enforce in certain discretization settings, such as hybrid and non-conforming meshes arises.
+This difficulty is a result of the choices of volume and surface quadrature which naturally arise in these settings.
+We first illustrate how specific pairings of volume and surface quadratures can result in the loss of the SBP property (\ref{eq:dsbp}).
+We then propose a skew-symmetric formulation which is entropy conservative without explicitly requiring operators which satisfy the SBP property.
 
-While the SBP property has been used to derive entropy stable schemes, the SBP property is difficult or impossible to enforce in certain discretization settings, such as hybrid and non-conforming meshes arises.
-This difficulty is a result of the choices of surface quadrature which naturally arise for these settings.
+\subsection{Loss of the SBP property}
+
+In this section, we give examples of specific pairings of volume and surface quadratures under which the decoupled SBP property does not hold (see Figure~\ref{fig:sbploss}).
+We consider two dimensional reference elements $\hat{D}$ with spatial coordinates $x,y$.
+\begin{figure}
+\centering
+\subfloat[GLL volume quadrature, Gauss surface quadrature]{\includegraphics[width=.4\textwidth]{figs/gllgauss.png}\label{subfig:gllgq}}
+\hspace{2em}
+\subfloat[Degree $2N$ volume quadrature, GLL surface quadrature]{\includegraphics[width=.4\textwidth]{figs/trigll.png}\label{subfig:trigll}}
+\caption{Volume and surface quadrature pairs which do not satisfy the assumptions of Lemma~\ref{lemma:dsbp}, and thus do not possess the decoupled SBP property (\ref{eq:dsbp}). }
+\label{fig:sbploss}
+\end{figure}
+
+\paragraph{Quadrilateral elements (Figure~\ref{subfig:gllgq})} We first consider a quadrilateral element $\hat{D}$ with an $(N+1)$ point tensor product GLL volume quadrature and $(N+1)$ point Gauss quadrature on each face.
+Let $u,v \in Q^N$ denote two arbitrary degree $N$ polynomials.
+The assumptions of Lemma~\ref{lemma:dsbp} are that the volume quadrature exactly integrates $\int_{\hat{D}} \pd{u}{x} v$ and that the surface quadrature exactly integrates $\int_{\partial \hat{D}} u v n_i$ on $\hat{D}$.
+Because the $(N+1)$-point Gauss rule is exact for polynomials of degree $2N+1$ and the product $uv \in P^{2N}$ on each face, the surface quadrature satisfies the assumptions of Lemma~\ref{lemma:dsbp}.
+However, the 1D GLL rule is only exact for polynomials of degree $(2N-1)$.
+The derivative $\pd{u}{x}$ is a polynomial of degree $(N-1)$ in $x$, but is degree $N$ in $y$.
+Thus, $\pd{u}{x}v$ is a polynomial of degree $(2N-1)$ in $x$ but degree $2N$ in $y$, and is not integrated exactly by the volume quadrature.
+
+\paragraph{Triangular elements (Figure~\ref{subfig:trigll})} We next consider a triangular element $\hat{D}$, where the volume quadrature is exact for degree $2N$ polynomials \cite{xiao2010quadrature} and an $(N+1)$-point GLL quadrature on each face.
+Let $u,v \in P^N$ denote two arbitrary degree $N$ polynomials.
+The derivative $\pd{u}{x} \in P^{(N-1})$, and $\pd{u}{x}v \in P^{(2N-1)}$, so the volume quadrature satisfies the assumptions of Lemma~\ref{lemma:dsbp}.
+However, because the surface quadrature is exact only degree $(2N-1)$ polynomials and the trace of $uv\in P^{2N}$, the surface quadrature does not satisfy the assumptions of Lemma~\ref{lemma:dsbp}.
 
@@
 \subfloat[Incompatible surface quadrature on the quadrilateral element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D_GQ.png}\label{subfig:hybrid2}}
-\caption{Examples of interface couplings which do not result in an SBP property and are not entropy stable.
-On the left, the surface quadrature is a $(N+1)$ point GLL rule, and results in a loss of the SBP property on the triangle.
-On the right, the surface quadrature is a $(N+1)$ point Gauss-Legendre rule, and results in a loss of the SBP property on the quadrilateral.
-}
+\caption{Examples of interface couplings which do not result in a decoupled SBP property (\ref{eq:dsbp}).  }
 \label{fig:hybrid}
@@
 
-\begin{figure}
-\centering
-\subfloat[Hex-pyramid coupling]{\includegraphics[width=.35\textwidth]{figs/hybrid3D.png}}
-\hspace{2em}
-\subfloat[Hex-prism coupling]{\includegraphics[width=.375\textwidth]{figs/hybrid3D_wedge.png}}
-\caption{Illustration of a 3D coupling between a GLL hexahedral element and a pyramid.  The SBP property does not hold on the pyramid due to the use of GLL quadrature on the quadrilateral face. }
-\label{fig:hybrid3d}
-\end{figure}
+These specific pairings of volume and surface quadratures appear naturally for hybrid meshes consisting of DG-SEM quadrilateral elements (using GLL volume quadrature) and triangular elements, as shown in Figure~\ref{fig:hybrid}.
+In Figure~\ref{subfig:hybrid1}, the surface quadrature is a $(N+1)$ point GLL rule, and results in a loss of the SBP property on the triangle.
+In Figure~\ref{subfig:hybrid2}, the surface quadrature is a $(N+1)$ point Gauss-Legendre rule, and results in a loss of the SBP property on the quadrilateral element.
 
-\begin{figure}[!h]
-\centering
-\begingroup
-\captionsetup[subfigure]{width=.45\textwidth}
-\subfloat[Entropy stable inter-element coupling in \cite{friedrich2017entropy} for a non-conforming interface.]{\raisebox{0em}{\includegraphics[width=.45\textwidth]{figs/nonconSBP.png}}\label{subfig:noncon2}}
-\hspace{2em}
-\subfloat[Entropy stable inter-element coupling in this work for a non-conforming interface.
+The goal of this work is to construct high order accurate discretizations which preserve entropy conservation for these situations in which the decoupled SBP property (\ref{eq:dsbp}) does not hold.
+
+%\begin{figure}
+%\centering
+%\subfloat[Hex-pyramid coupling]{\includegraphics[width=.35\textwidth]{figs/hybrid3D.png}}
+%\hspace{2em}
+%\subfloat[Hex-prism coupling]{\includegraphics[width=.375\textwidth]{figs/hybrid3D_wedge.png}}
+%\caption{Illustration of a 3D coupling between a GLL hexahedral element and a pyramid.  The SBP property does not hold on the pyramid due to the use of GLL quadrature on the quadrilateral face. }
+%\label{fig:hybrid3d}
+%\end{figure}
+%
+%\begin{figure}[!h]
+%\centering
+%\begingroup
+%\captionsetup[subfigure]{width=.45\textwidth}
+%\subfloat[Entropy stable inter-element coupling in \cite{friedrich2017entropy} for a non-conforming interface.]{\raisebox{0em}{\includegraphics[width=.45\textwidth]{figs/nonconSBP.png}}\label{subfig:noncon2}}
+%\hspace{2em}
+%\subfloat[Entropy stable inter-element coupling in this work for a non-conforming interface.
 The dotted black lines denote communication between neighboring elements.]{\raisebox{-0em}{\includegraphics[width=.45\textwidth]{figs/nonconQuad.png}}\label{subfig:noncon1}}
-\endgroup
-%\subfloat[3D hexahedra-pyramid coupling]{\includegraphics[width=.31\textwidth]{figs/hybrid3D.png}}
-\caption{Illustration of two different entropy stable inter-element couplings for $h$ non-conforming meshes.
+%\endgroup
+%%\subfloat[3D hexahedra-pyramid coupling]{\includegraphics[width=.31\textwidth]{figs/hybrid3D.png}}
+%\caption{Illustration of two different entropy stable inter-element couplings for $h$ non-conforming meshes.
 Each dashed red line indicates a dyadic flux computation required between two nodes.
@@
 }
-\label{fig:noncon}
-\end{figure}
+%\label{fig:noncon}
+%\end{figure}
 
@@
 
-It can be shown that the SBP property holds for $\bm{D}^i_N$ if the surface quadrature is exact for polynomials of degree $2N$ \cite{chan2017discretely}.
-The SBP property is lost if the surface quadrature is only exact for polynomial integrands of degree less than $2N$.
-However, one can still show that a variational version of the SBP property is satisfied.
+%The main restriction we wish to overcome
+
+The property (\ref{eq:dsbp}) (which we will refer to as the ``strong'' SBP property) relates the polynomial exactness of specific quadrature rules to algebraic properties of quadrature-based matrices.
+We will relax accuracy conditions on these quadrature rules by utilizing not the strong SBP property, but a weaker ``variational'' version of the SBP property.
 
 \begin{lemma}
-Let $v\in V^N$, and let the surface trace $\LRl{v}_{\partial \hat{D}} \in V^M\LRp{\partial \hat{D}}$ for some $M \leq N$.
-If the surface quadrature is exact for polynomials of degree $N+M$,  then for an arbitrary vector $\bm{u}_q$,
+Assume that $v \in V^M$.  %, and let the surface trace $\LRl{v}_{\partial \hat{D}} \in V^M\LRp{\partial \hat{D}}$ for some $M \leq N$. 
+If the volume quadrature is exact for polynomials of degree $N+M-1$ and the surface quadrature is exact for polynomials of degree $N+M$, then for an arbitrary vector $\bm{u}_q$, 
 \[
@@
 \]
-where $\bm{v}$ denotes the vector of values of $v$ of at both volume and surface quadrature points. 
+where $\bm{v}$ denotes the vector of values of $v(\bm{x})$ of at both volume and surface quadrature points. 
 \label{lemma:vsbp}
@@
 \begin{align*}
-\bm{v}^T\bm{Q}^i_N\bm{u} &= \bm{v}_q^T\bm{Q}_i \bm{u}_q - \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}} \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}} \bm{u}_f\\
-& - \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}} \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} + \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}} \bm{u}_f.
+\bm{v}^T\bm{Q}^i_N\bm{u} &= \bm{v}_q^T\bm{Q}_i \bm{u}_q - \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}}_i \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}}_i \bm{u}_f\\
+& - \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}}_i \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} + \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}}_i \bm{u}_f.
 \end{align*}
@@
 \begin{align*}
-\bm{v}^T\bm{Q}^i_N\bm{u} =& -\bm{v}_q^T\bm{Q}^T_i \bm{u}_q + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}} \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}} \bm{u}_f\\
-&\qquad - \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}} \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} - \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}} \bm{u}_f + \bm{v}_f^T\bm{W}_f\hat{\bm{n}} \bm{u}_f\\
+\bm{v}^T\bm{Q}^i_N\bm{u} =& -\bm{v}_q^T\bm{Q}^T_i \bm{u}_q + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}}_i \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}}_i \bm{u}_f\\
+&\qquad - \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}}_i \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} - \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}}_i \bm{u}_f + \bm{v}_f^T\bm{W}_f\hat{\bm{n}}_i \bm{u}_f\\
 =& \begin{pmatrix} \bm{v}_q\\ \bm{v}_f\end{pmatrix}^T 
@@
 \bm{0}& \\
-& \bm{W}_f\diag{\hat{\bm{n}}}
+& \bm{W}_f\diag{\hat{\bm{n}}_i}
 \end{pmatrix}\right.\\
@@
 \left.\begin{pmatrix}
--\bm{Q}_i^T + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q}^T\bm{W}_f\hat{\bm{n}} \bm{V}_f\bm{P}_q & -\frac{1}{2} \LRp{\bm{W}_f\hat{\bm{n}} \bm{V}_f\bm{P}_q}^T\\
-\frac{1}{2}\bm{W}_f\hat{\bm{n}} \bm{V}_f\bm{P}_q & -\frac{1}{2}\bm{W}_f\hat{\bm{n}}
+-\bm{Q}_i^T + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q}^T\bm{W}_f\hat{\bm{n}}_i \bm{V}_f\bm{P}_q & -\frac{1}{2} \LRp{\bm{W}_f\hat{\bm{n}}_i \bm{V}_f\bm{P}_q}^T\\
+\frac{1}{2}\bm{W}_f\hat{\bm{n}}_i \bm{V}_f\bm{P}_q & -\frac{1}{2}\bm{W}_f\hat{\bm{n}}_i
 \end{pmatrix}  \right)
 \begin{pmatrix} \bm{u}_q\\ \bm{u}_f\end{pmatrix}\\
-=& \bm{v}^T\LRp{\bm{B}_N - \LRp{\bm{Q}^i_N}^T}\bm{u}.
+=& \bm{v}^T\LRp{\bm{B}^i_N - \LRp{\bm{Q}^i_N}^T}\bm{u}.
 \end{align*}
@@
 \[
-\bm{1}^T\bm{Q}^i_N\bm{u} = \bm{1}^T\bm{B}_N\bm{u}, \qquad \bm{Q}^i_N\bm{1} = \bm{0},
+\bm{1}^T\bm{Q}^i_N\bm{u} = \bm{1}^T\bm{B}^i_N\bm{u}, \qquad \bm{Q}^i_N\bm{1} = \bm{0},
 \]
@@
 \[
-\bm{1}^T\LRp{\bm{Q}^i_N}\bm{u} = \bm{1}^T\LRp{\bm{B}_N - \LRp{\bm{Q}^i_N}^T}\bm{u} = \bm{1}^T{\bm{B}_N}\bm{u}.
+\bm{1}^T\LRp{\bm{Q}^i_N}\bm{u} = \bm{1}^T\LRp{\bm{B}^i_N - \LRp{\bm{Q}^i_N}^T}\bm{u} = \bm{1}^T{\bm{B}^i_N}\bm{u}.
 \]
 \end{proof}
+
 \subsection{Entropy stability on affine meshes}
 
-The high order methods in this work ensure that (\ref{eq:entropyineq}) is satisfied discretely by avoiding the use of the chain rule in the proof of entropy dissipation.
+The high order methods in this work ensure that the entropy inequality (\ref{eq:entropyineq}) is satisfied discretely by avoiding the use of the chain rule in the proof of entropy dissipation.
 These ``entropy stable'' schemes rely on two main ingredients: an entropy stable numerical flux as defined by Tadmor \cite{tadmor1987numerical} and a concept referred to as ``flux differencing''.
@@
 This formulation is then made entropy stable by adding interface dissipation, and can extended in a straightforward manner to affine elements.
-The skew-symmetric formulation for (\ref{eq:nonlineqs}) on $\hat{D}$ is
+
+Let $\bm{u}_q$ denote the values of the solution at volume quadrature points.
+We define the auxiliary conservative variables $\tilde{\bm{u}}$ in terms of the $L^2$ projections of the entropy variables
 \begin{gather}
-\pd{\bm{u}}{t} + \sum_{i=1}^d \LRs{\begin{array}{cc}
-\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{\hat{\bm{n}}}\bm{f}_i^* = 0,\\
-\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q,
-\label{eq:esdgSkew}
+\bm{v}_q = \bm{v}\LRp{\bm{u}_q}, \qquad \tilde{\bm{v}} = \begin{bmatrix}
+\bm{V}_q\\
+\bm{V}_f
+\end{bmatrix}\bm{P}_q\bm{v}_q, \qquad \tilde{\bm{u}} = \bm{u}\LRp{\tilde{\bm{v}}}.
 \end{gather}
-where $\bm{f}^*$ is some numerical flux.  Multiplying the formulation (\ref{eq:esdgSkew}) by $\bm{M}$ on both sides yields a weak form 
+The skew-symmetric formulation for (\ref{eq:nonlineqs}) on $\hat{D}$ is given in terms of $\tilde{\bm{u}}$
 \begin{gather}
-%\bm{M}\pd{\bm{u}}{t} + \sum_{i=1}^d \LRs{\begin{array}{c}
-%\bm{V}_q \\ \bm{V}_f\end{array}}^T \begin{pmatrix}
-%\bm{W}&\\
-%&\bm{W}_f
-%\end{pmatrix}
-%\LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} +  \bm{V}_f^T\bm{W}_f\diag{\hat{\bm{n}}}\bm{f}_i^*\label{eq:esdgSkewWeak}\\
-%= 
 \bm{M}\pd{\bm{u}}{t} + \sum_{i=1}^d\LRs{\begin{array}{c}
 \bm{V}_q \\ \bm{V}_f\end{array}}^T 
-\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{V}_f^T\bm{W}_f \diag{\hat{\bm{n}}}\bm{f}_i^* = 0.  \label{eq:esdgSkewWeak}
+\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{V}_f^T\bm{W}_f \diag{\hat{\bm{n}}_i}\bm{f}_i^* = 0,  \label{eq:esdgSkew}\\
+\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q,\nonumber
 \end{gather}
-where the matrix $\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}$ possesses the following structure
+where $\bm{f}^*$ is some numerical flux, and the matrix $\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}$ possesses the following block structure:
 \[
@@
 \bm{Q}_i-\bm{Q}_i^T & \LRp{\bm{V}_f \bm{P}_q}^T\bm{W}_f\diag{\bm{n}_i}\\
-\bm{W}_f\diag{\bm{n}_i}{\bm{V}_f \bm{P}_q} & \bm{0}
+-\bm{W}_f\diag{\bm{n}_i}{\bm{V}_f \bm{P}_q} & \bm{0}
 \end{pmatrix}.
 \]
-
+Multiplying the formulation (\ref{eq:esdgSkew}) by $\bm{M}^{-1}$ on both sides yields a strong form 
+\begin{gather*}
+\pd{\bm{u}}{t} + \sum_{i=1}^d \LRs{\begin{array}{cc}
+\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{\hat{\bm{n}}_i}\bm{f}_i^* = 0,
+%\\
+%\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q.
+%\label{eq:esdgSkewStrong}
+\end{gather*}
 We can now show that the skew-symmetric formulation is semi-discretely entropy conservative.  
@@
 \[
-\bm{1}^T\bm{W}\pd{U(\bm{u}_q)}{t} + \sum_{i=1}^d\bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}} \LRp{\psi_i(\tilde{\bm{u}}_f) - \tilde{\bm{v}}_f^T\bm{f}_i^*} = 0, \qquad \bm{u}_q = \bm{V}_q\bm{u}.
+\bm{1}^T\bm{W}\pd{U(\bm{u}_q)}{t} + \sum_{i=1}^d\bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}_i} \LRp{\psi_i(\tilde{\bm{u}}_f) - \tilde{\bm{v}}_f^T\bm{f}_i^*} = 0, \qquad \bm{u}_q = \bm{V}_q\bm{u}.
 \]
@@
 \begin{proof}
-Testing (\ref{eq:esdgSkewWeak}) by $\bm{v}_h = \bm{P}_q\bm{v}_q$ yields 
+Testing (\ref{eq:esdgSkew}) by $\bm{v}_h = \bm{P}_q\bm{v}_q$ yields 
 \begin{align}
 \bm{v}_q^T\bm{W}\pd{\LRp{\bm{V}_q\bm{u}}}{t} + \sum_{i=1}^d
-\tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \tilde{\bm{v}}_f^T \bm{W}_f \diag{\hat{\bm{n}}}\bm{f}_i^* = 0.
+\tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \tilde{\bm{v}}_f^T \bm{W}_f \diag{\hat{\bm{n}}_i}\bm{f}_i^* = 0.
 \end{align}
@@
 &= \bm{1}^T\LRp{\bm{Q}^i_N}\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i_N\psi_i(\tilde{\bm{u}}) \\
-&= \bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}} \psi_i(\tilde{\bm{u}}_f)
+&= \bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}_i} \psi_i(\tilde{\bm{u}}_f)
 \end{align*}
@@
 \begin{proof}
-To show local conservation, we test (\ref{eq:esdgSkewWeak}) with $1$  %Let $\bm{e}$ be the coefficients of $1$ under the basis $\phi_j(\bm{x})$.
+To show local conservation, we test (\ref{eq:esdgSkew}) with $1$
 \begin{align}
@@
 \]
-is a skew-symmetric matrix.  Using that $\bm{A}$, $\bm{x}^T\bm{A}\bm{x} = 0$ for any skew symmetric matrix $\bm{A}$, the volume term vanishes
+is a skew-symmetric matrix.  Using that $\bm{x}^T\bm{A}\bm{x} = 0$ for any skew symmetric matrix $\bm{A}$, the volume term vanishes
 \[
@@
 Next, we extend this formulation to affinely mapped elements.
-Derivatives with respect to physical coordinates on $D^k$ are computed in terms of a change of variables formula and derivatives with respect to reference coordinates
-\[
-\pd{u}{x_i} = \sum_{ij} \bm{G}^k_{ij}\pd{u}{\hat{x}_j}, \qquad \bm{G}^k_{ij} = J^k\pd{\hat{x}_j}{{x}_i}, 
-\]
+Let the domain be decomposed into non-overlapping elements $D^k$, such that $D^k$ is the image of the reference element $\hat{D}$ under an affine mapping $\bm{\Phi}^k$.
+We define geometric terms ${G}^k_{ij}$ as scaled derivatives of reference coordinates $\hat{\bm{x}}$ w.r.t.\ physical coordinates $\bm{x}$
+\begin{gather}
+\pd{u}{x_i} = \sum_{ij} {G}^k_{ij}\pd{u}{\hat{x}_j}, \qquad {G}^k_{ij} = J^k\pd{\hat{x}_j}{{x}_i}, 
+\label{eq:geofacs}
+\end{gather}
 where $J^k$ is the determinant of the Jacobian of the geometric mapping on the element $D^k$.
-For affine meshes, the scaled geometric terms $\bm{G}^k_{ij}$ are constant over each element.
+We also introduce the scaled outward normal components $n_iJ^k_f$, which can be computed in terms of (\ref{eq:geofacs}) and the reference normals $\hat{\bm{n}}$ on $\hat{D}$
+\begin{gather}
+n^k_i J^k_f = \sum_{j=1}^d G^k_{ij} \hat{{n}}_j.  
+\label{eq:normals}
+\end{gather}
+
+Derivatives with respect to physical coordinates on $D^k$ are computed in terms of a change of variables formula and geometric terms (\ref{eq:geofacs}).
+For affine meshes, the scaled geometric terms ${G}^k_{ij}$ are constant over each element.
 It was shown in \cite{chan2017discretely} that one can define a physical differentiation matrix as a linear combination of reference differentiation matrices.
-Let $\hat{\bm{D}}^i_N$ denote the decoupled SBP operator on the reference element $\hat{D}$, and let the decoupled SBP operators with respect to the physical coordinates on $D^k$ be defined as
+Let ${\bm{D}}^i_N$ denote the decoupled SBP operator on the reference element $\hat{D}$.
+Define the decoupled SBP operators $\bm{D}^i_k$ with respect to the physical coordinates on $D^k$ be defined as
 \begin{align}
-{\bm{D}}^i_N = \sum_{j=1}^d \bm{G}_{ij}\hat{\bm{D}}^j_N.
+{\bm{D}}^i_k = \sum_{j=1}^d {G}^k_{ij}{\bm{D}}^j_N.
 \end{align}
-%Then, $\bm{Q}^i_N = \bm{W} \bm{D}^i_N$ satisfies the SBP property on each element $D^k$
-%\begin{equation}
-%\bm{Q}^i_N + \LRp{\bm{Q}^i_N}^T = \bm{B}^i_N, \qquad \bm{B}^i_N = \begin{pmatrix}
-%\bm{0}&\\
-%& {J^k_f}\bm{W}_f \diag{{\bm{n}_i}}
-%\end{pmatrix}
-%\label{eq:physsbp}
-%\end{equation}
-%where $J^k_f$ are the Jacobian factors which result from mapping the faces of $D^k$ to reference faces.
-%\end{lemma}
+Let $\bm{n}^k_i$ be a vector containing concatenated values of the scaled outward normals $n^k_iJ^k_f$ at surface quadrature nodes.  
 Then, an entropy conservative formulation can be given on $D^k$ as follows:
 \begin{gather}
-\pd{\bm{u}}{t} + \sum_{i=1}^d \LRs{\begin{array}{cc}
-\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{{\bm{n}_i}}\bm{f}_i^* = 0, \label{eq:skewform}\\
+J^k\bm{M}\pd{\bm{u}}{t} + 
+\sum_{i=1}^d \LRs{\begin{array}{cc}
+\bm{V}_q \\
+\bm{V}_f\end{array}}^T \LRp{\LRp{\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T} \circ \bm{F}_S}\bm{1} + \bm{V}_f^T\bm{W}_f \diag{{\bm{n}^k_i}}\bm{f}_i^* = 0, \label{eq:skewform}\\
+%\sum_{i=1}^d \LRs{\begin{array}{cc}
+%\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{{\bm{n}_i}}\bm{f}_i^* = 0, \label{eq:skewform}\\
 \LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q, \nonumber\\
@@
 where $\tilde{\bm{u}}_f^+$ denotes the face value of the entropy-projected conservative variables $\tilde{\bm{u}}_f$ on the neighboring element.
-The formulation can be made entropy stable by adding an appropriate penalization term, such as a Lax-Friedrichs or matrix-based dissipation \cite{winters2017uniquely, chen2017entropy, chan2017discretely}.
+The entropy conservative formulation can be made entropy stable by adding appropriate interface dissipation, such as Lax-Friedrichs or matrix-based penalization terms \cite{winters2017uniquely, chen2017entropy, chan2017discretely}.
 
@@
 On affine meshes, it is possible to show entropy stability of the skew-symmetric formulation (\ref{eq:esdgSkew}) under a surface quadrature which is only exact for degree $N$ polynomials.
-However, on curved meshes, a stronger surface quadrature rule is required to demonstrate entropy stability, where the strength of the surface rule depends on the order and approximation of geometric terms.
+However, on curved meshes, a stronger surface quadrature rule is required to guarantee entropy stability, where the strength of the surface rule depends on the degree of approximation of the geometric terms, which now vary spatially over each element.
 
+\subsection{Discrete geometric conservation law and surface quadrature accuracy}
+
+\note{Add description of how normals factor into this and note that they're computed based on $\bm{G}_{ij}$.}
 Let $\bm{J}^k\bm{G}^k_{ij}$ denote the vector of scaled geometric terms ${J}^k\bm{G}^k_{ij}$ evaluated at volume and surface quadrature points.
@@
 However, additional assumptions must be satisfied in order to prove that the resulting formulation is entropy stable.
-
-\subsection{Discrete geometric conservation law and surface quadrature accuracy}
-
-\note{Add description of how normals factor into this and note that they're computed based on $\bm{G}_{ij}$.}
 
@@
 
-%\end{proof}
-%Thus, if geometric terms are approximated using degree $N$ polynomials as in \cite{kopriva2006metric}, then the skew-symmetric formulation with composite GLL quadrature can be used.
-
-%\subsection{Simplicial vs tensor product elements}
-%
-%It was shown in \cite{chan2018discretely} that the approximation error is $O\LRp{h^{R+1}}$, and that taking $R = (N+1)$ approximates the geometric terms with an error of $O(h^{N+2})$.
-%
-%\note{Need to note image of curl is in different spaces for simplicial vs tensor product elements.  Influences proofs that geometric terms satisfy appropriate assumptions in those cases.}
-%%\subsection{On implementation and computational cost}
-%
 \note{Kopriva trick to enforcing GCL still results in $Q^N$ normals 3D.
@@
 
-In this section, we present two-dimensional experiments on hybrid meshes consisting of quadrilaterals and triangles, as well as on non-conforming meshes of quadrilateral elements.
+In this section, we present two-dimensional experiments which verify the theoretical results presented.
+%on hybrid meshes consisting of quadrilaterals and triangles, as well as on non-conforming meshes of quadrilateral elements.
 
-\subsection{Reduced surface quadrature}
+\subsection{Entropy stability under reduced surface quadrature}
 
@@
 
-\subsection{Hybrid meshes}
+\subsection{GLL quadrilaterals with Gauss surface quadrature}
 
-\subsection{Non-conforming meshes}
+\note{Explain that for GLL quadratures, the decoupled SBP property doesn't hold when Gauss points are used.  }
+
+\subsection{Hybrid quadrilateral-triangular meshes}
+
+%\subsection{Non-conforming meshes}
 

### 2018-08-13 matlab/BBtests/BB_curved_mesh.m

@@ -1,6 +1,6 @@
 Globals2D;
 
-N = 4;
+N = 3;
 
 Nq = 2*N+1;
 
@@ -35,7 +35,7 @@ x1 = x; y1 = y;
 xe1 = Ve*x; ye1 = Ve*y;
 
 opt = 1;
-displacement = .3*sin(pi*x(Fmask(:,1),:));
+displacement = .25*cos(pi/2*x(Fmask(:,1),:));
 
 
 % xeC = xe1; yeC = ye1; yeC(3) = yeC(3) + .6;
@@ -78,11 +78,11 @@ if opt==1
     for e = 1:size(tri,1)
         A(tri(e,:),tri(e,:)) = 1; % adjacency matrix
     end
-    %     A = DBr*DBr' + DBs*DBs';
+    %     A = DBr*DBr' + DBs*DBs';    
     A = E'*E;
-    %     A = abs(A)>1e-8;
+%         A = double(abs(A)>.1);
     
-    %     keyboard
+% keyboard
     nbrs = {};
     weights = {};
     for i = 1:Np

### 2018-08-14 docs/skewform/main.tex
updating skewform paper

@@
 \author{Jesse Chan}%, David C.\ Del Rey Fernandez}
-\title{Skew-symmetric entropy stable discontinuous Galerkin formulations with applications to hybrid and non-conforming meshes}
+\title{Skew-symmetric entropy stable discontinuous Galerkin formulations}
 
@@
 The construction of such schemes has relied on the use of carefully chosen collocation points \cite{gassner2013skew, fisher2013high, carpenter2014entropy, chan2018efficient} or volume and surface quadrature rules \cite{chan2017discretely, chan2018discretely} to produce operators which satisfy a summation-by-parts (SBP) property.
-In this work, we show how to generalize the construction of semi-discretely entropy stable schemes to accommodate more general pairs of volume and surface quadratures under which an SBP property does not hold.
-We conclude by discussing the applicability of these new operators for entropy stable schemes on hybrid and non-conforming meshes.
+In this work, we show how to construct skew-symmetric schemes which are entropy stable even for volume and surface quadratures under which an SBP property does not hold.
+These skew-symmetric formulations avoid the use of a ``strong'' SBP property, and require only that operators exactly differentiate constants and satisfy a discrete form of the fundamental theorem of calculus.
+We conclude by discussing the applicability of these new operators for entropy stable schemes on hybrid meshes.% and non-conforming meshes.
 \end{abstract}
@@
 \begin{equation}
-\pd{\bm{u}}{t}  + \sum_{j=1}^d\pd{\bm{f}_j(\bm{u})}{x_j} = 0, \qquad \bm{u}\in \mathbb{R}^n, \qquad \bm{f}:\mathbb{R}^n\rightarrow\mathbb{R}^n,
+\pd{\bm{u}_h}{t}  + \sum_{j=1}^d\pd{\bm{f}_j(\bm{u})}{x_j} = 0, \qquad \bm{u}\in \mathbb{R}^n, \qquad \bm{f}:\mathbb{R}^n\rightarrow\mathbb{R}^n,
 \label{eq:nonlineqs}
@@
 
+The proofs presented in this work will also use anisotropic tensor product polynomial spaces, where the maximum polynomial degree varies depending on the coordinate direction.
+We denote such spaces by $Q^{N_1, \ldots, N_d}$, where $N_k$ are non-negative integers and
+\[
+Q^{N_1, N_2, \ldots, N_d}\LRp{\hat{D}} = \LRc{\hat{x}_1^{i_1}\ldots\hat{x}_d^{i_d}, \quad \hat{\bm{x}} \in \hat{D}, \quad 0\leq i_k \leq N_k, \quad k = 1,\ldots, d}.
+\]
+For example, the isotropic tensor product space $Q^N$ is the same as $Q^{N,\ldots,N}$.
+
 We also define trace spaces for each reference element.
@@
 \[
-V^N \LRp{\hat{f}} = \LRc{ \left.u\right|_{\hat{f}}, \quad u \in V^N\LRp{\hat{D}}}, \qquad \hat{f}\in \partial\hat{D}.
+V^N \LRp{\hat{f}} = \LRc{ \left.u\right|_{\hat{f}}, \quad u \in V^N\LRp{\hat{D}}, \quad \hat{f}\in \partial\hat{D}}.
 \]
@@
 
-\note{TODO: Standardize hat notation for operators/normals: can refer to operators on reference element and do $k$ superscript for physical elements, or do hats on reference element.}
+%\note{TODO: Standardize hat notation for operators/normals: can refer to operators on reference element and do $k$ superscript for physical elements, or do hats on reference element.}
 Let $\hat{D} \subset\mathbb{R}^d$ denote a reference element with surface $\partial \hat{D}$.  
@@
 
-The matrix $\bm{Q}^i_N$ also satisfies a decoupled summation-by-parts (SBP) property, which is used to prove semi-discrete entropy stability for nonlinear conservation laws.  
+For sufficiently accurate quadrature rules, the matrix $\bm{Q}^i_N$ also satisfies a decoupled summation-by-parts (SBP) property, which is used to prove semi-discrete entropy stability for nonlinear conservation laws.
 \begin{lemma}
@@
 For example, on simplicial elements, (\ref{eq:dsbp}) holds if the volume quadrature is exact for polynomial integrands of total degree $(2N-1)$, and the surface integral is exact for degree $2N$ polynomials on each face.
-Tensor product elements require stricter conditions: (\ref{eq:dsbp}) holds if the volume quadrature is exact for polynomial integrands of degree $2N$ in each coordinate, due to the fact that derivatives of $u\in Q^N$ are degree $(N-1)$ with respect to one coordinate and degree $N$ with respect to others.
+Tensor product elements require stricter conditions: (\ref{eq:dsbp}) holds if both the volume and surface quadratures are exact for polynomial integrands of degree $2N$ in each coordinate, due to the fact that derivatives of $u\in Q^N$ are degree $(N-1)$ with respect to one coordinate and degree $N$ with respect to others.
 
@@
 Let $u,v \in P^N$ denote two arbitrary degree $N$ polynomials.
-The derivative $\pd{u}{x} \in P^{(N-1})$, and $\pd{u}{x}v \in P^{(2N-1)}$, so the volume quadrature satisfies the assumptions of Lemma~\ref{lemma:dsbp}.
+The derivative $\pd{u}{x} \in P^{(N-1)}$, and $\pd{u}{x}v \in P^{(2N-1)}$, so the volume quadrature satisfies the assumptions of Lemma~\ref{lemma:dsbp}.
 However, because the surface quadrature is exact only degree $(2N-1)$ polynomials and the trace of $uv\in P^{2N}$, the surface quadrature does not satisfy the assumptions of Lemma~\ref{lemma:dsbp}.
@@
 In Figure~\ref{subfig:hybrid2}, the surface quadrature is a $(N+1)$ point Gauss-Legendre rule, and results in a loss of the SBP property on the quadrilateral element.
-
-The goal of this work is to construct high order accurate discretizations which preserve entropy conservation for these situations in which the decoupled SBP property (\ref{eq:dsbp}) does not hold.
+The goal of this work is to construct high order accurate discretizations which preserve entropy conservation for situations in which the decoupled SBP property (\ref{eq:dsbp}) does not hold.
 
@@
 We will relax accuracy conditions on these quadrature rules by utilizing not the strong SBP property, but a weaker ``variational'' version of the SBP property.
-
+This variational property relies on the assumption that the volume and surface quadrature rules are exact for polynomials of certain degrees.
+
+\begin{assumption}
+\label{ass:quad}
+Let $v(\bm{x})$ be a polynomial such that $v \in V^{M_{\rm vol}}$ with surface trace $\LRl{v}_{\partial \hat{D}} \in V^{M_{\rm face}}\LRp{f}$ for $f \in \partial \hat{D}$, where $M_{\rm vol}, M_{\rm face} \leq N$ are two non-negative integers.
+We assume that the volume and surface quadrature rules are sufficiently accurate such that
+\begin{enumerate}
+\item the mass matrix $\bm{M}$ is positive definite, 
+\item the volume and surface quadratures exactly integrate the quantities $\int_{\hat{D}} \pd{u}{x_i} v$ and $\int_{\partial \hat{D}} u v n_i$ for all $u \in V^N\LRp{\hat{D}}$, $i = 1,\ldots, d$, and $f \in \partial \hat{D}$.
+\end{enumerate}
+\end{assumption}
+
+This section presumes that Assumption~\ref{ass:quad} is satisfied for general  $M_{\rm vol}, M_{\rm surf}$.
+In Sections~\ref{sec:affine} and \ref{sec:curved}, specific values of $M_{\rm vol}, M_{\rm surf}$ will be motivated by proofs of entropy stability, and Assumption~\ref{ass:quad} will be verified for specific choices of volume and surface quadrature on simplicial and tensor product elements.
+%Suppose Assumption~\ref{ass:quad} is satisfied, and let $\bm{v}_q$ denote the vector of values of $v(\bm{x})$ at volume quadrature points.  Then, for an arbitrary vector $\bm{u}_q$, 
+%\[
+%\bm{v}_q^T \bm{Q}_i \bm{u}_q = \int_{\hat{D}} \pd{u}{x_i} v, \qquad \LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f \diag{\hat{\bm{n}_i}} \bm{V}_f\bm{P}_q\bm{u}_q =  \int_{\partial \hat{D}} \Pi_N u v n_i,
+%\]
+%where we have used that, because $v$ is a polynomial of degree $N$, it is equal to its own $L^2$ projection.
+
+The conditions in Assumption~\ref{ass:quad} are sufficient to ensure that the following variational SBP property holds:
 \begin{lemma}
-Assume that $v \in V^M$.  %, and let the surface trace $\LRl{v}_{\partial \hat{D}} \in V^M\LRp{\partial \hat{D}}$ for some $M \leq N$. 
-If the volume quadrature is exact for polynomials of degree $N+M-1$ and the surface quadrature is exact for polynomials of degree $N+M$, then for an arbitrary vector $\bm{u}_q$, 
-\[
-\bm{v}_q^T\bm{Q}_i \bm{u}_q = \bm{v}_q^T\LRp{ \LRp{\bm{V}_f\bm{P}_q}^T \bm{W}_f \diag{{\bm{n}}_i}\bm{V}_f\bm{P}_q - \bm{Q}_i^T}\bm{u}_q
-\]
-where $\bm{v}$ denotes the vector of values of $v(\bm{x})$ of at both volume and surface quadrature points. 
 \label{lemma:vsbp}
+%Assume that $v \in V^M$.  
+%If the volume quadrature is exact for polynomials of degree $N+M-1$ and the surface quadrature is exact for polynomials of degree $N+M$, 
+Suppose Assumption~\ref{ass:quad} holds, and let $\bm{v}_q$ denote the vector of values of $v(\bm{x})$ of at volume quadrature points.
+Let $u(\bm{x})$ be an $L^2$ integrable function, and let $\bm{u}_q$ denote the values of $u(\bm{x})$ at quadrature points.
+Then,
+\[
+\bm{v}_q^T\bm{Q}_i \bm{u}_q = \bm{v}_q^T\LRp{ \LRp{\bm{V}_f\bm{P}_q}^T \bm{W}_f \diag{\hat{\bm{n}}_i}\bm{V}_f\bm{P}_q - \bm{Q}_i^T}\bm{u}_q.
+\]
 \end{lemma}
 \begin{proof}
-Recall that $\bm{Q}_i = \bm{W} \bm{V}_q \bm{D}^i\bm{P}_q$.  Since the volume quadrature is exact for degree $2N-1$ polynomials, we have that
+Recall that $\bm{Q}_i = \bm{W} \bm{V}_q \bm{D}^i\bm{P}_q$.
+The quadrature-based $L^2$ projection $\Pi_Nu$ is a polynomial of degree $N$, and is computed discretely as $\bm{P}_q\bm{u}_q$.
+By the exactness of the volume quadrature in Assumption~\ref{ass:quad},
 \begin{align*}
@@
 \end{align*}
-Since the volume integrand $\LRp{\Pi_N u} \pd{v}{\hat{x}_i} \in V^{2N-1}$, it is exactly integrated using the volume quadrature.
-Additionally, since $\Pi_N u \in V^N$ and $v\in V^M$ on the surface of $\hat{D}$, the surface quadrature is exact for the surface integral and
+%Since the volume integrand $\LRp{\Pi_N u} \pd{v}{\hat{x}_i} \in V^{N+M-1}$, it is exactly integrated using the volume quadrature.  
+%Additionally, since $\Pi_N u \in V^N$ and $v\in V^M$ on the surface of $\hat{D}$, the surface quadrature is exact for the surface integral and 
+By the exactness of the surface quadrature in Assumption~\ref{ass:quad}, 
 \begin{align*}
@@
 \end{align*}
-where we have used that, since $v\in V^N$, $\bm{v}_f = \bm{V}_f\bm{P}_q\bm{v}_q$.  The proof is completed by noting that 
+where we have used that, since $v \in V^N$, $\bm{v}_f = \bm{V}_f\bm{P}_q\bm{v}_q$.  The proof is completed by noting that 
 \[
@@
 
-The variational SBP property also extends to the decoupled SBP operator.  This property, along with the exact differentiation of constants, is necessary for the proof of entropy stability.  
+The variational SBP property in Lemma~\ref{lemma:vsbp} can be used to prove a similar property for the decoupled SBP operator.
+This property, along with the exact differentiation of constants, is necessary for the proof of entropy stability.
 \begin{lemma}
-Let $\bm{D}^i_N$ be a decoupled SBP operator on the reference element $\hat{D}$, and let the surface quadrature be exact for polynomials of degree $N+M$ for some $M \leq N$.
-Suppose $v\in V^N$ and that $\LRl{v}_{\partial \hat{D}} \in P^M\LRp{\partial \hat{D}}$, Then,
+Suppose Assumption~\ref{ass:quad} holds.  Let $\bm{D}^i_N$ be a decoupled SBP operator on the reference element $\hat{D}$.  
+%Let the volume quadrature be exact for polynomials of degree $N+M-1$, and let the surface quadrature be exact for polynomials of degree $N+M$ for some $M \leq N$.
+Suppose $v\in V^N$ and that $\LRl{v}_{\partial \hat{D}} \in P^M\LRp{\partial \hat{D}}$.
+\note{Fix this to match assumptions in previous lemma.}
+Then, the following variational SBP property holds:
 \[
@@
 \]
-where $\bm{v}$ denotes the values of $v$ of at both volume and surface quadrature points.  
+where $\bm{v}, \bm{u}$ denotes the values of $v$ and $u$ at both volume and surface quadrature points.  
 \label{lemma:vdsbp}
@@
 \begin{proof}
-For convenience, let $\bm{u}_q, \bm{u}_f$ denote evaluations of $u$ at volume and surface points, such that 
+For convenience, let $\bm{u}_q, \bm{u}_f$ and $\bm{v}_q, \bm{v}_f$ denote evaluations of $u$ and $v$ at volume and surface points, such that 
 \[
@@
 \end{proof}
-Using Lemma~\ref{lemma:vdsbp}, we will determine the accuracy of the surface quadrature necessary to ensure the proof of semi-discrete entropy stability is valid.
-This minimal degree of surface quadrature accuracy will depend on the nature of the reference-to-physical mapping (e.g.\ affine vs curved elements).
-
-\begin{corollary}
+
+\subsection{Entropy stability on affine meshes}
+\label{sec:affine}
+
+In this section, we construct so-called ``entropy stable'' schemes on the reference element $\hat{D}$ and affine meshes.
+These methods ensure that the entropy inequality (\ref{eq:entropyineq}) is satisfied discretely by avoiding the use of the chain rule in the proof of entropy dissipation.
+Entropy stable schemes rely on two main ingredients: an entropy stable numerical flux as defined by Tadmor \cite{tadmor1987numerical} and a concept referred to as ``flux differencing''.
+Let $\bm{f}_S\LRp{\bm{u}_L,\bm{u}_R}$ be a numerical flux function which is a function of ``left'' and ``right'' states $\bm{u}_L,\bm{u}_R$.
+The numerical flux $\bm{f}_S$ is \textit{entropy conservative} if it satisfies the following three conditions:
+\begin{gather}
+\bm{f}_S(\bm{u},\bm{u}) = \bm{f}(\bm{u}), \qquad \text{(consistency)}\\
+\bm{f}_S(\bm{u}_L,\bm{u}_R) = \bm{f}_S(\bm{u}_R,\bm{u}_R), \qquad \text{(symmetry)}\nonumber\\
+\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}_S(\bm{u}_L,\bm{u}_R) = \psi(\bm{u}_L) - \psi(\bm{u}_R), \qquad \text{(conservation)}\nonumber
+\label{eq:esflux}
+\end{gather}
+The construction of entropy stable schemes will utilize (\ref{eq:esflux}) in discretizations of both volume and surface terms in a DG formulation.
+
+Proving a semi-discrete entropy inequality typically involves the ``strong'' SBP condition.
+The schemes presented in this work will relax this, and use only two (weaker) conditions: that the decoupled SBP operator $\bm{D}^i_N$ exactly differentiates constants and that a discrete form of the fundamental theorem of calculus (FTC) holds.
+The latter property can be derived from the variational SBP property in Lemma~\ref{lemma:vdsbp}.
+
+\begin{corollary}{(Discrete FTC and exact differentiation of constants)}
 \label{lemma:sbpcor}
-Let the surface quadrature be exact for polynomials of degree $N$.  Then, 
-\[
-\bm{1}^T\bm{Q}^i_N\bm{u} = \bm{1}^T\bm{B}^i_N\bm{u}, \qquad \bm{Q}^i_N\bm{1} = \bm{0},
-\]
-where $\bm{u}$ is a vector of values of some function at volume and surface quadrature points. %is an arbitrary vector of appropriate size.  
+Suppose Assumption~\ref{ass:quad} holds for $M_{\rm vol} = M_{\rm surf} = 0$ (i.e.\ for $v(\bm{x})$ constant).  
+Let $u(\bm{x})$ be some $L^2$ integrable function, and let $\bm{u}$ denote the vector of values of $u(\bm{x})$ at volume and surface quadrature points. Then, 
+\[
+\bm{1}^T\bm{Q}^i_N\bm{u} = \bm{1}^T\bm{B}^i_N\bm{u}, \qquad \bm{Q}^i_N\bm{1} = \bm{0}.
+\]
 \end{corollary}
@@
 
-\subsection{Entropy stability on affine meshes}
-
-The high order methods in this work ensure that the entropy inequality (\ref{eq:entropyineq}) is satisfied discretely by avoiding the use of the chain rule in the proof of entropy dissipation.
-These ``entropy stable'' schemes rely on two main ingredients: an entropy stable numerical flux as defined by Tadmor \cite{tadmor1987numerical} and a concept referred to as ``flux differencing''.
-% entropy stable schemes in one-dimension.
-Let $\bm{f}_S\LRp{\bm{u}_L,\bm{u}_R}$ be a numerical flux function which is a function of ``left'' and ``right'' states $\bm{u}_L,\bm{u}_R$.  
-The numerical flux $\bm{f}_S$ is \textit{entropy conservative} if it satisfies the following three conditions:  
-%The numerical flux $\bm{f}_S$ is an \textit{entropy stable} flux function if it satisfies the following three conditions:  
-\begin{gather}
-\bm{f}_S(\bm{u},\bm{u}) = \bm{f}(\bm{u}), \qquad \text{(consistency)}\\
-\bm{f}_S(\bm{u}_L,\bm{u}_R) = \bm{f}_S(\bm{u}_R,\bm{u}_R), \qquad \text{(symmetry)}\nonumber\\
-\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}_S(\bm{u}_L,\bm{u}_R) = \psi(\bm{u}_L) - \psi(\bm{u}_R), \qquad \text{(conservation)}\nonumber
-%\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}_S(\bm{u}_L,\bm{u}_R) \leq \psi(\bm{u}_L) - \psi(\bm{u}_R), \qquad \text{(entropy dissipation)}\nonumber
-\label{eq:esflux}
-\end{gather}
-%If instead of the third condition, $\bm{f}_S$ satisfies a \emph{stability} property 
-%\[
-%\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}_S(\bm{u}_L,\bm{u}_R) \leq \psi(\bm{u}_L) - \psi(\bm{u}_R)$, the flux $\bm{f}_S$ is referred to as \textit{entropy stable}.
-
-We begin by deriving a skew-symmetric formulation on the reference element $\hat{D}$, and showing it is entropy conservative.
-This formulation is then made entropy stable by adding interface dissipation, and can extended in a straightforward manner to affine elements.
-
-Let $\bm{u}_q$ denote the values of the solution at volume quadrature points.
+We can now construct a skew-symmetric formulation on the reference element $\hat{D}$ and show that it is semi-discretely entropy conservative.
+We will then extend this formulation to affine elements.
+In both cases, this formulation can be made entropy stable by adding interface dissipation.
+
+Let $\bm{u}_h$ denote the discrete solution, and let $\bm{u}_q$ denote the values of the solution at volume quadrature points.
 We define the auxiliary conservative variables $\tilde{\bm{u}}$ in terms of the $L^2$ projections of the entropy variables
@@
 \begin{gather}
-\bm{M}\pd{\bm{u}}{t} + \sum_{i=1}^d\LRs{\begin{array}{c}
+\bm{M}\pd{\bm{u}_h}{t} + \sum_{i=1}^d\LRs{\begin{array}{c}
 \bm{V}_q \\ \bm{V}_f\end{array}}^T 
@@
 Multiplying the formulation (\ref{eq:esdgSkew}) by $\bm{M}^{-1}$ on both sides yields a strong form 
-\begin{gather*}
-\pd{\bm{u}}{t} + \sum_{i=1}^d \LRs{\begin{array}{cc}
-\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{\hat{\bm{n}}_i}\bm{f}_i^* = 0,
+\begin{gather}
+\pd{\bm{u}_h}{t} + \sum_{i=1}^d \LRs{\begin{array}{cc}
+\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{\hat{\bm{n}}_i}\bm{f}_i^* = 0,\label{eq:esdg}
 %\\
@@
 %\label{eq:esdgSkewStrong}
-\end{gather*}
-We can now show that the skew-symmetric formulation is semi-discretely entropy conservative.  
+\end{gather}
+We can now show that the skew-symmetric formulation is entropy conservative.  
 \begin{theorem}
-Let the surface quadrature be exact for polynomials of degree $N$.  Then, the formulation (\ref{eq:esdgSkew}) is entropy conservative such that
+Suppose Assumption~\ref{ass:quad} holds for $M_{\rm vol} = M_{\rm surf} = 0$.  
+Then, the formulation (\ref{eq:esdgSkew}) is entropy conservative such that
 \[
@@
 &= \bm{1}^T\LRp{\bm{Q}^i_N}\psi_i(\tilde{\bm{u}}) - \psi_i(\tilde{\bm{u}})^T\LRp{\bm{Q}^i_N}\bm{1} \\
-&= \bm{1}^T\LRp{\bm{Q}^i_N}\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i_N\psi_i(\tilde{\bm{u}}) \\
-&= \bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}_i} \psi_i(\tilde{\bm{u}}_f)
+&= \bm{1}^T\LRp{\bm{Q}^i_N}\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i_N\psi_i(\tilde{\bm{u}}),
 \end{align*}
-where we have used Corollary~\ref{lemma:sbpcor} in the second to last equality.
+where we have used Corollary~\ref{lemma:sbpcor} in the last equality.
+Noting that $\bm{1}^T\bm{B}^i_N\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}_i} \psi_i(\tilde{\bm{u}}_f)$ completes the proof.
 %The final step of the proof is  where we have used Lemma~\ref{lemma:vdsbp} to integrate by parts and conclude that $\bm{1}^T\LRp{\bm{Q}^i_N}^T = 0$.
@@
 \begin{remark}
-The only result necessary to prove Theorem~\ref{thm:esdg} is Corollary~\ref{lemma:sbpcor}.
-Thus, when generalizing to curved elements, the proof of entropy stability for the skew-symmetric form will depend only on the curved version of Corollary~\ref{lemma:sbpcor}.
+It should be emphasized that the only result necessary to prove Theorem~\ref{thm:esdg} is Corollary~\ref{lemma:sbpcor}.
+The decoupled SBP property (\ref{eq:dsbp}) is not utilized, and is not necessary to guarantee a semi-discrete conservation of entropy.
 \end{remark}
@@
 \begin{align}
-\bm{1}^T\bm{W}\bm{V}_q\pd{\bm{u}}{t} + \sum_{i=1}^d
+\bm{1}^T\bm{W}\bm{V}_q\pd{\bm{u}_h}{t} + \sum_{i=1}^d
 \bm{1}^T
@@
 \begin{gather}
-J^k\bm{M}\pd{\bm{u}}{t} + 
+J^k\bm{M}\pd{\bm{u}_h}{t} + 
 \sum_{i=1}^d \LRs{\begin{array}{cc}
@@
 
-\subsection{Curvilinear meshes}
+\subsubsection{On conditions for Assumption~\ref{ass:quad}}
+
+Apart from algebraic manipulations, only Corollary~\ref{lemma:sbpcor} is necessary to prove entropy conservation for (\ref{eq:esdg})  and (\ref{eq:skewform}).
+Corollary~\ref{lemma:sbpcor} requires that Assumption~\ref{ass:quad} holds for $M_{\rm vol}=M_{\rm surf}=0$.
+Thus, the volume and surface quadratures must be sufficiently accurate to guarantee that the mass matrix is positive definite and to integrate
+\begin{equation}
+\int_{\hat{D}} \pd{u}{x_i}, \qquad \int_{\partial \hat{D}} u n_i. \label{eq:affineints}
+\end{equation}
+On simplicial elements, the mass matrix is guaranteed to be positive definite for any volume quadrature which is exact for degree $2N$ polynomial integrands.
+This choice of volume quadrature also guarantees that the volume term in (\ref{eq:affineints}) is integrated exactly.
+The surface quadrature can thus be taken to be any quadrature rule which is exact for only degree $N$ integrands on faces.
+In contrast, the construction of simplicial SBP operators has generally required face quadratures which are accurate for degree $2N$ polynomials \cite{hicken2016multidimensional, crean2018entropy, chan2018discretely}.
+
+On tensor product elements, we consider isotropic quadrature rules which are tensor products of one-dimensional quadrature rules.
+For example, we can use a quadrature rule constructed through a tensor product of one-dimensional $(N+1)$-point GLL quadrature rules, which is sufficient to guarantee that the mass matrix is positive definite \cite{canuto2007spectral}.
+This quadrature rule is also exact for integrands in $Q^{2N-1}$, and thus exactly integrates the volume term in (\ref{eq:affineints}).
+This leaves the surface quadrature, which can again be taken to be any quadrature rule exact for degree $N$ integrands.
+For example, on a quadrilateral element, one can use either $\left\lceil\frac{N+1}{2}\right\rceil$-point Gauss quadrature rule or a $\left\lceil\frac{N+3}{2}\right\rceil$-point GLL rule as face quadratures for a degree $N$ scheme.
+%This minimal degree of surface quadrature accuracy will depend on the nature of the physical mesh, and will differ between affine and curved elements.
+
+\subsection{Entropy stability on curvilinear meshes}
+\label{sec:curved}
 
 On affine meshes, it is possible to show entropy stability of the skew-symmetric formulation (\ref{eq:esdgSkew}) under a surface quadrature which is only exact for degree $N$ polynomials.
-However, on curved meshes, a stronger surface quadrature rule is required to guarantee entropy stability, where the strength of the surface rule depends on the degree of approximation of the geometric terms, which now vary spatially over each element.
+However, on curved meshes, stronger conditions are required to guarantee entropy stability.
+This is due to the fact that the geometric terms are now high order polynomials which vary spatially over each element.
+Moreover, Corollary~\ref{lemma:sbpcor} assumes affine geometric mappings, and does not hold on curved elements.
+In this section, we discuss how to extend Corollary~\ref{lemma:sbpcor} to curved simplicial and tensor product elements.
 
@@
 
-\note{Add description of how normals factor into this and note that they're computed based on $\bm{G}_{ij}$.}
-Let $\bm{J}^k\bm{G}^k_{ij}$ denote the vector of scaled geometric terms ${J}^k\bm{G}^k_{ij}$ evaluated at volume and surface quadrature points.
+Let the domain now be triangulated by a watertight mesh (as defined in \cite{chan2018discretely}) consisting of curved elements $D^k$.  
+In contrast to the affine case, geometric terms $G^k_{ij}$ for a curved element are now high order polynomials over the reference domain $\hat{D}$.
+We assume that the scaled normals are still computed from $G^k_{ij}$ using (\ref{eq:normals}).
+In this section, we describe how to construct appropriate decoupled SBP operators on curved meshes, and give conditions on the volume and surface quadrature rules under which a semi-discretely entropy stable scheme can be constructed.
+
+We first show how to construct decoupled SBP operators on curved elements.
+Let $\bm{G}^k_{ij}$ denote the vector of scaled geometric terms ${G}^k_{ij}$ evaluated at both volume and surface quadrature points.
 Decoupled SBP operators on a curved element $D^k$ can be defined as in \cite{chan2018discretely} by
 \begin{equation}
-\bm{D}^i_N = \sum_{j=1}^d \diag{\bm{G}^k_{ij}}\hat{\bm{D}}^j_N + \hat{\bm{D}}^j_N\diag{\bm{G}^k_{ij}}.
+\bm{D}^i_k = \frac{1}{2}\sum_{j=1}^d \LRp{\diag{\bm{G}^k_{ij}}{\bm{D}}^j_N + {\bm{D}}^j_N\diag{\bm{G}^k_{ij}}}, \qquad \bm{Q}^i_k = \bm{W} \bm{D}^i_k.
 \label{eq:dncurved}
 \end{equation}
-The skew-symmetric formulation (\ref{eq:skewform}) can be extended to curvilinear meshes using the definition of $\bm{D}^i_N$ in (\ref{eq:dncurved}).
+It can be shown that, if $\bm{D}^j_N$ satisfies a strong summation by parts property, $\bm{D}^i_k$ satisfies an analogous strong summation by parts property.
+%The construction of $\bm{D}^i_k$ in (\ref{eq:dncurved}) 
+The skew-symmetric formulation (\ref{eq:skewform}) can be extended to curvilinear meshes by replacing the reference element operators $\bm{Q}^i_N$ with the physical operators $\bm{Q}^i_k$ defined by (\ref{eq:dncurved}).
 However, additional assumptions must be satisfied in order to prove that the resulting formulation is entropy stable.
@@
 For curved elements, Lemma~\ref{lemma:vdsbp} and Corollary~\ref{lemma:sbpcor} do not necessarily hold at the discrete level.
-For example, expanding out the condition $\bm{Q}^i_N\bm{1} = \bm{0}$ in terms of (\ref{eq:dncurved}) yields
+For example, expanding out the condition $\bm{Q}^i_k\bm{1} = \bm{0}$ in terms of (\ref{eq:dncurved}) yields
 \begin{align}
-\bm{D}^i_N \bm{1} = \sum_{j=1}^d \diag{\bm{G}^k_{ij}}\hat{\bm{D}}^j_N \bm{1} + \hat{\bm{D}}^j_N\diag{\bm{G}^k_{ij}}\bm{1} = \sum_{j=1}^d \hat{\bm{D}}^j_N\LRp{\bm{G}^k_{ij}} = 0,
+\bm{Q}^i_k \bm{1} = \frac{1}{2} \bm{W}\sum_{j=1}^d \diag{\bm{G}^k_{ij}}{\bm{D}}^j_N \bm{1} + {\bm{D}}^j_N\diag{\bm{G}^k_{ij}}\bm{1} = \frac{1}{2}\bm{W}\sum_{j=1}^d {\bm{D}}^j_N\LRp{\bm{G}^k_{ij}} = 0,
 \label{eq:dgcl}
 \end{align}
-where we have used that $\hat{\bm{D}}^j_N \bm{1} = 0$.
-For degree $N$ isoparametric mappings, the GCL is automatically satisfied in two dimensions due to the fact that the exact geometric terms $\bm{G}^k_{ij}\in P^{N-1}$ \cite{kopriva2006metric}.
-However, in three dimensions, the GCL is not automatically preserved due to the fact that $\bm{G}^k_{ij}\in P^{2N-2}$, and thus cannot be represented exactly using degree $N$ polynomials.
-As a consequence, (\ref{eq:dgcl}) must be enforced through an alternative construction of $\bm{G}^k_{ij}$.
+where we have used that ${\bm{D}}^j_N \bm{1} = 0$.
+For degree $N$ isoparametric mappings, the GCL is automatically satisfied in two dimensions due to the fact that the exact geometric terms ${G}^k_{ij}$ are polynomials of degree $N$ \cite{kopriva2006metric}.
+However, in three dimensions, the GCL is not automatically satisfied due to the fact that the degree of $G^k_{ij}$ is larger than $N$.
+Thus, the geometric terms cannot be represented exactly using degree $N$ polynomials, and (\ref{eq:dgcl}) must be enforced through an alternative construction of ${G}^k_{ij}$.
 
@@
 \bm{G}^k_{i2}\\
-\bm{G}^k_{i3}\end{array}} = -\frac{1}{2}\LRp{\pd{I_{R}\bm{r}^k}{\hat{x}_j}-\pd{I_{R}\bm{r}^j}{\hat{x}_k}}, 
+\bm{G}^k_{i3}\end{array}} = -\frac{1}{2}\LRp{\pd{I_{N_{\rm geo}}\bm{r}^k}{\hat{x}_j}-\pd{I_{N_{\rm geo}}\bm{r}^j}{\hat{x}_k}}, 
 \label{eq:iconscurl}
 \end{align}
-where $I_{R}$ denotes a degree ${R}$ polynomial interpolation operator with appropriate interpolation nodes.\footnote{This interpolation step must be performed using interpolation points with an appropriate number of nodes on each boundary \cite{chan2018discretely}.
-These include, for example, GLL nodes on tensor product elements, as well as Warp and Blend nodes on non-tensor product elements \cite{warburton2006explicit, chan2015comparison}.}  Because the geometric terms are computed by applying the curl, both $\bm{G}^k_{ij}$ and $n_iJ^f$ are approximated using degree $(R-1)$ polynomials.
-
-Because $\bm{D}^i_N$ are now defined through (\ref{eq:dncurved}), Corollary~\ref{lemma:sbpcor} and the proof of entropy stability may not hold for curved elements and must be modified.
-The introduction of curvilinear meshes will impose slightly different conditions on the accuracy of the surface quadrature, which are summarized in the following curved version of Corollary~\ref{lemma:sbpcor}.
-\begin{lemma}
-Assume that the geometric terms $\bm{G}_{ij} \in V^N$ satisfy the discrete GCL (\ref{eq:dgcl}), that $\LRl{\bm{G}_{ij}}_{\partial \hat{D}}\hat{n}_j \in V^R\LRp{\partial \hat{D}}$, and that the surface quadrature is exact for polynomials of degree $N+R$.
-Then,
-\[
-\bm{1}^T\bm{Q}^i_N\bm{u} = \bm{1}^T\bm{B}^i_N \bm{u}, %\qquad \bm{B}^i_N = \begin{pmatrix}
-%\bm{0}& \\
-%& \bm{W}_f\diag{\bm{n}_i}
-%\end{pmatrix}, 
-\qquad \bm{Q}^i_N \bm{1} = 0, 
-\]
-where 
-\[
-\bm{B}^i_N =  \begin{pmatrix}
-\bm{0}&\\
-& {J^k_f}\bm{W}_f \diag{{\bm{n}_i}}
-\end{pmatrix}.
-\]
-\label{lemma:vdsbpcurved}
+where $I_{N_{\rm geo}}$ denotes a degree $N_{\rm geo}$ polynomial interpolation operator with appropriate interpolation nodes.\footnote{This interpolation step must be performed using interpolation points with an appropriate number of nodes on each boundary \cite{chan2018discretely}.
+These include, for example, GLL nodes on tensor product elements, as well as Warp and Blend nodes on non-tensor product elements \cite{warburton2006explicit, chan2015comparison}.}
+
+Because the decoupled SBP operators $\bm{D}^i_k$ are now defined through (\ref{eq:dncurved}), Corollary~\ref{lemma:sbpcor} and the proof of entropy stability no longer hold for curved elements and must be modified.
+The introduction of curvilinear meshes will impose slightly different conditions on the accuracy of the surface quadrature.
+We discuss simplicial and tensor product elements separately, as differences in the natural polynomial approximation spaces will result in different assumptions for each proof.
+
+\begin{lemma}{(Triangles and tetrahedra)}
+Let $D^k$ be a curved simplicial element, and let the geometric terms $G^k_{ij}$ be constructed .
+If the volume quadrature is exact for polynomials of degree $(N+N_{\rm geo}-1)$ and the surface quadrature is exact for polynomials of degree $(N+N_{\rm geo})$, then
+\[
+\bm{1}^T\bm{Q}^i_k\bm{u} = \bm{1}^T\bm{B}^i_k\bm{u}, \qquad \bm{Q}^i_k\bm{1} = \bm{0}.
+\]
+where $\bm{u}$ is a vector of values of some function at volume and surface quadrature points. 
 \end{lemma}
 \begin{proof}
-The second equality $\bm{Q}^i_N \bm{1} = 0$ is a consequence of the discrete GCL (\ref{eq:dgcl}), and the proof is identical to that of \cite{chan2018discretely}.
-For the first equality, if the surface quadrature is exact for polynomials of degree $2N$ on the trace space, then the stronger matrix SBP property $\bm{Q}^i_N = \bm{B}^i_N - \LRp{\bm{Q}^i_N}^T$ holds \cite{chan2018discretely}.
-Combining this with $\bm{Q}^i_N\bm{1} = 0$ yields the desired result.
+d
+\end{proof}
+
+\begin{lemma}{(Quadrilaterals and hexahedra)}
+Let $D^k$ be a curved tensor product element.  Then, 
+\end{lemma}
+\begin{proof}
+\note{Proof relies on two facts: that geometric terms are products of }
+\end{proof}
+
+%, which are summarized in the following curved version of Corollary~\ref{lemma:sbpcor}.
+%\note{Maybe provide 2 lemmas - one for triangles and one for quadrilaterals?  Polynomial degree of geofacs is slightly different between the two.  } 
+%\begin{lemma}
+%Assume that the geometric terms $\bm{G}^k_{ij} \in V^{N_{\rm geo}}$ satisfy the discrete GCL (\ref{eq:dgcl}), that the volume quadrature is exact for polynomials of degree $N+N_{\rm geo}-1$, and that the surface quadrature is exact for polynomials of degree $N+N_{\rm geo}$.
+Then,
+%\[
+%\bm{1}^T\bm{Q}^i_k\bm{u} = \bm{1}^T\bm{B}^i_k \bm{u}, 
+%\qquad \bm{Q}^i_k \bm{1} = 0, 
+%\]
+%where 
+%\[
+%\bm{B}^i_k =  \begin{pmatrix}
+%\bm{0}&\\
+%& \bm{W}_f \diag{{\bm{n}^k_i}}
+%\end{pmatrix}.
+%\]
+%\label{lemma:vdsbpcurved}
+%\end{lemma}
+%\begin{proof}
+%The second equality $\bm{Q}^i_k \bm{1} = 0$ is a consequence of the discrete GCL (\ref{eq:dgcl}), and the proof is identical to that of \cite{chan2018discretely}.  
+%%For the first equality, if the surface quadrature is exact for polynomials of degree $2N$ on the trace space, then the matrix SBP property $\bm{Q}^i_N = \bm{B}^i_N - \LRp{\bm{Q}^i_N}^T$ holds \cite{chan2018discretely}.
+Combining this with $\bm{Q}^i_k\bm{1} = 0$ yields the desired result.
 We focus on the proof of the variational SBP property for surface quadrature which are exact for polynomials of degree $N+R < 2N$.
-
-Expanding $\bm{1}^T\bm{Q}^i_N\bm{u}$ yields
-\[
-\bm{1}^T\bm{Q}^i_N\bm{u} = \frac{1}{2}\sum_{j=1}^d \bm{1}^T \diag{\bm{G}_{ij}} \hat{\bm{Q}}^j_N \bm{u} + \bm{1}^T\hat{\bm{Q}}^j_N \diag{\bm{G}_{ij}} \bm{u}.
+%For the first equality, we utilize the variational SBP property.  
+%Expanding $\bm{1}^T\bm{Q}^i_k\bm{u}$ yields
+%\[
+%\bm{1}^T\bm{Q}^i_k\bm{u} = \frac{1}{2}\sum_{j=1}^d \bm{1}^T \diag{\bm{G}^k_{ij}} {\bm{Q}}^j_N \bm{u} + \bm{1}^T{\bm{Q}}^j_N \diag{\bm{G}^k_{ij}} \bm{u}.
 %=  \frac{1}{2}\sum_{j=1}^d \bm{1}^T \diag{\bm{G}_{ij}} \hat{\bm{Q}}^j_N \bm{u},
-\]
-For the latter term in the sum, Lemma~\ref{lemma:vdsbp} holds and
-\begin{align}
-\sum_{j=1}^d\bm{1}^T\hat{\bm{Q}}^j_N \diag{\bm{G}_{ij}} \bm{u} &= 
-\sum_{j=1}^d\bm{1}^T\LRp{\hat{\bm{B}}^j_N - \LRp{\hat{\bm{Q}}^j_N}^T} \diag{\bm{G}_{ij}} \bm{u} \\
-&= \sum_{j=1}^d\bm{1}^T\hat{\bm{B}}^j_N\diag{\bm{G}_{ij}}\bm{u}, \nonumber
-\label{eq:vsbp1}
-\end{align}
-where we have used that $\hat{\bm{Q}}^j_N \bm{1} = \bm{0}$.  For the former term in the sum, we expand out 
-\begin{gather}
-\sum_{j=1}^d\bm{1}^T \diag{\bm{G}_{ij}} \hat{\bm{Q}}^j_N \bm{u} = 
-\sum_{j=1}^d{\bm{G}_{ij}}^T \hat{\bm{Q}}^j_N \bm{u} \\
-= \sum_{j=1}^d\begin{bmatrix}
-\bm{G}^q_{ij}\\
-\bm{G}^f_{ij}
-\end{bmatrix}^T \begin{bmatrix} 
-\hat{\bm{Q}}_i - \frac{1}{2}\LRp{\bm{V}_f \bm{P}_q}^T  \bm{W}_f{\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q &  \frac{1}{2}\LRp{\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q}^T\label{eq:ibpg}\\
--\frac{1}{2}\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q & \frac{1}{2}\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)
-\end{bmatrix} \begin{bmatrix}\bm{u}_q\\
-\bm{u}_f\end{bmatrix}.\nonumber
-\end{gather}
-where $\bm{G}^q_{ij},\bm{G}^f_{ij}$ denote the values of $\bm{G}_{ij}$ at volume and surface quadrature points.
-
-Because $\bm{G}_{ij} \in V^N$, it is equal to its own $L^2$ projection, and the values at volume and surface quadrature points are related by $\bm{G}^f_{ij} = \bm{V}_f\bm{P}_q\bm{G}^q_{ij}$.
+%\]
+%For the latter term in the sum, Lemma~\ref{lemma:vdsbp} holds and
+%\begin{align}
+%\sum_{j=1}^d\bm{1}^T{\bm{Q}}^j_N \diag{\bm{G}^k_{ij}} \bm{u} &= 
+%\sum_{j=1}^d\bm{1}^T\LRp{{\bm{B}}^j_N - \LRp{{\bm{Q}}^j_N}^T} \diag{\bm{G}^k_{ij}} \bm{u} \\
+%&= \sum_{j=1}^d\bm{1}^T{\bm{B}}^j_N\diag{\bm{G}^k_{ij}}\bm{u}, \nonumber
+%\label{eq:vsbp1}
+%\end{align}
+%where we have used that ${\bm{Q}}^j_N \bm{1} = \bm{0}$.  
+%
+%For the former term in the sum, we use the fact that $G^k_{ij}$, when constructed using (\ref{eq:iconscurl}), is a polynomial of degree $N$.
+Thus, $\bm{G}^k_{ij} = \begin{bmatrix}\bm{V}_q\\ \bm{V}_f\end{bmatrix}\bm{g}$ for some vector of polynomial coefficients $\bm{g}$.
+We can then rewrite the former term as
+%\begin{gather}
+%\sum_{j=1}^d\bm{1}^T \diag{\bm{G}^k_{ij}} {\bm{Q}}^j_N \bm{u} = 
+%\sum_{j=1}^d{\bm{G}^k_{ij}}^T {\bm{Q}}^j_N \bm{u}.% = \sum_{j=1}^d \bm{g}^T \begin{bmatrix}\bm{V}_q\\ \bm{V}_f\end{bmatrix}^T{\bm{Q}}^j_N \bm{u}.
+%%= \sum_{j=1}^d\begin{bmatrix}
+%%\bm{G}^q_{ij}\\
+%%\bm{G}^f_{ij}
+%%\end{bmatrix}^T \begin{bmatrix} 
+%%\hat{\bm{Q}}_i - \frac{1}{2}\LRp{\bm{V}_f \bm{P}_q}^T  \bm{W}_f{\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q &  \frac{1}{2}\LRp{\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q}^T\label{eq:ibpg}\\
+%%-\frac{1}{2}\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q & \frac{1}{2}\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)
+%%\end{bmatrix} \begin{bmatrix}\bm{u}_q\\
+%%\bm{u}_f\end{bmatrix}.\nonumber
+%\end{gather}
+%%where $\bm{G}^q_{ij},\bm{G}^f_{ij}$ denote the values of $\bm{G}_{ij}$ at volume and surface quadrature points.  
+%%Because $\bm{P}_q\bm{V}_q = \bm{I}$, straightforward computations give that %simplify the final expression 
+%%\[
+%%\begin{bmatrix}\bm{V}_q\\ \bm{V}_f\end{bmatrix}^T{\bm{Q}}^j_N = \begin{bmatrix}
+%%\bm{Q}^i-\bm{V}_f^T\bm{W}_f\diag{\hat{\bm{n}}_j}\bm{V}_f\bm{P}_q & \bm{V}_f^T\bm{W}_f\diag{\hat{\bm{n}}_j}
+%%\end{bmatrix}
+%%\]
+%
+%Because $\bm{G}_{ij} \in V^N$, it is equal to its own $L^2$ projection, and the values at volume and surface quadrature points are related by $\bm{G}^f_{ij} = \bm{V}_f\bm{P}_q\bm{G}^q_{ij}$.
 We can use this to simplify (\ref{eq:ibpg})
-\begin{align*}
-\bm{1}^T \diag{\bm{G}_{ij}} \hat{\bm{Q}}^j_N \bm{u} =&  \LRp{\bm{G}^q_{ij}}^T { \bm{Q}_i}  \bm{u}_q - \frac{1}{2} \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q+\bm{G}^f_{ij}}^T{\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q}\bm{u}_q\\
-&+ \frac{1}{2} \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q+\bm{G}^f_{ij}}^T\frac{1}{2}\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{u}_f\\
-=& \LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}}_i - \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q}^T{\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q}}\bm{u}_q \\
-&+ \LRp{\bm{G}^f_{ij}}^T\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{u}_f. %\\
-%&= \LRp{\bm{G}^q_{ij}}^T \LRp{ \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q}^T{\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q}}\bm{u}_q + \LRp{\bm{G}^f_{ij}}^T\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{u}_f 
-\end{align*}
-The first term can be further simplified using Lemma~\ref{lemma:vsbp}
-\[
-\LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}}_i - \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q}^T{\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q}}\bm{u}_q = \LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}}_i}^T\bm{u}_q.
-\]
-The latter term can be combined with the surface term of (\ref{eq:vsbp1}) by noting that 
-\[
-\LRp{\bm{G}^f_{ij}}^T\bm{W}_f \diag{\hat{\bm{n}}_i}\bm{u}_f =\bm{1}^T\hat{\bm{B}}^j_N\diag{\bm{G}_{ij}}\bm{u} = \bm{1}^T\diag{\bm{G}_{ij}}\hat{\bm{B}}^j_N\bm{u},
-\]
-where we have used that $\hat{\bm{B}}^j_N$ is diagonal and commutes with diagonal scaling by $\bm{G}_{ij}$.
-%as $\LRp{\bm{G}^f_{ij}}^T\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{u}_f = \bm{1}^T\hat{\bm{B}}^j_N \bm{u}$.  
-We can use polynomial exactness and the fact that the geometric terms satisfy the continuous GCL by construction \cite{chan2018discretely} to show that $\sum_{j=1}^d  \hat{\bm{Q}}_j {\bm{G}^q_{ij}} = 0$.
-Combining this with (\ref{eq:vsbp1}) and expanding out $\hat{\bm{B}}^j_N$ yields that
-\begin{align*}
-\bm{1}^T\bm{Q}^i_N\bm{u} &= \sum_{j=1}^d \LRp{
-\frac{1}{2}\LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}_i}}^T \bm{u}_q + \bm{1}^T\begin{pmatrix}
-\bm{0} &\\
-& \bm{W}_f \diag{\bm{G}_{ij} \circ \hat{\bm{n}}_j}
-\end{pmatrix}
-\bm{u}} \\
-&=  \bm{1}^T\begin{pmatrix}
-\bm{0} &\\
-& \bm{W}_f \diag{\sum_{j=1}^d \bm{G}_{ij} \circ \hat{\bm{n}}_j}
-\end{pmatrix}
-\bm{u} = \bm{1}^T\bm{B}^i_N\bm{u}.
-\end{align*}
-where we have used in the final step that $\sum_{j=1}^d\bm{G}_{ij}\hat{n}_j = \bm{n}_i J^k_f$ \cite{ciarlet1978finite, chan2018discretely}
-\end{proof}
+%\begin{align*}
+%\bm{1}^T \diag{\bm{G}_{ij}} \hat{\bm{Q}}^j_N \bm{u} =&  \LRp{\bm{G}^q_{ij}}^T { \bm{Q}_i}  \bm{u}_q - \frac{1}{2} \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q+\bm{G}^f_{ij}}^T{\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q}\bm{u}_q\\
+%&+ \frac{1}{2} \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q+\bm{G}^f_{ij}}^T\frac{1}{2}\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{u}_f\\
+%=& \LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}}_i - \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q}^T{\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q}}\bm{u}_q \\
+%&+ \LRp{\bm{G}^f_{ij}}^T\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{u}_f. %\\
+%%&= \LRp{\bm{G}^q_{ij}}^T \LRp{ \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q}^T{\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q}}\bm{u}_q + \LRp{\bm{G}^f_{ij}}^T\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{u}_f 
+%\end{align*}
+%The first term can be further simplified using Lemma~\ref{lemma:vsbp}
+%\[
+%\LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}}_i - \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q}^T{\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q}}\bm{u}_q = \LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}}_i}^T\bm{u}_q.
+%\]
+%The latter term can be combined with the surface term of (\ref{eq:vsbp1}) by noting that 
+%\[
+%\LRp{\bm{G}^f_{ij}}^T\bm{W}_f \diag{\hat{\bm{n}}_i}\bm{u}_f =\bm{1}^T\hat{\bm{B}}^j_N\diag{\bm{G}_{ij}}\bm{u} = \bm{1}^T\diag{\bm{G}_{ij}}\hat{\bm{B}}^j_N\bm{u},
+%\]
+%where we have used that $\hat{\bm{B}}^j_N$ is diagonal and commutes with diagonal scaling by $\bm{G}_{ij}$.
+%%as $\LRp{\bm{G}^f_{ij}}^T\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{u}_f = \bm{1}^T\hat{\bm{B}}^j_N \bm{u}$.  
+%We can use polynomial exactness and the fact that the geometric terms satisfy the continuous GCL by construction \cite{chan2018discretely} to show that $\sum_{j=1}^d  \hat{\bm{Q}}_j {\bm{G}^q_{ij}} = 0$.
+%Combining this with (\ref{eq:vsbp1}) and expanding out $\hat{\bm{B}}^j_N$ yields that
+%\begin{align*}
+%\bm{1}^T\bm{Q}^i_N\bm{u} &= \sum_{j=1}^d \LRp{
+%\frac{1}{2}\LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}_i}}^T \bm{u}_q + \bm{1}^T\begin{pmatrix}
+%\bm{0} &\\
+%& \bm{W}_f \diag{\bm{G}_{ij} \circ \hat{\bm{n}}_j}
+%\end{pmatrix}
+%\bm{u}} \\
+%&=  \bm{1}^T\begin{pmatrix}
+%\bm{0} &\\
+%& \bm{W}_f \diag{\sum_{j=1}^d \bm{G}_{ij} \circ \hat{\bm{n}}_j}
+%\end{pmatrix}
+%\bm{u} = \bm{1}^T\bm{B}^i_N\bm{u}.
+%\end{align*}
+%where we have used in the final step that $\sum_{j=1}^d\bm{G}_{ij}\hat{n}_j = \bm{n}_i J^k_f$ \cite{ciarlet1978finite, chan2018discretely}
+%\end{proof}
 
 Lemma~\ref{lemma:vdsbpcurved} implies that the polynomial degree $R$ of the surface geometric terms must be compatible with the accuracy of the surface quadrature.
-This, in turn, will depend on the way in which geometric terms are approximated.
-\note{Finish.}
+This, in turn, depends on the polynomial degree $N_{\rm geo}$ to which geometric terms are approximated.
 
@@
 Note - this still works regardless for conforming hexes because of the full matrix SBP property.}
+
+\begin{theorem}
+\label{thm:skewformcurved}
+Let the geometric terms $G^k_{ij}$ be computed from (\ref{eq:iconscurl}), let the scaled outward normals $\bm{n}^k_i$ be computed pointwise from (\ref{eq:normals}), and let $\bm{Q}^i_k$ be given by (\ref{eq:dncurved}).
+Let $\bm{M}^k = \bm{V}_q^T\bm{W}\diag{\bm{J}^k}\bm{V}_q$ be the curved mass matrix on $D^k$, and let the auxiliary quantities $\tilde{\bm{u}}$ be defined by
+\begin{gather*}
+\bm{v}_q = \bm{v}\LRp{\bm{u}_q}, \qquad \tilde{\bm{v}} = \begin{bmatrix}
+\bm{V}_q\\
+\bm{V}_f
+\end{bmatrix}\bm{P}^k_q\bm{v}_q, \qquad \tilde{\bm{u}} = \bm{u}\LRp{\tilde{\bm{v}}}.
+\end{gather*}
+where $\bm{P}^k_q = \LRp{\bm{M}^k}^{-1}\bm{V}_q^T\bm{W}\diag{\bm{J}^k}$.  Then, the formulation
+\begin{gather}
+\bm{M}^k\pd{\bm{u}_h}{t} + 
+\sum_{i=1}^d \LRs{\begin{array}{cc}
+\bm{V}_q \\
+\bm{V}_f\end{array}}^T \LRp{\LRp{\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T} \circ \bm{F}_S}\bm{1} + \bm{V}_f^T\bm{W}_f \diag{{\bm{n}^k_i}}\bm{f}_i^* = 0, \label{eq:skewformcurved}\\
+%\sum_{i=1}^d \LRs{\begin{array}{cc}
+%\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{{\bm{n}_i}}\bm{f}_i^* = 0, \label{eq:skewform}\\
+\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q, \nonumber\\
+\bm{f}^* = \bm{f}_S(\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f), \qquad \text{ on interior interfaces,} \nonumber
+\end{gather}
+is semi-discretely entropy conservative on curved meshes.  
+\end{theorem}
+
+\begin{remark}
+It is possible to replace the curved mass matrix $\bm{M}^k$ with a more easily invertible weight-adjusted approximation while maintaining high order accuracy, entropy stability, and local conservation \cite{chan2018discretely}.
+\end{remark}
 

### 2018-08-30 docs/ESDG_GQ/esdg_gq.tex

@@
 \renewcommand{\L}{L^2\LRp{\Omega}}
-\newcommand{\LK}{L^2\LRp{D^k}}
-\newcommand{\LdK}{L^2\LRp{\partial D^k}}
+\newcommand{\LK}{L^2\LRp{\Omega^k}}
+\newcommand{\LdK}{L^2\LRp{\partial \Omega^k}}
 \newcommand{\Dhat}{\widehat{D}}
@@
 
-\newcommand{\LinfDk}{L^{\infty}\LRp{D^k}}
+\newcommand{\LinfDk}{L^{\infty}\LRp{\Omega^k}}
 \newcommand{\diag}[1]{{\rm diag}\LRp{#1}}
@@
 \makeatother
+
+\newcommand\myeq{\mathrel{\stackrel{\makebox[0pt]{\mbox{\normalfont\tiny def}}}{=}}}
+%\newcommand\myeq{{=}}
 
@@
 In this work, we show how to efficiently generalize the construction of semi-discretely entropy stable schemes on tensor product elements to Gauss points and generalized SBP operators.
-Numerical experiments suggest that the use of Gauss points can significant improve accuracy on curved meshes.
+Numerical experiments suggest that the use of Gauss points significantly improves accuracy on curved meshes.
 \end{abstract}
@@
 The utility of these simulations depends both on the reliability and accuracy of numerical methods for the solution of the associated nonlinear partial differential equations (PDEs).
-Time dependent nonlinear conservation laws are ubiquitous in computational fluid dynamics (CFD), for which high order methods are increasingly of interest.
+Time dependent nonlinear conservation laws are ubiquitous in computational fluid dynamics, for which high order methods are increasingly of interest.
 Such methods are more accurate per degree of freedom than low order methods, while also possessing much smaller numerical dispersion and dissipation errors.
@@
 The instability of high order methods is rooted in the fact that discretizations of nonlinear conservation laws do not typically satisfy a discrete analogue of the conservation or dissipation of energy (entropy).
-While the numerical dissipation present for low order methods serves as a stabilizing mechanism, the low of numerical dissipation in high order methods reveals the lack of an inherent statement of stability.
+For low order methods, the lack of discrete stability can be offset by the presence of numerical dissipation, which serves as a stabilization mechanism.
+Because high order methods possess low numerical dissipation, the absence of a discrete stability property becomes more noticeable, manifesting itself through increased sensitivity and instability.
+
 The dissipation of entropy serves as an energetic principle for nonlinear conservation laws \cite{dafermos2005compensated}, and requires the use of the chain rule in its proof.
-The instability of high order methods is tied to the fact that, when discretizing systems of nonlinear PDEs, the chain rule does not typically hold at the discrete level.
-
+Discrete instability is typically tied to the fact that, when discretizing systems of nonlinear PDEs, the chain rule does not typically hold at the discrete level.
 The lack of a chain rule was circumvented by using a non-standard ``flux differencing'' formulation \cite{fisher2013high, carpenter2014entropy, gassner2016split, gassner2017br1}, which is key to constructing semi-discretely entropy stable high order schemes on unstructured quadrilateral and hexahedral meshes.
@@
 Entropy stability schemes on simplicial elements require evaluating two-point fluxes between solution states at all quadrature points on an element.
-The number of quadrature points on a simplex scales as $O(N^d)$ in $d$ dimensions, and results in $O(N^{2d})$ two-point flux evaluations per element.
+For a degree $N$ approximation, the number of quadrature points on a simplex scales as $O(N^d)$ in $d$ dimensions, and results in $O(N^{2d})$ two-point flux evaluations per element.
 In contrast, entropy stable schemes on quadrilateral and hexahedral elements require only the evaluation of two-point fluxes along lines of nodes due to a tensor product structure, resulting in $O(N^{d+1})$ evaluations in $d$ dimensions.
@@
 \[
-\bm{v} = \pd{S(\bm{u})}{\bm{u}}.  
+\bm{v} \myeq \pd{S(\bm{u})}{\bm{u}}.  
 \]
@@
 \begin{gather}
-\pd{S(\bm{u})}{t} + \sum_{i=1}^d \pd{F_i(\bm{u})}{x_i} \leq 0,
+\pd{S(\bm{u})}{t} + \sum_{i=1}^d \pd{F_i(\bm{u})}{x_i} \leq 0, \qquad F_i(\bm{u}) = \bm{v}^T\pd{\bm{f}_i}{\bm{u}}, 
 \label{eq:entropyineq}
 \end{gather}
-where $F_i$ denotes the $i$th scalar entropy flux function and $\psi_i(\bm{u})$ denotes the $i$th entropy potential
-\[
-F_i(\bm{u}) = \bm{v}^T\pd{\bm{f}_i}{x_i}, \qquad \psi_i(\bm{u}) = \bm{v}^T\bm{f}_i(\bm{u}) - F_i(\bm{u}). 
-\]
-Integrating (\ref{eq:entropyineq}) over a domain $\Omega$ and applying the divergence theorem yields an integrated entropy inequality
+where $F_i$ denotes the $i$th scalar entropy flux function.  Integrating (\ref{eq:entropyineq}) over a domain $\Omega$ and applying the divergence theorem yields an integrated entropy inequality
 \begin{equation}
@@
 \end{equation}
-where $\partial \Omega$ denotes the boundary of $\Omega$ and $\bm{n}_i$ denotes the $i$th component of the outward normal on $\partial \Omega$.
+where $\psi_i(\bm{u}) = \bm{v}^T\bm{f}_i(\bm{u}) - F_i(\bm{u})$ denotes the $i$th entropy potential, $\partial \Omega$ denotes the boundary of $\Omega$ and $\bm{n}_i$ denotes the $i$th component of the outward normal on $\partial \Omega$.
 Roughly speaking, this implies that the time rate of change of entropy is less than or equal to the flux of entropy through the boundary.
@@
 This higher accuracy of Gauss quadrature has been shown to translate to lower errors and slightly improved rates of convergence in simulations of wave propagation and fluid flow \cite{kopriva2010quadrature, hindenlang2012explicit, chan2015gpu}.
-However, Gauss points have not been widely used to construct entropy stable discretizations due to the lack of efficient, stable, and high order accurate inter-element coupling terms, known as simultaneous approximation terms (SBP-SAT) in the finite difference literature \cite{fernandez2014review, crean2017high, fernandez2018simultaneous}.
-SBP-SATs for Gauss points are non-compact, in the sense that they introduce all-to-all coupling between degrees of freedom on neighboring elements.
+However, Gauss points have not been widely used to construct entropy stable discretizations due to the lack of efficient, stable, and high order accurate inter-element coupling terms, known as simultaneous approximation terms (SAT) in the finite difference literature \cite{fernandez2014review, crean2017high, fernandez2018simultaneous}.
+SATs for Gauss points are non-compact, in the sense that they introduce all-to-all coupling between degrees of freedom on neighboring elements.
 This results in greater communication between elements, as well as a significantly larger number of two-point flux evaluations and floating point operations.
 
-It is possible to realize the improved accuracy of Gauss points while avoiding non-compact SBP-SATs through a staggered grid formulation, where the solution is stored at Gauss nodes but interpolated to a set of higher degree $(N+2)$ GLL ``flux'' points for computation \cite{parsani2016entropy}.
-Because GLL nodes include boundary points, compact and high order accurate SBP-SAT terms can be constructed for the flux points.
+It is possible to realize the improved accuracy of Gauss points while avoiding non-compact SATs through a staggered grid formulation, where the solution is stored at Gauss nodes but interpolated to a set of higher degree $(N+2)$ GLL ``flux'' points for computation \cite{parsani2016entropy}.
+Because GLL nodes include boundary points, compact and high order accurate SAT terms can be constructed for the flux points.
 After performing computations on the flux points, the results are interpolated back to Gauss points and used to evolve the solution forward in time.
@@
 \subfloat[Gauss nodes]{\includegraphics[width=.3\textwidth]{figs/gsbp.png}}
-\caption{Examples of nodal sets under which efficient entropy stable schemes can be constructed.  This work focuses on the construction of efficient and accurate SBP-SAT terms for Gauss nodal sets.}
+\caption{Examples of nodal sets under which efficient entropy stable schemes can be constructed.  This work focuses on the construction of efficient and accurate SAT terms for Gauss nodal sets.}
 \label{fig:nodesets}
@@
 \[
-\bm{D}_{ij} = \LRl{\pd{\ell_j}{x}}_{x = x_i}, \qquad \LRp{\bm{V}_f}_{i1} = \ell_i(-1), \qquad  \LRp{\bm{V}_f}_{i2} = \ell_i(1).
-\]
-We also introduce the diagonal matrix of quadrature weights $\bm{W}_{ij} = \delta_{ij} w_i$.
-In this work, $\bm{W}$ is identical to the mass matrix, whose entries are evaluated as $L^2$ inner products of basis functions.
+\bm{D}_{ij} = \LRl{\pd{\ell_j}{x}}_{x = x_i}, \qquad \LRp{\bm{V}_f}_{1i} = \ell_i(-1), \qquad  \LRp{\bm{V}_f}_{2i} = \ell_i(1).
+\]
+We also introduce the diagonal matrix of quadrature weights $\bm{W}_{ij} = \delta_{ij} w_i$, as well as the one-dimensional mass matrix $\bm{M}$ whose entries are $L^2$ inner products of basis functions.
+We assume that these inner products are computed using the quadrature rule $(x_i, w_i)$ at which the solution is collocated.
+Under such an assumption, the mass matrix is diagonal with entries equal to the quadrature weights
+\[
+\bm{M}_{ij} = \int_{-1}^1 \ell_i(x)\ell_j(x) \approx \sum_{k=1}^{N+1} \ell_i(x_k)\ell_j(x_k) w_k = \delta_{ij} w_i = \bm{W}_{ij}.
+\]
+Since $\bm{M} = \bm{W}$ under a collocation assumption, we utilize $\bm{W}$ for the remainder of this work to emphasize that the mass matrix is diagonal and related to the quadrature weights $w_i$.
+The treatment of non-diagonal mass matrices is covered in \cite{chan2017discretely, chan2018discretely}.
+
+%In this work, we assume that $\bm{W}$ is identical to the mass matrix, which \note{fix and mention mass lumping} whose entries are evaluated as $L^2$ inner products of basis functions.
 This equality holds so long as the basis functions are Lagrange polynomials and integrals are approximated using quadrature (either GLL or Gauss).
@@
 %\begin{proof}
-The proof is a direct restatement of integration by parts, and can be found in \cite{ortleb2016kinetic, ortleb2017kinetic, ranocha2018generalised}.  
+The proof is a direct restatement of integration by parts, and can be found in \cite{fernandez2014generalized, ortleb2016kinetic, ortleb2017kinetic, ranocha2018generalised}.  
 %We reproduce it here for completeness.  Let $\bm{u},\bm{v}$ be two arbitrary vectors, and let $u(x), v(x)$ denote the polynomials whose nodal values are given by $\bm{u},\bm{v}$.  Then, 
@@
 
-\subsection{Existing entropy stable SBP-SATs for generalized SBP operators}
+\subsection{Existing entropy stable SATs for generalized SBP operators}
 \label{sec:gsbpsat}
@@
 
-The construction of entropy conservative schemes relies on appropriate SBP-SATs for Gauss collocation schemes \cite{fernandez2014review, crean2017high, fernandez2018simultaneous}.
-Let the rows of $\bm{V}_f$ be denoted as $\bm{t}_L, \bm{t}_R$
+The construction of entropy conservative schemes relies on appropriate SATs for Gauss collocation schemes \cite{fernandez2014review, crean2017high, fernandez2018simultaneous}.
+Let the rows of $\bm{V}_f$ be denoted by column vectors $\bm{t}_L, \bm{t}_R$
 \[
 \bm{V}_f = \begin{bmatrix}
-\horzbar & \bm{t}_L & \horzbar\\
-\horzbar & \bm{t}_R & \horzbar
+%\horzbar & \bm{t}_L^T & \horzbar\\
+%\horzbar & \bm{t}_R^T & \horzbar
+\LRp{\bm{t}_L}_1, & \ldots, & \LRp{\bm{t}_L}_{N+1}\\
+\LRp{\bm{t}_R}_1, & \ldots, & \LRp{\bm{t}_R}_{N+1}
 \end{bmatrix}, \qquad \LRp{\bm{t}_L}_j = \ell_j(-1), \qquad \LRp{\bm{t}_R}_j = \ell_j(1).
 \]
-The inter-element coupling terms in \cite{fernandez2014review, crean2017high, fernandez2018simultaneous} utilize a decomposition of the surface mass matrix $\bm{V}_f^T\bm{B}\bm{V}_f$ as  
+The inter-element coupling terms in \cite{fernandez2014review, crean2017high, fernandez2018simultaneous} utilize a decomposition of the surface matrix $\bm{V}_f^T\bm{B}\bm{V}_f$ as  
 \begin{equation}
@@
 \end{equation}
-The construction of entropy conservative schemes on multiple elements requires appropriate inter-element coupling terms (SBP-SATs) involving $\bm{t}_L, \bm{t}_R$.
-We consider a two element mesh, and show how SBP-SATs can be interpreted as constructing a global SBP operator.
+The construction of entropy conservative schemes on multiple elements requires appropriate inter-element coupling terms (SATs) involving $\bm{t}_L, \bm{t}_R$.
+We consider a two element mesh, and show how when coupled with SATs, the resulting discretization matrices can be interpreted as constructing a global SBP operator.
 
 Let $\bm{u}^1_N, \bm{u}^2_N$ denote nodal degrees of freedom of the vector valued solution $\bm{u}(x)$ on the first and second element, respectively.
-To simplify notation, we assume that all following operators are defined in terms of Kronecker products, such that they are applied to each component of $\bm{u}^1_N, \bm{u}^2_N$.
+To simplify notation, we assume that all following operators are defined in terms of Kronecker products \cite{chen2017entropy}, such that they are applied to each component of $\bm{u}^1_N, \bm{u}^2_N$.
 We first define the matrix 
 \begin{equation}
-\bm{S} = \bm{Q} - \frac{1}{2}\bm{V}_f^T\bm{B}\bm{V}_f.
+\bm{S} \myeq \bm{Q} - \frac{1}{2}\bm{V}_f^T\bm{B}\bm{V}_f.
 \label{eq:Smat}
@@
 \begin{equation}
-\bm{Q}_{h} = 
+\bm{Q}_{h} \myeq 
 \underbrace{\begin{bmatrix}
@@
 , \qquad 
-\bm{W}_h = \begin{bmatrix}\bm{W} &\\
+\bm{W}_h \myeq \begin{bmatrix}\bm{W} &\\
 & \bm{W}
@@
 \end{bmatrix} \label{eq:ESgsbp} \\
-\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\LRp{\bm{u}_h}_i,\LRp{\bm{u}_h}_j}, \qquad 1 \leq  i,j \leq 2(N+1). \nonumber
+\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\LRp{\bm{u}_h}_i,\LRp{\bm{u}_h}_j}, \qquad 1 \leq  i,j \leq 2(N+1), \nonumber
 \end{gather}
-where $\circ$ denotes the Hadamard product \cite{horn2012matrix}.  Multiplying (\ref{eq:ESgsbp}) by $\bm{v}_h^T = \bm{v}\LRp{\bm{u}_h}^T$ 
+where $\circ$ denotes the Hadamard product \cite{horn2012matrix}.
+{It should be emphasized that here, $(\bm{u}_h)_i, (\bm{u}_h)_j$ denote vectors containing solution components at nodes $i,j$, and that (because $\bm{f}_S$ is a vector-valued flux) the term $\LRp{\bm{F}_S}_{ij}$ should be interpreted as a diagonal matrix whose diagonal entries consist of the components of $\bm{f}_S\LRp{(\bm{u}_h)_i, (\bm{u}_h)_j}$.}
+
+%\note{Note that $\bm{F}_S$ is a tensor \ldots add diag notation assuming a certain ordering of degrees of freedom.}
+Multiplying (\ref{eq:ESgsbp}) by $\bm{v}_h^T = \bm{v}\LRp{\bm{u}_h}^T$ 
 %\begin{align*}
@@
 \end{equation}
-We refer to \cite{crean2017high,crean2018entropy} for the proof of (\ref{eq:consentropyGSBP}).
-
-The drawback of the SBP-SATs introduced in this section lies in the nature of the off-diagonal matrices $\bm{t}_R\bm{t}_L$ and $-\bm{t}_L\bm{t}_R$.
+We refer to \cite{crean2017high, crean2018entropy} for the proof of (\ref{eq:consentropyGSBP}).
+
+The drawback of the SATs introduced in this section lies in the nature of the off-diagonal matrices $\bm{t}_R\bm{t}_L$ and $-\bm{t}_L\bm{t}_R$.
 For Gauss nodes, these blocks are dense, which implies that inter-element coupling terms produce a non-compact stencil.
 Evaluating (\ref{eq:ESgsbp}) requires computing two-point fluxes $\bm{f}_S$ between all nodes on two neighboring elements, which significantly increases both the computational work, as well as communication between neighboring elements.
-The main goal of this work is to circumvent the all-to-all coupling of degrees of freedom introduced by the SBP-SATs described in this section, which can be done through the use of ``decoupled'' SBP operators.
+The main goal of this work is to circumvent the all-to-all coupling of degrees of freedom introduced by the SATs described in this section, which can be done through the use of ``decoupled'' SBP operators.
 
@@
 \begin{equation}
-\bm{Q}_N  = \begin{bmatrix}
+\bm{Q}_N \myeq \begin{bmatrix}
 \bm{Q} - \frac{1}{2}\bm{V}_f^T\bm{B}\bm{V}_f & \frac{1}{2}\bm{V}_f^T\bm{B}\\
@@
 
-The matrix $\bm{Q}_N$ acts not only on volume nodes, but on both volume and surface nodes.
-However, $\bm{Q}_N$ does not directly define an approximation to the derivative.
+We note that the matrix $\bm{Q}_N$ acts not only on volume nodes, but on both volume and surface nodes.
+Thus, it is not immediately clear how to apply this operator to GSBP discretizations of nonlinear conservation laws.
+It is straightforward to evaluate the nonlinear flux at volume nodes since the solution is collocated at these points; however, evaluating the nonlinear flux at surface nodes is less straightforward.
+Moreover, $\bm{Q}_N$ does not directly define a difference operator, and must be combined with another operation to produce an approximation to the derivative.
+We will discuss how to apply $\bm{Q}_N$ in two steps.
+First, we will show how to approximate the derivative of an arbitrary function using $\bm{Q}_N$ given function values at both volume and surface nodes.
+Then, we will describe how to apply this approximation to compute derivatives of nonlinear flux functions given collocated solution values at volume nodes.
+
 Let $f(x), g(x)$ denote two functions, and let $\bm{f}, \bm{g}$ denote the values of $f,g$ at interior nodal points.
@@
 
+The approximation (\ref{eq:qn}) can also be applied to Gauss collocation schemes for nonlinear conservation laws.
+Let $u \in P^N$ be represented by the vector $\bm{u}$ of values at Gauss points, and let $f(x), g(x)$ denote two nonlinear functions.
+The operator $\bm{Q}_N$ can be used to approximate the quantity $f(u)\pd{g(u)}{x}$ using (\ref{eq:qn}) if $\bm{f}, \bm{g}$ are defined as
+\[
+\bm{f}_N = \begin{bmatrix}
+f\LRp{\bm{u}_1}\\
+\vdots\\
+f\LRp{\bm{u}_{N+1}}\\
+f\LRp{\bm{t}_L^T\bm{u}}\\
+f\LRp{\bm{t}_R^T\bm{u}}
+\end{bmatrix} = \begin{bmatrix}
+f(\bm{u}) \\
+f(\bm{u}_f)
+\end{bmatrix}, \qquad
+\bm{g}_N = \begin{bmatrix}
+g\LRp{\bm{u}_1}\\
+\vdots\\
+g\LRp{\bm{u}_{N+1}}\\
+g\LRp{\bm{t}_L^T\bm{u}}\\
+g\LRp{\bm{t}_R^T\bm{u}}
+\end{bmatrix} = \begin{bmatrix}
+g(\bm{u}) \\
+g(\bm{u}_f)
+\end{bmatrix}, \qquad \bm{u}_f = \bm{V}_f\bm{u}.
+\]
+%The key observation is that (\ref{eq:qn}) utilizes new information
+
 It was shown in \cite{chan2017discretely} that $\bm{u}$ is a high order accurate approximation to the quantity $f\pd{g}{x}$.
@@
 However, when applied to non-polynomial functions, the decoupled SBP operator $\bm{Q}_N$ improves accuracy near the boundaries.
-Figure~\ref{fig:dsbpcorrect} illustrates this by using both operators to approximate the derivative of a Gaussian $e^{-4x^2}$.
-The decoupled SBP operator results in an improved approximation near the boundary.
+Figure~\ref{fig:dsbpcorrect} illustrates this by using both operators to approximate the derivative of a Gaussian $e^{-4x^2}$ on $[-1,1]$.
+The decoupled SBP operator results in an improved approximation at all orders of approximation.
 %This can be interpreted as  \cite{brenner2007mathematical}
@@
 \centering
+\subfloat[$N=5$ approximation]{
+\label{subfig:dsbp1}
 \begin{tikzpicture}
 \begin{axis}[
-    width=.6\textwidth,
+    width=.51\textwidth,
     xlabel={$x$ coordinate},
@@
 %    xmin=.0125, xmax=.75,
-    ymin=-2, ymax=2.25,
+    ymin=-2, ymax=2.5,
     legend pos=north east, legend cell align=left, legend style={font=\tiny},	
     xmajorgrids=true, ymajorgrids=true, grid style=dashed,
-    legend entries={GSBP, Decoupled SBP, Exact}    
+    legend entries={GSBP, Decoupled, Exact}    
 ]
@@
 \end{tikzpicture}
-
-\caption{Degree $N= 5$ approximations of derivatives of a Gaussian $e^{-4x^2}$ using the generalized SBP operator $\bm{D}$ and the decoupled SBP operator $\bm{Q}_N$ via (\ref{eq:qn}).
-The colored circles and squares denote values at Gauss points.
+}
+\subfloat[$L^2$ errors]{
+\label{subfig:dsbp2}
+\begin{tikzpicture}
+\begin{semilogyaxis}[
+    width=.51\textwidth,
+    xlabel={Degree $N$},
+    ylabel={$L^2$ errors}, 
+%    xmin=.0125, xmax=.75,
+%    ymin=-2, ymax=2.25,
+    legend pos=north east, legend cell align=left, legend style={font=\tiny},	
+    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
+    legend entries={GSBP, Decoupled SBP}    
+]
+\pgfplotsset{
+cycle list={{blue, mark=*}, {red, mark=square*}}
+}
+%\addlegendimage{blue, mark=*}
+%\addlegendimage{red, mark=square*}
+%\addlegendimage{black, dashed}
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,1.67087)(2,1.84788)(3,1.29689)(4,1.14406)(5,0.67189)(6,0.4428)(7,0.242532)(8,0.124111)(9,0.0657786)(10,0.0272413)(11,0.0141812)(12,0.00491087)(13,0.00252903)(14,0.000750851)(15,0.000383985)};
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,1.37796)(2,1.23235)(3,0.769706)(4,0.682916)(5,0.348496)(6,0.252957)(7,0.120617)(8,0.0691975)(9,0.0323286)(10,0.0149469)(11,0.00695358)(12,0.00266351)(13,0.00124093)(14,0.000403653)(15,0.000188708)};
+
+\end{semilogyaxis}
+\end{tikzpicture}
+
+}
+
+\caption{Approximations of derivatives of a Gaussian $e^{-4x^2}$ using the generalized SBP operator $\bm{D}$ and the decoupled SBP operator $\bm{Q}_N$ via (\ref{eq:qn}).
+In Figure~\ref{subfig:dsbp1}, the colored circles and squares denote values at Gauss points for a degree $N = 5$ approximation.
+Figure~\ref{subfig:dsbp2} shows the convergence of $L^2$ errors as $N$ increases.
 }
@@
 
-We can now construct an entropy conservative Gauss collocation scheme with compact SBP-SATs using decoupled SBP operators.
+We can now construct an entropy conservative Gauss collocation scheme with compact SATs using decoupled SBP operators.
 As in Section~\ref{sec:gsbpsat}, we will construct a Gauss collocation scheme and provide a proof of semi-discrete entropy conservation for a two-element mesh.
-
-%(see Figure~\ref{fig:twoelem})
-%\note{Add pics of two-element mesh}
 
@@
 \label{eq:et}
-\bm{B}\bm{V}_f\bm{1} = \LRp{\bm{e}_R\bm{e}_R^T - \bm{e}_L\bm{e}_L}\bm{1} = \bm{e}_R-\bm{e}_L.
+\bm{B}\bm{V}_f\bm{1} = \LRp{\bm{e}_R\bm{e}_R^T - \bm{e}_L\bm{e}_L^T}\bm{1} = \bm{e}_R-\bm{e}_L.
 \end{equation}
@@
 \begin{align}
-\bm{Q}_h =& \LRs{\begin{array}{cc|cc}
+\bm{Q}_h \myeq& \LRs{\begin{array}{cc|cc}
 \bm{S} &  \frac{1}{2}\bm{V}_f^T\bm{B} & & \\
- -\frac{1}{2}\bm{B}\bm{V}_f & -\frac{1}{2}\bm{e}_L\bm{e}_L^T & & \frac{1}{2}\bm{e}_R\bm{e}_L^T\\
- \hline
+ -\frac{1}{2}\bm{B}\bm{V}_f & -\frac{1}{2}\bm{e}_L\bm{e}_L^T & & \frac{1}{2}\bm{e}_R\bm{e}_L^T\\ [.2em]
+ \hline\\[-1.0em]
 && \bm{S} & \frac{1}{2}\bm{V}_f^T\bm{B} \\
 & -\frac{1}{2}\bm{e}_L\bm{e}_R^T & - \frac{1}{2}\bm{B}\bm{V}_f & \frac{1}{2}\bm{e}_R\bm{e}_R^T
-\end{array}} \label{eq:decoupledsbp2elem}
+\end{array}}, \label{eq:decoupledsbp2elem}
 \end{align}
-We can also show that $\bm{Q}_h\bm{1} = 0$.  Using (\ref{eq:et}), we have that
+where we have abused notation and redefined $\bm{Q}_h$.  We can also show that $\bm{Q}_h\bm{1} = 0$.  Using (\ref{eq:et}), we have that
 \begin{align}
@@
 \LRp{\bm{S} + \frac{1}{2}\bm{V}_f^T\bm{B} }\bm{1}\\
-\frac{1}{2}\bm{B}\bm{V}_f\bm{1} + \frac{1}{2}\LRp{\bm{e}_R-\bm{e}_L}\\
+-\frac{1}{2}\bm{B}\bm{V}_f\bm{1} + \frac{1}{2}\LRp{\bm{e}_R-\bm{e}_L}\\
 \LRp{\bm{S} + \frac{1}{2}\bm{V}_f^T\bm{B} }\bm{1}\\
-\frac{1}{2}\bm{B}\bm{V}_f\bm{1} + \frac{1}{2}\LRp{\bm{e}_R-\bm{e}_L}
+-\frac{1}{2}\bm{B}\bm{V}_f\bm{1} + \frac{1}{2}\LRp{\bm{e}_R-\bm{e}_L}
 \end{bmatrix} 
@@
 \bm{Q}_h = \bm{S}_h + \frac{1}{2}\bm{B}_h, \qquad
-\bm{S}_h= \LRs{\begin{array}{cc|cc}
+\bm{S}_h \myeq \LRs{\begin{array}{cc|cc}
 \bm{S} &  \frac{1}{2}\bm{V}_f^T\bm{B} & & \\
@@
 \end{array}}\\
-\bm{B}_h = \LRs{\begin{array}{cc|cc}
+\bm{B}_h \myeq \LRs{\begin{array}{cc|cc}
 &  & & \\
@@
 Note that here, $\bm{B}_h$ is diagonal, unlike the boundary operator defined in Section~\ref{sec:gsbpsat}.
-
-The two-element decoupled SBP operator $\bm{Q}_h$ in (\ref{eq:decoupledsbp2elem}) is related to the GSBP operator in (\ref{eq:gsbp}) through a block interpolation matrix
-\[
-\bm{I}_h = \begin{bmatrix}
+It should also be noted that the two-element decoupled SBP operator $\bm{Q}_h$ in (\ref{eq:decoupledsbp2elem}) is related to the GSBP operator in (\ref{eq:gsbp}) through a block interpolation matrix
+\[
+\bm{I}_h \myeq \begin{bmatrix}
 \bm{I} &\\
@@
 
+We will now utilize the two-element operators described in this section to construct an entropy stable Gauss collocation scheme on two elements.
+This scheme will differ from that of Section~\ref{sec:gsbpsat} in that neighboring elements will only be coupled together through face values.
 Let $\bm{u}^1_N, \bm{u}^2_N$ denote the values of the conservative variables at Gauss points on elements 1 and 2, respectively, and let $\bm{u}_h$ denote their concatenation as defined in (\ref{eq:ESgsbp}).
@@
 \[
-\tilde{\bm{u}}^i_f = \bm{u}(\bm{v}^i_f), \qquad \bm{v}^i_f = \bm{V}_f\bm{v}(\bm{u}^i_N), \qquad i = 1,2. 
+\tilde{\bm{u}}^i_f \myeq \bm{u}(\bm{v}^i_f), \qquad \bm{v}^i_f \myeq \bm{V}_f\bm{v}(\bm{u}^i_N), \qquad i = 1,2. 
 \]
@@
 \[
-\bm{v}_h = \bm{v}\LRp{\bm{u}_h} = \bm{v}\LRp{\begin{bmatrix}
+\bm{v}_h \myeq \bm{v}\LRp{\bm{u}_h} = \bm{v}\LRp{\begin{bmatrix}
 {\bm{u}^1_N}\\
 {\bm{u}^2_N}
-\end{bmatrix}}, \qquad \tilde{\bm{v}} = \bm{I}_h\bm{v}_h, \qquad \tilde{\bm{u}} = \bm{u}\LRp{\tilde{\bm{v}}} = 
+\end{bmatrix}}, \qquad \tilde{\bm{v}} \myeq \bm{I}_h\bm{v}_h, \qquad \tilde{\bm{u}} \myeq \bm{u}\LRp{\tilde{\bm{v}}} = 
 %\bm{u}\LRp{
@@
 \[
-\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \quad 1\leq i,j \leq 2(N+3).
+\LRp{\bm{F}_S}_{ij} \myeq \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \quad 1\leq i,j \leq 2(N+3).
 \]
@@
 \label{eq:diagFS}
-\LRp{\bm{F}_S}_{ii} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j} = \bm{f}\LRp{\tilde{\bm{u}}_i}.  
+\LRp{\bm{F}_S}_{ii} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_i} = \bm{f}\LRp{\tilde{\bm{u}}_i}.  
 \end{equation}
@@
 \label{thm:consentropy}
-Let $\bm{Q}_h$ be defined by (\ref{eq:decoupledsbp2elem}), and let $\bm{u}_h$ denote the two-element solution of the following formulation
+Let $\bm{Q}_h$ be defined by (\ref{eq:decoupledsbp2elem}), and let $\bm{u}_h$ denote the two-element solution of the following formulation:
 \begin{gather}
@@
 \begin{gather*}
-\bm{1}^T\bm{W}_h\td{S(\bm{u}_h)}{t} + \bm{1}^T\bm{B}_h\LRp{\bm{v}_f^T\bm{f}(\tilde{\bm{u}}_f) - \psi\LRp{\tilde{\bm{u}}_f}} = 0\\
-\bm{v}_f = \begin{bmatrix}
-\bm{t}_L \bm{v}\LRp{\bm{u}_N^1}\\
-\bm{t}_R \bm{v}\LRp{\bm{u}_N^2}
-\end{bmatrix}, \qquad \tilde{\bm{u}}_f = \bm{u}(\bm{v}_f).
+\bm{1}^T\bm{W}_h\td{S(\bm{u}_h)}{t} + \bm{1}^T\bm{B}\LRp{\bm{v}_f^T\bm{f}(\tilde{\bm{u}}_f) - \psi\LRp{\tilde{\bm{u}}_f}} = 0\\
+\bm{v}_f \myeq \begin{bmatrix}
+\bm{t}_L^T \bm{v}\LRp{\bm{u}_N^1}\\
+\bm{t}_R^T \bm{v}\LRp{\bm{u}_N^2}
+\end{bmatrix}, \qquad \tilde{\bm{u}}_f \myeq \bm{u}(\bm{v}_f).
 \end{gather*}
@@
 \[
-\tilde{\bm{v}}^T\LRp{\bm{B}_h \circ \bm{F}_S}\bm{1} = \bm{1}^T\bm{B}_h\tilde{\bm{v}}_f^T\bm{f}\LRp{\tilde{\bm{u}}},
+\tilde{\bm{v}}^T\LRp{\bm{B}_h \circ \bm{F}_S}\bm{1} = \bm{1}^T\bm{B}_h\tilde{\bm{v}}^T\bm{f}\LRp{\tilde{\bm{u}}}= \bm{1}^T\bm{B}\tilde{\bm{v}}_f^T\bm{f}\LRp{\tilde{\bm{u}_f}},
 \]
 where we have used (\ref{eq:diagFS}) and the fact that $\bm{B}_h$ is diagonal.
-Here, $\tilde{\bm{v}}_f^T\bm{f}\LRp{\tilde{\bm{u}}}$ denotes a vector whose entries are the inner product of $\tilde{\bm{v}}$ and $\bm{f}\LRp{\tilde{\bm{u}}}$ at face points.
+Here, $\tilde{\bm{v}}^T\bm{f}\LRp{\tilde{\bm{u}}}$, $\tilde{\bm{v}}_f^T\bm{f}\LRp{\tilde{\bm{u}}_f}$ denote vectors whose entries are the vector inner products of components of $\tilde{\bm{v}}$ and $\bm{f}\LRp{\tilde{\bm{u}}}$ at volume and face points, respectively.
 
@@
 &= \sum_{ij} \LRp{\bm{S}_h}_{ij} \LRp{\psi\LRp{\tilde{\bm{u}}_i} - \psi\LRp{\tilde{\bm{u}}_j}}\\
-&= \psi\LRp{\tilde{\bm{u}}}^T \bm{S}_h \bm{1}- \bm{1}^T \bm{S}_h \psi\LRp{\tilde{\bm{u}}} = 2\LRp{ \psi\LRp{\tilde{\bm{u}}}^T\bm{S}_h\bm{1}}.
+&= \psi\LRp{\tilde{\bm{u}}}^T \bm{S}_h \bm{1}- \bm{1}^T \bm{S}_h \psi\LRp{\tilde{\bm{u}}} = 2\LRp{ \psi\LRp{\tilde{\bm{u}}}^T\bm{S}_h\bm{1}},
 \end{align*}
 where we have again used the skew-symmetry of $\bm{S}_h$ in the last step.  Substituting $\bm{S}_h = \bm{Q}_h - \frac{1}{2}\bm{B}_h$ and using (\ref{eq:Q1zero}) yields
-\[
-2\LRp{ \psi\LRp{\tilde{\bm{u}}}^T\bm{S}_h\bm{1}} = \psi\LRp{\tilde{\bm{u}}}^T\LRp{2\bm{Q}_h - \bm{B}_h} \bm{1} =-\psi\LRp{\tilde{\bm{u}}}^T\bm{B}_h \bm{1}  = -\bm{1}^T\bm{B}_h \psi\LRp{\tilde{\bm{u}}},
-\]
-where we have used the symmetry of $\bm{B}_h$ in the last step.
+\begin{align*} 
+2\LRp{ \psi\LRp{\tilde{\bm{u}}}^T\bm{S}_h\bm{1}} &= \psi\LRp{\tilde{\bm{u}}}^T\LRp{2\bm{Q}_h - \bm{B}_h} \bm{1} \\
+&=-\psi\LRp{\tilde{\bm{u}}}^T\bm{B}_h \bm{1}  = -\bm{1}^T\bm{B}_h \psi\LRp{\tilde{\bm{u}}} =  -\bm{1}^T\bm{B} \psi\LRp{\tilde{\bm{u}}_f}, 
+\end{align*} 
+where we have used the symmetry of $\bm{B}_h$ in the second to last step.
 \end{proof}
@@
 The formulation in Theorem~\ref{thm:consentropy} can be naturally extended to Cartesian meshes in higher dimensions through a tensor product construction.
-We consider first the construction of higher dimensional differentiation matrices on a two-dimensional reference element $\hat{D}$, assuming a two dimensional tensor product grid of quadrature nodes (the construction of decoupled SBP operators in three dimensions is straightforward and similar to the two-dimensional case).
-We then construct physical differentiation matrices on mapped elements $D^k$, through which we construct an entropy conservative scheme.
+We first consider the construction of higher dimensional differentiation matrices on a two-dimensional reference element $\hat{\Omega}$, assuming a two dimensional tensor product grid of quadrature nodes (the construction of decoupled SBP operators in three dimensions is straightforward and similar to the two-dimensional case).
+We then construct physical differentiation matrices on mapped elements $\Omega^k$, through which we construct an entropy conservative scheme.
 
@@
 \[
-{\bm{D}}^1 = \bm{D}_{\rm 1D} \otimes \bm{I}_{N+1}, \qquad {\bm{D}}^2 = \bm{I}_{N+1} \otimes \bm{D}_{\rm 1D}, \qquad {\bm{W}} = \bm{W}_{\rm 1D} \otimes  \bm{W}_{\rm 1D}.
+{\bm{D}}^1 \myeq \bm{D}_{\rm 1D} \otimes \bm{I}_{N+1}, \qquad {\bm{D}}^2  \myeq \bm{I}_{N+1} \otimes \bm{D}_{\rm 1D}, \qquad {\bm{W}} \myeq \bm{W}_{\rm 1D} \otimes  \bm{W}_{\rm 1D}, 
 \]
@@
 \[
-\bm{V}_f = \begin{bmatrix}
+\bm{V}_f \myeq \begin{bmatrix}
 \bm{V}_{f, {\rm 1D}} \otimes \bm{I}_{2}\\
@@
 \end{bmatrix}, \qquad 
-{\bm{B}}^1 = \begin{bmatrix}
+{\bm{B}}^1 \myeq \begin{bmatrix}
 \bm{B}_{\rm 1D} \otimes \bm{I}_{2} & \\
@@
 \end{bmatrix}, \qquad 
-{\bm{B}}^2 = \begin{bmatrix}
+{\bm{B}}^2 \myeq \begin{bmatrix}
 \bm{0} &\\
@@
 \[
-{\bm{Q}}^i_N = \begin{bmatrix}
+{\bm{Q}}^i_N \myeq \begin{bmatrix}
 {\bm{Q}}^i - \frac{1}{2}\bm{V}_f^T{\bm{B}}^i\bm{V}_f & \frac{1}{2}\bm{V}_f^T{\bm{B}}^i\\
 -\frac{1}{2}{\bm{B}}^i\bm{V}_f & \frac{1}{2}{\bm{B}}^i
-\end{bmatrix}, \qquad {\bm{Q}}^i = {\bm{W}}{\bm{D}}^i.
-\]
-
-Let the domain now be decomposed into non-overlapping elements $D^k$, such that $D^k$ is the image of $\hat{D}$ under a degree $N$ polynomial mapping $\bm{\Phi}^k$.
-We define geometric terms ${G}^k_{ij}$ as scaled derivatives of reference coordinates $\hat{\bm{x}}$ with respect to physical coordinates $\bm{x}$
-\[
-{G}^k_{ij} = J^k\pd{\hat{x}_j}{\bm{x}_j},
-\]
-where $J^k$ is the determinant of the Jacobian of $\bm{\Phi}^k$.
-We also introduce the face Jacobian $J^k_f$, which is computed from the mapping of a physical face to a reference face.
-We define matrices over each physical element $D^k$
+\end{bmatrix}, \qquad {\bm{Q}}^i \myeq {\bm{W}}{\bm{D}}^i.
+\]
+
+Let the domain now be decomposed into non-overlapping elements $\Omega^k$, such that $\Omega^k$ is the image of $\hat{\Omega}$ under a degree $N$ polynomial mapping $\bm{\Phi}^k$.
+We define geometric terms ${G}^k_{ij} = J^k\pd{\hat{x}_j}{\bm{x}_i}$ as scaled derivatives of reference coordinates $\hat{\bm{x}}$ with respect to physical coordinates $\bm{x}$.
+We also introduce the scaled normals $n_iJ^k_f$, which can be computed on quadrilateral and tensor product elements via
+\[
+n_i J^k_f =\pm \sum_{j=1}^d G^k_{ij} = \pm \sum_{j=1}^d  J^k\pd{\hat{x}_j}{\bm{x}_i},
+\]
+where the sign of $n_i J^k_f$ is negative for a ``left'' face and positive for a ``right'' face.  
+These geometric terms introduce scalings $J^k, J^k_f$, where $J^k$ is the determinant of the Jacobian of $\bm{\Phi}^k$ and $J^k_f$ denotes the determinant of the Jacobian of the mapping from a physical face to a reference face.
+The quantities $G^k_{ij}, n_iJ^k_f$ can now be used to define matrices over each physical element $\Omega^k$
 \begin{gather}
-\bm{W}_k = \bm{W} \diag{\bm{J}^k}, \qquad \bm{B}^i_k = \bm{B}^i_N \diag{\bm{n}_i \circ\bm{J}^f } \nonumber\\
-\bm{Q}^i_k = \frac{1}{2} \sum_{j=1}^d \LRp{\diag{{\bm{G}}^k_{ij}}\bm{Q}^i_N  + \bm{Q}^i_N \diag{{\bm{G}}^k_{ij}} } \label{eq:splitcurv},
+\bm{W}_k \myeq \bm{W} \diag{\bm{J}^k}, \qquad \bm{B}^i_k \myeq \bm{B}^i \diag{\bm{n}^k_i }, \nonumber\\
+\bm{Q}^i_k \myeq \frac{1}{2} \sum_{j=1}^d \LRp{\diag{{\bm{G}}^k_{ij}}\bm{Q}^i_N  + \bm{Q}^i_N \diag{{\bm{G}}^k_{ij}} } \label{eq:splitcurv},
 \end{gather}
-where we have discretized the curved differentiation matrix in split form \cite{nordstrom2006conservative}.  
-Here, $\bm{J}^k$ denotes the vector of values of $J^k$ at volume quadrature points, ${\bm{G}}^k_{ij}$ denotes the vector of values of ${G}^k_{ij}$ at volume and face quadrature points, and $\bm{n}^k_i, \bm{J}^k_f$ denote the $i$th outward unit normal and $J^k_f$ at face quadrature points.
-Then, an entropy conservative scheme in two dimensions is given as follows:
+where we have discretized the curved differentiation matrix in split form \cite{nordstrom2006conservative, kopriva2016geometry}.  
+Here, $\bm{J}^k$ denotes the vector of values of $J^k$ at volume quadrature points, ${\bm{G}}^k_{ij}$ denotes the vector of values of ${G}^k_{ij}$ at volume and face quadrature points, and $\bm{n}^k_i$ denotes the $i$th scaled outward normal $n_i J^k_f$ at face quadrature points.
+
+A 2D Gauss collocation scheme can now be given in terms of $\bm{W}_k, \bm{Q}^i_k, \bm{B}^i_k$.   Let $\bm{u}^k_N$ denote the vector of solution values on $\Omega^k$, and define the quantities 
+\begin{gather*}
+%{\bm{W}_k\td{\bm{u}^k_N}{t} + \sum_{i=1}^d \LRp{2\LRp{\bm{Q}^i_k\circ\bm{F}^i_S}\bm{1} + \bm{V}_f^T\bm{B}_k^i\LRp{\bm{f}^i_S\LRp{{\tilde{\bm{u}}_f}^+,\tilde{\bm{u}}^k_f}-\bm{f}^i\LRp{\tilde{\bm{u}}_f}}}} = 0\\
+\LRp{\bm{F}^i_S}_{mn} \myeq \bm{f}^i_S\LRp{\tilde{\bm{u}}_m,\tilde{\bm{u}}_n}, \qquad 1 \leq m,n \leq (N+1)^2 + 4(N+1) \\
+\tilde{\bm{u}} \myeq \begin{bmatrix}
+\bm{u}^k_N\\
+\tilde{\bm{u}}^k_f
+\end{bmatrix}, \qquad \tilde{\bm{u}}^k_f \myeq \bm{u}\LRp{\bm{v}^k_f}, \qquad 
+\bm{v}^k_f \myeq {\bm{V}_f\bm{v}\LRp{\bm{u}^k_N}},
+\end{gather*}
+where $(N+1)^2 + 4(N+1)$ is the total number of volume and face points.  
+We then have the following theorem on the semi-discrete conservation of entropy:
 \begin{theorem}
 \label{thm:esdg2d}
-Assume that $\bm{Q}^k\bm{1} = 0$, and that $D^k$ are mapped curvilinear elements.  Let $\bm{u}^k_N$ denote the solution on $D^k$ satisfying
-\begin{gather*}
-\sum_{k}\LRp{\bm{W}_k\td{\bm{u}^k_N}{t} + \sum_{i=1}^d \LRp{2\LRp{\bm{Q}^i_k\circ\bm{F}_S} + \bm{V}_f^T\bm{B}_k^i\bm{f}_S\LRp{{\tilde{\bm{u}}_f}^+,\tilde{\bm{u}}^k_f}}} = 0\\
-\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad \tilde{\bm{u}} = \begin{bmatrix}
-\bm{u}^k_N\\
-\tilde{\bm{u}}^k_f
-\end{bmatrix}, \qquad \tilde{\bm{u}}^k_f = \bm{u}\LRp{\bm{v}^k_f}, \qquad 
-\bm{v}^k_f = {\bm{V}_f\bm{v}\LRp{\bm{u}^k_N}},
-\end{gather*}
+Assume that $\bm{Q}^k\bm{1} = 0$, and that $\Omega^k$ are mapped curvilinear elements.  Let $\bm{u}^k_N$ satisfy the local formulation on $\Omega^k$
+\begin{equation}
+{\bm{W}_k\td{\bm{u}^k_N}{t} + \sum_{i=1}^d \LRp{2\LRp{\bm{Q}^i_k\circ\bm{F}^i_S}\bm{1} + \bm{V}_f^T\bm{B}_k^i\LRp{\bm{f}^i_S\LRp{{\tilde{\bm{u}}_f}^+,\tilde{\bm{u}}^k_f}-\bm{f}^i\LRp{\tilde{\bm{u}}_f}}}} = 0,
+\end{equation}
 where $\tilde{\bm{u}}^+$ denotes boundary values of the entropy-projected conservative variables on neighboring elements.  Then, $\bm{u}^k_N$ satisfies the discrete conservation of entropy 
 \[
-\sum_{k} \LRp{\bm{1}^T\bm{W}\td{\bm{u}^k_N}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}^i\LRp{\bm{v}_f^T\bm{f}(\tilde{\bm{u}}_f) - \psi\LRp{\tilde{\bm{u}}_f}}} = 0.
+\sum_{k} \LRp{\bm{1}^T\bm{W}_k\td{S\LRp{\bm{u}^k_N}}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}^i_k\LRp{\LRp{\bm{v}^k_f}^T\bm{f}^i(\tilde{\bm{u}}^k_f) - \psi\LRp{\tilde{\bm{u}}^k_f}}} = 0.
 \]
@@
 The operator $\bm{Q}^i_k$  in (\ref{eq:splitcurv}) can be applied without needing to explicitly store geometric terms $G^k_{ij}$ at face quadrature points.
-By using the structure of the boundary matrix $\bm{B}^i$, expressions involving geometric terms on faces can be replaced by expressions involving components of unit outward normals $\bm{n}_i$ on $D^k$.
+By using the structure of the boundary matrix $\bm{B}^i$, expressions involving geometric terms on faces can be replaced by expressions involving components of scaled outward normals ${n}_i J^k_f$ on $\Omega^k$.
 \end{remark}
+
+\subsection{Discrete geometric conservation law}
 
@@
 Let $N_{\rm geo}$ denote the degree of a polynomial geometric mapping.
-In three dimensions, the exact geometric terms for a degree $N_{\rm geo}$ polynomial mapping are polynomials of degree $(2N_{\rm geo} - 2)$ \cite{kopriva2006metric, hindenlang2012explicit, crean2018entropy}.
-If $(2N_{\rm geo}-2) \leq N$, or if $N_{\rm geo} \leq \left\lfloor\frac{N}{2}\right\rfloor + 1$, then the discrete GCL is automatically satisfied.
-
-For $N_{\rm geo} \geq \left\lfloor\frac{N}{2}\right\rfloor + 1$, modifications to the computation of geometric terms are required to ensure that the GCL is satisfied at the discrete level.
+In three dimensions, the exact geometric terms for a degree $N_{\rm geo}$ polynomial mapping are polynomials of degree $2N_{\rm geo}$ \cite{kopriva2006metric, hindenlang2012explicit, crean2018entropy}.
+If $2N_{\rm geo} \leq N$, or if $N_{\rm geo} \leq \left\lfloor\frac{N}{2}\right\rfloor$, then the discrete GCL is automatically satisfied.\footnote{We note that this condition is specific to three-dimensional hexahedral elements, and does not necessarily hold for other element types.
+For example, the discrete GCL is satisfied if $2N_{\rm geo}-2 \leq N$ for isoparametric tetrahedral elements \cite{chan2018discretely}.}
+
+For $N_{\rm geo} \geq \left\lfloor\frac{N}{2}\right\rfloor$, modifications to the computation of geometric terms are required to ensure that the GCL is satisfied at the discrete level.
 For general SBP operators, the discrete GCL can be enforced through the solution of a local least squares problem \cite{crean2018entropy}.
@@
 \begin{equation}
-\label{eq:conscurl}
 \begin{bmatrix}
@@
 \end{bmatrix}, \qquad j = 1,2,3.
+\label{eq:conscurl}
 \end{equation}
@@
 However, there is only a single unique entropy which appropriately treats the viscous heat conduction term in the compressible Navier-Stokes equations \cite{hughes1986new}.
-This entropy $S(\bm{u})$ is given in $d$ dimensions by
+This entropy $S(\bm{u})$ is given by
 \begin{equation*}
@@
 \end{equation*}
-where $s = \log\LRp{\frac{p}{\rho^\gamma}}$ is the physical specific entropy.  The entropy variables in $d$ dimensions are given by
+where $s = \log\LRp{\frac{p}{\rho^\gamma}}$ is the physical specific entropy, and the dimension $d = 1,2,3$.  The entropy variables in $d$ dimensions are given by
 \begin{align*}
 v_1 &= \frac{\rho e (\gamma + 1 - s) - E}{\rho e}, \\
-v_{1+i}&= \frac{\rho \bm{u}_i}{\rho e}, \qquad i = 1,\ldots, d,\\
+v_{1+ i}&= \frac{\rho \bm{u}_i}{\rho e}, \qquad i = 1,\ldots, d,\\
 v_{d+2} &= -\frac{\rho}{\rho e},
@@
 (E_{\rm avg}+ p_{\rm avg})\avg{w}\\
-\end{array}}.\nonumber
+\end{array}},\nonumber
 %&f^1_{1,S}(\bm{u}_L,\bm{u}_R) = \avg{\rho}^{\log} \avg{u},& &f^1_{2,S}(\bm{u}_L,\bm{u}_R) = \avg{\rho}^{\log} \avg{v},&\\
@@
 \begin{align*}
-p_{\rm avg} &= \frac{\avg{\rho}}{2\avg{\beta}}, \qquad E_{\rm avg} = \frac{\avg{\rho}^{\log}}{2(\gamma-1)\avg{\beta}^{\log}} + \frac{1}{2}\avg{\rho}^{\log}\nor{\bm{u}}^2_{\rm avg}\\
+p_{\rm avg} &= \frac{\avg{\rho}}{2\avg{\beta}}, \qquad E_{\rm avg} = \frac{\avg{\rho}^{\log}}{2(\gamma-1)\avg{\beta}^{\log}} + \frac{1}{2}\avg{\rho}^{\log}\nor{\bm{u}}^2_{\rm avg}, \\
 \nor{\bm{u}}^2_{\rm avg} &= 2(\avg{u}^2 + \avg{v}^2 + \avg{w}^2) - \LRp{\avg{u^2} +\avg{v^2} + \avg{w^2}}.\nonumber
@@
 
-In all problems, we estimate the timestep size $dt$ using $J, J^f$, and degree-dependent $L^2$ trace constants $C_N$ 
-\[
-dt = C_{\rm CFL}\frac{ h}{C_N}, \qquad h = \frac{1}{{\nor{J^{-1}}_{L^{\infty}}}\nor{J^f}_{L^{\infty}}},
-\]
-where $C_{\rm CFL}$ is some user-defined CFL constant.
-For isotropic elements, the ratio of $J$ to $J^f$ scales as the mesh size $h$, while $C_N$ captures the dependence of the maximum stable timestep on the polynomial degree $N$.
+In all problems, we estimate the timestep size $dt$ using $J, J^k_f$, and degree-dependent $L^2$ trace constants $C_N$ 
+\[
+dt = C_{\rm CFL}\frac{ h}{ a C_N}, \qquad h = \frac{1}{{\nor{J^{-1}}_{L^{\infty}}}\nor{J^k_f}_{L^{\infty}}},
+\]
+where $a$ is an estimate of the maximum wave speed, $h$ estimates the mesh size, and $C_{\rm CFL}$ is some user-defined CFL constant.
+For isotropic elements, the ratio of $J^k$ to $J^k_f$ scales as the mesh size $h$, while $C_N$ captures the dependence of the maximum stable timestep on the polynomial degree $N$.
 For hexahedral elements, $C_N$ varies depending on the choice of quadrature.
@@
 \rho(\bm{x},t) &= \LRp{1 - \frac{\frac{1}{2}(\gamma-1)(\beta e^{1-r(\bm{x},t)^2})^2}{8\gamma \pi^2}}^{\frac{1}{\gamma-1}}, \qquad p = \rho^{\gamma},\\
-u(\bm{x},t) &= 1 - \frac{\beta}{2\pi} e^{1-r(\bm{x},t)^2}(y-y_0), \qquad v(\bm{x},t) = \frac{\beta}{2\pi} e^{1-r(\bm{x},t)^2}(y-y_0),\nonumber
+u(\bm{x},t) &= 1 - \frac{\beta}{2\pi} e^{1-r(\bm{x},t)^2}(y-y_0), \qquad v(\bm{x},t) = \frac{\beta}{2\pi} e^{1-r(\bm{x},t)^2}(x-x_0-t),\nonumber
 \end{align}
@@
 \begin{align*}
-\tilde{x} &= x + \alpha\cos\LRp{\frac{\pi}{20}(x-10)}\cos\LRp{\frac{3\pi}{10}y}\\
+\tilde{x} &= x + \alpha\cos\LRp{\frac{\pi}{20}(x-10)}\cos\LRp{\frac{3\pi}{10}y}, \\
 \tilde{y} &= y + \alpha\sin\LRp{\frac{\pi}{5}(\tilde{x}-10)}\cos\LRp{\frac{\pi}{10}y}.
@@
 
-We note that $L^2$ approximation estimates on curved meshes \cite{lenoir1986optimal, bernardi1989optimal, warburton2013low} assume that the mesh size is small enough to be in the asymptotic regime (such that asymptotic error estimates hold).
+We note that $L^2$ approximation estimates on curved meshes \cite{lenoir1986optimal, warburton2013low} assume that the mesh size is small enough to be in the asymptotic regime (such that asymptotic error estimates hold).
 On curved meshes, this requires that the mesh is sufficiently fine to resolve both the solution and geometric mapping.
-The numerical results in Figure~\ref{fig:err2d} suggest that, compared to Gauss collocation schemes, under-integrated GLL collocation schemes may require a much finer mesh resolution to reach the asymptotic regime.
-
-%\note{Investigate robustness with respect to mesh distortion: GLL vs GQ errors.
-GQ is better for more curved meshes.
-Note that there is little difference between bilinear/trilinear and high order curved.
-}
+The results in Figure~\ref{fig:err2d} suggest that, compared to Gauss collocation schemes, under-integrated GLL collocation schemes require a finer mesh resolution to reach the asymptotic regime.
 
@@
 We also compared $L^2$ errors for both isoparametric and sub-parametric geometric mappings.
+In both cases, the discrete GCL is satisfied.
 For sub-parametric mappings, we chose the degree of approximation of geometry $N_{\rm geo} = \left\lfloor\frac{N}{2} \right\rfloor + 1$, such that the geometric terms are computed exactly and the GCL is satisfied by default.
@@
 \end{align*}
-where $\bar{a}, \bar{u}_n$ are defined as
+where $\bar{a}, \bar{u}_n$ are defined in 2D as
 \begin{gather*}
@@
 %It is known that the Lax-Friedrichs flux is fairly dissipative compared to more nuanced fluxes \cite{harten1983upstream}, though this dissipation is less pronounced at higher polynomial degrees in practice.
-Figure~\ref{fig:shockvort} shows solutions for with $N=4$ and $h = 1/100$ Gauss collocation schemes using a non-dissipative entropy conservative flux, a dissipative Lax-Friedrichs flux, and a matrix dissipation flux.
+Figure~\ref{fig:shockvort} shows solutions for $N=4$ and $h = 1/100$ Gauss collocation schemes using a non-dissipative entropy conservative flux, a dissipative Lax-Friedrichs flux, and a matrix dissipation flux.
 In all cases, the vortex passes through the shock stably without the use of additional slope limiting, filtering, or artificial viscosity.
@@
 This problem is posed on the periodic box $[-\pi,\pi]^3$, with initial conditions
-\begin{align*}
-\rho &= 1\\
-u &= \sin(x_1)\cos(x_2)\cos(x_3),\\
-v &= -\cos(x_1)\sin(x_2)\cos(x_3),\\ 
-w &= 0,\\
-p &= \frac{100}{\gamma} + \frac{1}{16} \LRp{\cos(2x_1) + \cos(2x_2)}\LRp{2+\cos(2x_3)}.
-\end{align*}
-The Taylor--Green vortex is used to study the transition and decay of turbulence \cite{debonis2013solutions}.
+\begin{gather*}
+\rho = 1, \qquad p = \frac{100}{\gamma} + \frac{1}{16} \LRp{\cos(2x_1) + \cos(2x_2)}\LRp{2+\cos(2x_3)},\\
+u = \sin(x_1)\cos(x_2)\cos(x_3), \qquad
+v = -\cos(x_1)\sin(x_2)\cos(x_3),\qquad
+w= 0.
+\end{gather*}
+The Taylor--Green vortex is used to study the transition and decay of turbulence.
 In the absence of viscosity, the Taylor--Green vortex develops successively smaller scales as time increases.
 As a result, the solution is guaranteed to contain under-resolved features after a sufficiently large time.
-We study the evolution of the kinetic energy $\kappa(t)$
+We study the evolution of kinetic energy $\kappa(t)$
 \[
@@
 as well as the kinetic energy dissipation rate $-\pd{\kappa}{t}$, which is approximated by differencing $\kappa(t)$.
+For both GLL and Gauss collocation schemes, integrals in the kinetic energy formula are evaluated using an $(N+1)$-point Gauss quadrature rule.
 
@@
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0,-0)(0.1,-0)(0.2,-1.1e-05)(0.3,-1e-05)(0.4,-1.4e-05)(0.5,-1.6e-05)(0.6,-2.5e-05)(0.7,-1.5e-05)(0.8,-2.8e-05)(0.9,-2.4e-05)(1,-2.6e-05)(1.1,-2.4e-05)(1.2,-3.2e-05)(1.3,-2.1e-05)(1.4,-2.5e-05)(1.5,-2.6e-05)(1.6,-2.1e-05)(1.7,-1.9e-05)(1.8,-2.6e-05)(1.9,-1.1e-05)(2,-2.5e-05)(2.1,-8e-06)(2.2,-2.5e-05)(2.3,-3e-06)(2.4,-2.3e-05)(2.5,-3e-06)(2.6,-1.7e-05)(2.7,-5e-06)(2.8,-1.2e-05)(2.9,-8e-06)(3,1e-06)(3.1,-1.4e-05)(3.2,1.7e-05)(3.3,2e-06)(3.4,3.8e-05)(3.5,5.3e-05)(3.6,0.000105)(3.7,0.000139)(3.8,0.000221)(3.9,0.000276)(4,0.000342)(4.1,0.000419)(4.2,0.000521)(4.3,0.000678)(4.4,0.000837)(4.5,0.000981)(4.6,0.001062)(4.7,0.001087)(4.8,0.00113)(4.9,0.001166)(5,0.001202)(5.1,0.001253)(5.2,0.001359)(5.3,0.001478)(5.4,0.001732)(5.5,0.002089)(5.6,0.002263)(5.7,0.002498)(5.8,0.002994)(5.9,0.003125)(6,0.003437)(6.1,0.004011)(6.2,0.004722)(6.3,0.005376)(6.4,0.006272)(6.5,0.007188)(6.6,0.00777)(6.7,0.007294)(6.8,0.008198)(6.9,0.009458)(7,0.009693)(7.1,0.011262)(7.2,0.010684)(7.3,0.011183)(7.4,0.012658)(7.5,0.013435)(7.6,0.012556)(7.7,0.012895)(7.8,0.012983)(7.9,0.011737)(8,0.011435)(8.1,0.01177)(8.2,0.011534)(8.3,0.011911)(8.4,0.011895)(8.5,0.011791)(8.6,0.011377)(8.7,0.009956)(8.8,0.010736)(8.9,0.011693)(9,0.009853)(9.1,0.009799)(9.2,0.009839)(9.3,0.010534)(9.4,0.010571)(9.5,0.009947)(9.6,0.010718)(9.7,0.011076)(9.8,0.01093)(9.9,0.011329)(10,0.011703)(10.1,0.01164)(10.2,0.012807)(10.3,0.012246)(10.4,0.010467)(10.5,0.010364)(10.6,0.010194)(10.7,0.008977)(10.8,0.009781)(10.9,0.010854)(11,0.010576)(11.1,0.009041)(11.2,0.008563)(11.3,0.007785)(11.4,0.007869)(11.5,0.008274)(11.6,0.008207)(11.7,0.007814)(11.8,0.007557)(11.9,0.00751)(12,0.0082)(12.1,0.007652)(12.2,0.007354)(12.3,0.007167)(12.4,0.007458)(12.5,0.007237)(12.6,0.007162)(12.7,0.006955)(12.8,0.006604)(12.9,0.006469)(13,0.006581)(13.1,0.006469)(13.2,0.00626)(13.3,0.006146)(13.4,0.006075)(13.5,0.00601)(13.6,0.005883)(13.7,0.005825)(13.8,0.005701)(13.9,0.005585)(14,0.005407)(14.1,0.005251)(14.2,0.005187)(14.3,0.005117)(14.4,0.005088)(14.5,0.00498)(14.6,0.004898)(14.7,0.004772)(14.8,0.004717)(14.9,0.004651)(15,0.004603)(15.1,0.004514)(15.2,0.004408)(15.3,0.004296)(15.4,0.004199)(15.5,0.004117)(15.6,0.004069)(15.7,0.003965)(15.8,0.003854)(15.9,0.003777)(16,0.003716)(16.1,0.003633)(16.2,0.003617)(16.3,0.003594)(16.4,0.003547)(16.5,0.003478)(16.6,0.003385)(16.7,0.003317)(16.8,0.003279)(16.9,0.003207)(17,0.003162)(17.1,0.003144)(17.2,0.003097)(17.3,0.00307)(17.4,0.003008)(17.5,0.002939)(17.6,0.00292)(17.7,0.002873)(17.8,0.002805)(17.9,0.002742)(18,0.00271)(18.1,0.00266)(18.2,0.002629)(18.3,0.002588)(18.4,0.002547)(18.5,0.002515)(18.6,0.002468)(18.7,0.002415)(18.8,0.002386)(18.9,0.002366)(19,0.002338)(19.1,0.002301)(19.2,0.002265)(19.3,0.002243)(19.4,0.002204)(19.5,0.002174)(19.6,0.002139)(19.7,0.002107)(19.8,0.002075)(19.9,0.002052)(20,0.00202)};
-
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0,-0)(0.1,-0)(0.2,-1.1e-05)(0.3,-1e-05)(0.4,-1.3e-05)(0.5,-1.6e-05)(0.6,-2.5e-05)(0.7,-1.6e-05)(0.8,-2.8e-05)(0.9,-2.5e-05)(1,-2.5e-05)(1.1,-2.4e-05)(1.2,-3.1e-05)(1.3,-2.2e-05)(1.4,-2.4e-05)(1.5,-2.7e-05)(1.6,-2.1e-05)(1.7,-1.9e-05)(1.8,-2.6e-05)(1.9,-1.2e-05)(2,-2.5e-05)(2.1,-8e-06)(2.2,-2.6e-05)(2.3,-2e-06)(2.4,-2.4e-05)(2.5,-3e-06)(2.6,-1.8e-05)(2.7,-4e-06)(2.8,-1.5e-05)(2.9,-5e-06)(3,-3e-06)(3.1,-1.4e-05)(3.2,1.1e-05)(3.3,-5e-06)(3.4,2.5e-05)(3.5,3.2e-05)(3.6,7.1e-05)(3.7,9e-05)(3.8,0.000156)(3.9,0.000201)(4,0.000279)(4.1,0.000375)(4.2,0.00051)(4.3,0.000691)(4.4,0.000864)(4.5,0.001022)(4.6,0.00114)(4.7,0.001224)(4.8,0.001302)(4.9,0.001404)(5,0.001526)(5.1,0.001685)(5.2,0.001869)(5.3,0.002032)(5.4,0.002409)(5.5,0.002752)(5.6,0.003089)(5.7,0.003588)(5.8,0.004093)(5.9,0.004445)(6,0.004825)(6.1,0.005375)(6.2,0.005868)(6.3,0.006183)(6.4,0.006331)(6.5,0.006797)(6.6,0.007712)(6.7,0.008319)(6.8,0.00896)(6.9,0.009326)(7,0.009777)(7.1,0.010393)(7.2,0.010403)(7.3,0.010421)(7.4,0.010724)(7.5,0.011246)(7.6,0.011484)(7.7,0.011616)(7.8,0.012531)(7.9,0.012689)(8,0.012545)(8.1,0.012473)(8.2,0.012667)(8.3,0.012824)(8.4,0.012941)(8.5,0.0127)(8.6,0.012743)(8.7,0.012718)(8.8,0.012434)(8.9,0.011995)(9,0.011842)(9.1,0.01197)(9.2,0.011787)(9.3,0.011592)(9.4,0.011345)(9.5,0.010925)(9.6,0.011148)(9.7,0.01112)(9.8,0.010876)(9.9,0.010962)(10,0.010975)(10.1,0.010899)(10.2,0.010607)(10.3,0.010373)(10.4,0.01008)(10.5,0.009831)(10.6,0.009418)(10.7,0.009322)(10.8,0.009141)(10.9,0.009067)(11,0.009091)(11.1,0.008885)(11.2,0.008732)(11.3,0.008681)(11.4,0.008587)(11.5,0.008343)(11.6,0.008185)(11.7,0.008042)(11.8,0.008009)(11.9,0.007849)(12,0.007513)(12.1,0.007208)(12.2,0.007146)(12.3,0.007062)(12.4,0.007009)(12.5,0.006919)(12.6,0.006761)(12.7,0.006594)(12.8,0.006558)(12.9,0.00642)(13,0.00625)(13.1,0.006121)(13.2,0.006015)(13.3,0.00599)(13.4,0.00586)(13.5,0.005745)(13.6,0.005636)(13.7,0.005559)(13.8,0.005431)(13.9,0.005396)(14,0.005265)(14.1,0.0052)(14.2,0.005097)(14.3,0.004995)(14.4,0.004901)(14.5,0.004814)(14.6,0.0047)(14.7,0.004622)(14.8,0.004574)(14.9,0.004471)(15,0.004354)(15.1,0.004272)(15.2,0.00426)(15.3,0.004171)(15.4,0.004091)(15.5,0.004049)(15.6,0.003979)(15.7,0.003897)(15.8,0.003842)(15.9,0.003795)(16,0.003726)(16.1,0.003677)(16.2,0.003605)(16.3,0.003528)(16.4,0.003477)(16.5,0.003454)(16.6,0.003389)(16.7,0.003346)(16.8,0.003298)(16.9,0.003229)(17,0.003182)(17.1,0.003154)(17.2,0.003085)(17.3,0.003019)(17.4,0.002948)(17.5,0.002911)(17.6,0.002874)(17.7,0.002836)(17.8,0.002812)(17.9,0.002746)(18,0.0027)(18.1,0.002667)(18.2,0.002635)(18.3,0.002613)(18.4,0.002582)(18.5,0.002542)(18.6,0.002495)(18.7,0.002458)(18.8,0.002435)(18.9,0.002405)(19,0.002376)(19.1,0.002342)(19.2,0.002314)(19.3,0.002278)(19.4,0.00225)(19.5,0.00221)(19.6,0.002181)(19.7,0.002155)(19.8,0.002114)(19.9,0.002085)(20,0.002057)};
+coordinates{(0,-0)(0.02,-0)(0.04,-1e-06)(0.06,-2e-06)(0.08,-1e-06)(0.1,-0)(0.12,-0)(0.14,-2e-06)(0.16,-5e-06)(0.18,-8e-06)(0.2,-1.1e-05)(0.22,-1.3e-05)(0.24,-1.5e-05)(0.26,-1.5e-05)(0.28,-1.3e-05)(0.3,-1e-05)(0.32,-8e-06)(0.34,-8e-06)(0.36,-9e-06)(0.38,-1.1e-05)(0.4,-1.4e-05)(0.42,-1.7e-05)(0.44,-2e-05)(0.46,-2e-05)(0.48,-1.8e-05)(0.5,-1.6e-05)(0.52,-1.6e-05)(0.54,-1.7e-05)(0.56,-1.9e-05)(0.58,-2.2e-05)(0.6,-2.5e-05)(0.62,-2.7e-05)(0.64,-2.7e-05)(0.66,-2.3e-05)(0.68,-1.9e-05)(0.7,-1.5e-05)(0.72,-1.4e-05)(0.74,-1.5e-05)(0.76,-1.8e-05)(0.78,-2.3e-05)(0.8,-2.8e-05)(0.82,-3.2e-05)(0.84,-3.3e-05)(0.86,-3.1e-05)(0.88,-2.7e-05)(0.9,-2.4e-05)(0.92,-2.2e-05)(0.94,-2e-05)(0.96,-2e-05)(0.98,-2.2e-05)(1,-2.6e-05)(1.02,-2.9e-05)(1.04,-2.9e-05)(1.06,-2.8e-05)(1.08,-2.6e-05)(1.1,-2.4e-05)(1.12,-2.2e-05)(1.14,-2.2e-05)(1.16,-2.4e-05)(1.18,-2.8e-05)(1.2,-3.2e-05)(1.22,-3.2e-05)(1.24,-3e-05)(1.26,-2.7e-05)(1.28,-2.4e-05)(1.3,-2.1e-05)(1.32,-2e-05)(1.34,-2e-05)(1.36,-2.2e-05)(1.38,-2.3e-05)(1.4,-2.5e-05)(1.42,-2.6e-05)(1.44,-2.6e-05)(1.46,-2.7e-05)(1.48,-2.7e-05)(1.5,-2.6e-05)(1.52,-2.5e-05)(1.54,-2.3e-05)(1.56,-2.1e-05)(1.58,-2.1e-05)(1.6,-2.1e-05)(1.62,-2.1e-05)(1.64,-2.2e-05)(1.66,-2.1e-05)(1.68,-2e-05)(1.7,-1.9e-05)(1.72,-1.8e-05)(1.74,-1.8e-05)(1.76,-2e-05)(1.78,-2.4e-05)(1.8,-2.6e-05)(1.82,-2.7e-05)(1.84,-2.4e-05)(1.86,-2e-05)(1.88,-1.5e-05)(1.9,-1.1e-05)(1.92,-1.1e-05)(1.94,-1.4e-05)(1.96,-1.9e-05)(1.98,-2.3e-05)(2,-2.5e-05)(2.02,-2.3e-05)(2.04,-1.8e-05)(2.06,-1.2e-05)(2.08,-8e-06)(2.1,-8e-06)(2.12,-1.2e-05)(2.14,-1.8e-05)(2.16,-2.4e-05)(2.18,-2.7e-05)(2.2,-2.5e-05)(2.22,-1.9e-05)(2.24,-1.1e-05)(2.26,-5e-06)(2.28,-2e-06)(2.3,-3e-06)(2.32,-8e-06)(2.34,-1.5e-05)(2.36,-2.2e-05)(2.38,-2.5e-05)(2.4,-2.3e-05)(2.42,-1.8e-05)(2.44,-1.2e-05)(2.46,-6e-06)(2.48,-3e-06)(2.5,-3e-06)(2.52,-8e-06)(2.54,-1.5e-05)(2.56,-1.9e-05)(2.58,-2e-05)(2.6,-1.6e-05)(2.62,-1.1e-05)(2.64,-6e-06)(2.66,-2e-06)(2.68,-1e-06)(2.7,-5e-06)(2.72,-1.2e-05)(2.74,-1.9e-05)(2.76,-2.1e-05)(2.78,-1.8e-05)(2.8,-1.2e-05)(2.82,-5e-06)(2.84,1e-06)(2.86,3e-06)(2.88,0)(2.9,-7e-06)(2.92,-1.4e-05)(2.94,-1.6e-05)(2.96,-1.3e-05)(2.98,-6e-06)(3,2e-06)(3.02,7e-06)(3.04,8e-06)(3.06,3e-06)(3.08,-4e-06)(3.1,-1.1e-05)(3.12,-1.2e-05)(3.14,-6e-06)(3.16,5e-06)(3.18,1.6e-05)(3.2,2.2e-05)(3.22,2.3e-05)(3.24,2e-05)(3.26,1.5e-05)(3.28,1.1e-05)(3.3,1.2e-05)(3.32,1.8e-05)(3.34,2.9e-05)(3.36,3.9e-05)(3.38,4.7e-05)(3.4,5.3e-05)(3.42,5.7e-05)(3.44,6.1e-05)(3.46,6.5e-05)(3.48,6.9e-05)(3.5,7.5e-05)(3.52,8.4e-05)(3.54,9.5e-05)(3.56,0.000108)(3.58,0.000123)(3.6,0.000137)(3.62,0.000151)(3.64,0.00016)(3.66,0.000167)(3.68,0.000175)(3.7,0.000188)(3.72,0.000207)(3.74,0.00023)(3.76,0.000254)(3.78,0.000276)(3.8,0.000293)(3.82,0.000307)(3.84,0.00032)(3.86,0.000335)(3.88,0.000356)(3.9,0.000382)(3.92,0.00041)(3.94,0.000434)(3.96,0.000454)(3.98,0.000471)(4,0.000488)(4.02,0.000506)(4.04,0.000527)(4.06,0.00055)(4.08,0.000576)(4.1,0.000602)(4.12,0.000627)(4.14,0.000652)(4.16,0.000678)(4.18,0.000709)(4.2,0.000743)(4.22,0.00078)(4.24,0.000818)(4.26,0.000859)(4.28,0.000902)(4.3,0.000945)(4.32,0.000985)(4.34,0.001023)(4.36,0.001058)(4.38,0.001092)(4.4,0.001123)(4.42,0.001152)(4.44,0.001183)(4.46,0.001212)(4.48,0.001239)(4.5,0.001262)(4.52,0.001281)(4.54,0.001301)(4.56,0.001326)(4.58,0.001354)(4.6,0.001382)(4.62,0.001409)(4.64,0.001432)(4.66,0.001451)(4.68,0.00147)(4.7,0.001492)(4.72,0.001521)(4.74,0.001559)(4.76,0.001603)(4.78,0.001651)(4.8,0.001698)(4.82,0.001746)(4.84,0.001792)(4.86,0.001836)(4.88,0.001873)(4.9,0.001904)(4.92,0.001928)(4.94,0.001948)(4.96,0.001965)(4.98,0.001981)(5,0.001996)(5.02,0.002011)(5.04,0.002019)(5.06,0.002018)(5.08,0.002007)(5.1,0.00199)(5.12,0.001975)(5.14,0.001965)(5.16,0.00196)(5.18,0.001957)(5.2,0.001953)(5.22,0.001946)(5.24,0.001939)(5.26,0.001939)(5.28,0.001956)(5.3,0.001996)(5.32,0.002061)(5.34,0.002145)(5.36,0.00224)(5.38,0.002341)(5.4,0.002443)(5.42,0.002543)(5.44,0.002638)(5.46,0.002726)(5.48,0.002809)(5.5,0.002885)(5.52,0.002949)(5.54,0.002996)(5.56,0.003018)(5.58,0.003014)(5.6,0.002992)(5.62,0.002973)(5.64,0.002982)(5.66,0.003043)(5.68,0.00317)(5.7,0.003359)(5.72,0.003591)(5.74,0.003839)(5.76,0.004073)(5.78,0.004266)(5.8,0.004396)(5.82,0.004441)(5.84,0.004401)(5.86,0.004294)(5.88,0.00415)(5.9,0.003999)(5.92,0.003873)(5.94,0.003811)(5.96,0.003847)(5.98,0.003993)(6,0.004232)(6.02,0.004527)(6.04,0.004833)(6.06,0.005103)(6.08,0.005334)(6.1,0.005569)(6.12,0.005837)(6.14,0.006136)(6.16,0.006463)(6.18,0.006803)(6.2,0.007136)(6.22,0.00744)(6.24,0.007702)(6.26,0.007926)(6.28,0.008125)(6.3,0.008313)(6.32,0.008495)(6.34,0.008648)(6.36,0.00871)(6.38,0.008605)(6.4,0.008321)(6.42,0.007965)(6.44,0.007715)(6.46,0.007709)(6.48,0.007986)(6.5,0.008491)(6.52,0.009104)(6.54,0.009684)(6.56,0.010117)(6.58,0.010317)(6.6,0.010234)(6.62,0.009882)(6.64,0.009345)(6.66,0.008749)(6.68,0.008205)(6.7,0.007768)(6.72,0.007448)(6.74,0.007279)(6.76,0.007351)(6.78,0.007768)(6.8,0.008528)(6.82,0.009469)(6.84,0.010375)(6.86,0.011063)(6.88,0.011398)(6.9,0.011387)(6.92,0.01123)(6.94,0.011075)(6.96,0.01089)(6.98,0.010733)(7,0.010765)(7.02,0.01106)(7.04,0.011626)(7.06,0.012321)(7.08,0.012964)(7.1,0.013498)(7.12,0.013808)(7.14,0.013787)(7.16,0.013581)(7.18,0.013369)(7.2,0.013109)(7.22,0.012843)(7.24,0.012799)(7.26,0.01303)(7.28,0.013428)(7.3,0.013887)(7.32,0.014317)(7.34,0.014632)(7.36,0.014756)(7.38,0.014686)(7.4,0.014579)(7.42,0.014635)(7.44,0.014862)(7.46,0.015073)(7.48,0.015116)(7.5,0.014986)(7.52,0.014736)(7.54,0.014511)(7.56,0.014461)(7.58,0.014532)(7.6,0.014612)(7.62,0.014665)(7.64,0.014642)(7.66,0.014475)(7.68,0.014125)(7.7,0.013612)(7.72,0.012984)(7.74,0.012235)(7.76,0.01157)(7.78,0.011386)(7.8,0.01153)(7.82,0.011749)(7.84,0.012047)(7.86,0.012372)(7.88,0.012527)(7.9,0.012431)(7.92,0.012218)(7.94,0.012038)(7.96,0.01189)(7.98,0.011736)(8,0.011595)(8.02,0.011484)(8.04,0.01137)(8.06,0.011233)(8.08,0.011115)(8.1,0.011065)(8.12,0.011145)(8.14,0.011418)(8.16,0.011912)(8.18,0.012598)(8.2,0.013376)(8.22,0.014119)(8.24,0.014773)(8.26,0.01528)(8.28,0.015472)(8.3,0.015214)(8.32,0.014377)(8.34,0.013059)(8.36,0.01173)(8.38,0.010901)(8.4,0.010437)(8.42,0.01036)(8.44,0.010646)(8.46,0.010979)(8.48,0.011134)(8.5,0.011135)(8.52,0.011176)(8.54,0.011402)(8.56,0.011752)(8.58,0.012005)(8.6,0.011974)(8.62,0.011609)(8.64,0.011016)(8.66,0.010389)(8.68,0.009869)(8.7,0.009497)(8.72,0.009264)(8.74,0.009179)(8.76,0.00931)(8.78,0.009735)(8.8,0.010461)(8.82,0.011337)(8.84,0.012075)(8.86,0.012488)(8.88,0.01273)(8.9,0.012901)(8.92,0.012632)(8.94,0.011704)(8.96,0.010529)(8.98,0.009686)(9,0.009349)(9.02,0.009326)(9.04,0.009464)(9.06,0.009729)(9.08,0.010032)(9.1,0.010342)(9.12,0.010679)(9.14,0.011074)(9.16,0.011527)(9.18,0.01185)(9.2,0.011981)(9.22,0.011963)(9.24,0.011825)(9.26,0.011552)(9.28,0.011139)(9.3,0.010649)(9.32,0.010249)(9.34,0.010109)(9.36,0.010196)(9.38,0.010281)(9.4,0.01023)(9.42,0.010108)(9.44,0.010037)(9.46,0.010002)(9.48,0.009945)(9.5,0.009917)(9.52,0.009937)(9.54,0.01007)(9.56,0.010359)(9.58,0.010654)(9.6,0.010845)(9.62,0.010977)(9.64,0.011131)(9.66,0.011326)(9.68,0.011527)(9.7,0.011704)(9.72,0.011897)(9.74,0.012042)(9.76,0.01182)(9.78,0.011497)(9.8,0.011604)(9.82,0.011925)(9.84,0.012054)(9.86,0.011902)(9.88,0.011722)(9.9,0.01176)(9.92,0.011688)(9.94,0.011222)(9.96,0.010615)(9.98,0.010046)(10,0.009708)(10.02,0.009592)(10.04,0.009614)(10.06,0.0098)(10.08,0.010226)(10.1,0.01091)(10.12,0.011762)(10.14,0.012632)(10.16,0.013402)(10.18,0.014051)(10.2,0.014566)(10.22,0.014773)(10.24,0.014611)(10.26,0.014166)(10.28,0.013528)(10.3,0.012797)(10.32,0.012079)(10.34,0.01142)(10.36,0.010847)(10.38,0.010396)(10.4,0.010075)(10.42,0.009875)(10.44,0.009809)(10.46,0.009902)(10.48,0.010097)(10.5,0.010291)(10.52,0.010446)(10.54,0.01056)(10.56,0.010619)(10.58,0.010602)(10.6,0.010464)(10.62,0.010122)(10.64,0.009546)(10.66,0.008828)(10.68,0.00822)(10.7,0.008002)(10.72,0.008277)(10.74,0.008912)(10.76,0.009681)(10.78,0.010442)(10.8,0.011071)(10.82,0.011359)(10.84,0.011321)(10.86,0.011296)(10.88,0.011403)(10.9,0.011317)(10.92,0.010943)(10.94,0.010665)(10.96,0.010806)(10.98,0.011148)(11,0.01127)(11.02,0.011025)(11.04,0.010503)(11.06,0.009812)(11.08,0.00844)(11.1,0.007333)(11.12,0.00716)(11.14,0.007428)(11.16,0.007757)(11.18,0.008093)(11.2,0.008383)(11.22,0.008495)(11.24,0.008377)(11.26,0.008048)(11.28,0.007577)(11.3,0.007108)(11.32,0.006797)(11.34,0.006703)(11.36,0.006759)(11.38,0.00683)(11.4,0.006833)(11.42,0.006775)(11.44,0.00672)(11.46,0.006754)(11.48,0.006958)(11.5,0.007347)(11.52,0.007844)(11.54,0.008334)(11.56,0.008709)(11.58,0.008896)(11.6,0.008851)(11.62,0.008624)(11.64,0.0083)(11.66,0.007893)(11.68,0.007493)(11.7,0.007164)(11.72,0.006927)(11.74,0.006763)(11.76,0.006643)(11.78,0.006571)(11.8,0.006539)(11.82,0.006512)(11.84,0.006463)(11.86,0.006386)(11.88,0.006291)(11.9,0.006217)(11.92,0.006222)(11.94,0.006345)(11.96,0.006593)(11.98,0.006922)(12,0.007269)(12.02,0.007568)(12.04,0.007771)(12.06,0.007867)(12.08,0.007873)(12.1,0.007762)(12.12,0.007555)(12.14,0.007355)(12.16,0.007185)(12.18,0.007101)(12.2,0.007112)(12.22,0.007139)(12.24,0.007146)(12.26,0.007131)(12.28,0.007078)(12.3,0.006993)(12.32,0.00691)(12.34,0.006855)(12.36,0.006819)(12.38,0.006825)(12.4,0.006894)(12.42,0.007001)(12.44,0.007116)(12.46,0.007196)(12.48,0.007199)(12.5,0.007131)(12.52,0.007041)(12.54,0.00696)(12.56,0.006897)(12.58,0.006861)(12.6,0.006834)(12.62,0.00681)(12.64,0.006783)(12.66,0.006736)(12.68,0.006668)(12.7,0.006582)(12.72,0.006488)(12.74,0.006406)(12.76,0.00634)(12.78,0.006278)(12.8,0.006213)(12.82,0.006165)(12.84,0.006164)(12.86,0.006201)(12.88,0.006247)(12.9,0.006271)(12.92,0.006281)(12.94,0.0063)(12.96,0.006331)(12.98,0.006345)(13,0.006334)(13.02,0.006316)(13.04,0.006303)(13.06,0.006297)(13.08,0.006299)(13.1,0.006311)(13.12,0.006319)(13.14,0.006317)(13.16,0.006296)(13.18,0.006245)(13.2,0.006177)(13.22,0.006111)(13.24,0.006052)(13.26,0.005995)(13.28,0.005946)(13.3,0.005916)(13.32,0.005906)(13.34,0.0059)(13.36,0.005889)(13.38,0.005874)(13.4,0.005855)(13.42,0.005836)(13.44,0.005821)(13.46,0.005804)(13.48,0.005784)(13.5,0.005765)(13.52,0.005748)(13.54,0.005726)(13.56,0.005707)(13.58,0.005675)(13.6,0.00562)(13.62,0.005591)(13.64,0.005579)(13.66,0.005573)(13.68,0.00556)(13.7,0.00553)(13.72,0.005485)(13.74,0.005437)(13.76,0.005391)(13.78,0.005346)(13.8,0.005305)(13.82,0.00527)(13.84,0.005242)(13.86,0.005223)(13.88,0.005218)(13.9,0.005224)(13.92,0.005237)(13.94,0.00525)(13.96,0.005256)(13.98,0.005251)(14,0.005236)(14.02,0.005214)(14.04,0.005181)(14.06,0.005133)(14.08,0.005074)(14.1,0.005016)(14.12,0.004962)(14.14,0.004918)(14.16,0.004881)(14.18,0.004843)(14.2,0.004804)(14.22,0.004764)(14.24,0.004727)(14.26,0.004696)(14.28,0.004675)(14.3,0.004665)(14.32,0.004663)(14.34,0.004661)(14.36,0.004655)(14.38,0.004644)(14.4,0.004626)(14.42,0.004606)(14.44,0.004589)(14.46,0.004577)(14.48,0.00457)(14.5,0.004566)(14.52,0.00457)(14.54,0.004578)(14.56,0.004585)(14.58,0.004589)(14.6,0.004585)(14.62,0.004569)(14.64,0.004545)(14.66,0.004519)(14.68,0.004498)(14.7,0.004486)(14.72,0.004479)(14.74,0.004472)(14.76,0.00446)(14.78,0.004439)(14.8,0.004406)(14.82,0.004366)(14.84,0.004326)(14.86,0.004295)(14.88,0.004274)(14.9,0.004262)(14.92,0.004253)(14.94,0.004243)(14.96,0.004236)(14.98,0.004231)(15,0.004229)(15.02,0.004229)(15.04,0.004229)(15.06,0.004228)(15.08,0.004224)(15.1,0.004209)(15.12,0.004177)(15.14,0.004134)(15.16,0.004084)(15.18,0.004035)(15.2,0.003992)(15.22,0.003961)(15.24,0.003942)(15.26,0.003937)(15.28,0.003942)(15.3,0.003955)(15.32,0.003972)(15.34,0.003988)(15.36,0.003998)(15.38,0.004001)(15.4,0.003988)(15.42,0.003954)(15.44,0.003908)(15.46,0.003863)(15.48,0.003819)(15.5,0.003774)(15.52,0.003733)(15.54,0.003696)(15.56,0.00366)(15.58,0.00363)(15.6,0.003607)(15.62,0.003594)(15.64,0.003588)(15.66,0.003588)(15.68,0.003592)(15.7,0.003599)(15.72,0.003607)(15.74,0.003618)(15.76,0.003629)(15.78,0.003637)(15.8,0.003646)(15.82,0.003657)(15.84,0.003667)(15.86,0.003672)(15.88,0.003669)(15.9,0.003656)(15.92,0.003637)(15.94,0.003611)(15.96,0.003581)(15.98,0.00355)(16,0.00352)(16.02,0.00349)(16.04,0.003458)(16.06,0.003425)(16.08,0.003395)(16.1,0.003373)(16.12,0.003359)(16.14,0.003351)(16.16,0.003345)(16.18,0.003343)(16.2,0.003341)(16.22,0.003337)(16.24,0.00333)(16.26,0.003321)(16.28,0.003311)(16.3,0.003304)(16.32,0.003299)(16.34,0.003293)(16.36,0.003287)(16.38,0.003277)(16.4,0.003266)(16.42,0.003253)(16.44,0.003234)(16.46,0.003213)(16.48,0.003194)(16.5,0.003182)(16.52,0.003173)(16.54,0.003168)(16.56,0.003166)(16.58,0.003164)(16.6,0.003158)(16.62,0.003145)(16.64,0.003125)(16.66,0.003099)(16.68,0.003069)(16.7,0.003038)(16.72,0.003009)(16.74,0.002984)(16.76,0.002961)(16.78,0.002938)(16.8,0.002917)(16.82,0.002895)(16.84,0.002874)(16.86,0.002856)(16.88,0.002842)(16.9,0.002831)(16.92,0.002822)(16.94,0.002817)(16.96,0.002816)(16.98,0.002821)(17,0.00283)(17.02,0.00284)(17.04,0.002851)(17.06,0.002861)(17.08,0.002868)(17.1,0.00287)(17.12,0.002867)(17.14,0.00286)(17.16,0.002848)(17.18,0.002836)(17.2,0.002828)(17.22,0.002824)(17.24,0.002822)(17.26,0.002825)(17.28,0.002831)(17.3,0.002835)(17.32,0.002836)(17.34,0.00283)(17.36,0.002817)(17.38,0.002798)(17.4,0.002776)(17.42,0.002754)(17.44,0.002733)(17.46,0.002714)(17.48,0.002698)(17.5,0.002685)(17.52,0.002674)(17.54,0.002664)(17.56,0.002652)(17.58,0.002641)(17.6,0.002629)(17.62,0.002617)(17.64,0.002607)(17.66,0.002597)(17.68,0.002584)(17.7,0.002569)(17.72,0.002553)(17.74,0.002539)(17.76,0.002525)(17.78,0.002513)(17.8,0.002503)(17.82,0.002498)(17.84,0.002499)(17.86,0.002501)(17.88,0.002505)(17.9,0.002508)(17.92,0.002512)(17.94,0.002511)(17.96,0.002509)(17.98,0.002508)(18,0.002508)(18.02,0.002507)(18.04,0.002505)(18.06,0.002496)(18.08,0.002483)(18.1,0.002469)(18.12,0.002455)(18.14,0.002442)(18.16,0.002428)(18.18,0.002412)(18.2,0.002398)(18.22,0.002387)(18.24,0.002375)(18.26,0.002362)(18.28,0.00235)(18.3,0.002341)(18.32,0.002331)(18.34,0.002319)(18.36,0.002308)(18.38,0.0023)(18.4,0.002297)(18.42,0.002293)(18.44,0.002287)(18.46,0.002283)(18.48,0.002279)(18.5,0.002274)(18.52,0.002266)(18.54,0.002258)(18.56,0.00225)(18.58,0.002244)(18.6,0.00224)(18.62,0.002236)(18.64,0.002231)(18.66,0.002228)(18.68,0.002225)(18.7,0.002222)(18.72,0.002216)(18.74,0.002211)(18.76,0.002208)(18.78,0.002206)(18.8,0.002203)(18.82,0.002194)(18.84,0.002181)(18.86,0.002168)(18.88,0.002159)(18.9,0.002154)(18.92,0.002153)(18.94,0.002155)(18.96,0.002157)(18.98,0.002158)(19,0.002158)(19.02,0.002156)(19.04,0.002153)(19.06,0.00215)(19.08,0.002146)(19.1,0.002141)(19.12,0.002135)(19.14,0.002129)(19.16,0.002123)(19.18,0.002115)(19.2,0.002105)(19.22,0.002094)(19.24,0.002085)(19.26,0.002078)(19.28,0.00207)(19.3,0.002063)(19.32,0.002058)(19.34,0.002052)(19.36,0.002046)(19.38,0.002039)(19.4,0.002031)(19.42,0.002023)(19.44,0.002014)(19.46,0.002004)(19.48,0.001992)(19.5,0.001978)(19.52,0.001963)(19.54,0.001951)(19.56,0.00194)(19.58,0.001931)(19.6,0.001924)(19.62,0.00192)(19.64,0.001919)(19.66,0.001919)(19.68,0.001919)(19.7,0.001921)(19.72,0.001925)(19.74,0.001928)(19.76,0.001928)(19.78,0.001925)(19.8,0.001919)(19.82,0.001912)(19.84,0.001905)(19.86,0.001898)(19.88,0.001892)(19.9,0.001886)(19.92,0.001881)(19.94,0.001877)(19.96,0.001873)(19.98,0.001867)(20,0.001861)};
+
+%coordinates{(0,-0)(0.1,-0)(0.2,-1.1e-05)(0.3,-1e-05)(0.4,-1.4e-05)(0.5,-1.6e-05)(0.6,-2.5e-05)(0.7,-1.5e-05)(0.8,-2.8e-05)(0.9,-2.4e-05)(1,-2.6e-05)(1.1,-2.4e-05)(1.2,-3.2e-05)(1.3,-2.1e-05)(1.4,-2.5e-05)(1.5,-2.6e-05)(1.6,-2.1e-05)(1.7,-1.9e-05)(1.8,-2.6e-05)(1.9,-1.1e-05)(2,-2.5e-05)(2.1,-8e-06)(2.2,-2.5e-05)(2.3,-3e-06)(2.4,-2.3e-05)(2.5,-3e-06)(2.6,-1.7e-05)(2.7,-5e-06)(2.8,-1.2e-05)(2.9,-8e-06)(3,1e-06)(3.1,-1.4e-05)(3.2,1.7e-05)(3.3,2e-06)(3.4,3.8e-05)(3.5,5.3e-05)(3.6,0.000105)(3.7,0.000139)(3.8,0.000221)(3.9,0.000276)(4,0.000342)(4.1,0.000419)(4.2,0.000521)(4.3,0.000678)(4.4,0.000837)(4.5,0.000981)(4.6,0.001062)(4.7,0.001087)(4.8,0.00113)(4.9,0.001166)(5,0.001202)(5.1,0.001253)(5.2,0.001359)(5.3,0.001478)(5.4,0.001732)(5.5,0.002089)(5.6,0.002263)(5.7,0.002498)(5.8,0.002994)(5.9,0.003125)(6,0.003437)(6.1,0.004011)(6.2,0.004722)(6.3,0.005376)(6.4,0.006272)(6.5,0.007188)(6.6,0.00777)(6.7,0.007294)(6.8,0.008198)(6.9,0.009458)(7,0.009693)(7.1,0.011262)(7.2,0.010684)(7.3,0.011183)(7.4,0.012658)(7.5,0.013435)(7.6,0.012556)(7.7,0.012895)(7.8,0.012983)(7.9,0.011737)(8,0.011435)(8.1,0.01177)(8.2,0.011534)(8.3,0.011911)(8.4,0.011895)(8.5,0.011791)(8.6,0.011377)(8.7,0.009956)(8.8,0.010736)(8.9,0.011693)(9,0.009853)(9.1,0.009799)(9.2,0.009839)(9.3,0.010534)(9.4,0.010571)(9.5,0.009947)(9.6,0.010718)(9.7,0.011076)(9.8,0.01093)(9.9,0.011329)(10,0.011703)(10.1,0.01164)(10.2,0.012807)(10.3,0.012246)(10.4,0.010467)(10.5,0.010364)(10.6,0.010194)(10.7,0.008977)(10.8,0.009781)(10.9,0.010854)(11,0.010576)(11.1,0.009041)(11.2,0.008563)(11.3,0.007785)(11.4,0.007869)(11.5,0.008274)(11.6,0.008207)(11.7,0.007814)(11.8,0.007557)(11.9,0.00751)(12,0.0082)(12.1,0.007652)(12.2,0.007354)(12.3,0.007167)(12.4,0.007458)(12.5,0.007237)(12.6,0.007162)(12.7,0.006955)(12.8,0.006604)(12.9,0.006469)(13,0.006581)(13.1,0.006469)(13.2,0.00626)(13.3,0.006146)(13.4,0.006075)(13.5,0.00601)(13.6,0.005883)(13.7,0.005825)(13.8,0.005701)(13.9,0.005585)(14,0.005407)(14.1,0.005251)(14.2,0.005187)(14.3,0.005117)(14.4,0.005088)(14.5,0.00498)(14.6,0.004898)(14.7,0.004772)(14.8,0.004717)(14.9,0.004651)(15,0.004603)(15.1,0.004514)(15.2,0.004408)(15.3,0.004296)(15.4,0.004199)(15.5,0.004117)(15.6,0.004069)(15.7,0.003965)(15.8,0.003854)(15.9,0.003777)(16,0.003716)(16.1,0.003633)(16.2,0.003617)(16.3,0.003594)(16.4,0.003547)(16.5,0.003478)(16.6,0.003385)(16.7,0.003317)(16.8,0.003279)(16.9,0.003207)(17,0.003162)(17.1,0.003144)(17.2,0.003097)(17.3,0.00307)(17.4,0.003008)(17.5,0.002939)(17.6,0.00292)(17.7,0.002873)(17.8,0.002805)(17.9,0.002742)(18,0.00271)(18.1,0.00266)(18.2,0.002629)(18.3,0.002588)(18.4,0.002547)(18.5,0.002515)(18.6,0.002468)(18.7,0.002415)(18.8,0.002386)(18.9,0.002366)(19,0.002338)(19.1,0.002301)(19.2,0.002265)(19.3,0.002243)(19.4,0.002204)(19.5,0.002174)(19.6,0.002139)(19.7,0.002107)(19.8,0.002075)(19.9,0.002052)(20,0.00202)};
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0,-0)(0.0199601,-0)(0.0399202,-1e-06)(0.0598802,-2e-06)(0.0798403,-2e-06)(0.0998004,-0)(0.11976,-0)(0.139721,-2e-06)(0.159681,-5e-06)(0.179641,-8e-06)(0.199601,-1.1e-05)(0.219561,-1.3e-05)(0.239521,-1.5e-05)(0.259481,-1.6e-05)(0.279441,-1.4e-05)(0.299401,-1.1e-05)(0.319361,-8e-06)(0.339321,-8e-06)(0.359281,-9e-06)(0.379242,-1.1e-05)(0.399202,-1.4e-05)(0.419162,-1.7e-05)(0.439122,-2e-05)(0.459082,-2e-05)(0.479042,-1.9e-05)(0.499002,-1.7e-05)(0.518962,-1.6e-05)(0.538922,-1.7e-05)(0.558882,-1.9e-05)(0.578842,-2.2e-05)(0.598802,-2.6e-05)(0.618762,-2.8e-05)(0.638723,-2.8e-05)(0.658683,-2.5e-05)(0.678643,-2e-05)(0.698603,-1.7e-05)(0.718563,-1.6e-05)(0.738523,-1.6e-05)(0.758483,-1.9e-05)(0.778443,-2.4e-05)(0.798403,-3e-05)(0.818363,-3.4e-05)(0.838323,-3.5e-05)(0.858283,-3.3e-05)(0.878244,-2.9e-05)(0.898204,-2.6e-05)(0.918164,-2.4e-05)(0.938124,-2.2e-05)(0.958084,-2.1e-05)(0.978044,-2.4e-05)(0.998004,-2.7e-05)(1.01796,-3e-05)(1.03792,-3.1e-05)(1.05788,-3e-05)(1.07784,-2.8e-05)(1.0978,-2.6e-05)(1.11776,-2.4e-05)(1.13772,-2.4e-05)(1.15768,-2.6e-05)(1.17764,-3e-05)(1.1976,-3.3e-05)(1.21756,-3.4e-05)(1.23752,-3.2e-05)(1.25749,-2.9e-05)(1.27745,-2.6e-05)(1.29741,-2.3e-05)(1.31737,-2.2e-05)(1.33733,-2.2e-05)(1.35729,-2.3e-05)(1.37725,-2.5e-05)(1.39721,-2.6e-05)(1.41717,-2.7e-05)(1.43713,-2.8e-05)(1.45709,-2.8e-05)(1.47705,-2.8e-05)(1.49701,-2.8e-05)(1.51697,-2.6e-05)(1.53693,-2.4e-05)(1.55689,-2.3e-05)(1.57685,-2.2e-05)(1.59681,-2.2e-05)(1.61677,-2.2e-05)(1.63673,-2.2e-05)(1.65669,-2.2e-05)(1.67665,-2.1e-05)(1.69661,-1.9e-05)(1.71657,-1.8e-05)(1.73653,-1.8e-05)(1.75649,-2e-05)(1.77645,-2.3e-05)(1.79641,-2.6e-05)(1.81637,-2.6e-05)(1.83633,-2.4e-05)(1.85629,-2e-05)(1.87625,-1.5e-05)(1.89621,-1.1e-05)(1.91617,-1e-05)(1.93613,-1.2e-05)(1.95609,-1.6e-05)(1.97605,-2.1e-05)(1.99601,-2.4e-05)(2.01597,-2.2e-05)(2.03593,-1.7e-05)(2.05589,-1.1e-05)(2.07585,-7e-06)(2.09581,-5e-06)(2.11577,-8e-06)(2.13573,-1.4e-05)(2.15569,-2.1e-05)(2.17565,-2.4e-05)(2.19561,-2.3e-05)(2.21557,-1.7e-05)(2.23553,-1e-05)(2.25549,-3e-06)(2.27545,1e-06)(2.29541,1e-06)(2.31537,-3e-06)(2.33533,-1e-05)(2.35529,-1.6e-05)(2.37525,-2e-05)(2.39521,-2e-05)(2.41517,-1.6e-05)(2.43513,-9e-06)(2.45509,-3e-06)(2.47505,1e-06)(2.49501,2e-06)(2.51497,-2e-06)(2.53493,-8e-06)(2.55489,-1.4e-05)(2.57485,-1.5e-05)(2.59481,-1.2e-05)(2.61477,-7e-06)(2.63473,-1e-06)(2.65469,3e-06)(2.67465,5e-06)(2.69461,2e-06)(2.71457,-4e-06)(2.73453,-1.1e-05)(2.75449,-1.5e-05)(2.77445,-1.3e-05)(2.79441,-7e-06)(2.81437,0)(2.83433,7e-06)(2.85429,1e-05)(2.87425,9e-06)(2.89421,3e-06)(2.91417,-4e-06)(2.93413,-9e-06)(2.95409,-7e-06)(2.97405,-0)(2.99401,8e-06)(3.01397,1.5e-05)(3.03393,1.7e-05)(3.05389,1.4e-05)(3.07385,8e-06)(3.09381,0)(3.11377,-3e-06)(3.13373,1e-06)(3.15369,1.1e-05)(3.17365,2.2e-05)(3.19361,3.1e-05)(3.21357,3.4e-05)(3.23353,3.2e-05)(3.25349,2.7e-05)(3.27345,2.2e-05)(3.29341,2e-05)(3.31337,2.5e-05)(3.33333,3.4e-05)(3.35329,4.5e-05)(3.37325,5.4e-05)(3.39321,5.9e-05)(3.41317,6.4e-05)(3.43313,6.7e-05)(3.45309,7e-05)(3.47305,7.4e-05)(3.49301,7.8e-05)(3.51297,8.6e-05)(3.53293,9.5e-05)(3.55289,0.000107)(3.57285,0.00012)(3.59281,0.000133)(3.61277,0.000145)(3.63273,0.000155)(3.65269,0.00016)(3.67265,0.000165)(3.69261,0.000172)(3.71257,0.000186)(3.73253,0.000206)(3.7525,0.000228)(3.77246,0.000249)(3.79242,0.000267)(3.81238,0.000281)(3.83234,0.000292)(3.8523,0.000304)(3.87226,0.000321)(3.89222,0.000344)(3.91218,0.000371)(3.93214,0.000397)(3.9521,0.000419)(3.97206,0.000438)(3.99202,0.000456)(4.01198,0.000476)(4.03194,0.000499)(4.0519,0.000526)(4.07186,0.000557)(4.09182,0.000591)(4.11178,0.000625)(4.13174,0.000659)(4.1517,0.000693)(4.17166,0.00073)(4.19162,0.000771)(4.21158,0.000815)(4.23154,0.00086)(4.2515,0.000906)(4.27146,0.000954)(4.29142,0.001002)(4.31138,0.00105)(4.33134,0.001095)(4.3513,0.001138)(4.37126,0.001179)(4.39122,0.001219)(4.41118,0.001257)(4.43114,0.001296)(4.4511,0.001334)(4.47106,0.001372)(4.49102,0.001406)(4.51098,0.001435)(4.53094,0.001463)(4.5509,0.001494)(4.57086,0.00153)(4.59082,0.001569)(4.61078,0.001607)(4.63074,0.001644)(4.6507,0.001676)(4.67066,0.001706)(4.69062,0.001734)(4.71058,0.001765)(4.73054,0.001801)(4.7505,0.00184)(4.77046,0.00188)(4.79042,0.001918)(4.81038,0.001954)(4.83034,0.001992)(4.8503,0.002033)(4.87026,0.002076)(4.89022,0.002119)(4.91018,0.002159)(4.93014,0.002197)(4.9501,0.002232)(4.97006,0.002267)(4.99002,0.002306)(5.00998,0.00235)(5.02994,0.002397)(5.0499,0.002446)(5.06986,0.002492)(5.08982,0.002537)(5.10978,0.002584)(5.12974,0.002636)(5.1497,0.002693)(5.16966,0.002751)(5.18962,0.002802)(5.20958,0.002843)(5.22954,0.002876)(5.2495,0.002904)(5.26946,0.002939)(5.28942,0.00299)(5.30938,0.003061)(5.32934,0.003148)(5.3493,0.003245)(5.36926,0.003342)(5.38922,0.003433)(5.40918,0.003518)(5.42914,0.003596)(5.4491,0.003671)(5.46906,0.003744)(5.48902,0.003812)(5.50898,0.003874)(5.52894,0.003946)(5.5489,0.004029)(5.56886,0.00411)(5.58882,0.004192)(5.60878,0.004287)(5.62874,0.004405)(5.6487,0.004542)(5.66866,0.004679)(5.68862,0.004792)(5.70858,0.004871)(5.72854,0.004926)(5.7485,0.004983)(5.76846,0.005073)(5.78842,0.00521)(5.80838,0.005363)(5.82834,0.005491)(5.8483,0.005607)(5.86826,0.005737)(5.88822,0.005864)(5.90818,0.005972)(5.92814,0.00607)(5.9481,0.006177)(5.96806,0.006295)(5.98802,0.006413)(6.00798,0.006521)(6.02794,0.006612)(6.0479,0.006694)(6.06786,0.006786)(6.08782,0.006894)(6.10778,0.007008)(6.12774,0.007101)(6.1477,0.007152)(6.16766,0.007172)(6.18762,0.007191)(6.20758,0.007216)(6.22754,0.007238)(6.2475,0.007253)(6.26747,0.007275)(6.28743,0.007316)(6.30739,0.007364)(6.32735,0.007411)(6.34731,0.007471)(6.36727,0.00756)(6.38723,0.007678)(6.40719,0.007808)(6.42715,0.007932)(6.44711,0.008059)(6.46707,0.008226)(6.48703,0.008455)(6.50699,0.008733)(6.52695,0.009014)(6.54691,0.009261)(6.56687,0.009464)(6.58683,0.009619)(6.60679,0.009746)(6.62675,0.009886)(6.64671,0.010054)(6.66667,0.010231)(6.68663,0.010408)(6.70659,0.010661)(6.72655,0.010973)(6.74651,0.011103)(6.76647,0.011172)(6.78643,0.01126)(6.80639,0.011311)(6.82635,0.011331)(6.84631,0.011337)(6.86627,0.011308)(6.88623,0.011217)(6.90619,0.011086)(6.92615,0.010956)(6.94611,0.010822)(6.96607,0.010705)(6.98603,0.010689)(7.00599,0.010798)(7.02595,0.010995)(7.04591,0.011235)(7.06587,0.011469)(7.08583,0.011682)(7.10579,0.011886)(7.12575,0.012018)(7.14571,0.011997)(7.16567,0.011844)(7.18563,0.0117)(7.20559,0.011624)(7.22555,0.011613)(7.24551,0.011661)(7.26547,0.011736)(7.28543,0.011784)(7.30539,0.011775)(7.32535,0.011715)(7.34531,0.011626)(7.36527,0.011528)(7.38523,0.011455)(7.40519,0.011437)(7.42515,0.011492)(7.44511,0.011601)(7.46507,0.011721)(7.48503,0.01182)(7.50499,0.011897)(7.52495,0.011971)(7.54491,0.012047)(7.56487,0.012092)(7.58483,0.012083)(7.60479,0.012026)(7.62475,0.011969)(7.64471,0.011992)(7.66467,0.012141)(7.68463,0.012396)(7.70459,0.012705)(7.72455,0.01302)(7.74451,0.013287)(7.76447,0.013446)(7.78443,0.013486)(7.80439,0.01343)(7.82435,0.013322)(7.84431,0.01326)(7.86427,0.013299)(7.88423,0.013392)(7.90419,0.013474)(7.92415,0.013537)(7.94411,0.013593)(7.96407,0.01362)(7.98403,0.013589)(8.00399,0.013514)(8.02395,0.013423)(8.04391,0.013344)(8.06387,0.013306)(8.08383,0.013303)(8.10379,0.013281)(8.12375,0.013184)(8.14371,0.012992)(8.16367,0.012774)(8.18363,0.012635)(8.20359,0.01263)(8.22355,0.012749)(8.24351,0.012938)(8.26347,0.013106)(8.28343,0.013176)(8.30339,0.01312)(8.32335,0.012971)(8.34331,0.012822)(8.36327,0.01278)(8.38323,0.012867)(8.40319,0.012987)(8.42315,0.013036)(8.44311,0.012996)(8.46307,0.012924)(8.48303,0.012907)(8.50299,0.012989)(8.52295,0.013142)(8.54291,0.013301)(8.56287,0.013392)(8.58283,0.01335)(8.60279,0.01321)(8.62275,0.013128)(8.64271,0.013227)(8.66267,0.013442)(8.68263,0.013568)(8.70259,0.013543)(8.72255,0.013503)(8.74251,0.013375)(8.76248,0.013106)(8.78244,0.012738)(8.8024,0.012366)(8.82236,0.012168)(8.84232,0.012145)(8.86228,0.01216)(8.88224,0.012192)(8.9022,0.012258)(8.92216,0.012312)(8.94212,0.012334)(8.96208,0.012352)(8.98204,0.012384)(9.002,0.012425)(9.02196,0.012481)(9.04192,0.012562)(9.06188,0.012657)(9.08184,0.012724)(9.1018,0.012704)(9.12176,0.012568)(9.14172,0.012341)(9.16168,0.012065)(9.18164,0.01177)(9.2016,0.011481)(9.22156,0.011248)(9.24152,0.011131)(9.26148,0.011144)(9.28144,0.01123)(9.3014,0.0113)(9.32136,0.011313)(9.34132,0.011273)(9.36128,0.011183)(9.38124,0.011049)(9.4012,0.010906)(9.42116,0.010804)(9.44112,0.010771)(9.46108,0.010824)(9.48104,0.010974)(9.501,0.011212)(9.52096,0.011492)(9.54092,0.01174)(9.56088,0.011883)(9.58084,0.011894)(9.6008,0.011792)(9.62076,0.01163)(9.64072,0.011484)(9.66068,0.011411)(9.68064,0.011421)(9.7006,0.011454)(9.72056,0.011447)(9.74052,0.011379)(9.76048,0.01126)(9.78044,0.011128)(9.8004,0.01102)(9.82036,0.010973)(9.84032,0.011012)(9.86028,0.011103)(9.88024,0.011151)(9.9002,0.011101)(9.92016,0.010978)(9.94012,0.010823)(9.96008,0.010657)(9.98004,0.010505)(10,0.010386)(10.02,0.01031)(10.0399,0.010298)(10.0599,0.010353)(10.0798,0.010436)(10.0998,0.010489)(10.1198,0.010507)(10.1397,0.010528)(10.1597,0.010575)(10.1796,0.010634)(10.1996,0.010664)(10.2196,0.010647)(10.2395,0.010595)(10.2595,0.010516)(10.2794,0.010414)(10.2994,0.010296)(10.3194,0.010168)(10.3393,0.010032)(10.3593,0.0099)(10.3792,0.009793)(10.3992,0.009726)(10.4192,0.00969)(10.4391,0.009673)(10.4591,0.00967)(10.479,0.009683)(10.499,0.009703)(10.519,0.009704)(10.5389,0.009676)(10.5589,0.009618)(10.5788,0.009525)(10.5988,0.009401)(10.6188,0.009274)(10.6387,0.009191)(10.6587,0.00918)(10.6786,0.009225)(10.6986,0.009276)(10.7186,0.009298)(10.7385,0.009281)(10.7585,0.009228)(10.7784,0.009132)(10.7984,0.008988)(10.8184,0.00882)(10.8383,0.008653)(10.8583,0.008513)(10.8782,0.008413)(10.8982,0.008363)(10.9182,0.008365)(10.9381,0.008395)(10.9581,0.008424)(10.978,0.008449)(10.998,0.008493)(11.018,0.008578)(11.0379,0.008693)(11.0579,0.008803)(11.0778,0.008868)(11.0978,0.00886)(11.1178,0.008782)(11.1377,0.008664)(11.1577,0.008549)(11.1776,0.008457)(11.1976,0.008396)(11.2176,0.008371)(11.2375,0.008366)(11.2575,0.008369)(11.2774,0.008387)(11.2974,0.008431)(11.3174,0.008493)(11.3373,0.008553)(11.3573,0.008601)(11.3772,0.008616)(11.3972,0.008576)(11.4172,0.008478)(11.4371,0.008351)(11.4571,0.008217)(11.477,0.008087)(11.497,0.007972)(11.517,0.007883)(11.5369,0.007832)(11.5569,0.007823)(11.5768,0.007847)(11.5968,0.00788)(11.6168,0.007887)(11.6367,0.007849)(11.6567,0.007768)(11.6766,0.007667)(11.6966,0.007577)(11.7166,0.007519)(11.7365,0.007498)(11.7565,0.007511)(11.7764,0.007547)(11.7964,0.007591)(11.8164,0.007634)(11.8363,0.007675)(11.8563,0.007701)(11.8762,0.007699)(11.8962,0.007664)(11.9162,0.007603)(11.9361,0.007525)(11.9561,0.007444)(11.976,0.007367)(11.996,0.007292)(12.016,0.007222)(12.0359,0.007157)(12.0559,0.007101)(12.0758,0.007055)(12.0958,0.007014)(12.1158,0.006972)(12.1357,0.006937)(12.1557,0.006912)(12.1756,0.006899)(12.1956,0.006899)(12.2156,0.006905)(12.2355,0.006903)(12.2555,0.006885)(12.2754,0.006852)(12.2954,0.00681)(12.3154,0.006769)(12.3353,0.006739)(12.3553,0.00672)(12.3752,0.006703)(12.3952,0.006679)(12.4152,0.006647)(12.4351,0.006603)(12.4551,0.006548)(12.475,0.006487)(12.495,0.006428)(12.515,0.00638)(12.5349,0.006342)(12.5549,0.006308)(12.5749,0.00627)(12.5948,0.006229)(12.6148,0.006193)(12.6347,0.006169)(12.6547,0.006158)(12.6747,0.006159)(12.6946,0.006163)(12.7146,0.006169)(12.7345,0.006175)(12.7545,0.006173)(12.7745,0.006163)(12.7944,0.00615)(12.8144,0.006137)(12.8343,0.006122)(12.8543,0.006111)(12.8743,0.00611)(12.8942,0.006118)(12.9142,0.006126)(12.9341,0.006125)(12.9541,0.006106)(12.9741,0.00607)(12.994,0.006026)(13.014,0.005978)(13.0339,0.00593)(13.0539,0.005893)(13.0739,0.005875)(13.0938,0.005865)(13.1138,0.005852)(13.1337,0.005838)(13.1537,0.005827)(13.1737,0.005813)(13.1936,0.005793)(13.2136,0.005768)(13.2335,0.00574)(13.2535,0.005709)(13.2735,0.005681)(13.2934,0.00567)(13.3134,0.005675)(13.3333,0.005685)(13.3533,0.005687)(13.3733,0.005669)(13.3932,0.005632)(13.4132,0.005586)(13.4331,0.005539)(13.4531,0.005497)(13.4731,0.005466)(13.493,0.005442)(13.513,0.005417)(13.5329,0.005392)(13.5529,0.005369)(13.5729,0.005352)(13.5928,0.005334)(13.6128,0.005312)(13.6327,0.005288)(13.6527,0.005259)(13.6727,0.005223)(13.6926,0.005183)(13.7126,0.005147)(13.7325,0.005124)(13.7525,0.005114)(13.7725,0.005113)(13.7924,0.005121)(13.8124,0.005132)(13.8323,0.005141)(13.8523,0.005145)(13.8723,0.005142)(13.8922,0.005127)(13.9122,0.005103)(13.9321,0.005073)(13.9521,0.005042)(13.9721,0.005011)(13.992,0.004977)(14.012,0.004941)(14.0319,0.004907)(14.0519,0.004879)(14.0719,0.004861)(14.0918,0.004849)(14.1118,0.004838)(14.1317,0.004827)(14.1517,0.004815)(14.1717,0.004804)(14.1916,0.004792)(14.2116,0.004779)(14.2315,0.004768)(14.2515,0.004762)(14.2715,0.004759)(14.2914,0.004754)(14.3114,0.004743)(14.3313,0.004721)(14.3513,0.00469)(14.3713,0.004652)(14.3912,0.004613)(14.4112,0.004577)(14.4311,0.004544)(14.4511,0.004516)(14.4711,0.00449)(14.491,0.004465)(14.511,0.004443)(14.5309,0.004422)(14.5509,0.004403)(14.5709,0.004385)(14.5908,0.004368)(14.6108,0.004351)(14.6307,0.004333)(14.6507,0.004313)(14.6707,0.004293)(14.6906,0.004274)(14.7106,0.004257)(14.7305,0.004239)(14.7505,0.004221)(14.7705,0.004204)(14.7904,0.004192)(14.8104,0.004186)(14.8303,0.004183)(14.8503,0.00418)(14.8703,0.004175)(14.8902,0.004167)(14.9102,0.004152)(14.9301,0.004129)(14.9501,0.004104)(14.9701,0.004085)(14.99,0.004075)(15.01,0.00407)(15.0299,0.004067)(15.0499,0.004065)(15.0699,0.004062)(15.0898,0.004059)(15.1098,0.004053)(15.1297,0.004043)(15.1497,0.004028)(15.1697,0.004012)(15.1896,0.003996)(15.2096,0.003981)(15.2295,0.003971)(15.2495,0.003962)(15.2695,0.003955)(15.2894,0.003948)(15.3094,0.003938)(15.3293,0.003924)(15.3493,0.003908)(15.3693,0.003891)(15.3892,0.003872)(15.4092,0.003854)(15.4291,0.003842)(15.4491,0.00384)(15.4691,0.00385)(15.489,0.003863)(15.509,0.003872)(15.5289,0.003873)(15.5489,0.003866)(15.5689,0.003853)(15.5888,0.003833)(15.6088,0.003809)(15.6287,0.003778)(15.6487,0.003746)(15.6687,0.003715)(15.6886,0.003685)(15.7086,0.003655)(15.7285,0.003623)(15.7485,0.003591)(15.7685,0.003559)(15.7884,0.00353)(15.8084,0.003504)(15.8283,0.003479)(15.8483,0.003455)(15.8683,0.003431)(15.8882,0.003409)(15.9082,0.00339)(15.9281,0.003377)(15.9481,0.00337)(15.9681,0.003364)(15.988,0.003359)(16.008,0.003355)(16.0279,0.003353)(16.0479,0.003353)(16.0679,0.003356)(16.0878,0.003364)(16.1078,0.003375)(16.1277,0.003384)(16.1477,0.003389)(16.1677,0.003387)(16.1876,0.003381)(16.2076,0.003367)(16.2275,0.003346)(16.2475,0.003316)(16.2675,0.003276)(16.2874,0.003231)(16.3074,0.003188)(16.3273,0.003152)(16.3473,0.003128)(16.3673,0.003115)(16.3872,0.003112)(16.4072,0.003113)(16.4271,0.003115)(16.4471,0.003114)(16.4671,0.003111)(16.487,0.003106)(16.507,0.003097)(16.5269,0.003086)(16.5469,0.003074)(16.5669,0.003062)(16.5868,0.003051)(16.6068,0.003042)(16.6267,0.003036)(16.6467,0.003029)(16.6667,0.003021)(16.6866,0.003011)(16.7066,0.002998)(16.7265,0.002984)(16.7465,0.002968)(16.7665,0.002954)(16.7864,0.00294)(16.8064,0.002925)(16.8263,0.002905)(16.8463,0.002882)(16.8663,0.002858)(16.8862,0.00284)(16.9062,0.002829)(16.9261,0.002827)(16.9461,0.002829)(16.9661,0.002835)(16.986,0.002842)(17.006,0.002848)(17.0259,0.002849)(17.0459,0.002846)(17.0659,0.00284)(17.0858,0.002832)(17.1058,0.002822)(17.1257,0.002813)(17.1457,0.002808)(17.1657,0.002807)(17.1856,0.002808)(17.2056,0.00281)(17.2255,0.00281)(17.2455,0.002809)(17.2655,0.002808)(17.2854,0.002808)(17.3054,0.002808)(17.3253,0.00281)(17.3453,0.002815)(17.3653,0.002821)(17.3852,0.002825)(17.4052,0.002829)(17.4251,0.002832)(17.4451,0.002833)(17.4651,0.002833)(17.485,0.002832)(17.505,0.002827)(17.525,0.002819)(17.5449,0.002808)(17.5649,0.002792)(17.5848,0.00277)(17.6048,0.002741)(17.6248,0.002711)(17.6447,0.002681)(17.6647,0.002653)(17.6846,0.002629)(17.7046,0.002608)(17.7246,0.002591)(17.7445,0.002575)(17.7645,0.002559)(17.7844,0.002544)(17.8044,0.002533)(17.8244,0.002525)(17.8443,0.002518)(17.8643,0.002509)(17.8842,0.002498)(17.9042,0.002488)(17.9242,0.002483)(17.9441,0.002481)(17.9641,0.002484)(17.984,0.00249)(18.004,0.002499)(18.024,0.002507)(18.0439,0.00251)(18.0639,0.002507)(18.0838,0.002499)(18.1038,0.002488)(18.1238,0.002475)(18.1437,0.002459)(18.1637,0.002442)(18.1836,0.002424)(18.2036,0.002408)(18.2236,0.002394)(18.2435,0.002384)(18.2635,0.002376)(18.2834,0.002372)(18.3034,0.002369)(18.3234,0.002367)(18.3433,0.002367)(18.3633,0.00237)(18.3832,0.002374)(18.4032,0.002379)(18.4232,0.00238)(18.4431,0.002378)(18.4631,0.002373)(18.483,0.002365)(18.503,0.002354)(18.523,0.002343)(18.5429,0.00233)(18.5629,0.002317)(18.5828,0.002304)(18.6028,0.002291)(18.6228,0.002279)(18.6427,0.002268)(18.6627,0.002256)(18.6826,0.002243)(18.7026,0.002229)(18.7226,0.002213)(18.7425,0.002198)(18.7625,0.002187)(18.7824,0.002183)(18.8024,0.002182)(18.8224,0.002185)(18.8423,0.00219)(18.8623,0.002199)(18.8822,0.002208)(18.9022,0.002216)(18.9222,0.00222)(18.9421,0.002223)(18.9621,0.002224)(18.982,0.002221)(19.002,0.002214)(19.022,0.002203)(19.0419,0.002191)(19.0619,0.00218)(19.0818,0.00217)(19.1018,0.002163)(19.1218,0.002158)(19.1417,0.002155)(19.1617,0.002154)(19.1816,0.002152)(19.2016,0.002149)(19.2216,0.002145)(19.2415,0.002139)(19.2615,0.002129)(19.2814,0.002116)(19.3014,0.0021)(19.3214,0.002082)(19.3413,0.002065)(19.3613,0.002051)(19.3812,0.002037)(19.4012,0.002023)(19.4212,0.00201)(19.4411,0.001999)(19.4611,0.00199)(19.481,0.001983)(19.501,0.001978)(19.521,0.001976)(19.5409,0.001977)(19.5609,0.001979)(19.5808,0.001981)(19.6008,0.001985)(19.6208,0.001991)(19.6407,0.001998)(19.6607,0.002003)(19.6806,0.002005)(19.7006,0.002005)(19.7206,0.002004)(19.7405,0.002003)(19.7605,0.002001)(19.7804,0.001998)(19.8004,0.001993)(19.8204,0.001987)(19.8403,0.001979)(19.8603,0.00197)(19.8802,0.00196)(19.9002,0.00195)(19.9202,0.001941)(19.9401,0.001933)(19.9601,0.001925)(19.98,0.001917)(20,0.001908)};
+
+%coordinates{(0,-0)(0.1,-0)(0.2,-1.1e-05)(0.3,-1e-05)(0.4,-1.3e-05)(0.5,-1.6e-05)(0.6,-2.5e-05)(0.7,-1.6e-05)(0.8,-2.8e-05)(0.9,-2.5e-05)(1,-2.5e-05)(1.1,-2.4e-05)(1.2,-3.1e-05)(1.3,-2.2e-05)(1.4,-2.4e-05)(1.5,-2.7e-05)(1.6,-2.1e-05)(1.7,-1.9e-05)(1.8,-2.6e-05)(1.9,-1.2e-05)(2,-2.5e-05)(2.1,-8e-06)(2.2,-2.6e-05)(2.3,-2e-06)(2.4,-2.4e-05)(2.5,-3e-06)(2.6,-1.8e-05)(2.7,-4e-06)(2.8,-1.5e-05)(2.9,-5e-06)(3,-3e-06)(3.1,-1.4e-05)(3.2,1.1e-05)(3.3,-5e-06)(3.4,2.5e-05)(3.5,3.2e-05)(3.6,7.1e-05)(3.7,9e-05)(3.8,0.000156)(3.9,0.000201)(4,0.000279)(4.1,0.000375)(4.2,0.00051)(4.3,0.000691)(4.4,0.000864)(4.5,0.001022)(4.6,0.00114)(4.7,0.001224)(4.8,0.001302)(4.9,0.001404)(5,0.001526)(5.1,0.001685)(5.2,0.001869)(5.3,0.002032)(5.4,0.002409)(5.5,0.002752)(5.6,0.003089)(5.7,0.003588)(5.8,0.004093)(5.9,0.004445)(6,0.004825)(6.1,0.005375)(6.2,0.005868)(6.3,0.006183)(6.4,0.006331)(6.5,0.006797)(6.6,0.007712)(6.7,0.008319)(6.8,0.00896)(6.9,0.009326)(7,0.009777)(7.1,0.010393)(7.2,0.010403)(7.3,0.010421)(7.4,0.010724)(7.5,0.011246)(7.6,0.011484)(7.7,0.011616)(7.8,0.012531)(7.9,0.012689)(8,0.012545)(8.1,0.012473)(8.2,0.012667)(8.3,0.012824)(8.4,0.012941)(8.5,0.0127)(8.6,0.012743)(8.7,0.012718)(8.8,0.012434)(8.9,0.011995)(9,0.011842)(9.1,0.01197)(9.2,0.011787)(9.3,0.011592)(9.4,0.011345)(9.5,0.010925)(9.6,0.011148)(9.7,0.01112)(9.8,0.010876)(9.9,0.010962)(10,0.010975)(10.1,0.010899)(10.2,0.010607)(10.3,0.010373)(10.4,0.01008)(10.5,0.009831)(10.6,0.009418)(10.7,0.009322)(10.8,0.009141)(10.9,0.009067)(11,0.009091)(11.1,0.008885)(11.2,0.008732)(11.3,0.008681)(11.4,0.008587)(11.5,0.008343)(11.6,0.008185)(11.7,0.008042)(11.8,0.008009)(11.9,0.007849)(12,0.007513)(12.1,0.007208)(12.2,0.007146)(12.3,0.007062)(12.4,0.007009)(12.5,0.006919)(12.6,0.006761)(12.7,0.006594)(12.8,0.006558)(12.9,0.00642)(13,0.00625)(13.1,0.006121)(13.2,0.006015)(13.3,0.00599)(13.4,0.00586)(13.5,0.005745)(13.6,0.005636)(13.7,0.005559)(13.8,0.005431)(13.9,0.005396)(14,0.005265)(14.1,0.0052)(14.2,0.005097)(14.3,0.004995)(14.4,0.004901)(14.5,0.004814)(14.6,0.0047)(14.7,0.004622)(14.8,0.004574)(14.9,0.004471)(15,0.004354)(15.1,0.004272)(15.2,0.00426)(15.3,0.004171)(15.4,0.004091)(15.5,0.004049)(15.6,0.003979)(15.7,0.003897)(15.8,0.003842)(15.9,0.003795)(16,0.003726)(16.1,0.003677)(16.2,0.003605)(16.3,0.003528)(16.4,0.003477)(16.5,0.003454)(16.6,0.003389)(16.7,0.003346)(16.8,0.003298)(16.9,0.003229)(17,0.003182)(17.1,0.003154)(17.2,0.003085)(17.3,0.003019)(17.4,0.002948)(17.5,0.002911)(17.6,0.002874)(17.7,0.002836)(17.8,0.002812)(17.9,0.002746)(18,0.0027)(18.1,0.002667)(18.2,0.002635)(18.3,0.002613)(18.4,0.002582)(18.5,0.002542)(18.6,0.002495)(18.7,0.002458)(18.8,0.002435)(18.9,0.002405)(19,0.002376)(19.1,0.002342)(19.2,0.002314)(19.3,0.002278)(19.4,0.00225)(19.5,0.00221)(19.6,0.002181)(19.7,0.002155)(19.8,0.002114)(19.9,0.002085)(20,0.002057)};
 
@@
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0,-0)(0.1,-0)(0.2,-1.1e-05)(0.3,-1e-05)(0.4,-1.4e-05)(0.5,-1.6e-05)(0.6,-2.5e-05)(0.7,-1.5e-05)(0.8,-2.8e-05)(0.9,-2.4e-05)(1,-2.6e-05)(1.1,-2.4e-05)(1.2,-3.2e-05)(1.3,-2.1e-05)(1.4,-2.5e-05)(1.5,-2.6e-05)(1.6,-2.1e-05)(1.7,-1.9e-05)(1.8,-2.6e-05)(1.9,-1.1e-05)(2,-2.5e-05)(2.1,-8e-06)(2.2,-2.5e-05)(2.3,-3e-06)(2.4,-2.3e-05)(2.5,-3e-06)(2.6,-1.7e-05)(2.7,-5e-06)(2.8,-1.2e-05)(2.9,-8e-06)(3,0)(3.1,-1.4e-05)(3.2,1.6e-05)(3.3,2e-06)(3.4,3.8e-05)(3.5,5.5e-05)(3.6,0.000113)(3.7,0.000157)(3.8,0.000259)(3.9,0.000347)(4,0.000462)(4.1,0.00061)(4.2,0.000794)(4.3,0.001034)(4.4,0.001274)(4.5,0.001479)(4.6,0.001593)(4.7,0.001646)(4.8,0.001711)(4.9,0.001762)(5,0.001796)(5.1,0.001797)(5.2,0.001863)(5.3,0.00193)(5.4,0.002103)(5.5,0.002362)(5.6,0.002622)(5.7,0.002965)(5.8,0.0034)(5.9,0.003864)(6,0.004146)(6.1,0.004136)(6.2,0.004996)(6.3,0.005703)(6.4,0.007004)(6.5,0.008301)(6.6,0.009652)(6.7,0.010951)(6.8,0.011407)(6.9,0.013516)(7,0.013684)(7.1,0.014113)(7.2,0.013742)(7.3,0.013059)(7.4,0.014071)(7.5,0.013728)(7.6,0.012821)(7.7,0.012544)(7.8,0.013137)(7.9,0.01424)(8,0.0126)(8.1,0.012763)(8.2,0.012715)(8.3,0.013327)(8.4,0.014155)(8.5,0.013513)(8.6,0.01272)(8.7,0.012531)(8.8,0.013325)(8.9,0.013302)(9,0.013029)(9.1,0.012178)(9.2,0.012225)(9.3,0.013543)(9.4,0.014134)(9.5,0.013937)(9.6,0.012882)(9.7,0.012053)(9.8,0.013117)(9.9,0.012565)(10,0.012059)(10.1,0.011629)(10.2,0.011937)(10.3,0.012583)(10.4,0.012349)(10.5,0.011896)(10.6,0.012309)(10.7,0.011838)(10.8,0.011069)(10.9,0.01059)(11,0.009978)(11.1,0.01026)(11.2,0.010219)(11.3,0.010047)(11.4,0.009645)(11.5,0.009094)(11.6,0.008931)(11.7,0.009278)(11.8,0.009438)(11.9,0.008792)(12,0.008361)(12.1,0.00806)(12.2,0.007456)(12.3,0.007163)(12.4,0.007304)(12.5,0.007684)(12.6,0.00785)(12.7,0.007487)(12.8,0.007027)(12.9,0.006911)(13,0.006836)(13.1,0.00682)(13.2,0.006961)(13.3,0.006948)(13.4,0.006935)(13.5,0.007177)(13.6,0.007091)(13.7,0.006754)(13.8,0.006527)(13.9,0.006203)(14,0.006094)(14.1,0.005863)(14.2,0.005841)(14.3,0.005897)(14.4,0.005707)(14.5,0.005541)(14.6,0.005396)(14.7,0.00524)(14.8,0.005027)(14.9,0.004758)(15,0.004552)(15.1,0.00432)(15.2,0.00422)(15.3,0.004206)(15.4,0.004258)(15.5,0.004133)(15.6,0.003853)(15.7,0.003705)(15.8,0.003694)(15.9,0.003643)(16,0.003556)(16.1,0.003453)(16.2,0.003301)(16.3,0.003159)(16.4,0.003127)(16.5,0.003173)(16.6,0.003171)(16.7,0.003147)(16.8,0.003069)(16.9,0.002928)(17,0.002782)(17.1,0.00268)(17.2,0.00265)(17.3,0.002667)(17.4,0.002667)(17.5,0.00264)(17.6,0.002643)(17.7,0.002728)(17.8,0.002793)(17.9,0.002758)(18,0.002614)(18.1,0.00245)(18.2,0.002369)(18.3,0.002311)(18.4,0.002249)(18.5,0.002216)(18.6,0.002239)(18.7,0.002253)(18.8,0.002197)(18.9,0.002097)(19,0.002026)(19.1,0.00198)(19.2,0.001938)(19.3,0.001898)(19.4,0.001831)(19.5,0.001766)(19.6,0.001727)(19.7,0.001723)(19.8,0.001719)(19.9,0.001707)(20,0.001698)};
-
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0,-0)(0.1,-0)(0.2,-1.1e-05)(0.3,-1e-05)(0.4,-1.3e-05)(0.5,-1.6e-05)(0.6,-2.5e-05)(0.7,-1.6e-05)(0.8,-2.8e-05)(0.9,-2.5e-05)(1,-2.5e-05)(1.1,-2.4e-05)(1.2,-3.1e-05)(1.3,-2.2e-05)(1.4,-2.4e-05)(1.5,-2.7e-05)(1.6,-2.1e-05)(1.7,-1.9e-05)(1.8,-2.6e-05)(1.9,-1.2e-05)(2,-2.5e-05)(2.1,-8e-06)(2.2,-2.6e-05)(2.3,-2e-06)(2.4,-2.4e-05)(2.5,-3e-06)(2.6,-1.8e-05)(2.7,-3e-06)(2.8,-1.4e-05)(2.9,-5e-06)(3,-2e-06)(3.1,-1.3e-05)(3.2,1.3e-05)(3.3,-2e-06)(3.4,3.1e-05)(3.5,4.2e-05)(3.6,8.7e-05)(3.7,0.000117)(3.8,0.000203)(3.9,0.000272)(4,0.000381)(4.1,0.000525)(4.2,0.000714)(4.3,0.000946)(4.4,0.001174)(4.5,0.001385)(4.6,0.001526)(4.7,0.001622)(4.8,0.001724)(4.9,0.001872)(5,0.002067)(5.1,0.002293)(5.2,0.002575)(5.3,0.002922)(5.4,0.003262)(5.5,0.003483)(5.6,0.003769)(5.7,0.004017)(5.8,0.004277)(5.9,0.004577)(6,0.005016)(6.1,0.005492)(6.2,0.006201)(6.3,0.006815)(6.4,0.007136)(6.5,0.008004)(6.6,0.008172)(6.7,0.008672)(6.8,0.009029)(6.9,0.009603)(7,0.010511)(7.1,0.010773)(7.2,0.01077)(7.3,0.011114)(7.4,0.011948)(7.5,0.012385)(7.6,0.012737)(7.7,0.012609)(7.8,0.013135)(7.9,0.0134)(8,0.01394)(8.1,0.014723)(8.2,0.014402)(8.3,0.01389)(8.4,0.013833)(8.5,0.014437)(8.6,0.014767)(8.7,0.014866)(8.8,0.014848)(8.9,0.014442)(9,0.014686)(9.1,0.014743)(9.2,0.014514)(9.3,0.013853)(9.4,0.013283)(9.5,0.013316)(9.6,0.013374)(9.7,0.013275)(9.8,0.013436)(9.9,0.013322)(10,0.012779)(10.1,0.01246)(10.2,0.012098)(10.3,0.011781)(10.4,0.011687)(10.5,0.011505)(10.6,0.011334)(10.7,0.011239)(10.8,0.010745)(10.9,0.010255)(11,0.010106)(11.1,0.009861)(11.2,0.009627)(11.3,0.009393)(11.4,0.009309)(11.5,0.009199)(11.6,0.00884)(11.7,0.00852)(11.8,0.008338)(11.9,0.008167)(12,0.008026)(12.1,0.007844)(12.2,0.007667)(12.3,0.007473)(12.4,0.007233)(12.5,0.007066)(12.6,0.006955)(12.7,0.00674)(12.8,0.006575)(12.9,0.006464)(13,0.006253)(13.1,0.005983)(13.2,0.005785)(13.3,0.005656)(13.4,0.005644)(13.5,0.005576)(13.6,0.005468)(13.7,0.005338)(13.8,0.005177)(13.9,0.004984)(14,0.004959)(14.1,0.004922)(14.2,0.004777)(14.3,0.004656)(14.4,0.004532)(14.5,0.004364)(14.6,0.004265)(14.7,0.004247)(14.8,0.004179)(14.9,0.004047)(15,0.003934)(15.1,0.003867)(15.2,0.003779)(15.3,0.003693)(15.4,0.003662)(15.5,0.003621)(15.6,0.003529)(15.7,0.003449)(15.8,0.003409)(15.9,0.003362)(16,0.003323)(16.1,0.003276)(16.2,0.003213)(16.3,0.003145)(16.4,0.003069)(16.5,0.003003)(16.6,0.00294)(16.7,0.002864)(16.8,0.002815)(16.9,0.002753)(17,0.002676)(17.1,0.002615)(17.2,0.002597)(17.3,0.002599)(17.4,0.002565)(17.5,0.002509)(17.6,0.002445)(17.7,0.002379)(17.8,0.002328)(17.9,0.002295)(18,0.002251)(18.1,0.002218)(18.2,0.002207)(18.3,0.002183)(18.4,0.002146)(18.5,0.002114)(18.6,0.002093)(18.7,0.002071)(18.8,0.002029)(18.9,0.001986)(19,0.001957)(19.1,0.001951)(19.2,0.001951)(19.3,0.001943)(19.4,0.001932)(19.5,0.001914)(19.6,0.001889)(19.7,0.001859)(19.8,0.001846)(19.9,0.001833)(20,0.001798)};
+coordinates{(0,-0)(0.02,-0)(0.04,-1e-06)(0.06,-2e-06)(0.08,-1e-06)(0.1,-0)(0.12,-0)(0.14,-2e-06)(0.16,-5e-06)(0.18,-8e-06)(0.2,-1.1e-05)(0.22,-1.3e-05)(0.24,-1.5e-05)(0.26,-1.5e-05)(0.28,-1.3e-05)(0.3,-1e-05)(0.32,-8e-06)(0.34,-8e-06)(0.36,-9e-06)(0.38,-1.1e-05)(0.4,-1.4e-05)(0.42,-1.7e-05)(0.44,-2e-05)(0.46,-2e-05)(0.48,-1.8e-05)(0.5,-1.6e-05)(0.52,-1.6e-05)(0.54,-1.7e-05)(0.56,-1.9e-05)(0.58,-2.2e-05)(0.6,-2.5e-05)(0.62,-2.7e-05)(0.64,-2.7e-05)(0.66,-2.3e-05)(0.68,-1.9e-05)(0.7,-1.5e-05)(0.72,-1.4e-05)(0.74,-1.5e-05)(0.76,-1.8e-05)(0.78,-2.3e-05)(0.8,-2.8e-05)(0.82,-3.2e-05)(0.84,-3.3e-05)(0.86,-3.1e-05)(0.88,-2.7e-05)(0.9,-2.4e-05)(0.92,-2.2e-05)(0.94,-2e-05)(0.96,-2e-05)(0.98,-2.2e-05)(1,-2.6e-05)(1.02,-2.9e-05)(1.04,-2.9e-05)(1.06,-2.8e-05)(1.08,-2.6e-05)(1.1,-2.4e-05)(1.12,-2.2e-05)(1.14,-2.2e-05)(1.16,-2.4e-05)(1.18,-2.8e-05)(1.2,-3.2e-05)(1.22,-3.2e-05)(1.24,-3e-05)(1.26,-2.7e-05)(1.28,-2.4e-05)(1.3,-2.1e-05)(1.32,-2e-05)(1.34,-2e-05)(1.36,-2.2e-05)(1.38,-2.3e-05)(1.4,-2.5e-05)(1.42,-2.6e-05)(1.44,-2.6e-05)(1.46,-2.7e-05)(1.48,-2.7e-05)(1.5,-2.6e-05)(1.52,-2.5e-05)(1.54,-2.3e-05)(1.56,-2.1e-05)(1.58,-2.1e-05)(1.6,-2.1e-05)(1.62,-2.1e-05)(1.64,-2.2e-05)(1.66,-2.2e-05)(1.68,-2e-05)(1.7,-1.9e-05)(1.72,-1.8e-05)(1.74,-1.8e-05)(1.76,-2e-05)(1.78,-2.4e-05)(1.8,-2.6e-05)(1.82,-2.7e-05)(1.84,-2.4e-05)(1.86,-2e-05)(1.88,-1.5e-05)(1.9,-1.1e-05)(1.92,-1.1e-05)(1.94,-1.4e-05)(1.96,-1.9e-05)(1.98,-2.3e-05)(2,-2.5e-05)(2.02,-2.3e-05)(2.04,-1.8e-05)(2.06,-1.2e-05)(2.08,-8e-06)(2.1,-8e-06)(2.12,-1.2e-05)(2.14,-1.8e-05)(2.16,-2.4e-05)(2.18,-2.7e-05)(2.2,-2.5e-05)(2.22,-1.9e-05)(2.24,-1.1e-05)(2.26,-5e-06)(2.28,-2e-06)(2.3,-3e-06)(2.32,-8e-06)(2.34,-1.5e-05)(2.36,-2.2e-05)(2.38,-2.5e-05)(2.4,-2.3e-05)(2.42,-1.9e-05)(2.44,-1.2e-05)(2.46,-6e-06)(2.48,-3e-06)(2.5,-3e-06)(2.52,-8e-06)(2.54,-1.5e-05)(2.56,-2e-05)(2.58,-2e-05)(2.6,-1.7e-05)(2.62,-1.1e-05)(2.64,-6e-06)(2.66,-2e-06)(2.68,-1e-06)(2.7,-5e-06)(2.72,-1.2e-05)(2.74,-1.9e-05)(2.76,-2.1e-05)(2.78,-1.8e-05)(2.8,-1.2e-05)(2.82,-5e-06)(2.84,0)(2.86,2e-06)(2.88,-1e-06)(2.9,-8e-06)(2.92,-1.5e-05)(2.94,-1.8e-05)(2.96,-1.4e-05)(2.98,-7e-06)(3,0)(3.02,5e-06)(3.04,5e-06)(3.06,0)(3.08,-7e-06)(3.1,-1.4e-05)(3.12,-1.6e-05)(3.14,-1e-05)(3.16,0)(3.18,1e-05)(3.2,1.6e-05)(3.22,1.7e-05)(3.24,1.3e-05)(3.26,6e-06)(3.28,2e-06)(3.3,2e-06)(3.32,7e-06)(3.34,1.7e-05)(3.36,2.7e-05)(3.38,3.3e-05)(3.4,3.8e-05)(3.42,4.2e-05)(3.44,4.5e-05)(3.46,4.7e-05)(3.48,5e-05)(3.5,5.5e-05)(3.52,6.3e-05)(3.54,7.4e-05)(3.56,8.6e-05)(3.58,9.9e-05)(3.6,0.000113)(3.62,0.000125)(3.64,0.000134)(3.66,0.000139)(3.68,0.000145)(3.7,0.000157)(3.72,0.000175)(3.74,0.000197)(3.76,0.00022)(3.78,0.000241)(3.8,0.000259)(3.82,0.000272)(3.84,0.000284)(3.86,0.000299)(3.88,0.00032)(3.9,0.000347)(3.92,0.000375)(3.94,0.0004)(3.96,0.000422)(3.98,0.000442)(4,0.000462)(4.02,0.000485)(4.04,0.000511)(4.06,0.000542)(4.08,0.000576)(4.1,0.00061)(4.12,0.000644)(4.14,0.000678)(4.16,0.000713)(4.18,0.000752)(4.2,0.000794)(4.22,0.000839)(4.24,0.000885)(4.26,0.000932)(4.28,0.000983)(4.3,0.001034)(4.32,0.001084)(4.34,0.001132)(4.36,0.00118)(4.38,0.001227)(4.4,0.001274)(4.42,0.001319)(4.44,0.001365)(4.46,0.001409)(4.48,0.001448)(4.5,0.001479)(4.52,0.001504)(4.54,0.001526)(4.56,0.001548)(4.58,0.001572)(4.6,0.001593)(4.62,0.001611)(4.64,0.001625)(4.66,0.001633)(4.68,0.001639)(4.7,0.001646)(4.72,0.001656)(4.74,0.00167)(4.76,0.001686)(4.78,0.0017)(4.8,0.001711)(4.82,0.001721)(4.84,0.001732)(4.86,0.001743)(4.88,0.001754)(4.9,0.001762)(4.92,0.001769)(4.94,0.001774)(4.96,0.00178)(4.98,0.001787)(5,0.001796)(5.02,0.001806)(5.04,0.001812)(5.06,0.001811)(5.08,0.001805)(5.1,0.001797)(5.12,0.001795)(5.14,0.001804)(5.16,0.001821)(5.18,0.001842)(5.2,0.001863)(5.22,0.001878)(5.24,0.001888)(5.26,0.001896)(5.28,0.001909)(5.3,0.00193)(5.32,0.001959)(5.34,0.001994)(5.36,0.002029)(5.38,0.002065)(5.4,0.002103)(5.42,0.002146)(5.44,0.002194)(5.46,0.002248)(5.48,0.002305)(5.5,0.002362)(5.52,0.002418)(5.54,0.002471)(5.56,0.002521)(5.58,0.002571)(5.6,0.002622)(5.62,0.002677)(5.64,0.002736)(5.66,0.002803)(5.68,0.00288)(5.7,0.002965)(5.72,0.003054)(5.74,0.003142)(5.76,0.003228)(5.78,0.003314)(5.8,0.0034)(5.82,0.003489)(5.84,0.003578)(5.86,0.003669)(5.88,0.003764)(5.9,0.003864)(5.92,0.00396)(5.94,0.004043)(5.96,0.004104)(5.98,0.004138)(6,0.004146)(6.02,0.004134)(6.04,0.004113)(6.06,0.004097)(6.08,0.004101)(6.1,0.004136)(6.12,0.004212)(6.14,0.004337)(6.16,0.004517)(6.18,0.004744)(6.2,0.004996)(6.22,0.005239)(6.24,0.005437)(6.26,0.005568)(6.28,0.005643)(6.3,0.005703)(6.32,0.005804)(6.34,0.005992)(6.36,0.006275)(6.38,0.006628)(6.4,0.007004)(6.42,0.007354)(6.44,0.007652)(6.46,0.007894)(6.48,0.008101)(6.5,0.008301)(6.52,0.008515)(6.54,0.008758)(6.56,0.009031)(6.58,0.009332)(6.6,0.009652)(6.62,0.009985)(6.64,0.010314)(6.66,0.01061)(6.68,0.010831)(6.7,0.010951)(6.72,0.010982)(6.74,0.010981)(6.76,0.011019)(6.78,0.011153)(6.8,0.011407)(6.82,0.011781)(6.84,0.012246)(6.86,0.012745)(6.88,0.013194)(6.9,0.013516)(6.92,0.013678)(6.94,0.013701)(6.96,0.013658)(6.98,0.013635)(7,0.013684)(7.02,0.013795)(7.04,0.013922)(7.06,0.014027)(7.08,0.014096)(7.1,0.014113)(7.12,0.014057)(7.14,0.013939)(7.16,0.013814)(7.18,0.013747)(7.2,0.013742)(7.22,0.013726)(7.24,0.013619)(7.26,0.013414)(7.28,0.013189)(7.3,0.013059)(7.32,0.013094)(7.34,0.013284)(7.36,0.013556)(7.38,0.013834)(7.4,0.014071)(7.42,0.014236)(7.44,0.014291)(7.46,0.014208)(7.48,0.014001)(7.5,0.013728)(7.52,0.013455)(7.54,0.013221)(7.56,0.013039)(7.58,0.01291)(7.6,0.012821)(7.62,0.012753)(7.64,0.01269)(7.66,0.012624)(7.68,0.012564)(7.7,0.012544)(7.72,0.012597)(7.74,0.012718)(7.76,0.012866)(7.78,0.013005)(7.8,0.013137)(7.82,0.013294)(7.84,0.013512)(7.86,0.0138)(7.88,0.01409)(7.9,0.01424)(7.92,0.01413)(7.94,0.013761)(7.96,0.013269)(7.98,0.012841)(8,0.0126)(8.02,0.012558)(8.04,0.012636)(8.06,0.012737)(8.08,0.01279)(8.1,0.012763)(8.12,0.012669)(8.14,0.012562)(8.16,0.012513)(8.18,0.012566)(8.2,0.012715)(8.22,0.012905)(8.24,0.01307)(8.26,0.013178)(8.28,0.013247)(8.3,0.013327)(8.32,0.013459)(8.34,0.013653)(8.36,0.013873)(8.38,0.014057)(8.4,0.014155)(8.42,0.014146)(8.44,0.014037)(8.46,0.013864)(8.48,0.013677)(8.5,0.013513)(8.52,0.013375)(8.54,0.013243)(8.56,0.013093)(8.58,0.012915)(8.6,0.01272)(8.62,0.012536)(8.64,0.0124)(8.66,0.012347)(8.68,0.012392)(8.7,0.012531)(8.72,0.012731)(8.74,0.012946)(8.76,0.01313)(8.78,0.013257)(8.8,0.013325)(8.82,0.013349)(8.84,0.013347)(8.86,0.013332)(8.88,0.013315)(8.9,0.013302)(8.92,0.013291)(8.94,0.01327)(8.96,0.013227)(8.98,0.01315)(9,0.013029)(9.02,0.012865)(9.04,0.012673)(9.06,0.012479)(9.08,0.01231)(9.1,0.012178)(9.12,0.012083)(9.14,0.012027)(9.16,0.012022)(9.18,0.012085)(9.2,0.012225)(9.22,0.012436)(9.24,0.012698)(9.26,0.012986)(9.28,0.013276)(9.3,0.013543)(9.32,0.01377)(9.34,0.013946)(9.36,0.014065)(9.38,0.014125)(9.4,0.014134)(9.42,0.014107)(9.44,0.014066)(9.46,0.014029)(9.48,0.013994)(9.5,0.013937)(9.52,0.013831)(9.54,0.013661)(9.56,0.013431)(9.58,0.013164)(9.6,0.012882)(9.62,0.01261)(9.64,0.012366)(9.66,0.012174)(9.68,0.012061)(9.7,0.012053)(9.72,0.012173)(9.74,0.012411)(9.76,0.01271)(9.78,0.012975)(9.8,0.013117)(9.82,0.013106)(9.84,0.012978)(9.86,0.012812)(9.88,0.012668)(9.9,0.012565)(9.92,0.012487)(9.94,0.012405)(9.96,0.012305)(9.98,0.012186)(10,0.012059)(10.02,0.01194)(10.04,0.011836)(10.06,0.01175)(10.08,0.01168)(10.1,0.011629)(10.12,0.011603)(10.14,0.011615)(10.16,0.011675)(10.18,0.011786)(10.2,0.011937)(10.22,0.012107)(10.24,0.012272)(10.26,0.012413)(10.28,0.012518)(10.3,0.012583)(10.32,0.012611)(10.34,0.012604)(10.36,0.012561)(10.38,0.012474)(10.4,0.012349)(10.42,0.012211)(10.44,0.012083)(10.46,0.011978)(10.48,0.011908)(10.5,0.011896)(10.52,0.011949)(10.54,0.012046)(10.56,0.012154)(10.58,0.012246)(10.6,0.012309)(10.62,0.012327)(10.64,0.012289)(10.66,0.012188)(10.68,0.012031)(10.7,0.011838)(10.72,0.011638)(10.74,0.011454)(10.76,0.011297)(10.78,0.01117)(10.8,0.011069)(10.82,0.010986)(10.84,0.010909)(10.86,0.010824)(10.88,0.01072)(10.9,0.01059)(10.92,0.010438)(10.94,0.010275)(10.96,0.010126)(10.98,0.010022)(11,0.009978)(11.02,0.009995)(11.04,0.010053)(11.06,0.010128)(11.08,0.010201)(11.1,0.01026)(11.12,0.010294)(11.14,0.010299)(11.16,0.01028)(11.18,0.01025)(11.2,0.010219)(11.22,0.010192)(11.24,0.010165)(11.26,0.010134)(11.28,0.010096)(11.3,0.010047)(11.32,0.009986)(11.34,0.009912)(11.36,0.009829)(11.38,0.00974)(11.4,0.009645)(11.42,0.009542)(11.44,0.00943)(11.46,0.009313)(11.48,0.009197)(11.5,0.009094)(11.52,0.009012)(11.54,0.008957)(11.56,0.008928)(11.58,0.00892)(11.6,0.008931)(11.62,0.008964)(11.64,0.009019)(11.66,0.009094)(11.68,0.009183)(11.7,0.009278)(11.72,0.009369)(11.74,0.009442)(11.76,0.009484)(11.78,0.009484)(11.8,0.009438)(11.82,0.009348)(11.84,0.009224)(11.86,0.009082)(11.88,0.008935)(11.9,0.008792)(11.92,0.00866)(11.94,0.008548)(11.96,0.00846)(11.98,0.0084)(12,0.008361)(12.02,0.008332)(12.04,0.008297)(12.06,0.008241)(12.08,0.008161)(12.1,0.00806)(12.12,0.007947)(12.14,0.007827)(12.16,0.007702)(12.18,0.007577)(12.2,0.007456)(12.22,0.00735)(12.24,0.007266)(12.26,0.007208)(12.28,0.007175)(12.3,0.007163)(12.32,0.007167)(12.34,0.007184)(12.36,0.007213)(12.38,0.007253)(12.4,0.007304)(12.42,0.007368)(12.44,0.007442)(12.46,0.007522)(12.48,0.007604)(12.5,0.007684)(12.52,0.007756)(12.54,0.007814)(12.56,0.007851)(12.58,0.007864)(12.6,0.00785)(12.62,0.007813)(12.64,0.007753)(12.66,0.007675)(12.68,0.007585)(12.7,0.007487)(12.72,0.007387)(12.74,0.007285)(12.76,0.007187)(12.78,0.007099)(12.8,0.007027)(12.82,0.006974)(12.84,0.00694)(12.86,0.006921)(12.88,0.006913)(12.9,0.006911)(12.92,0.006907)(12.94,0.006897)(12.96,0.006881)(12.98,0.006859)(13,0.006836)(13.02,0.006813)(13.04,0.006797)(13.06,0.006792)(13.08,0.006801)(13.1,0.00682)(13.12,0.006846)(13.14,0.006875)(13.16,0.006904)(13.18,0.006934)(13.2,0.006961)(13.22,0.006981)(13.24,0.006991)(13.26,0.006988)(13.28,0.006973)(13.3,0.006948)(13.32,0.006921)(13.34,0.0069)(13.36,0.006894)(13.38,0.006905)(13.4,0.006935)(13.42,0.006978)(13.44,0.007031)(13.46,0.007087)(13.48,0.007138)(13.5,0.007177)(13.52,0.007201)(13.54,0.007207)(13.56,0.007192)(13.58,0.007154)(13.6,0.007091)(13.62,0.007012)(13.64,0.006928)(13.66,0.006852)(13.68,0.006794)(13.7,0.006754)(13.72,0.006725)(13.74,0.006695)(13.76,0.006654)(13.78,0.006597)(13.8,0.006527)(13.82,0.00645)(13.84,0.006373)(13.86,0.006304)(13.88,0.006246)(13.9,0.006203)(13.92,0.006172)(13.94,0.006151)(13.96,0.006134)(13.98,0.006117)(14,0.006094)(14.02,0.006063)(14.04,0.006021)(14.06,0.005971)(14.08,0.005916)(14.1,0.005863)(14.12,0.005822)(14.14,0.005799)(14.16,0.005797)(14.18,0.005813)(14.2,0.005841)(14.22,0.005873)(14.24,0.005901)(14.26,0.005916)(14.28,0.005914)(14.3,0.005897)(14.32,0.005867)(14.34,0.005829)(14.36,0.005787)(14.38,0.005746)(14.4,0.005707)(14.42,0.005672)(14.44,0.005639)(14.46,0.005607)(14.48,0.005574)(14.5,0.005541)(14.52,0.005509)(14.54,0.00548)(14.56,0.005451)(14.58,0.005423)(14.6,0.005396)(14.62,0.005369)(14.64,0.00534)(14.66,0.005309)(14.68,0.005275)(14.7,0.00524)(14.72,0.005204)(14.74,0.005166)(14.76,0.005124)(14.78,0.005077)(14.8,0.005027)(14.82,0.004972)(14.84,0.004914)(14.86,0.004857)(14.88,0.004804)(14.9,0.004758)(14.92,0.004717)(14.94,0.004678)(14.96,0.004638)(14.98,0.004597)(15,0.004552)(15.02,0.004504)(15.04,0.004453)(15.06,0.004403)(15.08,0.004359)(15.1,0.00432)(15.12,0.004289)(15.14,0.004263)(15.16,0.004245)(15.18,0.004231)(15.2,0.00422)(15.22,0.00421)(15.24,0.004201)(15.26,0.004197)(15.28,0.004199)(15.3,0.004206)(15.32,0.004217)(15.34,0.004231)(15.36,0.004245)(15.38,0.004255)(15.4,0.004258)(15.42,0.004251)(15.44,0.004235)(15.46,0.00421)(15.48,0.004176)(15.5,0.004133)(15.52,0.004082)(15.54,0.004025)(15.56,0.003966)(15.58,0.003908)(15.6,0.003853)(15.62,0.003805)(15.64,0.003768)(15.66,0.003739)(15.68,0.003719)(15.7,0.003705)(15.72,0.003696)(15.74,0.003693)(15.76,0.003693)(15.78,0.003694)(15.8,0.003694)(15.82,0.003692)(15.84,0.003687)(15.86,0.003677)(15.88,0.003662)(15.9,0.003643)(15.92,0.003624)(15.94,0.003606)(15.96,0.003589)(15.98,0.003572)(16,0.003556)(16.02,0.00354)(16.04,0.003522)(16.06,0.003502)(16.08,0.003479)(16.1,0.003453)(16.12,0.003426)(16.14,0.003397)(16.16,0.003366)(16.18,0.003333)(16.2,0.0033)(16.22,0.003268)(16.24,0.003237)(16.26,0.003208)(16.28,0.003181)(16.3,0.003159)(16.32,0.003142)(16.34,0.003131)(16.36,0.003124)(16.38,0.003123)(16.4,0.003127)(16.42,0.003136)(16.44,0.003147)(16.46,0.003157)(16.48,0.003166)(16.5,0.003174)(16.52,0.003178)(16.54,0.00318)(16.56,0.003178)(16.58,0.003175)(16.6,0.003172)(16.62,0.003168)(16.64,0.003165)(16.66,0.00316)(16.68,0.003155)(16.7,0.003148)(16.72,0.003139)(16.74,0.003125)(16.76,0.003109)(16.78,0.003091)(16.8,0.00307)(16.82,0.003046)(16.84,0.003019)(16.86,0.002989)(16.88,0.002959)(16.9,0.002928)(16.92,0.002896)(16.94,0.002865)(16.96,0.002835)(16.98,0.002808)(17,0.002781)(17.02,0.002756)(17.04,0.002733)(17.06,0.002712)(17.08,0.002694)(17.1,0.002679)(17.12,0.002666)(17.14,0.002656)(17.16,0.002651)(17.18,0.00265)(17.2,0.00265)(17.22,0.002652)(17.24,0.002656)(17.26,0.002661)(17.28,0.002667)(17.3,0.002671)(17.32,0.002673)(17.34,0.002675)(17.36,0.002677)(17.38,0.002676)(17.4,0.002673)(17.42,0.002668)(17.44,0.002664)(17.46,0.002659)(17.48,0.002652)(17.5,0.002645)(17.52,0.00264)(17.54,0.002637)(17.56,0.002638)(17.58,0.00264)(17.6,0.002647)(17.62,0.002658)(17.64,0.002672)(17.66,0.002689)(17.68,0.002707)(17.7,0.002725)(17.72,0.002743)(17.74,0.002759)(17.76,0.002772)(17.78,0.002782)(17.8,0.002788)(17.82,0.002791)(17.84,0.00279)(17.86,0.002783)(17.88,0.002772)(17.9,0.002756)(17.92,0.002736)(17.94,0.002712)(17.96,0.002682)(17.98,0.00265)(18,0.002615)(18.02,0.002579)(18.04,0.002544)(18.06,0.002509)(18.08,0.002478)(18.1,0.002451)(18.12,0.002429)(18.14,0.00241)(18.16,0.002395)(18.18,0.002382)(18.2,0.00237)(18.22,0.002359)(18.24,0.002348)(18.26,0.002336)(18.28,0.002323)(18.3,0.00231)(18.32,0.002298)(18.34,0.002285)(18.36,0.002272)(18.38,0.002259)(18.4,0.002248)(18.42,0.002237)(18.44,0.002228)(18.46,0.00222)(18.48,0.002216)(18.5,0.002214)(18.52,0.002215)(18.54,0.002219)(18.56,0.002223)(18.58,0.002229)(18.6,0.002236)(18.62,0.002241)(18.64,0.002245)(18.66,0.002247)(18.68,0.002248)(18.7,0.002247)(18.72,0.002242)(18.74,0.002233)(18.76,0.002221)(18.78,0.002206)(18.8,0.002188)(18.82,0.002168)(18.84,0.002146)(18.86,0.002125)(18.88,0.002106)(18.9,0.002089)(18.92,0.002072)(18.94,0.002057)(18.96,0.002044)(18.98,0.002033)(19,0.002023)(19.02,0.002013)(19.04,0.002003)(19.06,0.001996)(19.08,0.001989)(19.1,0.001982)(19.12,0.001974)(19.14,0.001966)(19.16,0.001958)(19.18,0.00195)(19.2,0.001941)(19.22,0.001932)(19.24,0.001923)(19.26,0.001916)(19.28,0.001908)(19.3,0.001897)(19.32,0.001885)(19.34,0.001873)(19.36,0.001859)(19.38,0.001844)(19.4,0.001828)(19.42,0.001813)(19.44,0.001799)(19.46,0.001786)(19.48,0.001774)(19.5,0.001762)(19.52,0.001752)(19.54,0.001743)(19.56,0.001735)(19.58,0.001727)(19.6,0.00172)(19.62,0.001716)(19.64,0.001714)(19.66,0.001713)(19.68,0.001713)(19.7,0.001714)(19.72,0.001715)(19.74,0.001715)(19.76,0.001714)(19.78,0.001711)(19.8,0.001709)(19.82,0.001707)(19.84,0.001704)(19.86,0.001702)(19.88,0.001699)(19.9,0.001697)(19.92,0.001696)(19.94,0.001694)(19.96,0.001692)(19.98,0.001691)(20,0.001692)};
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0,-0)(0.0199601,-0)(0.0399202,-1e-06)(0.0598802,-2e-06)(0.0798403,-1e-06)(0.0998004,-0)(0.11976,-0)(0.139721,-2e-06)(0.159681,-5e-06)(0.179641,-8e-06)(0.199601,-1.1e-05)(0.219561,-1.3e-05)(0.239521,-1.5e-05)(0.259481,-1.5e-05)(0.279441,-1.3e-05)(0.299401,-1e-05)(0.319361,-8e-06)(0.339321,-8e-06)(0.359281,-9e-06)(0.379242,-1.1e-05)(0.399202,-1.3e-05)(0.419162,-1.7e-05)(0.439122,-1.9e-05)(0.459082,-2e-05)(0.479042,-1.8e-05)(0.499002,-1.6e-05)(0.518962,-1.6e-05)(0.538922,-1.7e-05)(0.558882,-1.9e-05)(0.578842,-2.2e-05)(0.598802,-2.5e-05)(0.618762,-2.7e-05)(0.638723,-2.7e-05)(0.658683,-2.4e-05)(0.678643,-1.9e-05)(0.698603,-1.6e-05)(0.718563,-1.4e-05)(0.738523,-1.5e-05)(0.758483,-1.8e-05)(0.778443,-2.3e-05)(0.798403,-2.8e-05)(0.818363,-3.2e-05)(0.838323,-3.3e-05)(0.858283,-3.1e-05)(0.878244,-2.8e-05)(0.898204,-2.5e-05)(0.918164,-2.2e-05)(0.938124,-2e-05)(0.958084,-2e-05)(0.978044,-2.2e-05)(0.998004,-2.5e-05)(1.01796,-2.8e-05)(1.03792,-2.9e-05)(1.05788,-2.8e-05)(1.07784,-2.6e-05)(1.0978,-2.4e-05)(1.11776,-2.2e-05)(1.13772,-2.2e-05)(1.15768,-2.4e-05)(1.17764,-2.8e-05)(1.1976,-3.1e-05)(1.21756,-3.2e-05)(1.23752,-3e-05)(1.25749,-2.7e-05)(1.27745,-2.4e-05)(1.29741,-2.2e-05)(1.31737,-2e-05)(1.33733,-2e-05)(1.35729,-2.1e-05)(1.37725,-2.3e-05)(1.39721,-2.4e-05)(1.41717,-2.5e-05)(1.43713,-2.6e-05)(1.45709,-2.7e-05)(1.47705,-2.7e-05)(1.49701,-2.7e-05)(1.51697,-2.5e-05)(1.53693,-2.3e-05)(1.55689,-2.2e-05)(1.57685,-2.1e-05)(1.59681,-2.1e-05)(1.61677,-2.1e-05)(1.63673,-2.2e-05)(1.65669,-2.2e-05)(1.67665,-2.1e-05)(1.69661,-1.9e-05)(1.71657,-1.8e-05)(1.73653,-1.8e-05)(1.75649,-2e-05)(1.77645,-2.3e-05)(1.79641,-2.6e-05)(1.81637,-2.7e-05)(1.83633,-2.5e-05)(1.85629,-2.1e-05)(1.87625,-1.6e-05)(1.89621,-1.2e-05)(1.91617,-1.1e-05)(1.93613,-1.3e-05)(1.95609,-1.7e-05)(1.97605,-2.2e-05)(1.99601,-2.5e-05)(2.01597,-2.4e-05)(2.03593,-1.9e-05)(2.05589,-1.4e-05)(2.07585,-9e-06)(2.09581,-8e-06)(2.11577,-1.1e-05)(2.13573,-1.7e-05)(2.15569,-2.3e-05)(2.17565,-2.7e-05)(2.19561,-2.6e-05)(2.21557,-2.1e-05)(2.23553,-1.3e-05)(2.25549,-7e-06)(2.27545,-3e-06)(2.29541,-2e-06)(2.31537,-6e-06)(2.33533,-1.3e-05)(2.35529,-2e-05)(2.37525,-2.5e-05)(2.39521,-2.4e-05)(2.41517,-2e-05)(2.43513,-1.4e-05)(2.45509,-8e-06)(2.47505,-4e-06)(2.49501,-3e-06)(2.51497,-6e-06)(2.53493,-1.3e-05)(2.55489,-1.9e-05)(2.57485,-2e-05)(2.59481,-1.8e-05)(2.61477,-1.3e-05)(2.63473,-8e-06)(2.65469,-3e-06)(2.67465,-1e-06)(2.69461,-3e-06)(2.71457,-1e-05)(2.73453,-1.7e-05)(2.75449,-2.1e-05)(2.77445,-2e-05)(2.79441,-1.4e-05)(2.81437,-8e-06)(2.83433,-1e-06)(2.85429,2e-06)(2.87425,1e-06)(2.89421,-5e-06)(2.91417,-1.3e-05)(2.93413,-1.7e-05)(2.95409,-1.6e-05)(2.97405,-1e-05)(2.99401,-2e-06)(3.01397,4e-06)(3.03393,6e-06)(3.05389,3e-06)(3.07385,-5e-06)(3.09381,-1.3e-05)(3.11377,-1.7e-05)(3.13373,-1.4e-05)(3.15369,-5e-06)(3.17365,6e-06)(3.19361,1.3e-05)(3.21357,1.6e-05)(3.23353,1.3e-05)(3.25349,7e-06)(3.27345,1e-06)(3.29341,-2e-06)(3.31337,1e-06)(3.33333,9e-06)(3.35329,1.9e-05)(3.37325,2.6e-05)(3.39321,3.1e-05)(3.41317,3.4e-05)(3.43313,3.6e-05)(3.45309,3.8e-05)(3.47305,3.9e-05)(3.49301,4.2e-05)(3.51297,4.7e-05)(3.53293,5.5e-05)(3.55289,6.5e-05)(3.57285,7.5e-05)(3.59281,8.7e-05)(3.61277,9.8e-05)(3.63273,0.000105)(3.65269,0.000109)(3.67265,0.000111)(3.69261,0.000117)(3.71257,0.000129)(3.73253,0.000147)(3.7525,0.000167)(3.77246,0.000186)(3.79242,0.000203)(3.81238,0.000215)(3.83234,0.000225)(3.8523,0.000235)(3.87226,0.000251)(3.89222,0.000272)(3.91218,0.000297)(3.93214,0.000322)(3.9521,0.000344)(3.97206,0.000363)(3.99202,0.000381)(4.01198,0.000402)(4.03194,0.000426)(4.0519,0.000455)(4.07186,0.000489)(4.09182,0.000525)(4.11178,0.000562)(4.13174,0.000597)(4.1517,0.000633)(4.17166,0.000672)(4.19162,0.000714)(4.21158,0.000758)(4.23154,0.000803)(4.2515,0.00085)(4.27146,0.000898)(4.29142,0.000946)(4.31138,0.000994)(4.33134,0.001041)(4.3513,0.001086)(4.37126,0.00113)(4.39122,0.001174)(4.41118,0.001218)(4.43114,0.001262)(4.4511,0.001306)(4.47106,0.001348)(4.49102,0.001385)(4.51098,0.001415)(4.53094,0.001441)(4.5509,0.001468)(4.57086,0.001496)(4.59082,0.001526)(4.61078,0.001553)(4.63074,0.001576)(4.6507,0.001595)(4.67066,0.001609)(4.69062,0.001622)(4.71058,0.001636)(4.73054,0.001655)(4.7505,0.001677)(4.77046,0.001701)(4.79042,0.001724)(4.81038,0.001747)(4.83034,0.001773)(4.8503,0.001803)(4.87026,0.001837)(4.89022,0.001872)(4.91018,0.001908)(4.93014,0.001944)(4.9501,0.001982)(4.97006,0.002022)(4.99002,0.002067)(5.00998,0.002115)(5.02994,0.002165)(5.0499,0.002213)(5.06986,0.002256)(5.08982,0.002293)(5.10978,0.002333)(5.12974,0.002379)(5.1497,0.002436)(5.16966,0.002503)(5.18962,0.002575)(5.20958,0.002648)(5.22954,0.002719)(5.2495,0.002786)(5.26946,0.002853)(5.28942,0.002922)(5.30938,0.002996)(5.32934,0.003072)(5.3493,0.003144)(5.36926,0.003208)(5.38922,0.003262)(5.40918,0.003308)(5.42914,0.00335)(5.4491,0.003391)(5.46906,0.003435)(5.48902,0.003483)(5.50898,0.003537)(5.52894,0.003595)(5.5489,0.003654)(5.56886,0.003713)(5.58882,0.003769)(5.60878,0.003823)(5.62874,0.003875)(5.6487,0.003925)(5.66866,0.003971)(5.68862,0.004017)(5.70858,0.004065)(5.72854,0.004117)(5.7485,0.004171)(5.76846,0.004225)(5.78842,0.004277)(5.80838,0.004328)(5.82834,0.004379)(5.8483,0.004434)(5.86826,0.0045)(5.88822,0.004577)(5.90818,0.004665)(5.92814,0.004757)(5.9481,0.004848)(5.96806,0.004934)(5.98802,0.005016)(6.00798,0.005093)(6.02794,0.005172)(6.0479,0.00526)(6.06786,0.005365)(6.08782,0.005492)(6.10778,0.005635)(6.12774,0.005784)(6.1477,0.005932)(6.16766,0.006071)(6.18762,0.006201)(6.20758,0.006326)(6.22754,0.006455)(6.2475,0.00659)(6.26747,0.006717)(6.28743,0.006815)(6.30739,0.00687)(6.32735,0.006897)(6.34731,0.006931)(6.36727,0.007005)(6.38723,0.007136)(6.40719,0.007314)(6.42715,0.007512)(6.44711,0.007704)(6.46707,0.007871)(6.48703,0.008004)(6.50699,0.008101)(6.52695,0.008161)(6.54691,0.008183)(6.56687,0.008179)(6.58683,0.008172)(6.60679,0.008192)(6.62675,0.008262)(6.64671,0.008386)(6.66667,0.008536)(6.68663,0.008672)(6.70659,0.008769)(6.72655,0.008829)(6.74651,0.008873)(6.76647,0.008935)(6.78643,0.009029)(6.80639,0.00914)(6.82635,0.009241)(6.84631,0.009337)(6.86627,0.009453)(6.88623,0.009603)(6.90619,0.009778)(6.92615,0.009965)(6.94611,0.010158)(6.96607,0.010349)(6.98603,0.010511)(7.00599,0.010622)(7.02595,0.010685)(7.04591,0.010725)(7.06587,0.010754)(7.08583,0.010773)(7.10579,0.010782)(7.12575,0.010785)(7.14571,0.010784)(7.16567,0.010778)(7.18563,0.01077)(7.20559,0.01077)(7.22555,0.010792)(7.24551,0.010851)(7.26547,0.010956)(7.28543,0.011114)(7.30539,0.011307)(7.32535,0.011494)(7.34531,0.01165)(7.36527,0.011793)(7.38523,0.011948)(7.40519,0.012106)(7.42515,0.012233)(7.44511,0.012309)(7.46507,0.012349)(7.48503,0.012385)(7.50499,0.012442)(7.52495,0.01253)(7.54491,0.012631)(7.56487,0.012709)(7.58483,0.012737)(7.60479,0.012712)(7.62475,0.012656)(7.64471,0.012599)(7.66467,0.012577)(7.68463,0.012609)(7.70459,0.012691)(7.72455,0.012801)(7.74451,0.012921)(7.76447,0.013037)(7.78443,0.013135)(7.80439,0.013208)(7.82435,0.013259)(7.84431,0.013298)(7.86427,0.013341)(7.88423,0.0134)(7.90419,0.013479)(7.92415,0.013573)(7.94411,0.013679)(7.96407,0.013799)(7.98403,0.01394)(8.00399,0.014104)(8.02395,0.014284)(8.04391,0.014465)(8.06387,0.014621)(8.08383,0.014723)(8.10379,0.014756)(8.12375,0.014724)(8.14371,0.014644)(8.16367,0.014532)(8.18363,0.014402)(8.20359,0.014271)(8.22355,0.014154)(8.24351,0.014054)(8.26347,0.013965)(8.28343,0.01389)(8.30339,0.01384)(8.32335,0.013817)(8.34331,0.013811)(8.36327,0.013812)(8.38323,0.013833)(8.40319,0.013896)(8.42315,0.014009)(8.44311,0.014156)(8.46307,0.014306)(8.48303,0.014437)(8.50299,0.014541)(8.52295,0.014624)(8.54291,0.014687)(8.56287,0.014731)(8.58283,0.014767)(8.60279,0.014801)(8.62275,0.014831)(8.64271,0.014852)(8.66267,0.014863)(8.68263,0.014866)(8.70259,0.014867)(8.72255,0.014876)(8.74251,0.01489)(8.76248,0.014888)(8.78244,0.014848)(8.8024,0.014772)(8.82236,0.014676)(8.84232,0.014581)(8.86228,0.0145)(8.88224,0.014442)(8.9022,0.014418)(8.92216,0.014439)(8.94212,0.014507)(8.96208,0.014602)(8.98204,0.014686)(9.002,0.014731)(9.02196,0.014741)(9.04192,0.014739)(9.06188,0.014742)(9.08184,0.014743)(9.1018,0.014733)(9.12176,0.014704)(9.14172,0.014656)(9.16168,0.014592)(9.18164,0.014514)(9.2016,0.014415)(9.22156,0.014291)(9.24152,0.014146)(9.26148,0.013995)(9.28144,0.013853)(9.3014,0.013721)(9.32136,0.013596)(9.34132,0.013476)(9.36128,0.013367)(9.38124,0.013283)(9.4012,0.013234)(9.42116,0.013224)(9.44112,0.013246)(9.46108,0.013282)(9.48104,0.013316)(9.501,0.01334)(9.52096,0.013355)(9.54092,0.013368)(9.56088,0.013377)(9.58084,0.013374)(9.6008,0.013353)(9.62076,0.01332)(9.64072,0.013289)(9.66068,0.013272)(9.68064,0.013275)(9.7006,0.013292)(9.72056,0.013317)(9.74052,0.013352)(9.76048,0.013394)(9.78044,0.013436)(9.8004,0.013465)(9.82036,0.013468)(9.84032,0.013443)(9.86028,0.013393)(9.88024,0.013322)(9.9002,0.013233)(9.92016,0.013129)(9.94012,0.013014)(9.96008,0.012895)(9.98004,0.012779)(10,0.012676)(10.02,0.012594)(10.0399,0.012536)(10.0599,0.012495)(10.0798,0.01246)(10.0998,0.012416)(10.1198,0.012357)(10.1397,0.01228)(10.1597,0.012191)(10.1796,0.012098)(10.1996,0.012007)(10.2196,0.011929)(10.2395,0.011865)(10.2595,0.011817)(10.2794,0.011781)(10.2994,0.011755)(10.3194,0.011735)(10.3393,0.01172)(10.3593,0.011705)(10.3792,0.011687)(10.3992,0.011663)(10.4192,0.011633)(10.4391,0.011596)(10.4591,0.011553)(10.479,0.011505)(10.499,0.011455)(10.519,0.011409)(10.5389,0.011373)(10.5589,0.011347)(10.5788,0.011334)(10.5988,0.01133)(10.6188,0.011328)(10.6387,0.011317)(10.6587,0.011289)(10.6786,0.011239)(10.6986,0.011167)(10.7186,0.011078)(10.7385,0.010975)(10.7585,0.010863)(10.7784,0.010745)(10.7984,0.010626)(10.8184,0.010512)(10.8383,0.01041)(10.8583,0.010323)(10.8782,0.010255)(10.8982,0.010206)(10.9182,0.010173)(10.9381,0.010149)(10.9581,0.010129)(10.978,0.010106)(10.998,0.010078)(11.018,0.01004)(11.0379,0.009989)(11.0579,0.009928)(11.0778,0.009861)(11.0978,0.009799)(11.1178,0.009745)(11.1377,0.009701)(11.1577,0.009663)(11.1776,0.009627)(11.1976,0.009589)(11.2176,0.009545)(11.2375,0.009494)(11.2575,0.009441)(11.2774,0.009393)(11.2974,0.009353)(11.3174,0.009325)(11.3373,0.00931)(11.3573,0.009306)(11.3772,0.009309)(11.3972,0.009313)(11.4172,0.009307)(11.4371,0.009287)(11.4571,0.00925)(11.477,0.009199)(11.497,0.009135)(11.517,0.009063)(11.5369,0.008987)(11.5569,0.008912)(11.5768,0.00884)(11.5968,0.008771)(11.6168,0.008703)(11.6367,0.008638)(11.6567,0.008576)(11.6766,0.00852)(11.6966,0.008471)(11.7166,0.008429)(11.7365,0.008394)(11.7565,0.008364)(11.7764,0.008338)(11.7964,0.008311)(11.8164,0.008279)(11.8363,0.008244)(11.8563,0.008205)(11.8762,0.008167)(11.8962,0.008131)(11.9162,0.0081)(11.9361,0.008074)(11.9561,0.008051)(11.976,0.008026)(11.996,0.007996)(12.016,0.00796)(12.0359,0.007921)(12.0559,0.007882)(12.0758,0.007844)(12.0958,0.007806)(12.1158,0.00777)(12.1357,0.007735)(12.1557,0.007701)(12.1756,0.007667)(12.1956,0.007631)(12.2156,0.007593)(12.2355,0.007555)(12.2555,0.007515)(12.2754,0.007473)(12.2954,0.007428)(12.3154,0.00738)(12.3353,0.007331)(12.3553,0.007282)(12.3752,0.007233)(12.3952,0.007187)(12.4152,0.007147)(12.4351,0.007114)(12.4551,0.007087)(12.475,0.007066)(12.495,0.007048)(12.515,0.007032)(12.5349,0.007013)(12.5549,0.006988)(12.5749,0.006955)(12.5948,0.006916)(12.6148,0.006874)(12.6347,0.00683)(12.6547,0.006784)(12.6747,0.00674)(12.6946,0.006699)(12.7146,0.006662)(12.7345,0.006629)(12.7545,0.0066)(12.7745,0.006575)(12.7944,0.006553)(12.8144,0.006534)(12.8343,0.006514)(12.8543,0.006491)(12.8743,0.006464)(12.8942,0.006433)(12.9142,0.006397)(12.9341,0.006354)(12.9541,0.006306)(12.9741,0.006253)(12.994,0.006198)(13.014,0.006142)(13.0339,0.006087)(13.0539,0.006033)(13.0739,0.005983)(13.0938,0.005939)(13.1138,0.005898)(13.1337,0.005859)(13.1537,0.005821)(13.1737,0.005785)(13.1936,0.005751)(13.2136,0.005719)(13.2335,0.005691)(13.2535,0.005669)(13.2735,0.005656)(13.2934,0.005651)(13.3134,0.005649)(13.3333,0.005648)(13.3533,0.005647)(13.3733,0.005644)(13.3932,0.005638)(13.4132,0.005627)(13.4331,0.005612)(13.4531,0.005594)(13.4731,0.005576)(13.493,0.005556)(13.513,0.005536)(13.5329,0.005515)(13.5529,0.005492)(13.5729,0.005468)(13.5928,0.005442)(13.6128,0.005414)(13.6327,0.005387)(13.6527,0.005361)(13.6727,0.005338)(13.6926,0.005314)(13.7126,0.005287)(13.7325,0.005255)(13.7525,0.005219)(13.7725,0.005177)(13.7924,0.005132)(13.8124,0.005087)(13.8323,0.005046)(13.8523,0.005011)(13.8723,0.004984)(13.8922,0.004965)(13.9122,0.004955)(13.9321,0.004953)(13.9521,0.004956)(13.9721,0.004959)(13.992,0.004961)(14.012,0.004959)(14.0319,0.004952)(14.0519,0.00494)(14.0719,0.004922)(14.0918,0.004897)(14.1118,0.004869)(14.1317,0.004839)(14.1517,0.004807)(14.1717,0.004777)(14.1916,0.004749)(14.2116,0.004723)(14.2315,0.004699)(14.2515,0.004678)(14.2715,0.004656)(14.2914,0.004634)(14.3114,0.004612)(14.3313,0.004588)(14.3513,0.004561)(14.3713,0.004532)(14.3912,0.0045)(14.4112,0.004466)(14.4311,0.004431)(14.4511,0.004397)(14.4711,0.004364)(14.491,0.004334)(14.511,0.004309)(14.5309,0.004289)(14.5509,0.004275)(14.5709,0.004265)(14.5908,0.004258)(14.6108,0.004255)(14.6307,0.004254)(14.6507,0.004252)(14.6707,0.004247)(14.6906,0.004239)(14.7106,0.004229)(14.7305,0.004216)(14.7505,0.004199)(14.7705,0.004179)(14.7904,0.004156)(14.8104,0.004131)(14.8303,0.004104)(14.8503,0.004076)(14.8703,0.004047)(14.8902,0.004019)(14.9102,0.003995)(14.9301,0.003972)(14.9501,0.003952)(14.9701,0.003934)(14.99,0.00392)(15.01,0.003907)(15.0299,0.003894)(15.0499,0.003881)(15.0699,0.003867)(15.0898,0.003853)(15.1098,0.003837)(15.1297,0.003819)(15.1497,0.003799)(15.1697,0.003779)(15.1896,0.003759)(15.2096,0.003741)(15.2295,0.003722)(15.2495,0.003706)(15.2695,0.003693)(15.2894,0.003684)(15.3094,0.003676)(15.3293,0.003669)(15.3493,0.003665)(15.3693,0.003662)(15.3892,0.003658)(15.4092,0.003652)(15.4291,0.003644)(15.4491,0.003633)(15.4691,0.00362)(15.489,0.003605)(15.509,0.003587)(15.5289,0.003567)(15.5489,0.003548)(15.5689,0.003529)(15.5888,0.003509)(15.6088,0.003491)(15.6287,0.003474)(15.6487,0.00346)(15.6687,0.003449)(15.6886,0.00344)(15.7086,0.003431)(15.7285,0.003423)(15.7485,0.003416)(15.7685,0.003409)(15.7884,0.0034)(15.8084,0.00339)(15.8283,0.00338)(15.8483,0.003371)(15.8683,0.003362)(15.8882,0.003353)(15.9082,0.003345)(15.9281,0.003337)(15.9481,0.003331)(15.9681,0.003323)(15.988,0.003315)(16.008,0.003306)(16.0279,0.003297)(16.0479,0.003287)(16.0679,0.003276)(16.0878,0.003263)(16.1078,0.00325)(16.1277,0.003238)(16.1477,0.003226)(16.1677,0.003213)(16.1876,0.0032)(16.2076,0.003186)(16.2275,0.003174)(16.2475,0.00316)(16.2675,0.003145)(16.2874,0.003129)(16.3074,0.003114)(16.3273,0.003099)(16.3473,0.003084)(16.3673,0.003069)(16.3872,0.003054)(16.4072,0.00304)(16.4271,0.003028)(16.4471,0.003016)(16.4671,0.003003)(16.487,0.002992)(16.507,0.002981)(16.5269,0.002969)(16.5469,0.002956)(16.5669,0.00294)(16.5868,0.002924)(16.6068,0.002908)(16.6267,0.002892)(16.6467,0.002878)(16.6667,0.002864)(16.6866,0.002853)(16.7066,0.002844)(16.7265,0.002835)(16.7465,0.002826)(16.7665,0.002816)(16.7864,0.002806)(16.8064,0.002796)(16.8263,0.002783)(16.8463,0.002768)(16.8663,0.002753)(16.8862,0.002738)(16.9062,0.002723)(16.9261,0.002707)(16.9461,0.002691)(16.9661,0.002676)(16.986,0.002663)(17.006,0.00265)(17.0259,0.002637)(17.0459,0.002625)(17.0659,0.002615)(17.0858,0.002607)(17.1058,0.002601)(17.1257,0.002597)(17.1457,0.002595)(17.1657,0.002596)(17.1856,0.002599)(17.2056,0.0026)(17.2255,0.002601)(17.2455,0.0026)(17.2655,0.002599)(17.2854,0.002596)(17.3054,0.00259)(17.3253,0.002582)(17.3453,0.002574)(17.3653,0.002565)(17.3852,0.002556)(17.4052,0.002544)(17.4251,0.002532)(17.4451,0.002521)(17.4651,0.00251)(17.485,0.002497)(17.505,0.002484)(17.525,0.002471)(17.5449,0.002458)(17.5649,0.002445)(17.5848,0.002431)(17.6048,0.002417)(17.6248,0.002404)(17.6447,0.002392)(17.6647,0.002379)(17.6846,0.002367)(17.7046,0.002355)(17.7246,0.002345)(17.7445,0.002336)(17.7645,0.002328)(17.7844,0.002321)(17.8044,0.002314)(17.8244,0.002308)(17.8443,0.002302)(17.8643,0.002295)(17.8842,0.002287)(17.9042,0.002278)(17.9242,0.002269)(17.9441,0.00226)(17.9641,0.00225)(17.984,0.002241)(18.004,0.002233)(18.024,0.002227)(18.0439,0.002222)(18.0639,0.002217)(18.0838,0.002214)(18.1038,0.002211)(18.1238,0.00221)(18.1437,0.002209)(18.1637,0.002206)(18.1836,0.002203)(18.2036,0.0022)(18.2236,0.002196)(18.2435,0.00219)(18.2635,0.002183)(18.2834,0.002176)(18.3034,0.002168)(18.3234,0.002161)(18.3433,0.002153)(18.3633,0.002145)(18.3832,0.002138)(18.4032,0.002132)(18.4232,0.002125)(18.4431,0.002119)(18.4631,0.002113)(18.483,0.002108)(18.503,0.002104)(18.523,0.0021)(18.5429,0.002097)(18.5629,0.002093)(18.5828,0.00209)(18.6028,0.002087)(18.6228,0.002083)(18.6427,0.002077)(18.6627,0.00207)(18.6826,0.002063)(18.7026,0.002056)(18.7226,0.002048)(18.7425,0.002038)(18.7625,0.002029)(18.7824,0.00202)(18.8024,0.002012)(18.8224,0.002003)(18.8423,0.001994)(18.8623,0.001985)(18.8822,0.001978)(18.9022,0.001972)(18.9222,0.001966)(18.9421,0.00196)(18.9621,0.001956)(18.982,0.001954)(19.002,0.001953)(19.022,0.001951)(19.0419,0.00195)(19.0619,0.00195)(19.0818,0.001952)(19.1018,0.001952)(19.1218,0.001952)(19.1417,0.001952)(19.1617,0.001952)(19.1816,0.001952)(19.2016,0.00195)(19.2216,0.001948)(19.2415,0.001946)(19.2615,0.001945)(19.2814,0.001944)(19.3014,0.001941)(19.3214,0.001939)(19.3413,0.001936)(19.3613,0.001934)(19.3812,0.001931)(19.4012,0.001926)(19.4212,0.001922)(19.4411,0.001918)(19.4611,0.001914)(19.481,0.00191)(19.501,0.001904)(19.521,0.001899)(19.5409,0.001894)(19.5609,0.001888)(19.5808,0.001882)(19.6008,0.001874)(19.6208,0.001868)(19.6407,0.001863)(19.6607,0.001858)(19.6806,0.001854)(19.7006,0.00185)(19.7206,0.001848)(19.7405,0.001847)(19.7605,0.001846)(19.7804,0.001843)(19.8004,0.001841)(19.8204,0.001839)(19.8403,0.001837)(19.8603,0.001834)(19.8802,0.001829)(19.9002,0.001823)(19.9202,0.001816)(19.9401,0.001809)(19.9601,0.0018)(19.98,0.00179)(20,0.001781)};
 \end{axis}
@@
 We do not count inter-element communication or flux computations, as these are typically less expensive than volume operations (especially for higher polynomial degrees).
-In three dimensions, a degree $N$ GLL collocation scheme contains $(N+1)^3$ nodes.
-Two-point fluxes are computed between states at one node and states at $(N+1)$ additional nodes, resulting in a total of $(N+1)^4$ two-point flux evaluations in each dimension.
-For a staggered grid scheme, two-point fluxes are computed on a degree $(N+1)$ GLL grid consisting of $(N+2)^3$ nodes, resulting in $(N+2)^4$ flux evaluations per dimension.
-
-Gauss collocation schemes compute two-point fluxes on a grid of Gauss nodes, resulting in $(N+1)^4$ flux evaluations.
-However, one must also evaluate two-point fluxes between face nodes and volume nodes, and vice versa.
-Specifically, Gauss schemes perform two evaluations of two-point fluxes between each face node and the line of $(N+1)$ volume nodes normal to that face, resulting in an extra $4(N+1)^3$ flux evaluations per dimension.
-Because these flux evaluations correspond to the computation of $\bm{Q}^i_N\circ \bm{F}_S$, one can also exploit the skew symmetric nature of the off-diagonal blocks of $\bm{Q}^i_N$ to reduce the number of two-point flux evaluations to $2(N+1)^3$.
+The total number of flux computations for each scheme can also be reduced using the symmetry of $\bm{f}_S$; however, this does not affect relative differences in the number of two-point flux evaluations between GLL, Gauss, and staggered grid schemes.
+However, exploiting symmetry in the Gauss scheme is slightly more straightforward (compared to GLL and staggered grid schemes) due to the block structure of the decoupled SBP operator.
+
+In 3D, a degree $N$ GLL collocation scheme contains $(N+1)^3$ nodes.
+Two-point fluxes are computed between states at one node and states at $(N+1)$ additional nodes, resulting in a total of $3(N+1)^4$ two-point flux evaluations.
+For a staggered grid scheme, two-point fluxes are computed on a degree $(N+1)$ GLL grid consisting of $(N+2)^3$ nodes, resulting in $3(N+2)^4$ evaluations.
+
+Gauss collocation schemes compute two-point fluxes on a grid of Gauss nodes, resulting in $3(N+1)^4$ flux evaluations in 3D.
+One must also evaluate two-point fluxes between face nodes and volume nodes, and vice versa.
+As a result, Gauss schemes evaluate two-point fluxes twice between each face node and the line of $(N+1)$ volume nodes normal to that face, resulting in an extra $12(N+1)^3$ flux evaluations in 3D.
+%Because these flux evaluations correspond to the computation of $\bm{Q}^i_N\circ \bm{F}_S$, one can also exploit the skew symmetric nature of the off-diagonal blocks of $\bm{Q}^i_N$ to reduce the number of two-point flux evaluations to $2(N+1)^3$.
 
 We also consider costs associated with applying operator matrices.
-GLL collocation schemes require applying a one-dimensional differentiation matrix in each dimension, resulting in $O\LRp{(N+1)^4}$ operations per dimension in 3D.
-Staggered-grid schemes require $O\LRp{(N+2)^4}$ operations per dimension (corresponding to differentiation on a degree $(N+1)$ GLL grid) as well as $O\LRp{2(N+2)(N+1)^3}$ operations required for interpolation to and from a $(N+1)$ point Gauss grid to a $(N+2)$ point GLL grid in three dimensions.
-Gauss collocation schemes require only interpolation to and from face points, resulting in an additional $O\LRp{2(N+1)^3}$ operations per dimension per element.
+GLL and Gauss collocation schemes require applying a one-dimensional differentiation matrix to each line of nodes, resulting in $O\LRp{3(N+1)^4}$ operations in 3D.
+Staggered-grid schemes require $O\LRp{3(N+2)^4}$ operations (corresponding to differentiation in each coordinate on a degree $(N+1)$ GLL grid) as well as $O\LRp{6(N+2)(N+1)^3}$ operations required for interpolation to and from a $(N+1)$ point Gauss grid to a $(N+2)$ point GLL grid in three dimensions.
+Gauss collocation schemes require only interpolation to and from points on 6 faces, resulting in an additional $O\LRp{12(N+1)^3}$ operations per dimension per element.
+%\note{Maybe factor in dimensions.}
 
@@
 \centering
-\subfloat[Two-point flux evaluations]{
+\subfloat[Two-point flux evaluations per dimension]{
 \begin{tikzpicture}
@@
              ybar=2*\pgflinewidth,
-	bar width=2.5pt,
+	bar width=4pt,
 %	nodes near coords,
@@
 ] 
-%nodal runtimes
 \addplot coordinates{(1,48)(2,243)(3,768)(4,1875)(5,3888)(6,7203)(7,12288)};
@@
 \addplot coordinates{(1,144)(2,567)(3,1536)(4,3375)(5,6480)(6,11319)(7,18432)};
-\addplot coordinates{(1,96)(2,405)(3,1152)(4,2625)(5,5184)(6,9261)(7,15360)};
-\legend{GLL, Staggered, Gauss, Gauss (w/sym.)}
+%\addplot coordinates{(1,96)(2,405)(3,1152)(4,2625)(5,5184)(6,9261)(7,15360)};
+\legend{GLL, Staggered, Gauss}%, Gauss (w/sym.)}
 \end{axis}
@@
 }
-\subfloat[Matrix computation operations]{
+\subfloat[Est.\ matrix computation operations ]{
 \begin{tikzpicture}
@@
 %nodal runtimes
-\addplot coordinates{(1,96)(2,486)(3,1536)(4,3750)(5,7776)(6,14406)(7,24576)};
-\addplot coordinates{(1,582)(2,1968)(3,5030)(4,10776)(5,20454)(6,35552)(7,57798)};
-\addplot coordinates{(1,192)(2,810)(3,2304)(4,5250)(5,10368)(6,18522)(7,30720)};
+\addplot coordinates{(1,48)(2,243)(3,768)(4,1875)(5,3888)(6,7203)(7,12288)};
+\addplot coordinates{(1,387)(2,1416)(3,3795)(4,8388)(5,16275)(6,28752)(7,47331)};
+\addplot coordinates{(1,144)(2,567)(3,1536)(4,3375)(5,6480)(6,11319)(7,18432)};
 
@@
 \end{tikzpicture}}
-\caption{Comparison of GLL, staggered grid, and Gauss collocation schemes in terms of number of two-point flux evaluations and operations associated with matrix computations. }
+\caption{Comparison of GLL, staggered grid, and Gauss collocation schemes in terms of number of two-point flux evaluations and operations associated with matrix computations in three dimensions. }
 \label{fig:cost}
@@
 Figure~\ref{fig:cost} shows the estimated number of two-point flux evaluations and matrix operations for GLL, staggered grid, and Gauss collocation schemes in three dimensions.
-We observe that a straightforward implementation of Gauss collocation does not significantly reduce the number of flux evaluations compared to staggered grid schemes, though this can be improved by exploiting symmetry in the off-diagonal blocks of the decoupled SBP operator.
-Gauss collocation schemes see a much more significant gain over staggered grid schemes in terms of matrix computations.
+We observe that a straightforward implementation of Gauss collocation does not significantly reduce the number of flux evaluations compared to staggered grid schemes, though Gauss collocation schemes result in a  significantly smaller number of operations from matrix computations.
 
@@
 
-We note that results for Gauss collocation are similar to those attained by entropy stable staggered-grid schemes \cite{parsani2016entropy}.
-However, Gauss collocation schemes result in a lower number of two-point flux evaluations and floating point operations from matrix computations compared to staggered-grid methods.
+We note that results for Gauss collocation are similar to those attained by entropy stable staggered-grid schemes \cite{parsani2016entropy}, and require a similar number of two-point flux evaluations.
+However, Gauss collocation schemes result in a lower number floating point operations from matrix computations compared to staggered-grid methods.
 While a rigorous computational comparison between GLL and Gauss collocation schemes remains to be done, Gauss collocation schemes show significant improvements in accuracy compared to GLL collocation schemes on non-Cartesian meshes.
@@
 
+%\note{Mention possible application to non-conforming.}
+
 \section{Acknowledgments}
 
-Jesse Chan is supported by NSF DMS-1719818 and DMS-1712639.
+Jesse Chan is supported by NSF DMS-1719818 and DMS-1712639.  The authors thank Florian Hindenlang for helpful discussions and suggestions.
 
@@
 Let $\LRc{\phi_j}_{j=1}^{N_p}$ denote a set of $N_p$ basis functions, and let $\LRc{\bm{x}_i, \bm{w}_i}_{i = 1}^{N_q}$ denote a set of $N_q$ volume quadrature points and weights in $d$ dimensions.
+We also introduce the set of $N^f_q$ surface quadrature points and weights $\LRc{\bm{x}^f_i, \bm{w}^f_i}_{i=1}^{N^f_q}$.
 Then, $\bm{V}_q, \bm{V}_f$ are given as 
 \begin{align*}
-\LRp{\bm{V}_q}_{ij} &= \phi_j(\bm{x}_i), \quad 1 \leq i,j \leq N_q,\\
-\LRp{\bm{V}_f}_{ij} &= \phi_j\LRp{\bm{x}^f_i}, \quad 1 \leq i \leq N_q, \quad 1\leq j \leq N^f_q.
+\LRp{\bm{V}_q}_{ij} &= \phi_j(\bm{x}_i), \qquad 1 \leq i \leq N_q, \qquad 1 \leq j \leq N_p,\\
+\LRp{\bm{V}_f}_{ij} &= \phi_j\LRp{\bm{x}^f_i}, \qquad 1 \leq i \leq N^f_q, \qquad 1\leq j \leq N_p.
 \end{align*}
-These matrices can be used to define the quadrature-based $L^2$ projection matrix $\bm{P}_q$.  Let $\bm{W}$ denote the diagonal matrix of quadrature weights.  Then, 
+These matrices can be used to define the quadrature-based $L^2$ projection matrix $\bm{P}_q$.
+Let $\bm{W}, \bm{W}_f$ denote the diagonal matrix of volume and surface quadrature weights, respectively.
+Then,
 \[
@@
 By composing this matrix with interpolation and projection matrices, one can define differencing operators $\bm{D}_q^i = \bm{V}_q\bm{D}^i\bm{P}_q$ which map values at quadrature points to values of approximate derivatives at quadrature points.
-It was shown in \cite{chan2017discretely} that $\bm{Q}^i = \bm{W}\bm{D}_q^i$ satisfies a generalized SBP property involving the face interpolation and projection matrices $\bm{V}_f, \bm{P}_q$.
+Moreover, $\bm{Q}^i = \bm{W}\bm{D}_q^i$ satisfies a generalized SBP property involving the face interpolation and projection matrices $\bm{V}_f, \bm{P}_q$ \cite{chan2017discretely}.
 

### 2018-09-08 docs/ESDG_GQ/esdg_gq.tex
updating other experiments and docs

@@
 
-While entropy stable collocation schemes have been constructed on Gauss-like quadrature points without boundary nodes \cite{crean2017high}, the inter-element coupling terms for such schemes introduce an ``all-to-all'' coupling between degrees of freedom on two neighboring elements.
+While entropy stable collocation schemes have been constructed on Gauss-like quadrature points without boundary nodes \cite{crean2017high}, the inter-element coupling terms for such schemes introduce an ``all-to-all'' coupling between degrees of freedom on two neighboring elements in one dimension (on tensor product elements in higher dimensions, these coupling terms couple together lines of nodes).
 These coupling terms require evaluating two-point fluxes between solution states at all collocation nodes on two neighboring elements, resulting in significantly more communication and computation compared to collocation schemes based on point sets containing boundary nodes.
@@
 However, Gauss points have not been widely used to construct entropy stable discretizations due to the lack of efficient, stable, and high order accurate inter-element coupling terms, known as simultaneous approximation terms (SAT) in the finite difference literature \cite{fernandez2014review, crean2017high, fernandez2018simultaneous}.
-SATs for Gauss points are non-compact, in the sense that they introduce all-to-all coupling between degrees of freedom on neighboring elements.
+SATs for Gauss points are non-compact, in the sense that they introduce all-to-all coupling between degrees of freedom on neighboring elements in one dimension.
 This results in greater communication between elements, as well as a significantly larger number of two-point flux evaluations and floating point operations.
@@
 Evaluating (\ref{eq:ESgsbp}) requires computing two-point fluxes $\bm{f}_S$ between all nodes on two neighboring elements, which significantly increases both the computational work, as well as communication between neighboring elements.
-The main goal of this work is to circumvent the all-to-all coupling of degrees of freedom introduced by the SATs described in this section, which can be done through the use of ``decoupled'' SBP operators.
+This leads to all-to-all coupling between degrees of freedom in 1D, and to coupling along one-dimensional lines of nodes in higher dimensions due to the tensor product structure.
+
+The main goal of this work is to circumvent this tighter coupling of degrees of freedom introduced by the SATs described in this section, which can be done through the use of ``decoupled'' SBP operators.
 
@@
 \bm{S} &  \frac{1}{2}\bm{V}_f^T\bm{B} & & \\
- - \frac{1}{2}\bm{B}\bm{V}_f &  & & \frac{1}{2}\bm{e}_R\bm{e}_L^T\\
- \hline
+ - \frac{1}{2}\bm{B}\bm{V}_f &  & & \frac{1}{2}\bm{e}_R\bm{e}_L^T\\ [.2em]
+ \hline\\[-1.0em]
 && \bm{S} &  \frac{1}{2}\bm{V}_f^T\bm{B} \\
@@
 %    xmin=.0125, 
-    ymin=2.5e-8, ymax=5,
+    ymin=2.5e-11, ymax=5,
     legend pos=south east, legend cell align=left, legend style={font=\tiny},	
@@
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.553435)(0.0625,0.0625441)(0.03125,0.0117156)(0.015625,0.0022841)};
+coordinates{(0.125,0.553435)(0.0625,0.0625441)(0.03125,0.0117156)(0.015625,0.0022841)}; %(0.0078125,0.000362526)
 
@@
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.25082)(0.0625,0.0198863)(0.03125,0.000673152)(0.015625,2.98263e-05)};
+coordinates{(0.125,0.25082)(0.0625,0.0198863)(0.03125,0.000673152)(0.015625,2.98263e-05)}; %(0.0078125, 1.77421e-06)
 
@@
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.0768018)(0.0625,0.0020579)(0.03125,5.55322e-05)(0.015625,2.36108e-06)};
+coordinates{(0.125,0.0768018)(0.0625,0.0020579)(0.03125,5.55322e-05)(0.015625,2.36108e-06)}; % (0.0078125,8.55848e-08)
 
@@
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.0151231)(0.0625,0.000454595)(0.03125,1.00203e-05)(0.015625,6.57796e-08)};
-
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.0106752)(0.0625,0.000133284)(0.03125,7.97032e-07)(0.015625,NaN)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.00657096)(0.0625,9.22845e-05)(0.03125,3.60025e-07)(0.015625,NaN)};
-
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.00330917)(0.0625,1.67825e-05)(0.03125,1.32696e-07)(0.015625,NaN)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.00185273)(0.0625,1.21465e-05)(0.03125,1.14289e-07)(0.015625,NaN)};
+coordinates{(0.125,0.0151231)(0.0625,0.000454595)(0.03125,1.00203e-05)(0.015625,6.57796e-08)}; %(0.0078125, 9.32809e-10)
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.0106752)(0.0625,0.000133284)(0.03125,7.97032e-07)(0.015625,9.35152e-09)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.00657096)(0.0625,9.22845e-05)(0.03125,3.60025e-07)(0.015625,3.63974e-09)}; %(0.0078125, 1.23726e-10)
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.00330917)(0.0625,1.67825e-05)(0.03125,1.32696e-07)(0.015625,4.73518e-10)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.00185273)(0.0625,1.21465e-05)(0.03125,1.14289e-07)(0.015625,1.72358e-10)}; 
 \end{loglogaxis}
@@
 %    xmin=.0125, 
-    ymin=2.5e-8, ymax=5,
+    ymin=2.5e-11, ymax=5,
       legend pos=south east, legend cell align=left, legend style={font=\tiny},	
@@
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.0112838)(0.0625,0.000219049)(0.03125,1.84304e-06)(0.015625,NaN)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.00809138)(0.0625,8.23415e-05)(0.03125,3.95565e-07)(0.015625,NaN)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.00388316)(0.0625,2.55074e-05)(0.03125,2.68102e-07)(0.015625,NaN)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.00222029)(0.0625,1.33111e-05)(0.03125,9.72405e-08)(0.015625,NaN)};
+coordinates{(0.125,0.0112838)(0.0625,0.000219049)(0.03125,1.84304e-06)(0.015625,2.40402e-08)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.00809138)(0.0625,8.23415e-05)(0.03125,3.95565e-07)(0.015625,4.01268e-09)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.00388316)(0.0625,2.55074e-05)(0.03125,2.68102e-07)(0.015625,9.13754e-10)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.00222029)(0.0625,1.33111e-05)(0.03125,9.72405e-08)(0.015625,1.81386e-10)};
 
@@
 %    xmin=.0125,
-    ymin=2.5e-8, ymax=5,    
+    ymin=2.5e-11, ymax=5,
     legend pos=south east, legend cell align=left, legend style={font=\tiny},	
@@
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.0378013)(0.0625,0.0007649)(0.03125,4.15848e-06)(0.015625,NaN)};
-
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.026577)(0.0625,0.000589552)(0.03125,5.3039e-06)(0.015625,NaN)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.0154563)(0.0625,0.000146207)(0.03125,8.83156e-07)(0.015625,NaN)};
+coordinates{(0.125,0.0378013)(0.0625,0.0007649)(0.03125,4.15848e-06)(0.015625,1.8685e-08)};
+
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.026577)(0.0625,0.000589552)(0.03125,5.3039e-06)(0.015625,2.23719e-08)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.125,0.0154563)(0.0625,0.000146207)(0.03125,8.83156e-07)(0.015625,1.18829e-09)};
 
@@
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.125,0.150916)(0.0625,0.00538908)(0.03125,2.79182e-05)};
+coordinates{(0.125,0.150916)(0.0625,0.00538908)(0.03125,2.79182e-05)(.015625,1.98845e-07)};
 \end{loglogaxis}
@@
 
-Figure~\ref{fig:err3d} shows the $L^2$ errors for degrees $N = 2, 3, 4$.
+Figure~\ref{fig:err3d} shows the $L^2$ errors for degrees $N = 2,\ldots, 7$.
 As in the 2D case, Gauss collocation schemes produce smaller errors than GLL collocation schemes, especially on on curved meshes.
@@
 coordinates{(0.25,1.53203)(0.125,0.23283)(0.0625,0.0417033)(0.03125,0.00806707)};
+
 \addplot+[semithick, mark options={solid, fill=markercolor}]
@@
 coordinates{(0.25,0.562488)(0.125,0.0864096)(0.0625,0.00268061)(0.03125,0.000136725)};
+
 \addplot+[semithick, mark options={solid, fill=markercolor}]
@@
 coordinates{(0.25,0.239027)(0.125,0.0077433)(0.0625,0.000219705)(0.03125,2.67589e-05)};
+
+% N = 5
 \addplot+[semithick, mark options={solid, fill=markercolor}]
@@
 coordinates{(0.25,0.0547127)(0.125,0.00242675)(0.0625,4.2962e-05)(0.03125,NaN)};
+% N = 6
 \addplot+[semithick, mark options={solid, fill=markercolor}]
@@
 coordinates{(0.25,0.0246519)(0.125,0.000324802)(0.0625,2.44009e-05)(0.03125,NaN)};
+% N = 7
 \addplot+[semithick, mark options={solid, fill=markercolor}]
@@
 However, Gauss collocation schemes result in a lower number floating point operations from matrix computations compared to staggered-grid methods.
-While a rigorous computational comparison between GLL and Gauss collocation schemes remains to be done, Gauss collocation schemes show significant improvements in accuracy compared to GLL collocation schemes on non-Cartesian meshes.
+Finally, while a rigorous computational comparison between GLL and Gauss collocation schemes remains to be done, Gauss collocation schemes show significant improvements in accuracy compared to GLL collocation schemes on non-Cartesian meshes.
 In particular, for sufficiently warped curvilinear mappings, degree $N$ Gauss collocation schemes achieve an accuracy comparable to degree $(N+1)$ GLL collocation schemes in two and three dimensions.

### 2018-09-08 docs/skewform/main.tex
updating other experiments and docs

@@
 \newtheorem{assumption}{Assumption}
+
+%\documentclass[preprint,10pt]{elsarticle}
+%\usepackage{amsmath,amssymb,amsfonts,amsthm}
+%\theoremstyle{definition}
+%\newtheorem{definition}{Definition}
+%\theoremstyle{lemma}
+%\newtheorem{lemma}{Lemma}
+%\newtheorem*{remark}{Remark}
+%\theoremstyle{theorem}
+%\newtheorem{theorem}{Theorem}
+%\theoremstyle{assumption}
+%\newtheorem{assumption}{Assumption}
+%\usepackage{fullpage}
 
@@
 These skew-symmetric formulations avoid the use of a ``strong'' SBP property, and require only that operators exactly differentiate constants and satisfy a discrete form of the fundamental theorem of calculus.
-We conclude by discussing the applicability of these new operators for entropy stable schemes on hybrid meshes.% and non-conforming meshes.
+We conclude by discussing mortar-based implementations of skew-symmetric formulations for entropy stable schemes on hybrid meshes.
 \end{abstract}
 
+%\tableofcontents
+
 \section{Introduction}
@@
 High order methods for the simulation of time-dependent compressible flow have the potential to achieve higher levels of accuracy at lower costs compared to current low order schemes \cite{wang2013high}.
-In addition to superior accuracy and efficiency on modern computing architectures, the low numerical dispersion and dissipation of high order methods \cite{ainsworth2004dispersive} enables the accurate propagation of waves over long distances and time scales.
-The same properties also make high order methods attractive for the resolution of unsteady phenomena such as vorticular and turbulent flows, which are sensitive to numerical dissipation \cite{visbal1999high, wang2013high}.
-However, while high order methods are often provably stable for wave problems, high order schemes for nonlinear conservation laws have consistently been hampered by problems of instability.
-
-When applied to nonlinear conservation laws, high order methods can experience artificial growth and blow-up near under-resolved features such as shocks, turbulence, or boundary layers.
+In addition to superior accuracy, % and efficiency on modern computing architectures,
+the low numerical dispersion and dissipation of high order methods \cite{ainsworth2004dispersive} enables the accurate propagation of waves over long distances and time scales.
+The same properties also make high order methods attractive for unsteady phenomena such as vorticular and turbulent flows, which are sensitive to numerical dissipation \cite{visbal1999high, wang2013high}.
+
+However, 
+%while high order methods are often provably stable for wave problems, high order schemes for nonlinear conservation laws have consistently been hampered by problems of instability.  
+when applied to nonlinear conservation laws, high order methods can experience artificial growth and blow-up near under-resolved features such as shocks, turbulence, or boundary layers.
 In practice, the application of high order methods to practical problems requires shock capturing and stabilization techniques (such as artificial viscosity) or solution regularization (such as filtering or limiting) to prevent solution blow-up.
@@
 These issues have motivated the introduction of high order \textit{entropy stable} discretizations, which satisfy a semi-discrete entropy inequality while maintaining high order accuracy in smooth regions.
-Proofs of continuous entropy inequalities rely on the chain rule.
-In contrast, these discrete entropy inequalities account for the loss of the chain rule due to effects such as quadrature errors, which are incurred when applying polynomially exact quadrature rules to nonlinear and rational integrands within the DG formulation.
+Proofs of continuous entropy inequalities rely on the chain rule, which does not hold discretely due to effects such as quadrature error.
+By using an approach referred to as ``flux differencing'', entropy stable schemes account for the loss of the chain rule while maintaining a semi-discrete analogue of the continuous entropy inequality.
 These schemes were first introduced as high order collocation methods on tensor product elements in \cite{fisher2013high, carpenter2014entropy, gassner2016split, gassner2017br1}, and were extended to simplicial elements in \cite{crean2017high, chen2017entropy, crean2018entropy, chan2017discretely, chan2018discretely}.
-These methods have also been extended to staggered grid \cite{parsani2016entropy}, generalized SBP and Gauss nodes \cite{chan2018efficient} and non-conforming meshes \cite{friedrich2017entropy}.
-Entropy stable boundary conditions have also been determined for the compressible Euler and Navier-Stokes equations \cite{parsani2015entropy, svard2018entropy}.
-
-\note{Finish: add Section description.
+Entropy stable methods have also been extended to a variety of other discretization settings, including staggered grids \cite{parsani2016entropy}, Gauss-Legendre collocation \cite{chan2018efficient}, and non-conforming meshes \cite{friedrich2017entropy}.
+
+Entropy stable discretizations are built upon flux differencing and a summation-by-parts (SBP) property, which serves as a discrete analogue of integration by parts for quadrature-based discretization matrices.
+However, the SBP property does not hold for some under-integrated quadrature rules, which naturally arise in certain discretization settings.
+For example, on hybrid meshes consisting of both quadrilateral and triangular elements, it is convenient to utilize the same quadrature rule on shared faces between different element types.
+On degree $N$ tensor product elements, a popular choice of quadrature is an $(N+1)$-point Gauss-Legendre-Lobatto (GLL) rule.
+When both volume and surface integrals are approximated using $(N+1)$ point GLL quadrature rules, the SBP property holds, despite the fact that GLL quadrature is inexact for the integrands which appear in finite element formulations \cite{fisher2013high}.
+However, while GLL quadrature induces an SBP property on quadrilateral elements, but does not always result in an SBP property if used for triangular elements \cite{chan2017discretely}.
+%A similar issue arises on non-conforming meshes, where a composite ``split-side'' quadrature naturally when coupling larger elements to smaller elements through a non-conforming interface \cite{kozdon2018energy}.
+For quadrilaterals utilizing a GLL volume quadrature, the use of a composite quadrature on a non-conforming face also results in the loss of the SBP property \cite{chan2017discretely}.
+
+This work proposes an alternative formulation which utilizes a skew-symmetric form of flux differencing.
+Under such a formulation, the proof of entropy stability requires only a weaker ``variational'' form of the SBP property, which holds under a more general class of quadrature rules compared to the decoupled SBP property introduced in \cite{chan2017discretely, chan2018discretely}.
+We show that the skew-symmetric formulation is entropy stable, locally conservative, and free-stream preserving on curved elements, and confirm theoretical results with numerical experiments on triangular, quadrilateral, and hybrid triangular-quadrilateral meshes.
+
+It should be noted that a similar approach to entropy stable discretizations was introduced within a finite difference framework \cite{chen2017entropy, crean2018entropy} using multidimensional differencing operators which are constructed to satisfy accuracy conditions and an SBP property \cite{hicken2016multidimensional}.
+These operators do not correspond to any specific basis or approximation space, but can be shown to exist for nodal points corresponding to sufficiently accurate choices of volume and surface quadrature.
+The formulations presented in this work require similar accuracy conditions, but differ from those of \cite{chen2017entropy, crean2018entropy} in that they build upon an explicit approximation space.
+In contrast with traditional SBP coupling terms \cite{crean2017high}, the use of decoupled SBP operators also accomodates general choices of volume and surface quadrature (e.g.\ volume quadratures without boundary nodes and over-integrated rules) while ensuring compact coupling conditions between neighboring elements.
+
+The structure of the paper is as follows: Section~\ref{sec:nonlin} describes the continuous entropy inequality which we aim to replicate discretely.
+Section~\ref{sec:approx} and Section~\ref{sec:sbp} introduce polynomial approximation spaces and quadrature-based SBP operators on simplicial and tensor product elements.
+Section~\ref{sec:skew1} then introduces the variational SBP property and an entropy conservative skew-symmetric formulation.
+Section~\ref{sec:skew2} extends the skew-symmetric formulation to curved elements, and Section~\ref{sec:num} presents numerical experiments which verify the theoretical assumptions, stability, and accuracy of the proposed formulations.
+
+%\note{Finish: add Section description.
 Talk about situations where the SBP or decoupled SBP property doesn't hold: triangles with reduced surface quadrature, GLL quads with GQ face quadratures, and hybrid couplings.}
@@
 \section{Entropy stability for systems of nonlinear conservation laws}
+\label{sec:nonlin}
 
@@
 \section{Polynomial approximation spaces}
+\label{sec:approx}
 
@@
 \section{Quadrature-based matrices and decoupled SBP operators}
-
+\label{sec:sbp}
 %\note{TODO: Standardize hat notation for operators/normals: can refer to operators on reference element and do $k$ superscript for physical elements, or do hats on reference element.}
@@
 \begin{gather}
-\bm{D}^i_N = \bm{W}_N^{-1} \bm{Q}^i_N, \qquad \bm{W}_N = \LRp{\begin{array}{cc}\bm{W}&\\ & \bm{W}_f\end{array}},\label{eq:QN}\\
-\bm{Q}^i_N  = \LRs{
+\bm{D}^i_N = \bm{W}_N^{-1} \bm{Q}^i_N, \qquad \bm{W}_N = \LRp{\begin{array}{cc}\bm{W}&\\ & \bm{W}_f\end{array}}, \label{eq:QN}\\
+\bm{E} = \bm{V}_f\bm{P}_q, \qquad \bm{B}^i = \bm{W}_f \diag{\hat{\bm{n}}_i}, \qquad \bm{Q}^i_N  = \LRs{
 \begin{array}{cc}
-\bm{Q}^i - \frac{1}{2}\LRp{\bm{V}_f \bm{P}_q}^T \bm{W}_f {\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q &  \frac{1}{2}\LRp{\bm{V}_f \bm{P}_q}^T \bm{W}_f {\rm diag}(\hat{\bm{n}}_i)\\
--\frac{1}{2}\bm{W}_f{\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q & \frac{1}{2} \bm{W}_f{\rm diag}(\hat{\bm{n}}_i)
+\bm{Q}^i - \frac{1}{2}\bm{E}^T\bm{B}^i\bm{E} &  \frac{1}{2}\bm{E}^T\bm{B}^i\\
+-\frac{1}{2}\bm{B}^i\bm{E} & \frac{1}{2} \bm{B}^i
 \end{array}}.  \nonumber
 \end{gather}
-When combined with projection and lifting matrices, $\bm{D}^i_N$ produces a high order polynomial approximation of $f\pd{g}{x_i}$. 
+Here, $\bm{B}^i$ is a boundary matrix, and $\bm{E}$ denotes the extrapolation matrix which maps values at volume quadrature points to values at surface quadrature points using quadrature-based $L^2$ projection and polynomial interpolation.
+
+It was shown in \cite{chan2017discretely, chan2018efficient} that, when combined with projection and lifting matrices, $\bm{D}^i_N$ produces a high order polynomial approximation of $f\pd{g}{x_i}$. 
 Let $f, g$ be differentiable functions, and let $\bm{f}_i = f(\bm{x}_i)$, $\bm{g}_i = g(\bm{x}_i)$ denote values of $f,g$ at volume and surface quadrature points.  Then,
@@
 \end{gather}
-The approximation can also be interpreted as solving a variational problem.
+%The approximation can also be interpreted as solving a variational problem.
 Let $n_i$ be the $i$th component of the unit normal on $\hat{D}$.
 Then (\ref{eq:dnapprox}) is equivalent to finding $u(\bm{x}) \approx f\pd{g}{x_i}$ such that, for all $v\in V^N\LRp{\hat{D}}$,
-\begin{align*}
-\int_{\hat{D}} u(\bm{x})v(\bm{x}) = \int_{\hat{D}}{g\pd{\Pi_Nf}{x_i}v} + \int_{\partial \hat{D}}{(f-\Pi_Nf)\frac{\LRp{gv + \Pi_N(gv)}}{2}}\hat{n}_i\hat{J}_f.
-\end{align*}
+%\begin{align*}
+%\int_{\hat{D}} u(\bm{x})v(\bm{x}) = \int_{\hat{D}}{g\pd{\Pi_Nf}{x_i}v} + \int_{\partial \hat{D}}{(f-\Pi_Nf)\frac{\LRp{gv + \Pi_N(gv)}}{2}}\hat{n}_i\hat{J}_f.
+%\end{align*}
 
@@
 \]
-are integrated exactly for all $u,v \in V^N\LRp{\hat{D}}$ and $i = 1,\ldots, d$.  Then, the decoupled SBP operator $\bm{D}^i_N$ satisfies a summation by parts property:
+are integrated exactly for all $u,v \in V^N\LRp{\hat{D}}$ and $i = 1,\ldots, d$.  Then, the decoupled SBP operator $\bm{D}^i_N$ (\ref{eq:QN}) satisfies a summation by parts property:
 \begin{gather}
-\bm{Q}^i_N+\LRp{\bm{Q}^i_N}^T = \LRp{\begin{array}{cc}\bm{0}&\\ & \bm{W}_f {\rm diag}(\bm{n}_i)\end{array}} = \bm{B}_N\label{eq:dsbp}
+\bm{Q}^i_N+\LRp{\bm{Q}^i_N}^T = \bm{B}^i_N, \qquad \bm{B}^i_N = \LRp{\begin{array}{cc}\bm{0}&\\ & \bm{B}^i\end{array}}.\label{eq:dsbp}
 \end{gather}
@@
 For example, on simplicial elements, (\ref{eq:dsbp}) holds if the volume quadrature is exact for polynomial integrands of total degree $(2N-1)$, and the surface integral is exact for degree $2N$ polynomials on each face.
-Tensor product elements require stricter conditions: (\ref{eq:dsbp}) holds if both the volume and surface quadratures are exact for polynomial integrands of degree $2N$ in each coordinate, due to the fact that derivatives of $u\in Q^N$ are degree $(N-1)$ with respect to one coordinate and degree $N$ with respect to others.
+Tensor product elements require stricter conditions: (\ref{eq:dsbp}) holds if both the volume and surface quadratures are exact for polynomial integrands of degree $2N$ in each coordinate, due to the fact that derivatives of $u\in Q^N$ are degree $(N-1)$ polynomials with respect to one coordinate and degree $N$ with respect to others.
 
@@
 
-\section{Skew-symmetric entropy conservative formulations}
-
-While the SBP property has been used to derive entropy stable schemes, the SBP property is difficult or impossible to enforce in certain discretization settings, such as hybrid and non-conforming meshes arises.
+\section{Skew-symmetric entropy conservative formulations on a single element}
+\label{sec:skew1}
+While the SBP property has been used to derive entropy stable schemes, the strong SBP property (\ref{eq:dsbp}) is difficult to enforce in certain discretization settings, such as hybrid and non-conforming meshes.
 This difficulty is a result of the choices of volume and surface quadrature which naturally arise in these settings.
@@
 \subfloat[Degree $2N$ volume quadrature, GLL surface quadrature]{\includegraphics[width=.4\textwidth]{figs/trigll.png}\label{subfig:trigll}}
-\caption{Volume and surface quadrature pairs which do not satisfy the assumptions of Lemma~\ref{lemma:dsbp}, and thus do not possess the decoupled SBP property (\ref{eq:dsbp}). }
+\caption{Volume and surface quadrature pairs which do not satisfy the assumptions of Lemma~\ref{lemma:dsbp}, and thus do not possess the decoupled SBP property (\ref{eq:dsbp}).
+Volume quadrature nodes are drawn as circles, while surface quadrature nodes are drawn as squares.}
 \label{fig:sbploss}
@@
 \subfloat[Incompatible surface quadrature on the quadrilateral element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D_GQ.png}\label{subfig:hybrid2}}
-\caption{Examples of interface couplings which do not result in a decoupled SBP property (\ref{eq:dsbp}).  }
+\caption{Examples of interface couplings which do not result in a decoupled SBP property (\ref{eq:dsbp}).
+Volume quadrature nodes are drawn as circles, while surface quadrature nodes are drawn as  squares.
+}
 \label{fig:hybrid}
@@
 The property (\ref{eq:dsbp}) (which we will refer to as the ``strong'' SBP property) relates the polynomial exactness of specific quadrature rules to algebraic properties of quadrature-based matrices.
-We will relax accuracy conditions on these quadrature rules by utilizing not the strong SBP property, but a weaker ``variational'' version of the SBP property.
-This variational property relies on the assumption that the volume and surface quadrature rules are exact for polynomials of certain degrees.
-
+We will relax accuracy conditions on these quadrature rules by utilizing a weaker ``variational'' version of the SBP property variational, which relies on the exactness of volume and surface quadrature rules only for specific polynomials.
+
+On tensor product elements, we restrict ourselves to isotropic volume quadrature rules which are construced from tensor products of one-dimensional quadrature formulas.
+For the remainder of this work, the degree of the multi-dimensional quadrature rule on tensor product elements will refer to the degree of exactness of the one-dimensional rule.
+For example, we refer to the quadrature rule constructed through a tensor product of one-dimensional $(N+1)$-point GLL quadrature rules as a degree $(2N-1)$ quadrature rule.
+This choice of quadrature is sufficient to guarantee that the mass matrix is positive definite \cite{canuto2007spectral}.
+
+Throughout the remainder of this work, we make the following assumptions:
 \begin{assumption}
 \label{ass:quad}
-Let $v(\bm{x})$ be a polynomial such that $v \in V^{M_{\rm vol}}$ with surface trace $\LRl{v}_{\partial \hat{D}} \in V^{M_{\rm face}}\LRp{f}$ for $f \in \partial \hat{D}$, where $M_{\rm vol}, M_{\rm face} \leq N$ are two non-negative integers.
-We assume that the volume and surface quadrature rules are sufficiently accurate such that
+Let $v \in V^{N}$ denote a fixed polynomial.  We assume that: 
 \begin{enumerate}
-\item the mass matrix $\bm{M}$ is positive definite, 
-\item the volume and surface quadratures exactly integrate the quantities $\int_{\hat{D}} \pd{u}{x_i} v$ and $\int_{\partial \hat{D}} u v n_i$ for all $u \in V^N\LRp{\hat{D}}$, $i = 1,\ldots, d$, and $f \in \partial \hat{D}$.
+\item the mass matrix $\bm{M}$ is positive definite under the volume quadrature rule,
+\item the volume quadrature rule is exact for integrals of the form\\$\int_{\hat{D}} \pd{u}{\hat{x}_j} v$ for all $u \in V^N\LRp{\hat{D}}$, $j = 1,\ldots, d$.
+\item the surface quadrature rule is exact for integrals of the form\\$\int_{\partial \hat{D}} u v n_j$ for all $u \in V^N\LRp{\hat{D}}$, $j = 1,\ldots, d$, and $f \in \partial \hat{D}$.  
 \end{enumerate}
@@
 
-This section presumes that Assumption~\ref{ass:quad} is satisfied for general  $M_{\rm vol}, M_{\rm surf}$.
-In Sections~\ref{sec:affine} and \ref{sec:curved}, specific values of $M_{\rm vol}, M_{\rm surf}$ will be motivated by proofs of entropy stability, and Assumption~\ref{ass:quad} will be verified for specific choices of volume and surface quadrature on simplicial and tensor product elements.
-%Suppose Assumption~\ref{ass:quad} is satisfied, and let $\bm{v}_q$ denote the vector of values of $v(\bm{x})$ at volume quadrature points.  Then, for an arbitrary vector $\bm{u}_q$, 
-%\[
-%\bm{v}_q^T \bm{Q}_i \bm{u}_q = \int_{\hat{D}} \pd{u}{x_i} v, \qquad \LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f \diag{\hat{\bm{n}_i}} \bm{V}_f\bm{P}_q\bm{u}_q =  \int_{\partial \hat{D}} \Pi_N u v n_i,
-%\]
-%where we have used that, because $v$ is a polynomial of degree $N$, it is equal to its own $L^2$ projection.
-
-The conditions in Assumption~\ref{ass:quad} are sufficient to ensure that the following variational SBP property holds:
+The conditions of Assumption~\ref{ass:quad} are relatively standard within the SBP literature \cite{hicken2016multidimensional, chan2017discretely, crean2018entropy}, though they have not previously depended on the specific choice of polynomial $v(\bm{x})$.
+The strong SBP property (\ref{eq:dsbp}) can be derived if Assumption~\ref{ass:quad} is satisfied for all $v \in V^N$.
+In contrast, we will derive a weaker SBP property which assumes only that Assumption~\ref{ass:quad} holds for a few specific polynomials $v$.
+In Sections~\ref{sec:singleelem} and \ref{sec:curved}, specific polynomials $v$ will be motivated by proofs of entropy stability, and we will present examples of volume and surface quadrature rules on simplicial and tensor product elements which satisfy Assumption~\ref{ass:quad} for these choices of $v$.
+% the quantities $\int_{\hat{D}} \pd{u}{x_i} v$ and
+
+%This section presumes that Assumption~\ref{ass:quad} is satisfied for general  $M_{\rm vol}, M_{\rm surf}$.
+In Sections~\ref{sec:singleelem} and \ref{sec:curved}, specific values of $M_{\rm vol}, M_{\rm surf}$ will be motivated by proofs of entropy stability, and Assumption~\ref{ass:quad} will be verified for specific choices of volume and surface quadrature on simplicial and tensor product elements.
+
+The conditions of Assumption~\ref{ass:quad} imply the following variational SBP property:
 \begin{lemma}
@@
 \[
-\bm{v}_q^T\bm{Q}_i \bm{u}_q = \bm{v}_q^T\LRp{ \LRp{\bm{V}_f\bm{P}_q}^T \bm{W}_f \diag{\hat{\bm{n}}_i}\bm{V}_f\bm{P}_q - \bm{Q}_i^T}\bm{u}_q.
+\bm{v}_q^T\bm{Q}_i \bm{u}_q = \bm{v}_q^T\LRp{ {\bm{E}}^T \bm{B}^i\bm{E} - \bm{Q}_i^T}\bm{u}_q.
 \]
@@
 &\int_{\partial \hat{D}} (\Pi_N u) v \hat{n}_i -   \int_{\hat{D}} \LRp{\Pi_N u} \pd{v}{\hat{x}_i} =\\
- &\bm{v}_f^T\bm{W}_f{\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q\bm{u}_q - \LRp{\bm{V}_q\bm{D}^i\bm{P}_q\bm{v}_q}^T\bm{W} \bm{V}_q\bm{P}_q\bm{u}_q\\
-=& \bm{v}_q^T \LRp{\bm{V}_f\bm{P}_q}^T\bm{W}_f{\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q\bm{u}_q - \bm{v}_q^T \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T \bm{W} \bm{V}_q\bm{P}_q\bm{u}_q,
-\end{align*}
-where we have used that, since $v \in V^N$, $\bm{v}_f = \bm{V}_f\bm{P}_q\bm{v}_q$.  The proof is completed by noting that 
+ &\bm{v}_f^T\bm{B}^i \bm{E}\bm{u}_q - \LRp{\bm{V}_q\bm{D}^i\bm{P}_q\bm{v}_q}^T\bm{W} \bm{V}_q\bm{P}_q\bm{u}_q\\
+=& \bm{v}_q^T {\bm{E}}^T\bm{B}^i \bm{E}\bm{u}_q - \bm{v}_q^T \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T \bm{W} \bm{V}_q\bm{P}_q\bm{u}_q,
+\end{align*}
+where we have used that, since $v \in V^N$, $\bm{v}_f = \bm{V}_f\bm{P}_q\bm{v}_q = \bm{E}\bm{v}_q$.  The proof is completed by noting that 
 \[
@@
 \begin{align*}
-\bm{v}^T\bm{Q}^i_N\bm{u} &= \bm{v}_q^T\bm{Q}_i \bm{u}_q - \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}}_i \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}}_i \bm{u}_f\\
-& - \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}}_i \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} + \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}}_i \bm{u}_f.
+\bm{v}^T\bm{Q}^i_N\bm{u} &= \bm{v}_q^T\bm{Q}_i \bm{u}_q - \frac{1}{2}\LRp{\bm{E}\bm{v}_q}^T\bm{B}^i \LRp{\bm{E}\bm{u}_q} + \frac{1}{2}\LRp{\bm{E}\bm{v}_q}^T\bm{B}^i \bm{u}_f\\
+& - \frac{1}{2}\bm{v}_f^T\bm{B}^i \LRp{\bm{E}\bm{u}_q} + \frac{1}{2}\bm{v}_f^T\bm{B}^i \bm{u}_f.
 \end{align*}
@@
 \begin{align*}
-\bm{v}^T\bm{Q}^i_N\bm{u} =& -\bm{v}_q^T\bm{Q}^T_i \bm{u}_q + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}}_i \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}}_i \bm{u}_f\\
-&\qquad - \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}}_i \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} - \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}}_i \bm{u}_f + \bm{v}_f^T\bm{W}_f\hat{\bm{n}}_i \bm{u}_f\\
+\bm{v}^T\bm{Q}^i_N\bm{u} =& -\bm{v}_q^T\bm{Q}^T_i \bm{u}_q + \frac{1}{2}\LRp{\bm{E}\bm{v}_q}^T\bm{B}^i \LRp{\bm{E}\bm{u}_q} + \frac{1}{2}\LRp{\bm{E}\bm{v}_q}^T\bm{B}^i \bm{u}_f\\
+&\qquad - \frac{1}{2}\bm{v}_f^T\bm{B}^i \LRp{\bm{E}\bm{u}_q} - \frac{1}{2}\bm{v}_f^T\bm{B}^i \bm{u}_f + \bm{v}_f^T\bm{B}^i \bm{u}_f\\
 =& \begin{pmatrix} \bm{v}_q\\ \bm{v}_f\end{pmatrix}^T 
@@
 \bm{0}& \\
-& \bm{W}_f\diag{\hat{\bm{n}}_i}
-\end{pmatrix}\right.\\
-&+
+& \bm{B}^i
+\end{pmatrix}\right.+
 \left.\begin{pmatrix}
--\bm{Q}_i^T + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q}^T\bm{W}_f\hat{\bm{n}}_i \bm{V}_f\bm{P}_q & -\frac{1}{2} \LRp{\bm{W}_f\hat{\bm{n}}_i \bm{V}_f\bm{P}_q}^T\\
-\frac{1}{2}\bm{W}_f\hat{\bm{n}}_i \bm{V}_f\bm{P}_q & -\frac{1}{2}\bm{W}_f\hat{\bm{n}}_i
+-\bm{Q}_i^T + \frac{1}{2}{\bm{E}}^T\bm{B}^i \bm{E} & -\frac{1}{2} \LRp{\bm{B}^i \bm{E}}^T\\
+\frac{1}{2}\bm{B}^i \bm{E} & -\frac{1}{2}\bm{B}^i
 \end{pmatrix}  \right)
@@
 
-\subsection{Entropy stability on affine meshes}
-\label{sec:affine}
-
-In this section, we construct so-called ``entropy stable'' schemes on the reference element $\hat{D}$ and affine meshes.
+\subsection{Entropy stability on a reference element}
+\label{sec:singleelem}
+
+In this section, we construct so-called ``entropy stable'' schemes on the reference element $\hat{D}$.
 These methods ensure that the entropy inequality (\ref{eq:entropyineq}) is satisfied discretely by avoiding the use of the chain rule in the proof of entropy dissipation.
@@
 \begin{gather}
-\bm{f}_S(\bm{u},\bm{u}) = \bm{f}(\bm{u}), \qquad \text{(consistency)}\\
-\bm{f}_S(\bm{u}_L,\bm{u}_R) = \bm{f}_S(\bm{u}_R,\bm{u}_R), \qquad \text{(symmetry)}\nonumber\\
-\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}_S(\bm{u}_L,\bm{u}_R) = \psi(\bm{u}_L) - \psi(\bm{u}_R), \qquad \text{(conservation)}\nonumber
+\bm{f}^i_S(\bm{u},\bm{u}) = \bm{f}_i(\bm{u}), \qquad \text{(consistency)}\\
+\bm{f}^i_S(\bm{u}_L,\bm{u}_R) = \bm{f}^i_S(\bm{u}_R,\bm{u}_R), \qquad \text{(symmetry)}\nonumber\\
+\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}^i_S(\bm{u}_L,\bm{u}_R) = \psi_i(\bm{u}_L) - \psi_i(\bm{u}_R), \qquad \text{(conservation)}\nonumber
 \label{eq:esflux}
 \end{gather}
-The construction of entropy stable schemes will utilize (\ref{eq:esflux}) in discretizations of both volume and surface terms in a DG formulation.
+for $j = 1,\ldots, d$.  The construction of entropy stable schemes will utilize (\ref{eq:esflux}) in discretizations of both volume and surface terms in a DG formulation.
 
@@
 
-\begin{corollary}{(Discrete FTC and exact differentiation of constants)}
+\begin{lemma}{(Discrete FTC and exact differentiation of constants)}
 \label{lemma:sbpcor}
-Suppose Assumption~\ref{ass:quad} holds for $M_{\rm vol} = M_{\rm surf} = 0$ (i.e.\ for $v(\bm{x})$ constant).  
+Suppose Assumption~\ref{ass:quad} holds for $v(\bm{x}) = 1$.  
 Let $u(\bm{x})$ be some $L^2$ integrable function, and let $\bm{u}$ denote the vector of values of $u(\bm{x})$ at volume and surface quadrature points. Then, 
@@
 \]
-\end{corollary}
+\end{lemma}
 \begin{proof}
 The proof that $\bm{Q}^i_N \bm{1} = \bm{0}$ follows from the property that polynomials are equal to their $L^2$ projection, and is identical to that of \cite{chan2017discretely,chan2018discretely}.
-The proof of the first equality follows from $\bm{Q}^i_N \bm{1} = \bm{0}$ and Lemma~\ref{lemma:vdsbp} with $M=0$
+The proof of the first equality follows from $\bm{Q}^i_N \bm{1} = \bm{0}$ and Lemma~\ref{lemma:vdsbp}
 \[
@@
 We can now construct a skew-symmetric formulation on the reference element $\hat{D}$ and show that it is semi-discretely entropy conservative.
-We will then extend this formulation to affine elements.
-In both cases, this formulation can be made entropy stable by adding interface dissipation.
-
+This formulation can be made entropy stable by adding interface dissipation.
 Let $\bm{u}_h$ denote the discrete solution, and let $\bm{u}_q$ denote the values of the solution at volume quadrature points.
@@
 \bm{V}_q \\ \bm{V}_f\end{array}}^T 
-\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{V}_f^T\bm{W}_f \diag{\hat{\bm{n}}_i}\bm{f}_i^* = 0,  \label{eq:esdgSkew}\\
-\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q,\nonumber
+\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}^i_S}\bm{1} + \bm{V}_f^T\bm{B}^i\bm{f}_i^* = 0,  \label{eq:esdgSkew}\\
+\LRp{\bm{F}^i_S}_{ij} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q,\nonumber
 \end{gather}
@@
 \LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} = \begin{pmatrix}
-\bm{Q}_i-\bm{Q}_i^T & \LRp{\bm{V}_f \bm{P}_q}^T\bm{W}_f\diag{\bm{n}_i}\\
--\bm{W}_f\diag{\bm{n}_i}{\bm{V}_f \bm{P}_q} & \bm{0}
+\bm{Q}_i-\bm{Q}_i^T & {\bm{E}}^T \bm{B}^i\\
+-\bm{B}^i\bm{E} & \bm{0}
 \end{pmatrix}.
@@
 \pd{\bm{u}_h}{t} + \sum_{i=1}^d \LRs{\begin{array}{cc}
-\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{\hat{\bm{n}}_i}\bm{f}_i^* = 0,\label{eq:esdg}
+\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}^i_S}\bm{1} + \bm{L}_f \diag{\hat{\bm{n}}_i}\bm{f}_i^* = 0,\label{eq:esdg}
 %\\
@@
 \begin{theorem}
-Suppose Assumption~\ref{ass:quad} holds for $M_{\rm vol} = M_{\rm surf} = 0$.  
+Suppose Assumption~\ref{ass:quad} holds for $v(\bm{x}) = 1$.  
 Then, the formulation (\ref{eq:esdgSkew}) is entropy conservative such that
-\[
-\bm{1}^T\bm{W}\pd{U(\bm{u}_q)}{t} + \sum_{i=1}^d\bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}_i} \LRp{\psi_i(\tilde{\bm{u}}_f) - \tilde{\bm{v}}_f^T\bm{f}_i^*} = 0, \qquad \bm{u}_q = \bm{V}_q\bm{u}.
-\]
+\begin{equation}
+\bm{1}^T\bm{W}\pd{U(\bm{u}_q)}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}^i \LRp{\psi_i(\tilde{\bm{u}}_f) - \tilde{\bm{v}}_f^T\bm{f}_i^*} = 0, \qquad \bm{u}_q = \bm{V}_q\bm{u}.
+\label{eq:esdgthm}
+\end{equation}
 \label{thm:esdg}
@@
 \bm{v}_q^T\bm{W}\pd{\LRp{\bm{V}_q\bm{u}}}{t} + \sum_{i=1}^d
-\tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \tilde{\bm{v}}_f^T \bm{W}_f \diag{\hat{\bm{n}}_i}\bm{f}_i^* = 0.
+\tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}^i_S}\bm{1} + \tilde{\bm{v}}_f^T \bm{B}^i\bm{f}_i^* = 0.
 \end{align}
@@
 \begin{align*}
-\tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} &= \tilde{\bm{v}}^T \LRp{\bm{Q}^i_N \circ \bm{F}_S}\bm{1} - \tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N}^T \circ \bm{F}_S}\bm{1}\\
-&= \tilde{\bm{v}}^T \LRp{\bm{Q}^i_N \circ \bm{F}_S}\bm{1} - \bm{1}^T \LRp{{\bm{Q}^i_N} \circ \bm{F}_S}\tilde{\bm{v}}.
-\end{align*}
-Where we have used that $\bm{F}_S$ is symmetric and that the Hadamard product commutes.  Applying the conservation condition on $\bm{f}_S$ in (\ref{eq:esflux}) then yields
-\begin{align*}
-\tilde{\bm{v}}^T \LRp{\bm{Q}^i_N \circ \bm{F}_S}\bm{1} - \bm{1}^T \LRp{{\bm{Q}^i_N} \circ \bm{F}_S}\tilde{\bm{v}} &= \sum_{jk} \LRp{\bm{Q}^i_N}_{jk} \LRp{\tilde{\bm{v}}_j-\tilde{\bm{v}}_k} \bm{f}_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k} \\
-&= \sum_{ij} \LRp{\bm{Q}^i_N}_{jk} \LRp{\psi_i(\tilde{\bm{u}}_j) - \psi_i(\tilde{\bm{u}}_k)}\\
-&= \bm{1}^T\LRp{\bm{Q}^i_N}\psi_i(\tilde{\bm{u}}) - \psi_i(\tilde{\bm{u}})^T\LRp{\bm{Q}^i_N}\bm{1} \\
-&= \bm{1}^T\LRp{\bm{Q}^i_N}\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i_N\psi_i(\tilde{\bm{u}}),
-\end{align*}
-where we have used Corollary~\ref{lemma:sbpcor} in the last equality.
-Noting that $\bm{1}^T\bm{B}^i_N\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}_i} \psi_i(\tilde{\bm{u}}_f)$ completes the proof.
+\tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}^i_S}\bm{1} &= \tilde{\bm{v}}^T \LRp{\bm{Q}^i_N \circ \bm{F}^i_S}\bm{1} - \tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N}^T \circ \bm{F}^i_S}\bm{1}\\
+&= \tilde{\bm{v}}^T \LRp{\bm{Q}^i_N \circ \bm{F}^i_S}\bm{1} - \bm{1}^T \LRp{{\bm{Q}^i_N} \circ \bm{F}^i_S}\tilde{\bm{v}}.
+\end{align*}
+Where we have used that $\bm{F}^i_S$ is symmetric and that the Hadamard product commutes.  Applying the conservation condition on $\bm{f}^i_S$ in (\ref{eq:esflux}) then yields
+\begin{align*}
+&\tilde{\bm{v}}^T \LRp{\bm{Q}^i_N \circ \bm{F}^i_S}\bm{1} - \bm{1}^T \LRp{{\bm{Q}^i_N} \circ \bm{F}^i_S}\tilde{\bm{v}} \\
+&= \sum_{jk} \LRp{\bm{Q}^i_N}_{jk} \LRp{\tilde{\bm{v}}_j-\tilde{\bm{v}}_k} \bm{f}_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k} = \sum_{jk} \LRp{\bm{Q}^i_N}_{jk} \LRp{\psi_i(\tilde{\bm{u}}_j) - \psi_i(\tilde{\bm{u}}_k)}\\
+&= \bm{1}^T\LRp{\bm{Q}^i_N}\psi_i(\tilde{\bm{u}}) - \psi_i(\tilde{\bm{u}})^T\LRp{\bm{Q}^i_N}\bm{1} = \bm{1}^T\LRp{\bm{Q}^i_N}\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i_N\psi_i(\tilde{\bm{u}}),
+\end{align*}
+where we have used Corollary~\ref{lemma:sbpcor} in the last equality.  Noting that $\bm{1}^T\bm{B}^i_N\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i \psi_i(\tilde{\bm{u}}_f)$ completes the proof.
 %The final step of the proof is  where we have used Lemma~\ref{lemma:vdsbp} to integrate by parts and conclude that $\bm{1}^T\LRp{\bm{Q}^i_N}^T = 0$.
@@
 \begin{remark}
-It should be emphasized that the only result necessary to prove Theorem~\ref{thm:esdg} is Corollary~\ref{lemma:sbpcor}.
-The decoupled SBP property (\ref{eq:dsbp}) is not utilized, and is not necessary to guarantee a semi-discrete conservation of entropy.
+The primary difference between this formulation and the formulation presented in \cite{chan2017discretely} is the use of the skew-symmetric volume operator $\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}$.
+When the strong SBP property (\ref{eq:dsbp}) holds, this formulation is exactly equivalent to the strong formulation presented in \cite{chan2017discretely}.
+However, the only result necessary to prove Theorem~\ref{thm:esdg} is Corollary~\ref{lemma:sbpcor}.
+The decoupled SBP property (\ref{eq:dsbp}) is not utilized, and is not necessary to guarantee a semi-discrete conservation of entropy for the skew-symmetric formulation.
 \end{remark}
@@
 \begin{align}
-\bm{1}^T\bm{W}\pd{\LRp{\bm{V}_q\bm{u}}}{t} + \sum_{i=1}^d\bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}_i}\bm{f}_i^* = 0. 
+\bm{1}^T\bm{W}\pd{\LRp{\bm{V}_q\bm{u}}}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}^i\bm{f}_i^* = 0. 
 \end{align}
@@
 \end{align}
-Because $\bm{F}_S$ is symmetric and $\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}$ is skew-symmetric, the term 
-\[
-\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}
-\]
-is a skew-symmetric matrix.  Using that $\bm{x}^T\bm{A}\bm{x} = 0$ for any skew symmetric matrix $\bm{A}$, the volume term vanishes
+Because $\bm{F}_S$ is symmetric and $\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}$ is skew-symmetric, their Hadamard product is also skew-symmetric.
+Using that $\bm{x}^T\bm{A}\bm{x} = 0$ for any skew symmetric matrix $\bm{A}$, the volume term vanishes
 \[
@@
 
-Next, we extend this formulation to affinely mapped elements.
-Let the domain be decomposed into non-overlapping elements $D^k$, such that $D^k$ is the image of the reference element $\hat{D}$ under an affine mapping $\bm{\Phi}^k$.
-We define geometric terms ${G}^k_{ij}$ as scaled derivatives of reference coordinates $\hat{\bm{x}}$ w.r.t.\ physical coordinates $\bm{x}$
+\subsection{On quadrature conditions for Assumption~\ref{ass:quad} with $v = 1$}
+\label{sec:assump1}
+Apart from algebraic manipulations, only Corollary~\ref{lemma:sbpcor} is necessary to prove entropy conservation for (\ref{eq:esdg}).
+Corollary~\ref{lemma:sbpcor} requires that Assumption~\ref{ass:quad} holds for $v=1$.
+Thus, the volume and surface quadratures must be sufficiently accurate to guarantee that the mass matrix is positive definite and to integrate
+\begin{equation}
+\int_{\hat{D}} \pd{u}{x_i}, \qquad \int_{\partial \hat{D}} u n_i. \label{eq:affineints}
+\end{equation}
+On simplicial elements, the mass matrix is guaranteed to be positive definite for any volume quadrature which is exact for degree $2N$ polynomial integrands.
+This choice of volume quadrature also guarantees that the volume term in (\ref{eq:affineints}) is integrated exactly.
+The surface quadrature can thus be taken to be any quadrature rule which is exact for only degree $N$ integrands on faces.
+In contrast, the construction of simplicial decoupled SBP operators has required face quadratures which are accurate for degree $2N$ polynomials \cite{chan2017discretely, chan2018discretely}.
+
+On tensor product elements, we can take any degree $(2N-1)$ quadrature rule which ensures a positive definite mass matrix (e.g.\ a $(N+1)$-point GLL quadrature), as a quadrature of this accuracy is sufficient to exactly integrate the volume term in (\ref{eq:affineints}).
+For the surface quadrature, we can again take any quadrature rule which is exact for degree $N$ polynomial integrands.
+For example, on a quadrilateral element, one can use $\left\lceil\frac{N+1}{2}\right\rceil$-point Gauss quadrature rule or a $\left\lceil\frac{N+3}{2}\right\rceil$-point GLL rule as face quadratures for a degree $N$ scheme.
+
+%\note{link SBP accuracy section} %These issues will be discussed further in Section~\ref{sec:num}.
+%This minimal degree of surface quadrature accuracy will depend on the nature of the physical mesh, and will differ between affine and curved elements.
+
+\subsection{On the accuracy of the skew-symmetric formulation}
+
+We note that, while the proofs in this Section show that it is possible to maintain entropy stability under reduced accuracy surface quadrature rules, the accuracy of the resulting schemes depend on the accuracy of the volume and surface quadrature.
+%We briefly remark on the accuracy of (\ref{eq:esdgSkew}).
+
+It was shown in \cite{chan2017discretely} that, when combined with quadrature-based $L^2$ projection and lifting operators, the decoupled SBP operator (\ref{eq:dsbp}) is a degree $N$ approximation to the derivative.
+This approximation can be interpreted as augmenting a volume approximation of the derivative with boundary correction terms \cite{chan2018efficient}.
+Let $f(\bm{x}),g(\bm{x})$ denote two $L^2$ integrable functions, and let $\bm{f}_N, \bm{g}_N$ denote the vectors of values of $f,g$ at both volume and surface quadrature points.
+A degree $N$ approximation $u\in V^N$ to $f\pd{g}{x_i}$ can be constructed via
+\[
+\bm{M}\bm{u} = \LRs{\begin{array}{c}
+\bm{V}_q \\ \bm{V}_f\end{array}}^T \diag{\bm{f}_N}\bm{Q}^i_N\bm{g}_N, \qquad \bm{f}_N = \begin{bmatrix} \bm{f}_q\\ \bm{f}_f\end{bmatrix}, \quad \bm{g}_N = \begin{bmatrix} \bm{g}_q\\ \bm{g}_f\end{bmatrix},
+\]
+where $\bm{u}$ denotes the vector of coefficients for $u$.  Expanding terms using (\ref{eq:dsbp}) yields 
+%\note{fix!  add $E = V_fP_q$}
+\begin{align*}
+\bm{M}\bm{u} &= \bm{V}_q^T \diag{\bm{f}_q} \bm{Q}^i \bm{g}_q + \frac{1}{2}\bm{V}_q^T\diag{\bm{f}_q}\bm{E}^T\bm{B}^i \LRp{\bm{g}_f -\bm{E}\bm{g}_q}  \\
+&\qquad + \frac{1}{2}\bm{V}_f^T\diag{\bm{f}_f}\bm{B}^i \LRp{\bm{g}_f -\bm{E}\bm{g}_q}.
+\end{align*}
+where we have used that $\bm{B}^i$ is diagonal and that diagonal matrices commute.
+Let $\bm{v}$ are coefficients of some test function $v\in V^N$, and let $\bm{v}_q = \bm{V}_q\bm{v}$ and $\bm{v}_f = \bm{V}_f \bm{v}$.
+Multiplying by $\bm{v}^T$ on both sides yields that $\bm{v}^T\bm{M}\bm{u}$ is equal to
+\begin{align*}
+\LRp{\diag{\bm{f}_q}\bm{v}_q}^T \bm{Q}^i \bm{g}_q + \frac{1}{2}\LRp{\diag{\bm{f}_f}\bm{v}_f+\bm{E}\diag{\bm{f}_q}\bm{v}_q}^T\bm{B}^i \LRp{\bm{g}_f -\bm{E}\bm{g}_q} .
+\end{align*}
+This algebraic expression can be reinterpreted as a quadrature approximation of the following variational problem
+\begin{equation}
+\int_{\hat{D}} u v = \int_{\hat{D}} f\pd{\Pi_N g}{x_i}v + \int_{\partial \hat{D}} \LRp{g - \Pi_Ng} \LRp{\frac{fv + \Pi_N(fv)}{2}} \hat{n}_i, \qquad \forall v\in V^N.
+\label{eq:var}
+\end{equation}
+
+The skew-symmetric formulation can be interpreted as approximating a skew symmetric form of (\ref{eq:var}).  Integrating half of the volume term by parts yields 
+\begin{align}
+&\frac{1}{2} \int_{\hat{D}} \LRp{f\pd{\Pi_N g}{x_k}v - g \pd{\Pi_N\LRp{fv}}{x_i}} \label{eq:skewDN}\\
+&+\frac{1}{2} \int_{\partial \hat{D}} \LRp{fgv\hat{n}_i + \LRp{g - \Pi_Ng} \LRp{\frac{fv + \Pi_N(fv)}{2}} \hat{n}_i}\nonumber
+\end{align}
+Approximating integrals using quadrature allows us to express (\ref{eq:skewDN}) as a matrix equation using $\bm{B}^i_N$ in (\ref{eq:dsbp})
+%\begin{align*}
+%\bm{v}^T\bm{M}\bm{u} &= \bm{v}^T\LRp{\LRs{\begin{array}{c}
+%\bm{V}_q \\ \bm{V}_f\end{array}}}^T\diag{\bm{f}_N}\LRp{\bm{B}^i_N + \frac{1}{2}\LRp{\bm{Q}_N-\bm{Q}_N^T}}{\bm{g}_N}\\
+%&= \bm{v}^T \LRp{\frac{1}{2}\LRp{\LRs{\begin{array}{c}
+%\bm{V}_q \\ \bm{V}_f\end{array}}}^T \LRp{\diag{\bm{f}_N}{\LRp{\bm{Q}_N-\bm{Q}_N^T}}{\bm{g}_N}} +  \bm{V}_f^T \diag{\bm{f}_f}\bm{B}^i \bm{g}_f}.  
+%\end{align*}
+%Since this holds for all vectors $\bm{v}$, we recover the matrix equation
+\begin{align*}
+\bm{M}\bm{u} &= {\frac{1}{2}\LRp{\LRs{\begin{array}{c}
+\bm{V}_q \\ \bm{V}_f\end{array}}}^T \LRp{\diag{\bm{f}_N}{\LRp{\bm{Q}_N-\bm{Q}_N^T}}{\bm{g}_N}} +  \bm{V}_f^T \diag{\bm{f}_f}\bm{B}^i \bm{g}_f}.
+\end{align*}
+
+The accuracy of this approximation depends on the degree of accuracy of the volume and surface quadrature rules.
+
+The entropy conservative skew-symmetric formulation (\ref{eq:esdgSkew})
+
+\note{Fix this up: if IBP holds for $v\in P^M$, the approximation should be $O(h^M)$}.
+
+\note{Add in continuous formulation and note that skew-symmetry approximates it.}
+
+\section{Skew-symmetric entropy conservative formulations on mapped elements}
+\label{sec:skew2}
+
+We now construct skew-symmetric formulations on mapped elements.
+We assume some domain $\Omega$ is decomposed into non-overlapping elements $D^k$, such that $D^k$ is the image of the reference element $\hat{D}$ under an isoparametric mapping $\bm{\Phi}^k$.
+We define geometric change of variables terms ${G}^k_{ij}$ as scaled derivatives of reference coordinates $\hat{\bm{x}}$ w.r.t.\ physical coordinates $\bm{x}$
 \begin{gather}
@@
 \end{gather}
-
-Derivatives with respect to physical coordinates on $D^k$ are computed in terms of a change of variables formula and geometric terms (\ref{eq:geofacs}).
-For affine meshes, the scaled geometric terms ${G}^k_{ij}$ are constant over each element.
+We also define $\bm{n}^k_i$ as the vector containing concatenated values of the scaled outward normals $n^k_iJ^k_f$ at surface quadrature nodes.
+For the remainder of the work, we assume that the mesh is watertight \cite{chan2018discretely} such that at all points on any internal face, the scaled outward normals $n^k_iJ^k_f$ on the two elements sharing this face are equal and opposite.
+
+%Derivatives with respect to physical coordinates on $D^k$ are computed in terms of a change of variables formula and geometric terms (\ref{eq:geofacs}).  
+%%For affine meshes, the scaled geometric terms ${G}^k_{ij}$ are constant over each element.
 It was shown in \cite{chan2017discretely} that one can define a physical differentiation matrix as a linear combination of reference differentiation matrices.
-Let ${\bm{D}}^i_N$ denote the decoupled SBP operator on the reference element $\hat{D}$.
+%Let ${\bm{D}}^i_N$ denote the decoupled SBP operator on the reference element $\hat{D}$.
 Define the decoupled SBP operators $\bm{D}^i_k$ with respect to the physical coordinates on $D^k$ be defined as
-\begin{align}
-{\bm{D}}^i_k = \sum_{j=1}^d {G}^k_{ij}{\bm{D}}^j_N.
-\end{align}
-Let $\bm{n}^k_i$ be a vector containing concatenated values of the scaled outward normals $n^k_iJ^k_f$ at surface quadrature nodes.  
-Then, an entropy conservative formulation can be given on $D^k$ as follows:
-\begin{gather}
-J^k\bm{M}\pd{\bm{u}_h}{t} + 
-\sum_{i=1}^d \LRs{\begin{array}{cc}
-\bm{V}_q \\
-\bm{V}_f\end{array}}^T \LRp{\LRp{\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T} \circ \bm{F}_S}\bm{1} + \bm{V}_f^T\bm{W}_f \diag{{\bm{n}^k_i}}\bm{f}_i^* = 0, \label{eq:skewform}\\
+%\begin{align}
+%{\bm{D}}^i_k = \sum_{j=1}^d {G}^k_{ij}\hat{\bm{D}}^j_N.
+%\end{align}
+
+%\note{Maybe remove affinely mapped elements section.}  
+%
+%\subsection{Affinely mapped elements}
+%\label{sec:affine}
+%
+%We first extend this formulation to affinely mapped elements.  Let the domain be 
+%
+%Let Assumption~\ref{ass:quad} again hold for $v(\bm{x}) = 1$.  Then, it is straightforward to show that the following formulation on $D^k$ is locally entropy conservative: 
+%\begin{gather}
+%J^k\bm{M}\pd{\bm{u}_h}{t} + 
 %\sum_{i=1}^d \LRs{\begin{array}{cc}
-%\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{{\bm{n}_i}}\bm{f}_i^* = 0, \label{eq:skewform}\\
-\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q, \nonumber\\
-\bm{f}^* = \bm{f}_S(\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f), \qquad \text{ on interior interfaces,} \nonumber
-\end{gather}
-where $\tilde{\bm{u}}_f^+$ denotes the face value of the entropy-projected conservative variables $\tilde{\bm{u}}_f$ on the neighboring element.
-The entropy conservative formulation can be made entropy stable by adding appropriate interface dissipation, such as Lax-Friedrichs or matrix-based penalization terms \cite{winters2017uniquely, chen2017entropy, chan2017discretely}.
-
-\subsubsection{On conditions for Assumption~\ref{ass:quad}}
-
-Apart from algebraic manipulations, only Corollary~\ref{lemma:sbpcor} is necessary to prove entropy conservation for (\ref{eq:esdg})  and (\ref{eq:skewform}).
-Corollary~\ref{lemma:sbpcor} requires that Assumption~\ref{ass:quad} holds for $M_{\rm vol}=M_{\rm surf}=0$.
-Thus, the volume and surface quadratures must be sufficiently accurate to guarantee that the mass matrix is positive definite and to integrate
-\begin{equation}
-\int_{\hat{D}} \pd{u}{x_i}, \qquad \int_{\partial \hat{D}} u n_i. \label{eq:affineints}
-\end{equation}
-On simplicial elements, the mass matrix is guaranteed to be positive definite for any volume quadrature which is exact for degree $2N$ polynomial integrands.
-This choice of volume quadrature also guarantees that the volume term in (\ref{eq:affineints}) is integrated exactly.
-The surface quadrature can thus be taken to be any quadrature rule which is exact for only degree $N$ integrands on faces.
-In contrast, the construction of simplicial SBP operators has generally required face quadratures which are accurate for degree $2N$ polynomials \cite{hicken2016multidimensional, crean2018entropy, chan2018discretely}.
-
-On tensor product elements, we consider isotropic quadrature rules which are tensor products of one-dimensional quadrature rules.
-For example, we can use a quadrature rule constructed through a tensor product of one-dimensional $(N+1)$-point GLL quadrature rules, which is sufficient to guarantee that the mass matrix is positive definite \cite{canuto2007spectral}.
-This quadrature rule is also exact for integrands in $Q^{2N-1}$, and thus exactly integrates the volume term in (\ref{eq:affineints}).
-This leaves the surface quadrature, which can again be taken to be any quadrature rule exact for degree $N$ integrands.
-For example, on a quadrilateral element, one can use either $\left\lceil\frac{N+1}{2}\right\rceil$-point Gauss quadrature rule or a $\left\lceil\frac{N+3}{2}\right\rceil$-point GLL rule as face quadratures for a degree $N$ scheme.
-%This minimal degree of surface quadrature accuracy will depend on the nature of the physical mesh, and will differ between affine and curved elements.
-
-\subsection{Entropy stability on curvilinear meshes}
-\label{sec:curved}
-
-On affine meshes, it is possible to show entropy stability of the skew-symmetric formulation (\ref{eq:esdgSkew}) under a surface quadrature which is only exact for degree $N$ polynomials.
+%\bm{V}_q \\
+%\bm{V}_f\end{array}}^T \LRp{\LRp{\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T} \circ \bm{F}_S}\bm{1} + \bm{V}_f^T\bm{W}_f \diag{{\bm{n}^k_i}}\bm{f}_i^* = 0, \label{eq:skewform}\\
+%%\sum_{i=1}^d \LRs{\begin{array}{cc}
+%%\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{{\bm{n}_i}}\bm{f}_i^* = 0, \label{eq:skewform}\\
+%\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q, \nonumber\\
+%\bm{f}^* = \bm{f}_S(\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f), \qquad \text{ on interior interfaces,} \nonumber
+%\end{gather}
+%where $\tilde{\bm{u}}_f^+$ denotes the face value of the entropy-projected conservative variables $\tilde{\bm{u}}_f$ on the neighboring element.
+The proof of global entropy conservation follows as in \cite{chan2017discretely}, and results from summing up (\ref{eq:esdgthm}) over all elements and noting that the surface terms cancel due to the symmetry and conservation properties of the Tadmor flux (\ref{eq:esflux}).
+The entropy conservative formulation (\ref{eq:skewform}) can be made entropy stable by adding appropriate interface dissipation, such as Lax-Friedrichs or matrix-based penalization terms \cite{winters2017uniquely, chen2017entropy, chan2017discretely}.
+
+%On affine meshes, it is possible to show entropy stability of the skew-symmetric formulation (\ref{eq:esdgSkew}) under a surface quadrature which is only exact for degree $N$ polynomials.  
+On a single element (and on affine meshes), it is possible to show entropy stability of the skew-symmetric formulation (\ref{eq:esdgSkew}) under a surface quadrature which is only exact for degree $N$ polynomials.
 However, on curved meshes, stronger conditions are required to guarantee entropy stability.
@@
 
-\subsection{Discrete geometric conservation law and surface quadrature accuracy}
-
-Let the domain now be triangulated by a watertight mesh (as defined in \cite{chan2018discretely}) consisting of curved elements $D^k$.  
-In contrast to the affine case, geometric terms $G^k_{ij}$ for a curved element are now high order polynomials over the reference domain $\hat{D}$.
-We assume that the scaled normals are still computed from $G^k_{ij}$ using (\ref{eq:normals}).
+\subsection{Curved elements and the geometric conservation law}
+\label{sec:curved}
+
+%Let the domain now be triangulated by a  consisting of curved elements $D^k$.  
 In this section, we describe how to construct appropriate decoupled SBP operators on curved meshes, and give conditions on the volume and surface quadrature rules under which a semi-discretely entropy stable scheme can be constructed.
 
+%For a curved element $D^k$, geometric terms $G^k_{ij}$ are high order polynomials over the reference domain $\hat{D}$.
+
 We first show how to construct decoupled SBP operators on curved elements.
-Let $\bm{G}^k_{ij}$ denote the vector of scaled geometric terms ${G}^k_{ij}$ evaluated at both volume and surface quadrature points.
+Let $\bm{G}^k_{ij}$ denote the vector of scaled geometric terms ${G}^k_{ij}$ evaluated at both volume and surface quadrature points, and let $\hat{\bm{D}}^j_N$ denote the decoupled SBP operator for the $j$th reference coordinate.
 Decoupled SBP operators on a curved element $D^k$ can be defined as in \cite{chan2018discretely} by
 \begin{equation}
-\bm{D}^i_k = \frac{1}{2}\sum_{j=1}^d \LRp{\diag{\bm{G}^k_{ij}}{\bm{D}}^j_N + {\bm{D}}^j_N\diag{\bm{G}^k_{ij}}}, \qquad \bm{Q}^i_k = \bm{W} \bm{D}^i_k.
+\bm{D}^i_k = \frac{1}{2}\sum_{j=1}^d \LRp{\diag{\bm{G}^k_{ij}}\hat{\bm{D}}^j_N + \hat{\bm{D}}^j_N\diag{\bm{G}^k_{ij}}}, \qquad \bm{Q}^i_k = \bm{W} \bm{D}^i_k.
 \label{eq:dncurved}
 \end{equation}
-It can be shown that, if $\bm{D}^j_N$ satisfies a strong summation by parts property, $\bm{D}^i_k$ satisfies an analogous strong summation by parts property.
-%The construction of $\bm{D}^i_k$ in (\ref{eq:dncurved}) 
-The skew-symmetric formulation (\ref{eq:skewform}) can be extended to curvilinear meshes by replacing the reference element operators $\bm{Q}^i_N$ with the physical operators $\bm{Q}^i_k$ defined by (\ref{eq:dncurved}).
-However, additional assumptions must be satisfied in order to prove that the resulting formulation is entropy stable.
-
-The first assumption which must be satisfied is the discrete geometric conservation law (GCL) \cite{thomas1979geometric, kopriva2006metric}.
-For curved elements, Lemma~\ref{lemma:vdsbp} and Corollary~\ref{lemma:sbpcor} do not necessarily hold at the discrete level.
-For example, expanding out the condition $\bm{Q}^i_k\bm{1} = \bm{0}$ in terms of (\ref{eq:dncurved}) yields
+It can be shown that, if $\hat{\bm{D}}^j_N$ satisfies a strong summation by parts property, $\bm{D}^i_k$ satisfies an analogous strong summation by parts property.
+
+For curved elements, steps must also be taken to ensure that the analogue of Lemma~\ref{lemma:vdsbp} and Corollary~\ref{lemma:sbpcor} hold at the discrete level.
+Expanding out the condition $\bm{Q}^i_k\bm{1} = \bm{0}$ in terms of (\ref{eq:dncurved}) yields
 \begin{align}
-\bm{Q}^i_k \bm{1} = \frac{1}{2} \bm{W}\sum_{j=1}^d \diag{\bm{G}^k_{ij}}{\bm{D}}^j_N \bm{1} + {\bm{D}}^j_N\diag{\bm{G}^k_{ij}}\bm{1} = \frac{1}{2}\bm{W}\sum_{j=1}^d {\bm{D}}^j_N\LRp{\bm{G}^k_{ij}} = 0,
+\bm{Q}^i_k \bm{1} = \frac{1}{2} \bm{W}\sum_{j=1}^d \diag{\bm{G}^k_{ij}}\hat{\bm{D}}^j_N \bm{1} + \hat{\bm{D}}^j_N\diag{\bm{G}^k_{ij}}\bm{1} = \frac{1}{2}\bm{W}\sum_{j=1}^d \hat{\bm{D}}^j_N\LRp{\bm{G}^k_{ij}} = 0,
 \label{eq:dgcl}
 \end{align}
-where we have used that ${\bm{D}}^j_N \bm{1} = 0$.
+where we have used that $\hat{\bm{D}}^j_N \bm{1} = 0$.
+We refer to the condition $\bm{Q}^i_k\bm{1} = 0$ as the discrete geometric conservation law (GCL) \cite{thomas1979geometric, kopriva2006metric}.
 For degree $N$ isoparametric mappings, the GCL is automatically satisfied in two dimensions due to the fact that the exact geometric terms ${G}^k_{ij}$ are polynomials of degree $N$ \cite{kopriva2006metric}.
@@
 
-A common approach is to rewrite the geometric terms as the curl of some quantity $\bm{r}^i$, but to interpolate $\bm{r}^i$ before applying the curl \cite{visbal2002use, kopriva2006metric, hindenlang2012explicit}:
-\begin{align}
+A common approach is to rewrite the geometric terms as the curl of some quantity $\bm{r}^i$, but to interpolate $\bm{r}^i$ before applying the curl \cite{visbal2002use, kopriva2006metric, hindenlang2012explicit, chan2018discretely}:
+\begin{gather}
 \bm{r}^i = { \pd{\bm{x}}{\hat{x}_i}\times \bm{x}}, \qquad
 \LRs{\begin{array}{c}
-\bm{G}^k_{i1}\\
-\bm{G}^k_{i2}\\
-\bm{G}^k_{i3}\end{array}} = -\frac{1}{2}\LRp{\pd{I_{N_{\rm geo}}\bm{r}^k}{\hat{x}_j}-\pd{I_{N_{\rm geo}}\bm{r}^j}{\hat{x}_k}}, 
+{G}^k_{i1}\\
+{G}^k_{i2}\\
+{G}^k_{i3}\end{array}} = -\frac{1}{2}\LRp{\pd{I_{N_{\rm geo}}\bm{r}^k}{\hat{x}_j}-\pd{I_{N_{\rm geo}}\bm{r}^j}{\hat{x}_k}},\\
+N_{\rm geo} \leq \begin{cases}
+N+1 & \text{(tetrahedra)}\\
+N & \text{(hexahedra)}
+\end{cases},\nonumber
 \label{eq:iconscurl}
-\end{align}
+\end{gather}
 where $I_{N_{\rm geo}}$ denotes a degree $N_{\rm geo}$ polynomial interpolation operator with appropriate interpolation nodes.\footnote{This interpolation step must be performed using interpolation points with an appropriate number of nodes on each boundary \cite{chan2018discretely}.
-These include, for example, GLL nodes on tensor product elements, as well as Warp and Blend nodes on non-tensor product elements \cite{warburton2006explicit, chan2015comparison}.}
+These include, for example, GLL nodes on tensor product elements, as well as Warp and Blend nodes on non-tensor product elements \cite{warburton2006explicit, chan2015comparison}.}  The restriction on the maximum value of $N_{\rm geo}$ ensures that $G^k_{ij} \in V^N$ (e.g.
+$G^k_{ij} \in P^N$ on tetrahedral elements and $G^k_{ij}\in Q^N$ on hexahedral elements), which is necessary to guarantee (\ref{eq:dgcl}).
 
@@
 
-\begin{lemma}{(Triangles and tetrahedra)}
-Let $D^k$ be a curved simplicial element, and let the geometric terms $G^k_{ij}$ be constructed .
-If the volume quadrature is exact for polynomials of degree $(N+N_{\rm geo}-1)$ and the surface quadrature is exact for polynomials of degree $(N+N_{\rm geo})$, then
-\[
-\bm{1}^T\bm{Q}^i_k\bm{u} = \bm{1}^T\bm{B}^i_k\bm{u}, \qquad \bm{Q}^i_k\bm{1} = \bm{0}.
-\]
-where $\bm{u}$ is a vector of values of some function at volume and surface quadrature points. 
+\begin{lemma}{(Discrete FTC and exact constant differentiation on curved elements)}
+\label{lemma:vdsbpcurved} 
+Let $D^k$ be a curved element, and let the geometric terms $G^k_{ij}$ be constructed using (\ref{eq:iconscurl}).
+Let Assumption~\ref{ass:quad} hold for $v = 1$ and $v = G^k_{ij}$ for all $i,j = 1,\ldots, d$, and let $\bm{u}$ denote a vector of values of some function at volume and surface quadrature points.
+Then,
+%If the volume quadrature is exact for polynomials of degree $(N+N_{\rm geo}-1)$ and the surface quadrature is exact for polynomials of degree $(N+N_{\rm geo})$, then 
+\[
+\bm{1}^T\bm{Q}^i_k\bm{u} = \bm{1}^T\bm{B}^i_k\bm{u}, \qquad \bm{Q}^i_k\bm{1} = \bm{0},
+\]
+where the physical boundary matrix $\bm{B}^i_k$ is defined as
+\[
+\bm{B}^i_k = \sum_{j=1}^d \bm{B}^j_N \diag{\bm{G}^k_{ij}} = \sum_{j=1}^d \diag{\bm{G}^k_{ij}}\bm{B}^j_N  = \begin{pmatrix}
+\bm{0}&\\
+& \bm{W}_f \diag{\bm{n}^k_i}\end{pmatrix}.
+\]
 \end{lemma}
 \begin{proof}
-d
+The proof of $\bm{Q}^i_k\bm{1} = \bm{0}$ was shown for for tensor product elements in \cite{kopriva2006metric} and for simplicial elements in \cite{chan2018discretely}, and relies only on the fact that $G^k_{ij} \in V^N$.
+We focus on showing that the discrete fundamental theorem of calculus $\bm{1}^T\bm{Q}^i_k\bm{u} = \bm{1}^T\bm{B}^i_k\bm{u}$ holds.
+Using (\ref{eq:dncurved}), we have that
+\begin{align*}
+\bm{1}^T\bm{Q}^i_k\bm{u} &= \frac{1}{2}\sum_{j=1}^d \bm{1}^T{\LRp{\diag{\bm{G}^k_{ij}}\hat{\bm{Q}}^j_N + \hat{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}}} \bm{u} \\
+&= \frac{1}{2}\sum_{j=1}^d \LRp{\LRp{\bm{G}^k_{ij}}^T \hat{\bm{Q}}^j_N\bm{u} + \bm{1}^T\hat{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}\bm{u}}.
+\end{align*}
+Applying Lemma~\ref{lemma:vdsbp} to $\sum_{j=1}^d\bm{1}^T\hat{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}\bm{u}$ yields
+\begin{align*}
+\sum_{j=1}^d\bm{1}^T\hat{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}\bm{u} &= \sum_{j=1}^d\bm{1}^T\LRp{\bm{B}^j_N - \LRp{\hat{\bm{Q}}^j_N}^T}\diag{\bm{G}^k_{ij}}\bm{u} \\
+&= \sum_{j=1}^d\bm{1}^T\bm{B}^j_N \diag{\bm{G}^k_{ij}}\bm{u},
+\end{align*}
+where we have used Lemma~\ref{lemma:sbpcor} to conclude that $\hat{\bm{Q}}^j_N \bm{1} = 0$.
+Repeating the procedure for $\sum_{j=1}^d \LRp{\bm{G}^k_{ij}}^T \hat{\bm{Q}}^j_N\bm{u}$ yields
+\begin{align*}
+\sum_{j=1}^d \LRp{\bm{G}^k_{ij}}^T \hat{\bm{Q}}^j_N\bm{u} &= \sum_{j=1}^d \LRp{\bm{G}^k_{ij}}^T \LRp{\bm{B}^j_N - \LRp{\hat{\bm{Q}}^j_N}^T}\bm{u} \\
+&= \sum_{j=1}^d \LRp{\bm{G}^k_{ij}}^T \bm{B}^j_N \bm{u} - \sum_{j=1}^d \bm{u}^T{\hat{\bm{Q}}^j_N}{\bm{G}^k_{ij}} = \sum_{j=1}^d \LRp{\bm{G}^k_{ij}}^T \bm{B}^j_N \bm{u},  
+\end{align*}
+where we have used that the $G^k_{ij}$ satisfies the discrete GCL.  The proof is completed by noting that $\bm{B}^j_N$ is diagonal and thus commutes with $\diag{\bm{G}^k_{ij}}$, such that
+\[
+\bm{1}^T\bm{Q}^i_k\bm{u} = \sum_{j=1}^d\bm{1}^T\LRp{\bm{B}^j_N \diag{\bm{G}^k_{ij}}}\bm{u} = \bm{1}^T\bm{B}^i_k \bm{u}.
+\]
 \end{proof}
 
-\begin{lemma}{(Quadrilaterals and hexahedra)}
-Let $D^k$ be a curved tensor product element.  Then, 
-\end{lemma}
-\begin{proof}
-\note{Proof relies on two facts: that geometric terms are products of }
-\end{proof}
-
-%, which are summarized in the following curved version of Corollary~\ref{lemma:sbpcor}.
-%\note{Maybe provide 2 lemmas - one for triangles and one for quadrilaterals?  Polynomial degree of geofacs is slightly different between the two.  } 
-%\begin{lemma}
-%Assume that the geometric terms $\bm{G}^k_{ij} \in V^{N_{\rm geo}}$ satisfy the discrete GCL (\ref{eq:dgcl}), that the volume quadrature is exact for polynomials of degree $N+N_{\rm geo}-1$, and that the surface quadrature is exact for polynomials of degree $N+N_{\rm geo}$.
-Then,
-%\[
-%\bm{1}^T\bm{Q}^i_k\bm{u} = \bm{1}^T\bm{B}^i_k \bm{u}, 
-%\qquad \bm{Q}^i_k \bm{1} = 0, 
-%\]
-%where 
-%\[
-%\bm{B}^i_k =  \begin{pmatrix}
-%\bm{0}&\\
-%& \bm{W}_f \diag{{\bm{n}^k_i}}
-%\end{pmatrix}.
-%\]
-%\label{lemma:vdsbpcurved}
-%\end{lemma}
-%\begin{proof}
-%The second equality $\bm{Q}^i_k \bm{1} = 0$ is a consequence of the discrete GCL (\ref{eq:dgcl}), and the proof is identical to that of \cite{chan2018discretely}.  
-%%For the first equality, if the surface quadrature is exact for polynomials of degree $2N$ on the trace space, then the matrix SBP property $\bm{Q}^i_N = \bm{B}^i_N - \LRp{\bm{Q}^i_N}^T$ holds \cite{chan2018discretely}.
-Combining this with $\bm{Q}^i_k\bm{1} = 0$ yields the desired result.
-We focus on the proof of the variational SBP property for surface quadrature which are exact for polynomials of degree $N+R < 2N$.
-%For the first equality, we utilize the variational SBP property.  
-%Expanding $\bm{1}^T\bm{Q}^i_k\bm{u}$ yields
-%\[
-%\bm{1}^T\bm{Q}^i_k\bm{u} = \frac{1}{2}\sum_{j=1}^d \bm{1}^T \diag{\bm{G}^k_{ij}} {\bm{Q}}^j_N \bm{u} + \bm{1}^T{\bm{Q}}^j_N \diag{\bm{G}^k_{ij}} \bm{u}.
-%=  \frac{1}{2}\sum_{j=1}^d \bm{1}^T \diag{\bm{G}_{ij}} \hat{\bm{Q}}^j_N \bm{u},
-%\]
-%For the latter term in the sum, Lemma~\ref{lemma:vdsbp} holds and
-%\begin{align}
-%\sum_{j=1}^d\bm{1}^T{\bm{Q}}^j_N \diag{\bm{G}^k_{ij}} \bm{u} &= 
-%\sum_{j=1}^d\bm{1}^T\LRp{{\bm{B}}^j_N - \LRp{{\bm{Q}}^j_N}^T} \diag{\bm{G}^k_{ij}} \bm{u} \\
-%&= \sum_{j=1}^d\bm{1}^T{\bm{B}}^j_N\diag{\bm{G}^k_{ij}}\bm{u}, \nonumber
-%\label{eq:vsbp1}
-%\end{align}
-%where we have used that ${\bm{Q}}^j_N \bm{1} = \bm{0}$.  
-%
-%For the former term in the sum, we use the fact that $G^k_{ij}$, when constructed using (\ref{eq:iconscurl}), is a polynomial of degree $N$.
-Thus, $\bm{G}^k_{ij} = \begin{bmatrix}\bm{V}_q\\ \bm{V}_f\end{bmatrix}\bm{g}$ for some vector of polynomial coefficients $\bm{g}$.
-We can then rewrite the former term as
-%\begin{gather}
-%\sum_{j=1}^d\bm{1}^T \diag{\bm{G}^k_{ij}} {\bm{Q}}^j_N \bm{u} = 
-%\sum_{j=1}^d{\bm{G}^k_{ij}}^T {\bm{Q}}^j_N \bm{u}.% = \sum_{j=1}^d \bm{g}^T \begin{bmatrix}\bm{V}_q\\ \bm{V}_f\end{bmatrix}^T{\bm{Q}}^j_N \bm{u}.
-%%= \sum_{j=1}^d\begin{bmatrix}
-%%\bm{G}^q_{ij}\\
-%%\bm{G}^f_{ij}
-%%\end{bmatrix}^T \begin{bmatrix} 
-%%\hat{\bm{Q}}_i - \frac{1}{2}\LRp{\bm{V}_f \bm{P}_q}^T  \bm{W}_f{\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q &  \frac{1}{2}\LRp{\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q}^T\label{eq:ibpg}\\
-%%-\frac{1}{2}\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q & \frac{1}{2}\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)
-%%\end{bmatrix} \begin{bmatrix}\bm{u}_q\\
-%%\bm{u}_f\end{bmatrix}.\nonumber
-%\end{gather}
-%%where $\bm{G}^q_{ij},\bm{G}^f_{ij}$ denote the values of $\bm{G}_{ij}$ at volume and surface quadrature points.  
-%%Because $\bm{P}_q\bm{V}_q = \bm{I}$, straightforward computations give that %simplify the final expression 
-%%\[
-%%\begin{bmatrix}\bm{V}_q\\ \bm{V}_f\end{bmatrix}^T{\bm{Q}}^j_N = \begin{bmatrix}
-%%\bm{Q}^i-\bm{V}_f^T\bm{W}_f\diag{\hat{\bm{n}}_j}\bm{V}_f\bm{P}_q & \bm{V}_f^T\bm{W}_f\diag{\hat{\bm{n}}_j}
-%%\end{bmatrix}
-%%\]
-%
-%Because $\bm{G}_{ij} \in V^N$, it is equal to its own $L^2$ projection, and the values at volume and surface quadrature points are related by $\bm{G}^f_{ij} = \bm{V}_f\bm{P}_q\bm{G}^q_{ij}$.
-We can use this to simplify (\ref{eq:ibpg})
-%\begin{align*}
-%\bm{1}^T \diag{\bm{G}_{ij}} \hat{\bm{Q}}^j_N \bm{u} =&  \LRp{\bm{G}^q_{ij}}^T { \bm{Q}_i}  \bm{u}_q - \frac{1}{2} \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q+\bm{G}^f_{ij}}^T{\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q}\bm{u}_q\\
-%&+ \frac{1}{2} \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q+\bm{G}^f_{ij}}^T\frac{1}{2}\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{u}_f\\
-%=& \LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}}_i - \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q}^T{\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q}}\bm{u}_q \\
-%&+ \LRp{\bm{G}^f_{ij}}^T\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{u}_f. %\\
-%%&= \LRp{\bm{G}^q_{ij}}^T \LRp{ \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q}^T{\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q}}\bm{u}_q + \LRp{\bm{G}^f_{ij}}^T\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{u}_f 
-%\end{align*}
-%The first term can be further simplified using Lemma~\ref{lemma:vsbp}
-%\[
-%\LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}}_i - \LRp{{\bm{V}_f \bm{P}_q}\bm{G}_{ij}^q}^T{\bm{W}_f{\rm diag}(\hat{\bm{n}}_i)\bm{V}_f\bm{P}_q}}\bm{u}_q = \LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}}_i}^T\bm{u}_q.
-%\]
-%The latter term can be combined with the surface term of (\ref{eq:vsbp1}) by noting that 
-%\[
-%\LRp{\bm{G}^f_{ij}}^T\bm{W}_f \diag{\hat{\bm{n}}_i}\bm{u}_f =\bm{1}^T\hat{\bm{B}}^j_N\diag{\bm{G}_{ij}}\bm{u} = \bm{1}^T\diag{\bm{G}_{ij}}\hat{\bm{B}}^j_N\bm{u},
-%\]
-%where we have used that $\hat{\bm{B}}^j_N$ is diagonal and commutes with diagonal scaling by $\bm{G}_{ij}$.
-%%as $\LRp{\bm{G}^f_{ij}}^T\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{u}_f = \bm{1}^T\hat{\bm{B}}^j_N \bm{u}$.  
-%We can use polynomial exactness and the fact that the geometric terms satisfy the continuous GCL by construction \cite{chan2018discretely} to show that $\sum_{j=1}^d  \hat{\bm{Q}}_j {\bm{G}^q_{ij}} = 0$.
-%Combining this with (\ref{eq:vsbp1}) and expanding out $\hat{\bm{B}}^j_N$ yields that
-%\begin{align*}
-%\bm{1}^T\bm{Q}^i_N\bm{u} &= \sum_{j=1}^d \LRp{
-%\frac{1}{2}\LRp{\bm{G}^q_{ij}}^T \LRp{ \hat{\bm{Q}_i}}^T \bm{u}_q + \bm{1}^T\begin{pmatrix}
-%\bm{0} &\\
-%& \bm{W}_f \diag{\bm{G}_{ij} \circ \hat{\bm{n}}_j}
-%\end{pmatrix}
-%\bm{u}} \\
-%&=  \bm{1}^T\begin{pmatrix}
-%\bm{0} &\\
-%& \bm{W}_f \diag{\sum_{j=1}^d \bm{G}_{ij} \circ \hat{\bm{n}}_j}
-%\end{pmatrix}
-%\bm{u} = \bm{1}^T\bm{B}^i_N\bm{u}.
-%\end{align*}
-%where we have used in the final step that $\sum_{j=1}^d\bm{G}_{ij}\hat{n}_j = \bm{n}_i J^k_f$ \cite{ciarlet1978finite, chan2018discretely}
-%\end{proof}
-
-Lemma~\ref{lemma:vdsbpcurved} implies that the polynomial degree $R$ of the surface geometric terms must be compatible with the accuracy of the surface quadrature.
-This, in turn, depends on the polynomial degree $N_{\rm geo}$ to which geometric terms are approximated.
-
-\note{Kopriva trick to enforcing GCL still results in $Q^N$ normals 3D.
-May not satisfy requirement that $\bm{1}^T\bm{Q}^i_N \psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i_N \psi_i(\tilde{\bm{u}})$ because of quadrature inaccuracy.
-Note - this still works regardless for conforming hexes because of the full matrix SBP property.}
-
-\begin{theorem}
-\label{thm:skewformcurved}
-Let the geometric terms $G^k_{ij}$ be computed from (\ref{eq:iconscurl}), let the scaled outward normals $\bm{n}^k_i$ be computed pointwise from (\ref{eq:normals}), and let $\bm{Q}^i_k$ be given by (\ref{eq:dncurved}).
-Let $\bm{M}^k = \bm{V}_q^T\bm{W}\diag{\bm{J}^k}\bm{V}_q$ be the curved mass matrix on $D^k$, and let the auxiliary quantities $\tilde{\bm{u}}$ be defined by
+The property $\bm{Q}^i_k\bm{1} = 0$ is also related to free-stream preservation \cite{kopriva2006metric}.
+Constant solutions are stationary solutions of systems of conservation laws.
+However, on curved meshes, the presence of spatially varying geometric terms can result in the production of spurious transient waves.
+The construction of geometric terms through (\ref{eq:iconscurl}) guarantees that the resulting methods are free-stream preserving, and that constant solutions remain stationary solutions of discretizations of (\ref{eq:nonlineqs}).
+
+We can now construct and prove entropy conservation and free stream preservation for a skew-symmetric formulation on a curved element $D^k$.
+Let $\bm{Q}^i_k$ be given by (\ref{eq:dncurved}), and define the curved mass matrix $\bm{M}^k = \bm{V}_q^T\bm{W}\diag{\bm{J}^k}\bm{V}_q$.
+Note that $\bm{M}^k$ is positive-definite so long as $J^k$ is positive at all quadrature points.
+We define the auxiliary quantities $\tilde{\bm{u}}$
 \begin{gather*}
@@
 \end{gather*}
-where $\bm{P}^k_q = \LRp{\bm{M}^k}^{-1}\bm{V}_q^T\bm{W}\diag{\bm{J}^k}$.  Then, the formulation
+where $\bm{P}^k_q = \LRp{\bm{M}^k}^{-1}\bm{V}_q^T\bm{W}\diag{\bm{J}^k}$.  Then, we have the following theorem:
+\begin{theorem}
+\label{thm:skewformcurved}
+Let the geometric terms $G^k_{ij}$ be computed from (\ref{eq:iconscurl}), let the scaled outward normals $\bm{n}^k_i$ be computed pointwise from (\ref{eq:normals}), and suppose Assumption~\ref{ass:quad} holds for $v = 1$ and $v= G^k_{ij}$.
+Let $\tilde{\bm{u}}_f^+$ denote the face value of the entropy-projected conservative variables $\tilde{\bm{u}}_f$ on the neighboring element.
+Then, the formulation
 \begin{gather}
@@
 \bm{V}_q \\
-\bm{V}_f\end{array}}^T \LRp{\LRp{\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T} \circ \bm{F}_S}\bm{1} + \bm{V}_f^T\bm{W}_f \diag{{\bm{n}^k_i}}\bm{f}_i^* = 0, \label{eq:skewformcurved}\\
+\bm{V}_f\end{array}}^T \LRp{\LRp{\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T} \circ \bm{F}^i_S}\bm{1} + \bm{V}_f^T \bm{W}_f \diag{{\bm{n}^k_i}}\bm{f}_i^* = 0, \label{eq:skewformcurved}\\
 %\sum_{i=1}^d \LRs{\begin{array}{cc}
 %\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{{\bm{n}_i}}\bm{f}_i^* = 0, \label{eq:skewform}\\
-\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q, \nonumber\\
-\bm{f}^* = \bm{f}_S(\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f), \qquad \text{ on interior interfaces,} \nonumber
+\LRp{\bm{F}^i_S}_{ij} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q, \nonumber\\
+\bm{f}_i^* = \bm{f}^i_S(\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f), \qquad \text{ on interior interfaces,} \nonumber
 \end{gather}
-is semi-discretely entropy conservative on curved meshes.  
+is semi-discretely entropy conservative on $D^k$ such that for $\bm{u}_q = \bm{V}_q\bm{u}$,
+\begin{gather*}
+\bm{1}^T\bm{W}\diag{\bm{J}^k}\pd{U(\bm{u}_q)}{t} + \sum_{i=1}^d\bm{1}^T\bm{W}_f \diag{\bm{n}^k_i} \LRp{\psi_i(\tilde{\bm{u}}_f) - \tilde{\bm{v}}_f^T\bm{f}_i^*} = 0.
+\end{gather*}
+Additionally, the method is free-stream preserving such that $\pd{\bm{u}_h}{t} = 0$ for constant solutions.
 \end{theorem}
+\begin{proof}
+The proof of entropy stability is identical to that of Theorem~\ref{thm:esdg}, except that Lemma~\ref{lemma:vdsbpcurved} is used instead of Lemma~\ref{lemma:sbpcor}.
+However, the proof of free-stream preservation differs from proofs found in the literature, which typically show that the right hand side is zero for $\bm{u}_h$ constant \cite{crean2018entropy, chan2018discretely}.
+This strategy for is appropriate for ``strong'' formulations, but must be modified for the skew-symmetric formulation (\ref{eq:skewformcurved}).
+
+We have that $\bm{f}_i^* = \bm{f}^i_S(\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f) = \bm{f}_i(\bm{u}_f)$ using the consistency property of the Tadmor flux (\ref{eq:esflux}) and the fact that the solution is globally constant.
+For $\bm{u}_h$ constant, $\bm{F}_S$ is also constant.
+We assume (without loss of generality) that $\bm{F}_S$ and $\bm{f}_i(\bm{u}_f)$ are the matrix and vector of ones, respectively.
+Then,
+\[
+\LRp{\LRp{\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T} \circ \bm{F}_S}\bm{1} = \LRp{\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T}\bm{1} = \LRp{\bm{Q}^i_k}^T\bm{1},
+\]
+where we have used that $\bm{Q}^i_k\bm{1} = 0$ by Lemma~\ref{lemma:vdsbpcurved}.  Thus, the right hand side reduces to 
+\begin{align*}
+&\sum_{j=1}^d \LRs{\begin{array}{cc}
+\bm{V}_q \\
+\bm{V}_f\end{array}}^T \LRp{-\LRp{\bm{Q}^i_k}^T + \bm{V}_f^T \bm{W}_f \diag{{\bm{n}^k_i}}}\bm{1}\\
+&= \sum_{j=1}^d \LRs{\begin{array}{cc}
+\bm{V}_q \\
+\bm{V}_f\end{array}}^T \LRp{\LRp{ -\frac{1}{2}\LRp{\diag{\bm{G}^k_{ij}}\hat{\bm{Q}}^j_N + \hat{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}}^T + \hat{\bm{B}}^j_N \diag{\bm{G}^k_{ij}}}}\bm{1},
+%&= \sum_{j=1}^d \LRs{\begin{array}{cc}
+%\bm{V}_q \\
+%\bm{V}_f\end{array}}^T  -\frac{1}{2}\LRp{\hat{\bm{Q}}^j_N}^T{\bm{G}^k_{ij}} + 
+\end{align*}
+where we have used the definition of $\bm{Q}^i_k$ in (\ref{eq:dncurved}).  Rearranging terms gives that, for a constant solution, (\ref{eq:skewformcurved}) reduces to
+\begin{align*}
+\bm{M}^k\pd{\bm{u}_h}{t} &+ \frac{1}{2} \sum_{j=1}^d \LRs{\begin{array}{cc}
+\bm{V}_q \\
+\bm{V}_f\end{array}}^T \LRp{-\LRp{\hat{\bm{Q}}^j_N}^T +\hat{\bm{B}}^j_N}\bm{G}^k_{ij}\\
+&+ \frac{1}{2} \sum_{j=1}^d \LRs{\begin{array}{cc}
+\bm{V}_q \\
+\bm{V}_f\end{array}}^T\diag{\bm{G}^k_{ij}} \LRp{-\LRp{\hat{\bm{Q}}^j_N}^T +\hat{\bm{B}}^j_N}\bm{1} = 0.
+\end{align*}
+Let $u$ denote the constant value of the solution such that the values of the solution at quadrature points is $u\bm{1}$.  Multiplying both sides by the constant solution $\bm{u}_h^T$ yields
+\begin{align*}
+&\frac{1}{2}\pd{}{t}\LRp{\bm{u}_h^T\bm{M}^k\bm{u}_h} + \\
+&\frac{u}{2} \sum_{j=1}^d \LRp{\bm{1}^T \LRp{-\LRp{\hat{\bm{Q}}^j_N}^T +\hat{\bm{B}}^j_N}\bm{G}^k_{ij}
++\LRp{\bm{G}^k_{ij}}^T \LRp{-\LRp{\hat{\bm{Q}}^j_N}^T +\hat{\bm{B}}^j_N}\bm{1}} = 0.
+\end{align*}
+By assumption, Lemma~\ref{lemma:vdsbp} holds for test functions $\bm{1}$ and $\bm{G}^k_{ij}$, and the formulation reduces to
+\begin{align*}
+\frac{1}{2}\pd{}{t}\LRp{\bm{u}_h^T\bm{M}^k\bm{u}_h} + \frac{u}{2} \sum_{j=1}^d \LRp{\bm{1}^T \hat{\bm{Q}}^j_N\bm{G}^k_{ij}
++\LRp{\bm{G}^k_{ij}}^T \hat{\bm{Q}}^j_N\bm{1}} = 0.
+\end{align*}
+We have that $\hat{\bm{Q}}^j_N\bm{1} = 0$ by construction, and $\sum_{j=1}^d \bm{1}^T \hat{\bm{Q}}^j_N\bm{G}^k_{ij} = 0$ since $G^k_{ij}$ satisfies the GCL.
+This implies that $\pd{}{t}\LRp{\bm{u}_h^T\bm{M}^k\bm{u}_h} = 0$ on $D^k$.
+Since $\bm{M}^k$ is symmetric and positive-definite, $\bm{u}_h^T\bm{M}^k\bm{u}_h$ is a norm on $\bm{u}_h$.
+Summing $\pd{}{t}\LRp{\bm{u}_h^T\bm{M}^k\bm{u}_h} = 0$ over all elements ensures that global constants are stationary solutions.
+\end{proof}
+
+The proof of global entropy conservation follows from summing up (\ref{eq:esdgthm}) over all elements and noting that the surface terms cancel due to the symmetry and conservation properties of the Tadmor flux (\ref{eq:esflux}) \cite{chan2017discretely}.
+The entropy conservative formulations presented in this work can be made entropy stable by adding appropriate interface dissipation, such as Lax-Friedrichs or matrix-based penalization terms \cite{winters2017uniquely, chen2017entropy, chan2017discretely}.
 
 \begin{remark}
-It is possible to replace the curved mass matrix $\bm{M}^k$ with a more easily invertible weight-adjusted approximation while maintaining high order accuracy, entropy stability, and local conservation \cite{chan2018discretely}.
+It is also possible to replace the curved mass matrix $\bm{M}^k$ with a more easily invertible weight-adjusted approximation while maintaining high order accuracy, entropy stability, and local conservation \cite{chan2018discretely}.
+This approximation avoids the inversion of dense weighted $L^2$ mass matrices $\bm{M}^k$ on curved simplicial elements.
+On tensor product elements, this approximation is generally unnecessary as common choices of volume quadrature result in a diagonal (lumped) mass matrix \cite{carpenter2014entropy, parsani2016entropy, chan2018efficient}.
 \end{remark}
 
+\subsection{On quadrature conditions for Assumption~\ref{ass:quad} for $v = 1$ and $v = G^k_{ij}$}
+
+Semi-discrete entropy conservation on curved meshes requires that Assumption~\ref{ass:quad} holds for $v = 1$ and $v = G^k_{ij}$.
+We discuss specific choices of volume and surface quadrature on for which this assumption is valid.
+\note{Emphasize we're interested in taking $N_{\rm geo}$ as large as possible for a given quadrature rule, and that there is a focus on degree $(2N-1)$ quadrature rules.}
+
+\subsubsection{Simplicial elements}
+
+On simplicial elements, for $u \in P^N$, $\pd{u}{\hat{x}_j}\in P^{N-1}$, and for $N_{\rm geo}\leq (N+1)$, $G^k_{ij}\in P^{N_{\rm geo}-1}$.
+Thus, the integrands in Assumption~\ref{ass:quad} $\pd{u}{\hat{x}_j} v \in P^{N+N_{\rm geo}-1}$ and $uv n_i \in P^{N+N_{\rm geo}}$ for $v = G^k_{ij}$.
+
+Thus, a volume quadrature rule of degree $(2N-1)$ and surface quadrature of degree $2N$ are sufficient to ensure semi-discrete entropy conservation for $N_{\rm geo}\leq N$.  
+\note{Add discussion on triangular elements.}
+
+\subsubsection{Tensor product elements}
+
+%Tensor product $(N+1)$-point GLL quadratures are a common choice for quadrilateral and hexahedral elements, and are exact for polynomials of degree $(2N-1)$ in each coordinate direction.  
+It was shown in Section~\ref{sec:assump1} that tensor product quadratures of degree $(2N-1)$ satisfy Assumption~\ref{ass:quad} holds for $v = 1$.
+However, in contrast to the simplicial case, it is not immediately clear that degree $(2N-1)$ volume quadratures exactly integrate $\int_{\hat{D}} \pd{u}{\hat{x}_j}v$ for $v = G^k_{ij}$ for tensor product elements.
+
+The key difference between simplicial and tensor product elements is the polynomial space in which the derivative lies.
+In contrast to the simplicial case, if $u \in Q^N$, $\pd{u}{\hat{x}_j} \not\in Q^{N-1}$.
+Consider the three-dimensional case with $i = 1$, $u, v \in Q^N$.
+Then, $\pd{u}{\hat{x}_1} \in Q^{N-1,N,N}$, and differentiation reduces the polynomial degree in one coordinate but not others.
+As a result, $\pd{u}{\hat{x}_j}v \not\in Q^{2N-1}$, and a tensor product quadrature of degree $(2N-1)$ does not exactly integrate $\int_{\hat{D}} \pd{u}{\hat{x}_j}v$ for general $v\in Q^N$.
+
+We first consider the two dimensional case, and show that Assumption~\ref{ass:quad} holds with $v = G^k_{ij}$ for all $N_{\rm geo} \leq N$.  On quadrilateral elements, $G^k_{ij}$ is
+\begin{align*}
+G^k_{11} &= \pd{x_2}{\hat{x}_2} \in Q^{N_{\rm geo}, N_{\rm geo}-1}, \qquad G^k_{12} = -\pd{x_2}{\hat{x}_1} \in Q^{N_{\rm geo}-1, N_{\rm geo}},\\
+G^k_{21} &= -\pd{x_1}{\hat{x}_2} \in Q^{N_{\rm geo}, N_{\rm geo}-1}, \qquad G^k_{22} = \pd{x_1}{\hat{x}_1}\in Q^{N_{\rm geo}-1, N_{\rm geo}}.
+\end{align*}
+Since $\pd{u}{\hat{x}_1} \in Q^{N-1,N}$ and $\pd{u}{\hat{x}_2} \in Q^{N,N-1}$
+\[
+\pd{u}{\hat{x}_i}G^k_{ij} \in Q^{N+N_{\rm geo}-1} \subset Q^{2N-1}, \qquad \forall N_{\rm geo}\leq N.
+\]
+Thus, any quadrature rule of degree $(2N-1)$ exactly integrates $\int_{\hat{D}}\pd{u}{\hat{x}_i}v$ for $v = G^k_{ij}$.
+
+We now consider the condition in Assumption~\ref{ass:quad} on the surface integrals $\int_{\partial \hat{D}} u v n_i$ for $v= G^k_{ij}$.  \note{Finish}
+
+We now consider the three-dimensional case.
+Unlike quadrilateral elements, Assumption~\ref{ass:quad} does not hold for all $N_{\rm geo} \leq N$ on general curved hexahedra.
+Expanding out (\ref{eq:iconscurl}), one can show that $G^k_{ij} \in Q^{N_{\rm geo}}$ on hexahedral elements.
+Thus, if $N_{\rm geo} = N$ and $u\in Q^N$, $\pd{u}{\hat{x}_1} G^k_{i1} \in Q^{N-1,N,N}$.
+\note{Finish}.
+
+\subsection{Sufficient conditions on quadrature and $N_{\rm geo}$ to guarantee entropy conservation}
+
+For convenience, this section summarizes conditions on the accuracy of volume and surface quadratures under which one can guarantee entropy conservation for a given degree $N_{\rm geo}$ of a geometric mapping.
+
+ \note{finish}
+
+%% $G^k_{1j} = \pd{\hat{x}_j}{x_1}$, $j = 1,2,3$ yields
+%%\begin{align*}
+%%G^k_{11} &= \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_2}x_3}}{\hat{x}_3}- \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_3}x_3}}{\hat{x}_2} \in Q^{N_{\rm geo}}\\
+%%G^k_{12} &= \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_3}x_3}}{\hat{x}_1}- \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_3}x_3}}{\hat{x}_1} \in Q^{N_{\rm geo}}\\
+%%G^k_{13} &= \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_1}x_3}}{\hat{x}_2}- \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_2}x_3}}{\hat{x}_1} \in Q^{N_{\rm geo}}.
+%%%Dt*((Ds*y).*z) - Ds*((Dt*y).*z)
+%%\end{align*}
+%
+%%Lemma~\ref{lemma:vdsbpcurved} implies that the polynomial degree $R$ of the surface geometric terms must be compatible with the accuracy of the surface quadrature.
+This, in turn, depends on the polynomial degree $N_{\rm geo}$ to which geometric terms are approximated.
+\note{Kopriva trick to enforcing GCL still results in $Q^N$ normals 3D.
+May not satisfy requirement that $\bm{1}^T\bm{Q}^i_N \psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i_N \psi_i(\tilde{\bm{u}})$ because of quadrature inaccuracy.
+Note - this still works regardless for conforming hexes because of the full matrix SBP property.}
+%
+%
+%\begin{lemma}
+%\label{lemma:curlPN}
+%Let $\bm{u} \in \mathbb{R}^3$ be represented on a hexahedron by polynomials of the form
+%\begin{gather*}
+%\bm{u} = \begin{pmatrix}u\\v\\w\end{pmatrix}, \qquad u \in Q^{N-1,N,N}, \qquad v \in Q^{N,N-1,N}, \qquad w \in Q^{N,N,N-1}.
+%\end{gather*}
+%Then 
+%\[
+%\Grad \times \bm{u} \in \begin{pmatrix} Q^{N,N-1,N-1}\\Q^{N-1,N,N-1}\\Q^{N-1,N-1,N}\end{pmatrix}
+%\]
+%and traces of $\LRl{\bm{u}}\cdot\bm{n} \in Q^{N-1}$ (\note{CHECK THIS}).
+%\end{lemma}
+%\begin{proof}
+%The proof is a straightforward computation
+%\end{proof}
+%
+%We summarize the
+
+\section{A mortar-based implementation}
+
+For hybrid meshes, the choice of surface quadrature for an interface may vary depending on the neighboring elements.
+Consider, for example, a mixed triangular-quadrilateral mesh where the quadrilaterals utilize GLL quadrature on each face.
+While te skew-symmetric formulation allows for the stable coupling of quadrilaterals and triangles using a GLL face quadrature, one may wish to utilize more accurate Gauss face quadratures for triangle-triangle interfaces.
+To accomodate arbitrary combinations of face quadrature, we introduce a mortar-based implementation.
+
+We assume that each element
+
+\note{Finish - figure out notation for face interp!}
+
 \section{Numerical experiments}
+\label{sec:num}
 
@@
 
-\subsection{Entropy stability under reduced surface quadrature}
+\subsection{Triangular meshes}
 
@@
 
-\subsection{GLL quadrilaterals with Gauss surface quadrature}
-
-\note{Explain that for GLL quadratures, the decoupled SBP property doesn't hold when Gauss points are used.  }
+\subsection{Quadrilateral meshes}
+
+\note{ GLL quadrilaterals with Gauss surface quadrature.  Explain that for GLL quadratures, the decoupled SBP property doesn't hold when Gauss points are used.  }
 
 \subsection{Hybrid quadrilateral-triangular meshes}
+
+It is not immediately clear that the outward normals as defined by (\ref{eq:normals}) will be equal and opposite across interfaces shared by elements of different types.
+Thus, we restrict ourselves to straight-sided meshes in this case, and will investigate the effect of curvilinear mappings on hybrid meshes in future work.
 
@@
 
+\section{Conclusions}
+
+\section{Acknowledgments}
+
+%\appendix
+%
+%\section{A modified Geometric terms}
+
 \bibliographystyle{unsrt}

### 2018-09-08 matlab/elas/ElasticityAcoustic2D_unified.m

@@ -1,24 +1,27 @@
-function ElasticAcoustic2D_unified
+% function ElasticAcoustic2D_unified
 
 % clear all, clear
 clear -global *
 
 Globals2D
+global tau
 
-K1D = 16
-N = 5;
+K1D = 8;
+N = 3;
 c_flag = 0;
-FinalTime = .5;
+FinalTime = 1.0;
+tau = 1;
 
 [Nv, VX, VY, K, EToV] = unif_tri_mesh(K1D);
 
 StartUp2D;
+% BuildPeriodicMaps2D(2,2);
 
 [rp sp] = EquiNodes2D(15); [rp sp] = xytors(rp,sp);
 Vp = Vandermonde2D(N,rp,sp)/V;
 xp = Vp*x; yp = Vp*y;
 
-Nq = 2*N+1;
+Nq = 2*N;
 [rq sq wq] = Cubature2D(Nq); % integrate u*v*c
 Vq = Vandermonde2D(N,rq,sq)/V;
 Pq = V*V'*Vq'*diag(wq); % J's cancel out
@@ -78,26 +81,25 @@ if ~isempty(Ke) && ~isempty(Ka)
 end
 
 %%
-global Nfld mu lambda Vq Pq tau c2
+global Nfld mu lambda Vq Pq c2
 Nfld = 5; %(u1,u2,sxx,syy,sxy)
 
 mu = ones(size(xq));
 lambda = ones(size(xq));
 c2 = ones(size(xq));
 
-k = 3;
-mu = 1 + .5*cos(k*pi*xq).*cos(k*pi*yq);
-c2 = 1 + .5*cos(k*pi*xq).*cos(k*pi*yq);
+% k = 1;
+% mu = 1 + .5*cos(k*pi*xq).*cos(k*pi*yq);
+% c2 = 1 + .5*cos(k*pi*xq).*cos(k*pi*yq);
 
-mu = V\(Pq*mu); 
-c2 = V\(Pq*c2);
-mu = repmat(mu(1,:),length(wq),1)/sqrt(2);
-c2 = repmat(c2(1,:),length(wq),1)/sqrt(2);
+% mu = V\(Pq*mu); 
+% c2 = V\(Pq*c2);
+% mu = repmat(mu(1,:),length(wq),1)/sqrt(2);
+% c2 = repmat(c2(1,:),length(wq),1)/sqrt(2);
 
[…]

### 2018-09-12 docs/skewform/main.tex
updating skew form paper

@@
 
-\documentclass[review]{siamart0216}
-
-\usepackage{amsmath,amssymb,amsfonts}
-\newtheorem*{remark}{Remark}
-\theoremstyle{assumption}
-\newtheorem{assumption}{Assumption}
-
-%\documentclass[preprint,10pt]{elsarticle}
-%\usepackage{amsmath,amssymb,amsfonts,amsthm}
-%\theoremstyle{definition}
-%\newtheorem{definition}{Definition}
-%\theoremstyle{lemma}
-%\newtheorem{lemma}{Lemma}
+%\documentclass[review]{siamart0216}
+\documentclass{svjour3}                     % onecolumn (standard format)
+%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
+%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
+%\documentclass[twocolumn]{svjour3}          % twocolumn
+
+\smartqed  % flush right qed marks, e.g. at end of proof
+
+\makeatletter
+\renewcommand*{\@thmcounterend}{}
+\makeatother
+%\spnewtheorem{lemma}{Lemma}{\bfseries}{\itshape}
+\spnewtheorem{assumption}{Assumption}{\bfseries}{\itshape}
+%\spnewtheorem{remark}{Remark}{\bfseries}{\itshape}
+\usepackage{cite}
+
+\usepackage{amsmath, amssymb, amsfonts}
 %\newtheorem*{remark}{Remark}
-%\theoremstyle{theorem}
-%\newtheorem{theorem}{Theorem}
 %\theoremstyle{assumption}
 %\newtheorem{assumption}{Assumption}
-%\usepackage{fullpage}
 
 \usepackage[titletoc,toc,title]{appendix}
-
-\usepackage{array} 
-\usepackage{listings}
-\usepackage{mathtools}
-\usepackage{pdfpages}
-\usepackage[textsize=footnotesize,color=green]{todonotes}
-\usepackage{bm}
-\usepackage{bbm}
-
 \usepackage{tikz}
-\usepackage[normalem]{ulem}
 \usepackage{hhline}
-
-\usepackage{graphicx}
+\usepackage{graphicx, color, bm}
 \usepackage{subfig}
-\usepackage{color}
-
-%% ====================================== graphics
-
 \usepackage{pgfplots}
 \usepackage{pgfplotstable}
-\definecolor{markercolor}{RGB}{124.9, 255, 160.65}
-\pgfplotsset{
-compat=1.3,
-width=10cm,
-tick label style={font=\small},
-label style={font=\small},
-legend style={font=\small}
-}
-
-\usetikzlibrary{calc}
-\usetikzlibrary{intersections}
-
-%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
-\newcommand{\logLogSlopeTriangle}[5]
-{
-    % #1. Relative offset in x direction.
-    % #2. Width in x direction, so xA-xB.
-    % #3. Relative offset in y direction.
-    % #4. Slope d(y)/d(log10(x)).
-    % #5. Plot options.
-
-    \pgfplotsextra
-    {
-        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
-        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
-        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
-        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}
-
-        % Calculate auxilliary quantities, in relative sense.
-        \pgfmathsetmacro{\xArel}{#1}
-        \pgfmathsetmacro{\yArel}{#3}
-        \pgfmathsetmacro{\xBrel}{#1-#2}
-        \pgfmathsetmacro{\yBrel}{\yArel}
-        \pgfmathsetmacro{\xCrel}{\xArel}
-
-        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
-        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
-        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.
-
-        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
-        \coordinate (A) at (rel axis cs:\xArel,\yArel);
-        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
-        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);
-
-        % Draw slope triangle.
-        \draw[#5]   (A)-- node[pos=0.5,anchor=north] {}
-                    (B)-- 
-                    (C)-- node[pos=0.5,anchor=west] {\textcolor{black}{#4}}
-                    cycle;
-    }
-}
-%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
-
-\newcommand{\logLogSlopeTriangleNeg}[5]
-{
-    % #1. Relative offset in x direction.
-    % #2. Width in x direction, so xA-xB.
-    % #3. Relative offset in y direction.
-    % #4. Slope d(y)/d(log10(x)).
-    % #5. Plot options.
-
-    \pgfplotsextra
-    {
-        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
-        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
-        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
-        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}
-
-        % Calculate auxilliary quantities, in relative sense.
-        \pgfmathsetmacro{\xArel}{#1}
-        \pgfmathsetmacro{\yArel}{#3}
-        \pgfmathsetmacro{\xBrel}{#1-#2}
-        \pgfmathsetmacro{\yBrel}{\yArel}
-        \pgfmathsetmacro{\xCrel}{\xArel}
-
-        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
-        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
-        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.
-
-        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
-        \coordinate (A) at (rel axis cs:\xArel,\yArel);
-        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
-        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);
-
-        % Draw slope triangle.
-        \draw[#5]   (A)-- node[pos=.5,anchor=south] {}
-                    (B)-- 
-                    (C)-- node[pos=0.5,anchor=west] {\textcolor{black}{#4}}
-                    cycle;
-    }
-}
-%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
-
-%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
-\newcommand{\logLogSlopeTriangleFlipNeg}[5]
-{
-    % #1. Relative offset in x direction.
-    % #2. Width in x direction, so xA-xB.
-    % #3. Relative offset in y direction.
-    % #4. Slope d(y)/d(log10(x)).
-    % #5. Plot options.
-
-    \pgfplotsextra
-    {
-        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
-        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
-        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
-        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}
-
-        % Calculate auxilliary quantities, in relative sense.
-        %\pgfmathsetmacro{\xArel}{#1}
-        %\pgfmathsetmacro{\yArel}{#3}
-        \pgfmathsetmacro{\xBrel}{#1-#2}
-        \pgfmathsetmacro{\yBrel}{#3}
-        \pgfmathsetmacro{\xCrel}{#1}
-
-        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
-        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
-        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.
-
-	\pgfmathsetmacro{\xArel}{\xBrel}
-        \pgfmathsetmacro{\yArel}{\yCrel}
-
-        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
-        \coordinate (A) at (rel axis cs:\xArel,\yArel);
-        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
-        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);
-
-        % Draw slope triangle.
-        \draw[#5]   (A)-- node[pos=0.5,anchor=east] {\textcolor{black}{#4}}
-                    (B)-- 
-                    (C)-- node[pos=0.5,anchor=north] {1}
-                    cycle;
-    }
-}
-%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
-
-%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
-\newcommand{\logLogSlopeTriangleFlip}[5]
-{
-    % #1. Relative offset in x direction.
-    % #2. Width in x direction, so xA-xB.
-    % #3. Relative offset in y direction.
-    % #4. Slope d(y)/d(log10(x)).
-    % #5. Plot options.
-
-    \pgfplotsextra
-    {
-        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
-        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
-        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
-        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}
-
-        % Calculate auxilliary quantities, in relative sense.
-        %\pgfmathsetmacro{\xArel}{#1}
-        %\pgfmathsetmacro{\yArel}{#3}
-        \pgfmathsetmacro{\xBrel}{#1-#2}
-        \pgfmathsetmacro{\yBrel}{#3}
-        \pgfmathsetmacro{\xCrel}{#1}
-
-        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
-        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
-        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.
-
-	\pgfmathsetmacro{\xArel}{\xBrel}
-        \pgfmathsetmacro{\yArel}{\yCrel}
-
-        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
-        \coordinate (A) at (rel axis cs:\xArel,\yArel);
-        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
-        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);
-
-        % Draw slope triangle.
-        \draw[#5]   (A)-- node[pos=0.5,anchor=east] {\textcolor{black}{#4}}
-                    (B)-- 
-                    (C)-- node[pos=0.5,anchor=south] {}
-                    cycle;
-    }
-}
-%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
-
-\usepackage{stmaryrd}
-
-\renewcommand{\hat}[1]{\hat{#1}}
-\renewcommand{\topfraction}{0.85}
-\renewcommand{\textfraction}{0.1}
-\renewcommand{\floatpagefraction}{0.75}
-
-\newcommand{\vect}[1]{\ensuremath\boldsymbol{#1}}
-\newcommand{\tensor}[1]{\underline{\bm{#1}}}
-\newcommand{\del}{\triangle}
-\newcommand{\curl}{\grad \times}
-\renewcommand{\div}{\grad \cdot}
-
-\newcommand{\bbm}[1]{\mathbbm{#1}}
-\newcommand{\bs}[1]{\boldsymbol{#1}}
-\newcommand{\equaldef}{\stackrel{\mathrm{def}}{=}}
-
+\usepackage{slopetri}
+
+\renewcommand{\hat}{\widehat}
+
+%% d in integrand
+\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}}
+\newcommand{\diag}[1]{{\rm diag}\LRp{#1}}
 \newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}{\rm #2}}}
@@
 \newcommand{\LRl}[1]{\left. #1 \right|}
-\newcommand{\pdd}[2]{\frac{\partial^2#1}{\partial#2^2}}
 \newcommand{\pdn}[3]{\frac{\partial^{#3}#1}{\partial#2^{#3}}}
-\newcommand{\mb}[1]{\mathbf{#1}}
-\newcommand{\mbb}[1]{\mathbb{#1}}
-\newcommand{\mc}[1]{\mathcal{#1}}
-\newcommand{\snor}[1]{\left| #1 \right|}
-
-%\newcommand{\cond}[1]{\kappa\LRp{#1}}
-\newcommand{\cond}[2]{\nor{#1}_{#2}\nor{{#1}^{-1}}_{#2}}
-
 \newcommand{\Grad} {\ensuremath{\nabla}}
-\newcommand{\Div} {\ensuremath{\nabla\cdot}}
 \newcommand{\jump}[1] {\ensuremath{\llbracket#1\rrbracket}}
@@
 
-\newcommand{\Oh}{{\Omega_h}}
-\renewcommand{\L}{L^2\LRp{\Omega}}
-\newcommand{\LK}{L^2\LRp{D^k}}
-\newcommand{\LdK}{L^2\LRp{\partial D^k}}
-\newcommand{\Dhat}{\hat{D}}
-\newcommand{\Lhat}{L^2\LRp{\Dhat}}
-
-\renewcommand{\hat}{\widehat}
-
-\newcommand{\eval}[2][\right]{\relax
-  \ifx#1\right\relax \left.\fi#2#1\rvert}
-
-\def\etal{{\it et al.~}}
-
-\newcommand{\note}[1]{{\color{blue}{#1}}}
-%\newcommand{\noteOne}[1]{{\color{blue}{#1}}}
-%\newcommand{\noteTwo}[1]{{\color{red}{#1}}}
-%\newcommand{\note}[1]{#1}
-%\newcommand{\noteOne}[1]{#1}
-%\newcommand{\noteTwo}[1]{#1}
-
-\newcommand{\LinfDk}{L^{\infty}\LRp{D^k}}
-
-\newcommand{\diag}[1]{{\rm diag}\LRp{#1}}
-
-\newcommand{\Ksub}{K_{\rm sub}}
-
-\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-
-%% d in integrand
-\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}}
-
-\makeatletter
-\renewcommand\d[1]{\mspace{6mu}\mathrm{d}#1\@ifnextchar\d{\mspace{-3mu}}{}}
-\makeatother
+\renewcommand{\note}[1]{{\color{blue}{#1}}}
+
+\graphicspath{{./figs/}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \date{}
-\author{Jesse Chan}%, David C.\ Del Rey Fernandez}
+\author{Jesse Chan}
 \title{Skew-symmetric entropy stable discontinuous Galerkin formulations}
-
-\graphicspath{{./figs/}}
 
@@
 These skew-symmetric formulations avoid the use of a ``strong'' SBP property, and require only that operators exactly differentiate constants and satisfy a discrete form of the fundamental theorem of calculus.
-We conclude by discussing mortar-based implementations of skew-symmetric formulations for entropy stable schemes on hybrid meshes.
+We conclude with an application of the skew-symmetric formulation for entropy stable schemes on mixed quadrilateral-triangle meshes.
 \end{abstract}
@@
 High order methods for the simulation of time-dependent compressible flow have the potential to achieve higher levels of accuracy at lower costs compared to current low order schemes \cite{wang2013high}.
-In addition to superior accuracy, % and efficiency on modern computing architectures,
-the low numerical dispersion and dissipation of high order methods \cite{ainsworth2004dispersive} enables the accurate propagation of waves over long distances and time scales.
+In addition to superior accuracy, the low numerical dispersion and dissipation of high order methods \cite{ainsworth2004dispersive} enables the accurate propagation of waves over long distances and time scales.
 The same properties also make high order methods attractive for unsteady phenomena such as vorticular and turbulent flows, which are sensitive to numerical dissipation \cite{visbal1999high, wang2013high}.
 
-However, 
-%while high order methods are often provably stable for wave problems, high order schemes for nonlinear conservation laws have consistently been hampered by problems of instability.  
-when applied to nonlinear conservation laws, high order methods can experience artificial growth and blow-up near under-resolved features such as shocks, turbulence, or boundary layers.
+However, when applied to nonlinear conservation laws, high order methods can experience artificial growth and blow-up near under-resolved features such as shocks, turbulence, or boundary layers.
 In practice, the application of high order methods to practical problems requires shock capturing and stabilization techniques (such as artificial viscosity) or solution regularization (such as filtering or limiting) to prevent solution blow-up.
@@
 It should be noted that a similar approach to entropy stable discretizations was introduced within a finite difference framework \cite{chen2017entropy, crean2018entropy} using multidimensional differencing operators which are constructed to satisfy accuracy conditions and an SBP property \cite{hicken2016multidimensional}.
-These operators do not correspond to any specific basis or approximation space, but can be shown to exist for nodal points corresponding to sufficiently accurate choices of volume and surface quadrature.
-The formulations presented in this work require similar accuracy conditions, but differ from those of \cite{chen2017entropy, crean2018entropy} in that they build upon an explicit approximation space.
-In contrast with traditional SBP coupling terms \cite{crean2017high}, the use of decoupled SBP operators also accomodates general choices of volume and surface quadrature (e.g.\ volume quadratures without boundary nodes and over-integrated rules) while ensuring compact coupling conditions between neighboring elements.
+These operators do not correspond to any specific basis or approximation space, but can be shown to exist for nodal points corresponding to sufficiently accurate choices of volume and surface quadrature and can be computed either algebraically or through an optimization problem.
+The formulations presented in this work require similar accuracy conditions, but differ from those of \cite{chen2017entropy, crean2018entropy} in that we start with an explicit approximation space, which then induce quadrature-based operators.
+The formulations presented here also accomodate general choices of volume and surface quadrature (e.g.\ volume quadratures without boundary nodes and over-integrated rules) while ensuring compact coupling conditions between neighboring elements.
 
@@
 For sufficiently accurate quadrature rules, the matrix $\bm{Q}^i_N$ also satisfies a decoupled summation-by-parts (SBP) property, which is used to prove semi-discrete entropy stability for nonlinear conservation laws.
-\begin{lemma}
+\begin{theorem}%[Theorem 1 in \cite{chan2017discretely}]
 Assume that the volume and surface quadrature rules are sufficiently accurate such that the quantities
@@
 \label{lemma:dsbp}
-\end{lemma}
+\end{theorem}
 \begin{proof}
@@
 
-The assumptions of Lemma~\ref{lemma:dsbp} are satisfied for sufficiently accurate volume and surface quadratures.
+The assumptions of Theorem~\ref{lemma:dsbp} are satisfied for sufficiently accurate volume and surface quadratures.
 For example, on simplicial elements, (\ref{eq:dsbp}) holds if the volume quadrature is exact for polynomial integrands of total degree $(2N-1)$, and the surface integral is exact for degree $2N$ polynomials on each face.
@@
 \begin{remark}
-It is worth noting that the assumptions of Lemma~\ref{lemma:dsbp} are sufficient but not necessary conditions.
+It should be stressed that the assumptions of Theorem~\ref{lemma:dsbp} are sufficient but not necessary conditions.
 For example, suppose $\hat{D}$ is a quadrilateral element, and that the volume quadrature is a tensor product of $(N+1)$ point one-dimensional Gauss-Legendre-Lobatto (GLL) rule.
-If the surface quadrature also taken to be an $(N+1)$-point GLL rule over each face, then $\bm{Q}^i$ satisfies a traditional SBP property, which can be used to prove the decoupled SBP property for $\bm{Q}^i_N$.
+If the surface quadrature also taken to be an $(N+1)$-point GLL rule over each face, then $\bm{Q}^i$ satisfies a traditional SBP property, which implies the decoupled SBP property for $\bm{Q}^i_N$.
 \end{remark}
@@
 \subfloat[Degree $2N$ volume quadrature, GLL surface quadrature]{\includegraphics[width=.4\textwidth]{figs/trigll.png}\label{subfig:trigll}}
-\caption{Volume and surface quadrature pairs which do not satisfy the assumptions of Lemma~\ref{lemma:dsbp}, and thus do not possess the decoupled SBP property (\ref{eq:dsbp}).
+\caption{Volume and surface quadrature pairs which do not satisfy the assumptions of Theorem~\ref{lemma:dsbp}, and thus do not possess the decoupled SBP property (\ref{eq:dsbp}).
 Volume quadrature nodes are drawn as circles, while surface quadrature nodes are drawn as squares.}
@@
 Let $u,v \in Q^N$ denote two arbitrary degree $N$ polynomials.
-The assumptions of Lemma~\ref{lemma:dsbp} are that the volume quadrature exactly integrates $\int_{\hat{D}} \pd{u}{x} v$ and that the surface quadrature exactly integrates $\int_{\partial \hat{D}} u v n_i$ on $\hat{D}$.
-Because the $(N+1)$-point Gauss rule is exact for polynomials of degree $2N+1$ and the product $uv \in P^{2N}$ on each face, the surface quadrature satisfies the assumptions of Lemma~\ref{lemma:dsbp}.
+The assumptions of Theorem~\ref{lemma:dsbp} are that the volume quadrature exactly integrates $\int_{\hat{D}} \pd{u}{x} v$ and that the surface quadrature exactly integrates $\int_{\partial \hat{D}} u v n_i$ on $\hat{D}$.
+Because the $(N+1)$-point Gauss rule is exact for polynomials of degree $2N+1$ and the product $uv \in P^{2N}$ on each face, the surface quadrature satisfies the assumptions of Theorem~\ref{lemma:dsbp}.
 However, the 1D GLL rule is only exact for polynomials of degree $(2N-1)$.
@@
 Let $u,v \in P^N$ denote two arbitrary degree $N$ polynomials.
-The derivative $\pd{u}{x} \in P^{(N-1)}$, and $\pd{u}{x}v \in P^{(2N-1)}$, so the volume quadrature satisfies the assumptions of Lemma~\ref{lemma:dsbp}.
-However, because the surface quadrature is exact only degree $(2N-1)$ polynomials and the trace of $uv\in P^{2N}$, the surface quadrature does not satisfy the assumptions of Lemma~\ref{lemma:dsbp}.
+The derivative $\pd{u}{x} \in P^{(N-1)}$, and $\pd{u}{x}v \in P^{(2N-1)}$, so the volume quadrature satisfies the assumptions of Theorem~\ref{lemma:dsbp}.
+However, because the surface quadrature is exact only degree $(2N-1)$ polynomials and the trace of $uv\in P^{2N}$, the surface quadrature does not satisfy the assumptions of Theorem~\ref{lemma:dsbp}.
 
@@
 \begin{align*}
-&\int_{\partial \hat{D}} (\Pi_N u) v \hat{n}_i -   \int_{\hat{D}} \LRp{\Pi_N u} \pd{v}{\hat{x}_i} =\\
- &\bm{v}_f^T\bm{B}^i \bm{E}\bm{u}_q - \LRp{\bm{V}_q\bm{D}^i\bm{P}_q\bm{v}_q}^T\bm{W} \bm{V}_q\bm{P}_q\bm{u}_q\\
+\int_{\partial \hat{D}} (\Pi_N u) v \hat{n}_i -   \int_{\hat{D}} \LRp{\Pi_N u} \pd{v}{\hat{x}_i} &=\bm{v}_f^T\bm{B}^i \bm{E}\bm{u}_q - \LRp{\bm{V}_q\bm{D}^i\bm{P}_q\bm{v}_q}^T\bm{W} \bm{V}_q\bm{P}_q\bm{u}_q\\
 =& \bm{v}_q^T {\bm{E}}^T\bm{B}^i \bm{E}\bm{u}_q - \bm{v}_q^T \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T \bm{W} \bm{V}_q\bm{P}_q\bm{u}_q,
@@
 
-\subsection{On the accuracy of the skew-symmetric formulation}
-
-We note that, while the proofs in this Section show that it is possible to maintain entropy stability under reduced accuracy surface quadrature rules, the accuracy of the resulting schemes depend on the accuracy of the volume and surface quadrature.
+%\subsection{On the accuracy of the skew-symmetric formulation}
+%
+%We note that, while the proofs in this Section show that it is possible to maintain entropy stability under reduced accuracy surface quadrature rules, the accuracy of the resulting schemes depend on the accuracy of the volume and surface quadrature.
 %We briefly remark on the accuracy of (\ref{eq:esdgSkew}).
-
-It was shown in \cite{chan2017discretely} that, when combined with quadrature-based $L^2$ projection and lifting operators, the decoupled SBP operator (\ref{eq:dsbp}) is a degree $N$ approximation to the derivative.
+%
+%It was shown in \cite{chan2017discretely} that, when combined with quadrature-based $L^2$ projection and lifting operators, the decoupled SBP operator (\ref{eq:dsbp}) is a degree $N$ approximation to the derivative.
 This approximation can be interpreted as augmenting a volume approximation of the derivative with boundary correction terms \cite{chan2018efficient}.
@@
 A degree $N$ approximation $u\in V^N$ to $f\pd{g}{x_i}$ can be constructed via
-\[
-\bm{M}\bm{u} = \LRs{\begin{array}{c}
-\bm{V}_q \\ \bm{V}_f\end{array}}^T \diag{\bm{f}_N}\bm{Q}^i_N\bm{g}_N, \qquad \bm{f}_N = \begin{bmatrix} \bm{f}_q\\ \bm{f}_f\end{bmatrix}, \quad \bm{g}_N = \begin{bmatrix} \bm{g}_q\\ \bm{g}_f\end{bmatrix},
-\]
-where $\bm{u}$ denotes the vector of coefficients for $u$.  Expanding terms using (\ref{eq:dsbp}) yields 
-%\note{fix!  add $E = V_fP_q$}
-\begin{align*}
-\bm{M}\bm{u} &= \bm{V}_q^T \diag{\bm{f}_q} \bm{Q}^i \bm{g}_q + \frac{1}{2}\bm{V}_q^T\diag{\bm{f}_q}\bm{E}^T\bm{B}^i \LRp{\bm{g}_f -\bm{E}\bm{g}_q}  \\
-&\qquad + \frac{1}{2}\bm{V}_f^T\diag{\bm{f}_f}\bm{B}^i \LRp{\bm{g}_f -\bm{E}\bm{g}_q}.
-\end{align*}
-where we have used that $\bm{B}^i$ is diagonal and that diagonal matrices commute.
+%\[
+%\bm{M}\bm{u} = \LRs{\begin{array}{c}
+%\bm{V}_q \\ \bm{V}_f\end{array}}^T \diag{\bm{f}_N}\bm{Q}^i_N\bm{g}_N, \qquad \bm{f}_N = \begin{bmatrix} \bm{f}_q\\ \bm{f}_f\end{bmatrix}, \quad \bm{g}_N = \begin{bmatrix} \bm{g}_q\\ \bm{g}_f\end{bmatrix},
+%\]
+%where $\bm{u}$ denotes the vector of coefficients for $u$.  Expanding terms using (\ref{eq:dsbp}) yields 
+%%\note{fix!  add $E = V_fP_q$}
+%\begin{align*}
+%\bm{M}\bm{u} &= \bm{V}_q^T \diag{\bm{f}_q} \bm{Q}^i \bm{g}_q + \frac{1}{2}\bm{V}_q^T\diag{\bm{f}_q}\bm{E}^T\bm{B}^i \LRp{\bm{g}_f -\bm{E}\bm{g}_q}  \\
+%&\qquad + \frac{1}{2}\bm{V}_f^T\diag{\bm{f}_f}\bm{B}^i \LRp{\bm{g}_f -\bm{E}\bm{g}_q}.
+%\end{align*}
+%where we have used that $\bm{B}^i$ is diagonal and that diagonal matrices commute.
 Let $\bm{v}$ are coefficients of some test function $v\in V^N$, and let $\bm{v}_q = \bm{V}_q\bm{v}$ and $\bm{v}_f = \bm{V}_f \bm{v}$.
 Multiplying by $\bm{v}^T$ on both sides yields that $\bm{v}^T\bm{M}\bm{u}$ is equal to
-\begin{align*}
-\LRp{\diag{\bm{f}_q}\bm{v}_q}^T \bm{Q}^i \bm{g}_q + \frac{1}{2}\LRp{\diag{\bm{f}_f}\bm{v}_f+\bm{E}\diag{\bm{f}_q}\bm{v}_q}^T\bm{B}^i \LRp{\bm{g}_f -\bm{E}\bm{g}_q} .
-\end{align*}
-This algebraic expression can be reinterpreted as a quadrature approximation of the following variational problem
-\begin{equation}
-\int_{\hat{D}} u v = \int_{\hat{D}} f\pd{\Pi_N g}{x_i}v + \int_{\partial \hat{D}} \LRp{g - \Pi_Ng} \LRp{\frac{fv + \Pi_N(fv)}{2}} \hat{n}_i, \qquad \forall v\in V^N.
-\label{eq:var}
-\end{equation}
-
-The skew-symmetric formulation can be interpreted as approximating a skew symmetric form of (\ref{eq:var}).  Integrating half of the volume term by parts yields 
-\begin{align}
-&\frac{1}{2} \int_{\hat{D}} \LRp{f\pd{\Pi_N g}{x_k}v - g \pd{\Pi_N\LRp{fv}}{x_i}} \label{eq:skewDN}\\
-&+\frac{1}{2} \int_{\partial \hat{D}} \LRp{fgv\hat{n}_i + \LRp{g - \Pi_Ng} \LRp{\frac{fv + \Pi_N(fv)}{2}} \hat{n}_i}\nonumber
-\end{align}
-Approximating integrals using quadrature allows us to express (\ref{eq:skewDN}) as a matrix equation using $\bm{B}^i_N$ in (\ref{eq:dsbp})
 %\begin{align*}
-%\bm{v}^T\bm{M}\bm{u} &= \bm{v}^T\LRp{\LRs{\begin{array}{c}
-%\bm{V}_q \\ \bm{V}_f\end{array}}}^T\diag{\bm{f}_N}\LRp{\bm{B}^i_N + \frac{1}{2}\LRp{\bm{Q}_N-\bm{Q}_N^T}}{\bm{g}_N}\\
-%&= \bm{v}^T \LRp{\frac{1}{2}\LRp{\LRs{\begin{array}{c}
+%\LRp{\diag{\bm{f}_q}\bm{v}_q}^T \bm{Q}^i \bm{g}_q + \frac{1}{2}\LRp{\diag{\bm{f}_f}\bm{v}_f+\bm{E}\diag{\bm{f}_q}\bm{v}_q}^T\bm{B}^i \LRp{\bm{g}_f -\bm{E}\bm{g}_q} .
+%\end{align*}
+%This algebraic expression can be reinterpreted as a quadrature approximation of the following variational problem
+%\begin{equation}
+%\int_{\hat{D}} u v = \int_{\hat{D}} f\pd{\Pi_N g}{x_i}v + \int_{\partial \hat{D}} \LRp{g - \Pi_Ng} \LRp{\frac{fv + \Pi_N(fv)}{2}} \hat{n}_i, \qquad \forall v\in V^N.
+%\label{eq:var}
+%\end{equation}
+%
+%The skew-symmetric formulation can be interpreted as approximating a skew symmetric form of (\ref{eq:var}).  Integrating half of the volume term by parts yields 
+%\begin{align}
+%&\frac{1}{2} \int_{\hat{D}} \LRp{f\pd{\Pi_N g}{x_k}v - g \pd{\Pi_N\LRp{fv}}{x_i}} \label{eq:skewDN}\\
+%&+\frac{1}{2} \int_{\partial \hat{D}} \LRp{fgv\hat{n}_i + \LRp{g - \Pi_Ng} \LRp{\frac{fv + \Pi_N(fv)}{2}} \hat{n}_i}\nonumber
+%\end{align}
+%Approximating integrals using quadrature allows us to express (\ref{eq:skewDN}) as a matrix equation using $\bm{B}^i_N$ in (\ref{eq:dsbp})
+%%\begin{align*}
+%%\bm{v}^T\bm{M}\bm{u} &= \bm{v}^T\LRp{\LRs{\begin{array}{c}
+%%\bm{V}_q \\ \bm{V}_f\end{array}}}^T\diag{\bm{f}_N}\LRp{\bm{B}^i_N + \frac{1}{2}\LRp{\bm{Q}_N-\bm{Q}_N^T}}{\bm{g}_N}\\
+%%&= \bm{v}^T \LRp{\frac{1}{2}\LRp{\LRs{\begin{array}{c}
+%%\bm{V}_q \\ \bm{V}_f\end{array}}}^T \LRp{\diag{\bm{f}_N}{\LRp{\bm{Q}_N-\bm{Q}_N^T}}{\bm{g}_N}} +  \bm{V}_f^T \diag{\bm{f}_f}\bm{B}^i \bm{g}_f}.  
+%%\end{align*}
+%%Since this holds for all vectors $\bm{v}$, we recover the matrix equation
+%\begin{align*}
+%\bm{M}\bm{u} &= {\frac{1}{2}\LRp{\LRs{\begin{array}{c}
 %\bm{V}_q \\ \bm{V}_f\end{array}}}^T \LRp{\diag{\bm{f}_N}{\LRp{\bm{Q}_N-\bm{Q}_N^T}}{\bm{g}_N}} +  \bm{V}_f^T \diag{\bm{f}_f}\bm{B}^i \bm{g}_f}.
 %\end{align*}
-%Since this holds for all vectors $\bm{v}$, we recover the matrix equation
-\begin{align*}
-\bm{M}\bm{u} &= {\frac{1}{2}\LRp{\LRs{\begin{array}{c}
-\bm{V}_q \\ \bm{V}_f\end{array}}}^T \LRp{\diag{\bm{f}_N}{\LRp{\bm{Q}_N-\bm{Q}_N^T}}{\bm{g}_N}} +  \bm{V}_f^T \diag{\bm{f}_f}\bm{B}^i \bm{g}_f}.
-\end{align*}
-
-The accuracy of this approximation depends on the degree of accuracy of the volume and surface quadrature rules.
-
-The entropy conservative skew-symmetric formulation (\ref{eq:esdgSkew})
-
-\note{Fix this up: if IBP holds for $v\in P^M$, the approximation should be $O(h^M)$}.
-
-\note{Add in continuous formulation and note that skew-symmetry approximates it.}
+%
+%The accuracy of this approximation depends on the degree of accuracy of the volume and surface quadrature rules.  
+%
+%The entropy conservative skew-symmetric formulation (\ref{eq:esdgSkew})
+%
+%\note{Fix this up: if IBP holds for $v\in P^M$, the approximation should be $O(h^M)$}. 
+%
+%
+%\note{Add in continuous formulation and note that skew-symmetry approximates it.}
 
@@
 The proof of $\bm{Q}^i_k\bm{1} = \bm{0}$ was shown for for tensor product elements in \cite{kopriva2006metric} and for simplicial elements in \cite{chan2018discretely}, and relies only on the fact that $G^k_{ij} \in V^N$.
-We focus on showing that the discrete fundamental theorem of calculus $\bm{1}^T\bm{Q}^i_k\bm{u} = \bm{1}^T\bm{B}^i_k\bm{u}$ holds.
-Using (\ref{eq:dncurved}), we have that
+
+The discrete fundamental theorem of calculus $\bm{1}^T\bm{Q}^i_k\bm{u} = \bm{1}^T\bm{B}^i_k\bm{u}$ can be shown to hold using (\ref{eq:dncurved}):
 \begin{align*}
@@
 \end{align*}
-Let $u$ denote the constant value of the solution such that the values of the solution at quadrature points is $u\bm{1}$.  Multiplying both sides by the constant solution $\bm{u}_h^T$ yields
-\begin{align*}
-&\frac{1}{2}\pd{}{t}\LRp{\bm{u}_h^T\bm{M}^k\bm{u}_h} + \\
-&\frac{u}{2} \sum_{j=1}^d \LRp{\bm{1}^T \LRp{-\LRp{\hat{\bm{Q}}^j_N}^T +\hat{\bm{B}}^j_N}\bm{G}^k_{ij}
-+\LRp{\bm{G}^k_{ij}}^T \LRp{-\LRp{\hat{\bm{Q}}^j_N}^T +\hat{\bm{B}}^j_N}\bm{1}} = 0.
+Let $u$ denote the constant value of the solution such that the values of the solution at quadrature points is $u\bm{1}$.
+Multiplying both sides by the constant solution $\bm{u}_h^T$ yields that $\frac{1}{2}\pd{}{t}\LRp{\bm{u}_h^T\bm{M}^k\bm{u}_h}$ is equal to
+\begin{align*}
+-\frac{u}{2} \sum_{j=1}^d \LRp{\bm{1}^T \LRp{-\LRp{\hat{\bm{Q}}^j_N}^T +\hat{\bm{B}}^j_N}\bm{G}^k_{ij}
++\LRp{\bm{G}^k_{ij}}^T \LRp{-\LRp{\hat{\bm{Q}}^j_N}^T +\hat{\bm{B}}^j_N}\bm{1}}.% = 0.
 \end{align*}
@@
 \end{align*}
-We have that $\hat{\bm{Q}}^j_N\bm{1} = 0$ by construction, and $\sum_{j=1}^d \bm{1}^T \hat{\bm{Q}}^j_N\bm{G}^k_{ij} = 0$ since $G^k_{ij}$ satisfies the GCL.
-This implies that $\pd{}{t}\LRp{\bm{u}_h^T\bm{M}^k\bm{u}_h} = 0$ on $D^k$.
+By construction, $\hat{\bm{Q}}^j_N\bm{1} = 0$, and $\sum_{j=1}^d \bm{1}^T \hat{\bm{Q}}^j_N\bm{G}^k_{ij} = 0$ since $G^k_{ij}$ satisfies the GCL.
+This implies that the latter term vanishes and $\pd{}{t}\LRp{\bm{u}_h^T\bm{M}^k\bm{u}_h} = 0$ on $D^k$.
 Since $\bm{M}^k$ is symmetric and positive-definite, $\bm{u}_h^T\bm{M}^k\bm{u}_h$ is a norm on $\bm{u}_h$.
-Summing $\pd{}{t}\LRp{\bm{u}_h^T\bm{M}^k\bm{u}_h} = 0$ over all elements ensures that global constants are stationary solutions.
+Summing $\pd{}{t}\LRp{\bm{u}_h^T\bm{M}^k\bm{u}_h} = 0$ over all elements ensures that globally constant solutions are stationary and that $\pd{\bm{u}_h}{t} = 0$.
 \end{proof}
@@
 Semi-discrete entropy conservation on curved meshes requires that Assumption~\ref{ass:quad} holds for $v = 1$ and $v = G^k_{ij}$.
-We discuss specific choices of volume and surface quadrature on for which this assumption is valid.
-\note{Emphasize we're interested in taking $N_{\rm geo}$ as large as possible for a given quadrature rule, and that there is a focus on degree $(2N-1)$ quadrature rules.}
+We discuss specific choices of volume and surface quadrature on for which this assumption is valid, and summarize the maximum degree $N_{\rm geo}$ of the polynomial geometric approximation under which entropy stability holds for common choices of volume and surface quadrature.
+%\note{Emphasize we're interested in taking $N_{\rm geo}$ as large as possible for a given quadrature rule, and that there is a focus on degree $(2N-1)$ quadrature rules.}
 
@@
 Thus, the integrands in Assumption~\ref{ass:quad} $\pd{u}{\hat{x}_j} v \in P^{N+N_{\rm geo}-1}$ and $uv n_i \in P^{N+N_{\rm geo}}$ for $v = G^k_{ij}$.
-
-Thus, a volume quadrature rule of degree $(2N-1)$ and surface quadrature of degree $2N$ are sufficient to ensure semi-discrete entropy conservation for $N_{\rm geo}\leq N$.  
-\note{Add discussion on triangular elements.}
+A volume quadrature rule of degree $(2N-1)$ and surface quadrature of degree $2N$ are then sufficient to ensure semi-discrete entropy conservation for $N_{\rm geo}\leq N$.
+If the volume quadrature is instead exact for degree $2N$ polynomials, then entropy conservation is guaranteed up to $N_{\rm geo} \leq N+1$.
+%\note{Add discussion on triangular elements.}
 
@@
 However, in contrast to the simplicial case, it is not immediately clear that degree $(2N-1)$ volume quadratures exactly integrate $\int_{\hat{D}} \pd{u}{\hat{x}_j}v$ for $v = G^k_{ij}$ for tensor product elements.
-
-The key difference between simplicial and tensor product elements is the polynomial space in which the derivative lies.
+The difference between simplicial and tensor product elements is the polynomial space in which the derivative lies.
 In contrast to the simplicial case, if $u \in Q^N$, $\pd{u}{\hat{x}_j} \not\in Q^{N-1}$.
-Consider the three-dimensional case with $i = 1$, $u, v \in Q^N$.
-Then, $\pd{u}{\hat{x}_1} \in Q^{N-1,N,N}$, and differentiation reduces the polynomial degree in one coordinate but not others.
+Consider the three-dimensional case with $u, v \in Q^N$ and $i = 1$.
+Then, differentiation reduces the polynomial degree in one coordinate but not others and $\pd{u}{\hat{x}_1} \in Q^{N-1,N,N}$.
 As a result, $\pd{u}{\hat{x}_j}v \not\in Q^{2N-1}$, and a tensor product quadrature of degree $(2N-1)$ does not exactly integrate $\int_{\hat{D}} \pd{u}{\hat{x}_j}v$ for general $v\in Q^N$.
@@
 
-\section{A mortar-based implementation}
-
-For hybrid meshes, the choice of surface quadrature for an interface may vary depending on the neighboring elements.
+%\section{A mortar-based implementation}
+%
+%For hybrid meshes, the choice of surface quadrature for an interface may vary depending on the neighboring elements.
 Consider, for example, a mixed triangular-quadrilateral mesh where the quadrilaterals utilize GLL quadrature on each face.
@@
 To accomodate arbitrary combinations of face quadrature, we introduce a mortar-based implementation.
-
-We assume that each element
-
-\note{Finish - figure out notation for face interp!}
+%
+%We assume that each element 
+%
+%\note{Finish - figure out notation for face interp!}
 
@@
 
-In this section, we present two-dimensional experiments which verify the theoretical results presented.
+In this section, we present two-dimensional experiments which verify the theoretical results presented.  We first investigate triangular and quadrilateral meshes, and then   
 %on hybrid meshes consisting of quadrilaterals and triangles, as well as on non-conforming meshes of quadrilateral elements.
@@
 
-%\subsection{Non-conforming meshes}
-
-%\appendix
-%
-%\section{An explicit skew-symmetric entropy stable formulation for DG-SEM} 
-%
-%\note{$\bm{V}_q,\bm{P}_q = \bm{I}$, while $\bm{V}_f$ reduces to a permutation matrix.  }
-%
-%\note{For non-conforming faces, need to interpolate entropy variables to form fluxes.  Given this info, can rotate from skew form to strong form using mortar variables.  }
-
 \section{Conclusions}
+
+\note{Need mortar to .}
 

### 2018-09-28 docs/ESDG_mortar/main.tex
adding mortar ESDG draft

--- 2018-07-19 docs/ESDG_GQ/ext_decoupled_sbp.tex
@@
 %\newtheorem{corollary}{Corollary}
-%\newtheorem*{remark}{Remark}
+\newtheorem*{remark}{Remark}
 %\theoremstyle{theorem}
@@
 \author{Jesse Chan, David C.\ Del Rey Fernandez}
-\title{Entropy stable skew-symmetric discontinuous Galerkin formulations on non-conforming meshes}
+\title{Mortar-based entropy stable discontinuous Galerkin methods}
 \graphicspath{{./figs/}}
@@
 
-%\begin{frontmatter}
-%\title{Skew-symmetric formulations for entropy stable discontinuous Galerkin methods on non-conforming meshes}
-%
-%\author[rice]{Jesse Chan\corref{cor1}}
-%\ead{Jesse.Chan@caam.rice.edu}
-%\address[rice]{Department of Computational and Applied Mathematics, Rice University, 6100 Main St, Houston, TX, 77005}
-%
-%\begin{abstract}
-%Entropy stable high order methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.  
-%\end{abstract}
-%\end{frontmatter}
-
-%\tableofcontents
-
 \section{Introduction}
 
-\section{Entropy stable Gauss-Legendre collocation methods}
+\note{Boilerplate introduction on high order + stability}
 
-\note{Idea: decoupled SBP operators allow for the construction of entropy stable schemes by chaining together different sets of nodes through extrapolation (interpolation) operators.  }
+\section{Mortar methods for hybrid and non-conforming meshes}
 
-\section{Gauss-Legendre collocation on non-conforming meshes}
+This work is motivated by complications which arise when designing entropy stable couplings between elements which do not share the same boundary nodes.
+This can arise, for example, for hybrid and non-conforming meshes.
+
+\paragraph{Hybrid meshes:} It was shown in \cite{chan2018skew} that an entropy stable
+
+\paragraph{Non-conforming meshes}
 
@@
 
-\begin{figure}
+\section{Mortar formulation}
+
+\begin{figure}[!h]
 \centering
-\begingroup
-\captionsetup{width=.25\linewidth}
-\subfloat[Conforming faces]{\includegraphics[width=.25\textwidth]{figs/gauss1.png}}
-\hspace{4em}
-\subfloat[Non-conforming faces]{\includegraphics[width=.25\textwidth]{figs/gauss2.png}}
-\endgroup
-\caption{Examples of surface quadratures for conforming and non-conforming interfaces for Gauss collocation methods.  }
+\includegraphics[width=.65\textwidth]{figs/mortar.png}
+\caption{Illustration of mortar operators.
+The matrix $\bm{E}$ maps from volume quadrature points to surface quadrature points, $\bm{E}_m$ maps from surface to mortar surface points, and $\tilde{\bm{E}}_m$ maps from mortar surface points to surface points.
+}
 \label{fig:gqcon_noncon}
@@
 
-It is possible to efficiently interpolate from volume Gauss nodes to split-side Gauss nodes in two steps.
+Let $\bm{V}_q$ and $\bm{V}_f$ denote interpolation matrices which evaluate at volume and surface quadrature points, respectively, and let $\bm{W}, \bm{W}_f$ denote diagonal matrices whose entries consist of volume and surface quadrature weights
+\[
+\text{\note{add defs of matrices}}
+\]
+
+Let $\bm{E} = \bm{V}_f\bm{P}_q$ denote the polynomial mapping which extrapolates values from volume quadrature points to values at surface quadrature points.
+We also define $\bm{B}_i$ as the diagonal matrix containing products of quadrature weights and the $i$th component of the outward normal vector $\hat{\bm{n}}_i$
+\[
+\bm{B}_i = \bm{W}_f \diag{\hat{\bm{n}}_i}.
+\]
+
+An entropy stable skew-symmetric formulation can be given on the reference element $\hat{D}$ as follows:
+\[
+\bm{M}\td{\bm{u}_N}{t} + \begin{bmatrix} \bm{V}_q \\ \bm{V}_f \end{bmatrix}^T
+\LRp{\begin{bmatrix}
+\bm{Q}_i-\bm{Q}_i^T & \bm{E}^T\bm{B}_i\\
+-\bm{B}_i\bm{E} & \bm{0}
+\end{bmatrix} \circ \bm{F}_S}\bm{1} + \bm{V}_f^T\bm{B}_i\bm{f}_i^* = 0.  
+\]
+
+We incorporate mortars by modifying this formulation.
+Let $\bm{V}_m$ denote the matrix which maps volume quadrature points to values at surface \textit{mortar} points.
+We also need to introduce interpolation operators $\bm{T}_f, \bm{T}_m$ for surface trace spaces.
+Here, $\bm{T}_f$ maps from polynomials on the surface $\partial \hat{D}$ to surface quadrature points, while $\bm{T}_m$ maps from polynomials on $\partial \hat{D}$ to mortar quadrature points.
+We can then define surface and mortar mass and projection matrices
+\begin{align*}
+\bm{M}_m = \bm{T}_m^T\bm{W}_m\bm{T}_m, \qquad \bm{P}_m = \bm{M}_m^{-1}\bm{T}_m^T\bm{W}_m\\
+\bm{M}_f = \bm{T}_f^T\bm{W}_f\bm{T}_f, \qquad \bm{P}_f = \bm{M}_m^{-1}\bm{T}_f^T\bm{W}_f.
+\end{align*}
+Note that we have used the mortar mass matrix $\bm{M}_m$ in the above definition of the face projection.
+This is only necessary if the surface and mortar quadratures do not exactly integrate degree $2N$ polynomials.
+If both the surface and mortar quadratures are exact for polynomials of degree $2N$, then $\bm{M}_m = \bm{M}_f$, and no distinction is necessary between the two mass matrices.
+
+\begin{remark}
+In order to show stability, we require that both $\bm{P}_f$ and $\bm{P}_m$ are defined using the same mass matrix.
+We have used the mortar mass matrix $\bm{M}_m$ here; however, we could also use the surface mass matrix $\bm{M}_f$ in both $\bm{P}_f$ and $\bm{P}_m$.
+We use the mortar mass matrix in this work, as the mortar quadrature is generally more accurate than the surface quadrature for our use cases (e.g.\ hybrid and non-conforming meshes).
+\end{remark}
+
+We can now define operators which map between surface and mortar quadrature points.
+Let $\bm{E}_f$ denote the map from surface to mortar points, and let $\tilde{\bm{E}}_f $ denote the map from mortar to surface points.
+Both operators are defined through an $L^2$ projection to the trace space and interpolation to appropriate points
+\[
+\bm{E}_f = \bm{T}_m \bm{P}_f, \qquad \tilde{\bm{E}}_f = \bm{T}_f \bm{P}_m.
+\]  
+We also define a mortar boundary matrix
+\[
+\tilde{\bm{B}}_i = \bm{W}_m\diag{\hat{\bm{n}}_i}.
+\]
+Then, a mortar-based formulation can be given as follows:
+\begin{equation}
+\bm{M}\td{\bm{u}_N}{t} + \begin{bmatrix} \bm{V}_q \\ \bm{V}_f \\ \bm{V}_m \end{bmatrix}^T
+\LRp{\begin{bmatrix}
+\bm{Q}_i-\bm{Q}_i^T & \bm{E}^T\bm{B}_i &\\
+-\bm{B}_i\bm{E} &  & \bm{B}_i\tilde{\bm{E}}_f \\
+& -\tilde{\bm{B}}_i{\bm{E}}_f & 
+\end{bmatrix} \circ \bm{F}_S}\bm{1} + \bm{V}_m^T\tilde{\bm{B}}_i\bm{f}_i^* = 0.  
+\label{eq:mortar}
+\end{equation}
+Here, we have appended an extra row and column to the decoupled SBP matrix, and the inter-element numerical flux $\bm{f}^*$ is now computed in terms of the \emph{mortar} nodes.
+
+We first note that
+\begin{align*}
+\bm{B}_i\tilde{\bm{E}}_f &= \diag{\bm{W}_f} \bm{T}_f \bm{P}_m = \diag{\hat{\bm{n}}} \bm{W}_f \bm{T}_f \bm{M}_m^{-1}\bm{T}_m\bm{W}_m\\
+&=  \bm{W}_f \bm{T}_f \bm{M}_m^{-1}\bm{T}^T_m\bm{W}_m\diag{\hat{\bm{n}}_i} = \bm{P}_f^T\bm{T}^T_m \tilde{\bm{B}}_i = \LRp{\tilde{\bm{B}}_i{\bm{E}}_f}^T
+\end{align*}
+where we have used that $\bm{W}_m$ is diagonal and that the outward normal $\hat{\bm{n}}_i$ on the reference element is constant over each face.
+This implies skew-symmetry of the decoupled SBP operator and entropy stability of the overall system.
+
+\note{Finish}
+
+\section{A mortar-based implementation}
+
+While the formulation presented above is convenient for analysis, it is computationally expensive to implement.
+However, using properties of the mortar matrices $\bm{E}_f, \tilde{\bm{E}}_f$, we can rewrite the above formulation in a way which reflects a traditional mortar-based finite element implementation.
+In other words, we wish to implement (\ref{eq:mortar}) such that the only modification from the implementations in \cite{chan2017discretely} is a pre-processing step on mortar faces.
+
+We first note that, since $\bm{E}_f$ maps from surface nodes to mortar nodes, $\bm{V}_m = \bm{E}_f \bm{V}_f$.  Thus, we can rewrite the numerical flux contribution as
+\begin{align*}
+ \bm{V}_m^T\tilde{\bm{B}}_i\bm{f}_i^* &= \bm{V}_f^T \bm{E}_f^T \bm{W}_m\diag{\hat{\bm{n}}_i} \bm{f}_i^* = \bm{V}_f^T  \bm{P}_f^T\bm{T}_m^T\bm{W}_m\diag{\hat{\bm{n}}_i} \bm{f}_i^*\\
+ &= \bm{V}_f^T \bm{W}_f \diag{\hat{\bm{n}}_i} \bm{T}_m\bm{M}_m^{-1}\bm{T}_m^T\bm{W}_m \bm{f}_i^* =  \bm{V}_f^T \bm{B}_i \tilde{\bm{E}}_f\bm{f}_i^*.
+\end{align*}
+
+We can similarly reformulate the remaining mortar contributions within (\ref{eq:mortar}).  Multiplying out
+
+%\bm{M}\td{\bm{u}_N}{t} + \begin{bmatrix} \bm{V}_q \\ \bm{V}_f \\ \bm{V}_m \end{bmatrix}^T
+%\LRp{\begin{bmatrix}
+%\bm{Q}_i-\bm{Q}_i^T & \bm{E}^T\bm{B}_i &\\
+%-\bm{B}_i\bm{E} &  & \bm{B}_i\tilde{\bm{E}}_f \\
+%& -\tilde{\bm{B}}_i{\bm{E}}_f & 
+%\end{bmatrix} \circ \bm{F}_S}\bm{1} + \bm{V}_m^T\tilde{\bm{B}}_i\bm{f}_i^* = 0.
+
+%It is possible to efficiently interpolate from volume Gauss nodes to split-side Gauss nodes in two steps.
 We first interpolate to full-side Gauss nodes on the face, then interpolate from full-side  to split-side Gauss nodes.
 Each of these two operations can be done using only one-dimensional interpolation matrices, implying that the interpolation matrix from volume to split-side Gauss nodes can be decomposed into
-\[
-\bm{V}_{\rm vol-to-split} = \bm{V}_{\rm vol-to-full} \bm{V}_{\rm full-to-split}
-\]
-where each of the matrices in the decomposition possess either a sparse or Kronecker product structure.
-
-ESDG schemes based on decoupled SBP operators introduced in \cite{chan2017discretely} assume a single interpolation matrix which extrapolates from volume to surface nodes.
+%\[
+%\bm{V}_{\rm vol-to-split} = \bm{V}_{\rm vol-to-full} \bm{V}_{\rm full-to-split}
+%\]
+%where each of the matrices in the decomposition possess either a sparse or Kronecker product structure.  
+%
+%ESDG schemes based on decoupled SBP operators introduced in \cite{chan2017discretely} assume a single interpolation matrix which extrapolates from volume to surface nodes.
 These schemes require dyadic flux evaluations between all volume and surface nodes connected by this interpolation matrix.
 For non-conforming interfaces with split-side Gauss nodes on the face, this necessitates dyadic flux computations between all volume nodes and all split-side Gauss nodes.
-
-We show in this section how to construct an entropy stable Gauss collocation scheme based on two layers of surface nodes: full-side and split-side nodes.
+%
+%We show in this section how to construct an entropy stable Gauss collocation scheme based on two layers of surface nodes: full-side and split-side nodes.
 These surface nodes will be loosely through a skew-symmetric coupling term involving polynomial interpolation matrices between full-side and split-side nodes.
@@
 
-Let ${\bm{V}_f}\tilde{P}_f$ denote the matrix which projects to the trace space using split-side quadrature and evaluates at full-side quadrature points.
+%We first introduce notation which is helpful for operations involving surface traces.  Let the matrix $\bm{R}$ denote the matrix which maps coefficients of a function to coefficients of its trace.  
+%
+%Let ${\bm{V}_f}\tilde{P}_f$ denote the matrix which projects to the trace space using split-side quadrature and evaluates at full-side quadrature points.
 Similarly, let $\tilde{\bm{V}_f}{P}_f$ denote the matrix which projects to the trace space using full-side quadrature and evaluates at split-side quadrature points
-\begin{align*}
-\bm{C}^i_f &= \begin{pmatrix}
-\bm{0} & {\rm diag}\LRp{\hat{\bm{n}}_i}{\bm{V}_f}\tilde{\bm{P}}_f\\
--\tilde{\bm{V}_f}{\bm{P}}_f{\rm diag}\LRp{\hat{\bm{n}}_i} &\bm{0}
-\end{pmatrix}, \\
-\bm{S}^i_f &= \begin{pmatrix}
-\diag{\bm{w}_f} & \\
-& \diag{\tilde{\bm{w}}_f}
-\end{pmatrix}\bm{C}^i_f.
-\end{align*}
-It is straightforward to show that the matrix $\bm{S}_f$ is skew-symmetric.
-
-%{fig:gqcon_noncon}
-Let $\tilde{\bm{F}}_S$ denote the flux matrix, then 
-\[
-\begin{bmatrix}\bm{L}_f & \tilde{\bm{L}}_f\end{bmatrix} \LRp{\bm{C}^i_f\circ\tilde{\bm{F}}_S}\bm{1}
-\]
-is a skew-symmetric correction term which ensures entropy stability and high order accuracy (through consistency up to quadrature order).
-
-Since the difference is high order accurate, you can remove this correction term and still get high order accuracy.
+%\begin{align*}
+%\bm{C}^i_f &= \begin{pmatrix}
+%\bm{0} & {\rm diag}\LRp{\hat{\bm{n}}_i}{\bm{V}_f}\tilde{\bm{P}}_f\\
+%-\tilde{\bm{V}_f}{\bm{P}}_f{\rm diag}\LRp{\hat{\bm{n}}_i} &\bm{0}
+%\end{pmatrix}, \\
+%\bm{S}^i_f &= \begin{pmatrix}
+%\diag{\bm{w}_f} & \\
+%& \diag{\tilde{\bm{w}}_f}
+%\end{pmatrix}\bm{C}^i_f.
+%\end{align*}
+%It is straightforward to show that the matrix $\bm{S}_f$ is skew-symmetric.
+%
+%%{fig:gqcon_noncon}
+%Let $\tilde{\bm{F}}_S$ denote the flux matrix, then 
+%\[
+%\begin{bmatrix}\bm{L}_f & \tilde{\bm{L}}_f\end{bmatrix} \LRp{\bm{C}^i_f\circ\tilde{\bm{F}}_S}\bm{1}
+%\]
+%is a skew-symmetric correction term which ensures entropy stability and high order accuracy (through consistency up to quadrature order).   
+%
+%Since the difference is high order accurate, you can remove this correction term and still get high order accuracy.
 However, the spatial entropy RHS is no longer zero if you use inconsistent surface and coupling quadrature without the correction term (\note{prove this}).
-
-\note{Do we need 3 layers of surface nodes in 3D?  Interpolation to isotropic split-side nodes is a Kronecker product - does that reduce things down?  }
+%
+%\note{Do we need 3 layers of surface nodes in 3D?  Interpolation to isotropic split-side nodes is a Kronecker product - does that reduce things down?  }
 

### 2018-09-28 docs/skewform/main.tex
updating with new drafts

@@
 
+\usepackage{slashbox} % for slashbox in tables
+
 \usepackage{amsmath, amssymb, amsfonts}
@@
 %\newtheorem{assumption}{Assumption}
-
+\usepackage{hyperref}
 \usepackage[titletoc,toc,title]{appendix}
@@
 \renewcommand{\hat}{\widehat}
+\renewcommand{\tilde}{\widetilde}
 
@@
 \begin{document}
+
+\institute{J. Chan \at 
+Department of Computational and Applied Mathematics\\
+Rice University\\
+              6100 Main Street\\ 
+              Houston, TX, 77005 \\
+              \email{jesse.chan@rice.edu}
+}
 
@@
 where $\bm{W}, \bm{W}_f$ are diagonal matrices of volume and surface quadrature weights, respectively.
-The matrix $\bm{P}_q$ is a quadrature-based discretization of the $L^2$ projection operator $\Pi_N$ onto degree $N$ polynomials.
+The matrix $\bm{P}_q$ is a quadrature-based discretization of the $L^2$ projection operator $\Pi_N$ onto degree $N$ polynomials, which is given as follows: find $\Pi_N u \in V^N$ such that
+\begin{equation}
+\int_{\hat{D}} \Pi_N u v = \int_{\hat{D}} u v, \qquad \forall v \in V^N.
+\label{eq:l2proj}
+\end{equation}
 
@@
 By choosing specific quadrature rules, $\bm{D}^i_q$ recovers high order summation-by-parts finite difference operators in \cite{gassner2013skew, fernandez2014generalized, ranocha2018generalised} and certain operators in \cite{hicken2016multidimensional}.
-However, to address difficulties in designing efficient entropy stable interface terms for nonlinear conservation laws, the PI introduced in \cite{chan2017discretely} a new ``decoupled'' summation by parts matrix which builds interface terms directly into the approximation of the derivative.
+However, to address difficulties in designing efficient entropy stable interface terms for nonlinear conservation laws, a new ``decoupled'' summation by parts matrix was introduced in \cite{chan2017discretely} which builds interface terms directly into the approximation of the derivative.
 
@@
 \end{gather}
-Here, $\bm{B}^i$ is a boundary matrix, and $\bm{E}$ denotes the extrapolation matrix which maps values at volume quadrature points to values at surface quadrature points using quadrature-based $L^2$ projection and polynomial interpolation.
+Here, $\bm{B}^i$ is a boundary ``integration'' matrix, and $\bm{E}$ denotes the extrapolation matrix which maps values at volume quadrature points to values at surface quadrature points using quadrature-based $L^2$ projection and polynomial interpolation.
 
@@
 \[
-\int_{\hat{D}} \pd{u}{x_i} v, \qquad \int_{\partial \hat{D}} u v n_i
-\]
-are integrated exactly for all $u,v \in V^N\LRp{\hat{D}}$ and $i = 1,\ldots, d$.  Then, the decoupled SBP operator $\bm{D}^i_N$ (\ref{eq:QN}) satisfies a summation by parts property:
+\int_{\hat{D}} \pd{u}{\hat{x}_j} v, \qquad \int_{\partial \hat{D}} u v \hat{n}_j
+\]
+are integrated exactly for all $u,v \in V^N\LRp{\hat{D}}$ and $j = 1,\ldots, d$.  Then, the decoupled SBP operator $\bm{D}^i_N$ (\ref{eq:QN}) satisfies a summation by parts property:
 \begin{gather}
@@
 Let $u,v \in Q^N$ denote two arbitrary degree $N$ polynomials.
-The assumptions of Theorem~\ref{lemma:dsbp} are that the volume quadrature exactly integrates $\int_{\hat{D}} \pd{u}{x} v$ and that the surface quadrature exactly integrates $\int_{\partial \hat{D}} u v n_i$ on $\hat{D}$.
+The assumptions of Theorem~\ref{lemma:dsbp} are that the volume quadrature exactly integrates $\int_{\hat{D}} \pd{u}{x_j} v$ and that the surface quadrature exactly integrates $\int_{\partial \hat{D}} u v \hat{n}_j$ on $\hat{D}$.
 Because the $(N+1)$-point Gauss rule is exact for polynomials of degree $2N+1$ and the product $uv \in P^{2N}$ on each face, the surface quadrature satisfies the assumptions of Theorem~\ref{lemma:dsbp}.
@@
 
-These specific pairings of volume and surface quadratures appear naturally for hybrid meshes consisting of DG-SEM quadrilateral elements (using GLL volume quadrature) and triangular elements, as shown in Figure~\ref{fig:hybrid}.
+\paragraph{}These specific pairings of volume and surface quadratures appear naturally for hybrid meshes consisting of DG-SEM quadrilateral elements (using GLL volume quadrature) and triangular elements, as shown in Figure~\ref{fig:hybrid}.
 In Figure~\ref{subfig:hybrid1}, the surface quadrature is a $(N+1)$ point GLL rule, and results in a loss of the SBP property on the triangle.
@@
 \item the volume quadrature rule is exact for integrals of the form\\$\int_{\hat{D}} \pd{u}{\hat{x}_j} v$ for all $u \in V^N\LRp{\hat{D}}$, $j = 1,\ldots, d$.
-\item the surface quadrature rule is exact for integrals of the form\\$\int_{\partial \hat{D}} u v n_j$ for all $u \in V^N\LRp{\hat{D}}$, $j = 1,\ldots, d$, and $f \in \partial \hat{D}$.  
+\item the surface quadrature rule is exact for integrals of the form\\$\int_{\partial \hat{D}} u v \hat{n}_j$ for all $u \in V^N\LRp{\hat{D}}$, $j = 1,\ldots, d$, and $f \in \partial \hat{D}$.  
 \end{enumerate}
@@
 %If the volume quadrature is exact for polynomials of degree $N+M-1$ and the surface quadrature is exact for polynomials of degree $N+M$, 
-Suppose Assumption~\ref{ass:quad} holds, and let $\bm{v}_q$ denote the vector of values of $v(\bm{x})$ of at volume quadrature points.
+Suppose Assumption~\ref{ass:quad} holds for $v(\bm{x})\in V^N$, and let $\bm{v}_q$ denote the vector of values of $v(\bm{x})$ of at volume quadrature points.
 Let $u(\bm{x})$ be an $L^2$ integrable function, and let $\bm{u}_q$ denote the values of $u(\bm{x})$ at quadrature points.
@@
 \end{align*}
-where we have used Corollary~\ref{lemma:sbpcor} in the last equality.  Noting that $\bm{1}^T\bm{B}^i_N\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i \psi_i(\tilde{\bm{u}}_f)$ completes the proof.
+where we have used Lemma~\ref{lemma:sbpcor} in the last equality.  Noting that $\bm{1}^T\bm{B}^i_N\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i \psi_i(\tilde{\bm{u}}_f)$ completes the proof.
 %The final step of the proof is  where we have used Lemma~\ref{lemma:vdsbp} to integrate by parts and conclude that $\bm{1}^T\LRp{\bm{Q}^i_N}^T = 0$.
@@
 When the strong SBP property (\ref{eq:dsbp}) holds, this formulation is exactly equivalent to the strong formulation presented in \cite{chan2017discretely}.
-However, the only result necessary to prove Theorem~\ref{thm:esdg} is Corollary~\ref{lemma:sbpcor}.
+However, the only result necessary to prove Theorem~\ref{thm:esdg} is Lemma~\ref{lemma:sbpcor}.
 The decoupled SBP property (\ref{eq:dsbp}) is not utilized, and is not necessary to guarantee a semi-discrete conservation of entropy for the skew-symmetric formulation.
@@
 Because $\bm{F}_S$ is symmetric and $\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}$ is skew-symmetric, their Hadamard product is also skew-symmetric.
-Using that $\bm{x}^T\bm{A}\bm{x} = 0$ for any skew symmetric matrix $\bm{A}$, the volume term vanishes
-\[
-\bm{1}^T\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} = 0.
-\]
+Using that $\bm{x}^T\bm{A}\bm{x} = 0$ for any skew symmetric matrix $\bm{A}$, the volume term $\bm{1}^T\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1}$ vanishes.
 % and that the Hadamard product is commutable $\LRp{\bm{A}\circ\bm{B} }^T = \bm{A}^T\circ\bm{B}^T$
@@
 \label{sec:assump1}
-Apart from algebraic manipulations, only Corollary~\ref{lemma:sbpcor} is necessary to prove entropy conservation for (\ref{eq:esdg}).
-Corollary~\ref{lemma:sbpcor} requires that Assumption~\ref{ass:quad} holds for $v=1$.
+Apart from algebraic manipulations, only Lemma~\ref{lemma:sbpcor} is necessary to prove entropy conservation for (\ref{eq:esdg}).
+Lemma~\ref{lemma:sbpcor} requires that Assumption~\ref{ass:quad} holds for $v=1$.
 Thus, the volume and surface quadratures must be sufficiently accurate to guarantee that the mass matrix is positive definite and to integrate
 \begin{equation}
-\int_{\hat{D}} \pd{u}{x_i}, \qquad \int_{\partial \hat{D}} u n_i. \label{eq:affineints}
+\int_{\hat{D}} \pd{u}{x_i}, \qquad \int_{\partial \hat{D}} u \hat{n}_i. \label{eq:affineints}
 \end{equation}
@@
 %This minimal degree of surface quadrature accuracy will depend on the nature of the physical mesh, and will differ between affine and curved elements.
-
-%\subsection{On the accuracy of the skew-symmetric formulation}
-%
-%We note that, while the proofs in this Section show that it is possible to maintain entropy stability under reduced accuracy surface quadrature rules, the accuracy of the resulting schemes depend on the accuracy of the volume and surface quadrature.
-%We briefly remark on the accuracy of (\ref{eq:esdgSkew}).
-%
-%It was shown in \cite{chan2017discretely} that, when combined with quadrature-based $L^2$ projection and lifting operators, the decoupled SBP operator (\ref{eq:dsbp}) is a degree $N$ approximation to the derivative.
-This approximation can be interpreted as augmenting a volume approximation of the derivative with boundary correction terms \cite{chan2018efficient}.
-Let $f(\bm{x}),g(\bm{x})$ denote two $L^2$ integrable functions, and let $\bm{f}_N, \bm{g}_N$ denote the vectors of values of $f,g$ at both volume and surface quadrature points.
-A degree $N$ approximation $u\in V^N$ to $f\pd{g}{x_i}$ can be constructed via
-%\[
-%\bm{M}\bm{u} = \LRs{\begin{array}{c}
-%\bm{V}_q \\ \bm{V}_f\end{array}}^T \diag{\bm{f}_N}\bm{Q}^i_N\bm{g}_N, \qquad \bm{f}_N = \begin{bmatrix} \bm{f}_q\\ \bm{f}_f\end{bmatrix}, \quad \bm{g}_N = \begin{bmatrix} \bm{g}_q\\ \bm{g}_f\end{bmatrix},
-%\]
-%where $\bm{u}$ denotes the vector of coefficients for $u$.  Expanding terms using (\ref{eq:dsbp}) yields 
-%%\note{fix!  add $E = V_fP_q$}
-%\begin{align*}
-%\bm{M}\bm{u} &= \bm{V}_q^T \diag{\bm{f}_q} \bm{Q}^i \bm{g}_q + \frac{1}{2}\bm{V}_q^T\diag{\bm{f}_q}\bm{E}^T\bm{B}^i \LRp{\bm{g}_f -\bm{E}\bm{g}_q}  \\
-%&\qquad + \frac{1}{2}\bm{V}_f^T\diag{\bm{f}_f}\bm{B}^i \LRp{\bm{g}_f -\bm{E}\bm{g}_q}.
-%\end{align*}
-%where we have used that $\bm{B}^i$ is diagonal and that diagonal matrices commute.
-Let $\bm{v}$ are coefficients of some test function $v\in V^N$, and let $\bm{v}_q = \bm{V}_q\bm{v}$ and $\bm{v}_f = \bm{V}_f \bm{v}$.
-Multiplying by $\bm{v}^T$ on both sides yields that $\bm{v}^T\bm{M}\bm{u}$ is equal to
-%\begin{align*}
-%\LRp{\diag{\bm{f}_q}\bm{v}_q}^T \bm{Q}^i \bm{g}_q + \frac{1}{2}\LRp{\diag{\bm{f}_f}\bm{v}_f+\bm{E}\diag{\bm{f}_q}\bm{v}_q}^T\bm{B}^i \LRp{\bm{g}_f -\bm{E}\bm{g}_q} .
-%\end{align*}
-%This algebraic expression can be reinterpreted as a quadrature approximation of the following variational problem
-%\begin{equation}
-%\int_{\hat{D}} u v = \int_{\hat{D}} f\pd{\Pi_N g}{x_i}v + \int_{\partial \hat{D}} \LRp{g - \Pi_Ng} \LRp{\frac{fv + \Pi_N(fv)}{2}} \hat{n}_i, \qquad \forall v\in V^N.
-%\label{eq:var}
-%\end{equation}
-%
-%The skew-symmetric formulation can be interpreted as approximating a skew symmetric form of (\ref{eq:var}).  Integrating half of the volume term by parts yields 
-%\begin{align}
-%&\frac{1}{2} \int_{\hat{D}} \LRp{f\pd{\Pi_N g}{x_k}v - g \pd{\Pi_N\LRp{fv}}{x_i}} \label{eq:skewDN}\\
-%&+\frac{1}{2} \int_{\partial \hat{D}} \LRp{fgv\hat{n}_i + \LRp{g - \Pi_Ng} \LRp{\frac{fv + \Pi_N(fv)}{2}} \hat{n}_i}\nonumber
-%\end{align}
-%Approximating integrals using quadrature allows us to express (\ref{eq:skewDN}) as a matrix equation using $\bm{B}^i_N$ in (\ref{eq:dsbp})
-%%\begin{align*}
-%%\bm{v}^T\bm{M}\bm{u} &= \bm{v}^T\LRp{\LRs{\begin{array}{c}
-%%\bm{V}_q \\ \bm{V}_f\end{array}}}^T\diag{\bm{f}_N}\LRp{\bm{B}^i_N + \frac{1}{2}\LRp{\bm{Q}_N-\bm{Q}_N^T}}{\bm{g}_N}\\
-%%&= \bm{v}^T \LRp{\frac{1}{2}\LRp{\LRs{\begin{array}{c}
-%%\bm{V}_q \\ \bm{V}_f\end{array}}}^T \LRp{\diag{\bm{f}_N}{\LRp{\bm{Q}_N-\bm{Q}_N^T}}{\bm{g}_N}} +  \bm{V}_f^T \diag{\bm{f}_f}\bm{B}^i \bm{g}_f}.  
-%%\end{align*}
-%%Since this holds for all vectors $\bm{v}$, we recover the matrix equation
-%\begin{align*}
-%\bm{M}\bm{u} &= {\frac{1}{2}\LRp{\LRs{\begin{array}{c}
-%\bm{V}_q \\ \bm{V}_f\end{array}}}^T \LRp{\diag{\bm{f}_N}{\LRp{\bm{Q}_N-\bm{Q}_N^T}}{\bm{g}_N}} +  \bm{V}_f^T \diag{\bm{f}_f}\bm{B}^i \bm{g}_f}.
-%\end{align*}
-%
-%The accuracy of this approximation depends on the degree of accuracy of the volume and surface quadrature rules.  
-%
-%The entropy conservative skew-symmetric formulation (\ref{eq:esdgSkew})
-%
-%\note{Fix this up: if IBP holds for $v\in P^M$, the approximation should be $O(h^M)$}. 
-%
-%
-%\note{Add in continuous formulation and note that skew-symmetry approximates it.}
 
@@
 This is due to the fact that the geometric terms are now high order polynomials which vary spatially over each element.
-Moreover, Corollary~\ref{lemma:sbpcor} assumes affine geometric mappings, and does not hold on curved elements.
-In this section, we discuss how to extend Corollary~\ref{lemma:sbpcor} to curved simplicial and tensor product elements.
+Moreover, Lemma~\ref{lemma:sbpcor} assumes affine geometric mappings, and does not hold on curved elements.
+In this section, we discuss how to extend Lemma~\ref{lemma:sbpcor} to curved simplicial and tensor product elements.
 
@@
 
-For curved elements, steps must also be taken to ensure that the analogue of Lemma~\ref{lemma:vdsbp} and Corollary~\ref{lemma:sbpcor} hold at the discrete level.
+For curved elements, steps must also be taken to ensure that the analogue of Lemma~\ref{lemma:vdsbp} and Lemma~\ref{lemma:sbpcor} hold at the discrete level.
 Expanding out the condition $\bm{Q}^i_k\bm{1} = \bm{0}$ in terms of (\ref{eq:dncurved}) yields
@@
 \LRs{\begin{array}{c}
-{G}^k_{i1}\\
-{G}^k_{i2}\\
-{G}^k_{i3}\end{array}} = -\frac{1}{2}\LRp{\pd{I_{N_{\rm geo}}\bm{r}^k}{\hat{x}_j}-\pd{I_{N_{\rm geo}}\bm{r}^j}{\hat{x}_k}},\\
+{G}^k_{1j}\\
+{G}^k_{2j}\\
+{G}^k_{3j}\end{array}} = \begin{bmatrix}
+\LRp{-\hat{\Grad}\times I_{N_{\rm geo}}\LRp{x_3\hat{\Grad}x_2}}_j\\
+\LRp{\hat{\Grad}\times I_{N_{\rm geo}}\LRp{x_3\hat{\Grad}x_1}}_j\\
+\LRp{\hat{\Grad}\times I_{N_{\rm geo}}\LRp{x_1\hat{\Grad}x_2}}_j
+\end{bmatrix},\\
+%-\frac{1}{2}\LRp{\pd{I_{N_{\rm geo}}\bm{r}^k}{\hat{x}_j}-\pd{I_{N_{\rm geo}}\bm{r}^j}{\hat{x}_k}},\\
 N_{\rm geo} \leq \begin{cases}
@@
 These include, for example, GLL nodes on tensor product elements, as well as Warp and Blend nodes on non-tensor product elements \cite{warburton2006explicit, chan2015comparison}.}  The restriction on the maximum value of $N_{\rm geo}$ ensures that $G^k_{ij} \in V^N$ (e.g.
-$G^k_{ij} \in P^N$ on tetrahedral elements and $G^k_{ij}\in Q^N$ on hexahedral elements), which is necessary to guarantee (\ref{eq:dgcl}).
-
-Because the decoupled SBP operators $\bm{D}^i_k$ are now defined through (\ref{eq:dncurved}), Corollary~\ref{lemma:sbpcor} and the proof of entropy stability no longer hold for curved elements and must be modified.
+$G^k_{ij} \in P^N$ on tetrahedral elements and $G^k_{ij}\in Q^N$ on hexahedral elements), which is also necessary to guarantee (\ref{eq:dgcl}).
+
+Because the decoupled SBP operators $\bm{D}^i_k$ are now defined through (\ref{eq:dncurved}), Lemma~\ref{lemma:sbpcor} and the proof of entropy stability no longer hold for curved elements and must be modified.
 The introduction of curvilinear meshes will impose slightly different conditions on the accuracy of the surface quadrature.
@@
 It is also possible to replace the curved mass matrix $\bm{M}^k$ with a more easily invertible weight-adjusted approximation while maintaining high order accuracy, entropy stability, and local conservation \cite{chan2018discretely}.
-This approximation avoids the inversion of dense weighted $L^2$ mass matrices $\bm{M}^k$ on curved simplicial elements.
-On tensor product elements, this approximation is generally unnecessary as common choices of volume quadrature result in a diagonal (lumped) mass matrix \cite{carpenter2014entropy, parsani2016entropy, chan2018efficient}.
+This approximation avoids the inversion of dense weighted $L^2$ mass matrices $\bm{M}^k$ on curved simplicial elements, but is generally unnecessary on tensor product elements as common choices of volume quadrature result in a diagonal (lumped) mass matrix \cite{carpenter2014entropy, parsani2016entropy, chan2018efficient}.
 \end{remark}
@@
 A volume quadrature rule of degree $(2N-1)$ and surface quadrature of degree $2N$ are then sufficient to ensure semi-discrete entropy conservation for $N_{\rm geo}\leq N$.
+However, that in order to ensure that the mass matrix is positive-definite in Assumption~\ref{ass:quad}, the volume quadrature must be degree $2N$ in general on simplices.
 If the volume quadrature is instead exact for degree $2N$ polynomials, then entropy conservation is guaranteed up to $N_{\rm geo} \leq N+1$.
@@
 
-We first consider the two dimensional case, and show that Assumption~\ref{ass:quad} holds with $v = G^k_{ij}$ for all $N_{\rm geo} \leq N$.  On quadrilateral elements, $G^k_{ij}$ is
+We first consider the two dimensional case with volume quadrature of degree $(2N-1)$ and show that Assumption~\ref{ass:quad} holds with $v = G^k_{ij}$ for all $N_{\rm geo} \leq N$.
+On quadrilateral elements, $G^k_{ij}$ is
 \begin{align*}
@@
 
-We now consider the condition in Assumption~\ref{ass:quad} on the surface integrals $\int_{\partial \hat{D}} u v n_i$ for $v= G^k_{ij}$.  \note{Finish}
+We now consider the condition in Assumption~\ref{ass:quad} on the surface integrals $\int_{\partial \hat{D}} u v \hat{n}_j$ for $v= G^k_{ij}$.
+For left and right faces of the quadrilateral, $\hat{n}_2 = 0$, so this condition reduces to ensuring that the quantity $u G^k_{i1}$ is integrated exactly using quadrature for $i = 1,2$.
+Since $G^k_{i1}$ are degree $N_{\rm geo}-1$ in the $\hat{x}_2$ coordinate, $G^k_{i1}$ is degree $N_{\rm geo}-1$ and $uG^k_{i1} \hat{n}_1 \in Q^{N+N_{\rm geo}-1}$ along the left and right faces.
+Similarly, $uG^k_{i2} \hat{n}_2 \in Q^{N+N_{\rm geo}-1}$ along the top and bottom faces and are zero along the left and right faces.
+Thus, any surface quadrature rule exact for degree $N+N_{\rm geo}-1$ polynomials satisfies Assumption~\ref{ass:quad}, and for a degree $(2N-1)$ surface quadrature, entropy stability is guaranteed for any $N_{\rm geo}\leq N$.
 
 We now consider the three-dimensional case.
-Unlike quadrilateral elements, Assumption~\ref{ass:quad} does not hold for all $N_{\rm geo} \leq N$ on general curved hexahedra.
-Expanding out (\ref{eq:iconscurl}), one can show that $G^k_{ij} \in Q^{N_{\rm geo}}$ on hexahedral elements.
-Thus, if $N_{\rm geo} = N$ and $u\in Q^N$, $\pd{u}{\hat{x}_1} G^k_{i1} \in Q^{N-1,N,N}$.
-\note{Finish}.
-
-\subsection{Sufficient conditions on quadrature and $N_{\rm geo}$ to guarantee entropy conservation}
-
-For convenience, this section summarizes conditions on the accuracy of volume and surface quadratures under which one can guarantee entropy conservation for a given degree $N_{\rm geo}$ of a geometric mapping.
-
- \note{finish}
-
-%% $G^k_{1j} = \pd{\hat{x}_j}{x_1}$, $j = 1,2,3$ yields
-%%\begin{align*}
-%%G^k_{11} &= \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_2}x_3}}{\hat{x}_3}- \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_3}x_3}}{\hat{x}_2} \in Q^{N_{\rm geo}}\\
-%%G^k_{12} &= \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_3}x_3}}{\hat{x}_1}- \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_3}x_3}}{\hat{x}_1} \in Q^{N_{\rm geo}}\\
-%%G^k_{13} &= \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_1}x_3}}{\hat{x}_2}- \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_2}x_3}}{\hat{x}_1} \in Q^{N_{\rm geo}}.
-%%%Dt*((Ds*y).*z) - Ds*((Dt*y).*z)
-%%\end{align*}
-%
-%%Lemma~\ref{lemma:vdsbpcurved} implies that the polynomial degree $R$ of the surface geometric terms must be compatible with the accuracy of the surface quadrature.
+Unlike quadrilateral elements, Assumption~\ref{ass:quad} does not hold under degree $(2N-1)$ volume quadrature for $N_{\rm geo} = N$ on general curved hexahedra.
+Expanding out (\ref{eq:iconscurl}) gives
+\[
+G^k_{11} = \pd{}{\hat{x}_3} I_{N_{\rm geo}}\LRp{{x}_3 \pd{x_2}{\hat{x}_2}} - \pd{}{\hat{x}_2} I_{N_{\rm geo}}\LRp{{x}_3 \pd{x_2}{\hat{x}_3}} \in Q^{N_{\rm geo}}.
+\]
+Repeating for the other geometric terms, one can show that $G^k_{ij} \in Q^{N_{\rm geo}}$ on hexahedral elements.
+Thus, if $u\in Q^N$, $\pd{u}{\hat{x}_1} G^k_{i1} \in Q^{N+N_{\rm geo}-1,N+N_{\rm geo},N+N_{\rm geo}}$, and is only integrated exactly by volume quadratures of degree $(2N-1)$ for geometric degrees $N_{\rm geo}\leq (N-1)$.
+Similarly, Assumption~\ref{ass:quad} does not hold under degree $(2N-1)$ surface quadratures unless $N_{\rm geo} \leq (N-1)$, due to the fact that traces of $G^k_{ij}$ are degree $N_{\rm geo}$ polynomials in each coordinate.\footnote{It is possible to  construct the geometric terms for $N_{\rm geo} = N$ using a local $H_{\rm div}$ basis where
+\[
+\bm{r}^i \in Q^{N-1,N,N} \times Q^{N,N-1,N} \times Q^{N,N,N-1}.
+\]  
+Then, the geometric terms $\Grad \times \bm{r}^i \in Q^{N,N-1,N-1}\times Q^{N-1,N,N-1} \times Q^{N-1,N-1,N}$ with traces in $Q^{N-1}$, and Assumption~\ref{ass:quad} holds under degree $(2N-1)$ volume and surface quadrature.
+Numerical results show that this approach produces geometric terms which are up to an order of magnitude more accurate than those constructed simply by taking $N_{\rm geo} \leq (N-1)$.
+This approach will be investigated in more detail in future work.
+}
+
+Most implementations on tensor product elements utilize volume and surface quadratures of either degree $(2N-1)$ or $2N$.
+We summarize below for different pairings of volume and surface quadrature the maximum degree $N_{\rm geo}$ under which Assumption~\ref{ass:quad} is satisfied and entropy stability is guaranteed:
+\begin{enumerate}
+\item On quadrilateral elements, Assumption~\ref{ass:quad} holds for $N_{\rm geo} = N$ and any tensor product volume and surface quadratures of degree $(2N-1)$ 
+\item On hexahedral elements, Assumption~\ref{ass:quad} holds for $N_{\rm geo} = N-1$ and any tensor product volume and surface quadratures of degree $(2N-1)$.
+If the SBP property holds (for example, when the volume and surface quadratures are of degree $2N$) then Assumption~\ref{ass:quad} holds for $N_{\rm geo} = N$.
+\end{enumerate}
+
+On hexahedral elements, the condition $N_{\rm geo} = N-1$ is new.  \note{Finish}
+
+We note that this work assumes that the same quadrature is used to volume contributions in each direction, and can be refined slightly 
+\note{add line DG}
+
+% $G^k_{1j} = \pd{\hat{x}_j}{x_1}$, $j = 1,2,3$ yields
+%\begin{align*}
+%G^k_{11} &= \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_2}x_3}}{\hat{x}_3}- \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_3}x_3}}{\hat{x}_2} \in Q^{N_{\rm geo}}\\
+%G^k_{12} &= \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_3}x_3}}{\hat{x}_1}- \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_3}x_3}}{\hat{x}_1} \in Q^{N_{\rm geo}}\\
+%G^k_{13} &= \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_1}x_3}}{\hat{x}_2}- \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_2}x_3}}{\hat{x}_1} \in Q^{N_{\rm geo}}.
+%%Dt*((Ds*y).*z) - Ds*((Dt*y).*z)
+%\end{align*}
+
+%Lemma~\ref{lemma:vdsbpcurved} implies that the polynomial degree $R$ of the surface geometric terms must be compatible with the accuracy of the surface quadrature.
 This, in turn, depends on the polynomial degree $N_{\rm geo}$ to which geometric terms are approximated.
@@
 Note - this still works regardless for conforming hexes because of the full matrix SBP property.}
-%
-%
+
 %\begin{lemma}
@@
 
+\subsection{On the accuracy of the skew-symmetric formulation}
+
+We note that, while it is possible to maintain entropy stability under reduced accuracy quadratures, the accuracy of the resulting schemes depend on the accuracy of the volume and surface quadrature rules.
+In this section, we derive truncation error estimates which take into account the degree of accuracy of the volume and surface quadrature.
+
+It was shown in \cite{chan2017discretely} that, when combined with quadrature-based $L^2$ projection and lifting operators, the decoupled SBP operator (\ref{eq:dsbp}) is a degree $N$ approximation to the derivative.
+This approximation can be interpreted as augmenting a volume approximation of the derivative with boundary correction terms \cite{chan2018efficient}.
+Let $f(\bm{x}),g(\bm{x})$ denote two $L^2$ integrable functions, and let $\bm{f}_N, \bm{g}_N$ denote the vectors of values of $f,g$ at both volume and surface quadrature points.
+A degree $N$ approximation $u\in V^N$ to $f\pd{g}{x_i}$ can be constructed via
+\begin{equation}
+\bm{M}\bm{u} = \LRs{\begin{array}{c}
+\bm{V}_q \\ \bm{V}_f\end{array}}^T \diag{\bm{f}_N}\bm{Q}^i_N\bm{g}_N, \qquad \bm{f}_N = \begin{bmatrix} \bm{f}_q\\ \bm{f}_f\end{bmatrix}, \quad \bm{g}_N = \begin{bmatrix} \bm{g}_q\\ \bm{g}_f\end{bmatrix},
+\label{eq:dsbpmatrixform}
+\end{equation}
+where $\bm{u}$ denotes the vector of coefficients for $u$.  
+%Expanding terms using (\ref{eq:dsbp}) yields 
+%%\note{fix!  add $E = V_fP_q$}
+%\begin{align*}
+%\bm{M}\bm{u} &= \bm{V}_q^T \diag{\bm{f}_q} \bm{Q}^i \bm{g}_q + \frac{1}{2}\bm{V}_q^T\diag{\bm{f}_q}\bm{E}^T\bm{B}^i \LRp{\bm{g}_f -\bm{E}\bm{g}_q}  \\
+%&\qquad + \frac{1}{2}\bm{V}_f^T\diag{\bm{f}_f}\bm{B}^i \LRp{\bm{g}_f -\bm{E}\bm{g}_q}.
+%\end{align*}
+%where we have used that $\bm{B}^i$ is diagonal and that diagonal matrices commute.
+Let $\bm{v}$ are coefficients of some test function $v\in V^N$, and let $\bm{v}_q = \bm{V}_q\bm{v}$ and $\bm{v}_f = \bm{V}_f \bm{v}$.
+Multiplying by $\bm{v}^T$ on both sides yields that $\bm{v}^T\bm{M}\bm{u}$ is equal to
+%\begin{align*}
+%\LRp{\diag{\bm{f}_q}\bm{v}_q}^T \bm{Q}^i \bm{g}_q + \frac{1}{2}\LRp{\diag{\bm{f}_f}\bm{v}_f+\bm{E}\diag{\bm{f}_q}\bm{v}_q}^T\bm{B}^i \LRp{\bm{g}_f -\bm{E}\bm{g}_q} .
+%\end{align*}
+
+This algebraic expression (\ref{eq:dsbpmatrixform}) can be reinterpreted as a quadrature approximation of a variational problem, which can be mapped to a physical element $D^k$.
+We seek to approximate $f\pd{g}{x_i}$ by $u \in V^N$ such that, $\forall v\in V^N$
+\begin{equation}
+\int_{D^k} u v = \int_{D^k} f\pd{\Pi_N g}{{x}_i}v + \int_{\partial {D}^k} \LRp{g - \Pi_Ng} \LRp{\frac{fv + \Pi_N(fv)}{2}} {n}^k_i,\label{eq:var}
+\end{equation}
+where $\Pi_N$ is the $L^2$ projection operator (\ref{eq:l2proj}).  
+Integrating half of the volume term by parts yields the skew-symmetric form of (\ref{eq:var})
+\begin{align}
+\int_{D^k} u  v &=\frac{1}{2} \int_{{D^k}} \LRp{f\pd{\Pi_N g}{{x}_i}v - g \pd{\Pi_N\LRp{fv}}{{x}_i}} \label{eq:skewDN}\\
+&+\frac{1}{2} \int_{\partial {D}^k} \LRp{fgv + \LRp{g - \Pi_Ng} \LRp{{fv + \Pi_N(fv)}{}}} {n}^k_i \qquad \forall v\in V^N\nonumber,
+\end{align}
+which yields the matrix formulation
+\begin{align}
+\bm{M}\bm{u} =& \frac{1}{2} \LRs{\begin{array}{c}
+\bm{V}_q \\ \bm{V}_f\end{array}}^T \diag{\bm{f}_N}\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}\bm{g}_N \label{eq:dsbpmatrixform}\\
+&+ \frac{1}{2}\bm{V}_f^T\bm{W}_f\diag{\bm{n}^k_i} \LRp{\bm{f}_f \circ \bm{g}_f}.\nonumber
+\end{align}
+
+When $f(\bm{x}) = 1$, the formulations (\ref{eq:var}), (\ref{eq:skewDN}) can be recast in terms of linear operators $D^i_h, \tilde{D}^i_h$ which capture ``strong'' and ``weak'' formulations of the derivative.
+Let $u(\bm{x})$ be a sufficiently regular function.
+Then, the strong formulation operator $D^i_h: H^1\LRp{D^k}\rightarrow V^N$ can be defined through
+\begin{align*}
+\int_{D^k} D^i_h u  v = \int_{D^k} \pd{\Pi_N u}{x_i}v + \frac{1}{2}\int_{\partial D^k} \LRp{u - \Pi_Nu} v {n}^k_i, \qquad \forall v\in V^N.
+\end{align*}
+Similarly, the weak form operator $\tilde{D}^i_h: H^1\LRp{D^k}\rightarrow V^N$ can be defined such that
+\begin{align*}
+\int_{D^k} \tilde{D}^i_hu v &= \int_{D^k} { -u \pd{{v}}{x_i}} + \frac{1}{2}\int_{\partial D^k}  \LRp{u+\Pi_N u} v {n}^k_i, \qquad \forall v\in V^N.
+\end{align*}
+Then, for $f(\bm{x}) = 1$, the skew-symmetric formulation (\ref{eq:skewDN}) is then equivalent to approximating derivatives using the operator $\frac{1}{2}\LRp{D^i_h - \tilde{D}^i_h}$.
+
+These two operators $D^i_h, \tilde{D}^i_h$ are identical if the integrals are computed using sufficiently accurate volume and surface quadrature rules, e.g.\ if $\bm{Q}^i_N$ satisfies the SBP property (\ref{eq:dsbp}).
+If the SBP property (\ref{eq:dsbp}) does not hold, then $D^i_h, \tilde{D}^i_h$ satisfy different error estimates.
+
+Let $\LRp{\cdot,\cdot}, \LRa{\cdot,\cdot}$ now denote volume and surface $L^2$ inner products over $D^k$, and let $\LRp{\cdot,\cdot}_h, \LRa{\cdot,\cdot}_h$ denote their approximations using quadrature.
+We assume that the volume and surface quadrature rules satisfy Assumption~\ref{ass:quad} for $v = 1, G^k_{ij}$, and that they are additionally exact for polynomials of degree $r_{\rm vol}, r_{\rm surf}$.
+We furthermore assume that quadrature errors are of the form \cite{baker1976effect, brenner2007mathematical}
+\begin{align}
+\LRb{\LRp{u,v} - \LRp{u,v}_h} \leq C_{\rm vol} h^{\alpha} \nor{u}_{H^{\alpha}\LRp{D^k}}\nor{v}_{H^{\alpha}\LRp{D^k}},\qquad \alpha = r_{\rm vol}+1, \label{eq:quadacc}\\
+\LRb{\LRa{u,v} - \LRa{u,v}_h} \leq C_{\rm surf} h^{\beta} \nor{uv}_{H^{\beta}\LRp{\partial D^k}}, \qquad \beta = r_{\rm surf}+1.\nonumber
+\end{align}
+Then, we have the following error estimates for $D^i_h, \tilde{D}^i_h$ on affine meshes:
+\begin{lemma}
+Let $D^k$ be affine elements, let $u(\bm{x})$ be a sufficiently regular function, and suppose $r_{\rm vol} \geq 2N$ and $r_{\rm surf} \geq N$.
+Then, if the SBP property (\ref{eq:dsbp}) does not hold,
+\begin{align*}
+\nor{\pd{u}{x_i} - \tilde{D}^i_h u}_{L^2\LRp{\Omega}} \leq C_1 h^{ \gamma}, \qquad \nor{\pd{u}{x_i} - D^i_h u}_{L^2\LRp{\Omega}} \leq C_2 h^{\max\LRp{N, \gamma}},
+\end{align*}
+where $\gamma = \min\LRc{\LRp{N+1}, \LRp{r_{\rm vol} - N + 1}, \LRp{r_{\rm surf} - N}}$, and $C_1,C_2$ are independent of $h$.
+If the SBP property (\ref{eq:dsbp}) does hold, then both errors are bounded by $O\LRp{h^{\max\LRp{N,\gamma}}}$.
+\label{lemma:quadacc}
+\end{lemma}
+\begin{proof}
+The error can be decomposed as
+\[
+\pd{u}{x_i} - \tilde{D}^i_h u = \LRp{\pd{u}{x_i} - \Pi_N\pd{u}{x_i}} + \LRp{\Pi_N\pd{u}{x_i} - \tilde{D}^i_h u}.
+\]
+Standard finite element approximation estimates \cite{brenner2007mathematical} give that 
+\[
+\nor{\pd{u}{x_i} - \Pi_N\pd{u}{x_i}}_{L^2\LRp{\Omega}} \lesssim h^{N+1} \sqrt{\sum_k \nor{u}_{H^{N+1}\LRp{D^k}}}.
+\]
+The latter term can be estimated using (\ref{eq:quadacc}).  We will first consider the error for $\tilde{D}^i_h$.  Let $\delta = \Pi_N\pd{u}{x} - \tilde{D}^i_h$.  Then, on each element $D^k$, 
+\[
+\nor{\Pi_N\pd{u}{x_i} - \tilde{D}^i_hu}_{L^2\LRp{D^k}}^2 = \LRb{\LRp{\Pi_N\pd{u}{x_i} - \tilde{D}^i_h,\delta} } = \LRb{\LRp{\pd{u}{x_i}, \delta} - \LRp{\tilde{D}^i_h,\delta} },
+\]
+where we have used the definition of $\Pi_N$ (\ref{eq:l2proj}) and the fact that $\delta \in V^N$.  Integrating $\LRp{\pd{u}{x_i}, \delta}$ by parts and using the definition of $\tilde{D}^i_h$ gives
+\begin{gather*}
+\nor{\Pi_N\pd{u}{x_i} - \tilde{D}^i_hu}_{L^2\LRp{D^k}}^2 \leq  \LRb{\LRa{u n^k_i, v}- \LRa{un^k_i, v}_h} + \LRb{\LRp{u, \pd{v}{x_i}} - \LRp{u, \pd{v}{x_i}}_h}\\
+\lesssim \LRp{h^{\beta}\nor{uv}_{H^{\beta}\LRp{\partial D^k}} +  h^{\alpha}\nor{u}_{H^{\alpha}\LRp{D^k}} \nor{\pd{v}{x_i}}_{H^{\alpha}\LRp{D^k}}}.
+\end{gather*}
+The terms $\nor{uv}_{H^{\beta}\LRp{\partial D^k}}, \nor{\pd{v}{x_i}}_{H^{\alpha}\LRp{D^k}}$ can be bounded using inverse and inverse trace inequalities \cite{brenner2007mathematical, chan2018discretely} and the fact that $v \in V^N$ and $\beta \geq (N+1)$
+\begin{align*}
+\nor{uv}_{H^{\beta}\LRp{\partial D^k}} &\lesssim h^{-1/2} \nor{uv}_{H^{\beta}\LRp{D^k}} \lesssim h^{-1/2} \nor{u}_{H^{\beta}\LRp{D^k}}\nor{v}_{H^{\beta}\LRp{D^k}}\\
+&\lesssim h^{-N-1/2} \nor{u}_{H^{\beta}\LRp{D^k}}\nor{v}_{L^2\LRp{D^k}} \\
+\nor{\pd{v}{x_i}}_{H^{\alpha}\LRp{D^k}} &\lesssim h^{-(N-1)}\nor{v}_{L^2\LRp{D^k}}.
+\end{align*}
+
+Noting that $ \nor{v}_{L^2\LRp{D^k}} = \nor{\Pi_N\pd{u}{x_i} - \tilde{D}^i_hu}_{L^2\LRp{D^k}}$, dividing through by $\nor{v}_{L^2\LRp{D^k}}$ gives
+\[
+\nor{\Pi_N\pd{u}{x_i} - \tilde{D}^i_hu}_{L^2\LRp{D^k}} \lesssim h^{\beta - N} \nor{u}_{H^{\beta}\LRp{D^k}} + h^{\alpha - N + 1} \nor{u}_{H^{\alpha}\LRp{D^k}}.
+\]
+ \note{Finish - how to lose extra factor of $1/h$ in trace estimate?}
+\end{proof}
+
+\begin{remark}
+Numerical experiments indicate that the assumptions of Lemma~\ref{lemma:quadacc} are sufficient, but not necessary.
+For specific combinations of volume and surface quadrature, we observe that the errors in Lemma~\ref{lemma:quadacc} converge faster than expected given the degree of exactness of each rule.
+These special cases will be investigated in more detail in future work.
+\end{remark}
+
 \section{Numerical experiments}
@@
 
-In this section, we present two-dimensional experiments which verify the theoretical results presented.  We first investigate triangular and quadrilateral meshes, and then   
+In this section, we present two-dimensional experiments which verify the theoretical results presented.
+We begin by investigating the stability and accuracy of the skew-symmetric formulation on triangular and quadrilateral meshes, and conclude with two-dimensional experiments on a hybrid mesh containing mixed quadrilateral and triangular elements.
 %on hybrid meshes consisting of quadrilaterals and triangles, as well as on non-conforming meshes of quadrilateral elements.
+
+\subsection{Quadrature accuracy and rates of convergence}
+
+\subsection{Quadrature accuracy and timestep restrictions}
 
@@
 
+\begin{itemize}
+\item \note{GQ-GQ hexes, tris}
+\item \note{GLL-GQ hexes, tris}
+\item \note{GQ-GLL hexes, GLL tris}
+\end{itemize}
+
 \section{Conclusions}
+
+\note{GLL quadrature really is super convenient.}
 

### 2018-10-19 docs/noncon/ext_decoupled_sbp.tex
updating talks + matlab

--- 2018-07-19 docs/ESDG_GQ/ext_decoupled_sbp.tex
@@
 %\end{frontmatter}
-
 %\tableofcontents
@@
 
-\section{Entropy stable Gauss-Legendre collocation methods}
+\section{Entropy stable DG methods on tensor product elements}
 
@@
 
-\section{Gauss-Legendre collocation on non-conforming meshes}
+\section{Decoupled SBP operators for non-conforming meshes}
 
@@
 
-\begin{figure}
-\centering
-\begingroup
-\captionsetup{width=.25\linewidth}
-\subfloat[Conforming faces]{\includegraphics[width=.25\textwidth]{figs/gauss1.png}}
-\hspace{4em}
-\subfloat[Non-conforming faces]{\includegraphics[width=.25\textwidth]{figs/gauss2.png}}
-\endgroup
-\caption{Examples of surface quadratures for conforming and non-conforming interfaces for Gauss collocation methods.  }
-\label{fig:gqcon_noncon}
-\end{figure}
+%\begin{figure}
+%\centering
+%\begingroup
+%\captionsetup{width=.25\linewidth}
+%\subfloat[Conforming faces]{\includegraphics[width=.25\textwidth]{figs/gauss1.png}}
+%\hspace{4em}
+%\subfloat[Non-conforming faces]{\includegraphics[width=.25\textwidth]{figs/gauss2.png}}
+%\endgroup
+%\caption{Examples of surface quadratures for conforming and non-conforming interfaces for Gauss collocation methods.  }
+%\label{fig:gqcon_noncon}
+%\end{figure}
 
@@
 
-\note{Do we need 3 layers of surface nodes in 3D?  Interpolation to isotropic split-side nodes is a Kronecker product - does that reduce things down?  }
+\section{An efficient mortar-based implementation}
+
+\subsection{Three-dimensional considerations}
+
+\note{We need 3 layers of surface nodes in 3D.
+Interpolation to isotropic split-side nodes can be done in a tensor-product fashion.
+May not matter: interpolation on 2D faces to a refined set of 4 faces is $O(N^4)$ operations, so it doesn't change the overall cost of the code.
+}
+
+\section{Numerical experiments}
 

### 2018-10-19 docs/skewform/main.tex
updating talks + matlab

@@
 
-%\documentclass[review]{siamart0216}
 \documentclass{svjour3}                     % onecolumn (standard format)
@@
 \usepackage{pgfplotstable}
-\usepackage{slopetri}
+\usepackage{slopetri} % my own package for drawing triangle slopes
 
@@
 \author{Jesse Chan}
-\title{Skew-symmetric entropy stable discontinuous Galerkin formulations}
+\title{Skew-symmetric entropy stable discontinuous Galerkin formulations}% with applications to hybrid meshes}
+\titlerunning{Skew-symmetric entropy stable DG formulations}
 
@@
 Entropy stable high order methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.
-The construction of such schemes has relied on the use of carefully chosen collocation points \cite{gassner2013skew, fisher2013high, carpenter2014entropy, chan2018efficient} or volume and surface quadrature rules \cite{chan2017discretely, chan2018discretely} to produce operators which satisfy a summation-by-parts (SBP) property.
-In this work, we show how to construct skew-symmetric schemes which are entropy stable even for volume and surface quadratures under which an SBP property does not hold.
+The construction of such schemes has relied on the use of carefully chosen collocation points \cite{gassner2013skew, fisher2013high, carpenter2014entropy, crean2018entropy, chan2018efficient} or volume and surface quadrature rules \cite{chan2017discretely, chan2018discretely} to produce operators which satisfy a summation-by-parts (SBP) property.
+In this work, we show how to construct skew-symmetric schemes which are entropy stable even for volume and surface quadratures under which a traditional SBP property does not hold.
 These skew-symmetric formulations avoid the use of a ``strong'' SBP property, and require only that operators exactly differentiate constants and satisfy a discrete form of the fundamental theorem of calculus.
@@
 Let $\hat{\bm{n}}$ denote the scaled outward normal vector $\hat{\bm{n}} = \LRc{\hat{n}_1\hat{J}_f,\ldots,\hat{n}_d\hat{J}_f}$, where $\hat{J}_f$ is the determinant of the Jacobian of the mapping of a face of $\partial \hat{D}$ to a reference face.
-Let $\hat{\bm{n}}_i$ denote the vector containing values of the $i$th component $\hat{n}_i\hat{J}f$ at all surface quadrature points, and let $\bm{Q}^i = \bm{W}\bm{D}^i_q$.
+Let $\hat{\bm{n}}_i$ denote the vector containing values of the $i$th component $\hat{n}_i\hat{J}f$ at all surface quadrature points, and let $\bm{Q}^i = \bm{W}\bm{D}^i_q = \bm{W}\bm{V}_q\bm{D}^i\bm{P}_q$.
 The ``decoupled'' summation by parts operator $\bm{D}^i_N$ is defined as the block matrix involving both volume and surface quadratures
@@
 \begin{theorem}%[Theorem 1 in \cite{chan2017discretely}]
-Assume that the volume and surface quadrature rules are sufficiently accurate such that the quantities
-\[
-\int_{\hat{D}} \pd{u}{\hat{x}_j} v, \qquad \int_{\partial \hat{D}} u v \hat{n}_j
-\]
-are integrated exactly for all $u,v \in V^N\LRp{\hat{D}}$ and $j = 1,\ldots, d$.  Then, the decoupled SBP operator $\bm{D}^i_N$ (\ref{eq:QN}) satisfies a summation by parts property:
+Assume that $\bm{Q}^i$ satisfies a generalized SBP property
+\[
+\bm{Q}^i = \bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T.
+\]
+Then, the decoupled SBP operator $\bm{D}^i_N$ (\ref{eq:QN}) satisfies a summation by parts property:
 \begin{gather}
@@
 
-The assumptions of Theorem~\ref{lemma:dsbp} are satisfied for sufficiently accurate volume and surface quadratures.
+The matrix $\bm{Q}^i$ satisfies a generalized SBP property if the volume and surface quadrature rules are sufficiently accurate such that the quantities
+\[
+\int_{\hat{D}} \pd{u}{\hat{x}_j} v, \qquad \int_{\partial \hat{D}} u v \hat{n}_j
+\]
+are integrated exactly for all $u,v \in V^N\LRp{\hat{D}}$ and $j = 1,\ldots, d$.
+This implies that Theorem~\ref{lemma:dsbp} is satisfied for sufficiently accurate volume and surface quadratures.
 For example, on simplicial elements, (\ref{eq:dsbp}) holds if the volume quadrature is exact for polynomial integrands of total degree $(2N-1)$, and the surface integral is exact for degree $2N$ polynomials on each face.
@@
 \begin{remark}
-It should be stressed that the assumptions of Theorem~\ref{lemma:dsbp} are sufficient but not necessary conditions.
-For example, suppose $\hat{D}$ is a quadrilateral element, and that the volume quadrature is a tensor product of $(N+1)$ point one-dimensional Gauss-Legendre-Lobatto (GLL) rule.
+It should be stressed that sufficiently accurate volume and surface quadratures are not necessary conditions for Theorem~\ref{lemma:dsbp}.
+For example, it is well known that the use of $(N+1)$ point Gauss-Legendre-Lobatto (GLL) rules for both volume and surface quadratures result in a generalized SBP property, despite the fact that these rules are only accurate for degree $(2N-1)$ polynomials.
+%suppose $\hat{D}$ is a quadrilateral element, and that the volume quadrature is a tensor product of $(N+1)$ point one-dimensional Gauss-Legendre-Lobatto (GLL) rule.
 If the surface quadrature also taken to be an $(N+1)$-point GLL rule over each face, then $\bm{Q}^i$ satisfies a traditional SBP property, which implies the decoupled SBP property for $\bm{Q}^i_N$.
+More generally, it is straightforward to show that Theorem~\ref{lemma:dsbp} holds whenever $\bm{Q}^i$ satisfies a generalized SBP property
+%\[
+%\bm{Q}^i = \bm{E}^T\bm{B}^i\bm{E}^T - \LRp{\bm{Q}^i}^T.
+%\]
+\label{remark:dsbp}
 \end{remark}
+
+We are interested in addressing cases where the generalized SBP property (and as a result, the strong decoupled SBP property in Theorem~\ref{lemma:dsbp}) do not hold.
 
@@
 \label{ass:quad}
-Let $v \in V^{N}$ denote a fixed polynomial.  We assume that: 
+Let $v \in V^{N}$ denote some fixed polynomial.  We assume that: 
 \begin{enumerate}
@@
 \begin{remark}
+The skew-symmetric operator ${\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T}$ in (\ref{eq:skewformcurved}) can be rewritten as
+\begin{align*}
+%\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T =&
+\sum_{j=1}^d \diag{\bm{G}^k_{ij}}{ \frac{1}{2}\LRp{\hat{\bm{Q}}_N^j-\LRp{\hat{\bm{Q}}_N^j}^T}}+  \frac{1}{2} \LRp{\hat{\bm{Q}}_N^j-\LRp{\hat{\bm{Q}}_N^j}^T}\diag{\bm{G}^k_{ij}}.
+\end{align*}
+Thus, the volume terms in the skew-symmetric formulation may be applied using only the skew-symmetric part of the reference SBP operator $\hat{\bm{Q}}^j_N$.
+\end{remark}
+
+\begin{remark}
 It is also possible to replace the curved mass matrix $\bm{M}^k$ with a more easily invertible weight-adjusted approximation while maintaining high order accuracy, entropy stability, and local conservation \cite{chan2018discretely}.
@@
 \subsection{On quadrature conditions for Assumption~\ref{ass:quad} for $v = 1$ and $v = G^k_{ij}$}
-
+\label{sec:curvgeo}
 Semi-discrete entropy conservation on curved meshes requires that Assumption~\ref{ass:quad} holds for $v = 1$ and $v = G^k_{ij}$.
@@
 \]  
-Then, the geometric terms $\Grad \times \bm{r}^i \in Q^{N,N-1,N-1}\times Q^{N-1,N,N-1} \times Q^{N-1,N-1,N}$ with traces in $Q^{N-1}$, and Assumption~\ref{ass:quad} holds under degree $(2N-1)$ volume and surface quadrature.
-Numerical results show that this approach produces geometric terms which are up to an order of magnitude more accurate than those constructed simply by taking $N_{\rm geo} \leq (N-1)$.
+Then, the geometric terms satisfy $\Grad \times \bm{r}^i \in Q^{N,N-1,N-1}\times Q^{N-1,N,N-1} \times Q^{N-1,N-1,N}$ with traces in $Q^{N-1}$, and Assumption~\ref{ass:quad} holds under degree $(2N-1)$ volume and surface quadrature.
+This approach will be investigated in more detail in future work.
+%Numerical results show that this approach produces geometric terms which are up to an order of magnitude more accurate than those constructed simply by taking $N_{\rm geo} \leq (N-1)$.
 This approach will be investigated in more detail in future work.
@@
 
-On hexahedral elements, the condition $N_{\rm geo} = N-1$ is new.  \note{Finish}
-
-We note that this work assumes that the same quadrature is used to volume contributions in each direction, and can be refined slightly 
-\note{add line DG}
+We note that the condition $N_{\rm geo} = N-1$ is non-standard for tensor product elements.
+However, this condition is only necessary for entropy stability when $\bm{Q}^i$ does not satisfy a generalized SBP property (see Remark~\ref{remark:dsbp}).
+To the author's knowledge, this setting has not been considered within the literature on tensor product elements.
+
+%\note{Finish}
+%\begin{remark}
+%We note that this work assumes that the same quadrature is used to volume contributions in each direction, and can be refined slightly 
+%\note{add line DG}
 
@@
 
-%\section{A mortar-based implementation}
-%
-%For hybrid meshes, the choice of surface quadrature for an interface may vary depending on the neighboring elements.
-Consider, for example, a mixed triangular-quadrilateral mesh where the quadrilaterals utilize GLL quadrature on each face.
-While te skew-symmetric formulation allows for the stable coupling of quadrilaterals and triangles using a GLL face quadrature, one may wish to utilize more accurate Gauss face quadratures for triangle-triangle interfaces.
-To accomodate arbitrary combinations of face quadrature, we introduce a mortar-based implementation.
-%
-%We assume that each element 
-%
-%\note{Finish - figure out notation for face interp!}
-
 \subsection{On the accuracy of the skew-symmetric formulation}
@@
 We note that, while it is possible to maintain entropy stability under reduced accuracy quadratures, the accuracy of the resulting schemes depend on the accuracy of the volume and surface quadrature rules.
-In this section, we derive truncation error estimates which take into account the degree of accuracy of the volume and surface quadrature.
-
-It was shown in \cite{chan2017discretely} that, when combined with quadrature-based $L^2$ projection and lifting operators, the decoupled SBP operator (\ref{eq:dsbp}) is a degree $N$ approximation to the derivative.
-This approximation can be interpreted as augmenting a volume approximation of the derivative with boundary correction terms \cite{chan2018efficient}.
+In this section, we analyze the degree of accuracy expected for volume and surface quadrature rules of a given accuracy.
+
+It was shown in \cite{chan2017discretely, chan2018efficient} that the decoupled SBP operator can be interpreted as augmenting a volume approximation of the derivative with boundary correction terms.
 Let $f(\bm{x}),g(\bm{x})$ denote two $L^2$ integrable functions, and let $\bm{f}_N, \bm{g}_N$ denote the vectors of values of $f,g$ at both volume and surface quadrature points.
@@
 where $\bm{u}$ denotes the vector of coefficients for $u$.
-%Expanding terms using (\ref{eq:dsbp}) yields 
-%%\note{fix!  add $E = V_fP_q$}
-%\begin{align*}
-%\bm{M}\bm{u} &= \bm{V}_q^T \diag{\bm{f}_q} \bm{Q}^i \bm{g}_q + \frac{1}{2}\bm{V}_q^T\diag{\bm{f}_q}\bm{E}^T\bm{B}^i \LRp{\bm{g}_f -\bm{E}\bm{g}_q}  \\
-%&\qquad + \frac{1}{2}\bm{V}_f^T\diag{\bm{f}_f}\bm{B}^i \LRp{\bm{g}_f -\bm{E}\bm{g}_q}.
-%\end{align*}
-%where we have used that $\bm{B}^i$ is diagonal and that diagonal matrices commute.
-Let $\bm{v}$ are coefficients of some test function $v\in V^N$, and let $\bm{v}_q = \bm{V}_q\bm{v}$ and $\bm{v}_f = \bm{V}_f \bm{v}$.
-Multiplying by $\bm{v}^T$ on both sides yields that $\bm{v}^T\bm{M}\bm{u}$ is equal to
-%\begin{align*}
-%\LRp{\diag{\bm{f}_q}\bm{v}_q}^T \bm{Q}^i \bm{g}_q + \frac{1}{2}\LRp{\diag{\bm{f}_f}\bm{v}_f+\bm{E}\diag{\bm{f}_q}\bm{v}_q}^T\bm{B}^i \LRp{\bm{g}_f -\bm{E}\bm{g}_q} .
-%\end{align*}
 
@@
 \begin{align}
-\bm{M}\bm{u} =& \frac{1}{2} \LRs{\begin{array}{c}
-\bm{V}_q \\ \bm{V}_f\end{array}}^T \diag{\bm{f}_N}\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}\bm{g}_N \label{eq:dsbpmatrixform}\\
-&+ \frac{1}{2}\bm{V}_f^T\bm{W}_f\diag{\bm{n}^k_i} \LRp{\bm{f}_f \circ \bm{g}_f}.\nonumber
+\bm{M}\bm{u} =& \frac{1}{2}\LRs{\begin{array}{c}
+\bm{V}_q \\ \bm{V}_f\end{array}}^T \diag{\bm{f}_N}\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T + \bm{B}^i_N}\bm{g}_N \label{eq:dsbpmatrixform}
+%&+ \frac{1}{2}\bm{V}_f^T\bm{W}_f\diag{\bm{n}^k_i} \LRp{\bm{f}_f \circ \bm{g}_f}.\nonumber
 \end{align}
 
-When $f(\bm{x}) = 1$, the formulations (\ref{eq:var}), (\ref{eq:skewDN}) can be recast in terms of linear operators $D^i_h, \tilde{D}^i_h$ which capture ``strong'' and ``weak'' formulations of the derivative.
-Let $u(\bm{x})$ be a sufficiently regular function.
-Then, the strong formulation operator $D^i_h: H^1\LRp{D^k}\rightarrow V^N$ can be defined through
-\begin{align*}
-\int_{D^k} D^i_h u  v = \int_{D^k} \pd{\Pi_N u}{x_i}v + \frac{1}{2}\int_{\partial D^k} \LRp{u - \Pi_Nu} v {n}^k_i, \qquad \forall v\in V^N.
-\end{align*}
-Similarly, the weak form operator $\tilde{D}^i_h: H^1\LRp{D^k}\rightarrow V^N$ can be defined such that
-\begin{align*}
-\int_{D^k} \tilde{D}^i_hu v &= \int_{D^k} { -u \pd{{v}}{x_i}} + \frac{1}{2}\int_{\partial D^k}  \LRp{u+\Pi_N u} v {n}^k_i, \qquad \forall v\in V^N.
-\end{align*}
-Then, for $f(\bm{x}) = 1$, the skew-symmetric formulation (\ref{eq:skewDN}) is then equivalent to approximating derivatives using the operator $\frac{1}{2}\LRp{D^i_h - \tilde{D}^i_h}$.
-
-These two operators $D^i_h, \tilde{D}^i_h$ are identical if the integrals are computed using sufficiently accurate volume and surface quadrature rules, e.g.\ if $\bm{Q}^i_N$ satisfies the SBP property (\ref{eq:dsbp}).
-If the SBP property (\ref{eq:dsbp}) does not hold, then $D^i_h, \tilde{D}^i_h$ satisfy different error estimates.
-
-Let $\LRp{\cdot,\cdot}, \LRa{\cdot,\cdot}$ now denote volume and surface $L^2$ inner products over $D^k$, and let $\LRp{\cdot,\cdot}_h, \LRa{\cdot,\cdot}_h$ denote their approximations using quadrature.
-We assume that the volume and surface quadrature rules satisfy Assumption~\ref{ass:quad} for $v = 1, G^k_{ij}$, and that they are additionally exact for polynomials of degree $r_{\rm vol}, r_{\rm surf}$.
-We furthermore assume that quadrature errors are of the form \cite{baker1976effect, brenner2007mathematical}
+The accuracy of the formulations (\ref{eq:esdgSkew}) and (\ref{eq:skewformcurved}) can be understood by analyzing the degree of polynomial exactness of (\ref{eq:dsbpmatrixform}) as an approximation of the derivative.
+Let $u(\bm{x})$ be a polynomial of degree $\leq N$ with coefficients $\bm{u}$, and let $\bm{u}_N = \LRs{\bm{u}_q, \bm{u}_f}^T$ denote the values of $u(\bm{x})$ at volume and surface quadrature points.
+An approximation of $\pd{u}{x_i}$ can be computed by applying (\ref{eq:dsbpmatrixform})
 \begin{align}
-\LRb{\LRp{u,v} - \LRp{u,v}_h} \leq C_{\rm vol} h^{\alpha} \nor{u}_{H^{\alpha}\LRp{D^k}}\nor{v}_{H^{\alpha}\LRp{D^k}},\qquad \alpha = r_{\rm vol}+1, \label{eq:quadacc}\\
-\LRb{\LRa{u,v} - \LRa{u,v}_h} \leq C_{\rm surf} h^{\beta} \nor{uv}_{H^{\beta}\LRp{\partial D^k}}, \qquad \beta = r_{\rm surf}+1.\nonumber
+\pd{u}{x_i}\approx \frac{1}{2}\bm{M}^{-1}\LRs{\begin{array}{c}
+\bm{V}_q \\ \bm{V}_f\end{array}}^T \LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T + \bm{B}^i_N}\bm{u}_N.
+\label{eq:dsbpapprox}
 \end{align}
-Then, we have the following error estimates for $D^i_h, \tilde{D}^i_h$ on affine meshes:
+
+From (\ref{eq:dsbpapprox}), it can be shown that when $\bm{Q}^i$ satisfies a generalized SBP property, the decoupled SBP operator (\ref{eq:dsbp}) produces a degree $N$ approximation to the derivative \cite{chan2017discretely}.
+When $\bm{Q}^i$ does not satisfy a generalized SBP property, we have the following lemma on the accuracy of (\ref{eq:dsbpapprox}):
 \begin{lemma}
-Let $D^k$ be affine elements, let $u(\bm{x})$ be a sufficiently regular function, and suppose $r_{\rm vol} \geq 2N$ and $r_{\rm surf} \geq N$.
-Then, if the SBP property (\ref{eq:dsbp}) does not hold,
-\begin{align*}
-\nor{\pd{u}{x_i} - \tilde{D}^i_h u}_{L^2\LRp{\Omega}} \leq C_1 h^{ \gamma}, \qquad \nor{\pd{u}{x_i} - D^i_h u}_{L^2\LRp{\Omega}} \leq C_2 h^{\max\LRp{N, \gamma}},
-\end{align*}
-where $\gamma = \min\LRc{\LRp{N+1}, \LRp{r_{\rm vol} - N + 1}, \LRp{r_{\rm surf} - N}}$, and $C_1,C_2$ are independent of $h$.
-If the SBP property (\ref{eq:dsbp}) does hold, then both errors are bounded by $O\LRp{h^{\max\LRp{N,\gamma}}}$.
-\label{lemma:quadacc}
+Suppose that the volume quadrature is exact for degree $M+N-1$ polynomials on simplices, or for polynomials in $Q^{M+N-1,M+N,M+N}$ on tensor product elements.
+Furthermore, assume that the surface quadrature is exact for degree $M+N$ polynomials.
+Then, the skew-symmetric approximation of the $x$-derivative (\ref{eq:dsbpapprox}) is exact for all polynomials of degree $M$.
+\label{lemma:dsbpapprox}
 \end{lemma}
 \begin{proof}
-The error can be decomposed as
-\[
-\pd{u}{x_i} - \tilde{D}^i_h u = \LRp{\pd{u}{x_i} - \Pi_N\pd{u}{x_i}} + \LRp{\Pi_N\pd{u}{x_i} - \tilde{D}^i_h u}.
-\]
-Standard finite element approximation estimates \cite{brenner2007mathematical} give that 
-\[
-\nor{\pd{u}{x_i} - \Pi_N\pd{u}{x_i}}_{L^2\LRp{\Omega}} \lesssim h^{N+1} \sqrt{\sum_k \nor{u}_{H^{N+1}\LRp{D^k}}}.
-\]
-The latter term can be estimated using (\ref{eq:quadacc}).  We will first consider the error for $\tilde{D}^i_h$.  Let $\delta = \Pi_N\pd{u}{x} - \tilde{D}^i_h$.  Then, on each element $D^k$, 
-\[
-\nor{\Pi_N\pd{u}{x_i} - \tilde{D}^i_hu}_{L^2\LRp{D^k}}^2 = \LRb{\LRp{\Pi_N\pd{u}{x_i} - \tilde{D}^i_h,\delta} } = \LRb{\LRp{\pd{u}{x_i}, \delta} - \LRp{\tilde{D}^i_h,\delta} },
-\]
-where we have used the definition of $\Pi_N$ (\ref{eq:l2proj}) and the fact that $\delta \in V^N$.  Integrating $\LRp{\pd{u}{x_i}, \delta}$ by parts and using the definition of $\tilde{D}^i_h$ gives
-\begin{gather*}
-\nor{\Pi_N\pd{u}{x_i} - \tilde{D}^i_hu}_{L^2\LRp{D^k}}^2 \leq  \LRb{\LRa{u n^k_i, v}- \LRa{un^k_i, v}_h} + \LRb{\LRp{u, \pd{v}{x_i}} - \LRp{u, \pd{v}{x_i}}_h}\\
-\lesssim \LRp{h^{\beta}\nor{uv}_{H^{\beta}\LRp{\partial D^k}} +  h^{\alpha}\nor{u}_{H^{\alpha}\LRp{D^k}} \nor{\pd{v}{x_i}}_{H^{\alpha}\LRp{D^k}}}.
-\end{gather*}
-The terms $\nor{uv}_{H^{\beta}\LRp{\partial D^k}}, \nor{\pd{v}{x_i}}_{H^{\alpha}\LRp{D^k}}$ can be bounded using inverse and inverse trace inequalities \cite{brenner2007mathematical, chan2018discretely} and the fact that $v \in V^N$ and $\beta \geq (N+1)$
-\begin{align*}
-\nor{uv}_{H^{\beta}\LRp{\partial D^k}} &\lesssim h^{-1/2} \nor{uv}_{H^{\beta}\LRp{D^k}} \lesssim h^{-1/2} \nor{u}_{H^{\beta}\LRp{D^k}}\nor{v}_{H^{\beta}\LRp{D^k}}\\
-&\lesssim h^{-N-1/2} \nor{u}_{H^{\beta}\LRp{D^k}}\nor{v}_{L^2\LRp{D^k}} \\
-\nor{\pd{v}{x_i}}_{H^{\alpha}\LRp{D^k}} &\lesssim h^{-(N-1)}\nor{v}_{L^2\LRp{D^k}}.
-\end{align*}
-
-Noting that $ \nor{v}_{L^2\LRp{D^k}} = \nor{\Pi_N\pd{u}{x_i} - \tilde{D}^i_hu}_{L^2\LRp{D^k}}$, dividing through by $\nor{v}_{L^2\LRp{D^k}}$ gives
-\[
-\nor{\Pi_N\pd{u}{x_i} - \tilde{D}^i_hu}_{L^2\LRp{D^k}} \lesssim h^{\beta - N} \nor{u}_{H^{\beta}\LRp{D^k}} + h^{\alpha - N + 1} \nor{u}_{H^{\alpha}\LRp{D^k}}.
-\]
- \note{Finish - how to lose extra factor of $1/h$ in trace estimate?}
+Suppose $u \in P^M$.
+Let $\bm{u}$ denote the polynomial coefficients of $u$, and let $\bm{e}$ denote the difference between $\bm{D}^i\bm{u}$ (the exact coefficients of $\pd{u}{x_i}$) and the approximation (\ref{eq:dsbpapprox})
+\[
+\bm{e} = \bm{D}^i\bm{u} - \frac{1}{2}\bm{M}^{-1}\LRs{\begin{array}{c}
+\bm{V}_q \\ \bm{V}_f\end{array}}^T\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T + \bm{B}^i_N}\bm{u}_N,
+\]
+where $\bm{e}$ is a polynomial of degree $N$.  
+If Assumption~\ref{ass:quad} is satisfied and $u(\bm{x})$ is polynomial, then the values of $u$ at quadrature points are $\bm{u}_q = \bm{V}_q\bm{u}$ and $\bm{u} = \bm{P}_q\bm{u}_q$.
+This implies that $\bm{u}_f = \bm{V}_f \bm{P}_q\bm{u}_q =  \bm{E}\bm{u}_q$.
+Expanding the latter term yields
+\begin{align*}
+&\frac{1}{2}\bm{M}^{-1}\LRs{\begin{array}{c}
+\bm{V}_q \\ \bm{V}_f\end{array}}^T\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T + \bm{B}^i_N}\bm{u}_N \\
+&= \frac{1}{2}\bm{M}^{-1}\LRs{\begin{array}{c}
+\bm{V}_q \\ \bm{V}_f\end{array}}^T\begin{bmatrix}
+\bm{Q}^i - \LRp{\bm{Q}^i}^T & \bm{E}^T\bm{B}^i\\
+-\bm{B}^i\bm{E} & \bm{B}^i
+\end{bmatrix}\begin{bmatrix}
+\bm{u}_q\\
+\bm{u}_f
+\end{bmatrix}\\
+&= \frac{1}{2}\bm{M}^{-1}\LRs{\begin{array}{c}
+\bm{V}_q \\ \bm{V}_f\end{array}}^T\begin{bmatrix}
+\bm{Q}^i\bm{u}_q + \LRp{\bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T}\bm{u}_q\\
+\bm{B}^i\LRp{\bm{u}_f - \bm{E}\bm{u}_q}
+\end{bmatrix} 
+\\
+&=\frac{1}{2}\bm{M}^{-1}
+\bm{V}_q^T\LRp{\bm{Q}^i\bm{u}_q + \LRp{\bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T}\bm{u}_q}.
+%&=\frac{1}{2} \bm{D}^i\bm{u} + \frac{1}{2}\bm{M}^{-1}
+%\bm{V}_q^T{ \LRp{\bm{E}^T\bm{B}^i\bm{E} - \bm{Q}^i}\bm{u}_q}
+\end{align*}
+Since $\bm{Q}^i = \bm{W}\bm{V}_q\bm{D}^i\bm{P}_q$ and $\bm{M} = \bm{V}_q^T\bm{W}\bm{V}_q$, we have that $\bm{M}^{-1}\bm{V}_q^T\bm{Q}^i\bm{u}_q = \bm{D}^i\bm{u}$.
+This simplifies the expression for error to
+\begin{align}
+\bm{e}^T\bm{M}\bm{e} &= \frac{1}{2}
+\bm{e}_q^T\LRp{-\bm{Q}^i\bm{u}_q + \LRp{\bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T}\bm{u}_q} \label{eq:skewerr}\\
+&= -\frac{1}{2}
+\bm{e}_q^T\bm{Q}^i\bm{u}_q + \frac{1}{2}\bm{e}_q^T\LRp{\bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T}\bm{u}_q, \qquad \bm{e}_q = \bm{V}_q\bm{e}. \nonumber
+\end{align}
+The remainder of the proof is similar to the proof of Lemma~\ref{lemma:vsbp}.  Since $u \in P^M$ and $e \in P^N$, by exactness of the quadrature rules, 
+\[
+\bm{e}_q^T\LRp{\bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T}\bm{u}_q = \int_{\partial \hat{D}} u e n_i - \int_{\hat{D}} u \pd{e}{x_i}
+= \int_{\hat{D}} \pd{u}{x_i} e = \bm{e}_q^T\bm{Q}^i\bm{u}_q.
+\]
+Combining this with (\ref{eq:skewerr}) implies that $\bm{e}^T\bm{M}\bm{e} = 0$, and since $\bm{M}$ is symmetric positive definite, $\bm{e} = 0$.  
 \end{proof}
 
-\begin{remark}
-Numerical experiments indicate that the assumptions of Lemma~\ref{lemma:quadacc} are sufficient, but not necessary.
-For specific combinations of volume and surface quadrature, we observe that the errors in Lemma~\ref{lemma:quadacc} converge faster than expected given the degree of exactness of each rule.
-These special cases will be investigated in more detail in future work.
-\end{remark}
+Lemma~\ref{lemma:dsbpapprox} suggests that, when a generalized SBP property does not hold, the use of under-integrated quadratures results in a loss of one or more orders of accuracy.
+For example, using $(N+1)$ point GLL rules (which are exact for only polynomials of degree $2N-1$) for either volume or surface quadratures should result in a loss of one order of accuracy compared to the use of $(N+1)$-point Gauss rules (which are exact for polynomials of degree $2N$).
+This is indeed observed in numerical experiments.
 
@@
 
-In this section, we present two-dimensional experiments which verify the theoretical results presented.
-We begin by investigating the stability and accuracy of the skew-symmetric formulation on triangular and quadrilateral meshes, and conclude with two-dimensional experiments on a hybrid mesh containing mixed quadrilateral and triangular elements.
-%on hybrid meshes consisting of quadrilaterals and triangles, as well as on non-conforming meshes of quadrilateral elements.
-
-\subsection{Quadrature accuracy and rates of convergence}
-
-\subsection{Quadrature accuracy and timestep restrictions}
-
-\subsection{Triangular meshes}
-
-\note{Affine and curved triangles with GLL surface quadrature (under-integrated).}
-
-\subsection{Quadrilateral meshes}
-
-\note{ GLL quadrilaterals with Gauss surface quadrature.  Explain that for GLL quadratures, the decoupled SBP property doesn't hold when Gauss points are used.  }
+In this section, we present two-dimensional experiments which verify the theoretical results presented and qualify the accuracy of the proposed methods.
+We begin by investigating the maximum stable timestep, stability, and accuracy of the skew-symmetric formulation on triangular and quadrilateral meshes, and conclude with two-dimensional experiments on a hybrid mesh containing mixed quadrilateral and triangular elements.
+
+We consider numerical solutions of the 2D compressible Euler equations
+\begin{align*}
+\pd{\rho}{t} + \pd{\LRp{\rho u}}{x_1} + \pd{\LRp{\rho v}}{x_2} &= 0,\\
+\pd{\rho u}{t} + \pd{\LRp{\rho u^2 + p }}{x_1} + \pd{\LRp{\rho uv}}{x_2} &= 0,\nonumber\\
+\pd{\rho v}{t} + \pd{\LRp{\rho uv}}{x_1} + \pd{\LRp{\rho v^2 + p }}{x_2} &= 0,\nonumber\\
+\pd{E}{t} + \pd{\LRp{u(E+p)}}{x_1} + \pd{\LRp{v(E+p)}}{x_2}&= 0,\nonumber
+\end{align*}
+where we have introduced the pressure is $p = (\gamma-1)\LRp{E - \frac{1}{2}\rho (u^2+v^2)}$ and the specific internal energy $\rho e = E - \frac{1}{2}\rho (u^2+v^2)$.
+
+We seek entropy stability with respect to the entropy for the compressible Navier-Stokes equations \cite{hughes1986new}
+\begin{equation*}
+U(\bm{u}) = -\frac{\rho s}{\gamma-1},
+\label{eq:entropy2d}
+\end{equation*}
+where $s = \log\LRp{\frac{p}{\rho^\gamma}}$ denotes the specific entropy. The mappings between conservative and entropy variables in two dimensions are given by
+\begin{align*}
+v_1 &= \frac{\rho e (\gamma + 1 - s) - E}{\rho e}, \qquad v_2 = \frac{\rho u}{\rho e}, \qquad v_3 = \frac{\rho v}{\rho e}, \qquad v_4 = -\frac{\rho}{\rho e}\\
+\rho &= -(\rho e) v_4, \qquad \rho u = (\rho e) v_2, \qquad \rho v = (\rho e) v_3, \qquad E = (\rho e)\LRp{1 - \frac{{v_2^2+v_3^2}}{2 v_4}},
+\end{align*}
+where $\rho e$ and $s$ can be expressed in terms of the entropy variables as
+\begin{equation*}
+\rho e = \LRp{\frac{(\gamma-1)}{\LRp{-v_4}^{\gamma}}}^{1/(\gamma-1)}e^{\frac{-s}{\gamma-1}}, \qquad s = \gamma - v_1 + \frac{{v_2^2+v_3^2}}{2v_4}.
+\end{equation*}
+
+There exist several choices for entropy conservative fluxes \cite{ismail2009affordable, ranocha2018comparison,chandrashekar2013kinetic}.  For the following experiments, we utilize the
+the entropy conservative numerical fluxes given by Chandrashekar in \cite{chandrashekar2013kinetic}
+\begin{align*}
+&f^1_{1,S}(\bm{u}_L,\bm{u}_R) = \avg{\rho}^{\log} \avg{u},& &f^1_{2,S}(\bm{u}_L,\bm{u}_R) = \avg{\rho}^{\log} \avg{v},&\\
+&f^2_{1,S}(\bm{u}_L,\bm{u}_R) = f^1_{1,S} \avg{u} + p_{\rm avg},&  &f^2_{2,S}(\bm{u}_L,\bm{u}_R) = f^1_{2,S} \avg{u},&\nonumber\\
+&f^3_{1,S}(\bm{u}_L,\bm{u}_R) = f^2_{2,S},& &f^3_{2,S}(\bm{u}_L,\bm{u}_R) = f^1_{2,S} \avg{v} + p_{\rm avg},&\nonumber\\
+&f^4_{1,S}(\bm{u}_L,\bm{u}_R) = \LRp{E_{\rm avg} + p_{\rm avg}}\avg{u},& &f^4_{2,S}(\bm{u}_L,\bm{u}_R) = \LRp{E_{\rm avg} + p_{\rm avg} }\avg{v},& \nonumber
+\end{align*}
+where the quantities $p_{\rm avg}, E_{\rm avg},  \nor{\bm{u}}^2_{\rm avg}$ are defined as
+\begin{gather*}
+p_{\rm avg} = \frac{\avg{\rho}}{2\avg{\beta}}, \qquad E_{\rm avg} = \frac{\avg{\rho}^{\log}}{2\avg{\beta}^{\log}\LRp{\gamma -1}}   + \frac{\nor{\bm{u}}^2_{\rm avg}}{2}, \\
+ \nor{\bm{u}}^2_{\rm avg} = 2(\avg{u}^2 + \avg{v}^2) - \LRp{\avg{u^2} +\avg{v^2}} \nonumber.  
+\end{gather*}
+
+For all convergence experiments, we compare the numerical solution to analytic solution for the isentropic vortex problem \cite{shu1998essentially} 
+\begin{align*}
+\rho(\bm{x},t) &= \LRp{1 - \frac{\frac{1}{2}(\gamma-1)(\beta e^{1-r(\bm{x},t)^2})^2}{8\gamma \pi^2}}^{\frac{1}{\gamma-1}}, \qquad p = \rho^{\gamma},\\
+u(\bm{x},t) &= 1 - \frac{\beta}{2\pi} e^{1-r(\bm{x},t)^2}(x_2-c_2), \qquad v(\bm{x},t) = \frac{\beta}{2\pi} e^{1-r(\bm{x},t)^2}(x_2-c_2).\nonumber
+\end{align*}
+Here, $u, v$ are the $x_1$ and $x_2$ velocity and $r(\bm{x},t) = \sqrt{(x_1-c_1-t)^2 + (x_2-c_2)^2}$.  The following experiments utilize $c_1 = 5, c_2 = 0$ and $\beta = 5$.
+
+\subsection{Choices of volume and surface quadrature considered}
+\label{sec:opts}
+
+On quadrilaterals, we consider volume quadratures which are tensor products of one-dimensional quadrature rules, while for triangles we utilize volume quadratures from \cite{xiao2010quadrature}.
+Surface quadratures are constructed face-by-face, and we refer to surface quadrature rules by the specific quadrature used over each face.
+
+We consider three choices of volume and surface quadrature on quadrilaterals: 
+\begin{enumerate}
+\item $(N+1)$ point Gauss quadrature, $(N+1)$ point Gauss surface quadrature,
+\item $(N+1)$ point GLL volume quadrature, $(N+1)$ point Gauss surface quadrature,
+\item $(N+1)$ point GLL volume quadrature, $(N+1)$ point GLL surface quadrature.
+\end{enumerate}
+On triangles, we consider two cases:
+\begin{enumerate}
+\item degree $2N$ volume quadrature, $(N+1)$ point Gauss surface quadrature,
+\item degree $2N$ volume quadrature, $(N+1)$ point GLL surface quadrature.
+\end{enumerate}
+
+These choices can be combined to provide three different options on two-dimensional hybrid meshes of quadrilateral and triangular elements, which are motivated by balancing computational efficiency and accuracy:
+\begin{enumerate}
+\item Option 1: $(N+1)$ point GLL volume quadrature on quadrilaterals and $(N+1)$ point GLL surface quadrature on quadrilaterals and triangles.\label{opt:1}
+\item Option 2: $(N+1)$ point GLL volume quadrature on quadrilaterals and $(N+1)$ point Gauss surface quadrature on quadrilaterals and triangles.\label{opt:2}
+\item Option 3: $(N+1)$ point Gauss volume quadrature on quadrilaterals and $(N+1)$ point Gauss surface quadrature on quadrilaterals and triangles.\label{opt:3}
+\end{enumerate}
+All three options assume a triangular volume quadrature which is exact for all polynomials of degree $2N$ or less.
+
+All options result in similar computational costs on on triangles, but slight differences in computational cost and complexity on quadrilaterals.
+On quadrilaterals, Option~\ref{opt:1} is the most computationally efficient option, as the formulation (\ref{thm:skewformcurved}) reduces to a standard entropy stable DG-SEM \cite{gassner2016split} or spectral collocation method \cite{carpenter2014entropy}.
+Option~\ref{opt:2} is slightly more expensive, and requires interpolating the solution from GLL to Gauss nodes on the boundary, as well as computing two-point flux interactions between boundary GLL and Gauss nodes.
+Option 3 is most involved, and results in a Gauss collocation method on quadrilaterals, as described in \cite{chan2018efficient}, and requires both interpolation and two-point flux interactions between lines of volume quadrature nodes and boundary nodes.
+However, Option~\ref{opt:3} provides some advantages over Options~\ref{opt:1} and \ref{opt:2} in terms of accuracy and stable timestep restrictions.
+
+\subsection{Choice of quadrature and timestep restrictions}
+
+We first investigate how the maximum stable timestep depends on specific choices of volume and surface quadrature.
+The dependence of timestep on quadrature has been documented for tensor product elements in \cite{gassner2011comparison}, where they showed that for a high order Taylor method in time, the maximum stable timestep under  $(N+1)$-point GLL volume and surface quadratures is roughly twice as large as the maximum stable timestep when volume/surface integrals are approximated using $(N+1)$ point Gauss quadratures.
+
+This discrepancy can be understood in terms of constants in finite element inverse Markov and trace inequalities.
+It was shown in \cite{chan2015gpu, chan2018multi} that, for linear problems, the maximum stable time-step scales inversely with the order-dependent constants $C_M, C_T$, where
+\[
+\int_{\hat{D}} \LRb{\Grad u}^2 \leq C_M \int_{\hat{D}} u^2, \qquad \int_{\partial \hat{D}} u^2 \leq C_T  \int_{\hat{D}} u^2, \qquad \forall u \in V^N.
+\]
+Here, the integrals over $\hat{D}, \partial \hat{D}$ are computed using the same volume and surface quadrature rules used for computations.
+These constants can be used to bound surface integrals which appear in DG formulations, which can in turn be used to construct bounds on the spectral radius of DG discretization matrices.
+These result in the following estimate of the maximum stable timestep
+\[
+dt \leq \frac{h}{\LRb{\lambda}_{\max} C_T}
+\]
+
+While the Markov constant $C_M$ is not affected by the choices of quadrature described in Section~\ref{sec:opts}, the constant $C_T$ depends on the choice of surface quadrature on triangular elements.
+It is known that $L^2$ norm computed using GLL quadrature is weaker than the full $L^2$ norm \cite{quarteroni1994introduction,canuto2007spectral}.
+Specifically,
+\[
+\LRp{2+\frac{1}{N}}^{-1/2}\int_{\partial \hat{D}, {\rm GLL}} u^2 \leq \int_{\partial \hat{D}} u^2 \qquad \forall u \in P^{N}\LRp{[-1,1]}, 
+\]
+where the former integral is under-integrated using GLL quadrature.
+In other words, when using GLL surface quadrature, the trace constant is not $C_T$ but $C_T \LRp{2+1/N}^{d/2}$.
+As a result, the triangular elements in Option~\ref{opt:1} (which use GLL surface quadrature) will have a smaller maximum stable timestep than the triangular elements for Options~\ref{opt:2} or \ref{opt:3} (though the timestep restriction for hexahedral elements in Option~\ref{opt:1} remains identical to that of a standard DG-SEM scheme).
+
+For Option~\ref{opt:2}, the timestep restriction
+
+\begin{table}
+\centering
+\subfloat[Quadrilateral elements]{
+ \begin{tabular}{|c || c| c| c| c| c| c| c|} 
+ \hline
+ N & 1 & 2& 3 & 4 & 5 & 6& 7\\
+ \hline\hline
+GLL/GLL &&&&&&&\\
+ \hline
+Gauss/Gauss &&&&&&&\\
+ \hline
+Gauss/GLL &&&&&&&\\
+ \hline
+GLL/Gauss  &&&&&&&\\
+ \hline
+ \end{tabular}
+ }
+ \caption{Markov and trace constants for different configurations of triangular and quadrilateral elements.}
+\end{table}
+
+\note{Add experiments}
+
+%\subsection{Quadrature accuracy and rates of convergence}
+%
+%\subsection{Triangular meshes}
+%
+%\note{Affine and curved triangles with GLL surface quadrature (under-integrated).}
+%
+%\subsection{Quadrilateral meshes}
+%
+%\note{ GLL quadrilaterals with Gauss surface quadrature.  Explain that for GLL quadratures, the decoupled SBP property doesn't hold when Gauss points are used.  }
 
@@
 Thus, we restrict ourselves to straight-sided meshes in this case, and will investigate the effect of curvilinear mappings on hybrid meshes in future work.
+
+\begin{figure}
+\centering
+%\subfloat{\raisebox{2.85em}{
+\includegraphics[width=.45\textwidth]{figs/hybrid_mesh.png}
+%}}
+\caption{Coarsest hybrid mesh used for convergence studies. }
+\end{figure}
+
+\begin{figure}
+\centering
+%\subfloat{\raisebox{2.85em}{\includegraphics[width=.45\textwidth]{figs/hybrid_mesh.png}}}
+%\hspace{.25em}
+%\subfloat[Option~\ref{opt:1}]{
+\begin{tikzpicture}
+\begin{loglogaxis}[
+    width=.7\textwidth,
+    xlabel={Mesh size $h$},
+    ylabel={$L^2$ errors}, 
+    xmax=.7,
+    ymin=1e-5, ymax=5,
+    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
+    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
+    legend entries={Option~\ref{opt:1},Option~\ref{opt:2},Option~\ref{opt:3}}    
+]
+\pgfplotsset{
+cycle list={{blue, mark=*}, {red, dashed ,mark=triangle*}, {black, mark=square*}}
+}
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,2.29471)(0.166667,1.55868)(0.0833333,0.746603)(0.0416667,0.274017)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,2.39742)(0.166667,1.63628)(0.0833333,0.885414)(0.0416667,0.435973)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,1.74989)(0.166667,0.687578)(0.0833333,0.169682)(0.0416667,0.0350721)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,1.06222)(0.166667,0.267115)(0.0833333,0.0395222)(0.0416667,0.00569986)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,1.11474)(0.166667,0.265955)(0.0833333,0.0382799)(0.0416667,0.00492061)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,0.51272)(0.166667,0.0625001)(0.0833333,0.00835287)(0.0416667,0.00120101)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,0.321422)(0.166667,0.037836)(0.0833333,0.00336728)(0.0416667,0.000268691)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,0.364547)(0.166667,0.0393928)(0.0833333,0.00357584)(0.0416667,0.000392883)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,0.133053)(0.166667,0.0122055)(0.0833333,0.000730739)(0.0416667,4.37208e-05)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,0.10556)(0.166667,0.00654941)(0.0833333,0.000267376)(0.0416667,NaN)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,0.0955729)(0.166667,0.00649092)(0.0833333,0.000236043)(0.0416667,NaN)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,0.0525951)(0.166667,0.00229986)(0.0833333,6.74854e-05)(0.0416667,NaN)};
+\node at (axis cs:.45, 2.25) {$N = 1$};
+\node at (axis cs:.45,.8) {$N = 2$};
+\node at (axis cs:.45,.225) {$N = 3$};
+\node at (axis cs:.45,.075) {$N = 4$};
+\end{loglogaxis}
+\end{tikzpicture}
+
+%}
+%\subfloat[Option~\ref{opt:2}]{
+%\begin{tikzpicture}
+%\begin{loglogaxis}[
+%    width=.49\textwidth,
+%    xlabel={Mesh size $h$},
+%    ylabel={$L^2$ errors}, 
+%%    xmax=1,
+%    ymin=1e-5, ymax=5,
+%    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
+%    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
+%    legend entries={$N=1$, $N=2$, $N=3$, $N=4$}    
+%]
+%\pgfplotsset{
+%cycle list name=color
+%}
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(0.333333,2.39742)(0.166667,1.63628)(0.0833333,0.885414)(0.0416667,0.435973)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(0.333333,1.11474)(0.166667,0.265955)(0.0833333,0.0382799)(0.0416667,0.00492061)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(0.333333,0.364547)(0.166667,0.0393928)(0.0833333,0.00357584)(0.0416667,0.000392883)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(0.333333,0.0955729)(0.166667,0.00649092)(0.0833333,0.000236043)(0.0416667,NaN)};
+%\end{loglogaxis}
+%\end{tikzpicture}
+%}
+%
+%\subfloat[Option~\ref{opt:3}]{
+%\begin{tikzpicture}
+%\begin{loglogaxis}[
+%    width=.49\textwidth,
+%    xlabel={Mesh size $h$},
+%    ylabel={$L^2$ errors}, 
+%%    xmax=1,
+%    ymin=1e-5, ymax=5,
+%    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
+%    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
+%    legend entries={$N=1$, $N=2$, $N=3$, $N=4$}    
+%]
+%\pgfplotsset{
+%cycle list name=color
+%}
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(0.333333,1.74989)(0.166667,0.687578)(0.0833333,0.169682)(0.0416667,0.0350721)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(0.333333,0.51272)(0.166667,0.0625001)(0.0833333,0.00835287)(0.0416667,0.00120101)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(0.333333,0.133053)(0.166667,0.0122055)(0.0833333,0.000730739)(0.0416667,4.37208e-05)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(0.333333,0.0525951)(0.166667,0.00229986)(0.0833333,6.74854e-05)(0.0416667,NaN)};
+%\end{loglogaxis}
+%\end{tikzpicture}
+%}
+\caption{Convergence of $L^2$ errors for the isentropic vortex solution for Option~\ref{opt:1}, Option~\ref{opt:2}, and Option~\ref{opt:3} for $N = 1,\ldots, 4$.}
+\end{figure}
 
@@
 
-\section{Acknowledgments}
+%\section{Acknowledgments} The author thanks David C.\ Del Rey Fernandez for
 

### 2018-10-19 docs/ESDG_mortar/ext_decoupled_sbp.tex

--- 2018-07-19 docs/ESDG_GQ/ext_decoupled_sbp.tex
@@
 \maketitle
-
-%\begin{frontmatter}
-%\title{Skew-symmetric formulations for entropy stable discontinuous Galerkin methods on non-conforming meshes}
-%
-%\author[rice]{Jesse Chan\corref{cor1}}
-%\ead{Jesse.Chan@caam.rice.edu}
-%\address[rice]{Department of Computational and Applied Mathematics, Rice University, 6100 Main St, Houston, TX, 77005}
-%
-%\begin{abstract}
-%Entropy stable high order methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.  
-%\end{abstract}
-%\end{frontmatter}
-
-%\tableofcontents
 
@@
 
+\section{Multiple surface quadratures}
+
+We first introduce notation which is helpful for operations involving surface traces.  Let the matrix $\bm{R}$ denote the matrix which maps coefficients of a function to coefficients of its trace.
+
 Let ${\bm{V}_f}\tilde{P}_f$ denote the matrix which projects to the trace space using split-side quadrature and evaluates at full-side quadrature points.

### 2018-10-19 docs/ESDG_mortar/main.tex

@@
 
-\paragraph{Hybrid meshes:} It was shown in \cite{chan2018skew} that an entropy stable
+\paragraph{Hybrid meshes:} entropy stable DG methods on hybrid meshes were introduced in \cite{chan2018skew} using a skew-symmetric formulation.
+The resulting methods are stable for more arbitrary choices of surface quadrature, in particular when an SBP property may not hold.
 
@@
 We can now define operators which map between surface and mortar quadrature points.
-Let $\bm{E}_f$ denote the map from surface to mortar points, and let $\tilde{\bm{E}}_f $ denote the map from mortar to surface points.
+Let $\bm{E}_m$ denote the map from surface to mortar points, and let $\tilde{\bm{E}}_m $ denote the map from mortar to surface points.
 Both operators are defined through an $L^2$ projection to the trace space and interpolation to appropriate points
 \[
-\bm{E}_f = \bm{T}_m \bm{P}_f, \qquad \tilde{\bm{E}}_f = \bm{T}_f \bm{P}_m.
+\bm{E}_m = \bm{T}_m \bm{P}_f, \qquad \tilde{\bm{E}}_m = \bm{T}_f \bm{P}_m.
 \]  
@@
 \bm{Q}_i-\bm{Q}_i^T & \bm{E}^T\bm{B}_i &\\
--\bm{B}_i\bm{E} &  & \bm{B}_i\tilde{\bm{E}}_f \\
-& -\tilde{\bm{B}}_i{\bm{E}}_f & 
+-\bm{B}_i\bm{E} &  & \bm{B}_i\tilde{\bm{E}}_m \\
+& -\tilde{\bm{B}}_i{\bm{E}}_m & 
 \end{bmatrix} \circ \bm{F}_S}\bm{1} + \bm{V}_m^T\tilde{\bm{B}}_i\bm{f}_i^* = 0.  
@@
 \begin{align*}
-\bm{B}_i\tilde{\bm{E}}_f &= \diag{\bm{W}_f} \bm{T}_f \bm{P}_m = \diag{\hat{\bm{n}}} \bm{W}_f \bm{T}_f \bm{M}_m^{-1}\bm{T}_m\bm{W}_m\\
-&=  \bm{W}_f \bm{T}_f \bm{M}_m^{-1}\bm{T}^T_m\bm{W}_m\diag{\hat{\bm{n}}_i} = \bm{P}_f^T\bm{T}^T_m \tilde{\bm{B}}_i = \LRp{\tilde{\bm{B}}_i{\bm{E}}_f}^T
+\bm{B}_i\tilde{\bm{E}}_m &= \diag{\bm{W}_f} \bm{T}_f \bm{P}_m = \diag{\hat{\bm{n}}} \bm{W}_f \bm{T}_f \bm{M}_m^{-1}\bm{T}_m\bm{W}_m\\
+&=  \bm{W}_f \bm{T}_f \bm{M}_m^{-1}\bm{T}^T_m\bm{W}_m\diag{\hat{\bm{n}}_i} = \bm{P}_f^T\bm{T}^T_m \tilde{\bm{B}}_i = \LRp{\tilde{\bm{B}}_i{\bm{E}}_m}^T
 \end{align*}
@@
 While the formulation presented above is convenient for analysis, it is computationally expensive to implement.
-However, using properties of the mortar matrices $\bm{E}_f, \tilde{\bm{E}}_f$, we can rewrite the above formulation in a way which reflects a traditional mortar-based finite element implementation.
+However, using properties of the mortar matrices $\bm{E}_m, \tilde{\bm{E}}_m$, we can rewrite the above formulation in a way which reflects a traditional mortar-based finite element implementation.
 In other words, we wish to implement (\ref{eq:mortar}) such that the only modification from the implementations in \cite{chan2017discretely} is a pre-processing step on mortar faces.
 
-We first note that, since $\bm{E}_f$ maps from surface nodes to mortar nodes, $\bm{V}_m = \bm{E}_f \bm{V}_f$.  Thus, we can rewrite the numerical flux contribution as
+We first note that, since $\bm{E}_m$ maps from surface nodes to mortar nodes, $\bm{V}_m = \bm{E}_m \bm{V}_f$.  Thus, we can rewrite the numerical flux contribution as
 \begin{align*}
- \bm{V}_m^T\tilde{\bm{B}}_i\bm{f}_i^* &= \bm{V}_f^T \bm{E}_f^T \bm{W}_m\diag{\hat{\bm{n}}_i} \bm{f}_i^* = \bm{V}_f^T  \bm{P}_f^T\bm{T}_m^T\bm{W}_m\diag{\hat{\bm{n}}_i} \bm{f}_i^*\\
- &= \bm{V}_f^T \bm{W}_f \diag{\hat{\bm{n}}_i} \bm{T}_m\bm{M}_m^{-1}\bm{T}_m^T\bm{W}_m \bm{f}_i^* =  \bm{V}_f^T \bm{B}_i \tilde{\bm{E}}_f\bm{f}_i^*.
+ \bm{V}_m^T\tilde{\bm{B}}_i\bm{f}_i^* &= \bm{V}_f^T \bm{E}_m^T \bm{W}_m\diag{\hat{\bm{n}}_i} \bm{f}_i^* = \bm{V}_f^T  \bm{P}_f^T\bm{T}_m^T\bm{W}_m\diag{\hat{\bm{n}}_i} \bm{f}_i^*\\
+ &= \bm{V}_f^T \bm{W}_f \diag{\hat{\bm{n}}_i} \bm{T}_m\bm{M}_m^{-1}\bm{T}_m^T\bm{W}_m \bm{f}_i^* =  \bm{V}_f^T \bm{B}_i \tilde{\bm{E}}_m\bm{f}_i^*.
 \end{align*}
 
-We can similarly reformulate the remaining mortar contributions within (\ref{eq:mortar}).  Multiplying out
+We can similarly reformulate the remaining mortar contributions within (\ref{eq:mortar}).  We decompose $\bm{F}_S$ into interactions between volume nodes, surface nodes, and mortar nodes
+\[
+\bm{F}_S = \begin{bmatrix}
+\bm{F}_S^{vv} & \bm{F}_S^{vs} & \bm{F}_S^{vm}\\
+\bm{F}_S^{sv} & \bm{F}_S^{ss} & \bm{F}_S^{sm}\\
+\bm{F}_S^{mv} & \bm{F}_S^{ms} & \bm{F}_S^{mm}
+\end{bmatrix}
+\]
+Then, the on-element contributions to (\ref{eq:mortar}) can be expanded as follows
+\begin{align*}
+&\begin{bmatrix} \bm{V}_q \\ \bm{V}_f \\ \bm{V}_m \end{bmatrix}^T
+\LRp{\begin{bmatrix}
+\bm{Q}_i-\bm{Q}_i^T & \bm{E}^T\bm{B}_i &\\
+-\bm{B}_i\bm{E} &  & \bm{B}_i\tilde{\bm{E}}_m \\
+& -\tilde{\bm{B}}_i{\bm{E}}_m & 
+\end{bmatrix} \circ \bm{F}_S}\bm{1} \\
+&= 
+\begin{bmatrix} \bm{V}_q \\ \bm{V}_f \end{bmatrix}^T
+\LRp{\begin{bmatrix}
+\bm{Q}_i-\bm{Q}_i^T & \bm{E}^T\bm{B}_i\\
+-\bm{B}_i\bm{E} & \\
+\end{bmatrix} \circ \bm{F}_S}\bm{1} \\
+&+ \bm{V}_f^T \LRp{\bm{B}_i\tilde{\bm{E}}_m\circ \bm{F}_S^{sm}}\bm{1} - \bm{V}_m^T \LRp{ \tilde{\bm{B}}_i\bm{E}_m \circ \bm{F}_S^{ms}}\bm{1}
+\end{align*}
+Since multiplication by diagonal matrices $\bm{B}_i, \tilde{\bm{B}}_i$ is associative under the Hadamard product, the latter terms can be rewritten as 
+\begin{align*}
+ \bm{V}_f^T \LRp{\bm{B}_i\tilde{\bm{E}}_m\circ \bm{F}_S^{sm}}\bm{1} &=  \bm{V}_f^T \bm{B}_i \LRp{\tilde{\bm{E}}_m\circ \bm{F}_S^{sm}}\bm{1}\\
+\bm{V}_m^T \LRp{ \tilde{\bm{B}}_i\bm{E}_m \circ \bm{F}_S^{ms}}\bm{1} &= \bm{V}_m^T \tilde{\bm{B}}_i \LRp{ \bm{E}_m \circ \bm{F}_S^{ms}}\bm{1} = \bm{V}_f^T \bm{B}_i \tilde{\bm{E}}_m \LRp{ \bm{E}_m \circ \bm{F}_S^{ms}}\bm{1}
+\end{align*}
+Then, (\ref{eq:mortar}) can be rewritten as 
+\begin{align}
+\bm{M}\td{\bm{u}_N}{t} &+ 
+\begin{bmatrix} \bm{V}_q \\ \bm{V}_f \end{bmatrix}^T
+\LRp{\begin{bmatrix}
+\bm{Q}_i-\bm{Q}_i^T & \bm{E}^T\bm{B}_i\\
+-\bm{B}_i\bm{E} & \\
+\end{bmatrix} \circ \bm{F}_S}\bm{1} + \bm{V}_f^T\bm{B}_i \tilde{\bm{f}}^*_i = 0 \label{eq:mortar2}\\
+\tilde{\bm{f}}^*_i &= \tilde{\bm{E}}_m\bm{f}^*_i + \LRp{\tilde{\bm{E}}_m\circ \bm{F}_S^{sm}}\bm{1} - \tilde{\bm{E}}_m \LRp{ \bm{E}_m \circ \bm{F}_S^{ms}}\bm{1} 
+\nonumber
+\end{align}
+If the surface and mortar nodes are identical, then $\bm{E}_m = \tilde{\bm{E}}_m = \bm{I}$ and $\LRp{\tilde{\bm{E}}_m \circ \bm{F}_S^{sm}}\bm{1} = \LRp{ \bm{E}_m \circ \bm{F}_S^{ms}}\bm{1}$.
+The correction terms within $\tilde{\bm{f}}^*$ cancel out, and we recover the original scheme from \cite{chan2017discretely}.
 
+Note that in (\ref{eq:mortar2}), the only interactions between surface and mortar nodes occurs within the computation of $\tilde{\bm{f}}^*_i$, which can be done solely on mortar interfaces.
+The computational structure of
+\begin{enumerate}
+\item Compute volume contributions to the RHS
+\item Communicate surface contributions to mortar interfaces and compute $\tilde{\bm{f}}^*_i$.
+\item Communicate $\tilde{\bm{f}}^*_i$ back to each element.  
+\end{enumerate}
 %\bm{M}\td{\bm{u}_N}{t} + \begin{bmatrix} \bm{V}_q \\ \bm{V}_f \\ \bm{V}_m \end{bmatrix}^T
@@
 %\bm{Q}_i-\bm{Q}_i^T & \bm{E}^T\bm{B}_i &\\
-%-\bm{B}_i\bm{E} &  & \bm{B}_i\tilde{\bm{E}}_f \\
-%& -\tilde{\bm{B}}_i{\bm{E}}_f & 
+%-\bm{B}_i\bm{E} &  & \bm{B}_i\tilde{\bm{E}}_m \\
+%& -\tilde{\bm{B}}_i{\bm{E}}_m & 
 %\end{bmatrix} \circ \bm{F}_S}\bm{1} + \bm{V}_m^T\tilde{\bm{B}}_i\bm{f}_i^* = 0.
@@
 
+\section{Error estimates}
+
+\note{Weak derivative vs strong derivative error estimates.  SBP yields that the skew-symmetric form is equivalent to weak form.}
+
+\note{Line DG approach can work, but it's flipped: we increase quadrature strength in the direction \emph{orthogonal} to derivative instead.
+May improve accuracy over GLL by giving equivalence w/skew form.}
+
+\section{Numerical experiments}
+
+\subsection{Gauss Lobatto quadrature}
+
+\subsection{Gauss quadrature}
+
 \bibliographystyle{unsrt}

### 2018-10-19 docs/skewform/deriv_estimate.tex

--- 2018-09-12 docs/skewform/main.tex
@@
 
-%\documentclass[review]{siamart0216}
-\documentclass{svjour3}                     % onecolumn (standard format)
-%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
-%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
-%\documentclass[twocolumn]{svjour3}          % twocolumn
+\documentclass{article}
 
-\smartqed  % flush right qed marks, e.g. at end of proof
-
-\makeatletter
-\renewcommand*{\@thmcounterend}{}
-\makeatother
-%\spnewtheorem{lemma}{Lemma}{\bfseries}{\itshape}
-\spnewtheorem{assumption}{Assumption}{\bfseries}{\itshape}
-%\spnewtheorem{remark}{Remark}{\bfseries}{\itshape}
 \usepackage{cite}
@@
 
-\renewcommand{\note}[1]{{\color{blue}{#1}}}
+\newcommand{\note}[1]{{\color{blue}{#1}}}
 
@@
 \date{}
-\author{Jesse Chan}
-\title{Skew-symmetric entropy stable discontinuous Galerkin formulations}
+\author{}
+\title{Error estimates for derivatives}
 
@@
 
-\maketitle
+Let $\LRp{\cdot,\cdot}, \LRp{\cdot,\cdot}_h$ denote the exact and quadrature-based $L^2$ inner product on the volume, and let $\LRa{\cdot,\cdot}, \LRa{\cdot,\cdot}_h$ denote the exact and quadrature-based $L^2$ inner products on the surface.
+We will assume that, in $d > 1$ dimensions, the volume and surface quadratures satisfy
+\begin{align*}
+\LRb{\LRp{u,v} - \LRp{u,v}_h} \lesssim h^{r_{\rm vol}+1+d}  \nor{u}_{H^{r+1}\LRp{D^k}}\nor{v}_{H^{r+1}\LRp{D^k}}\\
+\LRb{\LRa{u,v} - \LRa{u,v}_h} \lesssim h^{r_{\rm surf}+d} \nor{u}_{H^{r+1}\LRp{\partial D^k}}\nor{v}_{H^{r+1}\LRp{\partial D^k}}.
+\end{align*}
 
-\begin{abstract}
-Entropy stable high order methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.
-The construction of such schemes has relied on the use of carefully chosen collocation points \cite{gassner2013skew, fisher2013high, carpenter2014entropy, chan2018efficient} or volume and surface quadrature rules \cite{chan2017discretely, chan2018discretely} to produce operators which satisfy a summation-by-parts (SBP) property.
-In this work, we show how to construct skew-symmetric schemes which are entropy stable even for volume and surface quadratures under which an SBP property does not hold.
-These skew-symmetric formulations avoid the use of a ``strong'' SBP property, and require only that operators exactly differentiate constants and satisfy a discrete form of the fundamental theorem of calculus.
-We conclude with an application of the skew-symmetric formulation for entropy stable schemes on mixed quadrilateral-triangle meshes.
-\end{abstract}
+We wish to show error estimates for certain finite difference approximations of derivatives.  More precisely, we will show that, for sufficiently regular $f$, the weak derivative $D_hf$ satisfies
+\[
+\nor{ \pd{f}{x} - {D}_h {f}}\leq Ch^{p+1} \nor{f}_{H^{p+1}}.
+\]
+More precisely, we will show that the $L^2$ projection satisfies
+\[
+\nor{ \Pi_N\pd{f}{x} - {D}_h {f}}\leq Ch^{\alpha} \nor{f}_{H^{p+1}}, \qquad \alpha = \min\LRp{r_{\rm vol} + 1 -p, r_{\rm surf} - p}.
+\]
 
-%\tableofcontents
-
-\section{Introduction}
-
-High order methods for the simulation of time-dependent compressible flow have the potential to achieve higher levels of accuracy at lower costs compared to current low order schemes \cite{wang2013high}.
-In addition to superior accuracy, the low numerical dispersion and dissipation of high order methods \cite{ainsworth2004dispersive} enables the accurate propagation of waves over long distances and time scales.
-The same properties also make high order methods attractive for unsteady phenomena such as vorticular and turbulent flows, which are sensitive to numerical dissipation \cite{visbal1999high, wang2013high}.
-
-However, when applied to nonlinear conservation laws, high order methods can experience artificial growth and blow-up near under-resolved features such as shocks, turbulence, or boundary layers.
-In practice, the application of high order methods to practical problems requires shock capturing and stabilization techniques (such as artificial viscosity) or solution regularization (such as filtering or limiting) to prevent solution blow-up.
-The resulting schemes for nonlinear conservation laws walk a fine line between stability, robustness, and accuracy.
-Aggressive stabilization or regularization can result in the loss of high order accuracy, while too little can result in instability \cite{wang2013high}.
-Moreover, it can be difficult to determine robust expressions for stabilization paramaters, as parameters which work for one simulation can fail when applied to a different physical regime or discretization setting.
-
-These issues have motivated the introduction of high order \textit{entropy stable} discretizations, which satisfy a semi-discrete entropy inequality while maintaining high order accuracy in smooth regions.
-Proofs of continuous entropy inequalities rely on the chain rule, which does not hold discretely due to effects such as quadrature error.
-By using an approach referred to as ``flux differencing'', entropy stable schemes account for the loss of the chain rule while maintaining a semi-discrete analogue of the continuous entropy inequality.
-These schemes were first introduced as high order collocation methods on tensor product elements in \cite{fisher2013high, carpenter2014entropy, gassner2016split, gassner2017br1}, and were extended to simplicial elements in \cite{crean2017high, chen2017entropy, crean2018entropy, chan2017discretely, chan2018discretely}.
-Entropy stable methods have also been extended to a variety of other discretization settings, including staggered grids \cite{parsani2016entropy}, Gauss-Legendre collocation \cite{chan2018efficient}, and non-conforming meshes \cite{friedrich2017entropy}.
-
-Entropy stable discretizations are built upon flux differencing and a summation-by-parts (SBP) property, which serves as a discrete analogue of integration by parts for quadrature-based discretization matrices.
-However, the SBP property does not hold for some under-integrated quadrature rules, which naturally arise in certain discretization settings.
-For example, on hybrid meshes consisting of both quadrilateral and triangular elements, it is convenient to utilize the same quadrature rule on shared faces between different element types.
-On degree $N$ tensor product elements, a popular choice of quadrature is an $(N+1)$-point Gauss-Legendre-Lobatto (GLL) rule.
-When both volume and surface integrals are approximated using $(N+1)$ point GLL quadrature rules, the SBP property holds, despite the fact that GLL quadrature is inexact for the integrands which appear in finite element formulations \cite{fisher2013high}.
-However, while GLL quadrature induces an SBP property on quadrilateral elements, but does not always result in an SBP property if used for triangular elements \cite{chan2017discretely}.
-%A similar issue arises on non-conforming meshes, where a composite ``split-side'' quadrature naturally when coupling larger elements to smaller elements through a non-conforming interface \cite{kozdon2018energy}.
-For quadrilaterals utilizing a GLL volume quadrature, the use of a composite quadrature on a non-conforming face also results in the loss of the SBP property \cite{chan2017discretely}.
-
-This work proposes an alternative formulation which utilizes a skew-symmetric form of flux differencing.
-Under such a formulation, the proof of entropy stability requires only a weaker ``variational'' form of the SBP property, which holds under a more general class of quadrature rules compared to the decoupled SBP property introduced in \cite{chan2017discretely, chan2018discretely}.
-We show that the skew-symmetric formulation is entropy stable, locally conservative, and free-stream preserving on curved elements, and confirm theoretical results with numerical experiments on triangular, quadrilateral, and hybrid triangular-quadrilateral meshes.
-
-It should be noted that a similar approach to entropy stable discretizations was introduced within a finite difference framework \cite{chen2017entropy, crean2018entropy} using multidimensional differencing operators which are constructed to satisfy accuracy conditions and an SBP property \cite{hicken2016multidimensional}.
-These operators do not correspond to any specific basis or approximation space, but can be shown to exist for nodal points corresponding to sufficiently accurate choices of volume and surface quadrature and can be computed either algebraically or through an optimization problem.
-The formulations presented in this work require similar accuracy conditions, but differ from those of \cite{chen2017entropy, crean2018entropy} in that we start with an explicit approximation space, which then induce quadrature-based operators.
-The formulations presented here also accomodate general choices of volume and surface quadrature (e.g.\ volume quadratures without boundary nodes and over-integrated rules) while ensuring compact coupling conditions between neighboring elements.
-
-The structure of the paper is as follows: Section~\ref{sec:nonlin} describes the continuous entropy inequality which we aim to replicate discretely.
-Section~\ref{sec:approx} and Section~\ref{sec:sbp} introduce polynomial approximation spaces and quadrature-based SBP operators on simplicial and tensor product elements.
-Section~\ref{sec:skew1} then introduces the variational SBP property and an entropy conservative skew-symmetric formulation.
-Section~\ref{sec:skew2} extends the skew-symmetric formulation to curved elements, and Section~\ref{sec:num} presents numerical experiments which verify the theoretical assumptions, stability, and accuracy of the proposed formulations.
-
-%\note{Finish: add Section description.
-Talk about situations where the SBP or decoupled SBP property doesn't hold: triangles with reduced surface quadrature, GLL quads with GQ face quadratures, and hybrid couplings.}
-
-\section{Entropy stability for systems of nonlinear conservation laws}
-\label{sec:nonlin}
-
-We begin by reviewing the dissipation of entropy for a $d$-dimensional system of nonlinear conservation laws on a domain $\Omega$
-\begin{equation}
-\pd{\bm{u}_h}{t}  + \sum_{j=1}^d\pd{\bm{f}_j(\bm{u})}{x_j} = 0, \qquad \bm{u}\in \mathbb{R}^n, \qquad \bm{f}:\mathbb{R}^n\rightarrow\mathbb{R}^n,
-\label{eq:nonlineqs}
-\end{equation}
-where $\bm{u}$ are the conservative variables and $\bm{f}(\bm{u})$ is a vector-valued nonlinear flux function.
-We are interested in nonlinear conservation laws for which a convex entropy function $U(\bm{u})$ exists.
-For such systems, the  \emph{entropy variables} are an invertible mapping $\bm{v}(\bm{u}):\mathbb{R}^n\rightarrow \mathbb{R}^n$ defined as the derivative of the entropy function with respect to the conservative variables
-\begin{align}
-\bm{v}(\bm{u}) = \pd{U}{\bm{u}}.%, \qquad  \text{ invertible}.%\bm{u}\LRp{\bm{v}(\bm{u})} = \bm{u}.
-\label{eq:entropyvarsmap}
-\end{align}
-Several widely used equations in fluid modeling (Burgers, shallow water, compressible Euler and Navier-Stokes equations) yield convex entropy functions $U(\bm{u})$ \cite{hughes1986new, chen2017entropy}.
-Let $\partial \Omega$ be the boundary of $\Omega$ with outward unit normal $\bm{n}$.
-By multiplying the equation (\ref{eq:nonlineqs}) with $\bm{v}(\bm{u})^T$, the solutions $\bm{u}$ of (\ref{eq:nonlineqs}) can be shown to satisfy an entropy inequality
-\begin{equation}
-\int_{\Omega}\pd{U(\bm{u})}{t}\diff{x} + \int_{\partial \Omega} \sum_{j=1}^d \LRp{\bm{v}(\bm{u})^T\bm{f}_j(\bm{u}) - \psi_j\LRp{\bm{v}(\bm{u})}}n_j \diff{x} \leq 0, 
-\label{eq:entropyineq}
-\end{equation}
-where $\bm{n} = \LRp{n_1,\ldots,n_d}$ denotes the outward unit normal, and $\psi_j(\bm{u})$ is some function referred to as the entropy potential.
-
-The proof of (\ref{eq:entropyineq}) requires the use of the chain rule \cite{mock1980systems, harten1983symmetric, dafermos2005compensated}.
-The instability-in-practice of high order schemes for (\ref{eq:nonlineqs}) can be attributed in part to the fact that the discrete form of the equations do not satisfy the chain rule, and thus do not satisfy (\ref{eq:entropyineq}).
-As a result, discretizations of (\ref{eq:nonlineqs}) do not typically possess an underlying statement of stability.
-For low order schemes, this can be offset in practice by the inherent numerical dissipation.
-However, because high order discretizations possess low numerical dissipation, the lack of an underlying discrete stability has been conflated with the idea that high order methods are inherently less stable than low order methods.
-
-\section{Polynomial approximation spaces}
-\label{sec:approx}
-
-In this work, we consider either simplicial reference elements (triangles and tetrahedra) or tensor product reference elements (quadrilaterals and hexahedra).
-We define an approximation space using degree $N$ polynomials on the reference element; however, the natural polynomial approximation space differs depending on the element type \cite{chan2015gpu}.
-
-On a $d$-dimensional reference simplex, the natural polynomial space are total degree $N$ polynomials 
-\[
-P^N\LRp{\hat{D}} = \LRc{\hat{x}_1^{i_1}\ldots\hat{x}_d^{i_d}, \quad \hat{\bm{x}} \in \hat{D}, \quad 0\leq \sum_{k=1}^d i_k \leq N}.
-\]
-In contrast, the natural polynomial space on a $d$-dimensional tensor product element is the space of maximum degree $N$ polynomials
-\[
-Q^N\LRp{\hat{D}} = \LRc{\hat{x}_1^{i_1}\ldots\hat{x}_d^{i_d}, \quad \hat{\bm{x}} \in \hat{D}, \quad 0\leq i_k \leq N, \quad k = 1,\ldots, d}.
-\]
-We denote the natural approximation space on a given reference element $\hat{D}$ by $V^N$.  Furthermore, we denote the dimension of $V^N$ as $N_p = {\rm dim}\LRp{V^N\LRp{\hat{D}}}$.
-
-The proofs presented in this work will also use anisotropic tensor product polynomial spaces, where the maximum polynomial degree varies depending on the coordinate direction.
-We denote such spaces by $Q^{N_1, \ldots, N_d}$, where $N_k$ are non-negative integers and
-\[
-Q^{N_1, N_2, \ldots, N_d}\LRp{\hat{D}} = \LRc{\hat{x}_1^{i_1}\ldots\hat{x}_d^{i_d}, \quad \hat{\bm{x}} \in \hat{D}, \quad 0\leq i_k \leq N_k, \quad k = 1,\ldots, d}.
-\]
-For example, the isotropic tensor product space $Q^N$ is the same as $Q^{N,\ldots,N}$.
-
-We also define trace spaces for each reference element.
-Let $\hat{f}$ be a face of the reference element $\hat{D}$.
-The trace space $V^N \LRp{\hat{f}}$ is defined as the restrictions of functions in $V^N\LRp{\hat{D}}$ to $\hat{f}$, and denote the dimension of the trace space as ${\rm dim}\LRp{V^N\LRp{{\hat{f}}}} = N^f_p$.
-\[
-V^N \LRp{\hat{f}} = \LRc{ \left.u\right|_{\hat{f}}, \quad u \in V^N\LRp{\hat{D}}, \quad \hat{f}\in \partial\hat{D}}.
-\]
-For example, on a $d$-dimensional simplex, $V^N \LRp{\partial \hat{D}}$ consists of total degree $N$ polynomials on simplices of dimension $(d-1)$.
-On a $d$-dimensional tensor product element, $V^N \LRp{\partial \hat{D}}$ consists of maximum degree $N$ polynomials on a tensor product element of dimension $(d-1)$.
-
-%For example, the trace space for degree $N$ polynomials on a quadrilateral face $\hat{f}$ of the bi-unit hexahedral element $[-1,1]^3$ is $Q^N\LRp{\hat{f}}$, while the trace space for degree $N$ polynomials on a triangular face $f$ of the tetrahedron is $P^N\LRp{\hat{f}}$.
-
-%We similarly define the trace space for the surface $\partial \hat{D}$ of $\hat{D}$
-%\[
-%V^N \LRp{\partial \hat{D}} = \LRc{ \left.u\right|_{\partial \hat{D}}, \quad u \in V^N\LRp{\hat{D}}}.
-%\]
-%For example, on a $d$-dimensional simplex, $V^N \LRp{\partial \hat{D}}$ consists of total degree $N$ polynomials on simplices of dimension $(d-1)$.
-On a $d$-dimensional tensor product element, $V^N \LRp{\partial \hat{D}}$ consists of maximum degree $N$ polynomials on a tensor product element of dimension $(d-1)$.
-
-\section{Quadrature-based matrices and decoupled SBP operators}
-\label{sec:sbp}
-%\note{TODO: Standardize hat notation for operators/normals: can refer to operators on reference element and do $k$ superscript for physical elements, or do hats on reference element.}
-Let $\hat{D} \subset\mathbb{R}^d$ denote a reference element with surface $\partial \hat{D}$.  
-The high order schemes in \cite{chan2017discretely, chan2018discretely} begin by approximating the solution in a degree $N$ polynomial basis $\LRc{\phi_j({\bm{x}})}_{i=1}^{N_p}$ on $\hat{D}$.
-These schemes also assume volume and surface quadrature rules $({\bm{x}}_i, w_i)$, $\LRp{{\bm{x}}^f_i,w^f_i}$ on $\hat{D}$.
-We will specify the accuracy of each quadrature rule later, and discuss how quadrature accuracy implies specific summation-by-parts properties.
-
-Let $\bm{V}_q,\bm{V}_f$ denote interpolation matrices, and let $\bm{D}^i$ be the differentiation matrix with respect to the $i$th coordinate such that
-\begin{gather}
-\LRp{\bm{V}_q}_{ij} = \phi_j(\bm{x}_i), \qquad \LRp{\bm{V}_f}_{ij} = \phi_j(\bm{x}^f_i), \qquad \pd{\phi_j(\bm{x})}{x_i} = \sum_{k=1}^{N_p} \LRp{\bm{D}^i_{jk}} \phi_k(\bm{x}).
-\end{gather}
-The interpolation matrices $\bm{V}_q,\bm{V}_f$ map basis coefficients to evaluations at volume and surface quadrature points respectively, while the differentiation matrix ${\bm{D}}_i$ maps basis coefficients of a function to the basis coefficients of its derivative with respect to $x_k$.
-The interpolation matrices are used to assemble the mass matrix $\bm{M}$, the quadrature-based projection matrix $\bm{P}_q$, and lifting matrix $\bm{L}_f$
-\begin{gather}
-\bm{M} = \bm{V}_q^T\bm{W}\bm{V}_q, \qquad \bm{P}_q = \bm{M}^{-1}\bm{V}_q^T\bm{W}, \qquad \bm{L}_f = \bm{M}^{-1}\bm{V}_f^T\bm{W}_f,
-\end{gather}
-where $\bm{W}, \bm{W}_f$ are diagonal matrices of volume and surface quadrature weights, respectively.
-The matrix $\bm{P}_q$ is a quadrature-based discretization of the $L^2$ projection operator $\Pi_N$ onto degree $N$ polynomials.
-
-Interpolation, differentiation, and $L^2$ projection matrices can be combined to construct finite difference operators.
-For example, the matrix $\bm{D}^i_q = \bm{V}_q\bm{D}^i\bm{P}_q$ maps function values at quadrature points to approximate values of the derivative at quadrature points.
-By choosing specific quadrature rules, $\bm{D}^i_q$ recovers high order summation-by-parts finite difference operators in \cite{gassner2013skew, fernandez2014generalized, ranocha2018generalised} and certain operators in \cite{hicken2016multidimensional}.
-However, to address difficulties in designing efficient entropy stable interface terms for nonlinear conservation laws, the PI introduced in \cite{chan2017discretely} a new ``decoupled'' summation by parts matrix which builds interface terms directly into the approximation of the derivative.
-
-Let $\hat{\bm{n}}$ denote the scaled outward normal vector $\hat{\bm{n}} = \LRc{\hat{n}_1\hat{J}_f,\ldots,\hat{n}_d\hat{J}_f}$, where $\hat{J}_f$ is the determinant of the Jacobian of the mapping of a face of $\partial \hat{D}$ to a reference face.
-Let $\hat{\bm{n}}_i$ denote the vector containing values of the $i$th component $\hat{n}_i\hat{J}f$ at all surface quadrature points, and let $\bm{Q}^i = \bm{W}\bm{D}^i_q$.
-The ``decoupled'' summation by parts operator $\bm{D}^i_N$ is defined as the block matrix involving both volume and surface quadratures
-\begin{gather}
-\bm{D}^i_N = \bm{W}_N^{-1} \bm{Q}^i_N, \qquad \bm{W}_N = \LRp{\begin{array}{cc}\bm{W}&\\ & \bm{W}_f\end{array}}, \label{eq:QN}\\
-\bm{E} = \bm{V}_f\bm{P}_q, \qquad \bm{B}^i = \bm{W}_f \diag{\hat{\bm{n}}_i}, \qquad \bm{Q}^i_N  = \LRs{
-\begin{array}{cc}
-\bm{Q}^i - \frac{1}{2}\bm{E}^T\bm{B}^i\bm{E} &  \frac{1}{2}\bm{E}^T\bm{B}^i\\
--\frac{1}{2}\bm{B}^i\bm{E} & \frac{1}{2} \bm{B}^i
-\end{array}}.  \nonumber
-\end{gather}
-Here, $\bm{B}^i$ is a boundary matrix, and $\bm{E}$ denotes the extrapolation matrix which maps values at volume quadrature points to values at surface quadrature points using quadrature-based $L^2$ projection and polynomial interpolation.
-
-It was shown in \cite{chan2017discretely, chan2018efficient} that, when combined with projection and lifting matrices, $\bm{D}^i_N$ produces a high order polynomial approximation of $f\pd{g}{x_i}$. 
-Let $f, g$ be differentiable functions, and let $\bm{f}_i = f(\bm{x}_i)$, $\bm{g}_i = g(\bm{x}_i)$ denote values of $f,g$ at volume and surface quadrature points.  Then,
-\begin{gather}
-f\pd{g}{x_i} \approx \LRs{\begin{array}{cc}
-\bm{P}_q & \bm{L}_f\end{array}} {\rm diag}\LRp{\bm{f}}\bm{D}^i_N \bm{g}.  
-\label{eq:dnapprox}
-\end{gather}
-%The approximation can also be interpreted as solving a variational problem.
-Let $n_i$ be the $i$th component of the unit normal on $\hat{D}$.
-Then (\ref{eq:dnapprox}) is equivalent to finding $u(\bm{x}) \approx f\pd{g}{x_i}$ such that, for all $v\in V^N\LRp{\hat{D}}$,
-%\begin{align*}
-%\int_{\hat{D}} u(\bm{x})v(\bm{x}) = \int_{\hat{D}}{g\pd{\Pi_Nf}{x_i}v} + \int_{\partial \hat{D}}{(f-\Pi_Nf)\frac{\LRp{gv + \Pi_N(gv)}}{2}}\hat{n}_i\hat{J}_f.
-%\end{align*}
-
-For sufficiently accurate quadrature rules, the matrix $\bm{Q}^i_N$ also satisfies a decoupled summation-by-parts (SBP) property, which is used to prove semi-discrete entropy stability for nonlinear conservation laws.
-\begin{theorem}%[Theorem 1 in \cite{chan2017discretely}]
-Assume that the volume and surface quadrature rules are sufficiently accurate such that the quantities
-\[
-\int_{\hat{D}} \pd{u}{x_i} v, \qquad \int_{\partial \hat{D}} u v n_i
-\]
-are integrated exactly for all $u,v \in V^N\LRp{\hat{D}}$ and $i = 1,\ldots, d$.  Then, the decoupled SBP operator $\bm{D}^i_N$ (\ref{eq:QN}) satisfies a summation by parts property:
-\begin{gather}
-\bm{Q}^i_N+\LRp{\bm{Q}^i_N}^T = \bm{B}^i_N, \qquad \bm{B}^i_N = \LRp{\begin{array}{cc}\bm{0}&\\ & \bm{B}^i\end{array}}.\label{eq:dsbp}
-\end{gather}
-%\begin{enumerate}
-%\item The mass matrix is positive definite, 
-%\item The quadrature integrates exactly $\int_{\hat{D}} \pd{u}{x_i} v$ for all $u,v \in V^N\LRp{\hat{D}}$ and $i = 1,\ldots, d$.
-%\end{enumerate}
-\label{lemma:dsbp}
-\end{theorem}
-\begin{proof}
-The proof is a straightforward extension of Theorem 1 in \cite{chan2017discretely} to polynomial approximation spaces on non-simplicial elements.  
-\end{proof}
-
-The assumptions of Theorem~\ref{lemma:dsbp} are satisfied for sufficiently accurate volume and surface quadratures.
-For example, on simplicial elements, (\ref{eq:dsbp}) holds if the volume quadrature is exact for polynomial integrands of total degree $(2N-1)$, and the surface integral is exact for degree $2N$ polynomials on each face.
-Tensor product elements require stricter conditions: (\ref{eq:dsbp}) holds if both the volume and surface quadratures are exact for polynomial integrands of degree $2N$ in each coordinate, due to the fact that derivatives of $u\in Q^N$ are degree $(N-1)$ polynomials with respect to one coordinate and degree $N$ with respect to others.
-
-\begin{remark}
-It should be stressed that the assumptions of Theorem~\ref{lemma:dsbp} are sufficient but not necessary conditions.
-For example, suppose $\hat{D}$ is a quadrilateral element, and that the volume quadrature is a tensor product of $(N+1)$ point one-dimensional Gauss-Legendre-Lobatto (GLL) rule.
-If the surface quadrature also taken to be an $(N+1)$-point GLL rule over each face, then $\bm{Q}^i$ satisfies a traditional SBP property, which implies the decoupled SBP property for $\bm{Q}^i_N$.
-\end{remark}
-
-\section{Skew-symmetric entropy conservative formulations on a single element}
-\label{sec:skew1}
-While the SBP property has been used to derive entropy stable schemes, the strong SBP property (\ref{eq:dsbp}) is difficult to enforce in certain discretization settings, such as hybrid and non-conforming meshes.
-This difficulty is a result of the choices of volume and surface quadrature which naturally arise in these settings.
-We first illustrate how specific pairings of volume and surface quadratures can result in the loss of the SBP property (\ref{eq:dsbp}).
-We then propose a skew-symmetric formulation which is entropy conservative without explicitly requiring operators which satisfy the SBP property.
-
-\subsection{Loss of the SBP property}
-
-In this section, we give examples of specific pairings of volume and surface quadratures under which the decoupled SBP property does not hold (see Figure~\ref{fig:sbploss}).
-We consider two dimensional reference elements $\hat{D}$ with spatial coordinates $x,y$.
-\begin{figure}
-\centering
-\subfloat[GLL volume quadrature, Gauss surface quadrature]{\includegraphics[width=.4\textwidth]{figs/gllgauss.png}\label{subfig:gllgq}}
-\hspace{2em}
-\subfloat[Degree $2N$ volume quadrature, GLL surface quadrature]{\includegraphics[width=.4\textwidth]{figs/trigll.png}\label{subfig:trigll}}
-\caption{Volume and surface quadrature pairs which do not satisfy the assumptions of Theorem~\ref{lemma:dsbp}, and thus do not possess the decoupled SBP property (\ref{eq:dsbp}).
-Volume quadrature nodes are drawn as circles, while surface quadrature nodes are drawn as squares.}
-\label{fig:sbploss}
-\end{figure}
-
-\paragraph{Quadrilateral elements (Figure~\ref{subfig:gllgq})} We first consider a quadrilateral element $\hat{D}$ with an $(N+1)$ point tensor product GLL volume quadrature and $(N+1)$ point Gauss quadrature on each face.
-Let $u,v \in Q^N$ denote two arbitrary degree $N$ polynomials.
-The assumptions of Theorem~\ref{lemma:dsbp} are that the volume quadrature exactly integrates $\int_{\hat{D}} \pd{u}{x} v$ and that the surface quadrature exactly integrates $\int_{\partial \hat{D}} u v n_i$ on $\hat{D}$.
-Because the $(N+1)$-point Gauss rule is exact for polynomials of degree $2N+1$ and the product $uv \in P^{2N}$ on each face, the surface quadrature satisfies the assumptions of Theorem~\ref{lemma:dsbp}.
-However, the 1D GLL rule is only exact for polynomials of degree $(2N-1)$.
-The derivative $\pd{u}{x}$ is a polynomial of degree $(N-1)$ in $x$, but is degree $N$ in $y$.
-Thus, $\pd{u}{x}v$ is a polynomial of degree $(2N-1)$ in $x$ but degree $2N$ in $y$, and is not integrated exactly by the volume quadrature.
-
-\paragraph{Triangular elements (Figure~\ref{subfig:trigll})} We next consider a triangular element $\hat{D}$, where the volume quadrature is exact for degree $2N$ polynomials \cite{xiao2010quadrature} and an $(N+1)$-point GLL quadrature on each face.
-Let $u,v \in P^N$ denote two arbitrary degree $N$ polynomials.
-The derivative $\pd{u}{x} \in P^{(N-1)}$, and $\pd{u}{x}v \in P^{(2N-1)}$, so the volume quadrature satisfies the assumptions of Theorem~\ref{lemma:dsbp}.
-However, because the surface quadrature is exact only degree $(2N-1)$ polynomials and the trace of $uv\in P^{2N}$, the surface quadrature does not satisfy the assumptions of Theorem~\ref{lemma:dsbp}.
-
-\begin{figure}
-\centering
-\begingroup
-\captionsetup[subfigure]{width=.425\textwidth}
-\subfloat[Insufficiently accurate surface quadrature on the triangle element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D.png}\label{subfig:hybrid1}}
-\endgroup
-\hspace{2em}
-\subfloat[Incompatible surface quadrature on the quadrilateral element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D_GQ.png}\label{subfig:hybrid2}}
-\caption{Examples of interface couplings which do not result in a decoupled SBP property (\ref{eq:dsbp}).
-Volume quadrature nodes are drawn as circles, while surface quadrature nodes are drawn as  squares.
-}
-\label{fig:hybrid}
-\end{figure}
-
-These specific pairings of volume and surface quadratures appear naturally for hybrid meshes consisting of DG-SEM quadrilateral elements (using GLL volume quadrature) and triangular elements, as shown in Figure~\ref{fig:hybrid}.
-In Figure~\ref{subfig:hybrid1}, the surface quadrature is a $(N+1)$ point GLL rule, and results in a loss of the SBP property on the triangle.
-In Figure~\ref{subfig:hybrid2}, the surface quadrature is a $(N+1)$ point Gauss-Legendre rule, and results in a loss of the SBP property on the quadrilateral element.
-The goal of this work is to construct high order accurate discretizations which preserve entropy conservation for situations in which the decoupled SBP property (\ref{eq:dsbp}) does not hold.
-
-%\begin{figure}
-%\centering
-%\subfloat[Hex-pyramid coupling]{\includegraphics[width=.35\textwidth]{figs/hybrid3D.png}}
-%\hspace{2em}
-%\subfloat[Hex-prism coupling]{\includegraphics[width=.375\textwidth]{figs/hybrid3D_wedge.png}}
-%\caption{Illustration of a 3D coupling between a GLL hexahedral element and a pyramid.  The SBP property does not hold on the pyramid due to the use of GLL quadrature on the quadrilateral face. }
-%\label{fig:hybrid3d}
-%\end{figure}
-%
-%\begin{figure}[!h]
-%\centering
-%\begingroup
-%\captionsetup[subfigure]{width=.45\textwidth}
-%\subfloat[Entropy stable inter-element coupling in \cite{friedrich2017entropy} for a non-conforming interface.]{\raisebox{0em}{\includegraphics[width=.45\textwidth]{figs/nonconSBP.png}}\label{subfig:noncon2}}
-%\hspace{2em}
-%\subfloat[Entropy stable inter-element coupling in this work for a non-conforming interface.
-The dotted black lines denote communication between neighboring elements.]{\raisebox{-0em}{\includegraphics[width=.45\textwidth]{figs/nonconQuad.png}}\label{subfig:noncon1}}
-%\endgroup
-%%\subfloat[3D hexahedra-pyramid coupling]{\includegraphics[width=.31\textwidth]{figs/hybrid3D.png}}
-%\caption{Illustration of two different entropy stable inter-element couplings for $h$ non-conforming meshes.
-Each dashed red line indicates a dyadic flux computation required between two nodes.
-} %The coupling terms in \cite{friedrich2017entropy} requires computing dyadic fluxes between \emph{each} pair of nodes on adjacent interfaces.
-The new approach results in a simpler communication pattern between elements.
-}
-%\label{fig:noncon}
-%\end{figure}
-
-\subsection{A variational SBP property}
-
-%The main restriction we wish to overcome
-
-The property (\ref{eq:dsbp}) (which we will refer to as the ``strong'' SBP property) relates the polynomial exactness of specific quadrature rules to algebraic properties of quadrature-based matrices.
-We will relax accuracy conditions on these quadrature rules by utilizing a weaker ``variational'' version of the SBP property variational, which relies on the exactness of volume and surface quadrature rules only for specific polynomials.
-
-On tensor product elements, we restrict ourselves to isotropic volume quadrature rules which are construced from tensor products of one-dimensional quadrature formulas.
-For the remainder of this work, the degree of the multi-dimensional quadrature rule on tensor product elements will refer to the degree of exactness of the one-dimensional rule.
-For example, we refer to the quadrature rule constructed through a tensor product of one-dimensional $(N+1)$-point GLL quadrature rules as a degree $(2N-1)$ quadrature rule.
-This choice of quadrature is sufficient to guarantee that the mass matrix is positive definite \cite{canuto2007spectral}.
-
-Throughout the remainder of this work, we make the following assumptions:
-\begin{assumption}
-\label{ass:quad}
-Let $v \in V^{N}$ denote a fixed polynomial.  We assume that: 
-\begin{enumerate}
-\item the mass matrix $\bm{M}$ is positive definite under the volume quadrature rule,
-\item the volume quadrature rule is exact for integrals of the form\\$\int_{\hat{D}} \pd{u}{\hat{x}_j} v$ for all $u \in V^N\LRp{\hat{D}}$, $j = 1,\ldots, d$.
-\item the surface quadrature rule is exact for integrals of the form\\$\int_{\partial \hat{D}} u v n_j$ for all $u \in V^N\LRp{\hat{D}}$, $j = 1,\ldots, d$, and $f \in \partial \hat{D}$.  
-\end{enumerate}
-\end{assumption}
-
-The conditions of Assumption~\ref{ass:quad} are relatively standard within the SBP literature \cite{hicken2016multidimensional, chan2017discretely, crean2018entropy}, though they have not previously depended on the specific choice of polynomial $v(\bm{x})$.
-The strong SBP property (\ref{eq:dsbp}) can be derived if Assumption~\ref{ass:quad} is satisfied for all $v \in V^N$.
-In contrast, we will derive a weaker SBP property which assumes only that Assumption~\ref{ass:quad} holds for a few specific polynomials $v$.
-In Sections~\ref{sec:singleelem} and \ref{sec:curved}, specific polynomials $v$ will be motivated by proofs of entropy stability, and we will present examples of volume and surface quadrature rules on simplicial and tensor product elements which satisfy Assumption~\ref{ass:quad} for these choices of $v$.
-% the quantities $\int_{\hat{D}} \pd{u}{x_i} v$ and
-
-%This section presumes that Assumption~\ref{ass:quad} is satisfied for general  $M_{\rm vol}, M_{\rm surf}$.
-In Sections~\ref{sec:singleelem} and \ref{sec:curved}, specific values of $M_{\rm vol}, M_{\rm surf}$ will be motivated by proofs of entropy stability, and Assumption~\ref{ass:quad} will be verified for specific choices of volume and surface quadrature on simplicial and tensor product elements.
-
-The conditions of Assumption~\ref{ass:quad} imply the following variational SBP property:
-\begin{lemma}
-\label{lemma:vsbp}
-%Assume that $v \in V^M$.  
-%If the volume quadrature is exact for polynomials of degree $N+M-1$ and the surface quadrature is exact for polynomials of degree $N+M$, 
-Suppose Assumption~\ref{ass:quad} holds, and let $\bm{v}_q$ denote the vector of values of $v(\bm{x})$ of at volume quadrature points.
-Let $u(\bm{x})$ be an $L^2$ integrable function, and let $\bm{u}_q$ denote the values of $u(\bm{x})$ at quadrature points.
-Then,
-\[
-\bm{v}_q^T\bm{Q}_i \bm{u}_q = \bm{v}_q^T\LRp{ {\bm{E}}^T \bm{B}^i\bm{E} - \bm{Q}_i^T}\bm{u}_q.
-\]
-\end{lemma}
-\begin{proof}
-Recall that $\bm{Q}_i = \bm{W} \bm{V}_q \bm{D}^i\bm{P}_q$.
-The quadrature-based $L^2$ projection $\Pi_Nu$ is a polynomial of degree $N$, and is computed discretely as $\bm{P}_q\bm{u}_q$.
-By the exactness of the volume quadrature in Assumption~\ref{ass:quad},
-\begin{align*}
-\bm{v}_q^T\bm{Q}_i \bm{u}_q &= \bm{v}_q^T\bm{W} \bm{D}^i \bm{P}_q \bm{u}_q = \int_{\hat{D}} \pd{\Pi_N u}{\hat{x}_i} v = \int_{\partial \hat{D}} (\Pi_N u) v \hat{n}_i - \int_{\hat{D}} \LRp{\Pi_N u} \pd{v}{\hat{x}_i}.
-\end{align*}
-%Since the volume integrand $\LRp{\Pi_N u} \pd{v}{\hat{x}_i} \in V^{N+M-1}$, it is exactly integrated using the volume quadrature.  
-%Additionally, since $\Pi_N u \in V^N$ and $v\in V^M$ on the surface of $\hat{D}$, the surface quadrature is exact for the surface integral and 
-By the exactness of the surface quadrature in Assumption~\ref{ass:quad}, 
-\begin{align*}
-\int_{\partial \hat{D}} (\Pi_N u) v \hat{n}_i -   \int_{\hat{D}} \LRp{\Pi_N u} \pd{v}{\hat{x}_i} &=\bm{v}_f^T\bm{B}^i \bm{E}\bm{u}_q - \LRp{\bm{V}_q\bm{D}^i\bm{P}_q\bm{v}_q}^T\bm{W} \bm{V}_q\bm{P}_q\bm{u}_q\\
-=& \bm{v}_q^T {\bm{E}}^T\bm{B}^i \bm{E}\bm{u}_q - \bm{v}_q^T \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T \bm{W} \bm{V}_q\bm{P}_q\bm{u}_q,
-\end{align*}
-where we have used that, since $v \in V^N$, $\bm{v}_f = \bm{V}_f\bm{P}_q\bm{v}_q = \bm{E}\bm{v}_q$.  The proof is completed by noting that 
-\[
-\bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T \bm{W} \bm{V}_q\bm{P}_q = \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{M}\bm{P}_q = \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T\bm{W} = \bm{Q}_i^T.
-\]
-\end{proof}
-
-The variational SBP property in Lemma~\ref{lemma:vsbp} can be used to prove a similar property for the decoupled SBP operator.
-This property, along with the exact differentiation of constants, is necessary for the proof of entropy stability.
-\begin{lemma}
-Suppose Assumption~\ref{ass:quad} holds.  Let $\bm{D}^i_N$ be a decoupled SBP operator on the reference element $\hat{D}$.  
-%Let the volume quadrature be exact for polynomials of degree $N+M-1$, and let the surface quadrature be exact for polynomials of degree $N+M$ for some $M \leq N$.
-Suppose $v\in V^N$ and that $\LRl{v}_{\partial \hat{D}} \in P^M\LRp{\partial \hat{D}}$.
-\note{Fix this to match assumptions in previous lemma.}
-Then, the following variational SBP property holds:
-\[
-\bm{v}^T\bm{Q}^i_N\bm{u} = \bm{v}^T\LRp{\bm{B}^i_N - \LRp{\bm{Q}^i_N}^T}\bm{u}.%, \qquad \bm{Q}^i_N \bm{1} = \bm{0}.
-\]
-where $\bm{v}, \bm{u}$ denotes the values of $v$ and $u$ at both volume and surface quadrature points.  
-\label{lemma:vdsbp}
-\end{lemma}
-\begin{proof}
-For convenience, let $\bm{u}_q, \bm{u}_f$ and $\bm{v}_q, \bm{v}_f$ denote evaluations of $u$ and $v$ at volume and surface points, such that 
-\[
-\bm{u} = \begin{pmatrix} \bm{u}_q\\ \bm{u}_f\end{pmatrix}, \qquad \bm{v} = \begin{pmatrix} \bm{v}_q\\ \bm{v}_f\end{pmatrix}.  
-\]
-The proof of the variational summation by parts property uses the definition of $\bm{Q}^i_N$ (\ref{eq:QN}), 
-\begin{align*}
-\bm{v}^T\bm{Q}^i_N\bm{u} &= \bm{v}_q^T\bm{Q}_i \bm{u}_q - \frac{1}{2}\LRp{\bm{E}\bm{v}_q}^T\bm{B}^i \LRp{\bm{E}\bm{u}_q} + \frac{1}{2}\LRp{\bm{E}\bm{v}_q}^T\bm{B}^i \bm{u}_f\\
-& - \frac{1}{2}\bm{v}_f^T\bm{B}^i \LRp{\bm{E}\bm{u}_q} + \frac{1}{2}\bm{v}_f^T\bm{B}^i \bm{u}_f.
-\end{align*}
-Applying Lemma~\ref{lemma:vsbp} then yields
-\begin{align*}
-\bm{v}^T\bm{Q}^i_N\bm{u} =& -\bm{v}_q^T\bm{Q}^T_i \bm{u}_q + \frac{1}{2}\LRp{\bm{E}\bm{v}_q}^T\bm{B}^i \LRp{\bm{E}\bm{u}_q} + \frac{1}{2}\LRp{\bm{E}\bm{v}_q}^T\bm{B}^i \bm{u}_f\\
-&\qquad - \frac{1}{2}\bm{v}_f^T\bm{B}^i \LRp{\bm{E}\bm{u}_q} - \frac{1}{2}\bm{v}_f^T\bm{B}^i \bm{u}_f + \bm{v}_f^T\bm{B}^i \bm{u}_f\\
-=& \begin{pmatrix} \bm{v}_q\\ \bm{v}_f\end{pmatrix}^T 
-\left(\begin{pmatrix}
-\bm{0}& \\
-& \bm{B}^i
-\end{pmatrix}\right.+
-\left.\begin{pmatrix}
--\bm{Q}_i^T + \frac{1}{2}{\bm{E}}^T\bm{B}^i \bm{E} & -\frac{1}{2} \LRp{\bm{B}^i \bm{E}}^T\\
-\frac{1}{2}\bm{B}^i \bm{E} & -\frac{1}{2}\bm{B}^i
-\end{pmatrix}  \right)
-\begin{pmatrix} \bm{u}_q\\ \bm{u}_f\end{pmatrix}\\
-=& \bm{v}^T\LRp{\bm{B}^i_N - \LRp{\bm{Q}^i_N}^T}\bm{u}.
-\end{align*}
-\end{proof}
-
-\subsection{Entropy stability on a reference element}
-\label{sec:singleelem}
-
-In this section, we construct so-called ``entropy stable'' schemes on the reference element $\hat{D}$.
-These methods ensure that the entropy inequality (\ref{eq:entropyineq}) is satisfied discretely by avoiding the use of the chain rule in the proof of entropy dissipation.
-Entropy stable schemes rely on two main ingredients: an entropy stable numerical flux as defined by Tadmor \cite{tadmor1987numerical} and a concept referred to as ``flux differencing''.
-Let $\bm{f}_S\LRp{\bm{u}_L,\bm{u}_R}$ be a numerical flux function which is a function of ``left'' and ``right'' states $\bm{u}_L,\bm{u}_R$.
-The numerical flux $\bm{f}_S$ is \textit{entropy conservative} if it satisfies the following three conditions:
-\begin{gather}
-\bm{f}^i_S(\bm{u},\bm{u}) = \bm{f}_i(\bm{u}), \qquad \text{(consistency)}\\
-\bm{f}^i_S(\bm{u}_L,\bm{u}_R) = \bm{f}^i_S(\bm{u}_R,\bm{u}_R), \qquad \text{(symmetry)}\nonumber\\
-\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}^i_S(\bm{u}_L,\bm{u}_R) = \psi_i(\bm{u}_L) - \psi_i(\bm{u}_R), \qquad \text{(conservation)}\nonumber
-\label{eq:esflux}
-\end{gather}
-for $j = 1,\ldots, d$.  The construction of entropy stable schemes will utilize (\ref{eq:esflux}) in discretizations of both volume and surface terms in a DG formulation.
-
-Proving a semi-discrete entropy inequality typically involves the ``strong'' SBP condition.
-The schemes presented in this work will relax this, and use only two (weaker) conditions: that the decoupled SBP operator $\bm{D}^i_N$ exactly differentiates constants and that a discrete form of the fundamental theorem of calculus (FTC) holds.
-The latter property can be derived from the variational SBP property in Lemma~\ref{lemma:vdsbp}.
-
-\begin{lemma}{(Discrete FTC and exact differentiation of constants)}
-\label{lemma:sbpcor}
-Suppose Assumption~\ref{ass:quad} holds for $v(\bm{x}) = 1$.  
-Let $u(\bm{x})$ be some $L^2$ integrable function, and let $\bm{u}$ denote the vector of values of $u(\bm{x})$ at volume and surface quadrature points. Then, 
-\[
-\bm{1}^T\bm{Q}^i_N\bm{u} = \bm{1}^T\bm{B}^i_N\bm{u}, \qquad \bm{Q}^i_N\bm{1} = \bm{0}.
-\]
-\end{lemma}
-\begin{proof}
-The proof that $\bm{Q}^i_N \bm{1} = \bm{0}$ follows from the property that polynomials are equal to their $L^2$ projection, and is identical to that of \cite{chan2017discretely,chan2018discretely}.
-The proof of the first equality follows from $\bm{Q}^i_N \bm{1} = \bm{0}$ and Lemma~\ref{lemma:vdsbp}
-\[
-\bm{1}^T\LRp{\bm{Q}^i_N}\bm{u} = \bm{1}^T\LRp{\bm{B}^i_N - \LRp{\bm{Q}^i_N}^T}\bm{u} = \bm{1}^T{\bm{B}^i_N}\bm{u}.
-\]
-\end{proof}
-
-We can now construct a skew-symmetric formulation on the reference element $\hat{D}$ and show that it is semi-discretely entropy conservative.
-This formulation can be made entropy stable by adding interface dissipation.
-Let $\bm{u}_h$ denote the discrete solution, and let $\bm{u}_q$ denote the values of the solution at volume quadrature points.
-We define the auxiliary conservative variables $\tilde{\bm{u}}$ in terms of the $L^2$ projections of the entropy variables
-\begin{gather}
-\bm{v}_q = \bm{v}\LRp{\bm{u}_q}, \qquad \tilde{\bm{v}} = \begin{bmatrix}
-\bm{V}_q\\
-\bm{V}_f
-\end{bmatrix}\bm{P}_q\bm{v}_q, \qquad \tilde{\bm{u}} = \bm{u}\LRp{\tilde{\bm{v}}}.
-\end{gather}
-The skew-symmetric formulation for (\ref{eq:nonlineqs}) on $\hat{D}$ is given in terms of $\tilde{\bm{u}}$
-\begin{gather}
-\bm{M}\pd{\bm{u}_h}{t} + \sum_{i=1}^d\LRs{\begin{array}{c}
-\bm{V}_q \\ \bm{V}_f\end{array}}^T 
-\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}^i_S}\bm{1} + \bm{V}_f^T\bm{B}^i\bm{f}_i^* = 0,  \label{eq:esdgSkew}\\
-\LRp{\bm{F}^i_S}_{ij} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q,\nonumber
-\end{gather}
-where $\bm{f}^*$ is some numerical flux, and the matrix $\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}$ possesses the following block structure:
-\[
-\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} = \begin{pmatrix}
-\bm{Q}_i-\bm{Q}_i^T & {\bm{E}}^T \bm{B}^i\\
--\bm{B}^i\bm{E} & \bm{0}
-\end{pmatrix}.
-\]
-Multiplying the formulation (\ref{eq:esdgSkew}) by $\bm{M}^{-1}$ on both sides yields a strong form 
-\begin{gather}
-\pd{\bm{u}_h}{t} + \sum_{i=1}^d \LRs{\begin{array}{cc}
-\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}^i_S}\bm{1} + \bm{L}_f \diag{\hat{\bm{n}}_i}\bm{f}_i^* = 0,\label{eq:esdg}
-%\\
-%\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q.
-%\label{eq:esdgSkewStrong}
-\end{gather}
-We can now show that the skew-symmetric formulation is entropy conservative.  
-\begin{theorem}
-Suppose Assumption~\ref{ass:quad} holds for $v(\bm{x}) = 1$.  
-Then, the formulation (\ref{eq:esdgSkew}) is entropy conservative such that
-\begin{equation}
-\bm{1}^T\bm{W}\pd{U(\bm{u}_q)}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}^i \LRp{\psi_i(\tilde{\bm{u}}_f) - \tilde{\bm{v}}_f^T\bm{f}_i^*} = 0, \qquad \bm{u}_q = \bm{V}_q\bm{u}.
-\label{eq:esdgthm}
-\end{equation}
-\label{thm:esdg}
-\end{theorem}
-\begin{proof}
-Testing (\ref{eq:esdgSkew}) by $\bm{v}_h = \bm{P}_q\bm{v}_q$ yields 
-\begin{align}
-\bm{v}_q^T\bm{W}\pd{\LRp{\bm{V}_q\bm{u}}}{t} + \sum_{i=1}^d
-\tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}^i_S}\bm{1} + \tilde{\bm{v}}_f^T \bm{B}^i\bm{f}_i^* = 0.
-\end{align}
-One can show that \cite{chan2017discretely}
-\begin{align*}
-\tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}^i_S}\bm{1} &= \tilde{\bm{v}}^T \LRp{\bm{Q}^i_N \circ \bm{F}^i_S}\bm{1} - \tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N}^T \circ \bm{F}^i_S}\bm{1}\\
-&= \tilde{\bm{v}}^T \LRp{\bm{Q}^i_N \circ \bm{F}^i_S}\bm{1} - \bm{1}^T \LRp{{\bm{Q}^i_N} \circ \bm{F}^i_S}\tilde{\bm{v}}.
-\end{align*}
-Where we have used that $\bm{F}^i_S$ is symmetric and that the Hadamard product commutes.  Applying the conservation condition on $\bm{f}^i_S$ in (\ref{eq:esflux}) then yields
-\begin{align*}
-&\tilde{\bm{v}}^T \LRp{\bm{Q}^i_N \circ \bm{F}^i_S}\bm{1} - \bm{1}^T \LRp{{\bm{Q}^i_N} \circ \bm{F}^i_S}\tilde{\bm{v}} \\
-&= \sum_{jk} \LRp{\bm{Q}^i_N}_{jk} \LRp{\tilde{\bm{v}}_j-\tilde{\bm{v}}_k} \bm{f}_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k} = \sum_{jk} \LRp{\bm{Q}^i_N}_{jk} \LRp{\psi_i(\tilde{\bm{u}}_j) - \psi_i(\tilde{\bm{u}}_k)}\\
-&= \bm{1}^T\LRp{\bm{Q}^i_N}\psi_i(\tilde{\bm{u}}) - \psi_i(\tilde{\bm{u}})^T\LRp{\bm{Q}^i_N}\bm{1} = \bm{1}^T\LRp{\bm{Q}^i_N}\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i_N\psi_i(\tilde{\bm{u}}),
-\end{align*}
-where we have used Corollary~\ref{lemma:sbpcor} in the last equality.  Noting that $\bm{1}^T\bm{B}^i_N\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i \psi_i(\tilde{\bm{u}}_f)$ completes the proof.
-%The final step of the proof is  where we have used Lemma~\ref{lemma:vdsbp} to integrate by parts and conclude that $\bm{1}^T\LRp{\bm{Q}^i_N}^T = 0$.
-\end{proof}
-
-\begin{remark}
-The primary difference between this formulation and the formulation presented in \cite{chan2017discretely} is the use of the skew-symmetric volume operator $\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}$.
-When the strong SBP property (\ref{eq:dsbp}) holds, this formulation is exactly equivalent to the strong formulation presented in \cite{chan2017discretely}.
-However, the only result necessary to prove Theorem~\ref{thm:esdg} is Corollary~\ref{lemma:sbpcor}.
-The decoupled SBP property (\ref{eq:dsbp}) is not utilized, and is not necessary to guarantee a semi-discrete conservation of entropy for the skew-symmetric formulation.
-\end{remark}
-
-The skew symmetric formulation can also be shown to be locally conservative in the sense of \cite{shi2017local}, which is sufficient to show the numerical solution convergences to the weak solution under mesh refinement.
-\begin{theorem}
-The formulation (\ref{eq:esdgSkew}) is locally conservative such that
-\begin{align}
-\bm{1}^T\bm{W}\pd{\LRp{\bm{V}_q\bm{u}}}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}^i\bm{f}_i^* = 0. 
-\end{align}
-\end{theorem}
-\begin{proof}
-To show local conservation, we test (\ref{eq:esdgSkew}) with $1$
-\begin{align}
-\bm{1}^T\bm{W}\bm{V}_q\pd{\bm{u}_h}{t} + \sum_{i=1}^d
-\bm{1}^T
-\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}}\bm{f}_i^* = 0. 
-\end{align}
-Because $\bm{F}_S$ is symmetric and $\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}$ is skew-symmetric, their Hadamard product is also skew-symmetric.
-Using that $\bm{x}^T\bm{A}\bm{x} = 0$ for any skew symmetric matrix $\bm{A}$, the volume term vanishes
-\[
-\bm{1}^T\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} = 0.
-\]
-% and that the Hadamard product is commutable $\LRp{\bm{A}\circ\bm{B} }^T = \bm{A}^T\circ\bm{B}^T$
-%\begin{align}
-%\bm{1}^T\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} &= \bm{1}^T\LRp{\bm{Q}^i_N\circ \bm{F}_S}\bm{1} - \bm{1}^T\LRp{\LRp{\bm{Q}^i_N}^T\circ \bm{F}_S}\bm{1}\\
-%&= \bm{1}^T\LRp{\bm{Q}^i_N\circ \bm{F}_S}\bm{1} - 
-%\bm{1}^T\LRp{\bm{Q}^i_N\circ \bm{F}_S}\bm{1} = 0,
-%\end{align}
-%where we have used that $\bm{F}_S$ is symmetric.
-%\begin{align}
-%\end{align}
-\end{proof}
-
-\subsection{On quadrature conditions for Assumption~\ref{ass:quad} with $v = 1$}
-\label{sec:assump1}
-Apart from algebraic manipulations, only Corollary~\ref{lemma:sbpcor} is necessary to prove entropy conservation for (\ref{eq:esdg}).
-Corollary~\ref{lemma:sbpcor} requires that Assumption~\ref{ass:quad} holds for $v=1$.
-Thus, the volume and surface quadratures must be sufficiently accurate to guarantee that the mass matrix is positive definite and to integrate
-\begin{equation}
-\int_{\hat{D}} \pd{u}{x_i}, \qquad \int_{\partial \hat{D}} u n_i. \label{eq:affineints}
-\end{equation}
-On simplicial elements, the mass matrix is guaranteed to be positive definite for any volume quadrature which is exact for degree $2N$ polynomial integrands.
-This choice of volume quadrature also guarantees that the volume term in (\ref{eq:affineints}) is integrated exactly.
-The surface quadrature can thus be taken to be any quadrature rule which is exact for only degree $N$ integrands on faces.
-In contrast, the construction of simplicial decoupled SBP operators has required face quadratures which are accurate for degree $2N$ polynomials \cite{chan2017discretely, chan2018discretely}.
-
-On tensor product elements, we can take any degree $(2N-1)$ quadrature rule which ensures a positive definite mass matrix (e.g.\ a $(N+1)$-point GLL quadrature), as a quadrature of this accuracy is sufficient to exactly integrate the volume term in (\ref{eq:affineints}).
-For the surface quadrature, we can again take any quadrature rule which is exact for degree $N$ polynomial integrands.
-For example, on a quadrilateral element, one can use $\left\lceil\frac{N+1}{2}\right\rceil$-point Gauss quadrature rule or a $\left\lceil\frac{N+3}{2}\right\rceil$-point GLL rule as face quadratures for a degree $N$ scheme.
-
-%\note{link SBP accuracy section} %These issues will be discussed further in Section~\ref{sec:num}.
-%This minimal degree of surface quadrature accuracy will depend on the nature of the physical mesh, and will differ between affine and curved elements.
-
-%\subsection{On the accuracy of the skew-symmetric formulation}
-%
-%We note that, while the proofs in this Section show that it is possible to maintain entropy stability under reduced accuracy surface quadrature rules, the accuracy of the resulting schemes depend on the accuracy of the volume and surface quadrature.
-%We briefly remark on the accuracy of (\ref{eq:esdgSkew}).
-%
-%It was shown in \cite{chan2017discretely} that, when combined with quadrature-based $L^2$ projection and lifting operators, the decoupled SBP operator (\ref{eq:dsbp}) is a degree $N$ approximation to the derivative.
-This approximation can be interpreted as augmenting a volume approximation of the derivative with boundary correction terms \cite{chan2018efficient}.
-Let $f(\bm{x}),g(\bm{x})$ denote two $L^2$ integrable functions, and let $\bm{f}_N, \bm{g}_N$ denote the vectors of values of $f,g$ at both volume and surface quadrature points.
-A degree $N$ approximation $u\in V^N$ to $f\pd{g}{x_i}$ can be constructed via
-%\[
-%\bm{M}\bm{u} = \LRs{\begin{array}{c}
-%\bm{V}_q \\ \bm{V}_f\end{array}}^T \diag{\bm{f}_N}\bm{Q}^i_N\bm{g}_N, \qquad \bm{f}_N = \begin{bmatrix} \bm{f}_q\\ \bm{f}_f\end{bmatrix}, \quad \bm{g}_N = \begin{bmatrix} \bm{g}_q\\ \bm{g}_f\end{bmatrix},
-%\]
-%where $\bm{u}$ denotes the vector of coefficients for $u$.  Expanding terms using (\ref{eq:dsbp}) yields 
-%%\note{fix!  add $E = V_fP_q$}
-%\begin{align*}
-%\bm{M}\bm{u} &= \bm{V}_q^T \diag{\bm{f}_q} \bm{Q}^i \bm{g}_q + \frac{1}{2}\bm{V}_q^T\diag{\bm{f}_q}\bm{E}^T\bm{B}^i \LRp{\bm{g}_f -\bm{E}\bm{g}_q}  \\
-%&\qquad + \frac{1}{2}\bm{V}_f^T\diag{\bm{f}_f}\bm{B}^i \LRp{\bm{g}_f -\bm{E}\bm{g}_q}.
-%\end{align*}
-%where we have used that $\bm{B}^i$ is diagonal and that diagonal matrices commute.
-Let $\bm{v}$ are coefficients of some test function $v\in V^N$, and let $\bm{v}_q = \bm{V}_q\bm{v}$ and $\bm{v}_f = \bm{V}_f \bm{v}$.
-Multiplying by $\bm{v}^T$ on both sides yields that $\bm{v}^T\bm{M}\bm{u}$ is equal to
-%\begin{align*}
-%\LRp{\diag{\bm{f}_q}\bm{v}_q}^T \bm{Q}^i \bm{g}_q + \frac{1}{2}\LRp{\diag{\bm{f}_f}\bm{v}_f+\bm{E}\diag{\bm{f}_q}\bm{v}_q}^T\bm{B}^i \LRp{\bm{g}_f -\bm{E}\bm{g}_q} .
-%\end{align*}
-%This algebraic expression can be reinterpreted as a quadrature approximation of the following variational problem
-%\begin{equation}
-%\int_{\hat{D}} u v = \int_{\hat{D}} f\pd{\Pi_N g}{x_i}v + \int_{\partial \hat{D}} \LRp{g - \Pi_Ng} \LRp{\frac{fv + \Pi_N(fv)}{2}} \hat{n}_i, \qquad \forall v\in V^N.
-%\label{eq:var}
-%\end{equation}
-%
-%The skew-symmetric formulation can be interpreted as approximating a skew symmetric form of (\ref{eq:var}).  Integrating half of the volume term by parts yields 
-%\begin{align}
-%&\frac{1}{2} \int_{\hat{D}} \LRp{f\pd{\Pi_N g}{x_k}v - g \pd{\Pi_N\LRp{fv}}{x_i}} \label{eq:skewDN}\\
-%&+\frac{1}{2} \int_{\partial \hat{D}} \LRp{fgv\hat{n}_i + \LRp{g - \Pi_Ng} \LRp{\frac{fv + \Pi_N(fv)}{2}} \hat{n}_i}\nonumber
-%\end{align}
-%Approximating integrals using quadrature allows us to express (\ref{eq:skewDN}) as a matrix equation using $\bm{B}^i_N$ in (\ref{eq:dsbp})
-%%\begin{align*}
-%%\bm{v}^T\bm{M}\bm{u} &= \bm{v}^T\LRp{\LRs{\begin{array}{c}
-%%\bm{V}_q \\ \bm{V}_f\end{array}}}^T\diag{\bm{f}_N}\LRp{\bm{B}^i_N + \frac{1}{2}\LRp{\bm{Q}_N-\bm{Q}_N^T}}{\bm{g}_N}\\
-%%&= \bm{v}^T \LRp{\frac{1}{2}\LRp{\LRs{\begin{array}{c}
-%%\bm{V}_q \\ \bm{V}_f\end{array}}}^T \LRp{\diag{\bm{f}_N}{\LRp{\bm{Q}_N-\bm{Q}_N^T}}{\bm{g}_N}} +  \bm{V}_f^T \diag{\bm{f}_f}\bm{B}^i \bm{g}_f}.  
-%%\end{align*}
-%%Since this holds for all vectors $\bm{v}$, we recover the matrix equation
-%\begin{align*}
-%\bm{M}\bm{u} &= {\frac{1}{2}\LRp{\LRs{\begin{array}{c}
-%\bm{V}_q \\ \bm{V}_f\end{array}}}^T \LRp{\diag{\bm{f}_N}{\LRp{\bm{Q}_N-\bm{Q}_N^T}}{\bm{g}_N}} +  \bm{V}_f^T \diag{\bm{f}_f}\bm{B}^i \bm{g}_f}.
-%\end{align*}
-%
-%The accuracy of this approximation depends on the degree of accuracy of the volume and surface quadrature rules.  
-%
-%The entropy conservative skew-symmetric formulation (\ref{eq:esdgSkew})
-%
-%\note{Fix this up: if IBP holds for $v\in P^M$, the approximation should be $O(h^M)$}. 
-%
-%
-%\note{Add in continuous formulation and note that skew-symmetry approximates it.}
-
-\section{Skew-symmetric entropy conservative formulations on mapped elements}
-\label{sec:skew2}
-
-We now construct skew-symmetric formulations on mapped elements.
-We assume some domain $\Omega$ is decomposed into non-overlapping elements $D^k$, such that $D^k$ is the image of the reference element $\hat{D}$ under an isoparametric mapping $\bm{\Phi}^k$.
-We define geometric change of variables terms ${G}^k_{ij}$ as scaled derivatives of reference coordinates $\hat{\bm{x}}$ w.r.t.\ physical coordinates $\bm{x}$
-\begin{gather}
-\pd{u}{x_i} = \sum_{ij} {G}^k_{ij}\pd{u}{\hat{x}_j}, \qquad {G}^k_{ij} = J^k\pd{\hat{x}_j}{{x}_i}, 
-\label{eq:geofacs}
-\end{gather}
-where $J^k$ is the determinant of the Jacobian of the geometric mapping on the element $D^k$.
-We also introduce the scaled outward normal components $n_iJ^k_f$, which can be computed in terms of (\ref{eq:geofacs}) and the reference normals $\hat{\bm{n}}$ on $\hat{D}$
-\begin{gather}
-n^k_i J^k_f = \sum_{j=1}^d G^k_{ij} \hat{{n}}_j.  
-\label{eq:normals}
-\end{gather}
-We also define $\bm{n}^k_i$ as the vector containing concatenated values of the scaled outward normals $n^k_iJ^k_f$ at surface quadrature nodes.
-For the remainder of the work, we assume that the mesh is watertight \cite{chan2018discretely} such that at all points on any internal face, the scaled outward normals $n^k_iJ^k_f$ on the two elements sharing this face are equal and opposite.
-
-%Derivatives with respect to physical coordinates on $D^k$ are computed in terms of a change of variables formula and geometric terms (\ref{eq:geofacs}).  
-%%For affine meshes, the scaled geometric terms ${G}^k_{ij}$ are constant over each element.
-It was shown in \cite{chan2017discretely} that one can define a physical differentiation matrix as a linear combination of reference differentiation matrices.
-%Let ${\bm{D}}^i_N$ denote the decoupled SBP operator on the reference element $\hat{D}$.
-Define the decoupled SBP operators $\bm{D}^i_k$ with respect to the physical coordinates on $D^k$ be defined as
-%\begin{align}
-%{\bm{D}}^i_k = \sum_{j=1}^d {G}^k_{ij}\hat{\bm{D}}^j_N.
-%\end{align}
-
-%\note{Maybe remove affinely mapped elements section.}  
-%
-%\subsection{Affinely mapped elements}
-%\label{sec:affine}
-%
-%We first extend this formulation to affinely mapped elements.  Let the domain be 
-%
-%Let Assumption~\ref{ass:quad} again hold for $v(\bm{x}) = 1$.  Then, it is straightforward to show that the following formulation on $D^k$ is locally entropy conservative: 
-%\begin{gather}
-%J^k\bm{M}\pd{\bm{u}_h}{t} + 
-%\sum_{i=1}^d \LRs{\begin{array}{cc}
-%\bm{V}_q \\
-%\bm{V}_f\end{array}}^T \LRp{\LRp{\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T} \circ \bm{F}_S}\bm{1} + \bm{V}_f^T\bm{W}_f \diag{{\bm{n}^k_i}}\bm{f}_i^* = 0, \label{eq:skewform}\\
-%%\sum_{i=1}^d \LRs{\begin{array}{cc}
-%%\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{{\bm{n}_i}}\bm{f}_i^* = 0, \label{eq:skewform}\\
-%\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q, \nonumber\\
-%\bm{f}^* = \bm{f}_S(\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f), \qquad \text{ on interior interfaces,} \nonumber
-%\end{gather}
-%where $\tilde{\bm{u}}_f^+$ denotes the face value of the entropy-projected conservative variables $\tilde{\bm{u}}_f$ on the neighboring element.
-The proof of global entropy conservation follows as in \cite{chan2017discretely}, and results from summing up (\ref{eq:esdgthm}) over all elements and noting that the surface terms cancel due to the symmetry and conservation properties of the Tadmor flux (\ref{eq:esflux}).
-The entropy conservative formulation (\ref{eq:skewform}) can be made entropy stable by adding appropriate interface dissipation, such as Lax-Friedrichs or matrix-based penalization terms \cite{winters2017uniquely, chen2017entropy, chan2017discretely}.
-
-%On affine meshes, it is possible to show entropy stability of the skew-symmetric formulation (\ref{eq:esdgSkew}) under a surface quadrature which is only exact for degree $N$ polynomials.  
-On a single element (and on affine meshes), it is possible to show entropy stability of the skew-symmetric formulation (\ref{eq:esdgSkew}) under a surface quadrature which is only exact for degree $N$ polynomials.
-However, on curved meshes, stronger conditions are required to guarantee entropy stability.
-This is due to the fact that the geometric terms are now high order polynomials which vary spatially over each element.
-Moreover, Corollary~\ref{lemma:sbpcor} assumes affine geometric mappings, and does not hold on curved elements.
-In this section, we discuss how to extend Corollary~\ref{lemma:sbpcor} to curved simplicial and tensor product elements.
-
-\subsection{Curved elements and the geometric conservation law}
-\label{sec:curved}
-
-%Let the domain now be triangulated by a  consisting of curved elements $D^k$.  
-In this section, we describe how to construct appropriate decoupled SBP operators on curved meshes, and give conditions on the volume and surface quadrature rules under which a semi-discretely entropy stable scheme can be constructed.
-
-%For a curved element $D^k$, geometric terms $G^k_{ij}$ are high order polynomials over the reference domain $\hat{D}$.
-
-We first show how to construct decoupled SBP operators on curved elements.
-Let $\bm{G}^k_{ij}$ denote the vector of scaled geometric terms ${G}^k_{ij}$ evaluated at both volume and surface quadrature points, and let $\hat{\bm{D}}^j_N$ denote the decoupled SBP operator for the $j$th reference coordinate.
-Decoupled SBP operators on a curved element $D^k$ can be defined as in \cite{chan2018discretely} by
-\begin{equation}
-\bm{D}^i_k = \frac{1}{2}\sum_{j=1}^d \LRp{\diag{\bm{G}^k_{ij}}\hat{\bm{D}}^j_N + \hat{\bm{D}}^j_N\diag{\bm{G}^k_{ij}}}, \qquad \bm{Q}^i_k = \bm{W} \bm{D}^i_k.
-\label{eq:dncurved}
-\end{equation}
-It can be shown that, if $\hat{\bm{D}}^j_N$ satisfies a strong summation by parts property, $\bm{D}^i_k$ satisfies an analogous strong summation by parts property.
-
-For curved elements, steps must also be taken to ensure that the analogue of Lemma~\ref{lemma:vdsbp} and Corollary~\ref{lemma:sbpcor} hold at the discrete level.
-Expanding out the condition $\bm{Q}^i_k\bm{1} = \bm{0}$ in terms of (\ref{eq:dncurved}) yields
-\begin{align}
-\bm{Q}^i_k \bm{1} = \frac{1}{2} \bm{W}\sum_{j=1}^d \diag{\bm{G}^k_{ij}}\hat{\bm{D}}^j_N \bm{1} + \hat{\bm{D}}^j_N\diag{\bm{G}^k_{ij}}\bm{1} = \frac{1}{2}\bm{W}\sum_{j=1}^d \hat{\bm{D}}^j_N\LRp{\bm{G}^k_{ij}} = 0,
-\label{eq:dgcl}
-\end{align}
-where we have used that $\hat{\bm{D}}^j_N \bm{1} = 0$.
-We refer to the condition $\bm{Q}^i_k\bm{1} = 0$ as the discrete geometric conservation law (GCL) \cite{thomas1979geometric, kopriva2006metric}.
-For degree $N$ isoparametric mappings, the GCL is automatically satisfied in two dimensions due to the fact that the exact geometric terms ${G}^k_{ij}$ are polynomials of degree $N$ \cite{kopriva2006metric}.
-However, in three dimensions, the GCL is not automatically satisfied due to the fact that the degree of $G^k_{ij}$ is larger than $N$.
-Thus, the geometric terms cannot be represented exactly using degree $N$ polynomials, and (\ref{eq:dgcl}) must be enforced through an alternative construction of ${G}^k_{ij}$.
-
-A common approach is to rewrite the geometric terms as the curl of some quantity $\bm{r}^i$, but to interpolate $\bm{r}^i$ before applying the curl \cite{visbal2002use, kopriva2006metric, hindenlang2012explicit, chan2018discretely}:
-\begin{gather}
-\bm{r}^i = { \pd{\bm{x}}{\hat{x}_i}\times \bm{x}}, \qquad
-\LRs{\begin{array}{c}
-{G}^k_{i1}\\
-{G}^k_{i2}\\
-{G}^k_{i3}\end{array}} = -\frac{1}{2}\LRp{\pd{I_{N_{\rm geo}}\bm{r}^k}{\hat{x}_j}-\pd{I_{N_{\rm geo}}\bm{r}^j}{\hat{x}_k}},\\
-N_{\rm geo} \leq \begin{cases}
-N+1 & \text{(tetrahedra)}\\
-N & \text{(hexahedra)}
-\end{cases},\nonumber
-\label{eq:iconscurl}
-\end{gather}
-where $I_{N_{\rm geo}}$ denotes a degree $N_{\rm geo}$ polynomial interpolation operator with appropriate interpolation nodes.\footnote{This interpolation step must be performed using interpolation points with an appropriate number of nodes on each boundary \cite{chan2018discretely}.
-These include, for example, GLL nodes on tensor product elements, as well as Warp and Blend nodes on non-tensor product elements \cite{warburton2006explicit, chan2015comparison}.}  The restriction on the maximum value of $N_{\rm geo}$ ensures that $G^k_{ij} \in V^N$ (e.g.
-$G^k_{ij} \in P^N$ on tetrahedral elements and $G^k_{ij}\in Q^N$ on hexahedral elements), which is necessary to guarantee (\ref{eq:dgcl}).
-
-Because the decoupled SBP operators $\bm{D}^i_k$ are now defined through (\ref{eq:dncurved}), Corollary~\ref{lemma:sbpcor} and the proof of entropy stability no longer hold for curved elements and must be modified.
-The introduction of curvilinear meshes will impose slightly different conditions on the accuracy of the surface quadrature.
-We discuss simplicial and tensor product elements separately, as differences in the natural polynomial approximation spaces will result in different assumptions for each proof.
-
-\begin{lemma}{(Discrete FTC and exact constant differentiation on curved elements)}
-\label{lemma:vdsbpcurved} 
-Let $D^k$ be a curved element, and let the geometric terms $G^k_{ij}$ be constructed using (\ref{eq:iconscurl}).
-Let Assumption~\ref{ass:quad} hold for $v = 1$ and $v = G^k_{ij}$ for all $i,j = 1,\ldots, d$, and let $\bm{u}$ denote a vector of values of some function at volume and surface quadrature points.
-Then,
-%If the volume quadrature is exact for polynomials of degree $(N+N_{\rm geo}-1)$ and the surface quadrature is exact for polynomials of degree $(N+N_{\rm geo})$, then 
-\[
-\bm{1}^T\bm{Q}^i_k\bm{u} = \bm{1}^T\bm{B}^i_k\bm{u}, \qquad \bm{Q}^i_k\bm{1} = \bm{0},
-\]
-where the physical boundary matrix $\bm{B}^i_k$ is defined as
-\[
-\bm{B}^i_k = \sum_{j=1}^d \bm{B}^j_N \diag{\bm{G}^k_{ij}} = \sum_{j=1}^d \diag{\bm{G}^k_{ij}}\bm{B}^j_N  = \begin{pmatrix}
-\bm{0}&\\
-& \bm{W}_f \diag{\bm{n}^k_i}\end{pmatrix}.
-\]
-\end{lemma}
-\begin{proof}
-The proof of $\bm{Q}^i_k\bm{1} = \bm{0}$ was shown for for tensor product elements in \cite{kopriva2006metric} and for simplicial elements in \cite{chan2018discretely}, and relies only on the fact that $G^k_{ij} \in V^N$.
-
-The discrete fundamental theorem of calculus $\bm{1}^T\bm{Q}^i_k\bm{u} = \bm{1}^T\bm{B}^i_k\bm{u}$ can be shown to hold using (\ref{eq:dncurved}):
-\begin{align*}
-\bm{1}^T\bm{Q}^i_k\bm{u} &= \frac{1}{2}\sum_{j=1}^d \bm{1}^T{\LRp{\diag{\bm{G}^k_{ij}}\hat{\bm{Q}}^j_N + \hat{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}}} \bm{u} \\
-&= \frac{1}{2}\sum_{j=1}^d \LRp{\LRp{\bm{G}^k_{ij}}^T \hat{\bm{Q}}^j_N\bm{u} + \bm{1}^T\hat{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}\bm{u}}.
-\end{align*}
-Applying Lemma~\ref{lemma:vdsbp} to $\sum_{j=1}^d\bm{1}^T\hat{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}\bm{u}$ yields
-\begin{align*}
-\sum_{j=1}^d\bm{1}^T\hat{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}\bm{u} &= \sum_{j=1}^d\bm{1}^T\LRp{\bm{B}^j_N - \LRp{\hat{\bm{Q}}^j_N}^T}\diag{\bm{G}^k_{ij}}\bm{u} \\
-&= \sum_{j=1}^d\bm{1}^T\bm{B}^j_N \diag{\bm{G}^k_{ij}}\bm{u},
-\end{align*}
-where we have used Lemma~\ref{lemma:sbpcor} to conclude that $\hat{\bm{Q}}^j_N \bm{1} = 0$.
-Repeating the procedure for $\sum_{j=1}^d \LRp{\bm{G}^k_{ij}}^T \hat{\bm{Q}}^j_N\bm{u}$ yields
-\begin{align*}
-\sum_{j=1}^d \LRp{\bm{G}^k_{ij}}^T \hat{\bm{Q}}^j_N\bm{u} &= \sum_{j=1}^d \LRp{\bm{G}^k_{ij}}^T \LRp{\bm{B}^j_N - \LRp{\hat{\bm{Q}}^j_N}^T}\bm{u} \\
-&= \sum_{j=1}^d \LRp{\bm{G}^k_{ij}}^T \bm{B}^j_N \bm{u} - \sum_{j=1}^d \bm{u}^T{\hat{\bm{Q}}^j_N}{\bm{G}^k_{ij}} = \sum_{j=1}^d \LRp{\bm{G}^k_{ij}}^T \bm{B}^j_N \bm{u},  
-\end{align*}
-where we have used that the $G^k_{ij}$ satisfies the discrete GCL.  The proof is completed by noting that $\bm{B}^j_N$ is diagonal and thus commutes with $\diag{\bm{G}^k_{ij}}$, such that
-\[
-\bm{1}^T\bm{Q}^i_k\bm{u} = \sum_{j=1}^d\bm{1}^T\LRp{\bm{B}^j_N \diag{\bm{G}^k_{ij}}}\bm{u} = \bm{1}^T\bm{B}^i_k \bm{u}.
-\]
-\end{proof}
-
-The property $\bm{Q}^i_k\bm{1} = 0$ is also related to free-stream preservation \cite{kopriva2006metric}.
-Constant solutions are stationary solutions of systems of conservation laws.
-However, on curved meshes, the presence of spatially varying geometric terms can result in the production of spurious transient waves.
-The construction of geometric terms through (\ref{eq:iconscurl}) guarantees that the resulting methods are free-stream preserving, and that constant solutions remain stationary solutions of discretizations of (\ref{eq:nonlineqs}).
-
-We can now construct and prove entropy conservation and free stream preservation for a skew-symmetric formulation on a curved element $D^k$.
-Let $\bm{Q}^i_k$ be given by (\ref{eq:dncurved}), and define the curved mass matrix $\bm{M}^k = \bm{V}_q^T\bm{W}\diag{\bm{J}^k}\bm{V}_q$.
-Note that $\bm{M}^k$ is positive-definite so long as $J^k$ is positive at all quadrature points.
-We define the auxiliary quantities $\tilde{\bm{u}}$
-\begin{gather*}
-\bm{v}_q = \bm{v}\LRp{\bm{u}_q}, \qquad \tilde{\bm{v}} = \begin{bmatrix}
-\bm{V}_q\\
-\bm{V}_f
-\end{bmatrix}\bm{P}^k_q\bm{v}_q, \qquad \tilde{\bm{u}} = \bm{u}\LRp{\tilde{\bm{v}}}.
-\end{gather*}
-where $\bm{P}^k_q = \LRp{\bm{M}^k}^{-1}\bm{V}_q^T\bm{W}\diag{\bm{J}^k}$.  Then, we have the following theorem:
-\begin{theorem}
-\label{thm:skewformcurved}
-Let the geometric terms $G^k_{ij}$ be computed from (\ref{eq:iconscurl}), let the scaled outward normals $\bm{n}^k_i$ be computed pointwise from (\ref{eq:normals}), and suppose Assumption~\ref{ass:quad} holds for $v = 1$ and $v= G^k_{ij}$.
-Let $\tilde{\bm{u}}_f^+$ denote the face value of the entropy-projected conservative variables $\tilde{\bm{u}}_f$ on the neighboring element.
-Then, the formulation
-\begin{gather}
-\bm{M}^k\pd{\bm{u}_h}{t} + 
-\sum_{i=1}^d \LRs{\begin{array}{cc}
-\bm{V}_q \\
-\bm{V}_f\end{array}}^T \LRp{\LRp{\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T} \circ \bm{F}^i_S}\bm{1} + \bm{V}_f^T \bm{W}_f \diag{{\bm{n}^k_i}}\bm{f}_i^* = 0, \label{eq:skewformcurved}\\
-%\sum_{i=1}^d \LRs{\begin{array}{cc}
-%\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{{\bm{n}_i}}\bm{f}_i^* = 0, \label{eq:skewform}\\
-\LRp{\bm{F}^i_S}_{ij} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q, \nonumber\\
-\bm{f}_i^* = \bm{f}^i_S(\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f), \qquad \text{ on interior interfaces,} \nonumber
-\end{gather}
-is semi-discretely entropy conservative on $D^k$ such that for $\bm{u}_q = \bm{V}_q\bm{u}$,
-\begin{gather*}
-\bm{1}^T\bm{W}\diag{\bm{J}^k}\pd{U(\bm{u}_q)}{t} + \sum_{i=1}^d\bm{1}^T\bm{W}_f \diag{\bm{n}^k_i} \LRp{\psi_i(\tilde{\bm{u}}_f) - \tilde{\bm{v}}_f^T\bm{f}_i^*} = 0.
-\end{gather*}
-Additionally, the method is free-stream preserving such that $\pd{\bm{u}_h}{t} = 0$ for constant solutions.
-\end{theorem}
-\begin{proof}
-The proof of entropy stability is identical to that of Theorem~\ref{thm:esdg}, except that Lemma~\ref{lemma:vdsbpcurved} is used instead of Lemma~\ref{lemma:sbpcor}.
-However, the proof of free-stream preservation differs from proofs found in the literature, which typically show that the right hand side is zero for $\bm{u}_h$ constant \cite{crean2018entropy, chan2018discretely}.
-This strategy for is appropriate for ``strong'' formulations, but must be modified for the skew-symmetric formulation (\ref{eq:skewformcurved}).
-
-We have that $\bm{f}_i^* = \bm{f}^i_S(\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f) = \bm{f}_i(\bm{u}_f)$ using the consistency property of the Tadmor flux (\ref{eq:esflux}) and the fact that the solution is globally constant.
-For $\bm{u}_h$ constant, $\bm{F}_S$ is also constant.
-We assume (without loss of generality) that $\bm{F}_S$ and $\bm{f}_i(\bm{u}_f)$ are the matrix and vector of ones, respectively.
-Then,
-\[
-\LRp{\LRp{\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T} \circ \bm{F}_S}\bm{1} = \LRp{\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T}\bm{1} = \LRp{\bm{Q}^i_k}^T\bm{1},
-\]
-where we have used that $\bm{Q}^i_k\bm{1} = 0$ by Lemma~\ref{lemma:vdsbpcurved}.  Thus, the right hand side reduces to 
-\begin{align*}
-&\sum_{j=1}^d \LRs{\begin{array}{cc}
-\bm{V}_q \\
-\bm{V}_f\end{array}}^T \LRp{-\LRp{\bm{Q}^i_k}^T + \bm{V}_f^T \bm{W}_f \diag{{\bm{n}^k_i}}}\bm{1}\\
-&= \sum_{j=1}^d \LRs{\begin{array}{cc}
-\bm{V}_q \\
-\bm{V}_f\end{array}}^T \LRp{\LRp{ -\frac{1}{2}\LRp{\diag{\bm{G}^k_{ij}}\hat{\bm{Q}}^j_N + \hat{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}}^T + \hat{\bm{B}}^j_N \diag{\bm{G}^k_{ij}}}}\bm{1},
-%&= \sum_{j=1}^d \LRs{\begin{array}{cc}
-%\bm{V}_q \\
-%\bm{V}_f\end{array}}^T  -\frac{1}{2}\LRp{\hat{\bm{Q}}^j_N}^T{\bm{G}^k_{ij}} + 
-\end{align*}
-where we have used the definition of $\bm{Q}^i_k$ in (\ref{eq:dncurved}).  Rearranging terms gives that, for a constant solution, (\ref{eq:skewformcurved}) reduces to
-\begin{align*}
-\bm{M}^k\pd{\bm{u}_h}{t} &+ \frac{1}{2} \sum_{j=1}^d \LRs{\begin{array}{cc}
-\bm{V}_q \\
-\bm{V}_f\end{array}}^T \LRp{-\LRp{\hat{\bm{Q}}^j_N}^T +\hat{\bm{B}}^j_N}\bm{G}^k_{ij}\\
-&+ \frac{1}{2} \sum_{j=1}^d \LRs{\begin{array}{cc}
-\bm{V}_q \\
-\bm{V}_f\end{array}}^T\diag{\bm{G}^k_{ij}} \LRp{-\LRp{\hat{\bm{Q}}^j_N}^T +\hat{\bm{B}}^j_N}\bm{1} = 0.
-\end{align*}
-Let $u$ denote the constant value of the solution such that the values of the solution at quadrature points is $u\bm{1}$.
-Multiplying both sides by the constant solution $\bm{u}_h^T$ yields that $\frac{1}{2}\pd{}{t}\LRp{\bm{u}_h^T\bm{M}^k\bm{u}_h}$ is equal to
-\begin{align*}
--\frac{u}{2} \sum_{j=1}^d \LRp{\bm{1}^T \LRp{-\LRp{\hat{\bm{Q}}^j_N}^T +\hat{\bm{B}}^j_N}\bm{G}^k_{ij}
-+\LRp{\bm{G}^k_{ij}}^T \LRp{-\LRp{\hat{\bm{Q}}^j_N}^T +\hat{\bm{B}}^j_N}\bm{1}}.% = 0.
-\end{align*}
-By assumption, Lemma~\ref{lemma:vdsbp} holds for test functions $\bm{1}$ and $\bm{G}^k_{ij}$, and the formulation reduces to
-\begin{align*}
-\frac{1}{2}\pd{}{t}\LRp{\bm{u}_h^T\bm{M}^k\bm{u}_h} + \frac{u}{2} \sum_{j=1}^d \LRp{\bm{1}^T \hat{\bm{Q}}^j_N\bm{G}^k_{ij}
-+\LRp{\bm{G}^k_{ij}}^T \hat{\bm{Q}}^j_N\bm{1}} = 0.
-\end{align*}
-By construction, $\hat{\bm{Q}}^j_N\bm{1} = 0$, and $\sum_{j=1}^d \bm{1}^T \hat{\bm{Q}}^j_N\bm{G}^k_{ij} = 0$ since $G^k_{ij}$ satisfies the GCL.
-This implies that the latter term vanishes and $\pd{}{t}\LRp{\bm{u}_h^T\bm{M}^k\bm{u}_h} = 0$ on $D^k$.
-Since $\bm{M}^k$ is symmetric and positive-definite, $\bm{u}_h^T\bm{M}^k\bm{u}_h$ is a norm on $\bm{u}_h$.
-Summing $\pd{}{t}\LRp{\bm{u}_h^T\bm{M}^k\bm{u}_h} = 0$ over all elements ensures that globally constant solutions are stationary and that $\pd{\bm{u}_h}{t} = 0$.
-\end{proof}
-
-The proof of global entropy conservation follows from summing up (\ref{eq:esdgthm}) over all elements and noting that the surface terms cancel due to the symmetry and conservation properties of the Tadmor flux (\ref{eq:esflux}) \cite{chan2017discretely}.
-The entropy conservative formulations presented in this work can be made entropy stable by adding appropriate interface dissipation, such as Lax-Friedrichs or matrix-based penalization terms \cite{winters2017uniquely, chen2017entropy, chan2017discretely}.
-
-\begin{remark}
-It is also possible to replace the curved mass matrix $\bm{M}^k$ with a more easily invertible weight-adjusted approximation while maintaining high order accuracy, entropy stability, and local conservation \cite{chan2018discretely}.
-This approximation avoids the inversion of dense weighted $L^2$ mass matrices $\bm{M}^k$ on curved simplicial elements.
-On tensor product elements, this approximation is generally unnecessary as common choices of volume quadrature result in a diagonal (lumped) mass matrix \cite{carpenter2014entropy, parsani2016entropy, chan2018efficient}.
-\end{remark}
-
-\subsection{On quadrature conditions for Assumption~\ref{ass:quad} for $v = 1$ and $v = G^k_{ij}$}
-
-Semi-discrete entropy conservation on curved meshes requires that Assumption~\ref{ass:quad} holds for $v = 1$ and $v = G^k_{ij}$.
-We discuss specific choices of volume and surface quadrature on for which this assumption is valid, and summarize the maximum degree $N_{\rm geo}$ of the polynomial geometric approximation under which entropy stability holds for common choices of volume and surface quadrature.
-%\note{Emphasize we're interested in taking $N_{\rm geo}$ as large as possible for a given quadrature rule, and that there is a focus on degree $(2N-1)$ quadrature rules.}
-
-\subsubsection{Simplicial elements}
-
-On simplicial elements, for $u \in P^N$, $\pd{u}{\hat{x}_j}\in P^{N-1}$, and for $N_{\rm geo}\leq (N+1)$, $G^k_{ij}\in P^{N_{\rm geo}-1}$.
-Thus, the integrands in Assumption~\ref{ass:quad} $\pd{u}{\hat{x}_j} v \in P^{N+N_{\rm geo}-1}$ and $uv n_i \in P^{N+N_{\rm geo}}$ for $v = G^k_{ij}$.
-A volume quadrature rule of degree $(2N-1)$ and surface quadrature of degree $2N$ are then sufficient to ensure semi-discrete entropy conservation for $N_{\rm geo}\leq N$.
-If the volume quadrature is instead exact for degree $2N$ polynomials, then entropy conservation is guaranteed up to $N_{\rm geo} \leq N+1$.
-%\note{Add discussion on triangular elements.}
-
-\subsubsection{Tensor product elements}
-
-%Tensor product $(N+1)$-point GLL quadratures are a common choice for quadrilateral and hexahedral elements, and are exact for polynomials of degree $(2N-1)$ in each coordinate direction.  
-It was shown in Section~\ref{sec:assump1} that tensor product quadratures of degree $(2N-1)$ satisfy Assumption~\ref{ass:quad} holds for $v = 1$.
-However, in contrast to the simplicial case, it is not immediately clear that degree $(2N-1)$ volume quadratures exactly integrate $\int_{\hat{D}} \pd{u}{\hat{x}_j}v$ for $v = G^k_{ij}$ for tensor product elements.
-The difference between simplicial and tensor product elements is the polynomial space in which the derivative lies.
-In contrast to the simplicial case, if $u \in Q^N$, $\pd{u}{\hat{x}_j} \not\in Q^{N-1}$.
-Consider the three-dimensional case with $u, v \in Q^N$ and $i = 1$.
-Then, differentiation reduces the polynomial degree in one coordinate but not others and $\pd{u}{\hat{x}_1} \in Q^{N-1,N,N}$.
-As a result, $\pd{u}{\hat{x}_j}v \not\in Q^{2N-1}$, and a tensor product quadrature of degree $(2N-1)$ does not exactly integrate $\int_{\hat{D}} \pd{u}{\hat{x}_j}v$ for general $v\in Q^N$.
-
-We first consider the two dimensional case, and show that Assumption~\ref{ass:quad} holds with $v = G^k_{ij}$ for all $N_{\rm geo} \leq N$.  On quadrilateral elements, $G^k_{ij}$ is
-\begin{align*}
-G^k_{11} &= \pd{x_2}{\hat{x}_2} \in Q^{N_{\rm geo}, N_{\rm geo}-1}, \qquad G^k_{12} = -\pd{x_2}{\hat{x}_1} \in Q^{N_{\rm geo}-1, N_{\rm geo}},\\
-G^k_{21} &= -\pd{x_1}{\hat{x}_2} \in Q^{N_{\rm geo}, N_{\rm geo}-1}, \qquad G^k_{22} = \pd{x_1}{\hat{x}_1}\in Q^{N_{\rm geo}-1, N_{\rm geo}}.
-\end{align*}
-Since $\pd{u}{\hat{x}_1} \in Q^{N-1,N}$ and $\pd{u}{\hat{x}_2} \in Q^{N,N-1}$
-\[
-\pd{u}{\hat{x}_i}G^k_{ij} \in Q^{N+N_{\rm geo}-1} \subset Q^{2N-1}, \qquad \forall N_{\rm geo}\leq N.
-\]
-Thus, any quadrature rule of degree $(2N-1)$ exactly integrates $\int_{\hat{D}}\pd{u}{\hat{x}_i}v$ for $v = G^k_{ij}$.
-
-We now consider the condition in Assumption~\ref{ass:quad} on the surface integrals $\int_{\partial \hat{D}} u v n_i$ for $v= G^k_{ij}$.  \note{Finish}
-
-We now consider the three-dimensional case.
-Unlike quadrilateral elements, Assumption~\ref{ass:quad} does not hold for all $N_{\rm geo} \leq N$ on general curved hexahedra.
-Expanding out (\ref{eq:iconscurl}), one can show that $G^k_{ij} \in Q^{N_{\rm geo}}$ on hexahedral elements.
-Thus, if $N_{\rm geo} = N$ and $u\in Q^N$, $\pd{u}{\hat{x}_1} G^k_{i1} \in Q^{N-1,N,N}$.
-\note{Finish}.
-
-\subsection{Sufficient conditions on quadrature and $N_{\rm geo}$ to guarantee entropy conservation}
-
-For convenience, this section summarizes conditions on the accuracy of volume and surface quadratures under which one can guarantee entropy conservation for a given degree $N_{\rm geo}$ of a geometric mapping.
-
- \note{finish}
-
-%% $G^k_{1j} = \pd{\hat{x}_j}{x_1}$, $j = 1,2,3$ yields
-%%\begin{align*}
-%%G^k_{11} &= \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_2}x_3}}{\hat{x}_3}- \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_3}x_3}}{\hat{x}_2} \in Q^{N_{\rm geo}}\\
-%%G^k_{12} &= \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_3}x_3}}{\hat{x}_1}- \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_3}x_3}}{\hat{x}_1} \in Q^{N_{\rm geo}}\\
-%%G^k_{13} &= \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_1}x_3}}{\hat{x}_2}- \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_2}x_3}}{\hat{x}_1} \in Q^{N_{\rm geo}}.
-%%%Dt*((Ds*y).*z) - Ds*((Dt*y).*z)
-%%\end{align*}
-%
-%%Lemma~\ref{lemma:vdsbpcurved} implies that the polynomial degree $R$ of the surface geometric terms must be compatible with the accuracy of the surface quadrature.
-This, in turn, depends on the polynomial degree $N_{\rm geo}$ to which geometric terms are approximated.
-\note{Kopriva trick to enforcing GCL still results in $Q^N$ normals 3D.
-May not satisfy requirement that $\bm{1}^T\bm{Q}^i_N \psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i_N \psi_i(\tilde{\bm{u}})$ because of quadrature inaccuracy.
-Note - this still works regardless for conforming hexes because of the full matrix SBP property.}
-%
-%
-%\begin{lemma}
-%\label{lemma:curlPN}
-%Let $\bm{u} \in \mathbb{R}^3$ be represented on a hexahedron by polynomials of the form
-%\begin{gather*}
-%\bm{u} = \begin{pmatrix}u\\v\\w\end{pmatrix}, \qquad u \in Q^{N-1,N,N}, \qquad v \in Q^{N,N-1,N}, \qquad w \in Q^{N,N,N-1}.
-%\end{gather*}
-%Then 
-%\[
-%\Grad \times \bm{u} \in \begin{pmatrix} Q^{N,N-1,N-1}\\Q^{N-1,N,N-1}\\Q^{N-1,N-1,N}\end{pmatrix}
-%\]
-%and traces of $\LRl{\bm{u}}\cdot\bm{n} \in Q^{N-1}$ (\note{CHECK THIS}).
-%\end{lemma}
-%\begin{proof}
-%The proof is a straightforward computation
-%\end{proof}
-%
-%We summarize the
-
-%\section{A mortar-based implementation}
-%
-%For hybrid meshes, the choice of surface quadrature for an interface may vary depending on the neighboring elements.
-Consider, for example, a mixed triangular-quadrilateral mesh where the quadrilaterals utilize GLL quadrature on each face.
-While te skew-symmetric formulation allows for the stable coupling of quadrilaterals and triangles using a GLL face quadrature, one may wish to utilize more accurate Gauss face quadratures for triangle-triangle interfaces.
-To accomodate arbitrary combinations of face quadrature, we introduce a mortar-based implementation.
-%
-%We assume that each element 
-%
-%\note{Finish - figure out notation for face interp!}
-
-\section{Numerical experiments}
-\label{sec:num}
-
-In this section, we present two-dimensional experiments which verify the theoretical results presented.  We first investigate triangular and quadrilateral meshes, and then   
-%on hybrid meshes consisting of quadrilaterals and triangles, as well as on non-conforming meshes of quadrilateral elements.
-
-\subsection{Triangular meshes}
-
-\note{Affine and curved triangles with GLL surface quadrature (under-integrated).}
-
-\subsection{Quadrilateral meshes}
-
-\note{ GLL quadrilaterals with Gauss surface quadrature.  Explain that for GLL quadratures, the decoupled SBP property doesn't hold when Gauss points are used.  }
-
-\subsection{Hybrid quadrilateral-triangular meshes}
-
-It is not immediately clear that the outward normals as defined by (\ref{eq:normals}) will be equal and opposite across interfaces shared by elements of different types.
-Thus, we restrict ourselves to straight-sided meshes in this case, and will investigate the effect of curvilinear mappings on hybrid meshes in future work.
-
-\section{Conclusions}
-
-\note{Need mortar to .}
-
-\section{Acknowledgments}
-
-%\appendix
-%
-%\section{A modified Geometric terms}
-
-\bibliographystyle{unsrt}
-\bibliography{dg}
+Main notes: if $\bm{Q}_N$ does not satisfy the SBP property, then the error estimate is limited by the accuracy of the weak derivative.
+The strong derivative is always $O(h^p)$, but the weak derivative is $O\LRp{h^{\left\lfloor\alpha/2\right\rfloor}}$, where $\alpha$ is defined above.
 

### 2019-02-14 docs/skewform/main.tex

@@
 \author{Jesse Chan}
-\title{Skew-symmetric entropy stable discontinuous Galerkin formulations}% with applications to hybrid meshes}
+\title{Skew-symmetric entropy stable discontinuous Galerkin formulations with applications to hybrid meshes}
 \titlerunning{Skew-symmetric entropy stable DG formulations}
@@
 
-%\tableofcontents
+\tableofcontents
 
@@
 
-However, when applied to nonlinear conservation laws, high order methods can experience artificial growth and blow-up near under-resolved features such as shocks, turbulence, or boundary layers.
+However, when applied to nonlinear conservation laws, high order methods can experience artificial growth and blow-up near under-resolved features such as shocks or turbulence.
 In practice, the application of high order methods to practical problems requires shock capturing and stabilization techniques (such as artificial viscosity) or solution regularization (such as filtering or limiting) to prevent solution blow-up.
@@
 
-We are interested in addressing cases where the generalized SBP property (and as a result, the strong decoupled SBP property in Theorem~\ref{lemma:dsbp}) do not hold.
+When a generalized SBP property holds for $\bm{Q}^i$, entropy stability can be proven using the strong SBP property in Theorem~\ref{lemma:dsbp} \cite{chan2017discretely, chan2018discretely}.
+The focus of this work is to address cases where the generalized SBP property (and as a result, the SBP property in Theorem~\ref{lemma:dsbp}) do not hold.
 
@@
 \begin{equation}
-\bm{1}^T\bm{W}\pd{U(\bm{u}_q)}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}^i \LRp{\psi_i(\tilde{\bm{u}}_f) - \tilde{\bm{v}}_f^T\bm{f}_i^*} = 0, \qquad \bm{u}_q = \bm{V}_q\bm{u}.
+\bm{1}^T\bm{W}\pd{U(\bm{u}_q)}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}^i \LRp{\tilde{\bm{v}}_f^T\bm{f}_i^* - \psi_i(\tilde{\bm{u}}_f)} = 0, \qquad \bm{u}_q = \bm{V}_q\bm{u}.
 \label{eq:esdgthm}
@@
 \tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}^i_S}\bm{1} + \tilde{\bm{v}}_f^T \bm{B}^i\bm{f}_i^* = 0.
+\label{eq:entropyrhs}
 \end{align}
-One can show that \cite{chan2017discretely}
+Algebraic manipulations (see \cite{chan2017discretely} for details) yields that 
 \begin{align*}
@@
 
-A common approach is to rewrite the geometric terms as the curl of some quantity $\bm{r}^i$, but to interpolate $\bm{r}^i$ before applying the curl \cite{visbal2002use, kopriva2006metric, hindenlang2012explicit, chan2018discretely}:
+To ensure that the geometric terms satisfy the GCL, we first rewrite the geometric terms as the curl of some quantity $\bm{r}^i$, but interpolate $\bm{r}^i$ before applying the curl \cite{visbal2002use, kopriva2006metric, hindenlang2012explicit, chan2018discretely}:
 \begin{gather}
@@
 where $I_{N_{\rm geo}}$ denotes a degree $N_{\rm geo}$ polynomial interpolation operator with appropriate interpolation nodes.\footnote{This interpolation step must be performed using interpolation points with an appropriate number of nodes on each boundary \cite{chan2018discretely}.
-These include, for example, GLL nodes on tensor product elements, as well as Warp and Blend nodes on non-tensor product elements \cite{warburton2006explicit, chan2015comparison}.}  The restriction on the maximum value of $N_{\rm geo}$ ensures that $G^k_{ij} \in V^N$ (e.g.
+These include, for example, GLL nodes on tensor product elements, and optimized interpolation nodes on non-tensor product elements \cite{hesthaven1998electrostatics, warburton2006explicit, chan2015comparison}.}  The restriction on the maximum value of $N_{\rm geo}$ ensures that $G^k_{ij} \in V^N$ (e.g.
 $G^k_{ij} \in P^N$ on tetrahedral elements and $G^k_{ij}\in Q^N$ on hexahedral elements), which is also necessary to guarantee (\ref{eq:dgcl}).
@@
 The proof of entropy stability is identical to that of Theorem~\ref{thm:esdg}, except that Lemma~\ref{lemma:vdsbpcurved} is used instead of Lemma~\ref{lemma:sbpcor}.
-However, the proof of free-stream preservation differs from proofs found in the literature, which typically show that the right hand side is zero for $\bm{u}_h$ constant \cite{crean2018entropy, chan2018discretely}.
+However, proving free-stream preservation differs from approaches in the literature, which typically show that the right hand side is zero for $\bm{u}_h$ constant \cite{crean2018entropy, chan2018discretely}.
 This strategy for is appropriate for ``strong'' formulations, but must be modified for the skew-symmetric formulation (\ref{eq:skewformcurved}).
@@
 
+\section{Conditions on the accuracy of volume and surface quadrature rules}
+
+The previous sections outline minimal conditions under which entropy stability is guaranteed under a skew-symmetric formulation and a polynomial geometric mapping.
+In Section~\ref{sec:curvgeo}, we translate these minimial conditions into conditions on quadrature accuracy.
+
+Additionally, while it is possible to maintain entropy stability under reduced accuracy quadratures, the accuracy of the resulting schemes depend on the accuracy of the volume and surface quadrature rules.
+We analyze the degree of accuracy expected based on the accuracy of different quadrature rules in Section~\ref{sec:accskew}.
+
 \subsection{On quadrature conditions for Assumption~\ref{ass:quad} for $v = 1$ and $v = G^k_{ij}$}
 \label{sec:curvgeo}
+
 Semi-discrete entropy conservation on curved meshes requires that Assumption~\ref{ass:quad} holds for $v = 1$ and $v = G^k_{ij}$.
@@
 
-\subsubsection{Simplicial elements}
-
-On simplicial elements, for $u \in P^N$, $\pd{u}{\hat{x}_j}\in P^{N-1}$, and for $N_{\rm geo}\leq (N+1)$, $G^k_{ij}\in P^{N_{\rm geo}-1}$.
-Thus, the integrands in Assumption~\ref{ass:quad} $\pd{u}{\hat{x}_j} v \in P^{N+N_{\rm geo}-1}$ and $uv n_i \in P^{N+N_{\rm geo}}$ for $v = G^k_{ij}$.
-A volume quadrature rule of degree $(2N-1)$ and surface quadrature of degree $2N$ are then sufficient to ensure semi-discrete entropy conservation for $N_{\rm geo}\leq N$.
-However, that in order to ensure that the mass matrix is positive-definite in Assumption~\ref{ass:quad}, the volume quadrature must be degree $2N$ in general on simplices.
-If the volume quadrature is instead exact for degree $2N$ polynomials, then entropy conservation is guaranteed up to $N_{\rm geo} \leq N+1$.
-%\note{Add discussion on triangular elements.}
-
-\subsubsection{Tensor product elements}
-
-%Tensor product $(N+1)$-point GLL quadratures are a common choice for quadrilateral and hexahedral elements, and are exact for polynomials of degree $(2N-1)$ in each coordinate direction.  
+In order to ensure that the mass matrix is positive-definite in Assumption~\ref{ass:quad}, the volume quadrature must be degree $2N$ in general on simplices.
+The following lemma summarizes expected behavior for surface quadrature rules of varying order:
+\begin{lemma}
+Let $\hat{D}$ be a simplex with volume quadrature which is exact for degree $2N$ polynomials.
+Let the surface quadrature be exact for polynomials of degree $M+N$.
+Then, the skew-symmetric formulation (\ref{eq:skewformcurved}) is entropy stable for $N_{\rm geo} \leq \min\LRp{N+1,M+1}$.
+\label{lemma:curvsimp}
+\end{lemma}
+\begin{proof}
+Entropy stability holds if Assumption~\ref{ass:quad} holds for $v = 1$ and $v = G^k_{ij}$.  
+Simplicial elements require $N_{\rm geo}\leq (N+1)$ in order to guarantee that $G^k_{ij}\in P^{N_{\rm geo}-1} \subset P^{N}$, which is necessary to satisfy the discrete GCL \cite{chan2018discretely}.
+Then, for $u \in P^N$, $\pd{u}{\hat{x}_j}\in P^{N-1}$, the integrands in Assumption~\ref{ass:quad} are $\pd{u}{\hat{x}_j} v \in P^{N+N_{\rm geo}-2}$ and $uv n_i \in P^{N+N_{\rm geo}-1}$ for $v = G^k_{ij}$.
+The volume quadrature exactly integrates the first integrand for $N_{\rm geo} \leq N+1$, while the surface quadrature exactly integrates the second integrand for $M \geq N_{\rm geo}-1$, or $N_{\rm geo} \leq M+1$.
+\end{proof}
+
+The situation is more complicated for curved tensor product elements.
 It was shown in Section~\ref{sec:assump1} that tensor product quadratures of degree $(2N-1)$ satisfy Assumption~\ref{ass:quad} holds for $v = 1$.
@@
 Then, differentiation reduces the polynomial degree in one coordinate but not others and $\pd{u}{\hat{x}_1} \in Q^{N-1,N,N}$.
-As a result, $\pd{u}{\hat{x}_j}v \not\in Q^{2N-1}$, and a tensor product quadrature of degree $(2N-1)$ does not exactly integrate $\int_{\hat{D}} \pd{u}{\hat{x}_j}v$ for general $v\in Q^N$.
-
-We first consider the two dimensional case with volume quadrature of degree $(2N-1)$ and show that Assumption~\ref{ass:quad} holds with $v = G^k_{ij}$ for all $N_{\rm geo} \leq N$.
-On quadrilateral elements, $G^k_{ij}$ is
+As a result, $\pd{u}{\hat{x}_j}v \not\in Q^{2N-1}$, and a tensor product quadrature of degree $(2N-1)$ (in each coordinate) does not exactly integrate $\int_{\hat{D}} \pd{u}{\hat{x}_j}v$ for general $v\in Q^N$.
+
+We address the quadrilateral case first:
+\begin{lemma}
+Let $\hat{D}$ be a quadrilateral.
+Suppose the volume quadrature be exact for degree $M+N$ polynomials, and that the surface quadrature be exact for polynomials of degree $M+N$.
+Then, the skew-symmetric formulation (\ref{eq:skewformcurved}) is entropy stable for $N_{\rm geo} \leq \min\LRp{N,M+1}$.
+\label{lemma:curvquad}
+\end{lemma}
+\begin{proof}
+As in Lemma~\ref{lemma:curvsimp}, entropy stability holds if Assumption~\ref{ass:quad} holds for $v = 1$ and $v = G^k_{ij}$.
+The case of $v = 1$ was addressed previously, and we focus on $v = G^k_{ij}$.
+We first characterize the polynomial degree of the geometric terms $G^k_{ij}$.
+In contrast to the simplicial case, tensor product elements require $N_{\rm geo}\leq N$ in order to ensure that $G^k_{ij}\in Q^{N,N}$ and that the discrete GCL is satisfied \cite{kopriva2006metric}.
+On a quadrilateral element with a degree $N_{\rm geo}$ geometric mapping, $G^k_{ij}$ is
+% with volume quadrature of degree $(2N-1)$ and show that Assumption~\ref{ass:quad} holds with $v = G^k_{ij}$ for all $N_{\rm geo} \leq N$.  
 \begin{align*}
 G^k_{11} &= \pd{x_2}{\hat{x}_2} \in Q^{N_{\rm geo}, N_{\rm geo}-1}, \qquad G^k_{12} = -\pd{x_2}{\hat{x}_1} \in Q^{N_{\rm geo}-1, N_{\rm geo}},\\
-G^k_{21} &= -\pd{x_1}{\hat{x}_2} \in Q^{N_{\rm geo}, N_{\rm geo}-1}, \qquad G^k_{22} = \pd{x_1}{\hat{x}_1}\in Q^{N_{\rm geo}-1, N_{\rm geo}}.
+G^k_{21} &= -\pd{x_1}{\hat{x}_2} \in Q^{N_{\rm geo}, N_{\rm geo}-1}, \qquad G^k_{22} = \pd{x_1}{\hat{x}_1}\in Q^{N_{\rm geo}-1, N_{\rm geo}}.\nonumber
 \end{align*}
@@
 \[
-\pd{u}{\hat{x}_i}G^k_{ij} \in Q^{N+N_{\rm geo}-1} \subset Q^{2N-1}, \qquad \forall N_{\rm geo}\leq N.
-\]
-Thus, any quadrature rule of degree $(2N-1)$ exactly integrates $\int_{\hat{D}}\pd{u}{\hat{x}_i}v$ for $v = G^k_{ij}$.
+\pd{u}{\hat{x}_i}G^k_{ij} \in Q^{N+N_{\rm geo}-1}.% \subset Q^{2N-1}, \qquad \forall N_{\rm geo}\leq N.
+\]
+The volume quadrature exactly integrates this integrand for $M \geq N_{\rm geo}-1$.
 
@@
 Similarly, $uG^k_{i2} \hat{n}_2 \in Q^{N+N_{\rm geo}-1}$ along the top and bottom faces and are zero along the left and right faces.
-Thus, any surface quadrature rule exact for degree $N+N_{\rm geo}-1$ polynomials satisfies Assumption~\ref{ass:quad}, and for a degree $(2N-1)$ surface quadrature, entropy stability is guaranteed for any $N_{\rm geo}\leq N$.
+The surface quadrature rule exactly integrates such integrands for $M \geq N_{\rm geo} - 1$, or $N_{\rm geo} \leq M+1$.
+\end{proof}
+
+Existing proofs of entropy stability on quadrilaterals rely on $(N+1)$-point GLL volume and surface quadratures, which are exact for degree $2N-1$ polynomials.
+The novelty of Lemma~\ref{lemma:curvquad} is that the proof holds for any combination of degree $2N-1$ volume and surface quadratures (for example, $(N+1)$-point GLL volume quadrature and an $(N-1)$-point Gauss surface quadrature).
 
 We now consider the three-dimensional case.
-Unlike quadrilateral elements, Assumption~\ref{ass:quad} does not hold under degree $(2N-1)$ volume quadrature for $N_{\rm geo} = N$ on general curved hexahedra.
-Expanding out (\ref{eq:iconscurl}) gives
+In contrast to the quadrilateral case, the GCL is not automatically satisfied for a degree $N_{\rm geo} \leq N$ geometric mapping.
+Instead, GCL-satisfying geometric terms are approximated using (\ref{eq:iconscurl}).
+Expanding out the expression for $G^k_{11}$ gives
 \[
@@
 }
+We summarize these findings in the following lemma for hexahedral elements:
+\begin{lemma}
+Let $\hat{D}$ be a hexahedral element, with geometric terms constructed using (\ref{eq:iconscurl}).
+Let the volume quadrature be exact for degree $M+N$ polynomials, and let the surface quadrature be exact for polynomials of degree $M+N$.
+Then, the skew-symmetric formulation (\ref{eq:skewformcurved}) is entropy stable for $N_{\rm geo} \leq \min\LRp{N,M}$.
+\label{lemma:curvquad}
+\end{lemma}
 
@@
 \begin{enumerate}
-\item On quadrilateral elements, Assumption~\ref{ass:quad} holds for $N_{\rm geo} = N$ and any tensor product volume and surface quadratures of degree $(2N-1)$ 
-\item On hexahedral elements, Assumption~\ref{ass:quad} holds for $N_{\rm geo} = N-1$ and any tensor product volume and surface quadratures of degree $(2N-1)$.
-If the SBP property holds (for example, when the volume and surface quadratures are of degree $2N$) then Assumption~\ref{ass:quad} holds for $N_{\rm geo} = N$.
+\item On quadrilateral elements, Assumption~\ref{ass:quad} holds for $N_{\rm geo} \leq N$ and any tensor product volume and surface quadratures of degree $(2N-1)$ 
+\item On hexahedral elements, Assumption~\ref{ass:quad} holds for $N_{\rm geo} \leq N-1$ and any tensor product volume and surface quadratures of degree $(2N-1)$.
+If the SBP property holds (e.g.\ for GLL quadrature, or for volume and surface quadratures of degree $2N$) then Assumption~\ref{ass:quad} holds for $N_{\rm geo} \leq N$.
 \end{enumerate}
 
-We note that the condition $N_{\rm geo} = N-1$ is non-standard for tensor product elements.
+We note that the condition $N_{\rm geo} \leq N-1$ is non-standard for tensor product elements.
 However, this condition is only necessary for entropy stability when $\bm{Q}^i$ does not satisfy a generalized SBP property (see Remark~\ref{remark:dsbp}).
-To the author's knowledge, this setting has not been considered within the literature on tensor product elements.
-
-%\note{Finish}
-%\begin{remark}
-%We note that this work assumes that the same quadrature is used to volume contributions in each direction, and can be refined slightly 
-%\note{add line DG}
-
-% $G^k_{1j} = \pd{\hat{x}_j}{x_1}$, $j = 1,2,3$ yields
-%\begin{align*}
-%G^k_{11} &= \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_2}x_3}}{\hat{x}_3}- \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_3}x_3}}{\hat{x}_2} \in Q^{N_{\rm geo}}\\
-%G^k_{12} &= \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_3}x_3}}{\hat{x}_1}- \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_3}x_3}}{\hat{x}_1} \in Q^{N_{\rm geo}}\\
-%G^k_{13} &= \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_1}x_3}}{\hat{x}_2}- \pd{I_{\rm Ngeo} \LRp{\pd{x_2}{\hat{x}_2}x_3}}{\hat{x}_1} \in Q^{N_{\rm geo}}.
-%%Dt*((Ds*y).*z) - Ds*((Dt*y).*z)
-%\end{align*}
-
-%Lemma~\ref{lemma:vdsbpcurved} implies that the polynomial degree $R$ of the surface geometric terms must be compatible with the accuracy of the surface quadrature.
-This, in turn, depends on the polynomial degree $N_{\rm geo}$ to which geometric terms are approximated.
-\note{Kopriva trick to enforcing GCL still results in $Q^N$ normals 3D.
-May not satisfy requirement that $\bm{1}^T\bm{Q}^i_N \psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i_N \psi_i(\tilde{\bm{u}})$ because of quadrature inaccuracy.
-Note - this still works regardless for conforming hexes because of the full matrix SBP property.}
-
-%\begin{lemma}
-%\label{lemma:curlPN}
-%Let $\bm{u} \in \mathbb{R}^3$ be represented on a hexahedron by polynomials of the form
-%\begin{gather*}
-%\bm{u} = \begin{pmatrix}u\\v\\w\end{pmatrix}, \qquad u \in Q^{N-1,N,N}, \qquad v \in Q^{N,N-1,N}, \qquad w \in Q^{N,N,N-1}.
-%\end{gather*}
-%Then 
-%\[
-%\Grad \times \bm{u} \in \begin{pmatrix} Q^{N,N-1,N-1}\\Q^{N-1,N,N-1}\\Q^{N-1,N-1,N}\end{pmatrix}
-%\]
-%and traces of $\LRl{\bm{u}}\cdot\bm{n} \in Q^{N-1}$ (\note{CHECK THIS}).
-%\end{lemma}
-%\begin{proof}
-%The proof is a straightforward computation
-%\end{proof}
-%
-%We summarize the
+To the author's knowledge, this setting has not been considered within the literature.
 
 \subsection{On the accuracy of the skew-symmetric formulation}
-
-We note that, while it is possible to maintain entropy stability under reduced accuracy quadratures, the accuracy of the resulting schemes depend on the accuracy of the volume and surface quadrature rules.
-In this section, we analyze the degree of accuracy expected for volume and surface quadrature rules of a given accuracy.
+\label{sec:accskew}
 
@@
 Suppose that the volume quadrature is exact for degree $M+N-1$ polynomials on simplices, or for polynomials in $Q^{M+N-1,M+N,M+N}$ on tensor product elements.
-Furthermore, assume that the surface quadrature is exact for degree $M+N$ polynomials.
-Then, the skew-symmetric approximation of the $x$-derivative (\ref{eq:dsbpapprox}) is exact for all polynomials of degree $M$.
+Furthermore, assume that the surface quadrature is exact for degree $M+N$ polynomials on simplices and $Q^{M+N,M+N}$ on tensor product elements.
+Then, the skew-symmetric approximation of the derivative (\ref{eq:dsbpapprox}) is exact for polynomials of degree $M$.
 \label{lemma:dsbpapprox}
@@
 where $\bm{e}$ is a polynomial of degree $N$.  
-If Assumption~\ref{ass:quad} is satisfied and $u(\bm{x})$ is polynomial, then the values of $u$ at quadrature points are $\bm{u}_q = \bm{V}_q\bm{u}$ and $\bm{u} = \bm{P}_q\bm{u}_q$.
+If Assumption~\ref{ass:quad} is satisfied and $u(\bm{x})$ is polynomial, then the values of $u(\bm{x})$ at quadrature points are $\bm{u}_q = \bm{V}_q\bm{u}$ and $\bm{u} = \bm{P}_q\bm{u}_q$.
 This implies that $\bm{u}_f = \bm{V}_f \bm{P}_q\bm{u}_q =  \bm{E}\bm{u}_q$.
@@
 
-There exist several choices for entropy conservative fluxes \cite{ismail2009affordable, ranocha2018comparison,chandrashekar2013kinetic}.  For the following experiments, we utilize the
+There exist several choices for entropy conservative fluxes \cite{ismail2009affordable, ranocha2018comparison,chandrashekar2013kinetic}.  We utilize the
 the entropy conservative numerical fluxes given by Chandrashekar in \cite{chandrashekar2013kinetic}
@@
 \begin{gather*}
-p_{\rm avg} = \frac{\avg{\rho}}{2\avg{\beta}}, \qquad E_{\rm avg} = \frac{\avg{\rho}^{\log}}{2\avg{\beta}^{\log}\LRp{\gamma -1}}   + \frac{\nor{\bm{u}}^2_{\rm avg}}{2}, \\
+p_{\rm avg} = \frac{\avg{\rho}}{2\avg{\beta}}, \qquad E_{\rm avg} = \frac{\avg{\rho}^{\log}}{2\avg{\beta}^{\log}\LRp{\gamma -1}}   + \frac{\nor{\bm{u}}^2_{\rm avg}}{2}, \qquad  \beta = \frac{\rho}{2p},\\
  \nor{\bm{u}}^2_{\rm avg} = 2(\avg{u}^2 + \avg{v}^2) - \LRp{\avg{u^2} +\avg{v^2}} \nonumber.
 \end{gather*}
+From here on, \emph{entropy conservative} refers to a scheme which uses these entropy conservative fluxes at inter-element interfaces.
+We will refer to schemes which add interface dissipation as \emph{entropy stable}.
+In this work, we utilize a local Lax-Friedrichs interface dissipation.
 
@@
 
+A low storage 4th order Runge-Kutta scheme \cite{carpenter1994fourth} is used for all numerical experiments.
+The time-step is estimated based on formulas derived for linear advection in \cite{chan2015gpu, chan2018multi}
+\[
+dt = C \frac{h}{c_{\max} \max\LRc{\frac{1}{2}C_T, C_I}}
+\]
+where $h$ is the mesh size, $c_{\max}$ is the maximum wave-speed, $C$ is a user-defined CFL constant, and $C_T, C_I$ are constants in finite element inverse and trace inequalities.
+These constants scale proportionally to $N^2$, though precise values of $C_I, C_T$ vary slightly depending on the choice of volume or surface quadrature used.
+The dependence of $C_I, C_T$ on quadrature is discussed in more detail in Appendix~\ref{sec:consts}, where computed values of $C_I, C_T$ are also given.
+
 \subsection{Choices of volume and surface quadrature considered}
@@
 \begin{enumerate}
+\item $(N+1)$ point GLL volume quadrature, $(N+1)$ point GLL surface quadrature.
+\item $(N+1)$ point GLL volume quadrature, $(N+1)$ point Gauss surface quadrature,
 \item $(N+1)$ point Gauss quadrature, $(N+1)$ point Gauss surface quadrature,
-\item $(N+1)$ point GLL volume quadrature, $(N+1)$ point Gauss surface quadrature,
-\item $(N+1)$ point GLL volume quadrature, $(N+1)$ point GLL surface quadrature.
 \end{enumerate}
@@
 
-All options result in similar computational costs on on triangles, but slight differences in computational cost and complexity on quadrilaterals.
+All three options result in similar computational costs on on triangles, but slight differences in computational cost and complexity on quadrilaterals.
 On quadrilaterals, Option~\ref{opt:1} is the most computationally efficient option, as the formulation (\ref{thm:skewformcurved}) reduces to a standard entropy stable DG-SEM \cite{gassner2016split} or spectral collocation method \cite{carpenter2014entropy}.
-Option~\ref{opt:2} is slightly more expensive, and requires interpolating the solution from GLL to Gauss nodes on the boundary, as well as computing two-point flux interactions between boundary GLL and Gauss nodes.
-Option 3 is most involved, and results in a Gauss collocation method on quadrilaterals, as described in \cite{chan2018efficient}, and requires both interpolation and two-point flux interactions between lines of volume quadrature nodes and boundary nodes.
-However, Option~\ref{opt:3} provides some advantages over Options~\ref{opt:1} and \ref{opt:2} in terms of accuracy and stable timestep restrictions.
-
-\subsection{Choice of quadrature and timestep restrictions}
-
-We first investigate how the maximum stable timestep depends on specific choices of volume and surface quadrature.
-The dependence of timestep on quadrature has been documented for tensor product elements in \cite{gassner2011comparison}, where they showed that for a high order Taylor method in time, the maximum stable timestep under  $(N+1)$-point GLL volume and surface quadratures is roughly twice as large as the maximum stable timestep when volume/surface integrals are approximated using $(N+1)$ point Gauss quadratures.
-
-This discrepancy can be understood in terms of constants in finite element inverse Markov and trace inequalities.
-It was shown in \cite{chan2015gpu, chan2018multi} that, for linear problems, the maximum stable time-step scales inversely with the order-dependent constants $C_M, C_T$, where
-\[
-\int_{\hat{D}} \LRb{\Grad u}^2 \leq C_M \int_{\hat{D}} u^2, \qquad \int_{\partial \hat{D}} u^2 \leq C_T  \int_{\hat{D}} u^2, \qquad \forall u \in V^N.
-\]
-Here, the integrals over $\hat{D}, \partial \hat{D}$ are computed using the same volume and surface quadrature rules used for computations.
-These constants can be used to bound surface integrals which appear in DG formulations, which can in turn be used to construct bounds on the spectral radius of DG discretization matrices.
-These result in the following estimate of the maximum stable timestep
-\[
-dt \leq \frac{h}{\LRb{\lambda}_{\max} C_T}
-\]
-
-While the Markov constant $C_M$ is not affected by the choices of quadrature described in Section~\ref{sec:opts}, the constant $C_T$ depends on the choice of surface quadrature on triangular elements.
-It is known that $L^2$ norm computed using GLL quadrature is weaker than the full $L^2$ norm \cite{quarteroni1994introduction,canuto2007spectral}.
-Specifically,
-\[
-\LRp{2+\frac{1}{N}}^{-1/2}\int_{\partial \hat{D}, {\rm GLL}} u^2 \leq \int_{\partial \hat{D}} u^2 \qquad \forall u \in P^{N}\LRp{[-1,1]}, 
-\]
-where the former integral is under-integrated using GLL quadrature.
-In other words, when using GLL surface quadrature, the trace constant is not $C_T$ but $C_T \LRp{2+1/N}^{d/2}$.
-As a result, the triangular elements in Option~\ref{opt:1} (which use GLL surface quadrature) will have a smaller maximum stable timestep than the triangular elements for Options~\ref{opt:2} or \ref{opt:3} (though the timestep restriction for hexahedral elements in Option~\ref{opt:1} remains identical to that of a standard DG-SEM scheme).
-
-For Option~\ref{opt:2}, the timestep restriction
+Option 3 is most involved, resulting in a Gauss collocation method on quadrilaterals \cite{chan2018efficient}, and requires interpolation and two-point flux interactions between lines of volume quadrature nodes and boundary nodes.
+%However, Option~\ref{opt:3} provides some advantages over Options~\ref{opt:1} and \ref{opt:2} in terms of accuracy and stable timestep restrictions.
+
+Option~\ref{opt:2} is slightly more expensive than Option~\ref{opt:1}, as the solution must be interpolated from GLL to Gauss nodes on the boundary.
+However, this is less expensive than Option~\ref{opt:3} since volume GLL nodes include GLL boundary nodes as a subset.
+This implies that the matrix $\bm{V}_f$ is sparse, such that interpolation to boundary Gauss nodes involves only nodal values at boundary GLL nodes.
+Thus, Option~\ref{opt:2} requires only two-point flux computations between boundary GLL and Gauss nodes.
+In contrast, the Gauss collocation scheme in Option~\ref{opt:3} computes two-point flux interactions through $\bm{f}_S$ between each boundary node and a line of volume nodes.
+
+%Using interpolation and projection matrices between boundary Gauss and GLL nodes, Option~\ref{opt:2} can also be reformulated and implemented in a way which more closely resembles a traditional DG-SEM method.
+For brevity, we avoid describing the details of this implementation in this work, and instead refer the reader to the manuscript \note{ADD CITATION TO MORTAR PAPER}.
+
+\subsection{Verification of discrete entropy conservation}
+
+We first verify that, for an entropy conservative flux and periodic domain, the spatial formulation tested against the projected entropy variables (\ref{eq:entropyrhs}) is numerically zero.
+We refer to this quantity as the entropy right-hand side (RHS).
+Section~\ref{sec:curvgeo} outlines conditions on quadrature accuracy which guarantee that the formulations (\ref{eq:esdgSkew}) and (\ref{eq:skewformcurved}) are discretely entropy conservative.
+These numerical results confirm that these conditions are tight.
+
+We induce a curved polynomial mapping by defining curved coordinates $\tilde{\bm{x}}$ through a mapping of Cartesian coordinates $\bm{x} \in [-1,1]^2$ 
+\begin{align*}
+\tilde{x}_1 &= x_1 + \alpha \cos\LRp{\frac{\pi}{2}x_1}\sin\LRp{\pi x_2}\\
+\tilde{x}_2 &= x_2 + \alpha \sin\LRp{{\pi}x_1}\cos\LRp{\frac{\pi}{2} x_2},
+\end{align*}
+where $\alpha = 1/8$ for the following experiments.
+We vary the geometric degree of this mapping from $N_{\rm geo} = 1$ to $N_{\rm geo} = N$, where $N_{\rm geo}$ denotes the polynomial degree of the geometric mapping on a quadrilateral or triangular element.
+
+Since Assumption~\ref{ass:quad} requires that the volume quadrature is sufficiently accurate to ensure that the mass matrix is positive-definite, we fix the volume to quadrature to be exact for degree $2N$ polynomials on triangles.
+On quadrilaterals, we fix the volume quadrature to be an $(N+1)$ point GLL quadrature.
+To verify the conditions given in Section~\ref{sec:curvgeo}, we vary the accuracy of the 2D surface quadrature rule.
+
+The initial condition is taken as the $L^2$ projection of the discontinuous profile
+\[
+    \rho = \begin{cases}
+    3 & |x-x_0| < 2.5\\
+    2 & \text{ otherwise}
+    \end{cases}, \qquad 
+    x_0 = 7.5, \qquad u = v = 0, \qquad p = \rho^\gamma.
+\]
+We evolve the solution until final time $T = 1.0$ on a domain $[0,15]\times[-.5,.5]$ using the skew symmetric formulation with $N = 6$ and a CFL of $1/2$.
+Table~\ref{table:ecrhs} shows the maximum entropy RHS over the duration of the simulation.We observe that for all $N_{\rm geo} \leq M+1$, the maximum entropy RHS is $O(10^{-14})$ and at the level of machine precision.
+When $N_{\rm geo}> M+1$, we observe that the maximum entropy RHS increases .
+We note that the case of $M= 5$ for the quadrilateral corresponds to the use of an $(N+1)$-point GLL rule for both volume and surface quadrature.
+The strong SBP property holds for this choice of quadrature, and the skew-symmetric formulation is equivalent to the entropy stable spectral collocation or DG-SEM  methods of \cite{carpenter2014entropy, gassner2016split}.
 
@@
 \centering
-\subfloat[Quadrilateral elements]{
- \begin{tabular}{|c || c| c| c| c| c| c| c|} 
- \hline
- N & 1 & 2& 3 & 4 & 5 & 6& 7\\
+\subfloat[Triangular mesh]{
+\begin{tabular}{c||c|c|c|c|c|c|}
+& $N_{\rm geo} = 1$ & $N_{\rm geo} = 2$ & $N_{\rm geo} = 3$ & $N_{\rm geo} = 4$ & $N_{\rm geo} = 5$ & $N_{\rm geo} = 6$ \\
 \hline\hline
-GLL/GLL &&&&&&&\\
- \hline
-Gauss/Gauss &&&&&&&\\
- \hline
-Gauss/GLL &&&&&&&\\
- \hline
-GLL/Gauss  &&&&&&&\\
+$M = 5 $ &  8.68e-14 & 9.41e-14 & 9.31e-14 & 9.10e-14& 9.92e-14 & 8.90e-14\\
+\hline
+$M = 3 $ &  1.01e-13 & 8.87e-14 & 8.79e-14 & 9.68e-14 & 0.00833 & 0.00967 \\
+\hline
+$M = 1 $ & 1.8e-13 & 1.82e-13 & 1.998 & 2.104 & 2.080 & 2.086\\
+\hline
+\end{tabular}
+}\\
+\subfloat[Quadrilateral mesh]{
+\begin{tabular}{c||c|c|c|c|c|c|}
+& $N_{\rm geo} = 1$ & $N_{\rm geo} = 2$ & $N_{\rm geo} = 3$ & $N_{\rm geo} = 4$ & $N_{\rm geo} = 5$ & $N_{\rm geo} = 6$ \\
+\hline\hline
+$M = 5$ & 1.74e-14 & 1.06e-14 & 1.31e-14 &  1.35e-14 & 1.06e-14 & 1.31e-14\\
+$M = 3$ & 2.89e-14 & 2.59e-14 &  2.83e-14 & 2.43e-14 & 3.20e-05 & 3.27e-05 \\
+\hline
+$M = 1$ &4.68e-14 & 3.77e-14 & 0.1548 &  0.1532 & 0.1517 & 0.1517\\
 \hline
@@
 }
- \caption{Markov and trace constants for different configurations of triangular and quadrilateral elements.}
+\caption{Maximum absolute value of the entropy RHS over $t\in [0,1]$ for degree $N=6$ approximations on triangular and quadrilateral meshes.
+The volume quadrature for the quadrilateral mesh is taken to be $(N+1)$-point GLL quadrature.
+The surface quadrature is taken to be a 1D GLL quadrature with a varying number of points, such that the rule is exact for degree $M+N$ polynomials.
+}
+\label{table:ecrhs}
 \end{table}
 
-\note{Add experiments}
-
-%\subsection{Quadrature accuracy and rates of convergence}
-%
-%\subsection{Triangular meshes}
-%
-%\note{Affine and curved triangles with GLL surface quadrature (under-integrated).}
-%
-%\subsection{Quadrilateral meshes}
-%
-%\note{ GLL quadrilaterals with Gauss surface quadrature.  Explain that for GLL quadratures, the decoupled SBP property doesn't hold when Gauss points are used.  }
+%On hybrid meshes, we observe for Options~\ref{opt:1}, \ref{opt:2} and \ref{opt:3} that the entropy RHS is $O(10^{-14})$ in double precision on both affine and non-affine meshes as well.
+This indicates that the semi-discrete formulations  are indeed discretely entropy conservative.
+We omit the results here for brevity, as they are qualitatively similar to the previous results for quad-only or triangle-only meshes.
+%in \cite{crean2018entropy, chan2017discretely}.
 
@@
 
-It is not immediately clear that the outward normals as defined by (\ref{eq:normals}) will be equal and opposite across interfaces shared by elements of different types.
-Thus, we restrict ourselves to straight-sided meshes in this case, and will investigate the effect of curvilinear mappings on hybrid meshes in future work.
+We conclude with experiments on a mixed mesh containing both quadrilateral and triangular elements (Figure~\ref{fig:hybridmesh}).
+Figure~\ref{fig:hybriderrors} shows $L^2$ errors for the isentropic vortex computed at $T = 5$ for Options~\ref{opt:1}, \ref{opt:2}, and \ref{opt:3}.
+
+We observe that, in all cases, Option~\ref{opt:1} is less accurate than Options~\ref{opt:2} and \ref{opt:3}, and that Option~\ref{opt:3} achieves a rate of convergence close to the optimal $O(h^{N+1})$ rate, while Option~\ref{opt:1} typically achieves rates of convergence between $O(h^N)$ and $O(h^{N+1/2})$.
+However, Option~\ref{opt:2} behaves differently depending on the order $N$.
+For $N = 1$, Option~\ref{opt:2} achieves an accuracy similar to Option~\ref{opt:1}.
+However, as $N$ increases, Option~\ref{opt:2} becomes more accurate.
+For $N=4$, Option~\ref{opt:2} achieves the same level of error observed for Option~\ref{opt:3}.
+This suggests that Lemma~\ref{lemma:dsbpapprox} may be sufficient but not necessary for full order accuracy.
+
+%\begin{figure}
+%\centering
+%\caption{Coarsest hybrid mesh used for convergence studies. }
+%\label{fig:hybridmesh}
+%\end{figure}
 
@@
 \centering
-%\subfloat{\raisebox{2.85em}{
-\includegraphics[width=.45\textwidth]{figs/hybrid_mesh.png}
-%}}
-\caption{Coarsest hybrid mesh used for convergence studies. }
-\end{figure}
-
-\begin{figure}
-\centering
-%\subfloat{\raisebox{2.85em}{\includegraphics[width=.45\textwidth]{figs/hybrid_mesh.png}}}
-%\hspace{.25em}
-%\subfloat[Option~\ref{opt:1}]{
+\subfloat{\raisebox{3em}{
+\includegraphics[width=.43\textwidth]{figs/hybrid_mesh.png}
+}\label{fig:hybridmesh}}
+\subfloat[Convergence for $N = 1,2,3,4$]{
 \begin{tikzpicture}
 \begin{loglogaxis}[
-    width=.7\textwidth,
+    width=.5\textwidth,
     xlabel={Mesh size $h$},
     ylabel={$L^2$ errors}, 
-    xmax=.7,
+    xmax=.75,
     ymin=1e-5, ymax=5,
@@
     xmajorgrids=true, ymajorgrids=true, grid style=dashed,
-    legend entries={Option~\ref{opt:1},Option~\ref{opt:2},Option~\ref{opt:3}}    
+    legend entries={GLL-GLL, GLL-Gauss, Gauss-Gauss} 
 ]
 \pgfplotsset{
-cycle list={{blue, mark=*}, {red, dashed ,mark=triangle*}, {black, mark=square*}}
+cycle list={{blue, mark=*}, {red, dashed ,mark=square*},{black ,mark=triangle*}}
 }
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.333333,2.29471)(0.166667,1.55868)(0.0833333,0.746603)(0.0416667,0.274017)};
+coordinates{(0.333333,2.50656)(0.166667,1.874)(0.0833333,1.05647)(0.0416667,0.438926)};
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.333333,2.39742)(0.166667,1.63628)(0.0833333,0.885414)(0.0416667,0.435973)};
+coordinates{(0.333333,2.42633)(0.166667,1.75142)(0.0833333,0.958902)(0.0416667,0.462137)};
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.333333,1.74989)(0.166667,0.687578)(0.0833333,0.169682)(0.0416667,0.0350721)};
+coordinates{(0.333333,2.09363)(0.166667,0.981664)(0.0833333,0.264928)(0.0416667,0.0536376)};
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.333333,1.06222)(0.166667,0.267115)(0.0833333,0.0395222)(0.0416667,0.00569986)};
+coordinates{(0.333333,1.36947)(0.166667,0.390076)(0.0833333,0.0623387)(0.0416667,0.011361)};
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.333333,1.11474)(0.166667,0.265955)(0.0833333,0.0382799)(0.0416667,0.00492061)};
+coordinates{(0.333333,1.28187)(0.166667,0.326338)(0.0833333,0.0427717)(0.0416667,0.00402484)};
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.333333,0.51272)(0.166667,0.0625001)(0.0833333,0.00835287)(0.0416667,0.00120101)};
+coordinates{(0.333333,0.859449)(0.166667,0.132239)(0.0833333,0.0187852)(0.0416667,0.0026063)};
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.333333,0.321422)(0.166667,0.037836)(0.0833333,0.00336728)(0.0416667,0.000268691)};
+coordinates{(0.333333,0.48545)(0.166667,0.0615538)(0.0833333,0.00665061)(0.0416667,0.000711264)};
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.333333,0.364547)(0.166667,0.0393928)(0.0833333,0.00357584)(0.0416667,0.000392883)};
+coordinates{(0.333333,0.446407)(0.166667,0.0477486)(0.0833333,0.00372164)(0.0416667,0.000309705)};
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.333333,0.133053)(0.166667,0.0122055)(0.0833333,0.000730739)(0.0416667,4.37208e-05)};
+coordinates{(0.333333,0.286258)(0.166667,0.0331728)(0.0833333,0.00239593)(0.0416667,0.000144346)};
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.333333,0.10556)(0.166667,0.00654941)(0.0833333,0.000267376)(0.0416667,NaN)};
+coordinates{(0.333333,0.211329)(0.166667,0.0179798)(0.0833333,0.000816429)(0.0416667, 4.583068530777824e-05)};
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.333333,0.0955729)(0.166667,0.00649092)(0.0833333,0.000236043)(0.0416667,NaN)};
+coordinates{(0.333333,0.160937)(0.166667,0.0110045)(0.0833333,0.00041413)(0.0416667, 1.438708299584088e-05)};
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.333333,0.0525951)(0.166667,0.00229986)(0.0833333,6.74854e-05)(0.0416667,NaN)};
-\node at (axis cs:.45, 2.25) {$N = 1$};
-\node at (axis cs:.45,.8) {$N = 2$};
-\node at (axis cs:.45,.225) {$N = 3$};
-\node at (axis cs:.45,.075) {$N = 4$};
+coordinates{(0.333333,0.13975)(0.166667,0.00977304)(0.0833333,0.00037731)(0.0416667, 1.337566281886131e-05)};
+\node at (axis cs:.5,2.5) {$N = 1$};
+\node at (axis cs:.5,1.2) {$N = 2$};
+\node at (axis cs:.5,.425) {$N = 3$};
+\node at (axis cs:.5,.185) {$N = 4$};
 \end{loglogaxis}
 \end{tikzpicture}
-
-%}
-%\subfloat[Option~\ref{opt:2}]{
-%\begin{tikzpicture}
-%\begin{loglogaxis}[
-%    width=.49\textwidth,
-%    xlabel={Mesh size $h$},
-%    ylabel={$L^2$ errors}, 
-%%    xmax=1,
-%    ymin=1e-5, ymax=5,
-%    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
-%    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
-%    legend entries={$N=1$, $N=2$, $N=3$, $N=4$}    
-%]
-%\pgfplotsset{
-%cycle list name=color
-%}
-%\addplot+[semithick, mark options={solid, fill=markercolor}]
-%coordinates{(0.333333,2.39742)(0.166667,1.63628)(0.0833333,0.885414)(0.0416667,0.435973)};
-%\addplot+[semithick, mark options={solid, fill=markercolor}]
-%coordinates{(0.333333,1.11474)(0.166667,0.265955)(0.0833333,0.0382799)(0.0416667,0.00492061)};
-%\addplot+[semithick, mark options={solid, fill=markercolor}]
-%coordinates{(0.333333,0.364547)(0.166667,0.0393928)(0.0833333,0.00357584)(0.0416667,0.000392883)};
-%\addplot+[semithick, mark options={solid, fill=markercolor}]
-%coordinates{(0.333333,0.0955729)(0.166667,0.00649092)(0.0833333,0.000236043)(0.0416667,NaN)};
-%\end{loglogaxis}
-%\end{tikzpicture}
-%}
-%
-%\subfloat[Option~\ref{opt:3}]{
-%\begin{tikzpicture}
-%\begin{loglogaxis}[
-%    width=.49\textwidth,
-%    xlabel={Mesh size $h$},
-%    ylabel={$L^2$ errors}, 
-%%    xmax=1,
-%    ymin=1e-5, ymax=5,
-%    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
-%    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
-%    legend entries={$N=1$, $N=2$, $N=3$, $N=4$}    
-%]
-%\pgfplotsset{
-%cycle list name=color
-%}
-%\addplot+[semithick, mark options={solid, fill=markercolor}]
-%coordinates{(0.333333,1.74989)(0.166667,0.687578)(0.0833333,0.169682)(0.0416667,0.0350721)};
-%\addplot+[semithick, mark options={solid, fill=markercolor}]
-%coordinates{(0.333333,0.51272)(0.166667,0.0625001)(0.0833333,0.00835287)(0.0416667,0.00120101)};
-%\addplot+[semithick, mark options={solid, fill=markercolor}]
-%coordinates{(0.333333,0.133053)(0.166667,0.0122055)(0.0833333,0.000730739)(0.0416667,4.37208e-05)};
-%\addplot+[semithick, mark options={solid, fill=markercolor}]
-%coordinates{(0.333333,0.0525951)(0.166667,0.00229986)(0.0833333,6.74854e-05)(0.0416667,NaN)};
-%\end{loglogaxis}
-%\end{tikzpicture}
-%}
-\caption{Convergence of $L^2$ errors for the isentropic vortex solution for Option~\ref{opt:1}, Option~\ref{opt:2}, and Option~\ref{opt:3} for $N = 1,\ldots, 4$.}
+}
+\caption{Coarse hybrid mesh and  $L^2$ errors for the isentropic vortex solution for Option~\ref{opt:1}, Option~\ref{opt:2}, and Option~\ref{opt:3} for $N = 1,\ldots, 4$.}
+\label{fig:hybriderrors}
 \end{figure}
 
-\begin{itemize}
-\item \note{GQ-GQ hexes, tris}
-\item \note{GLL-GQ hexes, tris}
-\item \note{GQ-GLL hexes, GLL tris}
-\end{itemize}
-
 \section{Conclusions}
 
-\note{GLL quadrature really is super convenient.}
-
-\note{Need mortar to .}
-
-%\section{Acknowledgments} The author thanks David C.\ Del Rey Fernandez for
-
-%\appendix
+We have constructed skew-symmetric ``modal'' DG formulations of nonlinear conservation laws which are entropy stable under less restrictive conditions on quadrature accuracy.
+These formulations are motivated by volume and surface quadratures which arise naturally on hybrid meshes.
+Because these quadrature rules do not necessarily induce operators which satisfy a ``strong'' matrix summation-by-parts property, we generalize proofs of entropy stability and free stream (constant) preservation using a weaker ``variational'' SBP property.
+This variational SBP property is also used to construct explicit conditions relating between the order of accuracy and degree of accuracy of each quadrature rule.
+Finally, we derive conditions between the degree of quadrature accuracy and polynomial degree of the geometric mapping under which the skew-symmetric formulation is entropy stable on curved meshes.
+Numerical experiments confirm the entropy stability and high order accuracy of the proposed schemes on triangular, quadrilateral, and 2D hybrid meshes.
+
+\section{Acknowledgments}
+
+The author thanks David C.\ Del Rey Fernandez for helpful discussions.  Jesse Chan is supported by the National Science Foundation under awards DMS-1719818 and DMS-1712639.
+
+\appendix
+\section{Dependence of inverse and trace constants on quadrature}
+\label{sec:consts}
+
+The maximum stable timestep under explicit time-stepping depends on specific choices of volume and surface quadrature.
+The dependence of timestep on quadrature has been documented for tensor product elements in \cite{gassner2011comparison}, where they showed that for a high order Taylor method in time, the maximum stable timestep under  $(N+1)$-point GLL volume and surface quadratures is roughly twice as large as the maximum stable timestep when volume/surface integrals are approximated using $(N+1)$ point Gauss quadratures.
+
+This discrepancy can be understood in terms of constants in finite element inverse and trace inequalities.
+It was shown in \cite{chan2015gpu, chan2018multi} that, for linear problems, the maximum stable time-step scales inversely with the order-dependent constants $C_I, C_T$, where
+\begin{align}
+\int_{\hat{D}} \LRb{\Grad u}^2 \leq C_I \int_{\hat{D}} u^2, \qquad \int_{\partial \hat{D}} u^2 \leq C_T  \int_{\hat{D}} u^2, \qquad \forall u \in V^N \label{eq:consts}.
+\end{align}
+Here, the integrals over $\hat{D}, \partial \hat{D}$ are computed using the same volume and surface quadrature rules used for computations.
+These constants can be used to bound surface integrals which appear in DG formulations, which can in turn be used to construct bounds on the spectral radius of DG discretization matrices.
+The maximum stable timestep $dt_{\max}$ is thus inversely proportional to the inverse and trace constants
+\[
+dt_{\max} \propto C_T^{-1}, C_I^{-1}
+\]
+The constants $C_I, C_T$ depend on the choices of volume and surface quadrature used to evaluate each of the integrals in (\ref{eq:consts}).
+It is known that $L^2$ norm computed using GLL quadrature is weaker than the full $L^2$ norm \cite{quarteroni1994introduction,canuto2007spectral}.
+For the domain $\hat{D} = [-1,1]^d$ in $d$ dimensions, it can be shown that
+\begin{equation}
+\int_{\hat{D}} u^2 \leq \int_{\hat{D}, {\rm GLL}} u^2 \leq \LRp{2+\frac{1}{N}}^{d/2}\int_{\hat{D}} u^2 \qquad \forall u \in Q^{N},%\LRp{[-1,1]^d}, 
+\label{eq:gllineq}
+\end{equation}
+where the middle integral is under-integrated using GLL quadrature.
+In other words, the discrete $L^2$ norm induced using GLL quadrature is weaker than the $L^2$ norm induced by a more accurate quadrature rule, which will be reflected in the trace and inverse constants.
+
+\begin{table}
+\centering
+\subfloat[$C_I$, quadrilateral elements]{
+ \begin{tabular}{|c || c| c| c| c| c| c| c|} 
+ \hline
+ N & 1 & 2& 3 & 4 & 5 & 6& 7\\
+ \hline\hline
+Volume GLL & 2 & 12 & 37.16 & 91.67 &195.98&374.78&657.28\\
+ \hline
+Volume Gauss &6&30& 85.06 & 190.12 & 369.45 &652.30 &1072.75\\
+ \hline
+ \end{tabular}
+ }\\
+ \subfloat[$C_T$, quadrilateral elements]{
+ \begin{tabular}{|c || c| c| c| c| c| c| c|} 
+ \hline
+ N & 1 & 2& 3 & 4 & 5 & 6& 7\\
+ \hline\hline
+Volume GLL, surface GLL & 2 & 6 & 12 & 20&30&42&56\\
+% \hline
+%Volume GLL, surface Gauss  & 2 & 6 & 12 & 20&30&42&56 \\
+ \hline
+Volume Gauss, surface Gauss &6&12&20&30&42&56&72\\
+ \hline
+ \end{tabular}
+ }\\
+ \subfloat[$C_I$, triangular elements]{
+ \begin{tabular}{|c || c| c| c| c| c| c| c|} 
+ \hline
+ N & 1 & 2& 3 & 4 & 5 & 6& 7\\
+ \hline\hline
+Deg.\ $2N$ vol.\ quad.\  \cite{xiao2010quadrature} &9&39.27&100.10 &213.28 &401.16 &695.48 &1127.48\\
+ \hline
+ \end{tabular}
+ }\\
+\subfloat[$C_T$, triangular elements]{
+ \begin{tabular}{|c || c| c| c| c| c| c| c|} 
+ \hline
+ N & 1 & 2& 3 & 4 & 5 & 6& 7\\
+ \hline\hline
+Surface GLL & 12 & 16.14 & 20.52 & 28.12 &35.42 & 45.97 & 55.76\\
+ \hline
+Surface Gauss  &6&10.90&16.29& 24 & 31.88 &42.42 &52.89\\
+ \hline
+ \end{tabular}
+ }
+ \caption{Inverse and trace constants for triangular and quadrilateral elements with different quadrature configurations.}
+ \label{tab:consts}
+\end{table}
+
+Table~\ref{tab:consts} shows trace and inverse constants for triangular and quadrilateral elements under several different configurations of quadrature.
+For quadrilateral elements at high orders, we observe that the degree $N$ inverse constants $C_I$ under Gauss quadrature are roughly as large as the degree $(N+1)$ inverse constants under (volume) GLL quadrature.
+The degree $N$ trace constants $C_T$ under Gauss quadrature are exactly equal to the degree $(N+1)$ trace constants under GLL quadrature, which was proven in \cite{chan2015gpu}.
+Trace constants under GLL volume and Gauss surface quadrature are also identical to trace constants computed using GLL quadrature for both volume and surface integrals, which is a consequence of the lower bound in (\ref{eq:gllineq}).
+
+Several observations can be made based on the values of $C_I, C_T$ presented in Table~\ref{tab:consts}.
+On quadrilaterals, the maximum stable time-step for a degree $N$ DG scheme using Gauss quadrature is expected to be smaller than that of a degree $N$ scheme using GLL quadrature, which matches observations in \cite{gassner2011comparison}.
+Additionally, the maximum stable timestep under GLL volume quadrature and Gauss surface quadrature should be the same as the maximum stable timestep when GLL quadrature is used for both volume and surface integrals (e.g.\ DG-SEM).
+For triangles, the maximum stable timestep should be smaller under surface GLL quadrature compared to surface Gauss quadrature.
+However, we note that, while bounds on the maximum stable time-step can be derived based on the constants $C_I, C_T$ \cite{chan2015gpu, chan2018multi}, these bounds are not tight for upwind or dissipative fluxes \cite{krivodonova2013analysis}.
+
 %

### 2019-03-07 docs/ESDG_GQ/reviewerResponse.tex
updating GQ paper

\pdfoutput=1

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{bm}
%\usepackage{bbm}
%\usepackage{tikz}
%\usepackage[normalem]{ulem}
%\usepackage{hhline}
%\usepackage{graphicx}
%\usepackage{subfig}
\usepackage{color}
%\usepackage{doi}

%\renewcommand{\topfraction}{0.85}
%\renewcommand{\textfraction}{0.1}
%\renewcommand{\floatpagefraction}{0.75}
%
%
%\newcommand{\bbm}[1]{\mathbbm{#1}}
%\newcommand{\bs}[1]{\boldsymbol{#1}}
%\newcommand{\equaldef}{\stackrel{\mathrm{def}}{=}}
%
%
%\newcommand{\mb}[1]{\mathbf{#1}}
%\newcommand{\mbb}[1]{\mathbb{#1}}
%\newcommand{\mc}[1]{\mathcal{#1}}
%
%\renewcommand{\hat}{\widehat}
%\renewcommand{\tilde}{\widetilde}
\newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}{\rm #2}}}
\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
%\newcommand{\pdn}[3]{\frac{\partial^{#3}#1}{\partial#2^{#3}}}
%\newcommand{\snor}[1]{\left| #1 \right|}
%\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
%\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
%\newcommand{\LRb}[1]{\left| #1 \right|}
%\newcommand{\LRc}[1]{\left\{ #1 \right\}}
%\newcommand{\LRceil}[1]{\left\lceil #1 \right\rceil}
%\newcommand{\LRl}[1]{\left. \LRp{#1} \right|}
%\newcommand{\jump}[1] {\ensuremath{\llbracket#1\rrbracket}}
%\newcommand{\avg}[1] {\ensuremath{\LRc{\!\{#1\}\!}}}
%\newcommand{\Grad} {\ensuremath{\nabla}}
%\renewcommand{\d}{\partial}
\newcommand{\diag}[1]{{\rm diag}\LRp{#1}}

\newcommand{\note}[1]{{\color{blue}{#1}}}
\newcommand{\bnote}[1]{{\color{blue}{#1}}}
\newcommand{\rnote}[1]{{\color{red}{#1}}}

\newcommand{\LK}{L^2\LRp{D^k}}
\newcommand{\LdK}{L^2\LRp{\partial D^k}}
\newcommand{\Dhat}{\widehat{D}}
\newcommand{\Lhat}{L^2\LRp{\Dhat}}

\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}} % d in integrand

\begin{document}

%\maketitle

\note{We thank both Reviewers 1 and 2 for their feedback.  We describe steps taken to address reviewer comments and suggestions, which are described in the following response.  Revisions in the manuscript are also colored for ease of identification.  We hope these revisions improve the readability of this paper and its suitability for the audience of SISC.}

\section{Reviewer 1}

\begin{itemize}
\item The paper is well-written for the most part (authors should adjust a few grammatical typos throughout their paper), and is interesting. 
\item It would be good if the authors would add a few more numerical tests with model problems in domains with complex geometry where curved meshes are actually needed (at this point, authors considered either rectangular or parallelepiped domain).
\end{itemize}

\section{Reviewer 2}

\begin{itemize}
\item Page 7, eqn (3.8): why this equation holds? Can you add some comment here?

\note{Certainly.  We've added a continuous description of the equation, and have interpreted it as a quadrature approximation of a specific variational problem.}

\item Page 10, Theorem 3.4: In the proof, this leads to a extra coefficient 1/2 in front of the second term of the equation below (3.14).
\item In Theorem 3.4, it presents an semi-discrete entropy conservation for a two-element mesh. How to extend this to general case with multiple elements?
\item In the statement, boundary condition is not considered. How to utilize this theorem to solve the initial-boundary value problem numerically?
\item A pseudo code showing how to solve 1D conservation law would be very helpful
\item Page 19, Section 5.3: this problem is solved on the triangular mesh. How to generalized the proposed scheme from quadrilateral and hexahedral meshes to triangular mesh?
\item Page 4, line 152: Does the matrix size of $V_f$ equal to $(N+1) \times 2$ or $2 \times (N+1)$?
\item Page 14. line 495: $\bm{u} = (\bm{u}_1, . . . , \bm{u}_d)$. Is $\bm{u}_i$ a vector or a scalar?  
\item Page 15, line 518-522: $u$ and $v$ here represent the velocity in $x$ and $y$-direction, respectively. These conflict with the notations in previous sections, where, $u$ is the conservative variable and $v$ is the entropy variable. Definition of $\{\{\cdot\}\}$ should be added here, too.
\end{itemize}

\end{document}

### 2019-03-07 docs/ESDG_GQ/letter.tex
updating GQ paper

\documentclass{letter}
\usepackage{fullpage}
\signature{Jesse Chan, David C.\ Del Rey Fernandez, Mark H.\ Carpenter}
\address{6100 Main St,\\
Houston, TX 77005}

\begin{document}
\begin{letter}{SIAM Journal on Scientific Computing}

\opening{Dear SISC Section Editor,}

Please find enclosed our manuscript, 
\begin{center}
Jesse Chan, David C.\ Del Rey Fernandez, Mark H.\ Carpenter\\
\textit{Efficient entropy stable Gauss collocation methods},
\end{center}
which we would like to submit for publication as a research article in SISC.

The submitted manuscript investigates high order discontinuous Galerkin (DG) methods which satisfy a discrete entropy inequality.  Entropy stable DG methods have been constructed on quadrilateral and hexahedral elements based on collocation at tensor product Gauss-Legendre-Lobatto rules, which contain quadrature points at element boundaries.  Entropy stable DG methods have also been generalized to collocation schemes based on Gauss-Legendre rules, which do not contain quadrature points on element boundaries.  However, the inter-element coupling terms for Gauss-Legendre collocation are non-compact, requiring coupling between lines of nodes on neighboring elements.

This work shows how to construct Gauss-Legendre collocation schemes with compact inter-element couplings on curvilinear quadrilateral and hexahedral meshes.  We demonstrate the high order accuracy of the resulting schemes for smooth solutions, and show that, on non-Cartesian meshes, Gauss-Legendre collocation schemes achieve error levels which correspond to Gauss-Legendre-Lobatto collocation schemes of one degree higher.  We conclude by illustrating the stability and robustness of the proposed scheme for a shock-vortex interaction problem and the inviscid Taylor-Green vortex.

We hope that the method and results discussed in this manuscript would appeal to the readership of SISC.  All authors have approved the manuscript and agree with its submission, and we confirm that this manuscript has not been published elsewhere and is not under consideration by another journal.  We look forward to hearing from you at your earliest convenience.

\closing{Best regards}

\end{letter}

\end{document}

### 2019-03-07 docs/ESDG_mortar/main.tex
updating skew sym esdg paper

@@
 \author{Jesse Chan, David C.\ Del Rey Fernandez}
-\title{Mortar-based entropy stable discontinuous Galerkin methods}
+\title{Mortar-based entropy stable modal discontinuous Galerkin methods}
 \graphicspath{{./figs/}}
@@
 
+\begin{abstract}
+Entropy stable schemes provide a semi-discrete statement of stability for nonlinear systems of conservation laws.
+Entropy stable nodal discontinuous Galerkin (DG) methods have been constructed using entropy conservative finite volume fluxes \cite{tadmor1987numerical} and a summation-by-parts (SBP) finite difference framework \cite{gassner2013skew, carpenter2014entropy, gassner2016split, chen2017entropy, crean2018entropy, chan2017discretely, chan2018efficient}.
+Such DG schemes require evaluating entropy conservative numerical fluxes between combinations of volume and surface quadrature nodes.
+\end{abstract}
+
 \section{Introduction}
@@
 \note{Boilerplate introduction on high order + stability}
+
+The work presented here builds upon the skew-symmetric formulation introduced in \cite{chan2019skew}, but focuses instead on geometrically non-conforming meshes of tensor product elements.
+Such meshes may arise when applying domain decomposition techniques to a complex geometry (e.g.
+meshing sub-domains independently)  \cite{maday1989non, bernardi1993domain} or performing local mesh refinement \cite{staten2010hexahedral}.
+Entropy stable GLL collocation schemes have been constructed on non-conforming meshes in \cite{friedrich2017entropy} using special SBP projection operators.
+We study an alternative mortar-based approach to non-conforming interfaces, which differs from \cite{friedrich2017entropy} in that it is applicable to a broader class of ESDG methods (such as Gauss collocation schemes \cite{chan2018efficient}).
+The proposed approach will also reduce costs for both GLL and Gauss collocation on 3D non-conforming hexahedral meshes.
+
+\section{Entropy stable modal discontinuous Galerkin formulations}
+
+The primary complication of non-conforming meshes is the presence of hanging nodes or T-junctions, which arise where one element shares a ``non-conforming'' face with two or more neighboring elements.
+These are handled naturally by DG discretizations through the use of composite quadrature rules on non-conforming faces \cite{bui2012analysis, Kozdon2018}, which the proposed approach will utilize.
+
+The skew-symmetric formulation (\ref{eq:esdgSkew}) can be stably applied to non-conforming meshes using composite quadratures, but results in high computational costs.
+These costs are due to the structure of matrices present in the formulation (\ref{eq:esdgSkew}), which can be rewritten in block matrix form as
+\begin{equation}
+\bm{M}\td{\bm{u}_h}{t} + \sum_{i=1}^d\LRp{\begin{bmatrix} \bm{V}_q \\ \bm{V}_f \end{bmatrix}^T
+\LRp{\begin{bmatrix}
+\bm{Q}^i-\LRp{\bm{Q}^i}^T & \bm{E}^T\bm{B}^i\\
+-\bm{B}^i\bm{E} & \bm{0}
+\end{bmatrix} \circ \bm{F}^i_S}\bm{1} + \bm{V}_f^T\bm{B}^i\bm{f}_i^*} = 0.  
+\label{eq:esdgSkewBlock}
+\end{equation}
+The matrix $\bm{F}^i_S$ is formed by computing $\LRp{\bm{F}^i_S}_{jk} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}$ on the fly.
+Each entry requires evaluating the entropy conservative dyadic flux $\bm{f}^i_S$, resulting in a large number of floating point operations \cite{wintermeyer2018entropy}.
+However, due to the Hadamard product in (\ref{eq:esdgSkewBlock}), dyadic flux evaluations are only necessary between pairs of nodes corresponding to non-zero entries of  $\LRp{\bm{Q}^i-(\bm{Q}^i)^T}$ and $\bm{E}$.
+
+For quad and hex elements under tensor product volume quadrature, $\bm{Q}^i$ is the Kronecker product of 1D differentiation and identity matrices, which induces sparsity.
+As a result, dyadic flux evaluations are only required between ``lines'' of volume nodes \cite{carpenter2014entropy, chan2018efficient}.
+Dyadic flux evaluations also follow the sparsity pattern of the matrix $\bm{E}$, which maps from volume nodes to surface nodes.
+For GLL or Gauss collocation methods on conforming quad/hex meshes, $\bm{E}$ is also sparse when surface quadrature nodes are aligned with volume quadrature nodes \cite{chan2018efficient}.
+In such cases, dyadic flux evaluations are required only between lines of volume nodes and adjacent surface nodes  (Figure~\ref{subfig:aligned}).
+
+Composite quadrature rules on non-conforming meshes, however, are not aligned with volume nodes.
+It can be shown that the resulting matrix $\bm{E}$ is fully dense, and evaluating the formulation (\ref{eq:esdgSkewBlock}) requires dyadic flux evaluations between each surface node and \textit{all} volume nodes (as illustrated in Figure~\ref{subfig:nonaligned}).
+This greatly increases computational costs, especially at high orders of approximation.
+The proposed work will reduce the number of dyadic flux evaluations required for such composite quadratures.
+The increased cost of (\ref{eq:esdgSkewBlock}) under composite quadrature is due to the lack of sparsity in $\bm{E}$, which interpolates from volume to surface (composite) nodes.
+The proposed work avoids this issue by treating composite quadrature nodes as ``mortar'' nodes which are coupled directly to surface nodes but not directly to volume nodes, as illustrated in Figure~\ref{subfig:mortar}.
+
+\begin{figure}
+\centering
+\subfloat[Dyadic flux evaluations (aligned surface nodes)]{\raisebox{.1em}{\includegraphics[width=.375\textwidth]{figs/aligned.png}}\label{subfig:aligned}}
+\hspace{4em}
+\subfloat[Dyadic flux evaluations (composite surface nodes)]{\includegraphics[width=.374\textwidth]{figs/nonaligned.png}\label{subfig:nonaligned}}
+\caption{Figure~\ref{subfig:aligned} and \ref{subfig:nonaligned} illustrate when dyadic flux evaluations $\bm{f}^i_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}$ between nodes are necessary (indicated by dashed red lines) for conforming interfaces (Figure~\ref{subfig:aligned}) and non-conforming interfaces (Figure~\ref{subfig:nonaligned}) under a Gauss collocation scheme.
+Conforming interfaces (aligned surface nodes) require evaluations between a surface node and a line of volume nodes, while non-conforming interfaces (nonaligned composite surface nodes) require flux evaluations between a surface node and \textit{all} volume nodes.
+The proposed mortar approach (Figure~\ref{subfig:mortar}) reduces the number of dyadic flux evaluations by introducing composite nodes as mortars which couple only to surface nodes.
+}
+\label{fig:fluxsparsity}
+\end{figure}
 
@@
 
-\paragraph{Hybrid meshes:} entropy stable DG methods on hybrid meshes were introduced in \cite{chan2018skew} using a skew-symmetric formulation.
+\paragraph{Hybrid meshes:} entropy stable DG methods on hybrid meshes were introduced in \cite{chan2019skew} using a skew-symmetric formulation.
 The resulting methods are stable for more arbitrary choices of surface quadrature, in particular when an SBP property may not hold.
@@
 
-For non-conforming meshes, it can be advantageous to use composite Gauss quadratures on non-conforming interfaces.
-\note{Reference JK and LCW's paper on full-side vs split side mortars.}  However, interpolating the solution at volume Gauss nodes to split-side Gauss nodes is no longer a one-dimensional operation.
+For non-conforming meshes, it can be advantageous to use composite Gauss quadratures on non-conforming interfaces \cite{Kozdon2018}.
+However, interpolating the solution at volume Gauss nodes to split-side Gauss nodes is no longer a one-dimensional operation.
 
@@
 \centering
-\includegraphics[width=.65\textwidth]{figs/mortar.png}
-\caption{Illustration of mortar operators.
+\includegraphics[width=.5\textwidth]{figs/mortar.png}
+\caption{Illustration of mortar operators for a Gauss collocation scheme.
 The matrix $\bm{E}$ maps from volume quadrature points to surface quadrature points, $\bm{E}_m$ maps from surface to mortar surface points, and $\tilde{\bm{E}}_m$ maps from mortar surface points to surface points.
@@
 However, using properties of the mortar matrices $\bm{E}_m, \tilde{\bm{E}}_m$, we can rewrite the above formulation in a way which reflects a traditional mortar-based finite element implementation.
-In other words, we wish to implement (\ref{eq:mortar}) such that the only modification from the implementations in \cite{chan2017discretely} is a pre-processing step on mortar faces.
+In other words, we wish to implement (\ref{eq:mortar}) such that the only modification from the implementations in \cite{chan2017discretely} is an intermediate step on mortar faces.
 
@@
 \end{align}
-If the surface and mortar nodes are identical, then $\bm{E}_m = \tilde{\bm{E}}_m = \bm{I}$ and $\LRp{\tilde{\bm{E}}_m \circ \bm{F}_S^{sm}}\bm{1} = \LRp{ \bm{E}_m \circ \bm{F}_S^{ms}}\bm{1}$.
-The correction terms within $\tilde{\bm{f}}^*$ cancel out, and we recover the original scheme from \cite{chan2017discretely}.
+If the surface and mortar nodes are identical, then 
+\[
+\bm{E}_m = \tilde{\bm{E}}_m = \bm{I}, \qquad \LRp{\tilde{\bm{E}}_m \circ \bm{F}_S^{sm}}\bm{1} = \LRp{ \bm{E}_m \circ \bm{F}_S^{ms}}\bm{1}
+\]
+and the correction terms within $\tilde{\bm{f}}^*$ cancel out, and we recover the original scheme from \cite{chan2017discretely}.
 
@@
 \end{enumerate}
-%\bm{M}\td{\bm{u}_N}{t} + \begin{bmatrix} \bm{V}_q \\ \bm{V}_f \\ \bm{V}_m \end{bmatrix}^T
-%\LRp{\begin{bmatrix}
-%\bm{Q}_i-\bm{Q}_i^T & \bm{E}^T\bm{B}_i &\\
-%-\bm{B}_i\bm{E} &  & \bm{B}_i\tilde{\bm{E}}_m \\
-%& -\tilde{\bm{B}}_i{\bm{E}}_m & 
-%\end{bmatrix} \circ \bm{F}_S}\bm{1} + \bm{V}_m^T\tilde{\bm{B}}_i\bm{f}_i^* = 0.
-
-%It is possible to efficiently interpolate from volume Gauss nodes to split-side Gauss nodes in two steps.
-We first interpolate to full-side Gauss nodes on the face, then interpolate from full-side  to split-side Gauss nodes.
-Each of these two operations can be done using only one-dimensional interpolation matrices, implying that the interpolation matrix from volume to split-side Gauss nodes can be decomposed into
-%\[
-%\bm{V}_{\rm vol-to-split} = \bm{V}_{\rm vol-to-full} \bm{V}_{\rm full-to-split}
-%\]
-%where each of the matrices in the decomposition possess either a sparse or Kronecker product structure.  
-%
-%ESDG schemes based on decoupled SBP operators introduced in \cite{chan2017discretely} assume a single interpolation matrix which extrapolates from volume to surface nodes.
-These schemes require dyadic flux evaluations between all volume and surface nodes connected by this interpolation matrix.
-For non-conforming interfaces with split-side Gauss nodes on the face, this necessitates dyadic flux computations between all volume nodes and all split-side Gauss nodes.
-%
-%We show in this section how to construct an entropy stable Gauss collocation scheme based on two layers of surface nodes: full-side and split-side nodes.
-These surface nodes will be loosely through a skew-symmetric coupling term involving polynomial interpolation matrices between full-side and split-side nodes.
-This scheme requires dyadic flux computations between volume and full-side surface nodes, and between full-side and split-side surface nodes.
-Each set of dyadic flux computations can be reduced to
-
-%We first introduce notation which is helpful for operations involving surface traces.  Let the matrix $\bm{R}$ denote the matrix which maps coefficients of a function to coefficients of its trace.  
-%
-%Let ${\bm{V}_f}\tilde{P}_f$ denote the matrix which projects to the trace space using split-side quadrature and evaluates at full-side quadrature points.
-Similarly, let $\tilde{\bm{V}_f}{P}_f$ denote the matrix which projects to the trace space using full-side quadrature and evaluates at split-side quadrature points
-%\begin{align*}
-%\bm{C}^i_f &= \begin{pmatrix}
-%\bm{0} & {\rm diag}\LRp{\hat{\bm{n}}_i}{\bm{V}_f}\tilde{\bm{P}}_f\\
-%-\tilde{\bm{V}_f}{\bm{P}}_f{\rm diag}\LRp{\hat{\bm{n}}_i} &\bm{0}
-%\end{pmatrix}, \\
-%\bm{S}^i_f &= \begin{pmatrix}
-%\diag{\bm{w}_f} & \\
-%& \diag{\tilde{\bm{w}}_f}
-%\end{pmatrix}\bm{C}^i_f.
-%\end{align*}
-%It is straightforward to show that the matrix $\bm{S}_f$ is skew-symmetric.
-%
-%%{fig:gqcon_noncon}
-%Let $\tilde{\bm{F}}_S$ denote the flux matrix, then 
-%\[
-%\begin{bmatrix}\bm{L}_f & \tilde{\bm{L}}_f\end{bmatrix} \LRp{\bm{C}^i_f\circ\tilde{\bm{F}}_S}\bm{1}
-%\]
-%is a skew-symmetric correction term which ensures entropy stability and high order accuracy (through consistency up to quadrature order).   
-%
-%Since the difference is high order accurate, you can remove this correction term and still get high order accuracy.
-However, the spatial entropy RHS is no longer zero if you use inconsistent surface and coupling quadrature without the correction term (\note{prove this}).
-%
-%\note{Do we need 3 layers of surface nodes in 3D?  Interpolation to isotropic split-side nodes is a Kronecker product - does that reduce things down?  }
 
@@
 
+Figure~\ref{subfig:noncon} shows preliminary results for the 2D compressible Euler equations by the PI on a manually constructed non-conforming quadrilateral mesh, with volume, surface, and mortar quadratures constructed from one-dimensional GLL or Gauss quadrature rules.
+The PI has verified that this preliminary implementation of the mortar-based formulation is discretely entropy stable.
+A preliminary accuracy analysis for the skew-symmetric formulation (\ref{eq:esdgSkew}) also suggests that GLL quadrature should achieve an $O(h^N)$ rate of convergence and that Gauss quadrature should achieve an $O(h^{N+1})$ rate of convergence.
+Both predicted rates are confirmed in Figure~\ref{subfig:noncon}.
+
+\begin{figure}
+\centering
+\subfloat[Coarse non-conforming mesh]{\raisebox{3em}{\includegraphics[width=.425\textwidth]{figs/noncon_mesh.png}}}
+\hspace{.25em}
+\subfloat[Sub-optimal rates if under-integrated]{
+\begin{tikzpicture}
+\begin{loglogaxis}[
+    width=.5\textwidth,
+    xlabel={Mesh size $h$},
+    ylabel={$L^2$ errors}, 
+    xmax=1,
+    ymin=5e-5, ymax=5,
+    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
+    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
+    legend entries={Lobatto, Gauss, Lobatto-Gauss}    
+]
+\pgfplotsset{
+cycle list={{blue, mark=*}, {red, dashed ,mark=square*},{black, dashdotted ,mark=triangle*}}
+}
+% N = 1
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,2.6993)(0.166667,1.6718)(0.0833333,0.7792)(0.0416667,0.29211)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,2.0534)(0.166667,0.7966)(0.0833333,0.2022)(0.0416667,0.0449413)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,2.94072)(0.166667,1.73057)(0.0833333,0.97069)(0.0416667,0.487387)};
+
+% N = 2
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,1.1609)(0.166667,0.3562)(0.0833333,0.0768)(0.0416667,0.0192197)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,0.7191)(0.166667,0.141)(0.0833333,0.0178)(0.0416667,0.00243087)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,1.20291)(0.166667,0.347365)(0.0833333,0.0606688)(0.0416667,0.0144843)};
+
+% N = 3
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,0.551145)(0.166667,0.0754237)(0.0833333,0.0100808)(0.0416667,0.00120633)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,0.352706)(0.166667,0.0258415)(0.0833333,0.00211953)(0.0416667,0.000132969)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,0.557084)(0.166667,0.0671373)(0.0833333,0.00744086)(0.0416667,0.000837905)};
+
+% N = 4
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,0.267668)(0.166667,0.0192603)(0.0833333,0.00140341)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,0.128217)(0.166667,0.00779943)(0.0833333,0.000270803)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(0.333333,0.212176)(0.166667,0.0151178)(0.0833333,0.000826307)};
+
+\node at (axis cs:.55,2.5) {$N = 1$};
+\node at (axis cs:.55,1.0) {$N = 2$};
+\node at (axis cs:.55,.42) {$N = 3$};
+\node at (axis cs:.55,.16) {$N = 4$};
+\end{loglogaxis}
+\end{tikzpicture}
+}
+\caption{Blah}
+\end{figure}
+
 \subsection{Gauss Lobatto quadrature}

### 2019-03-07 docs/ESDG_GQ/reviewerResponse.tex
more updates to revision

@@
 \item The paper is well-written for the most part (authors should adjust a few grammatical typos throughout their paper), and is interesting.
-\item It would be good if the authors would add a few more numerical tests with model problems in domains with complex geometry where curved meshes are actually needed (at this point, authors considered either rectangular or parallelepiped domain).
+It would be good if the authors would add a few more numerical tests with model problems in domains with complex geometry where curved meshes are actually needed (at this point, authors considered either rectangular or parallelepiped domain).
+
+\note{blah}
 \end{itemize}
@@
 
-\note{Certainly.  We've added a continuous description of the equation, and have interpreted it as a quadrature approximation of a specific variational problem.}
+\note{Certainly.
+We've added a continuous description of the equation, and have interpreted the matrix version as a quadrature approximation of a specific variational problem for the approximation of the derivative.
+This variational problem includes boundary correction terms, which correspond to the matrix ``strong form'' of (3.8).}
 
 \item Page 10, Theorem 3.4: In the proof, this leads to a extra coefficient 1/2 in front of the second term of the equation below (3.14).
+
+\note{We thank the reviewer for catching this, and have fixed this in the revised version.}
+
 \item In Theorem 3.4, it presents an semi-discrete entropy conservation for a two-element mesh. How to extend this to general case with multiple elements?
+
+\note{The block structure of the two-element mesh can be extended to multiple elements.
+We have added a discussion involving a three-element mesh, which illustrates the procedure without introducing too much additional notation.
+The approach simply requires a global SBP operator over the entire mesh, which is constructed in a block fashion.
+We hope this revision makes the extension to multiple elements more clear.
+}
+
 \item In the statement, boundary condition is not considered. How to utilize this theorem to solve the initial-boundary value problem numerically?
-\item A pseudo code showing how to solve 1D conservation law would be very helpful
+
+\note{We have added a short discussion describing how to impose boundary conditions in a weak fashion through a boundary numerical flux.
+If the boundary numerical flux satisfies an entropy stability property, then the discretization with boundary conditions is also entropy stable.}
+
+\item A pseudo code showing how to solve 1D conservation law would be very helpful.
+
+\note{We have added a description of how to numerically solve the proposed system of ODEs using explicit time-stepping.
+With the exception of the evaluation of the nonlinear term, the procedure is relatively standard for time-explicit DG formulations.
+}
+
 \item Page 19, Section 5.3: this problem is solved on the triangular mesh. How to generalized the proposed scheme from quadrilateral and hexahedral meshes to triangular mesh?
+
+\note{We apologize for the confusion.
+Here, ``triangulated'' simply means ``meshed''.
+The meshes considered are all quadrilateral and hexahedral, with triangular and simplicial meshes treated in a separate work.}
+
 \item Page 4, line 152: Does the matrix size of $V_f$ equal to $(N+1) \times 2$ or $2 \times (N+1)$?
+
+\note{We thank the reviewer for catching this - it is indeed $2\times (N+1)$.}
+
 \item Page 14. line 495: $\bm{u} = (\bm{u}_1, . . . , \bm{u}_d)$. Is $\bm{u}_i$ a vector or a scalar?
+
+\note{We apologize for the confusion.  $\bm{u}_i$ should be a scalar, and we have edited the paper to clear this up.  We are no longer using bold to denote velocity $u_i$.}
+
 \item Page 15, line 518-522: $u$ and $v$ here represent the velocity in $x$ and $y$-direction, respectively.
@@
 Definition of $\{\{\cdot\}\}$ should be added here, too.
+
+\note{We have added definitions of the average and logarithmic mean.
+We have also clarified the notation that $u,v$ denote $x,y$ velocity, while bolded $\bm{u},\bm{v}$ denote conservative and entropy variables.}
 \end{itemize}

### 2019-03-07 docs/ESDG_GQ/esdg_gq.tex
more updates to revision

@@
 
-Let $\bm{D}$ denote the nodal differentiation matrix, and let $\bm{V}_f$ denote the $(N+1)\times 2$ matrix which interpolates polynomials at Gauss nodes to values at endpoints.
+Let $\bm{D}$ denote the nodal differentiation matrix, and let $\bm{V}_f$ denote the \note{$2\times (N+1)$} matrix which interpolates polynomials at Gauss nodes to values at endpoints.
 These two matrices are defined entrywise as
@@
 \end{equation}
-The expression (\ref{eq:qn}) can be rewritten in ``strong'' form as follows
+The approximation (\ref{eq:qn}) can be rewritten in ``strong'' form as follows
 \begin{align*}
@@
 The decoupled SBP operator $\bm{Q}_N$ can thus be interpreted as adding boundary corrections to the GSBP operator $\bm{D}$ in a skew-symmetric fashion.
+\note{More specifically, the expression (\ref{eq:qn}) corresponds to a quadrature approximation of the following variational approximation of the derivative \cite{chan2017discretely}: find $u\in P^N([-1,1])$ such that},
+\[
+\note{\int_{-1}^1u({x})v({x})  = \int_{-1}^1 {I_N f\pd{I_Ng}{x}v} + \LRl{(g-I_Ng)\frac{\LRp{fv + I_N(fv)}}{2}}_{-1}^1,} \qquad \forall v\in P^N([-1,1]),
+\]
+\note{where $I_N$ denotes the degree $N$ polynomial approximation at the $(N+1)$ Gauss points.}
 
@@
 \label{eq:form}
-\bm{W}_h \td{}{t} \bm{u}_h + \bm{I}_h^T\LRp{\bm{Q}_h \circ \bm{F}_S}\bm{1} = 0.
+\bm{W}_h \td{}{t} \bm{u}_h + \note{2} \bm{I}_h^T\LRp{\bm{Q}_h \circ \bm{F}_S}\bm{1} = 0.
 \end{gather}
@@
 
+\note{Extending the two-element case to multiple elements can be done by defining analogous ``global'' differentiation matrices.
+Suppose that there are now three elements.
+Then, the matrices $\bm{S}_h, \bm{B}_h$ can be defined as}
+\begin{gather*}
+\note{\bm{S}_h \myeq \LRs{\begin{array}{cc|cc|cc}
+\bm{S} &  \frac{1}{2}\bm{V}_f^T\bm{B} & &&& \\
+ - \frac{1}{2}\bm{B}\bm{V}_f &  & & \frac{1}{2}\bm{e}_R\bm{e}_L^T &&\\ [.2em]
+ \hline\\[-1.0em]
+&& \bm{S} &  \frac{1}{2}\bm{V}_f^T\bm{B}&& \\
+& -\frac{1}{2}\bm{e}_L\bm{e}_R^T &  - \frac{1}{2}\bm{B}\bm{V}_f &&\frac{1}{2}\bm{e}_R\bm{e}_L^T&\\
+ \hline\\[-1.0em]
+&&&& \bm{S} &  \frac{1}{2}\bm{V}_f^T\bm{B} \\
+&&& -\frac{1}{2}\bm{e}_L\bm{e}_R^T &  - \frac{1}{2}\bm{B}\bm{V}_f &\\
+\end{array}}}
+\\
+\note{\bm{B}_h \myeq \LRs{\begin{array}{cc|c|cc}
+&  & && \\
+ & -\bm{e}_L\bm{e}_L^T & && \\
+ \hline
+&&  \bm{0} & & \\
+ \hline
+&&&&    \\
+&&&  & \bm{e}_R\bm{e}_R^T\\
+\end{array}}, \qquad \bm{I}_h = \begin{bmatrix}
+\bm{I} &&\\
+\bm{V}_f &&\\
+&\bm{I}&\\
+&\bm{V}_f&\\
+&&\bm{I}\\
+&&\bm{V}_f\\
+\end{bmatrix}.}
+\end{gather*}
+\note{These global operators again satisfy the SBP property $\bm{Q}_h + \bm{Q}_h^T = \bm{B}_h$, where $\bm{Q}_h = \bm{S}_h + \frac{1}{2}\bm{B}_h$.
+Moreover, $\bm{Q}_h$ can be used to produce high order accurate approximations of derivatives, and can be used to construct entropy conservative schemes as in Theorem~\ref{thm:consentropy}.
+The extension to a general number of elements is straightforward.
+}
+
 It is possible to convert the semi-discrete entropy equality in Theorem~\ref{thm:consentropy} to a semi-discrete entropy inequality by adding appropriate interface dissipation terms, such as Lax-Friedrichs or matrix dissipation \cite{winters2017uniquely}.
 We note that these terms must be computed in terms of $\tilde{\bm{u}}_f$ in order to ensure a discrete dissipation of entropy \cite{chen2017entropy, chan2017discretely}.
+\note{Boundary conditions can also be incorporated into the formulation (\ref{eq:form}) in a weak fashion.
+Let $\bm{u}_L, \bm{u}_R$ denote the values of the solution at the left and right domain boundaries of a 1D domain.
+Using the SBP property, we can modify (\ref{eq:form}) to }
+\[
+\note{\bm{W}_h\td{\bm{u}_h}{t} + 2\bm{I}_h^T \LRp{\bm{S}_h\circ \bm{F}_S}\bm{1} + \bm{I}_h^T\bm{B}_h \begin{bmatrix}\bm{0}\\ \bm{f}(\bm{u}_L) \\ \vdots \\ \bm{0}\\ \bm{f}(\bm{u}_R)\end{bmatrix}}= 0,
+\]
+\note{where we have used the consistency of $\bm{f}_S$ and the fact that $\bm{B}_h$ is diagonal to simplify the boundary term.
+We can now replace $\bm{f}(\bm{u}_L), \bm{f}(\bm{u}_R)$ with numerical fluxes $\bm{f}^*_L, \bm{f}^*_R$ as follows:}
+\begin{equation}
+\note{\bm{W}_h\td{\bm{u}_h}{t} + 2\bm{I}_h^T \LRp{\bm{S}_h\circ \bm{F}_S}\bm{1} + \bm{I}_h^T \bm{B}_h \begin{bmatrix}\bm{0}\\ \bm{f}^*_L \\ \vdots \\ \bm{0}\\ \bm{f}^*_R\end{bmatrix}}= 0.
+\label{eq:bform}
+\end{equation}
+\note{If the boundary numerical flux satisfies the entropy stability conditions}
+\[
+\note{\psi_L - \bm{v}_L^T\bm{f}^*_{L} \leq 0, \qquad {\psi}_R - \bm{v}_R^T\bm{f}^*_{R} \leq 0}
+\]
+\note{then the resulting scheme satisfies a global entropy inequality \cite{chen2017entropy}.  }
+
+\note{Assuming that the ODE system (\ref{eq:form}) exactly integrated in time and that entropy dissipative numerical fluxes and boundary conditions are used, the solution will satisfy a discrete entropy inequality (\ref{eq:weakentropyineq}).
+In practice, the system (\ref{eq:form}) is solved using an ODE time-stepper.
+For an explicit time-stepper, then all that is necessary is to invert the diagonal matrix $\bm{W}_h$ and evaluate the spatial terms in (\ref{eq:bform}).
+We note that the conservation or dissipation of entropy is guaranteed up to time-stepper accuracy.
+In practice, we observe that entropy is dissipated for all problems considered, despite the fact that the proof does not hold at the fully discrete level.}
 
@@
 \label{sec:3}
-The compressible Euler equations in $d$ dimensions are given as follows:
+
+\note{In this section, we present numerical examples for} the compressible Euler equations, which are given in $d$ dimensions as follows:
 \begin{align*}
-\pd{\rho}{t} &+ \sum_{j=1}^d \pd{\LRp{\rho \bm{u}_j}}{x_j} = 0,\\
-\pd{\rho \bm{u}_i}{t} &+ \sum_{j=1}^d \pd{\LRp{\rho \bm{u}_i\bm{u}_j + p\delta_{ij} }}{x_j} = 0, \qquad i = 1,\ldots,d\\
-\pd{E}{t} &+ \sum_{j=1}^d \pd{\LRp{\bm{u}_j(E+p)}}{x_j} = 0.\nonumber
+\pd{\rho}{t} &+ \note{\sum_{j=1}^d \pd{\LRp{\rho {u}_j}}{x_j}} = 0,\\
+\note{\pd{\rho {u}_i}{t}} &+ \note{\sum_{j=1}^d \pd{\LRp{\rho u_iu_j + p\delta_{ij} }}{x_j}} = 0, \qquad i = 1,\ldots,d\\
+\pd{E}{t} &+ \note{\sum_{j=1}^d \pd{\LRp{{u}_j(E+p)}}{x_j}} = 0.\nonumber
 \end{align*}
-Here, $\rho$ is density, $\bm{u} = \LRp{\bm{u}_1, \ldots,\bm{u}_d}$ is the vector of velocities, and $E$ is the total energy.  The pressure $p$ and specific internal energy $\rho e$ are given by 
-\[
-p = (\gamma-1)\LRp{E - \frac{1}{2}\rho \LRb{\bm{u}}^2}, \qquad \rho e = E - \frac{1}{2}\rho \LRb{\bm{u}}^2, \qquad \LRb{\bm{u}}^2 = \LRp{\sum_{j=1}^d \bm{u}_j^2}.
+Here, $\rho$ is density, \note{$\LRp{{u}_1, \ldots,{u}_d}$} is the vector of velocities, and $E$ is the total energy.  The pressure $p$ and specific internal energy $\rho e$ are given by 
+\[
+\note{p = (\gamma-1)\LRp{E - \frac{1}{2}\rho \LRp{\sum_{j=1}^d u_j^2}}}, \qquad \note{\rho e = E - \frac{1}{2}\rho \LRp{\sum_{j=1}^d u_j^2}}. %\qquad \LRb{\bm{u}}^2 = \LRp{\sum_{j=1}^d u_j^2}.
 \]
@@
 v_1 &= \frac{\rho e (\gamma + 1 - s) - E}{\rho e}, \\
-v_{1+ i}&= \frac{\rho \bm{u}_i}{\rho e}, \qquad i = 1,\ldots, d,\\
+v_{1+ i}&= \frac{\rho \note{{u}_i}}{\rho e}, \qquad i = 1,\ldots, d,\\
 v_{d+2} &= -\frac{\rho}{\rho e},
@@
 \rho &= -(\rho e) v_{d+2}, \\
- \rho \bm{u}_i &= (\rho e) v_{1+i}, \qquad i = 1,\ldots,d\\
+ \rho \note{u_i} &= (\rho e) v_{1+i}, \qquad i = 1,\ldots,d\\
  E &= (\rho e)\LRp{1 - \frac{\sum_{j=1}^d{v_{1+j}^2}}{2 v_{d+2}}},
@@
 \end{equation*}
+\note{Let $f$ denote some piecewise polynomial function, and let $f^+$ denote the exterior value of $f$ across an element face.  We define the average and logarithmic averages as follows:}
+\[
+\note{\avg{f} = \frac{f^+ + f}{2}}, \qquad \note{\avg{f}^{\log} = \frac{f+ - f}{\log\LRp{f^+}-\log\LRp{f}}}.
+\]
+\note{The average and logarithmic average are assumed to act component-wise on vector-valued functions.
+We evaluate the logarithmic average using the numerically stable approach of \cite{ismail2009affordable}.}
 Explicit expressions for entropy conservative numerical fluxes in two dimensions  are given by Chandrashekar \cite{chandrashekar2013kinetic}
 \begin{align*}
-&f^1_{1,S}(\bm{u}_L,\bm{u}_R) = \avg{\rho}^{\log} \avg{u},& &f^1_{2,S}(\bm{u}_L,\bm{u}_R) = \avg{\rho}^{\log} \avg{v},&\\
-&f^2_{1,S}(\bm{u}_L,\bm{u}_R) = f^1_{1,S} \avg{u} + p_{\rm avg},&  &f^2_{2,S}(\bm{u}_L,\bm{u}_R) = f^1_{2,S} \avg{u},&\nonumber\\
-&f^3_{1,S}(\bm{u}_L,\bm{u}_R) = f^2_{2,S},& &f^3_{2,S}(\bm{u}_L,\bm{u}_R) = f^1_{2,S} \avg{v} + p_{\rm avg},&\nonumber\\
-&f^4_{1,S}(\bm{u}_L,\bm{u}_R) = \LRp{E_{\rm avg} + p_{\rm avg}}\avg{u},& &f^4_{2,S}(\bm{u}_L,\bm{u}_R) = \LRp{E_{\rm avg} + p_{\rm avg} }\avg{v},& \nonumber
+&f^1_{1,S}(\bm{u}_L,\bm{u}_R) = \avg{\rho}^{\log} \avg{u_1},& &f^1_{2,S}(\bm{u}_L,\bm{u}_R) = \avg{\rho}^{\log} \avg{u_2},&\\
+&f^2_{1,S}(\bm{u}_L,\bm{u}_R) = f^1_{1,S} \avg{u_1} + p_{\rm avg},&  &f^2_{2,S}(\bm{u}_L,\bm{u}_R) = f^1_{2,S} \avg{u_1},&\nonumber\\
+&f^3_{1,S}(\bm{u}_L,\bm{u}_R) = f^2_{2,S},& &f^3_{2,S}(\bm{u}_L,\bm{u}_R) = f^1_{2,S} \avg{u_2} + p_{\rm avg},&\nonumber\\
+&f^4_{1,S}(\bm{u}_L,\bm{u}_R) = \LRp{E_{\rm avg} + p_{\rm avg}}\avg{u_1},& &f^4_{2,S}(\bm{u}_L,\bm{u}_R) = \LRp{E_{\rm avg} + p_{\rm avg} }\avg{u_2},& \nonumber
 \end{align*}
-where we have defined the auxiliary quantities 
+where we have introduced the auxiliary quantities 
 \begin{gather}
-p_{\rm avg} = \frac{\avg{\rho}}{2\avg{\beta}}, \qquad E_{\rm avg} = \frac{\avg{\rho}^{\log}}{2\avg{\beta}^{\log}\LRp{\gamma -1}}   + \frac{\nor{\bm{u}}^2_{\rm avg}}{2}, \label{eq:fluxaux} \\
- \nor{\bm{u}}^2_{\rm avg} = 2(\avg{u}^2 + \avg{v}^2) - \LRp{\avg{u^2} +\avg{v^2}} \nonumber.  
+\note{\beta = \frac{\rho}{2p}}, \qquad p_{\rm avg} = \frac{\avg{\rho}}{2\avg{\beta}}, \qquad \qquad E_{\rm avg} = \frac{\avg{\rho}^{\log}}{2\avg{\beta}^{\log}\LRp{\gamma -1}}   + \frac{\note{u_{\rm avg}^2}}{2}, \label{eq:fluxaux} \\
+\note{u^2_{\rm avg}} = 2(\avg{u_1}^2 + \avg{u_1}^2) - \LRp{\avg{u_1^2} +\avg{u_2^2}} \nonumber.  
 \end{gather}
@@
 \bm{f}_{1,S} = \LRp{\begin{array}{c}
-\avg{\rho}^{\log}\avg{u}\\
-\avg{\rho}^{\log}\avg{u}^2 + p_{\rm avg}\\
-\avg{\rho}^{\log}\avg{u}\avg{v}\\
-\avg{\rho}^{\log}\avg{u}\avg{w}\\
-(E_{\rm avg}+ p_{\rm avg})\avg{u}\\
+\avg{\rho}^{\log}\avg{u_1}\\
+\avg{\rho}^{\log}\avg{u_1}^2 + p_{\rm avg}\\
+\avg{\rho}^{\log}\avg{u_1}\avg{u_2}\\
+\avg{\rho}^{\log}\avg{u_1}\avg{u_3}\\
+(E_{\rm avg}+ p_{\rm avg})\avg{u_1}\\
 \end{array}}, 
@@
 \bm{f}_{2,S} = \LRp{\begin{array}{c}
-\avg{\rho}^{\log}\avg{v}\\
-\avg{\rho}^{\log}\avg{u}\avg{v}\\
-\avg{\rho}^{\log}\avg{v}^2 + p_{\rm avg}\\
-\avg{\rho}^{\log}\avg{v}\avg{w}\\
-(E_{\rm avg}+ p_{\rm avg})\avg{v}\\
+\avg{\rho}^{\log}\avg{u_2}\\
+\avg{\rho}^{\log}\avg{u_1}\avg{u_2}\\
+\avg{\rho}^{\log}\avg{u_2}^2 + p_{\rm avg}\\
+\avg{\rho}^{\log}\avg{u_2}\avg{u_3}\\
+(E_{\rm avg}+ p_{\rm avg})\avg{u_2}\\
 \end{array}},\\
 \bm{f}_{3,S} = \LRp{\begin{array}{c}
-\avg{\rho}^{\log}\avg{w}\\
-\avg{\rho}^{\log}\avg{u}\avg{w}\\
-\avg{\rho}^{\log}\avg{v}\avg{w}\\
-\avg{\rho}^{\log}\avg{w}^2 + p_{\rm avg}\\
-(E_{\rm avg}+ p_{\rm avg})\avg{w}\\
+\avg{\rho}^{\log}\avg{u_3}\\
+\avg{\rho}^{\log}\avg{u_1}\avg{u_3}\\
+\avg{\rho}^{\log}\avg{u_2}\avg{u_3}\\
+\avg{\rho}^{\log}\avg{u_3}^2 + p_{\rm avg}\\
+(E_{\rm avg}+ p_{\rm avg})\avg{u_3}\\
 \end{array}},\nonumber
@@
 \end{gather*}
-with auxiliary quantities 
-\begin{align*}
-p_{\rm avg} &= \frac{\avg{\rho}}{2\avg{\beta}}, \qquad E_{\rm avg} = \frac{\avg{\rho}^{\log}}{2(\gamma-1)\avg{\beta}^{\log}} + \frac{1}{2}\avg{\rho}^{\log}\nor{\bm{u}}^2_{\rm avg}, \\
-\nor{\bm{u}}^2_{\rm avg} &= 2(\avg{u}^2 + \avg{v}^2 + \avg{w}^2) - \LRp{\avg{u^2} +\avg{v^2} + \avg{w^2}}.\nonumber
-\end{align*}
-
-In all problems, we estimate the timestep size $dt$ using $J, J^k_f$, and degree-dependent $L^2$ trace constants $C_N$ 
+%\note{where $u,v,w$ denote the $x,y,z$ velocities, respectively, and} 
+where the auxiliary quantities are defined as
+\begin{gather*}
+\note{\beta = \frac{\rho}{2p}}, \qquad p_{\rm avg} = \frac{\avg{\rho}}{2\avg{\beta}}, \qquad E_{\rm avg} = \frac{\avg{\rho}^{\log}}{2(\gamma-1)\avg{\beta}^{\log}} + \frac{1}{2}\avg{\rho}^{\log}\note{{u}^2_{\rm avg}}, \\
+\note{u^2_{\rm avg}} = 2(\avg{u_1}^2 + \avg{u_2}^2 + \avg{u_3}^2) - \LRp{\avg{u_1^2} +\avg{u_2^2} + \avg{u_3^2}}.\nonumber
+\end{gather*}
+
+In all problems, \note{we use an explicit low storage RK-45 time-stepper \cite{carpenter1994fourth}} and estimate the timestep size $dt$ using $J, J^k_f$, and degree-dependent $L^2$ trace constants $C_N$
 \[
@@
 \rho(\bm{x},t) &= \LRp{1 - \frac{\frac{1}{2}(\gamma-1)(\beta e^{1-r(\bm{x},t)^2})^2}{8\gamma \pi^2}}^{\frac{1}{\gamma-1}}, \qquad p = \rho^{\gamma},\\
-u(\bm{x},t) &= 1 - \frac{\beta}{2\pi} e^{1-r(\bm{x},t)^2}(y-y_0), \qquad v(\bm{x},t) = \frac{\beta}{2\pi} e^{1-r(\bm{x},t)^2}(x-x_0-t),\nonumber
+u_1(\bm{x},t) &= 1 - \frac{\beta}{2\pi} e^{1-r(\bm{x},t)^2}(y-y_0), \qquad u_2(\bm{x},t) = \frac{\beta}{2\pi} e^{1-r(\bm{x},t)^2}(x-x_0-t),\nonumber
 \end{align}
-where $u, v$ are the $x$ and $y$ velocity and $r(\bm{x},t) = \sqrt{(x-x_0-t)^2 + (y-y_0)^2}$.  Here, we take $x_0 = 5, y_0 = 0$ and $\beta = 5$.
+where $u_1, u_2$ are the $x$ and $y$ velocity and $r(\bm{x},t) = \sqrt{(x-x_0-t)^2 + (y-y_0)^2}$.  Here, we take $x_0 = 5, y_0 = 0$ and $\beta = 5$.
 
@@
 These warpings are constructed by modifying nodal positions according to the following mapping
+\note{
 \begin{align*}
-\tilde{x} &= x + \alpha\cos\LRp{\frac{\pi}{20}(x-10)}\cos\LRp{\frac{3\pi}{10}y}, \\
-\tilde{y} &= y + \alpha\sin\LRp{\frac{\pi}{5}(\tilde{x}-10)}\cos\LRp{\frac{\pi}{10}y}.
+\tilde{x} &= x + L_x\alpha\cos\LRp{\frac{\pi}{L_x}\LRp{x-\frac{L_x}{2}}}\cos\LRp{\frac{3\pi}{L_y}y}, \\
+\tilde{y} &= y + L_y\alpha\sin\LRp{\frac{4\pi}{L_x}\LRp{\tilde{x}-\frac{L_x}{2}}}\cos\LRp{\frac{\pi}{L_y}y},
 \end{align*}
-The lightly warped mesh corresponds to $\alpha = 1/32$, the moderately warped mesh corresponds to $\alpha = 1/8$, and the heavily warped mesh corresponds to $\alpha = 1/4$.
+where $L_x = 20, L_y = 10$ denote the lengths of the domain in the $x$ and $y$ directions, respectively.  
+The lightly warped mesh corresponds to $\alpha = 1/64$, the moderately warped mesh corresponds to $\alpha = 1/16$, and the heavily warped mesh corresponds to $\alpha = 1/8$.
 All results are computed using $C_{\rm CFL} = 1/2$ and Lax-Friedrichs interface dissipation.
+}
 
@@
 \rho(\bm{x},t) &= \LRp{1-\frac{(\gamma-1)}{2}\Pi^2}^{\frac{1}{\gamma-1}}\\
-\bm{u}(\bm{x},t) &= \Pi \bm{r}, \\
-E(\bm{x},t) &= \frac{p_0}{\gamma-1}\LRp{1-\frac{\gamma-1}{2}\Pi^2}^{\frac{\gamma}{\gamma-1}} + \frac{\rho}{2}\LRb{\bm{u}}.
+u_i(\bm{x},t) &= \Pi r_i, \\
+E(\bm{x},t) &= \frac{p_0}{\gamma-1}\LRp{1-\frac{\gamma-1}{2}\Pi^2}^{\frac{\gamma}{\gamma-1}} + \frac{\rho}{2}\sum_{i=1}^d u_i^2.
 \end{align*}
-where $\bm{u} = (u,v,w)^T$ is the velocity vector and % $\Pi$, $\bm{r}$ are defined as
-\[
-\Pi = \Pi_{\max}e^{\frac{1-\bm{r}^T\bm{r}}{2}}, \qquad \bm{r} = \begin{pmatrix}
+where $(u_1,u_2,u_3)^T$ is the velocity vector and % $\Pi$, $\bm{r}$ are defined as
+\[
+\Pi = \Pi_{\max}e^{\frac{1-\sum_{i=1}^dr_i^2}{2}}, \qquad \begin{pmatrix}r_1\\r_2\\r_3\end{pmatrix} = \begin{pmatrix}
 -(x_2-c_2-t)\\
@@
 \]
-We take $c_1 = c_2 = 5$, $p_0 = {1}/{\gamma}$, and $\Pi_{\max} = 0.4$, and solve on the domain $[0,10]\times [0,20]\times [0,10]$.
+We take \note{$c_1 = c_2 = 7.5$}, $p_0 = {1}/{\gamma}$, and $\Pi_{\max} = 0.4$, and solve on the domain \note{$[0,15]\times [0,20]\times [0,5]$ until final time $T=5$.
+We decompose the domain into uniform hexahedral elements with edge length $h$}.
 As in the 2D case, we also examine the effect of curvilinear mesh warping.
-Following \cite{chan2018discretely}, we construct a curved warping of the initial Cartesian mesh by mapping nodes on each hexahedron to warped nodal positions $(\tilde{x},\tilde{y},\tilde{z})$ through the transformation
+We construct a curved warping of the initial Cartesian mesh by mapping nodes on each hexahedron to warped nodal positions $(\tilde{x},\tilde{y},\tilde{z})$ through the transformation
 \begin{align*}
-\tilde{\bm{x}} &= \bm{x} + \sin\LRp{\pi \frac{(x-5)}{5}}\sin\LRp{2\pi \frac{(y-10)}{10}}\sin\LRp{\pi \frac{(z-5)}{5}}.
+\tilde{y} &= y + \note{\frac{1}{8}L_y\cos\LRp{3\pi \frac{(x-7.5)}{15}}\cos\LRp{\pi \frac{(y-10)}{20}}\cos\LRp{\pi \frac{(z-2.5)}{5}}}\\
+\tilde{x} &= x + \note{\frac{1}{8}L_x\cos\LRp{\pi \frac{(x-7.5)}{15}}\sin\LRp{4\pi \frac{(\tilde{y}-10)}{20}}\cos\LRp{\pi \frac{(z-2.5)}{5}}}\\
+\tilde{z} &= z + \note{\frac{1}{8}L_z\cos\LRp{\pi \frac{(\tilde{x}-7.5)}{15}}\cos\LRp{2\pi \frac{(\tilde{y}-10)}{20}}\cos\LRp{\pi \frac{(z-2.5)}{5}}}.
 \end{align*} 
+where $L_x = 15, L_y = 20, L_z = 5$.  
 On curved meshes, the geometric terms are constructed using the approach of Kopriva described in Section~\ref{sec:2}.  A $C_{\rm CFL} = .75$ is used for all experiments.
 
-Figure~\ref{fig:err3d} shows the $L^2$ errors for degrees $N = 2,\ldots, 7$.
-As in the 2D case, Gauss collocation schemes produce smaller errors than GLL collocation schemes, especially on on curved meshes.
-The difference between the two schemes on affine meshes is slightly more pronounced than in 2D.
-The difference between the two schemes on curved meshes is qualitatively similar to the 2D case, with Gauss collocation achieving higher accuracy than GLL collocation schemes at all polynomial degrees and mesh resolutions.
-%, which may be due to the specific curvilinear warping applied in this work.
+Figure~\ref{fig:err3d} shows the $L^2$ errors for degrees \note{$N = 2,\ldots, 5$}.\footnote{For the $N = 2$ GLL collocation scheme on the coarsest 3D mesh, the error is not shown because the solution diverged.
+} As in the 2D case, Gauss collocation schemes produce smaller errors than GLL collocation.
+The difference between the two schemes on affine meshes is slightly more pronounced than in 2D, while the difference between the two schemes on curved meshes is less significant than observed in 2D experiments.
+\note{We expect that this difference is due to the specific curved mapping.
+The warped 2D mesh used in Figure~\ref{fig:err2d} was generated to mimic a severe ``vorticular'' warping.
+The 3D mapping is less severely warped, due to different  domain dimensions and difficulties ensuring invertibility of the map from the reference to physical element.}
 
@@
 %    xmin=.0125, 
-    ymin=1e-5, ymax=5,
-    legend pos=north west, legend cell align=left, legend style={font=\tiny},	
+    ymin=1e-7, ymax=5,
+    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
     xmajorgrids=true, ymajorgrids=true, grid style=dashed,
@@
 }
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,2.62759)(0.125,0.728523)(0.0625,0.112225)(0.03125,0.018711)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,1.53203)(0.125,0.23283)(0.0625,0.0417033)(0.03125,0.00806707)};
-
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,0.934758)(0.125,0.191949)(0.0625,0.0119788)(0.03125,0.000558681)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,0.562488)(0.125,0.0864096)(0.0625,0.00268061)(0.03125,0.000136725)};
-
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,0.460425)(0.125,0.02445)(0.0625,0.000821249)(0.03125,4.67491e-05)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,0.239027)(0.125,0.0077433)(0.0625,0.000219705)(0.03125,2.67589e-05)};
+% N = 2
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(2.5,1.85364)(1.25,0.514785)(0.625,0.0793524)(0.3125,0.0132308)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(2.5,1.08061)(1.25,0.164679)(0.625,0.0294895)(0.3125,0.00570425)};
+
+% N = 3
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(2.5,0.661158)(1.25,0.135728)(0.625,0.00847132)(0.3125,0.000394629)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(2.5,0.397063)(1.25,0.0610903)(0.625,0.00189633)(0.3125,9.49625e-05)};
+
+% N = 4
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(2.5,0.325704)(1.25,0.0172973)(0.625,0.000580552)(0.3125,2.81526e-05)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(2.5,0.169055)(1.25,0.00547459)(0.625,0.000154222)(0.3125,7.74137e-06)};
 
@@
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,0.138105)(0.125,0.00744055)(0.0625,0.000126057)(0.03125,NaN)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,0.0547127)(0.125,0.00242675)(0.0625,4.2962e-05)(0.03125,NaN)};
-% N = 6
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,0.0489719)(0.125,0.000897467)(0.0625,2.55644e-05)(0.03125,NaN)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,0.0246519)(0.125,0.000324802)(0.0625,2.44009e-05)(0.03125,NaN)};
-% N = 7
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,0.023319)(0.125,0.00021951)(0.0625,NaN)(0.03125,NaN)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,0.00801774)(0.125,7.39203e-05)(0.0625,NaN)(0.03125,NaN)};
+coordinates{(2.5,0.0977112)(1.25,0.00526152)(0.625,8.73468e-05)(0.3125,9.44473e-07)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(2.5,0.0386913)(1.25,0.00171584)(0.625,2.46858e-05)(0.3125,2.07682e-07)};
 
@@
 }
-%\subfloat[Sub-parametric mesh]{
-%\begin{tikzpicture}
-%\begin{loglogaxis}[
-%    width=.5\textwidth,
-%    xlabel={Mesh size $h$},
-%%    ylabel={$L^2$ errors}, 
-%%    xmin=.0125, 
-%    ymin=1e-6, ymax=5,
-%      legend pos=south east, legend cell align=left, legend style={font=\tiny},	
-%    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
-%    legend entries={GLL, Gauss}    
-%]
-%\pgfplotsset{
-%cycle list={{blue, mark=*}, {red, dashed ,mark=square*}}
-%}
-%\addplot+[semithick, mark options={solid, fill=markercolor}]
-%coordinates{(0.25,2.80762)(0.125,0.951269)(0.0625,0.173495)(0.03125,0.0295367)};
-%\addplot+[semithick, mark options={solid, fill=markercolor}]
-%coordinates{(0.25,1.89337)(0.125,0.366475)(0.0625,0.0599646)(0.03125,0.0133605)};
-%\addplot+[semithick, mark options={solid, fill=markercolor}]
-%coordinates{(0.25,1.21946)(0.125,0.257914)(0.0625,0.0494569)(0.03125,0.00380144)};
-%\addplot+[semithick, mark options={solid, fill=markercolor}]
-%coordinates{(0.25,0.708125)(0.125,0.139733)(0.0625,0.015052)(0.03125,0.000419943)};
-%\addplot+[semithick, mark options={solid, fill=markercolor}]
-%coordinates{(0.25,0.594597)(0.125,0.100782)(0.0625,0.0051111)(0.03125,0.000186621)};
-%\addplot+[semithick, mark options={solid, fill=markercolor}]
-%coordinates{(0.25,0.347316)(0.125,0.0351013)(0.0625,0.000702893)(0.03125,4.39052e-05)};
-%\addplot+[semithick, mark options={solid, fill=markercolor}]
-%coordinates{(0.25,0.306622)(0.125,0.0198205)(0.0625,0.00156957)(0.03125,NaN)};
-%\addplot+[semithick, mark options={solid, fill=markercolor}]
-%coordinates{(0.25,0.142279)(0.125,0.00620177)(0.0625,0.000311399)(0.03125,NaN)};
-%\addplot+[semithick, mark options={solid, fill=markercolor}]
-%coordinates{(0.25,0.130865)(0.125,0.0108333)(0.0625,0.000188656)(0.03125,NaN)};
-%\addplot+[semithick, mark options={solid, fill=markercolor}]
-%coordinates{(0.25,0.0518407)(0.125,0.00289859)(0.0625,3.71549e-05)(0.03125,NaN)};
-%\addplot+[semithick, mark options={solid, fill=markercolor}]
-%coordinates{(0.25,0.0558827)(0.125,0.00251998)(0.0625,NaN)(0.03125,NaN)};
-%\addplot+[semithick, mark options={solid, fill=markercolor}]
-%coordinates{(0.25,0.0208791)(0.125,0.000423102)(0.0625,NaN)(0.03125,NaN)};
-%\end{loglogaxis}
-%\end{tikzpicture}
-%}\\
 \subfloat[Curved mesh]{
@@
 }
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,2.93541)(0.125,1.00445)(0.0625,0.18488)(0.03125,0.0323814)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,1.99429)(0.125,0.423426)(0.0625,0.0787544)(0.03125,0.0173475)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,1.23231)(0.125,0.266638)(0.0625,0.0513328)(0.03125,0.00390276)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,0.727022)(0.125,0.148442)(0.0625,0.0158392)(0.03125,0.000446109)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,0.599851)(0.125,0.101098)(0.0625,0.00511836)(0.03125,0.00018702)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,0.351163)(0.125,0.0353206)(0.0625,0.000705351)(0.03125,4.41811e-05)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,0.310649)(0.125,0.0199095)(0.0625,0.00157312)(0.03125,NaN)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,0.144996)(0.125,0.00619848)(0.0625,0.000311659)(0.03125,NaN)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,0.130579)(0.125,0.0108408)(0.0625,0.000188745)(0.03125,NaN)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,0.0517392)(0.125,0.00290387)(0.0625,3.71606e-05)(0.03125,NaN)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,0.0558635)(0.125,0.00251764)(0.0625,NaN)(0.03125,NaN)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(0.25,0.020886)(0.125,0.000421991)(0.0625,NaN)(0.03125,NaN)};
+% N = 2
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(2.5,NaN)(1.25,1.57425)(0.625,0.448088)(0.3125,0.0871033)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(2.5,2.48256)(1.25,0.981474)(0.625,0.232946)(0.3125,0.0529084)};
+
+% N = 3
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(2.5,2.03576)(1.25,0.599779)(0.625,0.101389)(0.3125,0.0120874)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(2.5,1.62319)(1.25,0.336149)(0.625,0.0485175)(0.3125,0.00371689)};
+
+% N = 4
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(2.5,1.22642)(1.25,0.240606)(0.625,0.0221903)(0.3125,0.000696802)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(2.5,0.90428)(1.25,0.131914)(0.625,0.00712465)(0.3125,0.000207085)};
+
+% N = 5
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(2.5,0.720447)(1.25,0.0986167)(0.625,0.00367923)(0.3125,0.000120058)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(2.5,0.524023)(1.25,0.0500179)(0.625,0.00121588)(0.3125,3.23497e-05)};
 \end{loglogaxis}
@@
 }
-\caption{$L^2$ errors for the 3D isentropic vortex for $N = 2,\ldots,7$ on sequences of Cartesian and curved meshes.}
+\caption{$L^2$ errors for the 3D isentropic vortex for \note{$N = 2,\ldots,5$} on sequences of Cartesian and curved meshes.}
 \label{fig:err3d}
@@
 The initial condition is taken to be the superposition of a stationary shock and a vortex propagating towards the right.
-The stationary Mach $M_s = 1.1$ shock is positioned at $x = .5$ normal to the $x$ axis, with left state $\LRp{\rho_L, u_L, v_L, p_L} = \LRp{1, \sqrt{\gamma}, 0, 1}$.
+The stationary Mach $M_s = 1.1$ shock is positioned at $x = .5$ normal to the $x$ axis, with left state $\LRp{\rho_L, u_L, v_L, p_L} = \LRp{1, \sqrt{\gamma}, 0, 1}$, \note{where $u_L, v_L$ denote $x,y$ components of the left-state velocity}.
 The right state is a scaling of the left state computed using the Rankine-Hugoniot conditions, such that the ratio of upstream and downstream states is
@@
 \begin{gather*}
-\rho = \rho_s \LRp{\frac{T_{\rm vor}}{T_L}}^{\frac{1}{\gamma-1}}, \qquad u = u_s + \delta u, \qquad v = v_s + \delta v, \qquad p = p_s \LRp{\frac{T_{\rm vor}}{T_L}}^{\frac{1}{\gamma-1}},
+\rho = \rho_s \LRp{\frac{T_{\rm vor}}{T_L}}^{\frac{1}{\gamma-1}}, \qquad u_1 = u_s + \delta u, \qquad u_2 = v_s + \delta v, \qquad p = p_s \LRp{\frac{T_{\rm vor}}{T_L}}^{\frac{1}{\gamma-1}},
 \end{gather*}
@@
 1 & 1 & 0 & 1\\
-\avg{u}-\bar{a} n_x & \avg{u} & n_y & \avg{u}+\bar{a} n_x\\
-\avg{v}-\bar{a} n_y & \avg{v} & -n_x & \avg{v}+\bar{a} n_y\\
-h-\bar{a} \bar{u}_n & \frac{1}{2}\nor{\bm{u}}^2_{\rm avg} & \bar{u}_n n_y - \avg{v}n_x & h + \bar{a} \bar{u}_n
+\avg{u_1}-\bar{a} n_x & \avg{u_1} & n_y & \avg{u_1}+\bar{a} n_x\\
+\avg{u_2}-\bar{a} n_y & \avg{u_2} & -n_x & \avg{u_2}+\bar{a} n_y\\
+h-\bar{a} \bar{u}_n & \frac{1}{2}u^2_{\rm avg} & \bar{u}_n n_y - \avg{u_2}n_x & h + \bar{a} \bar{u}_n
 \end{bmatrix}\\
@@
 \begin{gather*}
-\bar{u}_n = \avg{u}n_x + \avg{v}n_y, \qquad \bar{a} = \sqrt{\frac{\gamma p_{\rm avg}}{\avg{\rho}^{\log}}}, \qquad h = \frac{\gamma}{2(\gamma-1)\avg{\beta}^{\log}} + \frac{1}{2}\bar{u}, 
+\bar{u}_n = \avg{u_1}n_x + \avg{u_2}n_y, \qquad \bar{a} = \sqrt{\frac{\gamma p_{\rm avg}}{\avg{\rho}^{\log}}}, \qquad h = \frac{\gamma}{2(\gamma-1)\avg{\beta}^{\log}} + \frac{1}{2}\bar{u}, 
 \end{gather*}
-and $p_{\rm avg}, \nor{\bm{u}}^2_{\rm avg}$ are defined as in (\ref{eq:fluxaux}).
+and $p_{\rm avg}, u^2_{\rm avg}$ are defined as in (\ref{eq:fluxaux}).
 
 %It is known that the Lax-Friedrichs flux is fairly dissipative compared to more nuanced fluxes \cite{harten1983upstream}, though this dissipation is less pronounced at higher polynomial degrees in practice.
-Figure~\ref{fig:shockvort} shows solutions for $N=4$ and $h = 1/100$ Gauss collocation schemes using a non-dissipative entropy conservative flux, a dissipative Lax-Friedrichs flux, and a matrix dissipation flux.
+Figure~\ref{fig:shockvort} shows density solutions for $N=4$ and $h = 1/100$ Gauss collocation schemes using a non-dissipative entropy conservative flux, a dissipative Lax-Friedrichs flux, and a matrix dissipation flux.
 In all cases, the vortex passes through the shock stably without the use of additional slope limiting, filtering, or artificial viscosity.
@@
 \rho = 1, \qquad p = \frac{100}{\gamma} + \frac{1}{16} \LRp{\cos(2x_1) + \cos(2x_2)}\LRp{2+\cos(2x_3)},\\
-u = \sin(x_1)\cos(x_2)\cos(x_3), \qquad
-v = -\cos(x_1)\sin(x_2)\cos(x_3),\qquad
-w= 0.
+u_1 = \sin(x_1)\cos(x_2)\cos(x_3), \qquad
+u_2 = -\cos(x_1)\sin(x_2)\cos(x_3),\qquad
+u_3 = 0.
 \end{gather*}
@@
 \[
-\kappa(t) =\frac{1}{\LRb{\Omega}} \int_{\Omega} \rho \bm{u}\cdot\bm{u} \diff{\bm{x}},
+\kappa(t) =\frac{1}{\LRb{\Omega}} \int_{\Omega} \rho \note{ \LRp{\sum_{i=1}^d u_i^2}} \diff{\bm{x}},
 \]
@@
 
-Jesse Chan is supported by NSF DMS-1719818 and DMS-1712639.  The authors thank Florian Hindenlang for helpful discussions and suggestions.
+Jesse Chan is supported by NSF DMS-1719818 and DMS-1712639.  The authors thank Florian Hindenlang and Jeremy Kozdon for helpful discussions and suggestions.
 

### 2019-03-07 docs/skewform/arxiv/main.tex

--- 2019-02-14 docs/skewform/main.tex
@@
-\pdfoutput=1
+%\pdfoutput=1
 
@@
 %\documentclass[twocolumn]{svjour3}          % twocolumn
+%\usepackage[T1]{fontenc}
 
@@
 %\spnewtheorem{remark}{Remark}{\bfseries}{\itshape}
+
+%\usepackage{fullpage}
 \usepackage{cite}
 
-\usepackage{slashbox} % for slashbox in tables
-
-\usepackage{amsmath, amssymb, amsfonts}
+%\usepackage{slashbox} % for slashbox in tables
+
+\usepackage{amsmath, amsfonts, amssymb}
 %\newtheorem*{remark}{Remark}
@@
 \usepackage{tikz}
-\usepackage{hhline}
+%\usepackage{hhline}
 \usepackage{graphicx, color, bm}
@@
 \usepackage{pgfplotstable}
-\usepackage{slopetri} % my own package for drawing triangle slopes
+\definecolor{markercolor}{RGB}{124.9, 255, 160.65}
+%\usepackage{slopetri} % my own package for drawing triangle slopes
 
@@
 \author{Jesse Chan}
-\title{Skew-symmetric entropy stable discontinuous Galerkin formulations with applications to hybrid meshes}
+\title{Skew-symmetric entropy stable modal discontinuous Galerkin formulations}% with applications to hybrid meshes}
 \titlerunning{Skew-symmetric entropy stable DG formulations}
@@
 \begin{abstract}
-Entropy stable high order methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.
-The construction of such schemes has relied on the use of carefully chosen collocation points \cite{gassner2013skew, fisher2013high, carpenter2014entropy, crean2018entropy, chan2018efficient} or volume and surface quadrature rules \cite{chan2017discretely, chan2018discretely} to produce operators which satisfy a summation-by-parts (SBP) property.
-In this work, we show how to construct skew-symmetric schemes which are entropy stable even for volume and surface quadratures under which a traditional SBP property does not hold.
-These skew-symmetric formulations avoid the use of a ``strong'' SBP property, and require only that operators exactly differentiate constants and satisfy a discrete form of the fundamental theorem of calculus.
-We conclude with an application of the skew-symmetric formulation for entropy stable schemes on mixed quadrilateral-triangle meshes.
+High order entropy stable discontinuous Galerkin (DG) methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.
+The construction of such schemes has relied on the use of carefully chosen nodal points \cite{gassner2013skew, fisher2013high, carpenter2014entropy, crean2018entropy, chan2018efficient} or volume and surface quadrature rules \cite{chan2017discretely, chan2018discretely} to produce operators which satisfy a summation-by-parts (SBP) property.
+In this work, we show how to construct ``modal'' skew-symmetric DG formulations which are entropy stable for volume and surface quadratures under which a traditional SBP property does not hold.
+These skew-symmetric formulations avoid the use of a ``strong'' matrix-based SBP property, and instead rely on the assumption that discrete operators exactly differentiate constants and satisfy a discrete form of the fundamental theorem of calculus.
+We conclude with numerical experiments verifying the accuracy and stability of the proposed formulations, and discuss an application of skew-symmetric formulations for entropy stable DG schemes on mixed quadrilateral-triangle meshes.
 \end{abstract}
 
-\tableofcontents
+%\tableofcontents
 
@@
 Proofs of continuous entropy inequalities rely on the chain rule, which does not hold discretely due to effects such as quadrature error.
-By using an approach referred to as ``flux differencing'', entropy stable schemes account for the loss of the chain rule while maintaining a semi-discrete analogue of the continuous entropy inequality.
+Entropy stable schemes circumvent the loss of the chain rule while maintaining a semi-discrete analogue of the continuous entropy inequality by using an approach referred to as ``flux differencing''.
 These schemes were first introduced as high order collocation methods on tensor product elements in \cite{fisher2013high, carpenter2014entropy, gassner2016split, gassner2017br1}, and were extended to simplicial elements in \cite{crean2017high, chen2017entropy, crean2018entropy, chan2017discretely, chan2018discretely}.
@@
 
-Entropy stable discretizations are built upon flux differencing and a summation-by-parts (SBP) property, which serves as a discrete analogue of integration by parts for quadrature-based discretization matrices.
-However, the SBP property does not hold for some under-integrated quadrature rules, which naturally arise in certain discretization settings.
+Entropy stable ``modal'' DG discretizations \cite{chan2017discretely, chan2018discretely} are built upon flux differencing and a summation-by-parts (SBP) property, which serves as a discrete analogue of integration by parts for quadrature-based discretization matrices.
+However, the SBP property does not hold for certain under-integrated quadrature rules, which arise naturally in some discretization settings.
 For example, on hybrid meshes consisting of both quadrilateral and triangular elements, it is convenient to utilize the same quadrature rule on shared faces between different element types.
-On degree $N$ tensor product elements, a popular choice of quadrature is an $(N+1)$-point Gauss-Legendre-Lobatto (GLL) rule.
+On degree $N$ quadrilateral elements, a popular choice of quadrature is an $(N+1)$-point Gauss-Legendre-Lobatto (GLL) rule.
 When both volume and surface integrals are approximated using $(N+1)$ point GLL quadrature rules, the SBP property holds, despite the fact that GLL quadrature is inexact for the integrands which appear in finite element formulations \cite{fisher2013high}.
-However, while GLL quadrature induces an SBP property on quadrilateral elements, but does not always result in an SBP property if used for triangular elements \cite{chan2017discretely}.
-%A similar issue arises on non-conforming meshes, where a composite ``split-side'' quadrature naturally when coupling larger elements to smaller elements through a non-conforming interface \cite{kozdon2018energy}.
-For quadrilaterals utilizing a GLL volume quadrature, the use of a composite quadrature on a non-conforming face also results in the loss of the SBP property \cite{chan2017discretely}.
+However, while GLL quadrature induces an SBP property on quadrilateral elements, it does not guarantee an SBP property if used on triangular elements \cite{chan2017discretely}.
 
@@
 
-It should be noted that a similar approach to entropy stable discretizations was introduced within a finite difference framework \cite{chen2017entropy, crean2018entropy} using multidimensional differencing operators which are constructed to satisfy accuracy conditions and an SBP property \cite{hicken2016multidimensional}.
-These operators do not correspond to any specific basis or approximation space, but can be shown to exist for nodal points corresponding to sufficiently accurate choices of volume and surface quadrature and can be computed either algebraically or through an optimization problem.
-The formulations presented in this work require similar accuracy conditions, but differ from those of \cite{chen2017entropy, crean2018entropy} in that we start with an explicit approximation space, which then induce quadrature-based operators.
-The formulations presented here also accomodate general choices of volume and surface quadrature (e.g.\ volume quadratures without boundary nodes and over-integrated rules) while ensuring compact coupling conditions between neighboring elements.
+It should be noted that a similar approach to entropy stable discretizations was introduced within a finite difference framework \cite{chen2017entropy, crean2018entropy} using multidimensional differencing operators which satisfy accuracy conditions and an SBP property \cite{hicken2016multidimensional}.
+These operators exist for nodal points corresponding to sufficiently accurate choices of volume and surface quadrature, but do not correspond to any specific basis or approximation space.
+The formulations in \cite{chen2017entropy, crean2018entropy} differ from the ones presented in this work in that they are based on SBP finite differences and ``nodal'' (rather than ``modal'') DG formulations, with differentiation operators computed algebraically or through an optimization problem for each specific choice of nodes.
+In contrast, ``modal'' formulations induce quadrature-based operators from an explicit approximation space, and accomodate general choices of volume and surface quadrature (e.g.\ volume quadratures without boundary nodes and over-integrated quadrature rules).
 
@@
 Section~\ref{sec:skew1} then introduces the variational SBP property and an entropy conservative skew-symmetric formulation.
-Section~\ref{sec:skew2} extends the skew-symmetric formulation to curved elements, and Section~\ref{sec:num} presents numerical experiments which verify the theoretical assumptions, stability, and accuracy of the proposed formulations.
+Section~\ref{sec:skew2} extends the skew-symmetric formulation to curved elements, and Section~\ref{sec:quadacc} provides explicit conditions for entropy stability in terms of quadrature accuracy and the polynomial degree of geometric mappings.
+Section~\ref{sec:num} concludes by presenting numerical experiments which verify the theoretical assumptions, stability, and accuracy of the proposed formulations.
 
@@
 As a result, discretizations of (\ref{eq:nonlineqs}) do not typically possess an underlying statement of stability.
-For low order schemes, this can be offset in practice by the inherent numerical dissipation.
-However, because high order discretizations possess low numerical dissipation, the lack of an underlying discrete stability has been conflated with the idea that high order methods are inherently less stable than low order methods.
+This can be offset in practice by the numerical dissipation inherent in lower order schemes.
+However, because high order discretizations possess low numerical dissipation, the lack of an underlying discrete stability has contributed to the perception that high order methods are inherently less stable than low order methods.
 
@@
 We define an approximation space using degree $N$ polynomials on the reference element; however, the natural polynomial approximation space differs depending on the element type \cite{chan2015gpu}.
-
-On a $d$-dimensional reference simplex, the natural polynomial space are total degree $N$ polynomials 
+On a $d$-dimensional reference simplex, the natural polynomial space consists of total degree $N$ polynomials
 \[
@@
 
-The proofs presented in this work will also use anisotropic tensor product polynomial spaces, where the maximum polynomial degree varies depending on the coordinate direction.
+The proofs presented in this work will also refer to anisotropic tensor product polynomial spaces, where the maximum polynomial degree varies depending on the coordinate direction.
 We denote such spaces by $Q^{N_1, \ldots, N_d}$, where $N_k$ are non-negative integers and
@@
 It was shown in \cite{chan2017discretely, chan2018efficient} that, when combined with projection and lifting matrices, $\bm{D}^i_N$ produces a high order polynomial approximation of $f\pd{g}{x_i}$. 
-Let $f, g$ be differentiable functions, and let $\bm{f}_i = f(\bm{x}_i)$, $\bm{g}_i = g(\bm{x}_i)$ denote values of $f,g$ at volume and surface quadrature points.  Then,
+Let $f, g$ be differentiable functions, and let $\bm{f}_i = f(\bm{x}_i)$, $\bm{g}_i = g(\bm{x}_i)$ denote values of $f,g$ at volume and surface quadrature points.
+Then, $f\pd{g}{x_i}$ can be approximated by the polynomial with modal coefficients given by
 \begin{gather}
@@
 \begin{theorem}%[Theorem 1 in \cite{chan2017discretely}]
-Assume that $\bm{Q}^i$ satisfies a generalized SBP property
-\[
-\bm{Q}^i = \bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T.
-\]
-Then, the decoupled SBP operator $\bm{D}^i_N$ (\ref{eq:QN}) satisfies a summation by parts property:
+If $\bm{Q}^i$ satisfies a generalized SBP property
+\[
+\bm{Q}^i = \bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T,
+\]
+then the decoupled SBP operator $\bm{D}^i_N$ (\ref{eq:QN}) satisfies a summation by parts property:
 \begin{gather}
@@
 The proof is a straightforward extension of Theorem 1 in \cite{chan2017discretely} to polynomial approximation spaces on non-simplicial elements.  
-\end{proof}
+\qed\end{proof}
 
@@
 \begin{remark}
-It should be stressed that sufficiently accurate volume and surface quadratures are not necessary conditions for Theorem~\ref{lemma:dsbp}.
+It should be stressed that the accuracy conditions on volume and surface quadratures are sufficient but not necessary conditions for Theorem~\ref{lemma:dsbp}.
 For example, it is well known that the use of $(N+1)$ point Gauss-Legendre-Lobatto (GLL) rules for both volume and surface quadratures result in a generalized SBP property, despite the fact that these rules are only accurate for degree $(2N-1)$ polynomials.
@@
 
-Throughout the remainder of this work, we make the following assumptions:
+Throughout the remainder of this work, we will make the following assumption for specific functions $v(\bm{x})$:
 \begin{assumption}
@@
 \]
-\end{proof}
+\qed\end{proof}
 
@@
 \end{align*}
-\end{proof}
+\qed\end{proof}
 
@@
 %The final step of the proof is  where we have used Lemma~\ref{lemma:vdsbp} to integrate by parts and conclude that $\bm{1}^T\LRp{\bm{Q}^i_N}^T = 0$.
-\end{proof}
+\qed\end{proof}
 
@@
 %\end{align}
-\end{proof}
+\qed\end{proof}
 
@@
 Summing $\pd{}{t}\LRp{\bm{u}_h^T\bm{M}^k\bm{u}_h} = 0$ over all elements ensures that globally constant solutions are stationary and that $\pd{\bm{u}_h}{t} = 0$.
-\end{proof}
+\qed\end{proof}
 
@@
 \section{Conditions on the accuracy of volume and surface quadrature rules}
+\label{sec:quadacc}
 
 The previous sections outline minimal conditions under which entropy stability is guaranteed under a skew-symmetric formulation and a polynomial geometric mapping.
-In Section~\ref{sec:curvgeo}, we translate these minimial conditions into conditions on quadrature accuracy.
+In Section~\ref{sec:quadacc}, we translate these minimial conditions into conditions on quadrature accuracy.
 
@@
 \subsection{On quadrature conditions for Assumption~\ref{ass:quad} for $v = 1$ and $v = G^k_{ij}$}
-\label{sec:curvgeo}
 
@@
 The volume quadrature exactly integrates the first integrand for $N_{\rm geo} \leq N+1$, while the surface quadrature exactly integrates the second integrand for $M \geq N_{\rm geo}-1$, or $N_{\rm geo} \leq M+1$.
-\end{proof}
+\qed\end{proof}
 
@@
 The surface quadrature rule exactly integrates such integrands for $M \geq N_{\rm geo} - 1$, or $N_{\rm geo} \leq M+1$.
-\end{proof}
+\qed\end{proof}
 
@@
 &= -\frac{1}{2}
-\bm{e}_q^T\bm{Q}^i\bm{u}_q + \frac{1}{2}\bm{e}_q^T\LRp{\bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T}\bm{u}_q, \qquad \bm{e}_q = \bm{V}_q\bm{e}. \nonumber
+\bm{e}_q^T\bm{Q}^i\bm{u}_q + \frac{1}{2}\bm{e}_q^T\LRp{\bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T}\bm{u}_q, \nonumber
 \end{align}
-The remainder of the proof is similar to the proof of Lemma~\ref{lemma:vsbp}.  Since $u \in P^M$ and $e \in P^N$, by exactness of the quadrature rules, 
+where we have introduced $\bm{e}_q = \bm{V}_q\bm{e}$.
+The remainder of the proof is similar to the proof of Lemma~\ref{lemma:vsbp}.
+Since $u \in P^M$ and $e \in P^N$, by exactness of the quadrature rules,
 \[
@@
 Combining this with (\ref{eq:skewerr}) implies that $\bm{e}^T\bm{M}\bm{e} = 0$, and since $\bm{M}$ is symmetric positive definite, $\bm{e} = 0$.  
-\end{proof}
+\qed\end{proof}
 
 Lemma~\ref{lemma:dsbpapprox} suggests that, when a generalized SBP property does not hold, the use of under-integrated quadratures results in a loss of one or more orders of accuracy.
-For example, using $(N+1)$ point GLL rules (which are exact for only polynomials of degree $2N-1$) for either volume or surface quadratures should result in a loss of one order of accuracy compared to the use of $(N+1)$-point Gauss rules (which are exact for polynomials of degree $2N$).
+For example, if the SBP property does not hold, then using $(N+1)$ point GLL rules (which are exact for only polynomials of degree $2N-1$) for either volume or surface quadratures should result in a loss of one order of accuracy compared to the use of $(N+1)$-point Gauss rules (which are exact for polynomials of degree $2N$).
 This is indeed observed in numerical experiments.
@@
 where we have introduced the pressure is $p = (\gamma-1)\LRp{E - \frac{1}{2}\rho (u^2+v^2)}$ and the specific internal energy $\rho e = E - \frac{1}{2}\rho (u^2+v^2)$.
-
 We seek entropy stability with respect to the entropy for the compressible Navier-Stokes equations \cite{hughes1986new}
@@
 These constants scale proportionally to $N^2$, though precise values of $C_I, C_T$ vary slightly depending on the choice of volume or surface quadrature used.
-The dependence of $C_I, C_T$ on quadrature is discussed in more detail in Appendix~\ref{sec:consts}, where computed values of $C_I, C_T$ are also given.
+The dependence of $C_I, C_T$ on quadrature is discussed in more detail in Appendix~\ref{sec:consts}, where computed values are also given.
 
@@
 We refer to this quantity as the entropy right-hand side (RHS).
-Section~\ref{sec:curvgeo} outlines conditions on quadrature accuracy which guarantee that the formulations (\ref{eq:esdgSkew}) and (\ref{eq:skewformcurved}) are discretely entropy conservative.
+Section~\ref{sec:quadacc} outlines conditions on quadrature accuracy which guarantee that the formulations (\ref{eq:esdgSkew}) and (\ref{eq:skewformcurved}) are discretely entropy conservative.
 These numerical results confirm that these conditions are tight.
@@
 On quadrilaterals, we fix the volume quadrature to be an $(N+1)$ point GLL quadrature.
-To verify the conditions given in Section~\ref{sec:curvgeo}, we vary the accuracy of the 2D surface quadrature rule.
+To verify the conditions given in Section~\ref{sec:quadacc}, we vary the accuracy of the 2D surface quadrature rule.
 
@@
 We evolve the solution until final time $T = 1.0$ on a domain $[0,15]\times[-.5,.5]$ using the skew symmetric formulation with $N = 6$ and a CFL of $1/2$.
-Table~\ref{table:ecrhs} shows the maximum entropy RHS over the duration of the simulation.We observe that for all $N_{\rm geo} \leq M+1$, the maximum entropy RHS is $O(10^{-14})$ and at the level of machine precision.
+Table~\ref{table:ecrhs} shows the maximum entropy RHS over the duration of the simulation.
+We observe that for all $N_{\rm geo} \leq M+1$, the maximum entropy RHS is $O(10^{-14})$ and at the level of machine precision.
 When $N_{\rm geo}> M+1$, we observe that the maximum entropy RHS increases .
@@
 $M = 5$ & 1.74e-14 & 1.06e-14 & 1.31e-14 &  1.35e-14 & 1.06e-14 & 1.31e-14\\
+\hline
 $M = 3$ & 2.89e-14 & 2.59e-14 &  2.83e-14 & 2.43e-14 & 3.20e-05 & 3.27e-05 \\
@@
 }
-\caption{Maximum absolute value of the entropy RHS over $t\in [0,1]$ for degree $N=6$ approximations on triangular and quadrilateral meshes.
+\caption{Maximum absolute value of the entropy RHS for degree $N=6$ over $t\in [0,1]$ on triangular and quadrilateral meshes.
 The volume quadrature for the quadrilateral mesh is taken to be $(N+1)$-point GLL quadrature.
@@
 For $N=4$, Option~\ref{opt:2} achieves the same level of error observed for Option~\ref{opt:3}.
-This suggests that Lemma~\ref{lemma:dsbpapprox} may be sufficient but not necessary for full order accuracy.
+This suggests that Lemma~\ref{lemma:dsbpapprox} may be sufficient but not necessary for full order accuracy.\footnote{We note that these results may also differ depending on the type of interface dissipation applied \cite{hindenlang2019order}.}
 
@@
 Because these quadrature rules do not necessarily induce operators which satisfy a ``strong'' matrix summation-by-parts property, we generalize proofs of entropy stability and free stream (constant) preservation using a weaker ``variational'' SBP property.
+
 This variational SBP property is also used to construct explicit conditions relating between the order of accuracy and degree of accuracy of each quadrature rule.
-Finally, we derive conditions between the degree of quadrature accuracy and polynomial degree of the geometric mapping under which the skew-symmetric formulation is entropy stable on curved meshes.
+We also derive conditions under which the skew-symmetric formulation is entropy stable on curved meshes in terms of the degree of quadrature accuracy and polynomial degree of the geometric mapping.
 Numerical experiments confirm the entropy stability and high order accuracy of the proposed schemes on triangular, quadrilateral, and 2D hybrid meshes.

### 2019-03-07 docs/skewform/letter.tex

\documentclass{letter}
\usepackage{fullpage}
\signature{Jesse Chan}
\address{6100 Main St,\\
Houston, TX 77005}

\begin{document}
\begin{letter}{Journal of Scientific Computing}

\opening{Dear editor,}

Please find enclosed the manuscript, 
\begin{center}
Jesse Chan\\
\textit{Skew-symmetric entropy stable modal discontinuous Galerkin formulations},
\end{center}
which we would like to submit for publication as a research paper in the Journal of Scientific Computing (JSC).

The submitted manuscript introduces a new modal formulation for discretely entropy stable discontinuous Galerkin (DG) methods.  Previous work by the author has constructed discretely entropy stable high order DG methods using operators which obey a summation-by-parts (SBP) property.  These operators are constructed using numerical quadrature, and sufficiently accurate quadrature rules are a necessary condition for the SBP property to hold.  The submitted work introduces a skew-symmetric formulation which achieves entropy stability under relaxed quadrature conditions, which appear in couplings on hybrid meshes between different types of elements.  The stability and accuracy of the new formulation is tested on 2D meshes consisting of both quadrilateral and triangular elements.

We hope that the method and results discussed in this manuscript would appeal to the readership of JSC.  The author confirms that this manuscript has not been published elsewhere and is not under consideration by another journal.  We look forward to hearing from you at your earliest convenience.

\closing{Best regards}

\end{letter}

\end{document}

### 2019-08-06 docs/ESDG_ROM/main.tex
updating paper

\pdfoutput=1
%\documentclass[review,onefignum,onetabnum,final]{siamart171218}
\documentclass[review]{siamart171218}
%\documentclass{article}

%\usepackage{fullpage}
\usepackage{cite}
%\usepackage{geometry}

\usepackage{amsmath, amsfonts, amssymb}
\DeclareMathOperator*{\argmin}{\arg\,\min}
\DeclareMathOperator*{\argmax}{\arg\,\max}
\usepackage{algorithmic}

\newtheorem*{remark}{Remark}
\theoremstyle{assumption}
\newtheorem{assumption}{Assumption}

%\newtheorem{theorem}{Theorem}
%\newtheorem{remark}{Remark}
%\newenvironment{proof}{\paragraph{Proof:}}{\hfill$\square$}
%\newtheorem{assumption}{Assumption}
%\newtheorem{lemma}{Lemma}

\usepackage{hyperref}
\usepackage[titletoc,toc,title]{appendix}
\usepackage{tikz}
%\usepackage{hhline}
\usepackage{graphicx, color, bm}
\usepackage{subfig}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{slopetri} % my own package for drawing triangle slopes

\renewcommand{\hat}{\widehat}
\renewcommand{\tilde}{\widetilde}

%% d in integrand
\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}}
\newcommand{\diag}[1]{{\rm diag}\LRp{#1}}
\newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}{\rm #2}}}
\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRb}[1]{\left| #1 \right|}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}
\newcommand{\LRceil}[1]{\left\lceil #1 \right\rceil}
\newcommand{\LRl}[1]{\left. #1 \right|}
\newcommand{\pdn}[3]{\frac{\partial^{#3}#1}{\partial#2^{#3}}}
\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\jump}[1] {\ensuremath{\llbracket#1\rrbracket}}
\newcommand{\avg}[1] {\ensuremath{\LRc{\!\{#1\}\!}}}

\newcommand{\note}[1]{{\color{blue}{#1}}}

\graphicspath{{./figs/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\date{}
\author{Jesse Chan}
\title{Entropy stable reduced order modeling of nonlinear conservation laws}

\begin{document}

\maketitle

\begin{abstract}
Reduced order models of nonlinear conservation laws in fluid dynamics do not typically inherit stability properties of the full order model. We discuss the construction of reduced models of nonlinear conservation laws which retain accuracy while inheriting a semi-discrete entropy inequality independently of discretization resolutions and parameters. 
\end{abstract}

\tableofcontents

%\begin{itemize}
%\item Intro
%\item Entropy conservation for FOM
%\item Entropy conservation for POD
%\item Entropy stability via entropy dissipation
%\item Entropy stable hyper-reduction
%\item Proof: accuracy of projected operator
%\item Proof: entropy stability for periodic domain
%\item Proof: conservation if $\bm{1}$ is contained in $\bm{V}_r$
%\item Non-periodic domains and constrained hyper-reduction

%\item Numerical experiments
%\begin{itemize}
%\item Singular values of conservative snapshots vs conservative + entropy variable snapshots.
%\item 1D experiments, shocks, error convergence plots
%\item Effect of different hyper-reduction techniques (optimizing for $V_{\rm test} V_{\rm test}$ and $V_r V_{\rm test}$ vs $V_r V_r$)
%\item Comparison of entropy stable viscosity vs regular hyper-reduced viscosity
%\item Comparison of snapshots from entropy stable FOM vs regular FOM.  
%\item 2D experiments: KH and acoustic pulse
%\end{itemize}
%\end{itemize}

\section{Introduction}

Projection-based model reduction constructs low-dimensional surrogate models for many-query scenarios (e.g., simulations over multiple parameter values) that can be evaluated over a range of parameters at a low \textit{online} cost in exchange for a more expensive \textit{offline} pre-computation step \cite{benner2015survey}.  %Reduced order models (ROMs) can be constructed, for example, by performing a Galerkin projection of the original full order model onto a reduced basis.  
The development of reduced order models (ROMs) is relatively mature for several classes of problems (e.g., linear time-invariant systems, coercive elliptic PDEs).  However, the construction of robust and stable ROMs for transient and convection-dominated problems remains an active area of research \cite{cagniart2019model}.

For certain PDEs, ``structure-preserving'' ROMs provide robustness and stability by preserving energetic properties of the full system at the discrete level.  ROMs which retain either Lagrangian \cite{lall2003structure, carlberg2015preserving} or Hamiltonian structure \cite{polyuga2010structure, benner2012interpolation, gugercin2012structure, peng2016symplectic, chaturantabut2016structure, gong2017structure, afkham2017structure, afkham2018structure} have been constructed by combining Galerkin projection with an appropriate formulation of the full order model, and similar energy-conserving ROMs have been constructed for the incompressible Navier-Stokes equations \cite{farhat2014dimensional, farhat2015structure}.  The construction of structure-preserving ROMs to nonlinear conservation laws in fluid flow, however, remains an open problem.  For example, Galerkin projection yields ROMs which become unstable as the number of reduced basis functions (modes) is increased \cite{bui2007goal, carlberg2013gnat}.  As a result, stabilized discretizations are often employed.  Petrov-Galerkin ROMs, which use an alternative test basis \cite{maday2002blackbox, rozza2007stability, bui2007goal, amsallem2012stabilization, rozza2013reduced, ballarin2015supremizer, carlberg2017galerkin}, are a popular alternative, as are additional residual-based stabilization or dissipation terms \cite{wang2012proper, kalashnikova2014stabilization, caiazzo2014numerical, balajewicz2016minimal}.  Such models improve robustness in practice, though they do not provide a theoretical foundation for stability.

The construction of stable ROMs for the compressible Navier-Stokes equations is complicated by their complex structure.  Thus, practitioners often develop structure-preserving entropy stable ROMs for simplifications of the underlying PDE (such as the time-dependent \textit{linearized} compressible Navier-Stokes equations \cite{barone2009stable, kalashnikova2010stability, serre2012reliable, amsallem2012stabilization, kalashnikova2014construction}) and extrapolate such models to the full nonlinear equations.  A promising alternative is to enforce physical conditions such as such as kinetic energy preservation, which are empirically related to stability \cite{maboudi2018conservative}.  These methods significantly improve the robustness of numerical methods in practice, and can guarantee a discrete entropy inequality for systems of nonlinear conservation laws which yield stable split formulations \cite{fisher2013discretely}.

The approach taken in this work differs in the treatment of nonlinear terms.  In \cite{maboudi2018conservative}, equations are rewritten such that the nonlinear terms involve only polynomial (quadratic and cubic) nonlinearities, which can be evaluated exactly using precomputed matrices.  In this work, we approximate nonlinear terms using hyper-reduction techniques based on empirically computed quadrature rules \cite{an2008optimizing, hernandez2017dimensional}.  The hyper-reduction approach incurs approximation error; however, it is allows for the generalization of the split forms in \cite{maboudi2018conservative} to nonlinearly entropy stable formulations.

%
%Hyper-reduction methods
%\begin{itemize}
%\item Interpolatory: DEIM and QR-DEIM \cite{chaturantabut2010nonlinear, drmac2016new}, gappy POD \cite{bui2004aerodynamic, carlberg2011efficient}
%\item Element-based approaches: ECSW \cite{farhat2014dimensional, farhat2015structure}, EQP \cite{ryu2015extensions, yano2019lp, patera2017lp}
%\item Empirical cubature: \cite{an2008optimizing, hernandez2017dimensional}.  
%\end{itemize}
%
%Structure-preserving ROMs: 
%\begin{itemize}
%\item Linear Hamlitonian \cite{beattie2009interpolatory} 
%\item Nonlinear Hamiltonian systems: Gugercin \cite{gugercin2012structure}, Carlberg \cite{carlberg2015preserving}, Chaturantabut \cite{chaturantabut2016structure}, Gong \cite{gong2017structure}, Afkham et al.\ \cite{afkham2017structure, afkham2018structure}
%\item Conservation through explicit constraints over sub-meshes \cite{carlberg2017structure, carlberg2018conservative}
%\end{itemize}

In this work, we directly constructing entropy stable ROMs for nonlinear conservation laws through a modified Galerkin projection of an appropriate full order model.  The approach taken in this paper combines techniques from entropy stable finite volume schemes \cite{tadmor1987numerical, tadmor2003entropy} and entropy stable summation-by-parts (SBP) discretizations \cite{fisher2013high, carpenter2014entropy, chen2017entropy, crean2018entropy, chan2017discretely, chan2018discretely, chan2019skew} to produce entropy stable reduced order models.  The paper is organized as follows.  Section~\ref{sec:1} discusses the entropy inequality associated with systems of nonlinear conservation laws.  Section~\ref{sec:2} introduces a full order model on 1D periodic domains based on entropy stable finite volume schemes.  Section~\ref{sec:3} describes how to construct an entropy stable reduced basis approximation, while Section~\ref{sec:4} discusses entropy stable hyper-reduction techniques to reduce costs associated with the evaluation of nonlinear terms.  Section~\ref{sec:6} describes how to extend the aforementioned approaches to non-periodic boundary conditions, and Section~\ref{sec:7} describes the extension to higher dimensions.  We conclude in Section~\ref{sec:8} with numerical experiments verifying theoretical results.

\section{Entropy inequalities for nonlinear conservation laws}
\label{sec:1}
Let $\Omega$ denote some domain with boundary $\partial \Omega$.  General nonlinear conservation laws are expressed as a system of partial differential equations (PDEs) 
\begin{equation}
\pd{\bm{u}}{t}  + \sum_{i=1}^d\pd{\bm{f}_i(\bm{u})}{x_i} = 0, \qquad 
S(\bm{u}) \text{ is a convex function}, \qquad
\bm{v}(\bm{u}) = \pd{S}{\bm{u}},
\label{eq:nonlineqs}
\end{equation}
where $\bm{u}\in \mathbb{R}^n$ are the conservative variables, $\bm{f}_j$ are nonlinear fluxes, and $\bm{v}(\bm{u})$ are the \textit{entropy variables}.  By multiplying (\ref{eq:nonlineqs}) by the entropy variables, viscosity solutions \cite{oleinik1957discontinuous, kruvzkov1970first}  of popular fluid systems (e.g., shallow water, compressible Euler and Navier-Stokes \cite{hughes1986new, chen2017entropy}) can be shown to satisfy 
\begin{equation}
\int_{\Omega}\pd{S(\bm{u})}{t}\diff{\bm{x}} + \int_{\partial \Omega} \LRp{\bm{v}^T\bm{f}(\bm{u}) - \psi_i(\bm{u})}n_i \leq 0\label{eq:entropyineq},
\end{equation}
where $n_i$ denotes the $i$th component of the outward normal vector.  
The entropy inequality (\ref{eq:entropyineq}) is the analogue of energy stability for nonlinear conservation laws \cite{mock1980systems, harten1983symmetric}.  
However, due to the use of inexact treatment of nonlinear terms (e.g., from collocation approximations, aliasing errors, or inexact quadrature rules), most numerical methods for nonlinear conservation laws do not satisfy a discrete analogue of this stability condition.

\section{The full order model: entropy stable finite volume schemes}
\label{sec:2}

The reduced order models in this work are constructed using full order models constructed from discretely entropy stable finite volume methods \cite{tadmor1987numerical}.  For simplicity, we illustrate the construction of full and reduced order models on a periodic 1D domain.  We note that, while the reduced order model will be constructed from an entropy stable scheme, the solution snapshots do not need to be generated by entropy stable full order model.  Throughout this work, we will assume that the both full and reduced models yield ``physically relevant'' solutions, such that the entropy is a convex function.  For example, for the compressible Euler and Navier-Stokes equations, we will assume that the thermodynamic variables (density and energy/temperature/pressure) are positive.  Guaranteeing both positivity and accuracy of numerical solutions for general discretizations remains challenging \cite{guermond2016invariant, guermond2019invariant}, and will be explored in future work.

A key ingredient of the full and reduced order models considered in this work are \textit{entropy conservative} numerical fluxes.  Let $\bm{u}_L, \bm{u}_R$ denote left and right states.  Then, a two-point numerical flux $\bm{f}_S(\bm{u}_L, \bm{u}_R)$ is entropy conservative if it satisfies the following three conditions
\begin{gather}
\bm{f}_S(\bm{u},\bm{u}) = \bm{f}(\bm{u}), \qquad \text{(consistency)} \label{eq:esflux}\\
\bm{f}_S(\bm{u}_L,\bm{u}_R) = \bm{f}_S(\bm{u}_R,\bm{u}_R), \qquad \text{(symmetry)} \nonumber\\
\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}_S(\bm{u}_L,\bm{u}_R) = \psi(\bm{u}_L) - \psi(\bm{u}_R), \qquad \text{(entropy conservation)}\nonumber,
\end{gather}
Let the domain be decomposed into $K$ cells of size $\Delta x$, and let ${\LRp{\bm{u}_h}_i}$ denote the vector containing mean values of the conservation variables over the $i$th cell.  An entropy conservative finite volume method results from discretizing the integral form of the conservation law as follows
\begin{align}
&\td{{\LRp{\bm{u}_h}_1}}{t} + \frac{\bm{f}_S({\LRp{\bm{u}_h}_2},\LRp{\bm{u}_h}_1)-\bm{f}_S({\LRp{\bm{u}_h}_1},{\LRp{\bm{u}_h}_K})}{\Delta x} = \bm{0}, \label{eq:nonlineqs1d}\\
&\td{{\LRp{\bm{u}_h}_i}}{t} + \frac{\bm{f}_S({\LRp{\bm{u}_h}_{i+1}},\LRp{\bm{u}_h}_i)-\bm{f}_S({\LRp{\bm{u}_h}_i},{\LRp{\bm{u}_h}_i-1})}{\Delta x} = \bm{0}, \qquad 1 < i < K \nonumber\\
&\td{{\LRp{\bm{u}_h}_K}}{t} + \frac{\bm{f}_S({\LRp{\bm{u}_h}_1},\LRp{\bm{u}_h}_K)-\bm{f}_S({\LRp{\bm{u}_h}_K},{\LRp{\bm{u}_h}_K-1})}{\Delta x} = \bm{0},\nonumber
\end{align}
where periodicity is imposed through the equations for ${\LRp{\bm{u}_h}_1}, {\LRp{\bm{u}_h}_K}$.

We first rewrite the system (\ref{eq:nonlineqs1d}) in a matrix form which will be more amenable to model reduction.  We define the skew-symmetric matrix $\bm{Q}$ and flux matrix $\bm{F}$ such that
\begin{equation}
\bm{Q} = \frac{1}{2}\begin{bmatrix}
0 & 1 & &\ldots & -1\\
-1 & 0 & 1 &&  \\
& -1 & 0 & 1 &  \\
 & & & \ddots &  \\
1 & &\ldots  & -1 & 0
\end{bmatrix}, \qquad \bm{F}_{ij} = \bm{f}_{S}\LRp{\LRp{\bm{u}_h}_i, \LRp{\bm{u}_h}_j}.
\label{eq:Qmat}
\end{equation}
Note that $\bm{Q}\bm{1} = \bm{0}$, the matrix $\bm{F}$ is symmetric due to flux symmetry, and the diagonal of $\bm{F}$ is equal to $\bm{f}(\bm{u})$ due to flux consistency.  For a scalar nonlinear conservation law, the matrix-based formulation of (\ref{eq:nonlineqs1d}) is then equivalent to
\begin{align}
\Delta x \td{\bm{u}_h}{t} + 2\LRp{{\bm{Q}}\circ \bm{F}}\bm{1} = \bm{0}.
\label{eq:es}
\end{align}
where $\circ$ denotes the Hadamard product, $\bm{1}$ denotes the vector of all ones, and each entry of $\bm{u}_h$ corresponds to a point $\bm{x}_i$ in the domain.  In this setting, $\bm{Q}\circ\bm{F}$ extracts and takes linear combinations of nonlinear flux interactions between different nodal values of $(\bm{u}_h)_i$ and $(\bm{u}_h)_j$.  The discretization of the nonlinear flux term using $\LRp{\bm{Q}\circ \bm{F}}\bm{1}$ is commonly referred to as \textit{flux differencing} \cite{carpenter2014entropy, gassner2016split, chen2017entropy, crean2018entropy, chan2017discretely}.\footnote{The factor of 1/2 present in the definition of $\bm{Q}$ is to make (\ref{eq:es}) consistent with the entropy stable SBP literature.  The factor of 2 in (\ref{eq:es}) can also be derived from the chain rule \cite{chen2017entropy, crean2018entropy}.  }

For a system of $n$ nonlinear conservation laws, all matrices are treated in a Kronecker product fashion.  Let the vector $\bm{u}_h$ now correspond to the concatenated vector of components of the solution $\bm{u}_h = [\bm{u}_h^1, \ldots, \bm{u}_h^n]$.  Then, the matrix formulation of (\ref{eq:nonlineqs1d}) is given by
\begin{gather*}
\Delta x\td{\bm{u}_h}{t} + \LRp{\LRp{\bm{I}_{n \times n} \otimes \bm{Q} }\circ \bm{F}}\bm{1} = \bm{0},\\
\bm{F} = \begin{bmatrix}
\bm{F}_{1} & &\\
& \ddots &\\
&& \bm{F}_{n} \\
\end{bmatrix}, \qquad 
\LRp{\bm{F}_{i}}_{jk} = \bm{f}_{i,S}\LRp{\LRp{\bm{u}_h}_j, \LRp{\bm{u}_h}_k}.
\label{eq:essys}
\end{gather*}
Here, $\bm{F}$ is a block diagonal matrix where the $i$th block corresponds to the evaluation of the $i$th component of the numerical flux $\bm{f}_{i,S}(\bm{u}_L,\bm{u}_R)$.  From this point on, we drop explicit references to the Kronecker product to simplify notation, so that (\ref{eq:es}) applies to both scalar nonlinear conservation laws and systems of nonlinear conservation laws.

We next show that the full order model satisfies a discrete conservation of entropy.  
\begin{theorem}
\label{thm:ecfom}
Let $\bm{u}_h(t)$ be a solution of (\ref{eq:essys}) for which the entropy $S(\bm{u}_h)$ is convex at each time $t$.  Then, $\bm{u}_h$ satisfies the semi-discrete conservation of entropy 
\[
\Delta x \bm{1}^T\td{S(\bm{u}_h)}{t} = 0.
\]
\end{theorem}
\begin{proof}
The proof can be found in the literature \cite{tadmor1987numerical, tadmor2003entropy, tadmor2016entropy}.  We present an alternative proof which utilizes matrix properties of entropy stable summation by parts (SBP) schemes \cite{carpenter2014entropy, gassner2016split, chen2017entropy, crean2018entropy, chan2017discretely}.  We test (\ref{eq:es}) with the vector of entropy variables $\bm{v}_h = \bm{v}(\bm{u}_h)$
\[
\Delta x \bm{v}_h^T\td{\bm{u}_h}{t} + \bm{v}_h^T2\LRp{{\bm{Q}}\circ \bm{F}}\bm{1} = 0.
\]
Assuming the chain rule in time, 
\begin{align*}
\bm{v}_h^T\td{\bm{u}_h}{t} &= {\td{S(\bm{u}_h)}{\bm{u}}}^T\td{\bm{u}_h}{t} = \sum_j \LRp{\td{S(\bm{u}_h)}{\bm{u}}}_j^T\td{\LRp{\bm{u}_h}_j}{t} \\
&=  \sum_j \td{S(\LRp{\bm{u}_h}_j)}{t}  =  \bm{1}^T\td{S(\bm{u}_h)}{t}.
\end{align*}
Using skew-symmetry of $\bm{Q}$, the flux term yields
\begin{align*}
\bm{v}_h^T2\LRp{{\bm{Q}}\circ \bm{F}}\bm{1} &= \sum_{ij} \LRp{\bm{v}_h}_i^T \bm{Q}_{ij} \bm{f}_{S}\LRp{\LRp{\bm{u}_h}_i, \LRp{\bm{u}_h}_j} \\
&= \sum_{ij} \LRp{\bm{Q}_{ij}-\bm{Q}_{ji}} \LRp{\bm{v}_h}_i^T  \bm{f}_{S}\LRp{\LRp{\bm{u}_h}_i, \LRp{\bm{u}_h}_j}.
\end{align*}
Rearranging indices and using the symmetry of $\bm{f}_S(\bm{u}_L,\bm{u}_R) = \bm{f}_S(\bm{u}_R,\bm{u}_L)$ exposes the entropy conservation condition in the sum
\begin{align*}
&\sum_{ij} \LRp{\bm{Q}_{ij}-\bm{Q}_{ji}} \LRp{\bm{v}_h}_i^T \bm{f}_{S}\LRp{\LRp{\bm{u}_h}_i, \LRp{\bm{u}_h}_j} \\
=&\sum_{ij} \bm{Q}_{ij} \LRp{\LRp{\bm{v}_h}_i-\LRp{\bm{v}_h}_j}^T  \bm{f}_{S}\LRp{\LRp{\bm{u}_h}_i, \LRp{\bm{u}_h}_j}\\
=&\sum_{ij} \bm{Q}_{ij} \LRp{\psi\LRp{\LRp{\bm{u}_h}_i} - \psi(\LRp{\bm{u}_h}_j)} = \bm{\psi}^T\bm{Q}\bm{1}-\bm{1}^T\bm{Q}\bm{\psi} = 0,
\end{align*}
where we have used that $\bm{Q}=-\bm{Q}^T$ and $\bm{Q}\bm{1} = \bm{0}$ in the final step.

\end{proof}

\subsection{Viscosity and entropy dissipation}

Theorem~\ref{thm:ecfom} shows that the formulation (\ref{eq:essys}) preserves a semi-discrete conservation of entropy.  However, in the presence of shock discontinuities, entropy should be dissipated instead of conserved.  To mimic this at the semi-discrete level, we add entropy dissipation to (\ref{eq:es}) through appropriate viscosity terms.  For example, it was shown in \cite{tadmor2006entropy} that a centered approximation of the Navier-Stokes viscosity is entropy dissipative.  In this work, we apply a simple Laplacian artificial viscosity to each component of the solution \cite{upperman2019entropy}
\begin{equation}
\Delta x \td{\bm{u}_h}{t} + 2\LRp{{\bm{Q}}\circ \bm{F}}\bm{1} + \epsilon \bm{K}\bm{u}_h= 0, \qquad \bm{K} = \frac{1}{\Delta x^2} \begin{bmatrix}
1 & -1 & & \\
-1 & 2 & \ddots &\\
 & \ddots & \ddots & -1\\
 &  & -1& 1\\
\end{bmatrix}
\label{eq:esvisc}
\end{equation}
where $\epsilon$ is the visosity coefficient.  This choice of dissipation simplifies descriptions of how to treat viscous regularizations of the solution in an entropy stable fashion.  Future work will explore both physically relevant viscosities and more nuanced artificial dissipation mechanisms.

If $\bm{v}_h^T\bm{K}\bm{u}_h \geq 0$, then the solution satisfies a discrete dissipation of entropy
\begin{equation}
\Delta x \bm{1}^T\td{S(\bm{u}_h)}{t} = -\epsilon\bm{v}_h^T\bm{K}\bm{u}_h \leq 0.
\label{eq:entropybalancefom}
\end{equation}
The proof is given in \cite{tadmor2006entropy}, which we reproduce here.  We rewrite the viscous term as
\[
\bm{v}_h^T\bm{K}\bm{u}_h = \bm{v}_h^T \frac{1}{\Delta x^2}\begin{bmatrix}
\LRp{\bm{u}_h}_1 - \LRp{\bm{u}_h}_2\\ 
%\LRp{\LRp{\bm{u}_h}_2 - \LRp{\bm{u}_h}_1} - \LRp{\LRp{\bm{u}_h}_3 - \LRp{\bm{u}_h}_2}\\
\vdots\\
\LRp{\LRp{\bm{u}_h}_i - \LRp{\bm{u}_h}_{i-1}} - \LRp{\LRp{\bm{u}_h}_{i+1} - \LRp{\bm{u}_h}_{i}}\\
\vdots\\
\LRp{\bm{u}_h}_K- \LRp{\bm{u}_h}_{K-1} 
\end{bmatrix}.
\]
Since $\bm{v}_h$ and $\bm{u}_h$ are related through a differentiable invertible mapping, the mean value theorem implies that
\[
\LRp{\bm{u}_h}_i - \LRp{\bm{u}_h}_{i-1} = \LRp{\pd{\bm{u}}{\bm{v}}}_{i,i-1} \LRp{\LRp{\bm{v}_h}_i - \LRp{\bm{v}_h}_{i-1}}
\]
where $\LRp{\pd{\bm{u}}{\bm{v}}}_{i,i-1}$ is the positive-definite Jacobian matrix evaluated at some state between $(\bm{u}_h)_i$ and $(\bm{u}_h)_j$.  Substituting these expressions into the viscous term yields 
\begin{gather}
\bm{v}_h^T\bm{K}\bm{u}_h = 
%\bm{v}_h^T \frac{1}{\Delta x^2}\begin{bmatrix}
%\LRp{\pd{\bm{u}}{\bm{v}}}_{2,1} \LRp{\LRp{\bm{v}_h}_1 - \LRp{\bm{v}_h}_{2}}\\ 
%\vdots\\
%\LRp{\pd{\bm{u}}{\bm{v}}}_{i,i-1} \LRp{\LRp{\bm{v}_h}_i - \LRp{\bm{v}_h}_{i-1}} - \LRp{\pd{\bm{u}}{\bm{v}}}_{i+1,i} \LRp{\LRp{\bm{v}_h}_{i+1} - \LRp{\bm{v}_h}_{i}}\\
%\vdots\\
%\LRp{\pd{\bm{u}}{\bm{v}}}_{K,K-1} \LRp{\LRp{\bm{v}_h}_K - \LRp{\bm{v}_h}_{K-1}}
%\end{bmatrix}\\
 \frac{1}{\Delta x^2}\sum_{i=1}^{K-1} \LRp{\pd{\bm{u}}{\bm{v}}}_{i+1,i} \LRb{\LRp{\bm{v}_h}_{i} - \LRp{\bm{v}_h}_{i+1}}^2 \geq 0.
 \label{eq:entropydissfom}
\end{gather}

\section{An entropy conservative reduced order model}
\label{sec:3}
We can now formulate a reduced basis approximation on periodic domains using the entropy stable full order models described in the previous section.  Boundary conditions will be addressed later in this work.  We also note that this reduced order model is not practical, since the cost of evaluating nonlinear terms scales with the size of the full order model.  Later sections will discuss how to reduce the cost of nonlinear evaluations using appropriate hyper-reduction techniques.

Let $\LRc{{\phi}_j(x)}_{j=1}^{N}$ denote a reduced basis for each component of the solution, which may be generated using principal orthogonal decomposition (POD) or the reduced basis procedure.  Let $\bm{V}$ denote the generalized Vandermonde matrix whose columns contain evaluations of $\phi_j$ at grid points $x_i$
\[
\bm{V}_{ij} = {\phi}_j(x_i).
\]
We approximate values of the solution by $\bm{u}_h = \bm{V}\bm{u}_N$, where $\bm{u}_N$ denote ``modal'' coefficients  of the solution in the reduced basis.  Ignoring viscosity terms for now, plugging this expression into (\ref{eq:es}) and enforcing that the residual is orthogonal to all columns of $\bm{V}$ (Galerkin projection) yields a reduced order system
\begin{align*}
\bm{V}^T\bm{V}\td{\bm{u}_N}{t} + 2\bm{V}^T\LRp{\bm{Q}\circ \bm{F}}\bm{1} = 0.
\end{align*}
We will show that, while this formulation is not entropy conservative, a slight modification recovers semi-discrete entropy conservation.  We motivate this modification by first considering the time derivative term.  Because of the Galerkin projection step, we can no longer test with the vector of entropy variables.  However, we can test with an appropriate \textit{projection} of the entropy variables.  Let $\bm{v}_N$ denote coefficients of the projection of the entropy variables
\[
\bm{v}_N = \LRp{\bm{V}^T\bm{V}}^{-1}\bm{V}^T \bm{v}\LRp{\bm{V}\bm{u}_N} = \bm{V}^{\dagger} \bm{v}\LRp{\bm{V}\bm{u}_N},
\]
where $\bm{V}^{\dagger}$ is the pseudoinverse of $\bm{V}$.  Then, we can recover the time derivative of the discrete entropy by testing with the coefficients $\bm{v}_N$
\begin{align}
\bm{v}_N^T\bm{V}^T\bm{V}\td{\bm{u}_N}{t} &= \bm{v}\LRp{\bm{V}\bm{u}_N}^T \bm{V}\LRp{\bm{V}^T\bm{V}}^{-1}\bm{V}^T\bm{V} \td{\bm{u}_N}{t} \label{eq:esromtime}\\
&= 
 \bm{v}\LRp{\bm{V}\bm{u}_N}^T \td{\bm{V}\bm{u}_N}{t} = \bm{1}^T\td{S(\bm{V}\bm{u}_N)}{t}.\nonumber
\end{align}
Define the grid values of the projected entropy variables as $\tilde{\bm{v}}$ 
\[
\tilde{\bm{v}} = \bm{V}\bm{v}_N = \bm{V}\bm{V}^{\dagger} \bm{v}\LRp{\bm{V}\bm{u}_N}.
\]
It is possible to repeat the steps of the proof of Theorem~\ref{thm:ecfom} up to the point when we invoke the entropy conservation condition of (\ref{eq:esflux})
\begin{align*}
\tilde{\bm{v}}^T\LRp{{\bm{Q}}\circ \bm{F}}\bm{1} &= \frac{1}{2}\sum_{ij} \bm{Q}_{ij} \LRp{\tilde{\bm{v}}_i-\tilde{\bm{v}}_j}^T \bm{f}_{S}\LRp{\LRp{\bm{u}_h}_i, \LRp{\bm{u}_h}_j}\\
&\neq \frac{1}{2}\sum_{ij} \bm{Q}_{ij} \LRp{ \psi((\bm{u}_h)_i)- \psi((\bm{u}_h)_j)}.
\end{align*}
This is not possible due to the fact that the projected entropy variables $\tilde{\bm{v}}$ are no longer mappings of the grid values of the conservative variables $\bm{u}_h = \bm{V}\bm{u}_N$.  To remedy this, we follow approaches taken in \cite{parsani2016entropy, chan2017discretely} and replace the values of ${\bm{u}_h}$ used to evaluate the flux matrix $\bm{F}$ with the \textit{entropy-projected conservative variables} $\tilde{\bm{u}}$
\[
\tilde{\bm{u}} = \bm{u}\LRp{\bm{V}\bm{V}^{\dagger}\bm{v}\LRp{\bm{V}\bm{u}_N}} = \bm{u}\LRp{\tilde{\bm{v}}}.
\]
The entropy projected conservative variables are thus mappings of the projected entropy variables.  We note that $\bm{V}\bm{V}^{\dagger}$ can also be interpreted as a discrete approximation of the $L^2$ projection operator onto the span of the reduced basis $\phi_j$. One can then construct a semi-discretely entropy conservative reduced model.
\begin{theorem} 
Let the coefficients $\bm{u}_N$ solve
\begin{gather}
\bm{V}^T\bm{V}\td{\bm{u}_N}{t} + 2\bm{V}^T\LRp{\bm{Q}\circ \bm{F}}\bm{1} = 0, \label{eq:esrom}\\
\LRp{\bm{F}}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad \tilde{\bm{u}} = \bm{u}\LRp{\bm{V}\bm{V}^{\dagger}\bm{v}\LRp{\bm{V}\bm{u}_N}}.\nonumber
\end{gather}
Then, the solution satisfies the following semi-discrete conservation of entropy
\[
\Delta x\bm{1}^T\td{S(\bm{V}\bm{u}_N)}{t} = 0.
\]
\label{thm:esrom}
\end{theorem}
\begin{proof}
The result follows from testing with $\bm{v}_N$ and applying (\ref{eq:esromtime}).  The remaining steps are identical to those of Theorem~\ref{thm:ecfom}.  
\end{proof}

\begin{remark}
In order for the formulation (\ref{eq:esrom}) to remain accurate, the entropy projected conservative variables must accurately approximate the conservative variables.  This requires that the reduced basis accurately approximates the entropy variables.  This can be taken into account, for example, by computing the POD basis vectors from snapshots of both the conservative and entropy variables.
\end{remark}
%Here, $\bm{u}_N$ are the coefficients of $u_N$ and $\bm{P} = \LRp{\bm{V}^T\bm{V}}^{-1}\bm{V}^T = \bm{V}^{\dagger}$ is the pseudo-inverse of $\bm{V}$.

\section{Entropy conservative hyper-reduction}
\label{sec:4}
While the Galerkin projected formulation (\ref{eq:esrom}) is entropy stable, the computational cost involved in solving the semi-discrete system scales with the size of the full order model rather than the dimension of the reduced basis.  For example, explicit time-stepping methods require the evaluation of the nonlinear term $\bm{V}^T\LRp{\bm{Q} \circ \bm{F}}\bm{1}$.  Since the matrices $\bm{Q}$ and $\bm{F}$ are the original full order model matrices, the cost of the reduced system is not lower than the cost of the full order model.

To reduce the cost of evaluating nonlinear terms, we introduce a second \textit{hyper-reduction} step \cite{ryckelynck2009hyper}.  Here, a vector of nonlinear function evaluations is replaced by an approximation constructed from nonlinear function evaluations at a smaller set of reduced points \cite{barrault2004empirical, bui2004aerodynamic, chaturantabut2010nonlinear, farhat2015structure, drmac2016new, hernandez2017dimensional, yano2019lp}.  In this work, we utilize a sampling and weighting hyper-reduction strategy \cite{farhat2015structure, hernandez2017dimensional, yano2019lp} which enables proofs of discrete entropy stability.  These hyper-reduction techniques can be interpreted as reduced quadratures, and produce approximations of the form
\[
\bm{V}^Tg(\bm{V}\bm{u}_N) \approx \bm{V}\LRp{\mathcal{I},:}^T \bm{W} g\LRp{\bm{V}\LRp{\mathcal{I},:}}.
\]
Here, $g(\bm{u})$ denotes a nonlinear function, $\mathcal{I}$ denotes a subset of $N_s$ row indices, $\bm{V}\LRp{\mathcal{I},:}$ denotes the matrix consisting of the $N_s$ sampled rows of $\bm{V}$, and $\bm{W} = {\rm diag}(\bm{w})$ is a $N_s\times N_s$ diagonal matrix of positive weights.  We will describe algorithms for computing hyper-reduced points and weights in Section~\ref{sec:hyperreducalgo}.

We will hyper-reduce the nonlinear term $\LRp{\bm{Q}\circ \bm{F}}$ by constructing a hyper-reduced sampled matrix $\bm{Q}_s$.  Recall that the proof of entropy conservation in Theorem~\ref{thm:ecfom} required that matrix $\bm{Q}$ is skew symmetric and has zero row sums.  If the hyper-reduced matrix $\bm{Q}_s$ also satisfies the properties
\[
\bm{Q}_s = -\bm{Q}_s^T, \qquad \bm{Q}_s\bm{1} = 0
\]
then the proof of entropy conservation will carry over to the hyper-reduced model.  Unfortunately, common hyper-reduction techniques preserve either skew-symmetry or zero row sums, but not both.  For example, $\bm{Q}$ can be decomposed into the sum of local skew-symmetric matrices.  The hyper-reduction techniques of \cite{farhat2015structure, yano2019discontinuous} approximate $\bm{Q}$ using a sparse linear combination of those local matrices.  This hyper-reduction technique preserves skew-symmetry, but does not necessarily preserve the zero row sum property.  Similarly, techniques such as gappy POD or empirical interpolation \cite{everson1995karhunen, barrault2004empirical, chaturantabut2010nonlinear} may preserve the zero row sum but not skew-symmetry.

\subsection{Two-step hyper-reduction: compress and project}

We apply a two-step hyper-reduction procedure which preserves both the zero row sums and skew-symmetry of the hyper-reduced matrix $\bm{Q}_s$.  Rather than directly hyper-reducing the full order matrix $\bm{Q}$, we first construct a compressed ``modal'' matrix, then combine this with a projection operator based on the hyper-reduced points.

In the first step, we construct a compressed intermediate operator by combining  Galerkin projection with the ``expanded basis'' approach of \cite{hernandez2017dimensional}.  Let $\bm{V}_t$ denote a \textit{test} basis through which we will extract the action of $\bm{Q}$.  We define the intermediate reduced operator 
\begin{equation}
\hat{\bm{Q}}_{t} = \bm{V}_{t}^T\bm{Q}\bm{V}_{t}.
\label{eq:compressQ}
\end{equation}
Let $\bm{c}$ be some vector in the span of the test basis such that $\bm{c}=\bm{V}_t \hat{\bm{c}}$. Solving the system $\bm{V}_t^T\bm{V}_t \hat{\bm{d}} = \hat{\bm{Q}}_t\hat{\bm{c}}$ yields that
\[
\bm{V}_t^T\bm{V}_t \hat{\bm{d}} = \bm{V}_t^T\bm{Q}\bm{V}_t\hat{\bm{c}} = \bm{V}_t^T\bm{Q}\bm{c} \qquad \Longrightarrow\qquad \bm{V}_t^T \LRp{\bm{V}_t\hat{\bm{d}}-\bm{Q}\bm{c}} = \bm{0}.
\]
Here, $\hat{\bm{d}}$ are the coefficients of the orthogonal projection of $\bm{Q}\bm{c}$ onto the test basis.  Note that $\hat{\bm{Q}}_t$ is skew-symmetric by construction.  Moreover, if the vector of all ones lies in the span of the test basis (e.g., $\bm{1} = \bm{V}_t\bm{e}$ for some coefficient vector $\bm{e}$), then 
\[
\hat{\bm{Q}}_t\bm{e} = \bm{0},
\]
since $\hat{\bm{Q}}_t$ exactly recovers the action of $\bm{Q}\bm{1} = \bm{0}$ and $\bm{0}$ lies within the span of any basis.

%\subsubsection{A ``nodal'' reduced operator}

Since $\hat{\bm{Q}}_{t}$ acts on coefficients of the test basis (a ``modal'' operator), it cannot be directly applied to evaluate nonlinear functions of the solution.  However, we can construct ``nodal'' operator by combining $\hat{\bm{Q}}_t$ with appropriate mappings from hyper-reduced points to coefficients of the test basis.  We make the following assumptions on the hyper-reduced points to enable the construction of such mappings: 
\begin{assumption}
Let $\mathcal{I}$ denote the index set of hyper-reduced points.  We assume that the row-sampled test matrix $\bm{V}_t\LRp{\mathcal{I},:}$ is non-singular.  
 \label{ass:quad}
\end{assumption}
Note that if Assumption~\ref{ass:quad} holds, then the matrix $\bm{V}\LRp{\mathcal{I},:}$ is also non-singular, since the column space of $\bm{V}$ is contained in the column space of $\bm{V}_t$.  Then, we can define the projection matrix $\bm{P}_t$ onto the test basis as follows:
\[
\bm{P}_t = \LRp{\bm{V}_t\LRp{\mathcal{I},:}^T\bm{W}\bm{V}_t\LRp{\mathcal{I},:}}^{-1}\bm{V}_t\LRp{\mathcal{I},:}^T\bm{W}.
\]
The projection matrix $\bm{P}_t$ maps values at hyper-reduced points to coefficients in the test basis, and is well-defined under Assumption~\ref{ass:quad}.  The key property of $\bm{P}_t$ which we will utilize is as follows: suppose that $\bm{f} = \bm{V}_t\bm{c}$ for some coefficients $\bm{c}$.  Then,
\begin{equation}
\bm{P}_t \bm{f} = \LRp{\bm{V}_t\LRp{\mathcal{I},:}^T\bm{W}\bm{V}_t\LRp{\mathcal{I},:}}^{-1}\bm{V}_t\LRp{\mathcal{I},:}^T\bm{W} \bm{V}_t\LRp{\mathcal{I},:}\bm{c} = \bm{c}.
\label{eq:preproduce}
\end{equation}
In other words, if a vector lies in the range of $\bm{V}_t$, the projection matrix $\bm{P}_t$ recovers the coefficients exactly.  We also note that if the hyper-reduced set $\mathcal{I}$ is taken to include all the points and $\bm{W} = \Delta x\bm{I}$, then $\bm{P}_t$ reduces to the pseudo-inverse $\bm{V}_t^{\dagger} = \LRp{\bm{V}_t\bm{V}_t}^{-1}\bm{V}_t^T$.

We can now construct a ``nodal'' differentiation operator which satisfies both skew-symmetry and zero row sum properties: 
\begin{lemma}
Assume that $\bm{1}$ is exactly representable in the test basis and that the columns of the submatrix $\bm{V}_t\LRp{\mathcal{I},:}$ are linearly independent.  Define the matrix $\bm{Q}_t$
\begin{equation}
\bm{Q}_t = \bm{P}_t^T\hat{\bm{Q}}_t\bm{P}_t = \bm{P}_t^T\LRp{\bm{V}_t^T\bm{Q}\bm{V}_t}\bm{P}_t.
\label{eq:nodalQ}
\end{equation}
Then, $\bm{Q}_t$ is skew-symmetric and satisfies $\bm{Q}_t\bm{1} = \bm{0}$.  
\label{thm:Qt}
\end{lemma}
\begin{proof}
Since $\bm{Q}$ is skew-symmetric, the matrix $\bm{Q}_t$ is skew-symmetric by construction as well.  By (\ref{eq:preproduce}), if $\bm{1}$ is exactly representable in the test basis (such that for some coefficient vector $\bm{e}$ $\bm{1} = \bm{V}_t\LRp{\mathcal{I},:}\bm{e}$ ), then $\bm{P}_t\bm{1} = \bm{e}$.  Then, $\bm{Q}_t\bm{1} = \bm{P}_t^T\hat{\bm{Q}}_t\bm{e} = \bm{0}$, since $\hat{\bm{Q}}_t$ exactly differentiates elements in the test basis.
\end{proof}

\subsubsection{Choices of test basis}

We have not yet specified the choice of test basis $\bm{V}_t$.  Common sense suggests that the span of the test basis should include the reduced basis $\bm{V}$.  Theorem~\ref{thm:Qt} implies that the span of the test basis should also contain $\bm{1}$.  However, these two conditions are insufficient to ensure accuracy for solutions of nonlinear conservation laws.  Consider, for example, the 1D periodic Burgers' equation with initial condition 
\[
u(x,0) = -\sin(\pi x).
\]
The solution $u(x,t)$ is a decaying stationary shock which is anti-symmetric across the origin at all times $t$.  The POD modes are thus also anti-symmetric across the origin; however, their derivatives $\bm{Q}\bm{V}$ are symmetric across the origin, as shown in Figure~\ref{fig:modeQ}.  As a result, since the entries of $\bm{V}^T\bm{Q}\bm{V}$ are inner products of symmetric and anti-symmetric functions, $\bm{V}^T\bm{Q}\bm{V} \approx \bm{0}$.
\begin{figure}[!h]
\centering
\subfloat[Solution snapshots]{\includegraphics[width=.3\textwidth]{figs/burg_snaps.png}}
\hspace{.5em}
\subfloat[Modes (columns of $\bm{V}$)]{\includegraphics[width=.3\textwidth]{figs/burg3modes.png}}
\hspace{.5em}
\subfloat[Mode derivatives $\bm{Q}\bm{V}$]{\includegraphics[width=.3\textwidth]{figs/burg3modes_diff.png}}
\caption{Solution snapshots, three POD modes, and mode derivatives for a shock solution of the viscous periodic Burgers' equation. }
\label{fig:modeQ}
\end{figure}

This example implies that the reduced basis matrix $\bm{V}$ can do a poor job of sampling the range of $\bm{Q}\bm{V}$, and that the test basis should contain additional vectors.  To remedy this issue, we borrow techniques from least squares Petrov-Galerkin ROMs \cite{carlberg2011efficient, carlberg2017galerkin} and enrich the test basis $\bm{V}_t$ with vectors spanning the range of $\bm{Q}\bm{V}$.  The test basis now spans a space which contains $\bm{1}, \bm{V}$, and $\bm{Q}\bm{V}$, and thus has a dimension of at most $2N+1$.  The dimension of $\bm{V}_t$ may also be smaller if the intersection of the range of the reduced basis matrix $\bm{V}$ and the range of $\bm{Q}\bm{V}$ and $\bm{1}$ is non-empty.

We illustrate why this choice of test basis restores accuracy.  The entropy conservative flux for Burgers' equation corresponds to the discretization of the split form \cite{canuto2006spectral, chan2017discretely, maboudi2018conservative}
\[
\pd{u}{t} + \frac{1}{3}\LRp{\pd{u^2}{x}  + u\pd{u}{x}} = 0.
\]
An entropy conservative ROM without hyper-reduction is then given by 
\[
\bm{V}^T\bm{V}\td{u}{t} + \frac{1}{3}\bm{V}^T\LRp{\bm{Q}_t\bm{u}^2  + {\rm diag}\LRp{\bm{u}}\bm{Q}_t\bm{u}} = \bm{0}, \qquad \bm{u} = \bm{V}\bm{u}_N.
\]
It was shown in \cite{chan2017discretely, chenreview} that, for certain entropy conservative fluxes, the flux differencing procedure is equivalent in discretization of the split form.  Thus, to ensure accurate evalution of the spatial formulation, the reduced matrix $\bm{Q}_t$ should accurately approximate mixed derivatives of the form ${\rm diag}(\bm{f})\bm{Q}\bm{g}$.  We quantify the accuracy of $\bm{Q}_t$ in approximating such terms with the following theorem:
\begin{theorem}
Let $\bm{f},\bm{g}$ denote vectors, and let $\bm{u}_1, \bm{u}_2$ denote solutions of the following problems
\begin{align*}
\bm{V}^T\bm{V}\bm{u}_1 &= \bm{V}^T \diag{\bm{f}}\bm{Q}\bm{g}\\
\bm{V}^T\bm{V}\bm{u}_2 &= \bm{V}^T \diag{\bm{f}}\bm{Q}_t\bm{g}.
\end{align*}
Then, the error $\bm{e} = \bm{u}_1-\bm{u}_2$ satisfies
\[
\nor{\bm{V}\bm{e}} \leq \nor{\bm{Q}} \LRp{\nor{\bm{f}-\bm{V}\bm{P}\bm{f}} \nor{\bm{g}} + \nor{\bm{f}}\nor{\bm{g}-\bm{V}\bm{P}\bm{g}}}.
\]
\label{thm:acc}
\end{theorem}
\begin{proof}
Since $\bm{Q}_t = \bm{P}_t^T\LRp{\bm{V}_t^T\bm{Q}\bm{V}_t}\bm{P}_t$, we can derive the estimate by adding and subtracting $\bm{P}_t^T\bm{V}_t^T \diag{\bm{f}}\bm{Q}\bm{g}$.  Straightforward manipulations give the following
\begin{align*}
\bm{V}^T\diag{\bm{f}}\LRp{\bm{Q}\bm{g} - \bm{Q}_t\bm{g}} =  
\bm{V}^T\diag{\bm{f}} \LRp{\bm{Q}\LRp{\bm{I}-\bm{V}_t\bm{P}_t}\bm{g} + \LRp{\bm{I}-\bm{V}_t\bm{P}_t}^T\bm{Q}\bm{g}}
\end{align*}
 \note{Add proof}
 \end{proof}

Theorem~\ref{thm:acc} does not account for hyper-reduction.

\subsection{Hyper-reduction algorithm and target space}
\label{sec:hyperreducalgo}

In this work, we utilize the greedy algorithm for computing empirical cubature points and weights from \cite{an2008optimizing, hernandez2017dimensional}, which is described in Algorithm~\ref{alg:hr}.  Let $\bm{V}_{\rm target}$ be some matrix whose columns span some target space of functions to be integrated.  The algorithm approximates $\bm{V}_{\rm target}^T\bm{1}$ by $\bm{V}_{\rm target}\LRp{\mathcal{I},:}^T\bm{w}$.  The index set is constructed in a greedy fashion by selecting the row index of $\bm{V}_{\rm target}$ which is most positively parallel to the residual, then computing the weight vector which minimizes the residual through a (non-negative) least squares solve.

\begin{algorithm}
\caption{Compute hyper-reduction s.t.\ $\bm{V}_{\rm target}^T\bm{w}_{\rm target} \approx \bm{V}_{\rm target}\LRp{\mathcal{I},:}^T\bm{w}$.}
\begin{algorithmic}[1]
\STATE \textbf{Input}: $\bm{V}_{\rm target}, \bm{w}_{\rm target}, tol$.
\STATE \textbf{Output}: $\mathcal{I}$, $\bm{w}$.
\STATE Set $\bm{b} = \bm{V}_{\rm target}^T\bm{w}_{\rm target}$.
\STATE Initialize residual $\bm{r} = \bm{b}$, $\mathcal{I} = \emptyset$.
\WHILE {$\nor{\bm{r}}/\nor{\bm{b}} > tol$}
        \STATE Select new index $i$ such that
        \[
        i = \argmin_i \tilde{\bm{V}}_{i}\bm{r}/\nor{\bm{r}}, \qquad
        \tilde{\bm{V}}_i = \bm{V}_{\rm target}(i,:) / \nor{\bm{V}_{\rm target}(i,:)}, \quad i \not\in \mathcal{I}.
        \]
        \STATE Set $\mathcal{I} = \mathcal{I} \cup \LRc{i}$.  
        \STATE Compute $\bm{w}$ using linear least squares
        \[
        \bm{w} = \argmin_{\bm{c}} \nor{\bm{V}_{\rm target}\LRp{\mathcal{I},:}^T\bm{c} - \bm{b}}.
        \]
        \IF {$\min \bm{w} \leq 0$}
                \STATE Recompute $\bm{w}$ using non-negative least squares 
                \[
                \argmin_{\bm{c} \geq 0} \nor{\bm{V}_{\rm target}\LRp{\mathcal{I},:}^T\bm{c} - \bm{b}}
                \]             
        \ENDIF
        \STATE Recompute $\bm{r} = \bm{b} - \bm{V}_{\rm target}\LRp{\mathcal{I},:}^T\bm{w}$.
\ENDWHILE
\end{algorithmic}
\label{alg:hr}
\end{algorithm}

Algorithm~\ref{alg:hr} requires both $\bm{V}_{\rm target}$ and $\bm{w}_{\rm target}$ as inputs. For all numerical experiments, we set $\bm{w}_{\rm target} = \Delta x \bm{1}$ and set $\bm{V}_{\rm target}$ to be the span of products of POD functions $\bm{V}$, such that
\[
\mathcal{R}\LRp{\bm{V}_{\rm target}} = {\rm span}\LRc{\bm{V}(:,i) \circ \bm{V}(:,j), \quad i,j = 1,\ldots, N}.
\]
This choice of target space ensures that the mass matrix $\Delta x \bm{V}^T\bm{V}$ is accurately approximated.  In practice, since many of the pointwise vector products $\bm{V}(:,i) \circ \bm{V}(:,j)$ are linearly dependent, we replace $\bm{V}_{\rm target}$ by a dimensionally reduced matrix computed using, for example, another application of the SVD.  %a randomized SVD \cite{halko2011finding, rsvd}.

\begin{remark}
We note that the target space is designed to accurately approximate the mass matrix $\Delta x \bm{V}^T\bm{V}$, but does not take the test basis $\bm{V}_{t}$ into account.  It is thus possible to construct a hyper-reduced quadrature which accurately integrates the mass matrix but violates Assumption~\ref{ass:quad}.  We have not encountered this situation in our experiments; however, one can ensure that Assumption~\ref{ass:quad} is satisfied by checking if $\bm{V}_t\LRp{\mathcal{I},:}$ is non-singular and selecting additional points based on $\bm{V}_t$ as necessary.  
\end{remark}

\subsection{An entropy conservative hyper-reduced model on periodic domains}

To summarize, an entropy conservative reduced order model on periodic domains can be constructed using the following ingredients and steps:
\begin{enumerate}
\item Compute a reduced basis matrix $\bm{V}$ from snapshots of both the conservative variables and the entropy variables.  
\item Construct ``test'' basis matrices $\bm{V}_t$ to produce a compressed ``modal'' differentiation matrix $\hat{\bm{Q}}_t = \bm{V}_t^T\bm{Q}\bm{V}_t$.
\item Compute a hyper-reduced quadrature using Algorithm~\ref{alg:hr}.  
\item Construct the hyper-reduced nodal differentiation matrix $\bm{Q}_t = \bm{P}_t^T\hat{\bm{Q}}_t\bm{P}_t$ using the projection matrix onto the test basis $\bm{P}_t$.
\end{enumerate}

%: \note{summarize restrictions and assumptions on hyper-reduction, POD, and test basis.}
\begin{theorem}
Suppose that the hyper-reduced quadrature satisfies Assumption~\ref{ass:quad}.  Then, the semi-discrete formulation 
\begin{gather}
\bm{M}_N\td{\bm{u}_N}{t} + 2\bm{V}\LRp{\mathcal{I},:}^T\LRp{\bm{Q}_t \circ \bm{F}}\bm{1} = \bm{0} \label{eq:hrrom}\\
\bm{M}_N = \bm{V}\LRp{\mathcal{I},:}^T\bm{W}\bm{V}\LRp{\mathcal{I},:}, \qquad \bm{P} = \bm{M}_N^{-1}\bm{V}\LRp{\mathcal{I},:}^T\bm{W},\nonumber\\ 
\bm{v}_N = \bm{P}\bm{v}\LRp{\bm{V}\LRp{\mathcal{I},:}\bm{u}_N}, \qquad \tilde{\bm{v}} = \bm{V}\LRp{\mathcal{I},:}\bm{v}_N, \nonumber\\
\tilde{\bm{u}} = \bm{u}\LRp{\tilde{\bm{v}}}, \qquad \bm{F}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j},\nonumber
\end{gather}
discretely conserves entropy 
\[
\Delta x \bm{1}^T\td{S\LRp{\bm{V}\bm{u}_N}}{t} = 0.  
\]
Additionally, if the reduced basis $\bm{V}$ exactly represents $\bm{1}$, solutions of (\ref{eq:hrrom}) conserve global averages of the conservative variables.
\label{thm:hrromes}
\end{theorem}
\begin{proof}
Since the hyper-reduction preserves the skew-symmetry and zero row sums of $\bm{Q}_t$, the proof is identical in structure to the proof of Theorem~\ref{thm:esrom}.  
\end{proof}

\begin{remark}
It is not strictly necessary to use the hyper-reduced mass matrix 
\[
\bm{M}_N = \bm{V}\LRp{\mathcal{I},:}^T\bm{W}\bm{V}\LRp{\mathcal{I},:}
\]
 to construct an entropy conservative ROM.  For example, we can use the exact mass matrix $\bm{M}_N = \Delta x \bm{V}^T\bm{V}$ rather than the hyper-reduced mass matrix.  If the projection matrix $\bm{P}$ is defined accordingly, the resulting ROM remains entropy conservative.  However, we observe significantly larger errors when using the exact mass matrix instead of the hyper-reduced mass matrix.  
\end{remark}

We note that the flux matrix $\bm{F}$ is not computed explicitly, but instead is usually formed on the fly.  The computation of $\bm{F}$ constitutes a significant portion of the computational cost, but can be reduced for the full order model by exploiting the sparsity of $\bm{Q}$.  However, since the hyper-reduced matrix $\bm{Q}_t$ is fully dense, the ROM (\ref{eq:hrrom}) requires $O(N_s^2)$ nonlinear flux evaluations, in contrast to the typical $O(N_s)$ nonlinear function evaluations required for a typical hyper-reduced model.

\subsection{Hyper-reduction and entropy dissipation}
\label{sec:diss}
Given an entropy conservative ROM with hyper-reduction, we can construct an entropy stable ROM by adding appropriate entropy-dissipative viscosity terms.  Recall that the artificial diffusion matrix $\bm{K}$ in (\ref{eq:esvisc}) does not depend nonlinearly on $\bm{u}$.  Thus, typical model reduction techniques compress such operators using Galerkin projection, e.g., $\bm{V}^T\bm{K}\bm{V}$.  However, since one cannot in general show that this operator dicretely dissipates entropy, we discuss two treatments of the viscous terms which are provably entropy dissipative.

This section introduces hyper-reduction procedures for $\bm{K}$ which provide a provable discrete dissipation of entropy.  We seek to mimic the entropy balance of the full order model in (\ref{eq:entropybalancefom}).  Suppose the reduced order model is given by 
\[
\bm{M}_N\td{\bm{u}_N}{t} + 2\bm{V}\LRp{\mathcal{I},:}^T\LRp{\bm{Q}_t \circ \bm{F}}\bm{1} + \bm{d}(\bm{u}_N)= \bm{0}
\]
where $\bm{d}(\bm{u})$ is some hyper-reduced approximation to the viscous terms.  Theorem~\ref{thm:hrromes} gives that the entropy balance of the reduced order model is
\[
\Delta x \bm{1}^T\td{S\LRp{\bm{V}\bm{u}_N}}{t} = -\bm{v}_N^T\bm{d}(\bm{u}_N).
\]
We will design hyper-reduced treatments of the viscous terms such that $\bm{v}_N^T\bm{d}(\bm{u}_N) \geq 0$, which implies that the reduced model is entropy stable.

Let $\bm{D}$ denote the $(K-1)\times K$ difference matrix
\[
\bm{D} = \frac{1}{\Delta x}\begin{bmatrix}
1 & -1 &&& \\
 & 1 & -1 && \\
 & & \ddots & \ddots&\\
&  & & 1 & -1
\end{bmatrix}.
\]
The artificial diffusion matrix in (\ref{eq:esvisc}) can be decomposed as $\bm{K} = \bm{D}^T\bm{D}$.  To construct hyper-reduced diffusion terms, we first compute (using the SVD) an auxiliary basis $\bm{V}_{\bm{D}}$ for $\bm{D}\bm{V}$, where $\bm{V}$ is the reduced basis matrix.  We then perform hyper-reduction to compute weights and sampled row indices such that 
\[
\bm{V}_{\bm{D}}^T\bm{V}_{\bm{D}} \approx \bm{V}_{\bm{D}}\LRp{\mathcal{I}_{\bm{D}},:}^T \bm{W}_{\bm{D}}\bm{V}_{\bm{D}}\LRp{\mathcal{I}_{\bm{D}},:}.
\]

We can now construct entropy-dissipative diffusion terms using this second hyper-reduction of $\bm{D}$.  The first treatment is relatively straightforward, and simply samples rows of $\bm{D}$ corresponding to indices $\mathcal{I}_{\bm{D}}$.  The viscous terms $\bm{K}\bm{u}$ in the full order model are approximated using the sparse hyper-reduced matrix $\bm{D}\LRp{\mathcal{I}_{\bm{D}},:}$
\begin{equation}
%\bm{K}\bm{u} \approx 
\bm{d}(\bm{u}_N) = \bm{D}\LRp{\mathcal{I}_{\bm{D}},:}^T \bm{W}_{\bm{D}} \bm{D}\LRp{\mathcal{I}_{\bm{D}},:}\tilde{\bm{u}}.
\label{eq:visc1}
\end{equation}
Since the entropy-projected conservative variables $\tilde{\bm{u}}$ are mappings of the projected entropy variables $\bm{V}\bm{v}_N$, repeating the steps used to derive (\ref{eq:entropydissfom}) yields that the entropy dissipation of (\ref{eq:visc1}) is 
\begin{align*}
\bm{v}_N^T\bm{V}^T\bm{D}^T\LRp{\mathcal{I}_{\bm{D}},:}^T \bm{W}_{\bm{D}} \bm{D}\LRp{\mathcal{I},:}\tilde{\bm{u}} &= 
\tilde{\bm{v}}^T\bm{D}^T\LRp{\mathcal{I}_{\bm{D}},:}^T \bm{W}_{\bm{D}} \bm{D}\LRp{\mathcal{I},:}\bm{u}\LRp{\tilde{\bm{v}}}
\\
&= \sum_{i\in \mathcal{I}_{\bm{D}}} \frac{w_{\bm{D},i}}{\Delta x^2} \LRl{\pd{\bm{u}}{\bm{v}}}_{i,i+1} \LRp{\tilde{\bm{v}}_i - \tilde{\bm{v}}_{i+1}}^2.
\end{align*}

The second treatment of the diffusion terms is motivated by approaches taken in \cite{carpenter2014entropy, chen2017entropy, zakerzadeh2017entropy, upperman2019entropy}.  Recall that, using the mean value theorem, the $i$th entry of $\bm{D}\bm{u}$ is
\begin{equation}
\LRp{\bm{D}\bm{u}}_i = \bm{u}_i - \bm{u}_{i+1} = \LRl{\pd{\bm{u}}{\bm{v}}}_{i,i+1} \LRp{\bm{v}(\bm{u}_i)-\bm{v}(\bm{u}_{i+1})} = \LRl{\pd{\bm{u}}{\bm{v}}}_{i,i+1}  \LRp{\bm{D}\bm{v}(\bm{u})}_i,
\label{eq:meanval}
\end{equation}
where $\LRl{\pd{\bm{u}}{\bm{v}}}_{i,i+1}$ is $\pd{\bm{u}}{\bm{v}}$ evaluated at some intermediate state between $\bm{u}_i$ and $\bm{u}_{i+1}$.  This motivates a hyper-reduction based on sampling and weighting rows of $\bm{D}$ applied to the vector of entropy variables $\bm{v}(\bm{u})$.  The diffusion terms can be approximated by 
\begin{equation}
%\bm{K}\bm{u} \approx 
\bm{d}(\bm{u}_N) = \bm{V}^T\bm{D}\LRp{\mathcal{I}_{\bm{D}},:}^T \bm{W}_{\bm{D}} \bm{H} \bm{D}\LRp{\mathcal{I}_{\bm{D}},:}\bm{V} \bm{v}_N
\label{eq:visc2}
\end{equation}
where $\bm{v}_N$ are coefficients of the projected entropy variables in (\ref{eq:hrrom}), $\bm{W}_{\bm{D}}$ is a diagonal matrix of positive weights $w_{\bm{D},i}$, and $\bm{H}$ is a diagonal matrix containing entries of the Jacobian $\pd{\bm{u}}{\bm{v}}$\footnote{Recall that matrices are treated as Kronecker products, and that the diagonal entries of $\bm{H}_{ii}$ correspond to block matrices acting on the vector of solution components at a point.} evaluated at all hyper-reduced points in $\mathcal{I}_{\bm{D}}$ 
\[
\bm{H}_{ii} = \LRl{\pd{\bm{u}}{\bm{v}}}_{\bm{u} = \bar{\bm{u}}_i}, \qquad \bar{\bm{u}}_i = \frac{\tilde{\bm{u}}_i + \tilde{\bm{u}}_{i+1}}{2}.
\]
Since we do not know the intermediate state used to evaluate $\LRl{\pd{\bm{u}}{\bm{v}}}_{i,i+1}$ in (\ref{eq:meanval}), 
we have arbitrarily chosen to evaluate $\pd{\bm{u}}{\bm{v}}$ at the average of the entropy-projected conservative values $\tilde{\bm{u}}_i$ and $\tilde{\bm{u}}_{i+1}$.  Since the diagonal entries of $\bm{W}_{\bm{D}}$ are positive and $\pd{\bm{u}}{\bm{v}}$ is symmetric positive definite (for physically relevant values of $\tilde{\bm{u}}$), the entropy dissipation of the hyper-reduced diffusion term is given by
\[
\bm{v}_N^T\bm{V}^T\bm{D}^T\LRp{\mathcal{I}_{\bm{D}},:}^T \bm{W}_{\bm{D}} \bm{H} \bm{D}\LRp{\mathcal{I},:}\bm{V} \bm{v}_N = 
\sum_{i\in \mathcal{I}_{\bm{D}}} \frac{w_{\bm{D},i}}{\Delta x^2} \LRl{\pd{\bm{u}}{\bm{v}}}_{\bm{u}=\bar{\bm{u}}_i} \LRp{\tilde{\bm{v}}_i - \tilde{\bm{v}}_{i+1}}^2.
\]
This implies that the dissipation of entropy produced by the hyper-reduced approximation of the viscous terms (\ref{eq:visc2}) also mimics the dissipation of entropy (\ref{eq:entropydissfom}) derived for the full order model.
The difference between the hyper-reduction treatments of viscosity (\ref{eq:visc1}) and (\ref{eq:visc2}) are that (\ref{eq:visc1}) relies explicitly on the fact that $\bm{D}$ is a two-point finite volume difference matrix, which enables the use of the mean value theorem in deriving a statement of entropy dissipation.   The first approach (\ref{eq:visc1}) results in a simpler online phase, though it involves larger matrices compared to the second approach.  The second approach (\ref{eq:visc2}) is not limited to finite volume methods, and is entropy dissipative for any choice of differentiation matrix $\bm{D}$.  Both approaches require computing an additional set of hyper-reduced points during the offline phase.

\begin{remark}
The hyper-reduction treatments of viscosity described in this section can also be applied to interface dissipation, such as local Lax-Friedrichs penalization.  %However, these dissipation terms depend on the resolution of the full order model.  
\end{remark}

We note that, in practice, a naive approximation of diffusive terms $\bm{K}\bm{u}$ by
\begin{equation}
%\bm{K}\bm{u} \approx 
\bm{d}(\bm{u}_N) = \bm{V}^T\bm{K}\bm{V}\bm{P}\tilde{\bm{u}},
\label{eq:visc3}
\end{equation}
is often effective.  This approach is simpler to implement than (\ref{eq:visc1}),(\ref{eq:visc2}), and does not require a second set of hyper-reduced points.  We are unable to show that this treatment is provably entropy dissipative.  However, in all experiments, we observe that $\bm{v}_N^T\bm{V}^T\bm{K}\bm{V}\bm{P}\tilde{\bm{u}} \geq 0$, which implies that entropy is discretely dissipated.

\section{Weak imposition of boundary conditions}
\label{sec:6}
Until now, we have assumed periodic domains.  In this section, we discuss how to extend the construction of entropy stable ROMs to the non-periodic case.  Boundary conditions are weakly imposed through a numerical flux and a ``hybridized'' summation by parts operator \cite{chan2017discretely}.

In the non-periodic case, the matrix $\bm{Q}$ is nearly skew-symmetric, such that it satisfies a summation by parts property
\begin{equation}
\bm{Q} = \frac{1}{2}\begin{bmatrix}
-1 & 1 & & \\
-1 & 0 & \ddots &\\
 &\ddots &\ddots &1\\
  & & -1 &1\\
\end{bmatrix}, \qquad \bm{Q} + \bm{Q}^T = \bm{B}_{\rm SBP} = 
\begin{bmatrix}
-1& && \\
&0 &&\\
& &\ddots &\\
& &&1\\
\end{bmatrix}
\label{eq:fomsbp}
\end{equation}
A non-periodic full order model can be expressed as 
\begin{align}
\Delta x \td{\bm{u}_h}{t} + 2\LRp{{\bm{Q}}\circ \bm{F}}\bm{1} + \bm{B}_{\rm SBP}\LRp{\bm{f}^* - \bm{f}(\bm{u})} = \bm{0},
\label{eq:esbc}
\end{align}
where $\bm{f}^* = \LRs{\bm{f}^*_0, 0,\ldots, 0, \bm{f}^*_K}^T$ is a vector containing values of the nonlinear flux at the boundary points.  
Following \cite{chen2017entropy}, a boundary numerical flux $\bm{f}^*$ is entropy stable if 
\begin{equation}
\psi\LRp{\bm{u}} - \bm{v}(\bm{u})^T\bm{f}^* \leq \bm{0}.%, \qquad \text{check this!}}
\label{eq:esbflux}
\end{equation}
The flux is referred to as entropy conservative if the inequality is an equality.  If an entropy stable or entropy conservative flux $\bm{f}_S(\bm{u}_L,\bm{u}_R)$ is used to evaluate both the volume and surface flux, then the resulting full order model given by (\ref{eq:esbc}) is also entropy stable or entropy conservative \cite{tadmor1987numerical, parsani2015entropy, chen2017entropy}.

Extending entropy conservative treatments of boundary conditions to the reduced order model is less straightforward due to fact that reduced matrices do not satisfy the same SBP property (\ref{eq:fomsbp}).  If the full order matrix $\bm{Q}$ satisfies the SBP property, then the hyper-reduction of $\bm{Q}$ satisfies a \textit{generalized} SBP property \cite{fernandez2014generalized} 
\[
\bm{Q}_t + \bm{Q}_t^T = \bm{E}^T\bm{B}\bm{E}, \qquad \bm{Q}_t = \bm{P}_t^T\bm{V}_t^T\bm{Q}\bm{V}_t\bm{P}_t.
\]
where the matrices $\bm{B}$ and $\bm{E}$ encode scaling by normals and evaluation at boundary points.  In 1D, these matrices are given by
\[
\bm{B} = \begin{bmatrix}-1 & \\ & 1\end{bmatrix}, \qquad \bm{E} = \bm{V}_{bt}\bm{P}_t, \qquad \bm{V}_{bt} = \begin{bmatrix}\bm{V}_t(1,:) \\ \bm{V}_t(N,:)\end{bmatrix}.
\]
where $\bm{V}_t(:,1), \bm{V}_t(:,K)$ denote the first and last columns (corresponding to points on the boundary) of the test basis matrix $\bm{V}_t$.

For linear problems, it is possible to impose energy stable boundary conditions by combining generalized SBP operators with appropriate simultaneous approximation terms (SATs) \cite{fernandez2014review, fernandez2018simultaneous}.  However, due to the presence of $\bm{E}$ within the boundary term $\bm{E}^T\bm{B}\bm{E}$, the accurate and entropy stable imposition of nonlinear boundary conditions using generalized SBP SATs remains an open problem \cite{crean2018entropy, chan2018efficient, chenreview}. We address this issue by constructing a \textit{hybridized} SBP operator (also referred to as a decoupled SBP operator) \cite{chan2017discretely, chan2019skew}
\[
\bm{Q}_{h} = \frac{1}{2}\begin{bmatrix}
\bm{Q}_t - \bm{Q}_t^T & \bm{E}^T\bm{B} \\
-\bm{B}\bm{E} & \bm{B}
\end{bmatrix}.
\]
Note that $\bm{Q}_h$ satisfies a block version of the standard summation by parts property by construction
\[
\bm{Q}_h + \bm{Q}_h^T = \bm{B}_h = \begin{bmatrix}
\bm{0} & \\
& \bm{B} \end{bmatrix},  
\]
where $\bm{B}_h$ is diagonal and does not involve the boundary evaluation matrix $\bm{E}$.

A more subtle property is that $\bm{Q}_h\bm{1} = \bm{0}$ \cite{chan2019skew}.  Note that the full order matrix $\bm{Q}$ in (\ref{eq:fomsbp}) satisfies $\bm{Q}\bm{1} = \bm{0}$.  Then, since $\bm{1}$ is exactly representable in the test basis $\bm{V}_t$, $\bm{V}_t\bm{P}_t\bm{1} = \bm{1}$, and
\[
\bm{Q}_t = \bm{P}_t^T\bm{V}_t^T\bm{Q}\bm{V}_t\bm{P}_t\bm{1} = \bm{P}_t^T\bm{V}_t^T\bm{Q}\bm{1} = \bm{0}.
\]
This implies that $\bm{Q}_h\bm{1}$ simplifies to
\[
\bm{Q}_h \bm{1} =  \frac{1}{2}\begin{bmatrix}
\bm{Q}_t \bm{1} - \bm{Q}_t^T \bm{1} + \bm{E}^T\bm{B} \bm{1} \\
- \bm{B}\bm{E} \bm{1} + \bm{B}
\end{bmatrix}=\frac{1}{2}\begin{bmatrix}
- \bm{Q}_t^T \bm{1} + \bm{E}^T\bm{B} \bm{1} \\
\bm{0}
\end{bmatrix},
\]
where we have used that $\bm{Q}_t\bm{1} = \bm{0}$ and that (\ref{eq:preproduce}) implies $\bm{E}\bm{1} = \bm{1}$.   Here, we abuse notation by letting $\bm{1}$ denote any arbitrary length vector of all ones.  Furthermore, since $\bm{Q}_t$ satisfies a generalized SBP property,
\[
- \bm{Q}_t^T \bm{1} + \bm{E}^T\bm{B} \bm{1} = \LRp{-\bm{Q}_t^T + \bm{E}^T\bm{B}\bm{E}} \bm{1} = \bm{Q}_t \bm{1} = \bm{0}.
\]
we conclude that $\bm{Q}_h\bm{1} = \bm{0}$.

By replacing $\bm{Q}_t$ in (\ref{eq:hrrom}) with $\bm{Q}_h$, we can impose boundary conditions weakly through a numerical flux $\bm{f}^*$.  We define $\bm{V}_b$ as the matrix which evaluates the reduced basis at boundary points $\LRc{-1,1}$, and define $\bm{V}_h$ as the matrix which evaluates the reduced basis at both boundary points and hyper-reduced volume points
\[
\bm{V}_{b} = \begin{bmatrix}\bm{V}(1,:) \\ \bm{V}(N,:)\end{bmatrix}, \qquad \bm{V}_h = \begin{bmatrix}
\bm{V}\LRp{\mathcal{I},:}\\
\bm{V}_{b}
\end{bmatrix}.
\]
We then have the following theorem:
\begin{theorem}
If the boundary flux $\bm{f}^*$ is entropy stable in the sense of (\ref{eq:esbflux}), then the following hyper-reduced ROM 
\begin{gather}
\bm{M}_N\td{\bm{u}_N}{t} + 2\bm{V}_h^T\LRp{\bm{Q}_h \circ \bm{F}}\bm{1} + \bm{V}_{b}^T\bm{B}\LRp{\bm{f}^* - \bm{f}(\tilde{\bm{u}}_b)} = \bm{0} \label{eq:esbchr}\\
\bm{v}_N = \bm{P}\bm{v}\LRp{\bm{V}\LRp{\mathcal{I},:}\bm{u}_N}, \qquad \tilde{\bm{v}} = 
\bm{V}_h
\bm{v}_N, \qquad \tilde{\bm{v}}_b = \bm{V}_b\bm{v}_N, \nonumber\\
\tilde{\bm{u}} = \bm{u}\LRp{\tilde{\bm{v}}}, \qquad \bm{F}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad  \tilde{\bm{u}}_b = \bm{u}\LRp{\tilde{\bm{v}}_b}\nonumber
\end{gather}
is also entropy stable such that
\[
\td{S\LRp{\bm{V}\bm{u}_N}}{t} \leq 0,
\]
with equality holding for an entropy conservative flux.
\label{thm:esbchr}
\end{theorem}
\begin{proof}
The proof is identical in structure to the proof of Theorem 1 in \cite{chan2017discretely}.  We reproduce it here for completeness.
Testing with $\bm{v}_N$ and using the summation by parts property of $\bm{Q}_h$ then yields 
\begin{align*}
\td{S\LRp{\bm{V}\bm{u}_N}}{t}  + \tilde{\bm{v}}^T\LRp{\LRp{\bm{Q}_h-\bm{Q}_h} \circ \bm{F}}\bm{1} + \tilde{\bm{v}}_b^T\bm{B}\bm{f}^*  = \bm{0}.
\end{align*}
Here, we have used that, by the consistency of the entropy conservative flux (\ref{eq:esflux}), the diagonal entries of $\bm{F}$ are $\bm{F}_{ii} = \bm{f}\LRp{\tilde{\bm{u}}_i}$.  Thus, since $\bm{B}_h$ is also diagonal,
\[
\LRp{\bm{B}_h\circ\bm{F}}\bm{1} = \bm{B} \bm{f}(\tilde{\bm{u}}_b).
\]
Expanding out the term $\tilde{\bm{v}}^T\LRp{\LRp{\bm{Q}_h-\bm{Q}_h^T} \circ \bm{F}}\bm{1}$ and using skew-symmetry of $\bm{Q}_h-\bm{Q}_h^T$ yields
\begin{align*}
\tilde{\bm{v}}^T\big(\underbrace{\LRp{\bm{Q}_h-\bm{Q}_h^T}}_{\bm{S}_h} \circ \bm{F}\big)\bm{1} &= 
%\sum_{ij} \LRp{\bm{S}_h}_{ij} \tilde{\bm{v}}_i^T\bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j} \\
%&= 
\frac{1}{2}\sum_{ij} \LRp{\bm{S}_h}_{ij} \LRp{\tilde{\bm{v}}_i-\tilde{\bm{v}}_j}^T\bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}\\
&= \frac{1}{2}\sum_{ij} \LRp{\bm{Q}_h-\bm{Q}_h^T}_{ij} \LRp{\psi(\tilde{\bm{u}}_i)-\psi(\tilde{\bm{u}}_j)}\\
&= \sum_{ij} \LRp{\bm{Q}_h}_{ij} \LRp{\psi(\tilde{\bm{u}}_i)-\psi(\tilde{\bm{u}}_j)} = {\psi(\tilde{\bm{u}})^T\bm{Q}_h\bm{1} - \bm{1}^T\bm{Q}_h\psi(\tilde{\bm{u}})}\\
& = -\bm{1}^T\bm{B}_h\psi(\tilde{\bm{u}}) = -\bm{1}^T\bm{B}\psi(\tilde{\bm{u}}_b),
\end{align*}
where we have used that $\bm{Q}_h\bm{1} = \bm{0}$ and the SBP property in the last two lines.  Using that $\bm{B}$ is diagonal, straightforward manipulations imply that
\[
\td{S\LRp{\bm{V}\bm{u}_N}}{t}  - \bm{1}^T\bm{B}\LRp{\psi(\tilde{\bm{u}}_b)-\tilde{\bm{v}}_b^T\bm{f}^*}  = \bm{0}.
\]
If the flux is entropy stable, then $\td{S\LRp{\bm{V}\bm{u}_N}}{t} \leq 0$, with equality holding if the flux is entropy conservative.
\end{proof}

%\note{Discuss issues with high order accuracy and entropy stability for SBP \cite{chan2018efficient}.}

\section{Extension to higher dimensions}
\label{sec:7}

\subsection{Periodic domains}
For periodic domains, the extension to higher spatial dimensions is relatively straightforward.  We provide a concrete construction of matrices and formulations in two dimensions.  For finite volume methods on a $K \times K$ structured quadrilateral elements of size $\Delta x \times \Delta x$, differentiation matrices along each coordinate direction can be constructed as Kronecker products of one-dimensional matrices.  Let $\bm{Q}_{\rm 1D}$ denote the one-dimensional periodic differentiation matrix defined in (\ref{eq:Qmat}).  The differentiation matrix along the $i$th coordinate direction $\bm{Q}^i$ can be constructed as follows
\[
\bm{Q}^1 = \bm{Q}_{\rm 1D}\otimes \Delta x \bm{I}, \qquad \bm{Q}^2 = \Delta x\bm{I} \otimes \bm{Q}_{\rm 1D}.
\]
Given a basis matrix $\bm{V}$ and indices of hyper-reduced points $\mathcal{I}$, we can introduce test basis matrices $\bm{V}^i_t$ such that the range of  $\bm{V}^i_t$ is equal to the direct sum of the ranges of $\bm{V}$ and $\bm{Q}^i\bm{V}$
\[
\mathcal{R}\LRp{\bm{V}^i_t} = \mathcal{R}\LRp{\begin{bmatrix}\bm{V} &\bm{Q}^i\bm{V}\end{bmatrix}}, \qquad i = 1,\ldots,d.
%\mathcal{R}\LRp{\bm{V}^y_t} = \mathcal{R}\LRp{\begin{bmatrix}\bm{V} &\bm{Q}^y\bm{V}\end{bmatrix}}.
\]
Note that, unlike the 1D case, the test bases differ along each coordinate direction.  However, the dimension of the range of each test basis $\bm{V}_t^i$ is still at most $2N+1$.  
We can now define hyper-reduced projection matrices
\begin{align*}
\bm{P}^i_t = \LRp{\bm{V}^i_t\LRp{\mathcal{I},:}^T\bm{W}\bm{V}^i_t\LRp{\mathcal{I},:}}^{-1}\bm{V}^i_t\LRp{\mathcal{I},:}^T\bm{W}.%,\\
%\bm{P}^y_t = \LRp{\bm{V}^y_t\LRp{\mathcal{I},:}^T\bm{W}\bm{V}^y_t\LRp{\mathcal{I},:}}^{-1}\bm{V}^y_t\LRp{\mathcal{I},:}^T\bm{W}.
\end{align*}
These test bases can be used to construct reduced matrices $\bm{Q}_t^i$ %, \bm{Q}_t^y$ 
\[
\bm{Q}_t^i = \LRp{\bm{P}^i_t}^T\LRp{\LRp{\bm{V}^i_t}^T\bm{Q}^i\bm{V}^i_t}\bm{P}^i_t,
% \qquad 
%\bm{Q}_t^y = \LRp{\bm{P}^y_t}^T\LRp{\LRp{\bm{V}^y_t}^T\bm{Q}^y\bm{V}^y_t}\bm{P}^y_t.
\]
which are used to construct an entropy stable reduced order model in $d$ dimensions
\begin{gather}
\bm{M}_N\td{\bm{u}_N}{t} + \sum_{i=1}^d 2\bm{V}\LRp{\mathcal{I},:}^T\LRp{\bm{Q}^i_t \circ \bm{F}^i}\bm{1} = \bm{0} \label{eq:hrromd}\\
\bm{M}_N = \Delta x^2 \bm{V}^T\bm{V}, \qquad \bm{F}^i_{jk} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}.\nonumber
%\bm{M}_N = \bm{V}\LRp{\mathcal{I},:}^T\bm{W}\bm{V}\LRp{\mathcal{I},:}, \qquad \bm{P} = \bm{M}_N^{-1}\bm{V}\LRp{\mathcal{I},:}^T\bm{W},\nonumber\\ 
%\bm{v}_N = \bm{P}\bm{v}\LRp{\bm{V}\LRp{\mathcal{I},:}\bm{u}_N}, \qquad \tilde{\bm{v}} = \bm{V}\LRp{\mathcal{I},:}\bm{v}_N, \nonumber\\
%\tilde{\bm{u}} = \bm{u}\LRp{\tilde{\bm{v}}}, \qquad ,\nonumber
\end{gather}
Here, $\bm{f}^i_S\LRp{\bm{u}_L,\bm{u}_R}$ is the entropy conservative flux in the $i$th coordinate direction and $\bm{M}_N$ is the mass matrix as defined in (\ref{eq:hrrom}).  We can show that (\ref{eq:hrromd}) is entropy conservative by repeating steps in the proof of Theorem~\ref{thm:hrromes} for each coordinate direction.

\subsection{Non-periodic domains}

Non-periodic domains and boundary conditions can be treated by combining hybridized SBP operators and a weak imposition of boundary conditions as was done in the 1D hyper-reduced formulation (\ref{eq:esbchr}).  Without loss of generality, we assume a square domain with $4$ boundary faces, each of which is discretized by $K$ intervals of size $\Delta x$.  We assume we are given values of the outward normal vector $\bm{n}$ on the domain boundary
\[
\bm{n} = \LRs{n_1, \ldots, n_d}^T.
\]
We can enforce non-periodic boundary conditions through introducing numerical fluxes at boundary points.  However, the number of boundary points scales with $K$, the number of elements in the full order model along one coordinate direction.  To ensure that the cost of the reduced order model scales with the number of modes $N$ rather than the dimension of the full order model, we must apply an additional hyper-reduction to approximate boundary terms by a weighted combination of sub-sampled boundary points.

Let $\bm{V}_b$ denote the matrix whose columns correspond to values of the reduced basis at boundary points.  Let $\mathcal{I}_b$ denote a sub-sampled set of $N_b$ boundary points, and let $g$ denote a nonlinear function.  We then approximate the boundary inner product 
\[
\bm{V}_b^Tg(\bm{V}_b\bm{u}_N) \approx \bm{V}_b\LRp{\mathcal{I}_b,:}^T \bm{W}_b g\LRp{\bm{V}_b\LRp{\mathcal{I}_b,:}}.
\]
where $\bm{W}_b$ is a diagonal matrix whose entries consist of positive weights
\[
\bm{W}_b = {\rm diag}\LRp{\bm{w}_b}.
\]
We can now define the $N_b\times N_b$ operators $\bm{B}^1, \bm{B}^2$ as the diagonal matrices whose entries consist of weighted component values of outward normals on the boundary 
\[
\bm{B}^i = {\rm diag}\LRp{\bm{n}^i} \bm{W}_b,
\]
where $\bm{n}^i$ is a vector containing the values of the component $n_i$ at boundary points.

We can now construct hybridized SBP operators under both volume and boundary hyper-reduction.  Let $\bm{V}_{bt}$ be the matrix which evaluates the $i$th test basis at face points.  Then, we can define $\bm{E}_i$ as the matrix which extrapolates from hyper-reduced volume points to boundary points through projection onto the $i$th test basis
\[
\bm{E}_i = \bm{V}_{bt}^i\LRp{\mathcal{I}_b,:}\bm{P}^i_t.  
\]
The hybridized SBP operator for differentiation along the $i$th coordinate is then
\[
\bm{Q}_h^i = \begin{bmatrix}
\bm{Q}^i_t - \LRp{\bm{Q}^i_t}^T & \bm{E}_i^T\bm{B}^i \\ 
-\bm{B}^i\bm{E}_i & \bm{B}^i
\end{bmatrix}.
\]

These operators can be used to construct a hyper-reduced formulation in higher dimensions
\begin{gather}
\bm{M}_N\td{\bm{u}_N}{t} + \sum_{i=1}^d \LRp{2\bm{V}\LRp{\mathcal{I},:}^T\LRp{\bm{Q}^i_h \circ \bm{F}^i}\bm{1} + \bm{V}_b\LRp{\mathcal{I}_b,:}^T\bm{B}^i\LRp{\bm{f}_i^* - \bm{f}_i(\bm{u}_b)}} = \bm{0} \label{eq:hrromdbc},%\\
\end{gather}
where $\bm{M}_N$ and $\bm{F}^i$ are as defined in (\ref{eq:hrromd}), and $\bm{f}_i(\bm{u}), \bm{f}_i^*$ are the $i$th components of the flux function and boundary numerical flux, respectively.

In contrast to the 1D case, additional steps are necessary to ensure that entropy stability is preserved under this boundary hyper-reduction.  Recall that the proof of entropy stability for 1D non-periodic domains in Theorem~\ref{thm:esbchr} requires that $\bm{Q}_h\bm{1} = \bm{0}$.  The analogous proof of entropy stability in multiple dimensions requires that $\bm{Q}^i_h\bm{1} = \bm{0}$ for $i = 1,\ldots, d$.  However, this is not automatically satisfied.  Expanding out $\bm{Q}^i_h\bm{1}$ yields
\[
\bm{Q}^i_h\bm{1} = \begin{bmatrix}
\bm{Q}^i_t\bm{1} - \LRp{\bm{Q}^i_t}^T\bm{1} + \bm{E}_i^T\bm{B}^i \bm{1}\\
\bm{0}
\end{bmatrix}= \begin{bmatrix}
-\LRp{\bm{Q}^i_t}^T\bm{1} + \bm{E}_i^T\bm{B}^i \bm{1}\\
\bm{0}
\end{bmatrix}.
\]
In general, $-\LRp{\bm{Q}^i_t}^T\bm{1} + \bm{E}^T\bm{B}^i \bm{1}\neq 0$, so $\bm{Q}^i_h\bm{1} \neq \bm{0}$ and the proof of entropy stability does not hold.  To enforce that $\bm{Q}^i_h\bm{1} = \bm{0}$, we impose constraints on the boundary weights $\bm{w}_b$.  Note that 
\begin{align*}
-\LRp{\bm{Q}^i_t}^T\bm{1} + \bm{E}^T\bm{B}^i \bm{1} &= -\LRp{\bm{P}^i_t}^T\bm{V}_t^T\bm{Q}\bm{V}_t\bm{P}^i_t\bm{1} + \LRp{\bm{P}^i_t}^T\LRp{\bm{V}^i_{bt}\LRp{\mathcal{I}_b,:}}^T\bm{B}^i\bm{1}\\
&= \LRp{\bm{P}^i_t}^T\LRp{-\LRp{\bm{V}^i_t}^T\bm{Q}\bm{1} + \LRp{\bm{V}^i_{bt}}\LRp{\mathcal{I}_b,:}^T{\rm diag}\LRp{\bm{n}^i}\bm{w}_b}.
\end{align*}
where we have used that $\bm{V}^i_t\bm{P}^i_t\bm{1} = \bm{1}$.  Thus, to ensure that $\bm{Q}^i_h\bm{1} = 0$, it is sufficient to guarantee that
\begin{equation}
\LRp{\bm{V}^i_{bt}}\LRp{\mathcal{I}_b,:}^T{\rm diag}\LRp{\bm{n}^i}\bm{w}_b = \LRp{\bm{V}^i_t}^T\bm{Q}\bm{1}, \qquad i = 1,\ldots, d.
\label{eq:sbpconstraints}
\end{equation}
The $dN$ constraints encoded in (\ref{eq:sbpconstraints}) can be interpreted as enforcing a discrete version of the fundamental theorem of calculus relating integrals of function derivatives to boundary integrals of function values.  These constraints on the boundary weights $\bm{w}_b$ can be enforced, for example, by augmenting the non-negative least squares solve in Algorithm~\ref{alg:hr} with equality constraints.  We take an alternative approach in this paper, and incorporate the constraints (\ref{eq:sbpconstraints}) directly into hyper-reduction approaches based on linear programming.  We refer the reader to  \cite{yano2019lp} for details.  %\note{Expand}

\subsection{Entropy dissipation}

\note{Discuss how to produce entropy dissipative hyper-reduced operators in 2D}.

%Weak BCs using decoupled SBP operators \cite{chan2017discretely}.  In 1D, the surface quadrature is exact, and one can show that Galerkin projections of diagonal norm SBP operators satisfy a generalized SBP property.  However, in 2D, the boundary is also hyper-reduced, and the surface quadrature becomes inexact.  Thus, the generalized SBP property is not satisfied and entropy stability is lost on non-periodic domains.  To address this, we relax conditions on quadrature accuracy using theory from \cite{chan2019skew}, and describe the use of a constrained least squares solve to ensure that the quadrature rule generated by the empirical cubature procedure leads to a GSBP property.
%
%Sufficient conditions for semi-discrete entropy stability are $\bm{1}^T\bm{Q}_N = \bm{1}^T\bm{B}\bm{E}$.  Given a surface quadrature rule and surface normals, $\bm{1}^T\bm{B}\bm{E}$ is given.  Assuming that the projection exactly recovers constants, $\bm{1}^T\bm{Q}_N = \bm{1}^T\bm{Q}\bm{V}\bm{P}$.  Multiplying both sides by $\bm{V}$ yields
%\[
%\bm{1}^T\bm{Q}\bm{V} = \bm{1}^T\bm{B}\bm{E}\bm{V}  = \bm{1}^T\bm{B}\bm{V}_f = \bm{w}_f^T\diag{\bm{n}}\bm{V}_f
%\]
%This condition depends only on the basis matrix $\bm{V}$, the face basis matrix $\bm{V}_f$, and the outward surface normals $\bm{n}$.  We can encode these as equality constraints within a linear program \cite{yano2019lp, patera2017lp}.  More computationally tractable because the optimization is only over the surface points, and LPs become more expensive with the number of constraints (e.g. number of modes $N$) too.

\section{Numerical experiments}
\label{sec:8}
{In this section, we study the behavior of entropy stable reduced order models for} the compressible Euler equations.  In $d$ dimensions, these are given by:
\begin{align*}
\pd{\rho}{t} &+ {\sum_{j=1}^d \pd{\LRp{\rho {u}_j}}{x_j}} = 0,\\
{\pd{\rho {u}_i}{t}} &+ {\sum_{j=1}^d \pd{\LRp{\rho u_iu_j + p\delta_{ij} }}{x_j}} = 0, \qquad i = 1,\ldots,d\\
\pd{E}{t} &+ {\sum_{j=1}^d \pd{\LRp{{u}_j(E+p)}}{x_j}} = 0.\nonumber
\end{align*}
Here, $\rho$ is density, {$u_i$ is the $i$th component of} velocity, and $E$ is the total energy.  The pressure $p$ and specific internal energy $\rho e$ are given by 
\[
{p = (\gamma-1)\LRp{E - \frac{1}{2}\rho \LRp{\sum_{j=1}^d u_j^2}}}, \qquad {\rho e = E - \frac{1}{2}\rho \LRp{\sum_{j=1}^d u_j^2}}. 
\]
%There exists an infinite family of suitable convex entropies for the compressible Euler equations \cite{harten1983symmetric}.  However, 
There is a unique entropy which symmetrrizes the viscous heat conduction term in the compressible Navier-Stokes equations \cite{hughes1986new}.  This entropy $S(\bm{u})$ is given by
\begin{equation*}
S(\bm{u}) = -\frac{\rho s}{\gamma-1}, 
\end{equation*}
where $s = \log\LRp{\frac{p}{\rho^\gamma}}$ is the physical specific entropy, and the dimension $d = 1,2,3$.  The entropy variables in $d$ dimensions are given by
\begin{gather*}
v_1 = \frac{\rho e (\gamma + 1 - s) - E}{\rho e}, \qquad v_{d+2} = -\frac{\rho}{\rho e},\\
v_{1+ i}= \frac{\rho {{u}_i}}{\rho e}, \qquad i = 1,\ldots, d.
\end{gather*}
while the conservation variables in terms of the entropy variables are given by
\begin{gather*}
\rho = -(\rho e) v_{d+2}, \qquad E = (\rho e)\LRp{1 - \frac{\sum_{j=1}^d{v_{1+j}^2}}{2 v_{d+2}}}\\
 \rho {u_i} = (\rho e) v_{1+i}, \qquad i = 1,\ldots,d,
\end{gather*}
where the quantities $\rho e$ and $s$ in terms of the entropy variables are 
\begin{equation*}
\rho e = \LRp{\frac{(\gamma-1)}{\LRp{-v_{d+2}}^{\gamma}}}^{1/(\gamma-1)}e^{\frac{-s}{\gamma-1}}, \qquad s = \gamma - v_1 + \frac{\sum_{j=1}^d{v_{1+j}^2}}{2v_{d+2}}.
\end{equation*}
{Let $f_i, f_j$ denote two arbitrary values.  We define the average and logarithmic averages as follows:}
\[
{\avg{f} = \frac{f_i + f_j}{2}}, \qquad {\avg{f}^{\log} = \frac{f_i - f_j}{\log\LRp{f_i}-\log\LRp{f_j}}}.
\]
To ensure numerical stability when the denominator is close to zero, we evaluate the logarithmic average using the algorithm of \cite{ismail2009affordable}.  
Explicit expressions for entropy conservative numerical fluxes in two dimensions  are given by Chandrashekar \cite{chandrashekar2013kinetic}
\begin{align*}
&f^1_{1,S}(\bm{u}_L,\bm{u}_R) = \avg{\rho}^{\log} \avg{u_1},& &f^1_{2,S}(\bm{u}_L,\bm{u}_R) = \avg{\rho}^{\log} \avg{u_2},&\\
&f^2_{1,S}(\bm{u}_L,\bm{u}_R) = f^1_{1,S} \avg{u_1} + p_{\rm avg},&  &f^2_{2,S}(\bm{u}_L,\bm{u}_R) = f^1_{2,S} \avg{u_1},&\nonumber\\
&f^3_{1,S}(\bm{u}_L,\bm{u}_R) = f^2_{2,S},& &f^3_{2,S}(\bm{u}_L,\bm{u}_R) = f^1_{2,S} \avg{u_2} + p_{\rm avg},&\nonumber\\
&f^4_{1,S}(\bm{u}_L,\bm{u}_R) = \LRp{E_{\rm avg} + p_{\rm avg}}\avg{u_1},& &f^4_{2,S}(\bm{u}_L,\bm{u}_R) = \LRp{E_{\rm avg} + p_{\rm avg} }\avg{u_2},& \nonumber
\end{align*}
where we have introduced the auxiliary quantities 
\begin{gather}
{\beta = \frac{\rho}{2p}}, \qquad p_{\rm avg} = \frac{\avg{\rho}}{2\avg{\beta}}, \qquad \qquad E_{\rm avg} = \frac{\avg{\rho}^{\log}}{2\avg{\beta}^{\log}\LRp{\gamma -1}}   + \frac{{u_{\rm avg}^2}}{2}, \label{eq:fluxaux} \\
{u^2_{\rm avg}} = 2(\avg{u_1}^2 + \avg{u_1}^2) - \LRp{\avg{u_1^2} +\avg{u_2^2}} \nonumber.  
\end{gather}

\note{Add equations for $\pd{\bm{u}}{\bm{v}}$. }

\subsection{1D Euler equations}

We begin by examining the 1D Euler equations with reflective wall boundary conditions.  We utilize a full order finite volume model with $2500$ cells on $[-1,1]$ with a CFL of $.75$, and set the artificial viscosity parameter to $\epsilon = 2\times 10^{-4}$.  Entropy stable reflective wall boundary conditions are imposed using ``mirror states'' at both endpoints, and the initial condition is set to be 
\[
\rho = 2 + \frac{1}{2}e^{-100\LRp{x-\frac{1}{2}}^2}, \qquad 
    u = \frac{1}{10}e^{-100\LRp{x-\frac{1}{2}}^2}, \qquad p = \rho^{\gamma}.
\]
The solution produces a shock some time after $T=.25$.

%We first examine the periodic 1D Euler equations.  We utilize a full order finite volume model with $1000$ cells on $[-1,1]$, and utilize .  We set the artificial viscosity parameter to be $\epsilon = 5\times 10^{-4}$, and use the intial condition from \cite{maboudi2018conservative}
%\[
%\rho = \frac{1}{2} + \frac{1}{5}\cos\LRp{2\pi x}, \qquad u = 1.5, \qquad p = \frac{1}{2} + \frac{1}{5}\sin\LRp{2 \pi x}.
%\]
We first examine the impact of enriching solution snapshots with snapshots of the entropy variables.  We run the full order model until final time $T=.7$ and store the solution at $2801$ time-steps.  We compute a reduced basis using SVD by subsampling every $10$ snapshots.  Figure~\ref{subfig:a} shows the decay of the singular values with and without entropy variable enrichment.  In both cases, the decay is similar, with entropy variable enrichment resulting in a slightly slower decay for singular values smaller than $10^{-8}$.  The difference becomes more pronounced for a coarser subsampling.  We note that the enrichment does affect the resulting singular vectors, however.  For example, enrichment results in differences in several computed singular vector (Figure~\ref{subfig:c}).  We also compute projection errors for each solution snapshot using $25$ modes in Figure~\ref{subfig:b}, along with differences between the projected solutions with and without entropy variable enrichment.  The projection errors are nearly indistinguishable.  %For this problem setup, the remaining singular vectors show little difference; however, for other settings, we have observed larger differences with and without enrichment for higher modes.
\begin{figure}
\centering
\subfloat[Singular values]{\raisebox{.24em}{\includegraphics[width=.33\textwidth]{UVsingularvalues.png}}\label{subfig:a}}
\hspace{.02em}
\subfloat[Fifth singular vector]{\raisebox{.615em}{\includegraphics[width=.322\textwidth]{UVsingularvector.png}}\label{subfig:c}}
\hspace{.02em}
\subfloat[Projection errors, 25 modes]{\includegraphics[width=.33\textwidth]{UVprojectionerrors.png}\label{subfig:b}}
\caption{Behavior of the SVD and reduced basis functions with and without entropy variable enrichment of the solution snapshots. }
\label{fig:svd}
\end{figure}

We next examine solutions produced by the reduced order models.  We observe that the reduced models can utilize a larger CFL; however, to ensure that temporal errors are small, all solutions are computed using the same CFL as the full order model.  Figure~\ref{fig:romsols} shows density computed using 25, 75, and 125 modes and the hyper-reduced treatment of viscosity in (\ref{eq:visc2}).  For all resolutions, the shock is under-resolved and the solution possesses Gibbs-type oscillations.  However, despite this under-resolution, the ROM remains stable and does not blow up.  In all cases, we compute hyper-reductions of both convective and viscous terms.  For 25 modes, 92 convective hyper-reduced points and 133 viscous hyper-reduced points are used.  For 75 modes, 245 convective hyper-reduced points and 242 viscous hyper-reduced points are used.  For 125 modes, 410 convective hyper-reduced points and 292 viscous hyper-reduced points are used.  
\begin{figure}
\centering
\subfloat[25 modes, $T=.25$]{\includegraphics[width=.32\textwidth]{euler1Dwall_t25_25modes.png}}
\hspace{.05em}
\subfloat[75 modes, $T=.25$]{\includegraphics[width=.32\textwidth]{euler1Dwall_t25_75modes.png}}
\hspace{.05em}
\subfloat[125 modes, $T=.25$]{\includegraphics[width=.32\textwidth]{euler1Dwall_t25_125modes.png}}\\
\subfloat[25 modes, $T=.25$]{\includegraphics[width=.32\textwidth]{euler1Dwall_t75_25modes.png}}
\hspace{.05em}
\subfloat[75 modes, $T=.25$]{\includegraphics[width=.32\textwidth]{euler1Dwall_t75_75modes.png}}
\hspace{.05em}
\subfloat[125 modes, $T=.25$]{\includegraphics[width=.32\textwidth]{euler1Dwall_t75_125modes.png}}
\caption{Density $\rho$ computed using 25, 75, and 125 modes at times $T=.25, .75$.   }
\label{fig:romsols}
\end{figure}

Next, we examine the difference in the hyper-reduced treatments of entropy dissipation (\ref{eq:visc1}), (\ref{eq:visc2}), and (\ref{eq:visc3}) described in Section~\ref{sec:diss} (recall that (\ref{eq:visc3}) is not provably entropy dissipative).  We have verified that the convective term $\bm{v}_N^T \bm{V}_h^T\LRp{\bm{Q}_h \circ \bm{F}}\bm{1}$ is $O\LRp{10^{-15}}$ (near machine precision) over the entire simulation, confirming that the formulation (\ref{eq:esbchr}) is discretely entropy conservative in the absence of viscosity and boundary flux penalization terms.  Thus, if the discrete entropy dissipation $\bm{v}_N^T\bm{d}(\bm{u}_N)$ is positive, then the simulation is entropy stable.  Figure~\ref{fig:entropydiss}  shows the computed entropy dissipation over the time interval $[0,.75]$.  All hyper-reduced treatments produce similar results, with the entropy dissipation produced by the naive approximation (\ref{eq:visc3}) differing most significantly.  The naive approximation differs from (\ref{eq:visc1}) and (\ref{eq:visc2}) by about $O\LRp{10^{-1}}$ to $O\LRp{10^{-2}}$, while (\ref{eq:visc1}) differs from (\ref{eq:visc2}) by $O\LRp{10^{-6}}$.  
\begin{figure}
\centering
\subfloat[25 modes]{\includegraphics[width=.45\textwidth]{visc25modes.png}}
\hspace{.5em}
%\subfloat[Difference, 25 modes]{\includegraphics[width=.32\textwidth]{viscdiff25modes.png}}
\subfloat[75 modes]{\includegraphics[width=.45\textwidth]{visc75modes.png}}
\caption{Discrete entropy dissipation $\bm{v}_N^T\bm{d}(\bm{u}_N)$ computed using (\ref{eq:visc1}), (\ref{eq:visc2}), and (\ref{eq:visc3}) for 25 and 75 modes.}
\label{fig:entropydiss}
\end{figure}

We now compare the evolution of errors and entropy, and examine the effect of hyper-reduction on the discrete solution in Figure~\ref{fig:entropyevo}.  We observe that the evolution of the ROM average discrete entropy converges to the average discrete entropy for the full order model as the number of modes increases from 25 to 75.  For 125 modes, the average entropies for the ROM and FOM are indistinguishable.  The errors behave similarly, decreasing as the number of modes increases.  For reference, we also include errors computed without hyper-reduction using (\ref{eq:esrom}).  We observe that the hyper-reduced errors are similar to errors without hyper-reduction in all cases, though for $75$ modes, hyper-reduction produces lower errors.  The reasons for this are unclear.
\begin{figure}
\centering
\subfloat[Evolution of entropy]{\includegraphics[height=.24\textheight]{entropyevolution.png}}
\hspace{.5em}
\subfloat[Error between ROM and FOM]{\includegraphics[height=.235\textheight]{euler_error_hypred_comparison_K2500}}
%\hspace{.1em}
%\subfloat[Difference, 25 modes]{\includegraphics[width=.32\textwidth]{viscdiff25modes.png}}
%\subfloat[Entropy 75 modes]{\includegraphics[width=.32\textwidth]{euler1Dwall_entropy_75modes.png}}
\caption{Evolution of error (with and without hyper-reduction) and entropy over time interval $[0,.75]$ for $25, 75, 125$ modes.}
\label{fig:entropyevo}
\end{figure}

%Next, we examine

%\subsection{1D Euler equations with wall boundary conditions}
%\subsection{Accuracy of entropy projection}

%\subsection{2D periodic Euler equations}

\subsection{2D Euler equations}% with wall boundary conditions}

We now study the Euler equations in 2D with wall boundary conditions.  It was shown in \cite{svard2014entropy, chen2017entropy} that wall boundary conditions can be imposed using a ``mirror state''.  Let $\bm{n}$ denote the outward normal at a wall, and let $\bm{u}_{\bm{n}} = un_x + vn_y$ and $\bm{u}_{\bm{n}^\perp} = un_y - vn_x$ denote the normal and tangential components of the velocity at the wall.  We use a boundary numerical flux $\bm{f}^* = \bm{f}_S\LRp{\bm{u}^+,\bm{u}}$ augmented with local Lax-Friedrichs penalization, where the exterior state $\bm{u}^+$ is defined by
\[
\rho^+ = \rho, \qquad \bm{u}_{\bm{n}}^+ = -\bm{u}_{\bm{n}}, \qquad \bm{u}_{\bm{n}^\perp}^+ = \bm{u}_{\bm{n}^\perp}, \qquad p^+ = p.  
\]
We set a Gaussian pulse initial condition 
\[
\rho = 1 + e^{-50\LRp{x^2+(y+1/2)^2}}, \qquad \bm{u} = \bm{0}, \qquad p = \rho^{\gamma}.  
\]
The full order model is computed on a $100\times 100$ grid with viscosity coefficient $\epsilon = 1e-3$, and the reduced order solution is shown in Figure~\ref{fig:pulse2d}. Since little difference is observed between the hyper-reducted treatments of viscosity,  the naive treatment (\ref{eq:visc3}) is used.  The solution does not form shock disccontinuities, and is chosen instead to test the imposition of boundary conditions.  We note that, due to the nature of the linear programming software used to compute the boundary weights, the constraints (\ref{eq:sbpconstraints}) on the boundary weights $\bm{w}_b$ are imposed to a tolerance of $5e-8$.  Despite this inexact enforcement of constraints, the computed entropy RHS remained small, oscillating around $10^{-11}$ in the absence of entropy dissipative terms.

\begin{figure}
\centering
\subfloat[Full order model]{\includegraphics[width=.32\textwidth]{pulse2d.png}}
\hspace{.1em}
\subfloat[Reduced order model]{\includegraphics[width=.32\textwidth]{pulse2d_ROM.png}}
\hspace{.1em}
\subfloat[Hyper-reduction points]{\includegraphics[width=.32\textwidth]{pulse2d_ROM_pts.png}}
\caption{Density computed using 25 modes, 303 hyper-reduced volume points, and 75 hyper-reduced boundary points.  The relative $L^2$ error is $.00934$.}
\label{fig:pulse2d}
\end{figure}

\section{Conclusion}

\section{Acknowledgments}

The author thanks Matthias Heinkenschloss, Matthew Zahr, and Masayuki Yano for informative discussions.

\bibliographystyle{unsrt}
\bibliography{refs}

\end{document}

### 2019-08-08 docs/ESDG_ROM/main.tex
updating paper

@@
 %\documentclass[review,onefignum,onetabnum,final]{siamart171218}
-\documentclass[review]{siamart171218}
-%\documentclass{article}
-
-%\usepackage{fullpage}
-\usepackage{cite}
-%\usepackage{geometry}
-
-\usepackage{amsmath, amsfonts, amssymb}
+%\documentclass[review]{siamart171218}
+\documentclass[preprint,10pt]{elsarticle}
+
+\usepackage{fullpage}
+%\usepackage{cite}
+%\usepackage{geometry}[scale=.8]
+
+\usepackage[colorlinks=true]{hyperref}
+
+\usepackage{amsmath,amssymb,amsfonts,amsthm}
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}
+\theoremstyle{lemma}
+\newtheorem{lemma}{Lemma}
+\newtheorem*{remark}{Remark}
+\theoremstyle{theorem}
+\newtheorem{theorem}{Theorem}
+\theoremstyle{assumption}
+\newtheorem{assumption}{Assumption}
+
 \DeclareMathOperator*{\argmin}{\arg\,\min}
 \DeclareMathOperator*{\argmax}{\arg\,\max}
+\usepackage{algorithm}
 \usepackage{algorithmic}
 
-\newtheorem*{remark}{Remark}
-\theoremstyle{assumption}
-\newtheorem{assumption}{Assumption}
+%\newtheorem*{remark}{Remark}
+%\theoremstyle{assumption}
+%\newtheorem{assumption}{Assumption}
 
@@
 
-\usepackage{hyperref}
 \usepackage[titletoc,toc,title]{appendix}
@@
 \usepackage{pgfplotstable}
-\usepackage{slopetri} % my own package for drawing triangle slopes
+%\usepackage{slopetri} % my own package for drawing triangle slopes
 
@@
 
-\date{}
-\author{Jesse Chan}
+%\date{}
+%\author{Jesse Chan}
+
+\begin{document}
+
+\begin{frontmatter}
 \title{Entropy stable reduced order modeling of nonlinear conservation laws}
 
-\begin{document}
-
-\maketitle
+\author[rice]{Jesse Chan\corref{cor1}}
+\ead{Jesse.Chan@caam.rice.edu}
+\address[rice]{Department of Computational and Applied Mathematics, Rice University, 6100 Main St, Houston, TX, 77005}
 
@@
 Reduced order models of nonlinear conservation laws in fluid dynamics do not typically inherit stability properties of the full order model.
-We discuss the construction of reduced models of nonlinear conservation laws which retain accuracy while inheriting a semi-discrete entropy inequality independently of discretization resolutions and parameters.
+We construct projection-based reduced models of nonlinear conservation laws based on flux differencing which retain accuracy while inheriting a semi-discrete entropy inequality independently of discretization resolutions and parameters.
 \end{abstract}
-
-\tableofcontents
-
-%\begin{itemize}
-%\item Intro
-%\item Entropy conservation for FOM
-%\item Entropy conservation for POD
-%\item Entropy stability via entropy dissipation
-%\item Entropy stable hyper-reduction
-%\item Proof: accuracy of projected operator
-%\item Proof: entropy stability for periodic domain
-%\item Proof: conservation if $\bm{1}$ is contained in $\bm{V}_r$
-%\item Non-periodic domains and constrained hyper-reduction
-
-%\item Numerical experiments
-%\begin{itemize}
-%\item Singular values of conservative snapshots vs conservative + entropy variable snapshots.
-%\item 1D experiments, shocks, error convergence plots
-%\item Effect of different hyper-reduction techniques (optimizing for $V_{\rm test} V_{\rm test}$ and $V_r V_{\rm test}$ vs $V_r V_r$)
-%\item Comparison of entropy stable viscosity vs regular hyper-reduced viscosity
-%\item Comparison of snapshots from entropy stable FOM vs regular FOM.  
-%\item 2D experiments: KH and acoustic pulse
-%\end{itemize}
-%\end{itemize}
+\end{frontmatter}
+
+%\maketitle
+
+%\tableofcontents
 
@@
 Projection-based model reduction constructs low-dimensional surrogate models for many-query scenarios (e.g., simulations over multiple parameter values) that can be evaluated over a range of parameters at a low \textit{online} cost in exchange for a more expensive \textit{offline} pre-computation step \cite{benner2015survey}.
-%Reduced order models (ROMs) can be constructed, for example, by performing a Galerkin projection of the original full order model onto a reduced basis.
 The development of reduced order models (ROMs) is relatively mature for several classes of problems (e.g., linear time-invariant systems, coercive elliptic PDEs).
@@
 For certain PDEs, ``structure-preserving'' ROMs provide robustness and stability by preserving energetic properties of the full system at the discrete level.
-ROMs which retain either Lagrangian \cite{lall2003structure, carlberg2015preserving} or Hamiltonian structure \cite{polyuga2010structure, benner2012interpolation, gugercin2012structure, peng2016symplectic, chaturantabut2016structure, gong2017structure, afkham2017structure, afkham2018structure} have been constructed by combining Galerkin projection with an appropriate formulation of the full order model, and similar energy-conserving ROMs have been constructed for the incompressible Navier-Stokes equations \cite{farhat2014dimensional, farhat2015structure}.
+ROMs which retain either Lagrangian \cite{lall2003structure, carlberg2015preserving} or Hamiltonian structure \cite{benner2012interpolation, gugercin2012structure, peng2016symplectic, chaturantabut2016structure, gong2017structure, afkham2017structure, afkham2018structure} have been constructed by combining Galerkin projection with an appropriate formulation of the full order model, and similar energy-conserving ROMs have been constructed for the incompressible Navier-Stokes equations \cite{farhat2014dimensional, farhat2015structure}.
 The construction of structure-preserving ROMs to nonlinear conservation laws in fluid flow, however, remains an open problem.
@@
 As a result, stabilized discretizations are often employed.
-Petrov-Galerkin ROMs, which use an alternative test basis \cite{maday2002blackbox, rozza2007stability, bui2007goal, amsallem2012stabilization, rozza2013reduced, ballarin2015supremizer, carlberg2017galerkin}, are a popular alternative, as are additional residual-based stabilization or dissipation terms \cite{wang2012proper, kalashnikova2014stabilization, caiazzo2014numerical, balajewicz2016minimal}.
+Petrov-Galerkin ROMs, which use an alternative test basis \cite{maday2002blackbox, rozza2007stability, bui2007goal, serre2012reliable, amsallem2012stabilization, rozza2013reduced, ballarin2015supremizer, carlberg2017galerkin}, are a popular alternative, as are additional residual-based stabilization or dissipation terms \cite{wang2012proper, kalashnikova2014stabilization, caiazzo2014numerical, balajewicz2016minimal}.
 Such models improve robustness in practice, though they do not provide a theoretical foundation for stability.
@@
 The construction of stable ROMs for the compressible Navier-Stokes equations is complicated by their complex structure.
-Thus, practitioners often develop structure-preserving entropy stable ROMs for simplifications of the underlying PDE (such as the time-dependent \textit{linearized} compressible Navier-Stokes equations \cite{barone2009stable, kalashnikova2010stability, serre2012reliable, amsallem2012stabilization, kalashnikova2014construction}) and extrapolate such models to the full nonlinear equations.
+Thus, practitioners often develop structure-preserving entropy stable ROMs for simplifications of the underlying PDE (such as the time-dependent \textit{linearized} compressible Navier-Stokes equations \cite{barone2009stable, kalashnikova2010stability, amsallem2012stabilization, kalashnikova2014construction}) and extrapolate such models to the full nonlinear equations.
 A promising alternative is to enforce physical conditions such as such as kinetic energy preservation, which are empirically related to stability \cite{maboudi2018conservative}.
@@
 In this work, we directly constructing entropy stable ROMs for nonlinear conservation laws through a modified Galerkin projection of an appropriate full order model.
-The approach taken in this paper combines techniques from entropy stable finite volume schemes \cite{tadmor1987numerical, tadmor2003entropy} and entropy stable summation-by-parts (SBP) discretizations \cite{fisher2013high, carpenter2014entropy, chen2017entropy, crean2018entropy, chan2017discretely, chan2018discretely, chan2019skew} to produce entropy stable reduced order models.
+The approach taken in this paper combines techniques from entropy stable finite volume schemes \cite{tadmor1987numerical, tadmor2003entropy} and entropy stable summation-by-parts (SBP) discretizations \cite{fisher2013high, carpenter2014entropy, chen2017entropy, crean2018entropy, chan2017discretely, chan2019skew} to produce entropy stable reduced order models.
+
 The paper is organized as follows.
@@
 \begin{equation}
-\int_{\Omega}\pd{S(\bm{u})}{t}\diff{\bm{x}} + \int_{\partial \Omega} \LRp{\bm{v}^T\bm{f}(\bm{u}) - \psi_i(\bm{u})}n_i \leq 0\label{eq:entropyineq},
+\int_{\Omega}\pd{S(\bm{u})}{t}\diff{\bm{x}} + \sum_{i=1}^d \int_{\partial \Omega} \LRp{\bm{v}^T\bm{f}_i(\bm{u}) - \psi_i(\bm{u})}n_i \leq 0\label{eq:entropyineq},
 \end{equation}
@@
 The entropy inequality (\ref{eq:entropyineq}) is the analogue of energy stability for nonlinear conservation laws \cite{mock1980systems, harten1983symmetric}.  
-However, due to the use of inexact treatment of nonlinear terms (e.g., from collocation approximations, aliasing errors, or inexact quadrature rules), most numerical methods for nonlinear conservation laws do not satisfy a discrete analogue of this stability condition.
+However, due to the use of inexact treatment of nonlinear terms (e.g., from collocation approximations or inexact quadrature), most numerical methods for nonlinear conservation laws do not satisfy a discrete analogue of this stability condition.
 
@@
 
-We first rewrite the system (\ref{eq:nonlineqs1d}) in a matrix form which will be more amenable to model reduction.  We define the skew-symmetric matrix $\bm{Q}$ and flux matrix $\bm{F}$ such that
+We first rewrite the system (\ref{eq:nonlineqs1d}) in a matrix form which will be more amenable to model reduction.
+We define the skew-symmetric differentiation matrix $\bm{Q}$ and flux matrix $\bm{F}$ such that
 \begin{equation}
@@
 \end{equation}
-Note that $\bm{Q}\bm{1} = \bm{0}$, the matrix $\bm{F}$ is symmetric due to flux symmetry, and the diagonal of $\bm{F}$ is equal to $\bm{f}(\bm{u})$ due to flux consistency.
+Note that $\bm{Q}\bm{1} = \bm{0}$, that the matrix $\bm{F}$ is symmetric (due to symmetry of $\bm{f}_S$), and that the diagonal of $\bm{F}$ is equal to $\bm{f}(\bm{u})$ due to flux consistency.
 For a scalar nonlinear conservation law, the matrix-based formulation of (\ref{eq:nonlineqs1d}) is then equivalent to
@@
 \]
-Assuming the chain rule in time, 
-\begin{align*}
-\bm{v}_h^T\td{\bm{u}_h}{t} &= {\td{S(\bm{u}_h)}{\bm{u}}}^T\td{\bm{u}_h}{t} = \sum_j \LRp{\td{S(\bm{u}_h)}{\bm{u}}}_j^T\td{\LRp{\bm{u}_h}_j}{t} \\
-&=  \sum_j \td{S(\LRp{\bm{u}_h}_j)}{t}  =  \bm{1}^T\td{S(\bm{u}_h)}{t}.
-\end{align*}
-Using skew-symmetry of $\bm{Q}$, the flux term yields
-\begin{align*}
-\bm{v}_h^T2\LRp{{\bm{Q}}\circ \bm{F}}\bm{1} &= \sum_{ij} \LRp{\bm{v}_h}_i^T \bm{Q}_{ij} \bm{f}_{S}\LRp{\LRp{\bm{u}_h}_i, \LRp{\bm{u}_h}_j} \\
-&= \sum_{ij} \LRp{\bm{Q}_{ij}-\bm{Q}_{ji}} \LRp{\bm{v}_h}_i^T  \bm{f}_{S}\LRp{\LRp{\bm{u}_h}_i, \LRp{\bm{u}_h}_j}.
+Assuming continuity in time and using the definition of the entropy variables in (\ref{eq:nonlineqs}), $\bm{v}_h^T\td{\bm{u}_h}{t}$ simplifies to
+\begin{align*}
+{\td{S(\bm{u}_h)}{\bm{u}}}^T\td{\bm{u}_h}{t} = \sum_j \LRp{\td{S(\bm{u}_h)}{\bm{u}}}_j^T\td{\LRp{\bm{u}_h}_j}{t} =  \sum_j \td{S(\LRp{\bm{u}_h}_j)}{t}  =  \bm{1}^T\td{S(\bm{u}_h)}{t}.
+\end{align*}
+Using skew-symmetry of $\bm{Q}$, the flux term $\bm{v}_h^T2\LRp{{\bm{Q}}\circ \bm{F}}\bm{1}$ yields
+\begin{align*}
+\sum_{ij} \LRp{\bm{v}_h}_i^T \bm{Q}_{ij} \bm{f}_{S}\LRp{\LRp{\bm{u}_h}_i, \LRp{\bm{u}_h}_j} = \sum_{ij} \LRp{\bm{Q}_{ij}-\bm{Q}_{ji}} \LRp{\bm{v}_h}_i^T  \bm{f}_{S}\LRp{\LRp{\bm{u}_h}_i, \LRp{\bm{u}_h}_j}.
 \end{align*}
@@
 \subsection{Viscosity and entropy dissipation} 
-
+\label{sec:entropydissipation}
 Theorem~\ref{thm:ecfom} shows that the formulation (\ref{eq:essys}) preserves a semi-discrete conservation of entropy.
@@
 \begin{align*}
-\bm{V}^T\bm{V}\td{\bm{u}_N}{t} + 2\bm{V}^T\LRp{\bm{Q}\circ \bm{F}}\bm{1} = 0.
+\Delta x\bm{V}^T\bm{V}\td{\bm{u}_N}{t} + 2\bm{V}^T\LRp{\bm{Q}\circ \bm{F}}\bm{1} = 0.
 \end{align*}
@@
 \begin{gather}
-\bm{V}^T\bm{V}\td{\bm{u}_N}{t} + 2\bm{V}^T\LRp{\bm{Q}\circ \bm{F}}\bm{1} = 0, \label{eq:esrom}\\
+\Delta x\bm{V}^T\bm{V}\td{\bm{u}_N}{t} + 2\bm{V}^T\LRp{\bm{Q}\circ \bm{F}}\bm{1} = \bm{0}, \label{eq:esrom}\\
 \LRp{\bm{F}}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad \tilde{\bm{u}} = \bm{u}\LRp{\bm{V}\bm{V}^{\dagger}\bm{v}\LRp{\bm{V}\bm{u}_N}}.\nonumber
@@
 \]
+Additionally, if the reduced basis $\bm{V}$ exactly represents $\bm{1}$, solutions of (\ref{eq:esrom}) conserve global averages of the conservative variables.
 \label{thm:esrom}
@@
 \begin{proof}
-The result follows from testing with $\bm{v}_N$ and applying (\ref{eq:esromtime}).  The remaining steps are identical to those of Theorem~\ref{thm:ecfom}.  
+The conservation of entropy follows from testing with $\bm{v}_N$ and applying (\ref{eq:esromtime}).
+The remaining steps are identical to those of Theorem~\ref{thm:ecfom}.
+The global conservation results from testing with $\bm{1}$.
+Let $\bm{V}\bm{e}=\bm{1}$, where $\bm{e}$ denote coefficients of $\bm{1}$ in the reduced basis matrix.
+Then,
+\begin{align*}
+\Delta x\bm{e}^T\bm{V}^T\bm{V}\td{\bm{u}_N}{t} + 2\bm{e}^T\bm{V}^T\LRp{\bm{Q}\circ \bm{F}}\bm{1}  = \Delta x\bm{1}^T\td{\LRp{\bm{V}\bm{u}_N}}{t} + 2\bm{1}^T\LRp{\bm{Q}\circ \bm{F}}\bm{1} = 0.
+\end{align*}
+Since $\bm{Q}$ is skew-symmetric and $\bm{F}$ is symmetric, $\bm{Q} \circ \bm{F}$ is skew-symmetric and $\bm{1}^T\LRp{\bm{Q}\circ \bm{F}}\bm{1} = \bm{0}$, 
+\[
+\Delta x\bm{1}^T\td{\LRp{\bm{V}\bm{u}_N}}{t} = 0,
+\]
+which is a discrete statement of conservation.
 \end{proof}
@@
 
-\subsubsection{Choices of test basis}
+\subsubsection{Choice of test basis}
 
 We have not yet specified the choice of test basis $\bm{V}_t$.
-Common sense suggests that the span of the test basis should include the reduced basis $\bm{V}$.
-Theorem~\ref{thm:Qt} implies that the span of the test basis should also contain $\bm{1}$.
+The span of the test basis should include the reduced basis $\bm{V}$, and Lemma~\ref{thm:Qt} implies that the span of the test basis should also contain $\bm{1}$.
 However, these two conditions are insufficient to ensure accuracy for solutions of nonlinear conservation laws.
@@
 \centering
-\subfloat[Solution snapshots]{\includegraphics[width=.3\textwidth]{figs/burg_snaps.png}}
-\hspace{.5em}
-\subfloat[Modes (columns of $\bm{V}$)]{\includegraphics[width=.3\textwidth]{figs/burg3modes.png}}
-\hspace{.5em}
-\subfloat[Mode derivatives $\bm{Q}\bm{V}$]{\includegraphics[width=.3\textwidth]{figs/burg3modes_diff.png}}
+\subfloat[Solution snapshots]{\includegraphics[width=.32\textwidth]{figs/burg_snaps.png}}
+\hspace{.1em}
+\subfloat[Modes (columns of $\bm{V}$)]{\includegraphics[width=.32\textwidth]{figs/burg3modes.png}}
+\hspace{.1em}
+\subfloat[Mode derivatives $\bm{Q}\bm{V}$]{\includegraphics[width=.32\textwidth]{figs/burg3modes_diff.png}}
 \caption{Solution snapshots, three POD modes, and mode derivatives for a shock solution of the viscous periodic Burgers' equation. }
@@
 
-We illustrate why this choice of test basis restores accuracy.
-The entropy conservative flux for Burgers' equation corresponds to the discretization of the split form \cite{canuto2006spectral, chan2017discretely, maboudi2018conservative}
-\[
-\pd{u}{t} + \frac{1}{3}\LRp{\pd{u^2}{x}  + u\pd{u}{x}} = 0.
-\]
-An entropy conservative ROM without hyper-reduction is then given by 
-\[
-\bm{V}^T\bm{V}\td{u}{t} + \frac{1}{3}\bm{V}^T\LRp{\bm{Q}_t\bm{u}^2  + {\rm diag}\LRp{\bm{u}}\bm{Q}_t\bm{u}} = \bm{0}, \qquad \bm{u} = \bm{V}\bm{u}_N.
-\]
-It was shown in \cite{chan2017discretely, chenreview} that, for certain entropy conservative fluxes, the flux differencing procedure is equivalent in discretization of the split form.
-Thus, to ensure accurate evalution of the spatial formulation, the reduced matrix $\bm{Q}_t$ should accurately approximate mixed derivatives of the form ${\rm diag}(\bm{f})\bm{Q}\bm{g}$.
-We quantify the accuracy of $\bm{Q}_t$ in approximating such terms with the following theorem:
-\begin{theorem}
-Let $\bm{f},\bm{g}$ denote vectors, and let $\bm{u}_1, \bm{u}_2$ denote solutions of the following problems
-\begin{align*}
-\bm{V}^T\bm{V}\bm{u}_1 &= \bm{V}^T \diag{\bm{f}}\bm{Q}\bm{g}\\
-\bm{V}^T\bm{V}\bm{u}_2 &= \bm{V}^T \diag{\bm{f}}\bm{Q}_t\bm{g}.
-\end{align*}
-Then, the error $\bm{e} = \bm{u}_1-\bm{u}_2$ satisfies
-\[
-\nor{\bm{V}\bm{e}} \leq \nor{\bm{Q}} \LRp{\nor{\bm{f}-\bm{V}\bm{P}\bm{f}} \nor{\bm{g}} + \nor{\bm{f}}\nor{\bm{g}-\bm{V}\bm{P}\bm{g}}}.
-\]
+We comment on why this choice of test basis restores accuracy.  
+In the absence of hyper-reduction (where $\bm{P}_t = \bm{V}_t^{\dagger}$), the matrix $\bm{Q}_t$ should be an accurate approximation to $\bm{Q}$.
+The following lemma quantifies the relationship between accuracy and choice of test space.
+%The entropy conservative flux for Burgers' equation corresponds to the discretization of the split form \cite{canuto2006spectral, chan2017discretely, maboudi2018conservative}
+%\[
+%\pd{u}{t} + \frac{1}{3}\LRp{\pd{u^2}{x}  + u\pd{u}{x}} = 0.
+%\]
+%An entropy conservative ROM without hyper-reduction is then given by 
+%\[
+%\Delta x\bm{V}^T\bm{V}\td{u}{t} + \frac{1}{3}\bm{V}^T\LRp{\bm{Q}\bm{u}^2  + {\rm diag}\LRp{\bm{u}}\bm{Q}\bm{u}} = \bm{0}, \qquad \bm{u} = \bm{V}\bm{u}_N.
+%\]
+%It was shown in \cite{chan2017discretely, chenreview} that, for certain entropy conservative fluxes, the flux differencing procedure is equivalent in discretization of the split form.
+Thus, to ensure accurate evalution of the spatial formulation, the reduced differentiation matrix should at least accurately approximate mixed derivatives of the form ${\rm diag}(\bm{f})\bm{Q}\bm{g}$.
+We assume no hyper-reduction is performed, such that $\bm{P}_t = \bm{V}_t^{\dagger}$.
+\begin{lemma}
 \label{thm:acc}
-\end{theorem}
+Let $\bm{f}$ denote a vector.  Then, if $\mathcal{R}\LRp{\bm{Q}\bm{V}} \subset \mathcal{R}\LRp{\bm{V}_t}$, then
+\begin{align*}
+\nor{\bm{Q}\bm{f} - \LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{f}} = \nor{\bm{Q}\LRp{\bm{f}-\bm{V}_t\bm{V}_t^{\dagger}\bm{f}}} \leq \nor{\bm{Q}}\nor{\bm{f}-\bm{V}_t\bm{V}_t^{\dagger}\bm{f}}.
+\end{align*}
+%Let $\bm{f},\bm{g}$ denote vectors, and let $\bm{u}_1, \bm{u}_2$ denote solutions of the following problems
+%\begin{align*}
+%\bm{V}^T\bm{V}\bm{u}_1 &= \bm{V}^T \diag{\bm{f}}\bm{Q}\bm{g}\\
+%\bm{V}^T\bm{V}\bm{u}_2 &= \bm{V}^T \diag{\bm{f}} \LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{g}.
+%\end{align*}
+%Let $\bm{f}_{\bm{v}} = {\rm diag}\LRp{\bm{f}}\bm{v}$.  Then, the error $\bm{e} = \bm{u}_1-\bm{u}_2$ satisfies
+%\[
+%\nor{\bm{V}\bm{e}} \leq \nor{\bm{Q}} \nor{\bm{V}^{\dagger}}\LRp{\nor{\bm{f}}_{\infty} + \nor{\bm{g}}} \LRp{\nor{\bm{g}-\bm{V}_t\bm{V}_t^{\dagger}\bm{g}} + \max_{\bm{v}\in \mathcal{R}(\bm{V})} \nor{\bm{f}_{\bm{v}} -\bm{V}_t\bm{V}_t^\dagger \bm{f}_{\bm{v}}} }.
+%\]
+\end{lemma}
 \begin{proof}
-Since $\bm{Q}_t = \bm{P}_t^T\LRp{\bm{V}_t^T\bm{Q}\bm{V}_t}\bm{P}_t$, we can derive the estimate by adding and subtracting $\bm{P}_t^T\bm{V}_t^T \diag{\bm{f}}\bm{Q}\bm{g}$.
-Straightforward manipulations give the following
-\begin{align*}
-\bm{V}^T\diag{\bm{f}}\LRp{\bm{Q}\bm{g} - \bm{Q}_t\bm{g}} =  
-\bm{V}^T\diag{\bm{f}} \LRp{\bm{Q}\LRp{\bm{I}-\bm{V}_t\bm{P}_t}\bm{g} + \LRp{\bm{I}-\bm{V}_t\bm{P}_t}^T\bm{Q}\bm{g}}
-\end{align*}
- \note{Add proof}
+Define $\bm{e} = \bm{Q}\bm{f} - \LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{f}$.  Adding and subtracting $\bm{Q}\LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{f}$ yields
+\[
+\bm{e} = \bm{Q}\LRp{\bm{f} - \LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{f}} + \underbrace{\bm{Q}{\bm{V}_t\bm{V}_t^{\dagger}}\bm{f} - \LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{f}}_{\bm{e}_2}.
+\]
+Since $\mathcal{R}\LRp{\bm{Q}\bm{V}} \subset \mathcal{R}\LRp{\bm{V}_t}$, $\bm{e}_2\in \mathcal{R}\LRp{\bm{V}_t}$ and $\bm{V}_t\bm{V}_t^\dagger\bm{e}_2 = \bm{e}_2$.
+Thus, $\bm{e}_2^T\bm{e}_2 = \nor{\bm{e}_2}^2$ gives
+\[
+\nor{\bm{e}_2}^2 = \bm{e}_2^T\bm{Q}{\bm{V}_t\bm{V}_t^{\dagger}}\bm{f} - \bm{e}_2^T\LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{f} 
+= \bm{e}_2^T\bm{Q}{\bm{V}_t\bm{V}_t^{\dagger}}\bm{f} - \bm{e}_2^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{f} = \bm{0}.
+\]
+%We can derive the estimate by adding and subtracting $\diag{\bm{f}}\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{g}$.  Straightforward manipulations then give 
+%\begin{align*}
+%\bm{V}^T\bm{V}\bm{e} = \LRp{\bm{V}^T\diag{\bm{f}}\bm{Q}\LRp{\bm{g}-\bm{V}_t\bm{V}_t^{\dagger}\bm{g}} + \LRp{\LRp{\bm{I} - \bm{V}_t\bm{V}_t^{\dagger}}\diag{\bm{f}}\bm{V}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{g}}
+%\end{align*}
+%Multiplying by $\bm{e}^T$ on both sides and using the triangle inequality yields
+%\begin{align*}
+%\nor{\bm{V}\bm{e}}^2 &\leq 
+%\nor{\bm{e}^T\bm{V}^T\diag{\bm{f}}\bm{Q}\LRp{\bm{g}-\bm{V}_t\bm{V}_t^{\dagger}\bm{g}}} \\
+%&+ \nor{\LRp{\LRp{\bm{I} - \bm{V}_t\bm{V}_t^{\dagger}}\diag{\bm{f}}\bm{V}\bm{e}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{g}}.
+%\end{align*}
+%The first term is bounded by Cauchy-Schwarz and the fact that $\nor{\diag{\bm{f}}} = \nor{\bm{f}}_{\infty}$
+%\[
+%\nor{\bm{e}^T\bm{V}^T\diag{\bm{f}}\bm{Q}\LRp{\bm{g}-\bm{V}_t\bm{V}_t^{\dagger}\bm{g}}}\leq \nor{\bm{V}\bm{e}} \nor{\bm{f}}_{\infty} \nor{\bm{Q}} \nor{\bm{g}-\bm{V}_t\bm{V}_t^{\dagger}\bm{g}}.
+%\]
+%The second term is bounded similarly.  Using that $\bm{V}^{\dagger}\bm{V} = \bm{I}$, 
+%\begin{align*}
+%&\nor{\LRp{\LRp{\bm{I} - \bm{V}_t\bm{V}_t^{\dagger}}\diag{\bm{f}}\bm{V}\bm{e}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{g}}  \\
+%&\leq\nor{\LRp{\LRp{\bm{I} - \bm{V}_t\bm{V}_t^{\dagger}}\diag{\bm{f}}\bm{V}}} \nor{\bm{V}^{\dagger}}\nor{\bm{V}\bm{e}}\nor{\bm{Q}}\nor{\bm{V}_t \bm{V}_t^{\dagger}}\nor{\bm{g}} \\
+%&\leq \max_{\bm{v}\in \mathcal{R}(\bm{V})} \nor{\bm{f}_{\bm{v}} -\bm{V}_t\bm{V}_t^\dagger \bm{f}_{\bm{v}}}\nor{\bm{V}^{\dagger}} \nor{\bm{V}\bm{e}}\nor{\bm{Q}}\nor{\bm{g}} 
+%\end{align*}
+%where we have used that $\nor{\bm{V}^{\dagger}} \leq 1$ since $\bm{V}$ is orthonormal and $\nor{\bm{V}_t \bm{V}_t^{\dagger}} \leq 1$ since the normed quantity is an orthogonal projection.
+Dividing through by $\nor{\bm{V}\bm{e}}$ and an $\ell_2$ inequality complete the proof.
+We note that if $\bm{V}$ is orthonormal, then $\nor{\bm{V}^{\dagger}}\leq 1$.
  \end{proof}
 
-Theorem~\ref{thm:acc} does not account for hyper-reduction.
+%Lemma~\ref{thm:acc} does not account for hyper-reduction; however, it motivates what quantities hyper-reduction should aim to approximate accurately.
 
@@
 In this work, we utilize the greedy algorithm for computing empirical cubature points and weights from \cite{an2008optimizing, hernandez2017dimensional}, which is described in Algorithm~\ref{alg:hr}.
-Let $\bm{V}_{\rm target}$ be some matrix whose columns span some target space of functions to be integrated.
-The algorithm approximates $\bm{V}_{\rm target}^T\bm{1}$ by $\bm{V}_{\rm target}\LRp{\mathcal{I},:}^T\bm{w}$.
+Let $\bm{V}_{\rm target}$ be some matrix whose columns span some target space of functions to be integrated, and let $\bm{w}_{\rm target}$ denote some reference target weights.
+The algorithm approximates $\bm{V}_{\rm target}^T\bm{w}_{\rm target}$ by $\bm{V}_{\rm target}\LRp{\mathcal{I},:}^T\bm{w}$.
 The index set is constructed in a greedy fashion by selecting the row index of $\bm{V}_{\rm target}$ which is most positively parallel to the residual, then computing the weight vector which minimizes the residual through a (non-negative) least squares solve.
 
-\begin{algorithm}
+\begin{algorithm}[!h]
 \caption{Compute hyper-reduction s.t.\ $\bm{V}_{\rm target}^T\bm{w}_{\rm target} \approx \bm{V}_{\rm target}\LRp{\mathcal{I},:}^T\bm{w}$.}
@@
         \]
-        \STATE Set $\mathcal{I} = \mathcal{I} \cup \LRc{i}$.  
+        \STATE Add the new index to the set of hyper-reduced points $\mathcal{I} = \mathcal{I} \cup \LRc{i}$.  
         \STATE Compute $\bm{w}$ using linear least squares
         \[
-        \bm{w} = \argmin_{\bm{c}} \nor{\bm{V}_{\rm target}\LRp{\mathcal{I},:}^T\bm{c} - \bm{b}}.
+        \bm{w} = \argmin_{\bm{c}} \frac{1}{2}\nor{\bm{V}_{\rm target}\LRp{\mathcal{I},:}^T\bm{c} - \bm{b}}^2.
         \]
@@
                 \[
-                \argmin_{\bm{c} \geq 0} \nor{\bm{V}_{\rm target}\LRp{\mathcal{I},:}^T\bm{c} - \bm{b}}
+                \argmin_{\bm{c} \geq 0} \frac{1}{2}\nor{\bm{V}_{\rm target}\LRp{\mathcal{I},:}^T\bm{c} - \bm{b}}^2.
                 \]             
@@
 We note that the flux matrix $\bm{F}$ is not computed explicitly, but instead is usually formed on the fly.
-The computation of $\bm{F}$ constitutes a significant portion of the computational cost, but can be reduced for the full order model by exploiting the sparsity of $\bm{Q}$.
+The computation of $\bm{F}$ constitutes a significant portion of the computational cost, and is typically ameliorated by exploiting the sparsity of $\bm{Q}$ for the full order model.
 However, since the hyper-reduced matrix $\bm{Q}_t$ is fully dense, the ROM (\ref{eq:hrrom}) requires $O(N_s^2)$ nonlinear flux evaluations, in contrast to the typical $O(N_s)$ nonlinear function evaluations required for a typical hyper-reduced model.
@@
 
-For linear problems, it is possible to impose energy stable boundary conditions by combining generalized SBP operators with appropriate simultaneous approximation terms (SATs) \cite{fernandez2014review, fernandez2018simultaneous}.
+For linear problems, it is possible to impose energy stable boundary conditions by combining generalized SBP operators with appropriate simultaneous approximation terms (SATs) \cite{fernandez2018simultaneous}.
 However, due to the presence of $\bm{E}$ within the boundary term $\bm{E}^T\bm{B}\bm{E}$, the accurate and entropy stable imposition of nonlinear boundary conditions using generalized SBP SATs remains an open problem \cite{crean2018entropy, chan2018efficient, chenreview}.
@@
 \begin{theorem}
-If the boundary flux $\bm{f}^*$ is entropy stable in the sense of (\ref{eq:esbflux}), then the following hyper-reduced ROM 
+If the boundary flux $\bm{f}^*$ is entropy stable in the sense of (\ref{eq:esbflux}), then the hyper-reduced ROM 
 \begin{gather}
@@
 \]
-If the flux is entropy stable, then $\td{S\LRp{\bm{V}\bm{u}_N}}{t} \leq 0$, with equality holding if the flux is entropy conservative.
+If the flux is entropy stable, then $\td{S\LRp{\bm{V}\bm{u}_N}}{t} \leq 0$.  Equality holds if the flux is entropy conservative such that $\psi(\tilde{\bm{u}}_b)=\tilde{\bm{v}}_b^T\bm{f}^*$.
 \end{proof}
@@
 Note that, unlike the 1D case, the test bases differ along each coordinate direction.  However, the dimension of the range of each test basis $\bm{V}_t^i$ is still at most $2N+1$.  
-We can now define hyper-reduced projection matrices
+We can now define hyper-reduced projection matrices for the $i$th coordinate direction
 \begin{align*}
@@
 \]
-which are used to construct an entropy stable reduced order model in $d$ dimensions
+which are used to construct an entropy conservative reduced order model in $d$ dimensions
 \begin{gather}
@@
 Non-periodic domains and boundary conditions can be treated by combining hybridized SBP operators and a weak imposition of boundary conditions as was done in the 1D hyper-reduced formulation (\ref{eq:esbchr}).
-Without loss of generality, we assume a square domain with $4$ boundary faces, each of which is discretized by $K$ intervals of size $\Delta x$.
-We assume we are given values of the outward normal vector $\bm{n}$ on the domain boundary
-\[
-\bm{n} = \LRs{n_1, \ldots, n_d}^T.
-\]
-We can enforce non-periodic boundary conditions through introducing numerical fluxes at boundary points.
-However, the number of boundary points scales with $K$, the number of elements in the full order model along one coordinate direction.
-To ensure that the cost of the reduced order model scales with the number of modes $N$ rather than the dimension of the full order model, we must apply an additional hyper-reduction to approximate boundary terms by a weighted combination of sub-sampled boundary points.
+Without loss of generality, we assume a square domain with $4$ boundary faces, each of which is discretized by $K_{\rm 1D}$ intervals of size $\Delta x$.
+We assume we are given values of the outward normal vector $\bm{n} = \LRs{n_1, \ldots, n_d}^T$ on the domain boundary.
+We enforce non-periodic boundary conditions through introducing numerical fluxes at boundary points.
+However, the number of boundary points scales with $K_{\rm 1D}$.
+To ensure that the cost of the reduced order model scales with the number of modes $N$ rather than the dimension of the full order model, we must approximate boundary terms by a hyper-reduced weighted combination of sub-sampled boundary points.
 
@@
 
-\subsection{Entropy dissipation}
-
-\note{Discuss how to produce entropy dissipative hyper-reduced operators in 2D}.
+%\subsection{Entropy dissipation}
+%\note{Discuss how to produce entropy dissipative hyper-reduced operators in 2D}.
 
@@
 \begin{gather*}
-v_1 = \frac{\rho e (\gamma + 1 - s) - E}{\rho e}, \qquad v_{d+2} = -\frac{\rho}{\rho e},\\
+v_1 = \frac{\rho e (\gamma + 1 - s) - E}{\rho e}, \qquad v_{d+2} = -\frac{\rho}{\rho e}, \qquad
 v_{1+ i}= \frac{\rho {{u}_i}}{\rho e}, \qquad i = 1,\ldots, d.
@@
 \begin{gather*}
-\rho = -(\rho e) v_{d+2}, \qquad E = (\rho e)\LRp{1 - \frac{\sum_{j=1}^d{v_{1+j}^2}}{2 v_{d+2}}}\\
+\rho = -(\rho e) v_{d+2}, \qquad E = (\rho e)\LRp{1 - \frac{\sum_{j=1}^d{v_{1+j}^2}}{2 v_{d+2}}}, \qquad
  \rho {u_i} = (\rho e) v_{1+i}, \qquad i = 1,\ldots,d,
@@
 \end{equation*}
-{Let $f_i, f_j$ denote two arbitrary values.  We define the average and logarithmic averages as follows:}
+{Let $f_i, f_j$ denote two arbitrary values.  We define the average and logarithmic average}
 \[
@@
 To ensure numerical stability when the denominator is close to zero, we evaluate the logarithmic average using the algorithm of \cite{ismail2009affordable}.  
-Explicit expressions for entropy conservative numerical fluxes in two dimensions  are given by Chandrashekar \cite{chandrashekar2013kinetic}
-\begin{align*}
-&f^1_{1,S}(\bm{u}_L,\bm{u}_R) = \avg{\rho}^{\log} \avg{u_1},& &f^1_{2,S}(\bm{u}_L,\bm{u}_R) = \avg{\rho}^{\log} \avg{u_2},&\\
-&f^2_{1,S}(\bm{u}_L,\bm{u}_R) = f^1_{1,S} \avg{u_1} + p_{\rm avg},&  &f^2_{2,S}(\bm{u}_L,\bm{u}_R) = f^1_{2,S} \avg{u_1},&\nonumber\\
-&f^3_{1,S}(\bm{u}_L,\bm{u}_R) = f^2_{2,S},& &f^3_{2,S}(\bm{u}_L,\bm{u}_R) = f^1_{2,S} \avg{u_2} + p_{\rm avg},&\nonumber\\
-&f^4_{1,S}(\bm{u}_L,\bm{u}_R) = \LRp{E_{\rm avg} + p_{\rm avg}}\avg{u_1},& &f^4_{2,S}(\bm{u}_L,\bm{u}_R) = \LRp{E_{\rm avg} + p_{\rm avg} }\avg{u_2},& \nonumber
-\end{align*}
-where we have introduced the auxiliary quantities 
-\begin{gather}
-{\beta = \frac{\rho}{2p}}, \qquad p_{\rm avg} = \frac{\avg{\rho}}{2\avg{\beta}}, \qquad \qquad E_{\rm avg} = \frac{\avg{\rho}^{\log}}{2\avg{\beta}^{\log}\LRp{\gamma -1}}   + \frac{{u_{\rm avg}^2}}{2}, \label{eq:fluxaux} \\
-{u^2_{\rm avg}} = 2(\avg{u_1}^2 + \avg{u_1}^2) - \LRp{\avg{u_1^2} +\avg{u_2^2}} \nonumber.  
-\end{gather}
-
-\note{Add equations for $\pd{\bm{u}}{\bm{v}}$. }
+Explicit expressions for entropy conservative fluxes are given by Chandrashekar \cite{chandrashekar2013kinetic}.  In 1D, these fluxes are
+\begin{align*}
+f^1_S(\bm{u}_L,\bm{u}_R) &= \avg{\rho}^{\log} \avg{u}\\
+f^2_S(\bm{u}_L,\bm{u}_R) &= f^1_S(\bm{u}_L,\bm{u}_R)\avg{u} + p_{\rm avg}\\
+f^3_S(\bm{u}_L,\bm{u}_R) &= \LRp{E_{\rm avg} + p_{\rm avg}}\avg{u},
+\end{align*}
+where the auxiliary quantities are
+\begin{gather*}
+{\beta = \frac{\rho}{2p}}, \qquad p_{\rm avg} = \frac{\avg{\rho}}{2\avg{\beta}}, \qquad E_{\rm avg} = \frac{\avg{\rho}^{\log}}{2\avg{\beta}^{\log}\LRp{\gamma -1}}   + \avg{\rho}^{\log}\frac{{u_{\rm avg}^2}}{2}.
+\end{gather*}
+In 1D, $u_{\rm avg}^2 = 2\avg{u}^2-\avg{u^2}$.  
+In two dimensions, the $x$ and $y$ fluxes are given by
+\begin{align*}
+&f^1_{x,S}(\bm{u}_L,\bm{u}_R) = \avg{\rho}^{\log} \avg{u_1},& &f^1_{y,S}(\bm{u}_L,\bm{u}_R) = \avg{\rho}^{\log} \avg{u_2},&\\
+&f^2_{x,S}(\bm{u}_L,\bm{u}_R) = f^1_{1,S} \avg{u_1} + p_{\rm avg},&  &f^2_{y,S}(\bm{u}_L,\bm{u}_R) = f^1_{2,S} \avg{u_1},&\nonumber\\
+&f^3_{x,S}(\bm{u}_L,\bm{u}_R) = f^2_{2,S},& &f^3_{y,S}(\bm{u}_L,\bm{u}_R) = f^1_{y,S} \avg{u_2} + p_{\rm avg},&\nonumber\\
+&f^4_{x,S}(\bm{u}_L,\bm{u}_R) = \LRp{E_{\rm avg} + p_{\rm avg}}\avg{u_1},& &f^4_{y,S}(\bm{u}_L,\bm{u}_R) = \LRp{E_{\rm avg} + p_{\rm avg} }\avg{u_2},& \nonumber
+\end{align*}
+where we have redefined ${u^2_{\rm avg}} = 2(\avg{u_1}^2 + \avg{u_2}^2) - \LRp{\avg{u_1^2} +\avg{u_2^2}}.$
+
+The viscous terms are treated as described in Section~\ref{sec:entropydissipation}, and utilize evaluations of the Jacobian $\pd{\bm{u}}{\bm{v}}$.
+Explicit expressions for Jacobians in the compressible Euler and Navier-Stokes equations are given in \cite{hughes1986new}.
+%\note{Add equations for $\pd{\bm{u}}{\bm{v}}$. }
 
@@
 \hspace{.02em}
-\subfloat[Fifth singular vector]{\raisebox{.615em}{\includegraphics[width=.322\textwidth]{UVsingularvector.png}}\label{subfig:c}}
+\subfloat[Fifth singular vector]{\raisebox{.6em}{\includegraphics[width=.322\textwidth]{UVsingularvector.png}}\label{subfig:c}}
 \hspace{.02em}
-\subfloat[Projection errors, 25 modes]{\includegraphics[width=.33\textwidth]{UVprojectionerrors.png}\label{subfig:b}}
+\subfloat[Projection errors, 25 modes]{\raisebox{.05em}{\includegraphics[width=.33\textwidth]{UVprojectionerrors.png}\label{subfig:b}}}
 \caption{Behavior of the SVD and reduced basis functions with and without entropy variable enrichment of the solution snapshots. }
@@
 For 125 modes, 410 convective hyper-reduced points and 292 viscous hyper-reduced points are used.
-\begin{figure}
+
+\begin{figure}[!h]
 \centering
@@
 \subfloat[125 modes, $T=.25$]{\includegraphics[width=.32\textwidth]{euler1Dwall_t25_125modes.png}}\\
-\subfloat[25 modes, $T=.25$]{\includegraphics[width=.32\textwidth]{euler1Dwall_t75_25modes.png}}
+\subfloat[25 modes, $T=.75$]{\includegraphics[width=.32\textwidth]{euler1Dwall_t75_25modes.png}}
 \hspace{.05em}
-\subfloat[75 modes, $T=.25$]{\includegraphics[width=.32\textwidth]{euler1Dwall_t75_75modes.png}}
+\subfloat[75 modes, $T=.75$]{\includegraphics[width=.32\textwidth]{euler1Dwall_t75_75modes.png}}
 \hspace{.05em}
-\subfloat[125 modes, $T=.25$]{\includegraphics[width=.32\textwidth]{euler1Dwall_t75_125modes.png}}
-\caption{Density $\rho$ computed using 25, 75, and 125 modes at times $T=.25, .75$.   }
+\subfloat[125 modes, $T=.75$]{\includegraphics[width=.32\textwidth]{euler1Dwall_t75_125modes.png}}
+\caption{Density $\rho$ computed using 25, 75, and 125 modes at times $T=.25, .75$.
+The solution of the reduced order model is plotted at convective hyper-reduced points, shown in \textcolor{red}{red}.
+The viscous points are not shown.}
 \label{fig:romsols}
@@
 
+Next, we verify that (in the absence of viscosity), the ROM is semi-discretely entropy conservative.
+The proof of this property uses that the entropy contribution of the convective term $\bm{v}_N^T \bm{V}_h^T\LRp{\bm{Q}_h \circ \bm{F}}\bm{1} = 0$.
+Figure~\ref{fig:novisc} verifies this, plotting the absolute value of the entropy convective term over time.
+At all time-steps, $\LRb{\bm{v}_N^T \bm{V}_h^T\LRp{\bm{Q}_h \circ \bm{F}}\bm{1}}$ is near $O\LRp{10^{-14}}$ (close to machine precision), confirming that the formulation (\ref{eq:esbchr}) is discretely entropy conservative in the absence of viscosity and boundary flux penalization terms.
+We also run the ROM with the viscosity parameter $\epsilon = 0$.
+For 75 modes, the solution is more oscillatory than
+
+\begin{figure}[!h]
+\centering
+\subfloat[Entropy contribution $\LRb{\bm{v}_N^T \bm{V}_h^T\LRp{\bm{Q}_h \circ \bm{F}}\bm{1}}$]{\includegraphics[width=.4525\textwidth]{entropyrhs.png}}
+\hspace{2em}
+%\subfloat[Difference, 25 modes]{\includegraphics[width=.32\textwidth]{viscdiff25modes.png}}
+\subfloat[Density $\rho$ with 75 modes and no viscosity]{\includegraphics[width=.44\textwidth]{euler1Dwall_t75_75modes_novisc.png}}
+\caption{Entropy contribution of the convective term $\LRb{\bm{v}_N^T \bm{V}_h^T\LRp{\bm{Q}_h \circ \bm{F}}\bm{1}}$ over time and reduced order solution with viscosity parameter $\epsilon = 0$ at $T=.75$.
+The solution does not blow up despite the presence of large oscillations resulting from the shock.}
+\label{fig:novisc}
+\end{figure}
+
 Next, we examine the difference in the hyper-reduced treatments of entropy dissipation (\ref{eq:visc1}), (\ref{eq:visc2}), and (\ref{eq:visc3}) described in Section~\ref{sec:diss} (recall that (\ref{eq:visc3}) is not provably entropy dissipative).
-We have verified that the convective term $\bm{v}_N^T \bm{V}_h^T\LRp{\bm{Q}_h \circ \bm{F}}\bm{1}$ is $O\LRp{10^{-15}}$ (near machine precision) over the entire simulation, confirming that the formulation (\ref{eq:esbchr}) is discretely entropy conservative in the absence of viscosity and boundary flux penalization terms.
-Thus, if the discrete entropy dissipation $\bm{v}_N^T\bm{d}(\bm{u}_N)$ is positive, then the simulation is entropy stable.
+If the discrete entropy dissipation $\bm{v}_N^T\bm{d}(\bm{u}_N)$ is positive, then the simulation is entropy stable.
 Figure~\ref{fig:entropydiss}  shows the computed entropy dissipation over the time interval $[0,.75]$.
 All hyper-reduced treatments produce similar results, with the entropy dissipation produced by the naive approximation (\ref{eq:visc3}) differing most significantly.
-The naive approximation differs from (\ref{eq:visc1}) and (\ref{eq:visc2}) by about $O\LRp{10^{-1}}$ to $O\LRp{10^{-2}}$, while (\ref{eq:visc1}) differs from (\ref{eq:visc2}) by $O\LRp{10^{-6}}$.
-\begin{figure}
+The naive approximation differs from the provably entropy stable approximations (\ref{eq:visc1}) and (\ref{eq:visc2}) by about $O\LRp{10^{-1}}$ to $O\LRp{10^{-2}}$, while (\ref{eq:visc1}) differs from (\ref{eq:visc2}) by $O\LRp{10^{-6}}$.
+
+\begin{figure}[!h]
 \centering
@@
 \subfloat[75 modes]{\includegraphics[width=.45\textwidth]{visc75modes.png}}
-\caption{Discrete entropy dissipation $\bm{v}_N^T\bm{d}(\bm{u}_N)$ computed using (\ref{eq:visc1}), (\ref{eq:visc2}), and (\ref{eq:visc3}) for 25 and 75 modes.}
+\caption{Discrete entropy dissipation $\bm{v}_N^T\bm{d}(\bm{u}_N)$ computed using hyper-reduced treatments of viscous terms (\ref{eq:visc1}), (\ref{eq:visc2}), and (\ref{eq:visc3}) for 25 and 75 modes.}
 \label{fig:entropydiss}
@@
 We now compare the evolution of errors and entropy, and examine the effect of hyper-reduction on the discrete solution in Figure~\ref{fig:entropyevo}.
+Both entropy and error are computed on the original grid of the full order model (e.g., without hyper-reduction).
 We observe that the evolution of the ROM average discrete entropy converges to the average discrete entropy for the full order model as the number of modes increases from 25 to 75.
-For 125 modes, the average entropies for the ROM and FOM are indistinguishable.
+For 125 modes, the average entropies for the ROM and FOM are indistinguishable and are not shown.
 The errors behave similarly, decreasing as the number of modes increases.
 For reference, we also include errors computed without hyper-reduction using (\ref{eq:esrom}).
-We observe that the hyper-reduced errors are similar to errors without hyper-reduction in all cases, though for $75$ modes, hyper-reduction produces lower errors.
-The reasons for this are unclear.
-\begin{figure}
+We observe that the hyper-reduced errors are virtually identical to errors without hyper-reduction in all cases.
+
+\begin{figure}[!h]
 \centering
-\subfloat[Evolution of entropy]{\includegraphics[height=.24\textheight]{entropyevolution.png}}
+\subfloat[Entropy over time]{\includegraphics[height=.265\textheight]{entropyevolution.png}}
 \hspace{.5em}
-\subfloat[Error between ROM and FOM]{\includegraphics[height=.235\textheight]{euler_error_hypred_comparison_K2500}}
+\subfloat[Error between ROM and FOM]{\includegraphics[height=.26\textheight]{euler_error_hypred_comparison_K2500}}
 %\hspace{.1em}
@@
 %\subfloat[Entropy 75 modes]{\includegraphics[width=.32\textwidth]{euler1Dwall_entropy_75modes.png}}
-\caption{Evolution of error (with and without hyper-reduction) and entropy over time interval $[0,.75]$ for $25, 75, 125$ modes.}
+\caption{Evolution of error (with and without hyper-reduction) and entropy over time interval $[0,.75]$.}
 \label{fig:entropyevo}
@@
 Since little difference is observed between the hyper-reducted treatments of viscosity,  the naive treatment (\ref{eq:visc3}) is used.
-The solution does not form shock disccontinuities, and is chosen instead to test the imposition of boundary conditions.
+The solution does not form shock discontinuities, and is chosen instead to test the imposition of boundary conditions.
 We note that, due to the nature of the linear programming software used to compute the boundary weights, the constraints (\ref{eq:sbpconstraints}) on the boundary weights $\bm{w}_b$ are imposed to a tolerance of $5e-8$.
@@
 \subfloat[Hyper-reduction points]{\includegraphics[width=.32\textwidth]{pulse2d_ROM_pts.png}}
-\caption{Density computed using 25 modes, 303 hyper-reduced volume points, and 75 hyper-reduced boundary points.  The relative $L^2$ error is $.00934$.}
+\caption{Density computed using 25 modes, 303 hyper-reduced volume points (in \textcolor{red}{red}), and 75 hyper-reduced boundary points (in \textcolor{blue}{blue}).
+The relative $L^2$ error is $.00934$.}
 \label{fig:pulse2d}
@@
 
-\section{Conclusion}
+\section{Conclusion}  We have presented a methodology for constructing projection-based reduced order models for nonlinear conservation laws from a reduced basis, modified Galerkin projection, and tailored hyper-reduction techniques.
+The main novelty of these new models is the approximation of the nonlinear convective term, which combines a ``flux differencing'' approach with a hyper-reduced approximation of the differentiation matrix.
+
+Future work will aim to address computational costs associated with entropy stable reduced models.
+The main challenge is the fact that, unlike standard hyper-reduced models, the number of nonlinear evaluations necessary scales with $O(N_s^2)$ rather than $O(N_s)$, where $N_s$ is the number of hyper-reduced sampling points.
+These costs can be reduced using, for example, domain decomposition \cite{lucia2003reduced} or local $h$-refinement \cite{carlberg2015adaptive}, which will be explored in future work.
 

### 2019-09-06 docs/ESDG_GQ/reviewerResponse.tex
updating docs

@@
 %\newcommand{\LRceil}[1]{\left\lceil #1 \right\rceil}
-%\newcommand{\LRl}[1]{\left. \LRp{#1} \right|}
+\newcommand{\LRl}[1]{\left. \LRp{#1} \right|}
 %\newcommand{\jump}[1] {\ensuremath{\llbracket#1\rrbracket}}
@@
 We describe steps taken to address reviewer comments and suggestions, which are described in the following response.
-Revisions in the manuscript are also colored for ease of identification.
-We hope these revisions improve the readability of this paper and its suitability for the audience of SISC.}
+We have also edited a few additional sections of the manuscript to improve their readability, and have updated the numerical results of Figure 5, which were previously incorrectly reported.
+\\
+\\
+Revisions in the manuscript are also colored for ease of identification.  We hope these revisions improve the readability of this paper and its suitability for the audience of SISC.}
 
@@
 
-\note{blah}
+\note{We thank the reviewer for their comments.
+We agree that testing these methods on curved unstructured meshes is necessary to evaluate their performance, but since this is currently being done in the context of a separate ongoing project, we wish to restrict the scope of the numerical experiments in this paper.
+We have added the following remarks in the conclusion noting that the proposed method can be extended to unstructured curved meshes:
+\begin{quote}
+We note that, while the numerical experiments presented here consider only mapped Cartesian domains, the method is also applicable to complex geometries, and future work will focus on studying the performance of such methods on curvilinear quadrilateral and hexahedral unstructured meshes.
+\end{quote}
+}
+
 \end{itemize}
@@
 We've added a continuous description of the equation, and have interpreted the matrix version as a quadrature approximation of a specific variational problem for the approximation of the derivative.
-This variational problem includes boundary correction terms, which correspond to the matrix ``strong form'' of (3.8).}
+This variational problem includes boundary correction terms, which correspond to the matrix ``strong form'' of (3.8).
+The added text is as follows:}
+\begin{quote}
+\note{More specifically, the expression (3.8) corresponds to a quadrature approximation of the following variational approximation of the derivative: find $u\in P^N([-1,1])$ such that,}
+\[
+\note{\int_{-1}^1u({x})v({x})  = \int_{-1}^1 {I_N f\pd{I_Ng}{x}v} + \LRl{(g-I_Ng)\frac{\LRp{fv + I_N(fv)}}{2}}_{-1}^1, \qquad \forall v\in P^N([-1,1]),}
+\]
+\note{where $I_N$ denotes the degree $N$ polynomial approximation at the $(N+1)$ Gauss points.}
+\end{quote}
 
@@
 \note{The block structure of the two-element mesh can be extended to multiple elements.
-We have added a discussion involving a three-element mesh, which illustrates the procedure without introducing too much additional notation.
-The approach simply requires a global SBP operator over the entire mesh, which is constructed in a block fashion.
+We have added a discussion involving a three-element mesh on pages 11-12, which illustrates the procedure without introducing too much additional notation.
+We also describe how the extension to multiple elements approach requires a global SBP operator over the entire mesh, which is constructed in a similar block-wisee fashion.
 We hope this revision makes the extension to multiple elements more clear.
@@
 
-\note{We have added a short discussion describing how to impose boundary conditions in a weak fashion through a boundary numerical flux.
-If the boundary numerical flux satisfies an entropy stability property, then the discretization with boundary conditions is also entropy stable.}
+\note{We have added a short discussion on pages 12-13 describing how to impose boundary conditions in a weak fashion through a boundary numerical flux.
+If the boundary numerical flux satisfies an entropy stability property, then the discretization with boundary conditions is also entropy stable.
+The added text is as follows:}
+\begin{quote}
+\note{If the boundary numerical flux satisfies the entropy stability conditions}
+\[
+\note{\psi_L - \bm{v}_L^T\bm{f}^*_{L} \leq 0, \qquad {\psi}_R - \bm{v}_R^T\bm{f}^*_{R} \leq 0}
+\]
+\note{then the resulting scheme satisfies a global entropy inequality \cite{chen2017entropy}.}
+\end{quote}
 
@@
 With the exception of the evaluation of the nonlinear term, the procedure is relatively standard for time-explicit DG formulations.
+The added text is as follows:}
+
+\begin{quote}
+\note{Assuming that the ODE system (3.14) exactly integrated in time and that entropy dissipative numerical fluxes and boundary conditions are used, the solution will satisfy a discrete entropy inequality (2.3).
+In practice, the system (3.14) is solved using an ODE time-stepper.
+For an explicit time-stepper, then all that is necessary is to invert the diagonal matrix $\bm{W}_h$ and evaluate the spatial terms in (3.15).
 }
+\end{quote}
 
@@
 
-\note{We thank the reviewer for catching this - it is indeed $2\times (N+1)$.}
+\note{We thank the reviewer for catching this - it is indeed $2\times (N+1)$.  We have fixed this in the revised manuscript.}
 
@@
 
-\note{We apologize for the confusion.  $\bm{u}_i$ should be a scalar, and we have edited the paper to clear this up.  We are no longer using bold to denote velocity $u_i$.}
+\note{We apologize for the confusion.  $\bm{u}_i$ should be a scalar, and we have edited the paper to clear this up.  We no longer use bold to denote velocity $u_i$ in the revised manuscript.}
 
@@
 \note{We have added definitions of the average and logarithmic mean.
-We have also clarified the notation that $u,v$ denote $x,y$ velocity, while bolded $\bm{u},\bm{v}$ denote conservative and entropy variables.}
+We have also removed $u,v$ notation and replaced it with $u_i$ notation, where $u_i$ denotes the scalar velocity in the $i$th coordinate direction.
+}
 \end{itemize}
 
+\bibliographystyle{unsrt}
+\bibliography{dg}
+
 \end{document}

### 2019-09-06 docs/skewform/reviewerResponse.tex
updating docs

\pdfoutput=1

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{bm}
%\usepackage{bbm}
%\usepackage{tikz}
%\usepackage[normalem]{ulem}
%\usepackage{hhline}
%\usepackage{graphicx}
%\usepackage{subfig}
\usepackage{color}
%\usepackage{doi}

%\renewcommand{\topfraction}{0.85}
%\renewcommand{\textfraction}{0.1}
%\renewcommand{\floatpagefraction}{0.75}
%
%
%\newcommand{\bbm}[1]{\mathbbm{#1}}
%\newcommand{\bs}[1]{\boldsymbol{#1}}
%\newcommand{\equaldef}{\stackrel{\mathrm{def}}{=}}
%
%
%\newcommand{\mb}[1]{\mathbf{#1}}
%\newcommand{\mbb}[1]{\mathbb{#1}}
%\newcommand{\mc}[1]{\mathcal{#1}}
%
%\renewcommand{\hat}{\widehat}
%\renewcommand{\tilde}{\widetilde}
\newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}{\rm #2}}}
\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
%\newcommand{\pdn}[3]{\frac{\partial^{#3}#1}{\partial#2^{#3}}}
%\newcommand{\snor}[1]{\left| #1 \right|}
%\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
%\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
%\newcommand{\LRb}[1]{\left| #1 \right|}
%\newcommand{\LRc}[1]{\left\{ #1 \right\}}
%\newcommand{\LRceil}[1]{\left\lceil #1 \right\rceil}
\newcommand{\LRl}[1]{\left. \LRp{#1} \right|}
%\newcommand{\jump}[1] {\ensuremath{\llbracket#1\rrbracket}}
%\newcommand{\avg}[1] {\ensuremath{\LRc{\!\{#1\}\!}}}
%\newcommand{\Grad} {\ensuremath{\nabla}}
%\renewcommand{\d}{\partial}
\newcommand{\diag}[1]{{\rm diag}\LRp{#1}}

\newcommand{\note}[1]{{\color{blue}{#1}}}
\newcommand{\bnote}[1]{{\color{blue}{#1}}}
\newcommand{\rnote}[1]{{\color{red}{#1}}}

\newcommand{\LK}{L^2\LRp{D^k}}
\newcommand{\LdK}{L^2\LRp{\partial D^k}}
\newcommand{\Dhat}{\widehat{D}}
\newcommand{\Lhat}{L^2\LRp{\Dhat}}

\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}} % d in integrand

\begin{document}

%\maketitle

I am very grateful to both \bnote{Reviewer 1} and \rnote{Reviewer 2} for their feedback.  I describe steps taken to address reviewer comments and suggestions in the following response.  I have also edited a few additional sections of the manuscript to improve their readability, and have added some more recent references.
\\
\\
Revisions in the manuscript are also colored for ease of identification.  We hope these revisions improve the readability of this paper and its suitability for the audience of JSC.

\section{Reviewer 1}

\textit{My main concern with this paper is the discussion of what has been accomplished. Specifically, the paper reads as if the SBP property is not required; however, this is not the case and what the author has presented is a means of constructing decoupled SBP operators using nearly arbitrary surface and volume quadrature nodes. }

\bnote{I wholeheartedly agree with Reviewer 1's assessment of the structure of the paper, and thank them for bringing this to my attention.  I have restructured the paper to make this more clear, and have added references to make connections to similar approaches in the SBP literature.  The main restructuring shows how to construct the SBP operator, and utilizes the ``variational'' SBP property only to prove accuracy properties.  The paper is also shorter and more concise after these revisions.}

\begin{enumerate}
\item \textit{The major comment is that the discussion throughout the paper on the strong SBP property vs. the variational SBP property does not clearly transmit what the author has accomplished in this manuscript and is misleading. The author gives a systematic approach to the construction of modal decoupled SBP operators from nearly arbitrary sets of volume and surface quadrature rules. Note that the skew-symmetric operators that are used in the algorithm satisfy the strong SBP property. What the author needs to clearly state is that the construction of decoupled SBP operators proposed in (Chan 2018) can lose their SBP property for specific choices of volume and surface quadrature rules and that this paper shows how to remedy this issue.}.

\textit{It would be helpful [\ldots] to point out that the presented formulation is equivalent to using the newly defined decoupled SBP operator in computational space as applied to a skew-symmetric splitting of the conservation law.}

\bnote{I have rewritten significant portions of the manuscript.  The reviewer's observation that the skew-symmetric operators constructed satisfy the strong SBP property significantly simplifies proofs of the key lemmas and theorems.  The revised manuscript simply utilizes the strong SBP property now, and shifts the theoretical work in proving entropy conservation to showing that the reference skew-symmetric operators satisfy $\tilde{\bm{Q}}^i_N\bm{1} = \bm{0}$, and that the physical SBP operators in computational space satisfy an analogous condition.}

\end{enumerate}

\section{Reviewer 2}

\textit{The author generalizes his approach from [6,7] to handle cases where the discrete operator does not satisfy the decoupled SBP property.  The paper was nicely written and enjoyable to read.  I can find few faults with it.  The one suggestion I have is that the author bring up the constraints on accuracy sooner (such as in the abstract).  Only toward the end do we find out that, well, you probably still want the decoupled SBP property, or degree $2N-1$ volume and degree $2N$ surface quadratures, since accuracy will suffer otherwise.}

\rnote{I agree, and have added this information to both the introduction and conclusions.  The minor comments and typos have also been addressed; however, several no longer apply to the revised manuscript after the significant revisions made in light of Reviewer 1's comments.  }

\begin{itemize}
\item \rnote{I have corrected the following errors pointed out by Reviewer 2:}
\begin{itemize}
\item Page 2, Lines 16-19: I think Tadmor's 1987 work should be cited here too, no?
\item Page 3, Eq (1): subscript h on u, and should the zero on the RHS be bold?
\item Page 5, lines 5-6: The subscript on the basis does not match the index on the set.
\item Page 11, line 18: Here and elsewhere (5.3) seems to be a broken reference to (10).
\item Page 12, Eq (12): Ordinary derivative for the temporal derivative?
\item Page 15, line 43: As with (5.3), (6.1) seems to be a broken reference.
\item Page 16, line 47: Bold zero in GCL?
\item Page 19, line 34: "...surface quadrature on for which ..." typo  
\item Page 20, lines 7-8: "...satisfy Assumption 1 holds for v = 1."  typo
\item Page 26, line 13:  Search for "on on"
\item Page 26, line 50: "...we fix the volume to quadrature..."  Delete "to" ?
\end{itemize}
\item \textit{Page 12, Eq (14): I think $\psi_i$ should probably be a vector here for consistency.}

\rnote{I believe $\psi$ should remain a scalar here, since it's a function of a vector quantity.  To clarify this, however, I have added an extra line: ``Here, $\psi_i(\tilde{\bm{u}}_f)$ denotes the function $\psi_i$ evaluated at the face values of the entropy-projected conservative variables $\tilde{\bm{u}}_f$''.}

\item \textit{Page 15, line 21: Some ideas here go back to Thomas and Lombard, "Geometric conservation law and its application to flow computations on moving grids."}

\rnote{I have added this reference to the revised manuscript.}
\end{itemize}

\bibliographystyle{unsrt}
\bibliography{dg}

\end{document}

### 2019-09-06 docs/ESDG_GQ/arxiv_v2/esdg_gq.tex
updating docs

--- 2019-03-07 docs/ESDG_GQ/esdg_gq.tex
@@
 %\documentclass[review, 10pt]{siamart0216}
-%\documentclass{siamart0216}
-\documentclass[review,onefignum,onetabnum,final]{siamart171218}
+\documentclass{siamart0216}
+%\documentclass[review,onefignum,onetabnum,final]{siamart171218}
 
@@
 \newcommand{\LRl}[1]{\left. #1 \right|}
+
 \newcommand{\pdd}[2]{\frac{\partial^2#1}{\partial#2^2}}
@@
 
-\newcommand{\note}[1]{{\color{blue}{#1}}}
+\newcommand{\note}[1]{#1}
+%\newcommand{\note}[1]{{\color{blue}{#1}}}
 %\newcommand{\noteOne}[1]{{\color{blue}{#1}}}
@@
 
-This focus of this paper is on entropy stable high order collocation methods which satisfy a semi-discrete version of the entropy inequality (\ref{eq:weakentropyineq}).
+The focus of this paper is on entropy stable high order collocation methods which satisfy a semi-discrete version of the entropy inequality (\ref{eq:weakentropyineq}).
 These methods collocate the solution at some choice of collocation nodes, and are applicable to tensor product meshes consisting of quadrilateral and hexahedral elements.
@@
 %\]
-It can be show that the mass and differentiation matrices for Gauss nodes fall under the class of generalized SBP (GSBP) operators \cite{fernandez2014generalized}.  
+It can be shown that the mass and differentiation matrices for Gauss nodes fall under the class of generalized SBP (GSBP) operators \cite{fernandez2014generalized}.  
 \begin{lemma}
@@
 \]
-\note{where $I_N$ denotes the degree $N$ polynomial approximation at the $(N+1)$ Gauss points.}
+\note{where $I_N$ denotes the degree $N$ polynomial approximation at the $(N+1)$ Gauss points.  }
 
@@
 && \bm{S} &  \frac{1}{2}\bm{V}_f^T\bm{B}&& \\
-& -\frac{1}{2}\bm{e}_L\bm{e}_R^T &  - \frac{1}{2}\bm{B}\bm{V}_f &&\frac{1}{2}\bm{e}_R\bm{e}_L^T&\\
+& -\frac{1}{2}\bm{e}_L\bm{e}_R^T &  - \frac{1}{2}\bm{B}\bm{V}_f &&&\frac{1}{2}\bm{e}_R\bm{e}_L^T\\
  \hline\\[-1.0em]
@@
 \end{gather*}
-\note{These global operators again satisfy the SBP property $\bm{Q}_h + \bm{Q}_h^T = \bm{B}_h$, where $\bm{Q}_h = \bm{S}_h + \frac{1}{2}\bm{B}_h$.
+\note{These global operators again satisfy $\bm{Q}_h\bm{1} = 0$ and the SBP property $\bm{Q}_h + \bm{Q}_h^T = \bm{B}_h$, where $\bm{Q}_h = \bm{S}_h + \frac{1}{2}\bm{B}_h$.
 Moreover, $\bm{Q}_h$ can be used to produce high order accurate approximations of derivatives, and can be used to construct entropy conservative schemes as in Theorem~\ref{thm:consentropy}.
-The extension to a general number of elements is straightforward.
+The extension to a general number of elements is done in a similar fashion.
 }
@@
 \]
-\note{where we have used the consistency of $\bm{f}_S$ and the fact that $\bm{B}_h$ is diagonal to simplify the boundary term.
-We can now replace $\bm{f}(\bm{u}_L), \bm{f}(\bm{u}_R)$ with numerical fluxes $\bm{f}^*_L, \bm{f}^*_R$ as follows:}
+\note{where we have used the fact that $\bm{B}_h$ is diagonal and the consistency of $\bm{f}_S$ (implying that the diagonal of $\bm{F}_S$ reduces to the evaluation of the flux $\bm{f}(\bm{u})$) to simplify the boundary term $\LRp{\bm{B}_h\circ\bm{F}_S}\bm{1}$.
+Boundary conditions are incorporated by replacing $\bm{f}(\bm{u}_L), \bm{f}(\bm{u}_R)$ with numerical fluxes $\bm{f}^*_L, \bm{f}^*_R$ as follows:}
 \begin{equation}
@@
 For an explicit time-stepper, then all that is necessary is to invert the diagonal matrix $\bm{W}_h$ and evaluate the spatial terms in (\ref{eq:bform}).
+
 We note that the conservation or dissipation of entropy is guaranteed up to time-stepper accuracy.
-In practice, we observe that entropy is dissipated for all problems considered, despite the fact that the proof does not hold at the fully discrete level.}
+In practice, we observe that entropy is dissipated for all problems considered, despite the fact that the proof does not hold at the fully discrete level.
+A fully discrete entropy inequality can be guaranteed using implicit or space-time discretizations \cite{tadmor2003entropy, friedrich2018entropy}.}
 
@@
 
-Let $\bm{D}_{\rm 1D}, \bm{W}_{\rm 1D}$ denote the 1D differentiation and mass matrices, respectively, on on the reference interval $[-1,1]$.
+Let $\bm{D}_{\rm 1D}, \bm{W}_{\rm 1D}$ denote the 1D differentiation and mass matrices, respectively, on the reference interval $[-1,1]$.
 Let ${\bm{W}}$ denote the 2D reference mass matrix, and let ${\bm{D}}^i$ denote the differentiation matrices with respect to the $i$th reference coordinate.
@@
 \end{align*}
-Here, $\rho$ is density, \note{$\LRp{{u}_1, \ldots,{u}_d}$} is the vector of velocities, and $E$ is the total energy.  The pressure $p$ and specific internal energy $\rho e$ are given by 
+Here, $\rho$ is density, \note{$u_i$ denotes the $i$th component of} velocity, and $E$ is the total energy.  The pressure $p$ and specific internal energy $\rho e$ are given by 
 \[
@@
 \[
-\note{\avg{f} = \frac{f^+ + f}{2}}, \qquad \note{\avg{f}^{\log} = \frac{f+ - f}{\log\LRp{f^+}-\log\LRp{f}}}.
+\note{\avg{f} = \frac{f^+ + f}{2}}, \qquad \note{\avg{f}^{\log} = \frac{f^+ - f}{\log\LRp{f^+}-\log\LRp{f}}}.
 \]
 \note{The average and logarithmic average are assumed to act component-wise on vector-valued functions.
-We evaluate the logarithmic average using the numerically stable approach of \cite{ismail2009affordable}.}
+We evaluate the logarithmic average using the numerically stable algorithm of \cite{ismail2009affordable}.}
 Explicit expressions for entropy conservative numerical fluxes in two dimensions  are given by Chandrashekar \cite{chandrashekar2013kinetic}
@@
 \end{align*}
-where $(u_1,u_2,u_3)^T$ is the velocity vector and % $\Pi$, $\bm{r}$ are defined as
+where $(u_1,u_2,u_3)^T$ is the three-dimensional velocity vector and % $\Pi$, $\bm{r}$ are defined as
 \[
@@
 \rho = 1, \qquad p = \frac{100}{\gamma} + \frac{1}{16} \LRp{\cos(2x_1) + \cos(2x_2)}\LRp{2+\cos(2x_3)},\\
-u_1 = \sin(x_1)\cos(x_2)\cos(x_3), \qquad
-u_2 = -\cos(x_1)\sin(x_2)\cos(x_3),\qquad
-u_3 = 0.
+\note{u_1 = \sin(x_1)\cos(x_2)\cos(x_3),} \qquad
+\note{u_2 = -\cos(x_1)\sin(x_2)\cos(x_3),}\qquad
+\note{u_3 = 0.}
 \end{gather*}
@@
 Numerical experiments demonstrate both the stability and high order accuracy of the proposed Gauss collocation schemes on both affine and curvilinear meshes.
+\note{We note that, while the numerical experiments presented here consider only mapped Cartesian domains, the method is also applicable to complex geometries, and future work will focus on studying the performance of such methods on curvilinear quadrilateral and hexahedral unstructured meshes.}
 

### 2019-09-06 docs/ESDG_GQ/esdg_gq.tex
updating docs

@@
 \newcommand{\LRl}[1]{\left. #1 \right|}
+\newcommand{\note}[1]{{\color{blue}{#1}}}
+
 \newcommand{\pdd}[2]{\frac{\partial^2#1}{\partial#2^2}}
@@
 
-\newcommand{\note}[1]{{\color{blue}{#1}}}
+%\newcommand{\note}[1]{{\color{blue}{#1}}}
 %\newcommand{\noteOne}[1]{{\color{blue}{#1}}}

### 2019-09-06 docs/ESDG_GQ/esdg_gq_nasa.tex
updating docs

--- 2019-09-06 docs/ESDG_GQ/esdg_gq.tex
@@
 \begin{abstract}
-The construction of high order entropy stable collocation schemes on quadrilateral and hexahedral elements has relied on the use of Gauss-Legendre-Lobatto collocation points \cite{fisher2013high, carpenter2014entropy, gassner2016split} and their equivalence with summation-by-parts (SBP) finite difference operators \cite{gassner2013skew}.
-In this work, we show how to efficiently generalize the construction of semi-discretely entropy stable schemes on tensor product elements to Gauss points and generalized SBP operators.
+The construction of high-order entropy stable collocation schemes on quadrilateral and hexahedral elements has relied on the use of Gauss-Legendre-Lobatto collocation points \cite{fisher2013high, carpenter2014entropy, gassner2016split} and their equivalence with summation-by-parts (SBP) finite difference operators \cite{gassner2013skew}.
+In this work, we show how to efficiently generalize the construction of semidiscrete, entropy stable schemes on tensor product elements to Gauss points and generalized SBP operators.
 Numerical experiments suggest that the use of Gauss points significantly improves accuracy on curved meshes.
@@
 The utility of these simulations depends both on the reliability and accuracy of numerical methods for the solution of the associated nonlinear partial differential equations (PDEs).
-Time dependent nonlinear conservation laws are ubiquitous in computational fluid dynamics, for which high order methods are increasingly of interest.
-Such methods are more accurate per degree of freedom than low order methods, while also possessing much smaller numerical dispersion and dissipation errors.
-This makes high order methods especially well suited to time-dependent simulations.
+Time dependent nonlinear conservation laws are ubiquitous in computational fluid dynamics, for which high-order methods are increasingly of interest.
+Such methods are more accurate per degree of freedom than low-order methods, while also possessing much smaller numerical dispersion and dissipation errors.
+This makes high-order methods especially well suited to time-dependent simulations.
 In this work, we focus specifically on discontinuous Galerkin methods on unstructured quadrilateral and hexahedral meshes.
-These methods combine properties of high order approximations with the geometric flexibility of unstructured meshing.
+These methods combine properties of high-order approximations with the geometric flexibility of unstructured meshing.
 %When paired with explicit time-stepping methods, high order DG and finite element methods are also well-suited for many-core and GPU computing \cite{klockner2009nodal, swirydowicz2017acceleration, wintermeyer2018entropy}.
 
-However, high order methods are notorious for being more prone to instability compared to low order methods \cite{wang2013high}.
+However, high-order methods are notorious for being more prone to instability compared to low-order methods \cite{wang2013high}.
 This instability is addressed through various stabilization techniques (e.g.\ artificial viscosity, filtering, slope limiting).
-However, these techniques often reduce accuracy to first or second order, and can prevent solvers from realizing the advantages of high order approximations.
-Moreover, it is often not possible to prove that a high order scheme does not blow up even in the presence of stabilization.
-This ambiguity can necessitate the re-tuning of stabilization parameters, as parameters which are both stable and accurate for one problem or discretization setting  may provide either too little or too much numerical dissipation for another.
-
-The instability of high order methods is rooted in the fact that discretizations of nonlinear conservation laws do not typically satisfy a discrete analogue of the conservation or dissipation of energy (entropy).
-For low order methods, the lack of discrete stability can be offset by the presence of numerical dissipation, which serves as a stabilization mechanism.
-Because high order methods possess low numerical dissipation, the absence of a discrete stability property becomes more noticeable, manifesting itself through increased sensitivity and instability.
+However, these techniques often reduce accuracy to first or second order, and can prevent solvers from realizing the advantages of high-order approximations.
+Moreover, it is often not possible to prove that a high-order scheme does not blow up even in the presence of stabilization.
+This ambiguity can necessitate the retuning of stabilization parameters, as parameters which are both stable and accurate for one problem or discretization setting  may provide either too little or too much numerical dissipation for another.
+
+The instability of high-order methods is rooted in the fact that discretizations of nonlinear conservation laws do not typically satisfy a discrete analogue of the conservation or dissipation of energy (entropy).
+For low-order methods, the lack of discrete stability can be offset by the presence of numerical dissipation, which serves as a stabilization mechanism.
+Because high-order methods possess low numerical dissipation, the absence of a discrete stability property becomes more noticeable, manifesting itself through increased sensitivity and instability.
 
@@
 Discrete instability is typically tied to the fact that, when discretizing systems of nonlinear PDEs, the chain rule does not typically hold at the discrete level.
-The lack of a chain rule was circumvented by using a non-standard ``flux differencing'' formulation \cite{fisher2013high, carpenter2014entropy, gassner2016split, gassner2017br1}, which is key to constructing semi-discretely entropy stable high order schemes on unstructured quadrilateral and hexahedral meshes.
+The lack of a chain rule was circumvented by using a nonstandard ``flux differencing'' formulation \cite{fisher2013high, carpenter2014entropy, gassner2016split, gassner2017br1}, which is key to constructing semidiscrete, entropy stable, high-order schemes on unstructured quadrilateral and hexahedral meshes.
 Flux differencing replaces the derivative of the nonlinear flux with the discrete differentiation of an auxiliary quantity.
 This auxiliary quantity is computed through the evaluation of a two-point entropy conservative flux \cite{tadmor1987numerical} using pairs of solution values at quadrature points.
-These entropy stable schemes were later extended to non-tensor product elements using GLL-like quadrature points on triangles and tetrahedra \cite{chen2017entropy, crean2018entropy}.
+These entropy-stable schemes were later extended to nontensor product elements using GLL-like quadrature points on triangles and tetrahedra \cite{chen2017entropy, crean2018entropy}.
 More recently, the construction of efficient entropy stable schemes was extended to more arbitrary choices of basis and quadrature \cite{chan2017discretely, chan2018discretely}.
 
-While entropy stable collocation schemes have been constructed on Gauss-like quadrature points without boundary nodes \cite{crean2017high}, the inter-element coupling terms for such schemes introduce an ``all-to-all'' coupling between degrees of freedom on two neighboring elements in one dimension (on tensor product elements in higher dimensions, these coupling terms couple together lines of nodes).
+While entropy-stable collocation schemes have been constructed on Gauss-like quadrature points without boundary nodes \cite{crean2017high}, the interelement coupling terms for such schemes introduce an ``all-to-all'' coupling between degrees of freedom on two neighboring elements in one dimension (on tensor product elements in higher dimensions, these coupling terms couple together lines of nodes).
 These coupling terms require evaluating two-point fluxes between solution states at all collocation nodes on two neighboring elements, resulting in significantly more communication and computation compared to collocation schemes based on point sets containing boundary nodes.
-This work introduces efficient and entropy stable inter-element coupling terms for Gauss collocation schemes which require only communication of face values between neighboring elements.
+This work introduces efficient and entropy stable interelement coupling terms for Gauss collocation schemes, which require only communication of face values between neighboring elements.
 The construction of these terms follows the framework introduced in \cite{chan2017discretely, chan2018discretely} for triangles and tetrahedra.
 
-The main motivation for exploring tensor product (quadrilateral and hexahedral) elements is the significant reduction in the number of operations required compared to high order entropy stable schemes on simplicial meshes \cite{chan2017discretely, chan2018discretely}.
+The main motivation for exploring tensor product (quadrilateral and hexahedral) elements is the significant reduction in the number of operations required compared to high-order entropy stable schemes on simplicial meshes \cite{chan2017discretely, chan2018discretely}.
 Entropy stability schemes on simplicial elements require evaluating two-point fluxes between solution states at all quadrature points on an element.
@@
 In Section~\ref{sec:1}, we describe how to construct entropy stable discretizations of nonlinear conservation laws using different quadrature points on affine tensor product elements.
-In Section~\ref{sec:2}, we describe how to extend this construction to curvilinear elements, and Section~\ref{sec:3} presents numerical results which confirm the high order accuracy and stability of the proposed method for smooth, discontinuous, and under-resolved (turbulent) solutions of the compressible Euler equations in two and three dimensions.
+In Section~\ref{sec:2}, we describe how to extend this construction to curvilinear elements, and Section~\ref{sec:3} presents numerical results which confirm the high-order accuracy and stability of the proposed method for smooth, discontinuous, and underresolved (turbulent) solutions of the compressible Euler equations in two and three dimensions.
 
@@
 
-The focus of this paper is on entropy stable high order collocation methods which satisfy a semi-discrete version of the entropy inequality (\ref{eq:weakentropyineq}).
+The focus of this paper is on entropy stable high-order collocation methods that satisfy a semidiscrete version of the entropy inequality (\ref{eq:weakentropyineq}).
 These methods collocate the solution at some choice of collocation nodes, and are applicable to tensor product meshes consisting of quadrilateral and hexahedral elements.
@@
 The popularity of GLL nodes can be attributed in part to a connection made in \cite{gassner2013skew}, where it was shown by Gassner that collocation DG discretizations based on GLL nodes could be recast in terms of summation-by-parts (SBP) operators.
-This equivalence allowed Gassner to leverage existing finite difference formulations to produce stable high order discretizations of the nonlinear Burgers' equation.
-
-GLL quadratures contain boundary points, which greatly simplifies the construction of inter-element coupling terms for entropy stable collocation schemes.
-However, it is also known that the use of GLL quadrature within DG methods under-integrates the mass matrix, which can lead to solution ``aliasing'' and lower accuracy \cite{parsani2016entropy}.
+This equivalence allowed Gassner to leverage existing finite difference formulations to produce stable high-order discretizations of the nonlinear Burgers' equation.
+
+GLL quadratures contain boundary points, which greatly simplifies the construction of interelement coupling terms for entropy stable collocation schemes.
+However, it is also known that the use of GLL quadrature within DG methods underintegrates the mass matrix, which can lead to solution ``aliasing'' and lower accuracy \cite{parsani2016entropy}.
 In this work, we explore entropy stable collocation schemes based on Gauss quadrature points instead of GLL points.
@@
 This higher accuracy of Gauss quadrature has been shown to translate to lower errors and slightly improved rates of convergence in simulations of wave propagation and fluid flow \cite{kopriva2010quadrature, hindenlang2012explicit, chan2015gpu}.
-However, Gauss points have not been widely used to construct entropy stable discretizations due to the lack of efficient, stable, and high order accurate inter-element coupling terms, known as simultaneous approximation terms (SAT) in the finite difference literature \cite{fernandez2014review, crean2017high, fernandez2018simultaneous}.
-SATs for Gauss points are non-compact, in the sense that they introduce all-to-all coupling between degrees of freedom on neighboring elements in one dimension.
+However, Gauss points have not been widely used to construct entropy stable discretizations due to the lack of efficient, stable, and high-order accurate interelement coupling terms, known as simultaneous approximation terms (SAT) in the finite difference literature \cite{fernandez2014review, crean2017high, fernandez2018simultaneous}.
+SATs for Gauss points are noncompact, in the sense that they introduce all-to-all coupling between degrees of freedom on neighboring elements in one dimension.
 This results in greater communication between elements, as well as a significantly larger number of two-point flux evaluations and floating point operations.
 
-It is possible to realize the improved accuracy of Gauss points while avoiding non-compact SATs through a staggered grid formulation, where the solution is stored at Gauss nodes but interpolated to a set of higher degree $(N+2)$ GLL ``flux'' points for computation \cite{parsani2016entropy}.
-Because GLL nodes include boundary points, compact and high order accurate SAT terms can be constructed for the flux points.
+It is possible to realize the improved accuracy of Gauss points while avoiding noncompact SATs through a staggered grid formulation, where the solution is stored at Gauss nodes but interpolated to a set of higher degree $(N+2)$ GLL ``flux'' points for computation \cite{parsani2016entropy}.
+Because GLL nodes include boundary points, compact and high-order accurate SAT terms can be constructed for the flux points.
 After performing computations on the flux points, the results are interpolated back to Gauss points and used to evolve the solution forward in time.
@@
 
-The following sections will describe how to construct efficient high order entropy stable schemes using Gauss points.
+The following sections will describe how to construct efficient high-order entropy stable schemes using Gauss points.
 These schemes are based on ``decoupled'' SBP operators introduced in \cite{chan2017discretely, chan2018discretely}, which are applicable to general choices of basis and quadrature.
 By choosing a tensor product Lagrange polynomial basis and $(N+1)$ point Gauss quadrature rules, we recover a Gauss collocation scheme.
-The high order accuracy and entropy stability of this scheme are direct results of theorems presented in  \cite{chan2017discretely, chan2018discretely}.
+The high-order accuracy and entropy stability of this scheme are direct results of theorems presented in  \cite{chan2017discretely, chan2018discretely}.
 However, we will also present a different proof of entropy stability in one dimension for completeness.
@@
 Since $\bm{M} = \bm{W}$ under a collocation assumption, we utilize $\bm{W}$ for the remainder of this work to emphasize that the mass matrix is diagonal and related to the quadrature weights $w_i$.
-The treatment of non-diagonal mass matrices is covered in \cite{chan2017discretely, chan2018discretely}.
+The treatment of nondiagonal mass matrices is covered in \cite{chan2017discretely, chan2018discretely}.
 
@@
 
-In this section, we will review the construction of semi-discretely entropy stable discretizations.
+In this section, we will review the construction of semi-discrete entropy stable discretizations.
 Entropy stable discretizations can be constructed by first introducing an entropy conservative scheme, then adding appropriate interface dissipation to produce an entropy inequality.
@@
 \label{def:tadmor}
-Let $\bm{f}_S(\bm{u}_L,\bm{u}_R)$ be a bivariate function which is symmetric and consistent with the flux function $\bm{f}(\bm{u})$
+Let $\bm{f}_S(\bm{u}_L,\bm{u}_R)$ be a bivariate function that is symmetric and consistent with the flux function $\bm{f}(\bm{u})$
 \begin{align*}
@@
 
-For illustrative purposes, we will prove a semi-discrete entropy inequality on a one-dimensional mesh consisting of two elements of degree $N$.
+For illustrative purposes, we will prove a semidiscrete entropy inequality on a one-dimensional mesh consisting of two elements of degree $N$.
 We assume both meshes are translations of a reference element  $[-1,1]$, such that derivatives with respect to physical coordinates are identical to derivatives with respect to reference coordinates.
@@
 \]
-The inter-element coupling terms in \cite{fernandez2014review, crean2017high, fernandez2018simultaneous} utilize a decomposition of the surface matrix $\bm{V}_f^T\bm{B}\bm{V}_f$ as  
+The interelement coupling terms in \cite{fernandez2014review, crean2017high, fernandez2018simultaneous} utilize a decomposition of the surface matrix $\bm{V}_f^T\bm{B}\bm{V}_f$ as  
 \begin{equation}
@@
 \end{equation}
-The construction of entropy conservative schemes on multiple elements requires appropriate inter-element coupling terms (SATs) involving $\bm{t}_L, \bm{t}_R$.
+The construction of entropy conservative schemes on multiple elements requires appropriate interelement coupling terms (SATs) involving $\bm{t}_L, \bm{t}_R$.
 We consider a two element mesh, and show how when coupled with SATs, the resulting discretization matrices can be interpreted as constructing a global SBP operator.
@@
 \end{equation}
-It can be shown that $\bm{D}_h$ is high order accurate such that, if $\bm{u}_h$ is a polynomial of degree $N$, it is differentiated exactly.
+It can be shown that $\bm{D}_h$ is high-order accurate such that, if $\bm{u}_h$ is a polynomial of degree $N$, it is differentiated exactly.
 Straightforward computations show that $\bm{Q}_h$ also satisfies an SBP property $\bm{Q}_h + \bm{Q}_h^T = \bm{B}_h$.
@@
 %\end{align*}
-will yield a semi-discrete version of the conservation of entropy (mimicking (\ref{eq:weakentropyineq}) with the inequality replaced by an equality)
+will yield a semidiscrete version of the conservation of entropy (mimicking (\ref{eq:weakentropyineq}) with the inequality replaced by an equality)
 \begin{equation}
@@
 The drawback of the SATs introduced in this section lies in the nature of the off-diagonal matrices $\bm{t}_R\bm{t}_L$ and $-\bm{t}_L\bm{t}_R$.
-For Gauss nodes, these blocks are dense, which implies that inter-element coupling terms produce a non-compact stencil.
+For Gauss nodes, these blocks are dense, which implies that interelement coupling terms produce a noncompact stencil.
 Evaluating (\ref{eq:ESgsbp}) requires computing two-point fluxes $\bm{f}_S$ between all nodes on two neighboring elements, which significantly increases both the computational work, as well as communication between neighboring elements.
@@
 
-It was shown in \cite{chan2017discretely} that $\bm{u}$ is a high order accurate approximation to the quantity $f\pd{g}{x}$.
-Both the generalized SBP operator $\bm{D}$ and the expression in (\ref{eq:qn}) involving the decoupled SBP operator recover exact derivatives of high order polynomials.
+It was shown in \cite{chan2017discretely} that $\bm{u}$ is a high-order accurate approximation to the quantity $f\pd{g}{x}$.
+Both the generalized SBP operator $\bm{D}$ and the expression in (\ref{eq:qn}) involving the decoupled SBP operator recover exact derivatives of high-order polynomials.
 However, when applied to non-polynomial functions, the decoupled SBP operator $\bm{Q}_N$ improves accuracy near the boundaries.
@@
 \end{equation}
-We can now construct a semi-discretely entropy conservative formulation based on decoupled SBP operators:
+We can now construct a semidiscrete entropy conservative formulation based on decoupled SBP operators:
 \begin{theorem}
@@
 \end{gather}
-Then, $\bm{u}_h$ satisfies a semi-discrete conservation of entropy
+Then, $\bm{u}_h$ satisfies a semidiscrete conservation of entropy
 \begin{gather*}
@@
 The proof of Theorem~\ref{thm:consentropy} follows directly from choosing either GLL or Gauss quadratures in Theorem 4 of \cite{chan2017discretely}.
-The proof is reproduced here for clarity, as the two-element case illuminates the skew-symmetric nature and structure of the inter-element coupling more explicitly.
+The proof is reproduced here for clarity, as the two-element case illuminates the skew-symmetric nature and structure of the interelement coupling more explicitly.
 \end{remark}
@@
 \note{These global operators again satisfy $\bm{Q}_h\bm{1} = 0$ and the SBP property $\bm{Q}_h + \bm{Q}_h^T = \bm{B}_h$, where $\bm{Q}_h = \bm{S}_h + \frac{1}{2}\bm{B}_h$.
-Moreover, $\bm{Q}_h$ can be used to produce high order accurate approximations of derivatives, and can be used to construct entropy conservative schemes as in Theorem~\ref{thm:consentropy}.
+Moreover, $\bm{Q}_h$ can be used to produce high-order accurate approximations of derivatives, and can be used to construct entropy conservative schemes as in Theorem~\ref{thm:consentropy}.
 The extension to a general number of elements is done in a similar fashion.
@@
 
-It is possible to convert the semi-discrete entropy equality in Theorem~\ref{thm:consentropy} to a semi-discrete entropy inequality by adding appropriate interface dissipation terms, such as Lax-Friedrichs or matrix dissipation \cite{winters2017uniquely}.
+It is possible to convert the semidiscrete entropy equality in Theorem~\ref{thm:consentropy} to a semidiscrete entropy inequality by adding appropriate interface dissipation terms, such as Lax-Friedrichs or matrix dissipation \cite{winters2017uniquely}.
 We note that these terms must be computed in terms of $\tilde{\bm{u}}_f$ in order to ensure a discrete dissipation of entropy \cite{chen2017entropy, chan2017discretely}.
@@
 
-\section{Extension to higher dimensions and non-affine meshes}
+\section{Extension to higher dimensions and nonaffine meshes}
 \label{sec:2}
@@
 
-Let the domain now be decomposed into non-overlapping elements $\Omega^k$, such that $\Omega^k$ is the image of $\hat{\Omega}$ under a degree $N$ polynomial mapping $\bm{\Phi}^k$.
+Let the domain now be decomposed into nonoverlapping elements $\Omega^k$, such that $\Omega^k$ is the image of $\hat{\Omega}$ under a degree $N$ polynomial mapping $\bm{\Phi}^k$.
 We define geometric terms ${G}^k_{ij} = J^k\pd{\hat{x}_j}{\bm{x}_i}$ as scaled derivatives of reference coordinates $\hat{\bm{x}}$ with respect to physical coordinates $\bm{x}$.
@@
 where $(N+1)^2 + 4(N+1)$ is the total number of volume and face points.  
-We then have the following theorem on the semi-discrete conservation of entropy:
+We then have the following theorem on the semidiscrete conservation of entropy:
 \begin{theorem}
@@
 
-It is possible to ensure the satisfaction of a discrete GCL by using a sub-parametric polynomial geometric mapping.
+It is possible to ensure the satisfaction of a discrete GCL by using a subparametric polynomial geometric mapping.
 Let $N_{\rm geo}$ denote the degree of a polynomial geometric mapping.
@@
 Numerical results in 1D are similar to those presented in \cite{chan2017discretely}.
-Thus, we focus on two and three dimensional problems and comparisons of entropy stable GLL and Gauss collocation schemes.
+Thus, we focus on two- and three- dimensional problems and comparisons of entropy stable GLL and Gauss collocation schemes.
 
@@
 
-We begin by examining high order convergence of the proposed methods in two dimensions using the isentropic vortex problem \cite{shu1998essentially, crean2017high}.  The analytical solution is 
+We begin by examining high-order convergence of the proposed methods in two dimensions using the isentropic vortex problem \cite{shu1998essentially, crean2017high}.  The analytical solution is 
 \begin{align}
@@
 However, GLL and Gauss collocation differ more significantly on curved meshes.
-For both moderately and heavily warped meshes, the errors for a degree $N$ Gauss collocation scheme are nearly identical to errors for a higher order GLL collocation scheme of degree $(N+1)$.
-These results are in line with numerical experiments in \cite{parsani2016entropy}, which show that GLL collocation schemes lose one order of convergence in the $L^2$ norm on unstructured non-uniform meshes.
+For both moderately and heavily warped meshes, the errors for a degree $N$ Gauss collocation scheme are nearly identical to errors for a higher-order GLL collocation scheme of degree $(N+1)$.
+These results are in line with numerical experiments in \cite{parsani2016entropy}, which show that GLL collocation schemes lose one order of convergence in the $L^2$ norm on unstructured nonuniform meshes.
 Both results show that increasing quadrature accuracy significantly reduces the effect of polynomial aliasing due to curved meshes and spatially varying geometric terms.
@@
 On curved meshes, this requires that the mesh is sufficiently fine to resolve both the solution and geometric mapping.
-The results in Figure~\ref{fig:err2d} suggest that, compared to Gauss collocation schemes, under-integrated GLL collocation schemes require a finer mesh resolution to reach the asymptotic regime.
+The results in Figure~\ref{fig:err2d} suggest that, compared to Gauss collocation schemes, underintegrated GLL collocation schemes require a finer mesh resolution to reach the asymptotic regime.
 
@@
 
-We also compared $L^2$ errors for both isoparametric and sub-parametric geometric mappings.
+We also compared $L^2$ errors for both isoparametric and subparametric geometric mappings.
 In both cases, the discrete GCL is satisfied.
-For sub-parametric mappings, we chose the degree of approximation of geometry $N_{\rm geo} = \left\lfloor\frac{N}{2} \right\rfloor + 1$, such that the geometric terms are computed exactly and the GCL is satisfied by default.
+For subparametric mappings, we chose the degree of approximation of geometry $N_{\rm geo} = \left\lfloor\frac{N}{2} \right\rfloor + 1$, such that the geometric terms are computed exactly and the GCL is satisfied by default.
 This test is intended to address the fact that the GCL-preserving interpolation of Kopriva introduces a small approximation error, as the geometric terms are no longer exact.
-For these sub-parametric mappings, the gap between GLL and Gauss collocation widens slightly at $N=2$.
-However, the results for sub-parametric mappings are nearly identical to the results in the isoparametric case for higher polynomial degrees.
+For these subparametric mappings, the gap between GLL and Gauss collocation widens slightly at $N=2$.
+However, the results for subparametric mappings are nearly identical to the results in the isoparametric case for higher polynomial degrees.
 
@@
 %It is known that the Lax-Friedrichs flux is fairly dissipative compared to more nuanced fluxes \cite{harten1983upstream}, though this dissipation is less pronounced at higher polynomial degrees in practice.
-Figure~\ref{fig:shockvort} shows density solutions for $N=4$ and $h = 1/100$ Gauss collocation schemes using a non-dissipative entropy conservative flux, a dissipative Lax-Friedrichs flux, and a matrix dissipation flux.
+Figure~\ref{fig:shockvort} shows density solutions for $N=4$ and $h = 1/100$ Gauss collocation schemes using a nondissipative entropy conservative flux, a dissipative Lax-Friedrichs flux, and a matrix dissipation flux.
 In all cases, the vortex passes through the shock stably without the use of additional slope limiting, filtering, or artificial viscosity.
@@
 
-We note that this experiment also verifies that entropy stable decoupled SBP schemes (including the over-integrated case \cite{chan2017discretely}) are compatible with entropy stable wall boundary conditions.
-As far as the authors know, the stable and high order accurate imposition of such boundary conditions for existing GSBP couplings described in \cite{crean2017high} and Section~\ref{sec:gsbpsat} remains an open problem.
+We note that this experiment also verifies that entropy stable decoupled SBP schemes (including the overintegrated case \cite{chan2017discretely}) are compatible with entropy stable wall boundary conditions.
+As far as the authors know, the stable and high-order accurate imposition of such boundary conditions for existing GSBP couplings described in \cite{crean2017high} and Section~\ref{sec:gsbpsat} remains an open problem.
 
@@
 In the absence of viscosity, the Taylor--Green vortex develops successively smaller scales as time increases.
-As a result, the solution is guaranteed to contain under-resolved features after a sufficiently large time.
+As a result, the solution is guaranteed to contain underresolved features after a sufficiently large time.
 We study the evolution of kinetic energy $\kappa(t)$
@@
 Lax-Friedrichs dissipation is used for all simulations.
-For both affine and curved meshes, the presented Gauss collocation schemes remain stable in the presence of highly under-resolved solution features.
+For both affine and curved meshes, the presented Gauss collocation schemes remain stable in the presence of highly underresolved solution features.
 Kinetic energy dissipation rates for both GLL and Gauss collocation are qualitatively similar and are consistent with existing results in the literature for the inviscid Taylor-Green vortex \cite{gassner2016split,chan2018discretely}.
@@
 The main computational costs associated with entropy stable schemes are volume operations, which include evaluations of two-point fluxes and applications of one-dimensional differentiation and interpolation matrices.
-We do not count inter-element communication or flux computations, as these are typically less expensive than volume operations (especially for higher polynomial degrees).
+We do not count interelement communication or flux computations, as these are typically less expensive than volume operations (especially for higher polynomial degrees).
 The total number of flux computations for each scheme can also be reduced using the symmetry of $\bm{f}_S$; however, this does not affect relative differences in the number of two-point flux evaluations between GLL, Gauss, and staggered grid schemes.
@@
 
-This work shows how to construct efficient entropy stable high order Gauss collocation DG schemes on quadrilateral and hexahedral meshes.
-Key to the construction of efficient methods are decoupled SBP operators, which deliver entropy stability and high order accuracy while maintaining compact inter-element coupling terms.
+This work shows how to construct efficient entropy stable high-order Gauss collocation DG schemes on quadrilateral and hexahedral meshes.
+Key to the construction of efficient methods are decoupled SBP operators, which deliver entropy stability and high-order accuracy while maintaining compact interelement coupling terms.
 These operators are also compatible with existing entropy stable methods for applying interface dissipation \cite{winters2017uniquely} or imposing boundary conditions.
-Numerical experiments demonstrate both the stability and high order accuracy of the proposed Gauss collocation schemes on both affine and curvilinear meshes.
+Numerical experiments demonstrate both the stability and high-order accuracy of the proposed Gauss collocation schemes on both affine and curvilinear meshes.
 \note{We note that, while the numerical experiments presented here consider only mapped Cartesian domains, the method is also applicable to complex geometries, and future work will focus on studying the performance of such methods on curvilinear quadrilateral and hexahedral unstructured meshes.}

### 2019-09-06 docs/ESDG_ROM/main.tex
updating docs

@@
 \usepackage{amsmath,amssymb,amsfonts,amsthm}
+\usepackage{hhline}
 \theoremstyle{definition}
@@
 Reduced order models of nonlinear conservation laws in fluid dynamics do not typically inherit stability properties of the full order model.
-We construct projection-based reduced models of nonlinear conservation laws based on flux differencing which retain accuracy while inheriting a semi-discrete entropy inequality independently of discretization resolutions and parameters.
+We introduced projection-based reduced models of nonlinear conservation laws based on flux differencing which retain accuracy while inheriting a semi-discrete entropy inequality independently of the choice of basis, parameters, and discretization resolution.
 \end{abstract}
@@
 
-For certain PDEs, ``structure-preserving'' ROMs provide robustness and stability by preserving energetic properties of the full system at the discrete level.
+For certain PDEs, ``structure-preserving'' ROMs provide robustness and stability by reproducing energetic properties of the full system at the discrete level.
 ROMs which retain either Lagrangian \cite{lall2003structure, carlberg2015preserving} or Hamiltonian structure \cite{benner2012interpolation, gugercin2012structure, peng2016symplectic, chaturantabut2016structure, gong2017structure, afkham2017structure, afkham2018structure} have been constructed by combining Galerkin projection with an appropriate formulation of the full order model, and similar energy-conserving ROMs have been constructed for the incompressible Navier-Stokes equations \cite{farhat2014dimensional, farhat2015structure}.
@@
 
-The construction of stable ROMs for the compressible Navier-Stokes equations is complicated by their complex structure.
+The construction of stable ROMs for the compressible Navier-Stokes equations is further complicated by their non-trivial structure.
 Thus, practitioners often develop structure-preserving entropy stable ROMs for simplifications of the underlying PDE (such as the time-dependent \textit{linearized} compressible Navier-Stokes equations \cite{barone2009stable, kalashnikova2010stability, amsallem2012stabilization, kalashnikova2014construction}) and extrapolate such models to the full nonlinear equations.
@@
 In this work, we approximate nonlinear terms using hyper-reduction techniques based on empirically computed quadrature rules \cite{an2008optimizing, hernandez2017dimensional}.
-The hyper-reduction approach incurs approximation error; however, it is allows for the generalization of the split forms in \cite{maboudi2018conservative} to nonlinearly entropy stable formulations.
-
-%
-%Hyper-reduction methods
-%\begin{itemize}
-%\item Interpolatory: DEIM and QR-DEIM \cite{chaturantabut2010nonlinear, drmac2016new}, gappy POD \cite{bui2004aerodynamic, carlberg2011efficient}
-%\item Element-based approaches: ECSW \cite{farhat2014dimensional, farhat2015structure}, EQP \cite{ryu2015extensions, yano2019lp, patera2017lp}
-%\item Empirical cubature: \cite{an2008optimizing, hernandez2017dimensional}.  
-%\end{itemize}
-%
-%Structure-preserving ROMs: 
-%\begin{itemize}
-%\item Linear Hamlitonian \cite{beattie2009interpolatory} 
-%\item Nonlinear Hamiltonian systems: Gugercin \cite{gugercin2012structure}, Carlberg \cite{carlberg2015preserving}, Chaturantabut \cite{chaturantabut2016structure}, Gong \cite{gong2017structure}, Afkham et al.\ \cite{afkham2017structure, afkham2018structure}
-%\item Conservation through explicit constraints over sub-meshes \cite{carlberg2017structure, carlberg2018conservative}
-%\end{itemize}
-
-In this work, we directly constructing entropy stable ROMs for nonlinear conservation laws through a modified Galerkin projection of an appropriate full order model.
+The hyper-reduction approach incurs approximation error and additional computational cost; however, it is allows for the generalization of the split forms in \cite{maboudi2018conservative} to nonlinearly entropy stable formulations.
+We directly construct entropy stable ROMs for nonlinear conservation laws by combining hyper-reduction with a modified Galerkin projection of an appropriate full order model.
 The approach taken in this paper combines techniques from entropy stable finite volume schemes \cite{tadmor1987numerical, tadmor2003entropy} and entropy stable summation-by-parts (SBP) discretizations \cite{fisher2013high, carpenter2014entropy, chen2017entropy, crean2018entropy, chan2017discretely, chan2019skew} to produce entropy stable reduced order models.
@@
 The paper is organized as follows.
-Section~\ref{sec:1} discusses the entropy inequality associated with systems of nonlinear conservation laws.
+Section~\ref{sec:1} introduces entropy inequalities associated with systems of nonlinear conservation laws.
 Section~\ref{sec:2} introduces a full order model on 1D periodic domains based on entropy stable finite volume schemes.
@@
 \label{sec:1}
-Let $\Omega$ denote some domain with boundary $\partial \Omega$.  General nonlinear conservation laws are expressed as a system of partial differential equations (PDEs) 
+Let $\Omega$ denote some domain with boundary $\partial \Omega$.  Nonlinear conservation laws are expressed as a system of partial differential equations (PDEs) 
 \begin{equation}
@@
 \end{equation}
-where $\bm{u}\in \mathbb{R}^n$ are the conservative variables, $\bm{f}_j$ are nonlinear fluxes, and $\bm{v}(\bm{u})$ are the \textit{entropy variables}.
+where $\bm{u}\in \mathbb{R}^n$ are the conservative variables, $\bm{f}_i$ are nonlinear fluxes, and $\bm{v}(\bm{u})$ are the \textit{entropy variables}.
 By multiplying (\ref{eq:nonlineqs}) by the entropy variables, viscosity solutions \cite{oleinik1957discontinuous, kruvzkov1970first}  of popular fluid systems (e.g., shallow water, compressible Euler and Navier-Stokes \cite{hughes1986new, chen2017entropy}) can be shown to satisfy
@@
 
-The reduced order models in this work are constructed using full order models constructed from discretely entropy stable finite volume methods \cite{tadmor1987numerical}.
+The reduced order models in this work are constructed using full order models based on entropy stable finite volume methods \cite{tadmor1987numerical}.
 For simplicity, we illustrate the construction of full and reduced order models on a periodic 1D domain.
@@
 
-A key ingredient of the full and reduced order models considered in this work are \textit{entropy conservative} numerical fluxes.
+A key ingredient of the full and reduced order models considered in this work are \textit{entropy conservative} finite volume numerical fluxes.
 Let $\bm{u}_L, \bm{u}_R$ denote left and right states.
@@
 For a system of $n$ nonlinear conservation laws, all matrices are treated in a Kronecker product fashion.
-Let the vector $\bm{u}_h$ now correspond to the concatenated vector of components of the solution $\bm{u}_h = [\bm{u}_h^1, \ldots, \bm{u}_h^n]$.
+Let the vector $\bm{u}_h$ now correspond to the concatenated vector of components of the solution $\bm{u}_h = [\bm{u}_h^1, \ldots, \bm{u}_h^n]^T$.
 Then, the matrix formulation of (\ref{eq:nonlineqs1d}) is given by
@@
 \end{bmatrix}, \qquad 
-\LRp{\bm{F}_{i}}_{jk} = \bm{f}_{i,S}\LRp{\LRp{\bm{u}_h}_j, \LRp{\bm{u}_h}_k}.
+\LRp{\bm{F}_{i}}_{jk} = \bm{f}_{i,S}\LRp{\LRp{\bm{u}_h}_j, \LRp{\bm{u}_h}_k}, \qquad \LRp{\bm{u}_h}_j = \begin{bmatrix}\LRp{\bm{u}^1_h}_j & \LRp{\bm{u}^2_h}_j & \ldots & \LRp{\bm{u}^n_h}_j\end{bmatrix}^T.
 \label{eq:essys}
@@
 Here, $\bm{F}$ is a block diagonal matrix where the $i$th block corresponds to the evaluation of the $i$th component of the numerical flux $\bm{f}_{i,S}(\bm{u}_L,\bm{u}_R)$.
-From this point on, we drop explicit references to the Kronecker product to simplify notation, so that (\ref{eq:es}) applies to both scalar nonlinear conservation laws and systems of nonlinear conservation laws.
+From this point on, we drop explicit references to components and the Kronecker product to simplify notation, so that (\ref{eq:es}) applies to both scalar equations and systems.
 
@@
 \end{equation}
-The proof is given in \cite{tadmor2006entropy}, which we reproduce here.  We rewrite the viscous term as
+The proof is similar to the one given in \cite{tadmor2006entropy}.  We rewrite the viscous term as
 \[
@@
 \]
-where $\LRp{\pd{\bm{u}}{\bm{v}}}_{i,i-1}$ is the positive-definite Jacobian matrix evaluated at some state between $(\bm{u}_h)_i$ and $(\bm{u}_h)_j$.
-Substituting these expressions into the viscous term yields
+where $\LRp{\pd{\bm{u}}{\bm{v}}}_{i,i-1}$ is the Jacobian matrix evaluated at some state between $(\bm{u}_h)_i$ and $(\bm{u}_h)_j$.
+Since the Jacobian is related to the Hessian of the entropy through $\pdn{S(\bm{u})}{\bm{u}}{2} = \pd{\bm{v}}{\bm{u}} = \LRp{\pd{\bm{u}}{\bm{v}}}^{-1}$, it is positive definite so long as $S(\bm{u})$ is convex.
+Substituting these expressions into the viscous terms and using positive definiteness of $\pd{\bm{u}}{\bm{v}}$  yields
 \begin{gather}
@@
 We can now formulate a reduced basis approximation on periodic domains using the entropy stable full order models described in the previous section.
-Boundary conditions will be addressed later in this work.
+Boundary conditions will be addressed later in Section~\ref{sec:6}.
 We also note that this reduced order model is not practical, since the cost of evaluating nonlinear terms scales with the size of the full order model.
@@
 We motivate this modification by first considering the time derivative term.
-Because of the Galerkin projection step, we can no longer test with the vector of entropy variables.
+Because of the Galerkin projection, we can no longer directly test with the vector of entropy variables, which may not lie  in the span of the reduced basis functions.
 However, we can test with an appropriate \textit{projection} of the entropy variables.
@@
 &= 
- \bm{v}\LRp{\bm{V}\bm{u}_N}^T \td{\bm{V}\bm{u}_N}{t} = \bm{1}^T\td{S(\bm{V}\bm{u}_N)}{t}.\nonumber
+ \bm{v}\LRp{\bm{V}\bm{u}_N}^T \td{\LRp{\bm{V}\bm{u}_N}}{t} = \bm{1}^T\td{S(\bm{V}\bm{u}_N)}{t}.\nonumber
 \end{align}
-Define the grid values of the projected entropy variables as $\tilde{\bm{v}}$ 
-\[
-\tilde{\bm{v}} = \bm{V}\bm{v}_N = \bm{V}\bm{V}^{\dagger} \bm{v}\LRp{\bm{V}\bm{u}_N}.
-\]
-It is possible to repeat the steps of the proof of Theorem~\ref{thm:ecfom} up to the point when we invoke the entropy conservation condition of (\ref{eq:esflux})
+The remainder of the proof requires showing that 
+\[
+\bm{v}_N^T\bm{V}^T\LRp{\bm{Q}\circ \bm{F}}\bm{1} = \tilde{\bm{v}}^T\LRp{\bm{Q}\circ \bm{F}}\bm{1}  = 0, \qquad \tilde{\bm{v}} = \bm{V}\bm{v}_N = \bm{V}\bm{V}^{\dagger} \bm{v}\LRp{\bm{V}\bm{u}_N},
+\]
+where we have introduced the grid values of the projected entropy variables as $\tilde{\bm{v}}$.
+We can repeat the steps of the proof of Theorem~\ref{thm:ecfom} up to the point when we invoke the entropy conservation condition of (\ref{eq:esflux})
 \begin{align*}
@@
 To reduce the cost of evaluating nonlinear terms, we introduce a second \textit{hyper-reduction} step \cite{ryckelynck2009hyper}.
-Here, a vector of nonlinear function evaluations is replaced by an approximation constructed from nonlinear function evaluations at a smaller set of reduced points \cite{barrault2004empirical, bui2004aerodynamic, chaturantabut2010nonlinear, farhat2015structure, drmac2016new, hernandez2017dimensional, yano2019lp}.
+Here, a vector of nonlinear function evaluations is approximated by a vector constructed using nonlinear function evaluations at a subset of points \cite{barrault2004empirical, bui2004aerodynamic, chaturantabut2010nonlinear, farhat2015structure, drmac2016new, hernandez2017dimensional, yano2019lp}.
 In this work, we utilize a sampling and weighting hyper-reduction strategy \cite{farhat2015structure, hernandez2017dimensional, yano2019lp} which enables proofs of discrete entropy stability.
@@
 \]
-Here, $g(\bm{u})$ denotes a nonlinear function, $\mathcal{I}$ denotes a subset of $N_s$ row indices, $\bm{V}\LRp{\mathcal{I},:}$ denotes the matrix consisting of the $N_s$ sampled rows of $\bm{V}$, and $\bm{W} = {\rm diag}(\bm{w})$ is a $N_s\times N_s$ diagonal matrix of positive weights.
+Here, $g(\bm{u})$ denotes a nonlinear function, $\mathcal{I}$ denotes a subset of $N_s$ row indices corresponding to sampled points, $\bm{V}\LRp{\mathcal{I},:}$ denotes the sub-matrix consisting of the $N_s$ sampled rows of $\bm{V}$, and $\bm{W} = {\rm diag}(\bm{w})$ is a $N_s\times N_s$ diagonal matrix of positive weights.
 We will describe algorithms for computing hyper-reduced points and weights in Section~\ref{sec:hyperreducalgo}.
@@
 Rather than directly hyper-reducing the full order matrix $\bm{Q}$, we first construct a compressed ``modal'' matrix, then combine this with a projection operator based on the hyper-reduced points.
-
 In the first step, we construct a compressed intermediate operator by combining  Galerkin projection with the ``expanded basis'' approach of \cite{hernandez2017dimensional}.
-Let $\bm{V}_t$ denote a \textit{test} basis through which we will extract the action of $\bm{Q}$.
+Let $\bm{V}_t$ denote a \textit{test} basis through which to extract the action of $\bm{Q}$.
 We define the intermediate reduced operator
@@
 \begin{assumption}
-Let $\mathcal{I}$ denote the index set of hyper-reduced points.  We assume that the row-sampled test matrix $\bm{V}_t\LRp{\mathcal{I},:}$ is non-singular.  
+Let $\mathcal{I}$ denote the index set of hyper-reduced points.  We assume that the hyper-reduced test mass matrix $\bm{V}_t\LRp{\mathcal{I},:}^T\bm{W}\bm{V}_t\LRp{\mathcal{I},:}$ is non-singular.
+%row-sampled test matrix $\bm{V}_t\LRp{\mathcal{I},:}$ is non-singular.  
  \label{ass:quad}
 \end{assumption}
-Note that if Assumption~\ref{ass:quad} holds, then the matrix $\bm{V}\LRp{\mathcal{I},:}$ is also non-singular, since the column space of $\bm{V}$ is contained in the column space of $\bm{V}_t$.
-Then, we can define the projection matrix $\bm{P}_t$ onto the test basis as follows:
-\[
+Note that if Assumption~\ref{ass:quad} holds, then the mass matrix $\bm{V}\LRp{\mathcal{I},:}^T\bm{W}\bm{V}\LRp{\mathcal{I},:}$ is also non-singular, since the column space of $\bm{V}$ is contained in the column space of $\bm{V}_t$.
+
+If Assumption~\ref{ass:quad} holds, we can define the projection matrix $\bm{P}_t$ onto the test basis as follows:
+\begin{equation}
 \bm{P}_t = \LRp{\bm{V}_t\LRp{\mathcal{I},:}^T\bm{W}\bm{V}_t\LRp{\mathcal{I},:}}^{-1}\bm{V}_t\LRp{\mathcal{I},:}^T\bm{W}.
-\]
+\label{eq:Pt}
+\end{equation}
 The projection matrix $\bm{P}_t$ maps values at hyper-reduced points to coefficients in the test basis, and is well-defined under Assumption~\ref{ass:quad}.
@@
 \begin{lemma}
-Assume that $\bm{1}$ is exactly representable in the test basis and that the columns of the submatrix $\bm{V}_t\LRp{\mathcal{I},:}$ are linearly independent.  Define the matrix $\bm{Q}_t$
+Suppose that $\bm{1}$ is exactly representable in the test basis and that Assumption~\ref{ass:quad} holds.  Define the matrix $\bm{Q}_t$
 \begin{equation}
@@
 Since $\bm{Q}$ is skew-symmetric, the matrix $\bm{Q}_t$ is skew-symmetric by construction as well.
-By (\ref{eq:preproduce}), if $\bm{1}$ is exactly representable in the test basis (such that for some coefficient vector $\bm{e}$ $\bm{1} = \bm{V}_t\LRp{\mathcal{I},:}\bm{e}$ ), then $\bm{P}_t\bm{1} = \bm{e}$.
-Then, $\bm{Q}_t\bm{1} = \bm{P}_t^T\hat{\bm{Q}}_t\bm{e} = \bm{0}$, since $\hat{\bm{Q}}_t$ exactly differentiates elements in the test basis.
+By (\ref{eq:preproduce}), if $\bm{1}$ is exactly representable in the test basis (such that $\bm{1} = \bm{V}_t\LRp{\mathcal{I},:}\bm{e}$ for some coefficient vector $\bm{e}$  ), then $\bm{P}_t\bm{1} = \bm{e}$.
+Then, $\bm{Q}_t\bm{1} = \bm{P}_t^T\hat{\bm{Q}}_t\bm{e} = \bm{0}$, since $\hat{\bm{Q}}_t$ exactly differentiates elements of the test basis.
 \end{proof}
+
+\begin{remark}
+The projection matrix $\bm{P}_t$ can also be replaced, for example, by the gappy POD projection $\LRp{\bm{V}_t\LRp{\mathcal{I},:}}^{\dagger}$ \cite{willcox2006unsteady,astrid2008missing}.
+We do not observe significant differences in numerical results when doing so.
+In principle, any matrix operator which maps from nodal values to modal coefficients and reproduces the test basis exactly can be used for $\bm{P}_t$.
+In this work, we restrict ourselves to the projection matrix defined by (\ref{eq:Pt}), and will investigate other definitions of $\bm{P}_t$ in future work.
+\end{remark}
 
@@
 The solution $u(x,t)$ is a decaying stationary shock which is anti-symmetric across the origin at all times $t$.
-The POD modes are thus also anti-symmetric across the origin; however, their derivatives $\bm{Q}\bm{V}$ are symmetric across the origin, as shown in Figure~\ref{fig:modeQ}.
-As a result, since the entries of $\bm{V}^T\bm{Q}\bm{V}$ are inner products of symmetric and anti-symmetric functions, $\bm{V}^T\bm{Q}\bm{V} \approx \bm{0}$.
+The POD modes are thus also anti-symmetric across the origin; however, their derivatives $\bm{Q}\bm{V}$ are nearly symmetric across the origin, as shown in Figure~\ref{fig:modeQ}.
+As a result, since the entries of $\bm{V}^T\bm{Q}\bm{V}$ are inner products of nearly symmetric and anti-symmetric functions, $\bm{V}^T\bm{Q}\bm{V} \approx \bm{0}$.
 \begin{figure}[!h]
@@
 \subfloat[Mode derivatives $\bm{Q}\bm{V}$]{\includegraphics[width=.32\textwidth]{figs/burg3modes_diff.png}}
-\caption{Solution snapshots, three POD modes, and mode derivatives for a shock solution of the viscous periodic Burgers' equation. }
+\caption{Solution snapshots, three POD modes, and mode derivatives for a shock solution of the periodic Burgers' equation. }
 \label{fig:modeQ}
@@
 The index set is constructed in a greedy fashion by selecting the row index of $\bm{V}_{\rm target}$ which is most positively parallel to the residual, then computing the weight vector which minimizes the residual through a (non-negative) least squares solve.
+The iteration terminates when the norm of the residual is smaller than some user-defined tolerance $tol$.
+In this work, we utilize an $N$-mode POD basis, and $tol$ is determined by the singular values of the associated snapshot matrix.
+Suppose there are $M$ total singular values of the snapshot matrix; then, we define $tol$ as
+\[
+tol = \sqrt{\frac{\sum_{j=N+1}^M \sigma_j^2}{\sum_{j=1}^M \sigma_j^2}}.
+\]
 
 \begin{algorithm}[!h]
-\caption{Compute hyper-reduction s.t.\ $\bm{V}_{\rm target}^T\bm{w}_{\rm target} \approx \bm{V}_{\rm target}\LRp{\mathcal{I},:}^T\bm{w}$.}
+\caption{Compute hyper-reduction s.t.\ $\bm{V}_{\rm target}^T\bm{w}_{\rm target} \approx \bm{V}_{\rm target}\LRp{\mathcal{I},:}^T\bm{w}$ (from \cite{hernandez2017dimensional}).}
 \begin{algorithmic}[1]
@@
 \STATE \textbf{Output}: $\mathcal{I}$, $\bm{w}$.
-\STATE Set $\bm{b} = \bm{V}_{\rm target}^T\bm{w}_{\rm target}$.
-\STATE Initialize residual $\bm{r} = \bm{b}$, $\mathcal{I} = \emptyset$.
+\STATE Set $\bm{b} = \bm{V}_{\rm target}^T\bm{w}_{\rm target}$, initialize residual $\bm{r} = \bm{b}$, $\mathcal{I} = \emptyset$.
 \WHILE {$\nor{\bm{r}}/\nor{\bm{b}} > tol$}
@@
 For all numerical experiments, we set $\bm{w}_{\rm target} = \Delta x \bm{1}$ and set $\bm{V}_{\rm target}$ to be the span of products of POD functions $\bm{V}$, such that
-\[
+\begin{equation}
 \mathcal{R}\LRp{\bm{V}_{\rm target}} = {\rm span}\LRc{\bm{V}(:,i) \circ \bm{V}(:,j), \quad i,j = 1,\ldots, N}.
-\]
+\label{eq:Vtarget}
+\end{equation}
 This choice of target space ensures that the mass matrix $\Delta x \bm{V}^T\bm{V}$ is accurately approximated.
-In practice, since many of the pointwise vector products $\bm{V}(:,i) \circ \bm{V}(:,j)$ are linearly dependent, we replace $\bm{V}_{\rm target}$ by a dimensionally reduced matrix computed using, for example, another application of the SVD.
-%a randomized SVD \cite{halko2011finding, rsvd}.
-
-\begin{remark}
-We note that the target space is designed to accurately approximate the mass matrix $\Delta x \bm{V}^T\bm{V}$, but does not take the test basis $\bm{V}_{t}$ into account.
-It is thus possible to construct a hyper-reduced quadrature which accurately integrates the mass matrix but violates Assumption~\ref{ass:quad}.
-We have not encountered this situation in our experiments; however, one can ensure that Assumption~\ref{ass:quad} is satisfied by checking if $\bm{V}_t\LRp{\mathcal{I},:}$ is non-singular and selecting additional points based on $\bm{V}_t$ as necessary.
-\end{remark}
+In practice, since many of the pointwise vector products $\bm{V}(:,i) \circ \bm{V}(:,j)$ are linearly dependent, we replace $\bm{V}_{\rm target}$ by a dimensionally reduced matrix \cite{hernandez2017dimensional}.
+In this work, we compute this second reduction using another application of the SVD.
+Let $\mu_{1}, \ldots, \mu_{N_t}$ denote the singular values of $\bm{V}_{\rm target}$, and define the energy residual of the first $i$ modes
+\begin{equation}
+E_i = \sqrt{\frac{\sum_{j=i+1}^{N_t} \mu_j^2}{\sum_{j=1}^{N_t} \mu_j^2}}.  
+\label{eq:svdenergy}
+\end{equation}
+Note that $E_1 \geq E_2 \geq \ldots \geq E_{N_t}$.
+We replace the matrix $\bm{V}_{\rm target}$ with its $k$ leading left singular vectors, where $k$ is the smallest index such that $E_{k} \leq tol$.
+%\note{Add more clarifications on additional SVD for target matrix: tol or machine precision for floor?
+Use of relative SVD energy vs absolute.
+In our numerical experiments, all SVD truncations are performed such that the relative SVD energy is less than the tolerance defined by the number of modes $N$.}
+
+\subsection{Conditioning of the test mass matrix}
+\label{sec:condtest}
+We note that the target space is designed to accurately approximate the mass matrix $\Delta x \bm{V}^T\bm{V}$, and does not take the test basis matrix $\bm{V}_{t}$ into account.
+This is because incorporating the test basis by approximating either $\Delta x \bm{V}^T\bm{V}_{t}$ or $\Delta x \bm{V}_{t}^T\bm{V}_{t}$ produces a significantly larger number of hyper-reduced points without a noticable increase in accuracy.
+However, because the hyper-reduction strategy does not account for $\bm{V}_{t}$, it is possible to construct a hyper-reduced quadrature which accurately integrates the mass matrix but violates Assumption~\ref{ass:quad} by producing a singular test mass matrix $\bm{V}_t\LRp{\mathcal{I},:}^T\bm{W}\bm{V}_t\LRp{\mathcal{I},:}$.
+The challenge is then to construct a hyper-reduced volume quadrature which 1) accurately approximates the reduced basis mass matrix $\Delta x \bm{V}^T\bm{V}$ while 2) ensuring that the hyper-reduced test mass matrix is non-singular.
+We address this challenge by computing the spectra of the test mass matrix  and adding additional ``stabilizing'' points if the condition number is larger than a specified tolerance.
+
+Let $\bm{z}_j$ denote the $N_z$ eigenvectors corresponding to zero (small) eigenvalues of the hyper-reduced test mass matrix.
+The entries of $\bm{z}_j$ correspond to vectors of coefficients in the test basis $\bm{V}_t$.
+Let $\bm{Z}$ denote the matrix whose columns are point values of each eigenvector
+\[
+\bm{Z} = \begin{bmatrix}
+\bm{V}_t\bm{z}_1 & \ldots & \bm{V}_t\bm{z}_{N_z}
+\end{bmatrix}.
+\]
+To ensure that the hyper-reduced test mass matrix is non-singular, we will employ a greedy heuristic based on adding additional points to approximate $\Delta x\bm{Z}^T\bm{Z}$.
+Suppose that we have already computed hyper-reduced points and weights to approximate the reduced mass matrix $\Delta x\bm{V}^T\bm{V}$.
+Let $\mathcal{I}_{\bm{V}}$ denote the index set of these stabilizing points.
+We compute a second group of hyper-reduced points with index set $\mathcal{I}_{\bm{Z}}$  using the greedy hyper-reduction algorithm applied to the target space $\bm{Z}_{\rm target}$, which we define analogously to $\bm{V}_{\rm target}$ in (\ref{eq:Vtarget}).
+We note that, unlike the hyper-reduced points computed previously, we do not perform a dimensionality reduction on $\bm{Z}_{\rm target}$.
+
+Given $\mathcal{I}_{\bm{Z}}$, we then update the hyper-reduced index set as the union $\mathcal{I} = \mathcal{I}_{\bm{V}} \cup \mathcal{I}_{\bm{Z}}$, and compute a new set of hyper-reduced weights through the non-negative least squares solve
+\[
+\bm{w} = \argmin_{\bm{c} \geq 0} \frac{1}{2} \LRp{\nor{\bm{V}_{\rm target}\LRp{\mathcal{I},:}^T\bm{c} - \bm{b}}^2 + \alpha_Z\nor{\bm{Z}_{\rm target}\LRp{\mathcal{I},:}^T\bm{c} - \bm{d}}^2}, \qquad \bm{d} = \bm{Z}_{\rm target}^T\bm{w}_{\rm target},
+\]  
+where $\alpha_Z > 0$ is some scaling parameter which controls whether the weights $\bm{w}$ more accurately approximate the integration of the target space $\bm{V}_{\rm target}$ for the mass matrix, or $\bm{Z}_{\rm target}$ for the null space of the test mass matrix.
+We set $\alpha_Z = 10^{-2}$ in all experiments, which is empirically observed to control the condition number of the test mass matrix without significantly impacting the accuracy to which the reduced mass matrix $\Delta x\bm{V}^T\bm{V}$ is approximated.
+
+For the numerical experiments in this paper, it was sufficient to add one single set of stabilizing hyper-reduced points.
+However, this procedure can be repeated multiple times to further improve the condition number of the hyper-reduced test mass matrix.
+
+%Our approach involves three steps.
+First, we select hyper-reduced points and weights to directly approximate $\Delta x\bm{V}^T\bm{V}$.
+Next, we independently select another set of points and weights to approximate the null space of the hyper-reduced test mass matrix.
+Finally, we combine these two point sets together and compute new weights through a non-negative least squares solve.
+
+% is non-singular and selecting additional points until $\bm{V}_t\LRp{\mathcal{I},:}$ is non-singular (using, for example, DEIM).
+New weights associated with additional points are computed using the same non-negative least squares algorithm.
+
+%\note{Add new algo for new Vtest points.}
 
@@
 
-To summarize, an entropy conservative reduced order model on periodic domains can be constructed using the following ingredients and steps:
+To summarize, an entropy conservative reduced order model on periodic domains requires the following offline steps:
 \begin{enumerate}
-\item Compute a reduced basis matrix $\bm{V}$ from snapshots of both the conservative variables and the entropy variables.  
-\item Construct ``test'' basis matrices $\bm{V}_t$ to produce a compressed ``modal'' differentiation matrix $\hat{\bm{Q}}_t = \bm{V}_t^T\bm{Q}\bm{V}_t$.
-\item Compute a hyper-reduced quadrature using Algorithm~\ref{alg:hr}.  
-\item Construct the hyper-reduced nodal differentiation matrix $\bm{Q}_t = \bm{P}_t^T\hat{\bm{Q}}_t\bm{P}_t$ using the projection matrix onto the test basis $\bm{P}_t$.
+\item Compute a reduced basis matrix $\bm{V}$ from snapshots of both conservative variables and entropy variables.  
+\item Use the ``test'' basis matrix $\bm{V}_t$ to construct a ``modal'' differentiation matrix $\hat{\bm{Q}}_t = \bm{V}_t^T\bm{Q}\bm{V}_t$.
+\item Compute a hyper-reduced quadrature using Algorithm~\ref{alg:hr}, adding stabilizing points as necessary to ensure that the test mass matrix $\bm{V}_t\LRp{\mathcal{I},:}^T\bm{W}\bm{V}_t\LRp{\mathcal{I},:}$ is non-singular.
+\item Construct the hyper-reduced nodal differentiation matrix $\bm{Q}_t = \bm{P}_t^T\hat{\bm{Q}}_t\bm{P}_t$ using the projection matrix $\bm{P}_t = \LRp{\bm{V}_t\LRp{\mathcal{I},:}^T\bm{W}\bm{V}_t\LRp{\mathcal{I},:}}^{-1}\bm{V}_t\LRp{\mathcal{I},:}^T\bm{W}$ onto the test basis.
 \end{enumerate}
-
-%: \note{summarize restrictions and assumptions on hyper-reduction, POD, and test basis.}
+Then, an entropy conservative reduced order model is given by the following theorem:
 \begin{theorem}
@@
 \end{gather}
-discretely conserves entropy 
-\[
-\Delta x \bm{1}^T\td{S\LRp{\bm{V}\bm{u}_N}}{t} = 0.  
+semi-discretely conserves the average entropy 
+\[
+\bm{1}^T\bm{W}\td{S\LRp{\bm{V}\LRp{\mathcal{I},:}\bm{u}_N}}{t} = 0.  
 \]
@@
 \begin{proof}
-Since the hyper-reduction preserves the skew-symmetry and zero row sums of $\bm{Q}_t$, the proof is identical in structure to the proof of Theorem~\ref{thm:esrom}.  
+Testing the time derivative with $\bm{v}_N$ yields
+\begin{gather*}
+\bm{v}_N^T\bm{M}_N\td{\bm{u}_N}{t} = \bm{v}\LRp{\bm{V}\LRp{\mathcal{I},:}\bm{u}_h}^T\bm{P}^T\bm{M}_N\td{\bm{u}_N}{t} = \bm{v}\LRp{\bm{V}\LRp{\mathcal{I},:}\bm{u}_h}^T \bm{W}\bm{V}\LRp{\mathcal{I},:}\bm{M}_N^{-1}\bm{M}_N\td{\bm{u}_N}{t}
+\\
+= \bm{v}\LRp{\bm{V}\LRp{\mathcal{I},:}\bm{u}_h}^T \bm{W} \bm{V}\LRp{\mathcal{I},:}\td{\bm{u}_N}{t} = \bm{v}\LRp{\bm{V}\LRp{\mathcal{I},:}\bm{u}_h}^T \bm{W} \td{\LRp{\bm{V}\LRp{\mathcal{I},:}\bm{u}_N}}{t} = \bm{1}^T\bm{W}\td{S\LRp{\bm{V}\LRp{\mathcal{I},:}\bm{u}_N}}{t},
+\end{gather*}
+where we have used that $\bm{W}$ is diagonal and the temporal chain rule in the final step.  
+Since the hyper-reduction preserves the skew-symmetry and zero row sums of $\bm{Q}_t$, the remainder of the proof consists of showing that $\bm{v}_N^T\bm{V}\LRp{\mathcal{I},:}^T\LRp{\bm{Q}_t \circ \bm{F}}\bm{1} = 0$, and is identical in structure to the proof of Theorem~\ref{thm:esrom}.
 \end{proof}
@@
 
-We note that the flux matrix $\bm{F}$ is not computed explicitly, but instead is usually formed on the fly.
-The computation of $\bm{F}$ constitutes a significant portion of the computational cost, and is typically ameliorated by exploiting the sparsity of $\bm{Q}$ for the full order model.
-However, since the hyper-reduced matrix $\bm{Q}_t$ is fully dense, the ROM (\ref{eq:hrrom}) requires $O(N_s^2)$ nonlinear flux evaluations, in contrast to the typical $O(N_s)$ nonlinear function evaluations required for a typical hyper-reduced model.
+\subsection{Online cost estimates for hyper-reduction}
+\label{sec:cost}
+In practice, the flux matrix $\bm{F}$ is not computed explicitly, but instead is usually formed on the fly by evaluating the entropy stable flux $\bm{f}_S$.
+Due to the complexity of the formulas for $\bm{f}_S$, the computation of $\bm{F}$ constitutes a significant portion of the computational cost.
+For the full order model, this cost is typically ameliorated by exploiting the sparsity of $\bm{Q}$.
+However, since the hyper-reduced matrix $\bm{Q}_t$ is fully dense, the nonlinear convective term $\bm{Q}\circ\bm{F}$ in (\ref{eq:hrrom}) requires $O(N_s^2)$ nonlinear flux evaluations.
+In contrast, typical hyper-reduction approaches target a nonlinear \textit{vector}, which requires only $O(N_s)$ nonlinear function evaluations.
+We emphasize that this $O(N_s^2)$ online cost is specific only to the convective nonlinearity, and not to the hyper-reduced treatment of other nonlinear terms such as viscosity.
+
+While this cost is trivially parallelizable, it is a significant increase in online costs compared to traditional hyper-reduction procedures.
+However, we note that the cost can be comparable to techniques for exactly treating higher order polynomial nonlinearities for a large number of modes \cite{maboudi2018conservative}.
+For example, suppose one wishes to compute $\bm{V}^T f\LRp{\bm{V}\bm{u}_N}$, where $f(x) = x^m$ for $m \geq 1$ and $m$ integer.
+We can rewrite this as
+\[
+\bm{V}^T f(\bm{u}) = \bm{V}^T \diag{\bm{V}\bm{u}_N}^{m-1} \bm{V}\bm{u}_N.
+\]
+Suppose first that $m = 2$, such that the nonlinearity is quadratic.
+Define $\bm{V}_j = \diag{\bm{V}\LRp{:,j}}$ as the diagonal matrix with the $j$th column of $\bm{V}$ on the diagonal.
+Then, $\diag{\bm{V}\bm{u}_N} = \sum_{j=1}^N\bm{V}_j (\bm{u}_N)_j$, and
+\[
+\bm{V}^T f\LRp{\bm{V}\bm{u}_N} = \sum_{j=1}^N (\bm{u}_N)_j \bm{V}^T \bm{V}_j \bm{V}\bm{u}_N.
+\]
+The matrices $\bm{V}^T \bm{V}_j \bm{V}$ can be precomputed, such that the cost of evaluating the above expression is $N$ matrix-vector multiplications, or $O(N^3)$ operations.
+For $m = 3$, repeating the same steps yields
+\[
+\bm{V}^T f\LRp{\bm{V}\bm{u}_N} = \sum_{j=1}^N \sum_{k=1}^M (\bm{u}_N)_j  (\bm{u}_N)_k \bm{V}^T \bm{V}_j \bm{V}_k \bm{V}\bm{u}_N.
+\]
+which requires $N^2$ matrix-vector multiplications, or $O(N^4)$ operations.
+Generalizing this procedure to $m > 3$ yields that a degree $m$ polynomial nonlinearity can be evaluated in $O(N^{m+1})$ operations.
+
+Experimental results suggest that the number of hyper-reduced points $N_s$ scales as $O(N)$, though the constant increases with the spatial dimension (numerical experiments suggest the constant is approximately 2 in 1D and around 10 in 2D).
+Define constants $\alpha = N_s/ N$ and $\beta$, where $\beta$ denotes the number of operations required to evaluate the flux $\bm{f}_S$.
+We note that different choices of entropy stable flux functions (see, for example, \cite{ismail2009affordable, chandrashekar2013kinetic, hicken2018family}) produce significantly different values of $\beta$.
+Then, the hyper-reduced treatment of the convective nonlinearity requires $N_s^2$ flux evaluations.
+Assuming flux evaluations dominate costs, entropy stable schemes require $\alpha^2 \beta N^2$ operations.
+Thus, rather than increasing the asymptotic computational complexity, entropy stable hyper-reduction increases the associated constant.
+
+%\note{Compare to pre-computed matrices.
+For a nonlinearity $u^m$, the cost is $O(N^{m-1})$ $N\times N$ matrix multiplies, which scales as $O(N^{m+1})$.
+Assuming that the number of hyper-reduced points $N_s = O(N)$, then the cost scaling of $O(N_s^2) = O(N^2)$, where the constants depend on the ratio of $N_s/N$ and probably dimension.
+}
 
@@
 \label{sec:diss}
+
 Given an entropy conservative ROM with hyper-reduction, we can construct an entropy stable ROM by adding appropriate entropy-dissipative viscosity terms.
@@
 \end{equation}
-Since the entropy-projected conservative variables $\tilde{\bm{u}}$ are mappings of the projected entropy variables $\bm{V}\bm{v}_N$, repeating the steps used to derive (\ref{eq:entropydissfom}) yields that the entropy dissipation of (\ref{eq:visc1}) is
+Here, the entropy-projected conservative variables $\tilde{\bm{u}}$ are mappings of grid values of the projected entropy variables  $\bm{V}\bm{v}_N$.
+Repeating the steps used to derive (\ref{eq:entropydissfom}) yields that the entropy dissipation of (\ref{eq:visc1}) is
 \begin{align*}
@@
 \begin{remark}
-The hyper-reduction treatments of viscosity described in this section can also be applied to interface dissipation, such as local Lax-Friedrichs penalization.
-%However, these dissipation terms depend on the resolution of the full order model.
+The hyper-reduction treatments of viscosity described in this section can also be applied to nonlinear interface dissipation models, such as local Lax-Friedrichs penalization.
+We do not focus on such approximations for this paper, as the amount of dissipation applied is typically proportional to $\Delta x$ and thus depends on the dimension of the full order model.
 \end{remark}
@@
 is often effective.
-This approach is simpler to implement than (\ref{eq:visc1}),(\ref{eq:visc2}), and does not require a second set of hyper-reduced points.
+This approach is simpler to implement than (\ref{eq:visc1}), (\ref{eq:visc2}), and does not require a second set of hyper-reduced points.
 We are unable to show that this treatment is provably entropy dissipative.
@@
 \end{equation}
-The flux is referred to as entropy conservative if the inequality is an equality.
+The flux is referred to as entropy conservative if the inequality in (\ref{eq:esbflux}) is an equality.
 If an entropy stable or entropy conservative flux $\bm{f}_S(\bm{u}_L,\bm{u}_R)$ is used to evaluate both the volume and surface flux, then the resulting full order model given by (\ref{eq:esbc}) is also entropy stable or entropy conservative \cite{tadmor1987numerical, parsani2015entropy, chen2017entropy}.
@@
 Extending entropy conservative treatments of boundary conditions to the reduced order model is less straightforward due to fact that reduced matrices do not satisfy the same SBP property (\ref{eq:fomsbp}).
-If the full order matrix $\bm{Q}$ satisfies the SBP property, then the hyper-reduction of $\bm{Q}$ satisfies a \textit{generalized} SBP property \cite{fernandez2014generalized}
+If the full order matrix $\bm{Q}$ satisfies the SBP property, then the hyper-reduced matrix $\bm{Q}_t$ satisfies a \textit{generalized} SBP property \cite{fernandez2014generalized}
 \[
@@
 For linear problems, it is possible to impose energy stable boundary conditions by combining generalized SBP operators with appropriate simultaneous approximation terms (SATs) \cite{fernandez2018simultaneous}.
-However, due to the presence of $\bm{E}$ within the boundary term $\bm{E}^T\bm{B}\bm{E}$, the accurate and entropy stable imposition of nonlinear boundary conditions using generalized SBP SATs remains an open problem \cite{crean2018entropy, chan2018efficient, chenreview}.
+However, due to the presence of $\bm{E}$ within the boundary term $\bm{E}^T\bm{B}\bm{E}$, the accurate and entropy stable imposition of nonlinear boundary conditions using generalized SBP-SATs remains an open problem \cite{crean2018entropy, chan2018efficient, chenreview}.
 We address this issue by constructing a \textit{hybridized} SBP operator (also referred to as a decoupled SBP operator) \cite{chan2017discretely, chan2019skew}
@@
 \[
-\td{S\LRp{\bm{V}\bm{u}_N}}{t} \leq 0,
+\bm{1}^T\bm{W}\td{S\LRp{\bm{V}\bm{u}_N}}{t} \leq 0,
 \]
@@
 \begin{align*}
-\td{S\LRp{\bm{V}\bm{u}_N}}{t}  + \tilde{\bm{v}}^T\LRp{\LRp{\bm{Q}_h-\bm{Q}_h} \circ \bm{F}}\bm{1} + \tilde{\bm{v}}_b^T\bm{B}\bm{f}^*  = \bm{0}.
+\bm{1}^T\bm{W}\td{S\LRp{\bm{V}\bm{u}_N}}{t}  + \tilde{\bm{v}}^T\LRp{\LRp{\bm{Q}_h-\bm{Q}_h} \circ \bm{F}}\bm{1} + \tilde{\bm{v}}_b^T\bm{B}\bm{f}^*  = \bm{0}.
 \end{align*}
@@
 \begin{align*}
-\tilde{\bm{v}}^T\big(\underbrace{\LRp{\bm{Q}_h-\bm{Q}_h^T}}_{\bm{S}_h} \circ \bm{F}\big)\bm{1} &= 
+\tilde{\bm{v}}^T\big({\LRp{\bm{Q}_h-\bm{Q}_h^T}} \circ \bm{F}\big)\bm{1} &= 
 %\sum_{ij} \LRp{\bm{S}_h}_{ij} \tilde{\bm{v}}_i^T\bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j} \\
 %&= 
-\frac{1}{2}\sum_{ij} \LRp{\bm{S}_h}_{ij} \LRp{\tilde{\bm{v}}_i-\tilde{\bm{v}}_j}^T\bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}\\
+\frac{1}{2}\sum_{ij} \LRp{\bm{Q}_h-\bm{Q}_h^T}_{ij} \LRp{\tilde{\bm{v}}_i-\tilde{\bm{v}}_j}^T\bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}\\
 &= \frac{1}{2}\sum_{ij} \LRp{\bm{Q}_h-\bm{Q}_h^T}_{ij} \LRp{\psi(\tilde{\bm{u}}_i)-\psi(\tilde{\bm{u}}_j)}\\
-&= \sum_{ij} \LRp{\bm{Q}_h}_{ij} \LRp{\psi(\tilde{\bm{u}}_i)-\psi(\tilde{\bm{u}}_j)} = {\psi(\tilde{\bm{u}})^T\bm{Q}_h\bm{1} - \bm{1}^T\bm{Q}_h\psi(\tilde{\bm{u}})}\\
-& = -\bm{1}^T\bm{B}_h\psi(\tilde{\bm{u}}) = -\bm{1}^T\bm{B}\psi(\tilde{\bm{u}}_b),
+%&= \sum_{ij} \LRp{\bm{Q}_h}_{ij} \LRp{\psi(\tilde{\bm{u}}_i)-\psi(\tilde{\bm{u}}_j)} 
+&= {\psi(\tilde{\bm{u}})^T\bm{Q}_h\bm{1} - \bm{1}^T\bm{Q}_h\psi(\tilde{\bm{u}})} = -\bm{1}^T\bm{B}_h\psi(\tilde{\bm{u}}) = -\bm{1}^T\bm{B}\psi(\tilde{\bm{u}}_b),
 \end{align*}
@@
 \[
-\td{S\LRp{\bm{V}\bm{u}_N}}{t}  - \bm{1}^T\bm{B}\LRp{\psi(\tilde{\bm{u}}_b)-\tilde{\bm{v}}_b^T\bm{f}^*}  = \bm{0}.
-\]
-If the flux is entropy stable, then $\td{S\LRp{\bm{V}\bm{u}_N}}{t} \leq 0$.  Equality holds if the flux is entropy conservative such that $\psi(\tilde{\bm{u}}_b)=\tilde{\bm{v}}_b^T\bm{f}^*$.
+\bm{1}^T\bm{W}\td{S\LRp{\bm{V}\bm{u}_N}}{t}  - \bm{1}^T\bm{B}\LRp{\psi(\tilde{\bm{u}}_b)-\tilde{\bm{v}}_b^T\bm{f}^*}  = \bm{0}.
+\]
+If the flux is entropy stable, then $\bm{1}^T\bm{W}\td{S\LRp{\bm{V}\bm{u}_N}}{t} \leq 0$.
+Equality holds if the flux is entropy conservative such that $\psi(\tilde{\bm{u}}_b)=\tilde{\bm{v}}_b^T\bm{f}^*$.
 \end{proof}
@@
 \end{align*}
-These test bases can be used to construct reduced matrices $\bm{Q}_t^i$ %, \bm{Q}_t^y$ 
+These matrices can be used to construct reduced nodal differentiation matrices $\bm{Q}_t^i$ %, \bm{Q}_t^y$ 
 \[
@@
 \bm{M}_N\td{\bm{u}_N}{t} + \sum_{i=1}^d 2\bm{V}\LRp{\mathcal{I},:}^T\LRp{\bm{Q}^i_t \circ \bm{F}^i}\bm{1} = \bm{0} \label{eq:hrromd}\\
-\bm{M}_N = \Delta x^2 \bm{V}^T\bm{V}, \qquad \bm{F}^i_{jk} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}.\nonumber
+\bm{M}_N = \bm{V}\LRp{\mathcal{I},:}^T\bm{W}\bm{V}\LRp{\mathcal{I},:}, \qquad \bm{F}^i_{jk} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}.\nonumber
 %\bm{M}_N = \bm{V}\LRp{\mathcal{I},:}^T\bm{W}\bm{V}\LRp{\mathcal{I},:}, \qquad \bm{P} = \bm{M}_N^{-1}\bm{V}\LRp{\mathcal{I},:}^T\bm{W},\nonumber\\ 
@@
 
-\subsection{Non-periodic domains}
+\subsection{Non-periodic domains in higher dimensions}
 
@@
 However, the number of boundary points scales with $K_{\rm 1D}$.
-To ensure that the cost of the reduced order model scales with the number of modes $N$ rather than the dimension of the full order model, we must approximate boundary terms by a hyper-reduced weighted combination of sub-sampled boundary points.
+To ensure that the cost of the reduced order model scales with the number of modes $N$ rather than the dimension of the full order model, we approximate boundary terms by a hyper-reduced weighted combination of sampled boundary points.
 
@@
 \]
-
 These operators can be used to construct a hyper-reduced formulation in higher dimensions
@@
 These constraints on the boundary weights $\bm{w}_b$ can be enforced, for example, by augmenting the non-negative least squares solve in Algorithm~\ref{alg:hr} with equality constraints.
-We take an alternative approach in this paper, and incorporate the constraints (\ref{eq:sbpconstraints}) directly into hyper-reduction approaches based on linear programming.
+We take an alternative approach in this paper, and incorporate the constraints (\ref{eq:sbpconstraints}) directly into a hyper-reduction approach based on the solution of a linear programming problem using the dual simplex method.
 We refer the reader to \cite{yano2019lp} for details.
@@
 \[
-{p = (\gamma-1)\LRp{E - \frac{1}{2}\rho \LRp{\sum_{j=1}^d u_j^2}}}, \qquad {\rho e = E - \frac{1}{2}\rho \LRp{\sum_{j=1}^d u_j^2}}. 
+{p = (\gamma-1)\LRp{E - \frac{1}{2}\rho \LRb{\bm{u}}^2}}, \qquad {\rho e = E - \frac{1}{2}\rho\LRb{\bm{u}}^2}, \qquad \LRb{\bm{u}}^2 = {\sum_{j=1}^d u_j^2}.
 \]
 %There exists an infinite family of suitable convex entropies for the compressible Euler equations \cite{harten1983symmetric}.  However, 
-There is a unique entropy which symmetrrizes the viscous heat conduction term in the compressible Navier-Stokes equations \cite{hughes1986new}.  This entropy $S(\bm{u})$ is given by
+There is a unique entropy which symmetrizes the viscous heat conduction term in the compressible Navier-Stokes equations \cite{hughes1986new}.  This entropy $S(\bm{u})$ is given by
 \begin{equation*}
@@
 \end{align*}
-where we have redefined ${u^2_{\rm avg}} = 2(\avg{u_1}^2 + \avg{u_2}^2) - \LRp{\avg{u_1^2} +\avg{u_2^2}}.$
+where in 2D, we redefine ${u^2_{\rm avg}} = 2(\avg{u_1}^2 + \avg{u_2}^2) - \LRp{\avg{u_1^2} +\avg{u_2^2}}.$
 
@@
 Explicit expressions for Jacobians in the compressible Euler and Navier-Stokes equations are given in \cite{hughes1986new}.
+The solution is evolved using an explicit 5-stage 4th order Runge-Kutta scheme.
 %\note{Add equations for $\pd{\bm{u}}{\bm{v}}$. }
@@
 \]
-The solution produces a shock some time after $T=.25$.
+The solution exhibits a viscous shock some time after $T=.25$.
 
@@
 The difference becomes more pronounced for a coarser subsampling.
-We note that the enrichment does affect the resulting singular vectors, however.
-For example, enrichment results in differences in several computed singular vector (Figure~\ref{subfig:c}).
+We note that the enrichment does affect the form of the resulting singular vectors (Figure~\ref{subfig:c}).
 We also compute projection errors for each solution snapshot using $25$ modes in Figure~\ref{subfig:b}, along with differences between the projected solutions with and without entropy variable enrichment.
@@
 %For this problem setup, the remaining singular vectors show little difference; however, for other settings, we have observed larger differences with and without enrichment for higher modes.
+
 \begin{figure}
@@
 \subfloat[Projection errors, 25 modes]{\raisebox{.05em}{\includegraphics[width=.33\textwidth]{UVprojectionerrors.png}\label{subfig:b}}}
-\caption{Behavior of the SVD and reduced basis functions with and without entropy variable enrichment of the solution snapshots. }
+\caption{Snapshot singular values and reduced basis functions with and without entropy variable enrichment. }
 \label{fig:svd}
@@
 We next examine solutions produced by the reduced order models.
-We observe that the reduced models can utilize a larger CFL; however, to ensure that temporal errors are small, all solutions are computed using the same CFL as the full order model.
+It is known that reduced models can typically utilize a larger CFL \cite{lucia2004reduced, knezevic2011reduced}; however, to ensure that temporal errors are small, all solutions are computed using the same CFL as the full order model.
 Figure~\ref{fig:romsols} shows density computed using 25, 75, and 125 modes and the hyper-reduced treatment of viscosity in (\ref{eq:visc2}).
@@
 However, despite this under-resolution, the ROM remains stable and does not blow up.
-In all cases, we compute hyper-reductions of both convective and viscous terms.
-For 25 modes, 92 convective hyper-reduced points and 133 viscous hyper-reduced points are used.
-For 75 modes, 245 convective hyper-reduced points and 242 viscous hyper-reduced points are used.
-For 125 modes, 410 convective hyper-reduced points and 292 viscous hyper-reduced points are used.
 
@@
 \end{figure}
+%For 25 modes, 92 convective hyper-reduced points and 133 viscous hyper-reduced points are used.
+For 75 modes, 245 convective hyper-reduced points and 242 viscous hyper-reduced points are used.
+For 125 modes, 410 convective hyper-reduced points and 292 viscous hyper-reduced points are used.
+
+Table~\ref{tab:hrpts} shows the number of hyper-reduced points computed for different numbers of modes.
+We report on three different sets of points: empirical cubature points (used to approximate the mass matrix $\Delta x \bm{V}^T\bm{V}$), stabilizing points (added to control the condition number of the test mass matrix; see Section~\ref{sec:condtest}), and viscous points (used to approximate viscous terms; see Section~\ref{sec:diss}).
+All point sets are computed using Algorithm~\ref{alg:hr} with dimensionality reduction as described in Section~\ref{sec:hyperreducalgo}.
+In all cases, we observe that the number of empirical cubature and viscous points grows roughly as $2N$ (similar observations were made in \cite{hernandez2017dimensional}), while the number of stabilizing points hovers around $30$ for $N > 25$.
+
+\begin{table}[!h]
+\centering
+\begin{tabular}{|c || c | c | c |c |}
+\hline
+Number of modes $N$ & 25 & 75 & 125 & 175\\
+\hhline{|=|=|=|=|=|}
+Number of empirical cubature points &54 & 158 & 259 &355 \\
+\hline
+Number of stabilizing points & 3& 21 & 36 & 28 \\
+\hline
+Number of viscous points & 54 &159 & 259 & 366  \\
+\hline
+\end{tabular}
+\caption{Number of computed hyper-reduced points for the 1D Euler equations.}
+\label{tab:hrpts}
+\end{table}
 
@@
 We also run the ROM with the viscosity parameter $\epsilon = 0$.
-For 75 modes, the solution is more oscillatory than
+For 125 modes, the solution is significantly more oscillatory than the case of $\epsilon = 2\times 10^{-4}$; however, the reduced solution does not blow up.
 
@@
 \centering
-\subfloat[Entropy contribution $\LRb{\bm{v}_N^T \bm{V}_h^T\LRp{\bm{Q}_h \circ \bm{F}}\bm{1}}$]{\includegraphics[width=.4525\textwidth]{entropyrhs.png}}
+\subfloat[Convective entropy contribution $\LRb{\bm{v}_N^T \bm{V}_h^T\LRp{\bm{Q}_h \circ \bm{F}}\bm{1}}$]{\includegraphics[width=.4525\textwidth]{entropyrhs.png}}
 \hspace{2em}
 %\subfloat[Difference, 25 modes]{\includegraphics[width=.32\textwidth]{viscdiff25modes.png}}
-\subfloat[Density $\rho$ with 75 modes and no viscosity]{\includegraphics[width=.44\textwidth]{euler1Dwall_t75_75modes_novisc.png}}
-\caption{Entropy contribution of the convective term $\LRb{\bm{v}_N^T \bm{V}_h^T\LRp{\bm{Q}_h \circ \bm{F}}\bm{1}}$ over time and reduced order solution with viscosity parameter $\epsilon = 0$ at $T=.75$.
+\subfloat[Density $\rho$ with 125 modes and no viscosity]{\includegraphics[width=.432\textwidth]{euler1Dwall_t75_125modes_novisc.png}}
+\caption{Convective entropy contribution $\LRb{\bm{v}_N^T \bm{V}_h^T\LRp{\bm{Q}_h \circ \bm{F}}\bm{1}}$ over time and reduced order solution with viscosity parameter $\epsilon = 0$ at $T=.75$.
 The solution does not blow up despite the presence of large oscillations resulting from the shock.}
@@
 \centering
-\subfloat[25 modes]{\includegraphics[width=.45\textwidth]{visc25modes.png}}
-\hspace{.5em}
+\subfloat[Entropy dissipation for 25 modes]{\includegraphics[width=.45\textwidth]{visc25modes.png}}
+\hspace{2em}
 %\subfloat[Difference, 25 modes]{\includegraphics[width=.32\textwidth]{viscdiff25modes.png}}
-\subfloat[75 modes]{\includegraphics[width=.45\textwidth]{visc75modes.png}}
+\subfloat[Entropy dissipation for 75 modes]{\includegraphics[width=.45\textwidth]{visc75modes.png}}
 \caption{Discrete entropy dissipation $\bm{v}_N^T\bm{d}(\bm{u}_N)$ computed using hyper-reduced treatments of viscous terms (\ref{eq:visc1}), (\ref{eq:visc2}), and (\ref{eq:visc3}) for 25 and 75 modes.}
@@
 Both entropy and error are computed on the original grid of the full order model (e.g., without hyper-reduction).
+The error is computed as the discrete relative $L^2$ error over all the conservative variables.
 We observe that the evolution of the ROM average discrete entropy converges to the average discrete entropy for the full order model as the number of modes increases from 25 to 75.
@@
 \centering
-\subfloat[Entropy over time]{\includegraphics[height=.265\textheight]{entropyevolution.png}}
-\hspace{.5em}
-\subfloat[Error between ROM and FOM]{\includegraphics[height=.26\textheight]{euler_error_hypred_comparison_K2500}}
+\subfloat[Error between ROM and FOM]{\includegraphics[height=.282\textheight]{euler_error_hypred_comparison_K2500}}
+\hspace{2em}
+\subfloat[Entropy over time]{\includegraphics[height=.285\textheight]{entropyevolution.png}}
 %\hspace{.1em}
@@
 
+%\note{Finally, we investigate the role of viscosity.
+While the reduced order models constructed in this paper are entropy stable (or conservative) even for the inviscid regime, accuracy suffers as the viscosity coefficient decreases.
+We compute solutions using the same Gaussian pulse initial condition for $\epsilon = 2e-4$, $\epsilon = 1e-3$, and $\epsilon = 0$.
+For $\epsilon = 0$, we observe that the error does not converge as the number of modes increases.
+%We note that this degradation in accuracy appears to be related to the ratio of viscosity and grid spacing $\Delta x$, rather than simply the absolute magnitude of the viscosity coefficient $\epsilon$.
+}
+
 %Next, we examine
@@
 \]
-The full order model is computed on a $100\times 100$ grid with viscosity coefficient $\epsilon = 1e-3$, and the reduced order solution is shown in Figure~\ref{fig:pulse2d}.
+The full order model is computed on a $150\times 150$ grid with viscosity coefficient $\epsilon = 1e-3$.
+The POD basis is computed from 75 solution snapshots with a CFL of $.5$.
+The reduced order solutions are shown in Figures~\ref{fig:pulse2d}.
 Since little difference is observed between the hyper-reducted treatments of viscosity, the naive treatment (\ref{eq:visc3}) is used.
+The hyper-reduction algorithm adds one single stabilizing point to reduce the condition number of the $x$-coordinate test mass matrix from $O(10^{8})$ to approximately $5.71$.
+
 The solution does not form shock discontinuities, and is chosen instead to test the imposition of boundary conditions.
@@
 \centering
+%\subfloat[Full order model]{\includegraphics[width=.32\textwidth]{pulse2d.png}}
+%\hspace{.1em}
+%\subfloat[Reduced order model]{\includegraphics[width=.32\textwidth]{pulse2d_ROM.png}}
+%\hspace{.1em}
+%\subfloat[Hyper-reduction points]{\includegraphics[width=.32\textwidth]{pulse2d_ROM_pts.png}}
+%\caption{Density computed for $\epsilon = 1e-3$ using 20 modes, 194 hyper-reduced points (in \textcolor{red}{red}), and 68 hyper-reduced boundary points (in \textcolor{blue}{blue}).
+The relative $L^2$ error is $.0088005$.}
 \subfloat[Full order model]{\includegraphics[width=.32\textwidth]{pulse2d.png}}
@@
 \subfloat[Hyper-reduction points]{\includegraphics[width=.32\textwidth]{pulse2d_ROM_pts.png}}
-\caption{Density computed using 25 modes, 303 hyper-reduced volume points (in \textcolor{red}{red}), and 75 hyper-reduced boundary points (in \textcolor{blue}{blue}).
-The relative $L^2$ error is $.00934$.}
+\caption{Density computed for $\epsilon = 1e-3$ using 25 modes, 300 hyper-reduced points (in \textcolor{red}{red}), and 66 hyper-reduced boundary points (in \textcolor{blue}{blue}).
+The relative $L^2$ error is $.0071$.}
 \label{fig:pulse2d}
@@
 
+\note{We conclude by examining a 2D problem with a non-trivial shock structure.  }
+
+%Figures~\ref{fig:pulse2d_novisc} also shows reduced order solutions for 25 modes and $\epsilon = 0$.
+The number of hyper-reduced points does not change significantly; however, the results are noticably worse.
+%\begin{figure}
+%\centering
+%\subfloat[Full order model (inviscid)]{\includegraphics[width=.3\textwidth]{pulse2d_novisc.png}}
+%\hspace{.1em}
+%\subfloat[Reduced order model (inviscid)]{\includegraphics[width=.3\textwidth]{pulse2d_ROM_novisc.png}}
+%\hspace{.1em}
+%\subfloat[Singular values for $\epsilon = 1e-3$ and $\epsilon = 0$]{\includegraphics[width=.355\textwidth]{pulse2d_singvals.png}}
+%\caption{Density computed  for $\epsilon = 0$ using 25 modes, 302 hyper-reduced volume points (in \textcolor{red}{red}), and 79 hyper-reduced boundary points (in \textcolor{blue}{blue}).
+The relative $L^2$ error is $.01687$.}
+%\label{fig:pulse2d_novisc}
+%\end{figure}
+
 \section{Conclusion}  We have presented a methodology for constructing projection-based reduced order models for nonlinear conservation laws from a reduced basis, modified Galerkin projection, and tailored hyper-reduction techniques.
@@
 Future work will aim to address computational costs associated with entropy stable reduced models.
-The main challenge is the fact that, unlike standard hyper-reduced models, the number of nonlinear evaluations necessary scales with $O(N_s^2)$ rather than $O(N_s)$, where $N_s$ is the number of hyper-reduced sampling points.
-These costs can be reduced using, for example, domain decomposition \cite{lucia2003reduced} or local $h$-refinement \cite{carlberg2015adaptive}, which will be explored in future work.
+Unlike standard hyper-reduction techniques, the number of nonlinear evaluations necessary scales with $O(N_s^2)$ rather than $O(N_s)$, where $N_s$ is the number of hyper-reduced sampling points.
+This discrepancy is due to the fact that the hyper-reduction presented here approximates a nonlinear matrix, rather than a nonlinear vector.
+These costs can be reduced by combining domain decomposition \cite{lucia2003reduced} and a discontinuous Galerkin-type discretization \cite{chan2017discretely} to increase the sparsity of the resulting hyper-reduced matrices.
+We will also investigate implicit time-stepping for entropy stable reduced models, as the additional costs associated with the spatial formulation may be offset by the smaller matrix systems which must be solved in the linearization process.
 
@@
 
-The author thanks Matthias Heinkenschloss, Matthew Zahr, and Masayuki Yano for informative discussions.
+Jesse Chan gratefully acknowledges support from the National Science Foundation under awards DMS-1719818 and DMS-1712639, and thanks Matthias Heinkenschloss, Masayuki Yano, Matthew Zahr, and Irina Tezuar for informative discussions.
 

### 2019-09-06 docs/ESDG_mortar/main.tex
updating docs

@@
 %\date{}
-\author{Jesse Chan, David C.\ Del Rey Fernandez}
-\title{Mortar-based entropy stable modal discontinuous Galerkin methods}
+\author{Jesse Chan, Mario Bencomo, David C.\ Del Rey Fernandez}
+\title{On mortar-based entropy stable modal discontinuous Galerkin methods}
 \graphicspath{{./figs/}}
@@
 \begin{abstract}
-Entropy stable schemes provide a semi-discrete statement of stability for nonlinear systems of conservation laws.
+Entropy stable schemes provide a semi-discrete statement of stability for nonlinear conservation laws.
 Entropy stable nodal discontinuous Galerkin (DG) methods have been constructed using entropy conservative finite volume fluxes \cite{tadmor1987numerical} and a summation-by-parts (SBP) finite difference framework \cite{gassner2013skew, carpenter2014entropy, gassner2016split, chen2017entropy, crean2018entropy, chan2017discretely, chan2018efficient}.
-Such DG schemes require evaluating entropy conservative numerical fluxes between combinations of volume and surface quadrature nodes.
+Entropy stable DG schemes require evaluating entropy conservative numerical fluxes between combinations of volume and surface quadrature nodes.
+On tensor product elements with surface nodes aligned with volume nodes, flux evaluations are required only between ``lines'' of nodes, reducing the number of flux evaluations to $O(N)$ per node.
+However, because surface nodes are no longer aligned on non-conforming meshes, each volume node interacts each non-aligned surface node, increasing the number of flux evaluations to $O(N^{d-1})$ per node in $d$ dimensions.
 \end{abstract}
@@
 Note that we have used the mortar mass matrix $\bm{M}_m$ in the above definition of the face projection.
-This is only necessary if the surface and mortar quadratures do not exactly integrate degree $2N$ polynomials.
+%This is only necessary if the surface and mortar quadratures do not exactly integrate degree $2N$ polynomials.
 If both the surface and mortar quadratures are exact for polynomials of degree $2N$, then $\bm{M}_m = \bm{M}_f$, and no distinction is necessary between the two mass matrices.

### 2019-09-06 docs/GD lumping/main.tex
updating docs

\pdfoutput=1

%\documentclass[preprint,10pt]{elsarticle}
\documentclass[preprint,10pt]{amsart}
%\documentclass[review]{siamart0216}
%\documentclass{siamart0216}

%\usepackage{fullpage}
%\usepackage[colorlinks=true]{hyperref}

\usepackage{amsmath,amssymb,amsfonts}
%\usepackage{amsthm}
%\theoremstyle{definition}
%\newtheorem{definition}{Definition}
\theoremstyle{lemma}
\newtheorem{lemma}{Lemma}
\theoremstyle{corollary}
\newtheorem{corollary}{Corollary}
\newtheorem*{remark}{Remark}
\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}
%\theoremstyle{assumption}
%\newtheorem{assumption}{Assumption}

\usepackage[titletoc,toc,title]{appendix}

\usepackage{array} 
\usepackage{listings}
\usepackage{mathtools}
\usepackage{pdfpages}
\usepackage[textsize=footnotesize,color=green]{todonotes}
\usepackage{bm}
\usepackage{bbm}

\usepackage{tikz}
\usepackage[normalem]{ulem}
\usepackage{hhline}

\usepackage{graphicx}
\usepackage{subfig}
\usepackage{color}

%% ====================================== graphics

\usepackage{pgfplots}
\usepackage{pgfplotstable}
\definecolor{markercolor}{RGB}{124.9, 255, 160.65}
\pgfplotsset{
compat=1.3,
width=10cm,
tick label style={font=\small},
label style={font=\small},
legend style={font=\small}
}

\usetikzlibrary{calc}
\usetikzlibrary{intersections}

%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
\newcommand{\logLogSlopeTriangle}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        \pgfmathsetmacro{\xArel}{#1}
        \pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{\yArel}
        \pgfmathsetmacro{\xCrel}{\xArel}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=north] {}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=west] {\textcolor{black}{#4}}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

\newcommand{\logLogSlopeTriangleNeg}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        \pgfmathsetmacro{\xArel}{#1}
        \pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{\yArel}
        \pgfmathsetmacro{\xCrel}{\xArel}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=.5,anchor=south] {}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=west] {\textcolor{black}{#4}}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
\newcommand{\logLogSlopeTriangleFlipNeg}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        %\pgfmathsetmacro{\xArel}{#1}
        %\pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{#3}
        \pgfmathsetmacro{\xCrel}{#1}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

	\pgfmathsetmacro{\xArel}{\xBrel}
        \pgfmathsetmacro{\yArel}{\yCrel}

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=east] {\textcolor{black}{#4}}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=north] {1}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
\newcommand{\logLogSlopeTriangleFlip}[5]
{
    % #1. Relative offset in x direction.
    % #2. Width in x direction, so xA-xB.
    % #3. Relative offset in y direction.
    % #4. Slope d(y)/d(log10(x)).
    % #5. Plot options.

    \pgfplotsextra
    {
        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}

        % Calculate auxilliary quantities, in relative sense.
        %\pgfmathsetmacro{\xArel}{#1}
        %\pgfmathsetmacro{\yArel}{#3}
        \pgfmathsetmacro{\xBrel}{#1-#2}
        \pgfmathsetmacro{\yBrel}{#3}
        \pgfmathsetmacro{\xCrel}{#1}

        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.

	\pgfmathsetmacro{\xArel}{\xBrel}
        \pgfmathsetmacro{\yArel}{\yCrel}

        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
        \coordinate (A) at (rel axis cs:\xArel,\yArel);
        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=east] {\textcolor{black}{#4}}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=south] {}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

\usepackage{stmaryrd}

\renewcommand{\tilde}{\widetilde}
\renewcommand{\hat}{\widehat}
\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

\newcommand{\vect}[1]{\ensuremath\boldsymbol{#1}}
\newcommand{\tensor}[1]{\underline{\bm{#1}}}
\newcommand{\del}{\triangle}
\newcommand{\curl}{\grad \times}
\renewcommand{\div}{\grad \cdot}

\newcommand{\bbm}[1]{\mathbbm{#1}}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\equaldef}{\stackrel{\mathrm{def}}{=}}

\newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}{\rm #2}}}
\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRb}[1]{\left| #1 \right|}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}
\newcommand{\LRceil}[1]{\left\lceil #1 \right\rceil}
\newcommand{\LRl}[1]{\left. #1 \right|}
\newcommand{\pdd}[2]{\frac{\partial^2#1}{\partial#2^2}}
\newcommand{\pdn}[3]{\frac{\partial^{#3}#1}{\partial#2^{#3}}}
\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\snor}[1]{\left| #1 \right|}

%\newcommand{\cond}[1]{\kappa\LRp{#1}}
\newcommand{\cond}[2]{\nor{#1}_{#2}\nor{{#1}^{-1}}_{#2}}

\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\Div} {\ensuremath{\nabla\cdot}}
\newcommand{\jump}[1] {\ensuremath{\llbracket#1\rrbracket}}
\newcommand{\avg}[1] {\ensuremath{\LRc{\!\{#1\}\!}}}

\newcommand{\Oh}{{\Omega_h}}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\LK}{L^2\LRp{D^k}}
\newcommand{\LdK}{L^2\LRp{\partial D^k}}
\newcommand{\Dhat}{\widehat{D}}
\newcommand{\Lhat}{L^2\LRp{\Dhat}}

\newcommand{\eval}[2][\right]{\relax
  \ifx#1\right\relax \left.\fi#2#1\rvert}

\def\etal{{\it et al.~}}

\newcommand{\note}[1]{{\color{blue}{#1}}}
%\newcommand{\noteOne}[1]{{\color{blue}{#1}}}
%\newcommand{\noteTwo}[1]{{\color{red}{#1}}}
%\newcommand{\note}[1]{#1}
%\newcommand{\noteOne}[1]{#1}
%\newcommand{\noteTwo}[1]{#1}

\newcommand{\LinfDk}{L^{\infty}\LRp{D^k}}

\newcommand{\diag}[1]{{\rm diag}\LRp{#1}}

\newcommand{\Ksub}{K_{\rm sub}}

\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

%% d in integrand
\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}}

\makeatletter
\renewcommand\d[1]{\mspace{6mu}\mathrm{d}#1\@ifnextchar\d{\mspace{-3mu}}{}}
\makeatother

%\date{}
\author{}
\title{}
\graphicspath{{./figs/}}

\begin{document}

\section{Properties of GD functions}

Interior GD functions are defined as scalings and translations of a ``canonical'' GD function $\phi(x)$.  Assuming that the polynomial degree $p$ is odd, let $\ell_j(x)$ be the $(p+1)$ Lagrange functions defined using equispaced nodal points $x_j$ such that
\[
\LRc{x_0,\ldots, x_p} = \LRc{-(a-1), \ldots, a}, \qquad a = \frac{p+1}{2}.%\LRceil{\frac{p}{2}}.
\]
This choice of nodal points ensures that the interval $[0,1]$ lies is the middle of $[x_0, x_p]$.  We define the canonical GD function $\phi(x)$ piecewise on the intervals $I_0,\ldots, I_{2p-2}$, where $I_j = [-(p-1)+j,-p+j]$
\[
\phi(x) = \begin{cases}
\ell_j(x-j), & x \in I_j\\
0 & {\rm otherwise}
\end{cases}.
\]

Symbolic computations then yield the following lemma:
\begin{lemma}
Let $\phi$ be the canonical GD function.  For $p = 1,3,\ldots, 75$, 
\begin{enumerate}
\item $\int_{-\infty}^{\infty} \phi^2 \leq \int_{-\infty}^{\infty} \phi = 1$  \label{eq:lem1}
\item The GD basis functions satisfy \label{eq:lem2}
\[
\sum_{j=-p}^{p} \LRb{ \int \phi_0(x) \phi_0(x-j)} \leq 2.
\]
\item For any degree $p$ polynomial $u(x)$, \label{eq:lem3}
\[
\int_{-\infty}^{\infty} u\phi_i = u(0).
\]
\end{enumerate}
\label{lemma:gdprops}
\end{lemma}
\begin{proof}
We use symbolic software to verify the first part (\ref{eq:lem1}) and (\ref{eq:lem2}).  
The equality in (\ref{eq:lem1}) can be shown directly.  Because ${\rm supp}\LRp{\phi} = [-(p-1),p-1]$, $\int_{-\infty}^\infty \phi(x) = \int_{-(p-1)}^{p-1} \phi(x)$.  By the definition of $\phi(x)$, 
\[
\int_{-(p-1)}^{(p-1)} \phi(x) = \sum_{j=0}^p \int_{I_j} \ell_j(x-x_j) = \int_0^1 \sum_{j=0}^p \ell_j(x) = \int_0^1 1 = 1.
\]
To prove (\ref{eq:lem3}), we first show that $\int_{-(p-1)}^{p-1} x^k\phi = 0$ for $0 < k \leq p$.  For $k > 0$ odd, this holds by the symmetry of $\phi(x)$ across $0$.  For $0 < k \leq p$ even, this condition is equivalent to
\[
\int_{-(p-1)}^{p-1} x^k \phi_i(x) = 2 \int_{0}^{p-1} x^k \phi_i(x)= \sum_{j=0}^a \int_0^1 \ell_j(x) (x-j)^k = 0.
\]
which we verify for $p = 1,\ldots, 75$ using symbolic software.  
Then, (\ref{eq:lem3}) follows from (\ref{eq:lem1}) and a Taylor representation of $u(x)$ around $x = 0$.
\[
u(x) = u(0) + u'(0) x + u''(0) x^2 + \ldots + u^{(p)}(0) x^p.
\]
\end{proof}
We conjecture that Lemma~\ref{lemma:gdprops} holds for all $p > 0$ odd.  A translation and scaling of $\phi$ then provides the following corollary:
\begin{corollary}
Let $x_i$ be equispaced points with spacing $h$, and let $\phi_i(x) = \phi\LRp{(x - x_j)/h}$ be the GD function at $x_i$.  For any degree $p$ polynomial $u(x)$, 
\[
\frac{1}{h}\int_{-\infty}^{\infty} u\phi_i = u(x_i).
\]
\label{cor:acc}
\end{corollary}

\section{Accuracy and energy stability of the lumped GD mass matrix}

Using Lemma~\ref{lemma:gdprops}, we can show that the lumped GD mass matrix is high order accurate in the following sense:
\begin{lemma}
Let $\tilde{\bm{M}}$ denote the non-symmetric lumped GD mass matrix.  Then, $\tilde{\bm{M}}^{-1}\bm{Q}$ is a $(p+1)$ order accurate approximation to the first derivative.  
\end{lemma}
\begin{proof}
Let $u(x)$ be a degree $p$ polynomial with GD coefficients $\bm{u}_i = u(x_i)$.  Since the GD basis can reproduce polynomials of degree $p$, $\bm{\delta u} = \bm{M}^{-1}\bm{Q}\bm{u}$ are the GD coefficients of the exact derivative $\LRl{\pd{u}{x}}_{x_i}$.

The lumped GD mass matrix $\tilde{\bm{M}}$ is $(p+1)$ order accurate if $\tilde{\bm{M}} \bm{\delta u} = \bm{Q}\bm{u}$ as well.  Since $\bm{\delta u}$ is again polynomial, high order accuracy is ensured if $\bm{M}\bm{u} = \tilde{\bm{M}}\bm{u}$ for all coefficients $\bm{u}$ which correspond to polynomials of degree $p$.  Since the boundary rows of $\bm{M}$ and $\tilde{\bm{M}}$ are identical, $(\bm{M}\bm{u})_i = (\tilde{\bm{M}}\bm{u})_i$ for all indices $i$ corresponding to boundary GD functions.  For $\bm{u}$ polynomial and $i$ corresponding to interior GD functions, Corollary~\ref{cor:acc} guarantees $(\bm{M}\bm{u})_i = (\tilde{\bm{M}}\bm{u})_i$.  
\end{proof}

We now show that, for linear symmetric hyperbolic PDEs, the lumped GD mass matrix preserves semi-discrete energy stability.  We assume that there are sufficiently many elements relative to the order $p$.  Then, under a reordering of degrees of freedom, the GD mass matrix $\bm{M}$ is
\[
\bm{M} = \begin{bmatrix}
\bm{A} & \bm{B}\\
\bm{B}^T & \bm{C}
\end{bmatrix},
\]
where $\bm{A}$ is the sub-matrix consisting of rows and columns of $\bm{M}$ corresponding to boundary GD functions, and $\bm{C}$ is the sub-matrix corresponding to interior GD functions.  Since $\bm{M}$ is SPD, the matrices $\bm{A}, \bm{C}$ are also SPD.  We assume for simplicity that $h = 1$, such that the lumped mass matrix $\tilde{\bm{M}}$ is 
\[
{\tilde{\bm{M}}} = \begin{bmatrix}
\bm{A} & \bm{B}\\
\bm{0} & \bm{I}
\end{bmatrix}.
\]
We require that $\bm{x}^T\tilde{\bm{M}}\bm{x}$ induces a norm on $\bm{x}$, which holds if $\tilde{\bm{M}}$ is positive definite.

%We can prove positive-definiteness using 
%\begin{lemma}
%Let $\bm{A}$ be a symmetric banded Toeplitz matrix
%\[
%\begin{bmatrix}
%a_0 & a_1 & \ldots & a_p &&&\\
%a_1 & a_0 & a_1 & \ldots &a_p &&\\
%\vdots & \ddots & \ddots & \ddots& &&\\
%\end{bmatrix}
%\]
%The spectral radius of $\bm{A}$ is bounded by $\sum_{j=0}^p \LRb{a_j}$.
%\label{lemma:banded}
%\end{lemma}

For $h = 1$, we have the following property of $\bm{C}$:
\begin{lemma}
The maximum eigenvalue of $\LRb{\bm{C}}$ is bounded by $2$.
\label{lemma:Ceig}
\end{lemma}
\begin{proof}
The proof follows directly from bounding the Rayleigh quotient of $\LRb{\bm{C}}$ and using the banded Toeplitz nature of the matrix.  Expanding $\bm{u}^T\LRb{\bm{C}}\bm{u}$ out and using symmetry gives
\begin{align*}
\bm{u}^T\LRb{\bm{C}}\bm{u} &\leq \sum_{i=0}^K\sum_{j = \max(0,i-p)}^{\min(K,i+p)} \LRb{\bm{u}_i\bm{u}_j}\LRb{\int \phi_i\phi_j}\\
&\leq \sum_{i=0}^K \LRp{\bm{u}_i^2 \int\phi_i^2 + 2\sum_{j=i+1}^{\min(K,i+p)} \LRb{\bm{u}_i\bm{u}_j} \LRb{\int\phi_i\phi_j}}.
\end{align*}
Applying Young's inequality bounds this sum from above by
\begin{align}
&\leq\sum_{i=0}^K \LRp{\bm{u}_i^2 \int\phi_i^2 + \sum_{j=i+1}^{\min(K,i+p)} \LRp{\bm{u}_i^2 + \bm{u}_j^2} \LRb{\int\phi_i\phi_j}} \nonumber\\
&= \sum_{i=0}^K \bm{u}_i^2 \LRp{\sum_{j=i}^{\min(K,i+p)} \LRb{\int\phi_i\phi_{j}}} + \sum_{j=i+1}^{i+p} \bm{u}_j^2\LRb{\int \phi_i\phi_j}.\label{eq:tmp1}
\end{align}
Distributing the terms of the latter sum in (\ref{eq:tmp1}) yields that
\begin{align*}
\bm{u}^T\LRb{\bm{C}}\bm{u} &\leq \sum_{i=0}^K \bm{u}_i^2 \LRp{\sum_{j=\max(i-p,0)}^{\min(K,i+p)} \LRb{\int\phi_i\phi_{j}}}\\
&\leq  \sum_{i=0}^K \bm{u}_i^2\LRp{\sum_{j=-p}^{p} \LRb{\int\phi_0\phi_j}} \leq 2\sum_{i=0}^K \bm{u}_i^2  = 2\bm{u}^T\bm{u}
\end{align*}
where we have used translation invariance of the interior GD basis functions and property (\ref{eq:lem2}) of Lemma~\ref{lemma:gdprops}.
%\[
%\LRb{\int \phi_i\phi_{i+j}} = \LRb{\int \phi_0\phi_{j}}, \qquad -p \leq j \leq p.
%\]
%, using yields
%\[
%\bm{u}^T\LRb{\bm{C}}\bm{u} \leq \sum_{i=0}^K \bm{u}_i^2 \LRp{\sum_{j=-p}^{p} \LRb{\int\phi_0\phi_j}} \leq 2\sum_{i=0}^K \bm{u}_i^2  = 2\bm{u}^T\bm{u}.
%\]
\end{proof}

We can then show the following:
\begin{lemma}
The lumped GD mass matrix $\tilde{\bm{M}}$ is positive definite in the sense that
\[
0 < \bm{x}^T\tilde{\bm{M}}\bm{x}, \qquad \forall \bm{x}\in \mathbb{R}^n, \quad \bm{x}\neq 0.
\]
\end{lemma}
\begin{proof}
Let $\bm{x} = \LRs{\bm{u},\bm{v}}^T$.  Using that $\bm{u}^T\bm{C}\bm{u} \leq \bm{u}^T\LRb{\bm{C}}\bm{u}$ and Lemma~\ref{lemma:Ceig}
\begin{align*}
0 < \bm{x}^T{\bm{M}}\bm{x}&\leq \bm{u}^T\bm{A}\bm{u} + 2\bm{u}^T\bm{B}\bm{v} + \bm{v}^T\LRb{\bm{C}}\bm{v}\\
&\leq \bm{u}^T\bm{A}\bm{u} + 2\bm{u}^T\bm{B}\bm{v} + 2\bm{v}^T\bm{v}\\
&\leq 2\LRp{\bm{u}^T\bm{A}\bm{u} + \bm{u}^T\bm{B}\bm{v}+ \bm{v}^T\bm{v}} = 2 \bm{x}^T\tilde{\bm{M}}\bm{x}
\end{align*}
\end{proof}

\begin{remark}
The bound in Lemma~\ref{lemma:Ceig} is sufficient to show positive definiteness of the non-symmetric lumped GD mass matrix.  However, numerical experiments indicate that the maximum eigenvalue $\lambda_{\max}$ of $\bm{C}$ achieves a tighter bound $\lambda_{\max} \leq 1$ for all $p$ and $K$ tested.  
\end{remark}

Ideally, the norms induced by $\bm{M}$ and $\tilde{\bm{M}}$ should also be equivalent and induce equivalent measures of energy.  Numerical experiments that this is indeed the case.

%\section{Matlab code for symbolic computations}
%
%\begin{verbatim}
%clear
%syms x
%for p = 1:2:25
%    a = (p+1)/2;   
%    for k = 2:2:p-1
%        moment = sym(0);
%        phi2 = sym(0);
%        for j = -(a-1):0 % half of the points
%            ellj = sym(1);
%            denom = sym(1);
%            for i = -(a-1):a
%                if i~=j
%                    ellj = ellj * (x-i);
%                    denom = denom * (j-i);
%                end
%            end
%            ellj = ellj/denom;
%            moment = moment + int(ellj*(x-j)^k,0,1);
%            phi2 = phi2 + int(ellj*ellj,0,1);
%        end
%        if ~isequaln(moment,sym(0)) || (phi2 > 1.0)
%            fprintf('Condition violated for p = %d\n',p)
%            return
%        end
%    end
%    fprintf('Done with p = %d\n',p)
%end
%\end{verbatim}

\end{document}

### 2019-09-06 docs/elasticity_notes/arxiv/main.tex
updating docs for penalty note + elasticity
updating docs

--- 2017-01-17 docs/elasticity_notes/main.tex
@@
 }
+
+\usetikzlibrary{calc}
 
@@
 \LRa{\avg{\bm{\epsilon}_\sigma} - \frac{\tau_v}{2}\bm{A}_n\jump{\bm{\epsilon}_v},\bm{A}_n\jump{\bm{\eta}_v}}_{\Ldk} + \LRa{\avg{\bm{\epsilon}_v} - \frac{\tau_\sigma}{2}\bm{A}_n^T\jump{\bm{\epsilon}_\sigma},\bm{A}_n^T\jump{\bm{\eta}_{\sigma}}}_{\Ldk}}\\ %- \LRa{\bm{A}_n^T\jump{\bm{\epsilon}_\sigma},\frac{\tau_{\sigma}}{2}\bm{A}_n^T\jump{\bm{\eta}}_\sigma}_{\Ldk} }\\
-&\leq \frac{1}{2}\sum_{D^k\in \Oh}\nor{\avg{\bm{\epsilon}_\sigma} - \frac{\tau_v}{2}\bm{A}_n\jump{\bm{\epsilon}_v}}_{\Ldk} \nor{\bm{A}_n\jump{\bm{\eta}_v}}_{\Ldk} + \nor{\avg{\bm{\epsilon}_v} - \frac{\tau_\sigma}{2}\bm{A}_n^T\jump{\bm{\epsilon}_\sigma}}_{\Ldk}\nor{\bm{A}_n^T\jump{\bm{\eta}_{\sigma}}}_{\Ldk}\\
+&\leq \frac{1}{2}\sum_{D^k\in \Oh}\nor{\avg{\bm{\epsilon}_\sigma} - \frac{\tau_v}{2}\bm{A}_n\jump{\bm{\epsilon}_v}}_{\Ldk} \nor{\bm{A}_n\jump{\bm{\eta}_v}}_{\Ldk} \\
+&\hspace{10em} + \nor{\avg{\bm{\epsilon}_v} - \frac{\tau_\sigma}{2}\bm{A}_n^T\jump{\bm{\epsilon}_\sigma}}_{\Ldk}\nor{\bm{A}_n^T\jump{\bm{\eta}_{\sigma}}}_{\Ldk}\\ 
 %&\leq \LRp{\nor{\avg{\bm{\epsilon}_\sigma} - \frac{\tau_v}{2}\bm{A}_n\jump{\bm{\epsilon}_v}}_{\Ldk}  + \nor{\avg{\bm{\epsilon}_v} - \frac{\tau_\sigma}{2}\bm{A}_n^T\jump{\bm{\epsilon}_\sigma}}_{\LDk}} +  \LRp{\nor{\bm{A}_n\jump{\bm{\eta}_v}}_{\Ldk}^2 + \nor{\bm{A}_n^T\jump{\bm{\eta}_{\sigma}}}_{\Ldk}^2}\\
@@
 where $C_N = O(N^2)$ is the order-dependent constant in the surface polynomial trace inequality \cite{chan2015gpu} and $C_{\rm CFL}$ is a tunable global CFL constant.
-Finally, in all following experiments, we use $\tau_v = \tau_\sigma = 1$ unless specified otherwise.
+Finally, in all following experiments, we use $C_{\rm CFL} = 1$ and $\tau_v = \tau_\sigma = 1$ unless specified otherwise.
 
@@
 	legend cell align=left,
+    legend pos=south east,
 	width=.49\textwidth,
@@
     ymin=1e-12, ymax=1,
-    legend pos=south east,
     xmajorgrids=true,

### 2019-09-06 docs/skewform/arxiv_v2/main_v2.tex
updating docs

--- 2019-03-07 docs/skewform/arxiv/main.tex
@@
 %\usepackage[T1]{fontenc}
+
+%\usepackage[displaymath]{lineno}
 
@@
 \usepackage{cite}
-
-%\usepackage{slashbox} % for slashbox in tables
 
@@
 
-\renewcommand{\note}[1]{{\color{blue}{#1}}}
-
+%\renewcommand{\note}[1]{{\color{blue}{#1}}}
+%\newcommand{\bnote}[1]{{\color{blue}{#1}}}
+%\newcommand{\rnote}[1]{{\color{red}{#1}}}
+\renewcommand{\note}[1]{#1}
+\newcommand{\bnote}[1]{#1}
+\newcommand{\rnote}[1]{#1}
 \graphicspath{{./figs/}}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The construction of such schemes has relied on the use of carefully chosen nodal points \cite{gassner2013skew, fisher2013high, carpenter2014entropy, crean2018entropy, chan2018efficient} or volume and surface quadrature rules \cite{chan2017discretely, chan2018discretely} to produce operators which satisfy a summation-by-parts (SBP) property.
-In this work, we show how to construct ``modal'' skew-symmetric DG formulations which are entropy stable for volume and surface quadratures under which a traditional SBP property does not hold.
-These skew-symmetric formulations avoid the use of a ``strong'' matrix-based SBP property, and instead rely on the assumption that discrete operators exactly differentiate constants and satisfy a discrete form of the fundamental theorem of calculus.
-We conclude with numerical experiments verifying the accuracy and stability of the proposed formulations, and discuss an application of skew-symmetric formulations for entropy stable DG schemes on mixed quadrilateral-triangle meshes.
+In this work, we show how to construct ``modal'' DG formulations which are \bnote{entropy stable for volume and surface quadratures under which the SBP property in \cite{chan2017discretely} does not hold.
+These formulations rely on an alternative skew-symmetric construction of operators which automatically satisfy the SBP property.
+Entropy stability then follows for choices of volume and surface quadrature which satisfy sufficient accuracy conditions.}   \rnote{The accuracy of these new SBP operators depends on a separate set of conditions on quadrature accuracy, with design order accuracy recovered under the usual assumptions of degree $2N-1$ volume quadratures and degree $2N$ surface quadratures.
+}
+We conclude with numerical experiments verifying the accuracy and stability of the proposed formulations, and discuss an application of these formulations for entropy stable DG schemes on mixed quadrilateral-triangle meshes.
 \end{abstract}
@@
 Proofs of continuous entropy inequalities rely on the chain rule, which does not hold discretely due to effects such as quadrature error.
-Entropy stable schemes circumvent the loss of the chain rule while maintaining a semi-discrete analogue of the continuous entropy inequality by using an approach referred to as ``flux differencing''.
-These schemes were first introduced as high order collocation methods on tensor product elements in \cite{fisher2013high, carpenter2014entropy, gassner2016split, gassner2017br1}, and were extended to simplicial elements in \cite{crean2017high, chen2017entropy, crean2018entropy, chan2017discretely, chan2018discretely}.
-Entropy stable methods have also been extended to a variety of other discretization settings, including staggered grids \cite{parsani2016entropy}, Gauss-Legendre collocation \cite{chan2018efficient}, and non-conforming meshes \cite{friedrich2017entropy}.
-
-Entropy stable ``modal'' DG discretizations \cite{chan2017discretely, chan2018discretely} are built upon flux differencing and a summation-by-parts (SBP) property, which serves as a discrete analogue of integration by parts for quadrature-based discretization matrices.
+Entropy stable schemes \rnote{were originally introduced in the context of finite volume methods} \cite{tadmor1987numerical, tadmor2003entropy, fjordholm2012arbitrarily,chandrashekar2013kinetic, tadmor2016entropy, ray2016entropy}.
+They were then extended to high order collocation DG methods on tensor product elements in \cite{fisher2013high, carpenter2014entropy, gassner2016split, gassner2017br1} and to simplicial elements in \cite{crean2017high, chen2017entropy, crean2018entropy, chan2017discretely, chan2018discretely}.
+\rnote{These extensions combine summation-by-parts (SBP) differentiation operators, which satisfy a matrix analogue of integration by parts, with ``flux differencing'' for the discretization of nonlinear convective terms.
+Together, these techniques circumvents the loss of the chain rule while preserving a semi-discrete analogue of the continuous entropy inequality.}
+Entropy stable methods have also been extended to a variety of other discretization settings, including staggered grids \cite{parsani2016entropy, fernandez2019staggered}, Gauss-Legendre collocation \cite{chan2018efficient}, and non-conforming meshes \cite{friedrich2017entropy}.
+
+Entropy stable ``modal'' DG discretizations \cite{chan2017discretely, chan2018discretely} are built upon flux differencing and the SBP property.
 However, the SBP property does not hold for certain under-integrated quadrature rules, which arise naturally in some discretization settings.
@@
 
-This work proposes an alternative formulation which utilizes a skew-symmetric form of flux differencing.
-Under such a formulation, the proof of entropy stability requires only a weaker ``variational'' form of the SBP property, which holds under a more general class of quadrature rules compared to the decoupled SBP property introduced in \cite{chan2017discretely, chan2018discretely}.
-We show that the skew-symmetric formulation is entropy stable, locally conservative, and free-stream preserving on curved elements, and confirm theoretical results with numerical experiments on triangular, quadrilateral, and hybrid triangular-quadrilateral meshes.
-
-It should be noted that a similar approach to entropy stable discretizations was introduced within a finite difference framework \cite{chen2017entropy, crean2018entropy} using multidimensional differencing operators which satisfy accuracy conditions and an SBP property \cite{hicken2016multidimensional}.
+This work proposes an alternative formulation which utilizes a \note{skew-symmetric construction of the SBP operator which satisfies the SBP property by construction}.
+Under such a formulation, the proof of entropy stability \note{holds under weaker quadrature rules} compared to the SBP property introduced in \cite{chan2017discretely, chan2018discretely}.
+We show that this skew-symmetric formulation is entropy stable, locally conservative, and free-stream preserving on curved elements, and confirm theoretical results with numerical experiments on hybrid triangular-quadrilateral meshes.
+
+It should be noted that a similar approach to entropy stable discretizations was introduced within a finite difference framework \cite{chen2017entropy, crean2018entropy} using multidimensional differencing operators which satisfy similar accuracy conditions and an SBP property \cite{hicken2016multidimensional}.
 These operators exist for nodal points corresponding to sufficiently accurate choices of volume and surface quadrature, but do not correspond to any specific basis or approximation space.
@@
 Section~\ref{sec:approx} and Section~\ref{sec:sbp} introduce polynomial approximation spaces and quadrature-based SBP operators on simplicial and tensor product elements.
-Section~\ref{sec:skew1} then introduces the variational SBP property and an entropy conservative skew-symmetric formulation.
-Section~\ref{sec:skew2} extends the skew-symmetric formulation to curved elements, and Section~\ref{sec:quadacc} provides explicit conditions for entropy stability in terms of quadrature accuracy and the polynomial degree of geometric mappings.
+Section~\ref{sec:skew1} introduces \note{an alternative skew-symmetric construction of SBP operators and describes how to construct entropy stable formulations on a reference element.
+Connections between the accuracy of the new skew-symmetric SBP operators and quadrature accuracy are also discussed.}  Section~\ref{sec:skew2} extends the skew-symmetric formulation to curved elements, and provides explicit conditions for entropy stability in terms of quadrature accuracy and the polynomial degree of geometric mappings.
 Section~\ref{sec:num} concludes by presenting numerical experiments which verify the theoretical assumptions, stability, and accuracy of the proposed formulations.
@@
 \begin{equation}
-\pd{\bm{u}_h}{t}  + \sum_{j=1}^d\pd{\bm{f}_j(\bm{u})}{x_j} = 0, \qquad \bm{u}\in \mathbb{R}^n, \qquad \bm{f}:\mathbb{R}^n\rightarrow\mathbb{R}^n,
+\pd{\rnote{\bm{u}}}{t}  + \sum_{j=1}^d\pd{\bm{f}_j(\bm{u})}{x_j} = \bm{0}, \qquad \bm{u}\in \mathbb{R}^n, \qquad \bm{f}:\mathbb{R}^n\rightarrow\mathbb{R}^n,
 \label{eq:nonlineqs}
@@
 
-\section{Quadrature-based matrices and decoupled SBP operators}
+\section{Quadrature-based matrices and ``hybridized'' SBP operators}
 \label{sec:sbp}
@@
 Let $\hat{D} \subset\mathbb{R}^d$ denote a reference element with surface $\partial \hat{D}$.  
-The high order schemes in \cite{chan2017discretely, chan2018discretely} begin by approximating the solution in a degree $N$ polynomial basis $\LRc{\phi_j({\bm{x}})}_{i=1}^{N_p}$ on $\hat{D}$.
-These schemes also assume volume and surface quadrature rules $({\bm{x}}_i, w_i)$, $\LRp{{\bm{x}}^f_i,w^f_i}$ on $\hat{D}$.
-We will specify the accuracy of each quadrature rule later, and discuss how quadrature accuracy implies specific summation-by-parts properties.
+The high order schemes in \cite{chan2017discretely, chan2018discretely} begin by approximating the solution in a degree $N$ polynomial basis $\LRc{\phi_j(\hat{\bm{x}})}_{\rnote{j}=1}^{N_p}$ on $\hat{D}$.
+These schemes also assume volume and surface quadrature rules $(\hat{\bm{x}}_i, w_i)$, $\LRp{\hat{\bm{x}}^f_i,w^f_i}$ on $\hat{D}$.
+We will specify the accuracy of each quadrature rule later, and discuss how quadrature accuracy implies specific operator properties.
 
@@
 \begin{gather}
-\LRp{\bm{V}_q}_{ij} = \phi_j(\bm{x}_i), \qquad \LRp{\bm{V}_f}_{ij} = \phi_j(\bm{x}^f_i), \qquad \pd{\phi_j(\bm{x})}{x_i} = \sum_{k=1}^{N_p} \LRp{\bm{D}^i_{jk}} \phi_k(\bm{x}).
+\LRp{\bm{V}_q}_{ij} = \phi_j(\hat{\bm{x}}_i), \qquad \LRp{\bm{V}_f}_{ij} = \phi_j(\hat{\bm{x}}^f_i), \qquad \pd{\phi_j(\hat{\bm{x}})}{\hat{x}_i} = \sum_{k=1}^{N_p} \LRp{\bm{D}^i_{jk}} \phi_k(\hat{\bm{x}}).
 \end{gather}
-The interpolation matrices $\bm{V}_q,\bm{V}_f$ map basis coefficients to evaluations at volume and surface quadrature points respectively, while the differentiation matrix ${\bm{D}}_i$ maps basis coefficients of a function to the basis coefficients of its derivative with respect to $x_k$.
+The interpolation matrices $\bm{V}_q,\bm{V}_f$ map basis coefficients to evaluations at volume and surface quadrature points respectively, while the differentiation matrix ${\bm{D}}_i$ maps basis coefficients of a function to the basis coefficients of its derivative with respect to $\hat{x}_k$.
 The interpolation matrices are used to assemble the mass matrix $\bm{M}$, the quadrature-based projection matrix $\bm{P}_q$, and lifting matrix $\bm{L}_f$
@@
 where $\bm{W}, \bm{W}_f$ are diagonal matrices of volume and surface quadrature weights, respectively.
-The matrix $\bm{P}_q$ is a quadrature-based discretization of the $L^2$ projection operator $\Pi_N$ onto degree $N$ polynomials, which is given as follows: find $\Pi_N u \in V^N$ such that
+\bnote{We have also assumed that the volume quadrature is sufficiently accurate such that the mass matrix $\bm{M}$ is positive-definite and invertible.} The matrix $\bm{P}_q$ is a quadrature-based discretization of the $L^2$ projection operator $\Pi_N$ onto degree $N$ polynomials, which is given as follows: find $\Pi_N u \in V^N$ such that
 \begin{equation}
@@
 By choosing specific quadrature rules, $\bm{D}^i_q$ recovers high order summation-by-parts finite difference operators in \cite{gassner2013skew, fernandez2014generalized, ranocha2018generalised} and certain operators in \cite{hicken2016multidimensional}.
-However, to address difficulties in designing efficient entropy stable interface terms for nonlinear conservation laws, a new ``decoupled'' summation by parts matrix was introduced in \cite{chan2017discretely} which builds interface terms directly into the approximation of the derivative.
+However, to address difficulties in designing efficient entropy stable interface terms for nonlinear conservation laws, a new ``hybridized'' summation by parts matrix was introduced in \cite{chan2017discretely} which builds interface terms directly into the approximation of the derivative.\footnote{\note{The term ``hybridized'' SBP operator was introduced in the review paper \cite{chenreview}.
+These operators were originally referred to as ``decoupled'' SBP operators in \cite{chan2017discretely})}.}
 
 Let $\hat{\bm{n}}$ denote the scaled outward normal vector $\hat{\bm{n}} = \LRc{\hat{n}_1\hat{J}_f,\ldots,\hat{n}_d\hat{J}_f}$, where $\hat{J}_f$ is the determinant of the Jacobian of the mapping of a face of $\partial \hat{D}$ to a reference face.
-Let $\hat{\bm{n}}_i$ denote the vector containing values of the $i$th component $\hat{n}_i\hat{J}f$ at all surface quadrature points, and let $\bm{Q}^i = \bm{W}\bm{D}^i_q = \bm{W}\bm{V}_q\bm{D}^i\bm{P}_q$.
-The ``decoupled'' summation by parts operator $\bm{D}^i_N$ is defined as the block matrix involving both volume and surface quadratures
+Let $\hat{\bm{n}}_i$ denote the vector containing values of the $i$th component $\hat{n}_i\hat{J}_f$ at all surface quadrature points, and define the generalized SBP operator
+\[
+\bm{Q}^i = \bm{W}\bm{D}^i_q = \bm{W}\bm{V}_q\bm{D}^i\bm{P}_q.
+\]
+The ``hybridized'' summation by parts operator $\bm{Q}^i_N$ is defined as the block matrix involving both volume and surface quadratures
 \begin{gather}
-\bm{D}^i_N = \bm{W}_N^{-1} \bm{Q}^i_N, \qquad \bm{W}_N = \LRp{\begin{array}{cc}\bm{W}&\\ & \bm{W}_f\end{array}}, \label{eq:QN}\\
+%\bm{D}^i_N = \bm{W}_N^{-1} \bm{Q}^i_N, \qquad \bm{W}_N = \LRp{\begin{array}{cc}\bm{W}&\\ & \bm{W}_f\end{array}}, 
 \bm{E} = \bm{V}_f\bm{P}_q, \qquad \bm{B}^i = \bm{W}_f \diag{\hat{\bm{n}}_i}, \qquad \bm{Q}^i_N  = \LRs{
@@
 -\frac{1}{2}\bm{B}^i\bm{E} & \frac{1}{2} \bm{B}^i
-\end{array}}.  \nonumber
+\end{array}}.  \label{eq:QN}
 \end{gather}
@@
 
-It was shown in \cite{chan2017discretely, chan2018efficient} that, when combined with projection and lifting matrices, $\bm{D}^i_N$ produces a high order polynomial approximation of $f\pd{g}{x_i}$. 
-Let $f, g$ be differentiable functions, and let $\bm{f}_i = f(\bm{x}_i)$, $\bm{g}_i = g(\bm{x}_i)$ denote values of $f,g$ at volume and surface quadrature points.
-Then, $f\pd{g}{x_i}$ can be approximated by the polynomial with modal coefficients given by
-\begin{gather}
-f\pd{g}{x_i} \approx \LRs{\begin{array}{cc}
-\bm{P}_q & \bm{L}_f\end{array}} {\rm diag}\LRp{\bm{f}}\bm{D}^i_N \bm{g}.  
-\label{eq:dnapprox}
-\end{gather}
+%Let $f, g$ be differentiable functions, and let $\bm{f}_i = f(\bm{x}_i)$, $\bm{g}_i = g(\bm{x}_i)$ denote values of $f,g$ at volume and surface quadrature points.  Furthermore, define $\bm{D}^i_N$ as
+%\[
+%\bm{D}^i_N = \bm{W}_N^{-1} \bm{Q}^i_N, \qquad \bm{W}_N = \LRp{\begin{array}{cc}\bm{W}&\\ & \bm{W}_f\end{array}}, 
+%\]
+%Then, $f\pd{g}{x_i}$ can be approximated by a polynomial with modal coefficients given by the expression
+%\begin{gather}
+%f\pd{g}{x_i} \approx \LRs{\begin{array}{cc}
+%\bm{P}_q & \bm{L}_f\end{array}} {\rm diag}\LRp{\bm{f}}\bm{D}^i_N \bm{g}.  
+%\label{eq:dnapprox}
+%\end{gather}
 %The approximation can also be interpreted as solving a variational problem.
@@
 
-For sufficiently accurate quadrature rules, the matrix $\bm{Q}^i_N$ also satisfies a decoupled summation-by-parts (SBP) property, which is used to prove semi-discrete entropy stability for nonlinear conservation laws.
+For $\bm{Q}^i$ which satisfy a ``generalized'' SBP property, the matrix $\bm{Q}^i_N$ also satisfies a summation-by-parts (SBP) property, which is used to prove semi-discrete entropy stability for nonlinear conservation laws.
 \begin{theorem}%[Theorem 1 in \cite{chan2017discretely}]
-If $\bm{Q}^i$ satisfies a generalized SBP property
-\[
+If $\bm{Q}^i$ satisfies the generalized SBP property
+\begin{gather}
 \bm{Q}^i = \bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T,
-\]
-then the decoupled SBP operator $\bm{D}^i_N$ (\ref{eq:QN}) satisfies a summation by parts property:
+\label{eq:gsbp}
+\end{gather}
+then the hybridized SBP operator $\bm{Q}^i_N$ (\ref{eq:QN}) satisfies a summation by parts property:
 \begin{gather}
@@
 \[
-\int_{\hat{D}} \pd{u}{\hat{x}_j} v, \qquad \int_{\partial \hat{D}} u v \hat{n}_j
-\]
-are integrated exactly for all $u,v \in V^N\LRp{\hat{D}}$ and $j = 1,\ldots, d$.
+\int_{\hat{D}} \pd{u}{\hat{x}_i} v, \qquad \int_{\partial \hat{D}} u v \hat{n}_i
+\]
+are integrated exactly for all $u,v \in V^N\LRp{\hat{D}}$ and $i = 1,\ldots, d$.
 This implies that Theorem~\ref{lemma:dsbp} is satisfied for sufficiently accurate volume and surface quadratures.
@@
 
-When a generalized SBP property holds for $\bm{Q}^i$, entropy stability can be proven using the strong SBP property in Theorem~\ref{lemma:dsbp} \cite{chan2017discretely, chan2018discretely}.
+When a generalized SBP property holds for $\bm{Q}^i$, entropy stability can be proven using the SBP property in Theorem~\ref{lemma:dsbp} \cite{chan2017discretely, chan2018discretely}.
 The focus of this work is to address cases where the generalized SBP property (and as a result, the SBP property in Theorem~\ref{lemma:dsbp}) do not hold.
@@
 \label{sec:skew1}
-While the SBP property has been used to derive entropy stable schemes, the strong SBP property (\ref{eq:dsbp}) is difficult to enforce in certain discretization settings, such as hybrid and non-conforming meshes.
+While the SBP property has been used to derive entropy stable schemes,  \bnote{it is difficult to enforce the SBP property (\ref{eq:dsbp}) for $\bm{Q}^i_N$} in certain discretization settings, such as hybrid and non-conforming meshes.
 This difficulty is a result of the choices of volume and surface quadrature which naturally arise in these settings.
-We first illustrate how specific pairings of volume and surface quadratures can result in the loss of the SBP property (\ref{eq:dsbp}).
-We then propose a skew-symmetric formulation which is entropy conservative without explicitly requiring operators which satisfy the SBP property.
+We first illustrate how specific pairings of volume and surface quadratures can result in the loss of the SBP property (\ref{eq:dsbp}) \bnote{for $\bm{Q}^i_N$}.
+We then propose an \note{alternative skew-symmetric version of the hybridized SBP operator which satisfies the SBP property by construction}.
+The use of these operators results in formulations which are entropy conservative \bnote{under a wider range of quadratures.}
 
@@
 \centering
-\subfloat[GLL volume quadrature, Gauss surface quadrature]{\includegraphics[width=.4\textwidth]{figs/gllgauss.png}\label{subfig:gllgq}}
+\subfloat[GLL volume quadrature, Gauss surface quadrature]{\includegraphics[width=.4\textwidth]{gllgauss.png}\label{subfig:gllgq}}
 \hspace{2em}
-\subfloat[Degree $2N$ volume quadrature, GLL surface quadrature]{\includegraphics[width=.4\textwidth]{figs/trigll.png}\label{subfig:trigll}}
+\subfloat[Degree $2N$ volume quadrature, GLL surface quadrature]{\includegraphics[width=.4\textwidth]{trigll.png}\label{subfig:trigll}}
 \caption{Volume and surface quadrature pairs which do not satisfy the assumptions of Theorem~\ref{lemma:dsbp}, and thus do not possess the decoupled SBP property (\ref{eq:dsbp}).
@@
 Let $u,v \in Q^N$ denote two arbitrary degree $N$ polynomials.
-The assumptions of Theorem~\ref{lemma:dsbp} are that the volume quadrature exactly integrates $\int_{\hat{D}} \pd{u}{x_j} v$ and that the surface quadrature exactly integrates $\int_{\partial \hat{D}} u v \hat{n}_j$ on $\hat{D}$.
+The assumptions of Theorem~\ref{lemma:dsbp} are that the volume quadrature exactly integrates $\int_{\hat{D}} \pd{u}{x_i} v$ and that the surface quadrature exactly integrates $\int_{\partial \hat{D}} u v \hat{n}_i$ on $\hat{D}$.
 Because the $(N+1)$-point Gauss rule is exact for polynomials of degree $2N+1$ and the product $uv \in P^{2N}$ on each face, the surface quadrature satisfies the assumptions of Theorem~\ref{lemma:dsbp}.
@@
 \captionsetup[subfigure]{width=.425\textwidth}
-\subfloat[Insufficiently accurate surface quadrature on the triangle element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D.png}\label{subfig:hybrid1}}
+\subfloat[Insufficiently accurate surface quadrature on the triangle element.]{\includegraphics[width=.425\textwidth]{hybrid2D.png}\label{subfig:hybrid1}}
 \endgroup
 \hspace{2em}
-\subfloat[Incompatible surface quadrature on the quadrilateral element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D_GQ.png}\label{subfig:hybrid2}}
+\subfloat[Incompatible surface quadrature on the quadrilateral element.]{\includegraphics[width=.425\textwidth]{hybrid2D_GQ.png}\label{subfig:hybrid2}}
 \caption{Examples of interface couplings which do not result in a decoupled SBP property (\ref{eq:dsbp}).
@@
 
-%\begin{figure}
-%\centering
-%\subfloat[Hex-pyramid coupling]{\includegraphics[width=.35\textwidth]{figs/hybrid3D.png}}
-%\hspace{2em}
-%\subfloat[Hex-prism coupling]{\includegraphics[width=.375\textwidth]{figs/hybrid3D_wedge.png}}
-%\caption{Illustration of a 3D coupling between a GLL hexahedral element and a pyramid.  The SBP property does not hold on the pyramid due to the use of GLL quadrature on the quadrilateral face. }
-%\label{fig:hybrid3d}
-%\end{figure}
-%
-%\begin{figure}[!h]
-%\centering
-%\begingroup
-%\captionsetup[subfigure]{width=.45\textwidth}
-%\subfloat[Entropy stable inter-element coupling in \cite{friedrich2017entropy} for a non-conforming interface.]{\raisebox{0em}{\includegraphics[width=.45\textwidth]{figs/nonconSBP.png}}\label{subfig:noncon2}}
-%\hspace{2em}
-%\subfloat[Entropy stable inter-element coupling in this work for a non-conforming interface.
-The dotted black lines denote communication between neighboring elements.]{\raisebox{-0em}{\includegraphics[width=.45\textwidth]{figs/nonconQuad.png}}\label{subfig:noncon1}}
-%\endgroup
-%%\subfloat[3D hexahedra-pyramid coupling]{\includegraphics[width=.31\textwidth]{figs/hybrid3D.png}}
-%\caption{Illustration of two different entropy stable inter-element couplings for $h$ non-conforming meshes.
-Each dashed red line indicates a dyadic flux computation required between two nodes.
-} %The coupling terms in \cite{friedrich2017entropy} requires computing dyadic fluxes between \emph{each} pair of nodes on adjacent interfaces.
-The new approach results in a simpler communication pattern between elements.
+\subsection{\bnote{An alternative construction of hybridized SBP operators}} % variational SBP property}
+
+The property (\ref{eq:dsbp}) relates the polynomial exactness of specific quadrature rules to algebraic properties of quadrature-based matrices.
+We will relax accuracy conditions on these quadrature rules by utilizing \note{an alternative construction of $\bm{Q}^i_N$ based on the skew-symmetric matrix $\bm{Q}^i - \LRp{\bm{Q}^i}^T$.
+
+\begin{lemma}
+\bnote{Let $\tilde{\bm{Q}}^i_{N}$ denote the skew-hybridized SBP operator defined by
+\begin{equation}
+\tilde{\bm{Q}}^i_{N} = \frac{1}{2}\begin{bmatrix}
+\bm{Q}^i - \LRp{\bm{Q}^i}^T & \bm{E}^T\bm{B}^i\\
+-\bm{B}^i\bm{E} & \bm{B}^i
+\end{bmatrix}.
+\label{eq:skewQN} 
+\end{equation}
+Then, $\tilde{\bm{Q}}^i_{N}$ satisfies the SBP property (\ref{eq:dsbp}), and $\tilde{\bm{Q}}^i_N$ and ${\bm{Q}}^i_N$ are identical if $\bm{Q}^i$ satisfies a generalized SBP property (\ref{eq:gsbp}).}
+\end{lemma}
+\begin{proof}
+The SBP property (\ref{eq:dsbp}) holds by construction.  The equivalence between $\tilde{\bm{Q}}^i_N$ and ${\bm{Q}}^i_N$ requires that
+\[
+\frac{1}{2}\LRp{\bm{Q}^i + \LRp{\bm{Q}^i}^T} = \bm{Q}^i - \frac{1}{2}\bm{E}^T\bm{B}^i\bm{E}.
+\]
+Rearranging terms shows that this condition is equivalent to a scaling of the GSBP property (\ref{eq:gsbp})
+\[
+\frac{1}{2}\bm{Q}^i = \frac{1}{2}\LRp{\bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T}.
+\]
+\end{proof}
+While $\tilde{\bm{Q}}^i_N$ is guaranteed to satisfy the SBP property, the accuracy of $\tilde{\bm{Q}}^i_N$ as a differentiation operator now depends on the volume and surface quadrature rules.
+Before analyzing accuracy, we first derive conditions under which it is possible to use $\tilde{\bm{Q}}^i_N$ to construct entropy stable formulations of nonlinear conservation laws.
 }
-%\label{fig:noncon}
-%\end{figure}
-
-\subsection{A variational SBP property}
-
-%The main restriction we wish to overcome
-
-The property (\ref{eq:dsbp}) (which we will refer to as the ``strong'' SBP property) relates the polynomial exactness of specific quadrature rules to algebraic properties of quadrature-based matrices.
-We will relax accuracy conditions on these quadrature rules by utilizing a weaker ``variational'' version of the SBP property variational, which relies on the exactness of volume and surface quadrature rules only for specific polynomials.
-
-On tensor product elements, we restrict ourselves to isotropic volume quadrature rules which are construced from tensor products of one-dimensional quadrature formulas.
-For the remainder of this work, the degree of the multi-dimensional quadrature rule on tensor product elements will refer to the degree of exactness of the one-dimensional rule.
-For example, we refer to the quadrature rule constructed through a tensor product of one-dimensional $(N+1)$-point GLL quadrature rules as a degree $(2N-1)$ quadrature rule.
-This choice of quadrature is sufficient to guarantee that the mass matrix is positive definite \cite{canuto2007spectral}.
-
-Throughout the remainder of this work, we will make the following assumption for specific functions $v(\bm{x})$:
+
+\subsection{Entropy stability on a reference element}
+\label{sec:singleelem}
+
+In this section, we construct so-called ``entropy stable'' schemes on the reference element $\hat{D}$.
+These methods ensure that the entropy inequality (\ref{eq:entropyineq}) is satisfied discretely by avoiding the use of the chain rule in the proof of entropy dissipation.
+Entropy stable schemes rely on two main ingredients: an entropy stable numerical flux as defined by Tadmor \cite{tadmor1987numerical} and a concept referred to as ``flux differencing''.
+Let $\bm{f}_S\LRp{\bm{u}_L,\bm{u}_R}$ be a numerical flux function which is a function of ``left'' and ``right'' states $\bm{u}_L,\bm{u}_R$.
+The numerical flux $\bm{f}_S$ is \textit{entropy conservative} if it satisfies the following three conditions:
+\begin{gather}
+\bm{f}^i_S(\bm{u},\bm{u}) = \bm{f}_i(\bm{u}), \qquad \text{(consistency)}\label{eq:esflux}\\
+\bm{f}^i_S(\bm{u}_L,\bm{u}_R) = \bm{f}^i_S(\bm{u}_R,\bm{u}_R), \qquad \text{(symmetry)}\nonumber\\
+\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}^i_S(\bm{u}_L,\bm{u}_R) = \psi_i(\bm{u}_L) - \psi_i(\bm{u}_R), \qquad \text{(conservation)}\nonumber
+\end{gather}
+for $i = 1,\ldots, d$.  The construction of entropy stable schemes will utilize \rnote{(\ref{eq:esflux})} in discretizations of both volume and surface terms in a DG formulation.
+
+We can now construct a skew-symmetric formulation on the reference element $\hat{D}$ and show that it is semi-discretely entropy conservative \bnote{under one additional condition on $\tilde{\bm{Q}}^i_N$}.
+This formulation can be made entropy stable by adding interface dissipation.
+Let $\bm{u}_h$ denote the discrete solution, and let $\bm{u}_q$ denote the values of the solution at volume quadrature points.
+We define the auxiliary conservative variables $\tilde{\bm{u}}$ in terms of the $L^2$ projections of the entropy variables
+\begin{gather}
+\bm{v}_q = \bm{v}\LRp{\bm{u}_q}, \qquad \tilde{\bm{v}} = \begin{bmatrix}
+\bm{V}_q\\
+\bm{V}_f
+\end{bmatrix}\bm{P}_q\bm{v}_q, \qquad \tilde{\bm{u}} = \bm{u}\LRp{\tilde{\bm{v}}}.
+\end{gather}
+A matrix formulation for (\ref{eq:nonlineqs}) on $\hat{D}$ is given in terms of $\tilde{\bm{u}}$
+\bnote{\begin{gather}
+\bm{M}\td{\bm{u}_h}{t} + \sum_{i=1}^d\LRs{\begin{array}{c}
+\bm{V}_q \\ \bm{V}_f\end{array}}^T 
+\LRp{2\tilde{\bm{Q}}^i_N \circ \bm{F}^i_S}\bm{1} + \bm{V}_f^T\bm{B}^i\LRp{\bm{f}_i^*-\bm{f}(\tilde{\bm{u}}_f)} = 0,  \label{eq:esdgSkew}\\
+\LRp{\bm{F}^i_S}_{jk} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}, \qquad 1\leq j,k \leq N_q + N^f_q,\nonumber
+\end{gather}
+where $\tilde{\bm{u}}_f$ denotes the values of $\tilde{\bm{u}}$ on face nodes and $\bm{f}^*$ is some numerical flux, and $N_q, N^f_q$ denote the number of volume and face quadrature points, respectively.	.
+This formulation is identical to that of \cite{chan2017discretely}, except that the hybridized SBP operators $\bm{Q}^i_N$ are replaced with their skew-hybridized versions $\tilde{\bm{Q}}^i_N$.
+For this reason, we refer to (\ref{eq:esdgSkew}) as the ``skew-symmetric'' formulation.
+Under the condition that $\tilde{\bm{Q}}^i_N \bm{1} = \bm{0}$, the formulation (\ref{eq:esdgSkew}) is entropy conservative over $\hat{D}$:}
+\begin{theorem}
+%Suppose Assumption~\ref{ass:quad} holds for $v(\bm{x}) = 1$.  
+\bnote{Assume that $\tilde{\bm{Q}}^i_N \bm{1} = \bm{0}$.  }
+Then, the formulation (\ref{eq:esdgSkew}) is entropy conservative such that
+\begin{equation}
+\bm{1}^T\bm{W}\td{U(\bm{u}_q)}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}^i \LRp{\tilde{\bm{v}}_f^T\bm{f}_i^* - \psi_i(\tilde{\bm{u}}_f)} = 0, \qquad \bm{u}_q = \bm{V}_q\bm{u}_h.
+\label{eq:esdgthm}
+\end{equation}
+\rnote{Here, $\psi_i(\tilde{\bm{u}}_f)$ denotes the function $\psi_i$ evaluated at the face values of the entropy-projected conservative variables $\tilde{\bm{u}}_f$}.
+\label{thm:esdg}
+\end{theorem}
+The steps of the proof are identical to those of Theorem 2 in \cite{chan2017discretely}, and we skip them for brevity.
+
+%\begin{proof}
+%Testing (\ref{eq:esdgSkew}) by $\bm{v}_h = \bm{P}_q\bm{v}_q$ yields 
+%\begin{align}
+%\bm{v}_q^T\bm{W}\pd{\LRp{\bm{V}_q\bm{u}}}{t} + \sum_{i=1}^d
+%\tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}^i_S}\bm{1} + \tilde{\bm{v}}_f^T \bm{B}^i\bm{f}_i^* = 0.
+%\label{eq:entropyrhs}
+%\end{align}
+%Algebraic manipulations (see \cite{chan2017discretely} for details) yields that 
+%\begin{align*}
+%\tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}^i_S}\bm{1} &= \tilde{\bm{v}}^T \LRp{\bm{Q}^i_N \circ \bm{F}^i_S}\bm{1} - \tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N}^T \circ \bm{F}^i_S}\bm{1}\\
+%&= \tilde{\bm{v}}^T \LRp{\bm{Q}^i_N \circ \bm{F}^i_S}\bm{1} - \bm{1}^T \LRp{{\bm{Q}^i_N} \circ \bm{F}^i_S}\tilde{\bm{v}}.
+%\end{align*}
+%Where we have used that $\bm{F}^i_S$ is symmetric and that the Hadamard product commutes.  Applying the conservation condition on $\bm{f}^i_S$ in (\ref{eq:esflux}) then yields
+%\begin{align*}
+%&\tilde{\bm{v}}^T \LRp{\bm{Q}^i_N \circ \bm{F}^i_S}\bm{1} - \bm{1}^T \LRp{{\bm{Q}^i_N} \circ \bm{F}^i_S}\tilde{\bm{v}} \\
+%&= \sum_{jk} \LRp{\bm{Q}^i_N}_{jk} \LRp{\tilde{\bm{v}}_j-\tilde{\bm{v}}_k} \bm{f}_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k} = \sum_{jk} \LRp{\bm{Q}^i_N}_{jk} \LRp{\psi_i(\tilde{\bm{u}}_j) - \psi_i(\tilde{\bm{u}}_k)}\\
+%&= \bm{1}^T\LRp{\bm{Q}^i_N}\psi_i(\tilde{\bm{u}}) - \psi_i(\tilde{\bm{u}})^T\LRp{\bm{Q}^i_N}\bm{1} = \bm{1}^T\LRp{\bm{Q}^i_N}\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i_N\psi_i(\tilde{\bm{u}}),
+%\end{align*}
+%where we have used Lemma~\ref{lemma:sbpcor} in the last equality.  Noting that $\bm{1}^T\bm{B}^i_N\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i \psi_i(\tilde{\bm{u}}_f)$ completes the proof.
+%%The final step of the proof is  where we have used Lemma~\ref{lemma:vdsbp} to integrate by parts and conclude that $\bm{1}^T\LRp{\bm{Q}^i_N}^T = 0$.
+%\qed\end{proof}
+
+\begin{remark}
+We note that (\ref{eq:esdgSkew}) is also equivalent to the following skew-symmetric formulation:
+\begin{gather}
+\bm{M}\td{\bm{u}_h}{t} + \sum_{i=1}^d\LRs{\begin{array}{c}
+\bm{V}_q \\ \bm{V}_f\end{array}}^T 
+\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}^i_S}\bm{1} + \bm{V}_f^T\bm{B}^i\bm{f}_i^* = 0,  \label{eq:esdgSkew2}\\
+\LRp{\bm{F}^i_S}_{jk} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}, \qquad 1\leq j,k \leq N_q + N^f_q,\nonumber
+\end{gather}
+where the skew-symmetric matrix $\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}$ possesses the following block structure:
+\[
+\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} = \begin{pmatrix}
+\bm{Q}_i-\bm{Q}_i^T & {\bm{E}}^T \bm{B}^i\\
+-\bm{B}^i\bm{E} & \bm{0}
+\end{pmatrix}.
+\]
+\end{remark}
+
+The skew symmetric formulation can also be shown to be locally conservative in the sense of \cite{shi2017local}, which is necessary to prove that the numerical solution convergences to the weak solution under mesh refinement.
+\begin{theorem}
+The formulation (\ref{eq:esdgSkew}) is locally conservative such that
+\begin{align}
+\bm{1}^T\bm{W}\td{\LRp{\bm{V}_q\bm{u}}}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}^i\bm{f}_i^* = 0. 
+\end{align}
+\end{theorem}
+\begin{proof}
+To show local conservation, we test (\ref{eq:esdgSkew}) with $1$
+\begin{align}
+\bm{1}^T\bm{W}\bm{V}_q\td{\bm{u}_h}{t} + \sum_{i=1}^d
+\bm{1}^T
+\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}}\bm{f}_i^* = 0. 
+\end{align}
+Because $\bm{F}_S$ is symmetric and $\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}$ is skew-symmetric, their Hadamard product is also skew-symmetric.
+Using that $\bm{x}^T\bm{A}\bm{x} = 0$ for any skew symmetric matrix $\bm{A}$, the volume term $\bm{1}^T\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1}$ vanishes.
+\qed\end{proof}
+
+\subsection{\bnote{Properties of $\tilde{\bm{Q}}^i_N$ and quadrature accuracy}}
+
+\bnote{The proof of the semi-discrete entropy inequality in Theorem~\ref{thm:esdg} requires both the SBP condition (\ref{eq:dsbp}) and that $\tilde{\bm{Q}}^i_N\bm{1} = \bm{0}$.
+While the SBP condition is guaranteed by construction, $\tilde{\bm{Q}}^i_N\bm{1} = \bm{0}$ only holds under sufficiently accurate quadrature rules.
+In \cite{chan2017discretely}, it was shown that the hybridized SBP operator satisfies $\bm{Q}^i_N\bm{1} = \bm{0}$ for any volume quadrature such that the mass matrix $\bm{M}$ is positive-definite.
+However, ensuring that the skew-hybridized SBP operator satisfies $\tilde{\bm{Q}}^i_N\bm{1} = \bm{0}$ now requires conditions on both volume and surface quadratures which are related to a weak version of the generalized SBP condition (\ref{eq:gsbp}).
+}
+
+Throughout the remainder of this work, we will assume that the \bnote{volume and surface quadrature satisfy the following assumptions} for specific functions $v(\bm{x})$:
 \begin{assumption}
@@
 The conditions of Assumption~\ref{ass:quad} are relatively standard within the SBP literature \cite{hicken2016multidimensional, chan2017discretely, crean2018entropy}, though they have not previously depended on the specific choice of polynomial $v(\bm{x})$.
-The strong SBP property (\ref{eq:dsbp}) can be derived if Assumption~\ref{ass:quad} is satisfied for all $v \in V^N$.
+\bnote{The following theorem shows how these accuracy conditions are related to the condition $\tilde{\bm{Q}}^i_N\bm{1} = \bm{0}$.}%The SBP property (\ref{eq:dsbp}) can be derived if Assumption~\ref{ass:quad} is satisfied for all $v \in V^N$.
 In contrast, we will derive a weaker SBP property which assumes only that Assumption~\ref{ass:quad} holds for a few specific polynomials $v$.
-In Sections~\ref{sec:singleelem} and \ref{sec:curved}, specific polynomials $v$ will be motivated by proofs of entropy stability, and we will present examples of volume and surface quadrature rules on simplicial and tensor product elements which satisfy Assumption~\ref{ass:quad} for these choices of $v$.
-% the quantities $\int_{\hat{D}} \pd{u}{x_i} v$ and
-
-%This section presumes that Assumption~\ref{ass:quad} is satisfied for general  $M_{\rm vol}, M_{\rm surf}$.
-In Sections~\ref{sec:singleelem} and \ref{sec:curved}, specific values of $M_{\rm vol}, M_{\rm surf}$ will be motivated by proofs of entropy stability, and Assumption~\ref{ass:quad} will be verified for specific choices of volume and surface quadrature on simplicial and tensor product elements.
-
-The conditions of Assumption~\ref{ass:quad} imply the following variational SBP property:
-\begin{lemma}
-\label{lemma:vsbp}
-%Assume that $v \in V^M$.  
-%If the volume quadrature is exact for polynomials of degree $N+M-1$ and the surface quadrature is exact for polynomials of degree $N+M$, 
-Suppose Assumption~\ref{ass:quad} holds for $v(\bm{x})\in V^N$, and let $\bm{v}_q$ denote the vector of values of $v(\bm{x})$ of at volume quadrature points.
-Let $u(\bm{x})$ be an $L^2$ integrable function, and let $\bm{u}_q$ denote the values of $u(\bm{x})$ at quadrature points.
-Then,
-\[
-\bm{v}_q^T\bm{Q}_i \bm{u}_q = \bm{v}_q^T\LRp{ {\bm{E}}^T \bm{B}^i\bm{E} - \bm{Q}_i^T}\bm{u}_q.
-\]
+\begin{lemma}%{(Discrete FTC and exact differentiation of constants)}
+\label{lemma:sbpcor}
+\bnote{Suppose Assumption~\ref{ass:quad} holds for $v(\bm{x}) = 1$.  Then, $\tilde{\bm{Q}}^i_N\bm{1} = \bm{0}$.}
+%Suppose Assumption~\ref{ass:quad} holds for $v(\bm{x}) = 1$.  
+%Let $u(\bm{x})$ be some $L^2$ integrable function, and let $\bm{u}$ denote the vector of values of $u(\bm{x})$ at volume and surface quadrature points. Then, 
+%\[
+%\bm{1}^T\bm{Q}^i_N\bm{u} = \bm{1}^T\bm{B}^i_N\bm{u}, \qquad \bm{Q}^i_N\bm{1} = \bm{0}.
+%\]
 \end{lemma}
 \begin{proof}
-Recall that $\bm{Q}_i = \bm{W} \bm{V}_q \bm{D}^i\bm{P}_q$.
-The quadrature-based $L^2$ projection $\Pi_Nu$ is a polynomial of degree $N$, and is computed discretely as $\bm{P}_q\bm{u}_q$.
-By the exactness of the volume quadrature in Assumption~\ref{ass:quad},
-\begin{align*}
-\bm{v}_q^T\bm{Q}_i \bm{u}_q &= \bm{v}_q^T\bm{W} \bm{D}^i \bm{P}_q \bm{u}_q = \int_{\hat{D}} \pd{\Pi_N u}{\hat{x}_i} v = \int_{\partial \hat{D}} (\Pi_N u) v \hat{n}_i - \int_{\hat{D}} \LRp{\Pi_N u} \pd{v}{\hat{x}_i}.
-\end{align*}
-%Since the volume integrand $\LRp{\Pi_N u} \pd{v}{\hat{x}_i} \in V^{N+M-1}$, it is exactly integrated using the volume quadrature.  
-%Additionally, since $\Pi_N u \in V^N$ and $v\in V^M$ on the surface of $\hat{D}$, the surface quadrature is exact for the surface integral and 
-By the exactness of the surface quadrature in Assumption~\ref{ass:quad}, 
-\begin{align*}
-\int_{\partial \hat{D}} (\Pi_N u) v \hat{n}_i -   \int_{\hat{D}} \LRp{\Pi_N u} \pd{v}{\hat{x}_i} &=\bm{v}_f^T\bm{B}^i \bm{E}\bm{u}_q - \LRp{\bm{V}_q\bm{D}^i\bm{P}_q\bm{v}_q}^T\bm{W} \bm{V}_q\bm{P}_q\bm{u}_q\\
-=& \bm{v}_q^T {\bm{E}}^T\bm{B}^i \bm{E}\bm{u}_q - \bm{v}_q^T \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T \bm{W} \bm{V}_q\bm{P}_q\bm{u}_q,
-\end{align*}
-where we have used that, since $v \in V^N$, $\bm{v}_f = \bm{V}_f\bm{P}_q\bm{v}_q = \bm{E}\bm{v}_q$.  The proof is completed by noting that 
-\[
-\bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T \bm{W} \bm{V}_q\bm{P}_q = \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{M}\bm{P}_q = \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T\bm{W} = \bm{Q}_i^T.
-\]
+\bnote{
+Expanding out $\tilde{\bm{Q}}^i_N\bm{1}$ yields
+\[
+\tilde{\bm{Q}}^i_N\bm{1} = \frac{1}{2}\begin{bmatrix}
+\bm{Q}^i\bm{1} - \LRp{\bm{Q}^i}^T \bm{1} + \bm{E}^T\bm{B}^i \bm{1}\\
+-\bm{B}^i\bm{E}\bm{1} + \bm{B}^i\bm{1}\end{bmatrix}.
+\]
+Here, $\bm{1}$ denotes the appropriate length vector with all entries equal to one.  
+Since polynomials are equal to their $L^2$ projection, $\bm{E}\bm{1} = \bm{1}$ \cite{chan2017discretely,chan2018discretely}, and 
+\[
+-\bm{B}^i\bm{E}\bm{1} + \bm{B}^i\bm{1} = \bm{0}.  
+\]
+Moreover, since $\bm{Q}^i$ is a differentation matrix, $\bm{Q}^i\bm{1} = \bm{0}$, and showing $\tilde{\bm{Q}}^i_N\bm{1} = \bm{0}$ reduces to showing that 
+\[
+\LRp{\bm{Q}^i}^T \bm{1} = \bm{E}^T\bm{B}^i \bm{1}.
+\]
+However, under Assumption~\ref{ass:quad}, the entries of $\LRp{\bm{Q}^i}^T \bm{1}$ are exactly $\int_{\hat{D}} \pd{\phi_j}{\hat{\bm{x}}_i}$ and the entries of $\bm{E}^T\bm{B}^i \bm{1}$ are exactly $\int_{\partial \hat{D}} \phi_j(\bm{x}) \hat{n}_i$ since $\phi_j(\bm{x}) \in V^N$.
+These two terms are then identical by the exactness of integrals and fundamental theorem of calculus.
+}
+
+%The proof that $\bm{Q}^i_N \bm{1} = \bm{0}$ follows from the property that .     The proof of the first equality follows from $\bm{Q}^i_N \bm{1} = \bm{0}$ and Lemma~\ref{lemma:vdsbp} 
+%\[
+%\bm{1}^T\LRp{\bm{Q}^i_N}\bm{u} = \bm{1}^T\LRp{\bm{B}^i_N - \LRp{\bm{Q}^i_N}^T}\bm{u} = \bm{1}^T{\bm{B}^i_N}\bm{u}.
+%\]
 \qed\end{proof}
-
-The variational SBP property in Lemma~\ref{lemma:vsbp} can be used to prove a similar property for the decoupled SBP operator.
-This property, along with the exact differentiation of constants, is necessary for the proof of entropy stability.
-\begin{lemma}
-Suppose Assumption~\ref{ass:quad} holds.  Let $\bm{D}^i_N$ be a decoupled SBP operator on the reference element $\hat{D}$.  
-%Let the volume quadrature be exact for polynomials of degree $N+M-1$, and let the surface quadrature be exact for polynomials of degree $N+M$ for some $M \leq N$.
-Suppose $v\in V^N$ and that $\LRl{v}_{\partial \hat{D}} \in P^M\LRp{\partial \hat{D}}$.
-\note{Fix this to match assumptions in previous lemma.}
-Then, the following variational SBP property holds:
-\[
-\bm{v}^T\bm{Q}^i_N\bm{u} = \bm{v}^T\LRp{\bm{B}^i_N - \LRp{\bm{Q}^i_N}^T}\bm{u}.%, \qquad \bm{Q}^i_N \bm{1} = \bm{0}.
-\]
-where $\bm{v}, \bm{u}$ denotes the values of $v$ and $u$ at both volume and surface quadrature points.  
-\label{lemma:vdsbp}
-\end{lemma}
-\begin{proof}
-For convenience, let $\bm{u}_q, \bm{u}_f$ and $\bm{v}_q, \bm{v}_f$ denote evaluations of $u$ and $v$ at volume and surface points, such that 
-\[
-\bm{u} = \begin{pmatrix} \bm{u}_q\\ \bm{u}_f\end{pmatrix}, \qquad \bm{v} = \begin{pmatrix} \bm{v}_q\\ \bm{v}_f\end{pmatrix}.  
-\]
-The proof of the variational summation by parts property uses the definition of $\bm{Q}^i_N$ (\ref{eq:QN}), 
-\begin{align*}
-\bm{v}^T\bm{Q}^i_N\bm{u} &= \bm{v}_q^T\bm{Q}_i \bm{u}_q - \frac{1}{2}\LRp{\bm{E}\bm{v}_q}^T\bm{B}^i \LRp{\bm{E}\bm{u}_q} + \frac{1}{2}\LRp{\bm{E}\bm{v}_q}^T\bm{B}^i \bm{u}_f\\
-& - \frac{1}{2}\bm{v}_f^T\bm{B}^i \LRp{\bm{E}\bm{u}_q} + \frac{1}{2}\bm{v}_f^T\bm{B}^i \bm{u}_f.
-\end{align*}
-Applying Lemma~\ref{lemma:vsbp} then yields
-\begin{align*}
-\bm{v}^T\bm{Q}^i_N\bm{u} =& -\bm{v}_q^T\bm{Q}^T_i \bm{u}_q + \frac{1}{2}\LRp{\bm{E}\bm{v}_q}^T\bm{B}^i \LRp{\bm{E}\bm{u}_q} + \frac{1}{2}\LRp{\bm{E}\bm{v}_q}^T\bm{B}^i \bm{u}_f\\
-&\qquad - \frac{1}{2}\bm{v}_f^T\bm{B}^i \LRp{\bm{E}\bm{u}_q} - \frac{1}{2}\bm{v}_f^T\bm{B}^i \bm{u}_f + \bm{v}_f^T\bm{B}^i \bm{u}_f\\
-=& \begin{pmatrix} \bm{v}_q\\ \bm{v}_f\end{pmatrix}^T 
-\left(\begin{pmatrix}
-\bm{0}& \\
-& \bm{B}^i
-\end{pmatrix}\right.+
-\left.\begin{pmatrix}
--\bm{Q}_i^T + \frac{1}{2}{\bm{E}}^T\bm{B}^i \bm{E} & -\frac{1}{2} \LRp{\bm{B}^i \bm{E}}^T\\
-\frac{1}{2}\bm{B}^i \bm{E} & -\frac{1}{2}\bm{B}^i
-\end{pmatrix}  \right)
-\begin{pmatrix} \bm{u}_q\\ \bm{u}_f\end{pmatrix}\\
-=& \bm{v}^T\LRp{\bm{B}^i_N - \LRp{\bm{Q}^i_N}^T}\bm{u}.
-\end{align*}
-\qed\end{proof}
-
-\subsection{Entropy stability on a reference element}
-\label{sec:singleelem}
-
-In this section, we construct so-called ``entropy stable'' schemes on the reference element $\hat{D}$.
-These methods ensure that the entropy inequality (\ref{eq:entropyineq}) is satisfied discretely by avoiding the use of the chain rule in the proof of entropy dissipation.
-Entropy stable schemes rely on two main ingredients: an entropy stable numerical flux as defined by Tadmor \cite{tadmor1987numerical} and a concept referred to as ``flux differencing''.
-Let $\bm{f}_S\LRp{\bm{u}_L,\bm{u}_R}$ be a numerical flux function which is a function of ``left'' and ``right'' states $\bm{u}_L,\bm{u}_R$.
-The numerical flux $\bm{f}_S$ is \textit{entropy conservative} if it satisfies the following three conditions:
-\begin{gather}
-\bm{f}^i_S(\bm{u},\bm{u}) = \bm{f}_i(\bm{u}), \qquad \text{(consistency)}\\
-\bm{f}^i_S(\bm{u}_L,\bm{u}_R) = \bm{f}^i_S(\bm{u}_R,\bm{u}_R), \qquad \text{(symmetry)}\nonumber\\
-\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}^i_S(\bm{u}_L,\bm{u}_R) = \psi_i(\bm{u}_L) - \psi_i(\bm{u}_R), \qquad \text{(conservation)}\nonumber
-\label{eq:esflux}
-\end{gather}
-for $j = 1,\ldots, d$.  The construction of entropy stable schemes will utilize (\ref{eq:esflux}) in discretizations of both volume and surface terms in a DG formulation.
-
-Proving a semi-discrete entropy inequality typically involves the ``strong'' SBP condition.
-The schemes presented in this work will relax this, and use only two (weaker) conditions: that the decoupled SBP operator $\bm{D}^i_N$ exactly differentiates constants and that a discrete form of the fundamental theorem of calculus (FTC) holds.
-The latter property can be derived from the variational SBP property in Lemma~\ref{lemma:vdsbp}.
-
-\begin{lemma}{(Discrete FTC and exact differentiation of constants)}
-\label{lemma:sbpcor}
-Suppose Assumption~\ref{ass:quad} holds for $v(\bm{x}) = 1$.  
-Let $u(\bm{x})$ be some $L^2$ integrable function, and let $\bm{u}$ denote the vector of values of $u(\bm{x})$ at volume and surface quadrature points. Then, 
-\[
-\bm{1}^T\bm{Q}^i_N\bm{u} = \bm{1}^T\bm{B}^i_N\bm{u}, \qquad \bm{Q}^i_N\bm{1} = \bm{0}.
-\]
-\end{lemma}
-\begin{proof}
-The proof that $\bm{Q}^i_N \bm{1} = \bm{0}$ follows from the property that polynomials are equal to their $L^2$ projection, and is identical to that of \cite{chan2017discretely,chan2018discretely}.
-The proof of the first equality follows from $\bm{Q}^i_N \bm{1} = \bm{0}$ and Lemma~\ref{lemma:vdsbp}
-\[
-\bm{1}^T\LRp{\bm{Q}^i_N}\bm{u} = \bm{1}^T\LRp{\bm{B}^i_N - \LRp{\bm{Q}^i_N}^T}\bm{u} = \bm{1}^T{\bm{B}^i_N}\bm{u}.
-\]
-\qed\end{proof}
-
-We can now construct a skew-symmetric formulation on the reference element $\hat{D}$ and show that it is semi-discretely entropy conservative.
-This formulation can be made entropy stable by adding interface dissipation.
-Let $\bm{u}_h$ denote the discrete solution, and let $\bm{u}_q$ denote the values of the solution at volume quadrature points.
-We define the auxiliary conservative variables $\tilde{\bm{u}}$ in terms of the $L^2$ projections of the entropy variables
-\begin{gather}
-\bm{v}_q = \bm{v}\LRp{\bm{u}_q}, \qquad \tilde{\bm{v}} = \begin{bmatrix}
-\bm{V}_q\\
-\bm{V}_f
-\end{bmatrix}\bm{P}_q\bm{v}_q, \qquad \tilde{\bm{u}} = \bm{u}\LRp{\tilde{\bm{v}}}.
-\end{gather}
-The skew-symmetric formulation for (\ref{eq:nonlineqs}) on $\hat{D}$ is given in terms of $\tilde{\bm{u}}$
-\begin{gather}
-\bm{M}\pd{\bm{u}_h}{t} + \sum_{i=1}^d\LRs{\begin{array}{c}
-\bm{V}_q \\ \bm{V}_f\end{array}}^T 
-\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}^i_S}\bm{1} + \bm{V}_f^T\bm{B}^i\bm{f}_i^* = 0,  \label{eq:esdgSkew}\\
-\LRp{\bm{F}^i_S}_{ij} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q,\nonumber
-\end{gather}
-where $\bm{f}^*$ is some numerical flux, and the matrix $\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}$ possesses the following block structure:
-\[
-\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} = \begin{pmatrix}
-\bm{Q}_i-\bm{Q}_i^T & {\bm{E}}^T \bm{B}^i\\
--\bm{B}^i\bm{E} & \bm{0}
-\end{pmatrix}.
-\]
-Multiplying the formulation (\ref{eq:esdgSkew}) by $\bm{M}^{-1}$ on both sides yields a strong form 
-\begin{gather}
-\pd{\bm{u}_h}{t} + \sum_{i=1}^d \LRs{\begin{array}{cc}
-\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}^i_S}\bm{1} + \bm{L}_f \diag{\hat{\bm{n}}_i}\bm{f}_i^* = 0,\label{eq:esdg}
-%\\
-%\LRp{\bm{F}_S}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q.
-%\label{eq:esdgSkewStrong}
-\end{gather}
-We can now show that the skew-symmetric formulation is entropy conservative.  
-\begin{theorem}
-Suppose Assumption~\ref{ass:quad} holds for $v(\bm{x}) = 1$.  
-Then, the formulation (\ref{eq:esdgSkew}) is entropy conservative such that
-\begin{equation}
-\bm{1}^T\bm{W}\pd{U(\bm{u}_q)}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}^i \LRp{\tilde{\bm{v}}_f^T\bm{f}_i^* - \psi_i(\tilde{\bm{u}}_f)} = 0, \qquad \bm{u}_q = \bm{V}_q\bm{u}.
-\label{eq:esdgthm}
-\end{equation}
-\label{thm:esdg}
-\end{theorem}
-\begin{proof}
-Testing (\ref{eq:esdgSkew}) by $\bm{v}_h = \bm{P}_q\bm{v}_q$ yields 
-\begin{align}
-\bm{v}_q^T\bm{W}\pd{\LRp{\bm{V}_q\bm{u}}}{t} + \sum_{i=1}^d
-\tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}^i_S}\bm{1} + \tilde{\bm{v}}_f^T \bm{B}^i\bm{f}_i^* = 0.
-\label{eq:entropyrhs}
-\end{align}
-Algebraic manipulations (see \cite{chan2017discretely} for details) yields that 
-\begin{align*}
-\tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}^i_S}\bm{1} &= \tilde{\bm{v}}^T \LRp{\bm{Q}^i_N \circ \bm{F}^i_S}\bm{1} - \tilde{\bm{v}}^T \LRp{\LRp{\bm{Q}^i_N}^T \circ \bm{F}^i_S}\bm{1}\\
-&= \tilde{\bm{v}}^T \LRp{\bm{Q}^i_N \circ \bm{F}^i_S}\bm{1} - \bm{1}^T \LRp{{\bm{Q}^i_N} \circ \bm{F}^i_S}\tilde{\bm{v}}.
-\end{align*}
-Where we have used that $\bm{F}^i_S$ is symmetric and that the Hadamard product commutes.  Applying the conservation condition on $\bm{f}^i_S$ in (\ref{eq:esflux}) then yields
-\begin{align*}
-&\tilde{\bm{v}}^T \LRp{\bm{Q}^i_N \circ \bm{F}^i_S}\bm{1} - \bm{1}^T \LRp{{\bm{Q}^i_N} \circ \bm{F}^i_S}\tilde{\bm{v}} \\
-&= \sum_{jk} \LRp{\bm{Q}^i_N}_{jk} \LRp{\tilde{\bm{v}}_j-\tilde{\bm{v}}_k} \bm{f}_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k} = \sum_{jk} \LRp{\bm{Q}^i_N}_{jk} \LRp{\psi_i(\tilde{\bm{u}}_j) - \psi_i(\tilde{\bm{u}}_k)}\\
-&= \bm{1}^T\LRp{\bm{Q}^i_N}\psi_i(\tilde{\bm{u}}) - \psi_i(\tilde{\bm{u}})^T\LRp{\bm{Q}^i_N}\bm{1} = \bm{1}^T\LRp{\bm{Q}^i_N}\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i_N\psi_i(\tilde{\bm{u}}),
-\end{align*}
-where we have used Lemma~\ref{lemma:sbpcor} in the last equality.  Noting that $\bm{1}^T\bm{B}^i_N\psi_i(\tilde{\bm{u}}) = \bm{1}^T\bm{B}^i \psi_i(\tilde{\bm{u}}_f)$ completes the proof.
-%The final step of the proof is  where we have used Lemma~\ref{lemma:vdsbp} to integrate by parts and conclude that $\bm{1}^T\LRp{\bm{Q}^i_N}^T = 0$.
-\qed\end{proof}
-
-\begin{remark}
-The primary difference between this formulation and the formulation presented in \cite{chan2017discretely} is the use of the skew-symmetric volume operator $\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}$.
-When the strong SBP property (\ref{eq:dsbp}) holds, this formulation is exactly equivalent to the strong formulation presented in \cite{chan2017discretely}.
-However, the only result necessary to prove Theorem~\ref{thm:esdg} is Lemma~\ref{lemma:sbpcor}.
-The decoupled SBP property (\ref{eq:dsbp}) is not utilized, and is not necessary to guarantee a semi-discrete conservation of entropy for the skew-symmetric formulation.
-\end{remark}
-
-The skew symmetric formulation can also be shown to be locally conservative in the sense of \cite{shi2017local}, which is sufficient to show the numerical solution convergences to the weak solution under mesh refinement.
-\begin{theorem}
-The formulation (\ref{eq:esdgSkew}) is locally conservative such that
-\begin{align}
-\bm{1}^T\bm{W}\pd{\LRp{\bm{V}_q\bm{u}}}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}^i\bm{f}_i^* = 0. 
-\end{align}
-\end{theorem}
-\begin{proof}
-To show local conservation, we test (\ref{eq:esdgSkew}) with $1$
-\begin{align}
-\bm{1}^T\bm{W}\bm{V}_q\pd{\bm{u}_h}{t} + \sum_{i=1}^d
-\bm{1}^T
-\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{1}^T\bm{W}_f \diag{\hat{\bm{n}}}\bm{f}_i^* = 0. 
-\end{align}
-Because $\bm{F}_S$ is symmetric and $\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T}$ is skew-symmetric, their Hadamard product is also skew-symmetric.
-Using that $\bm{x}^T\bm{A}\bm{x} = 0$ for any skew symmetric matrix $\bm{A}$, the volume term $\bm{1}^T\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1}$ vanishes.
-% and that the Hadamard product is commutable $\LRp{\bm{A}\circ\bm{B} }^T = \bm{A}^T\circ\bm{B}^T$
-%\begin{align}
-%\bm{1}^T\LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} &= \bm{1}^T\LRp{\bm{Q}^i_N\circ \bm{F}_S}\bm{1} - \bm{1}^T\LRp{\LRp{\bm{Q}^i_N}^T\circ \bm{F}_S}\bm{1}\\
-%&= \bm{1}^T\LRp{\bm{Q}^i_N\circ \bm{F}_S}\bm{1} - 
-%\bm{1}^T\LRp{\bm{Q}^i_N\circ \bm{F}_S}\bm{1} = 0,
-%\end{align}
-%where we have used that $\bm{F}_S$ is symmetric.
-%\begin{align}
-%\end{align}
-\qed\end{proof}
+In Sections~\ref{sec:singleelem} and \ref{sec:curved}, specific polynomials $v(\bm{x})$ will be motivated by \note{the extension of the proof of entropy stability on curved elements}, and we will present examples of volume and surface quadrature rules on simplicial and tensor product elements which satisfy Assumption~\ref{ass:quad} for these choices of $v$.
 
@@
 \label{sec:assump1}
-Apart from algebraic manipulations, only Lemma~\ref{lemma:sbpcor} is necessary to prove entropy conservation for (\ref{eq:esdg}).
+Apart from algebraic manipulations, only Lemma~\ref{lemma:sbpcor} is necessary to prove entropy conservation in Theorem~\ref{thm:esdg}.
 Lemma~\ref{lemma:sbpcor} requires that Assumption~\ref{ass:quad} holds for $v=1$.
@@
 
-%\note{link SBP accuracy section} %These issues will be discussed further in Section~\ref{sec:num}.
-%This minimal degree of surface quadrature accuracy will depend on the nature of the physical mesh, and will differ between affine and curved elements.
+On tensor product elements, we restrict ourselves to isotropic volume quadrature rules which are construced from tensor products of one-dimensional quadrature formulas.
+For the remainder of this work, the degree of the multi-dimensional quadrature rule on tensor product elements will refer to the degree of exactness of the one-dimensional rule.
+For example, we refer to the quadrature rule constructed through a tensor product of one-dimensional $(N+1)$-point GLL quadrature rules as a degree $(2N-1)$ quadrature rule.
+This choice of quadrature is sufficient to guarantee that the mass matrix is positive definite \cite{canuto2007spectral}.
+
+% the quantities $\int_{\hat{D}} \pd{u}{x_i} v$ and
+
+%This section presumes that Assumption~\ref{ass:quad} is satisfied for general  $M_{\rm vol}, M_{\rm surf}$.
+In Sections~\ref{sec:singleelem} and \ref{sec:curved}, specific values of $M_{\rm vol}, M_{\rm surf}$ will be motivated by proofs of entropy stability, and Assumption~\ref{ass:quad} will be verified for specific choices of volume and surface quadrature on simplicial and tensor product elements.
+
+%The conditions of Assumption~\ref{ass:quad} imply the following variational SBP property:
+%\begin{lemma}
+%\label{lemma:vsbp}
+%%Assume that $v \in V^M$.  
+%%If the volume quadrature is exact for polynomials of degree $N+M-1$ and the surface quadrature is exact for polynomials of degree $N+M$, 
+%Suppose Assumption~\ref{ass:quad} holds for $v(\bm{x})\in V^N$, and let $\bm{v}_q$ denote the vector of values of $v(\bm{x})$ of at volume quadrature points.
+Let $u(\bm{x})$ be an $L^2$ integrable function, and let $\bm{u}_q$ denote the values of $u(\bm{x})$ at quadrature points.
+Then,
+%\[
+%\bm{v}_q^T\bm{Q}_i \bm{u}_q = \bm{v}_q^T\LRp{ {\bm{E}}^T \bm{B}^i\bm{E} - \bm{Q}_i^T}\bm{u}_q.
+%\]
+%\end{lemma}
+%\begin{proof}
+%Recall that $\bm{Q}_i = \bm{W} \bm{V}_q \bm{D}^i\bm{P}_q$.
+The quadrature-based $L^2$ projection $\Pi_Nu$ is a polynomial of degree $N$, and is computed discretely as $\bm{P}_q\bm{u}_q$.
+By the exactness of the volume quadrature in Assumption~\ref{ass:quad},
+%\begin{align*}
+%\bm{v}_q^T\bm{Q}_i \bm{u}_q &= \bm{v}_q^T\bm{W} \bm{D}^i \bm{P}_q \bm{u}_q = \int_{\hat{D}} \pd{\Pi_N u}{\hat{x}_i} v = \int_{\partial \hat{D}} (\Pi_N u) v \hat{n}_i - \int_{\hat{D}} \LRp{\Pi_N u} \pd{v}{\hat{x}_i}.
+%\end{align*}
+%%Since the volume integrand $\LRp{\Pi_N u} \pd{v}{\hat{x}_i} \in V^{N+M-1}$, it is exactly integrated using the volume quadrature.  
+%%Additionally, since $\Pi_N u \in V^N$ and $v\in V^M$ on the surface of $\hat{D}$, the surface quadrature is exact for the surface integral and 
+%By the exactness of the surface quadrature in Assumption~\ref{ass:quad}, 
+%\begin{align*}
+%\int_{\partial \hat{D}} (\Pi_N u) v \hat{n}_i -   \int_{\hat{D}} \LRp{\Pi_N u} \pd{v}{\hat{x}_i} &=\bm{v}_f^T\bm{B}^i \bm{E}\bm{u}_q - \LRp{\bm{V}_q\bm{D}^i\bm{P}_q\bm{v}_q}^T\bm{W} \bm{V}_q\bm{P}_q\bm{u}_q\\
+%=& \bm{v}_q^T {\bm{E}}^T\bm{B}^i \bm{E}\bm{u}_q - \bm{v}_q^T \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T \bm{W} \bm{V}_q\bm{P}_q\bm{u}_q,
+%\end{align*}
+%where we have used that, since $v \in V^N$, $\bm{v}_f = \bm{V}_f\bm{P}_q\bm{v}_q = \bm{E}\bm{v}_q$.  The proof is completed by noting that 
+%\[
+%\bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T \bm{W} \bm{V}_q\bm{P}_q = \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{M}\bm{P}_q = \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T\bm{W} = \bm{Q}_i^T.
+%\]
+%\qed\end{proof}
+%
+%The variational SBP property in Lemma~\ref{lemma:vsbp} can be used to prove a similar property for the decoupled SBP operator.
+This property, along with the exact differentiation of constants, is necessary for the proof of entropy stability.
+%\begin{lemma}
+%Suppose Assumption~\ref{ass:quad} holds.  Let $\bm{D}^i_N$ be a decoupled SBP operator on the reference element $\hat{D}$.  
+%%Let the volume quadrature be exact for polynomials of degree $N+M-1$, and let the surface quadrature be exact for polynomials of degree $N+M$ for some $M \leq N$.
+Suppose $v\in V^N$ and that $\LRl{v}_{\partial \hat{D}} \in P^M\LRp{\partial \hat{D}}$.
+\note{Fix this to match assumptions in previous lemma.}
+%Then, the following variational SBP property holds:
+%\[
+%\bm{v}^T\bm{Q}^i_N\bm{u} = \bm{v}^T\LRp{\bm{B}^i_N - \LRp{\bm{Q}^i_N}^T}\bm{u}.%, \qquad \bm{Q}^i_N \bm{1} = \bm{0}.
+%\]
+%where $\bm{v}, \bm{u}$ denotes the values of $v$ and $u$ at both volume and surface quadrature points.  
+%\label{lemma:vdsbp}
+%\end{lemma}
+%\begin{proof}
+%For convenience, let $\bm{u}_q, \bm{u}_f$ and $\bm{v}_q, \bm{v}_f$ denote evaluations of $u$ and $v$ at volume and surface points, such that 
+%\[
+%\bm{u} = \begin{pmatrix} \bm{u}_q\\ \bm{u}_f\end{pmatrix}, \qquad \bm{v} = \begin{pmatrix} \bm{v}_q\\ \bm{v}_f\end{pmatrix}.  
+%\]
+%The proof of the variational summation by parts property uses the definition of $\bm{Q}^i_N$ (\ref{eq:QN}), 
+%\begin{align*}
+%\bm{v}^T\bm{Q}^i_N\bm{u} &= \bm{v}_q^T\bm{Q}_i \bm{u}_q - \frac{1}{2}\LRp{\bm{E}\bm{v}_q}^T\bm{B}^i \LRp{\bm{E}\bm{u}_q} + \frac{1}{2}\LRp{\bm{E}\bm{v}_q}^T\bm{B}^i \bm{u}_f\\
+%& - \frac{1}{2}\bm{v}_f^T\bm{B}^i \LRp{\bm{E}\bm{u}_q} + \frac{1}{2}\bm{v}_f^T\bm{B}^i \bm{u}_f.
+%\end{align*}
+%Applying Lemma~\ref{lemma:vsbp} then yields
+%\begin{align*}
+%\bm{v}^T\bm{Q}^i_N\bm{u} =& -\bm{v}_q^T\bm{Q}^T_i \bm{u}_q + \frac{1}{2}\LRp{\bm{E}\bm{v}_q}^T\bm{B}^i \LRp{\bm{E}\bm{u}_q} + \frac{1}{2}\LRp{\bm{E}\bm{v}_q}^T\bm{B}^i \bm{u}_f\\
+%&\qquad - \frac{1}{2}\bm{v}_f^T\bm{B}^i \LRp{\bm{E}\bm{u}_q} - \frac{1}{2}\bm{v}_f^T\bm{B}^i \bm{u}_f + \bm{v}_f^T\bm{B}^i \bm{u}_f\\
+%=& \begin{pmatrix} \bm{v}_q\\ \bm{v}_f\end{pmatrix}^T 
+%\left(\begin{pmatrix}
+%\bm{0}& \\
+%& \bm{B}^i
+%\end{pmatrix}\right.+
+%\left.\begin{pmatrix}
+%-\bm{Q}_i^T + \frac{1}{2}{\bm{E}}^T\bm{B}^i \bm{E} & -\frac{1}{2} \LRp{\bm{B}^i \bm{E}}^T\\
+%\frac{1}{2}\bm{B}^i \bm{E} & -\frac{1}{2}\bm{B}^i
+%\end{pmatrix}  \right)
+%\begin{pmatrix} \bm{u}_q\\ \bm{u}_f\end{pmatrix}\\
+%=& \bm{v}^T\LRp{\bm{B}^i_N - \LRp{\bm{Q}^i_N}^T}\bm{u}.
+%\end{align*}
+%\qed\end{proof}
+
+\subsection{On the accuracy of \bnote{skew-hybridized SBP operators }}
+\label{sec:accskew}
+
+It was shown in \cite{chan2017discretely, chan2018efficient} that the \bnote{hybridized SBP operator $\bm{Q}^i_N$} can be interpreted as augmenting a volume approximation of the derivative with boundary correction terms.
+Let $f(\bm{x}),g(\bm{x})$ denote two $L^2$ integrable functions, and let $\bm{f}_N, \bm{g}_N$ denote the vectors of values of $f,g$ at both volume and surface quadrature points.
+A degree $N$ approximation $u\in V^N$ to $f\pd{g}{x_i}$ can be constructed via
+\begin{equation}
+\bm{M}\bm{u} = \LRs{\begin{array}{c}
+\bm{V}_q \\ \bm{V}_f\end{array}}^T \diag{\bm{f}_N}\bm{Q}^i_N\bm{g}_N, \qquad \bm{f}_N = \begin{bmatrix} \bm{f}_q\\ \bm{f}_f\end{bmatrix}, \quad \bm{g}_N = \begin{bmatrix} \bm{g}_q\\ \bm{g}_f\end{bmatrix},
+\label{eq:dsbpmatrixform}
+\end{equation}
+where $\bm{u}$ denotes the vector of coefficients for $u$.
+
+This algebraic expression (\ref{eq:dsbpmatrixform}) can be reinterpreted as a quadrature approximation of a variational problem, which can be mapped to a physical element $D^k$.
+We seek to approximate $f\pd{g}{x_i}$ by $u \in V^N$ such that, $\forall v\in V^N$
+\begin{equation}
+\int_{D^k} u v = \int_{D^k} f\pd{\Pi_N g}{{x}_i}v + \int_{\partial {D}^k} \LRp{g - \Pi_Ng} \LRp{\frac{fv + \Pi_N(fv)}{2}} {n}^k_i,\label{eq:var}
+\end{equation}
+where $\Pi_N$ is the $L^2$ projection operator (\ref{eq:l2proj}).  
+Integrating half of the volume term by parts yields the skew-symmetric form of (\ref{eq:var})
+\begin{align}
+\int_{D^k} u  v &=\frac{1}{2} \int_{{D^k}} \LRp{f\pd{\Pi_N g}{{x}_i}v - g \pd{\Pi_N\LRp{fv}}{{x}_i}} \label{eq:skewDN}\\
+&+\frac{1}{2} \int_{\partial {D}^k} \LRp{fgv + \LRp{g - \Pi_Ng} \LRp{{fv + \Pi_N(fv)}{}}} {n}^k_i \qquad \forall v\in V^N\nonumber,
+\end{align}
+which yields a matrix formulation \bnote{involving the skew-hybridized SBP operator $\tilde{\bm{Q}}^i_N$}
+\begin{align}
+\bm{M}\bm{u} =& \frac{1}{2}\LRs{\begin{array}{c}
+\bm{V}_q \\ \bm{V}_f\end{array}}^T \diag{\bm{f}_N}\underbrace{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T + \bm{B}^i_N}}_{\tilde{\bm{Q}}^i_N}\bm{g}_N \label{eq:dsbpmatrixform2}.
+%&+ \frac{1}{2}\bm{V}_f^T\bm{W}_f\diag{\bm{n}^k_i} \LRp{\bm{f}_f \circ \bm{g}_f}.\nonumber
+\end{align}
+
+The accuracy of the formulation (\ref{eq:esdgSkew}) can be understood by analyzing the degree of polynomial exactness of (\ref{eq:dsbpmatrixform2}) as an approximation of the derivative.
+Let $u(\bm{x})$ be a polynomial of degree $\leq N$ with coefficients $\bm{u}$, and let $\bm{u}_N = \LRs{\bm{u}_q, \bm{u}_f}^T$ denote the values of $u(\bm{x})$ at volume and surface quadrature points.
+An approximation of $\pd{u}{x_i}$ can be computed by applying (\ref{eq:dsbpmatrixform}) to compute
+\begin{align}
+\pd{u}{x_i}\approx 
+%\frac{1}{2}\bm{M}^{-1}\LRs{\begin{array}{c}
+%\bm{V}_q \\ \bm{V}_f\end{array}}^T \LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T + \bm{B}^i_N}\bm{u}_N = 
+\bm{M}^{-1}\LRs{\begin{array}{c}
+\bm{V}_q \\ \bm{V}_f\end{array}}^T \tilde{\bm{Q}}^i_N\bm{u}_N.
+\label{eq:dsbpapprox}
+\end{align}
+From (\ref{eq:dsbpapprox}), it can be shown that when $\bm{Q}^i$ satisfies a generalized SBP property, the hybridized SBP operator (\ref{eq:dsbp}) produces a degree $N$ approximation to the derivative \cite{chan2017discretely}.
+When $\bm{Q}^i$ does not satisfy a generalized SBP property, we have the following lemma on the accuracy of (\ref{eq:dsbpapprox}):
+\begin{lemma}
+Let $M\leq N$.
+Suppose that the volume quadrature is exact for degree $M+N-1$ polynomials on simplices, or for polynomials in $Q^{M+N-1,M+N,M+N}$ on tensor product elements.
+Furthermore, assume that the surface quadrature is exact for degree $M+N$ polynomials on simplices and $Q^{M+N,M+N}$ on tensor product elements.
+\bnote{Then, so long as the mass matrix is positive definite}, the skew-symmetric approximation of the $x$-derivative (\ref{eq:dsbpapprox}) is exact for polynomials of degree $M$.
+\label{lemma:dsbpapprox}
+\end{lemma}
+\begin{proof}
+\bnote{Suppose $u \in V^M$.
+Let $\bm{u}$ denote the polynomial coefficients of $u$, and let $\bm{e}$ denote the difference between $\bm{D}^i\bm{u}$ (the exact coefficients of $\pd{u}{x_i}$) and the approximation (\ref{eq:dsbpapprox})
+\[
+\bm{e} = \bm{D}^i\bm{u} - \bm{M}^{-1}\LRs{\begin{array}{c}
+\bm{V}_q \\ \bm{V}_f\end{array}}^T\tilde{\bm{Q}}^i_N\bm{u}_N%\frac{1}{2}\bm{M}^{-1}\LRs{\begin{array}{c}
+%\bm{V}_q \\ \bm{V}_f\end{array}}^T\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T + \bm{B}^i_N}\bm{u}_N,
+\]
+where $\bm{e}$ is a polynomial of degree $N$.  
+Since $u(\bm{x})$ is polynomial, the values of $u(\bm{x})$ at quadrature points are $\bm{u}_q = \bm{V}_q\bm{u}$ and $\bm{u} = \bm{P}_q\bm{u}_q$.
+This implies that $\bm{u}_f = \bm{V}_f \bm{P}_q\bm{u}_q =  \bm{E}\bm{u}_q$.
+Expanding the latter term yields
+\begin{align*}
+\bm{M}^{-1}\LRs{\begin{array}{c}
+\bm{V}_q \\ \bm{V}_f\end{array}}^T\tilde{\bm{Q}}^i_N\bm{u}_N &= \frac{1}{2}\bm{M}^{-1}\LRs{\begin{array}{c}
+\bm{V}_q \\ \bm{V}_f\end{array}}^T\begin{bmatrix}
+\bm{Q}^i - \LRp{\bm{Q}^i}^T & \bm{E}^T\bm{B}^i\\
+-\bm{B}^i\bm{E} & \bm{B}^i
+\end{bmatrix}\begin{bmatrix}
+\bm{u}_q\\
+\bm{u}_f
+\end{bmatrix}\\
+&= \frac{1}{2}\bm{M}^{-1}\LRs{\begin{array}{c}
+\bm{V}_q \\ \bm{V}_f\end{array}}^T\begin{bmatrix}
+\bm{Q}^i\bm{u}_q + \LRp{\bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T}\bm{u}_q\\
+\bm{B}^i\LRp{\bm{u}_f - \bm{E}\bm{u}_q}
+\end{bmatrix} 
+\\
+&=\frac{1}{2}\bm{M}^{-1}
+\bm{V}_q^T\LRp{\bm{Q}^i\bm{u}_q + \LRp{\bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T}\bm{u}_q}.
+%&=\frac{1}{2} \bm{D}^i\bm{u} + \frac{1}{2}\bm{M}^{-1}
+%\bm{V}_q^T{ \LRp{\bm{E}^T\bm{B}^i\bm{E} - \bm{Q}^i}\bm{u}_q}
+\end{align*}}
+Since $\bm{Q}^i = \bm{W}\bm{V}_q\bm{D}^i\bm{P}_q$ and $\bm{M} = \bm{V}_q^T\bm{W}\bm{V}_q$, we have that $\bm{M}^{-1}\bm{V}_q^T\bm{Q}^i\bm{u}_q = \bm{D}^i\bm{u}$.
+This simplifies the expression for error to
+\begin{align}
+\bm{e}^T\bm{M}\bm{e} &= \frac{1}{2}
+\bm{e}_q^T\LRp{-\bm{Q}^i\bm{u}_q + \LRp{\bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T}\bm{u}_q} \label{eq:skewerr}\\
+&= -\frac{1}{2}
+\bm{e}_q^T\bm{Q}^i\bm{u}_q + \frac{1}{2}\bm{e}_q^T\LRp{\bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T}\bm{u}_q, \nonumber
+\end{align}
+where we have introduced $\bm{e}_q = \bm{V}_q\bm{e}$.  %The remainder of the proof is similar to the proof of Lemma~\ref{lemma:vsbp}.  
+Since $u \in V^M$ and $e \in V^N$, by exactness of the quadrature rules, 
+\[
+\bm{e}_q^T\LRp{\bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T}\bm{u}_q = \int_{\partial \hat{D}} u e n_i - \int_{\hat{D}} u \pd{e}{x_i}
+= \int_{\hat{D}} \pd{u}{x_i} e = \bm{e}_q^T\bm{Q}^i\bm{u}_q.
+\]
+Combining this with (\ref{eq:skewerr}) implies that $\bm{e}^T\bm{M}\bm{e} = 0$, and since $\bm{M}$ is symmetric positive definite, $\bm{e} = 0$.  
+\qed\end{proof}
+
+%\begin{remark}
+%\end{remark}
+
+Lemma~\ref{lemma:dsbpapprox} suggests that, when a generalized SBP property does not hold, the use of under-integrated quadratures results in a loss of one or more orders of accuracy.
+For example, if the SBP property does not hold, then using $(N+1)$ point GLL rules (which are exact for only polynomials of degree $2N-1$) for either volume or surface quadratures should result in a loss of one order of accuracy compared to the use of $(N+1)$-point Gauss rules (which are exact for polynomials of degree $2N$).
+This is indeed observed in numerical experiments.
 
@@
 \begin{gather}
-\pd{u}{x_i} = \sum_{ij} {G}^k_{ij}\pd{u}{\hat{x}_j}, \qquad {G}^k_{ij} = J^k\pd{\hat{x}_j}{{x}_i}, 
+\pd{u}{x_i} = \sum_{j=1}^d {G}^k_{ij}\pd{u}{\hat{x}_j}, \qquad {G}^k_{ij} = J^k\pd{\hat{x}_j}{{x}_i}, 
 \label{eq:geofacs}
@@
 We also define $\bm{n}^k_i$ as the vector containing concatenated values of the scaled outward normals $n^k_iJ^k_f$ at surface quadrature nodes.
-For the remainder of the work, we assume that the mesh is watertight \cite{chan2018discretely} such that at all points on any internal face, the scaled outward normals $n^k_iJ^k_f$ on the two elements sharing this face are equal and opposite.
+For the remainder of the work, we assume that the mesh is watertight or \note{``well-constructed'' \cite{kopriva2016geometry, chan2018discretely, kopriva2019free}} such that at all points on any internal face, the scaled outward normals $n^k_iJ^k_f$ on the two elements sharing this face are equal and opposite.
 
@@
 %On affine meshes, it is possible to show entropy stability of the skew-symmetric formulation (\ref{eq:esdgSkew}) under a surface quadrature which is only exact for degree $N$ polynomials.  
-On a single element (and on affine meshes), it is possible to show entropy stability of the skew-symmetric formulation (\ref{eq:esdgSkew}) under a surface quadrature which is only exact for degree $N$ polynomials.
+As shown in the previous section, on a single element (and on affine meshes), it is possible to guarantee entropy stability of the skew-symmetric formulation (\ref{eq:esdgSkew}) under a surface quadrature which is only exact for degree $N$ polynomials.
 However, on curved meshes, stronger conditions are required to guarantee entropy stability.
@@
 %Let the domain now be triangulated by a  consisting of curved elements $D^k$.  
-In this section, we describe how to construct appropriate decoupled SBP operators on curved meshes, and give conditions on the volume and surface quadrature rules under which a semi-discretely entropy stable scheme can be constructed.
+In this section, we describe how to construct appropriate hybridized SBP operators on curved meshes, and give conditions on the volume and surface quadrature rules under which a semi-discretely entropy stable scheme can be constructed.
 
@@
 
-We first show how to construct decoupled SBP operators on curved elements.
-Let $\bm{G}^k_{ij}$ denote the vector of scaled geometric terms ${G}^k_{ij}$ evaluated at both volume and surface quadrature points, and let $\hat{\bm{D}}^j_N$ denote the decoupled SBP operator for the $j$th reference coordinate.
-Decoupled SBP operators on a curved element $D^k$ can be defined as in \cite{chan2018discretely} by
+We first show how to construct appropriate SBP operators on curved elements.
+Let $\bm{G}^k_{ij}$ denote the vector of scaled geometric terms ${G}^k_{ij}$ evaluated at both volume and surface quadrature points, and let $\tilde{\bm{Q}}^j_N$ now denote the skew-symmetric construction of the hybridized SBP operator for the $j$th reference coordinate.
+Hybridized SBP operators on a curved element $D^k$ can be defined as in \cite{chan2018discretely} by
 \begin{equation}
-\bm{D}^i_k = \frac{1}{2}\sum_{j=1}^d \LRp{\diag{\bm{G}^k_{ij}}\hat{\bm{D}}^j_N + \hat{\bm{D}}^j_N\diag{\bm{G}^k_{ij}}}, \qquad \bm{Q}^i_k = \bm{W} \bm{D}^i_k.
+\bm{Q}^i_k = \frac{1}{2}\sum_{j=1}^d \LRp{\diag{\bm{G}^k_{ij}}\tilde{\bm{Q}}^j_N + \tilde{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}}.%, \qquad \bm{Q}^i_k = \bm{W} \bm{D}^i_k.
 \label{eq:dncurved}
 \end{equation}
-It can be shown that, if $\hat{\bm{D}}^j_N$ satisfies a strong summation by parts property, $\bm{D}^i_k$ satisfies an analogous strong summation by parts property.
-
-For curved elements, steps must also be taken to ensure that the analogue of Lemma~\ref{lemma:vdsbp} and Lemma~\ref{lemma:sbpcor} hold at the discrete level.
-Expanding out the condition $\bm{Q}^i_k\bm{1} = \bm{0}$ in terms of (\ref{eq:dncurved}) yields
+\bnote{Since $\tilde{\bm{Q}}^j_N$ satisfies a summation by parts property on the reference element $\hat{D}$, then $\bm{Q}^i_k$ satisfies an analogous SBP property on the physical element $D^k$ \cite{chan2018discretely}.
+}
+
+We can now construct and prove entropy conservation and free stream preservation for a skew-symmetric formulation on a physical curved element $D^k$.
+\bnote{Free stream preservation is necessary to discretely preserve both entropy conservation and the condition that constant solutions are stationary solutions of systems of conservation laws.
+However, on curved meshes, the presence of spatially varying geometric terms can result in the production of spurious transient waves.
+The construction of geometric terms through (\ref{eq:iconscurl}) guarantees that the resulting methods are free-stream preserving, and that constant solutions remain stationary solutions of discretizations of (\ref{eq:nonlineqs}).
+}
+
+Let $\bm{Q}^i_k$ be given by (\ref{eq:dncurved}), and define the curved mass matrix 
+\[
+\bm{M}^k = \bm{V}_q^T\bm{W}\diag{\bm{J}^k}\bm{V}_q.
+\]
+Note that $\bm{M}^k$ is positive-definite so long as $J^k$ is positive at all quadrature points.  We define the auxiliary quantities $\tilde{\bm{u}}$ 
+\begin{gather*}
+\bm{v}_q = \bm{v}\LRp{\bm{u}_q}, \qquad \tilde{\bm{v}} = \begin{bmatrix}
+\bm{V}_q\\
+\bm{V}_f
+\end{bmatrix}\bm{P}^k_q\bm{v}_q, \qquad \tilde{\bm{u}} = \bm{u}\LRp{\tilde{\bm{v}}}.
+\end{gather*}
+where $\bm{P}^k_q = \LRp{\bm{M}^k}^{-1}\bm{V}_q^T\bm{W}\diag{\bm{J}^k}$.  Then, we have the following theorem:
+\begin{theorem}
+\label{thm:skewformcurved}
+%Let the geometric terms $G^k_{ij}$ be computed from (\ref{eq:iconscurl}), let the scaled outward normals $\bm{n}^k_i$ be computed pointwise from (\ref{eq:normals}), and suppose Assumption~\ref{ass:quad} holds for $v = 1$ and $v= G^k_{ij}$.
+\bnote{Assume that $\bm{Q}^i_k\bm{1}=\bm{0}$.}
+Let $\tilde{\bm{u}}_f^+$ denote the face value of the entropy-projected conservative variables $\tilde{\bm{u}}_f$ on the neighboring element.  Then, the formulation
+\begin{gather}
+%\bm{M}^k\pd{\bm{u}_h}{t} + 
+%\sum_{i=1}^d \LRs{\begin{array}{cc}
+%\bm{V}_q \\
+%\bm{V}_f\end{array}}^T \LRp{\LRp{\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T} \circ \bm{F}^i_S}\bm{1} + \bm{V}_f^T \bm{W}_f \diag{{\bm{n}^k_i}}\bm{f}_i^* = 0, 
+\bm{M}^k\pd{\bm{u}_h}{t} + 
+\sum_{i=1}^d \LRs{\begin{array}{cc}
+\bm{V}_q \\
+\bm{V}_f\end{array}}^T 2\LRp{\bm{Q}^i_k \circ \bm{F}^i_S}\bm{1} + \bm{V}_f^T \bm{W}_f \diag{\bm{n}^k_i}\LRp{\bm{f}_i^*-\bm{f}(\tilde{\bm{u}}_f)} = 0,
+\label{eq:skewformcurved}\\
+\LRp{\bm{F}^i_S}_{ij} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q, \nonumber\\
+\bm{f}_i^* = \bm{f}^i_S(\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f), \qquad \text{ on interior interfaces,} \nonumber
+\end{gather}
+is semi-discretely entropy conservative on $D^k$ such that for $\bm{u}_q = \bm{V}_q\bm{u}$,
+\begin{gather*}
+\bm{1}^T\bm{W}\diag{\bm{J}^k}\pd{U(\bm{u}_q)}{t} + \sum_{i=1}^d\bm{1}^T\bm{W}_f \diag{\bm{n}^k_i} \LRp{\psi_i(\tilde{\bm{u}}_f) - \tilde{\bm{v}}_f^T\bm{f}_i^*} = 0.
+\end{gather*}
+Additionally, the method is free-stream preserving such that $\pd{\bm{u}_h}{t} = 0$ for constant solutions.
+\label{thm:esdgCurved}
+\end{theorem}
+\bnote{We omit the proof of entropy conservation, since it is identical to the proofs in \cite{chan2017discretely, chan2018discretely}.
+Free-stream preservation follows directly from $\bm{Q}^i_k\bm{1} = \bm{0}$ and the fact that $\bm{F}_S$ is constant for constant solutions \cite{kopriva2006metric}.
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\bnote{The proof of Theorem~\ref{thm:esdgCurved} requires $\bm{Q}^i_k\bm{1} = \bm{0}$.
+For curved elements, additional steps must also be taken to ensure this condition.
+Assuming Assumption~\ref{ass:quad} holds for $v(\bm{x}) = 1$ and expanding out the expression for $\bm{Q}^i_k\bm{1} = \bm{0}$ using (\ref{eq:dncurved}) yields}
 \begin{align}
-\bm{Q}^i_k \bm{1} = \frac{1}{2} \bm{W}\sum_{j=1}^d \diag{\bm{G}^k_{ij}}\hat{\bm{D}}^j_N \bm{1} + \hat{\bm{D}}^j_N\diag{\bm{G}^k_{ij}}\bm{1} = \frac{1}{2}\bm{W}\sum_{j=1}^d \hat{\bm{D}}^j_N\LRp{\bm{G}^k_{ij}} = 0,
+\bm{Q}^i_k \bm{1} = \frac{1}{2} \sum_{j=1}^d \diag{\bm{G}^k_{ij}}\tilde{\bm{Q}}^j_N \bm{1} + \tilde{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}\bm{1} = \frac{1}{2}\sum_{j=1}^d \tilde{\bm{Q}}^j_N\LRp{\bm{G}^k_{ij}} = 0,
 \label{eq:dgcl}
 \end{align}
-where we have used that $\hat{\bm{D}}^j_N \bm{1} = 0$.
-We refer to the condition $\bm{Q}^i_k\bm{1} = 0$ as the discrete geometric conservation law (GCL) \cite{thomas1979geometric, kopriva2006metric}.
+\bnote{where we have used that $\tilde{\bm{Q}}^j_N \bm{1} = 0$ using Lemma~\ref{lemma:sbpcor}.}  We refer to the condition $\bm{Q}^i_k\bm{1} = 0$ as the discrete geometric conservation law (GCL) \cite{thomas1979geometric, kopriva2006metric}.
 For degree $N$ isoparametric mappings, the GCL is automatically satisfied in two dimensions due to the fact that the exact geometric terms ${G}^k_{ij}$ are polynomials of degree $N$ \cite{kopriva2006metric}.
@@
 
-To ensure that the geometric terms satisfy the GCL, we first rewrite the geometric terms as the curl of some quantity $\bm{r}^i$, but interpolate $\bm{r}^i$ before applying the curl \cite{visbal2002use, kopriva2006metric, hindenlang2012explicit, chan2018discretely}:
+To ensure that the geometric terms satisfy the GCL, we first rewrite the geometric terms as the curl of some quantity $\bm{r}^i$, but interpolate $\bm{r}^i$ before applying the curl \cite{thomas1979geometric, visbal2002use, kopriva2006metric, hindenlang2012explicit, chan2018discretely}:
 \begin{gather}
@@
 \LRp{\hat{\Grad}\times I_{N_{\rm geo}}\LRp{x_1\hat{\Grad}x_2}}_j
-\end{bmatrix},\\
+\end{bmatrix},\label{eq:iconscurl} \\
 %-\frac{1}{2}\LRp{\pd{I_{N_{\rm geo}}\bm{r}^k}{\hat{x}_j}-\pd{I_{N_{\rm geo}}\bm{r}^j}{\hat{x}_k}},\\
@@
 \end{cases},\nonumber
-\label{eq:iconscurl}
 \end{gather}
@@
 
-Because the decoupled SBP operators $\bm{D}^i_k$ are now defined through (\ref{eq:dncurved}), Lemma~\ref{lemma:sbpcor} and the proof of entropy stability no longer hold for curved elements and must be modified.
+Because the \bnote{skew-hybridized SBP operators $\bm{Q}^i_k$} are now defined through (\ref{eq:dncurved}), Lemma~\ref{lemma:sbpcor} and the proof of entropy stability no longer hold for curved elements and must be modified.
 The introduction of curvilinear meshes will impose slightly different conditions on the accuracy of the surface quadrature.
@@
 
-\begin{lemma}{(Discrete FTC and exact constant differentiation on curved elements)}
+\begin{lemma}
 \label{lemma:vdsbpcurved} 
-Let $D^k$ be a curved element, and let the geometric terms $G^k_{ij}$ be constructed using (\ref{eq:iconscurl}).
-Let Assumption~\ref{ass:quad} hold for $v = 1$ and $v = G^k_{ij}$ for all $i,j = 1,\ldots, d$, and let $\bm{u}$ denote a vector of values of some function at volume and surface quadrature points.
+Let $D^k$ be a curved element, and let the geometric terms $G^k_{ij}$ be constructed using \rnote{(\ref{eq:iconscurl})}.
+Let Assumption~\ref{ass:quad} hold for $v = 1$ and $v = G^k_{ij}$ for all $i,j = 1,\ldots, d$.
+Then,%, and let $\bm{u}$ denote a vector of values of some function at volume and surface quadrature points.
 Then,
@@
 \[
-\bm{1}^T\bm{Q}^i_k\bm{u} = \bm{1}^T\bm{B}^i_k\bm{u}, \qquad \bm{Q}^i_k\bm{1} = \bm{0},
-\]
-where the physical boundary matrix $\bm{B}^i_k$ is defined as
-\[
-\bm{B}^i_k = \sum_{j=1}^d \bm{B}^j_N \diag{\bm{G}^k_{ij}} = \sum_{j=1}^d \diag{\bm{G}^k_{ij}}\bm{B}^j_N  = \begin{pmatrix}
-\bm{0}&\\
-& \bm{W}_f \diag{\bm{n}^k_i}\end{pmatrix}.
-\]
+\qquad \bm{Q}^i_k\bm{1} = \bm{0},
+\]
+%where the physical boundary matrix $\bm{B}^i_k$ is defined as
+%\[
+%\bm{B}^i_k = \sum_{j=1}^d \bm{B}^j_N \diag{\bm{G}^k_{ij}} = \sum_{j=1}^d \diag{\bm{G}^k_{ij}}\bm{B}^j_N  = \begin{pmatrix}
+%\bm{0}&\\
+%& \bm{W}_f \diag{\bm{n}^k_i}\end{pmatrix}.
+%\]
 \end{lemma}
 \begin{proof}
-The proof of $\bm{Q}^i_k\bm{1} = \bm{0}$ was shown for for tensor product elements in \cite{kopriva2006metric} and for simplicial elements in \cite{chan2018discretely}, and relies only on the fact that $G^k_{ij} \in V^N$.
-
-The discrete fundamental theorem of calculus $\bm{1}^T\bm{Q}^i_k\bm{u} = \bm{1}^T\bm{B}^i_k\bm{u}$ can be shown to hold using (\ref{eq:dncurved}):
-\begin{align*}
-\bm{1}^T\bm{Q}^i_k\bm{u} &= \frac{1}{2}\sum_{j=1}^d \bm{1}^T{\LRp{\diag{\bm{G}^k_{ij}}\hat{\bm{Q}}^j_N + \hat{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}}} \bm{u} \\
-&= \frac{1}{2}\sum_{j=1}^d \LRp{\LRp{\bm{G}^k_{ij}}^T \hat{\bm{Q}}^j_N\bm{u} + \bm{1}^T\hat{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}\bm{u}}.
-\end{align*}
-Applying Lemma~\ref{lemma:vdsbp} to $\sum_{j=1}^d\bm{1}^T\hat{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}\bm{u}$ yields
-\begin{align*}
-\sum_{j=1}^d\bm{1}^T\hat{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}\bm{u} &= \sum_{j=1}^d\bm{1}^T\LRp{\bm{B}^j_N - \LRp{\hat{\bm{Q}}^j_N}^T}\diag{\bm{G}^k_{ij}}\bm{u} \\
-&= \sum_{j=1}^d\bm{1}^T\bm{B}^j_N \diag{\bm{G}^k_{ij}}\bm{u},
-\end{align*}
-where we have used Lemma~\ref{lemma:sbpcor} to conclude that $\hat{\bm{Q}}^j_N \bm{1} = 0$.
-Repeating the procedure for $\sum_{j=1}^d \LRp{\bm{G}^k_{ij}}^T \hat{\bm{Q}}^j_N\bm{u}$ yields
-\begin{align*}
-\sum_{j=1}^d \LRp{\bm{G}^k_{ij}}^T \hat{\bm{Q}}^j_N\bm{u} &= \sum_{j=1}^d \LRp{\bm{G}^k_{ij}}^T \LRp{\bm{B}^j_N - \LRp{\hat{\bm{Q}}^j_N}^T}\bm{u} \\
-&= \sum_{j=1}^d \LRp{\bm{G}^k_{ij}}^T \bm{B}^j_N \bm{u} - \sum_{j=1}^d \bm{u}^T{\hat{\bm{Q}}^j_N}{\bm{G}^k_{ij}} = \sum_{j=1}^d \LRp{\bm{G}^k_{ij}}^T \bm{B}^j_N \bm{u},  
-\end{align*}
-where we have used that the $G^k_{ij}$ satisfies the discrete GCL.  The proof is completed by noting that $\bm{B}^j_N$ is diagonal and thus commutes with $\diag{\bm{G}^k_{ij}}$, such that
-\[
-\bm{1}^T\bm{Q}^i_k\bm{u} = \sum_{j=1}^d\bm{1}^T\LRp{\bm{B}^j_N \diag{\bm{G}^k_{ij}}}\bm{u} = \bm{1}^T\bm{B}^i_k \bm{u}.
-\]
+The proof of $\bm{Q}^i_k\bm{1} = \bm{0}$ is analogous to the proof of Lemma~\ref{lemma:sbpcor}.
+The results follow for tensor product elements using results from \cite{kopriva2006metric} and for simplicial elements using results from \cite{chan2018discretely}.
+In both cases, the proof relies only on the fact that $G^k_{ij} \in V^N$.
+%where the physical boundary matrix $\bm{B}^i_k$ is defined as
+%\[
+%\bm{B}^i_k = \sum_{j=1}^d \bm{B}^j_N \diag{\bm{G}^k_{ij}} = \sum_{j=1}^d \diag{\bm{G}^k_{ij}}\bm{B}^j_N  = \begin{pmatrix}
+%\bm{0}&\\
+%& \bm{W}_f \diag{\bm{n}^k_i}\end{pmatrix}.
+%\]
+
+%The discrete fundamental theorem of calculus $\bm{1}^T\bm{Q}^i_k\bm{u} = \bm{1}^T\bm{B}^i_k\bm{u}$ can be shown to hold using (\ref{eq:dncurved}):
+%\begin{align*}
+%\bm{1}^T\bm{Q}^i_k\bm{u} &= \frac{1}{2}\sum_{j=1}^d \bm{1}^T{\LRp{\diag{\bm{G}^k_{ij}}\hat{\bm{Q}}^j_N + \hat{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}}} \bm{u} \\
+%&= \frac{1}{2}\sum_{j=1}^d \LRp{\LRp{\bm{G}^k_{ij}}^T \hat{\bm{Q}}^j_N\bm{u} + \bm{1}^T\hat{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}\bm{u}}.
+%\end{align*}
+%Applying Lemma~\ref{lemma:vdsbp} to $\sum_{j=1}^d\bm{1}^T\hat{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}\bm{u}$ yields
+%\begin{align*}
+%\sum_{j=1}^d\bm{1}^T\hat{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}\bm{u} &= \sum_{j=1}^d\bm{1}^T\LRp{\bm{B}^j_N - \LRp{\hat{\bm{Q}}^j_N}^T}\diag{\bm{G}^k_{ij}}\bm{u} \\
+%&= \sum_{j=1}^d\bm{1}^T\bm{B}^j_N \diag{\bm{G}^k_{ij}}\bm{u},
+%\end{align*}
+%where we have used Lemma~\ref{lemma:sbpcor} to conclude that $\hat{\bm{Q}}^j_N \bm{1} = 0$.
+%Repeating the procedure for $\sum_{j=1}^d \LRp{\bm{G}^k_{ij}}^T \hat{\bm{Q}}^j_N\bm{u}$ yields
+%\begin{align*}
+%\sum_{j=1}^d \LRp{\bm{G}^k_{ij}}^T \hat{\bm{Q}}^j_N\bm{u} &= \sum_{j=1}^d \LRp{\bm{G}^k_{ij}}^T \LRp{\bm{B}^j_N - \LRp{\hat{\bm{Q}}^j_N}^T}\bm{u} \\
+%&= \sum_{j=1}^d \LRp{\bm{G}^k_{ij}}^T \bm{B}^j_N \bm{u} - \sum_{j=1}^d \bm{u}^T{\hat{\bm{Q}}^j_N}{\bm{G}^k_{ij}} = \sum_{j=1}^d \LRp{\bm{G}^k_{ij}}^T \bm{B}^j_N \bm{u},  
+%\end{align*}
+%where we have used that the $G^k_{ij}$ satisfies the discrete GCL.  The proof is completed by noting that $\bm{B}^j_N$ is diagonal and thus commutes with $\diag{\bm{G}^k_{ij}}$, such that
+%\[
+%\bm{1}^T\bm{Q}^i_k\bm{u} = \sum_{j=1}^d\bm{1}^T\LRp{\bm{B}^j_N \diag{\bm{G}^k_{ij}}}\bm{u} = \bm{1}^T\bm{B}^i_k \bm{u}.
+%\]
 \qed\end{proof}
 
-The property $\bm{Q}^i_k\bm{1} = 0$ is also related to free-stream preservation \cite{kopriva2006metric}.
-Constant solutions are stationary solutions of systems of conservation laws.
-However, on curved meshes, the presence of spatially varying geometric terms can result in the production of spurious transient waves.
-The construction of geometric terms through (\ref{eq:iconscurl}) guarantees that the resulting methods are free-stream preserving, and that constant solutions remain stationary solutions of discretizations of (\ref{eq:nonlineqs}).
-
-We can now construct and prove entropy conservation and free stream preservation for a skew-symmetric formulation on a curved element $D^k$.
-Let $\bm{Q}^i_k$ be given by (\ref{eq:dncurved}), and define the curved mass matrix $\bm{M}^k = \bm{V}_q^T\bm{W}\diag{\bm{J}^k}\bm{V}_q$.
-Note that $\bm{M}^k$ is positive-definite so long as $J^k$ is positive at all quadrature points.
-We define the auxiliary quantities $\tilde{\bm{u}}$
-\begin{gather*}
-\bm{v}_q = \bm{v}\LRp{\bm{u}_q}, \qquad \tilde{\bm{v}} = \begin{bmatrix}
-\bm{V}_q\\
-\bm{V}_f
-\end{bmatrix}\bm{P}^k_q\bm{v}_q, \qquad \tilde{\bm{u}} = \bm{u}\LRp{\tilde{\bm{v}}}.
-\end{gather*}
-where $\bm{P}^k_q = \LRp{\bm{M}^k}^{-1}\bm{V}_q^T\bm{W}\diag{\bm{J}^k}$.  Then, we have the following theorem:
-\begin{theorem}
-\label{thm:skewformcurved}
-Let the geometric terms $G^k_{ij}$ be computed from (\ref{eq:iconscurl}), let the scaled outward normals $\bm{n}^k_i$ be computed pointwise from (\ref{eq:normals}), and suppose Assumption~\ref{ass:quad} holds for $v = 1$ and $v= G^k_{ij}$.
-Let $\tilde{\bm{u}}_f^+$ denote the face value of the entropy-projected conservative variables $\tilde{\bm{u}}_f$ on the neighboring element.
-Then, the formulation
-\begin{gather}
-\bm{M}^k\pd{\bm{u}_h}{t} + 
-\sum_{i=1}^d \LRs{\begin{array}{cc}
-\bm{V}_q \\
-\bm{V}_f\end{array}}^T \LRp{\LRp{\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T} \circ \bm{F}^i_S}\bm{1} + \bm{V}_f^T \bm{W}_f \diag{{\bm{n}^k_i}}\bm{f}_i^* = 0, \label{eq:skewformcurved}\\
-%\sum_{i=1}^d \LRs{\begin{array}{cc}
-%\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{{\bm{n}_i}}\bm{f}_i^* = 0, \label{eq:skewform}\\
-\LRp{\bm{F}^i_S}_{ij} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q, \nonumber\\
-\bm{f}_i^* = \bm{f}^i_S(\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f), \qquad \text{ on interior interfaces,} \nonumber
-\end{gather}
-is semi-discretely entropy conservative on $D^k$ such that for $\bm{u}_q = \bm{V}_q\bm{u}$,
-\begin{gather*}
-\bm{1}^T\bm{W}\diag{\bm{J}^k}\pd{U(\bm{u}_q)}{t} + \sum_{i=1}^d\bm{1}^T\bm{W}_f \diag{\bm{n}^k_i} \LRp{\psi_i(\tilde{\bm{u}}_f) - \tilde{\bm{v}}_f^T\bm{f}_i^*} = 0.
-\end{gather*}
-Additionally, the method is free-stream preserving such that $\pd{\bm{u}_h}{t} = 0$ for constant solutions.
-\end{theorem}
-\begin{proof}
-The proof of entropy stability is identical to that of Theorem~\ref{thm:esdg}, except that Lemma~\ref{lemma:vdsbpcurved} is used instead of Lemma~\ref{lemma:sbpcor}.
+%\begin{proof}
+%The proof of entropy stability is identical to that of Theorem~\ref{thm:esdg}, except that Lemma~\ref{lemma:vdsbpcurved} is used instead of Lemma~\ref{lemma:sbpcor}.
 However, proving free-stream preservation differs from approaches in the literature, which typically show that the right hand side is zero for $\bm{u}_h$ constant \cite{crean2018entropy, chan2018discretely}.
 This strategy for is appropriate for ``strong'' formulations, but must be modified for the skew-symmetric formulation (\ref{eq:skewformcurved}).
-
-We have that $\bm{f}_i^* = \bm{f}^i_S(\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f) = \bm{f}_i(\bm{u}_f)$ using the consistency property of the Tadmor flux (\ref{eq:esflux}) and the fact that the solution is globally constant.
+%
+%We have that $\bm{f}_i^* = \bm{f}^i_S(\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f) = \bm{f}_i(\bm{u}_f)$ using the consistency property of the Tadmor flux (\ref{eq:esflux}) and the fact that the solution is globally constant.
 For $\bm{u}_h$ constant, $\bm{F}_S$ is also constant.
@@
 Then,
-\[
-\LRp{\LRp{\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T} \circ \bm{F}_S}\bm{1} = \LRp{\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T}\bm{1} = \LRp{\bm{Q}^i_k}^T\bm{1},
-\]
-where we have used that $\bm{Q}^i_k\bm{1} = 0$ by Lemma~\ref{lemma:vdsbpcurved}.  Thus, the right hand side reduces to 
-\begin{align*}
-&\sum_{j=1}^d \LRs{\begin{array}{cc}
-\bm{V}_q \\
-\bm{V}_f\end{array}}^T \LRp{-\LRp{\bm{Q}^i_k}^T + \bm{V}_f^T \bm{W}_f \diag{{\bm{n}^k_i}}}\bm{1}\\
-&= \sum_{j=1}^d \LRs{\begin{array}{cc}
-\bm{V}_q \\
-\bm{V}_f\end{array}}^T \LRp{\LRp{ -\frac{1}{2}\LRp{\diag{\bm{G}^k_{ij}}\hat{\bm{Q}}^j_N + \hat{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}}^T + \hat{\bm{B}}^j_N \diag{\bm{G}^k_{ij}}}}\bm{1},
+%\[
+%\LRp{\LRp{\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T} \circ \bm{F}_S}\bm{1} = \LRp{\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T}\bm{1} = \LRp{\bm{Q}^i_k}^T\bm{1},
+%\]
+%where we have used that $\bm{Q}^i_k\bm{1} = 0$ by Lemma~\ref{lemma:vdsbpcurved}.  Thus, the right hand side reduces to 
+%\begin{align*}
+%&\sum_{j=1}^d \LRs{\begin{array}{cc}
+%\bm{V}_q \\
+%\bm{V}_f\end{array}}^T \LRp{-\LRp{\bm{Q}^i_k}^T + \bm{V}_f^T \bm{W}_f \diag{{\bm{n}^k_i}}}\bm{1}\\
 %&= \sum_{j=1}^d \LRs{\begin{array}{cc}
 %\bm{V}_q \\
-%\bm{V}_f\end{array}}^T  -\frac{1}{2}\LRp{\hat{\bm{Q}}^j_N}^T{\bm{G}^k_{ij}} + 
-\end{align*}
-where we have used the definition of $\bm{Q}^i_k$ in (\ref{eq:dncurved}).  Rearranging terms gives that, for a constant solution, (\ref{eq:skewformcurved}) reduces to
-\begin{align*}
-\bm{M}^k\pd{\bm{u}_h}{t} &+ \frac{1}{2} \sum_{j=1}^d \LRs{\begin{array}{cc}
-\bm{V}_q \\
-\bm{V}_f\end{array}}^T \LRp{-\LRp{\hat{\bm{Q}}^j_N}^T +\hat{\bm{B}}^j_N}\bm{G}^k_{ij}\\
-&+ \frac{1}{2} \sum_{j=1}^d \LRs{\begin{array}{cc}
-\bm{V}_q \\
-\bm{V}_f\end{array}}^T\diag{\bm{G}^k_{ij}} \LRp{-\LRp{\hat{\bm{Q}}^j_N}^T +\hat{\bm{B}}^j_N}\bm{1} = 0.
-\end{align*}
-Let $u$ denote the constant value of the solution such that the values of the solution at quadrature points is $u\bm{1}$.
+%\bm{V}_f\end{array}}^T \LRp{\LRp{ -\frac{1}{2}\LRp{\diag{\bm{G}^k_{ij}}\hat{\bm{Q}}^j_N + \hat{\bm{Q}}^j_N\diag{\bm{G}^k_{ij}}}^T + \hat{\bm{B}}^j_N \diag{\bm{G}^k_{ij}}}}\bm{1},
+%%&= \sum_{j=1}^d \LRs{\begin{array}{cc}
+%%\bm{V}_q \\
+%%\bm{V}_f\end{array}}^T  -\frac{1}{2}\LRp{\hat{\bm{Q}}^j_N}^T{\bm{G}^k_{ij}} + 
+%\end{align*}
+%where we have used the definition of $\bm{Q}^i_k$ in (\ref{eq:dncurved}).  Rearranging terms gives that, for a constant solution, (\ref{eq:skewformcurved}) reduces to
+%\begin{align*}
+%\bm{M}^k\pd{\bm{u}_h}{t} &+ \frac{1}{2} \sum_{j=1}^d \LRs{\begin{array}{cc}
+%\bm{V}_q \\
+%\bm{V}_f\end{array}}^T \LRp{-\LRp{\hat{\bm{Q}}^j_N}^T +\hat{\bm{B}}^j_N}\bm{G}^k_{ij}\\
+%&+ \frac{1}{2} \sum_{j=1}^d \LRs{\begin{array}{cc}
+%\bm{V}_q \\
+%\bm{V}_f\end{array}}^T\diag{\bm{G}^k_{ij}} \LRp{-\LRp{\hat{\bm{Q}}^j_N}^T +\hat{\bm{B}}^j_N}\bm{1} = 0.
+%\end{align*}
+%Let $u$ denote the constant value of the solution such that the values of the solution at quadrature points is $u\bm{1}$.
 Multiplying both sides by the constant solution $\bm{u}_h^T$ yields that $\frac{1}{2}\pd{}{t}\LRp{\bm{u}_h^T\bm{M}^k\bm{u}_h}$ is equal to
-\begin{align*}
--\frac{u}{2} \sum_{j=1}^d \LRp{\bm{1}^T \LRp{-\LRp{\hat{\bm{Q}}^j_N}^T +\hat{\bm{B}}^j_N}\bm{G}^k_{ij}
-+\LRp{\bm{G}^k_{ij}}^T \LRp{-\LRp{\hat{\bm{Q}}^j_N}^T +\hat{\bm{B}}^j_N}\bm{1}}.% = 0.
-\end{align*}
-By assumption, Lemma~\ref{lemma:vdsbp} holds for test functions $\bm{1}$ and $\bm{G}^k_{ij}$, and the formulation reduces to
-\begin{align*}
-\frac{1}{2}\pd{}{t}\LRp{\bm{u}_h^T\bm{M}^k\bm{u}_h} + \frac{u}{2} \sum_{j=1}^d \LRp{\bm{1}^T \hat{\bm{Q}}^j_N\bm{G}^k_{ij}
-+\LRp{\bm{G}^k_{ij}}^T \hat{\bm{Q}}^j_N\bm{1}} = 0.
-\end{align*}
-By construction, $\hat{\bm{Q}}^j_N\bm{1} = 0$, and $\sum_{j=1}^d \bm{1}^T \hat{\bm{Q}}^j_N\bm{G}^k_{ij} = 0$ since $G^k_{ij}$ satisfies the GCL.
+%\begin{align*}
+%-\frac{u}{2} \sum_{j=1}^d \LRp{\bm{1}^T \LRp{-\LRp{\hat{\bm{Q}}^j_N}^T +\hat{\bm{B}}^j_N}\bm{G}^k_{ij}
+%+\LRp{\bm{G}^k_{ij}}^T \LRp{-\LRp{\hat{\bm{Q}}^j_N}^T +\hat{\bm{B}}^j_N}\bm{1}}.% = 0.
+%\end{align*}
+%By assumption, Lemma~\ref{lemma:vdsbp} holds for test functions $\bm{1}$ and $\bm{G}^k_{ij}$, and the formulation reduces to
+%\begin{align*}
+%\frac{1}{2}\pd{}{t}\LRp{\bm{u}_h^T\bm{M}^k\bm{u}_h} + \frac{u}{2} \sum_{j=1}^d \LRp{\bm{1}^T \hat{\bm{Q}}^j_N\bm{G}^k_{ij}
+%+\LRp{\bm{G}^k_{ij}}^T \hat{\bm{Q}}^j_N\bm{1}} = 0.
+%\end{align*}
+%By construction, $\hat{\bm{Q}}^j_N\bm{1} = 0$, and $\sum_{j=1}^d \bm{1}^T \hat{\bm{Q}}^j_N\bm{G}^k_{ij} = 0$ since $G^k_{ij}$ satisfies the GCL.
 This implies that the latter term vanishes and $\pd{}{t}\LRp{\bm{u}_h^T\bm{M}^k\bm{u}_h} = 0$ on $D^k$.
@@
 Summing $\pd{}{t}\LRp{\bm{u}_h^T\bm{M}^k\bm{u}_h} = 0$ over all elements ensures that globally constant solutions are stationary and that $\pd{\bm{u}_h}{t} = 0$.
-\qed\end{proof}
+%\qed\end{proof}
 
@@
 
-\begin{remark}
-The skew-symmetric operator ${\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T}$ in (\ref{eq:skewformcurved}) can be rewritten as
-\begin{align*}
-%\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T =&
-\sum_{j=1}^d \diag{\bm{G}^k_{ij}}{ \frac{1}{2}\LRp{\hat{\bm{Q}}_N^j-\LRp{\hat{\bm{Q}}_N^j}^T}}+  \frac{1}{2} \LRp{\hat{\bm{Q}}_N^j-\LRp{\hat{\bm{Q}}_N^j}^T}\diag{\bm{G}^k_{ij}}.
-\end{align*}
-Thus, the volume terms in the skew-symmetric formulation may be applied using only the skew-symmetric part of the reference SBP operator $\hat{\bm{Q}}^j_N$.
-\end{remark}
+%\begin{remark}
+%The skew-symmetric operator ${\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T}$ in (\ref{eq:skewformcurved}) can be rewritten as
+%\begin{align*}
+%%\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T =&
+%\sum_{j=1}^d \diag{\bm{G}^k_{ij}}{ \frac{1}{2}\LRp{\tilde{\bm{Q}}_N^j-\LRp{\tilde{\bm{Q}}_N^j}^T}}+  \frac{1}{2} \LRp{\tilde{\bm{Q}}_N^j-\LRp{\tilde{\bm{Q}}_N^j}^T}\diag{\bm{G}^k_{ij}}.
+%\end{align*}
+%Thus, the volume terms in the skew-symmetric formulation may be applied using only the skew-symmetric part of the reference SBP operator $\tilde{\bm{Q}}^j_N$.
+%\end{remark}
 
@@
 
-\section{Conditions on the accuracy of volume and surface quadrature rules}
+\subsection{On quadrature conditions for Assumption~\ref{ass:quad} for $v = 1$ and $v = G^k_{ij}$}
 \label{sec:quadacc}
@@
 The previous sections outline minimal conditions under which entropy stability is guaranteed under a skew-symmetric formulation and a polynomial geometric mapping.
-In Section~\ref{sec:quadacc}, we translate these minimial conditions into conditions on quadrature accuracy.
-
-Additionally, while it is possible to maintain entropy stability under reduced accuracy quadratures, the accuracy of the resulting schemes depend on the accuracy of the volume and surface quadrature rules.
+In this section, we translate these minimial conditions into conditions on quadrature accuracy.
+%Additionally, while it is possible to maintain entropy stability under reduced accuracy quadratures, the accuracy of the resulting schemes depend on the accuracy of the volume and surface quadrature rules.
 We analyze the degree of accuracy expected based on the accuracy of different quadrature rules in Section~\ref{sec:accskew}.
 
-\subsection{On quadrature conditions for Assumption~\ref{ass:quad} for $v = 1$ and $v = G^k_{ij}$}
-
 Semi-discrete entropy conservation on curved meshes requires that Assumption~\ref{ass:quad} holds for $v = 1$ and $v = G^k_{ij}$.
-We discuss specific choices of volume and surface quadrature on for which this assumption is valid, and summarize the maximum degree $N_{\rm geo}$ of the polynomial geometric approximation under which entropy stability holds for common choices of volume and surface quadrature.
+We discuss specific choices of volume and surface quadrature for which this assumption is valid, and summarize the maximum degree $N_{\rm geo}$ of the polynomial geometric approximation under which entropy stability holds for common choices of volume and surface quadrature.
 %\note{Emphasize we're interested in taking $N_{\rm geo}$ as large as possible for a given quadrature rule, and that there is a focus on degree $(2N-1)$ quadrature rules.}
@@
 The situation is more complicated for curved tensor product elements.
-It was shown in Section~\ref{sec:assump1} that tensor product quadratures of degree $(2N-1)$ satisfy Assumption~\ref{ass:quad} holds for $v = 1$.
+It was shown in Section~\ref{sec:assump1} that tensor product quadratures of degree $(2N-1)$ satisfy Assumption~\ref{ass:quad} for $v = 1$.
 However, in contrast to the simplicial case, it is not immediately clear that degree $(2N-1)$ volume quadratures exactly integrate $\int_{\hat{D}} \pd{u}{\hat{x}_j}v$ for $v = G^k_{ij}$ for tensor product elements.
@@
 To the author's knowledge, this setting has not been considered within the literature.
-
-\subsection{On the accuracy of the skew-symmetric formulation}
-\label{sec:accskew}
-
-It was shown in \cite{chan2017discretely, chan2018efficient} that the decoupled SBP operator can be interpreted as augmenting a volume approximation of the derivative with boundary correction terms.
-Let $f(\bm{x}),g(\bm{x})$ denote two $L^2$ integrable functions, and let $\bm{f}_N, \bm{g}_N$ denote the vectors of values of $f,g$ at both volume and surface quadrature points.
-A degree $N$ approximation $u\in V^N$ to $f\pd{g}{x_i}$ can be constructed via
-\begin{equation}
-\bm{M}\bm{u} = \LRs{\begin{array}{c}
-\bm{V}_q \\ \bm{V}_f\end{array}}^T \diag{\bm{f}_N}\bm{Q}^i_N\bm{g}_N, \qquad \bm{f}_N = \begin{bmatrix} \bm{f}_q\\ \bm{f}_f\end{bmatrix}, \quad \bm{g}_N = \begin{bmatrix} \bm{g}_q\\ \bm{g}_f\end{bmatrix},
-\label{eq:dsbpmatrixform}
-\end{equation}
-where $\bm{u}$ denotes the vector of coefficients for $u$.
-
-This algebraic expression (\ref{eq:dsbpmatrixform}) can be reinterpreted as a quadrature approximation of a variational problem, which can be mapped to a physical element $D^k$.
-We seek to approximate $f\pd{g}{x_i}$ by $u \in V^N$ such that, $\forall v\in V^N$
-\begin{equation}
-\int_{D^k} u v = \int_{D^k} f\pd{\Pi_N g}{{x}_i}v + \int_{\partial {D}^k} \LRp{g - \Pi_Ng} \LRp{\frac{fv + \Pi_N(fv)}{2}} {n}^k_i,\label{eq:var}
-\end{equation}
-where $\Pi_N$ is the $L^2$ projection operator (\ref{eq:l2proj}).  
-Integrating half of the volume term by parts yields the skew-symmetric form of (\ref{eq:var})
-\begin{align}
-\int_{D^k} u  v &=\frac{1}{2} \int_{{D^k}} \LRp{f\pd{\Pi_N g}{{x}_i}v - g \pd{\Pi_N\LRp{fv}}{{x}_i}} \label{eq:skewDN}\\
-&+\frac{1}{2} \int_{\partial {D}^k} \LRp{fgv + \LRp{g - \Pi_Ng} \LRp{{fv + \Pi_N(fv)}{}}} {n}^k_i \qquad \forall v\in V^N\nonumber,
-\end{align}
-which yields the matrix formulation
-\begin{align}
-\bm{M}\bm{u} =& \frac{1}{2}\LRs{\begin{array}{c}
-\bm{V}_q \\ \bm{V}_f\end{array}}^T \diag{\bm{f}_N}\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T + \bm{B}^i_N}\bm{g}_N \label{eq:dsbpmatrixform}
-%&+ \frac{1}{2}\bm{V}_f^T\bm{W}_f\diag{\bm{n}^k_i} \LRp{\bm{f}_f \circ \bm{g}_f}.\nonumber
-\end{align}
-
-The accuracy of the formulations (\ref{eq:esdgSkew}) and (\ref{eq:skewformcurved}) can be understood by analyzing the degree of polynomial exactness of (\ref{eq:dsbpmatrixform}) as an approximation of the derivative.
-Let $u(\bm{x})$ be a polynomial of degree $\leq N$ with coefficients $\bm{u}$, and let $\bm{u}_N = \LRs{\bm{u}_q, \bm{u}_f}^T$ denote the values of $u(\bm{x})$ at volume and surface quadrature points.
-An approximation of $\pd{u}{x_i}$ can be computed by applying (\ref{eq:dsbpmatrixform})
-\begin{align}
-\pd{u}{x_i}\approx \frac{1}{2}\bm{M}^{-1}\LRs{\begin{array}{c}
-\bm{V}_q \\ \bm{V}_f\end{array}}^T \LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T + \bm{B}^i_N}\bm{u}_N.
-\label{eq:dsbpapprox}
-\end{align}
-
-From (\ref{eq:dsbpapprox}), it can be shown that when $\bm{Q}^i$ satisfies a generalized SBP property, the decoupled SBP operator (\ref{eq:dsbp}) produces a degree $N$ approximation to the derivative \cite{chan2017discretely}.
-When $\bm{Q}^i$ does not satisfy a generalized SBP property, we have the following lemma on the accuracy of (\ref{eq:dsbpapprox}):
-\begin{lemma}
-Suppose that the volume quadrature is exact for degree $M+N-1$ polynomials on simplices, or for polynomials in $Q^{M+N-1,M+N,M+N}$ on tensor product elements.
-Furthermore, assume that the surface quadrature is exact for degree $M+N$ polynomials on simplices and $Q^{M+N,M+N}$ on tensor product elements.
-Then, the skew-symmetric approximation of the derivative (\ref{eq:dsbpapprox}) is exact for polynomials of degree $M$.
-\label{lemma:dsbpapprox}
-\end{lemma}
-\begin{proof}
-Suppose $u \in P^M$.
-Let $\bm{u}$ denote the polynomial coefficients of $u$, and let $\bm{e}$ denote the difference between $\bm{D}^i\bm{u}$ (the exact coefficients of $\pd{u}{x_i}$) and the approximation (\ref{eq:dsbpapprox})
-\[
-\bm{e} = \bm{D}^i\bm{u} - \frac{1}{2}\bm{M}^{-1}\LRs{\begin{array}{c}
-\bm{V}_q \\ \bm{V}_f\end{array}}^T\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T + \bm{B}^i_N}\bm{u}_N,
-\]
-where $\bm{e}$ is a polynomial of degree $N$.  
-If Assumption~\ref{ass:quad} is satisfied and $u(\bm{x})$ is polynomial, then the values of $u(\bm{x})$ at quadrature points are $\bm{u}_q = \bm{V}_q\bm{u}$ and $\bm{u} = \bm{P}_q\bm{u}_q$.
-This implies that $\bm{u}_f = \bm{V}_f \bm{P}_q\bm{u}_q =  \bm{E}\bm{u}_q$.
-Expanding the latter term yields
-\begin{align*}
-&\frac{1}{2}\bm{M}^{-1}\LRs{\begin{array}{c}
-\bm{V}_q \\ \bm{V}_f\end{array}}^T\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T + \bm{B}^i_N}\bm{u}_N \\
-&= \frac{1}{2}\bm{M}^{-1}\LRs{\begin{array}{c}
-\bm{V}_q \\ \bm{V}_f\end{array}}^T\begin{bmatrix}
-\bm{Q}^i - \LRp{\bm{Q}^i}^T & \bm{E}^T\bm{B}^i\\
--\bm{B}^i\bm{E} & \bm{B}^i
-\end{bmatrix}\begin{bmatrix}
-\bm{u}_q\\
-\bm{u}_f
-\end{bmatrix}\\
-&= \frac{1}{2}\bm{M}^{-1}\LRs{\begin{array}{c}
-\bm{V}_q \\ \bm{V}_f\end{array}}^T\begin{bmatrix}
-\bm{Q}^i\bm{u}_q + \LRp{\bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T}\bm{u}_q\\
-\bm{B}^i\LRp{\bm{u}_f - \bm{E}\bm{u}_q}
-\end{bmatrix} 
-\\
-&=\frac{1}{2}\bm{M}^{-1}
-\bm{V}_q^T\LRp{\bm{Q}^i\bm{u}_q + \LRp{\bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T}\bm{u}_q}.
-%&=\frac{1}{2} \bm{D}^i\bm{u} + \frac{1}{2}\bm{M}^{-1}
-%\bm{V}_q^T{ \LRp{\bm{E}^T\bm{B}^i\bm{E} - \bm{Q}^i}\bm{u}_q}
-\end{align*}
-Since $\bm{Q}^i = \bm{W}\bm{V}_q\bm{D}^i\bm{P}_q$ and $\bm{M} = \bm{V}_q^T\bm{W}\bm{V}_q$, we have that $\bm{M}^{-1}\bm{V}_q^T\bm{Q}^i\bm{u}_q = \bm{D}^i\bm{u}$.
-This simplifies the expression for error to
-\begin{align}
-\bm{e}^T\bm{M}\bm{e} &= \frac{1}{2}
-\bm{e}_q^T\LRp{-\bm{Q}^i\bm{u}_q + \LRp{\bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T}\bm{u}_q} \label{eq:skewerr}\\
-&= -\frac{1}{2}
-\bm{e}_q^T\bm{Q}^i\bm{u}_q + \frac{1}{2}\bm{e}_q^T\LRp{\bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T}\bm{u}_q, \nonumber
-\end{align}
-where we have introduced $\bm{e}_q = \bm{V}_q\bm{e}$.
-The remainder of the proof is similar to the proof of Lemma~\ref{lemma:vsbp}.
-Since $u \in P^M$ and $e \in P^N$, by exactness of the quadrature rules,
-\[
-\bm{e}_q^T\LRp{\bm{E}^T\bm{B}^i\bm{E} - \LRp{\bm{Q}^i}^T}\bm{u}_q = \int_{\partial \hat{D}} u e n_i - \int_{\hat{D}} u \pd{e}{x_i}
-= \int_{\hat{D}} \pd{u}{x_i} e = \bm{e}_q^T\bm{Q}^i\bm{u}_q.
-\]
-Combining this with (\ref{eq:skewerr}) implies that $\bm{e}^T\bm{M}\bm{e} = 0$, and since $\bm{M}$ is symmetric positive definite, $\bm{e} = 0$.  
-\qed\end{proof}
-
-Lemma~\ref{lemma:dsbpapprox} suggests that, when a generalized SBP property does not hold, the use of under-integrated quadratures results in a loss of one or more orders of accuracy.
-For example, if the SBP property does not hold, then using $(N+1)$ point GLL rules (which are exact for only polynomials of degree $2N-1$) for either volume or surface quadratures should result in a loss of one order of accuracy compared to the use of $(N+1)$-point Gauss rules (which are exact for polynomials of degree $2N$).
-This is indeed observed in numerical experiments.
 
@@
 
-All three options result in similar computational costs on on triangles, but slight differences in computational cost and complexity on quadrilaterals.
+All three options result in similar computational costs on triangles, but slight differences in computational cost and complexity on quadrilaterals.
 On quadrilaterals, Option~\ref{opt:1} is the most computationally efficient option, as the formulation (\ref{thm:skewformcurved}) reduces to a standard entropy stable DG-SEM \cite{gassner2016split} or spectral collocation method \cite{carpenter2014entropy}.
@@
 
-We first verify that, for an entropy conservative flux and periodic domain, the spatial formulation tested against the projected entropy variables (\ref{eq:entropyrhs}) is numerically zero.
+We first verify that, for an entropy conservative flux and periodic domain, the spatial formulation tested against the projected entropy variables is numerically zero.
 We refer to this quantity as the entropy right-hand side (RHS).
@@
 We observe that for all $N_{\rm geo} \leq M+1$, the maximum entropy RHS is $O(10^{-14})$ and at the level of machine precision.
-When $N_{\rm geo}> M+1$, we observe that the maximum entropy RHS increases .
+When $N_{\rm geo}> M+1$, we observe that the maximum entropy RHS increases.
 We note that the case of $M= 5$ for the quadrilateral corresponds to the use of an $(N+1)$-point GLL rule for both volume and surface quadrature.
-The strong SBP property holds for this choice of quadrature, and the skew-symmetric formulation is equivalent to the entropy stable spectral collocation or DG-SEM  methods of \cite{carpenter2014entropy, gassner2016split}.
+For this choice of quadrature, the skew-symmetric formulation is equivalent to the entropy stable spectral collocation or DG-SEM  methods of \cite{carpenter2014entropy, gassner2016split}.
 
@@
 For $N=4$, Option~\ref{opt:2} achieves the same level of error observed for Option~\ref{opt:3}.
-This suggests that Lemma~\ref{lemma:dsbpapprox} may be sufficient but not necessary for full order accuracy.\footnote{We note that these results may also differ depending on the type of interface dissipation applied \cite{hindenlang2019order}.}
+This suggests that Lemma~\ref{lemma:dsbpapprox} may be sufficient but not necessary for full order accuracy.
+These results may also differ depending on the type of interface dissipation applied \cite{hindenlang2019order}.
 
@@
 \subfloat{\raisebox{3em}{
-\includegraphics[width=.43\textwidth]{figs/hybrid_mesh.png}
+\includegraphics[width=.43\textwidth]{hybrid_mesh.png}
 }\label{fig:hybridmesh}}
@@
 These formulations are motivated by volume and surface quadratures which arise naturally on hybrid meshes.
-Because these quadrature rules do not necessarily induce operators which satisfy a ``strong'' matrix summation-by-parts property, we generalize proofs of entropy stability and free stream (constant) preservation using a weaker ``variational'' SBP property.
-
-This variational SBP property is also used to construct explicit conditions relating between the order of accuracy and degree of accuracy of each quadrature rule.
-We also derive conditions under which the skew-symmetric formulation is entropy stable on curved meshes in terms of the degree of quadrature accuracy and polynomial degree of the geometric mapping.
+\bnote{Because these quadrature rules do not induce operators which satisfy properties necessary for entropy stability, we derive new ``skew-symmetric'' operators which satisfy necessary conditions under reduced restrictions on quadrature.
+We also derive a separate set of conditions relating the accuracy of the new operators and the degree of accuracy of each quadrature rule, and show that design order accuracy is recovered under the common assumptions of degree $2N-1$ volume quadratures and degree $2N$ surface quadratures.
+} Finally, we derive conditions under which the skew-symmetric formulation is entropy stable on curved meshes in terms of the degree of quadrature accuracy and polynomial degree of the geometric mapping.
 Numerical experiments confirm the entropy stability and high order accuracy of the proposed schemes on triangular, quadrilateral, and 2D hybrid meshes.
@@
 
-The author thanks David C.\ Del Rey Fernandez for helpful discussions.  Jesse Chan is supported by the National Science Foundation under awards DMS-1719818 and DMS-1712639.
+The author thanks David C.\ Del Rey Fernandez for helpful discussions, \rnote{as well as the two anonymous reviewers whose comments significantly improved the readabilty of this manuscript}.
+Jesse Chan is supported by the National Science Foundation under awards DMS-1719818 and DMS-1712639.
 

### 2019-09-06 docs/skewform/main.tex
updating docs

@@
 \renewcommand{\note}[1]{{\color{blue}{#1}}}
+\newcommand{\bnote}[1]{{\color{blue}{#1}}}
+\newcommand{\rnote}[1]{{\color{red}{#1}}}
 
@@
 \begin{abstract}
-Entropy stable high order methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.
-The construction of such schemes has relied on the use of carefully chosen collocation points \cite{gassner2013skew, fisher2013high, carpenter2014entropy, crean2018entropy, chan2018efficient} or volume and surface quadrature rules \cite{chan2017discretely, chan2018discretely} to produce operators which satisfy a summation-by-parts (SBP) property.
-In this work, we show how to construct skew-symmetric schemes which are entropy stable even for volume and surface quadratures under which a traditional SBP property does not hold.
-These skew-symmetric formulations avoid the use of a ``strong'' SBP property, and require only that operators exactly differentiate constants and satisfy a discrete form of the fundamental theorem of calculus.
-We conclude with an application of the skew-symmetric formulation for entropy stable schemes on mixed quadrilateral-triangle meshes.
+High order entropy stable discontinuous Galerkin (DG) methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.
+The construction of such schemes has relied on the use of carefully chosen nodal points \cite{gassner2013skew, fisher2013high, carpenter2014entropy, crean2018entropy, chan2018efficient} or volume and surface quadrature rules \cite{chan2017discretely, chan2018discretely} to produce operators which satisfy a summation-by-parts (SBP) property.
+In this work, we show how to construct ``modal'' DG formulations which are \bnote{entropy stable for volume and surface quadratures under which the SBP property in \cite{chan2017discretely} does not hold.
+These formulations rely on an alternative skew-symmetric construction of operators which automatically satisfy the SBP property.
+Entropy stability then follows for choices of volume and surface quadrature which satisfy sufficient accuracy conditions.}   \rnote{The accuracy of these new SBP operators is shown to depend on a separate set of accuracy conditions.
+}
+%avoid the use of a ``strong'' matrix-based SBP property, and instead rely on the assumption that discrete operators exactly differentiate constants and satisfy a discrete form of the fundamental theorem of calculus.
+We conclude with numerical experiments verifying the accuracy and stability of the proposed formulations, and discuss an application of these formulations for entropy stable DG schemes on mixed quadrilateral-triangle meshes.
 \end{abstract}
 
-\tableofcontents
+%\tableofcontents
 
@@
 
-Entropy stable discretizations are built upon flux differencing and a summation-by-parts (SBP) property, which serves as a discrete analogue of integration by parts for quadrature-based discretization matrices.
-However, the SBP property does not hold for some under-integrated quadrature rules, which naturally arise in certain discretization settings.
+Entropy stable ``modal'' DG discretizations \cite{chan2017discretely, chan2018discretely} are built upon flux differencing and a summation-by-parts (SBP) property, which serves as a discrete analogue of integration by parts for quadrature-based discretization matrices.
+However, the SBP property does not hold for certain under-integrated quadrature rules, which arise naturally in some discretization settings.
 For example, on hybrid meshes consisting of both quadrilateral and triangular elements, it is convenient to utilize the same quadrature rule on shared faces between different element types.
-On degree $N$ tensor product elements, a popular choice of quadrature is an $(N+1)$-point Gauss-Legendre-Lobatto (GLL) rule.
+On degree $N$ quadrilateral elements, a popular choice of quadrature is an $(N+1)$-point Gauss-Legendre-Lobatto (GLL) rule.
 When both volume and surface integrals are approximated using $(N+1)$ point GLL quadrature rules, the SBP property holds, despite the fact that GLL quadrature is inexact for the integrands which appear in finite element formulations \cite{fisher2013high}.
-However, while GLL quadrature induces an SBP property on quadrilateral elements, but does not always result in an SBP property if used for triangular elements \cite{chan2017discretely}.
-%A similar issue arises on non-conforming meshes, where a composite ``split-side'' quadrature naturally when coupling larger elements to smaller elements through a non-conforming interface \cite{kozdon2018energy}.
-For quadrilaterals utilizing a GLL volume quadrature, the use of a composite quadrature on a non-conforming face also results in the loss of the SBP property \cite{chan2017discretely}.
-
-This work proposes an alternative formulation which utilizes a skew-symmetric form of flux differencing.
-Under such a formulation, the proof of entropy stability requires only a weaker ``variational'' form of the SBP property, which holds under a more general class of quadrature rules compared to the decoupled SBP property introduced in \cite{chan2017discretely, chan2018discretely}.
-We show that the skew-symmetric formulation is entropy stable, locally conservative, and free-stream preserving on curved elements, and confirm theoretical results with numerical experiments on triangular, quadrilateral, and hybrid triangular-quadrilateral meshes.
-
-It should be noted that a similar approach to entropy stable discretizations was introduced within a finite difference framework \cite{chen2017entropy, crean2018entropy} using multidimensional differencing operators which are constructed to satisfy accuracy conditions and an SBP property \cite{hicken2016multidimensional}.
-These operators do not correspond to any specific basis or approximation space, but can be shown to exist for nodal points corresponding to sufficiently accurate choices of volume and surface quadrature and can be computed either algebraically or through an optimization problem.
-The formulations presented in this work require similar accuracy conditions, but differ from those of \cite{chen2017entropy, crean2018entropy} in that we start with an explicit approximation space, which then induce quadrature-based operators.
-The formulations presented here also accomodate general choices of volume and surface quadrature (e.g.\ volume quadratures without boundary nodes and over-integrated rules) while ensuring compact coupling conditions between neighboring elements.
+However, while GLL quadrature induces an SBP property on quadrilateral elements, it does not guarantee an SBP property if used on triangular elements \cite{chan2017discretely}.
+
+This work proposes an alternative formulation which utilizes a \note{skew-symmetric construction of the SBP operator}.
+Under such a formulation, the proof of entropy stability \note{holds under weaker quadrature rules} compared to the decoupled SBP property introduced in \cite{chan2017discretely, chan2018discretely}.
+We show that this skew-symmetric formulation is entropy stable, locally conservative, and free-stream preserving on curved elements, and confirm theoretical results with numerical experiments on triangular, quadrilateral, and hybrid triangular-quadrilateral meshes.
+
+It should be noted that a similar approach to entropy stable discretizations was introduced within a finite difference framework \cite{chen2017entropy, crean2018entropy} using multidimensional differencing operators which satisfy similar accuracy conditions and an SBP property \cite{hicken2016multidimensional}.
+These operators exist for nodal points corresponding to sufficiently accurate choices of volume and surface quadrature, but do not correspond to any specific basis or approximation space.
+The formulations in \cite{chen2017entropy, crean2018entropy} differ from the ones presented in this work in that they are based on SBP finite differences and ``nodal'' (rather than ``modal'') DG formulations, with differentiation operators computed algebraically or through an optimization problem for each specific choice of nodes.
+In contrast, ``modal'' formulations induce quadrature-based operators from an explicit approximation space, and accomodate general choices of volume and surface quadrature (e.g.\ volume quadratures without boundary nodes and over-integrated quadrature rules).
 
@@
 Let $\hat{\bm{n}}_i$ denote the vector containing values of the $i$th component $\hat{n}_i\hat{J}f$ at all surface quadrature points, and let $\bm{Q}^i = \bm{W}\bm{D}^i_q = \bm{W}\bm{V}_q\bm{D}^i\bm{P}_q$.
-The ``decoupled'' summation by parts operator $\bm{D}^i_N$ is defined as the block matrix involving both volume and surface quadratures
+The ``hybridized'' (originally referred to as ``decoupled'' in \cite{chan2017discretely}) summation by parts operator $\bm{D}^i_N$ is defined as the block matrix involving both volume and surface quadratures
 \begin{gather}
@@
 
-\subsection{A variational SBP property}
+\subsection{\bnote{An alternative construction of SBP operators}} % variational SBP property}
+
+%\subsection{A variational SBP property}
 
@@
 
-The property (\ref{eq:dsbp}) (which we will refer to as the ``strong'' SBP property) relates the polynomial exactness of specific quadrature rules to algebraic properties of quadrature-based matrices.
-We will relax accuracy conditions on these quadrature rules by utilizing a weaker ``variational'' version of the SBP property variational, which relies on the exactness of volume and surface quadrature rules only for specific polynomials.
+The property (\ref{eq:dsbp}) relates the polynomial exactness of specific quadrature rules to algebraic properties of quadrature-based matrices.
+We will relax accuracy conditions on these quadrature rules by utilizing \note{an alternative construction of $\bm{Q}^i_N$ based the skew-symmetric matrix $\bm{Q}^i - \LRp{\bm{Q}^i}^T$:}
+%a weaker ``variational'' version of the SBP property variational, which relies on the exactness of volume and surface quadrature rules only for specific polynomials.  
+\bnote{\begin{equation}
+\tilde{\bm{Q}}^i_{N} = \begin{bmatrix}
+\bm{Q}^i - \LRp{\bm{Q}^i}^T &  \frac{1}{2}\bm{E}^T\bm{B}^i\\
+-\frac{1}{2}\bm{B}^i\bm{E} & \frac{1}{2} \bm{B}^i
+\end{bmatrix}.
+\label{eq:skewQN} 
+\end{equation}
+It is clear to see that the skew-symmetric construction of $\tilde{\bm{Q}}^i_N$ as defined in (\ref{eq:skewQN}) satisfies the SBP property (\ref{eq:dsbp}) by construction.
+However, the accuracy of $\tilde{\bm{Q}}^i_N$ as a differentiation operator becomes slightly more difficult to analyze.
+}
 

### 2019-09-06 docs/skewform/main_v2.tex
updating docs

--- 2019-09-06 docs/skewform/arxiv_v2/main_v2.tex
@@
-%\pdfoutput=1
+\pdfoutput=1
 
@@
 \usepackage{cite}
+
+%\usepackage{slashbox} % for slashbox in tables
 
@@
 \usepackage{pgfplotstable}
-\definecolor{markercolor}{RGB}{124.9, 255, 160.65}
-%\usepackage{slopetri} % my own package for drawing triangle slopes
+\usepackage{slopetri} % my own package for drawing triangle slopes
 
@@
 
-%\renewcommand{\note}[1]{{\color{blue}{#1}}}
-%\newcommand{\bnote}[1]{{\color{blue}{#1}}}
-%\newcommand{\rnote}[1]{{\color{red}{#1}}}
-\renewcommand{\note}[1]{#1}
-\newcommand{\bnote}[1]{#1}
-\newcommand{\rnote}[1]{#1}
+\renewcommand{\note}[1]{{\color{blue}{#1}}}
+\newcommand{\bnote}[1]{{\color{blue}{#1}}}
+\newcommand{\rnote}[1]{{\color{red}{#1}}}
+
 \graphicspath{{./figs/}}
@@
 Let $\hat{D} \subset\mathbb{R}^d$ denote a reference element with surface $\partial \hat{D}$.  
-The high order schemes in \cite{chan2017discretely, chan2018discretely} begin by approximating the solution in a degree $N$ polynomial basis $\LRc{\phi_j(\hat{\bm{x}})}_{\rnote{j}=1}^{N_p}$ on $\hat{D}$.
-These schemes also assume volume and surface quadrature rules $(\hat{\bm{x}}_i, w_i)$, $\LRp{\hat{\bm{x}}^f_i,w^f_i}$ on $\hat{D}$.
+The high order schemes in \cite{chan2017discretely, chan2018discretely} begin by approximating the solution in a degree $N$ polynomial basis $\LRc{\phi_j({\bm{x}})}_{\rnote{j}=1}^{N_p}$ on $\hat{D}$.
+These schemes also assume volume and surface quadrature rules $({\bm{x}}_i, w_i)$, $\LRp{{\bm{x}}^f_i,w^f_i}$ on $\hat{D}$.
 We will specify the accuracy of each quadrature rule later, and discuss how quadrature accuracy implies specific operator properties.
@@
 \begin{gather}
-\LRp{\bm{V}_q}_{ij} = \phi_j(\hat{\bm{x}}_i), \qquad \LRp{\bm{V}_f}_{ij} = \phi_j(\hat{\bm{x}}^f_i), \qquad \pd{\phi_j(\hat{\bm{x}})}{\hat{x}_i} = \sum_{k=1}^{N_p} \LRp{\bm{D}^i_{jk}} \phi_k(\hat{\bm{x}}).
+\LRp{\bm{V}_q}_{ij} = \phi_j(\bm{x}_i), \qquad \LRp{\bm{V}_f}_{ij} = \phi_j(\bm{x}^f_i), \qquad \pd{\phi_j(\bm{x})}{x_i} = \sum_{k=1}^{N_p} \LRp{\bm{D}^i_{jk}} \phi_k(\bm{x}).
 \end{gather}
-The interpolation matrices $\bm{V}_q,\bm{V}_f$ map basis coefficients to evaluations at volume and surface quadrature points respectively, while the differentiation matrix ${\bm{D}}_i$ maps basis coefficients of a function to the basis coefficients of its derivative with respect to $\hat{x}_k$.
+The interpolation matrices $\bm{V}_q,\bm{V}_f$ map basis coefficients to evaluations at volume and surface quadrature points respectively, while the differentiation matrix ${\bm{D}}_i$ maps basis coefficients of a function to the basis coefficients of its derivative with respect to $x_k$.
 The interpolation matrices are used to assemble the mass matrix $\bm{M}$, the quadrature-based projection matrix $\bm{P}_q$, and lifting matrix $\bm{L}_f$
@@
 Let $\hat{\bm{n}}$ denote the scaled outward normal vector $\hat{\bm{n}} = \LRc{\hat{n}_1\hat{J}_f,\ldots,\hat{n}_d\hat{J}_f}$, where $\hat{J}_f$ is the determinant of the Jacobian of the mapping of a face of $\partial \hat{D}$ to a reference face.
-Let $\hat{\bm{n}}_i$ denote the vector containing values of the $i$th component $\hat{n}_i\hat{J}_f$ at all surface quadrature points, and define the generalized SBP operator
+Let $\hat{\bm{n}}_i$ denote the vector containing values of the $i$th component $\hat{n}_i\hat{J}f$ at all surface quadrature points, and define the generalized SBP operator
 \[
@@
 \end{gather}
-then the hybridized SBP operator $\bm{Q}^i_N$ (\ref{eq:QN}) satisfies a summation by parts property:
+then the decoupled SBP operator $\bm{Q}^i_N$ (\ref{eq:QN}) satisfies a summation by parts property:
 \begin{gather}
@@
 \[
-\int_{\hat{D}} \pd{u}{\hat{x}_i} v, \qquad \int_{\partial \hat{D}} u v \hat{n}_i
-\]
-are integrated exactly for all $u,v \in V^N\LRp{\hat{D}}$ and $i = 1,\ldots, d$.
+\int_{\hat{D}} \pd{u}{\hat{x}_j} v, \qquad \int_{\partial \hat{D}} u v \hat{n}_j
+\]
+are integrated exactly for all $u,v \in V^N\LRp{\hat{D}}$ and $j = 1,\ldots, d$.
 This implies that Theorem~\ref{lemma:dsbp} is satisfied for sufficiently accurate volume and surface quadratures.
@@
 We first illustrate how specific pairings of volume and surface quadratures can result in the loss of the SBP property (\ref{eq:dsbp}) \bnote{for $\bm{Q}^i_N$}.
-We then propose an \note{alternative skew-symmetric version of the hybridized SBP operator which satisfies the SBP property by construction}.
+We then propose a \note{alternative skew-symmetric version of the hybridized SBP operator which satisfies the SBP property by construction}.
 The use of these operators results in formulations which are entropy conservative \bnote{under a wider range of quadratures.}
@@
 \centering
-\subfloat[GLL volume quadrature, Gauss surface quadrature]{\includegraphics[width=.4\textwidth]{gllgauss.png}\label{subfig:gllgq}}
+\subfloat[GLL volume quadrature, Gauss surface quadrature]{\includegraphics[width=.4\textwidth]{figs/gllgauss.png}\label{subfig:gllgq}}
 \hspace{2em}
-\subfloat[Degree $2N$ volume quadrature, GLL surface quadrature]{\includegraphics[width=.4\textwidth]{trigll.png}\label{subfig:trigll}}
+\subfloat[Degree $2N$ volume quadrature, GLL surface quadrature]{\includegraphics[width=.4\textwidth]{figs/trigll.png}\label{subfig:trigll}}
 \caption{Volume and surface quadrature pairs which do not satisfy the assumptions of Theorem~\ref{lemma:dsbp}, and thus do not possess the decoupled SBP property (\ref{eq:dsbp}).
@@
 Let $u,v \in Q^N$ denote two arbitrary degree $N$ polynomials.
-The assumptions of Theorem~\ref{lemma:dsbp} are that the volume quadrature exactly integrates $\int_{\hat{D}} \pd{u}{x_i} v$ and that the surface quadrature exactly integrates $\int_{\partial \hat{D}} u v \hat{n}_i$ on $\hat{D}$.
+The assumptions of Theorem~\ref{lemma:dsbp} are that the volume quadrature exactly integrates $\int_{\hat{D}} \pd{u}{x_j} v$ and that the surface quadrature exactly integrates $\int_{\partial \hat{D}} u v \hat{n}_j$ on $\hat{D}$.
 Because the $(N+1)$-point Gauss rule is exact for polynomials of degree $2N+1$ and the product $uv \in P^{2N}$ on each face, the surface quadrature satisfies the assumptions of Theorem~\ref{lemma:dsbp}.
@@
 \captionsetup[subfigure]{width=.425\textwidth}
-\subfloat[Insufficiently accurate surface quadrature on the triangle element.]{\includegraphics[width=.425\textwidth]{hybrid2D.png}\label{subfig:hybrid1}}
+\subfloat[Insufficiently accurate surface quadrature on the triangle element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D.png}\label{subfig:hybrid1}}
 \endgroup
 \hspace{2em}
-\subfloat[Incompatible surface quadrature on the quadrilateral element.]{\includegraphics[width=.425\textwidth]{hybrid2D_GQ.png}\label{subfig:hybrid2}}
+\subfloat[Incompatible surface quadrature on the quadrilateral element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D_GQ.png}\label{subfig:hybrid2}}
 \caption{Examples of interface couplings which do not result in a decoupled SBP property (\ref{eq:dsbp}).
@@
 \end{gather}
-for $i = 1,\ldots, d$.  The construction of entropy stable schemes will utilize \rnote{(\ref{eq:esflux})} in discretizations of both volume and surface terms in a DG formulation.
+for $j = 1,\ldots, d$.  The construction of entropy stable schemes will utilize \rnote{(\ref{eq:esflux})} in discretizations of both volume and surface terms in a DG formulation.
 
@@
 \LRp{2\tilde{\bm{Q}}^i_N \circ \bm{F}^i_S}\bm{1} + \bm{V}_f^T\bm{B}^i\LRp{\bm{f}_i^*-\bm{f}(\tilde{\bm{u}}_f)} = 0,  \label{eq:esdgSkew}\\
-\LRp{\bm{F}^i_S}_{jk} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}, \qquad 1\leq j,k \leq N_q + N^f_q,\nonumber
+\LRp{\bm{F}^i_S}_{ij} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q,\nonumber
 \end{gather}
-where $\tilde{\bm{u}}_f$ denotes the values of $\tilde{\bm{u}}$ on face nodes and $\bm{f}^*$ is some numerical flux, and $N_q, N^f_q$ denote the number of volume and face quadrature points, respectively.	.
+where $\tilde{\bm{u}}_f$ denotes the values of $\tilde{\bm{u}}$ on face nodes and $\bm{f}^*$ is some numerical flux.
 This formulation is identical to that of \cite{chan2017discretely}, except that the hybridized SBP operators $\bm{Q}^i_N$ are replaced with their skew-hybridized versions $\tilde{\bm{Q}}^i_N$.
@@
 \begin{equation}
-\bm{1}^T\bm{W}\td{U(\bm{u}_q)}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}^i \LRp{\tilde{\bm{v}}_f^T\bm{f}_i^* - \psi_i(\tilde{\bm{u}}_f)} = 0, \qquad \bm{u}_q = \bm{V}_q\bm{u}_h.
+\bm{1}^T\bm{W}\td{U(\bm{u}_q)}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}^i \LRp{\tilde{\bm{v}}_f^T\bm{f}_i^* - \psi_i(\tilde{\bm{u}}_f)} = 0, \qquad \bm{u}_q = \bm{V}_q\bm{u}.
 \label{eq:esdgthm}
@@
 \LRp{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T} \circ \bm{F}^i_S}\bm{1} + \bm{V}_f^T\bm{B}^i\bm{f}_i^* = 0,  \label{eq:esdgSkew2}\\
-\LRp{\bm{F}^i_S}_{jk} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}, \qquad 1\leq j,k \leq N_q + N^f_q,\nonumber
+\LRp{\bm{F}^i_S}_{ij} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q,\nonumber
 \end{gather}
@@
 In \cite{chan2017discretely}, it was shown that the hybridized SBP operator satisfies $\bm{Q}^i_N\bm{1} = \bm{0}$ for any volume quadrature such that the mass matrix $\bm{M}$ is positive-definite.
-However, ensuring that the skew-hybridized SBP operator satisfies $\tilde{\bm{Q}}^i_N\bm{1} = \bm{0}$ now requires conditions on both volume and surface quadratures which are related to a weak version of the generalized SBP condition (\ref{eq:gsbp}).
+However, ensuring that the skew-hybridized SBP operator satisfies $\tilde{\bm{Q}}^i_N\bm{1} = \bm{0}$ now requires conditions on both volume and surface quadratures which are related to a weak version of the GSBP condition (\ref{eq:gsbp}).
 }
@@
 \]
-However, under Assumption~\ref{ass:quad}, the entries of $\LRp{\bm{Q}^i}^T \bm{1}$ are exactly $\int_{\hat{D}} \pd{\phi_j}{\hat{\bm{x}}_i}$ and the entries of $\bm{E}^T\bm{B}^i \bm{1}$ are exactly $\int_{\partial \hat{D}} \phi_j(\bm{x}) \hat{n}_i$ since $\phi_j(\bm{x}) \in V^N$.
+However, under Assumption~\ref{ass:quad}, the entries of $\LRp{\bm{Q}^i}^T \bm{1}$ are exactly $\int_{\hat{D}} \pd{\phi_j}{\bm{x}_i}$ and the entries of $\bm{E}^T\bm{B}^i \bm{1}$ are exactly $\int_{\partial \hat{D}} \phi_j(\bm{x}) \hat{n}_i$ since $\phi_j(\bm{x}) \in V^N$.
 These two terms are then identical by the exactness of integrals and fundamental theorem of calculus.
@@
 \bm{M}\bm{u} =& \frac{1}{2}\LRs{\begin{array}{c}
-\bm{V}_q \\ \bm{V}_f\end{array}}^T \diag{\bm{f}_N}\underbrace{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T + \bm{B}^i_N}}_{\tilde{\bm{Q}}^i_N}\bm{g}_N \label{eq:dsbpmatrixform2}.
+\bm{V}_q \\ \bm{V}_f\end{array}}^T \diag{\bm{f}_N}\underbrace{\LRp{\bm{Q}^i_N - \LRp{\bm{Q}^i_N}^T + \bm{B}^i_N}}_{\tilde{\bm{Q}}^i_N}\bm{g}_N \label{eq:dsbpmatrixform}.
 %&+ \frac{1}{2}\bm{V}_f^T\bm{W}_f\diag{\bm{n}^k_i} \LRp{\bm{f}_f \circ \bm{g}_f}.\nonumber
@@
 
-The accuracy of the formulation (\ref{eq:esdgSkew}) can be understood by analyzing the degree of polynomial exactness of (\ref{eq:dsbpmatrixform2}) as an approximation of the derivative.
+The accuracy of the formulation (\ref{eq:esdgSkew}) can be understood by analyzing the degree of polynomial exactness of (\ref{eq:dsbpmatrixform}) as an approximation of the derivative.
 Let $u(\bm{x})$ be a polynomial of degree $\leq N$ with coefficients $\bm{u}$, and let $\bm{u}_N = \LRs{\bm{u}_q, \bm{u}_f}^T$ denote the values of $u(\bm{x})$ at volume and surface quadrature points.
@@
 \end{align}
-From (\ref{eq:dsbpapprox}), it can be shown that when $\bm{Q}^i$ satisfies a generalized SBP property, the hybridized SBP operator (\ref{eq:dsbp}) produces a degree $N$ approximation to the derivative \cite{chan2017discretely}.
+From (\ref{eq:dsbpapprox}), it can be shown that when $\bm{Q}^i$ satisfies a generalized SBP property, the decoupled SBP operator (\ref{eq:dsbp}) produces a degree $N$ approximation to the derivative \cite{chan2017discretely}.
 When $\bm{Q}^i$ does not satisfy a generalized SBP property, we have the following lemma on the accuracy of (\ref{eq:dsbpapprox}):
 \begin{lemma}
-Let $M\leq N$.
 Suppose that the volume quadrature is exact for degree $M+N-1$ polynomials on simplices, or for polynomials in $Q^{M+N-1,M+N,M+N}$ on tensor product elements.
 Furthermore, assume that the surface quadrature is exact for degree $M+N$ polynomials on simplices and $Q^{M+N,M+N}$ on tensor product elements.
-\bnote{Then, so long as the mass matrix is positive definite}, the skew-symmetric approximation of the $x$-derivative (\ref{eq:dsbpapprox}) is exact for polynomials of degree $M$.
+\bnote{Then, so long as the mass matrix is positive definite}, the skew-symmetric approximation of the derivative (\ref{eq:dsbpapprox}) is exact for polynomials of degree $M$.
 \label{lemma:dsbpapprox}
@@
 \begin{gather}
-\pd{u}{x_i} = \sum_{j=1}^d {G}^k_{ij}\pd{u}{\hat{x}_j}, \qquad {G}^k_{ij} = J^k\pd{\hat{x}_j}{{x}_i}, 
+\pd{u}{x_i} = \sum_{ij} {G}^k_{ij}\pd{u}{\hat{x}_j}, \qquad {G}^k_{ij} = J^k\pd{\hat{x}_j}{{x}_i}, 
 \label{eq:geofacs}
@@
 %Let the domain now be triangulated by a  consisting of curved elements $D^k$.  
-In this section, we describe how to construct appropriate hybridized SBP operators on curved meshes, and give conditions on the volume and surface quadrature rules under which a semi-discretely entropy stable scheme can be constructed.
+In this section, we describe how to construct appropriate decoupled SBP operators on curved meshes, and give conditions on the volume and surface quadrature rules under which a semi-discretely entropy stable scheme can be constructed.
 
@@
 
-%\begin{remark}
-%The skew-symmetric operator ${\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T}$ in (\ref{eq:skewformcurved}) can be rewritten as
-%\begin{align*}
-%%\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T =&
-%\sum_{j=1}^d \diag{\bm{G}^k_{ij}}{ \frac{1}{2}\LRp{\tilde{\bm{Q}}_N^j-\LRp{\tilde{\bm{Q}}_N^j}^T}}+  \frac{1}{2} \LRp{\tilde{\bm{Q}}_N^j-\LRp{\tilde{\bm{Q}}_N^j}^T}\diag{\bm{G}^k_{ij}}.
-%\end{align*}
-%Thus, the volume terms in the skew-symmetric formulation may be applied using only the skew-symmetric part of the reference SBP operator $\tilde{\bm{Q}}^j_N$.
-%\end{remark}
+\begin{remark}
+The skew-symmetric operator ${\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T}$ in (\ref{eq:skewformcurved}) can be rewritten as
+\begin{align*}
+%\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T =&
+\sum_{j=1}^d \diag{\bm{G}^k_{ij}}{ \frac{1}{2}\LRp{\tilde{\bm{Q}}_N^j-\LRp{\tilde{\bm{Q}}_N^j}^T}}+  \frac{1}{2} \LRp{\tilde{\bm{Q}}_N^j-\LRp{\tilde{\bm{Q}}_N^j}^T}\diag{\bm{G}^k_{ij}}.
+\end{align*}
+Thus, the volume terms in the skew-symmetric formulation may be applied using only the skew-symmetric part of the reference SBP operator $\tilde{\bm{Q}}^j_N$.
+\end{remark}
 
@@
 \subfloat{\raisebox{3em}{
-\includegraphics[width=.43\textwidth]{hybrid_mesh.png}
+\includegraphics[width=.43\textwidth]{figs/hybrid_mesh.png}
 }\label{fig:hybridmesh}}

### 2019-09-06 docs/ESDG_GQ/figs/compute_op_counts.m

@@ -1,13 +1,25 @@
 N = 1:7;
 
-tp_GLL = 3*(N+1).^4;
-tp_GQ = tp_GLL + 12*(N+1).^3;
-tp_stag = 3*(N+2).^4;
+% num flux evals
+d = 3;
+tp_GLL = d*(N+1).^(d-1).*(N+1).^2;
+tp_GQ = d*(N+1).^(d-1).*((N+1).^2 + 4*(N+1)); % number of lines * 1D nnz
+tp_GQ_sym = d*((N+1).^(d-1)).*((N+1).^2 + 2*(N+1));
+tp_stag = d*(N+2).^(d-1).*(N+2).^2;
 
-% plot(tp_GLL,'o--')
-% hold on
-% plot(tp_GQ,'s--')
-% plot(tp_stag,'x--')
+bar(N,[tp_GLL; tp_GQ; tp_GQ_sym; tp_stag]','grouped')
+legend('Lobatto','Gauss','Sym Gauss','Staggered','Location','Best')
+grid on
+
+return
+
+plot(tp_GLL,'o--')
+hold on
+plot(tp_GQ,'s--')
+plot(tp_GQ_sym,'^--')
+plot(tp_stag,'x--')
+
+return
 
 op_GLL = 3*(N+1).^4;
 op_GQ = op_GLL + 12*(N+1).^3;

### 2019-09-06 docs/ESDG_GQ/sisc/figs/plotDN.m

@@ -1,0 +1,147 @@
+N = 5;
+[rq wq] = JacobiGQ(0,0,N);
+rp = linspace(-1,1,50)';
+Vq = Vandermonde1D(N,rq);
+D = GradVandermonde1D(N,rq)/Vq;
+Vf = Vandermonde1D(N,[-1;1])/Vq;
+Vp = Vandermonde1D(N,rp)/Vq;
+
+B = [-1 0;0 1];
+QN = [D - .5*Vf'*B*Vf .5*Vf'*B;
+    -.5*B*Vf .5*B];
+PN = [eye(N+1) diag(1./wq)*Vf'];
+
+% gex = @(x) exp(x).*sin(pi*x);%exp(-25*x.^2); %
+% gex = @(x) 1./(1+25*(x-0).^2);
+a = 4;
+gex = @(x) exp(-a*x.^2);
+dgex = @(x) -2*a*x.*exp(-a*x.^2);
+fex = @(x) 1 + 0*sin(pi*x);
+f = fex(rq);
+g = gex(rq);
+fdgex = @(x) fex(x).*dgex(x);
+
+fN = fex([rq;-1;1]);
+gN = gex([rq;-1;1]);
+u1 = f.*(D*g);
+u2 = PN*(fN.*(QN*gN));
+hold on
+p1 = plot(rp,Vp*u1,'b-','linewidth',2);
+p2 = plot(rp,Vp*u2,'r-','linewidth',2);
+plot(rq,u1,'bo','linewidth',2,'markersize',15,'MarkerFaceColor',[.49 1 .63])
+plot(rq,u2,'rs','linewidth',2,'markersize',15,'MarkerFaceColor',[.49 1 .63])
+plot(rp,fdgex(rp),'k--','linewidth',2)
+% legend('GSBP','Decoupled SBP')
+set(gca,'fontsize',16)
+grid on
+
+
+% print_pgf_coordinates(rq,u1)
+% print_pgf_coordinates(rp,Vp*u1)
+% print_pgf_coordinates(rq,u2)
+% print_pgf_coordinates(rp,Vp*u2)
+% print_pgf_coordinates(rp,fdgex(rp))
+
+%% convergence
+
+for N = 1:15
+    [rq wq] = JacobiGQ(0,0,N);
+    rp = linspace(-1,1,50)';
+    Vq = Vandermonde1D(N,rq);
+    D = GradVandermonde1D(N,rq)/Vq;
[…]

### 2019-09-06 docs/ESDG_mortar/figs/plot_mortar.m

@@ -18,29 +18,45 @@ sf2 = r1D2;
 % rf2 = [r1D2; e; r1D; -e];
 % sf2 = [e2; r1D; -e; r1D];
 
-a = .25;
-b = .25;
+% mortar
+a = .33;
+b = .75;
+
+% nonmortar values
+b = .33;
+
 plot([-1 1 1 -1 -1],[-1 -1 1 1 -1],'k-','linewidth',2)
 hold on
 plot(rq,sq,'ko','linewidth',2,'markersize',12,'MarkerFaceColor',[.49 1 .63])
 plot(rq,sq,'ko','linewidth',2,'markersize',16,'MarkerFaceColor',[.49 1 .63])
-plot(a+rf,sf,'k-','linewidth',2)
-plot(a+rf,sf,'rs','linewidth',2,'markersize',12,'MarkerFaceColor',[.49 1 .63])
+% plot(a+rf,sf,'k-','linewidth',2)
+% plot(a+rf,sf,'rs','linewidth',2,'markersize',12,'MarkerFaceColor',[.49 1 .63])
 
-% for i = 1:length(rf(:))
-%     for j = 1:length(rf2)
-%         plot([a+rf(i) b + rf2(j)],[sf(i) sf2(j)],'r--','linewidth',2)
-%     end
+
+plot(b+rf2(1:N+1),sf2(1:N+1),'k-','linewidth',2)
+plot(b+rf2(N+2:end),sf2(N+2:end),'k-','linewidth',2)
+plot(b+rf2,sf2,'bs','linewidth',2,'markersize',12,'MarkerFaceColor',[.49 1 .63])
+
+% non-mortar coupling
+rfi = rf2(2)+b;
+sfi = sf2(N+1+3);
+for i = 1:length(rq(:))
+    plot([rq(i) .97*rfi],[sq(i) sfi],'r--','linewidth',2)    
+end
+plot(rfi,sfi,'ro','linewidth',2,'markersize',20)
+
+% % mortar coupling
+% rfi = rf2(2)+b;
+% sfi = sf2(N+1+3);
+% for i = 1:N+1
+%     plot([a+rf(i) .97*rfi],[sf(i) sfi],'r--','linewidth',2)    
 % end
+% plot(rfi,sfi,'ro','linewidth',2,'markersize',20)
 
[…]

### 2019-09-06 talks/ICIAM2019/main.tex
updating rest of talks

\documentclass[compress]{beamer}
\mode<presentation>
\usetheme{Warsaw}
\usecolortheme{seagull}

%\useoutertheme[subsection=false]{smoothbars}

%\usepackage{stackengine}
%\setbeamertemplate{caption}{\raggedright\insertcaption\par}
\setbeamertemplate{caption}{\insertcaption} 
%\usepackage{caption}
%\usepackage{pgffor}

\usepackage{fancybox}
\usepackage{minibox}

%\usepackage{tcolorbox}
\usepackage{empheq}

%% ====================================== graphics

\usepackage{pgfplots}
\pgfplotsset{width=10cm,compat=1.9}
%\usepgfplotslibrary{external}
%\tikzexternalize
 \usepackage{pgfplotstable}

%\definecolor{markercolor}{RGB}{.49, 1, .63}
\definecolor{markercolor}{RGB}{124.9, 255, 160.65}

\pgfplotsset{
tick label style={font=\scriptsize},
label style={font=\scriptsize},
legend style={font=\scriptsize},
title style={font=\footnotesize}
}

%% ===========================================
\usepackage{stmaryrd}

\usepackage{amsmath,amsfonts,amsthm}

% for FD grid tikz
\usepackage{tikz,amssymb}
\usetikzlibrary{calc}
\newcommand\bound{10}

\usepackage{slopetri}

\useoutertheme{infolines}
\useinnertheme{rectangles}
\usepackage{hhline}
\setbeamercovered{dynamic}

\usepackage{soul}

\usepackage{array}
\usepackage{mathrsfs}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{cancel}

\usepackage{graphicx}
\usepackage{subfig}
\makeatletter
\let\@@magyar@captionfix\relax
\makeatother

\usepackage{color}

\usepackage{bibentry}
\nobibliography*

\theoremstyle{plain}

% vertical bars
\newcommand*{\vertbar}{\rule[-1ex]{0.5pt}{2.5ex}}

\renewcommand{\hat}{\widehat}
\renewcommand{\tilde}{\widetilde}

%% d in integrand
\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}}
\newcommand{\diag}[1]{{\rm diag}\LRp{#1}}
\newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}{\rm #2}}}
\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRb}[1]{\left| #1 \right|}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}
\newcommand{\LRceil}[1]{\left\lceil #1 \right\rceil}
\newcommand{\LRu}[1]{\left. #1 \right|}
\newcommand{\pdn}[3]{\frac{\partial^{#3}#1}{\partial#2^{#3}}}
\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\jump}[1] {\ensuremath{\llbracket#1\rrbracket}}
\newcommand{\avg}[1] {\ensuremath{\LRc{\!\{#1\}\!}}}

\graphicspath{{./figs/}}

\renewcommand{\note}[1]{\textcolor{red}{{#1}}}
%\renewcommand{\note}[1]{{\color{blue}{#1}}}

% removes nav symbols
\beamertemplatenavigationsymbolsempty
%\setbeamertemplate{caption}{\raggedright\insertcaption\par}

% defines newblock as null, giving compile issues otherwise
\let\newblock\relax

\title[Entropy stable ROMs]{	
Entropy stable reduced order modeling of nonlinear conservation laws}
\date[7/17/19]{ICIAM 2019, Valencia, Spain\\ \note{Reduced-order modeling} and data-driven estimation in waves and \note{fluids} \\July 17, 2019}
\author[J.\ Chan]{Jesse Chan}
\institute[Rice CAAM]{\inst{1}Department of Computational and Applied Mathematics}

\begin{document}
\makeatletter
\@addtoreset{subfigure}{framenumber}% subfigure counter resets every frame
\makeatother

\begin{frame}
\maketitle
\end{frame}

%% =================================================

%\frame{
%\frametitle{Discretely entropy stable schemes}
%
%Goal: address the stability and robustness of high order methods for nonlinear conservation laws.
%\begin{columns}
%\begin{column}{.5\textwidth}
%\begin{figure}
%%\vspace{-1em}
%\centering
%%\only<1>{\includegraphics[width=.95\textwidth]{figs/ndgFilter.pdf}}
%%\only<2>{\includegraphics[width=.95\textwidth]{figs/ndgFilter2.pdf}}
%%\hspace{1em}
%\raisebox{.5em}{\includegraphics[width=\textwidth]{figs/balancing.pdf}}
%\caption*{\tiny Image adapted from ``Man On Wire'' (2008)}
%%\visible<5>{\includegraphics[width=.475\textwidth]{figs/ductTape.png}}
%\end{figure}
%\end{column}
%\begin{column}{.475\textwidth}
%\begin{itemize}
%\item<1-> Aim for stability {independently} of artificial viscosity, limiters, discretization errors.
%\vspace{.75em}
%\item<1-> \textit{\note{Mechanical}} approach to stability based on algebraic properties of a discretization. 
%\vspace{.75em}
%\item<1-> Semi-discrete stability property for high order DG methods.  
%\end{itemize}
%\end{column}
%\end{columns}
%
%\let\thefootnote\relax\footnotetext{\tiny Finite volume methods: Tadmor, Chandrashekar, Ray, Svard, Fjordholm, Mishra, LeFloch, Rohde, \ldots}
%\let\thefootnote\relax\footnotetext{\tiny High order tensor product elements: Fisher, Carpenter, Gassner, Winters, Kopriva, Hindenlang, Persson, Pazner, \ldots}
%\let\thefootnote\relax\footnotetext{\tiny High order general elements: Chen and Shu, Crean, Hicken, DCDR Fernandez, Zingg, \ldots}
%}

\frame{
\frametitle{Projection-based reduced order models (ROMs)}
\vspace{-.5em}
\begin{figure}
\centering
\includegraphics[width=.65\textwidth]{figs/POD_illustration.png}
\end{figure}
\vspace{-.5em}
\begin{itemize}
\item Goal: reduce cost of many-query scenarios (design space exploration).
\item \note{Offline} compute reduced basis from snapshots of full order model (FOM), cheap \note{online} phase with reduced order model (ROM).
\end{itemize}

\uncover<2>{
\begin{center}
\minibox[frame]{Challenge: ROMs inherit FOM stability for elliptic\\
PDEs, but not for nonlinear conservation laws!}
\end{center}
}

\let\thefootnote\relax\footnotetext{\tiny Figure adapted from Brunton, Proctor, Kutz (2016), \textit{Discovering governing equations from data \ldots}.}
}

\frame{
\frametitle{Discrete instability for nonlinear PDEs}
\setcounter{subfigure}{0}
\vspace{-1em}
\begin{figure}
\begin{overprint}
\centering
\foreach \id in {1,2,3,4}{%
\only<\id>{
\captionsetup[subfloat]{width=.45\textwidth, justification=centering}
\subfloat[Exact solution]{%[$N = 7, K = 8$ (aligned mesh)]{
\makebox[.425\textwidth]{\includegraphics[width=.4\textwidth]{figs/burgersStable_\id.png}}}%
\hspace{1em}%
\subfloat[High order instability]{ %[$N = 7, K = 9$ (non-aligned mesh)]{
\makebox[.425\textwidth]{\includegraphics[width=.4\textwidth]{figs/burgersUnstable_\id.png}}}
} % only
} % foreach 
\end{overprint}
\end{figure}
\vspace{-.5em}
\begin{itemize}
\item For a larger number of modes, ROMs can blow up for under-resolved solutions of \note{nonlinear conservation laws} (shocks, turbulence).  
\vspace{.25em}
\item Instability tied to loss of the \note{chain rule} and inexact integration.  
\end{itemize}

\let\thefootnote\relax\footnotetext{\tiny Bui-Thanh, Willcox, et al.\ (2007). \textit{Goal-oriented, model-constrained optimization for reduction of large-scale systems}.}
}

\frame{
\frametitle{Entropy stability for nonlinear problems}
\vspace{-.5em}
\begin{itemize}
\item Generalizes energy stability to \note{nonlinear} systems of conservation laws (Burgers', shallow water, compressible Euler, MHD).  
\[
\pd{\bm{u}}{t} + \pd{\bm{f}(\bm{u})}{x} = 0.  
\]
\item Continuous entropy inequality: given a convex \note{entropy} function $S(\bm{u})$, entropy variables $\bm{v} = \bm{v}(\bm{u})$, and ``entropy potential'' $\psi(\bm{u})$, 
\begin{align*}
&\int_{\Omega} \bm{v}^T\LRp{\pd{\bm{u}}{t} + \pd{\bm{f}(\bm{u})}{x}} = 0, \qquad \boxed{\bm{v}(\bm{u}) = \LRu{\pd{S}{\bm{u}}}_{\bm{u}}} \\
&\Longrightarrow \int_{\Omega}\pd{S(\bm{u})}{t} + \LRu{\LRp{\bm{v}^T\bm{f}(\bm{u}) - \psi(\bm{u})}}_{-1}^1 \leq 0.
\end{align*}
\vspace{.0em}
\item Proof of entropy inequality relies on \note{chain rule}, integration by parts.  
\end{itemize}
}

\frame{
\frametitle{How is instability addressed in practice?}

\begin{overlayarea}{\textwidth}{.85\textheight}
\begin{itemize}
\item<1-> \textcolor{red}{Asymptotic} stability for \textcolor{red}{smooth} solutions (not shocks or turbulence!)
\item<3-> Common fix: \note{stabilize by regularizing} (filtering, added dissipation).  
\end{itemize}
\begin{figure}
\centering
\only<1>{
\vspace{.5em}
\includegraphics[width=1\textwidth]{figs/gassner_turb_01.pdf}
\caption*{Under-resolved solutions: turbulence (inviscid Taylor-Green vortex).}}
\only<2>{
\vspace{-.5em}
\includegraphics[width=.35\textwidth]{figs/shadowgraph.jpg}
\hspace{1em}
\includegraphics[width=.5\textwidth]{figs/trifoil.png}
\caption*{Under-resolved solutions: shock waves.}
}
\only<3->{
\vspace{-1em}
\includegraphics[width=.38\textwidth]{figs/ndgFilter.pdf}
\hspace{1em}
\visible<4>{\raisebox{2.5em}{\includegraphics[width=.55\textwidth]{figs/balancing.png}}}
%\visible<5>{\includegraphics[width=.475\textwidth]{figs/ductTape.png}}
}
\end{figure}
\end{overlayarea}
\let\thefootnote\relax\footnotetext{\tiny Figures from Beck and Gassner (2012), T.\ Warburton, \href{http://cirpwiki.info/wiki/CMS-Flow_Numerical_Methods}{\note{Coastal Inlets Research Program (CIRP)}}, ``Man on Wire'' (2008).}
}

\frame{
\frametitle{Scope of this talk, related work}

\begin{itemize}
\item Semi-discrete entropy inequality + retaining ROM accuracy.  
\vspace{1em}
\item Combinations of ideas from:
\vspace{.5em}
\begin{itemize}
\item<1-> entropy stable finite volume schemes,
\vspace{.5em}
\item<2-> high order entropy stable summation-by-parts (SBP) finite differences and nodal discontinuous Galerkin methods,
\vspace{.5em}
\item<3-> reduced order modeling and hyper-reduction.
\end{itemize}
\vspace{1em}
\item<4-> Does not address \textit{approximation issues} for transport problems!  
\end{itemize}

\let\thefootnote\relax\footnotetext{\tiny Finite volumes: Tadmor, Chandrashekar, Ray, Svard, Fjordholm, Mishra, LeFloch, Rohde, \ldots}
\let\thefootnote\relax\footnotetext{\tiny SBP: Fisher, Carpenter, Gassner, Winters, Kopriva, Hindenlang, Chen and Shu, Crean and Hicken, DCDR Fernandez, Zingg, Yamaleev, Parsani, Svard, Pazner and Persson, Wintermeyer, Bohm, Friedrich, Schnuke, \ldots.}
\let\thefootnote\relax\footnotetext{\tiny ROMs: Afkham, Ripamonti, Wang, Hesthaven (2018), Serre, Lafon, Gloerfelt, Bailly (2012), Kalashnikova, Barone, et al.\ (2014), Carlberg, Choi, Sargsyan (2018), Farhat, Chapman, Avery (2015), \ldots. }
}

\section{Starting point: an entropy stable full order model}

\frame[noframenumbering]{
\frametitle{Talk outline}

\only<1>{\tableofcontents}
\only<2>{\tableofcontents[currentsection,currentsubsection]}
}

\frame{
\frametitle{Full order model: entropy stable finite volumes}
\vspace{-.75em}
\begin{figure}
\centering
\includegraphics[width=.65\textwidth]{figs/fv_illustration.png}
\end{figure}
\vspace{-1.25em}
\begin{itemize}
\item Discretize integrated form of nonlinear conservation law
\[
\Delta x \td{\bm{u}_i}{t} + {\bm{f}_{S}(\bm{u}_{i+1},\bm{u}_i) - \bm{f}_{S}(\bm{u}_{i},\bm{u}_{i-1})} = 0.
\]
\item Matrix formulation:  let $\bm{F}_{ij} = \bm{f}_S\LRp{\bm{u}_i,\bm{u}_j}$, assume periodicity,
\[
\Delta x\td{\bm{u}}{t} + \LRp{\bm{Q}\circ \bm{F}}\bm{1}  = 0, \qquad \bm{Q} = \begin{bmatrix}
0 & 1 & & -1\\
-1 & 0 & 1 &\\
& \ddots & \ddots & 1 \\
1 & & -1  & 0
\end{bmatrix}
\]
\end{itemize}
}

\frame{
\frametitle{Recovering a semi-discrete entropy balance}
\vspace{-.5em}
\begin{itemize}
\item First, show that entropy is \textit{conserved} for the inviscid case.  Test with vector of entropy variables $\bm{v}$ 
\[
\bm{v}^T\LRp{\Delta x\td{\bm{u}}{t} + \LRp{\bm{Q}\circ \bm{F}}\bm{1}} = \bm{0}. 
\]
\item Use chain rule in time and definition of entropy variables
\[
\bm{v}^T\td{\bm{u}}{t} = \td{S(\bm{u})}{\bm{u}}^T \td{\bm{u}}{t} = \bm{1}^T \td{S(\bm{u})}{t}.
\]
\item If \note{$\bm{v}^T \LRp{\bm{Q}\circ \bm{F}}\bm{1} = 0$}, then the average entropy is conserved
\[
\Delta x \LRp{\bm{1}^T\td{S(\bm{u})}{t}} = 0.
\]
%\item Assume 
%\[
%\bm{v}^T\td{\bm{u}}{t} = \pd{S(\bm{u})}{\bm{u}}^T \td{\bm{u}}{t} = \pd{S(\bm{u})}{t} .
%\]
%\vspace{-.5em}
\end{itemize}

}

\frame{
\frametitle{Entropy conservation for the full order model}

\begin{itemize}
%\item Assume $\epsilon = 0$.  
\item Define $\bm{F}_{ij}$ using Tadmor's entropy conservative numerical flux
\begin{gather*}
\bm{f}_S(\bm{u},\bm{u}) = \bm{f}(\bm{u}), \qquad \text{(consistency)}\\
\bm{f}_S(\bm{u},\bm{v}) = \bm{f}_S(\bm{v},\bm{u}), \qquad \text{(symmetry)} \\
%,\qquad \text{(symmetry)} \\
\LRp{\bm{v}_L - \bm{v}_R}^T \bm{f}_S\LRp{\bm{u}_L,\bm{u}_R} = \psi_L - \psi_R, \qquad \text{(conservation)}.
\end{gather*}
\item Use \note{$\bm{Q} = -\bm{Q}^T$} and \note{$\bm{Q}\bm{1} = \bm{0}$} to show $\bm{v}^T\LRp{\bm{Q}\circ \bm{F}}\bm{1} = 0$.
%\[
%\bm{v}^T\LRp{\bm{Q}\circ \bm{F}}\bm{1} = 
%\only<1>{\sum_{ij} {\bm{Q}_{ij}} \bm{v}_i^T \bm{f}_S\LRp{\bm{u}_i,\bm{u}_j}}
%\only<2>{ \frac{1}{2}\sum_{ij} \note{\LRp{\bm{Q}_{ij} - \bm{Q}_{ji}}} \bm{v}_i^T \bm{f}_S\LRp{\bm{u}_i,\bm{u}_j}}
%\only<3>{ \frac{1}{2}\sum_{ij} \bm{Q}_{ij} \textcolor{red}{\LRp{\bm{v}_i - \bm{v}_j}^T \bm{f}_S\LRp{\bm{u}_i,\bm{u}_j}}}
%\only<4>{ \frac{1}{2}\sum_{ij} \bm{Q}_{ij} \LRp{\psi(\bm{u}_i) - \psi(\bm{u}_j)}}
%\only<5>{ \frac{1}{2}\LRp{\bm{\psi}^T\bm{Q}\bm{1} - \bm{1}^T\bm{Q}\bm{\psi}}}
%\only<6>{ \bm{0} \qquad \Longrightarrow \qquad \boxed{\td{S(\bm{u})}{t} = 0.}}
%\]
\begin{overlayarea}{\textwidth}{.425\textheight}
\vspace{-1em}
\begin{align*}
\bm{v}^T\LRp{\bm{Q}\circ \bm{F}}\bm{1} &= \sum_{ij} {\bm{Q}_{ij}} \bm{v}_i^T \bm{f}_S\LRp{\bm{u}_i,\bm{u}_j}\\
&= \only<1>{\frac{1}{2}\sum_{ij} \note{\LRp{\bm{Q}_{ij} - \bm{Q}_{ji}}} \bm{v}_i^T \bm{f}_S\LRp{\bm{u}_i,\bm{u}_j}}
\only<2->{\frac{1}{2}\sum_{ij} \bm{Q}_{ij} \textcolor{red}{\LRp{\bm{v}_i - \bm{v}_j}^T \bm{f}_S\LRp{\bm{u}_i,\bm{u}_j}}\\}
\only<3>{&=  \frac{1}{2}\sum_{ij} \bm{Q}_{ij} \LRp{\psi(\bm{u}_i) - \psi(\bm{u}_j)}}
\only<4>{&=  \frac{1}{2}\LRp{\bm{\psi}^T\bm{Q}\bm{1} - \bm{1}^T\bm{Q}\bm{\psi}} = \bm{0}}
%\only<5>{ \bm{0} \qquad \Longrightarrow \qquad \boxed{\td{S(\bm{u})}{t} = 0.}}
\end{align*}
\end{overlayarea}
\end{itemize}
\let\thefootnote\relax\footnotetext{\tiny Tadmor, Eitan (1987). \textit{The numerical viscosity of entropy stable schemes for systems of conservation laws. I.}}
}

\frame{
\frametitle{Entropy stability for the full order model}
\vspace{-.5em}
\begin{itemize}
\item Add appropriate \note{entropy dissipation} terms 
\begin{gather*}
\bm{v}^T\LRp{\Delta x\td{\bm{u}}{t} + \LRp{\bm{Q}\circ \bm{F}}\bm{1}  - \LRp{\text{viscous terms}}} = 0\\
\Longrightarrow \quad \Delta x \bm{1}^T \td{\bm{S}}{t} = \bm{v}^T \LRp{\text{viscous terms}} \overset{?}{\leq} 0.
\end{gather*}
\vspace{.0em}
\item Entropy is convex $\Longrightarrow$ Hessian inverse $\pd{\bm{u}}{\bm{v}}$ is SPD
\[
-\epsilon\pdn{\bm{u}}{x}{2} = -\epsilon\pd{}{x} \LRp{\pd{\bm{u}}{x}} = -\epsilon\pd{}{x}\LRp{{\pd{\bm{u}}{\bm{v}}}   \pd{\bm{v}}{x}}.
\]
Standard FEM/ROM treatment of diffusion tensor dissipates entropy.
\end{itemize}

\let\thefootnote\relax\footnotetext{\tiny Hughes, Mallet, Franca (1986). \textit{A new FE formulation for CFD: I. Sym.\ forms of the comp.\ Euler and NS equations \ldots}}
%\let\thefootnote\relax\footnotetext{\tiny Carpenter et al.\ (2014). \textit{Entropy stable spectral collocation schemes for the NS equations: discontinuous interfaces.}}
%\let\thefootnote\relax\footnotetext{\tiny Riviere and Wheeler (2000).  \textit{A discontinuous Galerkin method applied to nonlinear parabolic equations}}
}

\section{An expensive but entropy stable reduced order model}

\frame[noframenumbering]{
\frametitle{Talk outline}
\tableofcontents[currentsection,currentsubsection]
}

\frame{
\frametitle{Naive POD-Galerkin procedure}
\vspace{.5em}
\begin{itemize}
\item (Offline) compute POD basis from solution component snapshots.
\[
\bm{V} = {\begin{bmatrix} \vertbar & & \vertbar \\
\phi_1 & \ldots & \phi_N\\
\vertbar & & \vertbar
\end{bmatrix}}, \qquad \bm{u} \approx \bm{V} \bm{u}_N.
\]
\vspace{.0em}
\item (Online) Galerkin projection for coefficients $\bm{u}_N$
\[
\only<1>{{\bm{V}^T} \LRp{\Delta x\td{\LRp{\bm{V}\bm{u}_N}}{t} + \LRp{{\bm{Q}}\circ \bm{F}}\bm{1}} = 0.}
\only<2->{\Delta x{\bm{V}^T\bm{V}} \td{\bm{u}_N}{t} + \bm{V}^T \LRp{{\bm{Q}}\circ \bm{F}}\bm{1} = 0.}
%\only<2>{\underbrace{\bm{V}^T\bm{V}}_{\text{orthonormal}} \td{\bm{u}_N}{t} + \bm{V}^T \LRp{{\bm{Q}}\circ \bm{F}}\bm{1} = 0.}
%\only<3->{\td{\bm{u}_N}{t} + \bm{V}^T \LRp{{\bm{Q}}\circ \bm{F}}\bm{1} = 0.}
\]
\vspace{.0em}
\item<3-> Not discretely entropy stable!  Conservative variables are in the reduced space, entropy variables are not -- cannot test with $\bm{v}(\bm{V}\bm{u}_N)$.
\begin{gather*}
\bm{u} \in {\rm span}\LRc{\phi_1,\ldots,\phi_N}, \qquad \bm{v}(\bm{u}) \not\in {\rm span}\LRc{\phi_1,\ldots,\phi_N}.
\end{gather*}
\end{itemize}
}

\frame{
\frametitle{Entropy projection and discrete entropy stability }

\begin{itemize}
\item Entropy variables are not in ROM \textit{test space}.   Solution: test with the projection (pseudoinverse) of the entropy variables 
\begin{align*}
\LRp{\bm{V}^\dagger \bm{v}\LRp{\bm{V}\bm{u}_N}}^T\LRp{\Delta x  \bm{V}^T\bm{V}\td{\bm{u}_N}{t} + \bm{V}^T \LRp{{\bm{Q}}\circ \bm{F}}\bm{1}} = 0 
\end{align*}
\vspace{0em}
\item Projected entropy variables recover time derivative of discrete entropy 
\begin{align*}
\LRp{\bm{V}^\dagger \bm{v}\LRp{\bm{V}\bm{u}_N}}^T \bm{V}^T\bm{V}\td{\bm{u}_N}{t} &= \bm{v}\LRp{\bm{V}\bm{u}_N}^T \note{\underbrace{\LRp{\bm{V}^T\bm{V}\bm{V}^{\dagger}}^T}_{\bm{V}}} \td{\bm{u}_N}{t} \\
&= \LRu{\pd{S}{\bm{u}}}_{\LRp{\bm{V}\bm{u}_N}}^T \td{\LRp{\note{\bm{V}} \bm{u}_N}}{t} = \bm{1}^T\td{S\LRp{\bm{V}\bm{u}_N}}{t}
\end{align*}
%\vspace{0em}
%\item Entropy stability: evaluate $\bm{Q}\circ\bm{F}$ using projected entropy variables, enrich snapshots with entropy variables to ensure accuracy.
%\begin{gather*}
%\tilde{\bm{u}} = \bm{u}\LRp{\bm{V} \bm{V}^\dagger \bm{v}\LRp{\bm{V} \bm{u}_N}}, \qquad \LRp{\bm{F}}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}
%\end{gather*}
\end{itemize}
}

\frame{
\begin{itemize}
\item Projected entropy variables are no longer mappings of the conservative variables.  Denote $\tilde{\bm{v}} = \bm{V}\bm{V}^\dagger \bm{v}\LRp{\bm{V}\bm{u}_N}$, then
\begin{align*}
\tilde{\bm{v}}^T \LRp{\bm{Q}\circ \bm{F}}\bm{1} &= \frac{1}{2}\sum_{ij}
\bm{Q}_{ij}\note{\LRp{\tilde{\bm{v}}_i - \tilde{\bm{v}}_j}^T\bm{f}_S\LRp{{\bm{u}}_i,{\bm{u}}_j}} \\
&\neq \frac{1}{2}\sum_{ij} \bm{Q}_{ij} \LRp{\psi(\bm{u}_i) - \psi(\bm{u}_j)}.
\end{align*}
\item Fix: evaluate $\bm{Q}\circ\bm{F}$ in terms of projected entropy variables, enrich snapshots with entropy variables to ensure accuracy.
\begin{gather*}
\tilde{\bm{u}} = \bm{u}\LRp{\bm{V} \bm{V}^\dagger \bm{v}\LRp{\bm{V} \bm{u}_N}}.%, \qquad \LRp{\bm{F}}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}
\end{gather*}
\item Modified Galerkin projection ROM is discretely entropy stable  
\[
\Delta x {\bm{V}^T\bm{V}} \td{\bm{u}_N}{t} + \bm{V}^T \LRp{{\bm{Q}}\circ \bm{F}}\bm{1} = 0, \qquad \LRp{\bm{F}}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}
\]
\end{itemize}
}

\section{Reducing nonlinear costs: entropy stable hyper-reduction}

\frame[noframenumbering]{
\frametitle{Talk outline}
\tableofcontents[currentsection,currentsubsection]
}

\frame{
\frametitle{Evaluating nonlinear ROM terms dominates costs}
\vspace{-.75em}
\begin{center}
Problem: cost of evaluating ROM nonlinear terms scales with\\
\note{number of grid points (squared)}!  No improvement over full order model.
\[
{\tilde{\bm{u}} = \note{\bm{u}}\LRp{\bm{V} \bm{V}^\dagger \note{\bm{v}}\LRp{\bm{V} \bm{u}_N}}}, \qquad \LRp{{\bm{Q}}\circ \note{\bm{F}}}\bm{1} 
\]
\end{center}
\vspace{-1em}
\uncover<2>{
\begin{columns}
\begin{column}{.55\textwidth}
\begin{itemize}
\item \note{Hyper-reduction} reduces costs by approximating nonlinear evaluations.  
\begin{align*}
&\bm{V}^T\bm{g}(\bm{V}\bm{u}_N) \approx\\
&\underbrace{\bm{V}(\mathcal{I},:)^T}_{\text{sampled rows}}\bm{W}\bm{g}(\bm{V}(\mathcal{I},:)\bm{u}_N)
\end{align*}
%\vspace{1.5em}
\item Examples: gappy POD, DEIM, empirical cubature, ECSW, \ldots
\end{itemize}
\end{column}
}
\begin{column}{.45\textwidth}
\begin{figure}
%\centering
\begin{overlayarea}{\textwidth}{.5\textheight}
\only<1>{\includegraphics[width = .8\textwidth]{figs/hyperreduc_illustration0.png}}
\only<2>{\includegraphics[width = .8\textwidth]{figs/hyperreduc_illustration.png}}
\end{overlayarea}
\end{figure}
\end{column}
\end{columns}

\let\thefootnote\relax\footnotetext{\tiny Hernandez et al., An et al., Bui/Willcox, Chantarantabut /Sorensen, Drmac/Gugercin, Farhat et al.\, Patera/Yano, \ldots}
}

\frame{
\frametitle{Entropy stability and standard hyper-reduction}

\begin{itemize}
\item How to hyper-reduce $\LRp{\bm{Q}\circ \bm{F}}$?  Construct a \textit{sampled} matrix $\bm{Q}_s$ from $\bm{Q}$.  Need \note{$\bm{Q}_s = -\bm{Q}_s^T$} and \note{$\bm{Q}_s\bm{1} = \bm{0}$} for entropy stability.
\vspace{2em}
\item Options: sub-sample rows and columns of full matrix $\bm{Q}$ or approximate $\bm{Q}$ by weighted sum of local skew matrices $\bm{Q}_e$.  
\[
\bm{Q} = \sum_{e=1}^K \bm{Q}_e \approx \bm{Q}_s = \sum_{e = 1}^K \bm{w}_e \bm{Q}_e, \qquad \bm{w} \text{ sparse}.
\]
\uncover<2->{Problems: either \note{$\bm{Q}_s$ loses skew-symmetry} or \note{$\bm{Q}_s\bm{1}\neq \bm{0}$}.}
%\vspace{1.5em}
%\item<5-> Project $\bm{Q}$ onto reduced basis $\bm{V}^T\bm{Q} \bm{V}$.
\end{itemize}

\let\thefootnote\relax\footnotetext{\tiny Patera and Yano (2017). \textit{An LP empirical quadrature procedure for parametrized functions}.}
}

\frame{
\frametitle{Two-step hyper-reduction: compress and project}
%\begin{overlayarea}{\textwidth}{\textheight}
%\vspace{1em}
\begin{itemize}
\item Step 1: construct a \textit{modal} $\bm{Q}$ using an expanded ``test'' basis $\bm{V}_t$
\[
\bm{V}_t^T\bm{Q}\bm{V}_t, \qquad \bm{V}_t = {\rm orth}\LRp{
\begin{bmatrix} 
\bm{V} & \bm{1}
\end{bmatrix}}
\]
%\vspace{1em}
\item Step 2: determine modal coefficients using hyper-reduced quadrature.  
%\item<2-> Compute \note{empirical cubature} points + projection for $\bm{Q}_s$
\begin{gather*}
\bm{W}_{ij} = w_{ii}\delta_{ij}, \\%\quad \text{ (diagonal weighting matrix)} \\
\bm{M}_t = \bm{V}_t(\mathcal{I},:)^T\bm{W}\bm{V}_t(\mathcal{I},:), \\
\bm{P}_t = \bm{M}_t^{-1}\bm{V}_t(\mathcal{I},:)^T\bm{W}.%, \qquad \text{ discrete projection}
%\bm{Q}_s = \bm{P}_t^T \LRp{\bm{V}_t^T\bm{Q}\bm{V}_t} \bm{P}_t. 
\end{gather*}
$\bm{P}_t$ is a weighted orthogonal projection onto the test basis.  %(can also use Gappy POD).
%\item<2-> Compression + sampling ensure $\bm{Q}_t\bm{1} = 0$ and $\bm{Q}_t = -\bm{Q}_t^T$ (on periodic domains) for entropy stability.
%\[
%\underbrace{\bm{V}(\mathcal{I},:)^T\bm{W}\bm{V}(\mathcal{I},:)}_{\bm{M}_N} \td{\bm{u}_N}{t} + \bm{V}(\mathcal{I},:)^T \LRp{\bm{Q}_t\circ \bm{F}_S}\bm{1} = 0.
%\]
\vspace{1em}
\item Step 3: let $\bm{Q}_s = \bm{P}_t^T\LRp{\bm{V}_t^T\bm{Q}\bm{V}_t}\bm{P}_t$.  Note $\bm{Q}_s = -\bm{Q}_s^T$ and $\bm{Q}_s\bm{1} = \bm{0}$!
\end{itemize}
%\end{overlayarea}

\let\thefootnote\relax\footnotetext{\tiny Hernandez, Caicedo, Ferrer (2017).  \textit{Dimensional hyper-reduction of nonlinear finite element models via empirical cubature.}}
%\let\thefootnote\relax\footnotetext{\tiny Everson, Sirovich (1995).  \textit{Karhunen--Loeve procedure for gappy data.}}
}

\frame{
\begin{itemize}
\item Problem: modes $\bm{V}_t$ may sample $\bm{Q}\bm{V}$ very poorly, e.g., $\bm{V}_t^T\bm{Q}\bm{V}_t\approx \bm{0}$!
\vspace{-1.5em}
\begin{figure}
\centering
\subfloat[\scriptsize Shock snapshots]{\includegraphics[width=.3\textwidth]{figs/burg_snaps.png}}
\hspace{.1em}
\subfloat[\scriptsize Modes (columns of $\bm{V}$)]{\includegraphics[width=.3\textwidth]{figs/burg3modes.png}}
\hspace{.1em}
\subfloat[\scriptsize Mode derivatives $\bm{Q}\bm{V}$]{\includegraphics[width=.3\textwidth]{figs/burg3modes_diff.png}}
\end{figure}
\vspace{.5em}
\item Fix: expand basis $\bm{V}_t$ with ``least squares'' test functions
\begin{gather*}
\bm{V}_t = {\rm orth}\LRp{
\begin{bmatrix} 
\bm{V} & \bm{1} & \bm{Q}\bm{V} 
\end{bmatrix}}, \qquad
\bm{V}_t^T\bm{Q}\bm{V}_t \in \mathbb{R}^{(2N+1) \times (2N+1)}. %\mathbb{R}^{(d+1)N+1 \times (d+1)N+1}.
\end{gather*}
$\bm{V}_t^T\bm{Q}\bm{V}_t$ is accurate, skew-symmetric, and exact for constants.
\end{itemize}

\let\thefootnote\relax\footnotetext{\tiny Carlberg, Barone, Antil (2017).  \textit{Galerkin v. least-squares Petrov--Galerkin projection in nonlinear model reduction.}}
}

%\frame{
%\frametitle{Hyper-reduction through empirical cubature}
%
%\begin{itemize}
%%Reduced operator $\tilde{\bm{Q}}_N$ is skew-symmetric and satisfies $\tilde{\bm{Q}}_N\bm{1} = 0$.
%%\item Greedy algorithm for  approx.\ of $\int \phi_i(\bm{x}) \phi_j(\bm{x})$.  
%
%\let\thefootnote\relax\footnotetext{\tiny An, Kim, James (2009).  \textit{Optimizing cubature for efficient integration of subspace deformations}.}
%\let\thefootnote\relax\footnotetext{\tiny Hernandez, Caicedo, Ferrer (2017).  \textit{Dimensional hyper-reduction of nonlinear finite element models via empirical cubature.}}
%}

\frame{
\frametitle{Summary: entropy stable ROMs on periodic domains}

%\note{Summarize here}

\begin{itemize}
\item Entropy stable ROM using \textit{projected} entropy variables.  
\vspace{1em}
\item Two-step hyper-reduction of $\LRp{\bm{Q}\circ \bm{F}}$: compress and project
\begin{itemize}
\item Compress $\bm{Q}$ onto expanded test basis spanning $\bm{1}$, $\bm{Q}\bm{V}$.
\item Project hyper-reduced point values to modes of test basis 
\[
\bm{Q}_s = \bm{P}_t^T\LRp{\bm{V}_t^T\bm{Q}\bm{V}_t}\bm{P}_t
\]
%Reduced $\bm{Q}_s$ satisfies $\bm{Q}_s\bm{1} = \bm{0}$ and $\bm{Q}_s = -\bm{Q}_s^T$.
\end{itemize}
%\vspace{1em}
\item Entropy stable reduced order model with hyper-reduction:
\[
{\bm{V}(\mathcal{I},:)^T\bm{W}\bm{V}(\mathcal{I},:)} \td{\bm{u}_N}{t} + \bm{V}(\mathcal{I},:)^T \LRp{\bm{Q}_s\circ \bm{F}}\bm{1} = 0,
\]
where $\bm{P}$ is the hyper-reduced projection onto POD modes and
\begin{gather*}
%\underbrace{\bm{V}(\mathcal{I},:)^T\bm{W}\bm{V}(\mathcal{I},:)}_{\bm{M}_N} \td{\bm{u}_N}{t} + \bm{V}(\mathcal{I},:)^T \LRp{\bm{Q}_s\circ \bm{F}_S}\bm{1} = 0.
%{\bm{V}(\mathcal{I},:)^T\bm{W}\bm{V}(\mathcal{I},:)} \td{\bm{u}_N}{t} + \bm{V}(\mathcal{I},:)^T \LRp{\bm{Q}_s\circ \bm{F}}\bm{1} = 0,\\
{\bm{F}}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \quad \tilde{\bm{u}} = \bm{u}\LRp{\bm{V}(\mathcal{I},:)\bm{P} \bm{v}\LRp{\bm{V}\bm{u}_N}},
%\LRp{\bm{V}(\mathcal{I},:)^T\bm{W}\bm{V}(\mathcal{I},:)}^{-1}{\bm{V}(\mathcal{I},:)^T\bm{W}}.
\end{gather*}
\end{itemize}
}

\frame{
\frametitle{Non-periodic boundary conditions}

\begin{itemize}
%\item Non-periodic domains: full order matrix $\bm{Q}$ is \textit{nearly} skew-symmetric
%\[
%\bm{Q} = \bm{S} + \bm{B}, \qquad \bm{S} = \begin{bmatrix} 
% 0& 1 & 0\\
%-1& 0 &1\\
%0& -1 & 0
%\end{bmatrix}, \qquad \bm{B} = \begin{bmatrix} -1 & &\\
%&0&\\
%&&1
%\end{bmatrix}
%\]
\item Treat boundary conditions weakly using ``hybridized'' summation-by-parts (SBP) operators and numerical flux.  
\vspace{1em}
\item Can add entropy-dissipative penalization terms (e.g., Lax-Friedrichs).
%\[
%\bm{Q}_N = \begin{bmatrix}
%\bm{Q}-\bm{Q}^T & \bm{E}^T\bm{B}\\
%-\bm{B}\bm{E} & 
%\end{bmatrix}
%\]
\vspace{1em}
\item Let $\bm{V}_f$ be the surface interpolation matrix.  Entropy stability requires a weak SBP property for surface quadrature weights $\bm{w}_f$ 
\begin{align*}
\bm{V}_t^T \bm{Q}_x^T\bm{1} &= \bm{V}_f^T \LRp{\bm{n}_x \circ \note{\bm{w}_f}},\\
\bm{V}_t^T \bm{Q}_y^T\bm{1} &= \bm{V}_f^T \LRp{\bm{n}_y \circ \note{\bm{w}_f}}.
\end{align*}
Enforce conditions using constrained LP hyper-reduction.
\end{itemize}

\let\thefootnote\relax\footnotetext{\tiny Patera and Yano (2017). \textit{An LP empirical quadrature procedure for parametrized functions}.}
\let\thefootnote\relax\footnotetext{\tiny Chan (2018). \textit{On discretely entropy conservative and entropy stable discontinuous Galerkin methods.}}
\let\thefootnote\relax\footnotetext{\tiny Chan (2019). \textit{Skew-symmetric entropy stable modal discontinuous Galerkin formulations}.}
}

\section{Numerical experiments}

\frame[noframenumbering]{
\frametitle{Talk outline}
\tableofcontents[currentsection,currentsubsection]
}

\frame{
\frametitle{Numerical results: 1D Euler with reflective BCs + shock}
\setcounter{subfigure}{0}
\vspace{-1em}
\begin{figure}
\centering 
\only<1>{
\subfloat[25 modes, 100 ECP, $T=.25$]{\includegraphics[width=.45\textwidth]{figs/euler1Dwall_t25_25modes.png}}
\hspace{.5em}
\subfloat[25 modes, 100 ECP, $T=.75$]{\includegraphics[width=.45\textwidth]{figs/euler1Dwall_t75_25modes.png}}
\caption*{\footnotesize FOM with $2500$ grid points, viscosity coefficient $\epsilon = 2e-4$, ROM with 25 modes.}
}
\only<2>{
\subfloat[75 modes, 205 ECP, $T=.25$]{\includegraphics[width=.45\textwidth]{figs/euler1Dwall_t25_75modes.png}}
\hspace{.5em}
\subfloat[75 modes, 205 ECP, $T=.75$]{\includegraphics[width=.45\textwidth]{figs/euler1Dwall_t75_75modes.png}}
\caption*{\footnotesize FOM with $2500$ grid points, viscosity coefficient $\epsilon = 2e-4$, ROM with 75 modes.}
}
\only<3>{
\subfloat[125 modes, 304 ECP, $T=.25$]{\includegraphics[width=.45\textwidth]{figs/euler1Dwall_t25_125modes.png}}
\hspace{.5em}
\subfloat[125 modes, 304 ECP, $T=.75$]{\includegraphics[width=.45\textwidth]{figs/euler1Dwall_t75_125modes.png}}
\caption*{\footnotesize FOM with $2500$ grid points, viscosity coefficient $\epsilon = 2e-4$, ROM with 125 modes.}
}
\end{figure}

\begin{center}
All results use explicit RK-45 time-stepping with FOM time-step.
\end{center}
}

\frame{
\frametitle{Evolution of average entropy}

\begin{figure}
\centering 
\only<1>{
\includegraphics[width=.7\textwidth]{figs/euler1Dwall_entropy_25modes.png}
\caption*{Average entropy over time (25 modes).}
}
\only<2>{
\includegraphics[width=.7\textwidth]{figs/euler1Dwall_entropy_75modes.png}
\caption*{Average entropy over time (75 modes).}
}
\only<3>{
\includegraphics[width=.7\textwidth]{figs/euler1Dwall_entropy_125modes.png}
\caption*{Average entropy over time (125 modes).}
}
\end{figure}
}

\frame{
\frametitle{Error with and without hyper-reduction}
\vspace{-.5em}
\begin{figure}
\centering
\includegraphics[width=.75\textwidth]{figs/euler_error_hypred_comparison_K2500.png}
\caption{Error over time for a $K=2500$ FOM and ROM with $25, 75, 125$ modes.}
\end{figure}
}

\frame{
\frametitle{Numerical results: smoothed Kelvin-Helmholtz instability}
\vspace{-.5em}
\begin{figure}
\centering 
\only<1>{
\subfloat[Density, full order model]{\includegraphics[width=.45\textwidth]{figs/khfom.png}}
\hspace{.5em}
\subfloat[Reduced order model]{\includegraphics[width=.45\textwidth]{figs/khrom_75.png}}
}
\only<2>{
\subfloat[Density, full order model]{\includegraphics[width=.45\textwidth]{figs/khfom.png}}
\hspace{.5em}
\subfloat[ROM w/reduced quad.\ points]{\includegraphics[width=.45\textwidth]{figs/khrom_75_ecp865.png}}
}
\caption{Full order model with $40000$ points, viscosity $\epsilon = .5\Delta x$.  ROM with $75$ modes, $865$ reduced quadrature points, $1.11 \%$ relative $L^2$ error at $T = 2.5$.}
\end{figure}
}

\frame{
\frametitle{Numerical results: Gaussian pulse with reflective wall}

\begin{figure}
\centering 
\only<1>{
\subfloat[Density, full order model]{\includegraphics[width=.45\textwidth]{figs/pulse2d.png}}
\hspace{.5em}
\subfloat[Reduced order model]{\includegraphics[width=.45\textwidth]{figs/pulse2d_ROM.png}}
}
\only<2>{
\subfloat[Density, full order model]{\includegraphics[width=.45\textwidth]{figs/pulse2d.png}}
\hspace{.5em}
\subfloat[ROM w/reduced quad.\ points]{\includegraphics[width=.45\textwidth]{figs/pulse2d_ROM_pts.png}}
}
\caption{FOM with $10000$ points, viscosity $\epsilon = .5\Delta x$.  ROM with $25$ modes, $306$ reduced volume points, 82 reduced surface points, $.57 \%$ error at $T = .25$.}
\end{figure}

}
%% =================================================

\frame{
\frametitle{Summary and future work}

\begin{itemize}
\item Entropy stable reduced order models reproduces a semi-discrete evolution of entropy from the full order model.  
\item Open challenges: volume and surface hyper-reduction strategies, CFL and error analysis, approximation of convective solutions.
%\vspace{.25em}
%\item Current work: ROM, positivity preservation.  
\vspace{.25em}
%\item Current work: hybrid and non-conforming meshes, multi-GPU.
%\vspace{.25em}
\item This work is supported by DMS-1719818 and DMS-1712639. 
\end{itemize}
\vspace{.25em}
\begin{center}
Thank you!  Questions?
\vspace{.25em}

{\includegraphics[width=.15\textwidth]{figs/nsf.jpg}}
\end{center}

\let\thefootnote\relax\footnotetext{\tiny Chan (In preparation). \textit{Entropy stable reduced order modeling of nonlinear conservation laws}.}
\let\thefootnote\relax\footnotetext{\tiny Chan (2019). \textit{Skew-symmetric entropy stable modal discontinuous Galerkin formulations}.}
\let\thefootnote\relax\footnotetext{\tiny Chan (2018). \textit{On discretely entropy conservative and entropy stable discontinuous Galerkin methods.}}
}

%% =================================================

\frame[noframenumbering]{
\frametitle{Additional slides}
}

\frame[noframenumbering]{
\frametitle{Example of entropy conservative finite volume fluxes}

\begin{itemize}
\item Entropy conservative flux for Burgers' equation 
\[
f_S(u_L,u_R) = \frac{1}{6}\LRp{u_L^2 + u_Lu_R + u_R^2}.
\]
\item Flux differencing: let $u_L = u(x), u_R = u(y)$
\begin{align*}
\pd{{f}({u})}{x} &\Longrightarrow \note{\LRu{2\pd{f_S\LRp{u(x),u(y)}}{x}}_{y=x}}
\end{align*}
\item Recovering the Burgers' split formulation
\begin{align*}
f_S(u(x),u(y)) &= \frac{1}{6}\LRp{u(x)^2 + u(x)u(y) + u(y)^2}\\
\LRu{2\pd{f_S\LRp{u(x),u(y)}}{x}}_{y=x} &= \frac{1}{3}\pd{u^2}{x} + \frac{1}{3}u\pd{u}{x} + \frac{1}{3}u^2\cancel{\pd{1}{x}}.
\end{align*}
\end{itemize}
}

\frame[noframenumbering]{
\frametitle{Entropy conservative fluxes for compressible Euler}
\begin{itemize}
\item In one dimension:
\begin{align*}
f^1_S(\bm{u}_L,\bm{u}_R) &= \avg{\rho}^{\log} \avg{u}\\
f^2_S(\bm{u}_L,\bm{u}_R) &= \frac{\avg{\rho}}{2\avg{\beta}} + \avg{u}f^1_S\\
f^3_S(\bm{u}_L,\bm{u}_R) &= f^1_S\LRp{\frac{1}{2(\gamma-1)\avg{\beta}^{\log}} - \frac{1}{2}\avg{u^2}} + \avg{u}f^2_S,
\end{align*}
\vspace{.5em}
\item Involves logarithmic mean and ``inverse temperature'' $\beta$
\[
\avg{u}^{\log} = \frac{u_L - u_R}{\log{u_L}- \log{u_R}}, \qquad \beta = \frac{\rho}{2p}.
\]
\end{itemize}

\let\thefootnote\relax\footnotetext{\tiny Chandreshekar (2013),  \emph{Kinetic energy preserving and entropy stable FV schemes for comp.\ Euler and NS equations.}}
}

\frame[noframenumbering]{
\frametitle{Determining empirical cubature points (ECP) and weights}

\begin{itemize}
\item Goal: integrate a target basis to some accuracy.
\vspace{1em}
\item Offline step: greedy selection of hyper-reduction points
\begin{itemize}
\item Project residual onto remaining rows of basis matrix.
\item Find ``most positive'' point of projected residual.
\item Solve (nonlinear) least squares for (positive) quadrature weights.
\end{itemize}
\vspace{1em}
\item Target basis: products of modes
\[
{\rm span}\LRc{\phi_i(\bm{x})\phi_j(\bm{x}), \quad 1\leq i,j \leq N}.
\]
Reduce costs by substituting leading POD modes.  
%\vspace{1em}
%\item Number of EC points depends on FOM viscosity, smoothness.  
\end{itemize}

\let\thefootnote\relax\footnotetext{\tiny Hernandez, Caicedo, Ferrer (2017).  \textit{Dimensional hyper-reduction of nonlinear finite element models via empirical cubature.}}
}

\end{document}

### 2019-09-06 matlab/Codes1D/clenshaw_curtis.m

@@ -1,0 +1,21 @@
+function [r w] = clenshaw_curtis(N,kind)
+
+if nargin==1
+    kind=1;
+end
+
+j = (0:N)';
+if kind==1
+    r = cos((2*j+1)*pi/(2*(N+1))); % no boundary points
+elseif kind == 2
+    r = cos(pi*j/N); % boundary points    
+else
+    r = cos((2*j)*pi/(2*(N+1))); % nonsymmetric version
+end
+r = sort(r);
+
+% compute weights in a dumb way
+V = Vandermonde1D(N,r);
+[rq wq] = JacobiGQ(0,0,N);
+Vq = Vandermonde1D(N,rq)/V;
+w = Vq'*wq;

### 2019-09-12 matlab/cfd3D/check_noncon_gcl.m
updating from macbook 11

@@ -10,44 +10,43 @@ r = r(:); s = s(:); t = t(:);
 [wr ws wt] = meshgrid(w1D);
 w = wr(:).*ws(:).*wt(:);
 
-% quadrature nodes for collocation
-[rq1D wq1D] = JacobiGQ(0,0,N);
-
-[rq sq tq] = meshgrid(rq1D);
-rq = rq(:); sq = sq(:); tq = tq(:);
-[wr ws wt] = meshgrid(wq1D);
-wq = wr(:).*ws(:).*wt(:);
+% % quadrature nodes for collocation
+% [rq1D wq1D] = JacobiGQ(0,0,N);
+% [rq sq tq] = meshgrid(rq1D);
+% rq = rq(:); sq = sq(:); tq = tq(:);
+% [wr ws wt] = meshgrid(wq1D);
+% wq = wr(:).*ws(:).*wt(:);
 
 % 1D matrices
 V1D = Vandermonde1D(N,r1D);
 D1D = GradVandermonde1D(N,r1D)/V1D;
 I = eye(length(r1D));
-Vq1D = Vandermonde1D(N,rq1D)/V1D;
-M1D = Vq1D'*diag(wq1D)*Vq1D;
-Pq1D = M1D\(Vq1D'*diag(wq1D));
-Vf1D = Vandermonde1D(N,[-1;1])/Vandermonde1D(N,rq1D); % interp from quad to face points
+% Vq1D = Vandermonde1D(N,rq1D)/V1D;
+% M1D = Vq1D'*diag(wq1D)*Vq1D;
+% Pq1D = M1D\(Vq1D'*diag(wq1D));
+% Vf1D = Vandermonde1D(N,[-1;1])/Vandermonde1D(N,rq1D); % interp from quad to face points
 
 % 3D matrices
 V = kron(kron(V1D,V1D),V1D);
-Vq = kron(kron(Vq1D,Vq1D),Vq1D);
-M = Vq'*diag(wq)*Vq;
-Pq = (Vq'*diag(wq)*Vq)\(Vq'*diag(wq));
+% Vq = kron(kron(Vq1D,Vq1D),Vq1D);
+% M = Vq'*diag(wq)*Vq;
+% Pq = (Vq'*diag(wq)*Vq)\(Vq'*diag(wq));
 Dr = kron(kron(I,D1D),I);
[…]

### 2019-09-19 docs/ESDG_ROM/main.tex
updating ROM paper

@@
 
-\author[rice]{Jesse Chan\corref{cor1}}
-\ead{Jesse.Chan@caam.rice.edu}
+\author[rice]{Jesse Chan}
+\ead{Jesse.Chan@rice.edu}
 \address[rice]{Department of Computational and Applied Mathematics, Rice University, 6100 Main St, Houston, TX, 77005}
@@
 Reduced order models of nonlinear conservation laws in fluid dynamics do not typically inherit stability properties of the full order model.
-We introduced projection-based reduced models of nonlinear conservation laws based on flux differencing which retain accuracy while inheriting a semi-discrete entropy inequality independently of the choice of basis, parameters, and discretization resolution.
+We introduce projection-based reduced models of nonlinear conservation laws which are globally conservative and inherit a semi-discrete entropy inequality independently of the choice of basis and choice of parameters.
 \end{abstract}
@@
 ROMs which retain either Lagrangian \cite{lall2003structure, carlberg2015preserving} or Hamiltonian structure \cite{benner2012interpolation, gugercin2012structure, peng2016symplectic, chaturantabut2016structure, gong2017structure, afkham2017structure, afkham2018structure} have been constructed by combining Galerkin projection with an appropriate formulation of the full order model, and similar energy-conserving ROMs have been constructed for the incompressible Navier-Stokes equations \cite{farhat2014dimensional, farhat2015structure}.
-The construction of structure-preserving ROMs to nonlinear conservation laws in fluid flow, however, remains an open problem.
+The construction of structure-preserving ROMs for nonlinear conservation laws in fluid flow, however, remains an open problem.
 For example, Galerkin projection yields ROMs which become unstable as the number of reduced basis functions (modes) is increased \cite{bui2007goal, carlberg2013gnat}.
@@
 
+The diffusive matrices $\bm{K}$ can be similarly extended to higher dimensions via Kronecker products
+\[
+\bm{K} = \bm{K}_{\rm 1D}\otimes \bm{I} + \bm{I}\otimes \bm{K}_{\rm 1D}.  
+\]
+An hyper-reduction of $\bm{K}$ analogous to the one described in Section~\ref{sec:diss} can be used to guarantee a discrete dissipation of entropy.
+
 \subsection{Non-periodic domains in higher dimensions}
@@
 The $dN$ constraints encoded in (\ref{eq:sbpconstraints}) can be interpreted as enforcing a discrete version of the fundamental theorem of calculus relating integrals of function derivatives to boundary integrals of function values.
-These constraints on the boundary weights $\bm{w}_b$ can be enforced, for example, by augmenting the non-negative least squares solve in Algorithm~\ref{alg:hr} with equality constraints.
-We take an alternative approach in this paper, and incorporate the constraints (\ref{eq:sbpconstraints}) directly into a hyper-reduction approach based on the solution of a linear programming problem using the dual simplex method.
-We refer the reader to \cite{yano2019lp} for details.
-%\note{Expand}
+We enforce these constraints on the boundary weights $\bm{w}_b$ (\ref{eq:sbpconstraints}) directly into a hyper-reduction approach based on the solution of a linear programming problem using the dual simplex method, and refer the reader to \cite{yano2019lp} for details.
+We note that it may also be possible to enforce these constraints, for example, by augmenting the non-negative least squares solve in Algorithm~\ref{alg:hr} with equality constraints, and will explore these directions in future work.
 
@@
 
-\subsection{2D Euler equations}% with wall boundary conditions}
-
-We now study the Euler equations in 2D with wall boundary conditions.
+\subsection{2D Euler equations}
+
+We now study the Euler equations in 2D.
+We begin by considering periodic boundary conditions on the domain $[-1,1]^2$ and a smoothed version of the Kelvin-Helmholtz instability.
+The initial condition is adapted from \cite{munz1989numerical, maboudi2018conservative}
+\begin{gather*}
+%a = .1; sig = .1; %5*sqrt(2)*1e-3;
+%ff = 1./(1+exp(-(y+.5)./(sig.^2)))-1./(1+exp(-(y-.5)./(sig.^2)));
+\rho = 1 + \frac{1}{1+e^{-(y+1/2)/\sigma^2}} - \frac{1}{1+e^{-(y-1/2)/\sigma^2}}, \qquad
+u =  \frac{1}{1+e^{-(y+1/2)/\sigma^2}} - \frac{1}{1+e^{-(y-1/2)/\sigma^2}} - \frac{1}{2}, \\
+v = \alpha\sin(2\pi x) \LRp{e^{-(y+1/2)/\sigma^2} - e^{-(y-1/2)/\sigma^2}},\qquad
+p = 2.5
+\end{gather*}
+where $\alpha = .1$ and $\sigma = .1$.
+
+\begin{figure}
+\centering
+\includegraphics[width=.4\textwidth]{figs/khsvd.png}
+\caption{Decay of singular values for the 2D smoothed Kelvin-Helmholtz instability.}
+\label{fig:khsvd}
+\end{figure}
+We utilize a full order model with $200\times 200$ grid ($40000$ cells), which is run until time $T = 3$.
+We generate 301 solution snapshots with which to compute POD modes.
+The decay of the singular values of the solution snapshots with and without enrichment by the entropy variables is shown in Figure~\ref{fig:khsvd}.
+We observe that, compared to the 1D case, the singular values decay more slowly when enriched with snapshots of the entropy variables.
+However, as in 1D, the differences between the singular values with and without entropy variable enrichment are significantly larger for lower energy modes as compared to higher energy modes.
+
+%\note{Plot singular value decay, etc}
+\begin{figure}
+\centering
+\subfloat[Full order model]{\includegraphics[width=.32\textwidth]{figs/khfom.png}}
+\hspace{.175em}
+\subfloat[Reduced order model, 75 modes]{\includegraphics[width=.32\textwidth]{figs/khrom75.png}}
+\hspace{.1em}
+\subfloat[Hyper-reduced points]{\includegraphics[width=.32\textwidth]{figs/khrom75_hrpts.png}}
+\caption{Comparison of results from the full and reduced order models (75 modes) of the 2D smoothed Kelvin-Helmholtz instability.}
+\label{fig:khrom}
+\end{figure}
+The reduced order model utilizes $75$ POD modes to achieves a $1.02\%$ relative $L^2$ error.
+We note that the difference between the 75th singular value of the snapshot matrix with entropy variable enrichment is only about $6\%$ larger than the 75th singular value without entropy variable enrichment.
+This suggests that entropy variable enrichment does not significantly impact the accuracy of the 75-mode POD approximation.
+
+Since little difference is observed between the hyper-reducted treatments of viscosity, the naive treatment (\ref{eq:visc3}) is used.
+No additional stabilizing points were necessary for this setting, as the condition numbers of the hyper-reduced $x$ and $y$ test mass matrices were both $O(1)$.
+Figure~\ref{fig:khrom} compares results for the full and reduced order models, and shows the hyper-reduced points selected by the greedy empirical cubature algorithm.
+
+We now consider a 2D domain with wall boundary conditions.
 It was shown in \cite{svard2014entropy, chen2017entropy} that wall boundary conditions can be imposed using a ``mirror state''.
@@
 The POD basis is computed from 75 solution snapshots with a CFL of $.5$.
-The reduced order solutions are shown in Figures~\ref{fig:pulse2d}.
-Since little difference is observed between the hyper-reducted treatments of viscosity, the naive treatment (\ref{eq:visc3}) is used.
+Figure~\ref{fig:pulse2dsvd} shows the decay of the singular values with and without entropy variable enrichment.
+As before, we observe that the singular values decay more slowly under entropy variable enrichment, but that the difference in the singular values with and without enrichment is more evident for low energy modes.
+The 25th singular value with entropy variable enrichment is less than $5\%$ larger than the 25th singular value without entropy variable enrichment, suggesting that entropy variable enrichment does not significantly affect the accuracy of the 25-mode POD approximation.
+
+\begin{figure}
+\centering
+\includegraphics[width=.4\textwidth]{figs/pulse2dsvd.png}
+\caption{Decay of singular values for the 2D Gaussian pulse problem.}
+\label{fig:pulse2dsvd}
+\end{figure}
+
+A 25-mode reduced order solution is shown in Figures~\ref{fig:pulse2d}.
 The hyper-reduction algorithm adds one single stabilizing point to reduce the condition number of the $x$-coordinate test mass matrix from $O(10^{8})$ to approximately $5.71$.
-
 The solution does not form shock discontinuities, and is chosen instead to test the imposition of boundary conditions.
@@
 
-\note{We conclude by examining a 2D problem with a non-trivial shock structure.  }
+%\note{We conclude by examining a 2D problem with a non-trivial shock structure.  }
 

### 2019-09-19 matlab/Wave2D_osc_curv_bdry.m

@@ -1,0 +1,295 @@
+% clear all
+
+Globals2D
+
+N = 6; K1D = 10;
+a = 0;
+a = .05;
+% a = .25;
+CFL = .5;
+
+useCurved = 1;
+plotMesh = 0;
+
+computeEigs = 0;
+global tau
+tau = 1;
+
+FinalTime = 1;
+
+[Nv, VX, VY, K, EToV] = unif_tri_mesh(K1D);
+StartUp2D;
+
+% BuildPeriodicMaps2D(2,0);
+% PlotMesh2D; return
+
+[rp sp] = EquiNodes2D(25); [rp sp] = xytors(rp,sp);
+Vp = Vandermonde2D(N,rp,sp)/V;
+xp = Vp*x; yp = Vp*y;
+
+%% make curvilinear mesh (still unstable?)
+
+global Drq Dsq Vfqf Pfq Lq Prq Psq Vq Pq
+global rxJ sxJ ryJ syJ nxq nyq sJq Jq Nfq  rxq sxq ryq syq
+
+
+Nq = 2*N+1;
+[rq sq wq] = Cubature2D(Nq); % integrate u*v*c
+Vq = Vandermonde2D(N,rq,sq)/V;
+M = Vq'*diag(wq)*Vq;
+Pq = M\(Vq'*diag(wq)); % J's cancel out
+xq = Vq*x; yq = Vq*y;
+Jq = Vq*J;
+
+Drq = Vq*Dr; Dsq = Vq*Ds;
+[rq1D wq1D] = JacobiGQ(0,0,N);
+V1D = Vandermonde1D(N,JacobiGL(0,0,N));
+Vq1D = Vandermonde1D(N,rq1D)/V1D;
+M1D = Vq1D'*diag(wq1D)*Vq1D;
+Pq1D = M1D\(Vq1D'*diag(wq1D));
+Pfq = kron(eye(Nfaces),Pq1D);
+Vfqf = kron(eye(Nfaces),Vq1D);
+
+rfq = [rq1D; -rq1D; -ones(size(rq1D))];
+sfq = [-ones(size(rq1D)); rq1D; -rq1D];
+% rfq = Vfqf*r(Fmask(:),:);
+% sfq = Vfqf*s(Fmask(:),:);
+wfq = [wq1D; wq1D; wq1D];
+
+Vfq = Vandermonde2D(N,rfq,sfq)/V;
[…]

### 2019-09-19 docs/ESDG_ROM/main.tex
updating ROM paper

@@
 Reduced order models of nonlinear conservation laws in fluid dynamics do not typically inherit stability properties of the full order model.
-We introduce projection-based reduced models of nonlinear conservation laws which are globally conservative and inherit a semi-discrete entropy inequality independently of the choice of basis and choice of parameters.
+We introduce projection-based hyper-reduced models of nonlinear conservation laws which are globally conservative and inherit a semi-discrete entropy inequality independently of the choice of basis and choice of parameters.
 \end{abstract}
@@
 
-The construction of stable ROMs for the compressible Navier-Stokes equations is further complicated by their non-trivial structure.
-Thus, practitioners often develop structure-preserving entropy stable ROMs for simplifications of the underlying PDE (such as the time-dependent \textit{linearized} compressible Navier-Stokes equations \cite{barone2009stable, kalashnikova2010stability, amsallem2012stabilization, kalashnikova2014construction}) and extrapolate such models to the full nonlinear equations.
+The construction of stable ROMs for the compressible Navier-Stokes equations is further complicated by the non-trivial structure of the equations.
+In response, practitioners have developed structure-preserving entropy stable ROMs for simplifications of the underlying PDE (such as the time-dependent \textit{linearized} compressible Navier-Stokes equations \cite{barone2009stable, kalashnikova2010stability, amsallem2012stabilization, kalashnikova2014construction}) and extrapolate such models to the full nonlinear equations.
 A promising alternative is to enforce physical conditions such as such as kinetic energy preservation, which are empirically related to stability \cite{maboudi2018conservative}.
-These methods significantly improve the robustness of numerical methods in practice, and can guarantee a discrete entropy inequality for systems of nonlinear conservation laws which yield stable split formulations \cite{fisher2013discretely}.
-
-The approach taken in this work differs in the treatment of nonlinear terms.
+These methods significantly improve the robustness of numerical methods in practice, and guarantee a discrete entropy inequality for systems of nonlinear conservation laws which yield stable split formulations \cite{fisher2013discretely}.
+
+The approach taken in this work differs from the existing literature in the treatment of nonlinear terms.
 In \cite{maboudi2018conservative}, equations are rewritten such that the nonlinear terms involve only polynomial (quadratic and cubic) nonlinearities, which can be evaluated exactly using precomputed matrices.
@@
 The paper is organized as follows.
-Section~\ref{sec:1} introduces entropy inequalities associated with systems of nonlinear conservation laws.
 Section~\ref{sec:2} introduces a full order model on 1D periodic domains based on entropy stable finite volume schemes.
@@
 Section~\ref{sec:6} describes how to extend the aforementioned approaches to non-periodic boundary conditions, and Section~\ref{sec:7} describes the extension to higher dimensions.
-We conclude in Section~\ref{sec:8} with numerical experiments verifying theoretical results.
-
-\section{Entropy inequalities for nonlinear conservation laws}
-\label{sec:1}
-Let $\Omega$ denote some domain with boundary $\partial \Omega$.  Nonlinear conservation laws are expressed as a system of partial differential equations (PDEs) 
+We conclude in Section~\ref{sec:8} with numerical experiments which verifying the presented theoretical results.
+
+%\section{Entropy inequalities for nonlinear conservation laws}
+%\label{sec:1}
+
+\section{The full order model: entropy stable finite volume schemes}
+\label{sec:2}
+
+We briefly summarize entropy inequalities associated with systems of nonlinear conservation laws.
+Let $\Omega$ denote some domain with boundary $\partial \Omega$.
+Nonlinear conservation laws are expressed as a system of partial differential equations (PDEs)
 \begin{equation}
@@
 
-\section{The full order model: entropy stable finite volume schemes}
-\label{sec:2}
-
-The reduced order models in this work are constructed using full order models based on entropy stable finite volume methods \cite{tadmor1987numerical}.
+The reduced order models in this work are constructed using full order models based on entropy stable finite volume methods \cite{tadmor1987numerical}, which reproduce a discrete version of the continuous entropy inequality (\ref{eq:entropyineq}).
 For simplicity, we illustrate the construction of full and reduced order models on a periodic 1D domain.
@@
 \end{gather}
-Let the domain be decomposed into $K$ cells of size $\Delta x$, and let ${\LRp{\bm{u}_h}_i}$ denote the vector containing mean values of the conservation variables over the $i$th cell.
+These fluxes are used to construct entropy conservative and entropy stable finite volume schemes \cite{tadmor1987numerical, chandrashekar2013kinetic, ray2016entropy}.
+Let the domain be decomposed into $K$ cells of size $\Delta x$, and let ${\LRp{\bm{u}_h}_i}$ denote the vector containing mean values of the vector of conservation variables over the $i$th cell.
 An entropy conservative finite volume method results from discretizing the integral form of the conservation law as follows
@@
 
-We first rewrite the system (\ref{eq:nonlineqs1d}) in a matrix form which will be more amenable to model reduction.
+We first rewrite the system (\ref{eq:nonlineqs1d}) in a matrix form which is more amenable to model reduction.
 We define the skew-symmetric differentiation matrix $\bm{Q}$ and flux matrix $\bm{F}$ such that
@@
 The proof can be found in the literature \cite{tadmor1987numerical, tadmor2003entropy, tadmor2016entropy}.
-We present an alternative proof which utilizes matrix properties of entropy stable summation by parts (SBP) schemes \cite{carpenter2014entropy, gassner2016split, chen2017entropy, crean2018entropy, chan2017discretely}.
+Inspired by entropy stable summation by parts (SBP) schemes \cite{carpenter2014entropy, gassner2016split, chen2017entropy, crean2018entropy, chan2017discretely}, we present an alternative proof which relies only on matrix properties of $\bm{Q}$.
 We test (\ref{eq:es}) with the vector of entropy variables $\bm{v}_h = \bm{v}(\bm{u}_h)$
@@
 where $\epsilon$ is the visosity coefficient.
-This choice of dissipation simplifies descriptions of how to treat viscous regularizations of the solution in an entropy stable fashion.
-Future work will explore both physically relevant viscosities and more nuanced artificial dissipation mechanisms.
+This choice is intended to simplify the presentation of entropy stable treatments of diffusion terms.
+Future work will analyze both physically relevant viscosities and more nuanced artificial dissipation mechanisms.
 
@@
 
-Let $\LRc{{\phi}_j(x)}_{j=1}^{N}$ denote a reduced basis for each component of the solution, which may be generated using principal orthogonal decomposition (POD) or the reduced basis procedure.
+Let $\LRc{{\phi}_j(x)}_{j=1}^{N}$ denote a reduced basis for each component of the solution, which may be generated using (for example) principal orthogonal decomposition (POD) or the reduced basis procedure.
 Let $\bm{V}$ denote the generalized Vandermonde matrix whose columns contain evaluations of $\phi_j$ at grid points $x_i$
@@
 \]
-We approximate values of the solution by $\bm{u}_h = \bm{V}\bm{u}_N$, where $\bm{u}_N$ denote ``modal'' coefficients  of the solution in the reduced basis.
-Ignoring viscosity terms for now, plugging this expression into (\ref{eq:es}) and enforcing that the residual is orthogonal to all columns of $\bm{V}$ (Galerkin projection) yields a reduced order system
+We approximate grid values of the solution by $\bm{u}_h = \bm{V}\bm{u}_N$, where $\bm{u}_N$ denote ``modal'' coefficients  of the solution in the reduced basis.
+Ignoring viscosity terms for now, plugging this expression into (\ref{eq:es}) and enforcing that the residual is orthogonal to all columns of $\bm{V}$ (Galerkin projection) yields a reduced system
 \begin{align*}
@@
 This is not possible due to the fact that the projected entropy variables $\tilde{\bm{v}}$ are no longer mappings of the grid values of the conservative variables $\bm{u}_h = \bm{V}\bm{u}_N$.
-To remedy this, we follow approaches taken in \cite{parsani2016entropy, chan2017discretely} and replace the values of ${\bm{u}_h}$ used to evaluate the flux matrix $\bm{F}$ with the \textit{entropy-projected conservative variables} $\tilde{\bm{u}}$
+To remedy this, we follow approaches taken in \cite{parsani2016entropy, chan2017discretely} and replace the grid values of ${\bm{u}_h}$ used to evaluate the flux matrix $\bm{F}$ with values of the \textit{entropy-projected conservative variables} $\tilde{\bm{u}}$
 \[
@@
 \]
-The entropy projected conservative variables are thus mappings of the projected entropy variables.
+The entropy projected conservative variables $\tilde{\bm{u}}$ are thus mappings of the projected entropy variables.
 We note that $\bm{V}\bm{V}^{\dagger}$ can also be interpreted as a discrete approximation of the $L^2$ projection operator onto the span of the reduced basis $\phi_j$.
@@
 \]
-Additionally, if the reduced basis $\bm{V}$ exactly represents $\bm{1}$, solutions of (\ref{eq:esrom}) conserve global averages of the conservative variables.
+Additionally, if $\bm{1}$ lies within the range of the reduced basis matrix $\bm{V}$ (e.g., $1$ is in the span of the reduced basis functions $\phi_1, \ldots, \phi_N$), solutions of (\ref{eq:esrom}) conserve global averages of the conservative variables.
 \label{thm:esrom}
@@
 The global conservation results from testing with $\bm{1}$.
-Let $\bm{V}\bm{e}=\bm{1}$, where $\bm{e}$ denote coefficients of $\bm{1}$ in the reduced basis matrix.
+If $\bm{1}$ is in the range of $\bm{V}$, then the exist coefficients in the reduced basis matrix $\bm{e}$ such that $\bm{V}\bm{e}=\bm{1}$.
 Then,
@@
 To reduce the cost of evaluating nonlinear terms, we introduce a second \textit{hyper-reduction} step \cite{ryckelynck2009hyper}.
-Here, a vector of nonlinear function evaluations is approximated by a vector constructed using nonlinear function evaluations at a subset of points \cite{barrault2004empirical, bui2004aerodynamic, chaturantabut2010nonlinear, farhat2015structure, drmac2016new, hernandez2017dimensional, yano2019lp}.
-In this work, we utilize a sampling and weighting hyper-reduction strategy \cite{farhat2015structure, hernandez2017dimensional, yano2019lp} which enables proofs of discrete entropy stability.
+Here, terms involving vectors of nonlinear function evaluations are approximated by terms involving nonlinear function evaluations at a subset of points \cite{barrault2004empirical, bui2004aerodynamic, chaturantabut2010nonlinear, farhat2015structure, drmac2016new, hernandez2017dimensional, yano2019lp}.
+In this work, we utilize a sampling and weighting strategy \cite{farhat2015structure, hernandez2017dimensional, yano2019lp} which enables proofs of discrete entropy stability.
 These hyper-reduction techniques can be interpreted as reduced quadratures, and produce approximations of the form
@@
 \]
-then the proof of entropy conservation will carry over to the hyper-reduced model.
+then the proof of entropy conservation in Theorem~\ref{thm:esrom} will carry over to the hyper-reduced model.
 Unfortunately, common hyper-reduction techniques preserve either skew-symmetry or zero row sums, but not both.
@@
 This hyper-reduction technique preserves skew-symmetry, but does not necessarily preserve the zero row sum property.
-Similarly, techniques such as gappy POD or empirical interpolation \cite{everson1995karhunen, barrault2004empirical, chaturantabut2010nonlinear} may preserve the zero row sum but not skew-symmetry.
+Similarly, techniques such as gappy POD or empirical interpolation \cite{everson1995karhunen, barrault2004empirical, chaturantabut2010nonlinear} may preserve the zero row sum property but not skew-symmetry.
 
@@
 In other words, if a vector lies in the range of $\bm{V}_t$, the projection matrix $\bm{P}_t$ recovers the coefficients exactly.
-We also note that if the hyper-reduced set $\mathcal{I}$ is taken to include all the points and $\bm{W} = \Delta x\bm{I}$, then $\bm{P}_t$ reduces to the pseudo-inverse $\bm{V}_t^{\dagger} = \LRp{\bm{V}_t\bm{V}_t}^{-1}\bm{V}_t^T$.
+We also note that if the hyper-reduced set $\mathcal{I}$ is taken to include all the points and $\bm{W}$ is a multiple of the identity (for example, $\bm{W} = \Delta x \bm{I}$), then $\bm{P}_t$ reduces to the pseudo-inverse $\bm{V}_t^{\dagger} = \LRp{\bm{V}_t\bm{V}_t}^{-1}\bm{V}_t^T$.
 
@@
 \begin{lemma}
-Suppose that $\bm{1}$ is exactly representable in the test basis and that Assumption~\ref{ass:quad} holds.  Define the matrix $\bm{Q}_t$
+Suppose that $\bm{1}$ is exactly representable in the test basis $\bm{V}_t$ and that Assumption~\ref{ass:quad} holds.  Define the matrix $\bm{Q}_t$
 \begin{equation}
@@
 The POD modes are thus also anti-symmetric across the origin; however, their derivatives $\bm{Q}\bm{V}$ are nearly symmetric across the origin, as shown in Figure~\ref{fig:modeQ}.
-As a result, since the entries of $\bm{V}^T\bm{Q}\bm{V}$ are inner products of nearly symmetric and anti-symmetric functions, $\bm{V}^T\bm{Q}\bm{V} \approx \bm{0}$.
+Since the entries of $\bm{V}^T\bm{Q}\bm{V}$ are inner products of nearly symmetric and anti-symmetric functions, $\bm{V}^T\bm{Q}\bm{V} \approx \bm{0}$, and the resulting hyper-reduced model is highly inaccurate.
 \begin{figure}[!h]
@@
 In the absence of hyper-reduction (where $\bm{P}_t = \bm{V}_t^{\dagger}$), the matrix $\bm{Q}_t$ should be an accurate approximation to $\bm{Q}$.
-The following lemma quantifies the relationship between accuracy and choice of test space.
+The following lemma quantifies the relationship between accuracy in the $2$-norm and choice of test space.
 %The entropy conservative flux for Burgers' equation corresponds to the discretization of the split form \cite{canuto2006spectral, chan2017discretely, maboudi2018conservative}
@@
  \end{proof}
+ In other words, if $\bm{Q}\bm{V}$ is contained within the range of $\bm{V}_t$, the error in approximating $\bm{Q}\bm{f}$ using the compress-and-projected matrix $\LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}$ scales with the best approximation error $\nor{\bm{f}-\bm{V}_t\bm{V}_t^{\dagger}\bm{f}}$.
 
@@
 \label{sec:condtest}
-We note that the target space is designed to accurately approximate the mass matrix $\Delta x \bm{V}^T\bm{V}$, and does not take the test basis matrix $\bm{V}_{t}$ into account.
-This is because incorporating the test basis by approximating either $\Delta x \bm{V}^T\bm{V}_{t}$ or $\Delta x \bm{V}_{t}^T\bm{V}_{t}$ produces a significantly larger number of hyper-reduced points without a noticable increase in accuracy.
+We note that the choice of target space is designed to accurately approximate the mass matrix $\Delta x \bm{V}^T\bm{V}$, and does not take the test basis matrix $\bm{V}_{t}$ into account.
+This choice is motivated by the empirical observation that incorporating the test basis by approximating either $\Delta x \bm{V}^T\bm{V}_{t}$ or $\Delta x \bm{V}_{t}^T\bm{V}_{t}$ produces a significantly larger number of hyper-reduced points without any significant improvement in accuracy.
 However, because the hyper-reduction strategy does not account for $\bm{V}_{t}$, it is possible to construct a hyper-reduced quadrature which accurately integrates the mass matrix but violates Assumption~\ref{ass:quad} by producing a singular test mass matrix $\bm{V}_t\LRp{\mathcal{I},:}^T\bm{W}\bm{V}_t\LRp{\mathcal{I},:}$.
-The challenge is then to construct a hyper-reduced volume quadrature which 1) accurately approximates the reduced basis mass matrix $\Delta x \bm{V}^T\bm{V}$ while 2) ensuring that the hyper-reduced test mass matrix is non-singular.
+The challenge is then to construct a hyper-reduced volume quadrature which accurately approximates the reduced basis mass matrix $\Delta x \bm{V}^T\bm{V}$ while ensuring that the hyper-reduced test mass matrix is non-singular.
 We address this challenge by computing the spectra of the test mass matrix  and adding additional ``stabilizing'' points if the condition number is larger than a specified tolerance.
 
-Let $\bm{z}_j$ denote the $N_z$ eigenvectors corresponding to zero (small) eigenvalues of the hyper-reduced test mass matrix.
+Let $\bm{z}_j$ denote the $N_z$ eigenvectors corresponding to small eigenvalues of the hyper-reduced test mass matrix.
 The entries of $\bm{z}_j$ correspond to vectors of coefficients in the test basis $\bm{V}_t$.
@@
 We compute a second group of hyper-reduced points with index set $\mathcal{I}_{\bm{Z}}$  using the greedy hyper-reduction algorithm applied to the target space $\bm{Z}_{\rm target}$, which we define analogously to $\bm{V}_{\rm target}$ in (\ref{eq:Vtarget}).
-We note that, unlike the hyper-reduced points computed previously, we do not perform a dimensionality reduction on $\bm{Z}_{\rm target}$.
-
-Given $\mathcal{I}_{\bm{Z}}$, we then update the hyper-reduced index set as the union $\mathcal{I} = \mathcal{I}_{\bm{V}} \cup \mathcal{I}_{\bm{Z}}$, and compute a new set of hyper-reduced weights through the non-negative least squares solve
+We note that, unlike the hyper-reduced points computed previously, we do not perform a dimensionality reduction on $\bm{Z}_{\rm target}$ when computing this set of stabilizing points.
+
+Given $\mathcal{I}_{\bm{Z}}$, we update the hyper-reduced index set as the union $\mathcal{I} = \mathcal{I}_{\bm{V}} \cup \mathcal{I}_{\bm{Z}}$, and compute a new set of hyper-reduced weights through the non-negative least squares solve
 \[
@@
 \end{enumerate}
-Then, an entropy conservative reduced order model is given by the following theorem:
+Then, we have the following theorem:
 \begin{theorem}
@@
 \end{gather}
-semi-discretely conserves the average entropy 
+semi-discretely conserves the sampled and weighted average entropy  
 \[
@@
 
-\subsection{Online cost estimates for hyper-reduction}
+\subsection{Online cost estimates for the hyper-reduced system}
 \label{sec:cost}
-In practice, the flux matrix $\bm{F}$ is not computed explicitly, but instead is usually formed on the fly by evaluating the entropy stable flux $\bm{f}_S$.
+In practice, the flux matrix $\bm{F}$ is not computed explicitly, but is instead formed on the fly by evaluating the entropy stable flux $\bm{f}_S$.
 Due to the complexity of the formulas for $\bm{f}_S$, the computation of $\bm{F}$ constitutes a significant portion of the computational cost.
@@
 
-While this cost is trivially parallelizable, it is a significant increase in online costs compared to traditional hyper-reduction procedures.
+While the $O(N_s^2)$ nonlinear flux evaluations are trivially parallelizable, they constitute a significant increase in online costs compared to traditional hyper-reduction procedures.
 However, we note that the cost can be comparable to techniques for exactly treating higher order polynomial nonlinearities for a large number of modes \cite{maboudi2018conservative}.
-For example, suppose one wishes to compute $\bm{V}^T f\LRp{\bm{V}\bm{u}_N}$, where $f(x) = x^m$ for $m \geq 1$ and $m$ integer.
+For example, suppose one wishes to compute $\bm{V}^T g\LRp{\bm{V}\bm{u}_N}$, where $g(x) = x^m$ for $m \geq 1$ and $m$ integer.
 We can rewrite this as
 \[
-\bm{V}^T f(\bm{u}) = \bm{V}^T \diag{\bm{V}\bm{u}_N}^{m-1} \bm{V}\bm{u}_N.
+\bm{V}^T g(\bm{u}) = \bm{V}^T \diag{\bm{V}\bm{u}_N}^{m-1} \bm{V}\bm{u}_N.
 \]
@@
 \[
-\bm{V}^T f\LRp{\bm{V}\bm{u}_N} = \sum_{j=1}^N (\bm{u}_N)_j \bm{V}^T \bm{V}_j \bm{V}\bm{u}_N.
+\bm{V}^T g\LRp{\bm{V}\bm{u}_N} = \sum_{j=1}^N (\bm{u}_N)_j \bm{V}^T \bm{V}_j \bm{V}\bm{u}_N.
 \]
@@
 \[
-\bm{V}^T f\LRp{\bm{V}\bm{u}_N} = \sum_{j=1}^N \sum_{k=1}^M (\bm{u}_N)_j  (\bm{u}_N)_k \bm{V}^T \bm{V}_j \bm{V}_k \bm{V}\bm{u}_N.
+\bm{V}^T g\LRp{\bm{V}\bm{u}_N} = \sum_{j=1}^N \sum_{k=1}^M (\bm{u}_N)_j  (\bm{u}_N)_k \bm{V}^T \bm{V}_j \bm{V}_k \bm{V}\bm{u}_N.
 \]
@@
 Assuming flux evaluations dominate costs, entropy stable schemes require $\alpha^2 \beta N^2$ operations.
-Thus, rather than increasing the asymptotic computational complexity, entropy stable hyper-reduction increases the associated constant.
+Thus, rather than increasing the asymptotic computational complexity, the entropy stable hyper-reduction presented here increases the associated constant.
 
@@
 \[
-\Delta x \bm{1}^T\td{S\LRp{\bm{V}\bm{u}_N}}{t} = -\bm{v}_N^T\bm{d}(\bm{u}_N).
+\Delta x \bm{1}^T\td{S\LRp{\bm{V}\bm{u}_N}}{t} = -\bm{v}_N^T\bm{d}(\bm{u}_N), \qquad \bm{v}_N = \bm{P}\bm{v}\LRp{\bm{V}\LRp{\mathcal{I},:}\bm{u}_N}.
 \]
@@
 The hyper-reduction treatments of viscosity described in this section can also be applied to nonlinear interface dissipation models, such as local Lax-Friedrichs penalization.
-We do not focus on such approximations for this paper, as the amount of dissipation applied is typically proportional to $\Delta x$ and thus depends on the dimension of the full order model.
+We do not focus on such approximations for this paper, as the amount of dissipation applied is typically proportional to $\Delta x$ and thus changes based on the dimension of the full order model.
 \end{remark}
@@
 \]
-where $\bm{V}_t(:,1), \bm{V}_t(:,K)$ denote the first and last columns (corresponding to points on the boundary) of the test basis matrix $\bm{V}_t$.
+where $\bm{V}_{bt}$ interpolates test basis functions to boundary points, and $\bm{V}_t(:,1), \bm{V}_t(:,K)$ denote the first and last columns (corresponding to points on the boundary) of the test basis matrix $\bm{V}_t$.
 
@@
 However, due to the presence of $\bm{E}$ within the boundary term $\bm{E}^T\bm{B}\bm{E}$, the accurate and entropy stable imposition of nonlinear boundary conditions using generalized SBP-SATs remains an open problem \cite{crean2018entropy, chan2018efficient, chenreview}.
-We address this issue by constructing a \textit{hybridized} SBP operator (also referred to as a decoupled SBP operator) \cite{chan2017discretely, chan2019skew}
+We address this issue by constructing a \textit{hybridized} SBP operator \cite{chan2017discretely, chan2019skew} (also referred to as a decoupled SBP operator in \cite{chan2017discretely})
 \[
@@
 
-A more subtle property is that $\bm{Q}_h\bm{1} = \bm{0}$ \cite{chan2019skew}.
+It is less straightforward to show that $\bm{Q}_h\bm{1} = \bm{0}$, which is also necessary for the proof of entropy stability \cite{chan2019skew}.
 Note that the full order matrix $\bm{Q}$ in (\ref{eq:fomsbp}) satisfies $\bm{Q}\bm{1} = \bm{0}$.
@@
 where we have used that $\bm{Q}_t\bm{1} = \bm{0}$ and that (\ref{eq:preproduce}) implies $\bm{E}\bm{1} = \bm{1}$.
-Here, we abuse notation by letting $\bm{1}$ denote any arbitrary length vector of all ones.
+Here, we abuse notation by letting $\bm{1}$ denote the vector of ones with appropriate dimension.
 Furthermore, since $\bm{Q}_t$ satisfies a generalized SBP property,
@@
 
-Non-periodic domains and boundary conditions can be treated by combining hybridized SBP operators and a weak imposition of boundary conditions as was done in the 1D hyper-reduced formulation (\ref{eq:esbchr}).
+Non-periodic domains and boundary conditions can be treated by combining hybridized SBP operators and a weak imposition of boundary conditions as was done for the 1D hyper-reduced formulation (\ref{eq:esbchr}).
 Without loss of generality, we assume a square domain with $4$ boundary faces, each of which is discretized by $K_{\rm 1D}$ intervals of size $\Delta x$.
-We assume we are given values of the outward normal vector $\bm{n} = \LRs{n_1, \ldots, n_d}^T$ on the domain boundary.
-We enforce non-periodic boundary conditions through introducing numerical fluxes at boundary points.
+Denote values of the outward normal vector on the domain boundary by $\bm{n} = \LRs{n_1, \ldots, n_d}^T$.
+We enforce non-periodic boundary conditions through numerical fluxes at boundary points.
 However, the number of boundary points scales with $K_{\rm 1D}$.
@@
 \]
-where $\bm{W}_b$ is a diagonal matrix whose entries consist of positive weights
-\[
-\bm{W}_b = {\rm diag}\LRp{\bm{w}_b}.
-\]
-We can now define the $N_b\times N_b$ operators $\bm{B}^1, \bm{B}^2$ as the diagonal matrices whose entries consist of weighted component values of outward normals on the boundary 
+where $\bm{W}_b = {\rm diag}\LRp{\bm{w}_b}$ is a diagonal matrix whose entries consist of positive weights.  
+We also define $N_b\times N_b$ operators $\bm{B}^1, \bm{B}^2$ as the diagonal matrices whose entries consist of weighted component values of outward normals on the boundary 
 \[
@@
 
-We can now construct hybridized SBP operators under both volume and boundary hyper-reduction.
-Let $\bm{V}_{bt}$ be the matrix which evaluates the $i$th test basis at face points.
-Then, we can define $\bm{E}_i$ as the matrix which extrapolates from hyper-reduced volume points to boundary points through projection onto the $i$th test basis
+We can now construct hybridized SBP operators using hyper-reduced points on both the volume and boundary of the domain.
+Let $\bm{V}_{bt}$ be the matrix which evaluates the $i$th test basis at all points on the boundary.
+Then, we can define $\bm{E}_i$ as the matrix which extrapolates from hyper-reduced volume points to hyper-reduced boundary points through projection onto the $i$th test basis
 \[
@@
 The analogous proof of entropy stability in multiple dimensions requires that $\bm{Q}^i_h\bm{1} = \bm{0}$ for $i = 1,\ldots, d$.
-However, this is not automatically satisfied.
+However, this is not automatically satisfied under an arbitrary hyper-reduction of the boundary points.
 Expanding out $\bm{Q}^i_h\bm{1}$ yields
@@
 \end{equation}
-The $dN$ constraints encoded in (\ref{eq:sbpconstraints}) can be interpreted as enforcing a discrete version of the fundamental theorem of calculus relating integrals of function derivatives to boundary integrals of function values.
+The $dN$ constraints encoded in (\ref{eq:sbpconstraints}) can be interpreted as enforcing a discrete version of the fundamental theorem of calculus relating approximate integrals of reduced basis derivatives to boundary integrals of reduced basis values.
 We enforce these constraints on the boundary weights $\bm{w}_b$ (\ref{eq:sbpconstraints}) directly into a hyper-reduction approach based on the solution of a linear programming problem using the dual simplex method, and refer the reader to \cite{yano2019lp} for details.
@@
 We utilize a full order finite volume model with $2500$ cells on $[-1,1]$ with a CFL of $.75$, and set the artificial viscosity parameter to $\epsilon = 2\times 10^{-4}$.
-Entropy stable reflective wall boundary conditions are imposed using ``mirror states'' at both endpoints, and the initial condition is set to be
+It was shown in \cite{svard2014entropy, chen2017entropy} that wall boundary conditions can be imposed using a ``mirror state''.
+We use a boundary numerical flux $\bm{f}^* = \bm{f}_S\LRp{\bm{u}^+,\bm{u}}$ augmented with local Lax-Friedrichs penalization, where the exterior state $\bm{u}^+$ is defined by
+\[
+\rho^+ = \rho, \qquad {u}^+ = -u, \qquad p^+ = p.  
+\]  The initial condition is set to be 
 \[
@@
 In both cases, the decay is similar, with entropy variable enrichment resulting in a slightly slower decay for singular values smaller than $10^{-8}$.
-The difference becomes more pronounced for a coarser subsampling.
+The difference becomes more pronounced for a coarser subsampling of the snapshots.
 We note that the enrichment does affect the form of the resulting singular vectors (Figure~\ref{subfig:c}).
@@
 We next examine solutions produced by the reduced order models.
-It is known that reduced models can typically utilize a larger CFL \cite{lucia2004reduced, knezevic2011reduced}; however, to ensure that temporal errors are small, all solutions are computed using the same CFL as the full order model.
+It is known that reduced models can sometimes utilize a larger CFL compared the full order model \cite{lucia2004reduced, knezevic2011reduced}; however, to ensure that temporal errors are small, all solutions are computed using the same CFL used to generate the solution snapshots.
 Figure~\ref{fig:romsols} shows density computed using 25, 75, and 125 modes and the hyper-reduced treatment of viscosity in (\ref{eq:visc2}).
@@
 However, despite this under-resolution, the ROM remains stable and does not blow up.
+Moreover, the reduced order solution converges uniformly to the full order solution as the number of modes increases.
 
@@
 
-We now compare the evolution of errors and entropy, and examine the effect of hyper-reduction on the discrete solution in Figure~\ref{fig:entropyevo}.
-Both entropy and error are computed on the original grid of the full order model (e.g., without hyper-reduction).
-The error is computed as the discrete relative $L^2$ error over all the conservative variables.
-We observe that the evolution of the ROM average discrete entropy converges to the average discrete entropy for the full order model as the number of modes increases from 25 to 75.
+We now compare the evolution of both the $L^2$ error and discrete entropy, and examine the effect of hyper-reduction on the discrete solution in Figure~\ref{fig:entropyevo}.
+Both error and entropy are computed on the original grid of the full order model (e.g., without hyper-reduction), and the error is computed as the discrete relative $L^2$ error over all the conservative variables.
+We observe that the evolution of the average discrete entropy for the reduced model rapidly approaches the average discrete entropy for the full order model as the number of modes increases from 25 to 75.
 For 125 modes, the average entropies for the ROM and FOM are indistinguishable and are not shown.
@@
 v = \alpha\sin(2\pi x) \LRp{e^{-(y+1/2)/\sigma^2} - e^{-(y-1/2)/\sigma^2}},\qquad
-p = 2.5
+p = 2.5.
 \end{gather*}
-where $\alpha = .1$ and $\sigma = .1$.
+In the following experiments, we set $\alpha = .1$ and $\sigma = .1$.
 
@@
 \includegraphics[width=.4\textwidth]{figs/khsvd.png}
-\caption{Decay of singular values for the 2D smoothed Kelvin-Helmholtz instability.}
+\caption{Decay of singular values for snapshots of the 2D smoothed Kelvin-Helmholtz instability.}
 \label{fig:khsvd}
@@
 \subfloat[Hyper-reduced points]{\includegraphics[width=.32\textwidth]{figs/khrom75_hrpts.png}}
-\caption{Comparison of results from the full and reduced order models (75 modes) of the 2D smoothed Kelvin-Helmholtz instability.}
+\caption{Full and reduced order density for the 2D smoothed Kelvin-Helmholtz instability at $T = 3$ with $\epsilon = 1e-3$.
+The reduced order model uses 75 modes and $884$ hyper-reduced points (in \textcolor{red}{red}).
+The relative $L^2$ error is $.0102$.}
 \label{fig:khrom}
@@
 
-We now consider a 2D domain with wall boundary conditions.
-It was shown in \cite{svard2014entropy, chen2017entropy} that wall boundary conditions can be imposed using a ``mirror state''.
+We next consider a 2D domain with wall boundary conditions.
+Entropy stable reflective wall boundary conditions are again imposed using ``mirror states''.
 Let $\bm{n}$ denote the outward normal at a wall, and let $\bm{u}_{\bm{n}} = un_x + vn_y$ and $\bm{u}_{\bm{n}^\perp} = un_y - vn_x$ denote the normal and tangential components of the velocity at the wall.
@@
 \includegraphics[width=.4\textwidth]{figs/pulse2dsvd.png}
-\caption{Decay of singular values for the 2D Gaussian pulse problem.}
+\caption{Decay of singular values for snapshots of the 2D Gaussian pulse problem.}
 \label{fig:pulse2dsvd}
@@
 \subfloat[Hyper-reduction points]{\includegraphics[width=.32\textwidth]{pulse2d_ROM_pts.png}}
-\caption{Density computed for $\epsilon = 1e-3$ using 25 modes, 300 hyper-reduced points (in \textcolor{red}{red}), and 66 hyper-reduced boundary points (in \textcolor{blue}{blue}).
+\caption{Full and reduced order density at time $T = .25$ computed with $\epsilon = 1e-3$.
+The reduced order model uses 25 modes, 300 hyper-reduced points (in \textcolor{red}{red}), and 66 hyper-reduced boundary points (in \textcolor{blue}{blue}).
 The relative $L^2$ error is $.0071$.}
@@
 
-\section{Conclusion}  We have presented a methodology for constructing projection-based reduced order models for nonlinear conservation laws from a reduced basis, modified Galerkin projection, and tailored hyper-reduction techniques.
-The main novelty of these new models is the approximation of the nonlinear convective term, which combines a ``flux differencing'' approach with a hyper-reduced approximation of the differentiation matrix.
+\section{Conclusion}  We have presented a methodology for constructing projection-based reduced order models for nonlinear conservation laws by combining a reduced basis, a modified Galerkin projection, and tailored hyper-reduction techniques.
+The main novelty of these new reduced models is the approximation of the nonlinear convective term, which combines a ``flux differencing'' approach with an appropriate hyper-reduced approximation of the differentiation matrix.
 
@@
 This discrepancy is due to the fact that the hyper-reduction presented here approximates a nonlinear matrix, rather than a nonlinear vector.
-These costs can be reduced by combining domain decomposition \cite{lucia2003reduced} and a discontinuous Galerkin-type discretization \cite{chan2017discretely} to increase the sparsity of the resulting hyper-reduced matrices.
-We will also investigate implicit time-stepping for entropy stable reduced models, as the additional costs associated with the spatial formulation may be offset by the smaller matrix systems which must be solved in the linearization process.
+These costs can be reduced by combining domain decomposition \cite{lucia2003reduced} and a discontinuous Galerkin-type discretization \cite{chan2017discretely} to produce multi-domain reduced order models.
+Suppose there are $k = 1,\ldots, K$ subdomains.
+Then, this sort of multi-domain model is expected to reduce costs if the sum of the squares of the number of subdomain sampling points $\sum_k (N_s^k)^2$ is significantly smaller than the global number of sampling points $N_s^2$.
+Future work will also investigate implicit time-stepping for entropy stable reduced models, as additional costs associated with the spatial formulation may be offset by the reduction in cost for solving smaller matrix systems during the linearization process.
 

### 2019-10-14 docs/ESDG_mortar/main.tex
updating paper

--- 2018-10-19 docs/ESDG_mortar/main.tex
@@
 
-%\documentclass[preprint,10pt]{elsarticle}
-%\documentclass[preprint,10pt]{amsart}
-\documentclass[review]{siamart0216}
-%\documentclass{siamart0216}
+\documentclass{svjour3}                     % onecolumn (standard format)
+%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
+%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
+%\documentclass[twocolumn]{svjour3}          % twocolumn
+%\usepackage[T1]{fontenc}
+
+%\usepackage[displaymath]{lineno}
+
+\smartqed  % flush right qed marks, e.g. at end of proof
+
+\makeatletter
+\renewcommand*{\@thmcounterend}{}
+\makeatother
+%\spnewtheorem{lemma}{Lemma}{\bfseries}{\itshape}
+\spnewtheorem{assumption}{Assumption}{\bfseries}{\itshape}
+%\spnewtheorem{remark}{Remark}{\bfseries}{\itshape}
 
 %\usepackage{fullpage}
-%\usepackage[colorlinks=true]{hyperref}
+\usepackage{cite}
+\usepackage[numbers,sort&compress]{natbib}
 
-\usepackage{amsmath,amssymb,amsfonts}
-%\usepackage{amsthm}
-%\theoremstyle{definition}
-%\newtheorem{definition}{Definition}
-%\theoremstyle{lemma}
-%\newtheorem{lemma}{Lemma}
-%\theoremstyle{corollary}
-%\newtheorem{corollary}{Corollary}
-\newtheorem*{remark}{Remark}
-%\theoremstyle{theorem}
-%\newtheorem{theorem}{Theorem}
+%\usepackage{slashbox} % for slashbox in tables
+
+\usepackage{amsmath, amsfonts, amssymb}
+\usepackage{stmaryrd} % for llbracket
+
+%\newtheorem*{remark}{Remark}
 %\theoremstyle{assumption}
 %\newtheorem{assumption}{Assumption}
+\usepackage{hyperref}
+\hypersetup{colorlinks,
+citecolor=blue
+}
 
 \usepackage[titletoc,toc,title]{appendix}
-
-\usepackage{array} 
-\usepackage{listings}
-\usepackage{mathtools}
-\usepackage{pdfpages}
-\usepackage[textsize=footnotesize,color=green]{todonotes}
-\usepackage{bm}
-\usepackage{bbm}
-
 \usepackage{tikz}
-\usepackage[normalem]{ulem}
-\usepackage{hhline}
-
-\usepackage{graphicx}
+%\usepackage{hhline}
+\usepackage{graphicx, color, bm}
 \usepackage{subfig}
-\usepackage{color}
-
-%% ====================================== graphics
-
 \usepackage{pgfplots}
 \usepackage{pgfplotstable}
-\definecolor{markercolor}{RGB}{124.9, 255, 160.65}
-\pgfplotsset{
-compat=1.3,
-width=10cm,
-tick label style={font=\small},
-label style={font=\small},
-legend style={font=\small}
-}
+\usepackage{slopetri} % my own package for drawing triangle slopes
 
-\usetikzlibrary{calc}
-\usetikzlibrary{intersections}
+\renewcommand{\hat}{\widehat}
+\renewcommand{\tilde}{\widetilde}
 
-%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
-\newcommand{\logLogSlopeTriangle}[5]
-{
-    % #1. Relative offset in x direction.
-    % #2. Width in x direction, so xA-xB.
-    % #3. Relative offset in y direction.
-    % #4. Slope d(y)/d(log10(x)).
-    % #5. Plot options.
-
-    \pgfplotsextra
-    {
-        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
-        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
-        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
-        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}
-
-        % Calculate auxilliary quantities, in relative sense.
-        \pgfmathsetmacro{\xArel}{#1}
-        \pgfmathsetmacro{\yArel}{#3}
-        \pgfmathsetmacro{\xBrel}{#1-#2}
-        \pgfmathsetmacro{\yBrel}{\yArel}
-        \pgfmathsetmacro{\xCrel}{\xArel}
-
-        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
-        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
-        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.
-
-        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
-        \coordinate (A) at (rel axis cs:\xArel,\yArel);
-        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
-        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);
-
-        % Draw slope triangle.
-        \draw[#5]   (A)-- node[pos=0.5,anchor=north] {}
-                    (B)-- 
-                    (C)-- node[pos=0.5,anchor=west] {\textcolor{black}{#4}}
-                    cycle;
-    }
-}
-%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
-
-\newcommand{\logLogSlopeTriangleNeg}[5]
-{
-    % #1. Relative offset in x direction.
-    % #2. Width in x direction, so xA-xB.
-    % #3. Relative offset in y direction.
-    % #4. Slope d(y)/d(log10(x)).
-    % #5. Plot options.
-
-    \pgfplotsextra
-    {
-        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
-        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
-        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
-        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}
-
-        % Calculate auxilliary quantities, in relative sense.
-        \pgfmathsetmacro{\xArel}{#1}
-        \pgfmathsetmacro{\yArel}{#3}
-        \pgfmathsetmacro{\xBrel}{#1-#2}
-        \pgfmathsetmacro{\yBrel}{\yArel}
-        \pgfmathsetmacro{\xCrel}{\xArel}
-
-        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
-        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
-        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.
-
-        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
-        \coordinate (A) at (rel axis cs:\xArel,\yArel);
-        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
-        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);
-
-        % Draw slope triangle.
-        \draw[#5]   (A)-- node[pos=.5,anchor=south] {}
-                    (B)-- 
-                    (C)-- node[pos=0.5,anchor=west] {\textcolor{black}{#4}}
-                    cycle;
-    }
-}
-%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
-
-%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
-\newcommand{\logLogSlopeTriangleFlipNeg}[5]
-{
-    % #1. Relative offset in x direction.
-    % #2. Width in x direction, so xA-xB.
-    % #3. Relative offset in y direction.
-    % #4. Slope d(y)/d(log10(x)).
-    % #5. Plot options.
-
-    \pgfplotsextra
-    {
-        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
-        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
-        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
-        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}
-
-        % Calculate auxilliary quantities, in relative sense.
-        %\pgfmathsetmacro{\xArel}{#1}
-        %\pgfmathsetmacro{\yArel}{#3}
-        \pgfmathsetmacro{\xBrel}{#1-#2}
-        \pgfmathsetmacro{\yBrel}{#3}
-        \pgfmathsetmacro{\xCrel}{#1}
-
-        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
-        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
-        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.
-
-	\pgfmathsetmacro{\xArel}{\xBrel}
-        \pgfmathsetmacro{\yArel}{\yCrel}
-
-        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
-        \coordinate (A) at (rel axis cs:\xArel,\yArel);
-        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
-        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);
-
-        % Draw slope triangle.
-        \draw[#5]   (A)-- node[pos=0.5,anchor=east] {\textcolor{black}{#4}}
-                    (B)-- 
-                    (C)-- node[pos=0.5,anchor=north] {1}
-                    cycle;
-    }
-}
-%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
-
-%%% START MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
-\newcommand{\logLogSlopeTriangleFlip}[5]
-{
-    % #1. Relative offset in x direction.
-    % #2. Width in x direction, so xA-xB.
-    % #3. Relative offset in y direction.
-    % #4. Slope d(y)/d(log10(x)).
-    % #5. Plot options.
-
-    \pgfplotsextra
-    {
-        \pgfkeysgetvalue{/pgfplots/xmin}{\xmin}
-        \pgfkeysgetvalue{/pgfplots/xmax}{\xmax}
-        \pgfkeysgetvalue{/pgfplots/ymin}{\ymin}
-        \pgfkeysgetvalue{/pgfplots/ymax}{\ymax}
-
-        % Calculate auxilliary quantities, in relative sense.
-        %\pgfmathsetmacro{\xArel}{#1}
-        %\pgfmathsetmacro{\yArel}{#3}
-        \pgfmathsetmacro{\xBrel}{#1-#2}
-        \pgfmathsetmacro{\yBrel}{#3}
-        \pgfmathsetmacro{\xCrel}{#1}
-
-        \pgfmathsetmacro{\lnxB}{\xmin*(1-(#1-#2))+\xmax*(#1-#2)} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnxA}{\xmin*(1-#1)+\xmax*#1} % in [xmin,xmax].
-        \pgfmathsetmacro{\lnyA}{\ymin*(1-#3)+\ymax*#3} % in [ymin,ymax].
-        \pgfmathsetmacro{\lnyC}{\lnyA+#4*(\lnxA-\lnxB)}
-        \pgfmathsetmacro{\yCrel}{\lnyC-\ymin)/(\ymax-\ymin)} % THE IMPROVED EXPRESSION WITHOUT 'DIMENSION TOO LARGE' ERROR.
-
-	\pgfmathsetmacro{\xArel}{\xBrel}
-        \pgfmathsetmacro{\yArel}{\yCrel}
-
-        % Define coordinates for \draw. MIND THE 'rel axis cs' as opposed to the 'axis cs'.
-        \coordinate (A) at (rel axis cs:\xArel,\yArel);
-        \coordinate (B) at (rel axis cs:\xBrel,\yBrel);
-        \coordinate (C) at (rel axis cs:\xCrel,\yCrel);
-
-        % Draw slope triangle.
-        \draw[#5]   (A)-- node[pos=0.5,anchor=east] {\textcolor{black}{#4}}
-                    (B)-- 
-                    (C)-- node[pos=0.5,anchor=south] {}
-                    cycle;
-    }
-}
-%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.
-
-\usepackage{stmaryrd}
-
-\renewcommand{\tilde}{\widetilde}
-\renewcommand{\hat}{\widehat}
-\renewcommand{\topfraction}{0.85}
-\renewcommand{\textfraction}{0.1}
-\renewcommand{\floatpagefraction}{0.75}
-
-\newcommand{\vect}[1]{\ensuremath\boldsymbol{#1}}
-\newcommand{\tensor}[1]{\underline{\bm{#1}}}
-\newcommand{\del}{\triangle}
-\newcommand{\curl}{\grad \times}
-\renewcommand{\div}{\grad \cdot}
-
-\newcommand{\bbm}[1]{\mathbbm{#1}}
-\newcommand{\bs}[1]{\boldsymbol{#1}}
-\newcommand{\equaldef}{\stackrel{\mathrm{def}}{=}}
-
+%% d in integrand
+\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}}
+\newcommand{\diag}[1]{{\rm diag}\LRp{#1}}
 \newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}{\rm #2}}}
@@
 
-\newcommand{\Oh}{{\Omega_h}}
-\renewcommand{\L}{L^2\LRp{\Omega}}
-\newcommand{\LK}{L^2\LRp{D^k}}
-\newcommand{\LdK}{L^2\LRp{\partial D^k}}
-\newcommand{\Dhat}{\widehat{D}}
-\newcommand{\Lhat}{L^2\LRp{\Dhat}}
+\renewcommand{\note}[1]{{\color{blue}{#1}}}
+\newcommand{\bnote}[1]{{\color{blue}{#1}}}
+\newcommand{\rnote}[1]{{\color{red}{#1}}}
 
-\newcommand{\eval}[2][\right]{\relax
-  \ifx#1\right\relax \left.\fi#2#1\rvert}
+\graphicspath{{./figs/}}
 
-\def\etal{{\it et al.~}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\newcommand{\note}[1]{{\color{blue}{#1}}}
-%\newcommand{\noteOne}[1]{{\color{blue}{#1}}}
-%\newcommand{\noteTwo}[1]{{\color{red}{#1}}}
-%\newcommand{\note}[1]{#1}
-%\newcommand{\noteOne}[1]{#1}
-%\newcommand{\noteTwo}[1]{#1}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\newcommand{\LinfDk}{L^{\infty}\LRp{D^k}}
+\date{}
+%\title{Skew-symmetric entropy stable modal discontinuous Galerkin formulations}% with applications to hybrid meshes}
+%\titlerunning{Skew-symmetric entropy stable DG formulations}
 
-\newcommand{\diag}[1]{{\rm diag}\LRp{#1}}
+\author{Jesse Chan, David C.\ Del Rey Fernandez, Mario Bencomo}
+\title{Mortar-based entropy stable discontinuous Galerkin methods on non-conforming quadrilateral and hexahedral meshes}
+\titlerunning{Mortar-based entropy stable DG methods}
 
-\newcommand{\Ksub}{K_{\rm sub}}
-
-\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-
-%% d in integrand
-\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}}
-
-\makeatletter
-\renewcommand\d[1]{\mspace{6mu}\mathrm{d}#1\@ifnextchar\d{\mspace{-3mu}}{}}
-\makeatother
-
-%\date{}
-\author{Jesse Chan, David C.\ Del Rey Fernandez}
-\title{Mortar-based entropy stable discontinuous Galerkin methods}
 \graphicspath{{./figs/}}
@@
 
+\begin{abstract}
+High order entropy stable discontinuous Galerkin (DG) methods for nonlinear conservation laws reproduce a discrete entropy inequality by combining entropy conservative finite volume fluxes with summation-by-parts (SBP) discretization matrices.
+On tensor product (quadrilateral and hexahedral) elements, SBP matrices are typically constructed by collocating at Lobatto quadrature points.
+Recent work has extended the construction of entropy stable DG schemes to collocation at more accurate Gauss quadrature points \cite{chan2018efficient}.
+
+In this work, we extend entropy stable Gauss collocation schemes to non-conforming meshes.
+Entropy stable DG schemes require computing entropy conservative numerical fluxes between volume and surface quadrature nodes.
+On conforming tensor product meshes where volume and surface nodes are aligned, flux evaluations are required only between ``lines'' of nodes.
+However, on non-conforming meshes, volume and surface nodes are no longer aligned, resulting in a larger number of flux evaluations.
+We reduce this expense by introducing an entropy stable mortar-based treatment of non-conforming interfaces via a face-local correction term, and provide truncation error estimates for the resulting discretizations.
+Numerical experiments confirm the stability and accuracy of this approach, and we conclude by discussing an extension of this mortar treatment of non-conforming interfaces to general elements and arbitrary quadrature rules \cite{chan2017discretely}.
+\end{abstract}
+
+\begin{enumerate}
+\item Intro
+\item Gauss collocation schemes, conditions for entropy stability
+\item Increased cost on non-conforming meshes
+\item Mortar scheme and entropy stability
+\item Analysis of accuracy of Global SBP operator.
+Gauss nodes = exact for degree $N$, while Lobatto = exact for degree $(N-1)$.
+Approach in Friedrichs + DCDR for $hp$ meshes yields degree $(N-1)$ global operator.
+\item Extensions
+\begin{enumerate}
+\item Curved meshes
+\item $p$-nonconforming meshes
+\item General elements and why you'd want to implement this.
+\end{enumerate}
+\item Numerical experiments (2D vortex, 2D entropy conservation test, 3D convergence).
+Note that reduced rates of convergence are not observed for $N>1$ on $p$-nonconforming meshes, or on hybrid meshes with mixed quadrature.
+\end{enumerate}
+
 \section{Introduction}
 
-\note{Boilerplate introduction on high order + stability}
+%\note{Boilerplate introduction on high order + stability}
 
-\section{Mortar methods for hybrid and non-conforming meshes}
+Discretely entropy stability has emerged as a methodology for designing high order schemes for nonlinear conservation laws.
+Entropy stable discretizations ensure the satisfaction of a semi-discrete entropy inequality by combining specific finite volume numerical fluxes with summation-by-parts (SBP) discretization matrices.
+Compared to traditional high order methods, the resulting schemes demonstrate significantly improved robustness in the presence of under-resolved solution features such as shocks or turbulence while retaining high order accuracy.
 
-This work is motivated by complications which arise when designing entropy stable couplings between elements which do not share the same boundary nodes.
+Entropy stable discontinuous Galerkin (DG) methods were originally constructed for quadrilateral and hexahedral meshes based on nodal collocation at Lobatto quadrature points \cite{fisher2013high, carpenter2014entropy, gassner2016split}.
+Entropy stable schemes were later extended to simplicial and more general elements using tailored volume and surface quadrature rules \cite{chen2017entropy, crean2018entropy}.
+More general quadrature rules were addressed in \cite{chan2017discretely, chan2018discretely, chan2018efficient, chan2019skew}, including entropy stable collocation schemes on quadrilateral and hexahedral elements based on more accurate Gauss quadrature rules and generalized SBP operators \cite{chan2018efficient}.
+
+The work presented here focuses on geometrically non-conforming meshes.
+Such meshes may arise when applying domain decomposition techniques to a complex geometry (e.g.
+meshing sub-domains independently)  \cite{bernardi1993domain} or performing local mesh refinement.
+Entropy stable Lobatto collocation schemes have been constructed on non-conforming meshes in \cite{friedrich2017entropy} using SBP projection operators.
+In this work, we extend entropy stable Gauss collocation to non-conforming quadrilateral and hexahedral meshes.
+While it is straightforward to construct Gauss collocation schemes on non-conforming meshes, the treatment of non-conforming interfaces results in signficantly increased computational costs.
+We reduce such costs by adopting a mortar-based treatment of non-conforming interfaces.
+Moreover, while the cost of the proposed scheme is similar to that of  \cite{friedrich2017entropy} in 2D, the mortar-based approach is more computationally efficient for both Lobatto and Gauss collocation schemes on 3D non-conforming hexahedral meshes.
+
+The paper is organized as follows: Section~\ref{sec:0} briefly reviews the derivation of an entropy inequality for a system of nonlinear conservation laws.
+Section~\ref{sec:1} introduces ``hybridized'' SBP operators as a unified way to treat both Lobatto and Gauss collocation schemes on conforming meshes of tensor product elements.
+Section~\ref{sec:2} describes a naive extension to non-conforming meshes and illustrates why this formulation results in an increase in computational costs.
+Section~\ref{sec:3} introduces a mortar-based formulation which addresses such costs, and characterizes the accuracy of the resulting formulations.
+Section~\ref{sec:4} describes the extension to curved meshes, as well as more general elements and $p$ non-conformity.
+We conclude with numerical validation of theoretical results in Section~\ref{sec:5}.
+
+\section{Entropy stability for systems of nonlinear conservation laws}
+\label{sec:0}
+We are interested in the numerical approximation of solutions to systems of nonlinear conservation laws
+\begin{equation}
+\pd{\bm{u}}{t} + \sum_{i=1}^d \pd{\bm{f}_i\LRp{\bm{u}}}{x_i} = 0.
+\label{eq:nonlinpde}
+\end{equation}
+Here, $\bm{u}$ denotes the conservative variables and $\bm{f}_i(\bm{u})$ are nonlinear fluxes.
+We briefly review entropy stability for systems of conservation laws in $d$ dimensions.
+We assume there exists a convex scalar entropy $S(\bm{u})$ associated with (\ref{eq:nonlinpde}).
+We then define the entropy variables $\bm{v}(\bm{u})$ as the gradient of the entropy $S(\bm{u})$ with respect to the conservative variables
+\[
+\bm{v} = \pd{S(\bm{u})}{\bm{u}}.  
+\]
+For $S(\bm{u})$ convex, $\bm{v}(\bm{u})$ defines an invertible mapping between the conservative and entropy variables, whose inverse (from entropy to conservative variables) we denote by $\bm{u}(\bm{v})$.
+Viscosity solutions to (\ref{eq:nonlinpde}) satisfy an integrated form of the entropy inequality \cite{dafermos2005compensated}
+%\begin{gather}
+%\pd{S(\bm{u})}{t} + \sum_{i=1}^d \pd{F_i(\bm{u})}{x_i} \leq 0, \qquad F_i(\bm{u}) = \bm{v}^T\pd{\bm{f}_i}{\bm{u}}, 
+%\label{eq:entropyineq}
+%\end{gather}
+%Integrating (\ref{eq:entropyineq}) over a domain $\Omega$ and applying the divergence theorem yields 
+\begin{equation}
+\int_{\Omega} \pd{S(\bm{u})}{t} + \int_{\partial \Omega} \sum_{i=1}^d \bm{n}_i \LRp{\bm{v}^T\bm{f}_i(\bm{u}) - \psi_i(\bm{u})} \leq 0,
+\label{eq:weakentropyineq}
+\end{equation}
+where $F_i$ denotes the $i$th scalar entropy flux function, $\psi_i(\bm{u}) = \bm{v}^T\bm{f}_i(\bm{u}) - F_i(\bm{u})$ denotes the $i$th entropy potential, $\partial \Omega$ denotes the boundary of $\Omega$ and $\bm{n}_i$ denotes the $i$th component of the outward normal on $\partial \Omega$.
+This can be interpreted as implying that the time rate of change of entropy is bounded by the entropy flux through the boundary.
+
+\section{Entropy stable collocation DG methods and hybridized SBP operators}
+\label{sec:1}
+
+In this section, we summarize the work of \cite{chan2018efficient} on the construction of entropy stable collocation DG methods based on generalized summation by parts (GSBP) operators.
+These constructions are applicable to collocation schemes based on either Lobatto and Gauss nodes.
+
+\subsection{Entropy conservative schemes in 1D}
+
+We begin by introducing collocation discretization matrices on the reference interval $[-1,1]$.
+We assume the solution is collocated at $(N+1)$ quadrature points $x_i$ with associated quadrature weights $w_i$, and consider primarily Lobatto or Gauss quadrature points.
+The collocation assumption is equivalent to approximating the solution using a degree $N$ Lagrange basis $\ell_j(x)$ defined on the $(N+1)$ quadrature points.
+
+We define mass and integrated differentiation matrices $\bm{M}, \bm{Q}$
+\[
+\bm{M}_{ij} = \int_{-1}^1 \ell_i(x)\ell_j(x), \qquad \bm{Q}_{ij} = \int_{-1}^1 \pd{\ell_j}{x}\ell_i.
+\]
+We assume that all integrals are computed using the collocated quadrature rule, such that 
+\begin{gather*}
+\bm{M}_{ij} = \int_{-1}^1 \ell_i(x)\ell_j(x) \approx \sum_{k=1}^{N+1} \ell_i(x_k)\ell_j(x_k) w_k = \delta_{ij} w_i\\
+\bm{Q}_{ij} = \int_{-1}^1 \pd{\ell_j}{x}\ell_i \approx \sum_{k=1}^{N+1} \ell_i(x_k)\LRl{\pd{\ell_j}{x}}_{x_k} w_k.  
+\end{gather*}
+In other words, the mass matrix is diagonal with entries equal to the quadrature weights.
+Since the integrands of $\bm{M}$ are degree $2N$ polynomials, the collocation approximation of $\bm{M}$ is exact for Gauss quadrature, but not for Lobatto quadrature.
+For both Gauss and Lobatto quadrature, the matrix $\bm{Q}$ is exact under collocation quadrature.
+
+We introduce the $2\times (N+1)$ matrix $\bm{E}$ which interpolates values at collocation nodes to values at the endpoints $x = -1$ and $x=1$.  These two matrices are defined entrywise as
+\[
+\LRp{\bm{E}}_{1i} = \ell_i(-1), \qquad  \LRp{\bm{E}}_{2i} = \ell_i(1).
+\]
+The mass and differentiation matrices satisfy a generalized summation by parts property \cite{fernandez2014generalized}
+\begin{equation}
+\bm{Q} = \bm{E}^T \bm{B} \bm{E} - \bm{Q}^T, \qquad \bm{B} = \begin{bmatrix}-1 & \\ & 1\end{bmatrix}.
+\label{eq:gsbp}
+\end{equation}
+% The proof is a direct restatement of integration by parts, and can be found in \cite{fernandez2014generalized, ortleb2016kinetic, ortleb2017kinetic, ranocha2018generalised}.  
+The GSBP property holds for both Lobatto and Gauss nodes, and switching between these two nodal sets simply requires redefining the matrices $\bm{D}, \bm{E}$.
+For Gauss nodes, $\bm{E}$ is dense.
+For Lobatto nodes, since the collocation nodes include boundary points, the interpolation matrix $\bm{E}$ reduces to the matrix which extracts nodal values associated with the left and right endpoints
+\[
+\bm{E} = \begin{bmatrix}
+1 & 0 & \ldots & 0\\
+0 & \ldots & 0 & 1
+\end{bmatrix}.
+\]
+
+It is possible to construct energy stable discretizations of linear hyperbolic systems using GSBP operators based on Gauss nodes \cite{fernandez2014generalized}.
+However, for nonlinear conservation laws, the presence of the dense $\bm{E}$ matrix in the GSBP property (\ref{eq:gsbp}) complicates the imposition of boundary conditions and computation of inter-element numerical fluxes \cite{crean2018entropy, chan2017discretely, chan2018efficient}.
+
+We can avoid this issue by introducing ``hybridized'' (or decoupled) SBP operators \cite{chan2017discretely, chenreview}, which are defined as the block matrix $\bm{Q}_h$ \cite{chan2019skew}
+\[
+\bm{Q}_h = \frac{1}{2}\begin{bmatrix}
+\bm{Q}-\bm{Q}^T & \bm{E}^T\bm{B}\\
+-\bm{B}\bm{E} & \bm{B}
+\end{bmatrix}.
+\]
+Note that the hybridized SBP operator satisfies a block form of the SBP property
+\begin{equation}
+\bm{Q}_h + \bm{Q}_h^T = \begin{bmatrix}
+\bm{0} & \\
+& \bm{B}\end{bmatrix}.
+\label{eq:hsbp}
+\end{equation}
+where $\bm{E}$ does not appear in the block boundary matrix on the right hand side.  Additionally, it is straightforward to show that 
+\[
+\bm{Q}_h\bm{1} = \bm{0}
+\]
+using the GSBP property and properties of $\bm{Q}$.
+
+The aforementioned matrices can now be used to construct high order accurate entropy stable and entropy conservative schemes.
+We first introduce an entropy conservative numerical flux \cite{tadmor1987numerical}.
+Let $\bm{u}_L, \bm{u}_R$ denote left and right states in the conservative variables.
+Then, an entropy conservative numerical flux is a vector-valued function $\bm{f}_S$ which satisfies the following three properties
+\begin{gather*}
+\bm{f}_S\LRp{\bm{u},\bm{u}} = \bm{f}(\bm{u}), \qquad \text{(consistency)}\\
+\bm{f}_S\LRp{\bm{u}_L,\bm{u}_R} = \bm{f}_S(\bm{u}_R,\bm{u}_L), \qquad \text{(symmetry)}\\
+\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}_S\LRp{\bm{u}_L,\bm{u}_R} = \psi(\bm{u}_L) - \psi(\bm{u}_R), \qquad \text{(conservation)}.
+\end{gather*}
+Here, $\bm{v}_L, \bm{v}_R$ are the entropy variables evaluated at $\bm{u}_L, \bm{u}_R$, and $\psi$ is the entropy potential which appears in (\ref{eq:weakentropyineq}).
+Then, an entropy conservative scheme on the reference interval is given as
+\begin{gather}
+\bm{M}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T
+\LRp{\bm{Q}_h \circ \bm{F}}\bm{1} + \bm{E}^T\bm{B}\LRp{\bm{f}^*-\bm{f}(\tilde{\bm{u}}_f)} = 0 \label{eq:esdg1D}\\ 
+\tilde{\bm{u}} = \begin{bmatrix}
+\bm{u}_h\\
+\tilde{\bm{u}}_f
+\end{bmatrix}, \qquad \tilde{\bm{u}}_f = \bm{u}\LRp{\bm{E}\bm{v}(\bm{u}_h)}, \qquad \bm{f}^* = \bm{f}_S\LRp{\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f}.
+\nonumber
+\end{gather}
+Here, $\tilde{\bm{u}}_f^+$ denotes some exterior state which can be used to enforce boundary conditions or coupling between different elements, and $\circ$ denotes the matrix Hadamard product.
+
+\note{Maybe add proof here and show how rest of proofs just rely on SBP and $\bm{Q}_h\bm{1} = \bm{0}$?}.
+For proofs of high order accuracy and entropy conservation, we refer the reader to \cite{chan2018efficient}.
+
+The matrix $\bm{F}$ is formed by computing ${\bm{F}}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}$ on the fly.
+Each entry requires evaluating the entropy conservative numerical flux $\bm{f}_S$, resulting in a large number of floating point operations \cite{wintermeyer2018entropy}.
+However, due to properties of the Hadamard product in (\ref{eq:esdg1D}), flux evaluations are only necessary between pairs of nodes corresponding to non-zero entries of $\bm{Q}_h$.
+We will revisit this point later in the construction of entropy stable schemes on non-conforming meshes.
+
+\subsection{Entropy conservation formulations in higher dimensions}
+
+The formulation (\ref{eq:weakentropyineq}) can be extended to higher dimensions through a tensor product construction.
+For simplicity, we illustrate this in 2D (the extension to 3D is straightforward).
+Let $\bm{M}_{\rm 1D}, \bm{Q}_{\rm 1D}$ denote one-dimensional SBP norm (mass) and differentiation matrices, and let $\bm{E}_{\rm 1D}$ denote the 1D face interpolation matrix.
+We define multi-dimensional  mass and differentiation matrices in terms of Kronecker products
+\[
+\hat{\bm{Q}}_1 = \bm{Q}_{\rm 1D} \otimes \bm{M}_{\rm 1D}, \qquad \hat{\bm{Q}}_2  = \bm{M}_{\rm 1D} \otimes \bm{Q}_{\rm 1D}, \qquad \hat{\bm{M}} = \bm{M}_{\rm 1D} \otimes  \bm{M}_{\rm 1D}.
+\]
+We also construct 2D face interpolation matrices from Kronecker products.
+Let $\bm{E}_{\rm 1D}$ denote the one-dimensional face interpolation matrix, and let $\bm{B}_{\rm 1D}$ denote the 1D boundary matrix in (\ref{eq:gsbp}).
+For a specific ordering of the face points, the two-dimensional face interpolation matrix $\bm{E}$ and reference boundary matrices $\hat{\bm{B}}_1, \hat{\bm{B}}_2$ are given by
+\begin{equation}
+\bm{E} = \begin{bmatrix}
+\bm{E}_{\rm 1D} \otimes \bm{I}_{2}\\
+\bm{I}_{2} \otimes \bm{E}_{\rm 1D} 
+\end{bmatrix}, \qquad 
+\hat{\bm{B}}_1 = \begin{bmatrix}
+\bm{B}_{\rm 1D} \otimes \bm{M}_{\rm 1D} & \\
+& \bm{0}
+\end{bmatrix}, \qquad 
+\hat{\bm{B}}_2 = \begin{bmatrix}
+\bm{0} &\\
+& \bm{M}_{\rm 1D} \otimes \bm{B}_{\rm 1D} 
+\end{bmatrix},
+\label{eq:EBdefs}
+\end{equation}
+where $\bm{I}_2$ is the $2\times 2$ identity matrix.  Recall that $\bm{M}_{\rm 1D}$ is the diagonal matrix of quadrature weights, so $\hat{\bm{B}}_i$ are diagonal.
+
+The 2D differentiation and interpolation matrices $\hat{\bm{Q}}_i, \bm{E}$ can now be used to construct 2D hybridized SBP operators.
+Let $\hat{\bm{Q}}_{i,h}$ denote the hybridized SBP operator for the $i$th coordinate on the reference element, where $\hat{\bm{Q}}_{i,h}$ is defined as
+\begin{align*}
+\hat{\bm{Q}}_{i,h} = \frac{1}{2}\begin{bmatrix}
+\hat{\bm{Q}}_i - \hat{\bm{Q}}_i^T & \bm{E}^T\hat{\bm{B}}_i\\
+-\hat{\bm{B}}_i\bm{E} & \hat{\bm{B}}_i
+\end{bmatrix}.
+\end{align*}
+These matrices satisfy the multi-dimensional SBP property
+\begin{align}
+\hat{\bm{Q}}_{i,h} + \hat{\bm{Q}}_{i,h}^T = \begin{bmatrix}
+\bm{0} & \\
+& \hat{\bm{B}}_i
+\end{bmatrix}.
+\label{eq:hsbp}
+\end{align}
+The construction of 3D differentiation and interpolation matrices proceeds similarly.  Then, a $d$-dimensional entropy conservative formulation on the reference element $[-1,1]^d$ is given as 
+\begin{gather}
+\hat{\bm{M}}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T
+\sum_{i=1}^d \LRp{\hat{\bm{Q}}_{i,h} \circ \bm{F}_i}\bm{1} + \bm{E}^T\hat{\bm{B}}_i\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_f)} = 0 \label{eq:esdgref}\\
+\LRp{\bm{F}_i}_{jk} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}
+\end{gather}
+where $\bm{f}^i_S$ is the numerical flux for the $i$th coordinate direction, and $\tilde{\bm{u}},\tilde{\bm{u}}_f$ are defined as in (\ref{eq:esdg1D}).
+
+Finally, we can extend this formulation to the multi-element case.
+Suppose the domain $\Omega$ is decomposed into non-overlapping elements $D^k$ which are images of the reference mapping, such that $D^k$ is a differentiable mapping of the reference element $\hat{D} = [-1,1]^d$.
+Let $\bm{x}_i$ denote the $i$th physical coordinate on $D^k$ and let $\hat{\bm{x}}_j$ denote the $j$th reference coordinate.
+The mapping between reference and physical element induces scaled geometric terms
+\[
+G_{ij} = J\pd{\hat{\bm{x}}_j}{\bm{x}_i},
+\]
+where $J$ denotes the determinant of the Jacobian of the physical-to-reference mapping.  We refer to $J$ as the Jacobian from here onwards.
+These geometric terms also relate the normal vectors on reference and physical elements.
+Let $\hat{\bm{n}}\hat{J}_f$ denote the outward normal vector on $\hat{D}$, scaled by the reference face Jacobians.
+Then, the (scaled) physical normal vectors on $D^k$ are related to $\hat{\bm{n}}\hat{J}_f$ through
+\[
+\bm{n}J_f = G\hat{\bm{n}}\hat{J}_f.
+\]
+We assume that the mesh is watertight, such that the scaled normal vectors are equal and opposite across shared faces between two elements.
+
+We can now assemble physical mass and differentiation matrices on $D^k$ from reference mass and differentiation matrices using the chain rule.
+Let $\hat{\bm{Q}}_{j,h}$ denote the $j$th reference differentiation matrix on $\hat{D}$.
+We can define the physical differentiation matrices $\bm{Q}_{i,h}$ on $D^k$ via the skew-symmetric splitting
+\[
+\bm{Q}_{i,h} = \frac{1}{2}\sum_{j=1}^d \diag{\bm{g}_{ij}}\hat{\bm{Q}}_{j,h} + \hat{\bm{Q}}_{j,h} \diag{\bm{g}_{ij}},
+\]
+where $\bm{g}_{ij}$ denotes the vector of values of $G_{ij} = J\pd{\hat{\bm{x}}_j}{\bm{x}_i}$ at both volume and surface points.
+We also assume that the geometric terms $\bm{g}_{ij}$ are constructed such that they satisfy a discrete geometric conservation law (GCL)
+\[
+\sum_{j=1}^d \hat{\bm{Q}}_{j,h} \bm{g}_{ij} = \bm{0}.
+\]
+The discrete GCL is satisfied automatically for isoparametric mappings in 2D, and there exist several techniques to enforce the satisfaction of a discrete GCL on various three-dimensional domains \cite{thomas1979geometric, kopriva2006metric, crean2018entropy, chan2018discretely, kozdon2018energy, kopriva2019free}.
+
+Because the reference matrices $\hat{\bm{Q}}_{j,h}$ satisfy the SBP property (\ref{eq:hsbp}), one can show \cite{chan2018discretely} that the physical matrices $\bm{Q}_{i,h}$ satisfy a physical SBP property.
+\begin{lemma}
+\label{lemma:Qhprops}
+
+\[
+\bm{Q}_{i,h} + \bm{Q}_{i,h}^T = \begin{bmatrix}
+\bm{0} &\\
+& \bm{B}_i \end{bmatrix}, \qquad \bm{B}_i = \diag{\bm{n}_i \circ \bm{w}_f},
+\]
+where $\bm{n}_i$ is a vector whose entries consist of values of the $i$th component of the scaled normals $n_iJ_f$ at face points, and $\bm{w}_f$ is the vector of face quadrature weights $w_i$.  
+\end{lemma}
+
+Then, an entropy conservative scheme is given by
+\begin{gather}
+{\bm{M}}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T
+\sum_{i=1}^d \LRp{{\bm{Q}}_{i,h} \circ \bm{F}_i}\bm{1} + \bm{E}^T{\bm{B}}_i\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_f)} = 0 \label{eq:esdg}
+\end{gather}
+where $\bm{F}_i$ is defined as in (\ref{eq:esdgref}).
+Here, the exterior state $\tilde{\bm{u}}_f$ used to evaluate the numerical flux $\bm{f}_i^*$ corresponds either to interface values on a neighboring element or an exterior state used to enforce boundary conditions.
+
+All entropy conservative schemes described here can be made entropy stable by introducing entropy dissipation through mechanisms such as physical or artificial viscosity \cite{tadmor2006entropy,upperman2019entropy}.
+We introduce dissipation by incorporating a penalization term into the interface flux \cite{winters2017uniquely}.
+For example, Lax-Friedrichs dissipation can be added by modifying the interface flux $\bm{f}^*_i$
+\[
+\bm{f}^*_i \Longrightarrow \bm{f}^*_i - \frac{\lambda}{2} \jump{\tilde{\bm{u}}_f}, \qquad  \jump{\tilde{\bm{u}}_f} = \tilde{\bm{u}}_f^+ - \tilde{\bm{u}}_f,
+\]
+where $\lambda$ is an estimate of the maximum wave speed \cite{chen2017entropy, chan2017discretely}.
+
+\section{Non-conforming meshes}
+\label{sec:2}
+
+Section~\ref{sec:1} describes the construction of entropy stable schemes on geometrically conforming meshes, where each element shares at most one neighbor across a face.
+We extend the construction of stable schemes to meshes containing geometric non-conformity, where an element can share a face with two or more neighboring elements.
+To ensure stability, the coupling conditions imposed at this non-conforming face must be handled appropriately.
+For DG discretizations, this is most naturally achieved by combining composite quadrature rules on non-conforming faces with appropriate evaluations of volume terms \cite{kozdon2018energy}.
+However, for entropy stable schemes, the naive use of composite quadrature at non-conforming interfaces can significantly increase the computational cost.
+
+For quad and hex elements under tensor product volume quadrature, the differentiation matrices are Kronecker products of 1D differentiation matrices and diagonal mass matrices, which result in sparse operators.
+As a result, flux evaluations are only required between ``lines'' of volume nodes \cite{carpenter2014entropy, chan2018efficient}.
+Flux evaluations also follow the sparsity pattern of the matrix $\bm{E}$, which maps from volume nodes to surface nodes.
+For Lobatto or Gauss collocation methods on conforming quadrilateral and hexahedral meshes, $\bm{E}$ is sparse if the surface quadrature nodes are aligned with volume quadrature nodes \cite{chan2018efficient}.
+In such cases, flux evaluations are required only between lines of volume nodes and adjacent surface nodes, as shown in Figure~\ref{subfig:aligned}.
+
+Composite quadrature rules on non-conforming meshes, however, are not aligned with volume nodes.
+Suppose that $D^k$ is a quadrilateral element with two neighbors across each face, such that every face is non-conforming and utilizes a composite quadrature rule.
+The entropy conservative formulation (\ref{eq:esdg}) can be extended to the non-conforming case by redefining the interpolation matrix $\bm{E}$ as the matrix which interpolates from volume nodes to composite surface nodes.
+However, this version of $\bm{E}$ is fully dense, and evaluating the formulation (\ref{eq:esdg}) requires flux evaluations between each surface node and \textit{all} volume nodes, as illustrated in Figure~\ref{subfig:nonaligned}.
+This greatly increases computational costs, especially at high orders of approximation.
+%The proposed work will reduce the number of dyadic flux evaluations required for such composite quadratures.
+The increased cost of (\ref{eq:esdg}) under composite quadrature is due to the lack of sparsity in $\bm{E}$, which now interpolates from volume to \textit{composite} surface nodes.
+
+\begin{figure}
+\centering
+\subfloat[Flux evaluations required for aligned surface nodes]{\raisebox{.1em}{\includegraphics[width=.4\textwidth]{figs/aligned.png}}\label{subfig:aligned}}
+\hspace{4em}
+\subfloat[Flux evaluations required for non-aligned surface nodes]{\includegraphics[width=.39\textwidth]{figs/nonaligned.png}\label{subfig:nonaligned}}
+\caption{Nodes between which flux evaluations are required for Gauss nodes.
+Aligned surface nodes (conforming interfaces) require evaluations between each surface node and a line of volume nodes, while non-aligned surface nodes (non-conforming interfaces) require flux evaluations between a surface node and \textit{all} volume nodes.
+} %Figure~\ref{subfig:aligned} and \ref{subfig:nonaligned} illustrate when dyadic flux evaluations $\bm{f}^i_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}$ between nodes are necessary (indicated by dashed red lines) for conforming interfaces (Figure~\ref{subfig:aligned}) and non-conforming interfaces (Figure~\ref{subfig:nonaligned}) under a Gauss collocation scheme.
+Conforming interfaces (aligned surface nodes) require evaluations between a surface node and a line of volume nodes, while non-conforming interfaces (nonaligned composite surface nodes) require flux evaluations between a surface node and \textit{all} volume nodes.
+The proposed mortar approach (Figure~\ref{subfig:mortar}) reduces the number of dyadic flux evaluations by introducing composite nodes as mortars which couple only to surface nodes.
+}
+\label{fig:fluxsparsity}
+\end{figure}
+
+%This work is motivated by complications which arise when designing entropy stable couplings between elements which do not share the same boundary nodes.
 This can arise, for example, for hybrid and non-conforming meshes.
+%\paragraph{Hybrid meshes:} entropy stable DG methods on hybrid meshes were introduced in \cite{chan2019skew} using a skew-symmetric formulation.
+The resulting methods are stable for more arbitrary choices of surface quadrature, in particular when an SBP property may not hold.
+%
+%\paragraph{Non-conforming meshes}
 
-\paragraph{Hybrid meshes:} entropy stable DG methods on hybrid meshes were introduced in \cite{chan2018skew} using a skew-symmetric formulation.
-The resulting methods are stable for more arbitrary choices of surface quadrature, in particular when an SBP property may not hold.
+%On conforming meshes, it is most efficient to utilize both Gauss quadrature for volume integrals and Gauss quadrature for face or surface integrals.
+For solutions represented in terms of their values at tensor product volume Gauss nodes, extrapolation to face Gauss nodes can be done in an efficient line-by-line manner using one-dimensional interpolation matrices.
+%
+%For non-conforming meshes, it can be advantageous to use composite Gauss quadratures on non-conforming interfaces \cite{kozdon2018energy}.
+However, interpolating the solution at volume Gauss nodes to split-side Gauss nodes is no longer a one-dimensional operation.
 
-\paragraph{Non-conforming meshes}
+\section{An entropy stable mortar formulation}
+\label{sec:3}
 
-On conforming meshes, it is most efficient to utilize both Gauss quadrature for volume integrals and Gauss quadrature for face or surface integrals.
-For solutions represented in terms of their values at tensor product volume Gauss nodes, extrapolation to face Gauss nodes can be done in an efficient line-by-line manner using one-dimensional interpolation matrices.
-
-For non-conforming meshes, it can be advantageous to use composite Gauss quadratures on non-conforming interfaces.
-\note{Reference JK and LCW's paper on full-side vs split side mortars.}  However, interpolating the solution at volume Gauss nodes to split-side Gauss nodes is no longer a one-dimensional operation.
-
-\section{Mortar formulation}
-
-\begin{figure}[!h]
+The goal of this work is to reduce computational costs for Gauss collocation schemes in the presence of non-conforming interfaces.
+This can be done by treating composite quadrature nodes as a layer of ``mortar'' nodes which are coupled directly to surface nodes, but not directly to the volume nodes, as illustrated in Figure~\ref{fig:gqcon_noncon}.
+This results in modifications of the matrices involved in the entropy stable formulation (\ref{eq:esdg}).
+These modifications preserve both high order accuracy and entropy stablility, and yield an implementation which is identical to that of (\ref{eq:esdg}) except for a face-local correction to the numerical flux.
+\begin{figure}
 \centering
-\includegraphics[width=.65\textwidth]{figs/mortar.png}
-\caption{Illustration of mortar operators.
+\includegraphics[width=.6\textwidth]{figs/mortar.png}
+\caption{Illustration of mortar operators for a Gauss collocation scheme.
 The matrix $\bm{E}$ maps from volume quadrature points to surface quadrature points, $\bm{E}_m$ maps from surface to mortar surface points, and $\tilde{\bm{E}}_m$ maps from mortar surface points to surface points.
@@
 
-Let $\bm{V}_q$ and $\bm{V}_f$ denote interpolation matrices which evaluate at volume and surface quadrature points, respectively, and let $\bm{W}, \bm{W}_f$ denote diagonal matrices whose entries consist of volume and surface quadrature weights
+We introduce new matrices on the reference element $\hat{D}$.
+Let $\hat{\bm{x}}$ denote volume collocation points, and let $\hat{\bm{x}}_f$ denote face (surface) points on the boundary $\partial \hat{D}$.
+To simplify notation, we assume from this point onwards that the nodes are ordered face-by-face.
+In 2D, this implies that $\hat{\bm{x}}_f$ is
+\begin{equation}
+\hat{\bm{x}}_f = \begin{bmatrix}
+\hat{\bm{x}}_{f,1} &
+\hat{\bm{x}}_{f,2} &
+\hat{\bm{x}}_{f,3} &
+\hat{\bm{x}}_{f,4}
+\end{bmatrix}^T,
+\label{eq:facenodeordering}
+\end{equation}
+where $\hat{\bm{x}}_{f,j}$ denotes the vector of 1D nodal positions on the $j$th face of the reference quadrilateral $[-1,1]^2$.
+Recall that the matrix $\bm{E}$ as defined in (\ref{eq:EBdefs}) interpolates from volume collocation points $\hat{\bm{x}}$ to surface points $\hat{\bm{x}}_f$; the ordering (\ref{eq:facenodeordering}) simply corresponds to a permutation of the interpolation matrix $\bm{E}$.
+
+We now introduce a second set of $N_m$ mortar points $\hat{\bm{x}}_m$ on a face of the reference element.
+We assume these points also correspond to quadrature nodes with corresponding weights $\bm{w}_m$.
+For example, mortar nodes can be composite Gauss or Lobatto quadrature nodes (for $h$ non-conforming interfaces), higher degree Gauss or Lobatto nodes (for $p$ non-conforming interfaces), or identical to the surface nodes $\hat{\bm{x}}_f$ (non-mortar interface).
+For simplicity of notation, we assume that each face has the same set of mortar nodes; however, it is straightforward to extend this to the case when the mortar nodes vary face-by-face.
+
+We define the interpolation matrix $\bm{E}_{mf}$ as the operator which maps values at surface nodes to values at mortar nodes.
+In 2D, since the face of a quadrilateral is a 1D line, we can define $\bm{E}_{mf}$ as the block matrix acting on the surface nodes on the four faces
 \[
-\text{\note{add defs of matrices}}
+\bm{E}_{mf} = \bm{I}_{4\times 4}\otimes \bm{E}_m,
 \]
+where $\bm{I}_{4\times 4}$ is the 4-by-4 identity matrix and $\bm{E}_m$ is the mortar interpolation matrix over the reference face (interval) $[-1,1]$.
+The matrix $\bm{E}_m$ is defined in terms of $\hat{x}_{m,i}$, the mortar nodes mapped to the reference interval $[-1,1]$,
+\[
+\bm{E}_{mf} = \ell_j(\hat{x}_{m,i}), \qquad 1\leq i \leq N_m, \quad 1\leq j \leq (N+1).
+\]
+We can define an analogous matrix $\bm{E}_{fm}$ which maps from the mortar nodes back to the surface nodes.
+We first define the face mass matrix as the block diagonal matrix whose blocks are 1D diagonal mass matrices
+\[
+\bm{M}_f = \bm{I}_{4\times 4} \otimes \bm{M}_{\rm 1D}.
+\]
+These matrices are defined in 2D for simplicity, but are straightforward to extend to 3D.  
+The matrix $\bm{E}_{fm}$ can now be defined through a quadrature-based $L^2$ projection 
+\[
+\bm{E}_{fm} = \bm{M}_f^{-1}\bm{E}_{mf}^T\bm{M}_m, \qquad \bm{M}_m = \bm{I}_{4\times 4}\otimes \diag{\bm{w}_m}.
+\]
+Finally, we introduce diagonal boundary matrices on surface and mortar nodes 
+\begin{equation}
+\hat{\bm{B}}_{i,f} = \diag{\hat{\bm{n}}_{i,f}}\bm{M}_f %\circ \bm{w}_f}, 
+\qquad \hat{\bm{B}}_{i,m} = \diag{\hat{\bm{n}}_{i,m}}\bm{M}_m,% \circ \bm{w}_m},
+\label{eq:Bi}
+\end{equation}
+where $\hat{\bm{n}}_{i,f}, \hat{\bm{n}}_{i,m}$ are vectors containing components of the scaled reference normals at face and mortar points, respectively.
 
-Let $\bm{E} = \bm{V}_f\bm{P}_q$ denote the polynomial mapping which extrapolates values from volume quadrature points to values at surface quadrature points.
-We also define $\bm{B}_i$ as the diagonal matrix containing products of quadrature weights and the $i$th component of the outward normal vector $\hat{\bm{n}}_i$
+Note that $\bm{E}_{fm}$ exactly recovers polynomials of a certain degree on the reference face, where the degree of the polynomial is related to the accuracy of the surface and mortar quadratures.
+\begin{lemma}
+\label{lemma:Efm}
+Suppose the face (surface) quadrature is exact for polynomials of degree $N+M_f$ and mortar quadratures are exact for degree $N+M_m$ polynomials.
+Then, $\bm{E}_{fm}$ exactly recovers polynomials of degree $\min(N,M_f,M_m)$.
+\end{lemma}
+\begin{proof}
+Let $u(\bm{x})$ be a polynomial of degree $\min(N,M_f,M_m)$ or less, and let $\bm{u}_f$ be its values on the face (surface) nodes.
+Then, $\bm{u}_m = \bm{E}_{mf}\bm{u}_f$ are the interpolated values of the polynomial on the mortar nodes.
+Applying $\bm{E}_{fm}$ yields
 \[
-\bm{B}_i = \bm{W}_f \diag{\hat{\bm{n}}_i}.
+\bm{E}_{fm}\bm{u}_m = \bm{M}_f^{-1}\bm{E}_{mf}^T\bm{M}_m\bm{E}_{mf}\bm{u}_f.
 \]
+The entries of $\bm{E}_{mf}^T\bm{M}_m\bm{E}_{mf}$ are integrals of products of Lagrange basis functions with $u(\bm{x})$, where all integrals are approximated using mortar quadrature.
 
-An entropy stable skew-symmetric formulation can be given on the reference element $\hat{D}$ as follows:
+Since $u$ is degree $\min(N,M_f,M_m)$ and each Lagrange basis function is degree $N$, the integrand is computed exactly under the mortar quadrature.
+Moreover, since $u$ is degree $\min(N,M_f,M_m)$, it is also computed exactly using the face (surface) quadrature.
+Thus, $\bm{E}_{mf}^T\bm{M}_m\bm{E}_{mf} = \bm{M}_f \bm{u}_f$ by exactness of the face and mortar quadratures, and $\bm{E}_{fm}\bm{u}_m = \bm{u}_f$.
+\qed\end{proof}
+
+\subsection{Mortar-based hybridized SBP operators}
+
+The matrices $\bm{E}_{mf}, \bm{E}_{fm}, \hat{\bm{B}}_{i,f}, \hat{\bm{B}}_{i,m}$ can now be used to construct SBP operators which involve mortar nodes.
+We define the mortar-based hybridized SBP operator on the reference element $\hat{\bm{Q}}_{i,m}$ to be
+\begin{align}
+\hat{\bm{Q}}_{i,m} = \frac{1}{2}\begin{bmatrix}
+\hat{\bm{Q}}_i - \hat{\bm{Q}}_i^T & \bm{E}^T\hat{\bm{B}}_{i,f} & \\
+-\hat{\bm{B}}_{i,f}\bm{E} & & \hat{\bm{B}}_{i,f} \bm{E}_{fm}\\
+& -\hat{\bm{B}}_{i,m} \bm{E}_{mf} & \hat{\bm{B}}_{i,m}
+\end{bmatrix}.
+\label{eq:mhsbp}
+\end{align}
+
+Since it is not clear at first glance how the hybridized operator $\hat{\bm{Q}}_{i,h}$ or its mortar-based variant $\hat{\bm{Q}}_{i,m}$ can be used to perform differentiation, we review intuitive explanations of each operator.
+Let $\bm{u}$ denote basis coefficients for some function $u(\bm{x})$, and let $f, g$ denote two functions.
+Suppose $\bm{u}$ satisfies the following matrix system involving the hybridized operator $\hat{\bm{Q}}_{i,h}$
+\begin{align}
+\hat{\bm{M}}\bm{u} = \begin{bmatrix}
+\bm{I}\\
+\bm{E}
+\end{bmatrix}^T \diag{\bm{f}}\hat{\bm{Q}}_{i,h} \diag{\bm{g}}, \qquad \bm{f} = \begin{bmatrix} f(\hat{\bm{x}}) \\ f(\hat{\bm{x}}_f)\end{bmatrix}, \qquad \bm{g} = \begin{bmatrix} g(\hat{\bm{x}}) \\ g(\hat{\bm{x}}_f)\end{bmatrix}.
+\label{eq:Qh}
+\end{align}
+It was shown in \cite{chan2017discretely, chan2019skew} that this corresponds to a high order accurate approximation of $f\pd{g}{x_i}$.
+Let $\bm{v}$ denote basis coefficients for some arbitrary degree $N$ polynomial test function.
+Multiplying (\ref{eq:Qh}) by $\bm{v}^T$, expanding out the different blocks of $\bm{Q}_{i,h}$, and using generalized summation by parts yields
+\begin{align*}
+\bm{M}\bm{u} &= 
+\diag{\bm{f}}\bm{Q}_i\bm{g} + \frac{1}{2}\LRp{\diag{f(\hat{\bm{x}})}\bm{E}^T + \bm{E}^T f(\hat{\bm{x}}_f)}\bm{B}_i \LRp{g(\hat{\bm{x}}_f) - \bm{E} g(\hat{\bm{x}})}.
+\end{align*}
+The latter expression can be interpreted as a boundary correction term.
+Since $\bm{E}$ is a high order accurate boundary interpolation operator, this correction term vanishes if $g$ is a degree $N$ polynomial.
+%Replacing quadrature rules by the corresponding integrals, we can re-interpret (\ref{eq:Qh}) as the following variational problem: find a degree $N$ polynomial $u(\bm{x})$ such that, for all polynomials $v(\bm{x})$ of degree $N$,
+%\begin{align*}
+%\int_{\hat{D}} u(\bm{x}) v(\bm{x}) &= \frac{1}{2} \int_{\hat{D}} \LRp{f\pd{\Pi_N (g)}{x_i}v - g\pd{\Pi_N(fv)}{x_i}} \\
+%&+ \frac{1}{2}\int_{\partial \hat{D}} \LRp{g \Pi_N(fv) + \LRp{g - \Pi_N (g)} fv}\hat{n}_i
+%\end{align*}
+%where $\Pi_N$ denotes the (quadrature-based) $L^2$ projection operator onto degree $N$ polynomials on $\hat{D}$.  Integrating by parts yields
+%\begin{align*}
+%\int_{\hat{D}} u(\bm{x}) v(\bm{x}) &= \int_{\hat{D}} f\pd{\Pi_N (g)}{x_i}v + \int_{\partial \hat{D}} \LRp{g - \Pi_N(g)}\frac{\LRp{fv+\Pi_N(fv)}}{2}\hat{n}_i
+%\end{align*}
+%Thus, we can interpret $u(\bm{x})$ as the $L^2$ projection of $f\pd{\Pi_N (g)}{x_i}$ with additional boundary correction terms.
+These boundary correction terms vanish if $g$ is a polynomial of degree $N$.
+
+The mortar-based hybridized SBP operator can be interpreted similarly.
+Recall that $\bm{E}_{mf}$ interpolates from face nodes to mortar nodes.
+Thus, the matrix $\bm{E}_{mf}\bm{E}$ interpolates from volume nodes to mortar nodes.
+We can use this matrix to replicate (\ref{eq:Qh}) for the mortar-based operator $\hat{\bm{Q}}_{i,m}$.
+Let $\bm{u}$ solve the following system
+\begin{align}
+\hat{\bm{M}}\bm{u} = \begin{bmatrix}
+\bm{I}\\
+\bm{E}\\
+\bm{E}_{mf}\bm{E}
+\end{bmatrix}^T \diag{\bm{f}}\hat{\bm{Q}}_{i,m} \diag{\bm{g}}, \qquad \bm{f} = \begin{bmatrix} f(\hat{\bm{x}}) \\ f(\hat{\bm{x}}_f)\\ f(\hat{\bm{x}}_m)\end{bmatrix}, \quad \bm{g} = \begin{bmatrix} g(\hat{\bm{x}}) \\ g(\hat{\bm{x}}_f)\\ g(\hat{\bm{x}}_m)\end{bmatrix}.
+\label{eq:Qm}
+\end{align}
+Expanding out terms yields a similar expression involving multiple correction terms
+\begin{align}
+\bm{M}\bm{u} &= \diag{\bm{f}}\bm{Q}_i\bm{g} \label{eq:mhsbp_correction}\\
+&+ \frac{1}{2} \diag{f(\hat{\bm{x}})}\bm{E}^T\bm{B}_{i,f}\LRp{g(\hat{\bm{x}}_f) - \bm{E}g(\hat{\bm{x}})} \nonumber\\
+&+ \frac{1}{2} \bm{E}^T\diag{f(\hat{\bm{x}}_f)}\bm{B}_{i,f}\LRp{\bm{E}_{fm}g(\hat{\bm{x}}_m) - \bm{E}g(\hat{\bm{x}})} \nonumber\\
+&+ \frac{1}{2} \bm{E}^T\bm{E}_{mf}^T\diag{f(\hat{\bm{x}}_m)}\bm{B}_{i,m}\LRp{g(\hat{\bm{x}}_m) - \bm{E}_{mf}g(\hat{\bm{x}}_f)} \nonumber
+\end{align}
+We can show that these correction terms vanish for polynomial functions $g$.  %when $f = 1$ and $g$ is a degree $N$ polynomial.
+
+\begin{lemma}
+Suppose that the face (surface) quadrature is exact for polynomials of degree $N+M_f$ and mortar quadratures are exact for degree $N+M_m$ polynomials.
+Let $f = 1$ and $g$ be a degree $\min(N,M_f,M_m)$ polynomial.
+Then, the equations (\ref{eq:Qm}) and (\ref{eq:mhsbp_correction}) are exact.
+\end{lemma}
+\begin{proof}
+Following \cite{chan2017discretely}, if $f = 1$  and $g(\bm{x})$ is a degree $N$ polynomial, then  (\ref{eq:mhsbp_correction}) is exact if the correction terms vanish.
+The first and third correction terms are zero by the fact that $\bm{E}, \bm{E}_{mf}$ are degree $N$ interpolation operators, and the second correction term vanishes by Lemma~\ref{lemma:Efm}.
+\qed\end{proof}
+
+Finally, we note that the mortar-based hybridized SBP operators satisfy a summation by parts property.
+\begin{lemma}
+\label{lemma:mhsbp}
+Let $\bm{Q}_{i,m}$ be defined as in (\ref{eq:mhsbp}).  Then, 
 \[
-\bm{M}\td{\bm{u}_N}{t} + \begin{bmatrix} \bm{V}_q \\ \bm{V}_f \end{bmatrix}^T
-\LRp{\begin{bmatrix}
-\bm{Q}_i-\bm{Q}_i^T & \bm{E}^T\bm{B}_i\\
--\bm{B}_i\bm{E} & \bm{0}
-\end{bmatrix} \circ \bm{F}_S}\bm{1} + \bm{V}_f^T\bm{B}_i\bm{f}_i^* = 0.  
+\bm{Q}_{i,m} + \bm{Q}_{i,m}^T = \begin{bmatrix}
+\bm{0} && \\
+&\bm{0} &\\
+&& \bm{B}_{i,m}\end{bmatrix}, \qquad \bm{Q}_{i,m}\bm{1} = \bm{0}.
 \]
+\end{lemma}
+\begin{proof}
+The first property holds if 
+\begin{equation}
+\hat{\bm{B}}_{i,f} \bm{E}_{fm} = (\hat{\bm{B}}_{i,m} \bm{E}_{mf})^T = \bm{E}_{mf}^T\hat{\bm{B}}_{i,m},  
+\label{eq:mhsbpprop}
+\end{equation}
+where we have used that $\hat{\bm{B}}_{i,m}$ is diagonal.  Recall that $\bm{E}_{fm} = \bm{M}_f^{-1}\bm{E}_{mf}^T\bm{M}_m$.  
+Then, by the definition of $\bm{B}_{i,m},\bm{B}_{i,h}$ in (\ref{eq:Bi}), we have that
+\begin{align*}
+\hat{\bm{B}}_{i,f} \bm{E}_{fm} = \diag{\hat{\bm{n}}_{i,f}}\bm{M}_f \bm{M}_f^{-1} \bm{E}_{mf}^T\bm{M}_m 
+= \diag{\hat{\bm{n}}_{i,f}} \bm{E}_{mf}^T\bm{M}_m.
+\end{align*}
+where we have used that $\bm{M}_m$ is diagonal.
+Then, since the scaled outward normals $\hat{n}_i$  are constant over each face of the reference element and $\bm{E}_{mf}, \bm{M}_m$ are block matrices (with each block corresponding to a face), we have that
+\[
+\bm{E}_{mf}^T\bm{M}_m \diag{\hat{\bm{n}}_{i,m}}= \bm{E}_{mf}^T\bm{B}_{i,m}.
+\]  
+\qed\end{proof}
 
-We incorporate mortars by modifying this formulation.
+We can now construct an entropy stable formulation on the reference element using the mortar-based hybridized SBP operator
+\begin{gather}
+\hat{\bm{M}}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \\ \bm{E}_{mf}\bm{E} \end{bmatrix}^T
+\sum_{i=1}^d \LRp{\hat{\bm{Q}}_{i,m} \circ \bm{F}_i}\bm{1} + \bm{E}_{mf}^T\hat{\bm{B}}_{i,m}\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_m)} = 0 \label{eq:esdgmref}\\
+\tilde{\bm{u}} = \begin{bmatrix}
+\bm{u}_h\\
+\tilde{\bm{u}}_f\\
+\tilde{\bm{u}}_m
+\end{bmatrix}, \qquad \tilde{\bm{u}}_m = \bm{u}\LRp{\bm{E}_{mf}\bm{E}\bm{v}(\bm{u}_h)}, \qquad \bm{f}^* = \bm{f}_S\LRp{\tilde{\bm{u}}_m^+,\tilde{\bm{u}}_m},
+\end{gather}
+where $\LRp{\bm{F}_i}_{jk} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}$ for all indices $j,k$ of $\tilde{\bm{u}}$.  This can be extended to mapped elements
+
+%An entropy stable skew-symmetric formulation can be given on the reference element $\hat{D}$ as follows:
+%\[
+%\bm{M}\td{\bm{u}_N}{t} + \begin{bmatrix} \bm{V}_q \\ \bm{E} \end{bmatrix}^T
+%\LRp{\begin{bmatrix}
+%\bm{Q}_i-\bm{Q}_i^T & \bm{E}^T\bm{B}_i\\
+%-\bm{B}_i\bm{E} & \bm{0}
+%\end{bmatrix} \circ \bm{F}_S}\bm{1} + \bm{E}^T\bm{B}_i\bm{f}_i^* = 0.  
+%\]
+%
+%We incorporate mortars by modifying this formulation.
 Let $\bm{V}_m$ denote the matrix which maps volume quadrature points to values at surface \textit{mortar} points.
@@
 We can then define surface and mortar mass and projection matrices
-\begin{align*}
-\bm{M}_m = \bm{T}_m^T\bm{W}_m\bm{T}_m, \qquad \bm{P}_m = \bm{M}_m^{-1}\bm{T}_m^T\bm{W}_m\\
-\bm{M}_f = \bm{T}_f^T\bm{W}_f\bm{T}_f, \qquad \bm{P}_f = \bm{M}_m^{-1}\bm{T}_f^T\bm{W}_f.
-\end{align*}
-Note that we have used the mortar mass matrix $\bm{M}_m$ in the above definition of the face projection.
-This is only necessary if the surface and mortar quadratures do not exactly integrate degree $2N$ polynomials.
+%\begin{align*}
+%\bm{M}_m = \bm{T}_m^T\bm{M}_m\bm{T}_m, \qquad \bm{P}_m = \bm{M}_m^{-1}\bm{T}_m^T\bm{M}_m\\
+%\bm{M}_f = \bm{T}_f^T\bm{W}_f\bm{T}_f, \qquad \bm{P}_f = \bm{M}_m^{-1}\bm{T}_f^T\bm{W}_f.
+%\end{align*}
+%Note that we have used the mortar mass matrix $\bm{M}_m$ in the above definition of the face projection.
+%This is only necessary if the surface and mortar quadratures do not exactly integrate degree $2N$ polynomials.
 If both the surface and mortar quadratures are exact for polynomials of degree $2N$, then $\bm{M}_m = \bm{M}_f$, and no distinction is necessary between the two mass matrices.
-
-\begin{remark}
-In order to show stability, we require that both $\bm{P}_f$ and $\bm{P}_m$ are defined using the same mass matrix.
+%
+%\begin{remark}
+%In order to show stability, we require that both $\bm{P}_f$ and $\bm{P}_m$ are defined using the same mass matrix.
 We have used the mortar mass matrix $\bm{M}_m$ here; however, we could also use the surface mass matrix $\bm{M}_f$ in both $\bm{P}_f$ and $\bm{P}_m$.
 We use the mortar mass matrix in this work, as the mortar quadrature is generally more accurate than the surface quadrature for our use cases (e.g.\ hybrid and non-conforming meshes).
-\end{remark}
+%\end{remark}
 
-We can now define operators which map between surface and mortar quadrature points.
+%
+%We can now define operators which map between surface and mortar quadrature points.
 Let $\bm{E}_m$ denote the map from surface to mortar points, and let $\tilde{\bm{E}}_m $ denote the map from mortar to surface points.
 Both operators are defined through an $L^2$ projection to the trace space and interpolation to appropriate points
-\[
-\bm{E}_m = \bm{T}_m \bm{P}_f, \qquad \tilde{\bm{E}}_m = \bm{T}_f \bm{P}_m.
-\]  
-We also define a mortar boundary matrix
-\[
-\tilde{\bm{B}}_i = \bm{W}_m\diag{\hat{\bm{n}}_i}.
-\]
-Then, a mortar-based formulation can be given as follows:
-\begin{equation}
-\bm{M}\td{\bm{u}_N}{t} + \begin{bmatrix} \bm{V}_q \\ \bm{V}_f \\ \bm{V}_m \end{bmatrix}^T
-\LRp{\begin{bmatrix}
-\bm{Q}_i-\bm{Q}_i^T & \bm{E}^T\bm{B}_i &\\
--\bm{B}_i\bm{E} &  & \bm{B}_i\tilde{\bm{E}}_m \\
-& -\tilde{\bm{B}}_i{\bm{E}}_m & 
-\end{bmatrix} \circ \bm{F}_S}\bm{1} + \bm{V}_m^T\tilde{\bm{B}}_i\bm{f}_i^* = 0.  
-\label{eq:mortar}
-\end{equation}
-Here, we have appended an extra row and column to the decoupled SBP matrix, and the inter-element numerical flux $\bm{f}^*$ is now computed in terms of the \emph{mortar} nodes.
-
-We first note that
-\begin{align*}
-\bm{B}_i\tilde{\bm{E}}_m &= \diag{\bm{W}_f} \bm{T}_f \bm{P}_m = \diag{\hat{\bm{n}}} \bm{W}_f \bm{T}_f \bm{M}_m^{-1}\bm{T}_m\bm{W}_m\\
-&=  \bm{W}_f \bm{T}_f \bm{M}_m^{-1}\bm{T}^T_m\bm{W}_m\diag{\hat{\bm{n}}_i} = \bm{P}_f^T\bm{T}^T_m \tilde{\bm{B}}_i = \LRp{\tilde{\bm{B}}_i{\bm{E}}_m}^T
-\end{align*}
-where we have used that $\bm{W}_m$ is diagonal and that the outward normal $\hat{\bm{n}}_i$ on the reference element is constant over each face.
-This implies skew-symmetry of the decoupled SBP operator and entropy stability of the overall system.
-
-\note{Finish}
+%\[
+%\bm{E}_m = \bm{T}_m \bm{P}_f, \qquad \tilde{\bm{E}}_m = \bm{T}_f \bm{P}_m.
+%\]  
+%We also define a mortar boundary matrix
+%\[
+%\tilde{\bm{B}}_i = \bm{M}_m\diag{\hat{\bm{n}}_i}.
+%\]
+%Then, a mortar-based formulation can be given as follows:
+%\begin{equation}
+%\bm{M}\td{\bm{u}_N}{t} + \begin{bmatrix} \bm{V}_q \\ \bm{E} \\ \bm{V}_m \end{bmatrix}^T
+%\LRp{\begin{bmatrix}
+%\bm{Q}_i-\bm{Q}_i^T & \bm{E}^T\bm{B}_i &\\
+%-\bm{B}_i\bm{E} &  & \bm{B}_i\tilde{\bm{E}}_m \\
+%& -\tilde{\bm{B}}_i{\bm{E}}_m & 
+%\end{bmatrix} \circ \bm{F}_S}\bm{1} + \bm{V}_m^T\tilde{\bm{B}}_i\bm{f}_i^* = 0.  
+%\label{eq:mortar}
+%\end{equation}
+%Here, we have appended an extra row and column to the hybridized SBP matrix, and the inter-element numerical flux $\bm{f}^*$ is now computed in terms of the \emph{mortar} nodes.
 
@@
 However, using properties of the mortar matrices $\bm{E}_m, \tilde{\bm{E}}_m$, we can rewrite the above formulation in a way which reflects a traditional mortar-based finite element implementation.
-In other words, we wish to implement (\ref{eq:mortar}) such that the only modification from the implementations in \cite{chan2017discretely} is a pre-processing step on mortar faces.
+In other words, we wish to implement (\ref{eq:mortar}) such that the only modification from the implementations in \cite{chan2017discretely} is an intermediate step on mortar faces.
 
-We first note that, since $\bm{E}_m$ maps from surface nodes to mortar nodes, $\bm{V}_m = \bm{E}_m \bm{V}_f$.  Thus, we can rewrite the numerical flux contribution as
+We first note that, since $\bm{E}_m$ maps from surface nodes to mortar nodes, $\bm{V}_m = \bm{E}_m \bm{E}$.  Thus, we can rewrite the numerical flux contribution as
 \begin{align*}
- \bm{V}_m^T\tilde{\bm{B}}_i\bm{f}_i^* &= \bm{V}_f^T \bm{E}_m^T \bm{W}_m\diag{\hat{\bm{n}}_i} \bm{f}_i^* = \bm{V}_f^T  \bm{P}_f^T\bm{T}_m^T\bm{W}_m\diag{\hat{\bm{n}}_i} \bm{f}_i^*\\
- &= \bm{V}_f^T \bm{W}_f \diag{\hat{\bm{n}}_i} \bm{T}_m\bm{M}_m^{-1}\bm{T}_m^T\bm{W}_m \bm{f}_i^* =  \bm{V}_f^T \bm{B}_i \tilde{\bm{E}}_m\bm{f}_i^*.
+ \bm{V}_m^T\tilde{\bm{B}}_i\bm{f}_i^* &= \bm{E}^T \bm{E}_m^T \bm{M}_m\diag{\hat{\bm{n}}_i} \bm{f}_i^* = \bm{E}^T  \bm{P}_f^T\bm{T}_m^T\bm{M}_m\diag{\hat{\bm{n}}_i} \bm{f}_i^*\\
+ &= \bm{E}^T \bm{W}_f \diag{\hat{\bm{n}}_i} \bm{T}_m\bm{M}_m^{-1}\bm{T}_m^T\bm{M}_m \bm{f}_i^* =  \bm{E}^T \bm{B}_i \tilde{\bm{E}}_m\bm{f}_i^*.
 \end{align*}
@@
 \begin{align*}
-&\begin{bmatrix} \bm{V}_q \\ \bm{V}_f \\ \bm{V}_m \end{bmatrix}^T
+&\begin{bmatrix} \bm{V}_q \\ \bm{E} \\ \bm{V}_m \end{bmatrix}^T
 \LRp{\begin{bmatrix}
@@
 &= 
-\begin{bmatrix} \bm{V}_q \\ \bm{V}_f \end{bmatrix}^T
+\begin{bmatrix} \bm{V}_q \\ \bm{E} \end{bmatrix}^T
 \LRp{\begin{bmatrix}
@@
 \end{bmatrix} \circ \bm{F}_S}\bm{1} \\
-&+ \bm{V}_f^T \LRp{\bm{B}_i\tilde{\bm{E}}_m\circ \bm{F}_S^{sm}}\bm{1} - \bm{V}_m^T \LRp{ \tilde{\bm{B}}_i\bm{E}_m \circ \bm{F}_S^{ms}}\bm{1}
+&+ \bm{E}^T \LRp{\bm{B}_i\tilde{\bm{E}}_m\circ \bm{F}_S^{sm}}\bm{1} - \bm{V}_m^T \LRp{ \tilde{\bm{B}}_i\bm{E}_m \circ \bm{F}_S^{ms}}\bm{1}
 \end{align*}
@@
 \begin{align*}
- \bm{V}_f^T \LRp{\bm{B}_i\tilde{\bm{E}}_m\circ \bm{F}_S^{sm}}\bm{1} &=  \bm{V}_f^T \bm{B}_i \LRp{\tilde{\bm{E}}_m\circ \bm{F}_S^{sm}}\bm{1}\\
-\bm{V}_m^T \LRp{ \tilde{\bm{B}}_i\bm{E}_m \circ \bm{F}_S^{ms}}\bm{1} &= \bm{V}_m^T \tilde{\bm{B}}_i \LRp{ \bm{E}_m \circ \bm{F}_S^{ms}}\bm{1} = \bm{V}_f^T \bm{B}_i \tilde{\bm{E}}_m \LRp{ \bm{E}_m \circ \bm{F}_S^{ms}}\bm{1}
+ \bm{E}^T \LRp{\bm{B}_i\tilde{\bm{E}}_m\circ \bm{F}_S^{sm}}\bm{1} &=  \bm{E}^T \bm{B}_i \LRp{\tilde{\bm{E}}_m\circ \bm{F}_S^{sm}}\bm{1}\\
+\bm{V}_m^T \LRp{ \tilde{\bm{B}}_i\bm{E}_m \circ \bm{F}_S^{ms}}\bm{1} &= \bm{V}_m^T \tilde{\bm{B}}_i \LRp{ \bm{E}_m \circ \bm{F}_S^{ms}}\bm{1} = \bm{E}^T \bm{B}_i \tilde{\bm{E}}_m \LRp{ \bm{E}_m \circ \bm{F}_S^{ms}}\bm{1}
 \end{align*}
@@
 \bm{M}\td{\bm{u}_N}{t} &+ 
-\begin{bmatrix} \bm{V}_q \\ \bm{V}_f \end{bmatrix}^T
+\begin{bmatrix} \bm{V}_q \\ \bm{E} \end{bmatrix}^T
 \LRp{\begin{bmatrix}
@@
 -\bm{B}_i\bm{E} & \\
-\end{bmatrix} \circ \bm{F}_S}\bm{1} + \bm{V}_f^T\bm{B}_i \tilde{\bm{f}}^*_i = 0 \label{eq:mortar2}\\
+\end{bmatrix} \circ \bm{F}_S}\bm{1} + \bm{E}^T\bm{B}_i \tilde{\bm{f}}^*_i = 0 \label{eq:mortar2}\\
 \tilde{\bm{f}}^*_i &= \tilde{\bm{E}}_m\bm{f}^*_i + \LRp{\tilde{\bm{E}}_m\circ \bm{F}_S^{sm}}\bm{1} - \tilde{\bm{E}}_m \LRp{ \bm{E}_m \circ \bm{F}_S^{ms}}\bm{1} 
@@
 \end{enumerate}
-%\bm{M}\td{\bm{u}_N}{t} + \begin{bmatrix} \bm{V}_q \\ \bm{V}_f \\ \bm{V}_m \end{bmatrix}^T
-%\LRp{\begin{bmatrix}
-%\bm{Q}_i-\bm{Q}_i^T & \bm{E}^T\bm{B}_i &\\
-%-\bm{B}_i\bm{E} &  & \bm{B}_i\tilde{\bm{E}}_m \\
-%& -\tilde{\bm{B}}_i{\bm{E}}_m & 
-%\end{bmatrix} \circ \bm{F}_S}\bm{1} + \bm{V}_m^T\tilde{\bm{B}}_i\bm{f}_i^* = 0.
 
-%It is possible to efficiently interpolate from volume Gauss nodes to split-side Gauss nodes in two steps.
-We first interpolate to full-side Gauss nodes on the face, then interpolate from full-side  to split-side Gauss nodes.
-Each of these two operations can be done using only one-dimensional interpolation matrices, implying that the interpolation matrix from volume to split-side Gauss nodes can be decomposed into
-%\[
-%\bm{V}_{\rm vol-to-split} = \bm{V}_{\rm vol-to-full} \bm{V}_{\rm full-to-split}
-%\]
-%where each of the matrices in the decomposition possess either a sparse or Kronecker product structure.  
-%
-%ESDG schemes based on decoupled SBP operators introduced in \cite{chan2017discretely} assume a single interpolation matrix which extrapolates from volume to surface nodes.
-These schemes require dyadic flux evaluations between all volume and surface nodes connected by this interpolation matrix.
-For non-conforming interfaces with split-side Gauss nodes on the face, this necessitates dyadic flux computations between all volume nodes and all split-side Gauss nodes.
-%
-%We show in this section how to construct an entropy stable Gauss collocation scheme based on two layers of surface nodes: full-side and split-side nodes.
-These surface nodes will be loosely through a skew-symmetric coupling term involving polynomial interpolation matrices between full-side and split-side nodes.
-This scheme requires dyadic flux computations between volume and full-side surface nodes, and between full-side and split-side surface nodes.
-Each set of dyadic flux computations can be reduced to
-
-%We first introduce notation which is helpful for operations involving surface traces.  Let the matrix $\bm{R}$ denote the matrix which maps coefficients of a function to coefficients of its trace.  
-%
-%Let ${\bm{V}_f}\tilde{P}_f$ denote the matrix which projects to the trace space using split-side quadrature and evaluates at full-side quadrature points.
-Similarly, let $\tilde{\bm{V}_f}{P}_f$ denote the matrix which projects to the trace space using full-side quadrature and evaluates at split-side quadrature points
-%\begin{align*}
-%\bm{C}^i_f &= \begin{pmatrix}
-%\bm{0} & {\rm diag}\LRp{\hat{\bm{n}}_i}{\bm{V}_f}\tilde{\bm{P}}_f\\
-%-\tilde{\bm{V}_f}{\bm{P}}_f{\rm diag}\LRp{\hat{\bm{n}}_i} &\bm{0}
-%\end{pmatrix}, \\
-%\bm{S}^i_f &= \begin{pmatrix}
-%\diag{\bm{w}_f} & \\
-%& \diag{\tilde{\bm{w}}_f}
-%\end{pmatrix}\bm{C}^i_f.
-%\end{align*}
-%It is straightforward to show that the matrix $\bm{S}_f$ is skew-symmetric.
-%
-%%{fig:gqcon_noncon}
-%Let $\tilde{\bm{F}}_S$ denote the flux matrix, then 
-%\[
-%\begin{bmatrix}\bm{L}_f & \tilde{\bm{L}}_f\end{bmatrix} \LRp{\bm{C}^i_f\circ\tilde{\bm{F}}_S}\bm{1}
-%\]
-%is a skew-symmetric correction term which ensures entropy stability and high order accuracy (through consistency up to quadrature order).   
-%
-%Since the difference is high order accurate, you can remove this correction term and still get high order accuracy.
-However, the spatial entropy RHS is no longer zero if you use inconsistent surface and coupling quadrature without the correction term (\note{prove this}).
-%
-%\note{Do we need 3 layers of surface nodes in 3D?  Interpolation to isotropic split-side nodes is a Kronecker product - does that reduce things down?  }
+\begin{figure}
+\centering
+\includegraphics[width=.55\textwidth]{mortar_coupling.png}
+\caption{Blah}
+\end{figure}
 
@@
 \note{Line DG approach can work, but it's flipped: we increase quadrature strength in the direction \emph{orthogonal} to derivative instead.
-May improve accuracy over GLL by giving equivalence w/skew form.}
+May improve accuracy over Lobatto by giving equivalence w/skew form.}
 
 \section{Numerical experiments}
+
+Figure~\ref{subfig:noncon} shows preliminary results for the 2D compressible Euler equations by the PI on a manually constructed non-conforming quadrilateral mesh, with volume, surface, and mortar quadratures constructed from one-dimensional Lobatto or Gauss quadrature rules.
+The PI has verified that this preliminary implementation of the mortar-based formulation is discretely entropy stable.
+A preliminary accuracy analysis for the skew-symmetric formulation (\ref{eq:esdgSkew}) also suggests that Lobatto quadrature should achieve an $O(h^N)$ rate of convergence and that Gauss quadrature should achieve an $O(h^{N+1})$ rate of convergence.
+Both predicted rates are confirmed in Figure~\ref{subfig:noncon}.
+
+\begin{figure}
+\centering
+\includegraphics[width=.475\textwidth]{figs/mesh_curved0.png}
+\hspace{.5em}
+\includegraphics[width=.475\textwidth]{figs/mesh_curved.png}
+\caption{Two $N=3$ meshes used in the refinement and entropy conservation studies.}
+\end{figure}
+
+\begin{figure}
+\centering
+%\subfloat[Coarse $N=3$ non-conforming affine mesh]{\raisebox{2.6em}{\includegraphics[width=.475\textwidth]{figs/mesh_affine.png}}}
+\subfloat[Affine mesh]{
+\begin{tikzpicture}
+\begin{loglogaxis}[
+    width=.49\textwidth,
+    xlabel={Mesh size $h$},
+    ylabel={$L^2$ errors}, 
+    xmax=3.5,
+    ymin=1e-4, ymax=5,
+    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
+    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
+    legend entries={Lobatto, Gauss}%, Lobatto-Gauss}    
+]
+\pgfplotsset{
+cycle list={{blue, mark=*}, {red, dashed ,mark=square*}}%,{black, dashdotted ,mark=triangle*}}
+}
+% N = 1
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1.66667,2.78317)(0.833333,2.09894)(0.416667,1.25478)(0.208333,0.490133)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(1.66667,2.81566)(0.833333,2.2122)(0.416667,1.44626)(0.208333,0.801638)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1.66667,2.30772)(0.833333,1.16697)(0.416667,0.303428)(0.208333,0.0576807)};
+
+% N = 2
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1.66667,1.51766)(0.833333,0.431437)(0.416667,0.0792017)(0.208333,0.0167253)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(1.66667,1.52643)(0.833333,0.436466)(0.416667,0.0801575)(0.208333,0.0167043)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1.66667,1.06169)(0.833333,0.158534)(0.416667,0.0208737)(0.208333,0.00274974)};
+
+% N = 3
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1.66667,0.68797)(0.833333,0.0821894)(0.416667,0.008229)(0.208333,0.000856128)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(1.66667,0.685655)(0.833333,0.0815906)(0.416667,0.00838271)(0.208333,0.000889147)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1.66667,0.455651)(0.833333,0.0328805)(0.416667,0.00233608)(0.208333,0.000143112)};
+
+% N = 4
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1.66667,0.242584)(0.833333,0.0170238)(0.416667,0.000922618)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(1.66667,0.241312)(0.833333,0.0169965)(0.416667,0.000928106)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1.66667,0.173237)(0.833333,0.00858261)(0.416667,0.00029501)};
+
+\node at (axis cs:2.4,2.8) {$N = 1$};
+\node at (axis cs:2.4,1.3) {$N = 2$};
+\node at (axis cs:2.4,.6) {$N = 3$};
+\node at (axis cs:2.4,.22) {$N = 4$};
+\end{loglogaxis}
+\end{tikzpicture}
+}
+\hspace{.1em}
+\subfloat[Curved mesh]{
+\begin{tikzpicture}
+\begin{loglogaxis}[
+    width=.49\textwidth,
+    xlabel={Mesh size $h$},
+%    ylabel={$L^2$ errors}, 
+    xmax=3.5,
+    ymin=1e-4, ymax=5,
+    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
+    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
+    legend entries={Lobatto, Gauss}%, Lobatto-Gauss}    
+]
+\pgfplotsset{
+cycle list={{blue, mark=*}, {red, dashed ,mark=square*}}%,{black, dashdotted ,mark=triangle*}}
+}
+% N = 1
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1.66667,2.84018)(0.833333,2.32278)(0.416667,1.44878)(0.208333,0.608642)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(1.66667,2.8287)(0.833333,2.36062)(0.416667,1.61841)(0.208333,0.909621)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1.66667,2.42938)(0.833333,1.40381)(0.416667,0.504184)(0.208333,0.114213)};
+
+% N = 2
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1.66667,1.88997)(0.833333,0.710146)(0.416667,0.136728)(0.208333,0.0196411)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(1.66667,1.88857)(0.833333,0.709214)(0.416667,0.135323)(0.208333,0.0196621)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1.66667,1.32118)(0.833333,0.310973)(0.416667,0.0369794)(0.208333,0.00457492)};
+
+% N = 3
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1.66667,1.03021)(0.833333,0.211902)(0.416667,0.0195018)(0.208333,0.00184924)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(1.66667,1.03557)(0.833333,0.211255)(0.416667,0.0196214)(0.208333,0.0018744)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1.66667,0.6875)(0.833333,0.0837258)(0.416667,0.0052072)(0.208333,0.000298318)};
+
+% N = 4
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1.66667,0.558299)(0.833333,0.0594828)(0.416667,0.00284697)};
+%\addplot+[semithick, mark options={solid, fill=markercolor}]
+%coordinates{(1.66667,0.557251)(0.833333,0.0593767)(0.416667,0.00284163)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1.66667,0.330743)(0.833333,0.0204084)(0.416667,0.000691875)};
+
+\node at (axis cs:2.4,3.25) {$N = 1$};
+\node at (axis cs:2.4,1.75) {$N = 2$};
+\node at (axis cs:2.4,.9) {$N = 3$};
+\node at (axis cs:2.4,.425) {$N = 4$};
+\end{loglogaxis}
+\end{tikzpicture}
+}
+
+\caption{Blah}
+\end{figure}
 
@@
 \bibliographystyle{unsrt}
-\bibliography{dg}
+\bibliography{refs}
 

### 2019-12-20 docs/ESDG_mortar/main.tex
updating paper

@@
 
-\author{Jesse Chan, David C.\ Del Rey Fernandez, Mario Bencomo}
-\title{Mortar-based entropy stable discontinuous Galerkin methods on non-conforming quadrilateral and hexahedral meshes}
+\author{Jesse Chan, Mario Bencomo, David C.\ Del Rey Fernandez}
+\title{Mortar-based entropy stable discontinuous Galerkin methods on non-conforming meshes}
 \titlerunning{Mortar-based entropy stable DG methods}
@@
 
-\begin{enumerate}
-\item Intro
-\item Gauss collocation schemes, conditions for entropy stability
-\item Increased cost on non-conforming meshes
-\item Mortar scheme and entropy stability
-\item Analysis of accuracy of Global SBP operator.
-Gauss nodes = exact for degree $N$, while Lobatto = exact for degree $(N-1)$.
-Approach in Friedrichs + DCDR for $hp$ meshes yields degree $(N-1)$ global operator.
-\item Extensions
-\begin{enumerate}
-\item Curved meshes
-\item $p$-nonconforming meshes
-\item General elements and why you'd want to implement this.
-\end{enumerate}
-\item Numerical experiments (2D vortex, 2D entropy conservation test, 3D convergence).
-Note that reduced rates of convergence are not observed for $N>1$ on $p$-nonconforming meshes, or on hybrid meshes with mixed quadrature.
-\end{enumerate}
-
 \section{Introduction}
-
-%\note{Boilerplate introduction on high order + stability}
-
 Discretely entropy stability has emerged as a methodology for designing high order schemes for nonlinear conservation laws.
@@
 \begin{equation}
-\bm{Q}_h + \bm{Q}_h^T = \begin{bmatrix}
+\bm{Q}_h + \bm{Q}_h^T = \bm{B}_h, \qquad \bm{B}_h =  \begin{bmatrix}
 \bm{0} & \\
@@
 \end{equation}
-where $\bm{E}$ does not appear in the block boundary matrix on the right hand side.  Additionally, it is straightforward to show that 
+where $\bm{E}$ does not appear in the block boundary matrix on the right hand side.
+Here, we have used $\bm{0}$ to denote a matrix or vector of zeros, where the size is inferred from context.
+We will adopt a similar convention for the vector of ones $\bm{1}$.
+
+The matrix $\bm{Q}_h$ also satisfies
 \[
-\bm{Q}_h\bm{1} = \bm{0}
+\bm{Q}_h\bm{1} = 
+\frac{1}{2}\begin{bmatrix}
+\bm{Q}\bm{1}-\bm{Q}^T\bm{1} + \bm{E}^T\bm{B}\bm{1}\\
+-\bm{B}\bm{E}\bm{1} - \bm{B}\bm{1}
+\end{bmatrix}= 
+\frac{1}{2}\begin{bmatrix}
+-\bm{Q}^T\bm{1} + \bm{E}^T\bm{B}\bm{1}\\
+\bm{0}
+\end{bmatrix}
 \]
-using the GSBP property and properties of $\bm{Q}$.
+where we have used that $\bm{E}\bm{1} = \bm{1}$ (since $\bm{E}$ is a high order accurate boundary interpolation matrix) and $\bm{Q}\bm{1} = \bm{0}$ (since $\bm{Q}$ is a differentiation matrix).
+Combining the GSBP property $-\bm{Q}^T + \bm{E}^T\bm{B}\bm{E} = \bm{Q}$ and $\bm{E}\bm{1} = \bm{1}$ then yields that
+\begin{equation}
+\bm{Q}_h\bm{1} = 
+\frac{1}{2}\begin{bmatrix}
+-\bm{Q}^T\bm{1} + \bm{E}^T\bm{B}\bm{1}\\
+\bm{0}
+\end{bmatrix} = 
+\frac{1}{2}\begin{bmatrix}
+\bm{Q}\bm{1}\\
+\bm{0}
+\end{bmatrix} 
+= \frac{1}{2}\begin{bmatrix}
+\bm{0}\\
+\bm{0}
+\end{bmatrix} = \bm{0}.
+\label{eq:Qh1}
+\end{equation}
 
@@
 \bm{M}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T
-\LRp{\bm{Q}_h \circ \bm{F}}\bm{1} + \bm{E}^T\bm{B}\LRp{\bm{f}^*-\bm{f}(\tilde{\bm{u}}_f)} = 0 \label{eq:esdg1D}\\ 
+\LRp{2\bm{Q}_h \circ \bm{F}}\bm{1} + \bm{E}^T\bm{B}\LRp{\bm{f}^*-\bm{f}(\tilde{\bm{u}}_f)} = 0 \label{eq:esdg1D}\\ 
 \tilde{\bm{u}} = \begin{bmatrix}
@@
 
-\note{Maybe add proof here and show how rest of proofs just rely on SBP and $\bm{Q}_h\bm{1} = \bm{0}$?}.
+\begin{theorem}
+Assuming continuity in time, the formulation (\ref{eq:esdg1D}) is entropy conservative in the sense that
+\[
+\bm{1}^T\bm{M}\td{S(\bm{u}_h)}{t} + \bm{1}^T\bm{B}\LRp{\bm{v}_f^T\bm{f}^* - \psi(\tilde{\bm{u}}_f)} = 0, \qquad \bm{v}_f =\bm{E}\bm{v}(\bm{u}_h).
+\]
+\end{theorem}
+\begin{proof}
+The proof uses similar techniques as proofs in other papers \cite{chen2017entropy, crean2018entropy, chan2017discretely, chan2019skew}, and we repeat it for completeness.
+We begin by testing (\ref{eq:esdg1D}) with the vector of entropy variables evaluated at nodal points $\bm{v}(\bm{u}_h)$.
+We apply the chain rule in time and use that $\bm{M}$ is diagonal to show that
+\[
+\bm{v}(\bm{u}_h)^T\bm{M}\td{\bm{u}}{t} = \bm{1}^T\bm{M}\LRp{\bm{v}(\bm{u}_h)\circ \td{\bm{u}_h}{t}} =  \bm{1}^T\bm{M}\td{S(\bm{u}_h)}{t}.
+\]
+For the volume term, we use the SBP property to arrive at
+\begin{align*}
+\bm{v}(\bm{u}_h)^T\begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T\LRp{2\bm{Q}_h \circ \bm{F}}\bm{1} &= \tilde{\bm{v}}^T\LRp{\LRp{\bm{Q}_h-\bm{Q}_h^T} \circ \bm{F}}\bm{1}  + \LRp{\bm{E}\bm{v}(\bm{u}_h)}^T\LRp{\bm{B}_h \circ\bm{F}}\bm{1},
+\end{align*}
+where we have introduced $\tilde{\bm{v}} = \LRp{\begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}\bm{v}(\bm{u}_h)}$.  
+Note that, by the consistency of the entropy conservative flux $\bm{f}_S$, the diagonal of $\bm{F}$ is the vector $\bm{f}(\tilde{\bm{u}}_f)$.
+Using that $\bm{B}$ is diagonal, the latter term reduces to
+\[
+\LRp{\bm{E}\bm{v}(\bm{u}_h)}^T\LRp{\bm{B}_h \circ\bm{F}}\bm{1} = \bm{v}_f^T \bm{B}\bm{f}(\tilde{\bm{u}}_f) =  \bm{1}^T \bm{B}\LRp{\bm{v}_f^T\bm{f}(\tilde{\bm{u}}_f)}.
+\]
+The final part of the proof uses that 
+\begin{align*}
+\tilde{\bm{v}}^T\LRp{\LRp{\bm{Q}_h-\bm{Q}_h^T} \circ \bm{F}}\bm{1} &= \sum_j {\tilde{\bm{v}}}_j \LRp{\bm{Q}_h-\bm{Q}_h^T}_{ij} \bm{F}_{ij}\\
+&= \sum_j \LRp{\bm{Q}_h}_{ij} \LRp{{\tilde{\bm{v}}}_i-{\tilde{\bm{v}}}_j}^T\bm{f}_{S}\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}\\
+&= \sum_j \LRp{\bm{Q}_h}_{ij} \LRp{\psi(\tilde{\bm{u}}_i)-\psi(\tilde{\bm{u}}_j)} = \tilde{\bm{u}}^T\bm{Q}_h\bm{1}- \bm{1}^T\bm{Q}_h\tilde{\bm{u}}
+\end{align*}
+The term $\bm{Q}_h\bm{1} = \bm{0}$ by (\ref{eq:Qh1}).
+The proof is completed by applying the SBP property (\ref{eq:hsbp}) to the remaining term, which reduces to $-\bm{1}^T\bm{B}_h\tilde{\bm{u}}=-\bm{1}^T\bm{B}\tilde{\bm{u}_f}$.
+\qed\end{proof}
+
+Note that the proof of entropy conservation for (\ref{eq:esdg1D}) relies mainly on the fact that $\bm{Q}_h$ is conservative (e.g., exact for constants) (\ref{eq:Qh1}) and that $\bm{Q}_h$ satisfies the SBP property (\ref{eq:hsbp}).
+Extensions to the multi-element, multi-dimensional, and non-conforming setting will utilize the same properties.
+
+%\note{Maybe add proof here and show how rest of proofs just rely on SBP and $\bm{Q}_h\bm{1} = \bm{0}$?}.
 For proofs of high order accuracy and entropy conservation, we refer the reader to \cite{chan2018efficient}.
 
-The matrix $\bm{F}$ is formed by computing ${\bm{F}}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}$ on the fly.
+%The matrix $\bm{F}$ is formed by computing ${\bm{F}}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}$ on the fly.
 Each entry requires evaluating the entropy conservative numerical flux $\bm{f}_S$, resulting in a large number of floating point operations \cite{wintermeyer2018entropy}.
@@
 For simplicity, we illustrate this in 2D (the extension to 3D is straightforward).
-Let $\bm{M}_{\rm 1D}, \bm{Q}_{\rm 1D}$ denote one-dimensional SBP norm (mass) and differentiation matrices, and let $\bm{E}_{\rm 1D}$ denote the 1D face interpolation matrix.
+Let $\bm{M}_{\rm 1D}, \bm{Q}_{\rm 1D}$ denote one-dimensional generalized SBP norm (mass) and differentiation matrices, and let $\bm{E}_{\rm 1D}$ denote the 1D face interpolation matrix.
 We define multi-dimensional  mass and differentiation matrices in terms of Kronecker products
-\[
+\begin{equation}
 \hat{\bm{Q}}_1 = \bm{Q}_{\rm 1D} \otimes \bm{M}_{\rm 1D}, \qquad \hat{\bm{Q}}_2  = \bm{M}_{\rm 1D} \otimes \bm{Q}_{\rm 1D}, \qquad \hat{\bm{M}} = \bm{M}_{\rm 1D} \otimes  \bm{M}_{\rm 1D}.
-\]
+\label{eq:kron}
+\end{equation}
 We also construct 2D face interpolation matrices from Kronecker products.
@@
 Let $\hat{\bm{Q}}_{i,h}$ denote the hybridized SBP operator for the $i$th coordinate on the reference element, where $\hat{\bm{Q}}_{i,h}$ is defined as
-\begin{align*}
+\begin{align}
 \hat{\bm{Q}}_{i,h} = \frac{1}{2}\begin{bmatrix}
@@
 \end{bmatrix}.
-\end{align*}
-These matrices satisfy the multi-dimensional SBP property
+\label{eq:hsbpdef}
+\end{align}
+These matrices satisfy the following multi-dimensional properties:
+\begin{lemma}
+\label{lemma:hsbprefprops}
+Let $\hat{\bm{Q}}_{i}$ be defined as in (\ref{eq:kron}) and let $\hat{\bm{Q}}_{i,h}$ be defined as in (\ref{eq:hsbpdef}).  Then, 
 \begin{align}
@@
 & \hat{\bm{B}}_i
-\end{bmatrix}.
-\label{eq:hsbp}
+\end{bmatrix}, \qquad \hat{\bm{Q}}_{i,h}\bm{1} = \bm{0}.
 \end{align}
-The construction of 3D differentiation and interpolation matrices proceeds similarly.  Then, a $d$-dimensional entropy conservative formulation on the reference element $[-1,1]^d$ is given as 
+\end{lemma}
+The SBP property follows by construction of $\hat{\bm{Q}}_{i,h}$.
+The property $\hat{\bm{Q}}_{i,h}\bm{1} = \bm{0}$ is a consequence of the Kronecker product and the same arguments used to derive (\ref{eq:Qh1}).
+The construction of 3D differentiation and interpolation matrices proceeds similarly, and the matrices satisfy the same properties as described in Lemma~\ref{lemma:hsbprefprops}.
+
+Then, a $d$-dimensional entropy conservative formulation on the reference element $[-1,1]^d$ is given as 
 \begin{gather}
 \hat{\bm{M}}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T
-\sum_{i=1}^d \LRp{\hat{\bm{Q}}_{i,h} \circ \bm{F}_i}\bm{1} + \bm{E}^T\hat{\bm{B}}_i\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_f)} = 0 \label{eq:esdgref}\\
+\sum_{i=1}^d \LRp{2\hat{\bm{Q}}_{i,h} \circ \bm{F}_i}\bm{1} + \bm{E}^T\hat{\bm{B}}_i\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_f)} = 0 \label{eq:esdgref}\\
 \LRp{\bm{F}_i}_{jk} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}
@@
 Then, the (scaled) physical normal vectors on $D^k$ are related to $\hat{\bm{n}}\hat{J}_f$ through
-\[
+\begin{equation}
 \bm{n}J_f = G\hat{\bm{n}}\hat{J}_f.
-\]
+\label{eq:nJ_Gnhat}
+\end{equation}
 We assume that the mesh is watertight, such that the scaled normal vectors are equal and opposite across shared faces between two elements.
+We also assume that the geometric terms ${G}_{ij}$ are constructed such that they are polynomials of degree less than or equal to $N$ and satisfy a discrete geometric conservation law (GCL)
+\begin{equation}
+\sum_{j=1}^d \pd{}{\hat{x}_j} G_{ij} = 0.
+%\sum_{j=1}^d \hat{\bm{Q}}_{j,h} \bm{G}_{ij} = \bm{0}.
+\label{eq:dgcl}
+\end{equation}
+The discrete GCL is satisfied automatically for isoparametric mappings in 2D, and there exist several techniques to enforce the satisfaction of a discrete GCL on various three-dimensional domains \cite{thomas1979geometric, kopriva2006metric, crean2018entropy, chan2018discretely, kozdon2018energy, kopriva2019free}.
 
@@
 We can define the physical differentiation matrices $\bm{Q}_{i,h}$ on $D^k$ via the skew-symmetric splitting
-\[
-\bm{Q}_{i,h} = \frac{1}{2}\sum_{j=1}^d \diag{\bm{g}_{ij}}\hat{\bm{Q}}_{j,h} + \hat{\bm{Q}}_{j,h} \diag{\bm{g}_{ij}},
-\]
-where $\bm{g}_{ij}$ denotes the vector of values of $G_{ij} = J\pd{\hat{\bm{x}}_j}{\bm{x}_i}$ at both volume and surface points.
-We also assume that the geometric terms $\bm{g}_{ij}$ are constructed such that they satisfy a discrete geometric conservation law (GCL)
-\[
-\sum_{j=1}^d \hat{\bm{Q}}_{j,h} \bm{g}_{ij} = \bm{0}.
-\]
-The discrete GCL is satisfied automatically for isoparametric mappings in 2D, and there exist several techniques to enforce the satisfaction of a discrete GCL on various three-dimensional domains \cite{thomas1979geometric, kopriva2006metric, crean2018entropy, chan2018discretely, kozdon2018energy, kopriva2019free}.
+\begin{equation}
+\bm{Q}_{i,h} = \frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h} + \hat{\bm{Q}}_{j,h} \diag{\bm{G}_{ij}},
+\label{eq:curvedQ}
+\end{equation}
+where $\bm{G}_{ij}$ denotes the vector of values of $G_{ij} = J\pd{\hat{\bm{x}}_j}{\bm{x}_i}$ at both volume and surface points.
+Since $G_{ij}$ is assumed to be a polynomial of degree less than or equal to $N$, $\bm{G}_{ij}$ is constructed using polynomial interpolation.
 
-Because the reference matrices $\hat{\bm{Q}}_{j,h}$ satisfy the SBP property (\ref{eq:hsbp}), one can show \cite{chan2018discretely} that the physical matrices $\bm{Q}_{i,h}$ satisfy a physical SBP property.
+Because the reference matrices $\hat{\bm{Q}}_{j,h}$ satisfy the reference SBP properties (\ref{eq:hsbp}), one can show \cite{chan2018discretely} that the physical matrices $\bm{Q}_{i,h}$ satisfy analogous physical SBP properties.
 \begin{lemma}
 \label{lemma:Qhprops}
-
+Suppose the geometric terms $\bm{G}_{ij}$ satisfy the discrete GCL (\ref{eq:dgcl}), the normals are constructed via (\ref{eq:nJ_Gnhat}), and that $\bm{Q}_{i,h}$ is constructed using (\ref{eq:curvedQ}).
+Then, %the conditions for Lemma~\ref{lemma:hsbprefprops} hold.
+Then, %$\hat{\bm{Q}}_j$ satisfies the properties %a generalized SBP property% \note{finish here}
 \[
@@
 \bm{0} &\\
-& \bm{B}_i \end{bmatrix}, \qquad \bm{B}_i = \diag{\bm{n}_i \circ \bm{w}_f},
+& \bm{B}_i \end{bmatrix}, \qquad \bm{Q}_{i,h}\bm{1} = \bm{0},
 \]
-where $\bm{n}_i$ is a vector whose entries consist of values of the $i$th component of the scaled normals $n_iJ_f$ at face points, and $\bm{w}_f$ is the vector of face quadrature weights $w_i$.  
+where $\bm{n}_i$ is a vector whose entries consist of values of the $i$th component of the scaled normals $n_iJ_f$ at face points, $\bm{w}_f$ is the vector of face quadrature weights $w_i$, and $\bm{B}_i = \diag{\bm{n}_i \circ \bm{w}_f}$.
 \end{lemma}
+%\begin{proof}
+The proof is identical to that of \cite{chan2018discretely}.  
+%We review it here for completeness.  Expanding out $\bm{Q}_{i,h} + \bm{Q}_{i,h}^T$ yields
+%\[
+%\frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h} + \hat{\bm{Q}}_{j,h}^T\diag{\bm{G}_{ij}} + \hat{\bm{Q}}_{j,h} \diag{\bm{G}_{ij}} +  \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h}^T.
+%\]
+%For each term in the sum, we apply the SBP property of $\hat{\bm{Q}}_{j,h}$.  The volume terms cancel, leaving only surface terms
+%\[
+%\frac{1}{2}\sum_{j=1}^d {\hat{\bm{B}}_{j,h}\diag{\bm{G}_{ij}} +  \diag{\bm{G}_{ij}}\hat{\bm{B}}_{j,h}} = \begin{bmatrix}
+%\bm{0} &\\
+%& \bm{B}_i \end{bmatrix}.
+%\]
+%where the last step uses (\ref{eq:nJ_Gnhat}) and that $\hat{\bm{B}}_{j,h}$ and $\diag{\bm{G}_{ij}}$ are diagonal.  
+%
+%To show $\bm{Q}_{i,h}\bm{1} = \bm{0}$, we note that
+%\[
+%\bm{Q}_{i,h}\bm{1} = \frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h}\bm{1} + \hat{\bm{Q}}_{j,h} \diag{\bm{G}_{ij}}\bm{1} = \frac{1}{2}\sum_{j=1}^d \hat{\bm{Q}}_{j,h} \bm{G}_{ij}
+%\]
+%since $\hat{\bm{Q}}_{j,h}\bm{1} = \bm{0}$ by Lemma~\ref{lemma:hsbprefprops}.
+Note that, since $\bm{G}_{ij}$ is constructed using polynomial interpolation, $\bm{G}_{ij} = \LRs{\bm{g}_{ij}, \quad \bm{E}\bm{g}_{ij}}^T$, where $\bm{g}_{ij}$ denotes the values of $G_{ij}$ at volume points.
+The remaining term can be expanded out
+%\[
+%\sum_{j=1}^d \hat{\bm{Q}}_{j,h} \bm{G}_{ij} = \sum_{j=1}^d \begin{bmatrix} \hat{\bm{Q}}_j\bm{g}_{ij} - \hat{\bm{g}}_j^T\bm{G}_{ij} + \bm{E}^T\bm{B}\bm{E}\bm{g}_{ij}\\
+%\bm{B}\bm{E}\bm{g}_{ij} - \bm{B}\bm{E}\bm{g}_{ij}
+%\end{bmatrix} =
+%\sum_{j=1}^d \begin{bmatrix} - \hat{\bm{g}}_j^T\bm{G}_{ij} + \bm{E}^T\bm{B}\bm{E}\bm{g}_{ij}\\
+%\bm{0}
+%\end{bmatrix} 
+%\]
+%where we have used that $\sum_{j=1}^d\hat{\bm{Q}}_j\bm{g}_{ij} = \bm{0}$, since $G_{ij}$ satisfies the discrete GCL (\ref{eq:dgcl}), $\bm{g}_{ij}$ is a polynomial of degree less than or equal to $N$, and $\hat{\bm{Q}}_j$ exactly differentiates polynomials of degree $N$.
+The remaining terms vanish after an application of the SBP property and the GCL (\ref{eq:dgcl}).
+%\qed\end{proof}
 
@@
 {\bm{M}}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T
-\sum_{i=1}^d \LRp{{\bm{Q}}_{i,h} \circ \bm{F}_i}\bm{1} + \bm{E}^T{\bm{B}}_i\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_f)} = 0 \label{eq:esdg}
+\sum_{i=1}^d \LRp{2{\bm{Q}}_{i,h} \circ \bm{F}_i}\bm{1} + \bm{E}^T{\bm{B}}_i\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_f)} = 0 \label{eq:esdg}
 \end{gather}
@@
 
-\section{An entropy stable mortar formulation}
+\section{Entropy stable mortar formulations}
 \label{sec:3}
@@
 \label{lemma:Efm}
-Suppose the face (surface) quadrature is exact for polynomials of degree $N+M_f$ and mortar quadratures are exact for degree $N+M_m$ polynomials.
-Then, $\bm{E}_{fm}$ exactly recovers polynomials of degree $\min(N,M_f,M_m)$.
+Suppose the face (surface) quadrature is exact for polynomials of degree $N+N_f$ and mortar quadratures are exact for degree $N+N_m$ polynomials.
+Then, $\bm{E}_{fm}$ exactly recovers polynomials of degree $\min(N,N_f,N_m)$.
 \end{lemma}
 \begin{proof}
-Let $u(\bm{x})$ be a polynomial of degree $\min(N,M_f,M_m)$ or less, and let $\bm{u}_f$ be its values on the face (surface) nodes.
+Let $u(\bm{x})$ be a polynomial of degree $\min(N,N_f,N_m)$ or less, and let $\bm{u}_f$ be its values on the face (surface) nodes.
 Then, $\bm{u}_m = \bm{E}_{mf}\bm{u}_f$ are the interpolated values of the polynomial on the mortar nodes.
@@
 
-Since $u$ is degree $\min(N,M_f,M_m)$ and each Lagrange basis function is degree $N$, the integrand is computed exactly under the mortar quadrature.
-Moreover, since $u$ is degree $\min(N,M_f,M_m)$, it is also computed exactly using the face (surface) quadrature.
+Since $u$ is degree $\min(N,N_f,N_m)$ and each Lagrange basis function is degree $N$, the integrand is computed exactly under the mortar quadrature.
+Moreover, since $u$ is degree $\min(N,N_f,N_m)$, it is also computed exactly using the face (surface) quadrature.
 Thus, $\bm{E}_{mf}^T\bm{M}_m\bm{E}_{mf} = \bm{M}_f \bm{u}_f$ by exactness of the face and mortar quadratures, and $\bm{E}_{fm}\bm{u}_m = \bm{u}_f$.
@@
 Since it is not clear at first glance how the hybridized operator $\hat{\bm{Q}}_{i,h}$ or its mortar-based variant $\hat{\bm{Q}}_{i,m}$ can be used to perform differentiation, we review intuitive explanations of each operator.
-Let $\bm{u}$ denote basis coefficients for some function $u(\bm{x})$, and let $f, g$ denote two functions.
+Let $\bm{u}$ denote basis coefficients for some function $u(\bm{x})$, and let $f, g$ denote two functions on the reference element.
 Suppose $\bm{u}$ satisfies the following matrix system involving the hybridized operator $\hat{\bm{Q}}_{i,h}$
@@
 Let $\bm{v}$ denote basis coefficients for some arbitrary degree $N$ polynomial test function.
-Multiplying (\ref{eq:Qh}) by $\bm{v}^T$, expanding out the different blocks of $\bm{Q}_{i,h}$, and using generalized summation by parts yields
+Multiplying (\ref{eq:Qh}) by $\bm{v}^T$, expanding out the different blocks of $\hat{\bm{Q}}_{i,h}$, and using summation by parts yields
 \begin{align*}
-\bm{M}\bm{u} &= 
-\diag{\bm{f}}\bm{Q}_i\bm{g} + \frac{1}{2}\LRp{\diag{f(\hat{\bm{x}})}\bm{E}^T + \bm{E}^T f(\hat{\bm{x}}_f)}\bm{B}_i \LRp{g(\hat{\bm{x}}_f) - \bm{E} g(\hat{\bm{x}})}.
+\hat{\bm{M}}\bm{u} &= 
+\diag{\bm{f}}\hat{\bm{Q}}_i\bm{g} + \frac{1}{2}\LRp{\diag{f(\hat{\bm{x}})}\bm{E}^T + \bm{E}^T f(\hat{\bm{x}}_f)}\hat{\bm{B}}_i \LRp{g(\hat{\bm{x}}_f) - \bm{E} g(\hat{\bm{x}})}.
 \end{align*}
@@
 The mortar-based hybridized SBP operator can be interpreted similarly.
-Recall that $\bm{E}_{mf}$ interpolates from face nodes to mortar nodes.
+Recall that $\bm{E}_{mf}$ interpolates from face nodes to mortar nodes, and that $\bm{E}$ interpolates from volume nodes to face nodes.
 Thus, the matrix $\bm{E}_{mf}\bm{E}$ interpolates from volume nodes to mortar nodes.
@@
 \begin{align}
-\bm{M}\bm{u} &= \diag{\bm{f}}\bm{Q}_i\bm{g} \label{eq:mhsbp_correction}\\
-&+ \frac{1}{2} \diag{f(\hat{\bm{x}})}\bm{E}^T\bm{B}_{i,f}\LRp{g(\hat{\bm{x}}_f) - \bm{E}g(\hat{\bm{x}})} \nonumber\\
-&+ \frac{1}{2} \bm{E}^T\diag{f(\hat{\bm{x}}_f)}\bm{B}_{i,f}\LRp{\bm{E}_{fm}g(\hat{\bm{x}}_m) - \bm{E}g(\hat{\bm{x}})} \nonumber\\
-&+ \frac{1}{2} \bm{E}^T\bm{E}_{mf}^T\diag{f(\hat{\bm{x}}_m)}\bm{B}_{i,m}\LRp{g(\hat{\bm{x}}_m) - \bm{E}_{mf}g(\hat{\bm{x}}_f)} \nonumber
+\hat{\bm{M}}\bm{u} &= \diag{\bm{f}}\hat{\bm{Q}}_i\bm{g} \label{eq:mhsbp_correction}\\
+&+ \frac{1}{2} \diag{f(\hat{\bm{x}})}\bm{E}^T\hat{\bm{B}}_{i,f}\LRp{g(\hat{\bm{x}}_f) - \bm{E}g(\hat{\bm{x}})} \nonumber\\
+&+ \frac{1}{2} \bm{E}^T\diag{f(\hat{\bm{x}}_f)}\hat{\bm{B}}_{i,f}\LRp{\bm{E}_{fm}g(\hat{\bm{x}}_m) - \bm{E}g(\hat{\bm{x}})} \nonumber\\
+&+ \frac{1}{2} \bm{E}^T\bm{E}_{mf}^T\diag{f(\hat{\bm{x}}_m)}\hat{\bm{B}}_{i,m}\LRp{g(\hat{\bm{x}}_m) - \bm{E}_{mf}g(\hat{\bm{x}}_f)} \nonumber
 \end{align}
@@
 \begin{lemma}
-Suppose that the face (surface) quadrature is exact for polynomials of degree $N+M_f$ and mortar quadratures are exact for degree $N+M_m$ polynomials.
-Let $f = 1$ and $g$ be a degree $\min(N,M_f,M_m)$ polynomial.
-Then, the equations (\ref{eq:Qm}) and (\ref{eq:mhsbp_correction}) are exact.
+Let $f(\bm{x}) = 1$ and suppose that the face (surface) quadrature is exact for polynomials of degree $N+N_f$ and mortar quadratures are exact for degree $N+N_m$ polynomials.
+Then, derivative approximations given by the equations (\ref{eq:Qm}) and (\ref{eq:mhsbp_correction}) are exact if $g(\bm{x})$ is a degree $\min(N,N_f,N_m)$ polynomial.
 \end{lemma}
@@
 Following \cite{chan2017discretely}, if $f = 1$  and $g(\bm{x})$ is a degree $N$ polynomial, then  (\ref{eq:mhsbp_correction}) is exact if the correction terms vanish.
-The first and third correction terms are zero by the fact that $\bm{E}, \bm{E}_{mf}$ are degree $N$ interpolation operators, and the second correction term vanishes by Lemma~\ref{lemma:Efm}.
+The first and third correction terms are zero by the fact that $\bm{E}, \bm{E}_{mf}$ are degree $N$ interpolation operators, and the second correction term vanishes since Lemma~\ref{lemma:Efm} implies that $\bm{E}_{fm}$ exactly recovers polynomials of degree $\min(N,N_f,N_m)$.
 \qed\end{proof}
@@
 \label{lemma:mhsbp}
-Let $\bm{Q}_{i,m}$ be defined as in (\ref{eq:mhsbp}).  Then, 
+Let $\hat{\bm{Q}}_{i,m}$ be defined as in (\ref{eq:mhsbp}).  Then, 
 \[
-\bm{Q}_{i,m} + \bm{Q}_{i,m}^T = \begin{bmatrix}
+\hat{\bm{Q}}_{i,m} + \hat{\bm{Q}}_{i,m}^T = \begin{bmatrix}
 \bm{0} && \\
 &\bm{0} &\\
-&& \bm{B}_{i,m}\end{bmatrix}, \qquad \bm{Q}_{i,m}\bm{1} = \bm{0}.
+&& \hat{\bm{B}}_{i,m}\end{bmatrix}, \qquad \hat{\bm{Q}}_{i,m}\bm{1} = \bm{0}.
 \]
@@
 where we have used that $\hat{\bm{B}}_{i,m}$ is diagonal.  Recall that $\bm{E}_{fm} = \bm{M}_f^{-1}\bm{E}_{mf}^T\bm{M}_m$.  
-Then, by the definition of $\bm{B}_{i,m},\bm{B}_{i,h}$ in (\ref{eq:Bi}), we have that
+Then, by the definition of $\hat{\bm{B}}_{i,m}, \hat{\bm{B}}_{i,h}$ in (\ref{eq:Bi}), we have that
 \begin{align*}
@@
 \[
-\bm{E}_{mf}^T\bm{M}_m \diag{\hat{\bm{n}}_{i,m}}= \bm{E}_{mf}^T\bm{B}_{i,m}.
+\diag{\hat{\bm{n}}_{i,f}} \bm{E}_{mf}^T\bm{M}_m =\bm{E}_{mf}^T\bm{M}_m \diag{\hat{\bm{n}}_{i,m}}= \bm{E}_{mf}^T\hat{\bm{B}}_{i,m}.
 \]  
@@
 \hat{\bm{M}}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \\ \bm{E}_{mf}\bm{E} \end{bmatrix}^T
-\sum_{i=1}^d \LRp{\hat{\bm{Q}}_{i,m} \circ \bm{F}_i}\bm{1} + \bm{E}_{mf}^T\hat{\bm{B}}_{i,m}\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_m)} = 0 \label{eq:esdgmref}\\
+\sum_{i=1}^d \LRp{2\hat{\bm{Q}}_{i,m} \circ \bm{F}_i}\bm{1} + \bm{E}_{mf}^T\hat{\bm{B}}_{i,m}\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_m)} = 0 \label{eq:esdgmref}\\
 \tilde{\bm{u}} = \begin{bmatrix}
@@
 \end{gather}
-where $\LRp{\bm{F}_i}_{jk} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}$ for all indices $j,k$ of $\tilde{\bm{u}}$.  This can be extended to mapped elements
+where $\LRp{\bm{F}_i}_{jk} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}$ for all indices $j,k$ of $\tilde{\bm{u}}$.
 
-%An entropy stable skew-symmetric formulation can be given on the reference element $\hat{D}$ as follows:
-%\[
-%\bm{M}\td{\bm{u}_N}{t} + \begin{bmatrix} \bm{V}_q \\ \bm{E} \end{bmatrix}^T
-%\LRp{\begin{bmatrix}
-%\bm{Q}_i-\bm{Q}_i^T & \bm{E}^T\bm{B}_i\\
-%-\bm{B}_i\bm{E} & \bm{0}
-%\end{bmatrix} \circ \bm{F}_S}\bm{1} + \bm{E}^T\bm{B}_i\bm{f}_i^* = 0.  
-%\]
-%
-%We incorporate mortars by modifying this formulation.
-Let $\bm{V}_m$ denote the matrix which maps volume quadrature points to values at surface \textit{mortar} points.
-We also need to introduce interpolation operators $\bm{T}_f, \bm{T}_m$ for surface trace spaces.
-Here, $\bm{T}_f$ maps from polynomials on the surface $\partial \hat{D}$ to surface quadrature points, while $\bm{T}_m$ maps from polynomials on $\partial \hat{D}$ to mortar quadrature points.
-We can then define surface and mortar mass and projection matrices
-%\begin{align*}
-%\bm{M}_m = \bm{T}_m^T\bm{M}_m\bm{T}_m, \qquad \bm{P}_m = \bm{M}_m^{-1}\bm{T}_m^T\bm{M}_m\\
-%\bm{M}_f = \bm{T}_f^T\bm{W}_f\bm{T}_f, \qquad \bm{P}_f = \bm{M}_m^{-1}\bm{T}_f^T\bm{W}_f.
-%\end{align*}
-%Note that we have used the mortar mass matrix $\bm{M}_m$ in the above definition of the face projection.
-%This is only necessary if the surface and mortar quadratures do not exactly integrate degree $2N$ polynomials.
-If both the surface and mortar quadratures are exact for polynomials of degree $2N$, then $\bm{M}_m = \bm{M}_f$, and no distinction is necessary between the two mass matrices.
-%
-%\begin{remark}
-%In order to show stability, we require that both $\bm{P}_f$ and $\bm{P}_m$ are defined using the same mass matrix.
-We have used the mortar mass matrix $\bm{M}_m$ here; however, we could also use the surface mass matrix $\bm{M}_f$ in both $\bm{P}_f$ and $\bm{P}_m$.
-We use the mortar mass matrix in this work, as the mortar quadrature is generally more accurate than the surface quadrature for our use cases (e.g.\ hybrid and non-conforming meshes).
-%\end{remark}
+\subsection{Mortar-based formulations on mapped elements}
 
-%
-%We can now define operators which map between surface and mortar quadrature points.
-Let $\bm{E}_m$ denote the map from surface to mortar points, and let $\tilde{\bm{E}}_m $ denote the map from mortar to surface points.
-Both operators are defined through an $L^2$ projection to the trace space and interpolation to appropriate points
-%\[
-%\bm{E}_m = \bm{T}_m \bm{P}_f, \qquad \tilde{\bm{E}}_m = \bm{T}_f \bm{P}_m.
-%\]  
-%We also define a mortar boundary matrix
-%\[
-%\tilde{\bm{B}}_i = \bm{M}_m\diag{\hat{\bm{n}}_i}.
-%\]
-%Then, a mortar-based formulation can be given as follows:
-%\begin{equation}
-%\bm{M}\td{\bm{u}_N}{t} + \begin{bmatrix} \bm{V}_q \\ \bm{E} \\ \bm{V}_m \end{bmatrix}^T
-%\LRp{\begin{bmatrix}
-%\bm{Q}_i-\bm{Q}_i^T & \bm{E}^T\bm{B}_i &\\
-%-\bm{B}_i\bm{E} &  & \bm{B}_i\tilde{\bm{E}}_m \\
-%& -\tilde{\bm{B}}_i{\bm{E}}_m & 
-%\end{bmatrix} \circ \bm{F}_S}\bm{1} + \bm{V}_m^T\tilde{\bm{B}}_i\bm{f}_i^* = 0.  
-%\label{eq:mortar}
-%\end{equation}
-%Here, we have appended an extra row and column to the hybridized SBP matrix, and the inter-element numerical flux $\bm{f}^*$ is now computed in terms of the \emph{mortar} nodes.
+We can extend the above scheme to mapped curvilinear elements by constructing physical mortar-based hybridized SBP operators in a manner akin to (\ref{eq:curvedQ}).
+Abusing notation, we use $\bm{G}_{ij}$ to now denote the vector containing values of $G_{ij} = J\pd{\hat{\bm{x}}_j}{\bm{x}_i}$ at volume, surface, and mortar points.
+Then, we define physical operators $\bm{Q}_{i,m}$ via
+\begin{equation}
+\bm{Q}_{i,m} = \frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,m} + \hat{\bm{Q}}_{j,m} \diag{\bm{G}_{ij}}.
+\label{eq:curvedQ_mortar}
+\end{equation}
+Proofs of entropy stability for formulations (\ref{eq:esdg1D}), (\ref{eq:esdg}), and (\ref{eq:esdgmref}) require that the hybridized SBP operators are conservative (e.g., they are exact for constants) and satisfy the SBP property.
+Entropy conservation for mortar-based formulations on mapped elements will require similar properties.
+However, in contrast to hybridized SBP operators, conservation for mapped mortar-based operators requires additional constraints on the accuracy of surface and mortar quadratures relative to the polynomial degree of the geometric mapping.
+
+We assume now that each geometric mapping from reference element $\hat{D}$ to physical element $D^k$ is a polynomial of degree $N_{\rm geo}$.
+We also introduce the space of tensor product polynomials in $d$ dimensions $Q^{N_1,\ldots, N_d}$ as
+\[
+Q^{N_1,\ldots, N_d} = \LRc{ \hat{x}_1^{i_1}\hat{x}_2^{i_2}\ldots\hat{x}_d^{i_d}, \qquad 0\leq i_k \leq N_k, \quad k = 1,\ldots,d.}
+\]
+We denote $Q^{N,\ldots,N} = Q^N$ for conciseness.
+
+In \cite{chan2019skew}, it was shown that the geometric terms $G_{ij}$ are tensor product polynomials of specific degrees.  For mapped quadrilateral elements, $G_{ij}$ satisfy
+\begin{align*}
+&G_{i1} \in Q^{N_{\rm geo},N_{\rm geo}-1}\\
+&G_{i2} \in Q^{N_{\rm geo}-1,N_{\rm geo}}.
+\end{align*}
+for $i = 1,2$.  Moreover, the geometric terms naturally satisfy the GCL condition (\ref{eq:dgcl}).  We then have the following lemma:
+\begin{lemma}
+\label{lemma:Qmprops_2d}
+Suppose $D^k$ is a quadrilateral element, and suppose that the surface quadrature is exact for $Q^{N+N_f}$ and mortar quadratures are exact for degree $Q^{N+N_m}$ polynomials.
+Then, if $\bm{Q}_{i,m}$ is constructed using (\ref{eq:curvedQ_mortar}) and if $N_{\rm geo} \leq \min(N,N_f+1,N_m+1)$,
+\begin{align*}
+\bm{Q}_{i,m} + \bm{Q}_{i,m}^T &= \bm{B}_{i,m}, \quad \text{(SBP property)}\\
+\bm{Q}_{i,m}\bm{1} &= \bm{0}, \quad \text{(conservation)}.
+\end{align*}
+\end{lemma}
+\begin{proof}
+%The proof mimicks that of Lemma~\ref{lemma:Qhprops}.  
+The SBP property is straightforward to show using the SBP property of $\hat{\bm{Q}}_{i,m}$.  To show conservation, expanding out $\bm{Q}_{i,m}\bm{1} $ yields
+\[
+\bm{Q}_{i,m}\bm{1} = \frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,m}\bm{1} + \hat{\bm{Q}}_{j,m} \diag{\bm{G}_{ij}}\bm{1} = \frac{1}{2}\sum_{j=1}^d\hat{\bm{Q}}_{j,m} \bm{G}_{ij}
+\]
+since $\hat{\bm{Q}}_{j,m}\bm{1} = \bm{0}$ by Lemma~\ref{lemma:mhsbp}.
+We can expand out the remaining terms using (\ref{eq:mhsbp_correction}).
+However, since $\bm{G}_{ij}$ is constructed by interpolating the values of the polynomial $G_{ij}$, the boundary correction terms in (\ref{eq:mhsbp_correction}) vanish, reducing the expression to
+\[
+\sum_{j=1}^d\hat{\bm{Q}}_{j,m} \bm{G}_{ij} = \sum_{j=1}^d\hat{\bm{Q}}_{j} \bm{g}_{ij} = \bm{0},
+\]
+where $\bm{g}_{ij}$ denotes the values of the $G_{ij}$ at volume nodes, and the final sum vanishes due to the GCL (\ref{eq:dgcl}) and the fact that $\hat{\bm{Q}}_j$ is exact for degree $N$ polynomials.
+\note{Fix to add in tensor product quadrature.}
+\qed\end{proof}
+
+For hexahedral elements, it is more challenging to construct geometric terms $G_{ij}$ which satisfy the GCL while retaining high order accuracy.
+This is further complicated by the fact that we must take into account the polynomial degrees of $G_{ij}$ when proving conservation for mortar-based SBP operators.
+We consider two approaches for computing hexahedral geometric terms in this work:
+\begin{enumerate}
+\item Approach 1: the construction of $G_{ij}$ from \cite{kopriva2006metric, chan2018efficient}, which yields for $i = 1,2,3$
+\begin{align*}
+&G_{i1} \in Q^{N_{\rm geo}-1,N_{\rm geo},N_{\rm geo}}\\
+&G_{i2} \in Q^{N_{\rm geo},N_{\rm geo}-1,N_{\rm geo}}\\
+&G_{i3} \in Q^{N_{\rm geo},N_{\rm geo},N_{\rm geo}-1}.
+\end{align*}
+\item Approach 2: the construction of $G_{ij}$ from \cite{kozdon2018energy} (Appendix C.3), which yields for $i = 1,2,3$
+\begin{align*}
+&G_{i1} \in Q^{N_{\rm geo},N_{\rm geo}-1,N_{\rm geo}-1}\\
+&G_{i2} \in Q^{N_{\rm geo}-1,N_{\rm geo},N_{\rm geo}-1}\\
+&G_{i3} \in Q^{N_{\rm geo}-1,N_{\rm geo}-1,N_{\rm geo}}.
+\end{align*}
+\end{enumerate}
+These approaches are described in more detail in Appendix~\ref{app:A}.  
+\begin{lemma}
+\label{lemma:Qmprops_3d}
+Suppose $D^k$ is a hexahedral element, and suppose that the surface quadrature is exact for $Q^{N+N_f}$, mortar quadratures are exact for degree $Q^{N+N_m}$ polynomials, and that $\bm{Q}_{i,m}$ is constructed using (\ref{eq:curvedQ_mortar}).
+If either
+\begin{enumerate}
+\item $G_{ij}$ is constructed using Approach 1 \cite{kopriva2006metric, chan2018efficient} and $N_{\rm geo} \leq \min(N,N_f,N_m)$, or
+\item $G_{ij}$ is constructed using Approach 2 \cite{kozdon2018energy} and $N_{\rm geo} \leq \min(N,N_f-1,N_m-1)$, 
+\end{enumerate}
+then 
+\begin{align*}
+\bm{Q}_{i,m} + \bm{Q}_{i,m}^T &= \bm{B}_{i,m}, \quad \text{(SBP property)}\\
+\bm{Q}_{i,m}\bm{1} &= \bm{0}, \quad \text{(conservation)}.
+\end{align*}
+\end{lemma}
+\begin{proof}
+\note{Finish} The proof of Lemma~\ref{lemma:Qmprops_3d} is identical to that of Lemma~\ref{lemma:Qmprops_2d}
+\qed\end{proof}
+
+An entropy conservative mortar-based formulation on mapped elements is then
+\begin{gather}
+{\bm{M}}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \\ \bm{E}_{mf}\bm{E} \end{bmatrix}^T
+\sum_{i=1}^d \LRp{2{\bm{Q}}_{i,m} \circ \bm{F}_i}\bm{1} + \bm{E}_{mf}^T{\bm{B}}_{i,m}\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_m)} = 0 \label{eq:esdgm}\\
+\tilde{\bm{u}} = \begin{bmatrix}
+\bm{u}_h\\
+\tilde{\bm{u}}_f\\
+\tilde{\bm{u}}_m
+\end{bmatrix}, \qquad \tilde{\bm{u}}_m = \bm{u}\LRp{\bm{E}_{mf}\bm{E}\bm{v}(\bm{u}_h)}, \qquad \bm{f}^* = \bm{f}_S\LRp{\tilde{\bm{u}}_m^+,\tilde{\bm{u}}_m},
+\end{gather}
+where $\bm{B}_{i,m} = \diag{\bm{n}_{i,m}\circ \bm{w}_m}$ is the diagonal matrix whose entries consist of the scaled physical normals.
+Again, we assume geometric terms on curved elements are approximated using polynomials, and that the normals are constructed via (\ref{eq:nJ_Gnhat}) and polynomial interpolation.
+
+%\note{Note: line DG approach can work, but it's flipped: we increase quadrature strength in the direction \emph{orthogonal} to derivative instead.
+May improve accuracy over Lobatto by giving equivalence w/skew form.}
 
@@
 -\bm{B}_i\bm{E} & \\
-\end{bmatrix} \circ \bm{F}_S}\bm{1} + \bm{E}^T\bm{B}_i \tilde{\bm{f}}^*_i = 0 \label{eq:mortar2}\\
+\end{bmatrix} \circ \bm{F}_S}\bm{1} + \bm{E}^T\bm{B}_i \tilde{\bm{f}}^*_i = 0 \label{eq:mortarform}\\
 \tilde{\bm{f}}^*_i &= \tilde{\bm{E}}_m\bm{f}^*_i + \LRp{\tilde{\bm{E}}_m\circ \bm{F}_S^{sm}}\bm{1} - \tilde{\bm{E}}_m \LRp{ \bm{E}_m \circ \bm{F}_S^{ms}}\bm{1} 
@@
 \includegraphics[width=.55\textwidth]{mortar_coupling.png}
-\caption{Blah}
+\caption{Coupling between volume, surface, and mortar nodes for formulation (\ref{eq:mortarform}).}
 \end{figure}
 
-\section{Error estimates}
+\subsection{Extensions to $p$ non-conformity, general quadratures, and non-hexahedral elements}
 
-\note{Weak derivative vs strong derivative error estimates.  SBP yields that the skew-symmetric form is equivalent to weak form.}
+\note{Extension to $p$ non-conforming meshes is straightforward.  Simply replace the definition of a mortar.}
 
-\note{Line DG approach can work, but it's flipped: we increase quadrature strength in the direction \emph{orthogonal} to derivative instead.
-May improve accuracy over Lobatto by giving equivalence w/skew form.}
+\note{Extension to general quadratures just requires replacing $\bm{I}$ with $\bm{V}_q, \bm{P}_q$, and redefining $\bm{E}$ matrices.}
+
+\note{Explain why this isn't as much of an issue - Shadpey \cite{shadpey2019energy} use this for simplicial elements.
+The interpolation operators are dense, and using the mortar-based approach wouldn't actually decrease the number of operations.}
 
@@
 
+\subsection{2D results}
 Figure~\ref{subfig:noncon} shows preliminary results for the 2D compressible Euler equations by the PI on a manually constructed non-conforming quadrilateral mesh, with volume, surface, and mortar quadratures constructed from one-dimensional Lobatto or Gauss quadrature rules.
@@
 
+\subsection{3D results}
+\subsection{Gauss Lobatto quadrature}
+
+\subsection{Gauss quadrature}
+
+\appendix
+
+\section{Construction of geometric terms for curved hexahedral elements}
+\label{app:A}
+
 \bibliographystyle{unsrt}

### 2019-12-25 docs/ESDG_mortar/main.tex
updating paper

@@
 We reduce this expense by introducing an entropy stable mortar-based treatment of non-conforming interfaces via a face-local correction term, and provide truncation error estimates for the resulting discretizations.
-Numerical experiments confirm the stability and accuracy of this approach, and we conclude by discussing an extension of this mortar treatment of non-conforming interfaces to general elements and arbitrary quadrature rules \cite{chan2017discretely}.
+Numerical experiments confirm the stability and accuracy of this approach, and we conclude by discussing how to extend mortar treatments of non-conforming interfaces to general elements and arbitrary quadrature rules \cite{chan2017discretely}.
 \end{abstract}
@@
 
-\subsection{Entropy conservative schemes in 1D}
+\subsection{Hybridized SBP operators in 1D}
 
@@
 We assume the solution is collocated at $(N+1)$ quadrature points $x_i$ with associated quadrature weights $w_i$, and consider primarily Lobatto or Gauss quadrature points.
-The collocation assumption is equivalent to approximating the solution using a degree $N$ Lagrange basis $\ell_j(x)$ defined on the $(N+1)$ quadrature points.
+The collocation assumption is equivalent to approximating the solution using a degree $N$ Lagrange basis $\ell_j(x)$ at the $(N+1)$ quadrature points.
 
@@
 
-It is possible to construct energy stable discretizations of linear hyperbolic systems using GSBP operators based on Gauss nodes \cite{fernandez2014generalized}.
+It is possible to construct energy stable high order discretizations of linear hyperbolic systems using GSBP operators based on Gauss nodes \cite{fernandez2014generalized}.
 However, for nonlinear conservation laws, the presence of the dense $\bm{E}$ matrix in the GSBP property (\ref{eq:gsbp}) complicates the imposition of boundary conditions and computation of inter-element numerical fluxes \cite{crean2018entropy, chan2017discretely, chan2018efficient}.
-
-We can avoid this issue by introducing ``hybridized'' (or decoupled) SBP operators \cite{chan2017discretely, chenreview}, which are defined as the block matrix $\bm{Q}_h$ \cite{chan2019skew}
+This can be avoided by using ``hybridized'' (or decoupled) SBP operators \cite{chan2017discretely, chenreview, chan2019skew}, which are defined as the block matrix
 \[
@@
 \]
-Note that the hybridized SBP operator satisfies a block form of the SBP property
+The hybridized SBP operator satisfies a block form of the SBP property
 \begin{equation}
@@
 where $\bm{E}$ does not appear in the block boundary matrix on the right hand side.
-Here, we have used $\bm{0}$ to denote a matrix or vector of zeros, where the size is inferred from context.
-We will adopt a similar convention for the vector of ones $\bm{1}$.
-
-The matrix $\bm{Q}_h$ also satisfies
-\[
+Here, we have used $\bm{0}, \bm{1}$ to denote a matrix or vector of zeros or ones, where the size is inferred from context.
+Note that the matrix $\bm{Q}_h$ also satisfies
+\begin{equation}
 \bm{Q}_h\bm{1} = 
@@
 \frac{1}{2}\begin{bmatrix}
--\bm{Q}^T\bm{1} + \bm{E}^T\bm{B}\bm{1}\\
+\LRp{-\bm{Q}^T + \bm{E}^T\bm{B}}\bm{1}\\
 \bm{0}
-\end{bmatrix}
-\]
-where we have used that $\bm{E}\bm{1} = \bm{1}$ (since $\bm{E}$ is a high order accurate boundary interpolation matrix) and $\bm{Q}\bm{1} = \bm{0}$ (since $\bm{Q}$ is a differentiation matrix).
-Combining the GSBP property $-\bm{Q}^T + \bm{E}^T\bm{B}\bm{E} = \bm{Q}$ and $\bm{E}\bm{1} = \bm{1}$ then yields that
-\begin{equation}
-\bm{Q}_h\bm{1} = 
-\frac{1}{2}\begin{bmatrix}
--\bm{Q}^T\bm{1} + \bm{E}^T\bm{B}\bm{1}\\
-\bm{0}
-\end{bmatrix} = 
+\end{bmatrix}=
 \frac{1}{2}\begin{bmatrix}
@@
 \bm{0}
-\end{bmatrix} 
-= \frac{1}{2}\begin{bmatrix}
-\bm{0}\\
-\bm{0}
-\end{bmatrix} = \bm{0}.
+\end{bmatrix} = \bm{0}
 \label{eq:Qh1}
 \end{equation}
-
-The aforementioned matrices can now be used to construct high order accurate entropy stable and entropy conservative schemes.
-We first introduce an entropy conservative numerical flux \cite{tadmor1987numerical}.
-Let $\bm{u}_L, \bm{u}_R$ denote left and right states in the conservative variables.
-Then, an entropy conservative numerical flux is a vector-valued function $\bm{f}_S$ which satisfies the following three properties
-\begin{gather*}
-\bm{f}_S\LRp{\bm{u},\bm{u}} = \bm{f}(\bm{u}), \qquad \text{(consistency)}\\
-\bm{f}_S\LRp{\bm{u}_L,\bm{u}_R} = \bm{f}_S(\bm{u}_R,\bm{u}_L), \qquad \text{(symmetry)}\\
-\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}_S\LRp{\bm{u}_L,\bm{u}_R} = \psi(\bm{u}_L) - \psi(\bm{u}_R), \qquad \text{(conservation)}.
-\end{gather*}
-Here, $\bm{v}_L, \bm{v}_R$ are the entropy variables evaluated at $\bm{u}_L, \bm{u}_R$, and $\psi$ is the entropy potential which appears in (\ref{eq:weakentropyineq}).
-Then, an entropy conservative scheme on the reference interval is given as
-\begin{gather}
-\bm{M}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T
-\LRp{2\bm{Q}_h \circ \bm{F}}\bm{1} + \bm{E}^T\bm{B}\LRp{\bm{f}^*-\bm{f}(\tilde{\bm{u}}_f)} = 0 \label{eq:esdg1D}\\ 
-\tilde{\bm{u}} = \begin{bmatrix}
-\bm{u}_h\\
-\tilde{\bm{u}}_f
-\end{bmatrix}, \qquad \tilde{\bm{u}}_f = \bm{u}\LRp{\bm{E}\bm{v}(\bm{u}_h)}, \qquad \bm{f}^* = \bm{f}_S\LRp{\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f}.
-\nonumber
-\end{gather}
-Here, $\tilde{\bm{u}}_f^+$ denotes some exterior state which can be used to enforce boundary conditions or coupling between different elements, and $\circ$ denotes the matrix Hadamard product.
-
-\begin{theorem}
-Assuming continuity in time, the formulation (\ref{eq:esdg1D}) is entropy conservative in the sense that
-\[
-\bm{1}^T\bm{M}\td{S(\bm{u}_h)}{t} + \bm{1}^T\bm{B}\LRp{\bm{v}_f^T\bm{f}^* - \psi(\tilde{\bm{u}}_f)} = 0, \qquad \bm{v}_f =\bm{E}\bm{v}(\bm{u}_h).
-\]
-\end{theorem}
-\begin{proof}
-The proof uses similar techniques as proofs in other papers \cite{chen2017entropy, crean2018entropy, chan2017discretely, chan2019skew}, and we repeat it for completeness.
-We begin by testing (\ref{eq:esdg1D}) with the vector of entropy variables evaluated at nodal points $\bm{v}(\bm{u}_h)$.
-We apply the chain rule in time and use that $\bm{M}$ is diagonal to show that
-\[
-\bm{v}(\bm{u}_h)^T\bm{M}\td{\bm{u}}{t} = \bm{1}^T\bm{M}\LRp{\bm{v}(\bm{u}_h)\circ \td{\bm{u}_h}{t}} =  \bm{1}^T\bm{M}\td{S(\bm{u}_h)}{t}.
-\]
-For the volume term, we use the SBP property to arrive at
-\begin{align*}
-\bm{v}(\bm{u}_h)^T\begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T\LRp{2\bm{Q}_h \circ \bm{F}}\bm{1} &= \tilde{\bm{v}}^T\LRp{\LRp{\bm{Q}_h-\bm{Q}_h^T} \circ \bm{F}}\bm{1}  + \LRp{\bm{E}\bm{v}(\bm{u}_h)}^T\LRp{\bm{B}_h \circ\bm{F}}\bm{1},
-\end{align*}
-where we have introduced $\tilde{\bm{v}} = \LRp{\begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}\bm{v}(\bm{u}_h)}$.  
-Note that, by the consistency of the entropy conservative flux $\bm{f}_S$, the diagonal of $\bm{F}$ is the vector $\bm{f}(\tilde{\bm{u}}_f)$.
-Using that $\bm{B}$ is diagonal, the latter term reduces to
-\[
-\LRp{\bm{E}\bm{v}(\bm{u}_h)}^T\LRp{\bm{B}_h \circ\bm{F}}\bm{1} = \bm{v}_f^T \bm{B}\bm{f}(\tilde{\bm{u}}_f) =  \bm{1}^T \bm{B}\LRp{\bm{v}_f^T\bm{f}(\tilde{\bm{u}}_f)}.
-\]
-The final part of the proof uses that 
-\begin{align*}
-\tilde{\bm{v}}^T\LRp{\LRp{\bm{Q}_h-\bm{Q}_h^T} \circ \bm{F}}\bm{1} &= \sum_j {\tilde{\bm{v}}}_j \LRp{\bm{Q}_h-\bm{Q}_h^T}_{ij} \bm{F}_{ij}\\
-&= \sum_j \LRp{\bm{Q}_h}_{ij} \LRp{{\tilde{\bm{v}}}_i-{\tilde{\bm{v}}}_j}^T\bm{f}_{S}\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}\\
-&= \sum_j \LRp{\bm{Q}_h}_{ij} \LRp{\psi(\tilde{\bm{u}}_i)-\psi(\tilde{\bm{u}}_j)} = \tilde{\bm{u}}^T\bm{Q}_h\bm{1}- \bm{1}^T\bm{Q}_h\tilde{\bm{u}}
-\end{align*}
-The term $\bm{Q}_h\bm{1} = \bm{0}$ by (\ref{eq:Qh1}).
-The proof is completed by applying the SBP property (\ref{eq:hsbp}) to the remaining term, which reduces to $-\bm{1}^T\bm{B}_h\tilde{\bm{u}}=-\bm{1}^T\bm{B}\tilde{\bm{u}_f}$.
-\qed\end{proof}
-
-Note that the proof of entropy conservation for (\ref{eq:esdg1D}) relies mainly on the fact that $\bm{Q}_h$ is conservative (e.g., exact for constants) (\ref{eq:Qh1}) and that $\bm{Q}_h$ satisfies the SBP property (\ref{eq:hsbp}).
-Extensions to the multi-element, multi-dimensional, and non-conforming setting will utilize the same properties.
-
-%\note{Maybe add proof here and show how rest of proofs just rely on SBP and $\bm{Q}_h\bm{1} = \bm{0}$?}.
-For proofs of high order accuracy and entropy conservation, we refer the reader to \cite{chan2018efficient}.
-
-%The matrix $\bm{F}$ is formed by computing ${\bm{F}}_{ij} = \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}$ on the fly.
-Each entry requires evaluating the entropy conservative numerical flux $\bm{f}_S$, resulting in a large number of floating point operations \cite{wintermeyer2018entropy}.
-However, due to properties of the Hadamard product in (\ref{eq:esdg1D}), flux evaluations are only necessary between pairs of nodes corresponding to non-zero entries of $\bm{Q}_h$.
-We will revisit this point later in the construction of entropy stable schemes on non-conforming meshes.
-
-\subsection{Entropy conservation formulations in higher dimensions}
+where we have used that $\bm{E}\bm{1} = \bm{1}$ (since $\bm{E}$ is a high order accurate boundary interpolation matrix), $\bm{Q}\bm{1} = \bm{0}$ (since $\bm{Q}$ is a differentiation matrix), and the GSBP property $-\bm{Q}^T + \bm{E}^T\bm{B}\bm{E} = \bm{Q}$
+
+\subsection{Hybridized SBP operators in higher dimensions}
 
 The formulation (\ref{eq:weakentropyineq}) can be extended to higher dimensions through a tensor product construction.
-For simplicity, we illustrate this in 2D (the extension to 3D is straightforward).
+For simplicity, we illustrate this for 2D quadrilateral elements (the extension to 3D hexahedral elements is straightforward).
 Let $\bm{M}_{\rm 1D}, \bm{Q}_{\rm 1D}$ denote one-dimensional generalized SBP norm (mass) and differentiation matrices, and let $\bm{E}_{\rm 1D}$ denote the 1D face interpolation matrix.
@@
 where $\bm{I}_2$ is the $2\times 2$ identity matrix.  Recall that $\bm{M}_{\rm 1D}$ is the diagonal matrix of quadrature weights, so $\hat{\bm{B}}_i$ are diagonal.
+
+The construction of operators in (\ref{eq:kron}) and (\ref{eq:EBdefs}) correspond to the use of tensor product points quadrature points and \emph{aligned} surface points.
+For example, if Gauss points are used for volume quadrature, we assume that Gauss points are also used for surface quadrature.
+This ensures that the operators $\hat{\bm{Q}}_i, \hat{\bm{B}}_i, \bm{E}$ satisfy the analogous higher dimensional generalized SBP property $\hat{\bm{Q}}_i = \bm{E}^T\hat{\bm{B}}_i\bm{E} - \hat{\bm{Q}}_i^T$.
 
@@
 \end{lemma}
-The SBP property follows by construction of $\hat{\bm{Q}}_{i,h}$.
-The property $\hat{\bm{Q}}_{i,h}\bm{1} = \bm{0}$ is a consequence of the Kronecker product and the same arguments used to derive (\ref{eq:Qh1}).
+The proof is contained in \cite{chan2018efficient}.
+The SBP property follows from the construction of $\hat{\bm{Q}}_{i,h}$, and $\hat{\bm{Q}}_{i,h}\bm{1} = \bm{0}$ is a consequence of the Kronecker product and the same arguments used to derive (\ref{eq:Qh1}).
 The construction of 3D differentiation and interpolation matrices proceeds similarly, and the matrices satisfy the same properties as described in Lemma~\ref{lemma:hsbprefprops}.
 
-Then, a $d$-dimensional entropy conservative formulation on the reference element $[-1,1]^d$ is given as 
-\begin{gather}
-\hat{\bm{M}}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T
-\sum_{i=1}^d \LRp{2\hat{\bm{Q}}_{i,h} \circ \bm{F}_i}\bm{1} + \bm{E}^T\hat{\bm{B}}_i\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_f)} = 0 \label{eq:esdgref}\\
-\LRp{\bm{F}_i}_{jk} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}
-\end{gather}
-where $\bm{f}^i_S$ is the numerical flux for the $i$th coordinate direction, and $\tilde{\bm{u}},\tilde{\bm{u}}_f$ are defined as in (\ref{eq:esdg1D}).
-
-Finally, we can extend this formulation to the multi-element case.
+%Then, a $d$-dimensional entropy conservative formulation on the reference element $[-1,1]^d$ is given as 
+%\begin{gather}
+%\hat{\bm{M}}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T
+%\sum_{i=1}^d \LRp{2\hat{\bm{Q}}_{i,h} \circ \bm{F}_i}\bm{1} + \bm{E}^T\hat{\bm{B}}_i\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_f)} = 0 \label{eq:esdgref}\\
+%\LRp{\bm{F}_i}_{jk} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}
+%\end{gather}
+%where $\bm{f}^i_S$ is the numerical flux for the $i$th coordinate direction, and $\tilde{\bm{u}},\tilde{\bm{u}}_f$ are defined as in (\ref{eq:esdg1D}).
+The proof is given in \cite{chan2017discretely, chan2018efficient}, and follows by applying the proof of Theorem~\ref{thm:esdg1D} to each coordinate direction.
+
+%Finally, we can extend this formulation to the multi-element case.  
+We now construct an entropy conservative formulation on an unstructured curved mesh.  
 Suppose the domain $\Omega$ is decomposed into non-overlapping elements $D^k$ which are images of the reference mapping, such that $D^k$ is a differentiable mapping of the reference element $\hat{D} = [-1,1]^d$.
@@
 \sum_{j=1}^d \pd{}{\hat{x}_j} G_{ij} = 0.
-%\sum_{j=1}^d \hat{\bm{Q}}_{j,h} \bm{G}_{ij} = \bm{0}.
 \label{eq:dgcl}
@@
 
-Because the reference matrices $\hat{\bm{Q}}_{j,h}$ satisfy the reference SBP properties (\ref{eq:hsbp}), one can show \cite{chan2018discretely} that the physical matrices $\bm{Q}_{i,h}$ satisfy analogous physical SBP properties.
+Because the reference matrices $\hat{\bm{Q}}_{j,h}$ satisfy the reference SBP properties (\ref{eq:hsbp}), one can show \cite{chan2018discretely} that the physical matrices $\bm{Q}_{i,h}$ satisfy analogous SBP properties.
 \begin{lemma}
@@
 Suppose the geometric terms $\bm{G}_{ij}$ satisfy the discrete GCL (\ref{eq:dgcl}), the normals are constructed via (\ref{eq:nJ_Gnhat}), and that $\bm{Q}_{i,h}$ is constructed using (\ref{eq:curvedQ}).
-Then, %the conditions for Lemma~\ref{lemma:hsbprefprops} hold.
-Then, %$\hat{\bm{Q}}_j$ satisfies the properties %a generalized SBP property% \note{finish here}
+Then,
 \[
@@
 \]
-where $\bm{n}_i$ is a vector whose entries consist of values of the $i$th component of the scaled normals $n_iJ_f$ at face points, $\bm{w}_f$ is the vector of face quadrature weights $w_i$, and $\bm{B}_i = \diag{\bm{n}_i \circ \bm{w}_f}$.
+where $\bm{n}_i$ is a vector whose entries are values of the $i$th component of the scaled normals $n_iJ_f$ at face points, $\bm{w}_f$ is the vector of face quadrature weights $w_i$, and $\bm{B}_i = \diag{\bm{n}_i \circ \bm{w}_f}$.
 \end{lemma}
-%\begin{proof}
+\begin{proof}
 The proof is identical to that of \cite{chan2018discretely}.  
-%We review it here for completeness.  Expanding out $\bm{Q}_{i,h} + \bm{Q}_{i,h}^T$ yields
-%\[
-%\frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h} + \hat{\bm{Q}}_{j,h}^T\diag{\bm{G}_{ij}} + \hat{\bm{Q}}_{j,h} \diag{\bm{G}_{ij}} +  \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h}^T.
-%\]
-%For each term in the sum, we apply the SBP property of $\hat{\bm{Q}}_{j,h}$.  The volume terms cancel, leaving only surface terms
-%\[
-%\frac{1}{2}\sum_{j=1}^d {\hat{\bm{B}}_{j,h}\diag{\bm{G}_{ij}} +  \diag{\bm{G}_{ij}}\hat{\bm{B}}_{j,h}} = \begin{bmatrix}
-%\bm{0} &\\
-%& \bm{B}_i \end{bmatrix}.
-%\]
-%where the last step uses (\ref{eq:nJ_Gnhat}) and that $\hat{\bm{B}}_{j,h}$ and $\diag{\bm{G}_{ij}}$ are diagonal.  
-%
-%To show $\bm{Q}_{i,h}\bm{1} = \bm{0}$, we note that
-%\[
-%\bm{Q}_{i,h}\bm{1} = \frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h}\bm{1} + \hat{\bm{Q}}_{j,h} \diag{\bm{G}_{ij}}\bm{1} = \frac{1}{2}\sum_{j=1}^d \hat{\bm{Q}}_{j,h} \bm{G}_{ij}
-%\]
-%since $\hat{\bm{Q}}_{j,h}\bm{1} = \bm{0}$ by Lemma~\ref{lemma:hsbprefprops}.
-Note that, since $\bm{G}_{ij}$ is constructed using polynomial interpolation, $\bm{G}_{ij} = \LRs{\bm{g}_{ij}, \quad \bm{E}\bm{g}_{ij}}^T$, where $\bm{g}_{ij}$ denotes the values of $G_{ij}$ at volume points.
+We review it here for completeness.  Expanding out $\bm{Q}_{i,h} + \bm{Q}_{i,h}^T$ yields
+\[
+\frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h} + \hat{\bm{Q}}_{j,h}^T\diag{\bm{G}_{ij}} + \hat{\bm{Q}}_{j,h} \diag{\bm{G}_{ij}} +  \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h}^T.
+\]
+For each term in the sum, we apply the SBP property of $\hat{\bm{Q}}_{j,h}$.  The volume terms cancel, leaving only surface terms
+\[
+\frac{1}{2}\sum_{j=1}^d \LRp{\hat{\bm{B}}_{j,h}\diag{\bm{G}_{ij}} +  \diag{\bm{G}_{ij}}\hat{\bm{B}}_{j,h}} = \begin{bmatrix}
+\bm{0} &\\
+& \bm{B}_i \end{bmatrix}.
+\]
+where we have used (\ref{eq:nJ_Gnhat}) and that $\hat{\bm{B}}_{j,h}$ and $\diag{\bm{G}_{ij}}$ are diagonal matrices.
+
+To show $\bm{Q}_{i,h}\bm{1} = \bm{0}$, we note that
+\[
+\bm{Q}_{i,h}\bm{1} = \frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h}\bm{1} + \hat{\bm{Q}}_{j,h} \diag{\bm{G}_{ij}}\bm{1} = \frac{1}{2}\sum_{j=1}^d \hat{\bm{Q}}_{j,h} \bm{G}_{ij}
+\]
+since $\hat{\bm{Q}}_{j,h}\bm{1} = \bm{0}$ by Lemma~\ref{lemma:hsbprefprops}.
+Note that, since $\bm{G}_{ij}$ is constructed using polynomial interpolation, $\bm{G}_{ij} = \LRs{G_{ij}(\hat{\bm{x}}), \quad \bm{E}G_{ij}(\hat{\bm{x}})}^T$, where $G_{ij}(\hat{\bm{x}})$ denotes the values of $G_{ij}$ at volume points.
 The remaining term can be expanded out
-%\[
-%\sum_{j=1}^d \hat{\bm{Q}}_{j,h} \bm{G}_{ij} = \sum_{j=1}^d \begin{bmatrix} \hat{\bm{Q}}_j\bm{g}_{ij} - \hat{\bm{g}}_j^T\bm{G}_{ij} + \bm{E}^T\bm{B}\bm{E}\bm{g}_{ij}\\
-%\bm{B}\bm{E}\bm{g}_{ij} - \bm{B}\bm{E}\bm{g}_{ij}
-%\end{bmatrix} =
-%\sum_{j=1}^d \begin{bmatrix} - \hat{\bm{g}}_j^T\bm{G}_{ij} + \bm{E}^T\bm{B}\bm{E}\bm{g}_{ij}\\
-%\bm{0}
-%\end{bmatrix} 
-%\]
-%where we have used that $\sum_{j=1}^d\hat{\bm{Q}}_j\bm{g}_{ij} = \bm{0}$, since $G_{ij}$ satisfies the discrete GCL (\ref{eq:dgcl}), $\bm{g}_{ij}$ is a polynomial of degree less than or equal to $N$, and $\hat{\bm{Q}}_j$ exactly differentiates polynomials of degree $N$.
-The remaining terms vanish after an application of the SBP property and the GCL (\ref{eq:dgcl}).
-%\qed\end{proof}
-
-Then, an entropy conservative scheme is given by
+\begin{align*}
+\sum_{j=1}^d \hat{\bm{Q}}_{j,h} \bm{G}_{ij} &= \sum_{j=1}^d \begin{bmatrix} \hat{\bm{Q}}_jG_{ij}(\hat{\bm{x}}) - \hat{\bm{Q}}_j^TG_{ij}(\hat{\bm{x}}) + \bm{E}^T\bm{B}\bm{E}G_{ij}(\hat{\bm{x}})\\
+\bm{B}\bm{E}G_{ij}(\hat{\bm{x}}) - \bm{B}\bm{E}G_{ij}(\hat{\bm{x}})
+\end{bmatrix} \\
+&=
+\sum_{j=1}^d \begin{bmatrix} \LRp{- \hat{\bm{Q}}_j^T + \bm{E}^T\bm{B}\bm{E}}G_{ij}(\hat{\bm{x}})\\
+\bm{0}
+\end{bmatrix} = 
+\sum_{j=1}^d \begin{bmatrix} \hat{\bm{Q}}_jG_{ij}(\hat{\bm{x}})\\
+\bm{0}
+\end{bmatrix}  = \bm{0},
+\end{align*}
+where we have used  the generalized SBP property and that $\sum_{j=1}^d\hat{\bm{Q}}_jG_{ij}(\hat{\bm{x}}) = \bm{0}$, since $G_{ij}$ satisfies the discrete GCL (\ref{eq:dgcl}), $G_{ij}(\hat{\bm{x}})$ is a polynomial of degree less than or equal to $N$, and $\hat{\bm{Q}}_j$ exactly differentiates polynomials of degree $N$.
+\qed\end{proof}
+
+\subsection{Entropy conservative formulations on conforming meshes of mapped elements}
+
+The aforementioned matrices can now be used to construct high order accurate entropy stable and entropy conservative schemes.
+We first introduce an entropy conservative numerical flux \cite{tadmor1987numerical}.
+Let $\bm{u}_L, \bm{u}_R$ denote left and right states in the conservative variables.
+Then, an entropy conservative numerical flux is a vector-valued function $\bm{f}_{i,S}$ for $i = 1,\ldots,d$ which satisfies the following three properties
+\begin{gather*}
+\bm{f}_{i,S}\LRp{\bm{u},\bm{u}} = \bm{f}_i(\bm{u}), \qquad \text{(consistency)}\\
+\bm{f}_{i,S}\LRp{\bm{u}_L,\bm{u}_R} = \bm{f}_{i,S}(\bm{u}_R,\bm{u}_L), \qquad \text{(symmetry)}\\
+\LRp{\bm{v}_L-\bm{v}_R}^T\bm{f}_{i,S}\LRp{\bm{u}_L,\bm{u}_R} = \psi_i(\bm{u}_L) - \psi_i(\bm{u}_R), \qquad \text{(conservation)}.
+\end{gather*}
+Here, $\bm{v}_L, \bm{v}_R$ are the entropy variables evaluated at $\bm{u}_L, \bm{u}_R$, and $\psi_i$ is the $i$th entropy potential which appears in (\ref{eq:weakentropyineq}).  
+Then, an entropy conservative scheme on a mapped element is given by
 \begin{gather}
 {\bm{M}}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T
-\sum_{i=1}^d \LRp{2{\bm{Q}}_{i,h} \circ \bm{F}_i}\bm{1} + \bm{E}^T{\bm{B}}_i\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_f)} = 0 \label{eq:esdg}
+\sum_{i=1}^d \LRp{2{\bm{Q}}_{i,h} \circ \bm{F}_i}\bm{1} + \bm{E}^T{\bm{B}}_i\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_f)} = 0 \label{eq:esdg}\\
+\LRp{\bm{F}_i}_{jk} = \bm{f}_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k} \nonumber
 \end{gather}
-where $\bm{F}_i$ is defined as in (\ref{eq:esdgref}).
-Here, the exterior state $\tilde{\bm{u}}_f$ used to evaluate the numerical flux $\bm{f}_i^*$ corresponds either to interface values on a neighboring element or an exterior state used to enforce boundary conditions.
+Here, the exterior state $\tilde{\bm{u}}_f$ used to evaluate the numerical flux $\bm{f}_i^*$ corresponds either to interface values on a neighboring element or an exterior state used to enforce boundary conditions, and $\circ$ denotes the matrix Hadamard product.
+
+%\note{Mention proof}
+%Then, an entropy conservative scheme on the reference interval is given as
+%\begin{gather}
+%\bm{M}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T
+%\LRp{2\bm{Q}_h \circ \bm{F}}\bm{1} + \bm{E}^T\bm{B}\LRp{\bm{f}^*-\bm{f}(\tilde{\bm{u}}_f)} = 0 \label{eq:esdg1D}\\ 
+%\tilde{\bm{u}} = \begin{bmatrix}
+%\bm{u}_h\\
+%\tilde{\bm{u}}_f
+%\end{bmatrix}, \qquad \tilde{\bm{u}}_f = \bm{u}\LRp{\bm{E}\bm{v}(\bm{u}_h)},\nonumber \\ 
+%\LRp{\bm{F}}_{jk} = \bm{f}_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}, \qquad \bm{f}^* = \bm{f}_S\LRp{\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f}.\nonumber
+%\end{gather}
+%Here, $\tilde{\bm{u}}_f^+$ denotes some exterior state which can be used to enforce boundary conditions or coupling between different elements
+
+\begin{theorem}
+\label{thm:esdg}
+Assuming continuity in time, the formulation (\ref{eq:esdg}) is entropy conservative in the sense that
+\[
+\bm{1}^T\bm{M}\td{S(\bm{u}_h)}{t} + \bm{1}^T\bm{B}_i\LRp{\bm{v}_f^T\bm{f}_i^* - \psi_i(\tilde{\bm{u}}_f)} = 0, \qquad \bm{v}_f =\bm{E}\bm{v}(\bm{u}_h).
+\]
+\end{theorem}
+\begin{proof}
+The proof uses similar techniques as proofs in other papers \cite{chen2017entropy, crean2018entropy, chan2017discretely, chan2019skew}, and we repeat it for completeness.
+We begin by testing (\ref{eq:esdg}) with the vector of entropy variables evaluated at nodal points $\bm{v}(\bm{u}_h)$.
+We apply the chain rule in time and use that $\bm{M}$ is diagonal to show that
+\[
+\bm{v}(\bm{u}_h)^T\bm{M}\td{\bm{u}}{t} = \bm{1}^T\bm{M}\LRp{\bm{v}(\bm{u}_h)\circ \td{\bm{u}_h}{t}} =  \bm{1}^T\bm{M}\td{S(\bm{u}_h)}{t}.
+\]
+For the volume term, we use the SBP property to arrive at
+\begin{align*}
+&\bm{v}(\bm{u}_h)^T\begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T\LRp{2\bm{Q}_{i,h} \circ \bm{F}_i}\bm{1} = \\
+&\tilde{\bm{v}}^T\LRp{\LRp{\bm{Q}_{i,h}-\bm{Q}_{i,h}^T} \circ \bm{F}_i}\bm{1}  + \LRp{\bm{E}\bm{v}(\bm{u}_h)}^T\LRp{\bm{B}_{i,h} \circ\bm{F}_i}\bm{1},
+\end{align*}
+where we have introduced $\tilde{\bm{v}} = \LRp{\begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}\bm{v}(\bm{u}_h)}$.  
+Note that, by the consistency of the entropy conservative flux $\bm{f}_{i,S}$, the diagonal of $\bm{F}_i$ is the vector $\bm{f}_i(\tilde{\bm{u}}_f)$.
+Using that $\bm{B}_{i,h}$ is diagonal (with non-zero diagonal sub-block $\bm{B}_i$), the latter term reduces to
+\[
+\LRp{\bm{E}\bm{v}(\bm{u}_h)}^T\LRp{\bm{B}_{i,h} \circ\bm{F}_i}\bm{1} = \bm{v}_f^T \bm{B}_i\bm{f}_i(\tilde{\bm{u}}_f) =  \bm{1}^T \bm{B}_i\LRp{\bm{v}_f^T\bm{f}_i(\tilde{\bm{u}}_f)}.
+\]
+The final part of the proof uses that 
+\begin{align*}
+\tilde{\bm{v}}^T\LRp{\LRp{\bm{Q}_{i,h}-\bm{Q}_{i,h}^T} \circ \bm{F}_i}\bm{1} &= \sum_j {\tilde{\bm{v}}}_j \LRp{\bm{Q}_{i,h}-\bm{Q}_{i,h}^T}_{jk} \LRp{\bm{F}_{i}}_{jk}\\
+&= \sum_j \LRp{\bm{Q}_h}_{jk} \LRp{{\tilde{\bm{v}}}_j-{\tilde{\bm{v}}}_k}^T\bm{f}_{i,S}\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}\\
+&= \sum_j \LRp{\bm{Q}_h}_{jk} \LRp{\psi(\tilde{\bm{u}}_j)-\psi(\tilde{\bm{u}}_k)} = \tilde{\bm{u}}^T\bm{Q}_{i,h}\bm{1}- \bm{1}^T\bm{Q}_{i,h}\tilde{\bm{u}}
+\end{align*}
+The term $\bm{Q}_{i,h}\bm{1} = \bm{0}$ by (\ref{eq:Qh1}).
+The proof is completed by applying the SBP property (\ref{eq:hsbp}) to the remaining term, which reduces to $-\bm{1}^T\bm{B}_{i,h}\tilde{\bm{u}}=-\bm{1}^T\bm{B}_i\tilde{\bm{u}_f}$.
+\qed\end{proof}
+
+Note that the proof of entropy conservation relies mainly on Lemma~\ref{lemma:Qhprops}, which states that $\bm{Q}_{i,h}$ is conservative (e.g., exact for constants) and satisfies the SBP property.
+Extensions to the non-conforming setting will utilize the same properties.
 
@@
 Moreover, since $u$ is degree $\min(N,N_f,N_m)$, it is also computed exactly using the face (surface) quadrature.
-Thus, $\bm{E}_{mf}^T\bm{M}_m\bm{E}_{mf} = \bm{M}_f \bm{u}_f$ by exactness of the face and mortar quadratures, and $\bm{E}_{fm}\bm{u}_m = \bm{u}_f$.
+Thus, $\bm{E}_{mf}^T\bm{M}_m\bm{E}_{mf} \bm{u}_f = \bm{M}_f \bm{u}_f$ by exactness of the face and mortar quadratures, and $\bm{E}_{fm}\bm{u}_m = \bm{u}_f$.
 \qed\end{proof}
@@
 Since it is not clear at first glance how the hybridized operator $\hat{\bm{Q}}_{i,h}$ or its mortar-based variant $\hat{\bm{Q}}_{i,m}$ can be used to perform differentiation, we review intuitive explanations of each operator.
+
 Let $\bm{u}$ denote basis coefficients for some function $u(\bm{x})$, and let $f, g$ denote two functions on the reference element.
-Suppose $\bm{u}$ satisfies the following matrix system involving the hybridized operator $\hat{\bm{Q}}_{i,h}$
+Suppose $\bm{u}$ satisfies the following matrix system involving the standard hybridized operator $\hat{\bm{Q}}_{i,h}$
 \begin{align}
@@
 Let $\bm{v}$ denote basis coefficients for some arbitrary degree $N$ polynomial test function.
-Multiplying (\ref{eq:Qh}) by $\bm{v}^T$, expanding out the different blocks of $\hat{\bm{Q}}_{i,h}$, and using summation by parts yields
+Expanding out the different blocks of (\ref{eq:Qh}) and using generalized summation by parts for $\hat{\bm{Q}}_i$ yields
 \begin{align*}
 \hat{\bm{M}}\bm{u} &= 
-\diag{\bm{f}}\hat{\bm{Q}}_i\bm{g} + \frac{1}{2}\LRp{\diag{f(\hat{\bm{x}})}\bm{E}^T + \bm{E}^T f(\hat{\bm{x}}_f)}\hat{\bm{B}}_i \LRp{g(\hat{\bm{x}}_f) - \bm{E} g(\hat{\bm{x}})}.
+\diag{f(\hat{\bm{x}})}\hat{\bm{Q}}_i g(\hat{\bm{x}}) + \frac{1}{2}\LRp{\diag{f(\hat{\bm{x}})}\bm{E}^T + \bm{E}^T f(\hat{\bm{x}}_f)}\hat{\bm{B}}_i \LRp{g(\hat{\bm{x}}_f) - \bm{E} g(\hat{\bm{x}})}.
 \end{align*}
@@
 
-The mortar-based hybridized SBP operator can be interpreted similarly.
+The mortar-based hybridized SBP operator can be similarly interpreted as a high order differentiation operator.
 Recall that $\bm{E}_{mf}$ interpolates from face nodes to mortar nodes, and that $\bm{E}$ interpolates from volume nodes to face nodes.
@@
 \begin{align}
-\hat{\bm{M}}\bm{u} &= \diag{\bm{f}}\hat{\bm{Q}}_i\bm{g} \label{eq:mhsbp_correction}\\
+\hat{\bm{M}}\bm{u} &= \diag{f(\hat{\bm{x}})}\hat{\bm{Q}}_i{g}(\hat{\bm{x}}) \label{eq:mhsbp_correction}\\
 &+ \frac{1}{2} \diag{f(\hat{\bm{x}})}\bm{E}^T\hat{\bm{B}}_{i,f}\LRp{g(\hat{\bm{x}}_f) - \bm{E}g(\hat{\bm{x}})} \nonumber\\
@@
 
-We can now construct an entropy stable formulation on the reference element using the mortar-based hybridized SBP operator
-\begin{gather}
-\hat{\bm{M}}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \\ \bm{E}_{mf}\bm{E} \end{bmatrix}^T
-\sum_{i=1}^d \LRp{2\hat{\bm{Q}}_{i,m} \circ \bm{F}_i}\bm{1} + \bm{E}_{mf}^T\hat{\bm{B}}_{i,m}\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_m)} = 0 \label{eq:esdgmref}\\
-\tilde{\bm{u}} = \begin{bmatrix}
-\bm{u}_h\\
-\tilde{\bm{u}}_f\\
-\tilde{\bm{u}}_m
-\end{bmatrix}, \qquad \tilde{\bm{u}}_m = \bm{u}\LRp{\bm{E}_{mf}\bm{E}\bm{v}(\bm{u}_h)}, \qquad \bm{f}^* = \bm{f}_S\LRp{\tilde{\bm{u}}_m^+,\tilde{\bm{u}}_m},
-\end{gather}
-where $\LRp{\bm{F}_i}_{jk} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}$ for all indices $j,k$ of $\tilde{\bm{u}}$.
+%We can now construct an entropy conservative formulation on the reference element using the mortar-based hybridized SBP operator
+%\begin{gather}
+%\hat{\bm{M}}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \\ \bm{E}_{mf}\bm{E} \end{bmatrix}^T
+%\sum_{i=1}^d \LRp{2\hat{\bm{Q}}_{i,m} \circ \bm{F}_i}\bm{1} + \bm{E}_{mf}^T\hat{\bm{B}}_{i,m}\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_m)} = 0 \label{eq:esdgmref}\\
+%\tilde{\bm{u}} = \begin{bmatrix}
+%\bm{u}_h\\
+%\tilde{\bm{u}}_f\\
+%\tilde{\bm{u}}_m
+%\end{bmatrix}, \qquad \tilde{\bm{u}}_m = \bm{u}\LRp{\bm{E}_{mf}\bm{E}\bm{v}(\bm{u}_h)}%, \qquad \bm{f}^* = \bm{f}_S\LRp{\tilde{\bm{u}}_m^+,\tilde{\bm{u}}_m},
+%\end{gather}
+%where $\LRp{\bm{F}_i}_{jk} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}$ for all indices $j,k$ of $\tilde{\bm{u}}$, and $\bm{f}_i^*$ is a numerical flux used to impose boundary conditions.
+%
+%\note{Add statement of entropy conservation}
 
@@
 \end{equation}
-Proofs of entropy stability for formulations (\ref{eq:esdg1D}), (\ref{eq:esdg}), and (\ref{eq:esdgmref}) require that the hybridized SBP operators are conservative (e.g., they are exact for constants) and satisfy the SBP property.
+Proofs of entropy conservation (e.g., for the formulation (\ref{eq:esdg})) require that the hybridized SBP operators are conservative (e.g., exact for constants) and satisfy the SBP property.
 Entropy conservation for mortar-based formulations on mapped elements will require similar properties.
@@
 \]
-We denote $Q^{N,\ldots,N} = Q^N$ for conciseness.
+We denote the isotropic tensor product space $Q^N= Q^{N,\ldots,N}$ for conciseness.
 
@@
 \end{align*}
-for $i = 1,2$.  Moreover, the geometric terms naturally satisfy the GCL condition (\ref{eq:dgcl}).  We then have the following lemma:
-\begin{lemma}
-\label{lemma:Qmprops_2d}
-Suppose $D^k$ is a quadrilateral element, and suppose that the surface quadrature is exact for $Q^{N+N_f}$ and mortar quadratures are exact for degree $Q^{N+N_m}$ polynomials.
+for $i = 1,2$, and the geometric terms naturally satisfy the GCL condition (\ref{eq:dgcl}).  
+%We then have the following lemma:
+%\begin{lemma}
+%\label{lemma:Qmprops_2d}
+%Suppose $D^k$ is a quadrilateral element, and suppose that the surface quadrature is exact for degree ${N+N_f}$ polynomials and mortar quadratures are exact for degree ${N+N_m}$ polynomials.
 Then, if $\bm{Q}_{i,m}$ is constructed using (\ref{eq:curvedQ_mortar}) and if $N_{\rm geo} \leq \min(N,N_f+1,N_m+1)$,
-\begin{align*}
-\bm{Q}_{i,m} + \bm{Q}_{i,m}^T &= \bm{B}_{i,m}, \quad \text{(SBP property)}\\
-\bm{Q}_{i,m}\bm{1} &= \bm{0}, \quad \text{(conservation)}.
-\end{align*}
-\end{lemma}
-\begin{proof}
-%The proof mimicks that of Lemma~\ref{lemma:Qhprops}.  
-The SBP property is straightforward to show using the SBP property of $\hat{\bm{Q}}_{i,m}$.  To show conservation, expanding out $\bm{Q}_{i,m}\bm{1} $ yields
-\[
-\bm{Q}_{i,m}\bm{1} = \frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,m}\bm{1} + \hat{\bm{Q}}_{j,m} \diag{\bm{G}_{ij}}\bm{1} = \frac{1}{2}\sum_{j=1}^d\hat{\bm{Q}}_{j,m} \bm{G}_{ij}
-\]
-since $\hat{\bm{Q}}_{j,m}\bm{1} = \bm{0}$ by Lemma~\ref{lemma:mhsbp}.
+%\begin{align*}
+%\bm{Q}_{i,m} + \bm{Q}_{i,m}^T &= \bm{B}_{i,m}, \quad \text{(SBP property)}\\
+%\bm{Q}_{i,m}\bm{1} &= \bm{0}, \quad \text{(conservation)}.
+%\end{align*}
+%\end{lemma}
+%\begin{proof}
+%%The proof mimicks that of Lemma~\ref{lemma:Qhprops}.  
+%The SBP property is straightforward to show using the SBP property of $\hat{\bm{Q}}_{i,m}$.  To show conservation, expanding out $\bm{Q}_{i,m}\bm{1} $ yields
+%\[
+%\bm{Q}_{i,m}\bm{1} = \frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,m}\bm{1} + \hat{\bm{Q}}_{j,m} \diag{\bm{G}_{ij}}\bm{1} = \frac{1}{2}\sum_{j=1}^d\hat{\bm{Q}}_{j,m} \bm{G}_{ij}
+%\]
+%since $\hat{\bm{Q}}_{j,m}\bm{1} = \bm{0}$ by Lemma~\ref{lemma:mhsbp}.
 We can expand out the remaining terms using (\ref{eq:mhsbp_correction}).
 However, since $\bm{G}_{ij}$ is constructed by interpolating the values of the polynomial $G_{ij}$, the boundary correction terms in (\ref{eq:mhsbp_correction}) vanish, reducing the expression to
-\[
-\sum_{j=1}^d\hat{\bm{Q}}_{j,m} \bm{G}_{ij} = \sum_{j=1}^d\hat{\bm{Q}}_{j} \bm{g}_{ij} = \bm{0},
-\]
-where $\bm{g}_{ij}$ denotes the values of the $G_{ij}$ at volume nodes, and the final sum vanishes due to the GCL (\ref{eq:dgcl}) and the fact that $\hat{\bm{Q}}_j$ is exact for degree $N$ polynomials.
-\note{Fix to add in tensor product quadrature.}
-\qed\end{proof}
+%\[
+%\sum_{j=1}^d\hat{\bm{Q}}_{j,m} \bm{G}_{ij} = \sum_{j=1}^d\hat{\bm{Q}}_{j} G_{ij}(\hat{\bm{x}}) = \bm{0},
+%\]
+%where $G_{ij}(\hat{\bm{x}})$ denotes the values of the geometric terms at volume nodes $\hat{\bm{x}}$, and the final sum vanishes due to the GCL (\ref{eq:dgcl}) and the fact that $\hat{\bm{Q}}_j$ is exact for degree $N$ polynomials.
+%\qed\end{proof}
 
@@
 \begin{enumerate}
-\item Approach 1: the construction of $G_{ij}$ from \cite{kopriva2006metric, chan2018efficient}, which yields for $i = 1,2,3$
-\begin{align*}
-&G_{i1} \in Q^{N_{\rm geo}-1,N_{\rm geo},N_{\rm geo}}\\
-&G_{i2} \in Q^{N_{\rm geo},N_{\rm geo}-1,N_{\rm geo}}\\
-&G_{i3} \in Q^{N_{\rm geo},N_{\rm geo},N_{\rm geo}-1}.
-\end{align*}
-\item Approach 2: the construction of $G_{ij}$ from \cite{kozdon2018energy} (Appendix C.3), which yields for $i = 1,2,3$
+\item Approach 1: the construction of $G_{ij}$ from \cite{kopriva2006metric, chan2018efficient}, which yields 
+\begin{align*}
+&G_{ij} \in Q^{N_{\rm geo}}, \qquad i,j = 1,2,3.
+\end{align*}
+\item Approach 2: the construction of $G_{ij}$ from \cite{kozdon2018energy} (Appendix C.3), which yields
 \begin{align*}
@@
 &G_{i2} \in Q^{N_{\rm geo}-1,N_{\rm geo},N_{\rm geo}-1}\\
-&G_{i3} \in Q^{N_{\rm geo}-1,N_{\rm geo}-1,N_{\rm geo}}.
+&G_{i3} \in Q^{N_{\rm geo}-1,N_{\rm geo}-1,N_{\rm geo}}, \qquad i = 1,2,3.
 \end{align*}
@@
 \label{lemma:Qmprops_3d}
-Suppose $D^k$ is a hexahedral element, and suppose that the surface quadrature is exact for $Q^{N+N_f}$, mortar quadratures are exact for degree $Q^{N+N_m}$ polynomials, and that $\bm{Q}_{i,m}$ is constructed using (\ref{eq:curvedQ_mortar}).
-If either
+Suppose $D^k$ is a tensor product element element with tensor product surface quadratures and that $\bm{Q}_{i,m}$ is constructed using (\ref{eq:curvedQ_mortar}).
+Suppose also that the surface quadrature is exact for $Q^{N+N_f}$,  the mortar quadrature is exact for $Q^{N+N_m}$.
+If
 \begin{enumerate}
-\item $G_{ij}$ is constructed using Approach 1 \cite{kopriva2006metric, chan2018efficient} and $N_{\rm geo} \leq \min(N,N_f,N_m)$, or
-\item $G_{ij}$ is constructed using Approach 2 \cite{kozdon2018energy} and $N_{\rm geo} \leq \min(N,N_f-1,N_m-1)$, 
+\item $D^k$ is a quadrilateral element and $N_{\rm geo} \leq \min(N,N_f+1,N_m+1)$, or
+\item $D^k$ is a hexahedral element, $G_{ij}$ is constructed via Approach 1 \cite{kopriva2006metric}, and $N_{\rm geo} \leq \min(N,N_f,N_m)$, or
+\item $D^k$ is hexahedral element, $G_{ij}$ is constructed via Approach 2 \cite{kozdon2018energy}, and $N_{\rm geo} \leq \min(N,N_f+1,N_m+1)$, 
 \end{enumerate}
@@
 \begin{proof}
-\note{Finish} The proof of Lemma~\ref{lemma:Qmprops_3d} is identical to that of Lemma~\ref{lemma:Qmprops_2d}
+Using the SBP property of $\hat{\bm{Q}}_{i,m}$, proving the SBP property follows the same steps as the proof of Lemma~\ref{lemma:Qhprops}.
+To show conservation, expanding out $\bm{Q}_{i,m}\bm{1} $ yields
+\[
+\bm{Q}_{i,m}\bm{1} = \frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,m}\bm{1} + \hat{\bm{Q}}_{j,m} \diag{\bm{G}_{ij}}\bm{1} = \frac{1}{2}\sum_{j=1}^d\hat{\bm{Q}}_{j,m} \bm{G}_{ij}
+\]
+since $\hat{\bm{Q}}_{j,m}\bm{1} = \bm{0}$ by Lemma~\ref{lemma:mhsbp}.  Expanding out remaining terms using (\ref{eq:mhsbp_correction}) yields
+\begin{align*}
+\sum_{j=1}^d \hat{\bm{Q}}_j G_{ij}(\hat{\bm{x}}) &+ \frac{1}{2} \bm{E}^T\hat{\bm{B}}_{j,f}\LRp{G_{ij}(\hat{\bm{x}}_f) - \bm{E}G_{ij}(\hat{\bm{x}})} \\
+&+ \frac{1}{2} \bm{E}^T\hat{\bm{B}}_{j,f}\LRp{\bm{E}_{fm}G_{ij}(\hat{\bm{x}}_m) - \bm{E}G_{ij}(\hat{\bm{x}})} \\
+&+ \frac{1}{2} \bm{E}^T\bm{E}_{mf}^T\hat{\bm{B}}_{j,m}\LRp{G_{ij}(\hat{\bm{x}}_m) - \bm{E}_{mf}G_{ij}(\hat{\bm{x}}_f)}.
+\end{align*}
+where $G_{ij}(\hat{\bm{x}}), G_{ij}(\hat{\bm{x}}_f), G_{ij}(\hat{\bm{x}}_m)$ denote the values of $G_{ij}$ at volume, surface, and mortar points.
+By the fact that $G_{ij} \in Q^N$ and $\bm{E}, \bm{E}_{mf}$ are degree $N$ interpolation operators, the first and third correction terms vanish.
+The remaining boundary correction terms are
+\begin{equation}
+\bm{E}^T\hat{\bm{B}}_{j,f}\LRp{\bm{E}_{fm}G_{ij}(\hat{\bm{x}}_m) - \bm{E}G_{ij}(\hat{\bm{x}})}.  
+\label{eq:bcor}
+\end{equation}
+Recall that $\hat{\bm{B}}_{i,f}$ is a diagonal matrix whose entries are $\hat{\bm{n}}_{i,m}$ scaled by the mortar quadrature weights.
+On tensor product elements, $\hat{\bm{n}}_{i,m}=\pm 1$ on faces where $\hat{x}_i = \pm 1$ and zero otherwise.
+Thus, restricting $\hat{x}_1 = \pm 1$ yields the boundary values of $G_{11}$ on faces where the correction term (\ref{eq:bcor}) is non-zero.
+We can now show that (\ref{eq:bcor}) vanishes for $i = j = 1$ (the cases of $i = 2,\ldots,d$ are similar).
+\begin{enumerate}
+\item If $D^k$ is a quadrilateral element, $G_{11} \in Q^{N_{\rm geo},N_{\rm geo}-1}$, and the surface traces of $G_{11}$ are in the 1D trace space $P^{N_{\rm geo}-1}$.  
+\item If $D^k$ is a hexahedral element and $G_{ij}$ is constructed using Approach 1 \cite{kozdon2018energy}, $G_{11} \in Q^{N_{\rm geo}, N_{\rm geo}-1,N_{\rm geo}-1}$, and the surface traces of $G_{11}$ are in the quadrilateral trace space $Q^{N_{\rm geo}-1,N_{\rm geo}-1}$.
+\end{enumerate}
+Since $N_{\rm geo} \leq N_m+1$, by Lemma~\ref{lemma:Efm} and assumptions on $N_{\rm geo}, N_f, N_m$, $\bm{E}_{fm}G_{ij}(\hat{\bm{x}}_m) = \bm{E}G_{ij}(\hat{\bm{x}}$ and (\ref{eq:bcor}) vanishes.
+
+If Approach 1 \cite{kopriva2006metric} is used, then surface traces of $G_{11}$ are contained in $Q^{N_{\rm geo},N_{\rm geo}}$.
+Then, (\ref{eq:bcor}) vanishes since $N_{\rm geo} \leq \min(N,N_f,N_m)$.
+In both cases, the remaining terms vanish assuming that $G_{ij}$ satisfies the discrete GCL (\ref{eq:dgcl}).
 \qed\end{proof}
@@
 {\bm{M}}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \\ \bm{E}_{mf}\bm{E} \end{bmatrix}^T
-\sum_{i=1}^d \LRp{2{\bm{Q}}_{i,m} \circ \bm{F}_i}\bm{1} + \bm{E}_{mf}^T{\bm{B}}_{i,m}\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_m)} = 0 \label{eq:esdgm}\\
+\sum_{i=1}^d \LRp{2{\bm{Q}}_{i,m} \circ \bm{F}_i}\bm{1} + \bm{E}^T\bm{E}_{mf}^T{\bm{B}}_{i,m}\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_m)} = 0 \label{eq:esdgm}\\
 \tilde{\bm{u}} = \begin{bmatrix}
@@
 
+The proof of entropy conservation is algebraically identical to previous proofs, with Lemma~\ref{lemma:Qmprops_3d} used in place of Lemma~\ref{lemma:Qhprops}.
+
 %\note{Note: line DG approach can work, but it's flipped: we increase quadrature strength in the direction \emph{orthogonal} to derivative instead.
@@
 
-While the formulation presented above is convenient for analysis, it is computationally expensive to implement.
-However, using properties of the mortar matrices $\bm{E}_m, \tilde{\bm{E}}_m$, we can rewrite the above formulation in a way which reflects a traditional mortar-based finite element implementation.
-In other words, we wish to implement (\ref{eq:mortar}) such that the only modification from the implementations in \cite{chan2017discretely} is an intermediate step on mortar faces.
-
-We first note that, since $\bm{E}_m$ maps from surface nodes to mortar nodes, $\bm{V}_m = \bm{E}_m \bm{E}$.  Thus, we can rewrite the numerical flux contribution as
-\begin{align*}
- \bm{V}_m^T\tilde{\bm{B}}_i\bm{f}_i^* &= \bm{E}^T \bm{E}_m^T \bm{M}_m\diag{\hat{\bm{n}}_i} \bm{f}_i^* = \bm{E}^T  \bm{P}_f^T\bm{T}_m^T\bm{M}_m\diag{\hat{\bm{n}}_i} \bm{f}_i^*\\
- &= \bm{E}^T \bm{W}_f \diag{\hat{\bm{n}}_i} \bm{T}_m\bm{M}_m^{-1}\bm{T}_m^T\bm{M}_m \bm{f}_i^* =  \bm{E}^T \bm{B}_i \tilde{\bm{E}}_m\bm{f}_i^*.
-\end{align*}
-
-We can similarly reformulate the remaining mortar contributions within (\ref{eq:mortar}).  We decompose $\bm{F}_S$ into interactions between volume nodes, surface nodes, and mortar nodes
+While the formulation (\ref{eq:esdgm}) is convenient for analysis, it is less convenient to implement.
+However, using properties of the mortar matrices, we can show that the mortar-based formulation (\ref{eq:esdgm}) is equivalent to a conforming formulation (\ref{eq:esdg}) with a modified numerical flux involving face-local correction terms.
+
+We can reformulate the remaining mortar contributions within (\ref{eq:mortar}).  We decompose $\bm{F}_S$ into interactions between volume nodes, surface nodes, and mortar nodes
 \[
 \bm{F}_S = \begin{bmatrix}
-\bm{F}_S^{vv} & \bm{F}_S^{vs} & \bm{F}_S^{vm}\\
-\bm{F}_S^{sv} & \bm{F}_S^{ss} & \bm{F}_S^{sm}\\
-\bm{F}_S^{mv} & \bm{F}_S^{ms} & \bm{F}_S^{mm}
+\bm{F}_S^{vv} & \bm{F}_S^{vf} & \bm{F}_S^{vm}\\
+\bm{F}_S^{fv} & \bm{F}_S^{ff} & \bm{F}_S^{fm}\\
+\bm{F}_S^{mv} & \bm{F}_S^{mf} & \bm{F}_S^{mm}
 \end{bmatrix}
@@
 \begin{align*}
-&\begin{bmatrix} \bm{V}_q \\ \bm{E} \\ \bm{V}_m \end{bmatrix}^T
+&\begin{bmatrix} \bm{I} \\ \bm{E} \\ \bm{E}_m \end{bmatrix}^T
 \LRp{\begin{bmatrix}
 \bm{Q}_i-\bm{Q}_i^T & \bm{E}^T\bm{B}_i &\\
--\bm{B}_i\bm{E} &  & \bm{B}_i\tilde{\bm{E}}_m \\
-& -\tilde{\bm{B}}_i{\bm{E}}_m & 
+-\bm{B}_i\bm{E} &  & \bm{B}_{i}{\bm{E}}_m \\
+& -\bm{B}_{i,m}{\bm{E}}_m & {\bm{B}}_{i,m}
 \end{bmatrix} \circ \bm{F}_S}\bm{1} \\
 &= 
-\begin{bmatrix} \bm{V}_q \\ \bm{E} \end{bmatrix}^T
+\begin{bmatrix} \bm{I}_q \\ \bm{E} \end{bmatrix}^T
 \LRp{\begin{bmatrix}
@@
 
-\subsection{Extensions to $p$ non-conformity, general quadratures, and non-hexahedral elements}
+\subsection{Extensions of the mortar-based approach}
+
+\begin{figure}
+\centering
+\includegraphics[width=.5\textwidth]{noncon3D.png}
+\caption{Two-layer mortars for reduced $O(N^3)$ computational cost on non-conforming hexahedral interfaces.}
+\end{figure}
+
+\note{Extensions to 3D meshes: double anisotropic mortar for $O(N^3)$ operations vs $O(N^4)$.}
 
@@
 
-\note{Explain why this isn't as much of an issue - Shadpey \cite{shadpey2019energy} use this for simplicial elements.
+\note{Explain why non-hexahedral elements isn't as much of an issue - Shadpey \cite{shadpey2019energy} use this for simplicial elements.
 The interpolation operators are dense, and using the mortar-based approach wouldn't actually decrease the number of operations.}
@@
 
-\subsection{Gauss Lobatto quadrature}
-
-\subsection{Gauss quadrature}
+\textbf{Gauss Lobatto quadrature}
+
+\textbf{Gauss quadrature}
 
 \subsection{3D results}
-\subsection{Gauss Lobatto quadrature}
-
-\subsection{Gauss quadrature}
+\textbf{Gauss Lobatto quadrature}
+
+\textbf{Gauss quadrature}
 

### 2020-01-04 docs/ESDG_mortar/main.tex
updates to paper

@@
 \newcommand{\jump}[1] {\ensuremath{\llbracket#1\rrbracket}}
-\newcommand{\avg}[1] {\ensuremath{\LRc{\!\{#1\}\!}}}
+\newcommand{\avg}[1] {\ensuremath{\LRc{\!\LRc{#1}\!}}}
 
@@
 \author{Jesse Chan, Mario Bencomo, David C.\ Del Rey Fernandez}
-\title{Mortar-based entropy stable discontinuous Galerkin methods on non-conforming meshes}
+\title{Mortar-based entropy stable discontinuous Galerkin methods on non-conforming quadrilateral and hexahedral meshes}
 \titlerunning{Mortar-based entropy stable DG methods}
@@
 \begin{equation}
-\bm{n}J_f = G\hat{\bm{n}}\hat{J}_f.
+{n}_iJ_f = \sum_{j=1}^d{G}_{ij}\hat{{n}}_j\hat{J}_f.
 \label{eq:nJ_Gnhat}
@@
 \sum_{i=1}^d \LRp{2{\bm{Q}}_{i,h} \circ \bm{F}_i}\bm{1} + \bm{E}^T{\bm{B}}_i\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_f)} = 0 \label{eq:esdg}\\
-\LRp{\bm{F}_i}_{jk} = \bm{f}_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k} \nonumber
+\LRp{\bm{F}_i}_{jk} = \bm{f}_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}, \qquad \bm{f}_i^* = \bm{f}_{i,S}\LRp{\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f} \nonumber
 \end{gather}
-Here, the exterior state $\tilde{\bm{u}}_f$ used to evaluate the numerical flux $\bm{f}_i^*$ corresponds either to interface values on a neighboring element or an exterior state used to enforce boundary conditions, and $\circ$ denotes the matrix Hadamard product.
+Here, the exterior state $\tilde{\bm{u}}_f^+$ used to evaluate the numerical flux $\bm{f}_i^*$ corresponds either to interface values on a neighboring element or an exterior state used to enforce boundary conditions, and $\circ$ denotes the matrix Hadamard product.
 
@@
 \[
-\bm{1}^T\bm{M}\td{S(\bm{u}_h)}{t} + \bm{1}^T\bm{B}_i\LRp{\bm{v}_f^T\bm{f}_i^* - \psi_i(\tilde{\bm{u}}_f)} = 0, \qquad \bm{v}_f =\bm{E}\bm{v}(\bm{u}_h).
+\bm{1}^T\bm{M}\td{S(\bm{u}_h)}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}_i\LRp{\bm{v}_f^T\bm{f}_i^* - \psi_i(\tilde{\bm{u}}_f)} = 0, \qquad \bm{v}_f =\bm{E}\bm{v}(\bm{u}_h).
 \]
@@
 We introduce dissipation by incorporating a penalization term into the interface flux \cite{winters2017uniquely}.
-For example, Lax-Friedrichs dissipation can be added by modifying the interface flux $\bm{f}^*_i$
-\[
-\bm{f}^*_i \Longrightarrow \bm{f}^*_i - \frac{\lambda}{2} \jump{\tilde{\bm{u}}_f}, \qquad  \jump{\tilde{\bm{u}}_f} = \tilde{\bm{u}}_f^+ - \tilde{\bm{u}}_f,
+For example, Lax-Friedrichs dissipation can be added by modifying the interface flux term $\bm{B}_i\bm{f}^*_i$
+\[
+\bm{B}_i\bm{f}^*_i \Longrightarrow \bm{B}_i\bm{f}^*_i - \frac{\lambda}{2} \jump{\tilde{\bm{u}}_f}, \qquad  \jump{\tilde{\bm{u}}_f} = \tilde{\bm{u}}_f^+ - \tilde{\bm{u}}_f,
 \]
 where $\lambda$ is an estimate of the maximum wave speed \cite{chen2017entropy, chan2017discretely}.
+
+For the remainder of this work, we will construct formulations and prove they are entropy conservative, with the understanding that they can be made entropy stable by incorporating an entropy dissipative penalty term.
 
@@
 where $\hat{\bm{x}}_{f,j}$ denotes the vector of 1D nodal positions on the $j$th face of the reference quadrilateral $[-1,1]^2$.
-Recall that the matrix $\bm{E}$ as defined in (\ref{eq:EBdefs}) interpolates from volume collocation points $\hat{\bm{x}}$ to surface points $\hat{\bm{x}}_f$; the ordering (\ref{eq:facenodeordering}) simply corresponds to a permutation of the interpolation matrix $\bm{E}$.
-
-We now introduce a second set of $N_m$ mortar points $\hat{\bm{x}}_m$ on a face of the reference element.
+Recall that the matrix $\bm{E}$ as defined in (\ref{eq:EBdefs}) interpolates from volume collocation points $\hat{\bm{x}}$ to a specific ordering of the surface points $\hat{\bm{x}}_f$.
+The ordering (\ref{eq:facenodeordering}) simply corresponds to a permutation of the interpolation matrix $\bm{E}$.
+
+We now introduce a second set of mortar points $\hat{\bm{x}}_m$ on a face of the reference element.
 We assume these points also correspond to quadrature nodes with corresponding weights $\bm{w}_m$.
@@
 \[
-\bm{E}_{mf} = \ell_j(\hat{x}_{m,i}), \qquad 1\leq i \leq N_m, \quad 1\leq j \leq (N+1).
+\bm{E}_{mf} = \ell_j(\hat{x}_{m,i}), \qquad 1\leq i \leq \text{num.\ mortar points}, \quad 1\leq j \leq (N+1).
 \]
@@
 where $\hat{\bm{n}}_{i,f}, \hat{\bm{n}}_{i,m}$ are vectors containing components of the scaled reference normals at face and mortar points, respectively.
+
+%\begin{remark}
+%One could alternatively define $\bm{E}_{fm}$ via
+%\[
+%\bm{E}_{fm} = \LRp{\bm{E}_{mf}^T\bm{M}_m\bm{E}_{mf}}^{-1}\bm{E}_{mf}^T\bm{M}_m.
+%\]
+%However, it will turn out that this approach does not 
+%\end{remark}
 
@@
 \begin{proof}
-The first property holds if 
+The SBP property holds if 
 \begin{equation}
@@
 \end{align*}
-where we have used that $\bm{M}_m$ is diagonal.
-Then, since the scaled outward normals $\hat{n}_i$  are constant over each face of the reference element and $\bm{E}_{mf}, \bm{M}_m$ are block matrices (with each block corresponding to a face), we have that
+%where we have used that $\bm{M}_m$ is diagonal.  
+Then, since the scaled outward normals $\hat{n}_i\hat{J}_f$  are constant over each face of the reference element and $\bm{E}_{mf}, \bm{M}_m$ are block diagonal matrices (with each block corresponding to a face), $\diag{\hat{\bm{n}}_{i,f}}$ commutes and
 \[
@@
 
-\subsection{Mortar-based formulations on mapped elements}
+\subsection{Mortar-based entropy conservative formulations on mapped elements}
 
@@
 \begin{enumerate}
-\item Approach 1: the construction of $G_{ij}$ from \cite{kopriva2006metric, chan2018efficient}, which yields 
+\item Approach 1: the construction of $G_{ij}$ from \cite{kopriva2006metric}, which yields 
 \begin{align*}
@@
 \end{align*}
-\item Approach 2: the construction of $G_{ij}$ from \cite{kozdon2018energy} (Appendix C.3), which yields
+\item Approach 2: the construction of $G_{ij}$ from \cite{kozdon2018energy} (Appendix C.3, see also Footnote 3 in \cite{chan2019skew}), which yields
 \begin{align*}
@@
 \end{enumerate}
-Since $N_{\rm geo} \leq N_m+1$, by Lemma~\ref{lemma:Efm} and assumptions on $N_{\rm geo}, N_f, N_m$, $\bm{E}_{fm}G_{ij}(\hat{\bm{x}}_m) = \bm{E}G_{ij}(\hat{\bm{x}}$ and (\ref{eq:bcor}) vanishes.
+Since $N_{\rm geo} \leq N_m+1$, by Lemma~\ref{lemma:Efm} and the assumption that $N_{\rm geo} \leq \min(N, N_f+1, N_m+1)$, $\bm{E}_{fm}G_{ij}(\hat{\bm{x}}_m) = \bm{E}G_{ij}(\hat{\bm{x}})$ and (\ref{eq:bcor}) vanishes.
 
@@
 \tilde{\bm{u}}_m
-\end{bmatrix}, \qquad \tilde{\bm{u}}_m = \bm{u}\LRp{\bm{E}_{mf}\bm{E}\bm{v}(\bm{u}_h)}, \qquad \bm{f}^* = \bm{f}_S\LRp{\tilde{\bm{u}}_m^+,\tilde{\bm{u}}_m},
+\end{bmatrix}, \qquad \tilde{\bm{u}}_m = \bm{u}\LRp{\bm{E}_{mf}\bm{E}\bm{v}(\bm{u}_h)}, \qquad \bm{f}^*_i = \bm{f}_{i,S}\LRp{\tilde{\bm{u}}_m^+,\tilde{\bm{u}}_m},
 \end{gather}
@@
 
+\note{Add proof - include multi-element case}.
 The proof of entropy conservation is algebraically identical to previous proofs, with Lemma~\ref{lemma:Qmprops_3d} used in place of Lemma~\ref{lemma:Qhprops}.
 
+\subsection{Theoretical comparisons of Lobatto and Gauss nodes}
+
+High order entropy stable collocation schemes on tensor product elements most commonly utilize tensor product quadrature rules constructed from $(N+1)$-point Lobatto nodes \cite{carpenter2014entropy, gassner2016split} or $(N+1)$-point Gauss nodes \cite{chan2018efficient}.
+On non-conforming meshes, the most natural choice of mortar nodes for such collocation schemes is a composite Lobatto or Gauss rule.
+
+Recall that an $(N+1)$-point Lobatto rule is exact for polynomials of degree $(2N-1)$, while an $(N+1)$-point Gauss rule is exact for polynomials of degree $(2N+1)$.
+On non-conforming quadrilateral meshes, Theorem~\ref{lemma:Qmprops_3d} implies that both Lobatto and Gauss collocation schemes are stable for $N_{\rm geo}\leq N$, i.e., for isoparametric curved mappings.
+On non-conforming hexahedral meshes, Theorem~\ref{lemma:Qmprops_3d} implies that Gauss nodes are also stable for $N_{\rm geo}\leq N$, but that Lobatto nodes are stable only for $N_{\rm geo} \leq (N-1)$ if Approach 1 \cite{kopriva2006metric} is used to compute geometric terms.
+
+We note that the restriction on the degree of the geometric mapping $N_{\rm geo} \leq (N-1)$ for Lobatto nodes is not necessary on conforming hexahedral meshes.
+However, the introduction of non-aligned mortar nodes at non-conforming curved interfaces induces this additional restriction.
+Moreover, this restriction is consistent with Theorem 3 in \cite{fernandez2019entropy}.
+One can remove this constraint on the degree of the geometric mapping by using Approach 2 \cite{kozdon2018energy} to compute geometric terms.
+If Approach 2 is used, then Lobatto collocation schemes remain entropy conservative (stable) on curved hexahedral meshes for all $N_{\rm geo} \leq N$.
+
+Appendix~\ref{app:A} describes Approach 1 \cite{kopriva2006metric} and Approach 2 \cite{kozdon2018energy} in more detail, and demonstrates numerically that both approaches achieve similar high order accuracy in approximating geometric terms.
+
+%\begin{remark}
+%These results are , which states the 
+%\end{remark}
 %\note{Note: line DG approach can work, but it's flipped: we increase quadrature strength in the direction \emph{orthogonal} to derivative instead.
@@
 While the formulation (\ref{eq:esdgm}) is convenient for analysis, it is less convenient to implement.
-However, using properties of the mortar matrices, we can show that the mortar-based formulation (\ref{eq:esdgm}) is equivalent to a conforming formulation (\ref{eq:esdg}) with a modified numerical flux involving face-local correction terms.
-
-We can reformulate the remaining mortar contributions within (\ref{eq:mortar}).  We decompose $\bm{F}_S$ into interactions between volume nodes, surface nodes, and mortar nodes
-\[
-\bm{F}_S = \begin{bmatrix}
-\bm{F}_S^{vv} & \bm{F}_S^{vf} & \bm{F}_S^{vm}\\
-\bm{F}_S^{fv} & \bm{F}_S^{ff} & \bm{F}_S^{fm}\\
-\bm{F}_S^{mv} & \bm{F}_S^{mf} & \bm{F}_S^{mm}
+However, using properties of discretization matrices, we can show that the mortar-based formulation (\ref{eq:esdgm}) is equivalent to a conforming formulation (\ref{eq:esdg}) with a modified numerical flux involving face-local correction terms.
+
+First, we reformulate volume terms by incorporating geometric terms into the definition of the flux matrices.  Note that $\bm{Q}_{i,m}$ can alternatively be defined as \cite{gassner2016split}
+\[
+\bm{Q}_{i,m} = \sum_{j=1}^d \hat{\bm{Q}}_{j,m} \circ \avg{\bm{G}_{ij}}, \qquad \avg{\bm{G}_{ij}}_{kl} = \frac{1}{2}\LRp{ \LRp{\bm{G}_{ij}}_k + \LRp{\bm{G}_{ij}}_l},
+\]
+where $k,l$ are indices over the total number of volume, surface, and mortar points, and $\bm{G}_{ij}$ is the vector containing values of $G_{ij}$ at volume, surface, and mortar points.  
+The volume term of (\ref{eq:esdgm}) can then be rewritten as
+\begin{gather}
+\begin{bmatrix} \bm{I} \\ \bm{E} \\ \bm{E}_m\bm{E} \end{bmatrix}^T\sum_{i=1}^d \LRp{2{\bm{Q}}_{i,m} \circ \bm{F}_i}\bm{1} = 
+\begin{bmatrix} \bm{I} \\ \bm{E} \\ \bm{E}_m\bm{E} \end{bmatrix}^T\sum_{i=1}^d \LRp{2\hat{\bm{Q}}_{i,m} \circ \hat{\bm{F}}_i}\bm{1} \label{eq:Qmreform}\\
+\hat{\bm{F}}_i = \sum_{j=1}^d \avg{\bm{G}_{ij}}\circ \bm{F}_i.\nonumber
+\end{gather}
+We next decompose $\hat{\bm{F}}_i$ into interactions between volume nodes, surface nodes, and mortar nodes
+\[
+\hat{\bm{F}}_i = \begin{bmatrix}
+\hat{\bm{F}}_i^{vv} & \hat{\bm{F}}_i^{vf} & \hat{\bm{F}}_i^{vm}\\
+\hat{\bm{F}}_i^{fv} & \hat{\bm{F}}_i^{ff} & \hat{\bm{F}}_i^{fm}\\
+\hat{\bm{F}}_i^{mv} & \hat{\bm{F}}_i^{mf} & \hat{\bm{F}}_i^{mm}
 \end{bmatrix}
 \]
-Then, the on-element contributions to (\ref{eq:mortar}) can be expanded as follows
-\begin{align*}
-&\begin{bmatrix} \bm{I} \\ \bm{E} \\ \bm{E}_m \end{bmatrix}^T
+By expanding out the definition of the mortar-based SBP operator $\hat{\bm{Q}}_{i,m}$ and using properties of the matrices $\hat{\bm{B}}_{i,m}, \hat{\bm{F}}_i$, (\ref{eq:Qmreform}) can be rewritten as
+\begin{align*}
+\begin{bmatrix} \bm{I} \\ \bm{E} \\ \bm{E}_{mf}\bm{E} \end{bmatrix}^T
+&\LRp{\begin{bmatrix}
+\hat{\bm{Q}}_i-\hat{\bm{Q}}_i^T & \bm{E}^T\hat{\bm{B}}_i &\\
+-\hat{\bm{B}}_i\bm{E} &  & \hat{\bm{B}}_{i}{\bm{E}}_{fm} \\
+& -\hat{\bm{B}}_{i,m}{\bm{E}}_{mf} & 
+\end{bmatrix} \circ \hat{\bm{F}}_i}\bm{1} 
++ \bm{E}^T\bm{E}_{mf}^T \LRp{\hat{\bm{B}}_{i,m} \circ \hat{\bm{F}}_i^{mm}}\bm{1} 
+\\
+=& 
+\begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T
 \LRp{\begin{bmatrix}
-\bm{Q}_i-\bm{Q}_i^T & \bm{E}^T\bm{B}_i &\\
--\bm{B}_i\bm{E} &  & \bm{B}_{i}{\bm{E}}_m \\
-& -\bm{B}_{i,m}{\bm{E}}_m & {\bm{B}}_{i,m}
-\end{bmatrix} \circ \bm{F}_S}\bm{1} \\
-&= 
-\begin{bmatrix} \bm{I}_q \\ \bm{E} \end{bmatrix}^T
-\LRp{\begin{bmatrix}
-\bm{Q}_i-\bm{Q}_i^T & \bm{E}^T\bm{B}_i\\
--\bm{B}_i\bm{E} & \\
-\end{bmatrix} \circ \bm{F}_S}\bm{1} \\
-&+ \bm{E}^T \LRp{\bm{B}_i\tilde{\bm{E}}_m\circ \bm{F}_S^{sm}}\bm{1} - \bm{V}_m^T \LRp{ \tilde{\bm{B}}_i\bm{E}_m \circ \bm{F}_S^{ms}}\bm{1}
-\end{align*}
-Since multiplication by diagonal matrices $\bm{B}_i, \tilde{\bm{B}}_i$ is associative under the Hadamard product, the latter terms can be rewritten as 
-\begin{align*}
- \bm{E}^T \LRp{\bm{B}_i\tilde{\bm{E}}_m\circ \bm{F}_S^{sm}}\bm{1} &=  \bm{E}^T \bm{B}_i \LRp{\tilde{\bm{E}}_m\circ \bm{F}_S^{sm}}\bm{1}\\
-\bm{V}_m^T \LRp{ \tilde{\bm{B}}_i\bm{E}_m \circ \bm{F}_S^{ms}}\bm{1} &= \bm{V}_m^T \tilde{\bm{B}}_i \LRp{ \bm{E}_m \circ \bm{F}_S^{ms}}\bm{1} = \bm{E}^T \bm{B}_i \tilde{\bm{E}}_m \LRp{ \bm{E}_m \circ \bm{F}_S^{ms}}\bm{1}
-\end{align*}
-Then, (\ref{eq:mortar}) can be rewritten as 
-\begin{align}
-\bm{M}\td{\bm{u}_N}{t} &+ 
-\begin{bmatrix} \bm{V}_q \\ \bm{E} \end{bmatrix}^T
-\LRp{\begin{bmatrix}
-\bm{Q}_i-\bm{Q}_i^T & \bm{E}^T\bm{B}_i\\
--\bm{B}_i\bm{E} & \\
-\end{bmatrix} \circ \bm{F}_S}\bm{1} + \bm{E}^T\bm{B}_i \tilde{\bm{f}}^*_i = 0 \label{eq:mortarform}\\
-\tilde{\bm{f}}^*_i &= \tilde{\bm{E}}_m\bm{f}^*_i + \LRp{\tilde{\bm{E}}_m\circ \bm{F}_S^{sm}}\bm{1} - \tilde{\bm{E}}_m \LRp{ \bm{E}_m \circ \bm{F}_S^{ms}}\bm{1} 
-\nonumber
-\end{align}
-If the surface and mortar nodes are identical, then 
-\[
-\bm{E}_m = \tilde{\bm{E}}_m = \bm{I}, \qquad \LRp{\tilde{\bm{E}}_m \circ \bm{F}_S^{sm}}\bm{1} = \LRp{ \bm{E}_m \circ \bm{F}_S^{ms}}\bm{1}
-\]
-and the correction terms within $\tilde{\bm{f}}^*$ cancel out, and we recover the original scheme from \cite{chan2017discretely}.
-
-Note that in (\ref{eq:mortar2}), the only interactions between surface and mortar nodes occurs within the computation of $\tilde{\bm{f}}^*_i$, which can be done solely on mortar interfaces.
-The computational structure of
-\begin{enumerate}
-\item Compute volume contributions to the RHS
-\item Communicate surface contributions to mortar interfaces and compute $\tilde{\bm{f}}^*_i$.
-\item Communicate $\tilde{\bm{f}}^*_i$ back to each element.  
-\end{enumerate}
+\hat{\bm{Q}}_i-\hat{\bm{Q}}_i^T  & \bm{E}^T\hat{\bm{B}}_i\\
+-\hat{\bm{B}}_i\bm{E} & \hat{\bm{B}}_i\\
+\end{bmatrix} \circ \hat{\bm{F}}_i}\bm{1} \\
+&+ \bm{E}^T\LRp{\bm{E}_{mf}^T \LRp{\hat{\bm{B}}_{i,m} \circ \hat{\bm{F}}_i^{mm}}\bm{1} - \LRp{\bm{B}_i\circ \hat{\bm{F}}^{ff}_i}\bm{1} } \\
+&+ \bm{E}^T \LRp{\LRp{\hat{\bm{B}}_i\bm{E}_{fm}}\circ \hat{\bm{F}}_i^{fm}}\bm{1} - \bm{E}_{mf}^T \LRp{ \LRp{\hat{\bm{B}}_i\bm{E}_{mf}} \circ \hat{\bm{F}}_i^{mf}}\bm{1}.
+\end{align*}
+The above expressions can be simplified.  First, evaluating $\hat{\bm{F}}_i^{mm},\hat{\bm{F}}_i^{ff}$ yields
+\[
+\hat{\bm{F}}_i^{mm} = \diag{\sum_{j=1}^d G_{ij}(\hat{\bm{x}}_m) \circ \bm{f}_i(\tilde{\bm{u}}_m)}, 
+\quad 
+\hat{\bm{F}}_i^{ff} = \diag{\sum_{j=1}^d G_{ij}(\hat{\bm{x}}_f) \circ \bm{f}_i(\tilde{\bm{u}}_f)},
+\]
+Moreover, recall that the physical normals are defined as products of $G_{ij}$ with reference normals (\ref{eq:nJ_Gnhat}).
+Since $\hat{\bm{B}}_{i,m}, \hat{\bm{B}}_{i}$ contain scaled normals on the reference element on the diagonal, we can rewrite the correction terms using the physical normal boundary matrices $\bm{B}_{i,m}, \bm{B}_i$
+\begin{align*}
+&\bm{E}^T\LRp{\bm{E}_{mf}^T \LRp{\hat{\bm{B}}_{i,m} \circ \hat{\bm{F}}_i^{mm}}\bm{1} - \LRp{\bm{B}_i\circ \hat{\bm{F}}^{ff}_i}\bm{1} } \\
+&= \bm{E}^T\LRp{\bm{E}_{mf}^T \bm{B}_{i,m}\bm{f}_i(\tilde{\bm{u}}_m) - \bm{B}_{i}\bm{f}_i(\tilde{\bm{u}}_f) }.
+\end{align*}
+The remaining correction terms can also be simplified by noting that $\bm{E}_{fm}$ are block diagonal matrices, with each block corresponding to a face, and that the scaled reference normals (i.e., the diagonal entries of $\hat{\bm{B}}_{i,m}, \hat{\bm{B}}_{i}$) are constant over each face.
+%Define the physical flux matrices at surface-mortar nodes
+%\[
+%\bm{F}_i^{fm}, \qquad \bm{F}_i^{mf}
+%\]
+Then, using (\ref{eq:nJ_Gnhat}), we can rewrite the final set of correction terms as
+\begin{align*}
+&\bm{E}^T \LRp{\LRp{\hat{\bm{B}}_i\bm{E}_{fm}}\circ \hat{\bm{F}}_i^{fm}}\bm{1} - \bm{E}_{mf}^T \LRp{ \LRp{\hat{\bm{B}}_{i,m}\bm{E}_{mf}} \circ \hat{\bm{F}}_i^{mf}}\bm{1}\\
+& = \bm{E}^T \LRp{{\bm{E}_{fm}}\circ \bm{F}_i^{fm} \circ \avg{\bm{n}_i}_{fm}}\bm{1} - \bm{E}_{mf}^T \LRp{ {\bm{E}_{mf}} \circ {\bm{F}}_i^{mf} \circ \avg{\bm{n}_i}_{mf}}\bm{1}
+\end{align*}
+where $\bm{F}_i^{fm} = \LRp{\bm{F}_i^{mf}}^T$ are the flux matrices whose entries are evaluations of $\bm{f}_{i,S}$ between solution values at mortar and surface nodes, and $\avg{\bm{n}_i}_{fm} = \avg{\bm{n}_i}_{mf}^T$ are matrices whose entries are arithmetic averages of $n_iJ_f$ (the $i$th component of the scaled normal vector) between each mortar and surface point.
+
+Adding these correction terms back into the formulation (\ref{eq:esdgm}) yields 
+\begin{gather}
+{\bm{M}}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T
+\sum_{i=1}^d \LRp{2{\bm{Q}}_{i,h} \circ \bm{F}_i}\bm{1} + \bm{E}^T\LRp{\bm{f}_i^*- \bm{B}_i\bm{f}_i(\tilde{\bm{u}}_f)} = 0 \label{eq:esdgm_mortar}\\
+\bm{f}_i^* = \bm{E}_{mf}^T\LRp{{\bm{B}}_{i,m}\bm{f}_{i,m} + \LRp{{\bm{E}_{fm}}\circ \bm{F}_i^{fm} \circ \avg{\bm{n}_i}_{fm} - {\bm{E}_{mf}} \circ {\bm{F}}_i^{mf} \circ \avg{\bm{n}_i}_{mf}}\bm{1}},\nonumber
+\end{gather}
+where $\bm{f}_{i,m} = \bm{f}_{i,S}(\tilde{\bm{u}}_m^+,\tilde{\bm{u}}_m)$ is the numerical flux computed using the  at mortar points.
+The formulation (\ref{eq:esdgm_mortar}) is nearly equivalent to the hybridized mortar formulation, with the main difference being the presence of skew-symmetric correction terms in the definition of the numerical flux.
+Note that in this formulation, the only interactions between surface and mortar nodes occur within the computation of $\bm{f}^*_i$, and involve only computations which are local to mortar interfaces.
+\begin{remark}
+If the surface and mortar nodes are the same, then 
+\[
+\bm{E}_{mf} = \bm{E}_{fm} = \bm{I}, \qquad \avg{\bm{n}_i}_{fm} = \avg{\bm{n}_i}_{mf}
+\]
+and the correction terms in the definition of $\bm{f}^*_i$ cancel out.
+The formulation (\ref{eq:esdgm_mortar}) then reverts to a Gauss collocation scheme \cite{chan2017discretely} on conforming meshes.
+\end{remark}
 
@@
 
-\subsection{Extensions of the mortar-based approach}
-
-\begin{figure}
-\centering
-\includegraphics[width=.5\textwidth]{noncon3D.png}
-\caption{Two-layer mortars for reduced $O(N^3)$ computational cost on non-conforming hexahedral interfaces.}
-\end{figure}
-
-\note{Extensions to 3D meshes: double anisotropic mortar for $O(N^3)$ operations vs $O(N^4)$.}
-
-\note{Extension to $p$ non-conforming meshes is straightforward.  Simply replace the definition of a mortar.}
-
-\note{Extension to general quadratures just requires replacing $\bm{I}$ with $\bm{V}_q, \bm{P}_q$, and redefining $\bm{E}$ matrices.}
-
-\note{Explain why non-hexahedral elements isn't as much of an issue - Shadpey \cite{shadpey2019energy} use this for simplicial elements.
-The interpolation operators are dense, and using the mortar-based approach wouldn't actually decrease the number of operations.}
-
 \section{Numerical experiments}
@@
 \centering
-\includegraphics[width=.475\textwidth]{figs/mesh_curved0.png}
-\hspace{.5em}
-\includegraphics[width=.475\textwidth]{figs/mesh_curved.png}
-\caption{Two $N=3$ meshes used in the refinement and entropy conservation studies.}
+%\includegraphics[width=.475\textwidth]{figs/mesh_curved0.png}
+%\hspace{.5em}
+\includegraphics[width=.55\textwidth]{figs/noncon_curved_mesh.png}
+\caption{An $N=3$ mesh used in the refinement and entropy conservation studies.}
 \end{figure}
@@
 
+\section{Conclusions and extensions of the mortar-based approach}
+
+\begin{figure}
+\centering
+\includegraphics[width=.5\textwidth]{noncon3D.png}
+\caption{Two-layer mortars for reduced $O(N^3)$ computational cost on non-conforming hexahedral interfaces.}
+\end{figure}
+
+\note{Extensions to 3D meshes: double anisotropic mortar for $O(N^3)$ operations vs $O(N^4)$.}
+
+\note{Extension to $p$ non-conforming meshes is straightforward.  Simply replace the definition of a mortar.}
+
+\note{Extension to general quadratures just requires replacing $\bm{I}$ with $\bm{V}_q, \bm{P}_q$, and redefining $\bm{E}$ matrices.}
+
+\note{Explain why non-hexahedral elements isn't as much of an issue - Shadpey \cite{shadpey2019energy} use this for simplicial elements.
+The interpolation operators are dense, and using the mortar-based approach wouldn't actually decrease the number of operations.}
+
+Jesse Chan gratefully acknowledges support from the National Science Foundation under awards DMS-1719818 and DMS-1712639.
+
 \appendix
 
-\section{Construction of geometric terms for curved hexahedral elements}
+\section{Geometric terms for curved hexahedral meshes}
 \label{app:A}
+
+Entropy conservation (stability) requires that the geometric terms be constructed appropriately.
+In addition to requiring satisfaction of the discrete geometric conservation law \cite{kopriva2006metric, crean2018entropy, chan2018discretely}, the schemes derived in this work require that the polynomial degree of the geometric terms be related to the accuracy of the surface and mortar quadrature rules.
+We consider two methods to construct geometric terms: Approach 1 from \cite{kopriva2006metric}, and Approach 2 from \cite{kozdon2018energy}.
+
+Entropy conservation (or stability) on non-conforming meshes also require that the scaled outward normals are equal and opposite across each mortar node.
+This holds if the mesh is \emph{watertight} or \emph{well-constructed}, such that there are no gaps between neighboring elements.
+Several conditions which guarantee that a curved non-conforming mesh is well-constructed are described in \cite{kopriva2019free}.
+For the two-dimensional numerical experiments in this work, we construct watertight curved non-conforming meshes by first curving a conforming mesh.
+We then locally refine certain parent elements and transfer the geometric terms of the parent element to the child elements.
+For the three-dimensional numerical experiments in this work, \note{finish}.
+
+\subsection{Construction of geometric terms for curved hexahedra}
+
+Let $\bm{D}_{\rm 1D}$ denote the $(N_{\rm geo}+1) \times (N_{\rm geo}+1)$ one-dimensional nodal differentiation matrix at degree $N_{\rm geo}$ Lobatto nodes.
+Multi-dimensional differentiation matrices on the reference hexahedron can then be constructed via
+\begin{align*}
+\bm{D}_1 &= \bm{D}_{\rm 1D} \otimes \bm{I}_{\rm 1D}\otimes \bm{I}_{\rm 1D}\\
+\bm{D}_2 &= \bm{I}_{\rm 1D} \otimes \bm{D}_{\rm 1D}\otimes \bm{I}_{\rm 1D}\\
+\bm{D}_3 &= \bm{I}_{\rm 1D} \otimes \bm{I}_{\rm 1D}\otimes \bm{D}_{\rm 1D},
+\end{align*}
+where $\bm{I}_{\rm 1D}$ is the $(N_{\rm geo}+1) \times (N_{\rm geo}+1)$ identity matrix.
+
+Let $\bm{x},\bm{y},\bm{z}$ denote the vectors of physical nodal coordinates of a curved degree $N_{\rm geo}$ element.
+Approach 1 in \cite{kopriva2006metric} constructs scaled geometric terms $G_{ij} = J\pd{\hat{x}_j}{x_i}$ via
+\begin{align*}
+\bm{G}_{i1} &= \alpha_i\LRp{\bm{D}_3\bm{f}_{i2} - \bm{D}_2\bm{f}_{i3}},\\
+\bm{G}_{i2} &=\alpha_i \LRp{\bm{D}_1\bm{f}_{i3} - \bm{D}_3\bm{f}_{i1}},\\
+\bm{G}_{i3} &= \alpha_i\LRp{\bm{D}_2\bm{f}_{i1} - \bm{D}_1\bm{f}_{i2}}
+\end{align*}
+where $\alpha_1 = 1$ and $\alpha_2 = \alpha_3 = -1$, and the geometric ``potentials'' $\bm{f}_{ij}$ are defined as 
+\begin{align}
+\bm{f}_{1j} = \LRp{ \bm{D}_j \bm{y}} \circ \bm{z}, \qquad 
+\bm{f}_{2j} = \LRp{ \bm{D}_j \bm{x}} \circ \bm{z}, \qquad
+\bm{f}_{3j} = \LRp{ \bm{D}_j \bm{y}} \circ \bm{x}.  \label{eq:geopot}
+\end{align}
+The resulting geometric terms are tensor product polynomials of degree $G_{ij} \in Q^{N_{\rm geo}}$.
+If $N_{\rm geo} < N$, we use polynomial interpolation to compute values of the geometric terms at degree $N$ Lobatto nodes.
+
+Approach 2 in \cite{kozdon2018energy} (see also \cite{chan2019skew}, Footnote 3) modifies this approach slightly by reducing the degree of each geometric potentials along a coordinate direction.
+Let $\bm{I}_{N_{\rm geo}}^{N_{\rm geo}-1}$ denote the interpolation operator from degree $N_{\rm geo}$ to degree $N_{\rm geo}-1$ Lobatto nodes, and let $\bm{I}_{N_{\rm geo}-1}^{N_{\rm geo}}$ denote the interpolation operator from degree $N_{\rm geo}-1$ to degree $N_{\rm geo}$ Lobatto nodes.
+Define the degree reduction operator $\bm{F}_{\rm 1D}$ as the product of these interpolation operators
+\[
+\bm{F}_{\rm 1D} = \bm{I}_{N_{\rm geo}-1}^{N_{\rm geo}}\bm{I}_{N_{\rm geo}}^{N_{\rm geo}-1}.
+\]
+Multiplication by $\bm{F}_{\rm 1D}$ lowers the degree of a polynomial from $N_{\rm geo}$ to $N_{\rm geo}-1$ while leaving the boundary values unchanged.
+We can define multi-dimensional degree reduction operators
+\begin{align*}
+\bm{F}_1 &= \bm{F}_{\rm 1D} \otimes \bm{I}_{\rm 1D}\otimes \bm{I}_{\rm 1D}\\
+\bm{F}_2 &= \bm{I}_{\rm 1D} \otimes \bm{F}_{\rm 1D}\otimes \bm{I}_{\rm 1D}\\
+\bm{F}_3 &= \bm{I}_{\rm 1D} \otimes \bm{I}_{\rm 1D}\otimes \bm{F}_{\rm 1D}.
+\end{align*}
+Suppose $u \in Q^{N_{\rm geo}}$ is represented using nodal values $\bm{u}$; then, $\bm{F}_1\bm{u}$ corresponds to a polynomial in $Q^{N_{\rm geo}-1,N_{\rm geo},N_{\rm geo}}$.
+In other words, $\bm{F}_i$ lowers the polynomial degree in the $i$th coordinate by $1$.
+Approach 2 computes the geometric terms $\bm{G}_{ij}$ via
+\begin{align*}
+\bm{G}_{i1} &= \alpha_i\LRp{\bm{D}_3\tilde{\bm{f}}_{i2} - \bm{D}_2\tilde{\bm{f}}_{i3}},\\
+\bm{G}_{i2} &=\alpha_i \LRp{\bm{D}_1\tilde{\bm{f}}_{i3} - \bm{D}_3\tilde{\bm{f}}_{i1}},\\
+\bm{G}_{i3} &= \alpha_i\LRp{\bm{D}_2\tilde{\bm{f}}_{i1} - \bm{D}_1\tilde{\bm{f}}_{i2}}
+\end{align*}
+where $\tilde{\bm{f}}_{ij} = \bm{F}_j \bm{f}_{ij}$ are degree-reduced geometric potentials, and the original geometric potentials $\bm{f}_{ij}$ are computed via (\ref{eq:geopot}).
+Since multiplication by $\bm{D}_i$ reduces the degree by $1$ in the $i$th coordinate by $1$, the resulting geometric terms $G_{ij}$ satisfy
+\begin{align*}
+&G_{i1} \in Q^{N_{\rm geo},N_{\rm geo}-1,N_{\rm geo}-1}\\
+&G_{i2} \in Q^{N_{\rm geo}-1,N_{\rm geo},N_{\rm geo}-1}\\
+&G_{i3} \in Q^{N_{\rm geo}-1,N_{\rm geo}-1,N_{\rm geo}}, \qquad i = 1,2,3.
+\end{align*}
+
+While Approach 1 is nominally more accurate than Approach 2, both approaches achieve nearly identical accuracy and convergence rates.
+We analyze $L^2$ errors using Approach 1 and Approach 2 by  computing exact values of $G_{ij}$ using the cross product formula \cite{kopriva2006metric, hesthaven2007nodal}.
+We construct a curved mesh by transforming a Cartesian hexahedral grid on $[-1,1]^3$ by interpolating the global curved mapping at Lobatto nodes
+\begin{align*}
+    \tilde{x} &= x + \frac{1}{4}\cos(x)\sin(y)\sin(z)\\
+    \tilde{y} &= y + \frac{1}{4}\sin(x)\cos(y)\sin(z)\\
+    \tilde{z} &= z + \frac{1}{4}\sin(x)\sin(y)\cos(z),
+\end{align*}
+where $x,y,z$ denote coordinates on the Cartesian grid and $\tilde{x},\tilde{y},\tilde{z}$ denote coordinates on the mapped curved domain.
+Figure~\ref{fig:geo_err} shows the convergence of the $L^2$ error over all geometric terms under mesh refinement.
+Table~\ref{fig:geo_rates} shows computed convergence rates, which match the $O(h^{N+2})$ rates proven in \cite{chan2018discretely, crean2018entropy} for simplicial elements.
+\begin{figure}
+\centering
+%\subfloat[Coarse $N=3$ non-conforming affine mesh]{\raisebox{2.6em}{\includegraphics[width=.475\textwidth]{figs/mesh_affine.png}}}
+\begin{tikzpicture}
+\begin{loglogaxis}[
+    width=.6\textwidth,
+    xlabel={Mesh size $h$},
+    ylabel={$L^2$ error}, 
+    xmax=1.25,xmin=.0625,
+    ymin=2.5e-8, ymax=2.5,
+    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
+    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
+    legend entries={Approach 1 \cite{kopriva2006metric}, Approach 2 \cite{kozdon2018energy}}%, Lobatto-Gauss}    
+]
+\pgfplotsset{
+cycle list={{blue, mark=*}, {red, dashed ,mark=square*}}%,{black, dashdotted ,mark=triangle*}}
+}
+% N = 1
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.576948)(0.5,0.147753)(0.25,0.0217505)(0.125,0.00284466)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.812383)(0.5,0.207677)(0.25,0.0306257)(0.125,0.00401036)};
+
+% N = 2
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,1.14476)(0.5,0.0680173)(0.25,0.00414906)(0.125,0.000257077)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.885454)(0.5,0.0505198)(0.25,0.0029414)(0.125,0.000177322)};
+
+% N = 3
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.178578)(0.5,0.00711309)(0.25,0.000232189)(0.125,7.33457e-06)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.163661)(0.5,0.00592117)(0.25,0.000187325)(0.125,5.866e-06)};
+
+% N = 4
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.0419636)(0.5,0.000598623)(0.25,8.70891e-06)(0.125,1.31374e-07)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.0380338)(0.5,0.000530832)(0.25,7.64254e-06)(0.125,1.14732e-07)};
+
+\node at (axis cs:.09,3.5e-3) {$N = 1$};
+\node at (axis cs:.09,2.2e-4) {$N = 2$};
+\node at (axis cs:.09,6e-6) {$N = 3$};
+\node at (axis cs:.09,1.25e-7) {$N = 4$};
+\end{loglogaxis}
+\end{tikzpicture}
+\caption{Blah}
+\label{fig:geo_err}
+\end{figure}
+
+\begin{table}
+\centering
+\begin{tabular}{|c||c|c|c|c|}
+\hline
+$N$ & 1 & 2 & 3 & 4\\
+\hline
+Approach 1 &2.8472  & 4.0772 &4.9896 & 6.0879 \\
+\hline
+Approach 2 & 2.8494  & 4.0238 & 4.9608 & 6.0769  \\
+\hline
+\end{tabular}
+\caption{Computed $L^2$ error convergence rates for Approach 1 \cite{kopriva2006metric} and Approach 2 \cite{kozdon2018energy}.
+Both approaches achieve $O(h^{N+2})$ rates, which were proven in \cite{chan2018discretely, crean2018entropy} for simplicial elements.}
+\label{fig:geo_rates}
+\end{table}
 

### 2020-01-04 docs/ESDG_mortar/figs/compute_op_counts.m

@@ -1,28 +1,0 @@
-N = 1:7;
-
-tp_GLL = 3*(N+1).^4;
-tp_GQ = tp_GLL + 12*(N+1).^3;
-tp_stag = 3*(N+2).^4;
-
-% plot(tp_GLL,'o--')
-% hold on
-% plot(tp_GQ,'s--')
-% plot(tp_stag,'x--')
-
-op_GLL = 3*(N+1).^4;
-op_GQ = op_GLL + 12*(N+1).^3;
-op_stag = 3*(N+2).^4 + 6*(N+1).^3.*(N+2);
-
-plot(op_GLL,'o--')
-hold on
-plot(op_GQ,'s--')
-plot(op_stag,'x--')
-
-% fprintf('\n');
-% print_pgf_coordinates(tp_GLL)
-% print_pgf_coordinates(tp_GQ)
-% print_pgf_coordinates(tp_stag)
-fprintf('\n');
-print_pgf_coordinates(op_GLL)
-print_pgf_coordinates(op_stag)
-print_pgf_coordinates(op_GQ)

### 2020-02-27 docs/ESDG_ROM/main.tex
deleting a bunch of stuff and staging old changes

@@
 Suppose there are $k = 1,\ldots, K$ subdomains.
-Then, this sort of multi-domain model is expected to reduce costs if the sum of the squares of the number of subdomain sampling points $\sum_k (N_s^k)^2$ is significantly smaller than the global number of sampling points $N_s^2$.
+Then, a multi-domain reduced order model is expected to decrease costs if the sum of the squares of the number of subdomain sampling points $\sum_k (N_s^k)^2$ is significantly smaller than the global number of sampling points $N_s^2$.
 Future work will also investigate implicit time-stepping for entropy stable reduced models, as additional costs associated with the spatial formulation may be offset by the reduction in cost for solving smaller matrix systems during the linearization process.

### 2020-02-27 docs/ESDG_mortar/main.tex
deleting a bunch of stuff and staging old changes

@@
 
+\usepackage{stmaryrd} % for llbracket
 \usepackage{amsmath, amsfonts, amssymb}
-\usepackage{stmaryrd} % for llbracket
 
@@
 However, on non-conforming meshes, volume and surface nodes are no longer aligned, resulting in a larger number of flux evaluations.
-We reduce this expense by introducing an entropy stable mortar-based treatment of non-conforming interfaces via a face-local correction term, and provide truncation error estimates for the resulting discretizations.
-Numerical experiments confirm the stability and accuracy of this approach, and we conclude by discussing how to extend mortar treatments of non-conforming interfaces to general elements and arbitrary quadrature rules \cite{chan2017discretely}.
+We reduce this expense by introducing an entropy stable mortar-based treatment of non-conforming interfaces via a face-local correction term, and provide necessary conditions for high order accuracy.
+Numerical experiments confirm the stability and accuracy of this approach.
 \end{abstract}
@@
 Section~\ref{sec:2} describes a naive extension to non-conforming meshes and illustrates why this formulation results in an increase in computational costs.
-Section~\ref{sec:3} introduces a mortar-based formulation which addresses such costs, and characterizes the accuracy of the resulting formulations.
-Section~\ref{sec:4} describes the extension to curved meshes, as well as more general elements and $p$ non-conformity.
-We conclude with numerical validation of theoretical results in Section~\ref{sec:5}.
+Section~\ref{sec:3} introduces a mortar-based formulation on curved meshes which addresses such costs, and characterizes the accuracy of the resulting formulations.
+We conclude with numerical validation of theoretical results in Section~\ref{sec:4}.
 
@@
 However, for nonlinear conservation laws, the presence of the dense $\bm{E}$ matrix in the GSBP property (\ref{eq:gsbp}) complicates the imposition of boundary conditions and computation of inter-element numerical fluxes \cite{crean2018entropy, chan2017discretely, chan2018efficient}.
-This can be avoided by using ``hybridized'' (or decoupled) SBP operators \cite{chan2017discretely, chenreview, chan2019skew}, which are defined as the block matrix
+This can be avoided by using ``hybridized'' (also referred to as decoupled) SBP operators \cite{chan2017discretely, chenreview, chan2019skew}, which are defined as the block matrix
 \[
@@
 \end{lemma}
-The proof is contained in \cite{chan2018efficient}.
-The SBP property follows from the construction of $\hat{\bm{Q}}_{i,h}$, and $\hat{\bm{Q}}_{i,h}\bm{1} = \bm{0}$ is a consequence of the Kronecker product and the same arguments used to derive (\ref{eq:Qh1}).
+The proof is found in \cite{chan2018efficient}.
+The SBP property follows from the construction of $\hat{\bm{Q}}_{i,h}$, and $\hat{\bm{Q}}_{i,h}\bm{1} = \bm{0}$ is a consequence of properties of the Kronecker product and the same arguments used to derive (\ref{eq:Qh1}).
 The construction of 3D differentiation and interpolation matrices proceeds similarly, and the matrices satisfy the same properties as described in Lemma~\ref{lemma:hsbprefprops}.
@@
 The discrete GCL is satisfied automatically for isoparametric mappings in 2D, and there exist several techniques to enforce the satisfaction of a discrete GCL on various three-dimensional domains \cite{thomas1979geometric, kopriva2006metric, crean2018entropy, chan2018discretely, kozdon2018energy, kopriva2019free}.
+We will discuss specific approaches for non-conforming meshes in Section~\ref{sec:mapped} and Appendix~\ref{app:A}.
 
@@
 \end{lemma}
-\begin{proof}
-The proof is identical to that of \cite{chan2018discretely}.  
-We review it here for completeness.  Expanding out $\bm{Q}_{i,h} + \bm{Q}_{i,h}^T$ yields
-\[
-\frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h} + \hat{\bm{Q}}_{j,h}^T\diag{\bm{G}_{ij}} + \hat{\bm{Q}}_{j,h} \diag{\bm{G}_{ij}} +  \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h}^T.
-\]
-For each term in the sum, we apply the SBP property of $\hat{\bm{Q}}_{j,h}$.  The volume terms cancel, leaving only surface terms
-\[
-\frac{1}{2}\sum_{j=1}^d \LRp{\hat{\bm{B}}_{j,h}\diag{\bm{G}_{ij}} +  \diag{\bm{G}_{ij}}\hat{\bm{B}}_{j,h}} = \begin{bmatrix}
-\bm{0} &\\
-& \bm{B}_i \end{bmatrix}.
-\]
-where we have used (\ref{eq:nJ_Gnhat}) and that $\hat{\bm{B}}_{j,h}$ and $\diag{\bm{G}_{ij}}$ are diagonal matrices.
-
-To show $\bm{Q}_{i,h}\bm{1} = \bm{0}$, we note that
-\[
-\bm{Q}_{i,h}\bm{1} = \frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h}\bm{1} + \hat{\bm{Q}}_{j,h} \diag{\bm{G}_{ij}}\bm{1} = \frac{1}{2}\sum_{j=1}^d \hat{\bm{Q}}_{j,h} \bm{G}_{ij}
-\]
-since $\hat{\bm{Q}}_{j,h}\bm{1} = \bm{0}$ by Lemma~\ref{lemma:hsbprefprops}.
+The proof is given in \cite{chan2018discretely}.  
+%\begin{proof}
+%The proof is identical to that of \cite{chan2018discretely}.  
+%We review it here for completeness.  Expanding out $\bm{Q}_{i,h} + \bm{Q}_{i,h}^T$ yields
+%\[
+%\frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h} + \hat{\bm{Q}}_{j,h}^T\diag{\bm{G}_{ij}} + \hat{\bm{Q}}_{j,h} \diag{\bm{G}_{ij}} +  \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h}^T.
+%\]
+%For each term in the sum, we apply the SBP property of $\hat{\bm{Q}}_{j,h}$.  The volume terms cancel, leaving only surface terms
+%\[
+%\frac{1}{2}\sum_{j=1}^d \LRp{\hat{\bm{B}}_{j,h}\diag{\bm{G}_{ij}} +  \diag{\bm{G}_{ij}}\hat{\bm{B}}_{j,h}} = \begin{bmatrix}
+%\bm{0} &\\
+%& \bm{B}_i \end{bmatrix}.
+%\]
+%where we have used (\ref{eq:nJ_Gnhat}) and that $\hat{\bm{B}}_{j,h}$ and $\diag{\bm{G}_{ij}}$ are diagonal matrices.  
+%
+%To show $\bm{Q}_{i,h}\bm{1} = \bm{0}$, we note that
+%\[
+%\bm{Q}_{i,h}\bm{1} = \frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,h}\bm{1} + \hat{\bm{Q}}_{j,h} \diag{\bm{G}_{ij}}\bm{1} = \frac{1}{2}\sum_{j=1}^d \hat{\bm{Q}}_{j,h} \bm{G}_{ij}
+%\]
+%since $\hat{\bm{Q}}_{j,h}\bm{1} = \bm{0}$ by Lemma~\ref{lemma:hsbprefprops}.
 Note that, since $\bm{G}_{ij}$ is constructed using polynomial interpolation, $\bm{G}_{ij} = \LRs{G_{ij}(\hat{\bm{x}}), \quad \bm{E}G_{ij}(\hat{\bm{x}})}^T$, where $G_{ij}(\hat{\bm{x}})$ denotes the values of $G_{ij}$ at volume points.
 The remaining term can be expanded out
-\begin{align*}
-\sum_{j=1}^d \hat{\bm{Q}}_{j,h} \bm{G}_{ij} &= \sum_{j=1}^d \begin{bmatrix} \hat{\bm{Q}}_jG_{ij}(\hat{\bm{x}}) - \hat{\bm{Q}}_j^TG_{ij}(\hat{\bm{x}}) + \bm{E}^T\bm{B}\bm{E}G_{ij}(\hat{\bm{x}})\\
-\bm{B}\bm{E}G_{ij}(\hat{\bm{x}}) - \bm{B}\bm{E}G_{ij}(\hat{\bm{x}})
-\end{bmatrix} \\
-&=
-\sum_{j=1}^d \begin{bmatrix} \LRp{- \hat{\bm{Q}}_j^T + \bm{E}^T\bm{B}\bm{E}}G_{ij}(\hat{\bm{x}})\\
-\bm{0}
-\end{bmatrix} = 
-\sum_{j=1}^d \begin{bmatrix} \hat{\bm{Q}}_jG_{ij}(\hat{\bm{x}})\\
-\bm{0}
-\end{bmatrix}  = \bm{0},
-\end{align*}
-where we have used  the generalized SBP property and that $\sum_{j=1}^d\hat{\bm{Q}}_jG_{ij}(\hat{\bm{x}}) = \bm{0}$, since $G_{ij}$ satisfies the discrete GCL (\ref{eq:dgcl}), $G_{ij}(\hat{\bm{x}})$ is a polynomial of degree less than or equal to $N$, and $\hat{\bm{Q}}_j$ exactly differentiates polynomials of degree $N$.
-\qed\end{proof}
+%\begin{align*}
+%\sum_{j=1}^d \hat{\bm{Q}}_{j,h} \bm{G}_{ij} &= \sum_{j=1}^d \begin{bmatrix} \hat{\bm{Q}}_jG_{ij}(\hat{\bm{x}}) - \hat{\bm{Q}}_j^TG_{ij}(\hat{\bm{x}}) + \bm{E}^T\bm{B}\bm{E}G_{ij}(\hat{\bm{x}})\\
+%\bm{B}\bm{E}G_{ij}(\hat{\bm{x}}) - \bm{B}\bm{E}G_{ij}(\hat{\bm{x}})
+%\end{bmatrix} \\
+%&=
+%\sum_{j=1}^d \begin{bmatrix} \LRp{- \hat{\bm{Q}}_j^T + \bm{E}^T\bm{B}\bm{E}}G_{ij}(\hat{\bm{x}})\\
+%\bm{0}
+%\end{bmatrix} = 
+%\sum_{j=1}^d \begin{bmatrix} \hat{\bm{Q}}_jG_{ij}(\hat{\bm{x}})\\
+%\bm{0}
+%\end{bmatrix}  = \bm{0},
+%\end{align*}
+%where we have used  the generalized SBP property and that $\sum_{j=1}^d\hat{\bm{Q}}_jG_{ij}(\hat{\bm{x}}) = \bm{0}$, since $G_{ij}$ satisfies the discrete GCL (\ref{eq:dgcl}), $G_{ij}(\hat{\bm{x}})$ is a polynomial of degree less than or equal to $N$, and $\hat{\bm{Q}}_j$ exactly differentiates polynomials of degree $N$.
+%\qed\end{proof}
 
 \subsection{Entropy conservative formulations on conforming meshes of mapped elements}
+\label{sec:mapped}
 
@@
 \sum_{i=1}^d \LRp{2{\bm{Q}}_{i,h} \circ \bm{F}_i}\bm{1} + \bm{E}^T{\bm{B}}_i\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_f)} = 0 \label{eq:esdg}\\
-\LRp{\bm{F}_i}_{jk} = \bm{f}_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}, \qquad \bm{f}_i^* = \bm{f}_{i,S}\LRp{\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f} \nonumber
+\LRp{\bm{F}_i}_{jk} = \bm{f}_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}, \qquad \bm{f}_i^* = \bm{f}_{i,S}\LRp{\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f}, \nonumber
 \end{gather}
+where $\bm{M} = \hat{\bm{M}} \circ \diag{\bm{J}}$ is the diagonal mass matrix scaled by the determinant of the geometric Jacobian $J$ evaluated at nodal points.  
 Here, the exterior state $\tilde{\bm{u}}_f^+$ used to evaluate the numerical flux $\bm{f}_i^*$ corresponds either to interface values on a neighboring element or an exterior state used to enforce boundary conditions, and $\circ$ denotes the matrix Hadamard product.
-
-%\note{Mention proof}
-%Then, an entropy conservative scheme on the reference interval is given as
-%\begin{gather}
-%\bm{M}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T
-%\LRp{2\bm{Q}_h \circ \bm{F}}\bm{1} + \bm{E}^T\bm{B}\LRp{\bm{f}^*-\bm{f}(\tilde{\bm{u}}_f)} = 0 \label{eq:esdg1D}\\ 
-%\tilde{\bm{u}} = \begin{bmatrix}
-%\bm{u}_h\\
-%\tilde{\bm{u}}_f
-%\end{bmatrix}, \qquad \tilde{\bm{u}}_f = \bm{u}\LRp{\bm{E}\bm{v}(\bm{u}_h)},\nonumber \\ 
-%\LRp{\bm{F}}_{jk} = \bm{f}_S\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}, \qquad \bm{f}^* = \bm{f}_S\LRp{\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f}.\nonumber
-%\end{gather}
-%Here, $\tilde{\bm{u}}_f^+$ denotes some exterior state which can be used to enforce boundary conditions or coupling between different elements
-
-\begin{theorem}
-\label{thm:esdg}
 Assuming continuity in time, the formulation (\ref{eq:esdg}) is entropy conservative in the sense that
@@
 \]
-\end{theorem}
-\begin{proof}
-The proof uses similar techniques as proofs in other papers \cite{chen2017entropy, crean2018entropy, chan2017discretely, chan2019skew}, and we repeat it for completeness.
+The proof uses similar techniques as proofs in other papers \cite{chen2017entropy, crean2018entropy, chan2017discretely, chan2019skew}
+%\begin{theorem}
+%\label{thm:esdg}
+%Assuming continuity in time, the formulation (\ref{eq:esdg}) is entropy conservative in the sense that
+%\[
+%\bm{1}^T\bm{M}\td{S(\bm{u}_h)}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}_i\LRp{\bm{v}_f^T\bm{f}_i^* - \psi_i(\tilde{\bm{u}}_f)} = 0, \qquad \bm{v}_f =\bm{E}\bm{v}(\bm{u}_h).
+%\]
+%\end{theorem}
+%\begin{proof}
+%The proof uses similar techniques as proofs in other papers \cite{chen2017entropy, crean2018entropy, chan2017discretely, chan2019skew}, and we repeat it for completeness.
 We begin by testing (\ref{eq:esdg}) with the vector of entropy variables evaluated at nodal points $\bm{v}(\bm{u}_h)$.
 We apply the chain rule in time and use that $\bm{M}$ is diagonal to show that
-\[
-\bm{v}(\bm{u}_h)^T\bm{M}\td{\bm{u}}{t} = \bm{1}^T\bm{M}\LRp{\bm{v}(\bm{u}_h)\circ \td{\bm{u}_h}{t}} =  \bm{1}^T\bm{M}\td{S(\bm{u}_h)}{t}.
-\]
-For the volume term, we use the SBP property to arrive at
-\begin{align*}
-&\bm{v}(\bm{u}_h)^T\begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T\LRp{2\bm{Q}_{i,h} \circ \bm{F}_i}\bm{1} = \\
-&\tilde{\bm{v}}^T\LRp{\LRp{\bm{Q}_{i,h}-\bm{Q}_{i,h}^T} \circ \bm{F}_i}\bm{1}  + \LRp{\bm{E}\bm{v}(\bm{u}_h)}^T\LRp{\bm{B}_{i,h} \circ\bm{F}_i}\bm{1},
-\end{align*}
-where we have introduced $\tilde{\bm{v}} = \LRp{\begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}\bm{v}(\bm{u}_h)}$.  
-Note that, by the consistency of the entropy conservative flux $\bm{f}_{i,S}$, the diagonal of $\bm{F}_i$ is the vector $\bm{f}_i(\tilde{\bm{u}}_f)$.
+%\[
+%\bm{v}(\bm{u}_h)^T\bm{M}\td{\bm{u}}{t} = \bm{1}^T\bm{M}\LRp{\bm{v}(\bm{u}_h)\circ \td{\bm{u}_h}{t}} =  \bm{1}^T\bm{M}\td{S(\bm{u}_h)}{t}.
+%\]
+%For the volume term, we use the SBP property to arrive at
+%\begin{align*}
+%&\bm{v}(\bm{u}_h)^T\begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T\LRp{2\bm{Q}_{i,h} \circ \bm{F}_i}\bm{1} = \\
+%&\tilde{\bm{v}}^T\LRp{\LRp{\bm{Q}_{i,h}-\bm{Q}_{i,h}^T} \circ \bm{F}_i}\bm{1}  + \LRp{\bm{E}\bm{v}(\bm{u}_h)}^T\LRp{\bm{B}_{i,h} \circ\bm{F}_i}\bm{1},
+%\end{align*}
+%where we have introduced $\tilde{\bm{v}} = \LRp{\begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}\bm{v}(\bm{u}_h)}$.  
+%Note that, by the consistency of the entropy conservative flux $\bm{f}_{i,S}$, the diagonal of $\bm{F}_i$ is the vector $\bm{f}_i(\tilde{\bm{u}}_f)$.
 Using that $\bm{B}_{i,h}$ is diagonal (with non-zero diagonal sub-block $\bm{B}_i$), the latter term reduces to
-\[
-\LRp{\bm{E}\bm{v}(\bm{u}_h)}^T\LRp{\bm{B}_{i,h} \circ\bm{F}_i}\bm{1} = \bm{v}_f^T \bm{B}_i\bm{f}_i(\tilde{\bm{u}}_f) =  \bm{1}^T \bm{B}_i\LRp{\bm{v}_f^T\bm{f}_i(\tilde{\bm{u}}_f)}.
-\]
-The final part of the proof uses that 
-\begin{align*}
-\tilde{\bm{v}}^T\LRp{\LRp{\bm{Q}_{i,h}-\bm{Q}_{i,h}^T} \circ \bm{F}_i}\bm{1} &= \sum_j {\tilde{\bm{v}}}_j \LRp{\bm{Q}_{i,h}-\bm{Q}_{i,h}^T}_{jk} \LRp{\bm{F}_{i}}_{jk}\\
-&= \sum_j \LRp{\bm{Q}_h}_{jk} \LRp{{\tilde{\bm{v}}}_j-{\tilde{\bm{v}}}_k}^T\bm{f}_{i,S}\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}\\
-&= \sum_j \LRp{\bm{Q}_h}_{jk} \LRp{\psi(\tilde{\bm{u}}_j)-\psi(\tilde{\bm{u}}_k)} = \tilde{\bm{u}}^T\bm{Q}_{i,h}\bm{1}- \bm{1}^T\bm{Q}_{i,h}\tilde{\bm{u}}
-\end{align*}
-The term $\bm{Q}_{i,h}\bm{1} = \bm{0}$ by (\ref{eq:Qh1}).
+%\[
+%\LRp{\bm{E}\bm{v}(\bm{u}_h)}^T\LRp{\bm{B}_{i,h} \circ\bm{F}_i}\bm{1} = \bm{v}_f^T \bm{B}_i\bm{f}_i(\tilde{\bm{u}}_f) =  \bm{1}^T \bm{B}_i\LRp{\bm{v}_f^T\bm{f}_i(\tilde{\bm{u}}_f)}.
+%\]
+%The final part of the proof uses that 
+%\begin{align*}
+%\tilde{\bm{v}}^T\LRp{\LRp{\bm{Q}_{i,h}-\bm{Q}_{i,h}^T} \circ \bm{F}_i}\bm{1} &= \sum_j {\tilde{\bm{v}}}_j \LRp{\bm{Q}_{i,h}-\bm{Q}_{i,h}^T}_{jk} \LRp{\bm{F}_{i}}_{jk}\\
+%&= \sum_j \LRp{\bm{Q}_h}_{jk} \LRp{{\tilde{\bm{v}}}_j-{\tilde{\bm{v}}}_k}^T\bm{f}_{i,S}\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}\\
+%&= \sum_j \LRp{\bm{Q}_h}_{jk} \LRp{\psi(\tilde{\bm{u}}_j)-\psi(\tilde{\bm{u}}_k)} = \tilde{\bm{u}}^T\bm{Q}_{i,h}\bm{1}- \bm{1}^T\bm{Q}_{i,h}\tilde{\bm{u}}
+%\end{align*}
+%The term $\bm{Q}_{i,h}\bm{1} = \bm{0}$ by (\ref{eq:Qh1}).
 The proof is completed by applying the SBP property (\ref{eq:hsbp}) to the remaining term, which reduces to $-\bm{1}^T\bm{B}_{i,h}\tilde{\bm{u}}=-\bm{1}^T\bm{B}_i\tilde{\bm{u}_f}$.
-\qed\end{proof}
-
-Note that the proof of entropy conservation relies mainly on Lemma~\ref{lemma:Qhprops}, which states that $\bm{Q}_{i,h}$ is conservative (e.g., exact for constants) and satisfies the SBP property.
+%\qed\end{proof}
+The proof of entropy conservation relies mainly on Lemma~\ref{lemma:Qhprops}, which states that $\bm{Q}_{i,h}$ is conservative (e.g., exact for constants) and satisfies the SBP property.
 Extensions to the non-conforming setting will utilize the same properties.
@@
 \centering
-\subfloat[Flux evaluations required for aligned surface nodes]{\raisebox{.1em}{\includegraphics[width=.4\textwidth]{figs/aligned.png}}\label{subfig:aligned}}
+\subfloat[Flux evaluations required for aligned surface nodes]{\raisebox{.1em}{\includegraphics[width=.375\textwidth]{figs/aligned.png}}\label{subfig:aligned}}
 \hspace{4em}
-\subfloat[Flux evaluations required for non-aligned surface nodes]{\includegraphics[width=.39\textwidth]{figs/nonaligned.png}\label{subfig:nonaligned}}
+\subfloat[Flux evaluations required for non-aligned surface nodes]{\includegraphics[width=.37\textwidth]{figs/nonaligned.png}\label{subfig:nonaligned}}
 \caption{Nodes between which flux evaluations are required for Gauss nodes.
@@
 The matrix $\bm{E}$ maps from volume quadrature points to surface quadrature points, $\bm{E}_m$ maps from surface to mortar surface points, and $\tilde{\bm{E}}_m$ maps from mortar surface points to surface points.
-}
+\note{Fix notation after Mario standardizes it.}}
 \label{fig:gqcon_noncon}
@@
 The matrix $\bm{E}_{fm}$ can now be defined through a quadrature-based $L^2$ projection 
-\[
+\begin{equation}
 \bm{E}_{fm} = \bm{M}_f^{-1}\bm{E}_{mf}^T\bm{M}_m, \qquad \bm{M}_m = \bm{I}_{4\times 4}\otimes \diag{\bm{w}_m}.
-\]
+\label{eq:Efm}
+\end{equation}
 Finally, we introduce diagonal boundary matrices on surface and mortar nodes 
@@
 \begin{lemma}
+\label{lemma:accuracy}
 Let $f(\bm{x}) = 1$ and suppose that the face (surface) quadrature is exact for polynomials of degree $N+N_f$ and mortar quadratures are exact for degree $N+N_m$ polynomials.
@@
 
-\subsection{Mortar-based entropy conservative formulations on mapped elements}
-
-We can extend the above scheme to mapped curvilinear elements by constructing physical mortar-based hybridized SBP operators in a manner akin to (\ref{eq:curvedQ}).
+\subsection{Mortar-based SBP operators and entropy conservative formulations on mapped elements}
+
+We can extend mortar-based SBP operators to mapped curvilinear elements by constructing physical operators in a manner akin to (\ref{eq:curvedQ}).
 Abusing notation, we use $\bm{G}_{ij}$ to now denote the vector containing values of $G_{ij} = J\pd{\hat{\bm{x}}_j}{\bm{x}_i}$ at volume, surface, and mortar points.
@@
 
-We assume now that each geometric mapping from reference element $\hat{D}$ to physical element $D^k$ is a polynomial of degree $N_{\rm geo}$.
+We assume now that each geometric mapping from reference element $\hat{D}$ to physical element $D^k$ is a polynomial of degree $N_{\rm geo} \leq N$.
 We also introduce the space of tensor product polynomials in $d$ dimensions $Q^{N_1,\ldots, N_d}$ as
@@
 \end{enumerate}
-then 
+then the following properties hold:
 \begin{align*}
 \bm{Q}_{i,m} + \bm{Q}_{i,m}^T &= \bm{B}_{i,m}, \quad \text{(SBP property)}\\
-\bm{Q}_{i,m}\bm{1} &= \bm{0}, \quad \text{(conservation)}.
-\end{align*}
+\bm{Q}_{i,m}\bm{1} &= \bm{0}, \quad \text{(conservation)},
+\end{align*}
+
+where the boundary matrix 
+\begin{equation}
+\bm{B}_{i,m} = \diag{\bm{n}_{i,m}\circ \bm{w}_m}
+\end{equation}
+is the diagonal matrix whose entries consist of the scaled physical normals. 
 \end{lemma}
@@
 Using the SBP property of $\hat{\bm{Q}}_{i,m}$, proving the SBP property follows the same steps as the proof of Lemma~\ref{lemma:Qhprops}.  
+Expanding out $\bm{Q}_{i,m} + \bm{Q}_{i,m}^T$ yields
+\[
+\frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,m} + \hat{\bm{Q}}_{j,m}^T\diag{\bm{G}_{ij}} + \hat{\bm{Q}}_{j,m} \diag{\bm{G}_{ij}} +  \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,m}^T.
+\]
+For each term in the sum, we use the SBP property of $\hat{\bm{Q}}_{j,m}$.  The volume terms cancel, leaving only surface terms
+\[
+\frac{1}{2}\sum_{j=1}^d \LRp{\hat{\bm{B}}_{j,m}\diag{\bm{G}_{ij}} +  \diag{\bm{G}_{ij}}\hat{\bm{B}}_{j,m}} = \begin{bmatrix}
+\bm{0} & &\\
+& \bm{0} &\\
+& & \bm{B}_{i,m} \end{bmatrix}.
+\]
+where we have used (\ref{eq:nJ_Gnhat}) and that $\hat{\bm{B}}_{j,m}$ and $\diag{\bm{G}_{ij}}$ are diagonal matrices.
+
+%To show $\bm{Q}_{i,m}\bm{1} = \bm{0}$, we note that
+%\[
+%\bm{Q}_{i,m}\bm{1} = \frac{1}{2}\sum_{j=1}^d \diag{\bm{G}_{ij}}\hat{\bm{Q}}_{j,m}\bm{1} + \hat{\bm{Q}}_{j,m} \diag{\bm{G}_{ij}}\bm{1} = \frac{1}{2}\sum_{j=1}^d \hat{\bm{Q}}_{j,m} \bm{G}_{ij}
+%\]
+%since $\hat{\bm{Q}}_{j,m}\bm{1} = \bm{0}$ by Lemma~\ref{lemma:hsbprefprops}.
+Note that, since $\bm{G}_{ij}$ is constructed using polynomial interpolation, $\bm{G}_{ij} = \LRs{G_{ij}(\hat{\bm{x}}), \quad \bm{E}G_{ij}(\hat{\bm{x}})}^T$, where $G_{ij}(\hat{\bm{x}})$ denotes the values of $G_{ij}$ at volume points.
+The remaining term can be expanded out
+%\begin{align*}
+%\sum_{j=1}^d \hat{\bm{Q}}_{j,m} \bm{G}_{ij} &= \sum_{j=1}^d \begin{bmatrix} \hat{\bm{Q}}_jG_{ij}(\hat{\bm{x}}) - \hat{\bm{Q}}_j^TG_{ij}(\hat{\bm{x}}) + \bm{E}^T\bm{B}\bm{E}G_{ij}(\hat{\bm{x}})\\
+%\bm{B}\bm{E}G_{ij}(\hat{\bm{x}}) - \bm{B}\bm{E}G_{ij}(\hat{\bm{x}})
+%\end{bmatrix} \\
+%&=
+%\sum_{j=1}^d \begin{bmatrix} \LRp{- \hat{\bm{Q}}_j^T + \bm{E}^T\bm{B}\bm{E}}G_{ij}(\hat{\bm{x}})\\
+%\bm{0}
+%\end{bmatrix} = 
+%\sum_{j=1}^d \begin{bmatrix} \hat{\bm{Q}}_jG_{ij}(\hat{\bm{x}})\\
+%\bm{0}
+%\end{bmatrix}  = \bm{0},
+%\end{align*}
+%where we have used  the generalized SBP property and that $\sum_{j=1}^d\hat{\bm{Q}}_jG_{ij}(\hat{\bm{x}}) = \bm{0}$, since $G_{ij}$ satisfies the discrete GCL (\ref{eq:dgcl}), $G_{ij}(\hat{\bm{x}})$ is a polynomial of degree less than or equal to $N$, and $\hat{\bm{Q}}_j$ exactly differentiates polynomials of degree $N$.
+
 To show conservation, expanding out $\bm{Q}_{i,m}\bm{1} $ yields
@@
 
-An entropy conservative mortar-based formulation on mapped elements is then 
+An entropy conservative mortar formulation on a mapped element is then 
 \begin{gather}
@@
 \tilde{\bm{u}}_m
-\end{bmatrix}, \qquad \tilde{\bm{u}}_m = \bm{u}\LRp{\bm{E}_{mf}\bm{E}\bm{v}(\bm{u}_h)}, \qquad \bm{f}^*_i = \bm{f}_{i,S}\LRp{\tilde{\bm{u}}_m^+,\tilde{\bm{u}}_m},
+\end{bmatrix}, \qquad \tilde{\bm{u}}_m = \bm{u}\LRp{\bm{E}_{mf}\bm{E}\bm{v}(\bm{u}_h)}, \qquad \bm{f}^*_i = \bm{f}_{i,S}\LRp{\tilde{\bm{u}}_m^+,\tilde{\bm{u}}_m},\nonumber
 \end{gather}
-where $\bm{B}_{i,m} = \diag{\bm{n}_{i,m}\circ \bm{w}_m}$ is the diagonal matrix whose entries consist of the scaled physical normals.
 Again, we assume geometric terms on curved elements are approximated using polynomials, and that the normals are constructed via (\ref{eq:nJ_Gnhat}) and polynomial interpolation.
 
-\note{Add proof - include multi-element case}.
-The proof of entropy conservation is algebraically identical to previous proofs, with Lemma~\ref{lemma:Qmprops_3d} used in place of Lemma~\ref{lemma:Qhprops}.
-
-\subsection{Theoretical comparisons of Lobatto and Gauss nodes}
-
-High order entropy stable collocation schemes on tensor product elements most commonly utilize tensor product quadrature rules constructed from $(N+1)$-point Lobatto nodes \cite{carpenter2014entropy, gassner2016split} or $(N+1)$-point Gauss nodes \cite{chan2018efficient}.
-On non-conforming meshes, the most natural choice of mortar nodes for such collocation schemes is a composite Lobatto or Gauss rule.
-
+\begin{theorem}
+Assuming continuity in time, the local formulation (\ref{eq:esdgm}) satisfies 
+\begin{equation}
+\bm{1}^T\bm{M}\td{S(\bm{u}_h)}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}_{i,m}\LRp{\bm{v}_m^T\bm{f}_i^* - \psi_i(\tilde{\bm{u}}_m)} = 0, \qquad \bm{v}_m =\bm{E}_{mf}\bm{E}\bm{v}(\bm{u}_h).
+\label{eq:localec}
+\end{equation}
+\end{theorem}
+\begin{proof}
+The main steps are algebraically very similar to proofs in \cite{chan2018discretely, chan2018efficient, chan2019skew}, with Lemma~\ref{lemma:Qmprops_3d} used in place of Lemma~\ref{lemma:Qhprops}.
+We begin by testing (\ref{eq:esdg}) with the vector of entropy variables evaluated at nodal points $\bm{v}(\bm{u}_h)$.
+We apply the chain rule in time and use that $\bm{M}$ is diagonal to show that
+\[
+\bm{v}(\bm{u}_h)^T\bm{M}\td{\bm{u}}{t} = \bm{1}^T\bm{M}\LRp{\bm{v}(\bm{u}_h)\circ \td{\bm{u}_h}{t}} =  \bm{1}^T\bm{M}\td{S(\bm{u}_h)}{t}.
+\]
+For the volume term, we use the SBP property (Lemma~\ref{lemma:Qmprops_3d}) to arrive at
+\begin{align*}
+&\bm{v}(\bm{u}_h)^T\begin{bmatrix} \bm{I} \\ \bm{E} \\ \bm{E}_{mf}\bm{E} \end{bmatrix}^T\LRp{2\bm{Q}_{i,m} \circ \bm{F}_i}\bm{1} = \\
+&\tilde{\bm{v}}^T\LRp{\LRp{\bm{Q}_{i,m}-\bm{Q}_{i,m}^T} \circ \bm{F}_i}\bm{1}  + \LRp{\bm{E}\bm{v}(\bm{u}_h)}^T\LRp{\begin{bmatrix}\bm{0} && \\ & \bm{0} & \\ & & \bm{B}_{i,m}\end{bmatrix} \circ\bm{F}_i}\bm{1},
+\end{align*}
+where we have introduced 
+\[
+\tilde{\bm{v}} = {\begin{bmatrix} \bm{I} \\ \bm{E} \\ \bm{E}_{mf}\bm{E} \end{bmatrix}\bm{v}(\bm{u}_h)}.  
+\]
+Note that, by the consistency of the entropy conservative flux $\bm{f}_{i,S}$, the diagonal of $\bm{F}_i$ is the flux evaluated at volume, surface, and mortar points. Thus, the latter term reduces to 
+\[
+\LRp{\bm{E}\bm{v}(\bm{u}_h)}^T\LRp{\begin{bmatrix}\bm{0} && \\ & \bm{0} & \\ & & \bm{B}_{i,m}\end{bmatrix} \circ\bm{F}_i}\bm{1} = \bm{v}_m^T \bm{B}_{i,m}\bm{f}_i(\tilde{\bm{u}}_m) =  \bm{1}^T \bm{B}_{i,m}\LRp{\bm{v}_m^T\bm{f}_i(\tilde{\bm{u}}_m)}.
+\]
+where we have used that $\bm{B}_{i,m}$ is diagonal.  The final part of the proof uses that 
+\begin{align*}
+\tilde{\bm{v}}^T\LRp{\LRp{\bm{Q}_{i,m}-\bm{Q}_{i,m}^T} \circ \bm{F}_i}\bm{1} &= \sum_j {\tilde{\bm{v}}}_j \LRp{\bm{Q}_{i,m}-\bm{Q}_{i,m}^T}_{jk} \LRp{\bm{F}_{i}}_{jk}\\
+&= \sum_j \LRp{\bm{Q}_{i,m}}_{jk} \LRp{{\tilde{\bm{v}}}_j-{\tilde{\bm{v}}}_k}^T\bm{f}_{i,S}\LRp{\tilde{\bm{u}}_j,\tilde{\bm{u}}_k}\\
+&= \sum_j \LRp{\bm{Q}_{i,m}}_{jk} \LRp{\psi(\tilde{\bm{u}}_j)-\psi(\tilde{\bm{u}}_k)} = \tilde{\bm{u}}^T\bm{Q}_{i,m}\bm{1} - \bm{1}^T\bm{Q}_{i,m}\tilde{\bm{u}}
+\end{align*}
+By Lemma~\ref{lemma:Qmprops_3d}, $\bm{Q}_{i,m}\bm{1} = \bm{0}$.
+The proof is completed by applying the SBP property (\ref{eq:hsbp}) to the remaining term, which reduces to $- \bm{1}^T\bm{Q}_{i,m}\tilde{\bm{u}}=-\bm{1}^T\bm{B}_{i,m}\tilde{\bm{u}}_m$.
+\qed\end{proof}
+
+Global entropy conservation can be derived by summing up (\ref{eq:localec}) over each element
+\[
+\sum_k \bm{1}^T\bm{M}^k\td{S(\bm{u}^k_h)}{t} + \sum_k \sum_{i=1}^d \bm{1}^T\bm{B}^k_{i,m}\LRp{\LRp{\bm{v}^k_m}^T\bm{f}^{k,*}_i - \psi_i(\tilde{\bm{u}}^k_m)} = 0,
+\]
+where have introduced the superscript $k$ to denote operators and local solutions on the $k$th element $D^k$.
+The first term corresponds to the average entropy over the entire domain.
+Assume without loss of generality a periodic domain, such that all interfaces are interior interfaces.
+Sharing interface contributions involving the numerical flux between neighboring elements yields
+\begin{align}
+\sum_k &\sum_{i=1}^d \bm{1}^T\bm{B}^k_{i,m}\LRp{\LRp{\bm{v}^k_m}^T\bm{f}^{k,*}_i - \psi_i(\tilde{\bm{u}}^k_m)} = \label{eq:ecinterfaces}\\
+&\sum_k \sum_{i=1}^d \bm{1}^T\bm{B}^k_{i,m}\LRp{\frac{1}{2}\LRp{\bm{v}^k_m- \bm{v}^{k,+}_m }^T\bm{f}^{k,*}_i - \psi_i(\tilde{\bm{u}}^k_m)}\nonumber
+\end{align}
+where $\bm{v}^{k,+}_m$ denotes the exterior value of the entropy variables on the neighboring element.
+If the numerical flux is computed using the entropy conservative flux $\bm{f}^{k,*}_i = \bm{f}_{i,S}\LRp{\tilde{\bm{u}}^k_m,\tilde{\bm{u}}^{k,+}_m}$, then
+\[
+\LRp{\bm{v}^k_m- \bm{v}^{k,+}_m }^T\bm{f}_{i,S}\LRp{\tilde{\bm{u}}^k_m,\tilde{\bm{u}}^{k,+}_m} = \psi_i(\tilde{\bm{u}}^k_m) - \psi_i(\tilde{\bm{u}}^{k,+}_m).
+\]
+Returning the contributions $\psi_i(\tilde{\bm{u}}^{k,+}_m)$ to the neighboring element simplifies (\ref{eq:ecinterfaces}) as follows
+\begin{align*}
+&\sum_k \sum_{i=1}^d \bm{1}^T\bm{B}^k_{i,m}\LRp{\frac{1}{2}\LRp{\bm{v}^k_m- \bm{v}^{k,+}_m }^T\bm{f}^{k,*}_i - \psi_i(\tilde{\bm{u}}^k_m)}\\
+&= \sum_k \sum_{i=1}^d \bm{1}^T\bm{B}^k_{i,m}\LRp{\frac{1}{2}\LRp{\psi_i(\tilde{\bm{u}}^k_m) - \psi_i(\tilde{\bm{u}}^{k,+}_m) - \psi_i(\tilde{\bm{u}}^k_m)}}\\
+&= \sum_k \sum_{i=1}^d \bm{1}^T\bm{B}^k_{i,m}\LRp{\psi_i(\tilde{\bm{u}}^k_m) - \psi_i(\tilde{\bm{u}}^k_m)} = 0.
+\end{align*}
+When the domain is non-periodic, the interface contributions do not vanish at the boundary.
+These boundary terms can be made non-negative (e.g., entropy conservative or entropy stable) if boundary conditions are imposed in an appropriate manner \cite{svard2014entropy, chen2017entropy}.
+
+%\note{Add proof - include multi-element case}.
+
+\subsection{Comparison of Lobatto and Gauss nodes: requirements for entropy conservation (stability) and accuracy}
+\label{sec:compare}
+
+This section summarizes theoretical differences between entropy conservative (stable) Lobatto and Gauss collocation schemes on non-conforming meshes based on Lemma~\ref{lemma:Qmprops_3d} and Lemma~\ref{lemma:accuracy}.
+The main differences are that the use of Lobatto nodes induces additional requirements on the construction of geometric terms for curved meshes and lowers the expected order of accuracy by one degree.
+
+High order entropy stable collocation schemes on tensor product elements typically utilize tensor product quadrature rules based on $(N+1)$-point Lobatto nodes \cite{carpenter2014entropy, gassner2016split} or $(N+1)$-point Gauss nodes \cite{chan2018efficient}.
+On non-conforming meshes, the most natural choice of mortar nodes is a composite Lobatto or Gauss rule.
 Recall that an $(N+1)$-point Lobatto rule is exact for polynomials of degree $(2N-1)$, while an $(N+1)$-point Gauss rule is exact for polynomials of degree $(2N+1)$.
+
+First, we discuss the expected accuracy of each scheme.
+We have not conducted a rigorous error analysis; however, Lemma~\ref{lemma:accuracy} implies that a mortar-based SBP operator under Lobatto quadrature is exact for degree $(N-1)$ polynomials, while a mortar-based SBP operator under Gauss quadrature is exact for degree $N$ polynomials.
+In other words, the order of accuracy of a Gauss collocation scheme can be up to one degree higher than for a Lobatto collocation scheme.
+
+We note that order reduction for Lobatto nodes is not consistently evident in numerical experiments, both for conforming \cite{chan2018efficient, hindenlang2019order} and non-conforming meshes \cite{fernandez2019entropy}.
+In \cite{friedrich2017entropy}, order reduction with respect to the $L^2$ error was observed using Lobatto nodes on 2D non-conforming meshes.
+However, in \cite{fernandez2019entropy}, $L^2$ order reduction is not observed using Lobatto nodes on 3D non-conforming hexahedral meshes, though order reduction is observed for the $L^\infty$ error.
+Similar inconsistent behavior is observed for the 2D and 3D experiments in this paper when using Lobatto nodes.
+We do note observe order reduction for Gauss nodes in any of our experiments.
+
+Next, we discuss requirements on the construction of geometric terms.
 On non-conforming quadrilateral meshes, Theorem~\ref{lemma:Qmprops_3d} implies that both Lobatto and Gauss collocation schemes are stable for $N_{\rm geo}\leq N$, i.e., for isoparametric curved mappings.
@@
 
-We note that the restriction on the degree of the geometric mapping $N_{\rm geo} \leq (N-1)$ for Lobatto nodes is not necessary on conforming hexahedral meshes.
-However, the introduction of non-aligned mortar nodes at non-conforming curved interfaces induces this additional restriction.
-Moreover, this restriction is consistent with Theorem 3 in \cite{fernandez2019entropy}.
+The restriction on $N_{\rm geo} \leq (N-1)$ for Lobatto nodes is not necessary for entropy conservation (stability) on conforming hexahedral meshes.
+However, when using Approach 1 to compute geometric terms, the introduction of non-aligned mortar nodes at non-conforming curved interfaces induces this additional restriction.
+Moreover, this condition is consistent with Theorem 3 in \cite{fernandez2019entropy}.
 One can remove this constraint on the degree of the geometric mapping by using Approach 2 \cite{kozdon2018energy} to compute geometric terms.
 If Approach 2 is used, then Lobatto collocation schemes remain entropy conservative (stable) on curved hexahedral meshes for all $N_{\rm geo} \leq N$.
-
 Appendix~\ref{app:A} describes Approach 1 \cite{kopriva2006metric} and Approach 2 \cite{kozdon2018energy} in more detail, and demonstrates numerically that both approaches achieve similar high order accuracy in approximating geometric terms.
@@
 \section{A mortar-based implementation}
-
+\label{sec:mortarimplement}
 While the formulation (\ref{eq:esdgm}) is convenient for analysis, it is less convenient to implement.
@@
 &+ \bm{E}^T\LRp{\bm{E}_{mf}^T \LRp{\hat{\bm{B}}_{i,m} \circ \hat{\bm{F}}_i^{mm}}\bm{1} - \LRp{\bm{B}_i\circ \hat{\bm{F}}^{ff}_i}\bm{1} } \\
-&+ \bm{E}^T \LRp{\LRp{\hat{\bm{B}}_i\bm{E}_{fm}}\circ \hat{\bm{F}}_i^{fm}}\bm{1} - \bm{E}_{mf}^T \LRp{ \LRp{\hat{\bm{B}}_i\bm{E}_{mf}} \circ \hat{\bm{F}}_i^{mf}}\bm{1}.
+&+ \bm{E}^T \LRp{\LRp{\hat{\bm{B}}_i\bm{E}_{fm}}\circ \hat{\bm{F}}_i^{fm}}\bm{1} - \bm{E}^T\bm{E}_{mf}^T \LRp{ \LRp{\hat{\bm{B}}_i\bm{E}_{mf}} \circ \hat{\bm{F}}_i^{mf}}\bm{1}.
 \end{align*}
@@
 \begin{align*}
-&\bm{E}^T \LRp{\LRp{\hat{\bm{B}}_i\bm{E}_{fm}}\circ \hat{\bm{F}}_i^{fm}}\bm{1} - \bm{E}_{mf}^T \LRp{ \LRp{\hat{\bm{B}}_{i,m}\bm{E}_{mf}} \circ \hat{\bm{F}}_i^{mf}}\bm{1}\\
-& = \bm{E}^T \LRp{{\bm{E}_{fm}}\circ \bm{F}_i^{fm} \circ \avg{\bm{n}_i}_{fm}}\bm{1} - \bm{E}_{mf}^T \LRp{ {\bm{E}_{mf}} \circ {\bm{F}}_i^{mf} \circ \avg{\bm{n}_i}_{mf}}\bm{1}
+&\bm{E}^T \LRp{\LRp{\hat{\bm{B}}_i\bm{E}_{fm}}\circ \hat{\bm{F}}_i^{fm}}\bm{1} - \bm{E}^T\bm{E}_{mf}^T \LRp{ \LRp{\hat{\bm{B}}_{i,m}\bm{E}_{mf}} \circ \hat{\bm{F}}_i^{mf}}\bm{1}\\
+& = \bm{E}^T \LRp{{\bm{E}_{fm}}\circ \bm{F}_i^{fm} \circ \avg{\bm{n}_i}_{fm}}\bm{1} - \bm{E}^T\bm{E}_{mf}^T \LRp{ {\bm{E}_{mf}} \circ {\bm{F}}_i^{mf} \circ \avg{\bm{n}_i}_{mf}}\bm{1}
 \end{align*}
@@
 \sum_{i=1}^d \LRp{2{\bm{Q}}_{i,h} \circ \bm{F}_i}\bm{1} + \bm{E}^T\LRp{\bm{f}_i^*- \bm{B}_i\bm{f}_i(\tilde{\bm{u}}_f)} = 0 \label{eq:esdgm_mortar}\\
-\bm{f}_i^* = \bm{E}_{mf}^T\LRp{{\bm{B}}_{i,m}\bm{f}_{i,m} + \LRp{{\bm{E}_{fm}}\circ \bm{F}_i^{fm} \circ \avg{\bm{n}_i}_{fm} - {\bm{E}_{mf}} \circ {\bm{F}}_i^{mf} \circ \avg{\bm{n}_i}_{mf}}\bm{1}},\nonumber
+\bm{f}_i^* = \bm{E}_{mf}^T{\bm{B}}_{i,m}\bm{f}^*_{i,m} + \bm{\delta f},\nonumber\\
+\bm{\delta f} =   \LRp{{\bm{E}_{fm}}\circ \bm{F}_i^{fm} \circ \avg{\bm{n}_i}_{fm}}\bm{1} - \bm{E}_{mf}^T \LRp{ {\bm{E}_{mf}} \circ {\bm{F}}_i^{mf} \circ \avg{\bm{n}_i}_{mf}}\bm{1}\nonumber
 \end{gather}
-where $\bm{f}_{i,m} = \bm{f}_{i,S}(\tilde{\bm{u}}_m^+,\tilde{\bm{u}}_m)$ is the numerical flux computed using the  at mortar points.
+where $\bm{f}_{i,m} = \bm{f}_{i,S}(\tilde{\bm{u}}_m^+,\tilde{\bm{u}}_m)$ is the numerical flux computed at mortar points.
+
+\begin{figure}
+\centering
+\includegraphics[width=.55\textwidth]{mortar_coupling.png}
+\caption{Coupling between volume, surface, and mortar nodes for formulation (\ref{eq:esdgm_mortar}).}
+\end{figure}
 The formulation (\ref{eq:esdgm_mortar}) is nearly equivalent to the hybridized mortar formulation, with the main difference being the presence of skew-symmetric correction terms in the definition of the numerical flux.
@@
 \begin{remark}
-If the surface and mortar nodes are the same, then 
+On conforming meshes, the surface and mortar nodes are the same.  Then,  
 \[
@@
 \]
-and the correction terms in the definition of $\bm{f}^*_i$ cancel out.
+and the correction terms in the definition of $\bm{f}^*_i$ cancel each other out.
 The formulation (\ref{eq:esdgm_mortar}) then reverts to a Gauss collocation scheme \cite{chan2017discretely} on conforming meshes.
@@
 
-\begin{figure}
+\subsection{Improving efficiency on non-conforming hexahedral meshes}
+\label{sec:anisotropic}
+
+In 3D, the mortar-based approach enables a more efficient treatment of non-conforming hexahedral interfaces compared to existing approaches \cite{friedrich2017entropy}.
+The cost of entropy stable schemes based on flux differencing usually scales with the number of flux evaluations, which in turn scales by the number of non-zero entries present in differentiation and interpolation matrices.
+For a hexahedral element, both Gauss and Lobatto node differentiation operators contain $O(N^4)$ non-zeros due to their tensor product (Kronecker) structure.
+
+\begin{figure}[!h]
 \centering
-\includegraphics[width=.55\textwidth]{mortar_coupling.png}
-\caption{Coupling between volume, surface, and mortar nodes for formulation (\ref{eq:mortarform}).}
+\includegraphics[width=.5\textwidth]{noncon3D.png}
+\caption{Two-layer mortars for reduced $O(N^3)$ computational cost on non-conforming hexahedral interfaces.}
+\label{fig:aniso}
 \end{figure}
 
+The interpolation matrices are generally dense with $O(N^4)$ non-zero entries, as they interpolate from $(N+1)^2$ surface points to $4(N+1)^2$ non-aligned mortar points (for an isotropically refined face).
+However, we can expose sparsity by using a two-step interpolation from surface to mortar points.
+Figure~\ref{fig:aniso} illustrates this process.
+First, the $(N+1)^2$ surface points are interpolated to an anisotropically refined interface.
+Because the points are aligned along one coordinate direction, this requires interpolation only along lines of nodes and results in an interpolation matrix with $O(N^3)$ non-zero entries.
+The anisotropically refined nodes are then interpolated to the isotropically refined nodes.
+Again, because nodes are aligned, the resulting interpolation matrix has $O(N^3)$ non-zero entries.
+
+\begin{figure}[!h]
+\centering
+\subfloat[$\bm{E}_{m_1,f}$]{\includegraphics[width=.2\textwidth]{E12.png}}
+\hspace{4em}
+\subfloat[$\bm{E}_{m_2,m_1}$]{\includegraphics[width=.2045\textwidth]{E23.png}}
+\caption{Sparsity patterns of the interpolation operators $\bm{E}_{m_1,f}, \bm{E}_{m_2,f}$.}
+\label{fig:anisointerp}
+\end{figure}
+
+This two-step procedure can be interpreted as a sparse factorization of the interpolation matrix from $(N+1)^2$ surface points to $4(N+1)^2$ points.
+Let $\bm{x}_f$ denote the surface points, and let $\bm{x}_{m_1}, \bm{x}_{m_2}$ denote the anisotropically and isotropically refined mortar nodes (e.g., the first and second mortar layers), respectively.
+We first introduce $\bm{E}_{\rm 1D}$ be the 1D polynomial interpolation matrix which maps from Gauss or Lobatto points to split (composite) Gauss or Lobatto points.
+Let $\bm{E}_{m_1,f}$ denote the interpolation operator from surface to nodes in the first mortar layer, and let $\bm{E}_{m_2,m_1}$ denote the interpolation operator from nodes in the first mortar layer to the second mortar layer.
+These operators are defined
+\[
+\bm{E}_{m_1,f} = \bm{E}_{\rm 1D} \otimes \bm{I}_{N+1}, \qquad \bm{E}_{m_2,m_1} = \bm{I}_{2(N+1)} \otimes \bm{E}_{\rm 1D}.
+\]
+where $\bm{I}_{N+1},\bm{I}_{2(N+1)}$ denote the $(N+1)\times (N+1)$ and $2(N+1)\times 2(N+1)$ identity matrices, respectively.
+Then, $\bm{E}_{m_2,f} = \bm{E}_{m_2,m_1}\bm{E}_{m_1,f}$ interpolates from surface nodes to the second layer of mortar nodes.
+However, the operators $\bm{E}_{m_2,m_1}, \bm{E}_{m_1,f}$ are sparse (as shown in Figure~\ref{fig:anisointerp}), while $\bm{E}_{m_2,f}$ is completely dense.
+
+We can also define projection operators which transfer from the second to the first mortar layer and from the first mortar layer to surface nodes.
+Let $\bm{M}_{m_i}$ denote the diagonal matrix containing quadrature weights for the $i$th layer of mortar nodes, respectively.
+Then, we can define $\bm{E}_{f,m_1},\bm{E}_{m_1,m_2}$ via
+\begin{align*}
+\bm{E}_{f,m_1} = \bm{M}_f^{-1}\bm{E}_{m_1,f}^T\bm{M}_{m_1}\\
+\bm{E}_{m_1,m_2} = \bm{M}_{m_1}^{-1}\bm{E}_{m_2,m_1}^T\bm{M}_{m_2}.
+\end{align*}
+The definitions of $\bm{E}_{f,m_1}, \bm{E}_{m_1,m_2}$ are analogous to the definition of $\bm{E}_{fm}$ in (\ref{eq:Efm}).
+The operator $\bm{E}_{f,m_1}$ can be interpreted as projecting onto degree $N$ polynomials on a quadrilateral face, while $\bm{E}_{m_1,m_2}$ can be interpreted as projecting onto the space of piecewise degree $N$ polynomials on the two quadrilateral faces of the first mortar layer in Figure~\ref{fig:aniso}.
+%The definition of $\bm{E}_{m_1,m_2}$ is more complicated, and transfers between $\bm{x}_{m_1},\bm{x}_{m_2}$ by performing an $L^2$ projection onto the polynomial trace space of a hexahedron (using the second layer of mortar quadrature nodes $\bm{x}_{m_2}$), then interpolating to the first layer of mortar nodes $\bm{x}_{m_1}$.
+
+Putting these pieces together allows us to define two-layer mortar-based SBP operators.  We introduce boundary matrices 
+\[
+\hat{\bm{B}}_{j,m_i} = \diag{\hat{\bm{n}}_{i,m_i}}\bm{M}_{m_i}, \qquad i = 1,2, \qquad j = 1,\ldots, d.
+\]
+where $\hat{\bm{n}}_{i,m_i}$ denote scaled outward normals evaluated at the $i$th layer of mortar points.  Then, the two-layer mortar-based SBP operator is
+\begin{equation}
+\hat{\bm{Q}}_{i,m} = \frac{1}{2}\begin{bmatrix}
+\hat{\bm{Q}}_i - \hat{\bm{Q}}_i^T & \bm{E}^T\hat{\bm{B}}_{i,f} & & \\
+-\hat{\bm{B}}_{i,f}\bm{E} & & \hat{\bm{B}}_{i,f} \bm{E}_{f,m_1} &\\
+& -\hat{\bm{B}}_{i,m_1} \bm{E}_{m_1,f} & & \hat{\bm{B}}_{i,m_1}\bm{E}_{m_1,m_2} \\
+&& -\hat{\bm{B}}_{i,m_2} \bm{E}_{m_2,m_1} & \hat{\bm{B}}_{i,m_2} 
+\end{bmatrix}.
+\end{equation}
+It is straightforward to show that, for both Gauss and Lobatto nodes, $\hat{\bm{Q}}_{i,m}$ satisfies SBP and conservation properties
+\[
+\hat{\bm{Q}}_{i,m} + \hat{\bm{Q}}_{i,m}^T = \begin{bmatrix}
+\bm{0} & & &\\
+& \bm{0} & &\\ 
+& & \bm{0} & \\
+& & & \bm{B}_{i,m_2}
+\end{bmatrix}, \qquad \hat{\bm{Q}}_{i,m}\bm{1} = \bm{0}.
+\]
+The accuracy conditions for two-layer mortar operators are also identical to those of Lemma~\ref{lemma:accuracy}.
+We can now define physical SBP operators via the formula (\ref{eq:curvedQ_mortar}).
+These also satisfy SBP and conservation properties under the conditions of Lemma~\ref{lemma:Qmprops_3d}.
+
+Using the operators introduced so far, an entropy conservative formulation using two-layer mortar operators is given by  
+\begin{align}
+{\bm{M}}\td{\bm{u}_h}{t} &+ \begin{bmatrix} \bm{I} \\ \bm{E} \\ \bm{E}_{m_1,f}\bm{E} \\ \bm{E}_{m_2,m_1}\bm{E}_{m_1,f}\bm{E} \end{bmatrix}^T
+\sum_{i=1}^d \LRp{2{\bm{Q}}_{i,m} \circ \bm{F}_i}\bm{1} \label{eq:twomortar}\\
+&+ \bm{E}^T \bm{E}_{m_1,f}^T\bm{E}_{m_2, m_1}^T{\bm{B}}_{i,m_2}\LRp{\bm{f}_i^*-\bm{f}_i(\tilde{\bm{u}}_{m_2})} = 0.
+\end{align}
+These formulations look very similar to (\ref{eq:esdgm}).  The two-layer mortar formulation can also be reformulated in a way which involves only face-local correction terms to the numerical flux:
+\begin{gather}
+{\bm{M}}\td{\bm{u}_h}{t} + \begin{bmatrix} \bm{I} \\ \bm{E} \end{bmatrix}^T
+\sum_{i=1}^d \LRp{2{\bm{Q}}_{i,h} \circ \bm{F}_i}\bm{1} + \bm{E}^T\LRp{\bm{f}_i^*- \bm{B}_i\bm{f}_i(\tilde{\bm{u}}_f)} = 0\\
+\bm{f}_i^* = \bm{E}_{m_1,f}^T\bm{E}_{m_2,m_1}^T{\bm{B}}_{i,m_2}\bm{f}_{i,m_2} + \bm{\delta f},\nonumber\\
+\bm{\delta f} = \begin{bmatrix}
+\bm{I}\\
+\bm{E}_{m_1,f}\\
+\bm{E}_{m_2,m_1}\bm{E}_{m_1,f}
+\end{bmatrix}^T
+\LRp{\begin{bmatrix}
+& \bm{E}_{f,m_1} & \\
+-\bm{E}_{m_1,f} & & \bm{E}_{m_1,m_2}\\
+& -\bm{E}_{m_2,m_1}& 
+\end{bmatrix}\circ \bm{F}_i^{f,m_1,m_2} \circ \avg{\bm{n}_i}}\bm{1}\nonumber
+\end{gather}
+where $\bm{\delta f}$ is a face-local correction term for the numerical flux, and the matrix $\bm{F}_i^{f,m_1,m_2}$ is the matrix of entropy conservative flux evaluations between solution values at surface nodes and nodes in the two mortar layers.
+The term $\avg{\bm{n}_i}$ denote the averaged scaled normals between each surface point and each point in the two mortar layers.
+The derivation of this face-local formulation is similar to the derivation in Section~\ref{sec:mortarimplement}, and we omit it for brevity.
+
+%\note{Explain how to fit into mortar SBP operator.}
+
 \section{Numerical experiments}
-
-\subsection{2D results}
-Figure~\ref{subfig:noncon} shows preliminary results for the 2D compressible Euler equations by the PI on a manually constructed non-conforming quadrilateral mesh, with volume, surface, and mortar quadratures constructed from one-dimensional Lobatto or Gauss quadrature rules.
-The PI has verified that this preliminary implementation of the mortar-based formulation is discretely entropy stable.
-A preliminary accuracy analysis for the skew-symmetric formulation (\ref{eq:esdgSkew}) also suggests that Lobatto quadrature should achieve an $O(h^N)$ rate of convergence and that Gauss quadrature should achieve an $O(h^{N+1})$ rate of convergence.
-Both predicted rates are confirmed in Figure~\ref{subfig:noncon}.
+\label{sec:4}
+This section presents numerical examples for the compressible Euler equations, which are given in $d$ dimensions as follows:
+\begin{align*}
+\pd{\rho}{t} &+ {\sum_{j=1}^d \pd{\LRp{\rho {u}_j}}{x_j}} = 0,\\
+{\pd{\rho {u}_i}{t}} &+ {\sum_{j=1}^d \pd{\LRp{\rho u_iu_j + p\delta_{ij} }}{x_j}} = 0, \qquad i = 1,\ldots,d\\
+\pd{E}{t} &+ {\sum_{j=1}^d \pd{\LRp{{u}_j(E+p)}}{x_j}} = 0.\nonumber
+\end{align*}
+Here, $\rho$ is density, $u_i$ denotes the $i$th component of the velocity, and $E$ is the total energy.  The pressure $p$ and specific internal energy $\rho e$ are given by 
+\[
+{p = (\gamma-1)\LRp{E - \frac{1}{2}\rho \nor{\bm{u}}^2}}, \qquad {\rho e = E - \frac{1}{2}\rho \nor{\bm{u}}^2}, \qquad \nor{\bm{u}}^2 = {\sum_{j=1}^d u_j^2}.
+\]
+We enforce the entropy conservation (stability) with respect to the unique entropy $S(\bm{u})$ for the viscous compressible Navier-Stokes equations \cite{hughes1986new} given by
+\begin{equation*}
+S(\bm{u}) = -\frac{\rho s}{\gamma-1}, 
+\end{equation*}
+where $s = \log\LRp{\frac{p}{\rho^\gamma}}$ is the specific entropy, and $d = 2,3$ is the dimension.  The entropy variables with respect to $S(\bm{u})$ are also derived in \cite{hughes1986new} 
+\begin{align*}
+v_1 = \frac{\rho e (\gamma + 1 - s) - E}{\rho e}, \qquad v_{1+ i}= \frac{\rho {{u}_i}}{\rho e}, \qquad v_{d+2} = -\frac{\rho}{\rho e}
+\end{align*}
+for $i = 1,\ldots, d$.  The inverse mapping is given by
+\begin{align*}
+\rho = -(\rho e) v_{d+2}, \qquad 
+\rho {u_i} = (\rho e) v_{1+i}, \qquad 
+ E = (\rho e)\LRp{1 - \frac{\sum_{j=1}^d{v_{1+j}^2}}{2 v_{d+2}}},
+\end{align*}
+where $i = 1,\ldots,d$, and $\rho e$ and $s$ in terms of the entropy variables are 
+\begin{equation*}
+\rho e = \LRp{\frac{(\gamma-1)}{\LRp{-v_{d+2}}^{\gamma}}}^{1/(\gamma-1)}e^{\frac{-s}{\gamma-1}}, \qquad 
+s = \gamma - v_1 + \frac{\sum_{j=1}^d{v_{1+j}^2}}{2v_{d+2}}.
+\end{equation*}
+
+Entropy conservative fluxes involve both standard and logarithmic averages.
+Let $f$ denote some piecewise continuous function, and let $f^+$ denote the exterior value of $f$ across an element face.
+The average and logarithmic averages are
+\[
+{\avg{f} = \frac{f^+ + f}{2}}, \qquad {\avg{f}^{\log} = \frac{f^+ - f}{\log\LRp{f^+}-\log\LRp{f}}}.
+\]
+{The average and logarithmic average are assumed to act component-wise on vector-valued functions.
+We evaluate the logarithmic average using the numerically stable algorithm of \cite{winters2019entropy}.}
+Explicit expressions for entropy conservative numerical fluxes in two dimensions  are given by Chandrashekar \cite{chandrashekar2013kinetic}
+\begin{align*}
+&f^1_{1,S}(\bm{u}_L,\bm{u}_R) = \avg{\rho}^{\log} \avg{u_1},& &f^1_{2,S}(\bm{u}_L,\bm{u}_R) = \avg{\rho}^{\log} \avg{u_2},&\\
+&f^2_{1,S}(\bm{u}_L,\bm{u}_R) = f^1_{1,S} \avg{u_1} + p_{\rm avg},&  &f^2_{2,S}(\bm{u}_L,\bm{u}_R) = f^1_{2,S} \avg{u_1},&\nonumber\\
+&f^3_{1,S}(\bm{u}_L,\bm{u}_R) = f^2_{2,S},& &f^3_{2,S}(\bm{u}_L,\bm{u}_R) = f^1_{2,S} \avg{u_2} + p_{\rm avg},&\nonumber\\
+&f^4_{1,S}(\bm{u}_L,\bm{u}_R) = \LRp{E_{\rm avg} + p_{\rm avg}}\avg{u_1},& &f^4_{2,S}(\bm{u}_L,\bm{u}_R) = \LRp{E_{\rm avg} + p_{\rm avg} }\avg{u_2},& \nonumber
+\end{align*}
+where we have introduced the auxiliary quantities $\beta = \frac{\rho}{2p}$ and  
+\begin{gather}
+p_{\rm avg} = \frac{\avg{\rho}}{2\avg{\beta}}, \qquad E_{\rm avg} = \frac{\avg{\rho}^{\log}}{2\avg{\beta}^{\log}\LRp{\gamma -1}}   + \frac{{u_{\rm avg}^2}}{2}, \qquad 
+{u^2_{\rm avg}} = \sum_{j=1}^d u_j u_j^+ \label{eq:fluxaux}.
+\end{gather}
+where $u_j, u_j^+$ denote the trace and exterior values of $u_j$ across a face.  
+Expressions for entropy conservative numerical fluxes for the three-dimensional compressible Euler equations can also be explicitly written as
+\begin{gather*}
+\bm{f}_{1,S} = \LRp{\begin{array}{c}
+\avg{\rho}^{\log}\avg{u_1}\\
+\avg{\rho}^{\log}\avg{u_1}^2 + p_{\rm avg}\\
+\avg{\rho}^{\log}\avg{u_1}\avg{u_2}\\
+\avg{\rho}^{\log}\avg{u_1}\avg{u_3}\\
+(E_{\rm avg}+ p_{\rm avg})\avg{u_1}\\
+\end{array}}, 
+\qquad 
+\bm{f}_{2,S} = \LRp{\begin{array}{c}
+\avg{\rho}^{\log}\avg{u_2}\\
+\avg{\rho}^{\log}\avg{u_1}\avg{u_2}\\
+\avg{\rho}^{\log}\avg{u_2}^2 + p_{\rm avg}\\
+\avg{\rho}^{\log}\avg{u_2}\avg{u_3}\\
+(E_{\rm avg}+ p_{\rm avg})\avg{u_2}\\
+\end{array}},\\
+\bm{f}_{3,S} = \LRp{\begin{array}{c}
+\avg{\rho}^{\log}\avg{u_3}\\
+\avg{\rho}^{\log}\avg{u_1}\avg{u_3}\\
+\avg{\rho}^{\log}\avg{u_2}\avg{u_3}\\
+\avg{\rho}^{\log}\avg{u_3}^2 + p_{\rm avg}\\
+(E_{\rm avg}+ p_{\rm avg})\avg{u_3}\\
+\end{array}}. \nonumber
+\end{gather*}
+%where the auxiliary quantities are defined as
+%\begin{gather*}
+%{\beta = \frac{\rho}{2p}}, \qquad p_{\rm avg} = \frac{\avg{\rho}}{2\avg{\beta}}, \qquad E_{\rm avg} = \frac{\avg{\rho}^{\log}}{2(\gamma-1)\avg{\beta}^{\log}} + \frac{1}{2}\avg{\rho}^{\log}{{u}^2_{\rm avg}}.
+%\end{gather*}
+
+For all numerical experiments, we use an explicit low storage RK-45 time-stepper \cite{carpenter1994fourth} and estimate the timestep $dt$ using $J, J^k_f$, and degree-dependent $L^2$ trace constants $C_N$
+\[
+dt = C_{\rm CFL}\frac{ h}{ a C_N}, \qquad h = \frac{1}{{\nor{J^{-1}}_{L^{\infty}}}\nor{J^k_f}_{L^{\infty}}},
+\]
+where $a$ is an estimate of the maximum wave speed, $h$ estimates the mesh size, and $C_{\rm CFL}$ is some user-defined CFL constant.
+For isotropic elements, the ratio of $J^k$ to $J^k_f$ scales as the mesh size $h$, while $C_N$ captures the dependence of the maximum stable timestep on the polynomial degree $N$.
+For hexahedral elements, $C_N$ varies depending on the choice of quadrature.
+It was shown in \cite{chan2015gpu} that
+\[
+C_N =\begin{cases}
+ d\frac{N(N+1)}{2} & \text{for GLL nodes}\\
+d\frac{(N+1)(N+2)}{2} & \text{for Gauss nodes}
+\end{cases}.
+\]
+Thus, based on this conservative estimate of the maximum stable timestep, GLL collocation schemes should be able to take a timestep which is roughly $(1 + 2/N)$ times larger than the maximum stable timestep for Gauss collocation schemes.
+We do not account for this discrepancy in this work, and instead set the timestep for both GLL and Gauss collocation schemes based on the more conservative Gauss collocation estimate of $dt$.
+All results are computed using $C_{\rm CFL} = 1/2$ and Lax-Friedrichs interface dissipation.
+
+Suppose $\bm{u}_{\Omega}$ denotes the global solution vector; then, an entropy conservative DG scheme can be rewritten as
+\[
+\bm{M}_{\Omega}\td{\bm{u}_{\Omega}}{t} = \bm{F}(\bm{u}_{\Omega}).
+\]
+where $\bm{M}_{\Omega}$ is the global mass matrix, and $\bm{F}(\bm{u}_{\Omega})$ denotes the global right hand side.
+Let $\bm{v}_{\Omega}$ denote the global vector of entropy variables.
+The proof of entropy conservation implies that the ``spatial entropy'' $\bm{v}_{\Omega}^T\bm{F}(\bm{u}_{\Omega}) = 0$ \cite{crean2018entropy}.
+We verify that both 2D and 3D implementations are entropy conservative (and thus entropy stable if interface dissipation is added) by computing the spatial entropy for a discontinuous initial condition.
+In the absence of interface dissipation, all schemes yield a spatial entropy of magnitude $O\LRp{10^{-14}}$, which suggests that the implementations are semi-discretely entropy conservative up to machine precision.
+
+\subsection{Two-dimensional results}
+
+We begin by examining high order convergence of the proposed methods in two dimensions using the isentropic vortex problem \cite{shu1998essentially, crean2017high}.  The analytical solution is 
+\begin{align}
+\rho(\bm{x},t) &= \LRp{1 - \frac{\frac{1}{2}(\gamma-1)(\beta e^{1-r(\bm{x},t)^2})^2}{8\gamma \pi^2}}^{\frac{1}{\gamma-1}}, \qquad p = \rho^{\gamma},\\
+u_1(\bm{x},t) &= 1 - \frac{\beta}{2\pi} e^{1-r(\bm{x},t)^2}(y-y_0), \qquad u_2(\bm{x},t) = \frac{\beta}{2\pi} e^{1-r(\bm{x},t)^2}(x-x_0-t),\nonumber
+\end{align}
+where $u_1, u_2$ are the $x$ and $y$ velocity, $r(\bm{x},t) = \sqrt{(x-x_0-t)^2 + (y-y_0)^2}$$x_0 = 5, y_0 = 0$, and $\beta = 5$.
+We solve on a periodic rectangular domain $[0, 15] \times [-5,5]$ until final time $T=5$ and compute total $L^2$ errors over all solution fields.
+
+For a degree $N$ scheme, we approximate the $L^2$ error using an $(N+2)$ point Gauss quadrature rule.
+We also examine the influence of element curvature for both GLL and Gauss collocation schemes by examining $L^2$ errors on affine and curvilinear meshes (see Figure~\ref{fig:mesh2d}).
+These warpings are constructed by modifying nodal positions according to the following mapping
+\begin{align*}
+\tilde{x} &= x + L_x\alpha\cos\LRp{\frac{\pi}{L_x}\LRp{x-\frac{L_x}{2}}}\cos\LRp{\frac{3\pi}{L_y}y}, \\
+\tilde{y} &= y + L_y\alpha\sin\LRp{\frac{4\pi}{L_x}\LRp{\tilde{x}-\frac{L_x}{2}}}\cos\LRp{\frac{\pi}{L_y}y},
+\end{align*}
+where $\alpha = 1/16$, and $L_x = 15, L_y = 10$ denote the lengths of the domain in the $x$ and $y$ directions, respectively.
 
@@
 %\includegraphics[width=.475\textwidth]{figs/mesh_curved0.png}
-%\hspace{.5em}
-\includegraphics[width=.55\textwidth]{figs/noncon_curved_mesh.png}
-\caption{An $N=3$ mesh used in the refinement and entropy conservation studies.}
+\includegraphics[width=.45\textwidth]{figs/noncon_curved_mesh.png}
+\hspace{.5em}
+\includegraphics[width=.45\textwidth]{figs/noncon_curved_mesh2.png}
+\caption{Two curved $N=3$ meshes used in the grid refinement studies.}
+\label{fig:mesh2d}
 \end{figure}
+
+Figure~\ref{fig:2d} shows results for the 2D compressible Euler equations by the PI on non-conforming quadrilateral meshes constructed by refining in a checkerboard pattern (see Figure~\ref{fig:mesh2d}), with volume, surface, and mortar quadratures constructed from one-dimensional Lobatto or Gauss quadrature rules.
+As noted in Section~\ref{sec:compare} and Lemma~\ref{lemma:accuracy}, Gauss quadrature produces a degree $N$ differentiation operator, while Lobatto quadrature produces a degree $(N-1)$ differentiation operator.
+We observe a similar discrepancy in $L^2$ errors for Lobatto and Gauss quadratures on non-conforming meshes in Figure~\ref{fig:2d}.
 
@@
 }
-
-\caption{Blah}
+\caption{$L^2$ errors for Lobatto and Gauss collocation schemes on affine and curved meshes under $h$-refinement. }
+\label{fig:2d}
 \end{figure}
 
-\textbf{Gauss Lobatto quadrature}
-
-\textbf{Gauss quadrature}
-
-\subsection{3D results}
-\textbf{Gauss Lobatto quadrature}
-
-\textbf{Gauss quadrature}
-
-\section{Conclusions and extensions of the mortar-based approach}
+\subsection{Three-dimensional results}
+
+We test the accuracy of the mortar-based scheme in 3D using an extruded 2D vortex propagating in the $y$ direction, with analytic expressions given in \cite{williams2013nodal}
+\begin{align*}
+\rho(\bm{x},t) &= \LRp{1-\frac{(\gamma-1)}{2}\Pi^2}^{\frac{1}{\gamma-1}}\\
+u_i(\bm{x},t) &= \Pi r_i + \delta_{i2}, \\
+E(\bm{x},t) &= \frac{p_0}{\gamma-1}\LRp{1-\frac{\gamma-1}{2}\Pi^2}^{\frac{\gamma}{\gamma-1}} + \frac{\rho}{2}\sum_{i=1}^d u_i^2.
+\end{align*}
+where $(u_1,u_2,u_3)^T$ is the three-dimensional velocity vector and % $\Pi$, $\bm{r}$ are defined as
+\[
+\Pi = \Pi_{\max}e^{\frac{1-\sum_{i=1}^dr_i^2}{2}}, \qquad \begin{pmatrix}r_1\\r_2\\r_3\end{pmatrix} = \begin{pmatrix}
+-(x_2-c_2-t)\\
+x_1-c_1\\
+0
+\end{pmatrix}.
+\]
+We take {$c_1 = c_2 = 7.5$}, $p_0 = {1}/{\gamma}$, and $\Pi_{\max} = 0.4$, and solve on the domain $[0,15]\times [0,20]\times [0,1]$.
 
@@
 \centering
-\includegraphics[width=.5\textwidth]{noncon3D.png}
-\caption{Two-layer mortars for reduced $O(N^3)$ computational cost on non-conforming hexahedral interfaces.}
+\includegraphics[width=.65\textwidth]{figs/noncon_mesh3d.png}
+\caption{Construction of a non-conforming hexahedral mesh by combining two conforming Cartesian meshes. }
+\label{fig:noncon3dmesh}
 \end{figure}
 
-\note{Extensions to 3D meshes: double anisotropic mortar for $O(N^3)$ operations vs $O(N^4)$.}
-
-\note{Extension to $p$ non-conforming meshes is straightforward.  Simply replace the definition of a mortar.}
-
-\note{Extension to general quadratures just requires replacing $\bm{I}$ with $\bm{V}_q, \bm{P}_q$, and redefining $\bm{E}$ matrices.}
-
-\note{Explain why non-hexahedral elements isn't as much of an issue - Shadpey \cite{shadpey2019energy} use this for simplicial elements.
-The interpolation operators are dense, and using the mortar-based approach wouldn't actually decrease the number of operations.}
+We construct a non-conforming mesh by gluing together two uniform hexahedral meshes of different resolutions (see Figure~\ref{fig:noncon3dmesh}).
+The non-conforming interface runs parallel to the $y$-axis, such that the vortex propagates along the direction of the non-conforming interface.
+The solution is evolved until final time $T=1$.
+Figure~\ref{fig:3d_affine} shows computed estimates of $L^2$ and $L^{\infty}$ errors, where we estimate both by evaluating the solution using a tensor product Gauss quadrature rule with $N+3$ points in each coordinate direction.
+For the $L^2$ error, we observe that Gauss quadrature is more accurate than Lobatto quadrature.
+However, Gauss quadrature delivers significantly more accurate $L^\infty$ errors than Lobatto quadrature, which is consistent with numerical experiments in \cite{fernandez2019entropy}.
+
+Finally, we discuss computed $L^2$ convergence rates.
+For $N=1$, we observe convergence rates of $1.365$ and $1.939$ for Lobatto and Gauss, respectively.
+This order reduction is less prominent for $N=3$, with Lobatto and Gauss quadratures achieving rates of $3.468$ and $3.977$, respectively.
+Order reduction is observed for both Gauss and Lobatto nodes for $N=2$, with both choices achieving convergence rates of around $2.3$.
+This may indicate that the meshes have not yet been sufficiently refined to reach the asymptotic regime, or may be a consequence of the use of Lax-Friedrichs penalization \cite{hindenlang2019order}.
+
+\begin{figure}
+\centering
+%\subfloat[Coarse $N=3$ non-conforming affine mesh]{\raisebox{2.6em}{\includegraphics[width=.475\textwidth]{figs/mesh_affine.png}}}
+\subfloat[$L^2$ errors]{
+\begin{tikzpicture}
+\begin{loglogaxis}[
+    width=.49\textwidth,
+    xlabel={Mesh size $h$},
+    ylabel={Error}, 
+    xmax=1.5,
+    ymin=5e-6, ymax=.1,
+    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
+    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
+    legend entries={Lobatto, Gauss}%, Lobatto-Gauss}    
+]
+\pgfplotsset{
+cycle list={{blue, mark=*}, {red, dashed ,mark=square*}}%,{black, dashdotted ,mark=triangle*}}
+}
+% N = 1
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.0302287)(0.5,0.0204278)(0.25,0.00988213)(0.125,0.00383616)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.0312387)(0.5,0.020413)(0.25,0.00690005)(0.125,0.0017994)};
+
+% N = 2
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.025246)(0.5,0.00858607)(0.25,0.00156388)(0.125,0.000334682)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.0236979)(0.5,0.00616892)(0.25,0.000963938)(0.125,0.000249721)};
+
+% N = 3
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.0167185)(0.5,0.00255939)(0.25,0.000325533)(0.125,2.94146e-05)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.0131715)(0.5,0.0019141)(0.25,0.000161528)(0.125,1.02566e-05)};
+
+\node at (axis cs:2.4,2.8) {$N = 1$};
+\node at (axis cs:2.4,1.3) {$N = 2$};
+\node at (axis cs:2.4,.6) {$N = 3$};
+\end{loglogaxis}
+\end{tikzpicture}
+}
+\hspace{.1em}
+\subfloat[$L^{\infty}$ errors]{
+\begin{tikzpicture}
+\begin{loglogaxis}[
+    width=.49\textwidth,
+    xlabel={Mesh size $h$},
+%    ylabel={$L^2$ errors}, 
+    xmax=1.5,
+    ymin=1e-4, ymax=.5,
+    legend pos=south east, legend cell align=left, legend style={font=\tiny},	
+    xmajorgrids=true, ymajorgrids=true, grid style=dashed,
+    legend entries={Lobatto, Gauss}%, Lobatto-Gauss}    
+]
+\pgfplotsset{
+cycle list={{blue, mark=*}, {red, dashed ,mark=square*}}%,{black, dashdotted ,mark=triangle*}}
+}
+% N = 1
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.225344)(0.5,0.165699)(0.25,0.0820851)(0.125,0.0494277)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.222488)(0.5,0.144629)(0.25,0.073849)(0.125,0.0217714)};
+
+% N = 2
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.2011)(0.5,0.0658348)(0.25,0.0233793)(0.125,0.00727696)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.159118)(0.5,0.0478442)(0.25,0.0177047)(0.125,0.00421538)};
+
+% N = 3
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.113556)(0.5,0.0428492)(0.25,0.00686179)(0.125,0.00146676)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.0851151)(0.5,0.0386236)(0.25,0.00246054)(0.125,0.00027861)};
+
+\node at (axis cs:2.4,3.25) {$N = 1$};
+\node at (axis cs:2.4,1.75) {$N = 2$};
+\node at (axis cs:2.4,.9) {$N = 3$};
+\end{loglogaxis}
+\end{tikzpicture}
+}
+\caption{$L^2$ and $L^\infty$ errors for entropy stable Lobatto and Gauss collocation schemes on non-conforming Cartesian (affine) hexahedral meshes under $h$-refinement for $N=1,2,3$.
+Note the different axis limits.}
+\label{fig:3d_affine}
+\end{figure}
+
+%As in the 2D case, we also examine the effect of curvilinear mesh warping.
+We construct a curved warping of the initial Cartesian mesh by mapping nodes on each hexahedron to warped nodal positions $(\tilde{x},\tilde{y},\tilde{z})$ through the transformation
+%\begin{align*}
+%\tilde{y} &= y + {\frac{1}{8}L_y\cos\LRp{3\pi \frac{(x-7.5)}{15}}\cos\LRp{\pi \frac{(y-10)}{20}}\cos\LRp{\pi \frac{(z-2.5)}{5}}}\\
+%\tilde{x} &= x + {\frac{1}{8}L_x\cos\LRp{\pi \frac{(x-7.5)}{15}}\sin\LRp{4\pi \frac{(\tilde{y}-10)}{20}}\cos\LRp{\pi \frac{(z-2.5)}{5}}}\\
+%\tilde{z} &= z + {\frac{1}{8}L_z\cos\LRp{\pi \frac{(\tilde{x}-7.5)}{15}}\cos\LRp{2\pi \frac{(\tilde{y}-10)}{20}}\cos\LRp{\pi \frac{(z-2.5)}{5}}}.
+%\end{align*} 
+%where $L_x = 15, L_y = 20, L_z = 5$.
+
+\note{TODO: add curved results and results using two-layer mortars.}
+
+\section{Conclusions and extensions of the mortar-based approach}
+
+We have introduced an entropy stable mortar-based interface treatment which reduces the computational cost of high order entropy stable Lobatto and Gauss collocation DG schemes on non-conforming meshes.
+Mortars are incorporated into entropy stable DG schemes using mortar-based extensions of hybridized SBP operators introduced in \cite{chan2018discretely, chan2018efficient, chan2019skew}, and admit efficient face-local implementations.
+
+We conclude by discussing some extensions and generalizations of this mortar-based approach.
+First, the extension to $p$ and $hp$ non-conforming meshes is straightforward.
+As with $h$ non-conforming meshes, the mortar ndoes are taken to be the nodes on the refined side of a $hp$ non-conforming interfaces.
+The two-layer mortar approach of Section~\ref{sec:anisotropic} can also be extended to $hp$-nonconforming interfaces in a straightforward fashion.
+
+This approach can also be generalized to non-hexahedral elements.
+However, the mortar-based SBP operators in this paper require node alignment to expose more sparsity.
+Unlike tensor product elements, simplicial elements, neither mortar nor surface nodes are ``aligned'' with volume nodes.
+Thus, the mortar-based approach in this work is not expected to be more efficient than the direct use of hybridized SBP operators on non-tensor product elements \cite{shadpey2019energy}.
+
+\section{Acknowledgments}
 
 Jesse Chan gratefully acknowledges support from the National Science Foundation under awards DMS-1719818 and DMS-1712639.
+Mario Bencomo is supported partially by the Rice University CAAM Department Pfieffer Postdoctoral Fellowship.
 
@@
 
-\section{Geometric terms for curved hexahedral meshes}
+\section{Geometric terms for curved non-conforming meshes}
 \label{app:A}
@@
 Entropy conservation (stability) requires that the geometric terms be constructed appropriately.
-In addition to requiring satisfaction of the discrete geometric conservation law \cite{kopriva2006metric, crean2018entropy, chan2018discretely}, the schemes derived in this work require that the polynomial degree of the geometric terms be related to the accuracy of the surface and mortar quadrature rules.
-We consider two methods to construct geometric terms: Approach 1 from \cite{kopriva2006metric}, and Approach 2 from \cite{kozdon2018energy}.
-
-Entropy conservation (or stability) on non-conforming meshes also require that the scaled outward normals are equal and opposite across each mortar node.
+In addition to requiring satisfaction of the discrete geometric conservation law (GCL) \cite{kopriva2006metric, crean2018entropy, chan2018discretely}, the schemes derived in this work require that the polynomial degree of the geometric terms be related to the accuracy of the surface and mortar quadrature rules.
+
+As discussed in Section~\ref{sec:compare}, the standard construction of geometric terms (using a cross product formula \cite{kopriva2006metric, hesthaven2007nodal}) on 2D isoparametric curved elements are sufficient to ensure the GCL and thus entropy conservation (stability) for Lobatto and Gauss collocation schemes.
+For 3D isoparametric curved hexahedra, the standard construction of geometric terms does not ensure satisfaction of the GCL.
+We consider two alternative methods to construct geometric terms: Approach 1 from \cite{kopriva2006metric}, and Approach 2 from \cite{kozdon2018energy}.
+
+We note that entropy conservation (or stability) on non-conforming meshes also requires that the scaled outward normals are equal and opposite across each mortar face.
 This holds if the mesh is \emph{watertight} or \emph{well-constructed}, such that there are no gaps between neighboring elements.
@@
 For the two-dimensional numerical experiments in this work, we construct watertight curved non-conforming meshes by first curving a conforming mesh.
-We then locally refine certain parent elements and transfer the geometric terms of the parent element to the child elements.
-For the three-dimensional numerical experiments in this work, \note{finish}.
+We locally refine to construct a non-conforming mesh, with the refined child elements inheriting the geometric mapping of the parent elements.
+For the three-dimensional numerical experiments in this work, we adopt the following approach.
 
@@
 Let $\bm{x},\bm{y},\bm{z}$ denote the vectors of physical nodal coordinates of a curved degree $N_{\rm geo}$ element.
-Approach 1 in \cite{kopriva2006metric} constructs scaled geometric terms $G_{ij} = J\pd{\hat{x}_j}{x_i}$ via
+Approach 1 in \cite{kopriva2006metric} constructs scaled geometric terms $G_{ij} = J\pd{\hat{x}_j}{x_i}$ by computing the curl of geometric ``potentials'' $\bm{f}_{ij}$
 \begin{align*}
@@
 \end{align*}
-where $\alpha_1 = 1$ and $\alpha_2 = \alpha_3 = -1$, and the geometric ``potentials'' $\bm{f}_{ij}$ are defined as 
+where $\alpha_1 = 1$ and $\alpha_2 = \alpha_3 = -1$, and the potentials $\bm{f}_{ij}$ are defined as 
 \begin{align}
@@
 The resulting geometric terms are tensor product polynomials of degree $G_{ij} \in Q^{N_{\rm geo}}$.
-If $N_{\rm geo} < N$, we use polynomial interpolation to compute values of the geometric terms at degree $N$ Lobatto nodes.
-
-Approach 2 in \cite{kozdon2018energy} (see also \cite{chan2019skew}, Footnote 3) modifies this approach slightly by reducing the degree of each geometric potentials along a coordinate direction.
+We note that any geometric terms constructed by applying the discrete curl operator automatically satisfy the geometric conservation law.
+
+On non-conforming meshes, the resulting construction does not guarantee that the scaled normal vectors (\ref{eq:nJ_Gnhat}) are equal and opposite across faces \cite{kopriva2019free}.
+This is remedied by enforcing continuity of the geometric potentials $\bm{f}_{ij}$ across all non-conforming interfaces (for example, by interpolating values of $\bm{f}_{ij}$ from the coarse side to the refined side).
+
+Approach 2 in \cite{kozdon2018energy} (see also \cite{chan2019skew}, Footnote 3) further modifies this approach by slightly by reducing the degree of each geometric potential along the corresponding coordinate direction.
 Let $\bm{I}_{N_{\rm geo}}^{N_{\rm geo}-1}$ denote the interpolation operator from degree $N_{\rm geo}$ to degree $N_{\rm geo}-1$ Lobatto nodes, and let $\bm{I}_{N_{\rm geo}-1}^{N_{\rm geo}}$ denote the interpolation operator from degree $N_{\rm geo}-1$ to degree $N_{\rm geo}$ Lobatto nodes.
@@
 \end{align*} 
-
-While Approach 1 is nominally more accurate than Approach 2, both approaches achieve nearly identical accuracy and convergence rates.
+The difference between Approach 1 and Approach 2 appears in Lemma~\ref{lemma:Qmprops_3d}.
+For Lobatto nodes, the mortar-based operators $\bm{Q}_{i,m}$ satisfy an SBP property (and can be used to construct entropy stable formulations) for $N_{\rm geo} < N$ under Approach 1, while Approach 2 relaxes this condition to $N_{\rm geo} \leq N$.
+
+While Approach 1 is marginally more accurate than Approach 2 for $N > 1$, both approaches achieve nearly identical accuracy and convergence rates.
 We analyze $L^2$ errors using Approach 1 and Approach 2 by  computing exact values of $G_{ij}$ using the cross product formula \cite{kopriva2006metric, hesthaven2007nodal}.
@@
 Figure~\ref{fig:geo_err} shows the convergence of the $L^2$ error over all geometric terms under mesh refinement.
+Both Approach 1 and Approach 2 achieve very similar errors for $N=1,\ldots, 4$, with Approach 1 achieving slightly lower errors for all $N > 1$.
 Table~\ref{fig:geo_rates} shows computed convergence rates, which match the $O(h^{N+2})$ rates proven in \cite{chan2018discretely, crean2018entropy} for simplicial elements.
@@
 \begin{loglogaxis}[
-    width=.6\textwidth,
+    width=.5\textwidth,
     xlabel={Mesh size $h$},
     ylabel={$L^2$ error}, 
-    xmax=1.25,xmin=.0625,
+    xmax=1.7,xmin=.055,
     ymin=2.5e-8, ymax=2.5,
@@
 
-\node at (axis cs:.09,3.5e-3) {$N = 1$};
-\node at (axis cs:.09,2.2e-4) {$N = 2$};
-\node at (axis cs:.09,6e-6) {$N = 3$};
-\node at (axis cs:.09,1.25e-7) {$N = 4$};
+\node at (axis cs:.085,3.5e-3) {$N = 1$};
+\node at (axis cs:.085,2.2e-4) {$N = 2$};
+\node at (axis cs:.085,6e-6) {$N = 3$};
+\node at (axis cs:.085,1.25e-7) {$N = 4$};
 \end{loglogaxis}
 \end{tikzpicture}
-\caption{Blah}
+\caption{$L^2$ errors under mesh refinement for Approach 1 \cite{kopriva2006metric} and Approach 2 \cite{kozdon2018energy}.}
 \label{fig:geo_err}

### 2020-02-27 esdg_quadhex_cpp/main2d.cpp

@@ -198,8 +198,8 @@ int main(int argc, char **argv){
   App *app = new App;
   //app->device.setup("mode: 'Serial'");
   app->device.setup("mode: 'CUDA', device_id: 0");
-  //  app->props["compiler_flags"] = "-ftz=false -prec-div=true -prec-sqrt=true -arch=sm_30";
-  app->props["compiler_flags"] = "-arch=sm_30";
+  // app->props["compiler_flags"] = "-ftz=false -prec-div=true -prec-sqrt=true -arch=sm_30";
+  // app->props["compiler_flags"] = "-arch=sm_30";
   
   setupOccaMesh2d(mesh,app); // build mesh geofacs
 
@@ -232,9 +232,9 @@ int main(int argc, char **argv){
   string path = "okl/Euler2D.okl";
 
   //testing
-  app->volume = app->device.buildKernel(path.c_str(),"volume",app->props);
+  app->volume  = app->device.buildKernel(path.c_str(),"volume",app->props);  
   app->surface = app->device.buildKernel(path.c_str(),"surface",app->props);
-  app->update = app->device.buildKernel(path.c_str(),"update",app->props);
+  app->update  = app->device.buildKernel(path.c_str(),"update",app->props);
   app->eval_surface = app->device.buildKernel(path.c_str(),"eval_surface",app->props);
 
   // set initial condition

### 2020-02-27 matlab/Wave1D.m

@@ -1,7 +1,5 @@
 % wave
 
-function [l2err]= Wave1D(smax,delta,p0sigma,p0approx)
-
 % Driver script for solving the 1D advection equations
 Globals1D;
 
@@ -9,22 +7,23 @@ Globals1D;
 N = 5;
 
 % Generate simple mesh
-a = .1;
-K1D = 2*a/.0025;
-[Nv, VX, K, EToV] = MeshGen1D(-a,a,K1D);
+K1D = 32;
+
+CFL = .75; 
+
+[Nv, VX, K, EToV] = MeshGen1D(-1,1,K1D);
 
 % Initialize solver and construct grid and metric
 StartUp1D;
-
 
 rp = linspace(-1,1,100);
 Vp = Vandermonde1D(N,rp)/V;
 xp = Vp*x;
 
 % Set initial conditions
-% p = cos(pi/2*x);
+p = cos(pi*x);
 % p = exp(-100*x.^2);
-p = zeros(size(x));
+% p = zeros(size(x));
 u = zeros(size(x));
 
 % Solve Problem
@@ -38,8 +37,7 @@ resu = zeros(Np,K);
 
 % compute time step size
 xmin = min(abs(x(1,:)-x(2,:)));
-dt  = .5*xmin;
-% dt = min(dt,1/smax);
+dt  = CFL*xmin;
 Nsteps = ceil(FinalTime/dt);
 dt = FinalTime/Nsteps;
 
@@ -53,12 +51,11 @@ for tstep=1:Nsteps
         resu = rk4a(INTRK)*resu + dt*rhsu;
         p = p + rk4b(INTRK)*resp;
         u = u + rk4b(INTRK)*resu;
-
     end;
     % Increment time
     time = time+dt;
     
-    if 1 && mod(tstep,10)==0
+    if 1 && mod(tstep,25)==0
         clf
         pp = Vp*p;
         plot(xp,pp,'-');
@@ -69,6 +66,7 @@ for tstep=1:Nsteps
         drawnow
     end
 end;
+
 
 
 function [rhsp rhsu] = WaveRHS1D(p,u,t)
@@ -86,23 +84,14 @@ du(:) = u(vmapP)-u(vmapM);
 
 dp(mapB) = 0;
 du(mapB) = -2*u(vmapB);
-% if (t < .5)
[…]

### 2020-02-27 matlab/Wave2D_RK.m

@@ -7,8 +7,7 @@ Globals2D
 
 if nargin==0
     %     Nin = 4; K1D = 16;
-    Nin = 3; K1D = 2; c_flag = 0;
-    FinalTime = .5;
+    Nin = 4; K1D = 16; c_flag = 0;
     cfun = @(x,y) ones(size(x));
 %     cfun = @(x,y) 1 + .5*sin(pi*x).*sin(pi*y); % smooth velocity
 %     cfun = @(x,y) (1 + .5*sin(2*pi*x).*sin(2*pi*y) + (y > 0)); % piecewise smooth velocity
@@ -16,7 +15,8 @@ if nargin==0
 %     cfun = @(x,y) 2 + sin(2*pi*sin(x));
 end
 N = Nin;
-
+FinalTime = .5;
+    
 % filename = 'Grid/Other/block2.neu';
 % filename = 'Grid/Maxwell2D/Maxwell05.neu';
 % [Nv, VX, VY, K, EToV] = MeshReaderGambit2D(filename);
@@ -95,7 +95,7 @@ end
 
 
 %% check eigenvalues of DG matrix 
-if 1 && nargin==0
+if 0 && nargin==0
     e = zeros(3*Np*K,1);
     A = zeros(3*Np*K);
     for i = 1:3*Np*K
@@ -134,7 +134,7 @@ resu = zeros(Np,K); resv = zeros(Np,K); resp = zeros(Np,K);
 % compute time step size
 CN = (N+1)^2/2; % guessing...
 %dt = 1/(CN*max(abs(sJ(:)))*max(abs(1./J(:))));
-dt = 1/(max(cq(:))*CN*max(Fscale(:)));
+dt = 2/(max(cq(:))*CN*max(Fscale(:)));
 
 % outer time step loop
 tstep = 0;

### 2020-02-27 matlab/cfd2D/Euler2D_refactored.m

@@ -5,7 +5,7 @@ FinalTime = 1;
 
 N = 6;
 K1D = 10;
-CFL = .75;
+CFL = .5;
 plotMesh = 0;
 
 global tau
@@ -13,13 +13,9 @@ tau = .5;
 
 %% set up physical domain 
 
-% Lx = 7.5; Ly = 5; ratiox = 3/4; ratioy = .5;
 Lx = 10; Ly = 5; ratiox = 1; ratioy = Ly/Lx;
 [Nv, VX, VY, K, EToV] = unif_tri_mesh(round(ratiox*K1D),round(K1D*ratioy));
-% [Nv, VX, VY, K, EToV] = unif_tri_mesh(2,1);
-% VX = VX/max(abs(VX));  VY = VY/max(abs(VY));
 VX = (VX+1)*Lx; VY = VY*Ly;
-% [Nv, VX, VY, K, EToV] = unif_tri_mesh(2);
 
 StartUp2D;
 BuildPeriodicMaps2D(max(VX)-min(VX),max(VY)-min(VY));
@@ -83,6 +79,12 @@ WN = diag([wq;wf]);
 DNr = WN\QNr;
 DNs = WN\QNs;
 
+%% apply curved warping
+
+a = 0/8;
+x0 = Lx; y0 = 0;
+x = x + Lx*a*cos(1/2*pi*(x-x0)/Lx).*cos(3/2*pi*(y-y0)/Ly);
+y = y + Ly*a*sin(pi*(x-x0)/Lx).*cos(1/2*pi*(y-y0)/Ly);
 
 %% make geometric terms
 
@@ -94,13 +96,9 @@ xp = Vp*x; yp = Vp*y;
 rxJ = rx.*J;    sxJ = sx.*J;
 ryJ = ry.*J;    syJ = sy.*J;
 
-[rx,sx,ry,sy,Jf] = GeometricFactors2D(x,y,Vf*Dr,Vf*Ds);
-rxJf = rx.*Jf;    sxJf = sx.*Jf;
-ryJf = ry.*Jf;    syJf = sy.*Jf;
-
 % compute scaled physical normals
-nxJ = rxJf.*nrJ + sxJf.*nsJ;
-nyJ = ryJf.*nrJ + syJf.*nsJ;
+nxJ = (Vf*rxJ).*nrJ + (Vf*sxJ).*nsJ;
+nyJ = (Vf*ryJ).*nrJ + (Vf*syJ).*nsJ;
[…]

### 2020-02-27 matlab/cfd3D/check_gcl_convergence.m

@@ -1,6 +1,6 @@
 clear
 
-N = 3;
+N = 4;
 [r1D w1D] = JacobiGL(0,0,N);
 [r s t] = meshgrid(r1D);
 r = r(:); s = s(:); t = t(:);
@@ -27,148 +27,156 @@ Vq = kron(kron(Vq1D,Vq1D),Vq1D);
 r2 = r2(:); s2 = s2(:);
 e = ones(size(r2));
 
-rf = [-e; e; r2; r2; r2; r2];
-sf = [r2; r2; -e; e; s2; s2];
-tf = [s2; s2; s2; s2; -e; e];
-
-Vf = Vandermonde3DHex(N,rf,sf,tf)/Vandermonde3DHex(N,r,s,t);
-NODETOL = 1e-8;
-
-Kvec = 2.^(2:7);
-sk = 1;
-for K1D = Kvec 
-    h = 2/K1D;
-    x = h*r;
-    y = h*s;
-    z = h*t;
+if 0
+    [x,y,~] = meshgrid(linspace(-1,1,10));
     
-%     a = .125;
-%     dx = cos(pi/2*x).*sin(pi*y).*sin(pi*z);
-%     x = x + a*dx;
-%     dy = sin(pi*x).*cos(pi/2*y).*sin(pi*z);
-%     y = y + a*dy;
-%     dz = sin(pi*x).*sin(pi*y).*cos(pi/2*z);
-%     z = z + a*dz;
-
-a = 0.25;
-% d = (1+x).*(1-x).*(1-y).*(1+y).*(1+z).*(1-z);
-    dx = (1-y).*(1+y).*(1+z).*(1-z);
-    dy = (1+x).*(1-x).*(1+z).*(1-z);
-    dz = (1+x).*(1-x).*(1-y).*(1+y);
-% dx = d;     dy = d;     dz = d;
-x = x + a*dx;
-y = y + a*dy;
-z = z + a*dz;
+    a = .25;
+    dx = cos(x).*sin(y).*sin(z);
+    dy = sin(x).*cos(y).*sin(z);
+    dz = sin(x).*sin(y).*cos(z);
+    x = x + a*dx;
+    y = y + a*dy;
+    z = z + a*dz;
     
-    % a = .1;
-    % dx = exp(x+y+z);
-    % dy = exp(x+y+z);
-    % dz = exp(x+y+z);
-    % x = x + a*dx;
-    % y = y + a*dy;
-    % z = z + a*dz;
+    surf(x,y,z)
[…]

### 2020-02-27 matlab/cfd3D/hex_noncon.m

@@ -1,8 +1,8 @@
 %% set up reference hex
 clear
-N = 1;
+N = 7;
 CFL = .25;
-FinalTime = .33;
+FinalTime = .01;
 
 % interpolation nodes for geometric factors
 [r1D w1D] = JacobiGL(0,0,N);
@@ -13,7 +13,7 @@ w = wr(:).*ws(:).*wt(:);
 
 % quadrature nodes for collocation
 [rq1D wq1D] = JacobiGQ(0,0,N);
-% rq1D = rq1D*(1-1e-10);
+% [rq1D wq1D] = JacobiGL(0,0,N); rq1D = rq1D*(1-1e-11);
 [rq sq tq] = meshgrid(rq1D);
 rq = rq(:); sq = sq(:); tq = tq(:);
 [wr ws wt] = meshgrid(wq1D);
@@ -34,10 +34,11 @@ E1D = Vf1D;
 B1D = diag([-1,1]);
 QN = [Q1D-Q1D' E1D'*B1D;
     -B1D*E1D 0*B1D];
-VN = [eye(N+1);Vf1D];
+VN = [eye(N+1); Vf1D];
 invWVNT = diag(1./wq1D)*VN';
 
-% 3D matrices
+%% 3D matrices
+
 V = kron(kron(V1D,V1D),V1D);
 Vq = kron(kron(Vq1D,Vq1D),Vq1D);
 M = Vq'*diag(wq)*Vq;
@@ -74,6 +75,7 @@ ntJ = ntJ(:);
 
 %% compute face indices for face points on "lines" of nodes
 
+tol = 1e-12;
 fidr = zeros(2,(N+1)^2);
 fids = zeros(2,(N+1)^2);
 fidt = zeros(2,(N+1)^2);
@@ -89,22 +91,22 @@ for ii = 1:(N+1)^2
     [sf1 sf2] = meshgrid(sf,sfr);
     [tf1 tf2] = meshgrid(tf,tfr);
     D = abs(rf1 - rf2) + abs(sf1 - sf2) + abs(tf1 - tf2);
-    fidr(1,ii) = find(D(1,:) < 1e-10);
-    fidr(2,ii) = find(D(2,:) < 1e-10);
+    fidr(1,ii) = find(D(1,:) < tol);
+    fidr(2,ii) = find(D(2,:) < tol);
     
     [rf1 rf2] = meshgrid(rf,rfs);
[…]

### 2020-02-27 matlab/cfd_rom/fv_euler2d_periodic_fom.m

@@ -2,6 +2,7 @@ clear
 K = 200;
 %FinalTime = .35;
 FinalTime = 2.5;
+FinalTime = 3;
 CFL = 1;
 
 xv = linspace(-1,1,K+1)';

### 2020-02-27 matlab/corner.m

@@ -1,7 +1,7 @@
 function An = corner(A,n)
 
 if nargin==1
-    n = 10;
+    n = 5;
 end
 
 An = A(1:n,1:n);

### 2020-02-27 matlab/optPML1D.m

@@ -1,16 +1,0 @@
-clear
-clc
-smax = linspace(.1,200,10);
-for delta = .125;
-    for p0sigma = 0:1
-        for p0approx = 0%:p0sigma
-            for i = 1:length(smax)
-                l2err(i) = min(Wave1D(smax(i),delta,p0sigma,p0approx),1e1);
-                i
-            end
-            l2err
-            semilogy(smax,l2err,'o-','DisplayName',sprintf('delta = %f, p0sigma = %d, p0approx = %d',delta,p0sigma,p0approx));hold on            
-        end
-    end
-end
-legend show

### 2020-02-27 matlab/unif_tri_mesh.m

@@ -16,8 +16,8 @@ for ey = 1:Ky
         id2 = id(ex+1,ey);
         id3 = id(ex+1,ey+1);
         id4 = id(ex,ey+1);
-        VXe = VX([id1 id2 id3 id4]);
-        VYe = VY([id1 id2 id3 id4]);
+%         VXe = VX([id1 id2 id3 id4]);
+%         VYe = VY([id1 id2 id3 id4]);
         EToV(2*sk-1,:) = [id1 id2 id3];
         EToV(2*sk,:) = [id3 id4 id1];
         sk = sk + 1;                               

### 2020-02-27 matlab/wedge/smoothed_wedge_layers.m

@@ -5,11 +5,11 @@ if nargin==0
     filename = 'cube';
 end
 if nargin<2
-    surface_mesh = 'Grid/Maxwell2D/Maxwell2.neu';
+    surface_mesh = 'Grid/Maxwell2D/Maxwell0125.neu';
 end
 if nargin<3
     usurf = @(x,y) zeros(size(x)); % assumes wavy layer is centered around z = 0.
-%     usurf = @(x,y) .25*sin(2*x+2.5).*sin(3*y+3.5);
+    usurf = @(x,y) .25*sin(2*x+2.5).*sin(3*y+3.5);
 end
 if nargin<4
     nlevel = 1;
@@ -163,6 +163,7 @@ end
 % ue = [u;uold]; plot3(xe(ids,:),ye(ids,:),ue(ids,:),'ko-','linewidth',2);hold on
 % trisurf(tri,x,y,u)
 % axis equal
+return
 
 num_surface_verts = length(VX);
 EToV2D = EToV; 

### 2020-02-27 talks/WCCM2020/abstract.tex
adding more talks

--- 2018-06-07 talks/ECCM2018/abstract.tex
@@
 % Type down your paper title
-\title{Entropy stable discontinuous Galerkin methods using arbitrary bases and quadratures}
+\title{Extensions of entropy stable nodal discontinuous Galerkin schemes}
 
@@
 \begin{abstract}
-High order discontinuous Galerkin (DG) methods offer several advantages in the approximation of solutions of nonlinear conservation laws, such as improved accuracy and low numerical dispersion/dissipation.
+High order discontinuous Galerkin (DG) methods provide improved accuracy and low numerical dispersion/dissipation for simulations of nonlinear conservation laws.
 However, these methods also tend to suffer from instability in practice, requiring filtering, limiting, or artificial dissipation to prevent solution blow up.
-Provably stable DG methods based on summation-by-parts (SBP) operators and flux differencing address this inherent instability by ensuring that the solution satisfies a semi-discrete entropy inequality.
-
-Under specific combinations of basis functions or quadratures, nodal DG methods can be interpreted within an SBP framework.
-In this talk, we discuss how to construct discretely entropy stable discontinuous Galerkin methods for general choices of basis and quadrature using discrete $L^2$ projection operators and ``decoupled'' SBP operators.
-Extensions to curvilinear meshes and implementational aspects on Graphics Processing Units will be also discussed.
+\textit{Entropy stable} nodal DG methods based on summation-by-parts (SBP) operators and flux differencing address this instability by ensuring satisfaction of a semi-discrete entropy inequality.
+In this talk, we will present extensions of entropy stable DG methods to more general nodal points, which provide improved accuracy on warped and non-conforming meshes.
 \end{abstract}

### 2020-02-27 talks/WCCM2020/wccm2020-abstract.tex
adding more talks

\documentclass[a4paper]{wccm2020-abstract}
\usepackage{parskip}
\usepackage{pslatex}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\setlength{\parskip}{5pt}

\title{Mortar-based high order entropy stable discontinuous Galerkin schemes on non-conforming quadrilateral and hexahedral meshes}

\author{Jesse Chan$^{1}$}

\address{$^{1}$ Rice University (USA), 6100 Main St, Houston, TX, 77005, jesse.chan@rice.edu, jlchan.github.io}

\keywords{High order, discontinuous Galerkin, entropy stability} % Write down at least 3 Keywords

\begin{document}
\thispagestyle{empty}

High order discontinuous Galerkin (DG) methods provide improved accuracy and low numerical dispersion/dissipation for simulations of nonlinear conservation laws.  However, these methods also tend to suffer from instability in practice, requiring filtering, limiting, or artificial dissipation to prevent solution blow up.  \textit{Entropy stable} nodal DG methods based on summation-by-parts (SBP) operators and flux differencing address this instability by ensuring satisfaction of a semi-discrete entropy inequality.  Entropy stable DG methods typically utilize Lobatto quadrature points; however, recent work has extended such methods to more general nodal points \cite{esdg1,esdg3}.  In this talk, we will present entropy stable DG methods based on Gauss points, which can improve accuracy on warped and non-conforming meshes \cite{esdg2}.

\begin{thebibliography}{99}
\bibitem{esdg1} Chan, Jesse. "On discretely entropy conservative and entropy stable discontinuous Galerkin methods." Journal of Computational Physics 362 (2018): 346-374.
\bibitem{esdg2}  Chan, Jesse, David C. Del Rey Fern\'{a}ndez, and Mark H. Carpenter. ``Efficient entropy stable Gauss collocation methods.'' SIAM Journal on Scientific Computing 41.5 (2019): A2938-A2966.
\bibitem{esdg3} Chan, J. Skew-Symmetric Entropy Stable Modal Discontinuous Galerkin Formulations. Journal of Scientific Computing 81, 459–485 (2019) doi:10.1007/s10915-019-01026-w
\end{thebibliography}

\end{document}

### 2020-02-27 talks/UToronto_UH_2019/figs/plotTriQuadPts.m

@@ -1,0 +1,43 @@
+N = 3;
+[rq sq wq] = Cubature2D(2*N);
+
+[rq1D wq1D] = JacobiGQ(0,0,N);
+rf = [rq1D; -rq1D; -ones(size(rq1D))];
+sf = [-ones(size(rq1D)); rq1D; -rq1D];
+
+[rq sq] = rstoxy(rq,sq);
+[rf sf] = rstoxy(rf,sf);
+
+r1 = [-1 1 -1 -1];
+s1 = [-1 -1 1 -1];
+[r1 s1] = rstoxy(r1,s1);
+%plot([-1 1 -1 -1],[-1 -1 1 -1],'k-','linewidth',2)
+plot(r1,s1,'k-','linewidth',2)
+hold on
+plot(rq,sq,'ko','linewidth',2,'markersize',14,'MarkerFaceColor',[.49 1 .63]);
+plot(rf,sf,'rs','linewidth',2,'markersize',14,'MarkerFaceColor',[.49 1 .63]);
+axis off
+% set(h,'markersize',64)
+axis equal
+print(gcf,'-dpng','~/Desktop/bbwadg/talks/NAHOM_2019/figs/triQuadrature.png')
+
+return
+
+load ~/Downloads/QuadratureRules.mat
+
+quadpts = Q_GaussLegendre{N};
+[r1D,w1D] = JacobiGQ(0,0,N);
+% quadpts = Q_GaussLobatto{N};
+% [r1D,w1D] = JacobiGL(0,0,N+1); % 2*(N+1)-1 = 2*N+1
+
+rs = quadpts.Points;
+wq = quadpts.Weights;
+
+rq = rs(:,1);
+sq = rs(:,2);
+
+figure
+plot([-1 1 -1 -1],[-1 -1 1 -1],'k-','linewidth',2)
+hold on
+plot(rq,sq,'ko','linewidth',2,'markersize',14,'MarkerFaceColor',[.49 1 .63]);
+axis off

### 2020-02-28 docs/ESDG_jacobian/main.tex
new mortar figs and Jacobian draft

\documentclass{article}
\usepackage[utf8]{inputenc}

\usepackage{verbatimbox}

\usepackage{amsmath, amsfonts, amssymb, amsthm}
\usepackage{hyperref}
\hypersetup{colorlinks,
citecolor=blue
}
\usepackage{MnSymbol}% http://ctan.org/pkg/mnsymbol

% ===================================== 
\usepackage{stackengine}
\stackMath
\newlength\matfield
\newlength\tmplength
\def\matscale{1.}
\newcommand\dimbox[3]{%
  \setlength\matfield{\matscale\baselineskip}%
  \setbox0=\hbox{\vphantom{X}\smash{#3}}%
  \setlength{\tmplength}{#1\matfield-\ht0-\dp0}%
  \fboxrule=1pt\fboxsep=-\fboxrule\relax%
  \fbox{\makebox[#2\matfield]{\addstackgap[.5\tmplength]{\box0}}}%
}
\newcommand\raiserows[2]{%
   \setlength\matfield{\matscale\baselineskip}%
   \raisebox{#1\matfield}{#2}%
}
\newcommand\matbox[5]{
  \stackunder{\dimbox{#1}{#2}{$#5$}}{\scriptstyle(#3\times #4)}%
}
% ===================================== 
\usepackage{graphicx, color, bm}
\usepackage{subfig}
\usepackage{tikz,pgfplots,pgfplotstable}
%\usepackage{fullpage}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newenvironment{myproof}[1][\proofname]{%
  \proof Proof of #1%
}{\endproof}

\renewcommand{\hat}{\widehat}
\renewcommand{\tilde}{\widetilde}

%% d in integrand
\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}}
\newcommand{\diag}[1]{{\rm diag}\LRp{#1}}
\newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}{\rm #2}}}
\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRb}[1]{\left| #1 \right|}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}
\newcommand{\LRceil}[1]{\left\lceil #1 \right\rceil}
\newcommand{\LRl}[1]{\left. #1 \right|}
\newcommand{\pdn}[3]{\frac{\partial^{#3}#1}{\partial#2^{#3}}}
\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\jump}[1] {\ensuremath{\llbracket#1\rrbracket}}
\newcommand{\avg}[1] {\ensuremath{\LRc{\!\LRc{#1}\!}}}

\newcommand{\note}[1]{{\color{blue}{#1}}}
\newcommand{\bnote}[1]{{\color{blue}{#1}}}
\newcommand{\rnote}[1]{{\color{red}{#1}}}

\newcommand{\eq}[1]{\begin{align*}#1\end{align*}}
\newcommand{\eqlab}[1]{\begin{align}#1\end{align}}
\newcommand{\bmat}[1]{\begin{bmatrix}#1\end{bmatrix}}

\title{Explicit Jacobian matrix formulas for entropy stable summation-by-parts schemes}
\author{Jesse Chan, Christina Taylor}
\date{}

\begin{document}

\maketitle

\begin{abstract}
Entropy stable schemes replicate an entropy inequality at the semi-discrete level.  These schemes rely on an algebraic summation-by-parts (SBP) structure and a technique referred to as flux differencing, which has been explored and optimized in the context of matrix-free time-explicit solvers.  In this work, we provide simple and explicit formulas for Jacobian matrices for the semi-discrete systems of ODEs produced by entropy stable discretizations.  These formulas are derived based on the structure of flux differencing and automatic differentiation (AD).  Numerical results demonstrate the efficiency and utility of these Jacobian formulas, and apply them in the context of two-derivative explicit time-stepping schemes and implicit time-stepping.  
\end{abstract}

\section{Introduction}

This paper is concerned with the numerical solution of systems of nonlinear conservation laws.  Let $\Omega$ denote some domain with boundary $\partial \Omega$.  Nonlinear conservation laws are expressed as a system of nonlinear partial differential equations (PDEs) 
\begin{equation}
\pd{\bm{u}}{t}  + \sum_{i=1}^d\pd{\bm{f}_i(\bm{u})}{x_i} = 0, \qquad 
S(\bm{u}) \text{ convex}, \qquad
\bm{v}(\bm{u}) = \pd{S}{\bm{u}},
\label{eq:nonlineqs}
\end{equation}
where $\bm{u}\in \mathbb{R}^n$ are the conservative variables, $\bm{f}_i$ are nonlinear fluxes, and $\bm{v}(\bm{u})$ are the \textit{entropy variables} with respect to the entropy $S(\bm{u})$.  By multiplying (\ref{eq:nonlineqs}) by the entropy variables, vanishing viscosity solutions \cite{kruvzkov1970first} of many fluid systems \cite{hughes1986new, chen2017entropy} can be shown to satisfy the following entropy inequality
\begin{equation}
\int_{\Omega}\pd{S(\bm{u})}{t}\diff{\bm{x}} + \sum_{i=1}^d \int_{\partial \Omega} \LRp{\bm{v}^T\bm{f}_i(\bm{u}) - \psi_i(\bm{u})}n_i \leq 0\label{eq:entropyineq},
\end{equation}
where $n_i$ denotes the $i$th component of the outward normal vector.  The entropy inequality (\ref{eq:entropyineq}) is a statement of stability for nonlinear conservation laws \cite{mock1980systems, harten1983symmetric}.

High order entropy stable schemes (see for example \cite{carpenter2014entropy, gassner2016split, chen2017entropy, crean2018entropy, chan2017discretely}) reproduce this entropy inequality at the semi-discrete level.  The resulting methods display significantly improved robustness while retaining high order accuracy \cite{winters2018comparative, rojas2019robustness}.  These schemes are based on entropy conservative finite volume fluxes \cite{tadmor1987numerical}, which are extended to high order discretizations through a procedure referred to as flux differencing.  These methods have mainly been tested in the context of explicit time-stepping.  However, recent works have applied entropy stable methods to both the space-time and implicit settings \cite{friedrich2018entropy, hicken2020entropy}.

Both space-time and implicit time discretizations require the solution of a system of nonlinear equations.  This can be done using Newton's method, which involves the Jacobian matrix of the nonlinear equations.  While it is possible to compute the solution to the nonlinear system without explicitly computing the Jacobian matrix using Jacobian-free Newton-Krylov methods \cite{knoll2004jacobian, birken2019subcell}, the Jacobian matrix is also useful for preconditioning \cite{persson2006efficient, persson2008newton}.  In this work, we present explicit formulas for Jacobian matrices of systems resulting from entropy stable formulations.  We also show that computing the Jacobian matrix is not significantly more expensive than evaluating the residual of the nonlinear system.  Finally, we apply the new Jacobian formulas to both explicit two-derivative and implicit time-stepping schemes.

%, which are defined as follows.  Let $\bm{u}_L,\bm{u}_R$ denote left and right solution states.  Entropy conservative fluxes in the sense of Tadmor \cite{tadmor1987numerical} are defined as a set of bivariate symmetric and consistent functions $\bm{f}_{i,S}(\bm{u}_L,\bm{u}_R)$ such that 
%\[
%\LRp{\bm{v}_L -\bm{v}_R}^T\bm{f}_{i,S}(\bm{u}_L,\bm{u}_R) = \psi_i(\bm{u}_L) - \psi_i(\bm{u}_R), \qquad i = 1,\ldots,d.
%\]
%\note{Todo: finish}

\section{Explicit Jacobian formulas}

For clarity of presentation, we consider first a scalar nonlinear conservation law in one spatial dimension
\eqlab{
\pd{u}{t} + \pd{f(u)}{x} = 0. \label{eq:ncl}
}
We assume without loss of generality periodic boundary conditions, which will simplify the presentation of the main results.  Extensions to non-periodic boundaries are treated in Section~\ref{eq:npbc}.  The generalization to systems of nonlinear conservation laws and higher dimensions is also postponed until Section~\ref{sec:systems}.

Let $f_S(x,y)$ denote a bivariate scalar flux function which is symmetric and consistent.  Suppose $\bm{u}$ is a vector of nodal values of the solution.  Define the vector $\bm{f} = \bm{f}(\bm{u})$ approximating the flux derivative $\pd{f(u)}{x}$ as 
\eqlab{
\bm{f}(\bm{u}) = 2\LRp{\bm{Q}\circ \bm{F}}\bm{1}, \qquad \bm{F}_{ij} = f_S(\bm{u}_i,\bm{u}_j),
\label{eq:fu}
}
where $\bm{Q}$ is a discretization matrix to be specified later.  The simplest entropy stable numerical schemes based on flux differencing discretize (\ref{eq:ncl}) via the system of ODEs
\[
\bm{M}\td{\bm{u}}{t} + \bm{f}(\bm{u}) = \bm{0}.
\]
where $\bm{M}$ is a diagonal mass (norm) matrix with positive entries.  If $f_S(x,y)$ is entropy conservative and $\bm{Q}$ is skew-symmetric, then the resulting scheme is also discretely entropy conservative.  An entropy stable scheme can be constructed from an entropy conservative scheme by adding appropriate terms which dissipate entropy \cite{chen2017entropy, upperman2019entropy, hicken2020entropy}.

We are interested in computing the Jacobian matrix $\pd{\bm{f}}{\bm{u}}$. Let $\diag{\bm{x}}$ denote the diagonal matrix with the vector $\bm{x}$ on the diagonal, and let $\diag{\bm{A}}$ denote the vector diagonal of $\bm{A}$.  Then, we have the following lemma:
\begin{lemma}
\label{lemma:explicitJ}
Suppose that $\bm{Q} = \pm\bm{Q}^T$.  Then, the Jacobian matrix of (\ref{eq:fu}) can be expressed as either
\eq{
%\pd{\bm{f}}{\bm{u}} &= 2\LRp{\bm{Q}\circ \bm{F_y}} \pm \diag{\bm{1}^T\LRp{2\LRp{\bm{Q} + \diag{\diag{\bm{Q}}}}\circ \bm{F_y}}}\\
%\pd{\bm{f}}{\bm{u}} &= 2\LRp{\bm{Q}\circ \bm{F_x}^T} \pm \diag{\LRp{2\LRp{\bm{Q}+\diag{\diag{\bm{Q}}}}\circ \bm{F_x}}\bm{1}}
\pd{\bm{f}}{\bm{u}} &= 2\LRp{\bm{Q}\circ \bm{F_y}} \pm \diag{\bm{1}^T\LRp{2\bm{Q}\circ \bm{F_y}}}\\
\pd{\bm{f}}{\bm{u}} &= 2\LRp{\bm{Q}\circ \bm{F_x}^T} \pm \diag{\LRp{2\bm{Q} \circ \bm{F_x}}\bm{1}}
}
where the matrices $\bm{F}_x,\bm{F}_y$ is 
\[
\LRp{\bm{F}_x}_{ij} = \LRl{\pd{f_S}{x}}_{\bm{u}_i,\bm{u}_j}, \qquad \LRp{\bm{F}_y}_{ij} = \LRl{\pd{f_S}{y}}_{\bm{u}_i,\bm{u}_j}.
\]
\end{lemma}
\begin{proof}
We will prove the first formula involving $\bm{F}_y$.  The second formula follows via symmetry and similar steps.  By the chain rule,
\begin{align*}
\LRp{\pd{\bm{f}}{\bm{u}}}_{ij} &= \pd{\bm{f}_i}{\bm{u}_j} =  \sum_{k} 2\bm{Q}_{ik} \pd{}{\bm{u}_j}f_S\LRp{\bm{u}_i,\bm{u}_k}\\
&= \sum_{k} 2\bm{Q}_{ik} \LRp{\LRl{ \pd{f_S}{x}}_{\bm{u}_i,\bm{u}_k}\pd{\bm{u}_i}{\bm{u}_j} + \LRl{\pd{f_S}{y}}_{\bm{u}_i,\bm{u}_k} \pd{\bm{u}_k}{\bm{u}_j}}
\end{align*}
If $i\neq j$, then $\pd{\bm{u}_k}{\bm{u}_j} = \delta_{jk}$ and
\begin{align*}
\pd{\bm{f}_i}{\bm{u}_j} &= 2\bm{Q}_{ij} \LRl{\pd{f_S}{y}}_{\bm{u}_i,\bm{u}_j}.
\end{align*}
When $i=j$, $\pd{\bm{u}_i}{\bm{u}_j} = \pd{\bm{u}_i}{\bm{u}_i} = 1$, and 
\begin{align*}
\pd{\bm{f}_i}{\bm{u}_i} &= 
\LRp{\sum_{k} 2 \bm{Q}_{ik} \LRl{\pd{f_S}{x}}_{\bm{u}_i,\bm{u}_k}} + 2\bm{Q}_{ii}\LRl{\pd{f_S}{y}}_{\bm{u}_i,\bm{u}_i}.
\end{align*}
The term $2\bm{Q}_{ii}\LRl{\pd{f_S}{y}}_{\bm{u}_i,\bm{u}_i}$ is the diagonal of the matrix $2\LRp{\bm{Q}\circ \bm{F_y}}$, and we can simplify the first summation term.  By the symmetry of $f_S(x,y)$, we have that
\[
\LRl{\pd{f_S}{y}}_{x,y} = \LRl{\pd{f_S}{x}}_{y,x}
\]
Thus, by $\bm{Q} = \pm\bm{Q}^T$, 
\begin{align*}
\sum_k 2\bm{Q}_{ik} \LRl{\pd{f_S}{x}}_{\bm{u}_i,\bm{u}_k} &= \sum_k 2\bm{Q}_{ik} \LRl{\pd{f_S}{y}}_{\bm{u}_k,\bm{u}_i}\\
&= \LRp{\LRp{2\bm{Q}\circ \bm{F_y}^T} \bm{1}}_i = \LRp{\pm\bm{1}^T\LRp{2\bm{Q}\circ \bm{F_y}}}_i.
\end{align*}
\end{proof}
%Rearranging the steps of this proof implies that the Jacobian can also be expressed as 
While we consider only symmetric and skew-symmetric matrices $\bm{Q}$ in this work, one can use this lemma to compute the Jacobian $\pd{\bm{f}}{\bm{u}}$ for arbitrary matrices $\bm{Q}$ since any real matrix can be decomposed into symmetric and skew parts
\[
\bm{Q} = \frac{1}{2} \LRp{\bm{Q}+\bm{Q}^T} +  \frac{1}{2} \LRp{\bm{Q}-\bm{Q}^T}.
\]
Two applications of Lemma~\ref{lemma:explicitJ} then provide a formula for the Jacobian of (\ref{eq:fu}).

\subsection{Computing derivatives of bivariate flux functions}

The aforementioned proofs require partial derivatives of flux functions $f_S(u_L,u_R)$ with respect to at least one argument.  While this can be done by hand for simple fluxes, it rapidly becomes cumbersome for complex or piecewise-defined flux functions such as the logarithmic mean \cite{ismail2009affordable, winters2019entropy}.  This can be avoided by using Automatic Differentiation (AD) \cite{griewank2008evaluating}.  AD is distinct from both symbolic differentiation and finite difference approximations in that it does not return an explicit expression for derivatives, but is still able to evaluate derivatives accurately up to machine precision.  Given a computational function, AD returns a second black-box function which evaluates the derivative.

In this work, we utilize the Julia implementation of forward-mode automatic differentiation provided by \verb+ForwardDiff.jl+ \cite{RevelsLubinPapamarkou2016}.  The procedure is remarkably simple: given some flux function \verb+f(x,y)+, \verb+ForwardDiff.jl+ returns the derivative with respect to either $x$ or $y$ as another function.  For example, defining the function $\LRl{\pd{f}{y}}_{x,y}$ is a one-line operation:
\begin{verbbox}
dfdy(x,y) = ForwardDiff.derivative(y->f(x,y),y)
\end{verbbox}
\begin{figure}[!h]
\centering
\theverbbox
\end{figure}
\\
This simple API is made possible by the flexible Julia type system  \cite{bezanson2017julia}.

In principle, AD can be directly applied to $\bm{f}(\bm{u})$ to compute the Jacobian matrix.  However, because AD scales with the number of inputs and outputs, the cost of applying AD directly to $\bm{f}(\bm{u})$ increases as the discretization resolution increases.  In contrast, using the approach in this paper, the entries of the Jacobian matrix can be computed for roughly the same cost as a single evaluation of $\bm{f}(\bm{u})$.  Moreover, explicit Jacobian formulas enable implementations to take advantage of sparsity in $\bm{Q}$ when assembling the Jacobian matrix.

\section{Examples of matrices which appear in entropy stable numerical schemes}

In this section, we give some examples of matrices $\bm{Q}$ which appear in entropy stable numerical discretizations.  We assume periodicity, which ensures a skew-symmetric structure.  The treatment of non-periodic

\subsection{Finite volume methods}

The spatial discretization for a finite volume scheme can be reformulated in terms of (\ref{eq:fu}) \cite{chan2019entropy}.  Suppose that the 1D interval $[-1,1]$ is decomposed into $K$ non-overlapping elements of size $h$.  An entropy conservative finite volume scheme is given as
\eq{
\td{u_1}{t} &+ \frac{f_S(u_{2},u_1) - f_S(u_1,u_{K})}{h} = 0\\
\td{u_i}{t} &+ \frac{f_S(u_{i+1},u_i) - f_S(u_i,u_{i-1})}{h} = 0, \qquad i = 2,\ldots, K-1,\\
\td{u_K}{t} &+ \frac{f_S(u_{1},u_K) - f_S(u_K,u_{K-1})}{h} = 0,
}
where $u_i$ denotes the average value of the solution on each element and $f_S$ is an entropy conservative flux.  Let $\bm{M} = h\bm{I}$ and let $\bm{Q}$ be the periodic second-order central difference matrix
\[
\bm{Q} = \frac{1}{2}\begin{bmatrix}
0 & 1 & &\ldots & -1\\
-1 & 0 & 1 &&  \\
& -1 & 0 & 1 &  \\
 & & & \ddots &  \\
1 & &\ldots  & -1 & 0
\end{bmatrix}.
\]
Then, an entropy finite volume scheme is equivalent to 
\[
\td{\bm{u}}{t} + 2\LRp{\bm{Q}\circ\bm{F}}\bm{1} = \bm{0}, \qquad \bm{F}_{ij} = f_S(u_i,u_j)
\]
where $\bm{u} = \LRs{u_1,\ldots, u_K}^T$ is the vector of solution values.

\subsection{Multi-block summation-by-parts finite differences}

We next consider a multi-element summation-by-parts (SBP) finite element discretization \cite{kreiss1974finite, carpenter1999stable}.  Suppose again that the domain $[-1,1]$ is decomposed into $K$ non-overlapping elements $D^k$ of size $h$.  Let ${\bm{M}}, {\bm{Q}} \in \mathbb{R}^{N_p\times N_p}$ denote diagonal mass (norm) and nodal differentiation matrices, such that ${\bm{M}}^{-1}{\bm{Q}}$ approximates the first derivative on a reference interval and is exact for polynomials up to degree $N$.  The operators ${\bm{M}}, {\bm{Q}}$ satisfy an SBP property if
\eqlab{
{\bm{Q}}+{\bm{Q}}^T = {\bm{B}}, \qquad 
{\bm{B}} = \bmat{
-1 & & & \\
& 0 & & \\
& & \ddots & \\
& & & 1 
}.
\label{eq:fvQ}
}
We note that nodal discontinuous Galerkin spectral element (DG-SEM) discretizations \cite{kopriva2009implementing} also fall into a SBP framework \cite{gassner2013skew}, and are thus also included in this framework.

These matrices can be used to construct entropy conservative high order discretizations.  Let $J = h/2$ is the Jacobian of the mapping from the reference element $[-1,1]$ to a physical interval of size $h$ and let $\bm{F}^k_{ij} = f_S(u_{i,k},u_{j,k})$ denote the matrix of flux interactions between different nodes on the element $D^k$.  A local formulation on the element $D^k$ is given by
\eqlab{
J_k {\bm{M}} \td{\bm{u}_k}{t} + 2\LRp{{\bm{Q}} \circ \bm{F}^k}\bm{1} + \bm{B}\LRp{\bm{f}^*-f(\bm{u}_k)}   = \bm{0}, \qquad \bm{f}^* = \bmat{
f_S(u_{1,k}^+,u_{1,k})\\
0\\
\vdots\\
f_S(u_{N_p,k}^+,u_{N_p,k})
},
\label{eq:sbp1D}
}
where $u_{1,k}^+, u_{N_p,k}^+$ denote the exterior values of $u_{1,k}, u_{N_p,k}$ on neighboring elements.  Assuming that the elements are ordered from left to right in ascending order, for interior element indices $1 < k < K$, these are given by
\[
u_{1,k}^+ = u_{N_p,k-1}, \qquad u_{N_p,k}^+ = u_{1,k+1}.
\]
In other words, the first node on $D^k$ is connected to the last node on the previous element, and the last node on $D^k$ is connected to the first node on the next element.

For periodic boundary conditions, this local formulation can be understood as inducing a global skew-symmetric matrix.  To show this, we first use the SBP property to rewrite (\ref{eq:sbp1D}) in a skew-symmetric form \cite{chan2019skew}
\eqlab{
J_k {\bm{M}} \td{\bm{u}_k}{t} + \LRp{\LRp{\bm{Q}-\bm{Q}^T} \circ \bm{F}^k}\bm{1} + \bm{B}\bm{f}^* = \bm{0}.
%\qquad 
%\bm{f}^* = \bmat{
%f_S(u_{1,k}^+,u_{1,k})\\
%0\\
%\vdots\\
%f_S(u_{n,k}^+,u_{n,k})
%},
\label{eq:sbp1D}
}
We now define a global vector $\bm{u}_{\Omega} = \LRs{\bm{u}_1, \bm{u}_2, \ldots, \bm{u}_K}^T$.  Let the global flux matrix be defined as
\[
\bm{F} = \bmat{
\bm{F}_{11} & \ldots & \bm{F}_{1K}\\
\vdots & \ddots & \vdots\\
\bm{F}_{K1} & \ldots & \bm{F}_{KK}
}, \qquad \LRp{\bm{F}_{k_1,k_2}}_{ij} = f_S(\bm{u}_{k_1,i}, \bm{u}_{k_2,j}).
\]
The blocks of the matrix $\bm{F}$ capture flux interactions between solution values at different nodes and elements.  
The local formulations can now be concatenated into a single skew-symmetric matrix
\[
\bm{M}_{\Omega}\td{\bm{u}_{\Omega}}{t} + 2\LRp{\bm{Q}_{\Omega}\circ \bm{F}}\bm{1} = \bm{0},
\]
where $\bm{M}_{\Omega}$ is the block-diagonal matrix with blocks $J_k\bm{M}$, and 
\eqlab{
\bm{Q}_{\Omega} = \frac{1}{2}\bmat{
\bm{S} &\bm{B}_R & & -\bm{B}_L\\
-\bm{B}_L& \bm{S} & \bm{B}_R&\\
& -\bm{B}_L & \ddots & \bm{B}_R\\
\bm{B}_R& & -\bm{B}_L & \bm{S}
}, \qquad \bm{S} = \LRp{\bm{Q}-\bm{Q}^T},
\label{eq:sbpmat}
}
where the matrices $\bm{B}_L,\bm{B}_R$ are zeros except for a single entry
\[
\bm{B}_L = \bmat{
 & &1\\
& \udots & \\
0 & & 
}, \qquad \bm{B}_R = \bm{B}_L^T = \bmat{
 & & 0\\
& \udots & \\
1 & & 
}
\]
The matrix $\bm{Q}_{\Omega}$ can be considered a high order generalization of the finite volume matrix (\ref{eq:fvQ}).  Similar ``global SBP operator'' approaches were used to construct simultaneous approximation (SBP-SAT) interface coupling terms in \cite{crean2018entropy, chan2018efficient, fernandez2019entropy}.

The generalization to higher dimensional domains and curved geometric mappings is straightforward, but notationally much more complicated.  The construction of skew matrices $\bm{Q}_{\Omega}$ follows from approaches detailed in \cite{crean2018entropy, chan2017discretely, chan2018discretely, chan2018efficient, chan2019skew, hicken2020entropy}.  We omit it for conciseness in this work.

\subsection{Systems of conservation laws: vector-valued fluxes}
\label{sec:systems}

It is straightforward to extend the scheme from scalar nonlinear conservation laws to an $n\times n$ system of conservation laws.  Let $\bm{f}_{S}(\bm{u}_L,\bm{u}_R): \mathbb{R}^{n}\times \mathbb{R}^n \rightarrow \mathbb{R}^n$ denote an entropy conservative flux function for a 1D system of conservation laws.  We can modify the arrays and matrices in (\ref{eq:sbpmat}) to account for this.

Let $\bm{u}_{\Omega}$ now denote a vector of vectors 
\[
\bm{u}_{\Omega} = \bmat{\bm{u}_1\\
\bm{u}_2\\
\vdots\\
\bm{u}_K}, \qquad \bm{u}_k = \bmat{
\bm{u}_{k,1}\\
\bm{u}_{k,2}\\
\vdots\\
\bm{u}_{k,N_p}
}, \qquad
\bm{u}_{k,j} = \bmat{
\bm{u}_{k,j,1}\\
\bm{u}_{k,j,2}\\
\vdots\\
\bm{u}_{k,j,n}
}
\]
for $k = 1,\ldots, K$ and $j = 1,\ldots, N_p$.  Here, $\bm{u}_{k,j,i}$ denotes the $j$th degree of freedom for the $i$th component of the solution on the $k$th element.   Then, we can define global mass and differentiation matrices $\bm{M}_{\Omega}, \bm{Q}_{\Omega}$ as in (\ref{eq:sbpmat}), with the local blocks $\bm{M}, \bm{Q},\bm{B}_L, \bm{B}_R$ replaced by 
\eq{
\bm{M} &\longrightarrow \bm{M} \otimes \bm{I}_n, \qquad 
\bm{Q} \longrightarrow \bm{Q} \otimes \bm{I}_n,\\
\bm{B}_L &\longrightarrow \bm{B}_L \otimes \bm{I}_n, \qquad 
\bm{B}_R \longrightarrow \bm{B}_R \otimes \bm{I}_n,
}
where $\bm{I}_n$ is the $n\times n$ identity matrix.  Note that $\bm{Q}_{\Omega}$ as defined in (\ref{eq:sbpmat}) remains skew-symmetric under this new definition.  We also redefine the flux matrix to account for the vector-valued nature of the flux function $\bm{f}_S$ 
\eqlab{
\bm{F} = \bmat{
\bm{F}_{11} & \ldots & \bm{F}_{1K}\\
\vdots & \ddots & \vdots\\
\bm{F}_{K1} & \ldots & \bm{F}_{KK}
}, \qquad \LRp{\bm{F}_{k_1,k_2}}_{ij} = \diag{\bm{f}_S(\bm{u}_{k_1,i}, \bm{u}_{k_2,j})}.\label{eq:fluxsystem}
}
%\note{Redefine $\bm{M}_{\Omega}, \bm{Q}_{\Omega}$ using Kronecker product blocks. }
Then, an entropy conservative scheme is given by
\[
\bm{M}_{\Omega}\td{\bm{u}_{\Omega}}{t} + 2\LRp{ \bm{Q}_{\Omega} \circ \bm{F}}\bm{1} = \bm{0}.
\]
We note that when computing Jacobians for systems of nonlinear conservation laws $\pd{\bm{f}_S}{\bm{y}}$, the derivative of the flux function is computed with respect to each component of the solution.

\section{Generalizations to different numerical settings}

In this section, we discuss how to generalize the explicit construction of Jacobians to different numerical settings, including non-collocation schemes and non-periodic boundary conditions.

\subsection{Non-collocated schemes: hybridized SBP operators, entropy projection, and over-integration}

Most entropy stable schemes rely on ``collocated'' SBP operators (where the mass matrix $\bm{M}_{\Omega}$ is diagonal) built on nodal sets which include boundary nodes \cite{chen2017entropy, crean2018entropy}.  However, in certain cases, energy and entropy stable SBP schemes constructed using non-diagonal mass matrices \cite{chan2017discretely, chan2019entropy} and more general nodal sets \cite{fernandez2014review, ranocha2018generalised, crean2017high, chan2018efficient} achieve higher accuracy than SBP schemes built on nodal sets which include boundary nodes.  We discuss how to extend explicit Jacobian formulas to such ``modal'' settings.

\subsubsection{Entropy conservative schemes and hybridized SBP operators}

We now assume that the solution is approximated by 
\[
u(\bm{x},t) \approx \sum_{j=1}^{N_p} \hat{\bm{u}}_{k,i}(t) \phi_i(\bm{x}),
\]
where $\hat{\bm{u}}_{k,i}$ denotes the coefficients of the solution on an element $D^k$.  We assume two sets of quadrature points: volume quadrature points and weights $\LRc{w_i, \bm{x}_{q,i}}_{i=1}^{N_q}$ and surface quadrature points $\LRc{w_{f,i}, \bm{x}_{f,i}}_{i=1}^{N_f}$.  We assume both quadrature rules are exact for certain classes of integrands as detailed in \cite{chan2019skew, chan2019entropy}.

Evaluating $\bm{u}(\bm{x},t)$ at quadrature points is equivalent to multiplication by an interpolation matrix $\bm{V}$
\eq{
{\bm{V}}_{ij} = \phi_j(\bm{x}_i), \qquad i = 1,\ldots, N_q, \qquad j = 1,\ldots, N_p\\
\LRp{\bm{V}_f}_{ij} = \phi_j(\bm{x}_{f,i}), \qquad i = 1,\ldots, N_f, \qquad j = 1,\ldots, N_p.
}
We can similarly define mass and projection matrices $\bm{M}, \bm{P}$ 
\[
\bm{M} = \bm{V}^T\bm{W}\bm{V}, \qquad \bm{P} = \bm{M}^{-1}\bm{V}^T\bm{W},
\]
where $\bm{W}$ is a diagonal matrix whose entries are the quadrature weights $w_i$.  We also define a face interpolation matrix
\[
\bm{E} = \bm{V}_f\bm{P}
\]
which evaluates the solution at face quadrature points.  Finally, we define the matrix $\bm{V}_h$ as the mapping between local coefficients $\hat{\bm{u}}_k$ and combined volume and surface quadrature points
\[
\bm{V}_h = \bmat{\bm{V} \\ \bm{V}_f}.
\]
These matrices are involved in the application of hybridized SBP operators (originally referred to as decoupled SBP operators) \cite{chan2017discretely, chenreview}.  We present the main ideas in a 1D setting, and refer the reader to \cite{crean2018entropy, chan2017discretely, chan2019skew} for details on multi-dimensional settings.

Given some modal weak differentiation matrix $\hat{\bm{Q}}$ which acts on the basis coefficients $\hat{\bm{u}}_k$, we define a nodal differentiation matrix $\bm{Q} = \bm{P}^T\hat{\bm{Q}}\bm{P}$.  Then, we can define a hybridized SBP operator as 
\[
\bm{Q}_h = \frac{1}{2}\bmat{\bm{Q}-\bm{Q}^T & \bm{E}^T\bm{B} \\
\bm{B}\bm{E} & \bm{B}}, \qquad \bm{B} = \bmat{-1 & \\ & 1}.  
\]
The operator $\bm{Q}_h$ can be used to approximate coefficients of the derivative in the basis $\phi_i(\bm{x})$.  Let $f(u)$ denote some function of $u(\bm{x})$, and let $\hat{\bm{u}}$ denote the basis coefficients of $u(\bm{x})$.  Then, 
\[
\pd{f(u)}{x}\approx \sum_j {\hat{\bm{f}}}_j \phi_j(\bm{x}), \qquad \hat{\bm{f}} = \bm{M}^{-1}\bm{V}_h^T\bm{Q}_h f\LRp{\bm{V}_h\hat{\bm{u}}}
\]

We now construct global matrices for the multi-element (periodic) case.  We begin by concatenating the local coefficients $\hat{\bm{u}}_{k,i}$ into a global coefficient vector $\hat{\bm{u}}_{\Omega}$.  We also introduce boundary matrices $\bm{B}, \bm{B}_L,\bm{B}_R$ which enforce coupling between different elements and are defined as 
\[
\bm{B}_L = \bmat{ & 1\\0 &}, \qquad \bm{B}_R = \bmat{ &0 \\1  &}, \qquad \bm{B} = \bmat{-1 & \\ & 1}.
\]
In the multi-dimensional case, the entries of $\bm{B}_L, \bm{B}_R$ correspond instead to outward normals \cite{crean2018entropy, chan2017discretely}.

We now assume that the global solution $\bm{u}(\bm{x},t) \in \mathbb{R}^n$ is vector-valued, and redefine local matrices so that they act on the global vector of vector-valued coefficients
\eq{
\bm{V} &\longrightarrow \bm{I}_K \otimes \LRp{\bm{V} \otimes \bm{I}_n}, \qquad \bm{P} \longrightarrow \bm{I}_K \otimes \LRp{\bm{P}  \otimes \bm{I}_n}\\ 
\bm{E} &\longrightarrow \bm{I}_K \otimes \LRp{\bm{E}  \otimes \bm{I}_n}, \qquad \bm{V}_h \longrightarrow \bm{I}_K \otimes \LRp{\bm{V}_h  \otimes \bm{I}_n}\\
\bm{Q} &\longrightarrow \bm{I}_K \otimes \LRp{\bm{Q}  \otimes \bm{I}_n}, \qquad \bm{B} \longrightarrow \bm{I}_K \otimes \LRp{\bm{B}  \otimes \bm{I}_n}\\
\bm{B}_L &\longrightarrow \bm{I}_K \otimes \LRp{\bm{B}_L  \otimes \bm{I}_n}, \qquad \bm{B}_R \longrightarrow \bm{I}_K \otimes \LRp{\bm{B}_R  \otimes \bm{I}_n}.
}
We can also use $\bm{Q}_h$ to construct a globally skew-symmetric differentiation matrix (see also \cite{chan2018efficient}).  Define the matrix $\bm{S} = \bm{Q}-\bm{Q}^T$, and define $\bm{Q}_{\Omega}$ as the global block matrix
\[
\bm{Q}_{\Omega} = \frac{1}{2}\LRs{\begin{array}{cc|cc|cc|cc}
\bm{S} &  \bm{E}^T\bm{B} & &&& && \\
 - \bm{B}\bm{E} &  & & \bm{B}_R && && -\bm{B}_L\\ [.2em]
 \hline\\[-1.0em]
&& \bm{S} &  \bm{E}^T\bm{B}&&&& \\
& -\bm{B}_L &  - \bm{B}\bm{E} &&& \bm{B}_R&&\\
 \hline\\[-1.0em]
&&&& \ddots & \ddots&  \\
&&&   -\bm{B}_L & \ddots&\ddots&   \ddots&\\
 \hline\\[-1.0em]
&&&&&\ddots& \bm{S} &  \bm{E}^T\bm{B} \\
\bm{B}_R &&& &&-\bm{B}_L &  - \bm{B}\bm{E} &\\
\end{array}},
\]

It was shown in \cite{parsani2016entropy, chan2017discretely} that, when either the mass matrix is non-diagonal or the nodal set does not contain appropriate boundary points, it is necessary to perform an entropy projection (or extrapolation \cite{chenreview}) step to ensure discrete entropy stability.  Let $\bm{v}(\bm{u})$ denote the entropy variables as a function of the conservative variables, and let $\bm{u}(\bm{v})$ denote the inverse mapping.  We define the entropy projected variables $\tilde{\bm{u}}_{\Omega}$ as
\[
\tilde{\bm{u}}_{\Omega} = \bm{u}\LRp{\bm{V}_h\bm{P}\bm{v}\LRp{\bm{V}\hat{\bm{u}}_{\Omega}}}.
\]
We redefine the flux block $\bm{F}_{k_1,k_2}$ in (\ref{eq:fluxsystem}) using the entropy projected variables
\[
\LRp{\bm{F}_{k_1,k_2}}_{ij} = \diag{\bm{f}_S(\tilde{\bm{u}}_{k_1,i}, \tilde{\bm{u}}_{k_2,j})}
\]
Then, an entropy conservative method is given by 
\[
\bm{M}_{\Omega}\td{\bm{u}_{\Omega}}{t} + 2\bm{V}^T\LRp{ \bm{Q}_{\Omega} \circ \bm{F}}\bm{1} = \bm{0}.  
\]

\subsubsection{Jacobian matrices for hybridized SBP schemes}

Redefine the nonlinear term 
\[
\bm{f}(\hat{\bm{u}}) = 2\bm{V}^T\LRp{ \bm{Q}_{\Omega} \circ \bm{F}}\bm{1}.
\]
We introduce $\pd{\bm{u}}{\bm{v}}$ and $\pd{\bm{v}}{\bm{u}}$ as Jacobians of the conservative variables with respect to the entropy variables and vice versa.  These have been explicitly derived for several equations (for example, the Jacobians for the compressible Navier-Stokes equations are given in \cite{hughes1986new}).

We can compute the Jacobian of $\bm{f}(\hat{\bm{u}})$ via the chain rule, and illustrate this for an entry $i\neq j$ of the Jacobian
\[
\LRp{\pd{\bm{f}}{\hat{\bm{u}}_{\Omega}}}_{ij} =  2 \bm{V}_h^T \pd{}{\LRp{\hat{\bm{u}}_{\Omega}}_j} \LRp{\LRp{\bm{Q}_h\circ\bm{F}}\bm{1}}_i.
\]
We focus on the latter term $\pd{}{\hat{\bm{u}}_{\Omega}} \LRp{\bm{Q}_h\circ\bm{F}}\bm{1}$
\eq{
\LRp{\pd{}{\hat{\bm{u}}_{\Omega}} \LRp{\bm{Q}_h\circ\bm{F}}\bm{1}}_{ij} &= \pd{}{\hat{\bm{u}}_{\Omega,j}} \sum_{k} \LRp{\bm{Q}_h}_{ik} \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_k} \\
&=  2 \bm{V}_h^T \sum_{k} \LRp{\bm{Q}_h}_{ik} \LRl{\pd{\bm{f}_S}{y}}_{\tilde{\bm{u}}_i,\tilde{\bm{u}}_k} \pd{\tilde{\bm{u}}_i}{\hat{\bm{u}}_{\Omega,j}}
}
We define the matrix $\pd{\tilde{\bm{f}}}{\tilde{\bm{u}}}$ as
\[
\LRp{\pd{\tilde{\bm{f}}}{\tilde{\bm{u}}}}_{ij} = \LRp{\bm{Q}_h}_{ij} \LRl{\pd{\bm{f}_S}{y}}_{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j} = \LRp{\bm{Q}_h\circ \bm{F}_y}_{ij}.
\]
Let $\tilde{\bm{v}} = \bm{V}_h\bm{P}\bm{v}\LRp{\bm{V}\hat{\bm{u}}_{\Omega}}$ denote the projected entropy variables at evaluated at volume quadrature points.
The vector $\pd{\tilde{\bm{u}}}{\hat{\bm{u}}_{\Omega}}$  can be further expanded as 
\[
\pd{\tilde{\bm{u}}}{\hat{\bm{u}}_{\Omega}} = \LRl{\pd{\bm{u}}{\bm{v}}}_{\tilde{\bm{u}}} \bm{V}_h \bm{P} \LRl{\pd{\bm{v}}{\bm{u}}}_{\tilde{\bm{v}}} \bm{V}.
\]
where the Jacobian matrices for the maps between conservative and entropy variables are block diagonal matrices given by
\[
\LRl{\pd{\bm{u}}{\bm{v}}}_{\tilde{\bm{u}}} = \bmat{
\LRl{\pd{\bm{u}}{\bm{v}}}_{\tilde{\bm{u}}_1} &&\\
& \ddots &\\
&& \LRl{\pd{\bm{u}}{\bm{v}}}_{\tilde{\bm{u}}_K}
}, \qquad \LRl{\pd{\bm{u}}{\bm{v}}}_{\tilde{\bm{u}}_k} = \bmat{
\LRl{\pd{\bm{u}}{\bm{v}}}_{\tilde{\bm{u}}_{1,k}} &&\\
 &\ddots &\\
&& \LRl{\pd{\bm{u}}{\bm{v}}}_{\tilde{\bm{u}}_{N_p,k}}
},
\]
and the local block $\LRl{\pd{\bm{u}}{\bm{v}}}_{\tilde{\bm{u}}_{j,k}}$ is simply the Jacobian $\pd{\bm{u}}{\bm{v}}$ evaluated at the $j$th nodal value $\tilde{\bm{u}}_{j,k}$ on the $k$th element.

Let $N_p, N_q, N_f$ denote the number of total basis functions, quadrature points, and face quadrature points respectively, and define $N_{\rm total} = N_q + N_f$.  The structure and dimensions of matrices involved in constructing the Jacobian matrix are illustrated as follows:
\[
\pd{\bm{f}}{\hat{\bm{u}}_{\Omega}} = \renewcommand\matscale{.67}
\raiserows{2}{\matbox{3}{7}{N_p}{N_{\rm total}}{\bm{V}_h^T}}
\raiserows{0}{\matbox{7}{7}{N_{\rm total}}{N_{\rm total}}{\pd{\tilde{\bm{f}}}{\tilde{\bm{u}}}}}
\raiserows{0}{\matbox{7}{7}{N_{\rm total}}{N_{\rm total}}{\LRl{\pd{\bm{u}}{\bm{v}}}_{\tilde{\bm{u}}}}}
\raiserows{0}{\matbox{7}{5}{N_{\rm total}}{N_q}{ \bm{V}_h\bm{P} }}
\raiserows{1}{\matbox{5}{5}{N_q}{N_q}{ \LRl{\pd{\bm{v}}{\bm{u}}}_{\tilde{\bm{v}}} }}
\raiserows{1}{\matbox{5}{3}{N_q}{N_p}{\bm{V}}}
\]

\subsection{Non-periodic boundary conditions}
\label{eq:npbc}

Finally, we consider the case of non-periodic boundary conditions.  \note{Impose BCs by specifying exterior value.}

\section{Numerical experiments}

In this section, we compare the computational efficiency of the explicit formulas derived in this paper with other methods for computing the Jacobian.  Additionally, we apply explicit Jacobian formulas to enable two-derivative time-stepping methods \cite{chan2010explicit} and time-implicit discretizations.

\subsection{Comparisons of computational cost}

\begin{itemize}
\item Compare cost of evaluating scalar flux function vs cost of evaluating AD derivative (Burgers' flux, rational kernel $1/(a+b)$, and logmean).
\item Compare cost of computing Jacobian matrix via direct finite differences, \verb+FiniteDiff.jl+, AD applied to $\bm{f}(\bm{u})$, and the explicit Jacobian formula.  
\end{itemize}
These computations are performed using a single thread.  Because the steps are trivially parallelizable with high arithmetic intensity, the use of multi-threading or GPU acceleration is not expected to significantly change the relative costs.

\subsection{Two-derivative time-stepping methods}

Consider a general system of ODEs
\[
\td{\bm{u}}{t} + \bm{f}(\bm{u}) = \bm{0}.
\]
Two-derivative explicit time-stepping methods are constructed based on the assumption that second derivatives of $\bm{u}$ in time are available \cite{chan2010explicit, christlieb2016explicit}.  The resulting schemes can achieve higher order accuracy with fewer stages and function evaluations compared to standard Runge-Kutta methods.

Let $\bm{g}(\bm{u})$ denote the second derivative of $\bm{u}$ in time
\[
\bm{g}(\bm{u}) = \frac{{\rm d}^2 \bm{u}}{{\rm dt}^2} = \td{}{t} \bm{f}(\bm{u}) = \pd{\bm{f}}{\bm{u}}\td{\bm{u}}{t},
\]
where we have used the chain rule in the final step.  The simplest two-derivative method is the second order scheme given in \cite{chan2010explicit} as
\[
\bm{u}^{k+1} = \bm{u}^k + \Delta t \bm{f}(\bm{u}^k) + \frac{\Delta t^2}{2} \bm{g}(\bm{u}^k),
\]
where $\bm{u}^k$ denotes the solution at the $k$th time-step.

\subsection{Time-implicit discretizations}

Jacobian matrices also appear in time-implicit discretizations of nonlinear ODEs.  Consider the implicit midpoint rule 
\[
\bm{u}^{k+1} = \bm{u}^k + \Delta t \bm{f}\LRp{\frac{\bm{u}^{k+1} + \bm{u}^k}{2}}.
\]
This can be rewritten in the following form
\eq{
\bm{r} &= \bm{f}\LRp{\bm{u}^k + \frac{\Delta t}{2}\bm{r}}\\
\bm{u}^{k+1} &= \bm{u}^k + \Delta t \bm{r}.
}
Solving for $\bm{r}$ is a nonlinear equation and can be done via Newton's method
\eq{
\bm{u}^\ell &= \bm{u}^k + \frac{\Delta t}{2}\bm{r}^\ell\\
\bm{r}^{\ell + 1} &= \bm{r}^\ell - \LRp{\bm{I} - \frac{\Delta t}{2}\LRl{\pd{\bm{f}}{\bm{u}}}_{\bm{u}^{\ell}}}^{-1} \LRp{\bm{r}^\ell - \bm{f}\LRp{\bm{u}^\ell}}
}

\section{Conclusion and acknowledgments}

Jesse Chan gratefully acknowledges support from the National Science Foundation under awards DMS-1719818, DMS-1712639, and DMS-CAREER-1943186.

\bibliographystyle{unsrt}
\bibliography{refs}

\end{document}

### 2020-03-02 docs/ESDG_jacobian/main.tex
updating jacobian paper

@@
 }{\endproof}
+\newtheorem*{remark}{Remark}
 
@@
 \newcommand{\diag}[1]{{\rm diag}\LRp{#1}}
-\newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}{\rm #2}}}
+\newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}{ {#2}}}}
 \newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
@@
 
-\subsection{Systems of conservation laws: vector-valued fluxes}
+%\subsection{Dissipative terms}
+%
+%Entropy stable schemes are constructed by adding entropy dissipation to entropy conservative formulations.  We characterize such
+
+\section{Systems of conservation laws}
 \label{sec:systems}
 
-It is straightforward to extend the scheme from scalar nonlinear conservation laws to an $n\times n$ system of conservation laws.
+In this section, we extend the explicit Jacobian formulas from scalar nonlinear conservation laws to an $n\times n$ system of conservation laws.
 Let $\bm{f}_{S}(\bm{u}_L,\bm{u}_R): \mathbb{R}^{n}\times \mathbb{R}^n \rightarrow \mathbb{R}^n$ denote an entropy conservative flux function for a 1D system of conservation laws.
-We can modify the arrays and matrices in (\ref{eq:sbpmat}) to account for this.
+We first formulate a system of ODEs by modifying the definition of arrays and matrices in (\ref{eq:sbpmat}).
 
-Let $\bm{u}_{\Omega}$ now denote a vector of vectors 
+Let $\bm{u}_{\Omega}$ denote a vector of vectors 
 \[
@@
 \vdots\\
-\bm{u}_K}, \qquad \bm{u}_k = \bmat{
-\bm{u}_{k,1}\\
-\bm{u}_{k,2}\\
+\bm{u}_n}, \qquad \bm{u}_i = \bmat{
+\bm{u}_{i,1}\\
+\bm{u}_{i,2}\\
 \vdots\\
-\bm{u}_{k,N_p}
+\bm{u}_{i,K}
 }, \qquad
-\bm{u}_{k,j} = \bmat{
-\bm{u}_{k,j,1}\\
-\bm{u}_{k,j,2}\\
+\bm{u}_{i,k} = \bmat{
+\bm{u}_{i,k,1}\\
+\bm{u}_{i,k,2}\\
 \vdots\\
-\bm{u}_{k,j,n}
+\bm{u}_{i,k,N_p}
 }
 \]
-for $k = 1,\ldots, K$ and $j = 1,\ldots, N_p$.
+for $\ell = 1,\ldots, n$, $k = 1,\ldots, K$ and $j = 1,\ldots, N_p$.  Here, $\bm{u}_{\ell,k,j}$ denotes the $j$th degree of freedom for the $\ell$th component of the solution on the $k$th element.  
+%We redefine these matrices as the $n \times n$ block matrices 
+%\eq{
+%&\bm{M}_{\Omega} \longrightarrow \bm{I}_n\otimes \bm{M}_{\Omega} = \bmat{\bm{M}_{\Omega} &  & \\
+%& \ddots & \\
+% &  & \bm{M}_{\Omega}},
+%\\
+%&\bm{Q}_{\Omega} \longrightarrow \bm{e}_n\bm{e}_n^T\otimes \bm{Q}_{\Omega} = 
+%\bmat{\bm{Q}_{\Omega} & \ldots & \bm{Q}_{\Omega}\\
+%\vdots & \ddots & \vdots\\
+%\bm{Q}_{\Omega} & \ldots & \bm{Q}_{\Omega}},
+%}
+%where $\bm{e}_n\in \mathbb{R}^n$ is the vector of all ones, and $\bm{e}_n\bm{e}_n^T$ is the $n\times n$ matrix of all ones.  
+Let $(k_1, j_1)$ and $(k_2, j_2)$ be tuples which correspond to row and column indices of a matrix.  
+We now define the flux matrix $\bm{F}$ as the $n\times n$ block matrix containing flux interactions between different nodes
+\eqlab{
+\bm{F} = \bmat{
+\bm{F}_1 &&\\
+& \ddots &\\
+&& \bm{F}_n
+}, \qquad \LRp{\bm{F}_\ell}_{(k_1, j_1),(k_2,j_2)} = \LRp{\bm{f}_{S}\LRp{\bm{u}_{:, k_1,j_1},\bm{u}_{:, k_2,j_2}}}_{\ell}.
+\label{eq:fluxsys}
+}
+where $\bm{u}_{:,k,j}$ denotes the vector containing all solution components at the $k$th element and $j$th node, and each block $\bm{F}_\ell$ for $\ell = 1,\ldots, n$ corresponds to the $\ell$th component of the vector-valued flux evaluated at solution states $\bm{u}_{:,k_1,j_1},\bm{u}_{:,k_2,j_2}$.
+
+Let $\bm{M}_{\Omega}, \bm{Q}_{\Omega}$ denote the global mass and differentiation matrices in (\ref{eq:sbpmat}).  Then, an entropy conservative scheme is given by
+\[
+\LRp{\bm{I}_n\otimes \bm{M}_{\Omega}}\td{\bm{u}_{\Omega}}{t} + 2\LRp{ \LRp{\bm{e}_n\bm{e}_n^T\otimes \bm{Q}_{\Omega} } \circ \bm{F}}\bm{1} = \bm{0}.
+\]
+where $\bm{e}_n\in \mathbb{R}^n$ is the vector of all ones, and $\bm{e}_n\bm{e}_n^T$ is the $n\times n$ matrix of all ones.
+
+We now construct explicit Jacobian matrix formulas for systems of nonlinear conservation laws.  The right hand side function $\bm{f}(\bm{u})$ for systems can be rewritten as
+\[
+\bm{f}(\bm{u}) = 2\LRp{ \bm{Q}_{\Omega} \circ \bm{F}}\bm{1} 
+= 2\bmat{(\bm{Q}_{\Omega}\circ \bm{F}_1) \\ 
+\vdots\\
+(\bm{Q}_{\Omega}\circ \bm{F}_n)
+}\bm{1}
+\]
+Then, the Jacobian matrix is 
+\eqlab{
+\pd{\bm{f}}{\bm{u}} &= 2\bmat{
+\bm{Q}_{\Omega}\circ \bm{F}_{1,\bm{u}_1} & \ldots & \bm{Q}_{\Omega}\circ \bm{F}_{1,\bm{u}_n}\\ 
+\vdots & \ddots & \vdots\\
+\bm{Q}_{\Omega}\circ \bm{F}_{n,\bm{u}_1} & \ldots & \bm{Q}_{\Omega}\circ \bm{F}_{n,\bm{u}_n}
+} \label{eq:fluxjac}\\
+&= \LRp{\bm{e}_n\bm{e}_n^T\otimes \bm{Q}_{\Omega}} \circ 
+\bmat{
+\bm{F}_{1,\bm{u}_1} & \ldots &\bm{F}_{1,\bm{u}_{n}}\\
+\vdots & \ddots & \vdots \\
+\bm{F}_{n,\bm{u}_1} & \ldots &\bm{F}_{n,\bm{u}_n}\nonumber
+}.
+}
+Here, the flux matrices $\bm{F}_{i,\bm{u}_j}$ are evaluated via
+\[
+\LRp{\bm{F}_{i,\bm{u}_j}}_{j_1k_1,j_2k_2} = \LRl{\pd{\LRp{\bm{f}_S}_i}{\bm{u}_j}}_{\bm{u}_{:,k_1,j_1},\bm{u}_{:,k_2,j_2}},
+\]
+where $\pd{\LRp{\bm{f}_S}_i}{\bm{u}_{j}}$ denotes the derivative of the $i$th component of the flux $\bm{f}_S\LRp{\bm{u}_L,\bm{u}_R}$ with respect to the $j$th solution component of $\bm{u}_R$.
+Thus, each block $\bm{F}_{i,\bm{u}_j}$ corresponds to an entry of the Jacobian of $\bm{f}_S(\bm{u}_L,\bm{u}_R)$).
+
+\begin{remark}
+We note that an alternative ordering of degrees of freedom \cite{crean2018entropy} yields a Jacobian matrix with a more compact bandwidth.
+A smaller matrix bandwidth can also be achieved by bandwidth-minimizing algorithms such as the reverse Cuthill-McKee algorithm \cite{cuthill1969reducing}.
+%Let $\bm{u}_{\Omega}$ now denote a vector of vectors 
+%\[
+%\bm{u}_{\Omega} = \bmat{\bm{u}_1\\
+%\bm{u}_2\\
+%\vdots\\
+%\bm{u}_K}, \qquad \bm{u}_k = \bmat{
+%\bm{u}_{k,1}\\
+%\bm{u}_{k,2}\\
+%\vdots\\
+%\bm{u}_{k,N_p}
+%}, \qquad
+%\bm{u}_{k,j} = \bmat{
+%\bm{u}_{k,j,1}\\
+%\bm{u}_{k,j,2}\\
+%\vdots\\
+%\bm{u}_{k,j,n}
+%}
+%\]
+%for $k = 1,\ldots, K$ and $j = 1,\ldots, N_p$.
 Here, $\bm{u}_{k,j,i}$ denotes the $j$th degree of freedom for the $i$th component of the solution on the $k$th element.
-Then, we can define global mass and differentiation matrices $\bm{M}_{\Omega}, \bm{Q}_{\Omega}$ as in (\ref{eq:sbpmat}), with the local blocks $\bm{M}, \bm{Q},\bm{B}_L, \bm{B}_R$ replaced by
-\eq{
-\bm{M} &\longrightarrow \bm{M} \otimes \bm{I}_n, \qquad 
-\bm{Q} \longrightarrow \bm{Q} \otimes \bm{I}_n,\\
-\bm{B}_L &\longrightarrow \bm{B}_L \otimes \bm{I}_n, \qquad 
-\bm{B}_R \longrightarrow \bm{B}_R \otimes \bm{I}_n,
-}
-where $\bm{I}_n$ is the $n\times n$ identity matrix.
+Define global mass and differentiation matrices $\bm{M}_{\Omega}, \bm{Q}_{\Omega}$ as in (\ref{eq:sbpmat}), with the local blocks $\bm{M}, \bm{Q},\bm{B}_L, \bm{B}_R$ replaced by
+%\eq{
+%\bm{M} &\longrightarrow \bm{M} \otimes \bm{I}_n, \qquad 
+%\bm{Q} \longrightarrow \bm{Q} \otimes \bm{I}_n,\\
+%\bm{B}_L &\longrightarrow \bm{B}_L \otimes \bm{I}_n, \qquad 
+%\bm{B}_R \longrightarrow \bm{B}_R \otimes \bm{I}_n,
+%}
+%where $\bm{I}_n$ is the $n\times n$ identity matrix.
 Note that $\bm{Q}_{\Omega}$ as defined in (\ref{eq:sbpmat}) remains skew-symmetric under this new definition.
 We also redefine the flux matrix to account for the vector-valued nature of the flux function $\bm{f}_S$
-\eqlab{
-\bm{F} = \bmat{
-\bm{F}_{11} & \ldots & \bm{F}_{1K}\\
-\vdots & \ddots & \vdots\\
-\bm{F}_{K1} & \ldots & \bm{F}_{KK}
-}, \qquad \LRp{\bm{F}_{k_1,k_2}}_{ij} = \diag{\bm{f}_S(\bm{u}_{k_1,i}, \bm{u}_{k_2,j})}.\label{eq:fluxsystem}
-}
-%\note{Redefine $\bm{M}_{\Omega}, \bm{Q}_{\Omega}$ using Kronecker product blocks. }
-Then, an entropy conservative scheme is given by
-\[
-\bm{M}_{\Omega}\td{\bm{u}_{\Omega}}{t} + 2\LRp{ \bm{Q}_{\Omega} \circ \bm{F}}\bm{1} = \bm{0}.
-\]
-We note that when computing Jacobians for systems of nonlinear conservation laws $\pd{\bm{f}_S}{\bm{y}}$, the derivative of the flux function is computed with respect to each component of the solution.
+%\eqlab{
+%\bm{F} = \bmat{
+%\bm{F}_{11} & \ldots & \bm{F}_{1K}\\
+%\vdots & \ddots & \vdots\\
+%\bm{F}_{K1} & \ldots & \bm{F}_{KK}
+%}, \qquad \LRp{\bm{F}_{k_1,k_2}}_{ij} = \diag{\bm{f}_S(\bm{u}_{k_1,i}, \bm{u}_{k_2,j})}.\label{eq:fluxsystem}
+%}
+%Then, an entropy conservative formulation is given by
+%\[
+%\bm{M}_{\Omega}\td{\bm{u}_{\Omega}}{t} + 2\LRp{ \bm{Q}_{\Omega} \circ \bm{F}}\bm{1} = \bm{0}.
+%\]
+%Explicit Jacobians can also be constructed for this ordering.  We omit it for conciseness.
+\end{remark}
+
+%\subsection{Explicit Jacobian matrix formulas for systems of nonlinear conservation laws}
 
@@
 
-We now assume that the global solution $\bm{u}(\bm{x},t) \in \mathbb{R}^n$ is vector-valued, and redefine local matrices so that they act on the global vector of vector-valued coefficients
-\eq{
-\bm{V} &\longrightarrow \bm{I}_K \otimes \LRp{\bm{V} \otimes \bm{I}_n}, \qquad \bm{P} \longrightarrow \bm{I}_K \otimes \LRp{\bm{P}  \otimes \bm{I}_n}\\ 
-\bm{E} &\longrightarrow \bm{I}_K \otimes \LRp{\bm{E}  \otimes \bm{I}_n}, \qquad \bm{V}_h \longrightarrow \bm{I}_K \otimes \LRp{\bm{V}_h  \otimes \bm{I}_n}\\
-\bm{Q} &\longrightarrow \bm{I}_K \otimes \LRp{\bm{Q}  \otimes \bm{I}_n}, \qquad \bm{B} \longrightarrow \bm{I}_K \otimes \LRp{\bm{B}  \otimes \bm{I}_n}\\
-\bm{B}_L &\longrightarrow \bm{I}_K \otimes \LRp{\bm{B}_L  \otimes \bm{I}_n}, \qquad \bm{B}_R \longrightarrow \bm{I}_K \otimes \LRp{\bm{B}_R  \otimes \bm{I}_n}.
-}
 We can also use $\bm{Q}_h$ to construct a globally skew-symmetric differentiation matrix (see also \cite{chan2018efficient}).
@@
 \]
+We also construct global interpolation, projection, and extrapolation matrices
+\eq{
+\bm{V} &\longrightarrow \bm{I}_K \otimes \bm{V}, \qquad \bm{P} \longrightarrow \bm{I}_K \otimes \bm{P} \\ 
+\bm{E} &\longrightarrow \bm{I}_K \otimes \bm{E}, \qquad \bm{V}_h \longrightarrow \bm{I}_K \otimes \bm{V}_h
+}
+Finally, we assume that the global solution $\bm{u}(\bm{x},t) \in \mathbb{R}^n$ is vector-valued, and order the solution coefficients as in Section~\ref{sec:systems}.
 
@@
 \]
-We redefine the flux block $\bm{F}_{k_1,k_2}$ in (\ref{eq:fluxsystem}) using the entropy projected variables
+We evaluate each flux block $\bm{F}_{\ell}$ in (\ref{eq:fluxsys}) using the entropy projected variables
 \[
-\LRp{\bm{F}_{k_1,k_2}}_{ij} = \diag{\bm{f}_S(\tilde{\bm{u}}_{k_1,i}, \tilde{\bm{u}}_{k_2,j})}
+\LRp{\bm{F}_\ell}_{(k_1, j_1),(k_2,j_2)} = \LRp{\bm{f}_{S}\LRp{\tilde{\bm{u}}_{:, k_1,j_1},\tilde{\bm{u}}_{:, k_2,j_2}}}_{\ell}.
+%\LRp{\bm{F}_{i,\bm{u}_j}}_{j_1k_1,j_2k_2} = \LRl{\pd{\LRp{\bm{f}_S}_i}{\bm{u}_j}}_{\tilde{\bm{u}}_{:,k_1,j_1},\tilde{\bm{u}}_{:,k_2,j_2}},
 \]
@@
 \[
-\bm{M}_{\Omega}\td{\bm{u}_{\Omega}}{t} + 2\bm{V}^T\LRp{ \bm{Q}_{\Omega} \circ \bm{F}}\bm{1} = \bm{0}.  
+\LRp{\bm{I}_n\otimes \bm{M}_{\Omega}}\td{\bm{u}_{\Omega}}{t} + 2\LRp{\bm{I}_n\otimes \bm{V}_h}^T\LRp{ \LRp{\bm{e}_n\bm{e}_n^T\otimes \bm{Q}_{\Omega} } \circ \bm{F}}\bm{1} = \bm{0}.
+%\bm{M}_{\Omega}\td{\bm{u}_{\Omega}}{t} + 2\bm{V}^T\LRp{ \bm{Q}_{\Omega} \circ \bm{F}}\bm{1} = \bm{0}.  
 \]
@@
 \[
-\bm{f}(\hat{\bm{u}}) = 2\bm{V}^T\LRp{ \bm{Q}_{\Omega} \circ \bm{F}}\bm{1}.
+\bm{f}(\hat{\bm{u}}) = 2\LRp{\bm{I}_n\otimes \bm{V}_h}^T\LRp{ \LRp{\bm{e}_n\bm{e}_n^T\otimes \bm{Q}_{\Omega}} \circ \bm{F}}\bm{1}.
 \]
@@
 
-We can compute the Jacobian of $\bm{f}(\hat{\bm{u}})$ via the chain rule, and illustrate this for an entry $i\neq j$ of the Jacobian
+We can compute the Jacobian of $\bm{f}(\hat{\bm{u}})$ via the chain rule.  We assume $n=1$ for simplicity, and illustrate this for an entry $i\neq j$ of the Jacobian
 \[
-\LRp{\pd{\bm{f}}{\hat{\bm{u}}_{\Omega}}}_{ij} =  2 \bm{V}_h^T \pd{}{\LRp{\hat{\bm{u}}_{\Omega}}_j} \LRp{\LRp{\bm{Q}_h\circ\bm{F}}\bm{1}}_i.
+\LRp{\pd{\bm{f}}{\hat{\bm{u}}_{\Omega}}}_{ij} =  2 \bm{V}_h^T \pd{}{\LRp{\hat{\bm{u}}_{\Omega}}_j} \LRp{\LRp{\bm{Q}_{\Omega}\circ\bm{F}}\bm{1}}_i.
 \]
-We focus on the latter term $\pd{}{\hat{\bm{u}}_{\Omega}} \LRp{\bm{Q}_h\circ\bm{F}}\bm{1}$
+We focus on the latter term $\pd{}{\hat{\bm{u}}_{\Omega}} \LRp{\bm{Q}_{\Omega}\circ\bm{F}}\bm{1}$
 \eq{
-\LRp{\pd{}{\hat{\bm{u}}_{\Omega}} \LRp{\bm{Q}_h\circ\bm{F}}\bm{1}}_{ij} &= \pd{}{\hat{\bm{u}}_{\Omega,j}} \sum_{k} \LRp{\bm{Q}_h}_{ik} \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_k} \\
-&=  2 \bm{V}_h^T \sum_{k} \LRp{\bm{Q}_h}_{ik} \LRl{\pd{\bm{f}_S}{y}}_{\tilde{\bm{u}}_i,\tilde{\bm{u}}_k} \pd{\tilde{\bm{u}}_i}{\hat{\bm{u}}_{\Omega,j}}
+\LRp{\pd{}{\hat{\bm{u}}_{\Omega}} \LRp{\bm{Q}_{\Omega}\circ\bm{F}}\bm{1}}_{ij} &= \pd{}{\hat{\bm{u}}_{\Omega,j}} \sum_{k} \LRp{\bm{Q}_{\Omega}}_{ik} \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_k} \\
+&=  2 \bm{V}_h^T \sum_{k} \LRp{\bm{Q}_{\Omega}}_{ik} \LRl{\pd{\bm{f}_S}{y}}_{\tilde{\bm{u}}_i,\tilde{\bm{u}}_k} \pd{\tilde{\bm{u}}_i}{\hat{\bm{u}}_{\Omega,j}}
 }
-We define the matrix $\pd{\tilde{\bm{f}}}{\tilde{\bm{u}}}$ as
+We observe that the term $\pd{\tilde{\bm{u}}_i}{\hat{\bm{u}}_{\Omega,j}}$ does not disappear as it did in the proof of Lemma~\ref{lemma:explicitJ}.
+We thus treat the Jacobian matrix in two parts.
+First, we define the matrix $\pd{\tilde{\bm{f}}}{\tilde{\bm{u}}}$ as
 \[
-\LRp{\pd{\tilde{\bm{f}}}{\tilde{\bm{u}}}}_{ij} = \LRp{\bm{Q}_h}_{ij} \LRl{\pd{\bm{f}_S}{y}}_{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j} = \LRp{\bm{Q}_h\circ \bm{F}_y}_{ij}.
+\LRp{\pd{\tilde{\bm{f}}}{\tilde{\bm{u}}}}_{ij} = \LRp{\bm{Q}_{\Omega}}_{ij} \LRl{\pd{\bm{f}_S}{y}}_{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j} = \LRp{\bm{Q}_{\Omega}\circ \bm{F}_y}_{ij}.
 \]
-Let $\tilde{\bm{v}} = \bm{V}_h\bm{P}\bm{v}\LRp{\bm{V}\hat{\bm{u}}_{\Omega}}$ denote the projected entropy variables at evaluated at volume quadrature points.
+Now, let $\tilde{\bm{v}}$ denote the projected entropy variables at evaluated at volume quadrature points
+\[
+\tilde{\bm{v}} = \bm{V}_h\bm{P}\bm{v}\LRp{\bm{V}\hat{\bm{u}}_{\Omega}}.
+\]
 The vector $\pd{\tilde{\bm{u}}}{\hat{\bm{u}}_{\Omega}}$  can be further expanded as 
@@
 \]
-and the local block $\LRl{\pd{\bm{u}}{\bm{v}}}_{\tilde{\bm{u}}_{j,k}}$ is simply the Jacobian $\pd{\bm{u}}{\bm{v}}$ evaluated at the $j$th nodal value $\tilde{\bm{u}}_{j,k}$ on the $k$th element.
+and the local block $\LRl{\pd{\bm{u}}{\bm{v}}}_{\tilde{\bm{u}}_{j,k}}$ is the Jacobian matrix $\pd{\bm{u}}{\bm{v}}$ evaluated at the $j$th nodal solution value $\tilde{\bm{u}}_{j,k}$ on the $k$th element.
 
@@
 \raiserows{2}{\matbox{3}{7}{N_p}{N_{\rm total}}{\bm{V}_h^T}}
-\raiserows{0}{\matbox{7}{7}{N_{\rm total}}{N_{\rm total}}{\pd{\tilde{\bm{f}}}{\tilde{\bm{u}}}}}
+\raiserows{0}{\matbox{7}{7}{N_{\rm total}}{N_{\rm total}}{\bm{Q}_{\Omega}\circ\pd{\tilde{\bm{f}}}{\tilde{\bm{u}}}}}
 \raiserows{0}{\matbox{7}{7}{N_{\rm total}}{N_{\rm total}}{\LRl{\pd{\bm{u}}{\bm{v}}}_{\tilde{\bm{u}}}}}
@@
 \]
+
+We note that several matrices become simplified under common assumptions for entropy stable DG discretizations.
+These typically fall into two categories: collocated volume nodes and collocated volume and surface nodes \cite{shadpey2019energy}.
+For collocated volume nodes, the solution is represented using a nodal Lagrange basis constructed using $N_q = N_p$ volume quadrature nodes.
+Both collocated volume and surface nodes refers to the case when both the volume nodes collocated and the volume quadrature nodes also include surface quadrature nodes as a subset \cite{gassner2013skew, chen2017entropy}.
+For example, if only volume nodes are collocated \cite{chan2018efficient}, then $\bm{V} = \bm{I}$.
+If both volume and surface nodes are collocated, the system reduces to the simplified system described in Section~\ref{sec:systems} using the fact that $\pd{\bm{u}}{\bm{v}} = \LRp{\pd{\bm{v}}{\bm{u}}}^{-1}$.
 
@@
 \[
-\bm{u}^{k+1} = \bm{u}^k + \Delta t \bm{f}(\bm{u}^k) + \frac{\Delta t^2}{2} \bm{g}(\bm{u}^k),
+\bm{u}^{k+1} = \bm{u}^k - \Delta t \bm{f}(\bm{u}^k) - \frac{\Delta t^2}{2} \bm{g}(\bm{u}^k),
 \]
@@
 \[
-\bm{u}^{k+1} = \bm{u}^k + \Delta t \bm{f}\LRp{\frac{\bm{u}^{k+1} + \bm{u}^k}{2}}.
+\bm{u}^{k+1} = \bm{u}^k - \Delta t \bm{f}\LRp{\frac{\bm{u}^{k+1} + \bm{u}^k}{2}}.
 \]
-This can be rewritten in the following form
+This can be rewritten in the following form where $\bm{u}^{k+1/2} = \frac{\bm{u}^{k+1}+\bm{u}^k}{2}$
 \eq{
-\bm{r} &= \bm{f}\LRp{\bm{u}^k + \frac{\Delta t}{2}\bm{r}}\\
-\bm{u}^{k+1} &= \bm{u}^k + \Delta t \bm{r}.
+\bm{u}^{k+1/2} &= \bm{u}^k - \frac{\Delta t}{2}\bm{f}\LRp{\bm{u}^{k+1/2}}\\
+\bm{u}^{k+1} &= 2\bm{u}^{k+1/2} - \bm{u}^k.
 }
-Solving for $\bm{r}$ is a nonlinear equation and can be done via Newton's method
+Solving for $\bm{u}^{k+1/2}$ is a nonlinear equation and can be done via Newton's method
 \eq{
-\bm{u}^\ell &= \bm{u}^k + \frac{\Delta t}{2}\bm{r}^\ell\\
-\bm{r}^{\ell + 1} &= \bm{r}^\ell - \LRp{\bm{I} - \frac{\Delta t}{2}\LRl{\pd{\bm{f}}{\bm{u}}}_{\bm{u}^{\ell}}}^{-1} \LRp{\bm{r}^\ell - \bm{f}\LRp{\bm{u}^\ell}}
+%&\bm{u}^{k+1/2} + \frac{\Delta t}{2}\bm{f}\LRp{\bm{u}^{k+1/2}} - \bm{u}^k = \bm{0}\\
+&\bm{u}^{k+1/2,\ell+1} = \\
+&\bm{u}^{k+1/2,\ell} - \LRp{\bm{I} + \frac{\Delta t}{2}\LRl{\pd{\bm{f}}{\bm{u}}}_{\bm{u}^{k+1/2,\ell}}}^{-1}\LRp{\bm{u}^{k+1/2,\ell} + \frac{\Delta t}{2}\bm{f}\LRp{\bm{u}^{k+1/2,\ell}} - \bm{u}^k}.
 }
+
+%Solving for $\bm{r}$ is a nonlinear equation and can be done via Newton's method
+%\eq{
+%\bm{u}^\ell &= \bm{u}^k + \frac{\Delta t}{2}\bm{r}^\ell\\
+%\bm{r}^{\ell + 1} &= \bm{r}^\ell - \LRp{\bm{I} + \frac{\Delta t}{2}\LRl{\pd{\bm{f}}{\bm{u}}}_{\bm{u}^{\ell}}}^{-1} \LRp{\bm{r}^\ell + \bm{f}\LRp{\bm{u}^\ell}}
+%}
 

### 2020-04-03 docs/ESDG_ROM/reviewerResponse.tex
updating docs

\pdfoutput=1

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{bm}
%\usepackage{bbm}
%\usepackage{tikz}
%\usepackage[normalem]{ulem}
%\usepackage{hhline}
\usepackage{graphicx}
\usepackage{subfig}

\graphicspath{{./figs/}}

\usepackage{color}
%\usepackage{doi}

%\renewcommand{\topfraction}{0.85}
%\renewcommand{\textfraction}{0.1}
%\renewcommand{\floatpagefraction}{0.75}
%
%
%\newcommand{\bbm}[1]{\mathbbm{#1}}
%\newcommand{\bs}[1]{\boldsymbol{#1}}
%\newcommand{\equaldef}{\stackrel{\mathrm{def}}{=}}
%
%
%\newcommand{\mb}[1]{\mathbf{#1}}
%\newcommand{\mbb}[1]{\mathbb{#1}}
%\newcommand{\mc}[1]{\mathcal{#1}}
%
%\renewcommand{\hat}{\widehat}
%\renewcommand{\tilde}{\widetilde}
\newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}{\rm #2}}}
\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
%\newcommand{\pdn}[3]{\frac{\partial^{#3}#1}{\partial#2^{#3}}}
%\newcommand{\snor}[1]{\left| #1 \right|}
%\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
%\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
%\newcommand{\LRb}[1]{\left| #1 \right|}
%\newcommand{\LRc}[1]{\left\{ #1 \right\}}
%\newcommand{\LRceil}[1]{\left\lceil #1 \right\rceil}
\newcommand{\LRl}[1]{\left. \LRp{#1} \right|}
\newcommand{\nor}[1]{\left\|#1\right\|}
%\newcommand{\jump}[1] {\ensuremath{\llbracket#1\rrbracket}}
%\newcommand{\avg}[1] {\ensuremath{\LRc{\!\{#1\}\!}}}
%\newcommand{\Grad} {\ensuremath{\nabla}}
%\renewcommand{\d}{\partial}
\newcommand{\diag}[1]{{\rm diag}\LRp{#1}}

\newcommand{\note}[1]{{\color{blue}{#1}}}
\newcommand{\bnote}[1]{{\color{blue}{#1}}}
\newcommand{\rnote}[1]{{\color{red}{#1}}}

\newcommand{\LK}{L^2\LRp{D^k}}
\newcommand{\LdK}{L^2\LRp{\partial D^k}}
\newcommand{\Dhat}{\widehat{D}}
\newcommand{\Lhat}{L^2\LRp{\Dhat}}

\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}} % d in integrand

\begin{document}

%\maketitle

\note{We thank the two Reviewers for their feedback, and describe steps taken to address reviewer comments and suggestions in the following response.  We have also addressed the typos described in the minor comments by each reviewer.  For brevity, we have not included typo and minor comment responses in this response.
\\
\\
Revisions in the manuscript are also colored for ease of identification.  We hope these revisions improve the readability of this paper and its suitability for the audience of SISC.}

\section{Reviewer 2}

We thank reviewer 2 for their careful reading of the paper.  In particular, we are grateful for the minor comments provided, which fixed several small typos and errors in the submitted manuscript.

\begin{itemize}
\item In equation (2), the entropy potential $\psi_i(u)$ is used but never defined. I also noticed that while the entropy function $S(u)$ is mentioned, the corresponding entropy flux is never mentioned. Is this intentional?

\bnote{We thank the reviewer for pointing this out.  We have added a definition of $\psi_i(u)$ after equation (2).  The reviewer is also correct that we intentionally do not mention the entropy flux.  This is because it factors into the continuous derivation of an entropy inequality, but not in the discrete derivation.  In the continuous derivation, the entropy flux factors into the entropy inequality via the chain rule, which does not hold discretely. }

\item In equation (7), a $\Delta x$ term seems to be missing before $\bm{K}\bm{u}_h$. Furthermore, since the viscous term is a discrete version of the Laplacian with periodic boundary conditions, it is a bit strange that the first and last rows of K seem to represent the discretization of the second order derivative when Neumann/open boundary conditions are considered. Could the author comment on this?

\bnote{We thank the reviewer for pointing this out, and have fixed it in the revision.  We have also clarified that the choice of Neumann boundary conditions is arbitrary, and can be replaced with the periodic Laplacian.  However, we have not noticed a significant difference in numerical experiments between the use of periodic boundary conditions and open/Neumann boundary conditions for the Laplacian.  We adopt Neumann boundary conditions for all experiments because they appear to work for both periodic and reflective wall boundary conditions.  
\\
\\
Future works will investigate different viscous boundary conditions (such as solid wall conditions) and their impacts on reduced order models.
}

\item  In the first line after Assumption 1 on page 8, the author states that ”the column space of $V$ is contained in the column space of $V_t$”. However, no connection between these two spaces has been made up to this point. I presume this is true because of the way $V_t$ is constructed, which is mentioned much later in the text. I would recommend minor rearrangements to make this fact clearer.

\bnote{We thank the reviewer for pointing this out.  We have added a statement on the previous page when $\bm{V}_t$ is introduced stating that ``In this work, we assume that the span of the test basis includes the reduced basis, e.g., $\mathcal{R}(\bm{V})\subset \mathcal{R}(\bm{V}_t)$.''}

\item  In the proof for Lemma 2, it was not clear to me how $\mathcal{R}(QV) \subset \mathcal{R}(V_t)$ implies $e_2 \in \mathcal{R}(V_t)$. Could the author explain this? If it is straightforward, it need not be added to the proof.

\bnote{We thank the reviewer for catching this - it was an error on our part.  We have replaced the lemma with another discussion of how the choice of test space affects accuracy.  }
%\bnote{Certainly.  $\bm{e}_2$ is the linear combination of $\bm{Q}{\bm{V}_t\bm{V}_t^{\dagger}}\bm{f}$ and $\LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{f}$.  The first term $\bm{Q}{\bm{V}_t\bm{V}_t^{\dagger}}\bm{f}$ is $\bm{Q}$ applied to some vector, so the result is in the range of $\bm{Q}$ (and thus in the range of $\bm{V}_t$ by the assumption).  The latter term is $\LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T$ multiplied by some vector, and since $\bm{V}_t\bm{V}_t^{\dagger}$ is a symmetric projection matrix, the latter term is in the range of $\bm{V}_t$.  }

\item In Algorithm 1, what is the advantage of performing a linear least square approximation for $w$, followed by a non-negative least squares if a negative weight is obtained?  Why not directly use non-negative least squares to begin with?

\bnote{This approach follows the approach taken in \cite{hernandez2017dimensional}, where they noted that for many test cases, the non-negative least squares (NNLS) solve was not necessary to guarantee positivity of weights.  Replacing the NNLS solve with a standard least squares solve ended up being noticably less expensive.  }

\item In Section 4.3, additional stabilization points are added to ensure that the test mass matrix is non-singular. But is it guaranteed that this approach used will always lead to such a matrix?

\bnote{We thank the reviewer for bringing up this question.  Because the algorithm is greedy, it will continue to add points so long as the test mass matrix is near singular, so the test mass matrix will eventually end up being non-singular (in the limiting case, the algorithm will simply choose all FOM points).  However, it is unclear how many points need to be added to avoid near-singularity.  The choice of $\alpha$ and use of the near-null space of the test mass matrix are thus heuristic approaches to reduce how many stabilizing points we need to add. }

\item In Section 4.5, the author mentions tricks to improve the computational costs when polynomial-type function are considered. However, the entropy conservative fluxes used for the Euler equations makes use of logarithmic terms. Then would such tricks be helpful in the context of the experiments presented in this manuscript?

\bnote{They unfortunately would not be helpful for the experiments in this work, at least not without giving up discrete entropy conservation (dissipation).  We have added a footnote explaining this:
\\
\\
``Because entropy conservative fluxes for the compressible Euler equations include non-polynomial rational and logarithmic terms, it is not possible to apply this exact treatment of polynomial nonlinearities to general entropy conservative discretizations.''}

\item The expression for the entropy variable on page 20 all seem off by a factor of $(\gamma -1)$ compared with the expressions in the paper by Chandrashekar [66]. My guess is that the expression in this manuscript may be correct for the choice of the entropy function $S(u) = -\rho s$. Could the author verify this?

\bnote{We thank the reviewer for catching this.  We have corrected the entropy function in the revised manuscript.}

\item What was the value of $\epsilon$ for the experiments shown in Figure 3. Furthermore, what is the difference between running the model in the absence of viscosity and running the model by setting $\epsilon = 0$, as mentioned in the first paragraph on page 22?

\bnote{We have added $\epsilon = 2e-4$ to the description.  We have also removed the sentence with $\epsilon = 0$, which the reviewer has correctly pointed out is redundant.}

\item To truly judge the benefits of using ROMs, it would be useful if the actual computational times for the full model, and the time for the offline+online computation of the ROM is provided for at least one of the 2D experiments.

\bnote{We agree with the reviewer's comments, but note that for the 2D problems considered, the proposed entropy stable reduced order models are more expensive than the full order model due to the use of explicit time-stepping.  This is described some in the conclusion of the original submission, but we have significantly expanded the discussion in the revised paper in a new section (see Section 7.3).  The discussion is quoted below:
\\
\\
\textit{We note that the entropy stable reduced order models implemented in this work do not reduce computational cost compared to the original full order models due to the use of explicit time-stepping.  Explicit time-stepping is used in this paper to validate the proposed semi-discrete formulation.  However, we expect that for implicit time-stepping, entropy stable ROMs will see more significant efficiency gains.  
\\
\\
The cost of explicit time-stepping schemes scales with the cost of the ODE right hand side evaluation, and the cost of a right hand side evaluation for an entropy stable scheme scales with the number of entropy conservative flux evaluations required.  The number of entropy conservative flux evaluations is the same as the number of nonzero entries in the matrix $\bm{Q}^i$, since the nonlinear term is $\LRp{\bm{Q}^i\circ\bm{F}^i}\bm{1}$ and the entries of the matrix $\bm{F}^i$ (which are flux evaluations between different states) are evaluated on the fly.  
\\
\\
For the 2D full order methods in this paper, the differentiation matrices $\bm{Q}^i$ are sparse with $O(K^2)$ non-zero entries.  Thus, the cost of a 2D entropy stable finite volume method is $O(K^2)$ nonlinear flux evaluations for each coordinate direction.  For a reduced order model, however, the hyper-reduced matrices $\bm{Q}^i_t$ are dense, and evaluating $(\bm{Q}^i_t \circ \bm{F}^i)\bm{1}$  requires $O(N_s^2)$ flux evaluations where $N_s$ is the number of hyper-reduced points.  As an example, consider the Kelvin-Helmholtz results in Figure~8.  The full order model requires $O(40000)$ flux evaluations.  However, because the ROM has $884$ hyper-reduced points, $O(781,456)$ flux evaluations are required, resulting in a ROM which is more expensive per time-step than the full order model.  
\\
\\
While the proposed ROMs could still provide savings for sufficiently large full order models, this increased cost is a significant issue for explicit time-stepping.  The cost is offset slightly if the ROM has a larger CFL \cite{marley2015reduced}; however, this situation only occurs if a smaller number of modes are used to simulate the solution.  The situation changes for implicit time-stepping, as full order models require multiple solutions of large (but sparse) linear systems at each time-step.  For a $K\times K$ grid, these linear systems would require $O(K^2)$ flux evaluations to assemble, but we expect the solution of the linear system to be the dominant cost.  For a ROM using implicit time-stepping, the assembly of the linear system still requires $O(N_s^2)$ flux evaluations, but the size of the system scales with the number of modes, which is notably smaller than even the number of hyper-reduced points.  For example, the Kelvin-Helmholtz example uses only $75$ basis functions per component, compared to $884$ hyper-reduced points.  We thus expect that the use of entropy stable ROMs will result in a more significant speedup under implicit time-stepping.
}
}
\item It would be interesting to see how the proposed ROM works for a more complex 2D problem. Perhaps the author could also test with a 2D Riemann problem for the Euler equations where shocks and/or contact discontinuities interact with each other. If discontinuous initial conditions is a problem, the author could choose a smooth approximation connecting the 4 states.

\bnote{
We have included the results from a smoothed 2D Riemann problem simulation in this reviewer response (see Figure~\ref{fig:riemann}).  The problem is taken from \cite{chan2017discretely}, and is modified to use periodic boundary conditions (due to the fact that ``natural'' boundary conditions are not provably entropy stable).  The FOM is run until time $T=.25$ on a $200\times 200$ grid with a CFL of $.25$ and $\epsilon = 5e-3$.  The ROM uses $50$ modes and $812$ hyper-reduced points, and achieves a final $L^2$ error of $0.03278$.  The ROM runs stably despite the presence of oscillations along the shock traveling to the north-east.  The solution appears reasonably well-approximated in smoother areas.
\\
\\
We would like to run this experiment with a larger number of modes to ensure that the ROM solution converges to the FOM.  However, because our implementation for these early experiments is not well-optimized, the hyper-reduction step becomes very slow as the number of modes increases.  For this reason, we have not included the example in the paper, but would be happy to do so if the reviewer believes this will improve the quality of the paper.
}
\begin{figure}
\centering
\subfloat[FOM results]{\includegraphics[width=.3\textwidth]{riemannFOM.png}}
\hspace{.25em}
\subfloat[ROM results]{\includegraphics[width=.3\textwidth]{riemannROM.png}}
\hspace{.25em}
\subfloat[Hyper-reduced points]{\includegraphics[width=.3\textwidth]{riemannROM_hrpts.png}}
\caption{Comparison of ROM and FOM behavior for a 2D Riemann problem.}
\label{fig:riemann}
\end{figure}

\item In the expressions for the viscous terms in Section 4.6, the viscosity coefficient seems to be missing. Or rather, it has been implicitly taken as unity. Is this intentional?  Is the viscous term in equation (19) missing a $V^T$?

\bnote{We thank the reviewer for catching these issues.  The reviewer is correct that these are typos, and we have fixed this in the revised paper.}

\end{itemize}

\section{Reviewer 3}

We thank reviewer 2 for their feedback on the submitted paper, and for bringing up issues facing reduced basis approximations of transport-dominated problems.

\begin{itemize}
\item My main question is the following: the author construct a POD decomposition using shnapshots as usual. What I do not understand is why this POD basis should be good for long term unsteady solution. I am talking about section 3 where everything is done at the semi-discrete level, so time is missing.

Assume that one shock appears, then I guess that the POD data base should take into account this. However, if the data base is constructed with many parameters, there is no reason why apriori one should know if there will be shocks or not, so how do you proceed in practice. What are the limitations?
I have the same question for unsteady problem in general: if you take
$u_t+u_x =0$ on $[0,1]$ with inflow conditions, and $u(x,0)=$ heaviside, then the number of POD is approximately the number of grid points.

Here on the contrary, you need few modes. is it because of the structure of your problems (for example the Burgers one has a symmetry as you mention), or of the numerical dissipation you are adding (but does it solve the advection problem)???  There is nothing in the text that helps understanding why this works.

\rnote{We thank the reviewer for pointing this out, and agree that the POD basis is not ideal for transport-dominated problems.  However, POD based approaches can still be used in the model reduction of transport-type equations so long as the solution retains some structure (either symmetry in the Burgers' example, or a quasi-periodic structure for the captive-carry problem in \cite{carlberg2013gnat}).  These approaches are not optimal, but are still useful to some practitioners.  To clarify these points, we have added a paragraph to the introduction describing these approximation challenges.  
\\
\\
``We note that this work focuses on classical model reduction techniques, which lose effectiveness for general transport-type phenomena \cite{reiss2018shifted, abgrall2018model, rim2018transport, cagniart2019model}.  This is tied to difficulties in approximating convected solutions using a fixed reduced basis.  Despite these challenges, classical approaches are still in model reduction of transport-type equations for specific problem setups \cite{carlberg2013gnat}.  In this work, we restrict ourselves to classical model reduction techniques.  Challenges associated with the low-dimensional approximation of transport-type solutions will be addressed in future work.'' }

\item In section 4, you should give more details. For example the paragraph starting with "weigts in section 4.2" till just before the section 4.1 is really sketchy.

\rnote{We apologize for the confusion in this paragraph.  We have re-written it in the revised version.  In particular, we have added the following paragraph to outline our approach to hyper-reduction:
\\
\\
''We briefly outline our approach to hyper-reduction.  Rather than directly apply hyper-reduction on the nonlinear vector, we perform a matrix-based hyper-reduction which respects the matrix structure of the nonlinear term $\bm{V}^T\LRp{\bm{Q}\circ \bm{F}}$.  Specifically, we construct a smaller hyper-reduced matrix $\bm{Q}_s$ and approximate the term $\bm{V}^T\LRp{\bm{Q}\circ \bm{F}}$ with $\bm{V}\LRp{\mathcal{I},:}^T\bm{W}\LRp{\bm{Q}_s\circ \bm{F}_s}$, where $\bm{F}_s$ is a smaller hyper-reduced matrix containing flux evaluations between solution states at different hyper-reduced points.''
\\
\\
Additional changes were also made to clarify this paragraph. }

\item In 4.1, how do you ensure in practice the assumption 1?

\rnote{We have added a sentence explaining that a heuristic algorithm for ensuring Assumption 1 is provided in a later section (Section 4.3).}

\item What is new in lemma 2: if $A$ is a linear operator, we all know that $||Ax-Ay|| \leq ||A||  ||x-y||$.  Since I imagine that you want to say something different, maybe the formulation should be changed.  In fact you confuse the reader: you show one thing $Qf-(VV^\dag)^TQ(VV^\dag) f=Q(f-(VV^\dag) f )$, then the inequality is trivial.

\rnote{We thank the reviewer for pointing this out.  We have replaced the lemma (which also contained an error) with a simpler discussion of why the choice of test space reproduces the action of $\bm{Q}$ on the reduced basis matrix $\bm{V}$.  }

\item In 4.2, hyper reduction, I do not see/understand how you take into account the time: are the hyperreduced modes good for ever, and why, or not ? what are the limits of the method.

\rnote{We have added two sentences to clarify this point.  In Section 3, we specify that ``We assume that the solution is well-approximated in the reduced basis over the entire time-window of the simulation''.  
\\
\\
In Section 4.2, we have added ``Because we assume a fixed reduced basis in time, we compute a single set of empirical cubature points prior to the simulation and use those same points over the entire duration of the simulation''.  We also suggest approaches which can be combined with our proposed methodology to address this challenge in future work.  Section 4.2 includes the sentence ``Again, we emphasize that solution snapshots for transport-type equations may not be well-approximated by a low-dimensional reduced basis.  Future work will attempt to combine techniques introduced in this work with methods to address this approximation issue \cite{reiss2018shifted, rim2018transport, cagniart2019model}.''
}

\item page 22 you write ``We now compare the evolution of both the L2 error and discrete entropy,'' the L2 error of what?

\rnote{We have clarified that this is the discrete $L^2$ error between all components of the full and reduced order solutions. }

\item One things that makes me a bit confused: if I understand well, the norm that you use for the POD is related to the entropy (so again  my question about time). and many times in the text, you talk about L2 norm, is it the one related to the entropy, for example $\int_0^T <x, A_0 x> dx$, where $A_0$ is the Hessian of the entropy evaluated at $U(.,t)$, or something else, and what.

\rnote{We have added the following sentence clarifying the definition of the $L^2$ norm: ``Here, the discrete $L^2$ norm is defined as $\nor{\bm{u}}^2 = \Delta x^2 \bm{u}^T\bm{u}$ in 1D or $\nor{\bm{u}}^2 = (\Delta x \Delta y)^2 \bm{u}^T\bm{u}$ in 2D, and is the analogue of the continuous $L^2$ norm evaluated at discretization points for the full order model.''}

\item your description of the finite difference scheme, with Hadamard product, is not very classical in the hyperbolic community. I am not saying it is wrong, I say it is not conventional. So the reading starts to be hard after page 3, in the sense we need a dictionary. This should improved.

\rnote{We thank the reviewer for pointing this out.  We have added some sentences which try to point out non-standard notation, and try to provide some references to papers where a similar notation is used.  For example, on page 3, we have added the sentence ``We note that this reformulation using the Hadamard product is non-standard within the finite volume literature, but is more common in the SBP finite difference literature.''}

\item page 5, formula 2, before (9): you should write you use a path to compute the average Jacobian, and do it componentwise, other wise it does not make much sens.  take  (classical counter example) $f(x)=e^{i\pi x}, f(1)-f(0)=-2, f'(x)=i\pi f(x) and -2 =i\pi f(\xi) * (1-0)$ is not possible because $|-2|=2=\pi |f(\xi)|=\pi.$

\rnote{This has been clarified with a footnote in the revised version.}

\end{itemize}

\bibliographystyle{unsrt}
\bibliography{refs}

\end{document}

### 2020-04-03 docs/ESDG_ROM/letter.tex
updating docs

--- 2017-10-02 docs/splitform/letter.tex
@@
 \begin{document}
-\begin{letter}{Dr Mikhail Shashkov\\JCP Editor}
+\begin{letter}{Dr Remi Abgrall\\JCP Editor}
 
-\opening{Dear Dr Shashkov,}
+\opening{Dear Dr Abgrall,}
 
@@
 Jesse Chan\\
-\textit{On discretely entropy conservative and entropy stable discontinuous Galerkin methods},
+\textit{Entropy stable reduced order modeling of nonlinear conservation laws},
 \end{center}
@@
 
-The submitted manuscript describes the construction of high order discontinuous Galerkin (DG) methods which satisfy a discrete conservation or dissipation of entropy.
-Existing entropy conservative and entropy stable high order DG methods have been constructed based on diagonal norm summation-by-parts operators, or equivalently nodal collocation and under-integrated (lumped) mass matrices.
-This work extends the construction of entropy conservative and entropy stable schemes to non-diagonal mass matrices (or dense norm SBP operators) and over-integrated quadrature rules.
-Numerical experiments confirm the entropy conservation and high order accuracy of the new methods based on over-integrated quadrature, and comparisons are made with existing entropy stable methods based on under-integration.
-
-In contrast to existing entropy conservative DG schemes, these new methods are formulated at the continuous level, and the proofs of entropy conservation rely only on properties of quadrature-based integration and $L^2$ projection.
-This continuous formulation also makes it possible to construct entropy conservative and entropy-stable methods using non-standard bases and approximation spaces.
+The submitted manuscript describes the construction projection-based reduced order models which are globally conservative and satisfy a discrete conservation or dissipation of entropy.
+The reduced models are constructed by combining appropriate hyper-reduction strategies with techniques from entropy stable finite volume and summation by parts (SBP) schemes.
+The novelty of the proposed approach is in the discretization of the convective nonlinearity, which adapts the ``flux differencing'' technique from SBP schemes to reduced order models.
 

### 2020-04-03 docs/ESDG_ROM/main_revision1.tex
updating docs

--- 2020-02-27 docs/ESDG_ROM/main.tex
@@
 \usepackage{graphicx, color, bm}
+
+\graphicspath{{./figs/}}
+%\graphicspath{{../figs/}}
+
 \usepackage{subfig}
@@
 \newcommand{\note}[1]{{\color{blue}{#1}}}
-
-\graphicspath{{./figs/}}
+\newcommand{\bnote}[1]{{\color{blue}{#1}}}
+\newcommand{\rnote}[1]{{\color{red}{#1}}}
 
@@
 
+\rnote{We note that this work focuses on classical model reduction techniques, which lose effectiveness for general transport-type phenomena \cite{reiss2018shifted, abgrall2018model, rim2018transport, cagniart2019model}.
+This is tied to difficulties in approximating convected solutions using a fixed reduced basis.
+Despite these challenges, classical approaches are still in model reduction of transport-type equations for specific problem setups \cite{carlberg2013gnat}.
+In this work, we restrict ourselves to classical model reduction techniques.
+Challenges associated with the low-dimensional approximation of transport-type solutions will be addressed in future work.
+}
+
 The paper is organized as follows.
@@
 \end{equation}
-where $n_i$ denotes the $i$th component of the outward normal vector.  
+where $n_i$ denotes the $i$th component of the outward normal vector, \bnote{and $\psi_i(\bm{u})$ denotes the entropy potential for the $i$th coordinate}.  
 The entropy inequality (\ref{eq:entropyineq}) is the analogue of energy stability for nonlinear conservation laws \cite{mock1980systems, harten1983symmetric}.  
@@
 For simplicity, we illustrate the construction of full and reduced order models on a periodic 1D domain.
-We note that, while the reduced order model will be constructed from an entropy stable scheme, the solution snapshots do not need to be generated by entropy stable full order model.
+We note that, while the reduced order model will be constructed from an entropy stable scheme, the solution snapshots do not need to be generated by an entropy stable full order model.
 Throughout this work, we will assume that the both full and reduced models yield ``physically relevant'' solutions, such that the entropy is a convex function.
@@
 \end{equation}
-Note that $\bm{Q}\bm{1} = \bm{0}$, that the matrix $\bm{F}$ is symmetric (due to symmetry of $\bm{f}_S$), and that the diagonal of $\bm{F}$ is equal to $\bm{f}(\bm{u})$ due to flux consistency.
+\bnote{Let $\bm{1}$ denote the vector of all ones.}  Note that $\bm{Q}\bm{1} = \bm{0}$, that the matrix $\bm{F}$ is symmetric (due to symmetry of $\bm{f}_S$), and that the diagonal of $\bm{F}$ is equal to $\bm{f}(\bm{u})$ due to flux consistency.
 For a scalar nonlinear conservation law, the matrix-based formulation of (\ref{eq:nonlineqs1d}) is then equivalent to
@@
 \end{align}
-where $\circ$ denotes the Hadamard product, $\bm{1}$ denotes the vector of all ones, and each entry of $\bm{u}_h$ corresponds to a point $\bm{x}_i$ in the domain.
+where $\circ$ denotes the Hadamard product, and each entry of $\bm{u}_h$ corresponds to a point $\bm{x}_i$ in the domain.
 In this setting, $\bm{Q}\circ\bm{F}$ extracts and takes linear combinations of nonlinear flux interactions between different nodal values of $(\bm{u}_h)_i$ and $(\bm{u}_h)_j$.
+\rnote{We note that this reformulation using the Hadamard product is non-standard within the finite volume literature, but is more common in the SBP finite difference literature}.
 The discretization of the nonlinear flux term using $\LRp{\bm{Q}\circ \bm{F}}\bm{1}$ is commonly referred to as \textit{flux differencing} \cite{carpenter2014entropy, gassner2016split, chen2017entropy, crean2018entropy, chan2017discretely}.\footnote{The factor of 1/2 present in the definition of $\bm{Q}$ is to make (\ref{eq:es}) consistent with the entropy stable SBP literature.
-The factor of 2 in (\ref{eq:es}) can also be derived from the chain rule \cite{chen2017entropy, crean2018entropy}.
-}
+The factor of 2 in (\ref{eq:es}) can also be derived from the chain rule \cite{chen2017entropy, crean2018entropy}.}
 
@@
 \label{thm:ecfom}
-Let $\bm{u}_h(t)$ be a solution of (\ref{eq:essys}) for which the entropy $S(\bm{u}_h)$ is convex at each time $t$.  Then, $\bm{u}_h$ satisfies the semi-discrete conservation of entropy 
+Let $\bm{u}_h(t)$ be a solution of (\ref{eq:essys}) for which the entropy $S(\bm{u}_h)$ is convex at each time $t$, \bnote{and let the flux $\bm{f}_S(\bm{u}_L,\bm{u}_R)$ be entropy conservative as defined by (\ref{eq:esflux})}.
+Then, $\bm{u}_h$ satisfies the semi-discrete conservation of entropy
 \[
@@
 \begin{align*}
-\sum_{ij} \LRp{\bm{v}_h}_i^T \bm{Q}_{ij} \bm{f}_{S}\LRp{\LRp{\bm{u}_h}_i, \LRp{\bm{u}_h}_j} = \sum_{ij} \LRp{\bm{Q}_{ij}-\bm{Q}_{ji}} \LRp{\bm{v}_h}_i^T  \bm{f}_{S}\LRp{\LRp{\bm{u}_h}_i, \LRp{\bm{u}_h}_j}.
+\sum_{ij} \LRp{\bm{v}_h}_i^T \bnote{2}\bm{Q}_{ij} \bm{f}_{S}\LRp{\LRp{\bm{u}_h}_i, \LRp{\bm{u}_h}_j} = \sum_{ij} \LRp{\bm{Q}_{ij}-\bm{Q}_{ji}} \LRp{\bm{v}_h}_i^T  \bm{f}_{S}\LRp{\LRp{\bm{u}_h}_i, \LRp{\bm{u}_h}_j}.
 \end{align*}
@@
 \begin{equation}
-\Delta x \td{\bm{u}_h}{t} + 2\LRp{{\bm{Q}}\circ \bm{F}}\bm{1} + \epsilon \bm{K}\bm{u}_h= 0, \qquad \bm{K} = \frac{1}{\Delta x^2} \begin{bmatrix}
+\Delta x \td{\bm{u}_h}{t} + 2\LRp{{\bm{Q}}\circ \bm{F}}\bm{1} + \epsilon \bm{K}\bm{u}_h= 0, \qquad \bm{K} = \frac{1}{\bnote{\Delta x}} \begin{bmatrix}
 1 & -1 & & \\
@@
 Future work will analyze both physically relevant viscosities and more nuanced artificial dissipation mechanisms.
+
+\begin{remark}
+\bnote{The choice of Neumann boundary conditions for the Laplacian matrix $\bm{K}$ is arbitrary, and $\bm{K}$ can be replaced with the periodic Laplacian.
+Future works will investigate different viscous boundary conditions (such as solid wall conditions) and their impacts on reduced order models.}
+\end{remark}
 
@@
 \]
-where $\LRp{\pd{\bm{u}}{\bm{v}}}_{i,i-1}$ is the Jacobian matrix evaluated at some state between $(\bm{u}_h)_i$ and $(\bm{u}_h)_j$.
-Since the Jacobian is related to the Hessian of the entropy through $\pdn{S(\bm{u})}{\bm{u}}{2} = \pd{\bm{v}}{\bm{u}} = \LRp{\pd{\bm{u}}{\bm{v}}}^{-1}$, it is positive definite so long as $S(\bm{u})$ is convex.
+where $\LRp{\pd{\bm{u}}{\bm{v}}}_{i,i-1}$ is the Jacobian matrix evaluated at some state between $(\bm{u}_h)_i$ and $(\bm{u}_h)_{\bnote{i-1}}$.\footnote{\rnote{We note that this ``average'' Jacobian is evaluated componentwise via a path integral \cite{tadmor2003entropy}.}}  Since the Jacobian is related to the Hessian of the entropy through $\pdn{S(\bm{u})}{\bm{u}}{2} = \pd{\bm{v}}{\bm{u}} = \LRp{\pd{\bm{u}}{\bm{v}}}^{-1}$, it is positive definite so long as $S(\bm{u})$ is convex.
 Substituting these expressions into the viscous terms and using positive definiteness of $\pd{\bm{u}}{\bm{v}}$  yields
@@
 \]
-We approximate grid values of the solution by $\bm{u}_h = \bm{V}\bm{u}_N$, where $\bm{u}_N$ denote ``modal'' coefficients  of the solution in the reduced basis.
+\rnote{We assume that the solution is well-approximated in the reduced basis over the entire time-window of the simulation, } and approximate grid values of the solution by $\bm{u}_h = \bm{V}\bm{u}_N$.
+Here, $\bm{u}_N$ denote ``modal'' coefficients  of the solution in the reduced basis.
+\rnote{Again, we emphasize that solution snapshots for transport-type equations may not be well-approximated by a low-dimensional reduced basis.
+Future work will attempt to combine techniques introduced in this work with methods to address this approximation issue \cite{reiss2018shifted, rim2018transport, cagniart2019model}.}
+
 Ignoring viscosity terms for now, plugging this expression into (\ref{eq:es}) and enforcing that the residual is orthogonal to all columns of $\bm{V}$ (Galerkin projection) yields a reduced system
@@
 \[
-\bm{V}^Tg(\bm{V}\bm{u}_N) \approx \bm{V}\LRp{\mathcal{I},:}^T \bm{W} g\LRp{\bm{V}\LRp{\mathcal{I},:}}.
+\bm{V}^Tg(\bm{V}\bm{u}_N) \approx \bm{V}\LRp{\mathcal{I},:}^T \bm{W} g\LRp{\bm{V}\LRp{\mathcal{I},:}\bnote{\bm{u}_N}}.
 \]
@@
 
-We will hyper-reduce the nonlinear term $\LRp{\bm{Q}\circ \bm{F}}$ by constructing a hyper-reduced sampled matrix $\bm{Q}_s$.
-Recall that the proof of entropy conservation in Theorem~\ref{thm:ecfom} required that matrix $\bm{Q}$ is skew symmetric and has zero row sums.
-If the hyper-reduced matrix $\bm{Q}_s$ also satisfies the properties
+\rnote{We briefly outline our approach to hyper-reduction.
+Rather than directly apply hyper-reduction on the nonlinear vector, we perform a matrix-based hyper-reduction which respects the matrix structure of the nonlinear term $\bm{V}^T\LRp{\bm{Q}\circ \bm{F}}$.
+Specifically, we construct a smaller hyper-reduced matrix $\bm{Q}_s$ and approximate the term $\bm{V}^T\LRp{\bm{Q}\circ \bm{F}}$ with $\bm{V}\LRp{\mathcal{I},:}^T\bm{W}\LRp{\bm{Q}_s\circ \bm{F}_s}$, where $\bm{F}_s$ is a smaller hyper-reduced matrix containing flux evaluations between solution states at different hyper-reduced points.
+Recall that the proof of entropy conservation for the full order model in Theorem~\ref{thm:ecfom} required only that the matrix $\bm{Q}$ be skew symmetric and have zero row sum.
+} If the hyper-reduced matrix $\bm{Q}_s$ also satisfies those properties
 \[
@@
 \]
-then the proof of entropy conservation in Theorem~\ref{thm:esrom} will carry over to the hyper-reduced model.
-Unfortunately, common hyper-reduction techniques preserve either skew-symmetry or zero row sums, but not both.
-For example, $\bm{Q}$ can be decomposed into the sum of local skew-symmetric matrices.
-The hyper-reduction techniques of \cite{farhat2015structure, yano2019discontinuous} approximate $\bm{Q}$ using a sparse linear combination of those local matrices.
+then we will be able to extend the proof of entropy conservation in Theorem~\ref{thm:esrom} to the hyper-reduced model.
+
+Unfortunately, common hyper-reduction techniques \rnote{applied naively to the matrix $\bm{Q}$} preserve either skew-symmetry or zero row sums, but not both.
+For example, $\bm{Q}$ can be decomposed into the sum of local skew-symmetric matrices, and the hyper-reduction techniques of \cite{farhat2015structure, yano2019discontinuous} approximate $\bm{Q}$ using a sparse linear combination of those local matrices.
 This hyper-reduction technique preserves skew-symmetry, but does not necessarily preserve the zero row sum property.
 Similarly, techniques such as gappy POD or empirical interpolation \cite{everson1995karhunen, barrault2004empirical, chaturantabut2010nonlinear} may preserve the zero row sum property but not skew-symmetry.
+\rnote{The next section describes a two-step hyper-reduction which retains both skew-symmetry and zero row sums of $\bm{Q}_s$ (though at the cost of sparsity).}
 
@@
 Let $\bm{V}_t$ denote a \textit{test} basis through which to extract the action of $\bm{Q}$.
+\bnote{In this work, we assume that the span of the test basis includes the reduced basis, e.g., $\mathcal{R}(\bm{V})\subset \mathcal{R}(\bm{V}_t)$}.
 We define the intermediate reduced operator
@@
 \end{assumption}
-Note that if Assumption~\ref{ass:quad} holds, then the mass matrix $\bm{V}\LRp{\mathcal{I},:}^T\bm{W}\bm{V}\LRp{\mathcal{I},:}$ is also non-singular, since the column space of $\bm{V}$ is contained in the column space of $\bm{V}_t$.
+Note that if Assumption~\ref{ass:quad} holds, then the test mass matrix $\bm{V}\LRp{\mathcal{I},:}^T\bm{W}\bm{V}\LRp{\mathcal{I},:}$ is also non-singular, since the column space of $\bm{V}$ is contained in the column space of $\bm{V}_t$.
+\rnote{We will provide a heuristic algorithm for ensuring that the hyper-reduced test mass matrix is non-singular in Section~\ref{sec:condtest}.}
 
@@
 The projection matrix $\bm{P}_t$ maps values at hyper-reduced points to coefficients in the test basis, and is well-defined under Assumption~\ref{ass:quad}.
-The key property of $\bm{P}_t$ which we will utilize is as follows: suppose that $\bm{f} = \bm{V}_t\bm{c}$ for some coefficients $\bm{c}$.
+The key property of $\bm{P}_t$ which we will utilize is as follows: suppose that $\bm{f} = \bm{V}_t\bnote{\LRp{\mathcal{I},:}}\bm{c}$ for some coefficients $\bm{c}$.
 Then,
@@
 \end{equation}
-In other words, if a vector lies in the range of $\bm{V}_t$, the projection matrix $\bm{P}_t$ recovers the coefficients exactly.
-We also note that if the hyper-reduced set $\mathcal{I}$ is taken to include all the points and $\bm{W}$ is a multiple of the identity (for example, $\bm{W} = \Delta x \bm{I}$), then $\bm{P}_t$ reduces to the pseudo-inverse $\bm{V}_t^{\dagger} = \LRp{\bm{V}_t\bm{V}_t}^{-1}\bm{V}_t^T$.
+In other words, if a vector lies in the range of $\bm{V}_t\LRp{\mathcal{I},:}$, the projection matrix $\bm{P}_t$ recovers the coefficients exactly.
+We also note that if the hyper-reduced set $\mathcal{I}$ is taken to include all the points and $\bm{W}$ is a multiple of the identity (for example, $\bm{W} = \Delta x \bm{I}$), then $\bm{P}_t$ reduces to the pseudo-inverse $\bm{V}_t^{\dagger} = \LRp{\bm{V}_t\bnote{^T}\bm{V}_t}^{-1}\bm{V}_t^T$.
 
@@
 We have not yet specified the choice of test basis $\bm{V}_t$.
-The span of the test basis should include the reduced basis $\bm{V}$, and Lemma~\ref{thm:Qt} implies that the span of the test basis should also contain $\bm{1}$.
+The span of the test basis should include the reduced basis $\bm{V}$, and Lemma~\ref{thm:Qt} \bnote{assumes} that the span of the test basis should also contain $\bm{1}$.
 However, these two conditions are insufficient to ensure accuracy for solutions of nonlinear conservation laws.
@@
 \centering
-\subfloat[Solution snapshots]{\includegraphics[width=.32\textwidth]{figs/burg_snaps.png}}
+\subfloat[Solution snapshots]{\includegraphics[width=.32\textwidth]{burg_snaps.png}}
 \hspace{.1em}
-\subfloat[Modes (columns of $\bm{V}$)]{\includegraphics[width=.32\textwidth]{figs/burg3modes.png}}
+\subfloat[Modes (columns of $\bm{V}$)]{\includegraphics[width=.32\textwidth]{burg3modes.png}}
 \hspace{.1em}
-\subfloat[Mode derivatives $\bm{Q}\bm{V}$]{\includegraphics[width=.32\textwidth]{figs/burg3modes_diff.png}}
+\subfloat[Mode derivatives $\bm{Q}\bm{V}$]{\includegraphics[width=.32\textwidth]{burg3modes_diff.png}}
 \caption{Solution snapshots, three POD modes, and mode derivatives for a shock solution of the periodic Burgers' equation. }
@@
 
-We comment on why this choice of test basis restores accuracy.  
-In the absence of hyper-reduction (where $\bm{P}_t = \bm{V}_t^{\dagger}$), the matrix $\bm{Q}_t$ should be an accurate approximation to $\bm{Q}$.
-The following lemma quantifies the relationship between accuracy in the $2$-norm and choice of test space.
-%The entropy conservative flux for Burgers' equation corresponds to the discretization of the split form \cite{canuto2006spectral, chan2017discretely, maboudi2018conservative}
+We briefly comment on why this choice of test basis restores accuracy. 
+\bnote{The matrix $\bm{Q}_t$ should at minimum reproduce the action of $\bm{Q}$ on the reduced basis $\bm{V}$.  In the absence of hyper-reduction (where $\bm{P}_t = \bm{V}_t^{\dagger}$), we have that
+\[
+\bm{Q}_t \bm{V} = \LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{V} = \bm{Q}\bm{V}, 
+\]
+where we have used that ${\bm{V}_t\bm{V}_t^{\dagger}}$ is a symmetric projector onto $\mathcal{R}(\bm{V}_t) \supset \mathcal{R}(\bm{V}) \cup \mathcal{R}(\bm{QV})$.
+}  
+%In the absence of hyper-reduction (where $\bm{P}_t = \bm{V}_t^{\dagger}$), the matrix $\bm{Q}_t$ should be an accurate approximation to $\bm{Q}$.  
+%The following lemma quantifies the relationship between accuracy in the $2$-norm and choice of test space.  
+%\begin{lemma}
+%\label{thm:acc}
+%Let $\bm{f} \in \mathcal{R}\LRp{\bm{V}}$ denote a vector in the span of the reduced basis.  Then, if $\mathcal{R}\LRp{\bm{Q}\bm{V}} \subset \mathcal{R}\LRp{\bm{V}_t}$, then 
+%\begin{align*}
+%\bm{Q}\bm{f} = \LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{f} = \bm{0} % \nor{\bm{Q}\LRp{\bm{f}-\bm{V}_t\bm{V}_t^{\dagger}\bm{f}}}.
+%\leq \nor{\bm{Q}}\nor{\bm{f}-\bm{V}_t\bm{V}_t^{\dagger}\bm{f}}.
+%\end{align*}
+%%Let $\bm{f},\bm{g}$ denote vectors, and let $\bm{u}_1, \bm{u}_2$ denote solutions of the following problems
+%%\begin{align*}
+%%\bm{V}^T\bm{V}\bm{u}_1 &= \bm{V}^T \diag{\bm{f}}\bm{Q}\bm{g}\\
+%%\bm{V}^T\bm{V}\bm{u}_2 &= \bm{V}^T \diag{\bm{f}} \LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{g}.
+%%\end{align*}
+%%Let $\bm{f}_{\bm{v}} = {\rm diag}\LRp{\bm{f}}\bm{v}$.  Then, the error $\bm{e} = \bm{u}_1-\bm{u}_2$ satisfies
+%%\[
+%%\nor{\bm{V}\bm{e}} \leq \nor{\bm{Q}} \nor{\bm{V}^{\dagger}}\LRp{\nor{\bm{f}}_{\infty} + \nor{\bm{g}}} \LRp{\nor{\bm{g}-\bm{V}_t\bm{V}_t^{\dagger}\bm{g}} + \max_{\bm{v}\in \mathcal{R}(\bm{V})} \nor{\bm{f}_{\bm{v}} -\bm{V}_t\bm{V}_t^\dagger \bm{f}_{\bm{v}}} }.
+%%\]
+%\end{lemma}
+%\begin{proof}
+%Define $\bm{e} = \bm{Q}\bm{f} - \LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{f}$.  Adding and subtracting $\bm{Q}\LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{f}$ yields
 %\[
-%\pd{u}{t} + \frac{1}{3}\LRp{\pd{u^2}{x}  + u\pd{u}{x}} = 0.
+%\bm{e} = \bm{Q}\LRp{\bm{f} - \LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{f}} + \underbrace{\bm{Q}{\bm{V}_t\bm{V}_t^{\dagger}}\bm{f} - \LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{f}}_{\bm{e}_2}.
 %\]
-%An entropy conservative ROM without hyper-reduction is then given by 
+%Since $\mathcal{R}\LRp{\bm{Q}\bm{V}} \subset \mathcal{R}\LRp{\bm{V}_t}$ \bnote{and $\bm{V}_t\bm{V}_t^{\dagger}$ is a symmetric projection matrix onto the span of $\bm{V}_t$}, $\bm{e}_2\in \mathcal{R}\LRp{\bm{V}_t}$ and $\bm{V}_t\bm{V}_t^\dagger\bm{e}_2 = \bm{e}_2$.
+Thus, $\bm{e}_2^T\bm{e}_2 = \nor{\bm{e}_2}^2$ gives
 %\[
-%\Delta x\bm{V}^T\bm{V}\td{u}{t} + \frac{1}{3}\bm{V}^T\LRp{\bm{Q}\bm{u}^2  + {\rm diag}\LRp{\bm{u}}\bm{Q}\bm{u}} = \bm{0}, \qquad \bm{u} = \bm{V}\bm{u}_N.
+%\nor{\bm{e}_2}^2 = \bm{e}_2^T\bm{Q}{\bm{V}_t\bm{V}_t^{\dagger}}\bm{f} - \bm{e}_2^T\LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{f} 
+%= \bm{e}_2^T\bm{Q}{\bm{V}_t\bm{V}_t^{\dagger}}\bm{f} - \bm{e}_2^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{f} = \bm{0}.
 %\]
-%It was shown in \cite{chan2017discretely, chenreview} that, for certain entropy conservative fluxes, the flux differencing procedure is equivalent in discretization of the split form.
-Thus, to ensure accurate evalution of the spatial formulation, the reduced differentiation matrix should at least accurately approximate mixed derivatives of the form ${\rm diag}(\bm{f})\bm{Q}\bm{g}$.
-We assume no hyper-reduction is performed, such that $\bm{P}_t = \bm{V}_t^{\dagger}$.
-\begin{lemma}
-\label{thm:acc}
-Let $\bm{f}$ denote a vector.  Then, if $\mathcal{R}\LRp{\bm{Q}\bm{V}} \subset \mathcal{R}\LRp{\bm{V}_t}$, then
-\begin{align*}
-\nor{\bm{Q}\bm{f} - \LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{f}} = \nor{\bm{Q}\LRp{\bm{f}-\bm{V}_t\bm{V}_t^{\dagger}\bm{f}}} \leq \nor{\bm{Q}}\nor{\bm{f}-\bm{V}_t\bm{V}_t^{\dagger}\bm{f}}.
-\end{align*}
-%Let $\bm{f},\bm{g}$ denote vectors, and let $\bm{u}_1, \bm{u}_2$ denote solutions of the following problems
-%\begin{align*}
-%\bm{V}^T\bm{V}\bm{u}_1 &= \bm{V}^T \diag{\bm{f}}\bm{Q}\bm{g}\\
-%\bm{V}^T\bm{V}\bm{u}_2 &= \bm{V}^T \diag{\bm{f}} \LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{g}.
-%\end{align*}
-%Let $\bm{f}_{\bm{v}} = {\rm diag}\LRp{\bm{f}}\bm{v}$.  Then, the error $\bm{e} = \bm{u}_1-\bm{u}_2$ satisfies
-%\[
-%\nor{\bm{V}\bm{e}} \leq \nor{\bm{Q}} \nor{\bm{V}^{\dagger}}\LRp{\nor{\bm{f}}_{\infty} + \nor{\bm{g}}} \LRp{\nor{\bm{g}-\bm{V}_t\bm{V}_t^{\dagger}\bm{g}} + \max_{\bm{v}\in \mathcal{R}(\bm{V})} \nor{\bm{f}_{\bm{v}} -\bm{V}_t\bm{V}_t^\dagger \bm{f}_{\bm{v}}} }.
-%\]
-\end{lemma}
-\begin{proof}
-Define $\bm{e} = \bm{Q}\bm{f} - \LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{f}$.  Adding and subtracting $\bm{Q}\LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{f}$ yields
-\[
-\bm{e} = \bm{Q}\LRp{\bm{f} - \LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{f}} + \underbrace{\bm{Q}{\bm{V}_t\bm{V}_t^{\dagger}}\bm{f} - \LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{f}}_{\bm{e}_2}.
-\]
-Since $\mathcal{R}\LRp{\bm{Q}\bm{V}} \subset \mathcal{R}\LRp{\bm{V}_t}$, $\bm{e}_2\in \mathcal{R}\LRp{\bm{V}_t}$ and $\bm{V}_t\bm{V}_t^\dagger\bm{e}_2 = \bm{e}_2$.
-Thus, $\bm{e}_2^T\bm{e}_2 = \nor{\bm{e}_2}^2$ gives
-\[
-\nor{\bm{e}_2}^2 = \bm{e}_2^T\bm{Q}{\bm{V}_t\bm{V}_t^{\dagger}}\bm{f} - \bm{e}_2^T\LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{f} 
-= \bm{e}_2^T\bm{Q}{\bm{V}_t\bm{V}_t^{\dagger}}\bm{f} - \bm{e}_2^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{f} = \bm{0}.
-\]
-%We can derive the estimate by adding and subtracting $\diag{\bm{f}}\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{g}$.  Straightforward manipulations then give 
-%\begin{align*}
-%\bm{V}^T\bm{V}\bm{e} = \LRp{\bm{V}^T\diag{\bm{f}}\bm{Q}\LRp{\bm{g}-\bm{V}_t\bm{V}_t^{\dagger}\bm{g}} + \LRp{\LRp{\bm{I} - \bm{V}_t\bm{V}_t^{\dagger}}\diag{\bm{f}}\bm{V}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{g}}
-%\end{align*}
-%Multiplying by $\bm{e}^T$ on both sides and using the triangle inequality yields
-%\begin{align*}
-%\nor{\bm{V}\bm{e}}^2 &\leq 
-%\nor{\bm{e}^T\bm{V}^T\diag{\bm{f}}\bm{Q}\LRp{\bm{g}-\bm{V}_t\bm{V}_t^{\dagger}\bm{g}}} \\
-%&+ \nor{\LRp{\LRp{\bm{I} - \bm{V}_t\bm{V}_t^{\dagger}}\diag{\bm{f}}\bm{V}\bm{e}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{g}}.
-%\end{align*}
-%The first term is bounded by Cauchy-Schwarz and the fact that $\nor{\diag{\bm{f}}} = \nor{\bm{f}}_{\infty}$
-%\[
-%\nor{\bm{e}^T\bm{V}^T\diag{\bm{f}}\bm{Q}\LRp{\bm{g}-\bm{V}_t\bm{V}_t^{\dagger}\bm{g}}}\leq \nor{\bm{V}\bm{e}} \nor{\bm{f}}_{\infty} \nor{\bm{Q}} \nor{\bm{g}-\bm{V}_t\bm{V}_t^{\dagger}\bm{g}}.
-%\]
-%The second term is bounded similarly.  Using that $\bm{V}^{\dagger}\bm{V} = \bm{I}$, 
-%\begin{align*}
-%&\nor{\LRp{\LRp{\bm{I} - \bm{V}_t\bm{V}_t^{\dagger}}\diag{\bm{f}}\bm{V}\bm{e}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{g}}  \\
-%&\leq\nor{\LRp{\LRp{\bm{I} - \bm{V}_t\bm{V}_t^{\dagger}}\diag{\bm{f}}\bm{V}}} \nor{\bm{V}^{\dagger}}\nor{\bm{V}\bm{e}}\nor{\bm{Q}}\nor{\bm{V}_t \bm{V}_t^{\dagger}}\nor{\bm{g}} \\
-%&\leq \max_{\bm{v}\in \mathcal{R}(\bm{V})} \nor{\bm{f}_{\bm{v}} -\bm{V}_t\bm{V}_t^\dagger \bm{f}_{\bm{v}}}\nor{\bm{V}^{\dagger}} \nor{\bm{V}\bm{e}}\nor{\bm{Q}}\nor{\bm{g}} 
-%\end{align*}
-%where we have used that $\nor{\bm{V}^{\dagger}} \leq 1$ since $\bm{V}$ is orthonormal and $\nor{\bm{V}_t \bm{V}_t^{\dagger}} \leq 1$ since the normed quantity is an orthogonal projection.
+%%We can derive the estimate by adding and subtracting $\diag{\bm{f}}\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{g}$.  Straightforward manipulations then give 
+%%\begin{align*}
+%%\bm{V}^T\bm{V}\bm{e} = \LRp{\bm{V}^T\diag{\bm{f}}\bm{Q}\LRp{\bm{g}-\bm{V}_t\bm{V}_t^{\dagger}\bm{g}} + \LRp{\LRp{\bm{I} - \bm{V}_t\bm{V}_t^{\dagger}}\diag{\bm{f}}\bm{V}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{g}}
+%%\end{align*}
+%%Multiplying by $\bm{e}^T$ on both sides and using the triangle inequality yields
+%%\begin{align*}
+%%\nor{\bm{V}\bm{e}}^2 &\leq 
+%%\nor{\bm{e}^T\bm{V}^T\diag{\bm{f}}\bm{Q}\LRp{\bm{g}-\bm{V}_t\bm{V}_t^{\dagger}\bm{g}}} \\
+%%&+ \nor{\LRp{\LRp{\bm{I} - \bm{V}_t\bm{V}_t^{\dagger}}\diag{\bm{f}}\bm{V}\bm{e}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{g}}.
+%%\end{align*}
+%%The first term is bounded by Cauchy-Schwarz and the fact that $\nor{\diag{\bm{f}}} = \nor{\bm{f}}_{\infty}$
+%%\[
+%%\nor{\bm{e}^T\bm{V}^T\diag{\bm{f}}\bm{Q}\LRp{\bm{g}-\bm{V}_t\bm{V}_t^{\dagger}\bm{g}}}\leq \nor{\bm{V}\bm{e}} \nor{\bm{f}}_{\infty} \nor{\bm{Q}} \nor{\bm{g}-\bm{V}_t\bm{V}_t^{\dagger}\bm{g}}.
+%%\]
+%%The second term is bounded similarly.  Using that $\bm{V}^{\dagger}\bm{V} = \bm{I}$, 
+%%\begin{align*}
+%%&\nor{\LRp{\LRp{\bm{I} - \bm{V}_t\bm{V}_t^{\dagger}}\diag{\bm{f}}\bm{V}\bm{e}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}\bm{g}}  \\
+%%&\leq\nor{\LRp{\LRp{\bm{I} - \bm{V}_t\bm{V}_t^{\dagger}}\diag{\bm{f}}\bm{V}}} \nor{\bm{V}^{\dagger}}\nor{\bm{V}\bm{e}}\nor{\bm{Q}}\nor{\bm{V}_t \bm{V}_t^{\dagger}}\nor{\bm{g}} \\
+%%&\leq \max_{\bm{v}\in \mathcal{R}(\bm{V})} \nor{\bm{f}_{\bm{v}} -\bm{V}_t\bm{V}_t^\dagger \bm{f}_{\bm{v}}}\nor{\bm{V}^{\dagger}} \nor{\bm{V}\bm{e}}\nor{\bm{Q}}\nor{\bm{g}} 
+%%\end{align*}
+%%where we have used that $\nor{\bm{V}^{\dagger}} \leq 1$ since $\bm{V}$ is orthonormal and $\nor{\bm{V}_t \bm{V}_t^{\dagger}} \leq 1$ since the normed quantity is an orthogonal projection.
 Dividing through by $\nor{\bm{V}\bm{e}}$ and an $\ell_2$ inequality complete the proof.
 We note that if $\bm{V}$ is orthonormal, then $\nor{\bm{V}^{\dagger}}\leq 1$.
- \end{proof}
- In other words, if $\bm{Q}\bm{V}$ is contained within the range of $\bm{V}_t$, the error in approximating $\bm{Q}\bm{f}$ using the compress-and-projected matrix $\LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}$ scales with the best approximation error $\nor{\bm{f}-\bm{V}_t\bm{V}_t^{\dagger}\bm{f}}$.
+% \end{proof}
+% In other words, if $\bm{Q}\bm{V}$ is contained within the range of $\bm{V}_t$, the error in approximating $\bm{Q}\bm{f}$ using the compress-and-projected matrix $\LRp{\bm{V}_t\bm{V}_t^{\dagger}}^T\bm{Q}\bm{V}_t\bm{V}_t^{\dagger}$ scales with the best approximation error $\nor{\bm{f}-\bm{V}_t\bm{V}_t^{\dagger}\bm{f}}$.
 
@@
 In this work, we utilize the greedy algorithm for computing empirical cubature points and weights from \cite{an2008optimizing, hernandez2017dimensional}, which is described in Algorithm~\ref{alg:hr}.
+\rnote{Because we assume a fixed reduced basis in time, we compute a single set of empirical cubature points prior to the simulation and use those same points over the entire duration of the simulation.}
 Let $\bm{V}_{\rm target}$ be some matrix whose columns span some target space of functions to be integrated, and let $\bm{w}_{\rm target}$ denote some reference target weights.
@@
 \item Compute a reduced basis matrix $\bm{V}$ from snapshots of both conservative variables and entropy variables.  
-\item Use the ``test'' basis matrix $\bm{V}_t$ to construct a ``modal'' differentiation matrix $\hat{\bm{Q}}_t = \bm{V}_t^T\bm{Q}\bm{V}_t$.
+\item \bnote{Compute a ``test'' basis matrix $\bm{V}_t$ such that $\mathcal{R}\LRp{\bm{V}_t} = \mathcal{R}\LRp{\LRs{\bm{1}, \bm{V},\bm{Q}\bm{V}}}$}, and use $\bm{V}_t$ to construct a ``modal'' differentiation matrix $\hat{\bm{Q}}_t = \bm{V}_t^T\bm{Q}\bm{V}_t$.
 \item Compute a hyper-reduced quadrature using Algorithm~\ref{alg:hr}, adding stabilizing points as necessary to ensure that the test mass matrix $\bm{V}_t\LRp{\mathcal{I},:}^T\bm{W}\bm{V}_t\LRp{\mathcal{I},:}$ is non-singular.
@@
 For the full order model, this cost is typically ameliorated by exploiting the sparsity of $\bm{Q}$.
-However, since the hyper-reduced matrix $\bm{Q}_t$ is fully dense, the nonlinear convective term $\bm{Q}\circ\bm{F}$ in (\ref{eq:hrrom}) requires $O(N_s^2)$ nonlinear flux evaluations.
+However, since the hyper-reduced matrix $\bm{Q}_t$ is fully dense, the nonlinear convective term $\bm{Q}_{\bnote{t}}\circ\bm{F}$ in (\ref{eq:hrrom}) requires $O(N_s^2)$ nonlinear flux evaluations.
 In contrast, typical hyper-reduction approaches target a nonlinear \textit{vector}, which requires only $O(N_s)$ nonlinear function evaluations.
@@
 which requires $N^2$ matrix-vector multiplications, or $O(N^4)$ operations.
-Generalizing this procedure to $m > 3$ yields that a degree $m$ polynomial nonlinearity can be evaluated in $O(N^{m+1})$ operations.
+Generalizing this procedure to $m > 3$ yields that a degree $m$ polynomial nonlinearity can be evaluated in $O(N^{m+1})$ operations.\footnote{\bnote{Because entropy conservative fluxes for the compressible Euler equations include non-polynomial rational and logarithmic terms, it is not possible to apply this exact treatment of polynomial nonlinearities to general entropy conservative discretizations.
+}}
 
@@
 \[
-\bm{M}_N\td{\bm{u}_N}{t} + 2\bm{V}\LRp{\mathcal{I},:}^T\LRp{\bm{Q}_t \circ \bm{F}}\bm{1} + \bm{d}(\bm{u}_N)= \bm{0}
+\bm{M}_N\td{\bm{u}_N}{t} + 2\bm{V}\LRp{\mathcal{I},:}^T\LRp{\bm{Q}_t \circ \bm{F}}\bm{1} + \bnote{\epsilon}\bm{d}(\bm{u}_N)= \bm{0}
 \]
@@
 \[
-\Delta x \bm{1}^T\td{S\LRp{\bm{V}\bm{u}_N}}{t} = -\bm{v}_N^T\bm{d}(\bm{u}_N), \qquad \bm{v}_N = \bm{P}\bm{v}\LRp{\bm{V}\LRp{\mathcal{I},:}\bm{u}_N}.
+\Delta x \bm{1}^T\td{S\LRp{\bm{V}\bm{u}_N}}{t} = -\bnote{\epsilon}\bm{v}_N^T\bm{d}(\bm{u}_N), \qquad \bm{v}_N = \bm{P}\bm{v}\LRp{\bm{V}\LRp{\mathcal{I},:}\bm{u}_N}.
 \]
@@
 %\bm{K}\bm{u} \approx 
-\bm{d}(\bm{u}_N) = \bm{D}\LRp{\mathcal{I}_{\bm{D}},:}^T \bm{W}_{\bm{D}} \bm{D}\LRp{\mathcal{I}_{\bm{D}},:}\tilde{\bm{u}}.
+\bm{d}(\bm{u}_N) =\bnote{\bm{V}^T} \bm{D}\LRp{\mathcal{I}_{\bm{D}},:}^T \bm{W}_{\bm{D}} \bm{D}\LRp{\mathcal{I}_{\bm{D}},:}\tilde{\bm{u}}.
 \label{eq:visc1}
@@
 \begin{equation*}
-S(\bm{u}) = -\frac{\rho s}{\gamma-1}, 
+S(\bm{u}) = -\bnote{\rho s}, %\frac{\rho s}{\gamma-1}, 
 \end{equation*}
@@
 It is known that reduced models can sometimes utilize a larger CFL compared the full order model \cite{lucia2004reduced, knezevic2011reduced}; however, to ensure that temporal errors are small, all solutions are computed using the same CFL used to generate the solution snapshots.
-Figure~\ref{fig:romsols} shows density computed using 25, 75, and 125 modes and the hyper-reduced treatment of viscosity in (\ref{eq:visc2}).
+Figure~\ref{fig:romsols} shows density computed using 25, 75, and 125 modes and the hyper-reduced treatment of viscosity in (\ref{eq:visc2}) \bnote{using $\epsilon = 2e-4$}.
 For all resolutions, the shock is under-resolved and the solution possesses Gibbs-type oscillations.
@@
 \subfloat[125 modes, $T=.75$]{\includegraphics[width=.32\textwidth]{euler1Dwall_t75_125modes.png}}
-\caption{Density $\rho$ computed using 25, 75, and 125 modes at times $T=.25, .75$.
+\caption{Density $\rho$ computed using 25, 75, and 125 modes at times $T=.25, .75$ \bnote{for $\epsilon = 2e-4$}.
 The solution of the reduced order model is plotted at convective hyper-reduced points, shown in \textcolor{red}{red}.
@@
 At all time-steps, $\LRb{\bm{v}_N^T \bm{V}_h^T\LRp{\bm{Q}_h \circ \bm{F}}\bm{1}}$ is near $O\LRp{10^{-14}}$ (close to machine precision), confirming that the formulation (\ref{eq:esbchr}) is discretely entropy conservative in the absence of viscosity and boundary flux penalization terms.
-We also run the ROM with the viscosity parameter $\epsilon = 0$.
 For 125 modes, the solution is significantly more oscillatory than the case of $\epsilon = 2\times 10^{-4}$; however, the reduced solution does not blow up.
@@
 
-We now compare the evolution of both the $L^2$ error and discrete entropy, and examine the effect of hyper-reduction on the discrete solution in Figure~\ref{fig:entropyevo}.
+We now compare the evolution of both the discrete entropy \rnote{and the discrete $L^2$ error between all solution components of the full and reduced order models}, and examine the effect of hyper-reduction on the discrete solution in Figure~\ref{fig:entropyevo}.
+\rnote{Here, the discrete $L^2$ norm is defined as $\nor{\bm{u}}^2 = \Delta x^2 \bm{u}^T\bm{u}$ in 1D or $\nor{\bm{u}}^2 = (\Delta x \Delta y)^2 \bm{u}^T\bm{u}$ in 2D, and is the analogue of the continuous $L^2$ norm evaluated at discretization points for the full order model.}
 Both error and entropy are computed on the original grid of the full order model (e.g., without hyper-reduction), and the error is computed as the discrete relative $L^2$ error over all the conservative variables.
@@
 \centering
-\includegraphics[width=.4\textwidth]{figs/khsvd.png}
+\includegraphics[width=.4\textwidth]{khsvd.png}
 \caption{Decay of singular values for snapshots of the 2D smoothed Kelvin-Helmholtz instability.}
@@
 \centering
-\subfloat[Full order model]{\includegraphics[width=.32\textwidth]{figs/khfom.png}}
+\subfloat[Full order model]{\includegraphics[width=.32\textwidth]{khfom.png}}
 \hspace{.175em}
-\subfloat[Reduced order model, 75 modes]{\includegraphics[width=.32\textwidth]{figs/khrom75.png}}
+\subfloat[Reduced order model, 75 modes]{\includegraphics[width=.32\textwidth]{khrom75.png}}
 \hspace{.1em}
-\subfloat[Hyper-reduced points]{\includegraphics[width=.32\textwidth]{figs/khrom75_hrpts.png}}
+\subfloat[Hyper-reduced points]{\includegraphics[width=.32\textwidth]{khrom75_hrpts.png}}
 \caption{Full and reduced order density for the 2D smoothed Kelvin-Helmholtz instability at $T = 3$ with $\epsilon = 1e-3$.
@@
 \centering
-\includegraphics[width=.4\textwidth]{figs/pulse2dsvd.png}
+\includegraphics[width=.4\textwidth]{pulse2dsvd.png}
 \caption{Decay of singular values for snapshots of the 2D Gaussian pulse problem.}
@@
 
+\subsection{\bnote{On computational performance}}
+
+\bnote{We note that the entropy stable reduced order models implemented in this work do not reduce computational cost compared to the original full order models due to the use of explicit time-stepping.
+Explicit time-stepping is used in this paper to validate the proposed semi-discrete formulation.
+However, we expect that for implicit time-stepping, entropy stable ROMs will see more significant efficiency gains.
+
+The cost of explicit time-stepping schemes scales with the cost of the ODE right hand side evaluation, and the cost of a right hand side evaluation for an entropy stable scheme scales with the number of entropy conservative flux evaluations required.
+The number of entropy conservative flux evaluations is the same as the number of nonzero entries in the matrix $\bm{Q}^i$, since the nonlinear term is $\LRp{\bm{Q}^i\circ\bm{F}^i}\bm{1}$ and the entries of the matrix $\bm{F}^i$ (which are flux evaluations between different states) are evaluated on the fly.
+
+For the 2D full order methods in this paper, the differentiation matrices $\bm{Q}^i$ are sparse with $O(K^2)$ non-zero entries.
+Thus, the cost of a 2D entropy stable finite volume method is $O(K^2)$ nonlinear flux evaluations for each coordinate direction.
+For a reduced order model, however, the hyper-reduced matrices $\bm{Q}^i_t$ are dense, and evaluating $(\bm{Q}^i_t \circ \bm{F}^i)\bm{1}$  requires $O(N_s^2)$ flux evaluations where $N_s$ is the number of hyper-reduced points.
+As an example, consider the Kelvin-Helmholtz results in Figure~\ref{fig:khrom}.
+The full order model requires $O(40000)$ flux evaluations.
+However, because the ROM has $884$ hyper-reduced points, $O(781,456)$ flux evaluations are required, resulting in a ROM which is more expensive per time-step than the full order model.
+
+While the proposed ROMs could still provide savings for sufficiently large full order models, this increased cost is a significant issue for explicit time-stepping.
+The cost is offset slightly if the ROM has a larger CFL \cite{marley2015reduced}; however, this situation only occurs if a smaller number of modes are used to simulate the solution.
+The situation changes for implicit time-stepping, as full order models require multiple solutions of large (but sparse) linear systems at each time-step.
+For a $K\times K$ grid, these linear systems would require $O(K^2)$ flux evaluations to assemble, but we expect the solution of the linear system to be the dominant cost.
+For a ROM using implicit time-stepping, the assembly of the linear system still requires $O(N_s^2)$ flux evaluations, but the size of the system scales with the number of modes, which is notably smaller than even the number of hyper-reduced points.
+For example, the Kelvin-Helmholtz example uses only $75$ basis functions per component, compared to $884$ hyper-reduced points.
+We thus expect that the use of entropy stable ROMs will result in a more significant speedup under implicit time-stepping.
+}
+
 \section{Conclusion}  We have presented a methodology for constructing projection-based reduced order models for nonlinear conservation laws by combining a reduced basis, a modified Galerkin projection, and tailored hyper-reduction techniques.

### 2020-04-03 docs/ESDG_jacobian/main.tex
updating docs

@@
 This can be done using Newton's method, which involves the Jacobian matrix of the nonlinear equations.
-While it is possible to compute the solution to the nonlinear system without explicitly computing the Jacobian matrix using Jacobian-free Newton-Krylov methods \cite{knoll2004jacobian, birken2019subcell}, the Jacobian matrix is also useful for preconditioning \cite{persson2006efficient, persson2008newton}.
+While it is possible to compute the solution to the nonlinear system without explicitly computing the Jacobian matrix using Jacobian-free Newton-Krylov methods \cite{knoll2004jacobian, birken2019subcell}, the Jacobian matrix is commonly used to construct preconditioners \cite{persson2008newton}.
 In this work, we present explicit formulas for Jacobian matrices of systems resulting from entropy stable formulations.
@@
 We assume without loss of generality periodic boundary conditions, which will simplify the presentation of the main results.  
-Extensions to non-periodic boundaries are treated in Section~\ref{eq:npbc}.
-The generalization to systems of nonlinear conservation laws and higher dimensions is also postponed until Section~\ref{sec:systems}.
+%Extensions to non-periodic boundaries are treated in Section~\ref{eq:npbc}.  
+The generalization to systems of nonlinear conservation laws and higher dimensions is postponed until Section~\ref{sec:systems}.
 
@@
 AD is distinct from both symbolic differentiation and finite difference approximations in that it does not return an explicit expression for derivatives, but is still able to evaluate derivatives accurately up to machine precision.
-Given a computational function, AD returns a second black-box function which evaluates the derivative.
+Given a computational function, AD provides a black-box function which evaluates the derivative.
 
@@
 \\
+The computation of Jacobians for vector-valued flux functions can be performed similarly using \verb+ForwardDiff.jacobian+.
 This simple API is made possible by the flexible Julia type system \cite{bezanson2017julia}.
@@
 However, because AD scales with the number of inputs and outputs, the cost of applying AD directly to $\bm{f}(\bm{u})$ increases as the discretization resolution increases.
-In contrast, using the approach in this paper, the entries of the Jacobian matrix can be computed for roughly the same cost as a single evaluation of $\bm{f}(\bm{u})$.
+In contrast, using the approach in this paper, AD is applied only to the flux function, which has a small number of inputs and outputs which are fixed independently of the discretization resolution.
+As a result, the cost of evaluating derivatives of the flux function is roughly the same cost as evaluating the flux function, and entries of the Jacobian matrix can be computed for roughly the same cost as a single evaluation of the nonlinear flux $\bm{f}(\bm{u})$.
 Moreover, explicit Jacobian formulas enable implementations to take advantage of sparsity in $\bm{Q}$ when assembling the Jacobian matrix.
@@
 A local formulation on the element $D^k$ is given by
-\eqlab{
-J_k {\bm{M}} \td{\bm{u}_k}{t} + 2\LRp{{\bm{Q}} \circ \bm{F}^k}\bm{1} + \bm{B}\LRp{\bm{f}^*-f(\bm{u}_k)}   = \bm{0}, \qquad \bm{f}^* = \bmat{
+\begin{gather}
+J_k {\bm{M}} \td{\bm{u}_k}{t} + 2\LRp{{\bm{Q}} \circ \bm{F}^k}\bm{1} + \bm{B}\LRp{\bm{f}^*-f(\bm{u}_k)}  = \bm{0}, \nonumber\\
+\bm{f}^* = \bmat{
 f_S(u_{1,k}^+,u_{1,k})\\
@@
 \vdots\\
+0\\
 f_S(u_{N_p,k}^+,u_{N_p,k})
-},
+}.
 \label{eq:sbp1D}
-}
+\end{gather}
 where $u_{1,k}^+, u_{N_p,k}^+$ denote the exterior values of $u_{1,k}, u_{N_p,k}$ on neighboring elements.
@@
 The construction of skew matrices $\bm{Q}_{\Omega}$ follows from approaches detailed in \cite{crean2018entropy, chan2017discretely, chan2018discretely, chan2018efficient, chan2019skew, hicken2020entropy}.
-We omit it for conciseness in this work.
+We omit it for conciseness in this work, but note that the efficient assembly of DG matrices is discussed in detail elsewhere \cite{persson2008newton, hillewaert2011sharp, kempf2018automatic}.
 
@@
 %where $\bm{e}_n\in \mathbb{R}^n$ is the vector of all ones, and $\bm{e}_n\bm{e}_n^T$ is the $n\times n$ matrix of all ones.  
-Let $(k_1, j_1)$ and $(k_2, j_2)$ be tuples which correspond to row and column indices of a matrix.  
-We now define the flux matrix $\bm{F}$ as the $n\times n$ block matrix containing flux interactions between different nodes
+Let $(k_1, j_1)$ and $(k_2, j_2)$ be multi-indices which correspond to row and columns indices of a matrix, respectively.  
+We now define the flux matrix $\bm{F}$ as the $n\times n$ block diagonal matrix containing flux interactions between different nodes
 \eqlab{
@@
 }
-where $\bm{u}_{:,k,j}$ denotes the vector containing all solution components at the $k$th element and $j$th node, and each block $\bm{F}_\ell$ for $\ell = 1,\ldots, n$ corresponds to the $\ell$th component of the vector-valued flux evaluated at solution states $\bm{u}_{:,k_1,j_1},\bm{u}_{:,k_2,j_2}$.
+where $\bm{u}_{:,k,j}$ denotes the vector containing all solution components at the $k$th element and $j$th node, and each entry of the block $\bm{F}_\ell$ for $\ell = 1,\ldots, n$ corresponds to the $\ell$th component of the vector-valued flux evaluated at solution states $\bm{u}_{:,k_1,j_1},\bm{u}_{:,k_2,j_2}$.
 
@@
 \[
-\LRp{\bm{I}_n\otimes \bm{M}_{\Omega}}\td{\bm{u}_{\Omega}}{t} + 2\LRp{ \LRp{\bm{e}_n\bm{e}_n^T\otimes \bm{Q}_{\Omega} } \circ \bm{F}}\bm{1} = \bm{0}.
+\LRp{\bm{I}_n\otimes \bm{M}_{\Omega}}\td{\bm{u}_{\Omega}}{t} + 2\LRp{ \LRp{\bm{I}_n\otimes \bm{Q}_{\Omega} } \circ \bm{F}}\bm{1} = \bm{0}.
 \]
@@
 \[
-\bm{f}(\bm{u}) = 2\LRp{ \bm{Q}_{\Omega} \circ \bm{F}}\bm{1} 
+\bm{f}(\bm{u}) = 2\LRp{ \LRp{\bm{I}_n\otimes\bm{Q}_{\Omega}} \circ \bm{F}}\bm{1} 
 = 2\bmat{(\bm{Q}_{\Omega}\circ \bm{F}_1) \\ 
@@
 \eqlab{
-\pd{\bm{f}}{\bm{u}} &= 2\bmat{
-\bm{Q}_{\Omega}\circ \bm{F}_{1,\bm{u}_1} & \ldots & \bm{Q}_{\Omega}\circ \bm{F}_{1,\bm{u}_n}\\ 
+\pd{\bm{f}}{\bm{u}} &=
+\bmat{
+\bm{\partial F}_{1,\bm{u}_1} & \ldots & \bm{\partial F}_{1,\bm{u}_n}\\
 \vdots & \ddots & \vdots\\
-\bm{Q}_{\Omega}\circ \bm{F}_{n,\bm{u}_1} & \ldots & \bm{Q}_{\Omega}\circ \bm{F}_{n,\bm{u}_n}
-} \label{eq:fluxjac}\\
-&= \LRp{\bm{e}_n\bm{e}_n^T\otimes \bm{Q}_{\Omega}} \circ 
-\bmat{
-\bm{F}_{1,\bm{u}_1} & \ldots &\bm{F}_{1,\bm{u}_{n}}\\
-\vdots & \ddots & \vdots \\
-\bm{F}_{n,\bm{u}_1} & \ldots &\bm{F}_{n,\bm{u}_n}\nonumber
-}.
+\bm{\partial F}_{n,\bm{u}_1} & \ldots & \bm{\partial F}_{n,\bm{u}_n}
 }
-Here, the flux matrices $\bm{F}_{i,\bm{u}_j}$ are evaluated via
+% 2\bmat{
+%\bm{Q}_{\Omega}\circ \bm{F}_{1,\bm{u}_1} & \ldots & \bm{Q}_{\Omega}\circ \bm{F}_{1,\bm{u}_n}\\ 
+%\vdots & \ddots & \vdots\\
+%\bm{Q}_{\Omega}\circ \bm{F}_{n,\bm{u}_1} & \ldots & \bm{Q}_{\Omega}\circ \bm{F}_{n,\bm{u}_n}
+%} + \text{\note{add diagonal terms }}\label{eq:fluxjac}\\
+%&= \LRp{\bm{e}_n\bm{e}_n^T\otimes \bm{Q}_{\Omega}} \circ 
+%\bmat{
+%\bm{F}_{1,\bm{u}_1} & \ldots &\bm{F}_{1,\bm{u}_{n}}\\
+%\vdots & \ddots & \vdots \\
+%\bm{F}_{n,\bm{u}_1} & \ldots &\bm{F}_{n,\bm{u}_n}\nonumber
+%}.
+}
+where the Jacobian block $\bm{\partial F}_{i,\bm{u}_j}$ is evaluated as in Lemma~\ref{lemma:explicitJ}
 \[
-\LRp{\bm{F}_{i,\bm{u}_j}}_{j_1k_1,j_2k_2} = \LRl{\pd{\LRp{\bm{f}_S}_i}{\bm{u}_j}}_{\bm{u}_{:,k_1,j_1},\bm{u}_{:,k_2,j_2}},
+\bm{\partial F}_{i,\bm{u}_j} = 2 \LRp{\bm{Q}_{\Omega} \circ \bm{F}_{i,\bm{u}_j}} \pm \diag{\bm{1}^T\LRp{2\bm{Q}_{\Omega}\circ \bm{F}_{i,\bm{u}_j}}}.
 \]
-where $\pd{\LRp{\bm{f}_S}_i}{\bm{u}_{j}}$ denotes the derivative of the $i$th component of the flux $\bm{f}_S\LRp{\bm{u}_L,\bm{u}_R}$ with respect to the $j$th solution component of $\bm{u}_R$.
-Thus, each block $\bm{F}_{i,\bm{u}_j}$ corresponds to an entry of the Jacobian of $\bm{f}_S(\bm{u}_L,\bm{u}_R)$).
+for $\bm{Q}_{\Omega} = \pm \bm{Q}_{\Omega}^T$.  
+Here, the flux matrix $\bm{F}_{i,\bm{u}_j}$ is evaluted via
+\[
+\LRp{\bm{F}_{i,\bm{u}_j}}_{j_1k_1,j_2k_2} = \LRl{\pd{\LRp{\bm{f}_S}_i}{\bm{u}_{R,j}}}_{\bm{u}_{:,k_1,j_1},\bm{u}_{:,k_2,j_2}},
+\]
+where $\pd{\LRp{\bm{f}_S}_i}{\bm{u}_{R,j}}$ denotes the derivative of the $i$th component of the flux $\bm{f}_S\LRp{\bm{u}_L,\bm{u}_R}$ with respect to the $j$th solution component of $\bm{u}_R$.
+Thus, each entry of the block $\bm{F}_{i,\bm{u}_j}$ corresponds to an entry of the Jacobian (with respect to $\bm{u}_R$) of $\bm{f}_S(\bm{u}_L,\bm{u}_R)$ and an entry of the global differentiation matrix $\bm{Q}_{\Omega}$.
 
 \begin{remark}
-We note that an alternative ordering of degrees of freedom \cite{crean2018entropy} yields a Jacobian matrix with a more compact bandwidth.
-A smaller matrix bandwidth can also be achieved by bandwidth-minimizing algorithms such as the reverse Cuthill-McKee algorithm \cite{cuthill1969reducing}.
+An alternative ordering of degrees of freedom \cite{crean2018entropy} yields a Jacobian matrix with a more compact bandwidth.
+A smaller matrix bandwidth can also be achieved by precomputing bandwidth-minimizing row and column permutations using the reverse Cuthill-McKee algorithm \cite{cuthill1969reducing}.
 %Let $\bm{u}_{\Omega}$ now denote a vector of vectors 
@@
 
-\section{Generalizations to different numerical settings}
+%\section{Generalizations to different numerical settings}
 
-In this section, we discuss how to generalize the explicit construction of Jacobians to different numerical settings, including non-collocation schemes and non-periodic boundary conditions.
+%In this section, we discuss how to generalize the explicit construction of Jacobians to different numerical settings, including non-collocation schemes and non-periodic boundary conditions.
 
-\subsection{Non-collocated schemes: hybridized SBP operators, entropy projection, and over-integration}
+\section{Non-collocated schemes: hybridized SBP operators, entropy projection, over-integration}
 
@@
 
-\subsubsection{Entropy conservative schemes and hybridized SBP operators}
+\subsection{Entropy conservative schemes and hybridized SBP operators}
 
@@
 \]
-We also construct global interpolation, projection, and extrapolation matrices
+We also abuse notation and redefine $\bm{V},\bm{E},\bm{P},\bm{V}_h$ as \emph{global} interpolation, projection, and extrapolation matrices
 \eq{
@@
 We define the entropy projected variables $\tilde{\bm{u}}_{\Omega}$ as
-\[
+\eqlab{
 \tilde{\bm{u}}_{\Omega} = \bm{u}\LRp{\bm{V}_h\bm{P}\bm{v}\LRp{\bm{V}\hat{\bm{u}}_{\Omega}}}.
-\]
-We evaluate each flux block $\bm{F}_{\ell}$ in (\ref{eq:fluxsys}) using the entropy projected variables
-\[
+\label{eq:entropyprojection}
+}
+Let $\bm{F}$ again denote the block-diagonal flux matrix in (\ref{eq:fluxsys}).  We evaluate each flux block $\bm{F}_{\ell}$  using the entropy projected variables
+\eqlab{
 \LRp{\bm{F}_\ell}_{(k_1, j_1),(k_2,j_2)} = \LRp{\bm{f}_{S}\LRp{\tilde{\bm{u}}_{:, k_1,j_1},\tilde{\bm{u}}_{:, k_2,j_2}}}_{\ell}.
-%\LRp{\bm{F}_{i,\bm{u}_j}}_{j_1k_1,j_2k_2} = \LRl{\pd{\LRp{\bm{f}_S}_i}{\bm{u}_j}}_{\tilde{\bm{u}}_{:,k_1,j_1},\tilde{\bm{u}}_{:,k_2,j_2}},
-\]
+\label{eq:fluxell}
+}
 Then, an entropy conservative method is given by 
@@
 \]
+where $\bm{I}_n$ is the $n\times n$ identity matrix.
 
-\subsubsection{Jacobian matrices for hybridized SBP schemes}
+\subsection{Jacobian matrices for hybridized SBP schemes}
 
-Redefine the nonlinear term 
+We redefine the nonlinear term as
 \[
@@
 \]
-We introduce $\pd{\bm{u}}{\bm{v}}$ and $\pd{\bm{v}}{\bm{u}}$ as Jacobians of the conservative variables with respect to the entropy variables and vice versa.
+where the flux matrix $\bm{F}$ is computed using the entropy projected conservative variables (\ref{eq:entropyprojection}) via (\ref{eq:fluxell}).
+Let $\pd{\bm{u}}{\bm{v}}$ and $\pd{\bm{v}}{\bm{u}}$ denote Jacobians of the conservative variables with respect to the entropy variables and vice versa.
 These have been explicitly derived for several equations (for example, the Jacobians for the compressible Navier-Stokes equations are given in \cite{hughes1986new}).
 
-We can compute the Jacobian of $\bm{f}(\hat{\bm{u}})$ via the chain rule.  We assume $n=1$ for simplicity, and illustrate this for an entry $i\neq j$ of the Jacobian
+We can compute the Jacobian of $\bm{f}(\hat{\bm{u}})$ via the chain rule.  We assume a scalar equation $n=1$ for simplicity, and motivate our approach by considering an entry $i\neq j$ of the Jacobian
 \[
@@
 \LRp{\pd{}{\hat{\bm{u}}_{\Omega}} \LRp{\bm{Q}_{\Omega}\circ\bm{F}}\bm{1}}_{ij} &= \pd{}{\hat{\bm{u}}_{\Omega,j}} \sum_{k} \LRp{\bm{Q}_{\Omega}}_{ik} \bm{f}_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_k} \\
-&=  2 \bm{V}_h^T \sum_{k} \LRp{\bm{Q}_{\Omega}}_{ik} \LRl{\pd{\bm{f}_S}{y}}_{\tilde{\bm{u}}_i,\tilde{\bm{u}}_k} \pd{\tilde{\bm{u}}_i}{\hat{\bm{u}}_{\Omega,j}}
+&=  \sum_{k} \LRp{\bm{Q}_{\Omega}}_{ik} \LRl{\pd{\bm{f}_S}{y}}_{\tilde{\bm{u}}_i,\tilde{\bm{u}}_k} \pd{\tilde{\bm{u}}_i}{\hat{\bm{u}}_{\Omega,j}}
 }
@@
 \]
+The construction of $\pd{\tilde{\bm{f}}}{\tilde{\bm{u}}}$ for systems ($n > 1$) is similar to the procedure described in Section~\ref{sec:systems}.
 Now, let $\tilde{\bm{v}}$ denote the projected entropy variables at evaluated at volume quadrature points
@@
 \[
-\pd{\tilde{\bm{u}}}{\hat{\bm{u}}_{\Omega}} = \LRl{\pd{\bm{u}}{\bm{v}}}_{\tilde{\bm{u}}} \bm{V}_h \bm{P} \LRl{\pd{\bm{v}}{\bm{u}}}_{\tilde{\bm{v}}} \bm{V}.
+\pd{\tilde{\bm{u}}}{\hat{\bm{u}}_{\Omega}} = \LRl{\pd{\bm{u}}{\bm{v}}}_{\tilde{\bm{v}}} \bm{V}_h \bm{P} \LRl{\pd{\bm{v}}{\bm{u}}}_{\bm{V}\hat{\bm{u}}_{\Omega}} \bm{V}.
 \]
@@
 \raiserows{0}{\matbox{7}{7}{N_{\rm total}}{N_{\rm total}}{\bm{Q}_{\Omega}\circ\pd{\tilde{\bm{f}}}{\tilde{\bm{u}}}}}
-\raiserows{0}{\matbox{7}{7}{N_{\rm total}}{N_{\rm total}}{\LRl{\pd{\bm{u}}{\bm{v}}}_{\tilde{\bm{u}}}}}
+\raiserows{0}{\matbox{7}{7}{N_{\rm total}}{N_{\rm total}}{\LRl{\pd{\bm{u}}{\bm{v}}}_{\tilde{\bm{v}}}}}
 \raiserows{0}{\matbox{7}{5}{N_{\rm total}}{N_q}{ \bm{V}_h\bm{P} }}
-\raiserows{1}{\matbox{5}{5}{N_q}{N_q}{ \LRl{\pd{\bm{v}}{\bm{u}}}_{\tilde{\bm{v}}} }}
+\raiserows{1}{\matbox{5}{5}{N_q}{N_q}{ \LRl{\pd{\bm{v}}{\bm{u}}}_{\bm{V}\hat{\bm{u}}_{\Omega}} }}
 \raiserows{1}{\matbox{5}{3}{N_q}{N_p}{\bm{V}}}
@@
 
-\subsection{Non-periodic boundary conditions}
-\label{eq:npbc}
-
-Finally, we consider the case of non-periodic boundary conditions.  \note{Impose BCs by specifying exterior value.}
+%\subsection{Non-periodic boundary conditions}
+%\label{eq:npbc}
+%
+%Finally, we consider the case of non-periodic boundary conditions.  \note{Impose BCs by specifying exterior value.}
 

### 2020-04-03 docs/ESDG_mortar/main.tex
updating docs

@@
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(1,0.0302287)(0.5,0.0204278)(0.25,0.00988213)(0.125,0.00383616)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(1,0.0312387)(0.5,0.020413)(0.25,0.00690005)(0.125,0.0017994)};
+coordinates{(1,0.0300089)(0.5,0.02037)(0.25,0.0103349)(0.125,0.00395744)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.0309619)(0.5,0.0201996)(0.25,0.00698411)(0.125,0.0018293)};
 
@@
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(1,0.025246)(0.5,0.00858607)(0.25,0.00156388)(0.125,0.000334682)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(1,0.0236979)(0.5,0.00616892)(0.25,0.000963938)(0.125,0.000249721)};
+coordinates{(1,0.0252826)(0.5,0.00898116)(0.25,0.00161751)(0.125,0.000284006)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.0237929)(0.5,0.00638897)(0.25,0.000964247)(0.125,0.000126189)};
 
@@
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(1,0.225344)(0.5,0.165699)(0.25,0.0820851)(0.125,0.0494277)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(1,0.222488)(0.5,0.144629)(0.25,0.073849)(0.125,0.0217714)};
+coordinates{(1,0.225732)(0.5,0.175585)(0.25,0.0875643)(0.125,0.050868)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.221903)(0.5,0.144126)(0.25,0.0724546)(0.125,0.0219034)};
 
@@
 \addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(1,0.2011)(0.5,0.0658348)(0.25,0.0233793)(0.125,0.00727696)};
-\addplot+[semithick, mark options={solid, fill=markercolor}]
-coordinates{(1,0.159118)(0.5,0.0478442)(0.25,0.0177047)(0.125,0.00421538)};
+coordinates{(1,0.198065)(0.5,0.0693157)(0.25,0.024472)(0.125,0.00612038)};
+\addplot+[semithick, mark options={solid, fill=markercolor}]
+coordinates{(1,0.178736)(0.5,0.0575655)(0.25,0.016464)(0.125,0.0029831)};
 
